### 2021-12-09 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
gem cf analysis

---
title: "Pre/Post-Trikafta"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/Prepost Triakfta")
```

```{r import}
cgm_c <- read.csv("./Data_Raw/Data From CHCO/REDCap Upload_chco.csv")
```

### 2022-01-11 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
Update pre_post_trikafta.Rmd
genotypes

@@
 library(tidyverse)
+library(readxl)
 library(performance)
@@
 
-```{r import}
+```{r chco data}
+# Import CGM
 cgm_c <- read.csv("./Data_Raw/Data From CHCO/REDCap Upload_chco.csv")
+cgm_c$date_cgm <- as.Date(cgm_c$date_cgm_placement, "%Y-%m-%d")
+
+cgm_m <- read.csv("./Data_Raw/Data From Montana/CGM_Montana/REDCap Upload_mt.csv")
+cgm_m$date_cgm <- as.Date(cgm_m$date_cgm_placement, "%m/%d/%Y")
+
+cgm_w <- read.csv("./Data_Raw/Data From Washington/REDCap Upload_Washington.csv")
+cgm_w$date_cgm <- as.Date(cgm_w$date_cgm_placement, "%Y-%m-%d")
+
+cgm <- rbind(cgm_c, cgm_m,cgm_w )
+
+reg0 <- read.csv("./Data_Raw/Data From CHCO/Registry/Copy of Earliest Trikafta Dates in Registry_updated 2.2.21 cc modified_cfrdyn.csv")
+# Exclude participants
+reg <- subset(reg0,!reg0$MRN %in% c("1714343","787372"))
+# Dates and A1c
+reg$Earliest.Trikafta.Date.in.PortCF <- ifelse(reg$MRN =="1053930", "12/1/2019", 
+                                               ifelse(reg$MRN =="1695512", "12/12/2019", 
+                                                      ifelse(reg$MRN =="815274", "12/13/2019", 
+                                                             ifelse(reg$MRN =="1078408",   "3/9/2020",
+                                                                    ifelse(reg$MRN =="899717", "2/14/2020", as.character(reg$Earliest.Trikafta.Date.in.PortCF))))))
+
+a1c <- read.csv("./Data_Raw/Data From CHCO/Copy of A1C and OGTTs Before and After Trikafta_cc modified 8.25.20.csv")
+# Weight
+weight <- read.csv("./Data_Raw/Data From CHCO/Registry/weightdata.csv")
+
+weight$Earliest.Trikafta.Date.in.PCF <- 
+  ifelse(weight$MRN =="1053930", "12/1/2019", 
+         ifelse(weight$MRN =="1695512", "12/12/2019", 
+                ifelse(weight$MRN =="815274", "12/13/2019", 
+                       ifelse(weight$MRN =="1078408",   "3/9/2020",
+                              ifelse(weight$MRN =="899717", "2/14/2020", as.character(weight$Earliest.Trikafta.Date.in.PCF))))))
+
 ```
+
+```{r montana data}
+reg_montana <- 
+  read.csv("./Data_Raw/Data From Montana/Copy of Paper patient population glycemic (+hofer).csv")
+demo_adult <- read.csv("./Data_Raw/Data From Montana/demo_adults.csv")
+demo_kids <- read.csv("./Data_Raw/Data From Montana/demo_kids.csv")
+
+reg_montana$age <- ifelse(reg_montana$CFF.ID %in% demo_adult$CFF.ID, "Adult", 
+                          ifelse(reg_montana$CFF.ID %in% demo_kids$CFF.ID, "Kids", "Check"))
+
+reg_montana$trik <- as.Date(reg_montana$Trikafta.start.date,"%m/%d/%Y")
+
+reg_montana$test <- as.Date(reg_montana$Date.of.Encounter,  "%m/%d/%Y")
+reg_montana$prepost <- ifelse(reg_montana$trik > reg_montana$test, "before", "after")
+
+reg_montana$diff <- reg_montana$trik-reg_montana$test 
+reg_montana2 <- subset(reg_montana, abs(diff) <= 365)
+
+reg_montana3 <- reg_montana2 %>% group_by(CFF.ID, prepost) %>% count()
+reg_montana4 <- reg_montana3 %>% group_by(CFF.ID) %>% count()
+# Weight/FEV
+pft_adult <- read.csv("./Data_Raw/Data From Montana/Copy of PFTs growth adult (+hofer).csv")
+pft_kids <- read.csv("./Data_Raw/Data From Montana/PFTs kids.csv")
+
+pft_montana <- rbind(pft_adult, pft_kids)
+
+pft_montana_1 <- subset(pft_montana, pft_montana$CFF.ID %in% reg_montana$CFF.ID)
+# Age at pft
+pft_montana_1$pft <- as.Date(pft_montana_1$Date.of.Encounter,  "%m/%d/%Y")
+
+reg_montana_first <- reg_montana %>% group_by(CFF.ID) %>% filter(row_number()==1)
+
+reg_montana_first$trik <- ifelse(reg_montana_first$CFF.ID == "1044621", "2019-12-02", 
+                                 ifelse(reg_montana_first$CFF.ID == "1008380", "2020-1-1", as.character(reg_montana_first$trik)))
+
+pft_montana_2 <- merge(pft_montana_1, reg_montana_first[,c("CFF.ID" ,"Date.Of.Birth..mm.dd.yyyy.", "trik")])
+# Need pfts within a year of trik
+pft_montana_2$diff <-as.numeric(as.Date(pft_montana_2$pft) - as.Date(pft_montana_2$trik))
+pft_montana_3 <- subset(pft_montana_2, abs(diff) <= 365 )
+pft_montana_3$prepost <- ifelse(pft_montana_3$trik > pft_montana_3$pft, "before", "after")
+pft_montana_4 <- subset(pft_montana_3, pft_montana_3$prepost == "before")
+pft_montana_5 <- pft_montana_4 %>% group_by(CFF.ID) %>% arrange(diff) %>% filter(row_number()==1)
+
+pft_montana_5$DOB <- as.Date(pft_montana_5$Date.Of.Birth..mm.dd.yyyy.,  "%m/%d/%Y")
+
+pft_montana_5$age <- as.numeric(pft_montana_5$pft-pft_montana_5$DOB)/365.25
+
+pft_montana_5$BMI.Percentile
+# eed sex
+demo_montana <- rbind(demo_kids[,c("CFF.ID","Gender")], demo_adult[,c("CFF.ID","Gender")])
+
+analytic_montana <- merge(pft_montana_5, demo_montana ,by = "CFF.ID", all.x=T)
+# Need cfrd start date - don't have dates
+#gtube status
+gt_adult <- read.csv("./Data_Raw/Data From Montana/gtubes_adult.csv")
+gt_kids <- read.csv("./Data_Raw/Data From Montana/GT users kids.csv")
+
+gt_montana <- rbind(gt_adult, gt_kids)
+
+gt_montana_1 <- merge(gt_montana, reg_montana_first[,c("CFF.ID" , "trik")])
+gt_montana_1$gt <- as.Date(gt_montana_1$Date.of.Encounter,  "%m/%d/%Y")
+gt_montana_1$diff <-as.numeric(as.Date(gt_montana_1$gt) - as.Date(gt_montana_1$trik))
+gt_montana_2 <- subset(gt_montana_1, abs(diff) <= 365 )
+analytic_montana$gtube <- ifelse(analytic_montana$CFF.ID %in% gt_montana_2$CFF.ID, "Yes", "No")
+# Insufficent
+is_adult <- read.csv("./Data_Raw/Data From Montana/pancreatic insufficient adult.csv")
+is_kids <- read.csv("./Data_Raw/Data From Montana/pancreatic insufficient kids.csv")
+
+analytic_montana$is <- 
+  ifelse(analytic_montana$CFF.ID %in% c(is_adult$CFF.ID, is_kids$CFF.ID), "Insufficient", "Not")
+# Genotypes
+g_adult <- read.csv("./Data_Raw/Data From Montana/genotypes_adult.csv")
+g_kids <- read.csv("./Data_Raw/Data From Montana/genotypes kids.csv")[,1:3]
+geno_montana <- rbind(g_adult,g_kids)
+geno_montana$Genotypes1 <- geno_montana$Name.of.the.first.mutation
+geno_montana$Genotypes2 <- geno_montana$Name.of.the.second.mutation
+geno_montana$class2 <-
+  ifelse(geno_montana$Genotypes2%in%c('1717-1G->A','1717-1G-A','G542X','Q493X',
+                                      'R1162X','R553X','R553x','W1089X','W1282X',
+                                      '1078delT','R75X','3659delC','621+1G->T',
+                                      '621+1G>T','394delTT','3120+1g->A',
+                                      '1154InsTC','1154insTC','1213delT',
+                                      '1259insA','1288insTA','3791delC',
+                                      'E60X','K710X','2184delA','CFTRdele2,3',
+                                      '663delT','Glu528','1461ins4',
+                                      '306insA','R709X','CFTRdele22-24',
+                                      '2711delT','2183AA->G'),"I",
+         ifelse(geno_montana$Genotypes2 %in% c('F508','F508del','F508del ',
+                                               'G85E', 'I507','I 507',
+                                               'N1303K', 'A559T', 'R560T','A561E'),"II", 
+                ifelse(geno_montana$Genotypes2 %in% c('G551D', 'G551S','S549N'), "III", 
+                       ifelse(geno_montana$Genotypes2 %in% c('R334W','R347P', 
+                                                             'R117H','R117C', 
+                                                             'P67L','L206W', 
+                                                             'I206w','D614G', 
+                                                             'R347H','D1152H', 
+                                                             '3849+10kbC>T'),"IV", 
+                              ifelse(geno_montana$Genotypes2 %in% c('A455E', 
+                                                                    '2789+5G->A',
+                                                                    '3849+10CT', 
+                                                                    '3849+10kbC->T', 
+                                                                    '711+3A->G', 
+                                                                    '1898+5G->T',
+                                                                    'P574H', 
+                                                                    '3272-26A->G5'),"V","")))))   
+geno_montana$severity2 <- 
+  ifelse(geno_montana$Genotypes2%in%c('1717-1G->A','1717-1G-A','G542X','Q493X',
+                                      'R1162X','R553X','R553x','W1089X','W1282X',
+                                      '1078delT','R75X','3659delC','621+1G->T',
+                                      '621+1G>T','394delTT','3120+1g->A','1154InsTC',
+                                      '1154insTC','1213delT','1259insA','1288insTA',
+                                      '3791delC','E60X','K710X','2184delA',
+                                      'CFTRdele2,3','663delT','Glu528','1461ins4',
+                                      '306insA','R709X','CFTRdele22-24','2711delT',
+                                      '2183AA->G','F508','F508del','F508del','G85E',
+                                      'I507','I507','N1303K','A559T','R560T',
+                                      'A561E','G551D','G551S','S549N',
+                                      '1898+1G->A','2143delT','2183AA-G',
+                                      '2183delAA->G','2184insA','S434X',
+                                      '2585delT','3120+1G->A','3905insT','406-1G->A',
+                                      '712-1G->T','L1245X','Q2X','R1158X',
+                                      'R851X','V520F','L1077P','S489X','F311del',
+                                      'W846X','M1101K','2622+1G->A','S945L',
+                                      "Y1092X","R1066C","G178R","2957delT",
+                                      "CFTRdele1",'W1204X',"q493x","W57G",
+                                      "1585-8G>A","1717-8G->A","G1244E"),"Severe",
+         ifelse(geno_montana$Genotypes2%in%c('R334W','R347P','R117H','R117C',
+                                             'P67L','L206W','I206w','D614G',
+                                             'R347H','D1152H','3849+10kbC>T',
+                                             "D110H","R1070W",'A455E','2789+5G->A',
+                                             '3849+10CT','3849+10kbC->T','711+3A->G',
+                                             '1898+5G->T','P574H','3272-26A->G5',
+                                             '3849G->A','H199Y','H199y','S492F','S531P',
+                                             'L1335P','R352Q',"D1270N","2789+2insA",
+                                             "3849+10kbc->T","Q98R","Phe1052Val",
+                                             "F1052V","Tyr1014Cys","Y1014C","T338I",
+                                             "5T","c.2249C>T","P750L"),"Mild",
+                ifelse(geno_montana$Genotypes2%in%c('1262insA','I506T','3363delGT',
+                                                    'Q1209P','3349insT','G178R',
+                                                    'L467P','317insC','I336K',	
+                                                    'ex14a','3500-2A->G','c.3407_3422del16',
+                                                    '3349insT','1248+1G->A','M470V','1812-1G->A',
+                                                    'delX22-24','Q237H','Q237H','W57G',
+                                                    '1078delA','G178R','2957delT','G576A',
+                                                    'other','Other','Unknown','unknown',
+                                                    'CFTRdele1','1811+1.6kbA->','UNK',
+                                                    "I336K","M470V","S1235R","M470V","F508C",
+                                                    "R1066H","P205S"),"Unk","")))
+geno_montana$class1 <- 
+  ifelse(geno_montana$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                                             '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                                             '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                                             '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
+                              ifelse(geno_montana$Genotypes1 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
+                                     "II", 
+                                     ifelse(geno_montana$Genotypes1 %in% c('G551D', 'G551S', 'S549N'), "III", 
+                                            ifelse(geno_montana$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
+                                                   "IV", 
+                                                   ifelse(geno_montana$Genotypes1 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                                                         'P574H', '3272-26A->G5'), "V" , "")))))   
+geno_montana$severity1 <- ifelse(geno_montana$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                                                '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                                                '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                                                '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
+                                                                'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
+                                                                '2183delAA->G', '2184insA', 'S434X',
+                                                                '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
+                                                                'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
+                                                                "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
+                                 "Severe", 
+                                 ifelse(geno_montana$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
+                                                                       'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                                       'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
+                                                                       'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
+                                                                       "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
+
+                                        ifelse(geno_montana$Genotypes1 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
+                                                                                'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
+                                                                                'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
+                                                                                'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
+                                                                                "R1066H", "P205S"), 
+                                               "Unk","")))
+
+geno_montana$genoRisk <- ifelse(geno_montana$severity1 == "Mild" | geno_montana$severity2 == "Mild", "Low", 
+                                ifelse(geno_montana$severity1 == "Severe" & geno_montana$severity2 == "Severe", "High", ""))
+
+analytic_montana$genotype <- ifelse(analytic_montana$CFF.ID %in% geno_montana[geno_montana$genoRisk == "High", ]$CFF.ID, "High", 
+                                    ifelse(analytic_montana$CFF.ID %in% geno_montana[geno_montana$genoRisk == "Low", ]$CFF.ID, "Low",   ""   ))
+
+```

### 2022-01-11 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
Update pre_post_trikafta.Rmd

--- 2021-12-09 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
@@
 library(tidyverse)
-library(performance)
+library(readxl)
+library(AGD)
 library(knitr)
@@
 
-```{r import}
-cgm_c <- read.csv("./Data_Raw/Data From CHCO/REDCap Upload_chco.csv")
+```{r chco data}
+# CGM
+chco_cgm = read.csv("./Data_Cleaned/chco_cgm.csv")
+chco_cgm$timepoint = sub(".*_","",chco_cgm$subject_id)
+chco_cgm$timepoint = factor(chco_cgm$timepoint)
+levels(chco_cgm$timepoint) = c("Post","Post","Pre","Pre")
+chco_cgm$subject_id = sub("_.*","",chco_cgm$subject_id)
+colnames(chco_cgm)[which(colnames(chco_cgm) == "subject_id")] = "CFF.ID"
+# Start dates
+chco_dates = read.csv("./Data_Cleaned/chco_trikafta_dates.csv")
+# Fix select dates
+chco_dates$Earliest.Trikafta.Date.in.PortCF[chco_dates$MRN =="1053930"] = "12/1/2019"
+chco_dates$Earliest.Trikafta.Date.in.PortCF[chco_dates$MRN =="1695512"] = "12/12/2019"
+chco_dates$Earliest.Trikafta.Date.in.PortCF[chco_dates$MRN =="815274"] = "12/13/2019"
+chco_dates$Earliest.Trikafta.Date.in.PortCF[chco_dates$MRN =="1078408"] = "3/9/2020"
+chco_dates$Earliest.Trikafta.Date.in.PortCF[chco_dates$MRN =="899717"] = "2/14/2020"
+# A1C and OGTT
+chco_a1c = read.csv("./Data_Cleaned/chco_a1c_ogtt.csv")
+chco_a1c$CFF.ID = as.character(chco_a1c$CFF.ID)
+chco_pft = read.csv("./Data_Cleaned/chco_bmi_pft.csv")
+# Put everything together
+chco = left_join(chco_cgm,chco_a1c,by = "CFF.ID")
 ```

### 2022-01-13 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
Update pre_post_trikafta.Rmd

--- 2021-12-09 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
@@
 library(tidyverse)
-library(performance)
+library(readxl)
+library(PAutilities)
+library(AGD)
+library(rspiro)
+library(anthro)
 library(knitr)
@@
 
-```{r import}
-cgm_c <- read.csv("./Data_Raw/Data From CHCO/REDCap Upload_chco.csv")
+```{r code from maxie,eval=FALSE}
+# Source Maxie's code, and write final ID lists
+source("~/GitHub/CF/Christine Chan/Pre-post Trikafta/p001+cmw_UPDATED.R")
+ids = data.frame(unique(analytic_cm$CFF.ID))
+colnames(ids) = "CFF.ID"
+write.csv(ids,file = "./Data_Cleaned/final_ids.csv",row.names = F)
 ```
+
+# Table 1: Demographics at Trikafta Start
+
+```{r}
+# Start dates
+chco_dates = read.csv("./Data_Cleaned/chco_trikafta_dates.csv")
+# Fix select dates
+chco_dates$Earliest.Trikafta.Date.in.PortCF[chco_dates$MRN =="1053930"] = "12/1/2019"
+chco_dates$Earliest.Trikafta.Date.in.PortCF[chco_dates$MRN =="1695512"] = "12/12/2019"
+chco_dates$Earliest.Trikafta.Date.in.PortCF[chco_dates$MRN =="815274"] = "12/13/2019"
+chco_dates$Earliest.Trikafta.Date.in.PortCF[chco_dates$MRN =="1078408"] = "3/9/2020"
+chco_dates$Earliest.Trikafta.Date.in.PortCF[chco_dates$MRN =="899717"] = "2/14/2020"
+chco_dates$CFF.ID = as.character(chco_dates$CFF.ID)
+# A1C and OGTT
+chco_a1c = read.csv("./Data_Cleaned/chco_a1c_ogtt.csv")
+chco_a1c$CFF.ID = as.character(chco_a1c$CFF.ID)
+chco_pft = read.csv("./Data_Cleaned/chco_bmi_pft.csv")
+# Merge
+chco = left_join(chco_dates,chco_a1c,by = "CFF.ID")
+chco = full_join(chco,chco_pft,by = c("MRN", "Date"))
+# Diabetes only
+chco = chco %>% filter(!is.na(Diabetes)) %>% mutate(Location = "CHCO")
+# Montana
+montana = read.csv("./Data_Raw/Data From Montana/Copy of Paper patient population glycemic (+hofer).csv")
+demo_adult <- read.csv("./Data_Raw/Data From Montana/demo_adults.csv")
+demo_kids <- read.csv("./Data_Raw/Data From Montana/demo_kids.csv")
+montana
+```
+
+```{r cgm data,eval=FALSE}
+# IDs from Maxie
+ids = read.csv("./Data_Cleaned/final_ids.csv")
+# CGM
+chco_cgm = read.csv("./Data_Cleaned/chco_cgm.csv")
+chco_cgm$timepoint = sub(".*_","",chco_cgm$subject_id)
+chco_cgm$timepoint = factor(chco_cgm$timepoint)
+levels(chco_cgm$timepoint) = c("Post","Post","Pre","Pre")
+chco_cgm$subject_id = sub("_.*","",chco_cgm$subject_id)
+colnames(chco_cgm)[which(colnames(chco_cgm) == "subject_id")] = "CFF.ID"
+# First visit post if more than 1, and only those with pre and post
+chco_cgm = chco_cgm %>% group_by(CFF.ID,timepoint) %>%
+  slice(which.min(as.Date(date_cgm_placement, '%m/%d/%Y %H:%M')))
+
+# Put everything together
+chco = left_join(chco_cgm,chco_dates)
+
+chco = full_join(chco_dates,chco_a1c,by = "CFF.ID")
+chco = full_join(chco,chco_pft,by = c("MRN", "Date"))
+chco = full_join(chco,chco_cgm,by = c("CFF.ID", "Date"))
+# Filter and select columns
+chco = chco %>% filter(CFF.ID %in% ids$CFF.ID)
+```
+
+```{r montana data}
+
+```

### 2022-01-19 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
Update pre_post_trikafta.Rmd

--- 2022-01-11 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
@@
 library(tidyverse)
+library(lubridate)
 library(readxl)
+library(PAutilities)
 library(AGD)
+library(rspiro)
+library(anthro)
 library(knitr)
@@
 
-```{r chco data}
-# CGM
-chco_cgm = read.csv("./Data_Cleaned/chco_cgm.csv")
-chco_cgm$timepoint = sub(".*_","",chco_cgm$subject_id)
-chco_cgm$timepoint = factor(chco_cgm$timepoint)
-levels(chco_cgm$timepoint) = c("Post","Post","Pre","Pre")
-chco_cgm$subject_id = sub("_.*","",chco_cgm$subject_id)
-colnames(chco_cgm)[which(colnames(chco_cgm) == "subject_id")] = "CFF.ID"
+```{r code from maxie,eval=FALSE}
+# Source Maxie's code, and write final ID lists
+source("~/GitHub/CF/Christine Chan/Pre-post Trikafta/p001+cmw_UPDATED.R")
+ids = data.frame(unique(analytic_cm$CFF.ID))
+colnames(ids) = "CFF.ID"
+write.csv(ids,file = "./Data_Cleaned/final_ids.csv",row.names = F)
+```
+
+# Table 1: Demographics at Trikafta Start
+
+```{r eval=FALSE}
 # Start dates
-chco_dates = read.csv("./Data_Cleaned/chco_trikafta_dates.csv")
+chco_dates = read.csv("./Data_Cleaned/chco_trikafta_dates.csv",na.strings = "")
 # Fix select dates
@@
 chco_dates$Earliest.Trikafta.Date.in.PortCF[chco_dates$MRN =="899717"] = "2/14/2020"
+chco_dates$CFF.ID = as.character(chco_dates$CFF.ID)
 # A1C and OGTT
@@
 chco_pft = read.csv("./Data_Cleaned/chco_bmi_pft.csv")
-# Put everything together
-chco = left_join(chco_cgm,chco_a1c,by = "CFF.ID")
+# Merge
+chco = left_join(chco_dates,chco_a1c,by = "CFF.ID")
+chco = full_join(chco,chco_pft,by = c("MRN", "Date"))
+# Diabetes only
+chco = chco %>% filter(Diabetes == 1) %>% mutate(Location = "CHCO") %>%
+  select(CFF.ID,DOB,Earliest.Trikafta.Date.in.PortCF,
+         Sex,Race,Hispanic.Latinx,Genotypes1,Genotypes2,Pancreatic.Status,G.tube.in.past.12.months,
+         Date,HbA1c,OGTT_Fasting,OGTT_TwoHour,Height:Location) %>%
+  rename(Start = Earliest.Trikafta.Date.in.PortCF)
+# Montana
+montana = read.csv("./Data_Raw/Data From Montana/Copy of Paper patient population glycemic (+hofer).csv",na.strings = "")
+demo_adult = read.csv("./Data_Raw/Data From Montana/demo_adults.csv",na.strings = "")
+demo_kids = read.csv("./Data_Raw/Data From Montana/demo_kids.csv",na.strings = "")
+demo_kids$Date.Of.Birth..mm.dd.yyyy. = NULL
+demo_kids = demo_kids[,colnames(demo_adult)]
+demo = rbind(demo_adult,demo_kids)
+# Get within a year of Trikafta start
+montana$time_diff = mdy(montana$Trikafta.start.date) - mdy(montana$Date.of.Encounter)
+montana = montana %>% filter(abs(montana$time_diff)<=365)
+# Add demographics
+montana = left_join(montana,demo,by = "CFF.ID")
+# PFTs
+pft_adult <- read.csv("./Data_Raw/Data From Montana/Copy of PFTs growth adult (+hofer).csv")
+pft_kids <- read.csv("./Data_Raw/Data From Montana/PFTs kids.csv")
 ```
+
+```{r}
+
+```

### 2022-01-19 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
Update pre_post_trikafta.Rmd
just use Maxie's code

--- 2022-01-11 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
@@
 
-```{r chco data}
-# CGM
-chco_cgm = read.csv("./Data_Cleaned/chco_cgm.csv")
-chco_cgm$timepoint = sub(".*_","",chco_cgm$subject_id)
-chco_cgm$timepoint = factor(chco_cgm$timepoint)
-levels(chco_cgm$timepoint) = c("Post","Post","Pre","Pre")
-chco_cgm$subject_id = sub("_.*","",chco_cgm$subject_id)
-colnames(chco_cgm)[which(colnames(chco_cgm) == "subject_id")] = "CFF.ID"
-# Start dates
-chco_dates = read.csv("./Data_Cleaned/chco_trikafta_dates.csv")
-# Fix select dates
-chco_dates$Earliest.Trikafta.Date.in.PortCF[chco_dates$MRN =="1053930"] = "12/1/2019"
-chco_dates$Earliest.Trikafta.Date.in.PortCF[chco_dates$MRN =="1695512"] = "12/12/2019"
-chco_dates$Earliest.Trikafta.Date.in.PortCF[chco_dates$MRN =="815274"] = "12/13/2019"
-chco_dates$Earliest.Trikafta.Date.in.PortCF[chco_dates$MRN =="1078408"] = "3/9/2020"
-chco_dates$Earliest.Trikafta.Date.in.PortCF[chco_dates$MRN =="899717"] = "2/14/2020"
-# A1C and OGTT
-chco_a1c = read.csv("./Data_Cleaned/chco_a1c_ogtt.csv")
-chco_a1c$CFF.ID = as.character(chco_a1c$CFF.ID)
-chco_pft = read.csv("./Data_Cleaned/chco_bmi_pft.csv")
-# Put everything together
-chco = left_join(chco_cgm,chco_a1c,by = "CFF.ID")
+```{r}
+############pre/post trikaftka
+# c. chan
+##############################
+library(dplyr)
+setwd("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/")
+
+#cgm data
+# cgm_c <- read.csv("REDCap Upload.csv")
+cgm_c <- read.csv("Data From CHCO/REDCap Upload_chco.csv")
+cgm_c$date_cgm <- as.Date(cgm_c$date_cgm_placement, "%Y-%m-%d")
+
+cgm_m <- read.csv("Data From Montana/CGM_Montana/REDCap Upload_mt.csv")
+cgm_m$date_cgm <- as.Date(cgm_m$date_cgm_placement, "%m/%d/%Y")
+
+cgm_w <- read.csv("Data From Washington/REDCap Upload_Washington.csv")
+cgm_w$date_cgm <- as.Date(cgm_w$date_cgm_placement, "%Y-%m-%d")
+
+cgm <- rbind(cgm_c, cgm_m,cgm_w )
+
+`%!in%` = Negate(`%in%`)
+reg0 <- read.csv("Data From CHCO/Registry/Copy of Earliest Trikafta Dates in Registry_updated 2.2.21 cc modified_cfrdyn.csv")
+#need to get rid of those two
+reg <- subset(reg0, reg0$MRN %!in% c("1714343", "787372") )
+
+reg$Earliest.Trikafta.Date.in.PortCF <- ifelse(reg$MRN =="1053930", "12/1/2019", 
+                                               ifelse(reg$MRN =="1695512", "12/12/2019", 
+                                                      ifelse(reg$MRN =="815274", "12/13/2019", 
+                                                             ifelse(reg$MRN =="1078408",   "3/9/2020",
+                                       ifelse(reg$MRN =="899717", "2/14/2020", as.character(reg$Earliest.Trikafta.Date.in.PortCF))))))
+
+a1c <- read.csv("Data From CHCO/Copy of A1C and OGTTs Before and After Trikafta_cc modified 8.25.20.csv")
+
+weight <- read.csv("Data From CHCO/Registry/weightdata.csv")
+
+weight$Earliest.Trikafta.Date.in.PCF <- ifelse(weight$MRN =="1053930", "12/1/2019", 
+                                               ifelse(weight$MRN =="1695512", "12/12/2019", 
+                                                      ifelse(weight$MRN =="815274", "12/13/2019", 
+                                                             ifelse(weight$MRN =="1078408",   "3/9/2020",
+                                                                    ifelse(weight$MRN =="899717", "2/14/2020", as.character(weight$Earliest.Trikafta.Date.in.PCF))))))
+
+#montana data
+#total population
+
+library("readxl")
+
+reg_montana <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/Copy of Paper patient population glycemic (+hofer).csv")
+
+demo_adult <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/demo_adults.csv")
+demo_kids <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/demo_kids.csv")
+
+reg_montana$age <- ifelse(reg_montana$CFF.ID %in% demo_adult$CFF.ID, "Adult", 
+                          ifelse(reg_montana$CFF.ID %in% demo_kids$CFF.ID, "Kids", "Check"))
+
+reg_montana$trik <- as.Date(reg_montana$Trikafta.start.date,"%m/%d/%Y")
+
+reg_montana$test <- as.Date(reg_montana$Date.of.Encounter,  "%m/%d/%Y")
+reg_montana$prepost <- ifelse(reg_montana$trik > reg_montana$test, "before", "after")
+
+reg_montana$diff <- reg_montana$trik-reg_montana$test 
+reg_montana2 <- subset(reg_montana, abs(diff) <= 365)
+
+reg_montana3 <- reg_montana2 %>% group_by(CFF.ID, prepost) %>% count()
+reg_montana4 <- reg_montana3 %>% group_by(CFF.ID) %>% count()
+
+#weight/fev
+
+pft_adult <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/Copy of PFTs growth adult (+hofer).csv")
+pft_kids <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/PFTs kids.csv")
+
+pft_montana <- rbind(pft_adult, pft_kids)
+
+pft_montana_1 <- subset(pft_montana, pft_montana$CFF.ID %in% reg_montana$CFF.ID)
+
+#age at pft
+pft_montana_1$pft <- as.Date(pft_montana_1$Date.of.Encounter,  "%m/%d/%Y")
+
+reg_montana_first <- reg_montana %>% group_by(CFF.ID) %>% filter(row_number()==1)
+
+reg_montana_first$trik <- ifelse(reg_montana_first$CFF.ID == "1044621", "2019-12-02", 
+                                 ifelse(reg_montana_first$CFF.ID == "1008380", "2020-1-1", as.character(reg_montana_first$trik)))
+
+pft_montana_2 <- merge(pft_montana_1, reg_montana_first[,c("CFF.ID" ,"Date.Of.Birth..mm.dd.yyyy.", "trik")])
+
+#need pfts within a year of trik
+
+pft_montana_2$diff <-as.numeric(as.Date(pft_montana_2$pft) - as.Date(pft_montana_2$trik))
+pft_montana_3 <- subset(pft_montana_2, abs(diff) <= 365 )
+pft_montana_3$prepost <- ifelse(pft_montana_3$trik > pft_montana_3$pft, "before", "after")
+pft_montana_4 <- subset(pft_montana_3, pft_montana_3$prepost == "before")
+pft_montana_5 <- pft_montana_4 %>% group_by(CFF.ID) %>% arrange(diff) %>% filter(row_number()==1)
+
+pft_montana_5$DOB <- as.Date(pft_montana_5$Date.Of.Birth..mm.dd.yyyy.,  "%m/%d/%Y")
+
+pft_montana_5$age <- as.numeric(pft_montana_5$pft-pft_montana_5$DOB)/365.25
+
+pft_montana_5$BMI.Percentile
+
+#need sex
+demo_montana <- rbind(demo_kids[,c("CFF.ID","Gender")], demo_adult[,c("CFF.ID","Gender")])
+
+analytic_montana <- merge(pft_montana_5, demo_montana ,by = "CFF.ID", all.x=T)
+
+#need cfrd start date - don't have dates
+
+#gtube status
+
+gt_adult <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/gtubes_adult.csv")
+gt_kids <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/GT users kids.csv")
+
+gt_montana <- rbind(gt_adult, gt_kids)
+
+gt_montana_1 <- merge(gt_montana, reg_montana_first[,c("CFF.ID" , "trik")])
+gt_montana_1$gt <- as.Date(gt_montana_1$Date.of.Encounter,  "%m/%d/%Y")
+gt_montana_1$diff <-as.numeric(as.Date(gt_montana_1$gt) - as.Date(gt_montana_1$trik))
+gt_montana_2 <- subset(gt_montana_1, abs(diff) <= 365 )
+analytic_montana$gtube <- ifelse(analytic_montana$CFF.ID %in% gt_montana_2$CFF.ID, "Yes", "No")
+
+#Insufficent
+
+is_adult <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/pancreatic insufficient adult.csv")
+is_kids <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/pancreatic insufficient kids.csv")
+
+analytic_montana$is <- ifelse(analytic_montana$CFF.ID %in% c(is_adult$CFF.ID, is_kids$CFF.ID), "Insufficient", "Not")
+
+#genotypes
+
+g_adult <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/genotypes_adult.csv")
+g_kids <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/genotypes kids.csv")[,1:3]
+
+geno_montana <- rbind(g_adult,g_kids)
+
+geno_montana$Genotypes1 <- geno_montana$Name.of.the.first.mutation
+
+geno_montana$Genotypes2 <- geno_montana$Name.of.the.second.mutation
+
+geno_montana$class2 <- ifelse(geno_montana$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                                             '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                                             '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                                             '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
+                              ifelse(geno_montana$Genotypes2 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
+                                     "II", 
+                                     ifelse(geno_montana$Genotypes2 %in% c('G551D', 'G551S', 'S549N'), "III", 
+                                            ifelse(geno_montana$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
+                                                   "IV", 
+                                                   ifelse(geno_montana$Genotypes2 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                                                         'P574H', '3272-26A->G5'), "V" , "")))))
+
+geno_montana$severity2 <- ifelse(geno_montana$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                                                '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                                                '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                                                '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
+                                                                'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
+                                                                '2183delAA->G', '2184insA', 'S434X',
+                                                                '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
+                                                                'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
+                                                                "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
+                                 "Severe", 
+                                 ifelse(geno_montana$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
+                                                                       'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                                       'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
+                                                                       'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
+                                                                       "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
+
+                                        ifelse(geno_montana$Genotypes2 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
+                                                                                'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
+                                                                                'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
+                                                                                'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
+                                                                                "R1066H", "P205S"), 
+                                               "Unk","")))
+
+geno_montana$class1 <- ifelse(geno_montana$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                                             '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                                             '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                                             '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
+                              ifelse(geno_montana$Genotypes1 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
+                                     "II", 
+                                     ifelse(geno_montana$Genotypes1 %in% c('G551D', 'G551S', 'S549N'), "III", 
+                                            ifelse(geno_montana$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
+                                                   "IV", 
+                                                   ifelse(geno_montana$Genotypes1 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                                                         'P574H', '3272-26A->G5'), "V" , "")))))   
+geno_montana$severity1 <- ifelse(geno_montana$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                                                '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                                                '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                                                '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
+                                                                'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
+                                                                '2183delAA->G', '2184insA', 'S434X',
+                                                                '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
+                                                                'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
+                                                                "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
+                                 "Severe", 
+                                 ifelse(geno_montana$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
+                                                                       'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                                       'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
+                                                                       'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
+                                                                       "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
+
+                                        ifelse(geno_montana$Genotypes1 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
+                                                                                'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
+                                                                                'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
+                                                                                'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
+                                                                                "R1066H", "P205S"), 
+                                               "Unk","")))
+
+geno_montana$genoRisk <- ifelse(geno_montana$severity1 == "Mild" | geno_montana$severity2 == "Mild", "Low", 
+                                ifelse(geno_montana$severity1 == "Severe" & geno_montana$severity2 == "Severe", "High", ""))
+
+analytic_montana$genotype <- ifelse(analytic_montana$CFF.ID %in% geno_montana[geno_montana$genoRisk == "High", ]$CFF.ID, "High", 
+                                    ifelse(analytic_montana$CFF.ID %in% geno_montana[geno_montana$genoRisk == "Low", ]$CFF.ID, "Low",   ""   ))
+
+#washingotn
+
+reg_washington <- read.csv("Data From Washington/Copy of CFRD and trikafta Trikafta_on_Annual_Review-13-Apr-2021-0815.csv")
+reg_wash <- read.csv("Data From Washington/Copy of CFRD and trikafta Trikafta_on_Annual_Review-13-Apr-2021-0815.csv")
+
+analytic_wash <- reg_washington %>% group_by(CFF.ID) %>% filter(row_number()==1) %>% filter(CFF.ID != "1499770")
+analytic_wash$trik <- as.Date(analytic_wash$Approximate.date.patient.started.taking.this.drug,  "%m/%d/%Y")
+
+#aic 
+a1c_wash <- read.csv("Data From Washington/A1C_and_OGTT.csv")
+
+temp_wash <- reg_wash[,c("CFF.ID", "Approximate.date.patient.started.taking.this.drug")]
+temp_wash1 <- temp_wash %>% group_by(CFF.ID) %>% filter(row_number()==1)
+
+a1c_wash1 <- merge(a1c_wash, temp_wash1, by = c("CFF.ID"))
+a1c_wash1$pft <- as.Date(a1c_wash1$Date.of.Encounter,  "%m/%d/%Y")
+a1c_wash1$trik <- as.Date(a1c_wash1$Approximate.date.patient.started.taking.this.drug,  "%m/%d/%Y")
+a1c_wash1$diff <-as.numeric(a1c_wash1$pft - a1c_wash1$trik)
+a1c_wash2<- subset(a1c_wash1, abs(diff) <= 365 )
+
+#bmi
+bmi_wash <- read.csv("Data From Washington/BMIandWeight.csv")
+bmi_wash1 <- merge(bmi_wash, analytic_wash, by = "CFF.ID")
+bmi_wash1$Date <- as.Date(bmi_wash1$Date.of.Encounter.x,  "%m/%d/%Y")
+bmi_wash1$diff <-as.numeric(bmi_wash1$Date - bmi_wash1$trik)
+
+bmi_wash2 <- subset(bmi_wash1, abs(diff) <= 365 )
+bmi_wash2$prepost <- ifelse(bmi_wash2$trik > bmi_wash2$Date, "before", "after")
+
+#need age/bmi from first visit
+bmi_wash3 <- bmi_wash2 %>% group_by(CFF.ID) %>% filter(prepost == "before") %>% arrange(diff) %>%  filter(row_number()==1)
+bmi_wash3$first_encounter <- bmi_wash3$Date.of.Encounter.x
+
+analytic_wash <- merge(analytic_wash,bmi_wash3[,c("CFF.ID","BMI.Percentile" ,"Encounter.Age"  ,"Patients.height","BMI.Value","first_encounter")], by = "CFF.ID")
+
+#sex
+demo_wash <- read.csv("Data From Washington/Demographics.csv")
+
+analytic_wash$Sex <- ifelse(analytic_wash$CFF.ID %in% demo_wash[demo_wash$Gender== "Female",]$CFF.ID, "F", 
+                            ifelse(analytic_wash$CFF.ID %in% demo_wash[demo_wash$Gender== "Male",]$CFF.ID, "M",  "Check"))
+
+#Insufficent
+is_wash <- read.csv("Data From Washington/Pancreatic_Insufficient.csv")
+
+analytic_wash$is <- ifelse(analytic_wash$CFF.ID %in% is_wash$CFF.ID, "Insufficient", "Not")
+
+#gtube
+
+gt_wash <- read.csv("Data From Washington/G-tube.csv")
+
+analytic_wash$gt <- ifelse(analytic_wash$CFF.ID %in% gt_wash$CFF.ID, "GT", "Not")
+
+#genotypes
+
+geno_wash <- read.csv("Data From Washington/Genotypes.csv")
+
+geno_wash$Genotypes1 <- geno_wash$Name.of.the.first.mutation
+
+geno_wash$Genotypes2 <- geno_wash$Name.of.the.second.mutation
+
+#
+geno_wash$class2 <- ifelse(geno_wash$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                                       '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                                       '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                                       '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
+                           ifelse(geno_wash$Genotypes2 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
+                                  "II", 
+                                  ifelse(geno_wash$Genotypes2 %in% c('G551D', 'G551S', 'S549N'), "III", 
+                                         ifelse(geno_wash$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
+                                                "IV", 
+                                                ifelse(geno_wash$Genotypes2 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                                                   'P574H', '3272-26A->G5'), "V" , "")))))
+
+geno_wash$severity2 <- ifelse(geno_wash$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                                          '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                                          '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                                          '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
+                                                          'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
+                                                          '2183delAA->G', '2184insA', 'S434X',
+                                                          '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
+                                                          'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
+                                                          "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
+                              "Severe", 
+                              ifelse(geno_wash$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
+                                                                 'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                                 'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
+                                                                 'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
+                                                                 "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
+
+                                     ifelse(geno_wash$Genotypes2 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
+                                                                          'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
+                                                                          'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
+                                                                          'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
+                                                                          "R1066H", "P205S"), 
+                                            "Unk","")))
+
+geno_wash$class1 <- ifelse(geno_wash$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                                       '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                                       '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                                       '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
+                           ifelse(geno_wash$Genotypes1 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
+                                  "II", 
+                                  ifelse(geno_wash$Genotypes1 %in% c('G551D', 'G551S', 'S549N'), "III", 
+                                         ifelse(geno_wash$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
+                                                "IV", 
+                                                ifelse(geno_wash$Genotypes1 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                                                   'P574H', '3272-26A->G5'), "V" , "")))))   
+geno_wash$severity1 <- ifelse(geno_wash$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                                          '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                                          '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                                          '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
+                                                          'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
+                                                          '2183delAA->G', '2184insA', 'S434X',
+                                                          '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
+                                                          'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
+                                                          "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
+                              "Severe", 
+                              ifelse(geno_wash$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
+                                                                 'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                                 'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
+                                                                 'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
+                                                                 "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
+
+                                     ifelse(geno_wash$Genotypes1 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
+                                                                          'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
+                                                                          'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
+                                                                          'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
+                                                                          "R1066H", "P205S"), 
+                                            "Unk","")))
+
+geno_wash$genoRisk <- ifelse(geno_wash$severity1 == "Mild" | geno_wash$severity2 == "Mild", "Low", 
+                             ifelse(geno_wash$severity1 == "Severe" & geno_wash$severity2 == "Severe", "High", ""))
+
+analytic_wash$genotype <- ifelse(analytic_wash$CFF.ID %in% geno_wash[geno_wash$genoRisk == "High", ]$CFF.ID, "High", 
+                                    ifelse(analytic_wash$CFF.ID %in% geno_wash[geno_wash$genoRisk == "Low", ]$CFF.ID, "Low",   ""   ))
+
+#pfts
+
+pfts_wash <- read.csv("Data From Washington/PFTs.csv")
+
+pfts_wash1 <- merge(pfts_wash, analytic_wash, by = "CFF.ID")
+pfts_wash1$Date <- as.Date(pfts_wash1$Date.of.Encounter.x,  "%m/%d/%Y")
+pfts_wash1$diff <-as.numeric(pfts_wash1$Date - pfts_wash1$trik)
+
+pfts_wash2 <- subset(pfts_wash1, abs(diff) <= 365 )
+pfts_wash2$prepost <- ifelse(pfts_wash2$trik > pfts_wash2$Date, "before", "after")
+pfts_wash3 <- pfts_wash2 %>% group_by(CFF.ID) %>% filter(prepost == "before") %>% arrange(diff) %>%  filter(row_number()==1)
+
+analytic_wash <- merge(analytic_wash,pfts_wash3[,c("CFF.ID", "Predicted.Value.for.FVC", "Predicted.Value.for.FEV1")]  )
+
+library(arsenal)
+library(tidyverse)
+
+#getting analytic cohort
+reg1 <- subset(reg, reg$CFRD.yes.1 == 1)
+#there are 30 people with cfrd in the database
+
+a1c1 <- subset(a1c, a1c$Diabetes == 1)
+a1c2 <- subset(a1c1, a1c1$MRN %in% reg1$MRN)
+
+a1c2$trik <- as.Date(a1c2$Earliest.Trikafta.Date.in.PortCF, "%m/%d/%Y")
+
+a1c2$trik <- ifelse(a1c2$MRN =="1053930", "12/1/2019", 
+       ifelse(a1c2$MRN =="1695512", "12/12/2019", 
+               ifelse(a1c2$MRN =="815274", "12/13/2019", 
+                  ifelse(a1c2$MRN =="899717", "2/14/2020", as.character(a1c2$Earliest.Trikafta.Date.in.PortCF)))))
+
+a1c2$trik <- as.Date(a1c2$trik, "%m/%d/%Y")
+
+a1c2$test <- as.Date(a1c2$Date.of.A1C.or.OGTT,  "%m/%d/%Y")
+a1c2$prepost <- ifelse(a1c2$trik > a1c2$test, "before", "after")
+
+#who has before and after
+a1c3 <- a1c2 %>% group_by(MRN, prepost) %>% count()
+a1c4 <- a1c3 %>% group_by(MRN) %>% count()
+
+a1c5 <- subset(a1c2, a1c2$MRN %in% a1c4[a1c4$n == 2,]$MRN)
+
+#gcm cleaning
+
+out <- str_split(cgm$subject_id, "_")
+cgm$ID <- do.call(rbind, out)[,1]
+cgm$time <- tolower(do.call(rbind, out)[,2])
+
+cgm0 <- subset(cgm, cgm$ID %in% reg1$CFF.ID)
+
+cgm1 <- cgm0 %>% group_by(ID, time) %>% count()
+cgm2 <- cgm1 %>% group_by(ID) %>% count()
+
+#weight
+
+weight1 <- subset(weight, weight$MRN %in% reg1$MRN)
+
+reg1$a1c <- ifelse(reg1$MRN %in% a1c5$MRN, "Yes", "No")
+reg1$cgm <- ifelse(reg1$CFF.ID %in% cgm2$ID, "Yes", "No")
+
+######################
+#table1
+reg1$trek <-  as.Date(reg1$Earliest.Trikafta.Date.in.PortCF, "%m/%d/%Y")
+weight1$trek <- as.Date(weight1$Earliest.Trikafta.Date.in.PCF, "%m/%d/%Y")
+weight1$pft <- as.Date(weight1$PFT.Date, "%m/%d/%Y")
+weight1$diff <- weight1$pft-weight1$trek
+
+weight2 <- weight1  %>% arrange(as.numeric(diff))%>% group_by(MRN) %>%  filter(row_number()==1)
+
+reg2 <- merge(reg1, weight2, by = "MRN", all.x = T)
+
+reg2$age <- as.numeric(reg2$pft - as.Date(reg2$DOB, "%m/%d/%Y"))/365.25
+library(PAutilities)
+reg2$Sex_MM <- ifelse(reg2$Sex == "Male", "M", "F")
+
+library(AGD)
+
+bmi_levels <- c("Underweight", "Healthy weight", "Overweight", 
+                "Obesity", "Severe obesity")
+bmi_class <- c("Underweight", "Healthy weight", "Overweight","Class 1", "Class 2", "Class 3")
+reg2$agemos <- reg2$age*12
+
+reg2$Sex <- as.factor(reg2$Sex_MM)
+levels(reg2$Sex) <- c("F", "M")
+reg2$bmi <- reg2$BMI
+reg3 <- reg2 %>%
+  # the CDC growth charts are typically used for children 2 years or older
+  # filter(agemos >= 24) %>%
+  mutate(bmiz =  AGD::y2z(y = bmi, x = age, sex = Sex, ref = cdc.bmi),
+         bmipct = pnorm(bmiz),
+         z = qnorm(.95),
+         p95 = AGD::z2y(z = z, x= age, sex = Sex, ref = cdc.bmi),
+         # percentage of the 95th percentile
+         bmipct95 = bmi/p95,
+         # bmi category
+         bmicat = factor(case_when(bmipct < .05 ~ "Underweight",
+                                   bmipct < .85 ~ "Healthy weight",
+                                   bmipct < .95 ~ "Overweight",
+                                   bmipct95 < 1.2 ~ "Class 1",
+                                   bmipct95 < 1.4 ~ "Class 2",
+                                   bmipct95 >= 1.4 ~ "Class 3"),
+                         levels = bmi_class))
+
+#fev
+
+#BMI 
+reg2_BMI <- merge(reg1, weight1, by = "MRN", all.x = T)
+
+reg2_BMI$age <- as.numeric(reg2_BMI$pft - as.Date(reg2_BMI$DOB, "%m/%d/%Y"))/365.25
+library(PAutilities)
+reg2_BMI$Sex_MM <- ifelse(reg2_BMI$Sex == "Male", "M", "F")
+
+library(AGD)
+
+bmi_levels <- c("Underweight", "Healthy weight", "Overweight", 
+                "Obesity", "Severe obesity")
+bmi_class <- c("Underweight", "Healthy weight", "Overweight","Class 1", "Class 2", "Class 3")
+reg2_BMI$agemos <- reg2_BMI$age*12
+
+#
+reg2$class2 <- ifelse(reg2$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                             '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                             '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                             '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
+                      ifelse(reg2$Genotypes2 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
+                             "II", 
+                             ifelse(reg2$Genotypes2 %in% c('G551D', 'G551S', 'S549N'), "III", 
+                                    ifelse(reg2$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
+                                           "IV", 
+                                           ifelse(reg2$Genotypes2 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                                         'P574H', '3272-26A->G5'), "V" , "")))))
+
+reg2$severity2 <- ifelse(reg2$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                                '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                                '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                                '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
+                                                'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
+                                                '2183delAA->G', '2184insA', 'S434X',
+                                                '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
+                                                'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
+                                                "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
+                         "Severe", 
+                         ifelse(reg2$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
+                                                       'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                       'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
+                                                       'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
+                                                       "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
+
+                                ifelse(reg2$Genotypes2 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
+                                                                'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
+                                                                'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
+                                                                'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
+                                                                "R1066H", "P205S"), 
+                                       "Unk","")))
+
+reg2$class1 <- ifelse(reg2$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                             '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                             '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                             '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
+                      ifelse(reg2$Genotypes1 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
+                             "II", 
+                             ifelse(reg2$Genotypes1 %in% c('G551D', 'G551S', 'S549N'), "III", 
+                                    ifelse(reg2$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
+                                           "IV", 
+                                           ifelse(reg2$Genotypes1 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                                         'P574H', '3272-26A->G5'), "V" , "")))))   
+reg2$severity1 <- ifelse(reg2$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                                '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                                '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                                '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
+                                                'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
+                                                '2183delAA->G', '2184insA', 'S434X',
+                                                '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
+                                                'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
+                                                "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
+                         "Severe", 
+                         ifelse(reg2$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
+                                                       'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                       'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
+                                                       'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
+                                                       "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
+
+                                ifelse(reg2$Genotypes1 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
+                                                                'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
+                                                                'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
+                                                                'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
+                                                                "R1066H", "P205S"), 
+                                       "Unk","")))
+
+reg2$genoRisk <- ifelse(reg2$severity1 == "Mild" | reg2$severity2 == "Mild", "Low", 
+                        ifelse(reg2$severity1 == "Severe" & reg2$severity2 == "Severe", "High", ""))
+
+reg2_BMI$Sex <- as.factor(reg2_BMI$Sex_MM)
+levels(reg2_BMI$Sex) <- c("F", "M")
+reg2_BMI$bmi <- reg2_BMI$BMI
+
+reg2$ppFEV1 <- pctpred_GLI(reg2$age, reg2$Height/100, gender = ifelse(reg2$Sex_MM == "M", 1, 2) , 
+                           ethnicity = 1, 
+                           FEV1 = reg2$FEV1)
+
+reg2$ppFVC <- pctpred_GLI(reg2$age, reg2$Height/100, gender = ifelse(reg2$Sex_MM == "M", 1, 2) , 
+                          ethnicity = 1, 
+                          FVC = reg2$FVC)
+
+reg3_BMI <- reg2_BMI %>%
+  # the CDC growth charts are typically used for children 2 years or older
+  # filter(agemos >= 24) %>%
+  mutate(bmiz =  AGD::y2z(y = bmi, x = age, sex = Sex, ref = cdc.bmi),
+         bmipct = pnorm(bmiz),
+         z = qnorm(.95),
+         p95 = AGD::z2y(z = z, x= age, sex = Sex, ref = cdc.bmi),
+         # percentage of the 95th percentile
+         bmipct95 = bmi/p95,
+         # bmi category
+         bmicat = factor(case_when(bmipct < .05 ~ "Underweight",
+                                   bmipct < .85 ~ "Healthy weight",
+                                   bmipct < .95 ~ "Overweight",
+                                   bmipct95 < 1.2 ~ "Class 1",
+                                   bmipct95 < 1.4 ~ "Class 2",
+                                   bmipct95 >= 1.4 ~ "Class 3"),
+                         levels = bmi_class))
+
+#fev
+
+library(rspiro)
+
+reg3$ppFEV1 <- pctpred_GLI(reg3$age, reg3$Height/100, gender = ifelse(reg3$Sex_MM == "M", 1, 2) , 
+                           ethnicity = 1, 
+                           FEV1 = reg3$FEV1)
+
+reg3$ppFVC <- pctpred_GLI(reg3$age, reg3$Height/100, gender = ifelse(reg3$Sex_MM == "M", 1, 2) , 
+                           ethnicity = 1, 
+                           FVC = reg3$FVC)
+
+#
+reg3$CFRD <- as.Date(reg3$CFRD.Diagnosis.Date, "%m/%d/%Y")
+reg3$CFRD_duration <- as.numeric(reg3$trek.x-reg3$CFRD)/365.25
+
+#
+reg3$class2 <- ifelse(reg3$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                             '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                             '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                             '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
+                      ifelse(reg3$Genotypes2 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
+                             "II", 
+                             ifelse(reg3$Genotypes2 %in% c('G551D', 'G551S', 'S549N'), "III", 
+                                    ifelse(reg3$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
+                                           "IV", 
+                                           ifelse(reg3$Genotypes2 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                                         'P574H', '3272-26A->G5'), "V" , "")))))
+
+reg3$severity2 <- ifelse(reg3$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                                '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                                '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                                '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
+                                                'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
+                                                '2183delAA->G', '2184insA', 'S434X',
+                                                '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
+                                                'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
+                                                "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
+                         "Severe", 
+                         ifelse(reg3$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
+                                                       'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                       'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
+                                                       'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
+                                                       "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
+
+                                ifelse(reg3$Genotypes2 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
+                                                                'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
+                                                                'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
+                                                                'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
+                                                                "R1066H", "P205S"), 
+                                       "Unk","")))
+
+reg3$class1 <- ifelse(reg3$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                             '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                             '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                             '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
+                      ifelse(reg3$Genotypes1 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
+                             "II", 
+                             ifelse(reg3$Genotypes1 %in% c('G551D', 'G551S', 'S549N'), "III", 
+                                    ifelse(reg3$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
+                                           "IV", 
+                                           ifelse(reg3$Genotypes1 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                                         'P574H', '3272-26A->G5'), "V" , "")))))   
+reg3$severity1 <- ifelse(reg3$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                                '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                                '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                                '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
+                                                'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
+                                                '2183delAA->G', '2184insA', 'S434X',
+                                                '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
+                                                'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
+                                                "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
+                         "Severe", 
+                         ifelse(reg3$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
+                                                       'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                       'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
+                                                       'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
+                                                       "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
+
+                                ifelse(reg3$Genotypes1 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
+                                                                'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
+                                                                'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
+                                                                'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
+                                                                "R1066H", "P205S"), 
+                                       "Unk","")))
+
+reg3$genoRisk <- ifelse(reg3$severity1 == "Mild" | reg3$severity2 == "Mild", "Low", 
+                        ifelse(reg3$severity1 == "Severe" & reg3$severity2 == "Severe", "High", ""))
+
+# 
+# tab1 <- tableby(~age+bmipct + Weight + Sex + ppFEV1 + ppFVC + CFRD_duration +
+#                   G.tube.in.past.12.months + Pancreatic.Status +
+#                   genoRisk , data=reg3)
+
+reg3$location <- rep("reg3", nrow(reg3))
+
+reg_tab1 <- reg3[,c("CFF.ID", "age", "bmipct", "Weight", "Sex", "ppFEV1", "ppFVC", "G.tube.in.past.12.months", "Pancreatic.Status", "genoRisk",
+                    "location")]
+
+reg_tab1$BMI.Value <- rep(NA, nrow(reg_tab1))
+reg_tab1$age_category <- rep("Pediatric", nrow(reg_tab1))
+
+analytic_montana$bmipct <- analytic_montana$BMI.Percentile/100
+analytic_montana$Sex <- ifelse(analytic_montana$Gender == "Female", "F", "M")
+analytic_montana$Weight <- analytic_montana$Weight.of.Patient 
+analytic_montana$ppFEV1 <- analytic_montana$Percent.predicted.relative.to.predicted.FEV1..value
+analytic_montana$ppFVC <- analytic_montana$Percent.predicted.relative.to.predicted.FVC.value
+analytic_montana$G.tube.in.past.12.months <- analytic_montana$gtube 
+analytic_montana$Pancreatic.Status <- analytic_montana$is
+analytic_montana$genoRisk <- analytic_montana$genotype
+
+analytic_montana$location <- rep("Montana", nrow(analytic_montana))
+
+analytic_montana$age_category <- ifelse(analytic_montana$CFF.ID %in% demo_kids$CFF.ID, "Pediatric", "Adult")
+
+analytic_montana1 <- analytic_montana[,c("CFF.ID", "age", "bmipct", "Weight", "Sex", "ppFEV1", "ppFVC", "G.tube.in.past.12.months", "Pancreatic.Status", "genoRisk",
+                                         "location", "BMI.Value", "age_category")]
+
+analytic_wash$bmipct <- analytic_wash$BMI.Percentile/100
+analytic_wash$age <- analytic_wash$Encounter.Age
+analytic_wash$BMI.Value <- rep(NA, nrow(analytic_wash))
+analytic_wash$Weight <- rep(NA, nrow(analytic_wash))
+analytic_wash$ppFEV1 <- analytic_wash$Predicted.Value.for.FEV1
+analytic_wash$ppFVC <- analytic_wash$Predicted.Value.for.FVC
+analytic_wash$G.tube.in.past.12.months <- analytic_wash$gt
+analytic_wash$Pancreatic.Status <- analytic_wash$is
+analytic_wash$genoRisk <- analytic_wash$genotype
+analytic_wash$location <- rep("Washington", nrow(analytic_wash))
+analytic_wash$age_category <- rep("Pediatric", nrow(analytic_wash))
+analytic_wash$Earliest.Trikafta.Date.in.PortCF <-analytic_wash$trik
+
+analytic_wash1 <- analytic_wash[,c("CFF.ID", "age", "bmipct", "Weight", "Sex", "ppFEV1", "ppFVC", "G.tube.in.past.12.months", 
+                                   "Pancreatic.Status", "genoRisk",
+                                         "location", "BMI.Value", "age_category")]
+
+#
+
+uofm <- read.csv("Data from UofM/Moheet_Pre_Post_Trikafta_FinalData_20210914.csv")
+analytic_uofm <- subset(uofm, uofm$CFRD.Status == "CFRD with or without fasting hyperglycemia")
+analytic_uofm$age <- analytic_uofm$age.at.Trikafta.start
+
+bmi_levels <- c("Underweight", "Healthy weight", "Overweight", 
+                "Obesity", "Severe obesity")
+bmi_class <- c("Underweight", "Healthy weight", "Overweight","Class 1", "Class 2", "Class 3")
+analytic_uofm$agemos <- analytic_uofm$age*12
+
+analytic_uofm$Sex <- as.factor(analytic_uofm$Gender)
+levels(analytic_uofm$Sex) <- c("F", "M")
+analytic_uofm$bmi <- analytic_uofm$BMI.calc
+analytic_uofm$bmi_post <- analytic_uofm$BMI.calc.1
+analytic_uofm2 <- analytic_uofm %>%
+  # the CDC growth charts are typically used for children 2 years or older
+  # filter(agemos >= 24) %>%
+  mutate(bmiz =  AGD::y2z(y = bmi, x = age, sex = Sex, ref = cdc.bmi),
+         bmiz_post =  AGD::y2z(y = bmi_post, x = age, sex = Sex, ref = cdc.bmi),
+         bmipct = pnorm(bmiz),
+         bmipct_post = pnorm(bmiz_post))
+
+analytic_uofm2$ppFEV1 <- pctpred_GLI(analytic_uofm2$age, analytic_uofm2$Height..cm./100, gender = ifelse(analytic_uofm2$Sex== "M", 1, 2) , 
+                                     ethnicity = 1, 
+                                     FEV1 = analytic_uofm2$FEV1)
+
+analytic_uofm2$ppFEV1_post <- pctpred_GLI(analytic_uofm2$age, as.numeric(analytic_uofm2$Height..cm..1)/100, gender = ifelse(analytic_uofm2$Sex== "M", 1, 2) , 
+                                     ethnicity = 1, 
+                                     FEV1 = analytic_uofm2$FEV1.1)
+
+analytic_uofm2$ppFVC <- pctpred_GLI(analytic_uofm2$age, analytic_uofm2$Height..cm./100, gender = ifelse(analytic_uofm2$Sex == "M", 1, 2) , 
+                                    ethnicity = 1, 
+                                    FVC = analytic_uofm2$FVC)
+
+analytic_uofm2$ppFVC_post <- pctpred_GLI(analytic_uofm2$age, as.numeric(analytic_uofm2$Height..cm..1)/100, gender = ifelse(analytic_uofm2$Sex== "M", 1, 2) , 
+                                          ethnicity = 1, 
+                                          FEV1 = analytic_uofm2$FVC.1)
+
+analytic_uofm2$Weight <- analytic_uofm2$Weight..kg.
+
+analytic_uofm2$G.tube.in.past.12.months <- analytic_uofm2$Tube.Feeding
+
+analytic_uofm2$age_category <- ifelse(analytic_uofm2$age <= 20 ,"Pediatric", "Adult")
+analytic_uofm2$location <- rep("Minnesota", nrow(analytic_uofm2))
+analytic_uofm2$Pancreatic.Status <- ifelse(analytic_uofm2$Pancreatic.Sufficiency..on.enzymes.. == "Yes", "Insufficient", "Not")
+
+analytic_uofm2$Genotypes1 <- analytic_uofm2$Genotype1
+
+analytic_uofm2$Genotypes2 <- analytic_uofm2$Genotype2
+
+analytic_uofm2$class2 <- ifelse(analytic_uofm2$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                                                 '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                                                 '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                                                 '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G', '2183AA->G'), "I", 
+                                ifelse(analytic_uofm2$Genotypes2 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
+                                       "II", 
+                                       ifelse(analytic_uofm2$Genotypes2 %in% c('G551D', 'G551S', 'S549N'), "III", 
+                                              ifelse(analytic_uofm2$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
+                                                     "IV", 
+                                                     ifelse(analytic_uofm2$Genotypes2 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                                                             'P574H', '3272-26A->G5'), "V" , "")))))
+
+analytic_uofm2$severity2 <- ifelse(analytic_uofm2$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                                                    '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                                                    '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                                                    '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
+                                                                    'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
+                                                                    '2183delAA->G', '2184insA', 'S434X',
+                                                                    '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
+                                                                    'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
+                                                                    "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
+                                   "Severe", 
+                                   ifelse(analytic_uofm2$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
+                                                                           'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                                           'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
+                                                                           'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
+                                                                           "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
+
+                                          ifelse(analytic_uofm2$Genotypes2 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
+                                                                                    'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
+                                                                                    'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
+                                                                                    'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
+                                                                                    "R1066H", "P205S"), 
+                                                 "Unk","")))
+
+analytic_uofm2$class1 <- ifelse(analytic_uofm2$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                                       '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                                       '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                                       '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
+                           ifelse(analytic_uofm2$Genotypes1 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
+                                  "II", 
+                                  ifelse(analytic_uofm2$Genotypes1 %in% c('G551D', 'G551S', 'S549N'), "III", 
+                                         ifelse(analytic_uofm2$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
+                                                "IV", 
+                                                ifelse(analytic_uofm2$Genotypes1 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                                                   'P574H', '3272-26A->G5'), "V" , "")))))   
+analytic_uofm2$severity1 <- ifelse(analytic_uofm2$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
+                                                          '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
+                                                          '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
+                                                          '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
+                                                          'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
+                                                          '2183delAA->G', '2184insA', 'S434X',
+                                                          '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
+                                                          'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
+                                                          "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
+                              "Severe", 
+                              ifelse(analytic_uofm2$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
+                                                                 'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
+                                                                 'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
+                                                                 'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
+                                                                 "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
+
+                                     ifelse(analytic_uofm2$Genotypes1 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
+                                                                          'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
+                                                                          'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
+                                                                          'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
+                                                                          "R1066H", "P205S"), 
+                                            "Unk","")))
+
+analytic_uofm2$genoRisk <- ifelse(analytic_uofm2$severity1 == "Mild" | analytic_uofm2$severity2 == "Mild", "Low", 
+                             ifelse(analytic_uofm2$severity1 == "Severe" & analytic_uofm2$severity2 == "Severe", "High", ""))
+
+analytic_uofm2$genotype <- ifelse(analytic_uofm2$CFF.ID %in% analytic_uofm2[analytic_uofm2$genoRisk == "High", ]$CFF.ID, "High", 
+                                 ifelse(analytic_uofm2$CFF.ID %in% analytic_uofm2[analytic_uofm2$genoRisk == "Low", ]$CFF.ID, "Low",   ""   ))
+analytic_uofm2$bmipct <- ifelse(analytic_uofm2$age <=  20, analytic_uofm2$bmipct , NA)
+
+analytic_uofm2$BMI.Value <- analytic_uofm2$bmi
+analytic_uofm2$age_category
+analytic_uofm2_2 <- subset(analytic_uofm2, analytic_uofm2$Encounter....of.Days.POST.Trikafta.start <= 365 &
+                             analytic_uofm2$Encounter....of.Days.PRE.Trikafta.Start <= 365)
+
+analytic_uofm3 <- analytic_uofm2_2[,c("CFF.ID", "age", "bmipct", "Weight", "Sex", "ppFEV1", "ppFVC", "G.tube.in.past.12.months", 
+                                   "Pancreatic.Status", "genoRisk",
+                                   "location", "BMI.Value", "age_category")]
+
+analytic_cm <- rbind(reg_tab1 , analytic_montana1,analytic_wash1,analytic_uofm3 )
+
+tab1 <- tableby(~age+bmipct + Weight + Sex + ppFEV1 + ppFVC  +
+                  G.tube.in.past.12.months + Pancreatic.Status +
+                  genoRisk + location+BMI.Value + age_category, data=analytic_cm)
+
+#table 2
+
+#need a1c
+
+a1c_wash2
+
+reg_montana2
+
+#subsetfor u of m
+
+pre_minn <- analytic_uofm2_2[,c("CFF.ID", "A1C....of.Days.PRE.Trikafta.Start" , "A1C")]
+
+pre_minn$diff <- as.numeric(pre_minn$A1C....of.Days.PRE.Trikafta.Start)
+pre_minn$prepost <- rep("before", nrow(pre_minn))
+pre_minn$Hgb.A1C <- pre_minn$A1C
+pre_minn$trik <- rep(NA, nrow(pre_minn))
+pre_minn$test <- rep(NA, nrow(pre_minn))
+
+pre_minn1 <- pre_minn[,c("CFF.ID", "trik", "test", "prepost", "diff", "Hgb.A1C")]
+
+post_minn <- analytic_uofm2_2[,c("CFF.ID", "A1C....of.Days.POST.Trikafta.start", "A1C.1" )]
+post_minn$diff <- as.numeric(post_minn$A1C....of.Days.POST.Trikafta.start)
+post_minn$prepost <- rep("after", nrow(post_minn))
+post_minn$Hgb.A1C <- post_minn$A1C.1
+post_minn$trik <- rep(NA, nrow(post_minn))
+post_minn$test <- rep(NA, nrow(post_minn))
+post_minn1 <- post_minn[,c("CFF.ID", "trik", "test", "prepost", "diff", "Hgb.A1C")]
+
+minn <- subset(rbind(post_minn1, pre_minn1), abs(rbind(post_minn1, pre_minn1)$diff) <= 365)
+
+a1c5$diff <- as.numeric(a1c5$trik - a1c5$test)
+a1c6 <- subset(a1c5, abs(a1c5$diff) <= 365)
+
+#combining all into one
+
+a1c6_big <- a1c6[,c("CFF.ID", "trik", "test", "prepost", "diff", "Hgb.A1C")]
+reg_montana2$Hgb.A1C <- reg_montana2$Hgb.A1C.value...
+reg_montana2$trik <- ifelse(reg_montana2$CFF.ID == "1044621", "2019-12-02",
+                            ifelse(reg_montana2$CFF.ID == "1008380", "2020-1-1", as.character(reg_montana2$trik)))
+reg_montana2_big <- reg_montana2[,c("CFF.ID", "trik", "test", "prepost", "diff", "Hgb.A1C")]
+
+a1c_wash2$test <- a1c_wash2$pft
+a1c_wash2$Hgb.A1C <- a1c_wash2$Hgb.A1C.value...
+a1c_wash2$prepost <- ifelse(a1c_wash2$trik < a1c_wash2$test, "before", "after")
+a1c_wash2_big <- a1c_wash2[,c("CFF.ID", "trik", "test", "prepost", "diff", "Hgb.A1C")]
+
+a1c6_big2 <- rbind(a1c6_big,reg_montana2_big, a1c_wash2_big, minn)
+
+a1c7 <- a1c6_big2  %>% group_by(CFF.ID, prepost) %>%  count()
+
+mycontrols  <- tableby.control(test=FALSE, total=FALSE,
+                               numeric.test="kwt", cat.test="chisq",
+                               numeric.stats=c("N", "median", "q1q3"),
+                               cat.stats=c("countpct"),
+                               stats.labels=list(N='Count', median='Median', q1q3='Q1,Q3'))
+
+tab2 <- tableby(prepost ~ n, control=mycontrols, data=a1c7)
+
+before <- subset(a1c7, a1c7$prepost == "before")
+before$pre <- before$n
+
+after <- subset(a1c7, a1c7$prepost == "after")
+after$post <- after$n
+
+#number of values pre/post
+#row 1 in table 2
+a <- merge(before, after, by = "CFF.ID", all.y = T)
+wilcox.test(a$pre,a$post, paired = T)
+
+#test if difference
+
+#merge back in the a1c for pre and post 
+before <- subset(a1c6_big2, a1c6_big2$prepost == "before")
+before1 <- before %>% arrange(diff) %>% group_by(CFF.ID) %>% filter(row_number()==1)
+
+#time before
+before_max <- before %>% group_by(CFF.ID) %>% summarize(time = max(diff)/30.417)
+
+summary(before_max$time)
+sd(before_max$time)
+
+after <- subset(a1c6_big2, a1c6_big2$prepost == "after")
+after1 <- after %>% arrange(desc(diff)) %>% group_by(CFF.ID) %>% filter(row_number()==1)
+
+after_max <- after %>% group_by(CFF.ID) %>% summarize(time = max(abs(diff))/30.417)
+
+summary(after_max$time)
+sd(after_max$time)
+
+#test this
+t.test(before_max$time,after_max$time)
+
+time <- merge(before_max,after_max,by='CFF.ID',all.y=T)
+summary(time$time.x)
+sd(time$time.x, na.rm=T)
+summary(time[!is.na(time$time.x),]$time.y)
+sd(time[!is.na(time$time.x),]$time.y, na.rm=T)
+t.test(time$time.x,time$time.y, paired = T)
+
+time <- merge(before1,after1,by='CFF.ID',all.y=T)
+
+t.test(as.numeric(time$Hgb.A1C.x),as.numeric(time$Hgb.A1C.y), paired = T)
+
+summary(as.numeric(time$Hgb.A1C.x))
+sd(as.numeric(time$Hgb.A1C.x), na.rm=T)
+
+summary(as.numeric(time$Hgb.A1C.y))
+sd(as.numeric(time$Hgb.A1C.y),na.rm=T)
+
+temp <- rbind(before1,after1)
+temp$Hgb.A1C <- as.numeric(temp$Hgb.A1C)
+# tab3 <- tableby(prepost~Hgb.A1C, temp , control=mycontrols)
+
+time$diff_a1c <- as.numeric(time$Hgb.A1C.x) - as.numeric(time$Hgb.A1C.y)
+summary(time[time$Hgb.A1C.x < 6.5, ]$diff_a1c)
+
+t.test(time[time$Hgb.A1C.x < 6.5, ]$diff_a1c, mu=0)
+
+#use these numbers for pre/post cgm
+
+time2 <- subset(time, time$CFF.ID %in% cgm$ID)
+
+wilcox.test(as.numeric(time2$Hgb.A1C.x), as.numeric(time2$Hgb.A1C.y), paired = T)
+
+#DEMOGRAPHCIS
+# tab1 <- tableby(~age+bmipct + Weight + Sex + ppFEV1 + ppFVC  +
+#                   G.tube.in.past.12.months + Pancreatic.Status +
+#                   genoRisk + location+BMI.Value + age_category, data=analytic_cm[analytic_cm$CFF.ID %in% cgm0_small$CFF.ID,])
+#
+
+#insulin
+
+ins <- read.csv("insulindata.csv")
+
+ins2 <- merge(ins, time2[,c("CFF.ID", "trik.x")], by= "CFF.ID", all.x = T)
+
+ins2$Date_MM <- as.Date(ins2$Date, "%m/%d/%Y")
+
+ins2$diff <- as.numeric(ins2$Date_MM - ins2$trik.x)
+ins3 <- subset(ins2, abs(diff) <= 365)
+
+ins3$prepost <- ifelse(ins3$diff <= 0, "Pre", "Post")
+
+ins4 <- ins3 %>% group_by(CFF.ID, prepost) %>% count()
+ins5 <- ins4 %>% group_by(CFF.ID) %>% count()
+
+ins6 <- subset(ins3, ins3$CFF.ID %in% ins5[ins5$n > 1,]$CFF.ID)
+
+ins7 <- ins6 %>% group_by(CFF.ID, prepost) %>% summarise_at("Long.acting.or.total.basal.daily.dose..units.",mean)
+
+summary(ins7[ins7$prepost == "Post",]$Long.acting.or.total.basal.daily.dose..units.)
+
+ins8 <- merge(ins7[ins7$prepost == "Pre",], ins7[ins7$prepost == "Post",], by = "CFF.ID")
+
+wilcox.test(ins8$Long.acting.or.total.basal.daily.dose..units..x, ins8$Long.acting.or.total.basal.daily.dose..units..y, paired = T)
+
+##########################
+#weight data
+#need z score for weight
+library("anthro")
+weightz <- merge(weight1, reg1[,c("MRN", "CFF.ID", "Sex", "DOB")], by = "MRN", all.x=T)
+
+weightz$age <- as.numeric(weightz$pft - as.Date(weightz$DOB, "%m/%d/%Y"))/365.25
+library(PAutilities)
+weightz$Sex_MM <- ifelse(weightz$Sex == "Male", "M", "F")
+
+library(AGD)
+
+bmi_levels <- c("Underweight", "Healthy weight", "Overweight", 
+                "Obesity", "Severe obesity")
+bmi_class <- c("Underweight", "Healthy weight", "Overweight","Class 1", "Class 2", "Class 3")
+weightz$agemos <- weightz$age*12
+
+weightz$Sex <- as.factor(weightz$Sex_MM)
+levels(weightz$Sex) <- c("F", "M")
+weightz$bmi <- weightz$BMI
+weightz2 <- weightz %>%
+  # the CDC growth charts are typically used for children 2 years or older
+  # filter(agemos >= 24) %>%
+  mutate(bmiz =  AGD::y2z(y = bmi, x = age, sex = Sex, ref = cdc.bmi),
+         bmipct = pnorm(bmiz),
+         z = qnorm(.95),
+         p95 = AGD::z2y(z = z, x= age, sex = Sex, ref = cdc.bmi),
+         # percentage of the 95th percentile
+         bmipct95 = bmi/p95,
+         # bmi category
+         bmicat = factor(case_when(bmipct < .05 ~ "Underweight",
+                                   bmipct < .85 ~ "Healthy weight",
+                                   bmipct < .95 ~ "Overweight",
+                                   bmipct95 < 1.2 ~ "Class 1",
+                                   bmipct95 < 1.4 ~ "Class 2",
+                                   bmipct95 >= 1.4 ~ "Class 3"),
+                         levels = bmi_class))
+
+weightz2$ppFEV1 <- pctpred_GLI(weightz2$age, weightz2$Height/100, gender = ifelse(weightz2$Sex_MM == "M", 1, 2) , 
+                           ethnicity = 1, 
+                           FEV1 = weightz2$FEV1)
+
+weightz2$ppFVC <- pctpred_GLI(weightz2$age, weightz2$Height/100, gender = ifelse(weightz2$Sex_MM == "M", 1, 2) , 
+                          ethnicity = 1, 
+                          FVC = weightz2$FVC)
+
+weigh_before <- merge(weightz2, before1[,c("CFF.ID", "test")], by = "CFF.ID", all.x = T)
+weigh_before$weight_closest <- abs(as.numeric(weigh_before$pft-weigh_before$test))
+
+weigh_before1 <- weigh_before %>% group_by(MRN) %>% arrange(weight_closest) %>% filter(row_number()==1)
+weigh_before1$time <- rep("before",nrow(weigh_before1))
+
+weigh_after <- merge(weightz2, after1[,c("CFF.ID", "test")], by = "CFF.ID", all.x = T)
+weigh_after$weight_closest <- abs(as.numeric(weigh_after$pft-weigh_after$test))
+
+weigh_after1 <- weigh_after %>% group_by(MRN) %>% arrange(weight_closest) %>% filter(row_number()==1)
+
+weigh_after1$time <- rep("after",nrow(weigh_after1))
+
+timing <- rbind(weigh_before1,weigh_after1)
+
+tab2 <- tableby(time~Weight+ bmiz + ppFEV1+ppFVC, data=timing)
+
+timing2 <- merge(weigh_before1,weigh_after1, by = "MRN", all.y=T)
+wilcox.test(timing2$ppFEV1.x,timing2$ppFEV1.y, paired=T)
+
+wilcox.test(timing2$ppFVC.x,timing2$ppFVC.y, paired=T)
+
+#weight and fev data
+pfts_wash2$ppFVC <- pfts_wash2$Predicted.Value.for.FVC
+pfts_wash2$ppFEV1 <- pfts_wash2$Predicted.Value.for.FEV1
+pfts_wash2$bmipct <- pfts_wash2$BMI.Percentile
+pfts_wash2$Weight <- rep(NA, nrow(pfts_wash2))
+
+pft_montana_3$ppFVC <- pft_montana_3$Percent.predicted.relative.to.predicted.FVC.value
+pft_montana_3$ppFEV1 <- pft_montana_3$Percent.predicted.relative.to.predicted.FEV1..value
+pft_montana_3$bmipct <- pft_montana_3$BMI.Percentile
+pft_montana_3$Weight <- pft_montana_3$Weight.of.Patient
+reg3_BMI$prepost <- ifelse(reg3_BMI$diff < 0, "before","after")
+
+reg3_BMI$ppFEV1 <- pctpred_GLI(reg3_BMI$age, reg3_BMI$Height/100, gender = ifelse(reg3_BMI$Sex_MM == "M", 1, 2) , 
+                               ethnicity = 1, 
+                               FEV1 = reg3_BMI$FEV1)
+
+reg3_BMI$ppFVC <- pctpred_GLI(reg3_BMI$age, reg3_BMI$Height/100, gender = ifelse(reg3_BMI$Sex_MM == "M", 1, 2) , 
+                              ethnicity = 1, 
+                              FVC = reg3_BMI$FVC)
+
+pre_minn1 <- analytic_uofm2_2[,c("CFF.ID", "ppFVC",
+                    "ppFEV1","bmipct",
+                     "Weight", "A1C....of.Days.PRE.Trikafta.Start", "bmi")]
+pre_minn1$BMI.Value <- pre_minn1$bmi
+pre_minn1$diff <- as.numeric(pre_minn1$A1C....of.Days.PRE.Trikafta.Start)*-1
+pre_minn1$prepost <- rep("before", nrow(pre_minn1))
+
+post_minn1 <- analytic_uofm2_2[,c("CFF.ID", "ppFVC_post",
+                                 "ppFEV1_post","bmipct_post",
+                                 "Weight..kg..1", "A1C....of.Days.POST.Trikafta.start", 'bmi')]
+
+post_minn1$diff <- as.numeric(post_minn1$A1C....of.Days.POST.Trikafta.start)
+post_minn1$prepost <- rep("after", nrow(post_minn1))
+post_minn1$Weight <- post_minn1$Weight..kg..1
+post_minn1$ppFEV1 <-post_minn1$ppFEV1_post
+post_minn1$ppFVC <-post_minn1$ppFVC_post
+post_minn1$bmipct <- post_minn1$bmipct_post
+post_minn1$BMI.Value <- post_minn1$bmi
+
+reg3_BMI$BMI.Value <- reg3_BMI$BMI
+
+bmidat <- rbind(pfts_wash2[,c("CFF.ID", "ppFVC",
+              "ppFEV1","bmipct",
+              "diff","prepost", "Weight", "BMI.Value")],
+
+pft_montana_3[,c("CFF.ID", "ppFVC",
+                 "ppFEV1","bmipct",
+                 "diff","prepost", "Weight", "BMI.Value")],
+
+reg3_BMI[,c("CFF.ID", "ppFVC",
+            "ppFEV1","bmipct",
+            "diff","prepost", "Weight", "BMI.Value")],
+rbind(post_minn1[c("CFF.ID", "ppFVC",
+                   "ppFEV1","bmipct",
+                   "diff","prepost", "Weight", "BMI.Value")],
+      pre_minn1[c("CFF.ID", "ppFVC",
+                   "ppFEV1","bmipct",
+                   "diff","prepost", "Weight", "BMI.Value")]))
+
+#GET row closest to trik date
+
+bmidat_before <- bmidat %>% filter(diff != 0) %>% filter(prepost == "before") %>% 
+  group_by(CFF.ID) %>% arrange(desc(diff))%>% filter(row_number()==1)
+
+bmidat_before$time <- rep("before", nrow(bmidat_before))
+
+bmidat_after <- bmidat %>% filter(diff != 0) %>% filter(prepost == "after") %>% 
+  group_by(CFF.ID) %>% arrange(diff)%>% filter(row_number()==1)
+bmidat_after$time <- rep("post", nrow(bmidat_after))
+
+bmidat2 <- rbind(bmidat_before, bmidat_after)
+
+bmidat3 <- bmidat2 %>% group_by(CFF.ID) %>% count()
+bmidat4 <- subset(bmidat2, bmidat2$CFF.ID %in% bmidat3[bmidat3$n == 2,]$CFF.ID)
+
+bmidat4$ppFEV1 <- as.numeric(bmidat4$ppFEV1)
+bmidat4$ppFVC <- as.numeric(bmidat4$ppFVC)
+bmidat4$bmipct <- as.numeric(bmidat4$bmipct)
+bmidat4$Weight <- as.numeric(bmidat4$Weight)
+
+tab2b<- tableby(prepost~ppFVC + ppFEV1 + bmipct + Weight + BMI.Value, data=bmidat4[bmidat4$CFF.ID %in% time[!is.na(time$prepost.x),]$CFF.ID,])
+
+tab2b<- tableby(prepost~ppFVC + ppFEV1 + bmipct + Weight + BMI.Value, data=bmidat4[bmidat4$CFF.ID %in% a$CFF.ID,])
+
+bmidat4 <- merge(bmidat_before, bmidat_after, by = "CFF.ID")
+
+t.test(as.numeric(bmidat4$ppFVC.x),as.numeric(bmidat4$ppFVC.y), paired = T)
+t.test(bmidat4$ppFEV1.x,bmidat4$ppFEV1.y, paired = T)
+t.test(bmidat4$bmipct.x,bmidat4$bmipct.y , paired = T)
+t.test(as.numeric(bmidat4$Weight.x),as.numeric(bmidat4$Weight.y) , paired = T)
+t.test(as.numeric(bmidat4$BMI.Value.x),as.numeric(bmidat4$BMI.Value.y) , paired = T)
+
+#CGM data
+
+out <- str_split(cgm$subject_id, "_")
+cgm$ID <- do.call(rbind, out)[,1]
+cgm$time <- tolower(do.call(rbind, out)[,2])
+
+cgm0 <- subset(cgm, cgm$ID %in% a1c6_big2$CFF.ID)
+
+cgm_temp <- merge(cgm, a1c6_big2, by.x="ID", by.y="CFF.ID", all.x=T)
+
+cgm1 <- cgm0 %>% group_by(ID, time) %>% count()
+cgm2 <- cgm1 %>% group_by(ID) %>% count()
+
+View(cgm0[,c("subject_id", "ID", "time")])
+
+#pull first post and then do an overall
+
+cgm0$date <- as.Date(cgm0$date_cgm_placement)
+
+cgm02 <- cgm0 %>% filter(time=="post") %>%  group_by(ID) %>% 
+  arrange(date) %>% filter(row_number()==1)
+
+cgm0_small <- rbind(cgm0[cgm0$time == "pre",],cgm02)
+mycontrols  <- tableby.control(numeric.stats=c("medianq1q3"),
+                              stats.labels=list(medianq1q3="Median (Q1, Q3)"))
+
+a1c6_big2_tim <- a1c6_big2 %>% group_by(CFF.ID) %>% filter(row_number()==1)
+cgm0_time <- merge(cgm0_small, a1c6_big2_tim[,c("CFF.ID", "trik" )], by.y = "CFF.ID", by.x = "ID", all.x = T)
+
+cgm0_time$cgm_date <- as.Date(cgm0_time$date_cgm_placement, "%Y-%m-%d")
+cgm0_time$diff <- cgm0_time$date_cgm - as.Date(cgm0_time$trik, "%Y-%m-%d")
+
+View(cgm0_time[,c("date_cgm", "trik", "diff", "ID", "time")])
+
+tab3_extra <- tableby(time ~ diff, cgm0_time, control=mycontrols)
+
+tab3 <- tableby(time~ average_sensor
++ min_sensor
++ max_sensor
++ percent_time_over_120
++ percent_time_over_140
++ percent_time_over_200
++ avg_excur_over_140_per_day
++ avg_excur_over_200_per_day
++ percent_time_under_60
++ percent_time_under_70
++ average_auc_per_day
++ standard_deviation
++ r_mage, data=cgm0_small, control=mycontrols)
+
+tab3 <- tableby(~average_sensor
+                + min_sensor
+                + max_sensor
+                + percent_time_over_120
+                + percent_time_over_140
+                + percent_time_over_200
+                + avg_excur_over_140_per_day
+                + avg_excur_over_200_per_day
+                + percent_time_under_60
+                + percent_time_under_70
+                + average_auc_per_day
+                + standard_deviation
+                + r_mage, data=cgm0[cgm0$time == "post",]
+, control=mycontrols)
+
+temp <-merge(cgm0_small[cgm0_small$time=="pre",],
+cgm0_small[cgm0_small$time=="post",], by = "ID")
+
+wilcox.test(as.numeric(cgm0_time$diff )~cgm0_time$time, paired=T)
+wilcox.test(temp$average_sensor.x ,temp$average_sensor.y, paired=T)
+wilcox.test(temp$min_sensor.x   ,temp$min_sensor.y, paired=T)
+wilcox.test(temp$max_sensor.x   ,temp$max_sensor.y, paired=T)
+wilcox.test(temp$percent_time_over_120.x   ,temp$percent_time_over_120.y, paired=T)
+wilcox.test(temp$percent_time_over_140.x   ,temp$percent_time_over_140.y, paired=T)
+wilcox.test(temp$percent_time_over_200.x   ,temp$percent_time_over_200.y, paired=T)
+wilcox.test(temp$avg_excur_over_140_per_day.x   ,temp$avg_excur_over_140_per_day.y, paired=T)
+wilcox.test(temp$avg_excur_over_200_per_day.x   ,temp$avg_excur_over_200_per_day.y, paired=T)
+wilcox.test(temp$percent_time_under_60.x   ,temp$percent_time_under_60.y, paired=T)
+wilcox.test(temp$percent_time_under_70.x   ,temp$percent_time_under_70.y, paired=T)
+wilcox.test(temp$average_auc_per_day.x   ,temp$average_auc_per_day.y, paired=T)
+wilcox.test(temp$standard_deviation.x   ,temp$standard_deviation.y, paired=T)
+wilcox.test(temp$r_mage.x   ,temp$r_mage.y, paired=T)
+
+wilcox.test(cgm0_small$average_sensor~cgm0_small$time, paired=T)
+
+analytic_cm$ppFVC <- as.numeric(analytic_cm$ppFVC)
+
+tab1 <- tableby(~age+bmipct + Weight + Sex + ppFEV1 + ppFVC  +
+                  G.tube.in.past.12.months + Pancreatic.Status +
+                  genoRisk + location+BMI.Value + age_category, 
+                data=analytic_cm[analytic_cm$CFF.ID %in% cgm0_small$ID,])
 ```

### 2022-01-19 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
Update pre_post_trikafta.Rmd
nicer printing

@@
 knitr::opts_chunk$set(echo = FALSE)
-knitr::opts_knit$set(root.dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/Prepost Triakfta")
+knitr::opts_knit$set(root.dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Raw")
 ```
 
-```{r}
+```{r warning=FALSE,results='asis'}
 ############pre/post trikaftka
@@
 ##############################
-library(dplyr)
-setwd("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/")
 
@@
 # cgm_c <- read.csv("REDCap Upload.csv")
-cgm_c <- read.csv("Data From CHCO/REDCap Upload_chco.csv")
+cgm_c <- read.csv("./Data From CHCO/REDCap Upload_chco.csv")
 cgm_c$date_cgm <- as.Date(cgm_c$date_cgm_placement, "%Y-%m-%d")
 
-cgm_m <- read.csv("Data From Montana/CGM_Montana/REDCap Upload_mt.csv")
+cgm_m <- read.csv("./Data From Montana/CGM_Montana/REDCap Upload_mt.csv")
 cgm_m$date_cgm <- as.Date(cgm_m$date_cgm_placement, "%m/%d/%Y")
 
-cgm_w <- read.csv("Data From Washington/REDCap Upload_Washington.csv")
+cgm_w <- read.csv("./Data From Washington/REDCap Upload_Washington.csv")
 cgm_w$date_cgm <- as.Date(cgm_w$date_cgm_placement, "%Y-%m-%d")
@@
 `%!in%` = Negate(`%in%`)
-reg0 <- read.csv("Data From CHCO/Registry/Copy of Earliest Trikafta Dates in Registry_updated 2.2.21 cc modified_cfrdyn.csv")
+reg0 <- read.csv("./Data From CHCO/Registry/Copy of Earliest Trikafta Dates in Registry_updated 2.2.21 cc modified_cfrdyn.csv")
 #need to get rid of those two
@@
 
-a1c <- read.csv("Data From CHCO/Copy of A1C and OGTTs Before and After Trikafta_cc modified 8.25.20.csv")
-
-weight <- read.csv("Data From CHCO/Registry/weightdata.csv")
+a1c <- read.csv("./Data From CHCO/Copy of A1C and OGTTs Before and After Trikafta_cc modified 8.25.20.csv")
+
+weight <- read.csv("./Data From CHCO/Registry/weightdata.csv")
 
@@
 
-reg_montana <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/Copy of Paper patient population glycemic (+hofer).csv")
-
-demo_adult <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/demo_adults.csv")
-demo_kids <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/demo_kids.csv")
+reg_montana <- read.csv("./Data From Montana/Copy of Paper patient population glycemic (+hofer).csv")
+
+demo_adult <- read.csv("./Data From Montana/demo_adults.csv")
+demo_kids <- read.csv("./Data From Montana/demo_kids.csv")
 
@@
 
-pft_adult <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/Copy of PFTs growth adult (+hofer).csv")
-pft_kids <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/PFTs kids.csv")
+pft_adult <- read.csv("./Data From Montana/Copy of PFTs growth adult (+hofer).csv")
+pft_kids <- read.csv("./Data From Montana/PFTs kids.csv")
 
@@
 
-pft_montana_5$BMI.Percentile
-
 #need sex
@@
 
-gt_adult <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/gtubes_adult.csv")
-gt_kids <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/GT users kids.csv")
+gt_adult <- read.csv("./Data From Montana/gtubes_adult.csv")
+gt_kids <- read.csv("./Data From Montana/GT users kids.csv")
 
@@
 
-is_adult <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/pancreatic insufficient adult.csv")
-is_kids <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/pancreatic insufficient kids.csv")
+is_adult <- read.csv("./Data From Montana/pancreatic insufficient adult.csv")
+is_kids <- read.csv("./Data From Montana/pancreatic insufficient kids.csv")
 
@@
 
-g_adult <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/genotypes_adult.csv")
-g_kids <- read.csv("S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.02.09 - Prepost Triakfta/Data From Montana/genotypes kids.csv")[,1:3]
+g_adult <- read.csv("./Data From Montana/genotypes_adult.csv")
+g_kids <- read.csv("./Data From Montana/genotypes kids.csv")[,1:3]
 
@@
 
-reg_washington <- read.csv("Data From Washington/Copy of CFRD and trikafta Trikafta_on_Annual_Review-13-Apr-2021-0815.csv")
-reg_wash <- read.csv("Data From Washington/Copy of CFRD and trikafta Trikafta_on_Annual_Review-13-Apr-2021-0815.csv")
+reg_washington <- read.csv("./Data From Washington/Copy of CFRD and trikafta Trikafta_on_Annual_Review-13-Apr-2021-0815.csv")
+reg_wash <- read.csv("./Data From Washington/Copy of CFRD and trikafta Trikafta_on_Annual_Review-13-Apr-2021-0815.csv")
 
@@
 #aic 
-a1c_wash <- read.csv("Data From Washington/A1C_and_OGTT.csv")
+a1c_wash <- read.csv("./Data From Washington/A1C_and_OGTT.csv")
 
@@
 #bmi
-bmi_wash <- read.csv("Data From Washington/BMIandWeight.csv")
+bmi_wash <- read.csv("./Data From Washington/BMIandWeight.csv")
 bmi_wash1 <- merge(bmi_wash, analytic_wash, by = "CFF.ID")
@@
 #sex
-demo_wash <- read.csv("Data From Washington/Demographics.csv")
+demo_wash <- read.csv("./Data From Washington/Demographics.csv")
 
@@
 #Insufficent
-is_wash <- read.csv("Data From Washington/Pancreatic_Insufficient.csv")
+is_wash <- read.csv("./Data From Washington/Pancreatic_Insufficient.csv")
 
@@
 
-gt_wash <- read.csv("Data From Washington/G-tube.csv")
+gt_wash <- read.csv("./Data From Washington/G-tube.csv")
 
@@
 
-geno_wash <- read.csv("Data From Washington/Genotypes.csv")
+geno_wash <- read.csv("./Data From Washington/Genotypes.csv")
 
@@
 
-pfts_wash <- read.csv("Data From Washington/PFTs.csv")
+pfts_wash <- read.csv("./Data From Washington/PFTs.csv")
 
@@
 
-uofm <- read.csv("Data from UofM/Moheet_Pre_Post_Trikafta_FinalData_20210914.csv")
+uofm <- read.csv("./Data from UofM/Moheet_Pre_Post_Trikafta_FinalData_20210914.csv")
 analytic_uofm <- subset(uofm, uofm$CFRD.Status == "CFRD with or without fasting hyperglycemia")
@@
 analytic_uofm2$BMI.Value <- analytic_uofm2$bmi
-analytic_uofm2$age_category
 analytic_uofm2_2 <- subset(analytic_uofm2, analytic_uofm2$Encounter....of.Days.POST.Trikafta.start <= 365 &
@@
 #need a1c
-
-a1c_wash2
-
-reg_montana2
 
@@
                                stats.labels=list(N='Count', median='Median', q1q3='Q1,Q3'))
-
 tab2 <- tableby(prepost ~ n, control=mycontrols, data=a1c7)
@@
 a <- merge(before, after, by = "CFF.ID", all.y = T)
-wilcox.test(a$pre,a$post, paired = T)
 
@@
 
-summary(before_max$time)
-sd(before_max$time)
-
 after <- subset(a1c6_big2, a1c6_big2$prepost == "after")
@@
 
-summary(after_max$time)
-sd(after_max$time)
-
 #test this
-t.test(before_max$time,after_max$time)
 
 time <- merge(before_max,after_max,by='CFF.ID',all.y=T)
-summary(time$time.x)
-sd(time$time.x, na.rm=T)
-summary(time[!is.na(time$time.x),]$time.y)
-sd(time[!is.na(time$time.x),]$time.y, na.rm=T)
-t.test(time$time.x,time$time.y, paired = T)
 
 time <- merge(before1,after1,by='CFF.ID',all.y=T)
-
-t.test(as.numeric(time$Hgb.A1C.x),as.numeric(time$Hgb.A1C.y), paired = T)
-
-summary(as.numeric(time$Hgb.A1C.x))
-sd(as.numeric(time$Hgb.A1C.x), na.rm=T)
-
-summary(as.numeric(time$Hgb.A1C.y))
-sd(as.numeric(time$Hgb.A1C.y),na.rm=T)
 
@@
 time$diff_a1c <- as.numeric(time$Hgb.A1C.x) - as.numeric(time$Hgb.A1C.y)
-summary(time[time$Hgb.A1C.x < 6.5, ]$diff_a1c)
-
-t.test(time[time$Hgb.A1C.x < 6.5, ]$diff_a1c, mu=0)
 
@@
 time2 <- subset(time, time$CFF.ID %in% cgm$ID)
-
-wilcox.test(as.numeric(time2$Hgb.A1C.x), as.numeric(time2$Hgb.A1C.y), paired = T)
 
@@
 
-summary(ins7[ins7$prepost == "Post",]$Long.acting.or.total.basal.daily.dose..units.)
-
 ins8 <- merge(ins7[ins7$prepost == "Pre",], ins7[ins7$prepost == "Post",], by = "CFF.ID")
-
-wilcox.test(ins8$Long.acting.or.total.basal.daily.dose..units..x, ins8$Long.acting.or.total.basal.daily.dose..units..y, paired = T)
 
@@
 tab2 <- tableby(time~Weight+ bmiz + ppFEV1+ppFVC, data=timing)
-
 timing2 <- merge(weigh_before1,weigh_after1, by = "MRN", all.y=T)
-wilcox.test(timing2$ppFEV1.x,timing2$ppFEV1.y, paired=T)
-
-wilcox.test(timing2$ppFVC.x,timing2$ppFVC.y, paired=T)
 
@@
 
-tab2b<- tableby(prepost~ppFVC + ppFEV1 + bmipct + Weight + BMI.Value, data=bmidat4[bmidat4$CFF.ID %in% time[!is.na(time$prepost.x),]$CFF.ID,])
-
-tab2b<- tableby(prepost~ppFVC + ppFEV1 + bmipct + Weight + BMI.Value, data=bmidat4[bmidat4$CFF.ID %in% a$CFF.ID,])
+tab2b<- tableby(prepost~ppFVC + ppFEV1 + bmipct + Weight + BMI.Value, 
+                data=bmidat4[bmidat4$CFF.ID %in% time[!is.na(time$prepost.x),]$CFF.ID,])
+tab2b<- tableby(prepost~ppFVC + ppFEV1 + bmipct + Weight + BMI.Value, 
+                data=bmidat4[bmidat4$CFF.ID %in% a$CFF.ID,])
 
 bmidat4 <- merge(bmidat_before, bmidat_after, by = "CFF.ID")
-
-t.test(as.numeric(bmidat4$ppFVC.x),as.numeric(bmidat4$ppFVC.y), paired = T)
-t.test(bmidat4$ppFEV1.x,bmidat4$ppFEV1.y, paired = T)
-t.test(bmidat4$bmipct.x,bmidat4$bmipct.y , paired = T)
-t.test(as.numeric(bmidat4$Weight.x),as.numeric(bmidat4$Weight.y) , paired = T)
-t.test(as.numeric(bmidat4$BMI.Value.x),as.numeric(bmidat4$BMI.Value.y) , paired = T)
 
@@
 
-View(cgm0[,c("subject_id", "ID", "time")])
-
 #pull first post and then do an overall
@@
 cgm0_time$diff <- cgm0_time$date_cgm - as.Date(cgm0_time$trik, "%Y-%m-%d")
-
-View(cgm0_time[,c("date_cgm", "trik", "diff", "ID", "time")])
 
@@
                 + r_mage, data=cgm0_small, control=mycontrols)
-
 tab3 <- tableby(~average_sensor
@@
 
-wilcox.test(as.numeric(cgm0_time$diff )~cgm0_time$time, paired=T)
-wilcox.test(temp$average_sensor.x ,temp$average_sensor.y, paired=T)
-wilcox.test(temp$min_sensor.x   ,temp$min_sensor.y, paired=T)
-wilcox.test(temp$max_sensor.x   ,temp$max_sensor.y, paired=T)
-wilcox.test(temp$percent_time_over_120.x   ,temp$percent_time_over_120.y, paired=T)
-wilcox.test(temp$percent_time_over_140.x   ,temp$percent_time_over_140.y, paired=T)
-wilcox.test(temp$percent_time_over_200.x   ,temp$percent_time_over_200.y, paired=T)
-wilcox.test(temp$avg_excur_over_140_per_day.x   ,temp$avg_excur_over_140_per_day.y, paired=T)
-wilcox.test(temp$avg_excur_over_200_per_day.x   ,temp$avg_excur_over_200_per_day.y, paired=T)
-wilcox.test(temp$percent_time_under_60.x   ,temp$percent_time_under_60.y, paired=T)
-wilcox.test(temp$percent_time_under_70.x   ,temp$percent_time_under_70.y, paired=T)
-wilcox.test(temp$average_auc_per_day.x   ,temp$average_auc_per_day.y, paired=T)
-wilcox.test(temp$standard_deviation.x   ,temp$standard_deviation.y, paired=T)
-wilcox.test(temp$r_mage.x   ,temp$r_mage.y, paired=T)
-
-wilcox.test(cgm0_small$average_sensor~cgm0_small$time, paired=T)
-
 analytic_cm$ppFVC <- as.numeric(analytic_cm$ppFVC)
@@
                 data=analytic_cm[analytic_cm$CFF.ID %in% cgm0_small$ID,])
+summary(tab1)
 ```

### 2022-01-19 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
Update pre_post_trikafta.Rmd

@@
 analytic_cm <- rbind(reg_tab1 , analytic_montana1,analytic_wash1,analytic_uofm3 )
+# Add BMI category
+analytic_cm$bmicat = reg3_BMI$bmicat[match(analytic_cm$CFF.ID,reg3_BMI$CFF.ID)]
 
@@
 
-tab3 <- tableby(time~ average_sensor
+tab3a <- tableby(time~ average_sensor
                 + min_sensor
@@
                 + r_mage, data=cgm0_small, control=mycontrols)
-tab3 <- tableby(~average_sensor
+tab3c <- tableby(~average_sensor
                 + min_sensor
@@
 
-tab1 <- tableby(~age+bmipct + Weight + Sex + ppFEV1 + ppFVC  +
+tab3b <- tableby(~age+bmipct + Weight + Sex + ppFEV1 + ppFVC  +
                   G.tube.in.past.12.months + Pancreatic.Status +
@@
                 data=analytic_cm[analytic_cm$CFF.ID %in% cgm0_small$ID,])
-summary(tab1)
+summary(tab3a)
+summary(tab3b)
+summary(tab3c)
 ```

### 2022-01-19 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
Update pre_post_trikafta.Rmd
function to get closest visits to start date

--- 2022-01-19 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
@@
 knitr::opts_chunk$set(echo = FALSE)
-knitr::opts_knit$set(root.dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Raw")
+knitr::opts_knit$set(root.dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/Prepost Triakfta")
 ```
 
-```{r warning=FALSE,results='asis'}
-############pre/post trikaftka
-# c. chan
-##############################
+```{r code from maxie,eval=FALSE}
+# Source Maxie's code, and write final ID lists
+source("~/GitHub/CF/Christine Chan/Pre-post Trikafta/p001+cmw_UPDATED.R")
+ids = data.frame(unique(analytic_cm$CFF.ID))
+colnames(ids) = "CFF.ID"
+write.csv(ids,file = "./Data_Cleaned/final_ids.csv",row.names = F)
+```
 
-#cgm data
-# cgm_c <- read.csv("REDCap Upload.csv")
-cgm_c <- read.csv("./Data From CHCO/REDCap Upload_chco.csv")
-cgm_c$date_cgm <- as.Date(cgm_c$date_cgm_placement, "%Y-%m-%d")
+```{r closest date function}
+closest_dates = function(df,start = "Start",date = "Date"){
+  # Start and date columns must be formated for subtraction (ideally with lubridate)
+  df$Timepoint = as.numeric(df[,date] - df[,start])
+  # Remove visits on start date or > 1 year out
+  df = df %>% filter(Timepoint !=0 & abs(Timepoint) <= 365) %>% 
+    # Get closest visit before and closest after
+    group_by(CFF.ID,sign(Timepoint)) %>%
+    slice_min(Timepoint * sign(Timepoint),n=1) %>% 
+    # Format
+    ungroup %>% select(-Timepoint) %>% rename(Timepoint = `sign(Timepoint)`)
+  # Timepoint as factor
+  df$Timepoint = factor(df$Timepoint,labels = c("Before","After"))
+  # Delete unnecessary columns
+  df = df %>% select(-all_of(c(start,date)))
+  # Return
+  df = data.frame(df)
+  return(df)
+} 
+```
 
-cgm_m <- read.csv("./Data From Montana/CGM_Montana/REDCap Upload_mt.csv")
-cgm_m$date_cgm <- as.Date(cgm_m$date_cgm_placement, "%m/%d/%Y")
+# Table 1: Demographics at Trikafta Start
 
-cgm_w <- read.csv("./Data From Washington/REDCap Upload_Washington.csv")
-cgm_w$date_cgm <- as.Date(cgm_w$date_cgm_placement, "%Y-%m-%d")
+```{r}
+# Start dates
+chco_demographics = read.csv("./Data_Cleaned/chco_trikafta_dates.csv",na.strings = "")
+# Fix select dates
+chco_demographics$Earliest.Trikafta.Date.in.PortCF[chco_demographics$MRN =="1053930"] = "12/1/2019"
+chco_demographics$Earliest.Trikafta.Date.in.PortCF[chco_demographics$MRN =="1695512"] = "12/12/2019"
+chco_demographics$Earliest.Trikafta.Date.in.PortCF[chco_demographics$MRN =="815274"] = "12/13/2019"
+chco_demographics$Earliest.Trikafta.Date.in.PortCF[chco_demographics$MRN =="1078408"] = "3/9/2020"
+chco_demographics$Earliest.Trikafta.Date.in.PortCF[chco_demographics$MRN =="899717"] = "2/14/2020"
+chco_demographics$CFF.ID = as.character(chco_demographics$CFF.ID)
+# Diabetes and with start date only
+chco_demographics = chco_demographics %>% 
+  filter(!is.na(Earliest.Trikafta.Date.in.PortCF),CFRD.yes.1 == 1)
+chco_demographics$Earliest.Trikafta.Date.in.PortCF = 
+  mdy(chco_demographics$Earliest.Trikafta.Date.in.PortCF)
+chco_dates = chco_demographics %>% rename(Start = Earliest.Trikafta.Date.in.PortCF) %>%
+  select(CFF.ID,MRN,Start)
+# A1C and OGTT
+chco_a1c = read.csv("./Data_Cleaned/chco_a1c_ogtt.csv")
+## Format
+chco_a1c$CFF.ID = as.character(chco_a1c$CFF.ID)
+chco_a1c$Date = mdy(chco_a1c$Date)
+chco_a1c$Diabetes = NULL
+## Only those w/CFRD and known start
+chco_a1c = left_join(chco_dates,chco_a1c,by = "CFF.ID")
+## Get closest before and closest after dates
+chco_a1c = closest_dates(chco_a1c)
+# PFTs
+chco_pft = read.csv("./Data_Cleaned/chco_bmi_pft.csv")
+chco_pft$Date = mdy(chco_pft$Date)
+chco_pft = left_join(chco_dates,chco_pft,by = "MRN")
+chco_pft = closest_dates(chco_pft)
+# CGM
+chco_cgm = read.csv("./Data_Cleaned/chco_cgm.csv")
+chco_cgm$Date = date(mdy_hm(chco_cgm$date_cgm_placement))
+chco_cgm$CFF.ID = sub("_.*","",chco_cgm$subject_id)
+# chco_cgm$subject_id = NULL
+chco_cgm = left_join(chco_dates,chco_cgm,by = "CFF.ID")
+chco_cgm = closest_dates(chco_cgm)
+# Merge
+chco = full_join(chco_a1c,chco_pft,by = c("CFF.ID", "MRN", "Timepoint"))
+chco = full_join(t,chco_cgm,by = c("CFF.ID", "MRN", "Timepoint"))
 
-cgm <- rbind(cgm_c, cgm_m,cgm_w )
-
-`%!in%` = Negate(`%in%`)
-reg0 <- read.csv("./Data From CHCO/Registry/Copy of Earliest Trikafta Dates in Registry_updated 2.2.21 cc modified_cfrdyn.csv")
-#need to get rid of those two
-reg <- subset(reg0, reg0$MRN %!in% c("1714343", "787372") )
-
-reg$Earliest.Trikafta.Date.in.PortCF <- ifelse(reg$MRN =="1053930", "12/1/2019", 
-                                               ifelse(reg$MRN =="1695512", "12/12/2019", 
-                                                      ifelse(reg$MRN =="815274", "12/13/2019", 
-                                                             ifelse(reg$MRN =="1078408",   "3/9/2020",
-                                                                    ifelse(reg$MRN =="899717", "2/14/2020", as.character(reg$Earliest.Trikafta.Date.in.PortCF))))))
-
-a1c <- read.csv("./Data From CHCO/Copy of A1C and OGTTs Before and After Trikafta_cc modified 8.25.20.csv")
-
-weight <- read.csv("./Data From CHCO/Registry/weightdata.csv")
-
-weight$Earliest.Trikafta.Date.in.PCF <- ifelse(weight$MRN =="1053930", "12/1/2019", 
-                                               ifelse(weight$MRN =="1695512", "12/12/2019", 
-                                                      ifelse(weight$MRN =="815274", "12/13/2019", 
-                                                             ifelse(weight$MRN =="1078408",   "3/9/2020",
-                                                                    ifelse(weight$MRN =="899717", "2/14/2020", as.character(weight$Earliest.Trikafta.Date.in.PCF))))))
-
-#montana data
-#total population
-
-library("readxl")
-
-reg_montana <- read.csv("./Data From Montana/Copy of Paper patient population glycemic (+hofer).csv")
-
-demo_adult <- read.csv("./Data From Montana/demo_adults.csv")
-demo_kids <- read.csv("./Data From Montana/demo_kids.csv")
-
-reg_montana$age <- ifelse(reg_montana$CFF.ID %in% demo_adult$CFF.ID, "Adult", 
-                          ifelse(reg_montana$CFF.ID %in% demo_kids$CFF.ID, "Kids", "Check"))
-
-reg_montana$trik <- as.Date(reg_montana$Trikafta.start.date,"%m/%d/%Y")
-
-reg_montana$test <- as.Date(reg_montana$Date.of.Encounter,  "%m/%d/%Y")
-reg_montana$prepost <- ifelse(reg_montana$trik > reg_montana$test, "before", "after")
-
-reg_montana$diff <- reg_montana$trik-reg_montana$test 
-reg_montana2 <- subset(reg_montana, abs(diff) <= 365)
-
-reg_montana3 <- reg_montana2 %>% group_by(CFF.ID, prepost) %>% count()
-reg_montana4 <- reg_montana3 %>% group_by(CFF.ID) %>% count()
-
-#weight/fev
-
-pft_adult <- read.csv("./Data From Montana/Copy of PFTs growth adult (+hofer).csv")
-pft_kids <- read.csv("./Data From Montana/PFTs kids.csv")
-
-pft_montana <- rbind(pft_adult, pft_kids)
-
-pft_montana_1 <- subset(pft_montana, pft_montana$CFF.ID %in% reg_montana$CFF.ID)
-
-#age at pft
-pft_montana_1$pft <- as.Date(pft_montana_1$Date.of.Encounter,  "%m/%d/%Y")
-
-reg_montana_first <- reg_montana %>% group_by(CFF.ID) %>% filter(row_number()==1)
-
-reg_montana_first$trik <- ifelse(reg_montana_first$CFF.ID == "1044621", "2019-12-02", 
-                                 ifelse(reg_montana_first$CFF.ID == "1008380", "2020-1-1", as.character(reg_montana_first$trik)))
-
-pft_montana_2 <- merge(pft_montana_1, reg_montana_first[,c("CFF.ID" ,"Date.Of.Birth..mm.dd.yyyy.", "trik")])
-
-#need pfts within a year of trik
-
-pft_montana_2$diff <-as.numeric(as.Date(pft_montana_2$pft) - as.Date(pft_montana_2$trik))
-pft_montana_3 <- subset(pft_montana_2, abs(diff) <= 365 )
-pft_montana_3$prepost <- ifelse(pft_montana_3$trik > pft_montana_3$pft, "before", "after")
-pft_montana_4 <- subset(pft_montana_3, pft_montana_3$prepost == "before")
-pft_montana_5 <- pft_montana_4 %>% group_by(CFF.ID) %>% arrange(diff) %>% filter(row_number()==1)
-
-pft_montana_5$DOB <- as.Date(pft_montana_5$Date.Of.Birth..mm.dd.yyyy.,  "%m/%d/%Y")
-
-pft_montana_5$age <- as.numeric(pft_montana_5$pft-pft_montana_5$DOB)/365.25
-
-#need sex
-demo_montana <- rbind(demo_kids[,c("CFF.ID","Gender")], demo_adult[,c("CFF.ID","Gender")])
-
-analytic_montana <- merge(pft_montana_5, demo_montana ,by = "CFF.ID", all.x=T)
-
-#need cfrd start date - don't have dates
-
-#gtube status
-
-gt_adult <- read.csv("./Data From Montana/gtubes_adult.csv")
-gt_kids <- read.csv("./Data From Montana/GT users kids.csv")
-
-gt_montana <- rbind(gt_adult, gt_kids)
-
-gt_montana_1 <- merge(gt_montana, reg_montana_first[,c("CFF.ID" , "trik")])
-gt_montana_1$gt <- as.Date(gt_montana_1$Date.of.Encounter,  "%m/%d/%Y")
-gt_montana_1$diff <-as.numeric(as.Date(gt_montana_1$gt) - as.Date(gt_montana_1$trik))
-gt_montana_2 <- subset(gt_montana_1, abs(diff) <= 365 )
-analytic_montana$gtube <- ifelse(analytic_montana$CFF.ID %in% gt_montana_2$CFF.ID, "Yes", "No")
-
-#Insufficent
-
-is_adult <- read.csv("./Data From Montana/pancreatic insufficient adult.csv")
-is_kids <- read.csv("./Data From Montana/pancreatic insufficient kids.csv")
-
-analytic_montana$is <- ifelse(analytic_montana$CFF.ID %in% c(is_adult$CFF.ID, is_kids$CFF.ID), "Insufficient", "Not")
-
-#genotypes
-
-g_adult <- read.csv("./Data From Montana/genotypes_adult.csv")
-g_kids <- read.csv("./Data From Montana/genotypes kids.csv")[,1:3]
-
-geno_montana <- rbind(g_adult,g_kids)
-
-geno_montana$Genotypes1 <- geno_montana$Name.of.the.first.mutation
-
-geno_montana$Genotypes2 <- geno_montana$Name.of.the.second.mutation
-
-geno_montana$class2 <- ifelse(geno_montana$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-                                                             '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-                                                             '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-                                                             '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
-                              ifelse(geno_montana$Genotypes2 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
-                                     "II", 
-                                     ifelse(geno_montana$Genotypes2 %in% c('G551D', 'G551S', 'S549N'), "III", 
-                                            ifelse(geno_montana$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
-                                                   "IV", 
-                                                   ifelse(geno_montana$Genotypes2 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-                                                                                         'P574H', '3272-26A->G5'), "V" , "")))))
-
-geno_montana$severity2 <- ifelse(geno_montana$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-                                                                '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-                                                                '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-                                                                '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
-                                                                'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
-                                                                '2183delAA->G', '2184insA', 'S434X',
-                                                                '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
-                                                                'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
-                                                                "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
-                                 "Severe", 
-                                 ifelse(geno_montana$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
-                                                                       'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-                                                                       'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
-                                                                       'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
-                                                                       "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
-
-                                        ifelse(geno_montana$Genotypes2 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
-                                                                                'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
-                                                                                'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
-                                                                                'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
-                                                                                "R1066H", "P205S"), 
-                                               "Unk","")))
-
-geno_montana$class1 <- ifelse(geno_montana$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-                                                             '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-                                                             '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-                                                             '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
-                              ifelse(geno_montana$Genotypes1 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
-                                     "II", 
-                                     ifelse(geno_montana$Genotypes1 %in% c('G551D', 'G551S', 'S549N'), "III", 
-                                            ifelse(geno_montana$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
-                                                   "IV", 
-                                                   ifelse(geno_montana$Genotypes1 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-                                                                                         'P574H', '3272-26A->G5'), "V" , "")))))   
-geno_montana$severity1 <- ifelse(geno_montana$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-                                                                '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-                                                                '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-                                                                '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
-                                                                'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
-                                                                '2183delAA->G', '2184insA', 'S434X',
-                                                                '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
-                                                                'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
-                                                                "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
-                                 "Severe", 
-                                 ifelse(geno_montana$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
-                                                                       'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-                                                                       'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
-                                                                       'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
-                                                                       "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
-
-                                        ifelse(geno_montana$Genotypes1 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
-                                                                                'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
-                                                                                'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
-                                                                                'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
-                                                                                "R1066H", "P205S"), 
-                                               "Unk","")))
-
-geno_montana$genoRisk <- ifelse(geno_montana$severity1 == "Mild" | geno_montana$severity2 == "Mild", "Low", 
-                                ifelse(geno_montana$severity1 == "Severe" & geno_montana$severity2 == "Severe", "High", ""))
-
-analytic_montana$genotype <- ifelse(analytic_montana$CFF.ID %in% geno_montana[geno_montana$genoRisk == "High", ]$CFF.ID, "High", 
-                                    ifelse(analytic_montana$CFF.ID %in% geno_montana[geno_montana$genoRisk == "Low", ]$CFF.ID, "Low",   ""   ))
-
-#washingotn
-
-reg_washington <- read.csv("./Data From Washington/Copy of CFRD and trikafta Trikafta_on_Annual_Review-13-Apr-2021-0815.csv")
-reg_wash <- read.csv("./Data From Washington/Copy of CFRD and trikafta Trikafta_on_Annual_Review-13-Apr-2021-0815.csv")
-
-analytic_wash <- reg_washington %>% group_by(CFF.ID) %>% filter(row_number()==1) %>% filter(CFF.ID != "1499770")
-analytic_wash$trik <- as.Date(analytic_wash$Approximate.date.patient.started.taking.this.drug,  "%m/%d/%Y")
-
-#aic 
-a1c_wash <- read.csv("./Data From Washington/A1C_and_OGTT.csv")
-
-temp_wash <- reg_wash[,c("CFF.ID", "Approximate.date.patient.started.taking.this.drug")]
-temp_wash1 <- temp_wash %>% group_by(CFF.ID) %>% filter(row_number()==1)
-
-a1c_wash1 <- merge(a1c_wash, temp_wash1, by = c("CFF.ID"))
-a1c_wash1$pft <- as.Date(a1c_wash1$Date.of.Encounter,  "%m/%d/%Y")
-a1c_wash1$trik <- as.Date(a1c_wash1$Approximate.date.patient.started.taking.this.drug,  "%m/%d/%Y")
-a1c_wash1$diff <-as.numeric(a1c_wash1$pft - a1c_wash1$trik)
-a1c_wash2<- subset(a1c_wash1, abs(diff) <= 365 )
-
-#bmi
-bmi_wash <- read.csv("./Data From Washington/BMIandWeight.csv")
-bmi_wash1 <- merge(bmi_wash, analytic_wash, by = "CFF.ID")
-bmi_wash1$Date <- as.Date(bmi_wash1$Date.of.Encounter.x,  "%m/%d/%Y")
-bmi_wash1$diff <-as.numeric(bmi_wash1$Date - bmi_wash1$trik)
-
-bmi_wash2 <- subset(bmi_wash1, abs(diff) <= 365 )
-bmi_wash2$prepost <- ifelse(bmi_wash2$trik > bmi_wash2$Date, "before", "after")
-
-#need age/bmi from first visit
-bmi_wash3 <- bmi_wash2 %>% group_by(CFF.ID) %>% filter(prepost == "before") %>% arrange(diff) %>%  filter(row_number()==1)
-bmi_wash3$first_encounter <- bmi_wash3$Date.of.Encounter.x
-
-analytic_wash <- merge(analytic_wash,bmi_wash3[,c("CFF.ID","BMI.Percentile" ,"Encounter.Age"  ,"Patients.height","BMI.Value","first_encounter")], by = "CFF.ID")
-
-#sex
-demo_wash <- read.csv("./Data From Washington/Demographics.csv")
-
-analytic_wash$Sex <- ifelse(analytic_wash$CFF.ID %in% demo_wash[demo_wash$Gender== "Female",]$CFF.ID, "F", 
-                            ifelse(analytic_wash$CFF.ID %in% demo_wash[demo_wash$Gender== "Male",]$CFF.ID, "M",  "Check"))
-
-#Insufficent
-is_wash <- read.csv("./Data From Washington/Pancreatic_Insufficient.csv")
-
-analytic_wash$is <- ifelse(analytic_wash$CFF.ID %in% is_wash$CFF.ID, "Insufficient", "Not")
-
-#gtube
-
-gt_wash <- read.csv("./Data From Washington/G-tube.csv")
-
-analytic_wash$gt <- ifelse(analytic_wash$CFF.ID %in% gt_wash$CFF.ID, "GT", "Not")
-
-#genotypes
-
-geno_wash <- read.csv("./Data From Washington/Genotypes.csv")
-
-geno_wash$Genotypes1 <- geno_wash$Name.of.the.first.mutation
-
-geno_wash$Genotypes2 <- geno_wash$Name.of.the.second.mutation
-
-#
-geno_wash$class2 <- ifelse(geno_wash$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-                                                       '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-                                                       '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-                                                       '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
-                           ifelse(geno_wash$Genotypes2 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
-                                  "II", 
-                                  ifelse(geno_wash$Genotypes2 %in% c('G551D', 'G551S', 'S549N'), "III", 
-                                         ifelse(geno_wash$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
-                                                "IV", 
-                                                ifelse(geno_wash$Genotypes2 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-                                                                                   'P574H', '3272-26A->G5'), "V" , "")))))
-
-geno_wash$severity2 <- ifelse(geno_wash$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-                                                          '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-                                                          '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-                                                          '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
-                                                          'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
-                                                          '2183delAA->G', '2184insA', 'S434X',
-                                                          '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
-                                                          'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
-                                                          "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
-                              "Severe", 
-                              ifelse(geno_wash$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
-                                                                 'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-                                                                 'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
-                                                                 'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
-                                                                 "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
-
-                                     ifelse(geno_wash$Genotypes2 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
-                                                                          'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
-                                                                          'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
-                                                                          'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
-                                                                          "R1066H", "P205S"), 
-                                            "Unk","")))
-
-geno_wash$class1 <- ifelse(geno_wash$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-                                                       '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-                                                       '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-                                                       '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
-                           ifelse(geno_wash$Genotypes1 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
-                                  "II", 
-                                  ifelse(geno_wash$Genotypes1 %in% c('G551D', 'G551S', 'S549N'), "III", 
-                                         ifelse(geno_wash$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
-                                                "IV", 
-                                                ifelse(geno_wash$Genotypes1 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-                                                                                   'P574H', '3272-26A->G5'), "V" , "")))))   
-geno_wash$severity1 <- ifelse(geno_wash$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-                                                          '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-                                                          '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-                                                          '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
-                                                          'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
-                                                          '2183delAA->G', '2184insA', 'S434X',
-                                                          '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
-                                                          'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
-                                                          "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
-                              "Severe", 
-                              ifelse(geno_wash$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
-                                                                 'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-                                                                 'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
-                                                                 'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
-                                                                 "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
-
-                                     ifelse(geno_wash$Genotypes1 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
-                                                                          'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
-                                                                          'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
-                                                                          'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
-                                                                          "R1066H", "P205S"), 
-                                            "Unk","")))
-
-geno_wash$genoRisk <- ifelse(geno_wash$severity1 == "Mild" | geno_wash$severity2 == "Mild", "Low", 
-                             ifelse(geno_wash$severity1 == "Severe" & geno_wash$severity2 == "Severe", "High", ""))
-
-analytic_wash$genotype <- ifelse(analytic_wash$CFF.ID %in% geno_wash[geno_wash$genoRisk == "High", ]$CFF.ID, "High", 
-                                 ifelse(analytic_wash$CFF.ID %in% geno_wash[geno_wash$genoRisk == "Low", ]$CFF.ID, "Low",   ""   ))
-
-#pfts
-
-pfts_wash <- read.csv("./Data From Washington/PFTs.csv")
-
-pfts_wash1 <- merge(pfts_wash, analytic_wash, by = "CFF.ID")
-pfts_wash1$Date <- as.Date(pfts_wash1$Date.of.Encounter.x,  "%m/%d/%Y")
-pfts_wash1$diff <-as.numeric(pfts_wash1$Date - pfts_wash1$trik)
-
-pfts_wash2 <- subset(pfts_wash1, abs(diff) <= 365 )
-pfts_wash2$prepost <- ifelse(pfts_wash2$trik > pfts_wash2$Date, "before", "after")
-pfts_wash3 <- pfts_wash2 %>% group_by(CFF.ID) %>% filter(prepost == "before") %>% arrange(diff) %>%  filter(row_number()==1)
-
-analytic_wash <- merge(analytic_wash,pfts_wash3[,c("CFF.ID", "Predicted.Value.for.FVC", "Predicted.Value.for.FEV1")]  )
-
-library(arsenal)
-library(tidyverse)
-
-#getting analytic cohort
-reg1 <- subset(reg, reg$CFRD.yes.1 == 1)
-#there are 30 people with cfrd in the database
-
-a1c1 <- subset(a1c, a1c$Diabetes == 1)
-a1c2 <- subset(a1c1, a1c1$MRN %in% reg1$MRN)
-
-a1c2$trik <- as.Date(a1c2$Earliest.Trikafta.Date.in.PortCF, "%m/%d/%Y")
-
-a1c2$trik <- ifelse(a1c2$MRN =="1053930", "12/1/2019", 
-                    ifelse(a1c2$MRN =="1695512", "12/12/2019", 
-                           ifelse(a1c2$MRN =="815274", "12/13/2019", 
-                                  ifelse(a1c2$MRN =="899717", "2/14/2020", as.character(a1c2$Earliest.Trikafta.Date.in.PortCF)))))
-
-a1c2$trik <- as.Date(a1c2$trik, "%m/%d/%Y")
-
-a1c2$test <- as.Date(a1c2$Date.of.A1C.or.OGTT,  "%m/%d/%Y")
-a1c2$prepost <- ifelse(a1c2$trik > a1c2$test, "before", "after")
-
-#who has before and after
-a1c3 <- a1c2 %>% group_by(MRN, prepost) %>% count()
-a1c4 <- a1c3 %>% group_by(MRN) %>% count()
-
-a1c5 <- subset(a1c2, a1c2$MRN %in% a1c4[a1c4$n == 2,]$MRN)
-
-#gcm cleaning
-
-out <- str_split(cgm$subject_id, "_")
-cgm$ID <- do.call(rbind, out)[,1]
-cgm$time <- tolower(do.call(rbind, out)[,2])
-
-cgm0 <- subset(cgm, cgm$ID %in% reg1$CFF.ID)
-
-cgm1 <- cgm0 %>% group_by(ID, time) %>% count()
-cgm2 <- cgm1 %>% group_by(ID) %>% count()
-
-#weight
-
-weight1 <- subset(weight, weight$MRN %in% reg1$MRN)
-
-reg1$a1c <- ifelse(reg1$MRN %in% a1c5$MRN, "Yes", "No")
-reg1$cgm <- ifelse(reg1$CFF.ID %in% cgm2$ID, "Yes", "No")
-
-######################
-#table1
-reg1$trek <-  as.Date(reg1$Earliest.Trikafta.Date.in.PortCF, "%m/%d/%Y")
-weight1$trek <- as.Date(weight1$Earliest.Trikafta.Date.in.PCF, "%m/%d/%Y")
-weight1$pft <- as.Date(weight1$PFT.Date, "%m/%d/%Y")
-weight1$diff <- weight1$pft-weight1$trek
-
-weight2 <- weight1  %>% arrange(as.numeric(diff))%>% group_by(MRN) %>%  filter(row_number()==1)
-
-reg2 <- merge(reg1, weight2, by = "MRN", all.x = T)
-
-reg2$age <- as.numeric(reg2$pft - as.Date(reg2$DOB, "%m/%d/%Y"))/365.25
-library(PAutilities)
-reg2$Sex_MM <- ifelse(reg2$Sex == "Male", "M", "F")
-
-library(AGD)
-
-bmi_levels <- c("Underweight", "Healthy weight", "Overweight", 
-                "Obesity", "Severe obesity")
-bmi_class <- c("Underweight", "Healthy weight", "Overweight","Class 1", "Class 2", "Class 3")
-reg2$agemos <- reg2$age*12
-
-reg2$Sex <- as.factor(reg2$Sex_MM)
-levels(reg2$Sex) <- c("F", "M")
-reg2$bmi <- reg2$BMI
-reg3 <- reg2 %>%
-  # the CDC growth charts are typically used for children 2 years or older
-  # filter(agemos >= 24) %>%
-  mutate(bmiz =  AGD::y2z(y = bmi, x = age, sex = Sex, ref = cdc.bmi),
-         bmipct = pnorm(bmiz),
-         z = qnorm(.95),
-         p95 = AGD::z2y(z = z, x= age, sex = Sex, ref = cdc.bmi),
-         # percentage of the 95th percentile
-         bmipct95 = bmi/p95,
-         # bmi category
-         bmicat = factor(case_when(bmipct < .05 ~ "Underweight",
-                                   bmipct < .85 ~ "Healthy weight",
-                                   bmipct < .95 ~ "Overweight",
-                                   bmipct95 < 1.2 ~ "Class 1",
-                                   bmipct95 < 1.4 ~ "Class 2",
-                                   bmipct95 >= 1.4 ~ "Class 3"),
-                         levels = bmi_class))
-
-#fev
-
-#BMI 
-reg2_BMI <- merge(reg1, weight1, by = "MRN", all.x = T)
-
-reg2_BMI$age <- as.numeric(reg2_BMI$pft - as.Date(reg2_BMI$DOB, "%m/%d/%Y"))/365.25
-library(PAutilities)
-reg2_BMI$Sex_MM <- ifelse(reg2_BMI$Sex == "Male", "M", "F")
-
-library(AGD)
-
-bmi_levels <- c("Underweight", "Healthy weight", "Overweight", 
-                "Obesity", "Severe obesity")
-bmi_class <- c("Underweight", "Healthy weight", "Overweight","Class 1", "Class 2", "Class 3")
-reg2_BMI$agemos <- reg2_BMI$age*12
-
-#
-reg2$class2 <- ifelse(reg2$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-                                             '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-                                             '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-                                             '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
-                      ifelse(reg2$Genotypes2 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
-                             "II", 
-                             ifelse(reg2$Genotypes2 %in% c('G551D', 'G551S', 'S549N'), "III", 
-                                    ifelse(reg2$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
-                                           "IV", 
-                                           ifelse(reg2$Genotypes2 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-                                                                         'P574H', '3272-26A->G5'), "V" , "")))))
-
-reg2$severity2 <- ifelse(reg2$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-                                                '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-                                                '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-                                                '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
-                                                'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
-                                                '2183delAA->G', '2184insA', 'S434X',
-                                                '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
-                                                'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
-                                                "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
-                         "Severe", 
-                         ifelse(reg2$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
-                                                       'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-                                                       'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
-                                                       'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
-                                                       "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
-
-                                ifelse(reg2$Genotypes2 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
-                                                                'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
-                                                                'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
-                                                                'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
-                                                                "R1066H", "P205S"), 
-                                       "Unk","")))
-
-reg2$class1 <- ifelse(reg2$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-                                             '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-                                             '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-                                             '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
-                      ifelse(reg2$Genotypes1 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
-                             "II", 
-                             ifelse(reg2$Genotypes1 %in% c('G551D', 'G551S', 'S549N'), "III", 
-                                    ifelse(reg2$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
-                                           "IV", 
-                                           ifelse(reg2$Genotypes1 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-                                                                         'P574H', '3272-26A->G5'), "V" , "")))))   
-reg2$severity1 <- ifelse(reg2$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-                                                '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-                                                '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-                                                '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
-                                                'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
-                                                '2183delAA->G', '2184insA', 'S434X',
-                                                '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
-                                                'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
-                                                "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
-                         "Severe", 
-                         ifelse(reg2$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
-                                                       'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-                                                       'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
-                                                       'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
-                                                       "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
-
-                                ifelse(reg2$Genotypes1 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
-                                                                'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
-                                                                'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
-                                                                'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
-                                                                "R1066H", "P205S"), 
-                                       "Unk","")))
-
-reg2$genoRisk <- ifelse(reg2$severity1 == "Mild" | reg2$severity2 == "Mild", "Low", 
-                        ifelse(reg2$severity1 == "Severe" & reg2$severity2 == "Severe", "High", ""))
-
-reg2_BMI$Sex <- as.factor(reg2_BMI$Sex_MM)
-levels(reg2_BMI$Sex) <- c("F", "M")
-reg2_BMI$bmi <- reg2_BMI$BMI
-
-reg2$ppFEV1 <- pctpred_GLI(reg2$age, reg2$Height/100, gender = ifelse(reg2$Sex_MM == "M", 1, 2) , 
-                           ethnicity = 1, 
-                           FEV1 = reg2$FEV1)
-
-reg2$ppFVC <- pctpred_GLI(reg2$age, reg2$Height/100, gender = ifelse(reg2$Sex_MM == "M", 1, 2) , 
-                          ethnicity = 1, 
-                          FVC = reg2$FVC)
-
-reg3_BMI <- reg2_BMI %>%
-  # the CDC growth charts are typically used for children 2 years or older
-  # filter(agemos >= 24) %>%
-  mutate(bmiz =  AGD::y2z(y = bmi, x = age, sex = Sex, ref = cdc.bmi),
-         bmipct = pnorm(bmiz),
-         z = qnorm(.95),
-         p95 = AGD::z2y(z = z, x= age, sex = Sex, ref = cdc.bmi),
-         # percentage of the 95th percentile
-         bmipct95 = bmi/p95,
-         # bmi category
-         bmicat = factor(case_when(bmipct < .05 ~ "Underweight",
-                                   bmipct < .85 ~ "Healthy weight",
-                                   bmipct < .95 ~ "Overweight",
-                                   bmipct95 < 1.2 ~ "Class 1",
-                                   bmipct95 < 1.4 ~ "Class 2",
-                                   bmipct95 >= 1.4 ~ "Class 3"),
-                         levels = bmi_class))
-
-#fev
-
-library(rspiro)
-
-reg3$ppFEV1 <- pctpred_GLI(reg3$age, reg3$Height/100, gender = ifelse(reg3$Sex_MM == "M", 1, 2) , 
-                           ethnicity = 1, 
-                           FEV1 = reg3$FEV1)
-
-reg3$ppFVC <- pctpred_GLI(reg3$age, reg3$Height/100, gender = ifelse(reg3$Sex_MM == "M", 1, 2) , 
-                          ethnicity = 1, 
-                          FVC = reg3$FVC)
-
-#
-reg3$CFRD <- as.Date(reg3$CFRD.Diagnosis.Date, "%m/%d/%Y")
-reg3$CFRD_duration <- as.numeric(reg3$trek.x-reg3$CFRD)/365.25
-
-#
-reg3$class2 <- ifelse(reg3$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-                                             '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-                                             '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-                                             '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
-                      ifelse(reg3$Genotypes2 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
-                             "II", 
-                             ifelse(reg3$Genotypes2 %in% c('G551D', 'G551S', 'S549N'), "III", 
-                                    ifelse(reg3$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
-                                           "IV", 
-                                           ifelse(reg3$Genotypes2 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-                                                                         'P574H', '3272-26A->G5'), "V" , "")))))
-
-reg3$severity2 <- ifelse(reg3$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-                                                '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-                                                '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-                                                '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
-                                                'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
-                                                '2183delAA->G', '2184insA', 'S434X',
-                                                '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
-                                                'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
-                                                "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
-                         "Severe", 
-                         ifelse(reg3$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
-                                                       'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-                                                       'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
-                                                       'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
-                                                       "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
-
-                                ifelse(reg3$Genotypes2 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
-                                                                'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
-                                                                'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
-                                                                'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
-                                                                "R1066H", "P205S"), 
-                                       "Unk","")))
-
-reg3$class1 <- ifelse(reg3$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-                                             '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-                                             '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-                                             '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
-                      ifelse(reg3$Genotypes1 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
-                             "II", 
-                             ifelse(reg3$Genotypes1 %in% c('G551D', 'G551S', 'S549N'), "III", 
-                                    ifelse(reg3$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
-                                           "IV", 
-                                           ifelse(reg3$Genotypes1 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-                                                                         'P574H', '3272-26A->G5'), "V" , "")))))   
-reg3$severity1 <- ifelse(reg3$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-                                                '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-                                                '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-                                                '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
-                                                'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
-                                                '2183delAA->G', '2184insA', 'S434X',
-                                                '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
-                                                'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
-                                                "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
-                         "Severe", 
-                         ifelse(reg3$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
-                                                       'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-                                                       'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
-                                                       'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
-                                                       "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
-
-                                ifelse(reg3$Genotypes1 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
-                                                                'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
-                                                                'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
-                                                                'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
-                                                                "R1066H", "P205S"), 
-                                       "Unk","")))
-
-reg3$genoRisk <- ifelse(reg3$severity1 == "Mild" | reg3$severity2 == "Mild", "Low", 
-                        ifelse(reg3$severity1 == "Severe" & reg3$severity2 == "Severe", "High", ""))
-
-# 
-# tab1 <- tableby(~age+bmipct + Weight + Sex + ppFEV1 + ppFVC + CFRD_duration +
-#                   G.tube.in.past.12.months + Pancreatic.Status +
-#                   genoRisk , data=reg3)
-
-reg3$location <- rep("reg3", nrow(reg3))
-
-reg_tab1 <- reg3[,c("CFF.ID", "age", "bmipct", "Weight", "Sex", "ppFEV1", "ppFVC", "G.tube.in.past.12.months", "Pancreatic.Status", "genoRisk",
-                    "location")]
-
-reg_tab1$BMI.Value <- rep(NA, nrow(reg_tab1))
-reg_tab1$age_category <- rep("Pediatric", nrow(reg_tab1))
-
-analytic_montana$bmipct <- analytic_montana$BMI.Percentile/100
-analytic_montana$Sex <- ifelse(analytic_montana$Gender == "Female", "F", "M")
-analytic_montana$Weight <- analytic_montana$Weight.of.Patient 
-analytic_montana$ppFEV1 <- analytic_montana$Percent.predicted.relative.to.predicted.FEV1..value
-analytic_montana$ppFVC <- analytic_montana$Percent.predicted.relative.to.predicted.FVC.value
-analytic_montana$G.tube.in.past.12.months <- analytic_montana$gtube 
-analytic_montana$Pancreatic.Status <- analytic_montana$is
-analytic_montana$genoRisk <- analytic_montana$genotype
-
-analytic_montana$location <- rep("Montana", nrow(analytic_montana))
-
-analytic_montana$age_category <- ifelse(analytic_montana$CFF.ID %in% demo_kids$CFF.ID, "Pediatric", "Adult")
-
-analytic_montana1 <- analytic_montana[,c("CFF.ID", "age", "bmipct", "Weight", "Sex", "ppFEV1", "ppFVC", "G.tube.in.past.12.months", "Pancreatic.Status", "genoRisk",
-                                         "location", "BMI.Value", "age_category")]
-
-analytic_wash$bmipct <- analytic_wash$BMI.Percentile/100
-analytic_wash$age <- analytic_wash$Encounter.Age
-analytic_wash$BMI.Value <- rep(NA, nrow(analytic_wash))
-analytic_wash$Weight <- rep(NA, nrow(analytic_wash))
-analytic_wash$ppFEV1 <- analytic_wash$Predicted.Value.for.FEV1
-analytic_wash$ppFVC <- analytic_wash$Predicted.Value.for.FVC
-analytic_wash$G.tube.in.past.12.months <- analytic_wash$gt
-analytic_wash$Pancreatic.Status <- analytic_wash$is
-analytic_wash$genoRisk <- analytic_wash$genotype
-analytic_wash$location <- rep("Washington", nrow(analytic_wash))
-analytic_wash$age_category <- rep("Pediatric", nrow(analytic_wash))
-analytic_wash$Earliest.Trikafta.Date.in.PortCF <-analytic_wash$trik
-
-analytic_wash1 <- analytic_wash[,c("CFF.ID", "age", "bmipct", "Weight", "Sex", "ppFEV1", "ppFVC", "G.tube.in.past.12.months", 
-                                   "Pancreatic.Status", "genoRisk",
-                                   "location", "BMI.Value", "age_category")]
-
-#
-
-uofm <- read.csv("./Data from UofM/Moheet_Pre_Post_Trikafta_FinalData_20210914.csv")
-analytic_uofm <- subset(uofm, uofm$CFRD.Status == "CFRD with or without fasting hyperglycemia")
-analytic_uofm$age <- analytic_uofm$age.at.Trikafta.start
-
-bmi_levels <- c("Underweight", "Healthy weight", "Overweight", 
-                "Obesity", "Severe obesity")
-bmi_class <- c("Underweight", "Healthy weight", "Overweight","Class 1", "Class 2", "Class 3")
-analytic_uofm$agemos <- analytic_uofm$age*12
-
-analytic_uofm$Sex <- as.factor(analytic_uofm$Gender)
-levels(analytic_uofm$Sex) <- c("F", "M")
-analytic_uofm$bmi <- analytic_uofm$BMI.calc
-analytic_uofm$bmi_post <- analytic_uofm$BMI.calc.1
-analytic_uofm2 <- analytic_uofm %>%
-  # the CDC growth charts are typically used for children 2 years or older
-  # filter(agemos >= 24) %>%
-  mutate(bmiz =  AGD::y2z(y = bmi, x = age, sex = Sex, ref = cdc.bmi),
-         bmiz_post =  AGD::y2z(y = bmi_post, x = age, sex = Sex, ref = cdc.bmi),
-         bmipct = pnorm(bmiz),
-         bmipct_post = pnorm(bmiz_post))
-
-analytic_uofm2$ppFEV1 <- pctpred_GLI(analytic_uofm2$age, analytic_uofm2$Height..cm./100, gender = ifelse(analytic_uofm2$Sex== "M", 1, 2) , 
-                                     ethnicity = 1, 
-                                     FEV1 = analytic_uofm2$FEV1)
-
-analytic_uofm2$ppFEV1_post <- pctpred_GLI(analytic_uofm2$age, as.numeric(analytic_uofm2$Height..cm..1)/100, gender = ifelse(analytic_uofm2$Sex== "M", 1, 2) , 
-                                          ethnicity = 1, 
-                                          FEV1 = analytic_uofm2$FEV1.1)
-
-analytic_uofm2$ppFVC <- pctpred_GLI(analytic_uofm2$age, analytic_uofm2$Height..cm./100, gender = ifelse(analytic_uofm2$Sex == "M", 1, 2) , 
-                                    ethnicity = 1, 
-                                    FVC = analytic_uofm2$FVC)
-
-analytic_uofm2$ppFVC_post <- pctpred_GLI(analytic_uofm2$age, as.numeric(analytic_uofm2$Height..cm..1)/100, gender = ifelse(analytic_uofm2$Sex== "M", 1, 2) , 
-                                         ethnicity = 1, 
-                                         FEV1 = analytic_uofm2$FVC.1)
-
-analytic_uofm2$Weight <- analytic_uofm2$Weight..kg.
-
-analytic_uofm2$G.tube.in.past.12.months <- analytic_uofm2$Tube.Feeding
-
-analytic_uofm2$age_category <- ifelse(analytic_uofm2$age <= 20 ,"Pediatric", "Adult")
-analytic_uofm2$location <- rep("Minnesota", nrow(analytic_uofm2))
-analytic_uofm2$Pancreatic.Status <- ifelse(analytic_uofm2$Pancreatic.Sufficiency..on.enzymes.. == "Yes", "Insufficient", "Not")
-
-analytic_uofm2$Genotypes1 <- analytic_uofm2$Genotype1
-
-analytic_uofm2$Genotypes2 <- analytic_uofm2$Genotype2
-
-analytic_uofm2$class2 <- ifelse(analytic_uofm2$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-                                                                 '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-                                                                 '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-                                                                 '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G', '2183AA->G'), "I", 
-                                ifelse(analytic_uofm2$Genotypes2 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
-                                       "II", 
-                                       ifelse(analytic_uofm2$Genotypes2 %in% c('G551D', 'G551S', 'S549N'), "III", 
-                                              ifelse(analytic_uofm2$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
-                                                     "IV", 
-                                                     ifelse(analytic_uofm2$Genotypes2 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-                                                                                             'P574H', '3272-26A->G5'), "V" , "")))))
-
-analytic_uofm2$severity2 <- ifelse(analytic_uofm2$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-                                                                    '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-                                                                    '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-                                                                    '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
-                                                                    'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
-                                                                    '2183delAA->G', '2184insA', 'S434X',
-                                                                    '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
-                                                                    'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
-                                                                    "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
-                                   "Severe", 
-                                   ifelse(analytic_uofm2$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
-                                                                           'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-                                                                           'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
-                                                                           'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
-                                                                           "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
-
-                                          ifelse(analytic_uofm2$Genotypes2 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
-                                                                                    'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
-                                                                                    'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
-                                                                                    'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
-                                                                                    "R1066H", "P205S"), 
-                                                 "Unk","")))
-
-analytic_uofm2$class1 <- ifelse(analytic_uofm2$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-                                                                 '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-                                                                 '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-                                                                 '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
-                                ifelse(analytic_uofm2$Genotypes1 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
-                                       "II", 
-                                       ifelse(analytic_uofm2$Genotypes1 %in% c('G551D', 'G551S', 'S549N'), "III", 
-                                              ifelse(analytic_uofm2$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
-                                                     "IV", 
-                                                     ifelse(analytic_uofm2$Genotypes1 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-                                                                                             'P574H', '3272-26A->G5'), "V" , "")))))   
-analytic_uofm2$severity1 <- ifelse(analytic_uofm2$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-                                                                    '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-                                                                    '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-                                                                    '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
-                                                                    'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
-                                                                    '2183delAA->G', '2184insA', 'S434X',
-                                                                    '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
-                                                                    'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
-                                                                    "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
-                                   "Severe", 
-                                   ifelse(analytic_uofm2$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
-                                                                           'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-                                                                           'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
-                                                                           'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
-                                                                           "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,
-
-                                          ifelse(analytic_uofm2$Genotypes1 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
-                                                                                    'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
-                                                                                    'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
-                                                                                    'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
-                                                                                    "R1066H", "P205S"), 
-                                                 "Unk","")))
-
-analytic_uofm2$genoRisk <- ifelse(analytic_uofm2$severity1 == "Mild" | analytic_uofm2$severity2 == "Mild", "Low", 
-                                  ifelse(analytic_uofm2$severity1 == "Severe" & analytic_uofm2$severity2 == "Severe", "High", ""))
-
-analytic_uofm2$genotype <- ifelse(analytic_uofm2$CFF.ID %in% analytic_uofm2[analytic_uofm2$genoRisk == "High", ]$CFF.ID, "High", 
-                                  ifelse(analytic_uofm2$CFF.ID %in% analytic_uofm2[analytic_uofm2$genoRisk == "Low", ]$CFF.ID, "Low",   ""   ))
-analytic_uofm2$bmipct <- ifelse(analytic_uofm2$age <=  20, analytic_uofm2$bmipct , NA)
-
-analytic_uofm2$BMI.Value <- analytic_uofm2$bmi
-analytic_uofm2_2 <- subset(analytic_uofm2, analytic_uofm2$Encounter....of.Days.POST.Trikafta.start <= 365 &
-                             analytic_uofm2$Encounter....of.Days.PRE.Trikafta.Start <= 365)
-
-analytic_uofm3 <- analytic_uofm2_2[,c("CFF.ID", "age", "bmipct", "Weight", "Sex", "ppFEV1", "ppFVC", "G.tube.in.past.12.months", 
-                                      "Pancreatic.Status", "genoRisk",
-                                      "location", "BMI.Value", "age_category")]
-
-analytic_cm <- rbind(reg_tab1 , analytic_montana1,analytic_wash1,analytic_uofm3 )
-
-tab1 <- tableby(~age+bmipct + Weight + Sex + ppFEV1 + ppFVC  +
-                  G.tube.in.past.12.months + Pancreatic.Status +
-                  genoRisk + location+BMI.Value + age_category, data=analytic_cm)
-
-#table 2
-
-#need a1c
-
-#subsetfor u of m
-
-pre_minn <- analytic_uofm2_2[,c("CFF.ID", "A1C....of.Days.PRE.Trikafta.Start" , "A1C")]
-
-pre_minn$diff <- as.numeric(pre_minn$A1C....of.Days.PRE.Trikafta.Start)
-pre_minn$prepost <- rep("before", nrow(pre_minn))
-pre_minn$Hgb.A1C <- pre_minn$A1C
-pre_minn$trik <- rep(NA, nrow(pre_minn))
-pre_minn$test <- rep(NA, nrow(pre_minn))
-
-pre_minn1 <- pre_minn[,c("CFF.ID", "trik", "test", "prepost", "diff", "Hgb.A1C")]
-
-post_minn <- analytic_uofm2_2[,c("CFF.ID", "A1C....of.Days.POST.Trikafta.start", "A1C.1" )]
-post_minn$diff <- as.numeric(post_minn$A1C....of.Days.POST.Trikafta.start)
-post_minn$prepost <- rep("after", nrow(post_minn))
-post_minn$Hgb.A1C <- post_minn$A1C.1
-post_minn$trik <- rep(NA, nrow(post_minn))
-post_minn$test <- rep(NA, nrow(post_minn))
-post_minn1 <- post_minn[,c("CFF.ID", "trik", "test", "prepost", "diff", "Hgb.A1C")]
-
-minn <- subset(rbind(post_minn1, pre_minn1), abs(rbind(post_minn1, pre_minn1)$diff) <= 365)
-
-a1c5$diff <- as.numeric(a1c5$trik - a1c5$test)
-a1c6 <- subset(a1c5, abs(a1c5$diff) <= 365)
-
-#combining all into one
-
-a1c6_big <- a1c6[,c("CFF.ID", "trik", "test", "prepost", "diff", "Hgb.A1C")]
-reg_montana2$Hgb.A1C <- reg_montana2$Hgb.A1C.value...
-reg_montana2$trik <- ifelse(reg_montana2$CFF.ID == "1044621", "2019-12-02",
-                            ifelse(reg_montana2$CFF.ID == "1008380", "2020-1-1", as.character(reg_montana2$trik)))
-reg_montana2_big <- reg_montana2[,c("CFF.ID", "trik", "test", "prepost", "diff", "Hgb.A1C")]
-
-a1c_wash2$test <- a1c_wash2$pft
-a1c_wash2$Hgb.A1C <- a1c_wash2$Hgb.A1C.value...
-a1c_wash2$prepost <- ifelse(a1c_wash2$trik < a1c_wash2$test, "before", "after")
-a1c_wash2_big <- a1c_wash2[,c("CFF.ID", "trik", "test", "prepost", "diff", "Hgb.A1C")]
-
-a1c6_big2 <- rbind(a1c6_big,reg_montana2_big, a1c_wash2_big, minn)
-
-a1c7 <- a1c6_big2  %>% group_by(CFF.ID, prepost) %>%  count()
-
-mycontrols  <- tableby.control(test=FALSE, total=FALSE,
-                               numeric.test="kwt", cat.test="chisq",
-                               numeric.stats=c("N", "median", "q1q3"),
-                               cat.stats=c("countpct"),
-                               stats.labels=list(N='Count', median='Median', q1q3='Q1,Q3'))
-tab2 <- tableby(prepost ~ n, control=mycontrols, data=a1c7)
-
-before <- subset(a1c7, a1c7$prepost == "before")
-before$pre <- before$n
-
-after <- subset(a1c7, a1c7$prepost == "after")
-after$post <- after$n
-
-#number of values pre/post
-#row 1 in table 2
-a <- merge(before, after, by = "CFF.ID", all.y = T)
-
-#test if difference
-
-#merge back in the a1c for pre and post 
-before <- subset(a1c6_big2, a1c6_big2$prepost == "before")
-before1 <- before %>% arrange(diff) %>% group_by(CFF.ID) %>% filter(row_number()==1)
-
-#time before
-before_max <- before %>% group_by(CFF.ID) %>% summarize(time = max(diff)/30.417)
-
-after <- subset(a1c6_big2, a1c6_big2$prepost == "after")
-after1 <- after %>% arrange(desc(diff)) %>% group_by(CFF.ID) %>% filter(row_number()==1)
-
-after_max <- after %>% group_by(CFF.ID) %>% summarize(time = max(abs(diff))/30.417)
-
-#test this
-
-time <- merge(before_max,after_max,by='CFF.ID',all.y=T)
-
-time <- merge(before1,after1,by='CFF.ID',all.y=T)
-
-temp <- rbind(before1,after1)
-temp$Hgb.A1C <- as.numeric(temp$Hgb.A1C)
-# tab3 <- tableby(prepost~Hgb.A1C, temp , control=mycontrols)
-
-time$diff_a1c <- as.numeric(time$Hgb.A1C.x) - as.numeric(time$Hgb.A1C.y)
-
-#use these numbers for pre/post cgm
-
-time2 <- subset(time, time$CFF.ID %in% cgm$ID)
-
-#DEMOGRAPHCIS
-# tab1 <- tableby(~age+bmipct + Weight + Sex + ppFEV1 + ppFVC  +
-#                   G.tube.in.past.12.months + Pancreatic.Status +
-#                   genoRisk + location+BMI.Value + age_category, data=analytic_cm[analytic_cm$CFF.ID %in% cgm0_small$CFF.ID,])
-#
-
-#insulin
-
-ins <- read.csv("insulindata.csv")
-
-ins2 <- merge(ins, time2[,c("CFF.ID", "trik.x")], by= "CFF.ID", all.x = T)
-
-ins2$Date_MM <- as.Date(ins2$Date, "%m/%d/%Y")
-
-ins2$diff <- as.numeric(ins2$Date_MM - ins2$trik.x)
-ins3 <- subset(ins2, abs(diff) <= 365)
-
-ins3$prepost <- ifelse(ins3$diff <= 0, "Pre", "Post")
-
-ins4 <- ins3 %>% group_by(CFF.ID, prepost) %>% count()
-ins5 <- ins4 %>% group_by(CFF.ID) %>% count()
-
-ins6 <- subset(ins3, ins3$CFF.ID %in% ins5[ins5$n > 1,]$CFF.ID)
-
-ins7 <- ins6 %>% group_by(CFF.ID, prepost) %>% summarise_at("Long.acting.or.total.basal.daily.dose..units.",mean)
-
-ins8 <- merge(ins7[ins7$prepost == "Pre",], ins7[ins7$prepost == "Post",], by = "CFF.ID")
-
-##########################
-#weight data
-#need z score for weight
-library("anthro")
-weightz <- merge(weight1, reg1[,c("MRN", "CFF.ID", "Sex", "DOB")], by = "MRN", all.x=T)
-
-weightz$age <- as.numeric(weightz$pft - as.Date(weightz$DOB, "%m/%d/%Y"))/365.25
-library(PAutilities)
-weightz$Sex_MM <- ifelse(weightz$Sex == "Male", "M", "F")
-
-library(AGD)
-
-bmi_levels <- c("Underweight", "Healthy weight", "Overweight", 
-                "Obesity", "Severe obesity")
-bmi_class <- c("Underweight", "Healthy weight", "Overweight","Class 1", "Class 2", "Class 3")
-weightz$agemos <- weightz$age*12
-
-weightz$Sex <- as.factor(weightz$Sex_MM)
-levels(weightz$Sex) <- c("F", "M")
-weightz$bmi <- weightz$BMI
-weightz2 <- weightz %>%
-  # the CDC growth charts are typically used for children 2 years or older
-  # filter(agemos >= 24) %>%
-  mutate(bmiz =  AGD::y2z(y = bmi, x = age, sex = Sex, ref = cdc.bmi),
-         bmipct = pnorm(bmiz),
-         z = qnorm(.95),
-         p95 = AGD::z2y(z = z, x= age, sex = Sex, ref = cdc.bmi),
-         # percentage of the 95th percentile
-         bmipct95 = bmi/p95,
-         # bmi category
-         bmicat = factor(case_when(bmipct < .05 ~ "Underweight",
-                                   bmipct < .85 ~ "Healthy weight",
-                                   bmipct < .95 ~ "Overweight",
-                                   bmipct95 < 1.2 ~ "Class 1",
-                                   bmipct95 < 1.4 ~ "Class 2",
-                                   bmipct95 >= 1.4 ~ "Class 3"),
-                         levels = bmi_class))
-
-weightz2$ppFEV1 <- pctpred_GLI(weightz2$age, weightz2$Height/100, gender = ifelse(weightz2$Sex_MM == "M", 1, 2) , 
-                               ethnicity = 1, 
-                               FEV1 = weightz2$FEV1)
-
-weightz2$ppFVC <- pctpred_GLI(weightz2$age, weightz2$Height/100, gender = ifelse(weightz2$Sex_MM == "M", 1, 2) , 
-                              ethnicity = 1, 
-                              FVC = weightz2$FVC)
-
-weigh_before <- merge(weightz2, before1[,c("CFF.ID", "test")], by = "CFF.ID", all.x = T)
-weigh_before$weight_closest <- abs(as.numeric(weigh_before$pft-weigh_before$test))
-
-weigh_before1 <- weigh_before %>% group_by(MRN) %>% arrange(weight_closest) %>% filter(row_number()==1)
-weigh_before1$time <- rep("before",nrow(weigh_before1))
-
-weigh_after <- merge(weightz2, after1[,c("CFF.ID", "test")], by = "CFF.ID", all.x = T)
-weigh_after$weight_closest <- abs(as.numeric(weigh_after$pft-weigh_after$test))
-
-weigh_after1 <- weigh_after %>% group_by(MRN) %>% arrange(weight_closest) %>% filter(row_number()==1)
-
-weigh_after1$time <- rep("after",nrow(weigh_after1))
-
-timing <- rbind(weigh_before1,weigh_after1)
-
-tab2 <- tableby(time~Weight+ bmiz + ppFEV1+ppFVC, data=timing)
-timing2 <- merge(weigh_before1,weigh_after1, by = "MRN", all.y=T)
-
-#weight and fev data
-pfts_wash2$ppFVC <- pfts_wash2$Predicted.Value.for.FVC
-pfts_wash2$ppFEV1 <- pfts_wash2$Predicted.Value.for.FEV1
-pfts_wash2$bmipct <- pfts_wash2$BMI.Percentile
-pfts_wash2$Weight <- rep(NA, nrow(pfts_wash2))
-
-pft_montana_3$ppFVC <- pft_montana_3$Percent.predicted.relative.to.predicted.FVC.value
-pft_montana_3$ppFEV1 <- pft_montana_3$Percent.predicted.relative.to.predicted.FEV1..value
-pft_montana_3$bmipct <- pft_montana_3$BMI.Percentile
-pft_montana_3$Weight <- pft_montana_3$Weight.of.Patient
-reg3_BMI$prepost <- ifelse(reg3_BMI$diff < 0, "before","after")
-
-reg3_BMI$ppFEV1 <- pctpred_GLI(reg3_BMI$age, reg3_BMI$Height/100, gender = ifelse(reg3_BMI$Sex_MM == "M", 1, 2) , 
-                               ethnicity = 1, 
-                               FEV1 = reg3_BMI$FEV1)
-
-reg3_BMI$ppFVC <- pctpred_GLI(reg3_BMI$age, reg3_BMI$Height/100, gender = ifelse(reg3_BMI$Sex_MM == "M", 1, 2) , 
-                              ethnicity = 1, 
-                              FVC = reg3_BMI$FVC)
-
-pre_minn1 <- analytic_uofm2_2[,c("CFF.ID", "ppFVC",
-                                 "ppFEV1","bmipct",
-                                 "Weight", "A1C....of.Days.PRE.Trikafta.Start", "bmi")]
-pre_minn1$BMI.Value <- pre_minn1$bmi
-pre_minn1$diff <- as.numeric(pre_minn1$A1C....of.Days.PRE.Trikafta.Start)*-1
-pre_minn1$prepost <- rep("before", nrow(pre_minn1))
-
-post_minn1 <- analytic_uofm2_2[,c("CFF.ID", "ppFVC_post",
-                                  "ppFEV1_post","bmipct_post",
-                                  "Weight..kg..1", "A1C....of.Days.POST.Trikafta.start", 'bmi')]
-
-post_minn1$diff <- as.numeric(post_minn1$A1C....of.Days.POST.Trikafta.start)
-post_minn1$prepost <- rep("after", nrow(post_minn1))
-post_minn1$Weight <- post_minn1$Weight..kg..1
-post_minn1$ppFEV1 <-post_minn1$ppFEV1_post
-post_minn1$ppFVC <-post_minn1$ppFVC_post
-post_minn1$bmipct <- post_minn1$bmipct_post
-post_minn1$BMI.Value <- post_minn1$bmi
-
-reg3_BMI$BMI.Value <- reg3_BMI$BMI
-
-bmidat <- rbind(pfts_wash2[,c("CFF.ID", "ppFVC",
-                              "ppFEV1","bmipct",
-                              "diff","prepost", "Weight", "BMI.Value")],
-
-                pft_montana_3[,c("CFF.ID", "ppFVC",
-                                 "ppFEV1","bmipct",
-                                 "diff","prepost", "Weight", "BMI.Value")],
-
-                reg3_BMI[,c("CFF.ID", "ppFVC",
-                            "ppFEV1","bmipct",
-                            "diff","prepost", "Weight", "BMI.Value")],
-                rbind(post_minn1[c("CFF.ID", "ppFVC",
-                                   "ppFEV1","bmipct",
-                                   "diff","prepost", "Weight", "BMI.Value")],
-                      pre_minn1[c("CFF.ID", "ppFVC",
-                                  "ppFEV1","bmipct",
-                                  "diff","prepost", "Weight", "BMI.Value")]))
-
-#GET row closest to trik date
-
-bmidat_before <- bmidat %>% filter(diff != 0) %>% filter(prepost == "before") %>% 
-  group_by(CFF.ID) %>% arrange(desc(diff))%>% filter(row_number()==1)
-
-bmidat_before$time <- rep("before", nrow(bmidat_before))
-
-bmidat_after <- bmidat %>% filter(diff != 0) %>% filter(prepost == "after") %>% 
-  group_by(CFF.ID) %>% arrange(diff)%>% filter(row_number()==1)
-bmidat_after$time <- rep("post", nrow(bmidat_after))
-
-bmidat2 <- rbind(bmidat_before, bmidat_after)
-
-bmidat3 <- bmidat2 %>% group_by(CFF.ID) %>% count()
-bmidat4 <- subset(bmidat2, bmidat2$CFF.ID %in% bmidat3[bmidat3$n == 2,]$CFF.ID)
-
-bmidat4$ppFEV1 <- as.numeric(bmidat4$ppFEV1)
-bmidat4$ppFVC <- as.numeric(bmidat4$ppFVC)
-bmidat4$bmipct <- as.numeric(bmidat4$bmipct)
-bmidat4$Weight <- as.numeric(bmidat4$Weight)
-
-tab2b<- tableby(prepost~ppFVC + ppFEV1 + bmipct + Weight + BMI.Value, 
-                data=bmidat4[bmidat4$CFF.ID %in% time[!is.na(time$prepost.x),]$CFF.ID,])
-tab2b<- tableby(prepost~ppFVC + ppFEV1 + bmipct + Weight + BMI.Value, 
-                data=bmidat4[bmidat4$CFF.ID %in% a$CFF.ID,])
-
-bmidat4 <- merge(bmidat_before, bmidat_after, by = "CFF.ID")
-
-#CGM data
-
-out <- str_split(cgm$subject_id, "_")
-cgm$ID <- do.call(rbind, out)[,1]
-cgm$time <- tolower(do.call(rbind, out)[,2])
-
-cgm0 <- subset(cgm, cgm$ID %in% a1c6_big2$CFF.ID)
-
-cgm_temp <- merge(cgm, a1c6_big2, by.x="ID", by.y="CFF.ID", all.x=T)
-
-cgm1 <- cgm0 %>% group_by(ID, time) %>% count()
-cgm2 <- cgm1 %>% group_by(ID) %>% count()
-
-#pull first post and then do an overall
-
-cgm0$date <- as.Date(cgm0$date_cgm_placement)
-
-cgm02 <- cgm0 %>% filter(time=="post") %>%  group_by(ID) %>% 
-  arrange(date) %>% filter(row_number()==1)
-
-cgm0_small <- rbind(cgm0[cgm0$time == "pre",],cgm02)
-mycontrols  <- tableby.control(numeric.stats=c("medianq1q3"),
-                               stats.labels=list(medianq1q3="Median (Q1, Q3)"))
-
-a1c6_big2_tim <- a1c6_big2 %>% group_by(CFF.ID) %>% filter(row_number()==1)
-cgm0_time <- merge(cgm0_small, a1c6_big2_tim[,c("CFF.ID", "trik" )], by.y = "CFF.ID", by.x = "ID", all.x = T)
-
-cgm0_time$cgm_date <- as.Date(cgm0_time$date_cgm_placement, "%Y-%m-%d")
-cgm0_time$diff <- cgm0_time$date_cgm - as.Date(cgm0_time$trik, "%Y-%m-%d")
-
-tab3_extra <- tableby(time ~ diff, cgm0_time, control=mycontrols)
-
-tab3 <- tableby(time~ average_sensor
-                + min_sensor
-                + max_sensor
-                + percent_time_over_120
-                + percent_time_over_140
-                + percent_time_over_200
-                + avg_excur_over_140_per_day
-                + avg_excur_over_200_per_day
-                + percent_time_under_60
-                + percent_time_under_70
-                + average_auc_per_day
-                + standard_deviation
-                + r_mage, data=cgm0_small, control=mycontrols)
-tab3 <- tableby(~average_sensor
-                + min_sensor
-                + max_sensor
-                + percent_time_over_120
-                + percent_time_over_140
-                + percent_time_over_200
-                + avg_excur_over_140_per_day
-                + avg_excur_over_200_per_day
-                + percent_time_under_60
-                + percent_time_under_70
-                + average_auc_per_day
-                + standard_deviation
-                + r_mage, data=cgm0[cgm0$time == "post",]
-                , control=mycontrols)
-
-temp <-merge(cgm0_small[cgm0_small$time=="pre",],
-             cgm0_small[cgm0_small$time=="post",], by = "ID")
-
-analytic_cm$ppFVC <- as.numeric(analytic_cm$ppFVC)
-
-tab1 <- tableby(~age+bmipct + Weight + Sex + ppFEV1 + ppFVC  +
-                  G.tube.in.past.12.months + Pancreatic.Status +
-                  genoRisk + location+BMI.Value + age_category, 
-                data=analytic_cm[analytic_cm$CFF.ID %in% cgm0_small$ID,])
-summary(tab1)
+# Montana
+montana = read.csv("./Data_Raw/Data From Montana/Copy of Paper patient population glycemic (+hofer).csv",na.strings = "")
+demo_adult = read.csv("./Data_Raw/Data From Montana/demo_adults.csv",na.strings = "")
+demo_kids = read.csv("./Data_Raw/Data From Montana/demo_kids.csv",na.strings = "")
+demo_kids$Date.Of.Birth..mm.dd.yyyy. = NULL
+demo_kids = demo_kids[,colnames(demo_adult)]
+demo = rbind(demo_adult,demo_kids)
+# Get within a year of Trikafta start
+montana$time_diff = mdy(montana$Trikafta.start.date) - mdy(montana$Date.of.Encounter)
+montana = montana %>% filter(abs(montana$time_diff)<=365)
+# Add demographics
+montana = left_join(montana,demo,by = "CFF.ID")
+# PFTs
+pft_adult <- read.csv("./Data_Raw/Data From Montana/Copy of PFTs growth adult (+hofer).csv")
+pft_kids <- read.csv("./Data_Raw/Data From Montana/PFTs kids.csv")
+pft_adult = pft_adult[,colnames(pft_kids)]
+pfts = rbind(pft_adult,pft_kids)
+montana = full_join(montana,pfts,by = c("CFF.ID", "Date.of.Encounter"))
 ```

### 2022-01-19 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
Update pre_post_trikafta.Rmd
functions for averaging by pre and post

@@
 closest_dates = function(df,start = "Start",date = "Date"){
+  # Remove MRN
+  df$MRN = NULL
   # Start and date columns must be formated for subtraction (ideally with lubridate)
@@
   # Delete unnecessary columns
-  df = df %>% select(-all_of(c(start,date)))
+  df = df %>% select(-all_of(c(start,date))) %>% select(CFF.ID,Timepoint,everything())
   # Return
@@
   return(df)
+} 
+pre_post_average = function(df,start = "Start",date = "Date"){
+  # Remove MRN
+  df$MRN = NULL
+  # Start and date columns must be formated for subtraction (ideally with lubridate)
+  df$Timepoint = as.numeric(df[,date] - df[,start])
+  # Columns to exclude
+  excluded_vars = c(start,date,"CFF.ID","Timepoint","sign(Timepoint)")
+  # Remove visits on start date or > 1 year out
+  df = df %>% filter(Timepoint !=0 & abs(Timepoint) <= 365) %>% 
+    # Get closest visit before and closest after
+    group_by(CFF.ID,sign(Timepoint)) %>%
+    summarise(across(-any_of(excluded_vars), ~mean(.x,na.rm = T)),.groups = "drop") %>%
+    # Format
+    rename(Timepoint = `sign(Timepoint)`) %>%
+    mutate_all(~replace(., is.nan(.), NA)) # Replace NaN
+  # Timepoint as factor
+  df$Timepoint = factor(df$Timepoint,labels = c("Before","After"))
+  # Select
+  df = df %>% select(CFF.ID,Timepoint,everything())
+  # Return
+  df = data.frame(df)
 }

### 2022-01-20 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
Update pre_post_trikafta.Rmd
started piecewise regression

@@
 ```
+
+# Spline regression
+
+```{r}
+a1c = read.csv("./Data_Raw/Data From CHCO/Copy of A1C and OGTTs Before and After Trikafta_cc modified 8.25.20.csv")
+a1c$Earliest.Trikafta.Date.in.PortCF = mdy(a1c$Earliest.Trikafta.Date.in.PortCF)
+a1c$Date.of.A1C.or.OGTT = mdy(a1c$Date.of.A1C.or.OGTT)
+a1c = a1c %>% 
+  mutate(time = as.numeric(difftime(Date.of.A1C.or.OGTT,
+                                    Earliest.Trikafta.Date.in.PortCF,
+                                    units = "days"))) %>%
+  filter(!is.na(Hgb.A1C),abs(time)<=365)
+a1c$spline = ifelse(a1c$time>0,1,0)
+a1c$X = a1c$time * a1c$spline
+# Segmented model, knot at 0
+smod = lme(Hgb.A1C ~ time + X,random = ~1|MRN,data = a1c,na.action = na.omit)
+# Plot
+# Segments
+xs1 = c(min(a1c$time),0)
+beta1 = summary(smod)$tTable[1:2]
+ys1 = cbind(1, xs1) %*% beta1
+
+xs2 = c(0,max(a1c$time))
+beta2 = c(summary(smod)$tTable[1],sum(summary(smod)$tTable[2:3]))
+ys2 = cbind(1,xs2) %*% beta2
+
+p = ggplot(a1c,aes(x = time,y = Hgb.A1C)) + 
+  geom_point() + 
+  geom_segment(aes(x = xs1[1], xend = xs1[2], y = ys1[1], yend = ys1[2]),color = "blue") +
+  geom_segment(aes(x = xs2[1], xend = xs2[2], y = ys2[1], yend = ys2[2]),color = "blue") +
+  xlab("Days from Trikafta Start") + ylab("HbA1c") + theme_bw()
+p
+```

### 2022-01-26 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
Create pre_post_trikafta.Rmd

--- 2022-01-19 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
@@
 ---
-title: "Pre/Post-Trikafta"
-author: "Tim Vigers"
+title: ""
+author: "Tim Vigers & Laura Pyle"
 date: "`r format(Sys.time(), '%d %B %Y')`"
@@
 library(tidyverse)
-library(lubridate)
-library(readxl)
-library(PAutilities)
-library(AGD)
-library(rspiro)
-library(anthro)
+library(performance)
 library(knitr)
 knitr::opts_chunk$set(echo = FALSE)
-knitr::opts_knit$set(root.dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/Prepost Triakfta")
+if(Sys.info()["sysname"] == "Windows"){
+  home_dir = "S://PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan"
+} else if (Sys.info()["sysname"] == "Linux"){
+  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan"
+} else if (Sys.info()["sysname"] == "Darwin"){
+  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan"
+}
+knitr::opts_knit$set(root.dir = home_dir)
 ```
 
-```{r code from maxie,eval=FALSE}
-# Source Maxie's code, and write final ID lists
-source("~/GitHub/CF/Christine Chan/Pre-post Trikafta/p001+cmw_UPDATED.R")
-ids = data.frame(unique(analytic_cm$CFF.ID))
-colnames(ids) = "CFF.ID"
-write.csv(ids,file = "./Data_Cleaned/final_ids.csv",row.names = F)
+```{r}
+# Import 
+df = read.csv("./Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv")
 ```
-
-```{r closest date function}
-closest_dates = function(df,start = "Start",date = "Date"){
-  # Start and date columns must be formated for subtraction (ideally with lubridate)
-  df$Timepoint = as.numeric(df[,date] - df[,start])
-  # Remove visits on start date or > 1 year out
-  df = df %>% filter(Timepoint !=0 & abs(Timepoint) <= 365) %>% 
-    # Get closest visit before and closest after
-    group_by(CFF.ID,sign(Timepoint)) %>%
-    slice_min(Timepoint * sign(Timepoint),n=1) %>% 
-    # Format
-    ungroup %>% select(-Timepoint) %>% rename(Timepoint = `sign(Timepoint)`)
-  # Timepoint as factor
-  df$Timepoint = factor(df$Timepoint,labels = c("Before","After"))
-  # Delete unnecessary columns
-  df = df %>% select(-all_of(c(start,date)))
-  # Return
-  df = data.frame(df)
-  return(df)
-} 
-```
-
-# Table 1: Demographics at Trikafta Start
-
-```{r}
-# Start dates
-chco_demographics = read.csv("./Data_Cleaned/chco_trikafta_dates.csv",na.strings = "")
-# Fix select dates
-chco_demographics$Earliest.Trikafta.Date.in.PortCF[chco_demographics$MRN =="1053930"] = "12/1/2019"
-chco_demographics$Earliest.Trikafta.Date.in.PortCF[chco_demographics$MRN =="1695512"] = "12/12/2019"
-chco_demographics$Earliest.Trikafta.Date.in.PortCF[chco_demographics$MRN =="815274"] = "12/13/2019"
-chco_demographics$Earliest.Trikafta.Date.in.PortCF[chco_demographics$MRN =="1078408"] = "3/9/2020"
-chco_demographics$Earliest.Trikafta.Date.in.PortCF[chco_demographics$MRN =="899717"] = "2/14/2020"
-chco_demographics$CFF.ID = as.character(chco_demographics$CFF.ID)
-# Diabetes and with start date only
-chco_demographics = chco_demographics %>% 
-  filter(!is.na(Earliest.Trikafta.Date.in.PortCF),CFRD.yes.1 == 1)
-chco_demographics$Earliest.Trikafta.Date.in.PortCF = 
-  mdy(chco_demographics$Earliest.Trikafta.Date.in.PortCF)
-chco_dates = chco_demographics %>% rename(Start = Earliest.Trikafta.Date.in.PortCF) %>%
-  select(CFF.ID,MRN,Start)
-# A1C and OGTT
-chco_a1c = read.csv("./Data_Cleaned/chco_a1c_ogtt.csv")
-## Format
-chco_a1c$CFF.ID = as.character(chco_a1c$CFF.ID)
-chco_a1c$Date = mdy(chco_a1c$Date)
-chco_a1c$Diabetes = NULL
-## Only those w/CFRD and known start
-chco_a1c = left_join(chco_dates,chco_a1c,by = "CFF.ID")
-## Get closest before and closest after dates
-chco_a1c = closest_dates(chco_a1c)
-# PFTs
-chco_pft = read.csv("./Data_Cleaned/chco_bmi_pft.csv")
-chco_pft$Date = mdy(chco_pft$Date)
-chco_pft = left_join(chco_dates,chco_pft,by = "MRN")
-chco_pft = closest_dates(chco_pft)
-# CGM
-chco_cgm = read.csv("./Data_Cleaned/chco_cgm.csv")
-chco_cgm$Date = date(mdy_hm(chco_cgm$date_cgm_placement))
-chco_cgm$CFF.ID = sub("_.*","",chco_cgm$subject_id)
-# chco_cgm$subject_id = NULL
-chco_cgm = left_join(chco_dates,chco_cgm,by = "CFF.ID")
-chco_cgm = closest_dates(chco_cgm)
-# Merge
-chco = full_join(chco_a1c,chco_pft,by = c("CFF.ID", "MRN", "Timepoint"))
-chco = full_join(t,chco_cgm,by = c("CFF.ID", "MRN", "Timepoint"))
-
-# Montana
-montana = read.csv("./Data_Raw/Data From Montana/Copy of Paper patient population glycemic (+hofer).csv",na.strings = "")
-demo_adult = read.csv("./Data_Raw/Data From Montana/demo_adults.csv",na.strings = "")
-demo_kids = read.csv("./Data_Raw/Data From Montana/demo_kids.csv",na.strings = "")
-demo_kids$Date.Of.Birth..mm.dd.yyyy. = NULL
-demo_kids = demo_kids[,colnames(demo_adult)]
-demo = rbind(demo_adult,demo_kids)
-# Get within a year of Trikafta start
-montana$time_diff = mdy(montana$Trikafta.start.date) - mdy(montana$Date.of.Encounter)
-montana = montana %>% filter(abs(montana$time_diff)<=365)
-# Add demographics
-montana = left_join(montana,demo,by = "CFF.ID")
-# PFTs
-pft_adult <- read.csv("./Data_Raw/Data From Montana/Copy of PFTs growth adult (+hofer).csv")
-pft_kids <- read.csv("./Data_Raw/Data From Montana/PFTs kids.csv")
-pft_adult = pft_adult[,colnames(pft_kids)]
-pfts = rbind(pft_adult,pft_kids)
-montana = full_join(montana,pfts,by = c("CFF.ID", "Date.of.Encounter"))
-```

### 2022-01-27 Christine Chan/CF Metabolomics/chan_cfrd_metabolomics.Rmd
Create chan_cfrd_metabolomics.Rmd

---
title: ""
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD//PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
}
knitr::opts_knit$set(root.dir = home_dir)
```

```{r}
# Clinical data
source("/Users/timvigers/Dropbox/Mac/Documents/GitHub/CF/Christine Chan/CF Metabolomics/import_data.r")
# Metabolomics
metab_pos = read.csv("./Christine Chan/CF Metabolomics/Data_Raw/Christine Chan - CF-Diabetes - C18 Pos.csv")
```

### 2022-01-27 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
Form Combining
finished CHCO and started Montana, bit of work to be done on montana

@@
 
+NOTE: I HAVE THE SHARED DRIVE AS S:, to change if mapped to another letter
 ```{r setup, include=FALSE}
@@
 library(knitr)
+library(lubridate)
+library(stringr)
 knitr::opts_chunk$set(echo = FALSE)
@@
 
-```{r}
-# Import 
-df = read.csv("./Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv")
+Import the datasets (I'll be doing this one site at a time to look at the variables etc)
+```{r data import}
+###### Import CHCO data ###########
+
+# bmi, pulm function 577 obs
+# 8 vars MRN - ID, date for visits; start - ? start of trt?, fev, bmi, height, weight, fec 
+chco_bmi_pft = read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv")
+
+# continuous glucose monitoring, 23 obs
+# 105 vars subjectid - number_pre/post for visit and a date; numbers dont seem to correspond to mrn in bmi pft
+chco_cgm = read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/chco_cgm.csv")
+
+# chco glycemic, 261 obs
+# 7 var MRN and CFF id (=subjectid?), date for visits; start... etc
+chco_glycemic = read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/chco_glycemic.csv")
+
+# chco registry 108 obs
+# CFF ID and MRN, start date
+chco_registry = read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/chco_registry.csv")
+###################################
+
+###### IMPORT MONTANA DATA ########
+
+# NEED TO RENAME VARIABLES TO MATCH CHCO#
+# bmi, pulm function 975 obs
+montana_bmi_pft <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/montana_bmi_pft.csv")
+
+# cgm 16 obs
+montana_cgm <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/montana_cgm.csv")
+
+# registry: we need to combine demo (126), genotypes (126), pancreatic (241), gtubes (155)
+montana_demo <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/montana_demo.csv")
+montana_genotypes <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/montana_genotypes.csv")
+montana_gtubes <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/montana_gtubes.csv") # what are the dates here
+montana_pancreatic <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/montana_pancreatic.csv") # what is this? just dates and cffid
+# ask tim:what do we need to do with gtubes and pancreatic dates? dont look like chco data
+
+# glycemic 279
+montana_glycemic_registry <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv") # looks identical
+###################################
+
+###### IMPORT UW DATA #############
+###################################
+
+###### IMPORT ? DATA ########
+###################################
+
 ```
+
+Lets try and stitch these sets together by site (one site at a time as above)
+note: by default, the dplyr join functions join by all variables in common. So we need to make sure the common variables are same format and name
+```{r form stitching}
+####### CHCO ############
+# CGM Data Cleaning
+# get a CFF ID and visit variable by splitting subjectid variable into CFF ID and Visit
+chco_cgm$`CFF ID` = as.numeric(sub("_.*", "", chco_cgm$subject_id))
+chco_cgm$Visit = sub(".*_", "", chco_cgm$subject_id)
+
+# Get a Date variable by taking the date from "date time" char variable date_cgm_placement
+# weird because one obs isnt in a date time, just a date so ill just brute it
+chco_cgm = chco_cgm %>% mutate(datewithtime = str_detect(date_cgm_placement, " "),
+                               Date = ifelse(datewithtime,sub(" .*", "", chco_cgm$date_cgm_placement), date_cgm_placement))
+
+# per tim, its ok to have more than 2 dates per subject,
+# as it looks like a lot of this stuff was collected at different times for different subjects;
+# i expect this will make it pretty long
+
+chco_long = full_join(chco_glycemic, chco_bmi_pft) # will join by MRN, Start, Date
+
+chco_long = full_join( chco_registry, chco_long) # joins by MRN, CFFID, Start
+chco_long = full_join(chco_long, chco_cgm) # joins by cff id and date
+
+# add in a site id variable
+chco_long$site = "CHCO"
+
+#########################
+
+####### Montana ############
+# to do: registry var differences in name and type
+# rename the bmi_pft to chco names
+# cgm, glycemic look good
+
+# differences in montana and chco registry vars:
+# Race labels
+# make registry dataframe
+montana_registry = full_join(montana_demo, montana_genotypes)
+montana_registry = full_join(montana_registry, montana_pancreatic)
+montana_registry = full_join(montana_registry, montana_gtubes)
+
+# CGM looks like this is the same as CHCO (GROOVY)
+# cgm data cleaning
+montana_cgm$`CFF ID` = as.numeric(sub("_.*", "", montana_cgm$subject_id))
+montana_cgm$Visit = sub(".*_", "", montana_cgm$subject_id)
+
+# Get a Date variable by taking the date from "date time" char variable date_cgm_placement
+montana_cgm = montana_cgm %>% mutate(datewithtime = str_detect(date_cgm_placement, " "),
+                               Date = ifelse(datewithtime,sub(" .*", "", montana_cgm$date_cgm_placement), date_cgm_placement))
+
+#########################
+
+####### UW ############
+#########################
+
+####### UM? ############
+#########################
+
+```

### 2022-01-27 Christine Chan/CF Metabolomics/chan_cfrd_metabolomics.Rmd

@@
 ```{r setup, include=FALSE}
-library(arsenal)
+library(Hmisc)
+library(mixOmics)
 library(tidyverse)
-library(performance)
 library(knitr)
@@
 ```{r}
+clean_metab = function(data,prefix = "M"){
+  # Replace 1 with NA
+  data[data == 1] = NA
+  # Compound names
+  data$compound_id = paste0(prefix,1:nrow(data))
+  rownames(data) = data$compound_id
+  compounds = data[,c("compound_id","Compound")]
+  # Get samples only
+  data = data[,grep("area2height",colnames(data))]
+  data = data[,!grepl("blank",tolower(colnames(data)))]
+  colnames(data) = gsub("Lipid_Pos_|Lipid_Neg_|Hilic_Pos_|_area2height.raw.|_dup","",colnames(data))
+  # Transpose
+  data = data.frame(t(data))
+  # Fix subject names
+  data$subject_id = str_pad(rownames(data),3,"left","0")
+  # Return data and compounds
+  return(list(data,compounds))
+}
+```
+
+```{r}
 # Clinical data
 source("/Users/timvigers/Dropbox/Mac/Documents/GitHub/CF/Christine Chan/CF Metabolomics/import_data.r")
-# Metabolomics
-metab_pos = read.csv("./Christine Chan/CF Metabolomics/Data_Raw/Christine Chan - CF-Diabetes - C18 Pos.csv")
+df = data %>% group_by(medical_record_number) %>% filter(row_number() == 1) %>% ungroup() %>%
+  select(subject_id,group_category.factor,age_at_visit_1,gender.factor,race_ethnicity.factor,
+         average_bmi_z_score,fvc,fev1_percent,poc_a1c,ag,fructosamine,
+         percent_glycated_albumin)
+df$subject_id = as.character(sub("--\\d","",df$subject_id))
+
+# Import metabolomics
+metab_pos = read.csv("./Christine Chan/CF Metabolomics/Data_Raw/Christine Chan - CF-Diabetes - C18 Pos.csv",na.strings = "")
+metab_neg = read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/CF Metabolomics/Data_Raw/Christine Chan - CF-Diabetes - C18 Neg.csv",na.strings = "")
+hilic = read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/CF Metabolomics/Data_Raw/Christine Chan - CF-Diabetes - HILIC Pos.csv",na.strings = "")
+# Clean
+## Positive
+metab_pos = clean_metab(metab_pos,prefix = "P")
+compounds = metab_pos[[2]]
+## Negative
+metab_neg = clean_metab(metab_neg,prefix = "N")
+compounds = rbind(compounds,metab_neg[[2]])
+## HILIC
+hilic = clean_metab(hilic,prefix = "H")
+compounds = rbind(compounds,hilic[[2]])
+# Merge
+df = df %>% filter(subject_id %in% metab_pos[[1]]$subject_id) %>%
+  left_join(.,metab_pos[[1]],by = "subject_id") %>% 
+  left_join(.,metab_neg[[1]],by = "subject_id") %>%
+  left_join(.,hilic[[1]],by = "subject_id")
+# Larger groups
+df$combined = df$group_category.factor
+levels(df$combined) = c("HC & CFNGT","HC & CFNGT","CFAGT & CFRD", "CFAGT & CFRD")
+# Remove low variance
+df = df[,-nearZeroVar(df)$Position]
+# Matrices for mixOmics
+X = df %>% select(P1:H651)
+Y_all = df$group_category.factor
+Y_two = df$combined
 ```
+
+# PCA
+
+## Four groups
+
+```{r}
+pca_res = pca(X,center = T,scale = T)
+plotIndiv(pca_res,group = df$group_category.factor,legend = TRUE,pch = 16,
+          title = "PCA Components 1 - 2")
+```
+
+## Control and CFNGT vs. Abnormal Glycemia
+
+```{r}
+plotIndiv(pca_res,group = df$combined,legend = TRUE,pch = 16,
+          title = "PCA Components 1 - 2")
+```
+
+# PLS-DA
+
+```{r}
+plsda_res = splsda(X, Y_all)
+plotIndiv(plsda_res,group = df$group_category.factor,legend = TRUE,pch = 16,
+          title = "PCA Components 1 - 2")
+```

### 2022-01-27 R Markdown Template.Rmd

---
title: ""
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD//PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
}
knitr::opts_knit$set(root.dir = home_dir)
```

```{r}

```

### 2022-01-31 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
Form Combining
Montana, UW, need to work on UM after tim's advice and final form

@@
 ###### IMPORT UW DATA #############
+uw_registry <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/uw_registry.csv") #cffid and start
+uw_bmi <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/uw_bmi.csv") # date, cffid, height, bmi (height bmi need rename)
+uw_pft <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/uw_pft.csv") # datee, cffid, fev, fvc need renaming
+uw_cgm <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/uw_cgm.csv") # same code as montana
+uw_demo <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/uw_demo.csv") 
+uw_genotypes <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/uw_registry.csv") # no genotype data here?
+uw_glycemic <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/uw_glycemic.csv") # looks good, same as chco
+uw_gtubes <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/uw_gtubes.csv") # rename date of encounter to date
+uw_pancreatic <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/uw_pancreatic.csv") # same as montana, no pancreatic data though
+
 ###################################
 
-###### IMPORT ? DATA ########
+###### IMPORT UM DATA ########
+um_all <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/um_all.csv")
 ###################################
@@
 ####### Montana ############
-# to do: registry var differences in name and type
+
+# Race labels
+factor(chco_registry$Race) # Black, Caucasian, Other, Unknown
+factor(montana_demo$Race) # American Indian or Alaska Native, Black or African American, Some other race, Two or more races, White
+
+# keep original data, add in a new one to match CHCO
+montana_demo = montana_demo %>% rename(race_original = Race)
+montana_demo = montana_demo %>% mutate(Race = case_when(race_original == "White" ~ "Caucasian",
+                                                        race_original == "Black or African American" ~ "Black",
+                                                        race_original == NA ~ "Unknown",
+                                                        TRUE ~ "Other"))
 # rename the bmi_pft to chco names
-# cgm, glycemic look good
+montana_bmi_pft = montana_bmi_pft %>% rename(FVC = `Measure of FVC`,
+                                             FEV1 = `Measure of FEV1`,
+                                             Height = `Patients height`,
+                                             Weight = `Weight of Patient`)
 
-# differences in montana and chco registry vars:
-# Race labels
 # make registry dataframe
@@
 
+# combine forms
+montana_long = full_join(montana_glycemic_registry, montana_bmi_pft) # will join by cffid, Date
+
+montana_long = full_join(montana_registry, montana_long) # joins by MRN, CFFID, Start
+montana_long = full_join(montana_long, montana_cgm) # joins by cff id and date
+
+# add in a site id variable
+montana_long$site = "Montana"
 #########################
 
-####### UW ############
-#########################
+####### UM  ############
+# to do: rename variables to match CHCO
+# most variables have mult columns and no notation: FEV, A1C, FVC, CFRD Status, etc
+um_all %>% select(`CFF ID`, `CFRD Status...10`, `CFRD Status...34`) %>% filter(`CFRD Status...10` != `CFRD Status...34`)
 
-####### UM? ############
-#########################
+########################
 
+####### UW #############
+# Race labels
+factor(chco_registry$Race) # Black, Caucasian, Other, Unknown
+factor(uw_demo$Race) # American Indian or Alaska Native, Black or African American, Some other race, Two or more races, White
+
+# keep original data, add in a new one to match CHCO
+uw_demo = uw_demo %>% rename(race_original = Race)
+uw_demo = uw_demo %>% mutate(Race = case_when(race_original == "White" ~ "Caucasian",
+                                                        race_original == "Black or African American" ~ "Black",
+                                                        race_original == NA ~ "Unknown",
+                                                        TRUE ~ "Other"))
+# make bmi pft
+uw_bmi_pft = full_join(uw_bmi, uw_pft)
+# rename the bmi_pft to chco names
+uw_bmi_pft = uw_bmi_pft %>% rename(FVC = `Measure of FVC`,
+                                   FEV1 = `Measure of FEV1`,
+                                   Height = `Patients height`,
+                                   BMI = `BMI Value`)
+
+# CGM looks like this is the same as CHCO (GROOVY)
+# cgm data cleaning
+uw_cgm$`CFF ID` = as.numeric(sub("_.*", "", uw_cgm$subject_id))
+uw_cgm$Visit = sub(".*_", "", uw_cgm$subject_id)
+# Get a Date variable by taking the date from "date time" char variable date_cgm_placement
+uw_cgm = uw_cgm %>% mutate(datewithtime = str_detect(date_cgm_placement, " "),
+                               Date = ifelse(datewithtime,sub(" .*", "", uw_cgm$date_cgm_placement), date_cgm_placement))
+
+# combine registry (leaving pancreatic and genotypes out because only a id and date variable)
+uw_registry = full_join(uw_registry, uw_demo)
+# uw_registry = full_join(uw_registry, uw_genotypes)
+# uw_registry = full_join(uw_registry, uw_pancreatic)
+uw_registry = full_join(uw_registry, uw_gtubes)
+
+# combine forms
+uw_long = full_join(uw_glycemic, uw_bmi_pft) # will join by cffid, Date
+uw_long = full_join(uw_registry, uw_long) # joins by MRN, CFFID, Start
+uw_long = full_join(uw_long, uw_cgm) # joins by cff id and date
+
+# add in a site id variable
+uw_long$site = "UW"
+########################
 ```
+TO ASK TIM: 
+-significance of dates? adding a ton of different rows, and wondering which dates are significant
+-why do UW and Montana files (pancreatic) just have dates? are these different dates/ significant? do we need this file/ rename the date? gtubes similar
+
+bmi pft: didnt worry about percent predicted (included in uw and montana), wasnt in chco data

### 2022-02-02 Christine Chan/Pre-post Trikafta/pre_post_trikafta_merge.Rmd
separate merge and analysis files

--- 2022-01-31 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
@@
 title: ""
-author: "Tim Vigers & Laura Pyle"
+author: "Tim Vigers & Casey Sakamoto"
 date: "`r format(Sys.time(), '%d %B %Y')`"

### 2022-02-02 R Markdown Template.Rmd
Update R Markdown Template.Rmd

@@
 title: ""
-author: "Tim Vigers & Laura Pyle"
+author: "Tim Vigers"
 date: "`r format(Sys.time(), '%d %B %Y')`"
@@
 knitr::opts_chunk$set(echo = FALSE)
-if (Sys.info()["sysname"] == "Linux"){
-  home_dir = "~/UCD//PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
+if(Sys.info()["sysname"] == "Windows"){
+  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
+} else if (Sys.info()["sysname"] == "Linux"){
+  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
 } else if (Sys.info()["sysname"] == "Darwin"){

### 2022-02-02 Christine Chan/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
Create pre_post_trikafta_analysis.Rmd

---
title: "Pre-/Post-Trikafta in CFRD"
author: "Tim Vigers & Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "B:/Projects"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
}
knitr::opts_knit$set(root.dir = home_dir)
```

```{r}

```

### 2022-02-02 Christine Chan/Pre-post Trikafta/pre_post_trikafta_merge.Rmd

--- 2022-01-31 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
@@
 
+#hba1c should be a numeric, one with a note : pre trikafta cffid 1433411
+montana_long = montana_long %>% rename(HbA1c_original = HbA1c)
+montana_long = montana_long %>% mutate(hba1c_spc = str_detect(HbA1c_original, " "),
+                               HbA1c = as.numeric(ifelse(hba1c_spc,sub(" .*", "", montana_long$HbA1c_original), HbA1c_original)))
+
+montana_long = montana_long %>% select(-c(hba1c_spc, HbA1c_original))
 # add in a site id variable
@@
 bmi pft: didnt worry about percent predicted (included in uw and montana), wasnt in chco data
+
+```{r analysis dataset}
+# full join long datasets
+analysis_data = full_join(chco_long, montana_long)
+analysis_data = full_join(analysis_data, uw_long)
+
+# export csv analysis_data
+write.csv(analysis_data, file = "S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/analysis_data.csv")
+```

### 2022-02-02 Christine Chan/Pre-post Trikafta/pre_post_trikafta_merge.Rmd
Update pre_post_trikafta_merge.Rmd
relative paths

@@
 if(Sys.info()["sysname"] == "Windows"){
-  home_dir = "S://PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan"
+  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan"
 } else if (Sys.info()["sysname"] == "Linux"){
@@
 # 8 vars MRN - ID, date for visits; start - ? start of trt?, fev, bmi, height, weight, fec 
-chco_bmi_pft = read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv")
+chco_bmi_pft = read_csv("./Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv")
 
@@
 # 105 vars subjectid - number_pre/post for visit and a date; numbers dont seem to correspond to mrn in bmi pft
-chco_cgm = read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/chco_cgm.csv")
+chco_cgm = read_csv("./Prepost Triakfta/Data_Cleaned/chco_cgm.csv")
 
@@
 # 7 var MRN and CFF id (=subjectid?), date for visits; start... etc
-chco_glycemic = read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/chco_glycemic.csv")
+chco_glycemic = read_csv("./Prepost Triakfta/Data_Cleaned/chco_glycemic.csv")
 
@@
 # CFF ID and MRN, start date
-chco_registry = read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/chco_registry.csv")
+chco_registry = read_csv("./Prepost Triakfta/Data_Cleaned/chco_registry.csv")
 ###################################
@@
 # bmi, pulm function 975 obs
-montana_bmi_pft <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/montana_bmi_pft.csv")
+montana_bmi_pft <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_bmi_pft.csv")
 
 # cgm 16 obs
-montana_cgm <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/montana_cgm.csv")
+montana_cgm <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_cgm.csv")
 
 # registry: we need to combine demo (126), genotypes (126), pancreatic (241), gtubes (155)
-montana_demo <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/montana_demo.csv")
-montana_genotypes <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/montana_genotypes.csv")
-montana_gtubes <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/montana_gtubes.csv") # what are the dates here
-montana_pancreatic <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/montana_pancreatic.csv") # what is this? just dates and cffid
+montana_demo <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_demo.csv")
+montana_genotypes <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_genotypes.csv")
+montana_gtubes <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_gtubes.csv") # what are the dates here
+montana_pancreatic <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_pancreatic.csv") # what is this? just dates and cffid
 # ask tim:what do we need to do with gtubes and pancreatic dates? dont look like chco data
@@
 # glycemic 279
-montana_glycemic_registry <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv") # looks identical
+montana_glycemic_registry <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv") # looks identical
 ###################################
@@
 ###### IMPORT UW DATA #############
-uw_registry <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/uw_registry.csv") #cffid and start
-uw_bmi <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/uw_bmi.csv") # date, cffid, height, bmi (height bmi need rename)
-uw_pft <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/uw_pft.csv") # datee, cffid, fev, fvc need renaming
-uw_cgm <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/uw_cgm.csv") # same code as montana
-uw_demo <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/uw_demo.csv") 
-uw_genotypes <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/uw_registry.csv") # no genotype data here?
-uw_glycemic <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/uw_glycemic.csv") # looks good, same as chco
-uw_gtubes <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/uw_gtubes.csv") # rename date of encounter to date
-uw_pancreatic <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/uw_pancreatic.csv") # same as montana, no pancreatic data though
+uw_registry <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_registry.csv") #cffid and start
+uw_bmi <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_bmi.csv") # date, cffid, height, bmi (height bmi need rename)
+uw_pft <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_pft.csv") # datee, cffid, fev, fvc need renaming
+uw_cgm <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_cgm.csv") # same code as montana
+uw_demo <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_demo.csv") 
+uw_genotypes <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_registry.csv") # no genotype data here?
+uw_glycemic <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_glycemic.csv") # looks good, same as chco
+uw_gtubes <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_gtubes.csv") # rename date of encounter to date
+uw_pancreatic <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_pancreatic.csv") # same as montana, no pancreatic data though
 
@@
 ###### IMPORT UM DATA ########
-um_all <- read_csv("S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/um_all.csv")
+um_all <- read_csv("./Prepost Triakfta/Data_Cleaned/um_all.csv")
 ###################################
@@
 # export csv analysis_data
-write.csv(analysis_data, file = "S:/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Cleaned/analysis_data.csv")
+write.csv(analysis_data, file = "./Prepost Triakfta/Data_Cleaned/analysis_data.csv")
 ```

### 2022-02-02 Christine Chan/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
Update pre_post_trikafta_analysis.Rmd

@@
 if(Sys.info()["sysname"] == "Windows"){
-  home_dir = "B:/Projects"
+  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
 } else if (Sys.info()["sysname"] == "Linux"){
-  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
+  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
 } else if (Sys.info()["sysname"] == "Darwin"){
-  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
+  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
 }
@@
 ```{r}
+# Import cleaned data from Casey
+df = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/analysis_data.csv",na.strings = "NA")
+df$X = NULL
+# Fill in MRN and CFF ID for everyone
+df = df %>% 
+  group_by(CFF.ID) %>% fill(MRN,.direction = "downup") %>% 
+  group_by(MRN) %>% fill(CFF.ID,.direction = "downup") %>%
+  arrange(CFF.ID,Date)
+# Fill in constant variables for all rows
+df = df %>% group_by(MRN) %>% fill(CFF.ID,.direction = "downup") %>% 
+  group_by(CFF.ID) %>% fill(DOB,Start:Genotypes2,CFRD.Diagnosis.Date,.direction = "downup")
+# Remove those without a start date
+t = df %>% filter(is.na(Start)) %>% select(CFF.ID,site)
+```
 
+```{r}
+# Piecewise model function
+piecewise = function(outcome,data = df){
+
+}
 ```

### 2022-02-02 Christine Chan/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
Update pre_post_trikafta_analysis.Rmd

--- 2022-02-02 Christine Chan/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
@@
 library(performance)
+library(lubridate)
 library(knitr)
@@
 ```{r}
+# Import cleaned data from Casey
+df = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/analysis_data.csv",na.strings = "NA")
+df$X = NULL
+# Fill in MRN and CFF ID for everyone
+df = df %>% 
+  group_by(CFF.ID) %>% fill(MRN,.direction = "downup") %>% 
+  group_by(MRN) %>% fill(CFF.ID,.direction = "downup") %>%
+  arrange(CFF.ID,Date)
+# Fill in constant variables for all rows
+df = df %>% group_by(MRN) %>% fill(CFF.ID,.direction = "downup") %>% 
+  group_by(CFF.ID) %>% fill(DOB,Start:CFRD.Diagnosis.Date,.direction = "downup") %>% ungroup()
+df$Hispanic.Latinx[is.na(df$Hispanic.Latinx)] = FALSE
+# Assume that blanks for pancreatic and G tube indicate "no"
+df$Pancreatic.Status[is.na(df$Pancreatic.Status)] = "Sufficient"
+df$G.tube.in.past.12.months[is.na(df$G.tube.in.past.12.months)] = "No"
+# Convert dates
+date_vars = c("DOB","Start","Date")
+df[,date_vars] = lapply(df[,date_vars], function(d){
+  parse_date_time(d,orders = c("mdy","ymd","dmy"),quiet = T)
+})
+# Time from start and filtering
+df = df %>% mutate(time = as.numeric(difftime(Date,Start,units = "days"))) %>%
+  filter(!is.na(time),abs(time) <= 365,!is.na(CFF.ID),
+         !is.na(CFRD.Diagnosis.Date) | !is.na(CFRD.Status)) %>%
+  select(-MRN,CFF.ID,DOB,Start,everything())
+```
+
+# Table 1: Participant Characteristics
+
+```{r}
+demographics = df %>% group_by(CFF.ID) %>% filter(row_number() == 1)
 
 ```
+
+```{r}
+# Piecewise model function
+piecewise = function(outcome,data = df){
+
+}
+```

### 2022-02-02 Christine Chan/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
Update pre_post_trikafta_analysis.Rmd

--- 2022-02-02 Christine Chan/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
@@
 library(tidyverse)
+library(lmerTest)
+library(broom.mixed)
 library(performance)
+library(lubridate)
+library(plotly)
 library(knitr)
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE,fig.height = 8,fig.width = 10)
 if(Sys.info()["sysname"] == "Windows"){
-  home_dir = "B:/Projects"
+  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
 } else if (Sys.info()["sysname"] == "Linux"){
-  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
+  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
 } else if (Sys.info()["sysname"] == "Darwin"){
-  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
+  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
 }
@@
 
+```{r include=FALSE}
+# Import cleaned data from Casey
+df = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/analysis_data.csv",na.strings = "NA")
+df$X = NULL
+# Fill in MRN and CFF ID for everyone
+df = df %>% 
+  group_by(CFF.ID) %>% fill(MRN,.direction = "downup") %>% 
+  group_by(MRN) %>% fill(CFF.ID,.direction = "downup")
+# Assume that blanks for pancreatic and G tube indicate "no"
+df$Hispanic.Latinx[is.na(df$Hispanic.Latinx)] = FALSE
+df$Pancreatic.Status[is.na(df$Pancreatic.Status)] = "Sufficient"
+df$G.tube.in.past.12.months[is.na(df$G.tube.in.past.12.months)] = "No"
+# Convert dates
+date_vars = c("DOB","Start","Date")
+df[,date_vars] = lapply(df[,date_vars], function(d){
+  parse_date_time(d,orders = c("mdy","ymd","dmy"),quiet = T)
+})
+# Time from start and filtering
+df = df %>% mutate(Time = as.numeric(difftime(Date,Start,units = "days"))) 
+# Add UM (they did not provide dates, so need Time column)
+um_a1cs = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/um_a1cs.csv",na.strings = "")
+um_bmi = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/um_bmi.csv",na.strings = "")
+um_demo = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/um_demo.csv",na.strings = "")
+um_ogtt = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/um_ogtt.csv",na.strings = "")
+um_pfts = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/um_pfts.csv",na.strings = "")
+um = full_join(um_a1cs,um_bmi)
+um = full_join(um,um_ogtt)
+um = full_join(um,um_pfts)
+um = left_join(um,um_demo)
+um = um %>% filter(!is.na(CFRD.Status))
+df = full_join(df,um)
+# Filter and fill constants
+df = df %>% group_by(CFF.ID) %>% fill(DOB,Start:CFRD.Diagnosis.Date,.direction = "downup") %>%
+  ungroup() %>% filter(!is.na(Time),abs(Time) <= 365,!is.na(CFF.ID),
+                       (!is.na(CFRD.Diagnosis.Date) | !is.na(CFRD.Status) | site == "UW")) %>%
+  select(-MRN) %>% select(CFF.ID,site,DOB,Start,everything()) %>% rename(Site = "site")
+```
+
 ```{r}
+data = df
+outcome = "HbA1c"
+# Piecewise model function
+piecewise = function(outcome,data = df){
+  # Add spline column
+  data$Change = ifelse(data$Time < 0,0,1)
+  data$Change = data$Change * data$Time
+  # Model
+  f = as.formula(paste0(outcome,"~Time + Change + (1|CFF.ID)"))
+  mod = lmer(f,data = data)
+  res = tidy(mod,"fixed")
+  res$effect = NULL
+  # Coordinates for model plotting
+  x1_start = -365
+  y1_start = res$estimate[1] + res$estimate[2]*x1_start
+  x1_end = 0
+  y1_end = res$estimate[1] 
+  x2_start = 0
+  y2_start = res$estimate[1] 
+  x2_end = 365
+  y2_end = res$estimate[1] + (res$estimate[2]+res$estimate[3])*x2_end
+  # Plot
+  p = ggplot(data,aes_string(x = "Time",y = outcome,group = "CFF.ID",color = "Site")) +
+    geom_point(alpha = 0.5) + geom_line(alpha = 0.2) +
+    annotate("segment",x=x1_start,y=y1_start,xend=x1_end,yend=y1_end,color = "blue",size = 1) +
+    annotate("segment",x=x2_start,y=y2_start,xend=x2_end,yend=y2_end,color = "blue",size = 1) +
+    xlab("Days from Trikafta Start") + theme_bw()
+  # Interactive plot and model results
+  return(list(plot = ggplotly(p),table = res))
+}
+```
 
+# Methods
+
+# HbA1c
+
+```{r results='asis'}
+a1c = piecewise(outcome = "HbA1c",df)
+ggplotly(a1c$plot)
+kable(a1c$table,digits = 3)
 ```

### 2022-02-02 Christine Chan/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
Update pre_post_trikafta_analysis.Rmd

@@
 library(lmerTest)
+library(nlme)
 library(broom.mixed)
+library(segmented)
 library(performance)
@@
   res$effect = NULL
+  res$term[2] = "Slope Prior"
+  # Segmented fit
+  fit = lme(as.formula(paste0(outcome,"~Time + Change")),random = ~1|CFF.ID,data = data,na.action = na.omit)
+  seg = suppressWarnings(segmented(fit, seg.Z = ~Time, npsi=1))
+  # Model check if necessary
+  check = check_model(mod)
   # Coordinates for model plotting
@@
   # Interactive plot and model results
-  return(list(plot = ggplotly(p),table = res))
+  return(list(plot = ggplotly(p),table = res,check = check,seg = seg))
 }
@@
 
+For each outcome, we fit a piecewise linear mixed effects model with random intercept for each person and a "knot" at 0.
+This approach can be thought of as fitting two separate linear models, while ensuring that the two regression lines meet at Trikafta start.
+In the model results, "Slope Prior" indicates the slope prior to start of treatment (i.e.
+the average change in outcome for each "1 unit increase" in time).
+The "Change" term indicates the *change in slope* after medication start.
+To find the slope post-Trikafta, you add the "Slope Prior" and "Change" estimates.
+
 # HbA1c
 
-```{r results='asis'}
+```{r results='asis',warning=FALSE}
 a1c = piecewise(outcome = "HbA1c",df)
@@
 kable(a1c$table,digits = 3)
+a1c$check
 ```
+
+`r if(class(a1c$seg)=="lme"){"Automated R package was not able to find a better breakpoint."}`

### 2022-02-02 Christine Chan/Pre-post Trikafta/pre_post_trikafta_merge.Rmd
Update pre_post_trikafta_merge.Rmd

@@
 ###### IMPORT UM DATA ########
-um_all <- read_csv("./Prepost Triakfta/Data_Cleaned/um_all.csv")
+#um_all <- read_csv("./Prepost Triakfta/Data_Cleaned/um_all.csv")
 ###################################
@@
 # most variables have mult columns and no notation: FEV, A1C, FVC, CFRD Status, etc
-um_all %>% select(`CFF ID`, `CFRD Status...10`, `CFRD Status...34`) %>% filter(`CFRD Status...10` != `CFRD Status...34`)
+#um_all %>% select(`CFF ID`, `CFRD Status...10`, `CFRD Status...34`) %>% filter(`CFRD Status...10` != `CFRD Status...34`)
 

### 2022-02-02 Christine Chan/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
Update pre_post_trikafta_analysis.Rmd

--- 2022-02-02 Christine Chan/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
@@
 library(tidyverse)
+library(lmerTest)
+library(nlme)
+library(broom.mixed)
+library(segmented)
 library(performance)
 library(lubridate)
+library(plotly)
 library(knitr)
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE,fig.height = 8,fig.width = 10)
 if(Sys.info()["sysname"] == "Windows"){
@@
 
-```{r}
+```{r include=FALSE}
 # Import cleaned data from Casey
@@
   group_by(CFF.ID) %>% fill(MRN,.direction = "downup") %>% 
-  group_by(MRN) %>% fill(CFF.ID,.direction = "downup") %>%
-  arrange(CFF.ID,Date)
-# Fill in constant variables for all rows
-df = df %>% group_by(MRN) %>% fill(CFF.ID,.direction = "downup") %>% 
-  group_by(CFF.ID) %>% fill(DOB,Start:CFRD.Diagnosis.Date,.direction = "downup") %>% ungroup()
+  group_by(MRN) %>% fill(CFF.ID,.direction = "downup")
+# Assume that blanks for pancreatic and G tube indicate "no"
 df$Hispanic.Latinx[is.na(df$Hispanic.Latinx)] = FALSE
-# Assume that blanks for pancreatic and G tube indicate "no"
 df$Pancreatic.Status[is.na(df$Pancreatic.Status)] = "Sufficient"
@@
 # Time from start and filtering
-df = df %>% mutate(time = as.numeric(difftime(Date,Start,units = "days"))) %>%
-  filter(!is.na(time),abs(time) <= 365,!is.na(CFF.ID),
-         !is.na(CFRD.Diagnosis.Date) | !is.na(CFRD.Status)) %>%
-  select(-MRN,CFF.ID,DOB,Start,everything())
-```
-
-# Table 1: Participant Characteristics
-
-```{r}
-demographics = df %>% group_by(CFF.ID) %>% filter(row_number() == 1)
+df = df %>% mutate(Time = as.numeric(difftime(Date,Start,units = "days"))) 
+# Add UM (they did not provide dates, so need Time column)
+um_a1cs = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/um_a1cs.csv",na.strings = "")
+um_bmi = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/um_bmi.csv",na.strings = "")
+um_demo = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/um_demo.csv",na.strings = "")
+um_ogtt = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/um_ogtt.csv",na.strings = "")
+um_pfts = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/um_pfts.csv",na.strings = "")
+um = full_join(um_a1cs,um_bmi)
+um = full_join(um,um_ogtt)
+um = full_join(um,um_pfts)
+um = left_join(um,um_demo)
+um = um %>% filter(!is.na(CFRD.Status))
+df = full_join(df,um)
+# Filter and fill constants
+df = df %>% group_by(CFF.ID) %>% fill(DOB,Start:CFRD.Diagnosis.Date,.direction = "downup") %>%
+  ungroup() %>% filter(!is.na(Time),abs(Time) <= 365,!is.na(CFF.ID),
+                       (!is.na(CFRD.Diagnosis.Date) | !is.na(CFRD.Status) | site == "UW")) %>%
+  select(-MRN) %>% select(CFF.ID,site,DOB,Start,everything()) %>% rename(Site = "site")
 
@@
 ```{r}
+data = df
+outcome = "HbA1c"
 # Piecewise model function
 piecewise = function(outcome,data = df){
-
+  # Add spline column
+  data$Change = ifelse(data$Time < 0,0,1)
+  data$Change = data$Change * data$Time
+  # Model
+  f = as.formula(paste0(outcome,"~Time + Change + (1|CFF.ID)"))
+  mod = lmer(f,data = data)
+  res = tidy(mod,"fixed")
+  res$effect = NULL
+  #res$term[2] = "Slope Prior"
+  # Segmented fit
+  fit1 = lme(as.formula(paste0(outcome,"~Time + Change")),random = ~1|CFF.ID,data = data,na.action = na.omit)
+  seg = suppressWarnings(segmented(fit, seg.Z = ~Time, npsi=1))
+  # Model check if necessary
+  check = check_model(mod)
+  # Coordinates for model plotting
+  x1_start = -365
+  y1_start = res$estimate[1] + res$estimate[2]*x1_start
+  x1_end = 0
+  y1_end = res$estimate[1] 
+  x2_start = 0
+  y2_start = res$estimate[1] 
+  x2_end = 365
+  y2_end = res$estimate[1] + (res$estimate[2]+res$estimate[3])*x2_end
+  # Plot
+  p = ggplot(data,aes_string(x = "Time",y = outcome,group = "CFF.ID",color = "Site")) +
+    geom_point(alpha = 0.5) + geom_line(alpha = 0.2) +
+    annotate("segment",x=x1_start,y=y1_start,xend=x1_end,yend=y1_end,color = "blue",size = 1) +
+    annotate("segment",x=x2_start,y=y2_start,xend=x2_end,yend=y2_end,color = "blue",size = 1) +
+    xlab("Days from Trikafta Start") + theme_bw()
+  # Interactive plot and model results
+  return(list(plot = ggplotly(p),table = res,check = check,seg = seg))
 }
 ```
+
+# Methods
+
+For each outcome, we fit a piecewise linear mixed effects model with random intercept for each person and a "knot" at 0.
+This approach can be thought of as fitting two separate linear models, while ensuring that the two regression lines meet at Trikafta start.
+In the model results, "Slope Prior" indicates the slope prior to start of treatment (i.e.
+the average change in outcome for each "1 unit increase" in time).
+The "Change" term indicates the *change in slope* after medication start.
+To find the slope post-Trikafta, you add the "Slope Prior" and "Change" estimates.
+
+The random intercept model was selected because it performed better (by AIC) than a model with random intercept and continuous AR1 correlation structure.
+Random slope models did not converge for some outcomes, so were not used for the sake of consistency between results.
+
+# HbA1c
+
+```{r results='asis'}
+a1c = piecewise(outcome = "HbA1c",df)
+ggplotly(a1c$plot)
+kable(a1c$table,digits = 3)
+a1c$check
+```
+
+`r if(class(a1c$seg)=="lme"){"Automated R package was not able to find a better breakpoint."}`
+
+# BMI
+
+```{r results='asis'}
+bmi = piecewise(outcome = "BMI",df)
+ggplotly(bmi$plot)
+kable(bmi$table,digits = 3)
+bmi$check
+```
+
+`r if(class(bmi$seg)=="lme"){"Automated R package was not able to find a better breakpoint."}`

### 2022-02-07 Jordana Hoppe/PEx Microbiome/20200224_table1_misc_tests.Rmd
Create 20200224_table1_misc_tests.Rmd

---
title: "20200224_Table1_Final"
author: "Casey Sakamoto"
date: "2/24/2020"
output:
  pdf_document: default
  word_document: default
---
This is final version (instead of 20200101_table1.r)
(i messed up on version control)

```{r setup, include=FALSE}
#setwd("C:/Users/Casey/Desktop/research")
library(tidyverse)
library(tableone)
library(readxl)
library(knitr)
outpatient2020_01_25 <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordan Hoppe/PEX Microbiome/Data_Raw/outpatient2020_01_25.csv")
cult3 <- read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordan Hoppe/PEX Microbiome/Data_Raw/cult3.xlsx")
#outpatient2020_01_25 <- read.csv("C:/Users/Casey/Desktop/research/20200128plots/outpatient2020_01_25.csv")
#cult3 <- read_excel("C:/Users/Casey/Desktop/research/cult3.xlsx")
# age, gender, base fev1% (m r), fev1 pred(mr)
# change pred fev (mr), v1 pes score, days between visit (mr)
# ? genotyp, cf resp culture v1, positie viral pcr
# antibiotic class prescribed
outpatient_main = outpatient2020_01_25 %>% 
  select(sid,molecularid, perc_total, last, visit, lq_all, fev1_perc_6m_prior,sex , pes_total, age_yrs,
         fev1_perc_pre, fev1_perc_post,Antibiotic_class, visit_dif,pcr_neg, genotype1,genotype2, ShannonH_Median )%>%
  filter(!(sid == "JAR018" | sid == "BAR019" | sid == "NAO032" | sid == "EIL010" ))

outpatient_2vv = outpatient_main %>% group_by(sid,visit) %>% filter(n() > 224)
# take the max lq_all
outpatient_2vv = outpatient_2vv %>% group_by(sid, visit) %>% filter(lq_all == max(lq_all))

# combine this with our other patient data
tempdf = outpatient_main %>% group_by(sid,visit) %>% filter(n() <= 224)
outpatient_2vv = rbind(tempdf, outpatient_2vv)
rm(tempdf)
# subjects not in our 19
sublist = c("BAM014", "BSN024", "ESP008", "MEB022", "NDR020", "SKH012", "VAC001", "VCR009", "A_MO25")
outpatient_2vv = outpatient_2vv %>%  filter(!(sid %in% sublist))

summary(outpatient_2vv)

outpatient_2vv = outpatient_2vv %>% mutate(
  genotype = case_when(genotype1 == "F508del" & genotype2 == "F508del" ~ "F508del/F508del",
                       genotype1 == "F508del" & genotype2 != "F508del" ~ "F508del/Other",
                       visit==1 & sid =="SLP006"~"Other"))
# observations for table 1
unique_df = outpatient_2vv %>% select(-c(last, perc_total)) %>% unique.data.frame()
# sid and mid
molecular_id = unique_df[,1:2]
#write.csv(molecular_id, "molecularid.csv")
#
df_v1 = unique_df %>% filter(visit==1) %>% select(-visit_dif, -visit)
df_v2 = unique_df %>% filter(visit==2) 
df_v2 = df_v2 %>%  select(sid, visit_dif)

patient_t1_df = full_join(df_v1, df_v2, by = 'sid')
patient_t1_df = patient_t1_df %>% select(-visit.x, -visit.y)
rm(df_v1, df_v2, unique_df)
patient_t1_df = patient_t1_df %>% mutate(change_fev_perc = fev1_perc_pre - fev1_perc_6m_prior,
                                         change_fev_perc_v1v2 = fev1_perc_post - fev1_perc_pre)
patient_t1_df$age_yrs = floor(patient_t1_df$age_yrs)
# create table 1
all_var = c("age_yrs", "sex","genotype", "fev1_perc_6m_prior", "fev1_perc_pre", "change_fev_perc", "pes_total", "Antibiotic_class", "visit_dif","ShannonH_Median" ,"pcr_neg")
med_iqr_vars = c("age_yrs", "fev1_perc_6m_prior","fev1_perc_pre", "ShannonH_Median","change_fev_perc", "pes_total", "visit_dif")
cat_vars = c("sex","genotype", "antibiotic_class","pcr_neg")
t1 = CreateTableOne(data = patient_t1_df, vars = all_var, factorVars = cat_vars)
final = print(t1, nonnormal = med_iqr_vars, showAllLevels = TRUE, minMax = TRUE)

#write.csv(final, "table1.csv")
```

```{r table1 kable, echo=FALSE}
# names for rows in table order
t1_rownames = c("n",
                "Age in Years median[range]",
                "Sex n(%)",
                " ",
                "Genotype n(%)",
                " ",
                " ",
                "Baseline $FEV_1$% (highest predicted in prior 6 months) median[range]",
                "$FEV_1$% Predicted at Visit 1 median[range]",
                "Change in $FEV_1$% predicted from baseline to visit 1 median[range]",
                "Visit 1 PES Score median[range]",
                "Antibiotic Class n(%)",
                " ",
                " ",
                " ",
                " ",
                "Days Between Visits median[range]",
                "Shannon's H median[range]",
                "Positive Viral PCR at Visit 1 n(%)",
                " "
                )
t1_colnames = c("",
                "n = 19")
colnames(final) = t1_colnames
rownames(final) = t1_rownames
# remove bottom row (20), (19,2)
# change sex 1=male 2=female
final = final[-20,]
final[19,1] = ""
final[3,1] = "Male"
final[4,1] = "Female"
kable(final[-1,], caption = "Patient Characteristics at Baseline and PEx")
```

```{r misc hypothesis tests, include=FALSE}

# looking at lq_all
unique_df = outpatient_2vv %>% select(-c(last, perc_total)) %>% unique.data.frame()
df_v1 = unique_df %>% filter(visit==1) %>% select(-visit_dif, -visit)
df_v2 = unique_df %>% filter(visit==2)

df_v1 = df_v1 %>% select(sid, lq_all)
df_v2 = df_v2 %>% select(sid, lq_all)
load_df = full_join(df_v1,df_v2,by='sid') # x is visit 1, y is v2
load_df = load_df %>% select(sid, lq_all.x, lq_all.y) %>% mutate(dif = lq_all.y - lq_all.x)
mean(load_df$dif); sd(load_df$dif)
hist(df_v1$lq_all)
hist(df_v2$lq_all)

# wilcoxon signed rank test
wilcox.test(df_v1$lq_all, df_v2$lq_all, paired = TRUE)
#

rm(df_v1,df_v2,unique_df)
setwd("C:/Users/Casey/Desktop/WorkData/Microbiome Brandie 2019-2020")
write.csv(patient_t1_df, "patient_t1_df.csv")
```

### 2022-02-08 Christine Chan/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
Update pre_post_trikafta_analysis.Rmd

@@
 library(performance)
+library(AGD)
+library(rspiro)
 library(lubridate)
@@
   select(-MRN) %>% select(CFF.ID,site,DOB,Start,everything()) %>% rename(Site = "site")
-
+df$Sex = as.factor(df$Sex)
+levels(df$Sex) = c("F","M")
+# BMI category
+df$Age = pmax(as.numeric(difftime(df$Date,df$DOB,units = "days")),df$Age,na.rm = T)/365.25
+df$BMIz = y2z(df$BMI,x = df$Age,sex = df$Sex,ref=nl4.bmi)
+df$BMI_group = apply(df,1,function(r){
+  age = as.numeric(r["Age"])
+  if(!is.na(age) & age > 19){
+    bmi = as.numeric(r["BMI"])
+    group = cut(bmi,c(-Inf,18.5,25,30,35,40,Inf),
+        labels = c("Under Weight","Normal Weight","Overweight","Obese (Class I)",
+                   "Obese (Class II)","Obese (Class III+)"),right = F)
+  } else if (!is.na(age) & age <= 19){
+    z = as.numeric(r["BMIz"])
+    perc = pnorm(z)
+    group = cut(perc,c(-Inf,5,85,95,Inf),
+        labels = c("Under Weight","Normal Weight","Overweight","Obese (Class I)"),right = F)
+  } else {group = NA}
+  return(as.character(group))
+})
+df$BMI_group = as.factor(df$BMI_group)
+df$BMI_group = ordered(df$BMI_group,c("Under Weight","Normal Weight","Overweight",
+                                      "Obese (Class I)","Obese (Class II)","Obese (Class III+)"))
+# PFTs
+old <- c("Caucasian","Black","Other")
+new <- c(1,2,5)
+df$ppFEV = pctpred_GLI(df$Age,df$Height/100,gender = as.numeric(df$Sex),
+            ethnicity = new[match(df$Race, old)],
+            FEV1 = df$FEV1)
+df$ppFVC = pctpred_GLI(df$Age,df$Height/100,gender = as.numeric(df$Sex),
+            ethnicity = new[match(df$Race, old)],
+            FEV1 = df$FVC)
 ```
@@
 # Piecewise model function
-piecewise = function(outcome,data = df){
+piecewise = function(outcome,data = df,ordinal = F){
   # Add spline column
@@
   res$effect = NULL
-  #res$term[2] = "Slope Prior"
+  # Fit ordinal GEE (for example BMI group)
+  if(ordinal){
+    f = as.formula(paste0(outcome,"~Time + Change"))
+    mod = ordgee(f,data = data,id = data$CFF.ID)
+    res = summary(mod)
+  }
   # Segmented fit
-  fit1 = lme(as.formula(paste0(outcome,"~Time + Change")),random = ~1|CFF.ID,data = data,na.action = na.omit)
+  fit = lme(as.formula(paste0(outcome,"~Time + Change")),random = ~1|CFF.ID,
+            data = data,na.action = na.omit)
   seg = suppressWarnings(segmented(fit, seg.Z = ~Time, npsi=1))
@@
 
+Models did not converge with BMI converted to under weight, normal weight, etc.
+This is most likely due to lack of within-patient variability in terms of BMI group (i.e.
+people's BMIs aren't changing enough that they move from one group to another).
+
+# Insulin
+
+Most participants only have one observation of insulin dose, so these results would be the same as Maxie's previous reports.
+
+# ppFEV
+
 ```{r results='asis'}
-bmi = piecewise(outcome = "BMI",df)
-ggplotly(bmi$plot)
-kable(bmi$table,digits = 3)
-bmi$check
+ppFEV = piecewise(outcome = "ppFEV",df)
+ggplotly(ppFEV$plot)
+kable(ppFEV$table,digits = 3)
+ppFEV$check
 ```
 
-`r if(class(bmi$seg)=="lme"){"Automated R package was not able to find a better breakpoint."}`
+# ppFVC
+
+```{r results='asis'}
+ppFVC = piecewise(outcome = "ppFVC",df)
+ggplotly(ppFVC$plot)
+kable(ppFVC$table,digits = 3)
+ppFVC$check
+```

### 2022-02-08 Christine Chan/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
Update pre_post_trikafta_analysis.Rmd
rough draft done!

@@
 # Time from start and filtering
+df = df %>% group_by(CFF.ID) %>% fill(Start,.direction = "downup")
 df = df %>% mutate(Time = as.numeric(difftime(Date,Start,units = "days"))) 
@@
 # Piecewise model function
-piecewise = function(outcome,data = df,ordinal = F){
+piecewise = function(outcome,data = df,simple = F){
   # Add spline column
@@
   res$effect = NULL
-  # Fit ordinal GEE (for example BMI group)
-  if(ordinal){
-    f = as.formula(paste0(outcome,"~Time + Change"))
-    mod = ordgee(f,data = data,id = data$CFF.ID)
-    res = summary(mod)
-  }
   # Segmented fit
@@
 kable(a1c$table,digits = 3)
-a1c$check
 ```
@@
 kable(ppFEV$table,digits = 3)
-ppFEV$check
 ```
@@
 kable(ppFVC$table,digits = 3)
-ppFVC$check
 ```
+
+# CGM Metrics
+
+## Mean
+
+```{r results='asis'}
+meansg = piecewise(outcome = "average_sensor",df)
+ggplotly(meansg$plot)
+kable(meansg$table,digits = 3)
+```
+
+## Min
+
+```{r results='asis'}
+minsg = piecewise(outcome = "min_sensor",df)
+ggplotly(minsg$plot)
+kable(minsg$table,digits = 3)
+```
+
+## Max
+
+```{r results='asis'}
+maxsg = piecewise(outcome = "max_sensor",df)
+ggplotly(maxsg$plot)
+kable(maxsg$table,digits = 3)
+```
+
+## % Time > 120
+
+```{r results='asis'}
+over120 = piecewise(outcome = "percent_time_over_120",df)
+ggplotly(over120$plot)
+kable(over120$table,digits = 3)
+```
+
+## % Time > 140
+
+```{r results='asis'}
+over140 = piecewise(outcome = "percent_time_over_140",df)
+ggplotly(over140$plot)
+kable(over140$table,digits = 3)
+```
+
+## % Time > 200
+
+```{r results='asis'}
+over200 = piecewise(outcome = "percent_time_over_200",df)
+ggplotly(over200$plot)
+kable(over200$table,digits = 3)
+```
+
+## Mean excursions > 140
+
+```{r results='asis'}
+numover140 = piecewise(outcome = "avg_excur_over_140_per_day",df)
+ggplotly(numover140$plot)
+kable(numover140$table,digits = 3)
+```
+
+## Mean excursions > 200
+
+```{r results='asis'}
+numover200 = piecewise(outcome = "avg_excur_over_200_per_day",df)
+ggplotly(numover200$plot)
+kable(numover200$table,digits = 3)
+```
+
+## % Time < 70
+
+```{r results='asis'}
+under70 = piecewise(outcome = "percent_time_under_70",df)
+ggplotly(under70$plot)
+kable(under70$table,digits = 3)
+```
+
+## % Time < 60
+
+```{r results='asis'}
+under60 = piecewise(outcome = "percent_time_under_60",df)
+ggplotly(under60$plot)
+kable(under60$table,digits = 3)
+```
+
+## Average AUC per day
+
+```{r results='asis'}
+auc = piecewise(outcome = "average_auc_per_day",df)
+ggplotly(auc$plot)
+kable(auc$table,digits = 3)
+```
+
+## SD
+
+```{r results='asis'}
+sd = piecewise(outcome = "standard_deviation",df)
+ggplotly(sd$plot)
+kable(sd$table,digits = 3)
+```
+
+## MAGE
+
+```{r results='asis'}
+mage = piecewise(outcome = "r_mage",df)
+ggplotly(mage$plot)
+kable(mage$table,digits = 3)
+```

### 2022-02-08 Christine Chan/CF Metabolomics/chan_cfrd_metabolomics.Rmd
Update chan_cfrd_metabolomics.Rmd
PLSDA permutation tests

@@
 library(mixOmics)
+library(combinat)
 library(tidyverse)
+library(parallel)
 library(knitr)
@@
 
+## Four groups
+
 ```{r}
 plsda_res = splsda(X, Y_all)
-plotIndiv(plsda_res,group = df$group_category.factor,legend = TRUE,pch = 16,
-          title = "PCA Components 1 - 2")
+plotIndiv(plsda_res,group = df$group_category.factor,legend = TRUE,pch = 16)
 ```
+
+## Control and CFNGT vs. Abnormal Glycemia
+
+```{r}
+plsda_res = splsda(X, Y_two)
+plotIndiv(plsda_res,group = df$combined,legend = TRUE,pch = 16)
+```
+
+### Permutation testing
+
+```{r cache=TRUE}
+real_auc_1 = auroc(plsda_res,print = F,plot = F)$Comp1[1]
+real_auc_2 = auroc(plsda_res,print = F,plot = F)$Comp2[1]
+perms = 1000
+cl = makeCluster(8,type = "FORK")
+aucs = parLapply(cl,1:perms, function(s){
+  set.seed(s)
+  Y_perm = sample(df$combined,length(df$combined))
+  plsda_res = splsda(X, Y_perm)
+  aucs = auroc(plsda_res,print = F,plot = F)
+  comp1_auc = aucs$Comp1[1]
+  comp2_auc = aucs$Comp2[1]
+  return(c(comp1 = comp1_auc,comp2 = comp2_auc))
+})
+stopCluster(cl)
+aucs = as.data.frame(do.call(rbind,aucs))
+```
+
+Two group outcome was "shuffled" `r perms` times and sPLSDA re-fit on the permuted outcome.
+
+#### Component 1
+
+```{r}
+ggplot(aucs,aes(x = comp1)) + geom_histogram(binwidth = 0.02) + theme_bw() + 
+  geom_vline(xintercept = real_auc_1,color = "red") +
+  xlab("AUC") + ylab("Count")
+```
+
+Permutation p value for component 1 is `r round((1 + sum(aucs$comp1 > real_auc_1))/perms,3)`
+
+#### Component 2
+
+```{r}
+ggplot(aucs,aes(x = comp2)) + geom_histogram(binwidth = 0.02) + theme_bw() + 
+  geom_vline(xintercept = real_auc_2,color = "red") +
+  xlab("AUC") + ylab("Count")
+```
+
+Permutation p value for component 1 is `r round((1 + sum(aucs$comp2 > real_auc_2))/perms,3)`

### 2022-02-08 Christine Chan/CF Metabolomics/chan_cfrd_metabolomics.Rmd
Update chan_cfrd_metabolomics.Rmd
moderated t tests

@@
 ---
-title: ""
+title: "Metabolomics in CF and Healthy Controls"
 author: "Tim Vigers & Laura Pyle"
@@
 library(parallel)
+library(limma)
 library(knitr)
@@
 levels(df$combined) = c("HC & CFNGT","HC & CFNGT","CFAGT & CFRD", "CFAGT & CFRD")
+levels(df$group_category.factor) = c("HC","CFNGT","CFAGT","CFRD")
 # Remove low variance
@@
 
-Permutation p value for component 1 is `r round((1 + sum(aucs$comp1 > real_auc_1))/perms,3)`
+Permutation p value for component 1 is `r round(sum(aucs$comp1 > real_auc_1)/perms,3)`
 
@@
 
-Permutation p value for component 1 is `r round((1 + sum(aucs$comp2 > real_auc_2))/perms,3)`
+Permutation p value for component 1 is `r round(sum(aucs$comp2 >= real_auc_2)/perms,3)`
+
+# Moderated t tests
+
+## Four groups
+
+```{r warning=FALSE,message=FALSE}
+l = t(log(X))
+design = model.matrix(~group_category.factor,df)
+fit = lmFit(l,design)
+fit <- eBayes(fit)
+res = topTable(fit,number = nrow(l))
+colnames(res)[1:3] = c("CFNGT","CFAGT","CFRD")
+res = res[res$adj.P.Val < 0.05 & !is.na(res$adj.P.Val),]
+res$Compound = compounds$Compound[match(rownames(res),compounds$compound_id)]
+res = res %>% select(Compound,everything())
+kable(res,row.names = F)
+```
+
+`r nrow(res)` metabolites were different in at least one of the glycemia groups. Healthy controls are the reference group in the table above.
+
+## Control and CFNGT vs. Abnormal Glycemia
+
+```{r warning=FALSE,message=FALSE}
+design = model.matrix(~combined,df)
+fit = lmFit(l,design)
+fit <- eBayes(fit)
+res = topTable(fit,number = nrow(l))
+res = res[res$adj.P.Val < 0.05 & !is.na(res$adj.P.Val),]
+res$Compound = compounds$Compound[match(rownames(res),compounds$compound_id)]
+res = res %>% select(Compound,everything())
+kable(res,row.names = F)
+```
+
+`r nrow(res)` metabolites were different between the combined glycemia groups.
+Healthy controls and CFNGT is the reference group in the table above, and logFC refers to the log fold change between groups.

### 2022-02-09 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
reorganize Christine Chan's files

--- 2022-02-02 R Markdown Template.Rmd
@@
 ---
-title: ""
+title: "Pre-/Post-Trikafta (Prospective)"
 author: "Tim Vigers"

### 2022-02-09 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_merge.Rmd
reorganize Christine Chan's files

--- 2022-02-02 Christine Chan/Pre-post Trikafta/pre_post_trikafta_merge.Rmd
@@
 chco_cgm = read_csv("./Prepost Triakfta/Data_Cleaned/chco_cgm.csv")
-
+chco_cgm$Date = lubridate::date(parse_date_time(chco_cgm$date_cgm_placement,
+                                                orders = c("mdyHM","mdy")))
 # chco glycemic, 261 obs

### 2022-02-09 R Markdown Template.Rmd
Update R Markdown Template.Rmd

@@
 knitr::opts_knit$set(root.dir = home_dir)
+rm(home_dir)
 ```

### 2022-02-09 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
Update pre_post_trikafta_prospective_analysis.Rmd
table 1

@@
 library(performance)
+library(AGD)
 library(knitr)
@@
 ```{r}
+# Import
+df = read.csv("./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/trikafta_2022-02-8.csv",
+              na.strings = c(""))
+# Clean up
+df$Sex = factor(df$Sex..Male.1.)
+levels(df$Sex) = c("F","M")
+df = df %>% group_by(Study.ID) %>% mutate(trikafta_duration = sum(Time.on.trikafta.in.months))
+df[,c("Kalydeco","Orkambi","Symdeko")] = 
+  lapply(df[,c("Kalydeco.1","orkambi.1","tezacaftor.ivacaftor..symdeko..1")],function(c){
+    c = as.factor(c)
+    levels(c) = c("No","Yes")
+    return(c)
+  })
+# Genotype
+df$Genotype = apply(df,1,function(r){
+  if(tolower(r["CFTR.Mutation.1"]) == "f508del" & tolower(r["CFTR.Mutation.2"]) == "f508del"){
+    g = "Homozygous F508del"
+  } else if (tolower(r["CFTR.Mutation.1"]) == "f508del"){
+    g = "Heterozygous F508del"
+  } else {
+    g = "Other"
+  }
+  return(g)
+})
+# BMI z score (count everyone over 20 as 20)
+df$Age = ifelse(df$Age.at.visit. >= 20, 20, df$Age.at.visit.)
+df$BMIz = y2z(df$BMI,x = df$Age,sex = df$Sex,ref=nl4.bmi)
+```
 
+# Table 1: Demographics at Baseline
+
+For calculating BMI z scores for adults > 20, we used CDC references for 20 year olds.
+
+```{r results='asis'}
+t1 = tableby(~Age.at.visit.+Sex..Male.1.+BMIz+FEV1..+FVC....+trikafta_duration+
+               Genotype+Pancreatic.Status+Kalydeco+Orkambi+Symdeko,
+             data = df[df$Event.Name == "Visit 1",])
+summary(t1,digits = 2,labelTranslations = 
+          list(Age.at.visit. = "Age (years)",Sex..Male.1. = "Sex",
+               BMIz = "BMI z score",FEV1.. = "ppFEV1",FVC.... = "ppFVC",
+               trikafta_duration = "Duration on Trikafta (months) at Visit 2",
+               Pancreatic.Status = "Pancreatic Status"))
 ```

### 2022-02-09 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
Update pre_post_trikafta_prospective_analysis.Rmd

--- 2022-02-09 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
@@
 library(performance)
+library(AGD)
+library(pracma)
 library(knitr)
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE,warning = TRUE)
 if(Sys.info()["sysname"] == "Windows"){
@@
 ```{r}
+# Import
+df = read.csv("./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/trikafta_2022-02-8.csv",
+              na.strings = c(""))
+# Clean up
+df$Sex = factor(df$Sex..Male.1.)
+levels(df$Sex) = c("F","M")
+df = df %>% group_by(Study.ID) %>% 
+  mutate(trikafta_duration = sum(Time.on.trikafta.in.months)) %>% ungroup()
+df[,c("Kalydeco","Orkambi","Symdeko")] = 
+  lapply(df[,c("Kalydeco.1","orkambi.1","tezacaftor.ivacaftor..symdeko..1")],function(c){
+    c = as.factor(c)
+    levels(c) = c("No","Yes")
+    return(c)
+  })
+# Genotype
+df$Genotype = apply(df,1,function(r){
+  if(tolower(r["CFTR.Mutation.1"]) == "f508del" & tolower(r["CFTR.Mutation.2"]) == "f508del"){
+    g = "Homozygous F508del"
+  } else if (tolower(r["CFTR.Mutation.1"]) == "f508del"){
+    g = "Heterozygous F508del"
+  } else {
+    g = "Other"
+  }
+  return(g)
+})
+# BMI z score (count everyone over 20 as 20)
+df$Age = ifelse(df$Age.at.visit. >= 20, 20, df$Age.at.visit.)
+df$BMIz = y2z(df$BMI,x = df$Age,sex = df$Sex,ref=nl4.bmi)
+```
 
+# Table 1: Demographics at Baseline
+
+For calculating BMI z scores for adults > 20, we used CDC references for 20 year olds.
+
+```{r results='asis'}
+t1 = tableby(~Age.at.visit.+Sex..Male.1.+BMIz+FEV1..+FVC....+trikafta_duration+
+               Genotype+Pancreatic.Status+Kalydeco+Orkambi+Symdeko,
+             data = df[df$Event.Name == "Visit 1",])
+newlabels = list(Age.at.visit. = "Age (years)",Sex..Male.1. = "Sex",
+                 BMIz = "BMI z score",FEV1.. = "ppFEV1",FVC.... = "ppFVC",
+                 trikafta_duration = "Duration on Trikafta (months) at Visit 2",
+                 Pancreatic.Status = "Pancreatic Status")
+summary(t1,digits = 2,labelTranslations = newlabels)
 ```
+
+# Table 2: Pre-/Post-Trikafta Comparisons
+
+All metrics calculated based on 2 hour
+
+```{r}
+# convert Insulin to mcU/mL
+ins_cols = colnames(df)[grep("insulin",tolower(colnames(df)))]
+df[,ins_cols] = lapply(df[,ins_cols], function(c){
+  return(as.numeric(c) * 0.02869)
+})
+# cpeptide to pmol/L
+cpep_cols = colnames(df)[grep("c.peptide",tolower(colnames(df)))]
+df[,cpep_cols] = lapply(df[,cpep_cols], function(c){
+  c = as.numeric(c)/1000
+  return(c * 0.33)
+})
+# Calculate AUCs
+df$`Glucose AUC` = apply(df %>% select(Glucose.0.minute:Glucose.120.min),1,function(r){
+  trapz(c(0,10,30,60,90,120),as.numeric(r))
+})
+df$`C Peptide AUC` = apply(df %>% select(c.peptide.0.minute:c.peptide.120.min),1,function(r){
+  trapz(c(0,10,30,60,90,120),as.numeric(r))
+})
+# Other indices
+# Cpeptide index = delta cpeptide 0-30/delta glucose 0-30
+df$`C Peptide Index` = (df$c.peptide.30.min - df$c.peptide.0.minute)/
+  (df$Glucose.30.min - df$Glucose.0.minute)
+# Add labels
+newlabels["A1C.result"] = "HbA1c (%)"
+p <- paired(Event.Name ~ A1C.result+BMIz+FEV1..+FVC....+
+              `Glucose AUC`+`C Peptide AUC`+`C Peptide Index`, 
+            data = df, id = Study.ID)
+summary(p,pfootnote = T,labelTranslations = newlabels)
+```

### 2022-02-09 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
Update pre_post_trikafta_prospective_analysis.Rmd

@@
 
-All metrics calculated based on 2 hour
+All metrics calculated based on 2 hour OGTT (0, 10, 30, 60, 90, and 120 minutes).
 
-```{r}
+```{r results='asis'}
+# Glucose tolerance group
+glu_cols = c("Glucose.0.minute","Glucose.60.min","Glucose.120.min")
+df$Group = apply(df[,glu_cols],1,function(r){
+  if(r[1]<126 & r[2] < 200 & r[3] < 140){
+    return("NGT")
+  } else if (r[1] >= 126 | r[3] >= 200){
+    return("CFRD")
+  } else {
+    return("AGT")
+  }
+})
 # convert Insulin to mcU/mL
-ins_cols = colnames(df)[grep("insulin",tolower(colnames(df)))]
+ins_cols = colnames(df)[head(grep("insulin",tolower(colnames(df))),-2)]
 df[,ins_cols] = lapply(df[,ins_cols], function(c){
@@
 # cpeptide to pmol/L
-cpep_cols = colnames(df)[grep("c.peptide",tolower(colnames(df)))]
+cpep_cols = colnames(df)[head(grep("c.peptide",tolower(colnames(df))),-2)]
 df[,cpep_cols] = lapply(df[,cpep_cols], function(c){
@@
   (df$Glucose.30.min - df$Glucose.0.minute)
+# Matsuda c peptide (500,000/√[fasting glucose*fasting cpep*mean glucose*mean cpep])   [Glucose mg/dL, cpeptide pmol/L]
+glu_cols = colnames(df)[head(grep("glucose\\.\\d{1,}",tolower(colnames(df))),-2)]
+df$mean_glucose = rowMeans(df[,glu_cols])
+cpep_cols = head(cpep_cols,-2)
+df$mean_cpep = rowMeans(df[,cpep_cols])
+df$`Adapted Matsuda` = 500000/(sqrt(df$Glucose.0.minute*df$c.peptide.0.minute*
+                                      df$mean_glucose*df$mean_cpep))
+# C Pep ODI
+df$`C Peptide ODI` = df$`C Peptide Index`/df$`Adapted Matsuda`
 # Add labels
 newlabels["A1C.result"] = "HbA1c (%)"
-p <- paired(Event.Name ~ A1C.result+BMIz+FEV1..+FVC....+
-              `Glucose AUC`+`C Peptide AUC`+`C Peptide Index`, 
+newlabels["Glucose.0.minute"] = "Fasting glucose (mg/dL)"
+newlabels["Glucose.60.min"] = "1hr glucose (mg/dL)"
+newlabels["Glucose.120.min"] = "2hr glucose (mg/dL)"
+newlabels["c.peptide.0.minute"] = "Fasting C-peptide (pmol/L)"
+p <- paired(Event.Name ~ A1C.result+BMIz+FEV1..+FVC....+Group+Glucose.0.minute+
+              Glucose.60.min+Glucose.120.min+c.peptide.0.minute+
+              `Glucose AUC`+`C Peptide AUC`+`C Peptide Index`+`Adapted Matsuda`+
+              `C Peptide ODI`, 
             data = df, id = Study.ID)

### 2022-02-09 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
Update pre_post_trikafta_prospective_analysis.Rmd

--- 2022-02-09 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
@@
 
-All metrics calculated based on 2 hour
+All metrics calculated based on 2 hour OGTT (0, 10, 30, 60, 90, and 120 minutes).
 
-```{r}
+```{r results='asis'}
+# Make dataset for HOMA2 calculations - glucose in mmol/L and cpep in nmol/L
+homa = df %>% select(Study.ID,Event.Name,Glucose.0.minute:Glucose.180.min,
+                     c.peptide.0.minute:c.peptide.180.min)
+glu_cols = colnames(df)[grep("glucose\\.\\d{1,}",tolower(colnames(df)))]
+cpep_cols = colnames(df)[grep("c.peptide",tolower(colnames(df)))]
+homa[,glu_cols] = lapply(homa[,glu_cols], function(c){
+  c/18
+})
+homa[,cpep_cols] = lapply(homa[,cpep_cols], function(c){
+  c = suppressWarnings(as.numeric(c))/1000
+  c*0.33
+})
+write.csv(homa,"./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/homa_data.csv",
+          na = "",row.names = F)
+# Glucose tolerance group
+glu_cols = c("Glucose.0.minute","Glucose.60.min","Glucose.120.min")
+df$Group = apply(df[,glu_cols],1,function(r){
+  if(r[1]<126 & r[2] < 200 & r[3] < 140){
+    return("NGT")
+  } else if (r[1] >= 126 | r[3] >= 200){
+    return("CFRD")
+  } else {
+    return("AGT")
+  }
+})
 # convert Insulin to mcU/mL
-ins_cols = colnames(df)[grep("insulin",tolower(colnames(df)))]
+ins_cols = colnames(df)[head(grep("insulin",tolower(colnames(df))),-2)]
 df[,ins_cols] = lapply(df[,ins_cols], function(c){
@@
 # cpeptide to pmol/L
-cpep_cols = colnames(df)[grep("c.peptide",tolower(colnames(df)))]
+cpep_cols = colnames(df)[head(grep("c.peptide",tolower(colnames(df))),-2)]
 df[,cpep_cols] = lapply(df[,cpep_cols], function(c){
-  c = as.numeric(c)/1000
   return(c * 0.33)
@@
 })
+# Incremental AUCs
+df$`Glucose iAUC` = apply(df %>% select(Glucose.0.minute:Glucose.120.min),1,function(r){
+  r = r - r[1]
+  trapz(c(0,10,30,60,90,120),as.numeric(r))
+})
+df$`C Peptide iAUC` = apply(df %>% select(c.peptide.0.minute:c.peptide.120.min),1,function(r){
+  r = r - r[1]
+  trapz(c(0,10,30,60,90,120),as.numeric(r))
+})
 # Other indices
@@
   (df$Glucose.30.min - df$Glucose.0.minute)
+# Matsuda c peptide (500,000/√[fasting glucose*fasting cpep*mean glucose*mean cpep])   [Glucose mg/dL, cpeptide pmol/L]
+glu_cols = colnames(df)[head(grep("glucose\\.\\d{1,}",tolower(colnames(df))),-2)]
+df$mean_glucose = rowMeans(df[,glu_cols])
+df$mean_cpep = rowMeans(df[,cpep_cols])
+df$`Adapted Matsuda` = 500000/(sqrt(df$Glucose.0.minute*df$c.peptide.0.minute*
+                                      df$mean_glucose*df$mean_cpep))
+# C Pep ODI
+df$`C Peptide ODI` = df$`C Peptide Index`/df$`Adapted Matsuda`
+# iAUC ratio
+df$iAUC_ratio = df$`C Peptide iAUC`/df$`Glucose iAUC`
 # Add labels
 newlabels["A1C.result"] = "HbA1c (%)"
-p <- paired(Event.Name ~ A1C.result+BMIz+FEV1..+FVC....+
-              `Glucose AUC`+`C Peptide AUC`+`C Peptide Index`, 
+newlabels["Glucose.0.minute"] = "Fasting glucose (mg/dL)"
+newlabels["Glucose.60.min"] = "1hr glucose (mg/dL)"
+newlabels["Glucose.120.min"] = "2hr glucose (mg/dL)"
+newlabels["c.peptide.0.minute"] = "Fasting C-peptide (pmol/L)"
+newlabels["iAUC_ratio"] = "C pep iAUC:Glucose iAUC"
+p <- paired(Event.Name ~ A1C.result+BMIz+FEV1..+FVC....+Group+Glucose.0.minute+
+              Glucose.60.min+Glucose.120.min+c.peptide.0.minute+
+              `Glucose AUC`+`C Peptide AUC`+`C Peptide Index`+`Adapted Matsuda`+
+              `C Peptide ODI`+iAUC_ratio, 
             data = df, id = Study.ID)

### 2022-02-09 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
Update pre_post_trikafta_prospective_analysis.Rmd

--- 2022-02-09 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
@@
 library(knitr)
-knitr::opts_chunk$set(echo = FALSE,warning = TRUE)
+knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
 if(Sys.info()["sysname"] == "Windows"){
@@
 
-All metrics calculated based on 2 hour
+All metrics calculated based on 2 hour OGTT (0, 10, 30, 60, 90, and 120 minutes).
 
-```{r}
+```{r results='asis'}
+# Make dataset for HOMA2 calculations - glucose in mmol/L and cpep in nmol/L
+homa = df %>% select(Study.ID,Event.Name,Glucose.0.minute,c.peptide.0.minute)
+homa$Glucose.0.minute = homa$Glucose.0.minute/18
+homa$c.peptide.0.minute = homa$c.peptide.0.minute/1000*0.33
+write.csv(homa,"./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/homa_data.csv",
+          na = "",row.names = F)
+# Import HOMA2 data (calculated with https://www.dtu.ox.ac.uk/homacalculator/download.php)
+homa = read.csv("./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/homa_calculated.csv")
+df = left_join(df,homa,by = c("Study.ID", "Event.Name"))
+# Glucose tolerance group
+glu_cols = c("Glucose.0.minute","Glucose.60.min","Glucose.120.min")
+df$Group = apply(df[,glu_cols],1,function(r){
+  if(r[1]<126 & r[2] < 200 & r[3] < 140){
+    return("NGT")
+  } else if (r[1] >= 126 | r[3] >= 200){
+    return("CFRD")
+  } else {
+    return("AGT")
+  }
+})
 # convert Insulin to mcU/mL
-ins_cols = colnames(df)[grep("insulin",tolower(colnames(df)))]
+ins_cols = colnames(df)[head(grep("insulin",tolower(colnames(df))),-2)]
 df[,ins_cols] = lapply(df[,ins_cols], function(c){
@@
   (df$Glucose.30.min - df$Glucose.0.minute)
+# Matsuda c peptide (500,000/√[fasting glucose*fasting cpep*mean glucose*mean cpep])   [Glucose mg/dL, cpeptide pmol/L]
+glu_cols = colnames(df)[head(grep("glucose\\.\\d{1,}",tolower(colnames(df))),-2)]
+df$mean_glucose = rowMeans(df[,glu_cols])
+df$mean_cpep = rowMeans(df[,cpep_cols])
+df$`Adapted Matsuda` = 500000/(sqrt(df$Glucose.0.minute*df$c.peptide.0.minute*
+                                      df$mean_glucose*df$mean_cpep))
+# C Pep ODI
+df$`C Peptide ODI` = df$`C Peptide Index`/df$`Adapted Matsuda`
+# iAUC ratio
+df$iAUC_ratio = df$`C Peptide iAUC`/df$`Glucose iAUC`
 # Add labels
 newlabels["A1C.result"] = "HbA1c (%)"
-p <- paired(Event.Name ~ A1C.result+BMIz+FEV1..+FVC....+
-              `Glucose AUC`+`C Peptide AUC`+`C Peptide Index`, 
+newlabels["Glucose.0.minute"] = "Fasting glucose (mg/dL)"
+newlabels["Glucose.60.min"] = "1hr glucose (mg/dL)"
+newlabels["Glucose.120.min"] = "2hr glucose (mg/dL)"
+newlabels["c.peptide.0.minute"] = "Fasting C-peptide (pmol/L)"
+newlabels["iAUC_ratio"] = "C pep iAUC:Glucose iAUC"
+newlabels["homa2B"] = "HOMA2 %B"
+newlabels["homa2S"] = "HOMA2 %S"
+newlabels["homa2IR"] = "HOMA2 IR"
+p <- paired(Event.Name ~ A1C.result+BMIz+FEV1..+FVC....+Group+Glucose.0.minute+
+              Glucose.60.min+Glucose.120.min+c.peptide.0.minute+
+              `Glucose AUC`+`C Peptide AUC`+`C Peptide Index`+`Adapted Matsuda`+
+              `C Peptide ODI`+iAUC_ratio+homa2B+homa2S+homa2IR, 
             data = df, id = Study.ID)
@@
 ```
+
+# Figures
+
+## HbA1c
+
+```{r}
+ggplot(df,aes(x = Event.Name, y = A1C.result,fill = Event.Name)) + 
+  geom_boxplot() + geom_jitter(width = 0.25) + 
+  xlab("Visit") + ylab("HbA1c (%)") + theme_bw() + theme(legend.position = "none")
+```
+
+## Glucose
+
+```{r}
+plot_df = df %>% pivot_longer(all_of(glu_cols)) %>% select(Study.ID,Event.Name,name,value)
+plot_df$name = as.numeric(gsub("Glucose|\\.|min|minute","",plot_df$name))
+plot_df = plot_df %>% group_by(name,Event.Name) %>% 
+  summarise(m = mean(value,na.rm = T),
+            se = sd(value)/sqrt(n()),.groups = "drop")
+ggplot(plot_df,aes(x = name,y = m,color = Event.Name)) + 
+  geom_line() + geom_point() + 
+  geom_errorbar(aes(ymin=m-(1.96*se), ymax=m+(1.96*se)), width=2) +
+  xlab("Time (minutes)") + ylab("Glucose (mg/dL)") + ylim(0,200) + 
+  theme_bw()
+```
+
+## Glucose
+
+```{r}
+plot_df = df %>% pivot_longer(all_of(cpep_cols)) %>% select(Study.ID,Event.Name,name,value)
+plot_df$name = as.numeric(gsub("c.peptide|\\.|min|minute","",plot_df$name))
+plot_df = plot_df %>% group_by(name,Event.Name) %>% 
+  summarise(m = mean(value,na.rm = T),
+            se = sd(value)/sqrt(n()),.groups = "drop")
+ggplot(plot_df,aes(x = name,y = m,color = Event.Name)) + 
+  geom_line() + geom_point() + 
+  geom_errorbar(aes(ymin=m-(1.96*se), ymax=m+(1.96*se)), width=2) +
+  xlab("Time (minutes)") + ylab("C Peptide (pmol/L)") + 
+  theme_bw()
+```
+
+## Insulin
+
+```{r}
+plot_df = df %>% pivot_longer(all_of(ins_cols)) %>% select(Study.ID,Event.Name,name,value)
+plot_df$name = as.numeric(gsub("Insulin|insulin|\\.|min|minute","",plot_df$name))
+plot_df = plot_df %>% group_by(name,Event.Name) %>% 
+  summarise(m = mean(value,na.rm = T),
+            se = sd(value)/sqrt(n()),.groups = "drop")
+ggplot(plot_df,aes(x = name,y = m,color = Event.Name)) + 
+  geom_line() + geom_point() + 
+  geom_errorbar(aes(ymin=m-(1.96*se), ymax=m+(1.96*se)), width=2) +
+  xlab("Time (minutes)") + ylab("Insulin (mcU/mL)") + 
+  theme_bw()
+```

### 2022-02-15 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
Update pre_post_trikafta_prospective_analysis.Rmd
CGM comparisons and histograms

@@
 library(performance)
+library(cgmanalysis)
 library(AGD)
@@
 df$BMIz = y2z(df$BMI,x = df$Age,sex = df$Sex,ref=nl4.bmi)
+```
+
+```{r cgm data,include=FALSE}
+# Clean and calculate
+cleandata("./Christine Chan/Pre-post Trikafta Prospective/Data_Raw/CGM",
+          "./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/CGM",
+          verbose = T,removegaps = F)
+cgmvariables("./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/CGM",
+             "./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/",
+             outputname = "cgm_vars",printname = T,id_filename = T)
+# Merge
+cgm = read.csv("./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/cgm_vars.csv")
+cgm$Study.ID = str_extract(cgm$subject_id,"^[[:upper:]]{2}\\d{4}")
+cgm$Event.Name = 
+  str_extract(tolower(cgm$subject_id),"visit\\d{1}|visit \\d{1}|yr\\d{1}|year\\d{1}|year \\d{1}")
+cgm$Event.Name = paste("Visit",gsub("[[:alpha:]]|\\s","",cgm$Event.Name))
+df = left_join(df,cgm,by = c("Study.ID", "Event.Name"))
 ```
@@
 # C Pep ODI
-df$`C Peptide ODI` = df$`C Peptide Index`/df$`Adapted Matsuda`
+df$`C Peptide ODI` = df$`C Peptide Index`*df$`Adapted Matsuda`
 # iAUC ratio
@@
 newlabels["homa2IR"] = "HOMA2 IR"
-p <- paired(Event.Name ~ A1C.result+BMIz+FEV1..+FVC....+Group+Glucose.0.minute+
-              Glucose.60.min+Glucose.120.min+c.peptide.0.minute+
-              `Glucose AUC`+`C Peptide AUC`+`C Peptide Index`+`Adapted Matsuda`+
-              `C Peptide ODI`+iAUC_ratio+homa2B+homa2S+homa2IR, 
-            data = df, id = Study.ID)
+newlabels["mean_sensor"] = "Mean SG (mg/dL)"
+newlabels["max_sensor"] = "Max. SG (mg/dL)"
+newlabels["min_sensor"] = "Min SG (mg/dL)"
+newlabels["percent_time_70_180"] = "% TIR 70-180 mg/dL"
+newlabels["percent_time_over_140"] = "% Time >140 mg/dL"
+newlabels["percent_time_over_180"] = "% Time >180 mg/dL"
+newlabels["percent_time_over_250"] = "% Time >250 mg/dL"
+newlabels["percent_time_under_70"] = "% Time <70 mg/dL"
+newlabels["percent_time_under_54"] = "% Time <54 mg/dL"
+newlabels["standard_deviation"] = "SD"
+newlabels["cv"] = "CV"
+newlabels["r_mage"] = "MAGE"
+t2_vars = c("A1C.result","BMIz","FEV1..","FVC....","Group","Glucose.0.minute",
+            "Glucose.60.min","Glucose.120.min","c.peptide.0.minute",
+            "`Glucose AUC`","`C Peptide AUC`","`C Peptide Index`",
+            "`Adapted Matsuda`","`C Peptide ODI`","iAUC_ratio","homa2B","homa2S","homa2IR",
+            "mean_sensor","max_sensor","min_sensor","percent_time_70_180",
+            "percent_time_over_140","percent_time_over_180","percent_time_over_250",
+            "percent_time_under_70","percent_time_under_54","standard_deviation","cv","r_mage")
+f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
+p <- paired(f,data = df, id = Study.ID)
 summary(p,pfootnote = T,labelTranslations = newlabels)
+```
+
+```{r}
+for (v in t2_vars) {
+  v = gsub("`","",v)
+  if(is.numeric(df[,v])){
+    print(hist(df[,v]))
+  } else {
+    kable(table(df[,v]))
+  }
+}
 ```

### 2022-02-15 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
Update pre_post_trikafta_prospective_analysis.Rmd

@@
 })
+df$Group = factor(df$Group,levels = c("NGT","AGT","CFRD"),ordered = T)
 # convert Insulin to mcU/mL
@@
 f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
-p <- paired(f,data = df, id = Study.ID)
+p <- paired(f,data = df, id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F)
 summary(p,pfootnote = T,labelTranslations = newlabels)
@@
 
-```{r}
-for (v in t2_vars) {
-  v = gsub("`","",v)
-  if(is.numeric(df[,v])){
-    print(hist(df[,v]))
-  } else {
-    kable(table(df[,v]))
-  }
-}
-```
+Group was treated as an ordered variable with NGT < AGT < CFRD.
 

### 2022-03-01 Edith Zemanick/Sweat Test/p001.R
Maxie's sweat test code

@@ -1,0 +1,590 @@
+#################
+#sweat chloride
+#################
+library(tidyverse)
+library(arsenal)
+`%!in%` = Negate(`%in%`)
+
+
+setwd("//ucdenver.pvt/SOM/PEDS/RI Biostatistics Core/Shared/Shared Projects/Meier/CF/2020.10.27 - Zemanick - sweat test/data/")
+data_10 <- read.csv("Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv")
+data_11 <- subset(data_10, data_10$Age.at.Test.in.Years < 50)
+data_11$temp <- paste(data_11$PatientID, data_11$Diagnosis)
+
+data_12 <- subset(data_11, data_11$temp %!in% c("22 Non-CF","23 Non-CF",  "24 Non-CF" , "29 Non-CF" , "31 Non-CF" , "474 Non-CF" ,"483 Non-CF" ,"484 Non-CF" ,
+                                                "485 Non-CF", "505 Non-CF", "510 Non-CF", "785 Non-CF", "791 Non-CF", "792 Non-CF",
+                                                "793 Non-CF"))
+
+data_1_0 <- data_12 %>% distinct() %>% filter(Sex != "") %>% filter(Demographics.Available == "TRUE")
+
+#get rid of anyone with a time less than 2 days
+data_1_0$Age.at.Test.in.Days <- data_1_0$Age.at.Test.in.Weeks*7
+data_1_0$Na_Left <- ifelse(as.numeric(data_1_0$Na_Left) > 160, NA,as.numeric(data_1_0$Na_Left) )
+data_1_0$Na_Right <- ifelse(as.numeric(data_1_0$Na_Right) > 160, NA,as.numeric(data_1_0$Na_Right) )
+
+
+data_1 <- subset(data_1_0, data_1_0$Age.at.Test.in.Days >= 2)
+
+data_1$Diagnosis_MM <- ifelse(data_1$Diagnosis %in% c("Non-CF", "Normal"), "Non-CF",
+                              ifelse(data_1$Diagnosis %in% c("CFTR-related disorder", "CRMS"), "CRMS/CFTR", 
+                                     data_1$Diagnosis))
+
+
+#data cleaning 
+#get rid of first row for 22, 39, 
+
+
+
+#modulator data
+mod0 <- read.csv('Copy of Report CFTR Modulators.csv')
+mod <- mod0 %>% group_by(Patient.ID) %>%  arrange(Age.at.Start..years.) %>%  filter(row_number()==1)
+
[…]

### 2022-03-01 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Create sweat_test_analysis.Rmd

---
title: "Sweat electrolytes and CFTR activity in the era of CFTR modulators"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Edith Zemanick/Sweat Test"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Edith Zemanick/Sweat Test"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Edith Zemanick/Sweat Test"
}
knitr::opts_knit$set(root.dir = home_dir)
rm(home_dir)
```

```{r}
# Import data
df = read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv",na.strings = "")
# Filter per Maxie's code
df = df %>% 
  filter(Age.at.Test.in.Years < 50,
         !(Diagnosis == "Non-CF" & PatientID %in% 
             c(22,23,24,29,31,474,483,484,485,505,510,785,791,792,793))) %>% 
  distinct(.) %>% filter(Sex != "",Demographics.Available == "TRUE") %>%
  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks*7)
# Add modulator info

```

# Table 1: Patient Demographics and Sample Characteristics for sweat tests done on individuals not on CFTR modulators

```{r results='asis'}
t1_df = df %>% filter(Age.at.Test.in.Days >= 2) %>%
  group_by(PatientID) %>% arrange(Age.at.Test.in.Weeks) %>%  filter(row_number()==1)
t1 = tableby(Method ~ Sex,data = t1_df)
summary(t1)
```

### 2022-03-01 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd

@@
 rm(home_dir)
+source("~/GitHub/CF/get_genotype_class_severity.R")
 ```
@@
 df = read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv",na.strings = "")
+df$Diagnosis = factor(df$Diagnosis,
+                      levels = c("CFTR-related disorder","CRMS","Cystic Fibrosis","Non-CF","Normal"),
+                      labels = c("CFTR-related disorder","CRMS","Cystic Fibrosis","Non-CF","Non-CF"))
 # Filter per Maxie's code
@@
 # Add modulator info
-
+cftr = read.csv("./Data_Raw/Copy of Report CFTR Modulators.csv")
+df$Modulator = apply(df,1,function(r){
+  mod = cftr[cftr$Patient.ID == as.numeric(r["PatientID"]),]
+  if (any(duplicated(mod$Age.at.Start..years.))){
+    mod = mod[-which(duplicated(mod$Age.at.Start..years.)),]
+  }
+  mod = mod[order(mod$Age.at.Start..years.),]
+  age = as.numeric(r["Age.at.Test.in.Years"])
+  as.character(cut(age,c(-Inf,mod$Age.at.Start..years.,Inf),
+                   labels = c("None",mod$Modulator),right = F))
+})
+df = left_join(df,cftr,by = c("PatientID" = "Patient.ID"))
+# Genotype info
+df$geno1 = get_cf_genotype_class_severity(df$Genotypes1)
+df$geno2 = get_cf_genotype_class_severity(df$Genotypes2)
+df$geno1[df$Diagnosis == "Non-CF"] = "Not CF"
+df$geno2[df$Diagnosis == "Non-CF"] = "Not CF"
+df$geno_risk = paste(df$geno1,df$geno2,sep = "_")
+df$geno_risk = factor(df$geno_risk,
+                      levels = c("Mild_Severe","Mild_Unknown","Not CF_Not CF","Severe_Mild","Mild_Mild",
+                                 "Severe_Severe","Severe_Unknown","Unknown_Severe","Unknown_Unknown"),
+                      labels = c("Low","Low","Not CF","Low","Low","High","Unknown","Unknown","Unknown"))
 ```
@@
   group_by(PatientID) %>% arrange(Age.at.Test.in.Weeks) %>%  filter(row_number()==1)
-t1 = tableby(Method ~ Sex,data = t1_df)
+t1 = tableby(Method ~ Sex+Age.at.Test.in.Years+geno_risk,data = t1_df)
 summary(t1)

### 2022-03-01 R Markdown Template.Rmd
Update R Markdown Template.Rmd

--- 2022-02-02 R Markdown Template.Rmd
@@
 knitr::opts_knit$set(root.dir = home_dir)
+rm(home_dir)
+source("~/GitHub/CF/get_genotype_class.R")
 ```

### 2022-03-13 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd

--- 2022-03-01 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
@@
 library(knitr)
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
 if(Sys.info()["sysname"] == "Windows"){
@@
 # Add modulator info
-
+cftr = read.csv("./Data_Raw/Copy of Report CFTR Modulators.csv")
+df$Modulator = apply(df,1,function(r){
+  mod = cftr[cftr$Patient.ID == as.numeric(r["PatientID"]),]
+  if (any(duplicated(mod$Age.at.Start..years.))){
+    mod = mod[-which(duplicated(mod$Age.at.Start..years.)),]
+  }
+  mod = mod[order(mod$Age.at.Start..years.),]
+  age = as.numeric(r["Age.at.Test.in.Years"])
+  as.character(cut(age,c(-Inf,mod$Age.at.Start..years.,Inf),
+                   labels = c("None",mod$Modulator),right = F))
+})
+# Genotype info
+df$geno1 = get_cf_genotype_class_severity(df$Genotypes1)
+df$geno2 = get_cf_genotype_class_severity(df$Genotypes2)
+df$geno1[df$Diagnosis == "Non-CF"] = "Not CF"
+df$geno2[df$Diagnosis == "Non-CF"] = "Not CF"
+df$geno_risk = paste(df$geno1,df$geno2,sep = "_")
+df$geno_risk = factor(df$geno_risk,
+                      levels = c("Mild_Severe","Mild_Unknown","Not CF_Not CF",
+                                 "Severe_Mild","Mild_Mild",
+                                 "Severe_Severe","Severe_Unknown","Unknown_Severe",
+                                 "Unknown_Unknown"),
+                      labels = c("Low","Low","Not CF","Low","Low","High","Unknown",
+                                 "Unknown","Unknown"))
+# Merge left and right values
+df$Amount = apply(df[,c("Amount","Amount_Left","Amount_Right")],1,function(r){
+  amt = suppressWarnings(as.numeric(as.character(r)))
+  mean(amt,na.rm = T)
+})
+df$Cl = apply(df[,c("Cl","Cl_Left","Cl_Right")],1,function(r){
+  cl = suppressWarnings(as.numeric(as.character(r)))
+  mean(cl,na.rm = T)
+})
+df$Na = apply(df[,c("Na","Na_Left","Na_Right")],1,function(r){
+  na = suppressWarnings(as.numeric(as.character(r)))
+  mean(na,na.rm = T)
+})
+df$K = apply(df[,c("K","K_Left","K_Right")],1,function(r){
+  k = suppressWarnings(as.numeric(as.character(r)))
+  mean(k,na.rm = T)
+})
 ```
 
-# Table 1: Patient Demographics and Sample Characteristics for sweat tests done on individuals not on CFTR modulators
+# Table 1: Patient Demographics and Sample Characteristics
+
+For those with two sweat electrolyte measures (left and right), the average of the two was used.
 
@@
   group_by(PatientID) %>% arrange(Age.at.Test.in.Weeks) %>%  filter(row_number()==1)
-t1 = tableby(Method ~ Sex,data = t1_df)
-summary(t1)
+t1 = tableby(Diagnosis ~ Sex+Age.at.Test.in.Years+geno_risk+Amount+Cl+Na+K,
+             data = t1_df)
+summary(t1,labelTranslations = 
+          list(Age.at.Test.in.Years = "Age (Years)",geno_risk = "Genotype Severity"))
 ```
+
+# Chloride and Diagnosis
+
+## Table
+
+```{r}
+df$Cl_group = cut(df$Cl,c(0,30,60,Inf),right = F)
+t = table(df$Cl_group,df$Diagnosis)
+t = rownames_to_column(as.data.frame.matrix(t))
+colnames(t)[1] = "Chloride Range"
+kable(t,row.names = F)
+```
+
+## Plot
+
+```{r}
+ggplot(df,aes(x = Diagnosis,y = Cl)) + geom_boxplot() + theme_bw()
+```
+
+```{r}
+# Write groups of interest
+write.csv(df %>% filter(Diagnosis == "Non-CF",Cl >= 60),na = "",
+          file = "./Data_Cleaned/non_cf_cl_over_60.csv",row.names = F)
+write.csv(df %>% filter(Diagnosis == "Cystic Fibrosis",Cl < 60),na = "",
+          file = "./Data_Cleaned/cf_cl_under_60.csv",row.names = F)
+write.csv(df %>% filter(Diagnosis == "Cystic Fibrosis",Cl >= 30 & Cl < 60),na = "",
+          file = "./Data_Cleaned/cf_cl_30_59.csv",row.names = F)
+```

### 2022-03-14 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
Update pre_post_trikafta_analysis.Rmd

---
title: "Pre-/Post-Trikafta in CFRD"
author: "Tim Vigers & Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(lmerTest)
library(nlme)
library(broom.mixed)
library(segmented)
library(performance)
library(AGD)
library(rspiro)
library(lubridate)
library(plotly)
library(knitr)
knitr::opts_chunk$set(echo = FALSE,fig.height = 8,fig.width = 10)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
}
knitr::opts_knit$set(root.dir = home_dir)
```

```{r include=FALSE}
# Import cleaned data from Casey
df = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/analysis_data.csv",na.strings = "NA")
df$X = NULL
# Fill in MRN and CFF ID for everyone
df = df %>% 
  group_by(CFF.ID) %>% fill(MRN,.direction = "downup") %>% 
  group_by(MRN) %>% fill(CFF.ID,.direction = "downup")
# Assume that blanks for pancreatic and G tube indicate "no"
df$Hispanic.Latinx[is.na(df$Hispanic.Latinx)] = FALSE
df$Pancreatic.Status[is.na(df$Pancreatic.Status)] = "Sufficient"
df$G.tube.in.past.12.months[is.na(df$G.tube.in.past.12.months)] = "No"
# Convert dates
date_vars = c("DOB","Start","Date")
df[,date_vars] = lapply(df[,date_vars], function(d){
  parse_date_time(d,orders = c("mdy","ymd","dmy"),quiet = T)
})
# Time from start and filtering
df = df %>% group_by(CFF.ID) %>% fill(Start,.direction = "downup")
df = df %>% mutate(Time = as.numeric(difftime(Date,Start,units = "days"))) 
# Add UM (they did not provide dates, so need Time column)
um_a1cs = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_a1cs.csv",na.strings = "")
um_bmi = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_bmi.csv",na.strings = "")
um_demo = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_demo.csv",na.strings = "")
um_ogtt = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_ogtt.csv",na.strings = "")
um_pfts = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_pfts.csv",na.strings = "")
um = full_join(um_a1cs,um_bmi)
um = full_join(um,um_ogtt)
um = full_join(um,um_pfts)
um = left_join(um,um_demo)
um = um %>% filter(!is.na(CFRD.Status))
df = full_join(df,um)
# Filter and fill constants
df = df %>% group_by(CFF.ID) %>% fill(DOB,Start:CFRD.Diagnosis.Date,.direction = "downup") %>%
  ungroup() %>% filter(!is.na(Time),abs(Time) <= 365,!is.na(CFF.ID),
                       (!is.na(CFRD.Diagnosis.Date) | !is.na(CFRD.Status) | site == "UW")) %>%
  select(-MRN) %>% select(CFF.ID,site,DOB,Start,everything()) %>% rename(Site = "site")
df$Sex = as.factor(df$Sex)
levels(df$Sex) = c("F","M")
# BMI category
df$Age = pmax(as.numeric(difftime(df$Date,df$DOB,units = "days")),df$Age,na.rm = T)/365.25
# Treat adults as age = 20
df$Age_for_BMI = ifelse(df$Age > 20,20,df$Age)
df$BMIz = y2z(df$BMI,x = df$Age_for_BMI,sex = df$Sex,ref=nl4.bmi)
df$BMIz[is.infinite(df$BMIz)] = NA
df$BMI_group = apply(df,1,function(r){
  age = as.numeric(r["Age"])
  if(!is.na(age) & age > 19){
    bmi = as.numeric(r["BMI"])
    group = cut(bmi,c(-Inf,18.5,25,30,35,40,Inf),
        labels = c("Under Weight","Normal Weight","Overweight","Obese (Class I)",
                   "Obese (Class II)","Obese (Class III+)"),right = F)
  } else if (!is.na(age) & age <= 19){
    z = as.numeric(r["BMIz"])
    perc = pnorm(z)
    group = cut(perc,c(-Inf,5,85,95,Inf),
        labels = c("Under Weight","Normal Weight","Overweight","Obese (Class I)"),right = F)
  } else {group = NA}
  return(as.character(group))
})
df$BMI_group = as.factor(df$BMI_group)
df$BMI_group = ordered(df$BMI_group,c("Under Weight","Normal Weight","Overweight",
                                      "Obese (Class I)","Obese (Class II)","Obese (Class III+)"))
# PFTs
old <- c("Caucasian","Black","Other")
new <- c(1,2,5)
df$ppFEV = pctpred_GLI(df$Age,df$Height/100,gender = as.numeric(df$Sex),
            ethnicity = new[match(df$Race, old)],
            FEV1 = df$FEV1)
df$ppFVC = pctpred_GLI(df$Age,df$Height/100,gender = as.numeric(df$Sex),
            ethnicity = new[match(df$Race, old)],
            FEV1 = df$FVC)
```

```{r}
outcome = "HbA1c"
data = df
# Piecewise model function
piecewise = function(outcome = "HbA1c",data = df,simple = F){
  # Add spline column
  data$Change = ifelse(data$Time < 0,0,1)
  data$Change = data$Change * data$Time
  # Model
  f = as.formula(paste0(outcome,"~Time + Change + (1|CFF.ID)"))
  mod = lmer(f,data = data)
  res = tidy(mod,"fixed")
  res$effect = NULL
  # Segmented fit
  fit = lme(as.formula(paste0(outcome,"~Time + Change")),random = ~1|CFF.ID,
            data = data,na.action = na.omit)
  seg = suppressWarnings(segmented(fit, seg.Z = ~Time, npsi=1))
  # Model check if necessary
  check = check_model(mod)
  # Coordinates for model plotting
  x1_start = -365
  y1_start = res$estimate[1] + res$estimate[2]*x1_start
  x1_end = 0
  y1_end = res$estimate[1] 
  x2_start = 0
  y2_start = res$estimate[1] 
  x2_end = 365
  y2_end = res$estimate[1] + (res$estimate[2]+res$estimate[3])*x2_end
  # Plot
  p = ggplot(data,aes_string(x = "Time",y = outcome,group = "CFF.ID",color = "Site")) +
    geom_point(alpha = 0.5) + geom_line(alpha = 0.2) +
    annotate("segment",x=x1_start,y=y1_start,xend=x1_end,yend=y1_end,color = "blue",size = 1) +
    annotate("segment",x=x2_start,y=y2_start,xend=x2_end,yend=y2_end,color = "blue",size = 1) +
    xlab("Days from Trikafta Start") + theme_bw()
  # Interactive plot and model results
  return(list(plot = ggplotly(p),table = res,check = check,seg = seg))
}
```

# Methods

For each outcome, we fit a piecewise linear mixed effects model with random intercept for each person and a "knot" at 0. This approach can be thought of as fitting two separate linear models, while ensuring that the two regression lines meet at Trikafta start. In the model results, "Slope Prior" indicates the slope prior to start of treatment (i.e. the average change in outcome for each "1 unit increase" in time). The "Change" term indicates the *change in slope* after medication start. To find the slope post-Trikafta, you add the "Slope Prior" and "Change" estimates.

The random intercept model was selected because it performed better (by AIC) than a model with random intercept and continuous AR1 correlation structure. Random slope models did not converge for some outcomes, so were not used for the sake of consistency between results.

# HbA1c

```{r results='asis'}
a1c = piecewise(outcome = "HbA1c",data = df)
ggplotly(a1c$plot)
kable(a1c$table,digits = 3)
```

`r if(class(a1c$seg)=="lme"){"Automated R package was not able to find a better breakpoint."}`

# BMI

Models did not converge with BMI converted to under weight, normal weight, etc. This is most likely due to lack of within-patient variability in terms of BMI group (i.e. people's BMIs aren't changing enough that they move from one group to another).

# Insulin

Most participants only have one observation of insulin dose, so these results would be the same as Maxie's previous reports.

# ppFEV

```{r results='asis'}
ppFEV = piecewise(outcome = "ppFEV",data = df)
ggplotly(ppFEV$plot)
kable(ppFEV$table,digits = 3)
```

# ppFVC

```{r results='asis'}
ppFVC = piecewise(outcome = "ppFVC",data = df)
ggplotly(ppFVC$plot)
kable(ppFVC$table,digits = 3)
```

# CGM Metrics

## Mean

```{r results='asis'}
meansg = piecewise(outcome = "average_sensor",data = df)
ggplotly(meansg$plot)
kable(meansg$table,digits = 3)
```

## Min

```{r results='asis'}
minsg = piecewise(outcome = "min_sensor",data = df)
ggplotly(minsg$plot)
kable(minsg$table,digits = 3)
```

## Max

```{r results='asis'}
maxsg = piecewise(outcome = "max_sensor",data = df)
ggplotly(maxsg$plot)
kable(maxsg$table,digits = 3)
```

## % Time > 120

```{r results='asis'}
over120 = piecewise(outcome = "percent_time_over_120",data = df)
ggplotly(over120$plot)
kable(over120$table,digits = 3)
```

## % Time > 140

```{r results='asis'}
over140 = piecewise(outcome = "percent_time_over_140",data = df)
ggplotly(over140$plot)
kable(over140$table,digits = 3)
```

## % Time > 200

```{r results='asis'}
over200 = piecewise(outcome = "percent_time_over_200",data = df)
ggplotly(over200$plot)
kable(over200$table,digits = 3)
```

## Mean excursions > 140

```{r results='asis'}
numover140 = piecewise(outcome = "avg_excur_over_140_per_day",data = df)
ggplotly(numover140$plot)
kable(numover140$table,digits = 3)
```

## Mean excursions > 200

```{r results='asis'}
numover200 = piecewise(outcome = "avg_excur_over_200_per_day",data = df)
ggplotly(numover200$plot)
kable(numover200$table,digits = 3)
```

## % Time < 70

```{r results='asis'}
under70 = piecewise(outcome = "percent_time_under_70",data = df)
ggplotly(under70$plot)
kable(under70$table,digits = 3)
```

## % Time < 60

```{r results='asis'}
under60 = piecewise(outcome = "percent_time_under_60",data = df)
ggplotly(under60$plot)
kable(under60$table,digits = 3)
```

## Average AUC per day

```{r results='asis'}
auc = piecewise(outcome = "average_auc_per_day",data = df)
ggplotly(auc$plot)
kable(auc$table,digits = 3)
```

## SD

```{r results='asis'}
sd = piecewise(outcome = "standard_deviation",data = df)
ggplotly(sd$plot)
kable(sd$table,digits = 3)
```

## MAGE

```{r results='asis'}
mage = piecewise(outcome = "r_mage",data = df)
ggplotly(mage$plot)
kable(mage$table,digits = 3)
```

### 2022-03-14 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
Update pre_post_trikafta_analysis.Rmd

@@
             FEV1 = df$FVC)
+# Before or after start
+df$Before_After = cut(df$Time,c(-Inf,0,Inf),labels = c("Before","After"),right = F)
+# Most recent before/after
+most_recent = df %>% group_by(CFF.ID,Before_After) %>% slice_min(abs(Time))
 ```
 
+# Pre-post
+
 ```{r}
-outcome = "HbA1c"
-data = df
+summary(tableby( ~ ppFEV + ppFVC + BMIz,data = df[!is.na(df$average_sensor),]))
+```
+
+# Piecewise Models
+
+```{r}
 # Piecewise model function
@@
   res$effect = NULL
-  # Segmented fit
-  fit = lme(as.formula(paste0(outcome,"~Time + Change")),random = ~1|CFF.ID,
-            data = data,na.action = na.omit)
-  seg = suppressWarnings(segmented(fit, seg.Z = ~Time, npsi=1))
   # Model check if necessary
@@
   # Interactive plot and model results
-  return(list(plot = ggplotly(p),table = res,check = check,seg = seg))
+  return(list(plot = ggplotly(p),table = res,check = check))
 }

### 2022-03-14 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd

@@
           file = "./Data_Cleaned/non_cf_cl_over_60.csv",row.names = F)
+write.csv(df %>% filter(Diagnosis == "Non-CF",Cl >= 30 & Cl < 60),na = "",
+          file = "./Data_Cleaned/non_cf_cl_30_59.csv",row.names = F)
 write.csv(df %>% filter(Diagnosis == "Cystic Fibrosis",Cl < 60),na = "",

### 2022-03-15 Christine Chan/CFRD Screening Survey/cfrd_screening_analysis.Rmd
Create cfrd_screening_analysis.Rmd

--- 2022-03-01 R Markdown Template.Rmd
@@
 ---
-title: ""
+title: "CFRD Screening Survey"
 author: "Tim Vigers"

### 2022-03-28 Christine Chan/CFRD Screening Survey/cfrd_screening_analysis.Rmd
Create cfrd_screening_analysis.Rmd

--- 2022-02-09 R Markdown Template.Rmd
@@
 library(performance)
+library(broom)
+library(gtsummary)
 library(knitr)
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE,message = FALSE)
 if(Sys.info()["sysname"] == "Windows"){
@@
 ```{r}
+endo = read.csv("./Christine Chan/CFRD Screening/Data_Raw/endo.csv",na.strings = "")
+non_endo = read.csv("./Christine Chan/CFRD Screening/Data_Raw/non_endo.csv",na.strings = "")
+```
 
+# Hypothesis Tests
+
+## Centers with at least 50% of patients with state-funded health insurance have less than50% OGTT completion rates (CF center directors survey, questions #4 and #6).
+
+```{r}
+# Combine levels
+non_endo$state_funded = factor(non_endo$What.percent.of.your.CF.patients.have.state.funded.health.insurance..i.e...Medicaid..etc...)
+levels(non_endo$state_funded) = c('less than 50%','more than 50%','less than 50%','more than 50%',NA)
+non_endo$ogtt_completion = factor(non_endo$What.percent.of.eligible.CF.patients....10.years.old.at.your.center.were.screened.for.CFRD.with.an.OGTT.in.the.past.year.)
+levels(non_endo$ogtt_completion) = c('less than 50%','more than 50%','less than 50%','more than 50%',NA)
 ```
+
+### Cross-tab
+
+```{r}
+# Table
+tbl_cross(non_endo, row=state_funded, col=ogtt_completion, percent="row",missing = "no")
+```
+
+### Model results
+
+```{r}
+# Logistic model
+m = glm(ogtt_completion ~ state_funded,data = non_endo,family = binomial)
+kable(tidy(m,exponentiate = T),digits = 3)
+```
+
+There was no association between state-funded health insurance and OGTT completion (p = `r round(summary(m)$coefficients[2,4],3)`).
+
+## Centers with an identified Endocrinologist in coordinated clinic have >50% OGTT completion rates (CF center directors survey, questions #6 and #8)
+
+For this question, all "Other" responses were counted as "No."
+
+```{r}
+# Combine levels
+non_endo$endocrinologist = factor(non_endo$Does.your.center.have.an.identified.Endocrinologist.s..to.whom.you.refer.your.CF.patients.)
+levels(non_endo$endocrinologist) = c('No','No','Yes')
+```
+
+### Cross-tab
+
+```{r}
+# Table
+tbl_cross(non_endo, row=endocrinologist, col=ogtt_completion, percent="row",missing = "no")
+```
+
+### Model results
+
+```{r}
+# Logistic model
+m = glm(ogtt_completion ~ endocrinologist,data = non_endo,family = binomial)
+kable(tidy(m,exponentiate = T),digits = 3)
+```
+
+There was no association between having an endocrinologist and OGTT completion (p = `r round(summary(m)$coefficients[2,4],3)`).
+
+## Centers with a protocol or standardized method for ordering OGTTs are more likely to have >50% OGTT completion rates (CF center directors survey, questions #6 and #9)
+
+```{r}
+# Combine levels
+non_endo$protocol = factor(non_endo$Do.you.have.a.protocol.or.standardized.method.for.how.OGTTs.are.ordered.at.your.center.)
+levels(non_endo$protocol) = c('No','No','Yes','Yes')
+```
+
+### Cross-tab
+
+```{r}
+# Table
+tbl_cross(non_endo, row=protocol, col=ogtt_completion, percent="row",missing = "no")
+```
+
+### Model results
+
+Unable to perform a statistical tests because there were 0 centers without a protocol and over 50% OGTT completion.
+
+## Centers with an identified role for placing orders for OGTTs are more likely to have >50% OGTT completion rates (test for 1 role response vs 2 responses vs 3 vs 4 or more); CF center directors survey, questions #6 and #10)
+
+```{r}
+# Combine levels
+non_endo$role = apply(non_endo,1,function(r){
+  s = strsplit(as.character(r['Who.typically.places.the.orders..even.if.a.co.signer.is.needed..for.OGTTs.at.your.center...please.select.all.that.apply..']),",")
+  return(length(s[[1]]))
+})
+non_endo$role = factor(non_endo$role)
+```
+
+### Cross-tab
+
+```{r}
+# Table
+tbl_cross(non_endo, row=role, col=ogtt_completion, percent="row",missing = "no")
+```
+
+### Model results
+
+```{r}
+# Logistic model
+m = glm(ogtt_completion ~ role,data = non_endo,family = binomial)
+kable(tidy(m,exponentiate = T),digits = 3)
+```
+
+There was no association between OGTT ordering roles and OGTT completion.
+
+## Centers with an identified role for contacting patients to schedule the OGTT are more likely to have >50% completion rates (test for 1 role response vs 2 responses vs 3 vs 4 or more); CF center directors survey, questions #6 and #11
+
+```{r}
+# Combine levels
+non_endo$role = apply(non_endo,1,function(r){
+  s = strsplit(as.character(r['Who.typically.contacts.patients.to.schedule.the.OGTT...please.select.all.that.apply..']),",")
+  return(length(s[[1]]))
+})
+non_endo$role = factor(non_endo$role)
+```
+
+### Cross-tab
+
+```{r}
+# Table
+tbl_cross(non_endo, row=role, col=ogtt_completion, percent="row",missing = "no")
+```
+
+### Model results
+
+Unable to perform a statistical test due to 0 centers with 3 roles and < 50% completion.
+
+## Centers with an identified role for providing directions for patients re: how to prepare for the OGTT are more likely to have >50% completion rates (test for 1 role response vs 2 responses vs 3 vs 4 or more); CF center directors survey, questions #6 and #12
+
+```{r}
+# Combine levels
+non_endo$role = apply(non_endo,1,function(r){
+  s = strsplit(as.character(r['Who.typically.provides.directions.to.patients.regarding.how.to.prepare.for.the.OGTT...please.select.all.that.apply...']),",")
+  return(length(s[[1]]))
+})
+non_endo$role = factor(non_endo$role)
+```
+
+### Cross-tab
+
+```{r}
+# Table
+tbl_cross(non_endo, row=role, col=ogtt_completion, percent="row",missing = "no")
+```
+
+### Model results
+
+```{r}
+# Logistic model
+m = glm(ogtt_completion ~ role,data = non_endo,family = binomial)
+kable(tidy(m,exponentiate = T),digits = 3)
+```
+
+There was no association between OGTT direction roles and OGTT completion.
+
+## Centers that perform OGTTs in clinic are more likely to have >50% OGTT completion rates (CF center directors survey, questions #6 and #13)
+
+```{r}
+# Combine levels
+non_endo$location = apply(non_endo,1,function(r){
+  s = as.character(r['In.what.setting.s..do.your.patients.have.OGTTs.done..please.select.all.that.apply...'])
+  if(grepl("In our CF clinic",s)){
+    return("In clinic")
+  } else {
+    return("Other")
+  }
+})
+non_endo$location = as.factor(non_endo$location)
+```
+
+### Cross-tab
+
+```{r}
+# Table
+tbl_cross(non_endo, row=location, col=ogtt_completion, percent="row",missing = "no")
+```
+
+### Model results
+
+```{r}
+# Logistic model
+m = glm(ogtt_completion ~ state_funded,data = non_endo,family = binomial)
+kable(tidy(m,exponentiate = T),digits = 3)
+```
+
+There was no association between in-clinic OGTTs and OGTT completion (p = `r round(summary(m)$coefficients[2,4],3)`).
+
+## Centers that always/almost always provide reminders to patients for OGTTs are more likely to have >50% OGTT completion rates (CF center directors survey, questions #6 and #15)
+
+```{r}
+# Combine levels
+non_endo$reminders = as.factor(non_endo$How.often.does.your.center.provide.routine.reminders.to.patients.when.an.OGTT.is.upcoming..due..and.or.overdue..)
+levels(non_endo$reminders) = c("Other","Yes, always or almost always","Other")
+```
+
+### Cross-tab
+
+```{r}
+# Table
+tbl_cross(non_endo, row=reminders, col=ogtt_completion, percent="row",missing = "no")
+```
+
+### Model results
+
+```{r}
+# Logistic model
+m = glm(ogtt_completion ~ reminders,data = non_endo,family = binomial)
+kable(tidy(m,exponentiate = T),digits = 3)
+```
+
+There was a significant association between reminders and OGTT completion (p = `r round(summary(m)$coefficients[2,4],3)`).
+Centers with regular reminders were approximately 5 times more likely to have an over 50% OGTT completion rate.
+
+# Combined data
+
+```{r}
+endo = read.csv("./Christine Chan/CFRD Screening/Data_Raw/endo_checkboxes.csv",na.strings = "")
+non_endo = read.csv("./Christine Chan/CFRD Screening/Data_Raw/non_endo_checkboxes.csv",na.strings = "")
+```
+
+## Alternative OGTT sources
+
+```{r}
+# Combine and format
+endo_alt = factor(endo$How.often.have.you.recommended.using.an.alternative.glucose.source.for.OGTT.testing..i.e...juice..pop.soda..jelly.beans..licorice..etc...)
+non_endo_alt = factor(non_endo$How.often.does.your.center.use.alternative.glucose.sources.for.OGTT.testing..i.e...juice..pop.soda..jelly.beans..licorice..etc...)
+length(endo_alt) = length(non_endo_alt)
+levels(non_endo_alt) = c(NA,"Never","Rarely","Sometimes")
+plot_df = data.frame(list("Endocrinologist" = endo_alt,"Directors" = non_endo_alt))
+plot_df = plot_df %>% pivot_longer(cols = c("Endocrinologist","Directors")) %>%
+  group_by(name) %>% count(value) %>% drop_na(.)
+# Plot
+ggplot(plot_df,aes(x = value,y = n,group = name,fill = name)) + 
+  geom_bar(stat="identity",position=position_dodge()) +
+  theme_bw() + theme(legend.title = element_blank()) + 
+  xlab("Alternative Glucose Sources for OGTT")
+```
+
+Center directors and endocrinologists were not significantly different in terms of how often they recommend alternative glucose sources (p = 0.823).
+
+## OGTT Time Points Collected
+
+```{r}
+# Combine and format
+endo_alt = factor(endo$What.OGTT.glucose.time.points.are.collected.)
+non_endo_alt = non_endo %>% 
+  select(What.OGTT.time.points.for.glucose.levels.does.your.center.collect...choice.0..30..60..90..and.120.min.,
+         What.OGTT.time.points.for.glucose.levels.does.your.center.collect...choice.0..60..and.120.min.,
+         What.OGTT.time.points.for.glucose.levels.does.your.center.collect...choice.0.and.120.min.)
+non_endo_alt = factor(apply(non_endo_alt,1,function(r){
+  w = which(as.character(r) == "Checked")
+  if(length(w)>0){
+    names(r)[w[1]]
+  } else {NA}
+}))
+levels(non_endo_alt) = c("0, 30, 60, 90, and 120 min","0, 60, and 120 min","0 and 120 min")
+length(endo_alt) = length(non_endo_alt)
+plot_df = data.frame(list("Endocrinologist" = endo_alt,"Directors" = non_endo_alt))
+plot_df = plot_df %>% pivot_longer(cols = c("Endocrinologist","Directors")) %>%
+  group_by(name) %>% count(value) %>% drop_na(.)
+# Plot
+ggplot(plot_df,aes(x = value,y = n,group = name,fill = name)) + 
+  geom_bar(stat="identity",position=position_dodge()) +
+  theme_bw() + theme(legend.title = element_blank()) + 
+  xlab("OGTT Time Points Collected")
+```
+
+Center directors and endocrinologists were significantly different in terms of which timepoints were collected (p = 0.004).

### 2022-03-28 R Markdown Template.Rmd
Update R Markdown Template.Rmd

--- 2022-02-02 R Markdown Template.Rmd
@@
 knitr::opts_knit$set(root.dir = home_dir)
+rm(home_dir)
+source("~/GitHub/CF/get_genotype_class_severity.R")
 ```

### 2022-03-28 Christine Chan/CFRD Screening Survey/cfrd_screening_analysis.Rmd
Update cfrd_screening_analysis.Rmd

@@
 Center directors and endocrinologists were significantly different in terms of which timepoints were collected (p = 0.004).
+
+## Other Diabetes Screening Methods
+
+```{r}
+# Combine and format
+endo_alt = endo %>% select(Have.you.used.any.of.the.following.tests.to.screen.diagnose.CFRD.in.the.outpatient.setting...please.select.all.that.apply...choice.HbA1c.:Have.you.used.any.of.the.following.tests.to.screen.diagnose.CFRD.in.the.outpatient.setting...please.select.all.that.apply...choice.Other..specify.in.comment..) %>%
+  pivot_longer(everything()) %>% group_by(name) %>% summarise(n = sum(value == "Checked")) 
+endo_alt$name = factor(endo_alt$name)
+levels(endo_alt$name) = c("CGM","Fasting glucose","2 hr. glucose","HbA1c","Home fingerstick",
+                          "Other","Random glucose")
+endo_alt$type = "Endocrinologist"
+
+non_endo_alt = non_endo %>% 
+  select(Has.your.center.used.any.of.the.following.tests.to.screen.diagnose.CFRD.in.the.outpatient.setting...please.select.all.that.apply...choice.HbA1c.:
+           Has.your.center.used.any.of.the.following.tests.to.screen.diagnose.CFRD.in.the.outpatient.setting...please.select.all.that.apply...choice.None.of.the.above.) %>%
+  pivot_longer(everything()) %>% group_by(name) %>% summarise(n = sum(value == "Checked")) 
+non_endo_alt$name = factor(non_endo_alt$name)
+levels(non_endo_alt$name) = c("CGM","Fasting glucose","2 hr. glucose","HbA1c","Home fingerstick",
+                              "Other","Other","Random glucose")
+non_endo_alt = non_endo_alt %>% group_by(name) %>% summarise(n = sum(n))
+non_endo_alt$type = "Directors"
+
+plot_df = data.frame(rbind(endo_alt,non_endo_alt))
+t = plot_df %>% pivot_wider(id_cols = name,values_from = n,names_from = type) %>%
+  column_to_rownames(.,var = "name")
+# Plot
+ggplot(plot_df,aes(x = name,y = n,group = type,fill = type)) + 
+  geom_bar(stat="identity",position=position_dodge()) +
+  theme_bw() + theme(legend.title = element_blank()) + 
+  xlab("Other Diabetes Screening Methods")
+```
+
+Center directors and endocrinologists were not significantly different in terms of how often they use alternative screening measures (p = 0.116).

### 2022-03-30 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
still need DXA measure for Christine

--- 2022-02-09 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
@@
 df$BMIz = y2z(df$BMI,x = df$Age,sex = df$Sex,ref=nl4.bmi)
+# Weight z score
+df$weightz = y2z(df$Weight..kg.,x = df$Age,sex = df$Sex,ref=nl4.wgt)
+```
+
+```{r cgm data,include=FALSE}
+# Clean and calculate
+cleandata("./Christine Chan/Pre-post Trikafta Prospective/Data_Raw/CGM",
+          "./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/CGM",
+          removegaps = F)
+cgmvariables("./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/CGM",
+             "./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/",
+             outputname = "cgm_vars",id_filename = T)
+# Merge
+cgm = read.csv("./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/cgm_vars.csv")
+cgm$Study.ID = str_extract(cgm$subject_id,"^[[:upper:]]{2}\\d{4}")
+cgm$Event.Name = 
+  str_extract(tolower(cgm$subject_id),"visit\\d{1}|visit \\d{1}|yr\\d{1}|year\\d{1}|year \\d{1}")
+cgm$Event.Name = paste("Visit",gsub("[[:alpha:]]|\\s","",cgm$Event.Name))
+df = left_join(df,cgm,by = c("Study.ID", "Event.Name"))
 ```
@@
 
-```{r results='asis'}
-t1 = tableby(~Age.at.visit.+Sex..Male.1.+BMIz+FEV1..+FVC....+trikafta_duration+
-               Genotype+Pancreatic.Status+Kalydeco+Orkambi+Symdeko,
-             data = df[df$Event.Name == "Visit 1",])
+```{r}
+# Labels for all tables - not the prettiest code
 newlabels = list(Age.at.visit. = "Age (years)",Sex..Male.1. = "Sex",
@@
                  Pancreatic.Status = "Pancreatic Status")
+newlabels["A1C.result"] = "HbA1c (%)"
+newlabels["Weight..kg."] = "Weight (kg)"
+newlabels["weightz"] = "Weight Z"
+newlabels["Glucose.0.minute"] = "Fasting glucose (mg/dL)"
+newlabels["Glucose.60.min"] = "1hr glucose (mg/dL)"
+newlabels["Glucose.120.min"] = "2hr glucose (mg/dL)"
+newlabels["c.peptide.0.minute"] = "Fasting C-peptide (pmol/L)"
+newlabels["iAUC_ratio"] = "C pep iAUC:Glucose iAUC"
+newlabels["homa2B"] = "HOMA2 %B"
+newlabels["homa2S"] = "HOMA2 %S"
+newlabels["homa2IR"] = "HOMA2 IR"
+newlabels["mean_sensor"] = "Mean SG (mg/dL)"
+newlabels["max_sensor"] = "Max. SG (mg/dL)"
+newlabels["min_sensor"] = "Min SG (mg/dL)"
+newlabels["percent_time_70_180"] = "% TIR 70-180 mg/dL"
+newlabels["percent_time_over_140"] = "% Time >140 mg/dL"
+newlabels["percent_time_over_180"] = "% Time >180 mg/dL"
+newlabels["percent_time_over_250"] = "% Time >250 mg/dL"
+newlabels["percent_time_under_70"] = "% Time <70 mg/dL"
+newlabels["percent_time_under_54"] = "% Time <54 mg/dL"
+newlabels["standard_deviation"] = "SD"
+newlabels["cv"] = "CV"
+newlabels["r_mage"] = "MAGE"
+newlabels["trikafta_duration"] = "Modulator Duration"
+```
+
+## Full cohort
+
+```{r results='asis'}
+t1 = tableby(~Age.at.visit.+Weight..kg.+weightz+trikafta_duration+
+               Sex..Male.1.+BMIz+FEV1..+FVC....+trikafta_duration+
+               Genotype+Pancreatic.Status+Kalydeco+Orkambi+Symdeko,
+             data = df[df$Event.Name == "Visit 1",])
+summary(t1,digits = 2,labelTranslations = newlabels)
+```
+
+## Reduced Cohort
+
+```{r results='asis'}
+t1 = tableby(~Age.at.visit.+Weight..kg.+weightz+trikafta_duration+
+               Sex..Male.1.+BMIz+FEV1..+FVC....+trikafta_duration+
+               Genotype+Pancreatic.Status+Kalydeco+Orkambi+Symdeko,
+             data = df[df$Event.Name == "Visit 1" & 
+                         !is.na(df$date_cgm_placement) & 
+                         !is.na(df$A1C.result),])
 summary(t1,digits = 2,labelTranslations = newlabels)
@@
 
-All metrics calculated based on 2 hour OGTT (0, 10, 30, 60, 90, and 120 minutes).
+All metrics calculated based on 2 hour OGTT (0, 10, 30, 60, 90, and 120 minutes). Group was treated as an ordered variable with NGT < AGT < CFRD.
 
-```{r results='asis'}
+```{r}
 # Make dataset for HOMA2 calculations - glucose in mmol/L and cpep in nmol/L
@@
 })
+df$Group = factor(df$Group,levels = c("NGT","AGT","CFRD"),ordered = T)
 # convert Insulin to mcU/mL
-ins_cols = colnames(df)[head(grep("insulin",tolower(colnames(df))),-2)]
+ins_cols = colnames(df)[head(grep("insulin",tolower(colnames(df))),-1)]
 df[,ins_cols] = lapply(df[,ins_cols], function(c){
@@
 })
+df$`Insulin AUC` = apply(df %>% select(Insulin.0.minute:Insulin.120.min),1,function(r){
+  trapz(c(0,10,30,60,90,120),as.numeric(r))
+})
 # Incremental AUCs
@@
 })
+df$`Insulin iAUC` = apply(df %>% select(Insulin.0.minute:Insulin.120.min),1,function(r){
+  r = r - r[1]
+  trapz(c(0,10,30,60,90,120),as.numeric(r))
+})
 # Other indices
@@
 glu_cols = colnames(df)[head(grep("glucose\\.\\d{1,}",tolower(colnames(df))),-2)]
+glu_cols = glu_cols[-2]
+ins_cols = colnames(df)[head(grep("insulin\\.\\d{1,}",tolower(colnames(df))),-2)]
+ins_cols = ins_cols[-1]
 df$mean_glucose = rowMeans(df[,glu_cols])
@@
                                       df$mean_glucose*df$mean_cpep))
+df$Matsuda = apply(df,1,function(r){
+  glu = round(as.numeric(r[glu_cols]))
+  ins = round(as.numeric(r[ins_cols]),1)
+  matsuda = 10000/sqrt(glu[1]*ins[1]*(glu[1]+glu[2]*2+glu[3]*2+glu[4]*2+glu[5])/
+                         8*(ins[1]+ins[2]*2+ins[3]*2+ins[4]*2+ins[5])/8)
+})
 # C Pep ODI
-df$`C Peptide ODI` = df$`C Peptide Index`/df$`Adapted Matsuda`
+df$`C Peptide ODI` = df$`C Peptide Index`*df$`Adapted Matsuda`
+# Matsuda ODI
+df$`ODI (Cpeptide index x Matsuda)` = df$`C Peptide Index`*df$Matsuda
 # iAUC ratio
 df$iAUC_ratio = df$`C Peptide iAUC`/df$`Glucose iAUC`
-# Add labels
-newlabels["A1C.result"] = "HbA1c (%)"
-newlabels["Glucose.0.minute"] = "Fasting glucose (mg/dL)"
-newlabels["Glucose.60.min"] = "1hr glucose (mg/dL)"
-newlabels["Glucose.120.min"] = "2hr glucose (mg/dL)"
-newlabels["c.peptide.0.minute"] = "Fasting C-peptide (pmol/L)"
-newlabels["iAUC_ratio"] = "C pep iAUC:Glucose iAUC"
-newlabels["homa2B"] = "HOMA2 %B"
-newlabels["homa2S"] = "HOMA2 %S"
-newlabels["homa2IR"] = "HOMA2 IR"
-p <- paired(Event.Name ~ A1C.result+BMIz+FEV1..+FVC....+Group+Glucose.0.minute+
-              Glucose.60.min+Glucose.120.min+c.peptide.0.minute+
-              `Glucose AUC`+`C Peptide AUC`+`C Peptide Index`+`Adapted Matsuda`+
-              `C Peptide ODI`+iAUC_ratio+homa2B+homa2S+homa2IR, 
-            data = df, id = Study.ID)
+```
+
+## Full cohort
+
+```{r results='asis'}
+t2_vars = c("A1C.result","Weight..kg.","BMIz","FEV1..","FVC....","trikafta_duration",
+            "Group","Glucose.0.minute",
+            "Glucose.60.min","Glucose.120.min","c.peptide.0.minute",
+            "`Glucose iAUC`","`C Peptide iAUC`","`Insulin iAUC`",
+            "`C Peptide Index`","Matsuda",
+            "`Adapted Matsuda`","`C Peptide ODI`","`ODI (Cpeptide index x Matsuda)`",
+            "iAUC_ratio","homa2B","homa2S","homa2IR",
+            "mean_sensor","max_sensor","min_sensor","percent_time_70_180",
+            "percent_time_over_140","percent_time_over_180","percent_time_over_250",
+            "percent_time_under_70","percent_time_under_54","standard_deviation","cv","r_mage")
+f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
+p <- paired(f,data = df, id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F)
+summary(p,pfootnote = T,labelTranslations = newlabels)
+```
+
+## Reduced Cohort
+
+```{r results='asis'}
+f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
+p = paired(f,data = df[!is.na(df$date_cgm_placement) & !is.na(df$A1C.result),], 
+           id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F)
 summary(p,pfootnote = T,labelTranslations = newlabels)
@@
 ```
+
+# Notes
+
+1.
+The A1C numbers look correct to me.
+The p value in Table 2 is calculated for only those people with a measure at V1 and V2, so in this case it was calculated based on n = 15.
+Similarly, for "HbA1c (%) for those with CGM," there are 12 values missing out of 20, so the p value is for 8 participants.
+
+2. The citation for the trapezoidal method R package is:
+
+    Borchers, Hans W. Pracma: Practical Numerical Math Functions, 2022. https://CRAN.R-project.org/package=pracma.
+
+3. Citation for R (version `r paste0(R.version$major,".","1.3")`):
+
+    R Core Team. R: A Language and Environment for Statistical Computing. Vienna, Austria: R Foundation for Statistical Computing, 2022. https://www.R-project.org/.

### 2022-03-30 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
added DXA variables for Christine

@@
 ```{r cgm data,include=FALSE}
-# Clean and calculate
-cleandata("./Christine Chan/Pre-post Trikafta Prospective/Data_Raw/CGM",
-          "./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/CGM",
-          removegaps = F)
-cgmvariables("./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/CGM",
-             "./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/",
-             outputname = "cgm_vars",id_filename = T)
 # Merge
@@
 df = left_join(df,cgm,by = c("Study.ID", "Event.Name"))
+# Update DXA variables
+dxa_vars = c("Tissue..g.","Total.Fat.Mass....","Fat.Body.Mass..g.","Fat.Free.Mass..g.",
+            "Lean.Body.Mass..g.","Total.Mass..kg.","Fat.Free.Mass..",
+            "Fat.Mass.Index..FMI...","Lean.body.mass.fat.mass..LBM.FM.",
+            "Fat.Free.Mass.Index...FFMI.","Lean.Body.Mass.Index..LBMI.",
+            "BMD.Total.z.score","BMD.Subtotal.z.score",
+            "DXA.FMI.calculated.z.score","DXA.LBMI.calculated.z.score")
+dxa = read.csv("./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/dxa_data.csv")
+dxa = dxa %>% select(Study.ID,all_of(dxa_vars))
+df = left_join(df,dxa)
+keep = dxa$Study.ID
 ```
@@
 newlabels["trikafta_duration"] = "Modulator Duration"
+newlabels[dxa_vars] = 
+  c("Tissue (g)","Total Fat Mass (%)","Fat Body Mass (g)","Fat Free Mass (g)",
+    "Lean Body Mass (g)","Total Mass (kg)","Fat Free Mass %","Fat Mass Index (FMI)",
+    "Lean body mass/fat mass (LBM/FM)","Fat Free Mass Index (FFMI)","Lean Body Mass Index (LBMI)",
+    "BMD Total z-score","BMD Subtotal z-score","DXA FMI calculated z-score",
+    "DXA LBMI calculated z-score")
 ```
@@
                Genotype+Pancreatic.Status+Kalydeco+Orkambi+Symdeko,
+             control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),
              data = df[df$Event.Name == "Visit 1",])
@@
                Genotype+Pancreatic.Status+Kalydeco+Orkambi+Symdeko,
-             data = df[df$Event.Name == "Visit 1" & 
-                         !is.na(df$date_cgm_placement) & 
-                         !is.na(df$A1C.result),])
+             control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),
+             data = df[df$Event.Name == "Visit 1" & df$Study.ID %in% keep,])
 summary(t1,digits = 2,labelTranslations = newlabels)
@@
             "iAUC_ratio","homa2B","homa2S","homa2IR",
+            "Tissue..g.","Total.Fat.Mass....","Fat.Body.Mass..g.","Fat.Free.Mass..g.",
+            "Lean.Body.Mass..g.","Total.Mass..kg.","Fat.Free.Mass..",
+            "Fat.Mass.Index..FMI...","Lean.body.mass.fat.mass..LBM.FM.",
+            "Fat.Free.Mass.Index...FFMI.","Lean.Body.Mass.Index..LBMI.",
+            "BMD.Total.z.score","BMD.Subtotal.z.score",
+            "DXA.FMI.calculated.z.score","DXA.LBMI.calculated.z.score",
             "mean_sensor","max_sensor","min_sensor","percent_time_70_180",
@@
 f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
-p <- paired(f,data = df, id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F)
+p <- paired(f,data = df, id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F,
+            control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),)
 summary(p,pfootnote = T,labelTranslations = newlabels)
@@
 f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
-p = paired(f,data = df[!is.na(df$date_cgm_placement) & !is.na(df$A1C.result),], 
+p = paired(f,data = df[!is.na(df$date_cgm_placement) & df$Study.ID %in% keep,], 
+           control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),
            id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F)
@@
 ```
+
+# Table 3: Adults vs. Children
+
+```{r}
+df$`Age Group` = cut(df$Age.at.visit.,c(-Inf,18,Inf),labels = c("Child","Adult"))
+```
+
+The following table includes `r as.numeric(table(df[df$Event.Name == "Visit 1","Age Group"])["Child"])` children and `r as.numeric(table(df[df$Event.Name == "Visit 1","Age Group"])["Adult"])` adults.
+
+```{r results='asis'}
+f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
+p <- paired(f,data = df, id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F,
+            control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),
+            strata = `Age Group`)
+summary(p,pfootnote = T,labelTranslations = newlabels)
+```
+
+# Table 4: Prior Modulator Use
+
+Unable to perform statistical tests due to too many missing values.
 

### 2022-04-04 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
less confusing Ns for chan pre-/post-Trikafta

@@
 df = left_join(df,cgm,by = c("Study.ID", "Event.Name"))
+keep_a1c = df %>% group_by(Study.ID) %>% 
+  summarise(n = sum(!is.na(A1C.result))) %>%
+  filter(n == 2) %>% pull(Study.ID)
 # Update DXA variables
@@
 df = left_join(df,dxa)
-keep = dxa$Study.ID
+keep_dxa = dxa$Study.ID
 ```
@@
              control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),
-             data = df[df$Event.Name == "Visit 1",])
+             data = df[df$Event.Name == "Visit 1" & df$Study.ID %in% keep_a1c,])
 summary(t1,digits = 2,labelTranslations = newlabels)
@@
              control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),
-             data = df[df$Event.Name == "Visit 1" & df$Study.ID %in% keep,])
+             data = df[df$Event.Name == "Visit 1" & df$Study.ID %in% keep_dxa,])
 summary(t1,digits = 2,labelTranslations = newlabels)
@@
 f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
-p <- paired(f,data = df, id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F,
+p <- paired(f,data = df[df$Study.ID %in% keep_a1c,], id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F,
             control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),)
@@
 f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
-p = paired(f,data = df[!is.na(df$date_cgm_placement) & df$Study.ID %in% keep,], 
+p = paired(f,data = df[df$Study.ID %in% keep_dxa,], 
            control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),
@@
 f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
-p <- paired(f,data = df, id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F,
+p <- paired(f,data = df[df$Study.ID %in% keep_a1c,], id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F,
             control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),

### 2022-04-05 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
added weight z score

@@
 ```{r results='asis'}
-t2_vars = c("A1C.result","Weight..kg.","BMIz","FEV1..","FVC....","trikafta_duration",
-            "Group","Glucose.0.minute",
+t2_vars = c("A1C.result","Weight..kg.","weightz","BMIz","FEV1..","FVC....",
+            "trikafta_duration","Group","Glucose.0.minute",
             "Glucose.60.min","Glucose.120.min","c.peptide.0.minute",

### 2022-04-06 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
cftr nutrion in R for MCP package

---
title: "Serum Biomarkers of Nutrition Pre and Post CFTR Modulator Use in Children with Cystic Fibrosis"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(readxl)
library(performance)
library(mcp)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
}
knitr::opts_knit$set(root.dir = home_dir)
rm(home_dir)
source("~/GitHub/CF/get_genotype_class_severity.R")
```

```{r data cleaning}
# Import
df = read.csv("~/Dropbox/Work/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_dataset.csv")
# Only those on effective modulators
on_effective = unique(df$Patient.ID[which(df[,'Modulator'] %in% c('Kalydeco','Trikafta'))])
df = df %>% filter(Patient.ID %in% on_effective)
# Non-highy efective to "None
df$Modulator[!df$Modulator %in% c('Kalydeco','Trikafta')] = "None"
# Before or after
df$Before_After = df$Age.at.Test..Years. < df$Age.at.Start..years.
df$Before_After = factor(df$Before_After,levels = c(FALSE,TRUE),labels = c("After","Before"))
# Count number of measures 
num_values = df %>% group_by(Patient.ID,Before_After) %>% 
  summarise(n = sum(!is.na(X25OH.Vitamin.D)),.groups = "drop") %>%
  pivot_wider(id_cols = Patient.ID,names_from = Before_After,values_from = n)
num_values$`NA` = NULL
colnames(num_values)[2:3] = c("vit_measures_after","vit_measures_before")
df = left_join(df,num_values,by = "Patient.ID")
# At least two annual nutritional measurements prior to first modulator start date
two_measures = num_values$Patient.ID[which(num_values$`vit_measures_before` >= 2)]
# At least one measurement of nutritional markers and one weight/height measurement >= three months post-modulator
time_after = df %>% 
  filter(Age.at.Test..Years. >= (Age.at.Start..years. + 0.25),!is.na(X25OH.Vitamin.D)) %>%
  pull(Patient.ID) %>% unique(.)
# Filter out those without above criteria, get first modulator
df = df %>% filter(Patient.ID %in% intersect(two_measures,time_after)) %>%
  group_by(Patient.ID) %>% 
  mutate(first_mod = head(Modulator[Modulator != "None"],1)) %>% ungroup()
# Genotype group
geno = read_excel("/Users/timvigers/Dropbox/Work/CF/CF Info/Sweat Tests from Internal Database_genotype class.xlsx")
geno = geno[,c('PatientID','genoRisk')]
df = left_join(df,geno,by = c('Patient.ID' = 'PatientID'))
# Time prior to first modulator
years = df %>% mutate(years_to_start = Age.at.Test..Years. - Age.at.Start..years.) %>%
  group_by(Patient.ID,Before_After) %>% 
  mutate(years_to_start_max = max(abs(years_to_start))) %>% ungroup() %>%
  select(Patient.ID,Before_After,years_to_start_max) %>% distinct(.) %>%
  pivot_wider(id_cols = Patient.ID,names_from = Before_After,values_from = years_to_start_max)
years$`NA` = NULL
colnames(years)[2:3] = c('max_years_to_start','max_years_from_start')
df = left_join(df,years,by = "Patient.ID")
```

# Table 1: Participant Characteristics

```{r results='asis'}
df$Sex[df$Sex == ""] = NA
demo = df %>% group_by(Patient.ID) %>% fill(Sex,.direction = "downup") %>% 
  filter(row_number() == 1)
t1 = tableby(first_mod ~ Age.at.Start..years. + Sex + genoRisk + 
               vit_measures_after + vit_measures_before,
             data = demo)
summary(t1)
```

### 2022-04-06 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
table 1 finished

@@
 knitr::opts_chunk$set(echo = FALSE)
-if(Sys.info()["sysname"] == "Windows"){
-  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
-} else if (Sys.info()["sysname"] == "Linux"){
-  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
-} else if (Sys.info()["sysname"] == "Darwin"){
-  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
-}
-knitr::opts_knit$set(root.dir = home_dir)
-rm(home_dir)
 source("~/GitHub/CF/get_genotype_class_severity.R")
@@
 
-# Table 1: Participant Characteristics
+# Participant Characteristics
+
+For each person, we calculated the maximum amount of time from the earliest nutrition measure before modulator start to start date, and the amount of time from modulator start date to most recent nutrition measure.
+The ranges of these values are presented below ("Years to start" and "Years From start," respectively).
 
@@
 t1 = tableby(first_mod ~ Age.at.Start..years. + Sex + genoRisk + 
-               vit_measures_after + vit_measures_before,
-             data = demo)
-summary(t1)
+               vit_measures_after + vit_measures_before +
+               notest(max_years_to_start,"min","max") + 
+               notest(max_years_from_start,"min","max"),
+             data = demo,
+             control = tableby.control(numeric.stats=c("Nmiss", "median", "q1q3"),))
+summary(t1,test = F,
+        labelTranslations = 
+          list(Age.at.Start..years. = "Age at Start (years)",
+               genoRisk = "Genotype risk group",
+               vit_measures_after = "No. of vitamin measures post",
+               vit_measures_before = "No. of vitamin measures pre",
+               max_years_to_start = "Years to start",max_years_from_start = "Years From start"))
 ```
+
+# Vitamin measures before and after modulator initiation
+
+```{r results='asis'}
+
+```

### 2022-04-06 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
basic comparisons done

@@
 # Time prior to first modulator
-years = df %>% mutate(years_to_start = Age.at.Test..Years. - Age.at.Start..years.) %>%
-  group_by(Patient.ID,Before_After) %>% 
-  mutate(years_to_start_max = max(abs(years_to_start))) %>% ungroup() %>%
-  select(Patient.ID,Before_After,years_to_start_max) %>% distinct(.) %>%
+years = df %>% group_by(Patient.ID,Before_After) %>% 
+  mutate(years_to_start_max = max(abs(Age.at.Test..Years. - Age.at.Start..years.))) %>% 
+  select(Patient.ID,Before_After,years_to_start_max) %>% 
+  ungroup() %>% distinct(.) %>%
   pivot_wider(id_cols = Patient.ID,names_from = Before_After,values_from = years_to_start_max)
@@
 df = left_join(df,years,by = "Patient.ID")
+df$years = df$Age.at.Test..Years. - df$Age.at.Start..years.
 ```
@@
 
+For each participant, we selected the pre and post visit closest to modulator start (and at least 3 months after start) with both a vitamin A and FEV1 % predicted value measure.
+Some visits were on different days and had different measures, so some missing values were filled in with the closest value (within 1 month).
+
 ```{r results='asis'}
+nutrition_vars = c("VitaminA_Retinol","X25OH.Vitamin.D","Vitamin.E.Alpha",
+                   "BMI","FEV1...pred")
+# Get the most recent 
+paired_df = df %>% 
+  select(Patient.ID,Before_After,years,first_mod,all_of(nutrition_vars)) %>% 
+  filter(years >= 0.25 | years < 0) %>%
+  mutate(years = round(years)) %>%
+  group_by(Patient.ID,years) %>%
+  fill(all_of(nutrition_vars),.direction = "downup") %>%
+  drop_na(any_of(c("VitaminA_Retinol","FEV1...pred"))) %>%
+  group_by(Patient.ID,Before_After) %>%
+  slice_min(abs(years)) %>% distinct() %>% filter(row_number()==1) 
+paired_df$Before_After = relevel(paired_df$Before_After,ref = "Before")
 
+f = as.formula(paste0("Before_After ~ ",paste0(nutrition_vars,collapse = "+")))
+p = paired(f,data = paired_df,id = Patient.ID,strata = first_mod)
+summary(p,
+        labelTranslations = list(VitaminA_Retinol = "Vitamin A Retinol",
+                                 X25OH.Vitamin.D = "25OH-Vitamin D",
+                                 Vitamin.E.Alpha = "FEV1 % Predicted",
+                                 first_mod = "First Modulator"))
 ```
+
+# Piecewise longitudinal models
+
+```{r}
+# Set up data again
+df = read.csv("~/Dropbox/Work/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_dataset.csv")
+df$years = df$Age.at.Test..Years. - df$Age.at.Start..years.
+t = df %>% group_by(Patient.ID) %>% summarise(n = sum(!is.na(VitaminA_Retinol)))
+```
+
+```{r}
+outcome = "VitaminA_Retinol"
+piecewise_mod = function(outcome,data){
+  data = data %>% select(Patient.ID,years,all_of(outcome)) %>% drop_na()
+  plot = ggplot(data,aes_string(x = "years",y = outcome,group = "Patient.ID")) + 
+    geom_point() + geom_line()
+}
+```

### 2022-04-07 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
cf nutrition mcp models

@@
 library(mcp)
+library(lmerTest)
+library(plotly)
 library(knitr)
@@
 # Import
-df = read.csv("~/Dropbox/Work/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_dataset.csv")
+df = read.csv("~/Dropbox/Work/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_dataset.csv",na.strings = "")
 # Only those on effective modulators
@@
   filter(years >= 0.25 | years < 0) %>%
-  mutate(years = round(years)) %>%
+  mutate(years = round(years,1)) %>%
   group_by(Patient.ID,years) %>%
   fill(all_of(nutrition_vars),.direction = "downup") %>%
-  drop_na(any_of(c("VitaminA_Retinol","FEV1...pred"))) %>%
+  drop_na(any_of(c("X25OH.Vitamin.D"))) %>%
   group_by(Patient.ID,Before_After) %>%
-  slice_min(abs(years)) %>% distinct() %>% filter(row_number()==1) 
+  slice_min(abs(years)) %>% distinct() %>% 
+  arrange(Patient.ID,abs(years)) %>%filter(row_number()==1) 
 paired_df$Before_After = relevel(paired_df$Before_After,ref = "Before")
@@
 f = as.formula(paste0("Before_After ~ ",paste0(nutrition_vars,collapse = "+")))
-p = paired(f,data = paired_df,id = Patient.ID,strata = first_mod)
+p = paired(f,data = paired_df,id = Patient.ID)
 summary(p,
@@
 df = read.csv("~/Dropbox/Work/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_dataset.csv")
-df$years = df$Age.at.Test..Years. - df$Age.at.Start..years.
-t = df %>% group_by(Patient.ID) %>% summarise(n = sum(!is.na(VitaminA_Retinol)))
+df$Age.at.Start..years. = NULL
+# New age at start
+mods = read_excel("~/Dropbox/Work/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report CFTR Modulators_throughDecember2021.xlsx")
+mods = mods[mods$Modulator %in% c('Kalydeco','Trikafta'),]
+colnames(mods) = make.names(colnames(mods))
+mods = mods %>% select(Patient.ID,Modulator,Age.at.Start..years.) %>% 
+  pivot_wider(id_cols = Patient.ID,names_from = Modulator,
+              values_from = Age.at.Start..years.)
+colnames(mods)[2:ncol(mods)] = paste0("age_at_start_",colnames(mods)[2:ncol(mods)])
+# Age at first and second modulator
+mods$age_first = pmin(mods$age_at_start_Trikafta,mods$age_at_start_Kalydeco,na.rm = T)
+mods$age_second = pmax(mods$age_at_start_Trikafta,mods$age_at_start_Kalydeco)
+# Merge back in
+df = left_join(df,mods,by = "Patient.ID")
+df = df %>% group_by(Patient.ID) %>% 
+  fill(age_at_start_Trikafta,age_at_start_Kalydeco,.direction = "downup")
+# Time relative to first mod
+df$time = df$Age.at.Test..Years. - df$age_first
+# Calculate spline things
+df$c1 = 
+  (df$time-0)*(df$time>=0)
+df$c2 = 
+  (df$Age.at.Test..Years.-df$age_second)*(df$Age.at.Test..Years.>=df$age_second)
 ```
@@
 ```{r}
-outcome = "VitaminA_Retinol"
 piecewise_mod = function(outcome,data){
-  data = data %>% select(Patient.ID,years,all_of(outcome)) %>% drop_na()
-  plot = ggplot(data,aes_string(x = "years",y = outcome,group = "Patient.ID")) + 
-    geom_point() + geom_line()
+  data = data %>% select(Patient.ID,all_of(outcome),time,c1,c2) %>% drop_na()
+  # Base plot
+  plot = ggplot(data,aes_string(x = "time",y = outcome)) + 
+    geom_point(alpha = 0.2) + geom_line(alpha = 0.2) +
+    theme_bw() + theme(legend.position = "none") + xlab("Age at Test")
+  # Model
+  f = as.formula(paste0(outcome,"~ time + c1 + c2 + (1|Patient.ID)"))
+  m = lmer(f,data = data)
+  # Predicted values for plotting
+  pred = predict(m)
+
+  plot + geom_line(data=data, 
+            aes(y = pred, x=Age.at.Test..Years.), size = 1, col="blue") 
+  ggplotly(plot)
 }
 ```
+
+## Vitamin A
+
+```{r}
+mod_df = df %>% select(Patient.ID,VitaminA_Retinol,Age.at.Test..Years.,
+                       age_first,age_second) %>% drop_na()
+mod_df$c1 = 
+  (mod_df$Age.at.Test..Years.-mod_df$age_first)*(mod_df$Age.at.Test..Years.>=mod_df$age_first)
+mod_df$c2 = 
+  (mod_df$Age.at.Test..Years.-mod_df$age_second)*(mod_df$Age.at.Test..Years.>=mod_df$age_second)
+
+m = lmer(VitaminA_Retinol ~ c1+c2+(1|Patient.ID),data = mod_df)
+```
+
+```{r cache=TRUE}
+mod_df = df %>% select(Patient.ID,time,VitaminA_Retinol) %>% drop_na()
+model = list(
+  VitaminA_Retinol ~ 1,  # Plateau in the first segment (int_1)
+  1 + (1|Patient.ID) ~ 0 + time,    # Joined slope (time_2) at cp_1
+  1 + (1|Patient.ID) ~ 0 + time     # Joined slope (int_3, time_3) at cp_2
+)
+fit = mcp(model, data = mod_df)
+plot(fit) + theme_bw()
+```
+
+```{r eval=FALSE}
+library(MultiKink)
+data("triceps")
+pred7 <- lm(triceps~ age + I((age-5)*(age>=5)) +
+                                   I((age-10)*(age >= 10)),
+                    data = triceps)
+triceps$c1 = (triceps$age-5)*(triceps$age>=5)
+triceps$c2 = (triceps$age-10)*(triceps$age>=10)
+
+tri.age.plot <- ggplot(triceps, aes(x=age, y=triceps)) +
+                 geom_point(alpha=0.55, color="black") + 
+                 theme_minimal()
+tri.age.plot +
+  geom_line(data=triceps, 
+            aes(y = pred7, x=age), size = 1, col="blue") 
+```

### 2022-04-07 Christine Chan/CFRD Screening Survey/cfrd_screening_analysis.Rmd
CFRD screening proportions

@@
   column_to_rownames(.,var = "name")
+
+```
+
+### Percentages by responder type
+
+```{r}
+# Table
+kable(proportions(data.matrix(t),margin = 2)*100,digits = 2)
+```
+
+### Figure
+
+```{r}
 # Plot

### 2022-04-08 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
fixed paths

@@
 knitr::opts_chunk$set(echo = FALSE)
+if(Sys.info()["sysname"] == "Windows"){
+  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
+} else if (Sys.info()["sysname"] == "Linux"){
+  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
+} else if (Sys.info()["sysname"] == "Darwin"){
+  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
+}
+knitr::opts_knit$set(root.dir = home_dir)
+rm(home_dir)
 source("~/GitHub/CF/get_genotype_class_severity.R")
@@
 # Import
-df = read.csv("~/Dropbox/Work/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_dataset.csv",na.strings = "")
+df = read.csv("./Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_dataset.csv",na.strings = "")
 # Only those on effective modulators
@@
 # Genotype group
-geno = read_excel("/Users/timvigers/Dropbox/Work/CF/CF Info/Sweat Tests from Internal Database_genotype class.xlsx")
+geno = read_excel("./CF Info/Sweat Tests from Internal Database_genotype class.xlsx")
 geno = geno[,c('PatientID','genoRisk')]

### 2022-04-12 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
piecewise models by number of modulators

--- 2022-04-06 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
@@
 library(mcp)
+library(lmerTest)
+library(broom.mixed)
+library(plotly)
 library(knitr)
 knitr::opts_chunk$set(echo = FALSE)
+if(Sys.info()["sysname"] == "Windows"){
+  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
+} else if (Sys.info()["sysname"] == "Linux"){
+  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
+} else if (Sys.info()["sysname"] == "Darwin"){
+  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
+}
+knitr::opts_knit$set(root.dir = home_dir)
+rm(home_dir)
 source("~/GitHub/CF/get_genotype_class_severity.R")
@@
 # Import
-df = read.csv("~/Dropbox/Work/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_dataset.csv")
+df = read.csv("./Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_dataset.csv",na.strings = "")
 # Only those on effective modulators
@@
   filter(years >= 0.25 | years < 0) %>%
-  mutate(years = round(years)) %>%
+  mutate(years = round(years,1)) %>%
   group_by(Patient.ID,years) %>%
   fill(all_of(nutrition_vars),.direction = "downup") %>%
-  drop_na(any_of(c("VitaminA_Retinol","FEV1...pred"))) %>%
+  drop_na(any_of(c("VitaminA_Retinol"))) %>%
   group_by(Patient.ID,Before_After) %>%
-  slice_min(abs(years)) %>% distinct() %>% filter(row_number()==1) 
+  slice_min(abs(years)) %>% distinct() %>% 
+  arrange(Patient.ID,abs(years)) %>%filter(row_number()==1) 
 paired_df$Before_After = relevel(paired_df$Before_After,ref = "Before")
@@
                                  X25OH.Vitamin.D = "25OH-Vitamin D",
-                                 Vitamin.E.Alpha = "FEV1 % Predicted",
+                                 Vitamin.E.Alpha = "Vitamin E Alpha",
                                  first_mod = "First Modulator"))
@@
 # Set up data again
-df = read.csv("~/Dropbox/Work/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_dataset.csv")
-df$years = df$Age.at.Test..Years. - df$Age.at.Start..years.
-t = df %>% group_by(Patient.ID) %>% summarise(n = sum(!is.na(VitaminA_Retinol)))
+df = read.csv("./Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_dataset.csv")
+df$Age.at.Start..years. = NULL
+# New age at start
+mods = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report CFTR Modulators_throughDecember2021.xlsx")
+mods = mods[mods$Modulator %in% c('Kalydeco','Trikafta'),]
+colnames(mods) = make.names(colnames(mods))
+mods = mods %>% select(Patient.ID,Modulator,Age.at.Start..years.) %>% 
+  pivot_wider(id_cols = Patient.ID,names_from = Modulator,
+              values_from = Age.at.Start..years.)
+colnames(mods)[2:ncol(mods)] = paste0("age_at_start_",colnames(mods)[2:ncol(mods)])
+# Age at first and second modulator
+mods$age_first = pmin(mods$age_at_start_Trikafta,mods$age_at_start_Kalydeco,na.rm = T)
+mods$age_second = pmax(mods$age_at_start_Trikafta,mods$age_at_start_Kalydeco)
+# Merge back in
+df = left_join(df,mods,by = "Patient.ID")
+df = df %>% group_by(Patient.ID) %>% 
+  fill(age_at_start_Trikafta,age_at_start_Kalydeco,.direction = "downup")
+# Time relative to first mod
+df$time = df$Age.at.Test..Years. - df$age_first
+# Calculate spline things
+df$c1 = 
+  (df$time-0)*(df$time>=0)
+df$c2 = 
+  (df$Age.at.Test..Years.-df$age_second)*(df$Age.at.Test..Years.>=df$age_second)
+# Limit to 3 years prior, select columns
+df = df %>% filter(!is.na(time) & time >= -3) %>% 
+  select(Patient.ID,Age.at.Test..Years.,age_first,age_second,time,c1,c2,
+         all_of(nutrition_vars))
+df$c2[is.na(df$c2)] = 0
 ```
@@
 ```{r}
-outcome = "VitaminA_Retinol"
-piecewise_mod = function(outcome,data){
-  data = data %>% select(Patient.ID,years,all_of(outcome)) %>% drop_na()
-  plot = ggplot(data,aes_string(x = "years",y = outcome,group = "Patient.ID")) + 
-    geom_point() + geom_line()
+piecewise_mod = function(outcome,data,diagnostics = F){
+  data$Patient.ID = as.factor(data$Patient.ID)
+  data = data %>% select(Patient.ID,time,c1,c2,age_first,age_second,
+                         all_of(outcome)) %>% drop_na(all_of(outcome))
+  # Separate data for single and double mods
+  single = data %>% filter(is.na(age_second))
+  double = data %>% filter(!is.na(age_second))
+  # Models
+  f_single = as.formula(paste0(outcome,"~ time + c1 + (1|Patient.ID)"))
+  m_single = lmer(f_single,data = single)
+  f_double = as.formula(paste0(outcome,"~ time + c1 +c2 + (1|Patient.ID)"))
+  m_double = lmer(f_double,data = double)
+
+  # Diagnostics if required
+  if(diagnostics){
+    check_single = check_model(m_single)
+    check_double = check_model(m_double)
+  } else {
+    check_single = NULL
+    check_double = NULL
+  }
+  # Results
+  res_single = tidy(m_single,"fixed")
+  res_double = tidy(m_double,"fixed")
+  # Plots
+  single_plot_df = m_single@frame
+  single_plot_df$pred = predict(m_single)
+  single_plot = ggplot(single_plot_df,aes_string(x = "time",y = outcome,color = "Patient.ID")) + 
+    geom_line(aes(y = pred)) + 
+    theme_bw() + theme(legend.position = "none") + xlab("Time (years)")
+
+  double_plot_df = m_double@frame
+  double_plot_df$pred = predict(m_double)
+  double_plot = ggplot(double_plot_df,aes_string(x = "time",y = outcome,color = "Patient.ID")) + 
+    geom_line(aes(y = pred)) + 
+    theme_bw() + theme(legend.position = "none") + xlab("Time (years)")
+  # Return
+  return(list("results_single" = res_single,"results_double" = res_double,
+              "diagnostics_single" = check_single,"diagnostics_double" = check_double,
+              "plot_single" = single_plot,"plot_double" = double_plot))
 }
 ```
+
+## Vitamin A
+
+```{r}
+vita = piecewise_mod(outcome = "VitaminA_Retinol",data = df,diagnostics = T)
+```
+
+### Single modulator
+
+```{r results='asis'}
+vita$plot_single
+kable(vita$results_single,digits = 3)
+```
+
+### Double modulator
+
+```{r results='asis'}
+vita$plot_double
+kable(vita$results_double,digits = 3)
+```
+
+## BMI
+
+```{r}
+bmi = piecewise_mod(outcome = "BMI",data = df,diagnostics = T)
+```
+
+### Single modulator
+
+```{r results='asis'}
+bmi$plot_single
+kable(bmi$results_single,digits = 3)
+```
+
+### Double modulator
+
+```{r results='asis'}
+bmi$plot_double
+kable(bmi$results_double,digits = 3)
+```
+
+# Bayesian?
+
+```{r cache=TRUE,echo=TRUE}
+mod_df = df %>% select(Patient.ID,time,VitaminA_Retinol) %>% drop_na()
+model = list(
+  VitaminA_Retinol ~ 1, # Plateau 
+  1 + (1|Patient.ID) ~ 0 + time, # Joined slope
+  1 + (1|Patient.ID) ~ 0 + time, # Joined slope
+  1 + (1|Patient.ID) ~ 0 + time # Joined slope
+)
+fit = mcp(model, data = mod_df)
+```
+
+```{r}
+plot(fit) + theme_bw()
+```

### 2022-04-15 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
better CFTR nutrition data cleaning

@@
 # Set up data again
-df = read.csv("./Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_dataset.csv")
-df$Age.at.Start..years. = NULL
+vit = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report Vitamins_throughDecember2021.xlsx")
+vit = vit %>% select(`Patient ID`,Date,VitaminA_Retinol:`Vitamin E Gamma`)
 # New age at start
@@
 mods = mods[mods$Modulator %in% c('Kalydeco','Trikafta'),]
-colnames(mods) = make.names(colnames(mods))
-mods = mods %>% select(Patient.ID,Modulator,Age.at.Start..years.) %>% 
-  pivot_wider(id_cols = Patient.ID,names_from = Modulator,
-              values_from = Age.at.Start..years.)
-colnames(mods)[2:ncol(mods)] = paste0("age_at_start_",colnames(mods)[2:ncol(mods)])
-# Age at first and second modulator
-mods$age_first = pmin(mods$age_at_start_Trikafta,mods$age_at_start_Kalydeco,na.rm = T)
-mods$age_second = pmax(mods$age_at_start_Trikafta,mods$age_at_start_Kalydeco)
-# Merge back in
-df = left_join(df,mods,by = "Patient.ID")
-df = df %>% group_by(Patient.ID) %>% 
-  fill(age_at_start_Trikafta,age_at_start_Kalydeco,.direction = "downup")
+mods = mods %>% select(`Patient ID`,Modulator,`Earliest Date in Registry`) %>% 
+  pivot_wider(id_cols = `Patient ID`,names_from = Modulator,
+              values_from = `Earliest Date in Registry`)
+colnames(mods)[2:ncol(mods)] = paste0(colnames(mods)[2:ncol(mods)]," Date")
+mods$`First Mod Date` = pmin(mods$`Trikafta Date`,mods$`Kalydeco Date`,na.rm = T)
+mods$`Second Mod Date` = pmax(mods$`Trikafta Date`,mods$`Kalydeco Date`)
+# Add pancreatic status
+pancreatic = read_excel("./CF Info/Sweat Tests from Internal Database_genotype class_withPancStatus.xlsx")
+pancreatic = pancreatic %>% rename(`Patient ID` = PatientID) %>% select(`Patient ID`,PancreaticStatus)
+# FEV1 and BMI
+fev = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report_PFTs_ThroughDecember2021.xlsx")
+fev = fev %>% select(PatientID,Date,BMI,`FEV1 % pred`) %>% rename(`Patient ID` = PatientID)
+# Merge
+df = full_join(vit,fev) %>% filter(`Patient ID` %in% mods$`Patient ID`,)
+df = left_join(df,mods)
+df = left_join(df,pancreatic)
 # Time relative to first mod
-df$time = df$Age.at.Test..Years. - df$age_first
+df$time = as.numeric(difftime(df$Date,df$`First Mod Date`,units = "days"))/365.25
 # Calculate spline things
@@
 df$c2 = 
-  (df$Age.at.Test..Years.-df$age_second)*(df$Age.at.Test..Years.>=df$age_second)
+  (as.numeric(difftime(df$Date,df$`Second Mod Date`,units = "days"))/365.25)*(df$Date >= df$`Second Mod Date`)
 # Limit to 3 years prior, select columns
-df = df %>% filter(!is.na(time) & time >= -3) %>% 
-  select(Patient.ID,Age.at.Test..Years.,age_first,age_second,time,c1,c2,
-         all_of(nutrition_vars))
-df$c2[is.na(df$c2)] = 0
+df = df %>% filter(!is.na(time) & time >= -3) %>% distinct()
+# Vitamin A to numeric
+df$VitaminA_Retinol = suppressWarnings(as.numeric(df$VitaminA_Retinol))
 ```
@@
 piecewise_mod = function(outcome,data,diagnostics = F){
-  data$Patient.ID = as.factor(data$Patient.ID)
-  data = data %>% select(Patient.ID,time,c1,c2,age_first,age_second,
+  data$`Patient ID` = as.factor(data$`Patient ID`)
+  data = data %>% select(`Patient ID`,Date,`First Mod Date`,`Second Mod Date`,PancreaticStatus,BMI,
                          all_of(outcome)) %>% drop_na(all_of(outcome))
   # Separate data for single and double mods
-  single = data %>% filter(is.na(age_second))
-  double = data %>% filter(!is.na(age_second))
+  single = data %>% filter(c2 <= 0 | is.na(c2))
+  double = data %>% filter(!is.na(c2))
   # Models
-  f_single = as.formula(paste0(outcome,"~ time + c1 + (1|Patient.ID)"))
+  f_single = as.formula(paste0(outcome,"~ BMI + PancreaticStatus + time + c1 + (1|`Patient ID`)"))
+  f_double = as.formula(paste0(outcome,"~ BMI + PancreaticStatus + time + c1 +c2 + (1|`Patient ID`)"))
+  if(outcome == "BMI"){
+    f_single = update(f_single,.~.-BMI)
+    f_double = update(f_double,.~.-BMI)
+  }
   m_single = lmer(f_single,data = single)
-  f_double = as.formula(paste0(outcome,"~ time + c1 +c2 + (1|Patient.ID)"))
   m_double = lmer(f_double,data = double)
-
   # Diagnostics if required
@@
   single_plot_df$pred = predict(m_single)
-  single_plot = ggplot(single_plot_df,aes_string(x = "time",y = outcome,color = "Patient.ID")) + 
+  single_plot = ggplot(single_plot_df,aes_string(x = "time",y = outcome,color = "`Patient ID`")) + 
     geom_line(aes(y = pred)) + 
     theme_bw() + theme(legend.position = "none") + xlab("Time (years)")
-
   double_plot_df = m_double@frame
   double_plot_df$pred = predict(m_double)
-  double_plot = ggplot(double_plot_df,aes_string(x = "time",y = outcome,color = "Patient.ID")) + 
+  double_plot = ggplot(double_plot_df,aes_string(x = "time",y = outcome,color = "`Patient ID`")) + 
     geom_line(aes(y = pred)) + 
@@
 ```
+
+In the following models, the "time" estimate represents the slope prior to first modulator start, "c1" is the change in slope after first modulator start, and "c2" is the change in slope from first modulator to second.
+All models were adjusted for BMI and pancreatic status (except the BMI model, which was only adjusted for pancreatic status).
 
@@
 
+## BMI
+
 ```{r cache=TRUE,echo=TRUE}
-mod_df = df %>% select(Patient.ID,time,VitaminA_Retinol) %>% drop_na()
+mod_df = df %>% filter(!is.na(age_second)) %>% select(Patient.ID,time,BMI) %>% drop_na()
 model = list(
-  VitaminA_Retinol ~ 1, # Plateau 
+  BMI ~ 1, # Plateau 
   1 + (1|Patient.ID) ~ 0 + time, # Joined slope
@@
 )
-fit = mcp(model, data = mod_df)
+fit = mcp(model, data = mod_df,sample = "both")
 ```
 
+### Posterior means
+
 ```{r}
-plot(fit) + theme_bw()
+summary(fit)
 ```
+
+### Posterior fit
+
+```{r}
+plot(fit,q_fit = TRUE) + theme_bw()
+```
+
+### Parameter plot
+
+```{r}
+plot_pars(fit)
+```
+
+### Hypothesis tests
+
+```{r}
+hypothesis(fit,c("cp_1 = 0","cp_2 = 0","cp_3 = 0"))
+```

### 2022-04-15 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
flow chart working but needs editing

@@
 library(knitr)
+library(DiagrammeR)
+library(lubridate)
 knitr::opts_chunk$set(echo = FALSE)
@@
 
-```{r data cleaning}
-# Import
-df = read.csv("./Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_dataset.csv",na.strings = "")
-# Only those on effective modulators
-on_effective = unique(df$Patient.ID[which(df[,'Modulator'] %in% c('Kalydeco','Trikafta'))])
-df = df %>% filter(Patient.ID %in% on_effective)
-# Non-highy efective to "None
-df$Modulator[!df$Modulator %in% c('Kalydeco','Trikafta')] = "None"
-# Before or after
-df$Before_After = df$Age.at.Test..Years. < df$Age.at.Start..years.
-df$Before_After = factor(df$Before_After,levels = c(FALSE,TRUE),labels = c("After","Before"))
-# Count number of measures 
-num_values = df %>% group_by(Patient.ID,Before_After) %>% 
-  summarise(n = sum(!is.na(X25OH.Vitamin.D)),.groups = "drop") %>%
-  pivot_wider(id_cols = Patient.ID,names_from = Before_After,values_from = n)
-num_values$`NA` = NULL
-colnames(num_values)[2:3] = c("vit_measures_after","vit_measures_before")
-df = left_join(df,num_values,by = "Patient.ID")
-# At least two annual nutritional measurements prior to first modulator start date
-two_measures = num_values$Patient.ID[which(num_values$`vit_measures_before` >= 2)]
-# At least one measurement of nutritional markers and one weight/height measurement >= three months post-modulator
-time_after = df %>% 
-  filter(Age.at.Test..Years. >= (Age.at.Start..years. + 0.25),!is.na(X25OH.Vitamin.D)) %>%
-  pull(Patient.ID) %>% unique(.)
-# Filter out those without above criteria, get first modulator
-df = df %>% filter(Patient.ID %in% intersect(two_measures,time_after)) %>%
-  group_by(Patient.ID) %>% 
-  mutate(first_mod = head(Modulator[Modulator != "None"],1)) %>% ungroup()
-# Genotype group
-geno = read_excel("./CF Info/Sweat Tests from Internal Database_genotype class.xlsx")
-geno = geno[,c('PatientID','genoRisk')]
-df = left_join(df,geno,by = c('Patient.ID' = 'PatientID'))
-# Time prior to first modulator
-years = df %>% group_by(Patient.ID,Before_After) %>% 
-  mutate(years_to_start_max = max(abs(Age.at.Test..Years. - Age.at.Start..years.))) %>% 
-  select(Patient.ID,Before_After,years_to_start_max) %>% 
-  ungroup() %>% distinct(.) %>%
-  pivot_wider(id_cols = Patient.ID,names_from = Before_After,values_from = years_to_start_max)
-years$`NA` = NULL
-colnames(years)[2:3] = c('max_years_to_start','max_years_from_start')
-df = left_join(df,years,by = "Patient.ID")
-df$years = df$Age.at.Test..Years. - df$Age.at.Start..years.
+# Data cleaning
+
+```{r}
+# Vitamins
+vit = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report Vitamins_throughDecember2021.xlsx")
+vit = vit %>% select(`Patient ID`,Date,VitaminA_Retinol:`Vitamin E Gamma`)
+# New age at start
+mods = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report CFTR Modulators_throughDecember2021.xlsx")
+mods = mods[mods$Modulator %in% c('Kalydeco','Trikafta'),]
+mods = mods %>% select(`Patient ID`,Modulator,`Earliest Date in Registry`) %>% 
+  pivot_wider(id_cols = `Patient ID`,names_from = Modulator,
+              values_from = `Earliest Date in Registry`)
+colnames(mods)[2:ncol(mods)] = paste0(colnames(mods)[2:ncol(mods)]," Date")
+mods$`First Mod Date` = pmin(mods$`Trikafta Date`,mods$`Kalydeco Date`,na.rm = T)
+mods$`Second Mod Date` = pmax(mods$`Trikafta Date`,mods$`Kalydeco Date`)
+# Add pancreatic status
+pancreatic = read_excel("./CF Info/Sweat Tests from Internal Database_genotype class_withPancStatus.xlsx")
+pancreatic = pancreatic %>% rename(`Patient ID` = PatientID) %>% select(`Patient ID`,PancreaticStatus)
+# FEV1 and BMI
+fev = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report_PFTs_ThroughDecember2021.xlsx")
+fev = fev %>% select(PatientID,Date,BMI,`FEV1 % pred`) %>% rename(`Patient ID` = PatientID)
+# Merge
+df = full_join(vit,fev,by = c("Patient ID", "Date")) %>% filter(`Patient ID` %in% mods$`Patient ID`)
+df = left_join(df,mods,by = "Patient ID")
+df = left_join(df,pancreatic,by = "Patient ID")
+```
+
+```{r}
+# Counts from Elin
+# patients with CF in our clinical database from 2008- December 2021: 740 total, 474 in our dataset
+ncdb = 740
+# of patients treated with ivacaftor from 2008-2021: 58 total, 47 in dataset
+# of patients treated with trikafta (ETI) from 2008-2021: 228 total, 150 in dataset 
+# of patients treated with ivacaftor followed by ETI – these patients may be counted in both groups: 22 total, 17 in dataset
+# Tim counts
+df$Group = apply(df,1,function(r){
+  if(!is.na(r["Kalydeco Date"]) & is.na(r["Trikafta Date"])){
+    g = "Kalydeco"
+  } else if (is.na(r["Kalydeco Date"]) & !is.na(r["Trikafta Date"])){
+    g = "Trikafta"
+  } else if (!is.na(r["Kalydeco Date"]) & !is.na(r["Trikafta Date"])){
+    g = "Both"
+  }
+  return(g)
+})
+# Two nutrition measures before first mod start
+two_measures = df %>% 
+  mutate(before_first = Date < `First Mod Date` & !is.na(VitaminA_Retinol)) %>% 
+  group_by(`Patient ID`,Group) %>% count(before_first) %>% 
+  filter(before_first == T,n >= 2)
+# At least one nutritional measure and one weight/height measure >= 3 months after first
+after_measure = df %>% 
+  mutate(after_first = (Date >= `First Mod Date` %m+% months(3)) & 
+           !is.na(BMI) & 
+           !is.na(VitaminA_Retinol)) %>%
+  group_by(`Patient ID`,Group) %>% count(after_first) %>%
+  filter(after_first == T,n >= 1)
+# Filter 
+df = df %>% filter(`Patient ID` %in% two_measures & `Patient ID` %in% after_measure)
+```
+
+```{r}
+# Flowchart
+grViz("
+digraph graph2 {
+
+graph [layout = dot]
+
+# node definitions with substituted label text
+node [shape = rectangle, width = 4, fillcolor = Biege]
+a [label = '@@1']
+b [label = '@@2']
+c [label = '@@3']
+d [label = '@@4']
+e [label = '@@5']
+f [label = '@@6']
+g [label = '@@7']
+h [label = '@@8']
+
+a -> b
+b -> c
+c -> d
+c -> e
+c -> f
+}
+
+[1]: paste0('Clinical Database (n = ', ncdb, ')')
+[2]: paste0('Treated with ivacaftor or ETI (or both) from 2008-2021 (n = ', length(unique(mods$`Patient ID`)),')')
+[3]: paste0('In our dataset (n = ',length(unique(df$`Patient ID`)),')')
+[4]: paste0('Treated with ivacaftor alone from 2008-2021 (n = ', n_kaly,')')
+[5]: paste0('Treated with ETI alone from 2008-2021 (n = ', n_tri,')')
+[6]: paste0('Treated with ivacaftor AND ETI from 2008-2021 (n = ',')')
+[7]: paste0('At least two Vitamin A measures (n = ',')')
+[8]: paste0('At least two Vitamin A measures (n = ',')')
+[9]: paste0('At least two Vitamin A measures (n = ',')')
+")
 ```
@@
 
-# Vitamin measures before and after modulator initiation
-
-For each participant, we selected the pre and post visit closest to modulator start (and at least 3 months after start) with both a vitamin A and FEV1 % predicted value measure.
-Some visits were on different days and had different measures, so some missing values were filled in with the closest value (within 1 month).
-
-```{r results='asis'}
-nutrition_vars = c("VitaminA_Retinol","X25OH.Vitamin.D","Vitamin.E.Alpha",
-                   "BMI","FEV1...pred")
-# Get the most recent 
-paired_df = df %>% 
-  select(Patient.ID,Before_After,years,first_mod,all_of(nutrition_vars)) %>% 
-  filter(years >= 0.25 | years < 0) %>%
-  mutate(years = round(years,1)) %>%
-  group_by(Patient.ID,years) %>%
-  fill(all_of(nutrition_vars),.direction = "downup") %>%
-  drop_na(any_of(c("VitaminA_Retinol"))) %>%
-  group_by(Patient.ID,Before_After) %>%
-  slice_min(abs(years)) %>% distinct() %>% 
-  arrange(Patient.ID,abs(years)) %>%filter(row_number()==1) 
-paired_df$Before_After = relevel(paired_df$Before_After,ref = "Before")
-
-f = as.formula(paste0("Before_After ~ ",paste0(nutrition_vars,collapse = "+")))
-p = paired(f,data = paired_df,id = Patient.ID)
-summary(p,
-        labelTranslations = list(VitaminA_Retinol = "Vitamin A Retinol",
-                                 X25OH.Vitamin.D = "25OH-Vitamin D",
-                                 Vitamin.E.Alpha = "Vitamin E Alpha",
-                                 first_mod = "First Modulator"))
-```
-
 # Piecewise longitudinal models
@@
 ```{r}
-# Set up data again
-vit = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report Vitamins_throughDecember2021.xlsx")
-vit = vit %>% select(`Patient ID`,Date,VitaminA_Retinol:`Vitamin E Gamma`)
-# New age at start
-mods = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report CFTR Modulators_throughDecember2021.xlsx")
-mods = mods[mods$Modulator %in% c('Kalydeco','Trikafta'),]
-mods = mods %>% select(`Patient ID`,Modulator,`Earliest Date in Registry`) %>% 
-  pivot_wider(id_cols = `Patient ID`,names_from = Modulator,
-              values_from = `Earliest Date in Registry`)
-colnames(mods)[2:ncol(mods)] = paste0(colnames(mods)[2:ncol(mods)]," Date")
-mods$`First Mod Date` = pmin(mods$`Trikafta Date`,mods$`Kalydeco Date`,na.rm = T)
-mods$`Second Mod Date` = pmax(mods$`Trikafta Date`,mods$`Kalydeco Date`)
-# Add pancreatic status
-pancreatic = read_excel("./CF Info/Sweat Tests from Internal Database_genotype class_withPancStatus.xlsx")
-pancreatic = pancreatic %>% rename(`Patient ID` = PatientID) %>% select(`Patient ID`,PancreaticStatus)
-# FEV1 and BMI
-fev = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report_PFTs_ThroughDecember2021.xlsx")
-fev = fev %>% select(PatientID,Date,BMI,`FEV1 % pred`) %>% rename(`Patient ID` = PatientID)
-# Merge
-df = full_join(vit,fev) %>% filter(`Patient ID` %in% mods$`Patient ID`,)
-df = left_join(df,mods)
-df = left_join(df,pancreatic)
 # Time relative to first mod
@@
 df = df %>% filter(!is.na(time) & time >= -3) %>% distinct()
-# Vitamin A to numeric
-df$VitaminA_Retinol = suppressWarnings(as.numeric(df$VitaminA_Retinol))
 ```

### 2022-04-15 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
first flow chart draft done

@@
   group_by(`Patient ID`,Group) %>% count(after_first) %>%
-  filter(after_first == T,n >= 1)
+  filter(after_first == T,n >= 1,`Patient ID` %in% two_measures$`Patient ID`)
 # Filter
-df = df %>% filter(`Patient ID` %in% two_measures & `Patient ID` %in% after_measure)
+df = df %>% filter(`Patient ID` %in% after_measure$`Patient ID`)
 ```
@@
 h [label = '@@8']
+i [label = '@@9']
+j [label = '@@10']
+k [label = '@@11']
+l [label = '@@12']
+m [label = '@@13']
 
@@
 c -> f
+d -> g
+e -> h
+f -> i
+g -> j
+h -> k
+i -> l
+j -> m
+k -> m
+l -> m
 }
@@
 [1]: paste0('Clinical Database (n = ', ncdb, ')')
-[2]: paste0('Treated with ivacaftor or ETI (or both) from 2008-2021 (n = ', length(unique(mods$`Patient ID`)),')')
-[3]: paste0('In our dataset (n = ',length(unique(df$`Patient ID`)),')')
-[4]: paste0('Treated with ivacaftor alone from 2008-2021 (n = ', n_kaly,')')
-[5]: paste0('Treated with ETI alone from 2008-2021 (n = ', n_tri,')')
-[6]: paste0('Treated with ivacaftor AND ETI from 2008-2021 (n = ',')')
-[7]: paste0('At least two Vitamin A measures (n = ',')')
-[8]: paste0('At least two Vitamin A measures (n = ',')')
-[9]: paste0('At least two Vitamin A measures (n = ',')')
+[2]: paste0('Treated With Ivacaftor or Trikafta From 2008-2021 (n = ',length(unique(mods$`Patient ID`)), ')')
+[3]: paste0('In Our Dataset (n = ',length(unique(df$`Patient ID`)), ')')
+[4]: paste0('Treated With Ivacaftor Alone From 2008-2021 (n = ',as.numeric(table(groups$Group)['Kalydeco']), ')')
+[5]: paste0('Treated With Trikafta Alone From 2008-2021 (n = ',as.numeric(table(groups$Group)['Trikafta']), ')')
+[6]: paste0('Treated With Ivacaftor followed by Trikafta From 2008-2021 (n = ',as.numeric(table(groups$Group)['Both']), ')')
+[7]: paste0('At Least Two Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Kalydeco']), ')')
+[8]: paste0('At Least Two Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Trikafta']), ')')
+[9]: paste0('At Least Two Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Both']), ')')
+[10]: paste0('At Least One Vitamin A and One BMI Measure >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Kalydeco']), ')')
+[11]: paste0('At Least One Vitamin A and One BMI Measure >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Trikafta']), ')')
+[12]: paste0('At Least One Vitamin A and One BMI Measure >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Both']), ')')
+[13]: paste0('Final (n = ',length(unique(after_measure$`Patient ID`)), ')')
 ")

### 2022-04-15 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
write files and fixed graph size

--- 2022-04-15 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
@@
 })
+groups = df %>% group_by(`Patient ID`) %>% filter(row_number()==1)
 # Two nutrition measures before first mod start
@@
   group_by(`Patient ID`,Group) %>% count(after_first) %>%
-  filter(after_first == T,n >= 1)
-# Filter 
-df = df %>% filter(`Patient ID` %in% two_measures & `Patient ID` %in% after_measure)
+  filter(after_first == T,n >= 1,`Patient ID` %in% two_measures$`Patient ID`)
+# Write merged df
+write.csv(df,file = "./Edith Zemanick/CF Nutrition/Data_Cleaned/merged_data.csv",
+          row.names = F,na="")
+write.csv(mods,file = "./Edith Zemanick/CF Nutrition/Data_Cleaned/modulators.csv",
+          row.names = F,na="")
 ```
@@
 # Flowchart
-grViz("
+g = grViz("
 digraph graph2 {
-
 graph [layout = dot]
-
 # node definitions with substituted label text
-node [shape = rectangle, width = 4, fillcolor = Biege]
+node [shape = rectangle]
 a [label = '@@1']
@@
 [1]: paste0('Clinical Database (n = ', ncdb, ')')
-[2]: paste0('Treated with ivacaftor or ETI (or both) from 2008-2021 (n = ', length(unique(mods$`Patient ID`)),')')
-[3]: paste0('In our dataset (n = ',length(unique(df$`Patient ID`)),')')
-[4]: paste0('Treated with ivacaftor alone from 2008-2021 (n = ', n_kaly,')')
-[5]: paste0('Treated with ETI alone from 2008-2021 (n = ', n_tri,')')
-[6]: paste0('Treated with ivacaftor AND ETI from 2008-2021 (n = ',')')
-[7]: paste0('At least two Vitamin A measures (n = ',')')
-[8]: paste0('At least two Vitamin A measures (n = ',')')
-[9]: paste0('At least two Vitamin A measures (n = ',')')
+[2]: paste0('Treated With Ivacaftor or Trikafta From 2008-2021 (n = ',length(unique(mods$`Patient ID`)), ')')
+[3]: paste0('In Our Dataset (n = ',length(unique(df$`Patient ID`)), ')')
+[4]: paste0('Treated With Ivacaftor Alone From 2008-2021 (n = ',as.numeric(table(groups$Group)['Kalydeco']), ')')
+[5]: paste0('Treated With Trikafta Alone From 2008-2021 (n = ',as.numeric(table(groups$Group)['Trikafta']), ')')
+[6]: paste0('Treated With Ivacaftor followed by Trikafta From 2008-2021 (n = ',as.numeric(table(groups$Group)['Both']), ')')
+[7]: paste0('At Least Two Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Kalydeco']), ')')
+[8]: paste0('At Least Two Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Trikafta']), ')')
+[9]: paste0('At Least Two Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Both']), ')')
+[10]: paste0('At Least One Vitamin A and One BMI Measure >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Kalydeco']), ')')
+[11]: paste0('At Least One Vitamin A and One BMI Measure >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Trikafta']), ')')
+[12]: paste0('At Least One Vitamin A and One BMI Measure >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Both']), ')')
+[13]: paste0('Final (n = ',length(unique(after_measure$`Patient ID`)), ')')
 ")
+g
 ```

### 2022-04-18 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
groups of visits within 3 months of each other

@@
 df = left_join(df,pancreatic,by = "Patient ID")
+# Match close dates 
+df_list = split.data.frame(df,df$`Patient ID`)
+t = df_list[[1]]
+lapply(df_list, function(d){
+  d = d %>% arrange(Date)
+  # Calculate difference between all dates (in seconds)
+  dis = as.matrix(dist(d$Date))
+  # Which within 3 months
+  pairs = as.data.frame(which(dis <= 24*60*60*90 & lower.tri(dis),arr.ind = T))
+  # Groups
+  group = NA
+  length(group) = nrow(pairs)
+  g = 1
+  for (i in 1:length(group)) {
+    if (i == 1){
+      group[i] = 1 
+      next
+    } else {
+      k = j
+      j = as.numeric(pairs[i,])
+      if(length(intersect(j,k)) == 0){
+        g = g + 1
+      }
+      group[i] = g
+    }
+  }
+  pairs$group = group
+  # Add to d
+  d$group = seq(from = max(pairs$group)+1,by = 1,length.out = nrow(d))
+  d$group[pairs$row] = pairs$group
+})
 ```

### 2022-04-18 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
back to simpler data analysis

--- 2022-04-15 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
@@
   mutate(after_first = (Date >= `First Mod Date` %m+% months(3)) & 
-           !is.na(BMI) & 
            !is.na(VitaminA_Retinol)) %>%
   group_by(`Patient ID`,Group) %>% count(after_first) %>%
-  filter(after_first == T,n >= 1,`Patient ID` %in% two_measures$`Patient ID`)
+  filter(after_first == T,n >= 1)
 # Write merged df
@@
 [9]: paste0('At Least Two Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Both']), ')')
-[10]: paste0('At Least One Vitamin A and One BMI Measure >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Kalydeco']), ')')
-[11]: paste0('At Least One Vitamin A and One BMI Measure >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Trikafta']), ')')
-[12]: paste0('At Least One Vitamin A and One BMI Measure >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Both']), ')')
+[10]: paste0('At Least One Vitamin A >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Kalydeco']), ')')
+[11]: paste0('At Least One Vitamin A >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Trikafta']), ')')
+[12]: paste0('At Least One Vitamin A >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Both']), ')')
 [13]: paste0('Final (n = ',length(unique(after_measure$`Patient ID`)), ')')
@@
 
-# Participant Characteristics
+<!-- # Participant Characteristics -->
 
-For each person, we calculated the maximum amount of time from the earliest nutrition measure before modulator start to start date, and the amount of time from modulator start date to most recent nutrition measure.
+<!-- For each person, we calculated the maximum amount of time from the earliest nutrition measure before modulator start to start date, and the amount of time from modulator start date to most recent nutrition measure.
 The ranges of these values are presented below ("Years to start" and "Years From start," respectively).
+-->
 
-```{r results='asis'}
-df$Sex[df$Sex == ""] = NA
-demo = df %>% group_by(Patient.ID) %>% fill(Sex,.direction = "downup") %>%
-  filter(row_number() == 1)
-t1 = tableby(first_mod ~ Age.at.Start..years. + Sex + genoRisk +
-               vit_measures_after + vit_measures_before +
-               notest(max_years_to_start,"min","max") +
-               notest(max_years_from_start,"min","max"),
-             data = demo,
-             control = tableby.control(numeric.stats=c("Nmiss", "median", "q1q3"),))
-summary(t1,test = F,
-        labelTranslations =
-          list(Age.at.Start..years. = "Age at Start (years)",
-               genoRisk = "Genotype risk group",
-               vit_measures_after = "No. of vitamin measures post",
-               vit_measures_before = "No. of vitamin measures pre",
-               max_years_to_start = "Years to start",max_years_from_start = "Years From start"))
-```
+<!-- ```{r results='asis'} -->
+<!-- df$Sex[df$Sex == ""] = NA -->
+<!-- demo = df %>% group_by(Patient.ID) %>% fill(Sex,.direction = "downup") %>% -->
+<!--   filter(row_number() == 1) -->
+<!-- t1 = tableby(first_mod ~ Age.at.Start..years. + Sex + genoRisk + -->
+<!--                vit_measures_after + vit_measures_before + -->
+<!--                notest(max_years_to_start,"min","max") + -->
+<!--                notest(max_years_from_start,"min","max"), -->
+<!--              data = demo, -->
+<!--              control = tableby.control(numeric.stats=c("Nmiss", "median", "q1q3"),)) -->
+<!-- summary(t1,test = F, -->
+<!--         labelTranslations = -->
+<!--           list(Age.at.Start..years. = "Age at Start (years)", -->
+<!--                genoRisk = "Genotype risk group", -->
+<!--                vit_measures_after = "No. of vitamin measures post", -->
+<!--                vit_measures_before = "No. of vitamin measures pre", -->
+<!--                max_years_to_start = "Years to start",max_years_from_start = "Years From start")) -->
+<!-- ``` -->
 
-# Piecewise longitudinal models
+<!-- # Piecewise longitudinal models -->
 
-```{r}
-# Time relative to first mod
-df$time = as.numeric(difftime(df$Date,df$`First Mod Date`,units = "days"))/365.25
-# Calculate spline things
-df$c1 =
-  (df$time-0)*(df$time>=0)
-df$c2 =
-  (as.numeric(difftime(df$Date,df$`Second Mod Date`,units = "days"))/365.25)*(df$Date >= df$`Second Mod Date`)
-# Limit to 3 years prior, select columns
-df = df %>% filter(!is.na(time) & time >= -3) %>% distinct()
-```
+<!-- ```{r} -->
+<!-- # Time relative to first mod -->
+<!-- df$time = as.numeric(difftime(df$Date,df$`First Mod Date`,units = "days"))/365.25 -->
+<!-- # Calculate spline things -->
+<!-- df$c1 = -->
+<!--   (df$time-0)*(df$time>=0) -->
+<!-- df$c2 = -->
+<!--   (as.numeric(difftime(df$Date,df$`Second Mod Date`,units = "days"))/365.25)*(df$Date >= df$`Second Mod Date`) -->
+<!-- # Limit to 3 years prior, select columns -->
+<!-- df = df %>% filter(!is.na(time) & time >= -3) %>% distinct() -->
+<!-- ``` -->
 
-```{r}
-piecewise_mod = function(outcome,data,diagnostics = F){
-  data$`Patient ID` = as.factor(data$`Patient ID`)
-  data = data %>% select(`Patient ID`,Date,`First Mod Date`,`Second Mod Date`,PancreaticStatus,BMI,
-                         all_of(outcome)) %>% drop_na(all_of(outcome))
-  # Separate data for single and double mods
-  single = data %>% filter(c2 <= 0 | is.na(c2))
-  double = data %>% filter(!is.na(c2))
-  # Models
-  f_single = as.formula(paste0(outcome,"~ BMI + PancreaticStatus + time + c1 + (1|`Patient ID`)"))
-  f_double = as.formula(paste0(outcome,"~ BMI + PancreaticStatus + time + c1 +c2 + (1|`Patient ID`)"))
-  if(outcome == "BMI"){
-    f_single = update(f_single,.~.-BMI)
-    f_double = update(f_double,.~.-BMI)
-  }
-  m_single = lmer(f_single,data = single)
-  m_double = lmer(f_double,data = double)
-  # Diagnostics if required
-  if(diagnostics){
-    check_single = check_model(m_single)
-    check_double = check_model(m_double)
-  } else {
-    check_single = NULL
-    check_double = NULL
-  }
-  # Results
-  res_single = tidy(m_single,"fixed")
-  res_double = tidy(m_double,"fixed")
-  # Plots
-  single_plot_df = m_single@frame
-  single_plot_df$pred = predict(m_single)
-  single_plot = ggplot(single_plot_df,aes_string(x = "time",y = outcome,color = "`Patient ID`")) +
-    geom_line(aes(y = pred)) +
-    theme_bw() + theme(legend.position = "none") + xlab("Time (years)")
-  double_plot_df = m_double@frame
-  double_plot_df$pred = predict(m_double)
-  double_plot = ggplot(double_plot_df,aes_string(x = "time",y = outcome,color = "`Patient ID`")) +
-    geom_line(aes(y = pred)) +
-    theme_bw() + theme(legend.position = "none") + xlab("Time (years)")
-  # Return
-  return(list("results_single" = res_single,"results_double" = res_double,
-              "diagnostics_single" = check_single,"diagnostics_double" = check_double,
-              "plot_single" = single_plot,"plot_double" = double_plot))
-}
-```
+<!-- ```{r} -->
+<!-- piecewise_mod = function(outcome,data,diagnostics = F){ -->
+<!--   data$`Patient ID` = as.factor(data$`Patient ID`) -->
+<!--   data = data %>% select(`Patient ID`,Date,`First Mod Date`,`Second Mod Date`,PancreaticStatus,BMI, -->
+<!--                          all_of(outcome)) %>% drop_na(all_of(outcome)) -->
+<!--   # Separate data for single and double mods -->
+<!--   single = data %>% filter(c2 <= 0 | is.na(c2)) -->
+<!--   double = data %>% filter(!is.na(c2)) -->
+<!--   # Models -->
+<!--   f_single = as.formula(paste0(outcome,"~ BMI + PancreaticStatus + time + c1 + (1|`Patient ID`)")) -->
+<!--   f_double = as.formula(paste0(outcome,"~ BMI + PancreaticStatus + time + c1 +c2 + (1|`Patient ID`)")) -->
+<!--   if(outcome == "BMI"){ -->
+<!--     f_single = update(f_single,.~.-BMI) -->
+<!--     f_double = update(f_double,.~.-BMI) -->
+<!--   } -->
+<!--   m_single = lmer(f_single,data = single) -->
+<!--   m_double = lmer(f_double,data = double) -->
+<!--   # Diagnostics if required -->
+<!--   if(diagnostics){ -->
+<!--     check_single = check_model(m_single) -->
+<!--     check_double = check_model(m_double) -->
+<!--   } else { -->
+<!--     check_single = NULL -->
+<!--     check_double = NULL -->
+<!--   } -->
+<!--   # Results -->
+<!--   res_single = tidy(m_single,"fixed") -->
+<!--   res_double = tidy(m_double,"fixed") -->
+<!--   # Plots -->
+<!--   single_plot_df = m_single@frame -->
+<!--   single_plot_df$pred = predict(m_single) -->
+<!--   single_plot = ggplot(single_plot_df,aes_string(x = "time",y = outcome,color = "`Patient ID`")) + -->
+<!--     geom_line(aes(y = pred)) + -->
+<!--     theme_bw() + theme(legend.position = "none") + xlab("Time (years)") -->
+<!--   double_plot_df = m_double@frame -->
+<!--   double_plot_df$pred = predict(m_double) -->
+<!--   double_plot = ggplot(double_plot_df,aes_string(x = "time",y = outcome,color = "`Patient ID`")) + -->
+<!--     geom_line(aes(y = pred)) + -->
+<!--     theme_bw() + theme(legend.position = "none") + xlab("Time (years)") -->
+<!--   # Return -->
+<!--   return(list("results_single" = res_single,"results_double" = res_double, -->
+<!--               "diagnostics_single" = check_single,"diagnostics_double" = check_double, -->
+<!--               "plot_single" = single_plot,"plot_double" = double_plot)) -->
+<!-- } -->
+<!-- ``` -->
 
-In the following models, the "time" estimate represents the slope prior to first modulator start, "c1" is the change in slope after first modulator start, and "c2" is the change in slope from first modulator to second.
+<!-- In the following models, the "time" estimate represents the slope prior to first modulator start, "c1" is the change in slope after first modulator start, and "c2" is the change in slope from first modulator to second.
 All models were adjusted for BMI and pancreatic status (except the BMI model, which was only adjusted for pancreatic status).
+-->
 
-## Vitamin A
+<!-- ## Vitamin A -->
 
-```{r}
-vita = piecewise_mod(outcome = "VitaminA_Retinol",data = df,diagnostics = T)
-```
+<!-- ```{r} -->
+<!-- vita = piecewise_mod(outcome = "VitaminA_Retinol",data = df,diagnostics = T) -->
+<!-- ``` -->
 
-### Single modulator
+<!-- ### Single modulator -->
 
-```{r results='asis'}
-vita$plot_single
-kable(vita$results_single,digits = 3)
-```
+<!-- ```{r results='asis'} -->
+<!-- vita$plot_single -->
+<!-- kable(vita$results_single,digits = 3) -->
+<!-- ``` -->
 
-### Double modulator
+<!-- ### Double modulator -->
 
-```{r results='asis'}
-vita$plot_double
-kable(vita$results_double,digits = 3)
-```
+<!-- ```{r results='asis'} -->
+<!-- vita$plot_double -->
+<!-- kable(vita$results_double,digits = 3) -->
+<!-- ``` -->
 
-## BMI
+<!-- ## BMI -->
 
-```{r}
-bmi = piecewise_mod(outcome = "BMI",data = df,diagnostics = T)
-```
+<!-- ```{r} -->
+<!-- bmi = piecewise_mod(outcome = "BMI",data = df,diagnostics = T) -->
+<!-- ``` -->
 
-### Single modulator
+<!-- ### Single modulator -->
 
-```{r results='asis'}
-bmi$plot_single
-kable(bmi$results_single,digits = 3)
-```
+<!-- ```{r results='asis'} -->
+<!-- bmi$plot_single -->
+<!-- kable(bmi$results_single,digits = 3) -->
+<!-- ``` -->
 
-### Double modulator
+<!-- ### Double modulator -->
 
-```{r results='asis'}
-bmi$plot_double
-kable(bmi$results_double,digits = 3)
-```
+<!-- ```{r results='asis'} -->
+<!-- bmi$plot_double -->
+<!-- kable(bmi$results_double,digits = 3) -->
+<!-- ``` -->
 
-# Bayesian?
+<!-- # Bayesian? -->
 
-## BMI
+<!-- ## BMI -->
 
-```{r cache=TRUE,echo=TRUE}
-mod_df = df %>% filter(!is.na(age_second)) %>% select(Patient.ID,time,BMI) %>% drop_na()
-model = list(
-  BMI ~ 1, # Plateau
-  1 + (1|Patient.ID) ~ 0 + time, # Joined slope
-  1 + (1|Patient.ID) ~ 0 + time, # Joined slope
-  1 + (1|Patient.ID) ~ 0 + time # Joined slope
-)
-fit = mcp(model, data = mod_df,sample = "both")
-```
+<!-- ```{r cache=TRUE,echo=TRUE} -->
+<!-- mod_df = df %>% filter(!is.na(age_second)) %>% select(Patient.ID,time,BMI) %>% drop_na() -->
+<!-- model = list( -->
+<!--   BMI ~ 1, # Plateau -->
+<!--   1 + (1|Patient.ID) ~ 0 + time, # Joined slope -->
+<!--   1 + (1|Patient.ID) ~ 0 + time, # Joined slope -->
+<!--   1 + (1|Patient.ID) ~ 0 + time # Joined slope -->
+<!-- ) -->
+<!-- fit = mcp(model, data = mod_df,sample = "both") -->
+<!-- ``` -->
 
-### Posterior means
+<!-- ### Posterior means -->
 
-```{r}
-summary(fit)
-```
+<!-- ```{r} -->
+<!-- summary(fit) -->
+<!-- ``` -->
 
-### Posterior fit
+<!-- ### Posterior fit -->
 
-```{r}
-plot(fit,q_fit = TRUE) + theme_bw()
-```
+<!-- ```{r} -->
+<!-- plot(fit,q_fit = TRUE) + theme_bw() -->
+<!-- ``` -->
 
-### Parameter plot
+<!-- ### Parameter plot -->
 
-```{r}
-plot_pars(fit)
-```
+<!-- ```{r} -->
+<!-- plot_pars(fit) -->
+<!-- ``` -->
 
-### Hypothesis tests
+<!-- ### Hypothesis tests -->
 
-```{r}
-hypothesis(fit,c("cp_1 = 0","cp_2 = 0","cp_3 = 0"))
-```
+<!-- ```{r} -->
+<!-- hypothesis(fit,c("cp_1 = 0","cp_2 = 0","cp_3 = 0")) -->
+<!-- ``` -->
+
+<!-- t -->

### 2022-04-20 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
correlations

@@
 library(pracma)
+library(corrplot)
 library(knitr)
@@
 
-# Figures
+# Correlations
+
+White text indicates the correlation coefficient $R$, and results are only shown for correlations significant at the p < 0.05 level.
+
+```{r fig.width=10,fig.height=10}
+change_df = df %>% group_by(Study.ID) %>% 
+  summarise_at(vars(Weight..kg.,c.peptide.0.minute,`C Peptide Index`,
+                    `Adapted Matsuda`,`C Peptide ODI`,iAUC_ratio,homa2B:homa2IR),
+               .funs = diff) %>%
+  rename("Change in Weight (kg)" = Weight..kg.,
+         "Change in Fasting C-Peptide (pmol/L)" = c.peptide.0.minute,
+         "Change in C Peptide Index" = `C Peptide Index`,
+         "Change in Adapted Matsuda" = `Adapted Matsuda`,
+         "Change in C Peptide ODI" = `C Peptide ODI`,
+         "Change in C pep iAUC:Glucose iAUC" = iAUC_ratio,
+         "Change in HOMA2 %B" = homa2B, "Change in HOMA2 %S" = homa2S,
+         "Change in HOMA2 IR" = homa2IR)
+
+M = cor(change_df[,2:ncol(change_df)])
+testRes = cor.mtest(change_df[,2:ncol(change_df)])
+corrplot(M, p.mat = testRes$p, method = 'circle', type = 'lower', insig='blank',
+         addCoef.col ='white', number.cex = 0.8, order = 'AOE', diag=FALSE)
+```
+
+# Other Figures
 

### 2022-04-21 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
limit nutrition to those with sweat test data

@@
 })
+# "Our dataset"
+sweat = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report Electrolytes_throughDecember2021.xlsx")
+df = df %>% filter(`Patient ID` %in% sweat$`Patient ID`)
 groups = df %>% group_by(`Patient ID`) %>% filter(row_number()==1)
@@
   group_by(`Patient ID`,Group) %>% count(before_first) %>% 
-  filter(before_first == T,n >= 2)
+  filter(before_first == T,n >= 1)
 # At least one nutritional measure and one weight/height measure >= 3 months after first
@@
   filter(after_first == T,n >= 1)
-# Write merged df
-write.csv(df,file = "./Edith Zemanick/CF Nutrition/Data_Cleaned/merged_data.csv",
+# Write merged df - filter to just Vitamin A values
+t = df %>% 
+  mutate(`Pre/Post First Mod` = ifelse(Date > `First Mod Date`,"Post","Pre")) %>%
+  filter(!is.na(VitaminA_Retinol)) 
+write.csv(t,file = "./Edith Zemanick/CF Nutrition/Data_Cleaned/merged_data.csv",
           row.names = F,na="")
@@
 [2]: paste0('Treated With Ivacaftor or Trikafta From 2008-2021 (n = ',length(unique(mods$`Patient ID`)), ')')
-[3]: paste0('In Our Dataset (n = ',length(unique(df$`Patient ID`)), ')')
+[3]: paste0('In Our Dataset (n = ',length(intersect(mods$`Patient ID`,sweat$`Patient ID`)), ')')
 [4]: paste0('Treated With Ivacaftor Alone From 2008-2021 (n = ',as.numeric(table(groups$Group)['Kalydeco']), ')')
@@
 [6]: paste0('Treated With Ivacaftor followed by Trikafta From 2008-2021 (n = ',as.numeric(table(groups$Group)['Both']), ')')
-[7]: paste0('At Least Two Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Kalydeco']), ')')
-[8]: paste0('At Least Two Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Trikafta']), ')')
-[9]: paste0('At Least Two Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Both']), ')')
+[7]: paste0('At Least One Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Kalydeco']), ')')
+[8]: paste0('At Least One Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Trikafta']), ')')
+[9]: paste0('At Least One Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Both']), ')')
 [10]: paste0('At Least One Vitamin A >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Kalydeco']), ')')

### 2022-04-22 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
CGM pre/post trikafta data cleaning

@@
 df = left_join(df,cgm,by = c("Study.ID", "Event.Name"))
-keep_a1c = df %>% group_by(Study.ID) %>% 
-  summarise(n = sum(!is.na(A1C.result))) %>%
-  filter(n == 2) %>% pull(Study.ID)
 # Update DXA variables
@@
              control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),
-             data = df[df$Event.Name == "Visit 1" & df$Study.ID %in% keep_a1c,])
+             data = df[df$Event.Name == "Visit 1",])
 summary(t1,digits = 2,labelTranslations = newlabels)
@@
 f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
-p <- paired(f,data = df[df$Study.ID %in% keep_a1c,], id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F,
+p <- paired(f,data = df, id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F,
             control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),)
@@
 f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
-p <- paired(f,data = df[df$Study.ID %in% keep_a1c,], id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F,
+p <- paired(f,data = df, id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F,
             control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),

### 2022-04-22 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
CGM pre/post trikafta data cleaning

@@
 df = df %>% group_by(CFF.ID) %>% fill(DOB,Start:CFRD.Diagnosis.Date,.direction = "downup") %>%
-  ungroup() %>% filter(!is.na(Time),abs(Time) <= 365,!is.na(CFF.ID),
+  ungroup() %>% filter(!is.na(Time),abs(Time) <= 730,!is.na(CFF.ID),
                        (!is.na(CFRD.Diagnosis.Date) | !is.na(CFRD.Status) | site == "UW")) %>%
@@
 # Before or after start
-df$Before_After = cut(df$Time,c(-Inf,0,Inf),labels = c("Before","After"),right = F)
-# Most recent before/after
-most_recent = df %>% group_by(CFF.ID,Before_After) %>% slice_min(abs(Time))
+df$Before_After = cut(df$Time,c(-Inf,0,Inf),labels = c("Before","After"),right = T)
+# Sort
+df = df %>% arrange(Site,CFF.ID,Date)
+# Write sorted DF
+write.csv(df, file = "./Colleen Wood/Prepost Triakfta/Data_Cleaned/analysis_data_with_UM.csv",
+          row.names = F,na = "")
+# Most recent before/after - limit to 2 years before start
+# Want the value closest to start date before start, and most recent with respect to now from start
+# (i.e. the longest time after start)
+most_recent = df %>% group_by(CFF.ID,Site,Before_After) %>%
+  mutate(time_new = max(Time) - Time) %>%
+  arrange(time_new) %>% 
+  summarise(BMIz = first(na.omit(BMIz)),ppFEV = first(na.omit(ppFEV)),
+            ppFVC = first(na.omit(ppFVC))) %>%
+  arrange(Site,CFF.ID,Before_After)
+t = df %>% select(CFF.ID,Site,Date,Time,Before_After,BMIz,ppFEV,ppFVC) %>%
+  group_by(CFF.ID,Site,Before_After) %>%
+  arrange(Site,CFF.ID,Before_After)
+keep = most_recent %>% group_by(CFF.ID) %>% count() %>% filter(n == 2)
 ```
+
+# Participant Characteristics
 

### 2022-04-22 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_merge.Rmd
CGM pre/post trikafta data cleaning

--- 2022-02-02 Christine Chan/Pre-post Trikafta/pre_post_trikafta_merge.Rmd
@@
 if(Sys.info()["sysname"] == "Windows"){
-  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan"
+  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood"
 } else if (Sys.info()["sysname"] == "Linux"){
-  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan"
+  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood"
 } else if (Sys.info()["sysname"] == "Darwin"){
-  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan"
+  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood"
 }
@@
 chco_bmi_pft = read_csv("./Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv")
-
+chco_bmi_pft$Date = mdy(chco_bmi_pft$Date) 
 # continuous glucose monitoring, 23 obs
@@
 chco_cgm = read_csv("./Prepost Triakfta/Data_Cleaned/chco_cgm.csv")
-
+chco_cgm$Date = date(parse_date_time(chco_cgm$date_cgm_placement,
+                                                orders = c("mdyHM","mdy")))
 # chco glycemic, 261 obs
@@
 chco_glycemic = read_csv("./Prepost Triakfta/Data_Cleaned/chco_glycemic.csv")
-
+chco_glycemic$Date = mdy(chco_glycemic$Date)
 # chco registry 108 obs
@@
 montana_cgm <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_cgm.csv")
+montana_cgm$Date = date(parse_date_time(montana_cgm$date_cgm_placement,
+                                                orders = c("mdyHM","mdy")))
 
@@
 # glycemic 279
-montana_glycemic_registry <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv") # looks identical
+montana_glycemic_registry <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv")
+montana_glycemic_registry$Date = mdy(montana_glycemic_registry$Date)
+# looks identical
 ###################################
@@
 note: by default, the dplyr join functions join by all variables in common. So we need to make sure the common variables are same format and name
+
 ```{r form stitching}
@@
 
-# Get a Date variable by taking the date from "date time" char variable date_cgm_placement
-# weird because one obs isnt in a date time, just a date so ill just brute it
-chco_cgm = chco_cgm %>% mutate(datewithtime = str_detect(date_cgm_placement, " "),
-                               Date = ifelse(datewithtime,sub(" .*", "", chco_cgm$date_cgm_placement), date_cgm_placement))
-
 # per tim, its ok to have more than 2 dates per subject,
@@
 
-chco_long = full_join( chco_registry, chco_long) # joins by MRN, CFFID, Start
+chco_long = full_join(chco_registry, chco_long) # joins by MRN, CFFID, Start
 chco_long = full_join(chco_long, chco_cgm) # joins by cff id and date
@@
                                              Weight = `Weight of Patient`)
+montana_bmi_pft$Date = mdy(montana_bmi_pft$Date)
 
@@
 montana_registry = full_join(montana_registry, montana_gtubes)
+montana_registry$Date = mdy(montana_registry$Date)
 
@@
 # Get a Date variable by taking the date from "date time" char variable date_cgm_placement
-montana_cgm = montana_cgm %>% mutate(datewithtime = str_detect(date_cgm_placement, " "),
-                               Date = ifelse(datewithtime,sub(" .*", "", montana_cgm$date_cgm_placement), date_cgm_placement))
+montana_cgm$Date = mdy(montana_cgm$Date)
 
@@
 uw_long = full_join(uw_long, uw_cgm) # joins by cff id and date
-
+uw_long$Date = mdy(uw_long$Date)
 # add in a site id variable
@@
 ```
+
 TO ASK TIM: 

### 2022-04-28 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
fixed stupid DOB issue

@@
 })
+# Fix incorrect century (WHY LUBRIDATE WHY?!)
+df$DOB[which(df$DOB > lubridate::mdy("1/1/2022"))] = 
+  df$DOB[which(df$DOB > lubridate::mdy("1/1/2022"))] - years(100)
 # Time from start and filtering
@@
 um = um %>% filter(!is.na(CFRD.Status))
+um$Age = um$Age / 365.25
 df = full_join(df,um)
@@
 # BMI category
-df$Age = pmax(as.numeric(difftime(df$Date,df$DOB,units = "days")),df$Age,na.rm = T)/365.25
+df$Age = pmax(as.numeric(difftime(df$Date,df$DOB,units = "days"))/365.25,df$Age,na.rm = T)
 # Treat adults as age = 20
@@
 # Sort
-df = df %>% arrange(Site,CFF.ID,Date)
+df = df %>% arrange(Site,CFF.ID,Date) %>% distinct()
 # Write sorted DF
@@
           row.names = F,na = "")
+```
+
+```{r}
 # Most recent before/after - limit to 2 years before start

### 2022-04-28 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
colleen wood data mostly cleaned

@@
 }
+source("~/GitHub/CF/get_genotype_class_severity.R")
 knitr::opts_knit$set(root.dir = home_dir)
@@
 df$BMIz[is.infinite(df$BMIz)] = NA
+df$BMI.Percentile = pnorm(df$BMIz)*100
+df$Weightz = y2z(df$Weight,x = df$Age_for_BMI,sex = df$Sex,ref=nl4.wgt)
 df$BMI_group = apply(df,1,function(r){
@@
                                       "Obese (Class I)","Obese (Class II)","Obese (Class III+)"))
+# Get genotype severity
+df$Genotype1_sev = get_cf_genotype_class_severity(df$Genotypes1)
+df$Genotype2_sev = get_cf_genotype_class_severity(df$Genotypes2)
 # PFTs
@@
             FEV1 = df$FVC)
+# T1D duration
+df$CFRD.Diagnosis.Date = mdy(df$CFRD.Diagnosis.Date)
+df$T1D.Duration = as.numeric(difftime(df$Date,df$CFRD.Diagnosis.Date,units = "days"))/365.25
 # Before or after start
@@
 most_recent = df %>% group_by(CFF.ID,Site,Before_After) %>%
+  filter(Time > -2*365) %>%
   mutate(time_new = max(Time) - Time) %>%
   arrange(time_new) %>% 
-  summarise(BMIz = first(na.omit(BMIz)),ppFEV = first(na.omit(ppFEV)),
-            ppFVC = first(na.omit(ppFVC))) %>%
+  summarise(Age = first(na.omit(Age)),BMI.Percentile = first(na.omit(BMI.Percentile)),
+            Weight = first(na.omit(Weight)),Sex = first(na.omit(Sex)),
+            T1D.Duration = first(na.omit(T1D.Duration)),
+            G.tube.in.past.12.months = first(na.omit(G.tube.in.past.12.months)),
+            Pancreatic.Status = first(na.omit(Pancreatic.Status)),
+            Genotype1_sev = first(na.omit(Genotype1_sev)),
+            Genotype2_sev = first(na.omit(Genotype2_sev)),
+            BMIz = first(na.omit(BMIz)),Weightz = first(na.omit(Weightz)),
+            FEV1 = first(na.omit(FEV1)),FVC = first(na.omit(FVC)),
+            HbA1c = first(na.omit(HbA1c)),.groups = "keep") %>%
   arrange(Site,CFF.ID,Before_After)
-t = df %>% select(CFF.ID,Site,Date,Time,Before_After,BMIz,ppFEV,ppFVC) %>%
-  group_by(CFF.ID,Site,Before_After) %>%
-  arrange(Site,CFF.ID,Before_After)
-keep = most_recent %>% group_by(CFF.ID) %>% count() %>% filter(n == 2)
+# # Check
+# t = df %>% select(CFF.ID,Site,Date,Time,Before_After,BMIz,ppFEV,ppFVC) %>%
+#   group_by(CFF.ID,Site,Before_After) %>%
+#   arrange(Site,CFF.ID,Before_After)
+# Filter those without CFRD, and those without pre AND post
+keep = most_recent %>% filter(!is.na(HbA1c)) %>%
+  group_by(CFF.ID) %>% summarise(n = n()) %>%
+  filter(n == 2) %>% pull(CFF.ID)
+cfrd = unique(df$CFF.ID[!is.na(df$CFRD.Status) | !is.na(df$CFRD.Diagnosis.Date)])
+keep = unique(keep,cfrd)
+most_recent = most_recent %>% filter(CFF.ID %in% keep)
 ```

### 2022-04-28 get_genotype_class_severity.R

@@ -40,9 +40,6 @@ get_cf_genotype_class_severity = function(genotypes,out = "severity",na_unknown 
                            "Q1209P","Unknown","Unknown","S492F","Unknown","Mild",
                            "R1066H","Unknown","Unknown","M1101K","Unknown","Severe",
                            "CFTRdele1","Unknown","Severe"),ncol = 3,byrow = T)
-  
-  
-  
   colnames(severity_only) = c("Genotype","Class","Severity")
   severity_only[,1] = tolower(gsub("[[:punct:]]|\\s","",severity_only[,1]))
   geno = rbind(geno,severity_only)

### 2022-04-28 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
added dates to pre/post

@@
   arrange(time_new) %>% 
-  summarise(Age = first(na.omit(Age)),BMI.Percentile = first(na.omit(BMI.Percentile)),
-            Weight = first(na.omit(Weight)),Sex = first(na.omit(Sex)),
-            T1D.Duration = first(na.omit(T1D.Duration)),
+  summarise(Age = paste0(Date[which(Age == first(na.omit(Age)))],": ",round(first(na.omit(Age)),1)),
+            BMI.Percentile = paste0(Date[which(BMI.Percentile == first(na.omit(BMI.Percentile)))],
+                                    ": ",round(first(na.omit(BMI.Percentile)),1)),
+            Weight = paste0(Date[which(Weight == round(first(na.omit(Weight)),1))],
+                            ": ",first(na.omit(Weight))),
+            T1D.Duration = paste0(Date[which(T1D.Duration == first(na.omit(T1D.Duration)))],
+                                  ": ",round(first(na.omit(T1D.Duration)),1)),
             G.tube.in.past.12.months = first(na.omit(G.tube.in.past.12.months)),
@@
             Genotype2_sev = first(na.omit(Genotype2_sev)),
-            BMIz = first(na.omit(BMIz)),Weightz = first(na.omit(Weightz)),
-            FEV1 = first(na.omit(FEV1)),FVC = first(na.omit(FVC)),
-            HbA1c = first(na.omit(HbA1c)),.groups = "keep") %>%
+            BMIz = paste0(Date[which(BMIz == first(na.omit(BMIz)))],
+                          ": ",round(first(na.omit(BMIz)),1)),
+            FEV1 = paste0(Date[which(FEV1 == first(na.omit(FEV1)))[1]],
+                          ": ",round(first(na.omit(FEV1)),1)),
+            FVC = paste0(Date[which(FVC == first(na.omit(FVC)))],
+                         ": ",round(first(na.omit(FVC)),1)),
+            HbA1c = paste0(Date[which(HbA1c == first(na.omit(HbA1c)))],
+                           ": ",round(first(na.omit(HbA1c)),1)),
+            .groups = "keep") %>%
   arrange(Site,CFF.ID,Before_After)
+most_recent[most_recent == ": NA"] = NA
+most_recent[most_recent == "NA: NA"] = NA
+
 # # Check

### 2022-05-05 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
check Elin's updated data

@@
 # Vitamins
-vit = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report Vitamins_throughDecember2021.xlsx")
+vit = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report Vitamins_throughDecember2021_revised20220429.xlsx")
+vit$Date = date(ymd_hms(vit$Date))
+vit2 = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report Vitamins_throughDecember2021.xlsx")
+vit2$Date = date(ymd_hms(vit2$Date))
+diff = setdiff(vit[,c("Patient ID","Date")],vit2[,c("Patient ID","Date")])
+t = vit[vit$`Patient ID` %in% diff$`Patient ID`,]
+t2 = vit2[vit2$`Patient ID` %in% diff$`Patient ID`,]
+
 vit = vit %>% select(`Patient ID`,Date,VitaminA_Retinol:`Vitamin E Gamma`)

### 2022-05-05 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
two years pre CFTR

--- 2022-04-28 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
@@
 most_recent = df %>% group_by(CFF.ID,Site,Before_After) %>%
-  filter(Time > -2*365) %>%
+  filter(Time >= (-2*365)) %>%
   mutate(time_new = max(Time) - Time) %>%
   arrange(time_new) %>% 
-  summarise(Age = first(na.omit(Age)),BMI.Percentile = first(na.omit(BMI.Percentile)),
-            Weight = first(na.omit(Weight)),Sex = first(na.omit(Sex)),
-            T1D.Duration = first(na.omit(T1D.Duration)),
+  summarise(Start = first(na.omit(Start)),
+            Age = paste0(Date[which(Age == first(na.omit(Age)))],": ",round(first(na.omit(Age)),1)),
+            BMI.Percentile = paste0(Date[which(BMI.Percentile == first(na.omit(BMI.Percentile)))],
+                                    ": ",round(first(na.omit(BMI.Percentile)),1)),
+            Weight = paste0(Date[which(Weight == round(first(na.omit(Weight)),1))],
+                            ": ",first(na.omit(Weight))),
+            T1D.Duration = paste0(Date[which(T1D.Duration == first(na.omit(T1D.Duration)))],
+                                  ": ",round(first(na.omit(T1D.Duration)),1)),
             G.tube.in.past.12.months = first(na.omit(G.tube.in.past.12.months)),
@@
             Genotype2_sev = first(na.omit(Genotype2_sev)),
-            BMIz = first(na.omit(BMIz)),Weightz = first(na.omit(Weightz)),
-            FEV1 = first(na.omit(FEV1)),FVC = first(na.omit(FVC)),
-            HbA1c = first(na.omit(HbA1c)),.groups = "keep") %>%
+            BMIz = paste0(Date[which(BMIz == first(na.omit(BMIz)))],
+                          ": ",round(first(na.omit(BMIz)),1)),
+            FEV1 = paste0(Date[which(FEV1 == first(na.omit(FEV1)))[1]],
+                          ": ",round(first(na.omit(FEV1)),1)),
+            FVC = paste0(Date[which(FVC == first(na.omit(FVC)))],
+                         ": ",round(first(na.omit(FVC)),1)),
+            HbA1c = paste0(Date[which(HbA1c == first(na.omit(HbA1c)))],
+                           ": ",round(first(na.omit(HbA1c)),1)),
+            .groups = "keep") %>%
   arrange(Site,CFF.ID,Before_After)
+most_recent$Start = as.character(most_recent$Start)
+most_recent[most_recent == ": NA"] = NA
+most_recent[most_recent == "NA: NA"] = NA
+
 # # Check
@@
   filter(n == 2) %>% pull(CFF.ID)
-cfrd = unique(df$CFF.ID[!is.na(df$CFRD.Status) | !is.na(df$CFRD.Diagnosis.Date)])
+cfrd = unique(df$CFF.ID[!is.na(df$CFRD.Status) | 
+                          df$CFRD.Status != "Impaired Glucose Tolerance (FBG < 126, 2-h PG 140-199)" | 
+                          !is.na(df$CFRD.Diagnosis.Date)])
 keep = unique(keep,cfrd)
 most_recent = most_recent %>% filter(CFF.ID %in% keep)
+# Write 
+write.csv(most_recent,file = "./Colleen Wood/Prepost Triakfta/Data_Cleaned/pre_post.csv",
+          na = "",row.names = F)
 ```

### 2022-05-15 StatNet/knot_models.qmd
Create knot_models.qmd

---
title: "Piecewise Regression for Nutritional Outcomes Pre- and Post-CFTR Modulator"
author: "Tim Vigers"
format: revealjs
editor: visual
bibliography: /Users/timvigers/Dropbox/Miscellaneous/references.bib
csl: /Users/timvigers/GitHub/styles/american-medical-association-alphabetical.csl
---

```{r}
#| include: false
# Load libraries, data, etc.
library(DiagrammeR)
library(MultiKink)
library(ggplot2)
data("triceps")
```

## Study Overview

-   Do vitamin levels change after CFTR modulator start?

-   Registry data through 2021:

    -   Most patients have multiple nutrition measures before and after CFTR modulator start.
    -   A subset of patients were on ivacaftor followed by elexacaftor/tezacaftor/ivacaftor (ETI).

## Study Overview

-   The first approach was to limit the analysis dataset to those with two measures prior to modulator start, and at least one measure after.

    -   Then we did a t test on the differences between measures closest to modulator start.

-   However, this results in a lot of excluded data and it was unclear how to treat those on two different modulators.

-   Piecewise models allow us to use all the data and include all the participants in the same model.

## Piecewise Linear Regression Basics

-   See [module and presentation](https://bookdown.org/tpinto_home/Beyond-Linearity) by Jaroslaw Harezlak & Armando Teixeira-Pinto for the Biostatistics Collaboration of Australia for additional detail.

-   Will use the "triceps" data included in the `MultiKink` R package [@wanMultiKinkEstimationInference2020] because it's a little easier to see the trends.

-   892 measurements of the triceps skin fold.

    -   Interested in modeling changes with age.

---

Triceps data:

```{r}
tri.age.plot = ggplot(triceps,aes(x = age,y = triceps)) + geom_point(alpha = 0.5) +
  xlab("Age") + ylab("Triceps") + theme_bw()
tri.age.plot
```

---

Add knots at ages 5, 10, 20, 30, and 40:

```{r}
pred1 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age<5,]))
pred2 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age >=5 & triceps$age<10,]))
pred3 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age>=10 & triceps$age<20,]))
pred4 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age>=20 & triceps$age<30,]))
pred5 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age>=30 & triceps$age<40,]))
pred6 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age>=40,]))
tri.age.plot + 
  geom_line(data=triceps[triceps$age<5,], 
            aes(y = pred1, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age >=5 & triceps$age<10,], 
            aes(y = pred2, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age>=10 & triceps$age<20,], 
            aes(y = pred3, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age>=20 & triceps$age<30,], 
            aes(y = pred4, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age>=30 & triceps$age<40,], 
            aes(y = pred5, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age>=40,], 
            aes(y = pred6, x=age), size = 1, col="blue") 
```

## Continuity restriction

-   We can adapt the previous model to make sure that the regressions "join up" at each knot:

$$
y = \beta_0 + \beta_1 x+\beta_2(x-k_1)_+\beta_3(x-k_2)_+ +...+ \epsilon
$$

with

$$
(x-k)_+=
\left\{ 
  \begin{array}{ c l }
    0, & \quad \textrm{if } x < k \\
    x-k,                 & \quad \textrm{if } x \geq k 
  \end{array}
\right.
$$

```{r}
#| eval: false
grViz("
digraph graph2 {
graph [layout = dot]
# node definitions with substituted label text
node [shape = rectangle]
a [label = '@@1']
b [label = '@@2']
c [label = '@@3']
d [label = '@@4']
e [label = '@@5']
f [label = '@@6']
g [label = '@@7']
h [label = '@@8']
i [label = '@@9']
j [label = '@@10']
k [label = '@@11']
l [label = '@@12']
m [label = '@@13']

a -> b
b -> c
c -> d
c -> e
c -> f
d -> g
e -> h
f -> i
g -> j
h -> k
i -> l
j -> m
k -> m
l -> m
}

[1]: paste0('Clinical Database (n = ', ncdb, ')')
[2]: paste0('Treated With Ivacaftor or Trikafta From 2008-2021 (n = ',length(unique(mods$`Patient ID`)), ')')
[3]: paste0('In Our Dataset (n = ',length(intersect(mods$`Patient ID`,sweat$`Patient ID`)), ')')
[4]: paste0('Treated With Ivacaftor Alone From 2008-2021 (n = ',as.numeric(table(groups$Group)['Kalydeco']), ')')
[5]: paste0('Treated With Trikafta Alone From 2008-2021 (n = ',as.numeric(table(groups$Group)['Trikafta']), ')')
[6]: paste0('Treated With Ivacaftor followed by Trikafta From 2008-2021 (n = ',as.numeric(table(groups$Group)['Both']), ')')
[7]: paste0('At Least One Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Kalydeco']), ')')
[8]: paste0('At Least One Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Trikafta']), ')')
[9]: paste0('At Least One Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Both']), ')')
[10]: paste0('At Least One Vitamin A >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Kalydeco']), ')')
[11]: paste0('At Least One Vitamin A >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Trikafta']), ')')
[12]: paste0('At Least One Vitamin A >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Both']), ')')
[13]: paste0('Final (n = ',length(unique(after_measure$`Patient ID`)), ')')
")
```

# References

### 2022-05-15 StatNet/knot_models.qmd
Update knot_models.qmd

@@
 library(ggplot2)
+library(broom)
+library(knitr)
 data("triceps")
@@
 
-Add knots at ages 5, 10, 20, 30, and 40:
+Add knots at ages 5, 10, and 20:
 
@@
 pred4 <- predict(lm(triceps~age, 
-                    data = triceps[triceps$age>=20 & triceps$age<30,]))
-pred5 <- predict(lm(triceps~age, 
-                    data = triceps[triceps$age>=30 & triceps$age<40,]))
-pred6 <- predict(lm(triceps~age, 
-                    data = triceps[triceps$age>=40,]))
+                    data = triceps[triceps$age>=20,]))
 tri.age.plot + 
@@
             aes(y = pred3, x=age), size = 1, col="blue") +
-  geom_line(data=triceps[triceps$age>=20 & triceps$age<30,], 
-            aes(y = pred4, x=age), size = 1, col="blue") +
-  geom_line(data=triceps[triceps$age>=30 & triceps$age<40,], 
-            aes(y = pred5, x=age), size = 1, col="blue") +
-  geom_line(data=triceps[triceps$age>=40,], 
-            aes(y = pred6, x=age), size = 1, col="blue") 
+  geom_line(data=triceps[triceps$age>=20,], 
+            aes(y = pred4, x=age), size = 1, col="blue") 
 ```
@@
 
+## Manual coding
+
+```{r}
+#| echo: true
+# Make new variables
+triceps$c5 = (triceps$age-5)*(triceps$age>=5)
+triceps$c10 = (triceps$age-10)*(triceps$age>=10)
+triceps$c20 = (triceps$age-20)*(triceps$age>=20)
+```
+
+```{r}
+kable(triceps[c(which(triceps$age<5)[1],which(triceps$age<10&triceps$age>5)[1],
+          which(triceps$age>20&triceps$age>10)[1]),],row.names = F)
+```
+
+---
+
+Plot predicted values:
+
+```{r}
+# Linear model
+mod = lm(triceps ~ age + c5 + c10 + c20,data = triceps)
+# Plot
+pred = predict(mod)
+tri.age.plot +
+  geom_line(data=triceps, 
+            aes(y = pred, x=age), size = 1, col="blue") 
+```
+
+## Model interpretation
+
+```{r}
+kable(tidy(mod),digits = 3)
+```
+
+The estimates of c5, c10, and c20 indicate change in slope at the respective knot.
+
 ```{r}
@@
 grViz("
-digraph graph2 {
-graph [layout = dot]
-# node definitions with substituted label text
-node [shape = rectangle]
-a [label = '@@1']
-b [label = '@@2']
-c [label = '@@3']
-d [label = '@@4']
-e [label = '@@5']
-f [label = '@@6']
 g [label = '@@7']

### 2022-05-15 StatNet/knot_models.qmd
Update knot_models.qmd

@@
 data("triceps")
+cf = read.csv("/Users/timvigers/GitHub/CF/StatNet/data.csv",
+              stringsAsFactors = F,na.strings = "")
 ```
@@
 
----
+------------------------------------------------------------------------
 
@@
 ```{r}
-tri.age.plot = ggplot(triceps,aes(x = age,y = triceps)) + geom_point(alpha = 0.5) +
+tri.age.plot = ggplot(triceps,aes(x = age,y = triceps)) + geom_point(alpha = 0.6) +
   xlab("Age") + ylab("Triceps") + theme_bw()
@@
 
+-   Can add these values by hand easily (using the difference \* an indicator):
+
 ```{r}
@@
 
+-   Examples of what the new variables should look like:
+
 ```{r}
 kable(triceps[c(which(triceps$age<5)[1],which(triceps$age<10&triceps$age>5)[1],
-          which(triceps$age>20&triceps$age>10)[1]),],row.names = F)
+          which(triceps$age<20&triceps$age>10)[1],
+          which(triceps$age>=20)[1]),c("age","c5","c10","c20")],row.names = F)
 ```
 
----
+------------------------------------------------------------------------
 
@@
 
-The estimates of c5, c10, and c20 indicate change in slope at the respective knot.
+\
+The estimates of c5, c10, and c20 indicate change in slope at each respective knot.
+
+## Pre-/Post-CFTR Data
 
@@
 
+## Vitamin A Plot
+
+```{r}
+vita_plot = ggplot(cf,aes(x = Time,y = Vitamin.A,color = factor(Patient.ID))) + 
+  geom_point(alpha = 0.6) +
+  ylab("Vitamin A") + theme_bw() + theme(legend.position = "none")
+```
+
+```{r}
+mod = lmer(Vitamin.A ~ Time + c1 + c2 + (1|Patient.ID),data = cf)
+```
+
+### To Do
+
+-   Change missing c2 values to 0
+
+-   Write out mixed model with two knots
+
+-   Summarize results
+
+## A Bayesian Approach
+
 # References

### 2022-05-15 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
Update cf_nutrition.Rmd
clean data for statnet presentation

@@
 df = left_join(df,pancreatic,by = "Patient ID")
+# Time relative to first mod
+df$Time = as.numeric(difftime(df$Date,df$`First Mod Date`,units = "days"))/365.25
+# Calculate spline things
+df$c1 = (df$Time-0)*(df$Time>=0)
+df$c2 = as.numeric(difftime(df$Date,df$`Second Mod Date`,units = "days"))/365.25
+df$c2 = df$c2*(df$Date >= df$`Second Mod Date`)
+# Convert to numeric
+df$VitaminA_Retinol = suppressWarnings(as.numeric(df$VitaminA_Retinol))
+# Select columns
+df = df %>% filter(!is.na(VitaminA_Retinol)) %>%
+  select(`Patient ID`,`First Mod Date`,`Second Mod Date`,
+         Date,Time,c1,c2,VitaminA_Retinol) %>%
+  rename(`Vitamin A` = VitaminA_Retinol)
+# Write for presentation
+write.csv(df,file = "~/GitHub/CF/StatNet/data.csv",row.names = F,na = "")
 ```
@@
 <!--   (as.numeric(difftime(df$Date,df$`Second Mod Date`,units = "days"))/365.25)*(df$Date >= df$`Second Mod Date`) -->
-<!-- # Limit to 3 years prior, select columns -->
+<!-- # Limit to 3 years prior -->
 <!-- df = df %>% filter(!is.na(time) & time >= -3) %>% distinct() -->

### 2022-05-15 StatNet/knot_models.qmd
Update knot_models.qmd
bayesian approach on triceps data

@@
 csl: /Users/timvigers/GitHub/styles/american-medical-association-alphabetical.csl
+cache: false
 ---
@@
 library(broom)
+library(lmerTest)
+library(mcp)
 library(knitr)
@@
 
-```{r}
-#| eval: false
-grViz("
-g [label = '@@7']
-h [label = '@@8']
-i [label = '@@9']
-j [label = '@@10']
-k [label = '@@11']
-l [label = '@@12']
-m [label = '@@13']
-
-a -> b
-b -> c
-c -> d
-c -> e
-c -> f
-d -> g
-e -> h
-f -> i
-g -> j
-h -> k
-i -> l
-j -> m
-k -> m
-l -> m
-}
-
-[1]: paste0('Clinical Database (n = ', ncdb, ')')
-[2]: paste0('Treated With Ivacaftor or Trikafta From 2008-2021 (n = ',length(unique(mods$`Patient ID`)), ')')
-[3]: paste0('In Our Dataset (n = ',length(intersect(mods$`Patient ID`,sweat$`Patient ID`)), ')')
-[4]: paste0('Treated With Ivacaftor Alone From 2008-2021 (n = ',as.numeric(table(groups$Group)['Kalydeco']), ')')
-[5]: paste0('Treated With Trikafta Alone From 2008-2021 (n = ',as.numeric(table(groups$Group)['Trikafta']), ')')
-[6]: paste0('Treated With Ivacaftor followed by Trikafta From 2008-2021 (n = ',as.numeric(table(groups$Group)['Both']), ')')
-[7]: paste0('At Least One Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Kalydeco']), ')')
-[8]: paste0('At Least One Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Trikafta']), ')')
-[9]: paste0('At Least One Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Both']), ')')
-[10]: paste0('At Least One Vitamin A >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Kalydeco']), ')')
-[11]: paste0('At Least One Vitamin A >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Trikafta']), ')')
-[12]: paste0('At Least One Vitamin A >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Both']), ')')
-[13]: paste0('Final (n = ',length(unique(after_measure$`Patient ID`)), ')')
-")
-```
+-   Limited to 3 years pre/post
 
@@
 
+-   The `mcp` package provides an easy interface for fitting Bayesian regression models with multiple change points.
+
+## On the triceps data
+
+```{r}
+# Define the model
+model = list(
+  triceps ~ 1 + age,          # intercept + slope
+  ~ 0 + age,  # joined slope with 3 change points
+  ~ 0 + age,
+  ~ 0 + age
+)
+bayes_fit = mcp(model, data = triceps, sample = "both")
+plot(bayes_fit)
+```
+
+-   Visually, this approach seems to generally confirm the knots we used earlier.
+
+-   The model results bear this out as well:
+
+```{r}
+s = summary(bayes_fit)
+kable(s,digits = 3)
+```
+
+---
+
+-   `mcp` also allows us to test hypotheses about the location of these knots:
+
+```{r}
+hypothesis(bayes_fit, "cp_1 = 5")
+hypothesis(bayes_fit, "cp_2 = 10")
+hypothesis(bayes_fit, "cp_3 = 20")
+```
+
+- None of the hypothesis tests are significant, indicating that the knots found with `mcp` are reasonably close to the ones we pre-specified (the CI on cp_1 is fairly wide though).
+
+---
+
+```{r}
+# Define the model
+model = list(
+  Vitamin.A ~ 1 + Time, # intercept + slope
+  ~ 0 + Time  # joined slope
+)
+fit = mcp(model, data = cf)
+plot(fit)
+```
+
+-   What does this mean? Why is the change point estimated to be before 0?
+
+---
+
+```{r}
+# Dirichlet prior on change point
+d_prior = list(
+  cp_1 = "dnorm(0,1)"
+)
+fit_cp_prior = mcp(model, prior = d_prior,data = cf)
+plot(fit_cp_prior)
+```
+
+---
+
 # References

### 2022-05-18 StatNet/knot_models.qmd
Update knot_models.qmd

@@
 csl: /Users/timvigers/GitHub/styles/american-medical-association-alphabetical.csl
-cache: false
+cache: true
 ---
@@
 
-## Pre-/Post-CFTR Data
+## Pre-/Post-CFTR Modulator Data
 
--   Limited to 3 years pre/post
+-   Vitamin D measures at annual visits from CF registry and our site.
 
-## Vitamin A Plot
+-   Some patients started on ivacaftor who then transitioned to ETI.
+
+## Pre-/Post-CFTR Modulator Data
 
 ```{r}
-vita_plot = ggplot(cf,aes(x = Time,y = Vitamin.A,color = factor(Patient.ID))) + 
-  geom_point(alpha = 0.6) +
-  ylab("Vitamin A") + theme_bw() + theme(legend.position = "none")
+vita_plot = ggplot(cf,aes(x = Time,y = Vitamin.D,color = factor(Patient.ID))) + 
+  geom_point() + geom_line(alpha = 0.2) +
+  ylab("Vitamin D") + theme_bw() + theme(legend.position = "none")
+vita_plot
 ```
 
+## Pre-/Post-CFTR Modulator Model
+
+-   The model is similar to triceps data example, but the second knot varies by patient and we include a random intercept to account for repeated measures:
+
+$$
+\text{Vitamin D}_{ij} = \beta_0 + \beta_1 \text{Time}_{ij} +\\ \beta_2(\text{Time}_{ij}-k_1)_+ + \beta_3(\text{Time}_{ij}-k_{2i})_+ + b_i + \epsilon_i
+$$
+
+## Pre-/Post-CFTR Modulator Model
+
+with
+
+$$
+(x-k_1)_+=
+\left\{ 
+  \begin{array}{ c l }
+    0, & \quad \textrm{if } \text{Time} < 0 \\
+    x-k_1,                 & \quad \textrm{if } \text{Time} \geq 0 
+  \end{array}
+\right.
+$$
+
+and
+
+$$
+(x-k_2)_+=
+\left\{ 
+  \begin{array}{ c l }
+    0, & \quad \textrm{if } \text{time before }2^{nd}\text{ mod}  \\
+    \text{time after }2^{nd}\text{ mod},                 & \quad \textrm{if } \text{time after }2^{nd}\text{ mod}
+  \end{array}
+\right.
+$$
+
+and
+
+$$
+\epsilon \sim N(0,\sigma^2_{\epsilon})\text{ and } b\sim N(0,\sigma^2_b)
+$$
+
+## Pre-/Post-CFTR Modulator Results
+
 ```{r}
-mod = lmer(Vitamin.A ~ Time + c1 + c2 + (1|Patient.ID),data = cf)
+mod = lmer(Vitamin.D ~ Time + c1 + c2 + (1|Patient.ID),data = cf)
+plot_df = mod@frame
+plot_df$pred = predict(mod)
+ggplot(plot_df,aes(x = Time,y = Vitamin.D,group = factor(Patient.ID))) +
+    geom_line(aes(y = pred),alpha = 0.5) + geom_point(aes(y = pred),alpha = 0.5) +
+    theme_bw() + theme(legend.position = "none") + xlab("Time (years)")
 ```
 
-### To Do
+## Pre-/Post-CFTR Modulator Results
 
--   Change missing c2 values to 0
-
--   Write out mixed model with two knots
-
--   Summarize results
+```{r}
+kable(broom.mixed::tidy(mod,"fixed")[,-1],digits = 3)
+```
 
@@
 -   The `mcp` package provides an easy interface for fitting Bayesian regression models with multiple change points.
+    -   Essentially it translates `lme4` syntax into JAGS.
+-   Uninformative "t-tail" priors are used unless the user specifies something else.
+    -   "The default priors should be suitable for estimation in the absence of strong prior knowledge."[@lindelovMcpPackageRegression2020a]
 
-## On the triceps data
+## On the Triceps Data
 
 ```{r}
+#| echo: true
 # Define the model
@@
   triceps ~ 1 + age,          # intercept + slope
-  ~ 0 + age,  # joined slope with 3 change points
-  ~ 0 + age,
-  ~ 0 + age
+  ~ 0 + age,  # joined slope 
+  ~ 0 + age,  # joined slope 
+  ~ 0 + age   # joined slope 
 )
+```
+
+## On the Triceps Data
+
+```{r}
+#| include: false
 bayes_fit = mcp(model, data = triceps, sample = "both")
-plot(bayes_fit)
+s = summary(bayes_fit)
+p = plot(bayes_fit)
+```
+
+```{r}
+print(p)
 ```
@@
 
+## On the Triceps Data
+
 -   The model results bear this out as well:
@@
 ```{r}
-s = summary(bayes_fit)
 kable(s,digits = 3)
@@
 
----
+------------------------------------------------------------------------
+
+## On the Triceps Data
 
@@
 ```{r}
+#| echo: true
 hypothesis(bayes_fit, "cp_1 = 5")
@@
 
----
+## A Bayesian Model for Pre-/Post-CFTR Modulator
+
+-   `mcp` can model joined slopes with varying change points:
 
 ```{r}
+#| echo: true
 # Define the model
 model = list(
-  Vitamin.A ~ 1 + Time, # intercept + slope
+  Vitamin.D ~ 1 + Time, # intercept + slope
+  ~ 0 + Time,  # joined slope
+  1 + (1|Patient.ID) ~ 0 + Time
+)
+```
+
+## A Bayesian Model for Pre-/Post-CFTR Modulator
+
+```{r}
+#| include: false
+fit = mcp(model, data = cf)
+s = summary(fit)
+p = plot(fit)
+```
+
+```{r}
+kable(s,digits = 3)
+```
+
+## A Bayesian Model for Pre-/Post-CFTR Modulator
+
+```{r}
+print(p)
+```
+
+-   This looks terrible!
+
+## A Bayesian Model for Pre-/Post-First CFTR Modulator
+
+-   Try fitting on just those with one modulator, with a N(0,1) prior on a change at 0:
+
+```{r}
+#| echo: true
+# Define the model
+model = list(
+  Vitamin.D ~ 1 + Time, # intercept + slope
   ~ 0 + Time  # joined slope
 )
-fit = mcp(model, data = cf)
-plot(fit)
+priors = list(
+  cp_1 = "dnorm(0,1)"
+)
 ```
 
--   What does this mean? Why is the change point estimated to be before 0?
+```{r}
+#| include: false
+fit = mcp(model,prior = priors,data = cf)
+s = summary(fit)
+p = plot(fit)
+```
 
----
+## A Bayesian Model for Pre-/Post-First CFTR Modulator
 
 ```{r}
-# Dirichlet prior on change point
-d_prior = list(
-  cp_1 = "dnorm(0,1)"
-)
-fit_cp_prior = mcp(model, prior = d_prior,data = cf)
-plot(fit_cp_prior)
+kable(s,digits = 3)
 ```
 
----
+## A Bayesian Model for Pre-/Post-First CFTR Modulator
+
+```{r}
+print(p)
+```
+
+## Limitations
+
+-   Only 20 two-modulator participants with vitamin D values after starting ETI (21 measures among them).
+
+    -   Large change in slope after second modulator start is likely driven by a few influential observations.
+
+-   Somewhat arbitrarily limited pre-modulator values to \< 5 years from modulator start.
+
+-   General data quality issues that we are in the process of resolving.
+
+## Questions
+
+-   How to handle truncated data (e.g. a value of "\< 0.1")?
+
+-   Should I model the two-modulator and one-modulator groups separately, but still use piecewise models?
+
+-   Any general advice (particularly for Bayesian methods)?
 

### 2022-05-18 StatNet/knot_models.qmd
Update knot_models.qmd

@@
 
+## Introduction
+
+-   Exocrine pancreatic insufficiency (PI) is present from birth in up to two thirds of patients with CF
+-   85% of patients with CF ultimately have PI with associated fat malabsorption by one year of age
+-   Malabsorption puts persons with CF at risk for fat-soluble vitamin deficiencies.
+
+## Introduction
+
+-   People with CF rely on supplemental pancreatic enzyme replacement therapy (PERT) in addition to a diet high in protein, calories, fat-soluble vitamins, and salt.
+-   Highly effective CFTR modulators, ivacaftor and elexacaftor/tezacaftor/ivacaftor (ETI), substantially improve CFTR activity, lung function and nutritional status.
+-   Their impact on other markers of nutrition, particularly fat-soluble vitamin absorption, are still unclear.
+
 ## Study Overview
@@
     -   Most patients have multiple nutrition measures before and after CFTR modulator start.
-    -   A subset of patients were on ivacaftor followed by elexacaftor/tezacaftor/ivacaftor (ETI).
+    -   A subset of patients were on ivacaftor followed by ETI.
 
 ## Study Overview
+
+-   Retrospective study of children with CF included individuals ages 4-21 followed at Children's Hospital Colorado (CHCO) from January 1982 to July 2021.
+
+-   Criteria were a diagnosis of CF based on sweat chloride $\geq$ 60mmol/L or two CFTR gene mutations.
+
+## First Analysis
 

### 2022-05-19 StatNet/knot_models.qmd
Update knot_models.qmd
random slope

@@
 
+## Pre-/Post-CFTR Modulator Results With Random Slope
+
+```{r}
+mod = lmer(Vitamin.D ~ Time + c1 + c2 + (Time|Patient.ID),data = cf)
+plot_df = mod@frame
+plot_df$pred = predict(mod)
+ggplot(plot_df,aes(x = Time,y = Vitamin.D,group = factor(Patient.ID))) +
+    geom_line(aes(y = pred),alpha = 0.5) + geom_point(aes(y = pred),alpha = 0.5) +
+    theme_bw() + theme(legend.position = "none") + xlab("Time (years)")
+```
+
+## Pre-/Post-CFTR Modulator Results With Random Slope
+
+```{r}
+kable(broom.mixed::tidy(mod,"fixed")[,-1],digits = 3)
+```
+
 ## A Bayesian Approach
@@
 
+## Future Plans
+
+-   Try a random slope
+-   Apply similar models to other outcomes like sweat electrolytes, etc.
+
 ## Limitations
@@
 
--   How to handle truncated data (e.g. a value of "\< 0.1")?
+-   Other ways handle truncated data (e.g. a value of "\< 0.1")?
+
+    -   Tobit extension to the existing model?
 
@@
 
+    -   How informative can we make these priors?
+
 # References

### 2022-05-19 StatNet/knot_models.qmd
Update knot_models.qmd

@@
 -   The `mcp` package provides an easy interface for fitting Bayesian regression models with multiple change points.
+
     -   Essentially it translates `lme4` syntax into JAGS.
+
 -   Uninformative "t-tail" priors are used unless the user specifies something else.
+
     -   "The default priors should be suitable for estimation in the absence of strong prior knowledge."[@lindelovMcpPackageRegression2020a]
+    -   "Based on data to ensure that they remain uninformative across any order of magnitude." [@lindelovMcpPackageRegression2020a]
+
+## A Bayesian Approach
+
+-   Intercept priors are t(0, 3\*sd(y), 3)
+
+-   Slope priors are t(0, sd(y)/(max(x) − min(x)), 3).
+
+-   The prior $\sigma$ is N(0, sd(y)).
+
+-   U(−1, 1) prior is used on autoregressive coefficients.
 
@@
 
--   `mcp` can model joined slopes with varying change points:
+-   `mcp` can model joined slopes with varying change points.
+-   Package documentation recommends Dirichlet priors, but they sample slowly.
+    -   A multivariate generalization of the beta distribution, indexed by $\alpha$ ($\alpha$ = 1 is flat).
+    -   The conjugate prior of the categorical distribution and multinomial distribution.
+
+## A Bayesian Model for Pre-/Post-CFTR Modulator
 
@@
   1 + (1|Patient.ID) ~ 0 + Time
+)
+priors = list(
+  cp_1 = "dirichlet(1)"
 )
@@
 
--   This looks terrible!
+-   This looks terrible! `mcp` does seem to find a change point at time 0, but the change point at -3 doesn't make much sense.
 
@@
 
--   Try a random slope
 -   Apply similar models to other outcomes like sweat electrolytes, etc.
+-   Look into frequentist approaches for automatic detection (`segmented` package, etc.)
 
@@
     -   Large change in slope after second modulator start is likely driven by a few influential observations.
+    -   Posteriors may shrink towards priors too much if the data set is not large enough.
+        -   The likelihood does not overwhelm the prior.
+
+## Limitations
 

### 2022-07-16 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
vitamins

--- 2022-04-21 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
@@
 # Vitamins
-vit = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report Vitamins_throughDecember2021.xlsx")
-vit = vit %>% select(`Patient ID`,Date,VitaminA_Retinol:`Vitamin E Gamma`)
-# New age at start
+vit = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report Vitamins_throughDecember2021_revised20220712.xlsx")
+# Determine which modulator they were on
 mods = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report CFTR Modulators_throughDecember2021.xlsx")
@@
 # Merge
+df = left_join()
+# Merge
 df = full_join(vit,fev,by = c("Patient ID", "Date")) %>% filter(`Patient ID` %in% mods$`Patient ID`)

### 2022-08-08 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
Update cf_nutrition.Rmd

@@
 mods = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report CFTR Modulators_throughDecember2021.xlsx")
-mods = mods[mods$Modulator %in% c('Kalydeco','Trikafta'),]
-mods = mods %>% select(`Patient ID`,Modulator,`Earliest Date in Registry`) %>% 
+mods = mods %>% filter(Modulator %in% c('Kalydeco','Trikafta')) %>%
+  select(`Patient ID`,Modulator,`Earliest Date in Registry`) %>% 
   pivot_wider(id_cols = `Patient ID`,names_from = Modulator,
@@
 # Merge
-df = left_join()
-# Merge
 df = full_join(vit,fev,by = c("Patient ID", "Date")) %>% filter(`Patient ID` %in% mods$`Patient ID`)
@@
 df = left_join(df,pancreatic,by = "Patient ID")
-# Time relative to first mod
-df$Time = as.numeric(difftime(df$Date,df$`First Mod Date`,units = "days"))/365.25
-# Calculate spline things
-df$c1 = (df$Time-0)*(df$Time>=0)
-df$c2 = as.numeric(difftime(df$Date,df$`Second Mod Date`,units = "days"))/365.25
-df$c2 = df$c2*(df$Date >= df$`Second Mod Date`)
-# Convert to numeric
-df$VitaminA_Retinol = suppressWarnings(as.numeric(df$VitaminA_Retinol))
-# Select columns
-df = df %>% filter(!is.na(VitaminA_Retinol)) %>%
-  select(`Patient ID`,`First Mod Date`,`Second Mod Date`,
-         Date,Time,c1,c2,VitaminA_Retinol) %>%
-  rename(`Vitamin A` = VitaminA_Retinol)
-# Write for presentation
-write.csv(df,file = "~/GitHub/CF/StatNet/data.csv",row.names = F,na = "")
 ```
@@
 groups = df %>% group_by(`Patient ID`) %>% filter(row_number()==1)
+# Filter any measures within 3 months of modulator start
+df = df %>% 
+  filter(Date < `First Mod Date` - days(90) | 
+           ((Date > `First Mod Date` + days(90)) & (Date < `Second Mod Date` - days(90))) |
+           (Date > `Second Mod Date` + days(90) | is.na(`Second Mod Date`)))
 # Two nutrition measures before first mod start
 two_measures = df %>% 
-  mutate(before_first = Date < `First Mod Date` & !is.na(VitaminA_Retinol)) %>% 
+  mutate(before_first = Date < `First Mod Date` & 
+           (!is.na(VitaminA_Retinol) | !is.na(`25OH-Vitamin D`) | !is.na(`Vitamin E Alpha`))) %>% 
   group_by(`Patient ID`,Group) %>% count(before_first) %>% 
@@
   mutate(after_first = (Date >= `First Mod Date` %m+% months(3)) & 
-           !is.na(VitaminA_Retinol)) %>%
+           (!is.na(VitaminA_Retinol) | !is.na(`25OH-Vitamin D`) | !is.na(`Vitamin E Alpha`))) %>%
   group_by(`Patient ID`,Group) %>% count(after_first) %>%
   filter(after_first == T,n >= 1)
-# Write merged df - filter to just Vitamin A values
-t = df %>% 
-  mutate(`Pre/Post First Mod` = ifelse(Date > `First Mod Date`,"Post","Pre")) %>%
-  filter(!is.na(VitaminA_Retinol)) 
-write.csv(t,file = "./Edith Zemanick/CF Nutrition/Data_Cleaned/merged_data.csv",
-          row.names = F,na="")
-write.csv(mods,file = "./Edith Zemanick/CF Nutrition/Data_Cleaned/modulators.csv",
-          row.names = F,na="")
+# Filter
+df = df %>% filter(`Patient ID` %in% two_measures$`Patient ID` & 
+                     `Patient ID` %in% after_measure$`Patient ID`)
+# Get most recent pre-/post-first modulator values
+# Time from first mod
+df$Time = as.numeric(difftime(df$Date,df$`First Mod Date`,units = "days"))
+# Before or after start
+df$Before_After = cut(df$Time,c(-Inf,0,Inf),labels = c("Before","After"),right = T)
+t = df %>% filter(abs(Time) <= (3*365)) %>% # Within 3 years of start
+  group_by(`Patient ID`,Before_After) %>% 
+  arrange(abs(Time),.by_group = T) %>% # Sort so that the the first non-NA value is the closest to modulator start
+  summarise(`Vitamin A` = first(na.omit(VitaminA_Retinol)),
+            `Vitamin D` = first(na.omit(`25OH-Vitamin D`)),
+            `Vitamin E Alpha` = first(na.omit(`Vitamin E Alpha`)),
+            `Vitamin E Gamma` = first(na.omit(`Vitamin E Gamma`)),
+            BMI = first(na.omit(BMI)),.groups = "drop")
 ```
@@
 
-<!-- # Participant Characteristics -->
+# Participant Characteristics
 
-<!-- For each person, we calculated the maximum amount of time from the earliest nutrition measure before modulator start to start date, and the amount of time from modulator start date to most recent nutrition measure.
+For each person, we calculated the maximum amount of time from the earliest nutrition measure before modulator start to start date, and the amount of time from modulator start date to most recent nutrition measure.
 The ranges of these values are presented below ("Years to start" and "Years From start," respectively).
--->
 
-<!-- ```{r results='asis'} -->
-<!-- df$Sex[df$Sex == ""] = NA -->
-<!-- demo = df %>% group_by(Patient.ID) %>% fill(Sex,.direction = "downup") %>% -->
-<!--   filter(row_number() == 1) -->
-<!-- t1 = tableby(first_mod ~ Age.at.Start..years. + Sex + genoRisk + -->
-<!--                vit_measures_after + vit_measures_before + -->
-<!--                notest(max_years_to_start,"min","max") + -->
-<!--                notest(max_years_from_start,"min","max"), -->
-<!--              data = demo, -->
-<!--              control = tableby.control(numeric.stats=c("Nmiss", "median", "q1q3"),)) -->
-<!-- summary(t1,test = F, -->
-<!--         labelTranslations = -->
-<!--           list(Age.at.Start..years. = "Age at Start (years)", -->
-<!--                genoRisk = "Genotype risk group", -->
-<!--                vit_measures_after = "No. of vitamin measures post", -->
-<!--                vit_measures_before = "No. of vitamin measures pre", -->
-<!--                max_years_to_start = "Years to start",max_years_from_start = "Years From start")) -->
-<!-- ``` -->
+```{r results='asis'}
+# Import demo info and add
+
+df$Sex[df$Sex == ""] = NA
+demo = df %>% group_by(Patient.ID) %>% fill(Sex,.direction = "downup") %>%
+  filter(row_number() == 1)
+t1 = tableby(first_mod ~ Age.at.Start..years. + Sex + genoRisk +
+               vit_measures_after + vit_measures_before +
+               notest(max_years_to_start,"min","max") +
+               notest(max_years_from_start,"min","max"),
+             data = demo,
+             control = tableby.control(numeric.stats=c("Nmiss", "median", "q1q3"),))
+summary(t1,test = F,
+        labelTranslations =
+          list(Age.at.Start..years. = "Age at Start (years)",
+               genoRisk = "Genotype risk group",
+               vit_measures_after = "No. of vitamin measures post",
+               vit_measures_before = "No. of vitamin measures pre",
+               max_years_to_start = "Years to start",max_years_from_start = "Years From start"))
+```
 

### 2022-08-08 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
Update cf_nutrition.Rmd

@@
 
-```{r}
+```{r warning=FALSE}
 # Vitamins
 vit = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report Vitamins_throughDecember2021_revised20220712.xlsx")
+vit$VitaminA_Retinol = as.numeric(vit$VitaminA_Retinol)
 # Determine which modulator they were on
@@
 mods$`Second Mod Date` = pmax(mods$`Trikafta Date`,mods$`Kalydeco Date`)
+mods$`First Mod` = NA
+mods$`First Mod`[mods$`First Mod Date` == mods$`Trikafta Date`] = "Trikafta"
+mods$`First Mod`[mods$`First Mod Date` == mods$`Kalydeco Date`] = "Kalydeco"
 # Add pancreatic status
@@
 fev = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report_PFTs_ThroughDecember2021.xlsx")
-fev = fev %>% select(PatientID,Date,BMI,`FEV1 % pred`) %>% rename(`Patient ID` = PatientID)
+fev = fev %>% select(PatientID,Sex,Date,BMI,`FEV1 % pred`) %>% rename(`Patient ID` = PatientID)
 # Merge
@@
 df$Before_After = cut(df$Time,c(-Inf,0,Inf),labels = c("Before","After"),right = T)
-t = df %>% filter(abs(Time) <= (3*365)) %>% # Within 3 years of start
-  group_by(`Patient ID`,Before_After) %>% 
+before_after = df %>% filter(abs(Time) <= (3*365)) %>% # Within 3 years of start
+  group_by(`Patient ID`,Before_After) %>% fill(Sex) %>%
   arrange(abs(Time),.by_group = T) %>% # Sort so that the the first non-NA value is the closest to modulator start
-  summarise(`Vitamin A` = first(na.omit(VitaminA_Retinol)),
+  summarise(Sex = first(na.omit(Sex)),Age = first(na.omit(`Age at Test (years)`)),
+            `First Mod` = first(na.omit(`First Mod`)),
+            `Vitamin A` = first(na.omit(VitaminA_Retinol)),
             `Vitamin D` = first(na.omit(`25OH-Vitamin D`)),
@@
             `Vitamin E Gamma` = first(na.omit(`Vitamin E Gamma`)),
-            BMI = first(na.omit(BMI)),.groups = "drop")
+            BMI = first(na.omit(BMI)),
+            `FEV1 % pred` = first(na.omit(`FEV1 % pred`)),
+            .groups = "drop")
 ```
 
-```{r}
+```{r eval=FALSE}
 # Flowchart
@@
 [6]: paste0('Treated With Ivacaftor followed by Trikafta From 2008-2021 (n = ',as.numeric(table(groups$Group)['Both']), ')')
-[7]: paste0('At Least One Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Kalydeco']), ')')
-[8]: paste0('At Least One Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Trikafta']), ')')
-[9]: paste0('At Least One Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Both']), ')')
-[10]: paste0('At Least One Vitamin A >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Kalydeco']), ')')
-[11]: paste0('At Least One Vitamin A >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Trikafta']), ')')
-[12]: paste0('At Least One Vitamin A >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Both']), ')')
+[7]: paste0('At Least One Vitamin Measure Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Kalydeco']), ')')
+[8]: paste0('At Least One Vitamin Measure Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Trikafta']), ')')
+[9]: paste0('At Least One Vitamin Measure Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Both']), ')')
+[10]: paste0('At Least One Vitamin Measure >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Kalydeco']), ')')
+[11]: paste0('At Least One Vitamin Measure >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Trikafta']), ')')
+[12]: paste0('At Least One Vitamin Measure >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Both']), ')')
 [13]: paste0('Final (n = ',length(unique(after_measure$`Patient ID`)), ')')
@@
 
-For each person, we calculated the maximum amount of time from the earliest nutrition measure before modulator start to start date, and the amount of time from modulator start date to most recent nutrition measure.
-The ranges of these values are presented below ("Years to start" and "Years From start," respectively).
+```{r results='asis'}
+# Import demo info and add
+demo = before_after %>% group_by(`Patient ID`) %>% filter(row_number() == 1)
+t1 = tableby(`First Mod` ~ Age + Sex + `Vitamin A` + `Vitamin D` + 
+               `Vitamin E Alpha` + `Vitamin E Gamma` + BMI + `FEV1 % pred`,
+             data = demo,
+             control = tableby.control(numeric.stats=c("Nmiss", "median", "q1q3")))
+summary(t1)
+```
+
+# Pre-Post Copmparisons
+
+Continuous variables were compared using a Wilcoxon signed rank test and categorical variables were compared using McNemar’s test.
 
 ```{r results='asis'}
-# Import demo info and add
-
-df$Sex[df$Sex == ""] = NA
-demo = df %>% group_by(Patient.ID) %>% fill(Sex,.direction = "downup") %>%
-  filter(row_number() == 1)
-t1 = tableby(first_mod ~ Age.at.Start..years. + Sex + genoRisk +
-               vit_measures_after + vit_measures_before +
-               notest(max_years_to_start,"min","max") +
-               notest(max_years_from_start,"min","max"),
-             data = demo,
-             control = tableby.control(numeric.stats=c("Nmiss", "median", "q1q3"),))
-summary(t1,test = F,
-        labelTranslations =
-          list(Age.at.Start..years. = "Age at Start (years)",
-               genoRisk = "Genotype risk group",
-               vit_measures_after = "No. of vitamin measures post",
-               vit_measures_before = "No. of vitamin measures pre",
-               max_years_to_start = "Years to start",max_years_from_start = "Years From start"))
+t2 = paired(Before_After ~ `Vitamin A` + `Vitamin D` + 
+               `Vitamin E Alpha` + `Vitamin E Gamma` + BMI + `FEV1 % pred`,
+             data = before_after, id = `Patient ID`, 
+            numeric.test = "signed.rank",signed.rank.exact = FALSE)
+summary(t2)
 ```

### 2022-08-08 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
Update cf_nutrition.Rmd

@@
 
-# Data cleaning
-
 ```{r warning=FALSE}
@@
 fev = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report_PFTs_ThroughDecember2021.xlsx")
-fev = fev %>% select(PatientID,Sex,Date,BMI,`FEV1 % pred`) %>% rename(`Patient ID` = PatientID)
+fev = fev %>% select(PatientID,Sex,Date,`BMI percentile (CDC)`,`FEV1 % pred`) %>% rename(`Patient ID` = PatientID)
 # Merge
@@
             `Vitamin E Gamma` = first(na.omit(`Vitamin E Gamma`)),
-            BMI = first(na.omit(BMI)),
+            BMI = first(na.omit(`BMI percentile (CDC)`)),
             `FEV1 % pred` = first(na.omit(`FEV1 % pred`)),
@@
 
-# Pre-Post Copmparisons
+# Pre-Post Comparisons
 
@@
              data = before_after, id = `Patient ID`, 
-            numeric.test = "signed.rank",signed.rank.exact = FALSE)
+            numeric.test = "signed.rank",signed.rank.exact = FALSE,
+            control = paired.control(numeric.stats=c("Nmiss", "median", "q1q3")))
 summary(t2)

### 2022-08-21 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
remove participant flowchart

@@
 df = left_join(df,pancreatic,by = "Patient ID")
-```
-
-```{r}
-# Counts from Elin
-# patients with CF in our clinical database from 2008- December 2021: 740 total, 474 in our dataset
-ncdb = 740
-# of patients treated with ivacaftor from 2008-2021: 58 total, 47 in dataset
-# of patients treated with trikafta (ETI) from 2008-2021: 228 total, 150 in dataset 
-# of patients treated with ivacaftor followed by ETI – these patients may be counted in both groups: 22 total, 17 in dataset
-# Tim counts
-df$Group = apply(df,1,function(r){
-  if(!is.na(r["Kalydeco Date"]) & is.na(r["Trikafta Date"])){
-    g = "Kalydeco"
-  } else if (is.na(r["Kalydeco Date"]) & !is.na(r["Trikafta Date"])){
-    g = "Trikafta"
-  } else if (!is.na(r["Kalydeco Date"]) & !is.na(r["Trikafta Date"])){
-    g = "Both"
-  }
-  return(g)
-})
-# "Our dataset"
-sweat = read_excel("./Edith Zemanick/CF Nutrition/Data_Raw/Report Electrolytes_throughDecember2021.xlsx")
-df = df %>% filter(`Patient ID` %in% sweat$`Patient ID`)
-groups = df %>% group_by(`Patient ID`) %>% filter(row_number()==1)
-# Filter any measures within 3 months of modulator start
-df = df %>% 
-  filter(Date < `First Mod Date` - days(90) | 
-           ((Date > `First Mod Date` + days(90)) & (Date < `Second Mod Date` - days(90))) |
-           (Date > `Second Mod Date` + days(90) | is.na(`Second Mod Date`)))
-# Two nutrition measures before first mod start
-two_measures = df %>% 
-  mutate(before_first = Date < `First Mod Date` & 
-           (!is.na(VitaminA_Retinol) | !is.na(`25OH-Vitamin D`) | !is.na(`Vitamin E Alpha`))) %>% 
-  group_by(`Patient ID`,Group) %>% count(before_first) %>% 
-  filter(before_first == T,n >= 1)
-# At least one nutritional measure and one weight/height measure >= 3 months after first
-after_measure = df %>% 
-  mutate(after_first = (Date >= `First Mod Date` %m+% months(3)) & 
-           (!is.na(VitaminA_Retinol) | !is.na(`25OH-Vitamin D`) | !is.na(`Vitamin E Alpha`))) %>%
-  group_by(`Patient ID`,Group) %>% count(after_first) %>%
-  filter(after_first == T,n >= 1)
-# Filter
-df = df %>% filter(`Patient ID` %in% two_measures$`Patient ID` & 
-                     `Patient ID` %in% after_measure$`Patient ID`)
-# Get most recent pre-/post-first modulator values
-# Time from first mod
-df$Time = as.numeric(difftime(df$Date,df$`First Mod Date`,units = "days"))
-# Before or after start
-df$Before_After = cut(df$Time,c(-Inf,0,Inf),labels = c("Before","After"),right = T)
-before_after = df %>% filter(abs(Time) <= (3*365)) %>% # Within 3 years of start
-  group_by(`Patient ID`,Before_After) %>% fill(Sex) %>%
-  arrange(abs(Time),.by_group = T) %>% # Sort so that the the first non-NA value is the closest to modulator start
-  summarise(Sex = first(na.omit(Sex)),Age = first(na.omit(`Age at Test (years)`)),
-            `First Mod` = first(na.omit(`First Mod`)),
-            `Vitamin A` = first(na.omit(VitaminA_Retinol)),
-            `Vitamin D` = first(na.omit(`25OH-Vitamin D`)),
-            `Vitamin E Alpha` = first(na.omit(`Vitamin E Alpha`)),
-            `Vitamin E Gamma` = first(na.omit(`Vitamin E Gamma`)),
-            BMI = first(na.omit(`BMI percentile (CDC)`)),
-            `FEV1 % pred` = first(na.omit(`FEV1 % pred`)),
-            .groups = "drop")
-```
-
-```{r eval=FALSE}
-# Flowchart
-g = grViz("
-digraph graph2 {
-graph [layout = dot]
-# node definitions with substituted label text
-node [shape = rectangle]
-a [label = '@@1']
-b [label = '@@2']
-c [label = '@@3']
-d [label = '@@4']
-e [label = '@@5']
-f [label = '@@6']
-g [label = '@@7']
-h [label = '@@8']
-i [label = '@@9']
-j [label = '@@10']
-k [label = '@@11']
-l [label = '@@12']
-m [label = '@@13']
-
-a -> b
-b -> c
-c -> d
-c -> e
-c -> f
-d -> g
-e -> h
-f -> i
-g -> j
-h -> k
-i -> l
-j -> m
-k -> m
-l -> m
-}
-
-[1]: paste0('Clinical Database (n = ', ncdb, ')')
-[2]: paste0('Treated With Ivacaftor or Trikafta From 2008-2021 (n = ',length(unique(mods$`Patient ID`)), ')')
-[3]: paste0('In Our Dataset (n = ',length(intersect(mods$`Patient ID`,sweat$`Patient ID`)), ')')
-[4]: paste0('Treated With Ivacaftor Alone From 2008-2021 (n = ',as.numeric(table(groups$Group)['Kalydeco']), ')')
-[5]: paste0('Treated With Trikafta Alone From 2008-2021 (n = ',as.numeric(table(groups$Group)['Trikafta']), ')')
-[6]: paste0('Treated With Ivacaftor followed by Trikafta From 2008-2021 (n = ',as.numeric(table(groups$Group)['Both']), ')')
-[7]: paste0('At Least One Vitamin Measure Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Kalydeco']), ')')
-[8]: paste0('At Least One Vitamin Measure Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Trikafta']), ')')
-[9]: paste0('At Least One Vitamin Measure Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Both']), ')')
-[10]: paste0('At Least One Vitamin Measure >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Kalydeco']), ')')
-[11]: paste0('At Least One Vitamin Measure >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Trikafta']), ')')
-[12]: paste0('At Least One Vitamin Measure >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Both']), ')')
-[13]: paste0('Final (n = ',length(unique(after_measure$`Patient ID`)), ')')
-")
-g
 ```

### 2022-08-24 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
Update pre_post_trikafta_prospective_analysis.Rmd
minor figure changes

@@
 knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
-if(Sys.info()["sysname"] == "Windows"){
-  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
-} else if (Sys.info()["sysname"] == "Linux"){
-  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
-} else if (Sys.info()["sysname"] == "Darwin"){
-  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
-}
-knitr::opts_knit$set(root.dir = home_dir)
-rm(home_dir)
+knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/CF")
 ```
@@
 newlabels["homa2IR"] = "HOMA2 IR"
+newlabels["total_sensor_readings"] = "Total Sensor Readings"
 newlabels["mean_sensor"] = "Mean SG (mg/dL)"
@@
              data = df[df$Event.Name == "Visit 1",])
-summary(t1,digits = 2,labelTranslations = newlabels)
+summary(t1,digits = 2,labelTranslations = newlabels,pfootnote=TRUE)
 ```
@@
              data = df[df$Event.Name == "Visit 1" & df$Study.ID %in% keep_dxa,])
-summary(t1,digits = 2,labelTranslations = newlabels)
+summary(t1,digits = 2,labelTranslations = newlabels,pfootnote=TRUE)
 ```
@@
             "BMD.Total.z.score","BMD.Subtotal.z.score",
-            "DXA.FMI.calculated.z.score","DXA.LBMI.calculated.z.score",
+            "DXA.FMI.calculated.z.score","DXA.LBMI.calculated.z.score","total_sensor_readings",
             "mean_sensor","max_sensor","min_sensor","percent_time_70_180",
@@
 
+Line plots below indicate median with IQR represented by the error bars.
+
 ## HbA1c
@@
 ggplot(df,aes(x = Event.Name, y = A1C.result,fill = Event.Name)) + 
-  geom_boxplot() + geom_jitter(width = 0.25) + 
-  xlab("Visit") + ylab("HbA1c (%)") + theme_bw() + theme(legend.position = "none")
+  geom_boxplot(coef=0,alpha = 0.5,outlier.shape = NA) + geom_jitter(width = 0.15) + 
+  xlab("Visit") + ylab("HbA1c (%)") + 
+  theme_bw() + theme(legend.position = "none") +
+  scale_fill_grey()
 ```
@@
 plot_df = plot_df %>% group_by(name,Event.Name) %>% 
-  summarise(m = mean(value,na.rm = T),
-            se = sd(value)/sqrt(n()),.groups = "drop")
+  summarise(m = median(value,na.rm = T),
+            low = as.numeric(quantile(value,0.25)),
+            high = as.numeric(quantile(value,0.75)),
+            .groups = "drop")
 ggplot(plot_df,aes(x = name,y = m,color = Event.Name)) + 
   geom_line() + geom_point() + 
-  geom_errorbar(aes(ymin=m-(1.96*se), ymax=m+(1.96*se)), width=2) +
+  geom_errorbar(aes(ymin=low, ymax=high), width=2) +
   xlab("Time (minutes)") + ylab("Glucose (mg/dL)") + ylim(0,200) + 
-  theme_bw()
+  theme_bw() +
+  theme(legend.title=element_blank()) + scale_color_grey()
 ```
@@
 plot_df = plot_df %>% group_by(name,Event.Name) %>% 
-  summarise(m = mean(value,na.rm = T),
-            se = sd(value)/sqrt(n()),.groups = "drop")
+  summarise(m = median(value,na.rm = T),
+            low = as.numeric(quantile(value,0.25)),
+            high = as.numeric(quantile(value,0.75)),
+            .groups = "drop")
 ggplot(plot_df,aes(x = name,y = m,color = Event.Name)) + 
   geom_line() + geom_point() + 
-  geom_errorbar(aes(ymin=m-(1.96*se), ymax=m+(1.96*se)), width=2) +
+  geom_errorbar(aes(ymin=low, ymax=high), width=2) +
   xlab("Time (minutes)") + ylab("C Peptide (pmol/L)") + 
-  theme_bw()
+  theme_bw() +
+  theme(legend.title=element_blank()) + scale_color_grey()
 ```
@@
   summarise(m = mean(value,na.rm = T),
-            se = sd(value)/sqrt(n()),.groups = "drop")
+            low = as.numeric(quantile(value,0.25)),
+            high = as.numeric(quantile(value,0.75)),
+            .groups = "drop")
 ggplot(plot_df,aes(x = name,y = m,color = Event.Name)) + 
   geom_line() + geom_point() + 
-  geom_errorbar(aes(ymin=m-(1.96*se), ymax=m+(1.96*se)), width=2) +
+  geom_errorbar(aes(ymin=low, ymax=high), width=2) +
   xlab("Time (minutes)") + ylab("Insulin (mcU/mL)") + 
-  theme_bw()
+  theme_bw() + 
+  theme(legend.title=element_blank()) + scale_color_grey()
 ```

### 2022-08-25 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
Update pre_post_trikafta_prospective_analysis.Rmd
group changes adult vs. children

@@
 })
-df$Group = factor(df$Group,levels = c("NGT","AGT","CFRD"),ordered = T)
+df$Group = factor(df$Group,levels = c("NGT","AGT","CFRD"))
 # convert Insulin to mcU/mL
@@
 f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
-p <- paired(f,data = df, id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F,
-            control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),)
+p <- paired(f,data = df, id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F,cat.test = "mcnemar",
+            control = tableby.control(numeric.stats = c("Nmiss", "medianq1q3", "range")),)
 summary(p,pfootnote = T,labelTranslations = newlabels)
@@
 p = paired(f,data = df[df$Study.ID %in% keep_dxa,], 
-           control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),
+           control = tableby.control(numeric.stats = c("Nmiss", "medianq1q3", "range")),
            id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F)
@@
 df$`Age Group` = cut(df$Age.at.visit.,c(-Inf,18,Inf),labels = c("Child","Adult"))
+df = df %>% arrange(Study.ID,Event.Name) %>% group_by(Study.ID) %>% fill(`Age Group`,.direction = "down")
 ```
@@
 f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
-p <- paired(f,data = df, id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F,
-            control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),
+p <- paired(f,data = df, id = Study.ID,numeric.test = "signed.rank",cat.test = "mcnemar",signed.rank.exact = F,
+            control = tableby.control(numeric.stats = c("Nmiss", "medianq1q3", "range")),
             strata = `Age Group`)
@@
 
+# Group change tests
+
+## All
+
+```{r}
+groups = df %>% select(Study.ID,Event.Name,Group) %>%
+  pivot_wider(id_cols = Study.ID,names_from = Event.Name,values_from = Group) %>%
+  left_join(.,df[df$Event.Name == "Visit 1",c("Study.ID","Age Group")],by = "Study.ID")
+groups$Change = factor(as.numeric(groups$`Visit 2`) - as.numeric(groups$`Visit 1`),
+                       levels = c(-1:1),labels = c("Improved","Stayed Same or Got Worse",
+                                                   "Stayed Same or Got Worse"))
+p = prop.test(sum(groups$Change == "Improved"),n = nrow(groups))
+kable(table(groups$Change),col.names = c("Change","n"))
+```
+
+`r as.numeric(p$estimate*100)`% of all participants improved from visit 1 to visit 2 (95% CI: `r round(p$conf.int[1],3)*100` - `r round(p$conf.int[2],3)*100`).
+
+## Adults vs. Children
+
+```{r results='asis'}
+t = tableby(`Age Group`~Change,groups,cat.test="fe")
+summary(t,pfootnote = T)
+```
+
 # Notes

### 2022-09-06 Christine Chan/CF Registry/cf_registry_analysis.Rmd
really need to do this in R so it's fast for me

---
title: "CF Registry Analysis"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Documents/Work/CF/Christine Chan/CF Registry")
```

```{r}
# Read in CSVs
df = read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv")
annual = read.csv("./Data_Raw/CFF19_Annualized_Del1.csv")
colnames(annual)[colnames(annual)=="ReviewYear"] = "reviewyear"
# # Merge
# df = df.join(annual, on=['eDWID', 'reviewyear'],how="left")
# # Group years by 2003-2009, 2010-2013, and 2014-2018
# df["Period"] = pd.cut(df["reviewyear"], [2003, 2009, 2013, 2018])
# # Small dataframe for checking
# test = df.iloc[0:1000,:]
# test.to_csv("~/test.csv",index=False)
```

### 2022-09-06 Christine Chan/CF Registry/cf_registry_analysis.Rmd
Update cf_registry_analysis.Rmd

@@
 colnames(annual)[colnames(annual)=="ReviewYear"] = "reviewyear"
-# # Merge
-# df = df.join(annual, on=['eDWID', 'reviewyear'],how="left")
-# # Group years by 2003-2009, 2010-2013, and 2014-2018
-# df["Period"] = pd.cut(df["reviewyear"], [2003, 2009, 2013, 2018])
-# # Small dataframe for checking
-# test = df.iloc[0:1000,:]
-# test.to_csv("~/test.csv",index=False)
+demo = read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
+# Combine and format demographics columns 
+demo$Race = apply(demo,1,function(r){
+  race = r[paste0("Race",1:6)]
+  race = as.numeric(which(race == 1))
+  race = paste(race,collapse = ", ")
+})
+demo$Race[grep(", ",demo$Race)] = "More Than One"
+demo$Race = factor(demo$Race,levels = c(1:6,"More Than One"),
+                   labels = c("White","Black or African American",
+                              "American Indian or Alaska Native","Asian",
+                              "Native Hawaiian or Other Pacific Islander",
+                              "Some other race","More Than One"))
+demo$Hispanicrace = factor(demo$Hispanicrace,levels = 1:2,
+                           labels = c("Yes","No"))
+# Merge
+df = left_join(df,annual, by=c('eDWID', 'reviewyear'))
+df = left_join(df,demo,by = "eDWID")
+# Group years by 2003-2009, 2010-2013, and 2014-2018
+df$Period = cut(df$reviewyear,
+                breaks =c(2002, 2009, 2013, 2018),
+                labels = c("2003 - 2009","2010 - 2013","2014 - 2018"))
+df$smoking = factor(df$smoking,levels = 1:4,labels = c("No",
+                                                       "Occasionally",
+                                                       "Yes, Regularly, less than 1 ppd",
+                                                       "Yes, Regularly, 1 ppd or more"),ordered = T)
+df$second_smoke = factor(df$smoking,levels = 1:4,labels = c("No",
+                                                            "Occasionally",
+                                                            "Yes, Regularly, less than 1 ppd",
+                                                            "Yes, Regularly, 1 ppd or more"),ordered = T)
+# Factors
+ed_levels = c("Less than High School",
+              "High School diploma or equivalent",
+              "Some College",
+              "College Graduate",
+              "Masters/Doctoral level degree")
+df$patient_education = factor(df$patient_education,levels = c(1:5),
+                              labels = ed_levels)
+df$father_education = factor(df$father_education,levels = c(1:5),
+                             labels = ed_levels)
+df$mother_education = factor(df$mother_education,levels = c(1:5),
+                             labels = ed_levels)
+df$spouse_education = factor(df$spouse_education,levels = c(1:5),
+                             labels = ed_levels)
+df$pregnant = factor(df$pregnant,levels = 0:1,labels = c("No","Yes"),ordered = T)
+# Combine education checkboxes
+df = df[1:1000,]
+df$employment = apply(df,1,function(r){
+  emp = r[paste0("employment",1:7)]
+  emp = as.numeric(emp)
+  if(all(is.na(emp))){
+    return(NA)
+  } else {
+    emp = as.numeric(emp)
+    emp = which(emp == 1)
+    emp = paste(emp,collapse = ", ")
+    return(emp)
+  }
+})
+df$employment = sub(sub("1","Part Time",df$employment))
+df$employment = sub(sub("2","Full time homemaker",df$employment))
+df$employment = sub(sub("3","Full time employment",df$employment))
+df$employment = sub(sub("4","Unemployed",df$employment))
+df$employment = sub(sub("5","Student",df$employment))
+df$employment = sub(sub("6","Disabled",df$employment))
+df$employment = sub(sub("7","Retired",df$employment))
+# Small dataframe for checking
+test = df[1:1000,]
+write.csv(test,"~/test.csv",row.names = F)
 ```
+
+# Table 1: Participant characteristics at last visit in time period
+
+```{r warning=FALSE}
+t1_df = df %>% group_by(eDWID,Period) %>% 
+  summarise(Age = max(Age_YrEnd,na.rm = T),Race = last(na.omit(Race)),
+            Hispanic = last(na.omit(Hispanicrace)),Gender = last(na.omit(Gender)),
+            `Patient Education` = last(na.omit(patient_education)),
+            `Father Education` = last(na.omit(father_education)),
+            `Mother Education` = last(na.omit(mother_education)),
+            `Spouse Education` = last(na.omit(spouse_education)),
+            `Marital Status` = last(na.omit(marital_status)),
+            Employment = last(na.omit(employment)),
+            Pregnancy = max(pregnant,na.rm = T)
+  )
+# Replace infinite values with missing
+t1_df = data.frame(lapply(t1_df, function(c){
+  c[is.infinite(c)] = NA
+  return(c)
+}))
+
+cols = colnames(t1_df)[3:ncol(t1_df)]
+t1 = tableby(as.formula(paste0("Period~",paste0("`",cols,"`",collapse = "+"))),
+             data = t1_df)
+```

### 2022-09-06 Christine Chan/CF Registry/cf_registry_analysis.Rmd
Update cf_registry_analysis.Rmd

--- 2022-09-06 Christine Chan/CF Registry/cf_registry_analysis.Rmd
@@
 
-```{r}
+```{r cache=TRUE}
 # Read in CSVs
@@
 colnames(annual)[colnames(annual)=="ReviewYear"] = "reviewyear"
-# # Merge
-# df = df.join(annual, on=['eDWID', 'reviewyear'],how="left")
-# # Group years by 2003-2009, 2010-2013, and 2014-2018
-# df["Period"] = pd.cut(df["reviewyear"], [2003, 2009, 2013, 2018])
-# # Small dataframe for checking
-# test = df.iloc[0:1000,:]
-# test.to_csv("~/test.csv",index=False)
+demo = read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
+# Combine and format demographics columns 
+demo$Race = apply(demo,1,function(r){
+  race = r[paste0("Race",1:6)]
+  race = as.numeric(which(race == 1))
+  race = paste(race,collapse = ", ")
+})
+demo$Race[grep(", ",demo$Race)] = "More Than One"
+demo$Race = factor(demo$Race,levels = c(1:6,"More Than One"),
+                   labels = c("White","Black or African American",
+                              "American Indian or Alaska Native","Asian",
+                              "Native Hawaiian or Other Pacific Islander",
+                              "Some other race","More Than One"))
+demo$Hispanicrace = factor(demo$Hispanicrace,levels = 1:2,
+                           labels = c("Yes","No"))
+# Merge
+df = left_join(df,annual, by=c('eDWID', 'reviewyear'))
+df = left_join(df,demo,by = "eDWID")
+# Group years by 2003-2009, 2010-2013, and 2014-2018
+df$Period = cut(df$reviewyear,
+                breaks =c(2002, 2009, 2013, 2018),
+                labels = c("2003 - 2009","2010 - 2013","2014 - 2018"))
+df$smoking = factor(df$smoking,levels = 1:4,labels = c("No",
+                                                       "Occasionally",
+                                                       "Yes, Regularly, less than 1 ppd",
+                                                       "Yes, Regularly, 1 ppd or more"),ordered = T)
+df$second_smoke = factor(df$smoking,levels = 1:4,labels = c("No",
+                                                            "Occasionally",
+                                                            "Yes, Regularly, less than 1 ppd",
+                                                            "Yes, Regularly, 1 ppd or more"),ordered = T)
+# Factors
+ed_levels = c("Less than High School",
+              "High School diploma or equivalent",
+              "Some College",
+              "College Graduate",
+              "Masters/Doctoral level degree")
+df$patient_education = factor(df$patient_education,levels = c(1:5),
+                              labels = ed_levels)
+df$father_education = factor(df$father_education,levels = c(1:5),
+                             labels = ed_levels)
+df$mother_education = factor(df$mother_education,levels = c(1:5),
+                             labels = ed_levels)
+df$spouse_education = factor(df$spouse_education,levels = c(1:5),
+                             labels = ed_levels)
+df$pregnant = factor(df$pregnant,levels = 0:1,labels = c("No","Yes"),ordered = T)
+# Combine education checkboxes
+cl = makeCluster(detectCores()*0.5,type = "FORK")
+df$employment = parApply(cl,df,1,function(r){
+  emp = r[paste0("employment",1:7)]
+  emp = as.numeric(emp)
+  if(all(is.na(emp) | emp == 0)){
+    return(NA)
+  } else {
+    emp = as.numeric(emp)
+    emp = which(emp == 1)
+    emp = paste(emp,collapse = ", ")
+    return(emp)
+  }
+})
+stopC
+df$employment = sub("1","Part Time",df$employment)
+df$employment = sub("2","Full time homemaker",df$employment)
+df$employment = sub("3","Full time employment",df$employment)
+df$employment = sub("4","Unemployed",df$employment)
+df$employment = sub("5","Student",df$employment)
+df$employment = sub("6","Disabled",df$employment)
+df$employment = sub("7","Retired",df$employment)
+df$employment = as.factor(df$employment)
+# Small dataframe for checking
+test = df[1:1000,]
+write.csv(test,"~/test.csv",row.names = F)
 ```
+
+# Table 1: Participant characteristics at last visit in time period
+
+```{r warning=FALSE}
+t1_df = df %>% group_by(eDWID,Period) %>% 
+  summarise(Age = max(Age_YrEnd[is.finite(Age_YrEnd)]),Race = last(na.omit(Race)),
+            Hispanic = last(na.omit(Hispanicrace)),Gender = last(na.omit(Gender)),
+            `Patient Education` = last(na.omit(patient_education)),
+            `Father Education` = last(na.omit(father_education)),
+            `Mother Education` = last(na.omit(mother_education)),
+            `Spouse Education` = last(na.omit(spouse_education)),
+            `Marital Status` = last(na.omit(marital_status)),
+            Employment = last(na.omit(employment)),
+            Pregnancy = max(pregnant,na.rm = T)
+  )
+# Replace infinite values with missing
+t1_df = data.frame(lapply(t1_df, function(c){
+  c[is.infinite(c)] = NA
+  return(c)
+}))
+
+cols = colnames(t1_df)[3:ncol(t1_df)]
+t1 = tableby(as.formula(paste0("Period~",paste0("`",cols,"`",collapse = "+"))),
+             data = t1_df)
+```

### 2022-09-07 Christine Chan/CF Registry/cf_registry_analysis.Rmd
Update cf_registry_analysis.Rmd

--- 2022-09-06 Christine Chan/CF Registry/cf_registry_analysis.Rmd
@@
 library(knitr)
+library(parallel)
 knitr::opts_knit$set(root.dir = "/Users/timvigers/Documents/Work/CF/Christine Chan/CF Registry")
@@
 colnames(annual)[colnames(annual)=="ReviewYear"] = "reviewyear"
-# # Merge
-# df = df.join(annual, on=['eDWID', 'reviewyear'],how="left")
-# # Group years by 2003-2009, 2010-2013, and 2014-2018
-# df["Period"] = pd.cut(df["reviewyear"], [2003, 2009, 2013, 2018])
-# # Small dataframe for checking
-# test = df.iloc[0:1000,:]
-# test.to_csv("~/test.csv",index=False)
+demo = read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
+# Combine and format demographics columns 
+demo$Race = apply(demo,1,function(r){
+  race = r[paste0("Race",1:6)]
+  race = as.numeric(which(race == 1))
+  race = paste(race,collapse = ", ")
+})
+demo$Race[grep(", ",demo$Race)] = "More Than One"
+demo$Race = factor(demo$Race,levels = c(1:6,"More Than One"),
+                   labels = c("White","Black or African American",
+                              "American Indian or Alaska Native","Asian",
+                              "Native Hawaiian or Other Pacific Islander",
+                              "Some other race","More Than One"))
+demo$Hispanicrace = factor(demo$Hispanicrace,levels = 1:2,
+                           labels = c("Yes","No"))
+# Merge
+df = left_join(df,annual, by=c('eDWID', 'reviewyear'))
+df = left_join(df,demo,by = "eDWID")
+# Group years by 2003-2009, 2010-2013, and 2014-2018
+df$Period = cut(df$reviewyear,
+                breaks =c(2002, 2009, 2013, 2018),
+                labels = c("2003 - 2009","2010 - 2013","2014 - 2018"))
+df$smoking = factor(df$smoking,levels = 1:4,labels = c("No",
+                                                       "Occasionally",
+                                                       "Yes, Regularly, less than 1 ppd",
+                                                       "Yes, Regularly, 1 ppd or more"),ordered = T)
+df$second_smoke = factor(df$smoking,levels = 1:4,labels = c("No",
+                                                            "Occasionally",
+                                                            "Yes, Regularly, less than 1 ppd",
+                                                            "Yes, Regularly, 1 ppd or more"),ordered = T)
+# Factors
+ed_levels = c("Less than High School",
+              "High School diploma or equivalent",
+              "Some College",
+              "College Graduate",
+              "Masters/Doctoral level degree")
+df$patient_education = factor(df$patient_education,levels = c(1:5),
+                              labels = ed_levels)
+df$father_education = factor(df$father_education,levels = c(1:5),
+                             labels = ed_levels)
+df$mother_education = factor(df$mother_education,levels = c(1:5),
+                             labels = ed_levels)
+df$spouse_education = factor(df$spouse_education,levels = c(1:5),
+                             labels = ed_levels)
+df$pregnant = factor(df$pregnant,levels = 0:1,labels = c("No","Yes"),ordered = T)
+# Combine education checkboxes
+cl = makeCluster(detectCores()*0.5,type = "FORK")
+df$employment = parApply(cl,df,1,function(r){
+  emp = r[paste0("employment",1:7)]
+  emp = as.numeric(emp)
+  if(all(is.na(emp) | emp == 0)){
+    return(NA)
+  } else {
+    emp = as.numeric(emp)
+    emp = which(emp == 1)
+    emp = paste(emp,collapse = ", ")
+    return(emp)
+  }
+})
+stopCluster(cl)
+df$employment = sub("1","Part Time",df$employment)
+df$employment = sub("2","Full time homemaker",df$employment)
+df$employment = sub("3","Full time employment",df$employment)
+df$employment = sub("4","Unemployed",df$employment)
+df$employment = sub("5","Student",df$employment)
+df$employment = sub("6","Disabled",df$employment)
+df$employment = sub("7","Retired",df$employment)
+df$employment = as.factor(df$employment)
+# Small dataframe for checking
+test = df[1:1000,]
+write.csv(test,"~/test.csv",row.names = F,na = "")
 ```
+
+# Table 1: Participant characteristics at last visit in time period
+
+```{r warning=FALSE,results='asis'}
+# Summarize by ID and period
+t1_df = df %>% group_by(eDWID,Period) %>% 
+  summarise(Age = max(Age_YrEnd),Race = last(na.omit(Race)),
+            Hispanic = last(na.omit(Hispanicrace)),Gender = last(na.omit(Gender)),
+            `Patient Education` = last(na.omit(patient_education)),
+            `Father Education` = last(na.omit(father_education)),
+            `Mother Education` = last(na.omit(mother_education)),
+            `Spouse Education` = last(na.omit(spouse_education)),
+            `Marital Status` = last(na.omit(marital_status)),
+            Employment = last(na.omit(employment)),
+            Pregnancy = max(pregnant,na.rm = T)
+  )
+# Replace infinite values with missing
+t1_df = data.frame(lapply(t1_df, function(c){
+  c[is.infinite(c)] = NA
+  return(c)
+}))
+# Print table 1
+cols = colnames(t1_df)[3:ncol(t1_df)]
+t1 = tableby(as.formula(paste0("Period~",paste0("`",cols,"`",collapse = "+"))),
+             data = t1_df)
+summary(t1)
+```

### 2022-09-07 Christine Chan/CF Registry/cf_registry_analysis.Rmd
Update cf_registry_analysis.Rmd
tidy formatting

--- 2022-09-06 Christine Chan/CF Registry/cf_registry_analysis.Rmd
@@
 library(knitr)
+library(parallel)
+knitr::opts_chunk$set(echo = FALSE)
 knitr::opts_knit$set(root.dir = "/Users/timvigers/Documents/Work/CF/Christine Chan/CF Registry")
@@
 # Read in CSVs
-df = read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv")
-annual = read.csv("./Data_Raw/CFF19_Annualized_Del1.csv")
-colnames(annual)[colnames(annual)=="ReviewYear"] = "reviewyear"
-demo = read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
+df <- read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv")
+annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv")
+colnames(annual)[colnames(annual) == "ReviewYear"] <- "reviewyear"
+demo <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
 # Combine and format demographics columns
-demo$Race = apply(demo,1,function(r){
-  race = r[paste0("Race",1:6)]
-  race = as.numeric(which(race == 1))
-  race = paste(race,collapse = ", ")
+demo$Race <- apply(demo, 1, function(r) {
+  race <- r[paste0("Race", 1:6)]
+  race <- as.numeric(which(race == 1))
+  race <- paste(race, collapse = ", ")
 })
-demo$Race[grep(", ",demo$Race)] = "More Than One"
-demo$Race = factor(demo$Race,levels = c(1:6,"More Than One"),
-                   labels = c("White","Black or African American",
-                              "American Indian or Alaska Native","Asian",
+demo$Race[grep(", ", demo$Race)] <- "More Than One"
+demo$Race <- factor(demo$Race,
+  levels = c(1:6, "More Than One"),
+  labels = c(
+    "White", "Black or African American",
+    "American Indian or Alaska Native", "Asian",
     "Native Hawaiian or Other Pacific Islander",
-                              "Some other race","More Than One"))
-demo$Hispanicrace = factor(demo$Hispanicrace,levels = 1:2,
-                           labels = c("Yes","No"))
+    "Some other race", "More Than One"
+  )
+)
+demo$Hispanicrace <- factor(demo$Hispanicrace,
+  levels = 1:2,
+  labels = c("Yes", "No")
+)
 # Merge
-df = left_join(df,annual, by=c('eDWID', 'reviewyear'))
-df = left_join(df,demo,by = "eDWID")
+df <- left_join(df, annual, by = c("eDWID", "reviewyear"))
+df <- left_join(df, demo, by = "eDWID")
 # Group years by 2003-2009, 2010-2013, and 2014-2018
-df$Period = cut(df$reviewyear,
-                breaks =c(2002, 2009, 2013, 2018),
-                labels = c("2003 - 2009","2010 - 2013","2014 - 2018"))
-df$smoking = factor(df$smoking,levels = 1:4,labels = c("No",
+df$Period <- cut(df$reviewyear,
+  breaks = c(2002, 2009, 2013, 2018),
+  labels = c("2003 - 2009", "2010 - 2013", "2014 - 2018")
+)
+df$smoking <- factor(df$smoking, levels = 1:4, labels = c(
+  "No",
   "Occasionally",
   "Yes, Regularly, less than 1 ppd",
-                                                       "Yes, Regularly, 1 ppd or more"),ordered = T)
-df$second_smoke = factor(df$smoking,levels = 1:4,labels = c("No",
+  "Yes, Regularly, 1 ppd or more"
+), ordered = T)
+df$second_smoke <- factor(df$smoking, levels = 1:4, labels = c(
+  "No",
   "Occasionally",
   "Yes, Regularly, less than 1 ppd",
-                                                            "Yes, Regularly, 1 ppd or more"),ordered = T)
+  "Yes, Regularly, 1 ppd or more"
+), ordered = T)
 # Factors
-ed_levels = c("Less than High School",
+ed_levels <- c(
+  "Less than High School",
   "High School diploma or equivalent",
@@
   "College Graduate",
-              "Masters/Doctoral level degree")
-df$patient_education = factor(df$patient_education,levels = c(1:5),
-                              labels = ed_levels)
-df$father_education = factor(df$father_education,levels = c(1:5),
-                             labels = ed_levels)
-df$mother_education = factor(df$mother_education,levels = c(1:5),
-                             labels = ed_levels)
-df$spouse_education = factor(df$spouse_education,levels = c(1:5),
-                             labels = ed_levels)
-df$pregnant = factor(df$pregnant,levels = 0:1,labels = c("No","Yes"),ordered = T)
+  "Masters/Doctoral level degree"
+)
+df$patient_education <- factor(df$patient_education,
+  levels = c(1:5),
+  labels = ed_levels
+)
+df$father_education <- factor(df$father_education,
+  levels = c(1:5),
+  labels = ed_levels
+)
+df$mother_education <- factor(df$mother_education,
+  levels = c(1:5),
+  labels = ed_levels
+)
+df$spouse_education <- factor(df$spouse_education,
+  levels = c(1:5),
+  labels = ed_levels
+)
+df$pregnant <- factor(df$pregnant, levels = 0:1, labels = c("No", "Yes"), ordered = T)
 # Combine education checkboxes
-df = df[1:1000,]
-df$employment = apply(df,1,function(r){
-  emp = r[paste0("employment",1:7)]
-  emp = as.numeric(emp)
-  if(all(is.na(emp))){
+cl <- makeCluster(detectCores() * 0.5, type = "FORK")
+df$employment <- parApply(cl, df, 1, function(r) {
+  emp <- r[paste0("employment", 1:7)]
+  emp <- as.numeric(emp)
+  if (all(is.na(emp) | emp == 0)) {
     return(NA)
   } else {
-    emp = as.numeric(emp)
-    emp = which(emp == 1)
-    emp = paste(emp,collapse = ", ")
+    emp <- as.numeric(emp)
+    emp <- which(emp == 1)
+    emp <- paste(emp, collapse = ", ")
     return(emp)
@@
 })
-df$employment = sub(sub("1","Part Time",df$employment))
-df$employment = sub(sub("2","Full time homemaker",df$employment))
-df$employment = sub(sub("3","Full time employment",df$employment))
-df$employment = sub(sub("4","Unemployed",df$employment))
-df$employment = sub(sub("5","Student",df$employment))
-df$employment = sub(sub("6","Disabled",df$employment))
-df$employment = sub(sub("7","Retired",df$employment))
+stopCluster(cl)
+df$employment <- sub("1", "Part Time", df$employment)
+df$employment <- sub("2", "Full time homemaker", df$employment)
+df$employment <- sub("3", "Full time employment", df$employment)
+df$employment <- sub("4", "Unemployed", df$employment)
+df$employment <- sub("5", "Student", df$employment)
+df$employment <- sub("6", "Disabled", df$employment)
+df$employment <- sub("7", "Retired", df$employment)
+df$employment <- as.factor(df$employment)
 # Small dataframe for checking
-test = df[1:1000,]
-write.csv(test,"~/test.csv",row.names = F)
+test <- df[1:1000, ]
+write.csv(test, "~/test.csv", row.names = F, na = "")
 ```
@@
 
-```{r warning=FALSE}
-t1_df = df %>% group_by(eDWID,Period) %>% 
-  summarise(Age = max(Age_YrEnd,na.rm = T),Race = last(na.omit(Race)),
-            Hispanic = last(na.omit(Hispanicrace)),Gender = last(na.omit(Gender)),
+```{r warning=FALSE,results='asis'}
+# Summarize by ID and period
+t1_df <- df %>%
+  group_by(eDWID, Period) %>%
+  summarise(
+    Age = max(Age_YrEnd), Race = last(na.omit(Race)),
+    Hispanic = last(na.omit(Hispanicrace)), Gender = last(na.omit(Gender)),
     `Patient Education` = last(na.omit(patient_education)),
@@
     Employment = last(na.omit(employment)),
-            Pregnancy = max(pregnant,na.rm = T)
+    Pregnancy = max(pregnant, na.rm = T)
   )
 # Replace infinite values with missing
-t1_df = data.frame(lapply(t1_df, function(c){
-  c[is.infinite(c)] = NA
+t1_df <- data.frame(lapply(t1_df, function(c) {
+  c[is.infinite(c)] <- NA
   return(c)
 }))
-
-cols = colnames(t1_df)[3:ncol(t1_df)]
-t1 = tableby(as.formula(paste0("Period~",paste0("`",cols,"`",collapse = "+"))),
-             data = t1_df)
+# Print table 1
+cols <- colnames(t1_df)[3:ncol(t1_df)]
+t1 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
+  data = t1_df
+)
+summary(t1)
 ```

### 2022-09-07 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd

--- 2022-03-01 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
@@
 ```{r setup, include=FALSE}
+library(readxl)
 library(arsenal)
@@
 library(knitr)
-knitr::opts_chunk$set(echo = FALSE)
-if(Sys.info()["sysname"] == "Windows"){
-  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Edith Zemanick/Sweat Test"
-} else if (Sys.info()["sysname"] == "Linux"){
-  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Edith Zemanick/Sweat Test"
-} else if (Sys.info()["sysname"] == "Darwin"){
-  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Edith Zemanick/Sweat Test"
-}
-knitr::opts_knit$set(root.dir = home_dir)
-rm(home_dir)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
+knitr::opts_knit$set(root.dir = "/Users/timvigers/Documents/Work/CF/Edith Zemanick/Sweat Test")
 source("~/GitHub/CF/get_genotype_class_severity.R")
@@
 ```{r}
+# Import exclusion data from Excel
+diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
+exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude)]
 # Import data
-df = read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv",na.strings = "")
-df$Diagnosis = factor(df$Diagnosis,
-                      levels = c("CFTR-related disorder","CRMS","Cystic Fibrosis","Non-CF","Normal"),
-                      labels = c("CFTR-related disorder","CRMS","Cystic Fibrosis","Non-CF","Non-CF"))
+df <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv", na.strings = "")
+# Remove exclusions from Angela
+df <- df[!(df$PatientID %in% exclude), ]
+df$Diagnosis <- diagnoses$Diagnosis[match(df$PatientID, diagnoses$PatientID)]
 # Filter per Maxie's code
-df = df %>% 
-  filter(Age.at.Test.in.Years < 50,
+df <- df %>%
+  filter(
+    Age.at.Test.in.Years < 50,
     !(Diagnosis == "Non-CF" & PatientID %in%
-             c(22,23,24,29,31,474,483,484,485,505,510,785,791,792,793))) %>% 
-  distinct(.) %>% filter(Sex != "",Demographics.Available == "TRUE") %>%
-  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks*7)
+      c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793))
+  ) %>%
+  distinct(.) %>%
+  filter(Sex != "", Demographics.Available == "TRUE") %>%
+  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks * 7)
 # Add modulator info
-cftr = read.csv("./Data_Raw/Copy of Report CFTR Modulators.csv")
-df$Modulator = apply(df,1,function(r){
-  mod = cftr[cftr$Patient.ID == as.numeric(r["PatientID"]),]
-  if (any(duplicated(mod$Age.at.Start..years.))){
-    mod = mod[-which(duplicated(mod$Age.at.Start..years.)),]
+cftr <- read.csv("./Data_Raw/Copy of Report CFTR Modulators.csv")
+df$Modulator <- apply(df, 1, function(r) {
+  mod <- cftr[cftr$Patient.ID == as.numeric(r["PatientID"]), ]
+  if (any(duplicated(mod$Age.at.Start..years.))) {
+    mod <- mod[-which(duplicated(mod$Age.at.Start..years.)), ]
   }
-  mod = mod[order(mod$Age.at.Start..years.),]
-  age = as.numeric(r["Age.at.Test.in.Years"])
-  as.character(cut(age,c(-Inf,mod$Age.at.Start..years.,Inf),
-                   labels = c("None",mod$Modulator),right = F))
+  mod <- mod[order(mod$Age.at.Start..years.), ]
+  age <- as.numeric(r["Age.at.Test.in.Years"])
+  as.character(cut(age, c(-Inf, mod$Age.at.Start..years., Inf),
+    labels = c("None", mod$Modulator), right = F
+  ))
 })
-df = left_join(df,cftr,by = c("PatientID" = "Patient.ID"))
 # Genotype info
-df$geno1 = get_cf_genotype_class_severity(df$Genotypes1)
-df$geno2 = get_cf_genotype_class_severity(df$Genotypes2)
-df$geno1[df$Diagnosis == "Non-CF"] = "Not CF"
-df$geno2[df$Diagnosis == "Non-CF"] = "Not CF"
-df$geno_risk = paste(df$geno1,df$geno2,sep = "_")
-df$geno_risk = factor(df$geno_risk,
-                      levels = c("Mild_Severe","Mild_Unknown","Not CF_Not CF","Severe_Mild","Mild_Mild",
-                                 "Severe_Severe","Severe_Unknown","Unknown_Severe","Unknown_Unknown"),
-                      labels = c("Low","Low","Not CF","Low","Low","High","Unknown","Unknown","Unknown"))
+df$geno1 <- get_cf_genotype_class_severity(df$Genotypes1)
+df$geno2 <- get_cf_genotype_class_severity(df$Genotypes2)
+df$geno1[df$Diagnosis == "Non-CF"] <- "Not CF"
+df$geno2[df$Diagnosis == "Non-CF"] <- "Not CF"
+df$geno_risk <- paste(df$geno1, df$geno2, sep = "_")
+df$geno_risk <- factor(df$geno_risk,
+  levels = c(
+    "Mild_Severe", "Mild_Unknown", "Not CF_Not CF",
+    "Severe_Mild", "Mild_Mild",
+    "Severe_Severe", "Severe_Unknown", "Unknown_Severe",
+    "Unknown_Unknown"
+  ),
+  labels = c(
+    "Low", "Low", "Not CF", "Low", "Low", "High", "Unknown",
+    "Unknown", "Unknown"
+  )
+)
+# Merge left and right values
+df$Amount <- apply(df[, c("Amount", "Amount_Left", "Amount_Right")], 1, function(r) {
+  amt <- suppressWarnings(as.numeric(as.character(r)))
+  mean(amt, na.rm = T)
+})
+df$Cl <- apply(df[, c("Cl", "Cl_Left", "Cl_Right")], 1, function(r) {
+  cl <- suppressWarnings(as.numeric(as.character(r)))
+  mean(cl, na.rm = T)
+})
+df$Na <- apply(df[, c("Na", "Na_Left", "Na_Right")], 1, function(r) {
+  na <- suppressWarnings(as.numeric(as.character(r)))
+  mean(na, na.rm = T)
+})
+df$K <- apply(df[, c("K", "K_Left", "K_Right")], 1, function(r) {
+  k <- suppressWarnings(as.numeric(as.character(r)))
+  mean(k, na.rm = T)
+})
+# Sort
+df <- df %>% arrange(PatientID, Age.at.Test.in.Weeks)
 ```
 
-# Table 1: Patient Demographics and Sample Characteristics for sweat tests done on individuals not on CFTR modulators
+# Table 1: Patient Demographics and Sample Characteristics
+
+For those with two sweat electrolyte measures (left and right), the average of the two was used.
 
 ```{r results='asis'}
-t1_df = df %>% filter(Age.at.Test.in.Days >= 2) %>%
-  group_by(PatientID) %>% arrange(Age.at.Test.in.Weeks) %>%  filter(row_number()==1)
-t1 = tableby(Method ~ Sex+Age.at.Test.in.Years+geno_risk,data = t1_df)
-summary(t1)
+t1_df <- df %>%
+  filter(Age.at.Test.in.Days >= 2) %>%
+  group_by(PatientID) %>%
+  arrange(Age.at.Test.in.Weeks) %>%
+  filter(row_number() == 1)
+t1 <- tableby(Diagnosis ~ Sex + Age.at.Test.in.Years + geno_risk + Amount + Cl + Na + K,
+  data = t1_df
+)
+summary(t1,
+  labelTranslations =
+    list(Age.at.Test.in.Years = "Age (Years)", geno_risk = "Genotype Severity")
+)
 ```
+
+# Chloride and Diagnosis
+
+## Table
+
+```{r}
+df$Cl_group <- cut(df$Cl, c(0, 30, 60, Inf), right = F)
+t <- table(df$Cl_group, df$Diagnosis)
+t <- rownames_to_column(as.data.frame.matrix(t))
+colnames(t)[1] <- "Chloride Range"
+kable(t, row.names = F)
+```
+
+## Plot
+
+```{r}
+ggplot(df, aes(x = Diagnosis, y = Cl)) +
+  geom_boxplot() +
+  theme_bw()
+```
+
+```{r}
+# Write only those with multiple values
+mult <- unique(df$PatientID[which(duplicated(df$PatientID))])
+multiples <- df[df$PatientID %in% mult, ]
+write.csv(multiples, file = "./Data_Cleaned/multiple_values.csv", na = "", row.names = F)
+```

### 2022-09-08 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd

@@
 library(knitr)
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
+knitr::opts_chunk$set(echo = FALSE)
 knitr::opts_knit$set(root.dir = "/Users/timvigers/Documents/Work/CF/Edith Zemanick/Sweat Test")
@@
 df <- df %>%
+  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks * 7) %>%
   filter(
-    Age.at.Test.in.Years < 50,
+    Age.at.Test.in.Years < 50 & Age.at.Test.in.Days >= 2,
     !(Diagnosis == "Non-CF" & PatientID %in%
@@
   distinct(.) %>%
-  filter(Sex != "", Demographics.Available == "TRUE") %>%
-  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks * 7)
+  filter(Sex != "", Demographics.Available == "TRUE")
 # Add modulator info
@@
 )
+# Which arm
+df$Arm <- apply(df[, c("Cl", "Cl_Left", "Cl_Right", "Na", "Na_Left", "Na_Right", "K", "K_Left", "K_Right")], 1, function(r) {
+  k <- sum(!is.na(r))
+  if (k == 6) {
+    return("Both")
+  } else if (k == 3) {
+    return("Arm Not Noted")
+  } else {
+    return(NA)
+  }
+})
 # Merge left and right values
@@
 df <- df %>% arrange(PatientID, Age.at.Test.in.Weeks)
+# Get list of those multiple values
+mult <- unique(df$PatientID[which(duplicated(df$PatientID))])
 ```
 
-# Table 1: Patient Demographics and Sample Characteristics
+# Patient Demographics and Sample Characteristics
 
@@
 
+## By group
+
 ```{r results='asis'}
 t1_df <- df %>%
-  filter(Age.at.Test.in.Days >= 2) %>%
   group_by(PatientID) %>%
   arrange(Age.at.Test.in.Weeks) %>%
-  filter(row_number() == 1)
+  filter(row_number() == 1) # First measure for everyone
 t1 <- tableby(Diagnosis ~ Sex + Age.at.Test.in.Years + geno_risk + Amount + Cl + Na + K,
@@
 
+## By Method
+
+### All participants
+
+```{r results='asis'}
+t1 <- tableby(Method ~ Sex + Age.at.Test.in.Years + Diagnosis + geno_risk + Amount + Cl + Na + K + Arm,
+  data = t1_df
+)
+summary(t1,
+  labelTranslations =
+    list(Age.at.Test.in.Years = "Age (Years)", geno_risk = "Genotype Severity")
+)
+```
+
+### Participants with more than one measure
+
+```{r results='asis'}
+t1 <- tableby(Method ~ Sex + Age.at.Test.in.Years + Diagnosis + geno_risk + Amount + Cl + Na + K + Arm,
+  data = t1_df[t1_df$PatientID %in% mult, ]
+)
+summary(t1,
+  labelTranslations =
+    list(Age.at.Test.in.Years = "Age (Years)", geno_risk = "Genotype Severity")
+)
+```
+
 # Chloride and Diagnosis
-
-## Table
 
@@
 
-## Plot
+# Plots
 
-```{r}
+```{r warning=FALSE}
+ggplot(df, aes(x = Diagnosis, y = Na)) +
+  geom_boxplot() +
+  theme_bw() +
+  ggtitle("Sweat Sodium")
+
+ggplot(df, aes(x = Diagnosis, y = K)) +
+  geom_boxplot() +
+  theme_bw() +
+  ggtitle("Sweat Potassium")
+
 ggplot(df, aes(x = Diagnosis, y = Cl)) +
   geom_boxplot() +
-  theme_bw()
+  theme_bw() +
+  ggtitle("Sweat Chloride")
 ```
@@
 # Write only those with multiple values
-mult <- unique(df$PatientID[which(duplicated(df$PatientID))])
 multiples <- df[df$PatientID %in% mult, ]

### 2022-09-08 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd

@@
     toc: yes
-    toc_depth: 2
+    toc_depth: 5
     toc_float: yes
@@
 df$Diagnosis <- diagnoses$Diagnosis[match(df$PatientID, diagnoses$PatientID)]
-# Filter per Maxie's code
-df <- df %>%
-  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks * 7) %>%
-  filter(
-    Age.at.Test.in.Years < 50 & Age.at.Test.in.Days >= 2,
-    !(Diagnosis == "Non-CF" & PatientID %in%
-      c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793))
-  ) %>%
-  distinct(.) %>%
-  filter(Sex != "", Demographics.Available == "TRUE")
 # Add modulator info
@@
 })
+# Filter per Maxie's code
+df <- df %>%
+  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks * 7) %>%
+  filter(
+    Age.at.Test.in.Years < 50 & Age.at.Test.in.Days >= 2,
+    !(Diagnosis == "Non-CF" & PatientID %in%
+      c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793)),
+    Modulator == "None"
+  ) %>%
+  distinct(.) %>%
+  filter(Sex != "", Demographics.Available == "TRUE")
 # Genotype info
@@
   levels = c(
-    "Mild_Severe", "Mild_Unknown", "Not CF_Not CF",
-    "Severe_Mild", "Mild_Mild",
-    "Severe_Severe", "Severe_Unknown", "Unknown_Severe",
+    "Severe_Severe","Mild_Severe", "Mild_Unknown", "Severe_Mild", "Mild_Mild","Not CF_Not CF",
+
+     "Severe_Unknown", "Unknown_Severe",
     "Unknown_Unknown"
@@
   labels = c(
-    "Low", "Low", "Not CF", "Low", "Low", "High", "Unknown",
+    "High","Low", "Low", "Low", "Low", "Not CF",  "Unknown",
     "Unknown", "Unknown"
@@
 })
+# Sodium to chloride ratio
+df$`Na/Cl Ratio` = df$Na /df$Cl
 # Sort
@@
 mult <- unique(df$PatientID[which(duplicated(df$PatientID))])
+# Testing options (most variables are not normally distributed)
+mycontrols  <- tableby.control(numeric.test="kwt",numeric.stats=c("Nmiss", "median", "q1q3"))
 ```
@@
   filter(row_number() == 1) # First measure for everyone
-t1 <- tableby(Diagnosis ~ Sex + Age.at.Test.in.Years + geno_risk + Amount + Cl + Na + K,
-  data = t1_df
+t1 <- tableby(Diagnosis ~ Sex + Age.at.Test.in.Years + fe(geno_risk) + Amount + Cl + Na + K + `Na/Cl Ratio`,
+  data = t1_df,control = mycontrols,simulate.p.value=T
 )
-summary(t1,
+summary(t1,pfootnote = T,
   labelTranslations =
@@
 
-### All participants
-
 ```{r results='asis'}
-t1 <- tableby(Method ~ Sex + Age.at.Test.in.Years + Diagnosis + geno_risk + Amount + Cl + Na + K + Arm,
-  data = t1_df
+t1 <- tableby(Method ~ Sex + Age.at.Test.in.Years + Diagnosis + geno_risk + Amount + Cl + Na + K + `Na/Cl Ratio` + Arm,
+  data = t1_df,control = mycontrols,
 )
-summary(t1,
+summary(t1,pfootnote = T,
   labelTranslations =
@@
 
-### Participants with more than one measure
+# Electrolyte comparisons
+
+## By group
 
 ```{r results='asis'}
-t1 <- tableby(Method ~ Sex + Age.at.Test.in.Years + Diagnosis + geno_risk + Amount + Cl + Na + K + Arm,
-  data = t1_df[t1_df$PatientID %in% mult, ]
+t1 <- tableby(Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` + Arm,
+  data = t1_df,control = mycontrols,
 )
-summary(t1,
+summary(t1,pfootnote = T,
   labelTranslations =
@@
 
-# Chloride and Diagnosis
+### Gibson-Cooke only
 
-```{r}
-df$Cl_group <- cut(df$Cl, c(0, 30, 60, Inf), right = F)
-t <- table(df$Cl_group, df$Diagnosis)
-t <- rownames_to_column(as.data.frame.matrix(t))
-colnames(t)[1] <- "Chloride Range"
-kable(t, row.names = F)
+```{r results='asis'}
+t1 <- tableby(Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` + Arm,
+  data = t1_df[t1_df$Method == "Gibson-Cooke",],control = mycontrols,
+)
+summary(t1,pfootnote = T,
+  labelTranslations =
+    list(Age.at.Test.in.Years = "Age (Years)", geno_risk = "Genotype Severity")
+)
 ```
 
-# Plots
+### Macroduct only
+
+```{r results='asis'}
+t1 <- tableby(Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` + Arm,
+  data = t1_df[t1_df$Method == "Macroduct",],control = mycontrols,
+)
+summary(t1,pfootnote = T,
+  labelTranslations =
+    list(Age.at.Test.in.Years = "Age (Years)", geno_risk = "Genotype Severity")
+)
+```
+
+## By genotype risk
+
+```{r results='asis'}
+t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + Arm,
+  data = t1_df,control = mycontrols,
+)
+summary(t1,pfootnote = T,
+  labelTranslations =
+    list(Age.at.Test.in.Years = "Age (Years)", geno_risk = "Genotype Severity")
+)
+```
+
+## By genotype risk and diagnosis
+
+### Cystic fibrosis
+
+```{r results='asis'}
+t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + Arm,
+  data = t1_df[t1_df$Diagnosis == "Cystic Fibrosis",],control = mycontrols
+)
+summary(t1,pfootnote = T,
+  labelTranslations =
+    list(Age.at.Test.in.Years = "Age (Years)", geno_risk = "Genotype Severity")
+)
+```
+
+### CFTR-related disorder
+
+```{r results='asis'}
+t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + Arm,
+  data = t1_df[t1_df$Diagnosis == "CFTR-related disorder",],control = mycontrols
+)
+summary(t1,pfootnote = T,
+  labelTranslations =
+    list(Age.at.Test.in.Years = "Age (Years)", geno_risk = "Genotype Severity")
+)
+```
+
+### CRMS
+
+```{r results='asis'}
+t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + Arm,
+  data = t1_df[t1_df$Diagnosis == "CRMS",],control = mycontrols
+)
+summary(t1,pfootnote = T,
+  labelTranslations =
+    list(Age.at.Test.in.Years = "Age (Years)", geno_risk = "Genotype Severity")
+)
+```
+
+### Non-CF
+
+```{r results='asis'}
+t1 <- tableby( ~ Amount + Cl + Na + K + `Na/Cl Ratio` + Arm,
+  data = t1_df[t1_df$Diagnosis == "Non-CF",],control = mycontrols
+)
+summary(t1,pfootnote = T,
+  labelTranslations =
+    list(Age.at.Test.in.Years = "Age (Years)", geno_risk = "Genotype Severity")
+)
+```
+
+# Electrolyte plots
 
@@
   ggtitle("Sweat Chloride")
+
+ggplot(df, aes(x = Diagnosis, y = `Na/Cl Ratio`)) +
+  geom_boxplot() +
+  theme_bw() +
+  ggtitle("Na/Cl Ratio")
 ```
 
-```{r}
-# Write only those with multiple values
-multiples <- df[df$PatientID %in% mult, ]
-write.csv(multiples, file = "./Data_Cleaned/multiple_values.csv", na = "", row.names = F)
-```
+# Questions
+
+1. What are "CL diff" and "NA diff" (e.g. in Table 4)? If it's one sweat collection per patient, is is the average difference between people?

### 2022-09-12 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd

--- 2022-09-07 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
@@
 library(tidyverse)
-library(performance)
+library(pROC)
 library(knitr)
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
+knitr::opts_chunk$set(echo = FALSE)
 knitr::opts_knit$set(root.dir = "/Users/timvigers/Documents/Work/CF/Edith Zemanick/Sweat Test")
@@
 df$Diagnosis <- diagnoses$Diagnosis[match(df$PatientID, diagnoses$PatientID)]
-# Filter per Maxie's code
-df <- df %>%
-  filter(
-    Age.at.Test.in.Years < 50,
-    !(Diagnosis == "Non-CF" & PatientID %in%
-      c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793))
-  ) %>%
-  distinct(.) %>%
-  filter(Sex != "", Demographics.Available == "TRUE") %>%
-  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks * 7)
 # Add modulator info
@@
 })
+# Numeric variables
+num_vars = c("Amount","Na","K","Cl")
+num_vars = c(paste0(num_vars,"_Left"),paste0(num_vars,"_Right"))
+df[,num_vars] = suppressWarnings(lapply(df[,num_vars],as.numeric))
+# Filter per Maxie's code
+df <- df %>%
+  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks * 7) %>%
+  filter(
+    Age.at.Test.in.Years < 50 & Age.at.Test.in.Days >= 2,
+    !(Diagnosis == "Non-CF" & PatientID %in%
+      c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793)),
+    Modulator == "None"
+  ) %>%
+  distinct(.) %>%
+  filter(Sex != "", Demographics.Available == "TRUE")
 # Genotype info
@@
 )
+# Which arm?
+df$Arm <- apply(df[, c("Cl", "Cl_Left", "Cl_Right", "Na", "Na_Left", "Na_Right", "K", "K_Left", "K_Right")], 1, function(r) {
+  k <- sum(!is.na(r))
+  if (k == 6) {
+    return("Both")
+  } else if (k == 3) {
+    return("Arm Not Noted")
+  } else {
+    return(NA)
+  }
+})
+
+# Difference between arms
+df$`Cl Diff. Between Left and Right Arm` = as.numeric(df$Cl_Left) - as.numeric(df$Cl_Right)
+df$`Na Diff. Between Left and Right Arm` = as.numeric(df$Na_Left) - as.numeric(df$Na_Right)
 # Merge left and right values
@@
 })
+# Sodium to chloride ratio
+df$`Na/Cl Ratio` = df$Na /df$Cl
 # Sort
 df <- df %>% arrange(PatientID, Age.at.Test.in.Weeks)
+# Table options (most variables are not normally distributed)
+mycontrols  <- tableby.control(numeric.test="kwt",numeric.stats=c("Nmiss", "median", "q1q3"))
+labels = list(Age.at.Test.in.Years = "Age (Years)", geno_risk = "Genotype Severity")
 ```
 
-# Table 1: Patient Demographics and Sample Characteristics
+# Patient Demographics and Sample Characteristics
 
-For those with two sweat electrolyte measures (left and right), the average of the two was used.
+For those with two sweat electrolyte measures (left and right), the average of the two was used. Each participants' first sweat test value was used for the following tables
+
+## By diagnosis
 
@@
 t1_df <- df %>%
-  filter(Age.at.Test.in.Days >= 2) %>%
   group_by(PatientID) %>%
-  arrange(Age.at.Test.in.Weeks) %>%
-  filter(row_number() == 1)
-t1 <- tableby(Diagnosis ~ Sex + Age.at.Test.in.Years + geno_risk + Amount + Cl + Na + K,
-  data = t1_df
+  filter(row_number() == 1) # First measure for everyone
+t1 <- tableby(Diagnosis ~ Sex + Age.at.Test.in.Years + fe(geno_risk) + Amount + Cl + Na + K + `Na/Cl Ratio`,
+  data = t1_df,control = mycontrols,simulate.p.value=T
 )
-summary(t1,
-  labelTranslations =
-    list(Age.at.Test.in.Years = "Age (Years)", geno_risk = "Genotype Severity")
+summary(t1,pfootnote = T,
+  labelTranslations = labels
 )
@@
 
-# Chloride and Diagnosis
+## By method
 
-## Table
+```{r results='asis'}
+t1 <- tableby(Method ~ Sex + Age.at.Test.in.Years + Diagnosis + geno_risk + Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+  data = t1_df,control = mycontrols,
+)
+summary(t1,pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+# Electrolyte comparisons
+
+## By diagnosis
+
+```{r results='asis'}
+t1 <- tableby(Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+  data = t1_df,control = mycontrols,
+)
+summary(t1,pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### Gibson-Cooke only
+
+```{r results='asis'}
+t1 <- tableby(Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+  data = t1_df[t1_df$Method == "Gibson-Cooke",],control = mycontrols,
+)
+summary(t1,pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### Macroduct only
+
+```{r results='asis'}
+t1 <- tableby(Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+  data = t1_df[t1_df$Method == "Macroduct",],control = mycontrols,
+)
+summary(t1,pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+## By genotype risk
+
+```{r results='asis'}
+t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+  data = t1_df,control = mycontrols,
+)
+summary(t1,pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+## By genotype risk and diagnosis
+
+### Cystic fibrosis
+
+```{r results='asis'}
+t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+  data = t1_df[t1_df$Diagnosis == "Cystic Fibrosis",],control = mycontrols
+)
+summary(t1,pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### CFTR-related disorder
+
+```{r results='asis'}
+t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+  data = t1_df[t1_df$Diagnosis == "CFTR-related disorder",],control = mycontrols
+)
+summary(t1,pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### CRMS
+
+```{r results='asis'}
+t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+  data = t1_df[t1_df$Diagnosis == "CRMS",],control = mycontrols
+)
+summary(t1,pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### Non-CF
+
+```{r results='asis'}
+t1 <- tableby( ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+  data = t1_df[t1_df$Diagnosis == "Non-CF",],control = mycontrols
+)
+summary(t1,pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+## Plots
+
+```{r warning=FALSE}
+ggplot(df, aes(x = Diagnosis, y = Na)) +
+  geom_boxplot() +
+  theme_bw() +
+  ggtitle("Sweat Sodium")
+
+ggplot(df, aes(x = Diagnosis, y = K)) +
+  geom_boxplot() +
+  theme_bw() +
+  ggtitle("Sweat Potassium")
+
+ggplot(df, aes(x = Diagnosis, y = Cl)) +
+  geom_boxplot() +
+  theme_bw() +
+  ggtitle("Sweat Chloride")
+
+ggplot(df, aes(x = Diagnosis, y = `Na/Cl Ratio`)) +
+  geom_boxplot() +
+  theme_bw() +
+  ggtitle("Na/Cl Ratio")
+```
+
+# Between occasion
+
+For all participants with more than one sweat test, we compared their first and last values.
+
+```{r results='asis'}
+# Get first and last values for everyone with more than 1
+multiples = df %>% group_by(PatientID) %>% 
+  filter(n() > 1) %>% # Multiple values only
+  filter(row_number()==1 | row_number()==n()) %>% # First and last
+  mutate(time = row_number())
+multiples$time = factor(multiples$time,levels = 1:2,labels = c("First","Last"))
+# Paired testing options
+mycontrols  <- tableby.control(numeric.test="signed.rank",cat.test = "mcnemar",
+                               numeric.stats=c("Nmiss", "median", "q1q3"))
+```
+
+## Overall
+
+```{r results='asis'}
+p <- paired(time ~ Diagnosis +Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,signed.rank.exact = F,
+            data = multiples, id = PatientID,control = mycontrols)
+summary(p,pfootnote = T,labelTranslations = labels)
+```
+
+## By diagnosis
+
+Not enough observations in CRMS or CFTR-related disorder groups to calculate p values. Also not enough observations to further break down the table by sweat collection method.
+
+```{r results='asis'}
+p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,signed.rank.exact = F,strata = Diagnosis,
+            data = multiples %>% filter(Diagnosis %in% c("Cystic Fibrosis","Non-CF")), 
+            id = PatientID,control = mycontrols)
+summary(p,pfootnote = T,labelTranslations = labels)
+```
+
+## By genotype risk
+
+```{r results='asis'}
+p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,signed.rank.exact = F,strata = geno_risk,
+            data = multiples, 
+            id = PatientID,control = mycontrols)
+summary(p,pfootnote = T,labelTranslations = labels)
+```
+
+## By genotype risk and diagnosis
+
+Not enough paired observations in the CRMS or CFTR-related disorder groups.
+
+### Cystic fibrosis
+
+```{r results='asis'}
+p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio`,signed.rank.exact = F,strata = geno_risk,
+            data = multiples %>% filter(Diagnosis == "Cystic Fibrosis"), 
+            id = PatientID,control = mycontrols)
+summary(p,pfootnote = T,labelTranslations = labels)
+```
+
+### Non-CF
+
+```{r results='asis'}
+p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio`,signed.rank.exact = F,strata = geno_risk,
+            data = multiples %>% filter(Diagnosis == "Non-CF"), 
+            id = PatientID,control = mycontrols)
+summary(p,pfootnote = T,labelTranslations = labels)
+```
+
+# Chloride groups
+
+We performed receiver operating characteristic (ROC) analysis to find a sweat sodium cutoff that distinguishes between chloride groups.
+Only the first sweat test from each participant was used.
+Because the groups are based on chloride value, Na/Cl ratio is correlated with the outcome so Na alone was used instead.
+Also, ROC analyses generally assume a binary outcome (i.e.
+looking for a cutoff to distinguish between two groups), so
+
+In the following results, "Low" indicates a sweat chloride under 30, "Medium" indicates a sweat chloride of 30 - 59, and "High" indicates a sweat chloride greater than or equal to 60.
 
 ```{r}
-df$Cl_group <- cut(df$Cl, c(0, 30, 60, Inf), right = F)
-t <- table(df$Cl_group, df$Diagnosis)
-t <- rownames_to_column(as.data.frame.matrix(t))
-colnames(t)[1] <- "Chloride Range"
-kable(t, row.names = F)
+# Three groups
+t1_df$Group = cut(t1_df$Cl,breaks = c(0,30,60,Inf),right = F,labels = c("Low","Medium","High"))
+# Low vs. medium or high
+t1_df$Low = t1_df$Group
+levels(t1_df$Low) = c("Low","Medium or High","Medium or High")
+t1_df$Low = relevel(t1_df$Low,ref = "Medium or High")
+# High vs. medium or low
+t1_df$High = t1_df$Group
+levels(t1_df$High) = c("Low or Medium","Low or Medium","High")
+t1_df$High = relevel(t1_df$High,ref = "Low or Medium")
 ```
 
-## Plot
+## ROC analysis
 
 ```{r}
-ggplot(df, aes(x = Diagnosis, y = Cl)) +
-  geom_boxplot() +
-  theme_bw()
+roc_low <- roc(response = t1_df$Low,predictor = t1_df$Na,quiet = T)
+coords_low = round(coords(roc_low,"best",best.method="youden"),3)
+roc_high <- roc(response = t1_df$High,predictor = t1_df$Na,quiet = T)
+coords_high = round(coords(roc_high,"best",best.method="youden"),3)
+
+plot(roc_low,print.auc = TRUE, col = "blue") 
+plot(roc_high, print.auc = TRUE, col = "green", print.auc.y = .4, add = TRUE)
 ```
 
-```{r}
-# Write only those with multiple values
-mult <- unique(df$PatientID[which(duplicated(df$PatientID))])
-multiples <- df[df$PatientID %in% mult, ]
-write.csv(multiples, file = "./Data_Cleaned/multiple_values.csv", na = "", row.names = F)
+The blue curve shows the ROC for distinguishing between low vs.
+medium or high, and the green is the ROC curve for distinguishing between low or medium vs.
+high.
+The optimal threshold for distinguishing between low vs.
+medium or high was `r coords_low$threshold` (sensitivity = `r coords_low$sensitivity`, specificity = `r coords_low$specificity`) based on the Youden index.
+The optimal threshold for distinguishing between high vs.
+low or medium was `r coords_high$threshold` (sensitivity = `r coords_high$sensitivity`, specificity = `r coords_high$specificity`) based on the Youden index.
+The mean AUC for the two curves is `r round(mean(c(roc_low$auc,roc_high$auc)),3)`.
+
+## Histogram by group
+
+Vertical red lines indicate the thresholds identified above.
+
+```{r warning=FALSE}
+ggplot(t1_df[!is.na(t1_df$Group),],aes(Na,fill = Group)) + 
+  geom_histogram(alpha=0.3, position="identity",binwidth = 1) + 
+  geom_vline(xintercept = coords_low$threshold,color = "red") +
+  geom_vline(xintercept = coords_high$threshold,color = "red") +
+  ylab("Count") + theme_bw()
 ```

### 2022-09-13 Edith Zemanick/CF Nutrition/cf_nutrition.qmd
Create cf_nutrition.qmd

---
title: "CF Nutrition Pre- and Post-CFTR Modulator"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    self-contained: true
editor: visual
---

```{r include=FALSE}
library(readxl)
library(tidyverse)
```

```{r}
# Import
# Modulator dates
mods = read_excel("/Users/timvigers/Documents/Work/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report CFTR Modulators_throughDecember2021.xlsx")
# Sort and remove those where start and end date are the same
mods = mods %>% 
  arrange(`Patient ID`,`Earliest Date in Registry`) %>%
  filter(`Earliest Date in Registry` != `Last Date in Registry`)
# Import vitamin data
vit = read_excel("/Users/timvigers/Documents/Work/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report Vitamins_throughDecember2021_revised20220712.xlsx")
# Add FEV information
fev = read_excel("/Users/timvigers/Documents/Work/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report_PFTs_ThroughDecember2021.xlsx")
fev = fev %>%

fev = fev.rename(columns={'PatientID': 'Patient ID'})
df = pd.merge(vit, fev, how="outer", on=["Patient ID", "Date"])
```

### 2022-09-22 Edith Zemanick/CF Nutrition/cf_nutrition.qmd
prettier table 1

@@
 library(tidyverse)
+library(arsenal)
 ```
@@
 ```{r}
-# Import
-# Modulator dates
-mods = read_excel("/Users/timvigers/Documents/Work/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report CFTR Modulators_throughDecember2021.xlsx")
-# Sort and remove those where start and end date are the same
-mods = mods %>% 
-  arrange(`Patient ID`,`Earliest Date in Registry`) %>%
-  filter(`Earliest Date in Registry` != `Last Date in Registry`)
-# Import vitamin data
-vit = read_excel("/Users/timvigers/Documents/Work/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report Vitamins_throughDecember2021_revised20220712.xlsx")
-# Add FEV information
-fev = read_excel("/Users/timvigers/Documents/Work/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report_PFTs_ThroughDecember2021.xlsx")
-fev = fev %>%
+#
 
-fev = fev.rename(columns={'PatientID': 'Patient ID'})
-df = pd.merge(vit, fev, how="outer", on=["Patient ID", "Date"])
 ```
+
+# Data Cleaning
+
+-   All vitamin, PFT, and BMI percentile values were converted to numeric, so values such as "LESS THAN 0.4" were marked as missing.
+
+-   We excluded visits more than 3 years prior to effective modulator start and within 90 after effective modulator start.
+
+-   All values prior to effective modulator start were averaged to calculate the "pre" value.
+We selected the most recent vitamin values as the "post" value, and included any BMI percentile and FEV1 % predicted measures within a month of that visit.
+
+-   Sex was only included in the PFT datasheet, so those without PFT data were missing this variable.
+
+# Grouped by First Modulator
+
+# Pre/post comparison
+
+```{r results='asis'}
+p = paired(Timepoint~Vitamin.A+Vitamin.D+Vitamin.E.Alpha+
+           ppFEV1+BMI.Percentile,data = df,id=ID,control = list(numeric.test = "signed.rank"))
+summary(p,pfootnote = T)
+```

### 2022-09-22 Edith Zemanick/CF Nutrition/cf_nutrition.qmd
grammar mistakes and minor formatting

@@
 ```{r}
-#
-
+# Import - see cf_nutrition_clean_data.py
+df = read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis.csv")
+df$Timepoint = factor(df$Timepoint,levels = c("Pre","Post"))
 ```
@@
 
-# Grouped by First Modulator
+# Baseline value grouped by first modulator
+
+```{r}
+t1 = tableby(First.Modulator~Vitamin.A+Vitamin.D+Vitamin.E.Alpha+
+           ppFEV1+BMI.Percentile,data = df[df$Timepoint == "Pre",])
+summary(t1,pfootnote = T)
+```
 

### 2022-09-22 Edith Zemanick/CF Nutrition/cf_nutrition_clean_data.py

@@ -9,7 +9,7 @@ mods = pd.read_excel(
 mods.sort_values(by=["Patient ID", "Earliest Date in Registry"],
                  ascending=True, inplace=True)
 mods = mods[mods['Earliest Date in Registry'] != mods['Last Date in Registry']]
-# vitemins
+# Vitamins
 vit = pd.read_excel(
     wd+"Data_Raw/Report Vitamins_throughDecember2021_revised20220712.xlsx")
 # Add FEV information

### 2022-09-23 Edith Zemanick/CF Nutrition/cf_nutrition.qmd
added LFTs

--- 2022-09-22 Edith Zemanick/CF Nutrition/cf_nutrition.qmd
@@
 library(tidyverse)
+library(parsedate)
 library(arsenal)
@@
 ```{r}
-#
-
+# Import all data
+df = read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis.csv")
+df$Timepoint = factor(df$Timepoint,levels = c("Pre","Post"))
 ```
@@
 -   All values prior to effective modulator start were averaged to calculate the "pre" value.
-We selected the most recent vitamin values as the "post" value, and included any BMI percentile and FEV1 % predicted measures within a month of that visit.
+We selected the most recent vitamin values as the "post" value, and included any BMI percentile, FEV1 % predicted, ALT, AST, and GGTP measures within a month of that visit.
 
@@
 
-# Grouped by First Modulator
+# Baseline value grouped by first modulator
+
+```{r}
+t1 = tableby(First.Modulator~Vitamin.A+Vitamin.D+Vitamin.E.Alpha+
+           ppFEV1+BMI.Percentile+ALT+AST+GGTP,data = df[df$Timepoint == "Pre",])
+summary(t1,pfootnote = T)
+```
 
@@
 p = paired(Timepoint~Vitamin.A+Vitamin.D+Vitamin.E.Alpha+
-           ppFEV1+BMI.Percentile,data = df,id=ID,control = list(numeric.test = "signed.rank"))
+           ppFEV1+BMI.Percentile+ALT+AST+GGTP,data = df,id=ID,control = list(numeric.test = "signed.rank"))
 summary(p,pfootnote = T)

### 2022-09-23 Edith Zemanick/CF Nutrition/cf_nutrition_clean_data.py

@@ -16,6 +16,10 @@ vit = pd.read_excel(
 fev = pd.read_excel(wd+"Data_Raw/Report_PFTs_ThroughDecember2021.xlsx")
 fev = fev.rename(columns={'PatientID': 'Patient ID'})
 df = pd.merge(vit, fev, how="outer", on=["Patient ID", "Date"])
+# Add LFT info
+lft = pd.read_excel(wd+"Data_Raw/Report_LFTs_ThroughDecember2021.xlsx")
+lft = lft[["Patient ID", "Date", "ALT", "AST", "GGTP"]]
+df = pd.merge(df, lft, how="outer", on=["Patient ID", "Date"])
 # Data cleaning
 df.sort_values(by=["Patient ID", "Date"], ascending=True, inplace=True)
 # Get modulator at each vitamin measure, pre-/post- information, and start date
@@ -68,7 +72,7 @@ df["First Modulator"] = f
 # Replace missing ages with info from PFT visit
 df["Age at Test (years)"].fillna(df["Age At PFT (years)"], inplace=True)
 # Fill sex info
-df["Sex"] = df.groupby(["Patient ID"], sort=False,group_keys=False)[
+df["Sex"] = df.groupby(["Patient ID"], sort=False, group_keys=False)[
     'Sex'].apply(lambda x: x.ffill().bfill())
 # Drop those without modulator info
 df.dropna(subset="Modulator", inplace=True)
@@ -81,7 +85,7 @@ df.to_csv(wd+"Data_Cleaned/all.csv", index=False)
 # Analysis dataset
 # Convert to numeric
 cols = ["VitaminA_Retinol", "25OH-Vitamin D", "Vitamin E Alpha",
-        "Vitamin E Gamma", "FEV1 % pred", "BMI percentile (CDC)"]
+        "Vitamin E Gamma", "FEV1 % pred", "BMI percentile (CDC)", "ALT", "AST", "GGTP"]
 df[cols] = df[cols].apply(pd.to_numeric, errors='coerce')
 # For each person, average all their pre-effective mod values
 # but only use the most recent vitamin measure.
@@ -89,7 +93,7 @@ df[cols] = df[cols].apply(pd.to_numeric, errors='coerce')
 # use that otherwise missing
 # Place to store results
[…]

### 2022-09-23 Edith Zemanick/CF Nutrition/cf_nutrition.qmd
trying to figure out liver labs in python

@@
 # Import all data
-df = read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis.csv")
-df$Timepoint = factor(df$Timepoint,levels = c("Pre","Post"))
+vit = read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report Vitamins_throughDecember2021_revised20220712.xlsx") %>%
+  select(`Patient ID`,Date,VitaminA_Retinol:`Vitamin E Gamma`)
+fev = read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report_PFTs_ThroughDecember2021.xlsx") %>% 
+  rename(`Patient ID` = PatientID) %>%
+  select(`Patient ID`,Date,BMI:`FEF25-75 % pred`)
+lft = read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report_LFTs_ThroughDecember2021.xlsx") %>%
+  select(`Patient ID`,Date,ALT:GGTP)
+# Format dates
+vit$Date = as.Date(parse_date(vit$Date,approx = F))
+fev$Date = as.Date(parse_date(fev$Date,approx = F))
+lft$Date = as.Date(parse_date(lft$Date,approx = F))
+# Merge and sort
+df = full_join(vit,fev,by = c("Patient ID", "Date"))
+df = full_join(df,lft,by = c("Patient ID", "Date"))
+df = df %>% arrange(`Patient ID`,Date)
+# Conver to numeric
+df[,3:ncol(df)] = lapply(df[,3:ncol(df)],as.numeric)
+# Format modulator info
+mods = read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report CFTR Modulators_throughDecember2021.xlsx")
+mods$`Age at Start (years)` = NULL
+mods = mods %>% filter(Modulator %in% c("Kalydeco","Trikafta"))
+mods$`Earliest Date in Registry` = as.Date(parse_date(mods$`Earliest Date in Registry`,
+                                                      approx = T))
+mods$`Last Date in Registry` = as.Date(parse_date(mods$`Last Date in Registry`,
+                                                      approx = T))
+mods = mods %>% pivot_wider(id_cols = `Patient ID`,names_from = Modulator,
+                     values_from = `Earliest Date in Registry`)
+mods$`Effective Modulator Start` = pmin(mods$Trikafta,mods$Kalydeco,na.rm = T)
+mods$fmt = mods$`Effective Modulator Start` == mods$Trikafta
+mods$fmk = mods$`Effective Modulator Start` == mods$Kalydeco
+mods[,c("fmt","fmk")][is.na(mods[,c("fmt","fmk")])]=0
+mods$`First Modulator` = c("Trikafta","Kalydeco")[max.col(mods[,c("fmt","fmk")])]
+mods[,c("fmt","fmk")] = NULL
+# Put together
+df = df %>% filter(`Patient ID` %in% mods$`Patient ID`)
+df = left_join(df,mods,by = "Patient ID")
+df$`Pre or Post` = factor(df$Date <= df$`Effective Modulator Start`,
+                          levels = c("TRUE","FALSE"),labels = c("Pre","Post"))
+# Filter and sort again, just in case
+df$`Days from Start` = as.numeric(difftime(df$Date,df$`Effective Modulator Start`,units = "days"))
+df = df %>% filter(`Days from Start` >= -365*3,
+                   `Days from Start` < 0 | `Days from Start` >=90) %>% 
+  arrange(`Patient ID`,Date)
+# Average pre-modulator values
+pre = df %>% group_by(`Patient ID`,`Pre or Post`) %>% summarise_at(vars(BMI:GGTP),mean, na.rm = T) %>% filter(`Pre or Post` == "Pre")
+# Find most recent vitamin measure, get any other measures within 30 days
+post = df[df$`Pre or Post`=="Post",]
 ```

### 2022-09-23 Edith Zemanick/CF Nutrition/cf_nutrition_clean_data.py

@@ -92,8 +92,12 @@ df[cols] = df[cols].apply(pd.to_numeric, errors='coerce')
 # If they have FEV1 or BMI within a month of the vitamin measure,
 # use that otherwise missing
 # Place to store results
-analysis = {"ID": [], "Timepoint": [], "Vitamin A": [], "Vitamin D": [], "Vitamin E Alpha": [], "Vitamin E Gamma": [], "ppFEV1": [], "BMI Percentile": [],
-            "First Modulator": [], "Age Nearest Modulator Start": [], "Sex": [], "ALT": [], "AST": [], "GGTP": []}
+analysis = {"ID": [], "Timepoint": [], "Vitamin A": [], "Vitamin D": [], "Vitamin E Alpha": [], "Vitamin E Gamma": [], "ppFEV1": [], "BMI Percentile": [], "First Modulator": [], "Age Nearest Modulator Start": [
+], "Sex": [], "ALT": [], "High ALT (3x)": [], "High ALT (5x)": [], "AST": [], "High AST (3x)": [], "High AST (5x)": [], "GGTP": [], "High GGTP (3x)": [], "High GGTP (5x)": []}
+
+analysis = {"ID": [], "Timepoint": [], "Vitamin A": [], "Vitamin D": [], "Vitamin E Alpha": [], "Vitamin E Gamma": [], "ppFEV1": [
+], "BMI Percentile": [], "First Modulator": [], "Age Nearest Modulator Start": [], "Sex": [], "ALT": [], "High ALT (3x)": [],  "AST": [],  "GGTP": []}
+
 # Iterate through by ID
 ids = df["Patient ID"].unique().tolist()
 for id in ids:
@@ -132,9 +136,49 @@ for id in ids:
     analysis["ppFEV1"].append(pre["FEV1 % pred"].mean(skipna=True))
     analysis["BMI Percentile"].append(
         pre["BMI percentile (CDC)"].mean(skipna=True))
-    analysis["ALT"].append(pre["ALT"].mean(skipna=True))
-    analysis["AST"].append(pre["AST"].mean(skipna=True))
-    analysis["GGTP"].append(pre["GGTP"].mean(skipna=True))
+    # Check liver labs
+    alt_lim = 55
+    alt_value = pre["ALT"].mean(skipna=True)
+    analysis["ALT"].append(alt_value)
+    # if pd.isnull(alt_value):
+    #     analysis["High ALT (3x)"].append(np.nan)
[…]

### 2022-09-23 Edith Zemanick/CF Nutrition/cf_nutrition.qmd
liver labs mostly done

--- 2022-09-23 Edith Zemanick/CF Nutrition/cf_nutrition.qmd
@@
 # Import all data
-df = read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis.csv")
+df = read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis.csv",na.strings = "",check.names = F)
 df$Timepoint = factor(df$Timepoint,levels = c("Pre","Post"))
@@
 
-```{r}
-t1 = tableby(First.Modulator~Vitamin.A+Vitamin.D+Vitamin.E.Alpha+
-           ppFEV1+BMI.Percentile+ALT+AST+GGTP,data = df[df$Timepoint == "Pre",])
-summary(t1,pfootnote = T)
+```{r results='asis'}
+t1 <- tableby(`First Modulator` ~ `Vitamin A` + `Vitamin D` + `Vitamin E Alpha` +
+  `ppFEV1` + `BMI Percentile` +
+  `ALT` + `High ALT (3x)` + `High ALT (5x)` +
+  `AST` + `High AST (3x)` + `High AST (5x)` +
+  `GGTP` + `High GGTP (3x)` + `High GGTP (5x)`,
+data = df[df$Timepoint == "Pre", ]
+)
+summary(t1, pfootnote = T)
 ```
@@
 ```{r results='asis'}
-p = paired(Timepoint~Vitamin.A+Vitamin.D+Vitamin.E.Alpha+
-           ppFEV1+BMI.Percentile+ALT+AST+GGTP,data = df,id=ID,control = list(numeric.test = "signed.rank"))
-summary(p,pfootnote = T)
+p <- paired(Timepoint ~ `Vitamin A` + signed.rank(`Vitamin D`) + `Vitamin E Alpha` +
+  `ppFEV1` + `BMI Percentile` +
+  signed.rank(`ALT`) + `High ALT (3x)` + notest(`High ALT (5x)`) +
+  signed.rank(`AST`) + `High AST (3x)` + notest(`High AST (5x)`) +
+  signed.rank(`GGTP`) + notest(`High GGTP (3x)`) + notest(`High GGTP (5x)`),
+data = df, id = ID
+)
+summary(p, pfootnote = T)
 ```

### 2022-09-23 Edith Zemanick/CF Nutrition/cf_nutrition_clean_data.py

@@ -352,10 +352,6 @@ for id in ids:
         analysis["High GGTP (5x)"].append("Yes")
     if ggtp_value < ggtp_lim*5:
         analysis["High GGTP (5x)"].append("No")
-    # # Checks
-    # if len(analysis["ALT"]) != len(analysis["High ALT (3x)"]):
-    #     print(id)
-    #     break
 # Convert to DF
 analysis = pd.DataFrame(analysis)
 analysis.to_csv(wd+"Data_Cleaned/analysis.csv", index=False)

### 2022-09-29 Edith Zemanick/CF Nutrition/cf_nutrition.qmd
Update cf_nutrition.qmd

--- 2022-09-23 Edith Zemanick/CF Nutrition/cf_nutrition.qmd
@@
 ```{r}
-# Import all data
-df = read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis.csv")
+# Import datasets
+vit <- read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report Vitamins_throughDecember2021_revised20220712.xlsx")
+mods <- read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report CFTR Modulators_throughDecember2021.xlsx")
+pfts <- read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report_PFTs_ThroughDecember2021.xlsx")
+lfts <- read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report_LFTs_ThroughDecember2021.xlsx")
+pi = read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report Fecal Elastase_throughDecember2021.xlsx")
+# Format
+vit$VitaminA_Retinol = as.numeric(vit$VitaminA_Retinol)
+vit$`Age at Test (years)` <- NULL
+vit$Date = as.Date(vit$Date)
+vit$`Vitamin E Gamma` = NULL
+mods$`Earliest Date in Registry` = as.Date(mods$`Earliest Date in Registry`)
+pfts <- pfts %>%
+  select(PatientID, Sex, Date, `Height (cm)`:`FEV1 % pred`) %>%
+  rename(`Patient ID` = PatientID)
+pfts$Date = as.Date(pfts$Date)
+lfts <- lfts %>% select(`Patient ID`, Date:GGTP)
+lfts$Date = as.Date(lfts$Date)
+pi = pi %>% select(`Patient ID`,Date,`Fecal Elastase`)
+pi$`Fecal Elastase` = as.numeric(pi$`Fecal Elastase`)
+pi$Date = as.Date(pi$Date)
+# Get first modulator and start date
+mods$`Earliest Date in Registry` <- parse_date(mods$`Earliest Date in Registry`, approx = F)
+mods <- mods %>%
+  filter(Modulator %in% c("Kalydeco", "Trikafta")) %>%
+  pivot_wider(
+    id_cols = `Patient ID`, names_from = Modulator,
+    values_from = c(`Earliest Date in Registry`,`Age at Start (years)`)
+  )
+mods$Start <- pmin(mods$`Earliest Date in Registry_Trikafta`, 
+                   mods$`Earliest Date in Registry_Kalydeco`, na.rm = T)
+mods$tf <- mods$`Earliest Date in Registry_Trikafta` == mods$Start
+mods$kf <- mods$`Earliest Date in Registry_Kalydeco` == mods$Start
+mods <- mods %>% unite(First, tf, kf, na.rm = T)
+mods$`Age at Start` = pmin(mods$`Age at Start (years)_Trikafta`,mods$`Age at Start (years)_Kalydeco`,na.rm = T)
+mods$First <- factor(mods$First,
+  levels = c("FALSE_TRUE", "TRUE"),
+  labels = c("Kalydeco", "Trikafta")
+)
+# Combine
+vit = full_join(vit,pfts, by = c("Patient ID", "Date"))
+vit = full_join(vit,lfts, by = c("Patient ID", "Date"))
+vit = full_join(vit,pi,by = c("Patient ID", "Date"))
+vit <- left_join(vit, mods %>% select(`Patient ID`, Start, First,`Age at Start`), 
+                 by = "Patient ID")
+# Get pre vs. post info, time from start
+vit$`Pre-/Post-` <- factor(vit$Date >= vit$Start,
+  levels = c("FALSE", "TRUE"),
+  labels = c("Pre", "Post")
+)
+vit$`Days from Start` = as.numeric(difftime(vit$Date,vit$Start,units = "days"))
+# Filter those more than 3 years before start, and within 90 days after start
+vit = vit %>% filter(`Days from Start` > -3*365,
+                     `Days from Start` < 0 | `Days from Start` >= 90)
+# Order variables, fill sex
+vit = vit %>% select(`Patient ID`,Sex,Date,Start,First,`Age at Start`,
+                     `Pre-/Post-`,`Days from Start`,everything()) %>% 
+  group_by(`Patient ID`) %>% 
+  fill(Sex,.direction = "downup")
+# Average pre-mod values, get most recent vitamin values and and any other variables within 30 days
+pre = vit %>% group_by(`Patient ID`) %>% 
+  filter(`Pre-/Post-` == "Pre",
+         !is.na(VitaminA_Retinol) | !is.na(`25OH-Vitamin D`) | !is.na(`Vitamin E Alpha`)) %>%
+  mutate(n=n()) %>% filter(n>=2) %>% # At least two pre measures
+  summarise(`First Modulator` = as.character(First[1]),Timepoint = "Pre",
+            `Age at Start` = `Age at Start`[1],Sex = Sex[1],
+            across(`Days from Start`:`Fecal Elastase`, ~ mean(.x, na.rm = TRUE)))
+post = vit %>% group_by(`Patient ID`) %>% 
+  filter(`Pre-/Post-` == "Post",
+         !is.na(VitaminA_Retinol) | !is.na(`25OH-Vitamin D`) | !is.na(`Vitamin E Alpha`)) %>%
+  mutate(latest = max(Date),date_cutoff= latest - lubridate::days(30)) %>%
+  filter(Date >= date_cutoff) %>% 
+  arrange(`Patient ID`,Date) %>%
+  summarise(`First Modulator` = as.character(First[1]),Timepoint = "Post",
+            `Age at Start` = `Age at Start`[1],Sex = Sex[1],
+            across(`Days from Start`:`Fecal Elastase`,
+                   ~if(all(is.na(.))) NA else last(na.omit(.))))
+df = bind_rows(pre,post) %>%
+  mutate_all(~ifelse(is.nan(.), NA, .))
 df$Timepoint = factor(df$Timepoint,levels = c("Pre","Post"))
+# Pancreatic status and LFTs
+df$`Pancreatic Insufficient` = factor(df$`Fecal Elastase` < 128,
+                       levels = c("TRUE","FALSE"),labels = c("Yes","No"))
+df$`Elevated LFTs (3X)` = factor((!is.na(df$ALT) & df$ALT > 3*55) | 
+                                   !is.na(df$AST) & df$AST > 3*48 | 
+                                   !is.na(df$GGTP) & df$GGTP > 3*61,
+                                 levels = c("TRUE","FALSE"),labels = c("Yes","No"))
+df$`Elevated LFTs (5X)` = factor((!is.na(df$ALT) & df$ALT > 5*55) | 
+                                   !is.na(df$AST) & df$AST > 5*48 | 
+                                   !is.na(df$GGTP) & df$GGTP > 5*61,
+                                 levels = c("TRUE","FALSE"),labels = c("Yes","No"))
 ```
@@
 
--   We excluded visits more than 3 years prior to effective modulator start and within 90 after effective modulator start.
+-   We excluded visits more than 3 years prior to effective modulator start and within 90 days after effective modulator start.
 
 -   All values prior to effective modulator start were averaged to calculate the "pre" value.
-We selected the most recent vitamin values as the "post" value, and included any BMI percentile, FEV1 % predicted, ALT, AST, and GGTP measures within a month of that visit.
+We selected the most recent vitamin values as the "post" value, and included any BMI percentile, FEV1 % predicted, ALT, AST, and GGTP measures within 30 days of that visit.
 
@@
 
+-   Pancreatic insufficiency was determined based on a fecal elastase cutoff of \< 128 $\mu g/g$.
+
+-   Elevated liver labs were considered at two levels:
+
+    -   ALT, AST, or GGTP 3 times the normal level
+
+    -   ALT, AST, or GGTP 5 times the normal level
+
+-   Normal liver labs were considered:
+
+    -   ALT: 55 U/L
+
+    -   AST: 48 U/L
+
+    -   GGTP: 61 U/L
+
 # Baseline value grouped by first modulator
 
-```{r}
-t1 = tableby(First.Modulator~Vitamin.A+Vitamin.D+Vitamin.E.Alpha+
-           ppFEV1+BMI.Percentile+ALT+AST+GGTP,data = df[df$Timepoint == "Pre",])
-summary(t1,pfootnote = T)
+```{r results='asis'}
+t1 <- tableby(`First Modulator` ~ `Age at Start` + Sex + `VitaminA_Retinol` + `25OH-Vitamin D` + `Vitamin E Alpha` + `FEV1 % pred` + `BMI percentile (CDC)` + `ALT` + `AST` + `GGTP` + `Elevated LFTs (3X)` + `Elevated LFTs (5X)` + `Fecal Elastase` + `Pancreatic Insufficient`,data = df[df$Timepoint == "Pre", ],numeric.stats = c("Nmiss","medianrange"))
+summary(t1, pfootnote = T)
 ```
@@
 ```{r results='asis'}
-p = paired(Timepoint~Vitamin.A+Vitamin.D+Vitamin.E.Alpha+
-           ppFEV1+BMI.Percentile+ALT+AST+GGTP,data = df,id=ID,control = list(numeric.test = "signed.rank"))
-summary(p,pfootnote = T)
+p <- paired(Timepoint ~ `VitaminA_Retinol` + signed.rank(`25OH-Vitamin D`,"Nmiss","median","q1q3") + `Vitamin E Alpha` + `FEV1 % pred` + `BMI percentile (CDC)` + `ALT` + `AST` + `GGTP` + `Elevated LFTs (3X)` + `Elevated LFTs (5X)` + `Pancreatic Insufficient`,data = df, id = `Patient ID`)
+summary(p, pfootnote = T)
 ```

### 2022-09-30 Edith Zemanick/CF Nutrition/cf_nutrition.qmd
Update cf_nutrition.qmd

@@
     toc-float: true
+    code-fold: true
     self-contained: true
@@
 
-```{r}
+```{r warning=FALSE}
 # Import datasets
-vit <- read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report Vitamins_throughDecember2021_revised20220712.xlsx")
-mods <- read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report CFTR Modulators_throughDecember2021.xlsx")
-pfts <- read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report_PFTs_ThroughDecember2021.xlsx")
-lfts <- read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report_LFTs_ThroughDecember2021.xlsx")
-pi = read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report Fecal Elastase_throughDecember2021.xlsx")
+vit <- read_excel("/Users/timvigers/Dropbox/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report Vitamins_throughDecember2021_revised20220712.xlsx")
+mods <- read_excel("/Users/timvigers/Dropbox/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report CFTR Modulators_throughDecember2021.xlsx")
+pfts <- read_excel("/Users/timvigers/Dropbox/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report_PFTs_ThroughDecember2021.xlsx")
+lfts <- read_excel("/Users/timvigers/Dropbox/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report_LFTs_ThroughDecember2021.xlsx")
+geno = read_excel("/Users/timvigers/Dropbox/Work/Vigers/CF/CF Info/Sweat Tests from Internal Database_genotype class_withPancStatus.xlsx",na = ".")
 # Format
@@
 lfts$Date = as.Date(lfts$Date)
-pi = pi %>% select(`Patient ID`,Date,`Fecal Elastase`)
-pi$`Fecal Elastase` = as.numeric(pi$`Fecal Elastase`)
-pi$Date = as.Date(pi$Date)
+geno = geno %>% select(PatientID,genoRisk,PancreaticStatus) %>% distinct() %>% 
+  rename(`Patient ID`=PatientID,`Genotype Risk` = genoRisk,
+         `Pancreatic Status`=PancreaticStatus)
 # Get first modulator and start date
@@
 vit = full_join(vit,lfts, by = c("Patient ID", "Date"))
-vit = full_join(vit,pi,by = c("Patient ID", "Date"))
 vit <- left_join(vit, mods %>% select(`Patient ID`, Start, First,`Age at Start`), 
@@
             `Age at Start` = `Age at Start`[1],Sex = Sex[1],
-            across(`Days from Start`:`Fecal Elastase`, ~ mean(.x, na.rm = TRUE)))
+            across(`Days from Start`:GGTP, ~ mean(.x, na.rm = TRUE)))
 post = vit %>% group_by(`Patient ID`) %>% 
@@
             `Age at Start` = `Age at Start`[1],Sex = Sex[1],
-            across(`Days from Start`:`Fecal Elastase`,
+            across(`Days from Start`:GGTP,
                    ~if(all(is.na(.))) NA else last(na.omit(.))))
@@
 # Pancreatic status and LFTs
-df$`Pancreatic Insufficient` = factor(df$`Fecal Elastase` < 128,
-                       levels = c("TRUE","FALSE"),labels = c("Yes","No"))
 df$`Elevated LFTs (3X)` = factor((!is.na(df$ALT) & df$ALT > 3*55) | 
@@
                                  levels = c("TRUE","FALSE"),labels = c("Yes","No"))
+df$`Elevated LFTs (3X)`[rowSums(is.na(df[,c("ALT","AST","GGTP")]))==3]=NA
 df$`Elevated LFTs (5X)` = factor((!is.na(df$ALT) & df$ALT > 5*55) | 
@@
                                  levels = c("TRUE","FALSE"),labels = c("Yes","No"))
+df$`Elevated LFTs (5X)`[rowSums(is.na(df[,c("ALT","AST","GGTP")]))==3]=NA
+# Add genetic info
+df = left_join(df,geno,by = "Patient ID")
+# Delete those without paired obs.
+df = df %>%  group_by(`Patient ID`) %>% filter(n() == 2) %>% 
+  arrange(`Patient ID`,Timepoint)
+# Write data so Edith can check
+write.csv(df,file = "/Users/timvigers/Dropbox/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_data.csv",row.names = F,na="")
+write.csv(df[df$`First Modulator`=="Trikafta",],file = "/Users/timvigers/Dropbox/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_data_eti_only.csv",row.names = F,na="")
 ```
@@
 -   Sex was only included in the PFT datasheet, so those without PFT data were missing this variable.
-
--   Pancreatic insufficiency was determined based on a fecal elastase cutoff of \< 128 $\mu g/g$.
 
@@
 
-# Baseline value grouped by first modulator
+# Ivacaftor and ETI
+
+## Baseline value grouped by first modulator
 
 ```{r results='asis'}
-t1 <- tableby(`First Modulator` ~ `Age at Start` + Sex + `VitaminA_Retinol` + `25OH-Vitamin D` + `Vitamin E Alpha` + `FEV1 % pred` + `BMI percentile (CDC)` + `ALT` + `AST` + `GGTP` + `Elevated LFTs (3X)` + `Elevated LFTs (5X)` + `Fecal Elastase` + `Pancreatic Insufficient`,data = df[df$Timepoint == "Pre", ],numeric.stats = c("Nmiss","medianrange"))
+t1 <- tableby(`First Modulator` ~ `Age at Start` + Sex + `ALT` + `AST` + `GGTP` + `Elevated LFTs (3X)` + `Elevated LFTs (5X)`+`Pancreatic Status`,data = df[df$Timepoint == "Pre", ],numeric.stats = c("Nmiss","medianrange"),numeric.test="kwt")
 summary(t1, pfootnote = T)
@@
 
-# Pre/post comparison
+## Pre/post comparison
 
 ```{r results='asis'}
-p <- paired(Timepoint ~ `VitaminA_Retinol` + signed.rank(`25OH-Vitamin D`,"Nmiss","median","q1q3") + `Vitamin E Alpha` + `FEV1 % pred` + `BMI percentile (CDC)` + `ALT` + `AST` + `GGTP` + `Elevated LFTs (3X)` + `Elevated LFTs (5X)` + `Pancreatic Insufficient`,data = df, id = `Patient ID`)
+p <- paired(Timepoint ~ `VitaminA_Retinol` + `25OH-Vitamin D` + `Vitamin E Alpha` + `FEV1 % pred` + `BMI percentile (CDC)` + `ALT` + `AST` + `GGTP` + `Elevated LFTs (3X)` + `Elevated LFTs (5X)`,data = df, id = `Patient ID`,numeric.stats = c("Nmiss","medianrange"),numeric.test="signed.rank",signed.rank.exact = FALSE)
 summary(p, pfootnote = T)
 ```
+
+## Pre/post comparison by pancreatic status
+
+One participant with unknown pancreatic status was removed in order to calculate p values.
+
+```{r results='asis'}
+panc_df = df %>% filter(`Pancreatic Status`!= "Unknown")
+
+p <- paired(Timepoint ~ `VitaminA_Retinol` + `25OH-Vitamin D` + `Vitamin E Alpha` + `FEV1 % pred` + `BMI percentile (CDC)` + `ALT` + `AST` + `Elevated LFTs (3X)` + `Elevated LFTs (5X)`,data = panc_df, id = `Patient ID`,strata = `Pancreatic Status`,numeric.stats = c("Nmiss","medianrange"),numeric.test="signed.rank",signed.rank.exact = FALSE)
+summary(p, pfootnote = T)
+```
+
+# ETI only
+
+## Pre/post comparison
+
+```{r results='asis'}
+p <- paired(Timepoint ~ `VitaminA_Retinol` + `25OH-Vitamin D` + `Vitamin E Alpha` + `FEV1 % pred` + `BMI percentile (CDC)` + `ALT` + `AST` + `GGTP` + `Elevated LFTs (3X)` + `Elevated LFTs (5X)` + `Pancreatic Status`,data = df[df$`First Modulator`=="Trikafta",], id = `Patient ID`,numeric.stats = c("Nmiss","medianrange"),numeric.test="signed.rank",signed.rank.exact = FALSE)
+summary(p, pfootnote = T)
+```

### 2022-10-04 Edith Zemanick/CF Nutrition/cf_nutrition.qmd
Update cf_nutrition.qmd
missing data report

--- 2022-09-23 Edith Zemanick/CF Nutrition/cf_nutrition.qmd
@@
 
-```{r}
-# Import all data
-df = read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis.csv",na.strings = "",check.names = F)
-df$Timepoint = factor(df$Timepoint,levels = c("Pre","Post"))
+```{r warning=FALSE}
+# Import datasets
+vit <- read_excel("/Volumes/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report Vitamins_throughDecember2021_revised20220712.xlsx")
+mods <- read_excel("/Volumes/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report CFTR Modulators_throughDecember2021.xlsx")
+pfts <- read_excel("/Volumes/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report_PFTs_ThroughDecember2021.xlsx")
+lfts <- read_excel("/Volumes/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report_LFTs_ThroughDecember2021.xlsx")
+geno <- read_excel("/Volumes/Documents/Work/Vigers/CF/CF Info/Sweat Tests from Internal Database_genotype class_withPancStatus.xlsx", na = ".")
+# Format
+vit$VitaminA_Retinol <- as.numeric(vit$VitaminA_Retinol)
+vit$`Age at Test (years)` <- NULL
+vit$Date <- as.Date(vit$Date)
+vit$`Vitamin E Gamma` <- NULL
+mods$`Earliest Date in Registry` <- as.Date(mods$`Earliest Date in Registry`)
+pfts <- pfts %>%
+  select(PatientID, Sex, Date, `Height (cm)`:`FEV1 % pred`) %>%
+  rename(`Patient ID` = PatientID)
+pfts$Date <- as.Date(pfts$Date)
+lfts <- lfts %>% select(`Patient ID`, Date:GGTP)
+lfts$Date <- as.Date(lfts$Date)
+geno <- geno %>%
+  select(PatientID, genoRisk, PancreaticStatus) %>%
+  distinct() %>%
+  rename(
+    `Patient ID` = PatientID, `Genotype Risk` = genoRisk,
+    `Pancreatic Status` = PancreaticStatus
+  )
+# Get first modulator and start date
+mods <- mods %>%
+  filter(Modulator %in% c("Kalydeco", "Trikafta")) %>%
+  pivot_wider(
+    id_cols = `Patient ID`, names_from = Modulator,
+    values_from = c(`Earliest Date in Registry`, `Age at Start (years)`)
+  )
+mods$Start <- pmin(mods$`Earliest Date in Registry_Trikafta`,
+  mods$`Earliest Date in Registry_Kalydeco`,
+  na.rm = T
+)
+mods$tf <- mods$`Earliest Date in Registry_Trikafta` == mods$Start
+mods$kf <- mods$`Earliest Date in Registry_Kalydeco` == mods$Start
+mods <- mods %>% unite(First, tf, kf)
+mods$`Age at Start` <- pmin(mods$`Age at Start (years)_Trikafta`,
+  mods$`Age at Start (years)_Kalydeco`,
+  na.rm = T
+)
+mods$First <- factor(mods$First,
+  levels = c("FALSE_TRUE", "NA_TRUE", "TRUE_NA"),
+  labels = c("Kalydeco", "Kalydeco", "Trikafta")
+)
+# Combine
+vit <- full_join(vit, pfts, by = c("Patient ID", "Date"))
+vit <- full_join(vit, lfts, by = c("Patient ID", "Date"))
+vit <- left_join(vit, mods %>% select(`Patient ID`, Start, First, `Age at Start`),
+  by = "Patient ID"
+)
+# Get pre vs. post info, time from start
+vit$`Pre-/Post-` <- factor(vit$Date >= vit$Start,
+  levels = c("FALSE", "TRUE"),
+  labels = c("Pre", "Post")
+)
+vit$`Days from Start` <- as.numeric(difftime(vit$Date, vit$Start, units = "days"))
+# Filter out those more than 3 years before start, and within 90 days after start
+vit <- vit %>% filter(
+  `Days from Start` > -3 * 365,
+  `Days from Start` < 0 | `Days from Start` >= 90
+)
+# Order variables, fill sex
+vit <- vit %>%
+  select(
+    `Patient ID`, Sex, Date, Start, First, `Age at Start`,
+    `Pre-/Post-`, `Days from Start`, everything()
+  ) %>%
+  group_by(`Patient ID`) %>%
+  fill(Sex, .direction = "downup")
+# Average pre-mod values, get most recent vitamin values and and any other variables within 30 days
+pre_ids <- vit %>%
+  group_by(`Patient ID`) %>%
+  filter(`Pre-/Post-` == "Pre", !is.na(VitaminA_Retinol)) %>%
+  mutate(n = n()) %>%
+  filter(n >= 2) %>%
+  .$`Patient ID` %>%
+  unique() # At least two pre measures
+pre <- vit %>%
+  filter(`Patient ID` %in% pre_ids) %>%
+  summarise(
+    `First Modulator` = as.character(First[1]), Timepoint = "Pre",
+    `Age at Start` = `Age at Start`[1], Sex = Sex[1],
+    Start = as.character(Start[1]), `Latest Visit` = NA,
+    across(`Days from Start`:GGTP, ~ mean(.x, na.rm = TRUE))
+  )
+# Get visits within 30 days prior to latest vitamin measure post-CFTR modulator
+post_cutoff <- vit %>%
+  group_by(`Patient ID`) %>%
+  filter(`Pre-/Post-` == "Post", !is.na(VitaminA_Retinol)) %>%
+  mutate(latest = max(Date), date_cutoff = latest - lubridate::days(90)) %>%
+  select(`Patient ID`, latest, date_cutoff) %>%
+  distinct()
+vit <- left_join(vit, post_cutoff, by = "Patient ID")
+# Summarise visits into one row
+post <- vit %>%
+  filter(Date >= date_cutoff & Date <= latest) %>%
+  arrange(`Patient ID`, Date) %>%
+  summarise(
+    `First Modulator` = as.character(First[1]), Timepoint = "Post",
+    `Age at Start` = `Age at Start`[1], Sex = Sex[1],
+    Start = as.character(Start[1]),
+    `Latest Visit` = as.character(latest[1]),
+    across(
+      `Days from Start`:GGTP,
+      ~ if (all(is.na(.))) NA else last(na.omit(.))
+    )
+  ) %>%
+  rename()
+df <- bind_rows(pre, post) %>%
+  mutate_all(~ ifelse(is.nan(.), NA, .))
+df$Timepoint <- factor(df$Timepoint, levels = c("Pre", "Post"))
+# Pancreatic status and LFTs
+df$`Elevated LFTs (2X)` <- factor((!is.na(df$ALT) & df$ALT > 2 * 30) |
+  !is.na(df$AST) & df$AST > 2 * 30 |
+  !is.na(df$GGTP) & df$GGTP > 2 * 20,
+levels = c("TRUE", "FALSE"), labels = c("Yes", "No")
+)
+df$`Elevated LFTs (2X)`[rowSums(is.na(df[, c("ALT", "AST", "GGTP")])) == 3] <- NA
+df$`Elevated LFTs (3X)` <- factor((!is.na(df$ALT) & df$ALT > 3 * 30) |
+  !is.na(df$AST) & df$AST > 3 * 30 |
+  !is.na(df$GGTP) & df$GGTP > 3 * 20,
+levels = c("TRUE", "FALSE"), labels = c("Yes", "No")
+)
+df$`Elevated LFTs (3X)`[rowSums(is.na(df[, c("ALT", "AST", "GGTP")])) == 3] <- NA
+df$`Elevated LFTs (5X)` <- factor((!is.na(df$ALT) & df$ALT > 5 * 30) |
+  !is.na(df$AST) & df$AST > 5 * 30 |
+  !is.na(df$GGTP) & df$GGTP > 5 * 20,
+levels = c("TRUE", "FALSE"), labels = c("Yes", "No")
+)
+df$`Elevated LFTs (5X)`[rowSums(is.na(df[, c("ALT", "AST", "GGTP")])) == 3] <- NA
+# Add genetic info
+df <- left_join(df, geno, by = "Patient ID")
+# Delete those without paired obs.
+df <- df %>%
+  group_by(`Patient ID`) %>%
+  filter(n() == 2) %>%
+  arrange(`Patient ID`, Timepoint)
+# Write data so Edith can check
+write.csv(df, file = "/Volumes/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_data.csv", row.names = F, na = "")
+write.csv(df[df$`First Modulator` == "Trikafta", ], file = "/Volumes/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_data_eti_only.csv", row.names = F, na = "")
+# Write missing
+mrns <- read_excel("/Volumes/Documents/Work/Vigers/CF/CF Info/All Patients with CF or Non-CF_MRNs.xlsx")
+
+miss_post_pft <- df %>%
+  filter(Timepoint == "Post", is.na(`FEV1 % pred`)) %>%
+  mutate(
+    MRN = mrns$MRN[match(`Patient ID`, mrns$PatientID)],
+    `Start Plus 90` = parse_date(Start, approx = F) + lubridate::days(90),
+    `Latest Minus 30` = parse_date(`Latest Visit`, approx = F) - lubridate::days(30)
+  ) %>%
+  select(`Patient ID`, MRN, `First Modulator`, Start, `Start Plus 90`, `Latest Visit`, `Latest Minus 30`)
+
+write.csv(miss_post_pft, file = "/Volumes/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/missing_post_pft.csv", row.names = F, na = "")
+
+miss_post_lft <- df %>%
+  filter(Timepoint == "Post", is.na(ALT)) %>%
+  mutate(
+    MRN = mrns$MRN[match(`Patient ID`, mrns$PatientID)],
+    `Start Plus 90` = parse_date(Start, approx = F) + lubridate::days(90),
+    `Latest Minus 30` = parse_date(`Latest Visit`, approx = F) - lubridate::days(30)
+  ) %>%
+  select(`Patient ID`, MRN, `First Modulator`, Start, `Start Plus 90`, `Latest Visit`, `Latest Minus 30`)
+
+write.csv(miss_post_lft, file = "/Volumes/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/missing_post_lft.csv", row.names = F, na = "")
+
+miss_pre_pft <- df %>%
+  filter(Timepoint == "Pre", is.na(`FEV1 % pred`)) %>%
+  mutate(
+    MRN = mrns$MRN[match(`Patient ID`, mrns$PatientID)]
+  ) %>%
+  select(`Patient ID`, MRN, `First Modulator`, Start)
+
+write.csv(miss_pre_pft, file = "/Volumes/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/missing_pre_pft.csv", row.names = F, na = "")
+
+miss_pre_lft <- df %>%
+  filter(Timepoint == "Pre", is.na(ALT)) %>%
+  mutate(
+    MRN = mrns$MRN[match(`Patient ID`, mrns$PatientID)]
+  ) %>%
+  select(`Patient ID`, MRN, `First Modulator`, Start)
+
+write.csv(miss_pre_lft, file = "/Volumes/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/missing_pre_lft.csv", row.names = F, na = "")
 ```
@@
 
-# Baseline value grouped by first modulator
+-   Elevated liver labs were considered at two levels:
+
+    -   ALT, AST, or GGTP 3 times the normal level
+
+    -   ALT, AST, or GGTP 5 times the normal level
+
+-   Normal liver labs were considered:
+
+    -   ALT: 30 U/L
+
+    -   AST: 30 U/L
+
+    -   GGTP: 20 U/L
+
+# Ivacaftor and ETI
+
+## Baseline value grouped by first modulator
 
 ```{r results='asis'}
-t1 <- tableby(`First Modulator` ~ `Vitamin A` + `Vitamin D` + `Vitamin E Alpha` +
-  `ppFEV1` + `BMI Percentile` +
-  `ALT` + `High ALT (3x)` + `High ALT (5x)` +
-  `AST` + `High AST (3x)` + `High AST (5x)` +
-  `GGTP` + `High GGTP (3x)` + `High GGTP (5x)`,
-data = df[df$Timepoint == "Pre", ]
-)
+t1 <- tableby(`First Modulator` ~ `Age at Start` + Sex + `ALT` + `AST` + `GGTP` +`Elevated LFTs (2X)`+ `Elevated LFTs (3X)` + `Elevated LFTs (5X)`+`Pancreatic Status`,data = df[df$Timepoint == "Pre", ],numeric.stats = c("Nmiss","medianrange"),numeric.test="kwt")
 summary(t1, pfootnote = T)
@@
 
-# Pre/post comparison
+## Pre/post comparison
 
 ```{r results='asis'}
-p <- paired(Timepoint ~ `Vitamin A` + signed.rank(`Vitamin D`) + `Vitamin E Alpha` +
-  `ppFEV1` + `BMI Percentile` +
-  signed.rank(`ALT`) + `High ALT (3x)` + notest(`High ALT (5x)`) +
-  signed.rank(`AST`) + `High AST (3x)` + notest(`High AST (5x)`) +
-  signed.rank(`GGTP`) + notest(`High GGTP (3x)`) + notest(`High GGTP (5x)`),
-data = df, id = ID
-)
+p <- paired(Timepoint ~ `VitaminA_Retinol` + `25OH-Vitamin D` + `Vitamin E Alpha` + `FEV1 % pred` + `BMI percentile (CDC)` + `ALT` + `AST` + `GGTP` + `Elevated LFTs (2X)` + `Elevated LFTs (3X)`+`Elevated LFTs (5X)`,data = df, id = `Patient ID`,numeric.stats = c("Nmiss","medianrange"),numeric.test="signed.rank",signed.rank.exact = FALSE)
 summary(p, pfootnote = T)
 ```
+
+## Pre/post comparison by pancreatic status
+
+One participant with unknown pancreatic status was removed in order to calculate p values.
+
+```{r results='asis'}
+panc_df = df %>% filter(`Pancreatic Status`!= "Unknown")
+
+p <- paired(Timepoint ~ `VitaminA_Retinol` + `25OH-Vitamin D` + `Vitamin E Alpha` + `FEV1 % pred` + `BMI percentile (CDC)` + `ALT` + `AST` + `Elevated LFTs (3X)` + `Elevated LFTs (5X)`,data = panc_df, id = `Patient ID`,strata = `Pancreatic Status`,numeric.stats = c("Nmiss","medianrange"),numeric.test="signed.rank",signed.rank.exact = FALSE)
+summary(p, pfootnote = T)
+```
+
+# ETI only
+
+## Pre/post comparison
+
+```{r results='asis'}
+p <- paired(Timepoint ~ `VitaminA_Retinol` + `25OH-Vitamin D` + `Vitamin E Alpha` + `FEV1 % pred` + `BMI percentile (CDC)` + `ALT` + `AST` + `GGTP` + `Elevated LFTs (3X)` + `Elevated LFTs (5X)` + `Pancreatic Status`,data = df[df$`First Modulator`=="Trikafta",], id = `Patient ID`,numeric.stats = c("Nmiss","medianrange"),numeric.test="signed.rank",signed.rank.exact = FALSE)
+summary(p, pfootnote = T)
+```

### 2022-10-04 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
Update pre_post_trikafta_analysis.Rmd

--- 2022-04-28 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
@@
 library(tidyverse)
+library(readxl)
 library(lmerTest)
@@
 library(knitr)
-knitr::opts_chunk$set(echo = FALSE,fig.height = 8,fig.width = 10)
-if(Sys.info()["sysname"] == "Windows"){
-  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
-} else if (Sys.info()["sysname"] == "Linux"){
-  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
-} else if (Sys.info()["sysname"] == "Darwin"){
-  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
+knitr::opts_chunk$set(echo = FALSE, fig.height = 8, fig.width = 10)
+if (Sys.info()["sysname"] == "Windows") {
+  home_dir <- "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
+} else if (Sys.info()["sysname"] == "Linux") {
+  home_dir <- "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
+} else if (Sys.info()["sysname"] == "Darwin") {
+  home_dir <- "/Volumes/Documents/Work/Vigers/CF"
 }
@@
 # Import cleaned data from Casey
-df = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/analysis_data.csv",na.strings = "NA")
-df$X = NULL
+df <- read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/analysis_data.csv",
+  na.strings = "NA"
+)
+df$X <- NULL
 # Fill in MRN and CFF ID for everyone
-df = df %>% 
-  group_by(CFF.ID) %>% fill(MRN,.direction = "downup") %>% 
-  group_by(MRN) %>% fill(CFF.ID,.direction = "downup")
+df <- df %>%
+  group_by(CFF.ID) %>%
+  fill(MRN, .direction = "downup") %>%
+  group_by(MRN) %>%
+  fill(CFF.ID, .direction = "downup")
 # Assume that blanks for pancreatic and G tube indicate "no"
-df$Hispanic.Latinx[is.na(df$Hispanic.Latinx)] = FALSE
-df$Pancreatic.Status[is.na(df$Pancreatic.Status)] = "Sufficient"
-df$G.tube.in.past.12.months[is.na(df$G.tube.in.past.12.months)] = "No"
+df$Hispanic.Latinx[is.na(df$Hispanic.Latinx)] <- FALSE
+df$Pancreatic.Status[is.na(df$Pancreatic.Status)] <- "Sufficient"
+df$G.tube.in.past.12.months[is.na(df$G.tube.in.past.12.months)] <- "No"
 # Convert dates
-date_vars = c("DOB","Start","Date")
-df[,date_vars] = lapply(df[,date_vars], function(d){
-  parse_date_time(d,orders = c("mdy","ymd","dmy"),quiet = T)
+date_vars <- c("DOB", "Start", "Date")
+df[, date_vars] <- lapply(df[, date_vars], function(d) {
+  parse_date_time(d, orders = c("mdy", "ymd", "dmy"), quiet = T)
 })
 # Fix incorrect century (WHY LUBRIDATE WHY?!)
-df$DOB[which(df$DOB > lubridate::mdy("1/1/2022"))] = 
+df$DOB[which(df$DOB > lubridate::mdy("1/1/2022"))] <-
   df$DOB[which(df$DOB > lubridate::mdy("1/1/2022"))] - years(100)
-# Time from start and filtering
-df = df %>% group_by(CFF.ID) %>% fill(Start,.direction = "downup")
-df = df %>% mutate(Time = as.numeric(difftime(Date,Start,units = "days"))) 
 # Add UM (they did not provide dates, so need Time column)
-um_a1cs = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_a1cs.csv",na.strings = "")
-um_bmi = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_bmi.csv",na.strings = "")
-um_demo = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_demo.csv",na.strings = "")
-um_ogtt = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_ogtt.csv",na.strings = "")
-um_pfts = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_pfts.csv",na.strings = "")
-um = full_join(um_a1cs,um_bmi)
-um = full_join(um,um_ogtt)
-um = full_join(um,um_pfts)
-um = left_join(um,um_demo)
-um = um %>% filter(!is.na(CFRD.Status))
-um$Age = um$Age / 365.25
-df = full_join(df,um)
+df <- df %>%
+  group_by(CFF.ID) %>%
+  fill(Start, .direction = "downup")
+df <- df %>% mutate(Time = as.numeric(difftime(Date, Start, units = "days")))
+um_a1cs <- read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_a1cs.csv", na.strings = "")
+um_bmi <- read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_bmi.csv", na.strings = "")
+um_demo <- read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_demo.csv", na.strings = "")
+um_race <- read_excel("./Colleen Wood/Prepost Triakfta/Data_Raw/Data from UofM/CF_race_data.xlsx")
+um_race <- um_race %>%
+  select(`CFF ID`, Race) %>%
+  rename(CFF.ID = `CFF ID`)
+um_race$Race <- factor(um_race$Race,
+  levels = c(
+    "Black or African American", "Some other race",
+    "Two or more races", "White"
+  ),
+  labels = c("Black", "Other", "Other", "Caucasian")
+)
+um_ogtt <- read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_ogtt.csv", na.strings = "")
+um_pfts <- read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_pfts.csv", na.strings = "")
+um <- full_join(um_a1cs, um_bmi)
+um <- full_join(um, um_ogtt)
+um <- full_join(um, um_pfts)
+um <- left_join(um, um_demo)
+um <- left_join(um, um_race)
+um <- um %>% filter(!is.na(CFRD.Status),abs(Time) <= 365)
+um$Age <- um$Age / 365.25
+df <- full_join(df, um)
 # Filter and fill constants
-df = df %>% group_by(CFF.ID) %>% fill(DOB,Start:CFRD.Diagnosis.Date,.direction = "downup") %>%
-  ungroup() %>% filter(!is.na(Time),abs(Time) <= 730,!is.na(CFF.ID),
-                       (!is.na(CFRD.Diagnosis.Date) | !is.na(CFRD.Status) | site == "UW")) %>%
-  select(-MRN) %>% select(CFF.ID,site,DOB,Start,everything()) %>% rename(Site = "site")
-df$Sex = as.factor(df$Sex)
-levels(df$Sex) = c("F","M")
+df <- df %>%
+  group_by(CFF.ID) %>%
+  fill(DOB, Start:CFRD.Diagnosis.Date, .direction = "downup") %>%
+  ungroup() %>%
+  filter(
+    !is.na(Time), abs(Time) <= 730, !is.na(CFF.ID),
+    (!is.na(CFRD.Diagnosis.Date) | !is.na(CFRD.Status) | site == "UW")
+  ) %>%
+  select(-MRN) %>%
+  select(CFF.ID, site, DOB, Start, everything()) %>%
+  rename(Site = "site")
+df$Sex <- as.factor(df$Sex)
+levels(df$Sex) <- c("F", "M")
 # BMI category
-df$Age = pmax(as.numeric(difftime(df$Date,df$DOB,units = "days"))/365.25,df$Age,na.rm = T)
+df$Age <- pmax(as.numeric(difftime(df$Date, df$DOB, units = "days")) / 365.25, df$Age, na.rm = T)
 # Treat adults as age = 20
-df$Age_for_BMI = ifelse(df$Age > 20,20,df$Age)
-df$BMIz = y2z(df$BMI,x = df$Age_for_BMI,sex = df$Sex,ref=nl4.bmi)
-df$BMIz[is.infinite(df$BMIz)] = NA
-df$BMI.Percentile = pnorm(df$BMIz)*100
-df$Weightz = y2z(df$Weight,x = df$Age_for_BMI,sex = df$Sex,ref=nl4.wgt)
-df$BMI_group = apply(df,1,function(r){
-  age = as.numeric(r["Age"])
-  if(!is.na(age) & age > 19){
-    bmi = as.numeric(r["BMI"])
-    group = cut(bmi,c(-Inf,18.5,25,30,35,40,Inf),
-        labels = c("Under Weight","Normal Weight","Overweight","Obese (Class I)",
-                   "Obese (Class II)","Obese (Class III+)"),right = F)
-  } else if (!is.na(age) & age <= 19){
-    z = as.numeric(r["BMIz"])
-    perc = pnorm(z)
-    group = cut(perc,c(-Inf,5,85,95,Inf),
-        labels = c("Under Weight","Normal Weight","Overweight","Obese (Class I)"),right = F)
-  } else {group = NA}
+df$Age_for_BMI <- ifelse(df$Age > 20, 20, df$Age)
+df$BMIz <- y2z(df$BMI, x = df$Age_for_BMI, sex = df$Sex, ref = nl4.bmi)
+df$BMIz[is.infinite(df$BMIz)] <- NA
+df$BMI.Percentile <- pnorm(df$BMIz) * 100
+df$Weightz <- y2z(df$Weight, x = df$Age_for_BMI, sex = df$Sex, ref = nl4.wgt)
+df$BMI_group <- apply(df, 1, function(r) {
+  age <- as.numeric(r["Age"])
+  if (!is.na(age) & age > 19) {
+    bmi <- as.numeric(r["BMI"])
+    group <- cut(bmi, c(-Inf, 18.5, 25, 30, 35, 40, Inf),
+      labels = c(
+        "Under Weight", "Normal Weight", "Overweight", "Obese (Class I)",
+        "Obese (Class II)", "Obese (Class III+)"
+      ), right = F
+    )
+  } else if (!is.na(age) & age <= 19) {
+    z <- as.numeric(r["BMIz"])
+    perc <- pnorm(z)
+    group <- cut(perc, c(-Inf, 5, 85, 95, Inf),
+      labels = c("Under Weight", "Normal Weight", "Overweight", "Obese (Class I)"), right = F
+    )
+  } else {
+    group <- NA
+  }
   return(as.character(group))
 })
-df$BMI_group = as.factor(df$BMI_group)
-df$BMI_group = ordered(df$BMI_group,c("Under Weight","Normal Weight","Overweight",
-                                      "Obese (Class I)","Obese (Class II)","Obese (Class III+)"))
+df$BMI_group <- as.factor(df$BMI_group)
+df$BMI_group <- ordered(df$BMI_group, c(
+  "Under Weight", "Normal Weight", "Overweight",
+  "Obese (Class I)", "Obese (Class II)", "Obese (Class III+)"
+))
 # Get genotype severity
-df$Genotype1_sev = get_cf_genotype_class_severity(df$Genotypes1)
-df$Genotype2_sev = get_cf_genotype_class_severity(df$Genotypes2)
+df$Genotype1_sev <- get_cf_genotype_class_severity(df$Genotypes1)
+df$Genotype2_sev <- get_cf_genotype_class_severity(df$Genotypes2)
 # PFTs
-old <- c("Caucasian","Black","Other")
-new <- c(1,2,5)
-df$ppFEV = pctpred_GLI(df$Age,df$Height/100,gender = as.numeric(df$Sex),
+old <- c("Caucasian", "Black", "Other")
+new <- c(1, 2, 5)
+df$ppFEV <- pctpred_GLI(df$Age, df$Height / 100,
+  gender = as.numeric(df$Sex),
   ethnicity = new[match(df$Race, old)],
-            FEV1 = df$FEV1)
-df$ppFVC = pctpred_GLI(df$Age,df$Height/100,gender = as.numeric(df$Sex),
+  FEV1 = df$FEV1
+)
+df$ppFVC <- pctpred_GLI(df$Age, df$Height / 100,
+  gender = as.numeric(df$Sex),
   ethnicity = new[match(df$Race, old)],
-            FEV1 = df$FVC)
-# T1D duration
-df$CFRD.Diagnosis.Date = mdy(df$CFRD.Diagnosis.Date)
-df$T1D.Duration = as.numeric(difftime(df$Date,df$CFRD.Diagnosis.Date,units = "days"))/365.25
+  FEV1 = df$FVC
+)
+# uration
+df$CFRD.Diagnosis.Date <- mdy(df$CFRD.Diagnosis.Date)
+df$CFRD.Duration <- as.numeric(difftime(df$Date, df$CFRD.Diagnosis.Date, units = "days")) / 365.25
+df$Age_group <- cut(df$Age, c(-Inf, 18, Inf), right = F, labels = c("Peds", "Adult"))
 # Before or after start
-df$Before_After = cut(df$Time,c(-Inf,0,Inf),labels = c("Before","After"),right = T)
+df$Before_After <- cut(df$Time, c(-Inf, 0, Inf), labels = c("Before", "After"), right = T)
 # Sort
-df = df %>% arrange(Site,CFF.ID,Date) %>% distinct()
+df <- df %>%
+  arrange(Site, CFF.ID, Date) %>%
+  distinct()
 # Write sorted DF
-write.csv(df, file = "./Colleen Wood/Prepost Triakfta/Data_Cleaned/analysis_data_with_UM.csv",
-          row.names = F,na = "")
+write.csv(df,
+  file = "./Colleen Wood/Prepost Triakfta/Data_Cleaned/analysis_data_with_UM.csv",
+  row.names = F, na = ""
+)
 ```
@@
 # (i.e. the longest time after start)
-most_recent = df %>% group_by(CFF.ID,Site,Before_After) %>%
-  filter(Time > -2*365) %>%
-  mutate(time_new = max(Time) - Time) %>%
-  arrange(time_new) %>% 
-  summarise(Age = first(na.omit(Age)),BMI.Percentile = first(na.omit(BMI.Percentile)),
-            Weight = first(na.omit(Weight)),Sex = first(na.omit(Sex)),
-            T1D.Duration = first(na.omit(T1D.Duration)),
-            G.tube.in.past.12.months = first(na.omit(G.tube.in.past.12.months)),
-            Pancreatic.Status = first(na.omit(Pancreatic.Status)),
-            Genotype1_sev = first(na.omit(Genotype1_sev)),
-            Genotype2_sev = first(na.omit(Genotype2_sev)),
-            BMIz = first(na.omit(BMIz)),Weightz = first(na.omit(Weightz)),
-            FEV1 = first(na.omit(FEV1)),FVC = first(na.omit(FVC)),
-            HbA1c = first(na.omit(HbA1c)),.groups = "keep") %>%
-  arrange(Site,CFF.ID,Before_After)
-# # Check
-# t = df %>% select(CFF.ID,Site,Date,Time,Before_After,BMIz,ppFEV,ppFVC) %>%
-#   group_by(CFF.ID,Site,Before_After) %>%
-#   arrange(Site,CFF.ID,Before_After)
+pre_post <- df %>%
+  group_by(CFF.ID, Before_After) %>%
+  arrange(CFF.ID, Date) %>%
+  summarise(
+    Start = last(na.omit(Start)),
+    Age = last(na.omit(Age)),
+    BMI.Percentile = last(na.omit(BMI.Percentile)),
+    Weight = last(na.omit(Weight)),
+    CFRD.Duration = last(na.omit(CFRD.Duration)),
+    Age_group = last(na.omit(Age_group)),
+    Sex = last(na.omit(Sex)),
+    G.tube.in.past.12.months = last(na.omit(G.tube.in.past.12.months)),
+    Pancreatic.Status = last(na.omit(Pancreatic.Status)),
+    Genotype1_sev = last(na.omit(Genotype1_sev)),
+    Genotype2_sev = last(na.omit(Genotype2_sev)),
+    BMIz = last(na.omit(BMIz)),
+    ppFEV1 = last(na.omit(ppFEV)),
+    ppFVC = last(na.omit(ppFVC)),
+    HbA1c = last(na.omit(HbA1c)),
+    Site = last(na.omit(Site)),
+    .groups = "keep"
+  )
 # Filter those without CFRD, and those without pre AND post
-keep = most_recent %>% filter(!is.na(HbA1c)) %>%
-  group_by(CFF.ID) %>% summarise(n = n()) %>%
-  filter(n == 2) %>% pull(CFF.ID)
-cfrd = unique(df$CFF.ID[!is.na(df$CFRD.Status) | !is.na(df$CFRD.Diagnosis.Date)])
-keep = unique(keep,cfrd)
-most_recent = most_recent %>% filter(CFF.ID %in% keep)
+keep <- df %>%
+  filter(!is.na(HbA1c)) %>%
+  group_by(CFF.ID) %>%
+  summarise(n = n()) %>%
+  filter(n == 2) %>%
+  pull(CFF.ID)
+cfrd <- unique(df$CFF.ID[!is.na(df$CFRD.Status) |
+  df$CFRD.Status != "Impaired Glucose Tolerance (FBG < 126, 2-h PG 140-199)" |
+  !is.na(df$CFRD.Diagnosis.Date)])
+keep <- unique(keep, cfrd)
+pre_post <- pre_post %>% filter(CFF.ID %in% keep)
+two_obs <- pre_post %>%
+  group_by(CFF.ID) %>%
+  summarise(n = n()) %>%
+  filter(n == 2) %>%
+  pull(CFF.ID)
+pre_post <- pre_post %>% filter(CFF.ID %in% two_obs)
+# Write
+write.csv(pre_post,
+  file = "./Colleen Wood/Prepost Triakfta/Data_Cleaned/pre_post.csv",
+  na = "", row.names = F
+)
 ```
@@
 
+```{r results='asis'}
+t1 <- tableby(
+  ~ Age + BMI.Percentile + Weight + ppFEV1 + ppFVC + CFRD.Duration + Sex + G.tube.in.past.12.months +
+    Pancreatic.Status + Genotype1_sev + Genotype2_sev + Site,
+  pre_post[pre_post$Before_After == "Before", ]
+)
+summary(t1)
+```
+
 # Pre-post
@@
 ```{r}
-summary(tableby( ~ ppFEV + ppFVC + BMIz,data = df[!is.na(df$average_sensor),]))
+summary(tableby(~ ppFEV + ppFVC + BMIz, data = df[!is.na(df$average_sensor), ]))
 ```
@@
 # Piecewise model function
-piecewise = function(outcome = "HbA1c",data = df,simple = F){
+piecewise <- function(outcome = "HbA1c", data = df, simple = F) {
   # Add spline column
-  data$Change = ifelse(data$Time < 0,0,1)
-  data$Change = data$Change * data$Time
+  data$Change <- ifelse(data$Time < 0, 0, 1)
+  data$Change <- data$Change * data$Time
   # Model
-  f = as.formula(paste0(outcome,"~Time + Change + (1|CFF.ID)"))
-  mod = lmer(f,data = data)
-  res = tidy(mod,"fixed")
-  res$effect = NULL
+  f <- as.formula(paste0(outcome, "~Time + Change + (1|CFF.ID)"))
+  mod <- lmer(f, data = data)
+  res <- tidy(mod, "fixed")
+  res$effect <- NULL
   # Model check if necessary
-  check = check_model(mod)
+  check <- check_model(mod)
   # Coordinates for model plotting
-  x1_start = -365
-  y1_start = res$estimate[1] + res$estimate[2]*x1_start
-  x1_end = 0
-  y1_end = res$estimate[1] 
-  x2_start = 0
-  y2_start = res$estimate[1] 
-  x2_end = 365
-  y2_end = res$estimate[1] + (res$estimate[2]+res$estimate[3])*x2_end
+  x1_start <- -365
+  y1_start <- res$estimate[1] + res$estimate[2] * x1_start
+  x1_end <- 0
+  y1_end <- res$estimate[1]
+  x2_start <- 0
+  y2_start <- res$estimate[1]
+  x2_end <- 365
+  y2_end <- res$estimate[1] + (res$estimate[2] + res$estimate[3]) * x2_end
   # Plot
-  p = ggplot(data,aes_string(x = "Time",y = outcome,group = "CFF.ID",color = "Site")) +
-    geom_point(alpha = 0.5) + geom_line(alpha = 0.2) +
-    annotate("segment",x=x1_start,y=y1_start,xend=x1_end,yend=y1_end,color = "blue",size = 1) +
-    annotate("segment",x=x2_start,y=y2_start,xend=x2_end,yend=y2_end,color = "blue",size = 1) +
-    xlab("Days from Trikafta Start") + theme_bw()
+  p <- ggplot(data, aes_string(x = "Time", y = outcome, group = "CFF.ID", color = "Site")) +
+    geom_point(alpha = 0.5) +
+    geom_line(alpha = 0.2) +
+    annotate("segment", x = x1_start, y = y1_start, xend = x1_end, yend = y1_end, color = "blue", size = 1) +
+    annotate("segment", x = x2_start, y = y2_start, xend = x2_end, yend = y2_end, color = "blue", size = 1) +
+    xlab("Days from Trikafta Start") +
+    theme_bw()
   # Interactive plot and model results
-  return(list(plot = ggplotly(p),table = res,check = check))
+  return(list(plot = ggplotly(p), table = res, check = check))
 }
@@
 ```{r results='asis'}
-a1c = piecewise(outcome = "HbA1c",data = df)
+a1c <- piecewise(outcome = "HbA1c", data = df)
 ggplotly(a1c$plot)
-kable(a1c$table,digits = 3)
+kable(a1c$table, digits = 3)
 ```
@@
 ```{r results='asis'}
-ppFEV = piecewise(outcome = "ppFEV",data = df)
+ppFEV <- piecewise(outcome = "ppFEV", data = df)
 ggplotly(ppFEV$plot)
-kable(ppFEV$table,digits = 3)
+kable(ppFEV$table, digits = 3)
 ```
@@
 ```{r results='asis'}
-ppFVC = piecewise(outcome = "ppFVC",data = df)
+ppFVC <- piecewise(outcome = "ppFVC", data = df)
 ggplotly(ppFVC$plot)
-kable(ppFVC$table,digits = 3)
+kable(ppFVC$table, digits = 3)
 ```
@@
 ```{r results='asis'}
-meansg = piecewise(outcome = "average_sensor",data = df)
+meansg <- piecewise(outcome = "average_sensor", data = df)
 ggplotly(meansg$plot)
-kable(meansg$table,digits = 3)
+kable(meansg$table, digits = 3)
 ```
@@
 ```{r results='asis'}
-minsg = piecewise(outcome = "min_sensor",data = df)
+minsg <- piecewise(outcome = "min_sensor", data = df)
 ggplotly(minsg$plot)
-kable(minsg$table,digits = 3)
+kable(minsg$table, digits = 3)
 ```
@@
 ```{r results='asis'}
-maxsg = piecewise(outcome = "max_sensor",data = df)
+maxsg <- piecewise(outcome = "max_sensor", data = df)
 ggplotly(maxsg$plot)
-kable(maxsg$table,digits = 3)
+kable(maxsg$table, digits = 3)
 ```
@@
 ```{r results='asis'}
-over120 = piecewise(outcome = "percent_time_over_120",data = df)
+over120 <- piecewise(outcome = "percent_time_over_120", data = df)
 ggplotly(over120$plot)
-kable(over120$table,digits = 3)
+kable(over120$table, digits = 3)
 ```
@@
 ```{r results='asis'}
-over140 = piecewise(outcome = "percent_time_over_140",data = df)
+over140 <- piecewise(outcome = "percent_time_over_140", data = df)
 ggplotly(over140$plot)
-kable(over140$table,digits = 3)
+kable(over140$table, digits = 3)
 ```
@@
 ```{r results='asis'}
-over200 = piecewise(outcome = "percent_time_over_200",data = df)
+over200 <- piecewise(outcome = "percent_time_over_200", data = df)
 ggplotly(over200$plot)
-kable(over200$table,digits = 3)
+kable(over200$table, digits = 3)
 ```
@@
 ```{r results='asis'}
-numover140 = piecewise(outcome = "avg_excur_over_140_per_day",data = df)
+numover140 <- piecewise(outcome = "avg_excur_over_140_per_day", data = df)
 ggplotly(numover140$plot)
-kable(numover140$table,digits = 3)
+kable(numover140$table, digits = 3)
 ```
@@
 ```{r results='asis'}
-numover200 = piecewise(outcome = "avg_excur_over_200_per_day",data = df)
+numover200 <- piecewise(outcome = "avg_excur_over_200_per_day", data = df)
 ggplotly(numover200$plot)
-kable(numover200$table,digits = 3)
+kable(numover200$table, digits = 3)
 ```
@@
 ```{r results='asis'}
-under70 = piecewise(outcome = "percent_time_under_70",data = df)
+under70 <- piecewise(outcome = "percent_time_under_70", data = df)
 ggplotly(under70$plot)
-kable(under70$table,digits = 3)
+kable(under70$table, digits = 3)
 ```
@@
 ```{r results='asis'}
-under60 = piecewise(outcome = "percent_time_under_60",data = df)
+under60 <- piecewise(outcome = "percent_time_under_60", data = df)
 ggplotly(under60$plot)
-kable(under60$table,digits = 3)
+kable(under60$table, digits = 3)
 ```
@@
 ```{r results='asis'}
-auc = piecewise(outcome = "average_auc_per_day",data = df)
+auc <- piecewise(outcome = "average_auc_per_day", data = df)
 ggplotly(auc$plot)
-kable(auc$table,digits = 3)
+kable(auc$table, digits = 3)
 ```
@@
 ```{r results='asis'}
-sd = piecewise(outcome = "standard_deviation",data = df)
+sd <- piecewise(outcome = "standard_deviation", data = df)
 ggplotly(sd$plot)
-kable(sd$table,digits = 3)
+kable(sd$table, digits = 3)
 ```
@@
 ```{r results='asis'}
-mage = piecewise(outcome = "r_mage",data = df)
+mage <- piecewise(outcome = "r_mage", data = df)
 ggplotly(mage$plot)
-kable(mage$table,digits = 3)
+kable(mage$table, digits = 3)
 ```

### 2022-10-04 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_merge.Rmd
Update pre_post_trikafta_merge.Rmd

@@
 } else if (Sys.info()["sysname"] == "Darwin"){
-  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood"
+  home_dir = "/Volumes/Documents/Work/Vigers/CF/Colleen Wood"
 }

### 2022-10-04 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
Update pre_post_trikafta_analysis.Rmd

--- 2022-04-28 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
@@
 library(tidyverse)
-library(lmerTest)
-library(nlme)
-library(broom.mixed)
-library(segmented)
-library(performance)
-library(AGD)
-library(rspiro)
 library(lubridate)
-library(plotly)
-library(knitr)
-knitr::opts_chunk$set(echo = FALSE,fig.height = 8,fig.width = 10)
-if(Sys.info()["sysname"] == "Windows"){
-  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
-} else if (Sys.info()["sysname"] == "Linux"){
-  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
-} else if (Sys.info()["sysname"] == "Darwin"){
-  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
-}
 source("~/GitHub/CF/get_genotype_class_severity.R")
-knitr::opts_knit$set(root.dir = home_dir)
+variables = c("CFF ID","Site","Sex","Race","Genotypes1","Genotypes2","Pancreatic Status",
+              "G-tube in past 12 months","Start","Date","Age")
 ```
@@
 ```{r include=FALSE}
-# Import cleaned data from Casey
-df = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/analysis_data.csv",na.strings = "NA")
-df$X = NULL
-# Fill in MRN and CFF ID for everyone
-df = df %>% 
-  group_by(CFF.ID) %>% fill(MRN,.direction = "downup") %>% 
-  group_by(MRN) %>% fill(CFF.ID,.direction = "downup")
-# Assume that blanks for pancreatic and G tube indicate "no"
-df$Hispanic.Latinx[is.na(df$Hispanic.Latinx)] = FALSE
-df$Pancreatic.Status[is.na(df$Pancreatic.Status)] = "Sufficient"
-df$G.tube.in.past.12.months[is.na(df$G.tube.in.past.12.months)] = "No"
-# Convert dates
-date_vars = c("DOB","Start","Date")
-df[,date_vars] = lapply(df[,date_vars], function(d){
-  parse_date_time(d,orders = c("mdy","ymd","dmy"),quiet = T)
-})
-# Fix incorrect century (WHY LUBRIDATE WHY?!)
-df$DOB[which(df$DOB > lubridate::mdy("1/1/2022"))] = 
-  df$DOB[which(df$DOB > lubridate::mdy("1/1/2022"))] - years(100)
-# Time from start and filtering
-df = df %>% group_by(CFF.ID) %>% fill(Start,.direction = "downup")
-df = df %>% mutate(Time = as.numeric(difftime(Date,Start,units = "days"))) 
-# Add UM (they did not provide dates, so need Time column)
-um_a1cs = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_a1cs.csv",na.strings = "")
-um_bmi = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_bmi.csv",na.strings = "")
-um_demo = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_demo.csv",na.strings = "")
-um_ogtt = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_ogtt.csv",na.strings = "")
-um_pfts = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_pfts.csv",na.strings = "")
-um = full_join(um_a1cs,um_bmi)
-um = full_join(um,um_ogtt)
-um = full_join(um,um_pfts)
-um = left_join(um,um_demo)
-um = um %>% filter(!is.na(CFRD.Status))
-um$Age = um$Age / 365.25
-df = full_join(df,um)
-# Filter and fill constants
-df = df %>% group_by(CFF.ID) %>% fill(DOB,Start:CFRD.Diagnosis.Date,.direction = "downup") %>%
-  ungroup() %>% filter(!is.na(Time),abs(Time) <= 730,!is.na(CFF.ID),
-                       (!is.na(CFRD.Diagnosis.Date) | !is.na(CFRD.Status) | site == "UW")) %>%
-  select(-MRN) %>% select(CFF.ID,site,DOB,Start,everything()) %>% rename(Site = "site")
-df$Sex = as.factor(df$Sex)
-levels(df$Sex) = c("F","M")
-# BMI category
-df$Age = pmax(as.numeric(difftime(df$Date,df$DOB,units = "days"))/365.25,df$Age,na.rm = T)
-# Treat adults as age = 20
-df$Age_for_BMI = ifelse(df$Age > 20,20,df$Age)
-df$BMIz = y2z(df$BMI,x = df$Age_for_BMI,sex = df$Sex,ref=nl4.bmi)
-df$BMIz[is.infinite(df$BMIz)] = NA
-df$BMI.Percentile = pnorm(df$BMIz)*100
-df$Weightz = y2z(df$Weight,x = df$Age_for_BMI,sex = df$Sex,ref=nl4.wgt)
-df$BMI_group = apply(df,1,function(r){
-  age = as.numeric(r["Age"])
-  if(!is.na(age) & age > 19){
-    bmi = as.numeric(r["BMI"])
-    group = cut(bmi,c(-Inf,18.5,25,30,35,40,Inf),
-        labels = c("Under Weight","Normal Weight","Overweight","Obese (Class I)",
-                   "Obese (Class II)","Obese (Class III+)"),right = F)
-  } else if (!is.na(age) & age <= 19){
-    z = as.numeric(r["BMIz"])
-    perc = pnorm(z)
-    group = cut(perc,c(-Inf,5,85,95,Inf),
-        labels = c("Under Weight","Normal Weight","Overweight","Obese (Class I)"),right = F)
-  } else {group = NA}
-  return(as.character(group))
-})
-df$BMI_group = as.factor(df$BMI_group)
-df$BMI_group = ordered(df$BMI_group,c("Under Weight","Normal Weight","Overweight",
-                                      "Obese (Class I)","Obese (Class II)","Obese (Class III+)"))
-# Get genotype severity
-df$Genotype1_sev = get_cf_genotype_class_severity(df$Genotypes1)
-df$Genotype2_sev = get_cf_genotype_class_severity(df$Genotypes2)
-# PFTs
-old <- c("Caucasian","Black","Other")
-new <- c(1,2,5)
-df$ppFEV = pctpred_GLI(df$Age,df$Height/100,gender = as.numeric(df$Sex),
-            ethnicity = new[match(df$Race, old)],
-            FEV1 = df$FEV1)
-df$ppFVC = pctpred_GLI(df$Age,df$Height/100,gender = as.numeric(df$Sex),
-            ethnicity = new[match(df$Race, old)],
-            FEV1 = df$FVC)
-# T1D duration
-df$CFRD.Diagnosis.Date = mdy(df$CFRD.Diagnosis.Date)
-df$T1D.Duration = as.numeric(difftime(df$Date,df$CFRD.Diagnosis.Date,units = "days"))/365.25
-# Before or after start
-df$Before_After = cut(df$Time,c(-Inf,0,Inf),labels = c("Before","After"),right = T)
-# Sort
-df = df %>% arrange(Site,CFF.ID,Date) %>% distinct()
-# Write sorted DF
-write.csv(df, file = "./Colleen Wood/Prepost Triakfta/Data_Cleaned/analysis_data_with_UM.csv",
-          row.names = F,na = "")
+# CHCO
+chco_a1c = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_glycemic.csv",show_col_types = F)
+chco_a1c$Date = parse_date_time(chco_a1c$Date,orders = "mdy")
+chco_bmi = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv",show_col_types = F)
+chco_bmi$Date = parse_date_time(chco_bmi$Date,orders = "mdy")
+chco = full_join(chco_a1c,chco_bmi,by = c("MRN", "Date"))
+# Fill CFF ID by MRN
+chco= chco %>% group_by(MRN) %>% 
+  fill(`CFF ID`,.direction = "downup") %>% ungroup() %>%
+  filter(!is.na(`CFF ID`)) %>%
+  select(-MRN)
+# Add demographics
+chco_reg = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_registry.csv",show_col_types = F)
+chco_reg = chco_reg %>% select(-MRN)
+chco_reg$Start = parse_date_time(chco_reg$Start,orders = "mdy")
+chco = left_join(chco,chco_reg,by = "CFF ID")
+# Add Time
+chco$Time = as.numeric(difftime(chco$Date,chco$Start,units = "days"))
 ```
 
-```{r}
-# Most recent before/after - limit to 2 years before start
-# Want the value closest to start date before start, and most recent with respect to now from start
-# (i.e. the longest time after start)
-most_recent = df %>% group_by(CFF.ID,Site,Before_After) %>%
-  filter(Time > -2*365) %>%
-  mutate(time_new = max(Time) - Time) %>%
-  arrange(time_new) %>% 
-  summarise(Age = paste0(Date[which(Age == first(na.omit(Age)))],": ",round(first(na.omit(Age)),1)),
-            BMI.Percentile = paste0(Date[which(BMI.Percentile == first(na.omit(BMI.Percentile)))],
-                                    ": ",round(first(na.omit(BMI.Percentile)),1)),
-            Weight = paste0(Date[which(Weight == round(first(na.omit(Weight)),1))],
-                            ": ",first(na.omit(Weight))),
-            T1D.Duration = paste0(Date[which(T1D.Duration == first(na.omit(T1D.Duration)))],
-                                  ": ",round(first(na.omit(T1D.Duration)),1)),
-            G.tube.in.past.12.months = first(na.omit(G.tube.in.past.12.months)),
-            Pancreatic.Status = first(na.omit(Pancreatic.Status)),
-            Genotype1_sev = first(na.omit(Genotype1_sev)),
-            Genotype2_sev = first(na.omit(Genotype2_sev)),
-            BMIz = paste0(Date[which(BMIz == first(na.omit(BMIz)))],
-                          ": ",round(first(na.omit(BMIz)),1)),
-            FEV1 = paste0(Date[which(FEV1 == first(na.omit(FEV1)))[1]],
-                          ": ",round(first(na.omit(FEV1)),1)),
-            FVC = paste0(Date[which(FVC == first(na.omit(FVC)))],
-                         ": ",round(first(na.omit(FVC)),1)),
-            HbA1c = paste0(Date[which(HbA1c == first(na.omit(HbA1c)))],
-                           ": ",round(first(na.omit(HbA1c)),1)),
-            .groups = "keep") %>%
-  arrange(Site,CFF.ID,Before_After)
-most_recent[most_recent == ": NA"] = NA
-most_recent[most_recent == "NA: NA"] = NA
+```{r include=FALSE}
+# Montana
+mon_bmi = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_bmi_pft.csv",show_col_types = F)
+mon_bmi$Date = parse_date_time(mon_bmi$Date,orders = "mdy")
+mon_a1c = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv",show_col_types = F)
+mon_a1c$Date = parse_date_time(mon_a1c$Date,orders = "mdy")
+# Get start date and CFRD status
+mon_starts = mon_a1c %>% select(`CFF ID`,`CFRD Status`,Start) %>% 
+  distinct() %>% filter(`CFRD Status`=="CFRD with or without fasting hyperglycemia")
+mon_starts$Start = parse_date_time(mon_starts$Start,orders = "mdy")
+# Remove unnecessary columns
+mon_a1c = mon_a1c %>% select(`CFF ID`,Date,HbA1c,`OGTT Fasting`,`OGTT Two Hour`)
+# Merge and remove those without CFRD or start date
+mon = full_join(mon_bmi,mon_a1c,by = c("CFF ID", "Date"))
+mon= mon %>% filter(`CFF ID` %in% mon_starts$`CFF ID`)
+mon = left_join(mon,mon_starts %>% select(`CFF ID`,Start), by = "CFF ID")
+# Add Time
+mon$Time = as.numeric(difftime(mon$Date,mon$Start,units = "days"))
+```
 
-# # Check
-# t = df %>% select(CFF.ID,Site,Date,Time,Before_After,BMIz,ppFEV,ppFVC) %>%
-#   group_by(CFF.ID,Site,Before_After) %>%
-#   arrange(Site,CFF.ID,Before_After)
-# Filter those without CFRD, and those without pre AND post
-keep = most_recent %>% filter(!is.na(HbA1c)) %>%
-  group_by(CFF.ID) %>% summarise(n = n()) %>%
-  filter(n == 2) %>% pull(CFF.ID)
-cfrd = unique(df$CFF.ID[!is.na(df$CFRD.Status) | !is.na(df$CFRD.Diagnosis.Date)])
-keep = unique(keep,cfrd)
-most_recent = most_recent %>% filter(CFF.ID %in% keep)
+```{r include=FALSE}
+# UM
+um_a1c = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_a1cs.csv",show_col_types = F)
+um_a1c$Age = round(um_a1c$Age / 365.25,1)
+um_bmi = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_bmi.csv",show_col_types = F)
+um_bmi$Age = round(um_bmi$Age / 365.25,1)
+um_ogtt = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_ogtt.csv",show_col_types = F)
+um_ogtt$Age = round(um_ogtt$Age / 365.25,1)
+um_pft = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_pfts.csv",show_col_types = F)
+um_pft$Age = round(um_pft$Age / 365.25,1)
+um_race = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_race.csv",show_col_types = F)
+# Join
+um = reduce(list(um_a1c,um_bmi,um_ogtt,um_pft),full_join)
+um = left_join(um,um_race,by = "CFF ID")
+```
+
+```{r include=FALSE}
+# UW
+uw_bmi = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_bmi.csv",show_col_types = F)
+uw_bmi$Date = parse_date_time(uw_bmi$Date,orders = "mdy")
+uw_glyc = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_glycemic.csv",show_col_types = F)
+uw_glyc$Date = parse_date_time(uw_glyc$Date,orders = "mdy")
+uw_pft = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pft.csv",show_col_types = F)
+uw_pft$Date = parse_date_time(uw_pft$Date,orders = "mdy")
+uw_reg = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_registry.csv",show_col_types = F)
+uw_reg$Start = parse_date_time(uw_reg$Start,orders = "mdy")
+# Merge
+uw = reduce(list(uw_bmi,uw_glyc,uw_pft,uw_reg),full_join)
+```
+
+```{r include=FALSE}
+# Put everything together
+
 ```
@@
 # Participant Characteristics
+
+```{r results='asis'}
+t1 = tableby(~Age+BMI.Percentile+Weight+Sex+,pre_post[pre_post$Before_After == "Before",])
+summary(t1)
+```
 
@@
 ```
-
-# Piecewise Models
-
-```{r}
-# Piecewise model function
-piecewise = function(outcome = "HbA1c",data = df,simple = F){
-  # Add spline column
-  data$Change = ifelse(data$Time < 0,0,1)
-  data$Change = data$Change * data$Time
-  # Model
-  f = as.formula(paste0(outcome,"~Time + Change + (1|CFF.ID)"))
-  mod = lmer(f,data = data)
-  res = tidy(mod,"fixed")
-  res$effect = NULL
-  # Model check if necessary
-  check = check_model(mod)
-  # Coordinates for model plotting
-  x1_start = -365
-  y1_start = res$estimate[1] + res$estimate[2]*x1_start
-  x1_end = 0
-  y1_end = res$estimate[1] 
-  x2_start = 0
-  y2_start = res$estimate[1] 
-  x2_end = 365
-  y2_end = res$estimate[1] + (res$estimate[2]+res$estimate[3])*x2_end
-  # Plot
-  p = ggplot(data,aes_string(x = "Time",y = outcome,group = "CFF.ID",color = "Site")) +
-    geom_point(alpha = 0.5) + geom_line(alpha = 0.2) +
-    annotate("segment",x=x1_start,y=y1_start,xend=x1_end,yend=y1_end,color = "blue",size = 1) +
-    annotate("segment",x=x2_start,y=y2_start,xend=x2_end,yend=y2_end,color = "blue",size = 1) +
-    xlab("Days from Trikafta Start") + theme_bw()
-  # Interactive plot and model results
-  return(list(plot = ggplotly(p),table = res,check = check))
-}
-```
-
-# Methods
-
-For each outcome, we fit a piecewise linear mixed effects model with random intercept for each person and a "knot" at 0.
-This approach can be thought of as fitting two separate linear models, while ensuring that the two regression lines meet at Trikafta start.
-In the model results, "Slope Prior" indicates the slope prior to start of treatment (i.e.
-the average change in outcome for each "1 unit increase" in time).
-The "Change" term indicates the *change in slope* after medication start.
-To find the slope post-Trikafta, you add the "Slope Prior" and "Change" estimates.
-
-The random intercept model was selected because it performed better (by AIC) than a model with random intercept and continuous AR1 correlation structure.
-Random slope models did not converge for some outcomes, so were not used for the sake of consistency between results.
-
-# HbA1c
-
-```{r results='asis'}
-a1c = piecewise(outcome = "HbA1c",data = df)
-ggplotly(a1c$plot)
-kable(a1c$table,digits = 3)
-```
-
-`r if(class(a1c$seg)=="lme"){"Automated R package was not able to find a better breakpoint."}`
-
-# BMI
-
-Models did not converge with BMI converted to under weight, normal weight, etc.
-This is most likely due to lack of within-patient variability in terms of BMI group (i.e.
-people's BMIs aren't changing enough that they move from one group to another).
-
-# Insulin
-
-Most participants only have one observation of insulin dose, so these results would be the same as Maxie's previous reports.
-
-# ppFEV
-
-```{r results='asis'}
-ppFEV = piecewise(outcome = "ppFEV",data = df)
-ggplotly(ppFEV$plot)
-kable(ppFEV$table,digits = 3)
-```
-
-# ppFVC
-
-```{r results='asis'}
-ppFVC = piecewise(outcome = "ppFVC",data = df)
-ggplotly(ppFVC$plot)
-kable(ppFVC$table,digits = 3)
-```
-
-# CGM Metrics
-
-## Mean
-
-```{r results='asis'}
-meansg = piecewise(outcome = "average_sensor",data = df)
-ggplotly(meansg$plot)
-kable(meansg$table,digits = 3)
-```
-
-## Min
-
-```{r results='asis'}
-minsg = piecewise(outcome = "min_sensor",data = df)
-ggplotly(minsg$plot)
-kable(minsg$table,digits = 3)
-```
-
-## Max
-
-```{r results='asis'}
-maxsg = piecewise(outcome = "max_sensor",data = df)
-ggplotly(maxsg$plot)
-kable(maxsg$table,digits = 3)
-```
-
-## % Time > 120
-
-```{r results='asis'}
-over120 = piecewise(outcome = "percent_time_over_120",data = df)
-ggplotly(over120$plot)
-kable(over120$table,digits = 3)
-```
-
-## % Time > 140
-
-```{r results='asis'}
-over140 = piecewise(outcome = "percent_time_over_140",data = df)
-ggplotly(over140$plot)
-kable(over140$table,digits = 3)
-```
-
-## % Time > 200
-
-```{r results='asis'}
-over200 = piecewise(outcome = "percent_time_over_200",data = df)
-ggplotly(over200$plot)
-kable(over200$table,digits = 3)
-```
-
-## Mean excursions > 140
-
-```{r results='asis'}
-numover140 = piecewise(outcome = "avg_excur_over_140_per_day",data = df)
-ggplotly(numover140$plot)
-kable(numover140$table,digits = 3)
-```
-
-## Mean excursions > 200
-
-```{r results='asis'}
-numover200 = piecewise(outcome = "avg_excur_over_200_per_day",data = df)
-ggplotly(numover200$plot)
-kable(numover200$table,digits = 3)
-```
-
-## % Time < 70
-
-```{r results='asis'}
-under70 = piecewise(outcome = "percent_time_under_70",data = df)
-ggplotly(under70$plot)
-kable(under70$table,digits = 3)
-```
-
-## % Time < 60
-
-```{r results='asis'}
-under60 = piecewise(outcome = "percent_time_under_60",data = df)
-ggplotly(under60$plot)
-kable(under60$table,digits = 3)
-```
-
-## Average AUC per day
-
-```{r results='asis'}
-auc = piecewise(outcome = "average_auc_per_day",data = df)
-ggplotly(auc$plot)
-kable(auc$table,digits = 3)
-```
-
-## SD
-
-```{r results='asis'}
-sd = piecewise(outcome = "standard_deviation",data = df)
-ggplotly(sd$plot)
-kable(sd$table,digits = 3)
-```
-
-## MAGE
-
-```{r results='asis'}
-mage = piecewise(outcome = "r_mage",data = df)
-ggplotly(mage$plot)
-kable(mage$table,digits = 3)
-```

### 2022-10-06 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
Update pre_post_trikafta_analysis.Rmd

@@
 library(lubridate)
+library(rspiro)
 source("~/GitHub/CF/get_genotype_class_severity.R")
-variables = c("CFF ID","Site","Sex","Race","Genotypes1","Genotypes2","Pancreatic Status",
-              "G-tube in past 12 months","Start","Date","Age")
+t1_variables = c("CFF ID","Site","Sex","Race","Genotypes1","Genotypes2","Pancreatic Status",
+              "G-tube in past 12 months","Time","Age","BMI","Height","Weight","FEV1","FVC")
 ```
@@
 chco_bmi$Date = parse_date_time(chco_bmi$Date,orders = "mdy")
-chco = full_join(chco_a1c,chco_bmi,by = c("MRN", "Date"))
+chco = full_join(chco_a1c,chco_bmi)
 # Fill CFF ID by MRN
@@
 chco_reg = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_registry.csv",show_col_types = F)
-chco_reg = chco_reg %>% select(-MRN)
 chco_reg$Start = parse_date_time(chco_reg$Start,orders = "mdy")
-chco = left_join(chco,chco_reg,by = "CFF ID")
-# Add Time
+chco_reg$DOB = parse_date_time(chco_reg$DOB,orders = "mdy")
+keep = chco_reg %>% filter(!is.na(`CFRD Diagnosis Date`)) %>% pull(`CFF ID`)
+exclude = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv",show_col_types = F)
+exclude = exclude %>% filter(!is.na(`Remove (see notes)`)) %>% pull(MRN)
+exclude = chco_reg$`CFF ID`[match(exclude,chco_reg$MRN)]
+keep =setdiff(keep,exclude)
+chco = left_join(chco,chco_reg) %>% filter(`CFF ID` %in% keep)
+# Add missing variables
+chco$Age = round(as.numeric(difftime(chco$Date,chco$DOB,units = "days"))/365.25,1)
 chco$Time = as.numeric(difftime(chco$Date,chco$Start,units = "days"))
+chco$Site = "CHCO"
 ```
@@
 mon_a1c$Date = parse_date_time(mon_a1c$Date,orders = "mdy")
+mon_a1c$DOB = parse_date_time(mon_a1c$DOB,orders = "mdy")
 # Get start date and CFRD status
@@
 # Remove unnecessary columns
-mon_a1c = mon_a1c %>% select(`CFF ID`,Date,HbA1c,`OGTT Fasting`,`OGTT Two Hour`)
+mon_a1c = mon_a1c %>% select(`CFF ID`,DOB,Date,HbA1c,`OGTT Fasting`,`OGTT Two Hour`)
 # Merge and remove those without CFRD or start date
-mon = full_join(mon_bmi,mon_a1c,by = c("CFF ID", "Date"))
+mon = full_join(mon_bmi,mon_a1c)
 mon= mon %>% filter(`CFF ID` %in% mon_starts$`CFF ID`)
-mon = left_join(mon,mon_starts %>% select(`CFF ID`,Start), by = "CFF ID")
-# Add Time
+mon = left_join(mon,mon_starts %>% select(`CFF ID`,Start))
+# Add demographics
+mon_demo = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_demo.csv",show_col_types = F)
+mon_demo = mon_demo %>% select(`CFF ID`,Race,Sex)
+mon = left_join(mon,mon_demo)
+# Add genotypes
+mon_geno = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_genotypes.csv",show_col_types = F)
+mon = left_join(mon,mon_geno)
+# Add pancreatic and g tube status
+mon_pan = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_pancreatic_insufficient.csv",show_col_types = F)
+mon_pan = mon_pan %>% select(-Date) %>% distinct()
+mon_pan$`Pancreatic Status` = "Insufficient"
+mon_gtube = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_gtubes.csv",show_col_types = F)
+mon_gtube = mon_gtube %>% select(-Feeding,-Date) %>% distinct()
+mon_gtube$`G-tube in past 12 months` = "Yes"
+mon = left_join(mon,mon_pan)
+mon = left_join(mon,mon_gtube)
+# Add missing variables
+mon = mon %>% group_by(`CFF ID`) %>% fill(Start,DOB,.direction = "downup")
 mon$Time = as.numeric(difftime(mon$Date,mon$Start,units = "days"))
+mon$Age = round(as.numeric(difftime(mon$Date,mon$DOB))/365.25,1)
+mon$Site = "Montana"
 ```
@@
 um_race = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_race.csv",show_col_types = F)
+um_demo = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_demo.csv",show_col_types = F)
+keep = um_demo %>% 
+  filter(`CFRD Status` =="CFRD with or without fasting hyperglycemia") %>%
+  pull(`CFF ID`)
+um_demo = um_demo %>% select(-`CFRD Status`) %>% 
+  rename(`G-tube in past 12 months`=`G tube in past 12 months`)
 # Join
 um = reduce(list(um_a1c,um_bmi,um_ogtt,um_pft),full_join)
-um = left_join(um,um_race,by = "CFF ID")
+um = left_join(um,um_race)
+um = left_join(um,um_demo)
+um = um %>% filter(`CFF ID` %in% keep) %>% rename(Site = "site")
 ```
@@
 uw_pft$Date = parse_date_time(uw_pft$Date,orders = "mdy")
-uw_reg = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_registry.csv",show_col_types = F)
-uw_reg$Start = parse_date_time(uw_reg$Start,orders = "mdy")
+uw_demo = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_demo.csv",show_col_types = F)
+uw_demo = uw_demo %>% select(`CFF ID`,Race,Sex)
+# Get CFRD status and start date
+uw_start = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_mods.csv",show_col_types = F)
+uw_start$`Date of Encounter` = parse_date_time(uw_start$`Date of Encounter`,orders = "mdy")
+uw_start = uw_start %>% arrange(`CFF ID`,`Date of Encounter`) %>% 
+  group_by(`CFF ID`) %>% filter(row_number()==n()) %>%
+  select(`CFF ID`,`Date of Encounter`) %>% rename(Start = `Date of Encounter`)
+uw_cfrd = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_cfrd.csv",show_col_types = F)
+uw_cfrd = uw_cfrd %>% 
+  filter(`CFRD Status`=="CFRD with or without fasting hyperglycemia") %>% 
+  select(`CFF ID`) %>% distinct() %>% pull(`CFF ID`)
+keep = unique(uw_cfrd,uw_start$`CFF ID`)
+# Add genotypes
+uw_geno = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_genotypes.csv",show_col_types = F)
+# Pancreatic status
+uw_pan = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pancreatic.csv",show_col_types = F)
+uw_pan = uw_pan %>% select(-Date) %>% distinct() %>% mutate(`Pancreatic Status`="Insufficient")
+# G tubes
+uw_gtube = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_gtubes.csv",show_col_types = F)
+uw_gtube = uw_gtube %>% select(`CFF ID`) %>% distinct() %>% mutate(`G-tube in past 12 months`="Yes")
 # Merge
-uw = reduce(list(uw_bmi,uw_glyc,uw_pft,uw_reg),full_join)
+uw = reduce(list(uw_bmi,uw_glyc,uw_pft),full_join)
+uw = left_join(uw,uw_demo)
+uw = left_join(uw,uw_geno)
+uw = left_join(uw,uw_pan)
+uw = left_join(uw,uw_gtube)
+uw = left_join(uw,uw_start) %>% filter(`CFF ID` %in% keep)
+# Other variables (no weight for this site)
+uw$Time = as.numeric(difftime(uw$Date,uw$Start,units = "days"))
+uw$Site = "UW"
+uw$Weight = NA
 ```
@@
 # Put everything together
-
+chco = chco %>% select(all_of(t1_variables))
+mon = mon %>% select(all_of(t1_variables))
+um = um %>% select(all_of(t1_variables))
+uw = uw %>% select(all_of(t1_variables))
+df = bind_rows(chco,mon,um,uw) %>% arrange(`CFF ID`,Time)
+# Age group
+df$`Age Group` = cut(df$Age,breaks = c(-Inf,20,Inf),labels = c("Pedatric","Adult"),right = F)
+# Race and sex
+df$Race = factor(df$Race,levels = c("Black or African American","Caucasian",
+                                    "Other","Some other race","Two or more races","White"),
+                 labels = c("Black","Caucasian",
+                                    "Other","Other","Other","Caucasian"))
+df$Sex = factor(df$Sex,levels = c("Male","Female"))
+# BMI and FEV calculations
+age = df$Age
+height = df$Height/100
+race = df$Race
+old <- c("Caucasian","Black","Other")
+new <- c(1,2,5)
+race = new[match(race, old)]
+gender = as.numeric(df$Sex)
+df$ppFEV1  = pctpred_GLI(age,height,gender = gender,ethnicity = race,FEV1 = df$FEV1)
+df$ppFVC  = pctpred_GLI(age,height,gender = gender,ethnicity = race,FVC = df$FVC)
+# Genotype severity
+df$Genotypes1 = get_cf_genotype_class_severity(df$Genotypes1)
+df$Genotypes2 = get_cf_genotype_class_severity(df$Genotypes2)
+df = df %>% unite(Genotype,Genotypes1,Genotypes2)
+df$Genotype = factor(df$Genotype,levels = c("Mild_Severe","Severe_Mild","Severe_Severe",
+                                            "Severe_Unknown","Unknown_Unknown"),
+                     labels = c("Low","Low","High","Unknown","Unknown"))
 ```
 
-# Participant Characteristics
+# Participant characteristics at date closest to CFTR modulator start
 
 ```{r results='asis'}
+t1_data = df %>% group_by(`CFF ID`) %>% slice_min(abs(Time))
+
 t1 = tableby(~Age+BMI.Percentile+Weight+Sex+,pre_post[pre_post$Before_After == "Before",])

### 2022-10-06 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
a1c data cleaned and pre/post

--- 2022-10-04 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
@@
 library(lubridate)
+library(childsds)
+library(rspiro)
 source("~/GitHub/CF/get_genotype_class_severity.R")
-variables = c("CFF ID","Site","Sex","Race","Genotypes1","Genotypes2","Pancreatic Status",
-              "G-tube in past 12 months","Start","Date","Age")
+t1_variables = c("CFF ID","Site","Sex","Race","Genotypes1","Genotypes2","Pancreatic Status",
+              "G-tube in past 12 months","Time","Age",
+              "BMI Percentile","Height","Weight","FEV1","FVC","HbA1c")
 ```
@@
 chco_reg = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_registry.csv",show_col_types = F)
-chco_reg = chco_reg %>% select(-MRN)
 chco_reg$Start = parse_date_time(chco_reg$Start,orders = "mdy")
-chco = left_join(chco,chco_reg,by = "CFF ID")
-# Add Time
+chco_reg$DOB = parse_date_time(chco_reg$DOB,orders = "mdy")
+keep = chco_reg %>% filter(!is.na(`CFRD Diagnosis Date`)) %>% pull(`CFF ID`)
+exclude = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv",show_col_types = F)
+exclude = exclude %>% filter(!is.na(`Remove (see notes)`)) %>% pull(MRN)
+exclude = chco_reg$`CFF ID`[match(exclude,chco_reg$MRN)]
+keep =setdiff(keep,exclude)
+chco = left_join(chco,chco_reg) %>% filter(`CFF ID` %in% keep)
+# Add missing variables
+chco$Age = round(as.numeric(difftime(chco$Date,chco$DOB,units = "days"))/365.25,1)
 chco$Time = as.numeric(difftime(chco$Date,chco$Start,units = "days"))
+chco$Site = "CHCO"
+# BMI calculations
+chco$age_for_bmi = ifelse(chco$Age < 20,chco$Age,20)
+chco$`BMI Percentile` = round(sds(chco$BMI,age = chco$age_for_bmi,item = "bmi",type = "perc",
+                          sex = chco$Sex,male = "Male", female = "Female",
+                          ref = cdc.ref)*100,2)
 ```
@@
 mon_a1c$Date = parse_date_time(mon_a1c$Date,orders = "mdy")
+mon_a1c$DOB = parse_date_time(mon_a1c$DOB,orders = "mdy")
+mon_a1c$HbA1c = as.numeric(mon_a1c$HbA1c)
 # Get start date and CFRD status
@@
 # Remove unnecessary columns
-mon_a1c = mon_a1c %>% select(`CFF ID`,Date,HbA1c,`OGTT Fasting`,`OGTT Two Hour`)
+mon_a1c = mon_a1c %>% select(`CFF ID`,DOB,Date,HbA1c,`OGTT Fasting`,`OGTT Two Hour`)
 # Merge and remove those without CFRD or start date
-mon = full_join(mon_bmi,mon_a1c,by = c("CFF ID", "Date"))
+mon = full_join(mon_bmi,mon_a1c)
 mon= mon %>% filter(`CFF ID` %in% mon_starts$`CFF ID`)
-mon = left_join(mon,mon_starts %>% select(`CFF ID`,Start), by = "CFF ID")
-# Add Time
+mon = left_join(mon,mon_starts %>% select(`CFF ID`,Start))
+# Add demographics
+mon_demo = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_demo.csv",show_col_types = F)
+mon_demo = mon_demo %>% select(`CFF ID`,Race,Sex)
+mon = left_join(mon,mon_demo)
+# Add genotypes
+mon_geno = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_genotypes.csv",show_col_types = F)
+mon = left_join(mon,mon_geno)
+# Add pancreatic and g tube status
+mon_pan = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_pancreatic_insufficient.csv",show_col_types = F)
+mon_pan = mon_pan %>% select(-Date) %>% distinct()
+mon_pan$`Pancreatic Status` = "Insufficient"
+mon_gtube = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_gtubes.csv",show_col_types = F)
+mon_gtube = mon_gtube %>% select(-Feeding,-Date) %>% distinct()
+mon_gtube$`G-tube in past 12 months` = "Yes"
+mon = left_join(mon,mon_pan)
+mon = left_join(mon,mon_gtube)
+# Add missing variables
+mon = mon %>% group_by(`CFF ID`) %>% fill(Start,DOB,.direction = "downup")
 mon$Time = as.numeric(difftime(mon$Date,mon$Start,units = "days"))
+mon$Age = round(as.numeric(difftime(mon$Date,mon$DOB))/365.25,1)
+mon$Site = "Montana"
+# Fill in some missing 
+mon$`Pancreatic Status`[is.na(mon$`Pancreatic Status`)] = "Sufficient"
+# BMI calculations
+mon$age_for_bmi = ifelse(mon$Age < 20,mon$Age,20)
+mon$`BMI Percentile` = round(sds(mon$BMI,age = mon$age_for_bmi,item = "bmi",type = "perc",
+                          sex = mon$Sex,male = "Male", female = "Female",
+                          ref = cdc.ref)*100,2)
 ```
@@
 um_race = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_race.csv",show_col_types = F)
+um_demo = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_demo.csv",show_col_types = F)
+keep = um_demo %>% 
+  filter(`CFRD Status` =="CFRD with or without fasting hyperglycemia") %>%
+  pull(`CFF ID`)
+um_demo = um_demo %>% select(-`CFRD Status`) %>% 
+  rename(`G-tube in past 12 months`=`G tube in past 12 months`)
 # Join
 um = reduce(list(um_a1c,um_bmi,um_ogtt,um_pft),full_join)
-um = left_join(um,um_race,by = "CFF ID")
+um = left_join(um,um_race)
+um = left_join(um,um_demo)
+um = um %>% filter(`CFF ID` %in% keep,!is.na(Time),abs(Time) <= 365) %>% 
+  rename(Site = "site") %>% arrange(`CFF ID`,Time)
+# Filter per Maxie's previous work (unclear exactly how they got to this subset)
+um_ids = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_ids_from_maxie.csv",show_col_types = F)
+um = um %>% filter(`CFF ID` %in% um_ids$`CFF ID`)
+# BMI calculations
+um$age_for_bmi = ifelse(um$Age < 20,um$Age,20)
+um$`BMI Percentile` = round(sds(um$BMI,age = um$age_for_bmi,item = "bmi",type = "perc",
+                          sex = um$Sex,male = "Male", female = "Female",
+                          ref = cdc.ref)*100,2)
 ```
@@
 uw_pft$Date = parse_date_time(uw_pft$Date,orders = "mdy")
-uw_reg = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_registry.csv",show_col_types = F)
-uw_reg$Start = parse_date_time(uw_reg$Start,orders = "mdy")
+uw_demo = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_demo.csv",show_col_types = F)
+uw_demo = uw_demo %>% select(`CFF ID`,Race,Sex)
+# Get CFRD status and start date
+uw_start = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_mods.csv",show_col_types = F)
+uw_start$`Date of Encounter` = parse_date_time(uw_start$`Date of Encounter`,orders = "mdy")
+uw_start = uw_start %>% arrange(`CFF ID`,`Date of Encounter`) %>% 
+  group_by(`CFF ID`) %>% filter(row_number()==n()) %>%
+  select(`CFF ID`,`Date of Encounter`) %>% rename(Start = `Date of Encounter`)
+uw_cfrd = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_cfrd.csv",show_col_types = F)
+uw_cfrd = uw_cfrd %>% 
+  filter(`CFRD Status`=="CFRD with or without fasting hyperglycemia") %>% 
+  select(`CFF ID`) %>% distinct() %>% pull(`CFF ID`)
+keep = unique(uw_cfrd,uw_start$`CFF ID`)
+# Add genotypes
+uw_geno = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_genotypes.csv",show_col_types = F)
+# Pancreatic status
+uw_pan = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pancreatic.csv",show_col_types = F)
+uw_pan = uw_pan %>% select(-Date) %>% distinct() %>% mutate(`Pancreatic Status`="Insufficient")
+# G tubes
+uw_gtube = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_gtubes.csv",show_col_types = F)
+uw_gtube = uw_gtube %>% select(`CFF ID`) %>% distinct() %>% mutate(`G-tube in past 12 months`="Yes")
 # Merge
-uw = reduce(list(uw_bmi,uw_glyc,uw_pft,uw_reg),full_join)
+uw = reduce(list(uw_bmi,uw_glyc,uw_pft),full_join)
+uw = left_join(uw,uw_demo)
+uw = left_join(uw,uw_geno)
+uw = left_join(uw,uw_pan)
+uw = left_join(uw,uw_gtube)
+uw = left_join(uw,uw_start) %>% filter(`CFF ID` %in% keep)
+# Other variables (no weight for this site)
+uw$Time = as.numeric(difftime(uw$Date,uw$Start,units = "days"))
+uw$Site = "UW"
+uw$Weight = NA
+# BMI calculations
+uw$age_for_bmi = ifelse(uw$Age < 20,uw$Age,20)
+uw$`BMI Percentile 2` = round(sds(uw$BMI,age = uw$age_for_bmi,item = "bmi",type = "perc",
+                          sex = uw$Sex,male = "Male", female = "Female",
+                          ref = cdc.ref)*100,2)
+uw$`BMI Percentile` = coalesce(uw$`BMI Percentile`,uw$`BMI Percentile 2`)
 ```
@@
 # Put everything together
-
+chco = chco %>% select(all_of(t1_variables))
+mon = mon %>% select(all_of(t1_variables))
+um = um %>% select(all_of(t1_variables))
+uw = uw %>% select(all_of(t1_variables))
+df = bind_rows(chco,mon,um,uw) %>% arrange(`CFF ID`,Time)
+# Age group
+df$`Age Group` = cut(df$Age,breaks = c(-Inf,20,Inf),labels = c("Pedatric","Adult"),right = F)
+# Race and sex
+df$Race = factor(df$Race,levels = c("Black or African American","Caucasian",
+                                    "Other","Some other race","Two or more races","White"),
+                 labels = c("Black","Caucasian",
+                                    "Other","Other","Other","Caucasian"))
+df$Sex = factor(df$Sex,levels = c("Male","Female"))
+# FEV calculations
+age = df$Age
+height = df$Height/100
+race = df$Race
+old <- c("Caucasian","Black","Other")
+new <- c(1,2,5)
+race = new[match(race, old)]
+gender = as.numeric(df$Sex)
+df$ppFEV1  = pctpred_GLI(age,height,gender = gender,ethnicity = race,FEV1 = df$FEV1)
+df$ppFVC  = pctpred_GLI(age,height,gender = gender,ethnicity = race,FVC = df$FVC)
+# Genotype severity
+df$Genotypes1 = get_cf_genotype_class_severity(df$Genotypes1)
+df$Genotypes2 = get_cf_genotype_class_severity(df$Genotypes2)
+df = df %>% unite(Genotype,Genotypes1,Genotypes2)
+df$Genotype = factor(df$Genotype,levels = c("Mild_Severe","Severe_Mild","Severe_Severe",
+                                            "Severe_Unknown","Unknown_Unknown"),
+                     labels = c("Low","Low","High","Unknown","Unknown"))
+# Other little cleaning things
+df = df %>% filter(Age > 0,!is.na(Time))
+df$`G-tube in past 12 months`[is.na(df$`G-tube in past 12 months`)] = "No"
+# Pre-/post-modulator
+df$`Pre or Post` = factor(sign(df$Time),levels = -1:1,labels = c("Pre","Pre","Post"))
+# Get pre value closest to start, post value furthest from start
+df = df %>% arrange(`CFF ID`,Time) %>% group_by(`CFF ID`,`Pre or Post`) %>%
+  summarise(`A1c Values` = sum(!is.na(HbA1c)),
+            across(everything(),~ last(na.omit(.x))))
 ```
 
-# Participant Characteristics
+# Participant characteristics prior to CFTR modulator start
 
 ```{r results='asis'}
-t1 = tableby(~Age+BMI.Percentile+Weight+Sex+,pre_post[pre_post$Before_After == "Before",])
+# Table
+t1 = tableby(~Age+`BMI Percentile`+Weight+Sex+ppFEV1+ppFVC+`G-tube in past 12 months`+
+               `Pancreatic Status`+Site+Genotype+`Age Group`,
+             data = df %>% filter(`Pre or Post` == "Pre"))
 summary(t1)
@@
 ```{r}
-summary(tableby( ~ ppFEV + ppFVC + BMIz,data = df[!is.na(df$average_sensor),]))
+
 ```

### 2022-10-06 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
rough draft ready for Christine

--- 2022-10-04 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
@@
 title: "Pre-/Post-Trikafta in CFRD"
-author: "Tim Vigers & Casey Sakamoto"
+author: "Tim Vigers"
 date: "`r format(Sys.time(), '%d %B %Y')`"
@@
 ```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = FALSE)
 library(arsenal)
@@
 library(lubridate)
+library(childsds)
+library(rspiro)
 source("~/GitHub/CF/get_genotype_class_severity.R")
-variables = c("CFF ID","Site","Sex","Race","Genotypes1","Genotypes2","Pancreatic Status",
-              "G-tube in past 12 months","Start","Date","Age")
+t1_variables <- c(
+  "CFF ID", "Site", "Sex", "Race", "Genotypes1", "Genotypes2", "Pancreatic Status",
+  "G-tube in past 12 months", "Time", "Age",
+  "BMI Percentile", "Height", "Weight", "FEV1", "FVC", "HbA1c"
+)
 ```
@@
 # CHCO
-chco_a1c = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_glycemic.csv",show_col_types = F)
-chco_a1c$Date = parse_date_time(chco_a1c$Date,orders = "mdy")
-chco_bmi = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv",show_col_types = F)
-chco_bmi$Date = parse_date_time(chco_bmi$Date,orders = "mdy")
-chco = full_join(chco_a1c,chco_bmi,by = c("MRN", "Date"))
+chco_a1c <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_glycemic.csv", show_col_types = F)
+chco_a1c$Date <- parse_date_time(chco_a1c$Date, orders = "mdy")
+chco_bmi <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv", show_col_types = F)
+chco_bmi$Date <- parse_date_time(chco_bmi$Date, orders = "mdy")
+chco <- full_join(chco_a1c, chco_bmi)
 # Fill CFF ID by MRN
-chco= chco %>% group_by(MRN) %>% 
-  fill(`CFF ID`,.direction = "downup") %>% ungroup() %>%
+chco <- chco %>%
+  group_by(MRN) %>%
+  fill(`CFF ID`, .direction = "downup") %>%
+  ungroup() %>%
   filter(!is.na(`CFF ID`)) %>%
   select(-MRN)
-# Add demographics
-chco_reg = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_registry.csv",show_col_types = F)
-chco_reg = chco_reg %>% select(-MRN)
-chco_reg$Start = parse_date_time(chco_reg$Start,orders = "mdy")
-chco = left_join(chco,chco_reg,by = "CFF ID")
-# Add Time
-chco$Time = as.numeric(difftime(chco$Date,chco$Start,units = "days"))
+# Add demographics, filter
+chco_reg <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_registry.csv", show_col_types = F)
+chco_reg$Start <- parse_date_time(chco_reg$Start, orders = "mdy")
+chco_reg$DOB <- parse_date_time(chco_reg$DOB, orders = "mdy")
+keep <- chco_reg %>%
+  filter(!is.na(`CFRD Diagnosis Date`)) %>%
+  pull(`CFF ID`)
+exclude <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv", show_col_types = F)
+exclude <- exclude %>%
+  filter(!is.na(`Remove (see notes)`)) %>%
+  pull(MRN)
+exclude <- chco_reg$`CFF ID`[match(exclude, chco_reg$MRN)]
+keep <- setdiff(keep, exclude)
+chco <- left_join(chco, chco_reg) %>% filter(`CFF ID` %in% keep)
+# Add missing variables
+chco$Age <- round(as.numeric(difftime(chco$Date, chco$DOB, units = "days")) / 365.25, 1)
+chco$Time <- as.numeric(difftime(chco$Date, chco$Start, units = "days"))
+chco$Site <- "CHCO"
+# BMI calculations
+chco$age_for_bmi <- ifelse(chco$Age < 20, chco$Age, 20)
+chco$`BMI Percentile` <- round(sds(chco$BMI,
+  age = chco$age_for_bmi, item = "bmi", type = "perc",
+  sex = chco$Sex, male = "Male", female = "Female",
+  ref = cdc.ref
+) * 100, 2)
 ```
@@
 # Montana
-mon_bmi = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_bmi_pft.csv",show_col_types = F)
-mon_bmi$Date = parse_date_time(mon_bmi$Date,orders = "mdy")
-mon_a1c = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv",show_col_types = F)
-mon_a1c$Date = parse_date_time(mon_a1c$Date,orders = "mdy")
+mon_bmi <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_bmi_pft.csv", show_col_types = F)
+mon_bmi$Date <- parse_date_time(mon_bmi$Date, orders = "mdy")
+mon_a1c <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv", show_col_types = F)
+mon_a1c$Date <- parse_date_time(mon_a1c$Date, orders = "mdy")
+mon_a1c$DOB <- parse_date_time(mon_a1c$DOB, orders = "mdy")
+mon_a1c$HbA1c <- as.numeric(mon_a1c$HbA1c)
 # Get start date and CFRD status
-mon_starts = mon_a1c %>% select(`CFF ID`,`CFRD Status`,Start) %>% 
-  distinct() %>% filter(`CFRD Status`=="CFRD with or without fasting hyperglycemia")
-mon_starts$Start = parse_date_time(mon_starts$Start,orders = "mdy")
+mon_starts <- mon_a1c %>%
+  select(`CFF ID`, `CFRD Status`, Start) %>%
+  distinct() %>%
+  filter(`CFRD Status` == "CFRD with or without fasting hyperglycemia")
+mon_starts$Start <- parse_date_time(mon_starts$Start, orders = "mdy")
 # Remove unnecessary columns
-mon_a1c = mon_a1c %>% select(`CFF ID`,Date,HbA1c,`OGTT Fasting`,`OGTT Two Hour`)
+mon_a1c <- mon_a1c %>% select(`CFF ID`, DOB, Date, HbA1c, `OGTT Fasting`, `OGTT Two Hour`)
 # Merge and remove those without CFRD or start date
-mon = full_join(mon_bmi,mon_a1c,by = c("CFF ID", "Date"))
-mon= mon %>% filter(`CFF ID` %in% mon_starts$`CFF ID`)
-mon = left_join(mon,mon_starts %>% select(`CFF ID`,Start), by = "CFF ID")
-# Add Time
-mon$Time = as.numeric(difftime(mon$Date,mon$Start,units = "days"))
+mon <- full_join(mon_bmi, mon_a1c)
+mon <- mon %>% filter(`CFF ID` %in% mon_starts$`CFF ID`)
+mon <- left_join(mon, mon_starts %>% select(`CFF ID`, Start))
+# Add demographics
+mon_demo <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_demo.csv", show_col_types = F)
+mon_demo <- mon_demo %>% select(`CFF ID`, Race, Sex)
+mon <- left_join(mon, mon_demo)
+# Add genotypes
+mon_geno <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_genotypes.csv", show_col_types = F)
+mon <- left_join(mon, mon_geno)
+# Add pancreatic and g tube status
+mon_pan <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_pancreatic_insufficient.csv", show_col_types = F)
+mon_pan <- mon_pan %>%
+  select(-Date) %>%
+  distinct()
+mon_pan$`Pancreatic Status` <- "Insufficient"
+mon_gtube <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_gtubes.csv", show_col_types = F)
+mon_gtube <- mon_gtube %>%
+  select(-Feeding, -Date) %>%
+  distinct()
+mon_gtube$`G-tube in past 12 months` <- "Yes"
+mon <- left_join(mon, mon_pan)
+mon <- left_join(mon, mon_gtube)
+# Add missing variables
+mon <- mon %>%
+  group_by(`CFF ID`) %>%
+  fill(Start, DOB, .direction = "downup")
+mon$Time <- as.numeric(difftime(mon$Date, mon$Start, units = "days"))
+mon$Age <- round(as.numeric(difftime(mon$Date, mon$DOB)) / 365.25, 1)
+mon$Site <- "Montana"
+# Fill in some missing
+mon$`Pancreatic Status`[is.na(mon$`Pancreatic Status`)] <- "Sufficient"
+# BMI calculations
+mon$age_for_bmi <- ifelse(mon$Age < 20, mon$Age, 20)
+mon$`BMI Percentile` <- round(sds(mon$BMI,
+  age = mon$age_for_bmi, item = "bmi", type = "perc",
+  sex = mon$Sex, male = "Male", female = "Female",
+  ref = cdc.ref
+) * 100, 2)
 ```
@@
 # UM
-um_a1c = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_a1cs.csv",show_col_types = F)
-um_a1c$Age = round(um_a1c$Age / 365.25,1)
-um_bmi = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_bmi.csv",show_col_types = F)
-um_bmi$Age = round(um_bmi$Age / 365.25,1)
-um_ogtt = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_ogtt.csv",show_col_types = F)
-um_ogtt$Age = round(um_ogtt$Age / 365.25,1)
-um_pft = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_pfts.csv",show_col_types = F)
-um_pft$Age = round(um_pft$Age / 365.25,1)
-um_race = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_race.csv",show_col_types = F)
+um_a1c <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_a1cs.csv", show_col_types = F)
+um_a1c$Age <- round(um_a1c$Age / 365.25, 1)
+um_bmi <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_bmi.csv", show_col_types = F)
+um_bmi$Age <- round(um_bmi$Age / 365.25, 1)
+um_ogtt <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_ogtt.csv", show_col_types = F)
+um_ogtt$Age <- round(um_ogtt$Age / 365.25, 1)
+um_pft <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_pfts.csv", show_col_types = F)
+um_pft$Age <- round(um_pft$Age / 365.25, 1)
+um_race <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_race.csv", show_col_types = F)
+um_demo <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_demo.csv", show_col_types = F)
+keep <- um_demo %>%
+  filter(`CFRD Status` == "CFRD with or without fasting hyperglycemia") %>%
+  pull(`CFF ID`)
+um_demo <- um_demo %>%
+  select(-`CFRD Status`) %>%
+  rename(`G-tube in past 12 months` = `G tube in past 12 months`)
 # Join
-um = reduce(list(um_a1c,um_bmi,um_ogtt,um_pft),full_join)
-um = left_join(um,um_race,by = "CFF ID")
+um <- reduce(list(um_a1c, um_bmi, um_ogtt, um_pft), full_join)
+um <- left_join(um, um_race)
+um <- left_join(um, um_demo)
+um <- um %>%
+  filter(`CFF ID` %in% keep, !is.na(Time), abs(Time) <= 365) %>%
+  rename(Site = "site") %>%
+  arrange(`CFF ID`, Time)
+# Filter per Maxie's previous work (unclear exactly how they got to this subset)
+um_ids <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_ids_from_maxie.csv", show_col_types = F)
+um <- um %>% filter(`CFF ID` %in% um_ids$`CFF ID`)
+# BMI calculations
+um$age_for_bmi <- ifelse(um$Age < 20, um$Age, 20)
+um$`BMI Percentile` <- round(sds(um$BMI,
+  age = um$age_for_bmi, item = "bmi", type = "perc",
+  sex = um$Sex, male = "Male", female = "Female",
+  ref = cdc.ref
+) * 100, 2)
 ```
@@
 # UW
-uw_bmi = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_bmi.csv",show_col_types = F)
-uw_bmi$Date = parse_date_time(uw_bmi$Date,orders = "mdy")
-uw_glyc = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_glycemic.csv",show_col_types = F)
-uw_glyc$Date = parse_date_time(uw_glyc$Date,orders = "mdy")
-uw_pft = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pft.csv",show_col_types = F)
-uw_pft$Date = parse_date_time(uw_pft$Date,orders = "mdy")
-uw_reg = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_registry.csv",show_col_types = F)
-uw_reg$Start = parse_date_time(uw_reg$Start,orders = "mdy")
+uw_bmi <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_bmi.csv", show_col_types = F)
+uw_bmi$Date <- parse_date_time(uw_bmi$Date, orders = "mdy")
+uw_glyc <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_glycemic.csv", show_col_types = F)
+uw_glyc$Date <- parse_date_time(uw_glyc$Date, orders = "mdy")
+uw_pft <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pft.csv", show_col_types = F)
+uw_pft$Date <- parse_date_time(uw_pft$Date, orders = "mdy")
+uw_demo <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_demo.csv", show_col_types = F)
+uw_demo <- uw_demo %>% select(`CFF ID`, Race, Sex)
+# Get CFRD status and start date
+uw_start <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_mods.csv", show_col_types = F)
+uw_start$`Date of Encounter` <- parse_date_time(uw_start$`Date of Encounter`, orders = "mdy")
+uw_start <- uw_start %>%
+  arrange(`CFF ID`, `Date of Encounter`) %>%
+  group_by(`CFF ID`) %>%
+  filter(row_number() == n()) %>%
+  select(`CFF ID`, `Date of Encounter`) %>%
+  rename(Start = `Date of Encounter`)
+uw_cfrd <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_cfrd.csv", show_col_types = F)
+uw_cfrd <- uw_cfrd %>%
+  filter(`CFRD Status` == "CFRD with or without fasting hyperglycemia") %>%
+  select(`CFF ID`) %>%
+  distinct() %>%
+  pull(`CFF ID`)
+keep <- unique(uw_cfrd, uw_start$`CFF ID`)
+# Add genotypes
+uw_geno <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_genotypes.csv", show_col_types = F)
+# Pancreatic status
+uw_pan <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pancreatic.csv", show_col_types = F)
+uw_pan <- uw_pan %>%
+  select(-Date) %>%
+  distinct() %>%
+  mutate(`Pancreatic Status` = "Insufficient")
+# G tubes
+uw_gtube <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_gtubes.csv", show_col_types = F)
+uw_gtube <- uw_gtube %>%
+  select(`CFF ID`) %>%
+  distinct() %>%
+  mutate(`G-tube in past 12 months` = "Yes")
 # Merge
-uw = reduce(list(uw_bmi,uw_glyc,uw_pft,uw_reg),full_join)
+uw <- reduce(list(uw_bmi, uw_glyc, uw_pft), full_join)
+uw <- left_join(uw, uw_demo)
+uw <- left_join(uw, uw_geno)
+uw <- left_join(uw, uw_pan)
+uw <- left_join(uw, uw_gtube)
+uw <- left_join(uw, uw_start) %>% filter(`CFF ID` %in% keep)
+# Other variables (no weight for this site)
+uw$Time <- as.numeric(difftime(uw$Date, uw$Start, units = "days"))
+uw$Site <- "UW"
+uw$Weight <- NA
+# BMI calculations
+uw$age_for_bmi <- ifelse(uw$Age < 20, uw$Age, 20)
+uw$`BMI Percentile 2` <- round(sds(uw$BMI,
+  age = uw$age_for_bmi, item = "bmi", type = "perc",
+  sex = uw$Sex, male = "Male", female = "Female",
+  ref = cdc.ref
+) * 100, 2)
+uw$`BMI Percentile` <- coalesce(uw$`BMI Percentile`, uw$`BMI Percentile 2`)
 ```
@@
 # Put everything together
-
+chco <- chco %>% select(all_of(t1_variables))
+mon <- mon %>% select(all_of(t1_variables))
+um <- um %>% select(all_of(t1_variables))
+uw <- uw %>% select(all_of(t1_variables))
+df <- bind_rows(chco, mon, um, uw) %>% arrange(`CFF ID`, Time)
+# Age group
+df$`Age Group` <- cut(df$Age, breaks = c(-Inf, 20, Inf), labels = c("Pedatric", "Adult"), right = F)
+# Race and sex
+df$Race <- factor(df$Race,
+  levels = c(
+    "Black or African American", "Caucasian",
+    "Other", "Some other race", "Two or more races", "White"
+  ),
+  labels = c(
+    "Black", "Caucasian",
+    "Other", "Other", "Other", "Caucasian"
+  )
+)
+df$Sex <- factor(df$Sex, levels = c("Male", "Female"))
+# FEV calculations
+age <- df$Age
+height <- df$Height / 100
+race <- df$Race
+old <- c("Caucasian", "Black", "Other")
+new <- c(1, 2, 5)
+race <- new[match(race, old)]
+gender <- as.numeric(df$Sex)
+df$ppFEV1 <- pctpred_GLI(age, height, gender = gender, ethnicity = race, FEV1 = df$FEV1)
+df$ppFVC <- pctpred_GLI(age, height, gender = gender, ethnicity = race, FVC = df$FVC)
+# Genotype severity
+df$Genotypes1 <- get_cf_genotype_class_severity(df$Genotypes1)
+df$Genotypes2 <- get_cf_genotype_class_severity(df$Genotypes2)
+df <- df %>% unite(Genotype, Genotypes1, Genotypes2)
+df$Genotype <- factor(df$Genotype,
+  levels = c(
+    "Mild_Severe", "Severe_Mild", "Severe_Severe",
+    "Severe_Unknown", "Unknown_Unknown"
+  ),
+  labels = c("Low", "Low", "High", "Unknown", "Unknown")
+)
+# Other little cleaning things
+df <- df %>% filter(Age > 0, !is.na(Time))
+df$`G-tube in past 12 months`[is.na(df$`G-tube in past 12 months`)] <- "No"
+# Pre-/post-modulator
+df$`Pre or Post` <- factor(sign(df$Time), levels = -1:1, labels = c("Pre", "Pre", "Post"))
+# Get pre value closest to start, post value furthest from start
+df <- df %>%
+  arrange(`CFF ID`, Time) %>%
+  group_by(`CFF ID`, `Pre or Post`) %>%
+  summarise(
+    `A1c Values` = sum(!is.na(HbA1c)),
+    across(everything(), ~ last(na.omit(.x)))
+  )
+# Add insulin information
+chco_ins <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv", show_col_types = F)
+chco_ins$`CFF ID` <- chco_a1c$`CFF ID`[match(chco_ins$MRN, chco_a1c$MRN)]
+chco_ins <- chco_ins %>%
+  select(`CFF ID`, `Pre or Post`, `Total Daily Dose of insulin (units)`) %>%
+  rename(TDD = `Total Daily Dose of insulin (units)`) %>%
+  filter(!is.na(`CFF ID`))
+mon_ins <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_insulin.csv", show_col_types = F)
+mon_ins <- mon_ins %>%
+  select(`CFF ID`, `Pre or Post`, `Total Daily Dose of insulin (units)`) %>%
+  rename(TDD = `Total Daily Dose of insulin (units)`)
+uw_ins <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_insulin.csv", show_col_types = F)
+uw_ins <- uw_ins %>%
+  select(`CFF ID`, `Pre or Post`, `Total Daily Dose of insulin (units)`) %>%
+  rename(TDD = `Total Daily Dose of insulin (units)`)
+insulin <- bind_rows(chco_ins, mon_ins, uw_ins)
+# Join
+df <- left_join(df, insulin)
+df$`Pre or Post` <- factor(df$`Pre or Post`, levels = c("Pre", "Post"))
 ```
 
-# Participant Characteristics
+```{r include=FALSE}
+cgm_vars <- c(
+  "average_sensor", "min_sensor", "max_sensor", "percent_time_over_120",
+  "percent_time_over_140", "percent_time_over_200", "excursions_over_140",
+  "excursions_over_200", "percent_time_under_60", "percent_time_under_70",
+  "average_auc_per_day", "standard_deviation", "r_mage"
+)
+# Add CGM information
+# CHCO
+chco_cgm <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_cgm.csv", show_col_types = F)
+chco_cgm$`Pre or Post` <- factor(sapply(str_split(chco_cgm$subject_id, "_"), "[", 2))
+levels(chco_cgm$`Pre or Post`) <- c("Post", "Post", "Pre", "Pre")
+chco_cgm$`CFF ID` <- sapply(str_split(chco_cgm$subject_id, "_"), "[", 1)
+chco_cgm <- chco_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(cgm_vars))
+# Montana
+mon_cgm <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_cgm.csv", show_col_types = F)
+mon_cgm$`Pre or Post` <- factor(sapply(str_split(mon_cgm$subject_id, "_"), "[", 2))
+levels(mon_cgm$`Pre or Post`) <- c("Post", "Post", "Pre", "Pre")
+mon_cgm$`CFF ID` <- sapply(str_split(mon_cgm$subject_id, "_"), "[", 1)
+mon_cgm <- mon_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(cgm_vars))
+# UW
+uw_cgm <- read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_cgm.csv", show_col_types = F)
+uw_cgm$`Pre or Post` <- factor(sapply(str_split(uw_cgm$subject_id, "_"), "[", 2))
+levels(uw_cgm$`Pre or Post`) <- c("Post", "Pre")
+uw_cgm$`CFF ID` <- sapply(str_split(uw_cgm$subject_id, "_"), "[", 1)
+uw_cgm <- uw_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(cgm_vars))
+# Put together
+cgm <- bind_rows(chco_cgm, mon_cgm, uw_cgm)
+```
+
+# Participant characteristics prior to CFTR modulator start
 
 ```{r results='asis'}
-t1 = tableby(~Age+BMI.Percentile+Weight+Sex+,pre_post[pre_post$Before_After == "Before",])
+# Table
+t1 <- tableby(~ Age + `BMI Percentile` + Weight + Sex + ppFEV1 + ppFVC +
+  `G-tube in past 12 months` + `Pancreatic Status` + Site +
+  Genotype + `Age Group`,
+data = df %>% filter(`Pre or Post` == "Pre")
+)
 summary(t1)
@@
 
-```{r}
-summary(tableby( ~ ppFEV + ppFVC + BMIz,data = df[!is.na(df$average_sensor),]))
+## CGM
+
+```{r results='asis'}
+cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(cgm_vars, collapse = "+")))
+p <- paired(cgm_formula,
+  data = cgm, id = `CFF ID`,
+  signed.rank.exact = FALSE, numeric.test = "signed.rank",
+  numeric.stats = c("Nmiss", "medianrange", "q1q3")
+)
+summary(p, pfootnote = T, digits = 1)
 ```
+
+## Other outcomes
+
+```{r results='asis'}
+p <- paired(`Pre or Post` ~ `A1c Values` + HbA1c + Time + Weight + `BMI Percentile` +
+  ppFEV1 + ppFVC + TDD,
+data = df, id = `CFF ID`,
+signed.rank.exact = FALSE, numeric.test = "signed.rank",
+numeric.stats = c("Nmiss", "medianrange", "q1q3")
+)
+summary(p, pfootnote = T, digits = 1)
+```

### 2022-10-06 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
fixed UW filtering

--- 2022-10-06 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
@@
 library(lubridate)
+library(childsds)
 library(rspiro)
 source("~/GitHub/CF/get_genotype_class_severity.R")
-t1_variables = c("CFF ID","Site","Sex","Race","Genotypes1","Genotypes2","Pancreatic Status",
-              "G-tube in past 12 months","Time","Age","BMI","Height","Weight","FEV1","FVC")
+t1_variables <- c(
+  "CFF ID", "Site", "Sex", "Race", "Genotypes1", "Genotypes2", "Pancreatic Status",
+  "G-tube in past 12 months", "Time", "Age",
+  "BMI Percentile", "Height", "Weight", "FEV1", "FVC", "HbA1c"
+)
 ```
@@
 # CHCO
-chco_a1c = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_glycemic.csv",show_col_types = F)
-chco_a1c$Date = parse_date_time(chco_a1c$Date,orders = "mdy")
-chco_bmi = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv",show_col_types = F)
-chco_bmi$Date = parse_date_time(chco_bmi$Date,orders = "mdy")
-chco = full_join(chco_a1c,chco_bmi)
+chco_a1c <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_glycemic.csv", show_col_types = F)
+chco_a1c$Date <- parse_date_time(chco_a1c$Date, orders = "mdy")
+chco_bmi <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv", show_col_types = F)
+chco_bmi$Date <- parse_date_time(chco_bmi$Date, orders = "mdy")
+chco <- full_join(chco_a1c, chco_bmi)
 # Fill CFF ID by MRN
-chco= chco %>% group_by(MRN) %>% 
-  fill(`CFF ID`,.direction = "downup") %>% ungroup() %>%
+chco <- chco %>%
+  group_by(MRN) %>%
+  fill(`CFF ID`, .direction = "downup") %>%
+  ungroup() %>%
   filter(!is.na(`CFF ID`)) %>%
   select(-MRN)
-# Add demographics
-chco_reg = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_registry.csv",show_col_types = F)
-chco_reg$Start = parse_date_time(chco_reg$Start,orders = "mdy")
-chco_reg$DOB = parse_date_time(chco_reg$DOB,orders = "mdy")
-keep = chco_reg %>% filter(!is.na(`CFRD Diagnosis Date`)) %>% pull(`CFF ID`)
-exclude = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv",show_col_types = F)
-exclude = exclude %>% filter(!is.na(`Remove (see notes)`)) %>% pull(MRN)
-exclude = chco_reg$`CFF ID`[match(exclude,chco_reg$MRN)]
-keep =setdiff(keep,exclude)
-chco = left_join(chco,chco_reg) %>% filter(`CFF ID` %in% keep)
+# Add demographics, filter
+chco_reg <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_registry.csv", show_col_types = F)
+chco_reg$Start <- parse_date_time(chco_reg$Start, orders = "mdy")
+chco_reg$DOB <- parse_date_time(chco_reg$DOB, orders = "mdy")
+keep <- chco_reg %>%
+  filter(!is.na(`CFRD Diagnosis Date`)) %>%
+  pull(`CFF ID`)
+exclude <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv", show_col_types = F)
+exclude <- exclude %>%
+  filter(!is.na(`Remove (see notes)`)) %>%
+  pull(MRN)
+exclude <- chco_reg$`CFF ID`[match(exclude, chco_reg$MRN)]
+keep <- setdiff(keep, exclude)
+chco <- left_join(chco, chco_reg) %>% filter(`CFF ID` %in% keep)
 # Add missing variables
-chco$Age = round(as.numeric(difftime(chco$Date,chco$DOB,units = "days"))/365.25,1)
-chco$Time = as.numeric(difftime(chco$Date,chco$Start,units = "days"))
-chco$Site = "CHCO"
+chco$Age <- round(as.numeric(difftime(chco$Date, chco$DOB, units = "days")) / 365.25, 1)
+chco$Time <- as.numeric(difftime(chco$Date, chco$Start, units = "days"))
+chco$Site <- "CHCO"
+# BMI calculations
+chco$age_for_bmi <- ifelse(chco$Age < 20, chco$Age, 20)
+chco$`BMI Percentile` <- round(sds(chco$BMI,
+  age = chco$age_for_bmi, item = "bmi", type = "perc",
+  sex = chco$Sex, male = "Male", female = "Female",
+  ref = cdc.ref
+) * 100, 2)
 ```
@@
 # Montana
-mon_bmi = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_bmi_pft.csv",show_col_types = F)
-mon_bmi$Date = parse_date_time(mon_bmi$Date,orders = "mdy")
-mon_a1c = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv",show_col_types = F)
-mon_a1c$Date = parse_date_time(mon_a1c$Date,orders = "mdy")
-mon_a1c$DOB = parse_date_time(mon_a1c$DOB,orders = "mdy")
+mon_bmi <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_bmi_pft.csv", show_col_types = F)
+mon_bmi$Date <- parse_date_time(mon_bmi$Date, orders = "mdy")
+mon_a1c <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv", show_col_types = F)
+mon_a1c$Date <- parse_date_time(mon_a1c$Date, orders = "mdy")
+mon_a1c$DOB <- parse_date_time(mon_a1c$DOB, orders = "mdy")
+mon_a1c$HbA1c <- as.numeric(mon_a1c$HbA1c)
 # Get start date and CFRD status
-mon_starts = mon_a1c %>% select(`CFF ID`,`CFRD Status`,Start) %>% 
-  distinct() %>% filter(`CFRD Status`=="CFRD with or without fasting hyperglycemia")
-mon_starts$Start = parse_date_time(mon_starts$Start,orders = "mdy")
+mon_starts <- mon_a1c %>%
+  select(`CFF ID`, `CFRD Status`, Start) %>%
+  distinct() %>%
+  filter(`CFRD Status` == "CFRD with or without fasting hyperglycemia")
+mon_starts$Start <- parse_date_time(mon_starts$Start, orders = "mdy")
 # Remove unnecessary columns
-mon_a1c = mon_a1c %>% select(`CFF ID`,DOB,Date,HbA1c,`OGTT Fasting`,`OGTT Two Hour`)
+mon_a1c <- mon_a1c %>% select(`CFF ID`, DOB, Date, HbA1c, `OGTT Fasting`, `OGTT Two Hour`)
 # Merge and remove those without CFRD or start date
-mon = full_join(mon_bmi,mon_a1c)
-mon= mon %>% filter(`CFF ID` %in% mon_starts$`CFF ID`)
-mon = left_join(mon,mon_starts %>% select(`CFF ID`,Start))
+mon <- full_join(mon_bmi, mon_a1c)
+mon <- mon %>% filter(`CFF ID` %in% mon_starts$`CFF ID`)
+mon <- left_join(mon, mon_starts %>% select(`CFF ID`, Start))
 # Add demographics
-mon_demo = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_demo.csv",show_col_types = F)
-mon_demo = mon_demo %>% select(`CFF ID`,Race,Sex)
-mon = left_join(mon,mon_demo)
+mon_demo <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_demo.csv", show_col_types = F)
+mon_demo <- mon_demo %>% select(`CFF ID`, Race, Sex)
+mon <- left_join(mon, mon_demo)
 # Add genotypes
-mon_geno = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_genotypes.csv",show_col_types = F)
-mon = left_join(mon,mon_geno)
+mon_geno <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_genotypes.csv", show_col_types = F)
+mon <- left_join(mon, mon_geno)
 # Add pancreatic and g tube status
-mon_pan = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_pancreatic_insufficient.csv",show_col_types = F)
-mon_pan = mon_pan %>% select(-Date) %>% distinct()
-mon_pan$`Pancreatic Status` = "Insufficient"
-mon_gtube = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_gtubes.csv",show_col_types = F)
-mon_gtube = mon_gtube %>% select(-Feeding,-Date) %>% distinct()
-mon_gtube$`G-tube in past 12 months` = "Yes"
-mon = left_join(mon,mon_pan)
-mon = left_join(mon,mon_gtube)
+mon_pan <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_pancreatic_insufficient.csv", show_col_types = F)
+mon_pan <- mon_pan %>%
+  select(-Date) %>%
+  distinct()
+mon_pan$`Pancreatic Status` <- "Insufficient"
+mon_gtube <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_gtubes.csv", show_col_types = F)
+mon_gtube <- mon_gtube %>%
+  select(-Feeding, -Date) %>%
+  distinct()
+mon_gtube$`G-tube in past 12 months` <- "Yes"
+mon <- left_join(mon, mon_pan)
+mon <- left_join(mon, mon_gtube)
 # Add missing variables
-mon = mon %>% group_by(`CFF ID`) %>% fill(Start,DOB,.direction = "downup")
-mon$Time = as.numeric(difftime(mon$Date,mon$Start,units = "days"))
-mon$Age = round(as.numeric(difftime(mon$Date,mon$DOB))/365.25,1)
-mon$Site = "Montana"
+mon <- mon %>%
+  group_by(`CFF ID`) %>%
+  fill(Start, DOB, .direction = "downup")
+mon$Time <- as.numeric(difftime(mon$Date, mon$Start, units = "days"))
+mon$Age <- round(as.numeric(difftime(mon$Date, mon$DOB)) / 365.25, 1)
+mon$Site <- "Montana"
+# Fill in some missing
+mon$`Pancreatic Status`[is.na(mon$`Pancreatic Status`)] <- "Sufficient"
+# BMI calculations
+mon$age_for_bmi <- ifelse(mon$Age < 20, mon$Age, 20)
+mon$`BMI Percentile` <- round(sds(mon$BMI,
+  age = mon$age_for_bmi, item = "bmi", type = "perc",
+  sex = mon$Sex, male = "Male", female = "Female",
+  ref = cdc.ref
+) * 100, 2)
 ```
@@
 # UM
-um_a1c = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_a1cs.csv",show_col_types = F)
-um_a1c$Age = round(um_a1c$Age / 365.25,1)
-um_bmi = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_bmi.csv",show_col_types = F)
-um_bmi$Age = round(um_bmi$Age / 365.25,1)
-um_ogtt = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_ogtt.csv",show_col_types = F)
-um_ogtt$Age = round(um_ogtt$Age / 365.25,1)
-um_pft = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_pfts.csv",show_col_types = F)
-um_pft$Age = round(um_pft$Age / 365.25,1)
-um_race = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_race.csv",show_col_types = F)
-um_demo = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_demo.csv",show_col_types = F)
-keep = um_demo %>% 
-  filter(`CFRD Status` =="CFRD with or without fasting hyperglycemia") %>%
+um_a1c <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_a1cs.csv", show_col_types = F)
+um_a1c$Age <- round(um_a1c$Age / 365.25, 1)
+um_bmi <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_bmi.csv", show_col_types = F)
+um_bmi$Age <- round(um_bmi$Age / 365.25, 1)
+um_ogtt <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_ogtt.csv", show_col_types = F)
+um_ogtt$Age <- round(um_ogtt$Age / 365.25, 1)
+um_pft <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_pfts.csv", show_col_types = F)
+um_pft$Age <- round(um_pft$Age / 365.25, 1)
+um_race <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_race.csv", show_col_types = F)
+um_demo <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_demo.csv", show_col_types = F)
+keep <- um_demo %>%
+  filter(`CFRD Status` == "CFRD with or without fasting hyperglycemia") %>%
   pull(`CFF ID`)
-um_demo = um_demo %>% select(-`CFRD Status`) %>% 
-  rename(`G-tube in past 12 months`=`G tube in past 12 months`)
+um_demo <- um_demo %>%
+  select(-`CFRD Status`) %>%
+  rename(`G-tube in past 12 months` = `G tube in past 12 months`)
 # Join
-um = reduce(list(um_a1c,um_bmi,um_ogtt,um_pft),full_join)
-um = left_join(um,um_race)
-um = left_join(um,um_demo)
-um = um %>% filter(`CFF ID` %in% keep) %>% rename(Site = "site")
+um <- reduce(list(um_a1c, um_bmi, um_ogtt, um_pft), full_join)
+um <- left_join(um, um_race)
+um <- left_join(um, um_demo)
+um <- um %>%
+  filter(`CFF ID` %in% keep, !is.na(Time), abs(Time) <= 365) %>%
+  rename(Site = "site") %>%
+  arrange(`CFF ID`, Time)
+# BMI calculations
+um$age_for_bmi <- ifelse(um$Age < 20, um$Age, 20)
+um$`BMI Percentile` <- round(sds(um$BMI,
+  age = um$age_for_bmi, item = "bmi", type = "perc",
+  sex = um$Sex, male = "Male", female = "Female",
+  ref = cdc.ref
+) * 100, 2)
 ```
@@
 # UW
-uw_bmi = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_bmi.csv",show_col_types = F)
-uw_bmi$Date = parse_date_time(uw_bmi$Date,orders = "mdy")
-uw_glyc = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_glycemic.csv",show_col_types = F)
-uw_glyc$Date = parse_date_time(uw_glyc$Date,orders = "mdy")
-uw_pft = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pft.csv",show_col_types = F)
-uw_pft$Date = parse_date_time(uw_pft$Date,orders = "mdy")
-uw_demo = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_demo.csv",show_col_types = F)
-uw_demo = uw_demo %>% select(`CFF ID`,Race,Sex)
+uw_bmi <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_bmi.csv", show_col_types = F)
+uw_bmi$Date <- parse_date_time(uw_bmi$Date, orders = "mdy")
+uw_glyc <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_glycemic.csv", show_col_types = F)
+uw_glyc$Date <- parse_date_time(uw_glyc$Date, orders = "mdy")
+uw_pft <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pft.csv", show_col_types = F)
+uw_pft$Date <- parse_date_time(uw_pft$Date, orders = "mdy")
+uw_demo <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_demo.csv", show_col_types = F)
+uw_demo <- uw_demo %>% select(`CFF ID`, Race, Sex)
 # Get CFRD status and start date
-uw_start = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_mods.csv",show_col_types = F)
-uw_start$`Date of Encounter` = parse_date_time(uw_start$`Date of Encounter`,orders = "mdy")
-uw_start = uw_start %>% arrange(`CFF ID`,`Date of Encounter`) %>% 
-  group_by(`CFF ID`) %>% filter(row_number()==n()) %>%
-  select(`CFF ID`,`Date of Encounter`) %>% rename(Start = `Date of Encounter`)
-uw_cfrd = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_cfrd.csv",show_col_types = F)
-uw_cfrd = uw_cfrd %>% 
-  filter(`CFRD Status`=="CFRD with or without fasting hyperglycemia") %>% 
-  select(`CFF ID`) %>% distinct() %>% pull(`CFF ID`)
-keep = unique(uw_cfrd,uw_start$`CFF ID`)
+uw_start <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_mods.csv", show_col_types = F)
+uw_start$`Date of Encounter` <- parse_date_time(uw_start$`Date of Encounter`, orders = "mdy")
+uw_start <- uw_start %>%
+  arrange(`CFF ID`, `Date of Encounter`) %>%
+  group_by(`CFF ID`) %>%
+  filter(row_number() == 1) %>%
+  select(`CFF ID`, `Date of Encounter`) %>%
+  rename(Start = `Date of Encounter`)
+uw_cfrd <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_cfrd.csv", show_col_types = F)
+uw_cfrd <- uw_cfrd %>%
+  filter(`CFRD Status` == "CFRD with or without fasting hyperglycemia") %>%
+  select(`CFF ID`) %>%
+  distinct() %>%
+  pull(`CFF ID`)
+keep <- unique(uw_cfrd, uw_start$`CFF ID`)
 # Add genotypes
-uw_geno = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_genotypes.csv",show_col_types = F)
+uw_geno <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_genotypes.csv", show_col_types = F)
 # Pancreatic status
-uw_pan = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pancreatic.csv",show_col_types = F)
-uw_pan = uw_pan %>% select(-Date) %>% distinct() %>% mutate(`Pancreatic Status`="Insufficient")
+uw_pan <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pancreatic.csv", show_col_types = F)
+uw_pan <- uw_pan %>%
+  select(-Date) %>%
+  distinct() %>%
+  mutate(`Pancreatic Status` = "Insufficient")
 # G tubes
-uw_gtube = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_gtubes.csv",show_col_types = F)
-uw_gtube = uw_gtube %>% select(`CFF ID`) %>% distinct() %>% mutate(`G-tube in past 12 months`="Yes")
+uw_gtube <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_gtubes.csv", show_col_types = F)
+uw_gtube <- uw_gtube %>%
+  select(`CFF ID`) %>%
+  distinct() %>%
+  mutate(`G-tube in past 12 months` = "Yes")
 # Merge
-uw = reduce(list(uw_bmi,uw_glyc,uw_pft),full_join)
-uw = left_join(uw,uw_demo)
-uw = left_join(uw,uw_geno)
-uw = left_join(uw,uw_pan)
-uw = left_join(uw,uw_gtube)
-uw = left_join(uw,uw_start) %>% filter(`CFF ID` %in% keep)
+uw <- reduce(list(uw_bmi, uw_glyc, uw_pft), full_join)
+uw <- left_join(uw, uw_demo)
+uw <- left_join(uw, uw_geno)
+uw <- left_join(uw, uw_pan)
+uw <- left_join(uw, uw_gtube)
+uw <- left_join(uw, uw_start) %>% filter(`CFF ID` %in% keep)
 # Other variables (no weight for this site)
-uw$Time = as.numeric(difftime(uw$Date,uw$Start,units = "days"))
-uw$Site = "UW"
-uw$Weight = NA
+uw$Time <- as.numeric(difftime(uw$Date, uw$Start, units = "days"))
+uw$Site <- "UW"
+uw$Weight <- NA
+# BMI calculations
+uw$age_for_bmi <- ifelse(uw$Age < 20, uw$Age, 20)
+uw$`BMI Percentile 2` <- round(sds(uw$BMI,
+  age = uw$age_for_bmi, item = "bmi", type = "perc",
+  sex = uw$Sex, male = "Male", female = "Female",
+  ref = cdc.ref
+) * 100, 2)
+uw$`BMI Percentile` <- coalesce(uw$`BMI Percentile`, uw$`BMI Percentile 2`)
 ```
@@
 # Put everything together
-chco = chco %>% select(all_of(t1_variables))
-mon = mon %>% select(all_of(t1_variables))
-um = um %>% select(all_of(t1_variables))
-uw = uw %>% select(all_of(t1_variables))
-df = bind_rows(chco,mon,um,uw) %>% arrange(`CFF ID`,Time)
+chco <- chco %>% select(all_of(t1_variables))
+mon <- mon %>% select(all_of(t1_variables))
+um <- um %>% select(all_of(t1_variables))
+uw <- uw %>% select(all_of(t1_variables))
+df <- bind_rows(chco, mon, um, uw) %>% arrange(`CFF ID`, Time)
 # Age group
-df$`Age Group` = cut(df$Age,breaks = c(-Inf,20,Inf),labels = c("Pedatric","Adult"),right = F)
+df$`Age Group` <- cut(df$Age, breaks = c(-Inf, 20, Inf), 
+                      labels = c("Pedatric", "Adult"), right = F)
 # Race and sex
-df$Race = factor(df$Race,levels = c("Black or African American","Caucasian",
-                                    "Other","Some other race","Two or more races","White"),
-                 labels = c("Black","Caucasian",
-                                    "Other","Other","Other","Caucasian"))
-df$Sex = factor(df$Sex,levels = c("Male","Female"))
-# BMI and FEV calculations
-age = df$Age
-height = df$Height/100
-race = df$Race
-old <- c("Caucasian","Black","Other")
-new <- c(1,2,5)
-race = new[match(race, old)]
-gender = as.numeric(df$Sex)
-df$ppFEV1  = pctpred_GLI(age,height,gender = gender,ethnicity = race,FEV1 = df$FEV1)
-df$ppFVC  = pctpred_GLI(age,height,gender = gender,ethnicity = race,FVC = df$FVC)
+df$Race <- factor(df$Race,
+  levels = c(
+    "Black or African American", "Caucasian",
+    "Other", "Some other race", "Two or more races", "White"
+  ),
+  labels = c(
+    "Black", "Caucasian",
+    "Other", "Other", "Other", "Caucasian"
+  )
+)
+df$Sex <- factor(df$Sex, levels = c("Male", "Female"))
+# FEV calculations
+age <- df$Age
+height <- df$Height / 100
+race <- df$Race
+old <- c("Caucasian", "Black", "Other")
+new <- c(1, 2, 5)
+race <- new[match(race, old)]
+gender <- as.numeric(df$Sex)
+df$ppFEV1 <- pctpred_GLI(age, height, gender = gender, ethnicity = race, FEV1 = df$FEV1)
+df$ppFVC <- pctpred_GLI(age, height, gender = gender, ethnicity = race, FVC = df$FVC)
 # Genotype severity
-df$Genotypes1 = get_cf_genotype_class_severity(df$Genotypes1)
-df$Genotypes2 = get_cf_genotype_class_severity(df$Genotypes2)
-df = df %>% unite(Genotype,Genotypes1,Genotypes2)
-df$Genotype = factor(df$Genotype,levels = c("Mild_Severe","Severe_Mild","Severe_Severe",
-                                            "Severe_Unknown","Unknown_Unknown"),
-                     labels = c("Low","Low","High","Unknown","Unknown"))
+df$Genotypes1 <- get_cf_genotype_class_severity(df$Genotypes1)
+df$Genotypes2 <- get_cf_genotype_class_severity(df$Genotypes2)
+df <- df %>% unite(Genotype, Genotypes1, Genotypes2)
+df$Genotype <- factor(df$Genotype,
+  levels = c(
+    "Mild_Severe", "Severe_Mild", "Severe_Severe",
+    "Severe_Unknown", "Unknown_Unknown"
+  ),
+  labels = c("Low", "Low", "High", "Unknown", "Unknown")
+)
+# Other little cleaning things
+df <- df %>% filter(Age > 0, !is.na(Time))
+df$`G-tube in past 12 months`[is.na(df$`G-tube in past 12 months`)] <- "No"
+# Pre-/post-modulator
+df$`Pre or Post` <- factor(sign(df$Time), levels = -1:1, labels = c("Pre", "Pre", "Post"))
+# Add insulin information
+chco_ins <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv", show_col_types = F)
+chco_ins$`CFF ID` <- chco_a1c$`CFF ID`[match(chco_ins$MRN, chco_a1c$MRN)]
+chco_ins <- chco_ins %>%
+  select(`CFF ID`, `Pre or Post`, `Total Daily Dose of insulin (units)`) %>%
+  rename(TDD = `Total Daily Dose of insulin (units)`) %>%
+  filter(!is.na(`CFF ID`))
+mon_ins <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_insulin.csv", show_col_types = F)
+mon_ins <- mon_ins %>%
+  select(`CFF ID`, `Pre or Post`, `Total Daily Dose of insulin (units)`) %>%
+  rename(TDD = `Total Daily Dose of insulin (units)`)
+uw_ins <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_insulin.csv", show_col_types = F)
+uw_ins <- uw_ins %>%
+  select(`CFF ID`, `Pre or Post`, `Total Daily Dose of insulin (units)`) %>%
+  rename(TDD = `Total Daily Dose of insulin (units)`)
+insulin <- bind_rows(chco_ins, mon_ins, uw_ins)
+# Join
+df <- left_join(df, insulin)
+df$`Pre or Post` <- factor(df$`Pre or Post`, levels = c("Pre", "Post"))
+# Get pre value closest to start, post value closest to 1 year from start
+df$select_time = df$Time
+df$select_time[df$`Pre or Post`=="Post"] = 365 - df$select_time[df$`Pre or Post`=="Post"]
+df <- df %>%
+  arrange(`CFF ID`,`Pre or Post`, select_time) %>%
+  group_by(`CFF ID`, `Pre or Post`) %>%
+  summarise(
+    `Num. A1c Values` = sum(!is.na(HbA1c)),
+    across(everything(), ~ last(na.omit(.x))),
+  .groups = "keep") %>%
+  filter(!is.na(HbA1c))
+keep = df %>% group_by(`CFF ID`) %>% summarise(n=n()) %>% filter(n==2) %>% pull(`CFF ID`)
+df = df %>% filter(`CFF ID` %in% keep)
 ```
 
-# Participant characteristics at date closest to CFTR modulator start
+```{r include=FALSE}
+cgm_vars <- c(
+  "average_sensor", "min_sensor", "max_sensor", "percent_time_over_120",
+  "percent_time_over_140", "percent_time_over_200", "excursions_over_140",
+  "excursions_over_200", "percent_time_under_60", "percent_time_under_70",
+  "average_auc_per_day", "standard_deviation", "r_mage"
+)
+# Add CGM information
+# CHCO
+chco_cgm <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_cgm.csv", show_col_types = F)
+chco_cgm$`Pre or Post` <- factor(sapply(str_split(chco_cgm$subject_id, "_"), "[", 2))
+levels(chco_cgm$`Pre or Post`) <- c("Post", "Pre", "Pre")
+chco_cgm$`CFF ID` <- sapply(str_split(chco_cgm$subject_id, "_"), "[", 1)
+chco_cgm <- chco_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(cgm_vars))
+# Montana
+mon_cgm <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_cgm.csv", show_col_types = F)
+mon_cgm$`Pre or Post` <- factor(sapply(str_split(mon_cgm$subject_id, "_"), "[", 2))
+levels(mon_cgm$`Pre or Post`) <- c("Post", "Post", "Pre", "Pre")
+mon_cgm$`CFF ID` <- sapply(str_split(mon_cgm$subject_id, "_"), "[", 1)
+mon_cgm <- mon_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(cgm_vars))
+# UW
+uw_cgm <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_cgm.csv", show_col_types = F)
+uw_cgm$`Pre or Post` <- factor(sapply(str_split(uw_cgm$subject_id, "_"), "[", 2))
+levels(uw_cgm$`Pre or Post`) <- c("Post", "Pre")
+uw_cgm$`CFF ID` <- sapply(str_split(uw_cgm$subject_id, "_"), "[", 1)
+uw_cgm <- uw_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(cgm_vars))
+# Put together
+cgm <- bind_rows(chco_cgm, mon_cgm, uw_cgm)
+# Add to DF
+cgm$`CFF ID` = as.numeric(cgm$`CFF ID`)
+df = left_join(df,cgm)
+```
+
+# Participant characteristics prior to CFTR modulator start
 
 ```{r results='asis'}
-t1_data = df %>% group_by(`CFF ID`) %>% slice_min(abs(Time))
-
-t1 = tableby(~Age+BMI.Percentile+Weight+Sex+,pre_post[pre_post$Before_After == "Before",])
+# Table
+t1 <- tableby(~ Age + `BMI Percentile` + Weight + Sex + ppFEV1 + ppFVC +
+  `G-tube in past 12 months` + `Pancreatic Status` + Site +
+  Genotype + `Age Group`,
+data = df %>% filter(`Pre or Post` == "Pre")
+)
 summary(t1)
@@
 
-```{r}
-summary(tableby( ~ ppFEV + ppFVC + BMIz,data = df[!is.na(df$average_sensor),]))
+## HbA1c and others
+
+```{r results='asis'}
+p <- paired(`Pre or Post` ~ `Num. A1c Values` + HbA1c + Time + Weight + `BMI Percentile` +
+  ppFEV1 + ppFVC + TDD,
+data = df, id = `CFF ID`,
+signed.rank.exact = FALSE, numeric.test = "signed.rank",
+numeric.stats = c("Nmiss", "medianrange", "q1q3")
+)
+summary(p, pfootnote = T, digits = 1)
 ```
+
+## CGM
+
+Only 13 of those with CGM also had HbA1c values pre- and post-CFTR modulator.
+
+```{r results='asis'}
+cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(cgm_vars, collapse = "+")))
+p <- paired(cgm_formula,
+  data = df, id = `CFF ID`,
+  signed.rank.exact = FALSE, numeric.test = "signed.rank",
+  numeric.stats = c("Nmiss", "medianrange", "q1q3")
+)
+summary(p, pfootnote = T, digits = 1)
+```

### 2022-10-07 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
Update pre_post_trikafta_analysis.Rmd

@@
 # Age group
-df$`Age Group` <- cut(df$Age, breaks = c(-Inf, 20, Inf), 
-                      labels = c("Pedatric", "Adult"), right = F)
+df$`Age Group` <- cut(df$Age,
+  breaks = c(-Inf, 20, Inf),
+  labels = c("Pedatric", "Adult"), right = F
+)
 # Race and sex
@@
 # Get pre value closest to start, post value closest to 1 year from start
-df$select_time = df$Time
-df$select_time[df$`Pre or Post`=="Post"] = 365 - df$select_time[df$`Pre or Post`=="Post"]
+df$select_time <- df$Time
+df$select_time[df$`Pre or Post` == "Post"] <- 365 - df$select_time[df$`Pre or Post` == "Post"]
 df <- df %>%
-  arrange(`CFF ID`,`Pre or Post`, select_time) %>%
+  arrange(`CFF ID`, `Pre or Post`, select_time) %>%
   group_by(`CFF ID`, `Pre or Post`) %>%
@@
     across(everything(), ~ last(na.omit(.x))),
-  .groups = "keep") %>%
+    .groups = "keep"
+  ) %>%
   filter(!is.na(HbA1c))
-keep = df %>% group_by(`CFF ID`) %>% summarise(n=n()) %>% filter(n==2) %>% pull(`CFF ID`)
-df = df %>% filter(`CFF ID` %in% keep)
+keep <- df %>%
+  group_by(`CFF ID`) %>%
+  summarise(n = n()) %>%
+  filter(n == 2) %>%
+  pull(`CFF ID`)
+df <- df %>% filter(`CFF ID` %in% keep)
 ```
@@
 # Add to DF
-cgm$`CFF ID` = as.numeric(cgm$`CFF ID`)
-df = left_join(df,cgm)
+cgm$`CFF ID` <- as.numeric(cgm$`CFF ID`)
+df <- left_join(df, cgm)
 ```

### 2022-10-14 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd

@@
 knitr::opts_chunk$set(echo = FALSE)
-knitr::opts_knit$set(root.dir = "/Users/timvigers/Documents/Work/CF/Edith Zemanick/Sweat Test")
+if(Sys.info()["sysname"] == "Windows"){
+  home_dir = "C:/Users/timvigers/Dropbox/Work/Vigers/CF/Edith Zemanick/Sweat Test"
+  source("C:/Users/timvigers/Documents/GitHub/CF/get_genotype_class_severity.R")
+} else if (Sys.info()["sysname"] == "Linux"){
+  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Edith Zemanick/Sweat Test"
   source("~/GitHub/CF/get_genotype_class_severity.R")
+} else if (Sys.info()["sysname"] == "Darwin"){
+  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Edith Zemanick/Sweat Test"
+  source("~/GitHub/CF/get_genotype_class_severity.R")
+}
+knitr::opts_knit$set(root.dir = home_dir)
+rm(home_dir)
+
 ```

### 2022-10-19 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
CFRD missing data for Christine

@@
 library(rspiro)
-source("~/GitHub/CF/get_genotype_class_severity.R")
+source("C:/Users/timvigers/Documents/GitHub/CF/get_genotype_class_severity.R")
 t1_variables <- c(
   "CFF ID", "Site", "Sex", "Race", "Genotypes1", "Genotypes2", "Pancreatic Status",
-  "G-tube in past 12 months", "Time", "Age",
+  "G-tube in past 12 months","Time", "Age",
   "BMI Percentile", "Height", "Weight", "FEV1", "FVC", "HbA1c"
@@
 # CHCO
-chco_a1c <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_glycemic.csv", show_col_types = F)
+chco_a1c <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_glycemic.csv", show_col_types = F)
 chco_a1c$Date <- parse_date_time(chco_a1c$Date, orders = "mdy")
-chco_bmi <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv", show_col_types = F)
+chco_bmi <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv", show_col_types = F)
 chco_bmi$Date <- parse_date_time(chco_bmi$Date, orders = "mdy")
@@
 # Add demographics, filter
-chco_reg <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_registry.csv", show_col_types = F)
+chco_reg <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_registry.csv", show_col_types = F)
 chco_reg$Start <- parse_date_time(chco_reg$Start, orders = "mdy")
@@
   pull(`CFF ID`)
-exclude <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv", show_col_types = F)
+exclude <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv", show_col_types = F)
 exclude <- exclude %>%
@@
 # Montana
-mon_bmi <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_bmi_pft.csv", show_col_types = F)
+mon_bmi <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_bmi_pft.csv", show_col_types = F)
 mon_bmi$Date <- parse_date_time(mon_bmi$Date, orders = "mdy")
-mon_a1c <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv", show_col_types = F)
+mon_a1c <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv", show_col_types = F)
 mon_a1c$Date <- parse_date_time(mon_a1c$Date, orders = "mdy")
@@
 # Add demographics
-mon_demo <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_demo.csv", show_col_types = F)
+mon_demo <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_demo.csv", show_col_types = F)
 mon_demo <- mon_demo %>% select(`CFF ID`, Race, Sex)
@@
 # Add genotypes
-mon_geno <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_genotypes.csv", show_col_types = F)
+mon_geno <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_genotypes.csv", show_col_types = F)
 mon <- left_join(mon, mon_geno)
 # Add pancreatic and g tube status
-mon_pan <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_pancreatic_insufficient.csv", show_col_types = F)
+mon_pan <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_pancreatic_insufficient.csv", show_col_types = F)
 mon_pan <- mon_pan %>%
@@
 mon_pan$`Pancreatic Status` <- "Insufficient"
-mon_gtube <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_gtubes.csv", show_col_types = F)
+mon_gtube <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_gtubes.csv", show_col_types = F)
 mon_gtube <- mon_gtube %>%
@@
 # UM
-um_a1c <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_a1cs.csv", show_col_types = F)
+um_a1c <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_a1cs.csv", show_col_types = F)
 um_a1c$Age <- round(um_a1c$Age / 365.25, 1)
-um_bmi <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_bmi.csv", show_col_types = F)
+um_bmi <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_bmi.csv", show_col_types = F)
 um_bmi$Age <- round(um_bmi$Age / 365.25, 1)
-um_ogtt <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_ogtt.csv", show_col_types = F)
+um_ogtt <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_ogtt.csv", show_col_types = F)
 um_ogtt$Age <- round(um_ogtt$Age / 365.25, 1)
-um_pft <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_pfts.csv", show_col_types = F)
+um_pft <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_pfts.csv", show_col_types = F)
 um_pft$Age <- round(um_pft$Age / 365.25, 1)
-um_race <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_race.csv", show_col_types = F)
-um_demo <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_demo.csv", show_col_types = F)
+um_race <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_race.csv", show_col_types = F)
+um_demo <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_demo.csv", show_col_types = F)
 keep <- um_demo %>%
@@
 um <- um %>%
-  filter(`CFF ID` %in% keep, !is.na(Time), abs(Time) <= 365) %>%
+  filter(`CFF ID` %in% keep, !is.na(Time)) %>%
   rename(Site = "site") %>%
@@
 # UW
-uw_bmi <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_bmi.csv", show_col_types = F)
+uw_bmi <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_bmi.csv", show_col_types = F)
 uw_bmi$Date <- parse_date_time(uw_bmi$Date, orders = "mdy")
-uw_glyc <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_glycemic.csv", show_col_types = F)
+uw_glyc <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_glycemic.csv", show_col_types = F)
 uw_glyc$Date <- parse_date_time(uw_glyc$Date, orders = "mdy")
-uw_pft <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pft.csv", show_col_types = F)
+uw_pft <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pft.csv", show_col_types = F)
 uw_pft$Date <- parse_date_time(uw_pft$Date, orders = "mdy")
-uw_demo <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_demo.csv", show_col_types = F)
+uw_demo <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_demo.csv", show_col_types = F)
 uw_demo <- uw_demo %>% select(`CFF ID`, Race, Sex)
 # Get CFRD status and start date
-uw_start <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_mods.csv", show_col_types = F)
+uw_start <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_mods.csv", show_col_types = F)
 uw_start$`Date of Encounter` <- parse_date_time(uw_start$`Date of Encounter`, orders = "mdy")
@@
   rename(Start = `Date of Encounter`)
-uw_cfrd <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_cfrd.csv", show_col_types = F)
+uw_cfrd <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_cfrd.csv", show_col_types = F)
 uw_cfrd <- uw_cfrd %>%
@@
 # Add genotypes
-uw_geno <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_genotypes.csv", show_col_types = F)
+uw_geno <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_genotypes.csv", show_col_types = F)
 # Pancreatic status
-uw_pan <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pancreatic.csv", show_col_types = F)
+uw_pan <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pancreatic.csv", show_col_types = F)
 uw_pan <- uw_pan %>%
@@
 # G tubes
-uw_gtube <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_gtubes.csv", show_col_types = F)
+uw_gtube <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_gtubes.csv", show_col_types = F)
 uw_gtube <- uw_gtube %>%
@@
 uw <- uw %>% select(all_of(t1_variables))
-df <- bind_rows(chco, mon, um, uw) %>% arrange(`CFF ID`, Time)
+df <- bind_rows(chco, mon,um, uw) %>% arrange(`CFF ID`, Time)
 # Age group
@@
 # Add insulin information
-chco_ins <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv", show_col_types = F)
+chco_ins <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv", show_col_types = F)
 chco_ins$`CFF ID` <- chco_a1c$`CFF ID`[match(chco_ins$MRN, chco_a1c$MRN)]
@@
   filter(!is.na(`CFF ID`))
-mon_ins <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_insulin.csv", show_col_types = F)
+mon_ins <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_insulin.csv", show_col_types = F)
 mon_ins <- mon_ins %>%
@@
   rename(TDD = `Total Daily Dose of insulin (units)`)
-uw_ins <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_insulin.csv", show_col_types = F)
+uw_ins <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_insulin.csv", show_col_types = F)
 uw_ins <- uw_ins %>%
@@
 # CHCO
-chco_cgm <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_cgm.csv", show_col_types = F)
+chco_cgm <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_cgm.csv", show_col_types = F)
 chco_cgm$`Pre or Post` <- factor(sapply(str_split(chco_cgm$subject_id, "_"), "[", 2))
@@
 # Montana
-mon_cgm <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_cgm.csv", show_col_types = F)
+mon_cgm <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_cgm.csv", show_col_types = F)
 mon_cgm$`Pre or Post` <- factor(sapply(str_split(mon_cgm$subject_id, "_"), "[", 2))
@@
 # UW
-uw_cgm <- read_csv("~/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_cgm.csv", show_col_types = F)
+uw_cgm <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_cgm.csv", show_col_types = F)
 uw_cgm$`Pre or Post` <- factor(sapply(str_split(uw_cgm$subject_id, "_"), "[", 2))
@@
 df <- left_join(df, cgm)
+write.csv(df,file = "C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/analysis_data.csv",row.names = F,na="")
 ```
@@
   `G-tube in past 12 months` + `Pancreatic Status` + Site +
-  Genotype + `Age Group`,
+  Genotype + `Age Group` + Race,
 data = df %>% filter(`Pre or Post` == "Pre")
@@
   ppFEV1 + ppFVC + TDD,
-data = df, id = `CFF ID`,
-signed.rank.exact = FALSE, numeric.test = "signed.rank",
+data = df, id = `CFF ID`,strata = Site,
+signed.rank.exact = FALSE, numeric.test = "notest",
 numeric.stats = c("Nmiss", "medianrange", "q1q3")

### 2022-10-20 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
added tables by specific genotype

--- 2022-09-12 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
@@
 knitr::opts_chunk$set(echo = FALSE)
-knitr::opts_knit$set(root.dir = "/Users/timvigers/Documents/Work/CF/Edith Zemanick/Sweat Test")
+if(Sys.info()["sysname"] == "Windows"){
+  home_dir = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Test"
+  source("C:/Users/timvigers/Documents/GitHub/CF/get_genotype_class_severity.R")
+} else if (Sys.info()["sysname"] == "Linux"){
+  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Edith Zemanick/Sweat Test"
   source("~/GitHub/CF/get_genotype_class_severity.R")
+} else if (Sys.info()["sysname"] == "Darwin"){
+  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Edith Zemanick/Sweat Test"
+  source("~/GitHub/CF/get_genotype_class_severity.R")
+}
+knitr::opts_knit$set(root.dir = home_dir)
+rm(home_dir)
 ```
@@
 # Genotype info
+## By severity
 df$geno1 <- get_cf_genotype_class_severity(df$Genotypes1)
@@
   levels = c(
-    "Severe_Severe","Mild_Severe", "Mild_Unknown", "Severe_Mild", "Mild_Mild","Not CF_Not CF",
-
-     "Severe_Unknown", "Unknown_Severe",
-    "Unknown_Unknown"
+    "Severe_Severe","Mild_Severe", "Mild_Unknown", "Severe_Mild", "Mild_Mild",
+    "Not CF_Not CF","Severe_Unknown", "Unknown_Severe","Unknown_Unknown"
   ),
@@
 )
+## By function
+df$geno1_class <- get_cf_genotype_class_severity(df$Genotypes1,out = "class")
+df$geno1_class[grep("508",df$Genotypes1)] = "F508del"
+df$geno1_class = replace(df$geno1_class,df$geno1_class %in% c("I","II","III"),"MF")
+df$geno1_class = replace(df$geno1_class,df$geno1_class %in% c("IV","V"),"RF")
+
+df$geno2_class <- get_cf_genotype_class_severity(df$Genotypes2,out = "class")
+df$geno2_class[grep("508",df$Genotypes2)] = "F508del"
+df$geno2_class = replace(df$geno2_class,df$geno2_class %in% c("I","II","III"),"MF")
+df$geno2_class = replace(df$geno2_class,df$geno2_class %in% c("IV","V"),"RF")
+
+df$geno_class = factor(paste0(pmin(df$geno1_class,df$geno2_class),"/",
+                       pmax(df$geno1_class,df$geno2_class)))
 # Which arm?
@@
 mycontrols  <- tableby.control(numeric.test="kwt",numeric.stats=c("Nmiss", "median", "q1q3"))
-labels = list(Age.at.Test.in.Years = "Age (Years)", geno_risk = "Genotype Severity")
+labels = list(Age.at.Test.in.Years = "Age (Years)", 
+              geno_risk = "Genotype Severity",geno_class = "Specific Genotype")
 ```
@@
 
-For those with two sweat electrolyte measures (left and right), the average of the two was used. Each participants' first sweat test value was used for the following tables
+For those with two sweat electrolyte measures (left and right), the average of the two was used. Each participants' first sweat test value was used for the following tables.
 
@@
   filter(row_number() == 1) # First measure for everyone
-t1 <- tableby(Diagnosis ~ Sex + Age.at.Test.in.Years + fe(geno_risk) + Amount + Cl + Na + K + `Na/Cl Ratio`,
-  data = t1_df,control = mycontrols,simulate.p.value=T
+t1 <- tableby(Diagnosis ~ Sex + Age.at.Test.in.Years + fe(geno_risk) + fe(geno_class) + 
+                Amount + Cl + Na + K + `Na/Cl Ratio`,
+  data = t1_df,control = mycontrols,simulate.p.value = T
 )
@@
 ```{r results='asis'}
-t1 <- tableby(Method ~ Sex + Age.at.Test.in.Years + Diagnosis + geno_risk + Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
-  data = t1_df,control = mycontrols,
+t1 <- tableby(Method ~ Sex + Age.at.Test.in.Years + Diagnosis + fe(geno_risk) + fe(geno_class) +
+                Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + 
+                `Na Diff. Between Left and Right Arm`,
+  data = t1_df,control = mycontrols,simulate.p.value = T
 )
@@
 ```{r results='asis'}
-t1 <- tableby(Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+t1 <- tableby(Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` + 
+                `Cl Diff. Between Left and Right Arm` + 
+                `Na Diff. Between Left and Right Arm`,
   data = t1_df,control = mycontrols,
@@
 ```{r results='asis'}
-t1 <- tableby(Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+t1 <- tableby(Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` + 
+                `Cl Diff. Between Left and Right Arm` + 
+                `Na Diff. Between Left and Right Arm`,
   data = t1_df[t1_df$Method == "Gibson-Cooke",],control = mycontrols,
@@
 ```{r results='asis'}
-t1 <- tableby(Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+t1 <- tableby(Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` + 
+                `Cl Diff. Between Left and Right Arm` + 
+                `Na Diff. Between Left and Right Arm`,
   data = t1_df[t1_df$Method == "Macroduct",],control = mycontrols,
@@
 ```{r results='asis'}
-t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + 
+                `Cl Diff. Between Left and Right Arm` + 
+                `Na Diff. Between Left and Right Arm`,
   data = t1_df,control = mycontrols,
@@
 
-## Plots
+### Plots
 
@@
 
+## By specific genotype
+
+```{r results='asis'}
+t1 <- tableby(geno_class ~ Amount + Cl + Na + K + `Na/Cl Ratio` + 
+                `Cl Diff. Between Left and Right Arm` + 
+                `Na Diff. Between Left and Right Arm`,
+  data = t1_df,control = mycontrols,
+)
+summary(t1,pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+## By specific genotype and diagnosis
+
+### Cystic fibrosis
+
+```{r results='asis'}
+t1 <- tableby(geno_class ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+  data = t1_df[t1_df$Diagnosis == "Cystic Fibrosis",],control = mycontrols
+)
+summary(t1,pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### CFTR-related disorder
+
+```{r results='asis'}
+t1 <- tableby(geno_class ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+  data = t1_df[t1_df$Diagnosis == "CFTR-related disorder",],control = mycontrols
+)
+summary(t1,pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### CRMS
+
+```{r results='asis'}
+t1 <- tableby(geno_class ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+  data = t1_df[t1_df$Diagnosis == "CRMS",],control = mycontrols
+)
+summary(t1,pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### Non-CF
+
+```{r results='asis'}
+t1 <- tableby( ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+  data = t1_df[t1_df$Diagnosis == "Non-CF",],control = mycontrols
+)
+summary(t1,pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### Plots
+
+```{r warning=FALSE}
+ggplot(df, aes(x = geno_class, y = Na)) +
+  geom_boxplot() +
+  theme_bw() + xlab("Specific Genotype") +
+  ggtitle("Sweat Sodium")
+
+ggplot(df, aes(x = geno_class, y = K)) +
+  geom_boxplot() +
+  theme_bw() +
+  ggtitle("Sweat Potassium")
+
+ggplot(df, aes(x = geno_class, y = Cl)) +
+  geom_boxplot() +
+  theme_bw() + xlab("Specific Genotype") +
+  ggtitle("Sweat Chloride")
+
+ggplot(df, aes(x = geno_class, y = `Na/Cl Ratio`)) +
+  geom_boxplot() +
+  theme_bw() + xlab("Specific Genotype") +
+  ggtitle("Na/Cl Ratio")
+```
+
 # Between occasion
@@
 ```{r results='asis'}
-p <- paired(time ~ Diagnosis +Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,signed.rank.exact = F,
-            data = multiples, id = PatientID,control = mycontrols)
+p <- paired(time ~ Diagnosis + Cl + Na + K + `Na/Cl Ratio` + 
+              `Cl Diff. Between Left and Right Arm` + 
+              `Na Diff. Between Left and Right Arm`,
+            signed.rank.exact = F,data = multiples, 
+            id = PatientID,control = mycontrols)
 summary(p,pfootnote = T,labelTranslations = labels)
@@
 p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio`,signed.rank.exact = F,strata = geno_risk,
+            data = multiples %>% filter(Diagnosis == "Non-CF"), 
+            id = PatientID,control = mycontrols)
+summary(p,pfootnote = T,labelTranslations = labels)
+```
+
+## By specific genotype
+
+```{r results='asis'}
+p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio` + 
+              `Cl Diff. Between Left and Right Arm` + 
+              `Na Diff. Between Left and Right Arm`,
+            signed.rank.exact = F,strata = geno_class,numeric.test = "notest",
+            data = multiples, id = PatientID,control = mycontrols)
+summary(p,pfootnote = T,labelTranslations = labels)
+```
+
+## By genotype class and diagnosis
+
+Not enough paired observations in the CRMS or CFTR-related disorder groups.
+
+### Cystic fibrosis
+
+```{r results='asis'}
+p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio`,
+            signed.rank.exact = F,strata = geno_class,
+            data = multiples %>% filter(Diagnosis == "Cystic Fibrosis"), 
+            id = PatientID,control = mycontrols)
+summary(p,pfootnote = T,labelTranslations = labels)
+```
+
+### Non-CF
+
+```{r results='asis'}
+p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio`,signed.rank.exact = F,strata = geno_class,
             data = multiples %>% filter(Diagnosis == "Non-CF"), 

### 2022-10-24 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
colleen wood CGM report mostly done

@@
 # CHCO
-chco_a1c <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_glycemic.csv", show_col_types = F)
+chco_a1c <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_glycemic.csv", show_col_types = F)
 chco_a1c$Date <- parse_date_time(chco_a1c$Date, orders = "mdy")
-chco_bmi <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv", show_col_types = F)
+chco_bmi <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv", show_col_types = F)
 chco_bmi$Date <- parse_date_time(chco_bmi$Date, orders = "mdy")
@@
 # Add demographics, filter
-chco_reg <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_registry.csv", show_col_types = F)
+chco_reg <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_registry.csv", show_col_types = F)
 chco_reg$Start <- parse_date_time(chco_reg$Start, orders = "mdy")
@@
   pull(`CFF ID`)
-exclude <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv", show_col_types = F)
+exclude <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv", show_col_types = F)
 exclude <- exclude %>%
@@
 # Montana
-mon_bmi <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_bmi_pft.csv", show_col_types = F)
+mon_bmi <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_bmi_pft.csv", show_col_types = F)
 mon_bmi$Date <- parse_date_time(mon_bmi$Date, orders = "mdy")
-mon_a1c <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv", show_col_types = F)
+mon_a1c <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv", show_col_types = F)
 mon_a1c$Date <- parse_date_time(mon_a1c$Date, orders = "mdy")
@@
 # Add demographics
-mon_demo <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_demo.csv", show_col_types = F)
+mon_demo <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_demo.csv", show_col_types = F)
 mon_demo <- mon_demo %>% select(`CFF ID`, Race, Sex)
@@
 # Add genotypes
-mon_geno <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_genotypes.csv", show_col_types = F)
+mon_geno <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_genotypes.csv", show_col_types = F)
 mon <- left_join(mon, mon_geno)
 # Add pancreatic and g tube status
-mon_pan <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_pancreatic_insufficient.csv", show_col_types = F)
+mon_pan <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_pancreatic_insufficient.csv", show_col_types = F)
 mon_pan <- mon_pan %>%
@@
 mon_pan$`Pancreatic Status` <- "Insufficient"
-mon_gtube <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_gtubes.csv", show_col_types = F)
+mon_gtube <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_gtubes.csv", show_col_types = F)
 mon_gtube <- mon_gtube %>%
@@
 # UM
-um_a1c <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_a1cs.csv", show_col_types = F)
+um_a1c <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_a1cs.csv", show_col_types = F)
 um_a1c$Age <- round(um_a1c$Age / 365.25, 1)
-um_bmi <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_bmi.csv", show_col_types = F)
+um_bmi <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_bmi.csv", show_col_types = F)
 um_bmi$Age <- round(um_bmi$Age / 365.25, 1)
-um_ogtt <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_ogtt.csv", show_col_types = F)
+um_ogtt <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_ogtt.csv", show_col_types = F)
 um_ogtt$Age <- round(um_ogtt$Age / 365.25, 1)
-um_pft <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_pfts.csv", show_col_types = F)
+um_pft <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_pfts.csv", show_col_types = F)
 um_pft$Age <- round(um_pft$Age / 365.25, 1)
-um_race <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_race.csv", show_col_types = F)
-um_demo <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_demo.csv", show_col_types = F)
+um_race <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_race.csv", show_col_types = F)
+um_demo <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_demo.csv", show_col_types = F)
 keep <- um_demo %>%
@@
 # UW
-uw_bmi <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_bmi.csv", show_col_types = F)
+uw_bmi <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_bmi.csv", show_col_types = F)
 uw_bmi$Date <- parse_date_time(uw_bmi$Date, orders = "mdy")
-uw_glyc <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_glycemic.csv", show_col_types = F)
+uw_glyc <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_glycemic.csv", show_col_types = F)
 uw_glyc$Date <- parse_date_time(uw_glyc$Date, orders = "mdy")
-uw_pft <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pft.csv", show_col_types = F)
+uw_pft <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pft.csv", show_col_types = F)
 uw_pft$Date <- parse_date_time(uw_pft$Date, orders = "mdy")
-uw_demo <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_demo.csv", show_col_types = F)
+uw_demo <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_demo.csv", show_col_types = F)
 uw_demo <- uw_demo %>% select(`CFF ID`, Race, Sex)
 # Get CFRD status and start date
-uw_start <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_mods.csv", show_col_types = F)
+uw_start <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_mods.csv", show_col_types = F)
 uw_start$`Date of Encounter` <- parse_date_time(uw_start$`Date of Encounter`, orders = "mdy")
@@
   rename(Start = `Date of Encounter`)
-uw_cfrd <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_cfrd.csv", show_col_types = F)
+uw_cfrd <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_cfrd.csv", show_col_types = F)
 uw_cfrd <- uw_cfrd %>%
@@
 # Add genotypes
-uw_geno <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_genotypes.csv", show_col_types = F)
+uw_geno <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_genotypes.csv", show_col_types = F)
 # Pancreatic status
-uw_pan <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pancreatic.csv", show_col_types = F)
+uw_pan <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pancreatic.csv", show_col_types = F)
 uw_pan <- uw_pan %>%
@@
 # G tubes
-uw_gtube <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_gtubes.csv", show_col_types = F)
+uw_gtube <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_gtubes.csv", show_col_types = F)
 uw_gtube <- uw_gtube %>%
@@
 # Add insulin information
-chco_ins <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv", show_col_types = F)
+chco_ins <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv", show_col_types = F)
 chco_ins$`CFF ID` <- chco_a1c$`CFF ID`[match(chco_ins$MRN, chco_a1c$MRN)]
 chco_ins <- chco_ins %>%
-  select(`CFF ID`, `Pre or Post`, `Total Daily Dose of insulin (units)`) %>%
-  rename(TDD = `Total Daily Dose of insulin (units)`) %>%
+  rename(TDD = `Total Daily Dose of insulin (units)`,
+         `Long acting insulin dose` = `Long acting or total basal daily dose (units)`) %>%
+  select(`CFF ID`, `Pre or Post`,TDD,`Long acting insulin dose`) %>%
   filter(!is.na(`CFF ID`))
-mon_ins <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_insulin.csv", show_col_types = F)
+mon_ins <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_insulin.csv", show_col_types = F)
 mon_ins <- mon_ins %>%
-  select(`CFF ID`, `Pre or Post`, `Total Daily Dose of insulin (units)`) %>%
-  rename(TDD = `Total Daily Dose of insulin (units)`)
-uw_ins <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_insulin.csv", show_col_types = F)
+  rename(TDD = `Total Daily Dose of insulin (units)`,
+         `Long acting insulin dose` = `Long acting or total basal daily dose (units)`) %>%
+  select(`CFF ID`, `Pre or Post`, TDD,`Long acting insulin dose`)
+uw_ins <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_insulin.csv", show_col_types = F)
 uw_ins <- uw_ins %>%
-  select(`CFF ID`, `Pre or Post`, `Total Daily Dose of insulin (units)`) %>%
-  rename(TDD = `Total Daily Dose of insulin (units)`)
+  rename(TDD = `Total Daily Dose of insulin (units)`,
+         `Long acting insulin dose` = `Long acting or total basal daily dose (units)`) %>%
+  select(`CFF ID`, `Pre or Post`, TDD,`Long acting insulin dose`)
 insulin <- bind_rows(chco_ins, mon_ins, uw_ins)
@@
 cgm_vars <- c(
-  "average_sensor", "min_sensor", "max_sensor", "percent_time_over_120",
-  "percent_time_over_140", "percent_time_over_200", "excursions_over_140",
-  "excursions_over_200", "percent_time_under_60", "percent_time_under_70",
-  "average_auc_per_day", "standard_deviation", "r_mage"
+  "average_sensor", "min_sensor", "max_sensor", 
+  "percent_time_over_180","percent_time_70_180",
+  "percent_time_over_120",
+  "percent_time_over_140", "percent_time_over_200", 
+  "excursions_over_140","excursions_over_200", 
+  "percent_time_under_60", "percent_time_under_70",
+  "average_auc_per_day", "standard_deviation", "cv","r_mage"
+)
+day_cgm = c("daytime_avg_sensor_glucose",	"daytime_min_sensor_glucose",	
+            "daytime_max_sensor_glucose",	"percent_time_over_180_day",
+            "percent_time_70_180_day","percent_time_over_200_day",
+            "percent_time_under_60_day","percent_time_under_70_day","daytime_sd"
+            )
+night_cgm = c("nighttime_avg_sens_glucose",	"nighttime_min_sens_glucose",	
+            "nighttime_max_sens_glucose",	"percent_time_over_180_night",
+            "percent_time_70_180_night","percent_time_over_200_night",
+            "percent_time_under_60_night","percent_time_under_70_night","nighttime_sd"
             )
@@
 # CHCO
-chco_cgm <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_cgm.csv", show_col_types = F)
+chco_cgm <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_cgm.csv", show_col_types = F)
 chco_cgm$`Pre or Post` <- factor(sapply(str_split(chco_cgm$subject_id, "_"), "[", 2))
@@
 chco_cgm$`CFF ID` <- sapply(str_split(chco_cgm$subject_id, "_"), "[", 1)
-chco_cgm <- chco_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(cgm_vars))
+chco_cgm <- chco_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(c(cgm_vars,day_cgm,night_cgm)))
 # Montana
-mon_cgm <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_cgm.csv", show_col_types = F)
+mon_cgm <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_cgm.csv", show_col_types = F)
 mon_cgm$`Pre or Post` <- factor(sapply(str_split(mon_cgm$subject_id, "_"), "[", 2))
@@
 mon_cgm$`CFF ID` <- sapply(str_split(mon_cgm$subject_id, "_"), "[", 1)
-mon_cgm <- mon_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(cgm_vars))
+mon_cgm <- mon_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(c(cgm_vars,day_cgm,night_cgm)))
 # UW
-uw_cgm <- read_csv("C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_cgm.csv", show_col_types = F)
+uw_cgm <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_cgm.csv", show_col_types = F)
 uw_cgm$`Pre or Post` <- factor(sapply(str_split(uw_cgm$subject_id, "_"), "[", 2))
@@
 uw_cgm$`CFF ID` <- sapply(str_split(uw_cgm$subject_id, "_"), "[", 1)
-uw_cgm <- uw_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(cgm_vars))
+uw_cgm <- uw_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(c(cgm_vars,day_cgm,night_cgm)))
 # Put together
@@
 df <- left_join(df, cgm)
-write.csv(df,file = "C:/Users/timvigers/Dropbox/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/analysis_data.csv",row.names = F,na="")
+write.csv(df,file = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/analysis_data.csv",row.names = F,na="")
 ```
@@
 
+### All groups
+
+```{r results='asis'}
+p <- paired(`Pre or Post` ~ `Num. A1c Values` + HbA1c + Time + Weight + `BMI Percentile` +
+  ppFEV1 + ppFVC + TDD + `Long acting insulin dose`,
+data = df, id = `CFF ID`,
+signed.rank.exact = FALSE, 
+numeric.stats = c("Nmiss", "medianrange", "q1q3")
+)
+summary(p, pfootnote = T, digits = 1)
+```
+
+### By site
+
 ```{r results='asis'}
@@
 
+### Demographics
+
+```{r results='asis'}
+cgm = df[!is.na(df$average_sensor),]
+# Table
+t1 <- tableby(~ Age + `BMI Percentile` + Weight + Sex + ppFEV1 + ppFVC +
+  `G-tube in past 12 months` + `Pancreatic Status` + Site +
+  Genotype + `Age Group` + Race,
+data = cgm %>% filter(`Pre or Post` == "Pre")
+)
+summary(t1)
+```
+
+### Overall CGM variables
+
 ```{r results='asis'}
 cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(cgm_vars, collapse = "+")))
+cgm_formula = update(cgm_formula,.~.+TDD + `Long acting insulin dose`)
 p <- paired(cgm_formula,
-  data = df, id = `CFF ID`,
+  data = cgm, id = `CFF ID`,
   signed.rank.exact = FALSE, numeric.test = "signed.rank",
@@
 ```
+
+### Daytime CGM variables
+
+```{r results='asis'}
+cgm$daytime_cv = cgm$daytime_sd / cgm$daytime_avg_sensor_glucose
+day_cgm = c(day_cgm,"daytime_cv")
+cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(day_cgm, collapse = "+")))
+p <- paired(cgm_formula,
+  data = cgm, id = `CFF ID`,
+  signed.rank.exact = FALSE, numeric.test = "signed.rank",
+  numeric.stats = c("Nmiss", "medianrange", "q1q3")
+)
+summary(p, pfootnote = T, digits = 1)
+```
+
+### Nighttime CGM variables
+
+```{r results='asis'}
+cgm$nighttime_cv = cgm$nighttime_sd / cgm$nighttime_avg_sens_glucose
+night_cgm = c(night_cgm,"nighttime_cv")
+cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(night_cgm, collapse = "+")))
+p <- paired(cgm_formula,
+  data = cgm, id = `CFF ID`,
+  signed.rank.exact = FALSE, numeric.test = "signed.rank",
+  numeric.stats = c("Nmiss", "medianrange", "q1q3")
+)
+summary(p, pfootnote = T, digits = 1)
+```

### 2022-11-01 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
manuscript changes for CFRD and ETI

@@
 knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
-knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/CF")
+knitr::opts_knit$set(root.dir = "~/Dropbox/Work/Vigers/CF")
 ```
@@
 ```{r results='asis'}
-t2_vars = c("A1C.result","Weight..kg.","weightz","BMIz","FEV1..","FVC....",
-            "trikafta_duration","Group","Glucose.0.minute",
+t2_vars = c("Age.at.visit.","A1C.result","Weight..kg.","weightz","BMI","BMIz",
+            "FEV1..","FVC....","trikafta_duration","Group","Glucose.0.minute",
             "Glucose.60.min","Glucose.120.min","c.peptide.0.minute",
@@
 
-# Table 4: Prior Modulator Use
+# Prior Modulator Use
 
-Unable to perform statistical tests due to too many missing values.
+The following tables include 11 people previously on modulators and 9 not previously on modulators.
+
+## Table 4a: Pre-/Post-ETI Stratified by Previous Use
+
+```{r results='asis'}
+df$prev = 
+  factor(paste(df$Kalydeco,df$Symdeko, df$Orkambi),
+         levels = c("No No No","No No Yes","No Yes No","Yes No No"),
+         labels = c("No","Yes","Yes","Yes")
+  )
+df = df %>% group_by(Study.ID) %>% mutate(`Previous Modulator` = prev[Event.Name == "Visit 1"])
+
+prior_mod_vars = t2_vars[as.numeric(which(colSums(is.na(df[,gsub("`","",t2_vars)]))==0))]
+f = as.formula(paste0("Event.Name ~ ",paste0(prior_mod_vars,collapse = "+")))
+
+p <- paired(f,data = df, id = Study.ID,numeric.test = "signed.rank",cat.test = "mcnemar",signed.rank.exact = F,
+            strata = `Previous Modulator`,control = tableby.control(numeric.stats = c("Nmiss", "medianq1q3", "range")))
+summary(p,pfootnote = T,labelTranslations = newlabels)
+```
+
+## Table 4b: Previous Use vs. Not - Baseline
+
+```{r results='asis'}
+baseline = df %>% filter(Event.Name == "Visit 1")
+f = as.formula(paste0("`Previous Modulator` ~ ",paste0(prior_mod_vars,collapse = "+")))
+t = tableby(f,numeric.test = "kwt",data = baseline,
+            control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))))
+summary(t,digits = 2,labelTranslations = newlabels,pfootnote=TRUE)
+```
+
+## Table 4c: Previous Use vs. Not - Follow Up
+
+```{r results='asis'}
+visit2 = df %>% filter(Event.Name == "Visit 2")
+t = tableby(f,numeric.test = "kwt",data = visit2,
+            control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))))
+summary(t,digits = 2,labelTranslations = newlabels,pfootnote=TRUE)
+```
 
@@
 ggplot(df,aes(x = Event.Name, y = A1C.result,fill = Event.Name)) + 
-  geom_boxplot(coef=0,alpha = 0.5,outlier.shape = NA) + geom_jitter(width = 0.15) + 
+  geom_boxplot(coef=0,alpha = 0.5,outlier.shape = NA) + 
+  geom_point(alpha = 0.3) + 
+  geom_line(aes(group = Study.ID),alpha = 0.3) + 
   xlab("Visit") + ylab("HbA1c (%)") + 
-  theme_bw() + theme(legend.position = "none") +
+  theme_bw() + 
+  theme(legend.position = "none",axis.title.x=element_blank()) +
   scale_fill_grey()

### 2022-11-02 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
updated figure for pre/post ETI

@@
 
-## Glucose
+## C-Peptide
+
+### All
 
@@
   theme(legend.title=element_blank()) + scale_color_grey()
+```
+
+### Pediatric only
+
+```{r}
+plot_df = df %>% pivot_longer(all_of(cpep_cols)) %>% 
+  select(Study.ID,Event.Name,name,value,`Age Group`) %>%
+  filter(`Age Group` == "Child")
+plot_df$name = as.numeric(gsub("c.peptide|\\.|min|minute","",plot_df$name))
+plot_df = plot_df %>% group_by(name,Event.Name) %>% 
+  summarise(m = median(value,na.rm = T),
+            low = as.numeric(quantile(value,0.25)),
+            high = as.numeric(quantile(value,0.75)),
+            .groups = "drop")
+ggplot(plot_df,aes(x = name,y = m,color = Event.Name)) + 
+  geom_line() + geom_point() + 
+  geom_errorbar(aes(ymin=low, ymax=high), width=2) +
+  xlab("Time (minutes)") + ylab("C Peptide (pmol/L)") + 
+  theme_bw() +
+  theme(legend.title=element_blank()) + scale_color_grey()
+```
+
+### Pediatric vs. Adult
+
+```{r}
+plot_df = df %>% pivot_longer(all_of(cpep_cols)) %>% 
+  select(Study.ID,Event.Name,name,value,`Age Group`)
+plot_df$name = as.numeric(gsub("c.peptide|\\.|min|minute","",plot_df$name))
+plot_df = plot_df %>% group_by(name,Event.Name,`Age Group`) %>% 
+  summarise(m = median(value,na.rm = T),
+            low = as.numeric(quantile(value,0.25)),
+            high = as.numeric(quantile(value,0.75)),
+            .groups = "drop")
+ggplot(plot_df,aes(x = name,y = m,color = Event.Name)) + 
+  geom_line() + geom_point() + 
+  geom_errorbar(aes(ymin=low, ymax=high), width=2) +
+  xlab("Time (minutes)") + ylab("C Peptide (pmol/L)") + 
+  theme_bw() +
+  theme(legend.title=element_blank()) + scale_color_grey() + 
+  facet_grid(~`Age Group`)
 ```

### 2022-11-02 Christine Chan/CF Registry/cf_registry_analysis.Rmd
registry analysis formatting

@@
 knitr::opts_chunk$set(echo = FALSE)
-knitr::opts_knit$set(root.dir = "/Users/timvigers/Documents/Work/CF/Christine Chan/CF Registry")
+knitr::opts_knit$set(root.dir = "~/Dropbox/Work/CHCO/Christine Chan/CF registry data")
 ```
 
-```{r}
+```{r eval=FALSE}
 # Read in CSVs
-df <- read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv")
-annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv")
+df <- read.csv("CFF19_encountersMerged_Del1.csv")
+annual <- read.csv("CFF19_Annualized_Del1.csv")
 colnames(annual)[colnames(annual) == "ReviewYear"] <- "reviewyear"
-demo <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
+demo <- read.csv("CFF19_DemogCFDiag_Del1.csv")
 # Combine and format demographics columns
@@
 df$employment <- as.factor(df$employment)
-# Small dataframe for checking
-test <- df[1:1000, ]
-write.csv(test, "~/test.csv", row.names = F, na = "")
+# Save formatted data for re-import later
+save(df,file = "formatted_data.Rdata")
+```
+
+```{r}
+load("formatted_data.Rdata")
 ```

### 2022-11-16 Stacey Martiniano/ntm_agreement.qmd
started ntm agreement

---
title: "NTM CT Scoring - Initial Agreement"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    self-contained: true
editor: visual
---

```{r}
#| include: false
library(dplyr)
library(irr)
library(redcapAPI)
# Import data
api <- read.table("~/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt",header = F)
rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/",
                         token = api[1,1])
data = exportRecords(rcon)
# Make scorer IDs by initial
data$scorer = paste0(toupper(substr(data$reader_first_name,1,1)),
                     toupper(substr(data$reader_last_name,1,1)))
data = data %>% arrange(scan_id,scorer) %>% filter(scan_id>2) %>% 
  select(scorer,scan_id,bronchiectasis_rul:atelectasis___6)
```

# Scan 3

## Bronchiectasis

```{r}
# Get relevant columns and prepare for Fleiss' kappa calculation
bronch = data %>% filter(scan_id==3) %>% 
  select(scorer,bronchiectasis_rul:bronchiectasis_lll)
rows = as.character(label(bronch)[-1])
cols = bronch$scorer
bronch = data.frame(t(bronch[,-1]))
rownames(bronch) = rows
colnames(bronch) = cols
```

### 2022-11-28 Stacey Martiniano/ntm_agreement.qmd
better agreement plots

@@
   html:
+    code-fold: true
     toc: true
@@
     self-contained: true
+    fig-height: 8
+    fig-width: 10
 editor: visual
@@
 #| include: false
-library(dplyr)
+library(tidyverse)
 library(irr)
 library(redcapAPI)
+library(Hmisc)
+library(DT)
+library(RColorBrewer)
 # Import data
-api <- read.table("~/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt",header = F)
+api <- read.table("~/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt",
+                  header = F)
 rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/",
                          token = api[1,1])
-data = exportRecords(rcon)
+data = exportRecords(rcon,labels = F)
+data = data.frame(lapply(data,as.character))
 # Make scorer IDs by initial
@@
                      toupper(substr(data$reader_last_name,1,1)))
+# Select relevant columns, convert to long data
 data = data %>% arrange(scan_id,scorer) %>% filter(scan_id>2) %>% 
   select(scorer,scan_id,bronchiectasis_rul:atelectasis___6)
+data = data %>% pivot_longer(bronchiectasis_rul:atelectasis___6,
+                             names_to = c("feature","location"),
+                             names_pattern = "(.*)_(.*)$") %>%
+  mutate(feature=gsub("__","",feature),
+         feature = str_to_title(gsub("_"," ",feature)))
+data$location = factor(data$location,
+                       levels = c("1","2","3","4","5","6","ling","lll","lul","rll","rml","rul"),
+                       labels = c("RUL","RML","RLL","LUL","Ling","LLL","Ling","LLL","LUL","RLL","RML","RUL"))
+data$value = factor(data$value,levels = c("Checked","Unchecked","Absent","Cylindric","Varicose"))
+vals = levels(data$value)
+# One column per scorer
+data = data %>% pivot_wider(names_from = scorer,values_from = value)
+# Find disagreements
+data$disagree = !(data$AF == data$DL & data$DL == data$JW)
+# Set colors for whole document
+n = length(vals)
+name = "Accent"
+colors <- brewer.pal(n,name)
+names(colors) <- vals
+custom_colors <- scale_fill_manual(name = "", values = colors)
 ```
 
-# Scan 3
+# All disagreements
 
-## Bronchiectasis
+The table below is sortable and searchable, and includes only the features and locations where there was some disagreement.
 
 ```{r}
-# Get relevant columns and prepare for Fleiss' kappa calculation
-bronch = data %>% filter(scan_id==3) %>% 
-  select(scorer,bronchiectasis_rul:bronchiectasis_lll)
-rows = as.character(label(bronch)[-1])
-cols = bronch$scorer
-bronch = data.frame(t(bronch[,-1]))
-rownames(bronch) = rows
-colnames(bronch) = cols
+# Table
+data %>% filter(disagree == T) %>% select(-disagree) %>% datatable(.,rownames = F)
 ```
+# Plots
+
+## All scans
+
+```{r}
+data %>% 
+  pivot_longer(AF:JW) %>%
+  ggplot(.,aes(x=name,y=location,fill=value))+
+  geom_tile(color = "black") + 
+  theme_bw() + labs(fill="",x="Scorer",y="Location") +
+  custom_colors+
+  facet_grid(rows = vars(scan_id),cols = vars(feature))
+```
+
+## Scan 3
+
+```{r}
+scan = 3
+data %>% filter(scan_id == scan) %>%
+  select(feature,location,AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(.,aes(x=name,y=location,fill=value))+
+  geom_tile(color = "black") + 
+  theme_bw() + labs(fill="",x="Scorer",y="Location") +
+  custom_colors+
+  facet_wrap(~feature)
+```
+
+## Scan 4
+
+```{r}
+scan = 4
+data %>% filter(scan_id == scan) %>%
+  select(feature,location,AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(.,aes(x=name,y=location,fill=value))+
+  geom_tile(color = "black") + 
+  theme_bw() + labs(fill="",x="Scorer",y="Location") +
+  custom_colors+
+  facet_wrap(~feature)
+```
+
+## Scan 5
+
+```{r}
+scan = 5
+data %>% filter(scan_id == scan) %>%
+  select(feature,location,AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(.,aes(x=name,y=location,fill=value))+
+  geom_tile(color = "black") + 
+  theme_bw() + labs(fill="",x="Scorer",y="Location") +
+  custom_colors+
+  facet_wrap(~feature)
+```
+
+# Fleiss' Kappa by Feature
+
+```{r}
+
+```

### 2022-11-28 Stacey Martiniano/ntm_agreement.qmd
Update ntm_agreement.qmd
added kappas

@@
 
-# Fleiss' Kappa by Feature
+# Agreement by Feature
+
+## Atelectasis
 
 ```{r}
+feat = "Atelectasis"
+# Statistic
+k=data %>% filter(feature==feat) %>% select(AF:JW) %>% kappam.fleiss(.)
+ka= round(k$value,3)
+p = format.pval(k$p.value,eps=0.001,digits = 3)
+# Plot
+data %>% filter(feature==feat) %>%
+  select(scan_id,location,AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(.,aes(x=name,y=location,fill=value))+
+  geom_tile(color = "black") + 
+  theme_bw() + labs(fill="",x="Scorer",y="Location") +
+  custom_colors+
+  facet_wrap(~scan_id)
+```
 
+Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
+
+## Bronchiectasis
+
+```{r}
+feat = "Bronchiectasis"
+# Statistic
+k=data %>% filter(feature==feat) %>% select(AF:JW) %>% kappam.fleiss(.)
+ka= round(k$value,3)
+p = format.pval(k$p.value,eps=0.001,digits = 3)
+# Plot
+data %>% filter(feature==feat) %>%
+  select(scan_id,location,AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(.,aes(x=name,y=location,fill=value))+
+  geom_tile(color = "black") + 
+  theme_bw() + labs(fill="",x="Scorer",y="Location") +
+  custom_colors+
+  facet_wrap(~scan_id)
 ```
+
+Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
+
+## Cavity
+
+```{r}
+feat = "Cavity"
+# Statistic
+k=data %>% filter(feature==feat) %>% select(AF:JW) %>% kappam.fleiss(.)
+ka= round(k$value,3)
+p = format.pval(k$p.value,eps=0.001,digits = 3)
+# Plot
+data %>% filter(feature==feat) %>%
+  select(scan_id,location,AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(.,aes(x=name,y=location,fill=value))+
+  geom_tile(color = "black") + 
+  theme_bw() + labs(fill="",x="Scorer",y="Location") +
+  custom_colors+
+  facet_wrap(~scan_id)
+```
+
+Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
+
+## Consolidation
+
+```{r}
+feat = "Consolidation"
+# Statistic
+k=data %>% filter(feature==feat) %>% select(AF:JW) %>% kappam.fleiss(.)
+ka= round(k$value,3)
+p = format.pval(k$p.value,eps=0.001,digits = 3)
+# Plot
+data %>% filter(feature==feat) %>%
+  select(scan_id,location,AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(.,aes(x=name,y=location,fill=value))+
+  geom_tile(color = "black") + 
+  theme_bw() + labs(fill="",x="Scorer",y="Location") +
+  custom_colors+
+  facet_wrap(~scan_id)
+```
+
+Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
+
+## Cyst
+
+```{r}
+feat = "Cyst"
+# Statistic
+k=data %>% filter(feature==feat) %>% select(AF:JW) %>% kappam.fleiss(.)
+ka= round(k$value,3)
+p = format.pval(k$p.value,eps=0.001,digits = 3)
+# Plot
+data %>% filter(feature==feat) %>%
+  select(scan_id,location,AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(.,aes(x=name,y=location,fill=value))+
+  geom_tile(color = "black") + 
+  theme_bw() + labs(fill="",x="Scorer",y="Location") +
+  custom_colors+
+  facet_wrap(~scan_id)
+```
+
+Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
+
+## Ground Glass
+
+```{r}
+feat = "Ground Glass"
+# Statistic
+k=data %>% filter(feature==feat) %>% select(AF:JW) %>% kappam.fleiss(.)
+ka= round(k$value,3)
+p = format.pval(k$p.value,eps=0.001,digits = 3)
+# Plot
+data %>% filter(feature==feat) %>%
+  select(scan_id,location,AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(.,aes(x=name,y=location,fill=value))+
+  geom_tile(color = "black") + 
+  theme_bw() + labs(fill="",x="Scorer",y="Location") +
+  custom_colors+
+  facet_wrap(~scan_id)
+```
+
+Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
+
+## Mucoid Impaction
+
+```{r}
+feat = "Mucoid Impaction"
+# Statistic
+k=data %>% filter(feature==feat) %>% select(AF:JW) %>% kappam.fleiss(.)
+ka= round(k$value,3)
+p = format.pval(k$p.value,eps=0.001,digits = 3)
+# Plot
+data %>% filter(feature==feat) %>%
+  select(scan_id,location,AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(.,aes(x=name,y=location,fill=value))+
+  geom_tile(color = "black") + 
+  theme_bw() + labs(fill="",x="Scorer",y="Location") +
+  custom_colors+
+  facet_wrap(~scan_id)
+```
+
+Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
+
+## Nodule
+
+```{r}
+feat = "Nodule"
+# Statistic
+k=data %>% filter(feature==feat) %>% select(AF:JW) %>% kappam.fleiss(.)
+ka= round(k$value,3)
+p = format.pval(k$p.value,eps=0.001,digits = 3)
+# Plot
+data %>% filter(feature==feat) %>%
+  select(scan_id,location,AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(.,aes(x=name,y=location,fill=value))+
+  geom_tile(color = "black") + 
+  theme_bw() + labs(fill="",x="Scorer",y="Location") +
+  custom_colors+
+  facet_wrap(~scan_id)
+```
+
+Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
+
+## Tree in Bud
+
+```{r}
+feat = "Tree In Bud"
+# Statistics
+k=data %>% filter(feature==feat) %>% select(AF:JW) %>% kappam.fleiss(.)
+ka= round(k$value,3)
+p = format.pval(k$p.value,eps=0.001,digits = 3)
+# Plot
+data %>% filter(feature==feat) %>%
+  select(scan_id,location,AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(.,aes(x=name,y=location,fill=value))+
+  geom_tile(color = "black") + 
+  theme_bw() + labs(fill="",x="Scorer",y="Location") +
+  custom_colors+
+  facet_wrap(~scan_id)
+```
+
+Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.

### 2022-11-28 Stacey Martiniano/ntm_agreement.qmd
Update ntm_agreement.qmd
tidyverse styling

--- 2022-11-16 Stacey Martiniano/ntm_agreement.qmd
@@
 #| include: false
-library(dplyr)
+library(tidyverse)
 library(irr)
 library(redcapAPI)
+library(Hmisc)
+library(DT)
+library(RColorBrewer)
 # Import data
-api <- read.table("~/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt",header = F)
-rcon <- redcapConnection(url = "https://redcap.ucdenver.edu/api/",
-                         token = api[1,1])
-data = exportRecords(rcon)
+api <- read.table("~/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt",
+  header = F
+)
+rcon <- redcapConnection(
+  url = "https://redcap.ucdenver.edu/api/",
+  token = api[1, 1]
+)
+data <- exportRecords(rcon, labels = F)
+data <- data.frame(lapply(data, as.character))
 # Make scorer IDs by initial
-data$scorer = paste0(toupper(substr(data$reader_first_name,1,1)),
-                     toupper(substr(data$reader_last_name,1,1)))
-data = data %>% arrange(scan_id,scorer) %>% filter(scan_id>2) %>% 
-  select(scorer,scan_id,bronchiectasis_rul:atelectasis___6)
+data$scorer <- paste0(
+  toupper(substr(data$reader_first_name, 1, 1)),
+  toupper(substr(data$reader_last_name, 1, 1))
+)
+# Select relevant columns, convert to long data
+data <- data %>%
+  arrange(scan_id, scorer) %>%
+  filter(scan_id > 2) %>%
+  select(scorer, scan_id, bronchiectasis_rul:atelectasis___6)
+data <- data %>%
+  pivot_longer(bronchiectasis_rul:atelectasis___6,
+    names_to = c("feature", "location"),
+    names_pattern = "(.*)_(.*)$"
+  ) %>%
+  mutate(
+    feature = gsub("__", "", feature),
+    feature = str_to_title(gsub("_", " ", feature))
+  )
+data$location <- factor(data$location,
+  levels = c("1", "2", "3", "4", "5", "6", "ling", "lll", "lul", "rll", "rml", "rul"),
+  labels = c("RUL", "RML", "RLL", "LUL", "Ling", "LLL", "Ling", "LLL", "LUL", "RLL", "RML", "RUL")
+)
+data$value <- factor(data$value, levels = c("Checked", "Unchecked", "Absent", "Cylindric", "Varicose"))
+vals <- levels(data$value)
+# One column per scorer
+data <- data %>% pivot_wider(names_from = scorer, values_from = value)
+# Find disagreements
+data$disagree <- !(data$AF == data$DL & data$DL == data$JW)
+# Set colors for whole document
+n <- length(vals)
+name <- "Accent"
+colors <- brewer.pal(n, name)
+names(colors) <- vals
+custom_colors <- scale_fill_manual(name = "", values = colors)
 ```
 
-# Scan 3
+# All disagreements
+
+The table below is sortable and searchable, and includes only the features and locations where there was some disagreement.
+
+```{r}
+# Table
+data %>%
+  filter(disagree == T) %>%
+  select(-disagree) %>%
+  datatable(., rownames = F)
+```
+# Plots
+
+## All scans
+
+```{r}
+data %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(., aes(x = name, y = location, fill = value)) +
+  geom_tile(color = "black") +
+  theme_bw() +
+  labs(fill = "", x = "Scorer", y = "Location") +
+  custom_colors +
+  facet_grid(rows = vars(scan_id), cols = vars(feature))
+```
+
+## Scan 3
+
+```{r}
+scan <- 3
+data %>%
+  filter(scan_id == scan) %>%
+  select(feature, location, AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(., aes(x = name, y = location, fill = value)) +
+  geom_tile(color = "black") +
+  theme_bw() +
+  labs(fill = "", x = "Scorer", y = "Location") +
+  custom_colors +
+  facet_wrap(~feature)
+```
+
+## Scan 4
+
+```{r}
+scan <- 4
+data %>%
+  filter(scan_id == scan) %>%
+  select(feature, location, AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(., aes(x = name, y = location, fill = value)) +
+  geom_tile(color = "black") +
+  theme_bw() +
+  labs(fill = "", x = "Scorer", y = "Location") +
+  custom_colors +
+  facet_wrap(~feature)
+```
+
+## Scan 5
+
+```{r}
+scan <- 5
+data %>%
+  filter(scan_id == scan) %>%
+  select(feature, location, AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(., aes(x = name, y = location, fill = value)) +
+  geom_tile(color = "black") +
+  theme_bw() +
+  labs(fill = "", x = "Scorer", y = "Location") +
+  custom_colors +
+  facet_wrap(~feature)
+```
+
+# Agreement by Feature
+
+## Atelectasis
+
+```{r}
+feat <- "Atelectasis"
+# Statistic
+k <- data %>%
+  filter(feature == feat) %>%
+  select(AF:JW) %>%
+  kappam.fleiss(.)
+ka <- round(k$value, 3)
+p <- format.pval(k$p.value, eps = 0.001, digits = 3)
+# Plot
+data %>%
+  filter(feature == feat) %>%
+  select(scan_id, location, AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(., aes(x = name, y = location, fill = value)) +
+  geom_tile(color = "black") +
+  theme_bw() +
+  labs(fill = "", x = "Scorer", y = "Location") +
+  custom_colors +
+  facet_wrap(~scan_id)
+```
+
+Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
 
@@
 ```{r}
-# Get relevant columns and prepare for Fleiss' kappa calculation
-bronch = data %>% filter(scan_id==3) %>% 
-  select(scorer,bronchiectasis_rul:bronchiectasis_lll)
-rows = as.character(label(bronch)[-1])
-cols = bronch$scorer
-bronch = data.frame(t(bronch[,-1]))
-rownames(bronch) = rows
-colnames(bronch) = cols
+feat <- "Bronchiectasis"
+# Statistic
+k <- data %>%
+  filter(feature == feat) %>%
+  select(AF:JW) %>%
+  kappam.fleiss(.)
+ka <- round(k$value, 3)
+p <- format.pval(k$p.value, eps = 0.001, digits = 3)
+# Plot
+data %>%
+  filter(feature == feat) %>%
+  select(scan_id, location, AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(., aes(x = name, y = location, fill = value)) +
+  geom_tile(color = "black") +
+  theme_bw() +
+  labs(fill = "", x = "Scorer", y = "Location") +
+  custom_colors +
+  facet_wrap(~scan_id)
 ```
+
+Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
+
+## Cavity
+
+```{r}
+feat <- "Cavity"
+# Statistic
+k <- data %>%
+  filter(feature == feat) %>%
+  select(AF:JW) %>%
+  kappam.fleiss(.)
+ka <- round(k$value, 3)
+p <- format.pval(k$p.value, eps = 0.001, digits = 3)
+# Plot
+data %>%
+  filter(feature == feat) %>%
+  select(scan_id, location, AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(., aes(x = name, y = location, fill = value)) +
+  geom_tile(color = "black") +
+  theme_bw() +
+  labs(fill = "", x = "Scorer", y = "Location") +
+  custom_colors +
+  facet_wrap(~scan_id)
+```
+
+Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
+
+## Consolidation
+
+```{r}
+feat <- "Consolidation"
+# Statistic
+k <- data %>%
+  filter(feature == feat) %>%
+  select(AF:JW) %>%
+  kappam.fleiss(.)
+ka <- round(k$value, 3)
+p <- format.pval(k$p.value, eps = 0.001, digits = 3)
+# Plot
+data %>%
+  filter(feature == feat) %>%
+  select(scan_id, location, AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(., aes(x = name, y = location, fill = value)) +
+  geom_tile(color = "black") +
+  theme_bw() +
+  labs(fill = "", x = "Scorer", y = "Location") +
+  custom_colors +
+  facet_wrap(~scan_id)
+```
+
+Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
+
+## Cyst
+
+```{r}
+feat <- "Cyst"
+# Statistic
+k <- data %>%
+  filter(feature == feat) %>%
+  select(AF:JW) %>%
+  kappam.fleiss(.)
+ka <- round(k$value, 3)
+p <- format.pval(k$p.value, eps = 0.001, digits = 3)
+# Plot
+data %>%
+  filter(feature == feat) %>%
+  select(scan_id, location, AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(., aes(x = name, y = location, fill = value)) +
+  geom_tile(color = "black") +
+  theme_bw() +
+  labs(fill = "", x = "Scorer", y = "Location") +
+  custom_colors +
+  facet_wrap(~scan_id)
+```
+
+Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
+
+## Ground Glass
+
+```{r}
+feat <- "Ground Glass"
+# Statistic
+k <- data %>%
+  filter(feature == feat) %>%
+  select(AF:JW) %>%
+  kappam.fleiss(.)
+ka <- round(k$value, 3)
+p <- format.pval(k$p.value, eps = 0.001, digits = 3)
+# Plot
+data %>%
+  filter(feature == feat) %>%
+  select(scan_id, location, AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(., aes(x = name, y = location, fill = value)) +
+  geom_tile(color = "black") +
+  theme_bw() +
+  labs(fill = "", x = "Scorer", y = "Location") +
+  custom_colors +
+  facet_wrap(~scan_id)
+```
+
+Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
+
+## Mucoid Impaction
+
+```{r}
+feat <- "Mucoid Impaction"
+# Statistic
+k <- data %>%
+  filter(feature == feat) %>%
+  select(AF:JW) %>%
+  kappam.fleiss(.)
+ka <- round(k$value, 3)
+p <- format.pval(k$p.value, eps = 0.001, digits = 3)
+# Plot
+data %>%
+  filter(feature == feat) %>%
+  select(scan_id, location, AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(., aes(x = name, y = location, fill = value)) +
+  geom_tile(color = "black") +
+  theme_bw() +
+  labs(fill = "", x = "Scorer", y = "Location") +
+  custom_colors +
+  facet_wrap(~scan_id)
+```
+
+Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
+
+## Nodule
+
+```{r}
+feat <- "Nodule"
+# Statistic
+k <- data %>%
+  filter(feature == feat) %>%
+  select(AF:JW) %>%
+  kappam.fleiss(.)
+ka <- round(k$value, 3)
+p <- format.pval(k$p.value, eps = 0.001, digits = 3)
+# Plot
+data %>%
+  filter(feature == feat) %>%
+  select(scan_id, location, AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(., aes(x = name, y = location, fill = value)) +
+  geom_tile(color = "black") +
+  theme_bw() +
+  labs(fill = "", x = "Scorer", y = "Location") +
+  custom_colors +
+  facet_wrap(~scan_id)
+```
+
+Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
+
+## Tree in Bud
+
+```{r}
+feat <- "Tree In Bud"
+# Statistics
+k <- data %>%
+  filter(feature == feat) %>%
+  select(AF:JW) %>%
+  kappam.fleiss(.)
+ka <- round(k$value, 3)
+p <- format.pval(k$p.value, eps = 0.001, digits = 3)
+# Plot
+data %>%
+  filter(feature == feat) %>%
+  select(scan_id, location, AF:JW) %>%
+  pivot_longer(AF:JW) %>%
+  ggplot(., aes(x = name, y = location, fill = value)) +
+  geom_tile(color = "black") +
+  theme_bw() +
+  labs(fill = "", x = "Scorer", y = "Location") +
+  custom_colors +
+  facet_wrap(~scan_id)
+```
+
+Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.

### 2022-12-01 StatNet/cgm_fda.qmd
Create cgm_fda.qmd

---
title: "Functional Analysis of CGM Data"
author: "Tim Vigers"
format: revealjs
editor: visual
---

## The Question

-   Do nighttime glucose patterns differ between controls (HC), CF with normal glucose tolerance (CFNGT), CF with abnormal glucose tolerance (CFAGT), and people with CF related diabetes (CFRD)?

```{r}
#| include: false
# Libraries
library(tidyverse)
library(lubridate)
library(cgmanalysis)
library(tidyfun)
# Data import
sleep = read.csv("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Raw/sleep.csv",
                 na.strings = "")
# Get bedtime and wake times
sleep = sleep %>% 
  mutate(Day.1.End = Day.2.Date,Day.2.End = Day.3.Date,Day.3.End = Day.4.Date,
         Day.4.End = Day.5.Date,Day.5.End = Day.6.Date,Day.6.End = Day.7.Date) %>%
  select(Subject.ID:Average.BMI.Z.score,
                         matches("Day\\.\\d\\.Date"),
         matches("Day\\.\\d\\.Date"),
                         matches("Day\\.\\d\\.Bed.Time"),
                         matches("Day\\.\\d\\.Get.Up.Time"))
sleep = sleep %>% 
  pivot_longer(Day.1.Date:Day.7.Get.Up.Time,
               names_to = c("Day",".value"),names_pattern = "Day\\.(.)\\.(.*)") %>%
  filter(!is.na(Bed.Time) & !is.na(Get.Up.Time)) %>%
  mutate(Night.Start = ymd_hm(paste(Date,Bed.Time)),
         Night.End = ymd_hm(paste(Date,Get.Up.Time))+days(1)) %>%
  select(Subject.ID:Average.BMI.Z.score,Day,Night.Start,Night.End)
sleep$Subject.ID = sub("--1","",sleep$Subject.ID)
# Add a day for those going to sleep between 12am - 6am
sleep$Night.Start[hour(sleep$Night.Start) %in% 0:6] = 
  sleep$Night.Start[hour(sleep$Night.Start) %in% 0:6] + days(1)
# CGMs
# cleandata("~/Dropbox/Work/CHCO/Christine Chan/Functional data analysis/Raw CGM Data",
#           "~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean",removegaps = F)
cgms = list.files("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean",
                  full.names = T)
names(cgms) = unlist(lapply(cgms, function(f){
  df = read.csv(f)
  id = strsplit(basename(f),"_")[[1]][1]
  return(id)
}))
# For each actiwatch, pull out the relevant vector of CGM values
sleep = sleep[sleep$Subject.ID %in% names(cgms),]
vecs = apply(sleep,1,function(r){
  # Pull info
  id = as.character(r["Subject.ID"])
  start = ymd_hms(as.character(r["Night.Start"]))
  end = ymd_hms(as.character(r["Night.End"]))
  # Find CGM file
  df = read.csv(cgms[id])
  df$timestamp = round_date(ymd_hms(df$timestamp),"5 minutes")
  # Get the time between sensors
  interval = table(round(diff(df$timestamp)/60))
  interval = names(interval)[which.max(interval)]
  # Put glucose values on time grid, pull glucose values
  times = round_date(seq(start,end,by = paste(interval,"min")),"5 minutes")
  glucs = data.frame(cbind(times,df$sensorglucose[match(times,df$timestamp)]))
  glucs = glucs[,2]
  # Grid sequence for functional data
  arg = seq(0,1,length.out = length(glucs))
  return(list(glucs,arg))
})
# Pull out glucose values and sequences separately
args = sapply(vecs,"[[",2)
glucs = sapply(vecs,"[[",1)
# Find those where all glucose values are missing
keep = which(sapply(glucs, function(g){!all(is.na(g))}))
# Make functional object
fda_df = tfd(glucs[keep],args[keep])
```

## Questions for the Group

1.  How do we evaluate everyone over a common grid? Do we want to keep one night as the "grid unit?"
2.  How to handle missing data? Is linear approximation reasonable (I would assume not)?
3.  How to account for within-subject correlation?

### 2022-12-01 StatNet/cgm_fda.qmd
Update cgm_fda.qmd

@@
   sleep$Night.Start[hour(sleep$Night.Start) %in% 0:6] + days(1)
+# Shorter group names
+sleep$Combined.Group.Category = as.factor(sleep$Combined.Group.Category)
+levels(sleep$Combined.Group.Category) = c("CFAGT","CFNGT","CFRD","HC")
 # CGMs
@@
 # Find those where all glucose values are missing
-keep = which(sapply(glucs, function(g){!all(is.na(g))}))
-# Make functional object
-fda_df = tfd(glucs[keep],args[keep])
+keep = sapply(glucs, function(g){!all(is.na(g))})
+# Make functional object, add to demographic info
+fda_df = sleep %>% select(Subject.ID:Average.BMI.Z.score)
+fda_df = fda_df[keep,]
+fda_df$Glucose = tfd(glucs[keep],args[keep])
+```
+
+## Our dataset
+
+### FILL THIS IN
+
+```{r}
+fda_df %>% 
+  ggplot() + 
+  geom_spaghetti(aes(tf = Glucose, col = Combined.Group.Category)) +
+  facet_wrap(~ Gender) + 
+  scale_alpha(guide = 'none', range = c(.2, .4)) +
+  scale_color_discrete(name="") + theme_bw()
 ```

### 2022-12-02 StatNet/cgm_fda.qmd
Update cgm_fda.qmd
all data read in

--- 2022-12-01 StatNet/cgm_fda.qmd
@@
   sleep$Night.Start[hour(sleep$Night.Start) %in% 0:6] + days(1)
+# Shorter group names
+sleep$Combined.Group.Category = 
+  factor(sleep$Combined.Group.Category,
+         levels = c("Healthy Control",
+                    "Cystic Fibrosis Normal Glucose Tolerant",
+                    "Cystic Fibrosis Abnormal Glycemia",
+                    "Cystic Fibrosis Related Diabetes"),
+         labels = c("HC","CF-NGT","CF-AGT","CFRD"),ordered = T)
+levels(sleep$Combined.Group.Category) = c("CFAGT","CFNGT","CFRD","HC")
 # CGMs
-# cleandata("~/Dropbox/Work/CHCO/Christine Chan/Functional data analysis/Raw CGM Data",
-#           "~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean",removegaps = F)
-cgms = list.files("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean",
+# cleandata("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Raw/CGM",
+#           "~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",
+#           removegaps = F)
+cgms = list.files("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",
                   full.names = T)
@@
   df = read.csv(f)
-  id = strsplit(basename(f),"_")[[1]][1]
+  id = strsplit(basename(f)," ")[[1]][1]
   return(id)
@@
 # Find those where all glucose values are missing
-keep = which(sapply(glucs, function(g){!all(is.na(g))}))
-# Make functional object
-fda_df = tfd(glucs[keep],args[keep])
+keep = sapply(glucs, function(g){!all(is.na(g)) & !is.na(var(g,na.rm = T))})
+# Make functional object, add to demographic info
+fda_df = sleep %>% select(Subject.ID:Average.BMI.Z.score)
+fda_df = fda_df[keep,]
+fda_df$Glucose = tfd(glucs[keep],args[keep])
+```
+
+## Our dataset
+
+### FILL THIS IN
+
+```{r}
+fda_df %>% 
+  ggplot() + 
+  geom_spaghetti(aes(tf = Glucose, col = Combined.Group.Category)) +
+  facet_wrap(~ Gender) + 
+  scale_alpha(guide = 'none', range = c(.2, .4)) +
+  scale_color_discrete(name="") + theme_bw()
 ```

### 2022-12-02 StatNet/cgm_fda.qmd
better t1 formatting

--- 2022-12-01 StatNet/cgm_fda.qmd
@@
 author: "Tim Vigers"
-format: revealjs
+format: 
+  revealjs:
+    scrollable: true
 editor: visual
@@
 # Libraries
+library(arsenal)
 library(tidyverse)
@@
 library(tidyfun)
+library(refundr)
 # Data import
@@
                names_to = c("Day",".value"),names_pattern = "Day\\.(.)\\.(.*)") %>%
-  filter(!is.na(Bed.Time) & !is.na(Get.Up.Time)) %>%
+  filter(!is.na(Bed.Time) & !is.na(Get.Up.Time) & Day < 7) %>%
   mutate(Night.Start = ymd_hm(paste(Date,Bed.Time)),
@@
   sleep$Night.Start[hour(sleep$Night.Start) %in% 0:6] + days(1)
+# Shorter group names
+sleep$Combined.Group.Category = 
+  factor(sleep$Combined.Group.Category,
+         levels = c("Healthy Control",
+                    "Cystic Fibrosis Normal Glucose Tolerant",
+                    "Cystic Fibrosis Abnormal Glycemia",
+                    "Cystic Fibrosis Related Diabetes"),
+         labels = c("HC","CF-NGT","CF-AGT","CFRD"),ordered = T)
+# HC not on insulin
+sleep$CF.Insulin.Use[sleep$Combined.Group.Category == "HC"]="No"
 # CGMs
-# cleandata("~/Dropbox/Work/CHCO/Christine Chan/Functional data analysis/Raw CGM Data",
-#           "~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean",removegaps = F)
-cgms = list.files("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean",
+# cleandata("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Raw/CGM",
+#           "~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",
+#           removegaps = F)
+cgms = list.files("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",
                   full.names = T)
@@
 # Find those where all glucose values are missing
-keep = which(sapply(glucs, function(g){!all(is.na(g))}))
-# Make functional object
-fda_df = tfd(glucs[keep],args[keep])
+keep = sapply(glucs, function(g){!all(is.na(g)) & !is.na(var(g,na.rm = T))})
+# Make functional object, add to demographic info
+fda_df = sleep %>% select(Subject.ID:Average.BMI.Z.score)
+fda_df = fda_df[keep,]
+fda_df$Glucose = tfd(glucs[keep],args[keep],interpolate=T)
+```
+
+## Our dataset {.smaller}
+
+```{r results='asis'}
+demo = fda_df %>%
+  select(Subject.ID:Average.BMI.Z.score) %>% 
+  group_by(Subject.ID) %>% 
+  summarise(Age = Age[1],Gender = Gender[1],Race = Race.Ethnicity[1],
+            Insulin = CF.Insulin.Use[1],`BMI Z Score`= Average.BMI.Z.score[1],
+            `Days of Data` = n(),Group = Combined.Group.Category[1])
+
+t1 = 
+  tableby(Group~Age+Gender+Race+Insulin+`BMI Z Score`+`Days of Data`,
+             data = demo)
+summary(t1)
+```
+
+## Plot {.smaller}
+
+```{r}
+fda_df %>% 
+  ggplot() + 
+  geom_spaghetti(aes(tf = Glucose),alpha=0.2) +
+  scale_color_discrete(name="") + theme_bw() +
+  facet_wrap(~Combined.Group.Category)
+```
+
+```{r cache=TRUE}
+fpca = rfr_fpca(Y = "Glucose", data = fda_df)
+# dti_df %>% 
+#   modelr::add_predictions(fpca, var = "glucose_fits") %>% 
+#   ggplot(aes(y = glucose_fits)) + 
+#   geom_spaghetti() + 
+#   geom_spaghetti(aes(y = Glucose), color = "blue", alpha = .2)
 ```

### 2022-12-02 StatNet/cgm_fda.qmd
a bunch of different methods

--- 2022-12-01 StatNet/cgm_fda.qmd
@@
 library(cgmanalysis)
+library(refund)
+library(refund.shiny)
 library(tidyfun)
+library(refundr)
 # Data import
@@
 # Shorter group names
-sleep$Combined.Group.Category = as.factor(sleep$Combined.Group.Category)
-levels(sleep$Combined.Group.Category) = c("CFAGT","CFNGT","CFRD","HC")
+sleep$Combined.Group.Category = 
+  factor(sleep$Combined.Group.Category,
+         levels = c("Healthy Control",
+                    "Cystic Fibrosis Normal Glucose Tolerant",
+                    "Cystic Fibrosis Abnormal Glycemia",
+                    "Cystic Fibrosis Related Diabetes"),
+         labels = c("HC","CF-NGT","CF-AGT","CFRD"),ordered = T)
+# HC not on insulin
+sleep$CF.Insulin.Use[sleep$Combined.Group.Category == "HC"]="No"
 # CGMs
-# cleandata("~/Dropbox/Work/CHCO/Christine Chan/Functional data analysis/Raw CGM Data",
-#           "~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean",removegaps = F)
-cgms = list.files("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean",
+# cleandata("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Raw/CGM",
+#           "~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",
+#           removegaps = F)
+cgms = list.files("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",
                   full.names = T)
@@
 # Pull out glucose values and sequences separately
+glucs = sapply(vecs,"[[",1)
 args = sapply(vecs,"[[",2)
-glucs = sapply(vecs,"[[",1)
 # Find those where all glucose values are missing
-keep = sapply(glucs, function(g){!all(is.na(g))})
+keep = sapply(glucs, function(g){!all(is.na(g)) & !is.na(var(g,na.rm = T))})
 # Make functional object, add to demographic info
@@
 fda_df = fda_df[keep,]
-fda_df$Glucose = tfd(glucs[keep],args[keep])
+fda_df$Glucose = tfd(glucs[keep],args[keep],interpolate=T)
 ```
 
-## Our dataset
+## Our dataset {.smaller}
 
-### FILL THIS IN
+```{r results='asis'}
+demo = fda_df %>%
+  select(Subject.ID:Average.BMI.Z.score) %>% 
+  group_by(Subject.ID) %>% 
+  summarise(Age = Age[1],Gender = Gender[1],Race = Race.Ethnicity[1],
+            Insulin = CF.Insulin.Use[1],`BMI Z Score`= Average.BMI.Z.score[1],
+            `Days of Data` = n(),Group = Combined.Group.Category[1])
+t1 = 
+  tableby(Group~Age+Gender+Race+Insulin+`BMI Z Score`+`Days of Data`,
+             data = demo)
+summary(t1)
+```
+
+## Plot
 
@@
   ggplot() + 
-  geom_spaghetti(aes(tf = Glucose, col = Combined.Group.Category)) +
-  facet_wrap(~ Gender) + 
-  scale_alpha(guide = 'none', range = c(.2, .4)) +
-  scale_color_discrete(name="") + theme_bw()
+  geom_spaghetti(aes(tf = Glucose),alpha=0.2) +
+  scale_color_discrete(name="") + theme_bw() +
+  facet_wrap(~Combined.Group.Category)
+```
+
+```{r cache=TRUE}
+fpca = rfr_fpca(Y = "Glucose", data = fda_df)
+fda_df %>%
+  modelr::add_predictions(fpca, var = "glucose_fits") %>%
+  ggplot(aes(y = glucose_fits)) +
+  geom_spaghetti() +
+  geom_spaghetti(aes(y = Glucose), color = "blue", alpha = .2)
+```
+
+```{r}
+ys = apply(sleep,1,function(r){
+  # Pull info
+  id = as.character(r["Subject.ID"])
+  start = ymd_hms(as.character(r["Night.Start"]))
+  end = ymd_hms(as.character(r["Night.End"]))
+  # Find CGM file
+  df = read.csv(cgms[id])
+  id = paste(as.character(r["Subject.ID"]),as.character(r["Day"]),sep = "_")
+  df$timestamp = round_date(ymd_hms(df$timestamp),"5 minutes")
+  # Get the time between sensors
+  interval = table(round(diff(df$timestamp)/60))
+  interval = names(interval)[which.max(interval)]
+  # Put glucose values on time grid
+  times = round_date(seq(start,end,by = paste(interval,"min")),"5 minutes")
+  glucs = df$sensorglucose[match(times,df$timestamp)]
+  y = data.frame(".obs"=rep(id,length(glucs)),
+                 ".index"=seq(0,1,length.out=length(glucs)),
+                 ".value" = glucs)
+  return(y)
+})
+# Find those where all glucose values are missing
+# keep = sapply(ys, function(y){!all(is.na(y$.value)) & !is.na(var(y$.value,na.rm = T))})
+ydata = do.call(rbind,ys)
+colnames(ydata)[1] = ".id"
+fit = fpca.sc(ydata=na.omit(ydata[ydata$.id %in% names(table(ydata$.id)[1:10]),]),
+              var = TRUE, simul = FALSE)
+```
+
+```{r}
+library(fdapace)
+library(fastDummies)
+vecs = apply(sleep,1,function(r){
+  # Pull info
+  id = as.character(r["Subject.ID"])
+  start = ymd_hms(as.character(r["Night.Start"]))
+  end = ymd_hms(as.character(r["Night.End"]))
+  # Find CGM file
+  df = read.csv(cgms[id])
+  df$timestamp = round_date(ymd_hms(df$timestamp),"5 minutes")
+  # Get the time between sensors
+  interval = table(round(diff(df$timestamp)/60))
+  interval = names(interval)[which.max(interval)]
+  # Put glucose values on time grid, pull glucose values
+  times = round_date(seq(start,end,by = paste(interval,"min")),"5 minutes")
+  glucs = data.frame(cbind(times,df$sensorglucose[match(times,df$timestamp)]))
+  glucs = glucs[,2]
+  # Grid sequence for functional data
+  arg = seq(0,length(glucs))
+  return(list(glucs,arg))
+})
+ylist = sapply(vecs,"[[",1)
+tlist = sapply(vecs,"[[",2)
+keep = sapply(ylist, function(g){!all(is.na(g)) & !is.na(var(g,na.rm = T))})
+X = sleep %>% select(Subject.ID:Average.BMI.Z.score)
+X = X[keep,]
+X = dummy_cols(X[,"Combined.Group.Category"],remove_selected_columns=T,remove_first_dummy=T)
+FPCAobj <- FPCA(Ly=ylist[keep], Lt=tlist[keep],list(plot = TRUE))
+VCAM(Lt=tlist[keep], Ly=ylist[keep],X)
 ```

### 2022-12-02 StatNet/cgm_fda.qmd
focusing on Fourier basis

@@
 library(refund.shiny)
-library(tidyfun)
-library(refundr)
 # Data import
@@
 
+## Basis Function
+
+In general, the Fourier basis is the usual choice for periodic functions.
+
+\cite{ramsayFunctionalDataAnalysis2009}
+
+```{r}
+basisobj = create.fourier.basis(rangeval=c(0, 1))
+plot(basisobj)
+```
+
 ```{r cache=TRUE}
@@
 FPCAobj <- FPCA(Ly=ylist[keep], Lt=tlist[keep],list(plot = TRUE))
-VCAM(Lt=tlist[keep], Ly=ylist[keep],X)
+VCAM(Lt=tlist[1:10], Ly=ylist[1:10],X[1:10])
 ```
@@
 
-1.  How do we evaluate everyone over a common grid? Do we want to keep one night as the "grid unit?"
-2.  How to handle missing data? Is linear approximation reasonable (I would assume not)?
-3.  How to account for within-subject correlation?
+-   What is the best way to pick the order of basis functions, etc.?
+-   Currently I'm forcing the glucose measures to be on the interval \[0,1\], which means that the period for the Fourier basis is "nighttime." Since those with shorter sleep periods will have fewer observations, I think this means that

### 2022-12-02 StatNet/cgm_fda.qmd
Update cgm_fda.qmd

@@
 
+```{r}
+ys = apply(sleep,1,function(r){
+  # Pull info
+  id = as.character(r["Subject.ID"])
+  start = ymd_hms(as.character(r["Night.Start"]))
+  end = ymd_hms(as.character(r["Night.End"]))
+  # Find CGM file
+  df = read.csv(cgms[id])
+  id = paste(as.character(r["Subject.ID"]),as.character(r["Day"]),sep = "_")
+  df$timestamp = round_date(ymd_hms(df$timestamp),"5 minutes")
+  # Get the time between sensors
+  interval = table(round(diff(df$timestamp)/60))
+  interval = names(interval)[which.max(interval)]
+  # Put glucose values on time grid
+  times = round_date(seq(start,end,by = paste(interval,"min")),"5 minutes")
+  glucs = df$sensorglucose[match(times,df$timestamp)]
+  y = data.frame(".obs"=rep(id,length(glucs)),
+                 ".index"=seq(0,1,length.out=length(glucs)),
+                 ".value" = glucs)
+  return(y)
+})
+# Find those where all glucose values are missing
+# keep = sapply(ys, function(y){!all(is.na(y$.value)) & !is.na(var(y$.value,na.rm = T))})
+ydf = do.call(rbind,ys)
+colnames(ydf)[1] = ".id"
+```
+
+```{r}
+fit = fpca.sc(ydata = ydf[ydf$.id %in% names(table(ydf$.id))[1:5],],pve = 0.95)
+```
+
 ## Our dataset {.smaller}
@@
 ```{r}
+# Create a functional data object
+
 basisobj = create.fourier.basis(rangeval=c(0, 1))
@@
 
-```{r cache=TRUE}
-fpca = rfr_fpca(Y = "Glucose", data = fda_df)
-fda_df %>%
-  modelr::add_predictions(fpca, var = "glucose_fits") %>%
-  ggplot(aes(y = glucose_fits)) +
-  geom_spaghetti() +
-  geom_spaghetti(aes(y = Glucose), color = "blue", alpha = .2)
-```
-
-```{r}
-ys = apply(sleep,1,function(r){
-  # Pull info
-  id = as.character(r["Subject.ID"])
-  start = ymd_hms(as.character(r["Night.Start"]))
-  end = ymd_hms(as.character(r["Night.End"]))
-  # Find CGM file
-  df = read.csv(cgms[id])
-  id = paste(as.character(r["Subject.ID"]),as.character(r["Day"]),sep = "_")
-  df$timestamp = round_date(ymd_hms(df$timestamp),"5 minutes")
-  # Get the time between sensors
-  interval = table(round(diff(df$timestamp)/60))
-  interval = names(interval)[which.max(interval)]
-  # Put glucose values on time grid
-  times = round_date(seq(start,end,by = paste(interval,"min")),"5 minutes")
-  glucs = df$sensorglucose[match(times,df$timestamp)]
-  y = data.frame(".obs"=rep(id,length(glucs)),
-                 ".index"=seq(0,1,length.out=length(glucs)),
-                 ".value" = glucs)
-  return(y)
-})
-# Find those where all glucose values are missing
-# keep = sapply(ys, function(y){!all(is.na(y$.value)) & !is.na(var(y$.value,na.rm = T))})
-ydata = do.call(rbind,ys)
-colnames(ydata)[1] = ".id"
-fit = fpca.sc(ydata=na.omit(ydata[ydata$.id %in% names(table(ydata$.id)[1:10]),]),
-              var = TRUE, simul = FALSE)
-```
-
-```{r}
-library(fdapace)
-library(fastDummies)
-vecs = apply(sleep,1,function(r){
-  # Pull info
-  id = as.character(r["Subject.ID"])
-  start = ymd_hms(as.character(r["Night.Start"]))
-  end = ymd_hms(as.character(r["Night.End"]))
-  # Find CGM file
-  df = read.csv(cgms[id])
-  df$timestamp = round_date(ymd_hms(df$timestamp),"5 minutes")
-  # Get the time between sensors
-  interval = table(round(diff(df$timestamp)/60))
-  interval = names(interval)[which.max(interval)]
-  # Put glucose values on time grid, pull glucose values
-  times = round_date(seq(start,end,by = paste(interval,"min")),"5 minutes")
-  glucs = data.frame(cbind(times,df$sensorglucose[match(times,df$timestamp)]))
-  glucs = glucs[,2]
-  # Grid sequence for functional data
-  arg = seq(0,length(glucs))
-  return(list(glucs,arg))
-})
-ylist = sapply(vecs,"[[",1)
-tlist = sapply(vecs,"[[",2)
-keep = sapply(ylist, function(g){!all(is.na(g)) & !is.na(var(g,na.rm = T))})
-X = sleep %>% select(Subject.ID:Average.BMI.Z.score)
-X = X[keep,]
-X = dummy_cols(X[,"Combined.Group.Category"],remove_selected_columns=T,remove_first_dummy=T)
-FPCAobj <- FPCA(Ly=ylist[keep], Lt=tlist[keep],list(plot = TRUE))
-VCAM(Lt=tlist[1:10], Ly=ylist[1:10],X[1:10])
-```
-
 ## Questions for the Group
 
--   What is the best way to pick the order of basis functions, etc.?
+-   What is the best way to pick the order of basis functions, etc.? Visually?
 -   Currently I'm forcing the glucose measures to be on the interval \[0,1\], which means that the period for the Fourier basis is "nighttime." Since those with shorter sleep periods will have fewer observations, I think this means that

### 2022-12-02 StatNet/cgm_fda.qmd
Update cgm_fda.qmd

@@
 library(refund.shiny)
+library(tidyfun)
 # Data import
-sleep = read.csv("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Raw/sleep.csv",
+sleep = read.csv("C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Raw/sleep.csv",
                  na.strings = "")
@@
 # CGMs
-# cleandata("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Raw/CGM",
-#           "~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",
+# cleandata("C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Raw/CGM",
+#           "C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",
 #           removegaps = F)
-cgms = list.files("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",
+cgms = list.files("C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",
                   full.names = T)
@@
 # Find those where all glucose values are missing
-# keep = sapply(ys, function(y){!all(is.na(y$.value)) & !is.na(var(y$.value,na.rm = T))})
-ydf = do.call(rbind,ys)
+ydf = do.call(rbind,ys[keep])
 colnames(ydf)[1] = ".id"
@@
 
-```{r}
-fit = fpca.sc(ydata = ydf[ydf$.id %in% names(table(ydf$.id))[1:5],],pve = 0.95)
+```{r eval=FALSE}
+# FPCA on irregularly sampled data
+fit = fpca.sc(ydata = na.omit(ydf))
+# Save object
+save(fit,file = "C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/fpca_sc_fit.Rdata")
 ```

### 2022-12-02 StatNet/cgm_fda.qmd
Update cgm_fda.qmd

--- 2022-12-02 StatNet/cgm_fda.qmd
@@
 fda_df$Glucose = tfd(glucs[keep],args[keep],interpolate=T)
-```
-
-```{r}
+# Prepare for the refund package (tidyfun still pretty limited)
 ys = apply(sleep,1,function(r){
@@
 # Find those where all glucose values are missing
-# keep = sapply(ys, function(y){!all(is.na(y$.value)) & !is.na(var(y$.value,na.rm = T))})
-ydf = do.call(rbind,ys)
+ydf = do.call(rbind,ys[keep])
 colnames(ydf)[1] = ".id"
+# Save
+save(ydf,file = "C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/ydf.Rdata")
 ```
 
-```{r}
-fit = fpca.sc(ydata = ydf[ydf$.id %in% names(table(ydf$.id))[1:5],],pve = 0.95)
-```
-
-## Our dataset {.smaller}
+## Our Dataset {.smaller}
 
@@
 
-## Plot
+## Spaghetti Plot
 
@@
 
-## Basis Function
-
-In general, the Fourier basis is the usual choice for periodic functions.
-
-\cite{ramsayFunctionalDataAnalysis2009}
+## FPCA
 
 ```{r}
-# Create a functional data object
+#| eval: false
+#| echo: true
+# FPCA on irregularly sampled data
+fit = fpca.sc(ydata = na.omit(ydf))
+```
 
-basisobj = create.fourier.basis(rangeval=c(0, 1))
-plot(basisobj)
+```{r}
+#| include: false
+# load("C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/fpca_sc_fit.Rdata")
 ```

### 2022-12-02 StatNet/cgm_fda.qmd
Update cgm_fda.qmd

@@
     scrollable: true
+    smaller: true
 editor: visual
@@
 # Data import
-sleep = read.csv("C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Raw/sleep.csv",
+sleep = read.csv("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Raw/sleep.csv",
                  na.strings = "")
@@
 # CGMs
-# cleandata("C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Raw/CGM",
-#           "C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",
+# cleandata("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Raw/CGM",
+#           "~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",
 #           removegaps = F)
-cgms = list.files("C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",
+cgms = list.files("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",
                   full.names = T)
@@
 # Save
-save(ydf,file = "C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/ydf.Rdata")
+save(ydf,file = "~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/ydf.Rdata")
 ```
 
-## Our Dataset {.smaller}
+## What Is Functional Data Analysis (FDA)?
+
+- Most statistics use some sort of parametric model $p(x|\boldsymbol{\theta})$.
+  - We try to estimate the parameters of *p*.
+- Assume the data are normally distributed $N(\mu,\sigma^2)$.
+  - Can use the sample mean $\bar{x}$ as an estimate of the location parameter $\mu$.
+- Sometimes we do not want to assume a density function in advance.
+  - Interested in estimating $p$ itself.
+    - What is the smooth function that generates our data?
+
+## Goals and Uses
+
+- The goals of FDA are essentially the same as other areas of statistics:
+  1. Represent data to aid in analysis and to highlight characteristics
+  2. Study patterns and sources of variation
+  3. Explain variation in an outcome using independent variables
+
+- Common uses of FDA:
+  - Growth data
+  - Imaging data 
+  - Wearable device data
+  - Accelerometry (physical activity) 
+  - Heart rate
+  - **Continuous glucose monitoring** (CGM)
+
+## Semi- and Non-Parametric Regression
+
+- The basic linear model is: $y_i= \beta_0+\beta_1 x_i+\epsilon_i$ for $1\leq i \leq n$
+- Each predictor has a simple form
+  - The overall shape is determined by the model rather than the data.
+- The non-parametric version of this model is: $y_i = f(x_i)+\epsilon_i$
+
+## Semi- and Non-Parametric Regression
+
+- $f(x_i)$ is an arbitrary function.
+  - The only constraint is that $f(x_i)$ must be a smooth function.
+  - This model is much more flexible than parametric.
+- A semi-parametric model essentially uses parametric models for some predictors, and non-parametric models for others.
+
+## Basis Functions
+
+- How do we represent $f(x)$?
+- If $b_j(x)$ is the $j^{th}$ basis function, then: $$f(x)=\sum_{j=1}^k \beta_j b_j(x)$$
+- Polynomials are a simple form of basis function: $$f(x)=\beta_1+\beta_2x+\beta_3x^2+\beta_4x^3+\beta_5x^4$$ and $$y_i=\beta_1+\beta_2x+\beta_3x^2+\beta_4x^3+\beta_5x^4+\epsilon_i$$
+
+## Our Dataset
 
@@
 #| include: false
-# load("C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/fpca_sc_fit.Rdata")
+# load("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/fpca_sc_fit.Rdata")
 ```
@@
 
--   What is the best way to pick the order of basis functions, etc.? Visually?
--   Currently I'm forcing the glucose measures to be on the interval \[0,1\], which means that the period for the Fourier basis is "nighttime." Since those with shorter sleep periods will have fewer observations, I think this means that
+- I'm having some trouble running

### 2022-12-02 StatNet/cgm_fda.qmd
FPCA on first day only

@@
 })
-# Find those where all glucose values are missing
-ydf = do.call(rbind,ys[keep])
+sleep = sleep %>% unite("id_day",Subject.ID,Day,remove = F)
+names(ys) = sleep$id_day
+# Limit to first day with no missing
+sleep = sleep[keep,]
+sleep = sleep %>%
+  group_by(Subject.ID) %>%
+  filter(row_number()==1)
+ys = ys[sleep$id_day]
+# Combine
+ydf = do.call(rbind,ys)
 colnames(ydf)[1] = ".id"
@@
     3.  Explain variation in an outcome using independent variables
-
 -   Common uses of FDA:
@@
 ```{r results='asis'}
-demo = fda_df %>%
+sleep = sleep %>%
   select(Subject.ID:Average.BMI.Z.score) %>% 
-  group_by(Subject.ID) %>% 
-  summarise(Age = Age[1],Gender = Gender[1],Race = Race.Ethnicity[1],
-            Insulin = CF.Insulin.Use[1],`BMI Z Score`= Average.BMI.Z.score[1],
-            `Days of Data` = n(),Group = Combined.Group.Category[1])
-t1 = 
-  tableby(Group~Age+Gender+Race+Insulin+`BMI Z Score`+`Days of Data`,
-             data = demo)
+  rename(Age = Age,Gender = Gender,Race = Race.Ethnicity,
+            Insulin = CF.Insulin.Use,`BMI Z Score`= Average.BMI.Z.score,
+            Group = Combined.Group.Category)
+t1 = tableby(Group~Age+Gender+Race+Insulin+`BMI Z Score`,data = sleep)
 summary(t1)
@@
 ## Spaghetti Plot
+
+This plot includes all days of data, but the following analyses is only the first day due to computing limitations.
 

### 2022-12-02 Christine Chan/CGM and FDA/fpca.R

@@ -1,6 +1,6 @@
 library(refund)
 # FPCA on irregularly sampled data
-load("C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/ydf.Rdata")
+load("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/ydf.Rdata")
 fit = fpca.sc(ydata = na.omit(ydf))
 # Save object
-save(fit,file = "C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/fpca_sc_fit.Rdata")
+save(fit,file = "~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/fpca_sc_fit.Rdata")

### 2022-12-02 StatNet/cgm_fda.qmd
Update cgm_fda.qmd

@@
 ---
-
-## The Question
-
--   Do nighttime glucose patterns differ between controls (HC), CF with normal glucose tolerance (CFNGT), CF with abnormal glucose tolerance (CFAGT), and people with CF related diabetes (CFRD)?
 
@@
 names(ys) = sleep$id_day
-# Limit to first day with no missing
+# Limit to first day with enough data (50 obs. for now, but that's ad-hoc)
+keep = sapply(glucs, function(g){sum(!is.na(g))>=50})
 sleep = sleep[keep,]
@@
 
+## FPCA
+
+-   The FPCA model is of the form:
+
+$$Y_i(t)=\mu(t)+\sum_{k=1}^{K}c_{ik}\psi_k(t)+\epsilon_{i}(t) $$
+
+-   $\mu(t)$ is the population mean
+-   $\psi_k(t)$ are a set of population-level basis functions
+-   $c_{ik}$ are subject-specific scores with mean zero and variance $\lambda_k$
+-   $\epsilon_i(t)$ are residual curves
+-   There is an eigenfunction associated with each eigenvalue, rather than an eigenvector.
+-   These eigenfunctions describe major variational components, but in function form
+
+## Function-on-Scalar Regression
+
+-   The most common FoSR model is:
+
+$$Y_i(t)= \beta_0(t) + \sum_{k=1}^{p}x_{ik}\beta_k(t) + \epsilon_i(t)$$
+
+-   $\beta_k(t)$ are fixed effects associated with scalar covariates and the $\epsilon_i(t)$ are residual curves.
+-   $\beta_k(t)$ are interpreted analogously to coefficients in a multiple linear regression
+
+## The Question
+
+-   Do nighttime glucose patterns differ between controls (HC), CF with normal glucose tolerance (CFNGT), CF with abnormal glucose tolerance (CFAGT), and people with CF related diabetes (CFRD)?
+
 ## Our Dataset
@@
 
-## FPCA
+## FPCA on This Dataset
+
+```{r}
+#| include: false
+load("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/fpca_sc_fit.Rdata")
+```
+
+```{r}
+knitr::kable(head(ydf))
+```
 
@@
 
+## Function-on-Scalar Regression for Our Data
+
 ```{r}
 #| include: false
-# load("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/fpca_sc_fit.Rdata")
+ydf_wide = 
+  ydf %>% group_by(.id) %>% 
+  filter(row_number() <= 54) %>%
+  mutate(.index = row_number()) %>%
+  pivot_wider(id_cols = .id,names_from = .index,values_from = .value,names_prefix = "glucose_")
+ids = ydf_wide$.id
+ydf_wide$.id = NULL
+ydf_wide = data.matrix(ydf_wide)
+rownames(ydf_wide) = ids
+sleep$Glucose = ydf_wide
+sleep$Group = factor(sleep$Group,ordered = F)
+sleep = na.omit(sleep)
+```
+
+```{r}
+#| echo: true
+fosr_mod = bayes_fosr(Glucose ~ Group,data = sleep) 
+```
+
+```{r}
+data(DTI)
+DTI = subset(DTI, select = c(cca, case, pasat))
+DTI = DTI[complete.cases(DTI),]
+DTI$gender = factor(sample(c("male","female"), dim(DTI)[1], replace = TRUE))
+DTI$status = factor(sample(c("RRMS", "SPMS", "PPMS"), dim(DTI)[1], replace = TRUE))
+## fit models
+dti_fosr_mod = bayes_fosr(cca ~ gender, data = DTI)
 ```
@@
 
--   I'm having some trouble running
+-   Does this even seem like the correct approach?
+-   Really what we want to do is classify people based on CGM tracing. Is there a good package for functional discriminant analysis?
+-   FoSR in `refund` requires complete data, so how do we deal with irregular sampling across CGM tracings
+    -   There is between and within-subject variability in terms length of sleep and therefore number of glucose measures
+-   What does it mean that the FPCA eigenfunctions are essentially flat?
+-   I'm having some trouble running the full dataset, even on a relatively beefed up desktop. Any recommendations?
+-   Recommendations for easier to use packages?
+    -   Data formatting is a nightmare with the various FDA packages, and unfortunately `tidyfun` is pretty limited.

### 2022-12-11 Christine Chan/CF Registry/cf_registry_analysis.Rmd
pulmonary function

@@
 library(tidyverse)
-library(performance)
 library(knitr)
@@
 knitr::opts_chunk$set(echo = FALSE)
-knitr::opts_knit$set(root.dir = "~/Dropbox/Work/CHCO/Christine Chan/CF registry data")
+knitr::opts_knit$set(root.dir = "~/Work/CHCO/Christine Chan/CF registry data")
 ```
@@
 ), ordered = T)
-df$second_smoke <- factor(df$smoking, levels = 1:4, labels = c(
+df$second_smoke <- factor(df$second_smoke, levels = c(1:4), labels = c(
   "No",
@@
 # Save formatted data for re-import later
-save(df,file = "formatted_data.Rdata")
+save(df,file = "Data_Clean/formatted_data.Rdata")
 ```
@@
 ```{r}
-load("formatted_data.Rdata")
+load("Data_Clean/formatted_data.Rdata")
 ```
@@
 ```
+
+# Table 2: Pulmonary function
+
+```{r warning=FALSE,results='asis'}
+# Summarize by ID and period
+t2_df <- df %>%
+  group_by(eDWID, Period) %>%
+  summarise(
+    Smoking = max(smoking,na.rm = T),
+    `Secondhand smoke`=max(second_smoke,na.rm = T),
+    .groups = "keep"
+  )
+# Replace infinite values with missing
+t1_df <- data.frame(lapply(t1_df, function(c) {
+  c[is.infinite(c)] <- NA
+  return(c)
+}))
+# Print table 1
+cols <- colnames(t1_df)[3:ncol(t1_df)]
+t1 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
+  data = t1_df
+)
+summary(t1)
+```

### 2022-12-11 Christine Chan/CF Registry/cf_registry_analysis.Rmd
registry mostly done

--- 2022-11-02 Christine Chan/CF Registry/cf_registry_analysis.Rmd
@@
 knitr::opts_chunk$set(echo = FALSE)
-knitr::opts_knit$set(root.dir = "~/Dropbox/Work/CHCO/Christine Chan/CF registry data")
+knitr::opts_knit$set(root.dir = "/home/tim/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data")
 ```
@@
 # Combine education checkboxes
-cl <- makeCluster(detectCores() * 0.5, type = "FORK")
-df$employment <- parApply(cl, df, 1, function(r) {
+df$employment <- apply(df, 1, function(r) {
   emp <- r[paste0("employment", 1:7)]
@@
 })
-stopCluster(cl)
 df$employment <- sub("1", "Part Time", df$employment)
@@
 df$employment <- as.factor(df$employment)
+# Bronchodilator use
+df$bronchodilators <- apply(df, 1, function(r) {
+  bronch <- r[grep("bronchodilators",names(r))]
+  names(bronch) = c("Beta agonist",
+                    "Theophylline product",
+                    "Short acting beta agonist",
+                    "Long acting beta agonist",
+                    "Short acting anticholinergic",
+                    "Long acting anticholinergic",
+                    "Combination beta agonist and anticholinergic"
+  )
+  if (all(is.na(bronch))) {
+    return(NA)
+  } else {
+    bronch <- names(bronch)[which(bronch == 1)]
+    bronch <- paste(bronch, collapse = ", ")
+    return(bronch)
+  }
+})
+# Exacerbation medication
+df$act_primary <- apply(df, 1, function(r) {
+  act <- r[grep("A_act_primary",names(r))]
+  names(act) = c("Positive Expiratory Pressure (PEP)",
+                 "Postural drainage with clapping (CPT)",
+                 "Forced expiratory techniques",
+                 "Oscillating PEP",
+                 "High frequency chest wall oscillation",
+                 "Exercise",
+                 "None",
+                 "Other"
+  )
+  if (all(is.na(act) | act == 0)) {
+    return(NA)
+  } else {
+    act <- names(act)[which(act == 1)]
+    act <- paste(act, collapse = ", ")
+    return(act)
+  }
+})
+# Oxygen therapy
+df$oxygen_therapy <- factor(df$oxygen_therapy,levels = 1:5,
+                            labels = c("Yes, Continuously",
+                                       "Yes, Nocturnal and/or with exertion",
+                                       "Yes, During exacerbation",
+                                       "Yes, prn",
+                                       "No"))
 # Save formatted data for re-import later
-save(df,file = "formatted_data.Rdata")
+save(df,file = "Data_Clean/formatted_data.Rdata")
 ```
@@
 ```{r}
-load("formatted_data.Rdata")
+load("Data_Clean/formatted_data.Rdata")
+df = df %>% arrange(eDWID,reviewyear,encounternum)
 ```
@@
 ```
+
+# Table 2: Pulmonary function
+
+Continuous and count variables are from the last year of each period.
+
+```{r warning=FALSE,results='asis'}
+# Summarize by ID and period
+t2_df <- df %>%
+  group_by(eDWID, Period) %>%
+  summarise(
+    Smoking = max(smoking,na.rm = T),
+    `Secondhand smoke`=max(second_smoke,na.rm = T),
+    `FEV1 % Predicted` = last(na.omit(GLI_FEV1_pct_predicted)),
+    `FVC % Predicted` = last(na.omit(GLI_FVC_pct_predicted)),
+    `Pulm. Exacerbations` = last(na.omit(NumPulmExacerbation)),
+    .groups = "keep"
+  )
+# Replace infinite values with missing
+# t2_df <- data.frame(lapply(t2_df, function(c) {
+#   c[is.infinite(c)] <- NA
+#   return(c)
+# }))
+# Print table 1
+cols <- colnames(t2_df)[3:ncol(t2_df)]
+t2 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
+  data = t2_df
+)
+summary(t2)
+```
+
+# Questions
+
+1. Which variables do we include for microbiome stuff? Include "othermicroorganismstypes"?

### 2022-12-11 Christine Chan/CF Registry/cf_registry_analysis.Rmd
mutations added

@@
 knitr::opts_chunk$set(echo = FALSE)
-knitr::opts_knit$set(root.dir = "/home/tim/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data")
+knitr::opts_knit$set(root.dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data")
 ```
@@
 demo <- read.csv("CFF19_DemogCFDiag_Del1.csv")
+geno = read.csv("Data_Clean/genotypes.csv",header = F)
 # Combine and format demographics columns
@@
                                        "No"))
+# Genotypes
+df$Mutation1 = factor(df$Mutation1,levels = geno[geno$V1=="mutation1","V2"],
+                      labels = geno[geno$V1=="mutation1","V3"])
+df$Mutation2 = factor(df$Mutation2,levels = geno[geno$V1=="mutation2","V2"],
+                      labels = geno[geno$V1=="mutation2","V3"])
 # Save formatted data for re-import later

### 2022-12-12 Christine Chan/CF Registry/cf_registry_analysis.Rmd
might need to start over...
registry on last section

@@
 df$employment <- apply(df, 1, function(r) {
-  emp <- r[paste0("employment", 1:7)]
-  emp <- as.numeric(emp)
-  if (all(is.na(emp) | emp == 0)) {
+  emp <- r[paste0("employment",1:7)]
+  emp = as.numeric(emp)
+  names(emp) = c("Part Time",
+                    "Full time homemaker",
+                    "Full time employment",
+                    "Unemployed",
+                    "Student",
+                    "Disabled",
+                    "Retired"
+                    )
+  if (all(is.na(emp))) {
     return(NA)
   } else {
-    emp <- as.numeric(emp)
-    emp <- which(emp == 1)
+    emp <- names(emp)[which(emp == 1)]
     emp <- paste(emp, collapse = ", ")
@@
 })
-df$employment <- sub("1", "Part Time", df$employment)
-df$employment <- sub("2", "Full time homemaker", df$employment)
-df$employment <- sub("3", "Full time employment", df$employment)
-df$employment <- sub("4", "Unemployed", df$employment)
-df$employment <- sub("5", "Student", df$employment)
-df$employment <- sub("6", "Disabled", df$employment)
-df$employment <- sub("7", "Retired", df$employment)
-df$employment <- as.factor(df$employment)
 # Bronchodilator use
 df$bronchodilators <- apply(df, 1, function(r) {
-  bronch <- r[grep("bronchodilators",names(r))]
+  bronch <- r[grep("A_bronchodilators",names(r))]
+  bronch = as.numeric(bronch)
   names(bronch) = c("Beta agonist",
@@
   act <- r[grep("A_act_primary",names(r))]
+  act = as.numeric(act)
   names(act) = c("Positive Expiratory Pressure (PEP)",
@@
                       labels = geno[geno$V1=="mutation2","V3"])
+df$F508del = (df$Mutation1 =="F508del") + (df$Mutation2=="F508del")
+df$F508del = factor(df$F508del,levels = 0:2,labels = c("None","Heterozygous","Homozygous"))
+# Yes/no variables
+df$A_IsOnEnzymes = factor(df$A_IsOnEnzymes,levels = 0:1,labels = c("No","Yes"))
+df$A_salt_supplement = factor(df$A_salt_supplement,levels = 0:1,labels = c("No","Yes"))
+df$A_cf_vitamins = factor(df$A_cf_vitamins,levels = 0:1,labels = c("No","Yes"))
+df$A_dornasealfa = factor(df$A_dornasealfa,levels = 0:1,labels = c("No","Yes"))
+df$A_pulmonarycomplications4 = factor(df$A_pulmonarycomplications4,
+                                      levels = 0:1,labels = c("No","Yes"))
+df$A_pulmonarycomplications5 = factor(df$A_pulmonarycomplications5,
+                                      levels = 0:1,labels = c("No","Yes"))
+df$A_othercomplications9 = factor(df$A_othercomplications9,
+                                      levels = 0:1,labels = c("No","Yes"))
+df$A_hepatobiliary1_3 = factor(df$A_hepatobiliary1_3,
+                                      levels = 0:1,labels = c("No","Yes"))
 # Save formatted data for re-import later
@@
     `Pulm. Exacerbations` = last(na.omit(NumPulmExacerbation)),
+    `Bronchodilator` = last(na.omit(bronchodilators)),
+    `Primary ACT` = last(na.omit(act_primary)),
+    `Oxygen Therapy` = last(na.omit(oxygen_therapy)),
     .groups = "keep"
@@
 
+# Table 3: Panreactic function
+
+Continuous and count variables are from the last year of each period.
+
+```{r warning=FALSE,results='asis'}
+# Summarize by ID and period
+t3_df <- df %>%
+  group_by(eDWID, Period) %>%
+  summarise(
+    `On Enzymes` = last(na.omit(A_IsOnEnzymes)),
+    `Mutation Class` = last(na.omit(MutClass)),
+    `F508del`=last(na.omit(F508del)),
+    .groups = "keep"
+  )
+# Replace infinite values with missing
+# t3_df <- data.frame(lapply(t3_df, function(c) {
+#   c[is.infinite(c)] <- NA
+#   return(c)
+# }))
+# Print table 1
+cols <- colnames(t3_df)[3:ncol(t3_df)]
+t3 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
+  data = t3_df
+)
+summary(t3)
+```
+
+# Table 4: Nutrition
+
+Continuous and count variables are from the last year of each period.
+
+```{r warning=FALSE,results='asis'}
+# Summarize by ID and period
+t4_df <- df %>%
+  group_by(eDWID, Period) %>%
+  summarise(
+    `BMI %ile` = last(na.omit(A_bmipercentile)),
+    `Salt Supplement` = max(A_salt_supplement,na.rm = T),
+    `CF Vitamins` = max(A_cf_vitamins,na.rm = T),
+    `Hemoptysis` = max(A_pulmonarycomplications4,na.rm = T),
+    `Pneumothorax Chest Tube` = max(A_pulmonarycomplications5,na.rm = T),
+    `Sinus Disease` = max(A_othercomplications9,na.rm = T),
+    `Cirrhosis` = max(A_hepatobiliary1_3,na.rm = T),
+    .groups = "keep"
+  )
+# Replace infinite values with missing
+# t4_df <- data.frame(lapply(t4_df, function(c) {
+#   c[is.infinite(c)] <- NA
+#   return(c)
+# }))
+# Print table 1
+cols <- colnames(t4_df)[3:ncol(t4_df)]
+t4 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
+  data = t4_df
+)
+summary(t4)
+```
+
 # Questions

### 2022-12-12 Christine Chan/CF Registry/cf_registry_analysis.Rmd
might not need the apply stuff

@@
 # Yes/no variables
-df$A_IsOnEnzymes = factor(df$A_IsOnEnzymes,levels = 0:1,labels = c("No","Yes"))
-df$A_salt_supplement = factor(df$A_salt_supplement,levels = 0:1,labels = c("No","Yes"))
-df$A_cf_vitamins = factor(df$A_cf_vitamins,levels = 0:1,labels = c("No","Yes"))
-df$A_dornasealfa = factor(df$A_dornasealfa,levels = 0:1,labels = c("No","Yes"))
-df$A_pulmonarycomplications4 = factor(df$A_pulmonarycomplications4,
+df$A_IsOnEnzymes = factor(df$A_IsOnEnzymes,ordered = T,
                           levels = 0:1,labels = c("No","Yes"))
-df$A_pulmonarycomplications5 = factor(df$A_pulmonarycomplications5,
+df$A_salt_supplement = factor(df$A_salt_supplement,ordered = T,
                               levels = 0:1,labels = c("No","Yes"))
-df$A_othercomplications9 = factor(df$A_othercomplications9,
+df$A_cf_vitamins = factor(df$A_cf_vitamins,ordered = T,
                           levels = 0:1,labels = c("No","Yes"))
-df$A_hepatobiliary1_3 = factor(df$A_hepatobiliary1_3,
+df$A_dornasealfa = factor(df$A_dornasealfa,ordered = T,
+                          levels = 0:1,labels = c("No","Yes"))
+df$A_pulmonarycomplications4 = factor(df$A_pulmonarycomplications4,ordered = T,
+                                      levels = 0:1,labels = c("No","Yes"))
+df$A_pulmonarycomplications5 = factor(df$A_pulmonarycomplications5,ordered = T,
+                                      levels = 0:1,labels = c("No","Yes"))
+df$A_othercomplications9 = factor(df$A_othercomplications9,ordered = T,
+                                      levels = 0:1,labels = c("No","Yes"))
+df$A_hepatobiliary1_3 = factor(df$A_hepatobiliary1_3,ordered = T,
                                       levels = 0:1,labels = c("No","Yes"))
@@
     Employment = last(na.omit(employment)),
-    Pregnancy = max(pregnant, na.rm = T)
+    Pregnancy = max(pregnant, na.rm = T),
+    .groups = "keep"
   )

### 2022-12-12 Christine Chan/CF Registry/cf_registry_analysis.Rmd
6 registry tables done

--- 2022-12-11 Christine Chan/CF Registry/cf_registry_analysis.Rmd
@@
 library(parallel)
-knitr::opts_chunk$set(echo = FALSE)
-knitr::opts_knit$set(root.dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data")
+knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
+knitr::opts_knit$set(root.dir = "~/Work/CHCO/Christine Chan/CF registry data")
 ```
 
-```{r eval=FALSE}
+```{r}
 # Read in CSVs
@@
 demo <- read.csv("CFF19_DemogCFDiag_Del1.csv")
-geno = read.csv("Data_Clean/genotypes.csv",header = F)
+geno <- read.csv("Data_Clean/genotypes.csv", header = F)
 # Combine and format demographics columns
@@
 )
-df$pregnant <- factor(df$pregnant, levels = 0:1, labels = c("No", "Yes"), ordered = T)
-# Combine education checkboxes
-df$employment <- apply(df, 1, function(r) {
-  emp <- r[paste0("employment", 1:7)]
-  emp <- as.numeric(emp)
-  if (all(is.na(emp) | emp == 0)) {
-    return(NA)
-  } else {
-    emp <- as.numeric(emp)
-    emp <- which(emp == 1)
-    emp <- paste(emp, collapse = ", ")
-    return(emp)
-  }
-})
-df$employment <- sub("1", "Part Time", df$employment)
-df$employment <- sub("2", "Full time homemaker", df$employment)
-df$employment <- sub("3", "Full time employment", df$employment)
-df$employment <- sub("4", "Unemployed", df$employment)
-df$employment <- sub("5", "Student", df$employment)
-df$employment <- sub("6", "Disabled", df$employment)
-df$employment <- sub("7", "Retired", df$employment)
-df$employment <- as.factor(df$employment)
-# Bronchodilator use
-df$bronchodilators <- apply(df, 1, function(r) {
-  bronch <- r[grep("bronchodilators",names(r))]
-  names(bronch) = c("Beta agonist",
-                    "Theophylline product",
-                    "Short acting beta agonist",
-                    "Long acting beta agonist",
-                    "Short acting anticholinergic",
-                    "Long acting anticholinergic",
-                    "Combination beta agonist and anticholinergic"
-  )
-  if (all(is.na(bronch))) {
-    return(NA)
-  } else {
-    bronch <- names(bronch)[which(bronch == 1)]
-    bronch <- paste(bronch, collapse = ", ")
-    return(bronch)
-  }
-})
-# Exacerbation medication
-df$act_primary <- apply(df, 1, function(r) {
-  act <- r[grep("A_act_primary",names(r))]
-  names(act) = c("Positive Expiratory Pressure (PEP)",
-                 "Postural drainage with clapping (CPT)",
-                 "Forced expiratory techniques",
-                 "Oscillating PEP",
-                 "High frequency chest wall oscillation",
-                 "Exercise",
-                 "None",
-                 "Other"
-  )
-  if (all(is.na(act) | act == 0)) {
-    return(NA)
-  } else {
-    act <- names(act)[which(act == 1)]
-    act <- paste(act, collapse = ", ")
-    return(act)
-  }
-})
-# Oxygen therapy
-df$oxygen_therapy <- factor(df$oxygen_therapy,levels = 1:5,
-                            labels = c("Yes, Continuously",
+df$oxygen_therapy <- factor(df$oxygen_therapy,
+                            levels = 1:6,
+                            labels = c(
+                              "Yes, Continuously",
                               "Yes, Nocturnal and/or with exertion",
@@
                               "Yes, prn",
-                                       "No"))
+                              "No",
+                              "Unknown"
+                            )
+)
 # Genotypes
-df$Mutation1 = factor(df$Mutation1,levels = geno[geno$V1=="mutation1","V2"],
-                      labels = geno[geno$V1=="mutation1","V3"])
-df$Mutation2 = factor(df$Mutation2,levels = geno[geno$V1=="mutation2","V2"],
-                      labels = geno[geno$V1=="mutation2","V3"])
-# Save formatted data for re-import later
-save(df,file = "Data_Clean/formatted_data.Rdata")
-```
-
-```{r}
-load("Data_Clean/formatted_data.Rdata")
-df = df %>% arrange(eDWID,reviewyear,encounternum)
+df$Mutation1 <- factor(df$Mutation1,
+                       levels = geno[geno$V1 == "mutation1", "V2"],
+                       labels = geno[geno$V1 == "mutation1", "V3"]
+)
+df$Mutation2 <- factor(df$Mutation2,
+                       levels = geno[geno$V1 == "mutation2", "V2"],
+                       labels = geno[geno$V1 == "mutation2", "V3"]
+)
+df$F508del <- (df$Mutation1 == "F508del") + (df$Mutation2 == "F508del")
+df$F508del <- factor(df$F508del, levels = 0:2, 
+                     labels = c("None", "Heterozygous", "Homozygous"))
+# Percentiles for peds
+df$bmi_peds = df$A_bmipercentile * (df$Age_YrEnd < 18)
+df$bmi_adults = df$A_bmivalue * (df$Age_YrEnd >= 18)
+df$height_peds = df$A_heightpercentile * (df$Age_YrEnd < 18)
+df$height_adults = df$A_height * (df$Age_YrEnd >= 18)
+df$weight_peds = df$A_weightpercentile * (df$Age_YrEnd < 18)
+df$weight_adults = df$A_weight * (df$Age_YrEnd >= 18)
+# Yes/no variables
+yn <- c("A_IsOnEnzymes", "A_salt_supplement", "A_cf_vitamins", "A_dornasealfa", 
+        "A_pulmonarycomplications4", "A_pulmonarycomplications5", 
+        "A_othercomplications9", paste0("A_hepatobiliary1_",1:3), 
+        paste0("A_hepatobiliary2_",1:4),paste0("employment", 1:7),
+        "pregnant","dexa_results_osteoporosis","A_Vx770","A_VX809comb",
+        "A_feeding3","A_corticosteroids1","A_pulmonarycomplications2",
+        paste0("A_mycobacterials",c(1:5,7:10)),
+        "A_pseudomonasaeruginosa","A_staphylococcus_aureus",
+        "A_othermicroorganisms1","A_othermicroorganisms2",
+        "A_fungalyeast1","A_fungalyeast2","A_fungalyeast3")
+df[, yn] <- lapply(df[, yn], function(c) {
+  factor(c, levels = 0:1, labels = c("No", "Yes"), ordered = T)
+})
 ```
@@
 
-```{r warning=FALSE,results='asis'}
+```{r results='asis'}
 # Summarize by ID and period
@@
     `Marital Status` = last(na.omit(marital_status)),
-    Employment = last(na.omit(employment)),
-    Pregnancy = max(pregnant, na.rm = T)
+    `Part Time` = max(employment1, na.rm = F),
+    `Full time homemaker` = max(employment2, na.rm = F),
+    `Full time employment` = max(employment3, na.rm = F),
+    `Unemployed` = max(employment4, na.rm = F),
+    `Student` = max(employment5, na.rm = F),
+    `Disabled` = max(employment6, na.rm = F),
+    `Retired` = max(employment7, na.rm = F),
+    Pregnancy = max(pregnant, na.rm = T),
+    .groups = "keep"
   )
-# Replace infinite values with missing
-t1_df <- data.frame(lapply(t1_df, function(c) {
-  c[is.infinite(c)] <- NA
-  return(c)
-}))
-# Print table 1
+# Print table
 cols <- colnames(t1_df)[3:ncol(t1_df)]
@@
   summarise(
-    Smoking = max(smoking,na.rm = T),
-    `Secondhand smoke`=max(second_smoke,na.rm = T),
+    Smoking = max(smoking, na.rm = T),
+    `Secondhand smoke` = max(second_smoke, na.rm = T),
     `FEV1 % Predicted` = last(na.omit(GLI_FEV1_pct_predicted)),
     `FVC % Predicted` = last(na.omit(GLI_FVC_pct_predicted)),
-    `Pulm. Exacerbations` = last(na.omit(NumPulmExacerbation)),
+    `Pulm. Exacerbations` = sum(NumPulmExacerbation,na.rm = T),
+    `Beta agonist` = max(bronchodilators1, na.rm = T),
+    `Theophylline product` = max(bronchodilators2, na.rm = T),
+    `Short acting beta agonist` = max(bronchodilatorsinhaled1, na.rm = T),
+    `Long acting beta agonist` = max(bronchodilatorsinhaled2, na.rm = T),
+    `Short acting anticholinergic` = max(bronchodilatorsinhaled3, na.rm = T),
+    `Long acting anticholinergic` = max(bronchodilatorsinhaled4, na.rm = T),
+    `Combination beta agonist and anticholinergic` = max(bronchodilatorsinhaled5, na.rm = T),
+    `Positive Expiratory Pressure (PEP)` = max(act_primary1, na.rm = T),
+    `Postural drainage with clapping (CPT)` = max(act_primary2, na.rm = T),
+    `Forced expiratory techniques` = max(act_primary3, na.rm = T),
+    `Oscillating PEP` = max(act_primary4, na.rm = T),
+    `High frequency chest wall oscillation` = max(act_primary5, na.rm = T),
+    `Exercise` = max(act_primary6, na.rm = T),
+    `None` = max(act_primary7, na.rm = T),
+    `Other` = max(act_primary8, na.rm = T),
+    `Oxygen Therapy` = last(na.omit(oxygen_therapy)),
     .groups = "keep"
   )
-# Replace infinite values with missing
-# t2_df <- data.frame(lapply(t2_df, function(c) {
-#   c[is.infinite(c)] <- NA
-#   return(c)
-# }))
-# Print table 1
+# Print table
 cols <- colnames(t2_df)[3:ncol(t2_df)]
@@
 
-# Questions
+# Table 3: Panreactic function
+
+Continuous and count variables are from the last year of each period.
+
+```{r results='asis'}
+# Summarize by ID and period
+t3_df <- df %>%
+  group_by(eDWID, Period) %>%
+  summarise(
+    `On Enzymes` = last(na.omit(A_IsOnEnzymes)),
+    `Mutation Class` = last(na.omit(MutClass)),
+    `F508del` = last(na.omit(F508del)),
+    .groups = "keep"
+  )
+# Print table
+cols <- colnames(t3_df)[3:ncol(t3_df)]
+t3 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
+              data = t3_df
+)
+summary(t3)
+```
+
+# Table 4: Nutrition
+
+Continuous and count variables are from the last year of each period.
+
+```{r results='asis'}
+# Summarize by ID and period
+t4_df <- df %>%
+  group_by(eDWID, Period) %>%
+  summarise(
+    `BMI %ile (age <18)` = last(na.omit(bmi_peds)),
+    `BMI Value (age <= 18)` = last(na.omit(bmi_adults)),
+    `Salt Supplement` = max(A_salt_supplement, na.rm = T),
+    `CF Vitamins` = max(A_cf_vitamins, na.rm = T),
+    `Hemoptysis` = max(A_pulmonarycomplications4, na.rm = T),
+    `Pneumothorax Chest Tube` = max(A_pulmonarycomplications5, na.rm = T),
+    `Sinus Disease` = max(A_othercomplications9, na.rm = T),
+    `Cirrhosis` = max(A_hepatobiliary1_3, na.rm = T),
+    .groups = "keep"
+  )
+# Print table
+cols <- colnames(t4_df)[3:ncol(t4_df)]
+t4 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
+              data = t4_df
+)
+summary(t4)
+```
+
+# Table 5: Other
+
+Continuous and count variables are from the last year of each period.
+
+```{r results='asis'}
+# Summarize by ID and period
+t5_df <- df %>%
+  group_by(eDWID, Period) %>%
+  summarise(
+    `Osteoperosis`=max(dexa_results_osteoporosis,na.rm = T),
+    `Kalydeco`=max(A_Vx770,na.rm = T),
+    `Lumacaftor/Ivacaftor`=max(A_VX809comb,na.rm = T),
+    `Height %ile (age <18)` = last(na.omit(height_peds)),
+    `Height Value (age <= 18)` = last(na.omit(height_adults)),
+    `Weight %ile (age <18)` = last(na.omit(weight_peds)),
+    `Weight Value (age <= 18)` = last(na.omit(weight_adults)),
+    `Num. Hospitalizations` = sum(NumHosp,na.rm = T),
+    `Gall stones`=max(hepatobiliary1_1,na.rm=T),
+    `Gall stones, requiring surgery/procedure`=max(hepatobiliary1_2,na.rm=T),
+    `Liver disease, cirrhosis`=max(hepatobiliary1_3,na.rm=T),
+    `Liver disease, non- cirrhosis`=max(hepatobiliary2_1,na.rm=T),
+    `Hepatic Steatosis`=max(hepatobiliary2_2,na.rm=T),
+    `Liver disease, other`=max(hepatobiliary2_3,na.rm=T),
+    `Acute Liver Failure`=max(hepatobiliary2_4,na.rm=T),
+    `G-Tube`=max(A_feeding3,na.rm = T),
+    `Oral corticosteroids`=max(A_corticosteroids1,na.rm = T),
+    `Asthma`=max(A_pulmonarycomplications2,na.rm = T),
+    `Mycobacterial tuberculosis`=max(A_mycobacterials1,na.rm=T),
+    `Mycobacterium abscessus/chelonae`=max(A_mycobacterials2,na.rm=T),
+    `Mycobacterium avium complex (MAC)`=max(A_mycobacterials3,na.rm=T),
+    `Mycobacterium fortuitum group`=max(A_mycobacterials4,na.rm=T),
+    `Mycobacterium gordonae`=max(A_mycobacterials5,na.rm=T),
+    `Mycobacterium kansasii`=max(A_mycobacterials7,na.rm=T),
+    `Mycobacterium marinum`=max(A_mycobacterials8,na.rm=T),
+    `Mycobacterium terrae`=max(A_mycobacterials9,na.rm=T),
+    `Other Mycobacterium`=max(A_mycobacterials10,na.rm=T),
+    .groups = "keep"
+  )
+# Print table
+cols <- colnames(t5_df)[3:ncol(t5_df)]
+t5 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
+              data = t5_df
+)
+summary(t5)
+```
+
+# Table 6: Microbiome
+
+```{r results='asis'}
+# Summarize by ID and period
+t6_df <- df %>%
+  group_by(eDWID, Period) %>%
+  summarise(
+    `Pseudomonas`=max(A_pseudomonasaeruginosa,na.rm=T),
+    `Burkholderia`=max(A_staphylococcus_aureus,na.rm=T),
+    `Alcaligenes`=max(A_othermicroorganisms1,na.rm=T),
+    `Stenotrophomonas`=max(A_othermicroorganisms2,na.rm=T),
+    `Aspergillus`=max(A_fungalyeast1,na.rm=T),
+    `Candida`=max(A_fungalyeast2,na.rm=T),
+    `Scedosporium`=max(A_fungalyeast3,na.rm=T),
+    .groups = "keep"
+  )
+# Print table
+cols <- colnames(t6_df)[3:ncol(t6_df)]
+t6 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
+              data = t6_df
+)
+summary(t6)
+```
+
+# Questions/notes
 
 1. Which variables do we include for microbiome stuff? Include "othermicroorganismstypes"?
+2. We don't seem to have ETI data.

### 2022-12-12 Christine Chan/CF Registry/cf_registry_analysis.Rmd
lambda path

@@
 knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
-knitr::opts_knit$set(root.dir = "~/Work/CHCO/Christine Chan/CF registry data")
+knitr::opts_knit$set(root.dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data")
 ```

### 2022-12-13 Christine Chan/CF Registry/cf_registry_analysis.Rmd
CFRD registry

--- 2022-12-12 Christine Chan/CF Registry/cf_registry_analysis.Rmd
@@
     toc: yes
-    toc_depth: 2
-    toc_float: yes
+    toc_depth: 5
+    toc_float: no
+    page-layout: full
 ---
@@
 ```{r setup, include=FALSE}
+library(JM)
 library(arsenal)
@@
 library(knitr)
-library(parallel)
-knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
+library(pROC)
+library(lavaan)
+library(survMisc)
+knitr::opts_chunk$set(echo = FALSE,cache = TRUE,cache.lazy = FALSE)
 knitr::opts_knit$set(root.dir = "~/Work/CHCO/Christine Chan/CF registry data")
@@
 # Read in CSVs
-df <- read.csv("CFF19_encountersMerged_Del1.csv")
-annual <- read.csv("CFF19_Annualized_Del1.csv")
+df <- read.csv("CFF19_encountersMerged_Del1.csv",na.strings = c("","N/A"))
+annual <- read.csv("CFF19_Annualized_Del1.csv",na.strings = c("","N/A"))
 colnames(annual)[colnames(annual) == "ReviewYear"] <- "reviewyear"
-demo <- read.csv("CFF19_DemogCFDiag_Del1.csv")
-geno <- read.csv("Data_Clean/genotypes.csv", header = F)
+demo <- read.csv("CFF19_DemogCFDiag_Del1.csv",na.strings = c("","N/A"))
+geno <- read.csv("Data_Clean/genotypes.csv", header = F,na.strings = c("","N/A"))
 # Combine and format demographics columns
@@
 df$F508del <- (df$Mutation1 == "F508del") + (df$Mutation2 == "F508del")
-df$F508del <- factor(df$F508del, levels = 0:2, 
+df$F508del <- factor(df$F508del, levels = 0:2, ordered = T,
                      labels = c("None", "Heterozygous", "Homozygous"))
@@
         "A_othermicroorganisms1","A_othermicroorganisms2",
-        "A_fungalyeast1","A_fungalyeast2","A_fungalyeast3")
+        "A_fungalyeast1","A_fungalyeast2","A_fungalyeast3",
+        paste0("A_act_primary",1:8),paste0("A_bronchodilators",1:2),
+        paste0("A_bronchodilatorsinhaled",1:5))
 df[, yn] <- lapply(df[, yn], function(c) {
@@
 })
+df$Gender = as.factor(df$Gender)
+# CFRD diagnosis
+df$cfrd = df$cfrd_status == 3
+df$cfrd = factor(df$cfrd,levels = c(F,T),labels = c("CFRD-","CFRD+"),ordered = T)
+# Sort
+df = df %>% arrange(eDWID,reviewyear,encounternum)
 ```
 
-# Table 1: Participant characteristics at last visit in time period
+# Aim 1
 
-```{r results='asis'}
+## CFRD Incidence
+
+```{r warning=FALSE}
+df %>% group_by(reviewyear,eDWID) %>% 
+  summarise(cfrd = max(cfrd,na.rm = T),
+            age = max(Age_YrEnd,na.rm = T),
+            .groups = "drop_last") %>%
+  summarise(`CFRD Cases` = sum(cfrd=="CFRD+",na.rm = T),
+            `Total People with CF` = n(),
+            `CFRD Prevalence` =`CFRD Cases`/`Total People with CF`,
+            .groups = "drop") %>%
+  mutate(`New Cases` = `CFRD Cases`-`CFRD Cases`[1],
+         `CFRD Incidence` = `New Cases`/`Total People with CF`) %>%
+  kable(.)
+```
+
+## Table 1: Participant characteristics at last visit in time period
+
+```{r results='asis',warning=FALSE}
 # Summarize by ID and period
@@
   summarise(
+    CFRD = max(cfrd,na.rm = T),
     Age = max(Age_YrEnd), Race = last(na.omit(Race)),
@@
   )
+# Convert infinite values to missing
+t1_df = data.frame(lapply(t1_df, function(c){
+  c[is.infinite(c)] = NA
+  return(c)
+}))
 # Print table
-cols <- colnames(t1_df)[3:ncol(t1_df)]
-t1 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
-              data = t1_df
-)
+cols <- colnames(t1_df)[4:ncol(t1_df)]
+t1 = tableby(as.formula(paste0("CFRD~", paste0("`", cols, "`", collapse = "+"))),
+             strata = Period,data = t1_df)
 summary(t1)
@@
 
-# Table 2: Pulmonary function
+## Table 2: Pulmonary function
 
@@
 
-```{r results='asis'}
+```{r results='asis',warning=FALSE}
 # Summarize by ID and period
@@
   summarise(
+    CFRD = max(cfrd,na.rm = T),
     Smoking = max(smoking, na.rm = T),
     `Secondhand smoke` = max(second_smoke, na.rm = T),
-    `FEV1 % Predicted` = last(na.omit(GLI_FEV1_pct_predicted)),
-    `FVC % Predicted` = last(na.omit(GLI_FVC_pct_predicted)),
+    `FEV1 % Predicted` = last(na.omit(A_GLI_FEV1_pct_predicted)),
+    `FVC % Predicted` = last(na.omit(A_GLI_FVC_pct_predicted)),
     `Pulm. Exacerbations` = sum(NumPulmExacerbation,na.rm = T),
-    `Beta agonist` = max(bronchodilators1, na.rm = T),
-    `Theophylline product` = max(bronchodilators2, na.rm = T),
-    `Short acting beta agonist` = max(bronchodilatorsinhaled1, na.rm = T),
-    `Long acting beta agonist` = max(bronchodilatorsinhaled2, na.rm = T),
-    `Short acting anticholinergic` = max(bronchodilatorsinhaled3, na.rm = T),
-    `Long acting anticholinergic` = max(bronchodilatorsinhaled4, na.rm = T),
-    `Combination beta agonist and anticholinergic` = max(bronchodilatorsinhaled5, na.rm = T),
-    `Positive Expiratory Pressure (PEP)` = max(act_primary1, na.rm = T),
-    `Postural drainage with clapping (CPT)` = max(act_primary2, na.rm = T),
-    `Forced expiratory techniques` = max(act_primary3, na.rm = T),
-    `Oscillating PEP` = max(act_primary4, na.rm = T),
-    `High frequency chest wall oscillation` = max(act_primary5, na.rm = T),
-    `Exercise` = max(act_primary6, na.rm = T),
-    `None` = max(act_primary7, na.rm = T),
-    `Other` = max(act_primary8, na.rm = T),
+    `Beta agonist` = max(A_bronchodilators1, na.rm = T),
+    `Theophylline product` = max(A_bronchodilators2, na.rm = T),
+    `Short acting beta agonist` = max(A_bronchodilatorsinhaled1, na.rm = T),
+    `Long acting beta agonist` = max(A_bronchodilatorsinhaled2, na.rm = T),
+    `Short acting anticholinergic` = max(A_bronchodilatorsinhaled3, na.rm = T),
+    `Long acting anticholinergic` = max(A_bronchodilatorsinhaled4, na.rm = T),
+    `Combination beta agonist and anticholinergic` = max(A_bronchodilatorsinhaled5, na.rm = T),
+    `Positive Expiratory Pressure (PEP)` = max(A_act_primary1, na.rm = T),
+    `Postural drainage with clapping (CPT)` = max(A_act_primary2, na.rm = T),
+    `Forced expiratory techniques` = max(A_act_primary3, na.rm = T),
+    `Oscillating PEP` = max(A_act_primary4, na.rm = T),
+    `High frequency chest wall oscillation` = max(A_act_primary5, na.rm = T),
+    `Exercise` = max(A_act_primary6, na.rm = T),
+    `None` = max(A_act_primary7, na.rm = T),
+    `Other` = max(A_act_primary8, na.rm = T),
     `Oxygen Therapy` = last(na.omit(oxygen_therapy)),
@@
   )
+# Convert infinite values to missing
+t2_df = data.frame(lapply(t2_df, function(c){
+  c[is.infinite(c)] = NA
+  return(c)
+}))
 # Print table
-cols <- colnames(t2_df)[3:ncol(t2_df)]
-t2 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
-              data = t2_df
-)
+cols <- colnames(t2_df)[4:ncol(t2_df)]
+t2 = tableby(as.formula(paste0("CFRD~", paste0("`", cols, "`", collapse = "+"))),
+             strata = Period,data = t2_df)
 summary(t2)
@@
 
-# Table 3: Panreactic function
+## Table 3: Panreactic function
 
@@
   summarise(
+    CFRD = max(cfrd,na.rm = T),
     `On Enzymes` = last(na.omit(A_IsOnEnzymes)),
@@
   )
+# Convert infinite values to missing
+t3_df = data.frame(lapply(t3_df, function(c){
+  c[is.infinite(c)] = NA
+  return(c)
+}))
 # Print table
-cols <- colnames(t3_df)[3:ncol(t3_df)]
-t3 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
-              data = t3_df
-)
+cols <- colnames(t3_df)[4:ncol(t3_df)]
+t3 = tableby(as.formula(paste0("CFRD~", paste0("`", cols, "`", collapse = "+"))),
+             strata = Period,data = t3_df)
 summary(t3)
@@
 
-# Table 4: Nutrition
+## Table 4: Nutrition
 
@@
   summarise(
+    CFRD = max(cfrd,na.rm = T),
     `BMI %ile (age <18)` = last(na.omit(bmi_peds)),
@@
   )
+# Convert infinite values to missing
+t4_df = data.frame(lapply(t4_df, function(c){
+  c[is.infinite(c)] = NA
+  return(c)
+}))
 # Print table
-cols <- colnames(t4_df)[3:ncol(t4_df)]
-t4 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
-              data = t4_df
-)
+cols <- colnames(t4_df)[4:ncol(t4_df)]
+t4 = tableby(as.formula(paste0("CFRD~", paste0("`", cols, "`", collapse = "+"))),
+             strata = Period,data = t4_df)
 summary(t4)
@@
 
-# Table 5: Other
+## Table 5: Other
 
@@
   summarise(
+    CFRD = max(cfrd,na.rm = T),
     `Osteoperosis`=max(dexa_results_osteoporosis,na.rm = T),
@@
     `Num. Hospitalizations` = sum(NumHosp,na.rm = T),
-    `Gall stones`=max(hepatobiliary1_1,na.rm=T),
-    `Gall stones, requiring surgery/procedure`=max(hepatobiliary1_2,na.rm=T),
-    `Liver disease, cirrhosis`=max(hepatobiliary1_3,na.rm=T),
-    `Liver disease, non- cirrhosis`=max(hepatobiliary2_1,na.rm=T),
-    `Hepatic Steatosis`=max(hepatobiliary2_2,na.rm=T),
-    `Liver disease, other`=max(hepatobiliary2_3,na.rm=T),
-    `Acute Liver Failure`=max(hepatobiliary2_4,na.rm=T),
+    `Gall stones`=max(A_hepatobiliary1_1,na.rm=T),
+    `Liver disease, cirrhosis`=max(A_hepatobiliary1_3,na.rm=T),
+    `Liver disease, non- cirrhosis`=max(A_hepatobiliary2_1,na.rm=T),
+    `Hepatic Steatosis`=max(A_hepatobiliary2_2,na.rm=T),
+    `Liver disease, other`=max(A_hepatobiliary2_3,na.rm=T),
+    `Acute Liver Failure`=max(A_hepatobiliary2_4,na.rm=T),
     `G-Tube`=max(A_feeding3,na.rm = T),
@@
   )
+# Convert infinite values to missing
+t5_df = data.frame(lapply(t5_df, function(c){
+  c[is.infinite(c)] = NA
+  return(c)
+}))
 # Print table
-cols <- colnames(t5_df)[3:ncol(t5_df)]
-t5 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
-              data = t5_df
-)
+cols <- colnames(t5_df)[4:ncol(t5_df)]
+t5 = tableby(as.formula(paste0("CFRD~", paste0("`", cols, "`", collapse = "+"))),
+             strata = Period,data = t5_df)
 summary(t5)
@@
 
-# Table 6: Microbiome
+## Table 6: Microbiome
 
-```{r results='asis'}
+```{r results='asis',warning=FALSE}
 # Summarize by ID and period
@@
   summarise(
+    CFRD = max(cfrd,na.rm = T),
     `Pseudomonas`=max(A_pseudomonasaeruginosa,na.rm=T),
@@
   )
+# Convert infinite values to missing
+t6_df = data.frame(lapply(t6_df, function(c){
+  c[is.infinite(c)] = NA
+  return(c)
+}))
 # Print table
-cols <- colnames(t6_df)[3:ncol(t6_df)]
-t6 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
-              data = t6_df
-)
+cols <- colnames(t6_df)[4:ncol(t6_df)]
+t6 = tableby(as.formula(paste0("CFRD~", paste0("`", cols, "`", collapse = "+"))),
+             strata = Period,data = t6_df)
 summary(t6)
+```
+
+# Aim 2
+
+## Structural equation modeling (SEM)
+
+```{r echo=TRUE,eval=FALSE}
+model <- '
+  # regressions
+    GLI_FEV1_pct_predicted~cfrd
+    cfrd ~ Gender
+  # residual correlations
+    Gender ~~ corticosteroids1
+    Gender ~~ F508del
+    cfrd ~~ hgba1c
+    cfrd ~~ bmipercentile
+'
+fit <- sem(model, data = df)
+summary(fit, standardized = TRUE)
+```
+
+# Aim 3
+
+## ROC curves
+
+For each ROC curve, we used everyone's most recent visit where they had both a CFRD status and an A1c measure.
+Confidence intervals for the threshold that maximizes sensitivity and specificity (based on the Youden index) were calculated using 2000 bootstrap replicates.
+
+### CFRD
+
+```{r}
+# Clean data
+roc_df = df %>% group_by(eDWID) %>%
+  select(eDWID,Period,reviewyear,encounterid,encounternum,hgba1c,cfrd_status_annual) %>%
+  drop_na() %>% slice_max(reviewyear) %>% slice_max(encounternum)
+roc_df$cfrd_status = as.numeric(roc_df$cfrd_status_annual == 3)
+r = roc(cfrd_status~hgba1c,roc_df,plot=TRUE, print.auc=TRUE,ci=T,quiet=T)
+thresh.ci = ci.coords(r,"best",best.method="youden",progress="none")
+n = names(thresh.ci)
+thresh.ci = do.call(rbind,thresh.ci)
+rownames(thresh.ci)=n
+kable(thresh.ci,row.names = T)
+```
+
+### CFRD-related retinopathy
+
+```{r}
+# Clean data
+roc_df = df %>% group_by(eDWID) %>%
+  select(eDWID,Period,reviewyear,encounterid,encounternum,hgba1c,A_cfrdsecondarycomplications1) %>%
+  drop_na() %>% slice_max(reviewyear) %>% slice_max(encounternum)
+r = roc(A_cfrdsecondarycomplications1~hgba1c,roc_df,plot=TRUE, print.auc=TRUE,ci=T,quiet=T)
+thresh.ci = ci.coords(r,"best",best.method="youden",progress="none")
+n = names(thresh.ci)
+thresh.ci = do.call(rbind,thresh.ci)
+rownames(thresh.ci)=n
+kable(thresh.ci,row.names = T)
+```
+
+### CFRD-related microalbuminuria
+
+```{r}
+# Clean data
+roc_df = df %>% group_by(eDWID) %>%
+  select(eDWID,Period,reviewyear,encounterid,encounternum,hgba1c,A_cfrdsecondarycomplications2) %>%
+  drop_na() %>% slice_max(reviewyear) %>% slice_max(encounternum)
+r = roc(A_cfrdsecondarycomplications2~hgba1c,roc_df,plot=TRUE, print.auc=TRUE,ci=T,quiet=T)
+thresh.ci = ci.coords(r,"best",best.method="youden",progress="none")
+n = names(thresh.ci)
+thresh.ci = do.call(rbind,thresh.ci)
+rownames(thresh.ci)=n
+kable(thresh.ci,row.names = T)
+```
+
+### CFRD-related chronic renal insufficiency
+
+```{r}
+# Clean data
+roc_df = df %>% group_by(eDWID) %>%
+  select(eDWID,Period,reviewyear,encounterid,encounternum,hgba1c,A_cfrdsecondarycomplications3) %>%
+  drop_na() %>% slice_max(reviewyear) %>% slice_max(encounternum)
+r = roc(A_cfrdsecondarycomplications3~hgba1c,roc_df,plot=TRUE, print.auc=TRUE,ci=T,quiet=T)
+thresh.ci = ci.coords(r,"best",best.method="youden",progress="none")
+n = names(thresh.ci)
+thresh.ci = do.call(rbind,thresh.ci)
+rownames(thresh.ci)=n
+kable(thresh.ci,row.names = T)
+```
+
+### CFRD-related chronic renal insufficiency
+
+```{r}
+# Clean data
+roc_df = df %>% group_by(eDWID) %>%
+  select(eDWID,Period,reviewyear,encounterid,encounternum,hgba1c,A_cfrdsecondarycomplications3) %>%
+  drop_na() %>% slice_max(reviewyear) %>% slice_max(encounternum)
+r = roc(A_cfrdsecondarycomplications3~hgba1c,roc_df,plot=TRUE, print.auc=TRUE,ci=T,quiet=T)
+thresh.ci = ci.coords(r,"best",best.method="youden",progress="none")
+n = names(thresh.ci)
+thresh.ci = do.call(rbind,thresh.ci)
+rownames(thresh.ci)=n
+kable(thresh.ci,row.names = T)
+```
+
+### CFRD-related chronic renal failure requiring dialysis
+
+```{r}
+# Clean data
+roc_df = df %>% group_by(eDWID) %>%
+  select(eDWID,Period,reviewyear,encounterid,encounternum,hgba1c,A_cfrdsecondarycomplications4) %>%
+  drop_na() %>% slice_max(reviewyear) %>% slice_max(encounternum)
+r = roc(A_cfrdsecondarycomplications4~hgba1c,roc_df,plot=TRUE, print.auc=TRUE,ci=T,quiet=T)
+thresh.ci = ci.coords(r,"best",best.method="youden",progress="none")
+n = names(thresh.ci)
+thresh.ci = do.call(rbind,thresh.ci)
+rownames(thresh.ci)=n
+kable(thresh.ci,row.names = T)
+```
+
+### CFRD-related peripheral neuropathy
+
+```{r}
+# Clean data
+roc_df = df %>% group_by(eDWID) %>%
+  select(eDWID,Period,reviewyear,encounterid,encounternum,hgba1c,A_cfrdsecondarycomplications5) %>%
+  drop_na() %>% slice_max(reviewyear) %>% slice_max(encounternum)
+r = roc(A_cfrdsecondarycomplications5~hgba1c,roc_df,plot=TRUE, print.auc=TRUE,ci=T,quiet=T)
+thresh.ci = ci.coords(r,"best",best.method="youden",progress="none")
+n = names(thresh.ci)
+thresh.ci = do.call(rbind,thresh.ci)
+rownames(thresh.ci)=n
+kable(thresh.ci,row.names = T)
+```
+
+## Joint model
+
+```{r eval=FALSE,echo=TRUE}
+annual = left_join(annual,demo,by = "eDWID")
+# Super simple joint model to start with
+jm_df = annual %>% dplyr::select(eDWID,reviewyear,Gender,A_hgba1c,cfrd_status_annual) %>%
+  filter(cfrd_status_annual <= 3)
+jm_df$cfrd_status_annual = as.numeric(jm_df$cfrd_status_annual==3)
+jm_df = jm_df %>% group_by(eDWID) %>% 
+  mutate(cfrd = max(cfrd_status_annual),
+         reviewyear = reviewyear-2003)
+jm_df$stop = 
+# Survival model
+sf <- survfit(Surv(reviewyear, cfrd) ~ factor(Gender), data = jm_df)
+autoplot(sf)
 ```
@@
 
-1. Which variables do we include for microbiome stuff? Include "othermicroorganismstypes"?
-2. We don't seem to have ETI data.
+1. In the cleaned ROC data about 34% of people had CFRD. Does that seem correct?
+2. I also need to meet with Laura about the SEM since she's an expert.
+  - Do we want to include any latent variables?
+  - How should we treat BMI in the SEM? Percentile or raw value adjusted for age?
+3. Unfortunately some of the complicated models are not converging, so I need to investigate that a bit.
+4. What is the best way to treat time in these models? Should we use annualized data and assum even spacing?

### 2022-12-15 Christine Chan/CF Registry/cf_registry_analysis.Rmd
starting over CF registry

--- 2022-12-12 Christine Chan/CF Registry/cf_registry_analysis.Rmd
@@
 ```{r setup, include=FALSE}
-library(arsenal)
-library(tidyverse)
-library(knitr)
-library(parallel)
-knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
+library(table1)
+knitr::opts_chunk$set(echo = FALSE,cache = TRUE,cache.lazy = FALSE)
 knitr::opts_knit$set(root.dir = "~/Work/CHCO/Christine Chan/CF registry data")
@@
 ```{r}
-# Read in CSVs
-df <- read.csv("CFF19_encountersMerged_Del1.csv")
-annual <- read.csv("CFF19_Annualized_Del1.csv")
-colnames(annual)[colnames(annual) == "ReviewYear"] <- "reviewyear"
-demo <- read.csv("CFF19_DemogCFDiag_Del1.csv")
-geno <- read.csv("Data_Clean/genotypes.csv", header = F)
-# Combine and format demographics columns
-demo$Race <- apply(demo, 1, function(r) {
-  race <- r[paste0("Race", 1:6)]
-  race <- as.numeric(which(race == 1))
-  race <- paste(race, collapse = ", ")
-})
-demo$Race[grep(", ", demo$Race)] <- "More Than One"
-demo$Race <- factor(demo$Race,
-                    levels = c(1:6, "More Than One"),
-                    labels = c(
-                      "White", "Black or African American",
-                      "American Indian or Alaska Native", "Asian",
-                      "Native Hawaiian or Other Pacific Islander",
-                      "Some other race", "More Than One"
-                    )
-)
-demo$Hispanicrace <- factor(demo$Hispanicrace,
-                            levels = 1:2,
-                            labels = c("Yes", "No")
-)
-# Merge
-df <- left_join(df, annual, by = c("eDWID", "reviewyear"))
-df <- left_join(df, demo, by = "eDWID")
-# Group years by 2003-2009, 2010-2013, and 2014-2018
-df$Period <- cut(df$reviewyear,
-                 breaks = c(2002, 2009, 2013, 2018),
-                 labels = c("2003 - 2009", "2010 - 2013", "2014 - 2018")
-)
-df$smoking <- factor(df$smoking, levels = 1:4, labels = c(
-  "No",
-  "Occasionally",
-  "Yes, Regularly, less than 1 ppd",
-  "Yes, Regularly, 1 ppd or more"
-), ordered = T)
-df$second_smoke <- factor(df$second_smoke, levels = c(1:4), labels = c(
-  "No",
-  "Occasionally",
-  "Yes, Regularly, less than 1 ppd",
-  "Yes, Regularly, 1 ppd or more"
-), ordered = T)
-# Factors
-ed_levels <- c(
-  "Less than High School",
-  "High School diploma or equivalent",
-  "Some College",
-  "College Graduate",
-  "Masters/Doctoral level degree"
-)
-df$patient_education <- factor(df$patient_education,
-                               levels = c(1:5),
-                               labels = ed_levels
-)
-df$father_education <- factor(df$father_education,
-                              levels = c(1:5),
-                              labels = ed_levels
-)
-df$mother_education <- factor(df$mother_education,
-                              levels = c(1:5),
-                              labels = ed_levels
-)
-df$spouse_education <- factor(df$spouse_education,
-                              levels = c(1:5),
-                              labels = ed_levels
-)
-df$oxygen_therapy <- factor(df$oxygen_therapy,
-                            levels = 1:6,
-                            labels = c(
-                              "Yes, Continuously",
-                              "Yes, Nocturnal and/or with exertion",
-                              "Yes, During exacerbation",
-                              "Yes, prn",
-                              "No",
-                              "Unknown"
-                            )
-)
-# Genotypes
-df$Mutation1 <- factor(df$Mutation1,
-                       levels = geno[geno$V1 == "mutation1", "V2"],
-                       labels = geno[geno$V1 == "mutation1", "V3"]
-)
-df$Mutation2 <- factor(df$Mutation2,
-                       levels = geno[geno$V1 == "mutation2", "V2"],
-                       labels = geno[geno$V1 == "mutation2", "V3"]
-)
-df$F508del <- (df$Mutation1 == "F508del") + (df$Mutation2 == "F508del")
-df$F508del <- factor(df$F508del, levels = 0:2, 
-                     labels = c("None", "Heterozygous", "Homozygous"))
-# Percentiles for peds
-df$bmi_peds = df$A_bmipercentile * (df$Age_YrEnd < 18)
-df$bmi_adults = df$A_bmivalue * (df$Age_YrEnd >= 18)
-df$height_peds = df$A_heightpercentile * (df$Age_YrEnd < 18)
-df$height_adults = df$A_height * (df$Age_YrEnd >= 18)
-df$weight_peds = df$A_weightpercentile * (df$Age_YrEnd < 18)
-df$weight_adults = df$A_weight * (df$Age_YrEnd >= 18)
-# Yes/no variables
-yn <- c("A_IsOnEnzymes", "A_salt_supplement", "A_cf_vitamins", "A_dornasealfa", 
-        "A_pulmonarycomplications4", "A_pulmonarycomplications5", 
-        "A_othercomplications9", paste0("A_hepatobiliary1_",1:3), 
-        paste0("A_hepatobiliary2_",1:4),paste0("employment", 1:7),
-        "pregnant","dexa_results_osteoporosis","A_Vx770","A_VX809comb",
-        "A_feeding3","A_corticosteroids1","A_pulmonarycomplications2",
-        paste0("A_mycobacterials",c(1:5,7:10)),
-        "A_pseudomonasaeruginosa","A_staphylococcus_aureus",
-        "A_othermicroorganisms1","A_othermicroorganisms2",
-        "A_fungalyeast1","A_fungalyeast2","A_fungalyeast3")
-df[, yn] <- lapply(df[, yn], function(c) {
-  factor(c, levels = 0:1, labels = c("No", "Yes"), ordered = T)
-})
+
 ```
-
-# Table 1: Participant characteristics at last visit in time period
-
-```{r results='asis'}
-# Summarize by ID and period
-t1_df <- df %>%
-  group_by(eDWID, Period) %>%
-  summarise(
-    Age = max(Age_YrEnd), Race = last(na.omit(Race)),
-    Hispanic = last(na.omit(Hispanicrace)), Gender = last(na.omit(Gender)),
-    `Patient Education` = last(na.omit(patient_education)),
-    `Father Education` = last(na.omit(father_education)),
-    `Mother Education` = last(na.omit(mother_education)),
-    `Spouse Education` = last(na.omit(spouse_education)),
-    `Marital Status` = last(na.omit(marital_status)),
-    `Part Time` = max(employment1, na.rm = F),
-    `Full time homemaker` = max(employment2, na.rm = F),
-    `Full time employment` = max(employment3, na.rm = F),
-    `Unemployed` = max(employment4, na.rm = F),
-    `Student` = max(employment5, na.rm = F),
-    `Disabled` = max(employment6, na.rm = F),
-    `Retired` = max(employment7, na.rm = F),
-    Pregnancy = max(pregnant, na.rm = T),
-    .groups = "keep"
-  )
-# Print table
-cols <- colnames(t1_df)[3:ncol(t1_df)]
-t1 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
-              data = t1_df
-)
-summary(t1)
-```
-
-# Table 2: Pulmonary function
-
-Continuous and count variables are from the last year of each period.
-
-```{r results='asis'}
-# Summarize by ID and period
-t2_df <- df %>%
-  group_by(eDWID, Period) %>%
-  summarise(
-    Smoking = max(smoking, na.rm = T),
-    `Secondhand smoke` = max(second_smoke, na.rm = T),
-    `FEV1 % Predicted` = last(na.omit(GLI_FEV1_pct_predicted)),
-    `FVC % Predicted` = last(na.omit(GLI_FVC_pct_predicted)),
-    `Pulm. Exacerbations` = sum(NumPulmExacerbation,na.rm = T),
-    `Beta agonist` = max(bronchodilators1, na.rm = T),
-    `Theophylline product` = max(bronchodilators2, na.rm = T),
-    `Short acting beta agonist` = max(bronchodilatorsinhaled1, na.rm = T),
-    `Long acting beta agonist` = max(bronchodilatorsinhaled2, na.rm = T),
-    `Short acting anticholinergic` = max(bronchodilatorsinhaled3, na.rm = T),
-    `Long acting anticholinergic` = max(bronchodilatorsinhaled4, na.rm = T),
-    `Combination beta agonist and anticholinergic` = max(bronchodilatorsinhaled5, na.rm = T),
-    `Positive Expiratory Pressure (PEP)` = max(act_primary1, na.rm = T),
-    `Postural drainage with clapping (CPT)` = max(act_primary2, na.rm = T),
-    `Forced expiratory techniques` = max(act_primary3, na.rm = T),
-    `Oscillating PEP` = max(act_primary4, na.rm = T),
-    `High frequency chest wall oscillation` = max(act_primary5, na.rm = T),
-    `Exercise` = max(act_primary6, na.rm = T),
-    `None` = max(act_primary7, na.rm = T),
-    `Other` = max(act_primary8, na.rm = T),
-    `Oxygen Therapy` = last(na.omit(oxygen_therapy)),
-    .groups = "keep"
-  )
-# Print table
-cols <- colnames(t2_df)[3:ncol(t2_df)]
-t2 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
-              data = t2_df
-)
-summary(t2)
-```
-
-# Table 3: Panreactic function
-
-Continuous and count variables are from the last year of each period.
-
-```{r results='asis'}
-# Summarize by ID and period
-t3_df <- df %>%
-  group_by(eDWID, Period) %>%
-  summarise(
-    `On Enzymes` = last(na.omit(A_IsOnEnzymes)),
-    `Mutation Class` = last(na.omit(MutClass)),
-    `F508del` = last(na.omit(F508del)),
-    .groups = "keep"
-  )
-# Print table
-cols <- colnames(t3_df)[3:ncol(t3_df)]
-t3 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
-              data = t3_df
-)
-summary(t3)
-```
-
-# Table 4: Nutrition
-
-Continuous and count variables are from the last year of each period.
-
-```{r results='asis'}
-# Summarize by ID and period
-t4_df <- df %>%
-  group_by(eDWID, Period) %>%
-  summarise(
-    `BMI %ile (age <18)` = last(na.omit(bmi_peds)),
-    `BMI Value (age <= 18)` = last(na.omit(bmi_adults)),
-    `Salt Supplement` = max(A_salt_supplement, na.rm = T),
-    `CF Vitamins` = max(A_cf_vitamins, na.rm = T),
-    `Hemoptysis` = max(A_pulmonarycomplications4, na.rm = T),
-    `Pneumothorax Chest Tube` = max(A_pulmonarycomplications5, na.rm = T),
-    `Sinus Disease` = max(A_othercomplications9, na.rm = T),
-    `Cirrhosis` = max(A_hepatobiliary1_3, na.rm = T),
-    .groups = "keep"
-  )
-# Print table
-cols <- colnames(t4_df)[3:ncol(t4_df)]
-t4 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
-              data = t4_df
-)
-summary(t4)
-```
-
-# Table 5: Other
-
-Continuous and count variables are from the last year of each period.
-
-```{r results='asis'}
-# Summarize by ID and period
-t5_df <- df %>%
-  group_by(eDWID, Period) %>%
-  summarise(
-    `Osteoperosis`=max(dexa_results_osteoporosis,na.rm = T),
-    `Kalydeco`=max(A_Vx770,na.rm = T),
-    `Lumacaftor/Ivacaftor`=max(A_VX809comb,na.rm = T),
-    `Height %ile (age <18)` = last(na.omit(height_peds)),
-    `Height Value (age <= 18)` = last(na.omit(height_adults)),
-    `Weight %ile (age <18)` = last(na.omit(weight_peds)),
-    `Weight Value (age <= 18)` = last(na.omit(weight_adults)),
-    `Num. Hospitalizations` = sum(NumHosp,na.rm = T),
-    `Gall stones`=max(hepatobiliary1_1,na.rm=T),
-    `Gall stones, requiring surgery/procedure`=max(hepatobiliary1_2,na.rm=T),
-    `Liver disease, cirrhosis`=max(hepatobiliary1_3,na.rm=T),
-    `Liver disease, non- cirrhosis`=max(hepatobiliary2_1,na.rm=T),
-    `Hepatic Steatosis`=max(hepatobiliary2_2,na.rm=T),
-    `Liver disease, other`=max(hepatobiliary2_3,na.rm=T),
-    `Acute Liver Failure`=max(hepatobiliary2_4,na.rm=T),
-    `G-Tube`=max(A_feeding3,na.rm = T),
-    `Oral corticosteroids`=max(A_corticosteroids1,na.rm = T),
-    `Asthma`=max(A_pulmonarycomplications2,na.rm = T),
-    `Mycobacterial tuberculosis`=max(A_mycobacterials1,na.rm=T),
-    `Mycobacterium abscessus/chelonae`=max(A_mycobacterials2,na.rm=T),
-    `Mycobacterium avium complex (MAC)`=max(A_mycobacterials3,na.rm=T),
-    `Mycobacterium fortuitum group`=max(A_mycobacterials4,na.rm=T),
-    `Mycobacterium gordonae`=max(A_mycobacterials5,na.rm=T),
-    `Mycobacterium kansasii`=max(A_mycobacterials7,na.rm=T),
-    `Mycobacterium marinum`=max(A_mycobacterials8,na.rm=T),
-    `Mycobacterium terrae`=max(A_mycobacterials9,na.rm=T),
-    `Other Mycobacterium`=max(A_mycobacterials10,na.rm=T),
-    .groups = "keep"
-  )
-# Print table
-cols <- colnames(t5_df)[3:ncol(t5_df)]
-t5 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
-              data = t5_df
-)
-summary(t5)
-```
-
-# Table 6: Microbiome
-
-```{r results='asis'}
-# Summarize by ID and period
-t6_df <- df %>%
-  group_by(eDWID, Period) %>%
-  summarise(
-    `Pseudomonas`=max(A_pseudomonasaeruginosa,na.rm=T),
-    `Burkholderia`=max(A_staphylococcus_aureus,na.rm=T),
-    `Alcaligenes`=max(A_othermicroorganisms1,na.rm=T),
-    `Stenotrophomonas`=max(A_othermicroorganisms2,na.rm=T),
-    `Aspergillus`=max(A_fungalyeast1,na.rm=T),
-    `Candida`=max(A_fungalyeast2,na.rm=T),
-    `Scedosporium`=max(A_fungalyeast3,na.rm=T),
-    .groups = "keep"
-  )
-# Print table
-cols <- colnames(t6_df)[3:ncol(t6_df)]
-t6 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
-              data = t6_df
-)
-summary(t6)
-```
-
-# Questions/notes
-
-1. Which variables do we include for microbiome stuff? Include "othermicroorganismstypes"?
-2. We don't seem to have ETI data.

### 2022-12-15 Christine Chan/CF Registry/cf_registry_analysis.qmd
switching to .qmd file

---
title: "CF Registry Analysis"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
editor: visual
---

```{r}
library(table1)
```

### 2022-12-16 Christine Chan/CF Registry/cf_registry_analysis.qmd
CFRD incidence

@@
 ```{r}
+#| include: false
+library(tidyverse)
 library(table1)
+# All data
+df = read.csv("/home/timvigers/Work/CHCO/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",na.strings = c("","N/A"))
+df = df %>% arrange(eDWID,reviewyear,encounternum)
+# For those with CFRD by OGTT, find first visit with diagnosis
+first_cfrd = df %>% group_by(eDWID) %>% 
+  summarise(first_cfrd = which(twohour_bloodglucose >= 200),.groups = "drop")
+df = left_join(df,first_cfrd,by = "eDWID")
+df = df %>% group_by(eDWID) %>% 
+  mutate(cfrd_ogtt = case_when(
+    row_number() < first_cfrd ~ "No",
+    row_number() >= first_cfrd ~ "Yes"
+  )) %>%
+  ungroup()
+df$cfrd_ogtt[is.na(df$cfrd_ogtt)] = "No"
+df$cfrd_ogtt = factor(df$cfrd_ogtt,levels = c("No","Yes"),
+                      labels = c("No","Yes"),ordered = TRUE)
+# Testing
+t = df %>% select(eDWID,reviewyear,)
 ```
+
+```{r}
+#| @tbl-incidence
+t = df %>% 
+  group_by(eDWID,reviewyear) %>% 
+  summarise(cfrd_ogtt = max(cfrd_ogtt),.groups = "drop") %>% 
+  group_by(reviewyear) %>%
+  summarise(cfrd_rate = mean(as.numeric(cfrd_ogtt)-1))
+```

### 2022-12-20 Christine Chan/CF Registry/cf_registry_analysis.qmd
i think I have year on year incidence figured out?

--- 2022-12-15 Christine Chan/CF Registry/cf_registry_analysis.qmd
@@
 ```{r}
+#| include: false
+library(tidyverse)
 library(table1)
+# All data
+df = read.csv("/home/timvigers/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",na.strings = c("","N/A"))
+df = df %>% arrange(eDWID,reviewyear,encounternum)
+# For those with CFRD by OGTT, find first visit with diagnosis
+first_cfrd = df %>% group_by(eDWID) %>% 
+  summarise(first_cfrd = which(twohour_bloodglucose >= 200),.groups = "drop")
+df = left_join(df,first_cfrd,by = "eDWID")
+df = df %>% group_by(eDWID) %>% 
+  mutate(cfrd_ogtt = case_when(
+    row_number() < first_cfrd ~ "No",
+    row_number() >= first_cfrd ~ "Yes"
+  )) %>%
+  ungroup()
+df$cfrd_ogtt[is.na(df$cfrd_ogtt)] = "No"
+df$cfrd_ogtt = factor(df$cfrd_ogtt,levels = c("No","Yes"),
+                      labels = c("No","Yes"),ordered = TRUE)
+diagnosis = df %>% 
+  group_by(eDWID,reviewyear) %>% 
+  summarise(cfrd_ogtt = max(cfrd_ogtt),.groups = "drop")
 ```
+
+```{python}
+#| @tbl-incidence
+diagnosis = r.diagnosis
+years = diagnosis["reviewyear"].unique()
+cases = []
+at_risk = []
+incidence = []
+# Loop through years
+for y in years:
+  d = diagnosis[diagnosis["reviewyear"]==y]
+  this_year = d.loc[d["cfrd_ogtt"]=="Yes","eDWID"].to_list()
+  cases = set(cases).union(this_year)
+  new_cases = set(cases).difference(set(this_year))
+  at_risk = set(d["eDWID"].to_list()).difference(at_risk)
+  incidence.append(round(len(new_cases)/len(at_risk),3))
+```

### 2022-12-21 Christine Chan/CF Registry/cf_registry_analysis.qmd
Update cf_registry_analysis.qmd

--- 2022-12-15 Christine Chan/CF Registry/cf_registry_analysis.qmd
@@
 ```{r}
+#| include: false
+library(tidyverse)
 library(table1)
+# All data
+df = read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
+              na.strings = c("","N/A"))
+df = df %>% arrange(eDWID,ReviewYear) 
+# For those with CFRD by OGTT, find first visit with diagnosis
+first_cfrd = df %>% filter(A_twohour_bloodglucose >= 200) %>% 
+  group_by(eDWID) %>% slice_min(ReviewYear) %>% 
+  select(eDWID,ReviewYear) %>% rename(first_cfrd = ReviewYear)
+df = left_join(df,first_cfrd,by = "eDWID")
+df = df %>% group_by(eDWID) %>% 
+  mutate(cfrd_ogtt = case_when(
+    ReviewYear < first_cfrd ~ "No",
+    ReviewYear >= first_cfrd ~ "Yes"
+  )) %>%
+  ungroup()
+df$cfrd_ogtt[is.na(df$cfrd_ogtt)] = "No"
+diagnosis = df %>% 
+  select(eDWID,ReviewYear,Age_YrEnd,first_cfrd,cfrd_ogtt) %>%
+  group_by(eDWID) %>%
+  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd))
 ```
+
+```{python}
+#| @tbl-incidence
+import pandas as pd
+import numpy as np
+
+diagnosis = r.diagnosis
+years = diagnosis["ReviewYear"].unique()
+# Loop through years - all ages
+cases = []
+at_risk = []
+incidence = []
+for y in years:
+  d = diagnosis[diagnosis["ReviewYear"]==y]
+  this_year = d.loc[d["cfrd_ogtt"]=="Yes","eDWID"].to_list()
+  cases = set(cases).union(this_year)
+  new_cases = set(cases).difference(set(this_year))
+  at_risk = set(d["eDWID"].to_list()).difference(at_risk)
+  incidence.append(round(len(new_cases)/(len(at_risk)/1000),3))
+# Loop through years - age < 10
+cases = []
+at_risk = []
+incidence_10 = []
+for y in years:
+  d = diagnosis[(diagnosis["ReviewYear"]==y) & (diagnosis["Age in 2018"]<10)]
+  this_year = d.loc[d["cfrd_ogtt"]=="Yes","eDWID"].to_list()
+  cases = set(cases).union(this_year)
+  new_cases = set(cases).difference(set(this_year))
+  at_risk = set(d["eDWID"].to_list()).difference(at_risk)
+  if len(at_risk)==0:
+    incidence_10.append(np.nan)
+  else:
+    incidence_10.append(round(len(new_cases)/(len(at_risk)/1000),3))
+# Loop through years - age 10 - 18
+cases = []
+at_risk = []
+incidence_10_18 = []
+for y in years:
+  d = diagnosis[(diagnosis["ReviewYear"]==y) & (diagnosis["Age in 2018"]>=10) & (diagnosis["Age in 2018"]<=18)]
+  this_year = d.loc[d["cfrd_ogtt"]=="Yes","eDWID"].to_list()
+  cases = set(cases).union(this_year)
+  new_cases = set(cases).difference(set(this_year))
+  at_risk = set(d["eDWID"].to_list()).difference(at_risk)
+  if len(at_risk)==0:
+    incidence_10_18.append(np.nan)
+  else:
+    incidence_10_18.append(round(len(new_cases)/(len(at_risk)/1000),3))
+# Loop through years - age 19 - 30
+cases = []
+at_risk = []
+incidence_19_30 = []
+for y in years:
+  d = diagnosis[(diagnosis["ReviewYear"]==y) & (diagnosis["Age in 2018"]>=19) & (diagnosis["Age in 2018"]<=30)]
+  this_year = d.loc[d["cfrd_ogtt"]=="Yes","eDWID"].to_list()
+  cases = set(cases).union(this_year)
+  new_cases = set(cases).difference(set(this_year))
+  at_risk = set(d["eDWID"].to_list()).difference(at_risk)
+  if len(at_risk)==0:
+    incidence_19_30.append(np.nan)
+  else:
+    incidence_19_30.append(round(len(new_cases)/(len(at_risk)/1000),3))
+```

### 2022-12-21 Christine Chan/CF Registry/cf_registry_analysis.qmd
python section not necessary

@@
 #| include: false
+library(knitr)
 library(tidyverse)
@@
               na.strings = c("","N/A"))
-df = df %>% arrange(eDWID,ReviewYear) 
+df = df %>% arrange(eDWID,ReviewYear) %>% 
+  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd))
 # For those with CFRD by OGTT, find first visit with diagnosis
@@
   group_by(eDWID) %>% slice_min(ReviewYear) %>% 
-  select(eDWID,ReviewYear) %>% rename(first_cfrd = ReviewYear)
-df = left_join(df,first_cfrd,by = "eDWID")
+  select(eDWID,ReviewYear,`Age in 2018`) %>% rename(first_cfrd = ReviewYear)
+df = left_join(df,first_cfrd,by = c("eDWID","Age in 2018"))
 df = df %>% group_by(eDWID) %>% 
@@
 df$cfrd_ogtt[is.na(df$cfrd_ogtt)] = "No"
-diagnosis = df %>% 
-  select(eDWID,ReviewYear,Age_YrEnd,first_cfrd,cfrd_ogtt) %>%
-  group_by(eDWID) %>%
-  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd))
+df = df %>% 
+  select(eDWID,ReviewYear,`Age in 2018`,first_cfrd,cfrd_ogtt)
+ages = df %>% group_by(eDWID) %>% 
+  summarise(`Age in 2018` = names(sort(table(`Age in 2018`), decreasing = TRUE)[1]))
+ages$`Age in 2018` = round(as.numeric(ages$`Age in 2018`),1)
+```
+
+```{r}
+# Person-years variable
+py=1000
+# For each year, get the number of people diagnosed with CFRD that year
+years = 2003:2018
+cases = lapply(years, function(y){
+  total = unique(first_cfrd$eDWID[first_cfrd$first_cfrd==y])
+  cases_age_10 = total[total %in% ages$eDWID[ages$`Age in 2018` < 10]]
+  cases_age_10_18 = total[total %in% ages$eDWID[ages$`Age in 2018` >= 10 & ages$`Age in 2018` < 19]]
+  cases_age_19_30 = total[total %in% ages$eDWID[ages$`Age in 2018` >= 19 & ages$`Age in 2018` < 31]]
+  cases_age_30 = total[total %in% ages$eDWID[ages$`Age in 2018` >= 31]]
+  return(list("total"=total,"cases_age_10"=cases_age_10,
+              "cases_age_10_18"=cases_age_10_18,
+              "cases_age_19_30"=cases_age_19_30,
+              "cases_age_30"=cases_age_30))
+})
+# Double check that there is no overlap between sets (there isn't!)
+# lapply(1:length(cases), function(y){intersect(cases[[y]],unlist(cases[-y]))})
+# For each year, get the number of people at risk
+at_risk = lapply(years, function(y){
+  ar = unique(df$eDWID[df$ReviewYear==y])
+  ar = ar[!ar %in% unlist(cases[which(years<=y)]$total)]
+  ar_age_10 = intersect(df$eDWID[df$ReviewYear==y],ages$eDWID[ages$`Age in 2018`<10])
+  ar_age_10_18 = intersect(df$eDWID[df$ReviewYear==y],
+                           ages$eDWID[ages$`Age in 2018`>=10 & ages$`Age in 2018`<19])
+  ar_age_19_30 = intersect(df$eDWID[df$ReviewYear==y],
+                           ages$eDWID[ages$`Age in 2018`>=19 & ages$`Age in 2018`<31])
+  ar_age_30 = intersect(df$eDWID[df$ReviewYear==y],ages$eDWID[ages$`Age in 2018`>=31])
+  return(list("total"=ar,"ar_age_10"=ar_age_10,"ar_age_10_18"=ar_age_10_18,
+              "ar_age_19_30"=ar_age_19_30,"ar_age_30"=ar_age_30))
+})
+# Put everything in one table
+inc_df = lapply(years, function(y){
+  # Total
+  new_cases = length(cases[[which(years==y)]]$total)
+  total_at_risk = length(at_risk[[which(years==y)]]$total)
+  total_inc = new_cases/(total_at_risk/py)
+  # < 10
+  new_cases_10 = length(cases[[which(years==y)]]$cases_age_10)
+  at_risk_10 = length(at_risk[[which(years==y)]]$ar_age_10)
+  inc_10 = new_cases_10/(at_risk_10/py)
+  # 10 - 18
+  new_cases_10_18 = length(cases[[which(years==y)]]$cases_age_10_18)
+  at_risk_10_18 = length(at_risk[[which(years==y)]]$ar_age_10_18)
+  inc_10_18 = new_cases_10_18/(at_risk_10_18/py)
+  # 19 - 30
+  new_cases_19_30 = length(cases[[which(years==y)]]$cases_age_19_30)
+  at_risk_19_30 = length(at_risk[[which(years==y)]]$ar_age_19_30)
+  inc_19_30 = new_cases_19_30/(at_risk_19_30/py)
+  # > 30
+  new_cases_30 = length(cases[[which(years==y)]]$cases_age_30)
+  at_risk_30 = length(at_risk[[which(years==y)]]$ar_age_30)
+  inc_30 = new_cases_30/(at_risk_30/py)
+  return(list("Year" = y,
+              "Total New Cases" = new_cases,
+              "Total At Risk" = total_at_risk,
+              "Total Incidence" = total_inc,
+              "Total New Cases Age < 10" = new_cases_10,
+              "Total At Risk Age < 10" = at_risk_10,
+              "Total Incidence Age < 10" = inc_10,
+              "Total New Cases Age 10-18" = new_cases_10_18,
+              "Total At Risk Age 10-18" = at_risk_10_18,
+              "Total Incidence Age 10-18" = inc_10_18,
+              "Total New Cases Age 19-30" = new_cases_19_30,
+              "Total At Risk Age 19-30" = at_risk_19_30,
+              "Total Incidence Age 19-30" = inc_19_30,
+              "Total New Cases Age > 30" = new_cases_30,
+              "Total At Risk Age > 30" = at_risk_30,
+              "Total Incidence Age > 30" = inc_30))
+})
+inc_df = do.call(rbind,inc_df)
+kable(inc_df,digits=3)
 ```

### 2022-12-22 Christine Chan/CF Registry/cf_registry_analysis.qmd
Update cf_registry_analysis.qmd

--- 2022-12-20 Christine Chan/CF Registry/cf_registry_analysis.qmd
@@
     self-contained: true
+    page-layout: full
 editor: visual
@@
 #| include: false
+library(knitr)
 library(tidyverse)
 library(table1)
+library(pROC)
 # All data
-df = read.csv("/home/timvigers/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",na.strings = c("","N/A"))
-df = df %>% arrange(eDWID,reviewyear,encounternum)
+df = read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
+              na.strings = c("","N/A"))
+df = df %>% arrange(eDWID,ReviewYear) %>% 
+  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd))
 # For those with CFRD by OGTT, find first visit with diagnosis
-first_cfrd = df %>% group_by(eDWID) %>% 
-  summarise(first_cfrd = which(twohour_bloodglucose >= 200),.groups = "drop")
-df = left_join(df,first_cfrd,by = "eDWID")
+first_cfrd = df %>% filter(A_twohour_bloodglucose >= 200) %>% 
+  group_by(eDWID) %>% slice_min(ReviewYear) %>% 
+  select(eDWID,ReviewYear,`Age in 2018`) %>% rename(first_cfrd = ReviewYear)
+df = left_join(df,first_cfrd,by = c("eDWID","Age in 2018"))
 df = df %>% group_by(eDWID) %>% 
   mutate(cfrd_ogtt = case_when(
-    row_number() < first_cfrd ~ "No",
-    row_number() >= first_cfrd ~ "Yes"
+    ReviewYear < first_cfrd ~ "No",
+    ReviewYear >= first_cfrd ~ "Yes"
   )) %>%
@@
 df$cfrd_ogtt[is.na(df$cfrd_ogtt)] = "No"
-df$cfrd_ogtt = factor(df$cfrd_ogtt,levels = c("No","Yes"),
-                      labels = c("No","Yes"),ordered = TRUE)
-diagnosis = df %>% 
-  group_by(eDWID,reviewyear) %>% 
-  summarise(cfrd_ogtt = max(cfrd_ogtt),.groups = "drop")
+df = df %>% 
+  select(eDWID,ReviewYear,`Age in 2018`,first_cfrd,cfrd_ogtt,A_hgba1c)
+ages = df %>% group_by(eDWID) %>% 
+  summarise(`Age in 2018` = names(sort(table(`Age in 2018`), decreasing = TRUE)[1]))
+ages$`Age in 2018` = round(as.numeric(ages$`Age in 2018`),1)
 ```
 
-```{python}
-#| @tbl-incidence
-diagnosis = r.diagnosis
-years = diagnosis["reviewyear"].unique()
-cases = []
-at_risk = []
-incidence = []
-# Loop through years
-for y in years:
-  d = diagnosis[diagnosis["reviewyear"]==y]
-  this_year = d.loc[d["cfrd_ogtt"]=="Yes","eDWID"].to_list()
-  cases = set(cases).union(this_year)
-  new_cases = set(cases).difference(set(this_year))
-  at_risk = set(d["eDWID"].to_list()).difference(at_risk)
-  incidence.append(round(len(new_cases)/len(at_risk),3))
+```{r}
+#| label: tbl-incidence
+#| tbl-cap: "Year-on-Year Incidence of CFRD"
+# Person-years variable
+py=1000
+# For each year, get the number of people diagnosed with CFRD that year
+years = 2003:2018
+cases = lapply(years, function(y){
+  total = unique(first_cfrd$eDWID[first_cfrd$first_cfrd==y])
+  cases_age_10 = total[total %in% ages$eDWID[ages$`Age in 2018` < 10]]
+  cases_age_10_18 = total[total %in% ages$eDWID[ages$`Age in 2018` >= 10 & ages$`Age in 2018` < 19]]
+  cases_age_19_30 = total[total %in% ages$eDWID[ages$`Age in 2018` >= 19 & ages$`Age in 2018` < 31]]
+  cases_age_30 = total[total %in% ages$eDWID[ages$`Age in 2018` >= 31]]
+  return(list("total"=total,"cases_age_10"=cases_age_10,
+              "cases_age_10_18"=cases_age_10_18,
+              "cases_age_19_30"=cases_age_19_30,
+              "cases_age_30"=cases_age_30))
+})
+# Double check that there is no overlap between sets (there isn't!)
+# lapply(1:length(cases), function(y){intersect(cases[[y]],unlist(cases[-y]))})
+# For each year, get the number of people at risk
+at_risk = lapply(years, function(y){
+  ar = unique(df$eDWID[df$ReviewYear==y])
+  ar = ar[!ar %in% unlist(cases[which(years<=y)]$total)]
+  ar_age_10 = intersect(df$eDWID[df$ReviewYear==y],ages$eDWID[ages$`Age in 2018`<10])
+  ar_age_10_18 = intersect(df$eDWID[df$ReviewYear==y],
+                           ages$eDWID[ages$`Age in 2018`>=10 & ages$`Age in 2018`<19])
+  ar_age_19_30 = intersect(df$eDWID[df$ReviewYear==y],
+                           ages$eDWID[ages$`Age in 2018`>=19 & ages$`Age in 2018`<31])
+  ar_age_30 = intersect(df$eDWID[df$ReviewYear==y],ages$eDWID[ages$`Age in 2018`>=31])
+  return(list("total"=ar,"ar_age_10"=ar_age_10,"ar_age_10_18"=ar_age_10_18,
+              "ar_age_19_30"=ar_age_19_30,"ar_age_30"=ar_age_30))
+})
+# Put everything in one table
+inc_df = lapply(years, function(y){
+  # Total
+  new_cases = length(cases[[which(years==y)]]$total)
+  total_at_risk = length(at_risk[[which(years==y)]]$total)
+  total_inc = new_cases/(total_at_risk/py)
+  # < 10
+  new_cases_10 = length(cases[[which(years==y)]]$cases_age_10)
+  at_risk_10 = length(at_risk[[which(years==y)]]$ar_age_10)
+  inc_10 = new_cases_10/(at_risk_10/py)
+  # 10 - 18
+  new_cases_10_18 = length(cases[[which(years==y)]]$cases_age_10_18)
+  at_risk_10_18 = length(at_risk[[which(years==y)]]$ar_age_10_18)
+  inc_10_18 = new_cases_10_18/(at_risk_10_18/py)
+  # 19 - 30
+  new_cases_19_30 = length(cases[[which(years==y)]]$cases_age_19_30)
+  at_risk_19_30 = length(at_risk[[which(years==y)]]$ar_age_19_30)
+  inc_19_30 = new_cases_19_30/(at_risk_19_30/py)
+  # > 30
+  new_cases_30 = length(cases[[which(years==y)]]$cases_age_30)
+  at_risk_30 = length(at_risk[[which(years==y)]]$ar_age_30)
+  inc_30 = new_cases_30/(at_risk_30/py)
+  return(list(`Year` = y,
+              `Total New Cases` = new_cases,
+              `Total At Risk` = total_at_risk,
+              `Total Incidence` = total_inc,
+              `Total New Cases Age < 10` = new_cases_10,
+              `Total At Risk Age < 10` = at_risk_10,
+              `Total Incidence Age < 10` = inc_10,
+              `Total New Cases Age 10-18` = new_cases_10_18,
+              `Total At Risk Age 10-18` = at_risk_10_18,
+              `Total Incidence Age 10-18` = inc_10_18,
+              `Total New Cases Age 19-30` = new_cases_19_30,
+              `Total At Risk Age 19-30` = at_risk_19_30,
+              `Total Incidence Age 19-30` = inc_19_30,
+              `Total New Cases Age > 30` = new_cases_30,
+              `Total At Risk Age > 30` = at_risk_30,
+              `Total Incidence Age > 30` = inc_30))
+})
+inc_df = as.data.frame(do.call(rbind,inc_df))
+inc_df = data.frame(lapply(inc_df,as.numeric),check.names = F)
+kable(inc_df,digits=3)
 ```
+
+In @tbl-incidence, incidence per year is calculated as the number of new CFRD diagnoses per 1,000 person-years of those at risk (excluding previously identified cases).
+
+```{r}
+df$cfrd_ogtt = factor(df$cfrd_ogtt,levels = c("No","Yes"),ordered = T)
+cfrd = df %>% group_by(eDWID) %>% 
+  mutate(cfrd_ever = max(cfrd_ogtt),
+         time_from_diag = ReviewYear-first_cfrd) %>%
+  filter(cfrd_ever=="Yes",time_from_diag <= 0,time_from_diag >= -1) %>%
+  summarise(hba1c = last(na.omit(round(A_hgba1c,1)))) %>% drop_na()
+non_cfrd =  df %>% group_by(eDWID) %>% 
+  mutate(cfrd_ever = max(cfrd_ogtt)) %>% filter(cfrd_ever=="No") %>%
+  summarise(hba1c = last(na.omit(round(A_hgba1c,1)))) %>% drop_na()
+```
+
+```{r}
+#| label: fig-roc
+#| fig-cap: "ROC Curve"
+r = roc(cases = cfrd$hba1c,controls=non_cfrd$hba1c,quiet = T,plot=T,auc=T, ci=T,
+        print.auc=TRUE, show.thres=TRUE)
+```
+
+Because we do not have exact information regarding HbA1c timing (we only have encounter number within year), the annualized HbA1c from the year of diagnosis or from one year prior to diagnosis was used in the ROC analysis.
+For those with HbA1c measures the year of diagnosis and one year prior, the value from year of diagnosis was preferred.
+For those never diagnosed with CFRD, the most recent HbA1c was used.
+There were a total of `r nrow(cfrd)` HbA1c measures for people with CFRD and `r nrow(non_cfrd)` measures for those without.
+
+The AUC for this ROC curve was 0.61 (see @fig-roc).
+
+```{r}
+#| label: tbl-roc_thresh
+#| tbl-cap: "Best ROC Threshold (Youden Index)"
+kable(coords(r,"best"),digits = 3)
+```

### 2022-12-23 Quarto Template.qmd
Update Quarto Template.qmd

---
title: "Title"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: visual
---

# Table

```{r}
#| label: tbl-planets
#| tbl-cap: Planets

```

# figure

```{r}
#| label: fig-planets
#| fig-cap: Planets

```

### 2022-12-23 Christine Chan/CF Registry/cf_registry_analysis.qmd
Update cf_registry_analysis.qmd
formatting

--- 2022-12-21 Christine Chan/CF Registry/cf_registry_analysis.qmd
@@
 #| include: false
+library(knitr)
 library(tidyverse)
 library(table1)
+library(pROC)
+```
+
+```{r}
 # All data
-df = read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
-              na.strings = c("","N/A"))
-df = df %>% arrange(eDWID,ReviewYear) 
+df <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
+  na.strings = c("", "N/A")
+)
+df <- df %>%
+  arrange(eDWID, ReviewYear) %>%
+  mutate(`Age in 2018` = 2018 - (ReviewYear - Age_YrEnd))
 # For those with CFRD by OGTT, find first visit with diagnosis
-first_cfrd = df %>% filter(A_twohour_bloodglucose >= 200) %>% 
-  group_by(eDWID) %>% slice_min(ReviewYear) %>% 
-  select(eDWID,ReviewYear) %>% rename(first_cfrd = ReviewYear)
-df = left_join(df,first_cfrd,by = "eDWID")
-df = df %>% group_by(eDWID) %>% 
+first_cfrd <- df %>%
+  filter(A_twohour_bloodglucose >= 200) %>%
+  group_by(eDWID) %>%
+  slice_min(ReviewYear) %>%
+  select(eDWID, ReviewYear, `Age in 2018`) %>%
+  rename(first_cfrd = ReviewYear)
+df <- left_join(df, first_cfrd, by = c("eDWID", "Age in 2018"))
+df <- df %>%
+  group_by(eDWID) %>%
   mutate(cfrd_ogtt = case_when(
@@
   ungroup()
-df$cfrd_ogtt[is.na(df$cfrd_ogtt)] = "No"
-diagnosis = df %>% 
-  select(eDWID,ReviewYear,Age_YrEnd,first_cfrd,cfrd_ogtt) %>%
+df$cfrd_ogtt[is.na(df$cfrd_ogtt)] <- "No"
+df <- df %>%
+  select(eDWID, ReviewYear, `Age in 2018`, first_cfrd, cfrd_ogtt, A_hgba1c)
+ages <- df %>%
   group_by(eDWID) %>%
-  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd))
+  summarise(`Age in 2018` = names(sort(table(`Age in 2018`), 
+                                       decreasing = TRUE)[1]))
+ages$`Age in 2018` <- round(as.numeric(ages$`Age in 2018`), 1)
 ```
 
-```{python}
-#| @tbl-incidence
-import pandas as pd
-import numpy as np
+```{r}
+#| label: tbl-incidence
+#| tbl-cap: "Year-on-Year Incidence of CFRD"
+# Person-years variable
+py <- 1000
+# For each year, get the number of people diagnosed with CFRD that year
+years <- 2003:2018
+cases <- lapply(years, function(y) {
+  total <- unique(first_cfrd$eDWID[first_cfrd$first_cfrd == y])
+  cases_age_10 <- total[total %in% ages$eDWID[ages$`Age in 2018` < 10]]
+  cases_age_10_18 <- total[total %in% ages$eDWID[ages$`Age in 2018` >= 10 &
+    ages$`Age in 2018` < 19]]
+  cases_age_19_30 <- total[total %in% ages$eDWID[ages$`Age in 2018` >= 19 &
+    ages$`Age in 2018` < 31]]
+  cases_age_30 <- total[total %in% ages$eDWID[ages$`Age in 2018` >= 31]]
+  return(list(
+    "total" = total, "cases_age_10" = cases_age_10,
+    "cases_age_10_18" = cases_age_10_18,
+    "cases_age_19_30" = cases_age_19_30,
+    "cases_age_30" = cases_age_30
+  ))
+})
+# Double check that there is no overlap between sets (there isn't!)
+# lapply(1:length(cases), function(y){intersect(cases[[y]],unlist(cases[-y]))})
+# For each year, get the number of people at risk
+at_risk <- lapply(years, function(y) {
+  ar <- unique(df$eDWID[df$ReviewYear == y])
+  ar <- ar[!ar %in% unlist(cases[which(years <= y)]$total)]
+  ar_age_10 <- intersect(
+    df$eDWID[df$ReviewYear == y],
+    ages$eDWID[ages$`Age in 2018` < 10]
+  )
+  ar_age_10_18 <- intersect(
+    df$eDWID[df$ReviewYear == y],
+    ages$eDWID[ages$`Age in 2018` >= 10 & ages$`Age in 2018` < 19]
+  )
+  ar_age_19_30 <- intersect(
+    df$eDWID[df$ReviewYear == y],
+    ages$eDWID[ages$`Age in 2018` >= 19 & ages$`Age in 2018` < 31]
+  )
+  ar_age_30 <- intersect(
+    df$eDWID[df$ReviewYear == y],
+    ages$eDWID[ages$`Age in 2018` >= 31]
+  )
+  return(list(
+    "total" = ar, "ar_age_10" = ar_age_10, "ar_age_10_18" = ar_age_10_18,
+    "ar_age_19_30" = ar_age_19_30, "ar_age_30" = ar_age_30
+  ))
+})
+# Put everything in one table
+inc_df <- lapply(years, function(y) {
+  # Total
+  new_cases <- length(cases[[which(years == y)]]$total)
+  total_at_risk <- length(at_risk[[which(years == y)]]$total)
+  total_inc <- new_cases / (total_at_risk / py)
+  # < 10
+  new_cases_10 <- length(cases[[which(years == y)]]$cases_age_10)
+  at_risk_10 <- length(at_risk[[which(years == y)]]$ar_age_10)
+  inc_10 <- new_cases_10 / (at_risk_10 / py)
+  # 10 - 18
+  new_cases_10_18 <- length(cases[[which(years == y)]]$cases_age_10_18)
+  at_risk_10_18 <- length(at_risk[[which(years == y)]]$ar_age_10_18)
+  inc_10_18 <- new_cases_10_18 / (at_risk_10_18 / py)
+  # 19 - 30
+  new_cases_19_30 <- length(cases[[which(years == y)]]$cases_age_19_30)
+  at_risk_19_30 <- length(at_risk[[which(years == y)]]$ar_age_19_30)
+  inc_19_30 <- new_cases_19_30 / (at_risk_19_30 / py)
+  # > 30
+  new_cases_30 <- length(cases[[which(years == y)]]$cases_age_30)
+  at_risk_30 <- length(at_risk[[which(years == y)]]$ar_age_30)
+  inc_30 <- new_cases_30 / (at_risk_30 / py)
+  return(list(
+    `Year` = y,
+    `Total New Cases` = new_cases,
+    `Total At Risk` = total_at_risk,
+    `Total Incidence` = total_inc,
+    `Total New Cases Age < 10` = new_cases_10,
+    `Total At Risk Age < 10` = at_risk_10,
+    `Total Incidence Age < 10` = inc_10,
+    `Total New Cases Age 10-18` = new_cases_10_18,
+    `Total At Risk Age 10-18` = at_risk_10_18,
+    `Total Incidence Age 10-18` = inc_10_18,
+    `Total New Cases Age 19-30` = new_cases_19_30,
+    `Total At Risk Age 19-30` = at_risk_19_30,
+    `Total Incidence Age 19-30` = inc_19_30,
+    `Total New Cases Age > 30` = new_cases_30,
+    `Total At Risk Age > 30` = at_risk_30,
+    `Total Incidence Age > 30` = inc_30
+  ))
+})
+inc_df <- as.data.frame(do.call(rbind, inc_df))
+inc_df <- data.frame(lapply(inc_df, as.numeric), check.names = F)
+kable(inc_df, digits = 3)
+```
 
-diagnosis = r.diagnosis
-years = diagnosis["ReviewYear"].unique()
-# Loop through years - all ages
-cases = []
-at_risk = []
-incidence = []
-for y in years:
-  d = diagnosis[diagnosis["ReviewYear"]==y]
-  this_year = d.loc[d["cfrd_ogtt"]=="Yes","eDWID"].to_list()
-  cases = set(cases).union(this_year)
-  new_cases = set(cases).difference(set(this_year))
-  at_risk = set(d["eDWID"].to_list()).difference(at_risk)
-  incidence.append(round(len(new_cases)/(len(at_risk)/1000),3))
-# Loop through years - age < 10
-cases = []
-at_risk = []
-incidence_10 = []
-for y in years:
-  d = diagnosis[(diagnosis["ReviewYear"]==y) & (diagnosis["Age in 2018"]<10)]
-  this_year = d.loc[d["cfrd_ogtt"]=="Yes","eDWID"].to_list()
-  cases = set(cases).union(this_year)
-  new_cases = set(cases).difference(set(this_year))
-  at_risk = set(d["eDWID"].to_list()).difference(at_risk)
-  if len(at_risk)==0:
-    incidence_10.append(np.nan)
-  else:
-    incidence_10.append(round(len(new_cases)/(len(at_risk)/1000),3))
-# Loop through years - age 10 - 18
-cases = []
-at_risk = []
-incidence_10_18 = []
-for y in years:
-  d = diagnosis[(diagnosis["ReviewYear"]==y) & (diagnosis["Age in 2018"]>=10) & (diagnosis["Age in 2018"]<=18)]
-  this_year = d.loc[d["cfrd_ogtt"]=="Yes","eDWID"].to_list()
-  cases = set(cases).union(this_year)
-  new_cases = set(cases).difference(set(this_year))
-  at_risk = set(d["eDWID"].to_list()).difference(at_risk)
-  if len(at_risk)==0:
-    incidence_10_18.append(np.nan)
-  else:
-    incidence_10_18.append(round(len(new_cases)/(len(at_risk)/1000),3))
-# Loop through years - age 19 - 30
-cases = []
-at_risk = []
-incidence_19_30 = []
-for y in years:
-  d = diagnosis[(diagnosis["ReviewYear"]==y) & (diagnosis["Age in 2018"]>=19) & (diagnosis["Age in 2018"]<=30)]
-  this_year = d.loc[d["cfrd_ogtt"]=="Yes","eDWID"].to_list()
-  cases = set(cases).union(this_year)
-  new_cases = set(cases).difference(set(this_year))
-  at_risk = set(d["eDWID"].to_list()).difference(at_risk)
-  if len(at_risk)==0:
-    incidence_19_30.append(np.nan)
-  else:
-    incidence_19_30.append(round(len(new_cases)/(len(at_risk)/1000),3))
+In @tbl-incidence, incidence per year is calculated as the number of new CFRD diagnoses per 1,000 person-years of those at risk (excluding previously identified cases).
+
+```{r}
+df$cfrd_ogtt <- factor(df$cfrd_ogtt, levels = c("No", "Yes"), ordered = T)
+cfrd <- df %>%
+  group_by(eDWID) %>%
+  mutate(
+    cfrd_ever = max(cfrd_ogtt),
+    time_from_diag = ReviewYear - first_cfrd
+  ) %>%
+  filter(cfrd_ever == "Yes", time_from_diag <= 0, time_from_diag >= -1) %>%
+  summarise(hba1c = last(na.omit(round(A_hgba1c, 1)))) %>%
+  drop_na()
+non_cfrd <- df %>%
+  group_by(eDWID) %>%
+  mutate(cfrd_ever = max(cfrd_ogtt)) %>%
+  filter(cfrd_ever == "No") %>%
+  summarise(hba1c = last(na.omit(round(A_hgba1c, 1)))) %>%
+  drop_na()
 ```
+
+```{r}
+#| label: fig-roc
+#| fig-cap: "ROC Curve"
+r <- roc(
+  cases = cfrd$hba1c, controls = non_cfrd$hba1c, quiet = T, plot = T, auc = T, ci = T,
+  print.auc = TRUE, show.thres = TRUE
+)
+```
+
+Because we do not have exact information regarding HbA1c timing (we only have encounter number within year), the annualized HbA1c from the year of diagnosis or from one year prior to diagnosis was used in the ROC analysis.
+For those with HbA1c measures the year of diagnosis and one year prior, the value from year of diagnosis was preferred.
+For those never diagnosed with CFRD, the most recent HbA1c was used.
+There were a total of `r nrow(cfrd)` HbA1c measures for people with CFRD and `r nrow(non_cfrd)` measures for those without.
+
+The AUC for this ROC curve was 0.61 (see @fig-roc).
+
+```{r}
+#| label: tbl-roc_thresh
+#| tbl-cap: "Best ROC Threshold (Youden Index)"
+kable(coords(r, "best"), digits = 3)
+```

### 2022-12-23 Christine Chan/CF Registry/cf_registry_analysis.qmd
Update cf_registry_analysis.qmd

@@
 # All data
-df <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
+encounter <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
   na.strings = c("", "N/A")
 )
-df <- df %>%
-  arrange(eDWID, ReviewYear) %>%
-  mutate(`Age in 2018` = 2018 - (ReviewYear - Age_YrEnd))
-# For those with CFRD by OGTT, find first visit with diagnosis
-first_cfrd <- df %>%
-  filter(A_twohour_bloodglucose >= 200) %>%
-  group_by(eDWID) %>%
-  slice_min(ReviewYear) %>%
-  select(eDWID, ReviewYear, `Age in 2018`) %>%
-  rename(first_cfrd = ReviewYear)
-df <- left_join(df, first_cfrd, by = c("eDWID", "Age in 2018"))
-df <- df %>%
-  group_by(eDWID) %>%
-  mutate(cfrd_ogtt = case_when(
-    ReviewYear < first_cfrd ~ "No",
-    ReviewYear >= first_cfrd ~ "Yes"
-  )) %>%
-  ungroup()
-df$cfrd_ogtt[is.na(df$cfrd_ogtt)] <- "No"
-df <- df %>%
-  select(eDWID, ReviewYear, `Age in 2018`, first_cfrd, cfrd_ogtt, A_hgba1c)
-ages <- df %>%
+annual <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
+  na.strings = c("", "N/A")
+)
+encounter <- encounter %>%
+  arrange(eDWID, reviewyear, encounternum)
+# Add age
+ages <- annual %>%
   group_by(eDWID) %>%
@@
                                        decreasing = TRUE)[1]))
-ages$`Age in 2018` <- round(as.numeric(ages$`Age in 2018`), 1)
+encounter = left_join(encounter,ages,by = "eDWID")
+# For those with CFRD by OGTT, find first visit with diagnosis
+first_cfrd <- encounter %>%
+  filter(twohour_bloodglucose >= 200) %>%
+  group_by(eDWID) %>%
+  slice_min(reviewyear) %>%
+  select(eDWID, reviewyear, `Age in 2018`) %>%
+  rename(first_cfrd = reviewyear)
+encounter <- left_join(encounter, first_cfrd, by = c("eDWID", "Age in 2018"))
+encounter <- encounter %>%
+  group_by(eDWID) %>%
+  mutate(cfrd_ogtt = case_when(
+    reviewyear < first_cfrd ~ "No",
+    reviewyear >= first_cfrd ~ "Yes"
+  )) %>%
+  ungroup()
+encounter$cfrd_ogtt[is.na(df$cfrd_ogtt)] <- "No"
 ```
@@
 at_risk <- lapply(years, function(y) {
-  ar <- unique(df$eDWID[df$ReviewYear == y])
+  ar <- unique(encounter$eDWID[encounter$reviewyear == y])
   ar <- ar[!ar %in% unlist(cases[which(years <= y)]$total)]
   ar_age_10 <- intersect(
-    df$eDWID[df$ReviewYear == y],
+    encounter$eDWID[encounter$reviewyear == y],
     ages$eDWID[ages$`Age in 2018` < 10]
@@
   ar_age_10_18 <- intersect(
-    df$eDWID[df$ReviewYear == y],
+    encounter$eDWID[encounter$reviewyear == y],
     ages$eDWID[ages$`Age in 2018` >= 10 & ages$`Age in 2018` < 19]
@@
   ar_age_19_30 <- intersect(
-    df$eDWID[df$ReviewYear == y],
+    encounter$eDWID[encounter$reviewyear == y],
     ages$eDWID[ages$`Age in 2018` >= 19 & ages$`Age in 2018` < 31]
@@
   ar_age_30 <- intersect(
-    df$eDWID[df$ReviewYear == y],
+    encounter$eDWID[encounter$reviewyear == y],
     ages$eDWID[ages$`Age in 2018` >= 31]
@@
 ```{r}
+#| label: tbl-prevalence
+#| tbl-cap: "Yearly Prevalence of CFRD"
+df %>% group_by(ReviewYear) %>% 
+  summarise(Prevalance = mean(cfrd_status_annual==3,na.rm = T))
+
+t = df %>% select(eDWID,ReviewYear,A_twohour_bloodglucose,cfrd_ogtt,cfrd_status_annual)
+```
+
+In @tbl-prevalence, prevalence is calculated as the total number of CFRD cases seen in clinic divided by the total number of people seen in clinic in a given year.
+
+```{r}
+# Prepare data for ROC
 df$cfrd_ogtt <- factor(df$cfrd_ogtt, levels = c("No", "Yes"), ordered = T)
@@
 ```
+
+# Questions
+
+1.  For these calculations, do we need to just use those with an OGTT? Or everybody in the database?

### 2022-12-23 Christine Chan/CF Registry/cf_registry_analysis.qmd
cf registry from encounter level, remove ROC

--- 2022-12-22 Christine Chan/CF Registry/cf_registry_analysis.qmd
@@
 library(pROC)
+```
+
+```{r}
 # All data
-df = read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
-              na.strings = c("","N/A"))
-df = df %>% arrange(eDWID,ReviewYear) %>% 
-  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd))
+encounter <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
+                      na.strings = c("", "N/A")
+)
+annual <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
+                   na.strings = c("", "N/A")
+)
+encounter <- encounter %>%
+  arrange(eDWID, reviewyear, encounternum)
+# Add age
+ages <- annual %>% arrange(eDWID,ReviewYear) %>% 
+  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd)) %>%
+  group_by(eDWID) %>%
+  summarise(`Age in 2018` = as.numeric(names(sort(table(`Age in 2018`), 
+                                                  decreasing = TRUE)[1])))
+encounter = left_join(encounter,ages,by = "eDWID")
+encounter$`Age Group` = cut(encounter$`Age in 2018`,
+                            c(-Inf,10,19,31,Inf),
+                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
+                            right=F)
 # For those with CFRD by OGTT, find first visit with diagnosis
-first_cfrd = df %>% filter(A_twohour_bloodglucose >= 200) %>% 
-  group_by(eDWID) %>% slice_min(ReviewYear) %>% 
-  select(eDWID,ReviewYear,`Age in 2018`) %>% rename(first_cfrd = ReviewYear)
-df = left_join(df,first_cfrd,by = c("eDWID","Age in 2018"))
-df = df %>% group_by(eDWID) %>% 
+first_cfrd <- encounter %>%
+  filter(twohour_bloodglucose >= 200) %>%
+  group_by(eDWID) %>%
+  slice_min(reviewyear) %>%
+  select(eDWID, reviewyear, `Age in 2018`) %>%
+  rename(first_cfrd = reviewyear)
+encounter <- left_join(encounter, first_cfrd, by = c("eDWID", "Age in 2018"))
+encounter <- encounter %>%
+  group_by(eDWID) %>%
   mutate(cfrd_ogtt = case_when(
-    ReviewYear < first_cfrd ~ "No",
-    ReviewYear >= first_cfrd ~ "Yes"
+    reviewyear < first_cfrd ~ "No",
+    reviewyear >= first_cfrd ~ "Yes"
   )) %>%
   ungroup()
-df$cfrd_ogtt[is.na(df$cfrd_ogtt)] = "No"
-df = df %>% 
-  select(eDWID,ReviewYear,`Age in 2018`,first_cfrd,cfrd_ogtt,A_hgba1c)
-ages = df %>% group_by(eDWID) %>% 
-  summarise(`Age in 2018` = names(sort(table(`Age in 2018`), decreasing = TRUE)[1]))
-ages$`Age in 2018` = round(as.numeric(ages$`Age in 2018`),1)
+encounter$cfrd_ogtt[is.na(encounter$cfrd_ogtt)] <- "No"
 ```
@@
 #| tbl-cap: "Year-on-Year Incidence of CFRD"
-# Person-years variable
-py=1000
-# For each year, get the number of people diagnosed with CFRD that year
-years = 2003:2018
-cases = lapply(years, function(y){
-  total = unique(first_cfrd$eDWID[first_cfrd$first_cfrd==y])
-  cases_age_10 = total[total %in% ages$eDWID[ages$`Age in 2018` < 10]]
-  cases_age_10_18 = total[total %in% ages$eDWID[ages$`Age in 2018` >= 10 & ages$`Age in 2018` < 19]]
-  cases_age_19_30 = total[total %in% ages$eDWID[ages$`Age in 2018` >= 19 & ages$`Age in 2018` < 31]]
-  cases_age_30 = total[total %in% ages$eDWID[ages$`Age in 2018` >= 31]]
-  return(list("total"=total,"cases_age_10"=cases_age_10,
-              "cases_age_10_18"=cases_age_10_18,
-              "cases_age_19_30"=cases_age_19_30,
-              "cases_age_30"=cases_age_30))
-})
-# Double check that there is no overlap between sets (there isn't!)
-# lapply(1:length(cases), function(y){intersect(cases[[y]],unlist(cases[-y]))})
-# For each year, get the number of people at risk
-at_risk = lapply(years, function(y){
-  ar = unique(df$eDWID[df$ReviewYear==y])
-  ar = ar[!ar %in% unlist(cases[which(years<=y)]$total)]
-  ar_age_10 = intersect(df$eDWID[df$ReviewYear==y],ages$eDWID[ages$`Age in 2018`<10])
-  ar_age_10_18 = intersect(df$eDWID[df$ReviewYear==y],
-                           ages$eDWID[ages$`Age in 2018`>=10 & ages$`Age in 2018`<19])
-  ar_age_19_30 = intersect(df$eDWID[df$ReviewYear==y],
-                           ages$eDWID[ages$`Age in 2018`>=19 & ages$`Age in 2018`<31])
-  ar_age_30 = intersect(df$eDWID[df$ReviewYear==y],ages$eDWID[ages$`Age in 2018`>=31])
-  return(list("total"=ar,"ar_age_10"=ar_age_10,"ar_age_10_18"=ar_age_10_18,
-              "ar_age_19_30"=ar_age_19_30,"ar_age_30"=ar_age_30))
-})
-# Put everything in one table
-inc_df = lapply(years, function(y){
-  # Total
-  new_cases = length(cases[[which(years==y)]]$total)
-  total_at_risk = length(at_risk[[which(years==y)]]$total)
+# For each year, pull everyone with and without CFRD that year
+year_inc = function(df=encounter,year,py = 1000){
+  # Filter and count
+  inc = df %>% 
+    filter(reviewyear==year,is.na(first_cfrd) | first_cfrd == year) %>%
+    group_by(eDWID) %>% filter(row_number() == 1) %>%
+    select(eDWID,reviewyear,encounternum,`Age Group`,
+           twohour_bloodglucose,cfrd_ogtt,first_cfrd)
+  inc = inc %>% group_by(cfrd_ogtt) %>% count(`Age Group`,.drop = F)
+  # Calculations
+  new_cases = sum(inc$n[inc$cfrd_ogtt=="Yes"])
+  total_at_risk = sum(inc$n)
   total_inc = new_cases/(total_at_risk/py)
-  # < 10
-  new_cases_10 = length(cases[[which(years==y)]]$cases_age_10)
-  at_risk_10 = length(at_risk[[which(years==y)]]$ar_age_10)
+
+  new_cases_10 = sum(inc$n[inc$cfrd_ogtt=="Yes" & inc$`Age Group` == "< 10"])
+  at_risk_10 = sum(inc$n[inc$`Age Group` == "< 10"])
   inc_10 = new_cases_10/(at_risk_10/py)
-  # 10 - 18
-  new_cases_10_18 = length(cases[[which(years==y)]]$cases_age_10_18)
-  at_risk_10_18 = length(at_risk[[which(years==y)]]$ar_age_10_18)
+
+  new_cases_10_18 = sum(inc$n[inc$cfrd_ogtt=="Yes" & inc$`Age Group` == "10 - 18"])
+  at_risk_10_18 = sum(inc$n[inc$`Age Group` == "10 - 18"])
   inc_10_18 = new_cases_10_18/(at_risk_10_18/py)
-  # 19 - 30
-  new_cases_19_30 = length(cases[[which(years==y)]]$cases_age_19_30)
-  at_risk_19_30 = length(at_risk[[which(years==y)]]$ar_age_19_30)
+
+  new_cases_19_30 = sum(inc$n[inc$cfrd_ogtt=="Yes" & inc$`Age Group` == "19 - 30"])
+  at_risk_19_30 = sum(inc$n[inc$`Age Group` == "19 - 30"])
   inc_19_30 = new_cases_19_30/(at_risk_19_30/py)
-  # > 30
-  new_cases_30 = length(cases[[which(years==y)]]$cases_age_30)
-  at_risk_30 = length(at_risk[[which(years==y)]]$ar_age_30)
+
+  new_cases_30 = sum(inc$n[inc$cfrd_ogtt=="Yes" & inc$`Age Group` == "> 30"])
+  at_risk_30 = sum(inc$n[inc$`Age Group` == "> 30"])
   inc_30 = new_cases_30/(at_risk_30/py)
-  return(list(`Year` = y,
+
+  # Nice results
+  list(`Year` = year,
        `Total New Cases` = new_cases,
@@
        `Total Incidence` = total_inc,
+
        `Total New Cases Age < 10` = new_cases_10,
@@
        `Total Incidence Age < 10` = inc_10,
+
        `Total New Cases Age 10-18` = new_cases_10_18,
@@
        `Total Incidence Age 10-18` = inc_10_18,
+
        `Total New Cases Age 19-30` = new_cases_19_30,
@@
        `Total Incidence Age 19-30` = inc_19_30,
+
        `Total New Cases Age > 30` = new_cases_30,
        `Total At Risk Age > 30` = at_risk_30,
-              `Total Incidence Age > 30` = inc_30))
-})
-inc_df = as.data.frame(do.call(rbind,inc_df))
-inc_df = data.frame(lapply(inc_df,as.numeric),check.names = F)
-kable(inc_df,digits=3)
+       `Total Incidence Age > 30` = inc_30)
+}
+incidence = lapply(2003:2018, function(y){year_inc(year = y)})
+incidence = do.call(rbind,incidence)
+kable(incidence,digits = 3)
 ```
@@
 ```{r}
-df$cfrd_ogtt = factor(df$cfrd_ogtt,levels = c("No","Yes"),ordered = T)
-cfrd = df %>% group_by(eDWID) %>% 
-  mutate(cfrd_ever = max(cfrd_ogtt),
-         time_from_diag = ReviewYear-first_cfrd) %>%
-  filter(cfrd_ever=="Yes",time_from_diag <= 0,time_from_diag >= -1) %>%
-  summarise(hba1c = last(na.omit(round(A_hgba1c,1)))) %>% drop_na()
-non_cfrd =  df %>% group_by(eDWID) %>% 
-  mutate(cfrd_ever = max(cfrd_ogtt)) %>% filter(cfrd_ever=="No") %>%
-  summarise(hba1c = last(na.omit(round(A_hgba1c,1)))) %>% drop_na()
+#| label: tbl-prevalence
+#| tbl-cap: "Yearly Prevalence of CFRD"
+encounter %>% 
 ```
 
-```{r}
-#| label: fig-roc
-#| fig-cap: "ROC Curve"
-r = roc(cases = cfrd$hba1c,controls=non_cfrd$hba1c,quiet = T,plot=T,auc=T, ci=T,
-        print.auc=TRUE, show.thres=TRUE)
-```
+In @tbl-prevalence, prevalence is calculated as the total number of CFRD cases seen in clinic divided by the total number of people seen in clinic in a given year.
 
-Because we do not have exact information regarding HbA1c timing (we only have encounter number within year), the annualized HbA1c from the year of diagnosis or from one year prior to diagnosis was used in the ROC analysis.
-For those with HbA1c measures the year of diagnosis and one year prior, the value from year of diagnosis was preferred.
-For those never diagnosed with CFRD, the most recent HbA1c was used.
-There were a total of `r nrow(cfrd)` HbA1c measures for people with CFRD and `r nrow(non_cfrd)` measures for those without.
+# Questions
 
-The AUC for this ROC curve was 0.61 (see @fig-roc).
-
-```{r}
-#| label: tbl-roc_thresh
-#| tbl-cap: "Best ROC Threshold (Youden Index)"
-kable(coords(r,"best"),digits = 3)
-```
+1.  For these calculations, do we need to just use those with an OGTT? Or everybody in the database?

### 2022-12-24 Christine Chan/CF Registry/cf_registry_analysis.qmd
Update cf_registry_analysis.qmd

@@
 incidence = lapply(2003:2018, function(y){year_inc(year = y)})
-incidence = do.call(rbind,incidence)
+incidence = data.frame(do.call(rbind,incidence),check.names = F)
+incidence = data.frame(lapply(incidence,as.numeric),check.names = F)
 kable(incidence,digits = 3)
@@
 #| tbl-cap: "Yearly Prevalence of CFRD"
-encounter %>% 
+encounter$cfrd_ogtt = factor(encounter$cfrd_ogtt,levels = c("No","Yes"),ordered = T)
+encounter %>% group_by(eDWID,reviewyear) %>% 
+  summarise(highest_2_hour_bg = suppressWarnings(max(twohour_bloodglucose,na.rm = T)),
+            cfrd = max(cfrd_ogtt),
+            cfrd_status = suppressWarnings(max(cfrd_status,na.rm = T)),
+            .groups = "drop") %>%
+  group_by(reviewyear) %>%
+  summarise(`Prevalence by OGTT` = mean(as.numeric(cfrd)-1)*100,
+            `Prevalence by CFFPR` = mean(cfrd_status == 3,na.rm = T)*100) %>%
+  kable(.,digits = 3)
 ```
@@
 In @tbl-prevalence, prevalence is calculated as the total number of CFRD cases seen in clinic divided by the total number of people seen in clinic in a given year.
-
-# Questions
-
-1.  For these calculations, do we need to just use those with an OGTT? Or everybody in the database?

### 2022-12-26 Christine Chan/CF Registry/cf_registry_analysis.qmd
both CFRD definitions done (I think)

@@
   html:
-    toc: true
-    toc-depth: 5
-    toc-float: true
     code-fold: true
@@
     page-layout: full
-editor: visual
+editor: source
 ---
@@
 # For those with CFRD by OGTT, find first visit with diagnosis
-first_cfrd <- encounter %>%
+first_cfrd_ogtt <- encounter %>% 
   filter(twohour_bloodglucose >= 200) %>%
   group_by(eDWID) %>%
-  slice_min(reviewyear) %>%
+  filter(row_number()==1) %>%
   select(eDWID, reviewyear, `Age in 2018`) %>%
-  rename(first_cfrd = reviewyear)
-encounter <- left_join(encounter, first_cfrd, by = c("eDWID", "Age in 2018"))
+  rename(first_cfrd_ogtt = reviewyear)
+# For those with CFRD by registry criteria, find first visit with diagnosis
+first_cfrd_cffpr <- encounter %>% 
+  filter(cfrd_status==3) %>%
+  group_by(eDWID) %>%
+  filter(row_number()==1) %>%
+  select(eDWID, reviewyear, `Age in 2018`) %>%
+  rename(first_cfrd_cffpr = reviewyear)
+# Both
+first_cfrd_both <- encounter %>% 
+  filter(cfrd_status==3|twohour_bloodglucose >= 200) %>%
+  group_by(eDWID) %>%
+  filter(row_number()==1) %>%
+  select(eDWID, reviewyear, `Age in 2018`) %>%
+  rename(first_cfrd_both = reviewyear)
+# Add to encounter level data
+encounter <- left_join(encounter, first_cfrd_ogtt, by = c("eDWID", "Age in 2018"))
+encounter <- left_join(encounter, first_cfrd_cffpr, by = c("eDWID", "Age in 2018"))
+encounter <- left_join(encounter, first_cfrd_both,by = c("eDWID", "Age in 2018"))
+# Fill down CFRD Yes/No by year
 encounter <- encounter %>%
   group_by(eDWID) %>%
-  mutate(cfrd_ogtt = case_when(
-    reviewyear < first_cfrd ~ "No",
-    reviewyear >= first_cfrd ~ "Yes"
-  )) %>%
+  mutate(
+    cfrd_ogtt = case_when(
+      reviewyear < first_cfrd_ogtt ~ "No",
+      reviewyear >= first_cfrd_ogtt ~ "Yes"
+    ),
+    cfrd_cffpr = case_when(
+      reviewyear < first_cfrd_cffpr ~ "No",
+      reviewyear >= first_cfrd_cffpr ~ "Yes"
+    ),
+    cfrd_both = case_when(
+      reviewyear < first_cfrd_both ~ "No",
+      reviewyear >= first_cfrd_both ~ "Yes"
+    )
+  ) %>%
   ungroup()
 encounter$cfrd_ogtt[is.na(encounter$cfrd_ogtt)] <- "No"
+encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
+encounter$cfrd_both[is.na(encounter$cfrd_both)] <- "No"
+# Ordered factors
+encounter$cfrd_ogtt = factor(encounter$cfrd_ogtt,levels = c("No","Yes"),ordered = T)
+encounter$cfrd_cffpr = factor(encounter$cfrd_cffpr,levels = c("No","Yes"),ordered = T)
+encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
 ```
@@
 ```{r}
-#| label: tbl-incidence
-#| tbl-cap: "Year-on-Year Incidence of CFRD"
+#| label: tbl-incidence_cffpr
+#| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR Alone"
 # For each year, pull everyone with and without CFRD that year
@@
   inc = df %>%  
-    filter(reviewyear==year,is.na(first_cfrd) | first_cfrd == year) %>%
+    filter(reviewyear==year,is.na(first_cfrd_cffpr) | first_cfrd_cffpr == year) %>%
     group_by(eDWID) %>% filter(row_number() == 1) %>% 
-    select(eDWID,reviewyear,encounternum,`Age Group`,
-           twohour_bloodglucose,cfrd_ogtt,first_cfrd)
-  inc = inc %>% group_by(cfrd_ogtt) %>% count(`Age Group`,.drop = F)
+    select(eDWID,`Age Group`,cfrd_cffpr)
+  inc = inc %>% group_by(cfrd_cffpr) %>% count(`Age Group`,.drop = F)
   # Calculations
-  new_cases = sum(inc$n[inc$cfrd_ogtt=="Yes"])
+  new_cases = sum(inc$n[inc$cfrd_cffpr=="Yes"])
   total_at_risk = sum(inc$n)
@@
 
-  new_cases_10 = sum(inc$n[inc$cfrd_ogtt=="Yes" & inc$`Age Group` == "< 10"])
+  new_cases_10 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group` == "< 10"])
   at_risk_10 = sum(inc$n[inc$`Age Group` == "< 10"])
@@
 
-  new_cases_10_18 = sum(inc$n[inc$cfrd_ogtt=="Yes" & inc$`Age Group` == "10 - 18"])
+  new_cases_10_18 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group` == "10 - 18"])
   at_risk_10_18 = sum(inc$n[inc$`Age Group` == "10 - 18"])
@@
 
-  new_cases_19_30 = sum(inc$n[inc$cfrd_ogtt=="Yes" & inc$`Age Group` == "19 - 30"])
+  new_cases_19_30 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group` == "19 - 30"])
   at_risk_19_30 = sum(inc$n[inc$`Age Group` == "19 - 30"])
@@
 
-  new_cases_30 = sum(inc$n[inc$cfrd_ogtt=="Yes" & inc$`Age Group` == "> 30"])
+  new_cases_30 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group` == "> 30"])
   at_risk_30 = sum(inc$n[inc$`Age Group` == "> 30"])
@@
 
-In @tbl-incidence, incidence per year is calculated as the number of new CFRD diagnoses per 1,000 person-years of those at risk (excluding previously identified cases).
+In @tbl-incidence_cffpr, incidence per year is calculated as the number of new CFRD diagnoses (by registry definition alone) per 1,000 person-years of those at risk, excluding previously identified cases.
 
 ```{r}
-#| label: tbl-prevalence
-#| tbl-cap: "Yearly Prevalence of CFRD"
-encounter$cfrd_ogtt = factor(encounter$cfrd_ogtt,levels = c("No","Yes"),ordered = T)
+#| label: tbl-prevalence_cffpr
+#| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR Alone"
 encounter %>% group_by(eDWID,reviewyear) %>% 
-  summarise(highest_2_hour_bg = suppressWarnings(max(twohour_bloodglucose,na.rm = T)),
-            cfrd = max(cfrd_ogtt),
-            cfrd_status = suppressWarnings(max(cfrd_status,na.rm = T)),
-            .groups = "drop") %>%
+  summarise(cfrd = max(cfrd_cffpr),.groups = "drop") %>%
   group_by(reviewyear) %>%
-  summarise(`Prevalence by OGTT` = mean(as.numeric(cfrd)-1)*100,
-            `Prevalence by CFFPR` = mean(cfrd_status == 3,na.rm = T)*100) %>%
+  summarise(Prevalence = mean(as.numeric(cfrd)-1,na.rm = T)*100) %>%
   kable(.,digits = 3)
@@
 
-In @tbl-prevalence, prevalence is calculated as the total number of CFRD cases seen in clinic divided by the total number of people seen in clinic in a given year.
+In @tbl-prevalence_cffpr, prevalence is calculated as the total number of CFRD cases (by registry definition alone) seen in clinic divided by the total number of people seen in clinic in a given year.
+
+```{r}
+#| label: tbl-incidence_both
+#| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR and OGTT"
+# For each year, pull everyone with and without CFRD that year
+year_inc = function(df=encounter,year,py = 1000){
+  # Filter and count
+  inc = df %>% 
+    filter(reviewyear==year,is.na(first_cfrd_both) | first_cfrd_both == year) %>%
+    group_by(eDWID) %>% filter(row_number() == 1) %>%
+    select(eDWID,reviewyear,encounternum,`Age Group`,
+           twohour_bloodglucose,cfrd_both,first_cfrd_both)
+  inc = inc %>% group_by(cfrd_both) %>% count(`Age Group`,.drop = F)
+  # Calculations
+  new_cases = sum(inc$n[inc$cfrd_both=="Yes"])
+  total_at_risk = sum(inc$n)
+  total_inc = new_cases/(total_at_risk/py)
+
+  new_cases_10 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group` == "< 10"])
+  at_risk_10 = sum(inc$n[inc$`Age Group` == "< 10"])
+  inc_10 = new_cases_10/(at_risk_10/py)
+
+  new_cases_10_18 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group` == "10 - 18"])
+  at_risk_10_18 = sum(inc$n[inc$`Age Group` == "10 - 18"])
+  inc_10_18 = new_cases_10_18/(at_risk_10_18/py)
+
+  new_cases_19_30 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group` == "19 - 30"])
+  at_risk_19_30 = sum(inc$n[inc$`Age Group` == "19 - 30"])
+  inc_19_30 = new_cases_19_30/(at_risk_19_30/py)
+
+  new_cases_30 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group` == "> 30"])
+  at_risk_30 = sum(inc$n[inc$`Age Group` == "> 30"])
+  inc_30 = new_cases_30/(at_risk_30/py)
+
+  # Nice results
+  list(`Year` = year,
+       `Total New Cases` = new_cases,
+       `Total At Risk` = total_at_risk,
+       `Total Incidence` = total_inc,
+
+       `Total New Cases Age < 10` = new_cases_10,
+       `Total At Risk Age < 10` = at_risk_10,
+       `Total Incidence Age < 10` = inc_10,
+
+       `Total New Cases Age 10-18` = new_cases_10_18,
+       `Total At Risk Age 10-18` = at_risk_10_18,
+       `Total Incidence Age 10-18` = inc_10_18,
+
+       `Total New Cases Age 19-30` = new_cases_19_30,
+       `Total At Risk Age 19-30` = at_risk_19_30,
+       `Total Incidence Age 19-30` = inc_19_30,
+
+       `Total New Cases Age > 30` = new_cases_30,
+       `Total At Risk Age > 30` = at_risk_30,
+       `Total Incidence Age > 30` = inc_30)
+}
+incidence = lapply(2003:2018, function(y){year_inc(year = y)})
+incidence = data.frame(do.call(rbind,incidence),check.names = F)
+incidence = data.frame(lapply(incidence,as.numeric),check.names = F)
+kable(incidence,digits = 3)
+```
+
+In @tbl-incidence_both, incidence per year is calculated as the number of new CFRD diagnoses (by either registry definition or OGTT, whichever happened first) per 1,000 person-years of those at risk, excluding previously identified cases.
+
+```{r}
+#| label: tbl-prevalence_both
+#| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR and OGTT"
+encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
+encounter %>% group_by(eDWID,reviewyear) %>% 
+  summarise(cfrd = max(cfrd_both),.groups = "drop") %>%
+  group_by(reviewyear) %>%
+  summarise(`Prevalence` = mean(as.numeric(cfrd)-1,na.rm = T)*100) %>%
+  kable(.,digits = 3)
+```
+
+In @tbl-prevalence_both, prevalence is calculated as the total number of CFRD cases (by either registry definition or OGTT) seen in clinic divided by the total number of people seen in clinic in a given year.
+
+```{r}
+#| label: tbl-screening
+#| tbl-cap: "Yearly Screening Rates"
+
+```
+
+In @tbl-screening, people with CFRD (by either registry definition or OGTT) and those under 10 were excluded.
+Screening rate is calculated as the number of people with OGTTs divided by the total number of people seen in clinic that year.

### 2022-12-26 Christine Chan/CF Registry/cf_registry_analysis.qmd
CFRD registry screening and prevalence

--- 2022-12-23 Christine Chan/CF Registry/cf_registry_analysis.qmd
@@
   arrange(eDWID, reviewyear, encounternum)
+annual = annual %>% arrange(eDWID,ReviewYear)
 # Add age
-ages <- annual %>% arrange(eDWID,ReviewYear) %>% 
+ages <- annual %>% 
   mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd)) %>%
-  group_by(eDWID) %>%
+  rename(reviewyear = ReviewYear) %>%
+  group_by(eDWID,reviewyear) %>%
   summarise(`Age in 2018` = as.numeric(names(sort(table(`Age in 2018`), 
-                                                  decreasing = TRUE)[1])))
-encounter = left_join(encounter,ages,by = "eDWID")
-encounter$`Age Group` = cut(encounter$`Age in 2018`,
+                                                  decreasing = TRUE)[1])),
+            Age = Age_YrEnd[1],.groups = "drop")
+encounter = left_join(encounter,ages,by = c("eDWID", "reviewyear"))
+encounter$`Age Group 2018` = cut(encounter$`Age in 2018`,
+                            c(-Inf,10,19,31,Inf),
+                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
+                            right=F)
+encounter$`Age Group` = cut(encounter$Age,
                             c(-Inf,10,19,31,Inf),
@@
 # For those with CFRD by OGTT, find first visit with diagnosis
-first_cfrd <- encounter %>%
+first_cfrd_ogtt <- encounter %>% 
   filter(twohour_bloodglucose >= 200) %>%
   group_by(eDWID) %>%
-  slice_min(reviewyear) %>%
+  filter(row_number()==1) %>%
   select(eDWID, reviewyear, `Age in 2018`) %>%
-  rename(first_cfrd = reviewyear)
-encounter <- left_join(encounter, first_cfrd, by = c("eDWID", "Age in 2018"))
+  rename(first_cfrd_ogtt = reviewyear)
+# For those with CFRD by registry criteria, find first visit with diagnosis
+first_cfrd_cffpr <- encounter %>% 
+  filter(cfrd_status==3) %>%
+  group_by(eDWID) %>%
+  filter(row_number()==1) %>%
+  select(eDWID, reviewyear, `Age in 2018`) %>%
+  rename(first_cfrd_cffpr = reviewyear)
+# Both
+first_cfrd_both <- encounter %>% 
+  filter(cfrd_status==3|twohour_bloodglucose >= 200) %>%
+  group_by(eDWID) %>%
+  filter(row_number()==1) %>%
+  select(eDWID, reviewyear, `Age in 2018`) %>%
+  rename(first_cfrd_both = reviewyear)
+# Add to encounter level data
+encounter <- left_join(encounter, first_cfrd_ogtt, by = c("eDWID", "Age in 2018"))
+encounter <- left_join(encounter, first_cfrd_cffpr, by = c("eDWID", "Age in 2018"))
+encounter <- left_join(encounter, first_cfrd_both,by = c("eDWID", "Age in 2018"))
+# Fill down CFRD Yes/No by year
 encounter <- encounter %>%
   group_by(eDWID) %>%
-  mutate(cfrd_ogtt = case_when(
-    reviewyear < first_cfrd ~ "No",
-    reviewyear >= first_cfrd ~ "Yes"
-  )) %>%
+  mutate(
+    cfrd_ogtt = case_when(
+      reviewyear < first_cfrd_ogtt ~ "No",
+      reviewyear >= first_cfrd_ogtt ~ "Yes"
+    ),
+    cfrd_cffpr = case_when(
+      reviewyear < first_cfrd_cffpr ~ "No",
+      reviewyear >= first_cfrd_cffpr ~ "Yes"
+    ),
+    cfrd_both = case_when(
+      reviewyear < first_cfrd_both ~ "No",
+      reviewyear >= first_cfrd_both ~ "Yes"
+    )
+  ) %>%
   ungroup()
 encounter$cfrd_ogtt[is.na(encounter$cfrd_ogtt)] <- "No"
+encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
+encounter$cfrd_both[is.na(encounter$cfrd_both)] <- "No"
+# Ordered factors
+encounter$cfrd_ogtt = factor(encounter$cfrd_ogtt,levels = c("No","Yes"),ordered = T)
+encounter$cfrd_cffpr = factor(encounter$cfrd_cffpr,levels = c("No","Yes"),ordered = T)
+encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
+# Remove those missing age
+encounter = encounter %>% filter(!is.na(Age))
 ```
@@
   inc = df %>%  
-    filter(reviewyear==year,is.na(first_cfrd) | first_cfrd == year) %>%
+    filter(reviewyear==year,is.na(first_cfrd_cffpr) | first_cfrd_cffpr == year) %>%
     group_by(eDWID) %>% filter(row_number() == 1) %>% 
-    select(eDWID,reviewyear,encounternum,`Age Group`,
-           twohour_bloodglucose,cfrd_ogtt,first_cfrd)
-  inc = inc %>% group_by(cfrd_ogtt) %>% count(`Age Group`,.drop = F)
+    select(eDWID,`Age Group 2018`,cfrd_cffpr)
+  inc = inc %>% group_by(cfrd_cffpr) %>% count(`Age Group 2018`,.drop = F)
   # Calculations
-  new_cases = sum(inc$n[inc$cfrd_ogtt=="Yes"])
+  new_cases = sum(inc$n[inc$cfrd_cffpr=="Yes"])
   total_at_risk = sum(inc$n)
@@
 
-  new_cases_10 = sum(inc$n[inc$cfrd_ogtt=="Yes" & inc$`Age Group` == "< 10"])
-  at_risk_10 = sum(inc$n[inc$`Age Group` == "< 10"])
+  new_cases_10 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group 2018` == "< 10"])
+  at_risk_10 = sum(inc$n[inc$`Age Group 2018` == "< 10"])
   inc_10 = new_cases_10/(at_risk_10/py)
 
-  new_cases_10_18 = sum(inc$n[inc$cfrd_ogtt=="Yes" & inc$`Age Group` == "10 - 18"])
-  at_risk_10_18 = sum(inc$n[inc$`Age Group` == "10 - 18"])
+  new_cases_10_18 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group 2018` == "10 - 18"])
+  at_risk_10_18 = sum(inc$n[inc$`Age Group 2018` == "10 - 18"])
   inc_10_18 = new_cases_10_18/(at_risk_10_18/py)
 
-  new_cases_19_30 = sum(inc$n[inc$cfrd_ogtt=="Yes" & inc$`Age Group` == "19 - 30"])
-  at_risk_19_30 = sum(inc$n[inc$`Age Group` == "19 - 30"])
+  new_cases_19_30 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group 2018` == "19 - 30"])
+  at_risk_19_30 = sum(inc$n[inc$`Age Group 2018` == "19 - 30"])
   inc_19_30 = new_cases_19_30/(at_risk_19_30/py)
 
-  new_cases_30 = sum(inc$n[inc$cfrd_ogtt=="Yes" & inc$`Age Group` == "> 30"])
-  at_risk_30 = sum(inc$n[inc$`Age Group` == "> 30"])
+  new_cases_30 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group 2018` == "> 30"])
+  at_risk_30 = sum(inc$n[inc$`Age Group 2018` == "> 30"])
   inc_30 = new_cases_30/(at_risk_30/py)
@@
 
-       `Total New Cases Age < 10` = new_cases_10,
-       `Total At Risk Age < 10` = at_risk_10,
-       `Total Incidence Age < 10` = inc_10,
+       `Total New Cases Age in 2018 < 10` = new_cases_10,
+       `Total At Risk Age in 2018 < 10` = at_risk_10,
+       `Total Incidence Age in 2018 < 10` = inc_10,
 
-       `Total New Cases Age 10-18` = new_cases_10_18,
-       `Total At Risk Age 10-18` = at_risk_10_18,
-       `Total Incidence Age 10-18` = inc_10_18,
+       `Total New Cases Age in 2018 10-18` = new_cases_10_18,
+       `Total At Risk Age in 2018 10-18` = at_risk_10_18,
+       `Total Incidence Age in 2018 10-18` = inc_10_18,
 
-       `Total New Cases Age 19-30` = new_cases_19_30,
-       `Total At Risk Age 19-30` = at_risk_19_30,
-       `Total Incidence Age 19-30` = inc_19_30,
+       `Total New Cases Age in 2018 19-30` = new_cases_19_30,
+       `Total At Risk Age in 2018 19-30` = at_risk_19_30,
+       `Total Incidence Age in 2018 19-30` = inc_19_30,
 
-       `Total New Cases Age > 30` = new_cases_30,
-       `Total At Risk Age > 30` = at_risk_30,
-       `Total Incidence Age > 30` = inc_30)
+       `Total New Cases Age in 2018 > 30` = new_cases_30,
+       `Total At Risk Age in 2018 > 30` = at_risk_30,
+       `Total Incidence Age in 2018 > 30` = inc_30)
 }
 incidence = lapply(2003:2018, function(y){year_inc(year = y)})
-incidence = do.call(rbind,incidence)
+incidence = data.frame(do.call(rbind,incidence),check.names = F)
+incidence = data.frame(lapply(incidence,as.numeric),check.names = F)
 kable(incidence,digits = 3)
@@
 
-In @tbl-incidence, incidence per year is calculated as the number of new CFRD diagnoses per 1,000 person-years of those at risk (excluding previously identified cases).
+In @tbl-incidence_cffpr, incidence per year is calculated as the number of new CFRD diagnoses (by registry definition alone) per 1,000 person-years of those at risk, excluding previously identified cases.
 
 ```{r}
-#| label: tbl-prevalence
-#| tbl-cap: "Yearly Prevalence of CFRD"
-encounter %>% 
+#| label: tbl-prevalence_cffpr
+#| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR Alone"
+encounter %>% group_by(eDWID,reviewyear) %>% 
+  summarise(cfrd = max(cfrd_cffpr),.groups = "drop") %>%
+  group_by(reviewyear) %>%
+  summarise(Prevalence = mean(as.numeric(cfrd)-1,na.rm = T)*100) %>%
+  kable(.,digits = 3)
 ```
 
-In @tbl-prevalence, prevalence is calculated as the total number of CFRD cases seen in clinic divided by the total number of people seen in clinic in a given year.
+In @tbl-prevalence_cffpr, prevalence is calculated as the total number of CFRD cases (by registry definition alone) seen in clinic divided by the total number of people seen in clinic in a given year.
 
-# Questions
+```{r}
+#| label: tbl-incidence_both
+#| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR and OGTT"
+# For each year, pull everyone with and without CFRD that year
+year_inc = function(df=encounter,year,py = 1000){
+  # Filter and count
+  inc = df %>% 
+    filter(reviewyear==year,is.na(first_cfrd_both) | first_cfrd_both == year) %>%
+    group_by(eDWID) %>% filter(row_number() == 1) %>%
+    select(eDWID,reviewyear,encounternum,`Age Group 2018`,
+           twohour_bloodglucose,cfrd_both,first_cfrd_both)
+  inc = inc %>% group_by(cfrd_both) %>% count(`Age Group 2018`,.drop = F)
+  # Calculations
+  new_cases = sum(inc$n[inc$cfrd_both=="Yes"])
+  total_at_risk = sum(inc$n)
+  total_inc = new_cases/(total_at_risk/py)
 
-1.  For these calculations, do we need to just use those with an OGTT? Or everybody in the database?
+  new_cases_10 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group 2018` == "< 10"])
+  at_risk_10 = sum(inc$n[inc$`Age Group 2018` == "< 10"])
+  inc_10 = new_cases_10/(at_risk_10/py)
+
+  new_cases_10_18 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group 2018` == "10 - 18"])
+  at_risk_10_18 = sum(inc$n[inc$`Age Group 2018` == "10 - 18"])
+  inc_10_18 = new_cases_10_18/(at_risk_10_18/py)
+
+  new_cases_19_30 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group 2018` == "19 - 30"])
+  at_risk_19_30 = sum(inc$n[inc$`Age Group 2018` == "19 - 30"])
+  inc_19_30 = new_cases_19_30/(at_risk_19_30/py)
+
+  new_cases_30 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group 2018` == "> 30"])
+  at_risk_30 = sum(inc$n[inc$`Age Group 2018` == "> 30"])
+  inc_30 = new_cases_30/(at_risk_30/py)
+
+  # Nice results
+  list(`Year` = year,
+       `Total New Cases` = new_cases,
+       `Total At Risk` = total_at_risk,
+       `Total Incidence` = total_inc,
+
+       `Total New Cases Age in 2018 < 10` = new_cases_10,
+       `Total At Risk Age in 2018 < 10` = at_risk_10,
+       `Total Incidence Age in 2018 < 10` = inc_10,
+
+       `Total New Cases Age in 2018 10-18` = new_cases_10_18,
+       `Total At Risk Age in 2018 10-18` = at_risk_10_18,
+       `Total Incidence Age in 2018 10-18` = inc_10_18,
+
+       `Total New Cases Age in 2018 19-30` = new_cases_19_30,
+       `Total At Risk Age in 2018 19-30` = at_risk_19_30,
+       `Total Incidence Age in 2018 19-30` = inc_19_30,
+
+       `Total New Cases Age in 2018 > 30` = new_cases_30,
+       `Total At Risk Age in 2018 > 30` = at_risk_30,
+       `Total Incidence Age in 2018 > 30` = inc_30)
+}
+incidence = lapply(2003:2018, function(y){year_inc(year = y)})
+incidence = data.frame(do.call(rbind,incidence),check.names = F)
+incidence = data.frame(lapply(incidence,as.numeric),check.names = F)
+kable(incidence,digits = 3)
+```
+
+In @tbl-incidence_both, incidence per year is calculated as the number of new CFRD diagnoses (by either registry definition or OGTT, whichever happened first) per 1,000 person-years of those at risk, excluding previously identified cases.
+
+```{r}
+#| label: tbl-prevalence_both
+#| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR and OGTT"
+encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
+encounter %>% group_by(eDWID,reviewyear) %>% 
+  summarise(cfrd = max(cfrd_both),.groups = "drop") %>%
+  group_by(reviewyear) %>%
+  summarise(`Prevalence` = mean(as.numeric(cfrd)-1,na.rm = T)*100) %>%
+  kable(.,digits = 3)
+```
+
+In @tbl-prevalence_both, prevalence is calculated as the total number of CFRD cases (by either registry definition or OGTT) seen in clinic divided by the total number of people seen in clinic in a given year.
+
+```{r}
+#| label: tbl-screening
+#| tbl-cap: "Yearly Screening Rates"
+year_screen = function(df=encounter,year){
+  # Filter and count
+  inc = df %>% 
+    filter(reviewyear==year,first_cfrd_both >= year) %>%
+    mutate(ogtt = !(is.na(ogttfast_bloodglucose) | is.na(twohour_bloodglucose))) %>%
+    group_by(eDWID) %>%
+    summarise(ogtt = max(ogtt),`Age Group` = `Age Group`[1])
+  inc = inc %>% group_by(ogtt) %>% count(`Age Group`) %>% filter(`Age Group` != "< 10")
+  # Calculations
+  total_screened = sum(inc$n[inc$ogtt==1])
+  total_seen = sum(inc$n)
+  screen_rate = total_screened/total_seen * 100
+
+  screened_10_18 = inc$n[inc$ogtt==1 & inc$`Age Group` == "10 - 18"]
+  seen_10_18 = sum(inc$n[inc$`Age Group` == "10 - 18"])
+  rate_10_18 = screened_10_18/seen_10_18 *100
+
+  screened_19_30 = inc$n[inc$ogtt==1 & inc$`Age Group` == "19 - 30"]
+  seen_19_30 = sum(inc$n[inc$`Age Group` == "19 - 30"])
+  rate_19_30 = screened_19_30/seen_19_30 *100
+
+  screened_30 = inc$n[inc$ogtt==1 & inc$`Age Group` == "> 30"]
+  seen_30 = sum(inc$n[inc$`Age Group` == "> 30"])
+  rate_30 = screened_30/seen_30 *100
+
+  # Nice results
+  list(`Year` = year,
+       `Total Screened` = total_screened,
+       `Total Seen` = total_seen,
+       `Screen Rate (%)` = screen_rate,
+
+       `Total Screened Age 10-18` = screened_10_18,
+       `Total Seen Age 10-18` = seen_10_18,
+       `Screen Rate Age 10-18 (%)` = rate_10_18,
+
+       `Total Screened Age 19-30` = screened_19_30,
+       `Total Seen Age 19-30` = seen_19_30,
+       `Screen Rate Age 19-30 (%)` = rate_19_30,
+
+       `Total Screened Age > 30` = screened_30,
+       `Total Seen Age > 30` = seen_30,
+       `Screen Rate Age > 30 (%)` = rate_30)
+}
+screen = lapply(2003:2018, function(y){year_screen(year = y)})
+screen = data.frame(do.call(rbind,screen),check.names = F)
+screen = data.frame(lapply(screen,as.numeric),check.names = F)
+kable(screen,digits = 3)
+```
+
+In @tbl-screening, people with CFRD (by either registry definition or OGTT) and those under 10 were excluded.
+Screening rate is calculated as the number of people with an OGTT divided by the total number of people seen in clinic that year.

### 2022-12-27 Christine Chan/CF Registry/cf_registry_analysis.qmd
Update cf_registry_analysis.qmd

@@
 ages <- annual %>% 
-  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd)) %>%
   rename(reviewyear = ReviewYear) %>%
   group_by(eDWID,reviewyear) %>%
-  summarise(`Age in 2018` = as.numeric(names(sort(table(`Age in 2018`), 
-                                                  decreasing = TRUE)[1])),
-            Age = Age_YrEnd[1],.groups = "drop")
+  summarise(Age = as.numeric(Age_YrEnd[1]),.groups = "drop")
 encounter = left_join(encounter,ages,by = c("eDWID", "reviewyear"))
-encounter$`Age Group 2018` = cut(encounter$`Age in 2018`,
-                            c(-Inf,10,19,31,Inf),
-                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
-                            right=F)
 encounter$`Age Group` = cut(encounter$Age,
@@
   filter(row_number()==1) %>%
-  select(eDWID, reviewyear, `Age in 2018`) %>%
+  select(eDWID, reviewyear, `Age Group`) %>%
   rename(first_cfrd_ogtt = reviewyear)
@@
   filter(row_number()==1) %>%
-  select(eDWID, reviewyear, `Age in 2018`) %>%
+  select(eDWID, reviewyear, `Age Group`) %>%
   rename(first_cfrd_cffpr = reviewyear)
@@
   filter(row_number()==1) %>%
-  select(eDWID, reviewyear, `Age in 2018`) %>%
+  select(eDWID, reviewyear, `Age Group`) %>%
   rename(first_cfrd_both = reviewyear)
 # Add to encounter level data
-encounter <- left_join(encounter, first_cfrd_ogtt, by = c("eDWID", "Age in 2018"))
-encounter <- left_join(encounter, first_cfrd_cffpr, by = c("eDWID", "Age in 2018"))
-encounter <- left_join(encounter, first_cfrd_both,by = c("eDWID", "Age in 2018"))
+encounter <- left_join(encounter, first_cfrd_ogtt, by = c("eDWID", "Age Group"))
+encounter <- left_join(encounter, first_cfrd_cffpr, by = c("eDWID", "Age Group"))
+encounter <- left_join(encounter, first_cfrd_both,by = c("eDWID", "Age Group"))
 # Fill down CFRD Yes/No by year

### 2022-12-27 Christine Chan/CF Registry/cf_registry_analysis.qmd
a1c screening

@@
 Screening rate is calculated as the number of people with an OGTT divided by the total number of people seen in clinic that year.
+
+```{r}
+#| label: tbl-screening_a1c
+#| tbl-cap: "Yearly Screening Rates"
+year_screen = function(df=encounter,year){
+  # Filter and count
+  inc = df %>% 
+    filter(reviewyear==year,first_cfrd_both >= year) %>%
+    mutate(hba1c = !is.na(hgba1c)) %>%
+    group_by(eDWID) %>%
+    summarise(hba1c = max(hba1c),`Age Group` = `Age Group`[1])
+  inc = inc %>% group_by(hba1c) %>% count(`Age Group`) %>% filter(`Age Group` != "< 10")
+  # Calculations
+  total_screened = sum(inc$n[inc$hba1c==1])
+  total_seen = sum(inc$n)
+  screen_rate = total_screened/total_seen * 100
+
+  screened_10_18 = inc$n[inc$hba1c==1 & inc$`Age Group` == "10 - 18"]
+  seen_10_18 = sum(inc$n[inc$`Age Group` == "10 - 18"])
+  rate_10_18 = screened_10_18/seen_10_18 *100
+
+  screened_19_30 = inc$n[inc$hba1c==1 & inc$`Age Group` == "19 - 30"]
+  seen_19_30 = sum(inc$n[inc$`Age Group` == "19 - 30"])
+  rate_19_30 = screened_19_30/seen_19_30 *100
+
+  screened_30 = inc$n[inc$hba1c==1 & inc$`Age Group` == "> 30"]
+  seen_30 = sum(inc$n[inc$`Age Group` == "> 30"])
+  rate_30 = screened_30/seen_30 *100
+
+  # Nice results
+  list(`Year` = year,
+       `Total Screened` = total_screened,
+       `Total Seen` = total_seen,
+       `Screen Rate (%)` = screen_rate,
+
+       `Total Screened Age 10-18` = screened_10_18,
+       `Total Seen Age 10-18` = seen_10_18,
+       `Screen Rate Age 10-18 (%)` = rate_10_18,
+
+       `Total Screened Age 19-30` = screened_19_30,
+       `Total Seen Age 19-30` = seen_19_30,
+       `Screen Rate Age 19-30 (%)` = rate_19_30,
+
+       `Total Screened Age > 30` = screened_30,
+       `Total Seen Age > 30` = seen_30,
+       `Screen Rate Age > 30 (%)` = rate_30)
+}
+screen = lapply(2003:2018, function(y){year_screen(year = y)})
+screen = data.frame(do.call(rbind,screen),check.names = F)
+screen = data.frame(lapply(screen,as.numeric),check.names = F)
+kable(screen,digits = 3)
+```
+
+In @tbl-screening_a1c, people with CFRD (by either registry definition or OGTT) and those under 10 were excluded.
+Screening rate is calculated as the number of people with an HbA1c value divided by the total number of people seen in clinic that year.

### 2022-12-27 Christine Chan/CF Registry/cf_registry_analysis.qmd
Update cf_registry_analysis.qmd

@@
     group_by(eDWID) %>% filter(row_number() == 1) %>% 
-    select(eDWID,`Age Group 2018`,cfrd_cffpr)
-  inc = inc %>% group_by(cfrd_cffpr) %>% count(`Age Group 2018`,.drop = F)
+    select(eDWID,`Age Group`,cfrd_cffpr)
+  inc = inc %>% group_by(cfrd_cffpr) %>% count(`Age Group`,.drop = F)
   # Calculations
@@
 
-  new_cases_10 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group 2018` == "< 10"])
-  at_risk_10 = sum(inc$n[inc$`Age Group 2018` == "< 10"])
+  new_cases_10 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group` == "< 10"])
+  at_risk_10 = sum(inc$n[inc$`Age Group` == "< 10"])
   inc_10 = new_cases_10/(at_risk_10/py)
 
-  new_cases_10_18 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group 2018` == "10 - 18"])
-  at_risk_10_18 = sum(inc$n[inc$`Age Group 2018` == "10 - 18"])
+  new_cases_10_18 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group` == "10 - 18"])
+  at_risk_10_18 = sum(inc$n[inc$`Age Group` == "10 - 18"])
   inc_10_18 = new_cases_10_18/(at_risk_10_18/py)
 
-  new_cases_19_30 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group 2018` == "19 - 30"])
-  at_risk_19_30 = sum(inc$n[inc$`Age Group 2018` == "19 - 30"])
+  new_cases_19_30 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group` == "19 - 30"])
+  at_risk_19_30 = sum(inc$n[inc$`Age Group` == "19 - 30"])
   inc_19_30 = new_cases_19_30/(at_risk_19_30/py)
 
-  new_cases_30 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group 2018` == "> 30"])
-  at_risk_30 = sum(inc$n[inc$`Age Group 2018` == "> 30"])
+  new_cases_30 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group` == "> 30"])
+  at_risk_30 = sum(inc$n[inc$`Age Group` == "> 30"])
   inc_30 = new_cases_30/(at_risk_30/py)
@@
     group_by(eDWID) %>% filter(row_number() == 1) %>%
-    select(eDWID,reviewyear,encounternum,`Age Group 2018`,
+    select(eDWID,reviewyear,encounternum,`Age Group`,
            twohour_bloodglucose,cfrd_both,first_cfrd_both)
-  inc = inc %>% group_by(cfrd_both) %>% count(`Age Group 2018`,.drop = F)
+  inc = inc %>% group_by(cfrd_both) %>% count(`Age Group`,.drop = F)
   # Calculations
@@
 
-  new_cases_10 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group 2018` == "< 10"])
-  at_risk_10 = sum(inc$n[inc$`Age Group 2018` == "< 10"])
+  new_cases_10 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group` == "< 10"])
+  at_risk_10 = sum(inc$n[inc$`Age Group` == "< 10"])
   inc_10 = new_cases_10/(at_risk_10/py)
 
-  new_cases_10_18 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group 2018` == "10 - 18"])
-  at_risk_10_18 = sum(inc$n[inc$`Age Group 2018` == "10 - 18"])
+  new_cases_10_18 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group` == "10 - 18"])
+  at_risk_10_18 = sum(inc$n[inc$`Age Group` == "10 - 18"])
   inc_10_18 = new_cases_10_18/(at_risk_10_18/py)
 
-  new_cases_19_30 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group 2018` == "19 - 30"])
-  at_risk_19_30 = sum(inc$n[inc$`Age Group 2018` == "19 - 30"])
+  new_cases_19_30 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group` == "19 - 30"])
+  at_risk_19_30 = sum(inc$n[inc$`Age Group` == "19 - 30"])
   inc_19_30 = new_cases_19_30/(at_risk_19_30/py)
 
-  new_cases_30 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group 2018` == "> 30"])
-  at_risk_30 = sum(inc$n[inc$`Age Group 2018` == "> 30"])
+  new_cases_30 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group` == "> 30"])
+  at_risk_30 = sum(inc$n[inc$`Age Group` == "> 30"])
   inc_30 = new_cases_30/(at_risk_30/py)

### 2022-12-27 Christine Chan/CF Registry/cf_registry_analysis.qmd
updates finished!

@@
 ages <- annual %>% 
+  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd)) %>%
   rename(reviewyear = ReviewYear) %>%
   group_by(eDWID,reviewyear) %>%
-  summarise(Age = as.numeric(Age_YrEnd[1]),.groups = "drop")
+  summarise(`Age in 2018` = as.numeric(names(sort(table(`Age in 2018`), 
+                                                  decreasing = TRUE)[1])),
+            Age = Age_YrEnd[1],.groups = "drop")
 encounter = left_join(encounter,ages,by = c("eDWID", "reviewyear"))
+encounter$`Age Group 2018` = cut(encounter$`Age in 2018`,
+                            c(-Inf,10,19,31,Inf),
+                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
+                            right=F)
 encounter$`Age Group` = cut(encounter$Age,
@@
   filter(row_number()==1) %>%
-  select(eDWID, reviewyear, `Age Group`) %>%
+  select(eDWID, reviewyear, `Age in 2018`) %>%
   rename(first_cfrd_ogtt = reviewyear)
@@
   filter(row_number()==1) %>%
-  select(eDWID, reviewyear, `Age Group`) %>%
+  select(eDWID, reviewyear, `Age in 2018`) %>%
   rename(first_cfrd_cffpr = reviewyear)
@@
   filter(row_number()==1) %>%
-  select(eDWID, reviewyear, `Age Group`) %>%
+  select(eDWID, reviewyear, `Age in 2018`) %>%
   rename(first_cfrd_both = reviewyear)
 # Add to encounter level data
-encounter <- left_join(encounter, first_cfrd_ogtt, by = c("eDWID", "Age Group"))
-encounter <- left_join(encounter, first_cfrd_cffpr, by = c("eDWID", "Age Group"))
-encounter <- left_join(encounter, first_cfrd_both,by = c("eDWID", "Age Group"))
+encounter <- left_join(encounter, first_cfrd_ogtt, by = c("eDWID", "Age in 2018"))
+encounter <- left_join(encounter, first_cfrd_cffpr, by = c("eDWID", "Age in 2018"))
+encounter <- left_join(encounter, first_cfrd_both,by = c("eDWID", "Age in 2018"))
 # Fill down CFRD Yes/No by year
@@
   inc = inc %>% group_by(cfrd_cffpr) %>% count(`Age Group`,.drop = F)
+  inc = inc[!is.na(inc$`Age Group`),]
   # Calculations
@@
 
-       `Total New Cases Age in 2018 < 10` = new_cases_10,
-       `Total At Risk Age in 2018 < 10` = at_risk_10,
-       `Total Incidence Age in 2018 < 10` = inc_10,
+       `Total New Cases Age < 10` = new_cases_10,
+       `Total At Risk Age < 10` = at_risk_10,
+       `Total Incidence Age < 10` = inc_10,
 
-       `Total New Cases Age in 2018 10-18` = new_cases_10_18,
-       `Total At Risk Age in 2018 10-18` = at_risk_10_18,
-       `Total Incidence Age in 2018 10-18` = inc_10_18,
+       `Total New Cases Age 10-18` = new_cases_10_18,
+       `Total At Risk Age 10-18` = at_risk_10_18,
+       `Total Incidence Age 10-18` = inc_10_18,
 
-       `Total New Cases Age in 2018 19-30` = new_cases_19_30,
-       `Total At Risk Age in 2018 19-30` = at_risk_19_30,
-       `Total Incidence Age in 2018 19-30` = inc_19_30,
+       `Total New Cases Age 19-30` = new_cases_19_30,
+       `Total At Risk Age 19-30` = at_risk_19_30,
+       `Total Incidence Age 19-30` = inc_19_30,
 
-       `Total New Cases Age in 2018 > 30` = new_cases_30,
-       `Total At Risk Age in 2018 > 30` = at_risk_30,
-       `Total Incidence Age in 2018 > 30` = inc_30)
+       `Total New Cases Age > 30` = new_cases_30,
+       `Total At Risk Age > 30` = at_risk_30,
+       `Total Incidence Age > 30` = inc_30)
 }
@@
 ```{r}
+#| label: tbl-prevalence_cffpr_year
+#| tbl-cap: "Yearly Prevalence of CFRD by Age Group Based on CFFPR Alone"
+encounter %>% group_by(eDWID,reviewyear,`Age Group`) %>% 
+  summarise(cfrd = max(cfrd_cffpr),.groups = "drop") %>%
+  group_by(reviewyear,`Age Group`) %>%
+  summarise(Prevalence = mean(as.numeric(cfrd)-1,na.rm = T)*100,.groups = "drop") %>%
+  pivot_wider(names_from = `Age Group`,values_from = Prevalence) %>%
+  kable(.,digits = 3)
+```
+
+```{r}
 #| label: tbl-prevalence_cffpr
@@
   group_by(reviewyear) %>%
-  summarise(Prevalence = mean(as.numeric(cfrd)-1,na.rm = T)*100) %>%
+  summarise(Prevalence = mean(as.numeric(cfrd)-1,na.rm = T)*100,.groups = "drop") %>%
   kable(.,digits = 3)
@@
 
-In @tbl-prevalence_cffpr, prevalence is calculated as the total number of CFRD cases (by registry definition alone) seen in clinic divided by the total number of people seen in clinic in a given year.
+In @tbl-prevalence_cffpr_year and @tbl-prevalence_cffpr, prevalence is calculated as the total number of CFRD cases (by registry definition alone) seen in clinic divided by the total number of people seen in clinic in a given year.
 
@@
   inc = inc %>% group_by(cfrd_both) %>% count(`Age Group`,.drop = F)
+  inc = inc[!is.na(inc$`Age Group`),]
   # Calculations
@@
 ```{r}
+#| label: tbl-prevalence_both_year
+#| tbl-cap: "Yearly Prevalence of CFRD by Age Group Based on CFFPR and OGTT"
+encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
+encounter %>% group_by(eDWID,reviewyear,`Age Group`) %>% 
+  summarise(cfrd = max(cfrd_both),.groups = "drop") %>%
+  group_by(reviewyear,`Age Group`) %>%
+  summarise(`Prevalence` = mean(as.numeric(cfrd)-1,na.rm = T)*100,.groups = "drop") %>%
+  pivot_wider(names_from = `Age Group`,values_from = Prevalence)%>%
+  kable(.,digits = 3)
+```
+
+```{r}
 #| label: tbl-prevalence_both
 #| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR and OGTT"
-encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
 encounter %>% group_by(eDWID,reviewyear) %>% 
@@
   group_by(reviewyear) %>%
-  summarise(`Prevalence` = mean(as.numeric(cfrd)-1,na.rm = T)*100) %>%
+  summarise(`Prevalence` = mean(as.numeric(cfrd)-1,na.rm = T)*100,.groups = "drop") %>%
   kable(.,digits = 3)
@@
 
-In @tbl-prevalence_both, prevalence is calculated as the total number of CFRD cases (by either registry definition or OGTT) seen in clinic divided by the total number of people seen in clinic in a given year.
+In @tbl-prevalence_both_year and @tbl-prevalence_both, prevalence is calculated as the total number of CFRD cases (by either registry definition or OGTT) seen in clinic divided by the total number of people seen in clinic in a given year.
 
 ```{r}
-#| label: tbl-screening
-#| tbl-cap: "Yearly Screening Rates"
+#| label: tbl-screening_ogtt
+#| tbl-cap: "Yearly OGTT Screening Rates"
 year_screen = function(df=encounter,year){
@@
 
-In @tbl-screening, people with CFRD (by either registry definition or OGTT) and those under 10 were excluded.
-Screening rate is calculated as the number of people with an OGTT divided by the total number of people seen in clinic that year.
-
 ```{r}
 #| label: tbl-screening_a1c
-#| tbl-cap: "Yearly Screening Rates"
+#| tbl-cap: "Yearly HbA1c Screening Rates"
 year_screen = function(df=encounter,year){
@@
     filter(reviewyear==year,first_cfrd_both >= year) %>%
-    mutate(hba1c = !is.na(hgba1c)) %>%
+    mutate(a1c = !is.na(hgba1c)) %>%
     group_by(eDWID) %>%
-    summarise(hba1c = max(hba1c),`Age Group` = `Age Group`[1])
-  inc = inc %>% group_by(hba1c) %>% count(`Age Group`) %>% filter(`Age Group` != "< 10")
+    summarise(a1c = max(a1c),`Age Group` = `Age Group`[1])
+  inc = inc %>% group_by(a1c) %>% count(`Age Group`) %>% filter(`Age Group` != "< 10")
   # Calculations
-  total_screened = sum(inc$n[inc$hba1c==1])
+  total_screened = sum(inc$n[inc$a1c==1])
   total_seen = sum(inc$n)
@@
 
-  screened_10_18 = inc$n[inc$hba1c==1 & inc$`Age Group` == "10 - 18"]
+  screened_10_18 = inc$n[inc$a1c==1 & inc$`Age Group` == "10 - 18"]
   seen_10_18 = sum(inc$n[inc$`Age Group` == "10 - 18"])
@@
 
-  screened_19_30 = inc$n[inc$hba1c==1 & inc$`Age Group` == "19 - 30"]
+  screened_19_30 = inc$n[inc$a1c==1 & inc$`Age Group` == "19 - 30"]
   seen_19_30 = sum(inc$n[inc$`Age Group` == "19 - 30"])
@@
 
-  screened_30 = inc$n[inc$hba1c==1 & inc$`Age Group` == "> 30"]
+  screened_30 = inc$n[inc$a1c==1 & inc$`Age Group` == "> 30"]
   seen_30 = sum(inc$n[inc$`Age Group` == "> 30"])
@@
 
-In @tbl-screening_a1c, people with CFRD (by either registry definition or OGTT) and those under 10 were excluded.
-Screening rate is calculated as the number of people with an HbA1c value divided by the total number of people seen in clinic that year.
+In @tbl-screening_ogtt and @tbl-screening_a1c, people with CFRD (by either registry definition or OGTT) and those under age 10 were excluded.
+Screening rate is calculated as the number of people with an OGTT divided by the total number of people seen in clinic that year.

### 2022-12-27 Christine Chan/CF Registry/cf_registry_analysis.qmd
fixed incidence denominator

@@
   inc = df %>%  
-    filter(reviewyear==year,is.na(first_cfrd_cffpr) | first_cfrd_cffpr == year) %>%
+    filter(reviewyear==year,is.na(first_cfrd_cffpr) | first_cfrd_cffpr >= year) %>%
     group_by(eDWID) %>% filter(row_number() == 1) %>% 
@@
   inc = df %>% 
-    filter(reviewyear==year,is.na(first_cfrd_both) | first_cfrd_both == year) %>%
+    filter(reviewyear==year,is.na(first_cfrd_both) | first_cfrd_both >= year) %>%
     group_by(eDWID) %>% filter(row_number() == 1) %>%

### 2022-12-27 Christine Chan/CF Registry/cf_registry_analysis.qmd
Update cf_registry_analysis.qmd

@@
   inc = df %>% 
-    filter(reviewyear==year,first_cfrd_both >= year) %>%
-    mutate(ogtt = !(is.na(ogttfast_bloodglucose) | is.na(twohour_bloodglucose))) %>%
+    filter(reviewyear==year,first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>%
+    mutate(ogtt = !is.na(ogttfast_bloodglucose) | !is.na(twohour_bloodglucose)) %>%
     group_by(eDWID) %>%
@@
   inc = df %>% 
-    filter(reviewyear==year,first_cfrd_both >= year) %>%
+    filter(reviewyear==year,first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>%
     mutate(a1c = !is.na(hgba1c)) %>%
@@
 
-In @tbl-screening_ogtt and @tbl-screening_a1c, people with CFRD (by either registry definition or OGTT) and those under age 10 were excluded.
+In @tbl-screening_ogtt and @tbl-screening_a1c, people with CFRD (by registry definition) and those under age 10 were excluded.
 Screening rate is calculated as the number of people with an OGTT divided by the total number of people seen in clinic that year.

### 2023-01-03 Emily Muther/mental_health_screening.qmd
Create mental_health_screening.qmd

---
title: "Mental Health Screening in CF"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: visual
---

```{r}
#| include: false
library(tidyverse)
library(lubridate)
```

```{r}
# Data import
cf_gad = read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/GAD-7_General_Anxiety_Disorder_Full_Data_data_11-15-22.csv",na.strings = "")
cf_phq = read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/PHQ-9_Full_Data_data_11-11-22.csv",na.strings = "")
cf_psc = read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/PSC_Full_Data_data_11-11-22.csv",na.strings = "")
# Select and merge
cf_gad = cf_gad %>% select(PAT_MRN_ID,CONTACT_DATE,GAD.7.Total.Score,GAD.7.Totals) %>% drop_na()
cf_phq = cf_phq %>% 
  select(PAT_MRN_ID,CONTACT_DATE,PHQ.9.Total.Score..calculated.,PHQ.9.Totals) %>% drop_na()
cf_psc = cf_psc %>% select(PAT_MRN_ID,CONTACT_DATE,Score,PSC.Totals) %>% drop_na()
df = cf_gad %>% 
  full_join(cf_phq,by = c("PAT_MRN_ID", "CONTACT_DATE")) %>%
  full_join(cf_psc,by = c("PAT_MRN_ID", "CONTACT_DATE"))
# Get age
pa = read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/CF_Psychosocial_Assessment_Full_Data_data_11-15-22.csv",na.strings = "")
pa = pa %>% select(PAT_MRN_ID,CONTACT_DATE,AGE_AT_CONTACT_IN_YEARS)
# Rename columns, get distinct rows
df = left_join(df,pa,by = c("PAT_MRN_ID", "CONTACT_DATE")) %>% 
  select(PAT_MRN_ID,CONTACT_DATE,AGE_AT_CONTACT_IN_YEARS,everything()) %>%
  rename(id = PAT_MRN_ID,date = CONTACT_DATE,age=AGE_AT_CONTACT_IN_YEARS,
         `GAD-7 Total` = GAD.7.Total.Score,
         `GAD-7 Category`=GAD.7.Totals,
         `PHQ-9 Total`=PHQ.9.Total.Score..calculated.,
         `PHQ-9 Category`=PHQ.9.Totals,
         `PSC-17 Total`=Score,
         `PSC-17 Category`=PSC.Totals) %>%
  distinct()
df$date = parse_date(df$date,approx = F)
df = df %>% arrange(id,date)
```

# Table

```{r}
#| label: tbl-planets
#| tbl-cap: Planets

```

# Figure

```{r}
#| label: fig-planets
#| fig-cap: Planets

```

### 2023-01-04 Emily Muther/mental_health_screening.qmd
Update mental_health_screening.qmd
tidyverse formatting

@@
 #| include: false
+library(lubridate)
 library(tidyverse)
-library(lubridate)
 ```
@@
 # Data import
-cf_gad = read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/GAD-7_General_Anxiety_Disorder_Full_Data_data_11-15-22.csv",na.strings = "")
-cf_phq = read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/PHQ-9_Full_Data_data_11-11-22.csv",na.strings = "")
-cf_psc = read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/PSC_Full_Data_data_11-11-22.csv",na.strings = "")
+cf_gad <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/GAD-7_General_Anxiety_Disorder_Full_Data_data_11-15-22.csv", na.strings = "")
+cf_phq <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/PHQ-9_Full_Data_data_11-11-22.csv", na.strings = "")
+cf_psc <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/PSC_Full_Data_data_11-11-22.csv", na.strings = "")
 # Select and merge
-cf_gad = cf_gad %>% select(PAT_MRN_ID,CONTACT_DATE,GAD.7.Total.Score,GAD.7.Totals) %>% drop_na()
-cf_phq = cf_phq %>% 
-  select(PAT_MRN_ID,CONTACT_DATE,PHQ.9.Total.Score..calculated.,PHQ.9.Totals) %>% drop_na()
-cf_psc = cf_psc %>% select(PAT_MRN_ID,CONTACT_DATE,Score,PSC.Totals) %>% drop_na()
-df = cf_gad %>% 
-  full_join(cf_phq,by = c("PAT_MRN_ID", "CONTACT_DATE")) %>%
-  full_join(cf_psc,by = c("PAT_MRN_ID", "CONTACT_DATE"))
+cf_gad <- cf_gad %>%
+  select(PAT_MRN_ID, CONTACT_DATE, GAD.7.Total.Score, GAD.7.Totals) %>%
+  drop_na()
+cf_phq <- cf_phq %>%
+  select(PAT_MRN_ID, CONTACT_DATE, PHQ.9.Total.Score..calculated., PHQ.9.Totals) %>%
+  drop_na()
+cf_psc <- cf_psc %>%
+  select(PAT_MRN_ID, CONTACT_DATE, Score, PSC.Totals) %>%
+  drop_na()
+df <- cf_gad %>%
+  full_join(cf_phq, by = c("PAT_MRN_ID", "CONTACT_DATE")) %>%
+  full_join(cf_psc, by = c("PAT_MRN_ID", "CONTACT_DATE"))
 # Get age
-pa = read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/CF_Psychosocial_Assessment_Full_Data_data_11-15-22.csv",na.strings = "")
-pa = pa %>% select(PAT_MRN_ID,CONTACT_DATE,AGE_AT_CONTACT_IN_YEARS)
+pa <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/CF_Psychosocial_Assessment_Full_Data_data_11-15-22.csv", na.strings = "")
+pa <- pa %>% select(PAT_MRN_ID, CONTACT_DATE, AGE_AT_CONTACT_IN_YEARS)
 # Rename columns, get distinct rows
-df = left_join(df,pa,by = c("PAT_MRN_ID", "CONTACT_DATE")) %>% 
-  select(PAT_MRN_ID,CONTACT_DATE,AGE_AT_CONTACT_IN_YEARS,everything()) %>%
-  rename(id = PAT_MRN_ID,date = CONTACT_DATE,age=AGE_AT_CONTACT_IN_YEARS,
+df <- left_join(df, pa, by = c("PAT_MRN_ID", "CONTACT_DATE")) %>%
+  select(PAT_MRN_ID, CONTACT_DATE, AGE_AT_CONTACT_IN_YEARS, everything()) %>%
+  rename(
+    id = PAT_MRN_ID, date = CONTACT_DATE, age = AGE_AT_CONTACT_IN_YEARS,
     `GAD-7 Total` = GAD.7.Total.Score,
-         `GAD-7 Category`=GAD.7.Totals,
-         `PHQ-9 Total`=PHQ.9.Total.Score..calculated.,
-         `PHQ-9 Category`=PHQ.9.Totals,
-         `PSC-17 Total`=Score,
-         `PSC-17 Category`=PSC.Totals) %>%
+    `GAD-7 Category` = GAD.7.Totals,
+    `PHQ-9 Total` = PHQ.9.Total.Score..calculated.,
+    `PHQ-9 Category` = PHQ.9.Totals,
+    `PSC-17 Total` = Score,
+    `PSC-17 Category` = PSC.Totals
+  ) %>%
   distinct()
-df$date = parse_date(df$date,approx = F)
-df = df %>% arrange(id,date)
+# Format columns and sort
+df$date <- mdy(df$date)
+df <- df %>%
+  mutate(across(contains("Category"), factor)) %>%
+  arrange(id, date)
+# Write cleaned data
+write.csv(df, file = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Cleaned/cf_mental_health_screening.csv")
 ```
 
-# Table
-
 ```{r}
-#| label: tbl-planets
-#| tbl-cap: Planets
-
+skimr::skim(df)
 ```
-
-# Figure
-
-```{r}
-#| label: fig-planets
-#| fig-cap: Planets
-
-```

### 2023-01-05 Emily Muther/mental_health_screening.qmd
Update mental_health_screening.qmd

--- 2023-01-03 Emily Muther/mental_health_screening.qmd
@@
 author: "Tim Vigers"
-date: "`r format(Sys.time(), '%d %B %Y')`"
+date: last-modified
 format:
   html:
-    toc: true
-    toc-depth: 5
-    toc-float: true
     code-fold: true
     self-contained: true
-    fig-cap-location: top
     page-layout: full
-editor: visual
+    title-block-banner: true
+editor: source
 ---
 
-```{r}
-#| include: false
-library(tidyverse)
-library(lubridate)
+# Data Cleaning
+
+```{python}
+import pandas as pd
+# Data import
+gad7 = pd.read_csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/GAD-7_General_Anxiety_Disorder_Full_Data_data_11-15-22.csv")
+gad7 = gad7[['PAT_MRN_ID','CONTACT_DATE','AGE AT CONTACT ','GAD-7 Total Score','GAD-7 Totals']]
+gad7.dropna(axis=0,inplace=True)
+gad7.drop_duplicates(inplace=True)
+phq9 = pd.read_csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/PHQ-9_Full_Data_data_11-11-22.csv")
+phq9 = phq9[['PAT_MRN_ID','CONTACT_DATE','AGE AT CONTACT ','PHQ-9 Total Score (calculated)','PHQ-9 Totals']]
+phq9.dropna(axis=0,inplace=True)
+phq9.drop_duplicates(inplace=True)
+psc = pd.read_csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/PSC_Full_Data_data_11-11-22.csv")
+psc = psc[['PAT_MRN_ID','CONTACT_DATE','AGE AT CONTACT ','Score','PSC Totals']]
+psc.dropna(axis=0,inplace=True)
+psc.drop_duplicates(inplace=True)
+# Merge
+df = gad7.merge(phq9,on=['PAT_MRN_ID','CONTACT_DATE','AGE AT CONTACT '],how="outer")
+df = df.merge(psc,on=['PAT_MRN_ID','CONTACT_DATE','AGE AT CONTACT '],how="outer")
+# Rename
+df.rename({'PAT_MRN_ID':'ID','CONTACT_DATE':'Date','AGE AT CONTACT ':'Age','GAD-7 Total Score':'GAD-7 Total','GAD-7 Totals':'GAD-7 Group','PHQ-9 Total Score (calculated)':'PHQ-9 Total','PHQ-9 Totals':'PHQ-9 Group','Score':'PSC-17 Total','PSC Totals':'PSC-17 Group'},inplace=True,axis=1)
 ```
 
-```{r}
-# Data import
-cf_gad = read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/GAD-7_General_Anxiety_Disorder_Full_Data_data_11-15-22.csv",na.strings = "")
-cf_phq = read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/PHQ-9_Full_Data_data_11-11-22.csv",na.strings = "")
-cf_psc = read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/PSC_Full_Data_data_11-11-22.csv",na.strings = "")
-# Select and merge
-cf_gad = cf_gad %>% select(PAT_MRN_ID,CONTACT_DATE,GAD.7.Total.Score,GAD.7.Totals) %>% drop_na()
-cf_phq = cf_phq %>% 
-  select(PAT_MRN_ID,CONTACT_DATE,PHQ.9.Total.Score..calculated.,PHQ.9.Totals) %>% drop_na()
-cf_psc = cf_psc %>% select(PAT_MRN_ID,CONTACT_DATE,Score,PSC.Totals) %>% drop_na()
-df = cf_gad %>% 
-  full_join(cf_phq,by = c("PAT_MRN_ID", "CONTACT_DATE")) %>%
-  full_join(cf_psc,by = c("PAT_MRN_ID", "CONTACT_DATE"))
-# Get age
-pa = read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/CF_Psychosocial_Assessment_Full_Data_data_11-15-22.csv",na.strings = "")
-pa = pa %>% select(PAT_MRN_ID,CONTACT_DATE,AGE_AT_CONTACT_IN_YEARS)
-# Rename columns, get distinct rows
-df = left_join(df,pa,by = c("PAT_MRN_ID", "CONTACT_DATE")) %>% 
-  select(PAT_MRN_ID,CONTACT_DATE,AGE_AT_CONTACT_IN_YEARS,everything()) %>%
-  rename(id = PAT_MRN_ID,date = CONTACT_DATE,age=AGE_AT_CONTACT_IN_YEARS,
-         `GAD-7 Total` = GAD.7.Total.Score,
-         `GAD-7 Category`=GAD.7.Totals,
-         `PHQ-9 Total`=PHQ.9.Total.Score..calculated.,
-         `PHQ-9 Category`=PHQ.9.Totals,
-         `PSC-17 Total`=Score,
-         `PSC-17 Category`=PSC.Totals) %>%
-  distinct()
-df$date = parse_date(df$date,approx = F)
-df = df %>% arrange(id,date)
+- On first import, there are 2,646 observations each for GAD-7, PHQ-9, and PSC-17. 
+- After deleting rows missing any any of patient ID, contact date, age at contact, total score, or score group, there were: 
+  - 559 observations for GAD-7
+  - 548 for PHQ-9
+  - 308 for PSC-17
+- After dropping duplicate rows, there are: 
+  - 390 GAD-7 observations
+  - 383 for PHQ-9
+  - 216 for PSC-17
+- After merging these together based on ID, date, and age, there are 613 unique observations with at least one of the three screening measures.
+
+# Profiling
+
+```{python}
+from pandas_profiling import ProfileReport
+profile = ProfileReport(df, title="Data Profiling Report")
+profile.to_notebook_iframe()
 ```
-
-# Table
-
-```{r}
-#| label: tbl-planets
-#| tbl-cap: Planets
-
-```
-
-# Figure
-
-```{r}
-#| label: fig-planets
-#| fig-cap: Planets
-
-```

### 2023-01-10 Emily Muther/mental_health_screening.qmd
Update mental_health_screening.qmd

@@
 df.rename({'PAT_MRN_ID':'ID','CONTACT_DATE':'Date','AGE AT CONTACT ':'Age','GAD-7 Total Score':'GAD-7 Total','GAD-7 Totals':'GAD-7 Group','PHQ-9 Total Score (calculated)':'PHQ-9 Total','PHQ-9 Totals':'PHQ-9 Group','Score':'PSC-17 Total','PSC Totals':'PSC-17 Group'},inplace=True,axis=1)
+# Write to CSV
+df.to_csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Cleaned/cf_mental_health_screening.csv")
 ```

### 2023-01-11 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
trikafta individual plots

@@
 knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
-knitr::opts_knit$set(root.dir = "~/Dropbox/Work/Vigers/CF")
+knitr::opts_knit$set(root.dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF")
 ```
@@
 df = read.csv("./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/trikafta_2022-02-8.csv",
-              na.strings = c(""))
+              na.strings = c("","no serum"))
 # Clean up
@@
 
+# Individual plots
+
+```{r}
+# Small cohort
+reduced = df[df$Study.ID %in% keep_dxa,]
+# OGTT plots before and after
+plot_df = reduced %>% pivot_longer(all_of(glu_cols)) %>% select(Study.ID,Event.Name,name,value)
+plot_df$name = as.numeric(gsub("Glucose|\\.|min|minute","",plot_df$name))
+ggplot(plot_df,aes(x = name,y = value,color = Study.ID)) + 
+  geom_line() + geom_point() + 
+  xlab("Time (minutes)") + ylab("Glucose (mg/dL)") + 
+  theme_bw() +
+  theme(legend.position = "none") + 
+  facet_wrap(~Event.Name)
+```
+
+It was difficult to tell which patient was which in grayscale.
+
+```{r results="asis"}
+plot_vars = c("Weight..kg.","weightz","BMIz","Fat.Mass.Index..FMI...","Fat.Free.Mass.Index...FFMI.","`Glucose iAUC`","c.peptide.0.minute","`C Peptide iAUC`","iAUC_ratio","`C Peptide ODI`","`Adapted Matsuda`","homa2IR","FEV1..","FVC....")
+# Loop through variables and plot
+for (v in plot_vars) {
+  title = ifelse(v %in% names(newlabels),as.character(newlabels[v]),gsub("`","",v))
+  p = ggplot(reduced,aes_string(x="Event.Name",y = v,group="Study.ID")) + 
+    geom_point() + geom_line() + 
+    ylab(as.character(title)) + 
+    theme_bw() + theme(axis.title.x = element_blank())
+  cat("\n")
+  cat("## ",as.character(title))
+  cat("\n")
+  print(p)
+  cat("\n")
+}
+```
+
 # Notes

### 2023-01-12 Emily Muther/body_image_trikafta.qmd
Create body_image_trikafta.qmd

---
title: "Body Image and Trikafta"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: visual
---

```{r}
#| include: false
# Libraries
library(haven)
library(tidyverse)
library(arsenal)
# Import
df = read_sav("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Body Image and Trikafta/Data_Raw/TrikaftaBodyImage1.11.23.sav")
# Format variables
df$`Percent Adherence` = factor(df$PercentAdherence,levels = 1:11,
                             labels = paste0(seq(0,100,by=10),"%"),ordered = T)
df$`FRS Category` = factor(df$FRS_Category,levels = -1:1,
                           labels = c("want to be larger","neutral",
                                      "want to be smaller"))
df$`Doses Missed` = as.numeric(df$DosesMissed)
df$`Weight/Body Adherence` = factor(df$WeightBody_Adherence,levels = 1:5,
               labels = c("Strongly Agree","Slightly Agree","Neutral",
                          "Slightly Disagree","Strongly Disagree"))
df$`Weight/Body Adherence` = fct_rev(df$`Weight/Body Adherence`)
df$`Weight/Body Adherence` = factor(df$`Weight/Body Adherence`,ordered = T)
```

Because we want to keep the ordinal nature of some of these variables rather than simply treating them as categorical (i.e. with percent adherence it's meaningful that 10% \< 20% \< 30% ...), Emma is right in thinking that ordinal logistic regression is the right approach. Unfortunately, this requires estimating a lot of parameters, and we probably don't have the necessary numbers for it (for example, with percent adherence as the outcome we'd need to estimate 12 parameters, and the general rule of thumb is at least 10 observations per parameter).

So, after talking this through with another statistician in our group, I think the best way to approach this is to treat these variables as continuous, but apply non-parametric, rank-based tests between various groups. This keeps the information about the order of the categories but does not assume normal distribution, etc.

```{r results='asis'}
#| label: tbl-1
#| tbl-cap: Comparisons by FRS Category
t1 = tableby(`FRS Category`~kwt(`Percent Adherence`)+`Doses Missed` +
               kwt(`Weight/Body Adherence`),data = df,
             control=tableby.control(numeric.stats=c("Nmiss","median","q1q3","range"),
                                     numeric.test="kwt"))
summary(t1,pfootnote = T)
```

```{r}
#| label: fig-planets
#| fig-cap: Planets

```

# Questions

1. Which variables should be considered ordinal and which categorical?

### 2023-01-13 Emily Muther/body_image_trikafta.qmd
Update body_image_trikafta.qmd

@@
 # Libraries
+library(knitr)
+library(kableExtra)
 library(haven)
@@
 library(arsenal)
+library(psych)
+library(ggcorrplot)
 # Import
@@
 df$`Weight/Body Adherence` = factor(df$`Weight/Body Adherence`,ordered = T)
+df$`Do Not Want Trikafta` = factor(df$DoNotWantTrikafta,levels = 1:6,
+                                   labels = c("Strongly Agree","Slightly Agree",
+                                              "Neutral","Slightly Disagree",
+                                              "Strongly Disagree",NA))
+df$`Do Not Want Trikafta` = fct_rev(df$`Do Not Want Trikafta`)
+df$`Do Not Want Trikafta` = factor(df$`Do Not Want Trikafta`,ordered = T)
+df$`Continue Trikafta` = factor(df$ContinueTrikafta,levels = 0:6,
+                                labels = c(NA,"Strongly Agree",
+                                           "Slightly Agree","Neutral",  
+                                           "Slightly Disagree","Strongly Disagree",NA))
+df$`Continue Trikafta` = fct_rev(df$`Continue Trikafta`)
+df$`Continue Trikafta` = factor(df$`Continue Trikafta`,ordered = T)
 ```
@@
 #| tbl-cap: Comparisons by FRS Category
-t1 = tableby(`FRS Category`~kwt(`Percent Adherence`)+`Doses Missed` +
-               kwt(`Weight/Body Adherence`),data = df,
-             control=tableby.control(numeric.stats=c("Nmiss","median","q1q3","range"),
-                                     numeric.test="kwt"))
+rh = "~kwt(`Percent Adherence`)+`Doses Missed`+kwt(`Weight/Body Adherence`)+kwt(`Do Not Want Trikafta`)+kwt(`Continue Trikafta`)"
+t1 = tableby(as.formula(paste0("`FRS Category`",rh)),data = df,
+             control=tableby.control(numeric.stats=c("Nmiss","median","q1q3","range"),numeric.test="kwt"))
 summary(t1,pfootnote = T)
@@
 
+```{r results='asis'}
+#| label: tbl-2
+#| tbl-cap: Spearman's Rho
+x = df %>% select(BE_APPEARANCE_MEAN:BE_ATTRIBUTION_MEAN) %>% data.matrix(.)
+y = df %>% select(`Percent Adherence`:`Continue Trikafta`) %>% data.matrix(.)
+r = corr.test(x,y,adjust = "fdr")
+kable(r$r,digits=3)
+```
+
+```{r results='asis'}
+#| label: tbl-3
+#| tbl-cap: Spearman's Rho FDR-Adjusted p-values
+kable(r$p,digits=3)
+```
+
 ```{r}
-#| label: fig-planets
-#| fig-cap: Planets
-
+#| label: fig-heat
+#| fig-cap: Correlation Heatmap
+ggcorrplot(r$r,p.mat = r$p.adj)
 ```

### 2023-01-13 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
Update pre_post_trikafta_prospective_analysis.Rmd

@@
   })
+df$`Any Modulator` = rowSums(df[,c("Kalydeco.1","orkambi.1","tezacaftor.ivacaftor..symdeko..1")]) > 0
+df$`Any Modulator` = factor(df$`Any Modulator`,levels = c(F,T),labels = c("No","Yes"))
 # Genotype
@@
 newlabels["trikafta_duration"] = "Modulator Duration"
+newlabels["Data.Access.Group"] = "Site"
 newlabels[dxa_vars] = 
@@
                Sex..Male.1.+BMIz+FEV1..+FVC....+trikafta_duration+
-               Genotype+Pancreatic.Status+Kalydeco+Orkambi+Symdeko,
+               Genotype+Pancreatic.Status+Kalydeco+Orkambi+Symdeko+`Any Modulator`+Data.Access.Group,
              control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),
@@
   summarise_at(vars(Weight..kg.,c.peptide.0.minute,`C Peptide Index`,
-                    `Adapted Matsuda`,`C Peptide ODI`,iAUC_ratio,homa2B:homa2IR),
+                    `Adapted Matsuda`,`C Peptide ODI`,iAUC_ratio,homa2B:homa2IR,
+                    Fat.Mass.Index..FMI...),
                .funs = diff) %>%
@@
          "Change in HOMA2 %B" = homa2B, "Change in HOMA2 %S" = homa2S,
-         "Change in HOMA2 IR" = homa2IR)
+         "Change in HOMA2 IR" = homa2IR,
+         "Change in FMI"=Fat.Mass.Index..FMI...)
 
-M = cor(change_df[,2:ncol(change_df)])
+M = cor(change_df[,2:ncol(change_df)],use = "pairwise.complete.obs")
 testRes = cor.mtest(change_df[,2:ncol(change_df)])
@@
 ```{r results="asis"}
-plot_vars = c("Weight..kg.","weightz","BMIz","Fat.Mass.Index..FMI...","Fat.Free.Mass.Index...FFMI.","`Glucose iAUC`","c.peptide.0.minute","`C Peptide iAUC`","iAUC_ratio","`C Peptide ODI`","`Adapted Matsuda`","homa2IR","FEV1..","FVC....")
+plot_vars = c("Weight..kg.","weightz","BMIz","Fat.Mass.Index..FMI...","Fat.Free.Mass.Index...FFMI.","`Glucose iAUC`","c.peptide.0.minute","`C Peptide Index`","`C Peptide iAUC`","iAUC_ratio","`C Peptide ODI`","`Adapted Matsuda`","homa2IR","FEV1..","FVC....")
 # Loop through variables and plot

### 2023-01-13 Emily Muther/mental_health_screening.qmd
Update mental_health_screening.qmd

--- 2023-01-05 Emily Muther/mental_health_screening.qmd
@@
 df.rename({'PAT_MRN_ID':'ID','CONTACT_DATE':'Date','AGE AT CONTACT ':'Age','GAD-7 Total Score':'GAD-7 Total','GAD-7 Totals':'GAD-7 Group','PHQ-9 Total Score (calculated)':'PHQ-9 Total','PHQ-9 Totals':'PHQ-9 Group','Score':'PSC-17 Total','PSC Totals':'PSC-17 Group'},inplace=True,axis=1)
+# Pull in Trikafta data
+eti = pd.read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/Copy of Trikafta Earliest and Most Recent Dates for Patients Over 12.xlsx",usecols=["MRN","Earliest Trikafta in PortCF"])
+df = pd.merge(df,eti,left_on="ID",right_on="MRN")
+# Pre/post
+df["Post-Trikafta"] = df['Date'] > df['Earliest Trikafta in PortCF']
+df["Post-Trikafta"].replace({False:"Pre",True:"Post"},inplace=True)
+# Sort
+df["Date"] = pd.to_datetime(df["Date"])
+df.sort_values(by=["ID","Date"],inplace=True)
+# Write to CSV
+df.drop(["MRN"],axis=1,inplace=True)
+df.to_csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Cleaned/cf_mental_health_screening.csv",index=False)
 ```
@@
 
-# Profiling
+# Pre- vs. Post-ETI
 
-```{python}
-from pandas_profiling import ProfileReport
-profile = ProfileReport(df, title="Data Profiling Report")
-profile.to_notebook_iframe()
+```{r}
+#| include: false
+library(tidyverse)
+library(arsenal)
 ```
+
+```{r results="asis"}
+df = read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Cleaned/cf_mental_health_screening.csv",na.strings ="")
+
+means = df %>% group_by(ID,Post.Trikafta) %>% 
+  summarise(`GAD 7 Total` = mean(GAD.7.Total,na.rm = T),
+            `PHQ 9 Total` = mean(PHQ.9.Total,na.rm = T),
+            `PSC 17 Total` = mean(PSC.17.Total,na.rm = T),.groups = "drop")
+
+means$`GAD 7 Group` = cut(means$`GAD 7 Total`,c(-Inf,5,10,15,Inf),right = F,
+                          labels = c("0-4(Minimal)","5-9(Mild)",
+                          "10-14(Moderate)","15-21(Severe)"))
+
+means$`PHQ 9 Group` = cut(means$`PHQ 9 Total`,c(-Inf,5,10,15,20,Inf),right=F,
+                          labels = c("0-4(None)","5-9(Mild)","10-14(Moderate)",
+                          "15-19(Moderately Severe)","20-27(Severe)"))
+
+means$`PSC 17 Group` = cut(means$`PSC 17 Total`,c(-Inf,15,Inf),right = F,
+                          labels = c("0-14(Low Risk)","15-34(High Risk)"))
+
+# Remove missing and drop levels
+keep = means %>% group_by(ID) %>% summarise(n=n()) %>% filter(n==2) %>% pull(ID)
+means = means %>% filter(ID %in% keep)
+means$`GAD 7 Group` = droplevels(means$`GAD 7 Group`)
+means$`PHQ 9 Group` = droplevels(means$`PHQ 9 Group`)
+means$`PSC 17 Group` = droplevels(means$`PSC 17 Group`)    
+# Write 
+write.csv(means,file="/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Cleaned/means.csv",row.names=F)
+
+p <- paired(Post.Trikafta ~ `GAD 7 Total`+`GAD 7 Group`+
+  `PHQ 9 Total`+`PHQ 9 Group`+
+  `PSC 17 Total`+`PSC 17 Group`, 
+            data = means, id = ID,numeric.test = "signed.rank",signed.rank.exact = F,
+            numeric.stats = c("Nmiss","median","q1q3","range"))
+summary(p,pfootnote=T)
+```

### 2023-01-19 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
Update pre_post_trikafta_prospective_analysis.Rmd

@@
 library(corrplot)
+library(DiagrammeR)
 library(knitr)
@@
 
+# Flowchart
+
+```{r}
+nums = df %>% group_by(Event.Name) %>% count(Group)
+ndf <- create_node_df(n=6,
+                        type = paste(nums$Event.Name,nums$Group),
+                        label = paste(nums$Event.Name,nums$Group),
+                        value=nums$n,
+                      shape="rectangle",fixedsize=F)
+edges = df %>% group_by(Study.ID) %>% 
+  summarise(change = paste("Visit 1",Group[Event.Name=="Visit 1"],"->",
+                           "Visit 2",Group[Event.Name=="Visit 2"])) %>%
+  count(change)
+edges$from = match(sapply(strsplit(edges$change," -> "),"[",1),ndf$type)
+edges$to = match(sapply(strsplit(edges$change," -> "),"[",2),ndf$type)
+edf <-
+  create_edge_df(
+    from = edges$from,
+    to = edges$to,
+    rel = "a",
+    label=edges$n)
+
+graph <-
+  create_graph(
+    nodes_df = ndf,
+    edges_df = edf,
+    attr_theme = "lr")
+render_graph(graph)
+```
+
 # Notes

### 2023-02-03 StatNet/cgm_fda.qmd
Update cgm_fda.qmd

@@
     smaller: true
-editor: visual
+editor: source
 ---
@@
 # Data import
-sleep = read.csv("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Raw/sleep.csv",
+sleep = read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/CGM and FDA/Data_Raw/sleep.csv",
                  na.strings = "")
@@
 # CGMs
-# cleandata("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Raw/CGM",
-#           "~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",
+# cleandata("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/CGM and FDA/Data_Raw/CGM",
+#           "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",
 #           removegaps = F)
-cgms = list.files("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",
+cgms = list.files("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",
                   full.names = T)
@@
 # Save
-save(ydf,file = "~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/ydf.Rdata")
+save(ydf,file = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/ydf.Rdata")
 ```
@@
 #| include: false
-load("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/fpca_sc_fit.Rdata")
+load("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/fpca_sc_fit.Rdata")
 ```

### 2023-02-06 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
separate cleaning and analysis files Colleen Wood

@@
 knitr::opts_chunk$set(echo = FALSE)
+library(tidyverse)
 library(arsenal)
-library(tidyverse)
-library(lubridate)
-library(childsds)
-library(rspiro)
-source("C:/Users/timvigers/Documents/GitHub/CF/get_genotype_class_severity.R")
-t1_variables <- c(
-  "CFF ID", "Site", "Sex", "Race", "Genotypes1", "Genotypes2", "Pancreatic Status",
-  "G-tube in past 12 months","Time", "Age",
-  "BMI Percentile", "Height", "Weight", "FEV1", "FVC", "HbA1c"
-)
-```
-
-```{r include=FALSE}
-# CHCO
-chco_a1c <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_glycemic.csv", show_col_types = F)
-chco_a1c$Date <- parse_date_time(chco_a1c$Date, orders = "mdy")
-chco_bmi <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv", show_col_types = F)
-chco_bmi$Date <- parse_date_time(chco_bmi$Date, orders = "mdy")
-chco <- full_join(chco_a1c, chco_bmi)
-# Fill CFF ID by MRN
-chco <- chco %>%
-  group_by(MRN) %>%
-  fill(`CFF ID`, .direction = "downup") %>%
-  ungroup() %>%
-  filter(!is.na(`CFF ID`)) %>%
-  select(-MRN)
-# Add demographics, filter
-chco_reg <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_registry.csv", show_col_types = F)
-chco_reg$Start <- parse_date_time(chco_reg$Start, orders = "mdy")
-chco_reg$DOB <- parse_date_time(chco_reg$DOB, orders = "mdy")
-keep <- chco_reg %>%
-  filter(!is.na(`CFRD Diagnosis Date`)) %>%
-  pull(`CFF ID`)
-exclude <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv", show_col_types = F)
-exclude <- exclude %>%
-  filter(!is.na(`Remove (see notes)`)) %>%
-  pull(MRN)
-exclude <- chco_reg$`CFF ID`[match(exclude, chco_reg$MRN)]
-keep <- setdiff(keep, exclude)
-chco <- left_join(chco, chco_reg) %>% filter(`CFF ID` %in% keep)
-# Add missing variables
-chco$Age <- round(as.numeric(difftime(chco$Date, chco$DOB, units = "days")) / 365.25, 1)
-chco$Time <- as.numeric(difftime(chco$Date, chco$Start, units = "days"))
-chco$Site <- "CHCO"
-# BMI calculations
-chco$age_for_bmi <- ifelse(chco$Age < 20, chco$Age, 20)
-chco$`BMI Percentile` <- round(sds(chco$BMI,
-  age = chco$age_for_bmi, item = "bmi", type = "perc",
-  sex = chco$Sex, male = "Male", female = "Female",
-  ref = cdc.ref
-) * 100, 2)
-```
-
-```{r include=FALSE}
-# Montana
-mon_bmi <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_bmi_pft.csv", show_col_types = F)
-mon_bmi$Date <- parse_date_time(mon_bmi$Date, orders = "mdy")
-mon_a1c <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv", show_col_types = F)
-mon_a1c$Date <- parse_date_time(mon_a1c$Date, orders = "mdy")
-mon_a1c$DOB <- parse_date_time(mon_a1c$DOB, orders = "mdy")
-mon_a1c$HbA1c <- as.numeric(mon_a1c$HbA1c)
-# Get start date and CFRD status
-mon_starts <- mon_a1c %>%
-  select(`CFF ID`, `CFRD Status`, Start) %>%
-  distinct() %>%
-  filter(`CFRD Status` == "CFRD with or without fasting hyperglycemia")
-mon_starts$Start <- parse_date_time(mon_starts$Start, orders = "mdy")
-# Remove unnecessary columns
-mon_a1c <- mon_a1c %>% select(`CFF ID`, DOB, Date, HbA1c, `OGTT Fasting`, `OGTT Two Hour`)
-# Merge and remove those without CFRD or start date
-mon <- full_join(mon_bmi, mon_a1c)
-mon <- mon %>% filter(`CFF ID` %in% mon_starts$`CFF ID`)
-mon <- left_join(mon, mon_starts %>% select(`CFF ID`, Start))
-# Add demographics
-mon_demo <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_demo.csv", show_col_types = F)
-mon_demo <- mon_demo %>% select(`CFF ID`, Race, Sex)
-mon <- left_join(mon, mon_demo)
-# Add genotypes
-mon_geno <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_genotypes.csv", show_col_types = F)
-mon <- left_join(mon, mon_geno)
-# Add pancreatic and g tube status
-mon_pan <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_pancreatic_insufficient.csv", show_col_types = F)
-mon_pan <- mon_pan %>%
-  select(-Date) %>%
-  distinct()
-mon_pan$`Pancreatic Status` <- "Insufficient"
-mon_gtube <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_gtubes.csv", show_col_types = F)
-mon_gtube <- mon_gtube %>%
-  select(-Feeding, -Date) %>%
-  distinct()
-mon_gtube$`G-tube in past 12 months` <- "Yes"
-mon <- left_join(mon, mon_pan)
-mon <- left_join(mon, mon_gtube)
-# Add missing variables
-mon <- mon %>%
-  group_by(`CFF ID`) %>%
-  fill(Start, DOB, .direction = "downup")
-mon$Time <- as.numeric(difftime(mon$Date, mon$Start, units = "days"))
-mon$Age <- round(as.numeric(difftime(mon$Date, mon$DOB)) / 365.25, 1)
-mon$Site <- "Montana"
-# Fill in some missing
-mon$`Pancreatic Status`[is.na(mon$`Pancreatic Status`)] <- "Sufficient"
-# BMI calculations
-mon$age_for_bmi <- ifelse(mon$Age < 20, mon$Age, 20)
-mon$`BMI Percentile` <- round(sds(mon$BMI,
-  age = mon$age_for_bmi, item = "bmi", type = "perc",
-  sex = mon$Sex, male = "Male", female = "Female",
-  ref = cdc.ref
-) * 100, 2)
-```
-
-```{r include=FALSE}
-# UM
-um_a1c <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_a1cs.csv", show_col_types = F)
-um_a1c$Age <- round(um_a1c$Age / 365.25, 1)
-um_bmi <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_bmi.csv", show_col_types = F)
-um_bmi$Age <- round(um_bmi$Age / 365.25, 1)
-um_ogtt <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_ogtt.csv", show_col_types = F)
-um_ogtt$Age <- round(um_ogtt$Age / 365.25, 1)
-um_pft <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_pfts.csv", show_col_types = F)
-um_pft$Age <- round(um_pft$Age / 365.25, 1)
-um_race <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_race.csv", show_col_types = F)
-um_demo <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_demo.csv", show_col_types = F)
-keep <- um_demo %>%
-  filter(`CFRD Status` == "CFRD with or without fasting hyperglycemia") %>%
-  pull(`CFF ID`)
-um_demo <- um_demo %>%
-  select(-`CFRD Status`) %>%
-  rename(`G-tube in past 12 months` = `G tube in past 12 months`)
-# Join
-um <- reduce(list(um_a1c, um_bmi, um_ogtt, um_pft), full_join)
-um <- left_join(um, um_race)
-um <- left_join(um, um_demo)
-um <- um %>%
-  filter(`CFF ID` %in% keep, !is.na(Time)) %>%
-  rename(Site = "site") %>%
-  arrange(`CFF ID`, Time)
-# BMI calculations
-um$age_for_bmi <- ifelse(um$Age < 20, um$Age, 20)
-um$`BMI Percentile` <- round(sds(um$BMI,
-  age = um$age_for_bmi, item = "bmi", type = "perc",
-  sex = um$Sex, male = "Male", female = "Female",
-  ref = cdc.ref
-) * 100, 2)
-```
-
-```{r include=FALSE}
-# UW
-uw_bmi <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_bmi.csv", show_col_types = F)
-uw_bmi$Date <- parse_date_time(uw_bmi$Date, orders = "mdy")
-uw_glyc <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_glycemic.csv", show_col_types = F)
-uw_glyc$Date <- parse_date_time(uw_glyc$Date, orders = "mdy")
-uw_pft <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pft.csv", show_col_types = F)
-uw_pft$Date <- parse_date_time(uw_pft$Date, orders = "mdy")
-uw_demo <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_demo.csv", show_col_types = F)
-uw_demo <- uw_demo %>% select(`CFF ID`, Race, Sex)
-# Get CFRD status and start date
-uw_start <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_mods.csv", show_col_types = F)
-uw_start$`Date of Encounter` <- parse_date_time(uw_start$`Date of Encounter`, orders = "mdy")
-uw_start <- uw_start %>%
-  arrange(`CFF ID`, `Date of Encounter`) %>%
-  group_by(`CFF ID`) %>%
-  filter(row_number() == 1) %>%
-  select(`CFF ID`, `Date of Encounter`) %>%
-  rename(Start = `Date of Encounter`)
-uw_cfrd <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_cfrd.csv", show_col_types = F)
-uw_cfrd <- uw_cfrd %>%
-  filter(`CFRD Status` == "CFRD with or without fasting hyperglycemia") %>%
-  select(`CFF ID`) %>%
-  distinct() %>%
-  pull(`CFF ID`)
-keep <- unique(uw_cfrd, uw_start$`CFF ID`)
-# Add genotypes
-uw_geno <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_genotypes.csv", show_col_types = F)
-# Pancreatic status
-uw_pan <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pancreatic.csv", show_col_types = F)
-uw_pan <- uw_pan %>%
-  select(-Date) %>%
-  distinct() %>%
-  mutate(`Pancreatic Status` = "Insufficient")
-# G tubes
-uw_gtube <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_gtubes.csv", show_col_types = F)
-uw_gtube <- uw_gtube %>%
-  select(`CFF ID`) %>%
-  distinct() %>%
-  mutate(`G-tube in past 12 months` = "Yes")
-# Merge
-uw <- reduce(list(uw_bmi, uw_glyc, uw_pft), full_join)
-uw <- left_join(uw, uw_demo)
-uw <- left_join(uw, uw_geno)
-uw <- left_join(uw, uw_pan)
-uw <- left_join(uw, uw_gtube)
-uw <- left_join(uw, uw_start) %>% filter(`CFF ID` %in% keep)
-# Other variables (no weight for this site)
-uw$Time <- as.numeric(difftime(uw$Date, uw$Start, units = "days"))
-uw$Site <- "UW"
-uw$Weight <- NA
-# BMI calculations
-uw$age_for_bmi <- ifelse(uw$Age < 20, uw$Age, 20)
-uw$`BMI Percentile 2` <- round(sds(uw$BMI,
-  age = uw$age_for_bmi, item = "bmi", type = "perc",
-  sex = uw$Sex, male = "Male", female = "Female",
-  ref = cdc.ref
-) * 100, 2)
-uw$`BMI Percentile` <- coalesce(uw$`BMI Percentile`, uw$`BMI Percentile 2`)
-```
-
-```{r include=FALSE}
-# Put everything together
-chco <- chco %>% select(all_of(t1_variables))
-mon <- mon %>% select(all_of(t1_variables))
-um <- um %>% select(all_of(t1_variables))
-uw <- uw %>% select(all_of(t1_variables))
-df <- bind_rows(chco, mon,um, uw) %>% arrange(`CFF ID`, Time)
-# Age group
-df$`Age Group` <- cut(df$Age,
-  breaks = c(-Inf, 20, Inf),
-  labels = c("Pedatric", "Adult"), right = F
-)
-# Race and sex
-df$Race <- factor(df$Race,
-  levels = c(
-    "Black or African American", "Caucasian",
-    "Other", "Some other race", "Two or more races", "White"
-  ),
-  labels = c(
-    "Black", "Caucasian",
-    "Other", "Other", "Other", "Caucasian"
-  )
-)
-df$Sex <- factor(df$Sex, levels = c("Male", "Female"))
-# FEV calculations
-age <- df$Age
-height <- df$Height / 100
-race <- df$Race
-old <- c("Caucasian", "Black", "Other")
-new <- c(1, 2, 5)
-race <- new[match(race, old)]
-gender <- as.numeric(df$Sex)
-df$ppFEV1 <- pctpred_GLI(age, height, gender = gender, ethnicity = race, FEV1 = df$FEV1)
-df$ppFVC <- pctpred_GLI(age, height, gender = gender, ethnicity = race, FVC = df$FVC)
-# Genotype severity
-df$Genotypes1 <- get_cf_genotype_class_severity(df$Genotypes1)
-df$Genotypes2 <- get_cf_genotype_class_severity(df$Genotypes2)
-df <- df %>% unite(Genotype, Genotypes1, Genotypes2)
-df$Genotype <- factor(df$Genotype,
-  levels = c(
-    "Mild_Severe", "Severe_Mild", "Severe_Severe",
-    "Severe_Unknown", "Unknown_Unknown"
-  ),
-  labels = c("Low", "Low", "High", "Unknown", "Unknown")
-)
-# Other little cleaning things
-df <- df %>% filter(Age > 0, !is.na(Time))
-df$`G-tube in past 12 months`[is.na(df$`G-tube in past 12 months`)] <- "No"
-# Pre-/post-modulator
-df$`Pre or Post` <- factor(sign(df$Time), levels = -1:1, labels = c("Pre", "Pre", "Post"))
-# Add insulin information
-chco_ins <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv", show_col_types = F)
-chco_ins$`CFF ID` <- chco_a1c$`CFF ID`[match(chco_ins$MRN, chco_a1c$MRN)]
-chco_ins <- chco_ins %>%
-  rename(TDD = `Total Daily Dose of insulin (units)`,
-         `Long acting insulin dose` = `Long acting or total basal daily dose (units)`) %>%
-  select(`CFF ID`, `Pre or Post`,TDD,`Long acting insulin dose`) %>%
-  filter(!is.na(`CFF ID`))
-mon_ins <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_insulin.csv", show_col_types = F)
-mon_ins <- mon_ins %>%
-  rename(TDD = `Total Daily Dose of insulin (units)`,
-         `Long acting insulin dose` = `Long acting or total basal daily dose (units)`) %>%
-  select(`CFF ID`, `Pre or Post`, TDD,`Long acting insulin dose`)
-uw_ins <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_insulin.csv", show_col_types = F)
-uw_ins <- uw_ins %>%
-  rename(TDD = `Total Daily Dose of insulin (units)`,
-         `Long acting insulin dose` = `Long acting or total basal daily dose (units)`) %>%
-  select(`CFF ID`, `Pre or Post`, TDD,`Long acting insulin dose`)
-insulin <- bind_rows(chco_ins, mon_ins, uw_ins)
-# Join
-df <- left_join(df, insulin)
-df$`Pre or Post` <- factor(df$`Pre or Post`, levels = c("Pre", "Post"))
-# Get pre value closest to start, post value closest to 1 year from start
-df$select_time <- df$Time
-df$select_time[df$`Pre or Post` == "Post"] <- 365 - df$select_time[df$`Pre or Post` == "Post"]
-df <- df %>%
-  arrange(`CFF ID`, `Pre or Post`, select_time) %>%
-  group_by(`CFF ID`, `Pre or Post`) %>%
-  summarise(
-    `Num. A1c Values` = sum(!is.na(HbA1c)),
-    across(everything(), ~ last(na.omit(.x))),
-    .groups = "keep"
-  ) %>%
-  filter(!is.na(HbA1c))
-keep <- df %>%
-  group_by(`CFF ID`) %>%
-  summarise(n = n()) %>%
-  filter(n == 2) %>%
-  pull(`CFF ID`)
-df <- df %>% filter(`CFF ID` %in% keep)
-```
-
-```{r include=FALSE}
-cgm_vars <- c(
-  "average_sensor", "min_sensor", "max_sensor", 
-  "percent_time_over_180","percent_time_70_180",
-  "percent_time_over_120",
-  "percent_time_over_140", "percent_time_over_200", 
-  "excursions_over_140","excursions_over_200", 
-  "percent_time_under_60", "percent_time_under_70",
-  "average_auc_per_day", "standard_deviation", "cv","r_mage"
-)
-day_cgm = c("daytime_avg_sensor_glucose",	"daytime_min_sensor_glucose",	
-            "daytime_max_sensor_glucose",	"percent_time_over_180_day",
-            "percent_time_70_180_day","percent_time_over_200_day",
-            "percent_time_under_60_day","percent_time_under_70_day","daytime_sd"
-            )
-night_cgm = c("nighttime_avg_sens_glucose",	"nighttime_min_sens_glucose",	
-            "nighttime_max_sens_glucose",	"percent_time_over_180_night",
-            "percent_time_70_180_night","percent_time_over_200_night",
-            "percent_time_under_60_night","percent_time_under_70_night","nighttime_sd"
-            )
-# Add CGM information
-# CHCO
-chco_cgm <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_cgm.csv", show_col_types = F)
-chco_cgm$`Pre or Post` <- factor(sapply(str_split(chco_cgm$subject_id, "_"), "[", 2))
-levels(chco_cgm$`Pre or Post`) <- c("Post", "Pre", "Pre")
-chco_cgm$`CFF ID` <- sapply(str_split(chco_cgm$subject_id, "_"), "[", 1)
-chco_cgm <- chco_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(c(cgm_vars,day_cgm,night_cgm)))
-# Montana
-mon_cgm <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_cgm.csv", show_col_types = F)
-mon_cgm$`Pre or Post` <- factor(sapply(str_split(mon_cgm$subject_id, "_"), "[", 2))
-levels(mon_cgm$`Pre or Post`) <- c("Post", "Post", "Pre", "Pre")
-mon_cgm$`CFF ID` <- sapply(str_split(mon_cgm$subject_id, "_"), "[", 1)
-mon_cgm <- mon_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(c(cgm_vars,day_cgm,night_cgm)))
-# UW
-uw_cgm <- read_csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_cgm.csv", show_col_types = F)
-uw_cgm$`Pre or Post` <- factor(sapply(str_split(uw_cgm$subject_id, "_"), "[", 2))
-levels(uw_cgm$`Pre or Post`) <- c("Post", "Pre")
-uw_cgm$`CFF ID` <- sapply(str_split(uw_cgm$subject_id, "_"), "[", 1)
-uw_cgm <- uw_cgm %>% select(`CFF ID`, `Pre or Post`, all_of(c(cgm_vars,day_cgm,night_cgm)))
-# Put together
-cgm <- bind_rows(chco_cgm, mon_cgm, uw_cgm)
-# Add to DF
-cgm$`CFF ID` <- as.numeric(cgm$`CFF ID`)
-df <- left_join(df, cgm)
-write.csv(df,file = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/analysis_data.csv",row.names = F,na="")
+load("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/analysis_dataset.rdata")
 ```

### 2023-02-06 Colleen Wood/Pre-post Trikafta/clean_and_merge.R

@@ -1,0 +1,335 @@
+library(tidyverse)
+library(lubridate)
+library(childsds)
+library(rspiro)
+source("/Users/timvigers/GitHub/CF/genotype_class_severity.R")
+t1_variables <- c(
+  "CFF ID", "Site", "Sex", "Race", "Genotypes1", "Genotypes2", "Pancreatic Status",
+  "G-tube in past 12 months", "Time", "Age",
+  "BMI Percentile", "Height", "Weight", "FEV1", "FVC", "HbA1c"
+)
+# CHCO
+chco_a1c <- read_csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_glycemic.csv", show_col_types = F)
+chco_a1c$Date <- parse_date_time(chco_a1c$Date, orders = "mdy")
+chco_bmi <- read_csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv", show_col_types = F)
+chco_bmi$Date <- parse_date_time(chco_bmi$Date, orders = "mdy")
+chco <- full_join(chco_a1c, chco_bmi)
+# Fill CFF ID by MRN
+chco <- chco %>%
+  group_by(MRN) %>%
+  fill(`CFF ID`, .direction = "downup") %>%
+  ungroup() %>%
+  filter(!is.na(`CFF ID`)) %>%
+  select(-MRN)
+# Add demographics, filter
+chco_reg <- read_csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_registry.csv", show_col_types = F)
+chco_reg$Start <- parse_date_time(chco_reg$Start, orders = "mdy")
+chco_reg$DOB <- parse_date_time(chco_reg$DOB, orders = "mdy")
+keep <- chco_reg %>%
+  filter(!is.na(`CFRD Diagnosis Date`)) %>%
+  pull(`CFF ID`)
+exclude <- read_csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv", show_col_types = F)
[…]

### 2023-02-08 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd
tidyverse formatting

--- 2022-09-12 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
@@
 knitr::opts_chunk$set(echo = FALSE)
-knitr::opts_knit$set(root.dir = "/Users/timvigers/Documents/Work/CF/Edith Zemanick/Sweat Test")
+if (Sys.info()["sysname"] == "Windows") {
+  home_dir <- "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Test"
+  source("C:/Users/timvigers/Documents/GitHub/CF/get_genotype_class_severity.R")
+} else if (Sys.info()["sysname"] == "Linux") {
+  home_dir <- "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Edith Zemanick/Sweat Test"
   source("~/GitHub/CF/get_genotype_class_severity.R")
+} else if (Sys.info()["sysname"] == "Darwin") {
+  home_dir <- "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Edith Zemanick/Sweat Test"
+  source("~/GitHub/CF/genotype_class_severity.R")
+}
+knitr::opts_knit$set(root.dir = home_dir)
+rm(home_dir)
 ```
@@
 # Numeric variables
-num_vars = c("Amount","Na","K","Cl")
-num_vars = c(paste0(num_vars,"_Left"),paste0(num_vars,"_Right"))
-df[,num_vars] = suppressWarnings(lapply(df[,num_vars],as.numeric))
+num_vars <- c("Amount", "Na", "K", "Cl")
+num_vars <- c(paste0(num_vars, "_Left"), paste0(num_vars, "_Right"))
+df[, num_vars] <- suppressWarnings(lapply(df[, num_vars], as.numeric))
 # Filter per Maxie's code
@@
 # Genotype info
-df$geno1 <- get_cf_genotype_class_severity(df$Genotypes1)
-df$geno2 <- get_cf_genotype_class_severity(df$Genotypes2)
+## By severity
+df$geno1 <- cf_genotype_class_severity(df$Genotypes1)
+df$geno2 <- cf_genotype_class_severity(df$Genotypes2)
 df$geno1[df$Diagnosis == "Non-CF"] <- "Not CF"
@@
   levels = c(
-    "Severe_Severe","Mild_Severe", "Mild_Unknown", "Severe_Mild", "Mild_Mild","Not CF_Not CF",
-
-     "Severe_Unknown", "Unknown_Severe",
-    "Unknown_Unknown"
+    "Severe_Severe", "Mild_Severe", "Mild_Unknown", "Severe_Mild", "Mild_Mild",
+    "Not CF_Not CF", "Severe_Unknown", "Unknown_Severe", "Unknown_Unknown"
   ),
   labels = c(
-    "High","Low", "Low", "Low", "Low", "Not CF",  "Unknown",
+    "High", "Low", "Low", "Low", "Low", "Not CF", "Unknown",
     "Unknown", "Unknown"
@@
 )
+## By function
+df$geno1_class <- cf_genotype_class_severity(df$Genotypes1, out = "class")
+df$geno1_class[grep("508", df$Genotypes1)] <- "F508del"
+df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("I", "II", "III"), "MF")
+df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("IV", "V"), "RF")
+
+df$geno2_class <- cf_genotype_class_severity(df$Genotypes2, out = "class")
+df$geno2_class[grep("508", df$Genotypes2)] <- "F508del"
+df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("I", "II", "III"), "MF")
+df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("IV", "V"), "RF")
+
+df$geno_class <- factor(paste0(
+  pmin(df$geno1_class, df$geno2_class), "/",
+  pmax(df$geno1_class, df$geno2_class)
+))
 # Which arm?
@@
 # Difference between arms
-df$`Cl Diff. Between Left and Right Arm` = as.numeric(df$Cl_Left) - as.numeric(df$Cl_Right)
-df$`Na Diff. Between Left and Right Arm` = as.numeric(df$Na_Left) - as.numeric(df$Na_Right)
+df$`Cl Diff. Between Left and Right Arm` <- as.numeric(df$Cl_Left) - as.numeric(df$Cl_Right)
+df$`Na Diff. Between Left and Right Arm` <- as.numeric(df$Na_Left) - as.numeric(df$Na_Right)
 # Merge left and right values
@@
 # Sodium to chloride ratio
-df$`Na/Cl Ratio` = df$Na /df$Cl
+df$`Na/Cl Ratio` <- df$Na / df$Cl
 # Sort
@@
 # Table options (most variables are not normally distributed)
-mycontrols  <- tableby.control(numeric.test="kwt",numeric.stats=c("Nmiss", "median", "q1q3"))
-labels = list(Age.at.Test.in.Years = "Age (Years)", geno_risk = "Genotype Severity")
+mycontrols <- tableby.control(numeric.test = "kwt", numeric.stats = c("Nmiss", "median", "q1q3"))
+labels <- list(
+  Age.at.Test.in.Years = "Age (Years)",
+  geno_risk = "Genotype Severity", geno_class = "Specific Genotype"
+)
 ```
@@
   filter(row_number() == 1) # First measure for everyone
-t1 <- tableby(Diagnosis ~ Sex + Age.at.Test.in.Years + fe(geno_risk) + Amount + Cl + Na + K + `Na/Cl Ratio`,
-  data = t1_df,control = mycontrols,simulate.p.value=T
+t1 <- tableby(
+  Diagnosis ~ Sex + Age.at.Test.in.Years + fe(geno_risk) + fe(geno_class) +
+    Amount + Cl + Na + K + `Na/Cl Ratio`,
+  data = t1_df, control = mycontrols, simulate.p.value = T
 )
-summary(t1,pfootnote = T,
+summary(t1,
+  pfootnote = T,
   labelTranslations = labels
@@
 ```{r results='asis'}
-t1 <- tableby(Method ~ Sex + Age.at.Test.in.Years + Diagnosis + geno_risk + Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
-  data = t1_df,control = mycontrols,
+t1 <- tableby(
+  Method ~ Sex + Age.at.Test.in.Years + Diagnosis + fe(geno_risk) + fe(geno_class) +
+    Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` +
+    `Na Diff. Between Left and Right Arm`,
+  data = t1_df, control = mycontrols, simulate.p.value = T
 )
-summary(t1,pfootnote = T,
+summary(t1,
+  pfootnote = T,
   labelTranslations = labels
@@
 ```{r results='asis'}
-t1 <- tableby(Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
-  data = t1_df,control = mycontrols,
+t1 <- tableby(
+  Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` +
+    `Cl Diff. Between Left and Right Arm` +
+    `Na Diff. Between Left and Right Arm`,
+  data = t1_df, control = mycontrols,
 )
-summary(t1,pfootnote = T,
+summary(t1,
+  pfootnote = T,
   labelTranslations = labels
@@
 ```{r results='asis'}
-t1 <- tableby(Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
-  data = t1_df[t1_df$Method == "Gibson-Cooke",],control = mycontrols,
+t1 <- tableby(
+  Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` +
+    `Cl Diff. Between Left and Right Arm` +
+    `Na Diff. Between Left and Right Arm`,
+  data = t1_df[t1_df$Method == "Gibson-Cooke", ], control = mycontrols,
 )
-summary(t1,pfootnote = T,
+summary(t1,
+  pfootnote = T,
   labelTranslations = labels
@@
 ```{r results='asis'}
-t1 <- tableby(Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
-  data = t1_df[t1_df$Method == "Macroduct",],control = mycontrols,
+t1 <- tableby(
+  Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` +
+    `Cl Diff. Between Left and Right Arm` +
+    `Na Diff. Between Left and Right Arm`,
+  data = t1_df[t1_df$Method == "Macroduct", ], control = mycontrols,
 )
-summary(t1,pfootnote = T,
+summary(t1,
+  pfootnote = T,
   labelTranslations = labels
@@
 ```{r results='asis'}
-t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
-  data = t1_df,control = mycontrols,
+t1 <- tableby(
+  geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` +
+    `Cl Diff. Between Left and Right Arm` +
+    `Na Diff. Between Left and Right Arm`,
+  data = t1_df, control = mycontrols,
 )
-summary(t1,pfootnote = T,
+summary(t1,
+  pfootnote = T,
   labelTranslations = labels
@@
 t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
-  data = t1_df[t1_df$Diagnosis == "Cystic Fibrosis",],control = mycontrols
+  data = t1_df[t1_df$Diagnosis == "Cystic Fibrosis", ], control = mycontrols
 )
-summary(t1,pfootnote = T,
+summary(t1,
+  pfootnote = T,
   labelTranslations = labels
@@
 t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
-  data = t1_df[t1_df$Diagnosis == "CFTR-related disorder",],control = mycontrols
+  data = t1_df[t1_df$Diagnosis == "CFTR-related disorder", ], control = mycontrols
 )
-summary(t1,pfootnote = T,
+summary(t1,
+  pfootnote = T,
   labelTranslations = labels
@@
 t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
-  data = t1_df[t1_df$Diagnosis == "CRMS",],control = mycontrols
+  data = t1_df[t1_df$Diagnosis == "CRMS", ], control = mycontrols
 )
-summary(t1,pfootnote = T,
+summary(t1,
+  pfootnote = T,
   labelTranslations = labels
@@
 ```{r results='asis'}
-t1 <- tableby( ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
-  data = t1_df[t1_df$Diagnosis == "Non-CF",],control = mycontrols
+t1 <- tableby(~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+  data = t1_df[t1_df$Diagnosis == "Non-CF", ], control = mycontrols
 )
-summary(t1,pfootnote = T,
+summary(t1,
+  pfootnote = T,
   labelTranslations = labels
@@
 
+## By specific genotype
+
+```{r results='asis'}
+t1 <- tableby(
+  geno_class ~ Amount + Cl + Na + K + `Na/Cl Ratio` +
+    `Cl Diff. Between Left and Right Arm` +
+    `Na Diff. Between Left and Right Arm`,
+  data = t1_df, control = mycontrols,
+)
+summary(t1,
+  pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+## By specific genotype and diagnosis
+
+### Cystic fibrosis
+
+```{r results='asis'}
+t1 <- tableby(geno_class ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+  data = t1_df[t1_df$Diagnosis == "Cystic Fibrosis", ], control = mycontrols
+)
+summary(t1,
+  pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### CFTR-related disorder
+
+```{r results='asis'}
+t1 <- tableby(geno_class ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+  data = t1_df[t1_df$Diagnosis == "CFTR-related disorder", ], control = mycontrols
+)
+summary(t1,
+  pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### CRMS
+
+```{r results='asis'}
+t1 <- tableby(geno_class ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+  data = t1_df[t1_df$Diagnosis == "CRMS", ], control = mycontrols
+)
+summary(t1,
+  pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### Non-CF
+
+```{r results='asis'}
+t1 <- tableby(~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+  data = t1_df[t1_df$Diagnosis == "Non-CF", ], control = mycontrols
+)
+summary(t1,
+  pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### Plots
+
+```{r warning=FALSE}
+ggplot(df, aes(x = geno_class, y = Na)) +
+  geom_boxplot() +
+  theme_bw() +
+  xlab("Specific Genotype") +
+  ggtitle("Sweat Sodium")
+
+ggplot(df, aes(x = geno_class, y = K)) +
+  geom_boxplot() +
+  theme_bw() +
+  ggtitle("Sweat Potassium")
+
+ggplot(df, aes(x = geno_class, y = Cl)) +
+  geom_boxplot() +
+  theme_bw() +
+  xlab("Specific Genotype") +
+  ggtitle("Sweat Chloride")
+
+ggplot(df, aes(x = geno_class, y = `Na/Cl Ratio`)) +
+  geom_boxplot() +
+  theme_bw() +
+  xlab("Specific Genotype") +
+  ggtitle("Na/Cl Ratio")
+```
+
 # Between occasion
@@
 # Get first and last values for everyone with more than 1
-multiples = df %>% group_by(PatientID) %>% 
+multiples <- df %>%
+  group_by(PatientID) %>%
   filter(n() > 1) %>% # Multiple values only
-  filter(row_number()==1 | row_number()==n()) %>% # First and last
+  filter(row_number() == 1 | row_number() == n()) %>% # First and last
   mutate(time = row_number())
-multiples$time = factor(multiples$time,levels = 1:2,labels = c("First","Last"))
+multiples$time <- factor(multiples$time, levels = 1:2, labels = c("First", "Last"))
 # Paired testing options
-mycontrols  <- tableby.control(numeric.test="signed.rank",cat.test = "mcnemar",
-                               numeric.stats=c("Nmiss", "median", "q1q3"))
+mycontrols <- tableby.control(
+  numeric.test = "signed.rank", cat.test = "mcnemar",
+  numeric.stats = c("Nmiss", "median", "q1q3")
+)
 ```
@@
 ```{r results='asis'}
-p <- paired(time ~ Diagnosis +Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,signed.rank.exact = F,
-            data = multiples, id = PatientID,control = mycontrols)
-summary(p,pfootnote = T,labelTranslations = labels)
+p <- paired(
+  time ~ Diagnosis + Cl + Na + K + `Na/Cl Ratio` +
+    `Cl Diff. Between Left and Right Arm` +
+    `Na Diff. Between Left and Right Arm`,
+  signed.rank.exact = F, data = multiples,
+  id = PatientID, control = mycontrols
+)
+summary(p, pfootnote = T, labelTranslations = labels)
 ```
@@
 ```{r results='asis'}
-p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,signed.rank.exact = F,strata = Diagnosis,
-            data = multiples %>% filter(Diagnosis %in% c("Cystic Fibrosis","Non-CF")), 
-            id = PatientID,control = mycontrols)
-summary(p,pfootnote = T,labelTranslations = labels)
+p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+  signed.rank.exact = F, strata = Diagnosis,
+  data = multiples %>% filter(Diagnosis %in% c("Cystic Fibrosis", "Non-CF")),
+  id = PatientID, control = mycontrols
+)
+summary(p, pfootnote = T, labelTranslations = labels)
 ```
@@
 ```{r results='asis'}
-p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,signed.rank.exact = F,strata = geno_risk,
+p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+  signed.rank.exact = F, strata = geno_risk,
   data = multiples,
-            id = PatientID,control = mycontrols)
-summary(p,pfootnote = T,labelTranslations = labels)
+  id = PatientID, control = mycontrols
+)
+summary(p, pfootnote = T, labelTranslations = labels)
 ```
@@
 ```{r results='asis'}
-p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio`,signed.rank.exact = F,strata = geno_risk,
+p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio`,
+  signed.rank.exact = F, strata = geno_risk,
   data = multiples %>% filter(Diagnosis == "Cystic Fibrosis"),
-            id = PatientID,control = mycontrols)
-summary(p,pfootnote = T,labelTranslations = labels)
+  id = PatientID, control = mycontrols
+)
+summary(p, pfootnote = T, labelTranslations = labels)
 ```
@@
 ```{r results='asis'}
-p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio`,signed.rank.exact = F,strata = geno_risk,
+p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio`,
+  signed.rank.exact = F, strata = geno_risk,
   data = multiples %>% filter(Diagnosis == "Non-CF"),
-            id = PatientID,control = mycontrols)
-summary(p,pfootnote = T,labelTranslations = labels)
+  id = PatientID, control = mycontrols
+)
+summary(p, pfootnote = T, labelTranslations = labels)
+```
+
+## By specific genotype
+
+```{r results='asis'}
+p <- paired(
+  time ~ Cl + Na + K + `Na/Cl Ratio` +
+    `Cl Diff. Between Left and Right Arm` +
+    `Na Diff. Between Left and Right Arm`,
+  signed.rank.exact = F, strata = geno_class, numeric.test = "notest",
+  data = multiples, id = PatientID, control = mycontrols
+)
+summary(p, pfootnote = T, labelTranslations = labels)
+```
+
+## By genotype class and diagnosis
+
+Not enough paired observations in the CRMS or CFTR-related disorder groups.
+
+### Cystic fibrosis
+
+```{r results='asis'}
+p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio`,
+  signed.rank.exact = F, strata = geno_class,
+  data = multiples %>% filter(Diagnosis == "Cystic Fibrosis"),
+  id = PatientID, control = mycontrols
+)
+summary(p, pfootnote = T, labelTranslations = labels)
+```
+
+### Non-CF
+
+```{r results='asis'}
+p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio`,
+  signed.rank.exact = F, strata = geno_class,
+  data = multiples %>% filter(Diagnosis == "Non-CF"),
+  id = PatientID, control = mycontrols
+)
+summary(p, pfootnote = T, labelTranslations = labels)
 ```
@@
 # Three groups
-t1_df$Group = cut(t1_df$Cl,breaks = c(0,30,60,Inf),right = F,labels = c("Low","Medium","High"))
+t1_df$Group <- cut(t1_df$Cl, breaks = c(0, 30, 60, Inf), right = F, labels = c("Low", "Medium", "High"))
 # Low vs. medium or high
-t1_df$Low = t1_df$Group
-levels(t1_df$Low) = c("Low","Medium or High","Medium or High")
-t1_df$Low = relevel(t1_df$Low,ref = "Medium or High")
+t1_df$Low <- t1_df$Group
+levels(t1_df$Low) <- c("Low", "Medium or High", "Medium or High")
+t1_df$Low <- relevel(t1_df$Low, ref = "Medium or High")
 # High vs. medium or low
-t1_df$High = t1_df$Group
-levels(t1_df$High) = c("Low or Medium","Low or Medium","High")
-t1_df$High = relevel(t1_df$High,ref = "Low or Medium")
+t1_df$High <- t1_df$Group
+levels(t1_df$High) <- c("Low or Medium", "Low or Medium", "High")
+t1_df$High <- relevel(t1_df$High, ref = "Low or Medium")
 ```
@@
 ```{r}
-roc_low <- roc(response = t1_df$Low,predictor = t1_df$Na,quiet = T)
-coords_low = round(coords(roc_low,"best",best.method="youden"),3)
-roc_high <- roc(response = t1_df$High,predictor = t1_df$Na,quiet = T)
-coords_high = round(coords(roc_high,"best",best.method="youden"),3)
+roc_low <- roc(response = t1_df$Low, predictor = t1_df$Na, quiet = T)
+coords_low <- round(coords(roc_low, "best", best.method = "youden"), 3)
+roc_high <- roc(response = t1_df$High, predictor = t1_df$Na, quiet = T)
+coords_high <- round(coords(roc_high, "best", best.method = "youden"), 3)
 
-plot(roc_low,print.auc = TRUE, col = "blue") 
+plot(roc_low, print.auc = TRUE, col = "blue")
 plot(roc_high, print.auc = TRUE, col = "green", print.auc.y = .4, add = TRUE)
@@
 ```{r warning=FALSE}
-ggplot(t1_df[!is.na(t1_df$Group),],aes(Na,fill = Group)) + 
-  geom_histogram(alpha=0.3, position="identity",binwidth = 1) + 
-  geom_vline(xintercept = coords_low$threshold,color = "red") +
-  geom_vline(xintercept = coords_high$threshold,color = "red") +
-  ylab("Count") + theme_bw()
+ggplot(t1_df[!is.na(t1_df$Group), ], aes(Na, fill = Group)) +
+  geom_histogram(alpha = 0.3, position = "identity", binwidth = 1) +
+  geom_vline(xintercept = coords_low$threshold, color = "red") +
+  geom_vline(xintercept = coords_high$threshold, color = "red") +
+  ylab("Count") +
+  theme_bw()
 ```

### 2023-02-16 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd
just CF vs non-CF

@@
 
-# Chloride groups
+# ROC analyses
 
-We performed receiver operating characteristic (ROC) analysis to find a sweat sodium cutoff that distinguishes between chloride groups.
+We performed receiver operating characteristic (ROC) analysis to find a sweat sodium to sweat chloride cutoff that distinguishes between groups.
 Only the first sweat test from each participant was used.
-Because the groups are based on chloride value, Na/Cl ratio is correlated with the outcome so Na alone was used instead.
-Also, ROC analyses generally assume a binary outcome (i.e.
-looking for a cutoff to distinguish between two groups), so
 
-In the following results, "Low" indicates a sweat chloride under 30, "Medium" indicates a sweat chloride of 30 - 59, and "High" indicates a sweat chloride greater than or equal to 60.
+Due to low numbers of people with CFTR-related disorder or CRMS, this analysis focused on CF vs. non-CF until we receive additional data from NJH.
+
+## Diagnosis
+
+### All participants
 
 ```{r}
-# Three groups
-t1_df$Group <- cut(t1_df$Cl, breaks = c(0, 30, 60, Inf), right = F, labels = c("Low", "Medium", "High"))
-# Low vs. medium or high
-t1_df$Low <- t1_df$Group
-levels(t1_df$Low) <- c("Low", "Medium or High", "Medium or High")
-t1_df$Low <- relevel(t1_df$Low, ref = "Medium or High")
-# High vs. medium or low
-t1_df$High <- t1_df$Group
-levels(t1_df$High) <- c("Low or Medium", "Low or Medium", "High")
-t1_df$High <- relevel(t1_df$High, ref = "Low or Medium")
+# Exclude those with CFTR-related disorder or CRMS for now
+roc_df = t1_df %>% filter(Diagnosis %in% c("Cystic Fibrosis","Non-CF"))
+roc_df$Diagnosis = factor(roc_df$Diagnosis)
+roc_df$Diagnosis = relevel(roc_df$Diagnosis,ref = "Non-CF")
+
+roc_cf <- roc(response = roc_df$Diagnosis, predictor = roc_df$`Na/Cl Ratio`, quiet = T)
+coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)
+
+plot(roc_low, print.auc = T,print.thres=T)
 ```
 
-## ROC analysis
+The optimal Na/Cl threshold for distinguishing between CF and non-CF was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index, with and AUC of `r round(as.numeric(roc_cf$auc),3)`.
 
-```{r}
-roc_low <- roc(response = t1_df$Low, predictor = t1_df$Na, quiet = T)
-coords_low <- round(coords(roc_low, "best", best.method = "youden"), 3)
-roc_high <- roc(response = t1_df$High, predictor = t1_df$Na, quiet = T)
-coords_high <- round(coords(roc_high, "best", best.method = "youden"), 3)
-
-plot(roc_low, print.auc = TRUE, col = "blue")
-plot(roc_high, print.auc = TRUE, col = "green", print.auc.y = .4, add = TRUE)
-```
-
-The blue curve shows the ROC for distinguishing between low vs.
-medium or high, and the green is the ROC curve for distinguishing between low or medium vs.
-high.
-The optimal threshold for distinguishing between low vs.
-medium or high was `r coords_low$threshold` (sensitivity = `r coords_low$sensitivity`, specificity = `r coords_low$specificity`) based on the Youden index.
-The optimal threshold for distinguishing between high vs.
-low or medium was `r coords_high$threshold` (sensitivity = `r coords_high$sensitivity`, specificity = `r coords_high$specificity`) based on the Youden index.
-The mean AUC for the two curves is `r round(mean(c(roc_low$auc,roc_high$auc)),3)`.
-
-## Histogram by group
-
-Vertical red lines indicate the thresholds identified above.
-
-```{r warning=FALSE}
-ggplot(t1_df[!is.na(t1_df$Group), ], aes(Na, fill = Group)) +
-  geom_histogram(alpha = 0.3, position = "identity", binwidth = 1) +
-  geom_vline(xintercept = coords_low$threshold, color = "red") +
-  geom_vline(xintercept = coords_high$threshold, color = "red") +
-  ylab("Count") +
-  theme_bw()
-```
+###

### 2023-02-16 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd
better cutoff histogram

@@
 
-# ROC analyses
+# Cutoff analyses
 
@@
 
-Due to low numbers of people with CFTR-related disorder or CRMS, this analysis focused on CF vs. non-CF until we receive additional data from NJH.
-
 ## Diagnosis
 
-### All participants
+### ROC
 
 ```{r}
-# Exclude those with CFTR-related disorder or CRMS for now
-roc_df = t1_df %>% filter(Diagnosis %in% c("Cystic Fibrosis","Non-CF"))
-roc_df$Diagnosis = factor(roc_df$Diagnosis)
-roc_df$Diagnosis = relevel(roc_df$Diagnosis,ref = "Non-CF")
+# Re-format diagnosis column
+t1_df$`Combined Diagnosis` <- factor(t1_df$Diagnosis,
+                     levels = c("Non-CF","CFTR-related disorder","CRMS","Cystic Fibrosis"),
+                     labels = c("Non-CF","CFTR-related disorder or CRMS","CFTR-related disorder or CRMS","Cystic Fibrosis"),
+                     ordered = T)
 
-roc_cf <- roc(response = roc_df$Diagnosis, predictor = roc_df$`Na/Cl Ratio`, quiet = T)
+```
+
+```{r}
+roc_cf <- roc(response= t1_df$`Combined Diagnosis`,
+              predictor = t1_df$`Na/Cl Ratio`, 
+              levels = c("CFTR-related disorder or CRMS","Cystic Fibrosis"),
+              quiet = T)
 coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)
 
-plot(roc_low, print.auc = T,print.thres=T)
+roc_non_cf <- roc(response= t1_df$`Combined Diagnosis`,
+              predictor = t1_df$`Na/Cl Ratio`, 
+              levels = c("Non-CF","CFTR-related disorder or CRMS"),
+              quiet = T)
+coords_non_cf <- round(coords(roc_non_cf, "best", best.method = "youden"), 3)
+
+plot(roc_cf, print.auc = TRUE, col = "blue")
+plot(roc_non_cf, print.auc = TRUE, col = "green", print.auc.y = .4, add = TRUE)
 ```
 
-The optimal Na/Cl threshold for distinguishing between CF and non-CF was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index, with and AUC of `r round(as.numeric(roc_cf$auc),3)`.
+The blue curve shows the ROC for distinguishing between CFTR-related disorder or CRMS vs.
+CF, and the green is the ROC curve for distinguishing between non-CF vs.
+CFTR-related disorder or CRMS.
+The optimal threshold for distinguishing between CFTR-related disorder or CRMS vs.
+CF was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index.
+The optimal threshold for distinguishing between non-CF vs.
+CFTR-related disorder or CRMS was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index.
+The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.
 
-###
+### Histogram by group
+
+Vertical blue line indicates the threshold for CFTR-related disorder or CRMS vs.
+CF and green indicates the threshold for non-CF vs.
+CFTR-related disorder or CRMS.
+Na/Cl ratio was log-transformed for easier visualization.
+
+```{r warning=FALSE}
+ggplot(t1_df[!is.na(t1_df$`Combined Diagnosis`), ], 
+       aes(log(`Na/Cl Ratio`), fill = `Combined Diagnosis`)) +
+  geom_histogram(alpha = 0.3, position = "identity",bins = 50) +
+  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
+  geom_vline(xintercept = log(coords_non_cf$threshold), color = "green") +
+  ylab("Count") + xlab("ln(Na/Cl Ratio)") +
+  theme_bw()
+```

### 2023-02-16 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd
new ROC analyses done

@@
 
-## Diagnosis
+## Full cohort
 
-### ROC
+The full cohort analyses include all participants.
+
+### Diagnosis
+
+#### ROC
 
@@
 
-### Histogram by group
+#### Histogram by group
 
@@
 ```
+
+### Genotype
+
+Participants with unknown genotype risk were excluded from this analysis.
+
+#### ROC
+
+```{r}
+# Re-format diagnosis column
+t1_df$`Genotype Risk` <- factor(t1_df$geno_risk,
+                     levels = c("Not CF","Low","High","Unknown"),
+                     ordered = T)
+t1_df$`Genotype Risk`[t1_df$`Genotype Risk`=="Unknown"]= NA
+```
+
+```{r}
+roc_cf <- roc(response= t1_df$`Genotype Risk`,
+              predictor = t1_df$`Na/Cl Ratio`, 
+              levels = c("Low","High"),
+              quiet = T)
+coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)
+
+roc_non_cf <- roc(response= t1_df$`Genotype Risk`,
+              predictor = t1_df$`Na/Cl Ratio`, 
+              levels = c("Not CF","Low"),
+              quiet = T)
+coords_non_cf <- round(coords(roc_non_cf, "best", best.method = "youden"), 3)
+
+plot(roc_cf, print.auc = TRUE, col = "blue")
+plot(roc_non_cf, print.auc = TRUE, col = "green", print.auc.y = .4, add = TRUE)
+```
+
+The blue curve shows the ROC for distinguishing between low vs.
+high risk, and the green is the ROC curve for distinguishing between non-CF vs.
+low risk.
+The optimal threshold for distinguishing between low vs.
+high risk was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index.
+The optimal threshold for distinguishing between non-CF vs.
+low risk was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index.
+The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.
+
+#### Histogram by group
+
+Vertical blue line indicates the threshold for low vs. high risk and green indicates the threshold for non-CF vs. low risk. Na/Cl ratio was log-transformed for easier visualization.
+
+```{r warning=FALSE}
+ggplot(t1_df[!is.na(t1_df$`Genotype Risk`), ], 
+       aes(log(`Na/Cl Ratio`), fill = `Genotype Risk`)) +
+  geom_histogram(alpha = 0.3, position = "identity",bins = 50) +
+  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
+  geom_vline(xintercept = log(coords_non_cf$threshold), color = "green") +
+  ylab("Count") + xlab("ln(Na/Cl Ratio)") +
+  theme_bw()
+```
+
+## Limited cohort
+
+The limited cohort analyses only included participants with sweat chloride < 60.
+
+### Diagnosis
+
+#### ROC
+
+```{r}
+# Exclude those with Cl >= 60
+t1_df = t1_df %>% filter(Cl < 60)
+```
+
+```{r}
+roc_cf <- roc(response= t1_df$`Combined Diagnosis`,
+              predictor = t1_df$`Na/Cl Ratio`, 
+              levels = c("CFTR-related disorder or CRMS","Cystic Fibrosis"),
+              quiet = T)
+coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)
+
+roc_non_cf <- roc(response= t1_df$`Combined Diagnosis`,
+              predictor = t1_df$`Na/Cl Ratio`, 
+              levels = c("Non-CF","CFTR-related disorder or CRMS"),
+              quiet = T)
+coords_non_cf <- round(coords(roc_non_cf, "best", best.method = "youden"), 3)
+
+plot(roc_cf, print.auc = TRUE, col = "blue")
+plot(roc_non_cf, print.auc = TRUE, col = "green", print.auc.y = .4, add = TRUE)
+```
+
+The blue curve shows the ROC for distinguishing between CFTR-related disorder or CRMS vs.
+CF, and the green is the ROC curve for distinguishing between non-CF vs.
+CFTR-related disorder or CRMS.
+The optimal threshold for distinguishing between CFTR-related disorder or CRMS vs.
+CF was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index.
+The optimal threshold for distinguishing between non-CF vs.
+CFTR-related disorder or CRMS was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index.
+The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.
+
+#### Histogram by group
+
+Vertical blue line indicates the threshold for CFTR-related disorder or CRMS vs.
+CF and green indicates the threshold for non-CF vs.
+CFTR-related disorder or CRMS.
+Na/Cl ratio was log-transformed for easier visualization.
+
+```{r warning=FALSE}
+ggplot(t1_df[!is.na(t1_df$`Combined Diagnosis`), ], 
+       aes(log(`Na/Cl Ratio`), fill = `Combined Diagnosis`)) +
+  geom_histogram(alpha = 0.3, position = "identity",bins = 50) +
+  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
+  geom_vline(xintercept = log(coords_non_cf$threshold), color = "green") +
+  ylab("Count") + xlab("ln(Na/Cl Ratio)") +
+  theme_bw()
+```
+
+### Genotype
+
+Participants with unknown genotype risk were excluded from this analysis.
+
+#### ROC
+
+```{r}
+roc_cf <- roc(response= t1_df$`Genotype Risk`,
+              predictor = t1_df$`Na/Cl Ratio`, 
+              levels = c("Low","High"),
+              quiet = T)
+coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)
+
+roc_non_cf <- roc(response= t1_df$`Genotype Risk`,
+              predictor = t1_df$`Na/Cl Ratio`, 
+              levels = c("Not CF","Low"),
+              quiet = T)
+coords_non_cf <- round(coords(roc_non_cf, "best", best.method = "youden"), 3)
+
+plot(roc_cf, print.auc = TRUE, col = "blue")
+plot(roc_non_cf, print.auc = TRUE, col = "green", print.auc.y = .4, add = TRUE)
+```
+
+The blue curve shows the ROC for distinguishing between low vs.
+high risk, and the green is the ROC curve for distinguishing between non-CF vs.
+low risk.
+The optimal threshold for distinguishing between low vs.
+high risk was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index.
+The optimal threshold for distinguishing between non-CF vs.
+low risk was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index.
+The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.
+
+#### Histogram by group
+
+Vertical blue line indicates the threshold for low vs. high risk and green indicates the threshold for non-CF vs. low risk. Na/Cl ratio was log-transformed for easier visualization.
+
+```{r warning=FALSE}
+ggplot(t1_df[!is.na(t1_df$`Genotype Risk`), ], 
+       aes(log(`Na/Cl Ratio`), fill = `Genotype Risk`)) +
+  geom_histogram(alpha = 0.3, position = "identity",bins = 50) +
+  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
+  geom_vline(xintercept = log(coords_non_cf$threshold), color = "green") +
+  ylab("Count") + xlab("ln(Na/Cl Ratio)") +
+  theme_bw()
+```

### 2023-02-21 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd
flow chart finished

@@
 ```{r setup, include=FALSE}
+library(DiagrammeR)
 library(readxl)
@@
 df <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv", na.strings = "")
+n1 = nrow(df)
+n2 = length(unique(df$PatientID))
 # Remove exclusions from Angela
@@
 # Filter per Maxie's code
+exclude2 = c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793)
+no_dem = unique(df$PatientID[df$Demographics.Available!="TRUE" | is.na(df$Demographics.Available) | df$Sex==""])
 df <- df %>%
@@
     Age.at.Test.in.Years < 50 & Age.at.Test.in.Days >= 2,
-    !(Diagnosis == "Non-CF" & PatientID %in%
-      c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793)),
+    !(Diagnosis == "Non-CF" & PatientID %in% exclude2),
     Modulator == "None"
@@
 df$`Na/Cl Ratio` <- df$Na / df$Cl
+high_ratio = unique(which(df$`Na/Cl Ratio`>2))
+high_cl = unique(which(df$Cl>160))
+high_na = unique(which(df$Na>160))
+df = df[-unique(c(high_ratio,high_cl,high_na)),]
+# Sweat chloride group
+df$`Sweat Chloride Group` = cut(df$Cl,c(-Inf,30,60,Inf),right = F,
+                                labels = c("< 30","30-59",">60"))
 # Sort
@@
 )
+# First measure for everyone
+t1_df <- df %>%
+  group_by(PatientID) %>%
+  filter(row_number() == 1) 
+```
+
+# Flowchart
+
+```{r}
+grViz("digraph {
+
+# initiate graph
+graph [layout = dot]
+
+# global node settings
+node [shape = rectangle]
+
+# label nodes
+data1 [label = '@@1']
+data2 [label = '@@2']
+ineligible [label = '@@3']
+demo  [label = '@@4']
+highna  [label = '@@5']
+highcl  [label = '@@6']
+highnacl  [label = '@@7']
+review [label = '@@8']
+analysis [label = '@@9']
+
+# edge definitions with the node IDs
+data1  -> data2
+data2 -> ineligible
+ineligible -> {demo,highna,highcl,highnacl,review}
+data2  -> analysis
+highnacl -> analysis [style=invis]
+}
+[1]: paste0('Sweat tests performed \\n ','n = ',n1)
+[2]: paste0('Unique participants \\n ','n = ',n2)
+[3]: paste0('Ineligible \\n ','n = ',n2-nrow(t1_df))
+[4]: paste0('No demographics \\n ','n = ',length(no_dem))
+[5]: paste0('Na > 160 mmol/L \\n ','n = ',length(high_na))
+[6]: paste0('Cl > 160 mmol/L \\n ','n = ',length(high_cl))
+[7]: paste0('Na:Cl > 2 \\n ','n = ',length(high_ratio))
+[8]: paste0('Other \\n ','n = ',1)
+[9]: paste0('Analysis cohort \\n ','n = ',nrow(t1_df))
+")
 ```
@@
 ```{r results='asis'}
-t1_df <- df %>%
-  group_by(PatientID) %>%
-  filter(row_number() == 1) # First measure for everyone
 t1 <- tableby(

### 2023-02-21 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd
started simplifying tables

@@
 df <- df[!(df$PatientID %in% exclude), ]
-df$Diagnosis <- diagnoses$Diagnosis[match(df$PatientID, diagnoses$PatientID)]
+# Diagnosis per Angela
+df$Diagnosis <- factor(diagnoses$Diagnosis[match(df$PatientID, diagnoses$PatientID)])
+levels(df$Diagnosis) = c("CFTR-RD or CRMS","CFTR-RD or CRMS","Cystic Fibrosis","Non-CF")
 # Add modulator info
@@
 })
-
-# Difference between arms
-df$`Cl Diff. Between Left and Right Arm` <- as.numeric(df$Cl_Left) - as.numeric(df$Cl_Right)
-df$`Na Diff. Between Left and Right Arm` <- as.numeric(df$Na_Left) - as.numeric(df$Na_Right)
 # Merge left and right values
@@
 
-## By diagnosis
-
 ```{r results='asis'}
 t1 <- tableby(
-  Diagnosis ~ Sex + Age.at.Test.in.Years + fe(geno_risk) + fe(geno_class) +
-    Amount + Cl + Na + K + `Na/Cl Ratio`,
-  data = t1_df, control = mycontrols, simulate.p.value = T
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-## By method
-
-```{r results='asis'}
-t1 <- tableby(
-  Method ~ Sex + Age.at.Test.in.Years + Diagnosis + fe(geno_risk) + fe(geno_class) +
-    Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` +
-    `Na Diff. Between Left and Right Arm`,
+  Diagnosis ~ Sex + Age.at.Test.in.Years + Method+ fe(geno_risk) + fe(geno_class),
   data = t1_df, control = mycontrols, simulate.p.value = T
@@
 t1 <- tableby(
-  Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` +
-    `Cl Diff. Between Left and Right Arm` +
-    `Na Diff. Between Left and Right Arm`,
+  Diagnosis ~ Cl + Na + `Na/Cl Ratio` + `Sweat Chloride Group`,
   data = t1_df, control = mycontrols,
@@
 
-### Gibson-Cooke only
+## By sweat chloride group
 
@@
 t1 <- tableby(
-  Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` +
-    `Cl Diff. Between Left and Right Arm` +
-    `Na Diff. Between Left and Right Arm`,
-  data = t1_df[t1_df$Method == "Gibson-Cooke", ], control = mycontrols,
+  `Sweat Chloride Group` ~ fe(geno_risk) + fe(Diagnosis)+  Cl + Na + `Na/Cl Ratio`,
+  data = t1_df, control = mycontrols,simulate.p.value = T
 )
@@
 
-### Macroduct only
-
-```{r results='asis'}
-t1 <- tableby(
-  Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` +
-    `Cl Diff. Between Left and Right Arm` +
-    `Na Diff. Between Left and Right Arm`,
-  data = t1_df[t1_df$Method == "Macroduct", ], control = mycontrols,
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-## By genotype risk
-
-```{r results='asis'}
-t1 <- tableby(
-  geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` +
-    `Cl Diff. Between Left and Right Arm` +
-    `Na Diff. Between Left and Right Arm`,
-  data = t1_df, control = mycontrols,
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-## By genotype risk and diagnosis
+## By sweat chloride group and diagnosis
 
@@
 ```{r results='asis'}
-t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+t1 <- tableby(`Sweat Chloride Group` ~ Cl + Na + `Na/Cl Ratio`,
   data = t1_df[t1_df$Diagnosis == "Cystic Fibrosis", ], control = mycontrols
@@
 
-### CFTR-related disorder
+### CFTR-RD or CRMS
 
 ```{r results='asis'}
-t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
-  data = t1_df[t1_df$Diagnosis == "CFTR-related disorder", ], control = mycontrols
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-### CRMS
-
-```{r results='asis'}
-t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
-  data = t1_df[t1_df$Diagnosis == "CRMS", ], control = mycontrols
+t1 <- tableby(`Sweat Chloride Group` ~ Cl + Na + `Na/Cl Ratio`,
+  data = t1_df[t1_df$Diagnosis == "CFTR-RD or CRMS", ], 
+  control = mycontrols
 )
@@
 ```{r results='asis'}
-t1 <- tableby(~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
+t1 <- tableby(`Sweat Chloride Group` ~ Cl + Na + `Na/Cl Ratio`,
   data = t1_df[t1_df$Diagnosis == "Non-CF", ], control = mycontrols
@@
 
-### Plots
+# Variability
 
-```{r warning=FALSE}
-ggplot(df, aes(x = Diagnosis, y = Na)) +
-  geom_boxplot() +
-  theme_bw() +
-  ggtitle("Sweat Sodium")
+## Within occasion
 
-ggplot(df, aes(x = Diagnosis, y = K)) +
-  geom_boxplot() +
-  theme_bw() +
-  ggtitle("Sweat Potassium")
+```{r}
 
-ggplot(df, aes(x = Diagnosis, y = Cl)) +
-  geom_boxplot() +
-  theme_bw() +
-  ggtitle("Sweat Chloride")
-
-ggplot(df, aes(x = Diagnosis, y = `Na/Cl Ratio`)) +
-  geom_boxplot() +
-  theme_bw() +
-  ggtitle("Na/Cl Ratio")
 ```
 
-## By specific genotype
-
-```{r results='asis'}
-t1 <- tableby(
-  geno_class ~ Amount + Cl + Na + K + `Na/Cl Ratio` +
-    `Cl Diff. Between Left and Right Arm` +
-    `Na Diff. Between Left and Right Arm`,
-  data = t1_df, control = mycontrols,
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-## By specific genotype and diagnosis
-
-### Cystic fibrosis
-
-```{r results='asis'}
-t1 <- tableby(geno_class ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
-  data = t1_df[t1_df$Diagnosis == "Cystic Fibrosis", ], control = mycontrols
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-### CFTR-related disorder
-
-```{r results='asis'}
-t1 <- tableby(geno_class ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
-  data = t1_df[t1_df$Diagnosis == "CFTR-related disorder", ], control = mycontrols
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-### CRMS
-
-```{r results='asis'}
-t1 <- tableby(geno_class ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
-  data = t1_df[t1_df$Diagnosis == "CRMS", ], control = mycontrols
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-### Non-CF
-
-```{r results='asis'}
-t1 <- tableby(~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
-  data = t1_df[t1_df$Diagnosis == "Non-CF", ], control = mycontrols
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-### Plots
-
-```{r warning=FALSE}
-ggplot(df, aes(x = geno_class, y = Na)) +
-  geom_boxplot() +
-  theme_bw() +
-  xlab("Specific Genotype") +
-  ggtitle("Sweat Sodium")
-
-ggplot(df, aes(x = geno_class, y = K)) +
-  geom_boxplot() +
-  theme_bw() +
-  ggtitle("Sweat Potassium")
-
-ggplot(df, aes(x = geno_class, y = Cl)) +
-  geom_boxplot() +
-  theme_bw() +
-  xlab("Specific Genotype") +
-  ggtitle("Sweat Chloride")
-
-ggplot(df, aes(x = geno_class, y = `Na/Cl Ratio`)) +
-  geom_boxplot() +
-  theme_bw() +
-  xlab("Specific Genotype") +
-  ggtitle("Na/Cl Ratio")
-```
-
-# Between occasion
+## Between occasion
 
@@
 
-## Overall
-
 ```{r results='asis'}
 p <- paired(
-  time ~ Diagnosis + Cl + Na + K + `Na/Cl Ratio` +
-    `Cl Diff. Between Left and Right Arm` +
-    `Na Diff. Between Left and Right Arm`,
+  time ~ Cl + Na + `Na/Cl Ratio`,
   signed.rank.exact = F, data = multiples,
-  id = PatientID, control = mycontrols
-)
-summary(p, pfootnote = T, labelTranslations = labels)
-```
-
-## By diagnosis
-
-Not enough observations in CRMS or CFTR-related disorder groups to calculate p values. Also not enough observations to further break down the table by sweat collection method.
-
-```{r results='asis'}
-p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
-  signed.rank.exact = F, strata = Diagnosis,
-  data = multiples %>% filter(Diagnosis %in% c("Cystic Fibrosis", "Non-CF")),
-  id = PatientID, control = mycontrols
-)
-summary(p, pfootnote = T, labelTranslations = labels)
-```
-
-## By genotype risk
-
-```{r results='asis'}
-p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
-  signed.rank.exact = F, strata = geno_risk,
-  data = multiples,
-  id = PatientID, control = mycontrols
-)
-summary(p, pfootnote = T, labelTranslations = labels)
-```
-
-## By genotype risk and diagnosis
-
-Not enough paired observations in the CRMS or CFTR-related disorder groups.
-
-### Cystic fibrosis
-
-```{r results='asis'}
-p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio`,
-  signed.rank.exact = F, strata = geno_risk,
-  data = multiples %>% filter(Diagnosis == "Cystic Fibrosis"),
-  id = PatientID, control = mycontrols
-)
-summary(p, pfootnote = T, labelTranslations = labels)
-```
-
-### Non-CF
-
-```{r results='asis'}
-p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio`,
-  signed.rank.exact = F, strata = geno_risk,
-  data = multiples %>% filter(Diagnosis == "Non-CF"),
-  id = PatientID, control = mycontrols
-)
-summary(p, pfootnote = T, labelTranslations = labels)
-```
-
-## By specific genotype
-
-```{r results='asis'}
-p <- paired(
-  time ~ Cl + Na + K + `Na/Cl Ratio` +
-    `Cl Diff. Between Left and Right Arm` +
-    `Na Diff. Between Left and Right Arm`,
-  signed.rank.exact = F, strata = geno_class, numeric.test = "notest",
-  data = multiples, id = PatientID, control = mycontrols
-)
-summary(p, pfootnote = T, labelTranslations = labels)
-```
-
-## By genotype class and diagnosis
-
-Not enough paired observations in the CRMS or CFTR-related disorder groups.
-
-### Cystic fibrosis
-
-```{r results='asis'}
-p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio`,
-  signed.rank.exact = F, strata = geno_class,
-  data = multiples %>% filter(Diagnosis == "Cystic Fibrosis"),
-  id = PatientID, control = mycontrols
-)
-summary(p, pfootnote = T, labelTranslations = labels)
-```
-
-### Non-CF
-
-```{r results='asis'}
-p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio`,
-  signed.rank.exact = F, strata = geno_class,
-  data = multiples %>% filter(Diagnosis == "Non-CF"),
   id = PatientID, control = mycontrols
@@
 ```{r}
-# Re-format diagnosis column
-t1_df$`Combined Diagnosis` <- factor(t1_df$Diagnosis,
-                     levels = c("Non-CF","CFTR-related disorder","CRMS","Cystic Fibrosis"),
-                     labels = c("Non-CF","CFTR-related disorder or CRMS","CFTR-related disorder or CRMS","Cystic Fibrosis"),
-                     ordered = T)
-
-```
-
-```{r}
-roc_cf <- roc(response= t1_df$`Combined Diagnosis`,
+roc_cf <- roc(response= t1_df$`Diagnosis`,
               predictor = t1_df$`Na/Cl Ratio`, 
-              levels = c("CFTR-related disorder or CRMS","Cystic Fibrosis"),
+              levels = c("CFTR-RD or CRMS","Cystic Fibrosis"),
               quiet = T)
@@
 
-roc_non_cf <- roc(response= t1_df$`Combined Diagnosis`,
+roc_non_cf <- roc(response= t1_df$`Diagnosis`,
               predictor = t1_df$`Na/Cl Ratio`, 
-              levels = c("Non-CF","CFTR-related disorder or CRMS"),
+              levels = c("Non-CF","CFTR-RD or CRMS"),
               quiet = T)
@@
 
-The blue curve shows the ROC for distinguishing between CFTR-related disorder or CRMS vs.
+The blue curve shows the ROC for distinguishing between CFTR-RD or CRMS vs.
 CF, and the green is the ROC curve for distinguishing between non-CF vs.
-CFTR-related disorder or CRMS.
-The optimal threshold for distinguishing between CFTR-related disorder or CRMS vs.
+CFTR-RD or CRMS.
+The optimal threshold for distinguishing between CFTR-RD or CRMS vs.
 CF was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index.
 The optimal threshold for distinguishing between non-CF vs.
-CFTR-related disorder or CRMS was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index.
+CFTR-RD or CRMS was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index.
 The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.
@@
 
-Vertical blue line indicates the threshold for CFTR-related disorder or CRMS vs.
-CF and green indicates the threshold for non-CF vs.
-CFTR-related disorder or CRMS.
-Na/Cl ratio was log-transformed for easier visualization.
+Vertical blue line indicates the threshold for CFTR-RD or CRMS vs. CF and green indicates the threshold for non-CF vs. CFTR-RD or CRMS. Na/Cl ratio was log-transformed for easier visualization.
 
 ```{r warning=FALSE}
-ggplot(t1_df[!is.na(t1_df$`Combined Diagnosis`), ], 
-       aes(log(`Na/Cl Ratio`), fill = `Combined Diagnosis`)) +
+ggplot(t1_df[!is.na(t1_df$`Diagnosis`), ], 
+       aes(log(`Na/Cl Ratio`), fill = `Diagnosis`)) +
   geom_histogram(alpha = 0.3, position = "identity",bins = 50) +
@@
 ```{r}
-roc_cf <- roc(response= t1_df$`Combined Diagnosis`,
+roc_cf <- roc(response= t1_df$`Diagnosis`,
               predictor = t1_df$`Na/Cl Ratio`, 
-              levels = c("CFTR-related disorder or CRMS","Cystic Fibrosis"),
+              levels = c("CFTR-RD or CRMS","Cystic Fibrosis"),
               quiet = T)

### 2023-02-22 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
swaet test updates done

@@
 
+For all participants with more than one sweat test, we used only their first value to calculate the absolute difference (AD) between left and right arm.
+
 ```{r}
+t1_df$`Cl AD` = abs(t1_df$Cl_Left-t1_df$Cl_Right)
+t1_df$`Na AD` = abs(t1_df$Na_Left-t1_df$Na_Right)
+```
 
+### By diagnosis
+
+```{r results='asis'}
+t1 <- tableby(
+  Diagnosis ~ `Cl AD`+`Na AD`,
+  data = t1_df, control = mycontrols,
+)
+summary(t1,
+  pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### By sweat chloride group
+
+```{r results='asis'}
+t1 <- tableby(
+  `Sweat Chloride Group` ~ `Cl AD`+`Na AD`,
+  data = t1_df, control = mycontrols
+)
+summary(t1,
+  pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### By genotype risk
+
+```{r results='asis'}
+t1 <- tableby(
+  geno_risk ~ `Cl AD`+`Na AD`,
+  data = t1_df, control = mycontrols
+)
+summary(t1,
+  pfootnote = T,
+  labelTranslations = labels
+)
 ```
@@
 
-For all participants with more than one sweat test, we compared their first and last values.
+For all participants with more than one sweat test, we compared their first and last values by calculating the absolute difference (AD) between first and last measure.
 
@@
   filter(row_number() == 1 | row_number() == n()) %>% # First and last
-  mutate(time = row_number())
-multiples$time <- factor(multiples$time, levels = 1:2, labels = c("First", "Last"))
-# Paired testing options
-mycontrols <- tableby.control(
-  numeric.test = "signed.rank", cat.test = "mcnemar",
-  numeric.stats = c("Nmiss", "median", "q1q3")
+  summarise(`Cl AD` = abs(last(Cl) - first(Cl)), # Calculate AD
+            `Na AD` = abs(last(Na) - first(Na)),
+            Diagnosis = first(Diagnosis),
+            `Sweat Chloride Group` = first(`Sweat Chloride Group`),
+            geno_risk = first(geno_risk)) 
+```
+
+### By diagnosis
+
+```{r results='asis'}
+t1 <- tableby(
+  Diagnosis ~ `Cl AD`+`Na AD`,
+  data = multiples, control = mycontrols,
+)
+summary(t1,
+  pfootnote = T,
+  labelTranslations = labels
 )
@@
 
+### By sweat chloride group
+
 ```{r results='asis'}
-p <- paired(
-  time ~ Cl + Na + `Na/Cl Ratio`,
-  signed.rank.exact = F, data = multiples,
-  id = PatientID, control = mycontrols
+t1 <- tableby(
+  `Sweat Chloride Group` ~ `Cl AD`+`Na AD`,
+  data = multiples, control = mycontrols
 )
-summary(p, pfootnote = T, labelTranslations = labels)
+summary(t1,
+  pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### By genotype risk
+
+```{r results='asis'}
+t1 <- tableby(
+  geno_risk ~ `Cl AD`+`Na AD`,
+  data = multiples, control = mycontrols
+)
+summary(t1,
+  pfootnote = T,
+  labelTranslations = labels
+)
 ```

### 2023-03-09 Christine Chan/CF Registry/cf_registry_analysis.qmd
Update cf_registry_analysis.qmd

@@
 
+# Incidence and Prevalence
+
 ```{r}
 # All data
-encounter <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
+encounter <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
                       na.strings = c("", "N/A")
 )
-annual <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
+annual <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
                    na.strings = c("", "N/A")
@@
 Screening rate is calculated as the number of people with an OGTT divided by the total number of people seen in clinic that year.
+
+# ROC
+
+```{r}
+# Prepare data for ROC
+df$cfrd_ogtt <- factor(df$cfrd_ogtt, levels = c("No", "Yes"), ordered = T)
+cfrd <- df %>%
+  group_by(eDWID) %>%
+  mutate(
+    cfrd_ever = max(cfrd_ogtt),
+    time_from_diag = ReviewYear - first_cfrd
+  ) %>%
+  filter(cfrd_ever == "Yes", time_from_diag <= 0, time_from_diag >= -1) %>%
+  summarise(hba1c = last(na.omit(round(A_hgba1c, 1)))) %>%
+  drop_na()
+non_cfrd <- df %>%
+  group_by(eDWID) %>%
+  mutate(cfrd_ever = max(cfrd_ogtt)) %>%
+  filter(cfrd_ever == "No") %>%
+  summarise(hba1c = last(na.omit(round(A_hgba1c, 1)))) %>%
+  drop_na()
+```
+
+```{r}
+#| label: fig-roc
+#| fig-cap: "ROC Curve"
+r <- roc(
+  cases = cfrd$hba1c, controls = non_cfrd$hba1c, quiet = T, plot = T, auc = T, ci = T,
+  print.auc = TRUE, show.thres = TRUE
+)
+```
+
+Because we do not have exact information regarding HbA1c timing (we only have encounter number within year), the annualized HbA1c from the year of diagnosis or from one year prior to diagnosis was used in the ROC analysis.
+For those with HbA1c measures the year of diagnosis and one year prior, the value from year of diagnosis was preferred.
+For those never diagnosed with CFRD, the most recent HbA1c was used.
+There were a total of `r nrow(cfrd)` HbA1c measures for people with CFRD and `r nrow(non_cfrd)` measures for those without.
+
+The AUC for this ROC curve was 0.61 (see @fig-roc).
+
+```{r}
+#| label: tbl-roc_thresh
+#| tbl-cap: "Best ROC Threshold (Youden Index)"
+kable(coords(r, "best"), digits = 3)
+```

### 2023-03-10 Emily Muther/body_image_trikafta.qmd
started body image trikafta over

--- 2023-01-12 Emily Muther/body_image_trikafta.qmd
@@
 author: "Tim Vigers"
-date: "`r format(Sys.time(), '%d %B %Y')`"
+date: today
 format:
@@
 
-```{r}
-#| include: false
-# Libraries
-library(haven)
-library(tidyverse)
-library(arsenal)
-# Import
-df = read_sav("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Body Image and Trikafta/Data_Raw/TrikaftaBodyImage1.11.23.sav")
-# Format variables
-df$`Percent Adherence` = factor(df$PercentAdherence,levels = 1:11,
-                             labels = paste0(seq(0,100,by=10),"%"),ordered = T)
-df$`FRS Category` = factor(df$FRS_Category,levels = -1:1,
-                           labels = c("want to be larger","neutral",
-                                      "want to be smaller"))
-df$`Doses Missed` = as.numeric(df$DosesMissed)
-df$`Weight/Body Adherence` = factor(df$WeightBody_Adherence,levels = 1:5,
-               labels = c("Strongly Agree","Slightly Agree","Neutral",
-                          "Slightly Disagree","Strongly Disagree"))
-df$`Weight/Body Adherence` = fct_rev(df$`Weight/Body Adherence`)
-df$`Weight/Body Adherence` = factor(df$`Weight/Body Adherence`,ordered = T)
+```{python}
+# | include: false
+import pandas as pd
+# Read in data
+df = pd.read_spss(
+    "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Body Image and Trikafta/Data_Raw/TrikaftaBodyImage3.6.23.sav")
 ```
 
-Because we want to keep the ordinal nature of some of these variables rather than simply treating them as categorical (i.e.
-with percent adherence it's meaningful that 10% \< 20% \< 30% ...), Emma is right in thinking that ordinal logistic regression is the right approach.
-Unfortunately, this requires estimating a lot of parameters, and we probably don't have the necessary numbers for it (for example, with percent adherence as the outcome we'd need to estimate 12 parameters, and the general rule of thumb is at least 10 observations per parameter).
+# Data Cleaning Steps
 
-So, after talking this through with another statistician in our group, I think the best way to approach this is to treat these variables as continuous, but apply non-parametric, rank-based tests between various groups.
-This keeps the information about the order of the categories but does not assume normal distribution, etc.
+1. Participants who answered "I don't know" or "no" on the variable `TakingTrikafta` were excluded from these analyses.
 
-```{r results='asis'}
-#| label: tbl-1
-#| tbl-cap: Comparisons by FRS Category
-t1 = tableby(`FRS Category`~kwt(`Percent Adherence`)+`Doses Missed` +
-               kwt(`Weight/Body Adherence`),data = df,
-             control=tableby.control(numeric.stats=c("Nmiss","median","q1q3","range"),
-                                     numeric.test="kwt"))
-summary(t1,pfootnote = T)
+```{python}
+# | include: false
+# | eval: false
+from ydata_profiling import ProfileReport
+# EDA
+profile = ProfileReport(df, explorative=True)
+# Saving results to a HTML file
+profile.to_file(
+    "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Body Image and Trikafta/Reports/eda.html")
 ```
 
-```{r}
-#| label: fig-planets
-#| fig-cap: Planets
+```{python}
+# Filter
+df = df[df["TakingTrikafta"] == "yes"]
+```
+
+```{python}
+# | label: tbl-planets
+# | tbl-cap: Planets
 
@@
 
-# Questions
+# Figure
 
-1. Which variables should be considered ordinal and which categorical?
+```{python}
+# | label: fig-planets
+# | fig-cap: Planets
+
+```

### 2023-03-14 Quarto Template.qmd
Update Quarto Template.qmd

@@
 author: "Tim Vigers"
-date: "`r format(Sys.time(), '%d %B %Y')`"
+date: "today"
 format:
@@
 
-# figure
+# Figure
 

### 2023-03-24 Emily Muther/body_image_trikafta.qmd
Update body_image_trikafta.qmd
Latex not working for some reason

@@
 date: today
+date-format: long
 format:
-  html:
+  pdf:
     toc: true
     toc-depth: 5
-    toc-float: true
-    code-fold: true
-    self-contained: true
-    fig-cap-location: top
-    page-layout: full
-editor: visual
+    lot: true
+    lof: true
+execute:
+  echo: false
+pdf-engine: pdflatex
+editor: source
 ---
-
-```{python}
-# | include: false
-import pandas as pd
-# Read in data
-df = pd.read_spss(
-    "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Body Image and Trikafta/Data_Raw/TrikaftaBodyImage3.6.23.sav")
-```
 
@@
 
-```{python}
-# | include: false
-# | eval: false
-from ydata_profiling import ProfileReport
-# EDA
-profile = ProfileReport(df, explorative=True)
-# Saving results to a HTML file
-profile.to_file(
-    "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Body Image and Trikafta/Reports/eda.html")
+```{r warning=FALSE}
+#| include: false
+library(tidyverse)
+library(gtsummary)
+library(kableExtra)
+library(foreign)
+# Read in data
+df = read.spss("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Body Image and Trikafta/Data_Raw/TrikaftaBodyImage3.6.23.sav",to.data.frame = T)
+# Filter
+df = df %>% 
+  filter(TakingTrikafta=="yes")
+# Pick relevant columns
+df = df %>% 
+  select(ID,Age,Gender,PercentAdherence,DosesMissed,FeelTrikafta,
+         DifferencesTrikafta,BMI,FEV1,Race,Ethnicity,InsuranceType,
+         WeightBody_Adherence,DoNotWantTrikafta,ContinueTrikafta,
+         FRS_Category,contains("BE_"),contains("PHQ9"),contains("GAD7"),
+         -contains("MEAN"))
 ```
 
-```{python}
-# Filter
-df = df[df["TakingTrikafta"] == "yes"]
-```
+# Descriptive analyses
 
-```{python}
-# | label: tbl-planets
-# | tbl-cap: Planets
-
+```{r}
+#| label: tbl-1
+#| tbl-cap: Baseline characteristics by FRS category
+#| message: false
+df %>% select(-ID) %>%
+  tbl_summary(by = FRS_Category) %>%
+  as_kable_extra(
+    booktabs = TRUE,
+    longtable = TRUE
+  ) %>% 
+  kable_styling(
+    full_width = T,
+    latex_options = "repeat_header"
+  )
 ```
@@
 
-```{python}
+```{r}
 # | label: fig-planets

### 2023-03-25 Emily Muther/body_image_trikafta.qmd
Update body_image_trikafta.qmd

--- 2023-03-10 Emily Muther/body_image_trikafta.qmd
@@
 date: today
+date-format: long
 format:
@@
     page-layout: full
-editor: visual
+execute:
+  echo: false
+editor: source
 ---
-
-```{python}
-# | include: false
-import pandas as pd
-# Read in data
-df = pd.read_spss(
-    "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Body Image and Trikafta/Data_Raw/TrikaftaBodyImage3.6.23.sav")
-```
 
@@
 
-```{python}
-# | include: false
-# | eval: false
-from ydata_profiling import ProfileReport
-# EDA
-profile = ProfileReport(df, explorative=True)
-# Saving results to a HTML file
-profile.to_file(
-    "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Body Image and Trikafta/Reports/eda.html")
+```{r warning=FALSE}
+#| include: false
+library(tidyverse)
+library(gtsummary)
+library(kableExtra)
+library(foreign)
+# Read in data
+df = read.spss("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Body Image and Trikafta/Data_Raw/TrikaftaBodyImage3.6.23.sav",to.data.frame = T)
+# Filter
+df = df %>% 
+  filter(TakingTrikafta=="yes")
+# Trim white space
+df = df %>%
+  mutate(across(where(is.character),trimws))
+# Replace blanks with NA
+df[df==""]=NA
 ```
 
-```{python}
-# Filter
-df = df[df["TakingTrikafta"] == "yes"]
+# Descriptive analyses
+
+## Demographics
+
+```{r}
+#| label: demographics
+df %>% select(Age,Gender,Race,Ethnicity) %>%
+  tbl_summary(
+    by = Gender,
+    statistic = list(
+      all_continuous() ~ "{mean} ({sd})"
+    ),
+    missing_text = "(Missing)"
+  )
 ```
 
-```{python}
-# | label: tbl-planets
-# | tbl-cap: Planets
+## Frequency of reported changes
 
+```{r}
+#| label: changes
+df %>% select(ChangestoBreathing:MoodChangesSideEf,FeelTrikafta) %>%
+  tbl_summary(
+    type = list(
+      all_categorical() ~ "categorical"
+    ),
+    missing_text = "(Missing)"
+  )
 ```

### 2023-03-28 R Markdown Template.Rmd
Update R Markdown Template.Rmd

--- 2022-02-09 R Markdown Template.Rmd
@@
 rm(home_dir)
+source("~/GitHub/CF/genotype_class_severity.R")
 ```

### 2023-03-28 Scott Sagel/PEx in CF/nacfc_2023.qmd
Create nacfc_2023.qmd

---
title: "Changing face of pediatric pulmonary exacerbations in CF"
author: "Tim Vigers"
date: "today"
format:
  pdf:
    toc: true
    lot: true
    lof: true
editor: source
execute: 
  echo: false
---

```{r}
#| include: false
library(readxl)
library(tidyverse)
library(gtsummary)
source("~/GitHub/CF/genotype_class_severity.R")
# Import
chco = read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CHCO.xlsx",na = c("","NULL"))
cnh = read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CNH.xlsx")
# Format factor columns
chco$`Modulator at Time of Admit` = factor(chco$`Modulator at Time of Admit`,
                                           levels = c("No","Yes"),ordered = T)
```

# Data cleaning notes

1. Age at admission and days in hospital were averaged for each person with multiple admissions.

```{r}
#| label: tbl-chco
#| tbl-cap: CHCO Summary by Year
#| message: false
chco %>% group_by(`Unique number`,Year) %>%
  summarise(`# of Hospitalizations per Person` = n(),
            across(Sex:Genotypes2,~first(na.omit(.x))),
            across(`Modulator at Time of Admit`,~max(.x)),
            across(all_of(c("Age at Admit","# Days in Admit")),~mean(.x,na.rm=T)),
            .groups = "drop") %>%
  mutate(Genotypes1 = case_when(
    Genotypes1 == "F508del" | Genotypes1 == "delF508" ~ "F508del",
    .default = "Other"
  ),
  Genotypes2 = case_when(
    Genotypes2 == "F508del" | Genotypes2 == "delF508" ~ "F508del",
    .default = "Other"
  )) %>%
  unite(Genotype,Genotypes1,Genotypes2,sep = "/") %>%
  mutate(Genotype = factor(Genotype,
                           levels=c("F508del/F508del","F508del/Other",
                                             "Other/F508del","Other/Other"),
                           labels = c("F508del/F508del","F508del/Other",
                                             "F508del/Other","Other/Other")
                           )) %>%
  select(`# of Hospitalizations per Person`,`# Days in Admit`,`Age at Admit`,
         `Modulator at Time of Admit`,Sex:Genotype,Year) %>%
  tbl_summary(by=Year) %>%
  add_p() %>%
  as_gt()
```

The number of hospitalizations for PEx decreased from `r sum(chco$Year==2018)` in 2018, involving `r nrow(unique(chco[chco$Year==2018,"Unique number"]))` unique individuals, to `r sum(chco$Year==2022)` in 2022, involving `r nrow(unique(chco[chco$Year==2022,"Unique number"]))` unique individuals.

### 2023-03-29 Scott Sagel/PEx in CF/nacfc_2023.qmd
Update nacfc_2023.qmd

@@
 format:
-  pdf:
+  html:
     toc: true
-    lot: true
-    lof: true
+    toc-depth: 5
+    toc-float: true
+    code-fold: true
+    self-contained: true
+    fig-cap-location: top
+    page-layout: full
 editor: source
@@
 cnh = read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CNH.xlsx")
-# Format factor columns
-chco$`Modulator at Time of Admit` = factor(chco$`Modulator at Time of Admit`,
-                                           levels = c("No","Yes"),ordered = T)
-```
-
-# Data cleaning notes
-
-1. Age at admission and days in hospital were averaged for each person with multiple admissions.
-
-```{r}
-#| label: tbl-chco
-#| tbl-cap: CHCO Summary by Year
-#| message: false
-chco %>% group_by(`Unique number`,Year) %>%
-  summarise(`# of Hospitalizations per Person` = n(),
-            across(Sex:Genotypes2,~first(na.omit(.x))),
-            across(`Modulator at Time of Admit`,~max(.x)),
-            across(all_of(c("Age at Admit","# Days in Admit")),~mean(.x,na.rm=T)),
-            .groups = "drop") %>%
+# Format CHCO columns
+chco$`P. aeruginosa` = coalesce(chco$`P. aeruginosa mucoid`,chco$`P. aeruginosa non-mucoid`)
+chco = chco %>% rename(
+  Race = `Race (White/Black/Other)`,Ethnicity = `Hispanic/Latinx`
+) %>%
   mutate(Genotypes1 = case_when(
@@
     .default = "Other"
-  )) %>%
-  unite(Genotype,Genotypes1,Genotypes2,sep = "/") %>%
+  ),
+  `FEV1 % Predicted at Admit` = suppressWarnings(as.numeric(`FEV1 % Predicted at Admit`))) %>%
+  unite(Genotype,Genotypes1,Genotypes2,sep = "/")
+# Format CNH to match CHCO
+cnh = cnh %>%
+  rename(`# Days in Admit`=`Total hospitalization days`,
+         `Age at Admit`=`Age at admission`,
+         `FEV1 % Predicted at Discharge`=`End of treamtent ppFEV1`,
+         `Unique number` = `Study ID`) %>%
+  mutate(Site = "CNH",
+         `Modulator at Time of Admit` = case_when(
+           None == "Unchecked" ~ "Yes",
+           None =="Checked" ~"No"
+         ),
+         MRSA = case_when(
+           MRSA == "Unchecked" ~ NA,
+           MRSA =="Checked" ~ TRUE
+         ),
+         `Staph aureus` = case_when(
+           MSSA == "Unchecked" ~ NA,
+           MSSA =="Checked" ~ TRUE
+         ),
+         `P. aeruginosa` = case_when(
+           `P. aeruginosa` == "Unchecked" ~ NA,
+           `P. aeruginosa` =="Checked" ~ TRUE
+         ),
+         Genotype = case_when(
+           `CFTR genotype, groups` =="F508del heterozygous" ~ "F508del/Other",
+           `CFTR genotype, groups` =="F508del homozygous"~"F508del/F508del",
+           `CFTR genotype, groups` =="Other"~"Other/Other"
+         ),
+         `Specific Modulator at Admit` = case_when(
+           `Ivacaftor (Kalydeco)` == "Checked" ~ "Other",
+           `Ivacaftor/lumacaftor (Orkambi)` == "Checked" ~ "Other",
+           `Tezacaftor/ivacaftor (Symdeko)`=="Checked" ~ "Other",
+           `Elexacaftor/tezacaftor/ivacaftor (Trikafta)`=="Checked" ~ "Trikafta",
+           .default = "None"
+         ),
+         `FEV1 % Predicted at Admit` = as.numeric(`PEx ppFEV1`))
+# Merge
+df = full_join(chco,cnh)
+# Format column types
+df$`Modulator at Time of Admit` = factor(df$`Modulator at Time of Admit`,
+                                         levels = c("No","Yes"),ordered = T)
+df$`Specific Modulator at Admit`[df$`Modulator at Time of Admit`=="No"] = "None"
+df$`Specific Modulator at Admit` = factor(tolower(df$`Specific Modulator at Admit`),
+                                          levels = c("kalydeco","none","orkambi","symdeko","trikafta"),
+                                          labels = c("Other","None","Other","Other","Trikafta"))
+df$Race = factor(df$Race,levels = c("Amer. Indian or Alaskan Naitive","Asian",
+                                    "Black","Black or African American","Black/ AA",
+                                    "Decline to Answer","Hispanic","Other",
+                                    "Unknown / Not Reported","White"),
+                 labels = c("Other","Other","Black","Black","Black",
+                            "Unknown / Not Reported","Other","Other",
+                            "Unknown / Not Reported","White"))
+df$Ethnicity = factor(df$Ethnicity,levels = c("FALSE","Hispanic or Latino","HLS",
+                                              "NOT Hispanic or Latino","NOT HLS",
+                                              "TRUE","UNKNOWN"),
+                      labels = c("NOT Hispanic or Latino","Hispanic or Latino",
+                                 "Hispanic or Latino","NOT Hispanic or Latino",
+                                 "NOT Hispanic or Latino","Hispanic or Latino",
+                                 "Unknown"))
+bugs = c("P. aeruginosa","P. aeruginosa mucoid","P. aeruginosa non-mucoid","Staph aureus","MRSA")
+df[,bugs] = lapply(df[,bugs],function(c){
+  c[is.na(c)] = "No"
+  c = factor(c,levels = c("No",TRUE),labels = c("No","Yes"),ordered = T)
+})
+# Make summary table
+summ = df %>% group_by(`Unique number`,Site,Year) %>%
+  mutate(`Change in FEV1` = `FEV1 % Predicted at Discharge` - `FEV1 % Predicted at Admit`) %>%
+  summarise(`# of Hospitalizations per Person` = n(),
+            across(c(Sex:Genotype,`Specific Modulator at Admit`),~first(na.omit(.x))),
+            across(c(`Modulator at Time of Admit`,all_of(bugs)),~max(.x,na.rm = T)),
+            across(all_of(c("Age at Admit","# Days in Admit","Change in FEV1",
+                            "FEV1 % Predicted at Admit","FEV1 % Predicted at Discharge")),
+                   ~mean(.x,na.rm=T)),
+            .groups = "drop")  %>%
   mutate(Genotype = factor(Genotype,
@@
   )) %>%
-  select(`# of Hospitalizations per Person`,`# Days in Admit`,`Age at Admit`,
-         `Modulator at Time of Admit`,Sex:Genotype,Year) %>%
-  tbl_summary(by=Year) %>%
-  add_p() %>%
+  select(`Unique number`, `Site`,Year,`# of Hospitalizations per Person`,`# Days in Admit`,`Age at Admit`,
+         `Change in FEV1`,`FEV1 % Predicted at Admit`,`FEV1 % Predicted at Discharge`,
+         `Modulator at Time of Admit`,`Specific Modulator at Admit`,
+         all_of(bugs),Sex:Genotype)
+```
+
+# Data cleaning notes
+
+1. For each person with multiple admissions, continuous variables were averaged prior to generating summary tables.
+2.
+If a participant was on a modulators during any admission, they were counted as on modulators for that year.
+For example, participant 4 from CHOC was hospitalized 3 times in 2018 and was on a modulator during one of those admissions, so they were counted as on modulators in 2018.
+3. CNH only included combined P. aeruginosa information.
+
+# Summary Statistics
+
+```{r}
+#| label: tbl-chco
+#| tbl-cap: Summary by Year
+#| message: false
+summ %>%
+  select(-`Unique number`,-Site) %>%
+  tbl_summary(by=Year,
+              type = list(`# of Hospitalizations per Person` ~ 'continuous'),
+              statistic = list(
+                c("# of Hospitalizations per Person","# Days in Admit") ~ "{median} ({p25}, {p75})",
+                c("Age at Admit","Change in FEV1","FEV1 % Predicted at Admit",
+                  "FEV1 % Predicted at Discharge") ~ "{mean} ({sd})"
+              ),
+              missing_text = "(Missing)") %>%
+  add_ci(pattern = "{stat} [{ci}]") %>%
+  add_p(test = list(
+    `Age at Admit`~"t.test",
+    `Change in FEV1`~"t.test"
+  )) %>%
+  separate_p_footnotes() %>%
   as_gt()
@@
 
-The number of hospitalizations for PEx decreased from `r sum(chco$Year==2018)` in 2018, involving `r nrow(unique(chco[chco$Year==2018,"Unique number"]))` unique individuals, to `r sum(chco$Year==2022)` in 2022, involving `r nrow(unique(chco[chco$Year==2022,"Unique number"]))` unique individuals.
+The number of hospitalizations for PEx decreased from `r sum(summ$Year==2018)` in 2018 to `r sum(summ$Year==2022)` in 2022.

### 2023-03-29 Scott Sagel/PEx in CF/nacfc_2023.qmd
Update nacfc_2023.qmd

@@
   html:
-    toc: true
-    toc-depth: 5
-    toc-float: true
+    toc: false
     code-fold: true
@@
 library(tidyverse)
+library(gt)
 library(gtsummary)
@@
 3. CNH only included combined P. aeruginosa information.
+4. Participants were uniquely identified by ID, site, and year.
+
+The number of hospitalizations for PEx decreased from `r sum(df$Year==2018)` in 2018 to `r sum(df$Year==2022)` in 2022.
 
@@
 ```{r}
-#| label: tbl-chco
+#| label: tbl-all
 #| tbl-cap: Summary by Year
@@
   separate_p_footnotes() %>%
-  as_gt()
+  as_gt() %>%
+  as_raw_html()
 ```
 
-The number of hospitalizations for PEx decreased from `r sum(summ$Year==2018)` in 2018 to `r sum(summ$Year==2022)` in 2022.
+```{r}
+#| label: tbl-site
+#| tbl-cap: Stratified by Site
+#| message: false
+summ %>%
+  select(-`Unique number`) %>%
+  tbl_strata(
+    strata = Site,
+    .tbl_fun =
+      ~ .x %>%
+      tbl_summary(by=Year,
+              type = list(`# of Hospitalizations per Person` ~ 'continuous'),
+              statistic = list(
+                c("# of Hospitalizations per Person","# Days in Admit") ~ "{median} ({p25}, {p75})",
+                c("Age at Admit","Change in FEV1","FEV1 % Predicted at Admit",
+                  "FEV1 % Predicted at Discharge") ~ "{mean} ({sd})"
+              ),
+              missing_text = "(Missing)") %>%
+  add_ci(pattern = "{stat} [{ci}]") %>%
+  add_p(test = list(
+    `Age at Admit`~"t.test",
+    `Change in FEV1`~"t.test"
+  )) %>%
+  separate_p_footnotes()
+  ) %>%
+  as_gt() %>%
+  as_raw_html()
+```

### 2023-03-29 Quarto Template.qmd
Update Quarto Template.qmd

@@
 date: "today"
+date-format: long
 format:

### 2023-03-29 Scott Sagel/PEx in CF/nacfc_2023.qmd
Update nacfc_2023.qmd

@@
 ---
-title: "Changing face of pediatric pulmonary exacerbations in CF"
+title: "The Changing Face of Pediatric Pulmonary Exacerbations in CF (NACFC 2023)"
 author: "Tim Vigers"

### 2023-03-29 Jordana Hoppe/nacfc_2023.qmd
Create nacfc_2023.qmd

---
title: "Remote Collection of Respiratory Samples (NACFC 2023)"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r}
#| include: false
library(pander)
library(tidyverse)
library(caret)
library(gtsummary)
df = read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/clinical.csv",na.strings = "")
surveys = read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/surveys.csv",na.strings = "")
# Combined genotype
df$Mutation.1[df$Mutation.1 != "F508del"]="Other"
df$Mutation.2[df$Mutation.2 != "F508del"]="Other"
df$Genotype = paste(df$Mutation.1,df$Mutation.2,sep = "/")
# Age
df$Age = difftime(df$Date.of.clinic.culture.collection, df$Patient.DOB)/365.25
# Combine race columns
df = df %>% rowwise() %>%
  mutate(Race = paste(which(c_across(Race..choice.American.Indian.or.Alaska.Native.:
                                       Race..choice.Unknown.or.not.reported.)== "Checked"),
                      collapse = "+"),
         Race = factor(Race,levels = c("","1","5"),labels=c("Unknown","American Indian or Alaska Native","White")))
# Shorten organism names
colnames(df) = sub("If.culture.was.positive..select.all.organisms.isolated...choice.","",colnames(df))
```

# Table 1: Participant Characteristics

```{r}
#| label: tbl-1
df %>% select(Age,FEV1...predicted,FVC...predicted,
              Sex,Race,Hispanic.or.Latinx,Rural.location.) %>%
  tbl_summary(
    label = list(
      Age ~ "Age at clinic culture collection",
      FEV1...predicted~"ppFEV",FVC...predicted~"ppFVC",
      Hispanic.or.Latinx~"Hispanic or Latinx",
      Rural.location. ~"Rural location?"
    ),
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
    ))
```

# Sensitivity and specificity

```{r}
sens_spec = function(gold_standard,other_test,bug,type = "Home Collection"){
  t = table(other_test,gold_standard)
  if(all(dim(t)==c(1,1))){
    cat("\n")
    cat("###", bug)
    cat("\n")
    cat("#### Contingency table")
    cat("\n")
    cat("All samples marked as",unique(gold_standard))
    cat("\n")
  } else {
    sen = sensitivity(data = factor(other_test),reference = factor(gold_standard))
    spec = specificity(data = factor(other_test),reference = factor(gold_standard))

    dimnames(t)=lapply(dimnames(t),function(n){
      n = sub("Checked","Present",n)
      n = sub("Unchecked","Absent",n)
    })
    dimnames(t)$other_test = paste(type,dimnames(t)$other_test)
    dimnames(t)$gold_standard = paste("Clinic",dimnames(t)$gold_standard)

    cat("\n")
    cat("###", bug)
    cat("\n")
    cat("#### Contingency table")
    cat("\n")
    print(knitr::kable((as.data.frame.matrix(t))))
    cat("\n")
    cat("Sensitivity of the",tolower(type),"method was",round(sen,3),"and specificity was",round(spec,3))
    cat("\n")
  }
}
```

## Home Collection

```{r results='asis'}
sens_spec(df$Methicillin.sensitive.Staph.aureus.,
          df$Methicillin.sensitive.Staph.aureus..1,
          type = "Home Collection",bug = "MSSA")
```

```{r results='asis'}
sens_spec(df$Methicillin.resistant.Staph.aureus.,
          df$Methicillin.resistant.Staph.aureus..1,
          type = "Home Collection",bug = "MRSA")
```

```{r results='asis'}
sens_spec(df$A..xylosoxidans.,df$A..xylosoxidans..1,
          type = "Home Collection",bug = "A. xylosoxidans")
```

```{r results='asis'}
sens_spec(df$Achromobacter.not.xylosoxidans.,df$Achromobacter.not.xylosoxidans..1,
          type = "Home Collection",bug = "Achromobacter (not xylosoxidans)")
```

```{r results='asis'}
sens_spec(df$Acinetobacter.,df$Acinetobacter..1,
          type = "Home Collection",bug = "Acinetobacter")
```

```{r results='asis'}
sens_spec(df$H..influenzae.,df$H..influenzae..1,
          type = "Home Collection",bug = "H. influenzae")
```

```{r results='asis'}
sens_spec(df$P..aeruginosa.,df$P..aeruginosa..1,
          type = "Home Collection",bug = "P. aeruginosa")
```

```{r results='asis'}
sens_spec(df$P..aeruginosa..mucoid..,df$P..aeruginosa..mucoid...1,
          type = "Home Collection",bug = "P. aeruginosa (mucoid)")
```

```{r results='asis'}
sens_spec(df$Pseudomonas.species..not.aeruginosa.,df$Pseudomonas.species..not.aeruginosa..1,
          type = "Home Collection",bug = "Pseudomonas (not aeruginosa)")
```

```{r results='asis'}
sens_spec(df$B..cepacia.complex.,df$B..cepacia.complex..1,
          type = "Home Collection",bug = "B. cepacia complex")
```

```{r results='asis'}
sens_spec(df$Stenotrophomonas.,df$Stenotrophomonas..1,
          type = "Home Collection",bug = "Stenotrophomonas")
```

```{r results='asis'}
sens_spec(df$Aspergillus.,df$Aspergillus..1,
          type = "Home Collection",bug = "Aspergillus")
```

```{r results='asis'}
sens_spec(df$Penicillium.species.,df$Penicillium.species..1,
          type = "Home Collection",bug = "Penicillium species")
```

```{r results='asis'}
sens_spec(df$Sceosporium.species.,df$Sceosporium.species..1,
          type = "Home Collection",bug = "Sceosporium species")
```

```{r results='asis'}
sens_spec(df$M..avium.complex.,df$M..avium.complex..1,
          type = "Home Collection",bug = "M. avium complex")
```

```{r results='asis'}
sens_spec(df$M..abscessus.,df$M..abscessus..1,
          type = "Home Collection",bug = "M. abscessus")
```

### 2023-03-29 Scott Sagel/PEx in CF/nacfc_2023.qmd
Update nacfc_2023.qmd

@@
                                  "Unknown"))
-bugs = c("P. aeruginosa","P. aeruginosa mucoid","P. aeruginosa non-mucoid","Staph aureus","MRSA")
+bugs = c("P. aeruginosa","P. aeruginosa mucoid","P. aeruginosa non-mucoid","Staph aureus","MRSA","S. maltophilia")
 df[,bugs] = lapply(df[,bugs],function(c){
@@
 })
+df$`S. maltophilia`[df$Site == "CNH"] = NA
 # Make summary table
@@
 summ %>%
-  select(-`Unique number`) %>%
+  select(-`Unique number`,-`S. maltophilia`) %>%
   tbl_strata(

### 2023-03-29 Scott Sagel/PEx in CF/nacfc_2023.qmd
Update nacfc_2023.qmd
tidyverse styling

--- 2023-03-29 Scott Sagel/PEx in CF/nacfc_2023.qmd
@@
 # Import
-chco = read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CHCO.xlsx",na = c("","NULL"))
-cnh = read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CNH.xlsx")
+chco <- read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CHCO.xlsx", na = c("", "NULL"))
+cnh <- read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CNH.xlsx")
 # Format CHCO columns
-chco$`P. aeruginosa` = coalesce(chco$`P. aeruginosa mucoid`,chco$`P. aeruginosa non-mucoid`)
-chco = chco %>% rename(
-  Race = `Race (White/Black/Other)`,Ethnicity = `Hispanic/Latinx`
+chco$`P. aeruginosa` <- coalesce(chco$`P. aeruginosa mucoid`, chco$`P. aeruginosa non-mucoid`)
+chco <- chco %>%
+  rename(
+    Race = `Race (White/Black/Other)`, Ethnicity = `Hispanic/Latinx`
   ) %>%
-  mutate(Genotypes1 = case_when(
+  mutate(
+    Genotypes1 = case_when(
       Genotypes1 == "F508del" | Genotypes1 == "delF508" ~ "F508del",
@@
     ),
-  `FEV1 % Predicted at Admit` = suppressWarnings(as.numeric(`FEV1 % Predicted at Admit`))) %>%
-  unite(Genotype,Genotypes1,Genotypes2,sep = "/")
+    `FEV1 % Predicted at Admit` = suppressWarnings(as.numeric(`FEV1 % Predicted at Admit`))
+  ) %>%
+  unite(Genotype, Genotypes1, Genotypes2, sep = "/")
 # Format CNH to match CHCO
-cnh = cnh %>%
-  rename(`# Days in Admit`=`Total hospitalization days`,
-         `Age at Admit`=`Age at admission`,
-         `FEV1 % Predicted at Discharge`=`End of treamtent ppFEV1`,
-         `Unique number` = `Study ID`) %>%
-  mutate(Site = "CNH",
+cnh <- cnh %>%
+  rename(
+    `# Days in Admit` = `Total hospitalization days`,
+    `Age at Admit` = `Age at admission`,
+    `FEV1 % Predicted at Discharge` = `End of treamtent ppFEV1`,
+    `Unique number` = `Study ID`
+  ) %>%
+  mutate(
+    Site = "CNH",
     `Modulator at Time of Admit` = case_when(
       None == "Unchecked" ~ "Yes",
-           None =="Checked" ~"No"
+      None == "Checked" ~ "No"
     ),
@@
       MRSA == "Unchecked" ~ NA,
-           MRSA =="Checked" ~ TRUE
+      MRSA == "Checked" ~ TRUE
     ),
@@
       MSSA == "Unchecked" ~ NA,
-           MSSA =="Checked" ~ TRUE
+      MSSA == "Checked" ~ TRUE
     ),
@@
       `P. aeruginosa` == "Unchecked" ~ NA,
-           `P. aeruginosa` =="Checked" ~ TRUE
+      `P. aeruginosa` == "Checked" ~ TRUE
     ),
     Genotype = case_when(
-           `CFTR genotype, groups` =="F508del heterozygous" ~ "F508del/Other",
-           `CFTR genotype, groups` =="F508del homozygous"~"F508del/F508del",
-           `CFTR genotype, groups` =="Other"~"Other/Other"
+      `CFTR genotype, groups` == "F508del heterozygous" ~ "F508del/Other",
+      `CFTR genotype, groups` == "F508del homozygous" ~ "F508del/F508del",
+      `CFTR genotype, groups` == "Other" ~ "Other/Other"
     ),
@@
       `Ivacaftor/lumacaftor (Orkambi)` == "Checked" ~ "Other",
-           `Tezacaftor/ivacaftor (Symdeko)`=="Checked" ~ "Other",
-           `Elexacaftor/tezacaftor/ivacaftor (Trikafta)`=="Checked" ~ "Trikafta",
+      `Tezacaftor/ivacaftor (Symdeko)` == "Checked" ~ "Other",
+      `Elexacaftor/tezacaftor/ivacaftor (Trikafta)` == "Checked" ~ "Trikafta",
       .default = "None"
     ),
-         `FEV1 % Predicted at Admit` = as.numeric(`PEx ppFEV1`))
+    `FEV1 % Predicted at Admit` = as.numeric(`PEx ppFEV1`)
+  )
 # Merge
-df = full_join(chco,cnh)
+df <- full_join(chco, cnh)
 # Format column types
-df$`Modulator at Time of Admit` = factor(df$`Modulator at Time of Admit`,
-                                         levels = c("No","Yes"),ordered = T)
-df$`Specific Modulator at Admit`[df$`Modulator at Time of Admit`=="No"] = "None"
-df$`Specific Modulator at Admit` = factor(tolower(df$`Specific Modulator at Admit`),
-                                          levels = c("kalydeco","none","orkambi","symdeko","trikafta"),
-                                          labels = c("Other","None","Other","Other","Trikafta"))
-df$Race = factor(df$Race,levels = c("Amer. Indian or Alaskan Naitive","Asian",
-                                    "Black","Black or African American","Black/ AA",
-                                    "Decline to Answer","Hispanic","Other",
-                                    "Unknown / Not Reported","White"),
-                 labels = c("Other","Other","Black","Black","Black",
-                            "Unknown / Not Reported","Other","Other",
-                            "Unknown / Not Reported","White"))
-df$Ethnicity = factor(df$Ethnicity,levels = c("FALSE","Hispanic or Latino","HLS",
-                                              "NOT Hispanic or Latino","NOT HLS",
-                                              "TRUE","UNKNOWN"),
-                      labels = c("NOT Hispanic or Latino","Hispanic or Latino",
-                                 "Hispanic or Latino","NOT Hispanic or Latino",
-                                 "NOT Hispanic or Latino","Hispanic or Latino",
-                                 "Unknown"))
-bugs = c("P. aeruginosa","P. aeruginosa mucoid","P. aeruginosa non-mucoid","Staph aureus","MRSA")
-df[,bugs] = lapply(df[,bugs],function(c){
-  c[is.na(c)] = "No"
-  c = factor(c,levels = c("No",TRUE),labels = c("No","Yes"),ordered = T)
+df$`Modulator at Time of Admit` <- factor(df$`Modulator at Time of Admit`,
+  levels = c("No", "Yes"), ordered = T
+)
+df$`Specific Modulator at Admit`[df$`Modulator at Time of Admit` == "No"] <- "None"
+df$`Specific Modulator at Admit` <- factor(tolower(df$`Specific Modulator at Admit`),
+  levels = c("kalydeco", "none", "orkambi", "symdeko", "trikafta"),
+  labels = c("Other", "None", "Other", "Other", "Trikafta")
+)
+df$Race <- factor(df$Race,
+  levels = c(
+    "Amer. Indian or Alaskan Naitive", "Asian",
+    "Black", "Black or African American", "Black/ AA",
+    "Decline to Answer", "Hispanic", "Other",
+    "Unknown / Not Reported", "White"
+  ),
+  labels = c(
+    "Other", "Other", "Black", "Black", "Black",
+    "Unknown / Not Reported", "Other", "Other",
+    "Unknown / Not Reported", "White"
+  )
+)
+df$Ethnicity <- factor(df$Ethnicity,
+  levels = c(
+    "FALSE", "Hispanic or Latino", "HLS",
+    "NOT Hispanic or Latino", "NOT HLS",
+    "TRUE", "UNKNOWN"
+  ),
+  labels = c(
+    "NOT Hispanic or Latino", "Hispanic or Latino",
+    "Hispanic or Latino", "NOT Hispanic or Latino",
+    "NOT Hispanic or Latino", "Hispanic or Latino",
+    "Unknown"
+  )
+)
+bugs <- c("P. aeruginosa", "P. aeruginosa mucoid", "P. aeruginosa non-mucoid", "Staph aureus", "MRSA", "S. maltophilia")
+df[, bugs] <- lapply(df[, bugs], function(c) {
+  c[is.na(c)] <- "No"
+  c <- factor(c, levels = c("No", TRUE), labels = c("No", "Yes"), ordered = T)
 })
+df$`S. maltophilia`[df$Site == "CNH"] <- NA
 # Make summary table
-summ = df %>% group_by(`Unique number`,Site,Year) %>%
+summ <- df %>%
+  group_by(`Unique number`, Site, Year) %>%
   mutate(`Change in FEV1` = `FEV1 % Predicted at Discharge` - `FEV1 % Predicted at Admit`) %>%
-  summarise(`# of Hospitalizations per Person` = n(),
-            across(c(Sex:Genotype,`Specific Modulator at Admit`),~first(na.omit(.x))),
-            across(c(`Modulator at Time of Admit`,all_of(bugs)),~max(.x,na.rm = T)),
-            across(all_of(c("Age at Admit","# Days in Admit","Change in FEV1",
-                            "FEV1 % Predicted at Admit","FEV1 % Predicted at Discharge")),
-                   ~mean(.x,na.rm=T)),
-            .groups = "drop")  %>%
+  summarise(
+    `# of Hospitalizations per Person` = n(),
+    across(c(Sex:Genotype, `Specific Modulator at Admit`), ~ first(na.omit(.x))),
+    across(c(`Modulator at Time of Admit`, all_of(bugs)), ~ max(.x, na.rm = T)),
+    across(
+      all_of(c(
+        "Age at Admit", "# Days in Admit", "Change in FEV1",
+        "FEV1 % Predicted at Admit", "FEV1 % Predicted at Discharge"
+      )),
+      ~ mean(.x, na.rm = T)
+    ),
+    .groups = "drop"
+  ) %>%
   mutate(Genotype = factor(Genotype,
-                           levels=c("F508del/F508del","F508del/Other",
-                                    "Other/F508del","Other/Other"),
-                           labels = c("F508del/F508del","F508del/Other",
-                                      "F508del/Other","Other/Other")
+    levels = c(
+      "F508del/F508del", "F508del/Other",
+      "Other/F508del", "Other/Other"
+    ),
+    labels = c(
+      "F508del/F508del", "F508del/Other",
+      "F508del/Other", "Other/Other"
+    )
   )) %>%
-  select(`Unique number`, `Site`,Year,`# of Hospitalizations per Person`,`# Days in Admit`,`Age at Admit`,
-         `Change in FEV1`,`FEV1 % Predicted at Admit`,`FEV1 % Predicted at Discharge`,
-         `Modulator at Time of Admit`,`Specific Modulator at Admit`,
-         all_of(bugs),Sex:Genotype)
+  select(
+    `Unique number`, `Site`, Year, `# of Hospitalizations per Person`, `# Days in Admit`, `Age at Admit`,
+    `Change in FEV1`, `FEV1 % Predicted at Admit`, `FEV1 % Predicted at Discharge`,
+    `Modulator at Time of Admit`, `Specific Modulator at Admit`,
+    all_of(bugs), Sex:Genotype
+  )
 ```
@@
 summ %>%
-  select(-`Unique number`,-Site) %>%
-  tbl_summary(by=Year,
-              type = list(`# of Hospitalizations per Person` ~ 'continuous'),
+  select(-`Unique number`, -Site) %>%
+  tbl_summary(
+    by = Year,
+    type = list(`# of Hospitalizations per Person` ~ "continuous"),
     statistic = list(
-                c("# of Hospitalizations per Person","# Days in Admit") ~ "{median} ({p25}, {p75})",
-                c("Age at Admit","Change in FEV1","FEV1 % Predicted at Admit",
-                  "FEV1 % Predicted at Discharge") ~ "{mean} ({sd})"
+      c("# of Hospitalizations per Person", "# Days in Admit") ~ "{median} ({p25}, {p75})",
+      c(
+        "Age at Admit", "Change in FEV1", "FEV1 % Predicted at Admit",
+        "FEV1 % Predicted at Discharge"
+      ) ~ "{mean} ({sd})"
     ),
-              missing_text = "(Missing)") %>%
+    missing_text = "(Missing)"
+  ) %>%
   add_ci(pattern = "{stat} [{ci}]") %>%
   add_p(test = list(
-    `Age at Admit`~"t.test",
-    `Change in FEV1`~"t.test"
+    `Age at Admit` ~ "t.test",
+    `Change in FEV1` ~ "t.test"
   )) %>%
@@
 summ %>%
-  select(-`Unique number`) %>%
+  select(-`Unique number`, -`S. maltophilia`) %>%
   tbl_strata(
@@
       ~ .x %>%
-      tbl_summary(by=Year,
-              type = list(`# of Hospitalizations per Person` ~ 'continuous'),
+        tbl_summary(
+          by = Year,
+          type = list(`# of Hospitalizations per Person` ~ "continuous"),
           statistic = list(
-                c("# of Hospitalizations per Person","# Days in Admit") ~ "{median} ({p25}, {p75})",
-                c("Age at Admit","Change in FEV1","FEV1 % Predicted at Admit",
-                  "FEV1 % Predicted at Discharge") ~ "{mean} ({sd})"
+            c("# of Hospitalizations per Person", "# Days in Admit") ~ "{median} ({p25}, {p75})",
+            c(
+              "Age at Admit", "Change in FEV1", "FEV1 % Predicted at Admit",
+              "FEV1 % Predicted at Discharge"
+            ) ~ "{mean} ({sd})"
           ),
-              missing_text = "(Missing)") %>%
+          missing_text = "(Missing)"
+        ) %>%
         add_ci(pattern = "{stat} [{ci}]") %>%
         add_p(test = list(
-    `Age at Admit`~"t.test",
-    `Change in FEV1`~"t.test"
+          `Age at Admit` ~ "t.test",
+          `Change in FEV1` ~ "t.test"
         )) %>%

### 2023-03-29 Jordana Hoppe/nacfc_2023.qmd
Update nacfc_2023.qmd
tidyverse formatting

@@
 library(gtsummary)
-df = read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/clinical.csv",na.strings = "")
-surveys = read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/surveys.csv",na.strings = "")
+df <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/clinical.csv", na.strings = "")
+surveys <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/surveys.csv", na.strings = "")
 # Combined genotype
-df$Mutation.1[df$Mutation.1 != "F508del"]="Other"
-df$Mutation.2[df$Mutation.2 != "F508del"]="Other"
-df$Genotype = paste(df$Mutation.1,df$Mutation.2,sep = "/")
+df$Mutation.1[df$Mutation.1 != "F508del"] <- "Other"
+df$Mutation.2[df$Mutation.2 != "F508del"] <- "Other"
+df$Genotype <- paste(df$Mutation.1, df$Mutation.2, sep = "/")
 # Age
-df$Age = difftime(df$Date.of.clinic.culture.collection, df$Patient.DOB)/365.25
+df$Age <- difftime(df$Date.of.clinic.culture.collection, df$Patient.DOB) / 365.25
 # Combine race columns
-df = df %>% rowwise() %>%
-  mutate(Race = paste(which(c_across(Race..choice.American.Indian.or.Alaska.Native.:
-                                       Race..choice.Unknown.or.not.reported.)== "Checked"),
-                      collapse = "+"),
-         Race = factor(Race,levels = c("","1","5"),labels=c("Unknown","American Indian or Alaska Native","White")))
+df <- df %>%
+  rowwise() %>%
+  mutate(
+    Race = paste(
+      which(c_across(Race..choice.American.Indian.or.Alaska.Native.:
+      Race..choice.Unknown.or.not.reported.) == "Checked"),
+      collapse = "+"
+    ),
+    Race = factor(Race, levels = c("", "1", "5"), labels = c("Unknown", "American Indian or Alaska Native", "White"))
+  )
 # Shorten organism names
-colnames(df) = sub("If.culture.was.positive..select.all.organisms.isolated...choice.","",colnames(df))
+colnames(df) <- sub("If.culture.was.positive..select.all.organisms.isolated...choice.", "", colnames(df))
 ```
@@
 #| label: tbl-1
-df %>% select(Age,FEV1...predicted,FVC...predicted,
-              Sex,Race,Hispanic.or.Latinx,Rural.location.) %>%
+df %>%
+  select(
+    Age, FEV1...predicted, FVC...predicted,
+    Sex, Race, Hispanic.or.Latinx, Rural.location.
+  ) %>%
   tbl_summary(
@@
       Age ~ "Age at clinic culture collection",
-      FEV1...predicted~"ppFEV",FVC...predicted~"ppFVC",
-      Hispanic.or.Latinx~"Hispanic or Latinx",
-      Rural.location. ~"Rural location?"
+      FEV1...predicted ~ "ppFEV", FVC...predicted ~ "ppFVC",
+      Hispanic.or.Latinx ~ "Hispanic or Latinx",
+      Rural.location. ~ "Rural location?"
     ),
@@
       all_continuous() ~ "{mean} ({sd})"
-    ))
+    )
+  )
 ```
@@
 ```{r}
-sens_spec = function(gold_standard,other_test,bug,type = "Home Collection"){
-  t = table(other_test,gold_standard)
-  if(all(dim(t)==c(1,1))){
+sens_spec <- function(gold_standard, other_test, bug, type = "Home Collection") {
+  t <- table(other_test, gold_standard)
+  if (all(dim(t) == c(1, 1))) {
     cat("\n")
@@
     cat("\n")
-    cat("All samples marked as",unique(gold_standard))
+    cat("All samples marked as", unique(gold_standard))
     cat("\n")
   } else {
-    sen = sensitivity(data = factor(other_test),reference = factor(gold_standard))
-    spec = specificity(data = factor(other_test),reference = factor(gold_standard))
+    sen <- sensitivity(data = factor(other_test), reference = factor(gold_standard))
+    spec <- specificity(data = factor(other_test), reference = factor(gold_standard))
 
-    dimnames(t)=lapply(dimnames(t),function(n){
-      n = sub("Checked","Present",n)
-      n = sub("Unchecked","Absent",n)
+    dimnames(t) <- lapply(dimnames(t), function(n) {
+      n <- sub("Checked", "Present", n)
+      n <- sub("Unchecked", "Absent", n)
     })
-    dimnames(t)$other_test = paste(type,dimnames(t)$other_test)
-    dimnames(t)$gold_standard = paste("Clinic",dimnames(t)$gold_standard)
+    dimnames(t)$other_test <- paste(type, dimnames(t)$other_test)
+    dimnames(t)$gold_standard <- paste("Clinic", dimnames(t)$gold_standard)
 
@@
     cat("\n")
-    cat("Sensitivity of the",tolower(type),"method was",round(sen,3),"and specificity was",round(spec,3))
+    cat("Sensitivity of the", tolower(type), "method was", round(sen, 3), "and specificity was", round(spec, 3))
     cat("\n")
@@
   df$Methicillin.sensitive.Staph.aureus..1,
-          type = "Home Collection",bug = "MSSA")
+  type = "Home Collection", bug = "MSSA"
+)
 ```
@@
   df$Methicillin.resistant.Staph.aureus..1,
-          type = "Home Collection",bug = "MRSA")
+  type = "Home Collection", bug = "MRSA"
+)
 ```
@@
 ```{r results='asis'}
-sens_spec(df$A..xylosoxidans.,df$A..xylosoxidans..1,
-          type = "Home Collection",bug = "A. xylosoxidans")
+sens_spec(df$A..xylosoxidans., df$A..xylosoxidans..1,
+  type = "Home Collection", bug = "A. xylosoxidans"
+)
 ```
@@
 ```{r results='asis'}
-sens_spec(df$Achromobacter.not.xylosoxidans.,df$Achromobacter.not.xylosoxidans..1,
-          type = "Home Collection",bug = "Achromobacter (not xylosoxidans)")
+sens_spec(df$Achromobacter.not.xylosoxidans., df$Achromobacter.not.xylosoxidans..1,
+  type = "Home Collection", bug = "Achromobacter (not xylosoxidans)"
+)
 ```
@@
 ```{r results='asis'}
-sens_spec(df$Acinetobacter.,df$Acinetobacter..1,
-          type = "Home Collection",bug = "Acinetobacter")
+sens_spec(df$Acinetobacter., df$Acinetobacter..1,
+  type = "Home Collection", bug = "Acinetobacter"
+)
 ```
@@
 ```{r results='asis'}
-sens_spec(df$H..influenzae.,df$H..influenzae..1,
-          type = "Home Collection",bug = "H. influenzae")
+sens_spec(df$H..influenzae., df$H..influenzae..1,
+  type = "Home Collection", bug = "H. influenzae"
+)
 ```
@@
 ```{r results='asis'}
-sens_spec(df$P..aeruginosa.,df$P..aeruginosa..1,
-          type = "Home Collection",bug = "P. aeruginosa")
+sens_spec(df$P..aeruginosa., df$P..aeruginosa..1,
+  type = "Home Collection", bug = "P. aeruginosa"
+)
 ```
@@
 ```{r results='asis'}
-sens_spec(df$P..aeruginosa..mucoid..,df$P..aeruginosa..mucoid...1,
-          type = "Home Collection",bug = "P. aeruginosa (mucoid)")
+sens_spec(df$P..aeruginosa..mucoid.., df$P..aeruginosa..mucoid...1,
+  type = "Home Collection", bug = "P. aeruginosa (mucoid)"
+)
 ```
@@
 ```{r results='asis'}
-sens_spec(df$Pseudomonas.species..not.aeruginosa.,df$Pseudomonas.species..not.aeruginosa..1,
-          type = "Home Collection",bug = "Pseudomonas (not aeruginosa)")
+sens_spec(df$Pseudomonas.species..not.aeruginosa., df$Pseudomonas.species..not.aeruginosa..1,
+  type = "Home Collection", bug = "Pseudomonas (not aeruginosa)"
+)
 ```
@@
 ```{r results='asis'}
-sens_spec(df$B..cepacia.complex.,df$B..cepacia.complex..1,
-          type = "Home Collection",bug = "B. cepacia complex")
+sens_spec(df$B..cepacia.complex., df$B..cepacia.complex..1,
+  type = "Home Collection", bug = "B. cepacia complex"
+)
 ```
@@
 ```{r results='asis'}
-sens_spec(df$Stenotrophomonas.,df$Stenotrophomonas..1,
-          type = "Home Collection",bug = "Stenotrophomonas")
+sens_spec(df$Stenotrophomonas., df$Stenotrophomonas..1,
+  type = "Home Collection", bug = "Stenotrophomonas"
+)
 ```
@@
 ```{r results='asis'}
-sens_spec(df$Aspergillus.,df$Aspergillus..1,
-          type = "Home Collection",bug = "Aspergillus")
+sens_spec(df$Aspergillus., df$Aspergillus..1,
+  type = "Home Collection", bug = "Aspergillus"
+)
 ```
@@
 ```{r results='asis'}
-sens_spec(df$Penicillium.species.,df$Penicillium.species..1,
-          type = "Home Collection",bug = "Penicillium species")
+sens_spec(df$Penicillium.species., df$Penicillium.species..1,
+  type = "Home Collection", bug = "Penicillium species"
+)
 ```
@@
 ```{r results='asis'}
-sens_spec(df$Sceosporium.species.,df$Sceosporium.species..1,
-          type = "Home Collection",bug = "Sceosporium species")
+sens_spec(df$Sceosporium.species., df$Sceosporium.species..1,
+  type = "Home Collection", bug = "Sceosporium species"
+)
 ```
@@
 ```{r results='asis'}
-sens_spec(df$M..avium.complex.,df$M..avium.complex..1,
-          type = "Home Collection",bug = "M. avium complex")
+sens_spec(df$M..avium.complex., df$M..avium.complex..1,
+  type = "Home Collection", bug = "M. avium complex"
+)
 ```
@@
 ```{r results='asis'}
-sens_spec(df$M..abscessus.,df$M..abscessus..1,
-          type = "Home Collection",bug = "M. abscessus")
+sens_spec(df$M..abscessus., df$M..abscessus..1,
+  type = "Home Collection", bug = "M. abscessus"
+)
 ```

### 2023-03-30 Christine Chan/Hypoglycemia/tim_analysis.Rmd
Create tim_analysis.Rmd

---
title: "Hypoglycemia Analysis"
output:
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
    fig_height: 8
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nlme)
library(emmeans)
library(patchwork)
library(ggsignif)
library(arsenal)
library(zoo)
library(vcd)
library(kableExtra)
```

```{r}
getAUCs <- function(data,subject,time,vars){
  all_return <- NULL

  for (var in vars){
    subs <- unique(data[[subject]])
    return <- NULL
    for (sub in subs){
      try({
        df <- data[data[[subject]] == sub,]
        ogtt_0 <- df[df[[time]] == 0,var]
        ogtt_30 <- df[df[[time]] == 30,var]
        ogtt_60 <- df[df[[time]] == 60,var]
        ogtt_90 <- df[df[[time]] == 90,var]
        ogtt_120 <- df[df[[time]] == 120,var]
        ogtt_auc <- (ogtt_30[[var]] + ogtt_60[[var]] + ogtt_90[[var]])/2 + (ogtt_0[[var]] + ogtt_120[[var]])/4
        if(length(ogtt_auc) > 0){
          row <- c(sub,ogtt_auc)
          return <- rbind(return, row)
        }

      })

    }
    return <- as.data.frame(return)
    names(return) <- c(subject,paste0(var,"_auc"))
    if (is.null(all_return)){
      all_return <- return
    }
    else{
      all_return <- merge(all_return, return, all = T)
    }

  }
  all_return

}

```

```{r, warning = F, message = F}
dat <- read.csv("./data/CGMInHealthyControls_DATA_2021-05-25_0918.csv", na = "-999")

dat$group_category[dat$group_category == "1"] <- "Healthy"
dat$group_category[dat$group_category == "2"] <- "CF-NGT"
dat$group_category[dat$group_category == "3"] <- "CF-AG"
dat$group_category[dat$group_category == "4"] <- "CFRD"
dat$group_category <- factor(dat$group_category, levels = c("Healthy","CF-NGT","CF-AG","CFRD"))

dat$minute_60_glucose <- dat$lab_ogtt_1_hour_glucose
dat$minute_120_glucose <- dat$lab_ogtt_2_hour_glucose
dat$minute_0_glucose <- dat$lab_ogtt_fasting

dat$weight[is.na(dat$weight)] <- dat$weight_visit_2[is.na(dat$weight)]
dat$height[is.na(dat$height)] <- dat$height_visit_2[is.na(dat$height)]
dat$bmi[is.na(dat$bmi)] <- dat$bmi_visit_2[is.na(dat$bmi)]
dat$bmi_z_score[is.na(dat$bmi_z_score)] <- dat$bmi_z_score_visit_2[is.na(dat$bmi_z_score)]

dat <- dat[dat$date_of_visit != "",]

dat$subject_id <- gsub("*--1", "", dat$subject_id)
dat$subject_id <- gsub("*--2", "", dat$subject_id)
dat$subject_id <- gsub("*\\.1", "", dat$subject_id)
dat$subject_id <- gsub("*\\.2", "", dat$subject_id)
dat <- unique(dat)

dat$age <- as.numeric(dat$age)

dat$gender[dat$gender == 1] <- "Male"
dat$gender[dat$gender == 2] <- "Female"

dat$race_ethnicity[dat$race_ethnicity == 1] <- "White"
dat$race_ethnicity[dat$race_ethnicity == 2] <- "Black/African American"
dat$race_ethnicity[dat$race_ethnicity == 3] <- "Asian"
dat$race_ethnicity[dat$race_ethnicity == 5] <- "Hispanic"
dat$race_ethnicity[dat$race_ethnicity == 6] <- "Other/Multiple"

dat$tanner_stage[dat$tanner_stage___1 == 1] <- 1
dat$tanner_stage[dat$tanner_stage_male_testicul___1 == 1] <- 1
dat$tanner_stage[dat$tanner_stage_female_pubic___1 == 1] <- 1
dat$tanner_stage[dat$tanner_stage_female_breast___1 == 1] <- 1

dat$tanner_stage[dat$tanner_stage___2 == 1] <- 2
dat$tanner_stage[dat$tanner_stage_male_testicul___2 == 1] <- 2
dat$tanner_stage[dat$tanner_stage_female_pubic___2 == 1] <- 2
dat$tanner_stage[dat$tanner_stage_female_breast___2 == 1] <- 2

dat$tanner_stage[dat$tanner_stage___3 == 1] <- 3
dat$tanner_stage[dat$tanner_stage_male_testicul___3 == 1] <- 3
dat$tanner_stage[dat$tanner_stage_female_pubic___3 == 1] <- 3
dat$tanner_stage[dat$tanner_stage_female_breast___3 == 1] <- 3

dat$tanner_stage[dat$tanner_stage___4 == 1] <- 4
dat$tanner_stage[dat$tanner_stage_male_testicul___4 == 1] <- 4
dat$tanner_stage[dat$tanner_stage_female_pubic___4 == 1] <- 4
dat$tanner_stage[dat$tanner_stage_female_breast___4 == 1] <- 4

dat$tanner_stage[dat$tanner_stage___5 == 1] <- 5
dat$tanner_stage[dat$tanner_stage_male_testicul___5 == 1] <- 5
dat$tanner_stage[dat$tanner_stage_female_pubic___5 == 1] <- 5
dat$tanner_stage[dat$tanner_stage_female_breast___5 == 1] <- 5

dat$g_tube_feeds[dat$g_tube_feeds == 1] <- "Yes"
dat$g_tube_feeds[dat$g_tube_feeds == 0] <- "No"

gluc <- dat %>% select(subject_id, 
                       date_of_visit, 
                       age, 
                       gender, 
                       race_ethnicity, 
                       starts_with("min"),
                       -starts_with("min_spent"),
                       -contains("sensor"),
                       -contains("active"), 
                       -contains("glp"),
                       -contains("gip"))
gluc <- merge(gluc, dat %>% select(subject_id, date_of_visit, age, starts_with("ucd_")))
gluc <- unique(gluc)
gluc <- gluc[!is.na(gluc$min_30_glucose),]
gluc[] <- lapply(gluc, gsub, pattern="<", replacement = "")

gluc_names <- names(gluc)
gluc_names <- gsub("min_","",gluc_names)
gluc_names <- gsub("minute_","",gluc_names)
gluc_names <- gsub("minus_","-",gluc_names)
gluc_names <- gsub("ucd_","",gluc_names)
gluc_names <- gsub("_total","",gluc_names)
gluc_names <- gsub("glp_1","_glp1",gluc_names)
gluc_names <- gsub("c_peptide","cpeptide",gluc_names)
names(gluc) <- gluc_names
gluc[gluc == -999.99] <- NA
gluc <- unique(gluc)
new_gluc <- gluc[0,]

for (sub in unique(gluc$subject_id)){
  df <- gluc[gluc$subject_id == sub,]
  if (nrow(df) > 1){
    df <- df %>% arrange(`0_gip`,desc(date_of_visit))
    df <- df[1,]
  }
  new_gluc <- rbind(new_gluc,df)
}

gluc_wide <- new_gluc %>% select(-date_of_visit)
gluc <- gluc_wide %>% pivot_longer(cols = -c(subject_id), 
                                   names_to = c("time", ".value"), names_sep = "_")

#gluc <- gluc[!is.na(gluc$glucose) | !is.na(gluc$glucagon) | !is.na(gluc$insulin)| !is.na(gluc$cpeptide),]
#gluc <- gluc[!is.na(gluc$glucagon),]
gluc <- gluc[gluc$time != "-10",]
gluc$time <- as.factor(as.numeric(gluc$time))

gluc$glucose <- as.numeric(gluc$glucose)
gluc$insulin <- as.numeric(gluc$insulin)
gluc$glucagon <- as.numeric(gluc$glucagon)
gluc$cpeptide <- as.numeric(gluc$cpeptide)
gluc$gip <- as.numeric(gluc$gip)
gluc$glp1 <- as.numeric(gluc$glp1)

gluc$gip[gluc$gip < 0] <- NA
gluc$glp1[gluc$glp1 < 0] <- NA

gluc <- gluc %>% filter(!is.na(time))

aucs <- getAUCs(gluc, "subject_id","time",c("glucose","glucagon","insulin","cpeptide","gip","glp1"))

df_aucs <- merge(aucs,unique(gluc %>% select(subject_id)))
df_aucs$glucagon_auc <- as.numeric(df_aucs$glucagon_auc)
df_aucs$glucose_auc <- as.numeric(df_aucs$glucose_auc)
df_aucs$insulin_auc <- as.numeric(df_aucs$insulin_auc)
df_aucs$cpeptide_auc <- as.numeric(df_aucs$cpeptide_auc)
df_aucs$gip_auc <- as.numeric(df_aucs$gip_auc)
df_aucs$glp1_auc <- as.numeric(df_aucs$glp1_auc)

df_aucs <- df_aucs[!is.na(df_aucs$glucagon_auc) | !is.na(df_aucs$glucose_auc) |!is.na(df_aucs$insulin_auc) |!is.na(df_aucs$cpeptide_auc),]

df_aucs <- unique(df_aucs)

gluc_aucs <- df_aucs

gluc$hypo_60 <- ifelse(gluc$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 60])),"Yes","No")
gluc$hypo_70 <- ifelse(gluc$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 70])),"Yes","No")

gluc_aucs <- merge(gluc_aucs,unique(gluc %>% select(subject_id, hypo_60,hypo_70)))

glucose <- gluc_wide %>% select(subject_id, contains("glucose"),-starts_with("-10"))
glucose <- glucose[complete.cases(glucose),]
glucose[,2:10] <- sapply(glucose[,2:10], as.numeric)
glucose$glucose_AUC <- rowSums(glucose[,2:10])
glucose$glucose_mean <- glucose$glucose_AUC/9

insulin <- gluc_wide %>% select(subject_id, contains("insulin"),-starts_with("-10"))
insulin <- insulin[complete.cases(insulin),]
insulin[,2:10] <- sapply(insulin[,2:10], as.numeric)
insulin$insulin_AUC <- rowSums(insulin[,2:10])
insulin$insulin_mean <- insulin$insulin_AUC/9

cpeptide <- gluc_wide %>% select(subject_id, contains("cpeptide"),-starts_with("-10"))
cpeptide <- cpeptide[complete.cases(cpeptide),]
cpeptide[,2:10] <- sapply(cpeptide[,2:10], as.numeric)
cpeptide$cpeptide_AUC <- rowSums(cpeptide[,2:10])
cpeptide$cpeptide_mean <- cpeptide$cpeptide_AUC/9

glucagon <- gluc_wide %>% select(subject_id, contains("glucagon"),-starts_with("-10"))
glucagon <- glucagon[complete.cases(glucagon),]
glucagon[,2:10] <- sapply(glucagon[,2:10], as.numeric)
glucagon$glucagon_AUC <- rowSums(glucagon[,2:10])
glucagon$glucagon_mean <- glucagon$glucagon_AUC/9

gip <- gluc_wide %>% select(subject_id, contains("gip"),-starts_with("minus"))
gip <- gip[complete.cases(gip),]
gip[,2:10] <- sapply(gip[,2:10], as.numeric)
gip$gip_AUC <- rowSums(gip[,2:10])
gip$gip_mean <- gip$gip_AUC/9

glp1 <- gluc_wide %>% select(subject_id, contains("glp"),-starts_with("minus"))
glp1 <- glp1[complete.cases(glp1),]
glp1[,2:10] <- sapply(glp1[,2:10], as.numeric)
glp1$glp1_AUC <- rowSums(glp1[,2:10])
glp1$glp1_mean <- glp1$glp1_AUC/9

auc_df <- merge(glucose, insulin, all = T)
auc_df <- merge(auc_df, cpeptide, all = T)
auc_df <- merge(auc_df, glucagon, all = T)
auc_df <- merge(auc_df, gip, all = T)
auc_df <- merge(auc_df, glp1, all = T)

auc_df$hypo_60 <- ifelse(auc_df$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 60])),"Yes","No")
auc_df$hypo_70 <- ifelse(auc_df$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 70])),"Yes","No")

auc_df$matsuda_ins <- 10000/sqrt(auc_df$`0_glucose`*auc_df$`0_insulin`*auc_df$glucose_mean*auc_df$insulin_mean)
auc_df$matsuda_cpep <- 20/(auc_df$`0_cpeptide`*auc_df$`0_glucose`)
auc_df$homa_ir <- auc_df$`0_glucose`*auc_df$`0_insulin`/405
auc_df$igi <- (auc_df$`30_insulin` - auc_df$`0_insulin`)/(auc_df$`30_glucose` - auc_df$`0_glucose`)
auc_df$igi[is.infinite(auc_df$igi)] <- NA
auc_df$cpep_index <- 100*auc_df$`0_cpeptide`/auc_df$`0_glucose`
auc_df$odi_ins <- auc_df$igi * auc_df$matsuda_ins
auc_df$odi_cpep <- auc_df$igi * auc_df$matsuda_cpep
auc_df$cpep_ins0 <- auc_df$`0_cpeptide`/auc_df$`0_insulin`
auc_df$cpep_auc_30 <- (auc_df$`0_cpeptide`+auc_df$`10_cpeptide`+auc_df$`20_cpeptide`+auc_df$`30_cpeptide`)/4
auc_df$ins_auc_30 <- (auc_df$`0_insulin`+auc_df$`10_insulin`+auc_df$`20_insulin`+auc_df$`30_insulin`)/4
auc_df$cpep_ins_auc_30 <- auc_df$cpep_auc_30/auc_df$ins_auc_30
auc_df$cpep_ins_auc <- auc_df$cpeptide_AUC/auc_df$insulin_AUC

gluc_peak <- NULL
gluc_nadir <- NULL
gluc_peak_time <- NULL
gluc_nadir_time <- NULL

ins_peak <- NULL
ins_nadir <- NULL
ins_peak_time <- NULL
ins_nadir_time <- NULL

cpep_peak <- NULL
cpep_nadir <- NULL
cpep_peak_time <- NULL
cpep_nadir_time <- NULL

gluca_peak <- NULL
gluca_nadir <- NULL
gluca_peak_time <- NULL
gluca_nadir_time <- NULL

gip_peak <- NULL
gip_nadir <- NULL
gip_peak_time <- NULL
gip_nadir_time <- NULL

glp1_peak <- NULL
glp1_nadir <- NULL
glp1_peak_time <- NULL
glp1_nadir_time <- NULL

for (subject in auc_df$subject_id){
  sub_df <- auc_df %>% filter(subject_id == subject) %>% select(ends_with("glucose")) 
  gluc_peak <- c(gluc_peak,max(sub_df))
  gluc_nadir <- c(gluc_nadir,min(sub_df))
  gluc_peak_time <- c(gluc_peak_time,
                      strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")[[1]][1])
  gluc_nadir_time <- c(gluc_nadir_time,
                       strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")[[1]][1])

  sub_df <- auc_df %>% filter(subject_id == subject) %>% select(ends_with("insulin")) 
  ins_peak <- c(ins_peak,max(sub_df))
  ins_nadir <- c(ins_nadir,min(sub_df))
  ins_peak_time <- c(ins_peak_time,
                     strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")[[1]][1])
  ins_nadir_time <- c(ins_nadir_time,
                      strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")[[1]][1])

  sub_df <- auc_df %>% filter(subject_id == subject) %>% select(ends_with("cpeptide")) 
  cpep_peak <- c(cpep_peak,max(sub_df))
  cpep_nadir <- c(cpep_nadir,min(sub_df))

  cppt_temp <- strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")
  cpnt_temp <- strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")
  if(length(cppt_temp) > 0){
    cpep_peak_time <- c(cpep_peak_time,
                        cppt_temp[[1]][1])
  }
  else {
    cpep_peak_time <- c(cpep_peak_time,
                        NA)
  }
  if(length(cpnt_temp) > 0){
    cpep_nadir_time <- c(cpep_nadir_time,
                         cpnt_temp[[1]][1])
  }
  else {
    cpep_nadir_time <- c(cpep_nadir_time,
                         NA)
  }

  sub_df <- auc_df %>% filter(subject_id == subject) %>% select(ends_with("glucagon")) 
  gluca_peak <- c(gluca_peak,max(sub_df))
  gluca_nadir <- c(gluca_nadir,min(sub_df))
  glucap_temp <- strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")
  glucan_temp <- strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")
  if(length(glucap_temp) > 0){
    gluca_peak_time <- c(gluca_peak_time,
                         glucap_temp[[1]][1])
  }
  else {
    gluca_peak_time <- c(gluca_peak_time,
                         NA)
  }
  if(length(glucan_temp) > 0){
    gluca_nadir_time <- c(gluca_nadir_time,
                          glucan_temp[[1]][1])
  }
  else {
    gluca_nadir_time <- c(gluca_nadir_time,
                          NA)
  }

  sub_df <- auc_df %>% filter(subject_id == subject) %>% select(ends_with("gip")) 
  gip_peak <- c(gip_peak,max(sub_df))
  gip_nadir <- c(gip_nadir,min(sub_df))
  gipp_temp <- strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")
  gipn_temp <- strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")
  if(length(gipp_temp) > 0){
    gip_peak_time <- c(gip_peak_time,
                       gipp_temp[[1]][1])
  }
  else {
    gip_peak_time <- c(gip_peak_time,
                       NA)
  }
  if(length(gipn_temp) > 0){
    gip_nadir_time <- c(gip_nadir_time,
                        gipn_temp[[1]][1])
  }
  else {
    gip_nadir_time <- c(gip_nadir_time,
                        NA)
  }

  sub_df <- auc_df %>% filter(subject_id == subject) %>% select(ends_with("glp1")) 
  glp1_peak <- c(glp1_peak,max(sub_df))
  glp1_nadir <- c(glp1_nadir,min(sub_df))
  glp1p_temp <- strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")
  glp1n_temp <- strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")
  if(length(gipp_temp) > 0){
    glp1_peak_time <- c(glp1_peak_time,
                        glp1p_temp[[1]][1])
  }
  else {
    glp1_peak_time <- c(glp1_peak_time,
                        NA)
  }
  if(length(glp1n_temp) > 0){
    glp1_nadir_time <- c(glp1_nadir_time,
                         glp1n_temp[[1]][1])
  }
  else {
    glp1_nadir_time <- c(glp1_nadir_time,
                         NA)
  }
}

auc_df$gluc_peak <- gluc_peak
auc_df$gluc_nadir <- gluc_nadir
auc_df$gluc_peak_time <- gluc_peak_time
auc_df$gluc_nadir_time <- gluc_nadir_time
auc_df$ins_peak <- ins_peak
auc_df$ins_nadir <- ins_nadir
auc_df$ins_peak_time <- ins_peak_time
auc_df$ins_nadir_time <- ins_nadir_time
auc_df$cpep_peak <- cpep_peak
auc_df$cpep_nadir <- cpep_nadir
auc_df$cpep_peak_time <- cpep_peak_time
auc_df$cpep_nadir_time <- cpep_nadir_time
auc_df$gluca_peak <- gluca_peak
auc_df$gluca_nadir <- gluca_nadir
auc_df$gluca_peak_time <- gluca_peak_time
auc_df$gluca_nadir_time <- gluca_nadir_time

auc_df$gip_peak <- gip_peak
auc_df$gip_nadir <- gip_nadir
auc_df$gip_peak_time <- gip_peak_time
auc_df$gip_nadir_time <- gip_nadir_time

auc_df$glp1_peak <- glp1_peak
auc_df$glp1_nadir <- glp1_nadir
auc_df$glp1_peak_time <- glp1_peak_time
auc_df$glp1_nadir_time <- glp1_nadir_time

keep_ids <- gluc_aucs$subject_id

table_dat <- merge(new_gluc,gluc_aucs)
table_dat <- merge(table_dat, auc_df %>% select(subject_id, 
                                                matsuda_ins, 
                                                matsuda_cpep, 
                                                homa_ir, 
                                                igi, 
                                                cpep_index,
                                                odi_ins,
                                                odi_cpep,
                                                cpep_ins0,
                                                cpep_ins_auc_30,
                                                cpep_ins_auc,
                                                `0_glucose`,
                                                `0_insulin`,
                                                `0_cpeptide`,
                                                `0_glucagon`,
                                                `0_gip`,
                                                `0_glp1`,
                                                gluc_peak,
                                                gluc_nadir,
                                                gluc_peak_time,
                                                gluc_nadir_time,
                                                ins_peak,
                                                ins_nadir,
                                                ins_peak_time,
                                                ins_nadir_time,
                                                cpep_peak,
                                                cpep_nadir,
                                                cpep_peak_time,
                                                cpep_nadir_time,
                                                gluca_peak,
                                                gluca_nadir,
                                                gluca_peak_time,
                                                gluca_nadir_time,
                                                gip_peak,
                                                gip_nadir,
                                                gip_peak_time,
                                                gip_nadir_time,
                                                glp1_peak,
                                                glp1_nadir,
                                                glp1_peak_time,
                                                glp1_nadir_time
), all = T)

cgm <- dat %>% select(subject_id,
                      date_of_visit,
                      num_days_good_data:interquartile_range
) %>% 
  filter(subject_id %in% keep_ids) %>% 
  filter(!is.na(num_days_good_data)) %>%
  unique()

cgm$daytime_avg_sensor_glucose <- round(as.numeric(cgm$daytime_avg_sensor_glucose),4)
cgm$percent_time_over_140 <- round(as.numeric(cgm$percent_time_over_140),4)
cgm$percent_time_over_180 <- round(as.numeric(cgm$percent_time_over_180),4)
cgm$daytime_sd <- round(as.numeric(cgm$daytime_sd),4)
cgm$cov <- round(as.numeric(cgm$cov),4)
cgm$mag <- round(as.numeric(cgm$mag),4)
cgm <- unique(cgm)
cgm <- cgm[!duplicated(cgm[,1:2]),]
cgm <- cgm[!duplicated(cgm$subject_id),]

table_dat <- merge(table_dat, cgm %>% select(-date_of_visit), all = T)

dat$mutation_severity[dat$cf_mutation_1_class <=3 | dat$cf_mutation_2_class <=3] <-  "Severe"
dat$mutation_severity[is.na(dat$mutation_severity) & (dat$cf_mutation_1_class == 4 | 
                                                        dat$cf_mutation_2_class == 4 |
                                                        dat$cf_mutation_1_class <= 5 | 
                                                        dat$cf_mutation_2_class <=5)] <-  "Mild"
dat$mutation_severity[is.na(dat$mutation_severity) & (dat$cf_mutation_1_class == 6 | 
                                                        dat$cf_mutation_2_class == 6)] <-  "Other"

dat_demo <- dat %>% select(subject_id, 
                           date_of_visit,
                           group_category,
                           gender, 
                           race_ethnicity, 
                           tanner_stage,
                           weight, 
                           height, 
                           bmi, 
                           bmi_z_score, 
                           mutation_severity, 
                           fev1_percent,
                           fvc,
                           poc_a1c,
                           cf_pancreatic,
                           cf_insulin_use,
                           g_tube_feeds)
dat_demo <- dat_demo %>% filter(subject_id %in% keep_ids)

new_demo <- dat_demo[0,]
for (id in unique(dat_demo$subject_id)){
  id_df <- dat_demo %>% filter(subject_id == id) %>% arrange(date_of_visit) %>% na.locf(na.rm = FALSE)
  id_df <- id_df %>% arrange(desc(date_of_visit)) %>% na.locf(na.rm = FALSE)
  new_demo <- rbind(new_demo,id_df)
}

dat_demo <- new_demo %>% select(-date_of_visit) %>% unique()

dat_demo <- dat_demo[!duplicated(dat_demo$subject_id),]

table_dat <- merge(table_dat,dat_demo, all = T)

table_dat$`0_glucose` <- as.numeric(table_dat$`0_glucose`)
table_dat$`120_glucose` <- as.numeric(table_dat$`120_glucose`)
table_dat$`60_glucose` <- as.numeric(table_dat$`60_glucose`)

table_dat$`0_insulin` <- as.numeric(table_dat$`0_insulin`)
table_dat$`120_insulin` <- as.numeric(table_dat$`120_insulin`)
table_dat$`60_insulin` <- as.numeric(table_dat$`60_insulin`)

table_dat$`0_cpeptide` <- as.numeric(table_dat$`0_cpeptide`)
table_dat$`120_cpeptide` <- as.numeric(table_dat$`120_cpeptide`)
table_dat$`60_cpeptide` <- as.numeric(table_dat$`60_cpeptide`)

table_dat$`0_glucagon` <- as.numeric(table_dat$`0_glucagon`)
table_dat$`120_glucagon` <- as.numeric(table_dat$`120_glucagon`)
table_dat$`60_glucagon` <- as.numeric(table_dat$`60_glucagon`)

table_dat$`0_gip` <- as.numeric(table_dat$`0_gip`)
table_dat$`120_gip` <- as.numeric(table_dat$`120_gip`)
table_dat$`60_gip` <- as.numeric(table_dat$`60_gip`)

table_dat$`0_glp1` <- as.numeric(table_dat$`0_glp1`)
table_dat$`120_glp1` <- as.numeric(table_dat$`120_glp1`)
table_dat$`60_glp1` <- as.numeric(table_dat$`60_glp1`)

table_dat$ogtt_category[table_dat$`0_glucose` < 100 & table_dat$`60_glucose` < 155 & table_dat$`120_glucose` < 140] <- "NGT"
table_dat$ogtt_category[table_dat$`0_glucose` < 100 & table_dat$`60_glucose` > 155 & 
                          table_dat$`120_glucose` < 140 & table_dat$`60_glucose` < 200] <- "AGT" # previously EGI
table_dat$ogtt_category[(table_dat$`0_glucose` >= 100 & table_dat$`0_glucose` < 125) | 
                          (table_dat$`60_glucose` > 199) | 
                          (table_dat$`120_glucose` > 140 & table_dat$`120_glucose` <= 200)] <- "AGT" # previously IFG/IGT/INDET
table_dat$ogtt_category[table_dat$`0_glucose` >= 125 | table_dat$`120_glucose` >= 200] <- "CFRD"

cf_groups <- dat %>% select(subject_id,group_category) %>% filter(subject_id %in% keep_ids) %>% na.omit() %>% unique()
cf_groups$CF <- ifelse(cf_groups$group_category == "Healthy", "No","Yes")
cf_groups <- cf_groups %>% select(subject_id, CF) %>% unique()
table_dat <- merge(table_dat,cf_groups)

table_dat$CF_hypo_60[table_dat$CF == "Yes"  & table_dat$hypo_60 == "Yes"] <- "CF/Hypo"
table_dat$CF_hypo_60[table_dat$CF == "Yes"  & table_dat$hypo_60 == "No"] <- "CF/No-Hypo"
table_dat$CF_hypo_60[table_dat$CF == "No"  & table_dat$hypo_60 == "Yes"] <- "No-CF/Hypo"
table_dat$CF_hypo_60[table_dat$CF == "No"  & table_dat$hypo_60 == "No"] <- "No-CF/No-Hypo"

table_dat$CF_hypo_70[table_dat$CF == "Yes"  & table_dat$hypo_70 == "Yes"] <- "CF/Hypo"
table_dat$CF_hypo_70[table_dat$CF == "Yes"  & table_dat$hypo_70 == "No"] <- "CF/No-Hypo"
table_dat$CF_hypo_70[table_dat$CF == "No"  & table_dat$hypo_70 == "Yes"] <- "No-CF/Hypo"
table_dat$CF_hypo_70[table_dat$CF == "No"  & table_dat$hypo_70 == "No"] <- "No-CF/No-Hypo"

table_dat <- table_dat %>% filter(is.na(cf_pancreatic) | cf_pancreatic == 2)
table_dat$cf_insulin_use[table_dat$cf_insulin_use == 1] <- "Yes"
table_dat$cf_insulin_use[table_dat$cf_insulin_use == 0] <- "No"

table_dat$delta_glucagon <- as.numeric(table_dat$`180_glucagon`) - as.numeric(table_dat$`150_glucagon`)
table_dat$delta_glucose <- as.numeric(table_dat$`180_glucose`) - as.numeric(table_dat$`150_glucose`)
table_dat$delta_insulin <- as.numeric(table_dat$`180_insulin`) - as.numeric(table_dat$`150_insulin`)
table_dat$delta_cpeptide <- as.numeric(table_dat$`180_cpeptide`) - as.numeric(table_dat$`150_cpeptide`)
table_dat$delta_gip <- as.numeric(table_dat$`180_gip`) - as.numeric(table_dat$`150_gip`)
table_dat$delta_glp1 <- as.numeric(table_dat$`180_glp1`) - as.numeric(table_dat$`150_glp1`)

gluc <- merge(gluc,table_dat %>% select(subject_id, CF_hypo_60, CF_hypo_70))

auc_df <- merge(auc_df,table_dat %>% select(subject_id, CF_hypo_60, CF_hypo_70))

table_dat$age <- as.numeric(table_dat$age)

table_dat <- table_dat[!is.na(table_dat$date_of_visit),]

write.csv(table_dat, "analysis_data_incl_cfinsulinuse.csv")

```

# Methods

Global tests by group are Kruskal-Wallis (numeric) or Pearson Chi-square tests (count/categorical)

Pairwise-tests are exact wilcoxon rank sum tests with continuity correction (numeric) or chi-square (count/categorical)

p-values have not been adjusted for multiple comparisons

## After removing CF Insulin Use

Results did not change too much after removing participants with CF Insulin use. Some p-values that were close to 0.05 shifted from one side to the other

After removing CF Insulin Use:

- matsuda now significant on ANOVA with Hypo < 60
- OGTT Category pairwise comparison CF/Hypo60 vs CF/No-Hypo60 now significant
- HOMA-IR pairwise comparison CF/Hypo60 vs No-CF/No-Hypo60 now significant
- CGM Avg Glucose pairwise comparison CF/No-Hypo60 vs No-CF/Hypo60 no longer significant
- CGM Avg Glucose pairwise comparison CF/No-Hypo60 vs No-CF/No-Hypo60 no longer significant
- CGM MAG CF/No-Hypo60 vs No-CF/Hypo60 now significant
- Matsuda and Homa-IR pairwise comparisons CF/Hypo70 vs No-CF/No-Hypo70 now significant
- Delta Glucagon pairwise comparison CF/No-Hypo70 vs No-CF/Hypo70 no longer significant
- CGM Avg Glucose pairwise comparison CF/No-Hypo70 vs No-CF/Hypo70 no longer significant
- CGM Avg Glucose pairwise comparison CF/No-Hypo70 vs No-CF/No-Hypo70 no longer significant
- %time over 140 pairwise comparison CF/Hypo70 vs CF/No-Hypo no longer significant

# With CF Insulin

## Table 1 by CF/Hypo < 60

```{r, results='asis', warning = F, message = F}
control <- tableby.control(numeric.test = "kwt")

tab <- tableby(CF_hypo_60 ~ 
                 gender + 
                 age +
                 race_ethnicity+
                 tanner_stage+
                 weight+
                 height+
                 bmi+
                 bmi_z_score+
                 mutation_severity+
                 fev1_percent+
                 fvc+
                 poc_a1c+
                 ogtt_category+
                 cf_insulin_use +
                 g_tube_feeds +
                 matsuda_ins +
                 matsuda_cpep +
                 homa_ir +
                 igi +
                 cpep_index +
                 odi_ins +
                 odi_cpep +
                 cpep_ins0 +
                 cpep_ins_auc_30 +
                 cpep_ins_auc +
                 `0_glucose` +
                 gluc_peak +
                 gluc_nadir +
                 gluc_peak_time +
                 gluc_nadir_time +
                 glucose_auc +
                 delta_glucose +
                 `0_insulin` +
                 ins_peak +
                 ins_nadir +
                 ins_peak_time +
                 ins_nadir_time +
                 insulin_auc +
                 delta_insulin +
                 `0_cpeptide` +
                 cpep_peak +
                 cpep_nadir +
                 cpep_peak_time +
                 cpep_nadir_time +
                 cpeptide_auc +
                 delta_cpeptide +
                 `0_glucagon` +
                 gluca_peak +
                 gluca_nadir +
                 gluca_peak_time +
                 gluca_nadir_time +
                 glucagon_auc +
                 delta_glucagon +
                 `0_gip` +
                 gip_peak +
                 gip_nadir +
                 gip_peak_time +
                 gip_nadir_time +
                 gip_auc +
                 delta_gip +
                 `0_glp1` +
                 glp1_peak +
                 glp1_nadir +
                 glp1_peak_time +
                 glp1_nadir_time +
                 glp1_auc +
                 delta_glp1
               ,
               data = table_dat,
               control = control)
summary(tab)
```

```{r, warning = F, message = F}
pairwiseTableby <- function(tab){
  tbcall <- tab$Call
  data <- eval(tbcall$data)
  group_term <- tab$tables[[1]]$y$term
  group_levels <- head(names(tab$tables[[1]]$y$stats),-1)
  comps <- combn(group_levels, 2)

  for(i in seq_len(ncol(comps))){
    comp_name <- paste(comps[,i], collapse = " vs ")
    sub_df <- data %>% filter(get(group_term) %in% comps[,i])
    try({
      sub_tab <- tableby(eval(tbcall$formula),sub_df, numeric.test = "wt", wilcox.correct = TRUE, wilcox.exact = TRUE)
      p_list <- lapply(sub_tab$tables[[1]]$tables[[1]], function(x) x$test$p.value)
      if(exists("results_frame")){
        results_frame[comp_name] <- unlist(p_list)
      }
      else{
        results_frame <- data.frame(row.names = names(unlist(p_list)))
        results_frame[comp_name] <- unlist(p_list)
      }
    }, silent = TRUE)

  }
  results_frame
}
ptab <- pairwiseTableby(tab)

kbl(ptab, caption = "Pairwise p-values") %>% 
  column_spec(2,background = ifelse(ptab$`CF/Hypo vs CF/No-Hypo` < 0.05,"red","white")) %>%
  column_spec(3,background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/Hypo`< 0.05,"red","white"),"white)")) %>%
  column_spec(4,background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/No-Hypo` < 0.05,"red","white"),"white)"))  %>%
  column_spec(5,background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/Hypo` < 0.05,"red","white"),"white)"))  %>%
  column_spec(6,background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/No-Hypo` < 0.05,"red","white"),"white)")) 
```

## CGM with Hypo < 60

```{r, results='asis', warning = F, message = F}
cgm_table_dat <- table_dat %>% filter(!is.na(daytime_avg_sensor_glucose))

tab <- tableby(formula(
  paste("CF_hypo_60 ~", 
        paste(names(cgm_table_dat %>% 
                      select(num_days_good_data:interquartile_range)), 
              collapse = "+")))
  , 
  data = cgm_table_dat,
  control = control)
summary(tab)

ptab <- pairwiseTableby(tab)

kbl(ptab, caption = "Pairwise p-values") %>% 
  column_spec(2,background = ifelse(ptab$`CF/Hypo vs CF/No-Hypo` < 0.05,"red","white")) %>%
  column_spec(3,background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/Hypo`< 0.05,"red","white"),"white)")) %>%
  column_spec(4,background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/No-Hypo` < 0.05,"red","white"),"white)"))  %>%
  column_spec(5,background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/Hypo` < 0.05,"red","white"),"white)"))  %>%
  column_spec(6,background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/No-Hypo` < 0.05,"red","white"),"white)")) 
```

## OGTT with Hypo < 60

```{r, warning = F, message = F}
model <- lme(glucagon ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p1 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucagon (pg/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

model <- lme(glucose ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p2 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucose (mg/dL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

model <- lme(insulin ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p3 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Insulin (uIU/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

model <- lme(cpeptide ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p4 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "C-Peptide (ng/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

model <- lme(gip ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p5 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GIP (pmol/L)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

model <- lme(glp1 ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p6 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GLP-1 (pmol/L)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position='bottom')

png("ogtt_60.png", res = 300, width = 2000, height = 3000)
p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position='bottom')
dev.off()
```

```{r, warning = F, message = F}
model <- lme(glucagon ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p1 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucagon (pg/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(glucose ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p2 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucose (mg/dL)", linetype = "", shape = "") +
  theme_light()

model <- lme(insulin ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p3 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Insulin (uIU/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(cpeptide ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p4 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "C-Peptide (ng/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(gip ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p5 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GIP (pmol/L)", linetype = "", shape = "") +
  theme_light()

model <- lme(glp1 ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p6 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GLP-1 (pmol/L)", linetype = "", shape = "") +
  theme_light()

p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position='bottom')

png("ogtt_60_bw.png", res = 300, width = 2000, height = 3000)
p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position='bottom')
dev.off()
```

## Insulin Resistance with Hypo < 60

```{r, warning = F, message = F}
p1 <- ggplot(data = table_dat, aes(y = matsuda_ins, x = CF_hypo_60, color = CF_hypo_60)) +
  geom_boxplot() +
  labs(y = "Matsuda Index", x = "", color = "") +
  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo"),c("CF/Hypo","No-CF/No-Hypo")),
              color = "black",
              y_position = c(45,50),
              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  scale_color_brewer(palette="Dark2")

p2 <- ggplot(data = table_dat, aes(y = homa_ir, x = CF_hypo_60, color = CF_hypo_60)) +
  geom_boxplot() +
  labs(y = "HOMA IR", x = "", color = "") +
  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo")),
              color = "black",
              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  scale_color_brewer(palette="Dark2")

p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position='bottom')

png("resistance_60.png", height = 2000, width = 3000, res = 300)
p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position='bottom')
dev.off()
```

```{r, warning = F, message = F}
p1 <- ggplot(data = table_dat, aes(y = matsuda_ins, x = CF_hypo_60)) +
  geom_boxplot() +
  labs(y = "Matsuda Index", x = "", color = "") +
  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo"),c("CF/Hypo","No-CF/No-Hypo")),
              color = "black",
              y_position = c(45,50),
              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

p2 <- ggplot(data = table_dat, aes(y = homa_ir, x = CF_hypo_60)) +
  geom_boxplot() +
  labs(y = "HOMA IR", x = "", color = "") +
  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo")),
              color = "black",
              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position='bottom')

png("resistance_60_bw.png", height = 2000, width = 3000, res = 300)
p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position='bottom')
dev.off()
```

## Table 1 by CF/Hypo < 70

```{r, results='asis', warning = F, message = F}
tab <- tableby(CF_hypo_70 ~ 
                 gender + 
                 age +
                 race_ethnicity+
                 tanner_stage+
                 weight+
                 height+
                 bmi+
                 bmi_z_score+
                 mutation_severity+
                 fev1_percent+
                 fvc+
                 poc_a1c+
                 ogtt_category+
                 cf_insulin_use +
                 g_tube_feeds +
                 matsuda_ins +
                 matsuda_cpep +
                 homa_ir +
                 igi +
                 cpep_index +
                 odi_ins +
                 odi_cpep +
                 cpep_ins0 +
                 cpep_ins_auc_30 +
                 cpep_ins_auc +
                 `0_glucose` +
                 gluc_peak +
                 gluc_nadir +
                 gluc_peak_time +
                 gluc_nadir_time +
                 glucose_auc +
                 delta_glucose +
                 `0_insulin` +
                 ins_peak +
                 ins_nadir +
                 ins_peak_time +
                 ins_nadir_time +
                 insulin_auc +
                 delta_insulin +
                 `0_cpeptide` +
                 cpep_peak +
                 cpep_nadir +
                 cpep_peak_time +
                 cpep_nadir_time +
                 cpeptide_auc +
                 delta_cpeptide +
                 `0_glucagon` +
                 gluca_peak +
                 gluca_nadir +
                 gluca_peak_time +
                 gluca_nadir_time +
                 glucagon_auc +
                 delta_glucagon +
                 `0_gip` +
                 gip_peak +
                 gip_nadir +
                 gip_peak_time +
                 gip_nadir_time +
                 gip_auc +
                 delta_gip +
                 `0_glp1` +
                 glp1_peak +
                 glp1_nadir +
                 glp1_peak_time +
                 glp1_nadir_time +
                 glp1_auc +
                 delta_glp1
               ,
               data = table_dat,
               control = control)
summary(tab)

ptab <- pairwiseTableby(tab)

kbl(ptab, caption = "Pairwise p-values") %>% 
  column_spec(2,background = ifelse(ptab$`CF/Hypo vs CF/No-Hypo` < 0.05,"red","white")) %>%
  column_spec(3,background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/Hypo`< 0.05,"red","white"),"white)")) %>%
  column_spec(4,background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/No-Hypo` < 0.05,"red","white"),"white)"))  %>%
  column_spec(5,background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/Hypo` < 0.05,"red","white"),"white)"))  %>%
  column_spec(6,background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/No-Hypo` < 0.05,"red","white"),"white)")) 
```

## CGM with Hypo < 70

```{r, results='asis', warning = F, message = F}
tab <- tableby(formula(
  paste("CF_hypo_70 ~", 
        paste(names(cgm_table_dat %>% 
                      select(num_days_good_data:interquartile_range)), 
              collapse = "+")))
  , 
  data = cgm_table_dat,
  control = control)
summary(tab)

ptab <- pairwiseTableby(tab)

kbl(ptab, caption = "Pairwise p-values") %>% 
  column_spec(2,background = ifelse(ptab$`CF/Hypo vs CF/No-Hypo` < 0.05,"red","white")) %>%
  column_spec(3,background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/Hypo`< 0.05,"red","white"),"white)")) %>%
  column_spec(4,background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/No-Hypo` < 0.05,"red","white"),"white)"))  %>%
  column_spec(5,background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/Hypo` < 0.05,"red","white"),"white)"))  %>%
  column_spec(6,background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/No-Hypo` < 0.05,"red","white"),"white)")) 
```

## OGTT with Hypo < 70

```{r, warning = F, message = F}
model <- lme(glucagon ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p1 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucagon (pg/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

model <- lme(glucose ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p2 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucose (mg/dL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

model <- lme(insulin ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p3 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Insulin (uIU/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

model <- lme(cpeptide ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p4 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "C-Peptide (ng/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

model <- lme(gip ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p5 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GIP (pmol/L)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

model <- lme(glp1 ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p6 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GLP-1 (pmol/L)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position='bottom')

png("ogtt_70.png", res = 300, width = 2000, height = 3000)
p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position='bottom')
dev.off()
```

```{r, warning = F, message = F}
model <- lme(glucagon ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p1 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucagon (pg/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(glucose ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p2 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucose (mg/dL)", linetype = "", shape = "") +
  theme_light()

model <- lme(insulin ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p3 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Insulin (uIU/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(cpeptide ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p4 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "C-Peptide (ng/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(gip ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p5 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GIP (pmol/L)", linetype = "", shape = "") +
  theme_light()

model <- lme(glp1 ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p6 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GLP-1 (pmol/L)", linetype = "", shape = "") +
  theme_light()

p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position='bottom')

png("ogtt_70_bw.png", res = 300, width = 2000, height = 3000)
p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position='bottom')
dev.off()
```

## Insulin Resistance with Hypo < 70

```{r, warning = F, message = F}
p1 <- ggplot(data = table_dat, aes(y = matsuda_ins, x = CF_hypo_70, color = CF_hypo_70)) +
  geom_boxplot() +
  labs(y = "Matsuda Index", x = "", color = "") +
  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo")),
              color = "black",
              test = "wilcox.test",
              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  scale_color_brewer(palette="Dark2")

p2 <- ggplot(data = table_dat, aes(y = homa_ir, x = CF_hypo_70, color = CF_hypo_70)) +
  geom_boxplot() +
  labs(y = "HOMA IR", x = "", color = "") +
  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo")),
              color = "black",
              test = "wilcox.test",
              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  scale_color_brewer(palette="Dark2")

p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position='bottom')

png("resistance_70.png", height = 2000, width = 3000, res = 300)
p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position='bottom')
dev.off()
```

```{r, warning = F, message = F}
p1 <- ggplot(data = table_dat, aes(y = matsuda_ins, x = CF_hypo_70)) +
  geom_boxplot() +
  labs(y = "Matsuda Index", x = "", color = "") +
  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo")),
              color = "black",
              test = "wilcox.test",
              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

p2 <- ggplot(data = table_dat, aes(y = homa_ir, x = CF_hypo_70)) +
  geom_boxplot() +
  labs(y = "HOMA IR", x = "", color = "") +
  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo")),
              color = "black",
              test = "wilcox.test",
              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position='bottom')

png("resistance_70_bw.png", height = 2000, width = 3000, res = 300)
p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position='bottom')
dev.off()
```

## CGM Figures

```{r fig.height=5, results="asis"}
outcomes <- c(
  "daytime_avg_sensor_glucose",
  "percent_time_under_60",
  "percent_time_under_70",
  "percent_time_over_140",
  "percent_time_over_180",
  "daytime_sd",
  "cov",
  "mag"
)
outcomes_labs <- c(
  "Avg Glucose (Daytime)",
  "% Time Under 60",
  "% Time Under 70",
  "% Time Over 140",
  "% Time Over 180",
  "SD",
  "COV",
  "MAG"
)

for (i in seq_len(length(outcomes))){

  p70 <- ggplot(data = cgm_table_dat, aes_string(y = outcomes[i], x = "CF_hypo_70")) +
    labs(y = outcomes_labs[i], x= "") +
    geom_jitter(alpha = 0.5, width = 0.1) +
    geom_boxplot(alpha = 0.5, outlier.alpha = 0) +
    theme_light() +
    ggtitle("Hypoglycemia by OGTT (< 70)") +
    theme(axis.text.x = element_text(angle = 45, hjust=1))

  p60 <- ggplot(data = cgm_table_dat, aes_string(y = outcomes[i], x = "CF_hypo_60")) +
    labs(y = outcomes_labs[i], x= "") +
    geom_jitter(alpha = 0.5, width = 0.1) +
    geom_boxplot(alpha = 0.5, outlier.alpha = 0) +
    theme_light() +
    ggtitle("Hypoglycemia by OGTT (< 60)") +
    theme(axis.text.x = element_text(angle = 45, hjust=1))

  cat("\n")
  cat(paste("##",outcomes_labs[i]))
  cat("\n")
  print(p70 + p60)
  cat("\n")
}

```

## CGM Hypoglycemia

```{r,results="asis"}
tab <- tableby(group_category ~ percent_time_under_70 + percent_time_under_60, data = cgm_table_dat)
summary(tab)

under_70_cutoff <- 2*2.06
under_60_cutoff <- 2*0.687

cgm_table_dat$CGM_hypo_60 <- ifelse(cgm_table_dat$percent_time_under_60 > under_60_cutoff, "Yes", "No")
cgm_table_dat$CGM_hypo_70 <- ifelse(cgm_table_dat$percent_time_under_70 > under_70_cutoff, "Yes", "No")
```

Hypoglycemia by CGM is defined as a % time under 70 or 60 more than 2 SDs away from average among healthy controls

<60 cutoff

```{r}
under_60_cutoff
```

<70 cutoff

```{r}
under_70_cutoff
```

## Hypoglycemia by CGM by group

```{r, results="asis"}
cgm_tableby <- cgm_table_dat
cgm_tableby$CGM_hypo_70[cgm_tableby$CGM_hypo_70 == "Yes"] <- "Hypo by CGM"
cgm_tableby$CGM_hypo_70[cgm_tableby$CGM_hypo_70 == "No"] <- "Not Hypo by CGM"
cgm_tableby$hypo_70[cgm_tableby$hypo_70 == "Yes"] <- "Hypo by OGTT"
cgm_tableby$hypo_70[cgm_tableby$hypo_70 == "No"] <- "Not Hypo by OGTT"
cgm_tableby$CGM_hypo_60[cgm_tableby$CGM_hypo_60 == "Yes"] <- "Hypo by CGM"
cgm_tableby$CGM_hypo_60[cgm_tableby$CGM_hypo_60 == "No"] <- "Not Hypo by CGM"
cgm_tableby$hypo_60[cgm_tableby$hypo_60 == "Yes"] <- "Hypo by OGTT"
cgm_tableby$hypo_60[cgm_tableby$hypo_60 == "No"] <- "Not Hypo by OGTT"

tab <- tableby(group_category ~ CGM_hypo_60 + CGM_hypo_70, data = cgm_tableby)
summary(tab)
```

## Hypoglycemia Concordance

### 70/70

Hypoglycemia by OGTT with cutoff <70 and by CGM with cutoff < 70

Hypoglycemia by OGTT and by CGM have poor agreement with a Cohen's Kappa of 
-0.15 among participants with CF and 0.07 among those without CF

```{r, results = "asis"}
tab <- tableby(CGM_hypo_70 ~ hypo_70, 
               strata = CF,
               data = cgm_tableby, 
               test = F, 
               total = F)
summary(tab)
```
#### Concordance with CF
```{r}
mat <- matrix( data = c(4,12,16,23), nrow = 2, ncol = 2)
Kappa(mat)
```

#### Concordance without CF
```{r}
mat <- matrix( data = c(1,2,3,9), nrow = 2, ncol = 2)
Kappa(mat)
```

### 60/60

Hypoglycemia by OGTT with cutoff <60 and by CGM with cutoff < 60

Hypoglycemia by OGTT and by CGM have poor agreement with a Cohen's Kappa of
-0.17 among participants with CF. The concordance among those without CF was 
0.42, though due to a small sample size, this was not statistically significant.

```{r, results = "asis"}
tab <- tableby(CGM_hypo_60 ~ hypo_60, 
               strata = CF,
               data = cgm_tableby, 
               test = F, 
               total = F)
summary(tab)
```
#### Concordance with CF
```{r}
mat <- matrix( data = c(2,12,13,28), nrow = 2, ncol = 2)
Kappa(mat)
```

#### Concordance without CF
```{r}
mat <- matrix( data = c(1,1,1,12), nrow = 2, ncol = 2)
Kappa(mat)
```

### 70/60

Hypoglycemia by OGTT with cutoff <70 and by CGM with cutoff < 60

Hypoglycemia by OGTT and by CGM have poor agreement with a Cohen's Kappa of -0.17

```{r, results = "asis"}
tab <- tableby(CGM_hypo_60 ~ hypo_70, data = cgm_table_dat, test = F, total = F)
summary(tab)
```
```{r}
mat <- matrix( data = c(33,21,13,3), nrow = 2, ncol = 2)
Kappa(mat)
```

### 60/70

Hypoglycemia by OGTT with cutoff <60 and by CGM with cutoff < 70

Hypoglycemia by OGTT and by CGM have poor agreement with a Cohen's Kappa of 0.03

```{r, results = "asis"}
tab <- tableby(CGM_hypo_70 ~ hypo_60, data = cgm_table_dat, test = F, total = F)
summary(tab)
```
```{r}
mat <- matrix( data = c(39,12,14,5), nrow = 2, ncol = 2)
Kappa(mat)
```

```{r}
table_dat <- table_dat %>% filter(cf_insulin_use == "No" | is.na(cf_insulin_use))
gluc <- gluc %>% filter(subject_id %in% table_dat$subject_id)

write.csv(table_dat, "analysis_data_no_cfinsulinuse.csv")
```

# Removed CF Insulin Use

## Table 1 by CF/Hypo < 60

```{r, results='asis', warning = F, message = F}
control <- tableby.control(numeric.test = "kwt")

tab <- tableby(CF_hypo_60 ~ 
                 gender + 
                 age +
                 race_ethnicity+
                 tanner_stage+
                 weight+
                 height+
                 bmi+
                 bmi_z_score+
                 mutation_severity+
                 fev1_percent+
                 fvc+
                 poc_a1c+
                 ogtt_category+
                 cf_insulin_use +
                 g_tube_feeds +
                 matsuda_ins +
                 matsuda_cpep +
                 homa_ir +
                 igi +
                 cpep_index +
                 odi_ins +
                 odi_cpep +
                 cpep_ins0 +
                 cpep_ins_auc_30 +
                 cpep_ins_auc +
                 `0_glucose` +
                 gluc_peak +
                 gluc_nadir +
                 gluc_peak_time +
                 gluc_nadir_time +
                 glucose_auc +
                 delta_glucose +
                 `0_insulin` +
                 ins_peak +
                 ins_nadir +
                 ins_peak_time +
                 ins_nadir_time +
                 insulin_auc +
                 delta_insulin +
                 `0_cpeptide` +
                 cpep_peak +
                 cpep_nadir +
                 cpep_peak_time +
                 cpep_nadir_time +
                 cpeptide_auc +
                 delta_cpeptide +
                 `0_glucagon` +
                 gluca_peak +
                 gluca_nadir +
                 gluca_peak_time +
                 gluca_nadir_time +
                 glucagon_auc +
                 delta_glucagon +
                 `0_gip` +
                 gip_peak +
                 gip_nadir +
                 gip_peak_time +
                 gip_nadir_time +
                 gip_auc +
                 delta_gip +
                 `0_glp1` +
                 glp1_peak +
                 glp1_nadir +
                 glp1_peak_time +
                 glp1_nadir_time +
                 glp1_auc +
                 delta_glp1
               ,
               data = table_dat,
               control = control)
summary(tab)
ptab <- pairwiseTableby(tab)

kbl(ptab, caption = "Pairwise p-values") %>% 
  column_spec(2,background = ifelse(ptab$`CF/Hypo vs CF/No-Hypo` < 0.05,"red","white")) %>%
  column_spec(3,background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/Hypo`< 0.05,"red","white"),"white)")) %>%
  column_spec(4,background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/No-Hypo` < 0.05,"red","white"),"white)"))  %>%
  column_spec(5,background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/Hypo` < 0.05,"red","white"),"white)"))  %>%
  column_spec(6,background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/No-Hypo` < 0.05,"red","white"),"white)")) 
```

## CGM with Hypo < 60

```{r, results='asis', warning = F, message = F}
cgm_table_dat <- table_dat %>% filter(!is.na(daytime_avg_sensor_glucose))

tab <- tableby(formula(
  paste("CF_hypo_60 ~", 
        paste(names(cgm_table_dat %>% 
                      select(num_days_good_data:interquartile_range)), 
              collapse = "+")))
  , 
  data = cgm_table_dat,
  control = control)
summary(tab)

ptab <- pairwiseTableby(tab)

kbl(ptab, caption = "Pairwise p-values") %>% 
  column_spec(2,background = ifelse(ptab$`CF/Hypo vs CF/No-Hypo` < 0.05,"red","white")) %>%
  column_spec(3,background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/Hypo`< 0.05,"red","white"),"white)")) %>%
  column_spec(4,background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/No-Hypo` < 0.05,"red","white"),"white)"))  %>%
  column_spec(5,background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/Hypo` < 0.05,"red","white"),"white)"))  %>%
  column_spec(6,background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/No-Hypo` < 0.05,"red","white"),"white)")) 
```

## OGTT with Hypo < 60

```{r, warning = F, message = F}
model <- lme(glucagon ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p1 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucagon (pg/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

model <- lme(glucose ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p2 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucose (mg/dL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

model <- lme(insulin ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p3 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Insulin (uIU/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

model <- lme(cpeptide ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p4 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "C-Peptide (ng/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

model <- lme(gip ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p5 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GIP (pmol/L)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

model <- lme(glp1 ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p6 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GLP-1 (pmol/L)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position='bottom')

png("ogtt_60_noinsulin.png", res = 300, width = 2000, height = 3000)
p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position='bottom')
dev.off()
```

```{r, warning = F, message = F}
model <- lme(glucagon ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p1 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucagon (pg/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(glucose ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p2 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucose (mg/dL)", linetype = "", shape = "") +
  theme_light()

model <- lme(insulin ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p3 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Insulin (uIU/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(cpeptide ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p4 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "C-Peptide (ng/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(gip ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p5 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GIP (pmol/L)", linetype = "", shape = "") +
  theme_light()

model <- lme(glp1 ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))

p6 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GLP-1 (pmol/L)", linetype = "", shape = "") +
  theme_light()

p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position='bottom')

png("ogtt_60_noinsulin_bw.png", res = 300, width = 2000, height = 3000)
p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position='bottom')
dev.off()
```

## Insulin Resistance with Hypo < 60

```{r, warning = F, message = F}
p1 <- ggplot(data = table_dat, aes(y = matsuda_ins, x = CF_hypo_60, color = CF_hypo_60)) +
  geom_boxplot() +
  labs(y = "Matsuda Index", x = "", color = "") +
  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo"),c("CF/Hypo","No-CF/No-Hypo")),
              color = "black",
              y_position = c(45,50),
              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  scale_color_brewer(palette="Dark2")

p2 <- ggplot(data = table_dat, aes(y = homa_ir, x = CF_hypo_60, color = CF_hypo_60)) +
  geom_boxplot() +
  labs(y = "HOMA IR", x = "", color = "") +
  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo"),c("CF/Hypo","No-CF/No-Hypo")),
              color = "black",
              y_position = c(4.5,5),
              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  scale_color_brewer(palette="Dark2")

p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position='bottom')

png("resistance_60_noinsulin.png", height = 2000, width = 3000, res = 300)
p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position='bottom')
dev.off()
```

```{r, warning = F, message = F}
p1 <- ggplot(data = table_dat, aes(y = matsuda_ins, x = CF_hypo_60)) +
  geom_boxplot() +
  labs(y = "Matsuda Index", x = "", color = "") +
  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo"),c("CF/Hypo","No-CF/No-Hypo")),
              color = "black",
              y_position = c(45,50),
              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

p2 <- ggplot(data = table_dat, aes(y = homa_ir, x = CF_hypo_60)) +
  geom_boxplot() +
  labs(y = "HOMA IR", x = "", color = "") +
  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo"),c("CF/Hypo","No-CF/No-Hypo")),
              color = "black",
              y_position = c(4.5,5),
              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position='bottom')

png("resistance_60_noinsulin_bw.png", height = 2000, width = 3000, res = 300)
p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position='bottom')
dev.off()
```

## Table 1 by CF/Hypo < 70

```{r, results='asis', warning = F, message = F}
tab <- tableby(CF_hypo_70 ~ 
                 gender + 
                 age +
                 race_ethnicity+
                 tanner_stage+
                 weight+
                 height+
                 bmi+
                 bmi_z_score+
                 mutation_severity+
                 fev1_percent+
                 fvc+
                 poc_a1c+
                 ogtt_category+
                 cf_insulin_use +
                 g_tube_feeds +
                 matsuda_ins +
                 matsuda_cpep +
                 homa_ir +
                 igi +
                 cpep_index +
                 odi_ins +
                 odi_cpep +
                 cpep_ins0 +
                 cpep_ins_auc_30 +
                 cpep_ins_auc +
                 `0_glucose` +
                 gluc_peak +
                 gluc_nadir +
                 gluc_peak_time +
                 gluc_nadir_time +
                 glucose_auc +
                 delta_glucose +
                 `0_insulin` +
                 ins_peak +
                 ins_nadir +
                 ins_peak_time +
                 ins_nadir_time +
                 insulin_auc +
                 delta_insulin +
                 `0_cpeptide` +
                 cpep_peak +
                 cpep_nadir +
                 cpep_peak_time +
                 cpep_nadir_time +
                 cpeptide_auc +
                 delta_cpeptide +
                 `0_glucagon` +
                 gluca_peak +
                 gluca_nadir +
                 gluca_peak_time +
                 gluca_nadir_time +
                 glucagon_auc +
                 delta_glucagon +
                 `0_gip` +
                 gip_peak +
                 gip_nadir +
                 gip_peak_time +
                 gip_nadir_time +
                 gip_auc +
                 delta_gip +
                 `0_glp1` +
                 glp1_peak +
                 glp1_nadir +
                 glp1_peak_time +
                 glp1_nadir_time +
                 glp1_auc +
                 delta_glp1
               ,
               data = table_dat,
               control = control)
summary(tab)

ptab <- pairwiseTableby(tab)

kbl(ptab, caption = "Pairwise p-values") %>% 
  column_spec(2,background = ifelse(ptab$`CF/Hypo vs CF/No-Hypo` < 0.05,"red","white")) %>%
  column_spec(3,background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/Hypo`< 0.05,"red","white"),"white)")) %>%
  column_spec(4,background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/No-Hypo` < 0.05,"red","white"),"white)"))  %>%
  column_spec(5,background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/Hypo` < 0.05,"red","white"),"white)"))  %>%
  column_spec(6,background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/No-Hypo` < 0.05,"red","white"),"white)")) 
```

## CGM with Hypo < 70

```{r, results='asis', warning = F, message = F}
tab <- tableby(formula(
  paste("CF_hypo_70 ~", 
        paste(names(cgm_table_dat %>% 
                      select(num_days_good_data:interquartile_range)), 
              collapse = "+")))
  , 
  data = cgm_table_dat,
  control = control)
summary(tab)

ptab <- pairwiseTableby(tab)

kbl(ptab, caption = "Pairwise p-values") %>% 
  column_spec(2,background = ifelse(ptab$`CF/Hypo vs CF/No-Hypo` < 0.05,"red","white")) %>%
  column_spec(3,background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/Hypo`< 0.05,"red","white"),"white)")) %>%
  column_spec(4,background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/No-Hypo` < 0.05,"red","white"),"white)"))  %>%
  column_spec(5,background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/Hypo` < 0.05,"red","white"),"white)"))  %>%
  column_spec(6,background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/No-Hypo` < 0.05,"red","white"),"white)")) 
```

## OGTT with Hypo < 70

```{r, warning = F, message = F}
model <- lme(glucagon ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p1 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucagon (pg/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

model <- lme(glucose ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p2 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucose (mg/dL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

model <- lme(insulin ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p3 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Insulin (uIU/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

model <- lme(cpeptide ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p4 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "C-Peptide (ng/mL)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

model <- lme(gip ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p5 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GIP (pmol/L)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

model <- lme(glp1 ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p6 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, color = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GLP-1 (pmol/L)", color = "", shape = "") +
  theme_light() +
  scale_color_brewer(palette="Dark2")

p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position='bottom')

png("ogtt_70_noinsulin.png", res = 300, width = 2000, height = 3000)
p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position='bottom')
dev.off()
```

```{r, warning = F, message = F}
model <- lme(glucagon ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p1 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucagon (pg/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(glucose ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p2 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Glucose (mg/dL)", linetype = "", shape = "") +
  theme_light()

model <- lme(insulin ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p3 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "Insulin (uIU/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(cpeptide ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p4 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "C-Peptide (ng/mL)", linetype = "", shape = "") +
  theme_light()

model <- lme(gip ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p5 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GIP (pmol/L)", linetype = "", shape = "") +
  theme_light()

model <- lme(glp1 ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))

p6 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_70, linetype = CF_hypo_70, shape = CF_hypo_70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.4, position = "dodge") +
  geom_line() +
  labs(y = "GLP-1 (pmol/L)", linetype = "", shape = "") +
  theme_light()

p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position='bottom')

png("ogtt_70_noinsulin_bw.png", res = 300, width = 2000, height = 3000)
p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
  theme(legend.position='bottom')
dev.off()
```

## Insulin Resistance with Hypo < 70

```{r, warning = F, message = F}
p1 <- ggplot(data = table_dat, aes(y = matsuda_ins, x = CF_hypo_70, color = CF_hypo_70)) +
  geom_boxplot() +
  labs(y = "Matsuda Index", x = "", color = "") +
  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo"),c("CF/Hypo","No-CF/No-Hypo")),
              color = "black",
              y_position = c(45,50),
              test = "wilcox.test",
              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  scale_color_brewer(palette="Dark2")

p2 <- ggplot(data = table_dat, aes(y = homa_ir, x = CF_hypo_70, color = CF_hypo_70)) +
  geom_boxplot() +
  labs(y = "HOMA IR", x = "", color = "") +
  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo"),c("CF/Hypo","No-CF/No-Hypo")),
              color = "black",
              y_position = c(4.5,5),
              test = "wilcox.test",
              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  scale_color_brewer(palette="Dark2")

p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position='bottom')

png("resistance_70_noinsulin.png", height = 2000, width = 3000, res = 300)
p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position='bottom')
dev.off()
```

```{r, warning = F, message = F}
p1 <- ggplot(data = table_dat, aes(y = matsuda_ins, x = CF_hypo_70)) +
  geom_boxplot() +
  labs(y = "Matsuda Index", x = "", color = "") +
  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo"),c("CF/Hypo","No-CF/No-Hypo")),
              color = "black",
              y_position = c(45,50),
              test = "wilcox.test",
              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

p2 <- ggplot(data = table_dat, aes(y = homa_ir, x = CF_hypo_70)) +
  geom_boxplot() +
  labs(y = "HOMA IR", x = "", color = "") +
  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo"),c("CF/Hypo","No-CF/No-Hypo")),
              color = "black",
              y_position = c(4.5,5),
              test = "wilcox.test",
              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
  theme_light() +
  guides(color = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position='bottom')

png("resistance_70_noinsulin_bw.png", height = 2000, width = 3000, res = 300)
p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
  theme(legend.position='bottom')
dev.off()
```

## CGM Figures

```{r fig.height=5, results="asis"}
outcomes <- c(
  "daytime_avg_sensor_glucose",
  "percent_time_under_60",
  "percent_time_under_70",
  "percent_time_over_140",
  "percent_time_over_180",
  "daytime_sd",
  "cov",
  "mag"
)
outcomes_labs <- c(
  "Avg Glucose (Daytime)",
  "% Time Under 60",
  "% Time Under 70",
  "% Time Over 140",
  "% Time Over 180",
  "SD",
  "COV",
  "MAG"
)

for (i in seq_len(length(outcomes))){

  p70 <- ggplot(data = cgm_table_dat, aes_string(y = outcomes[i], x = "CF_hypo_70")) +
    labs(y = outcomes_labs[i], x= "") +
    geom_jitter(alpha = 0.5, width = 0.1) +
    geom_boxplot(alpha = 0.5, outlier.alpha = 0) +
    theme_light() +
    ggtitle("Hypoglycemia by OGTT (< 70)") +
    theme(axis.text.x = element_text(angle = 45, hjust=1))

  p60 <- ggplot(data = cgm_table_dat, aes_string(y = outcomes[i], x = "CF_hypo_60")) +
    labs(y = outcomes_labs[i], x= "") +
    geom_jitter(alpha = 0.5, width = 0.1) +
    geom_boxplot(alpha = 0.5, outlier.alpha = 0) +
    theme_light() +
    ggtitle("Hypoglycemia by OGTT (< 60)") +
    theme(axis.text.x = element_text(angle = 45, hjust=1))

  cat("\n")
  cat(paste("##",outcomes_labs[i]))
  cat("\n")
  print(p70 + p60)
  cat("\n")
}

```

## CGM Hypoglycemia

```{r,results="asis"}
tab <- tableby(group_category ~ percent_time_under_70 + percent_time_under_60, data = cgm_table_dat)
summary(tab)

under_70_cutoff <- 2*2.06
under_60_cutoff <- 2*0.687

cgm_table_dat$CGM_hypo_60 <- ifelse(cgm_table_dat$percent_time_under_60 > under_60_cutoff, "Yes", "No")
cgm_table_dat$CGM_hypo_70 <- ifelse(cgm_table_dat$percent_time_under_70 > under_70_cutoff, "Yes", "No")
```

Hypoglycemia by CGM is defined as a % time under 70 or 60 more than 2 SDs away from average among healthy controls

<60 cutoff

```{r}
under_60_cutoff
```

<70 cutoff

```{r}
under_70_cutoff
```

## Hypoglycemia by CGM by group

```{r, results="asis"}
cgm_tableby <- cgm_table_dat
cgm_tableby$CGM_hypo_70[cgm_tableby$CGM_hypo_70 == "Yes"] <- "Hypo by CGM"
cgm_tableby$CGM_hypo_70[cgm_tableby$CGM_hypo_70 == "No"] <- "Not Hypo by CGM"
cgm_tableby$hypo_70[cgm_tableby$hypo_70 == "Yes"] <- "Hypo by OGTT"
cgm_tableby$hypo_70[cgm_tableby$hypo_70 == "No"] <- "Not Hypo by OGTT"
cgm_tableby$CGM_hypo_60[cgm_tableby$CGM_hypo_60 == "Yes"] <- "Hypo by CGM"
cgm_tableby$CGM_hypo_60[cgm_tableby$CGM_hypo_60 == "No"] <- "Not Hypo by CGM"
cgm_tableby$hypo_60[cgm_tableby$hypo_60 == "Yes"] <- "Hypo by OGTT"
cgm_tableby$hypo_60[cgm_tableby$hypo_60 == "No"] <- "Not Hypo by OGTT"

tab <- tableby(group_category ~ CGM_hypo_60 + CGM_hypo_70, data = cgm_tableby)
summary(tab)
```

## Hypoglycemia Concordance

### 70/70

Hypoglycemia by OGTT with cutoff <70 and by CGM with cutoff < 70

Hypoglycemia by OGTT and by CGM have poor agreement with a Cohen's Kappa of 
-0.22 among participants with CF and 0.07 among those without CF

```{r, results = "asis"}
tab <- tableby(CGM_hypo_70 ~ hypo_70, 
               strata = CF,
               data = cgm_tableby, 
               test = F, 
               total = F)
summary(tab)
```

#### Concordance with CF
```{r}
mat <- matrix( data = c(3,10,15,16), nrow = 2, ncol = 2)
Kappa(mat)
```

#### Concordance without CF
```{r}
mat <- matrix( data = c(1,2,3,9), nrow = 2, ncol = 2)
Kappa(mat)
```

### 60/60

Hypoglycemia by OGTT with cutoff <60 and by CGM with cutoff < 60

Hypoglycemia by OGTT and by CGM have poor agreement with a Cohen's Kappa of
-0.17 among participants with CF. The concordance among those without CF was 
0.42, though due to a small sample size, this was not statistically significant.

```{r, results = "asis"}
tab <- tableby(CGM_hypo_60 ~ hypo_60, 
               strata = CF,
               data = cgm_tableby, 
               test = F, 
               total = F)
summary(tab)
```
#### Concordance with CF
```{r}
mat <- matrix( data = c(2,9,12,21), nrow = 2, ncol = 2)
Kappa(mat)
```

#### Concordance without CF
```{r}
mat <- matrix( data = c(1,1,1,12), nrow = 2, ncol = 2)
Kappa(mat)
```

### 2023-03-31 Christine Chan/Hypoglycemia/tim_analysis.Rmd
Update tim_analysis.Rmd

@@
 ---
-title: "Hypoglycemia Analysis"
+title: "Hypoglycemia Analysis - CFRD REMOVED"
 output:
@@
 ```{r}
-getAUCs <- function(data,subject,time,vars){
+getAUCs <- function(data, subject, time, vars) {
   all_return <- NULL
 
-  for (var in vars){
+  for (var in vars) {
     subs <- unique(data[[subject]])
     return <- NULL
-    for (sub in subs){
+    for (sub in subs) {
       try({
-        df <- data[data[[subject]] == sub,]
-        ogtt_0 <- df[df[[time]] == 0,var]
-        ogtt_30 <- df[df[[time]] == 30,var]
-        ogtt_60 <- df[df[[time]] == 60,var]
-        ogtt_90 <- df[df[[time]] == 90,var]
-        ogtt_120 <- df[df[[time]] == 120,var]
-        ogtt_auc <- (ogtt_30[[var]] + ogtt_60[[var]] + ogtt_90[[var]])/2 + (ogtt_0[[var]] + ogtt_120[[var]])/4
-        if(length(ogtt_auc) > 0){
-          row <- c(sub,ogtt_auc)
+        df <- data[data[[subject]] == sub, ]
+        ogtt_0 <- df[df[[time]] == 0, var]
+        ogtt_30 <- df[df[[time]] == 30, var]
+        ogtt_60 <- df[df[[time]] == 60, var]
+        ogtt_90 <- df[df[[time]] == 90, var]
+        ogtt_120 <- df[df[[time]] == 120, var]
+        ogtt_auc <- (ogtt_30[[var]] + ogtt_60[[var]] + ogtt_90[[var]]) / 
+          2 + (ogtt_0[[var]] + ogtt_120[[var]]) / 4
+        if (length(ogtt_auc) > 0) {
+          row <- c(sub, ogtt_auc)
           return <- rbind(return, row)
         }
-
       })
-
     }
     return <- as.data.frame(return)
-    names(return) <- c(subject,paste0(var,"_auc"))
-    if (is.null(all_return)){
+    names(return) <- c(subject, paste0(var, "_auc"))
+    if (is.null(all_return)) {
       all_return <- return
-    }
-    else{
+    } else {
       all_return <- merge(all_return, return, all = T)
     }
-
   }
   all_return
-
 }
-
 ```
@@
 ```{r, warning = F, message = F}
-dat <- read.csv("./data/CGMInHealthyControls_DATA_2021-05-25_0918.csv", na = "-999")
+dat <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/Hypoglycemia/Background/Previous Analyses/FsOGTT/data/CGMInHealthyControls_DATA_2021-05-25_0918.csv", na = "-999")
 
@@
 dat$group_category[dat$group_category == "4"] <- "CFRD"
-dat$group_category <- factor(dat$group_category, levels = c("Healthy","CF-NGT","CF-AG","CFRD"))
+dat$group_category <- factor(dat$group_category, levels = c("Healthy", "CF-NGT", "CF-AG", "CFRD"))
 
@@
 
-dat <- dat[dat$date_of_visit != "",]
+dat <- dat[dat$date_of_visit != "", ]
 
@@
 
-gluc <- dat %>% select(subject_id, 
+gluc <- dat %>% select(
+  subject_id,
   date_of_visit,
@@
   -contains("glp"),
-                       -contains("gip"))
+  -contains("gip")
+)
 gluc <- merge(gluc, dat %>% select(subject_id, date_of_visit, age, starts_with("ucd_")))
 gluc <- unique(gluc)
-gluc <- gluc[!is.na(gluc$min_30_glucose),]
-gluc[] <- lapply(gluc, gsub, pattern="<", replacement = "")
+gluc <- gluc[!is.na(gluc$min_30_glucose), ]
+gluc[] <- lapply(gluc, gsub, pattern = "<", replacement = "")
 
 gluc_names <- names(gluc)
-gluc_names <- gsub("min_","",gluc_names)
-gluc_names <- gsub("minute_","",gluc_names)
-gluc_names <- gsub("minus_","-",gluc_names)
-gluc_names <- gsub("ucd_","",gluc_names)
-gluc_names <- gsub("_total","",gluc_names)
-gluc_names <- gsub("glp_1","_glp1",gluc_names)
-gluc_names <- gsub("c_peptide","cpeptide",gluc_names)
+gluc_names <- gsub("min_", "", gluc_names)
+gluc_names <- gsub("minute_", "", gluc_names)
+gluc_names <- gsub("minus_", "-", gluc_names)
+gluc_names <- gsub("ucd_", "", gluc_names)
+gluc_names <- gsub("_total", "", gluc_names)
+gluc_names <- gsub("glp_1", "_glp1", gluc_names)
+gluc_names <- gsub("c_peptide", "cpeptide", gluc_names)
 names(gluc) <- gluc_names
@@
 gluc <- unique(gluc)
-new_gluc <- gluc[0,]
-
-for (sub in unique(gluc$subject_id)){
-  df <- gluc[gluc$subject_id == sub,]
-  if (nrow(df) > 1){
-    df <- df %>% arrange(`0_gip`,desc(date_of_visit))
-    df <- df[1,]
+new_gluc <- gluc[0, ]
+
+for (sub in unique(gluc$subject_id)) {
+  df <- gluc[gluc$subject_id == sub, ]
+  if (nrow(df) > 1) {
+    df <- df %>% arrange(`0_gip`, desc(date_of_visit))
+    df <- df[1, ]
   }
-  new_gluc <- rbind(new_gluc,df)
+  new_gluc <- rbind(new_gluc, df)
 }
@@
 gluc_wide <- new_gluc %>% select(-date_of_visit)
-gluc <- gluc_wide %>% pivot_longer(cols = -c(subject_id), 
-                                   names_to = c("time", ".value"), names_sep = "_")
-
-#gluc <- gluc[!is.na(gluc$glucose) | !is.na(gluc$glucagon) | !is.na(gluc$insulin)| !is.na(gluc$cpeptide),]
-#gluc <- gluc[!is.na(gluc$glucagon),]
-gluc <- gluc[gluc$time != "-10",]
+gluc <- gluc_wide %>% pivot_longer(
+  cols = -c(subject_id),
+  names_to = c("time", ".value"), names_sep = "_"
+)
+
+# gluc <- gluc[!is.na(gluc$glucose) | !is.na(gluc$glucagon) | !is.na(gluc$insulin)| !is.na(gluc$cpeptide),]
+# gluc <- gluc[!is.na(gluc$glucagon),]
+gluc <- gluc[gluc$time != "-10", ]
 gluc$time <- as.factor(as.numeric(gluc$time))
@@
 
-aucs <- getAUCs(gluc, "subject_id","time",c("glucose","glucagon","insulin","cpeptide","gip","glp1"))
-
-df_aucs <- merge(aucs,unique(gluc %>% select(subject_id)))
+aucs <- getAUCs(gluc, "subject_id", "time", c("glucose", "glucagon", "insulin", "cpeptide", "gip", "glp1"))
+
+df_aucs <- merge(aucs, unique(gluc %>% select(subject_id)))
 df_aucs$glucagon_auc <- as.numeric(df_aucs$glucagon_auc)
@@
 
-df_aucs <- df_aucs[!is.na(df_aucs$glucagon_auc) | !is.na(df_aucs$glucose_auc) |!is.na(df_aucs$insulin_auc) |!is.na(df_aucs$cpeptide_auc),]
+df_aucs <- df_aucs[!is.na(df_aucs$glucagon_auc) | !is.na(df_aucs$glucose_auc) | !is.na(df_aucs$insulin_auc) | !is.na(df_aucs$cpeptide_auc), ]
 
@@
 
-gluc$hypo_60 <- ifelse(gluc$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 60])),"Yes","No")
-gluc$hypo_70 <- ifelse(gluc$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 70])),"Yes","No")
-
-gluc_aucs <- merge(gluc_aucs,unique(gluc %>% select(subject_id, hypo_60,hypo_70)))
-
-glucose <- gluc_wide %>% select(subject_id, contains("glucose"),-starts_with("-10"))
-glucose <- glucose[complete.cases(glucose),]
-glucose[,2:10] <- sapply(glucose[,2:10], as.numeric)
-glucose$glucose_AUC <- rowSums(glucose[,2:10])
-glucose$glucose_mean <- glucose$glucose_AUC/9
-
-insulin <- gluc_wide %>% select(subject_id, contains("insulin"),-starts_with("-10"))
-insulin <- insulin[complete.cases(insulin),]
-insulin[,2:10] <- sapply(insulin[,2:10], as.numeric)
-insulin$insulin_AUC <- rowSums(insulin[,2:10])
-insulin$insulin_mean <- insulin$insulin_AUC/9
-
-cpeptide <- gluc_wide %>% select(subject_id, contains("cpeptide"),-starts_with("-10"))
-cpeptide <- cpeptide[complete.cases(cpeptide),]
-cpeptide[,2:10] <- sapply(cpeptide[,2:10], as.numeric)
-cpeptide$cpeptide_AUC <- rowSums(cpeptide[,2:10])
-cpeptide$cpeptide_mean <- cpeptide$cpeptide_AUC/9
-
-glucagon <- gluc_wide %>% select(subject_id, contains("glucagon"),-starts_with("-10"))
-glucagon <- glucagon[complete.cases(glucagon),]
-glucagon[,2:10] <- sapply(glucagon[,2:10], as.numeric)
-glucagon$glucagon_AUC <- rowSums(glucagon[,2:10])
-glucagon$glucagon_mean <- glucagon$glucagon_AUC/9
-
-gip <- gluc_wide %>% select(subject_id, contains("gip"),-starts_with("minus"))
-gip <- gip[complete.cases(gip),]
-gip[,2:10] <- sapply(gip[,2:10], as.numeric)
-gip$gip_AUC <- rowSums(gip[,2:10])
-gip$gip_mean <- gip$gip_AUC/9
-
-glp1 <- gluc_wide %>% select(subject_id, contains("glp"),-starts_with("minus"))
-glp1 <- glp1[complete.cases(glp1),]
-glp1[,2:10] <- sapply(glp1[,2:10], as.numeric)
-glp1$glp1_AUC <- rowSums(glp1[,2:10])
-glp1$glp1_mean <- glp1$glp1_AUC/9
+gluc$hypo_60 <- ifelse(gluc$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 60])), "Yes", "No")
+gluc$hypo_70 <- ifelse(gluc$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 70])), "Yes", "No")
+
+gluc_aucs <- merge(gluc_aucs, unique(gluc %>% select(subject_id, hypo_60, hypo_70)))
+
+glucose <- gluc_wide %>% select(subject_id, contains("glucose"), -starts_with("-10"))
+glucose <- glucose[complete.cases(glucose), ]
+glucose[, 2:10] <- sapply(glucose[, 2:10], as.numeric)
+glucose$glucose_AUC <- rowSums(glucose[, 2:10])
+glucose$glucose_mean <- glucose$glucose_AUC / 9
+
+insulin <- gluc_wide %>% select(subject_id, contains("insulin"), -starts_with("-10"))
+insulin <- insulin[complete.cases(insulin), ]
+insulin[, 2:10] <- sapply(insulin[, 2:10], as.numeric)
+insulin$insulin_AUC <- rowSums(insulin[, 2:10])
+insulin$insulin_mean <- insulin$insulin_AUC / 9
+
+cpeptide <- gluc_wide %>% select(subject_id, contains("cpeptide"), -starts_with("-10"))
+cpeptide <- cpeptide[complete.cases(cpeptide), ]
+cpeptide[, 2:10] <- sapply(cpeptide[, 2:10], as.numeric)
+cpeptide$cpeptide_AUC <- rowSums(cpeptide[, 2:10])
+cpeptide$cpeptide_mean <- cpeptide$cpeptide_AUC / 9
+
+glucagon <- gluc_wide %>% select(subject_id, contains("glucagon"), -starts_with("-10"))
+glucagon <- glucagon[complete.cases(glucagon), ]
+glucagon[, 2:10] <- sapply(glucagon[, 2:10], as.numeric)
+glucagon$glucagon_AUC <- rowSums(glucagon[, 2:10])
+glucagon$glucagon_mean <- glucagon$glucagon_AUC / 9
+
+gip <- gluc_wide %>% select(subject_id, contains("gip"), -starts_with("minus"))
+gip <- gip[complete.cases(gip), ]
+gip[, 2:10] <- sapply(gip[, 2:10], as.numeric)
+gip$gip_AUC <- rowSums(gip[, 2:10])
+gip$gip_mean <- gip$gip_AUC / 9
+
+glp1 <- gluc_wide %>% select(subject_id, contains("glp"), -starts_with("minus"))
+glp1 <- glp1[complete.cases(glp1), ]
+glp1[, 2:10] <- sapply(glp1[, 2:10], as.numeric)
+glp1$glp1_AUC <- rowSums(glp1[, 2:10])
+glp1$glp1_mean <- glp1$glp1_AUC / 9
 
@@
 
-auc_df$hypo_60 <- ifelse(auc_df$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 60])),"Yes","No")
-auc_df$hypo_70 <- ifelse(auc_df$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 70])),"Yes","No")
-
-auc_df$matsuda_ins <- 10000/sqrt(auc_df$`0_glucose`*auc_df$`0_insulin`*auc_df$glucose_mean*auc_df$insulin_mean)
-auc_df$matsuda_cpep <- 20/(auc_df$`0_cpeptide`*auc_df$`0_glucose`)
-auc_df$homa_ir <- auc_df$`0_glucose`*auc_df$`0_insulin`/405
-auc_df$igi <- (auc_df$`30_insulin` - auc_df$`0_insulin`)/(auc_df$`30_glucose` - auc_df$`0_glucose`)
+auc_df$hypo_60 <- ifelse(auc_df$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 60])), "Yes", "No")
+auc_df$hypo_70 <- ifelse(auc_df$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 70])), "Yes", "No")
+
+auc_df$matsuda_ins <- 10000 / sqrt(auc_df$`0_glucose` * auc_df$`0_insulin` * auc_df$glucose_mean * auc_df$insulin_mean)
+auc_df$matsuda_cpep <- 20 / (auc_df$`0_cpeptide` * auc_df$`0_glucose`)
+auc_df$homa_ir <- auc_df$`0_glucose` * auc_df$`0_insulin` / 405
+auc_df$igi <- (auc_df$`30_insulin` - auc_df$`0_insulin`) / (auc_df$`30_glucose` - auc_df$`0_glucose`)
 auc_df$igi[is.infinite(auc_df$igi)] <- NA
-auc_df$cpep_index <- 100*auc_df$`0_cpeptide`/auc_df$`0_glucose`
+auc_df$cpep_index <- 100 * auc_df$`0_cpeptide` / auc_df$`0_glucose`
 auc_df$odi_ins <- auc_df$igi * auc_df$matsuda_ins
 auc_df$odi_cpep <- auc_df$igi * auc_df$matsuda_cpep
-auc_df$cpep_ins0 <- auc_df$`0_cpeptide`/auc_df$`0_insulin`
-auc_df$cpep_auc_30 <- (auc_df$`0_cpeptide`+auc_df$`10_cpeptide`+auc_df$`20_cpeptide`+auc_df$`30_cpeptide`)/4
-auc_df$ins_auc_30 <- (auc_df$`0_insulin`+auc_df$`10_insulin`+auc_df$`20_insulin`+auc_df$`30_insulin`)/4
-auc_df$cpep_ins_auc_30 <- auc_df$cpep_auc_30/auc_df$ins_auc_30
-auc_df$cpep_ins_auc <- auc_df$cpeptide_AUC/auc_df$insulin_AUC
+auc_df$cpep_ins0 <- auc_df$`0_cpeptide` / auc_df$`0_insulin`
+auc_df$cpep_auc_30 <- (auc_df$`0_cpeptide` + auc_df$`10_cpeptide` + auc_df$`20_cpeptide` + auc_df$`30_cpeptide`) / 4
+auc_df$ins_auc_30 <- (auc_df$`0_insulin` + auc_df$`10_insulin` + auc_df$`20_insulin` + auc_df$`30_insulin`) / 4
+auc_df$cpep_ins_auc_30 <- auc_df$cpep_auc_30 / auc_df$ins_auc_30
+auc_df$cpep_ins_auc <- auc_df$cpeptide_AUC / auc_df$insulin_AUC
 
@@
 
-for (subject in auc_df$subject_id){
-  sub_df <- auc_df %>% filter(subject_id == subject) %>% select(ends_with("glucose")) 
-  gluc_peak <- c(gluc_peak,max(sub_df))
-  gluc_nadir <- c(gluc_nadir,min(sub_df))
-  gluc_peak_time <- c(gluc_peak_time,
-                      strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")[[1]][1])
-  gluc_nadir_time <- c(gluc_nadir_time,
-                       strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")[[1]][1])
-
-  sub_df <- auc_df %>% filter(subject_id == subject) %>% select(ends_with("insulin")) 
-  ins_peak <- c(ins_peak,max(sub_df))
-  ins_nadir <- c(ins_nadir,min(sub_df))
-  ins_peak_time <- c(ins_peak_time,
-                     strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")[[1]][1])
-  ins_nadir_time <- c(ins_nadir_time,
-                      strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")[[1]][1])
-
-  sub_df <- auc_df %>% filter(subject_id == subject) %>% select(ends_with("cpeptide")) 
-  cpep_peak <- c(cpep_peak,max(sub_df))
-  cpep_nadir <- c(cpep_nadir,min(sub_df))
+for (subject in auc_df$subject_id) {
+  sub_df <- auc_df %>%
+    filter(subject_id == subject) %>%
+    select(ends_with("glucose"))
+  gluc_peak <- c(gluc_peak, max(sub_df))
+  gluc_nadir <- c(gluc_nadir, min(sub_df))
+  gluc_peak_time <- c(
+    gluc_peak_time,
+    strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")[[1]][1]
+  )
+  gluc_nadir_time <- c(
+    gluc_nadir_time,
+    strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")[[1]][1]
+  )
+
+  sub_df <- auc_df %>%
+    filter(subject_id == subject) %>%
+    select(ends_with("insulin"))
+  ins_peak <- c(ins_peak, max(sub_df))
+  ins_nadir <- c(ins_nadir, min(sub_df))
+  ins_peak_time <- c(
+    ins_peak_time,
+    strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")[[1]][1]
+  )
+  ins_nadir_time <- c(
+    ins_nadir_time,
+    strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")[[1]][1]
+  )
+
+  sub_df <- auc_df %>%
+    filter(subject_id == subject) %>%
+    select(ends_with("cpeptide"))
+  cpep_peak <- c(cpep_peak, max(sub_df))
+  cpep_nadir <- c(cpep_nadir, min(sub_df))
 
@@
   cpnt_temp <- strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")
-  if(length(cppt_temp) > 0){
-    cpep_peak_time <- c(cpep_peak_time,
-                        cppt_temp[[1]][1])
+  if (length(cppt_temp) > 0) {
+    cpep_peak_time <- c(
+      cpep_peak_time,
+      cppt_temp[[1]][1]
+    )
+  } else {
+    cpep_peak_time <- c(
+      cpep_peak_time,
+      NA
+    )
   }
-  else {
-    cpep_peak_time <- c(cpep_peak_time,
-                        NA)
+  if (length(cpnt_temp) > 0) {
+    cpep_nadir_time <- c(
+      cpep_nadir_time,
+      cpnt_temp[[1]][1]
+    )
+  } else {
+    cpep_nadir_time <- c(
+      cpep_nadir_time,
+      NA
+    )
   }
-  if(length(cpnt_temp) > 0){
-    cpep_nadir_time <- c(cpep_nadir_time,
-                         cpnt_temp[[1]][1])
-  }
-  else {
-    cpep_nadir_time <- c(cpep_nadir_time,
-                         NA)
-  }
-
-  sub_df <- auc_df %>% filter(subject_id == subject) %>% select(ends_with("glucagon")) 
-  gluca_peak <- c(gluca_peak,max(sub_df))
-  gluca_nadir <- c(gluca_nadir,min(sub_df))
+
+  sub_df <- auc_df %>%
+    filter(subject_id == subject) %>%
+    select(ends_with("glucagon"))
+  gluca_peak <- c(gluca_peak, max(sub_df))
+  gluca_nadir <- c(gluca_nadir, min(sub_df))
   glucap_temp <- strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")
   glucan_temp <- strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")
-  if(length(glucap_temp) > 0){
-    gluca_peak_time <- c(gluca_peak_time,
-                         glucap_temp[[1]][1])
+  if (length(glucap_temp) > 0) {
+    gluca_peak_time <- c(
+      gluca_peak_time,
+      glucap_temp[[1]][1]
+    )
+  } else {
+    gluca_peak_time <- c(
+      gluca_peak_time,
+      NA
+    )
   }
-  else {
-    gluca_peak_time <- c(gluca_peak_time,
-                         NA)
+  if (length(glucan_temp) > 0) {
+    gluca_nadir_time <- c(
+      gluca_nadir_time,
+      glucan_temp[[1]][1]
+    )
+  } else {
+    gluca_nadir_time <- c(
+      gluca_nadir_time,
+      NA
+    )
   }
-  if(length(glucan_temp) > 0){
-    gluca_nadir_time <- c(gluca_nadir_time,
-                          glucan_temp[[1]][1])
-  }
-  else {
-    gluca_nadir_time <- c(gluca_nadir_time,
-                          NA)
-  }
-
-  sub_df <- auc_df %>% filter(subject_id == subject) %>% select(ends_with("gip")) 
-  gip_peak <- c(gip_peak,max(sub_df))
-  gip_nadir <- c(gip_nadir,min(sub_df))
+
+  sub_df <- auc_df %>%
+    filter(subject_id == subject) %>%
+    select(ends_with("gip"))
+  gip_peak <- c(gip_peak, max(sub_df))
+  gip_nadir <- c(gip_nadir, min(sub_df))
   gipp_temp <- strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")
   gipn_temp <- strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")
-  if(length(gipp_temp) > 0){
-    gip_peak_time <- c(gip_peak_time,
-                       gipp_temp[[1]][1])
+  if (length(gipp_temp) > 0) {
+    gip_peak_time <- c(
+      gip_peak_time,
+      gipp_temp[[1]][1]
+    )
+  } else {
+    gip_peak_time <- c(
+      gip_peak_time,
+      NA
+    )
   }
-  else {
-    gip_peak_time <- c(gip_peak_time,
-                       NA)
+  if (length(gipn_temp) > 0) {
+    gip_nadir_time <- c(
+      gip_nadir_time,
+      gipn_temp[[1]][1]
+    )
+  } else {
+    gip_nadir_time <- c(
+      gip_nadir_time,
+      NA
+    )
   }
-  if(length(gipn_temp) > 0){
-    gip_nadir_time <- c(gip_nadir_time,
-                        gipn_temp[[1]][1])
-  }
-  else {
-    gip_nadir_time <- c(gip_nadir_time,
-                        NA)
-  }
-
-  sub_df <- auc_df %>% filter(subject_id == subject) %>% select(ends_with("glp1")) 
-  glp1_peak <- c(glp1_peak,max(sub_df))
-  glp1_nadir <- c(glp1_nadir,min(sub_df))
+
+  sub_df <- auc_df %>%
+    filter(subject_id == subject) %>%
+    select(ends_with("glp1"))
+  glp1_peak <- c(glp1_peak, max(sub_df))
+  glp1_nadir <- c(glp1_nadir, min(sub_df))
   glp1p_temp <- strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")
   glp1n_temp <- strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")
-  if(length(gipp_temp) > 0){
-    glp1_peak_time <- c(glp1_peak_time,
-                        glp1p_temp[[1]][1])
+  if (length(gipp_temp) > 0) {
+    glp1_peak_time <- c(
+      glp1_peak_time,
+      glp1p_temp[[1]][1]
+    )
+  } else {
+    glp1_peak_time <- c(
+      glp1_peak_time,
+      NA
+    )
   }
-  else {
-    glp1_peak_time <- c(glp1_peak_time,
-                        NA)
-  }
-  if(length(glp1n_temp) > 0){
-    glp1_nadir_time <- c(glp1_nadir_time,
-                         glp1n_temp[[1]][1])
-  }
-  else {
-    glp1_nadir_time <- c(glp1_nadir_time,
-                         NA)
+  if (length(glp1n_temp) > 0) {
+    glp1_nadir_time <- c(
+      glp1_nadir_time,
+      glp1n_temp[[1]][1]
+    )
+  } else {
+    glp1_nadir_time <- c(
+      glp1_nadir_time,
+      NA
+    )
   }
@@
 
-table_dat <- merge(new_gluc,gluc_aucs)
-table_dat <- merge(table_dat, auc_df %>% select(subject_id, 
+table_dat <- merge(new_gluc, gluc_aucs)
+table_dat <- merge(table_dat, auc_df %>% select(
+  subject_id,
   matsuda_ins,
@@
 
-cgm <- dat %>% select(subject_id,
+cgm <- dat %>%
+  select(
+    subject_id,
     date_of_visit,
@@
 
-cgm$daytime_avg_sensor_glucose <- round(as.numeric(cgm$daytime_avg_sensor_glucose),4)
-cgm$percent_time_over_140 <- round(as.numeric(cgm$percent_time_over_140),4)
-cgm$percent_time_over_180 <- round(as.numeric(cgm$percent_time_over_180),4)
-cgm$daytime_sd <- round(as.numeric(cgm$daytime_sd),4)
-cgm$cov <- round(as.numeric(cgm$cov),4)
-cgm$mag <- round(as.numeric(cgm$mag),4)
+cgm$daytime_avg_sensor_glucose <- round(as.numeric(cgm$daytime_avg_sensor_glucose), 4)
+cgm$percent_time_over_140 <- round(as.numeric(cgm$percent_time_over_140), 4)
+cgm$percent_time_over_180 <- round(as.numeric(cgm$percent_time_over_180), 4)
+cgm$daytime_sd <- round(as.numeric(cgm$daytime_sd), 4)
+cgm$cov <- round(as.numeric(cgm$cov), 4)
+cgm$mag <- round(as.numeric(cgm$mag), 4)
 cgm <- unique(cgm)
-cgm <- cgm[!duplicated(cgm[,1:2]),]
-cgm <- cgm[!duplicated(cgm$subject_id),]
+cgm <- cgm[!duplicated(cgm[, 1:2]), ]
+cgm <- cgm[!duplicated(cgm$subject_id), ]
 
@@
 
-dat$mutation_severity[dat$cf_mutation_1_class <=3 | dat$cf_mutation_2_class <=3] <-  "Severe"
+dat$mutation_severity[dat$cf_mutation_1_class <= 3 | dat$cf_mutation_2_class <= 3] <- "Severe"
 dat$mutation_severity[is.na(dat$mutation_severity) & (dat$cf_mutation_1_class == 4 |
@@
   dat$cf_mutation_1_class <= 5 |
-                                                        dat$cf_mutation_2_class <=5)] <-  "Mild"
+  dat$cf_mutation_2_class <= 5)] <- "Mild"
 dat$mutation_severity[is.na(dat$mutation_severity) & (dat$cf_mutation_1_class == 6 |
@@
 
-dat_demo <- dat %>% select(subject_id, 
+dat_demo <- dat %>% select(
+  subject_id,
   date_of_visit,
@@
   cf_insulin_use,
-                           g_tube_feeds)
+  g_tube_feeds
+)
 dat_demo <- dat_demo %>% filter(subject_id %in% keep_ids)
 
-new_demo <- dat_demo[0,]
-for (id in unique(dat_demo$subject_id)){
-  id_df <- dat_demo %>% filter(subject_id == id) %>% arrange(date_of_visit) %>% na.locf(na.rm = FALSE)
-  id_df <- id_df %>% arrange(desc(date_of_visit)) %>% na.locf(na.rm = FALSE)
-  new_demo <- rbind(new_demo,id_df)
+new_demo <- dat_demo[0, ]
+for (id in unique(dat_demo$subject_id)) {
+  id_df <- dat_demo %>%
+    filter(subject_id == id) %>%
+    arrange(date_of_visit) %>%
+    na.locf(na.rm = FALSE)
+  id_df <- id_df %>%
+    arrange(desc(date_of_visit)) %>%
+    na.locf(na.rm = FALSE)
+  new_demo <- rbind(new_demo, id_df)
 }
 
-dat_demo <- new_demo %>% select(-date_of_visit) %>% unique()
-
-dat_demo <- dat_demo[!duplicated(dat_demo$subject_id),]
-
-table_dat <- merge(table_dat,dat_demo, all = T)
+dat_demo <- new_demo %>%
+  select(-date_of_visit) %>%
+  unique()
+
+dat_demo <- dat_demo[!duplicated(dat_demo$subject_id), ]
+
+table_dat <- merge(table_dat, dat_demo, all = T)
 
@@
 
-cf_groups <- dat %>% select(subject_id,group_category) %>% filter(subject_id %in% keep_ids) %>% na.omit() %>% unique()
-cf_groups$CF <- ifelse(cf_groups$group_category == "Healthy", "No","Yes")
-cf_groups <- cf_groups %>% select(subject_id, CF) %>% unique()
-table_dat <- merge(table_dat,cf_groups)
+cf_groups <- dat %>%
+  select(subject_id, group_category) %>%
+  filter(subject_id %in% keep_ids) %>%
+  na.omit() %>%
+  unique()
+cf_groups$CF <- ifelse(cf_groups$group_category == "Healthy", "No", "Yes")
+cf_groups <- cf_groups %>%
+  select(subject_id, CF) %>%
+  unique()
+table_dat <- merge(table_dat, cf_groups)
 
@@
 
-gluc <- merge(gluc,table_dat %>% select(subject_id, CF_hypo_60, CF_hypo_70))
-
-auc_df <- merge(auc_df,table_dat %>% select(subject_id, CF_hypo_60, CF_hypo_70))
+gluc <- merge(gluc, table_dat %>% select(subject_id, CF_hypo_60, CF_hypo_70))
+
+auc_df <- merge(auc_df, table_dat %>% select(subject_id, CF_hypo_60, CF_hypo_70))
 
@@
 
-table_dat <- table_dat[!is.na(table_dat$date_of_visit),]
+table_dat <- table_dat[!is.na(table_dat$date_of_visit), ]
 
@@
 
+# Remove CFRD participants (for reviewer comments)
+table_dat = table_dat %>% filter(group_category != "CFRD")
 ```
@@
 
-tab <- tableby(CF_hypo_60 ~ 
+tab <- tableby(
+  CF_hypo_60 ~
     gender +
     age +
-                 race_ethnicity+
-                 tanner_stage+
-                 weight+
-                 height+
-                 bmi+
-                 bmi_z_score+
-                 mutation_severity+
-                 fev1_percent+
-                 fvc+
-                 poc_a1c+
-                 ogtt_category+
+    race_ethnicity +
+    tanner_stage +
+    weight +
+    height +
+    bmi +
+    bmi_z_score +
+    mutation_severity +
+    fev1_percent +
+    fvc +
+    poc_a1c +
+    ogtt_category +
     cf_insulin_use +
@@
     glp1_auc +
-                 delta_glp1
-               ,
+    delta_glp1,
   data = table_dat,
-               control = control)
+  control = control
+)
 summary(tab)
@@
 ```{r, warning = F, message = F}
-pairwiseTableby <- function(tab){
+pairwiseTableby <- function(tab) {
   tbcall <- tab$Call
@@
   group_term <- tab$tables[[1]]$y$term
-  group_levels <- head(names(tab$tables[[1]]$y$stats),-1)
+  group_levels <- head(names(tab$tables[[1]]$y$stats), -1)
   comps <- combn(group_levels, 2)
 
-  for(i in seq_len(ncol(comps))){
-    comp_name <- paste(comps[,i], collapse = " vs ")
-    sub_df <- data %>% filter(get(group_term) %in% comps[,i])
-    try({
-      sub_tab <- tableby(eval(tbcall$formula),sub_df, numeric.test = "wt", wilcox.correct = TRUE, wilcox.exact = TRUE)
-      p_list <- lapply(sub_tab$tables[[1]]$tables[[1]], function(x) x$test$p.value)
-      if(exists("results_frame")){
+  for (i in seq_len(ncol(comps))) {
+    comp_name <- paste(comps[, i], collapse = " vs ")
+    sub_df <- data %>% filter(get(group_term) %in% comps[, i])
+    try(
+      {
+        sub_tab <- tableby(eval(tbcall$formula), sub_df, numeric.test = "wt", 
+                           wilcox.correct = TRUE, wilcox.exact = TRUE)
+        p_list <- 
+          lapply(sub_tab$tables[[1]]$tables[[1]], function(x) x$test$p.value)
+        if (exists("results_frame")) {
           results_frame[comp_name] <- unlist(p_list)
-      }
-      else{
+        } else {
           results_frame <- data.frame(row.names = names(unlist(p_list)))
@@
         }
-    }, silent = TRUE)
-
+      },
+      silent = TRUE
+    )
   }
@@
 kbl(ptab, caption = "Pairwise p-values") %>%
-  column_spec(2,background = ifelse(ptab$`CF/Hypo vs CF/No-Hypo` < 0.05,"red","white")) %>%
-  column_spec(3,background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/Hypo`< 0.05,"red","white"),"white)")) %>%
-  column_spec(4,background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/No-Hypo` < 0.05,"red","white"),"white)"))  %>%
-  column_spec(5,background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/Hypo` < 0.05,"red","white"),"white)"))  %>%
-  column_spec(6,background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/No-Hypo` < 0.05,"red","white"),"white)")) 
+  column_spec(2, background = 
+                ifelse(ptab$`CF/Hypo vs CF/No-Hypo` < 0.05, "red", "white")) %>%
+  column_spec(3, background = 
+                na.fill(ifelse(ptab$`CF/Hypo vs No-CF/Hypo` < 0.05, 
+                               "red", "white"), "white)")) %>%
+  column_spec(4, background = 
+                na.fill(ifelse(ptab$`CF/Hypo vs No-CF/No-Hypo` < 0.05, 
+                               "red", "white"), "white)")) %>%
+  column_spec(5, background = 
+                na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/Hypo` < 0.05, 
+                               "red", "white"), "white)")) %>%
+  column_spec(6, background = 
+                na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/No-Hypo` < 0.05, 
+                               "red", "white"), "white)"))
 ```
@@
 
-tab <- tableby(formula(
-  paste("CF_hypo_60 ~", 
-        paste(names(cgm_table_dat %>% 
+tab <- tableby(
+  formula(
+    paste(
+      "CF_hypo_60 ~",
+      paste(
+        names(cgm_table_dat %>%
           select(num_days_good_data:interquartile_range)),
-              collapse = "+")))
-  , 
+        collapse = "+"
+      )
+    )
+  ),
   data = cgm_table_dat,
-  control = control)
+  control = control
+)
 summary(tab)
@@
 ```{r, warning = F, message = F}
-model <- lme(glucagon ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
-
-p1 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
+model <- lme(glucagon ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
+             correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
+
+p1 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
+                               color = CF_hypo_60, shape = CF_hypo_60)) +
   geom_point() +
@@
   theme_light() +
-  scale_color_brewer(palette="Dark2")
-
-model <- lme(glucose ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
-
-p2 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
+  scale_color_brewer(palette = "Dark2")
+
+model <- lme(glucose ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
+             correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
+
+p2 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
+                               color = CF_hypo_60, shape = CF_hypo_60)) +
   geom_point() +
@@
   theme_light() +
-  scale_color_brewer(palette="Dark2")
-
-model <- lme(insulin ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
-
-p3 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
+  scale_color_brewer(palette = "Dark2")
+
+model <- lme(insulin ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
+             correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
+
+p3 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
+                               color = CF_hypo_60, shape = CF_hypo_60)) +
   geom_point() +
@@
   theme_light() +
-  scale_color_brewer(palette="Dark2")
-
-model <- lme(cpeptide ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
-
-p4 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
+  scale_color_brewer(palette = "Dark2")
+
+model <- lme(cpeptide ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
+             correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
+
+p4 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
+                               color = CF_hypo_60, shape = CF_hypo_60)) +
   geom_point() +
@@
   theme_light() +
-  scale_color_brewer(palette="Dark2")
-
-model <- lme(gip ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
-
-p5 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
+  scale_color_brewer(palette = "Dark2")
+
+model <- lme(gip ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
+             correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
+
+p5 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
+                               color = CF_hypo_60, shape = CF_hypo_60)) +
   geom_point() +
@@
   theme_light() +
-  scale_color_brewer(palette="Dark2")
-
-model <- lme(glp1 ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
-
-p6 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, color = CF_hypo_60, shape = CF_hypo_60)) +
+  scale_color_brewer(palette = "Dark2")
+
+model <- lme(glp1 ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
+             correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
+
+p6 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
+                               color = CF_hypo_60, shape = CF_hypo_60)) +
   geom_point() +
@@
   theme_light() +
-  scale_color_brewer(palette="Dark2")
-
-p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
-  theme(legend.position='bottom')
+  scale_color_brewer(palette = "Dark2")
+
+p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
+  theme(legend.position = "bottom")
 
 png("ogtt_60.png", res = 300, width = 2000, height = 3000)
-p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
-  theme(legend.position='bottom')
+p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
+  theme(legend.position = "bottom")
 dev.off()
@@
 ```{r, warning = F, message = F}
-model <- lme(glucagon ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
-
-p1 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
+model <- lme(glucagon ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
+             correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
+
+p1 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
+                               linetype = CF_hypo_60, shape = CF_hypo_60)) +
   geom_point() +
@@
 
-model <- lme(glucose ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
-
-p2 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
+model <- lme(glucose ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
+             correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
+
+p2 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
+                               linetype = CF_hypo_60, shape = CF_hypo_60)) +
   geom_point() +
@@
 
-model <- lme(insulin ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
-
-p3 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
+model <- lme(insulin ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
+             correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
+
+p3 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
+                               linetype = CF_hypo_60, shape = CF_hypo_60)) +
   geom_point() +
@@
 
-model <- lme(cpeptide ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
-
-p4 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
+model <- lme(cpeptide ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
+             correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
+
+p4 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
+                               linetype = CF_hypo_60, shape = CF_hypo_60)) +
   geom_point() +
@@
 
-model <- lme(gip ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
-
-p5 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
+model <- lme(gip ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
+             correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
+
+p5 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
+                               linetype = CF_hypo_60, shape = CF_hypo_60)) +
   geom_point() +
@@
 
-model <- lme(glp1 ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
-
-p6 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, linetype = CF_hypo_60, shape = CF_hypo_60)) +
+model <- lme(glp1 ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, 
+             correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
+
+p6 <- ggplot(data = means, aes(y = emmean, x = time, group = CF_hypo_60, 
+                               linetype = CF_hypo_60, shape = CF_hypo_60)) +
   geom_point() +
@@
 
-p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
-  theme(legend.position='bottom')
+p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
+  theme(legend.position = "bottom")
 
 png("ogtt_60_bw.png", res = 300, width = 2000, height = 3000)
-p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
-  theme(legend.position='bottom')
+p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
+  theme(legend.position = "bottom")
 dev.off()
@@
   labs(y = "Matsuda Index", x = "", color = "") +
-  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo"),c("CF/Hypo","No-CF/No-Hypo")),
+  geom_signif(
+    comparisons = list(c("CF/Hypo", "CF/No-Hypo"), c("CF/Hypo", "No-CF/No-Hypo")),
     color = "black",
-              y_position = c(45,50),
-              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
+    y_position = c(45, 50),
+    map_signif_level = function(p) sprintf("p = %.2g", p)
+  ) +
   theme_light() +
   guides(color = FALSE) +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  scale_color_brewer(palette="Dark2")
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_color_brewer(palette = "Dark2")
 
@@
   labs(y = "HOMA IR", x = "", color = "") +
-  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo")),
+  geom_signif(
+    comparisons = list(c("CF/Hypo", "CF/No-Hypo")),
     color = "black",
-              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
+    map_signif_level = function(p) sprintf("p = %.2g", p)
+  ) +
   theme_light() +
   guides(color = FALSE) +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  scale_color_brewer(palette="Dark2")
-
-p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
-  theme(legend.position='bottom')
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_color_brewer(palette = "Dark2")
+
+p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
+  theme(legend.position = "bottom")
 
 png("resistance_60.png", height = 2000, width = 3000, res = 300)
-p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
-  theme(legend.position='bottom')
+p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
+  theme(legend.position = "bottom")
 dev.off()
@@
   labs(y = "Matsuda Index", x = "", color = "") +
-  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo"),c("CF/Hypo","No-CF/No-Hypo")),
+  geom_signif(
+    comparisons = list(c("CF/Hypo", "CF/No-Hypo"), c("CF/Hypo", "No-CF/No-Hypo")),
     color = "black",
-              y_position = c(45,50),
-              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
+    y_position = c(45, 50),
+    map_signif_level = function(p) sprintf("p = %.2g", p)
+  ) +
   theme_light() +
   guides(color = FALSE) +
-  theme(axis.text.x = element_text(angle = 45, hjust=1))
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
@@
   labs(y = "HOMA IR", x = "", color = "") +
-  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo")),
+  geom_signif(
+    comparisons = list(c("CF/Hypo", "CF/No-Hypo")),
     color = "black",
-              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
+    map_signif_level = function(p) sprintf("p = %.2g", p)
+  ) +
   theme_light() +
   guides(color = FALSE) +
-  theme(axis.text.x = element_text(angle = 45, hjust=1))
-
-p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
-  theme(legend.position='bottom')
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
+  theme(legend.position = "bottom")
 
 png("resistance_60_bw.png", height = 2000, width = 3000, res = 300)
-p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
-  theme(legend.position='bottom')
+p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
+  theme(legend.position = "bottom")
 dev.off()
@@
 ```{r, results='asis', warning = F, message = F}
-tab <- tableby(CF_hypo_70 ~ 
+tab <- tableby(
+  CF_hypo_70 ~
     gender +
     age +
-                 race_ethnicity+
-                 tanner_stage+
-                 weight+
-                 height+
-                 bmi+
-                 bmi_z_score+
-                 mutation_severity+
-                 fev1_percent+
-                 fvc+
-                 poc_a1c+
-                 ogtt_category+
+    race_ethnicity +
+    tanner_stage +
+    weight +
+    height +
+    bmi +
+    bmi_z_score +
+    mutation_severity +
+    fev1_percent +
+    fvc +
+    poc_a1c +
+    ogtt_category +
     cf_insulin_use +
@@
 kbl(ptab, caption = "Pairwise p-values") %>%
-  column_spec(2,background = ifelse(ptab$`CF/Hypo vs CF/No-Hypo` < 0.05,"red","white")) %>%
-  column_spec(3,background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/Hypo`< 0.05,"red","white"),"white)")) %>%
-  column_spec(4,background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/No-Hypo` < 0.05,"red","white"),"white)"))  %>%
-  column_spec(5,background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/Hypo` < 0.05,"red","white"),"white)"))  %>%
-  column_spec(6,background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/No-Hypo` < 0.05,"red","white"),"white)")) 
+  column_spec(2, background = ifelse(ptab$`CF/Hypo vs CF/No-Hypo` < 0.05, "red", "white")) %>%
+  column_spec(3, background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/Hypo` < 0.05, "red", "white"), "white)")) %>%
+  column_spec(4, background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/No-Hypo` < 0.05, "red", "white"), "white)")) %>%
+  column_spec(5, background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/Hypo` < 0.05, "red", "white"), "white)")) %>%
+  column_spec(6, background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/No-Hypo` < 0.05, "red", "white"), "white)"))
 ```
@@
 ```{r, results='asis', warning = F, message = F}
-tab <- tableby(formula(
-  paste("CF_hypo_70 ~", 
-        paste(names(cgm_table_dat %>% 
+tab <- tableby(
+  formula(
+    paste(
+      "CF_hypo_70 ~",
+      paste(
+        names(cgm_table_dat %>%
           select(num_days_good_data:interquartile_range)),
-              collapse = "+")))
-  , 
+        collapse = "+"
+      )
+    )
+  ),
   data = cgm_table_dat,
-  control = control)
+  control = control
+)
 summary(tab)
@@
 ```{r, warning = F, message = F}
-model <- lme(glucagon ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))
+model <- lme(glucagon ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))
 
@@
   theme_light() +
-  scale_color_brewer(palette="Dark2")
-
-model <- lme(glucose ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))
+  scale_color_brewer(palette = "Dark2")
+
+model <- lme(glucose ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))
 
@@
   theme_light() +
-  scale_color_brewer(palette="Dark2")
-
-model <- lme(insulin ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))
+  scale_color_brewer(palette = "Dark2")
+
+model <- lme(insulin ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))
 
@@
   theme_light() +
-  scale_color_brewer(palette="Dark2")
-
-model <- lme(cpeptide ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))
+  scale_color_brewer(palette = "Dark2")
+
+model <- lme(cpeptide ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))
 
@@
   theme_light() +
-  scale_color_brewer(palette="Dark2")
-
-model <- lme(gip ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))
+  scale_color_brewer(palette = "Dark2")
+
+model <- lme(gip ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))
 
@@
   theme_light() +
-  scale_color_brewer(palette="Dark2")
-
-model <- lme(glp1 ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))
+  scale_color_brewer(palette = "Dark2")
+
+model <- lme(glp1 ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))
 
@@
   theme_light() +
-  scale_color_brewer(palette="Dark2")
-
-p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
-  theme(legend.position='bottom')
+  scale_color_brewer(palette = "Dark2")
+
+p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
+  theme(legend.position = "bottom")
 
 png("ogtt_70.png", res = 300, width = 2000, height = 3000)
-p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
-  theme(legend.position='bottom')
+p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
+  theme(legend.position = "bottom")
 dev.off()
@@
 
-model <- lme(glucose ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))
+model <- lme(glucose ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))
 
@@
 
-model <- lme(insulin ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))
+model <- lme(insulin ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))
 
@@
 
-model <- lme(cpeptide ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))
+model <- lme(cpeptide ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))
 
@@
 
-model <- lme(gip ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))
+model <- lme(gip ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))
 
@@
 
-model <- lme(glp1 ~ time*CF_hypo_70, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_70))
+model <- lme(glp1 ~ time * CF_hypo_70, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_70))
 
@@
 
-p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
-  theme(legend.position='bottom')
+p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
+  theme(legend.position = "bottom")
 
 png("ogtt_70_bw.png", res = 300, width = 2000, height = 3000)
-p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
-  theme(legend.position='bottom')
+p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
+  theme(legend.position = "bottom")
 dev.off()
@@
   labs(y = "Matsuda Index", x = "", color = "") +
-  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo")),
+  geom_signif(
+    comparisons = list(c("CF/Hypo", "CF/No-Hypo")),
     color = "black",
     test = "wilcox.test",
-              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
+    map_signif_level = function(p) sprintf("p = %.2g", p)
+  ) +
   theme_light() +
   guides(color = FALSE) +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  scale_color_brewer(palette="Dark2")
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_color_brewer(palette = "Dark2")
 
@@
   labs(y = "HOMA IR", x = "", color = "") +
-  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo")),
+  geom_signif(
+    comparisons = list(c("CF/Hypo", "CF/No-Hypo")),
     color = "black",
     test = "wilcox.test",
-              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
+    map_signif_level = function(p) sprintf("p = %.2g", p)
+  ) +
   theme_light() +
   guides(color = FALSE) +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  scale_color_brewer(palette="Dark2")
-
-p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
-  theme(legend.position='bottom')
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_color_brewer(palette = "Dark2")
+
+p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
+  theme(legend.position = "bottom")
 
 png("resistance_70.png", height = 2000, width = 3000, res = 300)
-p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
-  theme(legend.position='bottom')
+p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
+  theme(legend.position = "bottom")
 dev.off()
@@
   labs(y = "Matsuda Index", x = "", color = "") +
-  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo")),
+  geom_signif(
+    comparisons = list(c("CF/Hypo", "CF/No-Hypo")),
     color = "black",
     test = "wilcox.test",
-              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
+    map_signif_level = function(p) sprintf("p = %.2g", p)
+  ) +
   theme_light() +
   guides(color = FALSE) +
-  theme(axis.text.x = element_text(angle = 45, hjust=1))
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
@@
   labs(y = "HOMA IR", x = "", color = "") +
-  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo")),
+  geom_signif(
+    comparisons = list(c("CF/Hypo", "CF/No-Hypo")),
     color = "black",
     test = "wilcox.test",
-              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
+    map_signif_level = function(p) sprintf("p = %.2g", p)
+  ) +
   theme_light() +
   guides(color = FALSE) +
-  theme(axis.text.x = element_text(angle = 45, hjust=1))
-
-p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
-  theme(legend.position='bottom')
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
+  theme(legend.position = "bottom")
 
 png("resistance_70_bw.png", height = 2000, width = 3000, res = 300)
-p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
-  theme(legend.position='bottom')
+p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
+  theme(legend.position = "bottom")
 dev.off()
@@
 
-for (i in seq_len(length(outcomes))){
-
+for (i in seq_len(length(outcomes))) {
   p70 <- ggplot(data = cgm_table_dat, aes_string(y = outcomes[i], x = "CF_hypo_70")) +
-    labs(y = outcomes_labs[i], x= "") +
+    labs(y = outcomes_labs[i], x = "") +
     geom_jitter(alpha = 0.5, width = 0.1) +
@@
     ggtitle("Hypoglycemia by OGTT (< 70)") +
-    theme(axis.text.x = element_text(angle = 45, hjust=1))
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
   p60 <- ggplot(data = cgm_table_dat, aes_string(y = outcomes[i], x = "CF_hypo_60")) +
-    labs(y = outcomes_labs[i], x= "") +
+    labs(y = outcomes_labs[i], x = "") +
     geom_jitter(alpha = 0.5, width = 0.1) +
@@
     ggtitle("Hypoglycemia by OGTT (< 60)") +
-    theme(axis.text.x = element_text(angle = 45, hjust=1))
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
   cat("\n")
-  cat(paste("##",outcomes_labs[i]))
+  cat(paste("##", outcomes_labs[i]))
   cat("\n")
@@
 }
-
 ```
@@
 
-under_70_cutoff <- 2*2.06
-under_60_cutoff <- 2*0.687
+under_70_cutoff <- 2 * 2.06
+under_60_cutoff <- 2 * 0.687
 
@@
   test = F,
-               total = F)
+  total = F
+)
 summary(tab)
@@
 ```{r}
-mat <- matrix( data = c(4,12,16,23), nrow = 2, ncol = 2)
+mat <- matrix(data = c(4, 12, 16, 23), nrow = 2, ncol = 2)
 Kappa(mat)
@@
 ```{r}
-mat <- matrix( data = c(1,2,3,9), nrow = 2, ncol = 2)
+mat <- matrix(data = c(1, 2, 3, 9), nrow = 2, ncol = 2)
 Kappa(mat)
@@
 ```{r}
-mat <- matrix( data = c(2,12,13,28), nrow = 2, ncol = 2)
+mat <- matrix(data = c(2, 12, 13, 28), nrow = 2, ncol = 2)
 Kappa(mat)
@@
 ```{r}
-mat <- matrix( data = c(1,1,1,12), nrow = 2, ncol = 2)
+mat <- matrix(data = c(1, 1, 1, 12), nrow = 2, ncol = 2)
 Kappa(mat)
@@
 ```{r}
-mat <- matrix( data = c(33,21,13,3), nrow = 2, ncol = 2)
+mat <- matrix(data = c(33, 21, 13, 3), nrow = 2, ncol = 2)
 Kappa(mat)
@@
 ```{r}
-mat <- matrix( data = c(39,12,14,5), nrow = 2, ncol = 2)
+mat <- matrix(data = c(39, 12, 14, 5), nrow = 2, ncol = 2)
 Kappa(mat)
@@
 ```{r, warning = F, message = F}
-model <- lme(glucagon ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
+model <- lme(glucagon ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
 
@@
   theme_light() +
-  scale_color_brewer(palette="Dark2")
-
-model <- lme(glucose ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
+  scale_color_brewer(palette = "Dark2")
+
+model <- lme(glucose ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
 
@@
   theme_light() +
-  scale_color_brewer(palette="Dark2")
-
-model <- lme(insulin ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
+  scale_color_brewer(palette = "Dark2")
+
+model <- lme(insulin ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
 
@@
   theme_light() +
-  scale_color_brewer(palette="Dark2")
-
-model <- lme(cpeptide ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
+  scale_color_brewer(palette = "Dark2")
+
+model <- lme(cpeptide ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
 
@@
   theme_light() +
-  scale_color_brewer(palette="Dark2")
-
-model <- lme(gip ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
+  scale_color_brewer(palette = "Dark2")
+
+model <- lme(gip ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
 
@@
   theme_light() +
-  scale_color_brewer(palette="Dark2")
-
-model <- lme(glp1 ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
+  scale_color_brewer(palette = "Dark2")
+
+model <- lme(glp1 ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
 
@@
   theme_light() +
-  scale_color_brewer(palette="Dark2")
-
-p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
-  theme(legend.position='bottom')
+  scale_color_brewer(palette = "Dark2")
+
+p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
+  theme(legend.position = "bottom")
 
 png("ogtt_60_noinsulin.png", res = 300, width = 2000, height = 3000)
-p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
-  theme(legend.position='bottom')
+p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
+  theme(legend.position = "bottom")
 dev.off()
@@
 
-model <- lme(glucose ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
+model <- lme(glucose ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
 
@@
 
-model <- lme(insulin ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
+model <- lme(insulin ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
 
@@
 
-model <- lme(cpeptide ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
+model <- lme(cpeptide ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
 
@@
 
-model <- lme(gip ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
+model <- lme(gip ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
 
@@
 
-model <- lme(glp1 ~ time*CF_hypo_60, data = gluc, random = ~1|subject_id, correlation = corAR1(), na.action = na.omit)
-means <- as.data.frame(emmeans(model, ~ time|CF_hypo_60))
+model <- lme(glp1 ~ time * CF_hypo_60, data = gluc, random = ~ 1 | subject_id, correlation = corAR1(), na.action = na.omit)
+means <- as.data.frame(emmeans(model, ~ time | CF_hypo_60))
 
@@
 
-p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
-  theme(legend.position='bottom')
+p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
+  theme(legend.position = "bottom")
 
 png("ogtt_60_noinsulin_bw.png", res = 300, width = 2000, height = 3000)
-p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
-  theme(legend.position='bottom')
+p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
+  theme(legend.position = "bottom")
 dev.off()
@@
   labs(y = "HOMA IR", x = "", color = "") +
-  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo"),c("CF/Hypo","No-CF/No-Hypo")),
+  geom_signif(
+    comparisons = list(c("CF/Hypo", "CF/No-Hypo"), c("CF/Hypo", "No-CF/No-Hypo")),
     color = "black",
-              y_position = c(4.5,5),
-              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
+    y_position = c(4.5, 5),
+    map_signif_level = function(p) sprintf("p = %.2g", p)
+  ) +
   theme_light() +
   guides(color = FALSE) +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  scale_color_brewer(palette="Dark2")
-
-p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
-  theme(legend.position='bottom')
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_color_brewer(palette = "Dark2")
+
+p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
+  theme(legend.position = "bottom")
 
 png("resistance_60_noinsulin.png", height = 2000, width = 3000, res = 300)
-p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
-  theme(legend.position='bottom')
+p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
+  theme(legend.position = "bottom")
 dev.off()
@@
   labs(y = "HOMA IR", x = "", color = "") +
-  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo"),c("CF/Hypo","No-CF/No-Hypo")),
+  geom_signif(
+    comparisons = list(c("CF/Hypo", "CF/No-Hypo"), c("CF/Hypo", "No-CF/No-Hypo")),
     color = "black",
-              y_position = c(4.5,5),
-              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
+    y_position = c(4.5, 5),
+    map_signif_level = function(p) sprintf("p = %.2g", p)
+  ) +
   theme_light() +
   guides(color = FALSE) +
-  theme(axis.text.x = element_text(angle = 45, hjust=1))
-
-p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
-  theme(legend.position='bottom')
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
+  theme(legend.position = "bottom")
 
 png("resistance_60_noinsulin_bw.png", height = 2000, width = 3000, res = 300)
-p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
-  theme(legend.position='bottom')
+p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
+  theme(legend.position = "bottom")
 dev.off()
@@
   theme_light() +
-  scale_color_brewer(palette="Dark2")
-
-p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
-  theme(legend.position='bottom')
+  scale_color_brewer(palette = "Dark2")
+
+p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
+  theme(legend.position = "bottom")
 
 png("ogtt_70_noinsulin.png", res = 300, width = 2000, height = 3000)
-p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
-  theme(legend.position='bottom')
+p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
+  theme(legend.position = "bottom")
 dev.off()
@@
 
-p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
-  theme(legend.position='bottom')
+p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
+  theme(legend.position = "bottom")
 
 png("ogtt_70_noinsulin_bw.png", res = 300, width = 2000, height = 3000)
-p2+p3+p1+p4+p5+p6+plot_layout(guides = "collect", ncol = 2, nrow = 3) &
-  theme(legend.position='bottom')
+p2 + p3 + p1 + p4 + p5 + p6 + plot_layout(guides = "collect", ncol = 2, nrow = 3) &
+  theme(legend.position = "bottom")
 dev.off()
@@
   labs(y = "Matsuda Index", x = "", color = "") +
-  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo"),c("CF/Hypo","No-CF/No-Hypo")),
+  geom_signif(
+    comparisons = list(c("CF/Hypo", "CF/No-Hypo"), c("CF/Hypo", "No-CF/No-Hypo")),
     color = "black",
-              y_position = c(45,50),
+    y_position = c(45, 50),
     test = "wilcox.test",
-              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
+    map_signif_level = function(p) sprintf("p = %.2g", p)
+  ) +
   theme_light() +
   guides(color = FALSE) +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  scale_color_brewer(palette="Dark2")
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_color_brewer(palette = "Dark2")
 
@@
   labs(y = "HOMA IR", x = "", color = "") +
-  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo"),c("CF/Hypo","No-CF/No-Hypo")),
+  geom_signif(
+    comparisons = list(c("CF/Hypo", "CF/No-Hypo"), c("CF/Hypo", "No-CF/No-Hypo")),
     color = "black",
-              y_position = c(4.5,5),
+    y_position = c(4.5, 5),
     test = "wilcox.test",
-              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
+    map_signif_level = function(p) sprintf("p = %.2g", p)
+  ) +
   theme_light() +
   guides(color = FALSE) +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  scale_color_brewer(palette="Dark2")
-
-p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
-  theme(legend.position='bottom')
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_color_brewer(palette = "Dark2")
+
+p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
+  theme(legend.position = "bottom")
 
 png("resistance_70_noinsulin.png", height = 2000, width = 3000, res = 300)
-p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
-  theme(legend.position='bottom')
+p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
+  theme(legend.position = "bottom")
 dev.off()
@@
   labs(y = "Matsuda Index", x = "", color = "") +
-  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo"),c("CF/Hypo","No-CF/No-Hypo")),
+  geom_signif(
+    comparisons = list(c("CF/Hypo", "CF/No-Hypo"), c("CF/Hypo", "No-CF/No-Hypo")),
     color = "black",
-              y_position = c(45,50),
+    y_position = c(45, 50),
     test = "wilcox.test",
-              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
+    map_signif_level = function(p) sprintf("p = %.2g", p)
+  ) +
   theme_light() +
   guides(color = FALSE) +
-  theme(axis.text.x = element_text(angle = 45, hjust=1))
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
@@
   labs(y = "HOMA IR", x = "", color = "") +
-  geom_signif(comparisons = list(c("CF/Hypo","CF/No-Hypo"),c("CF/Hypo","No-CF/No-Hypo")),
+  geom_signif(
+    comparisons = list(c("CF/Hypo", "CF/No-Hypo"), c("CF/Hypo", "No-CF/No-Hypo")),
     color = "black",
-              y_position = c(4.5,5),
+    y_position = c(4.5, 5),
     test = "wilcox.test",
-              map_signif_level=function(p)sprintf("p = %.2g", p)) + 
+    map_signif_level = function(p) sprintf("p = %.2g", p)
+  ) +
   theme_light() +
   guides(color = FALSE) +
-  theme(axis.text.x = element_text(angle = 45, hjust=1))
-
-p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
-  theme(legend.position='bottom')
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
+  theme(legend.position = "bottom")
 
 png("resistance_70_noinsulin_bw.png", height = 2000, width = 3000, res = 300)
-p1+p2+plot_layout(guides = "collect", ncol = 2, nrow = 1) &
-  theme(legend.position='bottom')
+p1 + p2 + plot_layout(guides = "collect", ncol = 2, nrow = 1) &
+  theme(legend.position = "bottom")
 dev.off()
@@
 ```{r}
-mat <- matrix( data = c(3,10,15,16), nrow = 2, ncol = 2)
+mat <- matrix(data = c(3, 10, 15, 16), nrow = 2, ncol = 2)
 Kappa(mat)
@@
 ```{r}
-mat <- matrix( data = c(2,9,12,21), nrow = 2, ncol = 2)
+mat <- matrix(data = c(2, 9, 12, 21), nrow = 2, ncol = 2)
 Kappa(mat)

### 2023-03-31 Jordana Hoppe/nacfc_2023.qmd
Update nacfc_2023.qmd
added Cohen's kappa

@@
 library(caret)
+library(irr)
 library(gtsummary)
@@
 # Shorten organism names
+bugs = colnames(df)[grep("If.culture.was.positive..select.all.organisms.isolated...choice.",colnames(df))]
+bugs <- sub("If.culture.was.positive..select.all.organisms.isolated...choice.", "", bugs)
+bugs = bugs[1:16]
 colnames(df) <- sub("If.culture.was.positive..select.all.organisms.isolated...choice.", "", colnames(df))
@@
 
-# Sensitivity and specificity
+# Pairwise Comparisons
+
+Cohen's kappa is an agreement statistic ranging from -1 (perfect disagreement) to 1 (perfect agreement). A rough guideline for interpreting Cohen's kappa is:
+
+| Cohen's Kappa | Interpretation         |
+|---------------|------------------------|
+| 0             | No agreement           |
+| 0.10 - 0.20   | Slight agreement       |
+| 0.21 - 0.40   | Fair agreement         |
+| 0.41 - 0.60   | Moderate agreement     |
+| 0.61 - 0.80   | Substantial agreement  |
+| 0.81 - 0.99   | Near perfect agreement |
+| 1             | Perfect agreement      |
 
@@
     cat("\n")
-    cat("#### Contingency table")
-    cat("\n")
     cat("All samples marked as", unique(gold_standard))
@@
   } else {
+    # Sensitivity and specificity
     sen <- sensitivity(data = factor(other_test), reference = factor(gold_standard))
     spec <- specificity(data = factor(other_test), reference = factor(gold_standard))
-
+    # Kappa
+    k = kappa2(cbind(gold_standard,other_test))
+    # Formatting
     dimnames(t) <- lapply(dimnames(t), function(n) {
@@
     cat("\n")
-    cat("#### Contingency table")
-    cat("\n")
     print(knitr::kable((as.data.frame.matrix(t))))
     cat("\n")
-    cat("Sensitivity of the", tolower(type), "method was", round(sen, 3), "and specificity was", round(spec, 3))
+    cat("Sensitivity of the", tolower(type), "method was", round(sen, 3), "and specificity was", paste0(round(spec, 3),"."))
+    cat("\n")
+    cat("Cohen's kappa for the two tests was", round(k$value, 3), paste0("(p = ", format.pval(k$p.value,digits = 3, eps = 0.001),")"))
     cat("\n")
@@
 ```
+
+## Delayed Processing
+
+```{r results='asis'}
+sens_spec(df$Methicillin.sensitive.Staph.aureus.,
+  df$Methicillin.sensitive.Staph.aureus..2,
+  type = "Delayed Processing", bug = "MSSA"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$Methicillin.resistant.Staph.aureus.,
+  df$Methicillin.resistant.Staph.aureus..2,
+  type = "Delayed Processing", bug = "MRSA"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$A..xylosoxidans., df$A..xylosoxidans..2,
+  type = "Delayed Processing", bug = "A. xylosoxidans"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$Achromobacter.not.xylosoxidans., df$Achromobacter.not.xylosoxidans..2,
+  type = "Delayed Processing", bug = "Achromobacter (not xylosoxidans)"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$Acinetobacter., df$Acinetobacter..2,
+  type = "Delayed Processing", bug = "Acinetobacter"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$H..influenzae., df$H..influenzae..2,
+  type = "Delayed Processing", bug = "H. influenzae"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$P..aeruginosa., df$P..aeruginosa..2,
+  type = "Delayed Processing", bug = "P. aeruginosa"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$P..aeruginosa..mucoid.., df$P..aeruginosa..mucoid...2,
+  type = "Delayed Processing", bug = "P. aeruginosa (mucoid)"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$Pseudomonas.species..not.aeruginosa., df$Pseudomonas.species..not.aeruginosa..2,
+  type = "Delayed Processing", bug = "Pseudomonas (not aeruginosa)"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$B..cepacia.complex., df$B..cepacia.complex..2,
+  type = "Delayed Processing", bug = "B. cepacia complex"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$Stenotrophomonas., df$Stenotrophomonas..2,
+  type = "Delayed Processing", bug = "Stenotrophomonas"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$Aspergillus., df$Aspergillus..2,
+  type = "Delayed Processing", bug = "Aspergillus"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$Penicillium.species., df$Penicillium.species..2,
+  type = "Delayed Processing", bug = "Penicillium species"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$Sceosporium.species., df$Sceosporium.species..2,
+  type = "Delayed Processing", bug = "Sceosporium species"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$M..avium.complex., df$M..avium.complex..2,
+  type = "Delayed Processing", bug = "M. avium complex"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$M..abscessus., df$M..abscessus..2,
+  type = "Delayed Processing", bug = "M. abscessus"
+)
+```
+
+# Agreement between all 3 methods
+
+```{r}
+
+```

### 2023-04-03 Jordana Hoppe/nacfc_2023.qmd
Fleiss' kappa for Jordana

@@
     cat("\n")
-    print(knitr::kable((as.data.frame.matrix(t))))
+    print(knitr::kable(as.data.frame.matrix(t)))
     cat("\n")
@@
 
+## Delayed Processing
+
+```{r results='asis'}
+sens_spec(df$Methicillin.sensitive.Staph.aureus.,
+  df$Methicillin.sensitive.Staph.aureus..1,
+  type = "Delayed Processing", bug = "MSSA"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$Methicillin.resistant.Staph.aureus.,
+  df$Methicillin.resistant.Staph.aureus..1,
+  type = "Delayed Processing", bug = "MRSA"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$A..xylosoxidans., df$A..xylosoxidans..1,
+  type = "Delayed Processing", bug = "A. xylosoxidans"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$Achromobacter.not.xylosoxidans., df$Achromobacter.not.xylosoxidans..1,
+  type = "Delayed Processing", bug = "Achromobacter (not xylosoxidans)"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$Acinetobacter., df$Acinetobacter..1,
+  type = "Delayed Processing", bug = "Acinetobacter"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$H..influenzae., df$H..influenzae..1,
+  type = "Delayed Processing", bug = "H. influenzae"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$P..aeruginosa., df$P..aeruginosa..1,
+  type = "Delayed Processing", bug = "P. aeruginosa"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$P..aeruginosa..mucoid.., df$P..aeruginosa..mucoid...1,
+  type = "Delayed Processing", bug = "P. aeruginosa (mucoid)"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$Pseudomonas.species..not.aeruginosa., df$Pseudomonas.species..not.aeruginosa..1,
+  type = "Delayed Processing", bug = "Pseudomonas (not aeruginosa)"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$B..cepacia.complex., df$B..cepacia.complex..1,
+  type = "Delayed Processing", bug = "B. cepacia complex"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$Stenotrophomonas., df$Stenotrophomonas..1,
+  type = "Delayed Processing", bug = "Stenotrophomonas"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$Aspergillus., df$Aspergillus..1,
+  type = "Delayed Processing", bug = "Aspergillus"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$Penicillium.species., df$Penicillium.species..1,
+  type = "Delayed Processing", bug = "Penicillium species"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$Sceosporium.species., df$Sceosporium.species..1,
+  type = "Delayed Processing", bug = "Sceosporium species"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$M..avium.complex., df$M..avium.complex..1,
+  type = "Delayed Processing", bug = "M. avium complex"
+)
+```
+
+```{r results='asis'}
+sens_spec(df$M..abscessus., df$M..abscessus..1,
+  type = "Delayed Processing", bug = "M. abscessus"
+)
+```
+
 ## Home Collection
@@
 sens_spec(df$Methicillin.sensitive.Staph.aureus.,
-  df$Methicillin.sensitive.Staph.aureus..1,
+  df$Methicillin.sensitive.Staph.aureus..2,
   type = "Home Collection", bug = "MSSA"
@@
 sens_spec(df$Methicillin.resistant.Staph.aureus.,
-  df$Methicillin.resistant.Staph.aureus..1,
+  df$Methicillin.resistant.Staph.aureus..2,
   type = "Home Collection", bug = "MRSA"
@@
 ```{r results='asis'}
-sens_spec(df$A..xylosoxidans., df$A..xylosoxidans..1,
+sens_spec(df$A..xylosoxidans., df$A..xylosoxidans..2,
   type = "Home Collection", bug = "A. xylosoxidans"
@@
 ```{r results='asis'}
-sens_spec(df$Achromobacter.not.xylosoxidans., df$Achromobacter.not.xylosoxidans..1,
+sens_spec(df$Achromobacter.not.xylosoxidans., df$Achromobacter.not.xylosoxidans..2,
   type = "Home Collection", bug = "Achromobacter (not xylosoxidans)"
@@
 ```{r results='asis'}
-sens_spec(df$Acinetobacter., df$Acinetobacter..1,
+sens_spec(df$Acinetobacter., df$Acinetobacter..2,
   type = "Home Collection", bug = "Acinetobacter"
@@
 ```{r results='asis'}
-sens_spec(df$H..influenzae., df$H..influenzae..1,
+sens_spec(df$H..influenzae., df$H..influenzae..2,
   type = "Home Collection", bug = "H. influenzae"
@@
 ```{r results='asis'}
-sens_spec(df$P..aeruginosa., df$P..aeruginosa..1,
+sens_spec(df$P..aeruginosa., df$P..aeruginosa..2,
   type = "Home Collection", bug = "P. aeruginosa"
@@
 ```{r results='asis'}
-sens_spec(df$P..aeruginosa..mucoid.., df$P..aeruginosa..mucoid...1,
+sens_spec(df$P..aeruginosa..mucoid.., df$P..aeruginosa..mucoid...2,
   type = "Home Collection", bug = "P. aeruginosa (mucoid)"
@@
 ```{r results='asis'}
-sens_spec(df$Pseudomonas.species..not.aeruginosa., df$Pseudomonas.species..not.aeruginosa..1,
+sens_spec(df$Pseudomonas.species..not.aeruginosa., df$Pseudomonas.species..not.aeruginosa..2,
   type = "Home Collection", bug = "Pseudomonas (not aeruginosa)"
@@
 ```{r results='asis'}
-sens_spec(df$B..cepacia.complex., df$B..cepacia.complex..1,
+sens_spec(df$B..cepacia.complex., df$B..cepacia.complex..2,
   type = "Home Collection", bug = "B. cepacia complex"
@@
 ```{r results='asis'}
-sens_spec(df$Stenotrophomonas., df$Stenotrophomonas..1,
+sens_spec(df$Stenotrophomonas., df$Stenotrophomonas..2,
   type = "Home Collection", bug = "Stenotrophomonas"
@@
 ```{r results='asis'}
-sens_spec(df$Aspergillus., df$Aspergillus..1,
+sens_spec(df$Aspergillus., df$Aspergillus..2,
   type = "Home Collection", bug = "Aspergillus"
@@
 ```{r results='asis'}
-sens_spec(df$Penicillium.species., df$Penicillium.species..1,
+sens_spec(df$Penicillium.species., df$Penicillium.species..2,
   type = "Home Collection", bug = "Penicillium species"
@@
 ```{r results='asis'}
-sens_spec(df$Sceosporium.species., df$Sceosporium.species..1,
+sens_spec(df$Sceosporium.species., df$Sceosporium.species..2,
   type = "Home Collection", bug = "Sceosporium species"
@@
 ```{r results='asis'}
-sens_spec(df$M..avium.complex., df$M..avium.complex..1,
+sens_spec(df$M..avium.complex., df$M..avium.complex..2,
   type = "Home Collection", bug = "M. avium complex"
@@
 ```{r results='asis'}
-sens_spec(df$M..abscessus., df$M..abscessus..1,
+sens_spec(df$M..abscessus., df$M..abscessus..2,
   type = "Home Collection", bug = "M. abscessus"
-)
-```
-
-## Delayed Processing
-
-```{r results='asis'}
-sens_spec(df$Methicillin.sensitive.Staph.aureus.,
-  df$Methicillin.sensitive.Staph.aureus..2,
-  type = "Delayed Processing", bug = "MSSA"
-)
-```
-
-```{r results='asis'}
-sens_spec(df$Methicillin.resistant.Staph.aureus.,
-  df$Methicillin.resistant.Staph.aureus..2,
-  type = "Delayed Processing", bug = "MRSA"
-)
-```
-
-```{r results='asis'}
-sens_spec(df$A..xylosoxidans., df$A..xylosoxidans..2,
-  type = "Delayed Processing", bug = "A. xylosoxidans"
-)
-```
-
-```{r results='asis'}
-sens_spec(df$Achromobacter.not.xylosoxidans., df$Achromobacter.not.xylosoxidans..2,
-  type = "Delayed Processing", bug = "Achromobacter (not xylosoxidans)"
-)
-```
-
-```{r results='asis'}
-sens_spec(df$Acinetobacter., df$Acinetobacter..2,
-  type = "Delayed Processing", bug = "Acinetobacter"
-)
-```
-
-```{r results='asis'}
-sens_spec(df$H..influenzae., df$H..influenzae..2,
-  type = "Delayed Processing", bug = "H. influenzae"
-)
-```
-
-```{r results='asis'}
-sens_spec(df$P..aeruginosa., df$P..aeruginosa..2,
-  type = "Delayed Processing", bug = "P. aeruginosa"
-)
-```
-
-```{r results='asis'}
-sens_spec(df$P..aeruginosa..mucoid.., df$P..aeruginosa..mucoid...2,
-  type = "Delayed Processing", bug = "P. aeruginosa (mucoid)"
-)
-```
-
-```{r results='asis'}
-sens_spec(df$Pseudomonas.species..not.aeruginosa., df$Pseudomonas.species..not.aeruginosa..2,
-  type = "Delayed Processing", bug = "Pseudomonas (not aeruginosa)"
-)
-```
-
-```{r results='asis'}
-sens_spec(df$B..cepacia.complex., df$B..cepacia.complex..2,
-  type = "Delayed Processing", bug = "B. cepacia complex"
-)
-```
-
-```{r results='asis'}
-sens_spec(df$Stenotrophomonas., df$Stenotrophomonas..2,
-  type = "Delayed Processing", bug = "Stenotrophomonas"
-)
-```
-
-```{r results='asis'}
-sens_spec(df$Aspergillus., df$Aspergillus..2,
-  type = "Delayed Processing", bug = "Aspergillus"
-)
-```
-
-```{r results='asis'}
-sens_spec(df$Penicillium.species., df$Penicillium.species..2,
-  type = "Delayed Processing", bug = "Penicillium species"
-)
-```
-
-```{r results='asis'}
-sens_spec(df$Sceosporium.species., df$Sceosporium.species..2,
-  type = "Delayed Processing", bug = "Sceosporium species"
-)
-```
-
-```{r results='asis'}
-sens_spec(df$M..avium.complex., df$M..avium.complex..2,
-  type = "Delayed Processing", bug = "M. avium complex"
-)
-```
-
-```{r results='asis'}
-sens_spec(df$M..abscessus., df$M..abscessus..2,
-  type = "Delayed Processing", bug = "M. abscessus"
 )
@@
 
+Fleiss' kappa is interpreted analagously to Cohen's kappa, but applies to more than two tests. Tables with NaN and NAs indicate that all tests had the same values (i.e. all absent or all present).
+
 ```{r}
+fleiss = function(bug,bug_name){
+  fk = kappam.fleiss(df[,c(bug,paste0(bug,".1"),paste0(bug,".2"))],detail = T)
+  t = as.data.frame.matrix(fk$detail)
+  t$p.value = format.pval(t$p.value,digits = 3,eps = 0.001)
+  cat("\n")
+    cat("##", bug_name)
+    cat("\n")
+    print(knitr::kable(t))
+    cat("\n")
+    cat("Overall Fleiss' kappa for the three tests was", round(fk$value, 3), paste0("(p = ", format.pval(fk$p.value,digits = 3, eps = 0.001),")"))
+    cat("\n")
+}
+```
 
+```{r results='asis'}
+fleiss(bug = "Methicillin.sensitive.Staph.aureus.",bug_name = "MSSA")
 ```
+
+```{r results='asis'}
+fleiss(bug = "Methicillin.resistant.Staph.aureus.",bug_name = "MRSA")
+```
+
+```{r results='asis'}
+fleiss("A..xylosoxidans.", bug_name = "A. xylosoxidans")
+```
+
+```{r results='asis'}
+fleiss("Achromobacter.not.xylosoxidans.",bug_name = "Achromobacter (not xylosoxidans)")
+```
+
+```{r results='asis'}
+fleiss("Acinetobacter.", bug_name = "Acinetobacter")
+```
+
+```{r results='asis'}
+fleiss("H..influenzae.", bug_name = "H. influenzae")
+```
+
+```{r results='asis'}
+fleiss("P..aeruginosa.", bug_name = "P. aeruginosa")
+```
+
+```{r results='asis'}
+fleiss("P..aeruginosa..mucoid..",bug_name = "P. aeruginosa (mucoid)")
+```
+
+```{r results='asis'}
+fleiss("Pseudomonas.species..not.aeruginosa.",bug_name = "Pseudomonas (not aeruginosa)")
+```
+
+```{r results='asis'}
+fleiss("B..cepacia.complex.", bug_name = "B. cepacia complex")
+```
+
+```{r results='asis'}
+fleiss("Stenotrophomonas.", bug_name = "Stenotrophomonas")
+```
+
+```{r results='asis'}
+fleiss("Aspergillus.",bug_name = "Aspergillus")
+```
+
+```{r results='asis'}
+fleiss("Penicillium.species.",bug_name = "Penicillium species")
+```
+
+```{r results='asis'}
+fleiss("Sceosporium.species.",bug_name = "Sceosporium species")
+```
+
+```{r results='asis'}
+fleiss("M..avium.complex.",bug_name = "M. avium complex")
+```
+
+```{r results='asis'}
+fleiss("M..abscessus.", bug_name = "M. abscessus")
+```

### 2023-04-03 Stacey Martiniano/ntm_agreement.qmd
Update ntm_agreement.qmd
some agreement code from Agresti

@@
 author: "Tim Vigers"
-date: "`r format(Sys.time(), '%d %B %Y')`"
+date: "today"
+date-format: long
 format:
   html:
-    code-fold: true
     toc: true
@@
     toc-float: true
+    code-fold: true
     self-contained: true
-    fig-height: 8
-    fig-width: 10
-editor: visual
+    fig-cap-location: top
+    page-layout: full
+editor: source
 ---
@@
 # Import data
-api <- read.table("~/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt",
+api <- read.table("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt",
   header = F
@@
 
+```{r}
+# Agresti examples
+agt = matrix(c(22,2,2,0,
+               5,7,14,0,
+               0,2,36,0,
+               0,1,17,10),
+             byrow = T,ncol = 4)
+dimnames(agt) = list("A"=1:4,"B"=1:4)
+# As vectors
+c1 = rep(1:4,times=diag(agt))
+c1 = c(c1,rep(1:4,times=rowSums(agt*lower.tri(agt))))
+c1 = c(c1,rep(1:4,times=rowSums(agt*upper.tri(agt))))
+c2 = rep(1:4,times=diag(agt))
+c2 = c(c2,rep(1:4,times=colSums(agt*lower.tri(agt))))
+c2 = c(c2,rep(1:4,times=colSums(agt*upper.tri(agt))))
+# Check
+table(c1,c2)
+# Data from Agresti Biometrics 1988
+agt=matrix(c(1,1,22,1,2,2,1,3,2,1,4,0,1,5,0,
+2,1,5,2,2,7,2,3,14,2,4,0,2,5,0,
+3,1,0,3,2,2,3,3,36,3,4,0,3,5,0,
+4,1,0,4,2,1,4,3,14,4,4,7,4,5,0,
+5,1,0,5,2,0,5,3,3,5,4,0,5,5,3),byrow = T,ncol = 3)
+```
+
 # All disagreements
@@
 ```
+
 # Plots

### 2023-04-04 Stacey Martiniano/ntm_agreement.qmd
Update ntm_agreement.qmd
random effects approach

@@
 custom_colors <- scale_fill_manual(name = "", values = colors)
-```
-
-```{r}
-# Agresti examples
-agt = matrix(c(22,2,2,0,
-               5,7,14,0,
-               0,2,36,0,
-               0,1,17,10),
-             byrow = T,ncol = 4)
-dimnames(agt) = list("A"=1:4,"B"=1:4)
-# As vectors
-c1 = rep(1:4,times=diag(agt))
-c1 = c(c1,rep(1:4,times=rowSums(agt*lower.tri(agt))))
-c1 = c(c1,rep(1:4,times=rowSums(agt*upper.tri(agt))))
-c2 = rep(1:4,times=diag(agt))
-c2 = c(c2,rep(1:4,times=colSums(agt*lower.tri(agt))))
-c2 = c(c2,rep(1:4,times=colSums(agt*upper.tri(agt))))
-# Check
-table(c1,c2)
-# Data from Agresti Biometrics 1988
-agt=matrix(c(1,1,22,1,2,2,1,3,2,1,4,0,1,5,0,
-2,1,5,2,2,7,2,3,14,2,4,0,2,5,0,
-3,1,0,3,2,2,3,3,36,3,4,0,3,5,0,
-4,1,0,4,2,1,4,3,14,4,4,7,4,5,0,
-5,1,0,5,2,0,5,3,3,5,4,0,5,5,3),byrow = T,ncol = 3)
 ```
@@
 Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
+
+# Bootstrap
+
+To calculate bootstrap CIs for an kappa, we generated 1000 bootstrap samples clustered by scan ID and location, and calculated Fleiss' $\kappa$ for each bootstrap sample.
+Percentile and bias-corrected and accelerated (BCa) intervals are presented below.
+
+```{r}
+library(boot)
+boot_fleiss <- function(df, i) {
+  d <- df[i,c("AF","DL","JW")] # allows boot to select sample
+  k = kappam.fleiss(d)
+  return(k$value)
+}
+```
+
+```{r}
+# Remove scans and features with no variability
+df_list = split.data.frame(data,f = ~ scan_id + feature)
+perfect_agreement = lapply(df_list,function(d){
+  if(all(d$disagree==FALSE) | all(d$disagree==TRUE)){
+    return(names(d))
+  } else {
+    return(NA)
+  }
+})
+boot_df = data
+```
+
+## Kappa
+
+### Overall
+
+```{r}
+ka <- boot(data,boot_fleiss,R=1000,strata = interaction(factor(data$scan_id),factor(data$location)))
+ci=boot.ci(ka,type = c("perc","bca"))
+```
+
+Fleiss' $\kappa$ across all features was `r round(ci$t0,3)` (percentile CI: `r paste(round(ci$percent[4],3),"-",round(ci$percent[5],3))`, BCa CI: `r paste(round(ci$bca[4],3),"-",round(ci$bca[5],3))`)
+
+## Tetra/polychoric correlation
+
+```{r}
+
+```
+
+## Random effects
+
+In this scenario we have multiple crossed random effects. Each scan contains every level of feature and region, and each level of feature contains each level of region (or vice versa).
+
+```{r}
+library(lme4)
+library(performance)
+# Long data for lmer
+re_data_no_bronch = data %>% select(-disagree) %>% 
+  filter(feature != "Bronchiectasis") %>% # Remove bronchiectasis because it's on a different scale?
+  pivot_longer(AF:JW)
+re_data_no_bronch$value = factor(as.character(re_data_no_bronch$value),levels = c("Unchecked", "Checked"))
+# Logistic model
+m = glmer(value~name+(1|scan_id)+(1|feature)+(1|location),
+         data = re_data_no_bronch,family="binomial")
+summary(m)
+# Logistic model
+m2 = glmer(value~name+(1|scan_id)+(1|feature),
+         data = re_data_no_bronch,family="binomial")
+summary(m2)
+icc(m2)
+icc(m2,by_group = T)
+```
+
+There is essentially no variability in the random effect for location (which I think is good?).

### 2023-04-04 Stacey Martiniano/ntm_agreement.qmd
Update ntm_agreement.qmd
polychoric ready to go

@@
 })
-boot_df = data
 ```
@@
 ```{r}
+library(psych)
+boot_polychor <- function(df, i) {
+  d <- df[i,c("AF","DL","JW")] # allows boot to select sample
+  d = data.matrix(d)
+  p = suppressWarnings(polychoric(d,global = F,correct = F))
+  p = p$rho[lower.tri(p$rho)]
+  names(p)=c("AF-DL","AF-JW","DL-JW")
+  return(p)
+}
+```
 
+### Overall
+
+```{r}
+pa <- boot(data,boot_polychor,R=1000,strata = interaction(factor(data$scan_id),factor(data$location)))
+ci_af_dl=boot.ci(pa,index = 1,type = c("perc","bca"))
+ci_af_jw=boot.ci(pa,index = 2,type = c("perc","bca"))
+ci_dl_jw=boot.ci(pa,index = 3,type = c("perc","bca"))
 ```

### 2023-04-04 Christine Chan/Hypoglycemia/tim_analysis_no_CFRD.Rmd
rename tim analysis

--- 2023-03-31 Christine Chan/Hypoglycemia/tim_analysis.Rmd
@@
 table_dat$CF_hypo_60[table_dat$CF == "Yes" & table_dat$hypo_60 == "No"] <- "CF/No-Hypo"
-table_dat$CF_hypo_60[table_dat$CF == "No" & table_dat$hypo_60 == "Yes"] <- "No-CF/Hypo"
-table_dat$CF_hypo_60[table_dat$CF == "No" & table_dat$hypo_60 == "No"] <- "No-CF/No-Hypo"
+table_dat$CF_hypo_60[table_dat$CF == "No" & table_dat$hypo_60 == "Yes"] <- "HC/Hypo"
+table_dat$CF_hypo_60[table_dat$CF == "No" & table_dat$hypo_60 == "No"] <- "HC/No-Hypo"
 
@@
 table_dat$CF_hypo_70[table_dat$CF == "Yes" & table_dat$hypo_70 == "No"] <- "CF/No-Hypo"
-table_dat$CF_hypo_70[table_dat$CF == "No" & table_dat$hypo_70 == "Yes"] <- "No-CF/Hypo"
-table_dat$CF_hypo_70[table_dat$CF == "No" & table_dat$hypo_70 == "No"] <- "No-CF/No-Hypo"
+table_dat$CF_hypo_70[table_dat$CF == "No" & table_dat$hypo_70 == "Yes"] <- "HC/Hypo"
+table_dat$CF_hypo_70[table_dat$CF == "No" & table_dat$hypo_70 == "No"] <- "HC/No-Hypo"
 
@@
 - OGTT Category pairwise comparison CF/Hypo60 vs CF/No-Hypo60 now significant
-- HOMA-IR pairwise comparison CF/Hypo60 vs No-CF/No-Hypo60 now significant
-- CGM Avg Glucose pairwise comparison CF/No-Hypo60 vs No-CF/Hypo60 no longer significant
-- CGM Avg Glucose pairwise comparison CF/No-Hypo60 vs No-CF/No-Hypo60 no longer significant
-- CGM MAG CF/No-Hypo60 vs No-CF/Hypo60 now significant
-- Matsuda and Homa-IR pairwise comparisons CF/Hypo70 vs No-CF/No-Hypo70 now significant
-- Delta Glucagon pairwise comparison CF/No-Hypo70 vs No-CF/Hypo70 no longer significant
-- CGM Avg Glucose pairwise comparison CF/No-Hypo70 vs No-CF/Hypo70 no longer significant
-- CGM Avg Glucose pairwise comparison CF/No-Hypo70 vs No-CF/No-Hypo70 no longer significant
+- HOMA-IR pairwise comparison CF/Hypo60 vs HC/No-Hypo60 now significant
+- CGM Avg Glucose pairwise comparison CF/No-Hypo60 vs HC/Hypo60 no longer significant
+- CGM Avg Glucose pairwise comparison CF/No-Hypo60 vs HC/No-Hypo60 no longer significant
+- CGM MAG CF/No-Hypo60 vs HC/Hypo60 now significant
+- Matsuda and Homa-IR pairwise comparisons CF/Hypo70 vs HC/No-Hypo70 now significant
+- Delta Glucagon pairwise comparison CF/No-Hypo70 vs HC/Hypo70 no longer significant
+- CGM Avg Glucose pairwise comparison CF/No-Hypo70 vs HC/Hypo70 no longer significant
+- CGM Avg Glucose pairwise comparison CF/No-Hypo70 vs HC/No-Hypo70 no longer significant
 - %time over 140 pairwise comparison CF/Hypo70 vs CF/No-Hypo no longer significant
@@
   column_spec(3, background = 
-                na.fill(ifelse(ptab$`CF/Hypo vs No-CF/Hypo` < 0.05, 
+                na.fill(ifelse(ptab$`CF/Hypo vs HC/Hypo` < 0.05, 
                                "red", "white"), "white)")) %>%
   column_spec(4, background = 
-                na.fill(ifelse(ptab$`CF/Hypo vs No-CF/No-Hypo` < 0.05, 
+                na.fill(ifelse(ptab$`CF/Hypo vs HC/No-Hypo` < 0.05, 
                                "red", "white"), "white)")) %>%
   column_spec(5, background = 
-                na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/Hypo` < 0.05, 
+                na.fill(ifelse(ptab$`CF/No-Hypo vs HC/Hypo` < 0.05, 
                                "red", "white"), "white)")) %>%
   column_spec(6, background = 
-                na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/No-Hypo` < 0.05, 
+                na.fill(ifelse(ptab$`CF/No-Hypo vs HC/No-Hypo` < 0.05, 
                                "red", "white"), "white)"))
@@
   geom_signif(
-    comparisons = list(c("CF/Hypo", "CF/No-Hypo"), c("CF/Hypo", "No-CF/No-Hypo")),
+    comparisons = list(c("CF/Hypo", "CF/No-Hypo"), c("CF/Hypo", "HC/No-Hypo")),
     color = "black",
@@
   column_spec(2, background = ifelse(ptab$`CF/Hypo vs CF/No-Hypo` < 0.05, "red", "white")) %>%
-  column_spec(3, background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/Hypo` < 0.05, "red", "white"), "white)")) %>%
-  column_spec(4, background = na.fill(ifelse(ptab$`CF/Hypo vs No-CF/No-Hypo` < 0.05, "red", "white"), "white)")) %>%
-  column_spec(5, background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/Hypo` < 0.05, "red", "white"), "white)")) %>%
-  column_spec(6, background = na.fill(ifelse(ptab$`CF/No-Hypo vs No-CF/No-Hypo` < 0.05, "red", "white"), "white)"))
+  column_spec(3, background = na.fill(ifelse(ptab$`CF/Hypo vs HC/Hypo` < 0.05, "red", "white"), "white)")) %>%
+  column_spec(4, background = na.fill(ifelse(ptab$`CF/Hypo vs HC/No-Hypo` < 0.05, "red", "white"), "white)")) %>%
+  column_spec(5, background = na.fill(ifelse(ptab$`CF/No-Hypo vs HC/Hypo` < 0.05, "red", "white"), "white)")) %>%
+  column_spec(6, background = na.fill(ifelse(ptab$`CF/No-Hypo vs HC/No-Hypo` < 0.05, "red", "white"), "white)"))
 ```
@@
 outcomes_labs <- c(
-  "Avg Glucose (Daytime)",
-  "% Time Under 60",
-  "% Time Under 70",
-  "% Time Over 140",
-  "% Time Over 180",
+  "Avg Glucose (Daytime) [mg/dL]",
+  "% Time Under 60 (mg/dL)",
+  "% Time Under 70 (mg/dL)",
+  "% Time Over 140 (mg/dL)",
+  "% Time Over 180 (mg/dL)",
   "SD",
@@
   cat("\n")
-  cat(paste("##", outcomes_labs[i]))
+  cat(paste("###", outcomes_labs[i]))
   cat("\n")
@@
 ```
+
+# Tim's additional changes for reviewer 1
+
+```{r}
+# Plot
+ggplot(data = cgm_table_dat, aes_string(y = "percent_time_under_70", x = "CF_hypo_70")) +
+  labs(y = "% Time Under 70 (mg/dL)", x = "") +
+  geom_jitter(alpha = 0.5, width = 0.1) +
+  geom_boxplot(alpha = 0.5, outlier.alpha = 0) +
+  geom_hline(yintercept=4.1, linetype="dashed")+
+  annotate("text",x=4.4, y=4.5, label="4.1 %") +
+  theme_light() +
+  ggtitle("Hypoglycemia by OGTT (< 70)") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+```{r results='asis'}
+# Comparisons
+cgm_table_dat$perc_time_70_thresh = cut(cgm_table_dat$percent_time_under_70,
+                                        breaks = c(-Inf,4.1,Inf),
+                                        labels = c("No","Yes"))
+t = tableby(CF_hypo_70~fe(perc_time_70_thresh),data = cgm_table_dat)
+summary(t,pfootnote = T,
+        labelTranslations = 
+          list(perc_time_70_thresh="Percent time under 70 mg/dL > 4.1%"))
+```
+
+```{r results='asis'}
+cgm_table_dat$CF = factor(cgm_table_dat$CF,levels = c("Yes","No"),labels = c("CF","HC"))
+t = tableby(CF~fe(perc_time_70_thresh),data = cgm_table_dat)
+summary(t,pfootnote = T,
+        labelTranslations = 
+          list(perc_time_70_thresh="Percent time under 70 mg/dL > 4.1%"))
+```

### 2023-04-06 Jordana Hoppe/nacfc_2023.qmd
Update nacfc_2023.qmd

--- 2023-03-29 Jordana Hoppe/nacfc_2023.qmd
@@
     ),
-    Race = factor(Race, levels = c("", "1", "5"), labels = c("Unknown", "American Indian or Alaska Native", "White"))
+    Race = factor(Race, levels = c("1", "5"), labels = c("American Indian or Alaska Native", "White"))
   )
 # Shorten organism names
+bugs = colnames(df)[grep("If.culture.was.positive..select.all.organisms.isolated...choice.",colnames(df))]
+bugs <- sub("If.culture.was.positive..select.all.organisms.isolated...choice.", "", bugs)
+bugs = bugs[1:16]
 colnames(df) <- sub("If.culture.was.positive..select.all.organisms.isolated...choice.", "", colnames(df))
@@
 
-# Table 1: Participant Characteristics
-
 ```{r}
 #| label: tbl-1
+#| tbl-cap: Participant Characteristics
 df %>%
@@
 
-# Sensitivity and specificity
+### Agreement between all 3 methods
+
+Fleiss' kappa is interpreted analagously to Cohen's kappa, but applies to more than two tests. Tables with NaN and NAs indicate that all tests had the same values (i.e. all absent or all present).
+
+| Cohen's Kappa | Interpretation         |
+|---------------|------------------------|
+| 0             | No agreement           |
+| 0.10 - 0.20   | Slight agreement       |
+| 0.21 - 0.40   | Fair agreement         |
+| 0.41 - 0.60   | Moderate agreement     |
+| 0.61 - 0.80   | Substantial agreement  |
+| 0.81 - 0.99   | Near perfect agreement |
+| 1             | Perfect agreement      |
 
 ```{r}
-sens_spec <- function(gold_standard, other_test, bug, type = "Home Collection") {
-  t <- table(other_test, gold_standard)
-  if (all(dim(t) == c(1, 1))) {
+fleiss = function(bug,bug_name){
+  d = df[,c(bug,paste0(bug,".1"),paste0(bug,".2"))]
+  colnames(d) = c("Clinic","Delayed","Home")
+  d[d=="Checked"] = "Present"
+  d[d=="Unchecked"] = "Absent"
+  fk = kappam.fleiss(d)
   cat("\n")
-    cat("###", bug)
+    cat("##", bug_name)
     cat("\n")
-    cat("#### Contingency table")
+    print(ftable(d))
     cat("\n")
-    cat("All samples marked as", unique(gold_standard))
+    cat("Overall Fleiss' kappa for the three tests was", round(fk$value, 3), paste0("(p = ", format.pval(fk$p.value,digits = 3, eps = 0.001),")"))
     cat("\n")
-  } else {
-    sen <- sensitivity(data = factor(other_test), reference = factor(gold_standard))
-    spec <- specificity(data = factor(other_test), reference = factor(gold_standard))
-
-    dimnames(t) <- lapply(dimnames(t), function(n) {
-      n <- sub("Checked", "Present", n)
-      n <- sub("Unchecked", "Absent", n)
-    })
-    dimnames(t)$other_test <- paste(type, dimnames(t)$other_test)
-    dimnames(t)$gold_standard <- paste("Clinic", dimnames(t)$gold_standard)
-
-    cat("\n")
-    cat("###", bug)
-    cat("\n")
-    cat("#### Contingency table")
-    cat("\n")
-    print(knitr::kable((as.data.frame.matrix(t))))
-    cat("\n")
-    cat("Sensitivity of the", tolower(type), "method was", round(sen, 3), "and specificity was", round(spec, 3))
-    cat("\n")
-  }
 }
@@
 
-## Home Collection
-
-```{r results='asis'}
-sens_spec(df$Methicillin.sensitive.Staph.aureus.,
-  df$Methicillin.sensitive.Staph.aureus..1,
-  type = "Home Collection", bug = "MSSA"
-)
+```{r}
+fleiss(bug = "Methicillin.sensitive.Staph.aureus.",bug_name = "MSSA")
 ```
 
-```{r results='asis'}
-sens_spec(df$Methicillin.resistant.Staph.aureus.,
-  df$Methicillin.resistant.Staph.aureus..1,
-  type = "Home Collection", bug = "MRSA"
-)
+```{r}
+fleiss(bug = "Methicillin.resistant.Staph.aureus.",
+       bug_name = "MRSA")
 ```
 
-```{r results='asis'}
-sens_spec(df$A..xylosoxidans., df$A..xylosoxidans..1,
-  type = "Home Collection", bug = "A. xylosoxidans"
-)
+```{r}
+fleiss("A..xylosoxidans.", bug_name = "A. xylosoxidans")
 ```
 
-```{r results='asis'}
-sens_spec(df$Achromobacter.not.xylosoxidans., df$Achromobacter.not.xylosoxidans..1,
-  type = "Home Collection", bug = "Achromobacter (not xylosoxidans)"
-)
+```{r}
+fleiss("Achromobacter.not.xylosoxidans.",
+       bug_name = "Achromobacter (not xylosoxidans)")
 ```
 
-```{r results='asis'}
-sens_spec(df$Acinetobacter., df$Acinetobacter..1,
-  type = "Home Collection", bug = "Acinetobacter"
-)
+```{r}
+fleiss("Acinetobacter.", bug_name = "Acinetobacter")
 ```
 
-```{r results='asis'}
-sens_spec(df$H..influenzae., df$H..influenzae..1,
-  type = "Home Collection", bug = "H. influenzae"
-)
+```{r}
+fleiss("H..influenzae.", bug_name = "H. influenzae")
 ```
 
-```{r results='asis'}
-sens_spec(df$P..aeruginosa., df$P..aeruginosa..1,
-  type = "Home Collection", bug = "P. aeruginosa"
-)
+```{r}
+fleiss("P..aeruginosa.", bug_name = "P. aeruginosa")
 ```
 
-```{r results='asis'}
-sens_spec(df$P..aeruginosa..mucoid.., df$P..aeruginosa..mucoid...1,
-  type = "Home Collection", bug = "P. aeruginosa (mucoid)"
-)
+```{r}
+fleiss("P..aeruginosa..mucoid..",bug_name = "P. aeruginosa (mucoid)")
 ```
 
-```{r results='asis'}
-sens_spec(df$Pseudomonas.species..not.aeruginosa., df$Pseudomonas.species..not.aeruginosa..1,
-  type = "Home Collection", bug = "Pseudomonas (not aeruginosa)"
-)
+```{r}
+fleiss("Pseudomonas.species..not.aeruginosa.",
+       bug_name = "Pseudomonas (not aeruginosa)")
 ```
 
-```{r results='asis'}
-sens_spec(df$B..cepacia.complex., df$B..cepacia.complex..1,
-  type = "Home Collection", bug = "B. cepacia complex"
-)
+```{r}
+fleiss("B..cepacia.complex.", bug_name = "B. cepacia complex")
 ```
 
-```{r results='asis'}
-sens_spec(df$Stenotrophomonas., df$Stenotrophomonas..1,
-  type = "Home Collection", bug = "Stenotrophomonas"
-)
+```{r}
+fleiss("Stenotrophomonas.", bug_name = "Stenotrophomonas")
 ```
 
-```{r results='asis'}
-sens_spec(df$Aspergillus., df$Aspergillus..1,
-  type = "Home Collection", bug = "Aspergillus"
-)
+```{r}
+fleiss("Aspergillus.",bug_name = "Aspergillus")
 ```
 
-```{r results='asis'}
-sens_spec(df$Penicillium.species., df$Penicillium.species..1,
-  type = "Home Collection", bug = "Penicillium species"
-)
+```{r}
+fleiss("Penicillium.species.",bug_name = "Penicillium species")
 ```
 
-```{r results='asis'}
-sens_spec(df$Sceosporium.species., df$Sceosporium.species..1,
-  type = "Home Collection", bug = "Sceosporium species"
-)
+```{r}
+fleiss("Sceosporium.species.",bug_name = "Sceosporium species")
 ```
 
-```{r results='asis'}
-sens_spec(df$M..avium.complex., df$M..avium.complex..1,
-  type = "Home Collection", bug = "M. avium complex"
-)
+```{r}
+fleiss("M..avium.complex.",bug_name = "M. avium complex")
 ```
 
-```{r results='asis'}
-sens_spec(df$M..abscessus., df$M..abscessus..1,
-  type = "Home Collection", bug = "M. abscessus"
-)
+```{r}
+fleiss("M..abscessus.", bug_name = "M. abscessus")
 ```
+
+### Figure
+
+```{r message=FALSE}
+f = df %>% 
+  select(Subject.ID,contains("Methicillin.sensitive"),contains("H..influenzae."))
+colnames(f) = c("ID",paste("MSSA",c("Clinic","Delayed","Home")),
+                paste("H. influenzae",c("Clinic","Delayed","Home")))
+
+counts=f %>% 
+  pivot_longer(`MSSA Clinic`:`H. influenzae Home`,
+                       names_to = c("Bug","Test"),
+                       names_pattern = "(.*) (.*)") %>%
+  filter(value=="Checked") %>%
+  group_by(Bug,ID) %>%
+  summarise(Test=paste0(Test,collapse = "+")) %>%
+  count(Test) %>%
+  mutate(prop = prop.table(n))
+
+ggplot(counts,aes(x=Bug,y=n,fill=Test,label=n))+
+  geom_col(colour = "black", position = "fill") +
+  geom_text(aes(label = n), position = position_fill(vjust = 0.5)) +
+  scale_y_continuous(labels = scales::percent) +
+  scale_fill_brewer(palette = "Pastel1")+
+  xlab("Pathogen")+ylab("% of Samples")
+
+```

### 2023-04-10 Jordana Hoppe/nacfc_2023.qmd
nicer tables Hoppe NACFC 2023

@@
 colnames(df) <- sub("If.culture.was.positive..select.all.organisms.isolated...choice.", "", colnames(df))
+bug_names=c("Methicillin sensitive Staph aureus", 
+            "Methicillin resistant Staph aureus","A. xylosoxidans", 
+            "Achromobacter (not xylosoxidans)","Acinetobacter","H. influenzae", 
+            "P. aeruginosa", "P. aeruginosa (mucoid)", 
+            "Pseudomonas species (not aeruginosa)", "B. cepacia complex",
+            "Stenotrophomonas", "Aspergillus", "Penicillium species", 
+            "Sceosporium species", "M. avium complex", "M. abscessus" )
 ```
@@
 ```{r}
-fleiss = function(bug,bug_name){
+fleiss = function(bug){
   d = df[,c(bug,paste0(bug,".1"),paste0(bug,".2"))]
@@
   cat("\n")
-    cat("##", bug_name)
+    cat("##", bug_names[which(bugs==bug)])
     cat("\n")
@@
 ```{r}
-fleiss(bug = "Methicillin.sensitive.Staph.aureus.",bug_name = "MSSA")
+fleiss(bug = "Methicillin.sensitive.Staph.aureus.")
 ```
@@
 ```{r}
-fleiss(bug = "Methicillin.resistant.Staph.aureus.",
-       bug_name = "MRSA")
+fleiss(bug = "Methicillin.resistant.Staph.aureus.")
 ```
@@
 ```{r}
-fleiss("A..xylosoxidans.", bug_name = "A. xylosoxidans")
+fleiss("A..xylosoxidans.")
 ```
@@
 ```{r}
-fleiss("Achromobacter.not.xylosoxidans.",
-       bug_name = "Achromobacter (not xylosoxidans)")
+fleiss("Achromobacter.not.xylosoxidans.")
 ```
@@
 ```{r}
-fleiss("Acinetobacter.", bug_name = "Acinetobacter")
+fleiss("Acinetobacter.")
 ```
@@
 ```{r}
-fleiss("H..influenzae.", bug_name = "H. influenzae")
+fleiss("H..influenzae.")
 ```
@@
 ```{r}
-fleiss("P..aeruginosa.", bug_name = "P. aeruginosa")
+fleiss("P..aeruginosa.")
 ```
@@
 ```{r}
-fleiss("P..aeruginosa..mucoid..",bug_name = "P. aeruginosa (mucoid)")
+fleiss("P..aeruginosa..mucoid..")
 ```
@@
 ```{r}
-fleiss("Pseudomonas.species..not.aeruginosa.",
-       bug_name = "Pseudomonas (not aeruginosa)")
+fleiss("Pseudomonas.species..not.aeruginosa.")
 ```
@@
 ```{r}
-fleiss("B..cepacia.complex.", bug_name = "B. cepacia complex")
+fleiss("B..cepacia.complex.")
 ```
@@
 ```{r}
-fleiss("Stenotrophomonas.", bug_name = "Stenotrophomonas")
+fleiss("Stenotrophomonas.")
 ```
@@
 ```{r}
-fleiss("Aspergillus.",bug_name = "Aspergillus")
+fleiss("Aspergillus.")
 ```
@@
 ```{r}
-fleiss("Penicillium.species.",bug_name = "Penicillium species")
+fleiss("Penicillium.species.")
 ```
@@
 ```{r}
-fleiss("Sceosporium.species.",bug_name = "Sceosporium species")
+fleiss("Sceosporium.species.")
 ```
@@
 ```{r}
-fleiss("M..avium.complex.",bug_name = "M. avium complex")
+fleiss("M..avium.complex.")
 ```
@@
 ```{r}
-fleiss("M..abscessus.", bug_name = "M. abscessus")
+fleiss("M..abscessus.")
 ```
@@
   xlab("Pathogen")+ylab("% of Samples")
+```
 
+# No pathogens detected
+
+Among those with no pathogens detected in clinic (n = 14), 6 had pathogens detected in delayed samples, and 3 had pathogens detected at home.
+
+```{r results='asis'}
+# For those who tested negative for any bugs in clinic, did they also test
+# negative for all at home
+f = df %>% select(Subject.ID,contains (bugs)) %>% 
+  mutate(across(-Subject.ID,
+                ~as.numeric(factor(.x,levels=c("Unchecked","Checked")))-1)) %>%
+  rowwise() %>%
+  mutate(`Clinic % Positive` = mean(c_across(all_of(bugs)))*100,
+         `Delayed % Positive` = mean(c_across(all_of(paste0(bugs,".1"))))*100,
+         `Home % Positive` = mean(c_across(all_of(paste0(bugs,".2"))))*100) %>%
+  select(Subject.ID,`Clinic % Positive`:`Home % Positive`) %>%
+  filter(`Clinic % Positive`==0)
+knitr::kable(f)
 ```
+
+```{r}
+f = f %>%
+  mutate(across(c("Clinic % Positive","Delayed % Positive","Home % Positive"),~.x==0))
+kappam.fleiss(f[,2:4])
+```

### 2023-04-10 Christine Chan/CFRD Screening Survey/cfrd_screening_analysis.Rmd
center type comparisons CFRD screening

@@
 rm(home_dir)
-source("~/GitHub/CF/get_genotype_class_severity.R")
+source("~/GitHub/CF/genotype_class_severity.R")
 ```
@@
 
-# Hypothesis Tests
+# OGTT Completion Rate
 
-## Centers with at least 50% of patients with state-funded health insurance have less than50% OGTT completion rates (CF center directors survey, questions #4 and #6).
+## Centers with at least 50% of patients with state-funded health insurance have less than 50% OGTT completion rates (CF center directors survey, questions #4 and #6).
 
@@
 
+# Center Age Group
+
+```{r}
+non_endo$age_group = 
+  factor(non_endo$What.patient.group.does.your.CF.center.take.care.of.)
+endo$age_group = 
+  factor(endo$What.age.group.of.CF.patients.do.you.take.care.of...please.select.all.that.apply.)
+levels(endo$age_group) = c("Pediatric","Both","Both","Both","Adult")
+```
+
+## CF center directors
+
+```{r results='asis'}
+t1 = tableby(age_group ~ state_funded+ogtt_completion + endocrinologist + protocol + role + location +reminders,
+             data = non_endo,control = list(cat.test = "fe"))
+summary(t1,pfootnote = T)
+```
+
 # Combined data

### 2023-04-10 Christine Chan/Hypoglycemia/tim_analysis_no_CFRD.Rmd
Update tim_analysis_no_CFRD.Rmd
hypo +/- labels

@@
 
-table_dat$CF_hypo_60[table_dat$CF == "Yes" & table_dat$hypo_60 == "Yes"] <- "CF/Hypo"
-table_dat$CF_hypo_60[table_dat$CF == "Yes" & table_dat$hypo_60 == "No"] <- "CF/No-Hypo"
-table_dat$CF_hypo_60[table_dat$CF == "No" & table_dat$hypo_60 == "Yes"] <- "HC/Hypo"
-table_dat$CF_hypo_60[table_dat$CF == "No" & table_dat$hypo_60 == "No"] <- "HC/No-Hypo"
-
-table_dat$CF_hypo_70[table_dat$CF == "Yes" & table_dat$hypo_70 == "Yes"] <- "CF/Hypo"
-table_dat$CF_hypo_70[table_dat$CF == "Yes" & table_dat$hypo_70 == "No"] <- "CF/No-Hypo"
-table_dat$CF_hypo_70[table_dat$CF == "No" & table_dat$hypo_70 == "Yes"] <- "HC/Hypo"
-table_dat$CF_hypo_70[table_dat$CF == "No" & table_dat$hypo_70 == "No"] <- "HC/No-Hypo"
+table_dat$CF_hypo_60[table_dat$CF == "Yes" & table_dat$hypo_60 == "Yes"] <- "CF/Hypo+"
+table_dat$CF_hypo_60[table_dat$CF == "Yes" & table_dat$hypo_60 == "No"] <- "CF/Hypo-"
+table_dat$CF_hypo_60[table_dat$CF == "No" & table_dat$hypo_60 == "Yes"] <- "HC/Hypo+"
+table_dat$CF_hypo_60[table_dat$CF == "No" & table_dat$hypo_60 == "No"] <- "HC/Hypo-"
+
+table_dat$CF_hypo_70[table_dat$CF == "Yes" & table_dat$hypo_70 == "Yes"] <- "CF/Hypo+"
+table_dat$CF_hypo_70[table_dat$CF == "Yes" & table_dat$hypo_70 == "No"] <- "CF/Hypo-"
+table_dat$CF_hypo_70[table_dat$CF == "No" & table_dat$hypo_70 == "Yes"] <- "HC/Hypo+"
+table_dat$CF_hypo_70[table_dat$CF == "No" & table_dat$hypo_70 == "No"] <- "HC/Hypo-"
 
@@
 - CGM Avg Glucose pairwise comparison CF/No-Hypo70 vs HC/No-Hypo70 no longer significant
-- %time over 140 pairwise comparison CF/Hypo70 vs CF/No-Hypo no longer significant
+- %time over 140 pairwise comparison CF/Hypo70 vs CF/Hypo- no longer significant
 
@@
 
-## Table 1 by CF/Hypo < 60
+## Table 1 by CF/Hypo+ < 60
 
@@
   column_spec(2, background = 
-                ifelse(ptab$`CF/Hypo vs CF/No-Hypo` < 0.05, "red", "white")) %>%
+                ifelse(ptab$`CF/Hypo- vs CF/Hypo+` < 0.05, "red", "white")) %>%
   column_spec(3, background = 
-                na.fill(ifelse(ptab$`CF/Hypo vs HC/Hypo` < 0.05, 
+                na.fill(ifelse(ptab$`CF/Hypo+ vs HC/Hypo+` < 0.05, 
                                "red", "white"), "white)")) %>%
   column_spec(4, background = 
-                na.fill(ifelse(ptab$`CF/Hypo vs HC/No-Hypo` < 0.05, 
+                na.fill(ifelse(ptab$`CF/Hypo+ vs HC/Hypo-` < 0.05, 
                                "red", "white"), "white)")) %>%
   column_spec(5, background = 
-                na.fill(ifelse(ptab$`CF/No-Hypo vs HC/Hypo` < 0.05, 
+                na.fill(ifelse(ptab$`CF/Hypo- vs HC/Hypo+` < 0.05, 
                                "red", "white"), "white)")) %>%
   column_spec(6, background = 
-                na.fill(ifelse(ptab$`CF/No-Hypo vs HC/No-Hypo` < 0.05, 
+                na.fill(ifelse(ptab$`CF/Hypo- vs HC/Hypo-` < 0.05, 
                                "red", "white"), "white)"))
@@
   geom_signif(
-    comparisons = list(c("CF/Hypo", "CF/No-Hypo"), c("CF/Hypo", "HC/No-Hypo")),
+    comparisons = list(c("CF/Hypo+", "CF/Hypo-"), c("CF/Hypo+", "HC/Hypo-")),
     color = "black",
@@
   geom_signif(
-    comparisons = list(c("CF/Hypo", "CF/No-Hypo")),
+    comparisons = list(c("CF/Hypo+", "CF/Hypo-")),
     color = "black",
@@
 
-## Table 1 by CF/Hypo < 70
+## Table 1 by CF/Hypo+ < 70
 
@@
 kbl(ptab, caption = "Pairwise p-values") %>%
-  column_spec(2, background = ifelse(ptab$`CF/Hypo vs CF/No-Hypo` < 0.05, "red", "white")) %>%
-  column_spec(3, background = na.fill(ifelse(ptab$`CF/Hypo vs HC/Hypo` < 0.05, "red", "white"), "white)")) %>%
-  column_spec(4, background = na.fill(ifelse(ptab$`CF/Hypo vs HC/No-Hypo` < 0.05, "red", "white"), "white)")) %>%
-  column_spec(5, background = na.fill(ifelse(ptab$`CF/No-Hypo vs HC/Hypo` < 0.05, "red", "white"), "white)")) %>%
-  column_spec(6, background = na.fill(ifelse(ptab$`CF/No-Hypo vs HC/No-Hypo` < 0.05, "red", "white"), "white)"))
+  column_spec(2, background = ifelse(ptab$`CF/Hypo- vs CF/Hypo+` < 0.05, "red", "white")) %>%
+  column_spec(3, background = na.fill(ifelse(ptab$`CF/Hypo+ vs HC/Hypo+` < 0.05, "red", "white"), "white)")) %>%
+  column_spec(4, background = na.fill(ifelse(ptab$`CF/Hypo+ vs HC/Hypo-` < 0.05, "red", "white"), "white)")) %>%
+  column_spec(5, background = na.fill(ifelse(ptab$`CF/Hypo- vs HC/Hypo+` < 0.05, "red", "white"), "white)")) %>%
+  column_spec(6, background = na.fill(ifelse(ptab$`CF/Hypo- vs HC/Hypo-` < 0.05, "red", "white"), "white)"))
 ```

### 2023-04-14 Scott Sagel/PEx in CF/nacfc_2023.qmd
Update nacfc_2023.qmd

@@
 )
-bugs <- c("P. aeruginosa", "P. aeruginosa mucoid", "P. aeruginosa non-mucoid", "Staph aureus", "MRSA", "S. maltophilia")
+bugs <- c("Staph aureus", "MRSA", "Haemophilus influenzae", "P.
+aeruginosa non-mucoid", "P.
+aeruginosa mucoid", "Pseudomonas not aeruginosa", "Burkholderia", "burkhodescription_other", "A.
+xylosoxidans", "Achromobacter not xylosoxidans", "S.
+maltophilia", "Inquilinus limosus", "Pandoraea species", "Other bacteria", "Aspergillus species", "Candida species", "Pencillium species", "Scedosporium species", "Mycobacterium avium complex", "Other mycobacterium", "P.
+aeruginosa")
 df[, bugs] <- lapply(df[, bugs], function(c) {
@@
 df$`S. maltophilia`[df$Site == "CNH"] <- NA
+# Write cleaned data
+write.csv(df,file = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/pex_merged.csv",row.names = F,na = "")
 # Make summary table

### 2023-04-20 Emily Muther/mental_health_screening.qmd
master data restart

--- 2023-01-05 Emily Muther/mental_health_screening.qmd
@@
 author: "Tim Vigers"
-date: last-modified
+date: "today"
+date-format: long
 format:
   html:
+    toc: true
+    toc-depth: 5
+    toc-float: true
     code-fold: true
     self-contained: true
+    fig-cap-location: top
     page-layout: full
-    title-block-banner: true
 editor: source
@@
 
-# Data Cleaning
+# Data cleaning
 
-```{python}
-import pandas as pd
-# Data import
-gad7 = pd.read_csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/GAD-7_General_Anxiety_Disorder_Full_Data_data_11-15-22.csv")
-gad7 = gad7[['PAT_MRN_ID','CONTACT_DATE','AGE AT CONTACT ','GAD-7 Total Score','GAD-7 Totals']]
-gad7.dropna(axis=0,inplace=True)
-gad7.drop_duplicates(inplace=True)
-phq9 = pd.read_csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/PHQ-9_Full_Data_data_11-11-22.csv")
-phq9 = phq9[['PAT_MRN_ID','CONTACT_DATE','AGE AT CONTACT ','PHQ-9 Total Score (calculated)','PHQ-9 Totals']]
-phq9.dropna(axis=0,inplace=True)
-phq9.drop_duplicates(inplace=True)
-psc = pd.read_csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/PSC_Full_Data_data_11-11-22.csv")
-psc = psc[['PAT_MRN_ID','CONTACT_DATE','AGE AT CONTACT ','Score','PSC Totals']]
-psc.dropna(axis=0,inplace=True)
-psc.drop_duplicates(inplace=True)
-# Merge
-df = gad7.merge(phq9,on=['PAT_MRN_ID','CONTACT_DATE','AGE AT CONTACT '],how="outer")
-df = df.merge(psc,on=['PAT_MRN_ID','CONTACT_DATE','AGE AT CONTACT '],how="outer")
-# Rename
-df.rename({'PAT_MRN_ID':'ID','CONTACT_DATE':'Date','AGE AT CONTACT ':'Age','GAD-7 Total Score':'GAD-7 Total','GAD-7 Totals':'GAD-7 Group','PHQ-9 Total Score (calculated)':'PHQ-9 Total','PHQ-9 Totals':'PHQ-9 Group','Score':'PSC-17 Total','PSC Totals':'PSC-17 Group'},inplace=True,axis=1)
+1. For each participant, the first recorded visit was designated as baseline.
+
+```{r include=FALSE}
+library(tidyverse)
+library(plotly)
+library(gtsummary)
+library(GLMMadaptive)
+# Import
+df = read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Cleaned/MH master data.csv",check.names = F,na.strings = "")
+# Column types
+df$CONTACT_DATE = lubridate::mdy(df$CONTACT_DATE)
+df$`PHQ-9 Total Score (calculated)` = as.integer(df$`PHQ-9 Total Score (calculated)`)
+# Visits
+df = df %>% arrange(MRN,CONTACT_DATE) %>% group_by(MRN) %>%
+  mutate(`Visit Number` = row_number(), 
+         `Days From Baseline` = as.numeric(CONTACT_DATE - CONTACT_DATE[1]))
 ```
 
-- On first import, there are 2,646 observations each for GAD-7, PHQ-9, and PSC-17. 
-- After deleting rows missing any any of patient ID, contact date, age at contact, total score, or score group, there were: 
-  - 559 observations for GAD-7
-  - 548 for PHQ-9
-  - 308 for PSC-17
-- After dropping duplicate rows, there are: 
-  - 390 GAD-7 observations
-  - 383 for PHQ-9
-  - 216 for PSC-17
-- After merging these together based on ID, date, and age, there are 613 unique observations with at least one of the three screening measures.
+# Plots
 
-# Profiling
+```{r warning=FALSE}
+p_phq9_hist = ggplot(df,aes(x = `PHQ-9 Total Score (calculated)`)) + 
+  geom_histogram(binwidth = 1) +
+  scale_x_continuous(breaks = 0:27) +
+  theme_bw() 
+ggplotly(p_phq9_hist)
 
-```{python}
-from pandas_profiling import ProfileReport
-profile = ProfileReport(df, title="Data Profiling Report")
-profile.to_notebook_iframe()
+p_phq9_days = ggplot(df,aes(x=`Days From Baseline`,y = `PHQ-9 Total Score (calculated)`,group=MRN)) + 
+  geom_line(aes(color=factor(MRN)),alpha=0.2) +
+  theme_bw() +
+  theme(legend.position = "none")
+ggplotly(p_phq9_days)
+
+p_phq9_visit = ggplot(df,aes(x=`Visit Number`,y = `PHQ-9 Total Score (calculated)`,group=MRN)) + 
+  geom_line(aes(color=factor(MRN)),alpha=0.2) +
+  theme_bw() +
+  theme(legend.position = "none")
+ggplotly(p_phq9_visit)
 ```
+
+# Methods
+
+Because the PHQ-9 results are right-skewed with lots of 0 scores, we used a zero inflated Poisson mixed model to analyze change in score over time.

### 2023-05-01 Emily Muther/body_image_trikafta.qmd
ZINB for Emma Lyons

@@
 library(foreign)
+library(plotly)
+library(pscl)
+library(glmnet)
 # Read in data
@@
 df[df==""]=NA
+# Categorical variables
+df$FRS_Category = relevel(df$FRS_Category,ref = "neutral")
 ```
@@
 
-# Figure
+# Mental health scores by FRS category
+
+## PHQ9
+
+### Basic model
+
+Because almost 50% of respondents scored 0 on the PHQ9 and the variance appears to be much larger than the mean, we used a zero-inflated negative binomial (ZINB) model.
 
 ```{r}
-# | label: fig-planets
-# | fig-cap: Planets
+# Plot all the zeroes
+p = ggplot(df,aes(x=PHQ9_Total)) + geom_histogram(binwidth = 1) + theme_bw()
+ggplotly(p)
+# Fit ZIP model
+zinb1 <- zeroinfl(PHQ9_Total ~ FRS_Category,data = df,dist = "negbin")
+summary(zinb1)
+```
 
+The output above suggests that there are no differences between FRS categories in terms of the likelihood of scoring a 0 on the PHQ9, or in terms of score for those with a non-zero score.
+Because there may be other variables that predict a score of 0, we used the lasso to find a model for 0 scores.
+Parameters for the lasso were chosen based on 10-fold cross validation (CV).
+
+```{r}
+# Get variables with no missing data, and at least some variance.
+x = df[,which(colSums(is.na(df))==0)]
+x = x[,-caret::nearZeroVar(x)]
+# Remove variables that wouldn't make sense and other mental health variables
+x = x %>% select(-c(ID,DOB),-contains("PHQ9"),-contains("GAD7"),
+                 -contains("FRS"),-contains("bes"),-PercentAdherence)
+preds = colnames(x)
+# Outcome
+df$PHQ9_0 = as.factor(df$PHQ9_Total==0)
+# As a formula
+f = as.formula(paste0("PHQ9_0~",paste0(preds,collapse = "+")))
+# Lasso
+xm = model.matrix(update(f,.~.-1),df)
+y = df$PHQ9_0
+cvfit = cv.glmnet(xm, y,family="binomial",nlambda=1e6)
+# Results
+coef(cvfit, s = "lambda.min")
+coef(cvfit, s = "lambda.1se")
 ```
+
+Even with CV on $10^5$ lambda values, no variables are associated with the probability of scoring a 0.
+
+# Questions
+
+1.
+For the ZINB model, should we use different predictors for the logistic part of the model (i.e., the part that predicts whether someone will have a non-zero score) based on clinical knowledge?
+For example, would you expect that certain ages would be more likely to score a 0?

### 2023-05-10 Stacey Martiniano/ntm_agreement.qmd
matched up gold standard IDs

@@
 data <- data.frame(lapply(data, as.character))
+data$scan_id = as.numeric(data$scan_id)
 # Make scorer IDs by initial
@@
 )
+# Revert duplicate IDs back to original to match with gold standard
+training = read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/training.key.csv")
+dups = read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/duplicated.key.csv")
+data$scan_id[match(dups$duplicated_ssid..dates.shifted..17.days.,data$scan_id)] = dups$original_ssid
+training$Training_id = as.numeric(sub("T00","",training$Training_id))
+data$scan_id[data$scan_id %in% training$Training_id] <- 
+  training$Cffid.Scoring.Set[match(data$scan_id, training$Training_id, nomatch = NA)]
 # Select relevant columns, convert to long data
@@
   arrange(scan_id, scorer) %>%
-  filter(scan_id > 2) %>%
+  filter(as.numeric(scan_id) <= 5) %>%
   select(scorer, scan_id, bronchiectasis_rul:atelectasis___6)
@@
 custom_colors <- scale_fill_manual(name = "", values = colors)
+# Get gold standard scores
+gold_standard = read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Raw/gold_standard_scores.xlsx")
 ```

### 2023-05-17 Stacey Martiniano/ntm_agreement.qmd
GS Kappas done

--- 2023-04-04 Stacey Martiniano/ntm_agreement.qmd
@@
 # Import data
-api <- read.table("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt",
+api <- read.table("~/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt",
                   header = F
@@
 )
-# Select relevant columns, convert to long data
-data <- data %>%
+data$scorer[data$scorer=="GG"]="GS"
+# Revert duplicate IDs back to original to match with gold standard
+training = read.csv("~/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/training.key.csv")
+dups = read.csv("~/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/duplicated.key.csv")
+training$Training_id = as.numeric(sub("T00","",training$Training_id))
+# Change duplicates to OG IDs
+old <- dups$duplicated_ssid..dates.shifted..17.days.
+new <- dups$original_ssid
+data$scan_id[data$scan_id %in% old] <- new[match(data$scan_id, old, nomatch = 0)]
+# Change training IDs to OG IDs
+old <- training$Training_id
+new = training$Cffid.Scoring.Set
+data$scan_id[data$scan_id %in% old] <- new[match(data$scan_id, old, nomatch = 0)]
+```
+
+| Cohen's Kappa | Interpretation         |
+|---------------|------------------------|
+| 0             | No agreement           |
+| 0.10 - 0.20   | Slight agreement       |
+| 0.21 - 0.40   | Fair agreement         |
+| 0.41 - 0.60   | Moderate agreement     |
+| 0.61 - 0.80   | Substantial agreement  |
+| 0.81 - 0.99   | Near perfect agreement |
+| 1             | Perfect agreement      |
+
+# Comparison to gold standards
+
+```{r}
+# Gold standard data
+gs = data %>% 
   arrange(scan_id, scorer) %>%
-  filter(scan_id > 2) %>%
-  select(scorer, scan_id, bronchiectasis_rul:atelectasis___6)
-data <- data %>%
+  filter(scan_id %in% training$Cffid.Scoring.Set) %>%
+  group_by(scorer,scan_id) %>%
+  slice_max(scoresheet_timestamp) %>%
+  select(scorer, scan_id, bronchiectasis_rul:atelectasis___6) %>%
+  ungroup()
+# Plot colors
+name <- "Accent"
+colors <- brewer.pal(6, name)
+names(colors) <- c("Checked","Unchecked","Absent","Cylindric","Varicose",NA)
+custom_colors <- scale_fill_manual(name = "", values = colors)
+# Function
+compare = function(df,scan_num){
+  id = training$Cffid.Scoring.Set[training$Training_id == scan_num]
+  df = df %>% filter(scan_id == id)
+  df = df %>% pivot_longer(bronchiectasis_rul:atelectasis___6,
+                           names_to = c("feature", "location"),
+                           names_pattern = "(.*)_(.*)$"
+  ) %>% 
+    mutate(
+      feature = gsub("__", "", feature),
+      feature = str_to_title(gsub("_", " ", feature))
+    ) 
+  df$location <- factor(df$location,
+                        levels = c("1", "2", "3", "4", "5", "6", "ling", "lll", 
+                                   "lul", "rll", "rml", "rul"),
+                        labels = c("RUL", "RML", "RLL", "LUL", "Ling", "LLL", 
+                                   "Ling", "LLL", "LUL", "RLL", "RML", "RUL")
+  )
+  df$value <- factor(df$value, levels = c("Checked", "Unchecked", "Absent", "Cylindric", "Varicose"))
+  # One column per scorer
+  df <- df %>% 
+    pivot_wider(names_from = scorer, values_from = value) %>%
+    arrange(feature,location)
+  # Kappa for the whole scan
+  fk = kappam.fleiss(df[,c("AF","DL","GS","JW")])
+  af = kappa2(df[,c("AF","GS")])
+  dl = kappa2(df[,c("DL","GS")])
+  jw = kappa2(df[,c("JW","GS")])
+  cat("\n")
+  cat("### Overall")
+  cat("\n")
+  p = df %>%
+    select(feature, location, AF:JW) %>%
+    pivot_longer(AF:JW) %>%
+    ggplot(., aes(x = name, y = location, fill = value)) +
+    geom_tile(color = "black") +
+    theme_bw() +
+    labs(fill = "", x = "Scorer", y = "Location") +
+    custom_colors +
+    facet_wrap(~feature)
+  cat("\n")
+  print(p)
+  cat("\n")
+  cat("\n")
+  cat("Overall Fleiss' Kappa for this scan was ",round(fk$value,3)," (p = ",
+      format.pval(fk$p.value,eps = 0.001,digits = 3),")",sep = "")
+  cat("\n")
+  cat("\n")
+  cat("Cohen's Kappa for AF vs. the gold standard"," was ",round(af$value,3)," (p = ",
+      format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
+  cat("\n")
+  cat("\n")
+  cat("Cohen's Kappa for DL vs. the gold standard"," was ",round(dl$value,3)," (p = ",
+      format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
+  cat("\n")
+  cat("\n")
+  cat("Cohen's Kappa for JW vs. the gold standard"," was ",round(jw$value,3)," (p = ",
+      format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")
+  cat("\n")
+  # Analyze by feature
+  cat("\n")
+  cat("### By Feature and Scorer")
+  cat("\n")
+  invisible(lapply(levels(factor(df$feature)), function(f){
+    t = data.frame(df[df$feature == f,c("AF","DL","GS","JW")])
+    rownames(t) = as.character(df$location[df$feature == f])
+    fk = kappam.fleiss(t)
+    af = kappa2(t[,c("AF","GS")])
+    dl = kappa2(t[,c("DL","GS")])
+    jw = kappa2(t[,c("JW","GS")])
+    p = df %>%
+      filter(feature == f) %>%
+      select(location, AF:JW) %>%
+      pivot_longer(AF:JW) %>%
+      ggplot(., aes(x = name, y = location, fill = value)) +
+      geom_tile(color = "black") +
+      theme_bw() +
+      labs(fill = "", x = "Scorer", y = "Location") +
+      custom_colors +
+      ggtitle(f)
+    cat("\n")
+    cat("\n")
+    cat("####",f)
+    cat("\n")
+    cat("\n")
+    print(p)
+    cat("\n")
+    cat("\n")
+    cat("Fleiss' Kappa for ",f," was ",round(fk$value,3)," (p = ",
+        format.pval(fk$p.value,eps = 0.001,digits = 3),")",sep = "")
+    cat("\n")
+    cat("\n")
+    if(is.nan(af$value)){
+      cat("\n")
+      cat("Agreement between AF and the gold standard in",f,"was perfect.")
+      cat("\n")
+    } else {
+      cat("\n")
+      cat("Cohen's Kappa for AF vs. the gold standard in ",f," was ",round(af$value,3)," (p = ",
+          format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
+      cat("\n")
+    }
+    if(is.nan(dl$value)){
+      cat("\n")
+      cat("Agreement between DL and the gold standard in",f,"was perfect.")
+      cat("\n")
+    } else {
+      cat("\n")
+      cat("Cohen's Kappa for DL vs. the gold standard in ",f," was ",round(dl$value,3)," (p = ",
+          format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
+      cat("\n")
+    }
+    if(is.nan(jw$value)){
+      cat("\n")
+      cat("Agreement between JW and the gold standard in",f,"was perfect.")
+      cat("\n")
+    } else {
+      cat("\n")
+      cat("Cohen's Kappa for JW vs. the gold standard in ",f," was ",round(jw$value,3)," (p = ",
+          format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")
+      cat("\n")
+    }
+  }))
+}
+```
+
+## Scan 1
+
+```{r results='asis',warning=FALSE}
+compare(df = gs,scan_num = 1)
+```
+
+## Scan 2
+
+```{r results='asis',warning=FALSE}
+compare(df = gs,scan_num = 2)
+```
+
+## Scan 3
+
+```{r results='asis',warning=FALSE}
+compare(df = gs,scan_num = 3)
+```
+
+## Scan 4
+
+```{r results='asis',warning=FALSE}
+compare(df = gs,scan_num = 4)
+```
+
+## Scan 5
+
+```{r results='asis',warning=FALSE}
+compare(df = gs,scan_num = 5)
+```
+
+```{r}
+# Save dataset
+write.csv(gs,"~/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/gold_standard_comp.csv",row.names = F,na="")
+```
+
+# Duplicates
+
+```{r}
+# Gold standard data
+dup_df = data %>% 
+  arrange(scan_id, scorer) %>%
+  filter(scan_id %in% dups$original_ssid) %>%
+  select(scorer, scan_id, bronchiectasis_rul:atelectasis___6) %>%
+  ungroup()
+# Function
+compare = function(df,scan_num){
+  df = df %>% filter(scan_id == scan_num) %>%
+    group_by(scorer,scan_id) %>%
+    mutate(n = row_number()) %>% ungroup() %>%
     pivot_longer(bronchiectasis_rul:atelectasis___6,
@@
     ) 
-data$location <- factor(data$location,
-  levels = c("1", "2", "3", "4", "5", "6", "ling", "lll", "lul", "rll", "rml", "rul"),
-  labels = c("RUL", "RML", "RLL", "LUL", "Ling", "LLL", "Ling", "LLL", "LUL", "RLL", "RML", "RUL")
+  df$location <- factor(df$location,
+                        levels = c("1", "2", "3", "4", "5", "6", "ling", "lll", 
+                                   "lul", "rll", "rml", "rul"),
+                        labels = c("RUL", "RML", "RLL", "LUL", "Ling", "LLL", 
+                                   "Ling", "LLL", "LUL", "RLL", "RML", "RUL")
   )
-data$value <- factor(data$value, levels = c("Checked", "Unchecked", "Absent", "Cylindric", "Varicose"))
-vals <- levels(data$value)
+  df$value <- factor(df$value, levels = c("Checked", "Unchecked", "Absent", "Cylindric", "Varicose"))
   # One column per scorer
-data <- data %>% pivot_wider(names_from = scorer, values_from = value)
-# Find disagreements
-data$disagree <- !(data$AF == data$DL & data$DL == data$JW)
-# Set colors for whole document
-n <- length(vals)
-name <- "Accent"
-colors <- brewer.pal(n, name)
-names(colors) <- vals
-custom_colors <- scale_fill_manual(name = "", values = colors)
-```
-
-# All disagreements
-
-The table below is sortable and searchable, and includes only the features and locations where there was some disagreement.
-
-```{r}
-# Table
-data %>%
-  filter(disagree == T) %>%
-  select(-disagree) %>%
-  datatable(., rownames = F)
-```
-
-# Plots
-
-## All scans
-
-```{r}
-data %>%
-  pivot_longer(AF:JW) %>%
-  ggplot(., aes(x = name, y = location, fill = value)) +
-  geom_tile(color = "black") +
-  theme_bw() +
-  labs(fill = "", x = "Scorer", y = "Location") +
-  custom_colors +
-  facet_grid(rows = vars(scan_id), cols = vars(feature))
-```
-
-## Scan 3
-
-```{r}
-scan <- 3
-data %>%
-  filter(scan_id == scan) %>%
+  df <- df %>% 
+    pivot_wider(names_from = scorer, values_from = value) %>%
+    arrange(feature,n,location)
+  # Kappa for the whole scan
+  fk = kappam.fleiss(df[,c("AF","DL","GS","JW")])
+  af = kappa2(df[,c("AF","GS")])
+  dl = kappa2(df[,c("DL","GS")])
+  jw = kappa2(df[,c("JW","GS")])
+  cat("\n")
+  cat("### Overall")
+  cat("\n")
+  p = df %>%
     select(feature, location, AF:JW) %>%
@@
     facet_wrap(~feature)
-```
-
-## Scan 4
-
-```{r}
-scan <- 4
-data %>%
-  filter(scan_id == scan) %>%
-  select(feature, location, AF:JW) %>%
+  cat("\n")
+  print(p)
+  cat("\n")
+  cat("\n")
+  cat("Overall Fleiss' Kappa for this scan was ",round(fk$value,3)," (p = ",
+      format.pval(fk$p.value,eps = 0.001,digits = 3),")",sep = "")
+  cat("\n")
+  cat("\n")
+  cat("Cohen's Kappa for AF vs. the gold standard"," was ",round(af$value,3)," (p = ",
+      format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
+  cat("\n")
+  cat("\n")
+  cat("Cohen's Kappa for DL vs. the gold standard"," was ",round(dl$value,3)," (p = ",
+      format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
+  cat("\n")
+  cat("\n")
+  cat("Cohen's Kappa for JW vs. the gold standard"," was ",round(jw$value,3)," (p = ",
+      format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")
+  cat("\n")
+  # Analyze by feature
+  cat("\n")
+  cat("### By Feature and Scorer")
+  cat("\n")
+  invisible(lapply(levels(factor(df$feature)), function(f){
+    t = data.frame(df[df$feature == f,c("AF","DL","GS","JW")])
+    rownames(t) = as.character(df$location[df$feature == f])
+    fk = kappam.fleiss(t)
+    af = kappa2(t[,c("AF","GS")])
+    dl = kappa2(t[,c("DL","GS")])
+    jw = kappa2(t[,c("JW","GS")])
+    p = df %>%
+      filter(feature == f) %>%
+      select(location, AF:JW) %>%
       pivot_longer(AF:JW) %>%
@@
       custom_colors +
-  facet_wrap(~feature)
-```
-
-## Scan 5
-
-```{r}
-scan <- 5
-data %>%
-  filter(scan_id == scan) %>%
-  select(feature, location, AF:JW) %>%
-  pivot_longer(AF:JW) %>%
-  ggplot(., aes(x = name, y = location, fill = value)) +
-  geom_tile(color = "black") +
-  theme_bw() +
-  labs(fill = "", x = "Scorer", y = "Location") +
-  custom_colors +
-  facet_wrap(~feature)
-```
-
-# Agreement by Feature
-
-## Atelectasis
-
-```{r}
-feat <- "Atelectasis"
-# Statistic
-k <- data %>%
-  filter(feature == feat) %>%
-  select(AF:JW) %>%
-  kappam.fleiss(.)
-ka <- round(k$value, 3)
-p <- format.pval(k$p.value, eps = 0.001, digits = 3)
-# Plot
-data %>%
-  filter(feature == feat) %>%
-  select(scan_id, location, AF:JW) %>%
-  pivot_longer(AF:JW) %>%
-  ggplot(., aes(x = name, y = location, fill = value)) +
-  geom_tile(color = "black") +
-  theme_bw() +
-  labs(fill = "", x = "Scorer", y = "Location") +
-  custom_colors +
-  facet_wrap(~scan_id)
-```
-
-Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
-
-## Bronchiectasis
-
-```{r}
-feat <- "Bronchiectasis"
-# Statistic
-k <- data %>%
-  filter(feature == feat) %>%
-  select(AF:JW) %>%
-  kappam.fleiss(.)
-ka <- round(k$value, 3)
-p <- format.pval(k$p.value, eps = 0.001, digits = 3)
-# Plot
-data %>%
-  filter(feature == feat) %>%
-  select(scan_id, location, AF:JW) %>%
-  pivot_longer(AF:JW) %>%
-  ggplot(., aes(x = name, y = location, fill = value)) +
-  geom_tile(color = "black") +
-  theme_bw() +
-  labs(fill = "", x = "Scorer", y = "Location") +
-  custom_colors +
-  facet_wrap(~scan_id)
-```
-
-Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
-
-## Cavity
-
-```{r}
-feat <- "Cavity"
-# Statistic
-k <- data %>%
-  filter(feature == feat) %>%
-  select(AF:JW) %>%
-  kappam.fleiss(.)
-ka <- round(k$value, 3)
-p <- format.pval(k$p.value, eps = 0.001, digits = 3)
-# Plot
-data %>%
-  filter(feature == feat) %>%
-  select(scan_id, location, AF:JW) %>%
-  pivot_longer(AF:JW) %>%
-  ggplot(., aes(x = name, y = location, fill = value)) +
-  geom_tile(color = "black") +
-  theme_bw() +
-  labs(fill = "", x = "Scorer", y = "Location") +
-  custom_colors +
-  facet_wrap(~scan_id)
-```
-
-Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
-
-## Consolidation
-
-```{r}
-feat <- "Consolidation"
-# Statistic
-k <- data %>%
-  filter(feature == feat) %>%
-  select(AF:JW) %>%
-  kappam.fleiss(.)
-ka <- round(k$value, 3)
-p <- format.pval(k$p.value, eps = 0.001, digits = 3)
-# Plot
-data %>%
-  filter(feature == feat) %>%
-  select(scan_id, location, AF:JW) %>%
-  pivot_longer(AF:JW) %>%
-  ggplot(., aes(x = name, y = location, fill = value)) +
-  geom_tile(color = "black") +
-  theme_bw() +
-  labs(fill = "", x = "Scorer", y = "Location") +
-  custom_colors +
-  facet_wrap(~scan_id)
-```
-
-Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
-
-## Cyst
-
-```{r}
-feat <- "Cyst"
-# Statistic
-k <- data %>%
-  filter(feature == feat) %>%
-  select(AF:JW) %>%
-  kappam.fleiss(.)
-ka <- round(k$value, 3)
-p <- format.pval(k$p.value, eps = 0.001, digits = 3)
-# Plot
-data %>%
-  filter(feature == feat) %>%
-  select(scan_id, location, AF:JW) %>%
-  pivot_longer(AF:JW) %>%
-  ggplot(., aes(x = name, y = location, fill = value)) +
-  geom_tile(color = "black") +
-  theme_bw() +
-  labs(fill = "", x = "Scorer", y = "Location") +
-  custom_colors +
-  facet_wrap(~scan_id)
-```
-
-Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
-
-## Ground Glass
-
-```{r}
-feat <- "Ground Glass"
-# Statistic
-k <- data %>%
-  filter(feature == feat) %>%
-  select(AF:JW) %>%
-  kappam.fleiss(.)
-ka <- round(k$value, 3)
-p <- format.pval(k$p.value, eps = 0.001, digits = 3)
-# Plot
-data %>%
-  filter(feature == feat) %>%
-  select(scan_id, location, AF:JW) %>%
-  pivot_longer(AF:JW) %>%
-  ggplot(., aes(x = name, y = location, fill = value)) +
-  geom_tile(color = "black") +
-  theme_bw() +
-  labs(fill = "", x = "Scorer", y = "Location") +
-  custom_colors +
-  facet_wrap(~scan_id)
-```
-
-Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
-
-## Mucoid Impaction
-
-```{r}
-feat <- "Mucoid Impaction"
-# Statistic
-k <- data %>%
-  filter(feature == feat) %>%
-  select(AF:JW) %>%
-  kappam.fleiss(.)
-ka <- round(k$value, 3)
-p <- format.pval(k$p.value, eps = 0.001, digits = 3)
-# Plot
-data %>%
-  filter(feature == feat) %>%
-  select(scan_id, location, AF:JW) %>%
-  pivot_longer(AF:JW) %>%
-  ggplot(., aes(x = name, y = location, fill = value)) +
-  geom_tile(color = "black") +
-  theme_bw() +
-  labs(fill = "", x = "Scorer", y = "Location") +
-  custom_colors +
-  facet_wrap(~scan_id)
-```
-
-Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
-
-## Nodule
-
-```{r}
-feat <- "Nodule"
-# Statistic
-k <- data %>%
-  filter(feature == feat) %>%
-  select(AF:JW) %>%
-  kappam.fleiss(.)
-ka <- round(k$value, 3)
-p <- format.pval(k$p.value, eps = 0.001, digits = 3)
-# Plot
-data %>%
-  filter(feature == feat) %>%
-  select(scan_id, location, AF:JW) %>%
-  pivot_longer(AF:JW) %>%
-  ggplot(., aes(x = name, y = location, fill = value)) +
-  geom_tile(color = "black") +
-  theme_bw() +
-  labs(fill = "", x = "Scorer", y = "Location") +
-  custom_colors +
-  facet_wrap(~scan_id)
-```
-
-Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
-
-## Tree in Bud
-
-```{r}
-feat <- "Tree In Bud"
-# Statistics
-k <- data %>%
-  filter(feature == feat) %>%
-  select(AF:JW) %>%
-  kappam.fleiss(.)
-ka <- round(k$value, 3)
-p <- format.pval(k$p.value, eps = 0.001, digits = 3)
-# Plot
-data %>%
-  filter(feature == feat) %>%
-  select(scan_id, location, AF:JW) %>%
-  pivot_longer(AF:JW) %>%
-  ggplot(., aes(x = name, y = location, fill = value)) +
-  geom_tile(color = "black") +
-  theme_bw() +
-  labs(fill = "", x = "Scorer", y = "Location") +
-  custom_colors +
-  facet_wrap(~scan_id)
-```
-
-Fleiss' Kappa for multiple raters was `r ka` (p: `r p`) across all scans.
-
-# Bootstrap
-
-To calculate bootstrap CIs for an kappa, we generated 1000 bootstrap samples clustered by scan ID and location, and calculated Fleiss' $\kappa$ for each bootstrap sample.
-Percentile and bias-corrected and accelerated (BCa) intervals are presented below.
-
-```{r}
-library(boot)
-boot_fleiss <- function(df, i) {
-  d <- df[i,c("AF","DL","JW")] # allows boot to select sample
-  k = kappam.fleiss(d)
-  return(k$value)
+      ggtitle(f)
+    cat("\n")
+    cat("\n")
+    cat("####",f)
+    cat("\n")
+    cat("\n")
+    print(p)
+    cat("\n")
+    cat("\n")
+    cat("Fleiss' Kappa for ",f," was ",round(fk$value,3)," (p = ",
+        format.pval(fk$p.value,eps = 0.001,digits = 3),")",sep = "")
+    cat("\n")
+    cat("\n")
+    if(is.nan(af$value)){
+      cat("\n")
+      cat("Agreement between AF and the gold standard in",f,"was perfect.")
+      cat("\n")
+    } else {
+      cat("\n")
+      cat("Cohen's Kappa for AF vs. the gold standard in ",f," was ",round(af$value,3)," (p = ",
+          format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
+      cat("\n")
+    }
+    if(is.nan(dl$value)){
+      cat("\n")
+      cat("Agreement between DL and the gold standard in",f,"was perfect.")
+      cat("\n")
+    } else {
+      cat("\n")
+      cat("Cohen's Kappa for DL vs. the gold standard in ",f," was ",round(dl$value,3)," (p = ",
+          format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
+      cat("\n")
+    }
+    if(is.nan(jw$value)){
+      cat("\n")
+      cat("Agreement between JW and the gold standard in",f,"was perfect.")
+      cat("\n")
+    } else {
+      cat("\n")
+      cat("Cohen's Kappa for JW vs. the gold standard in ",f," was ",round(jw$value,3)," (p = ",
+          format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")
+      cat("\n")
+    }
+  }))
 }
 ```
-
-```{r}
-# Remove scans and features with no variability
-df_list = split.data.frame(data,f = ~ scan_id + feature)
-perfect_agreement = lapply(df_list,function(d){
-  if(all(d$disagree==FALSE) | all(d$disagree==TRUE)){
-    return(names(d))
-  } else {
-    return(NA)
-  }
-})
-boot_df = data
-```
-
-## Kappa
-
-### Overall
-
-```{r}
-ka <- boot(data,boot_fleiss,R=1000,strata = interaction(factor(data$scan_id),factor(data$location)))
-ci=boot.ci(ka,type = c("perc","bca"))
-```
-
-Fleiss' $\kappa$ across all features was `r round(ci$t0,3)` (percentile CI: `r paste(round(ci$percent[4],3),"-",round(ci$percent[5],3))`, BCa CI: `r paste(round(ci$bca[4],3),"-",round(ci$bca[5],3))`)
-
-## Tetra/polychoric correlation
-
-```{r}
-
-```
-
-## Random effects
-
-In this scenario we have multiple crossed random effects. Each scan contains every level of feature and region, and each level of feature contains each level of region (or vice versa).
-
-```{r}
-library(lme4)
-library(performance)
-# Long data for lmer
-re_data_no_bronch = data %>% select(-disagree) %>% 
-  filter(feature != "Bronchiectasis") %>% # Remove bronchiectasis because it's on a different scale?
-  pivot_longer(AF:JW)
-re_data_no_bronch$value = factor(as.character(re_data_no_bronch$value),levels = c("Unchecked", "Checked"))
-# Logistic model
-m = glmer(value~name+(1|scan_id)+(1|feature)+(1|location),
-         data = re_data_no_bronch,family="binomial")
-summary(m)
-# Logistic model
-m2 = glmer(value~name+(1|scan_id)+(1|feature),
-         data = re_data_no_bronch,family="binomial")
-summary(m2)
-icc(m2)
-icc(m2,by_group = T)
-```
-
-There is essentially no variability in the random effect for location (which I think is good?).

### 2023-05-17 Stacey Martiniano/ntm_agreement.qmd
Update ntm_agreement.qmd

@@
     toc-float: true
-    code-fold: true
+    code-fold: false
     self-contained: true
@@
 library(tidyverse)
-library(irr)
 library(redcapAPI)
-library(Hmisc)
-library(DT)
-library(RColorBrewer)
+library(lmerTest)
+library(performance)
+library(broom.mixed)
+library(knitr)
 # Import data
@@
 
-| Cohen's Kappa | Interpretation         |
-|---------------|------------------------|
-| 0             | No agreement           |
-| 0.10 - 0.20   | Slight agreement       |
-| 0.21 - 0.40   | Fair agreement         |
-| 0.41 - 0.60   | Moderate agreement     |
-| 0.61 - 0.80   | Substantial agreement  |
-| 0.81 - 0.99   | Near perfect agreement |
-| 1             | Perfect agreement      |
+# Gold standards
 
-# Comparison to gold standards
+## Excluding bronchiectasis
 
 ```{r}
+#| echo: false
 # Gold standard data
@@
   select(scorer, scan_id, bronchiectasis_rul:atelectasis___6) %>%
-  ungroup()
-# Plot colors
-name <- "Accent"
-colors <- brewer.pal(6, name)
-names(colors) <- c("Checked","Unchecked","Absent","Cylindric","Varicose",NA)
-custom_colors <- scale_fill_manual(name = "", values = colors)
-# Function
-compare = function(df,scan_num){
-  id = training$Cffid.Scoring.Set[training$Training_id == scan_num]
-  df = df %>% filter(scan_id == id)
-  df = df %>% pivot_longer(bronchiectasis_rul:atelectasis___6,
+  ungroup() %>% pivot_longer(bronchiectasis_rul:atelectasis___6,
                              names_to = c("feature", "location"),
@@
   ) 
-  df$location <- factor(df$location,
+gs$location <- factor(gs$location,
                       levels = c("1", "2", "3", "4", "5", "6", "ling", "lll", 
@@
 )
-  df$value <- factor(df$value, levels = c("Checked", "Unchecked", "Absent", "Cylindric", "Varicose"))
-  # One column per scorer
-  df <- df %>% 
-    pivot_wider(names_from = scorer, values_from = value) %>%
-    arrange(feature,location)
-  # Kappa for the whole scan
-  fk = kappam.fleiss(df[,c("AF","DL","GS","JW")])
-  af = kappa2(df[,c("AF","GS")])
-  dl = kappa2(df[,c("DL","GS")])
-  jw = kappa2(df[,c("JW","GS")])
-  cat("\n")
-  cat("### Overall")
-  cat("\n")
-  p = df %>%
-    select(feature, location, AF:JW) %>%
-    pivot_longer(AF:JW) %>%
-    ggplot(., aes(x = name, y = location, fill = value)) +
+# Exclude bronchiectasis
+gs = gs %>% filter(feature != "Bronchiectasis")
+gs$value = factor(gs$value,levels = c("Unchecked","Checked"),
+                  labels = c("Absent","Present"))
+# GS as reference group
+gs$scorer = factor(gs$scorer,levels = c("GS","AF","DL","JW"))
+gs$value = factor(gs$value)
+# Plot
+gs %>%
+  ggplot(., aes(x = scorer, y = location, fill = value)) +
   geom_tile(color = "black") +
@@
   labs(fill = "", x = "Scorer", y = "Location") +
-    custom_colors +
-    facet_wrap(~feature)
+  facet_grid(rows = vars(scan_id),cols = vars(feature))
+```
+
+### Comparison between raters
+
+```{r}
+# Mixed model
+mod1 <- glmer(value ~ scorer+(1|scan_id)+(1|feature)+(1|location), 
+              data=gs,family = "binomial")
+res = tidy(mod1,exponentiate = T,effects="fixed")
+kable(res %>% select(-effect),digits = 3)
+```
+
+Across all training scans, JW was significantly different from the gold standard ("Intercept"). Adjusted ICC was `r round(icc(mod1)$ICC_adjusted,3)`, indicating moderate agreement between raters.
+
+### Comparison within features
+
+```{r results='asis'}
+invisible(lapply(unique(gs$feature), function(f){
   cat("\n")
-  print(p)
+  cat("###",f)
   cat("\n")
   cat("\n")
-  cat("Overall Fleiss' Kappa for this scan was ",round(fk$value,3)," (p = ",
-      format.pval(fk$p.value,eps = 0.001,digits = 3),")",sep = "")
+  mod <- tryCatch(glmer(value ~ scorer+(1|scan_id)+(1|location), 
+                        data=gs[gs$feature==f,],family = "binomial"),
+                  error=function(err) NULL,warning=function(war) NULL)
+  if(!is.null(mod)){
+    res = tidy(mod,exponentiate = T,effects="fixed")
+    print(kable(res %>% select(-effect),digits = 3))
     cat("\n")
-  cat("\n")
-  cat("Cohen's Kappa for AF vs. the gold standard"," was ",round(af$value,3)," (p = ",
-      format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
-  cat("\n")
-  cat("\n")
-  cat("Cohen's Kappa for DL vs. the gold standard"," was ",round(dl$value,3)," (p = ",
-      format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
-  cat("\n")
-  cat("\n")
-  cat("Cohen's Kappa for JW vs. the gold standard"," was ",round(jw$value,3)," (p = ",
-      format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")
-  cat("\n")
-  # Analyze by feature
-  cat("\n")
-  cat("### By Feature and Scorer")
-  cat("\n")
-  invisible(lapply(levels(factor(df$feature)), function(f){
-    t = data.frame(df[df$feature == f,c("AF","DL","GS","JW")])
-    rownames(t) = as.character(df$location[df$feature == f])
-    fk = kappam.fleiss(t)
-    af = kappa2(t[,c("AF","GS")])
-    dl = kappa2(t[,c("DL","GS")])
-    jw = kappa2(t[,c("JW","GS")])
-    p = df %>%
-      filter(feature == f) %>%
-      select(location, AF:JW) %>%
-      pivot_longer(AF:JW) %>%
-      ggplot(., aes(x = name, y = location, fill = value)) +
-      geom_tile(color = "black") +
-      theme_bw() +
-      labs(fill = "", x = "Scorer", y = "Location") +
-      custom_colors +
-      ggtitle(f)
-    cat("\n")
-    cat("\n")
-    cat("####",f)
-    cat("\n")
-    cat("\n")
-    print(p)
-    cat("\n")
-    cat("\n")
-    cat("Fleiss' Kappa for ",f," was ",round(fk$value,3)," (p = ",
-        format.pval(fk$p.value,eps = 0.001,digits = 3),")",sep = "")
-    cat("\n")
-    cat("\n")
-    if(is.nan(af$value)){
-      cat("\n")
-      cat("Agreement between AF and the gold standard in",f,"was perfect.")
+    cat("Adjusted ICC was",round(icc(mod)$ICC_adjusted,3))
     cat("\n")
@@
     cat("\n")
-      cat("Cohen's Kappa for AF vs. the gold standard in ",f," was ",round(af$value,3)," (p = ",
-          format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
-      cat("\n")
-    }
-    if(is.nan(dl$value)){
-      cat("\n")
-      cat("Agreement between DL and the gold standard in",f,"was perfect.")
-      cat("\n")
-    } else {
-      cat("\n")
-      cat("Cohen's Kappa for DL vs. the gold standard in ",f," was ",round(dl$value,3)," (p = ",
-          format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
-      cat("\n")
-    }
-    if(is.nan(jw$value)){
-      cat("\n")
-      cat("Agreement between JW and the gold standard in",f,"was perfect.")
-      cat("\n")
-    } else {
-      cat("\n")
-      cat("Cohen's Kappa for JW vs. the gold standard in ",f," was ",round(jw$value,3)," (p = ",
-          format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")
+    cat("Model did not converge (likely due to low variability).")
     cat("\n")
@@
 }))
-}
 ```
-
-## Scan 1
-
-```{r results='asis',warning=FALSE}
-compare(df = gs,scan_num = 1)
-```
-
-## Scan 2
-
-```{r results='asis',warning=FALSE}
-compare(df = gs,scan_num = 2)
-```
-
-## Scan 3
-
-```{r results='asis',warning=FALSE}
-compare(df = gs,scan_num = 3)
-```
-
-## Scan 4
-
-```{r results='asis',warning=FALSE}
-compare(df = gs,scan_num = 4)
-```
-
-## Scan 5
-
-```{r results='asis',warning=FALSE}
-compare(df = gs,scan_num = 5)
-```
-
-```{r}
-# Save dataset
-write.csv(gs,"~/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/gold_standard_comp.csv",row.names = F,na="")
-```
-
-# Duplicates
-
-```{r}
-# Gold standard data
-dup_df = data %>% 
-  arrange(scan_id, scorer) %>%
-  filter(scan_id %in% dups$original_ssid) %>%
-  select(scorer, scan_id, bronchiectasis_rul:atelectasis___6) %>%
-  ungroup()
-# Function
-compare = function(df,scan_num){
-  df = df %>% filter(scan_id == scan_num) %>%
-    group_by(scorer,scan_id) %>%
-    mutate(n = row_number()) %>% ungroup() %>%
-    pivot_longer(bronchiectasis_rul:atelectasis___6,
-                           names_to = c("feature", "location"),
-                           names_pattern = "(.*)_(.*)$"
-  ) %>% 
-    mutate(
-      feature = gsub("__", "", feature),
-      feature = str_to_title(gsub("_", " ", feature))
-    ) 
-  df$location <- factor(df$location,
-                        levels = c("1", "2", "3", "4", "5", "6", "ling", "lll", 
-                                   "lul", "rll", "rml", "rul"),
-                        labels = c("RUL", "RML", "RLL", "LUL", "Ling", "LLL", 
-                                   "Ling", "LLL", "LUL", "RLL", "RML", "RUL")
-  )
-  df$value <- factor(df$value, levels = c("Checked", "Unchecked", "Absent", "Cylindric", "Varicose"))
-  # One column per scorer
-  df <- df %>% 
-    pivot_wider(names_from = scorer, values_from = value) %>%
-    arrange(feature,n,location)
-  # Kappa for the whole scan
-  fk = kappam.fleiss(df[,c("AF","DL","GS","JW")])
-  af = kappa2(df[,c("AF","GS")])
-  dl = kappa2(df[,c("DL","GS")])
-  jw = kappa2(df[,c("JW","GS")])
-  cat("\n")
-  cat("### Overall")
-  cat("\n")
-  p = df %>%
-    select(feature, location, AF:JW) %>%
-    pivot_longer(AF:JW) %>%
-    ggplot(., aes(x = name, y = location, fill = value)) +
-    geom_tile(color = "black") +
-    theme_bw() +
-    labs(fill = "", x = "Scorer", y = "Location") +
-    custom_colors +
-    facet_wrap(~feature)
-  cat("\n")
-  print(p)
-  cat("\n")
-  cat("\n")
-  cat("Overall Fleiss' Kappa for this scan was ",round(fk$value,3)," (p = ",
-      format.pval(fk$p.value,eps = 0.001,digits = 3),")",sep = "")
-  cat("\n")
-  cat("\n")
-  cat("Cohen's Kappa for AF vs. the gold standard"," was ",round(af$value,3)," (p = ",
-      format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
-  cat("\n")
-  cat("\n")
-  cat("Cohen's Kappa for DL vs. the gold standard"," was ",round(dl$value,3)," (p = ",
-      format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
-  cat("\n")
-  cat("\n")
-  cat("Cohen's Kappa for JW vs. the gold standard"," was ",round(jw$value,3)," (p = ",
-      format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")
-  cat("\n")
-  # Analyze by feature
-  cat("\n")
-  cat("### By Feature and Scorer")
-  cat("\n")
-  invisible(lapply(levels(factor(df$feature)), function(f){
-    t = data.frame(df[df$feature == f,c("AF","DL","GS","JW")])
-    rownames(t) = as.character(df$location[df$feature == f])
-    fk = kappam.fleiss(t)
-    af = kappa2(t[,c("AF","GS")])
-    dl = kappa2(t[,c("DL","GS")])
-    jw = kappa2(t[,c("JW","GS")])
-    p = df %>%
-      filter(feature == f) %>%
-      select(location, AF:JW) %>%
-      pivot_longer(AF:JW) %>%
-      ggplot(., aes(x = name, y = location, fill = value)) +
-      geom_tile(color = "black") +
-      theme_bw() +
-      labs(fill = "", x = "Scorer", y = "Location") +
-      custom_colors +
-      ggtitle(f)
-    cat("\n")
-    cat("\n")
-    cat("####",f)
-    cat("\n")
-    cat("\n")
-    print(p)
-    cat("\n")
-    cat("\n")
-    cat("Fleiss' Kappa for ",f," was ",round(fk$value,3)," (p = ",
-        format.pval(fk$p.value,eps = 0.001,digits = 3),")",sep = "")
-    cat("\n")
-    cat("\n")
-    if(is.nan(af$value)){
-      cat("\n")
-      cat("Agreement between AF and the gold standard in",f,"was perfect.")
-      cat("\n")
-    } else {
-      cat("\n")
-      cat("Cohen's Kappa for AF vs. the gold standard in ",f," was ",round(af$value,3)," (p = ",
-          format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
-      cat("\n")
-    }
-    if(is.nan(dl$value)){
-      cat("\n")
-      cat("Agreement between DL and the gold standard in",f,"was perfect.")
-      cat("\n")
-    } else {
-      cat("\n")
-      cat("Cohen's Kappa for DL vs. the gold standard in ",f," was ",round(dl$value,3)," (p = ",
-          format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
-      cat("\n")
-    }
-    if(is.nan(jw$value)){
-      cat("\n")
-      cat("Agreement between JW and the gold standard in",f,"was perfect.")
-      cat("\n")
-    } else {
-      cat("\n")
-      cat("Cohen's Kappa for JW vs. the gold standard in ",f," was ",round(jw$value,3)," (p = ",
-          format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")
-      cat("\n")
-    }
-  }))
-}
-```

### 2023-05-17 Stacey Martiniano/ntm_agreement.qmd
Update ntm_agreement.qmd
NTM duplicates done

--- 2023-05-17 Stacey Martiniano/ntm_agreement.qmd
@@
       labs(fill = "", x = "Scorer", y = "Location") +
-      custom_colors +
-      ggtitle(f)
+      custom_colors
     cat("\n")
@@
     pivot_wider(names_from = scorer, values_from = value) %>%
-    arrange(feature,n,location)
+    pivot_wider(names_from = n, values_from = AF:JW) %>%
+    arrange(feature,location)
   # Kappa for the whole scan
-  fk = kappam.fleiss(df[,c("AF","DL","GS","JW")])
-  af = kappa2(df[,c("AF","GS")])
-  dl = kappa2(df[,c("DL","GS")])
-  jw = kappa2(df[,c("JW","GS")])
+  af = kappa2(df[,c("AF_1","AF_2")])
+  dl = kappa2(df[,c("DL_1","DL_2")])
+  jw = kappa2(df[,c("JW_1","JW_2")])
   cat("\n")
@@
   p = df %>%
-    select(feature, location, AF:JW) %>%
-    pivot_longer(AF:JW) %>%
+    select(feature, location, AF_1:JW_2) %>%
+    pivot_longer(AF_1:JW_2) %>%
     ggplot(., aes(x = name, y = location, fill = value)) +
@@
   cat("\n")
-  cat("Overall Fleiss' Kappa for this scan was ",round(fk$value,3)," (p = ",
-      format.pval(fk$p.value,eps = 0.001,digits = 3),")",sep = "")
-  cat("\n")
-  cat("\n")
-  cat("Cohen's Kappa for AF vs. the gold standard"," was ",round(af$value,3)," (p = ",
+  cat("Cohen's Kappa for AF"," was ",round(af$value,3)," (p = ",
       format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
@@
   cat("\n")
-  cat("Cohen's Kappa for DL vs. the gold standard"," was ",round(dl$value,3)," (p = ",
+  cat("Cohen's Kappa for DL"," was ",round(dl$value,3)," (p = ",
       format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
@@
   cat("\n")
-  cat("Cohen's Kappa for JW vs. the gold standard"," was ",round(jw$value,3)," (p = ",
+  cat("Cohen's Kappa for JW"," was ",round(jw$value,3)," (p = ",
       format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")
@@
   invisible(lapply(levels(factor(df$feature)), function(f){
-    t = data.frame(df[df$feature == f,c("AF","DL","GS","JW")])
+    t = data.frame(df[df$feature == f,c("AF_1","AF_2","DL_1","DL_2","JW_1","JW_2")])
     rownames(t) = as.character(df$location[df$feature == f])
-    fk = kappam.fleiss(t)
-    af = kappa2(t[,c("AF","GS")])
-    dl = kappa2(t[,c("DL","GS")])
-    jw = kappa2(t[,c("JW","GS")])
+    af = kappa2(t[,c("AF_1","AF_2")])
+    dl = kappa2(t[,c("DL_1","DL_2")])
+    jw = kappa2(t[,c("JW_1","JW_2")])
     p = df %>%
       filter(feature == f) %>%
-      select(location, AF:JW) %>%
-      pivot_longer(AF:JW) %>%
+      select(location, AF_1:JW_2) %>%
+      pivot_longer(AF_1:JW_2) %>%
       ggplot(., aes(x = name, y = location, fill = value)) +
@@
     cat("\n")
-    cat("Fleiss' Kappa for ",f," was ",round(fk$value,3)," (p = ",
-        format.pval(fk$p.value,eps = 0.001,digits = 3),")",sep = "")
-    cat("\n")
-    cat("\n")
     if(is.nan(af$value)){
       cat("\n")
-      cat("Agreement between AF and the gold standard in",f,"was perfect.")
+      cat("Agreement within AF",f,"was perfect.")
       cat("\n")
@@
       cat("\n")
-      cat("Cohen's Kappa for AF vs. the gold standard in ",f," was ",round(af$value,3)," (p = ",
+      cat("Cohen's Kappa for AF",f," was ",round(af$value,3)," (p = ",
           format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
@@
       cat("\n")
-      cat("Agreement between DL and the gold standard in",f,"was perfect.")
+      cat("Agreement within DL",f,"was perfect.")
       cat("\n")
@@
       cat("\n")
-      cat("Cohen's Kappa for DL vs. the gold standard in ",f," was ",round(dl$value,3)," (p = ",
+      cat("Cohen's Kappa for DL",f," was ",round(dl$value,3)," (p = ",
           format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
@@
       cat("\n")
-      cat("Agreement between JW and the gold standard in",f,"was perfect.")
+      cat("Agreement within JW in",f,"was perfect.")
       cat("\n")
@@
       cat("\n")
-      cat("Cohen's Kappa for JW vs. the gold standard in ",f," was ",round(jw$value,3)," (p = ",
+      cat("Cohen's Kappa for JW",f," was ",round(jw$value,3)," (p = ",
           format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")
@@
 ```
+
+## Scan 1
+
+```{r results='asis',warning=FALSE}
+compare(df = dup_df,scan_num = 1062761)
+```
+
+## Scan 2
+
+```{r results='asis',warning=FALSE}
+compare(df = dup_df,scan_num = 1151991)
+```
+
+## Scan 3
+
+```{r results='asis',warning=FALSE}
+compare(df = dup_df,scan_num = 1340551)
+```
+
+## Scan 4
+
+```{r results='asis',warning=FALSE}
+compare(df = dup_df,scan_num = 1419420)
+```
+
+## Scan 5
+
+```{r results='asis',warning=FALSE}
+compare(df = dup_df,scan_num = 1468121)
+```

### 2023-05-18 Stacey Martiniano/ntm_agreement.qmd
Update ntm_agreement.qmd

--- 2023-05-17 Stacey Martiniano/ntm_agreement.qmd
@@
 data$scorer[data$scorer=="GG"]="GS"
+# Deidentify readers
+old <- c("AF","DL","JW")
+new <- c("A","B","C")
+data$scorer[data$scorer %in% old] <- new[match(data$scorer, old, nomatch = 0)]
 # Revert duplicate IDs back to original to match with gold standard
@@
   # Kappa for the whole scan
-  fk = kappam.fleiss(df[,c("AF","DL","GS","JW")])
-  af = kappa2(df[,c("AF","GS")])
-  dl = kappa2(df[,c("DL","GS")])
-  jw = kappa2(df[,c("JW","GS")])
+  fk = kappam.fleiss(df[,c("A","B","GS","C")])
+  af = kappa2(df[,c("A","GS")])
+  dl = kappa2(df[,c("B","GS")])
+  jw = kappa2(df[,c("C","GS")])
   cat("\n")
@@
   p = df %>%
-    select(feature, location, AF:JW) %>%
-    pivot_longer(AF:JW) %>%
+    select(feature, location, A:C) %>%
+    pivot_longer(A:C) %>%
     ggplot(., aes(x = name, y = location, fill = value)) +
@@
   cat("\n")
-  cat("Cohen's Kappa for AF vs. the gold standard"," was ",round(af$value,3)," (p = ",
+  cat("Cohen's Kappa for A vs. the gold standard"," was ",round(af$value,3)," (p = ",
       format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
@@
   cat("\n")
-  cat("Cohen's Kappa for DL vs. the gold standard"," was ",round(dl$value,3)," (p = ",
+  cat("Cohen's Kappa for B vs. the gold standard"," was ",round(dl$value,3)," (p = ",
       format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
@@
   cat("\n")
-  cat("Cohen's Kappa for JW vs. the gold standard"," was ",round(jw$value,3)," (p = ",
+  cat("Cohen's Kappa for C vs. the gold standard"," was ",round(jw$value,3)," (p = ",
       format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")
@@
   invisible(lapply(levels(factor(df$feature)), function(f){
-    t = data.frame(df[df$feature == f,c("AF","DL","GS","JW")])
+    t = data.frame(df[df$feature == f,c("A","B","GS","C")])
     rownames(t) = as.character(df$location[df$feature == f])
     fk = kappam.fleiss(t)
-    af = kappa2(t[,c("AF","GS")])
-    dl = kappa2(t[,c("DL","GS")])
-    jw = kappa2(t[,c("JW","GS")])
+    af = kappa2(t[,c("A","GS")])
+    dl = kappa2(t[,c("B","GS")])
+    jw = kappa2(t[,c("C","GS")])
     p = df %>%
       filter(feature == f) %>%
-      select(location, AF:JW) %>%
-      pivot_longer(AF:JW) %>%
+      select(location, A:C) %>%
+      pivot_longer(A:C) %>%
       ggplot(., aes(x = name, y = location, fill = value)) +
@@
       cat("\n")
-      cat("Agreement between AF and the gold standard in",f,"was perfect.")
+      cat("Agreement between A and the gold standard in",f,"was perfect.")
       cat("\n")
@@
       cat("\n")
-      cat("Cohen's Kappa for AF vs. the gold standard in ",f," was ",round(af$value,3)," (p = ",
+      cat("Cohen's Kappa for A vs. the gold standard in ",f," was ",round(af$value,3)," (p = ",
           format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
@@
       cat("\n")
-      cat("Agreement between DL and the gold standard in",f,"was perfect.")
+      cat("Agreement between B and the gold standard in",f,"was perfect.")
       cat("\n")
@@
       cat("\n")
-      cat("Cohen's Kappa for DL vs. the gold standard in ",f," was ",round(dl$value,3)," (p = ",
+      cat("Cohen's Kappa for B vs. the gold standard in ",f," was ",round(dl$value,3)," (p = ",
           format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
@@
       cat("\n")
-      cat("Agreement between JW and the gold standard in",f,"was perfect.")
+      cat("Agreement between C and the gold standard in",f,"was perfect.")
       cat("\n")
@@
       cat("\n")
-      cat("Cohen's Kappa for JW vs. the gold standard in ",f," was ",round(jw$value,3)," (p = ",
+      cat("Cohen's Kappa for C vs. the gold standard in ",f," was ",round(jw$value,3)," (p = ",
           format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")
@@
     pivot_wider(names_from = scorer, values_from = value) %>%
-    arrange(feature,n,location)
+    pivot_wider(names_from = n, values_from = A:C) %>%
+    arrange(feature,location)
   # Kappa for the whole scan
-  fk = kappam.fleiss(df[,c("AF","DL","GS","JW")])
-  af = kappa2(df[,c("AF","GS")])
-  dl = kappa2(df[,c("DL","GS")])
-  jw = kappa2(df[,c("JW","GS")])
+  af = kappa2(df[,c("A_1","A_2")])
+  dl = kappa2(df[,c("B_1","B_2")])
+  jw = kappa2(df[,c("C_1","C_2")])
   cat("\n")
@@
   p = df %>%
-    select(feature, location, AF:JW) %>%
-    pivot_longer(AF:JW) %>%
+    select(feature, location, A_1:C_2) %>%
+    pivot_longer(A_1:C_2) %>%
+    ggplot(., aes(x = name, y = location, fill = value)) +
+    geom_tile(color = "black") +
+    theme_bw() +
+    labs(fill = "", x = "Scorer", y = "Location") +
+    custom_colors +
+    facet_wrap(~feature)
+  cat("\n")
+  print(p)
+  cat("\n")
+  cat("\n")
+  cat("Cohen's Kappa for A"," was ",round(af$value,3)," (p = ",
+      format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
+  cat("\n")
+  cat("\n")
+  cat("Cohen's Kappa for B"," was ",round(dl$value,3)," (p = ",
+      format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
+  cat("\n")
+  cat("\n")
+  cat("Cohen's Kappa for C"," was ",round(jw$value,3)," (p = ",
+      format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")
+  cat("\n")
+  # Analyze by feature
+  cat("\n")
+  cat("### By Feature and Scorer")
+  cat("\n")
+  invisible(lapply(levels(factor(df$feature)), function(f){
+    t = data.frame(df[df$feature == f,c("A_1","A_2","B_1","B_2","C_1","C_2")])
+    rownames(t) = as.character(df$location[df$feature == f])
+    af = kappa2(t[,c("A_1","A_2")])
+    dl = kappa2(t[,c("B_1","B_2")])
+    jw = kappa2(t[,c("C_1","C_2")])
+    p = df %>%
+      filter(feature == f) %>%
+      select(location, A_1:C_2) %>%
+      pivot_longer(A_1:C_2) %>%
+      ggplot(., aes(x = name, y = location, fill = value)) +
+      geom_tile(color = "black") +
+      theme_bw() +
+      labs(fill = "", x = "Scorer", y = "Location") +
+      custom_colors
+    cat("\n")
+    cat("\n")
+    cat("####",f)
+    cat("\n")
+    cat("\n")
+    print(p)
+    cat("\n")
+    cat("\n")
+    if(is.nan(af$value)){
+      cat("\n")
+      cat("Agreement within A",f,"was perfect.")
+      cat("\n")
+    } else {
+      cat("\n")
+      cat("Cohen's Kappa for A",f," was ",round(af$value,3)," (p = ",
+          format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
+      cat("\n")
+    }
+    if(is.nan(dl$value)){
+      cat("\n")
+      cat("Agreement within B",f,"was perfect.")
+      cat("\n")
+    } else {
+      cat("\n")
+      cat("Cohen's Kappa for B",f," was ",round(dl$value,3)," (p = ",
+          format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
+      cat("\n")
+    }
+    if(is.nan(jw$value)){
+      cat("\n")
+      cat("Agreement within C in",f,"was perfect.")
+      cat("\n")
+    } else {
+      cat("\n")
+      cat("Cohen's Kappa for C",f," was ",round(jw$value,3)," (p = ",
+          format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")
+      cat("\n")
+    }
+  }))
+}
+```
+
+## Scan 1
+
+```{r results='asis',warning=FALSE}
+compare(df = dup_df,scan_num = 1062761)
+```
+
+## Scan 2
+
+```{r results='asis',warning=FALSE}
+compare(df = dup_df,scan_num = 1151991)
+```
+
+## Scan 3
+
+```{r results='asis',warning=FALSE}
+compare(df = dup_df,scan_num = 1340551)
+```
+
+## Scan 4
+
+```{r results='asis',warning=FALSE}
+compare(df = dup_df,scan_num = 1419420)
+```
+
+## Scan 5
+
+```{r results='asis',warning=FALSE}
+compare(df = dup_df,scan_num = 1468121)
+```
+
+# Singleton scans
+
+```{r}
+# Gold standard data
+other_df = data %>% 
+  arrange(scan_id, scorer) %>%
+  filter(!(scan_id %in% dups$original_ssid) & 
+           !(scan_id %in% training$Cffid.Scoring.Set )) %>%
+  select(scorer, scan_id, bronchiectasis_rul:atelectasis___6) %>%
+  ungroup()
+# Function
+compare = function(df,scan_num){
+  df = df %>% filter(scan_id == scan_num)
+  df = df %>% pivot_longer(bronchiectasis_rul:atelectasis___6,
+                           names_to = c("feature", "location"),
+                           names_pattern = "(.*)_(.*)$"
+  ) %>% 
+    mutate(
+      feature = gsub("__", "", feature),
+      feature = str_to_title(gsub("_", " ", feature))
+    ) 
+  df$location <- factor(df$location,
+                        levels = c("1", "2", "3", "4", "5", "6", "ling", "lll", 
+                                   "lul", "rll", "rml", "rul"),
+                        labels = c("RUL", "RML", "RLL", "LUL", "Ling", "LLL", 
+                                   "Ling", "LLL", "LUL", "RLL", "RML", "RUL")
+  )
+  df$value <- factor(df$value, levels = c("Checked", "Unchecked", 
+                                          "Absent", "Cylindric", "Varicose"))
+  # One column per scorer
+  df <- df %>% 
+    pivot_wider(names_from = scorer, values_from = value) %>%
+    arrange(feature,location)
+  # Kappa for the whole scan
+  fk = kappam.fleiss(df[,c("A","B","C")])
+  cat("\n")
+  cat("### Overall")
+  cat("\n")
+  p = df %>%
+    select(feature, location, A:C) %>%
+    pivot_longer(A:C) %>%
     ggplot(., aes(x = name, y = location, fill = value)) +
@@
   cat("\n")
-  cat("Cohen's Kappa for AF vs. the gold standard"," was ",round(af$value,3)," (p = ",
-      format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
-  cat("\n")
-  cat("\n")
-  cat("Cohen's Kappa for DL vs. the gold standard"," was ",round(dl$value,3)," (p = ",
-      format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
-  cat("\n")
-  cat("\n")
-  cat("Cohen's Kappa for JW vs. the gold standard"," was ",round(jw$value,3)," (p = ",
-      format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")
-  cat("\n")
   # Analyze by feature
   cat("\n")
-  cat("### By Feature and Scorer")
+  cat("### By Feature")
   cat("\n")
   invisible(lapply(levels(factor(df$feature)), function(f){
-    t = data.frame(df[df$feature == f,c("AF","DL","GS","JW")])
+    t = data.frame(df[df$feature == f,c("A","B","C")])
     rownames(t) = as.character(df$location[df$feature == f])
     fk = kappam.fleiss(t)
-    af = kappa2(t[,c("AF","GS")])
-    dl = kappa2(t[,c("DL","GS")])
-    jw = kappa2(t[,c("JW","GS")])
     p = df %>%
       filter(feature == f) %>%
-      select(location, AF:JW) %>%
-      pivot_longer(AF:JW) %>%
+      select(location, A:C) %>%
+      pivot_longer(A:C) %>%
       ggplot(., aes(x = name, y = location, fill = value)) +
@@
     cat("\n")
-    cat("\n")
-    if(is.nan(af$value)){
-      cat("\n")
-      cat("Agreement between AF and the gold standard in",f,"was perfect.")
-      cat("\n")
-    } else {
-      cat("\n")
-      cat("Cohen's Kappa for AF vs. the gold standard in ",f," was ",round(af$value,3)," (p = ",
-          format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
-      cat("\n")
-    }
-    if(is.nan(dl$value)){
-      cat("\n")
-      cat("Agreement between DL and the gold standard in",f,"was perfect.")
-      cat("\n")
-    } else {
-      cat("\n")
-      cat("Cohen's Kappa for DL vs. the gold standard in ",f," was ",round(dl$value,3)," (p = ",
-          format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
-      cat("\n")
-    }
-    if(is.nan(jw$value)){
-      cat("\n")
-      cat("Agreement between JW and the gold standard in",f,"was perfect.")
-      cat("\n")
-    } else {
-      cat("\n")
-      cat("Cohen's Kappa for JW vs. the gold standard in ",f," was ",round(jw$value,3)," (p = ",
-          format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")
-      cat("\n")
-    }
   }))
@@
 ```
+
+```{r results='asis'}
+invisible(lapply(unique(other_df$scan_id), function(s){
+  cat("\n")
+  cat("## Scan",s)
+  cat("\n")
+  compare(other_df,scan_num = s)
+  cat("\n")
+}))
+```

### 2023-05-18 Emily Muther/body_image_trikafta.qmd
Emma analysis done

--- 2023-03-25 Emily Muther/body_image_trikafta.qmd
@@
 author: "Tim Vigers"
-date: today
-date-format: long
+date: "May 12, 2023"
 format:
@@
 library(foreign)
+library(plotly)
+library(broom)
+library(report)
 # Read in data
@@
 
-# Figure
+# Mental health scores by FRS category
+
+## PHQ9
+
+### Odds of an elevated score
+
+Results are reported as odds ratios, with the FRS "neutral" category as the reference group.
 
 ```{r}
-# | label: fig-planets
-# | fig-cap: Planets
+mod = glm(PHQ9_elevated ~ FRS_Category,data = df,family = "binomial")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
 
+`r report(mod)`
+
+### Association with FRS category in those with elevated scores.
+
+Results are reported as expected counts, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod = glm(PHQ9_Total~ FRS_Category,data = df[df$PHQ9_elevated=="yes",],family = "poisson")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
 ```
+
+`r report(mod)`
+
+## GAD7
+
+### Odds of an elevated score
+
+Results are reported as odds ratios, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod = glm(GAD7_elevated ~ FRS_Category,data = df,family = "binomial")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+### Association with FRS category in those with elevated scores.
+
+Results are reported as expected counts, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod = glm(GAD7_Total~ FRS_Category,data = df[df$GAD7_elevated=="yes",],family = "poisson")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+# Mental health scores by BES
+
+## PHQ9
+
+### BE_APPEARANCE_MEAN
+
+#### Odds of an elevated score
+
+Results are reported as odds ratios for a one-unit change in the predictor.
+
+```{r}
+mod = glm(PHQ9_elevated ~ BE_APPEARANCE_MEAN,data = df,family = "binomial")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+#### Association with BE_APPEARANCE_MEAN in those with elevated scores.
+
+Results are reported as expected counts, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod = glm(PHQ9_Total~ BE_APPEARANCE_MEAN,data = df[df$PHQ9_elevated=="yes",],family = "poisson")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+### BE_ATTRIBUTION_MEAN
+
+#### Odds of an elevated score
+
+Results are reported as odds ratios for a one-unit change in the predictor.
+
+```{r}
+mod = glm(PHQ9_elevated ~ BE_ATTRIBUTION_MEAN,data = df,family = "binomial")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+#### Association with BE_ATTRIBUTION_MEAN in those with elevated scores.
+
+Results are reported as expected counts, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod = glm(PHQ9_Total~ BE_ATTRIBUTION_MEAN,data = df[df$PHQ9_elevated=="yes",],family = "poisson")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+### BE_WEIGHT_MEAN
+
+#### Odds of an elevated score
+
+Results are reported as odds ratios for a one-unit change in the predictor.
+
+```{r}
+mod = glm(PHQ9_elevated ~ BE_WEIGHT_MEAN,data = df,family = "binomial")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+#### Association with BE_WEIGHT_MEAN in those with elevated scores.
+
+Results are reported as expected counts, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod = glm(PHQ9_Total~ BE_WEIGHT_MEAN,data = df[df$PHQ9_elevated=="yes",],family = "poisson")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+## GAD7
+
+### BE_APPEARANCE_MEAN
+
+#### Odds of an elevated score
+
+Results are reported as odds ratios for a one-unit change in the predictor.
+
+```{r}
+mod = glm(GAD7_elevated ~ BE_APPEARANCE_MEAN,data = df,family = "binomial")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+#### Association with BE_APPEARANCE_MEAN in those with elevated scores.
+
+Results are reported as expected counts, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod = glm(GAD7_Total~ BE_APPEARANCE_MEAN,data = df[df$GAD7_elevated=="yes",],family = "poisson")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+### BE_ATTRIBUTION_MEAN
+
+#### Odds of an elevated score
+
+Results are reported as odds ratios for a one-unit change in the predictor.
+
+```{r}
+mod = glm(GAD7_elevated ~ BE_ATTRIBUTION_MEAN,data = df,family = "binomial")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+#### Association with BE_ATTRIBUTION_MEAN in those with elevated scores.
+
+Results are reported as expected counts, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod = glm(GAD7_Total~ BE_ATTRIBUTION_MEAN,data = df[df$GAD7_elevated=="yes",],family = "poisson")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+### BE_WEIGHT_MEAN
+
+#### Odds of an elevated score
+
+Results are reported as odds ratios for a one-unit change in the predictor.
+
+```{r}
+mod = glm(GAD7_elevated ~ BE_WEIGHT_MEAN,data = df,family = "binomial")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+#### Association with BE_WEIGHT_MEAN in those with elevated scores.
+
+Results are reported as expected counts, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod = glm(GAD7_Total~ BE_WEIGHT_MEAN,data = df[df$GAD7_elevated=="yes",],family = "poisson")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`

### 2023-05-18 Stacey Martiniano/ntm_agreement.qmd
last report with kappas for everything

@@
   p = df %>%
-    select(feature, location, A:C) %>%
-    pivot_longer(A:C) %>%
+    select(feature, location, A:GS) %>%
+    pivot_longer(A:GS) %>%
     ggplot(., aes(x = name, y = location, fill = value)) +
@@
       filter(feature == f) %>%
-      select(location, A:C) %>%
-      pivot_longer(A:C) %>%
+      select(location, A:GS) %>%
+      pivot_longer(A:GS) %>%
       ggplot(., aes(x = name, y = location, fill = value)) +
@@
       cat("\n")
-      cat("Agreement within A",f,"was perfect.")
+      cat("Agreement within A ",f,"was perfect.")
       cat("\n")
@@
       cat("\n")
-      cat("Cohen's Kappa for A",f," was ",round(af$value,3)," (p = ",
+      cat("Cohen's Kappa for A ",f," was ",round(af$value,3)," (p = ",
           format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
@@
       cat("\n")
-      cat("Agreement within B",f,"was perfect.")
+      cat("Agreement within B ",f,"was perfect.")
       cat("\n")
@@
       cat("\n")
-      cat("Cohen's Kappa for B",f," was ",round(dl$value,3)," (p = ",
+      cat("Cohen's Kappa for B ",f," was ",round(dl$value,3)," (p = ",
           format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
@@
       cat("\n")
-      cat("Cohen's Kappa for C",f," was ",round(jw$value,3)," (p = ",
+      cat("Cohen's Kappa for C ",f," was ",round(jw$value,3)," (p = ",
           format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")

### 2023-05-18 Stacey Martiniano/ntm_agreement.qmd
proportion of agreement instead of kappas

--- 2023-05-17 Stacey Martiniano/ntm_agreement.qmd
@@
 library(RColorBrewer)
+library(knitr)
+library(pheatmap)
 # Import data
@@
 data$scan_id[data$scan_id %in% old] <- new[match(data$scan_id, old, nomatch = 0)]
+# Plot colors
+name <- "Accent"
+colors <- brewer.pal(6, name)
+names(colors) <- c("Checked","Unchecked","Absent","Cylindric","Varicose",NA)
+custom_colors <- scale_fill_manual(name = "", values = colors)
 ```
@@
 )
-  df$value <- factor(df$value, levels = c("Checked", "Unchecked", "Absent", "Cylindric", "Varicose"))
 # One column per scorer
-  df <- df %>% 
+gs <- gs %>% 
   pivot_wider(names_from = scorer, values_from = value) %>%
   arrange(feature,location)
-  # Kappa for the whole scan
-  fk = kappam.fleiss(df[,c("AF","DL","GS","JW")])
-  af = kappa2(df[,c("AF","GS")])
-  dl = kappa2(df[,c("DL","GS")])
-  jw = kappa2(df[,c("JW","GS")])
-  cat("\n")
-  cat("### Overall")
-  cat("\n")
-  p = df %>%
-    select(feature, location, AF:JW) %>%
-    pivot_longer(AF:JW) %>%
+```
+
+## Bronciechtasis
+
+```{r}
+gs_bronch = gs %>% filter(feature=="Bronchiectasis")
+fk = kappam.fleiss(gs_bronch[,c("A","B","C","GS")])
+A = kappa2(gs_bronch[,c("A","GS")])
+B = kappa2(gs_bronch[,c("B","GS")])
+C = kappa2(gs_bronch[,c("C","GS")])
+# Plot
+gs_bronch %>%
+  select(scan_id,feature, location, A:GS) %>%
+  pivot_longer(A:GS) %>%
   ggplot(., aes(x = name, y = location, fill = value)) +
@@
   custom_colors +
-    facet_wrap(~feature)
-  cat("\n")
-  print(p)
-  cat("\n")
-  cat("\n")
-  cat("Overall Fleiss' Kappa for this scan was ",round(fk$value,3)," (p = ",
-      format.pval(fk$p.value,eps = 0.001,digits = 3),")",sep = "")
-  cat("\n")
-  cat("\n")
-  cat("Cohen's Kappa for AF vs. the gold standard"," was ",round(af$value,3)," (p = ",
-      format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
-  cat("\n")
-  cat("\n")
-  cat("Cohen's Kappa for DL vs. the gold standard"," was ",round(dl$value,3)," (p = ",
-      format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
-  cat("\n")
-  cat("\n")
-  cat("Cohen's Kappa for JW vs. the gold standard"," was ",round(jw$value,3)," (p = ",
-      format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")
-  cat("\n")
-  # Analyze by feature
-  cat("\n")
-  cat("### By Feature and Scorer")
-  cat("\n")
-  invisible(lapply(levels(factor(df$feature)), function(f){
-    t = data.frame(df[df$feature == f,c("AF","DL","GS","JW")])
-    rownames(t) = as.character(df$location[df$feature == f])
-    fk = kappam.fleiss(t)
-    af = kappa2(t[,c("AF","GS")])
-    dl = kappa2(t[,c("DL","GS")])
-    jw = kappa2(t[,c("JW","GS")])
-    p = df %>%
-      filter(feature == f) %>%
-      select(location, AF:JW) %>%
-      pivot_longer(AF:JW) %>%
+  facet_wrap(~scan_id)
+```
+
+Overall Fleiss' Kappa for all 5 gold standard scans was `r round(fk$value,3)` (p = `r format.pval(fk$p.value,eps = 0.001,digits = 3)`).
+
+Cohen's Kappa for A vs. the gold standard was `r round(A$value,3)` (p = `r format.pval(A$p.value,eps = 0.001,digits = 3)`).
+
+Cohen's Kappa for B vs. the gold standard was `r round(B$value,3)` (p = `r format.pval(B$p.value,eps = 0.001,digits = 3)`).
+
+Cohen's Kappa for C vs. the gold standard was `r round(C$value,3)` (p = `r format.pval(C$p.value,eps = 0.001,digits = 3)`).
+
+## Other features
+
+```{r}
+gs_other = gs %>% filter(feature!="Bronchiectasis")
+# Check agreement with GS column
+gs_other$A = factor(gs_other$A == gs_other$GS,levels = c(F,T),labels = c("Disagree","Agree"))
+gs_other$B = factor(gs_other$B == gs_other$GS,levels = c(F,T),labels = c("Disagree","Agree"))
+gs_other$C = factor(gs_other$C == gs_other$GS,levels = c(F,T),labels = c("Disagree","Agree"))
+# Plot
+gs_other %>%
+  select(scan_id,feature, location, A:GS) %>%
+  pivot_longer(A:C) %>%
   ggplot(., aes(x = name, y = location, fill = value)) +
@@
   labs(fill = "", x = "Scorer", y = "Location") +
-      custom_colors
-    cat("\n")
-    cat("\n")
-    cat("####",f)
-    cat("\n")
-    cat("\n")
-    print(p)
-    cat("\n")
-    cat("\n")
-    cat("Fleiss' Kappa for ",f," was ",round(fk$value,3)," (p = ",
-        format.pval(fk$p.value,eps = 0.001,digits = 3),")",sep = "")
-    cat("\n")
-    cat("\n")
-    if(is.nan(af$value)){
-      cat("\n")
-      cat("Agreement between AF and the gold standard in",f,"was perfect.")
-      cat("\n")
-    } else {
-      cat("\n")
-      cat("Cohen's Kappa for AF vs. the gold standard in ",f," was ",round(af$value,3)," (p = ",
-          format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
-      cat("\n")
-    }
-    if(is.nan(dl$value)){
-      cat("\n")
-      cat("Agreement between DL and the gold standard in",f,"was perfect.")
-      cat("\n")
-    } else {
-      cat("\n")
-      cat("Cohen's Kappa for DL vs. the gold standard in ",f," was ",round(dl$value,3)," (p = ",
-          format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
-      cat("\n")
-    }
-    if(is.nan(jw$value)){
-      cat("\n")
-      cat("Agreement between JW and the gold standard in",f,"was perfect.")
-      cat("\n")
-    } else {
-      cat("\n")
-      cat("Cohen's Kappa for JW vs. the gold standard in ",f," was ",round(jw$value,3)," (p = ",
-          format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")
-      cat("\n")
-    }
-  }))
-}
-```
-
-## Scan 1
-
-```{r results='asis',warning=FALSE}
-compare(df = gs,scan_num = 1)
-```
-
-## Scan 2
-
-```{r results='asis',warning=FALSE}
-compare(df = gs,scan_num = 2)
-```
-
-## Scan 3
-
-```{r results='asis',warning=FALSE}
-compare(df = gs,scan_num = 3)
-```
-
-## Scan 4
-
-```{r results='asis',warning=FALSE}
-compare(df = gs,scan_num = 4)
-```
-
-## Scan 5
-
-```{r results='asis',warning=FALSE}
-compare(df = gs,scan_num = 5)
-```
-
-```{r}
-# Save dataset
-write.csv(gs,"~/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/gold_standard_comp.csv",row.names = F,na="")
+  facet_grid(rows = vars(feature),cols = vars(scan_id))
+# Tables
+gs_other %>%
+  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "drop") %>%
+  kable(.,digits=3,caption = "Proportion of agreement with GS")
+gs_other %>%
+  group_by(scan_id) %>%
+  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "drop") %>%
+  kable(.,digits=3,caption = "Proportion of agreement with GS")
+gs_other %>%
+  group_by(feature) %>%
+  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "drop") %>%
+  kable(.,digits=3,caption = "Proportion of agreement with GS")
+gs_other %>%
+  group_by(scan_id,feature) %>%
+  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "drop") %>%
+  kable(.,digits=3,caption = "Proportion of agreement with GS")
+# Heatmaps
+gs_other %>%
+  group_by(scan_id) %>%
+  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "keep") %>%
+  column_to_rownames("scan_id") %>%
+  pheatmap(.,cluster_rows = F,cluster_cols = F)
+gs_other %>%
+  group_by(feature) %>%
+  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "keep") %>%
+  column_to_rownames("feature") %>%
+  pheatmap(.,cluster_rows = F,cluster_cols = F)
+gs_other %>%
+  group_by(scan_id,feature) %>%
+  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "keep") %>%
+  unite(ids,scan_id,feature,na.rm = T,sep = " ") %>%
+  column_to_rownames("ids") %>%
+  pheatmap(.,cluster_rows = F,cluster_cols = F)
 ```
@@
 # Gold standard data
-dup_df = data %>% 
+dup = data %>% 
   arrange(scan_id, scorer) %>%
   filter(scan_id %in% dups$original_ssid) %>%
-  select(scorer, scan_id, bronchiectasis_rul:atelectasis___6) %>%
-  ungroup()
-# Function
-compare = function(df,scan_num){
-  df = df %>% filter(scan_id == scan_num) %>%
   group_by(scorer,scan_id) %>%
-    mutate(n = row_number()) %>% ungroup() %>%
-    pivot_longer(bronchiectasis_rul:atelectasis___6,
+  mutate(n=row_number())%>%
+  select(scorer, scan_id,n ,bronchiectasis_rul:atelectasis___6) %>%
+  ungroup() %>% pivot_longer(bronchiectasis_rul:atelectasis___6,
                              names_to = c("feature", "location"),
@@
   ) 
-  df$location <- factor(df$location,
+dup$location <- factor(dup$location,
                        levels = c("1", "2", "3", "4", "5", "6", "ling", "lll", 
@@
 )
-  df$value <- factor(df$value, levels = c("Checked", "Unchecked", "Absent", "Cylindric", "Varicose"))
 # One column per scorer
-  df <- df %>% 
+dup <- dup %>% 
   pivot_wider(names_from = scorer, values_from = value) %>%
-    pivot_wider(names_from = n, values_from = AF:JW) %>%
+  pivot_wider(names_from = n, values_from = A:C) %>%
   arrange(feature,location)
-  # Kappa for the whole scan
-  af = kappa2(df[,c("AF_1","AF_2")])
-  dl = kappa2(df[,c("DL_1","DL_2")])
-  jw = kappa2(df[,c("JW_1","JW_2")])
-  cat("\n")
-  cat("### Overall")
-  cat("\n")
-  p = df %>%
-    select(feature, location, AF_1:JW_2) %>%
-    pivot_longer(AF_1:JW_2) %>%
+```
+
+## Bronciechtasis
+
+```{r}
+dup_bronch = dup %>% filter(feature=="Bronchiectasis")
+A = kappa2(dup_bronch[,c("A_1","A_2")])
+B = kappa2(dup_bronch[,c("B_1","B_2")])
+C = kappa2(dup_bronch[,c("C_1","C_2")])
+# Plot
+dup_bronch %>%
+  select(scan_id,feature, location,A_1:C_2) %>%
+  pivot_longer(A_1:C_2) %>%
   ggplot(., aes(x = name, y = location, fill = value)) +
@@
   custom_colors +
-    facet_wrap(~feature)
-  cat("\n")
-  print(p)
-  cat("\n")
-  cat("\n")
-  cat("Cohen's Kappa for AF"," was ",round(af$value,3)," (p = ",
-      format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
-  cat("\n")
-  cat("\n")
-  cat("Cohen's Kappa for DL"," was ",round(dl$value,3)," (p = ",
-      format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
-  cat("\n")
-  cat("\n")
-  cat("Cohen's Kappa for JW"," was ",round(jw$value,3)," (p = ",
-      format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")
-  cat("\n")
-  # Analyze by feature
-  cat("\n")
-  cat("### By Feature and Scorer")
-  cat("\n")
-  invisible(lapply(levels(factor(df$feature)), function(f){
-    t = data.frame(df[df$feature == f,c("AF_1","AF_2","DL_1","DL_2","JW_1","JW_2")])
-    rownames(t) = as.character(df$location[df$feature == f])
-    af = kappa2(t[,c("AF_1","AF_2")])
-    dl = kappa2(t[,c("DL_1","DL_2")])
-    jw = kappa2(t[,c("JW_1","JW_2")])
-    p = df %>%
-      filter(feature == f) %>%
-      select(location, AF_1:JW_2) %>%
-      pivot_longer(AF_1:JW_2) %>%
+  facet_wrap(~scan_id)
+```
+
+Cohen's Kappa for A vs. themselves was `r round(A$value,3)` (p = `r format.pval(A$p.value,eps = 0.001,digits = 3)`).
+
+Cohen's Kappa for B vs. themselves was `r round(B$value,3)` (p = `r format.pval(B$p.value,eps = 0.001,digits = 3)`).
+
+Cohen's Kappa for C vs. themselves was `r round(C$value,3)` (p = `r format.pval(C$p.value,eps = 0.001,digits = 3)`).
+
+## Other features
+
+```{r}
+dup_other = dup %>% filter(feature!="Bronchiectasis")
+# Check agreement with GS column
+dup_other$A = factor(dup_other$A_1 == dup_other$A_2,levels = c(F,T),labels = c("Disagree","Agree"))
+dup_other$B = factor(dup_other$B_1 == dup_other$B_2,levels = c(F,T),labels = c("Disagree","Agree"))
+dup_other$C = factor(dup_other$C_1 == dup_other$C_2,levels = c(F,T),labels = c("Disagree","Agree"))
+# Plot
+dup_other %>%
+  select(scan_id,feature, location, A:C) %>%
+  pivot_longer(A:C) %>%
   ggplot(., aes(x = name, y = location, fill = value)) +
@@
   labs(fill = "", x = "Scorer", y = "Location") +
-      custom_colors
-    cat("\n")
-    cat("\n")
-    cat("####",f)
-    cat("\n")
-    cat("\n")
-    print(p)
-    cat("\n")
-    cat("\n")
-    if(is.nan(af$value)){
-      cat("\n")
-      cat("Agreement within AF",f,"was perfect.")
-      cat("\n")
-    } else {
-      cat("\n")
-      cat("Cohen's Kappa for AF",f," was ",round(af$value,3)," (p = ",
-          format.pval(af$p.value,eps = 0.001,digits = 3),")",sep = "")
-      cat("\n")
-    }
-    if(is.nan(dl$value)){
-      cat("\n")
-      cat("Agreement within DL",f,"was perfect.")
-      cat("\n")
-    } else {
-      cat("\n")
-      cat("Cohen's Kappa for DL",f," was ",round(dl$value,3)," (p = ",
-          format.pval(dl$p.value,eps = 0.001,digits = 3),")",sep = "")
-      cat("\n")
-    }
-    if(is.nan(jw$value)){
-      cat("\n")
-      cat("Agreement within JW in",f,"was perfect.")
-      cat("\n")
-    } else {
-      cat("\n")
-      cat("Cohen's Kappa for JW",f," was ",round(jw$value,3)," (p = ",
-          format.pval(jw$p.value,eps = 0.001,digits = 3),")",sep = "")
-      cat("\n")
-    }
-  }))
-}
+  facet_grid(rows = vars(feature),cols = vars(scan_id))
+# Tables
+dup_other %>%
+  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "drop") %>%
+  kable(.,digits=3,caption = "Proportion of agreement with themselves")
+dup_other %>%
+  group_by(scan_id) %>%
+  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "drop") %>%
+  kable(.,digits=3,caption = "Proportion of agreement with themselves")
+dup_other %>%
+  group_by(feature) %>%
+  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "drop") %>%
+  kable(.,digits=3,caption = "Proportion of agreement with themselves")
+dup_other %>%
+  group_by(scan_id,feature) %>%
+  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "drop") %>%
+  kable(.,digits=3,caption = "Proportion of agreement with themselves")
+# Heatmaps
+dup_other %>%
+  group_by(scan_id) %>%
+  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "keep") %>%
+  column_to_rownames("scan_id") %>%
+  pheatmap(.,cluster_rows = F,cluster_cols = F)
+dup_other %>%
+  group_by(feature) %>%
+  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "keep") %>%
+  column_to_rownames("feature") %>%
+  pheatmap(.,cluster_rows = F,cluster_cols = F)
+dup_other %>%
+  group_by(scan_id,feature) %>%
+  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "keep") %>%
+  unite(ids,scan_id,feature,na.rm = T,sep = " ") %>%
+  column_to_rownames("ids") %>%
+  pheatmap(.,cluster_rows = F,cluster_cols = F)
 ```
-
-## Scan 1
-
-```{r results='asis',warning=FALSE}
-compare(df = dup_df,scan_num = 1062761)
-```
-
-## Scan 2
-
-```{r results='asis',warning=FALSE}
-compare(df = dup_df,scan_num = 1151991)
-```
-
-## Scan 3
-
-```{r results='asis',warning=FALSE}
-compare(df = dup_df,scan_num = 1340551)
-```
-
-## Scan 4
-
-```{r results='asis',warning=FALSE}
-compare(df = dup_df,scan_num = 1419420)
-```
-
-## Scan 5
-
-```{r results='asis',warning=FALSE}
-compare(df = dup_df,scan_num = 1468121)
-```

### 2023-05-18 Stacey Martiniano/ntm_agreement.qmd
singleton scans

@@
 
-Cohen's Kappa for A vs. the gold standard was `r round(A$value,3)` (p = `r format.pval(A$p.value,eps = 0.001,digits = 3)`).
+Overall Cohen's Kappa for A vs. the gold standard was `r round(A$value,3)` (p = `r format.pval(A$p.value,eps = 0.001,digits = 3)`).
 
-Cohen's Kappa for B vs. the gold standard was `r round(B$value,3)` (p = `r format.pval(B$p.value,eps = 0.001,digits = 3)`).
+Overall Cohen's Kappa for B vs. the gold standard was `r round(B$value,3)` (p = `r format.pval(B$p.value,eps = 0.001,digits = 3)`).
 
-Cohen's Kappa for C vs. the gold standard was `r round(C$value,3)` (p = `r format.pval(C$p.value,eps = 0.001,digits = 3)`).
+Overall Cohen's Kappa for C vs. the gold standard was `r round(C$value,3)` (p = `r format.pval(C$p.value,eps = 0.001,digits = 3)`).
 
@@
 
-Cohen's Kappa for A vs. themselves was `r round(A$value,3)` (p = `r format.pval(A$p.value,eps = 0.001,digits = 3)`).
+Overall Cohen's Kappa for A vs. themselves was `r round(A$value,3)` (p = `r format.pval(A$p.value,eps = 0.001,digits = 3)`).
 
-Cohen's Kappa for B vs. themselves was `r round(B$value,3)` (p = `r format.pval(B$p.value,eps = 0.001,digits = 3)`).
+Overall Cohen's Kappa for B vs. themselves was `r round(B$value,3)` (p = `r format.pval(B$p.value,eps = 0.001,digits = 3)`).
 
-Cohen's Kappa for C vs. themselves was `r round(C$value,3)` (p = `r format.pval(C$p.value,eps = 0.001,digits = 3)`).
+Overall Cohen's Kappa for C vs. themselves was `r round(C$value,3)` (p = `r format.pval(C$p.value,eps = 0.001,digits = 3)`).
 
@@
 ```
+
+# Singleton scans
+
+```{r}
+# Gold standard data
+df = data %>% 
+  arrange(scan_id, scorer) %>%
+  filter(!(scan_id %in% dups$original_ssid) & 
+           !(scan_id %in% training$Cffid.Scoring.Set)) %>%
+  select(scorer, scan_id,bronchiectasis_rul:atelectasis___6) %>%
+  pivot_longer(bronchiectasis_rul:atelectasis___6,
+                             names_to = c("feature", "location"),
+                             names_pattern = "(.*)_(.*)$"
+  ) %>% 
+  mutate(
+    feature = gsub("__", "", feature),
+    feature = str_to_title(gsub("_", " ", feature))
+  ) 
+df$location <- factor(df$location,
+                       levels = c("1", "2", "3", "4", "5", "6", "ling", "lll", 
+                                  "lul", "rll", "rml", "rul"),
+                       labels = c("RUL", "RML", "RLL", "LUL", "Ling", "LLL", 
+                                  "Ling", "LLL", "LUL", "RLL", "RML", "RUL")
+)
+# One column per scorer
+df <- df %>% 
+  pivot_wider(names_from = scorer, values_from = value) %>%
+  arrange(feature,location)
+```
+
+## Bronciechtasis
+
+```{r}
+df_bronch = df %>% filter(feature=="Bronchiectasis")
+fk = kappam.fleiss(df_bronch[,c("A","B","C")])
+# Plot
+df_bronch %>%
+  select(scan_id,feature, location,A:C) %>%
+  pivot_longer(A:C) %>%
+  ggplot(., aes(x = name, y = location, fill = value)) +
+  geom_tile(color = "black") +
+  theme_bw() +
+  labs(fill = "", x = "Scorer", y = "Location") +
+  custom_colors +
+  facet_wrap(~scan_id)
+```
+
+Overall Fleiss' Kappa for all 15 singleton scans was `r round(fk$value,3)` (p = `r format.pval(fk$p.value,eps = 0.001,digits = 3)`).
+
+## Other features
+
+```{r}
+df_other = df %>% filter(feature!="Bronchiectasis")
+# Check agreement with GS column
+df_other[,c("A","B","C")] = lapply(df_other[,c("A","B","C")],factor,
+                                   levels = c("Unchecked", "Checked"),
+                                   labels = c("Absent","Present"))
+# Three-way agreement
+df_other$`A vs. B`=factor(df_other$A == df_other$B,levels = c(F,T),labels = c("Disagree","Agree"))
+df_other$`A vs. C`=factor(df_other$A == df_other$C,levels = c(F,T),labels = c("Disagree","Agree"))
+df_other$`B vs. C`=factor(df_other$B == df_other$C,levels = c(F,T),labels = c("Disagree","Agree"))
+# Plot
+df_other %>%
+  select(scan_id,feature, location, `A vs. B`:`B vs. C`) %>%
+  pivot_longer(`A vs. B`:`B vs. C`) %>%
+  ggplot(., aes(x = name, y = location, fill = value)) +
+  geom_tile(color = "black") +
+  theme_bw() +
+  labs(fill = "", x = "Scorer", y = "Location") +
+  facet_grid(rows = vars(feature),cols = vars(scan_id))
+# Tables
+df_other %>%
+  summarise(across(`A vs. B`:`B vs. C`,~mean(.x == "Agree",na.rm = T)),.groups = "drop") %>%
+  kable(.,digits=3,caption = "Proportion of agreement with themselves")
+df_other %>%
+  group_by(scan_id) %>%
+  summarise(across(`A vs. B`:`B vs. C`,~mean(.x == "Agree",na.rm = T)),.groups = "drop") %>%
+  kable(.,digits=3,caption = "Proportion of agreement with themselves")
+df_other %>%
+  group_by(feature) %>%
+  summarise(across(`A vs. B`:`B vs. C`,~mean(.x == "Agree",na.rm = T)),.groups = "drop") %>%
+  kable(.,digits=3,caption = "Proportion of agreement with themselves")
+df_other %>%
+  group_by(scan_id,feature) %>%
+  summarise(across(`A vs. B`:`B vs. C`,~mean(.x == "Agree",na.rm = T)),.groups = "drop") %>%
+  kable(.,digits=3,caption = "Proportion of agreement with themselves")
+# Heatmaps
+df_other %>%
+  group_by(scan_id) %>%
+  summarise(across(`A vs. B`:`B vs. C`,~mean(.x == "Agree",na.rm = T)),.groups = "keep") %>%
+  column_to_rownames("scan_id") %>%
+  pheatmap(.,cluster_rows = F,cluster_cols = F)
+df_other %>%
+  group_by(feature) %>%
+  summarise(across(`A vs. B`:`B vs. C`,~mean(.x == "Agree",na.rm = T)),.groups = "keep") %>%
+  column_to_rownames("feature") %>%
+  pheatmap(.,cluster_rows = F,cluster_cols = F)
+df_other %>%
+  group_by(scan_id,feature) %>%
+  summarise(across(`A vs. B`:`B vs. C`,~mean(.x == "Agree",na.rm = T)),.groups = "keep") %>%
+  unite(ids,scan_id,feature,na.rm = T,sep = " ") %>%
+  column_to_rownames("ids") %>%
+  pheatmap(.,cluster_rows = F,cluster_cols = F)
+```

### 2023-05-18 Stacey Martiniano/ntm_agreement.qmd
tidyverse formatting

--- 2023-05-18 Stacey Martiniano/ntm_agreement.qmd
@@
 data <- data.frame(lapply(data, as.character))
-data$scan_id = as.numeric(data$scan_id)
+data$scan_id <- as.numeric(data$scan_id)
 # Make scorer IDs by initial
@@
 )
-data$scorer[data$scorer=="GG"]="GS"
+data$scorer[data$scorer == "GG"] <- "GS"
 # Deidentify readers
-old <- c("AF","DL","JW")
-new <- c("A","B","C")
+old <- c("AF", "DL", "JW")
+new <- c("A", "B", "C")
 data$scorer[data$scorer %in% old] <- new[match(data$scorer, old, nomatch = 0)]
 # Revert duplicate IDs back to original to match with gold standard
-training = read.csv("~/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/training.key.csv")
-dups = read.csv("~/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/duplicated.key.csv")
-training$Training_id = as.numeric(sub("T00","",training$Training_id))
+training <- read.csv("~/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/training.key.csv")
+dups <- read.csv("~/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/duplicated.key.csv")
+training$Training_id <- as.numeric(sub("T00", "", training$Training_id))
 # Change duplicates to OG IDs
@@
 old <- training$Training_id
-new = training$Cffid.Scoring.Set
+new <- training$Cffid.Scoring.Set
 data$scan_id[data$scan_id %in% old] <- new[match(data$scan_id, old, nomatch = 0)]
@@
 colors <- brewer.pal(6, name)
-names(colors) <- c("Checked","Unchecked","Absent","Cylindric","Varicose",NA)
+names(colors) <- c("Checked", "Unchecked", "Absent", "Cylindric", "Varicose", NA)
 custom_colors <- scale_fill_manual(name = "", values = colors)
@@
 # Gold standard data
-gs = data %>% 
+gs <- data %>%
   arrange(scan_id, scorer) %>%
   filter(scan_id %in% training$Cffid.Scoring.Set) %>%
-  group_by(scorer,scan_id) %>%
+  group_by(scorer, scan_id) %>%
   slice_max(scoresheet_timestamp) %>%
   select(scorer, scan_id, bronchiectasis_rul:atelectasis___6) %>%
-  ungroup() %>% pivot_longer(bronchiectasis_rul:atelectasis___6,
+  ungroup() %>%
+  pivot_longer(bronchiectasis_rul:atelectasis___6,
     names_to = c("feature", "location"),
@@
 gs$location <- factor(gs$location,
-                      levels = c("1", "2", "3", "4", "5", "6", "ling", "lll", 
-                                 "lul", "rll", "rml", "rul"),
-                      labels = c("RUL", "RML", "RLL", "LUL", "Ling", "LLL", 
-                                 "Ling", "LLL", "LUL", "RLL", "RML", "RUL")
+  levels = c(
+    "1", "2", "3", "4", "5", "6", "ling", "lll",
+    "lul", "rll", "rml", "rul"
+  ),
+  labels = c(
+    "RUL", "RML", "RLL", "LUL", "Ling", "LLL",
+    "Ling", "LLL", "LUL", "RLL", "RML", "RUL"
+  )
 )
@@
   pivot_wider(names_from = scorer, values_from = value) %>%
-  arrange(feature,location)
+  arrange(feature, location)
 ```
@@
 ```{r}
-gs_bronch = gs %>% filter(feature=="Bronchiectasis")
-fk = kappam.fleiss(gs_bronch[,c("A","B","C","GS")])
-A = kappa2(gs_bronch[,c("A","GS")])
-B = kappa2(gs_bronch[,c("B","GS")])
-C = kappa2(gs_bronch[,c("C","GS")])
+gs_bronch <- gs %>% filter(feature == "Bronchiectasis")
+fk <- kappam.fleiss(gs_bronch[, c("A", "B", "C", "GS")])
+A <- kappa2(gs_bronch[, c("A", "GS")])
+B <- kappa2(gs_bronch[, c("B", "GS")])
+C <- kappa2(gs_bronch[, c("C", "GS")])
 # Plot
 gs_bronch %>%
-  select(scan_id,feature, location, A:GS) %>%
+  select(scan_id, feature, location, A:GS) %>%
   pivot_longer(A:GS) %>%
@@
 ```{r}
-gs_other = gs %>% filter(feature!="Bronchiectasis")
+gs_other <- gs %>% filter(feature != "Bronchiectasis")
 # Check agreement with GS column
-gs_other$A = factor(gs_other$A == gs_other$GS,levels = c(F,T),labels = c("Disagree","Agree"))
-gs_other$B = factor(gs_other$B == gs_other$GS,levels = c(F,T),labels = c("Disagree","Agree"))
-gs_other$C = factor(gs_other$C == gs_other$GS,levels = c(F,T),labels = c("Disagree","Agree"))
+gs_other$A <- factor(gs_other$A == gs_other$GS, levels = c(F, T), labels = c("Disagree", "Agree"))
+gs_other$B <- factor(gs_other$B == gs_other$GS, levels = c(F, T), labels = c("Disagree", "Agree"))
+gs_other$C <- factor(gs_other$C == gs_other$GS, levels = c(F, T), labels = c("Disagree", "Agree"))
 # Plot
 gs_other %>%
-  select(scan_id,feature, location, A:GS) %>%
+  select(scan_id, feature, location, A:GS) %>%
   pivot_longer(A:C) %>%
@@
   labs(fill = "", x = "Scorer", y = "Location") +
-  facet_grid(rows = vars(feature),cols = vars(scan_id))
+  facet_grid(rows = vars(feature), cols = vars(scan_id))
 # Tables
 gs_other %>%
-  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "drop") %>%
-  kable(.,digits=3,caption = "Proportion of agreement with GS")
+  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
+  kable(., digits = 3, caption = "Proportion of agreement with GS")
 gs_other %>%
   group_by(scan_id) %>%
-  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "drop") %>%
-  kable(.,digits=3,caption = "Proportion of agreement with GS")
+  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
+  kable(., digits = 3, caption = "Proportion of agreement with GS")
 gs_other %>%
   group_by(feature) %>%
-  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "drop") %>%
-  kable(.,digits=3,caption = "Proportion of agreement with GS")
+  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
+  kable(., digits = 3, caption = "Proportion of agreement with GS")
 gs_other %>%
-  group_by(scan_id,feature) %>%
-  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "drop") %>%
-  kable(.,digits=3,caption = "Proportion of agreement with GS")
+  group_by(scan_id, feature) %>%
+  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
+  kable(., digits = 3, caption = "Proportion of agreement with GS")
 # Heatmaps
@@
   group_by(scan_id) %>%
-  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "keep") %>%
+  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
   column_to_rownames("scan_id") %>%
-  pheatmap(.,cluster_rows = F,cluster_cols = F)
+  pheatmap(., cluster_rows = F, cluster_cols = F)
 gs_other %>%
   group_by(feature) %>%
-  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "keep") %>%
+  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
   column_to_rownames("feature") %>%
-  pheatmap(.,cluster_rows = F,cluster_cols = F)
+  pheatmap(., cluster_rows = F, cluster_cols = F)
 gs_other %>%
-  group_by(scan_id,feature) %>%
-  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "keep") %>%
-  unite(ids,scan_id,feature,na.rm = T,sep = " ") %>%
+  group_by(scan_id, feature) %>%
+  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
+  unite(ids, scan_id, feature, na.rm = T, sep = " ") %>%
   column_to_rownames("ids") %>%
-  pheatmap(.,cluster_rows = F,cluster_cols = F)
+  pheatmap(., cluster_rows = F, cluster_cols = F)
 ```
@@
 # Gold standard data
-dup = data %>% 
+dup <- data %>%
   arrange(scan_id, scorer) %>%
   filter(scan_id %in% dups$original_ssid) %>%
-  group_by(scorer,scan_id) %>%
-  mutate(n=row_number())%>%
-  select(scorer, scan_id,n ,bronchiectasis_rul:atelectasis___6) %>%
-  ungroup() %>% pivot_longer(bronchiectasis_rul:atelectasis___6,
+  group_by(scorer, scan_id) %>%
+  mutate(n = row_number()) %>%
+  select(scorer, scan_id, n, bronchiectasis_rul:atelectasis___6) %>%
+  ungroup() %>%
+  pivot_longer(bronchiectasis_rul:atelectasis___6,
     names_to = c("feature", "location"),
@@
 dup$location <- factor(dup$location,
-                       levels = c("1", "2", "3", "4", "5", "6", "ling", "lll", 
-                                  "lul", "rll", "rml", "rul"),
-                       labels = c("RUL", "RML", "RLL", "LUL", "Ling", "LLL", 
-                                  "Ling", "LLL", "LUL", "RLL", "RML", "RUL")
+  levels = c(
+    "1", "2", "3", "4", "5", "6", "ling", "lll",
+    "lul", "rll", "rml", "rul"
+  ),
+  labels = c(
+    "RUL", "RML", "RLL", "LUL", "Ling", "LLL",
+    "Ling", "LLL", "LUL", "RLL", "RML", "RUL"
+  )
 )
@@
   pivot_wider(names_from = n, values_from = A:C) %>%
-  arrange(feature,location)
+  arrange(feature, location)
 ```
@@
 ```{r}
-dup_bronch = dup %>% filter(feature=="Bronchiectasis")
-A = kappa2(dup_bronch[,c("A_1","A_2")])
-B = kappa2(dup_bronch[,c("B_1","B_2")])
-C = kappa2(dup_bronch[,c("C_1","C_2")])
+dup_bronch <- dup %>% filter(feature == "Bronchiectasis")
+A <- kappa2(dup_bronch[, c("A_1", "A_2")])
+B <- kappa2(dup_bronch[, c("B_1", "B_2")])
+C <- kappa2(dup_bronch[, c("C_1", "C_2")])
 # Plot
 dup_bronch %>%
-  select(scan_id,feature, location,A_1:C_2) %>%
+  select(scan_id, feature, location, A_1:C_2) %>%
   pivot_longer(A_1:C_2) %>%
@@
 ```{r}
-dup_other = dup %>% filter(feature!="Bronchiectasis")
+dup_other <- dup %>% filter(feature != "Bronchiectasis")
 # Check agreement with GS column
-dup_other$A = factor(dup_other$A_1 == dup_other$A_2,levels = c(F,T),labels = c("Disagree","Agree"))
-dup_other$B = factor(dup_other$B_1 == dup_other$B_2,levels = c(F,T),labels = c("Disagree","Agree"))
-dup_other$C = factor(dup_other$C_1 == dup_other$C_2,levels = c(F,T),labels = c("Disagree","Agree"))
+dup_other$A <- factor(dup_other$A_1 == dup_other$A_2, levels = c(F, T), labels = c("Disagree", "Agree"))
+dup_other$B <- factor(dup_other$B_1 == dup_other$B_2, levels = c(F, T), labels = c("Disagree", "Agree"))
+dup_other$C <- factor(dup_other$C_1 == dup_other$C_2, levels = c(F, T), labels = c("Disagree", "Agree"))
 # Plot
 dup_other %>%
-  select(scan_id,feature, location, A:C) %>%
+  select(scan_id, feature, location, A:C) %>%
   pivot_longer(A:C) %>%
@@
   labs(fill = "", x = "Scorer", y = "Location") +
-  facet_grid(rows = vars(feature),cols = vars(scan_id))
+  facet_grid(rows = vars(feature), cols = vars(scan_id))
 # Tables
 dup_other %>%
-  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "drop") %>%
-  kable(.,digits=3,caption = "Proportion of agreement with themselves")
+  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
+  kable(., digits = 3, caption = "Proportion of agreement with themselves")
 dup_other %>%
   group_by(scan_id) %>%
-  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "drop") %>%
-  kable(.,digits=3,caption = "Proportion of agreement with themselves")
+  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
+  kable(., digits = 3, caption = "Proportion of agreement with themselves")
 dup_other %>%
   group_by(feature) %>%
-  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "drop") %>%
-  kable(.,digits=3,caption = "Proportion of agreement with themselves")
+  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
+  kable(., digits = 3, caption = "Proportion of agreement with themselves")
 dup_other %>%
-  group_by(scan_id,feature) %>%
-  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "drop") %>%
-  kable(.,digits=3,caption = "Proportion of agreement with themselves")
+  group_by(scan_id, feature) %>%
+  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
+  kable(., digits = 3, caption = "Proportion of agreement with themselves")
 # Heatmaps
@@
   group_by(scan_id) %>%
-  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "keep") %>%
+  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
   column_to_rownames("scan_id") %>%
-  pheatmap(.,cluster_rows = F,cluster_cols = F)
+  pheatmap(., cluster_rows = F, cluster_cols = F)
 dup_other %>%
   group_by(feature) %>%
-  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "keep") %>%
+  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
   column_to_rownames("feature") %>%
-  pheatmap(.,cluster_rows = F,cluster_cols = F)
+  pheatmap(., cluster_rows = F, cluster_cols = F)
 dup_other %>%
-  group_by(scan_id,feature) %>%
-  summarise(across(A:C,~mean(.x == "Agree",na.rm = T)),.groups = "keep") %>%
-  unite(ids,scan_id,feature,na.rm = T,sep = " ") %>%
+  group_by(scan_id, feature) %>%
+  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
+  unite(ids, scan_id, feature, na.rm = T, sep = " ") %>%
   column_to_rownames("ids") %>%
-  pheatmap(.,cluster_rows = F,cluster_cols = F)
+  pheatmap(., cluster_rows = F, cluster_cols = F)
 ```
+
+# Singleton scans
+
+```{r}
+# Gold standard data
+df <- data %>%
+  arrange(scan_id, scorer) %>%
+  filter(!(scan_id %in% dups$original_ssid) &
+    !(scan_id %in% training$Cffid.Scoring.Set)) %>%
+  select(scorer, scan_id, bronchiectasis_rul:atelectasis___6) %>%
+  pivot_longer(bronchiectasis_rul:atelectasis___6,
+    names_to = c("feature", "location"),
+    names_pattern = "(.*)_(.*)$"
+  ) %>%
+  mutate(
+    feature = gsub("__", "", feature),
+    feature = str_to_title(gsub("_", " ", feature))
+  )
+df$location <- factor(df$location,
+  levels = c(
+    "1", "2", "3", "4", "5", "6", "ling", "lll",
+    "lul", "rll", "rml", "rul"
+  ),
+  labels = c(
+    "RUL", "RML", "RLL", "LUL", "Ling", "LLL",
+    "Ling", "LLL", "LUL", "RLL", "RML", "RUL"
+  )
+)
+# One column per scorer
+df <- df %>%
+  pivot_wider(names_from = scorer, values_from = value) %>%
+  arrange(feature, location)
+```
+
+## Bronciechtasis
+
+```{r}
+df_bronch <- df %>% filter(feature == "Bronchiectasis")
+fk <- kappam.fleiss(df_bronch[, c("A", "B", "C")])
+# Plot
+df_bronch %>%
+  select(scan_id, feature, location, A:C) %>%
+  pivot_longer(A:C) %>%
+  ggplot(., aes(x = name, y = location, fill = value)) +
+  geom_tile(color = "black") +
+  theme_bw() +
+  labs(fill = "", x = "Scorer", y = "Location") +
+  custom_colors +
+  facet_wrap(~scan_id)
+```
+
+Overall Fleiss' Kappa for all 15 singleton scans was `r round(fk$value,3)` (p = `r format.pval(fk$p.value,eps = 0.001,digits = 3)`).
+
+## Other features
+
+```{r}
+df_other <- df %>% filter(feature != "Bronchiectasis")
+# Check agreement with GS column
+df_other[, c("A", "B", "C")] <- lapply(df_other[, c("A", "B", "C")], factor,
+  levels = c("Unchecked", "Checked"),
+  labels = c("Absent", "Present")
+)
+# Three-way agreement
+df_other$`A vs. B` <- factor(df_other$A == df_other$B, levels = c(F, T), labels = c("Disagree", "Agree"))
+df_other$`A vs. C` <- factor(df_other$A == df_other$C, levels = c(F, T), labels = c("Disagree", "Agree"))
+df_other$`B vs. C` <- factor(df_other$B == df_other$C, levels = c(F, T), labels = c("Disagree", "Agree"))
+# Plot
+df_other %>%
+  select(scan_id, feature, location, `A vs. B`:`B vs. C`) %>%
+  pivot_longer(`A vs. B`:`B vs. C`) %>%
+  ggplot(., aes(x = name, y = location, fill = value)) +
+  geom_tile(color = "black") +
+  theme_bw() +
+  labs(fill = "", x = "Scorer", y = "Location") +
+  facet_grid(rows = vars(feature), cols = vars(scan_id))
+# Tables
+df_other %>%
+  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
+  kable(., digits = 3, caption = "Proportion of agreement with themselves")
+df_other %>%
+  group_by(scan_id) %>%
+  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
+  kable(., digits = 3, caption = "Proportion of agreement with themselves")
+df_other %>%
+  group_by(feature) %>%
+  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
+  kable(., digits = 3, caption = "Proportion of agreement with themselves")
+df_other %>%
+  group_by(scan_id, feature) %>%
+  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
+  kable(., digits = 3, caption = "Proportion of agreement with themselves")
+# Heatmaps
+df_other %>%
+  group_by(scan_id) %>%
+  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
+  column_to_rownames("scan_id") %>%
+  pheatmap(., cluster_rows = F, cluster_cols = F)
+df_other %>%
+  group_by(feature) %>%
+  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
+  column_to_rownames("feature") %>%
+  pheatmap(., cluster_rows = F, cluster_cols = F)
+df_other %>%
+  group_by(scan_id, feature) %>%
+  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
+  unite(ids, scan_id, feature, na.rm = T, sep = " ") %>%
+  column_to_rownames("ids") %>%
+  pheatmap(., cluster_rows = F, cluster_cols = F)
+```
+
+# Comparison of scores between groups

### 2023-05-18 Christine Chan/CF Registry/cf_registry_analysis.qmd
added logistic regression to CFRD registry data

@@
 # All data
-encounter <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
+encounter <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
                       na.strings = c("", "N/A")
 )
-annual <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
+annual <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
                    na.strings = c("", "N/A")
@@
 ```
+
+# Logistic analyses
+
+Any participants with T1D or T2D were excluded from these analyses.
+
+```{r}
+# Import and clean data
+annual <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
+                   na.strings = c("", "N/A")
+)
+annual = annual %>% filter(cfrd_status_annual != 4,cfrd_status_annual != 5)
+annual$cfrd_status_annual = factor(annual$cfrd_status_annual,levels = 1:3,
+                                   labels = c("CFRD-","CFRD-","CFRD+"))
+
+demogs = read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/CFF19_DemogCFDiag_Del1.csv")
+demogs = demogs %>% select(eDWID,Gender,MutClass)
+
+annual = left_join(annual,demogs,by = join_by(eDWID))
+
+encounter <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
+                      na.strings = c("", "N/A")
+)
+# mods = encounter %>% group_by(eDWID,reviewyear) %>%
+#   summarise(across(c(Vx770,Vx809comb),~max(.x,na.rm = T)),.groups = "drop")
+# Analysis function
+log_reg = function(year = 2005){
+  # Clean data
+  df = annual %>% 
+    filter(ReviewYear==year) %>%
+    select(eDWID,Age_YrEnd,Gender,MutClass,A_IsOnEnzymes,A_FEV1_pct_predicted,
+           A_bmivalue,A_bmipercentile,A_height,A_heightpercentile,
+           NumPulmExacerbation,A_feeding2,A_pseudomonasaeruginosa,
+           A_fungalyeast1,A_pulmonarycomplications1,A_mycocultureresults1,
+           A_corticosteroids1,cfrd_status_annual)
+  # Format columns
+  df$MutClass = factor(df$MutClass,levels = c("1-3","4-5","Oth","N/A"),
+                       labels = c("1-3","4-5","Oth",NA))
+  df$NumPulmExacerbation=cut(df$NumPulmExacerbation,labels = c("0","1","2",">=3"),
+                             breaks = c(0,1,2,3,Inf),right = F)
+  df$A_feeding2 = factor(df$A_feeding2,levels = 0:1,labels = c("No","Yes"))
+  df$A_IsOnEnzymes = factor(df$A_IsOnEnzymes,levels = 0:1,labels = c("No","Yes"))
+  df$A_FEV1_pct_predicted = cut(df$A_FEV1_pct_predicted,breaks = c(0,40,80,100,Inf),right = F)
+  df$A_corticosteroids1 = factor(df$A_corticosteroids1,levels = 0:1,labels = c("No","Yes"))
+  df$A_fungalyeast1 = factor(df$A_fungalyeast1,levels = 0:1,labels = c("No","Yes"))
+  df$A_pulmonarycomplications1 = factor(df$A_pulmonarycomplications1,levels = 0:1,labels = c("No","Yes"))
+  df$A_mycocultureresults1 = factor(df$A_mycocultureresults1,levels = 0:1,labels = c("No","Yes"))
+  # Models
+  base_mod = as.formula("cfrd_status_annual~Gender+Age_YrEnd+A_IsOnEnzymes+MutClass+
+                   A_FEV1_pct_predicted+NumPulmExacerbation+A_feeding2")
+  adult_mod = update(base_mod,.~.+A_bmivalue+A_height)
+  peds_mod = update(base_mod,.~.+A_bmipercentile+A_heightpercentile)
+  # Simple models
+  adult_mod1 = glm(adult_mod,data = df,family = "binomial")
+  peds_mod1 = glm(peds_mod,data = df,family = "binomial")
+  # Add bugs
+  # P. aeruginosa
+  adult_mod2 = glm(update(adult_mod,.~.+A_pseudomonasaeruginosa),data = df,family = "binomial")
+  peds_mod2 = glm(update(peds_mod,.~.+A_pseudomonasaeruginosa),data = df,family = "binomial")
+  # Aspergillus
+  adult_mod3 = glm(update(adult_mod,.~.+A_fungalyeast1),data = df,family = "binomial")
+  peds_mod3 = glm(update(peds_mod,.~.+A_fungalyeast1),data = df,family = "binomial")
+  # ABPA
+  adult_mod4 = glm(update(adult_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
+  peds_mod4 = glm(update(peds_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
+  # Mycobacterium
+  adult_mod5 = glm(update(adult_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
+  peds_mod5 = glm(update(peds_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
+  # Print results
+  cat("\n")
+  cat("### Base Models")
+  cat("\n")
+  cat("#### Adult")
+  print(forest_model(adult_mod1))
+  cat("\n")
+  cat("#### Peds")
+  print(forest_model(peds_mod1))
+  cat("\n")
+  cat("\n")
+  cat("### P. aeruginosa")
+  cat("\n")
+  cat("#### Adult")
+  print(forest_model(adult_mod2))
+  cat("\n")
+  cat("#### Peds")
+  print(forest_model(peds_mod2))
+  cat("\n")
+  cat("\n")
+  cat("### Aspergillus")
+  cat("\n")
+  cat("#### Adult")
+  print(forest_model(adult_mod3))
+  cat("\n")
+  cat("#### Peds")
+  print(forest_model(peds_mod3))
+  cat("\n")
+  cat("\n")
+  cat("### ABPA")
+  cat("\n")
+  cat("#### Adult")
+  print(forest_model(adult_mod4))
+  cat("\n")
+  cat("#### Peds")
+  print(forest_model(peds_mod4))
+  cat("\n")
+  cat("\n")
+  cat("### Mycobacterium")
+  cat("\n")
+  cat("#### Adult")
+  print(forest_model(adult_mod5))
+  cat("\n")
+  cat("#### Peds")
+  print(forest_model(peds_mod5))
+  cat("\n")
+}
+```
+
+## 2005
+
+```{r}
+log_reg(2005)
+```
+
+# Questions
+
+1. What variables do we use for B. Cepacia and liver disease?
+2. Do we have CFTR modulator data aside from `Vx770` and `Vx809comb`?

### 2023-05-18 Christine Chan/CF Registry/cf_registry_analysis.qmd
removed cfrd registry analysis for now

--- 2022-12-27 Christine Chan/CF Registry/cf_registry_analysis.qmd
@@
 author: "Tim Vigers"
-date: "`r format(Sys.time(), '%d %B %Y')`"
+date: "today"
+date-format: long
 format:
   html:
+    toc: true
+    toc-depth: 5
+    toc-float: true
     code-fold: true
     self-contained: true
+    fig-cap-location: top
     page-layout: full
@@
 library(tidyverse)
-library(table1)
-library(pROC)
+library(forestmodel)
 ```
 
+# Logistic analyses
+
+Any participants with T1D or T2D were excluded from these analyses.
+
 ```{r}
-# All data
-encounter <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
+# Import and clean data
+annual <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
                    na.strings = c("", "N/A")
 )
-annual <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
+annual = annual %>% filter(cfrd_status_annual != 4,cfrd_status_annual != 5)
+annual$cfrd_status_annual = factor(annual$cfrd_status_annual,levels = 1:3,
+                                   labels = c("CFRD-","CFRD-","CFRD+"))
+
+demogs = read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/CFF19_DemogCFDiag_Del1.csv")
+demogs = demogs %>% select(eDWID,Gender,MutClass)
+
+annual = left_join(annual,demogs,by = join_by(eDWID))
+
+encounter <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
                       na.strings = c("", "N/A")
 )
-encounter <- encounter %>%
-  arrange(eDWID, reviewyear, encounternum)
-annual = annual %>% arrange(eDWID,ReviewYear)
-# Add age
-ages <- annual %>% 
-  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd)) %>%
-  rename(reviewyear = ReviewYear) %>%
-  group_by(eDWID,reviewyear) %>%
-  summarise(`Age in 2018` = as.numeric(names(sort(table(`Age in 2018`), 
-                                                  decreasing = TRUE)[1])),
-            Age = Age_YrEnd[1],.groups = "drop")
-encounter = left_join(encounter,ages,by = c("eDWID", "reviewyear"))
-encounter$`Age Group 2018` = cut(encounter$`Age in 2018`,
-                            c(-Inf,10,19,31,Inf),
-                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
-                            right=F)
-encounter$`Age Group` = cut(encounter$Age,
-                            c(-Inf,10,19,31,Inf),
-                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
-                            right=F)
-# For those with CFRD by OGTT, find first visit with diagnosis
-first_cfrd_ogtt <- encounter %>% 
-  filter(twohour_bloodglucose >= 200) %>%
-  group_by(eDWID) %>%
-  filter(row_number()==1) %>%
-  select(eDWID, reviewyear, `Age in 2018`) %>%
-  rename(first_cfrd_ogtt = reviewyear)
-# For those with CFRD by registry criteria, find first visit with diagnosis
-first_cfrd_cffpr <- encounter %>% 
-  filter(cfrd_status==3) %>%
-  group_by(eDWID) %>%
-  filter(row_number()==1) %>%
-  select(eDWID, reviewyear, `Age in 2018`) %>%
-  rename(first_cfrd_cffpr = reviewyear)
-# Both
-first_cfrd_both <- encounter %>% 
-  filter(cfrd_status==3|twohour_bloodglucose >= 200) %>%
-  group_by(eDWID) %>%
-  filter(row_number()==1) %>%
-  select(eDWID, reviewyear, `Age in 2018`) %>%
-  rename(first_cfrd_both = reviewyear)
-# Add to encounter level data
-encounter <- left_join(encounter, first_cfrd_ogtt, by = c("eDWID", "Age in 2018"))
-encounter <- left_join(encounter, first_cfrd_cffpr, by = c("eDWID", "Age in 2018"))
-encounter <- left_join(encounter, first_cfrd_both,by = c("eDWID", "Age in 2018"))
-# Fill down CFRD Yes/No by year
-encounter <- encounter %>%
-  group_by(eDWID) %>%
-  mutate(
-    cfrd_ogtt = case_when(
-      reviewyear < first_cfrd_ogtt ~ "No",
-      reviewyear >= first_cfrd_ogtt ~ "Yes"
-    ),
-    cfrd_cffpr = case_when(
-      reviewyear < first_cfrd_cffpr ~ "No",
-      reviewyear >= first_cfrd_cffpr ~ "Yes"
-    ),
-    cfrd_both = case_when(
-      reviewyear < first_cfrd_both ~ "No",
-      reviewyear >= first_cfrd_both ~ "Yes"
-    )
-  ) %>%
-  ungroup()
-encounter$cfrd_ogtt[is.na(encounter$cfrd_ogtt)] <- "No"
-encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
-encounter$cfrd_both[is.na(encounter$cfrd_both)] <- "No"
-# Ordered factors
-encounter$cfrd_ogtt = factor(encounter$cfrd_ogtt,levels = c("No","Yes"),ordered = T)
-encounter$cfrd_cffpr = factor(encounter$cfrd_cffpr,levels = c("No","Yes"),ordered = T)
-encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
-# Remove those missing age
-encounter = encounter %>% filter(!is.na(Age))
+mods = encounter %>% group_by(eDWID,reviewyear) %>%
+  summarise(across(c(Vx770,Vx809comb),~max(.x,na.rm = T)),.groups = "drop")
+annual = left_join(annual,mods)
+# Analysis function
+log_reg = function(year = 2005){
+  # Clean data
+  df = annual %>% 
+    filter(ReviewYear==year) %>%
+    select(eDWID,Age_YrEnd,Gender,MutClass,A_IsOnEnzymes,A_FEV1_pct_predicted,
+           A_bmivalue,A_bmipercentile,A_height,A_heightpercentile,
+           NumPulmExacerbation,A_feeding2,A_pseudomonasaeruginosa,
+           A_fungalyeast1,A_pulmonarycomplications1,A_mycocultureresults1,
+           A_corticosteroids1,cfrd_status_annual)
+  # Format columns
+  df$MutClass = factor(df$MutClass,levels = c("1-3","4-5","Oth","N/A"),
+                       labels = c("1-3","4-5","Oth",NA))
+  df$NumPulmExacerbation=cut(df$NumPulmExacerbation,labels = c("0","1","2",">=3"),
+                             breaks = c(0,1,2,3,Inf),right = F)
+  df$A_feeding2 = factor(df$A_feeding2,levels = 0:1,labels = c("No","Yes"))
+  df$A_IsOnEnzymes = factor(df$A_IsOnEnzymes,levels = 0:1,labels = c("No","Yes"))
+  df$A_FEV1_pct_predicted = cut(df$A_FEV1_pct_predicted,breaks = c(0,40,80,100,Inf),right = F)
+  df$A_corticosteroids1 = factor(df$A_corticosteroids1,levels = 0:1,labels = c("No","Yes"))
+  df$A_fungalyeast1 = factor(df$A_fungalyeast1,levels = 0:1,labels = c("No","Yes"))
+  df$A_pulmonarycomplications1 = factor(df$A_pulmonarycomplications1,levels = 0:1,labels = c("No","Yes"))
+  df$A_mycocultureresults1 = factor(df$A_mycocultureresults1,levels = 0:1,labels = c("No","Yes"))
+  # Models
+  base_mod = as.formula("cfrd_status_annual~Gender+Age_YrEnd+A_IsOnEnzymes+MutClass+
+                   A_FEV1_pct_predicted+NumPulmExacerbation+A_feeding2")
+  adult_mod = update(base_mod,.~.+A_bmivalue+A_height)
+  peds_mod = update(base_mod,.~.+A_bmipercentile+A_heightpercentile)
+  # Simple models
+  adult_mod1 = glm(adult_mod,data = df,family = "binomial")
+  peds_mod1 = glm(peds_mod,data = df,family = "binomial")
+  # Add bugs
+  # P. aeruginosa
+  adult_mod2 = glm(update(adult_mod,.~.+A_pseudomonasaeruginosa),data = df,family = "binomial")
+  peds_mod2 = glm(update(peds_mod,.~.+A_pseudomonasaeruginosa),data = df,family = "binomial")
+  # Aspergillus
+  adult_mod3 = glm(update(adult_mod,.~.+A_fungalyeast1),data = df,family = "binomial")
+  peds_mod3 = glm(update(peds_mod,.~.+A_fungalyeast1),data = df,family = "binomial")
+  # ABPA
+  adult_mod4 = glm(update(adult_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
+  peds_mod4 = glm(update(peds_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
+  # Mycobacterium
+  adult_mod5 = glm(update(adult_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
+  peds_mod5 = glm(update(peds_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
+  # Print results
+  cat("\n")
+  cat("### Base Models")
+  cat("\n")
+  cat("\n")
+  cat("#### Adult")
+  cat("\n")
+  cat("\n")
+  print(forest_model(adult_mod1))
+  cat("\n")
+  cat("\n")
+  cat("#### Peds")
+  cat("\n")
+  cat("\n")
+  print(forest_model(peds_mod1))
+  cat("\n")
+  cat("\n")
+  cat("### P. aeruginosa")
+  cat("\n")
+  cat("\n")
+  cat("#### Adult")
+  cat("\n")
+  cat("\n")
+  print(forest_model(adult_mod2))
+  cat("\n")
+  cat("\n")
+  cat("#### Peds")
+  cat("\n")
+  cat("\n")
+  print(forest_model(peds_mod2))
+  cat("\n")
+  cat("\n")
+  cat("### Aspergillus")
+  cat("\n")
+  cat("\n")
+  cat("#### Adult")
+  cat("\n")
+  cat("\n")
+  print(forest_model(adult_mod3))
+  cat("\n")
+  cat("\n")
+  cat("#### Peds")
+  cat("\n")
+  cat("\n")
+  print(forest_model(peds_mod3))
+  cat("\n")
+  cat("\n")
+  cat("### ABPA")
+  cat("\n")
+  cat("\n")
+  cat("#### Adult")
+  cat("\n")
+  cat("\n")
+  print(forest_model(adult_mod4))
+  cat("\n")
+  cat("\n")
+  cat("#### Peds")
+  cat("\n")
+  cat("\n")
+  print(forest_model(peds_mod4))
+  cat("\n")
+  cat("\n")
+  cat("### Mycobacterium")
+  cat("\n")
+  cat("\n")
+  cat("#### Adult")
+  cat("\n")
+  cat("\n")
+  print(forest_model(adult_mod5))
+  cat("\n")
+  cat("\n")
+  cat("#### Peds")
+  cat("\n")
+  cat("\n")
+  print(forest_model(peds_mod5))
+  cat("\n")
+}
 ```
 
-```{r}
-#| label: tbl-incidence_cffpr
-#| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR Alone"
-# For each year, pull everyone with and without CFRD that year
-year_inc = function(df=encounter,year,py = 1000){
-  # Filter and count
-  inc = df %>%  
-    filter(reviewyear==year,is.na(first_cfrd_cffpr) | first_cfrd_cffpr >= year) %>%
-    group_by(eDWID) %>% filter(row_number() == 1) %>% 
-    select(eDWID,`Age Group`,cfrd_cffpr)
-  inc = inc %>% group_by(cfrd_cffpr) %>% count(`Age Group`,.drop = F)
-  inc = inc[!is.na(inc$`Age Group`),]
-  # Calculations
-  new_cases = sum(inc$n[inc$cfrd_cffpr=="Yes"])
-  total_at_risk = sum(inc$n)
-  total_inc = new_cases/(total_at_risk/py)
+## 2005
 
-  new_cases_10 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group` == "< 10"])
-  at_risk_10 = sum(inc$n[inc$`Age Group` == "< 10"])
-  inc_10 = new_cases_10/(at_risk_10/py)
-
-  new_cases_10_18 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group` == "10 - 18"])
-  at_risk_10_18 = sum(inc$n[inc$`Age Group` == "10 - 18"])
-  inc_10_18 = new_cases_10_18/(at_risk_10_18/py)
-
-  new_cases_19_30 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group` == "19 - 30"])
-  at_risk_19_30 = sum(inc$n[inc$`Age Group` == "19 - 30"])
-  inc_19_30 = new_cases_19_30/(at_risk_19_30/py)
-
-  new_cases_30 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group` == "> 30"])
-  at_risk_30 = sum(inc$n[inc$`Age Group` == "> 30"])
-  inc_30 = new_cases_30/(at_risk_30/py)
-
-  # Nice results
-  list(`Year` = year,
-       `Total New Cases` = new_cases,
-       `Total At Risk` = total_at_risk,
-       `Total Incidence` = total_inc,
-
-       `Total New Cases Age < 10` = new_cases_10,
-       `Total At Risk Age < 10` = at_risk_10,
-       `Total Incidence Age < 10` = inc_10,
-
-       `Total New Cases Age 10-18` = new_cases_10_18,
-       `Total At Risk Age 10-18` = at_risk_10_18,
-       `Total Incidence Age 10-18` = inc_10_18,
-
-       `Total New Cases Age 19-30` = new_cases_19_30,
-       `Total At Risk Age 19-30` = at_risk_19_30,
-       `Total Incidence Age 19-30` = inc_19_30,
-
-       `Total New Cases Age > 30` = new_cases_30,
-       `Total At Risk Age > 30` = at_risk_30,
-       `Total Incidence Age > 30` = inc_30)
-}
-incidence = lapply(2003:2018, function(y){year_inc(year = y)})
-incidence = data.frame(do.call(rbind,incidence),check.names = F)
-incidence = data.frame(lapply(incidence,as.numeric),check.names = F)
-kable(incidence,digits = 3)
+```{r results='asis',fig.height=10,fig.width=10}
+log_reg(2005)
 ```
 
-In @tbl-incidence_cffpr, incidence per year is calculated as the number of new CFRD diagnoses (by registry definition alone) per 1,000 person-years of those at risk, excluding previously identified cases.
+# Questions
 
-```{r}
-#| label: tbl-prevalence_cffpr_year
-#| tbl-cap: "Yearly Prevalence of CFRD by Age Group Based on CFFPR Alone"
-encounter %>% group_by(eDWID,reviewyear,`Age Group`) %>% 
-  summarise(cfrd = max(cfrd_cffpr),.groups = "drop") %>%
-  group_by(reviewyear,`Age Group`) %>%
-  summarise(Prevalence = mean(as.numeric(cfrd)-1,na.rm = T)*100,.groups = "drop") %>%
-  pivot_wider(names_from = `Age Group`,values_from = Prevalence) %>%
-  kable(.,digits = 3)
-```
-
-```{r}
-#| label: tbl-prevalence_cffpr
-#| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR Alone"
-encounter %>% group_by(eDWID,reviewyear) %>% 
-  summarise(cfrd = max(cfrd_cffpr),.groups = "drop") %>%
-  group_by(reviewyear) %>%
-  summarise(Prevalence = mean(as.numeric(cfrd)-1,na.rm = T)*100,.groups = "drop") %>%
-  kable(.,digits = 3)
-```
-
-In @tbl-prevalence_cffpr_year and @tbl-prevalence_cffpr, prevalence is calculated as the total number of CFRD cases (by registry definition alone) seen in clinic divided by the total number of people seen in clinic in a given year.
-
-```{r}
-#| label: tbl-incidence_both
-#| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR and OGTT"
-# For each year, pull everyone with and without CFRD that year
-year_inc = function(df=encounter,year,py = 1000){
-  # Filter and count
-  inc = df %>% 
-    filter(reviewyear==year,is.na(first_cfrd_both) | first_cfrd_both >= year) %>%
-    group_by(eDWID) %>% filter(row_number() == 1) %>%
-    select(eDWID,reviewyear,encounternum,`Age Group`,
-           twohour_bloodglucose,cfrd_both,first_cfrd_both)
-  inc = inc %>% group_by(cfrd_both) %>% count(`Age Group`,.drop = F)
-  inc = inc[!is.na(inc$`Age Group`),]
-  # Calculations
-  new_cases = sum(inc$n[inc$cfrd_both=="Yes"])
-  total_at_risk = sum(inc$n)
-  total_inc = new_cases/(total_at_risk/py)
-
-  new_cases_10 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group` == "< 10"])
-  at_risk_10 = sum(inc$n[inc$`Age Group` == "< 10"])
-  inc_10 = new_cases_10/(at_risk_10/py)
-
-  new_cases_10_18 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group` == "10 - 18"])
-  at_risk_10_18 = sum(inc$n[inc$`Age Group` == "10 - 18"])
-  inc_10_18 = new_cases_10_18/(at_risk_10_18/py)
-
-  new_cases_19_30 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group` == "19 - 30"])
-  at_risk_19_30 = sum(inc$n[inc$`Age Group` == "19 - 30"])
-  inc_19_30 = new_cases_19_30/(at_risk_19_30/py)
-
-  new_cases_30 = sum(inc$n[inc$cfrd_both=="Yes" & inc$`Age Group` == "> 30"])
-  at_risk_30 = sum(inc$n[inc$`Age Group` == "> 30"])
-  inc_30 = new_cases_30/(at_risk_30/py)
-
-  # Nice results
-  list(`Year` = year,
-       `Total New Cases` = new_cases,
-       `Total At Risk` = total_at_risk,
-       `Total Incidence` = total_inc,
-
-       `Total New Cases Age < 10` = new_cases_10,
-       `Total At Risk Age < 10` = at_risk_10,
-       `Total Incidence Age < 10` = inc_10,
-
-       `Total New Cases Age 10-18` = new_cases_10_18,
-       `Total At Risk Age 10-18` = at_risk_10_18,
-       `Total Incidence Age 10-18` = inc_10_18,
-
-       `Total New Cases Age 19-30` = new_cases_19_30,
-       `Total At Risk Age 19-30` = at_risk_19_30,
-       `Total Incidence Age 19-30` = inc_19_30,
-
-       `Total New Cases Age > 30` = new_cases_30,
-       `Total At Risk Age > 30` = at_risk_30,
-       `Total Incidence Age > 30` = inc_30)
-}
-incidence = lapply(2003:2018, function(y){year_inc(year = y)})
-incidence = data.frame(do.call(rbind,incidence),check.names = F)
-incidence = data.frame(lapply(incidence,as.numeric),check.names = F)
-kable(incidence,digits = 3)
-```
-
-In @tbl-incidence_both, incidence per year is calculated as the number of new CFRD diagnoses (by either registry definition or OGTT, whichever happened first) per 1,000 person-years of those at risk, excluding previously identified cases.
-
-```{r}
-#| label: tbl-prevalence_both_year
-#| tbl-cap: "Yearly Prevalence of CFRD by Age Group Based on CFFPR and OGTT"
-encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
-encounter %>% group_by(eDWID,reviewyear,`Age Group`) %>% 
-  summarise(cfrd = max(cfrd_both),.groups = "drop") %>%
-  group_by(reviewyear,`Age Group`) %>%
-  summarise(`Prevalence` = mean(as.numeric(cfrd)-1,na.rm = T)*100,.groups = "drop") %>%
-  pivot_wider(names_from = `Age Group`,values_from = Prevalence)%>%
-  kable(.,digits = 3)
-```
-
-```{r}
-#| label: tbl-prevalence_both
-#| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR and OGTT"
-encounter %>% group_by(eDWID,reviewyear) %>% 
-  summarise(cfrd = max(cfrd_both),.groups = "drop") %>%
-  group_by(reviewyear) %>%
-  summarise(`Prevalence` = mean(as.numeric(cfrd)-1,na.rm = T)*100,.groups = "drop") %>%
-  kable(.,digits = 3)
-```
-
-In @tbl-prevalence_both_year and @tbl-prevalence_both, prevalence is calculated as the total number of CFRD cases (by either registry definition or OGTT) seen in clinic divided by the total number of people seen in clinic in a given year.
-
-```{r}
-#| label: tbl-screening_ogtt
-#| tbl-cap: "Yearly OGTT Screening Rates"
-year_screen = function(df=encounter,year){
-  # Filter and count
-  inc = df %>% 
-    filter(reviewyear==year,first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>%
-    mutate(ogtt = !is.na(ogttfast_bloodglucose) | !is.na(twohour_bloodglucose)) %>%
-    group_by(eDWID) %>%
-    summarise(ogtt = max(ogtt),`Age Group` = `Age Group`[1])
-  inc = inc %>% group_by(ogtt) %>% count(`Age Group`) %>% filter(`Age Group` != "< 10")
-  # Calculations
-  total_screened = sum(inc$n[inc$ogtt==1])
-  total_seen = sum(inc$n)
-  screen_rate = total_screened/total_seen * 100
-
-  screened_10_18 = inc$n[inc$ogtt==1 & inc$`Age Group` == "10 - 18"]
-  seen_10_18 = sum(inc$n[inc$`Age Group` == "10 - 18"])
-  rate_10_18 = screened_10_18/seen_10_18 *100
-
-  screened_19_30 = inc$n[inc$ogtt==1 & inc$`Age Group` == "19 - 30"]
-  seen_19_30 = sum(inc$n[inc$`Age Group` == "19 - 30"])
-  rate_19_30 = screened_19_30/seen_19_30 *100
-
-  screened_30 = inc$n[inc$ogtt==1 & inc$`Age Group` == "> 30"]
-  seen_30 = sum(inc$n[inc$`Age Group` == "> 30"])
-  rate_30 = screened_30/seen_30 *100
-
-  # Nice results
-  list(`Year` = year,
-       `Total Screened` = total_screened,
-       `Total Seen` = total_seen,
-       `Screen Rate (%)` = screen_rate,
-
-       `Total Screened Age 10-18` = screened_10_18,
-       `Total Seen Age 10-18` = seen_10_18,
-       `Screen Rate Age 10-18 (%)` = rate_10_18,
-
-       `Total Screened Age 19-30` = screened_19_30,
-       `Total Seen Age 19-30` = seen_19_30,
-       `Screen Rate Age 19-30 (%)` = rate_19_30,
-
-       `Total Screened Age > 30` = screened_30,
-       `Total Seen Age > 30` = seen_30,
-       `Screen Rate Age > 30 (%)` = rate_30)
-}
-screen = lapply(2003:2018, function(y){year_screen(year = y)})
-screen = data.frame(do.call(rbind,screen),check.names = F)
-screen = data.frame(lapply(screen,as.numeric),check.names = F)
-kable(screen,digits = 3)
-```
-
-```{r}
-#| label: tbl-screening_a1c
-#| tbl-cap: "Yearly HbA1c Screening Rates"
-year_screen = function(df=encounter,year){
-  # Filter and count
-  inc = df %>% 
-    filter(reviewyear==year,first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>%
-    mutate(a1c = !is.na(hgba1c)) %>%
-    group_by(eDWID) %>%
-    summarise(a1c = max(a1c),`Age Group` = `Age Group`[1])
-  inc = inc %>% group_by(a1c) %>% count(`Age Group`) %>% filter(`Age Group` != "< 10")
-  # Calculations
-  total_screened = sum(inc$n[inc$a1c==1])
-  total_seen = sum(inc$n)
-  screen_rate = total_screened/total_seen * 100
-
-  screened_10_18 = inc$n[inc$a1c==1 & inc$`Age Group` == "10 - 18"]
-  seen_10_18 = sum(inc$n[inc$`Age Group` == "10 - 18"])
-  rate_10_18 = screened_10_18/seen_10_18 *100
-
-  screened_19_30 = inc$n[inc$a1c==1 & inc$`Age Group` == "19 - 30"]
-  seen_19_30 = sum(inc$n[inc$`Age Group` == "19 - 30"])
-  rate_19_30 = screened_19_30/seen_19_30 *100
-
-  screened_30 = inc$n[inc$a1c==1 & inc$`Age Group` == "> 30"]
-  seen_30 = sum(inc$n[inc$`Age Group` == "> 30"])
-  rate_30 = screened_30/seen_30 *100
-
-  # Nice results
-  list(`Year` = year,
-       `Total Screened` = total_screened,
-       `Total Seen` = total_seen,
-       `Screen Rate (%)` = screen_rate,
-
-       `Total Screened Age 10-18` = screened_10_18,
-       `Total Seen Age 10-18` = seen_10_18,
-       `Screen Rate Age 10-18 (%)` = rate_10_18,
-
-       `Total Screened Age 19-30` = screened_19_30,
-       `Total Seen Age 19-30` = seen_19_30,
-       `Screen Rate Age 19-30 (%)` = rate_19_30,
-
-       `Total Screened Age > 30` = screened_30,
-       `Total Seen Age > 30` = seen_30,
-       `Screen Rate Age > 30 (%)` = rate_30)
-}
-screen = lapply(2003:2018, function(y){year_screen(year = y)})
-screen = data.frame(do.call(rbind,screen),check.names = F)
-screen = data.frame(lapply(screen,as.numeric),check.names = F)
-kable(screen,digits = 3)
-```
-
-In @tbl-screening_ogtt and @tbl-screening_a1c, people with CFRD (by registry definition) and those under age 10 were excluded.
-Screening rate is calculated as the number of people with an OGTT divided by the total number of people seen in clinic that year.
+1. What variables do we use for B. Cepacia and liver disease?
+2. Do we have CFTR modulator data aside from `Vx770` and `Vx809comb`?

### 2023-05-18 Christine Chan/CF Registry/cf_registry_analysis.qmd
CFRD registry logistic models mostly done

@@
                                    labels = c("CFRD-","CFRD-","CFRD+"))
-
 demogs = read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/CFF19_DemogCFDiag_Del1.csv")
 demogs = demogs %>% select(eDWID,Gender,MutClass)
-
 annual = left_join(annual,demogs,by = join_by(eDWID))
-
-encounter <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
-                      na.strings = c("", "N/A")
-)
-mods = encounter %>% group_by(eDWID,reviewyear) %>%
-  summarise(across(c(Vx770,Vx809comb),~max(.x,na.rm = T)),.groups = "drop")
-annual = left_join(annual,mods)
 # Analysis function
@@
            A_fungalyeast1,A_pulmonarycomplications1,A_mycocultureresults1,
-           A_corticosteroids1,cfrd_status_annual)
+           A_corticosteroids1,cfrd_status_annual,A_Vx770,A_VX809comb)
   # Format columns
@@
   df$A_mycocultureresults1 = factor(df$A_mycocultureresults1,levels = 0:1,labels = c("No","Yes"))
+  df$A_Vx770 = factor(df$A_Vx770,levels = 0:1,labels = c("No","Yes"))
+  df$A_VX809comb = factor(df$A_VX809comb,levels = 0:1,labels = c("No","Yes"))
+  # Remove all empty columns
+  df = df[,-which(colSums(is.na(df))==nrow(df))]
+  # Bugs
+  bugs = c("A_pseudomonasaeruginosa","A_fungalyeast1",
+           "A_pulmonarycomplications1","A_mycocultureresults1")
   # Models
-  base_mod = as.formula("cfrd_status_annual~Gender+Age_YrEnd+A_IsOnEnzymes+MutClass+
-                   A_FEV1_pct_predicted+NumPulmExacerbation+A_feeding2")
+  vars = colnames(df)[which(!colnames(df) %in% 
+                              c(bugs,"cfrd_status_annual","eDWID","A_bmivalue",
+                                "A_bmipercentile","A_height","A_heightpercentile"))]
+  base_mod = as.formula(paste0("cfrd_status_annual~",paste0(vars,collapse = "+")))
   adult_mod = update(base_mod,.~.+A_bmivalue+A_height)
@@
   peds_mod1 = glm(peds_mod,data = df,family = "binomial")
-  # Add bugs
-  # P. aeruginosa
-  adult_mod2 = glm(update(adult_mod,.~.+A_pseudomonasaeruginosa),data = df,family = "binomial")
-  peds_mod2 = glm(update(peds_mod,.~.+A_pseudomonasaeruginosa),data = df,family = "binomial")
-  # Aspergillus
-  adult_mod3 = glm(update(adult_mod,.~.+A_fungalyeast1),data = df,family = "binomial")
-  peds_mod3 = glm(update(peds_mod,.~.+A_fungalyeast1),data = df,family = "binomial")
-  # ABPA
-  adult_mod4 = glm(update(adult_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
-  peds_mod4 = glm(update(peds_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
-  # Mycobacterium
-  adult_mod5 = glm(update(adult_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
-  peds_mod5 = glm(update(peds_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
   # Print results
@@
   cat("\n")
+  if ("A_pseudomonasaeruginosa" %in% colnames(df)){
+    adult_mod2 = glm(update(adult_mod,.~.+A_pseudomonasaeruginosa),data = df,family = "binomial")
+    peds_mod2 = glm(update(peds_mod,.~.+A_pseudomonasaeruginosa),data = df,family = "binomial")
     cat("\n")
@@
     cat("\n")
+  }
+  if("A_fungalyeast1" %in% colnames(df)){
+    adult_mod3 = glm(update(adult_mod,.~.+A_fungalyeast1),data = df,family = "binomial")
+    peds_mod3 = glm(update(peds_mod,.~.+A_fungalyeast1),data = df,family = "binomial")
     cat("\n")
@@
     cat("\n")
+  }
+  if("A_pulmonarycomplications1" %in% colnames(df)){
+    adult_mod4 = glm(update(adult_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
+    peds_mod4 = glm(update(peds_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
     cat("\n")
@@
     cat("\n")
+  }
+  if("A_mycocultureresults1" %in% colnames(df)){
+    adult_mod5 = glm(update(adult_mod,.~.+A_mycocultureresults1),data = df,family = "binomial")
+    peds_mod5 = glm(update(peds_mod,.~.+A_mycocultureresults1),data = df,family = "binomial")
     cat("\n")
@@
   }
+}
 ```

### 2023-05-19 Christine Chan/CF Registry/cf_registry_analysis.qmd
Update cf_registry_analysis.qmd

--- 2023-05-18 Christine Chan/CF Registry/cf_registry_analysis.qmd
@@
 
-Any participants with T1D or T2D were excluded from these analyses.
+Any participants with T1D or T2D were excluded from these analyses. Variables that were entirely missing in a given year (e.g., `A_Vx770` and `A_VX809comb` in 2005) were excluded from those models.
 
@@
                              breaks = c(0,1,2,3,Inf),right = F)
-  df$A_feeding2 = factor(df$A_feeding2,levels = 0:1,labels = c("No","Yes"))
-  df$A_IsOnEnzymes = factor(df$A_IsOnEnzymes,levels = 0:1,labels = c("No","Yes"))
-  df$A_FEV1_pct_predicted = cut(df$A_FEV1_pct_predicted,breaks = c(0,40,80,100,Inf),right = F)
-  df$A_corticosteroids1 = factor(df$A_corticosteroids1,levels = 0:1,labels = c("No","Yes"))
-  df$A_fungalyeast1 = factor(df$A_fungalyeast1,levels = 0:1,labels = c("No","Yes"))
-  df$A_pulmonarycomplications1 = factor(df$A_pulmonarycomplications1,levels = 0:1,labels = c("No","Yes"))
-  df$A_mycocultureresults1 = factor(df$A_mycocultureresults1,levels = 0:1,labels = c("No","Yes"))
+  df$A_FEV1_pct_predicted = cut(df$A_FEV1_pct_predicted,
+                                breaks = c(0,40,80,100,Inf),right = F)
+  # Bugs
+  bugs = c("A_pseudomonasaeruginosa","A_fungalyeast1",
+           "A_pulmonarycomplications1","A_mycocultureresults1")
+  # Yes/No variables
+  yn_vars = c("A_feeding2","A_IsOnEnzymes","A_corticosteroids1",
+              "A_Vx770","A_VX809comb",bugs)
+
+  df[,yn_vars] = lapply(df[,yn_vars],factor,levels = 0:1,labels = c("No","Yes"))
+  # Remove all empty columns
+  empty = which(colSums(is.na(df))==nrow(df))
+  if(length(empty)>0){
+    df = df[,-empty]
+  }
   # Models
-  base_mod = as.formula("cfrd_status_annual~Gender+Age_YrEnd+A_IsOnEnzymes+MutClass+
-                   A_FEV1_pct_predicted+NumPulmExacerbation+A_feeding2")
+  vars = colnames(df)[which(!colnames(df) %in% 
+                              c(bugs,"cfrd_status_annual","eDWID","A_bmivalue",
+                                "A_bmipercentile","A_height","A_heightpercentile"))]
+  base_mod = as.formula(paste0("cfrd_status_annual~",paste0(vars,collapse = "+")))
   adult_mod = update(base_mod,.~.+A_bmivalue+A_height)
@@
 
+## 2013
+
+```{r results='asis',fig.height=10,fig.width=10}
+log_reg(2013)
+```
+
+## 2018
+
+```{r results='asis',fig.height=10,fig.width=10}
+log_reg(2018)
+```
+
 # Questions

### 2023-05-19 Emily Muther/body_image_trikafta.qmd
Update body_image_trikafta.qmd

--- 2023-03-25 Emily Muther/body_image_trikafta.qmd
@@
 author: "Tim Vigers"
-date: today
+date: "today"
 date-format: long
@@
 library(foreign)
+library(plotly)
+library(broom)
+library(report)
+library(sjPlot)
 # Read in data
-df = read.spss("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Body Image and Trikafta/Data_Raw/TrikaftaBodyImage3.6.23.sav",to.data.frame = T)
+df = read.spss("/Users/timvigers/Documents/Work/Vigers/CF/Emily Muther/Body Image and Trikafta/Data_Raw/TrikaftaBodyImage3.6.23.sav",to.data.frame = T)
 # Filter
@@
 
-# Figure
+# Mental health scores by FRS category
+
+## PHQ9
+
+### Odds of an elevated score
+
+Results are reported as odds ratios, with the FRS "neutral" category as the reference group.
 
 ```{r}
-# | label: fig-planets
-# | fig-cap: Planets
+mod = glm(PHQ9_elevated ~ FRS_Category,data = df,family = "binomial")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
 
+`r report(mod)`
+
+### Association with FRS category in those with elevated scores.
+
+Results are reported as expected counts, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod = glm(PHQ9_Total~ FRS_Category,data = df[df$PHQ9_elevated=="yes",],family = "poisson")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
 ```
+
+`r report(mod)`
+
+## GAD7
+
+### Odds of an elevated score
+
+Results are reported as odds ratios, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod = glm(GAD7_elevated ~ FRS_Category,data = df,family = "binomial")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+### Association with FRS category in those with elevated scores.
+
+Results are reported as expected counts, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod = glm(GAD7_Total~ FRS_Category,data = df[df$GAD7_elevated=="yes",],family = "poisson")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+# Mental health scores by BES
+
+## PHQ9
+
+### BE_APPEARANCE_MEAN
+
+#### Odds of an elevated score
+
+Results are reported as odds ratios for a one-unit change in the predictor.
+
+```{r}
+mod1 = glm(PHQ9_elevated ~ BE_APPEARANCE_MEAN,data = df,family = "binomial")
+res = tidy(mod1,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+#### Association with BE_APPEARANCE_MEAN in those with elevated scores.
+
+Results are reported as expected counts, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod2 = glm(PHQ9_Total~ BE_APPEARANCE_MEAN,data = df[df$PHQ9_elevated=="yes",],family = "poisson")
+res = tidy(mod2,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+### BE_ATTRIBUTION_MEAN
+
+#### Odds of an elevated score
+
+Results are reported as odds ratios for a one-unit change in the predictor.
+
+```{r}
+mod3 = glm(PHQ9_elevated ~ BE_ATTRIBUTION_MEAN,data = df,family = "binomial")
+res = tidy(mod3,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+#### Association with BE_ATTRIBUTION_MEAN in those with elevated scores.
+
+Results are reported as expected counts, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod4 = glm(PHQ9_Total~ BE_ATTRIBUTION_MEAN,data = df[df$PHQ9_elevated=="yes",],family = "poisson")
+res = tidy(mod4,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+### BE_WEIGHT_MEAN
+
+#### Odds of an elevated score
+
+Results are reported as odds ratios for a one-unit change in the predictor.
+
+```{r}
+mod5 = glm(PHQ9_elevated ~ BE_WEIGHT_MEAN,data = df,family = "binomial")
+res = tidy(mod5,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+#### Association with BE_WEIGHT_MEAN in those with elevated scores.
+
+Results are reported as expected counts, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod6 = glm(PHQ9_Total~ BE_WEIGHT_MEAN,data = df[df$PHQ9_elevated=="yes",],family = "poisson")
+res = tidy(mod6,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+### Forest plot of logistic regression results
+
+```{r}
+plot_models(mod1,mod3,mod5) + theme_bw() + theme(legend.position = "none")
+```
+
+## GAD7
+
+### BE_APPEARANCE_MEAN
+
+#### Odds of an elevated score
+
+Results are reported as odds ratios for a one-unit change in the predictor.
+
+```{r}
+mod1 = glm(GAD7_elevated ~ BE_APPEARANCE_MEAN,data = df,family = "binomial")
+res = tidy(mod1,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+#### Association with BE_APPEARANCE_MEAN in those with elevated scores.
+
+Results are reported as expected counts, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod2 = glm(GAD7_Total~ BE_APPEARANCE_MEAN,data = df[df$GAD7_elevated=="yes",],family = "poisson")
+res = tidy(mod2,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+### BE_ATTRIBUTION_MEAN
+
+#### Odds of an elevated score
+
+Results are reported as odds ratios for a one-unit change in the predictor.
+
+```{r}
+mod3 = glm(GAD7_elevated ~ BE_ATTRIBUTION_MEAN,data = df,family = "binomial")
+res = tidy(mod3,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+#### Association with BE_ATTRIBUTION_MEAN in those with elevated scores.
+
+Results are reported as expected counts, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod4 = glm(GAD7_Total~ BE_ATTRIBUTION_MEAN,data = df[df$GAD7_elevated=="yes",],family = "poisson")
+res = tidy(mod4,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+### BE_WEIGHT_MEAN
+
+#### Odds of an elevated score
+
+Results are reported as odds ratios for a one-unit change in the predictor.
+
+```{r}
+mod5 = glm(GAD7_elevated ~ BE_WEIGHT_MEAN,data = df,family = "binomial")
+res = tidy(mod5,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+#### Association with BE_WEIGHT_MEAN in those with elevated scores.
+
+Results are reported as expected counts, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod6 = glm(GAD7_Total~ BE_WEIGHT_MEAN,data = df[df$GAD7_elevated=="yes",],family = "poisson")
+res = tidy(mod6,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+### Forest plot of logistic regression results
+
+```{r}
+plot_models(mod1,mod3,mod5) + theme_bw() + theme(legend.position = "none")
+```

### 2023-05-19 Stacey Martiniano/ntm_agreement.qmd
Update ntm_agreement.qmd

@@
 library(pheatmap)
+library(arsenal)
 # Import data
@@
   column_to_rownames("scan_id") %>%
-  pheatmap(., cluster_rows = F, cluster_cols = F)
+  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
+           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
 gs_other %>%
@@
   column_to_rownames("feature") %>%
-  pheatmap(., cluster_rows = F, cluster_cols = F)
+  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
+           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
 gs_other %>%
@@
   column_to_rownames("ids") %>%
-  pheatmap(., cluster_rows = F, cluster_cols = F)
+  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
+           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
 ```
@@
   column_to_rownames("scan_id") %>%
-  pheatmap(., cluster_rows = F, cluster_cols = F)
+  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
+           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
 dup_other %>%
@@
   column_to_rownames("feature") %>%
-  pheatmap(., cluster_rows = F, cluster_cols = F)
+  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
+           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
 dup_other %>%
@@
   column_to_rownames("scan_id") %>%
-  pheatmap(., cluster_rows = F, cluster_cols = F)
+  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
+           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
 df_other %>%
@@
   column_to_rownames("feature") %>%
-  pheatmap(., cluster_rows = F, cluster_cols = F)
+  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
+           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
 df_other %>%
@@
 # Comparison of scores between groups
+
+For scans that were duplicated, we looked at the first reading and excluded the second.
+Scores were calculated by considering bronchiectasis as absent = 0, cylindric = 1, varicose = 2, cystic = 3.
+For all other features absence was counted as 0 and presence as 1.
+
+```{r}
+# Import disease state
+outcomes = read.csv("~/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/disease.key.csv",na.strings = "")
+outcomes$NTM.PD..Y.N. = factor(outcomes$NTM.PD..Y.N.,
+                               levels = c("No","Yes"),ordered = T)
+# Remove training scans, get first reading of duplicates
+scores = data %>% 
+  filter(scan_id %in% outcomes$cff_registry_id,scorer != "GS") %>%
+  group_by(scorer,scan_id) %>%
+  slice_min(scoresheet_timestamp) %>% 
+  select(scorer, scan_id, bronchiectasis_rul:atelectasis___6) %>%
+  mutate(across(contains("bronchiectasis"),
+                ~as.numeric(factor(.x,levels=c("Absent","Cylindric",
+                                               "Varicose","Cystic")))-1),
+         across(mucoid_impaction___1:atelectasis___6,
+                ~as.numeric(factor(.x,levels=c("Unchecked","Checked")))-1))
+scores = left_join(scores,outcomes %>% select(cff_registry_id,NTM.PD..Y.N.),
+                   by = join_by(scan_id == cff_registry_id)) %>%
+  rename(`NTM-PD` = NTM.PD..Y.N.)
+# Get scores
+majority = scores %>%
+  group_by(scan_id) %>%
+  summarise(across(bronchiectasis_rul:`NTM-PD`,
+                   ~tail(names(sort(table(.x))), 1)))
+max = scores %>%
+  group_by(scan_id) %>%
+  summarise(across(bronchiectasis_rul:`NTM-PD`,
+                   ~max(.x,na.rm = T)))
+```
+
+```{r}
+comp_score = function(df){
+  df = df %>% pivot_longer(bronchiectasis_rul:atelectasis___6,
+                           names_to = c("feature", "location"),
+                           names_pattern = "(.*)_(.*)$"
+  ) %>%
+    mutate(
+      feature = gsub("__", "", feature),
+      feature = str_to_title(gsub("_", " ", feature))
+    )
+  df$location <- factor(df$location,
+                        levels = c(
+                          "1", "2", "3", "4", "5", "6", "ling", "lll",
+                          "lul", "rll", "rml", "rul"
+                        ),
+                        labels = c(
+                          "RUL", "RML", "RLL", "LUL", "Ling", "LLL",
+                          "Ling", "LLL", "LUL", "RLL", "RML", "RUL"
+                        )
+  )
+  df$value = as.numeric(df$value)
+  # Tables
+  sums = df %>% group_by(scan_id,`NTM-PD`) %>%
+    summarise(Total = sum(value),.groups = "drop") 
+  t1 = tableby(`NTM-PD`~Total,data = sums,numeric.test = "kwt",
+               numeric.stats = c("Nmiss2","median","q1q3","range"))
+
+  sums = df %>% group_by(scan_id,`NTM-PD`,feature) %>%
+    summarise(score = sum(value),.groups = "drop") %>%
+    pivot_wider(names_from = feature,values_from = score)
+  f = as.formula(paste0("`NTM-PD`~",paste0(paste0("`",unique(df$feature),"`"),
+                                           collapse = "+")))
+  t2 = tableby(f,data = sums,numeric.test = "kwt",
+               numeric.stats = c("Nmiss2","median","q1q3","range"))
+  # Plots
+  total_plot = df %>% group_by(scan_id,`NTM-PD`,feature) %>%
+    summarise(score = sum(value),.groups = "drop") %>%
+    ggplot(.,aes(x = `NTM-PD`, y = score,fill = `NTM-PD`)) + 
+    geom_violin() +
+    theme_bw() + 
+    theme(legend.position = "none")
+  feature_plot = total_plot +
+    facet_wrap(~feature,scales = "free") 
+  # Print
+  cat("\n")
+  cat("### Total Scores")
+  cat("\n")
+  cat("\n")
+  print(summary(t1,pfootnote = T))
+  cat("\n")
+  cat("\n")
+  print(total_plot)
+  cat("\n")
+  cat("\n")
+  cat("### Scores by Feature")
+  cat("\n")
+  cat("\n")
+  print(summary(t2,pfootnote = T))
+  cat("\n")
+  cat("\n")
+  print(feature_plot)
+  cat("\n")
+  cat("\n")
+}
+```
+
+## Using majority score for each feature
+
+```{r results='asis'}
+comp_score(majority)
+```
+
+## Using majority score for each feature
+
+```{r results='asis'}
+comp_score(max)
+```

### 2023-05-20 Stacey Martiniano/ntm_agreement.qmd
scores by location

@@
                                levels = c("No","Yes"),ordered = T)
-# Remove training scans, get first reading of duplicates
+# Remove training scans, get first reading of duplicates, convert to numeric
 scores = data %>% 
@@
                 ~as.numeric(factor(.x,levels=c("Unchecked","Checked")))-1))
+# Add outcome info
 scores = left_join(scores,outcomes %>% select(cff_registry_id,NTM.PD..Y.N.),
@@
   rename(`NTM-PD` = NTM.PD..Y.N.)
-# Get scores
+# Calculate scores
 majority = scores %>%
@@
   # Tables
+  # Total score
   sums = df %>% group_by(scan_id,`NTM-PD`) %>%
@@
                numeric.stats = c("Nmiss2","median","q1q3","range"))
-
+  # By feature
   sums = df %>% group_by(scan_id,`NTM-PD`,feature) %>%
@@
   t2 = tableby(f,data = sums,numeric.test = "kwt",
+               numeric.stats = c("Nmiss2","median","q1q3","range"))
+  # By location
+  sums = df %>% group_by(scan_id,`NTM-PD`,location) %>%
+    summarise(score = sum(value),.groups = "drop") %>%
+    pivot_wider(names_from = location,values_from = score)
+  f = as.formula(paste0("`NTM-PD`~",paste0(paste0("`",unique(df$location),"`"),
+                                           collapse = "+")))
+  t3 = tableby(f,data = sums,numeric.test = "kwt",
                numeric.stats = c("Nmiss2","median","q1q3","range"))
@@
     facet_wrap(~feature,scales = "free") 
+  location_plot = df %>% group_by(scan_id,`NTM-PD`,location) %>%
+    summarise(score = sum(value),.groups = "drop") %>%
+    ggplot(.,aes(x = `NTM-PD`, y = score,fill = `NTM-PD`)) + 
+    geom_violin() +
+    theme_bw() + 
+    theme(legend.position = "none") +
+    facet_wrap(~location,scales = "free") 
   # Print
@@
   cat("\n")
+  cat("### Scores by Location")
+  cat("\n")
+  cat("\n")
+  print(summary(t3,pfootnote = T))
+  cat("\n")
+  cat("\n")
+  print(location_plot)
+  cat("\n")
+  cat("\n")
 }

### 2023-05-22 Christine Chan/CF Registry/cf_registry_analysis.qmd
calculate percentiles from raw values

--- 2023-05-18 Christine Chan/CF Registry/cf_registry_analysis.qmd
@@
 library(forestmodel)
+library(childsds)
 ```
@@
 # Analysis function
-log_reg = function(year = 2005){
+log_reg = function(year){
   # Clean data
@@
     select(eDWID,Age_YrEnd,Gender,MutClass,A_IsOnEnzymes,A_FEV1_pct_predicted,
-           A_bmivalue,A_bmipercentile,A_height,A_heightpercentile,
-           NumPulmExacerbation,A_feeding2,A_pseudomonasaeruginosa,
+           A_bmivalue,A_bmipercentile,A_height,A_weight,A_heightpercentile,
+           NumPulmExacerbation,A_pseudomonasaeruginosa,
            A_fungalyeast1,A_pulmonarycomplications1,A_mycocultureresults1,
            A_corticosteroids1,cfrd_status_annual,A_Vx770,A_VX809comb)
+  # Calculate percentiles
+  df$bmi = df$A_weight / ((df$A_height/100)^2)
+  df$age_perc = ifelse(df$Age_YrEnd < 20,df$Age_YrEnd,20)
+  df$height_perc = sds(df$A_height,age = df$age_perc,
+                    sex = df$Gender,male = "M",female = "F",
+                    item = "height2_20",ref = cdc.ref,type = "perc")*100
+  df$bmi_perc = sds(df$bmi,age = df$age_perc,
+                    sex = df$Gender,male = "M",female = "F",
+                    item = "bmi",ref = cdc.ref,type = "perc")*100
   # Format columns
   df$MutClass = factor(df$MutClass,levels = c("1-3","4-5","Oth","N/A"),
-                       labels = c("1-3","4-5","Oth",NA))
+                       labels = c("1-3","4-5","Oth","Unknown"))
   df$NumPulmExacerbation=cut(df$NumPulmExacerbation,labels = c("0","1","2",">=3"),
                              breaks = c(0,1,2,3,Inf),right = F)
-  df$A_feeding2 = factor(df$A_feeding2,levels = 0:1,labels = c("No","Yes"))
-  df$A_IsOnEnzymes = factor(df$A_IsOnEnzymes,levels = 0:1,labels = c("No","Yes"))
-  df$A_FEV1_pct_predicted = cut(df$A_FEV1_pct_predicted,breaks = c(0,40,80,100,Inf),right = F)
-  df$A_corticosteroids1 = factor(df$A_corticosteroids1,levels = 0:1,labels = c("No","Yes"))
-  df$A_fungalyeast1 = factor(df$A_fungalyeast1,levels = 0:1,labels = c("No","Yes"))
-  df$A_pulmonarycomplications1 = factor(df$A_pulmonarycomplications1,levels = 0:1,labels = c("No","Yes"))
-  df$A_mycocultureresults1 = factor(df$A_mycocultureresults1,levels = 0:1,labels = c("No","Yes"))
-  df$A_Vx770 = factor(df$A_Vx770,levels = 0:1,labels = c("No","Yes"))
-  df$A_VX809comb = factor(df$A_VX809comb,levels = 0:1,labels = c("No","Yes"))
-  # Remove all empty columns
-  df = df[,-which(colSums(is.na(df))==nrow(df))]
+  df$A_FEV1_pct_predicted = cut(df$A_FEV1_pct_predicted,
+                                breaks = c(0,40,80,100,Inf),right = F)
+  df$A_FEV1_pct_predicted = relevel(df$A_FEV1_pct_predicted,ref = "[100,Inf)")
+  df$height_perc = cut(df$height_perc,breaks = c(0,10,25,50,Inf))
+  df$height_perc = relevel(df$height_perc,ref = "(50,Inf]")
+  df$bmi_perc = cut(df$bmi_perc,breaks = c(0,10,50,85,Inf))
+  df$bmi_perc = relevel(df$bmi_perc,ref = "(50,85]")
+  df$age_group = cut(df$Age_YrEnd,breaks = c(0,10,18,30,Inf))
+  df$age_group = relevel(df$age_group,ref = "(10,18]")
   # Bugs
@@
            "A_pulmonarycomplications1","A_mycocultureresults1")
+  # Yes/No variables
+  yn_vars = c("A_IsOnEnzymes","A_corticosteroids1",
+              "A_Vx770","A_VX809comb",bugs)
+  df[,yn_vars] = lapply(df[,yn_vars],factor,levels = 0:1,labels = c("No","Yes"))
+  # Remove all empty columns
+  empty = which(colSums(is.na(df))==nrow(df))
+  if(length(empty)>0){
+    df = df[,-empty]
+  }
   # Models
@@
                               c(bugs,"cfrd_status_annual","eDWID","A_bmivalue",
-                                "A_bmipercentile","A_height","A_heightpercentile"))]
+                                "A_bmipercentile","A_height","A_heightpercentile",
+                                "Age_YrEnd","A_weight","bmi", "age_perc"))]
   base_mod = as.formula(paste0("cfrd_status_annual~",paste0(vars,collapse = "+")))
-  adult_mod = update(base_mod,.~.+A_bmivalue+A_height)
-  peds_mod = update(base_mod,.~.+A_bmipercentile+A_heightpercentile)
   # Simple models
-  adult_mod1 = glm(adult_mod,data = df,family = "binomial")
-  peds_mod1 = glm(peds_mod,data = df,family = "binomial")
+  mod1 = glm(base_mod,data = df,family = "binomial")
   # Print results
   cat("\n")
-  cat("### Base Models")
+  cat("### Base Model")
   cat("\n")
   cat("\n")
-  cat("#### Adult")
-  cat("\n")
-  cat("\n")
-  print(forest_model(adult_mod1))
-  cat("\n")
-  cat("\n")
-  cat("#### Peds")
-  cat("\n")
-  cat("\n")
-  print(forest_model(peds_mod1))
+  print(forest_model(mod1))
   cat("\n")
   if ("A_pseudomonasaeruginosa" %in% colnames(df)){
-    adult_mod2 = glm(update(adult_mod,.~.+A_pseudomonasaeruginosa),data = df,family = "binomial")
-    peds_mod2 = glm(update(peds_mod,.~.+A_pseudomonasaeruginosa),data = df,family = "binomial")
+    mod2 = glm(update(base_mod,.~.+A_pseudomonasaeruginosa),data = df,family = "binomial")
     cat("\n")
@@
     cat("\n")
-    cat("#### Adult")
-    cat("\n")
-    cat("\n")
-    print(forest_model(adult_mod2))
-    cat("\n")
-    cat("\n")
-    cat("#### Peds")
-    cat("\n")
-    cat("\n")
-    print(forest_model(peds_mod2))
+    print(forest_model(mod2))
     cat("\n")
@@
   if("A_fungalyeast1" %in% colnames(df)){
-    adult_mod3 = glm(update(adult_mod,.~.+A_fungalyeast1),data = df,family = "binomial")
-    peds_mod3 = glm(update(peds_mod,.~.+A_fungalyeast1),data = df,family = "binomial")
+    mod3 = glm(update(base_mod,.~.+A_fungalyeast1),data = df,family = "binomial")
     cat("\n")
@@
     cat("\n")
-    cat("#### Adult")
-    cat("\n")
-    cat("\n")
-    print(forest_model(adult_mod3))
-    cat("\n")
-    cat("\n")
-    cat("#### Peds")
-    cat("\n")
-    cat("\n")
-    print(forest_model(peds_mod3))
+    print(forest_model(mod3))
     cat("\n")
@@
   if("A_pulmonarycomplications1" %in% colnames(df)){
-    adult_mod4 = glm(update(adult_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
-    peds_mod4 = glm(update(peds_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
+    mod4 = glm(update(base_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
     cat("\n")
@@
     cat("\n")
-    cat("#### Adult")
-    cat("\n")
-    cat("\n")
-    print(forest_model(adult_mod4))
-    cat("\n")
-    cat("\n")
-    cat("#### Peds")
-    cat("\n")
-    cat("\n")
-    print(forest_model(peds_mod4))
+    print(forest_model(mod4))
     cat("\n")
@@
   if("A_mycocultureresults1" %in% colnames(df)){
-    adult_mod5 = glm(update(adult_mod,.~.+A_mycocultureresults1),data = df,family = "binomial")
-    peds_mod5 = glm(update(peds_mod,.~.+A_mycocultureresults1),data = df,family = "binomial")
+    mod5 = glm(update(base_mod,.~.+A_mycocultureresults1),data = df,family = "binomial")
     cat("\n")
@@
     cat("\n")
-    cat("#### Adult")
-    cat("\n")
-    cat("\n")
-    print(forest_model(adult_mod5))
-    cat("\n")
-    cat("\n")
-    cat("#### Peds")
-    cat("\n")
-    cat("\n")
-    print(forest_model(peds_mod5))
+    print(forest_model(mod5))
     cat("\n")
@@
 
-## 2005
+## 2006
 
 ```{r results='asis',fig.height=10,fig.width=10}
-log_reg(2005)
+log_reg(2006)
+```
+
+## 2013
+
+```{r results='asis',fig.height=10,fig.width=10}
+log_reg(2013)
+```
+
+## 2018
+
+```{r results='asis',fig.height=10,fig.width=10}
+log_reg(2018)
 ```

### 2023-05-26 Christine Chan/CF Registry/cf_registry_analysis.qmd
Update cf_registry_analysis.qmd

@@
 # Import and clean data
-annual <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
-                   na.strings = c("", "N/A")
+annual <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",na.strings = c("", "N/A")
 )
@@
            A_bmivalue,A_bmipercentile,A_height,A_weight,A_heightpercentile,
+           A_supplement_feeding,A_feeding2,
            NumPulmExacerbation,A_pseudomonasaeruginosa,
@@
            A_corticosteroids1,cfrd_status_annual,A_Vx770,A_VX809comb)
+  # If A_supplement_feeding is 0, then A_feeding2 is also 0 because they are not 
+  # on any supplemental feeding
+  df$A_feeding2[df$A_supplement_feeding == 0] = 0
   # Calculate percentiles
@@
   # Yes/No variables
-  yn_vars = c("A_IsOnEnzymes","A_corticosteroids1",
+  yn_vars = c("A_IsOnEnzymes","A_corticosteroids1","A_feeding2",
               "A_Vx770","A_VX809comb",bugs)
@@
                                 "A_bmipercentile","A_height","A_heightpercentile",
-                                "Age_YrEnd","A_weight","bmi", "age_perc"))]
+                                "Age_YrEnd","A_weight","bmi", "age_perc",
+                                "A_supplement_feeding"))]
   base_mod = as.formula(paste0("cfrd_status_annual~",paste0(vars,collapse = "+")))

### 2023-06-02 Christine Chan/CF Registry/cf_registry_analysis.qmd
CFRD registry

@@
 # Import and clean data
-annual <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",na.strings = c("", "N/A")
+annual <- read.csv("~/Documents/Work/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",na.strings = c("", "N/A")
 )
@@
                                    labels = c("CFRD-","CFRD-","CFRD+"))
-demogs = read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/CFF19_DemogCFDiag_Del1.csv")
+demogs = read.csv("~/Documents/Work/Christine Chan/CF registry data/CFF19_DemogCFDiag_Del1.csv")
 demogs = demogs %>% select(eDWID,Gender,MutClass)
@@
     filter(ReviewYear==year) %>%
-    select(eDWID,Age_YrEnd,Gender,MutClass,A_IsOnEnzymes,A_FEV1_pct_predicted,
+    select(eDWID,Age_YrEnd,Gender,MutClass,A_IsOnEnzymes,A_FEV1,A_FEV1_pct_predicted,
            A_bmivalue,A_bmipercentile,A_height,A_weight,A_heightpercentile,

### 2023-06-07 Christine Chan/CF Registry/cf_registry_analysis.qmd
CFRD registry abstract done

@@
     page-layout: full
+execute:
+  echo: false
+  warning: false
 editor: source
@@
 library(childsds)
+library(labelled)
 ```
@@
 
-Any participants with T1D or T2D were excluded from these analyses. Variables that were entirely missing in a given year (e.g., `A_Vx770` and `A_VX809comb` in 2005) were excluded from those models.
+Any participants with T1D or T2D were excluded from these analyses.
+Variables that were entirely missing in a given year (e.g., `A_Vx770` and `A_VX809comb` in 2005) were excluded from those models.
+Participants under age 10 at year end were also excluded due to missing PFT data.
 
@@
   df = annual %>% 
-    filter(ReviewYear==year) %>%
+    filter(ReviewYear==year,Age_YrEnd >= 10) %>%
     select(eDWID,Age_YrEnd,Gender,MutClass,A_IsOnEnzymes,A_FEV1,A_FEV1_pct_predicted,
            A_bmivalue,A_bmipercentile,A_height,A_weight,A_heightpercentile,
-           A_supplement_feeding,A_feeding2,
+           A_supplement_feeding,A_feeding2:A_feeding4,
            NumPulmExacerbation,A_pseudomonasaeruginosa,
-           A_fungalyeast1,A_pulmonarycomplications1,A_mycocultureresults1,
-           A_corticosteroids1,cfrd_status_annual,A_Vx770,A_VX809comb)
+           A_fungalyeast1,A_pulmonarycomplications1,A_mycocultureresults3,
+           A_corticosteroids1,cfrd_status_annual,A_Vx770,A_VX809comb,
+           A_hepatobiliary1_3,contains("A_hepatobiliary2_"))
   # If A_supplement_feeding is 0, then A_feeding2 is also 0 because they are not 
   # on any supplemental feeding
+  # Include all kinds of tube feeds
   df$A_feeding2[df$A_supplement_feeding == 0] = 0
+  df$A_feeding3[df$A_supplement_feeding == 0] = 0
+  df$A_feeding4[df$A_supplement_feeding == 0] = 0
+  df$tube_feeding = 
+    apply(df[,c("A_feeding2","A_feeding3","A_feeding4")],1,function(r){
+      if(all(is.na(r))){
+        return(NA)
+      } else {
+        return(max(as.numeric(r),na.rm = T))
+      }
+    })
+  # Liver disease
+  liver = c("A_hepatobiliary1_3",paste0("A_hepatobiliary2_",1:4))
+  df$liver_disease = 
+    apply(df[,liver],1,function(r){
+      if(all(is.na(r))){
+        return(NA)
+      } else {
+        return(max(as.numeric(r),na.rm = T))
+      }
+    })
   # Calculate percentiles
@@
   df$bmi_perc = relevel(df$bmi_perc,ref = "(50,85]")
-  df$age_group = cut(df$Age_YrEnd,breaks = c(0,10,18,30,Inf))
-  df$age_group = relevel(df$age_group,ref = "(10,18]")
+  df$age_group = cut(df$Age_YrEnd,breaks = c(10,18,30,Inf),right = F)
+  df$age_group = relevel(df$age_group,ref = "[10,18)")
   # Bugs
   bugs = c("A_pseudomonasaeruginosa","A_fungalyeast1",
-           "A_pulmonarycomplications1","A_mycocultureresults1")
+           "A_pulmonarycomplications1","A_mycocultureresults3")
   # Yes/No variables
   yn_vars = c("A_IsOnEnzymes","A_corticosteroids1","A_feeding2",
-              "A_Vx770","A_VX809comb",bugs)
+              "A_Vx770","A_VX809comb",bugs,"tube_feeding","liver_disease")
   df[,yn_vars] = lapply(df[,yn_vars],factor,levels = 0:1,labels = c("No","Yes"))
+  df$A_IsOnEnzymes = relevel(df$A_IsOnEnzymes,ref = "Yes")
   # Remove all empty columns
@@
                                 "Age_YrEnd","A_weight","bmi", "age_perc",
-                                "A_supplement_feeding"))]
+                                "A_supplement_feeding","A_feeding2","A_feeding3",
+                                "A_feeding4","A_FEV1",liver))]
   base_mod = as.formula(paste0("cfrd_status_annual~",paste0(vars,collapse = "+")))
+  # Labels
+  labels = list(cfrd_status_annual = "CFRD Status",MutClass = "Mutation class",
+                Gender = "Sex",A_IsOnEnzymes = "Pancreatic enzyme use",
+                A_FEV1_pct_predicted = "FEV1 percent predicted",
+                tube_feeding = "Enteric feedings",liver_disease = "Liver disease",
+                A_corticosteroids1 = "Use of corticosteroids",
+                NumPulmExacerbation = "Admissions for PEx",age_group = "Age group",
+                height_perc = "Height percentile",bmi_perc = "BMI percentile",
+                A_Vx770 = "Use of CFTR modulator Vx770",
+                A_VX809comb = "Use of CFTR modulator Vx809 combination",
+                A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
+                A_fungalyeast1 = "Aspergillus (any species)",
+                A_pulmonarycomplications1 = "ABPA",
+                A_mycocultureresults3 = "Microorganisms in myco culture")
+  var_label(df[,vars[which(vars %in% colnames(df))]]) = 
+    as.character(labels[vars[which(vars %in% colnames(df))]])
+  var_label(df[,bugs[which(bugs %in% colnames(df))]]) = 
+    as.character(labels[bugs[which(bugs %in% colnames(df))]])
   # Simple models
@@
   cat("\n")
-  print(forest_model(mod1))
+  print(forest_model(mod1,exclude_infinite_cis = F))
   cat("\n")
@@
     cat("\n")
-    print(forest_model(mod2))
+    print(forest_model(mod2,exclude_infinite_cis = F))
     cat("\n")
@@
     cat("\n")
-    print(forest_model(mod3))
+    print(forest_model(mod3,exclude_infinite_cis = F))
     cat("\n")
@@
     cat("\n")
-    print(forest_model(mod4))
+    print(forest_model(mod4,exclude_infinite_cis = F))
     cat("\n")
   }
-  if("A_mycocultureresults1" %in% colnames(df)){
-    mod5 = glm(update(base_mod,.~.+A_mycocultureresults1),data = df,family = "binomial")
+  if("A_mycocultureresults3" %in% colnames(df)){
+    mod5 = glm(update(base_mod,.~.+A_mycocultureresults3),data = df,family = "binomial")
     cat("\n")
@@
     cat("\n")
-    print(forest_model(mod5))
+    print(forest_model(mod5,exclude_infinite_cis = F))
     cat("\n")

### 2023-06-12 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
start racial differences in ETI

---
title: "Racial and Ethnic Differences in ETI"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r}
library(DataExplorer)
```

# EDA

```{r}
# Import
df = read.csv("~/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_Annualized_Del1.csv")
# Generate report
# df %>%
#   create_report(
#     output_file = paste0("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Reports/Annualized EDA ", 
#                          format(Sys.time(), "%Y-%m-%d"),".html"),
#     report_title = "Annualized EDA - Racial and Ethnic Differences in ETI"
#   )
skimr::skim(df)
```

### 2023-06-20 Jordana Hoppe/nacfc_2023.qmd
updated report for Jordana

@@
 library(gtsummary)
-df <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/clinical.csv", na.strings = "")
-surveys <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/surveys.csv", na.strings = "")
+df <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/clinical_2023-06-20.csv", na.strings = "")
+surveys <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/surveys_2023-06-20.csv", na.strings = "")
 # Combined genotype

### 2023-06-26 Emily Muther/body_image_trikafta.qmd
CF mental health screening

@@
 # Read in data
-df = read.spss("/Users/timvigers/Documents/Work/Vigers/CF/Emily Muther/Body Image and Trikafta/Data_Raw/TrikaftaBodyImage3.6.23.sav",to.data.frame = T)
+df = read.spss("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Emily Muther/Body Image and Trikafta/Data_Raw/TrikaftaBodyImage3.6.23.sav",to.data.frame = T)
 # Filter

### 2023-06-26 Emily Muther/mental_health_screening.qmd
CF mental health screening

@@
 
+```{r}
+#| include: false
+library(tidyverse)
+library(plotly)
+library(gtsummary)
+library(GLMMadaptive)
+library(lmerTest)
+library(sjPlot)
+```
+
 # Data cleaning
@@
 
-```{r include=FALSE}
-library(tidyverse)
-library(plotly)
-library(gtsummary)
-library(GLMMadaptive)
+```{r}
+#| warning: false
 # Import
@@
 # Column types
+df$MRN = factor(df$MRN)
 df$CONTACT_DATE = lubridate::mdy(df$CONTACT_DATE)
@@
 
-# Methods
+# Results
 
-Because the PHQ-9 results are right-skewed with lots of 0 scores, we used a zero inflated Poisson mixed model to analyze change in score over time.
+## ZINB mixed effects model
+
+Because the PHQ-9 results are right-skewed with lots of 0 scores, we first tried a zero inflated negative binomial (ZINB) mixed model to analyze change in score over time.
+At first we tried using categorical visit number as a time variable, but the model did not converge (likely due to the number of parameters).
+
+```{r}
+#| warning: false
+try(mixed_model(`PHQ-9 Total Score (calculated)` ~ factor(`Visit Number`),random = ~1|MRN,data = df,family = zi.negative.binomial(), zi_fixed = ~ 1))
+
+try(mixed_model(`PHQ-9 Total Score (calculated)` ~ `Days From Baseline`,
+                 random = ~1|MRN,data = df,family = zi.negative.binomial(), 
+                 zi_fixed = ~ 1))
+```
+
+Because the ZINB models were not converging, we instead tried to use a simpler Poisson mixed model (without accounting for the inflated 0s) with random intercept per participant and visit number treated as a continuous variable.
+
+```{r}
+#| warning: false
+m2 = glmer(`PHQ-9 Total Score (calculated)` ~ `Visit Number` + (1|MRN),data = df,family = "poisson")
+plot_model(m2,type="pred",
+           terms=c("Visit Number","MRN"),
+           pred.type="re", ci.lvl = NA) + 
+  theme(legend.position = "none")
+```
+
+## Simpler models
+
+Because the complex models were difficult to fit and not particularly informative, we instead modeled the odds of an elevated score for the full cohort, and the change in score over time for those with elevated scores.
+
+### PHQ9
+
+```{r}
+df$PHQ9_elevated = factor(df$`PHQ-9 Total Score (calculated)` >= 15,
+                          levels = c(F,T),labels = c("No","Yes"))
+```
+
+A PHQ9 score $\geq$ 15 was considered elevated.
+
+#### Odds of an elevated score
+
+```{r}
+mod = glmer(PHQ9_elevated ~ `Visit Number` + (1|MRN),data = df,family = "binomial")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+Results are reported as odds ratios, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod = glm(PHQ9_elevated ~ FRS_Category,data = df,family = "binomial")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+### Association with FRS category in those with elevated scores.
+
+Results are reported as expected counts, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod = glm(PHQ9_Total~ FRS_Category,data = df[df$PHQ9_elevated=="yes",],family = "poisson")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+## GAD7
+
+### Odds of an elevated score
+
+Results are reported as odds ratios, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod = glm(GAD7_elevated ~ FRS_Category,data = df,family = "binomial")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+### Association with FRS category in those with elevated scores.
+
+Results are reported as expected counts, with the FRS "neutral" category as the reference group.
+
+```{r}
+mod = glm(GAD7_Total~ FRS_Category,data = df[df$GAD7_elevated=="yes",],family = "poisson")
+res = tidy(mod,exponentiate = T)
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+# Questions
+
+1. Should we count moderate or severe scores as elevated?

### 2023-06-28 Stacey Martiniano/ntm_agreement.qmd
fixed section labels
compare scores by individual scorer

@@
 editor: source
+execute:
+  waring: false
 ---
@@
 # Import data
-api <- read.table("~/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt",
+api <- read.table("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt",
                   header = F
@@
 # Revert duplicate IDs back to original to match with gold standard
-training <- read.csv("~/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/training.key.csv")
-dups <- read.csv("~/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/duplicated.key.csv")
+training <- read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/training.key.csv")
+dups <- read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/duplicated.key.csv")
 training$Training_id <- as.numeric(sub("T00", "", training$Training_id))
@@
 # Import disease state
-outcomes = read.csv("~/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/disease.key.csv",na.strings = "")
+outcomes = read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/disease.key.csv",na.strings = "")
 outcomes$NTM.PD..Y.N. = factor(outcomes$NTM.PD..Y.N.,
@@
                    ~max(.x,na.rm = T)))
+a = scores %>%
+  filter(scorer == "A")
+b = scores %>%
+  filter(scorer == "B") 
+c = scores %>%
+  filter(scorer == "C")
 ```
@@
 
-## Using majority score for each feature
+## Using maximum score for each feature
 
@@
 ```
+
+## Scorer A
+
+```{r results='asis'}
+comp_score(a)
+```
+
+## Scorer B
+
+```{r results='asis'}
+comp_score(b)
+```
+
+## Scorer C
+
+```{r results='asis'}
+comp_score(c)
+```

### 2023-07-17 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
tidyverse formatting

@@
 ```{r}
-library(DataExplorer)
+library(readxl)
 ```
@@
 # Import
-df = read.csv("~/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_Annualized_Del1.csv")
-# Generate report
-# df %>%
-#   create_report(
-#     output_file = paste0("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Reports/Annualized EDA ", 
-#                          format(Sys.time(), "%Y-%m-%d"),".html"),
-#     report_title = "Annualized EDA - Racial and Ethnic Differences in ETI"
-#   )
-skimr::skim(df)
+df <- read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_Annualized_Del1.csv",na.strings = "")
+demo = read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_DemogCFDiag_Del1.csv",na.strings = "")
+dict = read_excel("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/Copy of Codes for CFFPR_2023.xlsx")
+# Merge demographic data
+
+# Factor variables
+df$smoking <- factor(df$smoking,
+  levels = 1:4,
+  labels = c(
+    "No", "Occasionally",
+    "Yes, Regularly, less than 1 ppd",
+    "Yes, Regularly, 1 ppd or more"
+  )
+)
+df$second_smoke <- factor(df$second_smoke,
+  levels = 1:4,
+  labels = c(
+    "Daily", "Several Times Per Week",
+    "Several Times Per Month or less",
+    "Never"
+  )
+)
+df$cfrd_status_annual <- factor(df$cfrd_status_annual,
+  levels = 1:5,
+  labels = c("NGT", "IGT", "CFRD", "T1D", "T2D")
+)
+df$Transplant_status <- factor(df$Transplant_status,
+  levels = 1:5,
+  labels = c(
+    "Not pertinent",
+    "Accepted, on waiting list",
+    "Evaluated, final decision pending",
+    "Evaluated, rejected",
+    "Had transplantation"
+  )
+)
+# Format binary variables
+bin_vars = c("insurance","employment")
+bin_vars = colnames(df)[which(sub("\\d","",colnames(df)) %in% bin_vars)]
+bin_vars = c(bin_vars,"pregnant")
+df[,bin_vars] = lapply(df[,bin_vars],factor,levels = 0:1,labels = c("No","Yes"))
 ```

### 2023-07-28 Stacey Martiniano/ntm_agreement.qmd
first pass at scoring across lobes

--- 2023-05-19 Stacey Martiniano/ntm_agreement.qmd
@@
 )
-data <- exportRecords(rcon, labels = F)
+data <- exportRecordsTyped(rcon, labels = F)
 data <- data.frame(lapply(data, as.character))
@@
 ```
-
-| Cohen's Kappa | Interpretation         |
-|---------------|------------------------|
-| 0             | No agreement           |
-| 0.10 - 0.20   | Slight agreement       |
-| 0.21 - 0.40   | Fair agreement         |
-| 0.41 - 0.60   | Moderate agreement     |
-| 0.61 - 0.80   | Substantial agreement  |
-| 0.81 - 0.99   | Near perfect agreement |
-| 1             | Perfect agreement      |
 
@@
   arrange(feature, location)
-```
-
-## Bronciechtasis
-
-```{r}
-gs_bronch <- gs %>% filter(feature == "Bronchiectasis")
-fk <- kappam.fleiss(gs_bronch[, c("A", "B", "C", "GS")])
-A <- kappa2(gs_bronch[, c("A", "GS")])
-B <- kappa2(gs_bronch[, c("B", "GS")])
-C <- kappa2(gs_bronch[, c("C", "GS")])
+# Check agreement with GS column
+gs$A <- factor(gs$A == gs$GS, levels = c(F, T), labels = c("Disagree", "Agree"))
+gs$B <- factor(gs$B == gs$GS, levels = c(F, T), labels = c("Disagree", "Agree"))
+gs$C <- factor(gs$C == gs$GS, levels = c(F, T), labels = c("Disagree", "Agree"))
 # Plot
-gs_bronch %>%
-  select(scan_id, feature, location, A:GS) %>%
-  pivot_longer(A:GS) %>%
-  ggplot(., aes(x = name, y = location, fill = value)) +
-  geom_tile(color = "black") +
-  theme_bw() +
-  labs(fill = "", x = "Scorer", y = "Location") +
-  custom_colors +
-  facet_wrap(~scan_id)
-```
-
-Overall Fleiss' Kappa for all 5 gold standard scans was `r round(fk$value,3)` (p = `r format.pval(fk$p.value,eps = 0.001,digits = 3)`).
-
-Overall Cohen's Kappa for A vs. the gold standard was `r round(A$value,3)` (p = `r format.pval(A$p.value,eps = 0.001,digits = 3)`).
-
-Overall Cohen's Kappa for B vs. the gold standard was `r round(B$value,3)` (p = `r format.pval(B$p.value,eps = 0.001,digits = 3)`).
-
-Overall Cohen's Kappa for C vs. the gold standard was `r round(C$value,3)` (p = `r format.pval(C$p.value,eps = 0.001,digits = 3)`).
-
-## Other features
-
-```{r}
-gs_other <- gs %>% filter(feature != "Bronchiectasis")
-# Check agreement with GS column
-gs_other$A <- factor(gs_other$A == gs_other$GS, levels = c(F, T), labels = c("Disagree", "Agree"))
-gs_other$B <- factor(gs_other$B == gs_other$GS, levels = c(F, T), labels = c("Disagree", "Agree"))
-gs_other$C <- factor(gs_other$C == gs_other$GS, levels = c(F, T), labels = c("Disagree", "Agree"))
-# Plot
-gs_other %>%
+gs %>%
   select(scan_id, feature, location, A:GS) %>%
@@
 # Tables
-gs_other %>%
+gs %>%
   summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
   kable(., digits = 3, caption = "Proportion of agreement with GS")
-gs_other %>%
+gs %>%
   group_by(scan_id) %>%
@@
   kable(., digits = 3, caption = "Proportion of agreement with GS")
-gs_other %>%
+gs %>%
   group_by(feature) %>%
@@
   kable(., digits = 3, caption = "Proportion of agreement with GS")
-gs_other %>%
+gs %>%
   group_by(scan_id, feature) %>%
@@
   kable(., digits = 3, caption = "Proportion of agreement with GS")
+gs %>%
+  group_by(location) %>%
+  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
+  kable(., digits = 3, caption = "Proportion of agreement with GS")
 # Heatmaps
-gs_other %>%
+gs %>%
   group_by(scan_id) %>%
@@
            colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
-gs_other %>%
+gs %>%
   group_by(feature) %>%
@@
            colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
-gs_other %>%
+gs %>%
   group_by(scan_id, feature) %>%
@@
   column_to_rownames("ids") %>%
+  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
+           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
+gs %>%
+  group_by(location) %>%
+  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
+  column_to_rownames("location") %>%
   pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
@@
   arrange(feature, location)
-```
-
-## Bronciechtasis
-
-```{r}
-dup_bronch <- dup %>% filter(feature == "Bronchiectasis")
-A <- kappa2(dup_bronch[, c("A_1", "A_2")])
-B <- kappa2(dup_bronch[, c("B_1", "B_2")])
-C <- kappa2(dup_bronch[, c("C_1", "C_2")])
+# Check agreement with GS column
+dup$A <- factor(dup$A_1 == dup$A_2, levels = c(F, T), labels = c("Disagree", "Agree"))
+dup$B <- factor(dup$B_1 == dup$B_2, levels = c(F, T), labels = c("Disagree", "Agree"))
+dup$C <- factor(dup$C_1 == dup$C_2, levels = c(F, T), labels = c("Disagree", "Agree"))
 # Plot
-dup_bronch %>%
-  select(scan_id, feature, location, A_1:C_2) %>%
-  pivot_longer(A_1:C_2) %>%
-  ggplot(., aes(x = name, y = location, fill = value)) +
-  geom_tile(color = "black") +
-  theme_bw() +
-  labs(fill = "", x = "Scorer", y = "Location") +
-  custom_colors +
-  facet_wrap(~scan_id)
-```
-
-Overall Cohen's Kappa for A vs. themselves was `r round(A$value,3)` (p = `r format.pval(A$p.value,eps = 0.001,digits = 3)`).
-
-Overall Cohen's Kappa for B vs. themselves was `r round(B$value,3)` (p = `r format.pval(B$p.value,eps = 0.001,digits = 3)`).
-
-Overall Cohen's Kappa for C vs. themselves was `r round(C$value,3)` (p = `r format.pval(C$p.value,eps = 0.001,digits = 3)`).
-
-## Other features
-
-```{r}
-dup_other <- dup %>% filter(feature != "Bronchiectasis")
-# Check agreement with GS column
-dup_other$A <- factor(dup_other$A_1 == dup_other$A_2, levels = c(F, T), labels = c("Disagree", "Agree"))
-dup_other$B <- factor(dup_other$B_1 == dup_other$B_2, levels = c(F, T), labels = c("Disagree", "Agree"))
-dup_other$C <- factor(dup_other$C_1 == dup_other$C_2, levels = c(F, T), labels = c("Disagree", "Agree"))
-# Plot
-dup_other %>%
+dup %>%
   select(scan_id, feature, location, A:C) %>%
@@
 # Tables
-dup_other %>%
+dup %>%
   summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
   kable(., digits = 3, caption = "Proportion of agreement with themselves")
-dup_other %>%
+dup %>%
   group_by(scan_id) %>%
@@
   kable(., digits = 3, caption = "Proportion of agreement with themselves")
-dup_other %>%
+dup %>%
   group_by(feature) %>%
@@
   kable(., digits = 3, caption = "Proportion of agreement with themselves")
-dup_other %>%
+dup %>%
   group_by(scan_id, feature) %>%
@@
   kable(., digits = 3, caption = "Proportion of agreement with themselves")
+dup %>%
+  group_by(location) %>%
+  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
+  kable(., digits = 3, caption = "Proportion of agreement with themselves")
 # Heatmaps
-dup_other %>%
+dup %>%
   group_by(scan_id) %>%
@@
            colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
-dup_other %>%
+dup %>%
   group_by(feature) %>%
@@
            colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
-dup_other %>%
+dup %>%
   group_by(scan_id, feature) %>%
@@
   column_to_rownames("ids") %>%
+  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
+           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
+dup %>%
+  group_by(location) %>%
+  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
+  column_to_rownames("location") %>%
   pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
@@
   arrange(feature, location)
-```
-
-## Bronciechtasis
-
-```{r}
-df_bronch <- df %>% filter(feature == "Bronchiectasis")
-fk <- kappam.fleiss(df_bronch[, c("A", "B", "C")])
-# Plot
-df_bronch %>%
-  select(scan_id, feature, location, A:C) %>%
-  pivot_longer(A:C) %>%
-  ggplot(., aes(x = name, y = location, fill = value)) +
-  geom_tile(color = "black") +
-  theme_bw() +
-  labs(fill = "", x = "Scorer", y = "Location") +
-  custom_colors +
-  facet_wrap(~scan_id)
-```
-
-Overall Fleiss' Kappa for all 15 singleton scans was `r round(fk$value,3)` (p = `r format.pval(fk$p.value,eps = 0.001,digits = 3)`).
-
-## Other features
-
-```{r}
-df_other <- df %>% filter(feature != "Bronchiectasis")
 # Check agreement with GS column
-df_other[, c("A", "B", "C")] <- lapply(df_other[, c("A", "B", "C")], factor,
-                                       levels = c("Unchecked", "Checked"),
-                                       labels = c("Absent", "Present")
+df[, c("A", "B", "C")] <- lapply(df[, c("A", "B", "C")], factor,
+                                       levels = c("Unchecked", "Checked",
+                                                  "Absent","Cylindric","Cystic",
+                                                  "Varicose"),
+                                       labels = c("Absent", "Present",
+                                                  "Absent", "Present", 
+                                                  "Present", "Present")
 )
 # Three-way agreement
-df_other$`A vs. B` <- factor(df_other$A == df_other$B, levels = c(F, T), labels = c("Disagree", "Agree"))
-df_other$`A vs. C` <- factor(df_other$A == df_other$C, levels = c(F, T), labels = c("Disagree", "Agree"))
-df_other$`B vs. C` <- factor(df_other$B == df_other$C, levels = c(F, T), labels = c("Disagree", "Agree"))
+df$`A vs. B` <- factor(df$A == df$B, levels = c(F, T), labels = c("Disagree", "Agree"))
+df$`A vs. C` <- factor(df$A == df$C, levels = c(F, T), labels = c("Disagree", "Agree"))
+df$`B vs. C` <- factor(df$B == df$C, levels = c(F, T), labels = c("Disagree", "Agree"))
 # Plot
-df_other %>%
+df %>%
   select(scan_id, feature, location, `A vs. B`:`B vs. C`) %>%
@@
 # Tables
-df_other %>%
+df %>%
   summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
   kable(., digits = 3, caption = "Proportion of agreement with themselves")
-df_other %>%
+df %>%
   group_by(scan_id) %>%
@@
   kable(., digits = 3, caption = "Proportion of agreement with themselves")
-df_other %>%
+df %>%
   group_by(feature) %>%
@@
   kable(., digits = 3, caption = "Proportion of agreement with themselves")
-df_other %>%
-  group_by(scan_id, feature) %>%
+df %>%
+  group_by(location) %>%
   summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
@@
 # Heatmaps
-df_other %>%
+df %>%
   group_by(scan_id) %>%
@@
            colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
-df_other %>%
+df %>%
   group_by(feature) %>%
@@
            colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
-df_other %>%
-  group_by(scan_id, feature) %>%
+df %>%
+  group_by(location) %>%
   summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
-  unite(ids, scan_id, feature, na.rm = T, sep = " ") %>%
-  column_to_rownames("ids") %>%
+  column_to_rownames("location") %>%
   pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
@@
 For scans that were duplicated, we looked at the first reading and excluded the second.
-Scores were calculated by considering bronchiectasis as absent = 0, cylindric = 1, varicose = 2, cystic = 3.
-For all other features absence was counted as 0 and presence as 1.
+For all features, absence was counted as 0 and presence as 1.
+Absence and presence refer to all lobes (e.g.
+bronchiectasis on RUL and RML would score a 1, and so would bronchiectasis on RUL, RML, and LUL).
 
@@
 # Import disease state
-outcomes = read.csv("~/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/disease.key.csv",na.strings = "")
+outcomes = read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/disease.key.csv",na.strings = "")
 outcomes$NTM.PD..Y.N. = factor(outcomes$NTM.PD..Y.N.,
                                levels = c("No","Yes"),ordered = T)
-# Remove training scans, get first reading of duplicates
+# Remove training scans, get first reading of duplicates, convert to numeric
 scores = data %>% 
@@
   slice_min(scoresheet_timestamp) %>% 
-  select(scorer, scan_id, bronchiectasis_rul:atelectasis___6) %>%
+  select(scorer, scan_id,bronchiectasis_rul:atelectasis___6) %>%
   mutate(across(contains("bronchiectasis"),
                 ~as.numeric(factor(.x,levels=c("Absent","Cylindric",
-                                               "Varicose","Cystic")))-1),
+                                               "Varicose","Cystic"),
+                                   labels = c("Unchecked","Checked",
+                                              "Checked","Checked")))-1),
          across(mucoid_impaction___1:atelectasis___6,
                 ~as.numeric(factor(.x,levels=c("Unchecked","Checked")))-1))
-scores = left_join(scores,outcomes %>% select(cff_registry_id,NTM.PD..Y.N.),
-                   by = join_by(scan_id == cff_registry_id)) %>%
-  rename(`NTM-PD` = NTM.PD..Y.N.)
-# Get scores
-majority = scores %>%
-  group_by(scan_id) %>%
-  summarise(across(bronchiectasis_rul:`NTM-PD`,
-                   ~tail(names(sort(table(.x))), 1)))
-max = scores %>%
-  group_by(scan_id) %>%
-  summarise(across(bronchiectasis_rul:`NTM-PD`,
-                   ~max(.x,na.rm = T)))
-```
-
-```{r}
-comp_score = function(df){
-  df = df %>% pivot_longer(bronchiectasis_rul:atelectasis___6,
+# Take the maximum across all locations
+scores = scores %>% pivot_longer(bronchiectasis_rul:atelectasis___6,
                names_to = c("feature", "location"),
@@
   ) %>%
-    mutate(
-      feature = gsub("__", "", feature),
-      feature = str_to_title(gsub("_", " ", feature))
-    )
-  df$location <- factor(df$location,
-                        levels = c(
-                          "1", "2", "3", "4", "5", "6", "ling", "lll",
-                          "lul", "rll", "rml", "rul"
-                        ),
-                        labels = c(
-                          "RUL", "RML", "RLL", "LUL", "Ling", "LLL",
-                          "Ling", "LLL", "LUL", "RLL", "RML", "RUL"
-                        )
-  )
-  df$value = as.numeric(df$value)
-  # Tables
-  sums = df %>% group_by(scan_id,`NTM-PD`) %>%
-    summarise(Total = sum(value),.groups = "drop") 
-  t1 = tableby(`NTM-PD`~Total,data = sums,numeric.test = "kwt",
-               numeric.stats = c("Nmiss2","median","q1q3","range"))
-
-  sums = df %>% group_by(scan_id,`NTM-PD`,feature) %>%
-    summarise(score = sum(value),.groups = "drop") %>%
-    pivot_wider(names_from = feature,values_from = score)
-  f = as.formula(paste0("`NTM-PD`~",paste0(paste0("`",unique(df$feature),"`"),
-                                           collapse = "+")))
-  t2 = tableby(f,data = sums,numeric.test = "kwt",
-               numeric.stats = c("Nmiss2","median","q1q3","range"))
-  # Plots
-  total_plot = df %>% group_by(scan_id,`NTM-PD`,feature) %>%
-    summarise(score = sum(value),.groups = "drop") %>%
-    ggplot(.,aes(x = `NTM-PD`, y = score,fill = `NTM-PD`)) + 
-    geom_violin() +
-    theme_bw() + 
-    theme(legend.position = "none")
-  feature_plot = total_plot +
-    facet_wrap(~feature,scales = "free") 
-  # Print
-  cat("\n")
-  cat("### Total Scores")
-  cat("\n")
-  cat("\n")
-  print(summary(t1,pfootnote = T))
-  cat("\n")
-  cat("\n")
-  print(total_plot)
-  cat("\n")
-  cat("\n")
-  cat("### Scores by Feature")
-  cat("\n")
-  cat("\n")
-  print(summary(t2,pfootnote = T))
-  cat("\n")
-  cat("\n")
-  print(feature_plot)
-  cat("\n")
-  cat("\n")
-}
+  group_by(scorer,scan_id,feature) %>%
+  summarise(value = max(value),.groups = "drop") 
+# Calculate scores
+majority = scores %>%
+  group_by(scan_id,feature) %>%
+  summarise(value = names(sort(-table(value)))[1],.groups = "drop") %>%
+  group_by(scan_id) %>%
+  summarise(score = sum(as.numeric(value)))
+max = scores %>%
+  group_by(scan_id,feature) %>%
+  summarise(value = max(value),.groups = "drop") %>%
+  group_by(scan_id) %>%
+  summarise(score = sum(value))
+# Outcomes
+majority = left_join(majority,outcomes %>% select(cff_registry_id,NTM.PD..Y.N.),
+                   by = join_by(scan_id == cff_registry_id)) %>%
+  rename(`NTM-PD` = NTM.PD..Y.N.)
+max = left_join(max,outcomes %>% select(cff_registry_id,NTM.PD..Y.N.),
+                   by = join_by(scan_id == cff_registry_id)) %>%
+  rename(`NTM-PD` = NTM.PD..Y.N.)
 ```
@@
 
-```{r results='asis'}
-comp_score(majority)
+```{r}
+majority %>% select(score,`NTM-PD`) %>%
+  tbl_summary(by = `NTM-PD`)
 ```
 
-## Using majority score for each feature
+## Using maximum score for each feature
 

### 2023-07-29 Stacey Martiniano/ntm_agreement.qmd
started analysis with bronch 0 and 1

@@
   slice_min(scoresheet_timestamp) %>% 
-  select(scorer, scan_id,bronchiectasis_rul:atelectasis___6) %>%
+  select(scorer, scan_id, bronchiectasis_rul:atelectasis___6) %>%
   mutate(across(contains("bronchiectasis"),
@@
                 ~as.numeric(factor(.x,levels=c("Unchecked","Checked")))-1))
-# Take the maximum across all locations
-scores = scores %>% pivot_longer(bronchiectasis_rul:atelectasis___6,
+# Add outcome info
+scores = left_join(scores,outcomes %>% select(cff_registry_id,NTM.PD..Y.N.),
+                   by = join_by(scan_id == cff_registry_id)) %>%
+  rename(`NTM-PD` = NTM.PD..Y.N.)
+# Calculate scores
+majority = scores %>%
+  group_by(scan_id) %>%
+  summarise(across(bronchiectasis_rul:`NTM-PD`,
+                   ~tail(names(sort(table(.x))), 1)))
+max = scores %>%
+  group_by(scan_id) %>%
+  summarise(across(bronchiectasis_rul:`NTM-PD`,
+                   ~max(.x,na.rm = T)))
+a = scores %>%
+  filter(scorer == "A")
+b = scores %>%
+  filter(scorer == "B") 
+c = scores %>%
+  filter(scorer == "C")
+```
+
+```{r}
+comp_score = function(df){
+  df = df %>% pivot_longer(bronchiectasis_rul:atelectasis___6,
                            names_to = c("feature", "location"),
@@
   ) %>%
-  group_by(scorer,scan_id,feature) %>%
-  summarise(value = max(value),.groups = "drop") 
-# Calculate scores
-majority = scores %>%
-  group_by(scan_id,feature) %>%
-  summarise(value = names(sort(-table(value)))[1],.groups = "drop") %>%
-  group_by(scan_id) %>%
-  summarise(score = sum(as.numeric(value)))
-max = scores %>%
-  group_by(scan_id,feature) %>%
-  summarise(value = max(value),.groups = "drop") %>%
-  group_by(scan_id) %>%
-  summarise(score = sum(value))
-# Outcomes
-majority = left_join(majority,outcomes %>% select(cff_registry_id,NTM.PD..Y.N.),
-                   by = join_by(scan_id == cff_registry_id)) %>%
-  rename(`NTM-PD` = NTM.PD..Y.N.)
-max = left_join(max,outcomes %>% select(cff_registry_id,NTM.PD..Y.N.),
-                   by = join_by(scan_id == cff_registry_id)) %>%
-  rename(`NTM-PD` = NTM.PD..Y.N.)
+    mutate(
+      feature = gsub("__", "", feature),
+      feature = str_to_title(gsub("_", " ", feature))
+    )
+  df$location <- factor(df$location,
+                        levels = c(
+                          "1", "2", "3", "4", "5", "6", "ling", "lll",
+                          "lul", "rll", "rml", "rul"
+                        ),
+                        labels = c(
+                          "RUL", "RML", "RLL", "LUL", "Ling", "LLL",
+                          "Ling", "LLL", "LUL", "RLL", "RML", "RUL"
+                        )
+  )
+  df$value = as.numeric(df$value)
+  # Tables
+  # Total score
+  sums = df %>% group_by(scan_id,`NTM-PD`) %>%
+    summarise(Total = sum(value),.groups = "drop") 
+  t1 = tableby(`NTM-PD`~Total,data = sums,numeric.test = "kwt",
+               numeric.stats = c("Nmiss2","median","q1q3","range"))
+  # By feature
+  sums = df %>% group_by(scan_id,`NTM-PD`,feature) %>%
+    summarise(score = sum(value),.groups = "drop") %>%
+    pivot_wider(names_from = feature,values_from = score)
+  f = as.formula(paste0("`NTM-PD`~",paste0(paste0("`",unique(df$feature),"`"),
+                                           collapse = "+")))
+  t2 = tableby(f,data = sums,numeric.test = "kwt",
+               numeric.stats = c("Nmiss2","median","q1q3","range"))
+  # By location
+  sums = df %>% group_by(scan_id,`NTM-PD`,location) %>%
+    summarise(score = sum(value),.groups = "drop") %>%
+    pivot_wider(names_from = location,values_from = score)
+  f = as.formula(paste0("`NTM-PD`~",paste0(paste0("`",unique(df$location),"`"),
+                                           collapse = "+")))
+  t3 = tableby(f,data = sums,numeric.test = "kwt",
+               numeric.stats = c("Nmiss2","median","q1q3","range"))
+  # Plots
+  total_plot = df %>% group_by(scan_id,`NTM-PD`,feature) %>%
+    summarise(score = sum(value),.groups = "drop") %>%
+    ggplot(.,aes(x = `NTM-PD`, y = score,fill = `NTM-PD`)) + 
+    geom_violin() +
+    theme_bw() + 
+    theme(legend.position = "none")
+  feature_plot = total_plot +
+    facet_wrap(~feature,scales = "free") 
+  location_plot = df %>% group_by(scan_id,`NTM-PD`,location) %>%
+    summarise(score = sum(value),.groups = "drop") %>%
+    ggplot(.,aes(x = `NTM-PD`, y = score,fill = `NTM-PD`)) + 
+    geom_violin() +
+    theme_bw() + 
+    theme(legend.position = "none") +
+    facet_wrap(~location,scales = "free") 
+  # Print
+  cat("\n")
+  cat("### Total Scores")
+  cat("\n")
+  cat("\n")
+  print(summary(t1,pfootnote = T))
+  cat("\n")
+  cat("\n")
+  print(total_plot)
+  cat("\n")
+  cat("\n")
+  cat("### Scores by Feature")
+  cat("\n")
+  cat("\n")
+  print(summary(t2,pfootnote = T))
+  cat("\n")
+  cat("\n")
+  print(feature_plot)
+  cat("\n")
+  cat("\n")
+  cat("### Scores by Location")
+  cat("\n")
+  cat("\n")
+  print(summary(t3,pfootnote = T))
+  cat("\n")
+  cat("\n")
+  print(location_plot)
+  cat("\n")
+  cat("\n")
+}
 ```
@@
 
-```{r}
-majority %>% select(score,`NTM-PD`) %>%
-  tbl_summary(by = `NTM-PD`)
+```{r results='asis'}
+comp_score(majority)
 ```

### 2023-07-31 Stacey Martiniano/ntm_agreement.qmd
NTM file paths

@@
 # Import data
-api <- read.table("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt",
+api <- read.table("/Users/timvigers/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt",
                   header = F
@@
 # Revert duplicate IDs back to original to match with gold standard
-training <- read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/training.key.csv")
-dups <- read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/duplicated.key.csv")
+training <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/training.key.csv")
+dups <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/duplicated.key.csv")
 training$Training_id <- as.numeric(sub("T00", "", training$Training_id))
@@
 # Import disease state
-outcomes = read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/disease.key.csv",na.strings = "")
+outcomes = read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/disease.key.csv",na.strings = "")
 outcomes$NTM.PD..Y.N. = factor(outcomes$NTM.PD..Y.N.,

### 2023-07-31 Stacey Martiniano/ntm_agreement.qmd
basically starting NTM over again

@@
     page-layout: full
+    theme:
+      light: flatly
+      dark: darkly
+bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
+csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
 editor: source
-execute:
-  waring: false
 ---
@@
 new <- dups$original_ssid
-data$scan_id[data$scan_id %in% old] <- new[match(data$scan_id, old, nomatch = 0)]
+data$scan_id[data$scan_id %in% old] <- new[na.omit(match(data$scan_id, old))]
 # Change training IDs to OG IDs
@@
 new <- training$Cffid.Scoring.Set
-data$scan_id[data$scan_id %in% old] <- new[match(data$scan_id, old, nomatch = 0)]
-# Plot colors
-name <- "Accent"
-colors <- brewer.pal(6, name)
-names(colors) <- c("Checked", "Unchecked", "Absent", "Cylindric", "Varicose", NA)
-custom_colors <- scale_fill_manual(name = "", values = colors)
-```
-
-# Comparison to gold standards
-
-```{r}
-# Gold standard data
-gs <- data %>%
-  arrange(scan_id, scorer) %>%
-  filter(scan_id %in% training$Cffid.Scoring.Set) %>%
-  group_by(scorer, scan_id) %>%
-  slice_max(scoresheet_timestamp) %>%
-  select(scorer, scan_id, bronchiectasis_rul:atelectasis___6) %>%
-  ungroup() %>%
+data$scan_id[data$scan_id %in% old] <- new[na.omit(match(data$scan_id, old))]
+# Convert to long
+data = data %>% select(scan_id,scorer,scoresheet_timestamp,
+                    bronchiectasis_rul:atelectasis___6) %>%
+  arrange(scan_id,scorer,scoresheet_timestamp) %>%
+  group_by(scan_id,scorer) %>%
+  mutate(scorer_repeat = row_number()) %>%
+  select(-scoresheet_timestamp) %>%
   pivot_longer(bronchiectasis_rul:atelectasis___6,
@@
   )
-gs$location <- factor(gs$location,
+data$location <- factor(data$location,
                       levels = c(
@@
 )
-# One column per scorer
-gs <- gs %>%
-  pivot_wider(names_from = scorer, values_from = value) %>%
-  arrange(feature, location)
-# Check agreement with GS column
-gs$A <- factor(gs$A == gs$GS, levels = c(F, T), labels = c("Disagree", "Agree"))
-gs$B <- factor(gs$B == gs$GS, levels = c(F, T), labels = c("Disagree", "Agree"))
-gs$C <- factor(gs$C == gs$GS, levels = c(F, T), labels = c("Disagree", "Agree"))
-# Plot
-gs %>%
-  select(scan_id, feature, location, A:GS) %>%
-  pivot_longer(A:C) %>%
-  ggplot(., aes(x = name, y = location, fill = value)) +
-  geom_tile(color = "black") +
-  theme_bw() +
-  labs(fill = "", x = "Scorer", y = "Location") +
-  facet_grid(rows = vars(feature), cols = vars(scan_id))
-# Tables
-gs %>%
-  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
-  kable(., digits = 3, caption = "Proportion of agreement with GS")
-gs %>%
-  group_by(scan_id) %>%
-  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
-  kable(., digits = 3, caption = "Proportion of agreement with GS")
-gs %>%
-  group_by(feature) %>%
-  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
-  kable(., digits = 3, caption = "Proportion of agreement with GS")
-gs %>%
-  group_by(scan_id, feature) %>%
-  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
-  kable(., digits = 3, caption = "Proportion of agreement with GS")
-gs %>%
-  group_by(location) %>%
-  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
-  kable(., digits = 3, caption = "Proportion of agreement with GS")
-# Heatmaps
-gs %>%
-  group_by(scan_id) %>%
-  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
-  column_to_rownames("scan_id") %>%
-  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
-           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
-gs %>%
-  group_by(feature) %>%
-  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
-  column_to_rownames("feature") %>%
-  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
-           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
-gs %>%
-  group_by(scan_id, feature) %>%
-  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
-  unite(ids, scan_id, feature, na.rm = T, sep = " ") %>%
-  column_to_rownames("ids") %>%
-  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
-           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
-gs %>%
-  group_by(location) %>%
-  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
-  column_to_rownames("location") %>%
-  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
-           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
+# Maybe not quite so long (one column per rater)
+data = data %>% 
+  pivot_wider(names_from = c(scorer,scorer_repeat), values_from = value,
+              names_sep = "_")
 ```
 
-# Duplicates
+Kappa interpretation table @mchughInterraterReliabilityKappa2012:
 
-```{r}
-# Gold standard data
-dup <- data %>%
-  arrange(scan_id, scorer) %>%
-  filter(scan_id %in% dups$original_ssid) %>%
-  group_by(scorer, scan_id) %>%
-  mutate(n = row_number()) %>%
-  select(scorer, scan_id, n, bronchiectasis_rul:atelectasis___6) %>%
-  ungroup() %>%
-  pivot_longer(bronchiectasis_rul:atelectasis___6,
-               names_to = c("feature", "location"),
-               names_pattern = "(.*)_(.*)$"
-  ) %>%
-  mutate(
-    feature = gsub("__", "", feature),
-    feature = str_to_title(gsub("_", " ", feature))
-  )
-dup$location <- factor(dup$location,
-                       levels = c(
-                         "1", "2", "3", "4", "5", "6", "ling", "lll",
-                         "lul", "rll", "rml", "rul"
-                       ),
-                       labels = c(
-                         "RUL", "RML", "RLL", "LUL", "Ling", "LLL",
-                         "Ling", "LLL", "LUL", "RLL", "RML", "RUL"
-                       )
-)
-# One column per scorer
-dup <- dup %>%
-  pivot_wider(names_from = scorer, values_from = value) %>%
-  pivot_wider(names_from = n, values_from = A:C) %>%
-  arrange(feature, location)
-# Check agreement with GS column
-dup$A <- factor(dup$A_1 == dup$A_2, levels = c(F, T), labels = c("Disagree", "Agree"))
-dup$B <- factor(dup$B_1 == dup$B_2, levels = c(F, T), labels = c("Disagree", "Agree"))
-dup$C <- factor(dup$C_1 == dup$C_2, levels = c(F, T), labels = c("Disagree", "Agree"))
-# Plot
-dup %>%
-  select(scan_id, feature, location, A:C) %>%
-  pivot_longer(A:C) %>%
-  ggplot(., aes(x = name, y = location, fill = value)) +
-  geom_tile(color = "black") +
-  theme_bw() +
-  labs(fill = "", x = "Scorer", y = "Location") +
-  facet_grid(rows = vars(feature), cols = vars(scan_id))
-# Tables
-dup %>%
-  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
-  kable(., digits = 3, caption = "Proportion of agreement with themselves")
-dup %>%
-  group_by(scan_id) %>%
-  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
-  kable(., digits = 3, caption = "Proportion of agreement with themselves")
-dup %>%
-  group_by(feature) %>%
-  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
-  kable(., digits = 3, caption = "Proportion of agreement with themselves")
-dup %>%
-  group_by(scan_id, feature) %>%
-  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
-  kable(., digits = 3, caption = "Proportion of agreement with themselves")
-dup %>%
-  group_by(location) %>%
-  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
-  kable(., digits = 3, caption = "Proportion of agreement with themselves")
-# Heatmaps
-dup %>%
-  group_by(scan_id) %>%
-  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
-  column_to_rownames("scan_id") %>%
-  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
-           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
-dup %>%
-  group_by(feature) %>%
-  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
-  column_to_rownames("feature") %>%
-  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
-           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
-dup %>%
-  group_by(scan_id, feature) %>%
-  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
-  unite(ids, scan_id, feature, na.rm = T, sep = " ") %>%
-  column_to_rownames("ids") %>%
-  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
-           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
-dup %>%
-  group_by(location) %>%
-  summarise(across(A:C, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
-  column_to_rownames("location") %>%
-  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
-           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
-```
-
-# Singleton scans
-
-```{r}
-# Gold standard data
-df <- data %>%
-  arrange(scan_id, scorer) %>%
-  filter(!(scan_id %in% dups$original_ssid) &
-           !(scan_id %in% training$Cffid.Scoring.Set)) %>%
-  select(scorer, scan_id, bronchiectasis_rul:atelectasis___6) %>%
-  pivot_longer(bronchiectasis_rul:atelectasis___6,
-               names_to = c("feature", "location"),
-               names_pattern = "(.*)_(.*)$"
-  ) %>%
-  mutate(
-    feature = gsub("__", "", feature),
-    feature = str_to_title(gsub("_", " ", feature))
-  )
-df$location <- factor(df$location,
-                      levels = c(
-                        "1", "2", "3", "4", "5", "6", "ling", "lll",
-                        "lul", "rll", "rml", "rul"
-                      ),
-                      labels = c(
-                        "RUL", "RML", "RLL", "LUL", "Ling", "LLL",
-                        "Ling", "LLL", "LUL", "RLL", "RML", "RUL"
-                      )
-)
-# One column per scorer
-df <- df %>%
-  pivot_wider(names_from = scorer, values_from = value) %>%
-  arrange(feature, location)
-# Check agreement with GS column
-df[, c("A", "B", "C")] <- lapply(df[, c("A", "B", "C")], factor,
-                                       levels = c("Unchecked", "Checked",
-                                                  "Absent","Cylindric","Cystic",
-                                                  "Varicose"),
-                                       labels = c("Absent", "Present",
-                                                  "Absent", "Present", 
-                                                  "Present", "Present")
-)
-# Three-way agreement
-df$`A vs. B` <- factor(df$A == df$B, levels = c(F, T), labels = c("Disagree", "Agree"))
-df$`A vs. C` <- factor(df$A == df$C, levels = c(F, T), labels = c("Disagree", "Agree"))
-df$`B vs. C` <- factor(df$B == df$C, levels = c(F, T), labels = c("Disagree", "Agree"))
-# Plot
-df %>%
-  select(scan_id, feature, location, `A vs. B`:`B vs. C`) %>%
-  pivot_longer(`A vs. B`:`B vs. C`) %>%
-  ggplot(., aes(x = name, y = location, fill = value)) +
-  geom_tile(color = "black") +
-  theme_bw() +
-  labs(fill = "", x = "Scorer", y = "Location") +
-  facet_grid(rows = vars(feature), cols = vars(scan_id))
-# Tables
-df %>%
-  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
-  kable(., digits = 3, caption = "Proportion of agreement with themselves")
-df %>%
-  group_by(scan_id) %>%
-  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
-  kable(., digits = 3, caption = "Proportion of agreement with themselves")
-df %>%
-  group_by(feature) %>%
-  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
-  kable(., digits = 3, caption = "Proportion of agreement with themselves")
-df %>%
-  group_by(location) %>%
-  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "drop") %>%
-  kable(., digits = 3, caption = "Proportion of agreement with themselves")
-# Heatmaps
-df %>%
-  group_by(scan_id) %>%
-  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
-  column_to_rownames("scan_id") %>%
-  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
-           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
-df %>%
-  group_by(feature) %>%
-  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
-  column_to_rownames("feature") %>%
-  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
-           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
-df %>%
-  group_by(location) %>%
-  summarise(across(`A vs. B`:`B vs. C`, ~ mean(.x == "Agree", na.rm = T)), .groups = "keep") %>%
-  column_to_rownames("location") %>%
-  pheatmap(., cluster_rows = F, cluster_cols = F,breaks = seq(0.6,1,length.out = 100),
-           colorRampPalette((brewer.pal(n = 7, name = "RdYlBu")))(100))
-```
-
-# Comparison of scores between groups
-
-For scans that were duplicated, we looked at the first reading and excluded the second.
-For all features, absence was counted as 0 and presence as 1.
-Absence and presence refer to all lobes (e.g.
-bronchiectasis on RUL and RML would score a 1, and so would bronchiectasis on RUL, RML, and LUL).
-
-```{r}
-# Import disease state
-outcomes = read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/disease.key.csv",na.strings = "")
-outcomes$NTM.PD..Y.N. = factor(outcomes$NTM.PD..Y.N.,
-                               levels = c("No","Yes"),ordered = T)
-# Remove training scans, get first reading of duplicates, convert to numeric
-scores = data %>% 
-  filter(scan_id %in% outcomes$cff_registry_id,scorer != "GS") %>%
-  group_by(scorer,scan_id) %>%
-  slice_min(scoresheet_timestamp) %>% 
-  select(scorer, scan_id, bronchiectasis_rul:atelectasis___6) %>%
-  mutate(across(contains("bronchiectasis"),
-                ~as.numeric(factor(.x,levels=c("Absent","Cylindric",
-                                               "Varicose","Cystic"),
-                                   labels = c("Unchecked","Checked",
-                                              "Checked","Checked")))-1),
-         across(mucoid_impaction___1:atelectasis___6,
-                ~as.numeric(factor(.x,levels=c("Unchecked","Checked")))-1))
-# Add outcome info
-scores = left_join(scores,outcomes %>% select(cff_registry_id,NTM.PD..Y.N.),
-                   by = join_by(scan_id == cff_registry_id)) %>%
-  rename(`NTM-PD` = NTM.PD..Y.N.)
-# Calculate scores
-majority = scores %>%
-  group_by(scan_id) %>%
-  summarise(across(bronchiectasis_rul:`NTM-PD`,
-                   ~tail(names(sort(table(.x))), 1)))
-max = scores %>%
-  group_by(scan_id) %>%
-  summarise(across(bronchiectasis_rul:`NTM-PD`,
-                   ~max(.x,na.rm = T)))
-a = scores %>%
-  filter(scorer == "A")
-b = scores %>%
-  filter(scorer == "B") 
-c = scores %>%
-  filter(scorer == "C")
-```
-
-```{r}
-comp_score = function(df){
-  df = df %>% pivot_longer(bronchiectasis_rul:atelectasis___6,
-                           names_to = c("feature", "location"),
-                           names_pattern = "(.*)_(.*)$"
-  ) %>%
-    mutate(
-      feature = gsub("__", "", feature),
-      feature = str_to_title(gsub("_", " ", feature))
-    )
-  df$location <- factor(df$location,
-                        levels = c(
-                          "1", "2", "3", "4", "5", "6", "ling", "lll",
-                          "lul", "rll", "rml", "rul"
-                        ),
-                        labels = c(
-                          "RUL", "RML", "RLL", "LUL", "Ling", "LLL",
-                          "Ling", "LLL", "LUL", "RLL", "RML", "RUL"
-                        )
-  )
-  df$value = as.numeric(df$value)
-  # Tables
-  # Total score
-  sums = df %>% group_by(scan_id,`NTM-PD`) %>%
-    summarise(Total = sum(value),.groups = "drop") 
-  t1 = tableby(`NTM-PD`~Total,data = sums,numeric.test = "kwt",
-               numeric.stats = c("Nmiss2","median","q1q3","range"))
-  # By feature
-  sums = df %>% group_by(scan_id,`NTM-PD`,feature) %>%
-    summarise(score = sum(value),.groups = "drop") %>%
-    pivot_wider(names_from = feature,values_from = score)
-  f = as.formula(paste0("`NTM-PD`~",paste0(paste0("`",unique(df$feature),"`"),
-                                           collapse = "+")))
-  t2 = tableby(f,data = sums,numeric.test = "kwt",
-               numeric.stats = c("Nmiss2","median","q1q3","range"))
-  # By location
-  sums = df %>% group_by(scan_id,`NTM-PD`,location) %>%
-    summarise(score = sum(value),.groups = "drop") %>%
-    pivot_wider(names_from = location,values_from = score)
-  f = as.formula(paste0("`NTM-PD`~",paste0(paste0("`",unique(df$location),"`"),
-                                           collapse = "+")))
-  t3 = tableby(f,data = sums,numeric.test = "kwt",
-               numeric.stats = c("Nmiss2","median","q1q3","range"))
-  # Plots
-  total_plot = df %>% group_by(scan_id,`NTM-PD`,feature) %>%
-    summarise(score = sum(value),.groups = "drop") %>%
-    ggplot(.,aes(x = `NTM-PD`, y = score,fill = `NTM-PD`)) + 
-    geom_violin() +
-    theme_bw() + 
-    theme(legend.position = "none")
-  feature_plot = total_plot +
-    facet_wrap(~feature,scales = "free") 
-  location_plot = df %>% group_by(scan_id,`NTM-PD`,location) %>%
-    summarise(score = sum(value),.groups = "drop") %>%
-    ggplot(.,aes(x = `NTM-PD`, y = score,fill = `NTM-PD`)) + 
-    geom_violin() +
-    theme_bw() + 
-    theme(legend.position = "none") +
-    facet_wrap(~location,scales = "free") 
-  # Print
-  cat("\n")
-  cat("### Total Scores")
-  cat("\n")
-  cat("\n")
-  print(summary(t1,pfootnote = T))
-  cat("\n")
-  cat("\n")
-  print(total_plot)
-  cat("\n")
-  cat("\n")
-  cat("### Scores by Feature")
-  cat("\n")
-  cat("\n")
-  print(summary(t2,pfootnote = T))
-  cat("\n")
-  cat("\n")
-  print(feature_plot)
-  cat("\n")
-  cat("\n")
-  cat("### Scores by Location")
-  cat("\n")
-  cat("\n")
-  print(summary(t3,pfootnote = T))
-  cat("\n")
-  cat("\n")
-  print(location_plot)
-  cat("\n")
-  cat("\n")
-}
-```
-
-## Using majority score for each feature
-
-```{r results='asis'}
-comp_score(majority)
-```
-
-## Using maximum score for each feature
-
-```{r results='asis'}
-comp_score(max)
-```
-
-## Scorer A
-
-```{r results='asis'}
-comp_score(a)
-```
-
-## Scorer B
-
-```{r results='asis'}
-comp_score(b)
-```
-
-## Scorer C
-
-```{r results='asis'}
-comp_score(c)
-```
+| Value of Kappa | Level of Agreement | % of Data that are Reliable |
+| -------------- | ------------------ | --------------------------- |
+| 0–.20          | None               | 0–4%                        |
+| .21–.39        | Minimal            | 4–15%                       |
+| .40–.59        | Weak               | 15–35%                      |
+| .60–.79        | Moderate           | 35–63%                      |
+| .80–.90        | Strong             | 64–81%                      |
+| Above.90       | Almost Perfect     | 82–100%                     |

### 2023-08-03 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
quarto template formatting

--- 2023-06-12 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
@@
     page-layout: full
+    theme:
+      light: flatly
+      dark: darkly
+bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
+csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
 editor: source
@@
 
-```{r}
-library(DataExplorer)
+```{r libraries}
+#| include: false
+library(tidyverse)
 ```
 
-# EDA
+```{r data cleaning}
+# Most cleaning performed in GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R
+load("/Volumes/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
+```
+
+# Table
 
 ```{r}
-# Import
-df = read.csv("~/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_Annualized_Del1.csv")
-# Generate report
-# df %>%
-#   create_report(
-#     output_file = paste0("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Reports/Annualized EDA ", 
-#                          format(Sys.time(), "%Y-%m-%d"),".html"),
-#     report_title = "Annualized EDA - Racial and Ethnic Differences in ETI"
-#   )
-skimr::skim(df)
+#| label: tbl-planets
+#| tbl-cap: Planets
+
 ```
+
+# Figure
+
+```{r}
+#| label: fig-planets
+#| fig-cap: Planets
+
+```

### 2023-08-03 Quarto Template.qmd
quarto template formatting

--- 2023-03-14 Quarto Template.qmd
@@
     page-layout: full
+    theme:
+      light: flatly
+      dark: darkly
+bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
+csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
 editor: source
 ---
+
+```{r libraries}
+#| include: false
+library(tidyverse)
+```
+
+```{r data cleaning}
+
+```
 

### 2023-08-03 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R

@@ -3,9 +3,9 @@ library(lubridate)
 library(readxl)
 library(Hmisc)
 # Import
-annual <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_Annualized_Del1.csv", na.strings = "")
-encounter <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_encountersMerged_Del1.csv", na.strings = "")
-demo <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_DemogCFDiag_Del1.csv", na.strings = "")
+annual <- read.csv("/Volumes/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_Annualized_Del1.csv", na.strings = "")
+encounter <- read.csv("/Volumes/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_encountersMerged_Del1.csv", na.strings = "")
+demo <- read.csv("/Volumes/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_DemogCFDiag_Del1.csv", na.strings = "")
 # Fix missing values for specific bacteria
 bugs <- c(
   "staphylococcus_aureus", "haemophilus_influenzae",
@@ -215,7 +215,7 @@ demo$Hispanicrace <- factor(demo$Hispanicrace,
 )
 demo$MutClass <- factor(demo$MutClass)
 # Get long mutation lists
-mutations <- read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/Copy of Codes for CFFPR_2023.xlsx")
+mutations <- read_excel("/Volumes/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/Copy of Codes for CFFPR_2023.xlsx")
 mutations <- mutations %>%
[…]

### 2023-08-04 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
ETI combined race numbers

@@
 
-```{r libraries}
+```{r data cleaning}
 #| include: false
-library(tidyverse)
-```
-
-```{r data cleaning}
-# Most cleaning performed in GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R
+source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R")
 load("/Volumes/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
@@
 
-# Table
+```{r libraries}
+#| include: false
+library(Hmisc)
+library(tidyverse)
+library(gtsummary)
+```
+
+# Race and ethnicity breakdown
+
+```{r table 1}
+#| label: tbl-1
+#| tbl-cap: Race by year
+annual %>%
+  select(ReviewYear, race, Hispanicrace) %>%
+  tbl_summary(by = ReviewYear, missing_text = "Missing")
+```
+
+```{r table 2}
+#| label: tbl-2
+#| tbl-cap: Race and ethnicity
+demo %>%
+  select(race, Hispanicrace) %>%
+  tbl_summary(by = Hispanicrace) %>%
+  modify_header(label = paste0("**", label(annual$Hispanicrace), "**")) %>%
+  add_overall()
+```
+
+Based on the above table, I decided to combine race and ethnicity with Hispanic "overriding" other racial identities.
+So if a participant was reported as both Black and Hispanic, they were consider Hispanic for the following analyses.
+Those with missing data for Hispanic ethnicity were moved to the "Mixed/Other/Unknown Race" category.
+
+```{r combine race and eth}
+demo$race <- as.character(demo$race)
+demo$race[demo$Hispanicrace == "Yes"] <- "Hispanic or Latino"
+demo$race[demo$Hispanicrace == "Unknown"] <- "Hispanic or Latino"
+demo$race <- factor(demo$race,
+  levels = c(
+    "White", "Black or African American", "Hispanic or Latino",
+    "American Indian or Alaska Native", "Asian",
+    "Native Hawaiian or Other Pacific Islander", "Mixed/Other/Unknown Race"
+  )
+)
+label(demo$race) <- "Race/Ethnicity"
+```
+
+# Participant characteristics
 
 ```{r}
-#| label: tbl-planets
-#| tbl-cap: Planets
-
+#| label: tbl-3
+#| tbl-cap: Characteristics by race
+demo %>%
+  select(race) %>%
+  tbl_summary()
 ```
@@
 #| fig-cap: Planets
-
 ```

### 2023-08-04 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
tidyverse formatting

@@
 #| include: false
-source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R")
+# source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R")
 load("/Volumes/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
@@
 
-```{r table 1}
-#| label: tbl-1
-#| tbl-cap: Race by year
+## Race by year
+
+```{r t1}
 annual %>%
   select(ReviewYear, race, Hispanicrace) %>%
-  tbl_summary(by = ReviewYear, missing_text = "Missing")
+  tbl_summary(
+    by = ReviewYear, missing_text = "Missing",
+    statistic = list(all_continuous() ~ "{mean} ({sd})")
+  ) %>%
+  modify_header(label = paste0("**Year**")) %>%
+  bold_labels()
 ```
 
-```{r table 2}
-#| label: tbl-2
-#| tbl-cap: Race and ethnicity
+## Race and ethnicity overall
+
+```{r t2}
 demo %>%
   select(race, Hispanicrace) %>%
-  tbl_summary(by = Hispanicrace) %>%
+  tbl_summary(
+    by = Hispanicrace,
+    statistic = list(all_continuous() ~ "{mean} ({sd})")
+  ) %>%
   modify_header(label = paste0("**", label(annual$Hispanicrace), "**")) %>%
-  add_overall()
+  add_overall() %>%
+  bold_labels()
 ```
@@
 
+## Race percentages overall
+
+```{r t3}
+demo %>%
+  select(race) %>%
+  tbl_summary(statistic = list(all_continuous() ~ "{mean} ({sd})")) %>%
+  modify_header(label = paste0("**Race**")) %>%
+  bold_labels()
+```
+
 # Participant characteristics
 
-```{r}
-#| label: tbl-3
-#| tbl-cap: Characteristics by race
+## Characteristics by race
+
+```{r t4}
 demo %>%
-  select(race) %>%
-  tbl_summary()
+  select(race, Gender, WasGenotyped, F508) %>%
+  tbl_summary(
+    by = race, missing_text = "Missing",
+    statistic = list(all_continuous() ~ "{mean} ({sd})")
+  ) %>%
+  modify_header(label = paste0("**Race**")) %>%
+  add_overall() %>%
+  bold_labels()
 ```
 
-# Figure
+```{r summarise by year}
+#| cache: true
+t5 <- encounter %>%
+  group_by(eDWID, reviewyear) %>%
+  summarise(across(c(race, Gender, WasGenotyped, F508, contains("insurance")), first),
+    `Age at Last Visit` = last(na.omit(encounterage)),
+    `Mean ppFEV` = mean(GLI_FEV1_pct_predicted, na.rm = T),
+    `Weight at Last Visit` = last(na.omit(weight)),
+    `Weight %ile at Last Visit` = last(na.omit(weightpercentile)),
+    `BMI at Last Visit` = last(na.omit(bmivalue)),
+    `BMI %ile at Last Visit` = last(na.omit(bmipercentile)),
+    .groups = "drop"
+  )
+label(t5)[colnames(t5) %in% names(labels)] <- labels[match(colnames(t5), names(labels), nomatch = 0)]
+```
+
+## Characteristics by year
+
+```{r t5}
+#| message: false
+t5 %>%
+  select(
+    -eDWID, reviewyear, race, Gender, WasGenotyped, F508, contains("insurance"),
+    `Age at Last Visit`, `Mean ppFEV`, `Weight at Last Visit`,
+    `Weight %ile at Last Visit`, `BMI at Last Visit`,
+    `BMI %ile at Last Visit`
+  ) %>%
+  tbl_summary(
+    by = reviewyear, missing_text = "Missing",
+    statistic = list(all_continuous() ~ "{mean} ({sd})")
+  ) %>%
+  modify_header(label = paste0("**Year**")) %>%
+  bold_labels()
+```
 

### 2023-08-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
table formatting

@@
   modify_header(label = paste0("**Year**")) %>%
-  bold_labels()
+  bold_labels() %>%
+  as_kable_extra()
 ```
@@
   add_overall() %>%
-  bold_labels()
+  bold_labels() %>%
+  as_kable_extra()
 ```
@@
   modify_header(label = paste0("**Race**")) %>%
-  bold_labels()
+  bold_labels() %>%
+  as_kable_extra()
 ```

### 2023-08-19 Emily Muther/mental_health_screening.qmd
cf mental health ZINB not working

--- 2023-04-20 Emily Muther/mental_health_screening.qmd
@@
 
+```{r}
+#| include: false
+library(tidyverse)
+library(plotly)
+library(gtsummary)
+library(GLMMadaptive)
+library(lmerTest)
+library(broom.mixed)
+library(knitr)
+library(sjPlot)
+library(report)
+```
+
 # Data cleaning
@@
 
-```{r include=FALSE}
-library(tidyverse)
-library(plotly)
-library(gtsummary)
-library(GLMMadaptive)
+```{r}
+#| warning: false
 # Import
-df = read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Cleaned/MH master data.csv",check.names = F,na.strings = "")
+df = read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Emily Muther/Mental Health Screening/Data_Cleaned/MH master data.csv",check.names = F,na.strings = "")
 # Column types
+df$MRN = factor(df$MRN)
 df$CONTACT_DATE = lubridate::mdy(df$CONTACT_DATE)
 df$`PHQ-9 Total Score (calculated)` = as.integer(df$`PHQ-9 Total Score (calculated)`)
+df$`GAD-7 Total Score` = as.integer(df$`GAD-7 Total Score`)
 # Visits
@@
 
-# Methods
+# Results
 
-Because the PHQ-9 results are right-skewed with lots of 0 scores, we used a zero inflated Poisson mixed model to analyze change in score over time.
+## ZINB mixed effects model
+
+Because the PHQ-9 results are right-skewed with lots of 0 scores, we first tried a zero inflated negative binomial (ZINB) mixed model to analyze change in score over time.
+At first we tried using categorical visit number as a time variable, but the model did not converge (likely due to the number of parameters).
+
+```{r}
+#| warning: false
+try(mixed_model(`PHQ-9 Total Score (calculated)` ~ factor(`Visit Number`),random = ~1|MRN,data = df,family = zi.negative.binomial(), zi_fixed = ~ 1))
+
+try(mixed_model(`PHQ-9 Total Score (calculated)` ~ `Days From Baseline`,
+                 random = ~1|MRN,data = df,family = zi.negative.binomial(), 
+                 zi_fixed = ~ 1))
+```
+
+Because the ZINB models were not converging, we instead tried to use a simpler Poisson mixed model (without accounting for the inflated 0s) with random intercept per participant.
+Visit number was treated as a continuous variable because this model would not converge with categorical visit numbers.
+
+```{r}
+#| warning: false
+m2 = glmer(`PHQ-9 Total Score (calculated)` ~ `Visit Number` + (1|MRN),data = df,family = "poisson")
+plot_model(m2,type="pred",
+           terms=c("Visit Number","MRN"),
+           pred.type="re", ci.lvl = NA) + 
+  theme(legend.position = "none")
+res = tidy(m2,exponentiate = T,effects="fixed")
+kable(res,digits = 3)
+```
+
+## Simpler models
+
+Because the complex models were difficult to fit and not particularly informative, we instead modeled the odds of an elevated score for the full cohort, and the change in score over time for those with elevated scores.
+Estimates are reported as odds ratios in tables, but are reported on the log scale in the text, just in case both are needed.
+
+### PHQ9
+
+```{r}
+df$PHQ9_elevated = factor(df$`PHQ-9 Total Score (calculated)` >= 5,
+                          levels = c(F,T),labels = c("No","Yes"))
+```
+
+A PHQ9 score $\geq$ 5 was considered elevated.
+
+#### Odds of an elevated score
+
+```{r}
+mod = glmer(PHQ9_elevated ~ `Visit Number` + (1|MRN),data = df,family = "binomial")
+res = tidy(mod,exponentiate = T,effects="fixed")
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+#### Association with visit number in those with elevated scores.
+
+Results are reported as expected counts.
+
+```{r}
+mod = glmer(`PHQ-9 Total Score (calculated)`~ `Visit Number` + (1|MRN),
+            data = df[df$PHQ9_elevated=="Yes",],family = "poisson")
+res = tidy(mod,exponentiate = T,effects="fixed")
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+### GAD7
+
+```{r}
+df$GAD7_elevated = factor(df$`GAD-7 Total Score` >= 4,
+                          levels = c(F,T),labels = c("No","Yes"))
+```
+
+A GAD7 score $\geq$ 4 was considered elevated.
+
+#### Odds of an elevated score
+
+```{r}
+mod = glmer(GAD7_elevated ~ `Visit Number` + (1|MRN),data = df,family = "binomial")
+res = tidy(mod,exponentiate = T,effects="fixed")
+kable(res,digits = 3)
+```
+
+`r report(mod)`
+
+#### Association with visit number in those with elevated scores.
+
+Results are reported as expected counts.
+
+```{r}
+mod = glmer(`GAD-7 Total Score`~ `Visit Number` + (1|MRN),
+            data = df[df$GAD7_elevated=="Yes",],family = "poisson")
+res = tidy(mod,exponentiate = T,effects="fixed")
+kable(res,digits = 3)
+```
+
+`r report(mod)`

### 2023-08-23 Stacey Martiniano/ntm_agreement.qmd
finished initial agreement, started LMMs

@@
       dark: darkly
-bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
+bibliography: /Users/timvigers/Library/CloudStorage/Dropbox/Miscellaneous/zotero.bib
+csl: /Users/timvigers/Library/CloudStorage/Dropbox/Miscellaneous/american-medical-association.csl
 editor: source
@@
 
-```{r}
-#| include: false
+```{r data cleaning}
+#| message: false
 library(tidyverse)
-library(irr)
 library(redcapAPI)
 library(Hmisc)
-library(DT)
-library(RColorBrewer)
+library(irr)
 library(knitr)
-library(pheatmap)
-library(arsenal)
 # Import data
-api <- read.table("/Users/timvigers/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt",
-                  header = F
-)
+api <- read.table("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt", header = F)
 rcon <- redcapConnection(
@@
 # Revert duplicate IDs back to original to match with gold standard
-training <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/training.key.csv")
-dups <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/duplicated.key.csv")
+training <- read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/training.key.csv")
+dups <- read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/duplicated.key.csv")
 training$Training_id <- as.numeric(sub("T00", "", training$Training_id))
@@
 # Convert to long
-data = data %>% select(scan_id,scorer,scoresheet_timestamp,
-                    bronchiectasis_rul:atelectasis___6) %>%
-  arrange(scan_id,scorer,scoresheet_timestamp) %>%
-  group_by(scan_id,scorer) %>%
+data <- data %>%
+  select(
+    scan_id, scorer, scoresheet_timestamp,
+    bronchiectasis_rul:atelectasis___6
+  ) %>%
+  arrange(scan_id, scorer, scoresheet_timestamp) %>%
+  group_by(scan_id, scorer) %>%
   mutate(scorer_repeat = row_number()) %>%
@@
 )
-# Maybe not quite so long (one column per rater)
-data = data %>% 
-  pivot_wider(names_from = c(scorer,scorer_repeat), values_from = value,
-              names_sep = "_")
+# Format values
+data$value <- factor(data$value,
+  levels = c(
+    "Absent", "Checked", "Cylindric", "Cystic", "Unchecked", "Varicose"
+  ),
+  labels = c("Absent", "Present", "Present", "Present", "Absent", "Present"),
+  ordered = T
+)
+# Separate out gold standards and duplicates
+data_dups <- data %>% filter(scan_id %in% dups$original_ssid)
+data_gs <- data %>% filter(scan_id %in% training$Cffid.Scoring.Set)
+data <- data %>%
+  filter(!scan_id %in% c(dups$original_ssid, training$Cffid.Scoring.Set))
+# Second readings only for comparison to GS
+data_gs <- data_gs %>%
+  group_by(scan_id, scorer, feature, location) %>%
+  slice_max(scorer_repeat)
 ```
 
-Kappa interpretation table @mchughInterraterReliabilityKappa2012:
+# Proprortion of agreement
+
+## Singletons
+
+### By feature
+
+```{r}
+#| message: false
+data %>%
+  group_by(feature, scan_id, scorer) %>%
+  summarise(v = max(value, na.rm = T)) %>%
+  summarise(a = sum(v == "Present")) %>%
+  summarise(
+    agreed = sum(a == 0 | a == 3),
+    disagreed = sum(a == 1 | a == 2),
+    proportion = round(agreed / (agreed + disagreed), 3)
+  ) %>%
+  kable()
+```
+
+### Feature by lobe
+
+```{r}
+#| message: false
+data %>%
+  group_by(feature, location, scan_id) %>%
+  summarise(a = factor(sum(value == "Present", na.rm = T))) %>%
+  summarise(
+    agreed = sum(a == 0 | a == 3),
+    disagreed = sum(a == 1 | a == 2),
+    proportion = round(agreed / (agreed + disagreed), 3)
+  ) %>%
+  kable()
+```
+
+## Duplicates
+
+### By feature
+
+```{r}
+#| message: false
+data_dups %>%
+  group_by(feature, scorer, scan_id, scorer_repeat) %>%
+  summarise(v = max(value, na.rm = T)) %>%
+  summarise(a = v[scorer_repeat == 1] == v[scorer_repeat == 2]) %>%
+  summarise(p = mean(a)) %>%
+  pivot_wider(
+    names_from = scorer, values_from = p,
+    names_prefix = "prop. agreed "
+  ) %>%
+  kable()
+```
+
+### Feature by lobe
+
+```{r}
+#| message: false
+#| warning: false
+data_dups %>%
+  group_by(feature, location, scorer, scan_id, scorer_repeat) %>%
+  summarise(v = max(value, na.rm = T)) %>%
+  summarise(a = v[scorer_repeat == 1] == v[scorer_repeat == 2]) %>%
+  summarise(p = mean(a, na.rm = T)) %>%
+  pivot_wider(
+    names_from = scorer, values_from = p,
+    names_prefix = "prop. agreed "
+  ) %>%
+  kable()
+```
+
+## Agreement with gold standard
+
+### By feature
+
+```{r}
+#| message: false
+data_gs %>%
+  group_by(feature, scan_id, scorer) %>%
+  summarise(v = max(value, na.rm = T)) %>%
+  summarise(
+    A = v[scorer == "A"] == v[scorer == "GS"],
+    B = v[scorer == "B"] == v[scorer == "GS"],
+    C = v[scorer == "C"] == v[scorer == "GS"]
+  ) %>%
+  summarise(
+    `prop. agreed A` = sum(A) / n(),
+    `prop. agreed B` = sum(B) / n(),
+    `prop. agreed C` = sum(C) / n()
+  ) %>%
+  kable()
+```
+
+### Feature by lobe
+
+Missing values were ignored in creating the table below.
+So if a scorer marked a feature and location blank on one scan and agreed with the gold standard on 3, their score will be 0.75 rather than 0.6.
+
+```{r}
+#| message: false
+data_gs %>%
+  group_by(feature, location, scan_id) %>%
+  summarise(
+    A = value[scorer == "A"] == value[scorer == "GS"],
+    B = value[scorer == "B"] == value[scorer == "GS"],
+    C = value[scorer == "C"] == value[scorer == "GS"]
+  ) %>%
+  summarise(
+    `prop. agreed A` = mean(A, na.rm = T),
+    `prop. agreed B` = mean(B, na.rm = T),
+    `prop. agreed C` = mean(C, na.rm = T)
+  ) %>%
+  kable()
+```
+
+# Using Cohen's kappa
 
@@
 | Above.90       | Almost Perfect     | 82–100%                     |
+
+: Kappa interpretation table @mchughInterraterReliabilityKappa2012a {#tbl-kappa}
+
+## Singletons
+
+### By feature
+
+```{r}
+#| message: false
+kappa <- data %>%
+  group_by(feature, scan_id, scorer) %>%
+  summarise(v = max(value, na.rm = T), .groups = "drop") %>%
+  pivot_wider(names_from = scorer, values_from = v)
+kappa <- split.data.frame(kappa, ~feature)
+kappas <- lapply(names(kappa), function(n) {
+  d <- kappa[[n]]
+  k <- kappam.fleiss(d[, c("A", "B", "C")])
+  return(c(n, k$value, k$p.value))
+})
+kappas <- data.frame(do.call(rbind, kappas))
+colnames(kappas) <- c("feature", "kappa", "p")
+kappas$kappa <- as.numeric(kappas$kappa)
+kappas$p <- format.pval(as.numeric(kappas$p), digits = 3, eps = 0.001)
+kable(kappas, digits = 3)
+```
+
+### Feature by lobe
+
+Because you technically cannot calculate Cohen's kappa for matrices where every cell is the same, values of `-Inf` can be interpreted as perfect agreement for a feature that had the same value across all scans.
+For example, none of the 15 scans showed consolidation in the	RML.
+
+```{r}
+#| message: false
+kappa <- data %>%
+  pivot_wider(names_from = scorer, values_from = value)
+kappa <- split.data.frame(kappa, ~ feature + location)
+kappas <- lapply(names(kappa), function(n) {
+  d <- kappa[[n]]
+  k <- kappam.fleiss(d[, c("A", "B", "C")])
+  return(c(strsplit(n, "\\.")[[1]], k$value, k$p.value))
+})
+kappas <- data.frame(do.call(rbind, kappas))
+colnames(kappas) <- c("feature", "location", "kappa", "p")
+kappas$kappa <- as.numeric(kappas$kappa)
+kappas$p <- format.pval(as.numeric(kappas$p), digits = 3, eps = 0.001)
+kable(kappas, digits = 3)
+```
+
+## Duplicates
+
+Unfortunately, because there weren't many duplicate scans it was difficult to calculate kappa for many features.
+
+### By feature
+
+```{r}
+#| message: false
+kappa <- data_dups %>%
+  group_by(feature, scan_id, scorer, scorer_repeat) %>%
+  summarise(v = max(value, na.rm = T), .groups = "drop") %>%
+  pivot_wider(names_from = c(scorer, scorer_repeat), values_from = v)
+kappa <- split.data.frame(kappa, ~feature)
+kappas <- lapply(names(kappa), function(n) {
+  d <- kappa[[n]]
+  ka <- kappa2(d[, c("A_1", "A_2")])
+  kb <- kappa2(d[, c("B_1", "B_2")])
+  kc <- kappa2(d[, c("C_1", "C_2")])
+  return(c(
+    strsplit(n, "\\.")[[1]],
+    ka$value, ka$p.value,
+    kb$value, kb$p.value,
+    kc$value, kb$p.value
+  ))
+})
+kappas <- data.frame(do.call(rbind, kappas))
+colnames(kappas) <- c(
+  "feature", "kappa A", "p A", "kappa B", "p B",
+  "kappa C", "p C"
+)
+kappas[, !grepl("feature", colnames(kappas))] <- lapply(kappas[, !grepl("feature", colnames(kappas))], as.numeric)
+kappas[, grep("p ", colnames(kappas))] <- lapply(kappas[, grep("p ", colnames(kappas))],
+  format.pval,
+  digits = 3, eps = 0.001
+)
+kable(kappas, digits = 3)
+```
+
+### Feature by lobe
+
+```{r}
+#| message: false
+kappa <- data_dups %>%
+  group_by(feature,location, scan_id, scorer, scorer_repeat) %>%
+  pivot_wider(names_from = c(scorer, scorer_repeat), values_from = value)
+kappa <- split.data.frame(kappa, ~feature+location)
+kappas <- lapply(names(kappa), function(n) {
+  d <- kappa[[n]]
+  ka <- kappa2(d[, c("A_1", "A_2")])
+  kb <- kappa2(d[, c("B_1", "B_2")])
+  kc <- kappa2(d[, c("C_1", "C_2")])
+  return(c(
+    strsplit(n, "\\.")[[1]], 
+    ka$value, ka$p.value,
+    kb$value, kb$p.value,
+    kc$value, kb$p.value
+  ))
+})
+kappas <- data.frame(do.call(rbind, kappas))
+colnames(kappas) <- c(
+  "feature", "location","kappa A", "p A", "kappa B", "p B",
+  "kappa C", "p C"
+)
+kappas[, !grepl("feature|location", colnames(kappas))] <- 
+  lapply(kappas[, !grepl("feature|location", colnames(kappas))], as.numeric)
+kappas[, grep("p ", colnames(kappas))] <- lapply(kappas[, grep("p ", colnames(kappas))],
+  format.pval,
+  digits = 3, eps = 0.001
+)
+kable(kappas, digits = 3)
+```
+
+## Agreement with gold standard
+
+Unfortunately, because there weren't many gold standard scans it was difficult to calculate kappa for many features.
+
+### By feature
+
+```{r}
+#| message: false
+kappa <- data_gs %>%
+  group_by(feature, scan_id, scorer) %>%
+  summarise(v = max(value, na.rm = T), .groups = "drop") %>%
+  pivot_wider(names_from = scorer, values_from = v)
+kappa <- split.data.frame(kappa, ~feature)
+kappas <- lapply(names(kappa), function(n) {
+  d <- kappa[[n]]
+  ka <- kappa2(d[, c("A", "GS")])
+  kb <- kappa2(d[, c("B", "GS")])
+  kc <- kappa2(d[, c("C", "GS")])
+  return(c(
+    strsplit(n, "\\.")[[1]],
+    ka$value, ka$p.value,
+    kb$value, kb$p.value,
+    kc$value, kb$p.value
+  ))
+})
+kappas <- data.frame(do.call(rbind, kappas))
+colnames(kappas) <- c(
+  "feature", "kappa A", "p A", "kappa B", "p B",
+  "kappa C", "p C"
+)
+kappas[, !grepl("feature", colnames(kappas))] <- lapply(kappas[, !grepl("feature", colnames(kappas))], as.numeric)
+kappas[, grep("p ", colnames(kappas))] <- lapply(kappas[, grep("p ", colnames(kappas))],
+  format.pval,
+  digits = 3, eps = 0.001
+)
+kable(kappas, digits = 3)
+```
+
+### Feature by lobe
+
+```{r}
+#| message: false
+kappa <- data_gs %>%
+  group_by(feature,location, scan_id, scorer) %>%
+  pivot_wider(names_from = scorer, values_from = value)
+kappa <- split.data.frame(kappa, ~feature+location)
+kappas <- lapply(names(kappa), function(n) {
+  d <- kappa[[n]]
+  ka <- kappa2(d[, c("A", "GS")])
+  kb <- kappa2(d[, c("B", "GS")])
+  kc <- kappa2(d[, c("C", "GS")])
+  return(c(
+    strsplit(n, "\\.")[[1]], 
+    ka$value, ka$p.value,
+    kb$value, kb$p.value,
+    kc$value, kb$p.value
+  ))
+})
+kappas <- data.frame(do.call(rbind, kappas))
+colnames(kappas) <- c(
+  "feature", "location","kappa A", "p A", "kappa B", "p B",
+  "kappa C", "p C"
+)
+kappas[, !grepl("feature|location", colnames(kappas))] <- 
+  lapply(kappas[, !grepl("feature|location", colnames(kappas))], as.numeric)
+kappas[, grep("p ", colnames(kappas))] <- lapply(kappas[, grep("p ", colnames(kappas))],
+  format.pval,
+  digits = 3, eps = 0.001
+)
+kable(kappas, digits = 3)
+```

### 2023-08-23 Stacey Martiniano/ntm_agreement_lmm.qmd
finished initial agreement, started LMMs

---
title: "NTM CT Scoring - Mixed Models"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Library/CloudStorage/Dropbox/Miscellaneous/zotero.bib
csl: /Users/timvigers/Library/CloudStorage/Dropbox/Miscellaneous/american-medical-association.csl
editor: source
---

```{r data cleaning}
#| message: false
library(tidyverse)
library(redcapAPI)
library(Hmisc)
library(irr)
library(knitr)
# Import data
api <- read.table("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt", header = F)
rcon <- redcapConnection(
  url = "https://redcap.ucdenver.edu/api/",
  token = api[1, 1]
)
data <- exportRecordsTyped(rcon, labels = F)
data <- data.frame(lapply(data, as.character))
data$scan_id <- as.numeric(data$scan_id)
# Make scorer IDs by initial
data$scorer <- paste0(
  toupper(substr(data$reader_first_name, 1, 1)),
  toupper(substr(data$reader_last_name, 1, 1))
)
data$scorer[data$scorer == "GG"] <- "GS"
# Deidentify readers
old <- c("AF", "DL", "JW")
new <- c("A", "B", "C")
data$scorer[data$scorer %in% old] <- new[match(data$scorer, old, nomatch = 0)]
# Revert duplicate IDs back to original to match with gold standard
training <- read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/training.key.csv")
dups <- read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/duplicated.key.csv")
training$Training_id <- as.numeric(sub("T00", "", training$Training_id))
# Change duplicates to OG IDs
old <- dups$duplicated_ssid..dates.shifted..17.days.
new <- dups$original_ssid
data$scan_id[data$scan_id %in% old] <- new[na.omit(match(data$scan_id, old))]
# Change training IDs to OG IDs
old <- training$Training_id
new <- training$Cffid.Scoring.Set
data$scan_id[data$scan_id %in% old] <- new[na.omit(match(data$scan_id, old))]
# Convert to long
data <- data %>%
  select(
    scan_id, scorer, scoresheet_timestamp,
    bronchiectasis_rul:atelectasis___6
  ) %>%
  arrange(scan_id, scorer, scoresheet_timestamp) %>%
  group_by(scan_id, scorer) %>%
  mutate(scorer_repeat = row_number()) %>%
  select(-scoresheet_timestamp) %>%
  pivot_longer(bronchiectasis_rul:atelectasis___6,
    names_to = c("feature", "location"),
    names_pattern = "(.*)_(.*)$"
  ) %>%
  mutate(
    feature = gsub("__", "", feature),
    feature = str_to_title(gsub("_", " ", feature))
  )
data$location <- factor(data$location,
  levels = c(
    "1", "2", "3", "4", "5", "6", "ling", "lll",
    "lul", "rll", "rml", "rul"
  ),
  labels = c(
    "RUL", "RML", "RLL", "LUL", "Ling", "LLL",
    "Ling", "LLL", "LUL", "RLL", "RML", "RUL"
  )
)
# Format values
data$value <- factor(data$value,
  levels = c(
    "Absent", "Checked", "Cylindric", "Cystic", "Unchecked", "Varicose"
  ),
  labels = c("Absent", "Present", "Present", "Present", "Absent", "Present")
)
data$value = as.numeric(data$value)-1
# Separate out gold standards and duplicates
data_dups <- data %>% filter(scan_id %in% dups$original_ssid)
data_gs <- data %>% filter(scan_id %in% training$Cffid.Scoring.Set)
data <- data %>%
  filter(!scan_id %in% c(dups$original_ssid, training$Cffid.Scoring.Set))
# Second readings only for comparison to GS
data_gs <- data_gs %>%
  group_by(scan_id, scorer, feature, location) %>%
  slice_max(scorer_repeat)
```

### 2023-08-24 Stacey Martiniano/ntm_agreement_lmm.qmd
mixed model convergence warnings?

@@
 library(Hmisc)
-library(irr)
+library(fastDummies)
+library(lmerTest)
 library(knitr)
@@
 )
-data$value = as.numeric(data$value)-1
+data$value <- as.numeric(data$value) - 1
 # Separate out gold standards and duplicates
@@
   slice_max(scorer_repeat)
+# Dummy variables for scorer, organize columns
+data <- data %>%
+  dummy_cols(select_columns = "scorer") %>%
+  select(scan_id, feature, location, scorer_A, scorer_B, scorer_C, value)
 ```
+
+# Data setup
+
+```{r}
+DT::datatable(data)
+```
+
+# Model
+
+- Readers $r=1,2,3$
+- Scans $s=1,2,...,S$
+- Features $f=1,2,...,F$
+- Lobes $l=1,2,...,6$
+- Outcome $y_{rsfl}$ is binary
+- Model notation: $$y_{rsf=1}=\alpha+\beta X_l + b_1+b_2+b_3+\epsilon_{rs}$$
+
+## Tree in Bud
+
+```{r}
+#| code-fold: false
+df <- data %>% filter(feature == "Tree In Bud")
+model <- glmer(value ~ location + (1 | scorer_A) + (1 | scorer_B) + (1 | scorer_C),
+  family = "binomial", data = df
+)
+```

### 2023-08-28 Stacey Martiniano/ntm_agreement_lmm.qmd
agreement LMMs

@@
       dark: darkly
-bibliography: /Users/timvigers/Library/CloudStorage/Dropbox/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Library/CloudStorage/Dropbox/Miscellaneous/american-medical-association.csl
+bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
+csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
 editor: source
@@
 # Import data
-api <- read.table("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt", header = F)
+api <- read.table("/Users/timvigers/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt", header = F)
 rcon <- redcapConnection(
@@
 
-## Tree in Bud
+## Bronchiectasis
 
 ```{r}
-#| code-fold: false
+df <- data %>% filter(feature == "Bronchiectasis")
+model <- glmer(value ~ location + (1 | scorer_A) + (1 | scorer_B) + (1 | scorer_C),
+  family = "binomial", data = df
+)
+summary(model)
+```
+
+## Mucoid Impaction
+
+```{r}
+df <- data %>% filter(feature == "Mucoid Impaction")
+model <- glmer(value ~ location + (1 | scorer_A) + (1 | scorer_B) + (1 | scorer_C),
+  family = "binomial", data = df
+)
+summary(model)
+```
+
+## Cyst
+
+```{r}
+df <- data %>% filter(feature == "Cyst")
+model <- glmer(value ~ location + (1 | scorer_A) + (1 | scorer_B) + (1 | scorer_C),
+  family = "binomial", data = df
+)
+summary(model)
+```
+
+## Cavity
+
+```{r}
+df <- data %>% filter(feature == "Cavity")
+model <- glmer(value ~ location + (1 | scorer_A) + (1 | scorer_B) + (1 | scorer_C),
+  family = "binomial", data = df
+)
+summary(model)
+```
+
+## Consolidation
+
+```{r}
+df <- data %>% filter(feature == "Consolidation")
+model <- glmer(value ~ location + (1 | scorer_A) + (1 | scorer_B) + (1 | scorer_C),
+  family = "binomial", data = df
+)
+summary(model)
+```
+
+## Ground Glass
+
+```{r}
+df <- data %>% filter(feature == "Ground Glass")
+model <- glmer(value ~ location + (1 | scorer_A) + (1 | scorer_B) + (1 | scorer_C),
+  family = "binomial", data = df
+)
+summary(model)
+```
+
+## Nodule
+
+```{r}
+df <- data %>% filter(feature == "Nodule")
+model <- glmer(value ~ location + (1 | scorer_A) + (1 | scorer_B) + (1 | scorer_C),
+  family = "binomial", data = df
+)
+summary(model)
+```
+
+## Tree In Bud
+
+```{r}
 df <- data %>% filter(feature == "Tree In Bud")
@@
 )
+summary(model)
 ```
+
+## Atelectasis
+
+```{r}
+df <- data %>% filter(feature == "Atelectasis")
+model <- glmer(value ~ location + (1 | scorer_A) + (1 | scorer_B) + (1 | scorer_C),
+  family = "binomial", data = df
+)
+summary(model)
+```

### 2023-08-29 Stacey Martiniano/ntm_agreement.qmd
combine NTM agreement files

--- 2023-07-31 Stacey Martiniano/ntm_agreement.qmd
@@
 ---
-title: "NTM CT Scoring - Initial Agreement"
+title: "NTM CT Scoring Agreement"
 author: "Tim Vigers"
@@
 
-```{r}
-#| include: false
+```{r data cleaning}
+#| message: false
 library(tidyverse)
-library(irr)
 library(redcapAPI)
 library(Hmisc)
-library(DT)
-library(RColorBrewer)
 library(knitr)
-library(pheatmap)
-library(arsenal)
+library(fastDummies)
+library(lmerTest)
+library(performance)
 # Import data
-api <- read.table("/Users/timvigers/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt",
-                  header = F
-)
+api <- read.table("/Users/timvigers/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt", header = F)
 rcon <- redcapConnection(
@@
 # Convert to long
-data = data %>% select(scan_id,scorer,scoresheet_timestamp,
-                    bronchiectasis_rul:atelectasis___6) %>%
-  arrange(scan_id,scorer,scoresheet_timestamp) %>%
-  group_by(scan_id,scorer) %>%
+data <- data %>%
+  select(
+    scan_id, scorer, scoresheet_timestamp,
+    bronchiectasis_rul:atelectasis___6
+  ) %>%
+  arrange(scan_id, scorer, scoresheet_timestamp) %>%
+  group_by(scan_id, scorer) %>%
+  mutate(scorer_repeat = row_number()) %>%
+  select(-scoresheet_timestamp) %>%
+  pivot_longer(bronchiectasis_rul:atelectasis___6,
+    names_to = c("Feature", "Location"),
+    names_pattern = "(.*)_(.*)$"
+  ) %>%
+  mutate(
+    Feature = gsub("__", "", Feature),
+    Feature = str_to_title(gsub("_", " ", Feature))
+  )
+data$Location <- factor(data$Location,
+  levels = c(
+    "1", "2", "3", "4", "5", "6", "ling", "lll",
+    "lul", "rll", "rml", "rul"
+  ),
+  labels = c(
+    "RUL", "RML", "RLL", "LUL", "Ling", "LLL",
+    "Ling", "LLL", "LUL", "RLL", "RML", "RUL"
+  )
+)
+# Format values
+data$value <- factor(data$value,
+  levels = c(
+    "Absent", "Checked", "Cylindric", "Cystic", "Unchecked", "Varicose"
+  ),
+  labels = c("Absent", "Present", "Present", "Present", "Absent", "Present"),
+  ordered = T
+)
+# Separate out gold standards and duplicates
+data_dups <- data %>% filter(scan_id %in% dups$original_ssid)
+data_gs <- data %>% filter(scan_id %in% training$Cffid.Scoring.Set)
+data <- data %>%
+  filter(!scan_id %in% c(dups$original_ssid, training$Cffid.Scoring.Set))
+# Second readings only for comparison to GS
+data_gs <- data_gs %>%
+  group_by(scan_id, scorer, Feature, Location) %>%
+  slice_max(scorer_repeat)
+```
+
+# Proportion of agreement
+
+## Singletons
+
+### By Feature
+
+```{r}
+#| message: false
+data %>%
+  group_by(Feature, scan_id, scorer) %>%
+  summarise(v = max(value, na.rm = T)) %>%
+  summarise(a = sum(v == "Present")) %>%
+  summarise(
+    `Agreed Feature Was Present` = sum(a == 3),
+    `Agreed Feature Was Absent` = sum(a == 0),
+    agreed = sum(a == 0 | a == 3),
+    disagreed = sum(a == 1 | a == 2),
+    proportion = round(agreed / (agreed + disagreed), 3)
+  ) %>%
+  rename(`Total Agreed` = agreed,`Total Disagreed`=disagreed,
+         `Proportion Agreed`=proportion) %>%
+  kable()
+```
+
+### Feature by lobe
+
+```{r}
+#| message: false
+data %>%
+  group_by(Feature, Location, scan_id) %>%
+  summarise(a = factor(sum(value == "Present", na.rm = T))) %>%
+  summarise(
+    `Agreed Feature Was Present` = sum(a == 3),
+    `Agreed Feature Was Absent` = sum(a == 0),
+    agreed = sum(a == 0 | a == 3),
+    disagreed = sum(a == 1 | a == 2),
+    proportion = round(agreed / (agreed + disagreed), 3)
+  ) %>%
+  rename(`Total Agreed` = agreed,`Total Disagreed`=disagreed,
+         `Proportion Agreed`=proportion) %>%
+  kable()
+```
+
+## Duplicates
+
+### By Feature
+
+```{r}
+#| message: false
+data_dups %>%
+  group_by(Feature, scorer, scan_id, scorer_repeat) %>%
+  summarise(v = max(value, na.rm = T)) %>%
+  summarise(
+    `Agreed Feature Was Present` = sum(v[scorer_repeat == 1] == v[scorer_repeat == 2] & 
+                                         v[scorer_repeat == 1] == "Present"),
+    `Agreed Feature Was Absent` = sum(v[scorer_repeat == 1] == v[scorer_repeat == 2] & 
+                                        v[scorer_repeat == 1] == "Absent"),
+    agreed = sum(v[scorer_repeat == 1] == v[scorer_repeat == 2]),
+    disagreed = sum(v[scorer_repeat == 1] != v[scorer_repeat == 2]),
+    proportion = round(agreed / (agreed + disagreed), 3)
+  ) %>% 
+  summarise(across(`Agreed Feature Was Present`:disagreed,~sum(.x,na.rm = T)),
+            proportion=mean(proportion)) %>%
+  rename(`Total Agreed` = agreed,`Total Disagreed`=disagreed,
+         `Proportion Agreed`=proportion) %>%
+  kable()
+```
+
+### Feature by lobe
+
+```{r}
+#| message: false
+#| warning: false
+data_dups %>%
+  group_by(Feature, Location, scorer, scan_id, scorer_repeat) %>%
+  summarise(v = max(value, na.rm = T)) %>%
+  summarise(
+    `Agreed Feature Was Present` = sum(v[scorer_repeat == 1] == v[scorer_repeat == 2] & 
+                                         v[scorer_repeat == 1] == "Present"),
+    `Agreed Feature Was Absent` = sum(v[scorer_repeat == 1] == v[scorer_repeat == 2] & 
+                                        v[scorer_repeat == 1] == "Absent"),
+    agreed = sum(v[scorer_repeat == 1] == v[scorer_repeat == 2]),
+    disagreed = sum(v[scorer_repeat == 1] != v[scorer_repeat == 2]),
+    proportion = round(agreed / (agreed + disagreed), 3)
+  ) %>% 
+  summarise(across(`Agreed Feature Was Present`:disagreed,~sum(.x,na.rm = T)),
+            proportion = mean(proportion)) %>%
+  rename(`Total Agreed` = agreed,`Total Disagreed`=disagreed,
+         `Proportion Agreed`=proportion) %>%
+  kable()
+```
+
+## Agreement with gold standard
+
+### By Feature
+
+```{r}
+#| message: false
+data_gs %>%
+  group_by(Feature, scan_id, scorer) %>%
+  summarise(v = max(value, na.rm = T)) %>%
+  summarise(
+    ap = v[scorer == "A"] == v[scorer == "GS"] & v[scorer == "A"]=="Present",
+    aa = v[scorer == "A"] == v[scorer == "GS"] & v[scorer == "A"]=="Absent",
+    pa = v[scorer == "A"] == v[scorer == "GS"],
+    bp = v[scorer == "B"] == v[scorer == "GS"] & v[scorer == "B"]=="Present",
+    ba = v[scorer == "B"] == v[scorer == "GS"] & v[scorer == "B"]=="Absent",
+    pb = v[scorer == "B"] == v[scorer == "GS"],
+    cp = v[scorer == "C"] == v[scorer == "GS"] & v[scorer == "C"]=="Present",
+    ca= v[scorer == "C"] == v[scorer == "GS"] & v[scorer == "C"]=="Absent",
+     pc= v[scorer == "C"] == v[scorer == "GS"]
+  ) %>% 
+  summarise(
+    `A Agreed Feature Was Present` = sum(ap,na.rm=T),
+    `A Agreed Feature Was Absent`=sum(aa,na.rm=T),
+    `Prop. A Agreed`=mean(pa,na.rm=T),
+    `B Agreed Feature Was Present`=sum(bp,na.rm=T),
+    `B Agreed Feature Was Absent`=sum(ba,na.rm=T),
+    `Prop. B Agreed`=mean(pb,na.rm=T),
+    `C Agreed Feature Was Present`=sum(cp,na.rm=T),
+    `C Agreed Feature Was Absent`=sum(ca,na.rm=T),
+    `Prop. C Agreed`=mean(pc,na.rm=T)
+  ) %>%
+  kable()
+```
+
+### Feature by lobe
+
+Missing values were ignored in creating the table below.
+So if a scorer marked a feature and location blank on one scan and agreed with the gold standard on 3, their score will be 0.75 rather than 0.6.
+
+```{r}
+#| message: false
+data_gs %>%
+  group_by(Feature, Location, scan_id) %>%
+  rename(v=value) %>%
+  summarise(
+    ap = v[scorer == "A"] == v[scorer == "GS"] & v[scorer == "A"]=="Present",
+    aa = v[scorer == "A"] == v[scorer == "GS"] & v[scorer == "A"]=="Absent",
+    pa = v[scorer == "A"] == v[scorer == "GS"],
+    bp = v[scorer == "B"] == v[scorer == "GS"] & v[scorer == "B"]=="Present",
+    ba = v[scorer == "B"] == v[scorer == "GS"] & v[scorer == "B"]=="Absent",
+    pb = v[scorer == "B"] == v[scorer == "GS"],
+    cp = v[scorer == "C"] == v[scorer == "GS"] & v[scorer == "C"]=="Present",
+    ca= v[scorer == "C"] == v[scorer == "GS"] & v[scorer == "C"]=="Absent",
+     pc= v[scorer == "C"] == v[scorer == "GS"]
+  ) %>% 
+  summarise(
+    `A Agreed Feature Was Present` = sum(ap,na.rm = T),
+    `A Agreed Feature Was Absent`=sum(aa,na.rm = T),
+    `Prop. A Agreed`=mean(pa,na.rm = T),
+    `B Agreed Feature Was Present`=sum(bp,na.rm = T),
+    `B Agreed Feature Was Absent`=sum(ba,na.rm = T),
+    `Prop. B Agreed`=mean(pb,na.rm = T),
+    `C Agreed Feature Was Present`=sum(cp,na.rm = T),
+    `C Agreed Feature Was Absent`=sum(ca,na.rm = T),
+    `Prop. C Agreed`=mean(pc,na.rm=T)
+  ) %>%
+  kable()
+```
+
+# Mixed Model Approach
+
+```{r data cleaning 2}
+# Import data
+api <- read.table("/Users/timvigers/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt", header = F)
+rcon <- redcapConnection(
+  url = "https://redcap.ucdenver.edu/api/",
+  token = api[1, 1]
+)
+data <- exportRecordsTyped(rcon, labels = F)
+data <- data.frame(lapply(data, as.character))
+data$scan_id <- as.numeric(data$scan_id)
+# Make scorer IDs by initial
+data$scorer <- paste0(
+  toupper(substr(data$reader_first_name, 1, 1)),
+  toupper(substr(data$reader_last_name, 1, 1))
+)
+data$scorer[data$scorer == "GG"] <- "GS"
+# Deidentify readers
+old <- c("AF", "DL", "JW")
+new <- c("A", "B", "C")
+data$scorer[data$scorer %in% old] <- new[match(data$scorer, old, nomatch = 0)]
+# Revert duplicate IDs back to original to match with gold standard
+training <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/training.key.csv")
+dups <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/duplicated.key.csv")
+training$Training_id <- as.numeric(sub("T00", "", training$Training_id))
+# Change duplicates to OG IDs
+old <- dups$duplicated_ssid..dates.shifted..17.days.
+new <- dups$original_ssid
+data$scan_id[data$scan_id %in% old] <- new[na.omit(match(data$scan_id, old))]
+# Change training IDs to OG IDs
+old <- training$Training_id
+new <- training$Cffid.Scoring.Set
+data$scan_id[data$scan_id %in% old] <- new[na.omit(match(data$scan_id, old))]
+# Convert to long
+data <- data %>%
+  select(
+    scan_id, scorer, scoresheet_timestamp,
+    bronchiectasis_rul:atelectasis___6
+  ) %>%
+  arrange(scan_id, scorer, scoresheet_timestamp) %>%
+  group_by(scan_id, scorer) %>%
   mutate(scorer_repeat = row_number()) %>%
@@
 )
-# Maybe not quite so long (one column per rater)
-data = data %>% 
-  pivot_wider(names_from = c(scorer,scorer_repeat), values_from = value,
-              names_sep = "_")
+# Format values
+data$value <- factor(data$value,
+  levels = c(
+    "Absent", "Checked", "Cylindric", "Cystic", "Unchecked", "Varicose"
+  ),
+  labels = c("Absent", "Present", "Present", "Present", "Absent", "Present")
+)
+data$value <- as.numeric(data$value) - 1
+# Separate out gold standards and duplicates
+data_dups <- data %>% filter(scan_id %in% dups$original_ssid)
+data_gs <- data %>% filter(scan_id %in% training$Cffid.Scoring.Set)
+data <- data %>%
+  filter(!scan_id %in% c(dups$original_ssid, training$Cffid.Scoring.Set))
+# Second readings only for comparison to GS
+data_gs <- data_gs %>%
+  group_by(scan_id, scorer, feature, location) %>%
+  slice_max(scorer_repeat)
+# Dummy variables for scorer, organize columns
+data <- data %>%
+  dummy_cols(select_columns = "scorer") %>%
+  select(scan_id, feature, location,scorer, scorer_A, scorer_B, scorer_C, value)
 ```
 
-Kappa interpretation table @mchughInterraterReliabilityKappa2012:
+## Data setup
 
-| Value of Kappa | Level of Agreement | % of Data that are Reliable |
-| -------------- | ------------------ | --------------------------- |
-| 0–.20          | None               | 0–4%                        |
-| .21–.39        | Minimal            | 4–15%                       |
-| .40–.59        | Weak               | 15–35%                      |
-| .60–.79        | Moderate           | 35–63%                      |
-| .80–.90        | Strong             | 64–81%                      |
-| Above.90       | Almost Perfect     | 82–100%                     |
+```{r}
+DT::datatable(data)
+```
+
+## Model
+
+- Readers $r=1,2,3$
+- Scans $s=1,2,...,S$
+- Features $f=1,2,...,F$
+- Lobes $l=1,2,...,6$
+- Outcome $y_{rsfl}$ is binary
+- Model notation: $$y_{rsf=1}=\alpha+\beta X_l + b_1+b_2+b_3+\epsilon_{rs}$$
+
+### Bronchiectasis
+
+```{r}
+#| code-fold: show
+df <- data %>% filter(feature == "Bronchiectasis")
+model <- glmer(value ~ location + (1 | scorer),
+  family = "binomial", data = df
+)
+summary(model)
+icc(model)
+```
+
+### Mucoid Impaction
+
+```{r}
+#| code-fold: show
+df <- data %>% filter(feature == "Mucoid Impaction")
+model <- glmer(value ~ location + (1 | scorer),
+  family = "binomial", data = df
+)
+summary(model)
+icc(model)
+```
+
+### Cyst
+
+```{r}
+#| code-fold: show
+df <- data %>% filter(feature == "Cyst")
+model <- glmer(value ~ location + (1 | scorer),
+  family = "binomial", data = df
+)
+summary(model)
+icc(model)
+```
+
+### Cavity
+
+```{r}
+#| code-fold: show
+df <- data %>% filter(feature == "Cavity")
+model <- glmer(value ~ location + (1 | scorer),
+  family = "binomial", data = df
+)
+summary(model)
+icc(model)
+```
+
+### Consolidation
+
+```{r}
+#| code-fold: show
+df <- data %>% filter(feature == "Consolidation")
+model <- glmer(value ~ location + (1 | scorer),
+  family = "binomial", data = df
+)
+summary(model)
+icc(model)
+```
+
+### Ground Glass
+
+```{r}
+#| code-fold: show
+df <- data %>% filter(feature == "Ground Glass")
+model <- glmer(value ~ location + (1 | scorer),
+  family = "binomial", data = df
+)
+summary(model)
+icc(model)
+```
+
+### Nodule
+
+```{r}
+#| code-fold: show
+df <- data %>% filter(feature == "Nodule")
+model <- glmer(value ~ location + (1 | scorer),
+  family = "binomial", data = df
+)
+summary(model)
+icc(model)
+```
+
+### Tree In Bud
+
+```{r}
+#| code-fold: show
+df <- data %>% filter(feature == "Tree In Bud")
+model <- glmer(value ~ location + (1 | scorer),
+  family = "binomial", data = df
+)
+summary(model)
+icc(model)
+```
+
+### Atelectasis
+
+```{r}
+#| code-fold: show
+df <- data %>% filter(feature == "Atelectasis")
+model <- glmer(value ~ location + (1 | scorer),
+  family = "binomial", data = df
+)
+summary(model)
+icc(model)
+```

### 2023-08-29 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
cahceing figured out race and eti

--- 2023-08-04 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
@@
 
-```{r data cleaning}
-#| include: false
-source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R")
-load("/Volumes/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
-```
-
 ```{r libraries}
@@
 library(tidyverse)
+library(multidplyr)
 library(gtsummary)
+library(lmerTest)
+library(broom.mixed)
+library(emmeans)
+library(knitr)
+```
+
+```{r data cleaning}
+#| include: false
+# source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R")
+load("~/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
+# Make modulators ordered factors
+encounter <- encounter %>%
+  mutate(across(contains("vx"), ~ factor(.x, ordered = T)))
 ```
@@
 
-```{r table 1}
-#| label: tbl-1
-#| tbl-cap: Race by year
+In the tables below, race and ethnicity are considered separately, so some participants may be double counted.
+For example, a participant identifying as Black and Hispanic would contribute to both categories.
+
+## Race and ethnicity overall
+
+The table below shows the race and ethnicity proportions for the entire cohort included in the CFFPR dataset.
+
+```{r}
+demo %>%
+  select(race, Hispanicrace) %>%
+  tbl_summary(
+    by = Hispanicrace,
+    statistic = list(all_continuous() ~ "{mean} ({sd})")
+  ) %>%
+  add_overall() %>%
+  bold_labels() %>%
+  as_kable_extra()
+```
+
+## Race by year
+
+The table below shows the proportions of race and ethnicities included in the dataset in a given year.
+Most people are seen at least once a year, but this table serves as a sanity check to make sure that groups were generally seen equal amounts each year.
+
+```{r}
 annual %>%
   select(ReviewYear, race, Hispanicrace) %>%
-  tbl_summary(by = ReviewYear, missing_text = "Missing")
+  tbl_summary(
+    by = ReviewYear, missing_text = "Missing",
+    statistic = list(all_continuous() ~ "{mean} ({sd})")
+  ) %>%
+  bold_labels() %>%
+  as_kable_extra()
 ```
 
-```{r table 2}
-#| label: tbl-2
-#| tbl-cap: Race and ethnicity
-demo %>%
-  select(race, Hispanicrace) %>%
-  tbl_summary(by = Hispanicrace) %>%
-  modify_header(label = paste0("**", label(annual$Hispanicrace), "**")) %>%
-  add_overall()
-```
+## Race percentages overall
 
-Based on the above table, I decided to combine race and ethnicity with Hispanic "overriding" other racial identities.
+Based on the above tables, I decided to combine race and ethnicity with Hispanic "overriding" other racial identities.
 So if a participant was reported as both Black and Hispanic, they were consider Hispanic for the following analyses.
-Those with missing data for Hispanic ethnicity were moved to the "Mixed/Other/Unknown Race" category.
+Those with missing data for Hispanic ethnicity were moved to the "Indigenous/Mixed/Other/Unknown Race" category.
 
-```{r combine race and eth}
+```{r}
 demo$race <- as.character(demo$race)
@@
   levels = c(
-    "White", "Black or African American", "Hispanic or Latino",
-    "American Indian or Alaska Native", "Asian",
+    "White", "Black or African American", "Hispanic or Latino", "Asian",
+    "American Indian or Alaska Native",
     "Native Hawaiian or Other Pacific Islander", "Mixed/Other/Unknown Race"
+  ),
+  labels = c(
+    "White", "Black or African American", "Hispanic or Latino", "Asian",
+    "Indigenous/Mixed/Other/Unknown Race",
+    "Indigenous/Mixed/Other/Unknown Race", "Indigenous/Mixed/Other/Unknown Race"
   )
 )
+encounter$race <- NULL
+encounter <- left_join(encounter, demo[, c("eDWID", "race")], by = join_by(eDWID))
+label(encounter$race) <- "Race/Ethnicity"
 label(demo$race) <- "Race/Ethnicity"
+```
+
+```{r}
+demo %>%
+  select(race) %>%
+  tbl_summary(statistic = list(all_continuous() ~ "{mean} ({sd})")) %>%
+  bold_labels() %>%
+  as_kable_extra()
 ```
@@
 
+## Characteristics by race
+
 ```{r}
-#| label: tbl-3
-#| tbl-cap: Characteristics by race
 demo %>%
-  select(race) %>%
-  tbl_summary()
+  select(race, Gender, WasGenotyped, F508_num) %>%
+  tbl_summary(
+    by = race, missing_text = "Missing",
+    statistic = list(all_continuous() ~ "{mean} ({sd})")
+  ) %>%
+  add_overall() %>%
+  bold_labels() %>%
+  as_kable_extra()
 ```
 
-# Figure
+## Characteristics by year
 
 ```{r}
-#| label: fig-planets
-#| fig-cap: Planets
+#| cache: true
+encounter %>%
+  group_by(eDWID, reviewyear) %>%
+  summarise(across(c(race, Gender, WasGenotyped, F508_num, contains("insurance")), first),
+    `Age at Last Visit` = last(na.omit(encounterage)),
+    `Mean ppFEV` = mean(GLI_FEV1_pct_predicted, na.rm = T),
+    `Weight at Last Visit` = last(na.omit(weight)),
+    `Weight %ile at Last Visit` = last(na.omit(weightpercentile)),
+    `BMI at Last Visit` = last(na.omit(bmivalue)),
+    `BMI %ile at Last Visit` = last(na.omit(bmipercentile)),
+    across(contains("Vx"), ~ max(.x)),
+    .groups = "drop"
+  ) %>%
+  select(
+    -eDWID, reviewyear, race, Gender, WasGenotyped, F508_num, contains("insurance"),
+    `Age at Last Visit`, `Mean ppFEV`, `Weight at Last Visit`,
+    `Weight %ile at Last Visit`, `BMI at Last Visit`,
+    `BMI %ile at Last Visit`, contains("Vx")
+  ) %>%
+  rename(any_of(tidy_labels)) %>%
+  tbl_summary(
+    by = Year, missing_text = "Missing",
+    statistic = list(all_continuous() ~ "{mean} ({sd})")
+  ) %>%
+  bold_labels() %>%
+  as_kable_extra()
 ```
+
+# Potential COVID effects
+
+## Spaghetti plot
+
+```{r}
+#| warning: false
+ggplot(encounter, aes(
+  y = GLI_FEV1_pct_predicted, x = encounterdate,
+  group = eDWID, color = factor(eDWID)
+)) +
+  geom_line(alpha = 0.02) +
+  theme_bw() +
+  theme(legend.position = "none")
+```
+
+## Visits prior to and after March 11, 2020
+
+The table below is limited to visits 2019 - 2021.
+
+```{r}
+encounter %>%
+  filter(encounterdate >= "2019-01-01") %>%
+  mutate(post_covid = encounterdate > "2020-03-11") %>%
+  group_by(eDWID) %>%
+  summarise(
+    `Number of Visits Pre-COVID` = sum(!post_covid),
+    `Number of Visits Post-COVID` = sum(post_covid),
+    race = first(race),
+    .groups = "drop"
+  ) %>%
+  rename(any_of(tidy_labels)) %>%
+  select(-eDWID) %>%
+  tbl_summary(by = `Race/Ethnicity`, missing_text = "Missing") %>%
+  bold_labels() %>%
+  as_kable_extra()
+```
+
+# Lung function over time
+
+## Annualized averages
+
+```{r}
+mod_df <- encounter %>%
+  group_by(eDWID, reviewyear) %>%
+  summarise(
+    across(
+      c(GLI_FEV1_pct_predicted, GLI_FVC_pct_predicted),
+      ~ mean(.x, na.rm = T)
+    ),
+    race = first(race),
+    .groups = "drop"
+  )
+```
+
+FEV1 and FVC percent predicted (based on GLI equations) were averaged by year for each person.
+
+```{r}
+#| warning: false
+encounter %>%
+  group_by(eDWID, reviewyear) %>%
+  summarise(across(c(GLI_FEV1_pct_predicted, GLI_FVC_pct_predicted), ~ mean(.x, na.rm = T)),
+    .groups = "drop"
+  ) %>%
+  pivot_longer(c("GLI_FEV1_pct_predicted", "GLI_FVC_pct_predicted")) %>%
+  ggplot(aes(x = reviewyear, y = value, color = factor(eDWID), group = eDWID)) +
+  geom_line(alpha = 0.01) +
+  theme_bw() +
+  theme(legend.position = "none") +
+  facet_wrap(~name)
+```
+
+```{r}
+#| cache: true
+# Most basic model of lung function over time by race
+mod <- lmer(
+  GLI_FEV1_pct_predicted ~ factor(reviewyear) * race + (1 | eDWID),
+  data = encounter %>% group_by(eDWID, reviewyear) %>%
+    summarise(across(c(GLI_FEV1_pct_predicted, GLI_FVC_pct_predicted), ~ mean(.x, na.rm = T)),
+      race = first(race),
+      .groups = "drop"
+    )
+)
+res <- tidy(mod, effects = "fixed")
+means <- data.frame(emmeans(mod, ~ race + reviewyear,
+  lmerTest.limit = 1e6,
+  pbkrtest.limit = 1e6
+))
+kable(means)
+```
+
+```{r}
+#| fig-width: 10
+#| fig-height: 10
+ggplot(means, aes(x = reviewyear, y = emmean, color = race)) +
+  geom_line() +
+  geom_point() +
+  xlab("Year") +
+  ylab("LS Mean FEV1") +
+  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE), width = .2) +
+  theme_bw()
+```

### 2023-08-30 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
started using table1

@@
 #| include: false
-library(Hmisc)
 library(tidyverse)
-library(multidplyr)
-library(gtsummary)
+library(table1)
 library(lmerTest)
@@
 load("~/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
-# Make modulators ordered factors
-encounter <- encounter %>%
-  mutate(across(contains("vx"), ~ factor(.x, ordered = T)))
 ```
@@
 ```{r}
-demo %>%
-  select(race, Hispanicrace) %>%
-  tbl_summary(
-    by = Hispanicrace,
-    statistic = list(all_continuous() ~ "{mean} ({sd})")
-  ) %>%
-  add_overall() %>%
-  bold_labels() %>%
-  as_kable_extra()
+table1(~ race | Hispanicrace, data=demo)
 ```
@@
 ```{r}
-annual %>%
-  select(ReviewYear, race, Hispanicrace) %>%
-  tbl_summary(
-    by = ReviewYear, missing_text = "Missing",
-    statistic = list(all_continuous() ~ "{mean} ({sd})")
-  ) %>%
-  bold_labels() %>%
-  as_kable_extra()
+table1(~race+Hispanicrace|factor(ReviewYear),data = annual)
 ```
@@
 ```{r}
-demo %>%
-  select(race) %>%
-  tbl_summary(statistic = list(all_continuous() ~ "{mean} ({sd})")) %>%
-  bold_labels() %>%
-  as_kable_extra()
+table1(~race,data = demo)
 ```
@@
 ```{r}
-demo %>%
-  select(race, Gender, WasGenotyped, F508_num) %>%
-  tbl_summary(
-    by = race, missing_text = "Missing",
-    statistic = list(all_continuous() ~ "{mean} ({sd})")
-  ) %>%
-  add_overall() %>%
-  bold_labels() %>%
-  as_kable_extra()
+table1(~Gender+WasGenotyped+F508_num|race,data=demo)
 ```
@@
 #| cache: true
-encounter %>%
+by_year = encounter %>%
   group_by(eDWID, reviewyear) %>%
@@
   as_kable_extra()
+```
+
+# ETI eligibility
+
+Trikafta originally approved October 21, 2019 for patients 12+ years of age. Approved for ages 6 - 11 on June 9, 2021.
+
+```{r}
+encounter %>%
+  mutate(eti_elig = factor(eti_elig, ordered = T)) %>%
+  group_by(eDWID, reviewyear) %>%
+  summarise(
+    eti_elig = max(eti_elig), race = first(na.omit(race)),
+    .groups = "drop"
+  ) %>%
+  select(-eDWID) %>%
+  rename(any_of(tidy_labels)) %>%
+  tbl_summary(by = `Race/Ethnicity`, missing = "Missing") %>%
+        bold_labels() %>%
+        as_kable_extra()
+  tbl_strata(
+    strata = Year,
+    .tbl_fun =
+      ~ .x %>%
+
+  )
 ```
@@
 #| warning: false
-encounter %>%
-  group_by(eDWID, reviewyear) %>%
-  summarise(across(c(GLI_FEV1_pct_predicted, GLI_FVC_pct_predicted), ~ mean(.x, na.rm = T)),
-    .groups = "drop"
-  ) %>%
+mod_df
 pivot_longer(c("GLI_FEV1_pct_predicted", "GLI_FVC_pct_predicted")) %>%
@@
   GLI_FEV1_pct_predicted ~ factor(reviewyear) * race + (1 | eDWID),
-  data = encounter %>% group_by(eDWID, reviewyear) %>%
-    summarise(across(c(GLI_FEV1_pct_predicted, GLI_FVC_pct_predicted), ~ mean(.x, na.rm = T)),
-      race = first(race),
-      .groups = "drop"
-    )
+  data = mod_df
 )

### 2023-08-30 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R

@@ -6,6 +6,17 @@ library(Hmisc)
 annual <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_Annualized_Del1.csv", na.strings = "")
 encounter <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_encountersMerged_Del1.csv", na.strings = "")
 demo <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_DemogCFDiag_Del1.csv", na.strings = "")
+# Dates
+demo$First_LungTransplantDate <- mdy(demo$First_LungTransplantDate)
+demo$Modulator_trikafta_first_date <- mdy(demo$Modulator_trikafta_first_date)
+encounter$encounterdate <- mdy(encounter$encounterdate)
+# ETI eligibility
+encounter$eti_elig <- "No"
+encounter$eti_elig[encounter$encounterdate >= ("2019-10-21") &
+  encounter$encounterage >= 12] <- "Yes"
+encounter$eti_elig[encounter$encounterdate >= ("2021-06-09") &
+  encounter$encounterage >= 6] <- "Yes"
+encounter$eti_elig <- factor(encounter$eti_elig, levels = c("No", "Yes"))
 # Fix missing values for specific bacteria
 bugs <- c(
   "staphylococcus_aureus", "haemophilus_influenzae",
@@ -257,10 +268,6 @@ demo$race <- factor(demo$race,
     "Native Hawaiian or Other Pacific Islander", "Mixed/Other/Unknown Race"
   )
 )
-# Dates
-demo$First_LungTransplantDate <- mdy(demo$First_LungTransplantDate)
-demo$Modulator_trikafta_first_date <- mdy(demo$Modulator_trikafta_first_date)
-encounter$encounterdate <- mdy(encounter$encounterdate)
 # Manage labels (tried to do this programmatically but the data dictionary has
 # some quirks and this ended up being faster)
 encounter_labels <-
@@ -303,7 +310,8 @@ encounter_labels <-
     "pulmonarycomplications1" = "ABPA",
[…]

### 2023-09-12 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
table1 instead of gtsummary

@@
   ) %>%
-  rename(any_of(tidy_labels)) %>%
-  tbl_summary(
-    by = Year, missing_text = "Missing",
-    statistic = list(all_continuous() ~ "{mean} ({sd})")
-  ) %>%
-  bold_labels() %>%
-  as_kable_extra()
+  rename(any_of(tidy_labels))
+
+```
+
+```{r}
+table1(~Gender+WasGenotyped+F508_num|race,data=by_year)
 ```

### 2023-09-15 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
race ETI all data manipulation in separate file

@@
 # source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R")
-load("~/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
+load("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
 ```
@@
 
-In the tables below, race and ethnicity are considered separately, so some participants may be double counted.
-For example, a participant identifying as Black and Hispanic would contribute to both categories.
-
-## Race and ethnicity overall
-
 The table below shows the race and ethnicity proportions for the entire cohort included in the CFFPR dataset.
+I decided to combine race and ethnicity with Hispanic "overriding" other racial identities.
+So if a participant was reported as both Black and Hispanic, they were consider Hispanic for the following analyses.
+Those with missing data for Hispanic ethnicity were moved to the "Indigenous/Mixed/Other/Unknown Race" category.
 
 ```{r}
-table1(~ race | Hispanicrace, data=demo)
+table1(~race, data = demo)
 ```
@@
 ```{r}
-table1(~race+Hispanicrace|factor(ReviewYear),data = annual)
-```
-
-## Race percentages overall
-
-Based on the above tables, I decided to combine race and ethnicity with Hispanic "overriding" other racial identities.
-So if a participant was reported as both Black and Hispanic, they were consider Hispanic for the following analyses.
-Those with missing data for Hispanic ethnicity were moved to the "Indigenous/Mixed/Other/Unknown Race" category.
-
-```{r}
-demo$race <- as.character(demo$race)
-demo$race[demo$Hispanicrace == "Yes"] <- "Hispanic or Latino"
-demo$race[demo$Hispanicrace == "Unknown"] <- "Hispanic or Latino"
-demo$race <- factor(demo$race,
-  levels = c(
-    "White", "Black or African American", "Hispanic or Latino", "Asian",
-    "American Indian or Alaska Native",
-    "Native Hawaiian or Other Pacific Islander", "Mixed/Other/Unknown Race"
-  ),
-  labels = c(
-    "White", "Black or African American", "Hispanic or Latino", "Asian",
-    "Indigenous/Mixed/Other/Unknown Race",
-    "Indigenous/Mixed/Other/Unknown Race", "Indigenous/Mixed/Other/Unknown Race"
-  )
-)
-encounter$race <- NULL
-encounter <- left_join(encounter, demo[, c("eDWID", "race")], by = join_by(eDWID))
-label(encounter$race) <- "Race/Ethnicity"
-label(demo$race) <- "Race/Ethnicity"
-```
-
-```{r}
-table1(~race,data = demo)
+table1(~ race | factor(ReviewYear), data = annual)
 ```
@@
 
+```{r}
+t1_vars <- c(
+  "Gender", "WasGenotyped", "F508_num", "mean_ppFEV", "Age_YrEnd",
+  "mean_ppFEV", "weight_last", "weight_perc_last", "bmi_last",
+  "bmi_perc_last", "Vx770", "Vx809comb", "Vx661comb", "Vx445comb"
+)
+t1_form <- as.formula(paste("~", paste(t1_vars, collapse = "+"), "|factor(ReviewYear)"))
+```
+
 ## Characteristics by race
@@
 ```{r}
-table1(~Gender+WasGenotyped+F508_num|race,data=demo)
+table1(~ Gender + WasGenotyped + F508_num | race, data = demo)
 ```
@@
 ```{r}
-#| cache: true
-by_year = encounter %>%
-  group_by(eDWID, reviewyear) %>%
-  summarise(across(c(race, Gender, WasGenotyped, F508_num, contains("insurance")), first),
-    `Age at Last Visit` = last(na.omit(encounterage)),
-    `Mean ppFEV` = mean(GLI_FEV1_pct_predicted, na.rm = T),
-    `Weight at Last Visit` = last(na.omit(weight)),
-    `Weight %ile at Last Visit` = last(na.omit(weightpercentile)),
-    `BMI at Last Visit` = last(na.omit(bmivalue)),
-    `BMI %ile at Last Visit` = last(na.omit(bmipercentile)),
-    across(contains("Vx"), ~ max(.x)),
-    .groups = "drop"
-  ) %>%
-  select(
-    -eDWID, reviewyear, race, Gender, WasGenotyped, F508_num, contains("insurance"),
-    `Age at Last Visit`, `Mean ppFEV`, `Weight at Last Visit`,
-    `Weight %ile at Last Visit`, `BMI at Last Visit`,
-    `BMI %ile at Last Visit`, contains("Vx")
-  ) %>%
-  rename(any_of(tidy_labels))
-
-```
-
-```{r}
-table1(~Gender+WasGenotyped+F508_num|race,data=by_year)
+table1(t1_form, data = annual)
 ```
@@
 
+## Overall by year
+
 ```{r}
-encounter %>%
-  mutate(eti_elig = factor(eti_elig, ordered = T)) %>%
-  group_by(eDWID, reviewyear) %>%
-  summarise(
-    eti_elig = max(eti_elig), race = first(na.omit(race)),
-    .groups = "drop"
-  ) %>%
-  select(-eDWID) %>%
-  rename(any_of(tidy_labels)) %>%
-  tbl_summary(by = `Race/Ethnicity`, missing = "Missing") %>%
-        bold_labels() %>%
-        as_kable_extra()
-  tbl_strata(
-    strata = Year,
-    .tbl_fun =
-      ~ .x %>%
+table1(~ eti_elig | factor(ReviewYear), data = annual, overall = F)
+```
 
-  )
+## By race and year
+
+```{r}
+
 ```

### 2023-09-15 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R

@@ -3,9 +3,9 @@ library(lubridate)
 library(readxl)
 library(Hmisc)
 # Import
-annual <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_Annualized_Del1.csv", na.strings = "")
-encounter <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_encountersMerged_Del1.csv", na.strings = "")
-demo <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_DemogCFDiag_Del1.csv", na.strings = "")
+annual <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_Annualized_Del1.csv", na.strings = "")
+encounter <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_encountersMerged_Del1.csv", na.strings = "")
+demo <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_DemogCFDiag_Del1.csv", na.strings = "")
 # Dates
 demo$First_LungTransplantDate <- mdy(demo$First_LungTransplantDate)
 demo$Modulator_trikafta_first_date <- mdy(demo$Modulator_trikafta_first_date)
@@ -228,7 +228,7 @@ demo$Hispanicrace <- factor(demo$Hispanicrace,
 )
 demo$MutClass <- factor(demo$MutClass)
 # Get long mutation lists - demographics
-mutations <- read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/Copy of Codes for CFFPR_2023.xlsx")
[…]

### 2023-09-15 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
back to gtsummary for race and ETI

@@
       dark: darkly
-bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
+bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
+csl: /Users/timvigers/Dropbox/Miscellaneous/american-medical-association.csl
 editor: source
@@
 
-```{r libraries}
+```{r setup}
 #| include: false
+# Libraries
+library(Hmisc)
 library(tidyverse)
-library(table1)
+library(gtsummary)
 library(lmerTest)
@@
 ```{r}
-table1(~race, data = demo)
+demo %>%
+  select(race) %>%
+  tbl_summary(missing_text = "Missing") %>%
+  bold_labels()
 ```
@@
 ```{r}
-table1(~ race | factor(ReviewYear), data = annual)
+annual %>%
+  select(race, ReviewYear) %>%
+  tbl_summary(by = ReviewYear, missing_text = "Missing") %>%
+  bold_labels()
 ```
@@
 )
-t1_form <- as.formula(paste("~", paste(t1_vars, collapse = "+"), "|factor(ReviewYear)"))
 ```
@@
 ```{r}
-table1(~ Gender + WasGenotyped + F508_num | race, data = demo)
+demo %>%
+  select(Gender, WasGenotyped, F508_num, race) %>%
+  tbl_summary(by = race, missing_text = "Missing") %>%
+  bold_labels()
 ```
@@
 
-```{r}
-table1(t1_form, data = annual)
+```{r results='asis'}
+set_gtsummary_theme(theme_gtsummary_compact(), quiet = T)
+years <- split.data.frame(annual, annual$ReviewYear)
+invisible(lapply(names(years), function(n) {
+  cat("\n")
+  cat("###", n)
+  cat("\n")
+  d <- years[[n]]
+  t <- d %>%
+    select(all_of(t1_vars), race) %>%
+    tbl_summary(by = race, missing_text = "Missing") %>%
+    bold_labels()
+  print(t)
+  cat("\n")
+}))
+reset_gtsummary_theme()
 ```
@@
 ```{r}
-table1(~ eti_elig | factor(ReviewYear), data = annual, overall = F)
+annual %>%
+  select(eti_elig, ReviewYear) %>%
+  tbl_summary(
+    by = ReviewYear, missing_text = "Missing",
+    type = all_dichotomous() ~ "categorical"
+  ) %>%
+  bold_labels()
 ```
@@
 
+```{r results='asis'}
+invisible(lapply(names(years), function(n) {
+  cat("\n")
+  cat("###", n)
+  cat("\n")
+  d <- years[[n]]
+  t <- d %>%
+    select(eti_elig, race) %>%
+    tbl_summary(
+      by = race, missing_text = "Missing",
+      type = all_dichotomous() ~ "categorical"
+    ) %>%
+    bold_labels()
+  print(t)
+  cat("\n")
+}))
+```
+
+# People without visits in a given year
+
 ```{r}
+# Which are missing from encounters data?
+a <- as.data.frame.matrix(table(annual$eDWID, annual$ReviewYear))
+a <- a %>% rownames_to_column("eDWID")
+e <- as.data.frame.matrix(table(by_year$eDWID, by_year$reviewyear))
+e <- e %>% rownames_to_column("eDWID")
+rowdiff_a <- setdiff(a, e)
+rowdiff_a$data <- "annual"
+rowdiff_e <- setdiff(e, a)
+rowdiff_e$data <- "encounter"
+diffs <- rbind(rowdiff_a, rowdiff_e)
+diffs <- diffs %>%
+  arrange(eDWID, desc(data)) %>%
+  select(-data)
+diffs <- diffs %>%
+  group_by(eDWID) %>%
+  reframe(across(`2015`:`2021`, ~ diff(as.numeric(.x))))
+diffs %>%
+  reframe(across(`2015`:`2021`, ~ sum(.x == 1))) %>%
+  kable()
+```
 
-```
+There were a total of `r length(setdiff(a$eDWID,e$eDWID))` included in the annual dataset that did not appear in the encounter-level dataset at all.
+There were an additional `r nrow(diffs)` patients who were included in both datasets, but had at least one year with no encounters.
+The table above shows the number of people missing an encounter in each year.
+
+If someone was not seen a given year, do we count them as eligible for that year? I am assuming they will be eligible but obviously will not have started any modulators.
 
@@
   ) %>%
-  rename(any_of(tidy_labels)) %>%
   select(-eDWID) %>%
-  tbl_summary(by = `Race/Ethnicity`, missing_text = "Missing") %>%
-  bold_labels() %>%
-  as_kable_extra()
+  tbl_summary(by = race, missing_text = "Missing") %>%
+  bold_labels()
 ```
@@
 
-```{r}
-mod_df <- encounter %>%
-  group_by(eDWID, reviewyear) %>%
-  summarise(
-    across(
-      c(GLI_FEV1_pct_predicted, GLI_FVC_pct_predicted),
-      ~ mean(.x, na.rm = T)
-    ),
-    race = first(race),
-    .groups = "drop"
-  )
-```
-
 FEV1 and FVC percent predicted (based on GLI equations) were averaged by year for each person.
@@
 #| warning: false
-mod_df
-pivot_longer(c("GLI_FEV1_pct_predicted", "GLI_FVC_pct_predicted")) %>%
-  ggplot(aes(x = reviewyear, y = value, color = factor(eDWID), group = eDWID)) +
+annual %>%
+  mutate(mean_ppFEV = as.numeric(mean_ppFEV), mean_ppFVC = as.numeric(mean_ppFVC)) %>%
+  pivot_longer(c("mean_ppFEV", "mean_ppFVC")) %>%
+  ggplot(aes(x = ReviewYear, y = value, color = factor(eDWID), group = eDWID)) +
   geom_line(alpha = 0.01) +
@@
 mod <- lmer(
-  GLI_FEV1_pct_predicted ~ factor(reviewyear) * race + (1 | eDWID),
-  data = mod_df
+  mean_ppFEV ~ factor(ReviewYear) * race + (1 | eDWID),
+  data = annual
 )
 res <- tidy(mod, effects = "fixed")
-means <- data.frame(emmeans(mod, ~ race + reviewyear,
+means <- data.frame(emmeans(mod, ~ race + ReviewYear,
   lmerTest.limit = 1e6,
@@
 ```{r}
-#| fig-width: 10
-#| fig-height: 10
-ggplot(means, aes(x = reviewyear, y = emmean, color = race)) +
+ggplot(means, aes(x = ReviewYear, y = emmean, color = race)) +
   geom_line() +

### 2023-09-15 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R

@@ -17,6 +17,10 @@ encounter$eti_elig[encounter$encounterdate >= ("2019-10-21") &
 encounter$eti_elig[encounter$encounterdate >= ("2021-06-09") &
   encounter$encounterage >= 6] <- "Yes"
 encounter$eti_elig <- factor(encounter$eti_elig, levels = c("No", "Yes"))
+annual$eti_elig <- "No"
+annual$eti_elig[annual$ReviewYear >= 2019 & annual$Age_YrEnd >= 12] <- "Yes"
+annual$eti_elig[annual$ReviewYear >= 2021 & annual$Age_YrEnd >= 6] <- "Yes"
+annual$eti_elig <- factor(annual$eti_elig, levels = c("No", "Yes"))
 # Fix missing values for specific bacteria
 bugs <- c(
   "staphylococcus_aureus", "haemophilus_influenzae",
@@ -277,23 +281,19 @@ demo$race <- factor(demo$race,
 )
 # Create some of our own annualized variables
 by_year <- encounter %>%
-  mutate(
-    eti_elig = factor(eti_elig, ordered = T),
-    across(contains("Vx"), ~ factor(.x, ordered = T))
-  ) %>%
+  mutate(across(contains("Vx"), ~ factor(.x, ordered = T))) %>%
   group_by(eDWID, reviewyear) %>%
   summarise(
     mean_ppFEV = mean(GLI_FEV1_pct_predicted, na.rm = T),
+    mean_ppFVC = mean(GLI_FVC_pct_predicted, na.rm = T),
     weight_last = last(na.omit(weight)),
     weight_perc_last = last(na.omit(weightpercentile)),
     bmi_last = last(na.omit(bmivalue)),
     bmi_perc_last = last(na.omit(bmipercentile)),
-    eti_elig = max(eti_elig),
     across(contains("Vx"), ~ max(.x)),
     .groups = "drop"
   ) %>%
   mutate(
-    eti_elig = factor(eti_elig, ordered = F),
     across(contains("Vx"), ~ factor(.x, ordered = F))
   )
 annual <- full_join(annual, by_year, by = join_by(eDWID, ReviewYear == reviewyear))
@@ -380,20 +380,24 @@ labels <-
     "Mutation3" = "Name of the third mutation (if any)",
     "First_LungTransplantDate" = "Date of First Lung Transplant",
     "Modulator_trikafta_first_date" = "Date of Trikafta Start",
[…]

### 2023-09-18 Stacey Martiniano/ntm_agreement.qmd
added tree in bud or nodule variable

--- 2023-07-31 Stacey Martiniano/ntm_agreement.qmd
@@
 
-```{r}
-#| include: false
+```{r data cleaning}
+#| message: false
 library(tidyverse)
-library(irr)
 library(redcapAPI)
 library(Hmisc)
-library(DT)
-library(RColorBrewer)
 library(knitr)
-library(pheatmap)
-library(arsenal)
+library(fastDummies)
+library(lmerTest)
+library(performance)
 # Import data
-api <- read.table("/Users/timvigers/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt",
-                  header = F
-)
+api <- read.table("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt", header = F)
 rcon <- redcapConnection(
@@
 # Revert duplicate IDs back to original to match with gold standard
-training <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/training.key.csv")
-dups <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/duplicated.key.csv")
+training <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/training.key.csv")
+dups <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/duplicated.key.csv")
 training$Training_id <- as.numeric(sub("T00", "", training$Training_id))
@@
 # Convert to long
-data = data %>% select(scan_id,scorer,scoresheet_timestamp,
-                    bronchiectasis_rul:atelectasis___6) %>%
-  arrange(scan_id,scorer,scoresheet_timestamp) %>%
-  group_by(scan_id,scorer) %>%
+data <- data %>%
+  select(
+    scan_id, scorer, scoresheet_timestamp,
+    bronchiectasis_rul:atelectasis___6
+  ) %>%
+  arrange(scan_id, scorer, scoresheet_timestamp) %>%
+  group_by(scan_id, scorer) %>%
+  mutate(scorer_repeat = row_number()) %>%
+  select(-scoresheet_timestamp) %>%
+  pivot_longer(bronchiectasis_rul:atelectasis___6,
+    names_to = c("Feature", "Location"),
+    names_pattern = "(.*)_(.*)$"
+  ) %>%
+  mutate(
+    Feature = gsub("__", "", Feature),
+    Feature = str_to_title(gsub("_", " ", Feature))
+  )
+data$Location <- factor(data$Location,
+  levels = c(
+    "1", "2", "3", "4", "5", "6", "ling", "lll",
+    "lul", "rll", "rml", "rul"
+  ),
+  labels = c(
+    "RUL", "RML", "RLL", "LUL", "Ling", "LLL",
+    "Ling", "LLL", "LUL", "RLL", "RML", "RUL"
+  )
+)
+# Add tree in bud OR nodule variable
+data <- data %>%
+  pivot_wider(names_from = Feature) %>%
+  mutate(
+    `Tree In Bud OR Nodule` =
+      factor(Nodule == "Checked" | `Tree In Bud` == "Checked",
+        levels = c(F, T), labels = c("Unchecked", "Checked")
+      )
+  ) %>%
+  pivot_longer(Bronchiectasis:last_col(), names_to = "Feature")
+# Format values
+data$value <- factor(data$value,
+  levels = c(
+    "Absent", "Checked", "Cylindric", "Cystic", "Unchecked", "Varicose"
+  ),
+  labels = c("Absent", "Present", "Present", "Present", "Absent", "Present"),
+  ordered = T
+)
+# Separate out gold standards and duplicates
+data_dups <- data %>% filter(scan_id %in% dups$original_ssid)
+data_gs <- data %>% filter(scan_id %in% training$Cffid.Scoring.Set)
+data <- data %>%
+  filter(!scan_id %in% c(dups$original_ssid, training$Cffid.Scoring.Set))
+# Second readings only for comparison to GS
+data_gs <- data_gs %>%
+  group_by(scan_id, scorer, Feature, Location) %>%
+  slice_max(scorer_repeat)
+```
+
+# Proportion of agreement
+
+## Singletons
+
+```{r}
+# Find scans with disagreement on certain features, so Stacey can discuss them with radiologists
+t <- data %>%
+  group_by(scan_id, Feature, scorer) %>%
+  summarise(v = max(value, na.rm = T)) %>%
+  summarise(a = sum(v == "Present")) %>%
+  filter(
+    Feature %in% c("Atelectasis", "Cavity", "Consolidation", "Ground Glass", "Nodule"),
+    a == 2
+  )
+```
+
+### By Feature
+
+```{r}
+#| message: false
+data %>%
+  group_by(Feature, scan_id, scorer) %>%
+  summarise(v = max(value, na.rm = T)) %>%
+  summarise(a = sum(v == "Present")) %>%
+  summarise(
+    `Agreed Feature Was Present` = sum(a == 3),
+    `Agreed Feature Was Absent` = sum(a == 0),
+    agreed = sum(a == 0 | a == 3),
+    disagreed = sum(a == 1 | a == 2),
+    proportion = round(agreed / (agreed + disagreed), 3)
+  ) %>%
+  rename(
+    `Total Agreed` = agreed, `Total Disagreed` = disagreed,
+    `Proportion Agreed` = proportion
+  ) %>%
+  kable()
+```
+
+### Feature by lobe
+
+```{r}
+#| message: false
+data %>%
+  group_by(Feature, Location, scan_id) %>%
+  summarise(a = factor(sum(value == "Present", na.rm = T))) %>%
+  summarise(
+    `Agreed Feature Was Present` = sum(a == 3),
+    `Agreed Feature Was Absent` = sum(a == 0),
+    agreed = sum(a == 0 | a == 3),
+    disagreed = sum(a == 1 | a == 2),
+    proportion = round(agreed / (agreed + disagreed), 3)
+  ) %>%
+  rename(
+    `Total Agreed` = agreed, `Total Disagreed` = disagreed,
+    `Proportion Agreed` = proportion
+  ) %>%
+  kable()
+```
+
+## Duplicates
+
+### By Feature
+
+```{r}
+#| message: false
+data_dups %>%
+  group_by(Feature, scorer, scan_id, scorer_repeat) %>%
+  summarise(v = max(value, na.rm = T)) %>%
+  summarise(
+    `Agreed Feature Was Present` = sum(v[scorer_repeat == 1] == v[scorer_repeat == 2] &
+      v[scorer_repeat == 1] == "Present"),
+    `Agreed Feature Was Absent` = sum(v[scorer_repeat == 1] == v[scorer_repeat == 2] &
+      v[scorer_repeat == 1] == "Absent"),
+    agreed = sum(v[scorer_repeat == 1] == v[scorer_repeat == 2]),
+    disagreed = sum(v[scorer_repeat == 1] != v[scorer_repeat == 2]),
+    proportion = round(agreed / (agreed + disagreed), 3)
+  ) %>%
+  summarise(across(`Agreed Feature Was Present`:disagreed, ~ sum(.x, na.rm = T)),
+    proportion = mean(proportion)
+  ) %>%
+  rename(
+    `Total Agreed` = agreed, `Total Disagreed` = disagreed,
+    `Proportion Agreed` = proportion
+  ) %>%
+  kable()
+```
+
+### Feature by lobe
+
+```{r}
+#| message: false
+#| warning: false
+data_dups %>%
+  group_by(Feature, Location, scorer, scan_id, scorer_repeat) %>%
+  summarise(v = max(value, na.rm = T)) %>%
+  summarise(
+    `Agreed Feature Was Present` = sum(v[scorer_repeat == 1] == v[scorer_repeat == 2] &
+      v[scorer_repeat == 1] == "Present"),
+    `Agreed Feature Was Absent` = sum(v[scorer_repeat == 1] == v[scorer_repeat == 2] &
+      v[scorer_repeat == 1] == "Absent"),
+    agreed = sum(v[scorer_repeat == 1] == v[scorer_repeat == 2]),
+    disagreed = sum(v[scorer_repeat == 1] != v[scorer_repeat == 2]),
+    proportion = round(agreed / (agreed + disagreed), 3)
+  ) %>%
+  summarise(across(`Agreed Feature Was Present`:disagreed, ~ sum(.x, na.rm = T)),
+    proportion = mean(proportion)
+  ) %>%
+  rename(
+    `Total Agreed` = agreed, `Total Disagreed` = disagreed,
+    `Proportion Agreed` = proportion
+  ) %>%
+  kable()
+```
+
+## Agreement with gold standard
+
+### By Feature
+
+```{r}
+#| message: false
+data_gs %>%
+  group_by(Feature, scan_id, scorer) %>%
+  summarise(v = max(value, na.rm = T)) %>%
+  summarise(
+    ap = v[scorer == "A"] == v[scorer == "GS"] & v[scorer == "A"] == "Present",
+    aa = v[scorer == "A"] == v[scorer == "GS"] & v[scorer == "A"] == "Absent",
+    pa = v[scorer == "A"] == v[scorer == "GS"],
+    bp = v[scorer == "B"] == v[scorer == "GS"] & v[scorer == "B"] == "Present",
+    ba = v[scorer == "B"] == v[scorer == "GS"] & v[scorer == "B"] == "Absent",
+    pb = v[scorer == "B"] == v[scorer == "GS"],
+    cp = v[scorer == "C"] == v[scorer == "GS"] & v[scorer == "C"] == "Present",
+    ca = v[scorer == "C"] == v[scorer == "GS"] & v[scorer == "C"] == "Absent",
+    pc = v[scorer == "C"] == v[scorer == "GS"]
+  ) %>%
+  summarise(
+    `A Agreed Feature Was Present` = sum(ap, na.rm = T),
+    `A Agreed Feature Was Absent` = sum(aa, na.rm = T),
+    `Prop. A Agreed` = mean(pa, na.rm = T),
+    `B Agreed Feature Was Present` = sum(bp, na.rm = T),
+    `B Agreed Feature Was Absent` = sum(ba, na.rm = T),
+    `Prop. B Agreed` = mean(pb, na.rm = T),
+    `C Agreed Feature Was Present` = sum(cp, na.rm = T),
+    `C Agreed Feature Was Absent` = sum(ca, na.rm = T),
+    `Prop. C Agreed` = mean(pc, na.rm = T)
+  ) %>%
+  kable()
+```
+
+### Feature by lobe
+
+Missing values were ignored in creating the table below.
+So if a scorer marked a feature and location blank on one scan and agreed with the gold standard on 3, their score will be 0.75 rather than 0.6.
+
+```{r}
+#| message: false
+data_gs %>%
+  group_by(Feature, Location, scan_id) %>%
+  rename(v = value) %>%
+  summarise(
+    ap = v[scorer == "A"] == v[scorer == "GS"] & v[scorer == "A"] == "Present",
+    aa = v[scorer == "A"] == v[scorer == "GS"] & v[scorer == "A"] == "Absent",
+    pa = v[scorer == "A"] == v[scorer == "GS"],
+    bp = v[scorer == "B"] == v[scorer == "GS"] & v[scorer == "B"] == "Present",
+    ba = v[scorer == "B"] == v[scorer == "GS"] & v[scorer == "B"] == "Absent",
+    pb = v[scorer == "B"] == v[scorer == "GS"],
+    cp = v[scorer == "C"] == v[scorer == "GS"] & v[scorer == "C"] == "Present",
+    ca = v[scorer == "C"] == v[scorer == "GS"] & v[scorer == "C"] == "Absent",
+    pc = v[scorer == "C"] == v[scorer == "GS"]
+  ) %>%
+  summarise(
+    `A Agreed Feature Was Present` = sum(ap, na.rm = T),
+    `A Agreed Feature Was Absent` = sum(aa, na.rm = T),
+    `Prop. A Agreed` = mean(pa, na.rm = T),
+    `B Agreed Feature Was Present` = sum(bp, na.rm = T),
+    `B Agreed Feature Was Absent` = sum(ba, na.rm = T),
+    `Prop. B Agreed` = mean(pb, na.rm = T),
+    `C Agreed Feature Was Present` = sum(cp, na.rm = T),
+    `C Agreed Feature Was Absent` = sum(ca, na.rm = T),
+    `Prop. C Agreed` = mean(pc, na.rm = T)
+  ) %>%
+  kable()
+```
+
+# Mixed Model Approach
+
+```{r data cleaning 2}
+# Import data
+api <- read.table("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt", header = F)
+rcon <- redcapConnection(
+  url = "https://redcap.ucdenver.edu/api/",
+  token = api[1, 1]
+)
+data <- exportRecordsTyped(rcon, labels = F)
+data <- data.frame(lapply(data, as.character))
+data$scan_id <- as.numeric(data$scan_id)
+# Make scorer IDs by initial
+data$scorer <- paste0(
+  toupper(substr(data$reader_first_name, 1, 1)),
+  toupper(substr(data$reader_last_name, 1, 1))
+)
+data$scorer[data$scorer == "GG"] <- "GS"
+# Deidentify readers
+old <- c("AF", "DL", "JW")
+new <- c("A", "B", "C")
+data$scorer[data$scorer %in% old] <- new[match(data$scorer, old, nomatch = 0)]
+# Revert duplicate IDs back to original to match with gold standard
+training <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/training.key.csv")
+dups <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/duplicated.key.csv")
+training$Training_id <- as.numeric(sub("T00", "", training$Training_id))
+# Change duplicates to OG IDs
+old <- dups$duplicated_ssid..dates.shifted..17.days.
+new <- dups$original_ssid
+data$scan_id[data$scan_id %in% old] <- new[na.omit(match(data$scan_id, old))]
+# Change training IDs to OG IDs
+old <- training$Training_id
+new <- training$Cffid.Scoring.Set
+data$scan_id[data$scan_id %in% old] <- new[na.omit(match(data$scan_id, old))]
+# Convert to long
+data <- data %>%
+  select(
+    scan_id, scorer, scoresheet_timestamp,
+    bronchiectasis_rul:atelectasis___6
+  ) %>%
+  arrange(scan_id, scorer, scoresheet_timestamp) %>%
+  group_by(scan_id, scorer) %>%
   mutate(scorer_repeat = row_number()) %>%
@@
 )
-# Maybe not quite so long (one column per rater)
-data = data %>% 
-  pivot_wider(names_from = c(scorer,scorer_repeat), values_from = value,
-              names_sep = "_")
+# Format values
+data$value <- factor(data$value,
+  levels = c(
+    "Absent", "Checked", "Cylindric", "Cystic", "Unchecked", "Varicose"
+  ),
+  labels = c("Absent", "Present", "Present", "Present", "Absent", "Present")
+)
+data$value <- as.numeric(data$value) - 1
+# Separate out gold standards and duplicates
+data_dups <- data %>% filter(scan_id %in% dups$original_ssid)
+data_gs <- data %>% filter(scan_id %in% training$Cffid.Scoring.Set)
+data <- data %>%
+  filter(!scan_id %in% c(dups$original_ssid, training$Cffid.Scoring.Set))
+# Second readings only for comparison to GS
+data_gs <- data_gs %>%
+  group_by(scan_id, scorer, feature, location) %>%
+  slice_max(scorer_repeat)
+# Dummy variables for scorer, organize columns
+data <- data %>%
+  dummy_cols(select_columns = "scorer") %>%
+  select(scan_id, feature, location, scorer, scorer_A, scorer_B, scorer_C, value)
 ```
 
-Kappa interpretation table @mchughInterraterReliabilityKappa2012:
+## Data setup
 
-| Value of Kappa | Level of Agreement | % of Data that are Reliable |
-| -------------- | ------------------ | --------------------------- |
-| 0–.20          | None               | 0–4%                        |
-| .21–.39        | Minimal            | 4–15%                       |
-| .40–.59        | Weak               | 15–35%                      |
-| .60–.79        | Moderate           | 35–63%                      |
-| .80–.90        | Strong             | 64–81%                      |
-| Above.90       | Almost Perfect     | 82–100%                     |
+```{r}
+DT::datatable(data)
+```
+
+## Model
+
+- Readers $r=1,2,3$
+- Scans $s=1,2,...,S$
+- Features $f=1,2,...,F$
+- Lobes $l=1,2,...,6$
+- Outcome $y_{rsfl}$ is binary
+- Model notation: $$y_{rsf=1}=\alpha+\beta X_l + b_1+b_2+b_3+\epsilon_{rs}$$
+
+### Bronchiectasis
+
+```{r}
+#| code-fold: show
+df <- data %>% filter(feature == "Bronchiectasis")
+model <- glmer(value ~ location + (1 | scorer),
+  family = "binomial", data = df
+)
+summary(model)
+icc(model)
+```
+
+### Mucoid Impaction
+
+```{r}
+#| code-fold: show
+df <- data %>% filter(feature == "Mucoid Impaction")
+model <- glmer(value ~ location + (1 | scorer),
+  family = "binomial", data = df
+)
+summary(model)
+icc(model)
+```
+
+### Cyst
+
+```{r}
+#| code-fold: show
+df <- data %>% filter(feature == "Cyst")
+model <- glmer(value ~ location + (1 | scorer),
+  family = "binomial", data = df
+)
+summary(model)
+icc(model)
+```
+
+### Cavity
+
+```{r}
+#| code-fold: show
+df <- data %>% filter(feature == "Cavity")
+model <- glmer(value ~ location + (1 | scorer),
+  family = "binomial", data = df
+)
+summary(model)
+icc(model)
+```
+
+### Consolidation
+
+```{r}
+#| code-fold: show
+df <- data %>% filter(feature == "Consolidation")
+model <- glmer(value ~ location + (1 | scorer),
+  family = "binomial", data = df
+)
+summary(model)
+icc(model)
+```
+
+### Ground Glass
+
+```{r}
+#| code-fold: show
+df <- data %>% filter(feature == "Ground Glass")
+model <- glmer(value ~ location + (1 | scorer),
+  family = "binomial", data = df
+)
+summary(model)
+icc(model)
+```
+
+### Nodule
+
+```{r}
+#| code-fold: show
+df <- data %>% filter(feature == "Nodule")
+model <- glmer(value ~ location + (1 | scorer),
+  family = "binomial", data = df
+)
+summary(model)
+icc(model)
+```
+
+### Tree In Bud
+
+```{r}
+#| code-fold: show
+df <- data %>% filter(feature == "Tree In Bud")
+model <- glmer(value ~ location + (1 | scorer),
+  family = "binomial", data = df
+)
+summary(model)
+icc(model)
+```
+
+### Atelectasis
+
+```{r}
+#| code-fold: show
+df <- data %>% filter(feature == "Atelectasis")
+model <- glmer(value ~ location + (1 | scorer),
+  family = "binomial", data = df
+)
+summary(model)
+icc(model)
+```

### 2023-09-19 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
started change point analysis

@@
 library(lmerTest)
+library(nlme)
 library(broom.mixed)
 library(emmeans)
-library(knitr)
+library(gt)
+library(performance)
+library(mcp)
 ```
@@
   reframe(across(`2015`:`2021`, ~ sum(.x == 1))) %>%
-  kable()
+  gt()
 ```
@@
 
-```{r}
-#| warning: false
-annual %>%
-  mutate(mean_ppFEV = as.numeric(mean_ppFEV), mean_ppFVC = as.numeric(mean_ppFVC)) %>%
-  pivot_longer(c("mean_ppFEV", "mean_ppFVC")) %>%
-  ggplot(aes(x = ReviewYear, y = value, color = factor(eDWID), group = eDWID)) +
-  geom_line(alpha = 0.01) +
-  theme_bw() +
-  theme(legend.position = "none") +
-  facet_wrap(~name)
-```
+### FEV1
+
+#### Basic model
+
+The most basic model we can fit to these data uses the average values as the outcome, with terms for review year, race, and their interaction and a random intercept for each patient.
 
@@
 # Most basic model of lung function over time by race
+# Try a few different models
 mod <- lmer(
@@
 ))
-kable(means)
+label(means) = list("Race/Ethnicity","Year","Mean ppFEV","SE","df","Lower CL", "Upper CL")
+gt(means) %>% fmt_number(emmean:last_col(),decimals = 3)
 ```
@@
 ```{r}
-ggplot(means, aes(x = ReviewYear, y = emmean, color = race)) +
-  geom_line() +
-  geom_point() +
+ggplot(left_join(annual, means, by = join_by(race, ReviewYear))) +
+  geom_line(mapping = aes(x = ReviewYear, y = mean_ppFEV, group = eDWID),alpha=0.005,linewidth=0.5) +
+  geom_point(mapping = aes(x = ReviewYear, y = emmean,color=race)) +
+  geom_line(mapping = aes(x = ReviewYear, y = emmean,group = race,color=race)) +
+  geom_errorbar(mapping=aes(x = ReviewYear,ymin = emmean - SE, ymax = emmean + SE,color=race), width = .2) +
   xlab("Year") +
   ylab("LS Mean FEV1") +
-  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE), width = .2) +
-  theme_bw()
+  theme_bw() +
+  scale_color_discrete(name = "Race/Ethnicity")
 ```
+
+```{r}
+ggplot(left_join(annual, means, by = join_by(race, ReviewYear))) +
+  geom_line(mapping = aes(x = ReviewYear, y = mean_ppFEV, group = eDWID),alpha=0.005,linewidth=0.5) +
+  geom_point(mapping = aes(x = ReviewYear, y = emmean,color=race)) +
+  geom_line(mapping = aes(x = ReviewYear, y = emmean,group = race,color=race)) +
+  geom_errorbar(mapping=aes(x = ReviewYear,ymin = emmean - SE, ymax = emmean + SE,color=race), width = .2) +
+  xlab("Year") +
+  ylab("LS Mean FEV1") +
+  theme_bw() +
+  scale_color_discrete(name = "Race/Ethnicity") +
+  facet_wrap(~race)
+```
+
+#### Change point model
+
+```{r}
+#| eval: false
+t = annual %>% select(eDWID,ReviewYear,mean_ppFEV) %>% na.omit(.)
+model = list(mean_ppFEV ~ 0 + ReviewYear, 
+              ~ 0 + ReviewYear)
+fit = mcp(model,data = t,cores=4,chains = 4)
+plot(fit)
+```

### 2023-09-20 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
moving race and ETI to lambda

@@
 ```{r}
-#| eval: false
+#| cache: true
 t = annual %>% select(eDWID,ReviewYear,mean_ppFEV) %>% na.omit(.)

### 2023-09-22 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
KM and first cox model done

@@
 
-## Overall by year
+The following tables have been aggregated by year, and therefore may miss some information regarding eligibility and ETI start.
+For example, if someone turned 12 on 12/30/2019, they would technically be considered eligible for ETI, even though realistically we probably wouldn't expect them to start that year.
+
+## Proportions
+
+### Overall by year
 
@@
 annual %>%
-  select(eti_elig, ReviewYear) %>%
+  select(eti_elig, contains("Vx"), ReviewYear) %>%
   tbl_summary(
@@
 
-## By race and year
+### By race and year
 
@@
   t <- d %>%
-    select(eti_elig, race) %>%
+    select(eti_elig, contains("Vx"), race) %>%
     tbl_summary(
@@
 }))
+```
+
+## Survival analysis
+
+For the following analysis, the start time ($t=0$) was the approximate date that each person became eligible for ETI.
+Because we don't have exact dates of birth, I calculated an approximate DOB based on age and encounter dates.
+
+```{r}
+# Create survival dataset
+surv_dat <- encounter %>%
+  mutate(stop = as.numeric(encounterdate - eti_elig_date)) %>%
+  filter(stop >= 0) %>%
+  group_by(eDWID) %>%
+  mutate(start = lag(stop)) %>%
+  select(eDWID, race, encounterdate, encounterage, eti_elig_date, Vx445comb, start, stop) %>%
+  filter(stop <= first(stop[Vx445comb == "Yes"]))
+surv_dat$Vx445comb <- as.numeric(factor(surv_dat$Vx445comb, levels = c("No", "Yes")))
+# Filter for KM
+km_dat <- surv_dat %>%
+  group_by(eDWID) %>%
+  slice_max(stop)
+kmfit <- survfit(Surv(stop, Vx445comb) ~ race, data = km_dat[km_dat$race != "Asian", ])
+autoplot(kmfit) + theme_bw()
+# Cox PH model
+fit <- coxph(Surv(start, stop, Vx445comb) ~ race, id = eDWID, data = surv_dat)
 ```
@@
 ))
-label(means) = list("Race/Ethnicity","Year","Mean ppFEV","SE","df","Lower CL", "Upper CL")
-gt(means) %>% fmt_number(emmean:last_col(),decimals = 3)
+label(means) <- list("Race/Ethnicity", "Year", "Mean ppFEV", "SE", "df", "Lower CL", "Upper CL")
+gt(means) %>% fmt_number(emmean:last_col(), decimals = 3)
 ```
@@
 ggplot(left_join(annual, means, by = join_by(race, ReviewYear))) +
-  geom_line(mapping = aes(x = ReviewYear, y = mean_ppFEV, group = eDWID),alpha=0.005,linewidth=0.5) +
-  geom_point(mapping = aes(x = ReviewYear, y = emmean,color=race)) +
-  geom_line(mapping = aes(x = ReviewYear, y = emmean,group = race,color=race)) +
-  geom_errorbar(mapping=aes(x = ReviewYear,ymin = emmean - SE, ymax = emmean + SE,color=race), width = .2) +
+  geom_point(mapping = aes(x = ReviewYear, y = emmean, color = race)) +
+  geom_line(mapping = aes(x = ReviewYear, y = emmean, group = race, color = race)) +
+  geom_errorbar(mapping = aes(x = ReviewYear, ymin = emmean - SE, ymax = emmean + SE, color = race), width = .2) +
   xlab("Year") +
@@
 ggplot(left_join(annual, means, by = join_by(race, ReviewYear))) +
-  geom_line(mapping = aes(x = ReviewYear, y = mean_ppFEV, group = eDWID),alpha=0.005,linewidth=0.5) +
-  geom_point(mapping = aes(x = ReviewYear, y = emmean,color=race)) +
-  geom_line(mapping = aes(x = ReviewYear, y = emmean,group = race,color=race)) +
-  geom_errorbar(mapping=aes(x = ReviewYear,ymin = emmean - SE, ymax = emmean + SE,color=race), width = .2) +
+  geom_line(mapping = aes(x = ReviewYear, y = mean_ppFEV, group = eDWID), alpha = 0.005, linewidth = 0.5) +
+  geom_point(mapping = aes(x = ReviewYear, y = emmean, color = race)) +
+  geom_line(mapping = aes(x = ReviewYear, y = emmean, group = race, color = race)) +
+  geom_errorbar(mapping = aes(x = ReviewYear, ymin = emmean - SE, ymax = emmean + SE, color = race), width = .2) +
+  xlab("Year") +
+  ylab("LS Mean FEV1") +
+  theme_bw() +
+  scale_color_discrete(name = "Race/Ethnicity")
+```
+
+```{r}
+ggplot(left_join(annual, means, by = join_by(race, ReviewYear))) +
+  geom_line(mapping = aes(x = ReviewYear, y = mean_ppFEV, group = eDWID), alpha = 0.005, linewidth = 0.5) +
+  geom_point(mapping = aes(x = ReviewYear, y = emmean, color = race)) +
+  geom_line(mapping = aes(x = ReviewYear, y = emmean, group = race, color = race)) +
+  geom_errorbar(mapping = aes(x = ReviewYear, ymin = emmean - SE, ymax = emmean + SE, color = race), width = .2) +
   xlab("Year") +
@@
 #| cache: true
-t = annual %>% select(eDWID,ReviewYear,mean_ppFEV) %>% na.omit(.)
-model = list(mean_ppFEV ~ 0 + ReviewYear, 
-              ~ 0 + ReviewYear)
-fit = mcp(model,data = t,cores=4,chains = 4)
+#| eval: false
+t <- annual %>%
+  select(eDWID, ReviewYear, mean_ppFEV) %>%
+  na.omit(.)
+model <- list(
+  mean_ppFEV ~ 0 + ReviewYear,
+  ~ 0 + ReviewYear
+)
+fit <- mcp(model, data = t, cores = 4, chains = 4)
 plot(fit)

### 2023-09-22 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R

@@ -28,7 +28,8 @@ encounter$approx_dob <- encounter$encounterdate %m-%
     as.integer((encounter$encounterage * 12) %% 12))
 encounter <- encounter %>%
   group_by(eDWID) %>%
-  mutate(approx_dob = mean.Date(approx_dob))
+  mutate(approx_dob = mean.Date(approx_dob)) %>%
+  ungroup()
 encounter$eti_elig_date <- NA
 # If 12 years or older on 2019-10-21, eligible right away
 encounter$eti_elig_date[encounter$approx_dob <= "2007-10-21"] <- as.numeric(ymd("2019-10-21"))
@@ -45,8 +46,8 @@ encounter$eti_elig_date[encounter$approx_dob > "2009-06-09" &
 encounter$eti_elig_date[encounter$approx_dob > "2015-06-09"] <-
   encounter$approx_dob[encounter$approx_dob > "2015-06-09"] + years(6)
 # Convert back to date from numeric
-encounter$eti_elig_date = as.Date(encounter$eti_elig_date,origin ="1970-01-01")
-
+encounter$eti_elig_date <- as.Date(encounter$eti_elig_date, origin = "1970-01-01")
+encounter$approx_dob <- NULL
 # Fix missing values for specific bacteria
 bugs <- c(
   "staphylococcus_aureus", "haemophilus_influenzae",
@@ -416,7 +417,8 @@ labels <-
     "weight_last" = "Weight at Last Visit",
     "weight_perc_last" = "Weight %ile at Last Visit",
     "bmi_last" = "BMI at Last Visit", "bmi_perc_last" = "BMI %ile at Last Visit",
-    "baseline_FEV" = "Baseline FEV1", "baseline_FVC" = "Baseline FVC"
+    "baseline_FEV" = "Baseline FEV1", "baseline_FVC" = "Baseline FVC",
+    "eti_elig_date" = "Approximate Date of ETI Eligibility"
   )
 # Merge
 annual <- full_join(demo, annual, by = join_by(eDWID))

### 2023-09-25 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
initial survival analysis done

@@
 library(mcp)
+library(ggsurvfit)
 ```
@@
 # source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R")
-load("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
+# load("UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
+load("/Users/timvigers/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
 ```
@@
 Because we don't have exact dates of birth, I calculated an approximate DOB based on age and encounter dates.
+Encounters prior to ETI eligibility were excluded, and only time to first encounter on ETI was considered (i.e.
+we did not fit a multi-state model to account for those who started and then stopped the medication).
 
@@
   group_by(eDWID) %>%
-  mutate(start = lag(stop)) %>%
+  mutate(start = lag(stop, default = 0)) %>%
   select(eDWID, race, encounterdate, encounterage, eti_elig_date, Vx445comb, start, stop) %>%
   filter(stop <= first(stop[Vx445comb == "Yes"]))
-surv_dat$Vx445comb <- as.numeric(factor(surv_dat$Vx445comb, levels = c("No", "Yes")))
-# Filter for KM
-km_dat <- surv_dat %>%
-  group_by(eDWID) %>%
-  slice_max(stop)
-kmfit <- survfit(Surv(stop, Vx445comb) ~ race, data = km_dat[km_dat$race != "Asian", ])
-autoplot(kmfit) + theme_bw()
-# Cox PH model
-fit <- coxph(Surv(start, stop, Vx445comb) ~ race, id = eDWID, data = surv_dat)
+surv_dat$Vx445comb <- as.numeric(factor(surv_dat$Vx445comb, levels = c("No", "Yes"))) - 1
+# Fit
+fit <- survfit(Surv(start, stop, Vx445comb, type = "interval") ~ race,
+  id = eDWID, data = surv_dat
+)
+# Plot
+names(fit$strata) <- gsub("race=", "", names(fit$strata))
+fit %>%
+  ggsurvfit(type = "risk") +
+  add_confidence_interval() +
+  add_risktable(
+    risktable_stats =
+      c("{round(n.risk)}", "{round(cum.event)}"),
+    stats_label = c("Eligible", "Num. Started")
+  ) +
+  add_risktable_strata_symbol() +
+  add_quantile() +
+  ylab("Proportion on Vx445comb") + xlab("Days From Eligibility")
+```
+
+Without Asian subgroup for clearer plot:
+
+```{r}
+fit <- survfit(Surv(start, stop, Vx445comb, type = "interval") ~ race,
+  id = eDWID, data = surv_dat[surv_dat$race != "Asian", ]
+)
+# Plot
+names(fit$strata) <- gsub("race=", "", names(fit$strata))
+fit %>%
+  ggsurvfit(type = "risk") +
+  add_confidence_interval() +
+  add_risktable(
+    risktable_stats =
+      c("{round(n.risk)}", "{round(cum.event)}"),
+    stats_label = c("Eligible", "Num. Started")
+  ) +
+  add_risktable_strata_symbol() +
+  add_quantile() +
+  ylab("Proportion on Vx445comb") + xlab("Days From Eligibility")
 ```
@@
 ```{r}
+#| warning: false
 ggplot(left_join(annual, means, by = join_by(race, ReviewYear))) +

### 2023-09-26 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
file paths

@@
 # source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R")
-# load("UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
-load("/Users/timvigers/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
+load("UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
 ```

### 2023-09-26 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R

@@ -4,9 +4,9 @@ library(readxl)
 library(rspiro)
 library(Hmisc)
 # Import
-annual <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_Annualized_Del1.csv", na.strings = "")
-encounter <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_encountersMerged_Del1.csv", na.strings = "")
-demo <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_DemogCFDiag_Del1.csv", na.strings = "")
+annual <- read.csv("UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_Annualized_Del1.csv", na.strings = "")
+encounter <- read.csv("UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_encountersMerged_Del1.csv", na.strings = "")
+demo <- read.csv("UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_DemogCFDiag_Del1.csv", na.strings = "")
 # Dates
 demo$First_LungTransplantDate <- mdy(demo$First_LungTransplantDate)
 demo$Modulator_trikafta_first_date <- mdy(demo$Modulator_trikafta_first_date)
@@ -17,7 +17,7 @@ encounter$eti_elig[encounter$encounterdate >= ("2019-10-21") &
   encounter$encounterage >= 12] <- "Yes"
 encounter$eti_elig[encounter$encounterdate >= ("2021-06-09") &
[…]

### 2023-09-29 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
mutations in eligibility (need to update survival)

@@
 # source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R")
-load("UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
+load("/Users/timvigers/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
 ```
@@
 ```{r results='asis'}
+#| message: false
 set_gtsummary_theme(theme_gtsummary_compact(), quiet = T)
@@
 
-## Survival analysis
+<!-- ## Survival analysis -->
 
-For the following analysis, the start time ($t=0$) was the approximate date that each person became eligible for ETI.
+<!-- For the following analysis, the start time ($t=0$) was the approximate date that each person became eligible for ETI.
 Because we don't have exact dates of birth, I calculated an approximate DOB based on age and encounter dates.
@@
 we did not fit a multi-state model to account for those who started and then stopped the medication).
+-->
 
-```{r}
-# Create survival dataset
-surv_dat <- encounter %>%
-  mutate(stop = as.numeric(encounterdate - eti_elig_date)) %>%
-  filter(stop >= 0) %>%
-  group_by(eDWID) %>%
-  mutate(start = lag(stop, default = 0)) %>%
-  select(eDWID, race, encounterdate, encounterage, eti_elig_date, Vx445comb, start, stop) %>%
-  filter(stop <= first(stop[Vx445comb == "Yes"]))
-surv_dat$Vx445comb <- as.numeric(factor(surv_dat$Vx445comb, levels = c("No", "Yes"))) - 1
-# Fit
-fit <- survfit(Surv(start, stop, Vx445comb, type = "interval") ~ race,
-  id = eDWID, data = surv_dat
-)
-# Plot
-names(fit$strata) <- gsub("race=", "", names(fit$strata))
-fit %>%
-  ggsurvfit(type = "risk") +
-  add_confidence_interval() +
-  add_risktable(
-    risktable_stats =
-      c("{round(n.risk)}", "{round(cum.event)}"),
-    stats_label = c("Eligible", "Num. Started")
-  ) +
-  add_risktable_strata_symbol() +
-  add_quantile() +
-  ylab("Proportion on Vx445comb") + xlab("Days From Eligibility")
-```
+<!-- ```{r} -->
+<!-- # Create survival dataset -->
+<!-- surv_dat <- encounter %>% -->
+<!--   mutate(stop = as.numeric(encounterdate - eti_elig_date)) %>% -->
+<!--   filter(stop >= 0) %>% -->
+<!--   group_by(eDWID) %>% -->
+<!--   mutate(start = lag(stop, default = 0)) %>% -->
+<!--   select(eDWID, race, encounterdate, encounterage, eti_elig_date, Vx445comb, start, stop) %>% -->
+<!--   filter(stop <= first(stop[Vx445comb == "Yes"])) -->
+<!-- surv_dat$Vx445comb <- as.numeric(factor(surv_dat$Vx445comb, levels = c("No", "Yes"))) - 1 -->
+<!-- # Fit -->
+<!-- fit <- survfit(Surv(start, stop, Vx445comb, type = "interval") ~ race, -->
+<!--   id = eDWID, data = surv_dat -->
+<!-- ) -->
+<!-- # Plot -->
+<!-- names(fit$strata) <- gsub("race=", "", names(fit$strata)) -->
+<!-- fit %>% -->
+<!--   ggsurvfit(type = "risk") + -->
+<!--   add_confidence_interval() + -->
+<!--   add_risktable( -->
+<!--     risktable_stats = -->
+<!--       c("{round(n.risk)}", "{round(cum.event)}"), -->
+<!--     stats_label = c("Eligible", "Num. Started") -->
+<!--   ) + -->
+<!--   add_risktable_strata_symbol() + -->
+<!--   add_quantile() + -->
+<!--   ylab("Proportion on Vx445comb") + xlab("Days From Eligibility") -->
+<!-- ``` -->
 
-Without Asian subgroup for clearer plot:
+<!-- Without Asian subgroup for clearer plot: -->
 
-```{r}
-fit <- survfit(Surv(start, stop, Vx445comb, type = "interval") ~ race,
-  id = eDWID, data = surv_dat[surv_dat$race != "Asian", ]
-)
-# Plot
-names(fit$strata) <- gsub("race=", "", names(fit$strata))
-fit %>%
-  ggsurvfit(type = "risk") +
-  add_confidence_interval() +
-  add_risktable(
-    risktable_stats =
-      c("{round(n.risk)}", "{round(cum.event)}"),
-    stats_label = c("Eligible", "Num. Started")
-  ) +
-  add_risktable_strata_symbol() +
-  add_quantile() +
-  ylab("Proportion on Vx445comb") + xlab("Days From Eligibility")
-```
+<!-- ```{r} -->
+<!-- fit <- survfit(Surv(start, stop, Vx445comb, type = "interval") ~ race, -->
+<!--   id = eDWID, data = surv_dat[surv_dat$race != "Asian", ] -->
+<!-- ) -->
+<!-- # Plot -->
+<!-- names(fit$strata) <- gsub("race=", "", names(fit$strata)) -->
+<!-- fit %>% -->
+<!--   ggsurvfit(type = "risk") + -->
+<!--   add_confidence_interval() + -->
+<!--   add_risktable( -->
+<!--     risktable_stats = -->
+<!--       c("{round(n.risk)}", "{round(cum.event)}"), -->
+<!--     stats_label = c("Eligible", "Num. Started") -->
+<!--   ) + -->
+<!--   add_risktable_strata_symbol() + -->
+<!--   add_quantile() + -->
+<!--   ylab("Proportion on Vx445comb") + xlab("Days From Eligibility") -->
+<!-- ``` -->
 

### 2023-09-29 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R

@@ -4,50 +4,19 @@ library(readxl)
 library(rspiro)
 library(Hmisc)
 # Import
-annual <- read.csv("UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_Annualized_Del1.csv", na.strings = "")
-encounter <- read.csv("UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_encountersMerged_Del1.csv", na.strings = "")
-demo <- read.csv("UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_DemogCFDiag_Del1.csv", na.strings = "")
+annual <- read.csv("/Users/timvigers/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_Annualized_Del1.csv", na.strings = "")
+encounter <- read.csv("/Users/timvigers/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_encountersMerged_Del1.csv", na.strings = "")
+demo <- read.csv("/Users/timvigers/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_DemogCFDiag_Del1.csv", na.strings = "")
+# Import responsive mutations list
+responsive_mutations <- read.csv("/Users/timvigers/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/eti_responsive_mutations.csv", header = F)
+responsive_mutations <- unique(responsive_mutations$V1)
 # Dates
 demo$First_LungTransplantDate <- mdy(demo$First_LungTransplantDate)
 demo$Modulator_trikafta_first_date <- mdy(demo$Modulator_trikafta_first_date)
[…]

### 2023-09-29 Christine Chan/CF Registry/cf_registry_analysis.qmd
cf registry analyses updated

--- 2023-06-02 Christine Chan/CF Registry/cf_registry_analysis.qmd
@@
 # Import and clean data
-annual <- read.csv("~/Documents/Work/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",na.strings = c("", "N/A")
+annual <- read.csv("/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data/Data_Raw/CFF19_Annualized_Del1.csv",
+  na.strings = c("", "N/A")
 )
-annual = annual %>% filter(cfrd_status_annual != 4,cfrd_status_annual != 5)
-annual$cfrd_status_annual = factor(annual$cfrd_status_annual,levels = 1:3,
-                                   labels = c("CFRD-","CFRD-","CFRD+"))
-demogs = read.csv("~/Documents/Work/Christine Chan/CF registry data/CFF19_DemogCFDiag_Del1.csv")
-demogs = demogs %>% select(eDWID,Gender,MutClass)
-annual = left_join(annual,demogs,by = join_by(eDWID))
+annual <- annual %>% filter(cfrd_status_annual != 4, cfrd_status_annual != 5)
+annual$cfrd_status_annual <- factor(annual$cfrd_status_annual,
+  levels = 1:3,
+  labels = c("CFRD-", "CFRD-", "CFRD+")
+)
+demogs <- read.csv("/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data/Data_Raw/CFF19_DemogCFDiag_Del1.csv")
+demogs <- demogs %>% select(eDWID, Gender, MutClass)
+annual <- left_join(annual, demogs, by = join_by(eDWID))
 # Analysis function
-log_reg = function(year){
+log_reg <- function(year) {
   # Clean data
-  df = annual %>% 
-    filter(ReviewYear==year) %>%
-    select(eDWID,Age_YrEnd,Gender,MutClass,A_IsOnEnzymes,A_FEV1,A_FEV1_pct_predicted,
-           A_bmivalue,A_bmipercentile,A_height,A_weight,A_heightpercentile,
-           A_supplement_feeding,A_feeding2,
-           NumPulmExacerbation,A_pseudomonasaeruginosa,
-           A_fungalyeast1,A_pulmonarycomplications1,A_mycocultureresults1,
-           A_corticosteroids1,cfrd_status_annual,A_Vx770,A_VX809comb)
+  df <- annual %>%
+    filter(ReviewYear == year, Age_YrEnd >= 10) %>%
+    select(
+      eDWID, Age_YrEnd, Gender, MutClass, A_IsOnEnzymes, A_FEV1, A_FEV1_pct_predicted,
+      A_bmivalue, A_bmipercentile, A_height, A_weight, A_heightpercentile,
+      A_supplement_feeding, A_feeding2:A_feeding4,
+      NumPulmExacerbation, A_pseudomonasaeruginosa,
+      A_fungalyeast1, A_pulmonarycomplications1, A_mycocultureresults3,
+      A_corticosteroids1, cfrd_status_annual, A_Vx770, A_VX809comb,
+      A_hepatobiliary1_3, contains("A_hepatobiliary2_")
+    )
   # If A_supplement_feeding is 0, then A_feeding2 is also 0 because they are not
   # on any supplemental feeding
-  df$A_feeding2[df$A_supplement_feeding == 0] = 0
+  # Include all kinds of tube feeds
+  df$A_feeding2[df$A_supplement_feeding == 0] <- 0
+  df$A_feeding3[df$A_supplement_feeding == 0] <- 0
+  df$A_feeding4[df$A_supplement_feeding == 0] <- 0
+  df$tube_feeding <-
+    apply(df[, c("A_feeding2", "A_feeding3", "A_feeding4")], 1, function(r) {
+      if (all(is.na(r))) {
+        return(NA)
+      } else {
+        return(max(as.numeric(r), na.rm = T))
+      }
+    })
+  # Liver disease
+  liver <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
+  df$liver_disease <-
+    apply(df[, liver], 1, function(r) {
+      if (all(is.na(r))) {
+        return(NA)
+      } else {
+        return(max(as.numeric(r), na.rm = T))
+      }
+    })
   # Calculate percentiles
-  df$bmi = df$A_weight / ((df$A_height/100)^2)
-  df$age_perc = ifelse(df$Age_YrEnd < 20,df$Age_YrEnd,20)
-  df$height_perc = sds(df$A_height,age = df$age_perc,
-                    sex = df$Gender,male = "M",female = "F",
-                    item = "height2_20",ref = cdc.ref,type = "perc")*100
-  df$bmi_perc = sds(df$bmi,age = df$age_perc,
-                    sex = df$Gender,male = "M",female = "F",
-                    item = "bmi",ref = cdc.ref,type = "perc")*100
+  df$bmi <- df$A_weight / ((df$A_height / 100)^2)
+  df$age_perc <- ifelse(df$Age_YrEnd < 20, df$Age_YrEnd, 20)
+  df$height_perc <- sds(df$A_height,
+    age = df$age_perc,
+    sex = df$Gender, male = "M", female = "F",
+    item = "height2_20", ref = cdc.ref, type = "perc"
+  ) * 100
+  df$bmi_perc <- sds(df$bmi,
+    age = df$age_perc,
+    sex = df$Gender, male = "M", female = "F",
+    item = "bmi", ref = cdc.ref, type = "perc"
+  ) * 100
   # Format columns
-  df$MutClass = factor(df$MutClass,levels = c("1-3","4-5","Oth","N/A"),
-                       labels = c("1-3","4-5","Oth","Unknown"))
-  df$NumPulmExacerbation=cut(df$NumPulmExacerbation,labels = c("0","1","2",">=3"),
-                             breaks = c(0,1,2,3,Inf),right = F)
-  df$A_FEV1_pct_predicted = cut(df$A_FEV1_pct_predicted,
-                                breaks = c(0,40,80,100,Inf),right = F)
-  df$A_FEV1_pct_predicted = relevel(df$A_FEV1_pct_predicted,ref = "[100,Inf)")
-  df$height_perc = cut(df$height_perc,breaks = c(0,10,25,50,Inf))
-  df$height_perc = relevel(df$height_perc,ref = "(50,Inf]")
-  df$bmi_perc = cut(df$bmi_perc,breaks = c(0,10,50,85,Inf))
-  df$bmi_perc = relevel(df$bmi_perc,ref = "(50,85]")
-  df$age_group = cut(df$Age_YrEnd,breaks = c(0,10,18,30,Inf))
-  df$age_group = relevel(df$age_group,ref = "(10,18]")
+  df$MutClass <- factor(df$MutClass,
+    levels = c("1-3", "4-5", "Oth", "N/A"),
+    labels = c("1-3", "4-5", "Oth", "Unknown")
+  )
+  df$NumPulmExacerbation <- cut(df$NumPulmExacerbation,
+    labels = c("0", "1", "2", ">=3"),
+    breaks = c(0, 1, 2, 3, Inf), right = F
+  )
+  df$A_FEV1_pct_predicted <- cut(df$A_FEV1_pct_predicted,
+    breaks = c(0, 40, 80, 100, Inf), right = F
+  )
+  df$A_FEV1_pct_predicted <- relevel(df$A_FEV1_pct_predicted, ref = "[100,Inf)")
+  df$height_perc <- cut(df$height_perc, breaks = c(0, 10, 25, 50, Inf))
+  df$height_perc <- relevel(df$height_perc, ref = "(50,Inf]")
+  df$bmi_perc <- cut(df$bmi_perc, breaks = c(0, 10, 50, 85, Inf))
+  df$bmi_perc <- relevel(df$bmi_perc, ref = "(50,85]")
+  df$age_group <- cut(df$Age_YrEnd, breaks = c(10, 18, 30, Inf), right = F)
+  df$age_group <- relevel(df$age_group, ref = "[10,18)")
   # Bugs
-  bugs = c("A_pseudomonasaeruginosa","A_fungalyeast1",
-           "A_pulmonarycomplications1","A_mycocultureresults1")
+  bugs <- c(
+    "A_pseudomonasaeruginosa", "A_fungalyeast1",
+    "A_pulmonarycomplications1", "A_mycocultureresults3"
+  )
   # Yes/No variables
-  yn_vars = c("A_IsOnEnzymes","A_corticosteroids1","A_feeding2",
-              "A_Vx770","A_VX809comb",bugs)
-  df[,yn_vars] = lapply(df[,yn_vars],factor,levels = 0:1,labels = c("No","Yes"))
+  yn_vars <- c(
+    "A_IsOnEnzymes", "A_corticosteroids1", "A_feeding2",
+    "A_Vx770", "A_VX809comb", bugs, "tube_feeding", "liver_disease"
+  )
+  df[, yn_vars] <- lapply(df[, yn_vars], factor, levels = 0:1, labels = c("No", "Yes"))
+  df$A_IsOnEnzymes <- relevel(df$A_IsOnEnzymes, ref = "Yes")
   # Remove all empty columns
-  empty = which(colSums(is.na(df))==nrow(df))
-  if(length(empty)>0){
-    df = df[,-empty]
+  empty <- which(colSums(is.na(df)) == nrow(df))
+  if (length(empty) > 0) {
+    df <- df[, -empty]
   }
   # Models
-  vars = colnames(df)[which(!colnames(df) %in% 
-                              c(bugs,"cfrd_status_annual","eDWID","A_bmivalue",
-                                "A_bmipercentile","A_height","A_heightpercentile",
-                                "Age_YrEnd","A_weight","bmi", "age_perc",
-                                "A_supplement_feeding"))]
-  base_mod = as.formula(paste0("cfrd_status_annual~",paste0(vars,collapse = "+")))
+  vars <- colnames(df)[which(!colnames(df) %in%
+    c(
+      bugs, "cfrd_status_annual", "eDWID", "A_bmivalue",
+      "A_bmipercentile", "A_height", "A_heightpercentile",
+      "Age_YrEnd", "A_weight", "bmi", "age_perc",
+      "A_supplement_feeding", "A_feeding2", "A_feeding3",
+      "A_feeding4", "A_FEV1", liver
+    ))]
+  base_mod <- as.formula(paste0("cfrd_status_annual~", paste0(vars, collapse = "+")))
+  # Labels
+  labels <- list(
+    cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
+    Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
+    A_FEV1_pct_predicted = "FEV1 percent predicted",
+    tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
+    A_corticosteroids1 = "Use of corticosteroids",
+    NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
+    height_perc = "Height percentile", bmi_perc = "BMI percentile",
+    A_Vx770 = "Use of CFTR modulator Vx770",
+    A_VX809comb = "Use of CFTR modulator Vx809 combination",
+    A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
+    A_fungalyeast1 = "Aspergillus (any species)",
+    A_pulmonarycomplications1 = "ABPA",
+    A_mycocultureresults3 = "Microorganisms in myco culture"
+  )
+  var_label(df[, vars[which(vars %in% colnames(df))]]) <-
+    as.character(labels[vars[which(vars %in% colnames(df))]])
+  var_label(df[, bugs[which(bugs %in% colnames(df))]]) <-
+    as.character(labels[bugs[which(bugs %in% colnames(df))]])
   # Simple models
-  mod1 = glm(base_mod,data = df,family = "binomial")
+  mod1 <- glm(base_mod, data = df, family = "binomial")
   # Print results
@@
   cat("\n")
-  print(forest_model(mod1))
+  print(forest_model(mod1, exclude_infinite_cis = F))
   cat("\n")
-  if ("A_pseudomonasaeruginosa" %in% colnames(df)){
-    mod2 = glm(update(base_mod,.~.+A_pseudomonasaeruginosa),data = df,family = "binomial")
+  if ("A_pseudomonasaeruginosa" %in% colnames(df)) {
+    mod2 <- glm(update(base_mod, . ~ . + A_pseudomonasaeruginosa), data = df, family = "binomial")
     cat("\n")
@@
     cat("\n")
-    print(forest_model(mod2))
+    print(forest_model(mod2, exclude_infinite_cis = F))
     cat("\n")
   }
-  if("A_fungalyeast1" %in% colnames(df)){
-    mod3 = glm(update(base_mod,.~.+A_fungalyeast1),data = df,family = "binomial")
+  if ("A_fungalyeast1" %in% colnames(df)) {
+    mod3 <- glm(update(base_mod, . ~ . + A_fungalyeast1), data = df, family = "binomial")
     cat("\n")
@@
     cat("\n")
-    print(forest_model(mod3))
+    print(forest_model(mod3, exclude_infinite_cis = F))
     cat("\n")
   }
-  if("A_pulmonarycomplications1" %in% colnames(df)){
-    mod4 = glm(update(base_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
+  if ("A_pulmonarycomplications1" %in% colnames(df)) {
+    mod4 <- glm(update(base_mod, . ~ . + A_pulmonarycomplications1), data = df, family = "binomial")
     cat("\n")
@@
     cat("\n")
-    print(forest_model(mod4))
+    print(forest_model(mod4, exclude_infinite_cis = F))
     cat("\n")
   }
-  if("A_mycocultureresults1" %in% colnames(df)){
-    mod5 = glm(update(base_mod,.~.+A_mycocultureresults1),data = df,family = "binomial")
+  if ("A_mycocultureresults3" %in% colnames(df)) {
+    mod5 <- glm(update(base_mod, . ~ . + A_mycocultureresults3), data = df, family = "binomial")
     cat("\n")
@@
     cat("\n")
-    print(forest_model(mod5))
+    print(forest_model(mod5, exclude_infinite_cis = F))
     cat("\n")

### 2023-09-29 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R

@@ -275,30 +275,6 @@ demo$race <- factor(demo$race,
     "Indigenous/Mixed/Other/Unknown Race", "Indigenous/Mixed/Other/Unknown Race"
   )
 )
-# Create some of our own annualized variables
-by_year <- encounter %>%
-  mutate(across(contains("Vx"), ~ factor(.x, ordered = T))) %>%
-  group_by(eDWID, reviewyear) %>%
-  summarise(
-    mean_ppFEV = mean(GLI_FEV1_pct_predicted, na.rm = T),
-    mean_ppFVC = mean(GLI_FVC_pct_predicted, na.rm = T),
-    weight_last = last(na.omit(weight)),
-    weight_perc_last = last(na.omit(weightpercentile)),
-    bmi_last = last(na.omit(bmivalue)),
-    bmi_perc_last = last(na.omit(bmipercentile)),
-    across(contains("Vx"), ~ max(.x)),
-    .groups = "drop"
-  ) %>%
-  group_by(eDWID) %>%
-  mutate(
-    baseline_FEV = first(na.omit(mean_ppFEV)),
-    baseline_FVC = first(na.omit(mean_ppFVC))
-  ) %>%
-  ungroup() %>%
-  mutate(
-    across(contains("Vx"), ~ factor(.x, ordered = F))
-  )
-annual <- full_join(annual, by_year, by = join_by(eDWID, ReviewYear == reviewyear))
 # Manage labels (tried to do this programmatically but the data dictionary has
 # some quirks and this ended up being faster)
 labels <-
@@ -323,8 +299,10 @@ labels <-
     "pulmonarymeds_notonany" = "Not on any pulmonary meds",
     "tobi" = "Tobramycin", "tobifrequency" = "Tobramycin frequency",
     "aztreonam" = "Aztreonam", "aztreonam_freq" = "Aztreonam frequency",
-    "Vx770" = "Vx770", "Vx809comb" = "Vx809comb", "Vx661comb" = "Vx661comb",
-    "Vx445comb" = "Vx445comb", "dornasealfa" = "Dornase alfa",
+    "Vx770" = "Prescribed Ivacaftor", "Vx809comb" = "Prescribed Lumacaftor/Ivacaftor", 
+    "Vx661comb" = "Prescribed Tezacaftor/Ivacaftor",
+    "Vx445comb" = "Prescribed Elexacaftor/Tezacaftor/Ivacaftor", 
+    "dornasealfa" = "Dornase alfa",
[…]

### 2023-10-03 Christine Chan/CF Registry/cf_registry_analysis.qmd
updated CFRD registry tables

@@
 #| include: false
-library(knitr)
+library(Hmisc)
 library(tidyverse)
@@
 library(childsds)
+library(gt)
+library(labelled)
 ```
 
-# Logistic analyses
-
-Any participants with T1D or T2D were excluded from these analyses.
-Variables that were entirely missing in a given year (e.g., `A_Vx770` and `A_VX809comb` in 2005) were excluded from those models.
-Participants under age 10 at year end were also excluded due to missing PFT data.
-
 ```{r}
-# Import and clean data
-annual <- read.csv("/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data/Data_Raw/CFF19_Annualized_Del1.csv",
+# Import data and codebook
+annual <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_Annualized_Del1.csv",
   na.strings = c("", "N/A")
 )
+data_dict <- readxl::read_excel("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
+data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
+# Remove those with T1D or T2D
 annual <- annual %>% filter(cfrd_status_annual != 4, cfrd_status_annual != 5)
@@
 )
-demogs <- read.csv("/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data/Data_Raw/CFF19_DemogCFDiag_Del1.csv")
+# Get mutations
+demogs <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_DemogCFDiag_Del1.csv")
 demogs <- demogs %>% select(eDWID, Gender, MutClass)
 annual <- left_join(annual, demogs, by = join_by(eDWID))
-# Analysis function
-log_reg <- function(year) {
-  # Clean data
-  df <- annual %>%
-    filter(ReviewYear == year, Age_YrEnd >= 10) %>%
-    select(
-      eDWID, Age_YrEnd, Gender, MutClass, A_IsOnEnzymes, A_FEV1, A_FEV1_pct_predicted,
-      A_bmivalue, A_bmipercentile, A_height, A_weight, A_heightpercentile,
-      A_supplement_feeding, A_feeding2:A_feeding4,
-      NumPulmExacerbation, A_pseudomonasaeruginosa,
-      A_fungalyeast1, A_pulmonarycomplications1, A_mycocultureresults3,
-      A_corticosteroids1, cfrd_status_annual, A_Vx770, A_VX809comb,
-      A_hepatobiliary1_3, contains("A_hepatobiliary2_")
-    )
+# Variable formatting
 # If A_supplement_feeding is 0, then A_feeding2 is also 0 because they are not
@@
 # Include all kinds of tube feeds
-  df$A_feeding2[df$A_supplement_feeding == 0] <- 0
-  df$A_feeding3[df$A_supplement_feeding == 0] <- 0
-  df$A_feeding4[df$A_supplement_feeding == 0] <- 0
-  df$tube_feeding <-
-    apply(df[, c("A_feeding2", "A_feeding3", "A_feeding4")], 1, function(r) {
+annual$A_feeding2[annual$A_supplement_feeding == 0] <- 0
+annual$A_feeding3[annual$A_supplement_feeding == 0] <- 0
+annual$A_feeding4[annual$A_supplement_feeding == 0] <- 0
+tube_fields <- c("A_feeding2", "A_feeding3", "A_feeding4")
+annual$tube_feeding <-
+  apply(annual[, tube_fields], 1, function(r) {
     if (all(is.na(r))) {
@@
 # Liver disease
-  liver <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
-  df$liver_disease <-
-    apply(df[, liver], 1, function(r) {
+liver_fields <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
+annual$liver_disease <-
+  apply(annual[, liver_fields], 1, function(r) {
     if (all(is.na(r))) {
@@
 # Calculate percentiles
-  df$bmi <- df$A_weight / ((df$A_height / 100)^2)
-  df$age_perc <- ifelse(df$Age_YrEnd < 20, df$Age_YrEnd, 20)
-  df$height_perc <- sds(df$A_height,
-    age = df$age_perc,
-    sex = df$Gender, male = "M", female = "F",
+annual$bmi <- annual$A_weight / ((annual$A_height / 100)^2)
+annual$age_perc <- ifelse(annual$Age_YrEnd < 20, annual$Age_YrEnd, 20)
+annual$height_perc <- sds(annual$A_height,
+  age = annual$age_perc,
+  sex = annual$Gender, male = "M", female = "F",
   item = "height2_20", ref = cdc.ref, type = "perc"
 ) * 100
-  df$bmi_perc <- sds(df$bmi,
-    age = df$age_perc,
-    sex = df$Gender, male = "M", female = "F",
+annual$bmi_perc <- sds(annual$bmi,
+  age = annual$age_perc,
+  sex = annual$Gender, male = "M", female = "F",
   item = "bmi", ref = cdc.ref, type = "perc"
@@
 # Format columns
-  df$MutClass <- factor(df$MutClass,
+annual$MutClass <- factor(annual$MutClass,
   levels = c("1-3", "4-5", "Oth", "N/A"),
@@
 )
-  df$NumPulmExacerbation <- cut(df$NumPulmExacerbation,
+annual$NumPulmExacerbation <- cut(annual$NumPulmExacerbation,
   labels = c("0", "1", "2", ">=3"),
@@
 )
-  df$A_FEV1_pct_predicted <- cut(df$A_FEV1_pct_predicted,
+annual$A_FEV1_pct_predicted <- cut(annual$A_FEV1_pct_predicted,
   breaks = c(0, 40, 80, 100, Inf), right = F
 )
-  df$A_FEV1_pct_predicted <- relevel(df$A_FEV1_pct_predicted, ref = "[100,Inf)")
-  df$height_perc <- cut(df$height_perc, breaks = c(0, 10, 25, 50, Inf))
-  df$height_perc <- relevel(df$height_perc, ref = "(50,Inf]")
-  df$bmi_perc <- cut(df$bmi_perc, breaks = c(0, 10, 50, 85, Inf))
-  df$bmi_perc <- relevel(df$bmi_perc, ref = "(50,85]")
-  df$age_group <- cut(df$Age_YrEnd, breaks = c(10, 18, 30, Inf), right = F)
-  df$age_group <- relevel(df$age_group, ref = "[10,18)")
+annual$A_FEV1_pct_predicted <- relevel(annual$A_FEV1_pct_predicted, ref = "[100,Inf)")
+annual$height_perc <- cut(annual$height_perc, breaks = c(0, 10, 25, 50, Inf))
+annual$height_perc <- relevel(annual$height_perc, ref = "(50,Inf]")
+annual$bmi_perc <- cut(annual$bmi_perc, breaks = c(0, 10, 50, 85, Inf))
+annual$bmi_perc <- relevel(annual$bmi_perc, ref = "(50,85]")
+annual$age_group <- cut(annual$Age_YrEnd, breaks = c(10, 18, 30, Inf), right = F)
+annual$age_group <- relevel(annual$age_group, ref = "[10,18)")
 # Bugs
@@
 )
-  df[, yn_vars] <- lapply(df[, yn_vars], factor, levels = 0:1, labels = c("No", "Yes"))
-  df$A_IsOnEnzymes <- relevel(df$A_IsOnEnzymes, ref = "Yes")
+annual[, yn_vars] <- lapply(annual[, yn_vars], factor, levels = 0:1, labels = c("No", "Yes"))
+annual$A_IsOnEnzymes <- relevel(annual$A_IsOnEnzymes, ref = "Yes")
+# Analysis function
+log_reg <- function(year, missingness = F) {
+  # Clean data
+  df <- annual %>%
+    filter(ReviewYear == year, Age_YrEnd >= 10) %>%
+    select(
+      eDWID, Age_YrEnd, Gender, MutClass, A_IsOnEnzymes, A_FEV1, A_FEV1_pct_predicted,
+      A_bmivalue, A_bmipercentile, A_height, A_weight, A_heightpercentile,
+      A_supplement_feeding, tube_feeding,
+      NumPulmExacerbation, A_pseudomonasaeruginosa,
+      A_fungalyeast1, A_pulmonarycomplications1, A_mycocultureresults3,
+      A_corticosteroids1, cfrd_status_annual, A_Vx770, A_VX809comb,
+      liver_disease
+    )
   # Remove all empty columns
@@
       "A_supplement_feeding", "A_feeding2", "A_feeding3",
-      "A_feeding4", "A_FEV1", liver
+      "A_feeding4", "A_FEV1", liver_fields
     ))]
@@
     as.character(labels[bugs[which(bugs %in% colnames(df))]])
+  # Visualize missingness if required
+  if (missingness) {
+    cat("\n")
+    cat("### Missingness")
+    cat("\n")
+    cat("\n")
+    naniar::vis_miss(df)
+    cat("\n")
+  }
   # Simple models
@@
 
+# Data cleaning
+
+Registry fields included in our liver disease yes/no variable were:
+
+```{r}
+data.frame(data_dict$code_meaning[data_dict$var %in% sub("A_", "", liver_fields)]) %>%
+  gt() %>%
+  tab_options(column_labels.hidden = TRUE)
+```
+
+Registry fields included in our tube feeding yes/no variable were:
+
+```{r}
+data.frame(data_dict$code_meaning[
+  paste0(data_dict$field_name, data_dict$code_value) %in%
+    sub("A_", "", tube_fields)
+]) %>%
+  gt() %>%
+  tab_options(column_labels.hidden = TRUE)
+```
+
+# CFRD validity check
+
+```{r}
+annual %>%
+  mutate(cfrd = as.numeric(cfrd_status_annual == "CFRD+")) %>%
+  group_by(eDWID) %>%
+  summarise(
+    `CFRD Status` = any(rle(cfrd)$lengths[rle(cfrd)$values == 1] >= 2),
+    .groups = "drop"
+  ) %>%
+  mutate(`CFRD Status` = factor(`CFRD Status`,
+    levels = c(F, T),
+    labels = c("CFRD-", "CFRD+")
+  )) %>%
+  count(`CFRD Status`) %>%
+  gt()
+```
+
+# HbA1c by year
+
+```{r}
+annual %>%
+  group_by(ReviewYear) %>%
+  summarise(
+    `Num. HbA1c Tests` = sum(!is.na(A_hgba1c)),
+    `Total n` = n(),
+    `Percent with HbA1c` = round(sum(!is.na(A_hgba1c)) / n() * 100, 2)
+  ) %>%
+  gt()
+```
+
+# Logistic analyses
+
+Any participants with T1D or T2D were excluded from these analyses.
+Variables that were entirely missing in a given year (e.g., `A_Vx770` and `A_VX809comb` in 2005) were excluded from those models.
+Participants under age 10 at year end were also excluded due to missing PFT data.
+
+## 2004
+
+```{r results='asis',fig.height=10,fig.width=10}
+log_reg(2004,missingness = T)
+```
+
 ## 2006

### 2023-10-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
race and eti (numbers look correct to Scott)

--- 2023-09-26 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
@@
       dark: darkly
-bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Dropbox/Miscellaneous/american-medical-association.csl
+bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
+csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
 editor: source
@@
 # source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R")
-load("UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
+load("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
 ```
@@
 t1_vars <- c(
-  "Gender", "WasGenotyped", "F508_num", "mean_ppFEV", "Age_YrEnd",
+  "Gender", "F508_num", "mean_ppFEV", "Age_YrEnd",
   "mean_ppFEV", "weight_last", "weight_perc_last", "bmi_last",
-  "bmi_perc_last", "Vx770", "Vx809comb", "Vx661comb", "Vx445comb"
+  "bmi_perc_last","Vx445comb"
 )
@@
 demo %>%
-  select(Gender, WasGenotyped, F508_num, race) %>%
+  select(Gender, F508_num, race) %>%
   tbl_summary(by = race, missing_text = "Missing") %>%

### 2023-10-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R

@@ -4,11 +4,11 @@ library(readxl)
 library(rspiro)
 library(Hmisc)
 # Import
-annual <- read.csv("/Users/timvigers/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_Annualized_Del1.csv", na.strings = "")
-encounter <- read.csv("/Users/timvigers/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_encountersMerged_Del1.csv", na.strings = "")
-demo <- read.csv("/Users/timvigers/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_DemogCFDiag_Del1.csv", na.strings = "")
+annual <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_Annualized_Del1.csv", na.strings = "")
+encounter <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_encountersMerged_Del1.csv", na.strings = "")
+demo <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_DemogCFDiag_Del1.csv", na.strings = "")
 # Import responsive mutations list
-responsive_mutations <- read.csv("/Users/timvigers/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/eti_responsive_mutations.csv", header = F)
+responsive_mutations <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/eti_responsive_mutations.csv", header = F)
 responsive_mutations <- unique(responsive_mutations$V1)
 # Dates
 demo$First_LungTransplantDate <- mdy(demo$First_LungTransplantDate)
[…]

### 2023-10-05 Scott Sagel/PEx in CF/nacfc_2023.qmd
nacfc presentation for Mo

@@
 # Import
-chco <- read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CHCO.xlsx", na = c("", "NULL"))
-cnh <- read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CNH.xlsx")
+chco <- read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CHCO.xlsx", na = c("", "NULL"))
+cnh <- read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CNH.xlsx")
 # Format CHCO columns
@@
 # Write cleaned data
-write.csv(df,file = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/pex_merged.csv",row.names = F,na = "")
+write.csv(df, file = "/Users/timvigers/Documents/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/pex_merged.csv", row.names = F, na = "")
 # Make summary table
@@
   )
+# Import REDCap data from Eli
+data = read.csv("~/Documents/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Raw/ExacerbationsInCF222_DATA_2023-10-04_1459.csv",na.strings = "")
+# Demographics
+demo = data %>% filter(redcap_event_name == "demographics_arm_1")
 ```

### 2023-10-05 Scott Sagel/PEx in CF/ExacerbationsInCF222_R_2023-10-04_1459.r

@@ -1,0 +1,647 @@
+#Clear existing data and graphics
+rm(list=ls())
+graphics.off()
+#Load Hmisc library
+library(Hmisc)
+#Read Data
+data=read.csv('~/Documents/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Raw/ExacerbationsInCF222_DATA_2023-10-04_1459.csv')
+#Setting Labels
+
+label(data$sid)="Subject ID"
+label(data$redcap_event_name)="Event Name"
+label(data$redcap_repeat_instrument)="Repeat Instrument"
+label(data$redcap_repeat_instance)="Repeat Instance"
+label(data$redcap_data_access_group)="Data Access Group"
+label(data$site)="Site"
+label(data$yob)="Year of birth"
+label(data$sex)="Sex assigned at birth"
+label(data$race)="Race"
+label(data$mixed_race___1)="Specify races, if two or more (choice=White)"
+label(data$mixed_race___2)="Specify races, if two or more (choice=Black or African-American)"
+label(data$mixed_race___3)="Specify races, if two or more (choice=American Indian or Alaska Native)"
+label(data$mixed_race___4)="Specify races, if two or more (choice=Asian)"
+label(data$mixed_race___5)="Specify races, if two or more (choice=Native Hawaiian or Pacific Islander)"
+label(data$ethnicity)="Ethnicity"
+label(data$insurance___1)="Insurance (choice=Private insurance)"
+label(data$insurance___2)="Insurance (choice=Medicare)"
+label(data$insurance___3)="Insurance (choice=Medicaid)"
+label(data$insurance___4)="Insurance (choice=State special needs program (e.g. CHP+))"
+label(data$insurance___5)="Insurance (choice=Tricare/other military health plan)"
+label(data$insurance___6)="Insurance (choice=Indian Health Service)"
+label(data$insurance___7)="Insurance (choice=Other)"
+label(data$insurance___8)="Insurance (choice=None)"
+label(data$mutation1)="Mutation 1"
+label(data$mutation1_spec)="Specify mutation 1, if Other"
+label(data$mutation2)="Mutation 2"
+label(data$mutation2_spec)="Specify mutation 2, if Other"
+label(data$pancreatic_status)="Pancreatic status"
+label(data$demographics_complete)="Complete?"
+label(data$admit_year)="Year of admission"
+label(data$days_admit)="# of days admitted"
+label(data$age_at_admit)="Patients age at admission (in years)"
[…]

### 2023-10-05 Christine Chan/CF Registry/cf_registry_analysis.qmd
cfrd registry analysis

@@
 # Import data and codebook
-annual <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_Annualized_Del1.csv",
+annual <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/Data_Raw/CFF19_Annualized_Del1.csv",
   na.strings = c("", "N/A")
 )
-data_dict <- readxl::read_excel("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
+data_dict <- readxl::read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
 data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
@@
 # Get mutations
-demogs <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_DemogCFDiag_Del1.csv")
+demogs <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/Data_Raw/CFF19_DemogCFDiag_Del1.csv")
 demogs <- demogs %>% select(eDWID, Gender, MutClass)

### 2023-10-06 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
eti eligibility

@@
   "mean_ppFEV", "weight_last", "weight_perc_last", "bmi_last",
-  "bmi_perc_last","Vx445comb"
+  "bmi_perc_last", "Vx445comb"
 )
@@
 ```{r results='asis'}
-invisible(lapply(names(years), function(n) {
+invisible(lapply(c("2019", "2020", "2021"), function(n) {
   cat("\n")

### 2023-10-06 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R

@@ -378,8 +378,8 @@ annual <- annual %>% filter(WasGenotyped == "Yes", is.na(First_LungTransplantDat
 # ETI eligibility
 encounter$eti_elig <- apply(encounter, 1, function(r) {
   # Format variables
-  age = as.numeric(r["encounterage"])
-  date = r["encounterdate"]
+  age <- as.numeric(r["encounterage"])
+  date <- r["encounterdate"]
   # Start with not eligible by default
   elig <- "No"
   # Get mutations
@@ -409,16 +409,16 @@ app_dobs <- encounter %>%
   group_by(eDWID) %>%
   filter(row_number() == 1) %>%
   select(eDWID, approx_dob, contains("Mutation"))
-app_dobs$age_at_first = as.numeric(eti_12_up - app_dobs$approx_dob)/365.25
-app_dobs$age_at_expansion = as.numeric(eti_12_up_expansion - app_dobs$approx_dob)/365.25
-app_dobs$age_at_second = as.numeric(eti_6_up - app_dobs$approx_dob)/365.25
+app_dobs$age_at_first <- as.numeric(eti_12_up - app_dobs$approx_dob) / 365.25
+app_dobs$age_at_expansion <- as.numeric(eti_12_up_expansion - app_dobs$approx_dob) / 365.25
+app_dobs$age_at_second <- as.numeric(eti_6_up - app_dobs$approx_dob) / 365.25
 # For each person, get date they became eligible
 app_dobs$eti_elig_date <- apply(app_dobs, 1, function(r) {
   # Format date
-  dob = ymd(r["approx_dob"])
-  age_first = as.numeric(r["age_at_first"])
-  age_expansion = as.numeric(r["age_at_expansion"])
-  age_second = as.numeric(r["age_at_second"])
+  dob <- ymd(r["approx_dob"])
+  age_first <- as.numeric(r["age_at_first"])
+  age_expansion <- as.numeric(r["age_at_expansion"])
+  age_second <- as.numeric(r["age_at_second"])
   # Start with not eligible by default
   date_elig <- NA
   # Get mutations
@@ -464,14 +464,14 @@ app_dobs$eti_elig_date <- apply(app_dobs, 1, function(r) {
   return(date_elig)
 }, simplify = T)
 # Convert back to date from numeric
-app_dobs$eti_elig_date = as.Date(app_dobs$eti_elig_date, origin = "1970-01-01")
[…]

### 2023-10-10 Scott Sagel/PEx in CF/nacfc_2023.qmd
pex separate data cleaning file

--- 2023-03-29 Scott Sagel/PEx in CF/nacfc_2023.qmd
@@
 #| include: false
-library(readxl)
 library(tidyverse)
-library(gt)
 library(gtsummary)
-source("~/GitHub/CF/genotype_class_severity.R")
-# Import
-chco <- read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CHCO.xlsx", na = c("", "NULL"))
-cnh <- read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CNH.xlsx")
-# Format CHCO columns
-chco$`P. aeruginosa` <- coalesce(chco$`P. aeruginosa mucoid`, chco$`P. aeruginosa non-mucoid`)
-chco <- chco %>%
-  rename(
-    Race = `Race (White/Black/Other)`, Ethnicity = `Hispanic/Latinx`
-  ) %>%
-  mutate(
-    Genotypes1 = case_when(
-      Genotypes1 == "F508del" | Genotypes1 == "delF508" ~ "F508del",
-      .default = "Other"
-    ),
-    Genotypes2 = case_when(
-      Genotypes2 == "F508del" | Genotypes2 == "delF508" ~ "F508del",
-      .default = "Other"
-    ),
-    `FEV1 % Predicted at Admit` = suppressWarnings(as.numeric(`FEV1 % Predicted at Admit`))
-  ) %>%
-  unite(Genotype, Genotypes1, Genotypes2, sep = "/")
-# Format CNH to match CHCO
-cnh <- cnh %>%
-  rename(
-    `# Days in Admit` = `Total hospitalization days`,
-    `Age at Admit` = `Age at admission`,
-    `FEV1 % Predicted at Discharge` = `End of treamtent ppFEV1`,
-    `Unique number` = `Study ID`
-  ) %>%
-  mutate(
-    Site = "CNH",
-    `Modulator at Time of Admit` = case_when(
-      None == "Unchecked" ~ "Yes",
-      None == "Checked" ~ "No"
-    ),
-    MRSA = case_when(
-      MRSA == "Unchecked" ~ NA,
-      MRSA == "Checked" ~ TRUE
-    ),
-    `Staph aureus` = case_when(
-      MSSA == "Unchecked" ~ NA,
-      MSSA == "Checked" ~ TRUE
-    ),
-    `P. aeruginosa` = case_when(
-      `P. aeruginosa` == "Unchecked" ~ NA,
-      `P. aeruginosa` == "Checked" ~ TRUE
-    ),
-    Genotype = case_when(
-      `CFTR genotype, groups` == "F508del heterozygous" ~ "F508del/Other",
-      `CFTR genotype, groups` == "F508del homozygous" ~ "F508del/F508del",
-      `CFTR genotype, groups` == "Other" ~ "Other/Other"
-    ),
-    `Specific Modulator at Admit` = case_when(
-      `Ivacaftor (Kalydeco)` == "Checked" ~ "Other",
-      `Ivacaftor/lumacaftor (Orkambi)` == "Checked" ~ "Other",
-      `Tezacaftor/ivacaftor (Symdeko)` == "Checked" ~ "Other",
-      `Elexacaftor/tezacaftor/ivacaftor (Trikafta)` == "Checked" ~ "Trikafta",
-      .default = "None"
-    ),
-    `FEV1 % Predicted at Admit` = as.numeric(`PEx ppFEV1`)
-  )
-# Merge
-df <- full_join(chco, cnh)
-# Format column types
-df$`Modulator at Time of Admit` <- factor(df$`Modulator at Time of Admit`,
-  levels = c("No", "Yes"), ordered = T
-)
-df$`Specific Modulator at Admit`[df$`Modulator at Time of Admit` == "No"] <- "None"
-df$`Specific Modulator at Admit` <- factor(tolower(df$`Specific Modulator at Admit`),
-  levels = c("kalydeco", "none", "orkambi", "symdeko", "trikafta"),
-  labels = c("Other", "None", "Other", "Other", "Trikafta")
-)
-df$Race <- factor(df$Race,
-  levels = c(
-    "Amer. Indian or Alaskan Naitive", "Asian",
-    "Black", "Black or African American", "Black/ AA",
-    "Decline to Answer", "Hispanic", "Other",
-    "Unknown / Not Reported", "White"
-  ),
-  labels = c(
-    "Other", "Other", "Black", "Black", "Black",
-    "Unknown / Not Reported", "Other", "Other",
-    "Unknown / Not Reported", "White"
-  )
-)
-df$Ethnicity <- factor(df$Ethnicity,
-  levels = c(
-    "FALSE", "Hispanic or Latino", "HLS",
-    "NOT Hispanic or Latino", "NOT HLS",
-    "TRUE", "UNKNOWN"
-  ),
-  labels = c(
-    "NOT Hispanic or Latino", "Hispanic or Latino",
-    "Hispanic or Latino", "NOT Hispanic or Latino",
-    "NOT Hispanic or Latino", "Hispanic or Latino",
-    "Unknown"
-  )
-)
-bugs <- c("P. aeruginosa", "P. aeruginosa mucoid", "P. aeruginosa non-mucoid", "Staph aureus", "MRSA", "S. maltophilia")
-df[, bugs] <- lapply(df[, bugs], function(c) {
-  c[is.na(c)] <- "No"
-  c <- factor(c, levels = c("No", TRUE), labels = c("No", "Yes"), ordered = T)
-})
-df$`S. maltophilia`[df$Site == "CNH"] <- NA
+load("/Users/timvigers/Dropbox/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Cleaned/analysis_dataset.RData")
+```
+
+# Data cleaning notes
+
+1. FEV1 % predicted at discharge was calculated using the `rspiro` package and GLI-2012 equations.
+  - We did not distinguish between NE and SE Asian when capturing race, so all Asian participants were assumed to be NE Asian for these calculations (this was picked randomly).
+  - Some participants are under age 3 and/or under 1 meter in height, so the ppFEV1 calculations may not be exactly accurate for these people.
+
+```{r}
 # Make summary table
 summ <- df %>%
-  group_by(`Unique number`, Site, Year) %>%
-  mutate(`Change in FEV1` = `FEV1 % Predicted at Discharge` - `FEV1 % Predicted at Admit`) %>%
+  group_by(sid,redcap_data_access_group,admit_year) %>%
+  mutate(`Change in FEV1` = ppfev1_discharge - ppfev1_admit) %>%
   summarise(
-    `# of Hospitalizations per Person` = n(),
-    across(c(Sex:Genotype, `Specific Modulator at Admit`), ~ first(na.omit(.x))),
+    `# of Hospitalizations` = n(),
+    across(c(sex:genotype), ~ first(na.omit(.x))),
     across(c(`Modulator at Time of Admit`, all_of(bugs)), ~ max(.x, na.rm = T)),
@@
   ) %>%
-  mutate(Genotype = factor(Genotype,
-    levels = c(
-      "F508del/F508del", "F508del/Other",
-      "Other/F508del", "Other/Other"
-    ),
-    labels = c(
-      "F508del/F508del", "F508del/Other",
-      "F508del/Other", "Other/Other"
-    )
-  )) %>%
   select(

### 2023-10-10 Scott Sagel/PEx in CF/ExacerbationsInCF222_R_2023-10-04_1459.r

@@ -1,647 +1,0 @@
-#Clear existing data and graphics
-rm(list=ls())
-graphics.off()
-#Load Hmisc library
-library(Hmisc)
-#Read Data
-data=read.csv('~/Documents/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Raw/ExacerbationsInCF222_DATA_2023-10-04_1459.csv')
-#Setting Labels
-
-label(data$sid)="Subject ID"
-label(data$redcap_event_name)="Event Name"
-label(data$redcap_repeat_instrument)="Repeat Instrument"
-label(data$redcap_repeat_instance)="Repeat Instance"
-label(data$redcap_data_access_group)="Data Access Group"
-label(data$site)="Site"
-label(data$yob)="Year of birth"
-label(data$sex)="Sex assigned at birth"
-label(data$race)="Race"
-label(data$mixed_race___1)="Specify races, if two or more (choice=White)"
-label(data$mixed_race___2)="Specify races, if two or more (choice=Black or African-American)"
-label(data$mixed_race___3)="Specify races, if two or more (choice=American Indian or Alaska Native)"
-label(data$mixed_race___4)="Specify races, if two or more (choice=Asian)"
-label(data$mixed_race___5)="Specify races, if two or more (choice=Native Hawaiian or Pacific Islander)"
-label(data$ethnicity)="Ethnicity"
-label(data$insurance___1)="Insurance (choice=Private insurance)"
-label(data$insurance___2)="Insurance (choice=Medicare)"
-label(data$insurance___3)="Insurance (choice=Medicaid)"
-label(data$insurance___4)="Insurance (choice=State special needs program (e.g. CHP+))"
-label(data$insurance___5)="Insurance (choice=Tricare/other military health plan)"
-label(data$insurance___6)="Insurance (choice=Indian Health Service)"
-label(data$insurance___7)="Insurance (choice=Other)"
-label(data$insurance___8)="Insurance (choice=None)"
-label(data$mutation1)="Mutation 1"
-label(data$mutation1_spec)="Specify mutation 1, if Other"
-label(data$mutation2)="Mutation 2"
-label(data$mutation2_spec)="Specify mutation 2, if Other"
-label(data$pancreatic_status)="Pancreatic status"
-label(data$demographics_complete)="Complete?"
-label(data$admit_year)="Year of admission"
-label(data$days_admit)="# of days admitted"
-label(data$age_at_admit)="Patients age at admission (in years)"
[…]

### 2023-10-10 Scott Sagel/PEx in CF/create_pex_analysis_dataset.R

@@ -1,0 +1,119 @@
+library(tidyverse)
+library(rspiro)
+library(Hmisc)
+# Import REDCap data
+redcap <- read.csv("/Users/timvigers/Dropbox/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Raw/ExacerbationsInCF222_DATA_2023-10-04_1459.csv", na.strings = "")
+# Separate demographics dataset
+demo <- redcap %>%
+  filter(redcap_event_name == "demographics_arm_1") %>%
+  select(sid, redcap_data_access_group, sex, race, ethnicity, mutation1, mutation2)
+# Format columns - race and sex should be formatted for ppFEV1 calculations
+demo$sex <- factor(demo$sex, levels = c(2, 1), labels = c("Male", "Female"))
+demo$race <- factor(demo$race,
+  levels = c(1, 2, 4, 9, 3, 5:8),
+  labels = c(
+    "White", "Black or African-American", "NE Asian", "SE Asian",
+    "Other/mixed", "Other/mixed", "Other/mixed", "Other/mixed", "Other/mixed"
+  )
+)
+demo$ethnicity <- factor(demo$ethnicity,
+  levels = 1:3,
+  labels = c(
+    "Hispanic or Latino", "Not Hispanic or Latino", "Unknown or not reported"
+  )
+)
+demo$genotype <- rowSums(demo[, c("mutation1", "mutation2")] == 1)
+demo$genotype <- factor(demo$genotype,
+  levels = 0:2,
+  labels = c("No F508del", "F508del heterozygous", "F508del homozygous")
+)
+demo[, c("mutation1", "mutation2")] <- NULL
+# Remove CHCO REDCap data - there is some weird issue with the year of admission
+# field, so Elin sent a complete file separately
+chco <- redcap %>%
+  filter(
+    redcap_data_access_group == "chco",
+    redcap_event_name != "demographics_arm_1"
+  ) %>%
+  select(
+    sid, redcap_data_access_group, redcap_repeat_instance,
+    admit_cx_results___1:admit_cx_results___9
+  )
+redcap <- redcap %>%
+  filter(redcap_data_access_group != "chco") %>%
+  # Also remove demographics
+  filter(redcap_event_name != "demographics_arm_1") %>%
+  # Limit data to necessary columns
+  select(
[…]

### 2023-10-11 Scott Sagel/PEx in CF/nacfc_2023.qmd
basic PEx tables done (need p values)

@@
 library(gtsummary)
-load("/Users/timvigers/Dropbox/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Cleaned/analysis_dataset.RData")
+# source("~/GitHub/CF/Scott Sagel/PEx in CF/create_pex_analysis_dataset.R")
+load("~/Dropbox/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Cleaned/analysis_dataset.RData")
 ```
@@
   - Some participants are under age 3 and/or under 1 meter in height, so the ppFEV1 calculations may not be exactly accurate for these people.
-
-```{r}
-# Make summary table
-summ <- df %>%
-  group_by(sid,redcap_data_access_group,admit_year) %>%
-  mutate(`Change in FEV1` = ppfev1_discharge - ppfev1_admit) %>%
-  summarise(
-    `# of Hospitalizations` = n(),
-    across(c(sex:genotype), ~ first(na.omit(.x))),
-    across(c(`Modulator at Time of Admit`, all_of(bugs)), ~ max(.x, na.rm = T)),
-    across(
-      all_of(c(
-        "Age at Admit", "# Days in Admit", "Change in FEV1",
-        "FEV1 % Predicted at Admit", "FEV1 % Predicted at Discharge"
-      )),
-      ~ mean(.x, na.rm = T)
-    ),
-    .groups = "drop"
-  ) %>%
-  select(
-    `Unique number`, `Site`, Year, `# of Hospitalizations per Person`, `# Days in Admit`, `Age at Admit`,
-    `Change in FEV1`, `FEV1 % Predicted at Admit`, `FEV1 % Predicted at Discharge`,
-    `Modulator at Time of Admit`, `Specific Modulator at Admit`,
-    all_of(bugs), Sex:Genotype
-  )
-```
-
-# Data cleaning notes
-
-1. For each person with multiple admissions, continuous variables were averaged prior to generating summary tables.
-2.
-If a participant was on a modulators during any admission, they were counted as on modulators for that year.
-For example, participant 4 from CHOC was hospitalized 3 times in 2018 and was on a modulator during one of those admissions, so they were counted as on modulators in 2018.
-3. CNH only included combined P. aeruginosa information.
-4. Participants were uniquely identified by ID, site, and year.
+2. For each person with multiple admissions, continuous variables were averaged prior to generating summary tables.
+3. If a participant was on a modulators during any admission, they were counted as on modulators for that year.
 
@@
 #| message: false
-summ %>%
-  select(-`Unique number`, -Site) %>%
+summary_table %>%
+  select(-sid, -redcap_data_access_group) %>%
+  select(num_hosp, days_admit:ppfev1_diff, age_at_admit, everything()) %>%
   tbl_summary(
-    by = Year,
-    type = list(`# of Hospitalizations per Person` ~ "continuous"),
+    by = admit_year,
+    type = list(num_hosp ~ "continuous"),
     statistic = list(
-      c("# of Hospitalizations per Person", "# Days in Admit") ~ "{median} ({p25}, {p75})",
-      c(
-        "Age at Admit", "Change in FEV1", "FEV1 % Predicted at Admit",
-        "FEV1 % Predicted at Discharge"
-      ) ~ "{mean} ({sd})"
+      all_continuous() ~ "{mean} ({sd})",
+      num_hosp ~ "{median} [{p25}, {p75}]",
+      days_admit ~ "{median} [{p25}, {p75}]"
     ),
-    missing_text = "(Missing)"
-  ) %>%
-  add_ci(pattern = "{stat} [{ci}]") %>%
-  add_p(test = list(
-    `Age at Admit` ~ "t.test",
-    `Change in FEV1` ~ "t.test"
-  )) %>%
-  separate_p_footnotes() %>%
-  as_gt() %>%
-  as_raw_html()
+    missing_text = "(Missing)",
+    label = labels
+  )
 ```
@@
 #| message: false
-summ %>%
-  select(-`Unique number`, -`S. maltophilia`) %>%
+summary_table %>%
+  select(-sid, -cftr_mod_spec) %>%
+  select(num_hosp, days_admit:ppfev1_diff, age_at_admit, everything()) %>%
   tbl_strata(
-    strata = Site,
+    strata = redcap_data_access_group,
     .tbl_fun =
@@
         tbl_summary(
-          by = Year,
-          type = list(`# of Hospitalizations per Person` ~ "continuous"),
+          by = admit_year,
+          type = list(num_hosp ~ "continuous"),
           statistic = list(
-            c("# of Hospitalizations per Person", "# Days in Admit") ~ "{median} ({p25}, {p75})",
-            c(
-              "Age at Admit", "Change in FEV1", "FEV1 % Predicted at Admit",
-              "FEV1 % Predicted at Discharge"
-            ) ~ "{mean} ({sd})"
+            all_continuous() ~ "{mean} ({sd})",
+            num_hosp ~ "{median} [{p25}, {p75}]",
+            days_admit ~ "{median} [{p25}, {p75}]"
           ),
-          missing_text = "(Missing)"
-        ) %>%
-        add_ci(pattern = "{stat} [{ci}]") %>%
-        add_p(test = list(
-          `Age at Admit` ~ "t.test",
-          `Change in FEV1` ~ "t.test"
-        )) %>%
-        separate_p_footnotes()
-  ) %>%
-  as_gt() %>%
-  as_raw_html()
+          missing_text = "(Missing)",
+          label = labels
+        )
+  )
 ```

### 2023-10-11 Scott Sagel/PEx in CF/create_pex_analysis_dataset.R

@@ -1,8 +1,7 @@
 library(tidyverse)
 library(rspiro)
-library(Hmisc)
 # Import REDCap data
-redcap <- read.csv("/Users/timvigers/Dropbox/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Raw/ExacerbationsInCF222_DATA_2023-10-04_1459.csv", na.strings = "")
+redcap <- read.csv("~/Dropbox/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Raw/ExacerbationsInCF222_DATA_2023-10-04_1459.csv", na.strings = "")
 # Separate demographics dataset
 demo <- redcap %>%
   filter(redcap_event_name == "demographics_arm_1") %>%
@@ -46,20 +45,21 @@ redcap <- redcap %>%
   # Limit data to necessary columns
   select(
     sid, redcap_data_access_group, redcap_repeat_instance,
+    cftr_mod_spec, cftr_mod,
     admit_year:age_at_admit, height, fev1_admit, fev1_discharge,
     admit_cx_results___1:admit_cx_results___9
   )
 # Import supplemental data from Elin
-supplement <- readxl::read_excel("/Users/timvigers/Dropbox/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Raw/ExacerbationsInCF222-Admissions and Inpatient Meds All Fields 2023 09 29.xlsx")
+supplement <- readxl::read_excel("~/Dropbox/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Raw/ExacerbationsInCF222-Admissions and Inpatient Meds All Fields 2023 09 29.xlsx")
 supplement <- supplement %>% select(
-  sid, redcap_repeat_instance,
+  sid, redcap_repeat_instance, cftr_mod_spec, cftr_mod,
   admit_year:age_at_admit, height, fev1_admit, fev1_discharge
 )
 # Merge everything together and sort columns
 chco <- left_join(chco, supplement, by = join_by(sid, redcap_repeat_instance))
 df <- bind_rows(chco, redcap)
 df <- left_join(df, demo)
-df <- df %>% select(all_of(colnames(demo)), everything())
+df <- df %>% select(all_of(colnames(demo)), admit_year, everything())
 # Calculate FEV1 % predicted
 df$ppfev1_admit <- pctpred_GLI(
   age = as.numeric(df$age_at_admit),
[…]

### 2023-10-12 Scott Sagel/PEx in CF/nacfc_2023.qmd
rough draft of nacfc presentation done

@@
 editor: source
-execute: 
-  echo: false
 ---
@@
 load("~/Dropbox/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Cleaned/analysis_dataset.RData")
+# Colors for plots
+yellow <- "#f8b80a"
+red <- "#e30022"
+purple <- "#7b2670"
+blue <- "#022451"
 ```
@@
 
-The number of hospitalizations for PEx decreased from `r sum(df$Year==2018)` in 2018 to `r sum(df$Year==2022)` in 2022.
-
-# Summary Statistics
+# Summary Statistics (Per Person)
 
@@
     label = labels
-  )
+  ) %>%
+  add_p(
+    test = list(
+      all_continuous() ~ "t.test",
+      num_hosp ~ "wilcox.test",
+      days_admit ~ "wilcox.test"
+    ),
+    test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
+  ) %>%
+  separate_p_footnotes()
 ```
@@
           label = labels
-        )
+        ) %>%
+        add_p(
+          test = list(
+            all_continuous() ~ "t.test",
+            num_hosp ~ "wilcox.test",
+            days_admit ~ "wilcox.test"
+          ),
+          test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
+        ) %>%
+        separate_p_footnotes()
   )
 ```
+
+# Plots
+
+## Slide 9
+
+```{r}
+# Number of hospitalizations
+df %>%
+  group_by(redcap_data_access_group, admit_year) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  ggplot(aes(x = redcap_data_access_group, y = n, fill = factor(admit_year))) +
+  geom_bar(position = "dodge", stat = "identity") +
+  xlab("Site") +
+  ylab("Num. of Hospitalizations") +
+  theme_bw() +
+  theme(legend.title = element_blank()) +
+  scale_fill_manual(values = c(purple, yellow))
+# % hospitalized 2 or more times in both years
+df %>%
+  group_by(sid, redcap_data_access_group, admit_year) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  group_by(redcap_data_access_group, admit_year) %>%
+  summarise(perc = mean(n >= 2) * 100, .groups = "drop") %>%
+  ggplot(aes(x = redcap_data_access_group, y = perc, fill = factor(admit_year))) +
+  geom_bar(position = "dodge", stat = "identity") +
+  xlab("Site") +
+  ylab("Percent With >= 2 Hospitalizations") +
+  theme_bw() +
+  theme(legend.title = element_blank()) +
+  scale_fill_manual(values = c(purple, yellow))
+```
+
+## Slide 10
+
+For the box and whisker plot of age at admission, we used the average age for people who were admitted more than once in a given year.
+
+```{r}
+#| message: false
+#| warning: false
+# Age
+df %>%
+  group_by(sid, redcap_data_access_group, admit_year) %>%
+  summarise(age_at_admit = mean(age_at_admit, na.rm = T), .groups = "drop") %>%
+  ggplot(aes(x = redcap_data_access_group, y = age_at_admit, 
+             fill = factor(admit_year))) +
+  geom_boxplot() +
+  xlab("Site") +
+  ylab("Age at Admission") +
+  theme_bw() +
+  theme(legend.title = element_blank()) +
+  scale_fill_manual(values = c(purple, red))
+# Sex
+df %>%
+  group_by(redcap_data_access_group, admit_year, sid) %>%
+  summarise(sex = unique(sex)) %>%
+  summarise(Male = mean(sex == "Male") * 100, Female = mean(sex == "Female") * 100) %>%
+  pivot_longer(c(Male, Female), names_to = "sex", values_to = "perc") %>%
+  ggplot(aes(x = factor(admit_year), y = perc, fill = factor(sex))) +
+  geom_bar(position = "stack", stat = "identity") +
+  geom_text(aes(label = paste(round(perc, 1), "%")),
+    position = position_stack(vjust = 0.5), size = 2
+  ) +
+  ylab("%") +
+  facet_grid(~redcap_data_access_group) +
+  theme_bw() +
+  theme(legend.title = element_blank(), axis.title.x = element_blank()) +
+  scale_fill_manual(values = c(yellow, red))
+# Race
+df %>%
+  group_by(redcap_data_access_group, admit_year, sid) %>%
+  summarise(race = unique(race)) %>%
+  summarise(
+    White = mean(race == "White") * 100,
+    `Black or African-American` = mean(race == "Black or African-American") * 100,
+    Asian = mean(race == "NE Asian") * 100,
+    `Other/mixed` = mean(race == "Other/mixed") * 100
+  ) %>%
+  pivot_longer(c(
+    White, `Black or African-American`,
+    Asian, `Other/mixed`
+  ), names_to = "race", values_to = "perc") %>%
+  ggplot(aes(x = factor(admit_year), y = perc, fill = factor(race))) +
+  geom_bar(position = "stack", stat = "identity") +
+  geom_text(aes(label = paste(round(perc, 1), "%")),
+    position = position_stack(vjust = 0.5), size = 2
+  ) +
+  ylab("%") +
+  facet_grid(~redcap_data_access_group) +
+  theme_bw() +
+  theme(legend.title = element_blank(), axis.title.x = element_blank()) +
+  scale_fill_manual(values = c(red, purple, blue, yellow))
+# Ethnicity
+df %>%
+  group_by(redcap_data_access_group, admit_year, sid) %>%
+  summarise(ethnicity = unique(ethnicity)) %>%
+  summarise(
+    `Hispanic or Latino` = mean(ethnicity == "Hispanic or Latino") * 100,
+    `Not Hispanic or Latino` = mean(ethnicity == "Not Hispanic or Latino") * 100,
+    `Unknown or not reported` = mean(ethnicity == "Unknown or not reported") * 100
+  ) %>%
+  pivot_longer(c(
+    `Hispanic or Latino`, `Not Hispanic or Latino`, `Unknown or not reported`
+  ), names_to = "ethnicity", values_to = "perc") %>%
+  ggplot(aes(x = factor(admit_year), y = perc, fill = factor(ethnicity))) +
+  geom_bar(position = "stack", stat = "identity") +
+  geom_text(aes(label = paste(round(perc, 1), "%")),
+    position = position_stack(vjust = 0.5), size = 2
+  ) +
+  ylab("%") +
+  facet_grid(~redcap_data_access_group) +
+  theme_bw() +
+  theme(legend.title = element_blank(), axis.title.x = element_blank()) +
+  scale_fill_manual(values = c(red, purple, yellow))
+```
+
+## Slide 11
+
+```{r}
+#| message: false
+# Ethnicity
+df %>%
+  group_by(redcap_data_access_group, admit_year, sid) %>%
+  summarise(genotype = unique(genotype)) %>%
+  summarise(
+    `No F508del` = mean(genotype == "No F508del") * 100,
+    `F508del heterozygous` = mean(genotype == "F508del heterozygous") * 100,
+    `F508del homozygous` = mean(genotype == "F508del homozygous") * 100
+  ) %>%
+  pivot_longer(c(
+    `No F508del`, `F508del heterozygous`, `F508del homozygous`
+  ), names_to = "genotype", values_to = "perc") %>%
+  ggplot(aes(x = factor(admit_year), y = perc, fill = factor(genotype))) +
+  geom_bar(position = "stack", stat = "identity") +
+  geom_text(aes(label = paste(round(perc, 1), "%")),
+    position = position_stack(vjust = 0.5), size = 2
+  ) +
+  ylab("%") +
+  facet_grid(~redcap_data_access_group) +
+  theme_bw() +
+  theme(legend.title = element_blank(), axis.title.x = element_blank()) +
+  scale_fill_manual(values = c(red, purple, yellow))
+```
+
+## Slide 12
+
+For people with multiple admissions, FEV1 % predicted was averaged at admission and at discharge. Baseline FEV1 % predicted is the average of up to 6 months of FEV1 measures prior to first admission.
+
+```{r}
+#| warning: false
+df %>% group_by(redcap_data_access_group,admit_year,sid) %>%
+  summarise(across(ppfev1_admit:ppfev1_baseline,~mean(.x,na.rm=T)),
+            .groups = "drop") %>%
+  rename(Admission = ppfev1_admit,Discharge = ppfev1_discharge,
+         Baseline = ppfev1_baseline) %>%
+  pivot_longer(c(Admission, Discharge, Baseline)) %>%
+  mutate(name = factor(name,levels=c("Baseline","Admission", "Discharge"))) %>%
+  ggplot(aes(x = name, y = value, 
+             fill = factor(admit_year))) +
+  geom_boxplot() + ylab("FEV1 % predicted") +
+  facet_wrap(~redcap_data_access_group)+
+  theme_bw() +
+  theme(legend.title = element_blank(),legend.position = "bottom",
+        axis.title.x = element_blank()) +
+  scale_fill_manual(values = c(purple, yellow))
+```

### 2023-10-12 Scott Sagel/PEx in CF/create_pex_analysis_dataset.R

@@ -1,11 +1,20 @@
 library(tidyverse)
 library(rspiro)
 # Import REDCap data
-redcap <- read.csv("~/Dropbox/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Raw/ExacerbationsInCF222_DATA_2023-10-04_1459.csv", na.strings = "")
-# Separate demographics dataset
-demo <- redcap %>%
+df <- readxl::read_excel("~/Library/CloudStorage/Dropbox/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Raw/ExacerbationsInCF222_DATA_2023-10-11_1252 All Data.xlsx")
+# Separate datasets
+demo <- df %>%
   filter(redcap_event_name == "demographics_arm_1") %>%
   select(sid, redcap_data_access_group, sex, race, ethnicity, mutation1, mutation2)
+fevs <- c("fev1_admit", "fev1_discharge", paste0("fev1_baseline", 1:6))
+df <- df %>%
+  filter(redcap_event_name != "demographics_arm_1") %>%
+  select(
+    sid, redcap_repeat_instance, cftr_mod, cftr_mod_spec,
+    admit_year:age_at_admit, height, height_baseline1,
+    all_of(fevs),
+    admit_cx_results___1:admit_cx_results___9
+  )
 # Format columns - race and sex should be formatted for ppFEV1 calculations
 demo$sex <- factor(demo$sex, levels = c(2, 1), labels = c("Male", "Female"))
 demo$race <- factor(demo$race,
@@ -27,56 +36,43 @@ demo$genotype <- factor(demo$genotype,
   labels = c("No F508del", "F508del heterozygous", "F508del homozygous")
 )
 demo[, c("mutation1", "mutation2")] <- NULL
-# Remove CHCO REDCap data - there is some weird issue with the year of admission
-# field, so Elin sent a complete file separately
-chco <- redcap %>%
-  filter(
-    redcap_data_access_group == "chco",
-    redcap_event_name != "demographics_arm_1"
-  ) %>%
-  select(
-    sid, redcap_data_access_group, redcap_repeat_instance,
-    admit_cx_results___1:admit_cx_results___9
-  )
-redcap <- redcap %>%
-  filter(redcap_data_access_group != "chco") %>%
-  # Also remove demographics
[…]

### 2023-10-19 Scott Sagel/PEx in CF/nacfc_2023.qmd
PEx added insurance

--- 2023-10-11 Scott Sagel/PEx in CF/nacfc_2023.qmd
@@
 library(gtsummary)
-# source("~/GitHub/CF/Scott Sagel/PEx in CF/create_pex_analysis_dataset.R")
-load("~/Dropbox/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Cleaned/analysis_dataset.RData")
+library(ggpubr)
+source("~/GitHub/CF/Scott Sagel/PEx in CF/create_pex_analysis_dataset.R")
+load("~/Library/CloudStorage/Dropbox/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Cleaned/analysis_dataset.RData")
+# Colors for plots
+yellow <- "#f8b80a"
+red <- "#e30022"
+purple <- "#7b2670"
+blue <- "#022451"
 ```
@@
           label = labels
-        )
+        ) %>%
+        add_p(
+          test = list(
+            all_continuous() ~ "t.test",
+            num_hosp ~ "wilcox.test",
+            days_admit ~ "wilcox.test"
+          ),
+          test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
+        ) %>%
+        separate_p_footnotes()
   )
 ```
+
+# Plots
+
+## Slide 9
+
+```{r}
+# Number of hospitalizations
+df %>%
+  ggplot(aes(x = factor(admit_year), fill = factor(admit_year))) +
+  geom_bar() +
+  xlab("Site") +
+  ylab("Num. of Hospitalizations") +
+  theme_bw() +
+  theme(legend.position = "none") +
+  scale_fill_manual(values = c(purple, yellow))
+# % hospitalized 2 or more times in both years
+df %>%
+  group_by(sid, admit_year) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  group_by(admit_year) %>%
+  summarise(perc = mean(n >= 2) * 100, .groups = "drop") %>%
+  ggplot(aes(x = factor(admit_year), y = perc, fill = factor(admit_year))) +
+  geom_bar(position = "dodge", stat = "identity") +
+  geom_signif(
+    xmin = 1, xmax = 2, y_position = 50, map_signif_level = T,
+    annotation = "<0.001"
+  ) +
+  xlab("Site") +
+  ylab("Percent With >= 2 Hospitalizations") +
+  theme_bw() +
+  theme(legend.position = "none") +
+  scale_fill_manual(values = c(purple, yellow))
+```
+
+## Slide 10
+
+For the box and whisker plot of age at admission, we used the average age for people who were admitted more than once in a given year.
+
+```{r}
+#| warning: false
+#| message: false
+# Age
+df %>%
+  group_by(sid, redcap_data_access_group, admit_year) %>%
+  summarise(age_at_admit = mean(age_at_admit, na.rm = T), .groups = "drop") %>%
+  ggplot(aes(x = factor(admit_year), y = age_at_admit, fill = factor(admit_year))) +
+  geom_boxplot() +
+  geom_signif(
+    xmin = 1, xmax = 2, y_position = 28, map_signif_level = T,
+    annotation = 0.15
+  ) +
+  xlab("Site") +
+  ylab("Age at Admission") +
+  theme_bw() +
+  theme(legend.position = "none") +
+  scale_fill_manual(values = c(purple, red))
+# Sex
+df %>%
+  group_by(admit_year, sid) %>%
+  summarise(sex = unique(sex)) %>%
+  summarise(Male = mean(sex == "Male") * 100, Female = mean(sex == "Female") * 100) %>%
+  pivot_longer(c(Male, Female), names_to = "sex", values_to = "perc") %>%
+  ggplot(aes(x = factor(admit_year), y = perc, fill = factor(sex))) +
+  geom_bar(position = "stack", stat = "identity") +
+  geom_signif(
+    xmin = 1, xmax = 2, y_position = 105, map_signif_level = T,
+    annotation = 0.12
+  ) +
+  ylab("%") +
+  theme_bw() +
+  theme(legend.title = element_blank(), axis.title.x = element_blank()) +
+  scale_fill_manual(values = c(yellow, red))
+# Race
+df %>%
+  group_by(admit_year, sid) %>%
+  summarise(race = unique(race)) %>%
+  summarise(
+    White = mean(race == "White") * 100,
+    `Black or African-American` = mean(race == "Black or African-American") * 100,
+    Asian = mean(race == "NE Asian") * 100,
+    `Other/mixed` = mean(race == "Other/mixed") * 100
+  ) %>%
+  pivot_longer(c(
+    White, `Black or African-American`,
+    Asian, `Other/mixed`
+  ), names_to = "race", values_to = "perc") %>%
+  ggplot(aes(x = factor(admit_year), y = perc, fill = factor(race))) +
+  geom_bar(position = "stack", stat = "identity") +
+  geom_signif(
+    xmin = 1, xmax = 2, y_position = 105, map_signif_level = T,
+    annotation = 0.012
+  ) +
+  ylab("%") +
+  theme_bw() +
+  theme(legend.title = element_blank(), axis.title.x = element_blank()) +
+  scale_fill_manual(values = c(red, purple, blue, yellow))
+# Ethnicity
+df %>%
+  group_by(admit_year, sid) %>%
+  summarise(ethnicity = unique(ethnicity)) %>%
+  summarise(
+    `Hispanic or Latino` = mean(ethnicity == "Hispanic or Latino") * 100,
+    `Not Hispanic or Latino` = mean(ethnicity == "Not Hispanic or Latino") * 100,
+    `Unknown or not reported` = mean(ethnicity == "Unknown or not reported") * 100
+  ) %>%
+  pivot_longer(c(
+    `Hispanic or Latino`, `Not Hispanic or Latino`, `Unknown or not reported`
+  ), names_to = "ethnicity", values_to = "perc") %>%
+  ggplot(aes(x = factor(admit_year), y = perc, fill = factor(ethnicity))) +
+  geom_bar(position = "stack", stat = "identity") +
+  geom_signif(
+    xmin = 1, xmax = 2, y_position = 105, map_signif_level = T,
+    annotation = 0.088
+  ) +
+  ylab("%") +
+  theme_bw() +
+  theme(legend.title = element_blank(), axis.title.x = element_blank()) +
+  scale_fill_manual(values = c(red, purple, yellow))
+```
+
+## Slide 11
+
+```{r}
+#| message: false
+# Genotype
+df %>%
+  group_by(admit_year, sid) %>%
+  summarise(genotype = unique(genotype)) %>%
+  summarise(
+    `No F508del` = mean(genotype == "No F508del") * 100,
+    `F508del heterozygous` = mean(genotype == "F508del heterozygous") * 100,
+    `F508del homozygous` = mean(genotype == "F508del homozygous") * 100
+  ) %>%
+  pivot_longer(c(
+    `No F508del`, `F508del heterozygous`, `F508del homozygous`
+  ), names_to = "genotype", values_to = "perc") %>%
+  ggplot(aes(x = factor(admit_year), y = perc, fill = factor(genotype))) +
+  geom_bar(position = "stack", stat = "identity") +
+  geom_signif(
+    xmin = 1, xmax = 2, y_position = 105, map_signif_level = T,
+    annotation = "<0.001"
+  ) +
+  ylab("%") +
+  theme_bw() +
+  theme(legend.title = element_blank(), axis.title.x = element_blank()) +
+  scale_fill_manual(values = c(red, purple, yellow))
+```
+
+## Slide 12
+
+For people with multiple admissions, FEV1 % predicted was averaged at admission and at discharge. Baseline FEV1 % predicted is the average of up to 6 months of FEV1 measures prior to first admission.
+
+```{r}
+#| warning: false
+df %>%
+  group_by(admit_year, sid) %>%
+  summarise(across(ppfev1_admit:ppfev1_baseline, ~ mean(.x, na.rm = T)),
+    .groups = "drop"
+  ) %>%
+  rename(
+    Admission = ppfev1_admit, Discharge = ppfev1_discharge,
+    Baseline = ppfev1_baseline
+  ) %>%
+  pivot_longer(c(Admission, Discharge, Baseline)) %>%
+  mutate(name = factor(name, levels = c("Baseline", "Admission", "Discharge"))) %>%
+  ggplot(aes(
+    x = name, y = value,
+    fill = factor(admit_year)
+  )) +
+  geom_boxplot() +
+  geom_pwc(
+    aes(group = admit_year),
+    tip.length = 0.01,
+    method = "t_test", label = "p.format"
+  ) +
+  ylab("FEV1 % predicted") +
+  theme_bw() +
+  theme(
+    legend.title = element_blank(), legend.position = "bottom",
+    axis.title.x = element_blank()
+  ) +
+  scale_fill_manual(values = c(purple, yellow))
+```
+
+```{r}
+# Proportions who returned to >= 90% and >= 100% at discharge
+df %>%
+  mutate(
+    returned_90 = ppfev1_discharge >= ppfev1_baseline * 0.9,
+    returned_100 = ppfev1_discharge >= ppfev1_baseline
+  ) %>%
+  select(sid, admit_year, redcap_repeat_instance, contains("returned")) %>%
+  group_by(admit_year) %>%
+  summarise(
+    returned_90 = mean(returned_90, na.rm = T) * 100,
+    returned_100 = mean(returned_100, na.rm = T) * 100
+  ) %>%
+  pivot_longer(c(returned_90, returned_100)) %>%
+  mutate(name = factor(name,
+    levels = c("returned_90", "returned_100"),
+    labels = c(
+      "Returned to 90% of Baseline",
+      "Returned to 100% of Baseline"
+    )
+  )) %>%
+  ggplot(aes(x = factor(name), y = value, fill = factor(admit_year))) +
+  geom_bar(position = "dodge", stat = "identity") +
+  geom_signif(
+    xmin = c(0.75, 1.75), xmax = c(1.25, 2.25), y_position = 95, map_signif_level = T,
+    annotation = c(0.7, 0.2)
+  ) +
+  ylab("Percent of Admissions") +
+  theme_bw() +
+  theme(axis.title.x = element_blank(), legend.title = element_blank()) +
+  scale_fill_manual(values = c(purple, yellow))
+```
+
+The above plot represents the proportion of admissions where the patient returned to close to baseline lung function, and was not aggregated by patient.
+In other words, multiple admissions for the same patient were considered independent in this case.

### 2023-10-19 Scott Sagel/PEx in CF/create_pex_analysis_dataset.R

@@ -5,7 +5,10 @@ df <- readxl::read_excel("~/Library/CloudStorage/Dropbox/Work/Vigers/CF/Scott Sa
 # Separate datasets
 demo <- df %>%
   filter(redcap_event_name == "demographics_arm_1") %>%
-  select(sid, redcap_data_access_group, sex, race, ethnicity, mutation1, mutation2)
+  select(
+    sid, redcap_data_access_group, sex, race, ethnicity,
+    mutation1, mutation2, contains("insurance")
+  )
 fevs <- c("fev1_admit", "fev1_discharge", paste0("fev1_baseline", 1:6))
 df <- df %>%
   filter(redcap_event_name != "demographics_arm_1") %>%
@@ -36,6 +39,19 @@ demo$genotype <- factor(demo$genotype,
   labels = c("No F508del", "F508del heterozygous", "F508del homozygous")
 )
 demo[, c("mutation1", "mutation2")] <- NULL
+# Combine insurance columns
+ins <- c(
+  "Private insurance", "Medicare", "Medicaid",
+  "State special needs program (e.g. CHP+)",
+  "Tricare/other military health plan", "Indian Health Service", "Other", "None"
+)
+demo$insurance <- apply(demo[, grep("insurance", colnames(demo))], 1, function(r) {
+  i <- which(r == 1)
+  if(all(r == 0)){i = length(ins)}
+  return(paste0(ins[i], collapse = "/"))
+})
+demo$insurance <- as.factor(demo$insurance)
+demo[, paste0("insurance___", 1:8)] <- NULL
 # Merge everything back together and sort columns
 df <- left_join(df, demo, join_by(sid))
 df <- df %>% select(all_of(colnames(demo)), admit_year, everything())
@@ -96,7 +112,7 @@ summary_table <- df %>%
   group_by(sid, redcap_data_access_group, admit_year) %>%
   summarise(
     num_hosp = n(),
-    across(c(sex:genotype), ~ first(na.omit(.x))),
+    across(c(sex:insurance), ~ first(na.omit(.x))),
     across(admit_cx_results___1:admit_cx_results___9, ~ any(.x == "Present")),
     cftr_mod = any(cftr_mod == "Yes"),
     cftr_mod_spec = paste0(sort(unique(cftr_mod_spec)), collapse = "/"),
@@ -138,9 +154,11 @@ labels <- list(
[…]

### 2023-10-25 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
smoothed plot by age
eligible only

@@
 # source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R")
-load("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
+load("~/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
 ```
@@
 
+#### All patients
+
 ```{r}
@@
   select(eti_elig, contains("Vx"), ReviewYear) %>%
+  tbl_summary(
+    by = ReviewYear, missing_text = "Missing",
+    type = all_dichotomous() ~ "categorical"
+  ) %>%
+  bold_labels()
+```
+
+#### Eligible only
+
+```{r}
+annual %>%
+  filter(eti_elig == "Yes") %>%
+  select(contains("Vx"), ReviewYear) %>%
   tbl_summary(
@@
   cat("\n")
+  cat("\n")
+  cat("#### All patients")
+  cat("\n")
   d <- years[[n]]
@@
     select(eti_elig, contains("Vx"), race) %>%
+    tbl_summary(
+      by = race, missing_text = "Missing",
+      type = all_dichotomous() ~ "categorical"
+    ) %>%
+    bold_labels()
+  print(t)
+  cat("\n")
+  cat("\n")
+  cat("#### Eligible only")
+  cat("\n")
+  t <- d %>%
+    filter(eti_elig == "Yes") %>%
+    select(contains("Vx"), race) %>%
     tbl_summary(
@@
 ```
+
+```{r}
+ggplot(annual, aes(x = Age_YrEnd, y = mean_ppFEV)) +
+  geom_smooth(aes(color = race), se = F)
+```

### 2023-10-25 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R

@@ -483,10 +483,32 @@ by_year <- encounter %>%
     bmi_last = last(na.omit(bmivalue)),
     bmi_perc_last = last(na.omit(bmipercentile)),
     Vx445comb = case_when("Yes" %in% Vx445comb ~ "Yes", .default = "No"),
-    eti_elig = case_when("Yes" %in% eti_elig ~ "Yes", .default = "No"),
     .groups = "drop"
   )
 annual <- full_join(annual, by_year, by = join_by(eDWID, ReviewYear == reviewyear))
+# ETI eligibility by year
+annual$eti_elig <- apply(annual, 1, function(r) {
+  # Format variables
+  age <- as.numeric(r["Age_YrEnd"])
+  date <- r["ReviewYear"]
+  # Start with not eligible by default
+  elig <- "No"
+  # Get mutations
+  muts <- r[grep("Mutation", names(r))]
+  # Start with first approval
+  if (date >= year(eti_12_up) & age >= 12 & "F508del" %in% muts) {
+    elig <- "Yes"
+  }
+  # Next, eligible if over 12 and mutation in expanded list
+  else if (date >= year(eti_12_up_expansion) & age >= 12 & any(muts %in% responsive_mutations)) {
+    elig <- "Yes"
+  }
+  # Next approved for 6 and up with F508del or other responsive mutation
+  else if (date >= year(eti_6_up) & age >= 6 & any(muts %in% responsive_mutations)) {
+    elig <- "Yes"
+  }
+  return(elig)
+}, simplify = T)
 # Labels
 label(annual) <- labels[colnames(annual)]
 label(demo) <- labels[colnames(demo)]

### 2023-10-29 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
Update racial_and_ethnic_differences_in_eti.qmd

@@
     toc-float: true
+    number-sections: true
     code-fold: true
@@
 
-#### Eligible only
+#### Eligible and prescribed off-label only
 
@@
 annual %>%
-  filter(eti_elig == "Yes") %>%
-  select(contains("Vx"), ReviewYear) %>%
+  filter(eti_elig == "Yes" | Vx445comb == "Yes") %>%
+  select(eti_elig, contains("Vx"), ReviewYear) %>%
   tbl_summary(
@@
   ) %>%
+  add_ci() %>%
+  add_p(test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)) %>%
+  separate_p_footnotes() %>%
   bold_labels()
@@
     ) %>%
+    add_p(test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)) %>%
+    separate_p_footnotes() %>%
     bold_labels()
@@
   cat("\n")
-  cat("#### Eligible only")
+  cat("#### Eligible and prescribed off-label only")
   cat("\n")
   t <- d %>%
-    filter(eti_elig == "Yes") %>%
-    select(contains("Vx"), race) %>%
+    filter(eti_elig == "Yes" | Vx445comb == "Yes") %>%
+    select(eti_elig, contains("Vx"), race) %>%
     tbl_summary(
@@
     ) %>%
+    add_ci() %>%
+    add_p(test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)) %>%
+    separate_p_footnotes() %>%
     bold_labels()
@@
 }))
+```
+
+### Figure
+
+The figure below includes those eligible for ETI and anyone who may have been prescribed ETI off label. Error bars represent 95% confidence intervals.
+
+```{r}
+annual %>%
+  filter(eti_elig == "Yes" | Vx445comb == "Yes") %>%
+  select(eDWID, race, ReviewYear, contains("Vx")) %>%
+  group_by(ReviewYear, race) %>%
+  summarise(
+    perc = mean(Vx445comb == "Yes", na.rm = T),
+    se = sqrt(perc * (1 - perc) / n()),
+    lower = perc - (1.96 * se), upper = perc + (1.96 * se),
+    .groups = "drop"
+  ) %>%
+  ggplot(aes(x = ReviewYear, y = perc * 100, fill = race)) +
+  geom_bar(position = "dodge", stat = "identity") +
+  geom_errorbar(aes(ymin = lower * 100, ymax = upper * 100),
+    position = "dodge", linewidth = 0.7
+  ) +
+  xlab("Year") +
+  ylab("Percentage Prescribed ETI") +
+  theme_bw() +
+  theme(legend.title = element_blank()) +
+  scale_fill_manual(values = c(
+    "#264653", "#2A9D8F", "#E9C46A",
+    "#F4A261", "#E76F51"
+  ))
 ```

### 2023-10-29 Jordana Hoppe/nacfc_2023.qmd
Update nacfc_2023.qmd

--- 2023-04-10 Jordana Hoppe/nacfc_2023.qmd
@@
 library(gtsummary)
-df <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/clinical.csv", na.strings = "")
-surveys <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/surveys.csv", na.strings = "")
+df <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/clinical_2023-06-20.csv", na.strings = "")
+surveys <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/surveys_2023-06-20.csv", na.strings = "")
 # Combined genotype
@@
 # Shorten organism names
-bugs = colnames(df)[grep("If.culture.was.positive..select.all.organisms.isolated...choice.",colnames(df))]
+bugs <- colnames(df)[grep("If.culture.was.positive..select.all.organisms.isolated...choice.", colnames(df))]
 bugs <- sub("If.culture.was.positive..select.all.organisms.isolated...choice.", "", bugs)
-bugs = bugs[1:16]
+bugs <- bugs[1:16]
 colnames(df) <- sub("If.culture.was.positive..select.all.organisms.isolated...choice.", "", colnames(df))
-bug_names=c("Methicillin sensitive Staph aureus", 
-            "Methicillin resistant Staph aureus","A. xylosoxidans", 
-            "Achromobacter (not xylosoxidans)","Acinetobacter","H. influenzae", 
+bug_names <- c(
+  "Methicillin sensitive Staph aureus",
+  "Methicillin resistant Staph aureus", "A. xylosoxidans",
+  "Achromobacter (not xylosoxidans)", "Acinetobacter", "H. influenzae",
   "P. aeruginosa", "P. aeruginosa (mucoid)",
@@
   "Stenotrophomonas", "Aspergillus", "Penicillium species",
-            "Sceosporium species", "M. avium complex", "M. abscessus" )
+  "Sceosporium species", "M. avium complex", "M. abscessus"
+)
 ```
@@
 ```{r}
-fleiss = function(bug){
-  d = df[,c(bug,paste0(bug,".1"),paste0(bug,".2"))]
-  colnames(d) = c("Clinic","Delayed","Home")
-  d[d=="Checked"] = "Present"
-  d[d=="Unchecked"] = "Absent"
-  fk = kappam.fleiss(d)
+fleiss <- function(bug) {
+  d <- df[, c(bug, paste0(bug, ".1"), paste0(bug, ".2"))]
+  colnames(d) <- c("Clinic", "Delayed", "Home")
+  d[d == "Checked"] <- "Present"
+  d[d == "Unchecked"] <- "Absent"
+  fk <- kappam.fleiss(d)
   cat("\n")
-    cat("##", bug_names[which(bugs==bug)])
+  cat("##", bug_names[which(bugs == bug)])
   cat("\n")
@@
   cat("\n")
-    cat("Overall Fleiss' kappa for the three tests was", round(fk$value, 3), paste0("(p = ", format.pval(fk$p.value,digits = 3, eps = 0.001),")"))
+  cat("Overall Fleiss' kappa for the three tests was", round(fk$value, 3), paste0("(p = ", format.pval(fk$p.value, digits = 3, eps = 0.001), ")"))
   cat("\n")
@@
 ```{r message=FALSE}
-f = df %>% 
-  select(Subject.ID,contains("Methicillin.sensitive"),contains("H..influenzae."))
-colnames(f) = c("ID",paste("MSSA",c("Clinic","Delayed","Home")),
-                paste("H. influenzae",c("Clinic","Delayed","Home")))
+f <- df %>%
+  select(Subject.ID, contains("Methicillin.sensitive"), contains("H..influenzae."))
+colnames(f) <- c(
+  "ID", paste("MSSA", c("Clinic", "Delayed", "Home")),
+  paste("H. influenzae", c("Clinic", "Delayed", "Home"))
+)
 
-counts=f %>% 
+counts <- f %>%
   pivot_longer(`MSSA Clinic`:`H. influenzae Home`,
-                       names_to = c("Bug","Test"),
-                       names_pattern = "(.*) (.*)") %>%
-  filter(value=="Checked") %>%
-  group_by(Bug,ID) %>%
-  summarise(Test=paste0(Test,collapse = "+")) %>%
+    names_to = c("Bug", "Test"),
+    names_pattern = "(.*) (.*)"
+  ) %>%
+  filter(value == "Checked") %>%
+  group_by(Bug, ID) %>%
+  summarise(Test = paste0(Test, collapse = "+")) %>%
   count(Test) %>%
@@
 
-ggplot(counts,aes(x=Bug,y=n,fill=Test,label=n))+
+ggplot(counts, aes(x = Bug, y = n, fill = Test, label = n)) +
   geom_col(colour = "black", position = "fill") +
-  geom_text(aes(label = n), position = position_fill(vjust = 0.5)) +
+  geom_text(aes(label = n), position = position_fill(vjust = 0.5),size=5) +
   scale_y_continuous(labels = scales::percent) +
-  scale_fill_brewer(palette = "Pastel1")+
-  xlab("Pathogen")+ylab("% of Samples")
+  scale_fill_brewer(palette = "Pastel1") +
+  xlab("Pathogen") +
+  ylab("% of Samples") +
+  theme_bw() +
+  theme(text = element_text(size = 20))
 ```
@@
 # negative for all at home
-f = df %>% select(Subject.ID,contains (bugs)) %>% 
-  mutate(across(-Subject.ID,
-                ~as.numeric(factor(.x,levels=c("Unchecked","Checked")))-1)) %>%
+f <- df %>%
+  select(Subject.ID, contains(bugs)) %>%
+  mutate(across(
+    -Subject.ID,
+    ~ as.numeric(factor(.x, levels = c("Unchecked", "Checked"))) - 1
+  )) %>%
   rowwise() %>%
-  mutate(`Clinic % Positive` = mean(c_across(all_of(bugs)))*100,
-         `Delayed % Positive` = mean(c_across(all_of(paste0(bugs,".1"))))*100,
-         `Home % Positive` = mean(c_across(all_of(paste0(bugs,".2"))))*100) %>%
-  select(Subject.ID,`Clinic % Positive`:`Home % Positive`) %>%
-  filter(`Clinic % Positive`==0)
+  mutate(
+    `Clinic % Positive` = mean(c_across(all_of(bugs))) * 100,
+    `Delayed % Positive` = mean(c_across(all_of(paste0(bugs, ".1")))) * 100,
+    `Home % Positive` = mean(c_across(all_of(paste0(bugs, ".2")))) * 100
+  ) %>%
+  select(Subject.ID, `Clinic % Positive`:`Home % Positive`) %>%
+  filter(`Clinic % Positive` == 0)
 knitr::kable(f)
@@
 ```{r}
-f = f %>%
-  mutate(across(c("Clinic % Positive","Delayed % Positive","Home % Positive"),~.x==0))
-kappam.fleiss(f[,2:4])
+f <- f %>%
+  mutate(across(c("Clinic % Positive", "Delayed % Positive", "Home % Positive"), ~ .x == 0))
+kappam.fleiss(f[, 2:4])
 ```

### 2023-11-03 Christine Chan/One-hour OGTT/tim_1_hour_ogtt_analysis.qmd
CF update

---
title: "1 hour OGTT glucose in CFRD"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
editor: source
---

```{r libraries}
#| include: false
library(tidyverse)
```

```{r data cleaning}

```

# Table

```{r}
#| label: tbl-planets
#| tbl-cap: Planets

```

# Figure

```{r}
#| label: fig-planets
#| fig-cap: Planets

```

### 2023-11-03 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
CF update

@@
 # source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R")
-load("~/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
+load("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
 ```

### 2023-11-03 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R

@@ -4,11 +4,11 @@ library(readxl)
 library(rspiro)
 library(Hmisc)
 # Import
-annual <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_Annualized_Del1.csv", na.strings = "")
-encounter <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_encountersMerged_Del1.csv", na.strings = "")
-demo <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_DemogCFDiag_Del1.csv", na.strings = "")
+annual <- read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_Annualized_Del1.csv", na.strings = "")
+encounter <- read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_encountersMerged_Del1.csv", na.strings = "")
+demo <- read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_DemogCFDiag_Del1.csv", na.strings = "")
 # Import responsive mutations list
-responsive_mutations <- read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/eti_responsive_mutations.csv", header = F)
+responsive_mutations <- read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/eti_responsive_mutations.csv", header = F)
 responsive_mutations <- unique(responsive_mutations$V1)
 # Dates
[…]

### 2023-11-29 Christine Chan/CF Registry/cf_registry_analysis.qmd
Update cf_registry_analysis.qmd
started survival analysis

--- 2023-10-03 Christine Chan/CF Registry/cf_registry_analysis.qmd
@@
 # Import data and codebook
-annual <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_Annualized_Del1.csv",
+annual <- read.csv("~/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_Annualized_Del1.csv",
   na.strings = c("", "N/A")
 )
-data_dict <- readxl::read_excel("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
+data_dict <- readxl::read_excel("~/Documents/Work/Christine Chan/CF registry data/Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
 data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
@@
 # Get mutations
-demogs <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_DemogCFDiag_Del1.csv")
+demogs <- read.csv("~/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_DemogCFDiag_Del1.csv")
 demogs <- demogs %>% select(eDWID, Gender, MutClass)
@@
 annual$A_IsOnEnzymes <- relevel(annual$A_IsOnEnzymes, ref = "Yes")
-# Analysis function
-log_reg <- function(year, missingness = F) {
-  # Clean data
-  df <- annual %>%
-    filter(ReviewYear == year, Age_YrEnd >= 10) %>%
-    select(
-      eDWID, Age_YrEnd, Gender, MutClass, A_IsOnEnzymes, A_FEV1, A_FEV1_pct_predicted,
-      A_bmivalue, A_bmipercentile, A_height, A_weight, A_heightpercentile,
-      A_supplement_feeding, tube_feeding,
-      NumPulmExacerbation, A_pseudomonasaeruginosa,
-      A_fungalyeast1, A_pulmonarycomplications1, A_mycocultureresults3,
-      A_corticosteroids1, cfrd_status_annual, A_Vx770, A_VX809comb,
-      liver_disease
-    )
-  # Remove all empty columns
-  empty <- which(colSums(is.na(df)) == nrow(df))
-  if (length(empty) > 0) {
-    df <- df[, -empty]
-  }
-  # Models
-  vars <- colnames(df)[which(!colnames(df) %in%
-    c(
-      bugs, "cfrd_status_annual", "eDWID", "A_bmivalue",
-      "A_bmipercentile", "A_height", "A_heightpercentile",
-      "Age_YrEnd", "A_weight", "bmi", "age_perc",
-      "A_supplement_feeding", "A_feeding2", "A_feeding3",
-      "A_feeding4", "A_FEV1", liver_fields
-    ))]
-  base_mod <- as.formula(paste0("cfrd_status_annual~", paste0(vars, collapse = "+")))
-  # Labels
-  labels <- list(
-    cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
-    Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
-    A_FEV1_pct_predicted = "FEV1 percent predicted",
-    tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
-    A_corticosteroids1 = "Use of corticosteroids",
-    NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
-    height_perc = "Height percentile", bmi_perc = "BMI percentile",
-    A_Vx770 = "Use of CFTR modulator Vx770",
-    A_VX809comb = "Use of CFTR modulator Vx809 combination",
-    A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
-    A_fungalyeast1 = "Aspergillus (any species)",
-    A_pulmonarycomplications1 = "ABPA",
-    A_mycocultureresults3 = "Microorganisms in myco culture"
-  )
-  var_label(df[, vars[which(vars %in% colnames(df))]]) <-
-    as.character(labels[vars[which(vars %in% colnames(df))]])
-  var_label(df[, bugs[which(bugs %in% colnames(df))]]) <-
-    as.character(labels[bugs[which(bugs %in% colnames(df))]])
-  # Visualize missingness if required
-  if (missingness) {
-    cat("\n")
-    cat("### Missingness")
-    cat("\n")
-    cat("\n")
-    naniar::vis_miss(df)
-    cat("\n")
-  }
-  # Simple models
-  mod1 <- glm(base_mod, data = df, family = "binomial")
-  # Print results
-  cat("\n")
-  cat("### Base Model")
-  cat("\n")
-  cat("\n")
-  print(forest_model(mod1, exclude_infinite_cis = F))
-  cat("\n")
-  if ("A_pseudomonasaeruginosa" %in% colnames(df)) {
-    mod2 <- glm(update(base_mod, . ~ . + A_pseudomonasaeruginosa), data = df, family = "binomial")
-    cat("\n")
-    cat("### P. aeruginosa")
-    cat("\n")
-    cat("\n")
-    print(forest_model(mod2, exclude_infinite_cis = F))
-    cat("\n")
-  }
-  if ("A_fungalyeast1" %in% colnames(df)) {
-    mod3 <- glm(update(base_mod, . ~ . + A_fungalyeast1), data = df, family = "binomial")
-    cat("\n")
-    cat("### Aspergillus")
-    cat("\n")
-    cat("\n")
-    print(forest_model(mod3, exclude_infinite_cis = F))
-    cat("\n")
-  }
-  if ("A_pulmonarycomplications1" %in% colnames(df)) {
-    mod4 <- glm(update(base_mod, . ~ . + A_pulmonarycomplications1), data = df, family = "binomial")
-    cat("\n")
-    cat("### ABPA")
-    cat("\n")
-    cat("\n")
-    print(forest_model(mod4, exclude_infinite_cis = F))
-    cat("\n")
-  }
-  if ("A_mycocultureresults3" %in% colnames(df)) {
-    mod5 <- glm(update(base_mod, . ~ . + A_mycocultureresults3), data = df, family = "binomial")
-    cat("\n")
-    cat("### Mycobacterium")
-    cat("\n")
-    cat("\n")
-    print(forest_model(mod5, exclude_infinite_cis = F))
-    cat("\n")
-  }
-}
 ```
@@
 
+Why are the percentages so low in 2006 - 2009? These appear to be correct based on the data we have, so could it be to do with how the data were collected or pulled by the CFFPR?
+
+# CFRD incidence and prevalence
+
+Based on CFFPR annualized data.
+
+```{r}
+# Find approximate birth date
+annual$approx_birth_year = annual$ReviewYear - ceiling(annual$Age_YrEnd)
+annual = annual %>% group_by(eDWID) %>% 
+  mutate(approx_birth_year = round(mean(approx_birth_year)))
+# Create survival dataset
+s <- annual %>%
+  rename(end = ReviewYear) %>% # ReviewYear is the end of the interval
+  group_by(eDWID) %>%
+  mutate(start = lag(end)) %>% # Get interval start
+  select(eDWID, start,end, Age_YrEnd, approx_birth_year,cfrd_status_annual) # Sort columns
+s$start[is.na(s$start)] = s$approx_birth_year[is.na(s$start)]
+# Age group as time-varying covariate
+s$age_group <- cut(s$Age_YrEnd,
+  breaks = c(-Inf, 10, 19, 31, Inf), right = F,
+  labels = c("< 10", "[10,19)", "[19,31)", ">= 31")
+)
+# Fill down diagnosis
+s$cfrd_status_annual[s$cfrd_status_annual=="CFRD-"] = NA
+s = s %>% group_by(eDWID) %>% fill(cfrd_status_annual,.direction = "down")
+s$cfrd_status_annual[is.na(s$cfrd_status_annual)] = "CFRD-"
+# Remove records where start == end
+s = s[s$end > s$start,]
+# Survival fit
+surv_obj = Surv(time = as.numeric(s$start),time2 = as.numeric(s$end),
+                event = as.numeric(s$cfrd_status_annual)-1)
+fit <- survfit(surv_obj ~ age_group,id = eDWID,data = s)
+ggsurvplot(fit,risk.table = T,cumevents = T,
+  cumcensor = T,
+           fun = "event",xlim = c(2003, 2018))
+```
+
 # Logistic analyses
@@
 
+```{r}
+# Analysis function
+log_reg <- function(year, missingness = T) {
+  # Clean data
+  df <- annual %>%
+    filter(ReviewYear == year, Age_YrEnd >= 10) %>%
+    select(
+      eDWID, age_group, Gender, bmi_perc, MutClass, A_IsOnEnzymes, A_FEV1_pct_predicted,
+      tube_feeding, NumPulmExacerbation, A_pseudomonasaeruginosa,
+      A_fungalyeast1, A_pulmonarycomplications1, A_mycocultureresults3,
+      A_corticosteroids1, cfrd_status_annual, A_Vx770, A_VX809comb,
+      liver_disease
+    )
+  # Remove all empty columns
+  empty <- which(colSums(is.na(df)) == nrow(df))
+  if (length(empty) > 0) {
+    df <- df[, -empty]
+  }
+  # Models
+  vars <- colnames(df)[which(!colnames(df) %in%
+    c(bugs, liver_fields, "cfrd_status_annual", "eDWID"))]
+  base_mod <- as.formula(paste0("cfrd_status_annual~", paste0(vars, collapse = "+")))
+  # Labels
+  labels <- list(
+    cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
+    Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
+    A_FEV1_pct_predicted = "FEV1 percent predicted",
+    tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
+    A_corticosteroids1 = "Use of corticosteroids",
+    NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
+    height_perc = "Height percentile", bmi_perc = "BMI percentile",
+    A_Vx770 = "Use of CFTR modulator Vx770",
+    A_VX809comb = "Use of CFTR modulator Vx809 combination",
+    A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
+    A_fungalyeast1 = "Aspergillus (any species)",
+    A_pulmonarycomplications1 = "ABPA",
+    A_mycocultureresults3 = "Microorganisms in myco culture"
+  )
+  var_label(df[, vars[which(vars %in% colnames(df))]]) <-
+    as.character(labels[vars[which(vars %in% colnames(df))]])
+  var_label(df[, bugs[which(bugs %in% colnames(df))]]) <-
+    as.character(labels[bugs[which(bugs %in% colnames(df))]])
+  # Visualize missingness if required
+  if (missingness) {
+    cat("\n")
+    cat("### Missingness")
+    cat("\n")
+    cat("\n")
+    print(naniar::vis_miss(df))
+    cat("\n")
+  }
+  # Simple models
+  mod1 <- glm(base_mod, data = df, family = "binomial")
+  # Print results
+  cat("\n")
+  cat("### Base Model")
+  cat("\n")
+  cat("\n")
+  print(forest_model(mod1, exclude_infinite_cis = F))
+  cat("\n")
+  if ("A_pseudomonasaeruginosa" %in% colnames(df)) {
+    mod2 <- glm(update(base_mod, . ~ . + A_pseudomonasaeruginosa), data = df, family = "binomial")
+    cat("\n")
+    cat("### P. aeruginosa")
+    cat("\n")
+    cat("\n")
+    print(forest_model(mod2, exclude_infinite_cis = F))
+    cat("\n")
+  }
+  if ("A_fungalyeast1" %in% colnames(df)) {
+    mod3 <- glm(update(base_mod, . ~ . + A_fungalyeast1), data = df, family = "binomial")
+    cat("\n")
+    cat("### Aspergillus")
+    cat("\n")
+    cat("\n")
+    print(forest_model(mod3, exclude_infinite_cis = F))
+    cat("\n")
+  }
+  if ("A_pulmonarycomplications1" %in% colnames(df)) {
+    mod4 <- glm(update(base_mod, . ~ . + A_pulmonarycomplications1), data = df, family = "binomial")
+    cat("\n")
+    cat("### ABPA")
+    cat("\n")
+    cat("\n")
+    print(forest_model(mod4, exclude_infinite_cis = F))
+    cat("\n")
+  }
+  if ("A_mycocultureresults3" %in% colnames(df)) {
+    mod5 <- glm(update(base_mod, . ~ . + A_mycocultureresults3), data = df, family = "binomial")
+    cat("\n")
+    cat("### Mycobacterium")
+    cat("\n")
+    cat("\n")
+    print(forest_model(mod5, exclude_infinite_cis = F))
+    cat("\n")
+  }
+}
+```
+
 ## 2004
@@
 ```{r results='asis',fig.height=10,fig.width=10}
-log_reg(2004,missingness = T)
+log_reg(2004)
 ```

### 2023-11-30 Christine Chan/CF Registry/cf_registry_analysis.qmd
CF registry survival analysis

@@
 # Find approximate birth date
-annual$approx_birth_year = annual$ReviewYear - ceiling(annual$Age_YrEnd)
-annual = annual %>% group_by(eDWID) %>% 
+annual$approx_birth_year <- annual$ReviewYear - ceiling(annual$Age_YrEnd)
+annual <- annual %>%
+  group_by(eDWID) %>%
   mutate(approx_birth_year = round(mean(approx_birth_year)))
@@
   mutate(start = lag(end)) %>% # Get interval start
-  select(eDWID, start,end, Age_YrEnd, approx_birth_year,cfrd_status_annual) # Sort columns
-s$start[is.na(s$start)] = s$approx_birth_year[is.na(s$start)]
+  select(eDWID, start, end, Age_YrEnd, approx_birth_year, cfrd_status_annual) # Sort columns
+s$start[is.na(s$start)] <- s$approx_birth_year[is.na(s$start)]
 # Age group as time-varying covariate
@@
 # Fill down diagnosis
-s$cfrd_status_annual[s$cfrd_status_annual=="CFRD-"] = NA
-s = s %>% group_by(eDWID) %>% fill(cfrd_status_annual,.direction = "down")
-s$cfrd_status_annual[is.na(s$cfrd_status_annual)] = "CFRD-"
+s$cfrd_status_annual[s$cfrd_status_annual == "CFRD-"] <- NA
+s <- s %>%
+  group_by(eDWID) %>%
+  fill(cfrd_status_annual, .direction = "down")
+s$cfrd_status_annual[is.na(s$cfrd_status_annual)] <- "CFRD-"
 # Remove records where start == end
-s = s[s$end > s$start,]
+s <- s[s$end > s$start, ]
 # Survival fit
-surv_obj = Surv(time = as.numeric(s$start),time2 = as.numeric(s$end),
-                event = as.numeric(s$cfrd_status_annual)-1)
-fit <- survfit(surv_obj ~ age_group,id = eDWID,data = s)
-ggsurvplot(fit,risk.table = T,cumevents = T,
-  cumcensor = T,
-           fun = "event",xlim = c(2003, 2018))
+surv_obj <- Surv(
+  time = as.numeric(s$start), time2 = as.numeric(s$end),
+  event = as.numeric(s$cfrd_status_annual) - 1
+)
+fit <- survfit(surv_obj ~ age_group, id = eDWID, data = s)
+ggsurvplot(fit,
+  risk.table = T, cumevents = T,
+  fun = "event", xlim = c(2002, 2018)
+)
 ```

### 2023-11-30 Christine Chan/CF Registry/cf_registry_analysis.qmd
fixed really stupid error that was removing a bunch of NAs

--- 2023-11-29 Christine Chan/CF Registry/cf_registry_analysis.qmd
@@
 # Remove those with T1D or T2D
-annual <- annual %>% filter(cfrd_status_annual != 4, cfrd_status_annual != 5)
+annual <- annual %>% arrange(eDWID, ReviewYear)
+annual = annual[!annual$cfrd_status_annual %in% c(4,5),]
 annual$cfrd_status_annual <- factor(annual$cfrd_status_annual,
@@
 annual$bmi_perc <- relevel(annual$bmi_perc, ref = "(50,85]")
-annual$age_group <- cut(annual$Age_YrEnd, breaks = c(10, 18, 30, Inf), right = F)
+annual$age_group <- cut(annual$Age_YrEnd, breaks = c(-Inf,10, 18, 30, Inf), right = F)
 annual$age_group <- relevel(annual$age_group, ref = "[10,18)")
@@
 
-Based on CFFPR annualized data.
+## Overall
 
 ```{r}
-# Find approximate birth date
-annual$approx_birth_year = annual$ReviewYear - ceiling(annual$Age_YrEnd)
-annual = annual %>% group_by(eDWID) %>% 
-  mutate(approx_birth_year = round(mean(approx_birth_year)))
-# Create survival dataset
-s <- annual %>%
-  rename(end = ReviewYear) %>% # ReviewYear is the end of the interval
+# Find the year that each person was diagnosed, make wide
+first_cfrd <- annual %>%
+  filter(cfrd_status_annual == "CFRD+") %>%
   group_by(eDWID) %>%
-  mutate(start = lag(end)) %>% # Get interval start
-  select(eDWID, start,end, Age_YrEnd, approx_birth_year,cfrd_status_annual) # Sort columns
-s$start[is.na(s$start)] = s$approx_birth_year[is.na(s$start)]
-# Age group as time-varying covariate
-s$age_group <- cut(s$Age_YrEnd,
-  breaks = c(-Inf, 10, 19, 31, Inf), right = F,
-  labels = c("< 10", "[10,19)", "[19,31)", ">= 31")
+  slice_head(n = 1) %>%
+  select(eDWID, ReviewYear) %>%
+  mutate(v = 1) %>%
+  pivot_wider(
+    id_cols = eDWID, names_from = ReviewYear, values_from = v,
+    values_fill = 0
   )
-# Fill down diagnosis
-s$cfrd_status_annual[s$cfrd_status_annual=="CFRD-"] = NA
-s = s %>% group_by(eDWID) %>% fill(cfrd_status_annual,.direction = "down")
-s$cfrd_status_annual[is.na(s$cfrd_status_annual)] = "CFRD-"
-# Remove records where start == end
-s = s[s$end > s$start,]
-# Survival fit
-surv_obj = Surv(time = as.numeric(s$start),time2 = as.numeric(s$end),
-                event = as.numeric(s$cfrd_status_annual)-1)
-fit <- survfit(surv_obj ~ age_group,id = eDWID,data = s)
-ggsurvplot(fit,risk.table = T,cumevents = T,
-  cumcensor = T,
-           fun = "event",xlim = c(2003, 2018))
+# Make a dataframe for who was at risk each year
+at_risk <- annual %>%
+  select(eDWID, ReviewYear) %>%
+  mutate(v = 1) %>%
+  pivot_wider(
+    id_cols = eDWID, names_from = ReviewYear, values_from = v,
+    values_fill = 0
+  )
+# Add age information
+ages <- annual %>%
+  pivot_wider(
+    id_cols = eDWID, names_from = ReviewYear, values_from = age_group
+  )
 ```

### 2023-11-30 Christine Chan/CF Registry/cf_registry_analysis.qmd
incidence data cleaning mostly done
need to remove people from at risk after diagnosis

@@
 )
+annual$age_2018 = 2018 - (annual$ReviewYear - annual$Age_YrEnd)
+annual = annual %>% 
+  group_by(eDWID) %>%
+  mutate(age_2018 = as.numeric(names(sort(table(age_2018),decreasing = TRUE)[1])))
 # Get mutations
@@
   slice_head(n = 1) %>%
-  select(eDWID, ReviewYear) %>%
-  mutate(v = 1) %>%
-  pivot_wider(
-    id_cols = eDWID, names_from = ReviewYear, values_from = v,
-    values_fill = 0
-  )
+  select(eDWID,ReviewYear) %>% rename(first_cfrd=ReviewYear)
 # Make a dataframe for who was at risk each year
-at_risk <- annual %>%
-  select(eDWID, ReviewYear) %>%
-  mutate(v = 1) %>%
-  pivot_wider(
-    id_cols = eDWID, names_from = ReviewYear, values_from = v,
-    values_fill = 0
-  )
-# Add age information
-ages <- annual %>%
-  pivot_wider(
-    id_cols = eDWID, names_from = ReviewYear, values_from = age_group
-  )
+at_risk = expand.grid(eDWID = unique(annual$eDWID), ReviewYear=2003:2018)
+at_risk = at_risk %>% arrange(eDWID)
+at_risk = at_risk %>% 
+  full_join(.,annual %>% select(eDWID, ReviewYear,age_2018,A_hgba1c),
+            by = join_by(eDWID, ReviewYear))
+# Combine
+inc = left_join(at_risk,first_cfrd,by = join_by(eDWID))
+# Calculate all missing ages
+inc = inc %>% 
+  group_by(eDWID) %>%
+  mutate(age = age_2018[1] - (2018 - ReviewYear)) %>%
+  # No missing age in 2018 field, so this indicates whether they were at risk 
+  # (i.e. seen that year)
+  mutate(age_2018 = !is.na(age_2018)) %>%
+  rename(at_risk = age_2018)
+# Age group
+inc$age_group = cut(inc$age,c(-Inf,10,19,31,Inf),right=F)
+# Use table1
+table1(ReviewYear~)
 ```

### 2023-12-01 Christine Chan/CF Registry/cf_registry_analysis.qmd
CFRD registry analysis done

--- 2023-11-30 Christine Chan/CF Registry/cf_registry_analysis.qmd
@@
 library(gt)
+library(gtsummary)
+library(forcats)
 library(labelled)
@@
 annual <- annual %>% arrange(eDWID, ReviewYear)
-annual = annual[!annual$cfrd_status_annual %in% c(4,5),]
+annual <- annual[!annual$cfrd_status_annual %in% c(4, 5), ]
 annual$cfrd_status_annual <- factor(annual$cfrd_status_annual,
@@
 )
+annual$age_2018 <- 2018 - (annual$ReviewYear - annual$Age_YrEnd)
+annual <- annual %>%
+  group_by(eDWID) %>%
+  mutate(age_2018 = as.numeric(names(sort(table(age_2018),
+    decreasing = TRUE
+  )[1])))
 # Get mutations
@@
 )
-annual$A_FEV1_pct_predicted <- relevel(annual$A_FEV1_pct_predicted, ref = "[100,Inf)")
+annual$A_FEV1_pct_predicted <- relevel(annual$A_FEV1_pct_predicted,
+  ref = "[100,Inf)"
+)
 annual$height_perc <- cut(annual$height_perc, breaks = c(0, 10, 25, 50, Inf))
@@
 annual$bmi_perc <- relevel(annual$bmi_perc, ref = "(50,85]")
-annual$age_group <- cut(annual$Age_YrEnd, breaks = c(-Inf,10, 18, 30, Inf), right = F)
+annual$age_group <- cut(annual$Age_YrEnd,
+  breaks = c(-Inf, 10, 18, 30, Inf),
+  right = F
+)
 annual$age_group <- relevel(annual$age_group, ref = "[10,18)")
@@
 )
-annual[, yn_vars] <- lapply(annual[, yn_vars], factor, levels = 0:1, labels = c("No", "Yes"))
+annual[, yn_vars] <- lapply(annual[, yn_vars], factor,
+  levels = 0:1,
+  labels = c("No", "Yes")
+)
 annual$A_IsOnEnzymes <- relevel(annual$A_IsOnEnzymes, ref = "Yes")
@@
 ```{r}
-data.frame(data_dict$code_meaning[data_dict$var %in% sub("A_", "", liver_fields)]) %>%
+data.frame(data_dict$code_meaning[data_dict$var %in%
+  sub("A_", "", liver_fields)]) %>%
   gt() %>%
@@
 
-# CFRD validity check
-
-```{r}
-annual %>%
-  mutate(cfrd = as.numeric(cfrd_status_annual == "CFRD+")) %>%
-  group_by(eDWID) %>%
-  summarise(
-    `CFRD Status` = any(rle(cfrd)$lengths[rle(cfrd)$values == 1] >= 2),
-    .groups = "drop"
-  ) %>%
-  mutate(`CFRD Status` = factor(`CFRD Status`,
-    levels = c(F, T),
-    labels = c("CFRD-", "CFRD+")
-  )) %>%
-  count(`CFRD Status`) %>%
-  gt()
-```
-
-# HbA1c by year
-
-```{r}
-annual %>%
-  group_by(ReviewYear) %>%
-  summarise(
-    `Num. HbA1c Tests` = sum(!is.na(A_hgba1c)),
-    `Total n` = n(),
-    `Percent with HbA1c` = round(sum(!is.na(A_hgba1c)) / n() * 100, 2)
-  ) %>%
-  gt()
-```
-
-Why are the percentages so low in 2006 - 2009? These appear to be correct based on the data we have, so could it be to do with how the data were collected or pulled by the CFFPR?
-
 # CFRD incidence and prevalence
 
-## Overall
+## By year
 
-```{r}
+```{r results='asis'}
 # Find the year that each person was diagnosed, make wide
@@
   select(eDWID, ReviewYear) %>%
-  mutate(v = 1) %>%
-  pivot_wider(
-    id_cols = eDWID, names_from = ReviewYear, values_from = v,
-    values_fill = 0
+  rename(first_cfrd = ReviewYear)
+# Make a dataframe for who was at risk each year
+at_risk <- expand.grid(eDWID = unique(annual$eDWID), ReviewYear = 2003:2018)
+at_risk <- at_risk %>% arrange(eDWID)
+at_risk <- at_risk %>%
+  full_join(., annual %>%
+    select(eDWID, ReviewYear, age_2018, A_hgba1c, cfrd_status_annual),
+  by = join_by(eDWID, ReviewYear)
   )
-# Make a dataframe for who was at risk each year
-at_risk <- annual %>%
-  select(eDWID, ReviewYear) %>%
-  mutate(v = 1) %>%
-  pivot_wider(
-    id_cols = eDWID, names_from = ReviewYear, values_from = v,
-    values_fill = 0
-  )
-# Add age information
-ages <- annual %>%
-  pivot_wider(
-    id_cols = eDWID, names_from = ReviewYear, values_from = age_group
-  )
+# Combine
+inc <- left_join(at_risk, first_cfrd, by = join_by(eDWID))
+# Calculate all missing ages
+inc <- inc %>%
+  group_by(eDWID) %>%
+  mutate(age = first(na.omit(age_2018)) - (2018 - ReviewYear)) %>%
+  # No missing age in 2018 field, so this indicates whether they were at risk
+  # (i.e. seen that year)
+  mutate(age_2018 = !is.na(age_2018)) %>%
+  rename(at_risk = age_2018) %>%
+  ungroup()
+# Age group
+inc$age_group <- cut(inc$age, c(-Inf, 10, 19, 31, Inf), right = F)
+# Remove people from at risk after diagnosis
+inc$at_risk[!is.na(inc$first_cfrd) & inc$ReviewYear > inc$first_cfrd] <- FALSE
+# Use gtsummary to make a list of tables by year
+invisible(lapply(unique(inc$ReviewYear), function(y) {
+  d <- inc %>%
+    filter(ReviewYear == y, at_risk) %>%
+    # Make missing values explicit for the tables
+    mutate(
+      a1c = factor(!is.na(A_hgba1c), levels = c(F, T), labels = c("No", "Yes")),
+      cfrd_status_annual = fct_explicit_na(cfrd_status_annual),
+    ) %>%
+    select(age_group, cfrd_status_annual, a1c, A_hgba1c) %>%
+    tbl_summary(
+      by = age_group, missing_text = "(Missing)",
+      label = list(
+        cfrd_status_annual ~ "Annualized CFRD Status",
+        A_hgba1c ~ "Annualized HbA1c", a1c ~ "A1c Collected"
+      ),
+      statistic = list(all_continuous() ~ "{mean} ({sd})")
+    ) %>%
+    add_overall()
+  cat("\n")
+  cat("### ", y)
+  cat("\n")
+  print(d)
+  cat("\n")
+}))
 ```
@@
     select(
-      eDWID, age_group, Gender, bmi_perc, MutClass, A_IsOnEnzymes, A_FEV1_pct_predicted,
-      tube_feeding, NumPulmExacerbation, A_pseudomonasaeruginosa,
-      A_fungalyeast1, A_pulmonarycomplications1, A_mycocultureresults3,
+      eDWID, age_group, Gender, bmi_perc, MutClass, A_IsOnEnzymes,
+      A_FEV1_pct_predicted, tube_feeding, NumPulmExacerbation,
+      A_pseudomonasaeruginosa, A_fungalyeast1, A_mycocultureresults3,
+      A_pulmonarycomplications1,
       A_corticosteroids1, cfrd_status_annual, A_Vx770, A_VX809comb,
@@
     )
+  # Assume that missing are CFRD- for these models
+  df$cfrd_status_annual[is.na(df$cfrd_status_annual)] <- "CFRD-"
+  # Relevel age group since we're excluding kids < 10
+  df$age_group <- droplevels(df$age_group)
   # Remove all empty columns
@@
     c(bugs, liver_fields, "cfrd_status_annual", "eDWID"))]
-  base_mod <- as.formula(paste0("cfrd_status_annual~", paste0(vars, collapse = "+")))
+  base_mod <- as.formula(paste0(
+    "cfrd_status_annual~",
+    paste0(vars, collapse = "+")
+  ))
   # Labels
@@
   if ("A_pseudomonasaeruginosa" %in% colnames(df)) {
-    mod2 <- glm(update(base_mod, . ~ . + A_pseudomonasaeruginosa), data = df, family = "binomial")
+    mod2 <- glm(update(base_mod, . ~ . + A_pseudomonasaeruginosa),
+      data = df,
+      family = "binomial"
+    )
     cat("\n")
@@
   if ("A_fungalyeast1" %in% colnames(df)) {
-    mod3 <- glm(update(base_mod, . ~ . + A_fungalyeast1), data = df, family = "binomial")
+    mod3 <- glm(update(base_mod, . ~ . + A_fungalyeast1),
+      data = df,
+      family = "binomial"
+    )
     cat("\n")
@@
   if ("A_pulmonarycomplications1" %in% colnames(df)) {
-    mod4 <- glm(update(base_mod, . ~ . + A_pulmonarycomplications1), data = df, family = "binomial")
+    mod4 <- glm(update(base_mod, . ~ . + A_pulmonarycomplications1),
+      data = df,
+      family = "binomial"
+    )
     cat("\n")
@@
   if ("A_mycocultureresults3" %in% colnames(df)) {
-    mod5 <- glm(update(base_mod, . ~ . + A_mycocultureresults3), data = df, family = "binomial")
+    mod5 <- glm(update(base_mod, . ~ . + A_mycocultureresults3),
+      data = df,
+      family = "binomial"
+    )
     cat("\n")
@@
 
-## 2004
+## By year
 
 ```{r results='asis',fig.height=10,fig.width=10}
-log_reg(2004)
+invisible(lapply(unique(annual$ReviewYear), function(y) {
+  cat("\n")
+  cat("### ", y)
+  cat("\n")
+  log_reg(y)
+  cat("\n")
+}))
 ```
-
-## 2006
-
-```{r results='asis',fig.height=10,fig.width=10}
-log_reg(2006)
-```
-
-## 2013
-
-```{r results='asis',fig.height=10,fig.width=10}
-log_reg(2013)
-```
-
-## 2018
-
-```{r results='asis',fig.height=10,fig.width=10}
-log_reg(2018)
-```
-
-# Questions
-
-1. What variables do we use for B. Cepacia and liver disease?
-2. Do we have CFTR modulator data aside from `Vx770` and `Vx809comb`?

### 2023-12-01 Christine Chan/CF Registry/cf_registry_analysis.qmd
parallel registry (needs more RAM)

@@
 library(labelled)
+library(parallel)
 ```
@@
 ```{r results='asis',fig.height=10,fig.width=10}
-invisible(lapply(unique(annual$ReviewYear), function(y) {
+cl = makeCluster(16,type = "FORK")
+invisible(parLapply(cl,unique(annual$ReviewYear), function(y) {
   cat("\n")
@@
 }))
+stopCluster(cl)
 ```

### 2023-12-01 Christine Chan/CF Registry/cf_registry_analysis.qmd
output turned out to be a little tricky in parallel

@@
 # Analysis function
-log_reg <- function(year, missingness = T) {
+log_reg <- function(year, missingness = F) {
   # Clean data
@@
 
-## By year
-
 ```{r results='asis',fig.height=10,fig.width=10}
-cl = makeCluster(16,type = "FORK")
-invisible(parLapply(cl,unique(annual$ReviewYear), function(y) {
+invisible(lapply(unique(annual$ReviewYear), function(y) {
   cat("\n")
-  cat("### ", y)
+  cat("## ", y)
   cat("\n")
@@
 }))
-stopCluster(cl)
 ```

### 2023-12-06 Christine Chan/CF Registry/cf_registry_analysis.qmd
Update cf_registry_analysis.qmd

@@
 library(parallel)
+switch(Sys.info()[["sysname"]],
+  Windows = {
+    home_dir <- "OneDrive - The University of Colorado Denver"
+  },
+  Darwin = {
+    home_dir <- "~/"
+  }
+)
+knitr::opts_knit$set(root.dir = home_dir)
 ```

### 2023-12-07 Christine Chan/CF Registry/cf_registry_analysis.qmd
original incidence and prevalence back

--- 2023-12-01 Christine Chan/CF Registry/cf_registry_analysis.qmd
@@
 library(parallel)
+switch(Sys.info()[["sysname"]],
+  Windows = {
+    home_dir <- "C:/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data"
+  },
+  Darwin = {
+    home_dir <- "/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data"
+  }
+)
+knitr::opts_knit$set(root.dir = home_dir)
+```
+
+# CFRD incidence and prevalence
+
+```{r}
+# All data
+encounter <- read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv",
+                      na.strings = c("", "N/A")
+)
+annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
+                   na.strings = c("", "N/A")
+)
+encounter <- encounter %>%
+  arrange(eDWID, reviewyear, encounternum)
+annual = annual %>% arrange(eDWID,ReviewYear)
+# Add age
+ages <- annual %>% 
+  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd)) %>%
+  rename(reviewyear = ReviewYear) %>%
+  group_by(eDWID,reviewyear) %>%
+  summarise(`Age in 2018` = as.numeric(names(sort(table(`Age in 2018`), 
+                                                  decreasing = TRUE)[1])),
+            Age = Age_YrEnd[1],.groups = "drop")
+encounter = left_join(encounter,ages,by = c("eDWID", "reviewyear"))
+encounter$`Age Group 2018` = cut(encounter$`Age in 2018`,
+                            c(-Inf,10,19,31,Inf),
+                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
+                            right=F)
+encounter$`Age Group` = cut(encounter$Age,
+                            c(-Inf,10,19,31,Inf),
+                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
+                            right=F)
+# For those with CFRD by OGTT, find first visit with diagnosis
+first_cfrd_ogtt <- encounter %>% 
+  filter(twohour_bloodglucose >= 200) %>%
+  group_by(eDWID) %>%
+  filter(row_number()==1) %>%
+  select(eDWID, reviewyear, `Age in 2018`) %>%
+  rename(first_cfrd_ogtt = reviewyear)
+# For those with CFRD by registry criteria, find first visit with diagnosis
+first_cfrd_cffpr <- encounter %>% 
+  filter(cfrd_status==3) %>%
+  group_by(eDWID) %>%
+  filter(row_number()==1) %>%
+  select(eDWID, reviewyear, `Age in 2018`) %>%
+  rename(first_cfrd_cffpr = reviewyear)
+# Both
+first_cfrd_both <- encounter %>% 
+  filter(cfrd_status==3|twohour_bloodglucose >= 200) %>%
+  group_by(eDWID) %>%
+  filter(row_number()==1) %>%
+  select(eDWID, reviewyear, `Age in 2018`) %>%
+  rename(first_cfrd_both = reviewyear)
+# Add to encounter level data
+encounter <- left_join(encounter, first_cfrd_ogtt, by = c("eDWID", "Age in 2018"))
+encounter <- left_join(encounter, first_cfrd_cffpr, by = c("eDWID", "Age in 2018"))
+encounter <- left_join(encounter, first_cfrd_both,by = c("eDWID", "Age in 2018"))
+# Fill down CFRD Yes/No by year
+encounter <- encounter %>%
+  group_by(eDWID) %>%
+  mutate(
+    cfrd_ogtt = case_when(
+      reviewyear < first_cfrd_ogtt ~ "No",
+      reviewyear >= first_cfrd_ogtt ~ "Yes"
+    ),
+    cfrd_cffpr = case_when(
+      reviewyear < first_cfrd_cffpr ~ "No",
+      reviewyear >= first_cfrd_cffpr ~ "Yes"
+    ),
+    cfrd_both = case_when(
+      reviewyear < first_cfrd_both ~ "No",
+      reviewyear >= first_cfrd_both ~ "Yes"
+    )
+  ) %>%
+  ungroup()
+encounter$cfrd_ogtt[is.na(encounter$cfrd_ogtt)] <- "No"
+encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
+encounter$cfrd_both[is.na(encounter$cfrd_both)] <- "No"
+# Ordered factors
+encounter$cfrd_ogtt = factor(encounter$cfrd_ogtt,levels = c("No","Yes"),ordered = T)
+encounter$cfrd_cffpr = factor(encounter$cfrd_cffpr,levels = c("No","Yes"),ordered = T)
+encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
+# Remove those missing age
+encounter = encounter %>% filter(!is.na(Age))
 ```
@@
 ```{r}
+# For each year, pull everyone with and without CFRD that year
+year_inc = function(df=encounter,year,py = 1000){
+  # Filter and count
+  inc = df %>%  
+    filter(reviewyear==year,is.na(first_cfrd_cffpr) | first_cfrd_cffpr >= year) %>%
+    group_by(eDWID) %>% filter(row_number() == 1) %>% 
+    select(eDWID,`Age Group`,cfrd_cffpr)
+  inc = inc %>% group_by(cfrd_cffpr) %>% count(`Age Group`,.drop = F)
+  inc = inc[!is.na(inc$`Age Group`),]
+  # Calculations
+  new_cases = sum(inc$n[inc$cfrd_cffpr=="Yes"])
+  total_at_risk = sum(inc$n)
+  total_inc = new_cases/(total_at_risk/py)
+
+  new_cases_10 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group` == "< 10"])
+  at_risk_10 = sum(inc$n[inc$`Age Group` == "< 10"])
+  inc_10 = new_cases_10/(at_risk_10/py)
+
+  new_cases_10_18 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group` == "10 - 18"])
+  at_risk_10_18 = sum(inc$n[inc$`Age Group` == "10 - 18"])
+  inc_10_18 = new_cases_10_18/(at_risk_10_18/py)
+
+  new_cases_19_30 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group` == "19 - 30"])
+  at_risk_19_30 = sum(inc$n[inc$`Age Group` == "19 - 30"])
+  inc_19_30 = new_cases_19_30/(at_risk_19_30/py)
+
+  new_cases_30 = sum(inc$n[inc$cfrd_cffpr=="Yes" & inc$`Age Group` == "> 30"])
+  at_risk_30 = sum(inc$n[inc$`Age Group` == "> 30"])
+  inc_30 = new_cases_30/(at_risk_30/py)
+
+  # Nice results
+  list(`Year` = year,
+       `Total New Cases` = new_cases,
+       `Total At Risk` = total_at_risk,
+       `Total Incidence` = total_inc,
+
+       `Total New Cases Age < 10` = new_cases_10,
+       `Total At Risk Age < 10` = at_risk_10,
+       `Total Incidence Age < 10` = inc_10,
+
+       `Total New Cases Age 10-18` = new_cases_10_18,
+       `Total At Risk Age 10-18` = at_risk_10_18,
+       `Total Incidence Age 10-18` = inc_10_18,
+
+       `Total New Cases Age 19-30` = new_cases_19_30,
+       `Total At Risk Age 19-30` = at_risk_19_30,
+       `Total Incidence Age 19-30` = inc_19_30,
+
+       `Total New Cases Age > 30` = new_cases_30,
+       `Total At Risk Age > 30` = at_risk_30,
+       `Total Incidence Age > 30` = inc_30)
+}
+incidence = lapply(2003:2018, function(y){year_inc(year = y)})
+incidence = data.frame(do.call(rbind,incidence),check.names = F)
+incidence = data.frame(lapply(incidence,as.numeric),check.names = F)
+kable(incidence,digits = 3)
+```
+
+# Data cleaning
+
+```{r}
 # Import data and codebook
-annual <- read.csv("~/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_Annualized_Del1.csv",
+annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
   na.strings = c("", "N/A")
 )
-data_dict <- readxl::read_excel("~/Documents/Work/Christine Chan/CF registry data/Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
+data_dict <- readxl::read_excel("./Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
 data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
@@
 # Get mutations
-demogs <- read.csv("~/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_DemogCFDiag_Del1.csv")
+demogs <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
 demogs <- demogs %>% select(eDWID, Gender, MutClass)
@@
 
-# Data cleaning
-
 Registry fields included in our liver disease yes/no variable were:
@@
   tab_options(column_labels.hidden = TRUE)
-```
-
-# CFRD incidence and prevalence
-
-## By year
-
-```{r results='asis'}
-# Find the year that each person was diagnosed, make wide
-first_cfrd <- annual %>%
-  filter(cfrd_status_annual == "CFRD+") %>%
-  group_by(eDWID) %>%
-  slice_head(n = 1) %>%
-  select(eDWID, ReviewYear) %>%
-  rename(first_cfrd = ReviewYear)
-# Make a dataframe for who was at risk each year
-at_risk <- expand.grid(eDWID = unique(annual$eDWID), ReviewYear = 2003:2018)
-at_risk <- at_risk %>% arrange(eDWID)
-at_risk <- at_risk %>%
-  full_join(., annual %>%
-    select(eDWID, ReviewYear, age_2018, A_hgba1c, cfrd_status_annual),
-  by = join_by(eDWID, ReviewYear)
-  )
-# Combine
-inc <- left_join(at_risk, first_cfrd, by = join_by(eDWID))
-# Calculate all missing ages
-inc <- inc %>%
-  group_by(eDWID) %>%
-  mutate(age = first(na.omit(age_2018)) - (2018 - ReviewYear)) %>%
-  # No missing age in 2018 field, so this indicates whether they were at risk
-  # (i.e. seen that year)
-  mutate(age_2018 = !is.na(age_2018)) %>%
-  rename(at_risk = age_2018) %>%
-  ungroup()
-# Age group
-inc$age_group <- cut(inc$age, c(-Inf, 10, 19, 31, Inf), right = F)
-# Remove people from at risk after diagnosis
-inc$at_risk[!is.na(inc$first_cfrd) & inc$ReviewYear > inc$first_cfrd] <- FALSE
-# Use gtsummary to make a list of tables by year
-invisible(lapply(unique(inc$ReviewYear), function(y) {
-  d <- inc %>%
-    filter(ReviewYear == y, at_risk) %>%
-    # Make missing values explicit for the tables
-    mutate(
-      a1c = factor(!is.na(A_hgba1c), levels = c(F, T), labels = c("No", "Yes")),
-      cfrd_status_annual = fct_explicit_na(cfrd_status_annual),
-    ) %>%
-    select(age_group, cfrd_status_annual, a1c, A_hgba1c) %>%
-    tbl_summary(
-      by = age_group, missing_text = "(Missing)",
-      label = list(
-        cfrd_status_annual ~ "Annualized CFRD Status",
-        A_hgba1c ~ "Annualized HbA1c", a1c ~ "A1c Collected"
-      ),
-      statistic = list(all_continuous() ~ "{mean} ({sd})")
-    ) %>%
-    add_overall()
-  cat("\n")
-  cat("### ", y)
-  cat("\n")
-  print(d)
-  cat("\n")
-}))
 ```

### 2023-12-07 Christine Chan/CF Registry/cf_registry_analysis.qmd
registry survival analysis data formatting

--- 2023-12-01 Christine Chan/CF Registry/cf_registry_analysis.qmd
@@
 library(labelled)
-library(parallel)
+library(survival)
+library(survminer)
+switch(Sys.info()[["sysname"]],
+  Windows = {
+    home_dir <- "C:/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data"
+  },
+  Darwin = {
+    home_dir <- "/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data"
+  }
+)
+knitr::opts_knit$set(root.dir = home_dir)
+```
+
+# CFRD incidence and prevalence
+
+```{r}
+# All data
+encounter <- read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv",
+  na.strings = c("", "N/A")
+)
+annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
+  na.strings = c("", "N/A")
+)
+encounter <- encounter %>%
+  arrange(eDWID, reviewyear, encounternum)
+annual <- annual %>% arrange(eDWID, ReviewYear)
+# Add age
+ages <- annual %>%
+  mutate(`Age in 2018` = 2018 - (ReviewYear - Age_YrEnd)) %>%
+  rename(reviewyear = ReviewYear) %>%
+  group_by(eDWID, reviewyear) %>%
+  summarise(
+    `Age in 2018` = as.numeric(names(sort(table(`Age in 2018`),
+      decreasing = TRUE
+    )[1])),
+    Age = Age_YrEnd[1], .groups = "drop"
+  )
+encounter <- left_join(encounter, ages, by = c("eDWID", "reviewyear"))
+encounter$`Age Group 2018` <- cut(encounter$`Age in 2018`,
+  c(-Inf, 10, 19, 31, Inf),
+  labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
+  right = F
+)
+encounter$`Age Group` <- cut(encounter$Age,
+  c(-Inf, 10, 19, 31, Inf),
+  labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
+  right = F
+)
+# For those with CFRD by registry criteria, find first visit with diagnosis
+first_cfrd_cffpr <- encounter %>%
+  filter(cfrd_status == 3) %>%
+  group_by(eDWID) %>%
+  filter(row_number() == 1) %>%
+  select(eDWID, reviewyear, `Age in 2018`) %>%
+  rename(first_cfrd_cffpr = reviewyear)
+# Add to encounter level data
+encounter <- left_join(encounter, first_cfrd_ogtt, by = c("eDWID", "Age in 2018"))
+encounter <- left_join(encounter, first_cfrd_cffpr, by = c("eDWID", "Age in 2018"))
+encounter <- left_join(encounter, first_cfrd_both, by = c("eDWID", "Age in 2018"))
+# Fill down CFRD Yes/No by year
+encounter <- encounter %>%
+  group_by(eDWID) %>%
+  mutate(
+    cfrd_cffpr = case_when(
+      reviewyear < first_cfrd_cffpr ~ "No",
+      reviewyear >= first_cfrd_cffpr ~ "Yes"
+    )
+  ) %>%
+  ungroup()
+encounter$cfrd_ogtt[is.na(encounter$cfrd_ogtt)] <- "No"
+encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
+encounter$cfrd_both[is.na(encounter$cfrd_both)] <- "No"
+# Ordered factors
+encounter$cfrd_ogtt <- factor(encounter$cfrd_ogtt, levels = c("No", "Yes"), ordered = T)
+encounter$cfrd_cffpr <- factor(encounter$cfrd_cffpr, levels = c("No", "Yes"), ordered = T)
+encounter$cfrd_both <- factor(encounter$cfrd_both, levels = c("No", "Yes"), ordered = T)
+# Remove those missing age
+encounter <- encounter %>% filter(!is.na(Age))
 ```
@@
 ```{r}
+#| label: tbl-incidence_cffpr
+#| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR Alone"
+# For each year, pull everyone with and without CFRD that year
+year_inc <- function(df = encounter, year, py = 1000) {
+  # Filter and count
+  inc <- df %>%
+    filter(reviewyear == year, is.na(first_cfrd_cffpr) | first_cfrd_cffpr >= year) %>%
+    group_by(eDWID) %>%
+    filter(row_number() == 1) %>%
+    select(eDWID, `Age Group`, cfrd_cffpr)
+  inc <- inc %>%
+    group_by(cfrd_cffpr) %>%
+    count(`Age Group`, .drop = F)
+  inc <- inc[!is.na(inc$`Age Group`), ]
+  # Calculations
+  new_cases <- sum(inc$n[inc$cfrd_cffpr == "Yes"])
+  total_at_risk <- sum(inc$n)
+  total_inc <- new_cases / (total_at_risk / py)
+
+  new_cases_10 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "< 10"])
+  at_risk_10 <- sum(inc$n[inc$`Age Group` == "< 10"])
+  inc_10 <- new_cases_10 / (at_risk_10 / py)
+
+  new_cases_10_18 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "10 - 18"])
+  at_risk_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
+  inc_10_18 <- new_cases_10_18 / (at_risk_10_18 / py)
+
+  new_cases_19_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "19 - 30"])
+  at_risk_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
+  inc_19_30 <- new_cases_19_30 / (at_risk_19_30 / py)
+
+  new_cases_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "> 30"])
+  at_risk_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
+  inc_30 <- new_cases_30 / (at_risk_30 / py)
+
+  # Nice results
+  list(
+    `Year` = year,
+    `Total New Cases` = new_cases,
+    `Total At Risk` = total_at_risk,
+    `Total Incidence` = total_inc,
+    `Total New Cases Age < 10` = new_cases_10,
+    `Total At Risk Age < 10` = at_risk_10,
+    `Total Incidence Age < 10` = inc_10,
+    `Total New Cases Age 10-18` = new_cases_10_18,
+    `Total At Risk Age 10-18` = at_risk_10_18,
+    `Total Incidence Age 10-18` = inc_10_18,
+    `Total New Cases Age 19-30` = new_cases_19_30,
+    `Total At Risk Age 19-30` = at_risk_19_30,
+    `Total Incidence Age 19-30` = inc_19_30,
+    `Total New Cases Age > 30` = new_cases_30,
+    `Total At Risk Age > 30` = at_risk_30,
+    `Total Incidence Age > 30` = inc_30
+  )
+}
+incidence <- lapply(2003:2018, function(y) {
+  year_inc(year = y)
+})
+incidence <- data.frame(do.call(rbind, incidence), check.names = F)
+incidence <- data.frame(lapply(incidence, as.numeric), check.names = F)
+kable(incidence, digits = 3)
+```
+
+In @tbl-incidence_cffpr, incidence per year is calculated as the number of new CFRD diagnoses (by registry definition alone) per 1,000 person-years of those at risk, excluding previously identified cases.
+
+```{r}
+#| label: tbl-prevalence_cffpr
+#| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR Alone"
+encounter %>%
+  group_by(eDWID, reviewyear) %>%
+  summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>%
+  group_by(reviewyear) %>%
+  summarise(Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100, .groups = "drop") %>%
+  kable(., digits = 3)
+```
+
+```{r}
+#| label: tbl-prevalence_cffpr_year
+#| tbl-cap: "Yearly Prevalence of CFRD by Age Group Based on CFFPR Alone"
+encounter %>%
+  group_by(eDWID, reviewyear, `Age Group`) %>%
+  summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>%
+  group_by(reviewyear, `Age Group`) %>%
+  summarise(Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100, .groups = "drop") %>%
+  pivot_wider(names_from = `Age Group`, values_from = Prevalence) %>%
+  kable(., digits = 3)
+```
+
+In @tbl-prevalence_cffpr_year and @tbl-prevalence_cffpr, prevalence is calculated as the total number of CFRD cases (by registry definition alone) seen in clinic divided by the total number of people seen in clinic in a given year.
+
+# CFRD screening rates
+
+```{r}
+#| label: tbl-screening_ogtt
+#| tbl-cap: "Yearly OGTT Screening Rates"
+year_screen <- function(df = encounter, year) {
+  # Filter and count
+  inc <- df %>%
+    filter(reviewyear == year, first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>%
+    mutate(ogtt = !is.na(ogttfast_bloodglucose) | !is.na(twohour_bloodglucose)) %>%
+    group_by(eDWID) %>%
+    summarise(ogtt = max(ogtt), `Age Group` = `Age Group`[1])
+  inc <- inc %>%
+    group_by(ogtt) %>%
+    count(`Age Group`) %>%
+    filter(`Age Group` != "< 10")
+  # Calculations
+  total_screened <- sum(inc$n[inc$ogtt == 1])
+  total_seen <- sum(inc$n)
+  screen_rate <- total_screened / total_seen * 100
+
+  screened_10_18 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "10 - 18"]
+  seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
+  rate_10_18 <- screened_10_18 / seen_10_18 * 100
+
+  screened_19_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "19 - 30"]
+  seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
+  rate_19_30 <- screened_19_30 / seen_19_30 * 100
+
+  screened_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "> 30"]
+  seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
+  rate_30 <- screened_30 / seen_30 * 100
+
+  # Nice results
+  list(
+    `Year` = year,
+    `Total Screened` = total_screened,
+    `Total Seen` = total_seen,
+    `Screen Rate (%)` = screen_rate,
+    `Total Screened Age 10-18` = screened_10_18,
+    `Total Seen Age 10-18` = seen_10_18,
+    `Screen Rate Age 10-18 (%)` = rate_10_18,
+    `Total Screened Age 19-30` = screened_19_30,
+    `Total Seen Age 19-30` = seen_19_30,
+    `Screen Rate Age 19-30 (%)` = rate_19_30,
+    `Total Screened Age > 30` = screened_30,
+    `Total Seen Age > 30` = seen_30,
+    `Screen Rate Age > 30 (%)` = rate_30
+  )
+}
+screen <- lapply(2003:2018, function(y) {
+  year_screen(year = y)
+})
+screen <- data.frame(do.call(rbind, screen), check.names = F)
+screen <- data.frame(lapply(screen, as.numeric), check.names = F)
+kable(screen, digits = 3)
+```
+
+```{r}
+#| label: tbl-screening_a1c
+#| tbl-cap: "Yearly HbA1c Screening Rates"
+year_screen <- function(df = encounter, year) {
+  # Filter and count
+  inc <- df %>%
+    filter(reviewyear == year, first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>%
+    mutate(a1c = !is.na(hgba1c)) %>%
+    group_by(eDWID) %>%
+    summarise(a1c = max(a1c), `Age Group` = `Age Group`[1])
+  inc <- inc %>%
+    group_by(a1c) %>%
+    count(`Age Group`) %>%
+    filter(`Age Group` != "< 10")
+  # Calculations
+  total_screened <- sum(inc$n[inc$a1c == 1])
+  total_seen <- sum(inc$n)
+  screen_rate <- total_screened / total_seen * 100
+
+  screened_10_18 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "10 - 18"]
+  seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
+  rate_10_18 <- screened_10_18 / seen_10_18 * 100
+
+  screened_19_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "19 - 30"]
+  seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
+  rate_19_30 <- screened_19_30 / seen_19_30 * 100
+
+  screened_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "> 30"]
+  seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
+  rate_30 <- screened_30 / seen_30 * 100
+
+  # Nice results
+  list(
+    `Year` = year,
+    `Total Screened` = total_screened,
+    `Total Seen` = total_seen,
+    `Screen Rate (%)` = screen_rate,
+    `Total Screened Age 10-18` = screened_10_18,
+    `Total Seen Age 10-18` = seen_10_18,
+    `Screen Rate Age 10-18 (%)` = rate_10_18,
+    `Total Screened Age 19-30` = screened_19_30,
+    `Total Seen Age 19-30` = seen_19_30,
+    `Screen Rate Age 19-30 (%)` = rate_19_30,
+    `Total Screened Age > 30` = screened_30,
+    `Total Seen Age > 30` = seen_30,
+    `Screen Rate Age > 30 (%)` = rate_30
+  )
+}
+screen <- lapply(2003:2018, function(y) {
+  year_screen(year = y)
+})
+screen <- data.frame(do.call(rbind, screen), check.names = F)
+screen <- data.frame(lapply(screen, as.numeric), check.names = F)
+kable(screen, digits = 3)
+```
+
+In @tbl-screening_ogtt and @tbl-screening_a1c, people with CFRD (by registry definition) and those under age 10 were excluded.
+Screening rate is calculated as the number of people with an OGTT or HbA1c, respectively, divided by the total number of people seen in clinic that year.
+
+# Data cleaning
+
+```{r}
 # Import data and codebook
-annual <- read.csv("~/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_Annualized_Del1.csv",
+annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
   na.strings = c("", "N/A")
 )
-data_dict <- readxl::read_excel("~/Documents/Work/Christine Chan/CF registry data/Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
+data_dict <- readxl::read_excel("./Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
 data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
@@
 annual$A_IsOnEnzymes <- relevel(annual$A_IsOnEnzymes, ref = "Yes")
+# To have CFRD, must have two consecutive visits with CFRD
+# Once diagnosed, always in CFRD group.
+annual = annual %>%
+  group_by(eDWID) %>%
+  mutate(two_consec = 0)
 ```
-
-# Data cleaning
 
@@
 
-# CFRD incidence and prevalence
+# Survival analysis
 
-## By year
+Based on CFFPR annualized data.
 
-```{r results='asis'}
-# Find the year that each person was diagnosed, make wide
-first_cfrd <- annual %>%
-  filter(cfrd_status_annual == "CFRD+") %>%
+```{r}
+# Create survival dataset
+s <- annual %>%
+  rename(end = ReviewYear) %>% # ReviewYear is the end of the interval
   group_by(eDWID) %>%
-  slice_head(n = 1) %>%
-  select(eDWID, ReviewYear) %>%
-  rename(first_cfrd = ReviewYear)
-# Make a dataframe for who was at risk each year
-at_risk <- expand.grid(eDWID = unique(annual$eDWID), ReviewYear = 2003:2018)
-at_risk <- at_risk %>% arrange(eDWID)
-at_risk <- at_risk %>%
-  full_join(., annual %>%
-    select(eDWID, ReviewYear, age_2018, A_hgba1c, cfrd_status_annual),
-  by = join_by(eDWID, ReviewYear)
+  mutate(start = lag(end,default = 2002)) %>% # Get interval start
+  select(eDWID, start, end, Age_YrEnd, cfrd_status_annual) # Sort columns
+# Age group
+s$age_group <- cut(s$Age_YrEnd,
+  breaks = c(-Inf, 10, 19, 31, Inf), right = F
 )
-# Combine
-inc <- left_join(at_risk, first_cfrd, by = join_by(eDWID))
-# Calculate all missing ages
-inc <- inc %>%
-  group_by(eDWID) %>%
-  mutate(age = first(na.omit(age_2018)) - (2018 - ReviewYear)) %>%
-  # No missing age in 2018 field, so this indicates whether they were at risk
-  # (i.e. seen that year)
-  mutate(age_2018 = !is.na(age_2018)) %>%
-  rename(at_risk = age_2018) %>%
-  ungroup()
-# Age group
-inc$age_group <- cut(inc$age, c(-Inf, 10, 19, 31, Inf), right = F)
-# Remove people from at risk after diagnosis
-inc$at_risk[!is.na(inc$first_cfrd) & inc$ReviewYear > inc$first_cfrd] <- FALSE
-# Use gtsummary to make a list of tables by year
-invisible(lapply(unique(inc$ReviewYear), function(y) {
-  d <- inc %>%
-    filter(ReviewYear == y, at_risk) %>%
-    # Make missing values explicit for the tables
-    mutate(
-      a1c = factor(!is.na(A_hgba1c), levels = c(F, T), labels = c("No", "Yes")),
-      cfrd_status_annual = fct_explicit_na(cfrd_status_annual),
-    ) %>%
-    select(age_group, cfrd_status_annual, a1c, A_hgba1c) %>%
-    tbl_summary(
-      by = age_group, missing_text = "(Missing)",
-      label = list(
-        cfrd_status_annual ~ "Annualized CFRD Status",
-        A_hgba1c ~ "Annualized HbA1c", a1c ~ "A1c Collected"
-      ),
-      statistic = list(all_continuous() ~ "{mean} ({sd})")
-    ) %>%
-    add_overall()
-  cat("\n")
-  cat("### ", y)
-  cat("\n")
-  print(d)
-  cat("\n")
-}))
+knitr::kable(head(s,40))
 ```
 
-# Logistic analyses
-
-Any participants with T1D or T2D were excluded from these analyses.
-Variables that were entirely missing in a given year (e.g., `A_Vx770` and `A_VX809comb` in 2005) were excluded from those models.
-Participants under age 10 at year end were also excluded due to missing PFT data.
+```{r}
+# New dataset approach
+cases = unique(annual$eDWID[annual$cfrd_status_annual == "CFRD+"])
+controls_df = annual[!annual$eDWID %in% cases,]
+cases_df = annual[annual$eDWID %in% cases,]
+# Controls
+tc = controls_df %>% 
+  select(eDWID,ReviewYear,cfrd_status_annual) %>%
+  filter(!is.na(cfrd_status_annual)) %>%
+  group_by(eDWID) %>%
+  summarise(start = first(ReviewYear),
+         end = last(ReviewYear),.groups = "drop") %>%
+  mutate(cfrd = 0)
+# Cases
+tc2 = cases_df %>% 
+  select(eDWID,ReviewYear,cfrd_status_annual) %>%
+  filter(!is.na(cfrd_status_annual)) %>%
+  mutate(neg_year = ReviewYear * (ifelse(cfrd_status_annual=="CFRD-",-1,1))) %>%
+  arrange(eDWID,cfrd_status_annual,neg_year) %>%
+  group_by(eDWID,cfrd_status_annual) %>% 
+  filter(row_number()==1)
+```
 
 ```{r}
-# Analysis function
-log_reg <- function(year, missingness = F) {
-  # Clean data
-  df <- annual %>%
-    filter(ReviewYear == year, Age_YrEnd >= 10) %>%
-    select(
-      eDWID, age_group, Gender, bmi_perc, MutClass, A_IsOnEnzymes,
-      A_FEV1_pct_predicted, tube_feeding, NumPulmExacerbation,
-      A_pseudomonasaeruginosa, A_fungalyeast1, A_mycocultureresults3,
-      A_pulmonarycomplications1,
-      A_corticosteroids1, cfrd_status_annual, A_Vx770, A_VX809comb,
-      liver_disease
+# Survival fit
+surv_obj <- Surv(
+  time = s$start, time2 = s$end,
+  event = as.numeric(s$cfrd_status_annual) - 1
 )
-  # Assume that missing are CFRD- for these models
-  df$cfrd_status_annual[is.na(df$cfrd_status_annual)] <- "CFRD-"
-  # Relevel age group since we're excluding kids < 10
-  df$age_group <- droplevels(df$age_group)
-  # Remove all empty columns
-  empty <- which(colSums(is.na(df)) == nrow(df))
-  if (length(empty) > 0) {
-    df <- df[, -empty]
-  }
-  # Models
-  vars <- colnames(df)[which(!colnames(df) %in%
-    c(bugs, liver_fields, "cfrd_status_annual", "eDWID"))]
-  base_mod <- as.formula(paste0(
-    "cfrd_status_annual~",
-    paste0(vars, collapse = "+")
-  ))
-  # Labels
-  labels <- list(
-    cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
-    Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
-    A_FEV1_pct_predicted = "FEV1 percent predicted",
-    tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
-    A_corticosteroids1 = "Use of corticosteroids",
-    NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
-    height_perc = "Height percentile", bmi_perc = "BMI percentile",
-    A_Vx770 = "Use of CFTR modulator Vx770",
-    A_VX809comb = "Use of CFTR modulator Vx809 combination",
-    A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
-    A_fungalyeast1 = "Aspergillus (any species)",
-    A_pulmonarycomplications1 = "ABPA",
-    A_mycocultureresults3 = "Microorganisms in myco culture"
+fit <- survfit(surv_obj ~ age_group, id = eDWID, data = s)
+ggsurvplot(fit,
+  risk.table = T, cumevents = T,
+  fun = "event", xlim = c(2002, 2018)
 )
-  var_label(df[, vars[which(vars %in% colnames(df))]]) <-
-    as.character(labels[vars[which(vars %in% colnames(df))]])
-  var_label(df[, bugs[which(bugs %in% colnames(df))]]) <-
-    as.character(labels[bugs[which(bugs %in% colnames(df))]])
-  # Visualize missingness if required
-  if (missingness) {
-    cat("\n")
-    cat("### Missingness")
-    cat("\n")
-    cat("\n")
-    print(naniar::vis_miss(df))
-    cat("\n")
-  }
-  # Simple models
-  mod1 <- glm(base_mod, data = df, family = "binomial")
-  # Print results
-  cat("\n")
-  cat("### Base Model")
-  cat("\n")
-  cat("\n")
-  print(forest_model(mod1, exclude_infinite_cis = F))
-  cat("\n")
-  if ("A_pseudomonasaeruginosa" %in% colnames(df)) {
-    mod2 <- glm(update(base_mod, . ~ . + A_pseudomonasaeruginosa),
-      data = df,
-      family = "binomial"
-    )
-    cat("\n")
-    cat("### P. aeruginosa")
-    cat("\n")
-    cat("\n")
-    print(forest_model(mod2, exclude_infinite_cis = F))
-    cat("\n")
-  }
-  if ("A_fungalyeast1" %in% colnames(df)) {
-    mod3 <- glm(update(base_mod, . ~ . + A_fungalyeast1),
-      data = df,
-      family = "binomial"
-    )
-    cat("\n")
-    cat("### Aspergillus")
-    cat("\n")
-    cat("\n")
-    print(forest_model(mod3, exclude_infinite_cis = F))
-    cat("\n")
-  }
-  if ("A_pulmonarycomplications1" %in% colnames(df)) {
-    mod4 <- glm(update(base_mod, . ~ . + A_pulmonarycomplications1),
-      data = df,
-      family = "binomial"
-    )
-    cat("\n")
-    cat("### ABPA")
-    cat("\n")
-    cat("\n")
-    print(forest_model(mod4, exclude_infinite_cis = F))
-    cat("\n")
-  }
-  if ("A_mycocultureresults3" %in% colnames(df)) {
-    mod5 <- glm(update(base_mod, . ~ . + A_mycocultureresults3),
-      data = df,
-      family = "binomial"
-    )
-    cat("\n")
-    cat("### Mycobacterium")
-    cat("\n")
-    cat("\n")
-    print(forest_model(mod5, exclude_infinite_cis = F))
-    cat("\n")
-  }
-}
 ```
-
-```{r results='asis',fig.height=10,fig.width=10}
-invisible(lapply(unique(annual$ReviewYear), function(y) {
-  cat("\n")
-  cat("## ", y)
-  cat("\n")
-  log_reg(y)
-  cat("\n")
-}))
-```

### 2023-12-07 Christine Chan/CF Registry/cf_registry_analysis.qmd
basic Cox PH done

@@
 library(Hmisc)
+library(zoo)
 library(tidyverse)
@@
 library(survminer)
+library(knitr)
 switch(Sys.info()[["sysname"]],
@@
 # Add to encounter level data
-encounter <- left_join(encounter, first_cfrd_ogtt, by = c("eDWID", "Age in 2018"))
-encounter <- left_join(encounter, first_cfrd_cffpr, by = c("eDWID", "Age in 2018"))
-encounter <- left_join(encounter, first_cfrd_both, by = c("eDWID", "Age in 2018"))
+encounter <- left_join(encounter, first_cfrd_cffpr,
+  by = c("eDWID", "Age in 2018")
+)
 # Fill down CFRD Yes/No by year
@@
   ungroup()
-encounter$cfrd_ogtt[is.na(encounter$cfrd_ogtt)] <- "No"
 encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
-encounter$cfrd_both[is.na(encounter$cfrd_both)] <- "No"
 # Ordered factors
-encounter$cfrd_ogtt <- factor(encounter$cfrd_ogtt, levels = c("No", "Yes"), ordered = T)
-encounter$cfrd_cffpr <- factor(encounter$cfrd_cffpr, levels = c("No", "Yes"), ordered = T)
-encounter$cfrd_both <- factor(encounter$cfrd_both, levels = c("No", "Yes"), ordered = T)
+encounter$cfrd_cffpr <- factor(encounter$cfrd_cffpr,
+  levels = c("No", "Yes"), ordered = T
+)
 # Remove those missing age
@@
 )
-annual$age_2018 <- 2018 - (annual$ReviewYear - annual$Age_YrEnd)
-annual <- annual %>%
-  group_by(eDWID) %>%
-  mutate(age_2018 = as.numeric(names(sort(table(age_2018),
-    decreasing = TRUE
-  )[1])))
 # Get mutations
@@
 )
-annual$A_FEV1_pct_predicted <- cut(annual$A_FEV1_pct_predicted,
-  breaks = c(0, 40, 80, 100, Inf), right = F
-)
-annual$A_FEV1_pct_predicted <- relevel(annual$A_FEV1_pct_predicted,
-  ref = "[100,Inf)"
-)
-annual$height_perc <- cut(annual$height_perc, breaks = c(0, 10, 25, 50, Inf))
-annual$height_perc <- relevel(annual$height_perc, ref = "(50,Inf]")
-annual$bmi_perc <- cut(annual$bmi_perc, breaks = c(0, 10, 50, 85, Inf))
-annual$bmi_perc <- relevel(annual$bmi_perc, ref = "(50,85]")
 annual$age_group <- cut(annual$Age_YrEnd,
@@
 annual$A_IsOnEnzymes <- relevel(annual$A_IsOnEnzymes, ref = "Yes")
-# To have CFRD, must have two consecutive visits with CFRD
-# Once diagnosed, always in CFRD group.
-annual = annual %>%
+# Approximate year of birth
+by <- annual %>%
+  mutate(birth_year = floor(ReviewYear - Age_YrEnd)) %>%
   group_by(eDWID) %>%
-  mutate(two_consec = 0)
+  count(birth_year) %>%
+  group_by(eDWID) %>%
+  slice_min(birth_year) %>%
+  select(-n)
+annual <- left_join(annual, by, by = join_by(eDWID))
 ```
@@
 ```{r}
-# Create survival dataset
+# List of covariates
+vars <- c(
+  "age_group", "Gender", "bmi_perc", "MutClass", "A_IsOnEnzymes",
+  "A_FEV1_pct_predicted", "tube_feeding", "NumPulmExacerbation",
+  "A_pseudomonasaeruginosa", "A_fungalyeast1", "A_mycocultureresults3",
+  "A_pulmonarycomplications1", "A_corticosteroids1", "cfrd_status_annual",
+  "A_Vx770", "A_VX809comb", "liver_disease"
+)
+labels <- list(
+  cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
+  Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
+  A_FEV1_pct_predicted = "FEV1 percent predicted",
+  tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
+  A_corticosteroids1 = "Use of corticosteroids",
+  NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
+  height_perc = "Height percentile", bmi_perc = "BMI percentile",
+  A_Vx770 = "Use of CFTR modulator Vx770",
+  A_VX809comb = "Use of CFTR modulator Vx809 combination",
+  A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
+  A_fungalyeast1 = "Aspergillus (any species)",
+  A_pulmonarycomplications1 = "ABPA",
+  A_mycocultureresults3 = "Microorganisms in myco culture"
+)
+# For now, let's just assume that everyone's first CFRD+ diagnosis is correct
 s <- annual %>%
-  rename(end = ReviewYear) %>% # ReviewYear is the end of the interval
+  filter(!is.na(cfrd_status_annual), Age_YrEnd != 0) %>%
+  select(
+    eDWID, ReviewYear, birth_year, Age_YrEnd,
+    all_of(vars), cfrd_status_annual
+  ) %>%
+  rename(tstop = ReviewYear) %>%
   group_by(eDWID) %>%
-  mutate(start = lag(end,default = 2002)) %>% # Get interval start
-  select(eDWID, start, end, Age_YrEnd, cfrd_status_annual) # Sort columns
-# Age group
-s$age_group <- cut(s$Age_YrEnd,
-  breaks = c(-Inf, 10, 19, 31, Inf), right = F
-)
-knitr::kable(head(s,40))
+  mutate(
+    tstart = lag(tstop, default = birth_year[1]),
+    age_at_entry = Age_YrEnd[1]
+  ) %>%
+  select(eDWID, birth_year, tstart, tstop, cfrd_status_annual, age_at_entry, everything())
+write.csv(s, file = "~/survival_data.csv", row.names = F, na = "")
 ```
@@
 ```{r}
-# New dataset approach
-cases = unique(annual$eDWID[annual$cfrd_status_annual == "CFRD+"])
-controls_df = annual[!annual$eDWID %in% cases,]
-cases_df = annual[annual$eDWID %in% cases,]
-# Controls
-tc = controls_df %>% 
-  select(eDWID,ReviewYear,cfrd_status_annual) %>%
-  filter(!is.na(cfrd_status_annual)) %>%
-  group_by(eDWID) %>%
-  summarise(start = first(ReviewYear),
-         end = last(ReviewYear),.groups = "drop") %>%
-  mutate(cfrd = 0)
-# Cases
-tc2 = cases_df %>% 
-  select(eDWID,ReviewYear,cfrd_status_annual) %>%
-  filter(!is.na(cfrd_status_annual)) %>%
-  mutate(neg_year = ReviewYear * (ifelse(cfrd_status_annual=="CFRD-",-1,1))) %>%
-  arrange(eDWID,cfrd_status_annual,neg_year) %>%
-  group_by(eDWID,cfrd_status_annual) %>% 
-  filter(row_number()==1)
-```
-
-```{r}
-# Survival fit
-surv_obj <- Surv(
-  time = s$start, time2 = s$end,
-  event = as.numeric(s$cfrd_status_annual) - 1
+cox_res <- coxph(
+  Surv(tstart, tstop, cfrd_status_annual == "CFRD+") ~
+    age_group + A_FEV1_pct_predicted,
+  data = s
 )
-fit <- survfit(surv_obj ~ age_group, id = eDWID, data = s)
-ggsurvplot(fit,
-  risk.table = T, cumevents = T,
-  fun = "event", xlim = c(2002, 2018)
+ggsurvplot(survfit(cox_res),
+  data = s,
+  risk.table = T, cumevents = T, fun = "event", xlim = c(2003, 2018)
 )

### 2023-12-08 Christine Chan/CF Registry/cf_registry_analysis.qmd
CF registry interval censored data ready

@@
 annual <- left_join(annual, by, by = join_by(eDWID))
+# Get first year of CFRD
+first_cfrd <- annual %>%
+  filter(cfrd_status_annual == "CFRD+") %>%
+  group_by(eDWID) %>%
+  summarise(first_cfrd = min(ReviewYear))
+annual <- left_join(annual, first_cfrd, by = join_by(eDWID))
 ```
@@
 s <- annual %>%
-  filter(!is.na(cfrd_status_annual), Age_YrEnd != 0) %>%
+  filter(
+    !is.na(cfrd_status_annual),
+    Age_YrEnd != 0,
+    is.na(first_cfrd) | ReviewYear <= first_cfrd
+  ) %>%
   select(
@@
   select(eDWID, birth_year, tstart, tstop, cfrd_status_annual, age_at_entry, everything())
-write.csv(s, file = "~/survival_data.csv", row.names = F, na = "")
 ```
@@
     age_group + A_FEV1_pct_predicted,
-  data = s
+  data = s, cluster = eDWID
 )
@@
 )
+surv_fit = survfit(Surv(tstart, tstop, cfrd_status_annual == "CFRD+") ~ age_group, 
+                   data = s)
+ggsurvplot(surv_fit,
+  risk.table = T, cumevents = T, fun = "event", xlim = c(2003, 2018)
+)
 ```

### 2023-12-13 Christine Chan/CF Registry/cf_registry_analysis.qmd
calculated how many had 1 row of data and how many had >=3 years between CFRD+

@@
 library(survminer)
+library(icenReg)
 library(knitr)
+library(eulerr)
 switch(Sys.info()[["sysname"]],
   Windows = {
-    home_dir <- "C:/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data"
+    home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Christine Chan/CF registry data"
   },
   Darwin = {
-    home_dir <- "/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data"
+    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Christine Chan/CF registry data"
   }
@@
 Based on CFFPR annualized data.
+These data are interval censored because we know the first visit at which people with CFRD were diagnosed, but we do not know the exact time that they developed CFRD.
+These data are also left truncated because not all participants enter the study at the same time.
+
+## Participants with a single observation
+
+```{r}
+# Get total number of people with only one row with CFRD info
+one_row <- annual %>%
+  group_by(eDWID) %>%
+  summarise(n = n()) %>%
+  filter(n == 1)
+one_row_cfrd <- annual %>%
+  filter(!is.na(cfrd_status_annual)) %>%
+  group_by(eDWID) %>%
+  summarise(n = n()) %>%
+  filter(n == 1)
+```
+
+There were `r nrow(one_row)` people with a single year of observation in the annual data (across all variables), and `r nrow(one_row_cfrd)` people with a single non-missing CFRD status.
+`r length(intersect(one_row$eDWID,one_row_cfrd$eDWID))` people fall into both of these categories:
+
+```{r}
+plot(
+  euler(list(
+    "One Observation (All)" = one_row$eDWID,
+    "One Observation (CFRD)" = one_row_cfrd$eDWID
+  )),
+  fills = c("lightblue2", "lightsalmon"),
+  quantities = T,
+  legend = list(side = "right")
+)
+```
+
+## Participants with $\geq$ 3 years between CFRD+
+
+```{r}
+# I can't think of a good tidyverse way to do this, so trying it myself
+dfs <- annual %>% select(eDWID, ReviewYear, cfrd_status_annual)
+dfs <- split.data.frame(dfs, dfs$eDWID)
+# For each ID, test whether they follow the pattern
+fit_pattern <- lapply(dfs, function(d) {
+  r <- rle(as.character(d$cfrd_status_annual))
+  # Check for CFRD+, CFRD-, CFRD+ pattern
+  # Can have NAs on either side of the CFRD- as well, so
+  # CFRD+, CFRD-, CFRD+ OR CFRD+, NA, CFRD-, CFRD+ OR CFRD+, CFRD-, NA, CFRD+
+  # OR CFRD+, NA,CFRD-, NA,CFRD+
+  if (
+    grepl(
+      paste("CFRD\\+,CFRD\\-,CFRD\\+|", "CFRD\\+,NA,CFRD\\-,CFRD\\+|",
+        "CFRD\\+,CFRD\\-,NA,CFRD\\+|", "CFRD\\+,NA,CFRD\\-,NA,CFRD\\+",
+        sep = "", collapse = ""
+      ),
+      paste(r$values, collapse = ",")
+    )) {
+    return(unique(d$eDWID))
+  } else {
+    return(NA)
+  }
+})
+fit_pattern = fit_pattern[!is.na(fit_pattern)]
+# Filter out those that don't follow the pattern, calculate largest gaps 
+# between CFRD+ diagnoses
+largest_gap = annual %>% 
+  filter(eDWID %in% names(t)) %>% 
+  select(eDWID,ReviewYear,cfrd_status_annual) %>%
+  group_by(eDWID) %>%
+  summarise(largest_gap = max(diff(ReviewYear)))
+```
+
+There were a total of `r nrow(largest_gap)` partipants who had a diagnosis of CFRD+ followed by at least one CFRD-, followed again by a CFRD+ diagnosis.
+For each person, we calculated the largest gap between CFRD+ diagnoses.
+
+```{r}
+kable(table(largest_gap$largest_gap),
+      col.names = c("Years between CFRD+","n"))
+```
+
+## Basic survival curve
 
@@
 )
-surv_fit = survfit(Surv(tstart, tstop, cfrd_status_annual == "CFRD+") ~ age_group, 
-                   data = s)
+surv_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual == "CFRD+") ~ age_group,
+  data = s
+)
 ggsurvplot(surv_fit,

### 2023-12-13 Christine Chan/CF Registry/cf_registry_analysis.qmd
corrected numbers for people with >= 3 year gaps

@@
   group_by(reviewyear) %>%
-  summarise(Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100, .groups = "drop") %>%
+  summarise(Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100, 
+            .groups = "drop") %>%
   kable(., digits = 3)
@@
 largest_gap = annual %>% 
-  filter(eDWID %in% names(t)) %>% 
+  filter(eDWID %in% names(fit_pattern),cfrd_status_annual=="CFRD+") %>% 
   select(eDWID,ReviewYear,cfrd_status_annual) %>%
   group_by(eDWID) %>%
-  summarise(largest_gap = max(diff(ReviewYear)))
+  summarise(largest_gap = max(diff(ReviewYear))-1)
+largest_gap = data.frame(table(largest_gap$largest_gap))
+colnames(largest_gap) = c("Years between CFRD+","n")
 ```
 
-There were a total of `r nrow(largest_gap)` partipants who had a diagnosis of CFRD+ followed by at least one CFRD-, followed again by a CFRD+ diagnosis.
+There were a total of `r sum(largest_gap$n)` participants who had a diagnosis of CFRD+ followed by at least one CFRD-, followed again by a CFRD+ diagnosis.
 For each person, we calculated the largest gap between CFRD+ diagnoses.
+A total of `r sum(largest_gap$n[as.numeric(largest_gap[,1])>=3])` people had a gap $\geq$ 3 years.
 
 ```{r}
-kable(table(largest_gap$largest_gap),
-      col.names = c("Years between CFRD+","n"))
+kable(largest_gap)
 ```

### 2023-12-14 Christine Chan/CF Registry/cf_registry_analysis.qmd
CFRD registry base cox model ready to go

@@
 library(eulerr)
+library(broom)
 switch(Sys.info()[["sysname"]],
@@
   group_by(reviewyear) %>%
-  summarise(Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100, 
-            .groups = "drop") %>%
+  summarise(
+    Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100,
+    .groups = "drop"
+  ) %>%
   kable(., digits = 3)
@@
 })
-fit_pattern = fit_pattern[!is.na(fit_pattern)]
+fit_pattern <- fit_pattern[!is.na(fit_pattern)]
 # Filter out those that don't follow the pattern, calculate largest gaps
 # between CFRD+ diagnoses
-largest_gap = annual %>% 
-  filter(eDWID %in% names(fit_pattern),cfrd_status_annual=="CFRD+") %>% 
-  select(eDWID,ReviewYear,cfrd_status_annual) %>%
+largest_gap <- annual %>%
+  filter(eDWID %in% names(fit_pattern), cfrd_status_annual == "CFRD+") %>%
+  select(eDWID, ReviewYear, cfrd_status_annual) %>%
   group_by(eDWID) %>%
-  summarise(largest_gap = max(diff(ReviewYear))-1)
-largest_gap = data.frame(table(largest_gap$largest_gap))
-colnames(largest_gap) = c("Years between CFRD+","n")
+  summarise(largest_gap = max(diff(ReviewYear)) - 1)
+largest_gap <- data.frame(table(largest_gap$largest_gap))
+colnames(largest_gap) <- c("Years between CFRD+", "n")
 ```
@@
 
-## Basic survival curve
-
 ```{r}
@@
 )
+# Nice labels for plots and tables
 labels <- list(
@@
   ) %>%
-  select(eDWID, birth_year, tstart, tstop, cfrd_status_annual, age_at_entry, everything())
+  select(
+    eDWID, birth_year, tstart, tstop, cfrd_status_annual,
+    age_at_entry, everything()
+  )
+# Convert from tibble to dataframe for survival packages
+s <- as.data.frame(s)
+# CFRD status to numeric
+s$cfrd_status_annual <- as.numeric(s$cfrd_status_annual) - 1
 ```
 
+## Kaplan Meier Analysis
+
 ```{r}
-cox_res <- coxph(
-  Surv(tstart, tstop, cfrd_status_annual == "CFRD+") ~
-    age_group + A_FEV1_pct_predicted,
-  data = s, cluster = eDWID
+# Whole cohort
+km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ 1,
+  data = s,cluster = eDWID
 )
+ggsurvplot(km_fit,
+  data = s, risk.table = T, cumevents = T,
+  fun = "event", xlim = c(2003, 2018),break.x.by=1
+)
+kable(tidy(km_fit))
+# By age group
+km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ age_group,
+  data = s,cluster = eDWID
+)
+ggsurvplot(km_fit,
+  data = s, risk.table = T, cumevents = T,
+  fun = "event", xlim = c(2003, 2018),break.x.by=1
+)
+km_table <- tidy(km_fit)
+km_table <- split.data.frame(km_table, km_table$strata)
+invisible(lapply(km_table, function(t) {
+  name <- t$strata[1]
+  t <- t %>% select(-strata)
+  kable(t, caption = name)
+}))
+```
+
+## Cox PH model
+
+```{r}
+# Formula for base model
+x <- c(
+  "age_group", "Gender", "bmi_perc", "MutClass", "A_IsOnEnzymes",
+  "A_FEV1_pct_predicted", "tube_feeding", "NumPulmExacerbation",
+  "A_pulmonarycomplications1", "A_corticosteroids1", "liver_disease"
+)
+y <- "Surv(tstart, tstop, cfrd_status_annual)"
+f <- as.formula(paste0(y, "~", paste0(x, collapse = "+")))
+cox_res <- coxph(f, data = s, cluster = eDWID)
+# Plot survival
 ggsurvplot(survfit(cox_res),
-  data = s,
-  risk.table = T, cumevents = T, fun = "event", xlim = c(2003, 2018)
+  data = s, risk.table = T, cumevents = T, fun = "event",
+  xlim = c(2003, 2018), break.x.by = 1
 )
-surv_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual == "CFRD+") ~ age_group,
-  data = s
-)
-ggsurvplot(surv_fit,
-  risk.table = T, cumevents = T, fun = "event", xlim = c(2003, 2018)
-)
+# Forest plot and coefficients table
+ggforest(cox_res,data = s)
+kable(tidy(cox_res))
+# Diagnostics
+test.ph= cox.zph(cox_res)
+ggcoxzph(test.ph)
+ggcoxdiagnostics(cox_res,type = "martingale")
 ```

### 2023-12-18 Christine Chan/CF Registry/cf_registry_analysis.qmd
CFRD registry first draft survival done

@@
   html:
+    fig-width: 10
+    fig-height: 10
     toc: true
@@
   Darwin = {
-    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Christine Chan/CF registry data"
+    home_dir <- "/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data"
   }
@@
     Age_YrEnd != 0,
-    is.na(first_cfrd) | ReviewYear <= first_cfrd
+    is.na(first_cfrd) | ReviewYear <= first_cfrd,
+    Age_YrEnd >= 10
   ) %>%
@@
 s$cfrd_status_annual <- as.numeric(s$cfrd_status_annual) - 1
+# Drop under 10s
+s$age_group <- droplevels(s$age_group)
 ```
@@
 km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ 1,
-  data = s,cluster = eDWID
+  data = s, cluster = eDWID
 )
@@
   data = s, risk.table = T, cumevents = T,
-  fun = "event", xlim = c(2003, 2018),break.x.by=1
+  fun = "event", xlim = c(2003, 2018), break.x.by = 1
 )
@@
 km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ age_group,
-  data = s,cluster = eDWID
+  data = s, cluster = eDWID
 )
@@
 
+### Base model
+
 ```{r}
@@
   "age_group", "Gender", "bmi_perc", "MutClass", "A_IsOnEnzymes",
-  "A_FEV1_pct_predicted", "tube_feeding", "NumPulmExacerbation",
-  "A_pulmonarycomplications1", "A_corticosteroids1", "liver_disease"
+  "tube_feeding", "liver_disease",
+  "A_FEV1_pct_predicted", "NumPulmExacerbation",
+  "A_pulmonarycomplications1", "A_corticosteroids1"
 )
@@
 # Forest plot and coefficients table
-ggforest(cox_res,data = s)
-kable(tidy(cox_res))
+ggforest(cox_res, data = s)
 # Diagnostics
-test.ph= cox.zph(cox_res)
-ggcoxzph(test.ph)
-ggcoxdiagnostics(cox_res,type = "martingale")
+# test.ph= cox.zph(cox_res)
+# ggcoxzph(test.ph)
+# ggcoxdiagnostics(cox_res,type = "martingale")
+# ggcoxdiagnostics(cox_res,type = "deviance")
+# ggcoxdiagnostics(cox_res,type = "score")
+# ggcoxdiagnostics(cox_res,type = "schoenfeld")
 ```
+
+#### P. aeruginosa
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```
+
+#### Aspergillus
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_fungalyeast1),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```
+
+#### ABPA
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```
+
+### Vx770
+
+```{r}
+f <- update(f, . ~ . + A_Vx770)
+cox_res <- coxph(f, data = s, cluster = eDWID)
+# Plot survival
+ggsurvplot(survfit(cox_res),
+  data = s, risk.table = T, cumevents = T, fun = "event",
+  xlim = c(2003, 2018), break.x.by = 1
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+# Diagnostics
+# test.ph= cox.zph(cox_res)
+# ggcoxzph(test.ph)
+# ggcoxdiagnostics(cox_res,type = "martingale")
+# ggcoxdiagnostics(cox_res,type = "deviance")
+# ggcoxdiagnostics(cox_res,type = "score")
+# ggcoxdiagnostics(cox_res,type = "schoenfeld")
+```
+
+#### P. aeruginosa
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```
+
+#### Aspergillus
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_fungalyeast1),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```
+
+#### ABPA
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```
+
+#### Mycobacterium
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_mycocultureresults3),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```
+
+### Vx809
+
+```{r}
+f <- update(f, . ~ . + A_VX809comb)
+cox_res <- coxph(f, data = s, cluster = eDWID)
+# Plot survival
+ggsurvplot(survfit(cox_res),
+  data = s, risk.table = T, cumevents = T, fun = "event",
+  xlim = c(2003, 2018), break.x.by = 1
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+# Diagnostics
+# test.ph= cox.zph(cox_res)
+# ggcoxzph(test.ph)
+# ggcoxdiagnostics(cox_res,type = "martingale")
+# ggcoxdiagnostics(cox_res,type = "deviance")
+# ggcoxdiagnostics(cox_res,type = "score")
+# ggcoxdiagnostics(cox_res,type = "schoenfeld")
+```
+
+#### P. aeruginosa
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```
+
+#### Aspergillus
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_fungalyeast1),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```
+
+#### ABPA
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```
+
+#### Mycobacterium
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_mycocultureresults3),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```

### 2023-12-18 Christine Chan/CF Registry/cf_registry_analysis.qmd
cfrd registry sensitivity analysis

@@
 
-# Survival analysis
-
-Based on CFFPR annualized data.
-These data are interval censored because we know the first visit at which people with CFRD were diagnosed, but we do not know the exact time that they developed CFRD.
-These data are also left truncated because not all participants enter the study at the same time.
-
 ## Participants with a single observation
@@
   summarise(largest_gap = max(diff(ReviewYear)) - 1)
+gap_ids = unique(largest_gap$eDWID[largest_gap$largest_gap>=4])
 largest_gap <- data.frame(table(largest_gap$largest_gap))
@@
 ```
+
+# Survival analysis - first CFRD diagnosis
+
+Based on CFFPR annualized data.
+These data are interval censored because we know the first visit at which people with CFRD were diagnosed, but we do not know the exact time that they developed CFRD.
+These data are also left truncated because not all participants enter the study at the same time.
 
@@
 ```
+
+# Survival analysis - second CFRD diagnosis (for those with $\geq$ 4 years between CFRD+ diagnoses)
+
+```{r}
+second_cfrd <- annual %>%
+  filter(eDWID %in% gap_ids,cfrd_status_annual=="CFRD+") %>%
+  group_by(eDWID) %>%
+  summarise(second_cfrd = ReviewYear[2])
+annual <- left_join(annual, second_cfrd, by = join_by(eDWID))
+annual$first_cfrd = pmax(annual$first_cfrd,annual$second_cfrd)
+```
+
+```{r}
+# For now, let's just assume that everyone's first CFRD+ diagnosis is correct
+s <- annual %>%
+  filter(
+    !is.na(cfrd_status_annual),
+    Age_YrEnd != 0,
+    is.na(first_cfrd) | ReviewYear <= first_cfrd,
+    Age_YrEnd >= 10
+  ) %>%
+  select(
+    eDWID, ReviewYear, birth_year, Age_YrEnd,
+    all_of(vars), cfrd_status_annual
+  ) %>%
+  rename(tstop = ReviewYear) %>%
+  group_by(eDWID) %>%
+  mutate(
+    tstart = lag(tstop, default = birth_year[1]),
+    age_at_entry = Age_YrEnd[1]
+  ) %>%
+  select(
+    eDWID, birth_year, tstart, tstop, cfrd_status_annual,
+    age_at_entry, everything()
+  )
+# Convert from tibble to dataframe for survival packages
+s <- as.data.frame(s)
+# CFRD status to numeric
+s$cfrd_status_annual <- as.numeric(s$cfrd_status_annual) - 1
+# Drop under 10s
+s$age_group <- droplevels(s$age_group)
+```
+
+## Kaplan Meier Analysis
+
+```{r}
+# Whole cohort
+km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ 1,
+  data = s, cluster = eDWID
+)
+ggsurvplot(km_fit,
+  data = s, risk.table = T, cumevents = T,
+  fun = "event", xlim = c(2003, 2018), break.x.by = 1
+)
+kable(tidy(km_fit))
+# By age group
+km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ age_group,
+  data = s, cluster = eDWID
+)
+ggsurvplot(km_fit,
+  data = s, risk.table = T, cumevents = T,
+  fun = "event", xlim = c(2003, 2018), break.x.by = 1
+)
+km_table <- tidy(km_fit)
+km_table <- split.data.frame(km_table, km_table$strata)
+invisible(lapply(km_table, function(t) {
+  name <- t$strata[1]
+  t <- t %>% select(-strata)
+  kable(t, caption = name)
+}))
+```
+
+## Cox PH model
+
+### Base model
+
+```{r}
+# Formula for base model
+x <- c(
+  "age_group", "Gender", "bmi_perc", "MutClass", "A_IsOnEnzymes",
+  "tube_feeding", "liver_disease",
+  "A_FEV1_pct_predicted", "NumPulmExacerbation",
+  "A_pulmonarycomplications1", "A_corticosteroids1"
+)
+y <- "Surv(tstart, tstop, cfrd_status_annual)"
+f <- as.formula(paste0(y, "~", paste0(x, collapse = "+")))
+cox_res <- coxph(f, data = s, cluster = eDWID)
+# Plot survival
+ggsurvplot(survfit(cox_res),
+  data = s, risk.table = T, cumevents = T, fun = "event",
+  xlim = c(2003, 2018), break.x.by = 1
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+# Diagnostics
+# test.ph= cox.zph(cox_res)
+# ggcoxzph(test.ph)
+# ggcoxdiagnostics(cox_res,type = "martingale")
+# ggcoxdiagnostics(cox_res,type = "deviance")
+# ggcoxdiagnostics(cox_res,type = "score")
+# ggcoxdiagnostics(cox_res,type = "schoenfeld")
+```
+
+#### P. aeruginosa
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```
+
+#### Aspergillus
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_fungalyeast1),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```
+
+#### ABPA
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```
+
+### Vx770
+
+```{r}
+f <- update(f, . ~ . + A_Vx770)
+cox_res <- coxph(f, data = s, cluster = eDWID)
+# Plot survival
+ggsurvplot(survfit(cox_res),
+  data = s, risk.table = T, cumevents = T, fun = "event",
+  xlim = c(2003, 2018), break.x.by = 1
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+# Diagnostics
+# test.ph= cox.zph(cox_res)
+# ggcoxzph(test.ph)
+# ggcoxdiagnostics(cox_res,type = "martingale")
+# ggcoxdiagnostics(cox_res,type = "deviance")
+# ggcoxdiagnostics(cox_res,type = "score")
+# ggcoxdiagnostics(cox_res,type = "schoenfeld")
+```
+
+#### P. aeruginosa
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```
+
+#### Aspergillus
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_fungalyeast1),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```
+
+#### ABPA
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```
+
+#### Mycobacterium
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_mycocultureresults3),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```
+
+### Vx809
+
+```{r}
+f <- update(f, . ~ . + A_VX809comb)
+cox_res <- coxph(f, data = s, cluster = eDWID)
+# Plot survival
+ggsurvplot(survfit(cox_res),
+  data = s, risk.table = T, cumevents = T, fun = "event",
+  xlim = c(2003, 2018), break.x.by = 1
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+# Diagnostics
+# test.ph= cox.zph(cox_res)
+# ggcoxzph(test.ph)
+# ggcoxdiagnostics(cox_res,type = "martingale")
+# ggcoxdiagnostics(cox_res,type = "deviance")
+# ggcoxdiagnostics(cox_res,type = "score")
+# ggcoxdiagnostics(cox_res,type = "schoenfeld")
+```
+
+#### P. aeruginosa
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```
+
+#### Aspergillus
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_fungalyeast1),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```
+
+#### ABPA
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```
+
+#### Mycobacterium
+
+```{r}
+# Update formula
+cox_res <- coxph(update(f, . ~ . + A_mycocultureresults3),
+  data = s, cluster = eDWID
+)
+# Forest plot and coefficients table
+ggforest(cox_res, data = s)
+```

### 2023-12-19 Christine Chan/CF Registry/cf_registry_analysis.qmd
reorganize home collection code

--- 2023-12-18 Christine Chan/CF Registry/cf_registry_analysis.qmd
@@
 
-# Survival analysis
+# Survival analysis - first CFRD diagnosis
 
@@
 These data are also left truncated because not all participants enter the study at the same time.
-
-## Participants with a single observation
-
-```{r}
-# Get total number of people with only one row with CFRD info
-one_row <- annual %>%
-  group_by(eDWID) %>%
-  summarise(n = n()) %>%
-  filter(n == 1)
-one_row_cfrd <- annual %>%
-  filter(!is.na(cfrd_status_annual)) %>%
-  group_by(eDWID) %>%
-  summarise(n = n()) %>%
-  filter(n == 1)
-```
-
-There were `r nrow(one_row)` people with a single year of observation in the annual data (across all variables), and `r nrow(one_row_cfrd)` people with a single non-missing CFRD status.
-`r length(intersect(one_row$eDWID,one_row_cfrd$eDWID))` people fall into both of these categories:
-
-```{r}
-plot(
-  euler(list(
-    "One Observation (All)" = one_row$eDWID,
-    "One Observation (CFRD)" = one_row_cfrd$eDWID
-  )),
-  fills = c("lightblue2", "lightsalmon"),
-  quantities = T,
-  legend = list(side = "right")
-)
-```
-
-## Participants with $\geq$ 3 years between CFRD+
-
-```{r}
-# I can't think of a good tidyverse way to do this, so trying it myself
-dfs <- annual %>% select(eDWID, ReviewYear, cfrd_status_annual)
-dfs <- split.data.frame(dfs, dfs$eDWID)
-# For each ID, test whether they follow the pattern
-fit_pattern <- lapply(dfs, function(d) {
-  r <- rle(as.character(d$cfrd_status_annual))
-  # Check for CFRD+, CFRD-, CFRD+ pattern
-  # Can have NAs on either side of the CFRD- as well, so
-  # CFRD+, CFRD-, CFRD+ OR CFRD+, NA, CFRD-, CFRD+ OR CFRD+, CFRD-, NA, CFRD+
-  # OR CFRD+, NA,CFRD-, NA,CFRD+
-  if (
-    grepl(
-      paste("CFRD\\+,CFRD\\-,CFRD\\+|", "CFRD\\+,NA,CFRD\\-,CFRD\\+|",
-        "CFRD\\+,CFRD\\-,NA,CFRD\\+|", "CFRD\\+,NA,CFRD\\-,NA,CFRD\\+",
-        sep = "", collapse = ""
-      ),
-      paste(r$values, collapse = ",")
-    )) {
-    return(unique(d$eDWID))
-  } else {
-    return(NA)
-  }
-})
-fit_pattern <- fit_pattern[!is.na(fit_pattern)]
-# Filter out those that don't follow the pattern, calculate largest gaps
-# between CFRD+ diagnoses
-largest_gap <- annual %>%
-  filter(eDWID %in% names(fit_pattern), cfrd_status_annual == "CFRD+") %>%
-  select(eDWID, ReviewYear, cfrd_status_annual) %>%
-  group_by(eDWID) %>%
-  summarise(largest_gap = max(diff(ReviewYear)) - 1)
-largest_gap <- data.frame(table(largest_gap$largest_gap))
-colnames(largest_gap) <- c("Years between CFRD+", "n")
-```
-
-There were a total of `r sum(largest_gap$n)` participants who had a diagnosis of CFRD+ followed by at least one CFRD-, followed again by a CFRD+ diagnosis.
-For each person, we calculated the largest gap between CFRD+ diagnoses.
-A total of `r sum(largest_gap$n[as.numeric(largest_gap[,1])>=3])` people had a gap $\geq$ 3 years.
-
-```{r}
-kable(largest_gap)
-```
 

### 2023-12-19 Jordana Hoppe/Home Collection/manuscript.qmd
reorganize home collection code

---
title: "Remote Collection of Respiratory Samples"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
csl: /Users/timvigers/Dropbox/Miscellaneous/american-medical-association.csl
editor: source
---

```{r libraries}
#| include: false
library(tidyverse)
library(lubridate)
library(gtsummary)
```

```{r data cleaning, include=FALSE}
# Use REDCap import
source("/Users/timvigers/GitHub/CF/Jordana Hoppe/Home Collection/home_collection_data.r")
# Fix missing labels
labels <- label(data)[which(label(data) != "")]
labels[paste0(names(labels), ".factor")] <- as.character(labels)
# Combine race columns
races <- c(
  "American Indian or Alaska Native", "Asian", "Black or African-American",
  "Native Hawaiian or Other Pacific Islander", "White", "More than one race",
  "Unknown or not reported"
)
data$race <- apply(data, 1, function(r) {
  w <- which(r[paste0("race___", 1:7)] == 1)
  paste0(races[w], collapse = "/")
})
# Age
data$age <- round(as.numeric(ymd(data$clinic_date) - ymd(data$pt_dob)) / 365.25, 1)
# Additional labels for tables
labels["race"] <- "Race"
labels["age"] <- "Age at Clinic Visit"
```

# Table

```{r}
#| label: tbl-1
#| tbl-cap: Participant Characteristics
data %>%
  select(
    age, pt_sex.factor, race, ethnicity.factor, mutation1.factor,
    mutation2.factor, rural_location.factor, fev1_pp, fvc_pp
  ) %>%
  mutate(
    mutation1.factor = droplevels(mutation1.factor),
    mutation2.factor = droplevels(mutation2.factor)
  ) %>%
  tbl_summary(
    label = labels, missing_text = "Missing",
    statistic = list(all_continuous() ~ "{mean} ({sd})")
  ) 
```

# Figure

```{r}
#| label: fig-planets
#| fig-cap: Planets
```

### 2023-12-19 Jordana Hoppe/Home Collection/home_collection_data.r

@@ -1,0 +1,287 @@
+# Load Hmisc library
+library(Hmisc)
+# Read Data
+data <- read.csv("/Users/timvigers/Dropbox/Work/Vigers/CF/Jordana Hoppe/Home Collection/Data_Raw/HomeCollection220661_DATA_2023-12-19_0917.csv")
+# Setting Labels
+label(data$subject_id) <- "Subject ID"
+label(data$pt_initials) <- "Patient Initials"
+label(data$pt_dob) <- "Patient DOB"
+label(data$pt_sex) <- "Sex"
+label(data$race___1) <- "Race (choice=American Indian or Alaska Native)"
+label(data$race___2) <- "Race (choice=Asian)"
+label(data$race___3) <- "Race (choice=Black or African-American)"
+label(data$race___4) <- "Race (choice=Native Hawaiian or Other Pacific Islander)"
+label(data$race___5) <- "Race (choice=White)"
+label(data$race___6) <- "Race (choice=More than one race)"
+label(data$race___7) <- "Race (choice=Unknown or not reported)"
+label(data$ethnicity) <- "Hispanic or Latinx"
+label(data$mutation1) <- "Mutation 1"
+label(data$mutation1_spec) <- "Specify mutation 1, if Other"
+label(data$tract_type1) <- "Tract type"
+label(data$mutation2) <- "Mutation 2"
+label(data$mutation2_spec) <- "Specify mutation 2, if Other"
+label(data$tract_type2) <- "Tract type "
+label(data$rural_location) <- "Rural location?"
+label(data$demographics_complete) <- "Complete?"
+label(data$clinic_date) <- "Date of clinic culture collection"
+label(data$clinic_cx_source) <- "Source of culture"
+label(data$clinic_negative_cx) <- "Negative culture?"
+label(data$clinic_cx_results___1) <- "If culture was positive, select all organisms isolated: (choice=Methicillin sensitive Staph aureus)"
+label(data$clinic_cx_results___2) <- "If culture was positive, select all organisms isolated: (choice=Methicillin resistant Staph aureus)"
+label(data$clinic_cx_results___16) <- "If culture was positive, select all organisms isolated: (choice=A. xylosoxidans)"
+label(data$clinic_cx_results___17) <- "If culture was positive, select all organisms isolated: (choice=Achromobacter not xylosoxidans)"
+label(data$clinic_cx_results___10) <- "If culture was positive, select all organisms isolated: (choice=Acinetobacter)"
[…]

### 2023-12-19 Jordana Hoppe/Home Collection/home_collection_surveys.r

@@ -1,0 +1,99 @@
+# Load Hmisc library
+library(Hmisc)
+# Read Data
+survey_data <- read.csv("/Users/timvigers/Dropbox/Work/Vigers/CF/Jordana Hoppe/Home Collection/Data_Raw/HomeCollection220661_DATA_2023-12-19_0911.csv")
+# Setting Labels
+label(survey_data$participant_id) <- "Participant ID"
+label(survey_data$redcap_survey_identifier) <- "Survey Identifier"
+label(survey_data$consent_timestamp) <- "Survey Timestamp"
+label(survey_data$parent_fname) <- "Your first name"
+label(survey_data$parent_lname) <- "Your last name"
+label(survey_data$email) <- "E-mail address"
+label(survey_data$consent___1) <- "By checking this box, I certify that I am at least 18 years old and that I give my consent freely to participate in this study. (choice=I consent)"
+label(survey_data$signature) <- "Please sign to indicate your consent to participate in this study."
+label(survey_data$consent_complete) <- "Complete?"
+label(survey_data$survey_timestamp) <- "Survey Timestamp"
+label(survey_data$q1) <- "Did you receive a home collection kit in the mail?"
+label(survey_data$q2) <- "How long did it take you to receive a home collection kit after your clinic visit?"
+label(survey_data$q3) <- "Were you able to obtain a sample at home? "
+label(survey_data$previous) <- "Have you previously collected a sample at home?"
+label(survey_data$goodsample) <- "Do you feel your home collection was a good sample? "
+label(survey_data$difficulty) <- "How would you rate the difficulty of the collection?"
+label(survey_data$barriers___1) <- "What barriers did you encounter with home collection?  (choice=Uncooperative child)"
+label(survey_data$barriers___2) <- "What barriers did you encounter with home collection?  (choice=Written instructions for collection were not clear)"
+label(survey_data$barriers___3) <- "What barriers did you encounter with home collection?  (choice=I did not think the sample I obtained was a good sample)"
+label(survey_data$barriers___4) <- "What barriers did you encounter with home collection?  (choice=Other, please specify: {barriers_spec})"
+label(survey_data$barriers_spec) <- "Please specify other barriers to home collection:"
[…]

### 2023-12-19 Jordana Hoppe/Home Collection/manuscript.qmd
home collection manuscript tables done

@@
 source("/Users/timvigers/GitHub/CF/Jordana Hoppe/Home Collection/home_collection_data.r")
+source("/Users/timvigers/GitHub/CF/Jordana Hoppe/Home Collection/home_collection_surveys.r")
 # Fix missing labels
@@
 labels[paste0(names(labels), ".factor")] <- as.character(labels)
+survey_labels <- label(survey_data)[which(label(survey_data) != "")]
+survey_labels[paste0(names(survey_labels), ".factor")] <- as.character(survey_labels)
+labels = c(labels,survey_labels)
 # Combine race columns
@@
 
-# Table
+# Table 1: Participant Characteristics
 
@@
     statistic = list(all_continuous() ~ "{mean} ({sd})")
-  ) 
+  ) %>%
+  as_kable_extra()
 ```
 
-# Figure
+It appears that the two participants who were previously marked as unknown race are now considered more than one race.
+
+# Table 2: Survey Results
 
 ```{r}
-#| label: fig-planets
-#| fig-cap: Planets
+#| label: tbl-2
+#| tbl-cap: Survey Results
+survey_data %>%
+  select(q1.factor:q6b___4.factor) %>%
+  tbl_summary(
+    label = labels, missing_text = "Missing"
+  ) %>%
+  as_kable_extra()
 ```
+
+# Table 3: Clinic Culture Results
+
+```{r}
+#| label: tbl-3
+#| tbl-cap: Clinic Culture Results
+data %>%
+  select(clinic_cx_source.factor:clinic_burk_sp.factor) %>%
+  tbl_summary(
+    label = labels, missing_text = "Missing"
+  ) %>%
+  as_kable_extra()
+```
+
+# Table 4: Research Culture Results
+
+```{r}
+#| label: tbl-4
+#| tbl-cap: Research Culture Results
+data %>%
+  select(research_cx.factor:res_burk_sp.factor) %>%
+  tbl_summary(
+    label = labels, missing_text = "Missing"
+  ) %>%
+  as_kable_extra()
+```
+
+For Yes/No variables in @tbl-2 through @tbl-4, the percentage of "Yes" answers is shown.

### 2023-12-19 Jordana Hoppe/Home Collection/home_collection_data.r

@@ -1,7 +1,7 @@
 # Load Hmisc library
 library(Hmisc)
 # Read Data
-data <- read.csv("/Users/timvigers/Dropbox/Work/Vigers/CF/Jordana Hoppe/Home Collection/Data_Raw/HomeCollection220661_DATA_2023-12-19_0917.csv")
+data <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Jordana Hoppe/Home Collection/Data_Raw/HomeCollection220661_DATA_2023-12-19_0917.csv")
 # Setting Labels
 label(data$subject_id) <- "Subject ID"
 label(data$pt_initials) <- "Patient Initials"

### 2023-12-19 Jordana Hoppe/Home Collection/home_collection_surveys.r

@@ -1,7 +1,7 @@
 # Load Hmisc library
 library(Hmisc)
 # Read Data
-survey_data <- read.csv("/Users/timvigers/Dropbox/Work/Vigers/CF/Jordana Hoppe/Home Collection/Data_Raw/HomeCollection220661_DATA_2023-12-19_0911.csv")
+survey_data <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Jordana Hoppe/Home Collection/Data_Raw/HomeCollection220661_DATA_2023-12-19_0911.csv")
 # Setting Labels
 label(survey_data$participant_id) <- "Participant ID"
 label(survey_data$redcap_survey_identifier) <- "Survey Identifier"

### 2023-12-19 Jordana Hoppe/Home Collection/manuscript.qmd
home collection basic euler diagrams

@@
 library(gtsummary)
+library(eulerr)
 ```
@@
 survey_labels[paste0(names(survey_labels), ".factor")] <- as.character(survey_labels)
-labels = c(labels,survey_labels)
+labels <- c(labels, survey_labels)
 # Combine race columns
@@
 data$age <- round(as.numeric(ymd(data$clinic_date) - ymd(data$pt_dob)) / 365.25, 1)
+# Find tests with no growth
+data$clinic_cx_no_growth <- apply(data, 1, function(r) {
+  sum(as.numeric(r[grep("clinic_cx_results___\\d$", names(r))]))
+})
+data$res_cx_no_growth <- apply(data, 1, function(r) {
+  sum(as.numeric(r[grep("res_cx_results___\\d$", names(r))]))
+})
+data$home_cx_no_growth <- apply(data, 1, function(r) {
+  sum(as.numeric(r[grep("home_cx_results___\\d$", names(r))]))
+})
 # Additional labels for tables
@@
 
-# Table 1: Participant Characteristics
+# Summary Tables
+
+## Table 1: Participant Characteristics
 
@@
 
-# Table 2: Survey Results
+## Table 2: Survey Results
 
@@
 
-# Table 3: Clinic Culture Results
+## Table 3: Clinic Culture Results
 
@@
 
-# Table 4: Research Culture Results
+## Table 4: Research Culture Results
 
@@
 
-For Yes/No variables in @tbl-2 through @tbl-4, the percentage of "Yes" answers is shown.
+## Table 5: Home Culture Results
+
+```{r}
+#| label: tbl-5
+#| tbl-cap: Home Culture Results
+data %>%
+  select(home_cx.factor:home_burk_sp.factor) %>%
+  tbl_summary(
+    label = labels, missing_text = "Missing"
+  ) %>%
+  as_kable_extra()
+```
+
+For Yes/No variables in @tbl-2 through @tbl-5, the percentage of "Yes" answers is shown.
+
+# Agreement
+
+## Euler Diagrams
+
+### MSSA
+
+```{r}
+data %>%
+  select(
+    clinic_cx_results___1.factor, res_cx_results___1.factor,
+    home_cx_results___1.factor
+  ) %>%
+  mutate(across(
+    everything(),
+    ~ factor(., levels = c("Checked", "Unchecked"), labels = c(1, 0))
+  )) %>%
+  mutate(across(
+    everything(), ~ as.numeric(as.character(.))
+  )) %>%
+  rename(
+    Clinic = clinic_cx_results___1.factor,
+    Research = res_cx_results___1.factor,
+    Home = home_cx_results___1.factor
+  ) %>%
+  euler(.) %>%
+  plot(., quantities = T)
+```
+
+### *H. flu*
+
+```{r}
+data %>%
+  select(
+    clinic_cx_results___3.factor, res_cx_results___3.factor,
+    home_cx_results___3.factor
+  ) %>%
+  mutate(across(
+    everything(),
+    ~ factor(., levels = c("Checked", "Unchecked"), labels = c(1, 0))
+  )) %>%
+  mutate(across(
+    everything(), ~ as.numeric(as.character(.))
+  )) %>%
+  rename(
+    Clinic = clinic_cx_results___3.factor,
+    Research = res_cx_results___3.factor,
+    Home = home_cx_results___3.factor
+  ) %>%
+  euler(.) %>%
+  plot(., quantities = T)
+```
+
+### No growth
+
+```{r}
+data %>%
+  select(clinic_cx_no_growth, res_cx_no_growth, home_cx_no_growth) %>%
+  mutate(across(
+    everything(), ~ .==0)
+  ) %>%
+  rename(
+    Clinic = clinic_cx_no_growth,
+    Research = res_cx_no_growth,
+    Home = home_cx_no_growth
+  ) %>%
+  euler(.) %>%
+  plot(., quantities = T)
+```

### 2023-12-19 StatNet/FDA/cgm_fda.qmd
reorganize StatNet folder

--- 2023-02-03 StatNet/cgm_fda.qmd
@@
 editor: source
+bibliography: FDA.bib
+csl: american-medical-association.csl
 ---
@@
 
-```{r}
-data(DTI)
-DTI = subset(DTI, select = c(cca, case, pasat))
-DTI = DTI[complete.cases(DTI),]
-DTI$gender = factor(sample(c("male","female"), dim(DTI)[1], replace = TRUE))
-DTI$status = factor(sample(c("RRMS", "SPMS", "PPMS"), dim(DTI)[1], replace = TRUE))
-## fit models
-dti_fosr_mod = bayes_fosr(cca ~ gender, data = DTI)
-```
+## Clustering and classification
+
+![](clusters.png)
+
+- $k$-means functional clustering is the most common approach. @wangFunctionalDataAnalysis2016
+  - Mean *functions* as cluster centers
+
+## Clustering via FPCA
+
+- Given $K$ FPCs, FPCA can be represented as:
+
+$$
+X_{iK}(t)=\mu(t)+\sum_{k=1}^K A_{ik}\phi_{k}(t)
+$$
+- ${A_{ik}}$ are the FPC scores and represent different cluster patterns.
+- Or, if you're using a basis expansion, you can use the projections onto the basis expansion space ${B_{ik}}$
+  - The advantage of FPCA is that the basis functions are determined by the data
+
+## Classification
+
+- Ideally, we would like to extend the clustering analysis to classification.
+- Functional linear discriminant analysis would allow us to apply Bayes' rule to new CGM tracings:
+
+$$
+Pr(Z=k|X_0)=\frac{\pi_k f_k(X_0)}{\sum_{j=1}^K\pi_j f_j(X_0)}
+$$
+
+- James and Hastie (2001) @jamesFunctionalLinearDiscriminant2001 proposed a method for this, @wangFunctionalDataAnalysis2016 but even the paper is difficult to access.
 
@@
 
--   Does this even seem like the correct approach?
--   Really what we want to do is classify people based on CGM tracing. Is there a good package for functional discriminant analysis?
--   FoSR in `refund` requires complete data, so how do we deal with irregular sampling across CGM tracings
+- Does this even seem like the correct approach? Is it worth the effort?
+- Is there a good package for functional discriminant analysis? 
+- FoSR in `refund` requires complete data, so how do we deal with irregular sampling across CGM tracings?
   - There is between and within-subject variability in terms length of sleep and therefore number of glucose measures
@@
   - Data formatting is a nightmare with the various FDA packages, and unfortunately `tidyfun` is pretty limited.
+
+## References

### 2024-01-02 Jordana Hoppe/Home Collection/manuscript.qmd
Jordana home swab manuscript analysis done

--- 2023-12-19 Jordana Hoppe/Home Collection/manuscript.qmd
@@
 library(gtsummary)
+library(irr)
+library(eulerr)
+library(kableExtra)
 ```
@@
 data$age <- round(as.numeric(ymd(data$clinic_date) - ymd(data$pt_dob)) / 365.25, 1)
+# Find tests with no growth
+data$clinic_cx_no_growth.factor <- apply(data, 1, function(r) {
+  sum(as.numeric(r[grep("clinic_cx_results___\\d$", names(r))]), na.rm = T)
+})
+data$clinic_cx_no_growth.factor <- factor(data$clinic_cx_no_growth.factor,
+  levels = 0:2,
+  labels = c("No Growth", "Some Growth", "Some Growth")
+)
+data$res_cx_no_growth.factor <- apply(data, 1, function(r) {
+  sum(as.numeric(r[grep("res_cx_results___\\d$", names(r))]), na.rm = T)
+})
+data$res_cx_no_growth.factor <- factor(data$res_cx_no_growth.factor,
+  levels = 0:2,
+  labels = c("No Growth", "Some Growth", "Some Growth")
+)
+data$home_cx_no_growth.factor <- apply(data, 1, function(r) {
+  sum(as.numeric(r[grep("home_cx_results___\\d$", names(r))]), na.rm = T)
+})
+data$home_cx_no_growth.factor <- factor(data$home_cx_no_growth.factor,
+  levels = 0:2,
+  labels = c("No Growth", "Some Growth", "Some Growth")
+)
 # Additional labels for tables
@@
 
-# Table 4: Research Culture Results
+## Table 4: Delayed Culture Results
 
@@
 #| label: tbl-4
-#| tbl-cap: Research Culture Results
+#| tbl-cap: Delayed Culture Results
 data %>%
@@
 
-For Yes/No variables in @tbl-2 through @tbl-4, the percentage of "Yes" answers is shown.
+## Table 5: Home Culture Results
+
+```{r}
+#| label: tbl-5
+#| tbl-cap: Home Culture Results
+data %>%
+  select(home_cx.factor:home_burk_sp.factor) %>%
+  tbl_summary(
+    label = labels, missing_text = "Missing"
+  ) %>%
+  as_kable_extra()
+```
+
+For Yes/No variables in @tbl-2 through @tbl-5, the percentage of "Yes" answers is shown.
+
+# Agreement
+
+Fleiss' kappa is interpreted analagously to Cohen's kappa, but applies to more than two tests. Tables with NaN and NAs indicate that all tests had the same values (i.e. all absent or all present).
+
+| Cohen's Kappa | Interpretation         |
+|---------------|------------------------|
+| 0             | No agreement           |
+| 0.10 - 0.20   | Slight agreement       |
+| 0.21 - 0.40   | Fair agreement         |
+| 0.41 - 0.60   | Moderate agreement     |
+| 0.61 - 0.80   | Substantial agreement  |
+| 0.81 - 0.99   | Near perfect agreement |
+| 1             | Perfect agreement      |
+
+```{r}
+agreement <- function(result_pattern, df = data) {
+  cols <- paste0(
+    c("clinic_cx_", "res_cx_", "home_cx_"),
+    paste0(result_pattern, ".factor")
+  )
+  d <- df[, cols]
+  colnames(d) <- c("Clinic", "Delayed", "Home")
+  d <- d %>%
+    mutate(across(everything(), ~ as.character(.)))
+  d[d == "NA"] <- NA
+  d[d == "Unchecked"] <- "Absent"
+  d[d == "Checked"] <- "Present"
+  fk <- kappam.fleiss(d)
+  cat("\n")
+  cat("### Agreement between all three methods")
+  cat("\n")
+  cat("\n")
+  cat("#### Fleiss' kappa")
+  cat("\n")
+  cat("\n")
+  f <- ftable(d)
+  f %>%
+    as.data.frame(.) %>%
+    kable(.) %>%
+    kable_styling(., "striped") %>%
+    print(.)
+  cat("\n")
+  cat("\n")
+  cat(
+    "Overall Fleiss' kappa for the three tests was", round(fk$value, 3),
+    paste0("(p = ", format.pval(fk$p.value, digits = 3, eps = 0.001), ")")
+  )
+  cat("\n")
+  cat("\n")
+  cat("#### Euler diagram")
+  cat("\n")
+  d %>%
+    mutate(across(everything(), ~ as.numeric(. == "Present" | . == "No Growth"))) %>%
+    drop_na(.) %>%
+    euler(.) %>%
+    plot(., quantities = T) %>%
+    print(.)
+  cat("\n")
+  cat("\n")
+  cat("### Pairwise agreement")
+  cat("\n")
+  clinic_vs_res <- xtabs(~ Clinic + Delayed, data = d)
+  clinic_vs_home <- xtabs(~ Clinic + Home, data = d)
+  home_vs_res <- xtabs(~ Home + Delayed, data = d)
+  cat("\n")
+  cat("\n")
+  cat("#### Clinic vs. delayed")
+  cat("\n")
+  cat("\n")
+  clinic_vs_res %>%
+    as.data.frame(.) %>%
+    kable(.) %>%
+    kable_styling("striped") %>%
+    print(.)
+  cat("\n")
+  clinic_vs_res %>%
+    prop.table(.) %>%
+    as.data.frame(.) %>%
+    rename("Percentage" = Freq) %>%
+    kable(., digits = 2) %>%
+    kable_styling("striped") %>%
+    print(.)
+  cat("\n")
+  cat("\n")
+  cat("#### Clinic vs. home")
+  cat("\n")
+  clinic_vs_home %>%
+    as.data.frame(.) %>%
+    kable(.) %>%
+    kable_styling("striped") %>%
+    print(.)
+  cat("\n")
+  clinic_vs_home %>%
+    prop.table(.) %>%
+    as.data.frame(.) %>%
+    rename("Percentage" = Freq) %>%
+    kable(., digits = 2) %>%
+    kable_styling("striped") %>%
+    print(.)
+  cat("\n")
+  cat("\n")
+  cat("#### Home vs. delayed")
+  cat("\n")
+  home_vs_res %>%
+    as.data.frame(.) %>%
+    kable(.) %>%
+    kable_styling("striped") %>%
+    print(.)
+  cat("\n")
+  home_vs_res %>%
+    prop.table(.) %>%
+    as.data.frame(.) %>%
+    rename("Percentage" = Freq) %>%
+    kable(., digits = 2) %>%
+    kable_styling("striped") %>%
+    print(.)
+  cat("\n")
+  cat("\n")
+}
+```
+
+## MSSA
+
+```{r results='asis'}
+agreement("results___1")
+```
+
+## MRSA
+
+```{r results='asis'}
+agreement("results___2")
+```
+
+## A. xylosoxidans
+
+```{r results='asis'}
+agreement("results___16")
+```
+
+## Achromobacter (not xylosoxidans)
+
+```{r results='asis'}
+agreement("results___17")
+```
+
+## Acinetobacter
+
+```{r results='asis'}
+agreement("results___10")
+```
+
+## H. influenzae
+
+```{r results='asis'}
+agreement("results___3")
+```
+
+## P. aeruginosa
+
+```{r results='asis'}
+agreement("results___4")
+```
+
+## P. aeruginosa (mucoid)
+
+```{r results='asis'}
+agreement("results___5")
+```
+
+## Pseudomonas species (not aeruginosa)
+
+```{r results='asis'}
+agreement("results___6")
+```
+
+## B. cepacia complex
+
+```{r results='asis'}
+agreement("results___7")
+```
+
+## Stenotrophomonas
+
+```{r results='asis'}
+agreement("results___8")
+```
+
+## Aspergillus
+
+```{r results='asis'}
+agreement("results___9")
+```
+
+## Penicillium species
+
+```{r results='asis'}
+agreement("results___12")
+```
+
+## Sceosporium species
+
+```{r results='asis'}
+agreement("results___13")
+```
+
+## M. avium complex
+
+```{r results='asis'}
+agreement("results___14")
+```
+
+## M. abscessus
+
+```{r results='asis'}
+agreement("results___15")
+```
+
+## No growth
+
+```{r results='asis'}
+agreement("no_growth")
+```

### 2024-01-02 StatNet/LMM Power/statnet_lmm_power.qmd
StatNet power presentation

---
title: "Power and Sample Size in General Linear Mixed Models"
format: revealjs
bibliography: zotero.bib
csl: american-medical-association.csl
---

```{r include=FALSE}
library(tidyverse)
```

## A quick refresher on terminology

- Type 1 error rate ($\alpha$): the probability that we incorrectly reject the null hypothesis ($H_0$).
- Type 2 error rate ($\beta$): the probability that we incorrectly fail to reject $H_0$.
- Power ($P$): the probability that we correctly reject $H_0$.
  - Can also be written as $1-\beta$

## A quick refresher on terminology

<br>
<br>

|                     | $H_0$ True  |$H_0$ False|
| ------------------- | ----------- |---------- |
| Reject $H_0$        | $\alpha$    | $1-\beta$ |
| Do Not Reject $H_0$ | $1-\alpha$  | $\beta$   |

## The difference between two independent means

<br>

![](t_test_indep_means.jpg) @faulStatisticalPowerAnalyses2009

## Next slide

- As the deviation from the null hypothesis ($d$) approaches 0, $P$ approaches $\alpha$. @diggleAnalysisLongitudinalData2013
- We typically choose $d$ to be the smallest clinically meaningful difference that we want to detect.
- Measurement variation ($\sigma^2$): the variance of the response.
  - Ideally approximated through previous studies, but sometimes estimated based on clinical knowledge.

## References

### 2024-01-02 Quarto Template.qmd
quarto template

@@
 
-```{r libraries}
+```{r setup}
 #| include: false
 library(tidyverse)
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
+)
+knitr::opts_knit$set(root.dir = home_dir)
 ```
@@
 #| tbl-cap: Planets
-
 ```

### 2024-01-02 Scott Sagel/PEx in CF/analysis_for_yaron.qmd
analysis for Yaron started

---
title: "Pulmonary Exacerbations in CF"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
csl: /Users/timvigers/Dropbox/Miscellaneous/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(lubridate)
library(gtsummary)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
# Import
df <- read.csv("./Scott Sagel/PEx in CF/Data_Raw/ExacerbationsInCF222_DATA_2024-01-02_0857.csv",
  na.strings = ""
)
# Get admit year (some missing in admit_year field)
df$year <- year(df$admit_date)
# Labels
labels <- list(
  "Aspergillus fumigatus" = "fungi___1",
  "Aspergillus (other species)" = "fungi___2",
  "Scedosporium apiospermum" = "fungi___3",
  "Candidia (any species)" = "fungi___4",
  "Other fungus" = "fungi___5",
  "NTM testing performed?" = "ntm_test",
  "M. abscessus species" = "ntm___1",
  "M. avium complex species" = "ntm___2",
  "Other NTM species" = "ntm___3",
  "A. xylosoxidans" = "bacteria_6months___1",
  "B. cepacia complex" = "bacteria_6months___2",
  "H. influenzae" = "bacteria_6months___3",
  "MRSA" = "bacteria_6months___4",
  "MSSA" = "bacteria_6months___5",
  "P. aeruginosa" = "bacteria_6months___6",
  "S. maltophilia" = "bacteria_6months___7",
  "Other bacteria" = "bacteria_6months___8"
)
labels <- setNames(names(labels), labels)
# Format
t <- df %>%
  select(
    sid, redcap_event_name, year, redcap_data_access_group,
    fungi___1:fungi___5,
    ntm_test, ntm___1:ntm___3,
    bacteria_6months___1:bacteria_6months___8
  ) %>%
  filter(!is.na(year))
```

```{r}
#| label: tbl-1
#| tbl-cap: Variable names
t %>%
  select(-sid, -redcap_event_name) %>%
  tbl_summary(by = year, missing_text = "Missing") %>%
  add_overall()
```

```{r}
#| label: tbl-2
#| tbl-cap: Labels
t %>%
  select(-sid, -redcap_event_name) %>%
  tbl_summary(
    by = year, missing_text = "Missing",
    label = labels
  ) %>%
  add_overall()
```

```{r include=FALSE}
# Write data to share with Yaron and Elin
write.csv(t,
  file = "./Scott Sagel/PEx in CF/Data_Cleaned/data_clean.csv",
  row.names = F, na = ""
)
```

### 2024-01-02 Scott Sagel/PEx in CF/nacfc_2023.qmd
analysis for Yaron started

@@
 source("~/GitHub/CF/Scott Sagel/PEx in CF/create_pex_analysis_dataset.R")
-load("~/Library/CloudStorage/Dropbox/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Cleaned/analysis_dataset.RData")
+load("~/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Scott Sagel/PEx in CF/Data_Cleaned/analysis_dataset.RData")
 # Colors for plots

### 2024-01-02 Scott Sagel/PEx in CF/create_pex_analysis_dataset.R

@@ -1,7 +1,7 @@
 library(tidyverse)
 library(rspiro)
 # Import REDCap data
-df <- readxl::read_excel("~/Library/CloudStorage/Dropbox/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Raw/ExacerbationsInCF222_DATA_2023-10-11_1252 All Data.xlsx")
+df <- readxl::read_excel("~/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Scott Sagel/PEx in CF/Data_Raw/ExacerbationsInCF222_DATA_2023-10-11_1252 All Data.xlsx")
 # Separate datasets
 demo <- df %>%
   filter(redcap_event_name == "demographics_arm_1") %>%
@@ -164,5 +164,5 @@ labels <- list(
 df <- df %>% filter(!is.na(admit_year))
 # Save
 save(df, summary_table, labels,
-  file = "~/Dropbox/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Cleaned/analysis_dataset.RData"
+  file = "~/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Scott Sagel/PEx in CF/Data_Cleaned/analysis_dataset.RData"
 )

### 2024-01-02 Scott Sagel/PEx in CF/analysis_for_yaron.qmd
combined labels

@@
 )
-labels <- setNames(names(labels), labels)
+labels <- setNames(paste0(names(labels)," (",labels,")"), labels)
 # Format
@@
     sid, redcap_event_name, year, redcap_data_access_group,
-    fungi___1:fungi___5,
+    fungal_test,fungi___1:fungi___5,
     ntm_test, ntm___1:ntm___3,
@@
 #| tbl-cap: Variable names
-t %>%
-  select(-sid, -redcap_event_name) %>%
-  tbl_summary(by = year, missing_text = "Missing") %>%
-  add_overall()
-```
-
-```{r}
-#| label: tbl-2
-#| tbl-cap: Labels
 t %>%

### 2024-01-02 Scott Sagel/PEx in CF/analysis_for_yaron.qmd
initial analysis for Yaron finished

@@
 df$year <- year(df$admit_date)
+# Rename centers
+df$redcap_data_access_group <- factor(df$redcap_data_access_group,
+  levels = c("chco", "childrens_national", "orange_county", "rady", "seattle"),
+  labels = c("CHCO", "Children's National", "Orange County", "Rady", "Seattle")
+)
+df$cftr_mod <- factor(df$cftr_mod, levels = 0:1, labels = c("No", "Yes"))
 # Labels
@@
   "Other NTM species" = "ntm___3",
-  "A. xylosoxidans" = "bacteria_6months___1",
-  "B. cepacia complex" = "bacteria_6months___2",
-  "H. influenzae" = "bacteria_6months___3",
-  "MRSA" = "bacteria_6months___4",
-  "MSSA" = "bacteria_6months___5",
-  "P. aeruginosa" = "bacteria_6months___6",
-  "S. maltophilia" = "bacteria_6months___7",
-  "Other bacteria" = "bacteria_6months___8"
+  "A. xylosoxidans" = "admit_cx_results___1",
+  "B. cepacia complex" = "admit_cx_results___2",
+  "H. influenzae" = "admit_cx_results___3",
+  "MRSA" = "admit_cx_results___4",
+  "Methicillin-susceptible Staph aureus" = "admit_cx_results___5",
+  "P. aeruginosa" = "admit_cx_results___6",
+  "P. aeruginosa mucoid" = "admit_cx_results___7",
+  "S. maltophilia" = "admit_cx_results___8",
+  "Other bacteria" = "admit_cx_results___9",
+  "CFTR Modulator" = "cftr_mod",
+  "Cultures Tested" = "redcap_data_access_group"
 )
-labels <- setNames(paste0(names(labels)," (",labels,")"), labels)
+labels <- setNames(names(labels), labels)
 # Format
-t <- df %>%
+df <- df %>%
   select(
-    sid, redcap_event_name, year, redcap_data_access_group,
-    fungal_test,fungi___1:fungi___5,
+    sid, redcap_event_name, redcap_data_access_group, year, cftr_mod,
+    fungal_test, fungi___1:fungi___5,
     ntm_test, ntm___1:ntm___3,
-    bacteria_6months___1:bacteria_6months___8
+    admit_cx_results___1:admit_cx_results___9
   ) %>%
@@
 
+# Fungi
+
+The tables below are limited to patients who had fungal testing performed.
+
 ```{r}
-#| label: tbl-1
-#| tbl-cap: Variable names
-t %>%
-  select(-sid, -redcap_event_name) %>%
-  tbl_summary(
-    by = year, missing_text = "Missing",
-    label = labels
-  ) %>%
-  add_overall()
+#| label: tbl-fungi
+#| tbl-cap: Fungi
+df %>%
+  filter(fungal_test == 1) %>%
+  select(year, redcap_data_access_group, fungi___1:fungi___5) %>%
+  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes()
 ```
 
-```{r include=FALSE}
-# Write data to share with Yaron and Elin
-write.csv(t,
-  file = "./Scott Sagel/PEx in CF/Data_Cleaned/data_clean.csv",
-  row.names = F, na = ""
+```{r}
+#| label: tbl-fungi_center
+#| tbl-cap: Fungi by Center
+#| message: false
+df %>%
+  filter(fungal_test == 1) %>%
+  select(year, redcap_data_access_group, fungi___1:fungi___5) %>%
+  tbl_strata(
+    strata = redcap_data_access_group,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
   )
 ```
+
+```{r}
+#| label: tbl-fungi_cftr_mod
+#| tbl-cap: Fungi by Modulator Status
+#| message: false
+df %>%
+  filter(fungal_test == 1) %>%
+  select(year, cftr_mod, fungi___1:fungi___5) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = cftr_mod, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+# NTM
+
+The tables below are limited to patients who had NTM testing performed.
+
+```{r}
+#| label: tbl-ntm
+#| tbl-cap: NTM
+df %>%
+  filter(ntm_test == 1) %>%
+  select(year, redcap_data_access_group, ntm___1:ntm___3) %>%
+  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes()
+```
+
+```{r}
+#| label: tbl-ntm_center
+#| tbl-cap: NTM by Center
+#| message: false
+df %>%
+  filter(ntm_test == 1) %>%
+  select(year, redcap_data_access_group, ntm___1:ntm___3) %>%
+  tbl_strata(
+    strata = redcap_data_access_group,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-ntm_cftr_mod
+#| tbl-cap: NTM by Modulator Status
+#| message: false
+df %>%
+  filter(ntm_test == 1) %>%
+  select(year, cftr_mod, ntm___1:ntm___3) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = cftr_mod, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+# Bacteria
+
+```{r}
+#| label: tbl-bacteria
+#| tbl-cap: Bacteria
+df %>%
+  select(
+    year, redcap_data_access_group, admit_cx_results___1:admit_cx_results___9
+  ) %>%
+  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes()
+```
+
+```{r}
+#| label: tbl-bacteria_center
+#| tbl-cap: Bacteria by Center
+#| message: false
+df %>%
+  select(
+    year, redcap_data_access_group, admit_cx_results___1:admit_cx_results___9
+  ) %>%
+  tbl_strata(
+    strata = redcap_data_access_group,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-bacteria_cftr_mod
+#| tbl-cap: Bacteria by Modulator Status
+#| message: false
+df %>%
+  select(year, cftr_mod, admit_cx_results___1:admit_cx_results___9) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = cftr_mod, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```

### 2024-01-02 Quarto Template.qmd
Update Quarto Template.qmd

@@
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Windows"="C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF"
 )

### 2024-01-02 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
Update racial_and_ethnic_differences_in_eti.qmd

@@
 library(ggsurvfit)
+# Paths
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Windows"="C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF"
+)
+knitr::opts_knit$set(root.dir = home_dir)
 ```
@@
 # source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R")
-load("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
+load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
 ```
@@
 For the following analysis, the start time ($t=0$) was the approximate date that each person became eligible for ETI.
-Because we don't have exact dates of birth, I calculated an approximate DOB based on age and encounter dates.
+Because we don't have exact dates of birth, I calculated an approximate DOB based on age and encounter dates, and then used the approximate DOB to determine date of eligibility.
 Encounters prior to ETI eligibility were excluded, and only time to first encounter on ETI was considered (i.e.
@@
 ```{r}
-#| cache: true
 # Most basic model of lung function over time by race

### 2024-01-03 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
race and ETI segmented analysis done

@@
       dark: darkly
-bibliography: /Users/timvigers/Library/CloudStorage/Dropbox/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Library/CloudStorage/Dropbox/Miscellaneous/american-medical-association.csl
+bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
+csl: /Users/timvigers/Dropbox/Miscellaneous/american-medical-association.csl
 editor: source
@@
 # Libraries
+library(segmented)
 library(Hmisc)
@@
   "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Windows"="C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF"
+  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF"
 )
@@
 
-# Lung function over time
+# Lung function
+
+## GLI-2012 vs. GLI-2022
+
+GLI-2012 spirometry equations include race/ethnicity in the formula, but GLI-2022 equations are race-free.
+The results of the GLI-2022 equations contain the "gl" suffix, which stands for "global" (so instead of "ppFEV1" it would be "ppFEV1gl").
+
+First, I wanted to ensure that the `rspiro` package is calculating FEV1 percent predicted correctly.
+In the graph below, ppFEV1 from the registry is shown on the x-axis and ppFEV1 as calculated with `rspiro` is shown on the y-axis.
+
+```{r}
+#| warning: false
+ggplot(encounter, aes(
+  x = GLI_FEV1_pct_predicted, y = GLI_FEV1_pct_predicted_check,
+  color = race
+)) +
+  geom_point() +
+  geom_abline(slope = 1, intercept = 0) +
+  theme_bw() +
+  xlab("Registry") +
+  ylab("rspiro") +
+  ylim(0, 150) +
+  facet_wrap(~race) +
+  ggtitle("Registry values vs. rspiro")
+```
+
+```{r}
+ggplot(encounter, aes(
+  x = ppFEV1gl, y = GLI_FEV1_pct_predicted_check,
+  color = race
+)) +
+  geom_point() +
+  geom_abline(slope = 1, intercept = 0) +
+  theme_bw() +
+  xlab("GLI-2022") +
+  ylab("GLI-2012") +
+  ylim(0, 150) +
+  xlim(0, 150) +
+  facet_wrap(~race) +
+  ggtitle("GLI-2022 vs. GLI-2012 (rspiro)")
+```
 
@@
 label(means) <- list("Race/Ethnicity", "Year", "Mean ppFEV", "SE", "df", "Lower CL", "Upper CL")
-gt(means) %>% fmt_number(emmean:last_col(), decimals = 3)
+gt(means) %>% fmt_number(emmean:last_col(), decimals = 1)
 ```
@@
 ggplot(left_join(annual, means, by = join_by(race, ReviewYear))) +
+  geom_point(mapping = aes(x = ReviewYear, y = emmean, color = race)) +
+  geom_line(mapping = aes(x = ReviewYear, y = emmean, group = race, color = race)) +
+  geom_errorbar(mapping = aes(
+    x = ReviewYear, ymin = emmean - SE,
+    ymax = emmean + SE, color = race
+  ), width = .2) +
+  xlab("Year") +
+  ylab("LS Mean FEV1") +
+  theme_bw() +
+  scale_color_discrete(name = "Race/Ethnicity")
+```
+
+```{r}
+#| warning: false
+ggplot(left_join(annual, means, by = join_by(race, ReviewYear))) +
+  geom_line(
+    mapping = aes(x = ReviewYear, y = mean_ppFEV, group = eDWID),
+    alpha = 0.005, linewidth = 0.5
+  ) +
   geom_point(mapping = aes(x = ReviewYear, y = emmean, color = race)) +
@@
 ggplot(left_join(annual, means, by = join_by(race, ReviewYear))) +
-  geom_line(mapping = aes(x = ReviewYear, y = mean_ppFEV, group = eDWID), alpha = 0.005, linewidth = 0.5) +
+  geom_line(
+    mapping = aes(x = ReviewYear, y = mean_ppFEV, group = eDWID),
+    alpha = 0.005, linewidth = 0.5
+  ) +
   geom_point(mapping = aes(x = ReviewYear, y = emmean, color = race)) +
-  geom_line(mapping = aes(x = ReviewYear, y = emmean, group = race, color = race)) +
-  geom_errorbar(mapping = aes(x = ReviewYear, ymin = emmean - SE, ymax = emmean + SE, color = race), width = .2) +
-  xlab("Year") +
-  ylab("LS Mean FEV1") +
-  theme_bw() +
-  scale_color_discrete(name = "Race/Ethnicity")
-```
-
-```{r}
-#| warning: false
-ggplot(left_join(annual, means, by = join_by(race, ReviewYear))) +
-  geom_line(mapping = aes(x = ReviewYear, y = mean_ppFEV, group = eDWID), alpha = 0.005, linewidth = 0.5) +
-  geom_point(mapping = aes(x = ReviewYear, y = emmean, color = race)) +
-  geom_line(mapping = aes(x = ReviewYear, y = emmean, group = race, color = race)) +
-  geom_errorbar(mapping = aes(x = ReviewYear, ymin = emmean - SE, ymax = emmean + SE, color = race), width = .2) +
+  geom_line(mapping = aes(
+    x = ReviewYear, y = emmean, group = race,
+    color = race
+  )) +
+  geom_errorbar(mapping = aes(
+    x = ReviewYear, ymin = emmean - SE,
+    ymax = emmean + SE, color = race
+  ), width = .2) +
   xlab("Year") +
@@
 ```{r}
+# Calculate time from first ETI use
+encounter <- encounter %>%
+  mutate(
+    days_from_start =
+      as.numeric(encounterdate - Modulator_trikafta_first_date)
+  )
+# Segmented regression by group
+lm_fixed_cp <- segreg(
+  GLI_FEV1_pct_predicted ~ race + seg(days_from_start, by = race),
+  data = encounter
+)
+# Results
+lm_fixed_cp$psi
+slope(lm_fixed_cp)
+# Plot
+new_df = lm_fixed_cp$model
+new_df$pred = predict(lm_fixed_cp)
+ggplot(new_df, aes(x = days_from_start, y = pred,group = race,color = race)) +
+  geom_line() +
+  ylim(0, 150)
+# # Random intercept version isn't converging
+# # Basic model
+# lmm_base <- lme(GLI_FEV1_pct_predicted ~ days_from_start,
+#   random = ~ 1 | eDWID, data = encounter, na.action = na.omit,method = "ML",
+#   control=list(msMaxIter=1000,msMaxEval = 1000)
+# )
+# # Segmented with fixed change point
+# lmm_fixed_cp <- segmented(lmm_base,
+#   seg.Z = ~days_from_start, psi = 0,
+#   random = list(eDWID = pdDiag(~ 1 + days_from_start + U))
+# )
+```
+
+```{r}
 #| cache: true
 #| eval: false
-t <- annual %>%
-  select(eDWID, ReviewYear, mean_ppFEV) %>%
-  na.omit(.)
+t <- encounter %>%
+  select(eDWID, GLI_FEV1_pct_predicted, days_from_start) %>%
+  drop_na()
+
 model <- list(
-  mean_ppFEV ~ 0 + ReviewYear,
-  ~ 0 + ReviewYear
+  GLI_FEV1_pct_predicted ~ 0 + days_from_start,
+  ~ 0 + days_from_start
 )
@@
 ```
-
-```{r}
-ggplot(annual, aes(x = Age_YrEnd, y = mean_ppFEV)) +
-  geom_smooth(aes(color = race), se = F)
-```

### 2024-01-03 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R

@@ -4,11 +4,11 @@ library(readxl)
 library(rspiro)
 library(Hmisc)
 # Import
-annual <- read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_Annualized_Del1.csv", na.strings = "")
-encounter <- read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_encountersMerged_Del1.csv", na.strings = "")
-demo <- read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_DemogCFDiag_Del1.csv", na.strings = "")
+annual <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_Annualized_Del1.csv", na.strings = "")
+encounter <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_encountersMerged_Del1.csv", na.strings = "")
+demo <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_DemogCFDiag_Del1.csv", na.strings = "")
 # Import responsive mutations list
-responsive_mutations <- read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/eti_responsive_mutations.csv", header = F)
[…]

### 2024-01-03 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
Create phase_1_analysis.qmd

---
title: "SSwAB-IT Phase 1 Analysis"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(redcapAPI)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows"="C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}

```

# Table

```{r}
#| label: tbl-planets
#| tbl-cap: Planets
```

# Figure

```{r}
#| label: fig-planets
#| fig-cap: Planets
```

### 2024-01-03 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
Update racial_and_ethnic_differences_in_eti.qmd

@@
       dark: darkly
-bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Dropbox/Miscellaneous/american-medical-association.csl
+bibliography: C:/Users/timvigers/Dropbox/Miscellaneous/zotero.bib
+csl: C:/Users/timvigers/Dropbox/Miscellaneous/american-medical-association.csl
 editor: source
@@
   ylim(0, 150)
-# # Random intercept version isn't converging
+# # Random intercept model isn't converging
 # # Basic model
@@
 #| cache: true
-#| eval: false
 t <- encounter %>%
@@
   drop_na()
-
 model <- list(

### 2024-01-03 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
SSwAB-IT prelim and ETI update

@@
 library(tidyverse)
+library(gtsummary)
 library(redcapAPI)
@@
 ```{r data cleaning}
-
+#| include: false
+# Import from REDCap
+unlockREDCap(c(rcon = "SSwAB-IT (New)"),
+  keyring = "API_KEYs",
+  envir = 1,
+  url = "https://redcap.ucdenver.edu/api/"
+)
+exportBulkRecords(list(db = rcon), envir = 1)
 ```
 
-# Table
+# Feasibility
 
 ```{r}
-#| label: tbl-planets
-#| tbl-cap: Planets
+#| message: false
+db_feasibility %>%
+  select(
+    redcap_data_access_group, able_to_collect_sample,
+    believe_collected_good_sample, sample_difficulty, willing_to_repeat,
+  ) %>%
+  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
+  add_overall()
 ```
-
-# Figure
-
-```{r}
-#| label: fig-planets
-#| fig-cap: Planets
-```

### 2024-01-03 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
SSwAB-IT prelim and ETI update

@@
       dark: darkly
-bibliography: C:/Users/timvigers/Dropbox/Miscellaneous/zotero.bib
-csl: C:/Users/timvigers/Dropbox/Miscellaneous/american-medical-association.csl
+bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
+csl: /Users/timvigers/Dropbox/Miscellaneous/american-medical-association.csl
 editor: source
@@
 #| cache: true
+#| eval: false
 t <- encounter %>%

### 2024-01-05 StatNet/LMM Power/statnet_lmm_power.qmd
statnet LMM power presentation almost done

@@
 title: "Power and Sample Size in General Linear Mixed Models"
-format: revealjs
+format: 
+  revealjs:
+    self-contained: true
 bibliography: zotero.bib
@@
 library(tidyverse)
+library(simr)
 ```
@@
 
-- Type 1 error rate ($\alpha$): the probability that we incorrectly reject the null hypothesis ($H_0$).
-- Type 2 error rate ($\beta$): the probability that we incorrectly fail to reject $H_0$.
-- Power ($P$): the probability that we correctly reject $H_0$.
+- Type 1 error rate ($\alpha$): 
+  - The probability that we incorrectly reject the null hypothesis ($H_0$).
+- Type 2 error rate ($\beta$): 
+  - The probability that we incorrectly fail to reject $H_0$.
+- Power ($P$): 
+  - The probability that we correctly reject $H_0$.
   - Can also be written as $1-\beta$
@@
 
-![](t_test_indep_means.jpg) @faulStatisticalPowerAnalyses2009
+![](t_test_indep_means.jpg)
 
-## Next slide
+## Power basics
 
-- As the deviation from the null hypothesis ($d$) approaches 0, $P$ approaches $\alpha$. @diggleAnalysisLongitudinalData2013
+- As the deviation from the null hypothesis ($d$) approaches 0, $P$ approaches $\alpha$.
 - We typically choose $d$ to be the smallest clinically meaningful difference that we want to detect.
-- Measurement variation ($\sigma^2$): the variance of the response.
+- Measurement variation ($\sigma^2$): 
+  - The variance of the response.
   - Ideally approximated through previous studies, but sometimes estimated based on clinical knowledge.
 
+## Power in longitudinal studies
+
+- Number of repeated observations per person ($n$): 
+  - The number of observations per person.
+  - May be constrained by practical matters like budget or participant burden.
+- Correlation among the repeated observations ($\rho$): 
+  - The pattern of correlation among repeated observations.
+  - Can sometimes be estimated based on previous studies.
+
+## Cough monitor study
+
+- Participants provided with AI-based cough monitor for 3 months.
+  - Can distinguish between coughs and other noises.
+- Cough frequency assessed by cough episodes per hour and cough seconds per hour.
+- Compare the frequency of nighttime cough in study participants with CF versus healthy controls.
+  - Controls will most likely be siblings of participants with CF.
+  - This is the aim that we are powering on.
+- Also interested in feasibility and the relationship between nighttime cough frequency and respiratory symptom scores.
+
+## Power for a continuous outcome
+
+- Given a fairly simple statistical model:
+
+$$
+Y_{ij} = \beta_0+\beta_1 x_i+\epsilon_{ij} \\
+j=1,...,n \\ 
+i=1,...,2m
+$$
+
+  - $x_i$ is the treatment indicator for the $ith$ subject.
+  - Assume that $Var(\epsilon_{ij})=\sigma^2$ and $Corr(Y_{ij},Y_{ik})=\rho$ for all $j\neq k$. 
+  - Also assume that each person has the same set of explanatory variables ($x_{ij}=x_{j}$).
+
+## Power for a continuous outcome
+
+- The required number of subjects per group is:
+
+$$
+m=\frac{2(z_\alpha+z_{1-P})^2\sigma^2\{1+(n-1)\rho\}}{nd^2}=\\
+$$
+
+$$
+\frac{2(z_\alpha+z_{1-P})^2\{1+(n-1)\rho\}}{n(\frac{d}{\sigma})^2}
+$$
+
+- Need to fill in $\rho$, $d$, and $\sigma^2$
+
+## Previous studies
+
+- Unfortunately there are very few previous studies on this.
+  - Only found one study that reported variability of night time coughs in CF.
+- According to Van Der Giessen et al.@vandergiessenNocturnalCoughChildren2009 the distributions of outcomes are skewed, so they report medians and IQR.
+  - For the purposes of this power calculation we are assuming normality since we don't have access to the data.
+- They report: 
+  - 0.9 csec/hr (IQR 0.3-6.1) the first night and 0.96 csec/hr (IQR 0.1-3.4) the second night
+  - Correlation between night 1 and night 2 was 0.29 for cough seconds and 0.35 for csec/hr.
+- Assume that controls will not cough and pwCF will cough approximately 0.9-1 times per hour (same for csec/hr). Also assume that $\sigma = \frac{IQR}{1.35}$ and $\rho=0.3$
+
+# Amalia's awesome workbook (demonstration)
+
+# GLIMMPSE demo
+
+## SIMR
+
+- Unfortunately not as easy to apply to this scenario since we don't have access to the prior data, but possible.
+- First, set up a basic data matrix with 10 participants (one pwCF and one control from 5 families). Each person starts with 10 nights of data.
+- Use the model:
+
+$$
+y∼group+time+group*time+(1|family/id)+\epsilon
+$$
+
+  - $(1|family/id)$ indicates random effects for participant nested within family.
+
+```{r}
+# Set up X matrix with 10 nights of data for 5 families
+nights <- 1:10
+groups <- c("Control", "CF")
+families <- as.character(1:5)
+X <- expand.grid(time = nights, group = groups, family = families)
+X$id <- paste0(X$group, X$family)
+# Set simulation parameters
+b <- c(0, 8, 0, 0) # Intercept and effect of night and group
+V1 <- list(0.5, 2) # Random effect variance
+# Create glmer object
+glmer_obj <- makeGlmer(y ~ group * time + (1 | family / id),
+  family = "poisson", fixef = b, VarCorr = V1, data = X
+)
+glmer_obj
+```
+
+```{r}
+#| cache: true
+powerSim(glmer_obj, nsim = 1000, test = "group", progress = F)
+```
+
+```{r}
+#| cache: true
+# Stable variables
+nights <- 1:10
+groups <- c("Control", "CF")
+b <- c(0, 8, 0, 0) # Intercept and effect of night and group
+V1 <- list(0.5, 2) # Random effect variance
+# Alter sample size (number of families) and plot results
+n <- lapply(2:100, function(f) {
+  # Set up data
+  families <- as.character(1:f)
+  X <- expand.grid(time = nights, group = groups, family = families)
+  X$id <- paste0(X$group, X$family)
+  # Create glmer object
+  glmer_obj <- makeGlmer(y ~ group * time + (1 | family / id),
+    family = "poisson", fixef = b, VarCorr = V1, data = X
+  )
+  ps = powerSim(glmer_obj, nsim = 1000, test = "group", progress = F)
+})
+
+```
+
+## Questions for the group
+
+- Do these assumptions seem reasonable given the limited prior knowledge?
+- Does anyone have code they'd be willing to share for these simulations?
+- Suggestions or software recommendations?
+
 ## References
+
+---
+nocite: |
+  @faulStatisticalPowerAnalyses2009, @diggleAnalysisLongitudinalData2013, @greenSIMRPackagePower2016
+---

### 2024-01-05 StatNet/LMM Power/statnet_lmm_power.qmd
LMM power presentation 99% done

@@
 ```{r include=FALSE}
+library(knitr)
 library(tidyverse)
@@
 - Cough frequency assessed by cough episodes per hour and cough seconds per hour.
+- First part of the study is feasibility.
+- Also interested in the relationship between nighttime cough frequency and respiratory symptom scores.
+
+## Cough monitor study
+
 - Compare the frequency of nighttime cough in study participants with CF versus healthy controls.
@@
   - This is the aim that we are powering on.
-- Also interested in feasibility and the relationship between nighttime cough frequency and respiratory symptom scores.
+- We have repeated measures for each participant, plus an additional layer of clustering by family.
+  - We expect some correlation between siblings.
 
-## Power for a continuous outcome
+## Power for a continuous outcome {.scrollable}
 
@@
 
-- Need to fill in $\rho$, $d$, and $\sigma^2$
+- Need to fill in $\rho$, $d$, and $\sigma^2$ based on prior knowledge.
 
@@
   - Only found one study that reported variability of night time coughs in CF.
-- According to Van Der Giessen et al.@vandergiessenNocturnalCoughChildren2009 the distributions of outcomes are skewed, so they report medians and IQR.
+- According to Van Der Giessen et al. @vandergiessenNocturnalCoughChildren2009 the distributions of outcomes are skewed, so they report medians and IQR.
   - For the purposes of this power calculation we are assuming normality since we don't have access to the data.
+
+## Previous studies {.scrollable}
+
 - They report: 
@@
   - Correlation between night 1 and night 2 was 0.29 for cough seconds and 0.35 for csec/hr.
-- Assume that controls will not cough and pwCF will cough approximately 0.9-1 times per hour (same for csec/hr). Also assume that $\sigma = \frac{IQR}{1.35}$ and $\rho=0.3$
 
-# Amalia's awesome workbook (demonstration)
+- We assume that controls will not cough @munyardHowMuchCoughing1996 and pwCF will cough approximately 0.9-1 times per hour (same for csec/hr).
+- Correlation between successive nights is $\rho=0.3$.
+- We use $\sigma = \frac{IQR}{1.35}$
+  - This feels a little hand-wavy so please suggest alternatives.
+
+# Amalia's awesome workbook
 
@@
 
-- Unfortunately not as easy to apply to this scenario since we don't have access to the prior data, but possible.
+- Unfortunately not as easy to apply to this scenario since we don't have access to the prior data, but theoretically possible.
 - First, set up a basic data matrix with 10 participants (one pwCF and one control from 5 families). Each person starts with 10 nights of data.
-- Use the model:
+
+## SIMR setup
+
+- Using the model:
 
@@
 
+## SIMR setup
+
 ```{r}
-# Set up X matrix with 10 nights of data for 5 families
-nights <- 1:10
+#| echo: TRUE
+# Set up X matrix with 5 nights of data for 5 families
+nights <- 1:5
 groups <- c("Control", "CF")
@@
 X$id <- paste0(X$group, X$family)
+```
+
+## SIMR setup {.scrollable}
+
+```{r}
+kable(head(X,20))
+```
+
+## SIMR setup {.scrollable}
+
+```{r}
+#| echo: TRUE
 # Set simulation parameters
-b <- c(0, 8, 0, 0) # Intercept and effect of night and group
+b <- c(0, 1, 0, 0) # Intercept and effects of night and group
 V1 <- list(0.5, 2) # Random effect variance
@@
 
+## SIMR results
+
 ```{r}
+#| echo: TRUE
 #| cache: true
 powerSim(glmer_obj, nsim = 1000, test = "group", progress = F)
-```
-
-```{r}
-#| cache: true
-# Stable variables
-nights <- 1:10
-groups <- c("Control", "CF")
-b <- c(0, 8, 0, 0) # Intercept and effect of night and group
-V1 <- list(0.5, 2) # Random effect variance
-# Alter sample size (number of families) and plot results
-n <- lapply(2:100, function(f) {
-  # Set up data
-  families <- as.character(1:f)
-  X <- expand.grid(time = nights, group = groups, family = families)
-  X$id <- paste0(X$group, X$family)
-  # Create glmer object
-  glmer_obj <- makeGlmer(y ~ group * time + (1 | family / id),
-    family = "poisson", fixef = b, VarCorr = V1, data = X
-  )
-  ps = powerSim(glmer_obj, nsim = 1000, test = "group", progress = F)
-})
-
 ```
@@
 - Do these assumptions seem reasonable given the limited prior knowledge?
-- Does anyone have code they'd be willing to share for these simulations?
-- Suggestions or software recommendations?
+- Is anyone familiar with `SIMR`? Would you be willing to share some code?
+- Other general suggestions or software recommendations?
 

### 2024-01-18 Quarto Template.qmd
updated paths

--- 2024-01-02 Quarto Template.qmd
@@
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
+  "Darwin" = "/Users/timvigers/Dropbox/Work/Vigers/CF",
+  "Windows" = "C:/Users/timvigers/Dropbox/Work/Vigers/CF"
 )
@@
 ```{r data cleaning}
-
 ```

### 2024-01-18 Christine Chan/CF Registry/cf_registry_analysis.qmd
Update cf_registry_analysis.qmd

@@
   Windows = {
-    home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Christine Chan/CF registry data"
+    home_dir <- "C:/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data"
   },
@@
 # Diagnostics
-# test.ph= cox.zph(cox_res)
-# ggcoxzph(test.ph)
+test.ph= cox.zph(cox_res)
+ggcoxzph(test.ph)
 # ggcoxdiagnostics(cox_res,type = "martingale")
@@
 # ggcoxdiagnostics(cox_res,type = "schoenfeld")
+```
+
+### Time varying coefficients
+
+```{r}
+s_split <- survSplit(Surv(tstart, tstop, cfrd_status_annual) ~ .,
+  data = s, cut = 2003:2018, episode = "tgroup"
+)
+cox_res_split <- coxph(update(f, . ~ Gender:strata(tgroup)), data = s_split)
+t = tidy(cox_res_split)
 ```

### 2024-01-22 Emily Muther/mental_health_screening_psc.qmd
mental health CF PSC

---
title: "Mental Health Screening in CF"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(plotly)
library(gtsummary)
library(GLMMadaptive)
library(lmerTest)
library(broom.mixed)
library(knitr)
library(sjPlot)
library(report)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| warning: false
# Import
df <- read.csv("./Emily Muther/General - Muther Research Group/Mental Health Screening/Data_Raw/PSC_Full_Data_data_11-11-22.csv",
  check.names = F, na.strings = ""
)
# Remove duplicate rows
df <- df %>% distinct()
# Sort by contact date
df$CONTACT_DATE <- lubridate::mdy(df$CONTACT_DATE)
df <- df %>% arrange(PAT_MRN_ID, CONTACT_DATE)
```

# Table 1: Demographics

```{r table 1}
cat_demo_vars <- c(
  "DEPARTMENT_NAME", "ENC_T", "PAYOR_NAME"
)
cont_demo_vars <- c("AGE AT CONTACT")
df %>%
  group_by(PAT_MRN_ID) %>%
  summarise(
    across(all_of(cont_demo_vars), ~ first(na.omit(.x))),
    across(all_of(cat_demo_vars), ~ first(na.omit(.x)))
  ) %>%
  select(-PAT_MRN_ID) %>%
  tbl_summary(missing_text = "Missing",label = list("Age at First Contact" = "AGE AT CONTACT"))
```

# Table 2: Prevalence of Behavioral Health Difficulties

```{r table 2}

```

### 2024-01-22 Jordana Hoppe/Home Collection/manuscript.qmd
home manuscript missing data

@@
 source("/Users/timvigers/GitHub/CF/Jordana Hoppe/Home Collection/home_collection_surveys.r")
+# Find missing subjects
+data_missing <- data %>%
+  filter(is.na(home_cx_source)) %>%
+  select(subject_id, pt_initials, home_cx:home_cx_results___17)
+survey_missing <- survey_data %>%
+  filter(is.na(q1)) %>%
+  select(participant_id, q1:q6b_spec)
+write.csv(data_missing, file = "~/missing_data.csv", row.names = F, na = "")
+write.csv(survey_missing, file = "~/missing_survey.csv", row.names = F, na = "")
 # Fix missing labels

### 2024-01-22 Jordana Hoppe/Home Collection/home_collection_data.r

@@ -1,7 +1,7 @@
 # Load Hmisc library
 library(Hmisc)
 # Read Data
-data <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Jordana Hoppe/Home Collection/Data_Raw/HomeCollection220661_DATA_2023-12-19_0917.csv")
+data <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Jordana Hoppe/Home Collection/Data_Raw/HomeCollection220661_DATA_2023-12-19_0917.csv",na.strings = "")
 # Setting Labels
 label(data$subject_id) <- "Subject ID"
 label(data$pt_initials) <- "Patient Initials"

### 2024-01-22 Jordana Hoppe/Home Collection/home_collection_surveys.r

@@ -1,7 +1,7 @@
 # Load Hmisc library
 library(Hmisc)
 # Read Data
-survey_data <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Jordana Hoppe/Home Collection/Data_Raw/HomeCollection220661_DATA_2023-12-19_0911.csv")
+survey_data <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Jordana Hoppe/Home Collection/Data_Raw/HomeCollection220661_DATA_2023-12-19_0911.csv",na.strings = "")
 # Setting Labels
 label(survey_data$participant_id) <- "Participant ID"
 label(survey_data$redcap_survey_identifier) <- "Survey Identifier"

### 2024-01-26 Christine Chan/CF Registry/cf_registry_analysis.qmd
added cluster argument
PSC

@@
   Windows = {
-    home_dir <- "C:/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data"
+    home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry"
   },
   Darwin = {
-    home_dir <- "/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data"
+    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry"
   }
@@
 )
-cox_res_split <- coxph(update(f, . ~ Gender:strata(tgroup)), data = s_split)
-t = tidy(cox_res_split)
+cox_res_split <- coxph(update(f, . ~ Gender:strata(tgroup)), data = s_split,
+                       cluster = eDWID)
+t = tidy(cox_res_split,exponentiate = T)
 ```

### 2024-01-26 Emily Muther/mental_health_screening_psc.qmd
PSC

@@
 ---
-title: "Mental Health Screening in CF"
+title: "Mental Health Screening in CF - PSC Screening"
 author: "Tim Vigers"
@@
 library(tidyverse)
-library(plotly)
 library(gtsummary)
-library(GLMMadaptive)
-library(lmerTest)
-library(broom.mixed)
-library(knitr)
-library(sjPlot)
-library(report)
+library(ggcorrplot)
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
   "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF"
@@
 df <- df %>% distinct()
-# Sort by contact date
+# Sort by ID and contact date
 df$CONTACT_DATE <- lubridate::mdy(df$CONTACT_DATE)
 df <- df %>% arrange(PAT_MRN_ID, CONTACT_DATE)
+# Exclude those with no PSC data
+df <- df %>% filter(!is.na(Score))
 ```
 
-# Table 1: Demographics
+# Data cleaning
+
+1. Duplicate rows were removed (n=21).
+2. All rows with missing data for `Score` were excluded (n=2319).
+3. In Table 1, demographics and first contact date are reported.
+
+# Demographics
 
@@
   select(-PAT_MRN_ID) %>%
-  tbl_summary(missing_text = "Missing",label = list("Age at First Contact" = "AGE AT CONTACT"))
+  tbl_summary(missing_text = "Missing", label = list("Age at First Contact" = "AGE AT CONTACT"))
 ```
 
-# Table 2: Prevalence of Behavioral Health Difficulties
+# Prevalence of behavioral health difficulties
+
+## All PSC screeners
 
 ```{r table 2}
+df %>%
+  select(
+    `Internalizing subscale`:`Externalizing subscale`, Score,
+    `PSC Attention`:`PSC Totals`
+  ) %>%
+  tbl_summary(
+    missing_text = "Missing",
+    type = list(`Internalizing subscale` ~ "continuous")
+  )
+```
 
+## By person
+
+The maximum score on each subscale was calculated for all participants.
+
+```{r}
+#| warning: false
+df %>%
+  mutate(across(`PSC Attention`:`PSC Totals`, ~ factor(.x, ordered = T))) %>%
+  group_by(PAT_MRN_ID) %>%
+  summarise(across(c(
+    `Internalizing subscale`:`Externalizing subscale`, Score,
+    `PSC Attention`:`PSC Totals`
+  ), ~ max(.x, na.rm = T))) %>%
+  mutate(across(
+    c(`Internalizing subscale`:`Externalizing subscale`, Score),
+    ~ case_when(.x == -Inf ~ NA, .default = .x)
+  )) %>%
+  select(-PAT_MRN_ID) %>%
+  tbl_summary(
+    missing_text = "Missing",
+    type = list(`Internalizing subscale` ~ "continuous")
+  )
 ```
+
+# Correlations
+
+```{r}
+X = df %>% 
+```

### 2024-01-29 Emily Muther/mental_health_screening_psc.qmd
PSC screening rough draft done

@@
 library(gtsummary)
-library(ggcorrplot)
+library(lme4)
+library(GLMMadaptive)
+library(plotly)
+library(knitr)
 home_dir <- switch(Sys.info()["sysname"],
@@
 df <- df %>% filter(!is.na(Score))
+# Make visit number
+df <- df %>%
+  group_by(PAT_MRN_ID) %>%
+  mutate(visit_num = row_number()) %>%
+  ungroup()
 ```
@@
   select(-PAT_MRN_ID) %>%
-  tbl_summary(missing_text = "Missing", label = list("Age at First Contact" = "AGE AT CONTACT"))
+  tbl_summary(
+    missing_text = "Missing",
+    label = list("Age at First Contact" = "AGE AT CONTACT")
+  )
 ```
@@
 
-# Correlations
+# Mixed models
 
-```{r}
-X = df %>% 
+Because some participants had multiple PSC scores, we used mixed effects models to account for correlation within each participant.
+Also, due to the large number of 0 scores, we used a zero-inflated negative binomial model to examine the association between PSC score and age.
+
+## Continuous scores
+
+### Total score
+
+```{r age vs. total psc}
+# Plot
+p <- ggplot(df, aes(
+  x = AGE_AT_CONTACT_IN_YEARS, y = Score,
+  group = factor(PAT_MRN_ID), color = factor(PAT_MRN_ID)
+)) +
+  geom_point() +
+  geom_line() +
+  theme_bw() +
+  theme(legend.position = "none")
+ggplotly(p)
+# Model
+gm <- mixed_model(Score ~ AGE_AT_CONTACT_IN_YEARS,
+  random = ~ 1 | PAT_MRN_ID, data = df,
+  family = zi.poisson(), zi_fixed = ~1
+)
+kable(summary(gm)$coef_table, digits = 3)
 ```
+
+### Internalizing subscale
+
+```{r age vs. internalizing psc}
+# Plot
+p <- ggplot(df, aes(
+  x = AGE_AT_CONTACT_IN_YEARS, y = `Internalizing subscale`,
+  group = factor(PAT_MRN_ID), color = factor(PAT_MRN_ID)
+)) +
+  geom_point() +
+  geom_line() +
+  theme_bw() +
+  theme(legend.position = "none")
+ggplotly(p)
+# Model
+gm <- mixed_model(`Internalizing subscale` ~ AGE_AT_CONTACT_IN_YEARS,
+  random = ~ 1 | PAT_MRN_ID, data = df,
+  family = zi.poisson(), zi_fixed = ~1, control = list(iter_EM = 0)
+)
+kable(summary(gm)$coef_table, digits = 3)
+```
+
+### Attention subscale
+
+```{r age vs. attention psc}
+# Plot
+p <- ggplot(df, aes(
+  x = AGE_AT_CONTACT_IN_YEARS, y = `Attention subscale`,
+  group = factor(PAT_MRN_ID), color = factor(PAT_MRN_ID)
+)) +
+  geom_point() +
+  geom_line() +
+  theme_bw() +
+  theme(legend.position = "none")
+ggplotly(p)
+# Model
+gm <- mixed_model(`Attention subscale` ~ AGE_AT_CONTACT_IN_YEARS,
+  random = ~ 1 | PAT_MRN_ID, data = df,
+  family = zi.poisson(), zi_fixed = ~1, control = list(iter_EM = 0)
+)
+kable(summary(gm)$coef_table, digits = 3)
+```
+
+### Externalizing subscale
+
+```{r age vs. externalizing psc}
+# Plot
+p <- ggplot(df, aes(
+  x = AGE_AT_CONTACT_IN_YEARS, y = `Externalizing subscale`,
+  group = factor(PAT_MRN_ID), color = factor(PAT_MRN_ID)
+)) +
+  geom_point() +
+  geom_line() +
+  theme_bw() +
+  theme(legend.position = "none")
+ggplotly(p)
+# Model
+gm <- mixed_model(`Externalizing subscale` ~ AGE_AT_CONTACT_IN_YEARS,
+  random = ~ 1 | PAT_MRN_ID, data = df,
+  family = zi.poisson(), zi_fixed = ~1, control = list(iter_EM = 0)
+)
+kable(summary(gm)$coef_table, digits = 3)
+```
+
+## Low vs. high risk
+
+### Total score
+
+```{r age vs. total psc categorical}
+# Model
+gm <- glmer(factor(`PSC Totals`) ~ AGE_AT_CONTACT_IN_YEARS + (1 | PAT_MRN_ID),
+  data = df, family = "binomial"
+)
+kable(summary(gm)$coefficients, digits = 3)
+```
+
+### Internalizing subscale
+
+```{r age vs. internalizing psc categorical}
+# Model
+gm <- glmer(
+  factor(`PSC Internalizing`) ~ AGE_AT_CONTACT_IN_YEARS + (1 | PAT_MRN_ID),
+  data = df, family = "binomial"
+)
+kable(table(factor(df$`PSC Internalizing`), df$AGE_AT_CONTACT_IN_YEARS))
+```
+
+The above model is not converging due to low number of high risk scores on this subscale.
+
+### Attention subscale
+
+```{r age vs. attention psc categorical}
+# Model
+gm <- glmer(
+  factor(`PSC Attention`) ~ AGE_AT_CONTACT_IN_YEARS + (1 | PAT_MRN_ID),
+  data = df, family = "binomial"
+)
+kable(summary(gm)$coefficients, digits = 3)
+```
+
+### Externalizing subscale
+
+```{r age vs. externalizing psc categorical}
+# Model
+gm <- glmer(
+  factor(`PSC Externalizing`) ~ AGE_AT_CONTACT_IN_YEARS + (1 | PAT_MRN_ID),
+  data = df, family = "binomial"
+)
+kable(summary(gm)$coefficients, digits = 3)
+```
+
+# Questions for Emily
+
+1. How do we handle people with multiple screenings at the same age?

### 2024-01-29 Christine Chan/CF Registry/cf_registry_analysis.qmd
formatting

@@
 # Diagnostics
-test.ph= cox.zph(cox_res)
+test.ph <- cox.zph(cox_res)
 ggcoxzph(test.ph)
@@
 )
-cox_res_split <- coxph(update(f, . ~ Gender:strata(tgroup)), data = s_split,
-                       cluster = eDWID)
-t = tidy(cox_res_split,exponentiate = T)
+cox_res_split <- coxph(update(f, . ~ Gender:strata(tgroup)),
+  data = s_split,
+  cluster = eDWID
+)
+t <- tidy(cox_res_split, exponentiate = T)
 ```

### 2024-01-29 Quarto Template.qmd
formatting

--- 2024-01-02 Quarto Template.qmd
@@
 library(tidyverse)
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
+switch(Sys.info()[["sysname"]],
+  Windows = {
+    home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry"
+  },
+  Darwin = {
+    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry"
+  }
 )

### 2024-01-29 Jordana Hoppe/Home Collection/manuscript.qmd
home collection file paths

@@
 library(kableExtra)
+switch(Sys.info()[["sysname"]],
+  Windows = {
+    home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
+  },
+  Darwin = {
+    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
+  }
+)
+knitr::opts_knit$set(root.dir = home_dir)
 ```

### 2024-01-29 Quarto Template.qmd
home collection file paths

--- 2024-01-18 Quarto Template.qmd
@@
 library(tidyverse)
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Dropbox/Work/Vigers/CF",
-  "Windows" = "C:/Users/timvigers/Dropbox/Work/Vigers/CF"
+switch(Sys.info()[["sysname"]],
+  Windows = {
+    home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
+  },
+  Darwin = {
+    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
+  }
 )
@@
 ```{r data cleaning}
+
 ```

### 2024-01-29 Jordana Hoppe/Home Collection/home_collection_data.r

@@ -1,7 +1,17 @@
 # Load Hmisc library
 library(Hmisc)
+# Set WD
+switch(Sys.info()[["sysname"]],
+       Windows = {
+         home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
+       },
+       Darwin = {
+         home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
+       }
+)
+setwd(home_dir)
 # Read Data
-data <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Jordana Hoppe/Home Collection/Data_Raw/HomeCollection220661_DATA_2023-12-19_0917.csv",na.strings = "")
+data <- read.csv("./Jordana Hoppe/Home Collection/Data_Raw/HomeCollection220661_DATA_2023-12-19_0917.csv",na.strings = "")
 # Setting Labels
 label(data$subject_id) <- "Subject ID"
 label(data$pt_initials) <- "Patient Initials"

### 2024-01-29 Jordana Hoppe/Home Collection/home_collection_surveys.r

@@ -1,7 +1,17 @@
 # Load Hmisc library
 library(Hmisc)
+# Set WD
+switch(Sys.info()[["sysname"]],
+       Windows = {
+         home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
+       },
+       Darwin = {
+         home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
+       }
+)
+setwd(home_dir)
 # Read Data
-survey_data <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Jordana Hoppe/Home Collection/Data_Raw/HomeCollection220661_DATA_2023-12-19_0911.csv",na.strings = "")
+survey_data <- read.csv("./Jordana Hoppe/Home Collection/Data_Raw/HomeCollection220661_DATA_2023-12-19_0911.csv",na.strings = "")
 # Setting Labels
 label(survey_data$participant_id) <- "Participant ID"
 label(survey_data$redcap_survey_identifier) <- "Survey Identifier"

### 2024-01-29 Jordana Hoppe/Home Collection/manuscript.qmd
excluding missing subjects home collection

@@
 source("/Users/timvigers/GitHub/CF/Jordana Hoppe/Home Collection/home_collection_surveys.r")
-# Find missing subjects
-data_missing <- data %>%
-  filter(is.na(home_cx_source)) %>%
-  select(subject_id, pt_initials, home_cx:home_cx_results___17)
-survey_missing <- survey_data %>%
-  filter(is.na(q1)) %>%
-  select(participant_id, q1:q6b_spec)
-write.csv(data_missing, file = "~/missing_data.csv", row.names = F, na = "")
-write.csv(survey_missing, file = "~/missing_survey.csv", row.names = F, na = "")
+# Exclude missing subjects
+data <- data %>%
+  filter(!subject_id %in% c("009-07", "009-17", "009-25", "009-37"))
+survey_data <- survey_data %>%
+  filter(!participant_id %in% c(15, 27, 30, 34, 47))
 # Fix missing labels

### 2024-01-29 Jordana Hoppe/Home Collection/home_collection_data.r

@@ -2,16 +2,15 @@
 library(Hmisc)
 # Set WD
 switch(Sys.info()[["sysname"]],
-       Windows = {
-         home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
-       },
-       Darwin = {
-         home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
-       }
+  Windows = {
+    home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
+  },
+  Darwin = {
+    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
+  }
 )
-setwd(home_dir)
 # Read Data
-data <- read.csv("./Jordana Hoppe/Home Collection/Data_Raw/HomeCollection220661_DATA_2023-12-19_0917.csv",na.strings = "")
+data <- read.csv(paste0(home_dir, "/", "Jordana Hoppe/Home Collection/Data_Raw/HomeCollection220661_DATA_2024-01-29_0945.csv"))
 # Setting Labels
 label(data$subject_id) <- "Subject ID"
 label(data$pt_initials) <- "Patient Initials"

### 2024-01-29 Jordana Hoppe/Home Collection/home_collection_surveys.r

@@ -2,16 +2,15 @@
 library(Hmisc)
 # Set WD
 switch(Sys.info()[["sysname"]],
-       Windows = {
-         home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
-       },
-       Darwin = {
-         home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
-       }
+  Windows = {
+    home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
+  },
+  Darwin = {
+    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
+  }
 )
-setwd(home_dir)
 # Read Data
-survey_data <- read.csv("./Jordana Hoppe/Home Collection/Data_Raw/HomeCollection220661_DATA_2023-12-19_0911.csv",na.strings = "")
+survey_data <- read.csv(paste0(home_dir, "/", "Jordana Hoppe/Home Collection/Data_Raw/HomeCollection220661_DATA_2024-01-29_0948.csv"))
 # Setting Labels
 label(survey_data$participant_id) <- "Participant ID"
 label(survey_data$redcap_survey_identifier) <- "Survey Identifier"

### 2024-01-29 Jordana Hoppe/Home Collection/manuscript.qmd
genotype to 3 categories

@@
 labels["age"] <- "Age at Clinic Visit"
+# Genotype format
+data$Genotype <- paste0(
+  data$mutation1.factor == "F508del", "/", data$mutation2.factor == "F508del"
+)
+data$Genotype <- factor(data$Genotype,
+  levels = c("TRUE/TRUE", "TRUE/FALSE", "FALSE/FALSE"),
+  labels = c("F508del/F508del", "F508del/Other", "Other/Other")
+)
 ```
@@
     age, pt_sex.factor, race, ethnicity.factor, mutation1.factor,
-    mutation2.factor, rural_location.factor, fev1_pp, fvc_pp
+    mutation2.factor, Genotype, rural_location.factor, fev1_pp, fvc_pp
   ) %>%

### 2024-01-31 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
kmeans for sweat test analysis

--- 2023-02-21 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
@@
 library(tidyverse)
+library(factoextra)
 library(pROC)
 library(knitr)
-knitr::opts_chunk$set(echo = FALSE)
-if (Sys.info()["sysname"] == "Windows") {
-  home_dir <- "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Test"
-  source("C:/Users/timvigers/Documents/GitHub/CF/get_genotype_class_severity.R")
-} else if (Sys.info()["sysname"] == "Linux") {
-  home_dir <- "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Edith Zemanick/Sweat Test"
-  source("~/GitHub/CF/get_genotype_class_severity.R")
-} else if (Sys.info()["sysname"] == "Darwin") {
-  home_dir <- "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Edith Zemanick/Sweat Test"
-  source("~/GitHub/CF/genotype_class_severity.R")
+source("/Users/timvigers/GitHub/CF/genotype_class_severity.R")
+switch(Sys.info()[["sysname"]],
+  Windows = {
+    home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Test"
+  },
+  Darwin = {
+    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Test"
   }
+)
 knitr::opts_knit$set(root.dir = home_dir)
-rm(home_dir)
 ```
@@
 df <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv", na.strings = "")
-n1 = nrow(df)
-n2 = length(unique(df$PatientID))
+n1 <- nrow(df)
+n2 <- length(unique(df$PatientID))
 # Remove exclusions from Angela
@@
 df$Diagnosis <- factor(diagnoses$Diagnosis[match(df$PatientID, diagnoses$PatientID)])
-levels(df$Diagnosis) = c("CFTR-RD or CRMS","CFTR-RD or CRMS","Cystic Fibrosis","Non-CF")
+levels(df$Diagnosis) <- c("CFTR-RD or CRMS", "CFTR-RD or CRMS", "Cystic Fibrosis", "Non-CF")
 # Add modulator info
@@
 # Filter per Maxie's code
-exclude2 = c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793)
-no_dem = unique(df$PatientID[df$Demographics.Available!="TRUE" | is.na(df$Demographics.Available) | df$Sex==""])
+exclude2 <- c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793)
+no_dem <- unique(df$PatientID[df$Demographics.Available != "TRUE" | is.na(df$Demographics.Available) | df$Sex == ""])
 df <- df %>%
@@
 df$`Na/Cl Ratio` <- df$Na / df$Cl
-high_ratio = unique(which(df$`Na/Cl Ratio`>2))
-high_cl = unique(which(df$Cl>160))
-high_na = unique(which(df$Na>160))
-df = df[-unique(c(high_ratio,high_cl,high_na)),]
+high_ratio <- unique(which(df$`Na/Cl Ratio` > 2))
+high_cl <- unique(which(df$Cl > 160))
+high_na <- unique(which(df$Na > 160))
+df <- df[-unique(c(high_ratio, high_cl, high_na)), ]
 # Sweat chloride group
-df$`Sweat Chloride Group` = cut(df$Cl,c(-Inf,30,60,Inf),right = F,
-                                labels = c("< 30","30-59",">60"))
+df$`Sweat Chloride Group` <- cut(df$Cl, c(-Inf, 30, 60, Inf),
+  right = F,
+  labels = c("< 30", "30-59", ">60")
+)
 # Sort
@@
 # Table options (most variables are not normally distributed)
-mycontrols <- tableby.control(numeric.test = "kwt", numeric.stats = c("Nmiss", "median", "q1q3"))
+mycontrols <- tableby.control(
+  numeric.test = "kwt",
+  numeric.stats = c("Nmiss", "median", "q1q3")
+)
 labels <- list(
@@
   filter(row_number() == 1)
+# Order groups
+df$Diagnosis <- factor(df$Diagnosis, levels = c("Cystic Fibrosis", "CFTR-RD or CRMS", "Non-CF"))
 ```
@@
 
+# Cohorts by Na/Cl ratio
+
+## Based on visual inspection
+
+```{r}
+# Visual inspection
+ggplot(df, aes(x = Diagnosis, y = `Na/Cl Ratio`)) +
+  geom_boxplot() +
+  geom_hline(yintercept = 1, color = "red") +
+  geom_hline(yintercept = 1.25, color = "red")
+df$nacl_ratio_cat <- cut(df$`Na/Cl Ratio`, breaks = c(-Inf, 1, 1.25, Inf), right = F)
+kable(table(df$nacl_ratio_cat, df$Diagnosis))
+```
+
+## Based on k-means clustering
+
+```{r}
+set.seed(1017)
+# K means
+km_df <- df %>%
+  select(Diagnosis, Na, Cl) %>%
+  drop_na()
+ggplot(km_df, aes(x = Na, y = Cl, color = Diagnosis)) +
+  geom_point()
+km <- kmeans(km_df[, c("Na", "Cl")], centers = 3)
+km_df$Cluster <- factor(km$cluster,
+  levels = 1:3,
+  labels = c("High", "Medium", "Low")
+)
+centers <- data.frame(km$centers)
+centers$Cluster <- rownames(km$centers)
+centers$Cluster <- factor(centers$Cluster,
+  levels = 1:3,
+  labels = c("High", "Medium", "Low")
+)
+ggplot(km_df, aes(x = Na, y = Cl, color = Cluster)) +
+  geom_point(shape = 20, alpha = 0.2) +
+  geom_point(data = centers, size = 5)
+rownames(centers) <- NULL
+centers <- centers %>% column_to_rownames("Cluster")
+kable(centers, caption = "K-means centers", row.names = T)
+kable(as.data.frame.matrix(table(km_df$Cluster, km_df$Diagnosis)), row.names = T)
+```
+
 # Patient Demographics and Sample Characteristics
@@
 t1 <- tableby(
-  Diagnosis ~ Sex + Age.at.Test.in.Years + Method+ fe(geno_risk) + fe(geno_class),
+  Diagnosis ~ Sex + Age.at.Test.in.Years + Method + fe(geno_risk) + fe(geno_class),
   data = t1_df, control = mycontrols, simulate.p.value = T
@@
 t1 <- tableby(
-  `Sweat Chloride Group` ~ fe(geno_risk) + fe(Diagnosis)+  Cl + Na + `Na/Cl Ratio`,
-  data = t1_df, control = mycontrols,simulate.p.value = T
+  `Sweat Chloride Group` ~ fe(geno_risk) + fe(Diagnosis) + Cl + Na + `Na/Cl Ratio`,
+  data = t1_df, control = mycontrols, simulate.p.value = T
 )
@@
 
+For all participants with more than one sweat test, we used only their first value to calculate the absolute difference (AD) between left and right arm.
+
 ```{r}
+t1_df$`Cl AD` <- abs(t1_df$Cl_Left - t1_df$Cl_Right)
+t1_df$`Na AD` <- abs(t1_df$Na_Left - t1_df$Na_Right)
+```
 
+### By diagnosis
+
+```{r results='asis'}
+t1 <- tableby(
+  Diagnosis ~ `Cl AD` + `Na AD`,
+  data = t1_df, control = mycontrols,
+)
+summary(t1,
+  pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### By sweat chloride group
+
+```{r results='asis'}
+t1 <- tableby(
+  `Sweat Chloride Group` ~ `Cl AD` + `Na AD`,
+  data = t1_df, control = mycontrols
+)
+summary(t1,
+  pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### By genotype risk
+
+```{r results='asis'}
+t1 <- tableby(
+  geno_risk ~ `Cl AD` + `Na AD`,
+  data = t1_df, control = mycontrols
+)
+summary(t1,
+  pfootnote = T,
+  labelTranslations = labels
+)
 ```
@@
   filter(row_number() == 1 | row_number() == n()) %>% # First and last
-  mutate(time = row_number())
-multiples$time <- factor(multiples$time, levels = 1:2, labels = c("First", "Last"))
-# Paired testing options
-mycontrols <- tableby.control(
-  numeric.test = "signed.rank", cat.test = "mcnemar",
-  numeric.stats = c("Nmiss", "median", "q1q3")
+  summarise(
+    `Cl AD` = abs(last(Cl) - first(Cl)), # Calculate AD
+    `Na AD` = abs(last(Na) - first(Na)),
+    Diagnosis = first(Diagnosis),
+    `Sweat Chloride Group` = first(`Sweat Chloride Group`),
+    geno_risk = first(geno_risk)
   )
@@
 
+### By diagnosis
+
 ```{r results='asis'}
-p <- paired(
-  time ~ Cl + Na + `Na/Cl Ratio`,
-  signed.rank.exact = F, data = multiples,
-  id = PatientID, control = mycontrols
+t1 <- tableby(
+  Diagnosis ~ `Cl AD` + `Na AD`,
+  data = multiples, control = mycontrols,
 )
-summary(p, pfootnote = T, labelTranslations = labels)
+summary(t1,
+  pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### By sweat chloride group
+
+```{r results='asis'}
+t1 <- tableby(
+  `Sweat Chloride Group` ~ `Cl AD` + `Na AD`,
+  data = multiples, control = mycontrols
+)
+summary(t1,
+  pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### By genotype risk
+
+```{r results='asis'}
+t1 <- tableby(
+  geno_risk ~ `Cl AD` + `Na AD`,
+  data = multiples, control = mycontrols
+)
+summary(t1,
+  pfootnote = T,
+  labelTranslations = labels
+)
 ```
@@
 ```{r}
-roc_cf <- roc(response= t1_df$`Diagnosis`,
+roc_cf <- roc(
+  response = t1_df$`Diagnosis`,
   predictor = t1_df$`Na/Cl Ratio`,
-              levels = c("CFTR-RD or CRMS","Cystic Fibrosis"),
-              quiet = T)
+  levels = c("CFTR-RD or CRMS", "Cystic Fibrosis"),
+  quiet = T
+)
 coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)
 
-roc_non_cf <- roc(response= t1_df$`Diagnosis`,
+roc_non_cf <- roc(
+  response = t1_df$`Diagnosis`,
   predictor = t1_df$`Na/Cl Ratio`,
-              levels = c("Non-CF","CFTR-RD or CRMS"),
-              quiet = T)
+  levels = c("Non-CF", "CFTR-RD or CRMS"),
+  quiet = T
+)
 coords_non_cf <- round(coords(roc_non_cf, "best", best.method = "youden"), 3)
@@
 ```{r warning=FALSE}
-ggplot(t1_df[!is.na(t1_df$`Diagnosis`), ], 
-       aes(log(`Na/Cl Ratio`), fill = `Diagnosis`)) +
-  geom_histogram(alpha = 0.3, position = "identity",bins = 50) +
+ggplot(
+  t1_df[!is.na(t1_df$`Diagnosis`), ],
+  aes(log(`Na/Cl Ratio`), fill = `Diagnosis`)
+) +
+  geom_histogram(alpha = 0.3, position = "identity", bins = 50) +
   geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
   geom_vline(xintercept = log(coords_non_cf$threshold), color = "green") +
-  ylab("Count") + xlab("ln(Na/Cl Ratio)") +
+  ylab("Count") +
+  xlab("ln(Na/Cl Ratio)") +
   theme_bw()
@@
 t1_df$`Genotype Risk` <- factor(t1_df$geno_risk,
-                     levels = c("Not CF","Low","High","Unknown"),
-                     ordered = T)
-t1_df$`Genotype Risk`[t1_df$`Genotype Risk`=="Unknown"]= NA
+  levels = c("Not CF", "Low", "High", "Unknown"),
+  ordered = T
+)
+t1_df$`Genotype Risk`[t1_df$`Genotype Risk` == "Unknown"] <- NA
 ```
@@
 ```{r}
-roc_cf <- roc(response= t1_df$`Genotype Risk`,
+roc_cf <- roc(
+  response = t1_df$`Genotype Risk`,
   predictor = t1_df$`Na/Cl Ratio`,
-              levels = c("Low","High"),
-              quiet = T)
+  levels = c("Low", "High"),
+  quiet = T
+)
 coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)
 
-roc_non_cf <- roc(response= t1_df$`Genotype Risk`,
+roc_non_cf <- roc(
+  response = t1_df$`Genotype Risk`,
   predictor = t1_df$`Na/Cl Ratio`,
-              levels = c("Not CF","Low"),
-              quiet = T)
+  levels = c("Not CF", "Low"),
+  quiet = T
+)
 coords_non_cf <- round(coords(roc_non_cf, "best", best.method = "youden"), 3)
@@
 ```{r warning=FALSE}
-ggplot(t1_df[!is.na(t1_df$`Genotype Risk`), ], 
-       aes(log(`Na/Cl Ratio`), fill = `Genotype Risk`)) +
-  geom_histogram(alpha = 0.3, position = "identity",bins = 50) +
+ggplot(
+  t1_df[!is.na(t1_df$`Genotype Risk`), ],
+  aes(log(`Na/Cl Ratio`), fill = `Genotype Risk`)
+) +
+  geom_histogram(alpha = 0.3, position = "identity", bins = 50) +
   geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
   geom_vline(xintercept = log(coords_non_cf$threshold), color = "green") +
-  ylab("Count") + xlab("ln(Na/Cl Ratio)") +
+  ylab("Count") +
+  xlab("ln(Na/Cl Ratio)") +
   theme_bw()
@@
 # Exclude those with Cl >= 60
-t1_df = t1_df %>% filter(Cl < 60)
+t1_df <- t1_df %>% filter(Cl < 60)
 ```

### 2024-01-31 Christine Chan/CF Registry/cf_registry_analysis.qmd
cf registry time varying covariates fixed

--- 2024-01-26 Christine Chan/CF Registry/cf_registry_analysis.qmd
@@
 
-# CFRD incidence and prevalence
+<!-- # CFRD incidence and prevalence -->
 
-```{r}
-# All data
-encounter <- read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv",
-  na.strings = c("", "N/A")
-)
-annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
-  na.strings = c("", "N/A")
-)
-encounter <- encounter %>%
-  arrange(eDWID, reviewyear, encounternum)
-annual <- annual %>% arrange(eDWID, ReviewYear)
-# Add age
-ages <- annual %>%
-  mutate(`Age in 2018` = 2018 - (ReviewYear - Age_YrEnd)) %>%
-  rename(reviewyear = ReviewYear) %>%
-  group_by(eDWID, reviewyear) %>%
-  summarise(
-    `Age in 2018` = as.numeric(names(sort(table(`Age in 2018`),
-      decreasing = TRUE
-    )[1])),
-    Age = Age_YrEnd[1], .groups = "drop"
-  )
-encounter <- left_join(encounter, ages, by = c("eDWID", "reviewyear"))
-encounter$`Age Group 2018` <- cut(encounter$`Age in 2018`,
-  c(-Inf, 10, 19, 31, Inf),
-  labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
-  right = F
-)
-encounter$`Age Group` <- cut(encounter$Age,
-  c(-Inf, 10, 19, 31, Inf),
-  labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
-  right = F
-)
-# For those with CFRD by registry criteria, find first visit with diagnosis
-first_cfrd_cffpr <- encounter %>%
-  filter(cfrd_status == 3) %>%
-  group_by(eDWID) %>%
-  filter(row_number() == 1) %>%
-  select(eDWID, reviewyear, `Age in 2018`) %>%
-  rename(first_cfrd_cffpr = reviewyear)
-# Add to encounter level data
-encounter <- left_join(encounter, first_cfrd_cffpr,
-  by = c("eDWID", "Age in 2018")
-)
-# Fill down CFRD Yes/No by year
-encounter <- encounter %>%
-  group_by(eDWID) %>%
-  mutate(
-    cfrd_cffpr = case_when(
-      reviewyear < first_cfrd_cffpr ~ "No",
-      reviewyear >= first_cfrd_cffpr ~ "Yes"
-    )
-  ) %>%
-  ungroup()
-encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
-# Ordered factors
-encounter$cfrd_cffpr <- factor(encounter$cfrd_cffpr,
-  levels = c("No", "Yes"), ordered = T
-)
-# Remove those missing age
-encounter <- encounter %>% filter(!is.na(Age))
-```
+<!-- ```{r} -->
+<!-- # All data -->
+<!-- encounter <- read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv", -->
+<!--   na.strings = c("", "N/A") -->
+<!-- ) -->
+<!-- annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv", -->
+<!--   na.strings = c("", "N/A") -->
+<!-- ) -->
+<!-- encounter <- encounter %>% -->
+<!--   arrange(eDWID, reviewyear, encounternum) -->
+<!-- annual <- annual %>% arrange(eDWID, ReviewYear) -->
+<!-- # Add age -->
+<!-- ages <- annual %>% -->
+<!--   mutate(`Age in 2018` = 2018 - (ReviewYear - Age_YrEnd)) %>% -->
+<!--   rename(reviewyear = ReviewYear) %>% -->
+<!--   group_by(eDWID, reviewyear) %>% -->
+<!--   summarise( -->
+<!--     `Age in 2018` = as.numeric(names(sort(table(`Age in 2018`), -->
+<!--       decreasing = TRUE -->
+<!--     )[1])), -->
+<!--     Age = Age_YrEnd[1], .groups = "drop" -->
+<!--   ) -->
+<!-- encounter <- left_join(encounter, ages, by = c("eDWID", "reviewyear")) -->
+<!-- encounter$`Age Group 2018` <- cut(encounter$`Age in 2018`, -->
+<!--   c(-Inf, 10, 19, 31, Inf), -->
+<!--   labels = c("< 10", "10 - 18", "19 - 30", "> 30"), -->
+<!--   right = F -->
+<!-- ) -->
+<!-- encounter$`Age Group` <- cut(encounter$Age, -->
+<!--   c(-Inf, 10, 19, 31, Inf), -->
+<!--   labels = c("< 10", "10 - 18", "19 - 30", "> 30"), -->
+<!--   right = F -->
+<!-- ) -->
+<!-- # For those with CFRD by registry criteria, find first visit with diagnosis -->
+<!-- first_cfrd_cffpr <- encounter %>% -->
+<!--   filter(cfrd_status == 3) %>% -->
+<!--   group_by(eDWID) %>% -->
+<!--   filter(row_number() == 1) %>% -->
+<!--   select(eDWID, reviewyear, `Age in 2018`) %>% -->
+<!--   rename(first_cfrd_cffpr = reviewyear) -->
+<!-- # Add to encounter level data -->
+<!-- encounter <- left_join(encounter, first_cfrd_cffpr, -->
+<!--   by = c("eDWID", "Age in 2018") -->
+<!-- ) -->
+<!-- # Fill down CFRD Yes/No by year -->
+<!-- encounter <- encounter %>% -->
+<!--   group_by(eDWID) %>% -->
+<!--   mutate( -->
+<!--     cfrd_cffpr = case_when( -->
+<!--       reviewyear < first_cfrd_cffpr ~ "No", -->
+<!--       reviewyear >= first_cfrd_cffpr ~ "Yes" -->
+<!--     ) -->
+<!--   ) %>% -->
+<!--   ungroup() -->
+<!-- encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No" -->
+<!-- # Ordered factors -->
+<!-- encounter$cfrd_cffpr <- factor(encounter$cfrd_cffpr, -->
+<!--   levels = c("No", "Yes"), ordered = T -->
+<!-- ) -->
+<!-- # Remove those missing age -->
+<!-- encounter <- encounter %>% filter(!is.na(Age)) -->
+<!-- ``` -->
 
-```{r}
-#| label: tbl-incidence_cffpr
-#| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR Alone"
-# For each year, pull everyone with and without CFRD that year
-year_inc <- function(df = encounter, year, py = 1000) {
-  # Filter and count
-  inc <- df %>%
-    filter(reviewyear == year, is.na(first_cfrd_cffpr) | first_cfrd_cffpr >= year) %>%
-    group_by(eDWID) %>%
-    filter(row_number() == 1) %>%
-    select(eDWID, `Age Group`, cfrd_cffpr)
-  inc <- inc %>%
-    group_by(cfrd_cffpr) %>%
-    count(`Age Group`, .drop = F)
-  inc <- inc[!is.na(inc$`Age Group`), ]
-  # Calculations
-  new_cases <- sum(inc$n[inc$cfrd_cffpr == "Yes"])
-  total_at_risk <- sum(inc$n)
-  total_inc <- new_cases / (total_at_risk / py)
+<!-- ```{r} -->
+<!-- #| label: tbl-incidence_cffpr -->
+<!-- #| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR Alone" -->
+<!-- # For each year, pull everyone with and without CFRD that year -->
+<!-- year_inc <- function(df = encounter, year, py = 1000) { -->
+<!--   # Filter and count -->
+<!--   inc <- df %>% -->
+<!--     filter(reviewyear == year, is.na(first_cfrd_cffpr) | first_cfrd_cffpr >= year) %>% -->
+<!--     group_by(eDWID) %>% -->
+<!--     filter(row_number() == 1) %>% -->
+<!--     select(eDWID, `Age Group`, cfrd_cffpr) -->
+<!--   inc <- inc %>% -->
+<!--     group_by(cfrd_cffpr) %>% -->
+<!--     count(`Age Group`, .drop = F) -->
+<!--   inc <- inc[!is.na(inc$`Age Group`), ] -->
+<!--   # Calculations -->
+<!--   new_cases <- sum(inc$n[inc$cfrd_cffpr == "Yes"]) -->
+<!--   total_at_risk <- sum(inc$n) -->
+<!--   total_inc <- new_cases / (total_at_risk / py) -->
 
-  new_cases_10 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "< 10"])
-  at_risk_10 <- sum(inc$n[inc$`Age Group` == "< 10"])
-  inc_10 <- new_cases_10 / (at_risk_10 / py)
+<!--   new_cases_10 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "< 10"]) -->
+<!--   at_risk_10 <- sum(inc$n[inc$`Age Group` == "< 10"]) -->
+<!--   inc_10 <- new_cases_10 / (at_risk_10 / py) -->
 
-  new_cases_10_18 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "10 - 18"])
-  at_risk_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
-  inc_10_18 <- new_cases_10_18 / (at_risk_10_18 / py)
+<!--   new_cases_10_18 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "10 - 18"]) -->
+<!--   at_risk_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"]) -->
+<!--   inc_10_18 <- new_cases_10_18 / (at_risk_10_18 / py) -->
 
-  new_cases_19_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "19 - 30"])
-  at_risk_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
-  inc_19_30 <- new_cases_19_30 / (at_risk_19_30 / py)
+<!--   new_cases_19_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "19 - 30"]) -->
+<!--   at_risk_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"]) -->
+<!--   inc_19_30 <- new_cases_19_30 / (at_risk_19_30 / py) -->
 
-  new_cases_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "> 30"])
-  at_risk_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
-  inc_30 <- new_cases_30 / (at_risk_30 / py)
+<!--   new_cases_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "> 30"]) -->
+<!--   at_risk_30 <- sum(inc$n[inc$`Age Group` == "> 30"]) -->
+<!--   inc_30 <- new_cases_30 / (at_risk_30 / py) -->
 
-  # Nice results
-  list(
-    `Year` = year,
-    `Total New Cases` = new_cases,
-    `Total At Risk` = total_at_risk,
-    `Total Incidence` = total_inc,
-    `Total New Cases Age < 10` = new_cases_10,
-    `Total At Risk Age < 10` = at_risk_10,
-    `Total Incidence Age < 10` = inc_10,
-    `Total New Cases Age 10-18` = new_cases_10_18,
-    `Total At Risk Age 10-18` = at_risk_10_18,
-    `Total Incidence Age 10-18` = inc_10_18,
-    `Total New Cases Age 19-30` = new_cases_19_30,
-    `Total At Risk Age 19-30` = at_risk_19_30,
-    `Total Incidence Age 19-30` = inc_19_30,
-    `Total New Cases Age > 30` = new_cases_30,
-    `Total At Risk Age > 30` = at_risk_30,
-    `Total Incidence Age > 30` = inc_30
-  )
-}
-incidence <- lapply(2003:2018, function(y) {
-  year_inc(year = y)
-})
-incidence <- data.frame(do.call(rbind, incidence), check.names = F)
-incidence <- data.frame(lapply(incidence, as.numeric), check.names = F)
-kable(incidence, digits = 3)
-```
+<!--   # Nice results -->
+<!--   list( -->
+<!--     `Year` = year, -->
+<!--     `Total New Cases` = new_cases, -->
+<!--     `Total At Risk` = total_at_risk, -->
+<!--     `Total Incidence` = total_inc, -->
+<!--     `Total New Cases Age < 10` = new_cases_10, -->
+<!--     `Total At Risk Age < 10` = at_risk_10, -->
+<!--     `Total Incidence Age < 10` = inc_10, -->
+<!--     `Total New Cases Age 10-18` = new_cases_10_18, -->
+<!--     `Total At Risk Age 10-18` = at_risk_10_18, -->
+<!--     `Total Incidence Age 10-18` = inc_10_18, -->
+<!--     `Total New Cases Age 19-30` = new_cases_19_30, -->
+<!--     `Total At Risk Age 19-30` = at_risk_19_30, -->
+<!--     `Total Incidence Age 19-30` = inc_19_30, -->
+<!--     `Total New Cases Age > 30` = new_cases_30, -->
+<!--     `Total At Risk Age > 30` = at_risk_30, -->
+<!--     `Total Incidence Age > 30` = inc_30 -->
+<!--   ) -->
+<!-- } -->
+<!-- incidence <- lapply(2003:2018, function(y) { -->
+<!--   year_inc(year = y) -->
+<!-- }) -->
+<!-- incidence <- data.frame(do.call(rbind, incidence), check.names = F) -->
+<!-- incidence <- data.frame(lapply(incidence, as.numeric), check.names = F) -->
+<!-- kable(incidence, digits = 3) -->
+<!-- ``` -->
 
-In @tbl-incidence_cffpr, incidence per year is calculated as the number of new CFRD diagnoses (by registry definition alone) per 1,000 person-years of those at risk, excluding previously identified cases.
+<!-- In @tbl-incidence_cffpr, incidence per year is calculated as the number of new CFRD diagnoses (by registry definition alone) per 1,000 person-years of those at risk, excluding previously identified cases.
+-->
 
-```{r}
-#| label: tbl-prevalence_cffpr
-#| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR Alone"
-encounter %>%
-  group_by(eDWID, reviewyear) %>%
-  summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>%
-  group_by(reviewyear) %>%
-  summarise(
-    Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100,
-    .groups = "drop"
-  ) %>%
-  kable(., digits = 3)
-```
+<!-- ```{r} -->
+<!-- #| label: tbl-prevalence_cffpr -->
+<!-- #| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR Alone" -->
+<!-- encounter %>% -->
+<!--   group_by(eDWID, reviewyear) %>% -->
+<!--   summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>% -->
+<!--   group_by(reviewyear) %>% -->
+<!--   summarise( -->
+<!--     Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100, -->
+<!--     .groups = "drop" -->
+<!--   ) %>% -->
+<!--   kable(., digits = 3) -->
+<!-- ``` -->
 
-```{r}
-#| label: tbl-prevalence_cffpr_year
-#| tbl-cap: "Yearly Prevalence of CFRD by Age Group Based on CFFPR Alone"
-encounter %>%
-  group_by(eDWID, reviewyear, `Age Group`) %>%
-  summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>%
-  group_by(reviewyear, `Age Group`) %>%
-  summarise(Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100, .groups = "drop") %>%
-  pivot_wider(names_from = `Age Group`, values_from = Prevalence) %>%
-  kable(., digits = 3)
-```
+<!-- ```{r} -->
+<!-- #| label: tbl-prevalence_cffpr_year -->
+<!-- #| tbl-cap: "Yearly Prevalence of CFRD by Age Group Based on CFFPR Alone" -->
+<!-- encounter %>% -->
+<!--   group_by(eDWID, reviewyear, `Age Group`) %>% -->
+<!--   summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>% -->
+<!--   group_by(reviewyear, `Age Group`) %>% -->
+<!--   summarise(Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100, .groups = "drop") %>% -->
+<!--   pivot_wider(names_from = `Age Group`, values_from = Prevalence) %>% -->
+<!--   kable(., digits = 3) -->
+<!-- ``` -->
 
-In @tbl-prevalence_cffpr_year and @tbl-prevalence_cffpr, prevalence is calculated as the total number of CFRD cases (by registry definition alone) seen in clinic divided by the total number of people seen in clinic in a given year.
+<!-- In @tbl-prevalence_cffpr_year and @tbl-prevalence_cffpr, prevalence is calculated as the total number of CFRD cases (by registry definition alone) seen in clinic divided by the total number of people seen in clinic in a given year.
+-->
 
-# CFRD screening rates
+<!-- # CFRD screening rates -->
 
-```{r}
-#| label: tbl-screening_ogtt
-#| tbl-cap: "Yearly OGTT Screening Rates"
-year_screen <- function(df = encounter, year) {
-  # Filter and count
-  inc <- df %>%
-    filter(reviewyear == year, first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>%
-    mutate(ogtt = !is.na(ogttfast_bloodglucose) | !is.na(twohour_bloodglucose)) %>%
-    group_by(eDWID) %>%
-    summarise(ogtt = max(ogtt), `Age Group` = `Age Group`[1])
-  inc <- inc %>%
-    group_by(ogtt) %>%
-    count(`Age Group`) %>%
-    filter(`Age Group` != "< 10")
-  # Calculations
-  total_screened <- sum(inc$n[inc$ogtt == 1])
-  total_seen <- sum(inc$n)
-  screen_rate <- total_screened / total_seen * 100
+<!-- ```{r} -->
+<!-- #| label: tbl-screening_ogtt -->
+<!-- #| tbl-cap: "Yearly OGTT Screening Rates" -->
+<!-- year_screen <- function(df = encounter, year) { -->
+<!--   # Filter and count -->
+<!--   inc <- df %>% -->
+<!--     filter(reviewyear == year, first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>% -->
+<!--     mutate(ogtt = !is.na(ogttfast_bloodglucose) | !is.na(twohour_bloodglucose)) %>% -->
+<!--     group_by(eDWID) %>% -->
+<!--     summarise(ogtt = max(ogtt), `Age Group` = `Age Group`[1]) -->
+<!--   inc <- inc %>% -->
+<!--     group_by(ogtt) %>% -->
+<!--     count(`Age Group`) %>% -->
+<!--     filter(`Age Group` != "< 10") -->
+<!--   # Calculations -->
+<!--   total_screened <- sum(inc$n[inc$ogtt == 1]) -->
+<!--   total_seen <- sum(inc$n) -->
+<!--   screen_rate <- total_screened / total_seen * 100 -->
 
-  screened_10_18 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "10 - 18"]
-  seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
-  rate_10_18 <- screened_10_18 / seen_10_18 * 100
+<!--   screened_10_18 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "10 - 18"] -->
+<!--   seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"]) -->
+<!--   rate_10_18 <- screened_10_18 / seen_10_18 * 100 -->
 
-  screened_19_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "19 - 30"]
-  seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
-  rate_19_30 <- screened_19_30 / seen_19_30 * 100
+<!--   screened_19_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "19 - 30"] -->
+<!--   seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"]) -->
+<!--   rate_19_30 <- screened_19_30 / seen_19_30 * 100 -->
 
-  screened_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "> 30"]
-  seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
-  rate_30 <- screened_30 / seen_30 * 100
+<!--   screened_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "> 30"] -->
+<!--   seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"]) -->
+<!--   rate_30 <- screened_30 / seen_30 * 100 -->
 
-  # Nice results
-  list(
-    `Year` = year,
-    `Total Screened` = total_screened,
-    `Total Seen` = total_seen,
-    `Screen Rate (%)` = screen_rate,
-    `Total Screened Age 10-18` = screened_10_18,
-    `Total Seen Age 10-18` = seen_10_18,
-    `Screen Rate Age 10-18 (%)` = rate_10_18,
-    `Total Screened Age 19-30` = screened_19_30,
-    `Total Seen Age 19-30` = seen_19_30,
-    `Screen Rate Age 19-30 (%)` = rate_19_30,
-    `Total Screened Age > 30` = screened_30,
-    `Total Seen Age > 30` = seen_30,
-    `Screen Rate Age > 30 (%)` = rate_30
-  )
-}
-screen <- lapply(2003:2018, function(y) {
-  year_screen(year = y)
-})
-screen <- data.frame(do.call(rbind, screen), check.names = F)
-screen <- data.frame(lapply(screen, as.numeric), check.names = F)
-kable(screen, digits = 3)
-```
+<!--   # Nice results -->
+<!--   list( -->
+<!--     `Year` = year, -->
+<!--     `Total Screened` = total_screened, -->
+<!--     `Total Seen` = total_seen, -->
+<!--     `Screen Rate (%)` = screen_rate, -->
+<!--     `Total Screened Age 10-18` = screened_10_18, -->
+<!--     `Total Seen Age 10-18` = seen_10_18, -->
+<!--     `Screen Rate Age 10-18 (%)` = rate_10_18, -->
+<!--     `Total Screened Age 19-30` = screened_19_30, -->
+<!--     `Total Seen Age 19-30` = seen_19_30, -->
+<!--     `Screen Rate Age 19-30 (%)` = rate_19_30, -->
+<!--     `Total Screened Age > 30` = screened_30, -->
+<!--     `Total Seen Age > 30` = seen_30, -->
+<!--     `Screen Rate Age > 30 (%)` = rate_30 -->
+<!--   ) -->
+<!-- } -->
+<!-- screen <- lapply(2003:2018, function(y) { -->
+<!--   year_screen(year = y) -->
+<!-- }) -->
+<!-- screen <- data.frame(do.call(rbind, screen), check.names = F) -->
+<!-- screen <- data.frame(lapply(screen, as.numeric), check.names = F) -->
+<!-- kable(screen, digits = 3) -->
+<!-- ``` -->
 
-```{r}
-#| label: tbl-screening_a1c
-#| tbl-cap: "Yearly HbA1c Screening Rates"
-year_screen <- function(df = encounter, year) {
-  # Filter and count
-  inc <- df %>%
-    filter(reviewyear == year, first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>%
-    mutate(a1c = !is.na(hgba1c)) %>%
-    group_by(eDWID) %>%
-    summarise(a1c = max(a1c), `Age Group` = `Age Group`[1])
-  inc <- inc %>%
-    group_by(a1c) %>%
-    count(`Age Group`) %>%
-    filter(`Age Group` != "< 10")
-  # Calculations
-  total_screened <- sum(inc$n[inc$a1c == 1])
-  total_seen <- sum(inc$n)
-  screen_rate <- total_screened / total_seen * 100
+<!-- ```{r} -->
+<!-- #| label: tbl-screening_a1c -->
+<!-- #| tbl-cap: "Yearly HbA1c Screening Rates" -->
+<!-- year_screen <- function(df = encounter, year) { -->
+<!--   # Filter and count -->
+<!--   inc <- df %>% -->
+<!--     filter(reviewyear == year, first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>% -->
+<!--     mutate(a1c = !is.na(hgba1c)) %>% -->
+<!--     group_by(eDWID) %>% -->
+<!--     summarise(a1c = max(a1c), `Age Group` = `Age Group`[1]) -->
+<!--   inc <- inc %>% -->
+<!--     group_by(a1c) %>% -->
+<!--     count(`Age Group`) %>% -->
+<!--     filter(`Age Group` != "< 10") -->
+<!--   # Calculations -->
+<!--   total_screened <- sum(inc$n[inc$a1c == 1]) -->
+<!--   total_seen <- sum(inc$n) -->
+<!--   screen_rate <- total_screened / total_seen * 100 -->
 
-  screened_10_18 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "10 - 18"]
-  seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
-  rate_10_18 <- screened_10_18 / seen_10_18 * 100
+<!--   screened_10_18 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "10 - 18"] -->
+<!--   seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"]) -->
+<!--   rate_10_18 <- screened_10_18 / seen_10_18 * 100 -->
 
-  screened_19_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "19 - 30"]
-  seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
-  rate_19_30 <- screened_19_30 / seen_19_30 * 100
+<!--   screened_19_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "19 - 30"] -->
+<!--   seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"]) -->
+<!--   rate_19_30 <- screened_19_30 / seen_19_30 * 100 -->
 
-  screened_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "> 30"]
-  seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
-  rate_30 <- screened_30 / seen_30 * 100
+<!--   screened_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "> 30"] -->
+<!--   seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"]) -->
+<!--   rate_30 <- screened_30 / seen_30 * 100 -->
 
-  # Nice results
-  list(
-    `Year` = year,
-    `Total Screened` = total_screened,
-    `Total Seen` = total_seen,
-    `Screen Rate (%)` = screen_rate,
-    `Total Screened Age 10-18` = screened_10_18,
-    `Total Seen Age 10-18` = seen_10_18,
-    `Screen Rate Age 10-18 (%)` = rate_10_18,
-    `Total Screened Age 19-30` = screened_19_30,
-    `Total Seen Age 19-30` = seen_19_30,
-    `Screen Rate Age 19-30 (%)` = rate_19_30,
-    `Total Screened Age > 30` = screened_30,
-    `Total Seen Age > 30` = seen_30,
-    `Screen Rate Age > 30 (%)` = rate_30
-  )
-}
-screen <- lapply(2003:2018, function(y) {
-  year_screen(year = y)
-})
-screen <- data.frame(do.call(rbind, screen), check.names = F)
-screen <- data.frame(lapply(screen, as.numeric), check.names = F)
-kable(screen, digits = 3)
-```
+<!--   # Nice results -->
+<!--   list( -->
+<!--     `Year` = year, -->
+<!--     `Total Screened` = total_screened, -->
+<!--     `Total Seen` = total_seen, -->
+<!--     `Screen Rate (%)` = screen_rate, -->
+<!--     `Total Screened Age 10-18` = screened_10_18, -->
+<!--     `Total Seen Age 10-18` = seen_10_18, -->
+<!--     `Screen Rate Age 10-18 (%)` = rate_10_18, -->
+<!--     `Total Screened Age 19-30` = screened_19_30, -->
+<!--     `Total Seen Age 19-30` = seen_19_30, -->
+<!--     `Screen Rate Age 19-30 (%)` = rate_19_30, -->
+<!--     `Total Screened Age > 30` = screened_30, -->
+<!--     `Total Seen Age > 30` = seen_30, -->
+<!--     `Screen Rate Age > 30 (%)` = rate_30 -->
+<!--   ) -->
+<!-- } -->
+<!-- screen <- lapply(2003:2018, function(y) { -->
+<!--   year_screen(year = y) -->
+<!-- }) -->
+<!-- screen <- data.frame(do.call(rbind, screen), check.names = F) -->
+<!-- screen <- data.frame(lapply(screen, as.numeric), check.names = F) -->
+<!-- kable(screen, digits = 3) -->
+<!-- ``` -->
 
-In @tbl-screening_ogtt and @tbl-screening_a1c, people with CFRD (by registry definition) and those under age 10 were excluded.
+<!-- In @tbl-screening_ogtt and @tbl-screening_a1c, people with CFRD (by registry definition) and those under age 10 were excluded.
 Screening rate is calculated as the number of people with an OGTT or HbA1c, respectively, divided by the total number of people seen in clinic that year.
+-->
 
@@
 annual <- left_join(annual, first_cfrd, by = join_by(eDWID))
+# BMI percentile to categorical
+annual$bmi_perc_cat <- cut(annual$bmi_perc, breaks = c(0, 10, 50, 85, Inf))
+annual$bmi_perc_cat <- relevel(annual$bmi_perc_cat, ref = "(50,85]")
+# FEV1 to categorical
+annual$ppFEV1_cat <- cut(annual$A_FEV1_pct_predicted,
+  breaks = c(0, 40, 80, 100, Inf), right = F
+)
+annual$ppFEV1_cat <- relevel(annual$ppFEV1_cat, ref = "[100,Inf)")
 ```
@@
 vars <- c(
-  "age_group", "Gender", "bmi_perc", "MutClass", "A_IsOnEnzymes",
-  "A_FEV1_pct_predicted", "tube_feeding", "NumPulmExacerbation",
+  "age_group", "Gender", "bmi_perc_cat", "MutClass", "A_IsOnEnzymes",
+  "ppFEV1_cat", "tube_feeding", "NumPulmExacerbation",
   "A_pseudomonasaeruginosa", "A_fungalyeast1", "A_mycocultureresults3",
@@
   Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
-  A_FEV1_pct_predicted = "FEV1 percent predicted",
+  ppFEV1_cat = "FEV1 percent predicted",
   tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
@@
   NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
-  height_perc = "Height percentile", bmi_perc = "BMI percentile",
+  height_perc = "Height percentile", bmi_perc_cat = "BMI percentile",
   A_Vx770 = "Use of CFTR modulator Vx770",
@@
 ```{r}
-# Whole cohort
-km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ 1,
-  data = s, cluster = eDWID
-)
-ggsurvplot(km_fit,
-  data = s, risk.table = T, cumevents = T,
-  fun = "event", xlim = c(2003, 2018), break.x.by = 1
-)
-kable(tidy(km_fit))
-# By age group
-km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ age_group,
-  data = s, cluster = eDWID
-)
-ggsurvplot(km_fit,
-  data = s, risk.table = T, cumevents = T,
-  fun = "event", xlim = c(2003, 2018), break.x.by = 1
-)
-km_table <- tidy(km_fit)
-km_table <- split.data.frame(km_table, km_table$strata)
-invisible(lapply(km_table, function(t) {
-  name <- t$strata[1]
-  t <- t %>% select(-strata)
-  kable(t, caption = name)
-}))
+# # Whole cohort
+# km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ 1,
+#   data = s, cluster = eDWID
+# )
+# ggsurvplot(km_fit,
+#   data = s, risk.table = T, cumevents = T,
+#   fun = "event", xlim = c(2003, 2018), break.x.by = 1
+# )
+# kable(tidy(km_fit))
+# # By age group
+# km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ age_group,
+#   data = s, cluster = eDWID
+# )
+# ggsurvplot(km_fit,
+#   data = s, risk.table = T, cumevents = T,
+#   fun = "event", xlim = c(2003, 2018), break.x.by = 1
+# )
+# km_table <- tidy(km_fit)
+# km_table <- split.data.frame(km_table, km_table$strata)
+# invisible(lapply(km_table, function(t) {
+#   name <- t$strata[1]
+#   t <- t %>% select(-strata)
+#   kable(t, caption = name)
+# }))
 ```
@@
 x <- c(
-  "age_group", "Gender", "bmi_perc", "MutClass", "A_IsOnEnzymes",
+  "age_group", "Gender", "bmi_perc_cat", "MutClass", "A_IsOnEnzymes",
   "tube_feeding", "liver_disease",
-  "A_FEV1_pct_predicted", "NumPulmExacerbation",
+  "ppFEV1_cat", "NumPulmExacerbation",
   "A_pulmonarycomplications1", "A_corticosteroids1"
@@
 # Plot survival
-ggsurvplot(survfit(cox_res),
-  data = s, risk.table = T, cumevents = T, fun = "event",
-  xlim = c(2003, 2018), break.x.by = 1
-)
+# ggsurvplot(survfit(cox_res),
+#   data = s, risk.table = T, cumevents = T, fun = "event",
+#   xlim = c(2003, 2018), break.x.by = 1
+# )
 # Forest plot and coefficients table
@@
 # Diagnostics
-test.ph= cox.zph(cox_res)
-ggcoxzph(test.ph)
+# test.ph <- cox.zph(cox_res)
+# ggcoxzph(test.ph)
 # ggcoxdiagnostics(cox_res,type = "martingale")
@@
 
-```{r}
+- `tgroup=1` indicates the years 2003 - 2008
+- `tgroup=2` indicates the years 2009 - 2013
+- `tgroup=3` indicates the years 2014 - 2018
+
+```{r results='asis'}
 s_split <- survSplit(Surv(tstart, tstop, cfrd_status_annual) ~ .,
-  data = s, cut = 2003:2018, episode = "tgroup"
+  data = s, cut = c(2008, 2013, 2018), episode = "tgroup"
 )
-cox_res_split <- coxph(update(f, . ~ Gender:strata(tgroup)), data = s_split,
-                       cluster = eDWID)
-t = tidy(cox_res_split,exponentiate = T)
+cox_res_split <- coxph(
+  Surv(tstart, tstop, cfrd_status_annual) ~ bmi_perc_cat*strata(tgroup),
+  data = s_split, cluster = eDWID
+)
+
+invisible(lapply(x, function(v) {
+  new_f <- as.formula(paste0(
+    y, "~", paste0(x[-which(x == v)], collapse = "+"),
+    paste0("+", v, "+", v, ":strata(tgroup)")
+  ))
+  cox_res_split <- coxph(new_f, data = s_split, cluster = eDWID)
+  t <- tidy(cox_res_split, exponentiate = T, conf.int = T)
+  cat("\n")
+  cat("####", v)
+  cat("\n")
+  print(kable(t, digits = 2))
+  cat("\n")
+  cat("\n")
+  t <- t %>%
+    filter(!is.na(estimate))
+  p <- t %>%
+    ggplot(aes(y = fct_rev(term))) +
+    geom_point(aes(x = estimate), shape = 15, size = 3) +
+    geom_linerange(aes(xmin = conf.low, xmax = conf.high)) +
+    geom_vline(xintercept = 1, linetype = "dashed") +
+    xlab("Hazard Ratio")
+  cat("\n")
+  print(p)
+  cat("\n")
+}))
 ```
 
-#### P. aeruginosa
+<!-- #### P. aeruginosa -->
 
-```{r}
-# Update formula
-cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa),
-  data = s, cluster = eDWID
-)
-# Forest plot and coefficients table
-ggforest(cox_res, data = s)
-```
+<!-- ```{r} -->
+<!-- # Update formula -->
+<!-- cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa), -->
+<!--   data = s, cluster = eDWID -->
+<!-- ) -->
+<!-- # Forest plot and coefficients table -->
+<!-- ggforest(cox_res, data = s) -->
+<!-- ``` -->
 
-#### Aspergillus
+<!-- #### Aspergillus -->
 
-```{r}
-# Update formula
-cox_res <- coxph(update(f, . ~ . + A_fungalyeast1),
-  data = s, cluster = eDWID
-)
-# Forest plot and coefficients table
-ggforest(cox_res, data = s)
-```
+<!-- ```{r} -->
+<!-- # Update formula -->
+<!-- cox_res <- coxph(update(f, . ~ . + A_fungalyeast1), -->
+<!--   data = s, cluster = eDWID -->
+<!-- ) -->
+<!-- # Forest plot and coefficients table -->
+<!-- ggforest(cox_res, data = s) -->
+<!-- ``` -->
 
-#### ABPA
+<!-- #### ABPA -->
 
-```{r}
-# Update formula
-cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1),
-  data = s, cluster = eDWID
-)
-# Forest plot and coefficients table
-ggforest(cox_res, data = s)
-```
+<!-- ```{r} -->
+<!-- # Update formula -->
+<!-- cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1), -->
+<!--   data = s, cluster = eDWID -->
+<!-- ) -->
+<!-- # Forest plot and coefficients table -->
+<!-- ggforest(cox_res, data = s) -->
+<!-- ``` -->
 
-### Vx770
+<!-- ### Vx770 -->
 
-```{r}
-f <- update(f, . ~ . + A_Vx770)
-cox_res <- coxph(f, data = s, cluster = eDWID)
-# Plot survival
-ggsurvplot(survfit(cox_res),
-  data = s, risk.table = T, cumevents = T, fun = "event",
-  xlim = c(2003, 2018), break.x.by = 1
-)
-# Forest plot and coefficients table
-ggforest(cox_res, data = s)
-# Diagnostics
-# test.ph= cox.zph(cox_res)
-# ggcoxzph(test.ph)
-# ggcoxdiagnostics(cox_res,type = "martingale")
-# ggcoxdiagnostics(cox_res,type = "deviance")
-# ggcoxdiagnostics(cox_res,type = "score")
-# ggcoxdiagnostics(cox_res,type = "schoenfeld")
-```
+<!-- ```{r} -->
+<!-- f <- update(f, . ~ . + A_Vx770) -->
+<!-- cox_res <- coxph(f, data = s, cluster = eDWID) -->
+<!-- # Plot survival -->
+<!-- ggsurvplot(survfit(cox_res), -->
+<!--   data = s, risk.table = T, cumevents = T, fun = "event", -->
+<!--   xlim = c(2003, 2018), break.x.by = 1 -->
+<!-- ) -->
+<!-- # Forest plot and coefficients table -->
+<!-- ggforest(cox_res, data = s) -->
+<!-- # Diagnostics -->
+<!-- # test.ph= cox.zph(cox_res) -->
+<!-- # ggcoxzph(test.ph) -->
+<!-- # ggcoxdiagnostics(cox_res,type = "martingale") -->
+<!-- # ggcoxdiagnostics(cox_res,type = "deviance") -->
+<!-- # ggcoxdiagnostics(cox_res,type = "score") -->
+<!-- # ggcoxdiagnostics(cox_res,type = "schoenfeld") -->
+<!-- ``` -->
 
-#### P. aeruginosa
+<!-- #### P. aeruginosa -->
 
-```{r}
-# Update formula
-cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa),
-  data = s, cluster = eDWID
-)
-# Forest plot and coefficients table
-ggforest(cox_res, data = s)
-```
+<!-- ```{r} -->
+<!-- # Update formula -->
+<!-- cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa), -->
+<!--   data = s, cluster = eDWID -->
+<!-- ) -->
+<!-- # Forest plot and coefficients table -->
+<!-- ggforest(cox_res, data = s) -->
+<!-- ``` -->
 
-#### Aspergillus
+<!-- #### Aspergillus -->
 
-```{r}
-# Update formula
-cox_res <- coxph(update(f, . ~ . + A_fungalyeast1),
-  data = s, cluster = eDWID
-)
-# Forest plot and coefficients table
-ggforest(cox_res, data = s)
-```
+<!-- ```{r} -->
+<!-- # Update formula -->
+<!-- cox_res <- coxph(update(f, . ~ . + A_fungalyeast1), -->
+<!--   data = s, cluster = eDWID -->
+<!-- ) -->
+<!-- # Forest plot and coefficients table -->
+<!-- ggforest(cox_res, data = s) -->
+<!-- ``` -->
 
-#### ABPA
+<!-- #### ABPA -->
 
-```{r}
-# Update formula
-cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1),
-  data = s, cluster = eDWID
-)
-# Forest plot and coefficients table
-ggforest(cox_res, data = s)
-```
+<!-- ```{r} -->
+<!-- # Update formula -->
+<!-- cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1), -->
+<!--   data = s, cluster = eDWID -->
+<!-- ) -->
+<!-- # Forest plot and coefficients table -->
+<!-- ggforest(cox_res, data = s) -->
+<!-- ``` -->
 
-#### Mycobacterium
+<!-- #### Mycobacterium -->
 
-```{r}
-# Update formula
-cox_res <- coxph(update(f, . ~ . + A_mycocultureresults3),
-  data = s, cluster = eDWID
-)
-# Forest plot and coefficients table
-ggforest(cox_res, data = s)
-```
+<!-- ```{r} -->
+<!-- # Update formula -->
+<!-- cox_res <- coxph(update(f, . ~ . + A_mycocultureresults3), -->
+<!--   data = s, cluster = eDWID -->
+<!-- ) -->
+<!-- # Forest plot and coefficients table -->
+<!-- ggforest(cox_res, data = s) -->
+<!-- ``` -->
 
-### Vx809
+<!-- ### Vx809 -->
 
-```{r}
-f <- update(f, . ~ . + A_VX809comb)
-cox_res <- coxph(f, data = s, cluster = eDWID)
-# Plot survival
-ggsurvplot(survfit(cox_res),
-  data = s, risk.table = T, cumevents = T, fun = "event",
-  xlim = c(2003, 2018), break.x.by = 1
-)
-# Forest plot and coefficients table
-ggforest(cox_res, data = s)
-# Diagnostics
-# test.ph= cox.zph(cox_res)
-# ggcoxzph(test.ph)
-# ggcoxdiagnostics(cox_res,type = "martingale")
-# ggcoxdiagnostics(cox_res,type = "deviance")
-# ggcoxdiagnostics(cox_res,type = "score")
-# ggcoxdiagnostics(cox_res,type = "schoenfeld")
-```
+<!-- ```{r} -->
+<!-- f <- update(f, . ~ . + A_VX809comb) -->
+<!-- cox_res <- coxph(f, data = s, cluster = eDWID) -->
+<!-- # Plot survival -->
+<!-- ggsurvplot(survfit(cox_res), -->
+<!--   data = s, risk.table = T, cumevents = T, fun = "event", -->
+<!--   xlim = c(2003, 2018), break.x.by = 1 -->
+<!-- ) -->
+<!-- # Forest plot and coefficients table -->
+<!-- ggforest(cox_res, data = s) -->
+<!-- # Diagnostics -->
+<!-- # test.ph= cox.zph(cox_res) -->
+<!-- # ggcoxzph(test.ph) -->
+<!-- # ggcoxdiagnostics(cox_res,type = "martingale") -->
+<!-- # ggcoxdiagnostics(cox_res,type = "deviance") -->
+<!-- # ggcoxdiagnostics(cox_res,type = "score") -->
+<!-- # ggcoxdiagnostics(cox_res,type = "schoenfeld") -->
+<!-- ``` -->
 
-#### P. aeruginosa
+<!-- #### P. aeruginosa -->
 
-```{r}
-# Update formula
-cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa),
-  data = s, cluster = eDWID
-)
-# Forest plot and coefficients table
-ggforest(cox_res, data = s)
-```
+<!-- ```{r} -->
+<!-- # Update formula -->
+<!-- cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa), -->
+<!--   data = s, cluster = eDWID -->
+<!-- ) -->
+<!-- # Forest plot and coefficients table -->
+<!-- ggforest(cox_res, data = s) -->
+<!-- ``` -->
 
-#### Aspergillus
+<!-- #### Aspergillus -->
 
-```{r}
-# Update formula
-cox_res <- coxph(update(f, . ~ . + A_fungalyeast1),
-  data = s, cluster = eDWID
-)
-# Forest plot and coefficients table
-ggforest(cox_res, data = s)
-```
+<!-- ```{r} -->
+<!-- # Update formula -->
+<!-- cox_res <- coxph(update(f, . ~ . + A_fungalyeast1), -->
+<!--   data = s, cluster = eDWID -->
+<!-- ) -->
+<!-- # Forest plot and coefficients table -->
+<!-- ggforest(cox_res, data = s) -->
+<!-- ``` -->
 
-#### ABPA
+<!-- #### ABPA -->
 
-```{r}
-# Update formula
-cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1),
-  data = s, cluster = eDWID
-)
-# Forest plot and coefficients table
-ggforest(cox_res, data = s)
-```
+<!-- ```{r} -->
+<!-- # Update formula -->
+<!-- cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1), -->
+<!--   data = s, cluster = eDWID -->
+<!-- ) -->
+<!-- # Forest plot and coefficients table -->
+<!-- ggforest(cox_res, data = s) -->
+<!-- ``` -->
 
-#### Mycobacterium
+<!-- #### Mycobacterium -->
 
-```{r}
-# Update formula
-cox_res <- coxph(update(f, . ~ . + A_mycocultureresults3),
-  data = s, cluster = eDWID
-)
-# Forest plot and coefficients table
-ggforest(cox_res, data = s)
-```
+<!-- ```{r} -->
+<!-- # Update formula -->
+<!-- cox_res <- coxph(update(f, . ~ . + A_mycocultureresults3), -->
+<!--   data = s, cluster = eDWID -->
+<!-- ) -->
+<!-- # Forest plot and coefficients table -->
+<!-- ggforest(cox_res, data = s) -->
+<!-- ``` -->
+
+<!-- t -->

### 2024-02-01 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd

@@
 
-```{r}
+```{r warning=FALSE}
 # Visual inspection
@@
 
-```{r}
+### By Na and Cl
+
+```{r warning=FALSE}
 set.seed(1017)
@@
   geom_point(data = centers, size = 5)
+rownames(centers) <- NULL
+centers <- centers %>% column_to_rownames("Cluster")
+kable(centers, caption = "K-means centers", row.names = T)
+kable(as.data.frame.matrix(table(km_df$Cluster, km_df$Diagnosis)), row.names = T)
+```
+
+### By Na/Cl ratio
+
+```{r warning=FALSE}
+set.seed(1017)
+# K means
+km_df <- df %>%
+  select(Diagnosis, `Na/Cl Ratio`) %>%
+  drop_na()
+km <- kmeans(km_df[, c("Na/Cl Ratio")], centers = 3)
+km_df$Cluster <- factor(km$cluster,
+  levels = 1:3,
+  labels = c("Low","Medium","High")
+)
+centers <- data.frame(km$centers)
+centers$Cluster <- rownames(km$centers)
+centers$Cluster <- factor(centers$Cluster,
+  levels = 1:3,
+  labels = c("Low","Medium","High")
+)
 rownames(centers) <- NULL

### 2024-02-08 Scott Sagel/PEx in CF/analysis_for_yaron.qmd
Yaron Trikafta fixed

@@
 library(gtsummary)
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
-)
-knitr::opts_knit$set(root.dir = home_dir)
+knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Scott Sagel/PEx in CF")
 ```
@@
 # Import
-df <- read.csv("./Scott Sagel/PEx in CF/Data_Raw/ExacerbationsInCF222_DATA_2024-01-02_0857.csv",
+df <- read.csv("./Data_Raw/ExacerbationsInCF222_DATA_2024-01-02_0857.csv",
   na.strings = ""
@@
 )
+# CFTR modulators
 df$cftr_mod <- factor(df$cftr_mod, levels = 0:1, labels = c("No", "Yes"))
+df$cftr_mod_spec <- factor(df$cftr_mod_spec,
+  levels = 1:4, labels = c("Kalydeco", "Orkambi", "Symdeko", "Trikafta")
+)
+df$Trikafta <- factor(df$cftr_mod_spec == "Trikafta",
+  levels = c(F, T), labels = c("No", "Yes")
+)
 # Labels
@@
   "Other bacteria" = "admit_cx_results___9",
-  "CFTR Modulator" = "cftr_mod",
+  "On CFTR Modulator?" = "cftr_mod",
+  "CFTR Modulator Type" = "cftr_mod_spec",
   "Cultures Tested" = "redcap_data_access_group"
@@
     sid, redcap_event_name, redcap_data_access_group, year, cftr_mod,
-    fungal_test, fungi___1:fungi___5,
+    cftr_mod_spec, Trikafta, fungal_test, fungi___1:fungi___5,
     ntm_test, ntm___1:ntm___3,
@@
   filter(fungal_test == 1) %>%
-  select(year, cftr_mod, fungi___1:fungi___5) %>%
+  select(year, cftr_mod_spec, fungi___1:fungi___5) %>%
   tbl_strata(
@@
       ~ .x %>%
-        tbl_summary(by = cftr_mod, missing_text = "Missing", label = labels) %>%
+        tbl_summary(by = cftr_mod_spec, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-fungi_trikafta
+#| tbl-cap: Fungi by Trikafta Status
+#| message: false
+df %>%
+  filter(fungal_test == 1) %>%
+  select(year, Trikafta, fungi___1:fungi___5) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
         add_overall() %>%
@@
   filter(ntm_test == 1) %>%
-  select(year, cftr_mod, ntm___1:ntm___3) %>%
+  select(year, cftr_mod_spec, ntm___1:ntm___3) %>%
   tbl_strata(
@@
       ~ .x %>%
-        tbl_summary(by = cftr_mod, missing_text = "Missing", label = labels) %>%
+        tbl_summary(by = cftr_mod_spec, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-ntm_trikafta
+#| tbl-cap: NTM by Trikafta Status
+#| message: false
+df %>%
+  filter(ntm_test == 1) %>%
+  select(year, Trikafta, ntm___1:ntm___3) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
         add_overall() %>%
@@
 df %>%
-  select(year, cftr_mod, admit_cx_results___1:admit_cx_results___9) %>%
+  select(year, cftr_mod_spec, admit_cx_results___1:admit_cx_results___9) %>%
   tbl_strata(
@@
       ~ .x %>%
-        tbl_summary(by = cftr_mod, missing_text = "Missing", label = labels) %>%
+        tbl_summary(by = cftr_mod_spec, missing_text = "Missing", label = labels) %>%
         add_overall() %>%
@@
 ```
+
+```{r}
+#| label: tbl-bacteria_trikafta
+#| tbl-cap: Bacteria by Trikafta Status
+#| message: false
+df %>%
+  select(year, Trikafta, admit_cx_results___1:admit_cx_results___9) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```

### 2024-02-08 Christine Chan/CF Registry/cf_registry_analysis.qmd
time varying survival models done

@@
 library(broom)
-switch(Sys.info()[["sysname"]],
-  Windows = {
-    home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry"
-  },
-  Darwin = {
-    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry"
-  }
-)
-knitr::opts_knit$set(root.dir = home_dir)
+knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry")
 ```
@@
 demogs <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
-demogs <- demogs %>% select(eDWID, Gender, MutClass)
+races <- c(
+  "White", "Black or African American", "American Indian or Alaska Native",
+  "Asian", "Native Hawaiian or Other Pacific ", "Some other race"
+)
+demogs$Race <- apply(demogs, 1, function(r) {
+  w <- which(r[paste0("Race", 1:6)] == 1)
+  if (length(w) > 1) {
+    return("More than one race")
+  } else {
+    return(races[w])
+  }
+})
+demogs$Race <- factor(demogs$Race,
+  levels = c(
+    "White", "American Indian or Alaska Native", "Asian",
+    "Black or African American", "More than one race",
+    "Native Hawaiian or Other Pacific ", "Some other race"
+  ),
+  labels = c(
+    "White", "Other race", "Other race",
+    "Black or African American", "More than one race",
+    "Other race", "Other race"
+  )
+)
+demogs$Hispanicrace <- factor(demogs$Hispanicrace,
+  levels = 1:2,
+  labels = c("Yes", "No")
+)
+demogs <- demogs %>% select(eDWID, Gender, MutClass, Race, Hispanicrace)
 annual <- left_join(annual, demogs, by = join_by(eDWID))
@@
 vars <- c(
-  "age_group", "Gender", "bmi_perc_cat", "MutClass", "A_IsOnEnzymes",
-  "ppFEV1_cat", "tube_feeding", "NumPulmExacerbation",
-  "A_pseudomonasaeruginosa", "A_fungalyeast1", "A_mycocultureresults3",
-  "A_pulmonarycomplications1", "A_corticosteroids1", "cfrd_status_annual",
-  "A_Vx770", "A_VX809comb", "liver_disease"
+  "age_group", "Gender", "bmi_perc", "bmi_perc_cat", "MutClass", "Race",
+  "Hispanicrace", "A_IsOnEnzymes", "ppFEV1_cat", "tube_feeding",
+  "NumPulmExacerbation", "A_pseudomonasaeruginosa", "A_fungalyeast1",
+  "A_mycocultureresults3", "A_pulmonarycomplications1", "A_corticosteroids1",
+  "cfrd_status_annual", "A_Vx770", "A_VX809comb", "liver_disease"
 )
@@
   cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
+  Race = "Race", Hispanicrace = "Hispanic",
   Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
@@
   A_pulmonarycomplications1 = "ABPA",
-  A_mycocultureresults3 = "Microorganisms in myco culture"
+  A_mycocultureresults3 = "Microorganisms in myco culture",
+  eDWID = "ID",birth_year="Birth year",tstart="Interval start",tstop="Interval stop",age_at_entry="Age at entry",Age_YrEnd="Age",bmi_perc = "BMI %ile"
 )
@@
 s$age_group <- droplevels(s$age_group)
+# Labels
+label(s) = labels[colnames(s)]
 ```
@@
 ```{r}
-# # Whole cohort
-# km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ 1,
-#   data = s, cluster = eDWID
-# )
-# ggsurvplot(km_fit,
-#   data = s, risk.table = T, cumevents = T,
-#   fun = "event", xlim = c(2003, 2018), break.x.by = 1
-# )
-# kable(tidy(km_fit))
-# # By age group
-# km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ age_group,
-#   data = s, cluster = eDWID
-# )
-# ggsurvplot(km_fit,
-#   data = s, risk.table = T, cumevents = T,
-#   fun = "event", xlim = c(2003, 2018), break.x.by = 1
-# )
-# km_table <- tidy(km_fit)
-# km_table <- split.data.frame(km_table, km_table$strata)
-# invisible(lapply(km_table, function(t) {
-#   name <- t$strata[1]
-#   t <- t %>% select(-strata)
-#   kable(t, caption = name)
-# }))
+# Whole cohort
+km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ 1,
+  data = s, cluster = eDWID
+)
+ggsurvplot(km_fit,
+  data = s, risk.table = T, cumevents = T,
+  fun = "event", xlim = c(2003, 2018), break.x.by = 1
+)
+kable(tidy(km_fit))
+# By age group
+km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ age_group,
+  data = s, cluster = eDWID
+)
+ggsurvplot(km_fit,
+  data = s, risk.table = T, cumevents = T,
+  fun = "event", xlim = c(2003, 2018), break.x.by = 1
+)
+km_table <- tidy(km_fit)
+km_table <- split.data.frame(km_table, km_table$strata)
+invisible(lapply(km_table, function(t) {
+  name <- t$strata[1]
+  t <- t %>% select(-strata)
+  kable(t, caption = name)
+}))
 ```
@@
 x <- c(
-  "age_group", "Gender", "bmi_perc_cat", "MutClass", "A_IsOnEnzymes",
-  "tube_feeding", "liver_disease",
-  "ppFEV1_cat", "NumPulmExacerbation",
-  "A_pulmonarycomplications1", "A_corticosteroids1"
+  "age_group", "Gender", "Race","Hispanicrace", "bmi_perc_cat", "MutClass", 
+  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat", 
+  "NumPulmExacerbation", "A_pulmonarycomplications1", "A_corticosteroids1"
 )
@@
 # Diagnostics
-# test.ph <- cox.zph(cox_res)
+test.ph <- data.frame(cox.zph(cox_res)$table)
+test.ph = test.ph[rownames(test.ph) != "GLOBAL",]
+# Find which variables technically violate PH assumption
+time_varying = rownames(test.ph)[test.ph$p < 0.05]
 # ggcoxzph(test.ph)
@@
 
-```{r results='asis'}
+The interpretation of these models can be a little bit tricky, so here is an example using BMI percentile group and not including the other covariates in the model:
+
+```{r}
 s_split <- survSplit(Surv(tstart, tstop, cfrd_status_annual) ~ .,
@@
 )
+s_split$tgroup <- factor(s_split$tgroup,
+  levels = 1:3, labels = c("[2003,2009)", "[2009,2014)", "[2014,2019)")
+)
 cox_res_split <- coxph(
-  Surv(tstart, tstop, cfrd_status_annual) ~ bmi_perc_cat*strata(tgroup),
+  Surv(tstart, tstop, cfrd_status_annual) ~ bmi_perc_cat * strata(tgroup),
   data = s_split, cluster = eDWID
 )
+kable(summary(cox_res_split)$coefficients, digits = 3)
+```
 
-invisible(lapply(x, function(v) {
+In the table above, the BMI reference group is the (50,85] group.
+So, in the years 2003 - 2008, the hazard ratio of the (0,10] BMI group was $exp(0.806)=2.238$ compared to the reference.
+Then, in the period of 2009 - 2013, the hazard ratio of the (0,10] BMI group was $exp(0.806+0.050)=2.354$ compared to the reference group, although this was not statistically significant.
+The HR for the (85,Inf] BMI group appears to increase over time, starting out at $exp(-0.511)=0.600$ in the first time period, then a statistically significant increase in the second time period to $exp(-0.511+0.282)=0.795$, with a final increase to $exp(-0.511+0.435)=0.927$ in the third period.
+The interpretation of the following results will follow these same formulas, and the time-varying results are shown at the bottom of each table.
+This report only includes variables that were statistically significant in the Cox proportional hazards test.
+
+```{r results='asis'}
+invisible(lapply(time_varying, function(v) {
   new_f <- as.formula(paste0(
     y, "~", paste0(x[-which(x == v)], collapse = "+"),
-    paste0("+", v, "+", v, ":strata(tgroup)")
+    paste0("+", v, "*strata(tgroup)")
   ))
   cox_res_split <- coxph(new_f, data = s_split, cluster = eDWID)
-  t <- tidy(cox_res_split, exponentiate = T, conf.int = T)
+  res <- summary(cox_res_split)$coefficients
   cat("\n")
@@
   cat("\n")
-  print(kable(t, digits = 2))
   cat("\n")
-  cat("\n")
-  t <- t %>%
-    filter(!is.na(estimate))
-  p <- t %>%
-    ggplot(aes(y = fct_rev(term))) +
-    geom_point(aes(x = estimate), shape = 15, size = 3) +
-    geom_linerange(aes(xmin = conf.low, xmax = conf.high)) +
-    geom_vline(xintercept = 1, linetype = "dashed") +
-    xlab("Hazard Ratio")
-  cat("\n")
-  print(p)
+  print(kable(res, digits = 3))
   cat("\n")

### 2024-02-16 Christine Chan/CF Registry/cf_registry_analysis.qmd
forest plot with mostly base R

@@
 ```{r}
+forest_plot <- function(mod, time_varying = NULL, labels = labels) {
+  coeff <- data.frame(summary(mod)$coefficients)
+  coeff$ll <- coeff$coef - (1.96 * coeff$se.coef.)
+  coeff$ul <- coeff$coef + (1.96 * coeff$se.coef.)
+  # Clean up variable names
+  l <- unlist(sapply(labels, paste0, ": "))
+  coeff$names <- rownames(coeff)
+  coeff$names <- str_replace_all(coeff$names, l)
+  coeff$names <- ordered(coeff$names, levels = unique(coeff$names))
+  # Exponentiate
+  coeff$exp.ll <- exp(coeff$ll)
+  coeff$exp.ul <- exp(coeff$ul)
+  # Plot
+  p <- ggplot(coeff, aes(y = fct_rev(names))) +
+    geom_point(aes(x = exp.coef.), shape = 15, size = 3) +
+    geom_linerange(aes(xmin = exp.ll, xmax = exp.ul)) +
+    geom_vline(xintercept = 1, linetype = "dashed") +
+    xlab("Hazard Ratio") +
+    theme_classic() +
+    theme(axis.ticks.y = element_blank(), axis.title.y = element_blank())
+}
+```
+
+```{r}
 # List of covariates
@@
   A_mycocultureresults3 = "Microorganisms in myco culture",
-  eDWID = "ID",birth_year="Birth year",tstart="Interval start",tstop="Interval stop",age_at_entry="Age at entry",Age_YrEnd="Age",bmi_perc = "BMI %ile"
+  eDWID = "ID", birth_year = "Birth year", tstart = "Interval start", tstop = "Interval stop", age_at_entry = "Age at entry", Age_YrEnd = "Age", bmi_perc = "BMI %ile"
 )
@@
 # Labels
-label(s) = labels[colnames(s)]
+label(s) <- labels[colnames(s)]
 ```
@@
 x <- c(
-  "age_group", "Gender", "Race","Hispanicrace", "bmi_perc_cat", "MutClass", 
+  "age_group", "Gender", "Race", "Hispanicrace", "bmi_perc_cat", "MutClass",
   "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
@@
 test.ph <- data.frame(cox.zph(cox_res)$table)
-test.ph = test.ph[rownames(test.ph) != "GLOBAL",]
+test.ph <- test.ph[rownames(test.ph) != "GLOBAL", ]
 # Find which variables technically violate PH assumption
-time_varying = rownames(test.ph)[test.ph$p < 0.05]
+time_varying <- rownames(test.ph)[test.ph$p < 0.05]
 # ggcoxzph(test.ph)

### 2024-02-16 Christine Chan/CF Registry/cf_registry_analysis.qmd
forest plot funciton mostly done

@@
 annual$ppFEV1_cat <- relevel(annual$ppFEV1_cat, ref = "[100,Inf)")
+# Males to reference group
+annual$Gender <- factor(annual$Gender, levels = c("M", "F"))
+# Hispanic no as reference
+annual$Hispanicrace <- factor(annual$Hispanicrace, levels = c("No", "Yes"))
 ```
@@
 forest_plot <- function(mod, time_varying = NULL, labels = labels) {
-  coeff <- data.frame(summary(mod)$coefficients)
-  coeff$ll <- coeff$coef - (1.96 * coeff$se.coef.)
-  coeff$ul <- coeff$coef + (1.96 * coeff$se.coef.)
-  # Clean up variable names
-  l <- unlist(sapply(labels, paste0, ": "))
-  coeff$names <- rownames(coeff)
-  coeff$names <- str_replace_all(coeff$names, l)
-  coeff$names <- ordered(coeff$names, levels = unique(coeff$names))
-  # Exponentiate
-  coeff$exp.ll <- exp(coeff$ll)
-  coeff$exp.ul <- exp(coeff$ul)
+  coeff <- tbl_regression(cox_res,
+    exponentiate = T,
+    pvalue_fun = function(x) x
+  ) %>%
+    add_n(location = "level") %>%
+    as_tibble()
+  coeff_n <- tbl_regression(cox_res, exponentiate = T) %>%
+    add_n(location = "label") %>%
+    as_tibble()
+  coeff$N2 <- coeff_n$`**N**`
+  # Combine Ns
+  coeff <- coeff %>% unite(N, `**N**`, N2, remove = F, na.rm = T)
+  # Separate CI column
+  coeff$hr <- as.numeric(coeff$`**HR**`)
+  coeff$ll <- as.numeric(sapply(str_split(coeff$`**95% CI**`, ", "), "[", 1))
+  coeff$ul <- as.numeric(sapply(str_split(coeff$`**95% CI**`, ", "), "[", 2))
+  # Add N to names
+  coeff$names <- paste0(coeff$`**Characteristic**`, " (N = ", coeff$N, ")")
+  coeff$names <- ordered(coeff$names,
+    levels = unique(coeff$names)
+  )
+  # P values
+  coeff$sig <- coeff$`**p-value**` < 0.05
+  coeff$p <- style_pvalue(coeff$`**p-value**`)
+  # Row types
+  coeff$type <- "level"
+  coeff$type[!is.na(coeff$N2)] <- "variable"
+  coeff$type[!is.na(coeff$`**N**`) & is.na(coeff$`**HR**`)] <- "reference"
+  # HR label
+  coeff$hr_lab <- paste0(coeff$`**HR**`, " (", coeff$`**95% CI**`, ")")
+  coeff$hr_lab[coeff$hr_lab == "NA (NA)"] <- NA
+  coeff$hr_lab[coeff$type == "reference"] <- "Reference"
   # Plot
+  p_left <- ggplot(coeff, aes(y = fct_rev(names))) +
+    geom_text(aes(x = 0, label = names),
+      hjust = 0,
+      fontface = ifelse(coeff$type == "variable", "bold", "plain")
+    ) +
+    geom_text(aes(x = 1, label = hr_lab), hjust = 0) +
+    theme_void() +
+    coord_cartesian(xlim = c(0, 2)) +
+    geom_stripes()
   p <- ggplot(coeff, aes(y = fct_rev(names))) +
-    geom_point(aes(x = exp.coef.), shape = 15, size = 3) +
-    geom_linerange(aes(xmin = exp.ll, xmax = exp.ul)) +
+    geom_point(aes(x = hr), shape = 15, size = 3) +
+    geom_linerange(aes(xmin = ll, xmax = ul)) +
     geom_vline(xintercept = 1, linetype = "dashed") +
@@
     theme_classic() +
-    theme(axis.ticks.y = element_blank(), axis.title.y = element_blank())
+    theme(
+      axis.line.y = element_blank(),
+      axis.ticks.y = element_blank(),
+      axis.text.y = element_blank(),
+      axis.title.y = element_blank(),
+      axis.line.x = element_blank()
+    ) +
+    geom_stripes()
+  # p values
+  p_right <- ggplot(coeff, aes(y = fct_rev(names))) +
+    geom_text(aes(x = 0, label = p),
+      hjust = 0,
+      fontface = ifelse(coeff$sig, "bold", "plain")
+    ) +
+    theme_void() +
+    geom_stripes()
+  # Layout
+  p_left + p + p_right +
+    plot_layout(widths = c(1, 2, 0.25), guides = "collect") &
+    theme(plot.margin = margin(0, 0, 0, 0, "pt"))
+  # Return each plot separately - will need to be combined outside of the function
+  return(c(p_left, p, p_right))
 }
@@
 # Forest plot and coefficients table
-ggforest(cox_res, data = s)
+forest_plot(cox_res)
 # Diagnostics

### 2024-02-16 Christine Chan/CF Registry/cf_registry_analysis.qmd
forest plot mostly done
had a better idea for easy forest plots, lol

@@
 forest_plot <- function(mod, time_varying = NULL, labels = labels) {
-  coeff <- tbl_regression(cox_res,
-    exponentiate = T,
+    coeff <- tbl_regression(mod,
         pvalue_fun = function(x) x
@@
         as_tibble()
-  coeff_n <- tbl_regression(cox_res, exponentiate = T) %>%
+    coeff_n <- tbl_regression(mod, exponentiate = T) %>%
         add_n(location = "label") %>%
@@
     coeff <- coeff %>% unite(N, `**N**`, N2, remove = F, na.rm = T)
-  # Separate CI column
-  coeff$hr <- as.numeric(coeff$`**HR**`)
-  coeff$ll <- as.numeric(sapply(str_split(coeff$`**95% CI**`, ", "), "[", 1))
-  coeff$ul <- as.numeric(sapply(str_split(coeff$`**95% CI**`, ", "), "[", 2))
+    # Exponentiate
+    coeff$hr <- exp(as.numeric(coeff$`**log(HR)**`))
+    coeff$exp.ll <- exp(as.numeric(sapply(str_split(coeff$`**95% CI**`, ", "), "[", 1)))
+    coeff$exp.ul <- exp(as.numeric(sapply(str_split(coeff$`**95% CI**`, ", "), "[", 2)))
     # Add N to names
@@
         theme_void() +
-    coord_cartesian(xlim = c(0, 2)) +
+        coord_cartesian(xlim = c(0, 1.5)) +
         geom_stripes()
@@
     # Layout
-  p_left + p + p_right +
+    final_p = p_left + p + p_right +
         plot_layout(widths = c(1, 2, 0.25), guides = "collect") &
         theme(plot.margin = margin(0, 0, 0, 0, "pt"))
-  # Return each plot separately - will need to be combined outside of the function
-  return(c(p_left, p, p_right))
+    # Return
+    return(final_p)
 }

### 2024-02-16 Jordana Hoppe/Home Collection/manuscript.qmd
jordana agreement manuscript

@@
 # Exclude missing subjects
-data <- data %>%
-  filter(!subject_id %in% c("009-07", "009-17", "009-25", "009-37"))
-survey_data <- survey_data %>%
-  filter(!participant_id %in% c(15, 27, 30, 34, 47))
+data_exclude <- c("009-07", "009-17", "009-25", "009-37")
+survey_exclude <- c(15, 27, 30, 34, 47)
 # Fix missing labels
@@
 survey_data %>%
+  filter(!participant_id %in% survey_exclude) %>%
   select(q1.factor:q6b___4.factor) %>%
@@
 data %>%
+  filter(!subject_id %in% data_exclude) %>%
   select(clinic_cx_source.factor:clinic_burk_sp.factor) %>%
@@
 data %>%
+  filter(!subject_id %in% data_exclude) %>%
   select(research_cx.factor:res_burk_sp.factor) %>%
@@
 data %>%
+  filter(!subject_id %in% data_exclude) %>%
   select(home_cx.factor:home_burk_sp.factor) %>%

### 2024-02-20 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
started SSwAB-IT analysis

@@
 library(redcapAPI)
+library(irr)
+library(knitr)
 home_dir <- switch(Sys.info()["sysname"],
@@
 exportBulkRecords(list(db = rcon), envir = 1)
+# Combine collection results
+df <- full_join(db_cos_collection_results, db_sos_collection_results)
+# Format
+df <- df %>%
+  select(
+    record_id, redcap_data_access_group,
+    staphylococcus_aureus_cos:other_result_cos,
+    staphylococcus_aureus_sos:other_result_sos
+  )
 ```
 
-# Feasibility
+# Agreement
+
+Cohen's kappa is interpreted analagously to a correlation coefficient, with -1 indicating perfect disagreement and 1 indicating perfect agreement.
+Tables with NaN and NAs indicate that all tests had the same values (i.e.
+all absent or all present).
+
+| Cohen's Kappa | Interpretation         |
+|---------------|------------------------|
+| 0             | No agreement           |
+| 0.10 - 0.20   | Slight agreement       |
+| 0.21 - 0.40   | Fair agreement         |
+| 0.41 - 0.60   | Moderate agreement     |
+| 0.61 - 0.80   | Substantial agreement  |
+| 0.81 - 0.99   | Near perfect agreement |
+| 1             | Perfect agreement      |
 
 ```{r}
-#| message: false
-db_feasibility %>%
-  select(
-    redcap_data_access_group, able_to_collect_sample,
-    believe_collected_good_sample, sample_difficulty, willing_to_repeat,
-  ) %>%
-  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
-  add_overall()
+agreement <- function(bug, data = df) {
+  cos <- paste0(bug, "_cos")
+  sos <- paste0(bug, "_sos")
+  d <- data[, c(cos, sos)]
+  a <- agree(d)
+  k <- kappa2(d)
+  ct <- table(d[, cos])
+  st <- table(d[, sos])
+  return(list(
+    "COS Absent" = ct[1], "COS Present" = ct[2],
+    "SOS Absent" = st[1], "SOS Present" = st[2],
+    "Percent Agreement" = a$value, "Cohen's Kappa" = k$value,
+    "p value" = k$p.value
+  ))
+}
 ```
+
+```{r}
+bugs <- c(
+  "staphylococcus_aureus", "haemophilus_influenza", "pseudomonas_aeruginosa",
+  "pseudomonas_non_aero", "s_maltophilia", "burkholderia", "aspergillus", "ntm",
+  "other"
+)
+bug_names <- c(
+  "S. aureus", "H. influenza", "P. aeruginosa", "Pseudomonas non-aeruginosa",
+  "S. maltophilia", "Burkholderia", "Aspergillus", "NTM", "Other"
+)
+bug_list = as.list(bug_names)
+names(bug_list) = bugs
+l <- lapply(bugs, function(b) {
+  agreement(b)
+})
+l <- data.frame(do.call(rbind, l), check.names = F)
+l <- data.frame(lapply(l, as.numeric), check.names = F)
+rownames(l) <- bug_names
+kable(l, digits = 2)
+```
+
+```{r results='asis'}
+invisible(lapply(names(bug_list),function(b){
+  cat("\n")
+  cat("##",as.character(bug_list[b]))
+  cat("\n")
+  cos <- paste0(b, "_cos")
+  sos <- paste0(b, "_sos")
+  d <- df[, c(cos, sos)]
+  t = data.frame(table(d))
+  colnames(t) = c("COS","SOS","n")
+  cat("\n")
+  print(kable(t))
+  cat("\n")
+}))
+```

### 2024-02-20 Christine Chan/CF Registry/cf_registry_analysis.qmd
CFRD registry with incidence

--- 2024-02-16 Christine Chan/CF Registry/cf_registry_analysis.qmd
@@
 library(survminer)
-library(icenReg)
 library(knitr)
@@
 library(broom)
+library(forestplot)
 knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry")
@@
 
-<!-- # CFRD incidence and prevalence -->
+# CFRD incidence and prevalence
 
-<!-- ```{r} -->
-<!-- # All data -->
-<!-- encounter <- read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv", -->
-<!--   na.strings = c("", "N/A") -->
-<!-- ) -->
-<!-- annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv", -->
-<!--   na.strings = c("", "N/A") -->
-<!-- ) -->
-<!-- encounter <- encounter %>% -->
-<!--   arrange(eDWID, reviewyear, encounternum) -->
-<!-- annual <- annual %>% arrange(eDWID, ReviewYear) -->
-<!-- # Add age -->
-<!-- ages <- annual %>% -->
-<!--   mutate(`Age in 2018` = 2018 - (ReviewYear - Age_YrEnd)) %>% -->
-<!--   rename(reviewyear = ReviewYear) %>% -->
-<!--   group_by(eDWID, reviewyear) %>% -->
-<!--   summarise( -->
-<!--     `Age in 2018` = as.numeric(names(sort(table(`Age in 2018`), -->
-<!--       decreasing = TRUE -->
-<!--     )[1])), -->
-<!--     Age = Age_YrEnd[1], .groups = "drop" -->
-<!--   ) -->
-<!-- encounter <- left_join(encounter, ages, by = c("eDWID", "reviewyear")) -->
-<!-- encounter$`Age Group 2018` <- cut(encounter$`Age in 2018`, -->
-<!--   c(-Inf, 10, 19, 31, Inf), -->
-<!--   labels = c("< 10", "10 - 18", "19 - 30", "> 30"), -->
-<!--   right = F -->
-<!-- ) -->
-<!-- encounter$`Age Group` <- cut(encounter$Age, -->
-<!--   c(-Inf, 10, 19, 31, Inf), -->
-<!--   labels = c("< 10", "10 - 18", "19 - 30", "> 30"), -->
-<!--   right = F -->
-<!-- ) -->
-<!-- # For those with CFRD by registry criteria, find first visit with diagnosis -->
-<!-- first_cfrd_cffpr <- encounter %>% -->
-<!--   filter(cfrd_status == 3) %>% -->
-<!--   group_by(eDWID) %>% -->
-<!--   filter(row_number() == 1) %>% -->
-<!--   select(eDWID, reviewyear, `Age in 2018`) %>% -->
-<!--   rename(first_cfrd_cffpr = reviewyear) -->
-<!-- # Add to encounter level data -->
-<!-- encounter <- left_join(encounter, first_cfrd_cffpr, -->
-<!--   by = c("eDWID", "Age in 2018") -->
-<!-- ) -->
-<!-- # Fill down CFRD Yes/No by year -->
-<!-- encounter <- encounter %>% -->
-<!--   group_by(eDWID) %>% -->
-<!--   mutate( -->
-<!--     cfrd_cffpr = case_when( -->
-<!--       reviewyear < first_cfrd_cffpr ~ "No", -->
-<!--       reviewyear >= first_cfrd_cffpr ~ "Yes" -->
-<!--     ) -->
-<!--   ) %>% -->
-<!--   ungroup() -->
-<!-- encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No" -->
-<!-- # Ordered factors -->
-<!-- encounter$cfrd_cffpr <- factor(encounter$cfrd_cffpr, -->
-<!--   levels = c("No", "Yes"), ordered = T -->
-<!-- ) -->
-<!-- # Remove those missing age -->
-<!-- encounter <- encounter %>% filter(!is.na(Age)) -->
-<!-- ``` -->
+```{r}
+# All data
+encounter <- read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv",
+  na.strings = c("", "N/A")
+)
+annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
+  na.strings = c("", "N/A")
+)
+encounter <- encounter %>%
+  arrange(eDWID, reviewyear, encounternum)
+annual <- annual %>% arrange(eDWID, ReviewYear)
+# Add age
+ages <- annual %>%
+  mutate(`Age in 2018` = 2018 - (ReviewYear - Age_YrEnd)) %>%
+  rename(reviewyear = ReviewYear) %>%
+  group_by(eDWID, reviewyear) %>%
+  summarise(
+    `Age in 2018` = as.numeric(names(sort(table(`Age in 2018`),
+      decreasing = TRUE
+    )[1])),
+    Age = Age_YrEnd[1], .groups = "drop"
+  )
+encounter <- left_join(encounter, ages, by = c("eDWID", "reviewyear"))
+encounter$`Age Group 2018` <- cut(encounter$`Age in 2018`,
+  c(-Inf, 10, 19, 31, Inf),
+  labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
+  right = F
+)
+encounter$`Age Group` <- cut(encounter$Age,
+  c(-Inf, 10, 19, 31, Inf),
+  labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
+  right = F
+)
+# For those with CFRD by registry criteria, find first visit with diagnosis
+first_cfrd_cffpr <- encounter %>%
+  filter(cfrd_status == 3) %>%
+  group_by(eDWID) %>%
+  filter(row_number() == 1) %>%
+  select(eDWID, reviewyear, `Age in 2018`) %>%
+  rename(first_cfrd_cffpr = reviewyear)
+# Add to encounter level data
+encounter <- left_join(encounter, first_cfrd_cffpr,
+  by = c("eDWID", "Age in 2018")
+)
+# Fill down CFRD Yes/No by year
+encounter <- encounter %>%
+  group_by(eDWID) %>%
+  mutate(
+    cfrd_cffpr = case_when(
+      reviewyear < first_cfrd_cffpr ~ "No",
+      reviewyear >= first_cfrd_cffpr ~ "Yes"
+    )
+  ) %>%
+  ungroup()
+encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
+# Ordered factors
+encounter$cfrd_cffpr <- factor(encounter$cfrd_cffpr,
+  levels = c("No", "Yes"), ordered = T
+)
+# Remove those missing age
+encounter <- encounter %>% filter(!is.na(Age))
+```
 
-<!-- ```{r} -->
-<!-- #| label: tbl-incidence_cffpr -->
-<!-- #| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR Alone" -->
-<!-- # For each year, pull everyone with and without CFRD that year -->
-<!-- year_inc <- function(df = encounter, year, py = 1000) { -->
-<!--   # Filter and count -->
-<!--   inc <- df %>% -->
-<!--     filter(reviewyear == year, is.na(first_cfrd_cffpr) | first_cfrd_cffpr >= year) %>% -->
-<!--     group_by(eDWID) %>% -->
-<!--     filter(row_number() == 1) %>% -->
-<!--     select(eDWID, `Age Group`, cfrd_cffpr) -->
-<!--   inc <- inc %>% -->
-<!--     group_by(cfrd_cffpr) %>% -->
-<!--     count(`Age Group`, .drop = F) -->
-<!--   inc <- inc[!is.na(inc$`Age Group`), ] -->
-<!--   # Calculations -->
-<!--   new_cases <- sum(inc$n[inc$cfrd_cffpr == "Yes"]) -->
-<!--   total_at_risk <- sum(inc$n) -->
-<!--   total_inc <- new_cases / (total_at_risk / py) -->
+```{r}
+#| label: tbl-incidence_cffpr
+#| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR Alone"
+# For each year, pull everyone with and without CFRD that year
+year_inc <- function(df = encounter, year, py = 1000) {
+  # Filter and count
+  inc <- df %>%
+    filter(reviewyear == year, is.na(first_cfrd_cffpr) | first_cfrd_cffpr >= year) %>%
+    group_by(eDWID) %>%
+    filter(row_number() == 1) %>%
+    select(eDWID, `Age Group`, cfrd_cffpr)
+  inc <- inc %>%
+    group_by(cfrd_cffpr) %>%
+    count(`Age Group`, .drop = F)
+  inc <- inc[!is.na(inc$`Age Group`), ]
+  # Calculations
+  new_cases <- sum(inc$n[inc$cfrd_cffpr == "Yes"])
+  total_at_risk <- sum(inc$n)
+  total_inc <- new_cases / (total_at_risk / py)
 
-<!--   new_cases_10 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "< 10"]) -->
-<!--   at_risk_10 <- sum(inc$n[inc$`Age Group` == "< 10"]) -->
-<!--   inc_10 <- new_cases_10 / (at_risk_10 / py) -->
+  new_cases_10 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "< 10"])
+  at_risk_10 <- sum(inc$n[inc$`Age Group` == "< 10"])
+  inc_10 <- new_cases_10 / (at_risk_10 / py)
 
-<!--   new_cases_10_18 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "10 - 18"]) -->
-<!--   at_risk_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"]) -->
-<!--   inc_10_18 <- new_cases_10_18 / (at_risk_10_18 / py) -->
+  new_cases_10_18 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "10 - 18"])
+  at_risk_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
+  inc_10_18 <- new_cases_10_18 / (at_risk_10_18 / py)
 
-<!--   new_cases_19_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "19 - 30"]) -->
-<!--   at_risk_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"]) -->
-<!--   inc_19_30 <- new_cases_19_30 / (at_risk_19_30 / py) -->
+  new_cases_19_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "19 - 30"])
+  at_risk_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
+  inc_19_30 <- new_cases_19_30 / (at_risk_19_30 / py)
 
-<!--   new_cases_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "> 30"]) -->
-<!--   at_risk_30 <- sum(inc$n[inc$`Age Group` == "> 30"]) -->
-<!--   inc_30 <- new_cases_30 / (at_risk_30 / py) -->
+  new_cases_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "> 30"])
+  at_risk_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
+  inc_30 <- new_cases_30 / (at_risk_30 / py)
 
-<!--   # Nice results -->
-<!--   list( -->
-<!--     `Year` = year, -->
-<!--     `Total New Cases` = new_cases, -->
-<!--     `Total At Risk` = total_at_risk, -->
-<!--     `Total Incidence` = total_inc, -->
-<!--     `Total New Cases Age < 10` = new_cases_10, -->
-<!--     `Total At Risk Age < 10` = at_risk_10, -->
-<!--     `Total Incidence Age < 10` = inc_10, -->
-<!--     `Total New Cases Age 10-18` = new_cases_10_18, -->
-<!--     `Total At Risk Age 10-18` = at_risk_10_18, -->
-<!--     `Total Incidence Age 10-18` = inc_10_18, -->
-<!--     `Total New Cases Age 19-30` = new_cases_19_30, -->
-<!--     `Total At Risk Age 19-30` = at_risk_19_30, -->
-<!--     `Total Incidence Age 19-30` = inc_19_30, -->
-<!--     `Total New Cases Age > 30` = new_cases_30, -->
-<!--     `Total At Risk Age > 30` = at_risk_30, -->
-<!--     `Total Incidence Age > 30` = inc_30 -->
-<!--   ) -->
-<!-- } -->
-<!-- incidence <- lapply(2003:2018, function(y) { -->
-<!--   year_inc(year = y) -->
-<!-- }) -->
-<!-- incidence <- data.frame(do.call(rbind, incidence), check.names = F) -->
-<!-- incidence <- data.frame(lapply(incidence, as.numeric), check.names = F) -->
-<!-- kable(incidence, digits = 3) -->
-<!-- ``` -->
+  # Nice results
+  list(
+    `Year` = year,
+    `Total New Cases` = new_cases,
+    `Total At Risk` = total_at_risk,
+    `Total Incidence` = total_inc,
+    `Total New Cases Age < 10` = new_cases_10,
+    `Total At Risk Age < 10` = at_risk_10,
+    `Total Incidence Age < 10` = inc_10,
+    `Total New Cases Age 10-18` = new_cases_10_18,
+    `Total At Risk Age 10-18` = at_risk_10_18,
+    `Total Incidence Age 10-18` = inc_10_18,
+    `Total New Cases Age 19-30` = new_cases_19_30,
+    `Total At Risk Age 19-30` = at_risk_19_30,
+    `Total Incidence Age 19-30` = inc_19_30,
+    `Total New Cases Age > 30` = new_cases_30,
+    `Total At Risk Age > 30` = at_risk_30,
+    `Total Incidence Age > 30` = inc_30
+  )
+}
+incidence <- lapply(2003:2018, function(y) {
+  year_inc(year = y)
+})
+incidence <- data.frame(do.call(rbind, incidence), check.names = F)
+incidence <- data.frame(lapply(incidence, as.numeric), check.names = F)
+kable(incidence, digits = 3)
+```
 
-<!-- In @tbl-incidence_cffpr, incidence per year is calculated as the number of new CFRD diagnoses (by registry definition alone) per 1,000 person-years of those at risk, excluding previously identified cases.
--->
+In @tbl-incidence_cffpr, incidence per year is calculated as the number of new CFRD diagnoses (by registry definition alone) per 1,000 person-years of those at risk, excluding previously identified cases.
 
-<!-- ```{r} -->
-<!-- #| label: tbl-prevalence_cffpr -->
-<!-- #| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR Alone" -->
-<!-- encounter %>% -->
-<!--   group_by(eDWID, reviewyear) %>% -->
-<!--   summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>% -->
-<!--   group_by(reviewyear) %>% -->
-<!--   summarise( -->
-<!--     Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100, -->
-<!--     .groups = "drop" -->
-<!--   ) %>% -->
-<!--   kable(., digits = 3) -->
-<!-- ``` -->
+```{r}
+#| label: tbl-prevalence_cffpr
+#| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR Alone"
+encounter %>%
+  group_by(eDWID, reviewyear) %>%
+  summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>%
+  group_by(reviewyear) %>%
+  summarise(
+    Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100,
+    .groups = "drop"
+  ) %>%
+  kable(., digits = 3)
+```
 
-<!-- ```{r} -->
-<!-- #| label: tbl-prevalence_cffpr_year -->
-<!-- #| tbl-cap: "Yearly Prevalence of CFRD by Age Group Based on CFFPR Alone" -->
-<!-- encounter %>% -->
-<!--   group_by(eDWID, reviewyear, `Age Group`) %>% -->
-<!--   summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>% -->
-<!--   group_by(reviewyear, `Age Group`) %>% -->
-<!--   summarise(Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100, .groups = "drop") %>% -->
-<!--   pivot_wider(names_from = `Age Group`, values_from = Prevalence) %>% -->
-<!--   kable(., digits = 3) -->
-<!-- ``` -->
+```{r}
+#| label: tbl-prevalence_cffpr_year
+#| tbl-cap: "Yearly Prevalence of CFRD by Age Group Based on CFFPR Alone"
+encounter %>%
+  group_by(eDWID, reviewyear, `Age Group`) %>%
+  summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>%
+  group_by(reviewyear, `Age Group`) %>%
+  summarise(Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100, .groups = "drop") %>%
+  pivot_wider(names_from = `Age Group`, values_from = Prevalence) %>%
+  kable(., digits = 3)
+```
 
-<!-- In @tbl-prevalence_cffpr_year and @tbl-prevalence_cffpr, prevalence is calculated as the total number of CFRD cases (by registry definition alone) seen in clinic divided by the total number of people seen in clinic in a given year.
--->
+In @tbl-prevalence_cffpr_year and @tbl-prevalence_cffpr, prevalence is calculated as the total number of CFRD cases (by registry definition alone) seen in clinic divided by the total number of people seen in clinic in a given year.
 
-<!-- # CFRD screening rates -->
+# CFRD screening rates
 
-<!-- ```{r} -->
-<!-- #| label: tbl-screening_ogtt -->
-<!-- #| tbl-cap: "Yearly OGTT Screening Rates" -->
-<!-- year_screen <- function(df = encounter, year) { -->
-<!--   # Filter and count -->
-<!--   inc <- df %>% -->
-<!--     filter(reviewyear == year, first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>% -->
-<!--     mutate(ogtt = !is.na(ogttfast_bloodglucose) | !is.na(twohour_bloodglucose)) %>% -->
-<!--     group_by(eDWID) %>% -->
-<!--     summarise(ogtt = max(ogtt), `Age Group` = `Age Group`[1]) -->
-<!--   inc <- inc %>% -->
-<!--     group_by(ogtt) %>% -->
-<!--     count(`Age Group`) %>% -->
-<!--     filter(`Age Group` != "< 10") -->
-<!--   # Calculations -->
-<!--   total_screened <- sum(inc$n[inc$ogtt == 1]) -->
-<!--   total_seen <- sum(inc$n) -->
-<!--   screen_rate <- total_screened / total_seen * 100 -->
+```{r}
+#| label: tbl-screening_ogtt
+#| tbl-cap: "Yearly OGTT Screening Rates"
+year_screen <- function(df = encounter, year) {
+  # Filter and count
+  inc <- df %>%
+    filter(reviewyear == year, first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>%
+    mutate(ogtt = !is.na(ogttfast_bloodglucose) | !is.na(twohour_bloodglucose)) %>%
+    group_by(eDWID) %>%
+    summarise(ogtt = max(ogtt), `Age Group` = `Age Group`[1])
+  inc <- inc %>%
+    group_by(ogtt) %>%
+    count(`Age Group`) %>%
+    filter(`Age Group` != "< 10")
+  # Calculations
+  total_screened <- sum(inc$n[inc$ogtt == 1])
+  total_seen <- sum(inc$n)
+  screen_rate <- total_screened / total_seen * 100
 
-<!--   screened_10_18 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "10 - 18"] -->
-<!--   seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"]) -->
-<!--   rate_10_18 <- screened_10_18 / seen_10_18 * 100 -->
+  screened_10_18 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "10 - 18"]
+  seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
+  rate_10_18 <- screened_10_18 / seen_10_18 * 100
 
-<!--   screened_19_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "19 - 30"] -->
-<!--   seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"]) -->
-<!--   rate_19_30 <- screened_19_30 / seen_19_30 * 100 -->
+  screened_19_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "19 - 30"]
+  seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
+  rate_19_30 <- screened_19_30 / seen_19_30 * 100
 
-<!--   screened_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "> 30"] -->
-<!--   seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"]) -->
-<!--   rate_30 <- screened_30 / seen_30 * 100 -->
+  screened_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "> 30"]
+  seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
+  rate_30 <- screened_30 / seen_30 * 100
 
-<!--   # Nice results -->
-<!--   list( -->
-<!--     `Year` = year, -->
-<!--     `Total Screened` = total_screened, -->
-<!--     `Total Seen` = total_seen, -->
-<!--     `Screen Rate (%)` = screen_rate, -->
-<!--     `Total Screened Age 10-18` = screened_10_18, -->
-<!--     `Total Seen Age 10-18` = seen_10_18, -->
-<!--     `Screen Rate Age 10-18 (%)` = rate_10_18, -->
-<!--     `Total Screened Age 19-30` = screened_19_30, -->
-<!--     `Total Seen Age 19-30` = seen_19_30, -->
-<!--     `Screen Rate Age 19-30 (%)` = rate_19_30, -->
-<!--     `Total Screened Age > 30` = screened_30, -->
-<!--     `Total Seen Age > 30` = seen_30, -->
-<!--     `Screen Rate Age > 30 (%)` = rate_30 -->
-<!--   ) -->
-<!-- } -->
-<!-- screen <- lapply(2003:2018, function(y) { -->
-<!--   year_screen(year = y) -->
-<!-- }) -->
-<!-- screen <- data.frame(do.call(rbind, screen), check.names = F) -->
-<!-- screen <- data.frame(lapply(screen, as.numeric), check.names = F) -->
-<!-- kable(screen, digits = 3) -->
-<!-- ``` -->
+  # Nice results
+  list(
+    `Year` = year,
+    `Total Screened` = total_screened,
+    `Total Seen` = total_seen,
+    `Screen Rate (%)` = screen_rate,
+    `Total Screened Age 10-18` = screened_10_18,
+    `Total Seen Age 10-18` = seen_10_18,
+    `Screen Rate Age 10-18 (%)` = rate_10_18,
+    `Total Screened Age 19-30` = screened_19_30,
+    `Total Seen Age 19-30` = seen_19_30,
+    `Screen Rate Age 19-30 (%)` = rate_19_30,
+    `Total Screened Age > 30` = screened_30,
+    `Total Seen Age > 30` = seen_30,
+    `Screen Rate Age > 30 (%)` = rate_30
+  )
+}
+screen <- lapply(2003:2018, function(y) {
+  year_screen(year = y)
+})
+screen <- data.frame(do.call(rbind, screen), check.names = F)
+screen <- data.frame(lapply(screen, as.numeric), check.names = F)
+kable(screen, digits = 3)
+```
 
-<!-- ```{r} -->
-<!-- #| label: tbl-screening_a1c -->
-<!-- #| tbl-cap: "Yearly HbA1c Screening Rates" -->
-<!-- year_screen <- function(df = encounter, year) { -->
-<!--   # Filter and count -->
-<!--   inc <- df %>% -->
-<!--     filter(reviewyear == year, first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>% -->
-<!--     mutate(a1c = !is.na(hgba1c)) %>% -->
-<!--     group_by(eDWID) %>% -->
-<!--     summarise(a1c = max(a1c), `Age Group` = `Age Group`[1]) -->
-<!--   inc <- inc %>% -->
-<!--     group_by(a1c) %>% -->
-<!--     count(`Age Group`) %>% -->
-<!--     filter(`Age Group` != "< 10") -->
-<!--   # Calculations -->
-<!--   total_screened <- sum(inc$n[inc$a1c == 1]) -->
-<!--   total_seen <- sum(inc$n) -->
-<!--   screen_rate <- total_screened / total_seen * 100 -->
+```{r}
+#| label: tbl-screening_a1c
+#| tbl-cap: "Yearly HbA1c Screening Rates"
+year_screen <- function(df = encounter, year) {
+  # Filter and count
+  inc <- df %>%
+    filter(reviewyear == year, first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>%
+    mutate(a1c = !is.na(hgba1c)) %>%
+    group_by(eDWID) %>%
+    summarise(a1c = max(a1c), `Age Group` = `Age Group`[1])
+  inc <- inc %>%
+    group_by(a1c) %>%
+    count(`Age Group`) %>%
+    filter(`Age Group` != "< 10")
+  # Calculations
+  total_screened <- sum(inc$n[inc$a1c == 1])
+  total_seen <- sum(inc$n)
+  screen_rate <- total_screened / total_seen * 100
 
-<!--   screened_10_18 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "10 - 18"] -->
-<!--   seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"]) -->
-<!--   rate_10_18 <- screened_10_18 / seen_10_18 * 100 -->
+  screened_10_18 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "10 - 18"]
+  seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
+  rate_10_18 <- screened_10_18 / seen_10_18 * 100
 
-<!--   screened_19_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "19 - 30"] -->
-<!--   seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"]) -->
-<!--   rate_19_30 <- screened_19_30 / seen_19_30 * 100 -->
+  screened_19_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "19 - 30"]
+  seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
+  rate_19_30 <- screened_19_30 / seen_19_30 * 100
 
-<!--   screened_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "> 30"] -->
-<!--   seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"]) -->
-<!--   rate_30 <- screened_30 / seen_30 * 100 -->
+  screened_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "> 30"]
+  seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
+  rate_30 <- screened_30 / seen_30 * 100
 
-<!--   # Nice results -->
-<!--   list( -->
-<!--     `Year` = year, -->
-<!--     `Total Screened` = total_screened, -->
-<!--     `Total Seen` = total_seen, -->
-<!--     `Screen Rate (%)` = screen_rate, -->
-<!--     `Total Screened Age 10-18` = screened_10_18, -->
-<!--     `Total Seen Age 10-18` = seen_10_18, -->
-<!--     `Screen Rate Age 10-18 (%)` = rate_10_18, -->
-<!--     `Total Screened Age 19-30` = screened_19_30, -->
-<!--     `Total Seen Age 19-30` = seen_19_30, -->
-<!--     `Screen Rate Age 19-30 (%)` = rate_19_30, -->
-<!--     `Total Screened Age > 30` = screened_30, -->
-<!--     `Total Seen Age > 30` = seen_30, -->
-<!--     `Screen Rate Age > 30 (%)` = rate_30 -->
-<!--   ) -->
-<!-- } -->
-<!-- screen <- lapply(2003:2018, function(y) { -->
-<!--   year_screen(year = y) -->
-<!-- }) -->
-<!-- screen <- data.frame(do.call(rbind, screen), check.names = F) -->
-<!-- screen <- data.frame(lapply(screen, as.numeric), check.names = F) -->
-<!-- kable(screen, digits = 3) -->
-<!-- ``` -->
+  # Nice results
+  list(
+    `Year` = year,
+    `Total Screened` = total_screened,
+    `Total Seen` = total_seen,
+    `Screen Rate (%)` = screen_rate,
+    `Total Screened Age 10-18` = screened_10_18,
+    `Total Seen Age 10-18` = seen_10_18,
+    `Screen Rate Age 10-18 (%)` = rate_10_18,
+    `Total Screened Age 19-30` = screened_19_30,
+    `Total Seen Age 19-30` = seen_19_30,
+    `Screen Rate Age 19-30 (%)` = rate_19_30,
+    `Total Screened Age > 30` = screened_30,
+    `Total Seen Age > 30` = seen_30,
+    `Screen Rate Age > 30 (%)` = rate_30
+  )
+}
+screen <- lapply(2003:2018, function(y) {
+  year_screen(year = y)
+})
+screen <- data.frame(do.call(rbind, screen), check.names = F)
+screen <- data.frame(lapply(screen, as.numeric), check.names = F)
+kable(screen, digits = 3)
+```
 
-<!-- In @tbl-screening_ogtt and @tbl-screening_a1c, people with CFRD (by registry definition) and those under age 10 were excluded.
+In @tbl-screening_ogtt and @tbl-screening_a1c, people with CFRD (by registry definition) and those under age 10 were excluded.
 Screening rate is calculated as the number of people with an OGTT or HbA1c, respectively, divided by the total number of people seen in clinic that year.
--->
 
@@
 ```{r}
-forest_plot <- function(mod, time_varying = NULL, labels = labels) {
-  coeff <- data.frame(summary(mod)$coefficients)
-  coeff$ll <- coeff$coef - (1.96 * coeff$se.coef.)
-  coeff$ul <- coeff$coef + (1.96 * coeff$se.coef.)
-  # Clean up variable names
-  l <- unlist(sapply(labels, paste0, ": "))
-  coeff$names <- rownames(coeff)
-  coeff$names <- str_replace_all(coeff$names, l)
+tv_forest_plot <- function(mod, time_varying = NULL, time_variable = "tgroup",
+                           labels = labels) {
+  coeff <- tbl_regression(mod, pvalue_fun = function(x) x) %>%
+    add_n(location = "level") %>%
+    add_nevent(location = "level") %>%
+    as_tibble()
+  # If time-varying, format appropriately
+  if (!is.null(time_varying)) {
+    # Split table into time-varying and not
+    w <- which(coeff$`**Characteristic**` == labels[time_varying])
+    coeff <- coeff[w:nrow(coeff), ]
+    t <- coeff %>% pivot_longer(`**Characteristic**`)
+  }
+  coeff_n <- tbl_regression(mod, exponentiate = T) %>%
+    add_n(location = "label") %>%
+    as_tibble()
+  coeff$N2 <- coeff_n$`**N**`
+  # Combine Ns
+  coeff$N <- pmax(coeff$`**N**`, coeff$N2, na.rm = T)
+
+  # Exponentiate
+  coeff$hr <- exp(as.numeric(coeff$`**log(HR)**`))
+  coeff$exp.ll <- exp(as.numeric(sapply(str_split(coeff$`**95% CI**`, ", "), "[", 1)))
+  coeff$exp.ul <- exp(as.numeric(sapply(str_split(coeff$`**95% CI**`, ", "), "[", 2)))
+  # Add N to names
+  coeff$names <- paste0(coeff$`**Characteristic**`, " (N = ", coeff$N, ")")
   coeff$names <- ordered(coeff$names, levels = unique(coeff$names))
-  # Exponentiate
-  coeff$exp.ll <- exp(coeff$ll)
-  coeff$exp.ul <- exp(coeff$ul)
-  # Plot
-  p <- ggplot(coeff, aes(y = fct_rev(names))) +
-    geom_point(aes(x = exp.coef.), shape = 15, size = 3) +
-    geom_linerange(aes(xmin = exp.ll, xmax = exp.ul)) +
-    geom_vline(xintercept = 1, linetype = "dashed") +
-    xlab("Hazard Ratio") +
-    theme_classic() +
-    theme(axis.ticks.y = element_blank(), axis.title.y = element_blank())
+  # P values
+  coeff$sig <- coeff$`**p-value**` < 0.05
+  coeff$p <- style_pvalue(coeff$`**p-value**`)
+  # Row types
+  coeff$type <- "level"
+  coeff$type[!is.na(coeff$N2)] <- "variable"
+  coeff$type[!is.na(coeff$`**N**`) & is.na(coeff$hr)] <- "reference"
+  # HR label
+  coeff$hr_lab <- paste0(
+    round(coeff$hr, 2), " (", round(coeff$exp.ll, 2), " - ",
+    round(coeff$exp.ul, 2), ")"
+  )
+  coeff$hr_lab[coeff$hr_lab == "NA (NA - NA)"] <- NA
+  coeff$hr_lab[coeff$type == "reference"] <- "Reference"
+  #
 }
@@
 # Forest plot and coefficients table
-ggforest(cox_res, data = s)
+plot_df <- tbl_regression(cox_res, exponentiate = T) %>%
+  add_n(location = "level") %>%
+  add_nevent(location = "level")
+plot_df <- plot_df %>% as_tibble()
+plot_df$expll <- as.numeric(sapply(strsplit(plot_df$`**95% CI**`, ", "), "[", 1))
+plot_df$expul <- as.numeric(sapply(strsplit(plot_df$`**95% CI**`, ", "), "[", 2))
+plot_df$hr <- as.numeric(plot_df$`**HR**`)
+plot_df %>%
+  forestplot(
+    labeltext = c(`**Characteristic**`, `**N**`, `**Event N**`, `**HR**`),
+    mean = "hr", lower = "expll", upper = "expul", xlog = T,
+    is.summary = is.na(plot_df$`**N**`)
+  ) %>%
+  fp_add_header(
+    `**Characteristic**` = c("", "Characteristic"),
+    `**N**` = c("", "N"),
+    `**Event N**` = c("", "Event N"),
+    `**HR**` = c("", "HR")
+  ) %>%
+  fp_add_lines() %>%
+  fp_set_zebra_style("#EFEFEF", ignore_subheaders = T)
 # Diagnostics
@@
 )
-kable(summary(cox_res_split)$coefficients, digits = 3)
+tbl_regression(cox_res_split) %>%
+  add_n(location = "level") %>%
+  add_nevent(location = "level")
 ```
@@
   cox_res_split <- coxph(new_f, data = s_split, cluster = eDWID)
-  res <- summary(cox_res_split)$coefficients
   cat("\n")
@@
   cat("\n")
-  print(kable(res, digits = 3))
+  plot_df <- tbl_regression(cox_res_split) %>%
+    add_n(location = "level") %>%
+    add_nevent(location = "level")%>% 
+    as_tibble()
+  plot_df$ll <- as.numeric(sapply(strsplit(plot_df$`**95% CI**`, ", "), "[", 1))
+  plot_df$ul <- as.numeric(sapply(strsplit(plot_df$`**95% CI**`, ", "), "[", 2))
+  plot_df$hr <- as.numeric(plot_df$`**log(HR)**`)
+  plot_df %>%
+    forestplot(
+      labeltext = c(`**Characteristic**`, `**N**`, `**Event N**`, `**log(HR)**`),
+      mean = "hr", lower = "ll", upper = "ul", 
+      is.summary = is.na(plot_df$`**N**`)
+    ) %>%
+    fp_add_header(
+      `**Characteristic**` = c("", "Characteristic"),
+      `**N**` = c("", "N"),
+      `**Event N**` = c("", "Event N"),
+      `**log(HR)**` = c("", "log(HR)")
+    ) %>%
+    fp_add_lines() %>%
+    fp_set_zebra_style("#EFEFEF", ignore_subheaders = T) %>%
+    print()
+  cat("\n")
   cat("\n")

### 2024-02-28 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
kam eti update

--- 2024-01-03 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
@@
       dark: darkly
-bibliography: C:/Users/timvigers/Dropbox/Miscellaneous/zotero.bib
-csl: C:/Users/timvigers/Dropbox/Miscellaneous/american-medical-association.csl
+bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
+csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
 editor: source
@@
 
-## GLI-2012 vs. GLI-2022
-
-GLI-2012 spirometry equations include race/ethnicity in the formula, but GLI-2022 equations are race-free.
-The results of the GLI-2022 equations contain the "gl" suffix, which stands for "global" (so instead of "ppFEV1" it would be "ppFEV1gl").
-
-First, I wanted to ensure that the `rspiro` package is calculating FEV1 percent predicted correctly.
-In the graph below, ppFEV1 from the registry is shown on the x-axis and ppFEV1 as calculated with `rspiro` is shown on the y-axis.
-
-```{r}
-#| warning: false
-ggplot(encounter, aes(
-  x = GLI_FEV1_pct_predicted, y = GLI_FEV1_pct_predicted_check,
-  color = race
-)) +
-  geom_point() +
-  geom_abline(slope = 1, intercept = 0) +
-  theme_bw() +
-  xlab("Registry") +
-  ylab("rspiro") +
-  ylim(0, 150) +
-  facet_wrap(~race) +
-  ggtitle("Registry values vs. rspiro")
-```
-
-```{r}
-ggplot(encounter, aes(
-  x = ppFEV1gl, y = GLI_FEV1_pct_predicted_check,
-  color = race
-)) +
-  geom_point() +
-  geom_abline(slope = 1, intercept = 0) +
-  theme_bw() +
-  xlab("GLI-2022") +
-  ylab("GLI-2012") +
-  ylim(0, 150) +
-  xlim(0, 150) +
-  facet_wrap(~race) +
-  ggtitle("GLI-2022 vs. GLI-2012 (rspiro)")
-```
-
 ## Annualized averages
@@
 The most basic model we can fit to these data uses the average values as the outcome, with terms for review year, race, and their interaction and a random intercept for each patient.
+This model is adjusted for age, private vs.
+other insurance, and parental education.
 
@@
 res <- tidy(mod, effects = "fixed")
-means <- data.frame(emmeans(mod, ~ race + ReviewYear,
+means <- data.frame(emmeans(mod, ~ race | ReviewYear,
   lmerTest.limit = 1e6,
@@
 ))
-label(means) <- list("Race/Ethnicity", "Year", "Mean ppFEV", "SE", "df", "Lower CL", "Upper CL")
-gt(means) %>% fmt_number(emmean:last_col(), decimals = 1)
 ```
-
-```{r}
-ggplot(left_join(annual, means, by = join_by(race, ReviewYear))) +
-  geom_point(mapping = aes(x = ReviewYear, y = emmean, color = race)) +
-  geom_line(mapping = aes(x = ReviewYear, y = emmean, group = race, color = race)) +
-  geom_errorbar(mapping = aes(
-    x = ReviewYear, ymin = emmean - SE,
-    ymax = emmean + SE, color = race
-  ), width = .2) +
-  xlab("Year") +
-  ylab("LS Mean FEV1") +
-  theme_bw() +
-  scale_color_discrete(name = "Race/Ethnicity")
-```
-
-```{r}
-#| warning: false
-ggplot(left_join(annual, means, by = join_by(race, ReviewYear))) +
-  geom_line(
-    mapping = aes(x = ReviewYear, y = mean_ppFEV, group = eDWID),
-    alpha = 0.005, linewidth = 0.5
-  ) +
-  geom_point(mapping = aes(x = ReviewYear, y = emmean, color = race)) +
-  geom_line(mapping = aes(x = ReviewYear, y = emmean, group = race, color = race)) +
-  geom_errorbar(mapping = aes(x = ReviewYear, ymin = emmean - SE, ymax = emmean + SE, color = race), width = .2) +
-  xlab("Year") +
-  ylab("LS Mean FEV1") +
-  theme_bw() +
-  scale_color_discrete(name = "Race/Ethnicity")
-```
-
-```{r}
-#| warning: false
-ggplot(left_join(annual, means, by = join_by(race, ReviewYear))) +
-  geom_line(
-    mapping = aes(x = ReviewYear, y = mean_ppFEV, group = eDWID),
-    alpha = 0.005, linewidth = 0.5
-  ) +
-  geom_point(mapping = aes(x = ReviewYear, y = emmean, color = race)) +
-  geom_line(mapping = aes(
-    x = ReviewYear, y = emmean, group = race,
-    color = race
-  )) +
-  geom_errorbar(mapping = aes(
-    x = ReviewYear, ymin = emmean - SE,
-    ymax = emmean + SE, color = race
-  ), width = .2) +
-  xlab("Year") +
-  ylab("LS Mean FEV1") +
-  theme_bw() +
-  scale_color_discrete(name = "Race/Ethnicity") +
-  facet_wrap(~race)
-```
-
-#### Change point model
-
-```{r}
-# Calculate time from first ETI use
-encounter <- encounter %>%
-  mutate(
-    days_from_start =
-      as.numeric(encounterdate - Modulator_trikafta_first_date)
-  )
-# Segmented regression by group
-lm_fixed_cp <- segreg(
-  GLI_FEV1_pct_predicted ~ race + seg(days_from_start, by = race),
-  data = encounter
-)
-# Results
-lm_fixed_cp$psi
-slope(lm_fixed_cp)
-# Plot
-new_df = lm_fixed_cp$model
-new_df$pred = predict(lm_fixed_cp)
-ggplot(new_df, aes(x = days_from_start, y = pred,group = race,color = race)) +
-  geom_line() +
-  ylim(0, 150)
-# # Random intercept model isn't converging
-# # Basic model
-# lmm_base <- lme(GLI_FEV1_pct_predicted ~ days_from_start,
-#   random = ~ 1 | eDWID, data = encounter, na.action = na.omit,method = "ML",
-#   control=list(msMaxIter=1000,msMaxEval = 1000)
-# )
-# # Segmented with fixed change point
-# lmm_fixed_cp <- segmented(lmm_base,
-#   seg.Z = ~days_from_start, psi = 0,
-#   random = list(eDWID = pdDiag(~ 1 + days_from_start + U))
-# )
-```
-
-```{r}
-#| cache: true
-t <- encounter %>%
-  select(eDWID, GLI_FEV1_pct_predicted, days_from_start) %>%
-  drop_na()
-model <- list(
-  GLI_FEV1_pct_predicted ~ 0 + days_from_start,
-  ~ 0 + days_from_start
-)
-fit <- mcp(model, data = t, cores = 4, chains = 4)
-plot(fit)
-```

### 2024-02-28 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R

@@ -4,9 +4,9 @@ library(readxl)
 library(rspiro)
 library(Hmisc)
 # Import
-annual <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_Annualized_Del1.csv", na.strings = "")
-encounter <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_encountersMerged_Del1.csv", na.strings = "")
-demo <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20230420/CFF21_DemogCFDiag_Del1.csv", na.strings = "")
+annual <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_Annualized_Del1.csv", na.strings = "")
+encounter <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_encountersMerged_Del1.csv", na.strings = "")
+demo <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_DemogCFDiag_Del1.csv", na.strings = "")
 # Import responsive mutations list
 responsive_mutations <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/eti_responsive_mutations.csv", header = F)
 responsive_mutations <- unique(responsive_mutations$V1)
[…]

### 2024-02-28 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
sswab it phase 1

@@
 
+## Overall
+
+## By bug
+
 ```{r}

### 2024-03-04 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
paths

@@
   levels = 1:3,
-  labels = c("Low","Medium","High")
+  labels = c("Low", "Medium", "High")
 )

### 2024-03-05 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
sweat test analysis formatting

@@
 )
+```
+
+### By genotype severity (CF only)
+
+```{r results='asis'}
+t1 <- tableby(
+  geno_risk ~ Cl + Na + `Na/Cl Ratio` + `Sweat Chloride Group`,
+  data = t1_df[t1_df$Diagnosis == "Cystic Fibrosis", ], control = mycontrols,
+)
+summary(t1,
+  pfootnote = T,
+  labelTranslations = labels
+)
+```
+
+### Pairwise comparisons
+
+Pairwise comparisons using Wilcoxon rank sum test with continuity correction
+
+```{r}
+pairwise.wilcox.test(t1_df$Cl, t1_df$Diagnosis, p.adjust.method = "fdr")
+pairwise.wilcox.test(t1_df$Na, t1_df$Diagnosis, p.adjust.method = "fdr")
+pairwise.wilcox.test(t1_df$`Na/Cl Ratio`, t1_df$Diagnosis, p.adjust.method = "fdr")
 ```

### 2024-03-05 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
sweat test analysis to Quarto

--- 2024-03-05 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
@@
 author: "Tim Vigers"
-date: "`r format(Sys.time(), '%d %B %Y')`"
-output:
-  html_document:
-    number_sections: yes
-    toc: yes
-    toc_depth: 5
-    toc_float: yes
+date: "today"
+date-format: long
+format:
+  html:
+    toc: true
+    toc-depth: 5
+    toc-float: true
+    code-fold: true
+    self-contained: true
+    fig-cap-location: top
+    page-layout: full
+    theme:
+      light: flatly
+      dark: darkly
+bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
+csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
+editor: source
 ---
 
-```{r setup, include=FALSE}
+```{r setup}
+#| include: false
 library(DiagrammeR)
@@
 
-```{r}
+```{r data cleaning}
 # Import exclusion data from Excel

### 2024-03-05 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
cotton conference changes done

@@
     toc: true
-    toc-depth: 5
+    toc-depth: 10
     toc-float: true
+    number-sections: true
     code-fold: true
@@
 
-# Flowchart
+<!-- # Flowchart -->
 
-```{r}
-grViz("digraph {
+<!-- ```{r} -->
+<!-- grViz("digraph { -->
 
-# initiate graph
-graph [layout = dot]
+<!-- # initiate graph -->
+<!-- graph [layout = dot] -->
 
-# global node settings
-node [shape = rectangle]
+<!-- # global node settings -->
+<!-- node [shape = rectangle] -->
 
-# label nodes
-data1 [label = '@@1']
-data2 [label = '@@2']
-ineligible [label = '@@3']
-demo  [label = '@@4']
-highna  [label = '@@5']
-highcl  [label = '@@6']
-highnacl  [label = '@@7']
-review [label = '@@8']
-analysis [label = '@@9']
+<!-- # label nodes -->
+<!-- data1 [label = '@@1'] -->
+<!-- data2 [label = '@@2'] -->
+<!-- ineligible [label = '@@3'] -->
+<!-- demo  [label = '@@4'] -->
+<!-- highna  [label = '@@5'] -->
+<!-- highcl  [label = '@@6'] -->
+<!-- highnacl  [label = '@@7'] -->
+<!-- review [label = '@@8'] -->
+<!-- analysis [label = '@@9'] -->
 
-# edge definitions with the node IDs
-data1  -> data2
-data2 -> ineligible
-ineligible -> {demo,highna,highcl,highnacl,review}
-data2  -> analysis
-highnacl -> analysis [style=invis]
-}
-[1]: paste0('Sweat tests performed \\n ','n = ',n1)
-[2]: paste0('Unique participants \\n ','n = ',n2)
-[3]: paste0('Ineligible \\n ','n = ',n2-nrow(t1_df))
-[4]: paste0('No demographics \\n ','n = ',length(no_dem))
-[5]: paste0('Na > 160 mmol/L \\n ','n = ',length(high_na))
-[6]: paste0('Cl > 160 mmol/L \\n ','n = ',length(high_cl))
-[7]: paste0('Na:Cl > 2 \\n ','n = ',length(high_ratio))
-[8]: paste0('Other \\n ','n = ',1)
-[9]: paste0('Analysis cohort \\n ','n = ',nrow(t1_df))
-")
-```
+<!-- # edge definitions with the node IDs -->
+<!-- data1  -> data2 -->
+<!-- data2 -> ineligible -->
+<!-- ineligible -> {demo,highna,highcl,highnacl,review} -->
+<!-- data2  -> analysis -->
+<!-- highnacl -> analysis [style=invis] -->
+<!-- } -->
+<!-- [1]: paste0('Sweat tests performed \\n ','n = ',n1) -->
+<!-- [2]: paste0('Unique participants \\n ','n = ',n2) -->
+<!-- [3]: paste0('Ineligible \\n ','n = ',n2-nrow(t1_df)) -->
+<!-- [4]: paste0('No demographics \\n ','n = ',length(no_dem)) -->
+<!-- [5]: paste0('Na > 160 mmol/L \\n ','n = ',length(high_na)) -->
+<!-- [6]: paste0('Cl > 160 mmol/L \\n ','n = ',length(high_cl)) -->
+<!-- [7]: paste0('Na:Cl > 2 \\n ','n = ',length(high_ratio)) -->
+<!-- [8]: paste0('Other \\n ','n = ',1) -->
+<!-- [9]: paste0('Analysis cohort \\n ','n = ',nrow(t1_df)) -->
+<!-- ") -->
+<!-- ``` -->
 
@@
 
+##### Na/Cl Ratio
+
 ```{r}
@@
 plot(roc_cf, print.auc = TRUE, col = "blue")
-plot(roc_non_cf, print.auc = TRUE, col = "green", print.auc.y = .4, add = TRUE)
+plot(roc_non_cf, print.auc = TRUE, col = "red")
 ```
@@
 The blue curve shows the ROC for distinguishing between CFTR-RD or CRMS vs.
-CF, and the green is the ROC curve for distinguishing between non-CF vs.
+CF, and the red is the ROC curve for distinguishing between non-CF vs.
 CFTR-RD or CRMS.
@@
 
+##### Cl Alone
+
+```{r}
+roc_cf_cl <- roc(
+  response = t1_df$`Diagnosis`,
+  predictor = t1_df$Cl,
+  levels = c("CFTR-RD or CRMS", "Cystic Fibrosis"),
+  quiet = T
+)
+coords_cf <- round(coords(roc_cf_cl, "best", best.method = "youden"), 3)
+
+roc_non_cf_cl <- roc(
+  response = t1_df$`Diagnosis`,
+  predictor = t1_df$Cl,
+  levels = c("Non-CF", "CFTR-RD or CRMS"),
+  quiet = T
+)
+coords_non_cf <- round(coords(roc_non_cf_cl, "best", best.method = "youden"), 3)
+
+plot(roc_cf_cl, print.auc = TRUE, col = "blue")
+plot(roc_non_cf_cl, print.auc = TRUE, col = "red")
+```
+
+The blue curve shows the ROC for distinguishing between CFTR-RD or CRMS vs.
+CF, and the red is the ROC curve for distinguishing between non-CF vs.
+CFTR-RD or CRMS.
+The optimal threshold for distinguishing between CFTR-RD or CRMS vs.
+CF was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index.
+The optimal threshold for distinguishing between non-CF vs.
+CFTR-RD or CRMS was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index.
+The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.
+
+##### ROC Comparison
+
+###### CFTR-RD or CRMS vs. CF
+
+```{r}
+roc.test(roc_cf,roc_cf_cl)
+```
+
+###### Non-CF vs. CFTR-RD or CRMS
+
+```{r}
+roc.test(roc_non_cf,roc_non_cf_cl)
+```
+
 #### Histogram by group
 
-Vertical blue line indicates the threshold for CFTR-RD or CRMS vs. CF and green indicates the threshold for non-CF vs. CFTR-RD or CRMS. Na/Cl ratio was log-transformed for easier visualization.
+Vertical blue line indicates the threshold for CFTR-RD or CRMS vs. CF and red indicates the threshold for non-CF vs. CFTR-RD or CRMS. Na/Cl ratio was log-transformed for easier visualization.
 
@@
   geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
-  geom_vline(xintercept = log(coords_non_cf$threshold), color = "green") +
+  geom_vline(xintercept = log(coords_non_cf$threshold), color = "red") +
   ylab("Count") +
@@
 plot(roc_cf, print.auc = TRUE, col = "blue")
-plot(roc_non_cf, print.auc = TRUE, col = "green", print.auc.y = .4, add = TRUE)
+plot(roc_non_cf, print.auc = TRUE, col = "red", print.auc.y = .4, add = TRUE)
 ```
@@
 The blue curve shows the ROC for distinguishing between low vs.
-high risk, and the green is the ROC curve for distinguishing between non-CF vs.
+high risk, and the red is the ROC curve for distinguishing between non-CF vs.
 low risk.
@@
 
-Vertical blue line indicates the threshold for low vs. high risk and green indicates the threshold for non-CF vs. low risk. Na/Cl ratio was log-transformed for easier visualization.
+Vertical blue line indicates the threshold for low vs. high risk and red indicates the threshold for non-CF vs. low risk. Na/Cl ratio was log-transformed for easier visualization.
 
@@
 
-Vertical blue line indicates the threshold for CFTR-RD or CRMS vs. CF and green indicates the threshold for non-CF vs. CFTR-RD or CRMS. Na/Cl ratio was log-transformed for easier visualization.
+Vertical blue line indicates the threshold for CFTR-RD or CRMS vs. CF and red indicates the threshold for non-CF vs. CFTR-RD or CRMS. Na/Cl ratio was log-transformed for easier visualization.
 

### 2024-03-06 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
Update phase_1_analysis.qmd

@@
       dark: darkly
-bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
+bibliography: /Volumes/Documents/Miscellaneous/zotero.bib
+csl: /Volumes/Documents/Miscellaneous/american-medical-association.csl
 editor: source
@@
 exportBulkRecords(list(db = rcon), envir = 1)
+# Exclude
+exclude <- c("01-030_W", "01-038_W", "02-0038", "02-0039", "04-001")
 # Combine collection results
@@
     staphylococcus_aureus_sos:other_result_sos
+  ) %>%
+  filter(!record_id %in% exclude)
+# Remove empty demographics rows
+db_demographics <- db_demographics %>%
+  filter(
+    redcap_event_name == "Clinic Visit (Arm 1: Eligible)",
+    !record_id %in% exclude
   )
+# Excluded participants
+db_medications <- db_medications %>% filter(!record_id %in% exclude)
+# Combine race and ethnicity columns
+races <- c(
+  "American Indian or Alaskan Native", "Asian", "Hawaiian or Pacific Islander",
+  "Black or African American", "White", "Unknown", "Other"
+)
+ethnicities <- c("Hispanic", "Non-Hispanic", "Unknown/Not Reported")
+db_demographics$Race <- apply(db_demographics, 1, function(r) {
+  w <- which(r[grep("race", names(r))] == "Checked")
+  if (length(w) == 1) {
+    return(races[w])
+  } else if (length(w) > 1) {
+    return("Multiple Races")
+  } else if (length(w) == 0) {
+    return(NA)
+  }
+})
+db_demographics$Ethnicity <- apply(db_demographics, 1, function(r) {
+  w <- which(r[grep("ethnicity", names(r))] == "Checked")
+  if (length(w) == 1) {
+    return(ethnicities[w])
+  } else if (length(w) > 1) {
+    return(paste(ethnicities[w], sep = "/"))
+  } else if (length(w) == 0) {
+    return(NA)
+  }
+})
+# # Drop levels
+# level_drop <- c("sex", "Race", "Ethnicity")
+# db_demographics[, level_drop] <- lapply(
+#   db_demographics[, level_drop],
+#   function(c) {
+#     droplevels(factor(c))
+#   }
+# )
+# Add groups to other dataframes
+db_medications <- left_join(db_medications,
+  db_demographics %>% select(record_id, group),
+  by = join_by(record_id)
+)
+```
+
+# Data cleaning
+
+Participants 01-030_W, 01-038_W, 02-0038, 02-0039, and 04-001 were excluded from all analyses.
+Participants missing either the COS or SOS sample were excluded from agreement analyses, but retained in feasibility and other analyses.
+
+# Demographics
+
+```{r}
+db_demographics %>%
+  select(group, sex, Race, Ethnicity) %>%
+  tbl_summary(by = group, label = list("sex" = "Sex")) %>%
+  add_overall() %>%
+  add_p(test = list(c("Race", "Ethnicity") ~ "fisher.test")) %>%
+  separate_p_footnotes() %>%
+  as_gt()
+```
+
+# Medications
+
+```{r}
+db_medications %>%
+  select(group, azithromycin:please_specify_mucolytic) %>%
+  select(
+    -inhaled_colisitin_status, -inhaled_vancomycin_status,
+    -other_inhaled_status, -please_specify_mucolytic
+  ) %>%
+  tbl_summary(by = group) %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes() %>%
+  as_gt()
 ```
@@
   d <- data[, c(cos, sos)]
+  d <- d[complete.cases(d), ]
   a <- agree(d)
@@
 
-## By bug
+To calculate overall agreement, we determined whether a given COS or SOS sample was positive for *any* organism, and then calculated agreement metrics.
 
@@
 )
-bug_list = as.list(bug_names)
-names(bug_list) = bugs
+bug_list <- as.list(bug_names)
+names(bug_list) <- bugs
+# Create overall columns (any bug detected in sample)
+df$overall_cos <- apply(df, 1, function(r) {
+  any(r[paste0(bugs, "_cos")] == "Yes")
+})
+df$overall_cos <- factor(df$overall_cos, levels = c(F, T), labels = c("Negative", "Positive"))
+
+df$overall_sos <- apply(df, 1, function(r) {
+  any(r[paste0(bugs, "_sos")] == "Yes")
+})
+df$overall_sos <- factor(df$overall_sos, levels = c(F, T), labels = c("Negative", "Positive"))
+# Agreement metrics
+d <- df[, c("overall_cos", "overall_sos")]
+d <- d[complete.cases(d), ]
+a <- agree(d)
+k <- kappa2(d)
+ct <- table(d[, "overall_cos"])
+st <- table(d[, "overall_sos"])
+kable(data.frame(
+  "COS Negative" = ct["Negative"], "COS Positive" = ct["Positive"],
+  "SOS Negative" = st["Negative"], "SOS Positive" = st["Positive"],
+  "Percent Agreement" = a$value, "Cohen's Kappa" = k$value,
+  "p value" = k$p.value, check.names = F
+), digits = 2)
+t <- data.frame(table(d))
+colnames(t) <- c("COS", "SOS", "n")
+kable(t)
+```
+
+## By bug
+
+```{r}
 l <- lapply(bugs, function(b) {
@@
 ```{r results='asis'}
-invisible(lapply(names(bug_list),function(b){
+invisible(lapply(names(bug_list), function(b) {
   cat("\n")
-  cat("##",as.character(bug_list[b]))
+  cat("##", as.character(bug_list[b]))
   cat("\n")
@@
   d <- df[, c(cos, sos)]
-  t = data.frame(table(d))
-  colnames(t) = c("COS","SOS","n")
+  d <- d[complete.cases(d), ]
+  t <- data.frame(table(d))
+  colnames(t) <- c("COS", "SOS", "n")
   cat("\n")

### 2024-03-07 Christine Chan/CF Registry/cf_registry_analysis.qmd
windows paths

@@
 library(forestplot)
-knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry")
+switch(Sys.info()[["sysname"]],
+  Windows = {
+    home_dir <- "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry"
+  },
+  Darwin = {
+    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry"
+  }
+)
+knitr::opts_knit$set(root.dir = home_dir)
 ```
@@
     add_n(location = "level") %>%
-    add_nevent(location = "level")%>% 
+    add_nevent(location = "level") %>%
     as_tibble()

### 2024-03-07 Quarto Template.qmd

--- 2024-01-18 Quarto Template.qmd
@@
 library(tidyverse)
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Dropbox/Work/Vigers/CF",
-  "Windows" = "C:/Users/timvigers/Dropbox/Work/Vigers/CF"
+switch(Sys.info()[["sysname"]],
+  Windows = {
+    home_dir <- "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF"
+  },
+  Darwin = {
+    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
+  }
 )

### 2024-03-08 Christine Chan/CF Registry/cf_registry_analysis.qmd
forest plots

@@
 ```{r}
-tv_forest_plot <- function(mod, time_varying = NULL, time_variable = "tgroup",
-                           labels = labels) {
+tv_forest_plot <- function(mod_coeff, time_variable = NULL, labels = labels) {
   coeff <- tbl_regression(mod, pvalue_fun = function(x) x) %>%
@@
   # If time-varying, format appropriately
-  if (!is.null(time_varying)) {
+  if (length(grep("strata", coeff$`**Characteristic**`)) > 0) {
     # Split table into time-varying and not
@@
 invisible(lapply(time_varying, function(v) {
+  # Formulas
   new_f <- as.formula(paste0(
@@
   ))
-  cox_res_split <- coxph(new_f, data = s_split, cluster = eDWID)
+  new_f_int <- as.formula(paste0(
+    y, "~", paste0(x[-which(x == v)], collapse = "+"),
+    paste0("+", v, "*tgroup")
+  ))
+  new_f_emm <- as.formula(paste0(
+    "~", paste0(v, "|tgroup")
+  ))
+  # This could be done more efficiently but it seems to work
+  # Fit one stratified model and one with a regular interaction term
+  cox_res_split <- coxph(new_f, data = s_split, cluster = eDWID, contrasts = "contr.sum")
+  cox_res_split_int <- coxph(new_f_int, data = s_split, cluster = eDWID)
+  # Use the regular interaction term to get means for each time period
   cat("\n")
@@
   cat("\n")
-  plot_df <- tbl_regression(cox_res_split) %>%
+  cox_res_split %>%
+    tbl_regression() %>%
     add_n(location = "level") %>%
-    add_nevent(location = "level") %>%
-    as_tibble()
-  plot_df$ll <- as.numeric(sapply(strsplit(plot_df$`**95% CI**`, ", "), "[", 1))
-  plot_df$ul <- as.numeric(sapply(strsplit(plot_df$`**95% CI**`, ", "), "[", 2))
-  plot_df$hr <- as.numeric(plot_df$`**log(HR)**`)
-  plot_df %>%
-    forestplot(
-      labeltext = c(`**Characteristic**`, `**N**`, `**Event N**`, `**log(HR)**`),
-      mean = "hr", lower = "ll", upper = "ul",
-      is.summary = is.na(plot_df$`**N**`)
-    ) %>%
-    fp_add_header(
-      `**Characteristic**` = c("", "Characteristic"),
-      `**N**` = c("", "N"),
-      `**Event N**` = c("", "Event N"),
-      `**log(HR)**` = c("", "log(HR)")
-    ) %>%
-    fp_add_lines() %>%
-    fp_set_zebra_style("#EFEFEF", ignore_subheaders = T) %>%
-    print()
+    add_nevent(location = "level")
+  cat("\n")
+  cat("\n")
+  cox_res_split %>%
+    tbl_regression() %>%
+    plot()
+  cat("\n")
+  cat("\n")
+  emm <- data.frame(emmeans(cox_res_split_int, new_f_emm, nuisance = x[-which(x == v)]))
+  plot_df <- data.frame(emm) %>% pivot_longer(cols = tgroup, value)
+
   cat("\n")

### 2024-03-08 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd

@@
       dark: darkly
-bibliography: /Volumes/Documents/Miscellaneous/zotero.bib
-csl: /Volumes/Documents/Miscellaneous/american-medical-association.csl
+
 editor: source
@@
 library(knitr)
+library(caret)
 home_dir <- switch(Sys.info()["sysname"],
@@
   cat("\n")
+  cat("\n")
+  c <- confusionMatrix(d[, sos], reference = d[, cos], mode = "everything")
+  cat("### 2x2 table")
+  print(kable(c$table))
+  cat("\n")
+  cat("\n")
+  cat("#### 2x2 table metrics")
+  print(kable(c$byClass))
+  cat("\n")
 }))

### 2024-03-08 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
cf registry separate ROC analysis

---
title: "CF Registry Analysis - ROC"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
csl: /Users/timvigers/Dropbox/Miscellaneous/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(gtsummary)
switch(Sys.info()[["sysname"]],
  Windows = {
    home_dir <- "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry"
  },
  Darwin = {
    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry"
  }
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
# Import data
encounter <- read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv",
  na.strings = c("", "N/A")
)
# Limit to those with two hour OGTT and Hba1c at the same encounter
roc_df <- encounter %>%
  select(eDWID, reviewyear, encounternum, twohour_bloodglucose, hgba1c) %>%
  drop_na() %>%
  arrange(eDWID, reviewyear, encounternum) %>%
  group_by(eDWID) %>%
  filter(row_number() == n()) %>%
  ungroup() %>%
  mutate(`OGTT 2-hour` = cut(twohour_bloodglucose,
    breaks = c(-Inf, 140, 200, Inf), labels = c("<140", "[140,200)", ">=200")
  ))
# Add demographic data
demo <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv",
  na.strings = c("", "N/A")
)
roc_df <- left_join(roc_df,
  demo %>% select(eDWID, Race1:Race6, Hispanicrace, Gender, MutClass),
  by = join_by(eDWID)
)
write.csv(roc_df, file = "~/ROC data.csv", row.names = F, na = "")
```

# Participant Characteristics

Race 1: White

Race 2: Black or African American

Race 3: American Indian or Alaska Native

Race 4: Asian

Race 5: Native Hawaiian or Other Pacific Islander

Race 6: Some other race

```{r}
roc_df %>%
  select(reviewyear, hgba1c, `OGTT 2-hour`, Race1:Race6, Hispanicrace, Gender, MutClass) %>%
  tbl_summary(by = `OGTT 2-hour`, missing_text = "Missing") %>%
  add_overall()
```

# ROC analysis

### 2024-03-08 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
cf registry separate ROC analysis

--- 2024-02-20 Christine Chan/CF Registry/cf_registry_analysis.qmd
@@
   warning: false
+  cache: true
 editor: source
@@
 library(broom)
-library(forestplot)
-knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry")
+library(emmeans)
+library(ggstats)
+source("/Users/timvigers/GitHub/shared-resources/Visualization/forest_plots.R")
+switch(Sys.info()[["sysname"]],
+  Windows = {
+    home_dir <- "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry"
+  },
+  Darwin = {
+    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry"
+  }
+)
+knitr::opts_knit$set(root.dir = home_dir)
 ```
@@
   inc <- df %>%
-    filter(reviewyear == year, is.na(first_cfrd_cffpr) | first_cfrd_cffpr >= year) %>%
+    filter(reviewyear == year, is.na(first_cfrd_cffpr) |
+      first_cfrd_cffpr >= year) %>%
     group_by(eDWID) %>%
@@
 
-  new_cases_10 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "< 10"])
+  new_cases_10 <-
+    sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "< 10"])
   at_risk_10 <- sum(inc$n[inc$`Age Group` == "< 10"])
@@
 
-  new_cases_10_18 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "10 - 18"])
+  new_cases_10_18 <- sum(inc$n[inc$cfrd_cffpr == "Yes" &
+    inc$`Age Group` == "10 - 18"])
   at_risk_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
@@
 
-  new_cases_19_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "19 - 30"])
+  new_cases_19_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" &
+    inc$`Age Group` == "19 - 30"])
   at_risk_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
@@
 
-  new_cases_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "> 30"])
+  new_cases_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" &
+    inc$`Age Group` == "> 30"])
   at_risk_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
   inc_30 <- new_cases_30 / (at_risk_30 / py)
-
   # Nice results
@@
   inc <- df %>%
-    filter(reviewyear == year, first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>%
-    mutate(ogtt = !is.na(ogttfast_bloodglucose) | !is.na(twohour_bloodglucose)) %>%
+    filter(reviewyear == year, first_cfrd_cffpr > year |
+      is.na(first_cfrd_cffpr)) %>%
+    mutate(ogtt = !is.na(ogttfast_bloodglucose) |
+      !is.na(twohour_bloodglucose)) %>%
     group_by(eDWID) %>%
@@
   inc <- df %>%
-    filter(reviewyear == year, first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>%
+    filter(reviewyear == year, first_cfrd_cffpr > year |
+      is.na(first_cfrd_cffpr)) %>%
     mutate(a1c = !is.na(hgba1c)) %>%
@@
 ```{r}
-tv_forest_plot <- function(mod, time_varying = NULL, time_variable = "tgroup",
-                           labels = labels) {
-  coeff <- tbl_regression(mod, pvalue_fun = function(x) x) %>%
-    add_n(location = "level") %>%
-    add_nevent(location = "level") %>%
-    as_tibble()
-  # If time-varying, format appropriately
-  if (!is.null(time_varying)) {
-    # Split table into time-varying and not
-    w <- which(coeff$`**Characteristic**` == labels[time_varying])
-    coeff <- coeff[w:nrow(coeff), ]
-    t <- coeff %>% pivot_longer(`**Characteristic**`)
-  }
-  coeff_n <- tbl_regression(mod, exponentiate = T) %>%
-    add_n(location = "label") %>%
-    as_tibble()
-  coeff$N2 <- coeff_n$`**N**`
-  # Combine Ns
-  coeff$N <- pmax(coeff$`**N**`, coeff$N2, na.rm = T)
-
-  # Exponentiate
-  coeff$hr <- exp(as.numeric(coeff$`**log(HR)**`))
-  coeff$exp.ll <- exp(as.numeric(sapply(str_split(coeff$`**95% CI**`, ", "), "[", 1)))
-  coeff$exp.ul <- exp(as.numeric(sapply(str_split(coeff$`**95% CI**`, ", "), "[", 2)))
-  # Add N to names
-  coeff$names <- paste0(coeff$`**Characteristic**`, " (N = ", coeff$N, ")")
-  coeff$names <- ordered(coeff$names, levels = unique(coeff$names))
-  # P values
-  coeff$sig <- coeff$`**p-value**` < 0.05
-  coeff$p <- style_pvalue(coeff$`**p-value**`)
-  # Row types
-  coeff$type <- "level"
-  coeff$type[!is.na(coeff$N2)] <- "variable"
-  coeff$type[!is.na(coeff$`**N**`) & is.na(coeff$hr)] <- "reference"
-  # HR label
-  coeff$hr_lab <- paste0(
-    round(coeff$hr, 2), " (", round(coeff$exp.ll, 2), " - ",
-    round(coeff$exp.ul, 2), ")"
-  )
-  coeff$hr_lab[coeff$hr_lab == "NA (NA - NA)"] <- NA
-  coeff$hr_lab[coeff$type == "reference"] <- "Reference"
-  #
-}
-```
-
-```{r}
 # List of covariates
@@
 # Formula for base model
-x <- c(
+xvars <- c(
   "age_group", "Gender", "Race", "Hispanicrace", "bmi_perc_cat", "MutClass",
@@
 y <- "Surv(tstart, tstop, cfrd_status_annual)"
-f <- as.formula(paste0(y, "~", paste0(x, collapse = "+")))
+f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
 cox_res <- coxph(f, data = s, cluster = eDWID)
@@
 # Forest plot and coefficients table
-plot_df <- tbl_regression(cox_res, exponentiate = T) %>%
+cox_res %>%
+  tbl_regression() %>%
   add_n(location = "level") %>%
   add_nevent(location = "level")
-plot_df <- plot_df %>% as_tibble()
-plot_df$expll <- as.numeric(sapply(strsplit(plot_df$`**95% CI**`, ", "), "[", 1))
-plot_df$expul <- as.numeric(sapply(strsplit(plot_df$`**95% CI**`, ", "), "[", 2))
-plot_df$hr <- as.numeric(plot_df$`**HR**`)
-plot_df %>%
-  forestplot(
-    labeltext = c(`**Characteristic**`, `**N**`, `**Event N**`, `**HR**`),
-    mean = "hr", lower = "expll", upper = "expul", xlog = T,
-    is.summary = is.na(plot_df$`**N**`)
-  ) %>%
-  fp_add_header(
-    `**Characteristic**` = c("", "Characteristic"),
-    `**N**` = c("", "N"),
-    `**Event N**` = c("", "Event N"),
-    `**HR**` = c("", "HR")
-  ) %>%
-  fp_add_lines() %>%
-  fp_set_zebra_style("#EFEFEF", ignore_subheaders = T)
+ggcoef_model(cox_res,
+  show_p_values = F, add_reference_rows = F,
+  signif_stars = F, colour = F,
+)
 # Diagnostics
@@
 
-- `tgroup=1` indicates the years 2003 - 2008
-- `tgroup=2` indicates the years 2009 - 2013
-- `tgroup=3` indicates the years 2014 - 2018
-
-The interpretation of these models can be a little bit tricky, so here is an example using BMI percentile group and not including the other covariates in the model:
-
 ```{r}
@@
 )
-cox_res_split <- coxph(
-  Surv(tstart, tstop, cfrd_status_annual) ~ bmi_perc_cat * strata(tgroup),
-  data = s_split, cluster = eDWID
-)
-tbl_regression(cox_res_split) %>%
-  add_n(location = "level") %>%
-  add_nevent(location = "level")
 ```
 
-In the table above, the BMI reference group is the (50,85] group.
-So, in the years 2003 - 2008, the hazard ratio of the (0,10] BMI group was $exp(0.806)=2.238$ compared to the reference.
-Then, in the period of 2009 - 2013, the hazard ratio of the (0,10] BMI group was $exp(0.806+0.050)=2.354$ compared to the reference group, although this was not statistically significant.
-The HR for the (85,Inf] BMI group appears to increase over time, starting out at $exp(-0.511)=0.600$ in the first time period, then a statistically significant increase in the second time period to $exp(-0.511+0.282)=0.795$, with a final increase to $exp(-0.511+0.435)=0.927$ in the third period.
-The interpretation of the following results will follow these same formulas, and the time-varying results are shown at the bottom of each table.
-This report only includes variables that were statistically significant in the Cox proportional hazards test.
+In the plots below, time-varying coefficients are excluded from the first forest plot, to show the average effect of the predictors.
+In the second forest plot, the estimated marginal means (EMMs) for the time-varying coefficients are shown.
+The EMMs are averaged over the levels of all the other variables in the model.
+In other words, the EMM represents the baseline hazard (not hazard ratio) at the average level of all the other variables in the model.
+For example, the log hazard for an average participant age [30, Inf) in the time period [2003,2009) is 1.887 and the log hazard for a similar participant in [2014,2019) is 1.401.
 
@@
 invisible(lapply(time_varying, function(v) {
+  # Formulas
   new_f <- as.formula(paste0(
-    y, "~", paste0(x[-which(x == v)], collapse = "+"),
+    y, "~", paste0(xvars[-which(xvars == v)], collapse = "+"),
     paste0("+", v, "*strata(tgroup)")
   ))
+  new_f_int <- as.formula(paste0(
+    y, "~", paste0(xvars[-which(xvars == v)], collapse = "+"),
+    paste0("+", v, "*tgroup")
+  ))
+  new_f_emm <- as.formula(paste0(
+    "~", paste0(v, "|tgroup")
+  ))
+  # This could be done more efficiently but it seems to work
+  # Fit one stratified model and one with a regular interaction term
   cox_res_split <- coxph(new_f, data = s_split, cluster = eDWID)
+  cox_res_split_int <- coxph(new_f_int, data = s_split, cluster = eDWID)
+  # Use the regular interaction term to get means for each time period
   cat("\n")
@@
   cat("\n")
-  plot_df <- tbl_regression(cox_res_split) %>%
+  cox_res_split %>%
+    tbl_regression() %>%
     add_n(location = "level") %>%
-    add_nevent(location = "level")%>% 
-    as_tibble()
-  plot_df$ll <- as.numeric(sapply(strsplit(plot_df$`**95% CI**`, ", "), "[", 1))
-  plot_df$ul <- as.numeric(sapply(strsplit(plot_df$`**95% CI**`, ", "), "[", 2))
-  plot_df$hr <- as.numeric(plot_df$`**log(HR)**`)
-  plot_df %>%
-    forestplot(
-      labeltext = c(`**Characteristic**`, `**N**`, `**Event N**`, `**log(HR)**`),
-      mean = "hr", lower = "ll", upper = "ul", 
-      is.summary = is.na(plot_df$`**N**`)
+    add_nevent(location = "level") %>%
+    print()
+  cat("\n")
+  cat("\n")
+  emm <- emmeans(cox_res_split_int, new_f_emm,
+    nuisance = xvars[-which(xvars == v)]
+  )
+  emm <- data.frame(emm)
+  # Plot base model (without time varying)
+  ggcoef_model(cox_res_split,
+    show_p_values = F, add_reference_rows = F, signif_stars = F, colour = F,
+    include = xvars[-which(xvars == v)]
   ) %>%
-    fp_add_header(
-      `**Characteristic**` = c("", "Characteristic"),
-      `**N**` = c("", "N"),
-      `**Event N**` = c("", "Event N"),
-      `**log(HR)**` = c("", "log(HR)")
-    ) %>%
-    fp_add_lines() %>%
-    fp_set_zebra_style("#EFEFEF", ignore_subheaders = T) %>%
+    print()
+  cat("\n")
+  cat("\n")
+  # Print EMMs
+  print(kable(emm, digits = 3))
+  cat("\n")
+  cat("\n")
+  # Plot marginal means
+  time_varying_forest_plot(emm, x = "emmean", var = v, var_label = labels[v]) %>%
     print()

### 2024-03-11 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
first attempt at registry ROC

@@
     fig-cap-location: top
+    fig-width: 8
+    fig-height: 8
     page-layout: full
-    theme:
-      light: flatly
-      dark: darkly
-bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Dropbox/Miscellaneous/american-medical-association.csl
+bibliography: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Miscellaneous/zotero.bib
+csl: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Miscellaneous/american-medical-association.csl
 editor: source
@@
 library(gtsummary)
+library(gt)
+library(pROC)
 switch(Sys.info()[["sysname"]],
@@
 )
-# Limit to those with two hour OGTT and Hba1c at the same encounter
+# Limit to those with two hour OGTT and HbA1c at the same encounter
 roc_df <- encounter %>%
@@
   ))
-# Add demographic data
+# Demographic data
 demo <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv",
@@
 )
+# Race
+races <- c(
+  "White", "Black or African American", "American Indian or Alaska Native",
+  "Asian", "Native Hawaiian or Other Pacific Islander", "Some other race"
+)
+demo$Race <- apply(demo, 1, function(r) {
+  w <- which(r[paste0("Race", 1:6)] == 1)
+  if (length(w) == 0) {
+    "Unknown"
+  } else if (length(w) > 1) {
+    "More than one race"
+  } else {
+    races[w]
+  }
+})
+demo$Race <- factor(demo$Race,
+  levels = c(
+    "White", "Black or African American", "American Indian or Alaska Native",
+    "Asian", "More than one race", "Native Hawaiian or Other Pacific Islander",
+    "Some other race"
+  ),
+  labels = c(
+    "White", "Black or African American", "Other race", "Other race",
+    "More than one race", "Other race", "Other race"
+  )
+)
+demo$Hispanicrace <- factor(demo$Hispanicrace,
+  levels = 1:2,
+  labels = c("Yes", "No")
+)
+# F508
+demo$F508 <- factor(demo$F508,
+  levels = 1:3,
+  labels = c("F508del/F508del", "F508del/Other", "Other/Other")
+)
+# Age at year end
+annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
+  na.strings = c("", "N/A")
+)
+annual <- annual %>% rename(reviewyear = ReviewYear)
+# Merge
 roc_df <- left_join(roc_df,
-  demo %>% select(eDWID, Race1:Race6, Hispanicrace, Gender, MutClass),
+  demo %>% select(eDWID, Race, Hispanicrace, Gender, F508, MutClass),
   by = join_by(eDWID)
 )
-write.csv(roc_df, file = "~/ROC data.csv", row.names = F, na = "")
+roc_df <- left_join(roc_df,
+  annual %>% select(eDWID, reviewyear, Age_YrEnd),
+  by = join_by(eDWID, reviewyear)
+)
 ```
@@
 
-Race 1: White
-
-Race 2: Black or African American
-
-Race 3: American Indian or Alaska Native
-
-Race 4: Asian
-
-Race 5: Native Hawaiian or Other Pacific Islander
-
-Race 6: Some other race
+## By two hour OGTT value
 
@@
 roc_df %>%
-  select(reviewyear, hgba1c, `OGTT 2-hour`, Race1:Race6, Hispanicrace, Gender, MutClass) %>%
-  tbl_summary(by = `OGTT 2-hour`, missing_text = "Missing") %>%
-  add_overall()
+  select(
+    reviewyear, Age_YrEnd, hgba1c, `OGTT 2-hour`, Race, Hispanicrace, Gender,
+    F508, MutClass
+  ) %>%
+  tbl_summary(
+    by = `OGTT 2-hour`, missing_text = "Missing",
+    type = list(reviewyear ~ "categorical"),
+    label = list(
+      hgba1c = "HbA1c", Hispanicrace = "Hispanic", MutClass = "Mutation Class",
+      reviewyear = "Year", Age_YrEnd = "Age at Year End"
+    )
+  ) %>%
+  add_overall() %>%
+  add_p(
+    test = list(reviewyear = "fisher.test"),
+    test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
+  ) %>%
+  separate_p_footnotes() %>%
+  as_gt()
+```
+
+## By age
+
+```{r}
+roc_df %>%
+  select(
+    reviewyear, Age_YrEnd, hgba1c, `OGTT 2-hour`, Race, Hispanicrace, Gender,
+    F508, MutClass
+  ) %>%
+  tbl_summary(
+    by = Age_YrEnd, missing_text = "Missing",
+    type = list(reviewyear ~ "categorical"),
+    label = list(
+      hgba1c = "HbA1c", Hispanicrace = "Hispanic", MutClass = "Mutation Class",
+      reviewyear = "Year", Age_YrEnd = "Age at Year End"
+    )
+  ) %>%
+  add_overall() %>%
+  add_p(
+    test = list(reviewyear = "fisher.test"),
+    test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
+  ) %>%
+  separate_p_footnotes() %>%
+  as_gt()
+```
+
+# HbA1c by diagnosis
+
+```{r}
+ggplot(roc_df, aes(x = `OGTT 2-hour`, y = hgba1c)) +
+  geom_boxplot() +
+  ylab("HbA1c") +
+  theme_classic()
 ```
@@
 # ROC analysis
+
+## Two hour OGTT value
+
+### Full cohort
+
+#### $\geq$ 200 vs. $<$ 200
+
+```{r}
+#| message: false
+roc_df$CFRD <- as.numeric(roc_df$twohour_bloodglucose >= 200)
+cfrd_roc <- roc(CFRD ~ hgba1c, data = roc_df)
+thresh <- coords(cfrd_roc, "best", best.method = "youden")
+plot(cfrd_roc, print.auc = T, print.thres = T)
+gt(coords(cfrd_roc)) %>% fmt_number(decimals = 3)
+```
+
+#### $\geq$ 140 vs. $<$ 140
+
+```{r}
+#| message: false
+roc_df$AGT <- as.numeric(roc_df$twohour_bloodglucose >= 140)
+agt_roc <- roc(AGT ~ hgba1c, data = roc_df)
+thresh <- coords(agt_roc, "best", best.method = "youden")
+plot(agt_roc, print.auc = T, print.thres = T)
+gt(coords(agt_roc)) %>% fmt_number(decimals = 3)
+```
+
+### Peds
+
+```{r}
+peds_df <- roc_df %>% filter(Age_YrEnd < 18)
+```
+
+#### $\geq$ 200 vs. $<$ 200
+
+```{r}
+#| message: false
+peds_df$CFRD <- as.numeric(peds_df$twohour_bloodglucose >= 200)
+cfrd_roc <- roc(CFRD ~ hgba1c, data = peds_df)
+thresh <- coords(cfrd_roc, "best", best.method = "youden")
+plot(cfrd_roc, print.auc = T, print.thres = T)
+gt(coords(cfrd_roc)) %>% fmt_number(decimals = 3)
+```
+
+#### $\geq$ 140 vs. $<$ 140
+
+```{r}
+#| message: false
+peds_df$AGT <- as.numeric(peds_df$twohour_bloodglucose >= 140)
+agt_roc <- roc(AGT ~ hgba1c, data = peds_df)
+thresh <- coords(agt_roc, "best", best.method = "youden")
+plot(agt_roc, print.auc = T, print.thres = T)
+gt(coords(agt_roc)) %>% fmt_number(decimals = 3)
+```
+
+### Adults
+
+```{r}
+adult_df <- roc_df %>% filter(Age_YrEnd >= 18)
+```
+
+#### $\geq$ 200 vs. $<$ 200
+
+```{r}
+#| message: false
+adult_df$CFRD <- as.numeric(adult_df$twohour_bloodglucose >= 200)
+cfrd_roc <- roc(CFRD ~ hgba1c, data = adult_df)
+thresh <- coords(cfrd_roc, "best", best.method = "youden")
+plot(cfrd_roc, print.auc = T, print.thres = T)
+gt(coords(cfrd_roc)) %>% fmt_number(decimals = 3)
+```
+
+#### $\geq$ 140 vs. $<$ 140
+
+```{r}
+#| message: false
+adult_df$AGT <- as.numeric(adult_df$twohour_bloodglucose >= 140)
+agt_roc <- roc(AGT ~ hgba1c, data = adult_df)
+thresh <- coords(agt_roc, "best", best.method = "youden")
+plot(agt_roc, print.auc = T, print.thres = T)
+gt(coords(agt_roc)) %>% fmt_number(decimals = 3)
+```

### 2024-03-11 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
CFRD registry rough draft

@@
 
-## Two hour OGTT value
+For the following analyses, we considered those with a two hour glucose value >= 140 and < 200 as a distinct group (AGT) and use receiver operating characteristic curves to find cutoffs that predict:
 
-### Full cohort
+1. NGT vs. AGT
 
-#### $\geq$ 200 vs. $<$ 200
+2. AGT vs. CFRD
+
+```{r}
+ngt_agt <- roc_df %>% filter(`OGTT 2-hour` != ">=200")
+ngt_agt$ogtt <- droplevels(ngt_agt$`OGTT 2-hour`)
+agt_cfrd <- roc_df %>% filter(`OGTT 2-hour` != "<140")
+agt_cfrd$ogtt <- droplevels(agt_cfrd$`OGTT 2-hour`)
+```
+
+## Peds and adults
+
+### NGT vs. AGT
 
@@
 #| message: false
-roc_df$CFRD <- as.numeric(roc_df$twohour_bloodglucose >= 200)
-cfrd_roc <- roc(CFRD ~ hgba1c, data = roc_df)
-thresh <- coords(cfrd_roc, "best", best.method = "youden")
-plot(cfrd_roc, print.auc = T, print.thres = T)
-gt(coords(cfrd_roc)) %>% fmt_number(decimals = 3)
+r <- roc(ogtt ~ hgba1c, data = ngt_agt)
+thresh <- coords(r, "best", best.method = "youden")
+plot(r, print.auc = T, print.thres = T)
+gt(coords(r)) %>% fmt_number(decimals = 3)
+summary(r)
 ```
 
-#### $\geq$ 140 vs. $<$ 140
+#### AGT vs. CFRD
 
@@
 #| message: false
-roc_df$AGT <- as.numeric(roc_df$twohour_bloodglucose >= 140)
-agt_roc <- roc(AGT ~ hgba1c, data = roc_df)
-thresh <- coords(agt_roc, "best", best.method = "youden")
-plot(agt_roc, print.auc = T, print.thres = T)
-gt(coords(agt_roc)) %>% fmt_number(decimals = 3)
+r <- roc(ogtt ~ hgba1c, data = agt_cfrd)
+thresh <- coords(r, "best", best.method = "youden")
+plot(r, print.auc = T, print.thres = T)
+gt(coords(r)) %>% fmt_number(decimals = 3)
+summary(r)
 ```
 
-### Peds
+## Peds
 
 ```{r}
-peds_df <- roc_df %>% filter(Age_YrEnd < 18)
+peds_ngt_agt <- ngt_agt %>% filter(Age_YrEnd < 18)
+peds_agt_cfrd <- agt_cfrd %>% filter(Age_YrEnd < 18)
 ```
 
-#### $\geq$ 200 vs. $<$ 200
+### NGT vs. AGT
 
@@
 #| message: false
-peds_df$CFRD <- as.numeric(peds_df$twohour_bloodglucose >= 200)
-cfrd_roc <- roc(CFRD ~ hgba1c, data = peds_df)
-thresh <- coords(cfrd_roc, "best", best.method = "youden")
-plot(cfrd_roc, print.auc = T, print.thres = T)
-gt(coords(cfrd_roc)) %>% fmt_number(decimals = 3)
+r <- roc(ogtt ~ hgba1c, data = peds_ngt_agt)
+thresh <- coords(r, "best", best.method = "youden")
+plot(r, print.auc = T, print.thres = T)
+gt(coords(r)) %>% fmt_number(decimals = 3)
+summary(r)
 ```
 
-#### $\geq$ 140 vs. $<$ 140
+#### AGT vs. CFRD
 
@@
 #| message: false
-peds_df$AGT <- as.numeric(peds_df$twohour_bloodglucose >= 140)
-agt_roc <- roc(AGT ~ hgba1c, data = peds_df)
-thresh <- coords(agt_roc, "best", best.method = "youden")
-plot(agt_roc, print.auc = T, print.thres = T)
-gt(coords(agt_roc)) %>% fmt_number(decimals = 3)
+r <- roc(ogtt ~ hgba1c, data = peds_agt_cfrd)
+thresh <- coords(r, "best", best.method = "youden")
+plot(r, print.auc = T, print.thres = T)
+gt(coords(r)) %>% fmt_number(decimals = 3)
+summary(r)
 ```
@@
 ```{r}
-adult_df <- roc_df %>% filter(Age_YrEnd >= 18)
+adult_ngt_agt <- ngt_agt %>% filter(Age_YrEnd >= 18)
+adult_agt_cfrd <- agt_cfrd %>% filter(Age_YrEnd >= 18)
 ```
 
-#### $\geq$ 200 vs. $<$ 200
+### NGT vs. AGT
 
@@
 #| message: false
-adult_df$CFRD <- as.numeric(adult_df$twohour_bloodglucose >= 200)
-cfrd_roc <- roc(CFRD ~ hgba1c, data = adult_df)
-thresh <- coords(cfrd_roc, "best", best.method = "youden")
-plot(cfrd_roc, print.auc = T, print.thres = T)
-gt(coords(cfrd_roc)) %>% fmt_number(decimals = 3)
+r <- roc(ogtt ~ hgba1c, data = adult_ngt_agt)
+thresh <- coords(r, "best", best.method = "youden")
+plot(r, print.auc = T, print.thres = T)
+gt(coords(r)) %>% fmt_number(decimals = 3)
+summary(r)
 ```
 
-#### $\geq$ 140 vs. $<$ 140
+#### AGT vs. CFRD
 
@@
 #| message: false
-adult_df$AGT <- as.numeric(adult_df$twohour_bloodglucose >= 140)
-agt_roc <- roc(AGT ~ hgba1c, data = adult_df)
-thresh <- coords(agt_roc, "best", best.method = "youden")
-plot(agt_roc, print.auc = T, print.thres = T)
-gt(coords(agt_roc)) %>% fmt_number(decimals = 3)
+r <- roc(ogtt ~ hgba1c, data = adult_agt_cfrd)
+thresh <- coords(r, "best", best.method = "youden")
+plot(r, print.auc = T, print.thres = T)
+gt(coords(r)) %>% fmt_number(decimals = 3)
+summary(r)
 ```

### 2024-03-12 Jordana Hoppe/Home Collection/manuscript.qmd
home collection manuscript

@@
       dark: darkly
-bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Dropbox/Miscellaneous/american-medical-association.csl
+bibliography: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Miscellaneous/zotero.bib
+csl: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Miscellaneous/american-medical-association.csl
 editor: source
@@
 library(kableExtra)
+library(caret)
+library(likert)
 switch(Sys.info()[["sysname"]],
   Windows = {
-    home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
+    home_dir <- "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF"
   },
@@
   cat("\n")
-  clinic_vs_res <- xtabs(~ Clinic + Delayed, data = d)
-  clinic_vs_home <- xtabs(~ Clinic + Home, data = d)
-  home_vs_res <- xtabs(~ Home + Delayed, data = d)
+  d <- data.frame(lapply(d, factor, levels = c("Absent", "Present")))
+  cat("\n")
+  cat("#### Clinic vs. delayed (clinic as reference)")
   cat("\n")
   cat("\n")
-  cat("#### Clinic vs. delayed")
+  cat("##### Confusion matrix")
   cat("\n")
   cat("\n")
-  clinic_vs_res %>%
-    as.data.frame(.) %>%
-    kable(.) %>%
-    kable_styling("striped") %>%
-    print(.)
-  cat("\n")
-  clinic_vs_res %>%
-    prop.table(.) %>%
-    as.data.frame(.) %>%
-    rename("Percentage" = Freq) %>%
-    kable(., digits = 2) %>%
-    kable_styling("striped") %>%
-    print(.)
+  c <- confusionMatrix((d[, "Delayed"]), reference = d[, "Clinic"], mode = "everything")
+  print(kable(c$table, digits = 3))
   cat("\n")
   cat("\n")
-  cat("#### Clinic vs. home")
-  cat("\n")
-  clinic_vs_home %>%
-    as.data.frame(.) %>%
-    kable(.) %>%
-    kable_styling("striped") %>%
-    print(.)
-  cat("\n")
-  clinic_vs_home %>%
-    prop.table(.) %>%
-    as.data.frame(.) %>%
-    rename("Percentage" = Freq) %>%
-    kable(., digits = 2) %>%
-    kable_styling("striped") %>%
-    print(.)
+  cat("##### Agreement")
   cat("\n")
   cat("\n")
-  cat("#### Home vs. delayed")
+  print(kable(c$overall, digits = 3, col.names = NULL))
   cat("\n")
-  home_vs_res %>%
-    as.data.frame(.) %>%
-    kable(.) %>%
-    kable_styling("striped") %>%
-    print(.)
   cat("\n")
-  home_vs_res %>%
-    prop.table(.) %>%
-    as.data.frame(.) %>%
-    rename("Percentage" = Freq) %>%
-    kable(., digits = 2) %>%
-    kable_styling("striped") %>%
-    print(.)
+  cat("##### Sensitivity, specificity, etc.")
+  cat("\n")
+  cat("\n")
+  print(kable(c$byClass, digits = 3, col.names = NULL))
+  cat("\n")
+  cat("\n")
+  cat("#### Clinic vs. home (clinic as reference)")
+  cat("\n")
+  cat("\n")
+  cat("##### Confusion matrix")
+  cat("\n")
+  cat("\n")
+  c <- confusionMatrix((d[, "Home"]), reference = d[, "Clinic"], mode = "everything")
+  print(kable(c$table, digits = 3))
+  cat("\n")
+  cat("\n")
+  cat("##### Agreement")
+  cat("\n")
+  cat("\n")
+  print(kable(c$overall, digits = 3, col.names = NULL))
+  cat("\n")
+  cat("\n")
+  cat("##### Sensitivity, specificity, etc.")
+  cat("\n")
+  cat("\n")
+  print(kable(c$byClass, digits = 3, col.names = NULL))
+  cat("\n")
+  cat("\n")
+  cat("#### Home vs. delayed (delayed as reference)")
+  cat("\n")
+  cat("\n")
+  cat("##### Confusion matrix")
+  cat("\n")
+  cat("\n")
+  c <- confusionMatrix((d[, "Home"]), reference = d[, "Delayed"], mode = "everything")
+  print(kable(c$table, digits = 3))
+  cat("\n")
+  cat("\n")
+  cat("##### Agreement")
+  cat("\n")
+  cat("\n")
+  print(kable(c$overall, digits = 3, col.names = NULL))
+  cat("\n")
+  cat("\n")
+  cat("##### Sensitivity, specificity, etc.")
+  cat("\n")
+  cat("\n")
+  print(kable(c$byClass, digits = 3, col.names = NULL))
   cat("\n")
@@
 ```
+
+# Survey results
+
+```{r}
+# List columns by number of possible answers
+levels3 <- "q2.factor"
+levels5 <- c("difficulty.factor", "shipping.factor")
+checkboxes <- c(
+  paste0("barriers___", 1:4, ".factor"),
+  paste0("q6a___", 1:5, ".factor"),
+  paste0("q6b___", 1:4, ".factor")
+)
+yn <- c(
+  "q1.factor", "q3.factor", "previous.factor", "goodsample.factor", "q4.factor",
+  "q5.factor", "q6.factor"
+)
+# Subset and label
+t <- survey_data %>%
+  filter(!participant_id %in% survey_exclude) %>%
+  select(participant_id, q1.factor:q6b___4.factor)
+labels <- as.character(label(survey_data)[sub("\\.factor", "", colnames(t))])
+names(labels) <- colnames(t)
+```
+
+## Plots
+
+```{r}
+yn_df = t %>% select(all_of(yn)) %>% rename(any_of(setNames(names(labels), labels)))
+yn_df = data.frame(lapply(yn_df,factor,levels=c("No","Yes")),check.names = F)
+plot(likert(yn_df))
+```
+
+```{r}
+check_df = t %>% select(all_of(checkboxes)) %>% rename(any_of(setNames(names(labels), labels)))
+plot(likert(check_df))
+```
+
+```{r}
+df3 = t %>% select(all_of(levels3)) %>% rename(any_of(setNames(names(labels), labels)))
+plot(likert(df3))
+```
+
+```{r}
+df5 = t %>% select(all_of(levels5)) %>% rename(any_of(setNames(names(labels), labels)))
+plot(likert(df5))
+```

### 2024-03-13 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
remove ABPA from survival models

@@
     page-layout: full
-execute:
-  echo: false
-  warning: false
-  cache: true
 editor: source
@@
   "NumPulmExacerbation", "A_pseudomonasaeruginosa", "A_fungalyeast1",
-  "A_mycocultureresults3", "A_pulmonarycomplications1", "A_corticosteroids1",
-  "cfrd_status_annual", "A_Vx770", "A_VX809comb", "liver_disease"
+  "A_mycocultureresults3", "A_corticosteroids1", "cfrd_status_annual", 
+  "A_Vx770", "A_VX809comb", "liver_disease"
 )
@@
   A_fungalyeast1 = "Aspergillus (any species)",
-  A_pulmonarycomplications1 = "ABPA",
   A_mycocultureresults3 = "Microorganisms in myco culture",
@@
   "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
-  "NumPulmExacerbation", "A_pulmonarycomplications1", "A_corticosteroids1"
+  "NumPulmExacerbation",  "A_corticosteroids1"
 )

### 2024-03-13 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
ROC update for NACFC abstract

--- 2024-03-08 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
@@
 )
-# Limit to those with two hour OGTT and Hba1c at the same encounter
+# Demographic data
+demo <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv",
+  na.strings = c("", "N/A")
+)
+# Race
+races <- c(
+  "White", "Black or African American", "American Indian or Alaska Native",
+  "Asian", "Native Hawaiian or Other Pacific Islander", "Some other race"
+)
+demo$Race <- apply(demo, 1, function(r) {
+  w <- which(r[paste0("Race", 1:6)] == 1)
+  if (length(w) == 0) {
+    "Unknown"
+  } else if (length(w) > 1) {
+    "More than one race"
+  } else {
+    races[w]
+  }
+})
+demo$Race <- factor(demo$Race,
+  levels = c(
+    "White", "Black or African American", "American Indian or Alaska Native",
+    "Asian", "More than one race", "Native Hawaiian or Other Pacific Islander",
+    "Some other race"
+  ),
+  labels = c(
+    "White", "Black or African American", "Other race", "Other race",
+    "More than one race", "Other race", "Other race"
+  )
+)
+demo$Hispanicrace <- factor(demo$Hispanicrace,
+  levels = 1:2,
+  labels = c("Yes", "No")
+)
+# F508
+demo$F508 <- factor(demo$F508,
+  levels = 1:3,
+  labels = c("F508del/F508del", "F508del/Other", "Other/Other")
+)
+# Age at year end
+annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
+  na.strings = c("", "N/A")
+)
+annual <- annual %>% rename(reviewyear = ReviewYear)
+# Merge
+encounter <- left_join(
+  encounter,
+  demo %>% select(eDWID, Race, Hispanicrace, Gender, F508, MutClass),
+  by = join_by(eDWID)
+)
+encounter <- left_join(
+  encounter,
+  annual %>% select(eDWID, reviewyear, Age_YrEnd),
+  by = join_by(eDWID, reviewyear)
+)
+# Limit to those with two hour OGTT and HbA1c at the same encounter
+# Also remove years before 2010 and those under age 10
 roc_df <- encounter %>%
-  select(eDWID, reviewyear, encounternum, twohour_bloodglucose, hgba1c) %>%
+  arrange(eDWID, reviewyear, encounternum) %>%
+  select(
+    eDWID, reviewyear, Age_YrEnd, Race, Hispanicrace, Gender, F508, MutClass,
+    twohour_bloodglucose, hgba1c
+  ) %>%
+  filter(reviewyear >= 2010, Age_YrEnd >= 10, hgba1c >= 3.5 & hgba1c <= 15) %>%
   drop_na() %>%
-  arrange(eDWID, reviewyear, encounternum) %>%
   group_by(eDWID) %>%
@@
   ungroup() %>%
-  mutate(`OGTT 2-hour` = cut(twohour_bloodglucose,
+  mutate(
+    `OGTT 2-hour` = cut(twohour_bloodglucose,
       breaks = c(-Inf, 140, 200, Inf), labels = c("<140", "[140,200)", ">=200")
-  ))
-# Add demographic data
-demo <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv",
-  na.strings = c("", "N/A")
+    ),
+    `Age Group` = cut(Age_YrEnd, breaks = c(10, 18, 30, Inf), right = F)
   )
-roc_df <- left_join(roc_df,
-  demo %>% select(eDWID, Race1:Race6, Hispanicrace, Gender, MutClass),
-  by = join_by(eDWID)
+# Metrics to report
+metrics <- c(
+  "thresh", "specificity", "sensitivity", "accuracy", "tn", "tp", "fn", "fp",
+  "npv", "ppv"
 )
-write.csv(roc_df, file = "~/ROC data.csv", row.names = F, na = "")
 ```
+
+# Data cleaning
+
+1.
+People who reported American Indian or Alaska Native, Asian, and Native Hawaiian or Other Pacific Islander for race were considered as "Other race" due to relatively small sample sizes for these groups.
+
+2. Years prior to 2010 and those under age 10 at the end of a given year were excluded from the analysis.
+
+3. HbA1cs < 3.5% or > 15% were excluded from the analysis, as these values were assumed to be entered incorrectly.
 
@@
 roc_df %>%
-  select(reviewyear, hgba1c, `OGTT 2-hour`, Race1:Race6, Hispanicrace, Gender, MutClass) %>%
-  tbl_summary(by = `OGTT 2-hour`, missing_text = "Missing") %>%
-  add_overall()
+  select(
+    reviewyear, hgba1c, Age_YrEnd, `Age Group`, `OGTT 2-hour`, Race,
+    Hispanicrace, Gender, F508, MutClass
+  ) %>%
+  tbl_summary(
+    by = `OGTT 2-hour`, missing_text = "Missing",
+    type = list(reviewyear ~ "categorical"),
+    label = list(
+      hgba1c = "HbA1c", Hispanicrace = "Hispanic", MutClass = "Mutation Class",
+      reviewyear = "Year", Age_YrEnd = "Age at Year End"
+    )
+  ) %>%
+  add_overall() %>%
+  add_p(
+    test = list(reviewyear = "fisher.test"),
+    test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
+  ) %>%
+  separate_p_footnotes() %>%
+  as_gt()
+```
+
+## By age
+
+```{r}
+roc_df %>%
+  select(
+    reviewyear, hgba1c, Age_YrEnd, `Age Group`, `OGTT 2-hour`, Race,
+    Hispanicrace, Gender, F508, MutClass
+  ) %>%
+  tbl_summary(
+    by = `Age Group`, missing_text = "Missing",
+    type = list(reviewyear ~ "categorical"),
+    label = list(
+      hgba1c = "HbA1c", Hispanicrace = "Hispanic", MutClass = "Mutation Class",
+      reviewyear = "Year", Age_YrEnd = "Age at Year End"
+    )
+  ) %>%
+  add_overall() %>%
+  add_p(
+    test = list(reviewyear = "fisher.test"),
+    test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
+  ) %>%
+  separate_p_footnotes() %>%
+  as_gt()
+```
+
+# HbA1c by diagnosis
+
+```{r}
+ggplot(roc_df, aes(x = `OGTT 2-hour`, y = hgba1c)) +
+  geom_boxplot() +
+  ylab("HbA1c") +
+  theme_classic()
 ```
@@
 # ROC analysis
+
+## Three groups, two cutoffs
+
+For the following analyses, we considered those with a two hour glucose value >= 140 and < 200 as a distinct group (AGT) and use receiver operating characteristic curves to find cutoffs that predict:
+
+1. NGT vs. AGT
+
+2. AGT vs. CFRD
+
+```{r}
+ngt_agt <- roc_df %>% filter(`OGTT 2-hour` != ">=200")
+ngt_agt$ogtt <- droplevels(ngt_agt$`OGTT 2-hour`)
+agt_cfrd <- roc_df %>% filter(`OGTT 2-hour` != "<140")
+agt_cfrd$ogtt <- droplevels(agt_cfrd$`OGTT 2-hour`)
+```
+
+### Peds and adults
+
+#### NGT vs. AGT
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = ngt_agt)
+plot(r, print.auc = T, print.thres = T)
+gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
+```
+
+#### AGT vs. CFRD
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = agt_cfrd)
+plot(r, print.auc = T, print.thres = T)
+gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
+```
+
+### Peds
+
+```{r}
+peds_ngt_agt <- ngt_agt %>% filter(Age_YrEnd < 18)
+peds_agt_cfrd <- agt_cfrd %>% filter(Age_YrEnd < 18)
+```
+
+#### NGT vs. AGT
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = peds_ngt_agt)
+plot(r, print.auc = T, print.thres = T)
+gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
+```
+
+#### AGT vs. CFRD
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = peds_agt_cfrd)
+plot(r, print.auc = T, print.thres = T)
+gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
+```
+
+### Adults
+
+```{r}
+adult_ngt_agt <- ngt_agt %>% filter(Age_YrEnd >= 18)
+adult_agt_cfrd <- agt_cfrd %>% filter(Age_YrEnd >= 18)
+```
+
+#### NGT vs. AGT
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = adult_ngt_agt)
+plot(r, print.auc = T, print.thres = T)
+gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
+```
+
+#### AGT vs. CFRD
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = adult_agt_cfrd)
+plot(r, print.auc = T, print.thres = T)
+gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
+```
+
+## Two groups, one cutoffs
+
+For the following analyses, we examined HbA1c cutoffs that best predict a two hour OGTT value either >= 140 or >= 200
+
+```{r}
+roc_df$agt <- as.numeric(roc_df$twohour_bloodglucose >= 140)
+roc_df$cfrd <- as.numeric(roc_df$twohour_bloodglucose >= 200)
+```
+
+### Peds and adults
+
+#### $\geq$ 140
+
+```{r}
+#| message: false
+r <- roc(agt ~ hgba1c, data = roc_df)
+plot(r, print.auc = T, print.thres = T)
+gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
+```
+
+#### $\geq$ 200
+
+```{r}
+#| message: false
+r <- roc(cfrd ~ hgba1c, data = roc_df)
+plot(r, print.auc = T, print.thres = T)
+gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
+```
+
+### Peds
+
+```{r}
+peds_roc <- roc_df %>% filter(Age_YrEnd < 18)
+```
+
+#### $\geq$ 140
+
+```{r}
+#| message: false
+r <- roc(agt ~ hgba1c, data = peds_roc)
+plot(r, print.auc = T, print.thres = T)
+gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
+```
+
+#### $\geq$ 200
+
+```{r}
+#| message: false
+r <- roc(cfrd ~ hgba1c, data = peds_roc)
+plot(r, print.auc = T, print.thres = T)
+gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
+```
+
+### Adults
+
+```{r}
+adult_roc <- roc_df %>% filter(Age_YrEnd >= 18)
+```
+
+#### $\geq$ 140
+
+```{r}
+#| message: false
+r <- roc(agt ~ hgba1c, data = adult_roc)
+plot(r, print.auc = T, print.thres = T)
+gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
+```
+
+#### $\geq$ 200
+
+```{r}
+#| message: false
+r <- roc(cfrd ~ hgba1c, data = adult_roc)
+plot(r, print.auc = T, print.thres = T)
+gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
+```

### 2024-03-13 Scott Sagel/PEx in CF/analysis_for_yaron.qmd
updated plots for Yaron

@@
       dark: darkly
-bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
+bibliography: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Miscellaneous/zotero.bib
+csl: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Miscellaneous/american-medical-association.csl
 editor: source
@@
 library(gtsummary)
+library(ggpubr)
+library(rspiro)
 knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Scott Sagel/PEx in CF")
+# Colors for plots
+yellow <- "#f8b80a"
+red <- "#e30022"
+purple <- "#7b2670"
+blue <- "#022451"
 ```
@@
 ```{r data cleaning}
+#| warning: false
 # Import
@@
   levels = c(F, T), labels = c("No", "Yes")
+)
+# % predicted FEV1
+df$sex <- factor(df$sex, levels = c(2, 1), labels = c("Male", "Female"))
+df$fev1_baseline <- rowMeans(df[, paste0("fev1_baseline", 1:6)], na.rm = T)
+df <- df %>%
+  group_by(sid) %>%
+  fill(sex) %>%
+  ungroup()
+df$ppfev1_admit <- pctpred_GLIgl(
+  age = df$age_at_admit, height = df$height / 100, gender = as.numeric(df$sex),
+  FEV1 = df$fev1_admit
+)
+df$ppfev1_baseline <- pctpred_GLIgl(
+  age = df$age_at_admit, height = df$height / 100, gender = as.numeric(df$sex),
+  FEV1 = df$fev1_baseline
+)
+df$ppfev1_discharge <- pctpred_GLIgl(
+  age = df$age_at_admit, height = df$height / 100, gender = as.numeric(df$sex),
+  FEV1 = df$fev1_discharge
 )
@@
   "CFTR Modulator Type" = "cftr_mod_spec",
-  "Cultures Tested" = "redcap_data_access_group"
+  "Cultures Tested" = "redcap_data_access_group",
+  "Length of Admission" = "days_admit",
+  "Systemic Steroids" = "steroids_system",
+  "Oral Steroids" = "steroids_oral",
+  "IV Steroids" = "steroids_iv"
 )
@@
   select(
-    sid, redcap_event_name, redcap_data_access_group, year, cftr_mod,
-    cftr_mod_spec, Trikafta, fungal_test, fungi___1:fungi___5,
+    sid, redcap_event_name, redcap_data_access_group, year, days_admit,
+    cftr_mod, cftr_mod_spec, Trikafta, fungal_test, fungi___1:fungi___5,
     ntm_test, ntm___1:ntm___3,
-    admit_cx_results___1:admit_cx_results___9
+    admit_cx_results___1:admit_cx_results___9,
+    ppfev1_admit, ppfev1_discharge, ppfev1_baseline,
+    contains("steroids")
   ) %>%
@@
 ```
+
+# Days of admission
+
+```{r}
+#| label: tbl-days
+#| tbl-cap: Length of Admission
+df %>%
+  select(year, days_admit) %>%
+  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes()
+```
+
+```{r}
+#| label: tbl-days_center
+#| tbl-cap: Length of Admission by Center
+#| message: false
+df %>%
+  select(year, redcap_data_access_group, days_admit) %>%
+  tbl_strata(
+    strata = redcap_data_access_group,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-days_cftr_mod
+#| tbl-cap: Length of Admission by Modulator Status
+#| message: false
+df %>%
+  select(year, cftr_mod_spec, days_admit) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(
+          by = cftr_mod_spec, missing_text = "Missing", label = labels
+        ) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-days_trikafta
+#| tbl-cap: Length of Admission by Trikafta Status
+#| message: false
+df %>%
+  select(year, Trikafta, days_admit) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+# Steroids
+
+```{r}
+#| label: tbl-steroids
+#| tbl-cap: Steroids
+df %>%
+  select(year, contains("steroids")) %>%
+  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes()
+```
+
+```{r}
+#| label: tbl-steroids_center
+#| tbl-cap: Steroids by Center
+#| message: false
+df %>%
+  select(year, redcap_data_access_group, contains("steroids")) %>%
+  tbl_strata(
+    strata = redcap_data_access_group,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-steroids_cftr_mod
+#| tbl-cap: Steroids by Modulator Status
+#| message: false
+df %>%
+  select(year, cftr_mod_spec, contains("steroids")) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(
+          by = cftr_mod_spec, missing_text = "Missing", label = labels
+        ) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-steroids_trikafta
+#| tbl-cap: Steroids by Trikafta Status
+#| message: false
+df %>%
+  select(year, Trikafta, contains("steroids")) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+# Plots
+
+```{r}
+# Number of hospitalizations
+df %>%
+  group_by(year) %>%
+  summarise(n = n()) %>%
+  ggplot(aes(x = factor(year), y = n, fill = factor(year))) +
+  geom_bar(position = "dodge", stat = "identity") +
+  geom_label(aes(label = n),nudge_y = -20, fill = "white") +
+  ylab("Num. of Hospitalizations") +
+  theme_bw() +
+  theme(legend.position = "none", axis.title.x = element_blank()) +
+  scale_fill_manual(values = c(purple, yellow))
+```
+
+```{r}
+# % hospitalized 2 or more times in both years
+df %>%
+  group_by(sid, year) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  group_by(year) %>%
+  summarise(
+    num = sum(n >= 2), total = n(),
+    perc = mean(n >= 2) * 100, .groups = "drop"
+  ) %>%
+  ggplot(aes(x = factor(year), y = perc, fill = factor(year))) +
+  geom_bar(position = "dodge", stat = "identity") +
+  geom_label(aes(label = scales::percent(perc, scale = 1, accuracy = 0.1)),
+    nudge_y = -5, fill = "white"
+  ) +
+  geom_label(aes(label = paste0(num, "/", total)),
+    nudge_y = -8, fill = "white"
+  ) +
+  geom_signif(
+    xmin = 1, xmax = 2, y_position = 50, map_signif_level = T,
+    annotation = "<0.001"
+  ) +
+  xlab("Site") +
+  ylab("Percent With >= 2 Hospitalizations") +
+  theme_bw() +
+  theme(legend.position = "none") +
+  scale_fill_manual(values = c(purple, yellow))
+```
+
+```{r}
+# Number hospitalized 2 or more times in both years
+df %>%
+  group_by(sid, year) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  group_by(year) %>%
+  summarise(
+    num = sum(n >= 2), total = n(),
+    .groups = "drop"
+  ) %>%
+  ggplot(aes(x = factor(year), y = num, fill = factor(year))) +
+  geom_bar(position = "dodge", stat = "identity") +
+  geom_label(aes(label = num), nudge_y = -5, fill = "white") +
+  xlab("Site") +
+  ylab("Number With >= 2 Hospitalizations") +
+  theme_bw() +
+  theme(legend.position = "none") +
+  scale_fill_manual(values = c(purple, yellow))
+```
+
+```{r}
+#| warning: false
+df %>%
+  group_by(year, sid) %>%
+  rename(
+    Admission = ppfev1_admit, Discharge = ppfev1_discharge,
+    Baseline = ppfev1_baseline
+  ) %>%
+  pivot_longer(c(Admission, Discharge, Baseline)) %>%
+  mutate(name = factor(name,
+    levels = c("Baseline", "Admission", "Discharge")
+  )) %>%
+  ggplot(aes(
+    x = name, y = value,
+    fill = factor(year)
+  )) +
+  geom_boxplot() +
+  geom_pwc(
+    aes(group = year),
+    tip.length = 0.01,
+    method = "t_test", label = "p.format"
+  ) +
+  ylab("FEV1 % predicted") +
+  theme_bw() +
+  theme(
+    legend.title = element_blank(), legend.position = "bottom",
+    axis.title.x = element_blank()
+  ) +
+  scale_fill_manual(values = c(purple, yellow))
+```
+
+FEV1 percent predicted was calculated using the GLI global (2022) equations with the `rspiro` R package.
+
+```{r}
+# Proportions who returned to >= 90% and >= 100% at discharge
+df %>%
+  mutate(
+    returned_90 = ppfev1_discharge >= ppfev1_baseline * 0.9,
+    returned_100 = ppfev1_discharge >= ppfev1_baseline
+  ) %>%
+  group_by(year) %>%
+  summarise(
+    n_returned_90 = sum(returned_90, na.rm = T),
+    n_returned_100 = sum(returned_100, na.rm = T),
+    total = sum(!is.na(ppfev1_discharge) & !is.na(ppfev1_baseline)),
+    returned_90 = mean(returned_90, na.rm = T) * 100,
+    returned_100 = mean(returned_100, na.rm = T) * 100,
+  ) %>%
+  ggplot(aes(x = factor(year))) +
+  geom_bar(aes(y = returned_90, color = "Returned to 90%"),
+    position = "dodge", stat = "identity", fill = yellow
+  ) +
+  geom_label(
+    aes(
+      y = returned_90,
+      label = scales::percent(returned_90, accuracy = 0.1, scale = 1)
+    ),
+    nudge_y = -5, fill = "white"
+  ) +
+  geom_label(aes(y = returned_90, label = paste0(n_returned_90, "/", total)),
+    nudge_y = -11, fill = "white"
+  ) +
+  geom_bar(aes(y = returned_100, color = "Returned to 100%"),
+    position = "dodge", stat = "identity", fill = purple
+  ) +
+  geom_label(
+    aes(
+      y = returned_100,
+      label = scales::percent(returned_100, accuracy = 0.1, scale = 1)
+    ),
+    nudge_y = -10, fill = "white"
+  ) +
+  geom_label(aes(y = returned_100, label = paste0(n_returned_100, "/", total)),
+    nudge_y = -16, fill = "white"
+  ) +
+  ylab("Percent") +
+  xlab("Year") +
+  theme_classic() +
+  scale_color_manual(
+    name = element_blank(), values = c(purple, yellow),
+    labels = c("Returned to 100%", "Returned to 90%")
+  )
+```
+
+The above plot represents the proportion of admissions where the patient returned to close to baseline lung function, and was not aggregated by patient.
+In other words, multiple admissions for the same patient were considered independent in this case.

### 2024-03-15 Jordana Hoppe/Home Collection/manuscript.qmd
fixed sensitivity and specificity

@@
   cat("\n")
-  c <- confusionMatrix((d[, "Delayed"]), reference = d[, "Clinic"], mode = "everything")
+  c <- confusionMatrix((d[, "Delayed"]),
+    reference = d[, "Clinic"], positive = "Present", mode = "everything"
+  )
   print(kable(c$table, digits = 3))
@@
   cat("\n")
-  c <- confusionMatrix((d[, "Home"]), reference = d[, "Clinic"], mode = "everything")
+  c <- confusionMatrix((d[, "Home"]),
+    reference = d[, "Clinic"], positive = "Present", mode = "everything"
+  )
   print(kable(c$table, digits = 3))
@@
   cat("\n")
-  c <- confusionMatrix((d[, "Home"]), reference = d[, "Delayed"], mode = "everything")
+  c <- confusionMatrix((d[, "Home"]),
+    reference = d[, "Delayed"], positive = "Present", mode = "everything"
+  )
   print(kable(c$table, digits = 3))
@@
 ```{r}
-yn_df = t %>% select(all_of(yn)) %>% rename(any_of(setNames(names(labels), labels)))
-yn_df = data.frame(lapply(yn_df,factor,levels=c("No","Yes")),check.names = F)
+yn_df <- t %>%
+  select(all_of(yn)) %>%
+  rename(any_of(setNames(names(labels), labels)))
+yn_df <- data.frame(lapply(yn_df, factor, levels = c("No", "Yes")), check.names = F)
 plot(likert(yn_df))
@@
 ```{r}
-check_df = t %>% select(all_of(checkboxes)) %>% rename(any_of(setNames(names(labels), labels)))
+check_df <- t %>%
+  select(all_of(checkboxes)) %>%
+  rename(any_of(setNames(names(labels), labels)))
 plot(likert(check_df))
@@
 ```{r}
-df3 = t %>% select(all_of(levels3)) %>% rename(any_of(setNames(names(labels), labels)))
+df3 <- t %>%
+  select(all_of(levels3)) %>%
+  rename(any_of(setNames(names(labels), labels)))
 plot(likert(df3))
@@
 ```{r}
-df5 = t %>% select(all_of(levels5)) %>% rename(any_of(setNames(names(labels), labels)))
+df5 <- t %>%
+  select(all_of(levels5)) %>%
+  rename(any_of(setNames(names(labels), labels)))
 plot(likert(df5))

### 2024-03-15 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
fixed sensitivity and specificity

--- 2024-02-28 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
@@
       dark: darkly
-bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
+
 editor: source
@@
 library(irr)
-library(knitr)
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF"
-)
-knitr::opts_knit$set(root.dir = home_dir)
+library(gt)
+library(caret)
 ```
@@
     staphylococcus_aureus_sos:other_result_sos
+  ) %>%
+  filter(!record_id %in% exclude)
+# Remove empty demographics rows
+db_demographics <- db_demographics %>%
+  filter(
+    redcap_event_name == "Clinic Visit (Arm 1: Eligible)",
+    !record_id %in% exclude
   )
+# Excluded participants
+db_medications <- db_medications %>% filter(!record_id %in% exclude)
+# Combine race and ethnicity columns
+races <- c(
+  "American Indian or Alaskan Native", "Asian", "Hawaiian or Pacific Islander",
+  "Black or African American", "White", "Unknown", "Other"
+)
+ethnicities <- c("Hispanic", "Non-Hispanic", "Unknown/Not Reported")
+db_demographics$Race <- apply(db_demographics, 1, function(r) {
+  w <- which(r[grep("race", names(r))] == "Checked")
+  if (length(w) == 1) {
+    return(races[w])
+  } else if (length(w) > 1) {
+    return("Multiple Races")
+  } else if (length(w) == 0) {
+    return(NA)
+  }
+})
+db_demographics$Ethnicity <- apply(db_demographics, 1, function(r) {
+  w <- which(r[grep("ethnicity", names(r))] == "Checked")
+  if (length(w) == 1) {
+    return(ethnicities[w])
+  } else if (length(w) > 1) {
+    return(paste(ethnicities[w], sep = "/"))
+  } else if (length(w) == 0) {
+    return(NA)
+  }
+})
+# # Drop levels
+# level_drop <- c("sex", "Race", "Ethnicity")
+# db_demographics[, level_drop] <- lapply(
+#   db_demographics[, level_drop],
+#   function(c) {
+#     droplevels(factor(c))
+#   }
+# )
+# Add groups to other dataframes
+db_medications <- left_join(db_medications,
+  db_demographics %>% select(record_id, group),
+  by = join_by(record_id)
+)
+# Bug list
+bugs <- c(
+  "staphylococcus_aureus", "haemophilus_influenza", "pseudomonas_aeruginosa",
+  "pseudomonas_non_aero", "s_maltophilia", "burkholderia", "aspergillus", "ntm",
+  "other"
+)
+bug_names <- c(
+  "S. aureus", "H. influenza", "P. aeruginosa", "Pseudomonas non-aeruginosa",
+  "S. maltophilia", "Burkholderia", "Aspergillus", "NTM", "Other"
+)
+# Convert yes/no to positive/negative
+df[, paste0(bugs, "_cos")] <- lapply(df[, paste0(bugs, "_cos")], factor,
+  levels = c("No", "Yes"), labels = c("Negative", "Positive")
+)
+df[, paste0(bugs, "_sos")] <- lapply(df[, paste0(bugs, "_sos")], factor,
+  levels = c("No", "Yes"), labels = c("Negative", "Positive")
+)
+# Create overall columns (any bug detected in sample)
+df$overall_cos <- apply(df, 1, function(r) {
+  any(r[paste0(bugs, "_cos")] == "Positive")
+})
+df$overall_cos <- factor(df$overall_cos,
+  levels = c(F, T), labels = c("Negative", "Positive")
+)
+df$overall_sos <- apply(df, 1, function(r) {
+  any(r[paste0(bugs, "_sos")] == "Positive")
+})
+df$overall_sos <- factor(df$overall_sos,
+  levels = c(F, T), labels = c("Negative", "Positive")
+)
+# Add overall to list
+bugs <- c(bugs, "overall")
+bug_names <- c(bug_names, "Overall")
+bug_list <- as.list(bug_names)
+names(bug_list) <- bugs
+```
+
+# Data cleaning
+
+Participants 01-030_W, 01-038_W, 02-0038, 02-0039, and 04-001 were excluded from all analyses.
+Participants missing either the COS or SOS sample were excluded from agreement analyses, but retained in feasibility and other analyses.
+
+# Demographics
+
+```{r}
+db_demographics %>%
+  select(group, sex, Race, Ethnicity) %>%
+  tbl_summary(by = group, label = list("sex" = "Sex")) %>%
+  add_overall() %>%
+  add_p(test = list(c("Race", "Ethnicity") ~ "fisher.test")) %>%
+  separate_p_footnotes() %>%
+  as_gt()
+```
+
+# Medications
+
+```{r}
+db_medications %>%
+  select(group, azithromycin:please_specify_mucolytic) %>%
+  select(
+    -inhaled_colisitin_status, -inhaled_vancomycin_status,
+    -other_inhaled_status, -please_specify_mucolytic
+  ) %>%
+  tbl_summary(by = group) %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes() %>%
+  as_gt()
 ```
@@
 
+Given a 2x2 table of the form
+
+| SOS      | COS      | n |
+|----------|----------|---|
+| Negative | Negative | D |
+| Positive | Negative | B |
+| Negative | Positive | C |
+| Positive | Positive | A |
+
+the accuracy metrics are calculated as follows:
+
+$Sensitivity=\frac{A}{A+C}$
+
+$Specificity=\frac{D}{B+D}$
+
+$Prevalence=\frac{A+C}{A+B+C+D}$
+
+$PPV=\frac{sensitivity∗prevalence}{(sensitivity∗prevalence)+((1−specificity)∗(1−prevalence))}$
+
+$NPV=\frac{specificity∗(1−prevalence)}{((1−sensitivity)∗prevalence)+((specificity)∗(1−prevalence))}$
+
+$DetectionRate=\frac{A}{A+B+C+D}$
+
+$DetectionPrevalence=\frac{A+B}{A+B+C+D}$
+
+$BalancedAccuracy=\frac{sensitivity+specificity}{2}$
+
+$Precision=\frac{A}{A+B}$
+
+$Recall=\frac{A}{A+C}$
+
+$F1=\frac{2∗precision∗recall}{precision+recall}$
+
 ```{r}
@@
   d <- data[, c(cos, sos)]
-  a <- agree(d)
-  k <- kappa2(d)
-  ct <- table(d[, cos])
-  st <- table(d[, sos])
-  return(list(
-    "COS Absent" = ct[1], "COS Present" = ct[2],
-    "SOS Absent" = st[1], "SOS Present" = st[2],
-    "Percent Agreement" = a$value, "Cohen's Kappa" = k$value,
-    "p value" = k$p.value
-  ))
+  d <- d[complete.cases(d), ]
+  # Agreement metrics
+  c <- confusionMatrix((d[, sos]),
+    reference = d[, cos], positive = "Positive", mode = "everything"
+  )
+  # Return
+  return(list(two_by_two = c$table, accuracy = c$overall, metrics = c$byClass))
 }
@@
 
-## Overall
-
-## By bug
-
 ```{r}
-bugs <- c(
-  "staphylococcus_aureus", "haemophilus_influenza", "pseudomonas_aeruginosa",
-  "pseudomonas_non_aero", "s_maltophilia", "burkholderia", "aspergillus", "ntm",
-  "other"
-)
-bug_names <- c(
-  "S. aureus", "H. influenza", "P. aeruginosa", "Pseudomonas non-aeruginosa",
-  "S. maltophilia", "Burkholderia", "Aspergillus", "NTM", "Other"
-)
-bug_list = as.list(bug_names)
-names(bug_list) = bugs
 l <- lapply(bugs, function(b) {
-  agreement(b)
+  a <- agreement(b)
+  return(list(
+    Accuracy = a$accuracy["Accuracy"],
+    `Accuracy Lower Limit` = a$accuracy["AccuracyLower"],
+    `Accuracy Upper Limit` = a$accuracy["AccuracyUpper"],
+    Kappa = a$accuracy["Kappa"], `Mcnemar P Value` = a$accuracy["McnemarPValue"]
+  ))
 })
@@
 rownames(l) <- bug_names
-kable(l, digits = 2)
+gt(l, rownames_to_stub = T) %>% fmt_number(n_sigfig = 3)
 ```
@@
   d <- df[, c(cos, sos)]
-  t = data.frame(table(d))
-  colnames(t) = c("COS","SOS","n")
+  d <- d[complete.cases(d), ]
   cat("\n")
-  print(kable(t))
+  c <- confusionMatrix(d[, sos],
+    reference = d[, cos], positive = "Positive", mode = "everything"
+  )
+  metrics <- data.frame(c$overall)
+  metrics2 <- data.frame(c$byClass)
+  cat("### 2x2 table")
+  cat("\n")
+  two_by_two <- data.frame(c$table)
+  colnames(two_by_two) <- c("SOS", "COS", "n")
+  print(gt(two_by_two))
+  cat("\n")
+  cat("### Metrics")
+  cat("\n")
+  print(gt(metrics, rownames_to_stub = T) %>%
+    fmt_number(n_sigfig = 3) %>%
+    tab_options(column_labels.hidden = TRUE))
+  cat("\n")
+  print(gt(metrics2, rownames_to_stub = T) %>%
+    fmt_number(n_sigfig = 3) %>%
+    tab_options(column_labels.hidden = TRUE))
   cat("\n")
@@
 ```
+
+# Questions
+
+1. Are we okay with calculating prevalence from the data for the PPV and NPV? Do these numbers seem pretty representative of what you'd expect in the population?

### 2024-03-18 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
ETI survival analysis

@@
       dark: darkly
-bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
 editor: source
@@
 library(mcp)
+library(survival)
 library(ggsurvfit)
+library(ggstats)
+library(naniar)
 # Paths
@@
   "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF"
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF"
 )
@@
 ```{r results='asis'}
-invisible(lapply(c("2019", "2020", "2021"), function(n) {
+invisible(lapply(c("2019", "2020", "2021", "2022"), function(n) {
   cat("\n")
@@
   mutate(stop = as.numeric(encounterdate - eti_elig_date)) %>%
-  filter(stop >= 0) %>%
+  filter(stop > 0) %>%
   group_by(eDWID) %>%
   mutate(start = lag(stop, default = 0)) %>%
-  select(eDWID, race, encounterdate, encounterage, eti_elig_date, Vx445comb, start, stop) %>%
+  select(
+    eDWID, race, Insurance, family_income, patient_education, father_education,
+    mother_education, encounterdate, encounterage, eti_elig_date, Vx445comb,
+    start, stop
+  ) %>%
   filter(stop <= first(stop[Vx445comb == "Yes"]))
@@
 # Fit
-fit <- survfit(Surv(start, stop, Vx445comb, type = "interval") ~ race,
-  id = eDWID, data = surv_dat
-)
+y <- "Surv(time = start, time2 = stop, event = Vx445comb)"
+surv_f <- as.formula(paste0(y, "~race"))
+fit <- survfit(surv_f, id = eDWID, data = surv_dat)
+```
+
+### Days to percentile started on ETI
+
+```{r}
+# Quantiles
+tbl_survfit(fit, probs = c(0.25, 0.5, 0.75))
+```
+
+### Plots
+
+```{r}
+# Plot
+names(fit$strata) <- gsub("race=", "", names(fit$strata))
+fit %>%
+  ggsurvfit(type = "risk") +
+  add_confidence_interval() +
+  add_risktable(
+    risktable_stats =
+      c("{round(n.risk)}", "{round(cum.event)}")
+  ) +
+  add_risktable_strata_symbol() +
+  add_quantile() +
+  ylab("Proportion on Vx445comb") + xlab("Days From Eligibility")
+```
+
+Without Asian subgroup for clearer plot:
+
+```{r}
+fit <- survfit(surv_f, id = eDWID, data = surv_dat[surv_dat$race != "Asian", ])
 # Plot
@@
 
-Without Asian subgroup for clearer plot:
+### Cox PH models
+
+#### Base model (unadjusted)
 
 ```{r}
-fit <- survfit(Surv(start, stop, Vx445comb, type = "interval") ~ race,
-  id = eDWID, data = surv_dat[surv_dat$race != "Asian", ]
+cox_res <- coxph(surv_f, data = surv_dat, cluster = eDWID)
+cox_res %>%
+  tbl_regression() %>%
+  add_n(location = "level") %>%
+  add_nevent(location = "level")
+ggcoef_model(cox_res,
+  show_p_values = F, add_reference_rows = F,
+  signif_stars = F, colour = F,
 )
-# Plot
-names(fit$strata) <- gsub("race=", "", names(fit$strata))
-fit %>%
-  ggsurvfit(type = "risk") +
-  add_confidence_interval() +
-  add_risktable(
-    risktable_stats =
-      c("{round(n.risk)}", "{round(cum.event)}"),
-    stats_label = c("Eligible", "Num. Started")
-  ) +
-  add_risktable_strata_symbol() +
-  add_quantile() +
-  ylab("Proportion on Vx445comb") + xlab("Days From Eligibility")
+```
+
+#### Adjusted for insurance status, family income, and education
+
+```{r}
+new_surv_f <- update(surv_f, . ~ . + Insurance + family_income +
+  patient_education + father_education + mother_education)
+cox_res <- coxph(new_surv_f, data = surv_dat, cluster = eDWID)
+cox_res %>%
+  tbl_regression() %>%
+  add_n(location = "level") %>%
+  add_nevent(location = "level")
+ggcoef_model(cox_res,
+  show_p_values = F, add_reference_rows = F,
+  signif_stars = F, colour = F,
+)
+```
+
+##### Missing data
+
+```{r}
+gg_miss_var(surv_dat, show_pct = T, facet = race)
+```
+
+#### Adjusted for insurance status only
+
+```{r}
+new_surv_f <- update(surv_f, . ~ . + Insurance)
+cox_res <- coxph(new_surv_f, data = surv_dat, cluster = eDWID)
+cox_res %>%
+  tbl_regression() %>%
+  add_n(location = "level") %>%
+  add_nevent(location = "level")
+ggcoef_model(cox_res,
+  show_p_values = F, add_reference_rows = F,
+  signif_stars = F, colour = F,
+)
+```
+
+#### Adjusted for combined insurance status only
+
+Health Insurance Policy (e.g. Private Insurance) = Private
+
+Indian Health Service = Public
+
+Medicaid = Public
+
+Medicare = Public
+
+More than one = More than one
+
+No Insurance = No Insurance
+
+Other = Other
+
+State special needs program, e.g., BCMH, CCS, CRS, GHPP, etc. = Other
+
+TriCare or other military health plan = Tricare
+
+```{r}
+surv_dat$Insurance <- factor(surv_dat$Insurance,
+  levels = c(
+    "Health Insurance Policy (e.g. Private Insurance)",
+    "Indian Health Service",
+    "Medicaid",
+    "Medicare",
+    "More than one",
+    "No Insurance",
+    "Other",
+    "State special needs program, e.g., BCMH, CCS, CRS, GHPP, etc.",
+    "TriCare or other military health plan"
+  ),
+  labels = c(
+    "Private", "Public", "Public", "Public", "More Than One", "No Insurance",
+    "Other", "Other", "Tricare"
+  )
+)
+cox_res <- coxph(new_surv_f, data = surv_dat, cluster = eDWID)
+cox_res %>%
+  tbl_regression() %>%
+  add_n(location = "level") %>%
+  add_nevent(location = "level")
+ggcoef_model(cox_res,
+  show_p_values = F, add_reference_rows = F,
+  signif_stars = F, colour = F,
+)
 ```

### 2024-03-18 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R

@@ -3,17 +3,272 @@ library(lubridate)
 library(readxl)
 library(rspiro)
 library(Hmisc)
+# Paths
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF"
+)
+setwd(home_dir)
 # Import
-annual <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_Annualized_Del1.csv", na.strings = "")
-encounter <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_encountersMerged_Del1.csv", na.strings = "")
-demo <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_DemogCFDiag_Del1.csv", na.strings = "")
+annual <- read.csv("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_Annualized_Del1.csv", na.strings = "")
+encounter <- read.csv("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_encountersMerged_Del1.csv", na.strings = "")
+demo <- read.csv("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_DemogCFDiag_Del1.csv", na.strings = "")
 # Import responsive mutations list
-responsive_mutations <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/eti_responsive_mutations.csv", header = F)
[…]

### 2024-03-18 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
minor chnges

@@
 #| include: false
+library(redcapAPI)
 library(tidyverse)
+library(gt)
 library(gtsummary)
-library(redcapAPI)
-library(irr)
-library(gt)
 library(caret)
@@
 # Add overall to list
-bugs <- c(bugs, "overall")
-bug_names <- c(bug_names, "Overall")
+bugs <- c("overall", bugs)
+bug_names <- c("Across all bugs", bug_names)
 bug_list <- as.list(bug_names)
 names(bug_list) <- bugs
+# Agreement columns
+df[, bugs] <- lapply(bugs, function(b) {
+  cos <- df[, paste0(b, "_cos")]
+  sos <- df[, paste0(b, "_sos")]
+  f <- factor(paste(cos, sos, sep = "_"),
+    exclude = c(
+      "NA_Negative", "Negative_NA", "NA_Positive", "Positive_NA", "NA_NA"
+    )
+  )
+  f <- factor(f,
+    levels = c(
+      "Negative_Negative", "Negative_Positive", "Positive_Negative",
+      "Positive_Positive"
+    ),
+    labels = c(
+      "COS Negative / SOS Negative", "COS Negative / SOS Positive",
+      "COS Positive / SOS Negative", "COS Positive / SOS Positive"
+    )
+  )
+  return(f)
+})
 ```
@@
 
-Cohen's kappa is interpreted analagously to a correlation coefficient, with -1 indicating perfect disagreement and 1 indicating perfect agreement.
-Tables with NaN and NAs indicate that all tests had the same values (i.e.
-all absent or all present).
-
-| Cohen's Kappa | Interpretation         |
-|---------------|------------------------|
-| 0             | No agreement           |
-| 0.10 - 0.20   | Slight agreement       |
-| 0.21 - 0.40   | Fair agreement         |
-| 0.41 - 0.60   | Moderate agreement     |
-| 0.61 - 0.80   | Substantial agreement  |
-| 0.81 - 0.99   | Near perfect agreement |
-| 1             | Perfect agreement      |
-
-Given a 2x2 table of the form
-
-| SOS      | COS      | n |
-|----------|----------|---|
-| Negative | Negative | D |
-| Positive | Negative | B |
-| Negative | Positive | C |
-| Positive | Positive | A |
-
-the accuracy metrics are calculated as follows:
-
-$Sensitivity=\frac{A}{A+C}$
-
-$Specificity=\frac{D}{B+D}$
-
-$Prevalence=\frac{A+C}{A+B+C+D}$
-
-$PPV=\frac{sensitivity∗prevalence}{(sensitivity∗prevalence)+((1−specificity)∗(1−prevalence))}$
-
-$NPV=\frac{specificity∗(1−prevalence)}{((1−sensitivity)∗prevalence)+((specificity)∗(1−prevalence))}$
-
-$DetectionRate=\frac{A}{A+B+C+D}$
-
-$DetectionPrevalence=\frac{A+B}{A+B+C+D}$
-
-$BalancedAccuracy=\frac{sensitivity+specificity}{2}$
-
-$Precision=\frac{A}{A+B}$
-
-$Recall=\frac{A}{A+C}$
-
-$F1=\frac{2∗precision∗recall}{precision+recall}$
+```{r}
+df %>%
+  select(all_of(bugs)) %>%
+  tbl_summary(label = bug_list, missing_text = "Missing") %>%
+  add_ci(statistic=list(all_categorical() ~ "{conf.low} - {conf.high}%"))
+```
 

### 2024-03-18 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
minor chnges

@@
 ```{r results='asis'}
-invisible(lapply(c("2019", "2020", "2021", "2022"), function(n) {
+invisible(lapply(c("2020", "2021", "2022"), function(n) {
   cat("\n")
@@
     eDWID, race, Insurance, family_income, patient_education, father_education,
-    mother_education, encounterdate, encounterage, eti_elig_date, Vx445comb,
-    start, stop
+    mother_education, encounterdate, encounterage, gli_fev1_ppred_rn,
+    eti_elig_date, Vx445comb, start, stop
   ) %>%
@@
 ```{r}
+#| warning: false
 # Plot
@@
   add_quantile() +
-  ylab("Proportion on Vx445comb") + xlab("Days From Eligibility")
+  ylab("Proportion on ETI") + xlab("Days From Eligibility")
 ```
@@
 ```{r}
+#| warning: false
 fit <- survfit(surv_f, id = eDWID, data = surv_dat[surv_dat$race != "Asian", ])
@@
   show_p_values = F, add_reference_rows = F,
-  signif_stars = F, colour = F,
-)
-```
-
-#### Adjusted for insurance status, family income, and education
-
-```{r}
-new_surv_f <- update(surv_f, . ~ . + Insurance + family_income +
-  patient_education + father_education + mother_education)
-cox_res <- coxph(new_surv_f, data = surv_dat, cluster = eDWID)
-cox_res %>%
-  tbl_regression() %>%
-  add_n(location = "level") %>%
-  add_nevent(location = "level")
-ggcoef_model(cox_res,
-  show_p_values = F, add_reference_rows = F,
-  signif_stars = F, colour = F,
+  signif_stars = F, colour = F, exponentiate = T
 )
@@
 gg_miss_var(surv_dat, show_pct = T, facet = race)
-```
-
-#### Adjusted for insurance status only
-
-```{r}
-new_surv_f <- update(surv_f, . ~ . + Insurance)
-cox_res <- coxph(new_surv_f, data = surv_dat, cluster = eDWID)
-cox_res %>%
-  tbl_regression() %>%
-  add_n(location = "level") %>%
-  add_nevent(location = "level")
-ggcoef_model(cox_res,
-  show_p_values = F, add_reference_rows = F,
-  signif_stars = F, colour = F,
-)
 ```
@@
 
-State special needs program, e.g., BCMH, CCS, CRS, GHPP, etc. = Other
+State special needs program, e.g., BCMH, CCS, CRS, GHPP, etc. = Public
 
@@
     "Private", "Public", "Public", "Public", "More Than One", "No Insurance",
-    "Other", "Other", "Tricare"
+    "Other", "Public", "Tricare"
   )
 )
+new_surv_f <- update(surv_f, . ~ . + Insurance)
 cox_res <- coxph(new_surv_f, data = surv_dat, cluster = eDWID)
@@
   show_p_values = F, add_reference_rows = F,
-  signif_stars = F, colour = F,
+  signif_stars = F, colour = F, exponentiate = T
 )
@@
 
-# People without visits in a given year
+#### Adjusted for insurance, age group, and FEV1
 
 ```{r}
-# Which are missing from encounters data?
-a <- as.data.frame.matrix(table(annual$eDWID, annual$ReviewYear))
-a <- a %>% rownames_to_column("eDWID")
-e <- as.data.frame.matrix(table(by_year$eDWID, by_year$reviewyear))
-e <- e %>% rownames_to_column("eDWID")
-rowdiff_a <- setdiff(a, e)
-rowdiff_a$data <- "annual"
-rowdiff_e <- setdiff(e, a)
-rowdiff_e$data <- "encounter"
-diffs <- rbind(rowdiff_a, rowdiff_e)
-diffs <- diffs %>%
-  arrange(eDWID, desc(data)) %>%
-  select(-data)
-diffs <- diffs %>%
-  group_by(eDWID) %>%
-  reframe(across(`2015`:`2021`, ~ diff(as.numeric(.x))))
-diffs %>%
-  reframe(across(`2015`:`2021`, ~ sum(.x == 1))) %>%
-  gt()
+surv_dat$`Age Group` <- cut(surv_dat$encounterage,
+  breaks = c(-Inf, 6, 12, 18, 30, Inf), right = F
+)
+new_surv_f <- update(surv_f, . ~ . + Insurance + `Age Group` +
+  gli_fev1_ppred_rn)
+cox_res <- coxph(new_surv_f, data = surv_dat, cluster = eDWID)
+cox_res %>%
+  tbl_regression() %>%
+  add_n(location = "level") %>%
+  add_nevent(location = "level")
+ggcoef_model(cox_res,
+  show_p_values = F, add_reference_rows = F,
+  signif_stars = F, colour = F, exponentiate = T
+)
 ```
-
-There were a total of `r length(setdiff(a$eDWID,e$eDWID))` included in the annual dataset that did not appear in the encounter-level dataset at all.
-There were an additional `r nrow(diffs)` patients who were included in both datasets, but had at least one year with no encounters.
-The table above shows the number of people missing an encounter in each year.
-
-If someone was not seen a given year, do we count them as eligible for that year? I am assuming they will be eligible but obviously will not have started any modulators.
-
-# Potential COVID effects
-
-## Spaghetti plot
-
-```{r}
-#| warning: false
-ggplot(encounter, aes(
-  y = GLI_FEV1_pct_predicted, x = encounterdate,
-  group = eDWID, color = factor(eDWID)
-)) +
-  geom_line(alpha = 0.02) +
-  theme_bw() +
-  theme(legend.position = "none")
-```
-
-## Visits prior to and after March 11, 2020
-
-The table below is limited to visits 2019 - 2021.
-
-```{r}
-encounter %>%
-  filter(encounterdate >= "2019-01-01") %>%
-  mutate(post_covid = encounterdate > "2020-03-11") %>%
-  group_by(eDWID) %>%
-  summarise(
-    `Number of Visits Pre-COVID` = sum(!post_covid),
-    `Number of Visits Post-COVID` = sum(post_covid),
-    race = first(race),
-    .groups = "drop"
-  ) %>%
-  select(-eDWID) %>%
-  tbl_summary(by = race, missing_text = "Missing") %>%
-  bold_labels()
-```
-
-# Lung function
-
-## Annualized averages
-
-FEV1 and FVC percent predicted (based on GLI equations) were averaged by year for each person.
-
-### FEV1
-
-#### Basic model
-
-The most basic model we can fit to these data uses the average values as the outcome, with terms for review year, race, and their interaction and a random intercept for each patient.
-This model is adjusted for age, private vs.
-other insurance, and parental education.
-
-```{r}
-# Most basic model of lung function over time by race
-# Try a few different models
-mod <- lmer(
-  mean_ppFEV ~ factor(ReviewYear) * race + (1 | eDWID),
-  data = annual
-)
-res <- tidy(mod, effects = "fixed")
-means <- data.frame(emmeans(mod, ~ race | ReviewYear,
-  lmerTest.limit = 1e6,
-  pbkrtest.limit = 1e6
-))
-```

### 2024-03-18 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
ETI Cox models mostly finished

@@
 So if a participant was reported as both Black and Hispanic, they were consider Hispanic for the following analyses.
-Those with missing data for Hispanic ethnicity were moved to the "Indigenous/Mixed/Other/Unknown Race" category.
+Those with missing data for Hispanic ethnicity were moved to the "Other" category.
 
@@
 ```{r}
-annual %>%
-  filter(eti_elig == "Yes" | Vx445comb == "Yes") %>%
+p <- annual %>%
+  filter(eti_elig == "Yes" | Vx445comb == "Yes", ReviewYear > 2019) %>%
   select(eDWID, race, ReviewYear, contains("Vx")) %>%
@@
   ylab("Percentage Prescribed ETI") +
-  theme_bw() +
-  theme(legend.title = element_blank()) +
+  theme_bw(base_size = 16) +
+  theme(legend.title = element_blank(), legend.text = element_text(size = 14)) +
   scale_fill_manual(values = c(
@@
   ))
+p
+```
+
+```{r}
+#| warning: false
+# Save image
+ggsave("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Figures/bar_chart.png",
+  plot = p, device = "png", width = 12, height = 8, units = "in"
+)
 ```
@@
 surv_dat <- encounter %>%
+  filter(eti_elig == "Yes") %>%
   mutate(stop = as.numeric(encounterdate - eti_elig_date)) %>%
@@
     eti_elig_date, Vx445comb, start, stop
-  ) %>%
-  filter(stop <= first(stop[Vx445comb == "Yes"]))
+  )
 surv_dat$Vx445comb <- as.numeric(factor(surv_dat$Vx445comb, levels = c("No", "Yes"))) - 1
@@
 # Quantiles
-tbl_survfit(fit, probs = c(0.25, 0.5, 0.75))
+tbl_survfit(fit, probs = c(0.25, 0.5, 0.75, 0.9))
 ```
@@
 names(fit$strata) <- gsub("race=", "", names(fit$strata))
-fit %>%
+p <- fit %>%
   ggsurvfit(type = "risk") +
+  xlim(c(0, 365 * 2)) +
   add_confidence_interval() +
-  add_risktable(
-    risktable_stats =
-      c("{round(n.risk)}", "{round(cum.event)}")
-  ) +
-  add_risktable_strata_symbol() +
   add_quantile() +
-  ylab("Proportion on ETI") + xlab("Days From Eligibility")
+  ylab("Proportion on ETI") + xlab("Days From Eligibility") +
+  theme_bw(base_size = 16) +
+  theme(legend.title = element_blank(), legend.text = element_text(size = 14))
+p
+```
+
+```{r}
+#| warning: false
+ggsave("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Figures/survival.png",
+  plot = p, device = "png", width = 12, height = 8, units = "in"
+)
 ```
@@
 names(fit$strata) <- gsub("race=", "", names(fit$strata))
-fit %>%
+p <- fit %>%
   ggsurvfit(type = "risk") +
+  xlim(c(0, 365 * 2)) +
   add_confidence_interval() +
-  add_risktable(
-    risktable_stats =
-      c("{round(n.risk)}", "{round(cum.event)}"),
-    stats_label = c("Eligible", "Num. Started")
-  ) +
-  add_risktable_strata_symbol() +
   add_quantile() +
-  ylab("Proportion on ETI") + xlab("Days From Eligibility")
+  ylab("Proportion on ETI") + xlab("Days From Eligibility") +
+  theme_bw(base_size = 16) +
+  theme(legend.title = element_blank(), legend.text = element_text(size = 14))
+p
+```
+
+```{r}
+#| warning: false
+ggsave("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Figures/survival_no_asian.png",
+  plot = p, device = "png", width = 12, height = 8, units = "in"
+)
 ```
@@
 ```{r}
-cox_res <- coxph(surv_f, data = surv_dat, cluster = eDWID)
+cox_res <- coxph(surv_f, data = surv_dat, id = eDWID)
 cox_res %>%
-  tbl_regression() %>%
-  add_n(location = "level") %>%
-  add_nevent(location = "level")
+  tbl_regression(exponentiate = T)
 ggcoef_model(cox_res,
@@
 cox_res %>%
-  tbl_regression() %>%
-  add_n(location = "level") %>%
-  add_nevent(location = "level")
+  tbl_regression(exponentiate = T)
 ggcoef_model(cox_res,
@@
 surv_dat$`Age Group` <- cut(surv_dat$encounterage,
-  breaks = c(-Inf, 6, 12, 18, 30, Inf), right = F
+  breaks = c(6, 12, 18, 30, Inf), right = F
 )
@@
 ```
+
+# Number of clinic visits
+
+For each participant, we divided the total number of visits by the number of unique years that they were included in the registry data. Visits per year are presented by race.
+
+```{r}
+encounter %>%
+  group_by(eDWID, race) %>%
+  summarise(
+    `Visits per Year` = n() / length(unique(reviewyear)),
+    .groups = "drop"
+  ) %>%
+  select(-eDWID) %>%
+  tbl_summary(by = race)
+```

### 2024-03-18 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R

@@ -295,9 +295,8 @@ demo$race <- factor(demo$race,
     "Native Hawaiian or Other Pacific Islander", "Mixed/Other/Unknown Race"
   ),
   labels = c(
-    "White", "Black or African American", "Hispanic or Latino", "Asian",
-    "Indigenous/Mixed/Other/Unknown Race",
-    "Indigenous/Mixed/Other/Unknown Race", "Indigenous/Mixed/Other/Unknown Race"
+    "Non-Hispanic White", "Black or African American", "Hispanic or Latino", "Asian",
+    "Other", "Other", "Other"
   )
 )
 # Merge
@@ -485,7 +484,7 @@ labels <-
     "GLI_FVC_pct_predicted" = "ppFVC", "FEV1" = "FEV1",
     "GLI_FEV1_pct_predicted" = "ppFEV1", "FEV1FVC" = "FEV1/FVC",
     "GLI_FEV1FVC_pct_predicted" = "ppFEV1/FVC",
-    "gli_fev1_ppred_rn" = "ppFEV1rn", "gli_fvc_ppred_rn" = "ppFVCrn",
+    "gli_fev1_ppred_rn" = "ppFEV1 (Race Neutral)", "gli_fvc_ppred_rn" = "ppFVCrn",
     "gli_fev1fvc_ppred_rn" = "ppFEV1/FVCrn",
     "hepatobiliary1_3" = "Liver disease, cirrhosis",
     "acutehepatitis" = "Acute hepatitis",
@@ -523,7 +522,7 @@ labels <-
     "pregnant" = "Was patient pregnant during the reporting year?",
     "Patient_Dx" = "Patient Diagnosis", "DOB_Year" = "Year of Birth",
     "Gender" = "Gender",
-    "Race1" = "White", "Race2" = "Black or African American",
+    "Race1" = "Non-Hispanic White", "Race2" = "Black or African American",
     "Race3" = "American Indian or Alaska Native", "Race4" = "Asian",
     "Race5" = "Native Hawaiian or Other Pacific Islander",
     "Race6" = "Some other race", "race" = "Race/Ethnicity",

### 2024-03-21 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
SSwAB-IT phase 1 excluded positive/positive

@@
       dark: darkly
-
+bibliography: /Users/timvigers/Documents/Notes/zotero.bib
+csl: /Users/timvigers/GitHub/styles/american-medical-association-10th-edition.csl
 editor: source
@@
 library(caret)
+library(vcd)
+library(eulerr)
 ```
@@
 
-```{r}
-df %>%
-  select(all_of(bugs)) %>%
-  tbl_summary(label = bug_list, missing_text = "Missing") %>%
-  add_ci(statistic=list(all_categorical() ~ "{conf.low} - {conf.high}%"))
-```
+For these agreement analyses, we reported accuracy metrics that include all samples.
+Accuracy is the proportion of samples where the two tests agreed (positive/positive and negative/negative).
+Cohen's kappa accounts for agreement between the two tests by chance, and is interpreted analagously to correlation where -1 indicates perfect disagreement and 1 indicates perfect agreement.
+Cohen's kapps of NaN or NA indicates a lack of variability between the two tests.
+
+| Cohen's Kappa | Interpretation         |
+|---------------|------------------------|
+| 0             | No agreement           |
+| 0.10 - 0.20   | Slight agreement       |
+| 0.21 - 0.40   | Fair agreement         |
+| 0.41 - 0.60   | Moderate agreement     |
+| 0.61 - 0.80   | Substantial agreement  |
+| 0.81 - 0.99   | Near perfect agreement |
+| 1             | Perfect agreement      |
+
+Bangdiwala's agreement plot @bangdiwalaAgreementChart2013 is a way of visualizing 2 x 2 contingency tables, and additional details are in the paper referenced at the end of this document.
 
@@
   c <- confusionMatrix((d[, sos]),
-    reference = d[, cos], positive = "Positive", mode = "everything"
+    reference = d[, cos], positive = "Positive", mode = "everything",
+    dnn = c("SOS", "COS")
   )
   # Return
-  return(list(two_by_two = c$table, accuracy = c$overall, metrics = c$byClass))
+  return(list(two_by_two = c$table, accuracy = c$overall))
 }
@@
   a <- agreement(b)
+  # Remove those where both tests were positive
+  n <- a$two_by_two
+  n[2, 2] <- 0
+  # Proportion tests for non-agreement in participants with one negative test
+  cos_p <- prop.test(n[1, 2], sum(n))
+  sos_p <- prop.test(n[2, 1], sum(n))
+  nn <- prop.test(n[1, 1], sum(n))
   return(list(
     Accuracy = a$accuracy["Accuracy"],
-    `Accuracy Lower Limit` = a$accuracy["AccuracyLower"],
-    `Accuracy Upper Limit` = a$accuracy["AccuracyUpper"],
-    Kappa = a$accuracy["Kappa"], `Mcnemar P Value` = a$accuracy["McnemarPValue"]
+    `Accuracy LL` = a$accuracy["AccuracyLower"],
+    `Accuracy UL` = a$accuracy["AccuracyUpper"],
+    Kappa = a$accuracy["Kappa"],
+    `Mcnemar P Value` = a$accuracy["McnemarPValue"],
+    `N With Negative Test on Either` = sum(n),
+    `N Negative on Both` = n[1, 1],
+    `Prop. Both Negative` = nn$estimate,
+    `Prop. Both Negative LL ` = nn$conf.int[1],
+    `Prop. Both Negative UL ` = nn$conf.int[2],
+    `N Postive on COS and Negative on SOS` = n[1, 2],
+    `Prop. Postive on COS` = cos_p$estimate,
+    `Prop. Postive on COS LL ` = cos_p$conf.int[1],
+    `Prop. Postive on COS UL ` = cos_p$conf.int[2],
+    `N Postive on SOS and Negative on COS` = n[2, 1],
+    `Prop. Postive on SOS` = sos_p$estimate,
+    `Prop. Postive on SOS LL ` = sos_p$conf.int[1],
+    `Prop. Postive on SOS UL ` = sos_p$conf.int[2]
   ))
@@
   d <- d[complete.cases(d), ]
+  colnames(d) <- c("COS", "SOS")
   cat("\n")
-  c <- confusionMatrix(d[, sos],
-    reference = d[, cos], positive = "Positive", mode = "everything"
+  c <- confusionMatrix(d[, "SOS"],
+    reference = d[, "COS"], positive = "Positive", mode = "everything"
   )
   metrics <- data.frame(c$overall)
-  metrics2 <- data.frame(c$byClass)
   cat("### 2x2 table")
@@
   two_by_two <- data.frame(c$table)
-  colnames(two_by_two) <- c("SOS", "COS", "n")
   print(gt(two_by_two))
   cat("\n")
-  cat("### Metrics")
+  cat("### Accuracy")
   cat("\n")
@@
   cat("\n")
-  print(gt(metrics2, rownames_to_stub = T) %>%
-    fmt_number(n_sigfig = 3) %>%
-    tab_options(column_labels.hidden = TRUE))
+  cat("### Agreement plot")
+  cat("\n")
+  print(agreementplot(table(d)))
+  cat("\n")
+  cat("### Positive/Positive Excluded")
+  cat("\n")
+  t <- d[rowSums(d == "Positive") < 2, ]
+  t$`Samples With a Negative Test` <- paste(
+    "COS", t[, "COS"], "/", "SOS", t[, "SOS"]
+  )
+  t %>%
+    select(`Samples With a Negative Test`) %>%
+    tbl_summary() %>%
+    add_ci() %>%
+    as_gt() %>%
+    print()
+  cat("\n")
+  cat("### Euler diagram of negative results")
+  cat("\n")
+  print(plot(euler(d == "Negative"), quantities = T))
   cat("\n")
@@
 
-# Questions
-
-1. Are we okay with calculating prevalence from the data for the PPV and NPV? Do these numbers seem pretty representative of what you'd expect in the population?
+# References

### 2024-03-21 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
CF research conference presentation rough draft

@@
       dark: darkly
-bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
+bibliography: /Users/timvigers/Documents/Notes/zotero.bib
+csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
 editor: source
@@
 library(knitr)
+library(caret)
 source("/Users/timvigers/GitHub/CF/genotype_class_severity.R")
@@
 diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
-exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude)]
+exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) | diagnoses$Code == "Exclude"]
+exclude <- exclude[!is.na(exclude)]
 # Import data
@@
 # Diagnosis per Angela
-df$Diagnosis <- factor(diagnoses$Diagnosis[match(df$PatientID, diagnoses$PatientID)])
-levels(df$Diagnosis) <- c("CFTR-RD or CRMS", "CFTR-RD or CRMS", "Cystic Fibrosis", "Non-CF")
+df$Diagnosis <- factor(diagnoses$Diagnosis[match(df$PatientID, diagnoses$PatientID)],
+  levels = c("CFTR-related disorder", "CRMS", "Cystic Fibrosis", "Non-CF"),
+  labels = c("CFTR-RD or CRMS", "CFTR-RD or CRMS", "Cystic Fibrosis", "Non-CF")
+)
 # Add modulator info
@@
 
-<!-- # Flowchart -->
+# Flowchart
 
-<!-- ```{r} -->
-<!-- grViz("digraph { -->
+```{r}
+grViz("digraph {
 
-<!-- # initiate graph -->
-<!-- graph [layout = dot] -->
+# initiate graph
+graph [layout = dot]
 
-<!-- # global node settings -->
-<!-- node [shape = rectangle] -->
+# global node settings
+node [shape = rectangle]
 
-<!-- # label nodes -->
-<!-- data1 [label = '@@1'] -->
-<!-- data2 [label = '@@2'] -->
-<!-- ineligible [label = '@@3'] -->
-<!-- demo  [label = '@@4'] -->
-<!-- highna  [label = '@@5'] -->
-<!-- highcl  [label = '@@6'] -->
-<!-- highnacl  [label = '@@7'] -->
-<!-- review [label = '@@8'] -->
-<!-- analysis [label = '@@9'] -->
+# label nodes
+data1 [label = '@@1']
+data2 [label = '@@2']
+ineligible [label = '@@3']
+demo  [label = '@@4']
+highna  [label = '@@5']
+highcl  [label = '@@6']
+highnacl  [label = '@@7']
+review [label = '@@8']
+analysis [label = '@@9']
+ratio [label = '@@10']
 
-<!-- # edge definitions with the node IDs -->
-<!-- data1  -> data2 -->
-<!-- data2 -> ineligible -->
-<!-- ineligible -> {demo,highna,highcl,highnacl,review} -->
-<!-- data2  -> analysis -->
-<!-- highnacl -> analysis [style=invis] -->
-<!-- } -->
-<!-- [1]: paste0('Sweat tests performed \\n ','n = ',n1) -->
-<!-- [2]: paste0('Unique participants \\n ','n = ',n2) -->
-<!-- [3]: paste0('Ineligible \\n ','n = ',n2-nrow(t1_df)) -->
-<!-- [4]: paste0('No demographics \\n ','n = ',length(no_dem)) -->
-<!-- [5]: paste0('Na > 160 mmol/L \\n ','n = ',length(high_na)) -->
-<!-- [6]: paste0('Cl > 160 mmol/L \\n ','n = ',length(high_cl)) -->
-<!-- [7]: paste0('Na:Cl > 2 \\n ','n = ',length(high_ratio)) -->
-<!-- [8]: paste0('Other \\n ','n = ',1) -->
-<!-- [9]: paste0('Analysis cohort \\n ','n = ',nrow(t1_df)) -->
-<!-- ") -->
-<!-- ``` -->
+# edge definitions with the node IDs
+data1  -> data2
+data2 -> ineligible
+ineligible -> {demo,highna,highcl,highnacl,review}
+data2  -> analysis -> ratio
+highnacl -> analysis [style=invis]
+}
+[1]: paste0('Sweat tests performed \\n ','n = ',n1)
+[2]: paste0('Unique participants \\n ','n = ',n2)
+[3]: paste0('Ineligible \\n ','n = ',n2-nrow(t1_df))
+[4]: paste0('No demographics \\n ','n = ',length(no_dem))
+[5]: paste0('Na > 160 mmol/L \\n ','n = ',length(high_na))
+[6]: paste0('Cl > 160 mmol/L \\n ','n = ',length(high_cl))
+[7]: paste0('Na:Cl > 2 \\n ','n = ',length(high_ratio))
+[8]: paste0('Other \\n ','n = ',1)
+[9]: paste0('Analysis cohort \\n ','n = ',nrow(t1_df))
+[10]: paste0('Both Na and Cl Measured \\n ','n = ',sum(!is.na(df$`Na/Cl Ratio`)))
+")
+```
 
@@
   geom_hline(yintercept = 1, color = "red") +
-  geom_hline(yintercept = 1.25, color = "red")
+  geom_hline(yintercept = 1.25, color = "red") +
+  theme_classic()
 df$nacl_ratio_cat <- cut(df$`Na/Cl Ratio`, breaks = c(-Inf, 1, 1.25, Inf), right = F)
@@
 ```{r}
-roc.test(roc_cf,roc_cf_cl)
+roc.test(roc_cf, roc_cf_cl)
 ```
@@
 ```{r}
-roc.test(roc_non_cf,roc_non_cf_cl)
+roc.test(roc_non_cf, roc_non_cf_cl)
 ```

### 2024-03-21 Edith Zemanick/Sweat Test/sweat_test_cf_conference.qmd
CF research conference presentation rough draft

--- 2024-03-21 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
@@
 ---
-title: "Sweat electrolytes and CFTR activity in the era of CFTR modulators"
+title: "Using Sweat Na/Cl Ratio to Predict CFTR-RD or CRMS"
 author: "Tim Vigers"
@@
 date-format: long
-format:
-  html:
-    toc: true
-    toc-depth: 10
-    toc-float: true
-    number-sections: true
-    code-fold: true
-    self-contained: true
-    fig-cap-location: top
-    page-layout: full
-    theme:
-      light: flatly
-      dark: darkly
-bibliography: /Users/timvigers/Documents/Notes/zotero.bib
+format: revealjs
+editor: source
+bibliography: Presentation.bib
 csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
-editor: source
 ---
@@
 library(knitr)
+library(kableExtra)
 library(caret)
+library(rpart)
+library(rpart.plot)
 source("/Users/timvigers/GitHub/CF/genotype_class_severity.R")
@@
 
-# Flowchart
+# Background
 
-```{r}
-grViz("digraph {
+# Our data
 
-# initiate graph
-graph [layout = dot]
+## Flowchart
 
-# global node settings
-node [shape = rectangle]
-
-# label nodes
-data1 [label = '@@1']
-data2 [label = '@@2']
-ineligible [label = '@@3']
-demo  [label = '@@4']
-highna  [label = '@@5']
-highcl  [label = '@@6']
-highnacl  [label = '@@7']
-review [label = '@@8']
-analysis [label = '@@9']
-ratio [label = '@@10']
-
-# edge definitions with the node IDs
-data1  -> data2
-data2 -> ineligible
-ineligible -> {demo,highna,highcl,highnacl,review}
-data2  -> analysis -> ratio
-highnacl -> analysis [style=invis]
-}
-[1]: paste0('Sweat tests performed \\n ','n = ',n1)
-[2]: paste0('Unique participants \\n ','n = ',n2)
-[3]: paste0('Ineligible \\n ','n = ',n2-nrow(t1_df))
-[4]: paste0('No demographics \\n ','n = ',length(no_dem))
-[5]: paste0('Na > 160 mmol/L \\n ','n = ',length(high_na))
-[6]: paste0('Cl > 160 mmol/L \\n ','n = ',length(high_cl))
-[7]: paste0('Na:Cl > 2 \\n ','n = ',length(high_ratio))
-[8]: paste0('Other \\n ','n = ',1)
-[9]: paste0('Analysis cohort \\n ','n = ',nrow(t1_df))
-[10]: paste0('Both Na and Cl Measured \\n ','n = ',sum(!is.na(df$`Na/Cl Ratio`)))
-")
-```
-
-# Cohorts by Na/Cl ratio
+![](flowchart.png)
 
@@
   theme_classic()
-df$nacl_ratio_cat <- cut(df$`Na/Cl Ratio`, breaks = c(-Inf, 1, 1.25, Inf), right = F)
-kable(table(df$nacl_ratio_cat, df$Diagnosis))
 ```
 
-## Based on k-means clustering
+## Confusion matrix
 
-### By Na and Cl
+```{r}
+df$nacl_ratio_cat <- cut(df$`Na/Cl Ratio`,
+  breaks = c(-Inf, 1, 1.25, Inf), right = F
+)
+t <- table(df$nacl_ratio_cat, df$Diagnosis)
+t <- addmargins(t)
+# colorize the diagonal elements in table with non-ints
+for (i in 1:3) {
+  t[i, i] <- cell_spec(t[i, i],
+    bold = T, color = "white", background = "darkgreen"
+  )
+}
+kable(t)
+```
+
+## Confusion matrix (proportions)
+
+```{r}
+t <- round(prop.table(table(df$nacl_ratio_cat, df$Diagnosis), margin = 2) * 100, 1)
+# colorize the diagonal elements in table with non-ints
+for (i in 1:(ncol(t))) {
+  t[i, i] <- cell_spec(t[i, i],
+    bold = T, color = "white", background = "darkgreen"
+  )
+}
+kable(t, digits = 3)
+```
+
+::: {.smaller}
+
+*Proportions in the table above are out of the column total (i.e. by diagnosis), not the total number of participants.*
+
+:::
+
+# ROC analyses
+
+## Na/Cl Ratio for CF vs. CFTR-RD or CRMS
+
+```{r}
+roc_cf <- roc(
+  response = df$`Diagnosis`,
+  predictor = df$`Na/Cl Ratio`,
+  levels = c("CFTR-RD or CRMS", "Cystic Fibrosis"),
+  quiet = T
+)
+coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)
+plot(roc_cf, print.auc = TRUE, col = "blue",print.thres=T)
+```
+
+## Na/Cl Ratio for Non-CF vs. CFTR-RD or CRMS
+
+```{r}
+roc_non_cf <- roc(
+  response = df$`Diagnosis`,
+  predictor = df$`Na/Cl Ratio`,
+  levels = c("Non-CF", "CFTR-RD or CRMS"),
+  quiet = T
+)
+coords_non_cf <- round(coords(roc_non_cf, "best", best.method = "youden"), 3)
+plot(roc_non_cf, print.auc = TRUE, col = "red",print.thres=T)
+```
+
+## Histograms by group
+
+Vertical blue line indicates the threshold for CFTR-RD or CRMS vs. CF and red indicates the threshold for non-CF vs. CFTR-RD or CRMS. Na/Cl ratio was log-transformed for easier visualization.
+
+```{r warning=FALSE}
+ggplot(
+  df[!is.na(df$`Diagnosis`), ],
+  aes(log(`Na/Cl Ratio`), fill = `Diagnosis`)
+) +
+  geom_histogram(alpha = 0.3, position = "identity", bins = 50) +
+  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
+  geom_vline(xintercept = log(coords_non_cf$threshold), color = "red") +
+  ylab("Count") +
+  xlab("ln(Na/Cl Ratio)") +
+  theme_bw()
+```
+
+# K-means clustering
+
+## Brief overview
+
+Detail here
+
+# K-means clustering by Na and Cl
 
@@
   drop_na()
-ggplot(km_df, aes(x = Na, y = Cl, color = Diagnosis)) +
-  geom_point()
 km <- kmeans(km_df[, c("Na", "Cl")], centers = 3)
@@
 )
+```
+
+## Na by Cl plot
+
+```{r}
+ggplot(km_df, aes(x = Na, y = Cl, color = Diagnosis)) +
+  geom_point(alpha = 0.5) +
+  theme_classic()
+```
+
+## Na by Cl plot with k-means centers
+
+```{r}
 ggplot(km_df, aes(x = Na, y = Cl, color = Cluster)) +
-  geom_point(shape = 20, alpha = 0.2) +
-  geom_point(data = centers, size = 5)
+  geom_point(shape = 20, alpha = 0.1) +
+  geom_point(data = centers, size = 5) +
+  theme_classic()
+```
+
+## K-means centers
+
+```{r}
 rownames(centers) <- NULL
 centers <- centers %>% column_to_rownames("Cluster")
-kable(centers, caption = "K-means centers", row.names = T)
-kable(as.data.frame.matrix(table(km_df$Cluster, km_df$Diagnosis)), row.names = T)
+kable(centers, row.names = T, digits = 1)
 ```
 
-### By Na/Cl ratio
+## Confusion matrix
 
-```{r warning=FALSE}
-set.seed(1017)
-# K means
-km_df <- df %>%
-  select(Diagnosis, `Na/Cl Ratio`) %>%
-  drop_na()
-km <- kmeans(km_df[, c("Na/Cl Ratio")], centers = 3)
-km_df$Cluster <- factor(km$cluster,
-  levels = 1:3,
-  labels = c("Low", "Medium", "High")
+```{r}
+t <- table(km_df$Cluster, km_df$Diagnosis)
+t <- addmargins(t)
+# colorize the diagonal elements in table with non-ints
+for (i in 1:3) {
+  t[i, i] <- cell_spec(t[i, i],
+    bold = T, color = "white", background = "darkgreen"
   )
-centers <- data.frame(km$centers)
-centers$Cluster <- rownames(km$centers)
-centers$Cluster <- factor(centers$Cluster,
-  levels = 1:3,
-  labels = c("Low", "Medium", "High")
-)
-rownames(centers) <- NULL
-centers <- centers %>% column_to_rownames("Cluster")
-kable(centers, caption = "K-means centers", row.names = T)
-kable(as.data.frame.matrix(table(km_df$Cluster, km_df$Diagnosis)), row.names = T)
+}
+kable(t, row.names = T)
 ```
 
-# Patient Demographics and Sample Characteristics
+## Confusion matrix (proportions)
 
-For those with two sweat electrolyte measures (left and right), the average of the two was used. Each participants' first sweat test value was used for the following tables.
-
-```{r results='asis'}
-t1 <- tableby(
-  Diagnosis ~ Sex + Age.at.Test.in.Years + Method + fe(geno_risk) + fe(geno_class),
-  data = t1_df, control = mycontrols, simulate.p.value = T
+```{r}
+t <- table(km_df$Cluster, km_df$Diagnosis)
+t <- round(prop.table(t, margin = 2) * 100, 1)
+# colorize the diagonal elements in table with non-ints
+for (i in 1:(ncol(t))) {
+  t[i, i] <- cell_spec(t[i, i],
+    bold = T, color = "white", background = "darkgreen"
   )
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
+}
+kable(t, digits = 3)
 ```
 
-# Electrolyte comparisons
+# Classification trees
 
-## By diagnosis
-
-```{r results='asis'}
-t1 <- tableby(
-  Diagnosis ~ Cl + Na + `Na/Cl Ratio` + `Sweat Chloride Group`,
-  data = t1_df, control = mycontrols,
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
+```{r}
+cart_df <- df %>%
+  select(Diagnosis, Na, Cl) %>%
+  drop_na()
 ```
 
-### By genotype severity (CF only)
+## Brief overview
 
-```{r results='asis'}
-t1 <- tableby(
-  geno_risk ~ Cl + Na + `Na/Cl Ratio` + `Sweat Chloride Group`,
-  data = t1_df[t1_df$Diagnosis == "Cystic Fibrosis", ], control = mycontrols,
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
+Detail here
+
+## All 3 categories
+
+```{r}
+set.seed(1017)
+cart_full <- rpart(Diagnosis ~ Na + Cl, data = cart_df, method = "class",
+                   control = rpart.control(cp=0))
+rpart.plot(cart_full)
 ```
 
-### Pairwise comparisons
-
-Pairwise comparisons using Wilcoxon rank sum test with continuity correction
+## Confusion matrix
 
 ```{r}
-pairwise.wilcox.test(t1_df$Cl, t1_df$Diagnosis, p.adjust.method = "fdr")
-pairwise.wilcox.test(t1_df$Na, t1_df$Diagnosis, p.adjust.method = "fdr")
-pairwise.wilcox.test(t1_df$`Na/Cl Ratio`, t1_df$Diagnosis, p.adjust.method = "fdr")
+t <- table(predict(cart_full,type = "class"),cart_df$Diagnosis)
+t <- addmargins(t)
+# colorize the diagonal elements in table with non-ints
+for (i in 1:3) {
+  t[i, i] <- cell_spec(t[i, i],
+    bold = T, color = "white", background = "darkgreen"
+  )
+}
+kable(t)
 ```
 
-## By sweat chloride group
+## Confusion matrix (proportions)
 
-```{r results='asis'}
-t1 <- tableby(
-  `Sweat Chloride Group` ~ fe(geno_risk) + fe(Diagnosis) + Cl + Na + `Na/Cl Ratio`,
-  data = t1_df, control = mycontrols, simulate.p.value = T
+```{r}
+t <- table(predict(cart_full,type = "class"),cart_df$Diagnosis)
+t <- round(prop.table(t, margin = 2) * 100, 1)
+# colorize the diagonal elements in table with non-ints
+for (i in 1:(ncol(t))) {
+  t[i, i] <- cell_spec(t[i, i],
+    bold = T, color = "white", background = "darkgreen"
   )
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
+}
+kable(t, digits = 3)
 ```
 
-## By sweat chloride group and diagnosis
+## CF vs. CFTR-RD or CRMS
 
-### Cystic fibrosis
-
-```{r results='asis'}
-t1 <- tableby(`Sweat Chloride Group` ~ Cl + Na + `Na/Cl Ratio`,
-  data = t1_df[t1_df$Diagnosis == "Cystic Fibrosis", ], control = mycontrols
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
+```{r}
+set.seed(1017)
+cf_crms <- cart_df[cart_df$Diagnosis != "Non-CF", ]
+cf_crms$Diagnosis <- droplevels(cf_crms$Diagnosis)
+cart_cf_crms <- rpart(Diagnosis ~ Na + Cl, data = cf_crms, method = "class")
+rpart.plot(cart_cf_crms)
 ```
 
-### CFTR-RD or CRMS
+## Confusion matrix
 
-```{r results='asis'}
-t1 <- tableby(`Sweat Chloride Group` ~ Cl + Na + `Na/Cl Ratio`,
-  data = t1_df[t1_df$Diagnosis == "CFTR-RD or CRMS", ],
-  control = mycontrols
+```{r}
+t <- table(predict(cart_cf_crms,type = "class"),cf_crms$Diagnosis)
+t <- addmargins(t)
+# colorize the diagonal elements in table with non-ints
+for (i in 1:2) {
+  t[i, i] <- cell_spec(t[i, i],
+    bold = T, color = "white", background = "darkgreen"
   )
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
+}
+kable(t)
 ```
 
-### Non-CF
+## Confusion matrix (proportions)
 
-```{r results='asis'}
-t1 <- tableby(`Sweat Chloride Group` ~ Cl + Na + `Na/Cl Ratio`,
-  data = t1_df[t1_df$Diagnosis == "Non-CF", ], control = mycontrols
+```{r}
+t <- table(predict(cart_cf_crms,type = "class"),cf_crms$Diagnosis)
+t <- round(prop.table(t, margin = 2) * 100, 1)
+# colorize the diagonal elements in table with non-ints
+for (i in 1:(ncol(t))) {
+  t[i, i] <- cell_spec(t[i, i],
+    bold = T, color = "white", background = "darkgreen"
   )
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
+}
+kable(t, digits = 3)
 ```
 
-# Variability
-
-## Within occasion
-
-For all participants with more than one sweat test, we used only their first value to calculate the absolute difference (AD) between left and right arm.
+## Non-CF vs. CFTR-RD or CRMS
 
 ```{r}
-t1_df$`Cl AD` <- abs(t1_df$Cl_Left - t1_df$Cl_Right)
-t1_df$`Na AD` <- abs(t1_df$Na_Left - t1_df$Na_Right)
+set.seed(1017)
+non_cf_crms <- cart_df[cart_df$Diagnosis != "Cystic Fibrosis", ]
+non_cf_crms$Diagnosis <- droplevels(non_cf_crms$Diagnosis)
+cart_non_cf_crms <- rpart(Diagnosis ~ Na + Cl, data = non_cf_crms,
+                          method = "class")
+rpart.plot(cart_non_cf_crms)
 ```
 
-### By diagnosis
+## Confusion matrix
 
-```{r results='asis'}
-t1 <- tableby(
-  Diagnosis ~ `Cl AD` + `Na AD`,
-  data = t1_df, control = mycontrols,
+```{r}
+t <- table(predict(cart_non_cf_crms,type = "class"),non_cf_crms$Diagnosis)
+t <- addmargins(t)
+# colorize the diagonal elements in table with non-ints
+for (i in 1:2) {
+  t[i, i] <- cell_spec(t[i, i],
+    bold = T, color = "white", background = "darkgreen"
   )
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
+}
+kable(t)
 ```
 
-### By sweat chloride group
+## Confusion matrix (proportions)
 
-```{r results='asis'}
-t1 <- tableby(
-  `Sweat Chloride Group` ~ `Cl AD` + `Na AD`,
-  data = t1_df, control = mycontrols
+```{r}
+t<- table(predict(cart_non_cf_crms,type = "class"),non_cf_crms$Diagnosis)
+t <- round(prop.table(t, margin = 2) * 100, 1)
+# colorize the diagonal elements in table with non-ints
+for (i in 1:(ncol(t))) {
+  t[i, i] <- cell_spec(t[i, i],
+    bold = T, color = "white", background = "darkgreen"
   )
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
+}
+kable(t, digits = 3)
 ```
 
-### By genotype risk
+## Future directions
 
-```{r results='asis'}
-t1 <- tableby(
-  geno_risk ~ `Cl AD` + `Na AD`,
-  data = t1_df, control = mycontrols
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
+1. Additional work on the prediction approaches to see if we can improve the results in our current data.
 
-## Between occasion
+  - Support vector machines may allow for more flexeible partitioning of the Na/Cl space.
 
-For all participants with more than one sweat test, we compared their first and last values by calculating the absolute difference (AD) between first and last measure.
+2. Validate final prediction models on new data.
 
-```{r results='asis'}
-# Get first and last values for everyone with more than 1
-multiples <- df %>%
-  group_by(PatientID) %>%
-  filter(n() > 1) %>% # Multiple values only
-  filter(row_number() == 1 | row_number() == n()) %>% # First and last
-  summarise(
-    `Cl AD` = abs(last(Cl) - first(Cl)), # Calculate AD
-    `Na AD` = abs(last(Na) - first(Na)),
-    Diagnosis = first(Diagnosis),
-    `Sweat Chloride Group` = first(`Sweat Chloride Group`),
-    geno_risk = first(geno_risk)
-  )
-```
+# Questions?
 
-### By diagnosis
+## References
 
-```{r results='asis'}
-t1 <- tableby(
-  Diagnosis ~ `Cl AD` + `Na AD`,
-  data = multiples, control = mycontrols,
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
+---
+nocite: |
+  @jamesIntroductionStatisticalLearning2021, @robinPROCOpensourcePackage2011, @therneau1997introduction
+---
 
-### By sweat chloride group
-
-```{r results='asis'}
-t1 <- tableby(
-  `Sweat Chloride Group` ~ `Cl AD` + `Na AD`,
-  data = multiples, control = mycontrols
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-### By genotype risk
-
-```{r results='asis'}
-t1 <- tableby(
-  geno_risk ~ `Cl AD` + `Na AD`,
-  data = multiples, control = mycontrols
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-# Cutoff analyses
-
-We performed receiver operating characteristic (ROC) analysis to find a sweat sodium to sweat chloride cutoff that distinguishes between groups.
-Only the first sweat test from each participant was used.
-
-## Full cohort
-
-The full cohort analyses include all participants.
-
-### Diagnosis
-
-#### ROC
-
-##### Na/Cl Ratio
-
-```{r}
-roc_cf <- roc(
-  response = t1_df$`Diagnosis`,
-  predictor = t1_df$`Na/Cl Ratio`,
-  levels = c("CFTR-RD or CRMS", "Cystic Fibrosis"),
-  quiet = T
-)
-coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)
-
-roc_non_cf <- roc(
-  response = t1_df$`Diagnosis`,
-  predictor = t1_df$`Na/Cl Ratio`,
-  levels = c("Non-CF", "CFTR-RD or CRMS"),
-  quiet = T
-)
-coords_non_cf <- round(coords(roc_non_cf, "best", best.method = "youden"), 3)
-
-plot(roc_cf, print.auc = TRUE, col = "blue")
-plot(roc_non_cf, print.auc = TRUE, col = "red")
-```
-
-The blue curve shows the ROC for distinguishing between CFTR-RD or CRMS vs.
-CF, and the red is the ROC curve for distinguishing between non-CF vs.
-CFTR-RD or CRMS.
-The optimal threshold for distinguishing between CFTR-RD or CRMS vs.
-CF was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index.
-The optimal threshold for distinguishing between non-CF vs.
-CFTR-RD or CRMS was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index.
-The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.
-
-##### Cl Alone
-
-```{r}
-roc_cf_cl <- roc(
-  response = t1_df$`Diagnosis`,
-  predictor = t1_df$Cl,
-  levels = c("CFTR-RD or CRMS", "Cystic Fibrosis"),
-  quiet = T
-)
-coords_cf <- round(coords(roc_cf_cl, "best", best.method = "youden"), 3)
-
-roc_non_cf_cl <- roc(
-  response = t1_df$`Diagnosis`,
-  predictor = t1_df$Cl,
-  levels = c("Non-CF", "CFTR-RD or CRMS"),
-  quiet = T
-)
-coords_non_cf <- round(coords(roc_non_cf_cl, "best", best.method = "youden"), 3)
-
-plot(roc_cf_cl, print.auc = TRUE, col = "blue")
-plot(roc_non_cf_cl, print.auc = TRUE, col = "red")
-```
-
-The blue curve shows the ROC for distinguishing between CFTR-RD or CRMS vs.
-CF, and the red is the ROC curve for distinguishing between non-CF vs.
-CFTR-RD or CRMS.
-The optimal threshold for distinguishing between CFTR-RD or CRMS vs.
-CF was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index.
-The optimal threshold for distinguishing between non-CF vs.
-CFTR-RD or CRMS was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index.
-The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.
-
-##### ROC Comparison
-
-###### CFTR-RD or CRMS vs. CF
-
-```{r}
-roc.test(roc_cf, roc_cf_cl)
-```
-
-###### Non-CF vs. CFTR-RD or CRMS
-
-```{r}
-roc.test(roc_non_cf, roc_non_cf_cl)
-```
-
-#### Histogram by group
-
-Vertical blue line indicates the threshold for CFTR-RD or CRMS vs. CF and red indicates the threshold for non-CF vs. CFTR-RD or CRMS. Na/Cl ratio was log-transformed for easier visualization.
-
-```{r warning=FALSE}
-ggplot(
-  t1_df[!is.na(t1_df$`Diagnosis`), ],
-  aes(log(`Na/Cl Ratio`), fill = `Diagnosis`)
-) +
-  geom_histogram(alpha = 0.3, position = "identity", bins = 50) +
-  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
-  geom_vline(xintercept = log(coords_non_cf$threshold), color = "red") +
-  ylab("Count") +
-  xlab("ln(Na/Cl Ratio)") +
-  theme_bw()
-```
-
-### Genotype
-
-Participants with unknown genotype risk were excluded from this analysis.
-
-#### ROC
-
-```{r}
-# Re-format diagnosis column
-t1_df$`Genotype Risk` <- factor(t1_df$geno_risk,
-  levels = c("Not CF", "Low", "High", "Unknown"),
-  ordered = T
-)
-t1_df$`Genotype Risk`[t1_df$`Genotype Risk` == "Unknown"] <- NA
-```
-
-```{r}
-roc_cf <- roc(
-  response = t1_df$`Genotype Risk`,
-  predictor = t1_df$`Na/Cl Ratio`,
-  levels = c("Low", "High"),
-  quiet = T
-)
-coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)
-
-roc_non_cf <- roc(
-  response = t1_df$`Genotype Risk`,
-  predictor = t1_df$`Na/Cl Ratio`,
-  levels = c("Not CF", "Low"),
-  quiet = T
-)
-coords_non_cf <- round(coords(roc_non_cf, "best", best.method = "youden"), 3)
-
-plot(roc_cf, print.auc = TRUE, col = "blue")
-plot(roc_non_cf, print.auc = TRUE, col = "red", print.auc.y = .4, add = TRUE)
-```
-
-The blue curve shows the ROC for distinguishing between low vs.
-high risk, and the red is the ROC curve for distinguishing between non-CF vs.
-low risk.
-The optimal threshold for distinguishing between low vs.
-high risk was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index.
-The optimal threshold for distinguishing between non-CF vs.
-low risk was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index.
-The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.
-
-#### Histogram by group
-
-Vertical blue line indicates the threshold for low vs. high risk and red indicates the threshold for non-CF vs. low risk. Na/Cl ratio was log-transformed for easier visualization.
-
-```{r warning=FALSE}
-ggplot(
-  t1_df[!is.na(t1_df$`Genotype Risk`), ],
-  aes(log(`Na/Cl Ratio`), fill = `Genotype Risk`)
-) +
-  geom_histogram(alpha = 0.3, position = "identity", bins = 50) +
-  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
-  geom_vline(xintercept = log(coords_non_cf$threshold), color = "red") +
-  ylab("Count") +
-  xlab("ln(Na/Cl Ratio)") +
-  theme_bw()
-```
-
-## Limited cohort
-
-The limited cohort analyses only included participants with sweat chloride < 60.
-
-### Diagnosis
-
-#### ROC
-
-```{r}
-# Exclude those with Cl >= 60
-t1_df <- t1_df %>% filter(Cl < 60)
-```
-
-```{r}
-roc_cf <- roc(
-  response = t1_df$`Diagnosis`,
-  predictor = t1_df$`Na/Cl Ratio`,
-  levels = c("CFTR-RD or CRMS", "Cystic Fibrosis"),
-  quiet = T
-)
-coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)
-
-roc_non_cf <- roc(
-  response = t1_df$`Diagnosis`,
-  predictor = t1_df$`Na/Cl Ratio`,
-  levels = c("Non-CF", "CFTR-RD or CRMS"),
-  quiet = T
-)
-coords_non_cf <- round(coords(roc_non_cf, "best", best.method = "youden"), 3)
-
-plot(roc_cf, print.auc = TRUE, col = "blue")
-plot(roc_non_cf, print.auc = TRUE, col = "red", print.auc.y = .4, add = TRUE)
-```
-
-The blue curve shows the ROC for distinguishing between CFTR-RD or CRMS vs.
-CF, and the red is the ROC curve for distinguishing between non-CF vs.
-CFTR-RD or CRMS.
-The optimal threshold for distinguishing between CFTR-RD or CRMS vs.
-CF was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index.
-The optimal threshold for distinguishing between non-CF vs.
-CFTR-RD or CRMS was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index.
-The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.
-
-#### Histogram by group
-
-Vertical blue line indicates the threshold for CFTR-RD or CRMS vs. CF and red indicates the threshold for non-CF vs. CFTR-RD or CRMS. Na/Cl ratio was log-transformed for easier visualization.
-
-```{r warning=FALSE}
-ggplot(
-  t1_df[!is.na(t1_df$`Diagnosis`), ],
-  aes(log(`Na/Cl Ratio`), fill = `Diagnosis`)
-) +
-  geom_histogram(alpha = 0.3, position = "identity", bins = 50) +
-  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
-  geom_vline(xintercept = log(coords_non_cf$threshold), color = "red") +
-  ylab("Count") +
-  xlab("ln(Na/Cl Ratio)") +
-  theme_bw()
-```
-
-### Genotype
-
-Participants with unknown genotype risk were excluded from this analysis.
-
-#### ROC
-
-```{r}
-roc_cf <- roc(
-  response = t1_df$`Genotype Risk`,
-  predictor = t1_df$`Na/Cl Ratio`,
-  levels = c("Low", "High"),
-  quiet = T
-)
-coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)
-
-roc_non_cf <- roc(
-  response = t1_df$`Genotype Risk`,
-  predictor = t1_df$`Na/Cl Ratio`,
-  levels = c("Not CF", "Low"),
-  quiet = T
-)
-coords_non_cf <- round(coords(roc_non_cf, "best", best.method = "youden"), 3)
-
-plot(roc_cf, print.auc = TRUE, col = "blue")
-plot(roc_non_cf, print.auc = TRUE, col = "red", print.auc.y = .4, add = TRUE)
-```
-
-The blue curve shows the ROC for distinguishing between low vs.
-high risk, and the red is the ROC curve for distinguishing between non-CF vs.
-low risk.
-The optimal threshold for distinguishing between low vs.
-high risk was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index.
-The optimal threshold for distinguishing between non-CF vs.
-low risk was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index.
-The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.
-
-#### Histogram by group
-
-Vertical blue line indicates the threshold for low vs. high risk and red indicates the threshold for non-CF vs. low risk. Na/Cl ratio was log-transformed for easier visualization.
-
-```{r warning=FALSE}
-ggplot(
-  t1_df[!is.na(t1_df$`Genotype Risk`), ],
-  aes(log(`Na/Cl Ratio`), fill = `Genotype Risk`)
-) +
-  geom_histogram(alpha = 0.3, position = "identity", bins = 50) +
-  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
-  geom_vline(xintercept = log(coords_non_cf$threshold), color = "red") +
-  ylab("Count") +
-  xlab("ln(Na/Cl Ratio)") +
-  theme_bw()
-```
+::: {#refs}
+:::

### 2024-03-21 Edith Zemanick/Sweat Test/sweat_test_cf_conference.qmd
method descriptions done

@@
 ---
-title: "Using Sweat Na/Cl Ratio to Predict CFTR-RD or CRMS"
+title: "Using Sweat Na/Cl Ratio to Predict CFTR-RD and CRMS"
 author: "Tim Vigers"
@@
 
-Detail here
+- An approach to splitting up a data set into K non-overlapping groups.
+
+- We want the within-cluster variation to be as low as possible.
+
+- There are several metrics for within-cluster variation, but we usually use the squared Euclidean distance between all pairs of points in a given cluster divided by the number of points in the cluster.
+
+## The K-means clustering algorithm
+
+1. Randomly assign each observation to clusters 1 through K.
+
+2. Iterate until the cluster assignments stop changing:
+
+  a. For each cluster, compute the cluster centroid (a vector of feature means).
+
+  b. Re-assign each point to the cluster corresponding with the closest centroid.
+
+## The K-means clustering algorithm
+
+![](kmeans.jpg)
 
@@
 
-Detail here
+- Essentially, a regression tree is a way of predicting which category an observation will fall into based on a series of data splitting rules.
+
+- We want to split the predictor space (the set of possible values for the variables we're interested in) into $R_1, R_2, ...
+R_J$ non-overlapping regions that minimize some measure of model predictive value (RSS, error rate, etc.).
+
+## Brief overview
+
+- Because in most cases it's impossible to look at every single partition of the predictor space, we use a top-down and greedy approach to splitting up the data.
+
+  - This is a fancy way of saying that we use the data split that gives us the best result at a given step, and don't look at future steps.
+
+## Potential issues
+
+- This kind of recursive splitting can results in trees that predict extremely well with the training data but are not generalizable.
+
+- To avoid this we essentially build one large, overfit tree, and then "prune" it down to its best subtree.
+
+  - Won't go into detail here, but the process is called *cost complexity pruning* and also utilizes cross validation to approximate the test error.
+
+## Baseball example
+
+![](hitters.jpg)
 
@@
 
+# Random forests
+
+## Brief overview
+
 ## Future directions

### 2024-03-25 Edith Zemanick/Sweat Test/sweat_test_cf_conference.qmd
CF research conference background

@@
 bibliography: Presentation.bib
-csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
+csl: /Users/timvigers/GitHub/styles/american-medical-association-10th-edition.csl
 ---
@@
 # Background
+
+## Sweat chloride testing
+
+- Sweat chloride testing is the gold standard for diagnosing Cystic Fibrosis (CF).
+
+| Chloride level (mmol/L) | Diagnosis                |
+|-------------------------|--------------------------|
+| >= 60                   | Cystic Fibrosis          |
+| 30 - 59                 | Intermediate diagnosis   |
+| < 30                    | Cystic Fibrosis unlikely |
+
+## Intermediate sweat test
+
+- Can be caused by:
+  - CF
+  - CF Carrier
+  - CFTR-Related Metabolic Syndrome (CRMS)
+    - Cystic Fibrosis Screen Positive, Inconclusive Diagnosis (CFSPID)
+  - CFTR-related disorder (CFTR-RD)
+
+## CRMS/CFSPID
+
+- Inconclusive CF diagnosis following newborn screen
+- Elevated IRT, intermediate sweat, and less than 2 disease-causing CFTR variants
+- No clinical signs of Cystic Fibrosis
+- Should be followed by CF provider to monitor for symptoms of CF @bombieriRecommendationsClassificationDiseases2011
+  - Close follow-up, repeat sweat testing, and CFTR sequencing in the first months of life 
+  - Then annual follow-up until age 6-7
+
+## CFTR-RD
+
+- Cystic Fibrosis Transmembrane Regulator (CFTR)-related disorders
+  - "A clinical entity associated with CFTR dysfunction that does not fulfill diagnostic criteria for CF" @bombieriRecommendationsClassificationDiseases2011
+  - Has symptoms, but no positive newborn screen, sweat test, and less than 2 disease-causing CFTR variants
+- Should be followed closely by PCP with consultation of a CF provider to monitor for symptoms of CF
+
+## Using sodium and chloride for diagnosis
+
+- Small studies have shown that 
+  - The ratio of sweat sodium to chloride (Na:Cl) is significantly different between pwCF and non-CF patients @hallSweatSodiumChloride1990
+  - Suggested cutoffs are anywhere from 1.0-1.2 @hallSweatSodiumChloride1990 @treggiariRoleSweatIon2021
+- This ratio may be more helpful in diagnosing patients with intermediate sweat chloride values @treggiariRoleSweatIon2021
+- *Can we use both sodium and chloride to better predict CFTR-RD/CRMS?*
 
@@
 nocite: |
-  @jamesIntroductionStatisticalLearning2021, @robinPROCOpensourcePackage2011, @therneau1997introduction
+  @jamesIntroductionStatisticalLearning2021, @robinPROCOpensourcePackage2011, @therneau1997introduction, @metcalf_utility_2024
 ---

### 2024-03-25 Edith Zemanick/Sweat Test/sweat_test_cf_conference.qmd
cf research conference pretty much final draft

@@
 ---
-title: "Using Sweat Na/Cl Ratio to Predict CFTR-RD and CRMS"
+title: "Using Sweat Na and Cl to Predict CFTR-RD and CRMS"
 author: "Tim Vigers"
@@
 library(rpart.plot)
+library(gtsummary)
 source("/Users/timvigers/GitHub/CF/genotype_class_severity.R")
@@
 diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
-exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) | diagnoses$Code == "Exclude"]
+exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) |
+  diagnoses$Code == "Exclude"]
 exclude <- exclude[!is.na(exclude)]
 # Import data
-df <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv", na.strings = "")
+df <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv",
+  na.strings = ""
+)
 n1 <- nrow(df)
@@
 # Filter per Maxie's code
-exclude2 <- c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793)
+exclude2 <- c(
+  22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793
+)
 no_dem <- unique(df$PatientID[df$Demographics.Available != "TRUE" | is.na(df$Demographics.Available) | df$Sex == ""])
@@
 df$geno1_class[grep("508", df$Genotypes1)] <- "F508del"
-df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("I", "II", "III"), "MF")
-df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("IV", "V"), "RF")
+df$geno1_class <- replace(
+  df$geno1_class,
+  df$geno1_class %in% c("I", "II", "III"), "Minimal function"
+)
+df$geno1_class <- replace(
+  df$geno1_class,
+  df$geno1_class %in% c("IV", "V"), "Residual function"
+)
 
@@
 df$geno2_class[grep("508", df$Genotypes2)] <- "F508del"
-df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("I", "II", "III"), "MF")
-df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("IV", "V"), "RF")
+df$geno2_class <- replace(
+  df$geno2_class,
+  df$geno2_class %in% c("I", "II", "III"), "Minimal function"
+)
+df$geno2_class <- replace(
+  df$geno2_class,
+  df$geno2_class %in% c("IV", "V"), "Residual function"
+)
 
@@
 # Which arm?
-df$Arm <- apply(df[, c("Cl", "Cl_Left", "Cl_Right", "Na", "Na_Left", "Na_Right", "K", "K_Left", "K_Right")], 1, function(r) {
+df$Arm <- apply(df[, c(
+  "Cl", "Cl_Left", "Cl_Right", "Na", "Na_Left", "Na_Right", "K", "K_Left",
+  "K_Right"
+)], 1, function(r) {
   k <- sum(!is.na(r))
@@
 # Merge left and right values
-df$Amount <- apply(df[, c("Amount", "Amount_Left", "Amount_Right")], 1, function(r) {
+df$Amount <- apply(
+  df[, c("Amount", "Amount_Left", "Amount_Right")], 1,
+  function(r) {
     amt <- suppressWarnings(as.numeric(as.character(r)))
     mean(amt, na.rm = T)
-})
+  }
+)
 df$Cl <- apply(df[, c("Cl", "Cl_Left", "Cl_Right")], 1, function(r) {
@@
 # Order groups
-df$Diagnosis <- factor(df$Diagnosis, levels = c("Cystic Fibrosis", "CFTR-RD or CRMS", "Non-CF"))
+df$Diagnosis <- factor(df$Diagnosis,
+  levels = c("Cystic Fibrosis", "CFTR-RD or CRMS", "Non-CF")
+)
 ```
@@
 ![](flowchart.png)
+
+## Table 1 {.scrollable}
+
+```{r}
+df %>%
+  filter(!is.na(Na) & !is.na(Cl)) %>%
+  select(Diagnosis, Age.at.Test.in.Weeks, Sex, Na, Cl, K, Method, geno_class) %>%
+  tbl_summary(
+    by = Diagnosis, missing_text = "Missing",
+    label = list(
+      Age.at.Test.in.Weeks = "Age (Weeks)",
+      geno_class = "Genotype Class"
+    )
+  ) %>%
+  add_overall() %>%
+  add_p(test = list(geno_class ~ "fisher.test"), test.args = all_tests("fisher.test") ~ list(simulate.p.value = TRUE)) %>%
+  separate_p_footnotes()
+```
 
@@
 coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)
-plot(roc_cf, print.auc = TRUE, col = "blue",print.thres=T)
+plot(roc_cf, print.auc = TRUE, col = "blue", print.thres = T)
 ```
@@
 coords_non_cf <- round(coords(roc_non_cf, "best", best.method = "youden"), 3)
-plot(roc_non_cf, print.auc = TRUE, col = "red",print.thres=T)
+plot(roc_non_cf, print.auc = TRUE, col = "red", print.thres = T)
 ```
@@
 ![](kmeans.jpg)
+
+## K-means pros and cons
+
+| Pros                                         | Cons                                                        |
+|----------------------------------------------|-------------------------------------------------------------|
+| Easy to implement and understand.            | Must choose k manually.                                     |
+| Scales well to large data sets.              | Has trouble when clusters are of varying sizes and density.                                                            |
+| Easily adapts to new examples.               | Centroids can be dragged by outliers. |
+| Generalizes to clusters of different shapes. | Does not scale well with more dimensions.                        |
 
@@
 
+## Classification tree pros and cons
+
+| Pros                                      | Cons                                                                |
+|-------------------------------------------|---------------------------------------------------------------------|
+| Easy to interpret.                        | Prone to overfitting.                                               |
+| Robust to outliers (except when overfit). | Unstable unless well- regularized.                                  |
+| Non-linear.                               | Greedy algorithm means the tree is not and optimal global solution. |
+| Non-parametric.                           | Can become unwieldy if not pruned correctly.                        |
+
 ## All 3 categories
@@
 set.seed(1017)
-cart_full <- rpart(Diagnosis ~ Na + Cl, data = cart_df, method = "class",
-                   control = rpart.control(cp=0))
-rpart.plot(cart_full)
+cart_full <- rpart(Diagnosis ~ Na + Cl,
+  data = cart_df, method = "class",
+  control = rpart.control(cp = 0)
+)
+rpart.plot(cart_full, type = 3, extra = 8)
 ```
@@
 ```{r}
-t <- table(predict(cart_full,type = "class"),cart_df$Diagnosis)
+t <- table(predict(cart_full, type = "class"), cart_df$Diagnosis)
 t <- addmargins(t)
@@
 ```{r}
-t <- table(predict(cart_full,type = "class"),cart_df$Diagnosis)
+t <- table(predict(cart_full, type = "class"), cart_df$Diagnosis)
 t <- round(prop.table(t, margin = 2) * 100, 1)
@@
 cart_cf_crms <- rpart(Diagnosis ~ Na + Cl, data = cf_crms, method = "class")
-rpart.plot(cart_cf_crms)
+rpart.plot(cart_cf_crms, type = 3, extra = 8)
 ```
@@
 ```{r}
-t <- table(predict(cart_cf_crms,type = "class"),cf_crms$Diagnosis)
+t <- table(predict(cart_cf_crms, type = "class"), cf_crms$Diagnosis)
 t <- addmargins(t)
@@
 ```{r}
-t <- table(predict(cart_cf_crms,type = "class"),cf_crms$Diagnosis)
+t <- table(predict(cart_cf_crms, type = "class"), cf_crms$Diagnosis)
 t <- round(prop.table(t, margin = 2) * 100, 1)
@@
 non_cf_crms$Diagnosis <- droplevels(non_cf_crms$Diagnosis)
-cart_non_cf_crms <- rpart(Diagnosis ~ Na + Cl, data = non_cf_crms,
-                          method = "class")
-rpart.plot(cart_non_cf_crms)
+cart_non_cf_crms <- rpart(Diagnosis ~ Na + Cl,
+  data = non_cf_crms,
+  method = "class"
+)
+rpart.plot(cart_non_cf_crms, type = 1, extra = 8)
 ```
@@
 ```{r}
-t <- table(predict(cart_non_cf_crms,type = "class"),non_cf_crms$Diagnosis)
+t <- table(predict(cart_non_cf_crms, type = "class"), non_cf_crms$Diagnosis)
 t <- addmargins(t)
@@
 ```{r}
-t<- table(predict(cart_non_cf_crms,type = "class"),non_cf_crms$Diagnosis)
+t <- table(predict(cart_non_cf_crms, type = "class"), non_cf_crms$Diagnosis)
 t <- round(prop.table(t, margin = 2) * 100, 1)
@@
 
-# Random forests
+# Support vector machines
 
 ## Brief overview
+
+- Aims to find a hyperplane that best separates our groups.
+  - In a p-dimensional space, a hyperplane is a flat affine subspace of dimension p − 1.
+    - When considering 2 dimensions, this is just a line.
+
+## Hyperplane classifiers
+
+![](hyperplane.png)
+
+## The maximal margin classifier
+
+![](mmc.png)
+
+The points with arrows to our dividing line are the *support vectors*.
+
+## But what if the boundary isn't just a simple line?
+
+![](nlb.png)
+
+## But what if the boundary isn't just a simple line?
+
+- The *support vector machine* (SVM) is an extension of the support vector classifier where the feature space is enlarged using kernels.
+  - This is a little complicated, but is analogous to adding polynomial terms to a regression model.
+    - In fact, SVMs are closely related to logistic regression.
+
+## SVM pros and cons
+
+- One of the best “out of the box” classifiers.
+- Works best when there's fairly clear separation between classes.
+
+## A simple SVM with radial kernel
+
+![](plot_zoom.png)
+
+## Confusion matrix
+
+```{r}
+library(e1071)
+set.seed(1017)
+svmfit <- svm(Diagnosis ~ Na + Cl, data = cart_df)
+```
+
+```{r}
+t <- table(predict(svmfit, type = "class"), cart_df$Diagnosis)
+t <- addmargins(t)
+# colorize the diagonal elements in table with non-ints
+for (i in 1:2) {
+  t[i, i] <- cell_spec(t[i, i],
+    bold = T, color = "white", background = "darkgreen"
+  )
+}
+kable(t)
+```
+
+## Confusion matrix (proportions)
+
+```{r}
+t <- table(predict(svmfit, type = "class"), cart_df$Diagnosis)
+t <- round(prop.table(t, margin = 2) * 100, 1)
+# colorize the diagonal elements in table with non-ints
+for (i in 1:(ncol(t))) {
+  t[i, i] <- cell_spec(t[i, i],
+    bold = T, color = "white", background = "darkgreen"
+  )
+}
+kable(t, digits = 3)
+```
+
+## Comparison of methods
+
+| Method                                           | % Correct CF | % Correct Non-CF | % Correct CFTR-RD or CRMS |
+|--------------------------------------------------|--------------|------------------|---------------------------|
+| K-means                                          | 93.1         | 81               | 78.6                      |
+| Classification tree (all)                        | 98.9         | 99.9             | 21.4                      |
+| Classification tree (CF vs. CFTR-RD or CRMS)     | 99.2         |                  | 92.9                      |
+| Classification tree (Non-CF vs. CFTR-RD or CRMS) | 39.3         | 99.5             |                           |
+| SVM                                              | 98.2         | 99.9             | 0                         |
 
@@
 
-  - Support vector machines may allow for more flexeible partitioning of the Na/Cl space.
+  - Bagging, boosting, and random forests are all extensions of regression trees that minimize overfitting.
 
-2. Validate final prediction models on new data.
+2. Scrap SVMs for now.
 
-# Questions?
+3. Would it be worth considering additional variables for these models? Sex, age at test, etc.?
+
+4. Validate final prediction models on new data.
+
+# Questions, comments, suggestions?
 

### 2024-03-25 Edith Zemanick/Sweat Test/sweat_test_cf_conference.qmd
cf research conference done

@@
 format: revealjs
+embed-resources: true
 editor: source
@@
 
-The points with arrows to our dividing line are the *support vectors*.
-
 ## But what if the boundary isn't just a simple line?
@@
     - In fact, SVMs are closely related to logistic regression.
+
+## But what if the boundary isn't just a simple line?
+
+![](svm.png)
 

### 2024-03-27 Christine Chan/EnVision CF/katie_nacfc_2024.qmd
katie NACFC basic tables done

---
title: "Title"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(gtsummary)
switch(Sys.info()[["sysname"]],
  Windows = {
    home_dir <- "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF"
  },
  Darwin = {
    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
  }
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
# Import
df <- read.csv("/Users/timvigers/Desktop/EnVision/EnvisionCF_DATA_2024-03-14_1325.csv",
  na.strings = ""
)
# Fix missing access group
df$redcap_data_access_group[is.na(df$redcap_data_access_group)] <- "iowa"
# Trim columns
glucose <- paste0("timepoint_", c(0, 10, 30, 60, 90, 120, 150, 180), "_min")
df <- df %>%
  select(
    study_id:redcap_data_access_group, age_at_current_date:origin_race,
    cftr_mutation_1, cftr_mutation_2, bmi_study_visit, bmi_stand_score,
    a1c_result, fev1, all_of(glucose)
  )
# Get diagnosis
df$Diagnosis <- NA
df$Diagnosis[df[, glucose[1]] < 100 &
  (rowSums(df[, glucose[2:5]] < 200, na.rm = T) == rowSums(!is.na(df[, glucose[2:5]]))) &
  (rowSums(df[, glucose[6:8]] < 140, na.rm = T) == rowSums(!is.na(df[, glucose[6:8]])))] <- "NGT"
df$Diagnosis[df[, glucose[1]] >= 100] <- "IFG"
df$Diagnosis[rowSums(df[, glucose[2:5]] >= 200, na.rm = T) > 0] <- "INDET"
df$Diagnosis[rowSums(df[, glucose[6:8]] >= 140, na.rm = T) > 0] <- "IGF"
df$Diagnosis[df$timepoint_0_min >= 126 | df$timepoint_120_min >= 200] <- "CFRD"
df$Diagnosis <- factor(df$Diagnosis, levels = c("NGT", "IFG", "INDET", "IGF", "CFRD"))
# Exclude those missing <= 3 missing OGTT values
df$Keep <- rowSums(is.na(df[, glucose])) <= 3
df <- df %>% filter(Keep)
# Write for checking
write.csv(df, file = "~/df.csv", row.names = F, na = "")
```

# Demographics at first visit

`r length(unique(df$study_id))` participants were included because they were missing <= 3 OGTT timepoints.

```{r}
#| label: tbl-1
#| tbl-cap: Participant characteristics (overall)
df %>%
  group_by(study_id) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(redcap_data_access_group:cftr_mutation_2) %>%
  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
  add_overall()
```

# Diagnoses

## All visits

```{r}
#| label: tbl-2
#| tbl-cap: Participant diagnoses based on OGTT
df %>%
  select(redcap_data_access_group, Diagnosis) %>%
  tbl_summary(by = redcap_data_access_group) %>%
  add_overall()
```

## Visit 1 only

```{r}
#| label: tbl-3
#| tbl-cap: Participant diagnoses based on OGTT (visit 1 only)
df %>%
  group_by(study_id) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(redcap_data_access_group, Diagnosis) %>%
  tbl_summary(by = redcap_data_access_group) %>%
  add_overall()
```

# Figure

```{r}
#| label: fig-1
#| fig-cap: Diagnoses over time
#| fig-width: 12
#| fig-height: 8
df %>%
  group_by(study_id) %>%
  mutate(`Visit Num.` = row_number()) %>%
  group_by(`Visit Num.`) %>%
  count(Diagnosis) %>%
  mutate(total = sum(n), perc = n / total) %>%
  ggplot(., aes(x = `Visit Num.`, y = perc, fill = Diagnosis)) +
  geom_bar(position = "stack", stat = "identity") +
  geom_text(aes(label = paste0(scales::percent(perc), " (n = ", n, ")")),
    position = position_stack(vjust = 0.5)
  ) +
  theme_bw() +
  theme(
    axis.text.y = element_blank(), axis.ticks.y = element_blank(),
    axis.title.y = element_blank()
  )
```

### 2024-03-28 Christine Chan/EnVision CF/amir_nacfc_2024.qmd
katie and amir NACFC updates

---
title: "Title"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(lmerTest)
library(emmeans)
switch(Sys.info()[["sysname"]],
  Windows = {
    home_dir <- "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF"
  },
  Darwin = {
    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
  }
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
# Import data cleaned for Katie's analysis
df <- read.csv("./Christine Chan/EnVision CF/Data_Clean/nacfc_2024_analysis_data.csv",
  na.strings = ""
)
glucose <- paste0("timepoint_", c(0, 10, 30, 60, 90, 120, 150, 180), "_min")
# Remove those missing all OGTT values
long_df <- df %>% select(study_id, redcap_event_name, all_of(glucose), Diagnosis)
long_df <- long_df[rowSums(is.na(long_df)) < length(glucose), ]
# Long OGTTs
long_df <- long_df %>%
  pivot_longer(timepoint_0_min:timepoint_180_min,
    names_to = "Time", values_to = "Glucose",
    names_pattern = "timepoint_(.*)_min"
  )
# Time as a factor
long_df$Time = factor(long_df$Time,levels = c(0, 10, 30, 60, 90, 120, 150, 180))
```

```{r mixed models}
mod <- lmer(Glucose ~ Time * Diagnosis + (1 | study_id), data = long_df)
means <- data.frame(emmeans(mod, ~ Time + Diagnosis, pbkrtest.limit = 1e4))
```

# Figure

```{r}
#| label: fig-mixed-model
#| fig-cap: Model means by timepoint
ggplot(means, aes(x = Time, y = emmean, color = Diagnosis, group = Diagnosis)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),width=0.2) +
  geom_line() +
  theme_classic() +
  ylab("Mean Glucose")
```

### 2024-03-28 Christine Chan/EnVision CF/katie_nacfc_2024.qmd
katie and amir NACFC updates

@@
 ---
-title: "Title"
+title: "Draft abstract for Envision primary outcome"
 author: "Tim Vigers"
@@
 library(gtsummary)
+library(pracma)
 switch(Sys.info()[["sysname"]],
@@
 # Import
-df <- read.csv("/Users/timvigers/Desktop/EnVision/EnvisionCF_DATA_2024-03-14_1325.csv",
+df <- read.csv("./Christine Chan/EnVision CF/Data_Raw/EnvisionCF_DATA_2024-03-14_1325.csv",
   na.strings = ""
@@
     study_id:redcap_data_access_group, age_at_current_date:origin_race,
-    cftr_mutation_1, cftr_mutation_2, bmi_study_visit, bmi_stand_score,
-    a1c_result, fev1, all_of(glucose)
+    cftr_mutation_1, cftr_mutation_2, date_visit, bmi_study_visit,
+    bmi_stand_score, a1c_result, fev1, all_of(glucose)
   )
@@
 df$Diagnosis <- factor(df$Diagnosis, levels = c("NGT", "IFG", "INDET", "IGF", "CFRD"))
+# iAUCs
+df$iAUC30 <- apply(df, 1, function(r) {
+  y <- as.numeric(r[glucose[1:3]]) - as.numeric(r[glucose[1]])
+  if (is.na(last(y))) {
+    return(NA)
+  } else {
+    x <- c(0, 10, 30)[!is.na(y)]
+    y <- y[!is.na(y)]
+    auc <- trapz(x, y)
+    return(auc)
+  }
+})
+df$iAUC120 <- apply(df, 1, function(r) {
+  y <- as.numeric(r[glucose[1:6]]) - as.numeric(r[glucose[1]])
+  if (is.na(last(y))) {
+    return(NA)
+  } else {
+    x <- c(0, 10, 30, 60, 90, 120)[!is.na(y)]
+    y <- y[!is.na(y)]
+    auc <- trapz(x, y)
+    return(auc)
+  }
+})
+df$iAUC180 <- apply(df, 1, function(r) {
+  y <- as.numeric(r[glucose]) - as.numeric(r[glucose[1]])
+  if (is.na(last(y))) {
+    return(NA)
+  } else {
+    x <- c(0, 10, 30, 60, 90, 120, 150, 180)[!is.na(y)]
+    y <- y[!is.na(y)]
+    auc <- trapz(x, y)
+    return(auc)
+  }
+})
+# Other formatting
+df$date_visit <- lubridate::ymd(df$date_visit)
+# Write for checking
+df$Keep <- NULL
+write.csv(df,
+  file = "./Christine Chan/EnVision CF/Data_Clean/nacfc_2024_analysis_data.csv",
+  row.names = F, na = ""
+)
 # Exclude those missing <= 3 missing OGTT values
@@
 df <- df %>% filter(Keep)
-# Write for checking
-write.csv(df, file = "~/df.csv", row.names = F, na = "")
 ```
@@
 
-# Diagnoses
+# Glycemic metrics
+
+iAUC was calculated as the area under the glucose curve from baseline to the timepoint of interest, excluding the area under the fasting glucose level.
 
@@
 df %>%
-  select(redcap_data_access_group, Diagnosis) %>%
-  tbl_summary(by = redcap_data_access_group) %>%
+  select(
+    redcap_data_access_group, Diagnosis, timepoint_0_min, timepoint_30_min,
+    timepoint_60_min, timepoint_120_min, timepoint_180_min, iAUC30, iAUC120,
+    iAUC180
+  ) %>%
+  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
   add_overall()
@@
   ungroup() %>%
-  select(redcap_data_access_group, Diagnosis) %>%
-  tbl_summary(by = redcap_data_access_group) %>%
+  select(
+    redcap_data_access_group, Diagnosis, timepoint_0_min, timepoint_30_min,
+    timepoint_60_min, timepoint_120_min, timepoint_180_min, iAUC30, iAUC120,
+    iAUC180
+  ) %>%
+  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
   add_overall()
@@
 #| label: fig-1
-#| fig-cap: Diagnoses over time
+#| fig-cap: Percentages by visit
 #| fig-width: 12

### 2024-03-30 Christine Chan/EnVision CF/amir_nacfc_2024.qmd
amir NACFC 2024

@@
 library(tidyverse)
+library(gt)
+library(gtsummary)
 library(lmerTest)
@@
 # Remove those missing all OGTT values
-long_df <- df %>% select(study_id, redcap_event_name, all_of(glucose), Diagnosis)
-long_df <- long_df[rowSums(is.na(long_df)) < length(glucose), ]
+df <- df[rowSums(is.na(df[, glucose])) < length(glucose), ]
+# Hypoglycemia
+df$Hypo70 <- factor(rowSums(df[, glucose] < 70, na.rm = T) > 0,
+  levels = c(F, T), labels = c("No", "Yes")
+)
+df$Hypo60 <- factor(rowSums(df[, glucose] < 60, na.rm = T) > 0,
+  levels = c(F, T), labels = c("No", "Yes")
+)
+```
+
+# Hypoglycemia defined as < 70 mg/dL at any point
+
+## Participant demographics
+
+```{r}
+#| label: tbl-70
+#| tbl-cap: BG <70 at Any Point
+df %>%
+  select(
+    age_at_current_date:cftr_mutation_2, all_of(glucose), iAUC30:iAUC180, Hypo70
+  ) %>%
+  tbl_summary(by = Hypo70, missing_text = "Missing") %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes()
+```
+
+## LS mean glucose by hypoglycemia status
+
+```{r mixed models 70}
 # Long OGTTs
-long_df <- long_df %>%
+long_df <- df %>%
   pivot_longer(timepoint_0_min:timepoint_180_min,
@@
 # Time as a factor
-long_df$Time = factor(long_df$Time,levels = c(0, 10, 30, 60, 90, 120, 150, 180))
+long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
+# Fit
+mod <- lmer(Glucose ~ Time * Hypo70 + (1 | study_id), data = long_df)
+means <- data.frame(emmeans(mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4))
+gt(means)
 ```
 
-```{r mixed models}
-mod <- lmer(Glucose ~ Time * Diagnosis + (1 | study_id), data = long_df)
-means <- data.frame(emmeans(mod, ~ Time + Diagnosis, pbkrtest.limit = 1e4))
-```
-
-# Figure
+## Figure
 
 ```{r}
-#| label: fig-mixed-model
+#| label: fig-mixed-model70
 #| fig-cap: Model means by timepoint
-ggplot(means, aes(x = Time, y = emmean, color = Diagnosis, group = Diagnosis)) +
+ggplot(means, aes(x = Time, y = emmean, color = Hypo70, group = Hypo70)) +
   geom_point() +
-  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),width=0.2) +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
   geom_line() +
@@
 ```
+
+# Hypoglycemia defined as < 60 mg/dL at any point
+
+## Participant demographics
+
+```{r}
+#| label: tbl-60
+#| tbl-cap: BG <60 at Any Point
+df %>%
+  select(
+    age_at_current_date:cftr_mutation_2, all_of(glucose), iAUC30:iAUC180, Hypo60
+  ) %>%
+  tbl_summary(by = Hypo60, missing_text = "Missing") %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes()
+```
+
+## LS mean glucose by hypoglycemia status
+
+```{r mixed models 60}
+# Long OGTTs
+long_df <- df %>%
+  pivot_longer(timepoint_0_min:timepoint_180_min,
+    names_to = "Time", values_to = "Glucose",
+    names_pattern = "timepoint_(.*)_min"
+  )
+# Time as a factor
+long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
+# Fit
+mod <- lmer(Glucose ~ Time * Hypo60 + (1 | study_id), data = long_df)
+means <- data.frame(emmeans(mod, ~ Hypo60 | Time, pbkrtest.limit = 1e4))
+gt(means)
+```
+
+## Figure
+
+```{r}
+#| label: fig-mixed-model60
+#| fig-cap: Model means by timepoint
+ggplot(means, aes(x = Time, y = emmean, color = Hypo60, group = Hypo60)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
+  geom_line() +
+  theme_classic() +
+  ylab("Mean Glucose")
+```

### 2024-04-01 Christine Chan/EnVision CF/amir_nacfc_2024.qmd
Update amir_nacfc_2024.qmd

@@
 )
+df$Hypo50 <- factor(rowSums(df[, glucose] <= 50, na.rm = T) > 0,
+  levels = c(F, T), labels = c("No", "Yes")
+)
 ```
@@
 #| label: tbl-70
-#| tbl-cap: BG <70 at Any Point
+#| tbl-cap: BG < 70 at Any Point
 df %>%
@@
 #| label: tbl-60
-#| tbl-cap: BG <60 at Any Point
+#| tbl-cap: BG < 60 at Any Point
 df %>%
@@
 ```
+
+# Hypoglycemia defined as <= 50 mg/dL at any point
+
+## Participant demographics
+
+```{r}
+#| label: tbl-50
+#| tbl-cap: BG <= 50 at Any Point
+df %>%
+  select(
+    age_at_current_date:cftr_mutation_2, all_of(glucose), iAUC30:iAUC180, Hypo50
+  ) %>%
+  tbl_summary(by = Hypo50, missing_text = "Missing") %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes()
+```
+
+## LS mean glucose by hypoglycemia status
+
+```{r mixed models 50}
+# Long OGTTs
+long_df <- df %>%
+  pivot_longer(timepoint_0_min:timepoint_180_min,
+    names_to = "Time", values_to = "Glucose",
+    names_pattern = "timepoint_(.*)_min"
+  )
+# Time as a factor
+long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
+# Fit
+mod <- lmer(Glucose ~ Time * Hypo50 + (1 | study_id), data = long_df)
+means <- data.frame(emmeans(mod, ~ Hypo50 | Time, pbkrtest.limit = 1e4))
+gt(means)
+```
+
+## Figure
+
+```{r}
+#| label: fig-mixed-model50
+#| fig-cap: Model means by timepoint
+ggplot(means, aes(x = Time, y = emmean, color = Hypo50, group = Hypo50)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
+  geom_line() +
+  theme_classic() +
+  ylab("Mean Glucose")
+```

### 2024-04-03 Christine Chan/EnVision CF/amir_nacfc_2024.qmd
Update amir_nacfc_2024.qmd

@@
   select(
-    age_at_current_date:cftr_mutation_2, all_of(glucose), iAUC30:iAUC180, Hypo70
+    age_at_current_date:cftr_mutation_2, all_of(glucose), iAUC30:iAUC180, 
+    Hypo70, 
   ) %>%

### 2024-04-03 Christine Chan/EnVision CF/katie_nacfc_2024.qmd
Update katie_nacfc_2024.qmd

@@
 df$redcap_data_access_group[is.na(df$redcap_data_access_group)] <- "iowa"
-# Trim columns
+# Glucose columns
 glucose <- paste0("timepoint_", c(0, 10, 30, 60, 90, 120, 150, 180), "_min")
-df <- df %>%
-  select(
-    study_id:redcap_data_access_group, age_at_current_date:origin_race,
-    cftr_mutation_1, cftr_mutation_2, date_visit, bmi_study_visit,
-    bmi_stand_score, a1c_result, fev1, all_of(glucose)
-  )
 # Get diagnosis
@@
   ungroup() %>%
-  select(redcap_data_access_group:cftr_mutation_2) %>%
+  select(
+    redcap_data_access_group, age_at_current_date:origin_race, cftr_mutation_1,
+    cftr_mutation_2
+  ) %>%
   tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%

### 2024-04-03 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
Update racial_and_ethnic_differences_in_eti.qmd

@@
 library(performance)
-library(mcp)
 library(survival)
@@
   "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF"
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
 )
@@
 
-Other = Other
+Other = No Insurance (per Liz Cromwell)
 
@@
     "Private", "Public", "Public", "Public", "More Than One", "No Insurance",
-    "Other", "Public", "Tricare"
+    "No Insurance", "Public", "Tricare"
   )
 )
+# FEV1 to categorical
+surv_dat$fev1_cat <- cut(surv_dat$gli_fev1_ppred_rn,
+  breaks = c(0, 40, 80, 100, Inf),right = F
+)
+label(surv_dat$fev1_cat)=label(surv_dat$gli_fev1_ppred_rn)
+# New fit
 new_surv_f <- update(surv_f, . ~ . + Insurance)

### 2024-04-03 Quarto Template.qmd
Update Quarto Template.qmd

@@
       dark: darkly
-bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
+bibliography: ~/Documents/Miscellaneous/zotero.bib
+csl: ~/Documents/Miscellaneous/american-medical-association.csl
 editor: source
@@
 library(tidyverse)
-switch(Sys.info()[["sysname"]],
-  Windows = {
-    home_dir <- "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF"
-  },
-  Darwin = {
-    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
-  }
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
 )

### 2024-04-03 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
Update racial_and_ethnic_differences_in_eti.qmd

@@
 
+### 2020 - 2022
+
+The table below includes everyone who was in the registry in 2020, 2021, or 2022. 
+People were considered eligible for ETI if they were eligible any of the three years, and were considered prescribed ETI if they had Vx445comb="Yes" for any year.
+So, a person who was eligible in 2020 but not prescribed until 2022 would count the same as someone who was eligible in 2020 and started immediately.
+
+```{r}
+annual %>%
+  filter(ReviewYear %in% 2020:2022) %>%
+  group_by(eDWID) %>%
+  summarise(race = race[1],eti_elig = any(eti_elig == "Yes"),
+            on_eti = any(Vx445comb=="Yes")) %>%
+  select(-eDWID) %>%
+  tbl_summary(by=race) %>%
+  add_overall() %>%
+  add_ci()
+```
+
 ### Figure
@@
   theme_bw(base_size = 16) +
-  theme(legend.title = element_blank(), legend.text = element_text(size = 14))
+  theme(legend.title = element_blank(), legend.text = element_text(size = 14),
+        legend.position = c(0.7, 0.3))
 p
@@
 surv_dat$fev1_cat <- cut(surv_dat$gli_fev1_ppred_rn,
-  breaks = c(0, 40, 80, 100, Inf),right = F
+  breaks = c(0, 40, 80, 100, Inf), right = F
 )
-label(surv_dat$fev1_cat)=label(surv_dat$gli_fev1_ppred_rn)
+label(surv_dat$fev1_cat) <- label(surv_dat$gli_fev1_ppred_rn)
 # New fit
@@
 new_surv_f <- update(surv_f, . ~ . + Insurance + `Age Group` +
-  gli_fev1_ppred_rn)
+  fev1_cat)
 cox_res <- coxph(new_surv_f, data = surv_dat, cluster = eDWID)
@@
 ggcoef_model(cox_res,
-  show_p_values = F, add_reference_rows = F,
+  show_p_values = F, add_reference_rows = T,
   signif_stars = F, colour = F, exponentiate = T
@@
   select(-eDWID) %>%
-  tbl_summary(by = race)
+  tbl_summary(by = race) %>%
+  add_p()
 ```

### 2024-04-04 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
Update phase_1_analysis.qmd

@@
       dark: darkly
-bibliography: /Users/timvigers/Documents/Notes/zotero.bib
-csl: /Users/timvigers/GitHub/styles/american-medical-association-10th-edition.csl
 editor: source
@@
 df <- full_join(db_cos_collection_results, db_sos_collection_results)
+# Add dates to demographics
+db_demographics <- left_join(db_demographics,
+  df %>% select(record_id, cos_result_date),
+  by = join_by(record_id)
+)
+db_demographics$age_at_cos <- difftime(db_demographics$cos_result_date,
+  db_demographics$dob,
+  units = "days"
+) / 365.25
 # Format
@@
 db_demographics %>%
-  select(group, sex, Race, Ethnicity) %>%
-  tbl_summary(by = group, label = list("sex" = "Sex")) %>%
+  select(group, age_at_cos, sex, Race, Ethnicity) %>%
+  tbl_summary(
+    by = group,
+    statistic = all_continuous()~c("{median} ({min}, {max})"),
+    label = list("sex" = "Sex", "age_at_cos" = "Age at COS")
+  ) %>%
   add_overall() %>%

### 2024-04-05 Christine Chan/EnVision CF/amir_nacfc_2024.qmd
Update amir_nacfc_2024.qmd

--- 2024-03-30 Christine Chan/EnVision CF/amir_nacfc_2024.qmd
@@
 )
+df$Hypo50 <- factor(rowSums(df[, glucose] <= 50, na.rm = T) > 0,
+  levels = c(F, T), labels = c("No", "Yes")
+)
+# Add insulin data
+insulin <- read.csv("./Christine Chan/EnVision CF/Data_Clean/insulin_wide.csv")
+df <- left_join(df, insulin, by = join_by(study_id, date_visit))
+insulin <- gsub("timepoint_", "insulin_", glucose)
 ```
@@
 #| label: tbl-70
-#| tbl-cap: BG <70 at Any Point
+#| tbl-cap: BG < 70 at Any Point
 df %>%
   select(
-    age_at_current_date:cftr_mutation_2, all_of(glucose), iAUC30:iAUC180, Hypo70
+    age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2,
+    all_of(glucose), iAUC30:iAUC180,
+    Hypo70, all_of(insulin), hypoglycemia_symptoms_yesno
   ) %>%
@@
 means <- data.frame(emmeans(mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4))
-gt(means)
+# Long insulins
+insulin_long_df <- df %>%
+  pivot_longer(all_of(insulin),
+    names_to = "Time", values_to = "Insulin",
+    names_pattern = "insulin_(.*)_min"
+  )
+# Time as a factor
+insulin_long_df$Time <- factor(insulin_long_df$Time,
+  levels = c(0, 10, 30, 60, 90, 120, 150, 180)
+)
+# Fit
+ins_mod <- lmer(Insulin ~ Time * Hypo70 + (1 | study_id), data = insulin_long_df)
+ins_means <- data.frame(emmeans(ins_mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4))
+# Combine
+means <- means %>% select(-SE,-df) 
+means$Variable = "Glucose"
+ins_means <- ins_means %>% select(-SE,-df)
+ins_means$Variable = "Insulin"
+plot_df = data.frame(rbind(means,ins_means))
+levels(plot_df$Hypo70) = c("No Hypoglycemia","Hypoglycemia")
 ```
-
-## Figure
 
@@
 #| fig-cap: Model means by timepoint
-ggplot(means, aes(x = Time, y = emmean, color = Hypo70, group = Hypo70)) +
+ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) +
   geom_point() +
@@
   geom_line() +
-  theme_classic() +
-  ylab("Mean Glucose")
+  ylab("EM Mean") +
+  theme_bw() +
+  facet_wrap(~Hypo70)
 ```
@@
 #| label: tbl-60
-#| tbl-cap: BG <60 at Any Point
+#| tbl-cap: BG < 60 at Any Point
 df %>%
   select(
-    age_at_current_date:cftr_mutation_2, all_of(glucose), iAUC30:iAUC180, Hypo60
+    age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2,
+    all_of(glucose), iAUC30:iAUC180,
+    Hypo60, all_of(insulin), hypoglycemia_symptoms_yesno
   ) %>%
@@
 means <- data.frame(emmeans(mod, ~ Hypo60 | Time, pbkrtest.limit = 1e4))
-gt(means)
+# Long insulins
+insulin_long_df <- df %>%
+  pivot_longer(all_of(insulin),
+    names_to = "Time", values_to = "Insulin",
+    names_pattern = "insulin_(.*)_min"
+  )
+# Time as a factor
+insulin_long_df$Time <- factor(insulin_long_df$Time,
+  levels = c(0, 10, 30, 60, 90, 120, 150, 180)
+)
+# Fit
+ins_mod <- lmer(Insulin ~ Time * Hypo60 + (1 | study_id), data = insulin_long_df)
+ins_means <- data.frame(emmeans(ins_mod, ~ Hypo60 | Time, pbkrtest.limit = 1e4))
+# Combine
+means <- means %>% select(-SE,-df) 
+means$Variable = "Glucose"
+ins_means <- ins_means %>% select(-SE,-df)
+ins_means$Variable = "Insulin"
+plot_df = data.frame(rbind(means,ins_means))
+levels(plot_df$Hypo60) = c("No Hypoglycemia","Hypoglycemia")
 ```
-
-## Figure
 
@@
 #| fig-cap: Model means by timepoint
-ggplot(means, aes(x = Time, y = emmean, color = Hypo60, group = Hypo60)) +
+ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) +
   geom_point() +
@@
   geom_line() +
-  theme_classic() +
-  ylab("Mean Glucose")
+  ylab("EM Mean") +
+  theme_bw() +
+  facet_wrap(~Hypo60)
 ```
+
+# Hypoglycemia defined as <= 50 mg/dL at any point
+
+## Participant demographics
+
+```{r}
+#| label: tbl-50
+#| tbl-cap: BG <= 50 at Any Point
+df %>%
+  select(
+    age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2,
+    all_of(glucose), iAUC30:iAUC180,
+    Hypo50, all_of(insulin), hypoglycemia_symptoms_yesno
+  ) %>%
+  tbl_summary(by = Hypo50, missing_text = "Missing") %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes()
+```
+
+## LS mean glucose by hypoglycemia status
+
+```{r mixed models 50}
+# Long OGTTs
+long_df <- df %>%
+  pivot_longer(timepoint_0_min:timepoint_180_min,
+    names_to = "Time", values_to = "Glucose",
+    names_pattern = "timepoint_(.*)_min"
+  )
+# Time as a factor
+long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
+# Fit
+mod <- lmer(Glucose ~ Time * Hypo50 + (1 | study_id), data = long_df)
+means <- data.frame(emmeans(mod, ~ Hypo50 | Time, pbkrtest.limit = 1e4))
+# Long insulins
+insulin_long_df <- df %>%
+  pivot_longer(all_of(insulin),
+    names_to = "Time", values_to = "Insulin",
+    names_pattern = "insulin_(.*)_min"
+  )
+# Time as a factor
+insulin_long_df$Time <- factor(insulin_long_df$Time,
+  levels = c(0, 10, 30, 60, 90, 120, 150, 180)
+)
+# Fit
+ins_mod <- lmer(Insulin ~ Time * Hypo50 + (1 | study_id), data = insulin_long_df)
+ins_means <- data.frame(emmeans(ins_mod, ~ Hypo50 | Time, pbkrtest.limit = 1e4))
+# Combine
+means <- means %>% select(-SE,-df) 
+means$Variable = "Glucose"
+ins_means <- ins_means %>% select(-SE,-df)
+ins_means$Variable = "Insulin"
+plot_df = data.frame(rbind(means,ins_means))
+levels(plot_df$Hypo50) = c("No Hypoglycemia","Hypoglycemia")
+```
+
+```{r}
+#| label: fig-mixed-model50
+#| fig-cap: Model means by timepoint
+ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
+  geom_line() +
+  ylab("EM Mean") +
+  theme_bw() +
+  facet_wrap(~Hypo50)
+```

### 2024-04-05 Emily Muther/body_image_trikafta.qmd
Update body_image_trikafta.qmd

--- 2023-05-19 Emily Muther/body_image_trikafta.qmd
@@
 library(sjPlot)
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+)
+knitr::opts_knit$set(root.dir = home_dir)
+```
+
+```{r}
 # Read in data
-df = read.spss("/Users/timvigers/Documents/Work/Vigers/CF/Emily Muther/Body Image and Trikafta/Data_Raw/TrikaftaBodyImage3.6.23.sav",to.data.frame = T)
+df = read.spss("./Emily Muther/General - Muther Research Group/BodyImage_Trikafta/Data/TrikaftaBodyImage3.6.23.sav",to.data.frame = T)
 # Filter
@@
 
-`r report(mod)`
+`r report(mod1)`
 
@@
 
-`r report(mod)`
+`r report(mod2)`
 
@@
 
-`r report(mod)`
+`r report(mod3)`
 
@@
 
-`r report(mod)`
+`r report(mod4)`
 
@@
 
-`r report(mod)`
+`r report(mod5)`
 
@@
 
-`r report(mod)`
+`r report(mod6)`
 

### 2024-04-05 Christine Chan/EnVision CF/amir_nacfc_2024.qmd
Update amir_nacfc_2024.qmd

--- 2024-04-01 Christine Chan/EnVision CF/amir_nacfc_2024.qmd
@@
 glucose <- paste0("timepoint_", c(0, 10, 30, 60, 90, 120, 150, 180), "_min")
+insulin <- gsub("timepoint_", "insulin_", glucose)
 # Remove those missing all OGTT values
 df <- df[rowSums(is.na(df[, glucose])) < length(glucose), ]
-# Hypoglycemia
-df$Hypo70 <- factor(rowSums(df[, glucose] < 70, na.rm = T) > 0,
-  levels = c(F, T), labels = c("No", "Yes")
-)
-df$Hypo60 <- factor(rowSums(df[, glucose] < 60, na.rm = T) > 0,
-  levels = c(F, T), labels = c("No", "Yes")
-)
-df$Hypo50 <- factor(rowSums(df[, glucose] <= 50, na.rm = T) > 0,
-  levels = c(F, T), labels = c("No", "Yes")
-)
+```
+
+# Demographics at first visit
+
+```{r}
+#| label: tbl-1
+#| tbl-cap: Participant characteristics (overall)
+df %>%
+  group_by(study_id) %>%
+  filter(row_number() == 1) %>%
+  ungroup() %>%
+  select(
+    redcap_data_access_group, age_at_current_date:origin_race, cftr_mutation_1,
+    cftr_mutation_2
+  ) %>%
+  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
+  add_overall()
 ```
@@
   select(
-    age_at_current_date:cftr_mutation_2, all_of(glucose), iAUC30:iAUC180, Hypo70
+    age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2,
+    all_of(glucose), Hypo70, all_of(insulin), hypoglycemia_symptoms_yesno,
+    iAUC30gluc:iAUC180ins, homa_ir
   ) %>%
@@
 means <- data.frame(emmeans(mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4))
-gt(means)
+# Long insulins
+insulin_long_df <- df %>%
+  pivot_longer(all_of(insulin),
+    names_to = "Time", values_to = "Insulin",
+    names_pattern = "insulin_(.*)_min"
+  )
+# Time as a factor
+insulin_long_df$Time <- factor(insulin_long_df$Time,
+  levels = c(0, 10, 30, 60, 90, 120, 150, 180)
+)
+# Fit
+ins_mod <- lmer(Insulin ~ Time * Hypo70 + (1 | study_id), data = insulin_long_df)
+ins_means <- data.frame(emmeans(ins_mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4))
+# Combine
+means <- means %>% select(-SE, -df)
+means$Variable <- "Glucose"
+ins_means <- ins_means %>% select(-SE, -df)
+ins_means$Variable <- "Insulin"
+plot_df <- data.frame(rbind(means, ins_means))
+levels(plot_df$Hypo70) <- c("No Hypoglycemia", "Hypoglycemia")
 ```
-
-## Figure
 
@@
   select(
-    age_at_current_date:cftr_mutation_2, all_of(glucose), iAUC30:iAUC180, Hypo60
+    age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2,
+    all_of(glucose), Hypo60, all_of(insulin), hypoglycemia_symptoms_yesno,
+    iAUC30gluc:iAUC180ins, homa_ir
   ) %>%
@@
 means <- data.frame(emmeans(mod, ~ Hypo60 | Time, pbkrtest.limit = 1e4))
-gt(means)
+# Long insulins
+insulin_long_df <- df %>%
+  pivot_longer(all_of(insulin),
+    names_to = "Time", values_to = "Insulin",
+    names_pattern = "insulin_(.*)_min"
+  )
+# Time as a factor
+insulin_long_df$Time <- factor(insulin_long_df$Time,
+  levels = c(0, 10, 30, 60, 90, 120, 150, 180)
+)
+# Fit
+ins_mod <- lmer(Insulin ~ Time * Hypo60 + (1 | study_id), data = insulin_long_df)
+ins_means <- data.frame(emmeans(ins_mod, ~ Hypo60 | Time, pbkrtest.limit = 1e4))
+# Combine
+means <- means %>% select(-SE, -df)
+means$Variable <- "Glucose"
+ins_means <- ins_means %>% select(-SE, -df)
+ins_means$Variable <- "Insulin"
+plot_df <- data.frame(rbind(means, ins_means))
+levels(plot_df$Hypo60) <- c("No Hypoglycemia", "Hypoglycemia")
 ```
-
-## Figure
 
@@
   geom_line() +
-  theme_classic() +
-  ylab("Mean Glucose")
+  ylab("EM Mean") +
+  theme_bw() +
+  facet_wrap(~Hypo60)
 ```
@@
   select(
-    age_at_current_date:cftr_mutation_2, all_of(glucose), iAUC30:iAUC180, Hypo50
+    age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2,
+    all_of(glucose), Hypo50, all_of(insulin), hypoglycemia_symptoms_yesno,
+    iAUC30gluc:iAUC180ins, homa_ir
   ) %>%
@@
 means <- data.frame(emmeans(mod, ~ Hypo50 | Time, pbkrtest.limit = 1e4))
-gt(means)
+# Long insulins
+insulin_long_df <- df %>%
+  pivot_longer(all_of(insulin),
+    names_to = "Time", values_to = "Insulin",
+    names_pattern = "insulin_(.*)_min"
+  )
+# Time as a factor
+insulin_long_df$Time <- factor(insulin_long_df$Time,
+  levels = c(0, 10, 30, 60, 90, 120, 150, 180)
+)
+# Fit
+ins_mod <- lmer(Insulin ~ Time * Hypo50 + (1 | study_id), data = insulin_long_df)
+ins_means <- data.frame(emmeans(ins_mod, ~ Hypo50 | Time, pbkrtest.limit = 1e4))
+# Combine
+means <- means %>% select(-SE, -df)
+means$Variable <- "Glucose"
+ins_means <- ins_means %>% select(-SE, -df)
+ins_means$Variable <- "Insulin"
+plot_df <- data.frame(rbind(means, ins_means))
+levels(plot_df$Hypo50) <- c("No Hypoglycemia", "Hypoglycemia")
 ```
-
-## Figure
 
@@
 #| fig-cap: Model means by timepoint
-ggplot(means, aes(x = Time, y = emmean, color = Hypo50, group = Hypo50)) +
+ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) +
   geom_point() +
@@
   geom_line() +
-  theme_classic() +
-  ylab("Mean Glucose")
+  ylab("EM Mean") +
+  theme_bw() +
+  facet_wrap(~Hypo50)
 ```

### 2024-04-05 Christine Chan/EnVision CF/katie_nacfc_2024.qmd
Update katie_nacfc_2024.qmd

@@
 df$Diagnosis <- factor(df$Diagnosis, levels = c("NGT", "IFG", "INDET", "IGF", "CFRD"))
+# Add insulin data
+insulin <- read.csv("./Christine Chan/EnVision CF/Data_Clean/insulin_wide.csv")
+df <- left_join(df, insulin, by = join_by(study_id, date_visit))
+insulin <- gsub("timepoint_", "insulin_", glucose)
 # iAUCs
-df$iAUC30 <- apply(df, 1, function(r) {
+df$iAUC30gluc <- apply(df, 1, function(r) {
   y <- as.numeric(r[glucose[1:3]]) - as.numeric(r[glucose[1]])
@@
 })
-df$iAUC120 <- apply(df, 1, function(r) {
+df$iAUC120gluc <- apply(df, 1, function(r) {
   y <- as.numeric(r[glucose[1:6]]) - as.numeric(r[glucose[1]])
@@
 })
-df$iAUC180 <- apply(df, 1, function(r) {
+df$iAUC180gluc <- apply(df, 1, function(r) {
   y <- as.numeric(r[glucose]) - as.numeric(r[glucose[1]])
@@
 })
-# Other formatting
-df$date_visit <- lubridate::ymd(df$date_visit)
-# Write for checking
-df$Keep <- NULL
+df$iAUC30ins <- apply(df, 1, function(r) {
+  y <- as.numeric(r[insulin[1:3]]) - as.numeric(r[insulin[1]])
+  if (is.na(last(y))) {
+    return(NA)
+  } else {
+    x <- c(0, 10, 30)[!is.na(y)]
+    y <- y[!is.na(y)]
+    auc <- trapz(x, y)
+    return(auc)
+  }
+})
+df$iAUC120ins <- apply(df, 1, function(r) {
+  y <- as.numeric(r[insulin[1:6]]) - as.numeric(r[insulin[1]])
+  if (is.na(last(y))) {
+    return(NA)
+  } else {
+    x <- c(0, 10, 30, 60, 90, 120)[!is.na(y)]
+    y <- y[!is.na(y)]
+    auc <- trapz(x, y)
+    return(auc)
+  }
+})
+df$iAUC180ins <- apply(df, 1, function(r) {
+  y <- as.numeric(r[insulin]) - as.numeric(r[insulin[1]])
+  if (is.na(last(y))) {
+    return(NA)
+  } else {
+    x <- c(0, 10, 30, 60, 90, 120, 150, 180)[!is.na(y)]
+    y <- y[!is.na(y)]
+    auc <- trapz(x, y)
+    return(auc)
+  }
+})
+# Hypoglycemia
+df$Hypo70 <- factor(rowSums(df[, glucose] < 70, na.rm = T) > 0,
+  levels = c(F, T), labels = c("No", "Yes")
+)
+df$Hypo60 <- factor(rowSums(df[, glucose] < 60, na.rm = T) > 0,
+  levels = c(F, T), labels = c("No", "Yes")
+)
+df$Hypo50 <- factor(rowSums(df[, glucose] <= 50, na.rm = T) > 0,
+  levels = c(F, T), labels = c("No", "Yes")
+)
+# HOMA IR
+df$homa_ir = (df$timepoint_0_min * df$insulin_0_min)/405
+# Write for Amir's abstract
 write.csv(df,
@@
 # Exclude those missing <= 3 missing OGTT values
+df$Keep <- NULL
 df$Keep <- rowSums(is.na(df[, glucose])) <= 3
@@
     redcap_data_access_group, Diagnosis, timepoint_0_min, timepoint_30_min,
-    timepoint_60_min, timepoint_120_min, timepoint_180_min, iAUC30, iAUC120,
-    iAUC180
+    timepoint_60_min, timepoint_120_min, timepoint_180_min, 
+    iAUC30gluc:iAUC180ins,homa_ir
   ) %>%

### 2024-04-05 Emily Muther/body_image_trikafta.qmd
Update body_image_trikafta.qmd

--- 2023-05-19 Emily Muther/body_image_trikafta.qmd
@@
 library(sjPlot)
+library(ggstats)
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+)
+knitr::opts_knit$set(root.dir = home_dir)
+```
+
+```{r}
+#| warning: false
 # Read in data
-df = read.spss("/Users/timvigers/Documents/Work/Vigers/CF/Emily Muther/Body Image and Trikafta/Data_Raw/TrikaftaBodyImage3.6.23.sav",to.data.frame = T)
+df <- read.spss("./Emily Muther/General - Muther Research Group/BodyImage_Trikafta/Data/TrikaftaBodyImage3.6.23.sav", to.data.frame = T)
 # Filter
-df = df %>% 
-  filter(TakingTrikafta=="yes")
+df <- df %>%
+  filter(TakingTrikafta == "yes")
 # Trim white space
-df = df %>%
-  mutate(across(where(is.character),trimws))
+df <- df %>%
+  mutate(across(where(is.character), trimws))
 # Replace blanks with NA
-df[df==""]=NA
+df[df == ""] <- NA
 # Categorical variables
-df$FRS_Category = relevel(df$FRS_Category,ref = "neutral")
+df$FRS_Category <- relevel(df$FRS_Category, ref = "neutral")
 ```
@@
 #| label: demographics
-df %>% select(Age,Gender,Race,Ethnicity) %>%
+df %>%
+  select(Age, Gender, Race, Ethnicity) %>%
   tbl_summary(
@@
 #| label: changes
-df %>% select(ChangestoBreathing:MoodChangesSideEf,FeelTrikafta) %>%
+df %>%
+  select(ChangestoBreathing:MoodChangesSideEf, FeelTrikafta) %>%
   tbl_summary(
@@
 ```{r}
-mod = glm(PHQ9_elevated ~ FRS_Category,data = df,family = "binomial")
-res = tidy(mod,exponentiate = T)
-kable(res,digits = 3)
+mod <- glm(PHQ9_elevated ~ FRS_Category, data = df, family = "binomial")
+tbl_regression(mod)
 ```
-
-`r report(mod)`
 
@@
 ```{r}
-mod = glm(PHQ9_Total~ FRS_Category,data = df[df$PHQ9_elevated=="yes",],family = "poisson")
-res = tidy(mod,exponentiate = T)
-kable(res,digits = 3)
+mod2 <- glm(PHQ9_Total ~ FRS_Category,
+  data = df[df$PHQ9_elevated == "yes", ],
+  family = "poisson"
+)
+tbl_regression(mod2)
 ```
-
-`r report(mod)`
 
@@
 ```{r}
-mod = glm(GAD7_elevated ~ FRS_Category,data = df,family = "binomial")
-res = tidy(mod,exponentiate = T)
-kable(res,digits = 3)
+mod3 <- glm(GAD7_elevated ~ FRS_Category, data = df, family = "binomial")
+tbl_regression(mod3)
 ```
-
-`r report(mod)`
 
@@
 ```{r}
-mod = glm(GAD7_Total~ FRS_Category,data = df[df$GAD7_elevated=="yes",],family = "poisson")
-res = tidy(mod,exponentiate = T)
-kable(res,digits = 3)
+mod4 <- glm(GAD7_Total ~ FRS_Category, data = df[df$GAD7_elevated == "yes", ], family = "poisson")
+tbl_regression(mod4)
 ```
-
-`r report(mod)`
 
@@
 ```{r}
-mod1 = glm(PHQ9_elevated ~ BE_APPEARANCE_MEAN,data = df,family = "binomial")
-res = tidy(mod1,exponentiate = T)
-kable(res,digits = 3)
+mod5 <- glm(PHQ9_elevated ~ BE_APPEARANCE_MEAN, data = df, family = "binomial")
+tbl_regression(mod5)
 ```
-
-`r report(mod)`
 
@@
 ```{r}
-mod2 = glm(PHQ9_Total~ BE_APPEARANCE_MEAN,data = df[df$PHQ9_elevated=="yes",],family = "poisson")
-res = tidy(mod2,exponentiate = T)
-kable(res,digits = 3)
+mod6 <- glm(PHQ9_Total ~ BE_APPEARANCE_MEAN,
+  data = df[df$PHQ9_elevated == "yes", ],
+  family = "poisson"
+)
+tbl_regression(mod6)
 ```
-
-`r report(mod)`
 
@@
 ```{r}
-mod3 = glm(PHQ9_elevated ~ BE_ATTRIBUTION_MEAN,data = df,family = "binomial")
-res = tidy(mod3,exponentiate = T)
-kable(res,digits = 3)
+mod7 <- glm(PHQ9_elevated ~ BE_ATTRIBUTION_MEAN, data = df, family = "binomial")
+tbl_regression(mod7)
 ```
-
-`r report(mod)`
 
@@
 ```{r}
-mod4 = glm(PHQ9_Total~ BE_ATTRIBUTION_MEAN,data = df[df$PHQ9_elevated=="yes",],family = "poisson")
-res = tidy(mod4,exponentiate = T)
-kable(res,digits = 3)
+mod8 <- glm(PHQ9_Total ~ BE_ATTRIBUTION_MEAN,
+  data = df[df$PHQ9_elevated == "yes", ],
+  family = "poisson"
+)
+tbl_regression(mod8)
 ```
-
-`r report(mod)`
 
@@
 ```{r}
-mod5 = glm(PHQ9_elevated ~ BE_WEIGHT_MEAN,data = df,family = "binomial")
-res = tidy(mod5,exponentiate = T)
-kable(res,digits = 3)
+mod9 <- glm(PHQ9_elevated ~ BE_WEIGHT_MEAN, data = df, family = "binomial")
+tbl_regression(mod9)
 ```
-
-`r report(mod)`
 
@@
 ```{r}
-mod6 = glm(PHQ9_Total~ BE_WEIGHT_MEAN,data = df[df$PHQ9_elevated=="yes",],family = "poisson")
-res = tidy(mod6,exponentiate = T)
-kable(res,digits = 3)
+mod10 <- glm(PHQ9_Total ~ BE_WEIGHT_MEAN,
+  data = df[df$PHQ9_elevated == "yes", ],
+  family = "poisson"
+)
+tbl_regression(mod10)
 ```
 
-`r report(mod)`
-
 ### Forest plot of logistic regression results
-
-```{r}
-plot_models(mod1,mod3,mod5) + theme_bw() + theme(legend.position = "none")
-```
 
@@
 ```{r}
-mod1 = glm(GAD7_elevated ~ BE_APPEARANCE_MEAN,data = df,family = "binomial")
-res = tidy(mod1,exponentiate = T)
-kable(res,digits = 3)
+mod11 <- glm(GAD7_elevated ~ BE_APPEARANCE_MEAN, data = df, family = "binomial")
+tbl_regression(mod11)
 ```
-
-`r report(mod)`
 
@@
 ```{r}
-mod2 = glm(GAD7_Total~ BE_APPEARANCE_MEAN,data = df[df$GAD7_elevated=="yes",],family = "poisson")
-res = tidy(mod2,exponentiate = T)
-kable(res,digits = 3)
+mod12 <- glm(GAD7_Total ~ BE_APPEARANCE_MEAN, data = df[df$GAD7_elevated == "yes", ], family = "poisson")
+tbl_regression(mod12)
 ```
-
-`r report(mod)`
 
@@
 ```{r}
-mod3 = glm(GAD7_elevated ~ BE_ATTRIBUTION_MEAN,data = df,family = "binomial")
-res = tidy(mod3,exponentiate = T)
-kable(res,digits = 3)
+mod13 <- glm(GAD7_elevated ~ BE_ATTRIBUTION_MEAN, data = df, family = "binomial")
+tbl_regression(mod13)
 ```
-
-`r report(mod)`
 
@@
 ```{r}
-mod4 = glm(GAD7_Total~ BE_ATTRIBUTION_MEAN,data = df[df$GAD7_elevated=="yes",],family = "poisson")
-res = tidy(mod4,exponentiate = T)
-kable(res,digits = 3)
+mod14 <- glm(GAD7_Total ~ BE_ATTRIBUTION_MEAN,
+  data = df[df$GAD7_elevated == "yes", ],
+  family = "poisson"
+)
+tbl_regression(mod14)
 ```
-
-`r report(mod)`
 
@@
 ```{r}
-mod5 = glm(GAD7_elevated ~ BE_WEIGHT_MEAN,data = df,family = "binomial")
-res = tidy(mod5,exponentiate = T)
-kable(res,digits = 3)
+mod15 <- glm(GAD7_elevated ~ BE_WEIGHT_MEAN, data = df, family = "binomial")
+tbl_regression(mod15)
 ```
-
-`r report(mod)`
 
@@
 ```{r}
-mod6 = glm(GAD7_Total~ BE_WEIGHT_MEAN,data = df[df$GAD7_elevated=="yes",],family = "poisson")
-res = tidy(mod6,exponentiate = T)
-kable(res,digits = 3)
+mod16 <- glm(GAD7_Total ~ BE_WEIGHT_MEAN,
+  data = df[df$GAD7_elevated == "yes", ],
+  family = "poisson"
+)
+tbl_regression(mod16)
 ```
-
-`r report(mod)`
 
@@
 ```{r}
-plot_models(mod1,mod3,mod5) + theme_bw() + theme(legend.position = "none")
+ggcoef_compare(list("PHQ9" = mod,
+  "PHQ9" = mod5, "PHQ9" = mod7, "PHQ9" = mod9,"GAD7" = mod3,
+  "GAD7" = mod11, "GAD7" = mod13, "GAD7" = mod15
+), exponentiate = T)
 ```

### 2024-04-07 Christine Chan/EnVision CF/amir_nacfc_2024.qmd
linux path

@@
 library(emmeans)
-switch(Sys.info()[["sysname"]],
-  Windows = {
-    home_dir <- "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF"
-  },
-  Darwin = {
-    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
-  }
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
 )

### 2024-04-08 Christine Chan/EnVision CF/katie_nacfc_2024.qmd
Update katie_nacfc_2024.qmd

@@
 library(pracma)
-switch(Sys.info()[["sysname"]],
-  Windows = {
-    home_dir <- "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF"
-  },
-  Darwin = {
-    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
-  }
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
 )

### 2024-04-08 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
Update sweat_test_analysis.qmd

@@
       dark: darkly
-bibliography: /Users/timvigers/Documents/Notes/zotero.bib
-csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
 editor: source
@@
 library(caret)
-source("/Users/timvigers/GitHub/CF/genotype_class_severity.R")
-switch(Sys.info()[["sysname"]],
-  Windows = {
-    home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Test"
-  },
-  Darwin = {
-    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Test"
-  }
+source("~/GitHub/CF/genotype_class_severity.R")
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Test",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Test",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Test"
 )
@@
 
-plot(roc_cf, print.auc = TRUE, col = "blue")
-plot(roc_non_cf, print.auc = TRUE, col = "red")
+plot(roc_cf, print.auc = TRUE,lty=1)
+plot(roc_non_cf, print.auc = TRUE,add=T,lty=2,print.auc.y=0.4)
+legend("bottomright",legend = c("CF vs. CFTR-RD or CRMS","Non-CF vs. CFTR-RD or CRMS"),lty = c(1,2))
 ```

### 2024-04-08 Christine Chan/EnVision CF/amir_nacfc_2024.qmd
test commit
adding iAUC 60

@@
 insulin <- gsub("timepoint_", "insulin_", glucose)
-# Remove those missing all OGTT values
-df <- df[rowSums(is.na(df[, glucose])) < length(glucose), ]
+# Remove those missing last 3 glucose timepoints
+df <- df[rowSums(!is.na(df[, tail(glucose, 3)])) > 0, ]
 ```

### 2024-04-08 Christine Chan/EnVision CF/katie_nacfc_2024.qmd
adding iAUC 60

--- 2024-04-03 Christine Chan/EnVision CF/katie_nacfc_2024.qmd
@@
 })
-df$iAUC120 <- apply(df, 1, function(r) {
+df$iAUC60gluc <- apply(df, 1, function(r) {
+  y <- as.numeric(r[glucose[1:4]]) - as.numeric(r[glucose[1]])
+  if (is.na(last(y))) {
+    return(NA)
+  } else {
+    x <- c(0, 10, 30, 60)[!is.na(y)]
+    y <- y[!is.na(y)]
+    auc <- trapz(x, y)
+    return(auc)
+  }
+})
+df$iAUC120gluc <- apply(df, 1, function(r) {
   y <- as.numeric(r[glucose[1:6]]) - as.numeric(r[glucose[1]])
@@
 })
-# Other formatting
-df$date_visit <- lubridate::ymd(df$date_visit)
-# Write for checking
-df$Keep <- NULL
+df$iAUC30ins <- apply(df, 1, function(r) {
+  y <- as.numeric(r[insulin[1:3]]) - as.numeric(r[insulin[1]])
+  if (is.na(last(y))) {
+    return(NA)
+  } else {
+    x <- c(0, 10, 30)[!is.na(y)]
+    y <- y[!is.na(y)]
+    auc <- trapz(x, y)
+    return(auc)
+  }
+})
+df$iAUC60ins <- apply(df, 1, function(r) {
+  y <- as.numeric(r[insulin[1:4]]) - as.numeric(r[insulin[1]])
+  if (is.na(last(y))) {
+    return(NA)
+  } else {
+    x <- c(0, 10, 30, 60)[!is.na(y)]
+    y <- y[!is.na(y)]
+    auc <- trapz(x, y)
+    return(auc)
+  }
+})
+df$iAUC120ins <- apply(df, 1, function(r) {
+  y <- as.numeric(r[insulin[1:6]]) - as.numeric(r[insulin[1]])
+  if (is.na(last(y))) {
+    return(NA)
+  } else {
+    x <- c(0, 10, 30, 60, 90, 120)[!is.na(y)]
+    y <- y[!is.na(y)]
+    auc <- trapz(x, y)
+    return(auc)
+  }
+})
+df$iAUC180ins <- apply(df, 1, function(r) {
+  y <- as.numeric(r[insulin]) - as.numeric(r[insulin[1]])
+  if (is.na(last(y))) {
+    return(NA)
+  } else {
+    x <- c(0, 10, 30, 60, 90, 120, 150, 180)[!is.na(y)]
+    y <- y[!is.na(y)]
+    auc <- trapz(x, y)
+    return(auc)
+  }
+})
+# Hypoglycemia
+df$Hypo70 <- factor(rowSums(df[, glucose] < 70, na.rm = T) > 0,
+  levels = c(F, T), labels = c("No", "Yes")
+)
+df$Hypo60 <- factor(rowSums(df[, glucose] < 60, na.rm = T) > 0,
+  levels = c(F, T), labels = c("No", "Yes")
+)
+df$Hypo50 <- factor(rowSums(df[, glucose] <= 50, na.rm = T) > 0,
+  levels = c(F, T), labels = c("No", "Yes")
+)
+# HOMA IR
+df$homa_ir <- (df$timepoint_0_min * df$insulin_0_min) / 405
+# Write for Amir's abstract
 write.csv(df,
@@
     redcap_data_access_group, age_at_current_date:origin_race, cftr_mutation_1,
-    cftr_mutation_2
+    cftr_mutation_2, weight
   ) %>%
@@
     redcap_data_access_group, Diagnosis, timepoint_0_min, timepoint_30_min,
-    timepoint_60_min, timepoint_120_min, timepoint_180_min, iAUC30, iAUC120,
-    iAUC180
+    timepoint_60_min, timepoint_120_min, timepoint_180_min,
+    iAUC30gluc:iAUC180ins, homa_ir
   ) %>%

### 2024-04-10 Emily Muther/body_image_trikafta.qmd
nicer table and forest plot

@@
 library(report)
-library(sjPlot)
 library(ggstats)
+library(Hmisc)
 home_dir <- switch(Sys.info()["sysname"],
@@
 # Read in data
-df <- read.spss("./Emily Muther/General - Muther Research Group/BodyImage_Trikafta/Data/TrikaftaBodyImage3.6.23.sav", to.data.frame = T)
+df <- read.spss("./Emily Muther/General/BodyImage_Trikafta/Data/TrikaftaBodyImage3.6.23.sav",
+  to.data.frame = T
+)
 # Filter
@@
 df$FRS_Category <- relevel(df$FRS_Category, ref = "neutral")
+levels(df$FRS_Category) <- c(
+  "Neutral", "Want to be larger", "Want to be smaller"
+)
+# Labels
+label(df$FRS_Category) <- "FRS Category"
+label(df$BE_APPEARANCE_MEAN) <- "Mean Body Esteem Scale (Appearance)"
+label(df$BE_ATTRIBUTION_MEAN) <- "Mean Body Esteem Scale (Attribution)"
+label(df$BE_WEIGHT_MEAN) <- "Mean Body Esteem Scale (Weight)"
 ```
@@
 
-### Forest plot of logistic regression results
+# Nicely formatted stuff
+
+## Table
 
 ```{r}
-ggcoef_compare(list("PHQ9" = mod,
-  "PHQ9" = mod5, "PHQ9" = mod7, "PHQ9" = mod9,"GAD7" = mod3,
+tbl_stack(
+  list(
+    mod %>%
+      tbl_regression(exponentiate = T,intercept = T),
+    mod5 %>%
+      tbl_regression(exponentiate = T,intercept = T),
+    mod7 %>%
+      tbl_regression(exponentiate = T,intercept = T),
+    mod9 %>%
+      tbl_regression(exponentiate = T,intercept = T),
+    mod3 %>%
+      tbl_regression(exponentiate = T,intercept = T),
+    mod11 %>%
+      tbl_regression(exponentiate = T,intercept = T),
+    mod13 %>%
+      tbl_regression(exponentiate = T,intercept = T),
+    mod15 %>%
+      tbl_regression(exponentiate = T,intercept = T)
+  ),
+  group_header = c(
+    rep("Elevated PHQ9", times = 4),
+    rep("Elevated GAD7", times = 4)
+  )
+)
+```
+
+This table shows the odds ratios and confidence intervals for an elevated score on the PHQ9 or GAD7.
+Odds ratios for FRS category represent the multiplicative difference in odds compared the reference group ("Neutral"), and estimates for BES mean scores represent the average multiplicative change in odds for a one unit increase in score.
+For example, the odds of an elevated score on the GAD for someone who endorsed wanting to be larger is 0.75 times those of someone who endorsed feeling neutral about their body (i.e.
+a decrease of 25%).
+
+## Forest plot of logistic regression results
+
+```{r}
+# Plot models together
+p <- ggcoef_compare(list(
+  "PHQ9" = mod,
+  "PHQ9" = mod5, "PHQ9" = mod7, "PHQ9" = mod9, "GAD7" = mod3,
   "GAD7" = mod11, "GAD7" = mod13, "GAD7" = mod15
-), exponentiate = T)
+), exponentiate = T, add_reference_rows = F) +
+  theme(text = element_text(size = 18))
+p
+# Save
+ggsave("./Emily Muther/General/BodyImage_Trikafta/Reports/forest_plot.png",
+       plot = p,width = 9,height = 6,units = "in")
 ```
+
+Figure 1 shows the unadjusted odds ratios and confidence intervals for an elevated score on the PHQ9 or GAD7.
+Filled circles indicate an odds ratio significantly different from 1.
+Estimates to the left of the vertical bar indicate lower odds of an elevated score.

### 2024-04-15 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
updates for Jordana

@@
 library(caret)
+library(twoxtwo)
 library(vcd)
@@
 Accuracy is the proportion of samples where the two tests agreed (positive/positive and negative/negative).
-Cohen's kappa accounts for agreement between the two tests by chance, and is interpreted analagously to correlation where -1 indicates perfect disagreement and 1 indicates perfect agreement.
-Cohen's kapps of NaN or NA indicates a lack of variability between the two tests.
+Cohen's kappa accounts for agreement between the two tests by chance, and is interpreted analogously to correlation where -1 indicates perfect disagreement and 1 indicates perfect agreement.
+A Cohen's kappa of NaN or NA indicates a lack of variability between the two tests.
 

### 2024-04-15 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
update for Jordana

@@
 library(caret)
-library(twoxtwo)
-library(vcd)
 library(eulerr)
@@
 )
-db_demographics$age_at_cos <- difftime(db_demographics$cos_result_date,
+db_demographics$age_at_cos <- as.numeric(difftime(db_demographics$cos_result_date,
   db_demographics$dob,
   units = "days"
-) / 365.25
+)) / 365.25
+db_demographics$age_group <- cut(db_demographics$age_at_cos, c(-Inf, 18, Inf),
+  right = F
+)
 # Format
@@
 })
-# # Drop levels
-# level_drop <- c("sex", "Race", "Ethnicity")
-# db_demographics[, level_drop] <- lapply(
-#   db_demographics[, level_drop],
-#   function(c) {
-#     droplevels(factor(c))
-#   }
-# )
 # Add groups to other dataframes
@@
 names(bug_list) <- bugs
-# Agreement columns
-df[, bugs] <- lapply(bugs, function(b) {
-  cos <- df[, paste0(b, "_cos")]
-  sos <- df[, paste0(b, "_sos")]
-  f <- factor(paste(cos, sos, sep = "_"),
-    exclude = c(
-      "NA_Negative", "Negative_NA", "NA_Positive", "Positive_NA", "NA_NA"
+# Add group to dataframe
+df <- left_join(df, db_demographics %>% select(record_id, group),
+  by = join_by(record_id)
 )
-  )
-  f <- factor(f,
-    levels = c(
-      "Negative_Negative", "Negative_Positive", "Positive_Negative",
-      "Positive_Positive"
-    ),
-    labels = c(
-      "COS Negative / SOS Negative", "COS Negative / SOS Positive",
-      "COS Positive / SOS Negative", "COS Positive / SOS Positive"
-    )
-  )
-  return(f)
-})
 ```
@@
 db_demographics %>%
-  select(group, age_at_cos, sex, Race, Ethnicity) %>%
+  select(group, age_at_cos, age_group, sex, Race, Ethnicity) %>%
   tbl_summary(
     by = group,
-    statistic = all_continuous()~c("{median} ({min}, {max})"),
-    label = list("sex" = "Sex", "age_at_cos" = "Age at COS")
+    statistic = all_continuous() ~ c("{median} ({min}, {max})"),
+    label = list(
+      "sex" = "Sex", "age_at_cos" = "Age at COS", "age_group" = "Age Group"
+    )
   ) %>%
@@
 
-Bangdiwala's agreement plot @bangdiwalaAgreementChart2013 is a way of visualizing 2 x 2 contingency tables, and additional details are in the paper referenced at the end of this document.
-
 ```{r}
@@
   d <- data[, c(cos, sos)]
+  colnames(d) <- c("COS", "SOS")
   d <- d[complete.cases(d), ]
   # Agreement metrics
-  c <- confusionMatrix((d[, sos]),
-    reference = d[, cos], positive = "Positive", mode = "everything",
+  c <- confusionMatrix(d$SOS,
+    reference = d$COS,
+    positive = "Positive", mode = "everything",
     dnn = c("SOS", "COS")
   )
-  # Return
-  return(list(two_by_two = c$table, accuracy = c$overall))
+  # Print
+  cat("\n")
+  cat("####", "2x2 Table")
+  cat("\n")
+  d %>%
+    tbl_cross(percent = "cell") %>%
+    print()
+  cat("\n")
+  cat("####", "2x2 Table (With CIs)")
+  cat("\n")
+  d %>%
+    tbl_cross(percent = "cell") %>%
+    add_ci() %>%
+    print()
+  cat("\n")
+  cat("Overall accuracy (both negative/negative and positive/positive) was ",
+    scales::percent(c$overall["Accuracy"], accuracy = 0.1), " (95% CI: ",
+    scales::percent(c$overall["AccuracyLower"], accuracy = 0.1), " - ",
+    scales::percent(c$overall["AccuracyUpper"], accuracy = 0.1), ").",
+    sep = ""
+  )
+  cat("\n")
+  cat("\n")
+  cat("Kappa was ", round(c$overall["Kappa"], 3), ".", sep = "")
+  cat("\n")
+  cat("\n")
+  cat("#### Euler diagram of negative results")
+  cat("\n")
+  print(plot(euler(d == "Negative"), quantities = T))
+  cat("\n")
+  cat("\n")
+  cat("#### Euler diagram of positive results")
+  cat("\n")
+  print(plot(euler(d == "Positive"), quantities = T))
+  cat("\n")
 }
-```
-
-```{r}
-l <- lapply(bugs, function(b) {
-  a <- agreement(b)
-  # Remove those where both tests were positive
-  n <- a$two_by_two
-  n[2, 2] <- 0
-  # Proportion tests for non-agreement in participants with one negative test
-  cos_p <- prop.test(n[1, 2], sum(n))
-  sos_p <- prop.test(n[2, 1], sum(n))
-  nn <- prop.test(n[1, 1], sum(n))
-  return(list(
-    Accuracy = a$accuracy["Accuracy"],
-    `Accuracy LL` = a$accuracy["AccuracyLower"],
-    `Accuracy UL` = a$accuracy["AccuracyUpper"],
-    Kappa = a$accuracy["Kappa"],
-    `Mcnemar P Value` = a$accuracy["McnemarPValue"],
-    `N With Negative Test on Either` = sum(n),
-    `N Negative on Both` = n[1, 1],
-    `Prop. Both Negative` = nn$estimate,
-    `Prop. Both Negative LL ` = nn$conf.int[1],
-    `Prop. Both Negative UL ` = nn$conf.int[2],
-    `N Postive on COS and Negative on SOS` = n[1, 2],
-    `Prop. Postive on COS` = cos_p$estimate,
-    `Prop. Postive on COS LL ` = cos_p$conf.int[1],
-    `Prop. Postive on COS UL ` = cos_p$conf.int[2],
-    `N Postive on SOS and Negative on COS` = n[2, 1],
-    `Prop. Postive on SOS` = sos_p$estimate,
-    `Prop. Postive on SOS LL ` = sos_p$conf.int[1],
-    `Prop. Postive on SOS UL ` = sos_p$conf.int[2]
-  ))
-})
-l <- data.frame(do.call(rbind, l), check.names = F)
-l <- data.frame(lapply(l, as.numeric), check.names = F)
-rownames(l) <- bug_names
-gt(l, rownames_to_stub = T) %>% fmt_number(n_sigfig = 3)
+agreement_test <- function(bug, group = "group", data = df) {
+  # Prepare data
+  cos <- paste0(bug, "_cos")
+  sos <- paste0(bug, "_sos")
+  d <- data[, c(cos, sos, group)]
+  d <- d[complete.cases(d), ]
+  colnames(d) <- c("COS", "SOS", "Group")
+  # Agreement and detailed agreement
+  d$agree <- d$COS == d$SOS
+  d$detailed_agree <- paste0("COS ", d$COS, "/", "SOS ", d$SOS)
+  # Print
+  cat("\n")
+  d %>%
+    select(agree, detailed_agree, Group) %>%
+    tbl_summary(by = Group, label = list(
+      agree ~ "Agree (Both Positive/Positive and Negative/Negative)",
+      detailed_agree ~ "Detailed Agreement"
+    )) %>%
+    add_overall() %>%
+    add_p(test.args = all_tests("fisher.test") ~ list(workspace = 2e6)) %>%
+    separate_p_footnotes() %>%
+    print()
+  cat("\n")
+}
 ```
@@
   cat("\n")
-  cos <- paste0(b, "_cos")
-  sos <- paste0(b, "_sos")
-  d <- df[, c(cos, sos)]
-  d <- d[complete.cases(d), ]
-  colnames(d) <- c("COS", "SOS")
+  cat("###", "Overall agreement")
   cat("\n")
-  c <- confusionMatrix(d[, "SOS"],
-    reference = d[, "COS"], positive = "Positive", mode = "everything"
-  )
-  metrics <- data.frame(c$overall)
-  cat("### 2x2 table")
+  agreement(b)
   cat("\n")
-  two_by_two <- data.frame(c$table)
-  print(gt(two_by_two))
+  cat("###", "Agreement by group")
   cat("\n")
-  cat("### Accuracy")
-  cat("\n")
-  print(gt(metrics, rownames_to_stub = T) %>%
-    fmt_number(n_sigfig = 3) %>%
-    tab_options(column_labels.hidden = TRUE))
-  cat("\n")
-  cat("### Agreement plot")
-  cat("\n")
-  print(agreementplot(table(d)))
-  cat("\n")
-  cat("### Positive/Positive Excluded")
-  cat("\n")
-  t <- d[rowSums(d == "Positive") < 2, ]
-  t$`Samples With a Negative Test` <- paste(
-    "COS", t[, "COS"], "/", "SOS", t[, "SOS"]
-  )
-  t %>%
-    select(`Samples With a Negative Test`) %>%
-    tbl_summary() %>%
-    add_ci() %>%
-    as_gt() %>%
-    print()
-  cat("\n")
-  cat("### Euler diagram of negative results")
-  cat("\n")
-  print(plot(euler(d == "Negative"), quantities = T))
+  agreement_test(b)
   cat("\n")
 }))
+
+# invisible(lapply(names(bug_list), function(b) {
+#   cat("\n")
+#   cat("##", as.character(bug_list[b]))
+#   cat("\n")
+#   cos <- paste0(b, "_cos")
+#   sos <- paste0(b, "_sos")
+#   d <- df[, c(cos, sos)]
+#   d <- d[complete.cases(d), ]
+#   colnames(d) <- c("COS", "SOS")
+#   cat("\n")
+#   c <- confusionMatrix(d[, "SOS"],
+#     reference = d[, "COS"], positive = "Positive", mode = "everything"
+#   )
+#   metrics <- data.frame(c$overall)
+#   cat("### 2x2 table")
+#   cat("\n")
+#   two_by_two <- data.frame(c$table)
+#   print(gt(two_by_two))
+#   cat("\n")
+#   cat("### Accuracy")
+#   cat("\n")
+#   print(gt(metrics, rownames_to_stub = T) %>%
+#     fmt_number(n_sigfig = 3) %>%
+#     tab_options(column_labels.hidden = TRUE))
+#   cat("\n")
+#   cat("### Agreement plot")
+#   cat("\n")
+#   print(agreementplot(table(d)))
+#   cat("\n")
+#   cat("### Positive/Positive Excluded")
+#   cat("\n")
+#   t <- d[rowSums(d == "Positive") < 2, ]
+#   t$`Samples With a Negative Test` <- paste(
+#     "COS", t[, "COS"], "/", "SOS", t[, "SOS"]
+#   )
+#   t %>%
+#     select(`Samples With a Negative Test`) %>%
+#     tbl_summary() %>%
+#     add_ci() %>%
+#     as_gt() %>%
+#     print()
+#   cat("\n")
+#   cat("### Euler diagram of negative results")
+#   cat("\n")
+#   print(plot(euler(d == "Negative"), quantities = T))
+#   cat("\n")
+# }))
 ```

### 2024-04-17 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
chan stratified by HbA1c

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
 library(tidyverse)
 library(arsenal)
-load("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/analysis_dataset.rdata")
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+)
+knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_knit$set(root.dir = home_dir)
+```
+
+```{r}
+load("./Colleen Wood/Prepost Trikafta/Data_Cleaned/analysis_dataset.rdata")
+df <- df %>%
+  group_by(`CFF ID`) %>%
+  mutate(`Baseline HbA1c` = HbA1c[`Pre or Post` == "Pre"]) %>%
+  ungroup()
+df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf), right = F)
 ```
@@
 # Table
-t1 <- tableby(~ Age + `BMI Percentile` + Weight + Sex + ppFEV1 + ppFVC +
+t1 <- tableby(
+  ~ Age + `BMI Percentile` + Weight + Sex + ppFEV1 + ppFVC +
     `G-tube in past 12 months` + `Pancreatic Status` + Site +
@@
 ```{r results='asis'}
-p <- paired(`Pre or Post` ~ `Num. A1c Values` + HbA1c + Time + Weight + `BMI Percentile` +
+p <- paired(
+  `Pre or Post` ~ `Num. A1c Values` + HbA1c + Time + Weight + `BMI Percentile` +
     ppFEV1 + ppFVC + TDD + `Long acting insulin dose`,
   data = df, id = `CFF ID`,
+  signed.rank.exact = FALSE,
+  numeric.stats = c("Nmiss", "medianrange", "q1q3")
+)
+summary(p, pfootnote = T, digits = 1)
+```
+
+### Stratified by baseline HbA1c
+
+```{r results='asis'}
+p <- paired(
+  `Pre or Post` ~ `Num. A1c Values` + HbA1c + Time + Weight + `BMI Percentile` +
+    ppFEV1 + ppFVC,
+  data = df, id = `CFF ID`, strata = `Baseline HbA1c`,
   signed.rank.exact = FALSE,
@@
 ```{r results='asis'}
-p <- paired(`Pre or Post` ~ `Num. A1c Values` + HbA1c + Time + Weight + `BMI Percentile` +
+p <- paired(
+  `Pre or Post` ~ `Num. A1c Values` + HbA1c + Time + Weight + `BMI Percentile` +
     ppFEV1 + ppFVC + TDD,
-data = df, id = `CFF ID`,strata = Site,
+  data = df, id = `CFF ID`, strata = Site,
   signed.rank.exact = FALSE, numeric.test = "notest",
@@
 ```{r results='asis'}
-cgm = df[!is.na(df$average_sensor),]
+cgm <- df[!is.na(df$average_sensor), ]
 # Table
-t1 <- tableby(~ Age + `BMI Percentile` + Weight + Sex + ppFEV1 + ppFVC +
+t1 <- tableby(
+  ~ Age + `BMI Percentile` + Weight + Sex + ppFEV1 + ppFVC +
     `G-tube in past 12 months` + `Pancreatic Status` + Site +
@@
 cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(cgm_vars, collapse = "+")))
-cgm_formula = update(cgm_formula,.~.+TDD + `Long acting insulin dose`)
+cgm_formula <- update(cgm_formula, . ~ . + TDD + `Long acting insulin dose`)
 p <- paired(cgm_formula,
   data = cgm, id = `CFF ID`,
+  signed.rank.exact = FALSE, numeric.test = "signed.rank",
+  numeric.stats = c("Nmiss", "medianrange", "q1q3")
+)
+summary(p, pfootnote = T, digits = 1)
+```
+
+#### Stratified by baseline HbA1c
+
+```{r results='asis'}
+cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(cgm_vars, collapse = "+")))
+cgm_formula <- update(cgm_formula, . ~ . + TDD + `Long acting insulin dose`)
+p <- paired(cgm_formula,
+  data = cgm, id = `CFF ID`, strata = `Baseline HbA1c`,
   signed.rank.exact = FALSE, numeric.test = "signed.rank",
@@
 ```{r results='asis'}
-cgm$daytime_cv = cgm$daytime_sd / cgm$daytime_avg_sensor_glucose
-day_cgm = c(day_cgm,"daytime_cv")
+cgm$daytime_cv <- cgm$daytime_sd / cgm$daytime_avg_sensor_glucose
+day_cgm <- c(day_cgm, "daytime_cv")
 cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(day_cgm, collapse = "+")))
@@
 
+#### Stratified by baseline HbA1c
+
+```{r results='asis'}
+cgm$daytime_cv <- cgm$daytime_sd / cgm$daytime_avg_sensor_glucose
+day_cgm <- c(day_cgm, "daytime_cv")
+cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(day_cgm, collapse = "+")))
+p <- paired(cgm_formula,
+  data = cgm, id = `CFF ID`, strata = `Baseline HbA1c`,
+  signed.rank.exact = FALSE, numeric.test = "signed.rank",
+  numeric.stats = c("Nmiss", "medianrange", "q1q3")
+)
+summary(p, pfootnote = T, digits = 1)
+```
+
 ### Nighttime CGM variables
@@
 ```{r results='asis'}
-cgm$nighttime_cv = cgm$nighttime_sd / cgm$nighttime_avg_sens_glucose
-night_cgm = c(night_cgm,"nighttime_cv")
+cgm$nighttime_cv <- cgm$nighttime_sd / cgm$nighttime_avg_sens_glucose
+night_cgm <- c(night_cgm, "nighttime_cv")
 cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(night_cgm, collapse = "+")))
@@
 ```
+
+#### Stratified by baseline HbA1c
+
+```{r results='asis'}
+cgm$nighttime_cv <- cgm$nighttime_sd / cgm$nighttime_avg_sens_glucose
+night_cgm <- c(night_cgm, "nighttime_cv")
+cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(night_cgm, collapse = "+")))
+p <- paired(cgm_formula,
+  data = cgm, id = `CFF ID`, strata = `Baseline HbA1c`,
+  signed.rank.exact = FALSE, numeric.test = "signed.rank",
+  numeric.stats = c("Nmiss", "medianrange", "q1q3")
+)
+summary(p, pfootnote = T, digits = 1)
+```

### 2024-04-17 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
race neutral lung function

@@
 library(arsenal)
+library(rspiro)
 home_dir <- switch(Sys.info()["sysname"],
@@
 load("./Colleen Wood/Prepost Trikafta/Data_Cleaned/analysis_dataset.rdata")
+# Baseline HbA1c
 df <- df %>%
@@
 df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf), right = F)
+# Race neutral equations
+df$ppFEV1gl <- pctpred_GLIgl(
+  age = df$Age, height = df$Height / 100, gender = df$Sex, FEV1 = df$FEV1
+)
+df$ppFVCgl <- pctpred_GLIgl(
+  age = df$Age, height = df$Height / 100, gender = df$Sex, FVC = df$FEV1
+)
 ```
@@
 t1 <- tableby(
-  ~ Age + `BMI Percentile` + Weight + Sex + ppFEV1 + ppFVC +
-    `G-tube in past 12 months` + `Pancreatic Status` + Site +
+  ~ Age + `BMI Percentile` + Weight + Sex + ppFEV1 + ppFEV1gl + ppFVC +
+    ppFVCgl + `G-tube in past 12 months` + `Pancreatic Status` + Site +
     Genotype + `Age Group` + Race,
@@
 p <- paired(
-  `Pre or Post` ~ `Num. A1c Values` + HbA1c + Time + Weight + `BMI Percentile` +
-    ppFEV1 + ppFVC + TDD + `Long acting insulin dose`,
+  `Pre or Post` ~ `Num. A1c Values` + HbA1c + Time + Weight +
+    `BMI Percentile` + ppFEV1 + ppFEV1gl + ppFVC + ppFVCgl + TDD +
+    `Long acting insulin dose`,
   data = df, id = `CFF ID`,
@@
 p <- paired(
-  `Pre or Post` ~ `Num. A1c Values` + HbA1c + Time + Weight + `BMI Percentile` +
-    ppFEV1 + ppFVC + TDD,
+  `Pre or Post` ~ `Num. A1c Values` + HbA1c + Time + Weight +
+    `BMI Percentile` + ppFEV1 + ppFEV1gl + ppFVC + ppFVCgl + TDD,
   data = df, id = `CFF ID`, strata = Site,

### 2024-04-22 Quarto Template.qmd
template for github

@@
 )
+github_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/GitHub",
+  "Windows" = "C:/Users/Tim/GitHub",
+  "Linux" = "/home/timvigers/GitHub"
+)
 knitr::opts_knit$set(root.dir = home_dir)

### 2024-04-22 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
formatting

@@
 library(ggstats)
-source("/Users/timvigers/GitHub/shared-resources/Visualization/forest_plots.R")
-switch(Sys.info()[["sysname"]],
-  Windows = {
-    home_dir <- "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry"
-  },
-  Darwin = {
-    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry"
-  }
+library(data.table)
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
 )
+github_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/GitHub",
+  "Windows" = "C:/Users/Tim/GitHub",
+  "Linux" = "/home/timvigers/GitHub"
+)
+source(paste0(github_dir, "/shared-resources/Visualization/forest_plots.R"))
 knitr::opts_knit$set(root.dir = home_dir)

### 2024-04-29 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
CHEC analysis started
sweat test project CHEC

@@
 # Import data
-df <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv", na.strings = "")
+df <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv",
+  na.strings = ""
+)
 n1 <- nrow(df)
@@
 
-# Flowchart
+<!-- # Flowchart -->
 
-```{r}
-grViz("digraph {
+<!-- ```{r} -->
+<!-- grViz("digraph { -->
 
-# initiate graph
-graph [layout = dot]
+<!-- # initiate graph -->
+<!-- graph [layout = dot] -->
 
-# global node settings
-node [shape = rectangle]
+<!-- # global node settings -->
+<!-- node [shape = rectangle] -->
 
-# label nodes
-data1 [label = '@@1']
-data2 [label = '@@2']
-ineligible [label = '@@3']
-demo  [label = '@@4']
-highna  [label = '@@5']
-highcl  [label = '@@6']
-highnacl  [label = '@@7']
-review [label = '@@8']
-analysis [label = '@@9']
-ratio [label = '@@10']
+<!-- # label nodes -->
+<!-- data1 [label = '@@1'] -->
+<!-- data2 [label = '@@2'] -->
+<!-- ineligible [label = '@@3'] -->
+<!-- demo  [label = '@@4'] -->
+<!-- highna  [label = '@@5'] -->
+<!-- highcl  [label = '@@6'] -->
+<!-- highnacl  [label = '@@7'] -->
+<!-- review [label = '@@8'] -->
+<!-- analysis [label = '@@9'] -->
+<!-- ratio [label = '@@10'] -->
 
-# edge definitions with the node IDs
-data1  -> data2
-data2 -> ineligible
-ineligible -> {demo,highna,highcl,highnacl,review}
-data2  -> analysis -> ratio
-highnacl -> analysis [style=invis]
-}
-[1]: paste0('Sweat tests performed \\n ','n = ',n1)
-[2]: paste0('Unique participants \\n ','n = ',n2)
-[3]: paste0('Ineligible \\n ','n = ',n2-nrow(t1_df))
-[4]: paste0('No demographics \\n ','n = ',length(no_dem))
-[5]: paste0('Na > 160 mmol/L \\n ','n = ',length(high_na))
-[6]: paste0('Cl > 160 mmol/L \\n ','n = ',length(high_cl))
-[7]: paste0('Na:Cl > 2 \\n ','n = ',length(high_ratio))
-[8]: paste0('Other \\n ','n = ',1)
-[9]: paste0('Analysis cohort \\n ','n = ',nrow(t1_df))
-[10]: paste0('Both Na and Cl Measured \\n ','n = ',sum(!is.na(df$`Na/Cl Ratio`)))
-")
-```
+<!-- # edge definitions with the node IDs -->
+<!-- data1  -> data2 -->
+<!-- data2 -> ineligible -->
+<!-- ineligible -> {demo,highna,highcl,highnacl,review} -->
+<!-- data2  -> analysis -> ratio -->
+<!-- highnacl -> analysis [style=invis] -->
+<!-- } -->
+<!-- [1]: paste0('Sweat tests performed \\n ','n = ',n1) -->
+<!-- [2]: paste0('Unique participants \\n ','n = ',n2) -->
+<!-- [3]: paste0('Ineligible \\n ','n = ',n2-nrow(t1_df)) -->
+<!-- [4]: paste0('No demographics \\n ','n = ',length(no_dem)) -->
+<!-- [5]: paste0('Na > 160 mmol/L \\n ','n = ',length(high_na)) -->
+<!-- [6]: paste0('Cl > 160 mmol/L \\n ','n = ',length(high_cl)) -->
+<!-- [7]: paste0('Na:Cl > 2 \\n ','n = ',length(high_ratio)) -->
+<!-- [8]: paste0('Other \\n ','n = ',1) -->
+<!-- [9]: paste0('Analysis cohort \\n ','n = ',nrow(t1_df)) -->
+<!-- [10]: paste0('Both Na and Cl Measured \\n ','n = ',sum(!is.na(df$`Na/Cl Ratio`))) -->
+<!-- ") -->
+<!-- ``` -->
 
@@
 
-plot(roc_cf, print.auc = TRUE,lty=1)
-plot(roc_non_cf, print.auc = TRUE,add=T,lty=2,print.auc.y=0.4)
-legend("bottomright",legend = c("CF vs. CFTR-RD or CRMS","Non-CF vs. CFTR-RD or CRMS"),lty = c(1,2))
+plot(roc_cf, print.auc = TRUE, lty = 1)
+plot(roc_non_cf, print.auc = TRUE, add = T, lty = 2, print.auc.y = 0.4)
+legend("bottomright", legend = c("CF vs. CFTR-RD or CRMS", "Non-CF vs. CFTR-RD or CRMS"), lty = c(1, 2))
 ```

### 2024-04-29 Edith Zemanick/Sweat Test/sweat_test_analysis_chec.qmd
CHEC analysis started
sweat test project CHEC

--- 2024-04-03 Quarto Template.qmd
@@
 library(tidyverse)
+library(readxl)
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Test",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Test",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Test"
+)
+github_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/GitHub",
+  "Windows" = "C:/Users/Tim/GitHub",
+  "Linux" = "/home/timvigers/GitHub"
 )
@@
 ```{r data cleaning}
+# Import exclusion data from Excel
+diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
+exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) | diagnoses$Code == "Exclude"]
+exclude <- exclude[!is.na(exclude)]
+# Import data
+df <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv",
+  na.strings = ""
+)
+# Remove exclusions from Angela
+df <- df[!(df$PatientID %in% exclude), ]
+# Diagnosis per Angela
+df$Diagnosis <- factor(diagnoses$Diagnosis[match(df$PatientID, diagnoses$PatientID)],
+  levels = c("CFTR-related disorder", "CRMS", "Cystic Fibrosis", "Non-CF"),
+  labels = c("CFTR-RD or CRMS", "CFTR-RD or CRMS", "Cystic Fibrosis", "Non-CF")
+)
+# Add modulator info
+cftr <- read.csv("./Data_Raw/Copy of Report CFTR Modulators.csv")
+df$Modulator <- apply(df, 1, function(r) {
+  mod <- cftr[cftr$Patient.ID == as.numeric(r["PatientID"]), ]
+  if (any(duplicated(mod$Age.at.Start..years.))) {
+    mod <- mod[-which(duplicated(mod$Age.at.Start..years.)), ]
+  }
+  mod <- mod[order(mod$Age.at.Start..years.), ]
+  age <- as.numeric(r["Age.at.Test.in.Years"])
+  as.character(cut(age, c(-Inf, mod$Age.at.Start..years., Inf),
+    labels = c("None", mod$Modulator), right = F
+  ))
+})
 ```

### 2024-04-29 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
survival code cleanup

--- 2024-03-13 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
@@
 library(ggstats)
-source("/Users/timvigers/GitHub/shared-resources/Visualization/forest_plots.R")
-switch(Sys.info()[["sysname"]],
-  Windows = {
-    home_dir <- "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry"
-  },
-  Darwin = {
-    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry"
-  }
+library(rprev)
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
 )
+github_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/GitHub",
+  "Windows" = "C:/Users/Tim/GitHub",
+  "Linux" = "/home/timvigers/GitHub"
+)
+source(paste0(github_dir, "/shared-resources/Visualization/forest_plots.R"))
 knitr::opts_knit$set(root.dir = home_dir)
 ```
-
-# CFRD incidence and prevalence
-
-```{r}
-# All data
-encounter <- read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv",
-  na.strings = c("", "N/A")
-)
-annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
-  na.strings = c("", "N/A")
-)
-encounter <- encounter %>%
-  arrange(eDWID, reviewyear, encounternum)
-annual <- annual %>% arrange(eDWID, ReviewYear)
-# Add age
-ages <- annual %>%
-  mutate(`Age in 2018` = 2018 - (ReviewYear - Age_YrEnd)) %>%
-  rename(reviewyear = ReviewYear) %>%
-  group_by(eDWID, reviewyear) %>%
-  summarise(
-    `Age in 2018` = as.numeric(names(sort(table(`Age in 2018`),
-      decreasing = TRUE
-    )[1])),
-    Age = Age_YrEnd[1], .groups = "drop"
-  )
-encounter <- left_join(encounter, ages, by = c("eDWID", "reviewyear"))
-encounter$`Age Group 2018` <- cut(encounter$`Age in 2018`,
-  c(-Inf, 10, 19, 31, Inf),
-  labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
-  right = F
-)
-encounter$`Age Group` <- cut(encounter$Age,
-  c(-Inf, 10, 19, 31, Inf),
-  labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
-  right = F
-)
-# For those with CFRD by registry criteria, find first visit with diagnosis
-first_cfrd_cffpr <- encounter %>%
-  filter(cfrd_status == 3) %>%
-  group_by(eDWID) %>%
-  filter(row_number() == 1) %>%
-  select(eDWID, reviewyear, `Age in 2018`) %>%
-  rename(first_cfrd_cffpr = reviewyear)
-# Add to encounter level data
-encounter <- left_join(encounter, first_cfrd_cffpr,
-  by = c("eDWID", "Age in 2018")
-)
-# Fill down CFRD Yes/No by year
-encounter <- encounter %>%
-  group_by(eDWID) %>%
-  mutate(
-    cfrd_cffpr = case_when(
-      reviewyear < first_cfrd_cffpr ~ "No",
-      reviewyear >= first_cfrd_cffpr ~ "Yes"
-    )
-  ) %>%
-  ungroup()
-encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
-# Ordered factors
-encounter$cfrd_cffpr <- factor(encounter$cfrd_cffpr,
-  levels = c("No", "Yes"), ordered = T
-)
-# Remove those missing age
-encounter <- encounter %>% filter(!is.na(Age))
-```
-
-```{r}
-#| label: tbl-incidence_cffpr
-#| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR Alone"
-# For each year, pull everyone with and without CFRD that year
-year_inc <- function(df = encounter, year, py = 1000) {
-  # Filter and count
-  inc <- df %>%
-    filter(reviewyear == year, is.na(first_cfrd_cffpr) |
-      first_cfrd_cffpr >= year) %>%
-    group_by(eDWID) %>%
-    filter(row_number() == 1) %>%
-    select(eDWID, `Age Group`, cfrd_cffpr)
-  inc <- inc %>%
-    group_by(cfrd_cffpr) %>%
-    count(`Age Group`, .drop = F)
-  inc <- inc[!is.na(inc$`Age Group`), ]
-  # Calculations
-  new_cases <- sum(inc$n[inc$cfrd_cffpr == "Yes"])
-  total_at_risk <- sum(inc$n)
-  total_inc <- new_cases / (total_at_risk / py)
-
-  new_cases_10 <-
-    sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "< 10"])
-  at_risk_10 <- sum(inc$n[inc$`Age Group` == "< 10"])
-  inc_10 <- new_cases_10 / (at_risk_10 / py)
-
-  new_cases_10_18 <- sum(inc$n[inc$cfrd_cffpr == "Yes" &
-    inc$`Age Group` == "10 - 18"])
-  at_risk_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
-  inc_10_18 <- new_cases_10_18 / (at_risk_10_18 / py)
-
-  new_cases_19_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" &
-    inc$`Age Group` == "19 - 30"])
-  at_risk_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
-  inc_19_30 <- new_cases_19_30 / (at_risk_19_30 / py)
-
-  new_cases_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" &
-    inc$`Age Group` == "> 30"])
-  at_risk_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
-  inc_30 <- new_cases_30 / (at_risk_30 / py)
-  # Nice results
-  list(
-    `Year` = year,
-    `Total New Cases` = new_cases,
-    `Total At Risk` = total_at_risk,
-    `Total Incidence` = total_inc,
-    `Total New Cases Age < 10` = new_cases_10,
-    `Total At Risk Age < 10` = at_risk_10,
-    `Total Incidence Age < 10` = inc_10,
-    `Total New Cases Age 10-18` = new_cases_10_18,
-    `Total At Risk Age 10-18` = at_risk_10_18,
-    `Total Incidence Age 10-18` = inc_10_18,
-    `Total New Cases Age 19-30` = new_cases_19_30,
-    `Total At Risk Age 19-30` = at_risk_19_30,
-    `Total Incidence Age 19-30` = inc_19_30,
-    `Total New Cases Age > 30` = new_cases_30,
-    `Total At Risk Age > 30` = at_risk_30,
-    `Total Incidence Age > 30` = inc_30
-  )
-}
-incidence <- lapply(2003:2018, function(y) {
-  year_inc(year = y)
-})
-incidence <- data.frame(do.call(rbind, incidence), check.names = F)
-incidence <- data.frame(lapply(incidence, as.numeric), check.names = F)
-kable(incidence, digits = 3)
-```
-
-In @tbl-incidence_cffpr, incidence per year is calculated as the number of new CFRD diagnoses (by registry definition alone) per 1,000 person-years of those at risk, excluding previously identified cases.
-
-```{r}
-#| label: tbl-prevalence_cffpr
-#| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR Alone"
-encounter %>%
-  group_by(eDWID, reviewyear) %>%
-  summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>%
-  group_by(reviewyear) %>%
-  summarise(
-    Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100,
-    .groups = "drop"
-  ) %>%
-  kable(., digits = 3)
-```
-
-```{r}
-#| label: tbl-prevalence_cffpr_year
-#| tbl-cap: "Yearly Prevalence of CFRD by Age Group Based on CFFPR Alone"
-encounter %>%
-  group_by(eDWID, reviewyear, `Age Group`) %>%
-  summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>%
-  group_by(reviewyear, `Age Group`) %>%
-  summarise(Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100, .groups = "drop") %>%
-  pivot_wider(names_from = `Age Group`, values_from = Prevalence) %>%
-  kable(., digits = 3)
-```
-
-In @tbl-prevalence_cffpr_year and @tbl-prevalence_cffpr, prevalence is calculated as the total number of CFRD cases (by registry definition alone) seen in clinic divided by the total number of people seen in clinic in a given year.
-
-# CFRD screening rates
-
-```{r}
-#| label: tbl-screening_ogtt
-#| tbl-cap: "Yearly OGTT Screening Rates"
-year_screen <- function(df = encounter, year) {
-  # Filter and count
-  inc <- df %>%
-    filter(reviewyear == year, first_cfrd_cffpr > year |
-      is.na(first_cfrd_cffpr)) %>%
-    mutate(ogtt = !is.na(ogttfast_bloodglucose) |
-      !is.na(twohour_bloodglucose)) %>%
-    group_by(eDWID) %>%
-    summarise(ogtt = max(ogtt), `Age Group` = `Age Group`[1])
-  inc <- inc %>%
-    group_by(ogtt) %>%
-    count(`Age Group`) %>%
-    filter(`Age Group` != "< 10")
-  # Calculations
-  total_screened <- sum(inc$n[inc$ogtt == 1])
-  total_seen <- sum(inc$n)
-  screen_rate <- total_screened / total_seen * 100
-
-  screened_10_18 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "10 - 18"]
-  seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
-  rate_10_18 <- screened_10_18 / seen_10_18 * 100
-
-  screened_19_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "19 - 30"]
-  seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
-  rate_19_30 <- screened_19_30 / seen_19_30 * 100
-
-  screened_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "> 30"]
-  seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
-  rate_30 <- screened_30 / seen_30 * 100
-
-  # Nice results
-  list(
-    `Year` = year,
-    `Total Screened` = total_screened,
-    `Total Seen` = total_seen,
-    `Screen Rate (%)` = screen_rate,
-    `Total Screened Age 10-18` = screened_10_18,
-    `Total Seen Age 10-18` = seen_10_18,
-    `Screen Rate Age 10-18 (%)` = rate_10_18,
-    `Total Screened Age 19-30` = screened_19_30,
-    `Total Seen Age 19-30` = seen_19_30,
-    `Screen Rate Age 19-30 (%)` = rate_19_30,
-    `Total Screened Age > 30` = screened_30,
-    `Total Seen Age > 30` = seen_30,
-    `Screen Rate Age > 30 (%)` = rate_30
-  )
-}
-screen <- lapply(2003:2018, function(y) {
-  year_screen(year = y)
-})
-screen <- data.frame(do.call(rbind, screen), check.names = F)
-screen <- data.frame(lapply(screen, as.numeric), check.names = F)
-kable(screen, digits = 3)
-```
-
-```{r}
-#| label: tbl-screening_a1c
-#| tbl-cap: "Yearly HbA1c Screening Rates"
-year_screen <- function(df = encounter, year) {
-  # Filter and count
-  inc <- df %>%
-    filter(reviewyear == year, first_cfrd_cffpr > year |
-      is.na(first_cfrd_cffpr)) %>%
-    mutate(a1c = !is.na(hgba1c)) %>%
-    group_by(eDWID) %>%
-    summarise(a1c = max(a1c), `Age Group` = `Age Group`[1])
-  inc <- inc %>%
-    group_by(a1c) %>%
-    count(`Age Group`) %>%
-    filter(`Age Group` != "< 10")
-  # Calculations
-  total_screened <- sum(inc$n[inc$a1c == 1])
-  total_seen <- sum(inc$n)
-  screen_rate <- total_screened / total_seen * 100
-
-  screened_10_18 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "10 - 18"]
-  seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
-  rate_10_18 <- screened_10_18 / seen_10_18 * 100
-
-  screened_19_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "19 - 30"]
-  seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
-  rate_19_30 <- screened_19_30 / seen_19_30 * 100
-
-  screened_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "> 30"]
-  seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
-  rate_30 <- screened_30 / seen_30 * 100
-
-  # Nice results
-  list(
-    `Year` = year,
-    `Total Screened` = total_screened,
-    `Total Seen` = total_seen,
-    `Screen Rate (%)` = screen_rate,
-    `Total Screened Age 10-18` = screened_10_18,
-    `Total Seen Age 10-18` = seen_10_18,
-    `Screen Rate Age 10-18 (%)` = rate_10_18,
-    `Total Screened Age 19-30` = screened_19_30,
-    `Total Seen Age 19-30` = seen_19_30,
-    `Screen Rate Age 19-30 (%)` = rate_19_30,
-    `Total Screened Age > 30` = screened_30,
-    `Total Seen Age > 30` = seen_30,
-    `Screen Rate Age > 30 (%)` = rate_30
-  )
-}
-screen <- lapply(2003:2018, function(y) {
-  year_screen(year = y)
-})
-screen <- data.frame(do.call(rbind, screen), check.names = F)
-screen <- data.frame(lapply(screen, as.numeric), check.names = F)
-kable(screen, digits = 3)
-```
-
-In @tbl-screening_ogtt and @tbl-screening_a1c, people with CFRD (by registry definition) and those under age 10 were excluded.
-Screening rate is calculated as the number of people with an OGTT or HbA1c, respectively, divided by the total number of people seen in clinic that year.
 
@@
 data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
+# Find each person's entry date
+annual = annual %>% group_by(eDWID) %>%
+  mutate(entry_date = min(ReviewYear))
 # Remove those with T1D or T2D
@@
 annual <- left_join(annual, by, by = join_by(eDWID))
-# Get first year of CFRD
-first_cfrd <- annual %>%
-  filter(cfrd_status_annual == "CFRD+") %>%
-  group_by(eDWID) %>%
-  summarise(first_cfrd = min(ReviewYear))
-annual <- left_join(annual, first_cfrd, by = join_by(eDWID))
 # BMI percentile to categorical
@@
 annual$Hispanicrace <- factor(annual$Hispanicrace, levels = c("No", "Yes"))
-```
-
-Registry fields included in our liver disease yes/no variable were:
-
-```{r}
-data.frame(data_dict$code_meaning[data_dict$var %in%
-  sub("A_", "", liver_fields)]) %>%
-  gt() %>%
-  tab_options(column_labels.hidden = TRUE)
-```
-
-Registry fields included in our tube feeding yes/no variable were:
-
-```{r}
-data.frame(data_dict$code_meaning[
-  paste0(data_dict$field_name, data_dict$code_value) %in%
-    sub("A_", "", tube_fields)
-]) %>%
-  gt() %>%
-  tab_options(column_labels.hidden = TRUE)
-```
-
-# Survival analysis - first CFRD diagnosis
-
-Based on CFFPR annualized data.
-These data are interval censored because we know the first visit at which people with CFRD were diagnosed, but we do not know the exact time that they developed CFRD.
-These data are also left truncated because not all participants enter the study at the same time.
-
-```{r}
+# Get first year of CFRD
+first_cfrd <- annual %>%
+  filter(cfrd_status_annual == "CFRD+") %>%
+  group_by(eDWID) %>%
+  summarise(first_cfrd = min(ReviewYear))
+annual <- left_join(annual, first_cfrd, by = join_by(eDWID))
+# Year prior to first CFRD in registry - assume negative for CFRD. Assume all
+# years following are CFRD+
+annual$CFRD <- NA
+annual$CFRD[annual$ReviewYear < annual$first_cfrd] <- "CFRD-"
+annual$CFRD[annual$ReviewYear >= annual$first_cfrd] <- "CFRD+"
+annual$CFRD[is.na(annual$first_cfrd)] <- "CFRD-"
+annual$CFRD <- factor(annual$CFRD, levels = c("CFRD-", "CFRD+"))
+# For each person, filter out all the years after first CFRD diagnosis
+annual <- annual %>% filter(ReviewYear <= first_cfrd | is.na(first_cfrd))
+# Remove those under 10
+annual <- annual %>% filter(Age_YrEnd != 0, Age_YrEnd >= 10)
+annual$age_group = droplevels(annual$age_group)
+# Add formatted columns for survival analysis
+annual = annual %>% 
+  mutate(tstop = ReviewYear) %>%
+  group_by(eDWID) %>%
+  mutate(
+    tstart = lag(tstop, default = birth_year[1]),
+    age_at_entry = Age_YrEnd[1]
+  )
 # List of covariates
@@
   A_mycocultureresults3 = "Microorganisms in myco culture",
-  eDWID = "ID", birth_year = "Birth year", tstart = "Interval start", tstop = "Interval stop", age_at_entry = "Age at entry", Age_YrEnd = "Age", bmi_perc = "BMI %ile"
+  eDWID = "ID", birth_year = "Birth year",
+  tstart = "Interval start", tstop = "Interval stop",
+  age_at_entry = "Age at entry", Age_YrEnd = "Age", bmi_perc = "BMI %ile"
 )
-# For now, let's just assume that everyone's first CFRD+ diagnosis is correct
-s <- annual %>%
-  filter(
-    !is.na(cfrd_status_annual),
-    Age_YrEnd != 0,
-    is.na(first_cfrd) | ReviewYear <= first_cfrd,
-    Age_YrEnd >= 10
-  ) %>%
-  select(
-    eDWID, ReviewYear, birth_year, Age_YrEnd,
-    all_of(vars), cfrd_status_annual
-  ) %>%
-  rename(tstop = ReviewYear) %>%
-  group_by(eDWID) %>%
-  mutate(
-    tstart = lag(tstop, default = birth_year[1]),
-    age_at_entry = Age_YrEnd[1]
-  ) %>%
-  select(
-    eDWID, birth_year, tstart, tstop, cfrd_status_annual,
-    age_at_entry, everything()
-  )
-# Convert from tibble to dataframe for survival packages
-s <- as.data.frame(s)
-# CFRD status to numeric
-s$cfrd_status_annual <- as.numeric(s$cfrd_status_annual) - 1
-# Drop under 10s
-s$age_group <- droplevels(s$age_group)
-# Labels
-label(s) <- labels[colnames(s)]
 ```
+
+Registry fields included in our liver disease yes/no variable were:
+
+```{r}
+data.frame(data_dict$code_meaning[data_dict$var %in%
+  sub("A_", "", liver_fields)]) %>%
+  gt() %>%
+  tab_options(column_labels.hidden = TRUE)
+```
+
+Registry fields included in our tube feeding yes/no variable were:
+
+```{r}
+data.frame(data_dict$code_meaning[
+  paste0(data_dict$field_name, data_dict$code_value) %in%
+    sub("A_", "", tube_fields)
+]) %>%
+  gt() %>%
+  tab_options(column_labels.hidden = TRUE)
+```
+
+# CFRD incidence and prevalence
+
+```{r}
+t2004 = annual %>% select()
+```
+
+```{r}
+# Use flexsurv to fit a survival model that can be passed to rprev
+llog <- flexsurv::flexsurvreg(Surv(tstop, as.numeric(CFRD)==2) ~ 1, 
+                              data=prevdata, dist='llogis')
+prev_llog <- prevalence(index='2004-01-01', 
+                        num_years_to_estimate=1, 
+                        data=annual, 
+                        inc_formula = entrydate ~ 1,
+                        surv_model=llog,
+                        death_column = "first_cfrd",
+                        incident_column = "first_cfrd",
+                        N_boot = 100)
+```
+
+# CFRD screening rates
+
+# Survival analysis - first CFRD diagnosis
+
+Based on CFFPR annualized data.
+These data are interval censored because we know the first visit at which people with CFRD were diagnosed, but we do not know the exact time that they developed CFRD.
+These data are also left truncated because not all participants enter the study at the same time.
 
@@
 # Whole cohort
-km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ 1,
-  data = s, cluster = eDWID
+km_fit <- survfit(Surv(tstart, tstop, as.numeric(CFRD)==2) ~ 1,
+  data = annual, id = eDWID
 )
 ggsurvplot(km_fit,
-  data = s, risk.table = T, cumevents = T,
+  data = annual, risk.table = T, cumevents = T,
   fun = "event", xlim = c(2003, 2018), break.x.by = 1
@@
 # By age group
-km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ age_group,
-  data = s, cluster = eDWID
+km_fit <- survfit(Surv(tstart, tstop, as.numeric(CFRD)==2) ~ age_group,
+  data = annual, id = eDWID
 )
 ggsurvplot(km_fit,
-  data = s, risk.table = T, cumevents = T,
+  data = annual, risk.table = T, cumevents = T,
   fun = "event", xlim = c(2003, 2018), break.x.by = 1
@@
 )
-y <- "Surv(tstart, tstop, cfrd_status_annual)"
+y <- "Surv(tstart, tstop, as.numeric(CFRD)==2)"
 f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
-cox_res <- coxph(f, data = s, cluster = eDWID)
+cox_res <- coxph(f, data = annual, id = eDWID)
 # Plot survival
@@
 time_varying <- rownames(test.ph)[test.ph$p < 0.05]
-# ggcoxzph(test.ph)
-# ggcoxdiagnostics(cox_res,type = "martingale")
-# ggcoxdiagnostics(cox_res,type = "deviance")
-# ggcoxdiagnostics(cox_res,type = "score")
-# ggcoxdiagnostics(cox_res,type = "schoenfeld")
 ```
@@
 ```{r}
-s_split <- survSplit(Surv(tstart, tstop, cfrd_status_annual) ~ .,
-  data = s, cut = c(2008, 2013, 2018), episode = "tgroup"
+annual_split <- survSplit(Surv(tstart, tstop, as.numeric(CFRD)==2) ~ .,
+  data = annual, cut = c(2008, 2013, 2018), episode = "tgroup"
 )
-s_split$tgroup <- factor(s_split$tgroup,
+annual_split$tgroup <- factor(annual_split$tgroup,
   levels = 1:3, labels = c("[2003,2009)", "[2009,2014)", "[2014,2019)")
@@
   # Fit one stratified model and one with a regular interaction term
-  cox_res_split <- coxph(new_f, data = s_split, cluster = eDWID)
-  cox_res_split_int <- coxph(new_f_int, data = s_split, cluster = eDWID)
+  cox_res_split <- coxph(new_f, data = annual_split, cluster = eDWID)
+  cox_res_split_int <- coxph(new_f_int, data = annual_split, cluster = eDWID)
   # Use the regular interaction term to get means for each time period
@@
 ```
-
-<!-- #### P. aeruginosa -->
-
-<!-- ```{r} -->
-<!-- # Update formula -->
-<!-- cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa), -->
-<!--   data = s, cluster = eDWID -->
-<!-- ) -->
-<!-- # Forest plot and coefficients table -->
-<!-- ggforest(cox_res, data = s) -->
-<!-- ``` -->
-
-<!-- #### Aspergillus -->
-
-<!-- ```{r} -->
-<!-- # Update formula -->
-<!-- cox_res <- coxph(update(f, . ~ . + A_fungalyeast1), -->
-<!--   data = s, cluster = eDWID -->
-<!-- ) -->
-<!-- # Forest plot and coefficients table -->
-<!-- ggforest(cox_res, data = s) -->
-<!-- ``` -->
-
-<!-- #### ABPA -->
-
-<!-- ```{r} -->
-<!-- # Update formula -->
-<!-- cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1), -->
-<!--   data = s, cluster = eDWID -->
-<!-- ) -->
-<!-- # Forest plot and coefficients table -->
-<!-- ggforest(cox_res, data = s) -->
-<!-- ``` -->
-
-<!-- ### Vx770 -->
-
-<!-- ```{r} -->
-<!-- f <- update(f, . ~ . + A_Vx770) -->
-<!-- cox_res <- coxph(f, data = s, cluster = eDWID) -->
-<!-- # Plot survival -->
-<!-- ggsurvplot(survfit(cox_res), -->
-<!--   data = s, risk.table = T, cumevents = T, fun = "event", -->
-<!--   xlim = c(2003, 2018), break.x.by = 1 -->
-<!-- ) -->
-<!-- # Forest plot and coefficients table -->
-<!-- ggforest(cox_res, data = s) -->
-<!-- # Diagnostics -->
-<!-- # test.ph= cox.zph(cox_res) -->
-<!-- # ggcoxzph(test.ph) -->
-<!-- # ggcoxdiagnostics(cox_res,type = "martingale") -->
-<!-- # ggcoxdiagnostics(cox_res,type = "deviance") -->
-<!-- # ggcoxdiagnostics(cox_res,type = "score") -->
-<!-- # ggcoxdiagnostics(cox_res,type = "schoenfeld") -->
-<!-- ``` -->
-
-<!-- #### P. aeruginosa -->
-
-<!-- ```{r} -->
-<!-- # Update formula -->
-<!-- cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa), -->
-<!--   data = s, cluster = eDWID -->
-<!-- ) -->
-<!-- # Forest plot and coefficients table -->
-<!-- ggforest(cox_res, data = s) -->
-<!-- ``` -->
-
-<!-- #### Aspergillus -->
-
-<!-- ```{r} -->
-<!-- # Update formula -->
-<!-- cox_res <- coxph(update(f, . ~ . + A_fungalyeast1), -->
-<!--   data = s, cluster = eDWID -->
-<!-- ) -->
-<!-- # Forest plot and coefficients table -->
-<!-- ggforest(cox_res, data = s) -->
-<!-- ``` -->
-
-<!-- #### ABPA -->
-
-<!-- ```{r} -->
-<!-- # Update formula -->
-<!-- cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1), -->
-<!--   data = s, cluster = eDWID -->
-<!-- ) -->
-<!-- # Forest plot and coefficients table -->
-<!-- ggforest(cox_res, data = s) -->
-<!-- ``` -->
-
-<!-- #### Mycobacterium -->
-
-<!-- ```{r} -->
-<!-- # Update formula -->
-<!-- cox_res <- coxph(update(f, . ~ . + A_mycocultureresults3), -->
-<!--   data = s, cluster = eDWID -->
-<!-- ) -->
-<!-- # Forest plot and coefficients table -->
-<!-- ggforest(cox_res, data = s) -->
-<!-- ``` -->
-
-<!-- ### Vx809 -->
-
-<!-- ```{r} -->
-<!-- f <- update(f, . ~ . + A_VX809comb) -->
-<!-- cox_res <- coxph(f, data = s, cluster = eDWID) -->
-<!-- # Plot survival -->
-<!-- ggsurvplot(survfit(cox_res), -->
-<!--   data = s, risk.table = T, cumevents = T, fun = "event", -->
-<!--   xlim = c(2003, 2018), break.x.by = 1 -->
-<!-- ) -->
-<!-- # Forest plot and coefficients table -->
-<!-- ggforest(cox_res, data = s) -->
-<!-- # Diagnostics -->
-<!-- # test.ph= cox.zph(cox_res) -->
-<!-- # ggcoxzph(test.ph) -->
-<!-- # ggcoxdiagnostics(cox_res,type = "martingale") -->
-<!-- # ggcoxdiagnostics(cox_res,type = "deviance") -->
-<!-- # ggcoxdiagnostics(cox_res,type = "score") -->
-<!-- # ggcoxdiagnostics(cox_res,type = "schoenfeld") -->
-<!-- ``` -->
-
-<!-- #### P. aeruginosa -->
-
-<!-- ```{r} -->
-<!-- # Update formula -->
-<!-- cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa), -->
-<!--   data = s, cluster = eDWID -->
-<!-- ) -->
-<!-- # Forest plot and coefficients table -->
-<!-- ggforest(cox_res, data = s) -->
-<!-- ``` -->
-
-<!-- #### Aspergillus -->
-
-<!-- ```{r} -->
-<!-- # Update formula -->
-<!-- cox_res <- coxph(update(f, . ~ . + A_fungalyeast1), -->
-<!--   data = s, cluster = eDWID -->
-<!-- ) -->
-<!-- # Forest plot and coefficients table -->
-<!-- ggforest(cox_res, data = s) -->
-<!-- ``` -->
-
-<!-- #### ABPA -->
-
-<!-- ```{r} -->
-<!-- # Update formula -->
-<!-- cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1), -->
-<!--   data = s, cluster = eDWID -->
-<!-- ) -->
-<!-- # Forest plot and coefficients table -->
-<!-- ggforest(cox_res, data = s) -->
-<!-- ``` -->
-
-<!-- #### Mycobacterium -->
-
-<!-- ```{r} -->
-<!-- # Update formula -->
-<!-- cox_res <- coxph(update(f, . ~ . + A_mycocultureresults3), -->
-<!--   data = s, cluster = eDWID -->
-<!-- ) -->
-<!-- # Forest plot and coefficients table -->
-<!-- ggforest(cox_res, data = s) -->
-<!-- ``` -->
-
-<!-- t -->

### 2024-04-29 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
incidence tables updated

@@
 data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
-# Find each person's entry date
-annual = annual %>% group_by(eDWID) %>%
-  mutate(entry_date = min(ReviewYear))
 # Remove those with T1D or T2D
@@
 annual$A_IsOnEnzymes <- relevel(annual$A_IsOnEnzymes, ref = "Yes")
-# Approximate year of birth
-by <- annual %>%
-  mutate(birth_year = floor(ReviewYear - Age_YrEnd)) %>%
-  group_by(eDWID) %>%
-  count(birth_year) %>%
-  group_by(eDWID) %>%
-  slice_min(birth_year) %>%
-  select(-n)
-annual <- left_join(annual, by, by = join_by(eDWID))
 # BMI percentile to categorical
@@
 annual$CFRD <- factor(annual$CFRD, levels = c("CFRD-", "CFRD+"))
+# Add formatted columns for survival analysis
+annual$tstop <- ymd(paste0(annual$ReviewYear, "-12-31"))
+annual$tstart <- ymd(paste0(annual$ReviewYear, "-01-01"))
+annual <- annual %>%
+  group_by(eDWID) %>%
+  mutate(entry_date = tstart[1]) %>%
+  ungroup()
 # For each person, filter out all the years after first CFRD diagnosis
@@
 annual <- annual %>% filter(Age_YrEnd != 0, Age_YrEnd >= 10)
-annual$age_group = droplevels(annual$age_group)
-# Add formatted columns for survival analysis
-annual = annual %>% 
-  mutate(tstop = ReviewYear) %>%
-  group_by(eDWID) %>%
-  mutate(
-    tstart = lag(tstop, default = birth_year[1]),
-    age_at_entry = Age_YrEnd[1]
-  )
+annual$age_group <- droplevels(annual$age_group)
 # List of covariates
@@
   Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
-  ppFEV1_cat = "FEV1 percent predicted",
+  ppFEV1_cat = "FEV1 percent predicted category",
+  A_FEV1_pct_predicted = "FEV1 percent predicted",
+  A_FVC_pct_predicted = "FVC percent predicted",
   tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
@@
   A_fungalyeast1 = "Aspergillus (any species)",
-  A_mycocultureresults3 = "Microorganisms in myco culture",
-  eDWID = "ID", birth_year = "Birth year",
+  A_mycocultureresults3 = "Microorganisms in myco culture", eDWID = "ID",
   tstart = "Interval start", tstop = "Interval stop",
-  age_at_entry = "Age at entry", Age_YrEnd = "Age", bmi_perc = "BMI %ile"
+  Age_YrEnd = "Age", bmi_perc = "BMI %ile"
 )
+label(annual[, which(colnames(annual) %in% names(labels))]) <-
+  labels[colnames(annual)[which(colnames(annual) %in% names(labels))]]
 ```
@@
 
-```{r}
-t2004 = annual %>% select()
-```
+## Demographics of Incident CFRD cases versus non-CFRD cases in 2004, 2011, and 2018
 
 ```{r}
-# Use flexsurv to fit a survival model that can be passed to rprev
-llog <- flexsurv::flexsurvreg(Surv(tstop, as.numeric(CFRD)==2) ~ 1, 
-                              data=prevdata, dist='llogis')
-prev_llog <- prevalence(index='2004-01-01', 
-                        num_years_to_estimate=1, 
-                        data=annual, 
-                        inc_formula = entrydate ~ 1,
-                        surv_model=llog,
-                        death_column = "first_cfrd",
-                        incident_column = "first_cfrd",
-                        N_boot = 100)
+annual %>%
+  filter(ReviewYear %in% c(2004, 2011, 2018)) %>%
+  select(
+    ReviewYear, CFRD, Age_YrEnd, Gender, Race, Hispanicrace, bmi_perc,
+    bmi_perc_cat, MutClass, tube_feeding, liver_disease, A_FEV1_pct_predicted,
+    A_FVC_pct_predicted, NumPulmExacerbation, A_corticosteroids1
+  ) %>%
+  tbl_strata(
+    strata = ReviewYear,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(
+          by = CFRD, missing_text = "Missing",
+          type = list(A_corticosteroids1 ~ "categorical")
+        ) %>%
+        add_p() %>%
+        add_overall() %>%
+        separate_p_footnotes()
+  )
 ```
@@
 # Whole cohort
-km_fit <- survfit(Surv(tstart, tstop, as.numeric(CFRD)==2) ~ 1,
+og <- as.numeric(ymd("2003-01-01"))
+km_fit <- survfit(
+  Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
+    origin = og
+  ) ~ 1,
   data = annual, id = eDWID
@@
 ggsurvplot(km_fit,
-  data = annual, risk.table = T, cumevents = T,
-  fun = "event", xlim = c(2003, 2018), break.x.by = 1
+  data = annual, risk.table = T, cumevents = T, fun = "event",
+  xlab = "Days from 2003-01-01"
 )
-kable(tidy(km_fit))
+cum_inc <- tidy(km_fit)
+cum_inc$time <- as.Date(cum_inc$time + og)
+kable(cum_inc)
 # By age group
-km_fit <- survfit(Surv(tstart, tstop, as.numeric(CFRD)==2) ~ age_group,
+km_fit <- survfit(
+  Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
+    origin = og
+  ) ~ age_group,
   data = annual, id = eDWID
@@
 ggsurvplot(km_fit,
-  data = annual, risk.table = T, cumevents = T,
-  fun = "event", xlim = c(2003, 2018), break.x.by = 1
+  data = annual, risk.table = T, cumevents = T, fun = "event",
+  xlab = "Days from 2003-01-01"
 )
 km_table <- tidy(km_fit)
+km_table$time <- as.Date(km_table$time + og)
 km_table <- split.data.frame(km_table, km_table$strata)
+```
+
+```{r results='asis'}
 invisible(lapply(km_table, function(t) {
@@
   t <- t %>% select(-strata)
-  kable(t, caption = name)
+  print(kable(t, caption = name))
 }))
@@
 )
-y <- "Surv(tstart, tstop, as.numeric(CFRD)==2)"
+y <- "Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2)"
 f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
@@
 ```{r}
-annual_split <- survSplit(Surv(tstart, tstop, as.numeric(CFRD)==2) ~ .,
+annual_split <- survSplit(
+  Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
+    origin = og
+  ) ~ .,
   data = annual, cut = c(2008, 2013, 2018), episode = "tgroup"
@@
 )
+y <- "Surv(as.numeric(tstart), as.numeric(tstop), event)"
 ```

### 2024-05-01 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
gtsummary for Colleen

@@
 library(tidyverse)
-library(arsenal)
+library(gtsummary)
 library(rspiro)
@@
 )
+# Columns formats
+df$`Num. A1c Values` <- as.numeric(df$`Num. A1c Values`)
 ```
@@
 summary(t1)
+
+df %>%
+  select(
+    Age, `BMI Percentile`, Weight, Sex, ppFEV1, ppFEV1gl, ppFVC,
+    ppFVCgl, `G-tube in past 12 months`, `Pancreatic Status`, Site,
+    Genotype, `Age Group`, Race
+  ) %>%
+  tbl_summary(
+    statistic = list(c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ "{mean} ± {sd}"),
+    missing_text = "Missing"
+  )
 ```
@@
 summary(p, pfootnote = T, digits = 1)
+
+df %>%
+  select(
+    `CFF ID`, `Pre or Post`, `Num. A1c Values`, HbA1c, Time, Weight,
+    `BMI Percentile`, ppFEV1, ppFEV1gl, ppFVC, ppFVCgl, TDD,
+  ) %>%
+  tbl_summary(by = `Pre or Post`,type = list(`Num. A1c Values`~"continuous"),
+              include = -`CFF ID`) %>%
+  add_p(test = list(
+      c("Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile", "TDD") ~
+        "paired.wilcox.test",
+      c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ "paired.t.test"
+    ),
+    group = `CFF ID`)%>%
+  separate_p_footnotes()
 ```

### 2024-05-06 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
tbl_strata not working for some reason

--- 2024-04-17 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
@@
 library(tidyverse)
+library(gtsummary)
 library(arsenal)
+library(rspiro)
 home_dir <- switch(Sys.info()["sysname"],
@@
 load("./Colleen Wood/Prepost Trikafta/Data_Cleaned/analysis_dataset.rdata")
+# Baseline HbA1c, BMI group pre- and post-modulator
+df <- df %>%
+  group_by(`CFF ID`) %>%
+  mutate(
+    `Baseline HbA1c` = HbA1c[`Pre or Post` == "Pre"],
+    `Baseline BMI` = `BMI Percentile`[`Pre or Post` == "Pre"],
+    `Post BMI` = `BMI Percentile`[`Pre or Post` == "Post"]
+  ) %>%
+  ungroup()
+df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf), right = F)
+df$`Baseline BMI` <- cut(df$`Baseline BMI`, c(-Inf, 85, Inf), right = F)
+df$`Post BMI` <- cut(df$`Post BMI`, c(-Inf, 85, Inf), right = F)
+# Pre-modulator BMI percentile
 df <- df %>%
@@
 df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf), right = F)
+# Race neutral equations
+df$ppFEV1gl <- pctpred_GLIgl(
+  age = df$Age, height = df$Height / 100, gender = df$Sex, FEV1 = df$FEV1
+)
+df$ppFVCgl <- pctpred_GLIgl(
+  age = df$Age, height = df$Height / 100, gender = df$Sex, FVC = df$FEV1
+)
+# Columns formats
+df$`Num. A1c Values` <- as.numeric(df$`Num. A1c Values`)
 ```
@@
 
-```{r results='asis'}
-# Table
-t1 <- tableby(
-  ~ Age + `BMI Percentile` + Weight + Sex + ppFEV1 + ppFVC +
-    `G-tube in past 12 months` + `Pancreatic Status` + Site +
-    Genotype + `Age Group` + Race,
-  data = df %>% filter(`Pre or Post` == "Pre")
+```{r}
+df %>%
+  filter(`Pre or Post` == "Pre") %>%
+  select(
+    Age, `BMI Percentile`, Weight, Sex, ppFEV1, ppFEV1gl, ppFVC,
+    ppFVCgl, `G-tube in past 12 months`, `Pancreatic Status`, Site,
+    Genotype, `Age Group`, Race
+  ) %>%
+  tbl_summary(
+    statistic = list(c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ "{mean} ± {sd}"),
+    missing_text = "Missing"
   )
-summary(t1)
 ```
@@
 
-```{r results='asis'}
-p <- paired(
-  `Pre or Post` ~ `Num. A1c Values` + HbA1c + Time + Weight + `BMI Percentile` +
-    ppFEV1 + ppFVC + TDD + `Long acting insulin dose`,
-  data = df, id = `CFF ID`,
-  signed.rank.exact = FALSE,
-  numeric.stats = c("Nmiss", "medianrange", "q1q3")
-)
-summary(p, pfootnote = T, digits = 1)
+```{r warning=FALSE, message=FALSE}
+df %>%
+  select(
+    `CFF ID`, `Pre or Post`, `Num. A1c Values`, HbA1c, Time, Weight,
+    `BMI Percentile`, ppFEV1, ppFEV1gl, ppFVC, ppFVCgl, TDD,
+  ) %>%
+  tbl_summary(
+    by = `Pre or Post`, type = list(
+      c(
+        "Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile",
+        "ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl", "TDD"
+      ) ~ "continuous2"
+    ),
+    statistic = list(
+      c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~
+        c("{mean} ± {sd}", "{min}, {max}"),
+      c("Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile") ~
+        c("{median} [{p25}, {p75}]", "{min}, {max}")
+    ),
+    include = -`CFF ID`, digits = list(everything() ~ 2),
+    missing_text = "Missing"
+  ) %>%
+  add_p(
+    test = list(
+      c("Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile", "TDD") ~
+        "paired.wilcox.test",
+      c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ "paired.t.test"
+    ),
+    group = `CFF ID`
+  ) %>%
+  separate_p_footnotes()
 ```
@@
 
-```{r results='asis'}
-p <- paired(
-  `Pre or Post` ~ `Num. A1c Values` + HbA1c + Time + Weight + `BMI Percentile` +
-    ppFEV1 + ppFVC,
-  data = df, id = `CFF ID`, strata = `Baseline HbA1c`,
-  signed.rank.exact = FALSE,
-  numeric.stats = c("Nmiss", "medianrange", "q1q3")
+```{r warning=FALSE, message=FALSE}
+df %>%
+  select(
+    `CFF ID`, `Pre or Post`, `Baseline HbA1c`, `Num. A1c Values`, HbA1c, Time,
+    Weight, `BMI Percentile`, ppFEV1, ppFEV1gl, ppFVC, ppFVCgl, TDD,
+  ) %>%
+  tbl_strata(
+    strata = `Baseline HbA1c`,
+    .tbl_fun = ~ .x %>%
+      tbl_summary(
+        by = `Pre or Post`, type = list(
+          c(
+            "Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile",
+            "ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl", "TDD"
+          ) ~ "continuous2"
+        ),
+        statistic = list(
+          c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~
+            c("{mean} ± {sd}", "{min}, {max}"),
+          c("Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile") ~
+            c("{median} [{p25}, {p75}]", "{min}, {max}")
+        ),
+        include = -`CFF ID`, digits = list(everything() ~ 2),
+        missing_text = "Missing"
+      ) %>%
+      add_p(
+        test = list(
+          c("Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile", "TDD") ~
+            "paired.wilcox.test",
+          c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ "paired.t.test"
+        ),
+        group = `CFF ID`
+      ) %>%
+      separate_p_footnotes()
   )
-summary(p, pfootnote = T, digits = 1)
 ```
@@
 
-```{r results='asis'}
-p <- paired(
-  `Pre or Post` ~ `Num. A1c Values` + HbA1c + Time + Weight + `BMI Percentile` +
-    ppFEV1 + ppFVC + TDD,
-  data = df, id = `CFF ID`, strata = Site,
-  signed.rank.exact = FALSE, numeric.test = "notest",
-  numeric.stats = c("Nmiss", "medianrange", "q1q3")
+```{r warning=FALSE, message=FALSE}
+df %>%
+  select(
+    `CFF ID`, `Pre or Post`, Site, `Num. A1c Values`, HbA1c, Time,
+    Weight, `BMI Percentile`, ppFEV1, ppFEV1gl, ppFVC, ppFVCgl, TDD,
+  ) %>%
+  tbl_strata(
+    strata = Site,
+    .tbl_fun = ~ .x %>%
+      tbl_summary(
+        by = `Pre or Post`, type = list(
+          c(
+            "Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile",
+            "ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl", "TDD"
+          ) ~ "continuous2"
+        ),
+        statistic = list(
+          c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~
+            c("{mean} ± {sd}", "{min}, {max}"),
+          c("Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile") ~
+            c("{median} [{p25}, {p75}]", "{min}, {max}")
+        ),
+        include = -`CFF ID`, digits = list(everything() ~ 2),
+        missing_text = "Missing"
+      ) %>%
+      add_p(
+        test = list(
+          c("Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile", "TDD") ~
+            "paired.wilcox.test",
+          c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ "paired.t.test"
+        ),
+        group = `CFF ID`
+      ) %>%
+      separate_p_footnotes()
   )
-summary(p, pfootnote = T, digits = 1)
+```
+
+### By BMI percentile pre-modulator
+
+```{r warning=FALSE, message=FALSE}
+df %>%
+  select(
+    `CFF ID`, `Pre or Post`, `Baseline BMI`, `Num. A1c Values`, HbA1c, Time,
+    Weight, `BMI Percentile`, ppFEV1, ppFEV1gl, ppFVC, ppFVCgl, TDD,
+  ) %>%
+  tbl_strata(
+    strata = `Baseline BMI`,
+    .tbl_fun = ~ .x %>%
+      tbl_summary(
+        by = `Pre or Post`, type = list(
+          c(
+            "Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile",
+            "ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl", "TDD"
+          ) ~ "continuous2"
+        ),
+        statistic = list(
+          c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~
+            c("{mean} ± {sd}", "{min}, {max}"),
+          c("Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile") ~
+            c("{median} [{p25}, {p75}]", "{min}, {max}")
+        ),
+        include = -`CFF ID`, digits = list(everything() ~ 2),
+        missing_text = "Missing"
+      ) %>%
+      add_p(
+        test = list(
+          c("Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile", "TDD") ~
+            "paired.wilcox.test",
+          c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ "paired.t.test"
+        ),
+        group = `CFF ID`
+      ) %>%
+      separate_p_footnotes()
+  )
+```
+
+### By BMI percentile post-modulator
+
+```{r warning=FALSE, message=FALSE}
+df %>%
+  select(
+    `CFF ID`, `Pre or Post`, `Post BMI`, `Num. A1c Values`, HbA1c, Time,
+    Weight, `BMI Percentile`, ppFEV1, ppFEV1gl, ppFVC, ppFVCgl, TDD,
+  ) %>%
+  filter(!is.na(`Post BMI`)) %>%
+  tbl_strata(
+    strata = `Post BMI`,
+    .tbl_fun = ~ .x %>%
+      tbl_summary(
+        by = `Pre or Post`,
+        type = list(
+          c(
+            "Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile",
+            "ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl", "TDD"
+          ) ~ "continuous2"
+        ),
+        statistic = list(
+          c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~
+            c("{mean} ± {sd}", "{min}, {max}"),
+          c("Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile") ~
+            c("{median} [{p25}, {p75}]", "{min}, {max}")
+        ),
+        include = -`CFF ID`, digits = list(everything() ~ 2),
+        missing_text = "Missing"
+      ) %>%
+      add_p(
+        test = list(
+          c(
+            "Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile",
+            "TDD"
+          ) ~ "paired.wilcox.test",
+          c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ "paired.t.test"
+        ),
+        group = `CFF ID`
+      ) %>%
+      separate_p_footnotes()
+  )
 ```
@@
 
-```{r results='asis'}
+```{r warning=FALSE, message=FALSE}
 cgm <- df[!is.na(df$average_sensor), ]
-# Table
-t1 <- tableby(
-  ~ Age + `BMI Percentile` + Weight + Sex + ppFEV1 + ppFVC +
-    `G-tube in past 12 months` + `Pancreatic Status` + Site +
-    Genotype + `Age Group` + Race,
-  data = cgm %>% filter(`Pre or Post` == "Pre")
+cgm %>%
+  filter(`Pre or Post` == "Pre") %>%
+  select(
+    Age, `BMI Percentile`, Weight, Sex, ppFEV1, ppFEV1gl, ppFVC,
+    ppFVCgl, `G-tube in past 12 months`, `Pancreatic Status`, Site,
+    Genotype, `Age Group`, Race
+  ) %>%
+  tbl_summary(
+    type = list(
+      c(
+        "Age", "BMI Percentile", "Weight", "ppFEV1", "ppFEV1gl", "ppFVC",
+        "ppFVCgl"
+      ) ~ "continuous2"
+    ),
+    statistic = list(all_continuous2() ~
+      c("{median} [{p25}, {p75}]", "{min}, {max}")),
+    digits = list(everything() ~ 2)
   )
-summary(t1)
 ```
@@
 cgm_formula <- update(cgm_formula, . ~ . + TDD + `Long acting insulin dose`)
-p <- paired(cgm_formula,
-  data = cgm, id = `CFF ID`,
-  signed.rank.exact = FALSE, numeric.test = "signed.rank",
-  numeric.stats = c("Nmiss", "medianrange", "q1q3")
-)
-summary(p, pfootnote = T, digits = 1)
+cgm_vars2 <- c(cgm_vars, "TDD", "Long acting insulin dose")
+
+cgm %>%
+  select(`CFF ID`, `Pre or Post`, all_of(cgm_vars2)) %>%
+  tbl_summary(
+    by = `Pre or Post`,
+    type = list(everything() ~ "continuous2"),
+    statistic = list(
+      everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
+    ),
+    include = -`CFF ID`, digits = list(everything() ~ 2),
+    missing_text = "Missing"
+  ) %>%
+  add_p(
+    test = list(everything() ~ "paired.wilcox.test"),
+    group = `CFF ID`
+  ) %>%
+  separate_p_footnotes()
 ```
@@
 summary(p, pfootnote = T, digits = 1)
+
+cgm %>%
+  select(`CFF ID`, `Pre or Post`, `Baseline HbA1c`, all_of(cgm_vars2)) %>%
+  tbl_strata(
+    by = `Baseline HbA1c`,
+    .tbl_fun = ~ .x %>%
+      tbl_summary(
+        by = `Pre or Post`,
+        type = list(everything() ~ "continuous2"),
+        statistic = list(
+          everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
+        ),
+        include = -`CFF ID`, digits = list(everything() ~ 2),
+        missing_text = "Missing"
+      ) %>%
+      add_p(
+        test = list(everything() ~ "paired.wilcox.test"),
+        group = `CFF ID`
+      ) %>%
+      separate_p_footnotes()
+  )
 ```

### 2024-05-06 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
stratification done, need change plots

@@
 )
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
 knitr::opts_knit$set(root.dir = home_dir)
@@
 
-```{r warning=FALSE, message=FALSE}
+```{r}
 df %>%
@@
 
-### Stratified by baseline HbA1c
+### By site
 
-```{r warning=FALSE, message=FALSE}
+```{r}
+df %>%
+  select(
+    `CFF ID`, `Pre or Post`, Site, `Num. A1c Values`, HbA1c, Time,
+    Weight, `BMI Percentile`, ppFEV1, ppFEV1gl, ppFVC, ppFVCgl, TDD,
+  ) %>%
+  tbl_strata(
+    strata = Site,
+    .tbl_fun = ~ .x %>%
+      tbl_summary(
+        by = `Pre or Post`, type = list(
+          c(
+            "Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile",
+            "ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl", "TDD"
+          ) ~ "continuous2"
+        ),
+        statistic = list(
+          c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~
+            c("{mean} ± {sd}", "{min}, {max}"),
+          c("Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile") ~
+            c("{median} [{p25}, {p75}]", "{min}, {max}")
+        ),
+        include = -`CFF ID`, digits = list(everything() ~ 2),
+        missing_text = "Missing"
+      ) %>%
+      add_p(
+        test = list(
+          c("Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile", "TDD") ~
+            "paired.wilcox.test",
+          c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ "paired.t.test"
+        ),
+        group = `CFF ID`
+      ) %>%
+      separate_p_footnotes()
+  )
+```
+
+### By baseline HbA1c
+
+```{r}
 df %>%
@@
 
-### By site
-
-```{r warning=FALSE, message=FALSE}
-df %>%
-  select(
-    `CFF ID`, `Pre or Post`, Site, `Num. A1c Values`, HbA1c, Time,
-    Weight, `BMI Percentile`, ppFEV1, ppFEV1gl, ppFVC, ppFVCgl, TDD,
-  ) %>%
-  tbl_strata(
-    strata = Site,
-    .tbl_fun = ~ .x %>%
-      tbl_summary(
-        by = `Pre or Post`, type = list(
-          c(
-            "Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile",
-            "ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl", "TDD"
-          ) ~ "continuous2"
-        ),
-        statistic = list(
-          c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~
-            c("{mean} ± {sd}", "{min}, {max}"),
-          c("Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile") ~
-            c("{median} [{p25}, {p75}]", "{min}, {max}")
-        ),
-        include = -`CFF ID`, digits = list(everything() ~ 2),
-        missing_text = "Missing"
-      ) %>%
-      add_p(
-        test = list(
-          c("Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile", "TDD") ~
-            "paired.wilcox.test",
-          c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ "paired.t.test"
-        ),
-        group = `CFF ID`
-      ) %>%
-      separate_p_footnotes()
-  )
-```
-
 ### By BMI percentile pre-modulator
 
-```{r warning=FALSE, message=FALSE}
+```{r}
 df %>%
@@
 
-```{r warning=FALSE, message=FALSE}
+```{r}
 cgm <- df[!is.na(df$average_sensor), ]
@@
 
-```{r results='asis'}
-cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(cgm_vars, collapse = "+")))
-cgm_formula <- update(cgm_formula, . ~ . + TDD + `Long acting insulin dose`)
-cgm_vars2 <- c(cgm_vars, "TDD", "Long acting insulin dose")
-
+```{r}
+cgm_vars <- c(cgm_vars, "TDD", "Long acting insulin dose")
 cgm %>%
-  select(`CFF ID`, `Pre or Post`, all_of(cgm_vars2)) %>%
+  select(`CFF ID`, `Pre or Post`, all_of(cgm_vars)) %>%
   tbl_summary(
@@
 
-#### Stratified by baseline HbA1c
+#### By baseline HbA1c
 
-```{r results='asis'}
-cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(cgm_vars, collapse = "+")))
-cgm_formula <- update(cgm_formula, . ~ . + TDD + `Long acting insulin dose`)
-p <- paired(cgm_formula,
-  data = cgm, id = `CFF ID`, strata = `Baseline HbA1c`,
-  signed.rank.exact = FALSE, numeric.test = "signed.rank",
-  numeric.stats = c("Nmiss", "medianrange", "q1q3")
+```{r}
+cgm %>%
+  select(
+    `CFF ID`, `Pre or Post`, `Baseline HbA1c`, all_of(cgm_vars)
+  ) %>%
+  tbl_strata(
+    strata = `Baseline HbA1c`,
+    .tbl_fun = ~ .x %>%
+      tbl_summary(
+        by = `Pre or Post`,
+        type = list(everything() ~ "continuous2"),
+        statistic = list(
+          everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
+        ),
+        include = -`CFF ID`, digits = list(everything() ~ 2),
+        missing_text = "Missing"
+      ) %>%
+      add_p(
+        test = list(everything() ~ "paired.wilcox.test"),
+        group = `CFF ID`
+      ) %>%
+      separate_p_footnotes()
   )
-summary(p, pfootnote = T, digits = 1)
+```
 
+#### By BMI percentile pre-modulator
+
+```{r}
 cgm %>%
-  select(`CFF ID`, `Pre or Post`, `Baseline HbA1c`, all_of(cgm_vars2)) %>%
+  select(
+    `CFF ID`, `Pre or Post`, `Baseline BMI`, all_of(cgm_vars)
+  ) %>%
   tbl_strata(
-    by = `Baseline HbA1c`,
+    strata = `Baseline BMI`,
+    .tbl_fun = ~ .x %>%
+      tbl_summary(
+        by = `Pre or Post`,
+        type = list(everything() ~ "continuous2"),
+        statistic = list(
+          everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
+        ),
+        include = -`CFF ID`, digits = list(everything() ~ 2),
+        missing_text = "Missing"
+      ) %>%
+      add_p(
+        test = list(everything() ~ "paired.wilcox.test"),
+        group = `CFF ID`
+      ) %>%
+      separate_p_footnotes()
+  )
+```
+
+#### By BMI percentile post-modulator
+
+```{r}
+cgm %>%
+  select(
+    `CFF ID`, `Pre or Post`, `Post BMI`, all_of(cgm_vars)
+  ) %>%
+  tbl_strata(
+    strata = `Post BMI`,
     .tbl_fun = ~ .x %>%
@@
 
-```{r results='asis'}
+```{r}
 cgm$daytime_cv <- cgm$daytime_sd / cgm$daytime_avg_sensor_glucose
 day_cgm <- c(day_cgm, "daytime_cv")
-cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(day_cgm, collapse = "+")))
-p <- paired(cgm_formula,
-  data = cgm, id = `CFF ID`,
-  signed.rank.exact = FALSE, numeric.test = "signed.rank",
-  numeric.stats = c("Nmiss", "medianrange", "q1q3")
-)
-summary(p, pfootnote = T, digits = 1)
+
+cgm %>%
+  select(`CFF ID`, `Pre or Post`, all_of(day_cgm)) %>%
+  tbl_summary(
+    by = `Pre or Post`,
+    type = list(everything() ~ "continuous2"),
+    statistic = list(
+      everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
+    ),
+    include = -`CFF ID`, digits = list(everything() ~ 2),
+    missing_text = "Missing"
+  ) %>%
+  add_p(
+    test = list(everything() ~ "paired.wilcox.test"),
+    group = `CFF ID`
+  ) %>%
+  separate_p_footnotes()
 ```
 
-#### Stratified by baseline HbA1c
+#### By baseline HbA1c
 
-```{r results='asis'}
-cgm$daytime_cv <- cgm$daytime_sd / cgm$daytime_avg_sensor_glucose
-day_cgm <- c(day_cgm, "daytime_cv")
-cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(day_cgm, collapse = "+")))
-p <- paired(cgm_formula,
-  data = cgm, id = `CFF ID`, strata = `Baseline HbA1c`,
-  signed.rank.exact = FALSE, numeric.test = "signed.rank",
-  numeric.stats = c("Nmiss", "medianrange", "q1q3")
+```{r}
+cgm %>%
+  select(
+    `CFF ID`, `Pre or Post`, `Baseline HbA1c`, all_of(day_cgm)
+  ) %>%
+  tbl_strata(
+    strata = `Baseline HbA1c`,
+    .tbl_fun = ~ .x %>%
+      tbl_summary(
+        by = `Pre or Post`,
+        type = list(everything() ~ "continuous2"),
+        statistic = list(
+          everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
+        ),
+        include = -`CFF ID`, digits = list(everything() ~ 2),
+        missing_text = "Missing"
+      ) %>%
+      add_p(
+        test = list(everything() ~ "paired.wilcox.test"),
+        group = `CFF ID`
+      ) %>%
+      separate_p_footnotes()
   )
-summary(p, pfootnote = T, digits = 1)
+```
+
+#### By BMI percentile pre-modulator
+
+```{r}
+cgm %>%
+  select(
+    `CFF ID`, `Pre or Post`, `Baseline BMI`, all_of(day_cgm)
+  ) %>%
+  tbl_strata(
+    strata = `Baseline BMI`,
+    .tbl_fun = ~ .x %>%
+      tbl_summary(
+        by = `Pre or Post`,
+        type = list(everything() ~ "continuous2"),
+        statistic = list(
+          everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
+        ),
+        include = -`CFF ID`, digits = list(everything() ~ 2),
+        missing_text = "Missing"
+      ) %>%
+      add_p(
+        test = list(everything() ~ "paired.wilcox.test"),
+        group = `CFF ID`
+      ) %>%
+      separate_p_footnotes()
+  )
+```
+
+#### By BMI percentile post-modulator
+
+```{r}
+cgm %>%
+  select(
+    `CFF ID`, `Pre or Post`, `Post BMI`, all_of(day_cgm)
+  ) %>%
+  tbl_strata(
+    strata = `Post BMI`,
+    .tbl_fun = ~ .x %>%
+      tbl_summary(
+        by = `Pre or Post`,
+        type = list(everything() ~ "continuous2"),
+        statistic = list(
+          everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
+        ),
+        include = -`CFF ID`, digits = list(everything() ~ 2),
+        missing_text = "Missing"
+      ) %>%
+      add_p(
+        test = list(everything() ~ "paired.wilcox.test"),
+        group = `CFF ID`
+      ) %>%
+      separate_p_footnotes()
+  )
 ```
@@
 
-```{r results='asis'}
+```{r}
 cgm$nighttime_cv <- cgm$nighttime_sd / cgm$nighttime_avg_sens_glucose
 night_cgm <- c(night_cgm, "nighttime_cv")
-cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(night_cgm, collapse = "+")))
-p <- paired(cgm_formula,
-  data = cgm, id = `CFF ID`,
-  signed.rank.exact = FALSE, numeric.test = "signed.rank",
-  numeric.stats = c("Nmiss", "medianrange", "q1q3")
-)
-summary(p, pfootnote = T, digits = 1)
+
+cgm %>%
+  select(`CFF ID`, `Pre or Post`, all_of(night_cgm)) %>%
+  tbl_summary(
+    by = `Pre or Post`,
+    type = list(everything() ~ "continuous2"),
+    statistic = list(
+      everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
+    ),
+    include = -`CFF ID`, digits = list(everything() ~ 2),
+    missing_text = "Missing"
+  ) %>%
+  add_p(
+    test = list(everything() ~ "paired.wilcox.test"),
+    group = `CFF ID`
+  ) %>%
+  separate_p_footnotes()
 ```
 
-#### Stratified by baseline HbA1c
+#### By baseline HbA1c
 
-```{r results='asis'}
-cgm$nighttime_cv <- cgm$nighttime_sd / cgm$nighttime_avg_sens_glucose
-night_cgm <- c(night_cgm, "nighttime_cv")
-cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(night_cgm, collapse = "+")))
-p <- paired(cgm_formula,
-  data = cgm, id = `CFF ID`, strata = `Baseline HbA1c`,
-  signed.rank.exact = FALSE, numeric.test = "signed.rank",
-  numeric.stats = c("Nmiss", "medianrange", "q1q3")
+```{r}
+cgm %>%
+  select(
+    `CFF ID`, `Pre or Post`, `Baseline HbA1c`, all_of(night_cgm)
+  ) %>%
+  tbl_strata(
+    strata = `Baseline HbA1c`,
+    .tbl_fun = ~ .x %>%
+      tbl_summary(
+        by = `Pre or Post`,
+        type = list(everything() ~ "continuous2"),
+        statistic = list(
+          everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
+        ),
+        include = -`CFF ID`, digits = list(everything() ~ 2),
+        missing_text = "Missing"
+      ) %>%
+      add_p(
+        test = list(everything() ~ "paired.wilcox.test"),
+        group = `CFF ID`
+      ) %>%
+      separate_p_footnotes()
   )
-summary(p, pfootnote = T, digits = 1)
 ```
+
+#### By BMI percentile pre-modulator
+
+```{r}
+cgm %>%
+  select(
+    `CFF ID`, `Pre or Post`, `Baseline BMI`, all_of(night_cgm)
+  ) %>%
+  tbl_strata(
+    strata = `Baseline BMI`,
+    .tbl_fun = ~ .x %>%
+      tbl_summary(
+        by = `Pre or Post`,
+        type = list(everything() ~ "continuous2"),
+        statistic = list(
+          everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
+        ),
+        include = -`CFF ID`, digits = list(everything() ~ 2),
+        missing_text = "Missing"
+      ) %>%
+      add_p(
+        test = list(everything() ~ "paired.wilcox.test"),
+        group = `CFF ID`
+      ) %>%
+      separate_p_footnotes()
+  )
+```
+
+#### By BMI percentile post-modulator
+
+```{r}
+cgm %>%
+  select(
+    `CFF ID`, `Pre or Post`, `Post BMI`, all_of(night_cgm)
+  ) %>%
+  tbl_strata(
+    strata = `Post BMI`,
+    .tbl_fun = ~ .x %>%
+      tbl_summary(
+        by = `Pre or Post`,
+        type = list(everything() ~ "continuous2"),
+        statistic = list(
+          everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
+        ),
+        include = -`CFF ID`, digits = list(everything() ~ 2),
+        missing_text = "Missing"
+      ) %>%
+      add_p(
+        test = list(everything() ~ "paired.wilcox.test"),
+        group = `CFF ID`
+      ) %>%
+      separate_p_footnotes()
+  )
+```

### 2024-05-06 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
base plots for colleen done

@@
 ```{r}
+# Table
 df %>%
@@
   separate_p_footnotes()
+# Plot
+ggplot(df, aes(x = `Pre or Post`, y = HbA1c, group = `CFF ID`)) +
+  geom_point(alpha = 0.2) +
+  geom_line(alpha = 0.2) +
+  theme_classic() +
+  theme(axis.title.x = element_blank())
 ```
@@
       tbl_summary(
-        by = `Pre or Post`, type = list(
+        by = `Pre or Post`,
+        type = list(
           c(
@@
         test = list(
-          c("Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile", "TDD") ~
-            "paired.wilcox.test",
+          c(
+            "Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile",
+            "TDD"
+          ) ~ "paired.wilcox.test",
           c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ "paired.t.test"
@@
   )
+# Plot
+ggplot(df, aes(x = `Pre or Post`, y = HbA1c, group = `CFF ID`)) +
+  geom_point(alpha = 0.2) +
+  geom_line(alpha = 0.2) +
+  theme_classic() +
+  theme(axis.title.x = element_blank()) +
+  facet_wrap(~Site)
 ```
@@
   )
+# Plot
+ggplot(df, aes(x = `Pre or Post`, y = HbA1c, group = `CFF ID`)) +
+  geom_point(alpha = 0.2) +
+  geom_line(alpha = 0.2) +
+  theme_classic() +
+  theme(axis.title.x = element_blank()) +
+  facet_wrap(~`Baseline HbA1c`)
 ```
@@
   )
+# Plot
+ggplot(df, aes(x = `Pre or Post`, y = HbA1c, group = `CFF ID`)) +
+  geom_point(alpha = 0.2) +
+  geom_line(alpha = 0.2) +
+  theme_classic() +
+  theme(axis.title.x = element_blank()) +
+  facet_wrap(~`Baseline BMI`)
 ```
@@
   )
+# Plot
+ggplot(
+  df[!is.na(df$`Post BMI`), ],
+  aes(x = `Pre or Post`, y = HbA1c, group = `CFF ID`)
+) +
+  geom_point(alpha = 0.2) +
+  geom_line(alpha = 0.2) +
+  theme_classic() +
+  theme(axis.title.x = element_blank()) +
+  facet_wrap(~`Post BMI`)
 ```

### 2024-05-07 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
tables and plots formatted

@@
 
-## Demographics of Incident CFRD cases versus non-CFRD cases in 2004, 2011, and 2018
+## Demographics of incident CFRD cases versus non-CFRD cases in 2004, 2011, and 2018
 
 ```{r}
+#| message: false
 annual %>%
@@
 
-## Kaplan Meier Analysis
+## Incidence and prevalence
+
+### Full cohort
 
 ```{r}
-# Whole cohort
-og <- as.numeric(ymd("2003-01-01"))
+#| message: false
+og <- ymd("2003-01-01")
 km_fit <- survfit(
   Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
-    origin = og
+    origin = as.numeric(og)
   ) ~ 1,
@@
 )
-ggsurvplot(km_fit,
-  data = annual, risk.table = T, cumevents = T, fun = "event",
-  xlab = "Days from 2003-01-01"
-)
-cum_inc <- tidy(km_fit)
-cum_inc$time <- as.Date(cum_inc$time + og)
-kable(cum_inc)
-# By age group
-km_fit <- survfit(
+# Plot
+surv_p <- ggsurvplot(km_fit, data = annual, fun = "event")
+surv_p$plot +
+  scale_x_continuous(
+    breaks = km_fit$time,
+    labels = year(ymd("2003-01-01") + days(km_fit$time))
+  ) +
+  theme(axis.title.x = element_blank(), legend.position = "none") +
+  ggtitle("Cumulative Incidence")
+# Table
+km_table <- tidy(km_fit)
+km_table$time <- year(og + days(km_table$time))
+gt(km_table) %>%
+  cols_hide("std.error") %>%
+  fmt_percent(columns = c("estimate", "conf.low", "conf.high")) %>%
+  cols_merge(
+    columns = c("estimate", "conf.low", "conf.high"),
+    pattern = "{1} [{2} - {3}]"
+  ) %>%
+  cols_label(
+    time = "Year", n.risk = "At Risk", n.event = "CFRD+",
+    n.censor = "Censored", estimate = "Estimated CFRD- [95% CI]"
+  ) %>%
+  opt_row_striping() %>%
+  opt_table_lines() %>%
+  opt_stylize(style = 6, color = "gray")
+```
+
+### By age group
+
+```{r}
+#| message: false
+# Plot
+km_age_fit <- survfit(
   Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
-    origin = og
+    origin = as.numeric(og)
   ) ~ age_group,
@@
 )
-ggsurvplot(km_fit,
-  data = annual, risk.table = T, cumevents = T, fun = "event",
-  xlab = "Days from 2003-01-01"
-)
-km_table <- tidy(km_fit)
-km_table$time <- as.Date(km_table$time + og)
+surv_p <- ggsurvplot(km_age_fit, data = annual, fun = "event")
+surv_p$plot +
+  scale_x_continuous(
+    breaks = km_age_fit$time,
+    labels = year(ymd("2003-01-01") + days(km_age_fit$time))
+  ) +
+  scale_color_discrete(name = "Age Group", labels = c("10-17", "18-29", "30+")) +
+  theme(axis.title.x = element_blank(), legend.title = element_blank()) +
+  ggtitle("Cumulative Incidence by Age Group")
+# Table
+km_table <- tidy(km_age_fit)
+km_table$time <- year(og + days(km_table$time))
 km_table <- split.data.frame(km_table, km_table$strata)
+
+invisible(lapply(km_table, function(km) {
+  title <- km$strata[1]
+  title <- sub("age_group=", "Ages ", title)
+  gt(km) %>%
+    cols_hide(c("std.error", "strata")) %>%
+    fmt_percent(columns = c("estimate", "conf.low", "conf.high")) %>%
+    cols_merge(
+      columns = c("estimate", "conf.low", "conf.high"),
+      pattern = "{1} [{2} - {3}]"
+    ) %>%
+    cols_label(
+      time = "Year", n.risk = "At Risk", n.event = "CFRD+",
+      n.censor = "Censored", estimate = "Estimated CFRD- [95% CI]"
+    ) %>%
+    opt_row_striping() %>%
+    opt_table_lines() %>%
+    opt_stylize(style = 6, color = "gray") %>%
+    tab_caption(title)
+}))
 ```
 
-```{r results='asis'}
-invisible(lapply(km_table, function(t) {
-  name <- t$strata[1]
-  t <- t %>% select(-strata)
-  print(kable(t, caption = name))
-}))
+## Screening rates
+
+```{r}
+screening <- annual %>%
+  group_by(ReviewYear, age_group) %>%
+  summarise(
+    `Screened HbA1c` = mean(!is.na(A_hgba1c)),
+    `Screened OGTT` = mean(!is.na(A_twohour_bloodglucose) |
+      !is.na(A_ogttfast_bloodglucose)),
+    `Screened Either` = mean(!is.na(A_twohour_bloodglucose) |
+      !is.na(A_ogttfast_bloodglucose) | !is.na(A_hgba1c)),
+    .groups = "drop"
+  ) %>%
+  pivot_longer(`Screened HbA1c`:`Screened Either`)
+ggplot(screening, aes(x = ReviewYear, y = value, color = age_group)) +
+  geom_line() +
+  facet_wrap(~name) +
+  theme_bw() +
+  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
+  scale_color_discrete(name = "Age Group", labels = c("10-17", "18-29", "30+"))
 ```
@@
   Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
-    origin = og
+    origin = as.numeric(og)
   ) ~ .,

### 2024-05-07 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
survival analysis mostly finalized

@@
 
-# CFRD screening rates
+# Incidence
 
-# Survival analysis - first CFRD diagnosis
-
-Based on CFFPR annualized data.
-These data are interval censored because we know the first visit at which people with CFRD were diagnosed, but we do not know the exact time that they developed CFRD.
-These data are also left truncated because not all participants enter the study at the same time.
-
-## Incidence and prevalence
-
-### Full cohort
+## Full cohort
 
@@
 
-### By age group
+## By age group
 
@@
 
-## Screening rates
+# Screening rates
+
+## By age group
 
@@
 ```
+
+## Who is getting OGTTs?
+
+```{r}
+# How many OGTTs are non-CFRD people getting 2016-2018
+ogtts <- annual %>%
+  filter(ReviewYear %in% 2016:2018, CFRD == "CFRD-") %>%
+  group_by(eDWID) %>%
+  summarise(
+    `Age Group` = last(na.omit(age_group)),
+    `OGTTs` = sum(!is.na(A_twohour_bloodglucose) | !is.na(A_ogttfast_bloodglucose))
+  )
+# Add demographics
+ogtts$eDWID <- as.numeric(ogtts$eDWID)
+ogtts <- left_join(ogtts, demogs, by = join_by(eDWID))
+label(ogtts$eDWID) <- "ID"
+# Format OGTT counts
+ogtts$OGTTs <- factor(ogtts$OGTTs,
+  levels = 0:3, labels = c(
+    "No OGTT", "At least 1 OGTT", "At least 1 OGTT", "At least 1 OGTT"
+  )
+)
+# Compare no OGTT to at least 1 OGTT
+ogtts %>% 
+  select(-eDWID) %>%
+  tbl_summary(by=OGTTs,missing_text = "Missing") %>%
+  add_p()
+```
+
+# Survival analysis - first CFRD diagnosis
+
+Based on CFFPR annualized data.
+These data are interval censored because we know the first visit at which people with CFRD were diagnosed, but we do not know the exact time that they developed CFRD.
+These data are also left truncated because not all participants enter the study at the same time.
 
@@
 ```{r}
+annual$tstart <- as.numeric(annual$tstart)
+annual$tstop <- as.numeric(annual$tstop)
 annual_split <- survSplit(
@@
   ) ~ .,
-  data = annual, cut = c(2008, 2013, 2018), episode = "tgroup"
+  data = annual,
+  cut = c(
+    as.numeric(ymd("2008-12-31")) - as.numeric(og),
+    as.numeric(ymd("2013-12-31")) - as.numeric(og)
+  ),
+  episode = "tgroup"
 )
@@
 )
-y <- "Surv(as.numeric(tstart), as.numeric(tstop), event)"
+y <- "Surv(tstart, tstop, event)"
 ```
@@
   ))
-  new_f_int <- as.formula(paste0(
-    y, "~", paste0(xvars[-which(xvars == v)], collapse = "+"),
-    paste0("+", v, "*tgroup")
-  ))
   new_f_emm <- as.formula(paste0(
@@
   cox_res_split <- coxph(new_f, data = annual_split, cluster = eDWID)
-  cox_res_split_int <- coxph(new_f_int, data = annual_split, cluster = eDWID)
   # Use the regular interaction term to get means for each time period
@@
   cat("\n")
-  emm <- emmeans(cox_res_split_int, new_f_emm,
-    nuisance = xvars[-which(xvars == v)]
-  )
+  emm <- emmeans(cox_res_split, new_f_emm, nuisance = xvars[-which(xvars == v)])
   emm <- data.frame(emm)

### 2024-05-10 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
separate analysis for outcomes

---
title: "Racial and Ethnic Differences in ETI: Outcomes"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    number-sections: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
# Libraries
library(nlme)
library(performance)
library(segmented)
library(tidyverse)
library(mcp)
options(mc.cores = 4)
set.seed(1017)
# Paths
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data setup}
#| include: false
# Import
encounter <- read.csv("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_encountersMerged_Del1.csv", na.strings = "")
annual <- read.csv("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_Annualized_Del1.csv", na.strings = "")
demo <- read.csv("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_DemogCFDiag_Del1.csv", na.strings = "")
# Race variable
races <- list(
  "Race1" = "White",
  "Race2" = "Black or African American",
  "Race3" = "American Indian or Alaska Native",
  "Race4" = "Asian",
  "Race5" = "Native Hawaiian or Other Pacific Islander",
  "Race6" = "Some other race"
)
demo$Race <- apply(demo[, paste0("Race", 1:6)], 1, function(r) {
  w <- which(r == 1)
  if (length(w) > 1 | 6 %in% w) {
    return("Mixed/Other/Unknown Race")
  } else {
    return(as.character(races[w]))
  }
})
demo$Race[demo$Hispanicrace == 1] <- "Hispanic or Latino"
demo$Race[is.na(demo$Hispanicrace)] <- "Mixed/Other/Unknown Race"
demo$Race <- factor(demo$Race,
  levels = c(
    "White", "Black or African American", "Hispanic or Latino", "Asian",
    "American Indian or Alaska Native",
    "Native Hawaiian or Other Pacific Islander", "Mixed/Other/Unknown Race"
  ),
  labels = c(
    "Non-Hispanic White", "Black or African American", "Hispanic or Latino", "Asian",
    "Other", "Other", "Other"
  )
)
# Insurance
ins_vars <- c(paste0("insurance", 1:7), "noinsurance")
ins <- c(
  "Health Insurance Policy (e.g. Private Insurance)", "Medicare", "Medicaid",
  "State special needs program, e.g., BCMH, CCS, CRS, GHPP, etc.",
  "TriCare or other military health plan",
  "Indian Health Service", "Other", "No Insurance"
)
annual$Insurance <- apply(annual[, ins_vars], 1, function(r) {
  w <- which(r == 1)
  if (length(w) == 0) {
    return(NA)
  } else if (length(w) > 1) {
    return("More than one")
  } else {
    return(ins[w])
  }
})
annual$Insurance <- factor(annual$Insurance,
  levels = c(
    "Health Insurance Policy (e.g. Private Insurance)",
    "Indian Health Service",
    "Medicaid",
    "Medicare",
    "More than one",
    "No Insurance",
    "Other",
    "State special needs program, e.g., BCMH, CCS, CRS, GHPP, etc.",
    "TriCare or other military health plan"
  ),
  labels = c(
    "Private", "Public", "Public", "Public", "More Than One", "No Insurance",
    "No Insurance", "Public", "Tricare"
  )
)
# Merge
encounter <- left_join(
  encounter,
  demo %>% select(eDWID, Race, Modulator_trikafta_first_date)
)
encounter <- left_join(encounter,
  annual %>% select(eDWID, ReviewYear, Insurance),
  by = join_by(eDWID, reviewyear == ReviewYear)
)
# Age group
encounter$age_group <- cut(encounter$encounterage, breaks = c(6, 12, 18, 30, Inf), right = F)
# Days between encounter and first ETI
encounter$Modulator_trikafta_first_date <-
  mdy(encounter$Modulator_trikafta_first_date)
encounter$encounterdate <- mdy(encounter$encounterdate)
encounter$Days <- as.numeric(encounter$encounterdate - encounter$Modulator_trikafta_first_date)
# Separate cleaned set
cleaned <- encounter %>% filter(gli_fev1_ppred_rn <= 200, !is.na(Days))
# List of outcomes
continuous_outcomes <- c("gli_fev1_ppred_rn", "gli_fvc_ppred_rn", "gli_fev1fvc_ppred_rn")
```

```{r functions}
# Plots
loess_plot <- function(df = cleaned, outcome, facet = F) {
  p <- ggplot(df, aes(x = Days, y = !!sym(outcome), color = Race)) +
    geom_point(shape = ".", aes(x = Days, y = !!sym(outcome)), inherit.aes = F) +
    geom_smooth() +
    theme_classic()
  if (facet) {
    p <- p + facet_wrap(~Race)
  }
  return(p)
}
```

# Data cleaning

1. For now, I have excluded any rows where the FEV1 predicted values are over 200.

# LOESS-smoothed outcome plots

```{r results='asis'}
#| warning: false
invisible(lapply(continuous_outcomes, function(o) {
  cat("\n")
  cat("##", o)
  cat("\n")
  print(loess_plot(outcome = o))
  print(loess_plot(outcome = o, facet = T))
}))
```

# Segmented models

The following models are adjusted for age group and insurance status.

## A Bayesian approach

### FEV1

```{r}
#| cache: true
model <- list(
  gli_fev1_ppred_rn ~ 1, # intercept
  gli_fev1_ppred_rn ~ 1 ~ 0 + Days # joined slope
)
fit <- mcp(model, data = cleaned)
```

```{r}
summary(fit)
plot(fit)
```

## Frequentist models

First, a Gaussian mixed model is fit with random intercept for patient ID. Then, the `segmented.lme` function is applied to this mixed model, with 0 as the starting value for finding the optimal breakpoint.

### FEV1

```{r}
#| cache: true
# Fit lme
mod1 <- lme(gli_fev1_ppred_rn ~ Days * Race + Insurance,
  random = ~ 1 | eDWID,
  data = cleaned, na.action = na.omit
)
mod1_seg <- segmented.lme(mod1, ~Days,
  psi = 0,
  random = list(eDWID = pdDiag(~ 1 + Days + U)) # ind. random eff. (changepoint fixed)
)
# mod1_seg <- segmented.lme(mod1, ~Days,
#   psi = 0,
#   random = list(eDWID = pdDiag(~ 1 + Days + U + G0))  #ind. random eff. (in the changepoint too)
# )
```

# Questions for Scott and Kam

1. Some of the predicted lung function measures are insanely high. How should we deal with outliers? Is there a limit that you would consider biologically impossible? Over 200%? What about FVC and FEV1/FVC?

2. Similarly, what would be a good cutoff for the lower end? I assume that a value of 0 is not possible.

3. I know we've discussed this before, but looking at the LOESS plots, do we want to cut off the pre-ETI at about 2000 days?

4. Are we interested in statistical inference on age group? Or are we simply adjusting for it? If we're not clinically interested in the age groups is would be worth treating it as a continuous variable.

```{r}
#| warning: false
loess_plot(df = encounter, outcome = "gli_fev1_ppred_rn")
```

### 2024-05-10 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
trying to speed up mcp

@@
 library(mcp)
-options(mc.cores = 4)
-set.seed(1017)
 # Paths
@@
 #| cache: true
+set.seed(1017)
 model <- list(
@@
 )
-fit <- mcp(model, data = cleaned)
+fit <- mcp(model, data = cleaned, chains = 4, cores = 4, iter = 1000, adapt = 0) # Increase iter and adapt later, because it will take forever to run
 ```
@@
 )
+# mod1_seg <- segmented.lme(mod1, ~Days,
+#   psi = 0,
+#   random = list(eDWID = pdDiag(~ 1 + Days + U)), # ind. random eff. (changepoint fixed)
+#   control = seg.control(n.boot = 0, display = TRUE)
+# )
 mod1_seg <- segmented.lme(mod1, ~Days,
   psi = 0,
-  random = list(eDWID = pdDiag(~ 1 + Days + U)) # ind. random eff. (changepoint fixed)
+  random = list(eDWID = pdDiag(~ 1 + Days + U + G0)), # ind. random eff. (in the changepoint too)
+  control = seg.control(n.boot = 0, display = TRUE)
 )
-# mod1_seg <- segmented.lme(mod1, ~Days,
-#   psi = 0,
-#   random = list(eDWID = pdDiag(~ 1 + Days + U + G0))  #ind. random eff. (in the changepoint too)
-# )
+```
+
+```{r}
 ```

### 2024-05-10 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
plots only

@@
     fig-cap-location: top
+    fig-height: 10
+    fig-width: 10
     page-layout: full
@@
 # List of outcomes
-continuous_outcomes <- c("gli_fev1_ppred_rn", "gli_fvc_ppred_rn", "gli_fev1fvc_ppred_rn")
+continuous_outcomes <- c(
+  "gli_fev1_ppred_rn", "gli_fvc_ppred_rn", "gli_fev1fvc_ppred_rn",
+  "bmipercentile", "heightpercentile", "weightpercentile"
+)
+binary_outcomes <- c(
+  "staphylococcus_aureus", "haemophilus_influenzae", "pseudomonasaeruginosa",
+  "burkho_complex"
+)
 ```
@@
   p <- ggplot(df, aes(x = Days, y = !!sym(outcome), color = Race)) +
-    geom_point(shape = ".", aes(x = Days, y = !!sym(outcome)), inherit.aes = F) +
+    geom_point(
+      shape = ".", aes(x = Days, y = !!sym(outcome)), inherit.aes = F,
+      alpha = 0.1
+    ) +
     geom_smooth() +
@@
   return(p)
+}
+# Basic segmented model
+basic_segmented <- function(df = cleaned, outcome, n_change_points = NULL,
+                            change_points = NULL) {
+  # Formula
+  f <- as.formula(paste0(outcome, "~ Days+Race + age_group + Insurance"))
+  # Basic linear model
+  lmod <- lm(f, data = df)
+  # Segmented model
+  lmod_seg <- segmented(lmod, seg.Z = ~Days, psi = 0, control = seg.control(it.max = 0))
+  # Plot
+  d <- lmod$model
+  d$pred <- fitted(lmod_seg)
 }
@@
   cat("\n")
+  cat("\n")
   cat("##", o)
+  cat("\n")
   cat("\n")
@@
 The following models are adjusted for age group and insurance status.
+
+## FEV1
+
+Based on visual inspection of the LOESS plots, I expect there to be a change in slope at around 0 and maybe a second one at around 1 year.
+
+```{r}
+```
 
@@
 )
-fit <- mcp(model, data = cleaned, chains = 4, cores = 4, iter = 1000, adapt = 0) # Increase iter and adapt later, because it will take forever to run
+fit <- mcp(model, data = cleaned, chains = 4, cores = 4)
 ```
@@
 
-## Frequentist models
+<!-- ## Frequentist models -->
 
-First, a Gaussian mixed model is fit with random intercept for patient ID.
+<!-- First, a Gaussian mixed model is fit with random intercept for patient ID and random slope for each patient.
 Then, the `segmented.lme` function is applied to this mixed model, with 0 as the starting value for finding the optimal breakpoint.
+The changepoint is also treated as a random effect and allowed to vary within a person.
+-->
 
-### FEV1
+<!-- ### FEV1 -->
 
-```{r}
-#| cache: true
-# Fit lme
-mod1 <- lme(gli_fev1_ppred_rn ~ Days * Race + Insurance,
-  random = ~ 1 | eDWID,
-  data = cleaned, na.action = na.omit
-)
-# mod1_seg <- segmented.lme(mod1, ~Days,
-#   psi = 0,
-#   random = list(eDWID = pdDiag(~ 1 + Days + U)), # ind. random eff. (changepoint fixed)
-#   control = seg.control(n.boot = 0, display = TRUE)
-# )
-mod1_seg <- segmented.lme(mod1, ~Days,
-  psi = 0,
-  random = list(eDWID = pdDiag(~ 1 + Days + U + G0)), # ind. random eff. (in the changepoint too)
-  control = seg.control(n.boot = 0, display = TRUE)
-)
-```
+<!-- ```{r} -->
+<!-- #| cache: true -->
+<!-- # Fit initial model -->
+<!-- ri_rs_mod_fev1 <- lme(gli_fev1_ppred_rn ~ Race * Days + age_group + Insurance, -->
+<!--   random = ~ 1 + Days | eDWID, data = cleaned, na.action = na.omit -->
+<!-- ) -->
+<!-- # Find change point -->
+<!-- ri_rs_mod_seg_fev <- segmented.lme(ri_rs_mod_fev1, ~Days, psi = 0, -->
+<!--   random = list(eDWID = pdDiag(~ 1 + Days + U + G0)), # ind. random eff. (in the changepoint too) -->
+<!--   control = seg.control(n.boot = 0, display = F) -->
+<!-- ) -->
+<!-- ``` -->
 
-```{r}
-```
+<!-- ```{r} -->
+<!-- plot(ri_rs_mod_seg_fev1) -->
+<!-- ``` -->
 
@@
 
+5. Do the linearity assumptions seem accurate to you?
+
+6. Why do you think the BMI, height, etc. percentiles look so strange? Do they normally look like this?
+
 ```{r}
@@
 ```
+
+- Is the second change point due to who is being collected after one year out?
+  - 
+- Kam will
+
+< 20% predicted and >=150 (Kam)
+
+Look at 3 years before and two after
+
+2 years for 12 and up
+1.5 years for 6

### 2024-05-12 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
reorganized race and ETI files

@@
 #| include: false
-# source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R")
-load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
+# source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_prescription_dataset.R")
+load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/prescription_dataset.RData")
 ```

### 2024-05-12 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
reorganized race and ETI files

@@
 )
+# Save dataset for mcp model script
+save(encounter, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/mcp_dataset.RData")
 ```
@@
 
+## Frequentist models
+
+First, a Gaussian mixed model is fit with random intercept for patient ID and random slope for each patient.
+Then, the `segmented.lme` function is applied to this mixed model, with 0 as the starting value for finding the optimal breakpoint.
+The changepoint is also treated as a random effect and allowed to vary within a person.
+
+### FEV1
+
 ```{r}
 #| cache: true
-set.seed(1017)
-model <- list(
-  gli_fev1_ppred_rn ~ 1, # intercept
-  gli_fev1_ppred_rn ~ 1 ~ 0 + Days # joined slope
+#| eval: false
+# Fit initial model
+ri_rs_mod_fev1 <- lme(gli_fev1_ppred_rn ~ Race * Days + age_group + Insurance,
+  random = ~ 1 + Days | eDWID, data = cleaned, na.action = na.omit
 )
-fit <- mcp(model, data = cleaned, chains = 4, cores = 4)
+# Find change point
+ri_rs_mod_seg_fev <- segmented.lme(ri_rs_mod_fev1, ~Days, psi = 0,
+  random = list(eDWID = pdDiag(~ 1 + Days + U + G0)), # ind. random eff. (in the changepoint too)
+  control = seg.control(n.boot = 0, display = F)
+)
 ```
@@
 ```{r}
-summary(fit)
-plot(fit)
+
 ```
-
-<!-- ## Frequentist models -->
-
-<!-- First, a Gaussian mixed model is fit with random intercept for patient ID and random slope for each patient.
-Then, the `segmented.lme` function is applied to this mixed model, with 0 as the starting value for finding the optimal breakpoint.
-The changepoint is also treated as a random effect and allowed to vary within a person.
--->
-
-<!-- ### FEV1 -->
-
-<!-- ```{r} -->
-<!-- #| cache: true -->
-<!-- # Fit initial model -->
-<!-- ri_rs_mod_fev1 <- lme(gli_fev1_ppred_rn ~ Race * Days + age_group + Insurance, -->
-<!--   random = ~ 1 + Days | eDWID, data = cleaned, na.action = na.omit -->
-<!-- ) -->
-<!-- # Find change point -->
-<!-- ri_rs_mod_seg_fev <- segmented.lme(ri_rs_mod_fev1, ~Days, psi = 0, -->
-<!--   random = list(eDWID = pdDiag(~ 1 + Days + U + G0)), # ind. random eff. (in the changepoint too) -->
-<!--   control = seg.control(n.boot = 0, display = F) -->
-<!-- ) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- plot(ri_rs_mod_seg_fev1) -->
-<!-- ``` -->
 

### 2024-05-12 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_prescription_dataset.R

@@ -554,5 +554,5 @@ tidy_labels <- names(labels)
 names(tidy_labels) <- sapply(labels, "[[", 1)
 # Save
 save(annual, encounter, demo, labels, tidy_labels, by_year,
-  file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData"
+  file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/prescription_dataset.RData"
 )

### 2024-05-12 Kamyron Jordan/Racial and Ethnic Differences in ETI/mcp_models.R

@@ -1,0 +1,33 @@
+# Setup
+library(tidyverse)
+library(mcp)
+# Paths
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+)
+setwd(home_dir)
+# Data import and seed.
+# Data cleaning code in racial_and_ethnic_differences_in_eti_outcomes.qmd
+load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/data_for_mcp.RData")
+set.seed(1017)
+# FEV1 single changepoint
+fev1_single_cp_model <- list(
+  gli_fev1_ppred_rn ~ 1, # intercept
+  ~ 0 + Days # joined slope
+)
+fev1_single_cp_fit <- mcp(fev1_single_cp_model,
+  data = cleaned, chains = 4, cores = 4
+)
+save(fev1_single_cp_fit, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/fev1_single_cp_fit.RData")
+# FEV1 double changepoint
+fev1_double_cp_model <- list(
+  gli_fev1_ppred_rn ~ 1, # intercept
+  ~ 0 + Days, # joined slope
+  ~ 0 + Days # joined slope
+)
+fev1_double_cp_fit <- mcp(fev1_double_cp_model,
+  data = cleaned, chains = 4, cores = 4
+)
+save(fev1_double_cp_fit, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/fev1_double_cp_fit.RData")

### 2024-05-13 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
mcp model script

@@
 # Save dataset for mcp model script
-save(encounter, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/mcp_dataset.RData")
+save(cleaned, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/mcp_dataset.RData")
 ```

### 2024-05-13 Kamyron Jordan/Racial and Ethnic Differences in ETI/mcp_models.R

@@ -10,7 +10,7 @@ home_dir <- switch(Sys.info()["sysname"],
 setwd(home_dir)
 # Data import and seed.
 # Data cleaning code in racial_and_ethnic_differences_in_eti_outcomes.qmd
-load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/data_for_mcp.RData")
+load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/mcp_dataset.RData")
 set.seed(1017)
 # FEV1 single changepoint
 fev1_single_cp_model <- list(

### 2024-05-16 Quarto Template.qmd
quarto template reference style

@@
       dark: darkly
-bibliography: ~/Documents/Miscellaneous/zotero.bib
-csl: ~/Documents/Miscellaneous/american-medical-association.csl
+bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
+csl: /home/timvigers/GitHub/styles/american-medical-association.csl
 editor: source

### 2024-05-16 Edith Zemanick/Sweat Test/sweat_test_analysis_chec.qmd
sweat test analyses with CHEC data

--- 2024-04-03 Quarto Template.qmd
@@
 ---
-title: "Title"
+title: "CHEC Study Sweat Chloride"
 author: "Tim Vigers"
@@
 library(tidyverse)
+library(readxl)
+library(ggpmisc)
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Test",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Test",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Test"
 )
+github_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/GitHub",
+  "Windows" = "C:/Users/Tim/GitHub",
+  "Linux" = "/home/timvigers/GitHub"
+)
+source(paste0(github_dir, "/CF/genotype_class_severity.R"))
 knitr::opts_knit$set(root.dir = home_dir)
@@
 
-```{r data cleaning}
+# CHCO data
+
+```{r}
+# Edith's data
+# Import exclusion data from Excel
+diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
+exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) | diagnoses$Code == "Exclude"]
+exclude <- exclude[!is.na(exclude)]
+# Import data
+df <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv",
+  na.strings = ""
+)
+# Remove exclusions from Angela
+df <- df[!(df$PatientID %in% exclude), ]
+# Diagnosis per Angela
+df$Diagnosis <- factor(diagnoses$Diagnosis[match(df$PatientID, diagnoses$PatientID)],
+  levels = c("CFTR-related disorder", "CRMS", "Cystic Fibrosis", "Non-CF"),
+  labels = c("CFTR-RD or CRMS", "CFTR-RD or CRMS", "Cystic Fibrosis", "Non-CF")
+)
+# Add modulator info
+cftr <- read.csv("./Data_Raw/Copy of Report CFTR Modulators.csv")
+df$Modulator <- apply(df, 1, function(r) {
+  mod <- cftr[cftr$Patient.ID == as.numeric(r["PatientID"]), ]
+  if (any(duplicated(mod$Age.at.Start..years.))) {
+    mod <- mod[-which(duplicated(mod$Age.at.Start..years.)), ]
+  }
+  mod <- mod[order(mod$Age.at.Start..years.), ]
+  age <- as.numeric(r["Age.at.Test.in.Years"])
+  as.character(cut(age, c(-Inf, mod$Age.at.Start..years., Inf),
+    labels = c("None", mod$Modulator), right = F
+  ))
+})
+# Numeric variables
+num_vars <- c("Amount", "Na", "K", "Cl")
+num_vars <- c(paste0(num_vars, "_Left"), paste0(num_vars, "_Right"))
+df[, num_vars] <- suppressWarnings(lapply(df[, num_vars], as.numeric))
+# Filter per Maxie's code
+exclude2 <- c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793)
+no_dem <- unique(df$PatientID[df$Demographics.Available != "TRUE" | is.na(df$Demographics.Available) | df$Sex == ""])
+df <- df %>%
+  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks * 7) %>%
+  filter(
+    Age.at.Test.in.Years < 50 & Age.at.Test.in.Days >= 2,
+    !(Diagnosis == "Non-CF" & PatientID %in% exclude2),
+    Modulator == "None"
+  ) %>%
+  distinct(.) %>%
+  filter(Sex != "", Demographics.Available == "TRUE")
+# Genotype info
+## By severity
+df$geno1 <- cf_genotype_class_severity(df$Genotypes1)
+df$geno2 <- cf_genotype_class_severity(df$Genotypes2)
+df$geno1[df$Diagnosis == "Non-CF"] <- "Not CF"
+df$geno2[df$Diagnosis == "Non-CF"] <- "Not CF"
+df$geno_risk <- paste(df$geno1, df$geno2, sep = "/")
+df$geno_risk <- factor(df$geno_risk)
+## By function
+df$geno1_class <- cf_genotype_class_severity(df$Genotypes1, out = "class")
+df$geno1_class[grep("508", df$Genotypes1)] <- "F508del"
+df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("I", "II", "III"), "MF")
+df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("IV", "V"), "RF")
+df$geno2_class <- cf_genotype_class_severity(df$Genotypes2, out = "class")
+df$geno2_class[grep("508", df$Genotypes2)] <- "F508del"
+df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("I", "II", "III"), "MF")
+df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("IV", "V"), "RF")
+df$geno_class <- factor(paste0(
+  pmin(df$geno1_class, df$geno2_class), "/",
+  pmax(df$geno1_class, df$geno2_class)
+))
+levels(df$geno_class) <- c(
+  "MF/MF", "MF/MF", "MF/RF", "MF/Unknown", "MF/MF",
+  "MF/RF", "MF/Unknown", "Both RF or Unknown", "Both RF or Unknown"
+)
+# Which arm?
+df$Arm <- apply(df[, c("Cl", "Cl_Left", "Cl_Right", "Na", "Na_Left", "Na_Right", "K", "K_Left", "K_Right")], 1, function(r) {
+  k <- sum(!is.na(r))
+  if (k == 6) {
+    return("Both")
+  } else if (k == 3) {
+    return("Arm Not Noted")
+  } else {
+    return(NA)
+  }
+})
+# Merge left and right values
+df$Amount <- apply(df[, c("Amount", "Amount_Left", "Amount_Right")], 1, function(r) {
+  amt <- suppressWarnings(as.numeric(as.character(r)))
+  mean(amt, na.rm = T)
+})
+df$Cl <- apply(df[, c("Cl", "Cl_Left", "Cl_Right")], 1, function(r) {
+  cl <- suppressWarnings(as.numeric(as.character(r)))
+  mean(cl, na.rm = T)
+})
+df$Na <- apply(df[, c("Na", "Na_Left", "Na_Right")], 1, function(r) {
+  na <- suppressWarnings(as.numeric(as.character(r)))
+  mean(na, na.rm = T)
+})
+df$K <- apply(df[, c("K", "K_Left", "K_Right")], 1, function(r) {
+  k <- suppressWarnings(as.numeric(as.character(r)))
+  mean(k, na.rm = T)
+})
+# Sodium to chloride ratio
+df$`Na/Cl Ratio` <- df$Na / df$Cl
+high_ratio <- unique(which(df$`Na/Cl Ratio` > 2))
+high_cl <- unique(which(df$Cl > 160))
+high_na <- unique(which(df$Na > 160))
+df <- df[-unique(c(high_ratio, high_cl, high_na)), ]
+# Sweat chloride group
+df$`Sweat Chloride Group` <- cut(df$Cl, c(-Inf, 30, 60, Inf),
+  right = F,
+  labels = c("< 30", "30-59", ">60")
+)
+# Sort
+df <- df %>% arrange(PatientID, Age.at.Test.in.Weeks)
+# First measure for everyone, remove those with no sweat chloride
+df <- df %>%
+  filter(!is.na(Cl)) %>%
+  group_by(PatientID) %>%
+  filter(row_number() == 1)
+# Order groups
+df$Diagnosis <- factor(df$Diagnosis, levels = c("Cystic Fibrosis", "CFTR-RD or CRMS", "Non-CF"))
 ```
 
-# Table
+## By sweat chloride group
 
 ```{r}
-#| label: tbl-planets
-#| tbl-cap: Planets
+#| warning: false
+ggplot(df, aes(x = Na, y = Cl, colour = `Sweat Chloride Group`)) +
+  geom_point(alpha=0.2) +
+  stat_poly_line(se=F) +
+  stat_poly_eq(use_label(c("eq", "R2"))) +
+  theme_bw()
+ggplot(df, aes(x = Na, y = Cl)) +
+  geom_point(alpha=0.2) +
+  stat_poly_line(se=F) +
+  stat_poly_eq(use_label(c("eq", "R2"))) +
+  theme_bw() +
+  facet_wrap(~`Sweat Chloride Group`, scales = "free", ncol = 2)
 ```
 
-# Figure
+## By classification
 
 ```{r}
-#| label: fig-planets
-#| fig-cap: Planets
+#| warning: false
+ggplot(df, aes(x = Na, y = Cl, colour = Diagnosis)) +
+  geom_point(alpha=0.2) +
+  stat_poly_line(se=F) +
+  stat_poly_eq(use_label(c("eq", "R2"))) +
+  theme_bw()
+ggplot(df, aes(x = Na, y = Cl)) +
+  geom_point(alpha=0.2) +
+  stat_poly_line(se=F) +
+  stat_poly_eq(use_label(c("eq", "R2"))) +
+  theme_bw() +
+  facet_wrap(~Diagnosis, scales = "free", ncol = 2)
 ```
+
+## By genotype severity
+
+```{r}
+#| warning: false
+ggplot(df, aes(x = Na, y = Cl, colour = geno_class)) +
+  geom_point(alpha=0.2) +
+  stat_poly_line(se=F) +
+  stat_poly_eq(use_label(c("eq", "R2"))) +
+  theme_bw()
+ggplot(df, aes(x = Na, y = Cl)) +
+  geom_point(alpha=0.2) +
+  stat_poly_line(se=F) +
+  stat_poly_eq(use_label(c("eq", "R2"))) +
+  theme_bw() +
+  facet_wrap(~geno_class, scales = "free", ncol = 2)
+```
+
+# CHEC data
+
+```{r data cleaning}
+# Import CHEC data
+chec <- read.csv("./Data_Raw/CHEC sweat sodium and chloride.csv")
+# Chloride groups
+chec$swc_group <- cut(chec$swcavg, c(-Inf, 30, 60, Inf),
+  right = F,
+  labels = c("< 30", "30-59", ">60")
+)
+```
+
+## By sweat chloride group
+
+```{r}
+#| warning: false
+ggplot(chec, aes(x = swsavg, y = swcavg, colour = swc_group)) +
+  geom_point(alpha=0.2) +
+  stat_poly_line(se=F) +
+  stat_poly_eq(use_label(c("eq", "R2"))) +
+  theme_bw()
+ggplot(chec, aes(x = swsavg, y = swcavg)) +
+  geom_point(alpha=0.2) +
+  stat_poly_line(se=F) +
+  stat_poly_eq(use_label(c("eq", "R2"))) +
+  theme_bw() +
+  facet_wrap(~swc_group, scales = "free", ncol = 2)
+```
+
+## By CHEC genotype group
+
+```{r}
+#| warning: false
+ggplot(chec, aes(x = swsavg, y = swcavg, colour = genotype_group)) +
+  geom_point(alpha=0.2) +
+  stat_poly_line(se=F) +
+  stat_poly_eq(use_label(c("eq", "R2"))) +
+  theme_bw()
+ggplot(chec, aes(x = swsavg, y = swcavg)) +
+  geom_point(alpha=0.2) +
+  stat_poly_line(se=F) +
+  stat_poly_eq(use_label(c("eq", "R2"))) +
+  theme_bw() +
+  facet_wrap(~genotype_group, scales = "free", ncol = 2)
+```

### 2024-05-17 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
separate scripts for mcp and segmented

@@
 encounter$Days <- as.numeric(encounter$encounterdate - encounter$Modulator_trikafta_first_date)
+# Remove impossible lung function values
+encounter$gli_fev1_ppred_rn[encounter$gli_fev1_ppred_rn < 20 | encounter$gli_fev1_ppred_rn > 150] <- NA
+encounter$gli_fvc_ppred_rn[encounter$gli_fvc_ppred_rn < 20 | encounter$gli_fvc_ppred_rn > 150] <- NA
+encounter$gli_fev1fvc_ppred_rn[encounter$gli_fev1fvc_ppred_rn < 20 | encounter$gli_fev1fvc_ppred_rn > 150] <- NA
 # Separate cleaned set
-cleaned <- encounter %>% filter(gli_fev1_ppred_rn <= 200, !is.na(Days))
+encounter <- encounter %>% filter(!is.na(Days))
 # List of outcomes
@@
 # Save dataset for mcp model script
-save(cleaned, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/mcp_dataset.RData")
+save(encounter, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/cleaned_outcomes_dataset.RData")
 ```
@@
 # Plots
-loess_plot <- function(df = cleaned, outcome, facet = F) {
+loess_plot <- function(df = encounter, outcome, facet = F) {
   p <- ggplot(df, aes(x = Days, y = !!sym(outcome), color = Race)) +
@@
 }
-# Basic segmented model
-basic_segmented <- function(df = cleaned, outcome, n_change_points = NULL,
-                            change_points = NULL) {
-  # Formula
-  f <- as.formula(paste0(outcome, "~ Days+Race + age_group + Insurance"))
-  # Basic linear model
-  lmod <- lm(f, data = df)
-  # Segmented model
-  lmod_seg <- segmented(lmod, seg.Z = ~Days, psi = 0, control = seg.control(it.max = 0))
-  # Plot
-  d <- lmod$model
-  d$pred <- fitted(lmod_seg)
-}
 ```
@@
 
-1. For now, I have excluded any rows where the FEV1 predicted values are over 200.
+1. Lung function percent predicted values < 20% or > 150% were excluded.
 
@@
 The following models are adjusted for age group and insurance status.
+First, a Gaussian mixed model is fit with random intercept for patient ID.
+Then, the `segmented.lme` function is applied to this mixed model, with no pre-specified starting values and `npsi = 2`.
+The change points are treated as random effects and allowed to vary within a person.
 
@@
 
-## Frequentist models
-
-First, a Gaussian mixed model is fit with random intercept for patient ID and random slope for each patient.
-Then, the `segmented.lme` function is applied to this mixed model, with 0 as the starting value for finding the optimal breakpoint.
-The changepoint is also treated as a random effect and allowed to vary within a person.
-
-### FEV1
-
 ```{r}
-#| cache: true
 #| eval: false
-# Fit initial model
-ri_rs_mod_fev1 <- lme(gli_fev1_ppred_rn ~ Race * Days + age_group + Insurance,
-  random = ~ 1 + Days | eDWID, data = cleaned, na.action = na.omit
-)
-# Find change point
-ri_rs_mod_seg_fev <- segmented.lme(ri_rs_mod_fev1, ~Days, psi = 0,
-  random = list(eDWID = pdDiag(~ 1 + Days + U + G0)), # ind. random eff. (in the changepoint too)
-  control = seg.control(n.boot = 0, display = F)
-)
-```
-
-```{r}
-
-```
-
-# Questions for Scott and Kam
-
-1.
-Some of the predicted lung function measures are insanely high.
-How should we deal with outliers?
-Is there a limit that you would consider biologically impossible?
-Over 200%?
-What about FVC and FEV1/FVC?
-
-2. Similarly, what would be a good cutoff for the lower end? I assume that a value of 0 is not possible.
-
-3. I know we've discussed this before, but looking at the LOESS plots, do we want to cut off the pre-ETI at about 2000 days?
-
-4.
-Are we interested in statistical inference on age group?
-Or are we simply adjusting for it?
-If we're not clinically interested in the age groups is would be worth treating it as a continuous variable.
-
-5. Do the linearity assumptions seem accurate to you?
-
-6. Why do you think the BMI, height, etc. percentiles look so strange? Do they normally look like this?
-
-```{r}
-#| warning: false
-loess_plot(df = encounter, outcome = "gli_fev1_ppred_rn")
+load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/fev1_single_cp_fit.RData")
+load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/fev1_double_cp_fit.RData")
 ```
@@
 - Is the second change point due to who is being collected after one year out?
-  - 
-- Kam will
-
-< 20% predicted and >=150 (Kam)
 

### 2024-05-17 Kamyron Jordan/Racial and Ethnic Differences in ETI/mcp_models.R

@@ -1,6 +1,7 @@
 # Setup
 library(tidyverse)
 library(mcp)
+set.seed(1017)
 # Paths
 home_dir <- switch(Sys.info()["sysname"],
   "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
@@ -8,26 +9,21 @@ home_dir <- switch(Sys.info()["sysname"],
   "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
 )
 setwd(home_dir)
-# Data import and seed.
+# Data import
 # Data cleaning code in racial_and_ethnic_differences_in_eti_outcomes.qmd
-load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/mcp_dataset.RData")
-set.seed(1017)
+load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/cleaned_outcomes_dataset.RData")
 # FEV1 single changepoint
 fev1_single_cp_model <- list(
   gli_fev1_ppred_rn ~ 1, # intercept
   ~ 0 + Days # joined slope
 )
-fev1_single_cp_fit <- mcp(fev1_single_cp_model,
-  data = cleaned, chains = 4, cores = 4
-)
-save(fev1_single_cp_fit, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/fev1_single_cp_fit.RData")
+fev1_single_cp_fit <- mcp(fev1_single_cp_model, data = encounter %>% filter(!is.na(gli_fev1_ppred_rn)))
+save(fev1_single_cp_fit, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/mcp/fev1_single_cp_fit.RData")
 # FEV1 double changepoint
 fev1_double_cp_model <- list(
   gli_fev1_ppred_rn ~ 1, # intercept
   ~ 0 + Days, # joined slope
   ~ 0 + Days # joined slope
 )
-fev1_double_cp_fit <- mcp(fev1_double_cp_model,
-  data = cleaned, chains = 4, cores = 4
-)
-save(fev1_double_cp_fit, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/fev1_double_cp_fit.RData")
[…]

### 2024-05-17 Kamyron Jordan/Racial and Ethnic Differences in ETI/segmented_models.R

@@ -1,0 +1,27 @@
+# Setup
+library(tidyverse)
+library(nlme)
+library(segmented)
+set.seed(1017)
+# Paths
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+)
+setwd(home_dir)
+# Data import
+# Data cleaning code in racial_and_ethnic_differences_in_eti_outcomes.qmd
+load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/cleaned_outcomes_dataset.RData")
+# FEV1
+# Basic linear mixed model
+fev1_mod <- lme(gli_fev1_ppred_rn ~ Days,
+  data = encounter, random = ~ 1 | eDWID,
+  na.action = na.omit
+)
+# Segmented model
+fev1_seg <- segmented(fev1_mod,
+  seg.Z = ~Days, npsi = 2, random = list(eDWID = pdDiag(~ 1 + Days + U + G0)),
+  control = seg.control(n.boot = 0, display = T)
+)
+save(fev1_seg, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/segmented/fev1_segmented_fit.RData")

### 2024-05-17 Edith Zemanick/Sweat Test/sweat_test_analysis_chec.qmd
added pre-modulator correlations

@@
 ggplot(df, aes(x = Na, y = Cl, colour = `Sweat Chloride Group`)) +
-  geom_point(alpha=0.2) +
-  stat_poly_line(se=F) +
+  geom_point(alpha = 0.2) +
+  stat_poly_line(se = F) +
   stat_poly_eq(use_label(c("eq", "R2"))) +
@@
 ggplot(df, aes(x = Na, y = Cl)) +
-  geom_point(alpha=0.2) +
-  stat_poly_line(se=F) +
+  geom_point(alpha = 0.2) +
+  stat_poly_line(se = F) +
   stat_poly_eq(use_label(c("eq", "R2"))) +
@@
 ggplot(df, aes(x = Na, y = Cl, colour = Diagnosis)) +
-  geom_point(alpha=0.2) +
-  stat_poly_line(se=F) +
+  geom_point(alpha = 0.2) +
+  stat_poly_line(se = F) +
   stat_poly_eq(use_label(c("eq", "R2"))) +
@@
 ggplot(df, aes(x = Na, y = Cl, colour = geno_class)) +
-  geom_point(alpha=0.2) +
-  stat_poly_line(se=F) +
+  geom_point(alpha = 0.2) +
+  stat_poly_line(se = F) +
   stat_poly_eq(use_label(c("eq", "R2"))) +
@@
 )
+chec$swc_group_pm <- cut(chec$swhcavgpm, c(-Inf, 30, 60, Inf),
+  right = F,
+  labels = c("< 30", "30-59", ">60")
+)
 ```
+
+# Pre-modulator
+
+## By sweat chloride group
+
+```{r}
+#| warning: false
+ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm, colour = swc_group_pm)) +
+  geom_point(alpha = 0.2) +
+  stat_poly_line(se = F) +
+  stat_poly_eq(use_label(c("eq", "R2"))) +
+  theme_bw()
+ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm)) +
+  geom_point(alpha = 0.2) +
+  stat_poly_line(se = F) +
+  stat_poly_eq(use_label(c("eq", "R2"))) +
+  theme_bw() +
+  facet_wrap(~swc_group_pm, scales = "free", ncol = 2)
+```
+
+## By CHEC genotype group
+
+```{r}
+#| warning: false
+ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm, colour = genotype_group)) +
+  geom_point(alpha = 0.2) +
+  stat_poly_line(se = F) +
+  stat_poly_eq(use_label(c("eq", "R2"))) +
+  theme_bw()
+ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm)) +
+  geom_point(alpha = 0.2) +
+  stat_poly_line(se = F) +
+  stat_poly_eq(use_label(c("eq", "R2"))) +
+  theme_bw() +
+  facet_wrap(~genotype_group, scales = "free", ncol = 2)
+```
+
+## By CHEC genotype group
+
+```{r}
+#| warning: false
+ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm, colour = genotype_group)) +
+  geom_point(alpha = 0.2) +
+  stat_poly_line(se = F) +
+  stat_poly_eq(use_label(c("eq", "R2"))) +
+  theme_bw()
+ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm)) +
+  geom_point(alpha = 0.2) +
+  stat_poly_line(se = F) +
+  stat_poly_eq(use_label(c("eq", "R2"))) +
+  theme_bw() +
+  facet_wrap(~genotype_group, scales = "free", ncol = 2)
+```
+
+# Post-modulator
 
@@
 ggplot(chec, aes(x = swsavg, y = swcavg, colour = swc_group)) +
-  geom_point(alpha=0.2) +
-  stat_poly_line(se=F) +
+  geom_point(alpha = 0.2) +
+  stat_poly_line(se = F) +
   stat_poly_eq(use_label(c("eq", "R2"))) +
@@
 ggplot(chec, aes(x = swsavg, y = swcavg)) +
-  geom_point(alpha=0.2) +
-  stat_poly_line(se=F) +
+  geom_point(alpha = 0.2) +
+  stat_poly_line(se = F) +
   stat_poly_eq(use_label(c("eq", "R2"))) +
@@
 ggplot(chec, aes(x = swsavg, y = swcavg, colour = genotype_group)) +
-  geom_point(alpha=0.2) +
-  stat_poly_line(se=F) +
+  geom_point(alpha = 0.2) +
+  stat_poly_line(se = F) +
   stat_poly_eq(use_label(c("eq", "R2"))) +
@@
 ```
+
+## By CHEC genotype group
+
+```{r}
+#| warning: false
+ggplot(chec, aes(x = swsavg, y = swcavg, colour = genotype_group)) +
+  geom_point(alpha = 0.2) +
+  stat_poly_line(se = F) +
+  stat_poly_eq(use_label(c("eq", "R2"))) +
+  theme_bw()
+ggplot(chec, aes(x = swsavg, y = swcavg)) +
+  geom_point(alpha = 0.2) +
+  stat_poly_line(se = F) +
+  stat_poly_eq(use_label(c("eq", "R2"))) +
+  theme_bw() +
+  facet_wrap(~genotype_group, scales = "free", ncol = 2)
+```
+
+## By modulator type
+
+```{r}
+#| warning: false
+ggplot(chec, aes(x = swsavg, y = swcavg, colour = modqual)) +
+  geom_point(alpha = 0.2) +
+  stat_poly_line(se = F) +
+  stat_poly_eq(use_label(c("eq", "R2"))) +
+  theme_bw()
+ggplot(chec, aes(x = swsavg, y = swcavg)) +
+  geom_point(alpha = 0.2) +
+  stat_poly_line(se = F) +
+  stat_poly_eq(use_label(c("eq", "R2"))) +
+  theme_bw() +
+  facet_wrap(~modqual, scales = "free", ncol = 2)
+```

### 2024-05-17 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
race and ETI update without mixed models

--- 2024-05-13 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
@@
 encounter$Days <- as.numeric(encounter$encounterdate - encounter$Modulator_trikafta_first_date)
-# Separate cleaned set
-cleaned <- encounter %>% filter(gli_fev1_ppred_rn <= 200, !is.na(Days))
+# Remove impossible lung function values
+encounter$gli_fev1_ppred_rn[encounter$gli_fev1_ppred_rn < 20 | encounter$gli_fev1_ppred_rn > 150] <- NA
+encounter$gli_fvc_ppred_rn[encounter$gli_fvc_ppred_rn < 20 | encounter$gli_fvc_ppred_rn > 150] <- NA
+encounter$gli_fev1fvc_ppred_rn[encounter$gli_fev1fvc_ppred_rn < 20 | encounter$gli_fev1fvc_ppred_rn > 150] <- NA
+# Filter
+encounter <- encounter %>%
+  filter(!is.na(Days))
 # List of outcomes
@@
 # Save dataset for mcp model script
-save(cleaned, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/mcp_dataset.RData")
+# save(encounter, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/cleaned_outcomes_dataset.RData")
 ```
@@
 # Plots
-loess_plot <- function(df = cleaned, outcome, facet = F) {
-  p <- ggplot(df, aes(x = Days, y = !!sym(outcome), color = Race)) +
+loess_plot <- function(data = encounter, outcome, facet = F, smooth = "bs",
+                       df = 10) {
+  p <- ggplot(data, aes(x = Days, y = !!sym(outcome), color = Race)) +
     geom_point(
@@
     ) +
-    geom_smooth() +
     theme_classic()
+  if (smooth == "bs") {
+    p <- p + geom_smooth(method = "lm", formula = y ~ splines::bs(x,
+      degree = 3, df = df
+    ))
+  }
+  if (smooth == "loess") {
+    p <- p + geom_smooth(method = "loess")
+  }
+  if (smooth == "gam") {
+    p <- p + geom_smooth(method = "gam")
+  }
   if (facet) {
@@
   cat("\n")
-  print(loess_plot(outcome = o))
-  print(loess_plot(outcome = o, facet = T))
+  print(loess_plot(outcome = o, smooth = "loess"))
+  print(loess_plot(outcome = o, facet = T, smooth = "loess"))
+}))
+```
+
+# B-spline smoothing (cubic, df=10)
+
+```{r results='asis'}
+#| warning: false
+invisible(lapply(continuous_outcomes, function(o) {
+  cat("\n")
+  cat("\n")
+  cat("##", o)
+  cat("\n")
+  cat("\n")
+  print(loess_plot(outcome = o, smooth = "bs"))
+  print(loess_plot(outcome = o, facet = T, smooth = "bs"))
+}))
+```
+
+# B-spline smoothing (cubic, df=3)
+
+```{r results='asis'}
+#| warning: false
+invisible(lapply(continuous_outcomes, function(o) {
+  cat("\n")
+  cat("\n")
+  cat("##", o)
+  cat("\n")
+  cat("\n")
+  print(loess_plot(outcome = o, smooth = "bs", df = 3))
+  print(loess_plot(outcome = o, facet = T, smooth = "bs", df = 3))
+}))
+```
+
+# GAM smoothing
+
+```{r results='asis'}
+#| warning: false
+invisible(lapply(continuous_outcomes, function(o) {
+  cat("\n")
+  cat("\n")
+  cat("##", o)
+  cat("\n")
+  cat("\n")
+  print(loess_plot(outcome = o, smooth = "gam"))
+  print(loess_plot(outcome = o, facet = T, smooth = "gam"))
 }))
@@
 ```{r}
+load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/segmented/fev1_single_cp_fit.RData")
+plot(fev1_single_cp_fit)
+summary(fev1_single_cp_fit)
+load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/segmented/fev1_double_cp_fit.RData")
+plot(fev1_double_cp_fit)
+summary(fev1_double_cp_fit)
 ```

### 2024-05-17 Kamyron Jordan/Racial and Ethnic Differences in ETI/mcp_models.R

@@ -14,16 +14,16 @@ setwd(home_dir)
 load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/cleaned_outcomes_dataset.RData")
 # FEV1 single changepoint
 fev1_single_cp_model <- list(
-  gli_fev1_ppred_rn ~ 1, # intercept
-  ~ 0 + Days # joined slope
+  gli_fev1_ppred_rn ~ 1 + Days,
+  1 + (1 | eDWID) ~ 0 + Days # Varying change point by person
 )
 fev1_single_cp_fit <- mcp(fev1_single_cp_model, data = encounter %>% filter(!is.na(gli_fev1_ppred_rn)))
 save(fev1_single_cp_fit, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/mcp/fev1_single_cp_fit.RData")
 # FEV1 double changepoint
 fev1_double_cp_model <- list(
-  gli_fev1_ppred_rn ~ 1, # intercept
-  ~ 0 + Days, # joined slope
-  ~ 0 + Days # joined slope
+  gli_fev1_ppred_rn ~ 1 + Days,
+  1 + (1 | eDWID) ~ 0 + Days, # Varying change point by person
+  1 + (1 | eDWID) ~ 0 + Days # Varying second change point by person (I think?)
 )
 fev1_double_cp_fit <- mcp(fev1_double_cp_model, data = encounter %>% filter(!is.na(gli_fev1_ppred_rn)))
 save(fev1_double_cp_fit, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/mcp/fev1_double_cp_fit.RData")

### 2024-05-17 Kamyron Jordan/Racial and Ethnic Differences in ETI/segmented_models.R

@@ -15,13 +15,14 @@ setwd(home_dir)
 load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/cleaned_outcomes_dataset.RData")
 # FEV1
 # Basic linear mixed model
-fev1_mod <- lme(gli_fev1_ppred_rn ~ Days,
-  data = encounter, random = ~ 1 | eDWID,
-  na.action = na.omit
+fev1_mod <- lm(gli_fev1_ppred_rn ~ Days, data = encounter, na.action = na.omit)
+# Segmented model, one change point
+fev1_single_cp_fit <- segmented(fev1_mod,
+  seg.Z = ~Days, npsi = 1, control = seg.control(display = T)
 )
-# Segmented model
-fev1_seg <- segmented(fev1_mod,
-  seg.Z = ~Days, npsi = 2, random = list(eDWID = pdDiag(~ 1 + Days + U + G0)),
-  control = seg.control(n.boot = 0, display = T)
+save(fev1_single_cp_fit, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/segmented/fev1_single_cp_fit.RData")
+# Segmented model, two change points
+fev1_double_cp_fit <- segmented(fev1_mod,
+  seg.Z = ~Days, npsi = 2, control = seg.control(display = T)
 )
-save(fev1_seg, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/segmented/fev1_segmented_fit.RData")
+save(fev1_double_cp_fit, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/segmented/fev1_double_cp_fit.RData")

### 2024-05-22 Emily Muther/mental_health_screening_psc.qmd
paths and sweat test tables plus distributions

@@
   "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF"
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+)
+github_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/GitHub",
+  "Windows" = "C:/Users/Tim/GitHub",
+  "Linux" = "/home/timvigers/GitHub"
 )
@@
 # Import
-df <- read.csv("./Emily Muther/General - Muther Research Group/Mental Health Screening/Data_Raw/PSC_Full_Data_data_11-11-22.csv",
+df <- read.csv("./Emily Muther/General/Mental Health Screening/Data_Raw/PSC_Full_Data_data_11-11-22.csv",
   check.names = F, na.strings = ""
@@
   )
+
+ggplot(df,aes(x=Score))+
+  geom_histogram(binwidth = 1) +
+  theme_bw()
 ```

### 2024-05-25 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
removed REDCap group

@@
   select(
-    record_id, redcap_data_access_group,
+    record_id,
     staphylococcus_aureus_cos:other_result_cos,

### 2024-05-28 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
sswab it feasibility

@@
 )
+# Feasibility
+db_feasibility <- db_feasibility %>% filter(!record_id %in% exclude)
 ```
@@
   as_gt()
+```
+
+# Acceptability and feasibility
+
+```{r}
+db_feasibility %>%
+  select(
+    previous_home_sample:if_no_why_not___4, -barrier_other,
+    -barriers___5, -if_no_why_not___4
+  ) %>%
+  tbl_summary(
+    missing_text = "Missing",
+    type = all_dichotomous() ~ "categorical"
+  )
 ```

### 2024-05-30 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
phase 1 update

@@
 
+## By randomization group
+
 ```{r}
@@
   cat("\n")
-  print(plot(euler(d == "Negative"), quantities = T))
+  print(plot(euler(d == "Negative"), quantities = list(cex = 3)))
   cat("\n")
@@
   cat("\n")
-  print(plot(euler(d == "Positive"), quantities = T))
+  print(plot(euler(d == "Positive"), quantities = list(cex = 3)))
   cat("\n")

### 2024-06-21 Emily Muther/body_image_trikafta.qmd
body image trikafta gender table

@@
 #| label: demographics
+normal_dist <- c("FRS_IdealBody")
 df %>%
-  select(Age, Gender, Race, Ethnicity) %>%
+  select(
+    PHQ9_Total, PHQ9_category, GAD7_Total, GAD7_category, FRStotal,
+    FRS_IdealBody, FRS_ThinkyouLook, FRS_Category, Age, Gender, Race,
+    Ethnicity
+  ) %>%
   tbl_summary(
     by = Gender,
-    statistic = list(
-      all_continuous() ~ "{mean} ({sd})"
-    ),
+    type = list(c("FRStotal", "FRS_IdealBody", "FRS_ThinkyouLook") ~
+      "continuous"),
+    statistic = list(all_of(normal_dist) ~ "{mean} [{sd}]"),
     missing_text = "(Missing)"
-  )
+  ) %>%
+  add_overall() %>%
+  add_p(
+    test = list(all_of(normal_dist) ~ "t.test"),
+    test.args = all_tests("wilcox.test") ~ list(exact = F)
+  ) %>%
+  separate_p_footnotes()
 ```
@@
     mod %>%
-      tbl_regression(exponentiate = T,intercept = T),
+      tbl_regression(exponentiate = T, intercept = T),
     mod5 %>%
-      tbl_regression(exponentiate = T,intercept = T),
+      tbl_regression(exponentiate = T, intercept = T),
     mod7 %>%
-      tbl_regression(exponentiate = T,intercept = T),
+      tbl_regression(exponentiate = T, intercept = T),
     mod9 %>%
-      tbl_regression(exponentiate = T,intercept = T),
+      tbl_regression(exponentiate = T, intercept = T),
     mod3 %>%
-      tbl_regression(exponentiate = T,intercept = T),
+      tbl_regression(exponentiate = T, intercept = T),
     mod11 %>%
-      tbl_regression(exponentiate = T,intercept = T),
+      tbl_regression(exponentiate = T, intercept = T),
     mod13 %>%
-      tbl_regression(exponentiate = T,intercept = T),
+      tbl_regression(exponentiate = T, intercept = T),
     mod15 %>%
-      tbl_regression(exponentiate = T,intercept = T)
+      tbl_regression(exponentiate = T, intercept = T)
   ),
@@
 ggsave("./Emily Muther/General/BodyImage_Trikafta/Reports/forest_plot.png",
-       plot = p,width = 9,height = 6,units = "in")
+  plot = p, width = 9, height = 6, units = "in"
+)
 ```

### 2024-06-21 Emily Muther/body_image_trikafta.qmd
trikafta adherence

@@
   )
+```
+
+## Adherence
+
+```{r}
+#| label: adherence
+df %>%
+  select(contains("Adherence"),FRS_Category) %>%
+  mutate(PercentAdherence=as.numeric(sub("%","",as.character(PercentAdherence)))) %>%
+  tbl_summary(
+    by = "FRS_Category",
+    type = list("PercentAdherence"~"continuous"),
+    missing_text = "(Missing)"
+  ) %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes()
 ```

### 2024-06-24 Emily Muther/body_image_trikafta.qmd
body image and trikafta added ranges, etc.

--- 2024-06-21 Emily Muther/body_image_trikafta.qmd
@@
 #| label: demographics
-normal_dist <- c("FRS_IdealBody")
+normal_dist <- c("FRS_IdealBody", "FEV1")
 df %>%
@@
     FRS_IdealBody, FRS_ThinkyouLook, FRS_Category, Age, Gender, Race,
-    Ethnicity
+    Ethnicity, InsuranceType, BMI, FEV1
   ) %>%
@@
     by = Gender,
-    type = list(c("FRStotal", "FRS_IdealBody", "FRS_ThinkyouLook") ~
-      "continuous"),
-    statistic = list(all_of(normal_dist) ~ "{mean} [{sd}]"),
+    type = list(
+      c("FRStotal", "FRS_IdealBody", "FRS_ThinkyouLook") ~ "continuous2",
+      all_continuous() ~ "continuous2"
+    ),
+    statistic = list(
+      all_continuous() ~ c(
+        "{median} ({p25}, {p75})",
+        "{min}, {max}"
+      ),
+      all_of(normal_dist) ~ "{mean} [{sd}]"
+    ),
     missing_text = "(Missing)"
@@
   separate_p_footnotes()
+# Skim (so Emily can see distributions)
+df %>%
+  select(
+    PHQ9_Total, PHQ9_category, GAD7_Total, GAD7_category, FRStotal,
+    FRS_IdealBody, FRS_ThinkyouLook, FRS_Category, Age, Gender, Race,
+    Ethnicity, InsuranceType, BMI, FEV1
+  ) %>%
+  skimr::skim()
 ```
@@
   )
+```
+
+## Adherence
+
+```{r}
+#| label: adherence
+df %>%
+  select(contains("Adherence"), FRS_Category) %>%
+  mutate(
+    PercentAdherence = as.numeric(sub("%", "", as.character(PercentAdherence))),
+    over80 = PercentAdherence >= 80
+  ) %>%
+  tbl_summary(
+    by = "FRS_Category",
+    type = list("PercentAdherence" ~ "continuous"),
+    missing_text = "(Missing)"
+  ) %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes()
 ```

### 2024-06-24 Emily Muther/body_image_trikafta.qmd
body image trikafta update

@@
 #| label: demographics
-normal_dist <- c("FRS_IdealBody", "FEV1")
+normal_dist <- c("FEV1", "Age", "BMI")
 df %>%
@@
   separate_p_footnotes()
-# Skim (so Emily can see distributions)
-df %>%
-  select(
-    PHQ9_Total, PHQ9_category, GAD7_Total, GAD7_category, FRStotal,
-    FRS_IdealBody, FRS_ThinkyouLook, FRS_Category, Age, Gender, Race,
-    Ethnicity, InsuranceType, BMI, FEV1
-  ) %>%
-  skimr::skim()
 ```

### 2024-06-25 Emily Muther/body_image_trikafta.qmd
range for all Table 1

@@
       all_continuous() ~ c(
-        "{median} ({p25}, {p75})",
+        "{median} [{p25}, {p75}]",
         "{min}, {max}"
       ),
-      all_of(normal_dist) ~ "{mean} [{sd}]"
+      all_of(normal_dist) ~ c("{mean} ({sd})","{min}, {max}")
     ),

### 2024-06-28 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
reference formatting

@@
       dark: darkly
+bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
+csl: /home/timvigers/GitHub/styles/american-medical-association.csl
 editor: source
@@
 Accuracy is the proportion of samples where the two tests agreed (positive/positive and negative/negative).
-Cohen's kappa accounts for agreement between the two tests by chance, and is interpreted analogously to correlation where -1 indicates perfect disagreement and 1 indicates perfect agreement.
+Cohen's kappa accounts for agreement between the two tests by chance, and is interpreted analogously to correlation where -1 indicates perfect disagreement and 1 indicates perfect agreement [@mchughInterraterReliabilityKappa2012;@ranganathanCommonPitfallsStatistical2017].
 A Cohen's kappa of NaN or NA indicates a lack of variability between the two tests.
@@
 }))
-
-# invisible(lapply(names(bug_list), function(b) {
-#   cat("\n")
-#   cat("##", as.character(bug_list[b]))
-#   cat("\n")
-#   cos <- paste0(b, "_cos")
-#   sos <- paste0(b, "_sos")
-#   d <- df[, c(cos, sos)]
-#   d <- d[complete.cases(d), ]
-#   colnames(d) <- c("COS", "SOS")
-#   cat("\n")
-#   c <- confusionMatrix(d[, "SOS"],
-#     reference = d[, "COS"], positive = "Positive", mode = "everything"
-#   )
-#   metrics <- data.frame(c$overall)
-#   cat("### 2x2 table")
-#   cat("\n")
-#   two_by_two <- data.frame(c$table)
-#   print(gt(two_by_two))
-#   cat("\n")
-#   cat("### Accuracy")
-#   cat("\n")
-#   print(gt(metrics, rownames_to_stub = T) %>%
-#     fmt_number(n_sigfig = 3) %>%
-#     tab_options(column_labels.hidden = TRUE))
-#   cat("\n")
-#   cat("### Agreement plot")
-#   cat("\n")
-#   print(agreementplot(table(d)))
-#   cat("\n")
-#   cat("### Positive/Positive Excluded")
-#   cat("\n")
-#   t <- d[rowSums(d == "Positive") < 2, ]
-#   t$`Samples With a Negative Test` <- paste(
-#     "COS", t[, "COS"], "/", "SOS", t[, "SOS"]
-#   )
-#   t %>%
-#     select(`Samples With a Negative Test`) %>%
-#     tbl_summary() %>%
-#     add_ci() %>%
-#     as_gt() %>%
-#     print()
-#   cat("\n")
-#   cat("### Euler diagram of negative results")
-#   cat("\n")
-#   print(plot(euler(d == "Negative"), quantities = T))
-#   cat("\n")
-# }))
 ```
 
+# Sub-group comparisons
+
 # References

### 2024-06-28 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
additional subgroup analyses

@@
 )) / 365.25
-db_demographics$age_group <- cut(db_demographics$age_at_cos, c(-Inf, 18, Inf),
+db_demographics$age_group <- cut(db_demographics$age_at_cos,
+  c(-Inf, 10, 18, 30, Inf),
   right = F
@@
 names(bug_list) <- bugs
-# Add group to dataframe
-df <- left_join(df, db_demographics %>% select(record_id, group),
+# Add groups to dataframe
+df <- left_join(df, db_demographics %>% select(record_id, group, age_group),
   by = join_by(record_id)
 )
+df <- left_join(df, db_medications, by = join_by(record_id, group))
 # Feasibility
@@
 
+## Age group
+
+```{r results='asis'}
+invisible(lapply(names(bug_list), function(b) {
+  cat("\n")
+  cat("###", as.character(bug_list[b]))
+  cat("\n")
+  cat("####", "Agreement by age group")
+  cat("\n")
+  agreement_test(b, group = "age_group")
+  cat("\n")
+}))
+```
+
+## Sex
+
+```{r results='asis'}
+invisible(lapply(names(bug_list), function(b) {
+  cat("\n")
+  cat("###", as.character(bug_list[b]))
+  cat("\n")
+  cat("####", "Agreement by gender assigned at birth")
+  cat("\n")
+  agreement_test(b, group = "age_group")
+  cat("\n")
+}))
+```
+
+## Azithromycin
+
+```{r results='asis'}
+invisible(lapply(names(bug_list), function(b) {
+  cat("\n")
+  cat("###", as.character(bug_list[b]))
+  cat("\n")
+  cat("####", "Agreement by Azithromycin status")
+  cat("\n")
+  agreement_test(b, group = "azithromycin")
+  cat("\n")
+}))
+```
+
+## Inhaled Tobramycin
+
+```{r results='asis'}
+invisible(lapply(names(bug_list), function(b) {
+  cat("\n")
+  cat("###", as.character(bug_list[b]))
+  cat("\n")
+  cat("####", "Agreement by inhaled Tobramycin status")
+  cat("\n")
+  agreement_test(b, group = "inhaled_tobramycin")
+  cat("\n")
+}))
+```
+
+## Inhaled Aztreonam
+
+```{r results='asis'}
+invisible(lapply(names(bug_list), function(b) {
+  cat("\n")
+  cat("###", as.character(bug_list[b]))
+  cat("\n")
+  cat("####", "Agreement by inhaled Aztreonam status")
+  cat("\n")
+  agreement_test(b, group = "inhaled_aztreonam")
+  cat("\n")
+}))
+```
+
+## Inhaled Colisitin
+
+```{r results='asis'}
+invisible(lapply(names(bug_list), function(b) {
+  cat("\n")
+  cat("###", as.character(bug_list[b]))
+  cat("\n")
+  cat("####", "Agreement by inhaled Colisitin status")
+  cat("\n")
+  agreement_test(b, group = "inhaled_colisitin")
+  cat("\n")
+}))
+```
+
+## Inhaled Vancomycin
+
+```{r results='asis'}
+invisible(lapply(names(bug_list), function(b) {
+  cat("\n")
+  cat("###", as.character(bug_list[b]))
+  cat("\n")
+  cat("####", "Agreement by inhaled Vancomycin status")
+  cat("\n")
+  agreement_test(b, group = "inhaled_vancomycin")
+  cat("\n")
+}))
+```
+
+## Other inhaled antibiotics
+
+```{r results='asis'}
+invisible(lapply(names(bug_list), function(b) {
+  cat("\n")
+  cat("###", as.character(bug_list[b]))
+  cat("\n")
+  cat("####", "Agreement by other inhaled antibiotics status")
+  cat("\n")
+  agreement_test(b, group = "other_inhaled_antibiotics")
+  cat("\n")
+}))
+```
+
+## Chronic oral antibiotic
+
+```{r results='asis'}
+invisible(lapply(names(bug_list), function(b) {
+  cat("\n")
+  cat("###", as.character(bug_list[b]))
+  cat("\n")
+  cat("####", "Agreement by chronic oral antibiotic status")
+  cat("\n")
+  agreement_test(b, group = "chronic_oral_antibiotic")
+  cat("\n")
+}))
+```
+
+## CFTR modulator status
+
+```{r results='asis'}
+invisible(lapply(names(bug_list), function(b) {
+  cat("\n")
+  cat("###", as.character(bug_list[b]))
+  cat("\n")
+  cat("####", "Agreement by CFTR modulator status")
+  cat("\n")
+  agreement_test(b, group = "cftr_modulator")
+  cat("\n")
+}))
+```
+
 # References

### 2024-07-07 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
separate data cleaning code for outcomes

@@
 # Libraries
-library(nlme)
+library(tidyverse)
+library(lmerTest)
 library(performance)
 library(segmented)
-library(tidyverse)
-library(mcp)
 # Paths
@@
 
-```{r data setup}
-#| include: false
-# Import
-encounter <- read.csv("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_encountersMerged_Del1.csv", na.strings = "")
-annual <- read.csv("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_Annualized_Del1.csv", na.strings = "")
-demo <- read.csv("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_DemogCFDiag_Del1.csv", na.strings = "")
-# Race variable
-races <- list(
-  "Race1" = "White",
-  "Race2" = "Black or African American",
-  "Race3" = "American Indian or Alaska Native",
-  "Race4" = "Asian",
-  "Race5" = "Native Hawaiian or Other Pacific Islander",
-  "Race6" = "Some other race"
-)
-demo$Race <- apply(demo[, paste0("Race", 1:6)], 1, function(r) {
-  w <- which(r == 1)
-  if (length(w) > 1 | 6 %in% w) {
-    return("Mixed/Other/Unknown Race")
-  } else {
-    return(as.character(races[w]))
-  }
-})
-demo$Race[demo$Hispanicrace == 1] <- "Hispanic or Latino"
-demo$Race[is.na(demo$Hispanicrace)] <- "Mixed/Other/Unknown Race"
-demo$Race <- factor(demo$Race,
-  levels = c(
-    "White", "Black or African American", "Hispanic or Latino", "Asian",
-    "American Indian or Alaska Native",
-    "Native Hawaiian or Other Pacific Islander", "Mixed/Other/Unknown Race"
-  ),
-  labels = c(
-    "Non-Hispanic White", "Black or African American", "Hispanic or Latino", "Asian",
-    "Other", "Other", "Other"
-  )
-)
-# Insurance
-ins_vars <- c(paste0("insurance", 1:7), "noinsurance")
-ins <- c(
-  "Health Insurance Policy (e.g. Private Insurance)", "Medicare", "Medicaid",
-  "State special needs program, e.g., BCMH, CCS, CRS, GHPP, etc.",
-  "TriCare or other military health plan",
-  "Indian Health Service", "Other", "No Insurance"
-)
-annual$Insurance <- apply(annual[, ins_vars], 1, function(r) {
-  w <- which(r == 1)
-  if (length(w) == 0) {
-    return(NA)
-  } else if (length(w) > 1) {
-    return("More than one")
-  } else {
-    return(ins[w])
-  }
-})
-annual$Insurance <- factor(annual$Insurance,
-  levels = c(
-    "Health Insurance Policy (e.g. Private Insurance)",
-    "Indian Health Service",
-    "Medicaid",
-    "Medicare",
-    "More than one",
-    "No Insurance",
-    "Other",
-    "State special needs program, e.g., BCMH, CCS, CRS, GHPP, etc.",
-    "TriCare or other military health plan"
-  ),
-  labels = c(
-    "Private", "Public", "Public", "Public", "More Than One", "No Insurance",
-    "No Insurance", "Public", "Tricare"
-  )
-)
-# Merge
-encounter <- left_join(
-  encounter,
-  demo %>% select(eDWID, Race, Modulator_trikafta_first_date)
-)
-encounter <- left_join(encounter,
-  annual %>% select(eDWID, ReviewYear, Insurance),
-  by = join_by(eDWID, reviewyear == ReviewYear)
-)
-# Age group
-encounter$age_group <- cut(encounter$encounterage, breaks = c(6, 12, 18, 30, Inf), right = F)
-# Days between encounter and first ETI
-encounter$Modulator_trikafta_first_date <-
-  mdy(encounter$Modulator_trikafta_first_date)
-encounter$encounterdate <- mdy(encounter$encounterdate)
-encounter$Days <- as.numeric(encounter$encounterdate - encounter$Modulator_trikafta_first_date)
-# Remove impossible lung function values
-encounter$gli_fev1_ppred_rn[encounter$gli_fev1_ppred_rn < 20 | encounter$gli_fev1_ppred_rn > 150] <- NA
-encounter$gli_fvc_ppred_rn[encounter$gli_fvc_ppred_rn < 20 | encounter$gli_fvc_ppred_rn > 150] <- NA
-encounter$gli_fev1fvc_ppred_rn[encounter$gli_fev1fvc_ppred_rn < 20 | encounter$gli_fev1fvc_ppred_rn > 150] <- NA
-# Filter
-encounter <- encounter %>%
-  filter(!is.na(Days))
-# List of outcomes
-continuous_outcomes <- c(
-  "gli_fev1_ppred_rn", "gli_fvc_ppred_rn", "gli_fev1fvc_ppred_rn",
-  "bmipercentile", "heightpercentile", "weightpercentile"
-)
-binary_outcomes <- c(
-  "staphylococcus_aureus", "haemophilus_influenzae", "pseudomonasaeruginosa",
-  "burkho_complex"
-)
-# Save dataset for mcp model script
-# save(encounter, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/cleaned_outcomes_dataset.RData")
+```{r data import}
+# source("/home/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R")
+load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData")
 ```
+
+# Data cleaning
+
+1. Lung function percent predicted values < 20% or > 150% were excluded.
+2. Limited to 3 years before and 2 years after ETI start.
+  - Based on `Modulator_trikafta_first_date` variable provided by CFFPR.
+3. Hospitalizations and home IV episodes excluded.
 
@@
 ```
-
-# Data cleaning
-
-1. Lung function percent predicted values < 20% or > 150% were excluded.
 
@@
 
-# Segmented models
+<!-- # Segmented models -->
 
-The following models are adjusted for age group and insurance status.
+<!-- The following models are adjusted for age group and insurance status.
 First, a Gaussian mixed model is fit with random intercept for patient ID.
@@
 The change points are treated as random effects and allowed to vary within a person.
+-->
 
-## FEV1
+<!-- ## FEV1 -->
 
-Based on visual inspection of the LOESS plots, I expect there to be a change in slope at around 0 and maybe a second one at around 1 year.
+<!-- Based on visual inspection of the LOESS plots, I expect there to be a change in slope at around 0 and maybe a second one at around 1 year. -->
 
-```{r}
-load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/segmented/fev1_single_cp_fit.RData")
-plot(fev1_single_cp_fit)
-summary(fev1_single_cp_fit)
-load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/segmented/fev1_double_cp_fit.RData")
-plot(fev1_double_cp_fit)
-summary(fev1_double_cp_fit)
-```
+<!-- ```{r} -->
+<!-- load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/segmented/fev1_single_cp_fit.RData") -->
+<!-- plot(fev1_single_cp_fit) -->
+<!-- summary(fev1_single_cp_fit) -->
+<!-- load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/segmented/fev1_double_cp_fit.RData") -->
+<!-- plot(fev1_double_cp_fit) -->
+<!-- summary(fev1_double_cp_fit) -->
+<!-- ``` -->
 
-## A Bayesian approach
+<!-- ## A Bayesian approach -->
 
-### FEV1
+<!-- ### FEV1 -->
 
-```{r}
-#| eval: false
-load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/fev1_single_cp_fit.RData")
-load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/fev1_double_cp_fit.RData")
-```
+<!-- ```{r} -->
+<!-- #| eval: false -->
+<!-- load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/fev1_single_cp_fit.RData") -->
+<!-- load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/fev1_double_cp_fit.RData") -->
+<!-- ``` -->
 
-- Is the second change point due to who is being collected after one year out?
+<!-- - Is the second change point due to who is being collected after one year out? -->
 
-Look at 3 years before and two after
+<!-- Look at 3 years before and two after -->
 
-2 years for 12 and up
-1.5 years for 6
+<!-- 2 years for 12 and up -->
+<!-- 1.5 years for 6 -->

### 2024-07-07 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R

@@ -1,0 +1,143 @@
+library(tidyverse)
+library(lubridate)
+library(parallel)
+# Paths
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+)
+setwd(home_dir)
+# Import
+encounter <- read.csv("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_encountersMerged_Del1.csv",
+  na.strings = ""
+)
+annual <- read.csv("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_Annualized_Del1.csv",
+  na.strings = ""
+)
+demo <- read.csv("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_DemogCFDiag_Del1.csv",
+  na.strings = ""
+)
+hospitalizations <- read.csv("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_CareEpisodes_Del1.csv",
+  na.strings = ""
+)
+# Race variable
+races <- list(
+  "Race1" = "White",
+  "Race2" = "Black or African American",
+  "Race3" = "American Indian or Alaska Native",
+  "Race4" = "Asian",
+  "Race5" = "Native Hawaiian or Other Pacific Islander",
+  "Race6" = "Some other race"
+)
+demo$Race <- apply(demo[, paste0("Race", 1:6)], 1, function(r) {
+  w <- which(r == 1)
+  if (length(w) > 1 | 6 %in% w) {
+    return("Mixed/Other/Unknown Race")
+  } else {
+    return(as.character(races[w]))
+  }
+})
+demo$Race[demo$Hispanicrace == 1] <- "Hispanic or Latino"
+demo$Race[is.na(demo$Hispanicrace)] <- "Mixed/Other/Unknown Race"
+demo$Race <- factor(demo$Race,
+  levels = c(
+    "White", "Black or African American", "Hispanic or Latino", "Asian",
+    "American Indian or Alaska Native",
[…]

### 2024-07-07 Kamyron Jordan/Racial and Ethnic Differences in ETI/mcp_models.R

@@ -1,29 +1,0 @@
-# Setup
-library(tidyverse)
-library(mcp)
-set.seed(1017)
-# Paths
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
-)
-setwd(home_dir)
-# Data import
-# Data cleaning code in racial_and_ethnic_differences_in_eti_outcomes.qmd
-load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/cleaned_outcomes_dataset.RData")
-# FEV1 single changepoint
-fev1_single_cp_model <- list(
-  gli_fev1_ppred_rn ~ 1 + Days,
-  1 + (1 | eDWID) ~ 0 + Days # Varying change point by person
-)
-fev1_single_cp_fit <- mcp(fev1_single_cp_model, data = encounter %>% filter(!is.na(gli_fev1_ppred_rn)))
-save(fev1_single_cp_fit, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/mcp/fev1_single_cp_fit.RData")
-# FEV1 double changepoint
-fev1_double_cp_model <- list(
-  gli_fev1_ppred_rn ~ 1 + Days,
-  1 + (1 | eDWID) ~ 0 + Days, # Varying change point by person
-  1 + (1 | eDWID) ~ 0 + Days # Varying second change point by person (I think?)
-)
-fev1_double_cp_fit <- mcp(fev1_double_cp_model, data = encounter %>% filter(!is.na(gli_fev1_ppred_rn)))
-save(fev1_double_cp_fit, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/mcp/fev1_double_cp_fit.RData")

### 2024-07-07 Kamyron Jordan/Racial and Ethnic Differences in ETI/segmented_models.R

@@ -1,28 +1,0 @@
-# Setup
-library(tidyverse)
-library(nlme)
-library(segmented)
-set.seed(1017)
-# Paths
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
-)
-setwd(home_dir)
-# Data import
-# Data cleaning code in racial_and_ethnic_differences_in_eti_outcomes.qmd
-load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/cleaned_outcomes_dataset.RData")
-# FEV1
-# Basic linear mixed model
-fev1_mod <- lm(gli_fev1_ppred_rn ~ Days, data = encounter, na.action = na.omit)
-# Segmented model, one change point
-fev1_single_cp_fit <- segmented(fev1_mod,
-  seg.Z = ~Days, npsi = 1, control = seg.control(display = T)
-)
-save(fev1_single_cp_fit, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/segmented/fev1_single_cp_fit.RData")
-# Segmented model, two change points
-fev1_double_cp_fit <- segmented(fev1_mod,
-  seg.Z = ~Days, npsi = 2, control = seg.control(display = T)
-)
-save(fev1_double_cp_fit, file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/segmented/fev1_double_cp_fit.RData")

### 2024-07-07 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
need desktop for some of this I think

@@
 library(lmerTest)
+library(splines)
+library(ggeffects)
 library(performance)
@@
 3. Hospitalizations and home IV episodes excluded.
+4. Years of pregnancy excluded (we only have annualized pregancy data, so don't know exact dates for this).
 
@@
 
-<!-- # Segmented models -->
+# Segmented models
 
-<!-- The following models are adjusted for age group and insurance status.
-First, a Gaussian mixed model is fit with random intercept for patient ID.
-Then, the `segmented.lme` function is applied to this mixed model, with no pre-specified starting values and `npsi = 2`.
-The change points are treated as random effects and allowed to vary within a person.
--->
+The following models are adjusted for age group and insurance status. We fit a Gaussian mixed model for continuous outcomes with random intercept for each subject.
 
-<!-- ## FEV1 -->
+```{r}
+invisible(lapply(continuous_outcomes), function(o) {
+  f <- as.formula(paste0(
+    o,
+    "~ age_group + Insurance + bs(Days, knots = 0, degree = 1) * Race + (1 | eDWID)"
+  ))
+  m = lmer(f,data = encounter)
+  predict_response(model = m, terms = c("Days[all]", "Race")) %>% plot()
+})
 
-<!-- Based on visual inspection of the LOESS plots, I expect there to be a change in slope at around 0 and maybe a second one at around 1 year. -->
-
-<!-- ```{r} -->
-<!-- load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/segmented/fev1_single_cp_fit.RData") -->
-<!-- plot(fev1_single_cp_fit) -->
-<!-- summary(fev1_single_cp_fit) -->
-<!-- load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/segmented/fev1_double_cp_fit.RData") -->
-<!-- plot(fev1_double_cp_fit) -->
-<!-- summary(fev1_double_cp_fit) -->
-<!-- ``` -->
-
-<!-- ## A Bayesian approach -->
-
-<!-- ### FEV1 -->
-
-<!-- ```{r} -->
-<!-- #| eval: false -->
-<!-- load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/fev1_single_cp_fit.RData") -->
-<!-- load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/fev1_double_cp_fit.RData") -->
-<!-- ``` -->
-
-<!-- - Is the second change point due to who is being collected after one year out? -->
-
-<!-- Look at 3 years before and two after -->
-
-<!-- 2 years for 12 and up -->
-<!-- 1.5 years for 6 -->
+```

### 2024-07-07 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R

@@ -92,7 +92,7 @@ encounter <- left_join(
   demo %>% select(eDWID, Race, Modulator_trikafta_first_date)
 )
 encounter <- left_join(encounter,
-  annual %>% select(eDWID, ReviewYear, Insurance),
+  annual %>% select(eDWID, ReviewYear, Insurance, pregnant),
   by = join_by(eDWID, reviewyear == ReviewYear)
 )
 # Age group
@@ -107,14 +107,16 @@ encounter$Days <- as.numeric(encounter$encounterdate - encounter$Modulator_trika
 # Hospitalization or home IVs
 hospitalizations$CareEpi_StartDt <- mdy(hospitalizations$CareEpi_StartDt)
 hospitalizations$CareEpi_EndDt <- mdy(hospitalizations$CareEpi_EndDt)
-cl <- makeForkCluster(8)
+cl <- makeForkCluster(detectCores() * (3 / 4))
 hosps <- parApply(cl, hospitalizations, 1, function(r) {
   which(encounter$eDWID == r["eDWID"] &
     encounter$encounterdate >= r["CareEpi_StartDt"] &
     encounter$encounterdate <= r["CareEpi_EndDt"])
 })
-hosps <- unique(unlist(hosps))
 stopCluster(cl)
+hosps <- unique(unlist(hosps))
+encounter$hospitalized <- "No"
+encounter$hospitalized[hosps] <- "Yes"
 # Remove impossible lung function values
 encounter$gli_fev1_ppred_rn[encounter$gli_fev1_ppred_rn < 20 |
   encounter$gli_fev1_ppred_rn > 150] <- NA
@@ -126,7 +128,9 @@ encounter$gli_fev1fvc_ppred_rn[encounter$gli_fev1fvc_ppred_rn < 20 |
 encounter <- encounter %>%
   filter(
     !is.na(Days),
-    Days >= -365.25 * 3 & Days <= 365.25 * 2
+    Days >= -365.25 * 3 & Days <= 365.25 * 2,
+    hospitalized == "No",
+    pregnant != 1
   )
 # List of outcomes
 continuous_outcomes <- c(
@@ -138,6 +142,6 @@ binary_outcomes <- c(
   "burkho_complex"
 )
 # Save
-save(encounter,
+save(encounter, continuous_outcomes, binary_outcomes,
   file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData"
 )

### 2024-07-07 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
maybe only use half the cores instead of 3/4

@@
   ))
-  m = lmer(f,data = encounter)
+  m <- lmer(f, data = encounter)
   predict_response(model = m, terms = c("Days[all]", "Race")) %>% plot()
 })
-
 ```

### 2024-07-07 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R

@@ -107,7 +107,7 @@ encounter$Days <- as.numeric(encounter$encounterdate - encounter$Modulator_trika
 # Hospitalization or home IVs
 hospitalizations$CareEpi_StartDt <- mdy(hospitalizations$CareEpi_StartDt)
 hospitalizations$CareEpi_EndDt <- mdy(hospitalizations$CareEpi_EndDt)
-cl <- makeForkCluster(detectCores() * (3 / 4))
+cl <- makeForkCluster(detectCores() * (2 / 4))
 hosps <- parApply(cl, hospitalizations, 1, function(r) {
   which(encounter$eDWID == r["eDWID"] &
     encounter$encounterdate >= r["CareEpi_StartDt"] &

### 2024-07-07 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
double check hospitalization code!

@@
 
-```{r data import}
+```{r data import and minor cleaning}
+# Run code for base dataset (takes forever)
 # source("/home/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R")
+# Import
 load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData")
+# Additional quick cleaning steps
+# Remove impossible lung function values
+encounter$gli_fev1_ppred_rn[encounter$gli_fev1_ppred_rn < 20 |
+  encounter$gli_fev1_ppred_rn > 150] <- NA
+encounter$gli_fvc_ppred_rn[encounter$gli_fvc_ppred_rn < 20 |
+  encounter$gli_fvc_ppred_rn > 150] <- NA
+encounter$gli_fev1fvc_ppred_rn[encounter$gli_fev1fvc_ppred_rn < 20 |
+  encounter$gli_fev1fvc_ppred_rn > 150] <- NA
+# Filter
+encounter <- encounter %>%
+  filter(
+    !is.na(Days),
+    Days >= -365.25 * 3 & Days <= 365.25 * 2,
+    hospitalized == "No",
+    pregnant != 1
+  )
 ```

### 2024-07-07 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
cache

@@
 #| warning: false
+#| cache: true
 invisible(lapply(continuous_outcomes, function(o) {
@@
 
+## With Asian subgroup
+
 ```{r}
-invisible(lapply(continuous_outcomes), function(o) {
+invisible(lapply(continuous_outcomes, function(o) {
   f <- as.formula(paste0(
@@
   m <- lmer(f, data = encounter)
-  predict_response(model = m, terms = c("Days[all]", "Race")) %>% plot()
-})
+  predict_response(model = m, terms = c("Days[all]", "Race")) %>%
+    plot() %>%
+    print()
+}))
 ```
+
+## Without Asian subgroup (for clearer plotting)
+
+```{r}
+invisible(lapply(continuous_outcomes, function(o) {
+  f <- as.formula(paste0(
+    o,
+    "~ age_group + Insurance + bs(Days, knots = 0, degree = 1) * Race + (1 | eDWID)"
+  ))
+  m <- lmer(f, data = encounter %>% filter(Race != "Asian"))
+  predict_response(model = m, terms = c("Days[all]", "Race")) %>%
+    plot() %>%
+    print()
+}))
+```

### 2024-07-08 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
data cleaned and tables done

@@
 library(ggeffects)
-library(performance)
-library(segmented)
+library(gtsummary)
 # Paths
@@
 
-```{r data import and minor cleaning}
+```{r data import}
 # Run code for base dataset (takes forever)
@@
 load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData")
+```
+
+# Data cleaning
+
+1. Lung function percent predicted values < 20% or > 150% were excluded.
+  - `r sum(encounter$gli_fev1_ppred_rn < 20,na.rm = T)` ppFEV1 values < 20% and `r sum(encounter$gli_fev1_ppred_rn > 150,na.rm = T)` over 150%.
+2. Limited to 3 years before and 2 years after ETI start.
+  - Based on `Modulator_trikafta_first_date` variable provided by CFFPR.
+3. Hospitalizations and home IV episodes excluded.
+  - Based on the care episodes dataset, there were `r length(unique(encounter$eDWID[encounter$hospitalized=="Yes"]))` unique participants with at least one care episode and a total of `r sum(encounter$hospitalized=="Yes")` encounters that were part of a care episode.
+  - After filtering out these episodes, there were still 914 encounters with `encounterlocation` variable indicating either hospitalization or home IV, so these were also excluded.
+4. Years of pregnancy excluded (we only have annualized pregnancy data, so don't know exact dates for this).
+  - `r length(unique(encounter$eDWID[encounter$pregnant==1]))` unique participants with at least one care episode and a total of `r sum(encounter$pregnant==1, na.rm = T)` encounters during a pregnancy year.
+5.
+Those missing `Modulator_trikafta_first_date` were automatically excluded from the models because they don't have a value for `Days`.
+This naturally means that people not eligible for Trikafta are not included in the models.
+6. 990 people with no encounters post-ETI were excluded.
+
+```{r filtering}
 # Additional quick cleaning steps
@@
     Days >= -365.25 * 3 & Days <= 365.25 * 2,
-    hospitalized == "No",
-    pregnant != 1
+    hospitalized == "No", !encounterlocation %in% c("Hospital", "Home IV"),
+    pregnant != 1,
+    !is.na(Modulator_trikafta_first_date)
   )
+# Get some summary statistics by participant
+t1_participant <- encounter %>%
+  group_by(eDWID) %>%
+  summarise(
+    `Years With Data` = length(unique(reviewyear)),
+    `Number of Encounters` = n(),
+    `Number of Encounters Pre-ETI` = sum(Days <= 0),
+    `Number of Encounters Post-ETI` = sum(Days > 0),
+    Age = mean(encounterage, na.rm = T),
+    across(all_of(continuous_outcomes), \(x) mean(x, na.rm = TRUE)),
+    Race = names(sort(table(Race), decreasing = TRUE))[1],
+    Insurance = names(sort(table(Insurance), decreasing = TRUE))[1],
+    .groups = "drop"
+  )
+no_post <- t1_participant$eDWID[t1_participant$`Number of Encounters Post-ETI` == 0]
+encounter <- encounter %>% filter(!eDWID %in% no_post)
+t1_participant <- t1_participant %>% filter(!eDWID %in% no_post)
 ```
 
-# Data cleaning
+# Table 1
 
-1. Lung function percent predicted values < 20% or > 150% were excluded.
-2. Limited to 3 years before and 2 years after ETI start.
-  - Based on `Modulator_trikafta_first_date` variable provided by CFFPR.
-3. Hospitalizations and home IV episodes excluded.
-4. Years of pregnancy excluded (we only have annualized pregancy data, so don't know exact dates for this).
+## By person
+
+Continuous variables were averaged across all years for each person, and categorical variables are reported as the most common value.
+
+```{r}
+t1_participant %>%
+  select(-eDWID) %>%
+  tbl_summary(
+    by = Race,
+    type = list(`Years With Data` ~ "continuous"),
+    missing_text = "Missing"
+  ) %>%
+  add_overall()
+```
+
+# Encounter-level
+
+```{r}
+encounter %>%
+  select(Race, encounterlocation, bacterialculturedone:othermicroorganisms1) %>%
+  tbl_summary(by = Race, missing_text = "Missing")
+```
+
+# LOESS-smoothed outcome plots
 
@@
 # Plots
-loess_plot <- function(data = encounter, outcome, facet = F, smooth = "bs",
-                       df = 10) {
+loess_plot <- function(data = encounter, outcome, facet = F) {
   p <- ggplot(data, aes(x = Days, y = !!sym(outcome), color = Race)) +
@@
     ) +
+    geom_smooth(method = "loess",se = F) +
     theme_classic()
-  if (smooth == "bs") {
-    p <- p + geom_smooth(method = "lm", formula = y ~ splines::bs(x,
-      degree = 3, df = df
-    ))
-  }
-  if (smooth == "loess") {
-    p <- p + geom_smooth(method = "loess")
-  }
-  if (smooth == "gam") {
-    p <- p + geom_smooth(method = "gam")
-  }
   if (facet) {
@@
 
-# LOESS-smoothed outcome plots
-
 ```{r results='asis'}
 #| warning: false
-#| cache: true
 invisible(lapply(continuous_outcomes, function(o) {
@@
   cat("\n")
-  print(loess_plot(outcome = o, smooth = "loess"))
-  print(loess_plot(outcome = o, facet = T, smooth = "loess"))
-}))
-```
-
-# B-spline smoothing (cubic, df=10)
-
-```{r results='asis'}
-#| warning: false
-#| cache: true
-invisible(lapply(continuous_outcomes, function(o) {
-  cat("\n")
-  cat("\n")
-  cat("##", o)
-  cat("\n")
-  cat("\n")
-  print(loess_plot(outcome = o, smooth = "bs"))
-  print(loess_plot(outcome = o, facet = T, smooth = "bs"))
-}))
-```
-
-# B-spline smoothing (cubic, df=3)
-
-```{r results='asis'}
-#| warning: false
-#| cache: true
-invisible(lapply(continuous_outcomes, function(o) {
-  cat("\n")
-  cat("\n")
-  cat("##", o)
-  cat("\n")
-  cat("\n")
-  print(loess_plot(outcome = o, smooth = "bs", df = 3))
-  print(loess_plot(outcome = o, facet = T, smooth = "bs", df = 3))
-}))
-```
-
-# GAM smoothing
-
-```{r results='asis'}
-#| warning: false
-#| cache: true
-invisible(lapply(continuous_outcomes, function(o) {
-  cat("\n")
-  cat("\n")
-  cat("##", o)
-  cat("\n")
-  cat("\n")
-  print(loess_plot(outcome = o, smooth = "gam"))
-  print(loess_plot(outcome = o, facet = T, smooth = "gam"))
+  print(loess_plot(outcome = o))
 }))
@@
 ```
+
+# Questions
+
+1. Do we want to combine the Asian and "Other" groups? It makes sense from a statistical point of view.

### 2024-07-08 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R

@@ -86,6 +86,18 @@ annual$Insurance <- factor(annual$Insurance,
     "No Insurance", "Public", "Tricare"
   )
 )
+# Other factor variables
+encounter$encounterlocation <- factor(encounter$encounterlocation,
+  levels = 1:6, labels = c("Clinic", "Hospital", "Home IV", "Other", "5", "6")
+)
+encounter$bacterialculturedone[is.na(encounter$bacterialculturedone)] <- 0
+encounter$bacterialculturedone <- factor(encounter$bacterialculturedone,
+  levels = 0:1, labels = c("No", "Yes")
+)
+encounter$cultureresults <- factor(encounter$cultureresults,
+  levels = 1:3,
+  labels = c("No growth/sterile culture", "Normal flora", "Microorganisms")
+)
 # Merge
 encounter <- left_join(
   encounter,

### 2024-07-09 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
started flowchart

@@
 ```{r data import}
-# Run code for base dataset (takes forever)
 # source("/home/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R")
-# Import
 load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData")
@@
 
-# Data cleaning
-
-1. Lung function percent predicted values < 20% or > 150% were excluded.
-  - `r sum(encounter$gli_fev1_ppred_rn < 20,na.rm = T)` ppFEV1 values < 20% and `r sum(encounter$gli_fev1_ppred_rn > 150,na.rm = T)` over 150%.
-2. Limited to 3 years before and 2 years after ETI start.
-  - Based on `Modulator_trikafta_first_date` variable provided by CFFPR.
-3. Hospitalizations and home IV episodes excluded.
-  - Based on the care episodes dataset, there were `r length(unique(encounter$eDWID[encounter$hospitalized=="Yes"]))` unique participants with at least one care episode and a total of `r sum(encounter$hospitalized=="Yes")` encounters that were part of a care episode.
-  - After filtering out these episodes, there were still 914 encounters with `encounterlocation` variable indicating either hospitalization or home IV, so these were also excluded.
-4. Years of pregnancy excluded (we only have annualized pregnancy data, so don't know exact dates for this).
-  - `r length(unique(encounter$eDWID[encounter$pregnant==1]))` unique participants with at least one care episode and a total of `r sum(encounter$pregnant==1, na.rm = T)` encounters during a pregnancy year.
-5.
-Those missing `Modulator_trikafta_first_date` were automatically excluded from the models because they don't have a value for `Days`.
-This naturally means that people not eligible for Trikafta are not included in the models.
-6. 990 people with no encounters post-ETI were excluded.
-
-```{r filtering}
-# Additional quick cleaning steps
+```{r filtering and flowchart}
+# Variables for flowchart
+n_enc_1 <- nrow(encounter)
+n_people_1 <- length(unique(encounter$eDWID))
+n_hosp_1 <- nrow(hospitalizations)
+n_people_hosp_1 <- length(unique(hospitalizations$eDWID))
+# Lung function cutoffs
+lower <- 20
+upper <- 150
 # Remove impossible lung function values
+low_fev1 = which(encounter$gli_fev1_ppred_rn)
 encounter$gli_fev1_ppred_rn[encounter$gli_fev1_ppred_rn < 20 |
@@
 t1_participant <- t1_participant %>% filter(!eDWID %in% no_post)
+```
+
+# Data cleaning
+
+1. Lung function percent predicted values < 20% or > 150% were excluded.
+  - `r sum(encounter$gli_fev1_ppred_rn < 20,na.rm = T)` ppFEV1 values < 20% and `r sum(encounter$gli_fev1_ppred_rn > 150,na.rm = T)` over 150%.
+2. Limited to 3 years before and 2 years after ETI start.
+  - Based on `Modulator_trikafta_first_date` variable provided by CFFPR.
+3. Hospitalizations and home IV episodes excluded.
+  - Based on the care episodes dataset, there were `r length(unique(encounter$eDWID[encounter$hospitalized=="Yes"]))` unique participants with at least one care episode and a total of `r sum(encounter$hospitalized=="Yes")` encounters that were part of a care episode.
+  - After filtering out these episodes, there were still 914 encounters with `encounterlocation` variable indicating either hospitalization or home IV, so these were also excluded.
+4. Years of pregnancy excluded (we only have annualized pregnancy data, so don't know exact dates for this).
+  - `r length(unique(encounter$eDWID[encounter$pregnant==1]))` unique participants with at least one care episode and a total of `r sum(encounter$pregnant==1, na.rm = T)` encounters during a pregnancy year.
+5.
+Those missing `Modulator_trikafta_first_date` were automatically excluded from the models because they don't have a value for `Days`.
+This naturally means that people not eligible for Trikafta are not included in the models.
+6. 990 people with no encounters post-ETI were excluded.
+
+```{r filtering}
+# Additional quick cleaning steps
+
 ```
@@
     ) +
-    geom_smooth(method = "loess",se = F) +
+    geom_smooth(method = "loess", se = F) +
     theme_classic()

### 2024-07-11 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
simplified incidence and prevalence using annualized data

@@
 annual$age_group <- cut(annual$Age_YrEnd,
-  breaks = c(-Inf, 10, 18, 30, Inf),
+  breaks = c(-Inf, 10, 19, 31, Inf),
   right = F
 )
-annual$age_group <- relevel(annual$age_group, ref = "[10,18)")
 # Bugs
@@
   ungroup()
-# For each person, filter out all the years after first CFRD diagnosis
-annual <- annual %>% filter(ReviewYear <= first_cfrd | is.na(first_cfrd))
-# Remove those under 10
-annual <- annual %>% filter(Age_YrEnd != 0, Age_YrEnd >= 10)
-annual$age_group <- droplevels(annual$age_group)
 # List of covariates
@@
 
-# CFRD incidence and prevalence
+# CFRD prevalence
+
+## By year
+
+```{r}
+annual %>%
+  group_by(ReviewYear) %>%
+  summarise(Prevalence = mean(CFRD == "CFRD+") * 100) %>%
+  kable(., digits = 2)
+```
+
+## By year and age group
+
+```{r}
+annual %>%
+  group_by(ReviewYear, age_group) %>%
+  summarise(Prevalence = mean(CFRD == "CFRD+") * 100, .groups = "drop") %>%
+  pivot_wider(names_from = age_group, values_from = Prevalence) %>%
+  kable(., digits = 2)
+```
+
+# CFRD incidence
+
+```{r}
+# For each person, filter out all the years after first CFRD diagnosis
+annual <- annual %>% filter(ReviewYear <= first_cfrd | is.na(first_cfrd))
+# For each year, pull everyone with and without CFRD that year
+year_inc <- function(df = annual, year, py = 1000) {
+  # Filter and count
+  inc <- df %>% filter(ReviewYear == year)
+  inc <- inc %>%
+    group_by(CFRD) %>%
+    count(age_group, .drop = F)
+  inc <- inc[!is.na(inc$age_group), ]
+  # Calculations
+  new_cases <- sum(inc$n[inc$CFRD == "CFRD+"])
+  total_at_risk <- sum(inc$n)
+  total_inc <- new_cases / (total_at_risk / py)
+
+  new_cases_10 <- sum(inc$n[inc$CFRD == "CFRD+" & inc$age_group == "[-Inf,10)"])
+  at_risk_10 <- sum(inc$n[inc$age_group == "[-Inf,10)"])
+  inc_10 <- new_cases_10 / (at_risk_10 / py)
+
+  new_cases_10_18 <- sum(inc$n[inc$CFRD == "CFRD+" & inc$age_group == "[10,19)"])
+  at_risk_10_18 <- sum(inc$n[inc$age_group == "[10,19)"])
+  inc_10_18 <- new_cases_10_18 / (at_risk_10_18 / py)
+
+  new_cases_19_30 <- sum(inc$n[inc$CFRD == "CFRD+" & inc$age_group == "[19,31)"])
+  at_risk_19_30 <- sum(inc$n[inc$age_group == "[19,31)"])
+  inc_19_30 <- new_cases_19_30 / (at_risk_19_30 / py)
+
+  new_cases_30 <- sum(inc$n[inc$CFRD == "CFRD+" & inc$age_group == "[31, Inf)"])
+  at_risk_30 <- sum(inc$n[inc$age_group == "[31, Inf)"])
+  inc_30 <- new_cases_30 / (at_risk_30 / py)
+
+  # Nice results
+  list(
+    `Year` = year,
+    `Total New Cases` = new_cases,
+    `Total At Risk` = total_at_risk,
+    `Total Incidence` = total_inc,
+    `Total New Cases Age < 10` = new_cases_10,
+    `Total At Risk Age < 10` = at_risk_10,
+    `Total Incidence Age < 10` = inc_10,
+    `Total New Cases Age 10-18` = new_cases_10_18,
+    `Total At Risk Age 10-18` = at_risk_10_18,
+    `Total Incidence Age 10-18` = inc_10_18,
+    `Total New Cases Age 19-30` = new_cases_19_30,
+    `Total At Risk Age 19-30` = at_risk_19_30,
+    `Total Incidence Age 19-30` = inc_19_30,
+    `Total New Cases Age >= 31` = new_cases_30,
+    `Total At Risk Age >= 31` = at_risk_30,
+    `Total Incidence Age >= 31` = inc_30
+  )
+}
+incidence <- lapply(2003:2018, function(y) {
+  year_inc(year = y)
+})
+incidence <- data.frame(do.call(rbind, incidence), check.names = F)
+incidence <- data.frame(lapply(incidence, as.numeric), check.names = F)
+kable(incidence, digits = 2)
+```
 
@@
   )
-```
-
-# Incidence
-
-## Full cohort
-
-```{r}
-#| message: false
-og <- ymd("2003-01-01")
-km_fit <- survfit(
-  Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
-    origin = as.numeric(og)
-  ) ~ 1,
-  data = annual, id = eDWID
-)
-# Plot
-surv_p <- ggsurvplot(km_fit, data = annual, fun = "event")
-surv_p$plot +
-  scale_x_continuous(
-    breaks = km_fit$time,
-    labels = year(ymd("2003-01-01") + days(km_fit$time))
-  ) +
-  theme(axis.title.x = element_blank(), legend.position = "none") +
-  ggtitle("Cumulative Incidence")
-# Table
-km_table <- tidy(km_fit)
-km_table$time <- year(og + days(km_table$time))
-gt(km_table) %>%
-  cols_hide("std.error") %>%
-  fmt_percent(columns = c("estimate", "conf.low", "conf.high")) %>%
-  cols_merge(
-    columns = c("estimate", "conf.low", "conf.high"),
-    pattern = "{1} [{2} - {3}]"
-  ) %>%
-  cols_label(
-    time = "Year", n.risk = "At Risk", n.event = "CFRD+",
-    n.censor = "Censored", estimate = "Estimated CFRD- [95% CI]"
-  ) %>%
-  opt_row_striping() %>%
-  opt_table_lines() %>%
-  opt_stylize(style = 6, color = "gray")
-```
-
-## By age group
-
-```{r}
-#| message: false
-# Plot
-km_age_fit <- survfit(
-  Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
-    origin = as.numeric(og)
-  ) ~ age_group,
-  data = annual, id = eDWID
-)
-surv_p <- ggsurvplot(km_age_fit, data = annual, fun = "event")
-surv_p$plot +
-  scale_x_continuous(
-    breaks = km_age_fit$time,
-    labels = year(ymd("2003-01-01") + days(km_age_fit$time))
-  ) +
-  scale_color_discrete(name = "Age Group", labels = c("10-17", "18-29", "30+")) +
-  theme(axis.title.x = element_blank(), legend.title = element_blank()) +
-  ggtitle("Cumulative Incidence by Age Group")
-# Table
-km_table <- tidy(km_age_fit)
-km_table$time <- year(og + days(km_table$time))
-km_table <- split.data.frame(km_table, km_table$strata)
-
-invisible(lapply(km_table, function(km) {
-  title <- km$strata[1]
-  title <- sub("age_group=", "Ages ", title)
-  gt(km) %>%
-    cols_hide(c("std.error", "strata")) %>%
-    fmt_percent(columns = c("estimate", "conf.low", "conf.high")) %>%
-    cols_merge(
-      columns = c("estimate", "conf.low", "conf.high"),
-      pattern = "{1} [{2} - {3}]"
-    ) %>%
-    cols_label(
-      time = "Year", n.risk = "At Risk", n.event = "CFRD+",
-      n.censor = "Censored", estimate = "Estimated CFRD- [95% CI]"
-    ) %>%
-    opt_row_striping() %>%
-    opt_table_lines() %>%
-    opt_stylize(style = 6, color = "gray") %>%
-    tab_caption(title)
-}))
 ```
@@
   select(-eDWID) %>%
-  tbl_summary(by=OGTTs,missing_text = "Missing") %>%
+  tbl_summary(by = OGTTs, missing_text = "Missing") %>%
   add_p()
@@
 ```{r}
+# Remove those under 10
+annual$age_group <- relevel(annual$age_group, ref = "[10,19)")
+annual <- annual %>% filter(Age_YrEnd != 0, Age_YrEnd >= 10)
+annual$age_group <- droplevels(annual$age_group)
+# Origin for survival analyses
+og <- ymd("2003-01-01")
 # Formula for base model

### 2024-07-11 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
exclude under 10 for all analyses

@@
 )
+# Age groups 10 - 18 inclusive, 19 - 30 inclusive, >= 31
 annual$age_group <- cut(annual$Age_YrEnd,
@@
   ungroup()
+# Remove those under 10
+annual$age_group <- relevel(annual$age_group, ref = "[10,19)")
+annual <- annual %>% filter(Age_YrEnd != 0, Age_YrEnd >= 10)
+annual$age_group <- droplevels(annual$age_group)
 # List of covariates
@@
 
-  new_cases_10 <- sum(inc$n[inc$CFRD == "CFRD+" & inc$age_group == "[-Inf,10)"])
-  at_risk_10 <- sum(inc$n[inc$age_group == "[-Inf,10)"])
-  inc_10 <- new_cases_10 / (at_risk_10 / py)
-
   new_cases_10_18 <- sum(inc$n[inc$CFRD == "CFRD+" & inc$age_group == "[10,19)"])
@@
     `Total Incidence` = total_inc,
-    `Total New Cases Age < 10` = new_cases_10,
-    `Total At Risk Age < 10` = at_risk_10,
-    `Total Incidence Age < 10` = inc_10,
     `Total New Cases Age 10-18` = new_cases_10_18,
@@
   scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
-  scale_color_discrete(name = "Age Group", labels = c("10-17", "18-29", "30+"))
+  scale_color_discrete(name = "Age Group")
 ```
@@
 ```{r}
-# Remove those under 10
-annual$age_group <- relevel(annual$age_group, ref = "[10,19)")
-annual <- annual %>% filter(Age_YrEnd != 0, Age_YrEnd >= 10)
-annual$age_group <- droplevels(annual$age_group)
 # Origin for survival analyses

### 2024-07-12 Jordana Hoppe/Home Collection/manuscript.qmd
figure updates for Jordana

@@
       dark: darkly
-bibliography: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Miscellaneous/american-medical-association.csl
 editor: source
@@
 library(likert)
-switch(Sys.info()[["sysname"]],
-  Windows = {
-    home_dir <- "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF"
-  },
-  Darwin = {
-    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
-  }
+library(patchwork)
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+)
+github_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/GitHub",
+  "Windows" = "C:/Users/Tim/GitHub",
+  "Linux" = "/home/timvigers/GitHub"
 )
@@
 # Use REDCap import
-source("/Users/timvigers/GitHub/CF/Jordana Hoppe/Home Collection/home_collection_data.r")
-source("/Users/timvigers/GitHub/CF/Jordana Hoppe/Home Collection/home_collection_surveys.r")
+source(paste0(github_dir, "/CF/Jordana Hoppe/Home Collection/home_collection_data.r"))
+source(paste0(github_dir, "/CF/Jordana Hoppe/Home Collection/home_collection_surveys.r"))
 # Exclude missing subjects
@@
 
-## Table 3: Clinic Culture Results
+## Table 3: Immediate Culture Results
 
@@
 #| label: tbl-3
-#| tbl-cap: Clinic Culture Results
+#| tbl-cap: Immediate Culture Results
 data %>%
@@
   d <- df[, cols]
-  colnames(d) <- c("Clinic", "Delayed", "Home")
+  colnames(d) <- c("Immediate", "Delayed", "Home")
   d <- d %>%
@@
   cat("\n")
-  cat("#### Clinic vs. delayed (clinic as reference)")
+  cat("#### Immediate vs. delayed (immediate as reference)")
   cat("\n")
@@
   c <- confusionMatrix((d[, "Delayed"]),
-    reference = d[, "Clinic"], positive = "Present", mode = "everything"
+    reference = d[, "Immediate"], positive = "Present", mode = "everything"
   )
@@
   cat("\n")
-  cat("#### Clinic vs. home (clinic as reference)")
+  cat("#### Immediate vs. home (immediate as reference)")
   cat("\n")
@@
   c <- confusionMatrix((d[, "Home"]),
-    reference = d[, "Clinic"], positive = "Present", mode = "everything"
+    reference = d[, "Immediate"], positive = "Present", mode = "everything"
   )
@@
 yn_df <- data.frame(lapply(yn_df, factor, levels = c("No", "Yes")), check.names = F)
-plot(likert(yn_df))
+p <- plot(likert(yn_df))
+ggsave("Jordana Hoppe/Home Collection/Dissemination/Figures/yn.png",
+  plot = p,
+  width = 8, height = 6, units = "in"
+)
 ```
@@
   rename(any_of(setNames(names(labels), labels)))
-plot(likert(check_df))
+colnames(check_df) <- c(
+  "Uncooperative child", "Written instructions for collection were not clear",
+  "I did not think the sample I obtained was a good sample", "Other",
+  "Decreased travel time", "Less missed work",
+  "Financial reasons - costs to travel to clinic can be a burden",
+  "Health concerns with coming to clinic", "Other",
+  "I do not feel comfortable with home collection",
+  "I do not feel collection at home is reliable",
+  "We prefer in-clinic visits only", "Other"
+)
+benefits_df <- check_df[, c(
+  "Decreased travel time", "Less missed work",
+  "Financial reasons - costs to travel to clinic can be a burden",
+  "Health concerns with coming to clinic"
+)]
+barriers_df <- check_df[, c(
+  "Uncooperative child",
+  "I did not think the sample I obtained was a good sample",
+  "Written instructions for collection were not clear",
+  "I do not feel collection at home is reliable",
+  "I do not feel comfortable with home collection"
+)]
+
+p_benefits <- plot(likert(benefits_df)) + ggtitle("Benefits") +
+  theme(
+    axis.text.x = element_blank(), axis.title.x = element_blank(),
+    legend.position = "none"
+  )
+p_barriers <- plot(likert(barriers_df)) + ggtitle("Barriers")
+p <- p_benefits / p_barriers
+ggsave("Jordana Hoppe/Home Collection/Dissemination/Figures/check.png",
+  plot = p,
+  width = 8, height = 6, units = "in"
+)
 ```
@@
   rename(any_of(setNames(names(labels), labels)))
-plot(likert(df3))
+p <- plot(likert(df3))
+ggsave("Jordana Hoppe/Home Collection/Dissemination/Figures/q2.png",
+  plot = p,
+  width = 8, height = 6, units = "in"
+)
 ```
@@
   rename(any_of(setNames(names(labels), labels)))
-plot(likert(df5))
+p <- plot(likert(df5))
+ggsave("Jordana Hoppe/Home Collection/Dissemination/Figures/difficulty.png",
+  plot = p,
+  width = 9, height = 6, units = "in"
+)
 ```

### 2024-07-16 Jordana Hoppe/SSwAB-IT/process_data.qmd
started process data run charts

---
title: "SSwAB-IT Process Data"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
csl: /home/timvigers/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(redcapAPI)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/Tim/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| include: false
# Import from REDCap
unlockREDCap(c(rcon = "SSwAB-IT Process Data"),
  keyring = "API_KEYs",
  envir = 1,
  url = "https://redcap.ucdenver.edu/api/"
)
exportBulkRecords(list(db = rcon), envir = 1)
# Select columns and convert to long
df <- db_sswabit_process_data %>%
  select(cf_center, week, total_swabs_ordered:total_swabs_resulted) %>%
  pivot_longer(cols = starts_with("total_"))
# Nicer names
df$name <- gsub("_", " ", df$name)
df$name <- str_to_title(df$name)
```

# Run chart

```{r}
#| label: fig-run
#| fig-width: 12
#| fig-height: 9
ggplot(df, aes(x = week, y = value, colour = cf_center, group = cf_center)) +
  geom_line() +
  geom_point() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  facet_wrap(vars(name), scales = "free_y") +
  theme(
    legend.title = element_blank(), axis.title.y = element_blank(),
    axis.title.x = element_blank(), text = element_text(size = 20)
  )
```

### 2024-07-17 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.qmd
pre post ETI Chan qmd instead of rmd

--- 2024-05-06 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
@@
 author: "Tim Vigers"
-date: "`r format(Sys.time(), '%d %B %Y')`"
-output:
-  html_document:
-    number_sections: yes
-    toc: yes
-    toc_depth: 5
-    toc_float: yes
+date: "today"
+date-format: long
+format:
+  html:
+    toc: true
+    toc-depth: 5
+    toc-float: true
+    code-fold: true
+    self-contained: true
+    fig-cap-location: top
+    page-layout: full
+    theme:
+      light: flatly
+      dark: darkly
+bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
+csl: /home/timvigers/GitHub/styles/american-medical-association.csl
+editor: source
 ---
@@
 library(tidyverse)
+library(lubridate)
 library(gtsummary)
@@
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Colleen Wood"
 )
@@
 
+```{r data cleaning}
+#| include: false
+# CHCO data
+chco_bmi_pft <- read.csv(
+  "./Prepost Trikafta/Data_Cleaned/chco_bmi_pft.csv",
+  na.strings = ""
+)
+chco_bmi_pft$Date <- mdy(chco_bmi_pft$Date)
+chco_cgm <- read.csv(
+  "./Prepost Trikafta/Data_Cleaned/chco_cgm.csv",
+  na.strings = ""
+)
+chco_cgm$Date <- date(parse_date_time(chco_cgm$date_cgm_placement,
+  orders = c("mdyHM", "mdy")
+))
+chco_glycemic <- read.csv(
+  "./Prepost Trikafta/Data_Cleaned/chco_glycemic.csv",
+  na.strings = ""
+)
+chco_glycemic$Date <- mdy(chco_glycemic$Date)
+chco_registry <- read.csv(
+  "./Prepost Trikafta/Data_Cleaned/chco_registry.csv",
+  na.strings = ""
+)
+# Montana
+montana_bmi_pft <- read.csv(
+  "./Prepost Trikafta/Data_Cleaned/montana_bmi_pft.csv",
+  na.strings = ""
+)
+montana_cgm <- read.csv(
+  "./Prepost Trikafta/Data_Cleaned/montana_cgm.csv",
+  na.strings = ""
+)
+montana_cgm$Date <- date(parse_date_time(montana_cgm$date_cgm_placement,
+  orders = c("mdyHM", "mdy")
+))
+montana_demo <- read.csv(
+  "./Prepost Trikafta/Data_Cleaned/montana_demo.csv",
+  na.strings = ""
+)
+montana_genotypes <- read.csv(
+  "./Prepost Trikafta/Data_Cleaned/montana_genotypes.csv",
+  na.strings = ""
+)
+montana_gtubes <- read.csv(
+  "./Prepost Trikafta/Data_Cleaned/montana_gtubes.csv",
+  na.strings = ""
+)
+montana_pancreatic <- read.csv(
+  "./Prepost Trikafta/Data_Cleaned/montana_pancreatic.csv",
+  na.strings = ""
+)
+montana_glycemic_registry <- read.csv(
+  "./Prepost Trikafta/Data_Cleaned/montana_glycemic_registry.csv",
+  na.strings = ""
+)
+montana_glycemic_registry$Date <- mdy(montana_glycemic_registry$Date)
+# UW
+uw_registry <- read.csv(
+  "./Prepost Trikafta/Data_Cleaned/uw_registry.csv",
+  na.strings = ""
+)
+uw_bmi <- read.csv(
+  "./Prepost Trikafta/Data_Cleaned/uw_bmi.csv",
+  na.strings = ""
+)
+uw_pft <- read.csv(
+  "./Prepost Trikafta/Data_Cleaned/uw_pft.csv",
+  na.strings = ""
+)
+uw_cgm <- read.csv(
+  "./Prepost Trikafta/Data_Cleaned/uw_cgm.csv",
+  na.strings = ""
+)
+uw_demo <- read.csv(
+  "./Prepost Trikafta/Data_Cleaned/uw_demo.csv",
+  na.strings = ""
+)
+uw_genotypes <- read.csv(
+  "./Prepost Trikafta/Data_Cleaned/uw_registry.csv",
+  na.strings = ""
+)
+uw_glycemic <- read.csv(
+  "./Prepost Trikafta/Data_Cleaned/uw_glycemic.csv",
+  na.strings = ""
+)
+uw_gtubes <- read.csv(
+  "./Prepost Trikafta/Data_Cleaned/uw_gtubes.csv",
+  na.strings = ""
+)
+uw_pancreatic <- read.csv(
+  "./Prepost Trikafta/Data_Cleaned/uw_pancreatic.csv",
+  na.strings = ""
+)
+```
+
 ```{r}
-load("./Colleen Wood/Prepost Trikafta/Data_Cleaned/analysis_dataset.rdata")
 # Baseline HbA1c, BMI group pre- and post-modulator

### 2024-07-17 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.qmd
updates for Colleen Wood finished

--- 2024-05-06 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
@@
 
+```{r data cleaning}
+#| include: false
+# # Import CHCO data
+# chco_bmi_pft <- read.csv(
+#   "./Prepost Trikafta/Data_Cleaned/chco_bmi_pft.csv",
+#   na.strings = "", check.names = F
+# )
+# chco_bmi_pft$Date <- mdy(chco_bmi_pft$Date)
+# chco_cgm <- read.csv(
+#   "./Prepost Trikafta/Data_Cleaned/chco_cgm.csv",
+#   na.strings = "", check.names = F
+# )
+# chco_cgm$Date <- date(parse_date_time(chco_cgm$date_cgm_placement,
+#   orders = c("mdyHM", "mdy")
+# ))
+# chco_cgm$`CFF ID` <- as.numeric(sub("_.*", "", chco_cgm$subject_id))
+# chco_cgm$Visit <- sub(".*_", "", chco_cgm$subject_id)
+# chco_glycemic <- read.csv(
+#   "./Prepost Trikafta/Data_Cleaned/chco_glycemic.csv",
+#   na.strings = "", check.names = F
+# )
+# chco_glycemic$Date <- mdy(chco_glycemic$Date)
+# chco_registry <- read.csv(
+#   "./Prepost Trikafta/Data_Cleaned/chco_registry.csv",
+#   na.strings = "", check.names = F
+# )
+# chco_registry$Start <- NULL
+# # Import Montana
+# montana_bmi_pft <- read.csv(
+#   "./Prepost Trikafta/Data_Cleaned/montana_bmi_pft.csv",
+#   na.strings = "", check.names = F
+# )
+# montana_bmi_pft$Date <- mdy(montana_bmi_pft$Date)
+# montana_cgm <- read.csv(
+#   "./Prepost Trikafta/Data_Cleaned/montana_cgm.csv",
+#   na.strings = "", check.names = F
+# )
+# montana_cgm$Date <- date(parse_date_time(montana_cgm$date_cgm_placement,
+#   orders = c("mdyHM", "mdy")
+# ))
+# montana_cgm$`CFF ID` <- as.numeric(sub("_.*", "", montana_cgm$subject_id))
+# montana_cgm$Visit <- sub(".*_", "", montana_cgm$subject_id)
+# montana_cgm$Date <- mdy(montana_cgm$Date)
+# montana_demo <- read.csv(
+#   "./Prepost Trikafta/Data_Cleaned/montana_demo.csv",
+#   na.strings = "", check.names = F
+# )
+# montana_genotypes <- read.csv(
+#   "./Prepost Trikafta/Data_Cleaned/montana_genotypes.csv",
+#   na.strings = "", check.names = F
+# )
+# montana_gtubes <- read.csv(
+#   "./Prepost Trikafta/Data_Cleaned/montana_gtubes.csv",
+#   na.strings = "", check.names = F
+# )
+# montana_gtubes$Date <- NULL
+# montana_pancreatic <- read.csv(
+#   "./Prepost Trikafta/Data_Cleaned/montana_pancreatic.csv",
+#   na.strings = "", check.names = F
+# )
+# montana_pancreatic$Date <- NULL
+# montana_pancreatic$`Pancreatic Status` <- "Insufficient"
+# montana_glycemic_registry <- read.csv(
+#   "./Prepost Trikafta/Data_Cleaned/montana_glycemic_registry.csv",
+#   na.strings = "", check.names = F
+# )
+# montana_glycemic_registry$Date <- mdy(montana_glycemic_registry$Date)
+# # Import UW
+# uw_registry <- read.csv(
+#   "./Prepost Trikafta/Data_Cleaned/uw_registry.csv",
+#   na.strings = "", check.names = F
+# )
+# uw_registry <- uw_registry %>% rename(Date = Start)
+# uw_registry$Date <- mdy(uw_registry$Date)
+# uw_bmi <- read.csv(
+#   "./Prepost Trikafta/Data_Cleaned/uw_bmi.csv",
+#   na.strings = "", check.names = F
+# )
+# uw_bmi$Date <- mdy(uw_bmi$Date)
+# uw_pft <- read.csv(
+#   "./Prepost Trikafta/Data_Cleaned/uw_pft.csv",
+#   na.strings = "", check.names = F
+# )
+# uw_pft$Date <- mdy(uw_pft$Date)
+# uw_cgm <- read.csv(
+#   "./Prepost Trikafta/Data_Cleaned/uw_cgm.csv",
+#   na.strings = "", check.names = F
+# )
+# uw_cgm$`CFF ID` <- as.numeric(sub("_.*", "", uw_cgm$subject_id))
+# uw_cgm$Visit <- sub(".*_", "", uw_cgm$subject_id)
+# uw_demo <- read.csv(
+#   "./Prepost Trikafta/Data_Cleaned/uw_demo.csv",
+#   na.strings = "", check.names = F
+# )
+# uw_genotypes <- read.csv(
+#   "./Prepost Trikafta/Data_Cleaned/uw_registry.csv",
+#   na.strings = "", check.names = F
+# )
+# uw_glycemic <- read.csv(
+#   "./Prepost Trikafta/Data_Cleaned/uw_glycemic.csv",
+#   na.strings = "", check.names = F
+# )
+# uw_glycemic$Date <- mdy(uw_glycemic$Date)
+# uw_gtubes <- read.csv(
+#   "./Prepost Trikafta/Data_Cleaned/uw_gtubes.csv",
+#   na.strings = "", check.names = F
+# )
+# uw_gtubes$`Date of Encounter` <- NULL
+# uw_pancreatic <- read.csv(
+#   "./Prepost Trikafta/Data_Cleaned/uw_pancreatic.csv",
+#   na.strings = "", check.names = F
+# )
+# uw_pancreatic$Date <- NULL
+# # Format and merge
+# # CHCO
+# chco_long <- full_join(chco_glycemic, chco_bmi_pft)
+# chco_long <- full_join(chco_registry, chco_long)
+# chco_long <- full_join(chco_long, chco_cgm)
+# chco_long$site <- "CHCO"
+# chco_long$`Hispanic/Latinx`[chco_long$`Hispanic/Latinx`] <- "Yes"
+# chco_long$`Hispanic/Latinx`[chco_long$`Hispanic/Latinx` == "FALSE"] <- "No"
+# # Montana
+# montana_registry <- full_join(montana_demo, montana_genotypes)
+# montana_registry <- full_join(montana_registry, montana_pancreatic)
+# montana_registry <- full_join(montana_registry, montana_gtubes)
+# montana_long <- full_join(montana_glycemic_registry, montana_bmi_pft)
+# montana_long <- full_join(montana_registry, montana_long)
+# montana_long <- full_join(montana_long, montana_cgm)
+# montana_long <- montana_long %>% rename(HbA1c_original = HbA1c)
+# montana_long <- montana_long %>% mutate(
+#   hba1c_spc = str_detect(HbA1c_original, " "),
+#   HbA1c = as.numeric(ifelse(hba1c_spc, sub(" .*", "", montana_long$HbA1c_original), HbA1c_original))
+# )
+# montana_long <- montana_long %>% select(-c(hba1c_spc, HbA1c_original))
+# montana_long$site <- "Montana"
+# montana_long <- montana_long %>%
+#   rename(`Hispanic/Latinx` = `Is Patient of Hispanic Origin`)
+# # UW
+# uw_bmi_pft <- full_join(uw_bmi, uw_pft)
+# uw_registry <- full_join(uw_registry, uw_demo)
+# uw_registry <- full_join(uw_registry, uw_gtubes)
+# uw_long <- full_join(uw_glycemic, uw_bmi_pft)
+# uw_long <- full_join(uw_registry, uw_long)
+# uw_long <- full_join(uw_long, uw_cgm)
+# uw_long$site <- "UW"
+# uw_long <- uw_long %>%
+#   rename(`Hispanic/Latinx` = `Is Patient of Hispanic Origin`)
+# # Put everything together
+# analysis_data <- full_join(chco_long, montana_long)
+# analysis_data <- full_join(analysis_data, uw_long)
+# # Race
+# analysis_data$Race <- factor(analysis_data$Race,
+#   levels = c(
+#     "American Indian or Alaska Native", "Black", "Black or African American",
+#     "Caucasian", "Other", "Some other race", "Two or more races", "Unknown",
+#     "White"
+#   ),
+#   labels = c(
+#     "Other", "Black", "Black", "White", "Other", "Other", "Other",
+#     "Unknown", "White"
+#   )
+# )
+# df <- analysis_data
+load("/home/timvigers/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/analysis_dataset.rdata")
+# Ethnicity
+chco_demo <- read.csv("/home/timvigers/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Raw/Data From CHCO/Registry/Copy of Earliest Trikafta Dates in Registry_updated 2.2.21 cc modified_cfrdyn.csv")
+chco_demo$Hispanic <- NA
+chco_demo$Hispanic[chco_demo$Hispanic.Latinx] <- "Yes"
+chco_demo$Hispanic[!chco_demo$Hispanic.Latinx] <- "No"
+chco_demo <- chco_demo[, c("CFF.ID", "Hispanic")]
+
+montana_demo <- read.csv("/home/timvigers/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/montana_demo.csv")
+montana_demo$Hispanic <- montana_demo$Is.Patient.of.Hispanic.Origin
+montana_demo <- montana_demo[, c("CFF.ID", "Hispanic")]
+
+uw_demo <- read.csv("/home/timvigers/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/uw_demo.csv")
+uw_demo$Hispanic <- uw_demo$Is.Patient.of.Hispanic.Origin
+uw_demo <- uw_demo[, c("CFF.ID", "Hispanic")]
+
+eth <- do.call(rbind, list(chco_demo, montana_demo, uw_demo))
+colnames(eth) <- c("CFF ID", "Hispanic")
+
+df <- left_join(df, eth)
+```
+
 ```{r}
-load("./Colleen Wood/Prepost Trikafta/Data_Cleaned/analysis_dataset.rdata")
 # Baseline HbA1c, BMI group pre- and post-modulator
@@
   ungroup()
-df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf), right = F)
+df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf),
+  right = F
+)
 df$`Baseline BMI` <- cut(df$`Baseline BMI`, c(-Inf, 85, Inf), right = F)
@@
     ppFVCgl, `G-tube in past 12 months`, `Pancreatic Status`, Site,
-    Genotype, `Age Group`, Race
+    Genotype, `Age Group`, Race,Hispanic
   ) %>%
   tbl_summary(
-    statistic = list(c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ "{mean} ± {sd}"),
+    statistic = list(c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ 
+                       "{mean} ± {sd}"),
     missing_text = "Missing"
@@
 ```{r}
-cgm <- df[!is.na(df$average_sensor), ]
+cgm <- df[!is.na(df$mean_sensor), ]
 cgm %>%

### 2024-07-17 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_merge.Rmd
updates for Colleen Wood finished

@@
 library(stringr)
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Colleen Wood"
+)
 knitr::opts_chunk$set(echo = FALSE)
-if(Sys.info()["sysname"] == "Windows"){
-  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood"
-} else if (Sys.info()["sysname"] == "Linux"){
-  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood"
-} else if (Sys.info()["sysname"] == "Darwin"){
-  home_dir = "/Volumes/Documents/Work/Vigers/CF/Colleen Wood"
-}
 knitr::opts_knit$set(root.dir = home_dir)
@@
 # 8 vars MRN - ID, date for visits; start - ? start of trt?, fev, bmi, height, weight, fec 
-chco_bmi_pft = read_csv("./Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv")
+chco_bmi_pft = read_csv("./Prepost Trikafta/Data_Cleaned/chco_bmi_pft.csv")
 chco_bmi_pft$Date = mdy(chco_bmi_pft$Date) 
@@
 # 105 vars subjectid - number_pre/post for visit and a date; numbers dont seem to correspond to mrn in bmi pft
-chco_cgm = read_csv("./Prepost Triakfta/Data_Cleaned/chco_cgm.csv")
+chco_cgm = read_csv("./Prepost Trikafta/Data_Cleaned/chco_cgm.csv")
 chco_cgm$Date = date(parse_date_time(chco_cgm$date_cgm_placement,
@@
 # 7 var MRN and CFF id (=subjectid?), date for visits; start... etc
-chco_glycemic = read_csv("./Prepost Triakfta/Data_Cleaned/chco_glycemic.csv")
+chco_glycemic = read_csv("./Prepost Trikafta/Data_Cleaned/chco_glycemic.csv")
 chco_glycemic$Date = mdy(chco_glycemic$Date)
@@
 # CFF ID and MRN, start date
-chco_registry = read_csv("./Prepost Triakfta/Data_Cleaned/chco_registry.csv")
+chco_registry = read_csv("./Prepost Trikafta/Data_Cleaned/chco_registry.csv")
 ###################################
@@
 # bmi, pulm function 975 obs
-montana_bmi_pft <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_bmi_pft.csv")
+montana_bmi_pft <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_bmi_pft.csv")
 
 # cgm 16 obs
-montana_cgm <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_cgm.csv")
+montana_cgm <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_cgm.csv")
 montana_cgm$Date = date(parse_date_time(montana_cgm$date_cgm_placement,
@@
 # registry: we need to combine demo (126), genotypes (126), pancreatic (241), gtubes (155)
-montana_demo <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_demo.csv")
-montana_genotypes <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_genotypes.csv")
-montana_gtubes <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_gtubes.csv") # what are the dates here
-montana_pancreatic <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_pancreatic.csv") # what is this? just dates and cffid
+montana_demo <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_demo.csv")
+montana_genotypes <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_genotypes.csv")
+montana_gtubes <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_gtubes.csv") # what are the dates here
+montana_pancreatic <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_pancreatic.csv") # what is this? just dates and cffid
 # ask tim:what do we need to do with gtubes and pancreatic dates? dont look like chco data
@@
 # glycemic 279
-montana_glycemic_registry <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv")
+montana_glycemic_registry <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_glycemic_registry.csv")
 montana_glycemic_registry$Date = mdy(montana_glycemic_registry$Date)
@@
 ###### IMPORT UW DATA #############
-uw_registry <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_registry.csv") #cffid and start
-uw_bmi <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_bmi.csv") # date, cffid, height, bmi (height bmi need rename)
-uw_pft <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_pft.csv") # datee, cffid, fev, fvc need renaming
-uw_cgm <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_cgm.csv") # same code as montana
-uw_demo <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_demo.csv") 
-uw_genotypes <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_registry.csv") # no genotype data here?
-uw_glycemic <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_glycemic.csv") # looks good, same as chco
-uw_gtubes <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_gtubes.csv") # rename date of encounter to date
-uw_pancreatic <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_pancreatic.csv") # same as montana, no pancreatic data though
+uw_registry <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_registry.csv") #cffid and start
+uw_bmi <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_bmi.csv") # date, cffid, height, bmi (height bmi need rename)
+uw_pft <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_pft.csv") # datee, cffid, fev, fvc need renaming
+uw_cgm <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_cgm.csv") # same code as montana
+uw_demo <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_demo.csv") 
+uw_genotypes <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_registry.csv") # no genotype data here?
+uw_glycemic <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_glycemic.csv") # looks good, same as chco
+uw_gtubes <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_gtubes.csv") # rename date of encounter to date
+uw_pancreatic <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_pancreatic.csv") # same as montana, no pancreatic data though
 
@@
 ###### IMPORT UM DATA ########
-#um_all <- read_csv("./Prepost Triakfta/Data_Cleaned/um_all.csv")
+#um_all <- read_csv("./Prepost Trikafta/Data_Cleaned/um_all.csv")
 ###################################
@@
 # rename the bmi_pft to chco names
-montana_bmi_pft = montana_bmi_pft %>% rename(FVC = `Measure of FVC`,
-                                             FEV1 = `Measure of FEV1`,
-                                             Height = `Patients height`,
-                                             Weight = `Weight of Patient`)
+# montana_bmi_pft = montana_bmi_pft %>% rename(FVC = `Measure of FVC`,
+#                                              FEV1 = `Measure of FEV1`,
+#                                              Height = `Patients height`,
+#                                              Weight = `Weight of Patient`)
 montana_bmi_pft$Date = mdy(montana_bmi_pft$Date)
@@
 # rename the bmi_pft to chco names
-uw_bmi_pft = uw_bmi_pft %>% rename(FVC = `Measure of FVC`,
-                                   FEV1 = `Measure of FEV1`,
-                                   Height = `Patients height`,
-                                   BMI = `BMI Value`)
+# uw_bmi_pft = uw_bmi_pft %>% rename(FVC = `Measure of FVC`,
+#                                    FEV1 = `Measure of FEV1`,
+#                                    Height = `Patients height`,
+#                                    BMI = `BMI Value`)
 
@@
 # export csv analysis_data
-write.csv(analysis_data, file = "./Prepost Triakfta/Data_Cleaned/analysis_data.csv")
+write.csv(analysis_data, file = "./Prepost Trikafta/Data_Cleaned/analysis_data.csv")
 ```

### 2024-07-17 Colleen Wood/Pre-post Trikafta/clean_and_merge.R

@@ -2,16 +2,16 @@ library(tidyverse)
 library(lubridate)
 library(childsds)
 library(rspiro)
-source("/Users/timvigers/GitHub/CF/genotype_class_severity.R")
+source("/home/timvigers/GitHub/CF/genotype_class_severity.R")
 t1_variables <- c(
   "CFF ID", "Site", "Sex", "Race", "Genotypes1", "Genotypes2", "Pancreatic Status",
   "G-tube in past 12 months", "Time", "Age",
   "BMI Percentile", "Height", "Weight", "FEV1", "FVC", "HbA1c"
 )
 # CHCO
-chco_a1c <- read_csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_glycemic.csv", show_col_types = F)
+chco_a1c <- read_csv("/home/timvigers/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/chco_glycemic.csv", show_col_types = F)
 chco_a1c$Date <- parse_date_time(chco_a1c$Date, orders = "mdy")
-chco_bmi <- read_csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv", show_col_types = F)
+chco_bmi <- read_csv("/home/timvigers/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/chco_bmi_pft.csv", show_col_types = F)
 chco_bmi$Date <- parse_date_time(chco_bmi$Date, orders = "mdy")
 chco <- full_join(chco_a1c, chco_bmi)
 # Fill CFF ID by MRN
@@ -22,13 +22,13 @@ chco <- chco %>%
   filter(!is.na(`CFF ID`)) %>%
   select(-MRN)
 # Add demographics, filter
-chco_reg <- read_csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_registry.csv", show_col_types = F)
[…]

### 2024-07-22 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
ETI outcomes flowchart

@@
 library(gtsummary)
+library(ggflowchart)
 # Paths
@@
 n_people_1 <- length(unique(encounter$eDWID))
-n_hosp_1 <- nrow(hospitalizations)
-n_people_hosp_1 <- length(unique(hospitalizations$eDWID))
-# Lung function cutoffs
-lower <- 20
-upper <- 150
-# Remove impossible lung function values
-low_fev1 = which(encounter$gli_fev1_ppred_rn)
-encounter$gli_fev1_ppred_rn[encounter$gli_fev1_ppred_rn < 20 |
-  encounter$gli_fev1_ppred_rn > 150] <- NA
-encounter$gli_fvc_ppred_rn[encounter$gli_fvc_ppred_rn < 20 |
-  encounter$gli_fvc_ppred_rn > 150] <- NA
-encounter$gli_fev1fvc_ppred_rn[encounter$gli_fev1fvc_ppred_rn < 20 |
-  encounter$gli_fev1fvc_ppred_rn > 150] <- NA
-# Filter
-encounter <- encounter %>%
-  filter(
-    !is.na(Days),
-    Days >= -365.25 * 3 & Days <= 365.25 * 2,
-    hospitalized == "No", !encounterlocation %in% c("Hospital", "Home IV"),
-    pregnant != 1,
-    !is.na(Modulator_trikafta_first_date)
-  )
+no_eti_date <- which(is.na(encounter$Modulator_trikafta_first_date))
+n_people_no_eti_date <- length(unique(encounter$eDWID[no_eti_date]))
+# Remove those without an ETI start date
+encounter <- encounter[-no_eti_date, ]
+n_enc_2 <- nrow(encounter)
+n_people_2 <- length(unique(encounter$eDWID))
+# Filter out hospitalizations
+hosp <- which(encounter$hospitalized == "Yes" |
+  encounter$encounterlocation %in% c("Hospital", "Home IV"))
+n_people_hosp <- length(unique(encounter$eDWID[hosp]))
+encounter <- encounter[-hosp, ]
+n_enc_3 <- nrow(encounter)
+n_people_3 <- length(unique(encounter$eDWID))
+# Pregnancies
+preg <- which(encounter$pregnant == 1)
+n_people_preg <- length(unique(encounter$eDWID[preg]))
+encounter <- encounter[-preg, ]
+n_enc_4 <- nrow(encounter)
+n_people_4 <- length(unique(encounter$eDWID))
+# Days from ETI start
+early <- which(encounter$Days < -365.25 * 3)
+n_people_early <- length(unique(encounter$eDWID[early]))
+late <- which(encounter$Days > 365.25 * 2)
+n_people_late <- length(unique(encounter$eDWID[late]))
+encounter <- encounter[-early, ]
+encounter <- encounter[-late, ]
+n_enc_5 <- nrow(encounter)
+n_people_5 <- length(unique(encounter$eDWID))
 # Get some summary statistics by participant
@@
 no_post <- t1_participant$eDWID[t1_participant$`Number of Encounters Post-ETI` == 0]
-encounter <- encounter %>% filter(!eDWID %in% no_post)
+no_post_visits <- which(encounter$eDWID %in% no_post)
+encounter <- encounter[-no_post_visits, ]
+n_enc_6 <- nrow(encounter)
+n_people_6 <- length(unique(encounter$eDWID))
 t1_participant <- t1_participant %>% filter(!eDWID %in% no_post)
+# Lung function cutoffs
+lower <- 20
+upper <- 150
+# Remove impossible lung function values
+low_fev1 <- which(encounter$gli_fev1_ppred_rn < lower)
+high_fev1 <- which(encounter$gli_fev1_ppred_rn > upper)
+encounter$gli_fev1_ppred_rn[low_fev1] <- NA
+encounter$gli_fev1_ppred_rn[high_fev1] <- NA
+
+low_fvc <- which(encounter$gli_fvc_ppred_rn < lower)
+high_fvc <- which(encounter$gli_fvc_ppred_rn > upper)
+encounter$gli_fvc_ppred_rn[low_fvc] <- NA
+encounter$gli_fvc_ppred_rn[high_fvc] <- NA
 ```
@@
 
-1. Lung function percent predicted values < 20% or > 150% were excluded.
-  - `r sum(encounter$gli_fev1_ppred_rn < 20,na.rm = T)` ppFEV1 values < 20% and `r sum(encounter$gli_fev1_ppred_rn > 150,na.rm = T)` over 150%.
-2. Limited to 3 years before and 2 years after ETI start.
-  - Based on `Modulator_trikafta_first_date` variable provided by CFFPR.
-3. Hospitalizations and home IV episodes excluded.
-  - Based on the care episodes dataset, there were `r length(unique(encounter$eDWID[encounter$hospitalized=="Yes"]))` unique participants with at least one care episode and a total of `r sum(encounter$hospitalized=="Yes")` encounters that were part of a care episode.
-  - After filtering out these episodes, there were still 914 encounters with `encounterlocation` variable indicating either hospitalization or home IV, so these were also excluded.
-4. Years of pregnancy excluded (we only have annualized pregnancy data, so don't know exact dates for this).
-  - `r length(unique(encounter$eDWID[encounter$pregnant==1]))` unique participants with at least one care episode and a total of `r sum(encounter$pregnant==1, na.rm = T)` encounters during a pregnancy year.
-5.
-Those missing `Modulator_trikafta_first_date` were automatically excluded from the models because they don't have a value for `Days`.
-This naturally means that people not eligible for Trikafta are not included in the models.
-6. 990 people with no encounters post-ETI were excluded.
-
-```{r filtering}
-# Additional quick cleaning steps
-
+```{r flowchart}
+# Make labels
+tots1 <- paste0(n_people_1, " People\n", n_enc_1, " Encounters")
+no_date <- paste0(
+  "No ETI Date:\n", n_people_no_eti_date, " People\n",
+  length(no_eti_date), " Encounters"
+)
+tots2 <- paste0(n_people_2, " People\n", n_enc_2, " Encounters")
+hosps <- paste0(
+  "Care Episodes:\n", n_people_hosp, " People\n",
+  length(hosp), " Encounters"
+)
+tots3 <- paste0(n_people_3, " People\n", n_enc_3, " Encounters")
+pregs <- paste0(
+  "Pregnancies:\n", n_people_preg, " People\n",
+  length(preg), " Encounters"
+)
+tots4 <- paste0(n_people_4, " People\n", n_enc_4, " Encounters")
+earlies <- paste0(
+  "Visits > ", 365.25 * 3, " Days BEFORE ETI:\n", n_people_early, " People\n",
+  length(early), " Encounters"
+)
+lates <- paste0(
+  "Visits > ", 365.25 * 2, " Days AFTER ETI:\n", n_people_late, " People\n",
+  length(late), " Encounters"
+)
+tots5 <- paste0(n_people_5, " People\n", n_enc_5, " Encounters")
+no_posts <- paste0(
+  "No Post-ETI Visits:\n", length(no_post), " People\n",
+  length(no_post_visits), " Encounters (Pre-ETI)"
+)
+tots6 <- paste0(n_people_6, " People\n", n_enc_6, " Encounters")
+fev1s <- paste0(
+  length(low_fev1), " FEV1 < ", lower, "\n",
+  length(high_fev1), " FEV1 > ", upper
+)
+fvcs <- paste0(
+  length(low_fvc), " FVC < ", lower, "\n",
+  length(high_fvc), " FVC > ", upper
+)
+# Assemble the chart!
+flow_chart <- tibble(
+  from = c(
+    tots1, tots1, tots2, tots2, tots3, tots3, tots4, tots4, tots4, tots5,
+    tots5, tots6, tots6
+  ),
+  to = c(
+    tots2, no_date, tots3, hosps, tots4, pregs, tots5, earlies, lates, tots6,
+    no_posts, fev1s, fvcs
+  )
+)
+ggflowchart(flow_chart)
 ```
@@
 
-# LOESS-smoothed outcome plots
+<!-- # LOESS-smoothed outcome plots -->
 
-```{r functions}
-# Plots
-loess_plot <- function(data = encounter, outcome, facet = F) {
-  p <- ggplot(data, aes(x = Days, y = !!sym(outcome), color = Race)) +
-    geom_point(
-      shape = ".", aes(x = Days, y = !!sym(outcome)), inherit.aes = F,
-      alpha = 0.1
-    ) +
-    geom_smooth(method = "loess", se = F) +
-    theme_classic()
-  if (facet) {
-    p <- p + facet_wrap(~Race)
-  }
-  return(p)
-}
-```
+<!-- ```{r functions} -->
+<!-- # Plots -->
+<!-- loess_plot <- function(data = encounter, outcome, facet = F) { -->
+<!--   p <- ggplot(data, aes(x = Days, y = !!sym(outcome), color = Race)) + -->
+<!--     geom_point( -->
+<!--       shape = ".", aes(x = Days, y = !!sym(outcome)), inherit.aes = F, -->
+<!--       alpha = 0.1 -->
+<!--     ) + -->
+<!--     geom_smooth(method = "loess", se = F) + -->
+<!--     theme_classic() -->
+<!--   if (facet) { -->
+<!--     p <- p + facet_wrap(~Race) -->
+<!--   } -->
+<!--   return(p) -->
+<!-- } -->
+<!-- ``` -->
 
-```{r results='asis'}
-#| warning: false
-invisible(lapply(continuous_outcomes, function(o) {
-  cat("\n")
-  cat("\n")
-  cat("##", o)
-  cat("\n")
-  cat("\n")
-  print(loess_plot(outcome = o))
-}))
-```
+<!-- ```{r results='asis'} -->
+<!-- #| warning: false -->
+<!-- invisible(lapply(continuous_outcomes, function(o) { -->
+<!--   cat("\n") -->
+<!--   cat("\n") -->
+<!--   cat("##", o) -->
+<!--   cat("\n") -->
+<!--   cat("\n") -->
+<!--   print(loess_plot(outcome = o)) -->
+<!-- })) -->
+<!-- ``` -->
 
@@
 The following models are adjusted for age group and insurance status. We fit a Gaussian mixed model for continuous outcomes with random intercept for each subject.
-
-## With Asian subgroup
-
-```{r}
-invisible(lapply(continuous_outcomes, function(o) {
-  f <- as.formula(paste0(
-    o,
-    "~ age_group + Insurance + bs(Days, knots = 0, degree = 1) * Race + (1 | eDWID)"
-  ))
-  m <- lmer(f, data = encounter)
-  predict_response(model = m, terms = c("Days[all]", "Race")) %>%
-    plot() %>%
-    print()
-}))
-```
-
-## Without Asian subgroup (for clearer plotting)
 

### 2024-07-22 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R

@@ -48,7 +48,7 @@ demo$Race <- factor(demo$Race,
   ),
   labels = c(
     "Non-Hispanic White", "Black or African American", "Hispanic or Latino",
-    "Asian", "Other", "Other", "Other"
+    "Other", "Other", "Other", "Other"
   )
 )
 # Insurance
@@ -140,6 +140,6 @@ binary_outcomes <- c(
   "burkho_complex"
 )
 # Save
-save(encounter, continuous_outcomes, binary_outcomes,
+save(encounter, hospitalizations, continuous_outcomes, binary_outcomes,
   file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData"
 )

### 2024-07-22 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
uncommented models section

@@
 
-<!-- # LOESS-smoothed outcome plots -->
+# LOESS-smoothed outcome plots
 
-<!-- ```{r functions} -->
-<!-- # Plots -->
-<!-- loess_plot <- function(data = encounter, outcome, facet = F) { -->
-<!--   p <- ggplot(data, aes(x = Days, y = !!sym(outcome), color = Race)) + -->
-<!--     geom_point( -->
-<!--       shape = ".", aes(x = Days, y = !!sym(outcome)), inherit.aes = F, -->
-<!--       alpha = 0.1 -->
-<!--     ) + -->
-<!--     geom_smooth(method = "loess", se = F) + -->
-<!--     theme_classic() -->
-<!--   if (facet) { -->
-<!--     p <- p + facet_wrap(~Race) -->
-<!--   } -->
-<!--   return(p) -->
-<!-- } -->
-<!-- ``` -->
+```{r functions}
+# Plots
+loess_plot <- function(data = encounter, outcome, facet = F) {
+  p <- ggplot(data, aes(x = Days, y = !!sym(outcome), color = Race)) +
+    geom_point(
+      shape = ".", aes(x = Days, y = !!sym(outcome)), inherit.aes = F,
+      alpha = 0.1
+    ) +
+    geom_smooth(method = "loess", se = F) +
+    theme_classic()
+  if (facet) {
+    p <- p + facet_wrap(~Race)
+  }
+  return(p)
+}
+```
 
-<!-- ```{r results='asis'} -->
-<!-- #| warning: false -->
-<!-- invisible(lapply(continuous_outcomes, function(o) { -->
-<!--   cat("\n") -->
-<!--   cat("\n") -->
-<!--   cat("##", o) -->
-<!--   cat("\n") -->
-<!--   cat("\n") -->
-<!--   print(loess_plot(outcome = o)) -->
-<!-- })) -->
-<!-- ``` -->
+```{r results='asis'}
+#| warning: false
+invisible(lapply(continuous_outcomes, function(o) {
+  cat("\n")
+  cat("\n")
+  cat("##", o)
+  cat("\n")
+  cat("\n")
+  print(loess_plot(outcome = o))
+}))
+```
 

### 2024-07-22 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
parallelize ETI outcomes

@@
 library(ggflowchart)
+library(parallel)
 # Paths
@@
 
-```{r results='asis'}
-#| warning: false
-invisible(lapply(continuous_outcomes, function(o) {
+```{r}
+cl <- makeForkCluster(detectCores() / 2)
+invisible(parLapply(cl, continuous_outcomes, function(o) {
   cat("\n")
@@
 ```{r}
-invisible(lapply(continuous_outcomes, function(o) {
+invisible(parLapply(cl, continuous_outcomes, function(o) {
   f <- as.formula(paste0(

### 2024-07-24 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
separate R script for mixed models

--- 2024-07-22 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
@@
 
-```{r filtering and flowchart}
-# Variables for flowchart
-n_enc_1 <- nrow(encounter)
-n_people_1 <- length(unique(encounter$eDWID))
-no_eti_date <- which(is.na(encounter$Modulator_trikafta_first_date))
-n_people_no_eti_date <- length(unique(encounter$eDWID[no_eti_date]))
-# Remove those without an ETI start date
-encounter <- encounter[-no_eti_date, ]
-n_enc_2 <- nrow(encounter)
-n_people_2 <- length(unique(encounter$eDWID))
-# Filter out hospitalizations
-hosp <- which(encounter$hospitalized == "Yes" |
-  encounter$encounterlocation %in% c("Hospital", "Home IV"))
-n_people_hosp <- length(unique(encounter$eDWID[hosp]))
-encounter <- encounter[-hosp, ]
-n_enc_3 <- nrow(encounter)
-n_people_3 <- length(unique(encounter$eDWID))
-# Pregnancies
-preg <- which(encounter$pregnant == 1)
-n_people_preg <- length(unique(encounter$eDWID[preg]))
-encounter <- encounter[-preg, ]
-n_enc_4 <- nrow(encounter)
-n_people_4 <- length(unique(encounter$eDWID))
-# Days from ETI start
-early <- which(encounter$Days < -365.25 * 3)
-n_people_early <- length(unique(encounter$eDWID[early]))
-late <- which(encounter$Days > 365.25 * 2)
-n_people_late <- length(unique(encounter$eDWID[late]))
-encounter <- encounter[-early, ]
-encounter <- encounter[-late, ]
-n_enc_5 <- nrow(encounter)
-n_people_5 <- length(unique(encounter$eDWID))
-# Get some summary statistics by participant
-t1_participant <- encounter %>%
-  group_by(eDWID) %>%
-  summarise(
-    `Years With Data` = length(unique(reviewyear)),
-    `Number of Encounters` = n(),
-    `Number of Encounters Pre-ETI` = sum(Days <= 0),
-    `Number of Encounters Post-ETI` = sum(Days > 0),
-    Age = mean(encounterage, na.rm = T),
-    across(all_of(continuous_outcomes), \(x) mean(x, na.rm = TRUE)),
-    Race = names(sort(table(Race), decreasing = TRUE))[1],
-    Insurance = names(sort(table(Insurance), decreasing = TRUE))[1],
-    .groups = "drop"
-  )
-no_post <- t1_participant$eDWID[t1_participant$`Number of Encounters Post-ETI` == 0]
-no_post_visits <- which(encounter$eDWID %in% no_post)
-encounter <- encounter[-no_post_visits, ]
-n_enc_6 <- nrow(encounter)
-n_people_6 <- length(unique(encounter$eDWID))
-t1_participant <- t1_participant %>% filter(!eDWID %in% no_post)
-# Lung function cutoffs
-lower <- 20
-upper <- 150
-# Remove impossible lung function values
-low_fev1 <- which(encounter$gli_fev1_ppred_rn < lower)
-high_fev1 <- which(encounter$gli_fev1_ppred_rn > upper)
-encounter$gli_fev1_ppred_rn[low_fev1] <- NA
-encounter$gli_fev1_ppred_rn[high_fev1] <- NA
-
-low_fvc <- which(encounter$gli_fvc_ppred_rn < lower)
-high_fvc <- which(encounter$gli_fvc_ppred_rn > upper)
-encounter$gli_fvc_ppred_rn[low_fvc] <- NA
-encounter$gli_fvc_ppred_rn[high_fvc] <- NA
-```
-
 # Data cleaning
@@
 ```{r flowchart}
-# Make labels
-tots1 <- paste0(n_people_1, " People\n", n_enc_1, " Encounters")
-no_date <- paste0(
-  "No ETI Date:\n", n_people_no_eti_date, " People\n",
-  length(no_eti_date), " Encounters"
-)
-tots2 <- paste0(n_people_2, " People\n", n_enc_2, " Encounters")
-hosps <- paste0(
-  "Care Episodes:\n", n_people_hosp, " People\n",
-  length(hosp), " Encounters"
-)
-tots3 <- paste0(n_people_3, " People\n", n_enc_3, " Encounters")
-pregs <- paste0(
-  "Pregnancies:\n", n_people_preg, " People\n",
-  length(preg), " Encounters"
-)
-tots4 <- paste0(n_people_4, " People\n", n_enc_4, " Encounters")
-earlies <- paste0(
-  "Visits > ", 365.25 * 3, " Days BEFORE ETI:\n", n_people_early, " People\n",
-  length(early), " Encounters"
-)
-lates <- paste0(
-  "Visits > ", 365.25 * 2, " Days AFTER ETI:\n", n_people_late, " People\n",
-  length(late), " Encounters"
-)
-tots5 <- paste0(n_people_5, " People\n", n_enc_5, " Encounters")
-no_posts <- paste0(
-  "No Post-ETI Visits:\n", length(no_post), " People\n",
-  length(no_post_visits), " Encounters (Pre-ETI)"
-)
-tots6 <- paste0(n_people_6, " People\n", n_enc_6, " Encounters")
-fev1s <- paste0(
-  length(low_fev1), " FEV1 < ", lower, "\n",
-  length(high_fev1), " FEV1 > ", upper
-)
-fvcs <- paste0(
-  length(low_fvc), " FVC < ", lower, "\n",
-  length(high_fvc), " FVC > ", upper
-)
-# Assemble the chart!
-flow_chart <- tibble(
-  from = c(
-    tots1, tots1, tots2, tots2, tots3, tots3, tots4, tots4, tots4, tots5,
-    tots5, tots6, tots6
-  ),
-  to = c(
-    tots2, no_date, tots3, hosps, tots4, pregs, tots5, earlies, lates, tots6,
-    no_posts, fev1s, fvcs
-  )
-)
 ggflowchart(flow_chart)
@@
 
-```{r results='asis'}
-#| warning: false
+```{r}
 invisible(lapply(continuous_outcomes, function(o) {
@@
 ```{r}
-invisible(lapply(continuous_outcomes, function(o) {
-  f <- as.formula(paste0(
-    o,
-    "~ age_group + Insurance + bs(Days, knots = 0, degree = 1) * Race + (1 | eDWID)"
-  ))
-  m <- lmer(f, data = encounter %>% filter(Race != "Asian"))
-  predict_response(model = m, terms = c("Days[all]", "Race")) %>%
-    plot() %>%
-    print()
-}))
+
 ```
@@
 
-1. Do we want to combine the Asian and "Other" groups? It makes sense from a statistical point of view.
+1.

### 2024-07-24 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R

@@ -21,6 +21,15 @@ demo <- read.csv("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw
 hospitalizations <- read.csv("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_CareEpisodes_Del1.csv",
   na.strings = ""
 )
+# List of outcomes
+continuous_outcomes <- c(
+  "gli_fev1_ppred_rn", "gli_fvc_ppred_rn", "gli_fev1fvc_ppred_rn",
+  "bmipercentile", "heightpercentile", "weightpercentile"
+)
+binary_outcomes <- c(
+  "staphylococcus_aureus", "haemophilus_influenzae", "pseudomonasaeruginosa",
+  "burkho_complex"
+)
 # Race variable
 races <- list(
   "Race1" = "White",
@@ -130,16 +139,122 @@ stopCluster(cl)
 hosps <- unique(unlist(hosps))
 encounter$hospitalized <- "No"
 encounter$hospitalized[hosps] <- "Yes"
-# List of outcomes
-continuous_outcomes <- c(
-  "gli_fev1_ppred_rn", "gli_fvc_ppred_rn", "gli_fev1fvc_ppred_rn",
-  "bmipercentile", "heightpercentile", "weightpercentile"
+# Variables for flowchart
+n_enc_1 <- nrow(encounter)
+n_people_1 <- length(unique(encounter$eDWID))
+no_eti_date <- which(is.na(encounter$Modulator_trikafta_first_date))
+n_people_no_eti_date <- length(unique(encounter$eDWID[no_eti_date]))
+# Remove those without an ETI start date
+encounter <- encounter[-no_eti_date, ]
+n_enc_2 <- nrow(encounter)
+n_people_2 <- length(unique(encounter$eDWID))
+# Filter out hospitalizations
+hosp <- which(encounter$hospitalized == "Yes" |
+  encounter$encounterlocation %in% c("Hospital", "Home IV"))
+n_people_hosp <- length(unique(encounter$eDWID[hosp]))
+encounter <- encounter[-hosp, ]
+n_enc_3 <- nrow(encounter)
+n_people_3 <- length(unique(encounter$eDWID))
+# Pregnancies
+preg <- which(encounter$pregnant == 1)
+n_people_preg <- length(unique(encounter$eDWID[preg]))
+encounter <- encounter[-preg, ]
[…]

### 2024-07-24 Kamyron Jordan/Racial and Ethnic Differences in ETI/outcomes_mixed_models.R

@@ -1,0 +1,22 @@
+library(lme4)
+library(ggeffects)
+# Load dataset
+load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData")
+# Make a dataframe for the levels we want predictions at
+new_df <- expand.grid(
+  Days = c(min(encounter$Days), 0, max(encounter$Days)),
+  Race = unique(encounter$Race)
+)
+# Fit models
+results <- lapply(continuous_outcomes, function(o) {
+  # Formula
+  f <-
+    as.formula(paste0(
+      o,
+      "~ age_group + Insurance + bs(Days, knots = 0, degree = 1) * Race + (1 | eDWID)"
+    ))
+  m <- lmer(f, data = encounter)
+  # Predicted values
+  pred <- predict_response(m, terms = new_df)
+  return(list(model = m, predictions = pred))
+})

### 2024-07-25 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
mixed model predictions still not working

@@
 library(tidyverse)
-library(lmerTest)
-library(splines)
-library(ggeffects)
 library(gtsummary)
@@
 ```{r}
-
 ```

### 2024-07-25 Kamyron Jordan/Racial and Ethnic Differences in ETI/outcomes_mixed_models.R

@@ -1,22 +1,29 @@
 library(lme4)
+library(splines)
 library(ggeffects)
+# Paths
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+)
+setwd(home_dir)
 # Load dataset
 load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData")
-# Make a dataframe for the levels we want predictions at
-new_df <- expand.grid(
-  Days = c(min(encounter$Days), 0, max(encounter$Days)),
-  Race = unique(encounter$Race)
-)
 # Fit models
 results <- lapply(continuous_outcomes, function(o) {
   # Formula
   f <-
     as.formula(paste0(
       o,
-      "~ age_group + Insurance + bs(Days, knots = 0, degree = 1) * Race + (1 | eDWID)"
+      "~ bs(Days, knots = 0, degree = 1) * Race + (1 | eDWID)"
     ))
   m <- lmer(f, data = encounter)
   # Predicted values
-  pred <- predict_response(m, terms = new_df)
+  pred <- predict_response(m, terms = c("Days[-1095,0,730]", "Race"))
   return(list(model = m, predictions = pred))
 })
+# Save
+save(results,
+  file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_lmm_results.RData"
+)

### 2024-07-30 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
screening rates figure

@@
   pivot_longer(`Screened HbA1c`:`Screened Either`)
-ggplot(screening, aes(x = ReviewYear, y = value, color = age_group)) +
+screening_plot <- ggplot(screening, aes(x = ReviewYear, y = value, color = age_group)) +
   geom_line() +
+  xlab("Year") +
   facet_wrap(~name) +
@@
   scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
-  scale_color_discrete(name = "Age Group")
+  scale_color_discrete(name = "Age Group") +
+  theme(text = element_text(size = 20), axis.title.y = element_blank())
+ggsave("./Dissemination/Figures/screening_rates.png",
+  plot = screening_plot, width = 12, height = 9, units = "in"
+)
 ```

### 2024-07-30 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
figure and table update

@@
 library(ggstats)
-library(rprev)
+library(forestmodel)
 home_dir <- switch(Sys.info()["sysname"],
@@
     ReviewYear, CFRD, Age_YrEnd, Gender, Race, Hispanicrace, bmi_perc,
-    bmi_perc_cat, MutClass, tube_feeding, liver_disease, A_FEV1_pct_predicted,
-    A_FVC_pct_predicted, NumPulmExacerbation, A_corticosteroids1
+    bmi_perc_cat, MutClass, tube_feeding, liver_disease,
+    A_FEV1_pct_predicted, A_FVC_pct_predicted, NumPulmExacerbation,
+    A_corticosteroids1
   ) %>%
@@
 
-```{r}
+```{r results='asis'}
 # How many OGTTs are non-CFRD people getting 2016-2018
@@
     `Age Group` = last(na.omit(age_group)),
-    `OGTTs` = sum(!is.na(A_twohour_bloodglucose) | !is.na(A_ogttfast_bloodglucose))
+    `OGTTs` = sum(!is.na(A_twohour_bloodglucose) |
+      !is.na(A_ogttfast_bloodglucose))
   )
@@
 )
-# Compare no OGTT to at least 1 OGTT
-ogtts %>%
-  select(-eDWID) %>%
-  tbl_summary(by = OGTTs, missing_text = "Missing") %>%
-  add_p()
+# Make table for each variable
+age <- ogtts %>%
+  select(`Age Group`, OGTTs) %>%
+  tbl_summary(by = `Age Group`) %>%
+  add_p() %>%
+  print()
+gender <- ogtts %>%
+  select(Gender, OGTTs) %>%
+  tbl_summary(by = Gender) %>%
+  add_p() %>%
+  print()
+muts <- ogtts %>%
+  select(MutClass, OGTTs) %>%
+  tbl_summary(by = MutClass) %>%
+  add_p() %>%
+  print()
+race <- ogtts %>%
+  select(Race, OGTTs) %>%
+  tbl_summary(by = Race) %>%
+  add_p() %>%
+  print()
+eth <- ogtts %>%
+  select(Hispanicrace, OGTTs) %>%
+  tbl_summary(by = Hispanicrace) %>%
+  add_p() %>%
+  print()
 ```
@@
 )
-y <- "Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2)"
+y <- "Surv(time = as.numeric(tstart), time2 = as.numeric(tstop), event = as.numeric(CFRD) == 2)"
 f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
-cox_res <- coxph(f, data = annual, id = eDWID)
-# Plot survival
-# ggsurvplot(survfit(cox_res),
-#   data = s, risk.table = T, cumevents = T, fun = "event",
-#   xlim = c(2003, 2018), break.x.by = 1
-# )
+cox_res <- coxph(f, data = annual, cluster = eDWID)
 # Forest plot and coefficients table
 cox_res %>%
-  tbl_regression() %>%
-  add_n(location = "level") %>%
-  add_nevent(location = "level")
-ggcoef_model(cox_res,
-  show_p_values = F, add_reference_rows = F,
-  signif_stars = F, colour = F,
-)
+  tbl_regression()
+ggcoef_model(cox_res, show_p_values = F, signif_stars = F, colour = F)
 # Diagnostics
@@
 time_varying <- rownames(test.ph)[test.ph$p < 0.05]
+```
+
+### Number of events by variable and year
+
+```{r results='asis'}
+invisible(lapply(xvars, function(v) {
+  cat("\n")
+  cat("#### By group")
+  cat("\n")
+  annual %>%
+    group_by_at(v) %>%
+    summarise(
+      n = length(unique(eDWID)),
+      `CFRD+` = sum(CFRD == "CFRD+"),
+      .groups = "drop"
+    ) %>%
+    gt() %>%
+    print()
+  cat("\n")
+  cat("#### By and year")
+  cat("\n")
+  annual %>%
+    group_by_at(c(v, "ReviewYear")) %>%
+    summarise(
+      n = length(unique(eDWID)),
+      `CFRD+` = sum(CFRD == "CFRD+"),
+      .groups = "drop"
+    ) %>%
+    gt() %>%
+    print()
+  cat("\n")
+}))
 ```
@@
     tbl_regression() %>%
-    add_n(location = "level") %>%
-    add_nevent(location = "level") %>%
     print()
@@
   # Plot marginal means
-  time_varying_forest_plot(emm, x = "emmean", var = v, var_label = labels[v]) %>%
-    print()
+  tvfp <- time_varying_forest_plot(emm,
+    x = "emmean", var = v,
+    var_label = labels[v]
+  ) +
+    theme(text = element_text(size = 20))
+  print(tvfp)
+  ggsave(
+    paste0(
+      "./Dissemination/Figures/", v,
+      "_time_varying_forest_plot.png"
+    ),
+    plot = tvfp, width = 12, height = 9, units = "in"
+  )
   cat("\n")

### 2024-07-30 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
vertex terms added

@@
 
-  new_cases_10_18 <- sum(inc$n[inc$CFRD == "CFRD+" & inc$age_group == "[10,19)"])
+  new_cases_10_18 <- sum(inc$n[inc$CFRD == "CFRD+" &
+    inc$age_group == "[10,19)"])
   at_risk_10_18 <- sum(inc$n[inc$age_group == "[10,19)"])
@@
 
-  new_cases_19_30 <- sum(inc$n[inc$CFRD == "CFRD+" & inc$age_group == "[19,31)"])
+  new_cases_19_30 <- sum(inc$n[inc$CFRD == "CFRD+" &
+    inc$age_group == "[19,31)"])
   at_risk_19_30 <- sum(inc$n[inc$age_group == "[19,31)"])
@@
 
-  new_cases_30 <- sum(inc$n[inc$CFRD == "CFRD+" & inc$age_group == "[31, Inf)"])
+  new_cases_30 <- sum(inc$n[inc$CFRD == "CFRD+" &
+    inc$age_group == "[31, Inf)"])
   at_risk_30 <- sum(inc$n[inc$age_group == "[31, Inf)"])
@@
     A_FEV1_pct_predicted, A_FVC_pct_predicted, NumPulmExacerbation,
-    A_corticosteroids1
+    A_corticosteroids1, A_IsOnEnzymes, A_Vx770, A_VX809comb
   ) %>%
@@
   pivot_longer(`Screened HbA1c`:`Screened Either`)
-screening_plot <- ggplot(screening, aes(x = ReviewYear, y = value, color = age_group)) +
+screening_plot <- ggplot(screening, aes(
+  x = ReviewYear, y = value,
+  color = age_group
+)) +
   geom_line() +
   xlab("Year") +
-  facet_wrap(~name) +
+  facet_wrap(~name, nrow = 3) +
   theme_bw() +
@@
   theme(text = element_text(size = 20), axis.title.y = element_blank())
+screening_plot
 ggsave("./Dissemination/Figures/screening_rates.png",
@@
   tbl_summary(by = `Age Group`) %>%
+  add_ci() %>%
   add_p() %>%
@@
   tbl_summary(by = Gender) %>%
+  add_ci() %>%
   add_p() %>%
@@
   tbl_summary(by = MutClass) %>%
+  add_ci() %>%
   add_p() %>%
@@
   tbl_summary(by = Race) %>%
+  add_ci() %>%
   add_p() %>%
@@
   tbl_summary(by = Hispanicrace) %>%
+  add_ci() %>%
   add_p() %>%
@@
 
-## Cox PH model
+## Cox PH models
 
@@
 
+### Vx770
+
+Years prior to 2011 were excluded from the following model
+
+```{r results='asis'}
+annual %>%
+  group_by(ReviewYear) %>%
+  count(A_Vx770) %>%
+  gt()
+```
+
+```{r results='asis'}
+f_vx770 <- update(f, . ~ . + A_Vx770)
+vx770_cox_res <- coxph(f_vx770,
+  data = annual[annual$ReviewYear >= 2011, ],
+  cluster = eDWID
+)
+vx770_cox_res %>%
+  tbl_regression()
+ggcoef_model(vx770_cox_res,
+  show_p_values = F, signif_stars = F, colour = F
+)
+```
+
+### Vx809comb
+
+Years prior to 2015 were excluded from the following model due to missing data.
+
+```{r results='asis'}
+annual %>%
+  group_by(ReviewYear) %>%
+  count(A_VX809comb) %>%
+  gt()
+```
+
+```{r}
+f_vx809comb <- update(f, . ~ . + A_VX809comb)
+vx809comb_cox_res <- coxph(f_vx809comb,
+  data = annual[annual$ReviewYear >= 2015, ],
+  cluster = eDWID
+)
+vx809comb_cox_res %>%
+  tbl_regression()
+ggcoef_model(vx809comb_cox_res,
+  show_p_values = F, signif_stars = F, colour = F
+)
+```
+
 ### Number of events by variable and year
@@
 ```{r results='asis'}
-invisible(lapply(xvars, function(v) {
+invisible(lapply(c(xvars, "A_Vx770", "A_VX809comb"), function(v) {
   cat("\n")

### 2024-07-31 Christine Chan/EnVision CF/catecholamines_analysis.qmd
started catecholamines analysis

--- 2024-04-07 Christine Chan/EnVision CF/amir_nacfc_2024.qmd
@@
 ---
-title: "Title"
+title: "Catecholamines and Hypoglycemia"
 author: "Tim Vigers"
@@
 ```{r data cleaning}
-# Import data cleaned for Katie's analysis
-df <- read.csv("./Christine Chan/EnVision CF/Data_Clean/nacfc_2024_analysis_data.csv",
+# Import cleaned data
+df <- read.csv("./Christine Chan/EnVision CF/Data_Clean/analysis_dataset.csv",
   na.strings = ""
 )
-glucose <- paste0("timepoint_", c(0, 10, 30, 60, 90, 120, 150, 180), "_min")
-insulin <- gsub("timepoint_", "insulin_", glucose)
-# Remove those missing all OGTT values
-df <- df[rowSums(is.na(df[, glucose])) < length(glucose), ]
+# For this analysis, we only want those with catecholamines 
+df = df %>% group_by(study_id,Date) %>%
+  mutate(catecholamines = any(!is.na(Epinephrine)|!is.na(Norepinephrine))) %>%
+  filter(catecholamines) %>% ungroup()
 ```
@@
 # Demographics at first visit
+
+BMI percentiles were calculated using the `childsds` R package, assuming that male = 1 and female = 2 for the `Sex` variable.
+Those with age > 20 were considered age = 20 for the purposes of percentile calculations.
 
@@
   select(
-    redcap_data_access_group, age_at_current_date:origin_race, cftr_mutation_1,
+    redcap_data_access_group, age_visit:ethnicity, cftr_mutation_1,
     cftr_mutation_2
@@
 
-## Participant demographics
+## Demographics
+
+IA0005, IA0006, IA0017, IA0042, and IA0079 had catecholamines measured at multiple timepoints but these are treated as independent in the table below.
 
@@
 df %>%
-  select(
-    age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2,
-    all_of(glucose), Hypo70, all_of(insulin), hypoglycemia_symptoms_yesno,
-    iAUC30gluc:iAUC180ins, homa_ir
+  filter(Timepoint %in% c(0, 120, 150, 180)) %>%
+  pivot_wider(
+    names_from = Timepoint, values_from = c(Epinephrine, Norepinephrine),
+    id_cols = c(study_id:cftr_mutation_2,Hypo60,Hypo70)
   ) %>%
+  select(Hypo70,age_visit:cftr_mutation_2,Epinephrine_0:Norepinephrine_180) %>%
   tbl_summary(by = Hypo70, missing_text = "Missing") %>%
@@
 ```{r mixed models 70}
-# Long OGTTs
-long_df <- df %>%
-  pivot_longer(timepoint_0_min:timepoint_180_min,
-    names_to = "Time", values_to = "Glucose",
-    names_pattern = "timepoint_(.*)_min"
-  )
 # Time as a factor
@@
 
-## Participant demographics
+## Demographics
 

### 2024-07-31 Christine Chan/EnVision CF/clean_envision_data.R

@@ -4,6 +4,7 @@
 # Libraries
 library(tidyverse)
 library(readxl)
+library(childsds)
 # Home directory
 home_dir <- switch(Sys.info()["sysname"],
   "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
@@ -42,12 +43,6 @@ insulin$Timepoint <- sub("Insulin ", "", insulin$Timepoint)
 insulin$Timepoint <- sub("Insulin-endo ", "", insulin$Timepoint)
 insulin$Timepoint <- sub(" min", "", insulin$Timepoint)
 insulin$Timepoint <- as.numeric(insulin$Timepoint)
-# Remove exact duplicates
-insulin <- insulin %>% distinct()
-# For duplicates with different insulin values, take the first (per Katie)
-insulin <- insulin %>%
-  group_by(study_id, Date, Timepoint) %>%
-  summarise(Insulin = first(na.omit(Insulin)))
 #-------------------------------------------------------------------------------
 # Catecholamines (Excel files needed some manual cleaning prior to this)
 #-------------------------------------------------------------------------------
@@ -87,7 +82,8 @@ glucose$Date <- ymd(glucose$Date)
 #-------------------------------------------------------------------------------
 demo <- redcap %>%
   select(
-    study_id, date_visit, age_visit, height, weight, sex, origin_race, ethnicity
+    study_id, date_visit, age_visit, height, weight, sex, origin_race,
+    ethnicity, redcap_data_access_group, cftr_mutation_1, cftr_mutation_2
   ) %>%
   filter(!is.na(date_visit)) %>%
   rename(Date = date_visit) %>%
@@ -159,6 +155,12 @@ ia005 <- data.frame(
   "Insulin" = as.character(c(6, NA, 40, 49, 31, 26, 7, 6))
 )
 insulin <- rbind(insulin, ia005)
+# Remove exact duplicates
+insulin <- insulin %>% distinct()
+# For duplicates with different insulin values, take the first (per Katie)
+insulin <- insulin %>%
+  group_by(study_id, Date, Timepoint) %>%
+  summarise(Insulin = first(na.omit(Insulin)))
 #-------------------------------------------------------------------------------
 # Combine everything
 #-------------------------------------------------------------------------------
[…]

### 2024-07-31 Christine Chan/EnVision CF/catecholamines_analysis.qmd
catecholamines tables done

@@
 # For this analysis, we only want those with catecholamines
-df = df %>% group_by(study_id,Date) %>%
-  mutate(catecholamines = any(!is.na(Epinephrine)|!is.na(Norepinephrine))) %>%
-  filter(catecholamines) %>% ungroup()
+df <- df %>%
+  group_by(study_id, Date) %>%
+  mutate(catecholamines = any(!is.na(Epinephrine) | !is.na(Norepinephrine))) %>%
+  filter(catecholamines) %>%
+  ungroup()
 ```
@@
 #| tbl-cap: BG < 70 at Any Point
+#| warning: false
 df %>%
@@
     names_from = Timepoint, values_from = c(Epinephrine, Norepinephrine),
-    id_cols = c(study_id:cftr_mutation_2,Hypo60,Hypo70)
+    id_cols = c(study_id:cftr_mutation_2, Hypo70)
   ) %>%
-  select(Hypo70,age_visit:cftr_mutation_2,Epinephrine_0:Norepinephrine_180) %>%
+  select(Hypo70, age_visit:cftr_mutation_2, Epinephrine_0:Norepinephrine_180) %>%
   tbl_summary(by = Hypo70, missing_text = "Missing") %>%
@@
 
-## LS mean glucose by hypoglycemia status
-
-```{r mixed models 70}
-# Time as a factor
-long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
-# Fit
-mod <- lmer(Glucose ~ Time * Hypo70 + (1 | study_id), data = long_df)
-means <- data.frame(emmeans(mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4))
-# Long insulins
-insulin_long_df <- df %>%
-  pivot_longer(all_of(insulin),
-    names_to = "Time", values_to = "Insulin",
-    names_pattern = "insulin_(.*)_min"
-  )
-# Time as a factor
-insulin_long_df$Time <- factor(insulin_long_df$Time,
-  levels = c(0, 10, 30, 60, 90, 120, 150, 180)
-)
-# Fit
-ins_mod <- lmer(Insulin ~ Time * Hypo70 + (1 | study_id), data = insulin_long_df)
-ins_means <- data.frame(emmeans(ins_mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4))
-# Combine
-means <- means %>% select(-SE, -df)
-means$Variable <- "Glucose"
-ins_means <- ins_means %>% select(-SE, -df)
-ins_means$Variable <- "Insulin"
-plot_df <- data.frame(rbind(means, ins_means))
-levels(plot_df$Hypo70) <- c("No Hypoglycemia", "Hypoglycemia")
-```
-
-```{r}
-#| label: fig-mixed-model70
-#| fig-cap: Model means by timepoint
-ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
-  geom_line() +
-  ylab("EM Mean") +
-  theme_bw() +
-  facet_wrap(~Hypo70)
-```
-
 # Hypoglycemia defined as < 60 mg/dL at any point
@@
 
+IA0005, IA0006, IA0017, IA0042, and IA0079 had catecholamines measured at multiple timepoints but these are treated as independent in the table below.
+
 ```{r}
@@
 #| tbl-cap: BG < 60 at Any Point
+#| warning: false
 df %>%
-  select(
-    age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2,
-    all_of(glucose), Hypo60, all_of(insulin), hypoglycemia_symptoms_yesno,
-    iAUC30gluc:iAUC180ins, homa_ir
+  filter(Timepoint %in% c(0, 120, 150, 180)) %>%
+  pivot_wider(
+    names_from = Timepoint, values_from = c(Epinephrine, Norepinephrine),
+    id_cols = c(study_id:cftr_mutation_2, Hypo60)
   ) %>%
+  select(Hypo60, age_visit:cftr_mutation_2, Epinephrine_0:Norepinephrine_180) %>%
   tbl_summary(by = Hypo60, missing_text = "Missing") %>%
@@
 
-## LS mean glucose by hypoglycemia status
+<!-- ## LS mean glucose by hypoglycemia status -->
 
-```{r mixed models 60}
-# Long OGTTs
-long_df <- df %>%
-  pivot_longer(timepoint_0_min:timepoint_180_min,
-    names_to = "Time", values_to = "Glucose",
-    names_pattern = "timepoint_(.*)_min"
-  )
-# Time as a factor
-long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
-# Fit
-mod <- lmer(Glucose ~ Time * Hypo60 + (1 | study_id), data = long_df)
-means <- data.frame(emmeans(mod, ~ Hypo60 | Time, pbkrtest.limit = 1e4))
-# Long insulins
-insulin_long_df <- df %>%
-  pivot_longer(all_of(insulin),
-    names_to = "Time", values_to = "Insulin",
-    names_pattern = "insulin_(.*)_min"
-  )
-# Time as a factor
-insulin_long_df$Time <- factor(insulin_long_df$Time,
-  levels = c(0, 10, 30, 60, 90, 120, 150, 180)
-)
-# Fit
-ins_mod <- lmer(Insulin ~ Time * Hypo60 + (1 | study_id), data = insulin_long_df)
-ins_means <- data.frame(emmeans(ins_mod, ~ Hypo60 | Time, pbkrtest.limit = 1e4))
-# Combine
-means <- means %>% select(-SE, -df)
-means$Variable <- "Glucose"
-ins_means <- ins_means %>% select(-SE, -df)
-ins_means$Variable <- "Insulin"
-plot_df <- data.frame(rbind(means, ins_means))
-levels(plot_df$Hypo60) <- c("No Hypoglycemia", "Hypoglycemia")
-```
+<!-- ```{r mixed models 70} -->
+<!-- # Time as a factor -->
+<!-- long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180)) -->
+<!-- # Fit -->
+<!-- mod <- lmer(Glucose ~ Time * Hypo70 + (1 | study_id), data = long_df) -->
+<!-- means <- data.frame(emmeans(mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4)) -->
+<!-- # Long insulins -->
+<!-- insulin_long_df <- df %>% -->
+<!--   pivot_longer(all_of(insulin), -->
+<!--     names_to = "Time", values_to = "Insulin", -->
+<!--     names_pattern = "insulin_(.*)_min" -->
+<!--   ) -->
+<!-- # Time as a factor -->
+<!-- insulin_long_df$Time <- factor(insulin_long_df$Time, -->
+<!--   levels = c(0, 10, 30, 60, 90, 120, 150, 180) -->
+<!-- ) -->
+<!-- # Fit -->
+<!-- ins_mod <- lmer(Insulin ~ Time * Hypo70 + (1 | study_id), data = insulin_long_df) -->
+<!-- ins_means <- data.frame(emmeans(ins_mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4)) -->
+<!-- # Combine -->
+<!-- means <- means %>% select(-SE, -df) -->
+<!-- means$Variable <- "Glucose" -->
+<!-- ins_means <- ins_means %>% select(-SE, -df) -->
+<!-- ins_means$Variable <- "Insulin" -->
+<!-- plot_df <- data.frame(rbind(means, ins_means)) -->
+<!-- levels(plot_df$Hypo70) <- c("No Hypoglycemia", "Hypoglycemia") -->
+<!-- ``` -->
 
-```{r}
-#| label: fig-mixed-model60
-#| fig-cap: Model means by timepoint
-ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
-  geom_line() +
-  ylab("EM Mean") +
-  theme_bw() +
-  facet_wrap(~Hypo60)
-```
+<!-- ```{r} -->
+<!-- #| label: fig-mixed-model70 -->
+<!-- #| fig-cap: Model means by timepoint -->
+<!-- ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) + -->
+<!--   geom_point() + -->
+<!--   geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) + -->
+<!--   geom_line() + -->
+<!--   ylab("EM Mean") + -->
+<!--   theme_bw() + -->
+<!--   facet_wrap(~Hypo70) -->
+<!-- ``` -->
 
-# Hypoglycemia defined as <= 50 mg/dL at any point
+<!-- # Hypoglycemia defined as < 60 mg/dL at any point -->
 
-## Participant demographics
+<!-- ## Demographics -->
 
-```{r}
-#| label: tbl-50
-#| tbl-cap: BG <= 50 at Any Point
-df %>%
-  select(
-    age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2,
-    all_of(glucose), Hypo50, all_of(insulin), hypoglycemia_symptoms_yesno,
-    iAUC30gluc:iAUC180ins, homa_ir
-  ) %>%
-  tbl_summary(by = Hypo50, missing_text = "Missing") %>%
-  add_overall() %>%
-  add_p() %>%
-  separate_p_footnotes()
-```
+<!-- ```{r} -->
+<!-- #| label: tbl-60 -->
+<!-- #| tbl-cap: BG < 60 at Any Point -->
+<!-- df %>% -->
+<!--   select( -->
+<!--     age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2, -->
+<!--     all_of(glucose), Hypo60, all_of(insulin), hypoglycemia_symptoms_yesno, -->
+<!--     iAUC30gluc:iAUC180ins, homa_ir -->
+<!--   ) %>% -->
+<!--   tbl_summary(by = Hypo60, missing_text = "Missing") %>% -->
+<!--   add_overall() %>% -->
+<!--   add_p() %>% -->
+<!--   separate_p_footnotes() -->
+<!-- ``` -->
 
-## LS mean glucose by hypoglycemia status
+<!-- ## LS mean glucose by hypoglycemia status -->
 
-```{r mixed models 50}
-# Long OGTTs
-long_df <- df %>%
-  pivot_longer(timepoint_0_min:timepoint_180_min,
-    names_to = "Time", values_to = "Glucose",
-    names_pattern = "timepoint_(.*)_min"
-  )
-# Time as a factor
-long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
-# Fit
-mod <- lmer(Glucose ~ Time * Hypo50 + (1 | study_id), data = long_df)
-means <- data.frame(emmeans(mod, ~ Hypo50 | Time, pbkrtest.limit = 1e4))
-# Long insulins
-insulin_long_df <- df %>%
-  pivot_longer(all_of(insulin),
-    names_to = "Time", values_to = "Insulin",
-    names_pattern = "insulin_(.*)_min"
-  )
-# Time as a factor
-insulin_long_df$Time <- factor(insulin_long_df$Time,
-  levels = c(0, 10, 30, 60, 90, 120, 150, 180)
-)
-# Fit
-ins_mod <- lmer(Insulin ~ Time * Hypo50 + (1 | study_id), data = insulin_long_df)
-ins_means <- data.frame(emmeans(ins_mod, ~ Hypo50 | Time, pbkrtest.limit = 1e4))
-# Combine
-means <- means %>% select(-SE, -df)
-means$Variable <- "Glucose"
-ins_means <- ins_means %>% select(-SE, -df)
-ins_means$Variable <- "Insulin"
-plot_df <- data.frame(rbind(means, ins_means))
-levels(plot_df$Hypo50) <- c("No Hypoglycemia", "Hypoglycemia")
-```
+<!-- ```{r mixed models 60} -->
+<!-- # Long OGTTs -->
+<!-- long_df <- df %>% -->
+<!--   pivot_longer(timepoint_0_min:timepoint_180_min, -->
+<!--     names_to = "Time", values_to = "Glucose", -->
+<!--     names_pattern = "timepoint_(.*)_min" -->
+<!--   ) -->
+<!-- # Time as a factor -->
+<!-- long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180)) -->
+<!-- # Fit -->
+<!-- mod <- lmer(Glucose ~ Time * Hypo60 + (1 | study_id), data = long_df) -->
+<!-- means <- data.frame(emmeans(mod, ~ Hypo60 | Time, pbkrtest.limit = 1e4)) -->
+<!-- # Long insulins -->
+<!-- insulin_long_df <- df %>% -->
+<!--   pivot_longer(all_of(insulin), -->
+<!--     names_to = "Time", values_to = "Insulin", -->
+<!--     names_pattern = "insulin_(.*)_min" -->
+<!--   ) -->
+<!-- # Time as a factor -->
+<!-- insulin_long_df$Time <- factor(insulin_long_df$Time, -->
+<!--   levels = c(0, 10, 30, 60, 90, 120, 150, 180) -->
+<!-- ) -->
+<!-- # Fit -->
+<!-- ins_mod <- lmer(Insulin ~ Time * Hypo60 + (1 | study_id), data = insulin_long_df) -->
+<!-- ins_means <- data.frame(emmeans(ins_mod, ~ Hypo60 | Time, pbkrtest.limit = 1e4)) -->
+<!-- # Combine -->
+<!-- means <- means %>% select(-SE, -df) -->
+<!-- means$Variable <- "Glucose" -->
+<!-- ins_means <- ins_means %>% select(-SE, -df) -->
+<!-- ins_means$Variable <- "Insulin" -->
+<!-- plot_df <- data.frame(rbind(means, ins_means)) -->
+<!-- levels(plot_df$Hypo60) <- c("No Hypoglycemia", "Hypoglycemia") -->
+<!-- ``` -->
 
-```{r}
-#| label: fig-mixed-model50
-#| fig-cap: Model means by timepoint
-ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
-  geom_line() +
-  ylab("EM Mean") +
-  theme_bw() +
-  facet_wrap(~Hypo50)
-```
+<!-- ```{r} -->
+<!-- #| label: fig-mixed-model60 -->
+<!-- #| fig-cap: Model means by timepoint -->
+<!-- ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) + -->
+<!--   geom_point() + -->
+<!--   geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) + -->
+<!--   geom_line() + -->
+<!--   ylab("EM Mean") + -->
+<!--   theme_bw() + -->
+<!--   facet_wrap(~Hypo60) -->
+<!-- ``` -->
+
+<!-- # Hypoglycemia defined as <= 50 mg/dL at any point -->
+
+<!-- ## Participant demographics -->
+
+<!-- ```{r} -->
+<!-- #| label: tbl-50 -->
+<!-- #| tbl-cap: BG <= 50 at Any Point -->
+<!-- df %>% -->
+<!--   select( -->
+<!--     age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2, -->
+<!--     all_of(glucose), Hypo50, all_of(insulin), hypoglycemia_symptoms_yesno, -->
+<!--     iAUC30gluc:iAUC180ins, homa_ir -->
+<!--   ) %>% -->
+<!--   tbl_summary(by = Hypo50, missing_text = "Missing") %>% -->
+<!--   add_overall() %>% -->
+<!--   add_p() %>% -->
+<!--   separate_p_footnotes() -->
+<!-- ``` -->
+
+<!-- ## LS mean glucose by hypoglycemia status -->
+
+<!-- ```{r mixed models 50} -->
+<!-- # Long OGTTs -->
+<!-- long_df <- df %>% -->
+<!--   pivot_longer(timepoint_0_min:timepoint_180_min, -->
+<!--     names_to = "Time", values_to = "Glucose", -->
+<!--     names_pattern = "timepoint_(.*)_min" -->
+<!--   ) -->
+<!-- # Time as a factor -->
+<!-- long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180)) -->
+<!-- # Fit -->
+<!-- mod <- lmer(Glucose ~ Time * Hypo50 + (1 | study_id), data = long_df) -->
+<!-- means <- data.frame(emmeans(mod, ~ Hypo50 | Time, pbkrtest.limit = 1e4)) -->
+<!-- # Long insulins -->
+<!-- insulin_long_df <- df %>% -->
+<!--   pivot_longer(all_of(insulin), -->
+<!--     names_to = "Time", values_to = "Insulin", -->
+<!--     names_pattern = "insulin_(.*)_min" -->
+<!--   ) -->
+<!-- # Time as a factor -->
+<!-- insulin_long_df$Time <- factor(insulin_long_df$Time, -->
+<!--   levels = c(0, 10, 30, 60, 90, 120, 150, 180) -->
+<!-- ) -->
+<!-- # Fit -->
+<!-- ins_mod <- lmer(Insulin ~ Time * Hypo50 + (1 | study_id), data = insulin_long_df) -->
+<!-- ins_means <- data.frame(emmeans(ins_mod, ~ Hypo50 | Time, pbkrtest.limit = 1e4)) -->
+<!-- # Combine -->
+<!-- means <- means %>% select(-SE, -df) -->
+<!-- means$Variable <- "Glucose" -->
+<!-- ins_means <- ins_means %>% select(-SE, -df) -->
+<!-- ins_means$Variable <- "Insulin" -->
+<!-- plot_df <- data.frame(rbind(means, ins_means)) -->
+<!-- levels(plot_df$Hypo50) <- c("No Hypoglycemia", "Hypoglycemia") -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- #| label: fig-mixed-model50 -->
+<!-- #| fig-cap: Model means by timepoint -->
+<!-- ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) + -->
+<!--   geom_point() + -->
+<!--   geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) + -->
+<!--   geom_line() + -->
+<!--   ylab("EM Mean") + -->
+<!--   theme_bw() + -->
+<!--   facet_wrap(~Hypo50) -->
+<!-- ``` -->

### 2024-07-31 Christine Chan/EnVision CF/clean_envision_data.R

@@ -175,7 +175,7 @@ final_df <- full_join(final_df, demo)
 final_df$bmi <- final_df$weight / ((final_df$height^2) / 10000)
 final_df$bmi_perc <- sds(
   value = final_df$bmi,
-  age = ifelse(final_df$age_visit<20,final_df$age_visit,20),
+  age = ifelse(final_df$age_visit < 20, final_df$age_visit, 20),
   sex = final_df$sex, male = 1, female = 2,
   ref = cdc.ref, item = "bmi", type = "perc"
 ) * 100
@@ -183,6 +183,14 @@ final_df$bmi_perc <- sds(
 final_df <- final_df %>%
   group_by(study_id, Date) %>%
   mutate(Hypo70 = any(Glucose < 70), Hypo60 = any(Glucose < 60))
+final_df$Hypo70 <- factor(final_df$Hypo70,
+  levels = c(T, F),
+  labels = c("Hypoglycemia < 70", "No Hypoglycemia < 70")
+)
+final_df$Hypo60 <- factor(final_df$Hypo60,
+  levels = c(T, F),
+  labels = c("Hypoglycemia < 60", "No Hypoglycemia < 60")
+)
 # Format
 final_df <- final_df %>%
   select(

### 2024-08-02 Christine Chan/EnVision CF/catecholamines_analysis.qmd
catecholamines prelim analysis done

--- 2024-07-31 Christine Chan/EnVision CF/catecholamines_analysis.qmd
@@
 
+# Data cleaning
+
+- BMI percentiles were calculated using the `childsds` R package, assuming that male = 1 and female = 2 for the `Sex` variable.
+Those with age > 20 were considered age = 20 for the purposes of percentile calculations.
+
+- Epinephrine and norepinephrine were log transformed for modeling purposes.
+
 ```{r data cleaning}
@@
 # For this analysis, we only want those with catecholamines
-df = df %>% group_by(study_id,Date) %>%
-  mutate(catecholamines = any(!is.na(Epinephrine)|!is.na(Norepinephrine))) %>%
-  filter(catecholamines) %>% ungroup()
+df <- df %>%
+  group_by(study_id, Date) %>%
+  mutate(catecholamines = any(!is.na(Epinephrine) | !is.na(Norepinephrine))) %>%
+  filter(catecholamines) %>%
+  ungroup()
+# # Time as a factor
+df$Timepoint <- factor(df$Timepoint, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
 ```
@@
 # Demographics at first visit
-
-BMI percentiles were calculated using the `childsds` R package, assuming that male = 1 and female = 2 for the `Sex` variable.
-Those with age > 20 were considered age = 20 for the purposes of percentile calculations.
 
@@
 
-## LS mean glucose by hypoglycemia status
+## Epinephrine
 
-```{r mixed models 70}
-# Time as a factor
-long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
+Epinephrine was modeled using a linear mixed effect model with random intercept for participant and time treated as a categorical variable.
+Plot error bars indicate 95% confidence intervals around the least square mean.
+
+```{r}
+#| warning: false
 # Fit
-mod <- lmer(Glucose ~ Time * Hypo70 + (1 | study_id), data = long_df)
-means <- data.frame(emmeans(mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4))
-# Long insulins
-insulin_long_df <- df %>%
-  pivot_longer(all_of(insulin),
-    names_to = "Time", values_to = "Insulin",
-    names_pattern = "insulin_(.*)_min"
-  )
-# Time as a factor
-insulin_long_df$Time <- factor(insulin_long_df$Time,
-  levels = c(0, 10, 30, 60, 90, 120, 150, 180)
-)
-# Fit
-ins_mod <- lmer(Insulin ~ Time * Hypo70 + (1 | study_id), data = insulin_long_df)
-ins_means <- data.frame(emmeans(ins_mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4))
-# Combine
-means <- means %>% select(-SE, -df)
-means$Variable <- "Glucose"
-ins_means <- ins_means %>% select(-SE, -df)
-ins_means$Variable <- "Insulin"
-plot_df <- data.frame(rbind(means, ins_means))
-levels(plot_df$Hypo70) <- c("No Hypoglycemia", "Hypoglycemia")
+mod <- lmer(log(Epinephrine) ~ Timepoint * Hypo70 + (1 | study_id), data = df)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"), type = "response"))
+means %>% gt(caption = "LS Means")
+# Plot
+emmip(mod, Hypo70 ~ Timepoint, CIs = T, type = "response") +
+  theme_bw() +
+  theme(legend.title = element_blank())
 ```
 
+## Norepinephrine
+
+Norepinephrine was modeled using a linear mixed effect model with random intercept for participant and time treated as a categorical variable.
+Plot error bars indicate 95% confidence intervals around the least square mean.
+
 ```{r}
-#| label: fig-mixed-model70
-#| fig-cap: Model means by timepoint
-ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
-  geom_line() +
-  ylab("EM Mean") +
+#| warning: false
+# Fit
+mod <- lmer(log(Norepinephrine) ~ Timepoint * Hypo70 + (1 | study_id), data = df)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"), type = "response"))
+means %>% gt(caption = "LS Means")
+# Plot
+emmip(mod, Hypo70 ~ Timepoint, CIs = T, type = "response") +
   theme_bw() +
-  facet_wrap(~Hypo70)
+  theme(legend.title = element_blank())
 ```
@@
 
-## LS mean glucose by hypoglycemia status
+## Epinephrine
 
-```{r mixed models 60}
-# Long OGTTs
-long_df <- df %>%
-  pivot_longer(timepoint_0_min:timepoint_180_min,
-    names_to = "Time", values_to = "Glucose",
-    names_pattern = "timepoint_(.*)_min"
-  )
-# Time as a factor
-long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
+Epinephrine was modeled using a linear mixed effect model with random intercept for participant and time treated as a categorical variable.
+Plot error bars indicate 95% confidence intervals around the least square mean.
+
+```{r}
+#| warning: false
 # Fit
-mod <- lmer(Glucose ~ Time * Hypo60 + (1 | study_id), data = long_df)
-means <- data.frame(emmeans(mod, ~ Hypo60 | Time, pbkrtest.limit = 1e4))
-# Long insulins
-insulin_long_df <- df %>%
-  pivot_longer(all_of(insulin),
-    names_to = "Time", values_to = "Insulin",
-    names_pattern = "insulin_(.*)_min"
-  )
-# Time as a factor
-insulin_long_df$Time <- factor(insulin_long_df$Time,
-  levels = c(0, 10, 30, 60, 90, 120, 150, 180)
-)
-# Fit
-ins_mod <- lmer(Insulin ~ Time * Hypo60 + (1 | study_id), data = insulin_long_df)
-ins_means <- data.frame(emmeans(ins_mod, ~ Hypo60 | Time, pbkrtest.limit = 1e4))
-# Combine
-means <- means %>% select(-SE, -df)
-means$Variable <- "Glucose"
-ins_means <- ins_means %>% select(-SE, -df)
-ins_means$Variable <- "Insulin"
-plot_df <- data.frame(rbind(means, ins_means))
-levels(plot_df$Hypo60) <- c("No Hypoglycemia", "Hypoglycemia")
+mod <- lmer(log(Epinephrine) ~ Timepoint * Hypo60 + (1 | study_id), data = df)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"), type = "response"))
+means %>% gt(caption = "LS Means")
+# Plot
+emmip(mod, Hypo60 ~ Timepoint, CIs = T, type = "response") +
+  theme_bw() +
+  theme(legend.title = element_blank())
 ```
 
-```{r}
-#| label: fig-mixed-model60
-#| fig-cap: Model means by timepoint
-ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
-  geom_line() +
-  ylab("EM Mean") +
-  theme_bw() +
-  facet_wrap(~Hypo60)
-```
+## Norepinephrine
 
-# Hypoglycemia defined as <= 50 mg/dL at any point
-
-## Participant demographics
+Norepinephrine was modeled using a linear mixed effect model with random intercept for participant and time treated as a categorical variable.
+Plot error bars indicate 95% confidence intervals around the least square mean.
 
 ```{r}
-#| label: tbl-50
-#| tbl-cap: BG <= 50 at Any Point
-df %>%
-  select(
-    age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2,
-    all_of(glucose), Hypo50, all_of(insulin), hypoglycemia_symptoms_yesno,
-    iAUC30gluc:iAUC180ins, homa_ir
-  ) %>%
-  tbl_summary(by = Hypo50, missing_text = "Missing") %>%
-  add_overall() %>%
-  add_p() %>%
-  separate_p_footnotes()
+#| warning: false
+# Fit
+mod <- lmer(log(Norepinephrine) ~ Timepoint * Hypo60 + (1 | study_id), data = df)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"), type = "response"))
+means %>% gt(caption = "LS Means")
+# Plot
+emmip(mod, Hypo60 ~ Timepoint, CIs = T, type = "response") +
+  theme_bw() +
+  theme(legend.title = element_blank())
 ```
-
-## LS mean glucose by hypoglycemia status
-
-```{r mixed models 50}
-# Long OGTTs
-long_df <- df %>%
-  pivot_longer(timepoint_0_min:timepoint_180_min,
-    names_to = "Time", values_to = "Glucose",
-    names_pattern = "timepoint_(.*)_min"
-  )
-# Time as a factor
-long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
-# Fit
-mod <- lmer(Glucose ~ Time * Hypo50 + (1 | study_id), data = long_df)
-means <- data.frame(emmeans(mod, ~ Hypo50 | Time, pbkrtest.limit = 1e4))
-# Long insulins
-insulin_long_df <- df %>%
-  pivot_longer(all_of(insulin),
-    names_to = "Time", values_to = "Insulin",
-    names_pattern = "insulin_(.*)_min"
-  )
-# Time as a factor
-insulin_long_df$Time <- factor(insulin_long_df$Time,
-  levels = c(0, 10, 30, 60, 90, 120, 150, 180)
-)
-# Fit
-ins_mod <- lmer(Insulin ~ Time * Hypo50 + (1 | study_id), data = insulin_long_df)
-ins_means <- data.frame(emmeans(ins_mod, ~ Hypo50 | Time, pbkrtest.limit = 1e4))
-# Combine
-means <- means %>% select(-SE, -df)
-means$Variable <- "Glucose"
-ins_means <- ins_means %>% select(-SE, -df)
-ins_means$Variable <- "Insulin"
-plot_df <- data.frame(rbind(means, ins_means))
-levels(plot_df$Hypo50) <- c("No Hypoglycemia", "Hypoglycemia")
-```
-
-```{r}
-#| label: fig-mixed-model50
-#| fig-cap: Model means by timepoint
-ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
-  geom_line() +
-  ylab("EM Mean") +
-  theme_bw() +
-  facet_wrap(~Hypo50)
-```

### 2024-08-02 Christine Chan/EnVision CF/clean_envision_data.R

@@ -182,7 +182,10 @@ final_df$bmi_perc <- sds(
 # Check for hypoglycemia
 final_df <- final_df %>%
   group_by(study_id, Date) %>%
-  mutate(Hypo70 = any(Glucose < 70), Hypo60 = any(Glucose < 60))
+  mutate(
+    Hypo70 = any(na.omit(Glucose) < 70),
+    Hypo60 = any(na.omit(Glucose) < 60)
+  )
 final_df$Hypo70 <- factor(final_df$Hypo70,
   levels = c(T, F),
   labels = c("Hypoglycemia < 70", "No Hypoglycemia < 70")

### 2024-08-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
knit report on the server

@@
 library(ggflowchart)
+library(lmerTest)
+library(splines)
 # Paths
@@
 
-```{r}
+```{r results='asis'}
+invisible(lapply(continuous_outcomes, function(o) {
+  # Formula
+  f <-
+    as.formula(paste0(
+      o,
+      "~ bs(Days, knots = 0, degree = 1) * Race + (1 | eDWID)"
+    ))
+  m <- lmer(f, data = encounter)
+  # Print table
+  m %>%
+    tbl_regression() %>%
+    print()
+  # Plot
+  predict_response(m, terms = c("Days[-1095,0,730]", "Race")) %>% 
+    plot()
+}))
 ```

### 2024-08-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/outcomes_mixed_models.R

@@ -1,29 +1,0 @@
-library(lme4)
-library(splines)
-library(ggeffects)
-# Paths
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/tim/CF"
-)
-setwd(home_dir)
-# Load dataset
-load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData")
-# Fit models
-results <- lapply(continuous_outcomes, function(o) {
-  # Formula
-  f <-
-    as.formula(paste0(
-      o,
-      "~ bs(Days, knots = 0, degree = 1) * Race + (1 | eDWID)"
-    ))
-  m <- lmer(f, data = encounter)
-  # Predicted values
-  pred <- predict_response(m, terms = c("Days[-1095,0,730]", "Race"))
-  return(list(model = m, predictions = pred))
-})
-# Save
-save(results,
-  file = "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_lmm_results.RData"
-)

### 2024-08-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
don't run LOESS for now

@@
 ```{r functions}
+#| eval: false
 # Plots
@@
 ```{r}
+#| eval: false
 invisible(lapply(continuous_outcomes, function(o) {

### 2024-08-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
better paths

@@
   "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+  "Linux" = "/home/timvigers/CF"
 )

### 2024-08-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
ugh, okay better paths again

@@
   "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/timvigers/CF"
+  "Linux" = "/home/tim/CF"
 )

### 2024-08-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
needed ggeffects

@@
 library(splines)
+library(ggeffects)
 # Paths

### 2024-08-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
added LOESS plots back

@@
 ```{r functions}
-#| eval: false
 # Plots
@@
 ```{r}
-#| eval: false
 invisible(lapply(continuous_outcomes, function(o) {
@@
 ```
-
-# Questions
-
-1.

### 2024-08-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
removed LOESS again

@@
 
-# LOESS-smoothed outcome plots
-
-```{r functions}
-# Plots
-loess_plot <- function(data = encounter, outcome, facet = F) {
-  p <- ggplot(data, aes(x = Days, y = !!sym(outcome), color = Race)) +
-    geom_point(
-      shape = ".", aes(x = Days, y = !!sym(outcome)), inherit.aes = F,
-      alpha = 0.1
-    ) +
-    geom_smooth(method = "loess", se = F) +
-    theme_classic()
-  if (facet) {
-    p <- p + facet_wrap(~Race)
-  }
-  return(p)
-}
-```
-
-```{r}
-invisible(lapply(continuous_outcomes, function(o) {
-  cat("\n")
-  cat("\n")
-  cat("##", o)
-  cat("\n")
-  cat("\n")
-  print(loess_plot(outcome = o))
-}))
-```
-
 # Segmented models
@@
   # Print table
+  cat("\n")
+  cat("##",o)
+  cat("\n")
+  cat("\n")
+  cat("### Model results")
+  cat("\n")
   m %>%
@@
     print()
+  cat("\n")
   # Plot
+  cat("\n")
+  cat("### Plot")
+  cat("\n")
   predict_response(m, terms = c("Days[-1095,0,730]", "Race")) %>% 
-    plot()
+    plot() %>%
+    print()
+  cat("\n")
 }))

### 2024-08-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
so many cat("\n")

@@
   cat("\n")
+  cat("\n")
   cat("##",o)
@@
   cat("\n")
+  cat("\n")
   m %>%
@@
     print()
+  cat("\n")
   cat("\n")
@@
   cat("\n")
+  cat("\n")
   predict_response(m, terms = c("Days[-1095,0,730]", "Race")) %>% 
@@
   cat("\n")
+  cat("\n")
 }))

### 2024-08-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
added predictions table

--- 2024-08-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
@@
 ```{r results='asis'}
+#| message: false
 invisible(lapply(continuous_outcomes, function(o) {
@@
   cat("\n")
-  cat("##",o)
+  cat("\n")
+  cat("##", o)
   cat("\n")
@@
   cat("### Model results")
+  cat("\n")
   cat("\n")
@@
   cat("\n")
+  cat("\n")
   # Plot
+  cat("\n")
+  cat("\n")
+  cat("## Model predictions")
+  cat("\n")
+  cat("\n")
+  pred <- predict_response(m, terms = c("Days[-1095,0,730]", "Race"))
+  pred %>% gt()
+  cat("\n")
   cat("\n")
@@
   cat("\n")
-  predict_response(m, terms = c("Days[-1095,0,730]", "Race")) %>% 
+  cat("\n")
+  pred %>%
     plot() %>%

### 2024-08-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
forgot print() again

@@
   pred <- predict_response(m, terms = c("Days[-1095,0,730]", "Race"))
-  pred %>% gt()
+  pred %>% gt() %>% print()
   cat("\n")

### 2024-08-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
make ETI outcomes dataset on Lambda

--- 2024-08-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
@@
   "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/tim/CF"
+  "Linux" = "/home/timvigers//OneDrive/Vigers/CF"
 )
@@
 
-Continuous variables were averaged across all years for each person, and categorical variables are reported as the most common value.
+Continuous variables are reported at the most recent visit prior to ETI start for each person, and categorical variables are reported as the most common value.
 
@@
 #| message: false
-invisible(lapply(continuous_outcomes, function(o) {
+invisible(lapply(continuous_outcomes[1:3], function(o) {
   # Formula
@@
   pred <- predict_response(m, terms = c("Days[-1095,0,730]", "Race"))
-  pred %>% gt()
+  pred %>% print()
   cat("\n")

### 2024-08-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R

@@ -5,7 +5,7 @@ library(parallel)
 home_dir <- switch(Sys.info()["sysname"],
   "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
   "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+  "Linux" = "/home/tim/CF"
 )
 setwd(home_dir)
 # Import
@@ -178,8 +178,9 @@ t1_participant <- encounter %>%
     `Number of Encounters` = n(),
     `Number of Encounters Pre-ETI` = sum(Days <= 0),
     `Number of Encounters Post-ETI` = sum(Days > 0),
-    Age = mean(encounterage, na.rm = T),
-    across(all_of(continuous_outcomes), \(x) mean(x, na.rm = TRUE)),
+    Age = max(encounterage[Days <= 0], na.rm = T),
+    across(all_of(continuous_outcomes), \(x) max(x[Days <= 0], na.rm = TRUE)),
+    Sex = names(sort(table(Sex), decreasing = TRUE))[1],
     Race = names(sort(table(Race), decreasing = TRUE))[1],
     Insurance = names(sort(table(Insurance), decreasing = TRUE))[1],
     .groups = "drop"

### 2024-08-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
adjusted models

@@
       o,
-      "~ bs(Days, knots = 0, degree = 1) * Race + (1 | eDWID)"
+      "~ bs(Days, knots = 0, degree = 1) * Race + age_group + Insurance + (1 | eDWID)"
     ))

### 2024-08-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
ugh, file paths

@@
   "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/timvigers//OneDrive/Vigers/CF"
+  "Linux" = "/home/tim/CF"
 )

### 2024-08-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
weord table formatting

@@
   pred <- predict_response(m, terms = c("Days[-1095,0,730]", "Race"))
-  pred %>% print()
+  pred %>% gt() %>% print()
   cat("\n")

### 2024-08-07 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
new home directory paths

@@
     page-layout: full
-bibliography: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Miscellaneous/american-medical-association.csl
+bibliography: /home/tim/Documents/Miscellaneous/zotero.bib
+csl: /home/tim/GitHub/styles/american-medical-association.csl
 editor: source
@@
 library(pROC)
-switch(Sys.info()[["sysname"]],
-  Windows = {
-    home_dir <- "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry"
-  },
-  Darwin = {
-    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry"
-  }
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
+  "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/CF Registry"
+)
+github_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/GitHub",
+  "Windows" = "C:/Users/Tim/GitHub",
+  "Linux" = "/home/tim/GitHub"
 )
@@
     `OGTT 2-hour` = cut(twohour_bloodglucose,
-      breaks = c(-Inf, 140, 200, Inf), labels = c("<140", "[140,200)", ">=200")
+      breaks = c(-Inf, 140, 200, Inf), 
+      labels = c("<140", "[140,200)", ">=200")
     ),
@@
 metrics <- c(
-  "thresh", "specificity", "sensitivity", "accuracy", "tn", "tp", "fn", "fp",
-  "npv", "ppv"
+  "thresh", "specificity", "sensitivity", "accuracy", "tn", "tp", "fn", 
+  "fp","npv", "ppv"
 )
@@
 
+### By race
+
 #### NGT vs. AGT

### 2024-08-07 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
new home directory paths

@@
   "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
+  "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/CF Registry"
 )
@@
   "Windows" = "C:/Users/Tim/GitHub",
-  "Linux" = "/home/timvigers/GitHub"
+  "Linux" = "/home/tim/GitHub"
 )

### 2024-08-07 Quarto Template.qmd
new home directory paths

--- 2024-04-03 Quarto Template.qmd
@@
       dark: darkly
-bibliography: ~/Documents/Miscellaneous/zotero.bib
-csl: ~/Documents/Miscellaneous/american-medical-association.csl
+bibliography: /home/tim/Documents/Miscellaneous/zotero.bib
+csl: /home/tim/GitHub/styles/american-medical-association.csl
 editor: source
@@
 library(tidyverse)
+library(redcapAPI)
 home_dir <- switch(Sys.info()["sysname"],
@@
   "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+  "Linux" = "/home/tim/OneDrive/Vigers/CF"
+)
+github_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/GitHub",
+  "Windows" = "C:/Users/Tim/GitHub",
+  "Linux" = "/home/tim/GitHub"
 )
@@
 ```{r data cleaning}
+#| message: false
+# Import from REDCap
+unlockREDCap(c(rcon = ""),
+  keyring = "API_KEYs",
+  envir = 1,
+  url = "https://redcap.ucdenver.edu/api/"
+)
+exportBulkRecords(list(db = rcon), envir = 1)
 ```

### 2024-08-07 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
ridiculously inefficient ROC code but it's working

--- 2024-03-13 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
@@
     `OGTT 2-hour` = cut(twohour_bloodglucose,
-      breaks = c(-Inf, 140, 200, Inf), labels = c("<140", "[140,200)", ">=200")
+      breaks = c(-Inf, 140, 200, Inf),
+      labels = c("<140", "[140,200)", ">=200")
     ),
@@
 metrics <- c(
-  "thresh", "specificity", "sensitivity", "accuracy", "tn", "tp", "fn", "fp",
-  "npv", "ppv"
+  "thresh", "specificity", "sensitivity", "accuracy", "tn", "tp", "fn",
+  "fp", "npv", "ppv"
 )
@@
 
-## Three groups, two cutoffs
+For each ROC curve, we performed 2000 bootstrap replicates.
+Here the median best threshold (based on Youden index), sensitivity, and specificity are presented along with the bootstrapped 95% confidence intervals.
+Column names are printed in terms of percentiles, so `threshold.2.5.` indicates the lower limit of the confidence interval, `threshold.50.` indicates the median best threshold, and `threshold.97.5.` indicates the upper limit of the confidence interval.
+
+## NGT vs. AGT and AGT vs. CFRD
 
@@
 
-### Peds and adults
+### Full cohort
 
@@
 plot(r, print.auc = T, print.thres = T)
-gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 plot(r, print.auc = T, print.thres = T)
-gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
 ```
 
-### Peds
+### Peds vs. Adults
+
+#### Peds
 
@@
 
-#### NGT vs. AGT
+##### NGT vs. AGT
 
@@
 plot(r, print.auc = T, print.thres = T)
-gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
 ```
 
-#### AGT vs. CFRD
+##### AGT vs. CFRD
 
@@
 plot(r, print.auc = T, print.thres = T)
-gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
 ```
 
-### Adults
+#### Adults
 
@@
 plot(r, print.auc = T, print.thres = T)
-gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
 ```
 
-## Two groups, one cutoffs
+### By race
+
+#### White
+
+```{r}
+white_ngt_agt <- ngt_agt %>% filter(Race == "White")
+white_agt_cfrd <- agt_cfrd %>% filter(Race == "White")
+```
+
+##### NGT vs. AGT
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = white_ngt_agt)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+```
+
+##### AGT vs. CFRD
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = white_agt_cfrd)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+```
+
+#### Black or African American
+
+```{r}
+black_ngt_agt <- ngt_agt %>% filter(Race == "Black or African American")
+black_agt_cfrd <- agt_cfrd %>% filter(Race == "Black or African American")
+```
+
+##### NGT vs. AGT
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = black_ngt_agt)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+```
+
+##### AGT vs. CFRD
+
+Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = black_agt_cfrd)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(coords(r, "best"))) %>%
+  fmt_number(n_sigfig = 3)
+gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+#### More than one race
+
+```{r}
+more_ngt_agt <- ngt_agt %>% filter(Race == "More than one race")
+more_agt_cfrd <- agt_cfrd %>% filter(Race == "More than one race")
+```
+
+##### NGT vs. AGT
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = more_ngt_agt)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+```
+
+##### AGT vs. CFRD
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = more_agt_cfrd)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+```
+
+#### Other race
+
+```{r}
+other_ngt_agt <- ngt_agt %>% filter(Race == "Other race")
+other_agt_cfrd <- agt_cfrd %>% filter(Race == "Other race")
+```
+
+##### NGT vs. AGT
+
+Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = other_ngt_agt)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(coords(r, "best"))) %>%
+  fmt_number(n_sigfig = 3)
+gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+##### AGT vs. CFRD
+
+Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = other_agt_cfrd)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(coords(r, "best"))) %>%
+  fmt_number(n_sigfig = 3)
+gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+## Two hour OGTT >= 140 or >= 200
 
@@
 
-#### $\geq$ 140
+##### $\geq$ 140
 
@@
 plot(r, print.auc = T, print.thres = T)
-gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
 ```
 
-#### $\geq$ 200
+##### $\geq$ 200
 
@@
 plot(r, print.auc = T, print.thres = T)
-gt(coords(r, ret = metrics)) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
 ```
+
+### By race
+
+#### White
+
+```{r}
+white_roc <- roc_df %>% filter(Race == "White")
+```
+
+##### $\geq$ 140
+
+```{r}
+#| message: false
+r <- roc(agt ~ hgba1c, data = white_roc)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+```
+
+##### $\geq$ 200
+
+```{r}
+#| message: false
+r <- roc(cfrd ~ hgba1c, data = white_roc)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+```
+
+#### Black or African American
+
+```{r}
+black_roc <- roc_df %>% filter(Race == "Black or African American")
+```
+
+##### $\geq$ 140
+
+```{r}
+#| message: false
+r <- roc(agt ~ hgba1c, data = black_roc)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+```
+
+##### $\geq$ 200
+
+```{r}
+#| message: false
+r <- roc(cfrd ~ hgba1c, data = black_roc)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+```
+
+#### More than one race
+
+```{r}
+more_roc <- roc_df %>% filter(Race == "More than one race")
+```
+
+##### $\geq$ 140
+
+```{r}
+#| message: false
+r <- roc(agt ~ hgba1c, data = more_roc)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+```
+
+##### $\geq$ 200
+
+```{r}
+#| message: false
+r <- roc(cfrd ~ hgba1c, data = more_roc)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+```
+
+#### Other race
+
+```{r}
+other_roc <- roc_df %>% filter(Race == "Other race")
+```
+
+##### $\geq$ 140
+
+Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.
+
+```{r}
+#| message: false
+r <- roc(agt ~ hgba1c, data = other_roc)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(coords(r, "best"))) %>%
+  fmt_number(n_sigfig = 3)
+gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+##### $\geq$ 200
+
+Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.
+
+```{r}
+#| message: false
+r <- roc(cfrd ~ hgba1c, data = other_roc)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(coords(r, "best"))) %>%
+  fmt_number(n_sigfig = 3)
+gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
+  fmt_number(n_sigfig = 3)
+```

### 2024-08-07 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
file paths for lambda

@@
   "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
-  "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/CF Registry"
+  "Linux" = "/home/tim/CF/Christine Chan/CF Registry"
 )

### 2024-08-08 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
set seed

@@
   "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
-  "Linux" = "/home/tim/CF/Christine Chan/CF Registry"
+  "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/CF Registry"
 )
@@
 knitr::opts_knit$set(root.dir = home_dir)
+set.seed(1017)
 ```

### 2024-08-19 Christine Chan/EnVision CF/christine_CGM_OGTT_category.qmd
preliminary CGM for EnVision

---
title: "CGM and OGTT Categories"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
library(gtsummary)
library(dplyr)
library(pracma)
library(ggplot2)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/tim/OneDrive/Vigers/CF"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
# Import
df <- read.csv("./Christine Chan/EnVision CF/Data_Clean/analysis_dataset.csv",
  na.strings = ""
)
# Columns
glucose <- paste0("Glucose_", c(0, 10, 30, 60, 90, 120, 150, 180))
insulin <- paste0("Insulin_", c(0, 10, 30, 60, 90, 120, 150, 180))
# Get diagnosis
df$Diagnosis <- NA
df$Diagnosis[df[, glucose[1]] < 100 &
  (rowSums(df[, glucose[2:5]] < 200, na.rm = T) == rowSums(!is.na(df[, glucose[2:5]]))) &
  (rowSums(df[, glucose[6:8]] < 140, na.rm = T) == rowSums(!is.na(df[, glucose[6:8]])))] <- "NGT"
df$Diagnosis[df[, glucose[1]] >= 100] <- "IFG"
df$Diagnosis[rowSums(df[, glucose[2:5]] >= 200, na.rm = T) > 0] <- "INDET"
df$Diagnosis[rowSums(df[, glucose[6:8]] >= 140, na.rm = T) > 0] <- "IGF"
df$Diagnosis[df$timepoint_0_min >= 126 | df$timepoint_120_min >= 200] <- "CFRD"
df$Diagnosis <- factor(df$Diagnosis, levels = c("NGT", "IFG", "INDET", "IGF", "CFRD"))
# iAUCs
df$iAUC30gluc <- apply(df, 1, function(r) {
  y <- as.numeric(r[glucose[1:3]]) - as.numeric(r[glucose[1]])
  if (is.na(last(y))) {
    return(NA)
  } else {
    x <- c(0, 10, 30)[!is.na(y)]
    y <- y[!is.na(y)]
    auc <- trapz(x, y)
    return(auc)
  }
})
df$iAUC60gluc <- apply(df, 1, function(r) {
  y <- as.numeric(r[glucose[1:4]]) - as.numeric(r[glucose[1]])
  if (is.na(last(y))) {
    return(NA)
  } else {
    x <- c(0, 10, 30, 60)[!is.na(y)]
    y <- y[!is.na(y)]
    auc <- trapz(x, y)
    return(auc)
  }
})
df$iAUC120gluc <- apply(df, 1, function(r) {
  y <- as.numeric(r[glucose[1:6]]) - as.numeric(r[glucose[1]])
  if (is.na(last(y))) {
    return(NA)
  } else {
    x <- c(0, 10, 30, 60, 90, 120)[!is.na(y)]
    y <- y[!is.na(y)]
    auc <- trapz(x, y)
    return(auc)
  }
})
df$iAUC180gluc <- apply(df, 1, function(r) {
  y <- as.numeric(r[glucose]) - as.numeric(r[glucose[1]])
  if (is.na(last(y))) {
    return(NA)
  } else {
    x <- c(0, 10, 30, 60, 90, 120, 150, 180)[!is.na(y)]
    y <- y[!is.na(y)]
    auc <- trapz(x, y)
    return(auc)
  }
})
df$iAUC30ins <- apply(df, 1, function(r) {
  y <- as.numeric(r[insulin[1:3]]) - as.numeric(r[insulin[1]])
  if (is.na(last(y))) {
    return(NA)
  } else {
    x <- c(0, 10, 30)[!is.na(y)]
    y <- y[!is.na(y)]
    auc <- trapz(x, y)
    return(auc)
  }
})
df$iAUC60ins <- apply(df, 1, function(r) {
  y <- as.numeric(r[insulin[1:4]]) - as.numeric(r[insulin[1]])
  if (is.na(last(y))) {
    return(NA)
  } else {
    x <- c(0, 10, 30, 60)[!is.na(y)]
    y <- y[!is.na(y)]
    auc <- trapz(x, y)
    return(auc)
  }
})
df$iAUC120ins <- apply(df, 1, function(r) {
  y <- as.numeric(r[insulin[1:6]]) - as.numeric(r[insulin[1]])
  if (is.na(last(y))) {
    return(NA)
  } else {
    x <- c(0, 10, 30, 60, 90, 120)[!is.na(y)]
    y <- y[!is.na(y)]
    auc <- trapz(x, y)
    return(auc)
  }
})
df$iAUC180ins <- apply(df, 1, function(r) {
  y <- as.numeric(r[insulin]) - as.numeric(r[insulin[1]])
  if (is.na(last(y))) {
    return(NA)
  } else {
    x <- c(0, 10, 30, 60, 90, 120, 150, 180)[!is.na(y)]
    y <- y[!is.na(y)]
    auc <- trapz(x, y)
    return(auc)
  }
})
# HOMA IR
df$homa_ir <- (df$Glucose_0 * df$Insulin_0) / 405
# Filter out those with no CGM data and those with no associated OGTT
df = df %>% filter(!is.na(mean_sensor),!is.na(Date))
```

# Demographics at first visit

`r sum(duplicated(df$study_id))` participants wore a CGM more than once.

```{r}
#| label: tbl-1
#| tbl-cap: Participant characteristics (overall)
df %>%
  group_by(study_id) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(
    redcap_data_access_group, age_visit:origin_race, cftr_mutation_1,
    cftr_mutation_2, weight
  ) %>%
  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
  add_overall()
```

# Glycemic metrics

iAUC was calculated as the area under the glucose curve from baseline to the timepoint of interest, excluding the area under the fasting glucose level.

## All visits

```{r}
#| label: tbl-2
#| tbl-cap: Participant diagnoses based on OGTT
df %>%
  select(
    redcap_data_access_group, Diagnosis, Glucose_0, Glucose_30,
    Glucose_60, Glucose_120, Glucose_180,
    iAUC30gluc:iAUC180ins, homa_ir
  ) %>%
  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
  add_overall()
```

## Visit 1 only

```{r}
#| label: tbl-3
#| tbl-cap: Participant diagnoses based on OGTT (visit 1 only)
df %>%
  group_by(study_id) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(
    redcap_data_access_group, Diagnosis, Glucose_0, Glucose_30,
    Glucose_60, Glucose_120, Glucose_180,
    iAUC30gluc:iAUC180ins, homa_ir
  ) %>%
  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
  add_overall()
```

# Figure

```{r}
#| label: fig-1
#| fig-cap: Percentages by visit
#| fig-width: 12
#| fig-height: 8
df %>%
  group_by(study_id) %>%
  mutate(`Visit Num.` = row_number()) %>%
  group_by(`Visit Num.`) %>%
  count(Diagnosis) %>%
  mutate(total = sum(n), perc = n / total) %>%
  ggplot(., aes(x = `Visit Num.`, y = perc, fill = Diagnosis)) +
  geom_bar(position = "stack", stat = "identity") +
  geom_text(aes(label = paste0(scales::percent(perc), " (n = ", n, ")")),
    position = position_stack(vjust = 0.5)
  ) +
  theme_bw() +
  theme(
    axis.text.y = element_blank(), axis.ticks.y = element_blank(),
    axis.title.y = element_blank()
  )
```

### 2024-08-19 Christine Chan/EnVision CF/clean_envision_data.R

@@ -2,14 +2,18 @@
 # Setup
 #-------------------------------------------------------------------------------
 # Libraries
-library(tidyverse)
+library(dplyr)
+library(tidyr)
+library(lubridate)
 library(readxl)
+library(stringi)
 library(childsds)
+library(cgmanalysis)
 # Home directory
 home_dir <- switch(Sys.info()["sysname"],
   "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
   "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+  "Linux" = "/home/tim/OneDrive/Vigers/CF"
 )
 setwd(home_dir)
 #-------------------------------------------------------------------------------
@@ -161,13 +165,56 @@ insulin <- insulin %>% distinct()
 insulin <- insulin %>%
   group_by(study_id, Date, Timepoint) %>%
   summarise(Insulin = first(na.omit(Insulin)))
+# Convert to numeric
+insulin$Insulin <- as.numeric(insulin$Insulin)
 #-------------------------------------------------------------------------------
-# Combine everything
+# CGM data
+#-------------------------------------------------------------------------------
+# Clean data (does not need to be run every time)
+# Many of these files appeared to be manually edited and are missing some or all
+# of the first 2 rows. Some were saved as XLSX or TXT files as well.
+# cleandata(
+#   inputdirectory = "./Christine Chan/EnVision CF/Data_Clean/CGM/Manually Edited",
+#   outputdirectory = "./Christine Chan/EnVision CF/Data_Clean/CGM/Cleaned",
+#   id_filename = T
+# )
+# cgmvariables(
+#   inputdirectory = "./Christine Chan/EnVision CF/Data_Clean/CGM/Cleaned",
+#   outputdirectory = "./Christine Chan/EnVision CF/Data_Clean/CGM",
+#   outputname = "cgm_variables"
+# )
+# Add to visit dates
+cgm <- read.csv("./Christine Chan/EnVision CF/Data_Clean/CGM/cgm_variables.csv",
+  na.strings = ""
+)
+cgm$subject_id <- sub("_sensor_raw_data_upload", "", cgm$subject_id)
[…]

### 2024-08-19 Christine Chan/EnVision CF/christine_CGM_OGTT_category.qmd
formatting

@@
 df$Diagnosis[df[, glucose[1]] < 100 &
-  (rowSums(df[, glucose[2:5]] < 200, na.rm = T) == rowSums(!is.na(df[, glucose[2:5]]))) &
-  (rowSums(df[, glucose[6:8]] < 140, na.rm = T) == rowSums(!is.na(df[, glucose[6:8]])))] <- "NGT"
+  (rowSums(df[, glucose[2:5]] < 200, na.rm = T) ==
+    rowSums(!is.na(df[, glucose[2:5]]))) &
+  (rowSums(df[, glucose[6:8]] < 140, na.rm = T) ==
+    rowSums(!is.na(df[, glucose[6:8]])))] <- "NGT"
 df$Diagnosis[df[, glucose[1]] >= 100] <- "IFG"
@@
 df$Diagnosis[df$timepoint_0_min >= 126 | df$timepoint_120_min >= 200] <- "CFRD"
-df$Diagnosis <- factor(df$Diagnosis, levels = c("NGT", "IFG", "INDET", "IGF", "CFRD"))
+df$Diagnosis <- factor(df$Diagnosis,
+  levels = c("NGT", "IFG", "INDET", "IGF", "CFRD")
+)
 # iAUCs
@@
 # Filter out those with no CGM data and those with no associated OGTT
-df = df %>% filter(!is.na(mean_sensor),!is.na(Date))
+df <- df %>% filter(!is.na(mean_sensor), !is.na(Date))
 ```

### 2024-08-22 Christine Chan/EnVision CF/christine_CGM_OGTT_category.qmd
EnVision CGM

--- 2024-08-19 Christine Chan/EnVision CF/christine_CGM_OGTT_category.qmd
@@
 )
-# Columns
-glucose <- paste0("Glucose_", c(0, 10, 30, 60, 90, 120, 150, 180))
-insulin <- paste0("Insulin_", c(0, 10, 30, 60, 90, 120, 150, 180))
-# Get diagnosis
-df$Diagnosis <- NA
-df$Diagnosis[df[, glucose[1]] < 100 &
-  (rowSums(df[, glucose[2:5]] < 200, na.rm = T) == rowSums(!is.na(df[, glucose[2:5]]))) &
-  (rowSums(df[, glucose[6:8]] < 140, na.rm = T) == rowSums(!is.na(df[, glucose[6:8]])))] <- "NGT"
-df$Diagnosis[df[, glucose[1]] >= 100] <- "IFG"
-df$Diagnosis[rowSums(df[, glucose[2:5]] >= 200, na.rm = T) > 0] <- "INDET"
-df$Diagnosis[rowSums(df[, glucose[6:8]] >= 140, na.rm = T) > 0] <- "IGF"
-df$Diagnosis[df$timepoint_0_min >= 126 | df$timepoint_120_min >= 200] <- "CFRD"
-df$Diagnosis <- factor(df$Diagnosis, levels = c("NGT", "IFG", "INDET", "IGF", "CFRD"))
-# iAUCs
-df$iAUC30gluc <- apply(df, 1, function(r) {
-  y <- as.numeric(r[glucose[1:3]]) - as.numeric(r[glucose[1]])
-  if (is.na(last(y))) {
-    return(NA)
-  } else {
-    x <- c(0, 10, 30)[!is.na(y)]
-    y <- y[!is.na(y)]
-    auc <- trapz(x, y)
-    return(auc)
-  }
-})
-df$iAUC60gluc <- apply(df, 1, function(r) {
-  y <- as.numeric(r[glucose[1:4]]) - as.numeric(r[glucose[1]])
-  if (is.na(last(y))) {
-    return(NA)
-  } else {
-    x <- c(0, 10, 30, 60)[!is.na(y)]
-    y <- y[!is.na(y)]
-    auc <- trapz(x, y)
-    return(auc)
-  }
-})
-df$iAUC120gluc <- apply(df, 1, function(r) {
-  y <- as.numeric(r[glucose[1:6]]) - as.numeric(r[glucose[1]])
-  if (is.na(last(y))) {
-    return(NA)
-  } else {
-    x <- c(0, 10, 30, 60, 90, 120)[!is.na(y)]
-    y <- y[!is.na(y)]
-    auc <- trapz(x, y)
-    return(auc)
-  }
-})
-df$iAUC180gluc <- apply(df, 1, function(r) {
-  y <- as.numeric(r[glucose]) - as.numeric(r[glucose[1]])
-  if (is.na(last(y))) {
-    return(NA)
-  } else {
-    x <- c(0, 10, 30, 60, 90, 120, 150, 180)[!is.na(y)]
-    y <- y[!is.na(y)]
-    auc <- trapz(x, y)
-    return(auc)
-  }
-})
-df$iAUC30ins <- apply(df, 1, function(r) {
-  y <- as.numeric(r[insulin[1:3]]) - as.numeric(r[insulin[1]])
-  if (is.na(last(y))) {
-    return(NA)
-  } else {
-    x <- c(0, 10, 30)[!is.na(y)]
-    y <- y[!is.na(y)]
-    auc <- trapz(x, y)
-    return(auc)
-  }
-})
-df$iAUC60ins <- apply(df, 1, function(r) {
-  y <- as.numeric(r[insulin[1:4]]) - as.numeric(r[insulin[1]])
-  if (is.na(last(y))) {
-    return(NA)
-  } else {
-    x <- c(0, 10, 30, 60)[!is.na(y)]
-    y <- y[!is.na(y)]
-    auc <- trapz(x, y)
-    return(auc)
-  }
-})
-df$iAUC120ins <- apply(df, 1, function(r) {
-  y <- as.numeric(r[insulin[1:6]]) - as.numeric(r[insulin[1]])
-  if (is.na(last(y))) {
-    return(NA)
-  } else {
-    x <- c(0, 10, 30, 60, 90, 120)[!is.na(y)]
-    y <- y[!is.na(y)]
-    auc <- trapz(x, y)
-    return(auc)
-  }
-})
-df$iAUC180ins <- apply(df, 1, function(r) {
-  y <- as.numeric(r[insulin]) - as.numeric(r[insulin[1]])
-  if (is.na(last(y))) {
-    return(NA)
-  } else {
-    x <- c(0, 10, 30, 60, 90, 120, 150, 180)[!is.na(y)]
-    y <- y[!is.na(y)]
-    auc <- trapz(x, y)
-    return(auc)
-  }
-})
-# HOMA IR
-df$homa_ir <- (df$Glucose_0 * df$Insulin_0) / 405
 # Filter out those with no CGM data and those with no associated OGTT
-df = df %>% filter(!is.na(mean_sensor),!is.na(Date))
+df <- df %>% filter(!is.na(mean_sensor), !is.na(Date))
+# Scully cutoffs
+df$scully140 <- cut(df$percent_time_over_140,
+  breaks = c(-Inf, 17.5, Inf), right = F
+)
+df$scully180 <- cut(df$percent_time_over_180,
+  breaks = c(-Inf, 3.4, Inf), right = F
+)
 ```
@@
 
+## Time spent >140 mg/dL >=17.5% vs <17.5%
+
+```{r}
+df %>%
+  group_by(study_id) %>%
+  filter(row_number() == 1) %>%
+  ungroup() %>%
+  select(
+    redcap_data_access_group, age_visit:origin_race, cftr_mutation_1,
+    cftr_mutation_2, weight, scully140
+  ) %>%
+  tbl_summary(by = scully140, missing_text = "Missing") %>%
+  add_p() %>%
+  separate_p_footnotes() %>%
+  add_overall()
+```
+
+## Time spent >180 mg/dL >=3.4% vs < 3.4%
+
+```{r}
+df %>%
+  group_by(study_id) %>%
+  filter(row_number() == 1) %>%
+  ungroup() %>%
+  select(
+    redcap_data_access_group, age_visit:origin_race, cftr_mutation_1,
+    cftr_mutation_2, weight, scully180
+  ) %>%
+  tbl_summary(by = scully180, missing_text = "Missing") %>%
+  add_p() %>%
+  separate_p_footnotes() %>%
+  add_overall()
+```
+
 # Glycemic metrics
@@
   select(
-    redcap_data_access_group, Diagnosis, Glucose_0, Glucose_30,
+    redcap_data_access_group, Diagnosis, a1c_result, Glucose_0, Glucose_30,
     Glucose_60, Glucose_120, Glucose_180,
@@
   tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
+  add_overall()
+```
+
+## Time spent >140 mg/dL >=17.5% vs <17.5%
+
+```{r}
+df %>%
+  select(
+    redcap_data_access_group, Diagnosis, a1c_result, Glucose_0, Glucose_30,
+    Glucose_60, Glucose_120, Glucose_180,
+    iAUC30gluc:iAUC180ins, homa_ir, scully140
+  ) %>%
+  tbl_summary(by = scully140, missing_text = "Missing") %>%
+  add_p() %>%
+  separate_p_footnotes() %>%
+  add_overall()
+```
+
+## Time spent >180 mg/dL >=3.4% vs < 3.4%
+
+```{r}
+df %>%
+  select(
+    redcap_data_access_group, Diagnosis, a1c_result, Glucose_0, Glucose_30,
+    Glucose_60, Glucose_120, Glucose_180,
+    iAUC30gluc:iAUC180ins, homa_ir, scully180
+  ) %>%
+  tbl_summary(by = scully180, missing_text = "Missing") %>%
+  add_p() %>%
+  separate_p_footnotes() %>%
   add_overall()
@@
   tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
+  add_overall()
+```
+
+## Time spent >140 mg/dL >=17.5% vs <17.5%
+
+```{r}
+df %>%
+  group_by(study_id) %>%
+  filter(row_number() == 1) %>%
+  ungroup() %>%
+  select(
+    redcap_data_access_group, Diagnosis, a1c_result, Glucose_0, Glucose_30,
+    Glucose_60, Glucose_120, Glucose_180,
+    iAUC30gluc:iAUC180ins, homa_ir, scully140
+  ) %>%
+  tbl_summary(by = scully140, missing_text = "Missing") %>%
+  add_p() %>%
+  separate_p_footnotes() %>%
+  add_overall()
+```
+
+## Time spent >180 mg/dL >=3.4% vs < 3.4%
+
+```{r}
+df %>%
+  group_by(study_id) %>%
+  filter(row_number() == 1) %>%
+  ungroup() %>%
+  select(
+    redcap_data_access_group, Diagnosis, a1c_result, Glucose_0, Glucose_30,
+    Glucose_60, Glucose_120, Glucose_180,
+    iAUC30gluc:iAUC180ins, homa_ir, scully180
+  ) %>%
+  tbl_summary(by = scully180, missing_text = "Missing") %>%
+  add_p() %>%
+  separate_p_footnotes() %>%
   add_overall()

### 2024-08-22 Christine Chan/EnVision CF/clean_envision_data.R

@@ -9,6 +9,7 @@ library(readxl)
 library(stringi)
 library(childsds)
 library(cgmanalysis)
+library(pracma)
 # Home directory
 home_dir <- switch(Sys.info()["sysname"],
   "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
@@ -183,10 +184,10 @@ insulin$Insulin <- as.numeric(insulin$Insulin)
 #   outputdirectory = "./Christine Chan/EnVision CF/Data_Clean/CGM",
 #   outputname = "cgm_variables"
 # )
-# Add to visit dates
 cgm <- read.csv("./Christine Chan/EnVision CF/Data_Clean/CGM/cgm_variables.csv",
   na.strings = ""
 )
+cgm = cgm[-which(cgm$percent_cgm_wear==0),]
 cgm$subject_id <- sub("_sensor_raw_data_upload", "", cgm$subject_id)
 cgm$study_id <- sapply(
   stri_split_fixed(str = cgm$subject_id, pattern = "_", n = 2),
@@ -201,6 +202,29 @@ cgm <- left_join(cgm, redcap %>% select(study_id, redcap_event_name, date_visit)
 )
 cgm <- cgm %>% rename(Date = date_visit)
 cgm$Date <- ymd(cgm$Date)
+# Manually fix dates for CGM replacements/unscheduled visits
+# CC0010 unscheduled_visit_arm_1 starts 2019-10-26. CC0010 visit_1_arm_1 on 2020-07-14 has a CGM associated with it. What is the CGM from 2019-10-26?
+
+# IA0003 unscheduled_visit_arm_1 starts 2019-12-10. IA0003 visit_1_arm_1 has a CGM associated with it on 2019-09-16. What is the CGM from 2019-12-10?
+
+
+cgm$Date[cgm$study_id == "IA0001" &
+  cgm$redcap_event_name == "visit_2_arm_1_replacement_cgm_raw_data_upload"] <-
+  "2020-11-09"
+
+cgm$Date[cgm$study_id == "IA0002" &
+  cgm$redcap_event_name == "visit_2_arm_1_replacement_cgm_raw_data_upload"] <-
+  "2020-08-10"
+
+cgm$Date[cgm$study_id == "IA0014" &
[…]

### 2024-08-22 Emily Muther/body_image_trikafta.qmd
collaped weight_adherence categories

@@
   "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+  "Linux" = "/home/tim/OneDrive/Vigers/CF"
 )
@@
 )
+levels(df$WeightBody_Adherence) = c("Agree","Agree","Neutral","Disagree","Disagree")
 # Labels

### 2024-08-23 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
basically starting over on sweat test analysis

--- 2024-04-08 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
@@
 #| include: false
-library(DiagrammeR)
+library(tidyverse)
 library(readxl)
-library(arsenal)
-library(tidyverse)
-library(factoextra)
-library(pROC)
-library(knitr)
-library(caret)
-source("~/GitHub/CF/genotype_class_severity.R")
 home_dir <- switch(Sys.info()["sysname"],
@@
   "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Test",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Test"
+  "Linux" = "/home/tim/OneDrive/Vigers/CF/Edith Zemanick/Sweat Test"
 )
@@
 ```{r data cleaning}
-# Import exclusion data from Excel
-diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
-exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) | diagnoses$Code == "Exclude"]
-exclude <- exclude[!is.na(exclude)]
-# Import data
-df <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv", na.strings = "")
-n1 <- nrow(df)
-n2 <- length(unique(df$PatientID))
-# Remove exclusions from Angela
-df <- df[!(df$PatientID %in% exclude), ]
-# Diagnosis per Angela
-df$Diagnosis <- factor(diagnoses$Diagnosis[match(df$PatientID, diagnoses$PatientID)],
-  levels = c("CFTR-related disorder", "CRMS", "Cystic Fibrosis", "Non-CF"),
-  labels = c("CFTR-RD or CRMS", "CFTR-RD or CRMS", "Cystic Fibrosis", "Non-CF")
-)
-# Add modulator info
-cftr <- read.csv("./Data_Raw/Copy of Report CFTR Modulators.csv")
-df$Modulator <- apply(df, 1, function(r) {
-  mod <- cftr[cftr$Patient.ID == as.numeric(r["PatientID"]), ]
-  if (any(duplicated(mod$Age.at.Start..years.))) {
-    mod <- mod[-which(duplicated(mod$Age.at.Start..years.)), ]
-  }
-  mod <- mod[order(mod$Age.at.Start..years.), ]
-  age <- as.numeric(r["Age.at.Test.in.Years"])
-  as.character(cut(age, c(-Inf, mod$Age.at.Start..years., Inf),
-    labels = c("None", mod$Modulator), right = F
-  ))
-})
-# Numeric variables
-num_vars <- c("Amount", "Na", "K", "Cl")
-num_vars <- c(paste0(num_vars, "_Left"), paste0(num_vars, "_Right"))
-df[, num_vars] <- suppressWarnings(lapply(df[, num_vars], as.numeric))
-# Filter per Maxie's code
-exclude2 <- c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793)
-no_dem <- unique(df$PatientID[df$Demographics.Available != "TRUE" | is.na(df$Demographics.Available) | df$Sex == ""])
-df <- df %>%
-  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks * 7) %>%
-  filter(
-    Age.at.Test.in.Years < 50 & Age.at.Test.in.Days >= 2,
-    !(Diagnosis == "Non-CF" & PatientID %in% exclude2),
-    Modulator == "None"
-  ) %>%
-  distinct(.) %>%
-  filter(Sex != "", Demographics.Available == "TRUE")
-# Genotype info
-## By severity
-df$geno1 <- cf_genotype_class_severity(df$Genotypes1)
-df$geno2 <- cf_genotype_class_severity(df$Genotypes2)
-df$geno1[df$Diagnosis == "Non-CF"] <- "Not CF"
-df$geno2[df$Diagnosis == "Non-CF"] <- "Not CF"
-df$geno_risk <- paste(df$geno1, df$geno2, sep = "_")
-df$geno_risk <- factor(df$geno_risk,
-  levels = c(
-    "Severe_Severe", "Mild_Severe", "Mild_Unknown", "Severe_Mild", "Mild_Mild",
-    "Not CF_Not CF", "Severe_Unknown", "Unknown_Severe", "Unknown_Unknown"
-  ),
-  labels = c(
-    "High", "Low", "Low", "Low", "Low", "Not CF", "Unknown",
-    "Unknown", "Unknown"
-  )
-)
-## By function
-df$geno1_class <- cf_genotype_class_severity(df$Genotypes1, out = "class")
-df$geno1_class[grep("508", df$Genotypes1)] <- "F508del"
-df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("I", "II", "III"), "MF")
-df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("IV", "V"), "RF")
 
-df$geno2_class <- cf_genotype_class_severity(df$Genotypes2, out = "class")
-df$geno2_class[grep("508", df$Genotypes2)] <- "F508del"
-df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("I", "II", "III"), "MF")
-df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("IV", "V"), "RF")
-
-df$geno_class <- factor(paste0(
-  pmin(df$geno1_class, df$geno2_class), "/",
-  pmax(df$geno1_class, df$geno2_class)
-))
-# Which arm?
-df$Arm <- apply(df[, c("Cl", "Cl_Left", "Cl_Right", "Na", "Na_Left", "Na_Right", "K", "K_Left", "K_Right")], 1, function(r) {
-  k <- sum(!is.na(r))
-  if (k == 6) {
-    return("Both")
-  } else if (k == 3) {
-    return("Arm Not Noted")
-  } else {
-    return(NA)
-  }
-})
-# Merge left and right values
-df$Amount <- apply(df[, c("Amount", "Amount_Left", "Amount_Right")], 1, function(r) {
-  amt <- suppressWarnings(as.numeric(as.character(r)))
-  mean(amt, na.rm = T)
-})
-df$Cl <- apply(df[, c("Cl", "Cl_Left", "Cl_Right")], 1, function(r) {
-  cl <- suppressWarnings(as.numeric(as.character(r)))
-  mean(cl, na.rm = T)
-})
-df$Na <- apply(df[, c("Na", "Na_Left", "Na_Right")], 1, function(r) {
-  na <- suppressWarnings(as.numeric(as.character(r)))
-  mean(na, na.rm = T)
-})
-df$K <- apply(df[, c("K", "K_Left", "K_Right")], 1, function(r) {
-  k <- suppressWarnings(as.numeric(as.character(r)))
-  mean(k, na.rm = T)
-})
-# Sodium to chloride ratio
-df$`Na/Cl Ratio` <- df$Na / df$Cl
-high_ratio <- unique(which(df$`Na/Cl Ratio` > 2))
-high_cl <- unique(which(df$Cl > 160))
-high_na <- unique(which(df$Na > 160))
-df <- df[-unique(c(high_ratio, high_cl, high_na)), ]
-# Sweat chloride group
-df$`Sweat Chloride Group` <- cut(df$Cl, c(-Inf, 30, 60, Inf),
-  right = F,
-  labels = c("< 30", "30-59", ">60")
-)
-# Sort
-df <- df %>% arrange(PatientID, Age.at.Test.in.Weeks)
-# Table options (most variables are not normally distributed)
-mycontrols <- tableby.control(
-  numeric.test = "kwt",
-  numeric.stats = c("Nmiss", "median", "q1q3")
-)
-labels <- list(
-  Age.at.Test.in.Years = "Age (Years)",
-  geno_risk = "Genotype Severity", geno_class = "Specific Genotype"
-)
-# First measure for everyone
-t1_df <- df %>%
-  group_by(PatientID) %>%
-  filter(row_number() == 1)
-# Order groups
-df$Diagnosis <- factor(df$Diagnosis, levels = c("Cystic Fibrosis", "CFTR-RD or CRMS", "Non-CF"))
 ```
-
-# Flowchart
-
-```{r}
-grViz("digraph {
-
-# initiate graph
-graph [layout = dot]
-
-# global node settings
-node [shape = rectangle]
-
-# label nodes
-data1 [label = '@@1']
-data2 [label = '@@2']
-ineligible [label = '@@3']
-demo  [label = '@@4']
-highna  [label = '@@5']
-highcl  [label = '@@6']
-highnacl  [label = '@@7']
-review [label = '@@8']
-analysis [label = '@@9']
-ratio [label = '@@10']
-
-# edge definitions with the node IDs
-data1  -> data2
-data2 -> ineligible
-ineligible -> {demo,highna,highcl,highnacl,review}
-data2  -> analysis -> ratio
-highnacl -> analysis [style=invis]
-}
-[1]: paste0('Sweat tests performed \\n ','n = ',n1)
-[2]: paste0('Unique participants \\n ','n = ',n2)
-[3]: paste0('Ineligible \\n ','n = ',n2-nrow(t1_df))
-[4]: paste0('No demographics \\n ','n = ',length(no_dem))
-[5]: paste0('Na > 160 mmol/L \\n ','n = ',length(high_na))
-[6]: paste0('Cl > 160 mmol/L \\n ','n = ',length(high_cl))
-[7]: paste0('Na:Cl > 2 \\n ','n = ',length(high_ratio))
-[8]: paste0('Other \\n ','n = ',1)
-[9]: paste0('Analysis cohort \\n ','n = ',nrow(t1_df))
-[10]: paste0('Both Na and Cl Measured \\n ','n = ',sum(!is.na(df$`Na/Cl Ratio`)))
-")
-```
-
-# Cohorts by Na/Cl ratio
-
-## Based on visual inspection
-
-```{r warning=FALSE}
-# Visual inspection
-ggplot(df, aes(x = Diagnosis, y = `Na/Cl Ratio`)) +
-  geom_boxplot() +
-  geom_hline(yintercept = 1, color = "red") +
-  geom_hline(yintercept = 1.25, color = "red") +
-  theme_classic()
-df$nacl_ratio_cat <- cut(df$`Na/Cl Ratio`, breaks = c(-Inf, 1, 1.25, Inf), right = F)
-kable(table(df$nacl_ratio_cat, df$Diagnosis))
-```
-
-## Based on k-means clustering
-
-### By Na and Cl
-
-```{r warning=FALSE}
-set.seed(1017)
-# K means
-km_df <- df %>%
-  select(Diagnosis, Na, Cl) %>%
-  drop_na()
-ggplot(km_df, aes(x = Na, y = Cl, color = Diagnosis)) +
-  geom_point()
-km <- kmeans(km_df[, c("Na", "Cl")], centers = 3)
-km_df$Cluster <- factor(km$cluster,
-  levels = 1:3,
-  labels = c("High", "Medium", "Low")
-)
-centers <- data.frame(km$centers)
-centers$Cluster <- rownames(km$centers)
-centers$Cluster <- factor(centers$Cluster,
-  levels = 1:3,
-  labels = c("High", "Medium", "Low")
-)
-ggplot(km_df, aes(x = Na, y = Cl, color = Cluster)) +
-  geom_point(shape = 20, alpha = 0.2) +
-  geom_point(data = centers, size = 5)
-rownames(centers) <- NULL
-centers <- centers %>% column_to_rownames("Cluster")
-kable(centers, caption = "K-means centers", row.names = T)
-kable(as.data.frame.matrix(table(km_df$Cluster, km_df$Diagnosis)), row.names = T)
-```
-
-### By Na/Cl ratio
-
-```{r warning=FALSE}
-set.seed(1017)
-# K means
-km_df <- df %>%
-  select(Diagnosis, `Na/Cl Ratio`) %>%
-  drop_na()
-km <- kmeans(km_df[, c("Na/Cl Ratio")], centers = 3)
-km_df$Cluster <- factor(km$cluster,
-  levels = 1:3,
-  labels = c("Low", "Medium", "High")
-)
-centers <- data.frame(km$centers)
-centers$Cluster <- rownames(km$centers)
-centers$Cluster <- factor(centers$Cluster,
-  levels = 1:3,
-  labels = c("Low", "Medium", "High")
-)
-rownames(centers) <- NULL
-centers <- centers %>% column_to_rownames("Cluster")
-kable(centers, caption = "K-means centers", row.names = T)
-kable(as.data.frame.matrix(table(km_df$Cluster, km_df$Diagnosis)), row.names = T)
-```
-
-# Patient Demographics and Sample Characteristics
-
-For those with two sweat electrolyte measures (left and right), the average of the two was used. Each participants' first sweat test value was used for the following tables.
-
-```{r results='asis'}
-t1 <- tableby(
-  Diagnosis ~ Sex + Age.at.Test.in.Years + Method + fe(geno_risk) + fe(geno_class),
-  data = t1_df, control = mycontrols, simulate.p.value = T
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-# Electrolyte comparisons
-
-## By diagnosis
-
-```{r results='asis'}
-t1 <- tableby(
-  Diagnosis ~ Cl + Na + `Na/Cl Ratio` + `Sweat Chloride Group`,
-  data = t1_df, control = mycontrols,
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-### By genotype severity (CF only)
-
-```{r results='asis'}
-t1 <- tableby(
-  geno_risk ~ Cl + Na + `Na/Cl Ratio` + `Sweat Chloride Group`,
-  data = t1_df[t1_df$Diagnosis == "Cystic Fibrosis", ], control = mycontrols,
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-### Pairwise comparisons
-
-Pairwise comparisons using Wilcoxon rank sum test with continuity correction
-
-```{r}
-pairwise.wilcox.test(t1_df$Cl, t1_df$Diagnosis, p.adjust.method = "fdr")
-pairwise.wilcox.test(t1_df$Na, t1_df$Diagnosis, p.adjust.method = "fdr")
-pairwise.wilcox.test(t1_df$`Na/Cl Ratio`, t1_df$Diagnosis, p.adjust.method = "fdr")
-```
-
-## By sweat chloride group
-
-```{r results='asis'}
-t1 <- tableby(
-  `Sweat Chloride Group` ~ fe(geno_risk) + fe(Diagnosis) + Cl + Na + `Na/Cl Ratio`,
-  data = t1_df, control = mycontrols, simulate.p.value = T
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-## By sweat chloride group and diagnosis
-
-### Cystic fibrosis
-
-```{r results='asis'}
-t1 <- tableby(`Sweat Chloride Group` ~ Cl + Na + `Na/Cl Ratio`,
-  data = t1_df[t1_df$Diagnosis == "Cystic Fibrosis", ], control = mycontrols
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-### CFTR-RD or CRMS
-
-```{r results='asis'}
-t1 <- tableby(`Sweat Chloride Group` ~ Cl + Na + `Na/Cl Ratio`,
-  data = t1_df[t1_df$Diagnosis == "CFTR-RD or CRMS", ],
-  control = mycontrols
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-### Non-CF
-
-```{r results='asis'}
-t1 <- tableby(`Sweat Chloride Group` ~ Cl + Na + `Na/Cl Ratio`,
-  data = t1_df[t1_df$Diagnosis == "Non-CF", ], control = mycontrols
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-# Variability
-
-## Within occasion
-
-For all participants with more than one sweat test, we used only their first value to calculate the absolute difference (AD) between left and right arm.
-
-```{r}
-t1_df$`Cl AD` <- abs(t1_df$Cl_Left - t1_df$Cl_Right)
-t1_df$`Na AD` <- abs(t1_df$Na_Left - t1_df$Na_Right)
-```
-
-### By diagnosis
-
-```{r results='asis'}
-t1 <- tableby(
-  Diagnosis ~ `Cl AD` + `Na AD`,
-  data = t1_df, control = mycontrols,
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-### By sweat chloride group
-
-```{r results='asis'}
-t1 <- tableby(
-  `Sweat Chloride Group` ~ `Cl AD` + `Na AD`,
-  data = t1_df, control = mycontrols
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-### By genotype risk
-
-```{r results='asis'}
-t1 <- tableby(
-  geno_risk ~ `Cl AD` + `Na AD`,
-  data = t1_df, control = mycontrols
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-## Between occasion
-
-For all participants with more than one sweat test, we compared their first and last values by calculating the absolute difference (AD) between first and last measure.
-
-```{r results='asis'}
-# Get first and last values for everyone with more than 1
-multiples <- df %>%
-  group_by(PatientID) %>%
-  filter(n() > 1) %>% # Multiple values only
-  filter(row_number() == 1 | row_number() == n()) %>% # First and last
-  summarise(
-    `Cl AD` = abs(last(Cl) - first(Cl)), # Calculate AD
-    `Na AD` = abs(last(Na) - first(Na)),
-    Diagnosis = first(Diagnosis),
-    `Sweat Chloride Group` = first(`Sweat Chloride Group`),
-    geno_risk = first(geno_risk)
-  )
-```
-
-### By diagnosis
-
-```{r results='asis'}
-t1 <- tableby(
-  Diagnosis ~ `Cl AD` + `Na AD`,
-  data = multiples, control = mycontrols,
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-### By sweat chloride group
-
-```{r results='asis'}
-t1 <- tableby(
-  `Sweat Chloride Group` ~ `Cl AD` + `Na AD`,
-  data = multiples, control = mycontrols
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-### By genotype risk
-
-```{r results='asis'}
-t1 <- tableby(
-  geno_risk ~ `Cl AD` + `Na AD`,
-  data = multiples, control = mycontrols
-)
-summary(t1,
-  pfootnote = T,
-  labelTranslations = labels
-)
-```
-
-# Cutoff analyses
-
-We performed receiver operating characteristic (ROC) analysis to find a sweat sodium to sweat chloride cutoff that distinguishes between groups.
-Only the first sweat test from each participant was used.
-
-## Full cohort
-
-The full cohort analyses include all participants.
-
-### Diagnosis
-
-#### ROC
-
-##### Na/Cl Ratio
-
-```{r}
-roc_cf <- roc(
-  response = t1_df$`Diagnosis`,
-  predictor = t1_df$`Na/Cl Ratio`,
-  levels = c("CFTR-RD or CRMS", "Cystic Fibrosis"),
-  quiet = T
-)
-coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)
-
-roc_non_cf <- roc(
-  response = t1_df$`Diagnosis`,
-  predictor = t1_df$`Na/Cl Ratio`,
-  levels = c("Non-CF", "CFTR-RD or CRMS"),
-  quiet = T
-)
-coords_non_cf <- round(coords(roc_non_cf, "best", best.method = "youden"), 3)
-
-plot(roc_cf, print.auc = TRUE,lty=1)
-plot(roc_non_cf, print.auc = TRUE,add=T,lty=2,print.auc.y=0.4)
-legend("bottomright",legend = c("CF vs. CFTR-RD or CRMS","Non-CF vs. CFTR-RD or CRMS"),lty = c(1,2))
-```
-
-The blue curve shows the ROC for distinguishing between CFTR-RD or CRMS vs.
-CF, and the red is the ROC curve for distinguishing between non-CF vs.
-CFTR-RD or CRMS.
-The optimal threshold for distinguishing between CFTR-RD or CRMS vs.
-CF was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index.
-The optimal threshold for distinguishing between non-CF vs.
-CFTR-RD or CRMS was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index.
-The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.
-
-##### Cl Alone
-
-```{r}
-roc_cf_cl <- roc(
-  response = t1_df$`Diagnosis`,
-  predictor = t1_df$Cl,
-  levels = c("CFTR-RD or CRMS", "Cystic Fibrosis"),
-  quiet = T
-)
-coords_cf <- round(coords(roc_cf_cl, "best", best.method = "youden"), 3)
-
-roc_non_cf_cl <- roc(
-  response = t1_df$`Diagnosis`,
-  predictor = t1_df$Cl,
-  levels = c("Non-CF", "CFTR-RD or CRMS"),
-  quiet = T
-)
-coords_non_cf <- round(coords(roc_non_cf_cl, "best", best.method = "youden"), 3)
-
-plot(roc_cf_cl, print.auc = TRUE, col = "blue")
-plot(roc_non_cf_cl, print.auc = TRUE, col = "red")
-```
-
-The blue curve shows the ROC for distinguishing between CFTR-RD or CRMS vs.
-CF, and the red is the ROC curve for distinguishing between non-CF vs.
-CFTR-RD or CRMS.
-The optimal threshold for distinguishing between CFTR-RD or CRMS vs.
-CF was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index.
-The optimal threshold for distinguishing between non-CF vs.
-CFTR-RD or CRMS was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index.
-The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.
-
-##### ROC Comparison
-
-###### CFTR-RD or CRMS vs. CF
-
-```{r}
-roc.test(roc_cf, roc_cf_cl)
-```
-
-###### Non-CF vs. CFTR-RD or CRMS
-
-```{r}
-roc.test(roc_non_cf, roc_non_cf_cl)
-```
-
-#### Histogram by group
-
-Vertical blue line indicates the threshold for CFTR-RD or CRMS vs. CF and red indicates the threshold for non-CF vs. CFTR-RD or CRMS. Na/Cl ratio was log-transformed for easier visualization.
-
-```{r warning=FALSE}
-ggplot(
-  t1_df[!is.na(t1_df$`Diagnosis`), ],
-  aes(log(`Na/Cl Ratio`), fill = `Diagnosis`)
-) +
-  geom_histogram(alpha = 0.3, position = "identity", bins = 50) +
-  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
-  geom_vline(xintercept = log(coords_non_cf$threshold), color = "red") +
-  ylab("Count") +
-  xlab("ln(Na/Cl Ratio)") +
-  theme_bw()
-```
-
-### Genotype
-
-Participants with unknown genotype risk were excluded from this analysis.
-
-#### ROC
-
-```{r}
-# Re-format diagnosis column
-t1_df$`Genotype Risk` <- factor(t1_df$geno_risk,
-  levels = c("Not CF", "Low", "High", "Unknown"),
-  ordered = T
-)
-t1_df$`Genotype Risk`[t1_df$`Genotype Risk` == "Unknown"] <- NA
-```
-
-```{r}
-roc_cf <- roc(
-  response = t1_df$`Genotype Risk`,
-  predictor = t1_df$`Na/Cl Ratio`,
-  levels = c("Low", "High"),
-  quiet = T
-)
-coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)
-
-roc_non_cf <- roc(
-  response = t1_df$`Genotype Risk`,
-  predictor = t1_df$`Na/Cl Ratio`,
-  levels = c("Not CF", "Low"),
-  quiet = T
-)
-coords_non_cf <- round(coords(roc_non_cf, "best", best.method = "youden"), 3)
-
-plot(roc_cf, print.auc = TRUE, col = "blue")
-plot(roc_non_cf, print.auc = TRUE, col = "red", print.auc.y = .4, add = TRUE)
-```
-
-The blue curve shows the ROC for distinguishing between low vs.
-high risk, and the red is the ROC curve for distinguishing between non-CF vs.
-low risk.
-The optimal threshold for distinguishing between low vs.
-high risk was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index.
-The optimal threshold for distinguishing between non-CF vs.
-low risk was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index.
-The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.
-
-#### Histogram by group
-
-Vertical blue line indicates the threshold for low vs. high risk and red indicates the threshold for non-CF vs. low risk. Na/Cl ratio was log-transformed for easier visualization.
-
-```{r warning=FALSE}
-ggplot(
-  t1_df[!is.na(t1_df$`Genotype Risk`), ],
-  aes(log(`Na/Cl Ratio`), fill = `Genotype Risk`)
-) +
-  geom_histogram(alpha = 0.3, position = "identity", bins = 50) +
-  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
-  geom_vline(xintercept = log(coords_non_cf$threshold), color = "red") +
-  ylab("Count") +
-  xlab("ln(Na/Cl Ratio)") +
-  theme_bw()
-```
-
-## Limited cohort
-
-The limited cohort analyses only included participants with sweat chloride < 60.
-
-### Diagnosis
-
-#### ROC
-
-```{r}
-# Exclude those with Cl >= 60
-t1_df <- t1_df %>% filter(Cl < 60)
-```
-
-```{r}
-roc_cf <- roc(
-  response = t1_df$`Diagnosis`,
-  predictor = t1_df$`Na/Cl Ratio`,
-  levels = c("CFTR-RD or CRMS", "Cystic Fibrosis"),
-  quiet = T
-)
-coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)
-
-roc_non_cf <- roc(
-  response = t1_df$`Diagnosis`,
-  predictor = t1_df$`Na/Cl Ratio`,
-  levels = c("Non-CF", "CFTR-RD or CRMS"),
-  quiet = T
-)
-coords_non_cf <- round(coords(roc_non_cf, "best", best.method = "youden"), 3)
-
-plot(roc_cf, print.auc = TRUE, col = "blue")
-plot(roc_non_cf, print.auc = TRUE, col = "red", print.auc.y = .4, add = TRUE)
-```
-
-The blue curve shows the ROC for distinguishing between CFTR-RD or CRMS vs.
-CF, and the red is the ROC curve for distinguishing between non-CF vs.
-CFTR-RD or CRMS.
-The optimal threshold for distinguishing between CFTR-RD or CRMS vs.
-CF was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index.
-The optimal threshold for distinguishing between non-CF vs.
-CFTR-RD or CRMS was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index.
-The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.
-
-#### Histogram by group
-
-Vertical blue line indicates the threshold for CFTR-RD or CRMS vs. CF and red indicates the threshold for non-CF vs. CFTR-RD or CRMS. Na/Cl ratio was log-transformed for easier visualization.
-
-```{r warning=FALSE}
-ggplot(
-  t1_df[!is.na(t1_df$`Diagnosis`), ],
-  aes(log(`Na/Cl Ratio`), fill = `Diagnosis`)
-) +
-  geom_histogram(alpha = 0.3, position = "identity", bins = 50) +
-  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
-  geom_vline(xintercept = log(coords_non_cf$threshold), color = "red") +
-  ylab("Count") +
-  xlab("ln(Na/Cl Ratio)") +
-  theme_bw()
-```
-
-### Genotype
-
-Participants with unknown genotype risk were excluded from this analysis.
-
-#### ROC
-
-```{r}
-roc_cf <- roc(
-  response = t1_df$`Genotype Risk`,
-  predictor = t1_df$`Na/Cl Ratio`,
-  levels = c("Low", "High"),
-  quiet = T
-)
-coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)
-
-roc_non_cf <- roc(
-  response = t1_df$`Genotype Risk`,
-  predictor = t1_df$`Na/Cl Ratio`,
-  levels = c("Not CF", "Low"),
-  quiet = T
-)
-coords_non_cf <- round(coords(roc_non_cf, "best", best.method = "youden"), 3)
-
-plot(roc_cf, print.auc = TRUE, col = "blue")
-plot(roc_non_cf, print.auc = TRUE, col = "red", print.auc.y = .4, add = TRUE)
-```
-
-The blue curve shows the ROC for distinguishing between low vs.
-high risk, and the red is the ROC curve for distinguishing between non-CF vs.
-low risk.
-The optimal threshold for distinguishing between low vs.
-high risk was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index.
-The optimal threshold for distinguishing between non-CF vs.
-low risk was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index.
-The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.
-
-#### Histogram by group
-
-Vertical blue line indicates the threshold for low vs. high risk and red indicates the threshold for non-CF vs. low risk. Na/Cl ratio was log-transformed for easier visualization.
-
-```{r warning=FALSE}
-ggplot(
-  t1_df[!is.na(t1_df$`Genotype Risk`), ],
-  aes(log(`Na/Cl Ratio`), fill = `Genotype Risk`)
-) +
-  geom_histogram(alpha = 0.3, position = "identity", bins = 50) +
-  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
-  geom_vline(xintercept = log(coords_non_cf$threshold), color = "red") +
-  ylab("Count") +
-  xlab("ln(Na/Cl Ratio)") +
-  theme_bw()
-```

### 2024-08-23 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
sweat test and body image trikafta

@@
 ```{r data cleaning}
-
+# Import data
+chco <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv",
+  na.strings = ""
+)
+njh <- read_excel("./Data_Raw/SweatVariability2025_CFgenotypes 2024-08-21.xlsx")
+chec <- read.csv("./Data_Raw/CHEC sweat sodium and chloride.csv")
+# Fix diagnoses/exclusion data for CHCO (based on Angela's review)
+diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
+exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) | diagnoses$Code == "Exclude"]
+exclude <- exclude[!is.na(exclude)]
+chco <- chco[!(chco$PatientID %in% exclude), ]
+chco$Diagnosis <- factor(
+  diagnoses$Diagnosis[match(
+    chco$PatientID,
+    diagnoses$PatientID
+  )],
+  levels = c("CFTR-related disorder", "CRMS", "Cystic Fibrosis", "Non-CF"),
+  labels = c("CFTR-RD or CRMS", "CFTR-RD or CRMS", "Cystic Fibrosis", "Non-CF")
+)
 ```

### 2024-08-23 Emily Muther/body_image_trikafta.qmd
sweat test and body image trikafta

@@
     page-layout: full
-execute:
-  echo: false
 editor: source
@@
 )
-levels(df$WeightBody_Adherence) = c("Agree","Agree","Neutral","Disagree","Disagree")
+levels(df$WeightBody_Adherence) <- c("Agree", "Agree", "Neutral", "Disagree", "Disagree")
 # Labels
@@
       ),
-      all_of(normal_dist) ~ c("{mean} ({sd})","{min}, {max}")
+      all_of(normal_dist) ~ c("{mean} ({sd})", "{min}, {max}")
     ),
@@
 
-## Adherence
+## Adherence by FRS category
 

### 2024-08-27 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
CFF GCT study

--- 2024-08-07 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
@@
 #| include: false
-library(tidyverse)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
 library(gtsummary)
@@
 Column names are printed in terms of percentiles, so `threshold.2.5.` indicates the lower limit of the confidence interval, `threshold.50.` indicates the median best threshold, and `threshold.97.5.` indicates the upper limit of the confidence interval.
+All non-bootstrapped thresholds are also reported.
 
@@
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
   fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all"))) %>% fmt_number(n_sigfig = 3)
 ```

### 2024-08-27 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
CFF GCT study

---
title: "CFF GCT Study"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-height: 9
    fig-width: 12
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /home/tim/Documents/Miscellaneous/zotero.bib
csl: /home/tim/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(pROC)
library(gt)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/tim/OneDrive/Vigers/CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/Tim/GitHub",
  "Linux" = "/home/tim/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| warning: false
# Import from REDCap
df <- read_excel("./Christine Chan/CFF GCT Study/Data_Raw/CFF Study REDCap Export 2021-09-09a LSP.xlsx")
# Get diagnosis
df$Diagnosis <- "NGT"
df$Diagnosis[df$plas_pre_gluc >= 100] <- "IFG"
df$Diagnosis[df$plasma_gluc_1_hr >= 200] <- "INDET"
df$Diagnosis[df$plas_gluc_2_hr >= 140] <- "IGF"
df$Diagnosis[df$plas_pre_gluc >= 126 | df$plas_gluc_2_hr >= 200] <- "CFRD"
df$Diagnosis <- factor(df$Diagnosis,
  levels = c("NGT", "IFG", "INDET", "IGF", "CFRD")
)
df$Dysglycemia <- factor(df$Diagnosis,
  levels = c("NGT", "IFG", "INDET", "IGF", "CFRD"),
  labels = c("NGT", "AGT", "AGT", "AGT", "AGT")
)
# Numeric variables
df$plasma_glu_val <- as.numeric(df$plasma_glu_val)
df$plasma_glucose <- as.numeric(df$plasma_glucose)
# Data were double entered but it wasn't completed so most participants have 3
# rows
df <- df %>%
  arrange(subject_id...1) %>%
  filter(grepl("--1|--2|-1|-2", subject_id...1)) %>%
  mutate(id = sub("--1|--2|-1|-2", "", subject_id...1)) %>%
  group_by(id) %>%
  filter(row_number() == 1) %>%
  select(
    id, gender2:ethnic_origin_race___6, age_at_enrollment_years,
    pancreatic_exocrine_insuff, Diagnosis, Dysglycemia,
    plasma_glucose, capillary, plasma_glu_val, hour_capillary_glucose_val,
    hemo_a1c, plas_pre_gluc, plasma_gluc_1_hr, plas_gluc_2_hr
  )
```

# OGTT Figures

```{r}
#| label: fig-boxplot
#| fig-cap: Boxplot of OGTT glucose values by diabetes status
#| warning: false
plot_df <- df %>% pivot_longer(plas_pre_gluc:plas_gluc_2_hr,
  names_to = "ogtt_timepoint", values_to = "glucose"
)
plot_df$ogtt_timepoint <- factor(plot_df$ogtt_timepoint,
  levels = c("plas_pre_gluc", "plasma_gluc_1_hr", "plas_gluc_2_hr")
)
box_p <- ggplot(
  plot_df,
  aes(x = ogtt_timepoint, y = glucose, fill = ogtt_timepoint)
) +
  geom_boxplot() +
  facet_wrap(~Diagnosis) +
  theme(legend.position = "none")
box_p
```

```{r}
#| label: fig-point
#| fig-cap: Point plot of OGTT glucose values by diabetes status
#| warning: false
plot_df <- df %>% pivot_longer(plas_pre_gluc:plas_gluc_2_hr,
  names_to = "ogtt_timepoint", values_to = "glucose"
)
plot_df$ogtt_timepoint <- factor(plot_df$ogtt_timepoint,
  levels = c("plas_pre_gluc", "plasma_gluc_1_hr", "plas_gluc_2_hr")
)
point_p <- ggplot(
  plot_df,
  aes(x = ogtt_timepoint, y = glucose, color = ogtt_timepoint)
) +
  geom_point() +
  facet_wrap(~Diagnosis) +
  theme(legend.position = "none")
point_p
```

```{r}
#| label: fig-violin
#| fig-cap: Violin plot of OGTT glucose values by diabetes status
#| warning: false
plot_df <- df %>% pivot_longer(plas_pre_gluc:plas_gluc_2_hr,
  names_to = "ogtt_timepoint", values_to = "glucose"
)
plot_df$ogtt_timepoint <- factor(plot_df$ogtt_timepoint,
  levels = c("plas_pre_gluc", "plasma_gluc_1_hr", "plas_gluc_2_hr")
)
violin_p <- ggplot(
  plot_df,
  aes(x = ogtt_timepoint, y = glucose, fill = ogtt_timepoint)
) +
  geom_violin() +
  facet_wrap(~Diagnosis) +
  theme(legend.position = "none")
violin_p
```

# ROC analyses

Dysglycemia was defined as any diagnosis on OGTT other than normal glucose tolerance (NGT).

```{r}
#| label: tbl-roc-dysglycemia
#| tbl-cap: Accuracy of different screening tests to predict dysglycemia using ROC curve analyses
#| message: false
gctpl_roc <- roc(Dysglycemia ~ plasma_glu_val, data = df)
gctcap_roc <- roc(Dysglycemia ~ hour_capillary_glucose_val, data = df)
ogtt1_roc <- roc(Dysglycemia ~ plasma_gluc_1_hr, data = df)
rpg_roc <- roc(Dysglycemia ~ plasma_glucose, data = df)
rcg_roc <- roc(Dysglycemia ~ capillary, data = df)
a1c_roc <- roc(Dysglycemia ~ hemo_a1c, data = df)
roc_table <-
  lapply(
    list(gctpl_roc, gctcap_roc, ogtt1_roc, rpg_roc, rcg_roc, a1c_roc),
    function(roc) {
      coords(roc, "best", ret = "all")
    }
  )
roc_table <- do.call(rbind, roc_table)
rownames(roc_table) <- c("GCTpl", "GCTcap", "OGTT 1 hr", "RPG", "RCG", "A1c")
gt(roc_table, rownames_to_stub = T) %>% fmt_number(decimals = 3)
```

```{r}
invisible(lapply(
  list(gctpl_roc, gctcap_roc, ogtt1_roc, rpg_roc, rcg_roc, a1c_roc),
  function(roc) {
    print(plot(roc, print.auc = T, print.thres = T))
  }
))
```

```{r}
#| label: tbl-roc-cutoffs
#| tbl-cap: Sensitivity and specificity (95% bootstrap confidence intervals) to measure the performance of the glucose challenge test plasma glucose (GCTpl) to predict dysglycemia at different thresholds
#| message: false
gctpl_140 <- as.data.frame(ci.coords(gctpl_roc,
  x = 140,
  ret = c("sensitivity", "specificity")
))
gctpl_150 <- as.data.frame(ci.coords(gctpl_roc,
  x = 150,
  ret = c("sensitivity", "specificity")
))
gctpl_160 <- as.data.frame(ci.coords(gctpl_roc,
  x = 160,
  ret = c("sensitivity", "specificity")
))
gctpl_table <- do.call(rbind, list(gctpl_140, gctpl_150, gctpl_160))
rownames(gctpl_table) <- c(140, 150, 160)
gt(gctpl_table, rownames_to_stub = T) %>% fmt_number(decimals = 3)
```

### 2024-08-27 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
GCT table 1

@@
 library(pROC)
+library(gtsummary)
 library(gt)
@@
   filter(row_number() == 1) %>%
+  ungroup() %>%
   select(
@@
   )
+```
+
+# Demographics
+
+```{r}
+#| label: table1a
+#| tbl-cap: Participant characteristics by diagnosis
+#| message: false
+df %>%
+  select(
+    age_at_enrollment_years, gender2:ethnic_origin_race___6,
+    pancreatic_exocrine_insuff, Diagnosis
+  ) %>%
+  tbl_summary(by = Diagnosis) %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes()
+```
+
+```{r}
+#| label: table1b
+#| tbl-cap: Participant characteristics by dysglycemia
+#| message: false
+df %>%
+  select(
+    age_at_enrollment_years, gender2:ethnic_origin_race___6,
+    pancreatic_exocrine_insuff, Dysglycemia
+  ) %>%
+  tbl_summary(by = Dysglycemia) %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes()
 ```

### 2024-08-28 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
initial GCT analysis finished

@@
 # Import from REDCap
-df <- read_excel("./Christine Chan/CFF GCT Study/Data_Raw/CFF Study REDCap Export 2021-09-09a LSP.xlsx")
+df <- read.csv("./Christine Chan/CFF GCT Study/Data_Raw/CysticFibrosisScreen_DATA_LABELS_2024-04-19_1740.csv",
+  na.strings = ""
+)
+# Create standardized study ID
+df$ID <- sub("--1|--2|--01|-1|-2", "", df[, 1])
+# Numeric variables
+df$Plasma.Glucose <- as.numeric(df$Plasma.Glucose)
+df$X1.hour.Plasma.Glucose.Value <- as.numeric(df$X1.hour.Plasma.Glucose.Value)
+# Combine race and ethnicity columns
+races <- c(
+  "White", "Black/African American", "American Indian/Alaska Native",
+  "Asian Non-Indian", "Asian Indian", "Native Hawaiian/Other Pacific Islander"
+)
+ethnicities <- c("Hispanic/Latino", "Not Hispanic/Latino", "N/A")
+df$Race <- apply(
+  df[, grep("Ethnic.Origin.Race", colnames(df))], 1,
+  function(r) {
+    paste0(races[which(r == "Checked")], collapse = "/")
+  }
+)
+df$Race <- factor(df$Race, levels = races)
+df$Ethnicity <- apply(
+  df[, grep("Ethnicity", colnames(df))], 1,
+  function(r) {
+    paste0(ethnicities[which(r == "Checked")], collapse = "/")
+  }
+)
+df$Ethnicity <- factor(df$Ethnicity, levels = ethnicities[1:2])
+# Fill in missing data for every column and take the first row, then select and
+# rename columns
+df <- df %>%
+  group_by(ID) %>%
+  fill(everything(), .direction = "downup") %>%
+  filter(row_number() == 1) %>%
+  ungroup() %>%
+  rename(
+    "OGTT 0 min" = Plasma.Glucose..pre.glucose.drink...mg.dl.,
+    "OGTT 60 min" = Plasma.Glucose..1.hour.after.glucose.drink...mg.dl.,
+    "OGTT 120 min" = Plasma.Glucose..2.hour.after.glucose.drink...mg.dl.,
+    "GCT 0 min plasma" = Plasma.Glucose,
+    "GCT 0 min capillary" = Capillary,
+    "GCT 60 min plasma" = X1.hour.Plasma.Glucose.Value,
+    "GCT 60 min capillary" = X1.hour.Capillary.Glucose.Value,
+    "HbA1c" = Hemoglobin.A1c..,
+    "Age" = Age.at.enrollment..years.,
+    "Pancreatic insufficient" = Pancreatic.exocrine.insufficient
+  ) %>%
+  select(
+    ID, HbA1c, Age, Gender, Race, Ethnicity, `Pancreatic insufficient`,
+    contains("GCT"), contains("OGTT ")
+  )
 # Get diagnosis
 df$Diagnosis <- "NGT"
-df$Diagnosis[df$plas_pre_gluc >= 100] <- "IFG"
-df$Diagnosis[df$plasma_gluc_1_hr >= 200] <- "INDET"
-df$Diagnosis[df$plas_gluc_2_hr >= 140] <- "IGF"
-df$Diagnosis[df$plas_pre_gluc >= 126 | df$plas_gluc_2_hr >= 200] <- "CFRD"
+df$Diagnosis[df$`OGTT 0 min` >= 100] <- "IFG"
+df$Diagnosis[df$`OGTT 60 min` >= 200] <- "INDET"
+df$Diagnosis[df$`OGTT 120 min` >= 140] <- "IGT"
+df$Diagnosis[df$`OGTT 120 min` >= 200] <- "CFRD"
+df[which(rowSums(is.na(df[, grep("OGTT", colnames(df))])) == 3), "Diagnosis"] <- NA
 df$Diagnosis <- factor(df$Diagnosis,
-  levels = c("NGT", "IFG", "INDET", "IGF", "CFRD")
+  levels = c("NGT", "IFG", "INDET", "IGT", "CFRD")
 )
 df$Dysglycemia <- factor(df$Diagnosis,
-  levels = c("NGT", "IFG", "INDET", "IGF", "CFRD"),
+  levels = c("NGT", "IFG", "INDET", "IGT", "CFRD"),
   labels = c("NGT", "AGT", "AGT", "AGT", "AGT")
 )
-# Numeric variables
-df$plasma_glu_val <- as.numeric(df$plasma_glu_val)
-df$plasma_glucose <- as.numeric(df$plasma_glucose)
-# Data were double entered but it wasn't completed so most participants have 3
-# rows
-df <- df %>%
-  arrange(subject_id...1) %>%
-  filter(grepl("--1|--2|-1|-2", subject_id...1)) %>%
-  mutate(id = sub("--1|--2|-1|-2", "", subject_id...1)) %>%
-  group_by(id) %>%
-  filter(row_number() == 1) %>%
-  ungroup() %>%
-  select(
-    id, gender2:ethnic_origin_race___6, age_at_enrollment_years,
-    pancreatic_exocrine_insuff, Diagnosis, Dysglycemia,
-    plasma_glucose, capillary, plasma_glu_val, hour_capillary_glucose_val,
-    hemo_a1c, plas_pre_gluc, plasma_gluc_1_hr, plas_gluc_2_hr
-  )
+# Write for Christine to check
+write.csv(df,
+          file = "./Christine Chan/CFF GCT Study/Data_Clean/gct_analysis_data.csv",
+          row.names = F,na = "")
+# Remove those without OGTT data
+no_ogtt <- df$ID[which(is.na(df$Diagnosis))]
+df <- df %>% filter(!ID %in% no_ogtt)
 ```
@@
 # Demographics
+
+A total of `r length(no_ogtt)` participants were excluded because they had no OGTT data.
 
@@
 df %>%
-  select(
-    age_at_enrollment_years, gender2:ethnic_origin_race___6,
-    pancreatic_exocrine_insuff, Diagnosis
-  ) %>%
-  tbl_summary(by = Diagnosis) %>%
+  select(-ID, -Dysglycemia) %>%
+  tbl_summary(by = Diagnosis, missing_text = "(Missing)") %>%
   add_overall() %>%
@@
 df %>%
-  select(
-    age_at_enrollment_years, gender2:ethnic_origin_race___6,
-    pancreatic_exocrine_insuff, Dysglycemia
-  ) %>%
+  select(-ID) %>%
   tbl_summary(by = Dysglycemia) %>%
@@
 #| warning: false
-plot_df <- df %>% pivot_longer(plas_pre_gluc:plas_gluc_2_hr,
-  names_to = "ogtt_timepoint", values_to = "glucose"
+plot_df <- df %>% pivot_longer(contains("OGTT"),
+  names_to = "OGTT Timepoint", values_to = "Glucose"
 )
-plot_df$ogtt_timepoint <- factor(plot_df$ogtt_timepoint,
-  levels = c("plas_pre_gluc", "plasma_gluc_1_hr", "plas_gluc_2_hr")
+plot_df$`OGTT Timepoint` <- factor(plot_df$`OGTT Timepoint`,
+  levels = c("OGTT 0 min", "OGTT 60 min", "OGTT 120 min")
 )
@@
   plot_df,
-  aes(x = ogtt_timepoint, y = glucose, fill = ogtt_timepoint)
+  aes(x = `OGTT Timepoint`, y = Glucose, fill = `OGTT Timepoint`)
 ) +
   geom_boxplot() +
+  scale_y_continuous(breaks = c(100, 140, 200, 300)) +
   facet_wrap(~Diagnosis) +
@@
 #| warning: false
-plot_df <- df %>% pivot_longer(plas_pre_gluc:plas_gluc_2_hr,
-  names_to = "ogtt_timepoint", values_to = "glucose"
-)
-plot_df$ogtt_timepoint <- factor(plot_df$ogtt_timepoint,
-  levels = c("plas_pre_gluc", "plasma_gluc_1_hr", "plas_gluc_2_hr")
-)
 point_p <- ggplot(
   plot_df,
-  aes(x = ogtt_timepoint, y = glucose, color = ogtt_timepoint)
+  aes(x = `OGTT Timepoint`, y = Glucose, color = `OGTT Timepoint`)
 ) +
-  geom_point() +
+  geom_jitter() +
+  scale_y_continuous(breaks = c(100, 140, 200, 300)) +
   facet_wrap(~Diagnosis) +
@@
 #| warning: false
-plot_df <- df %>% pivot_longer(plas_pre_gluc:plas_gluc_2_hr,
-  names_to = "ogtt_timepoint", values_to = "glucose"
-)
-plot_df$ogtt_timepoint <- factor(plot_df$ogtt_timepoint,
-  levels = c("plas_pre_gluc", "plasma_gluc_1_hr", "plas_gluc_2_hr")
-)
 violin_p <- ggplot(
   plot_df,
-  aes(x = ogtt_timepoint, y = glucose, fill = ogtt_timepoint)
+  aes(x = `OGTT Timepoint`, y = Glucose, fill = `OGTT Timepoint`)
 ) +
   geom_violin() +
+  scale_y_continuous(breaks = c(100, 140, 200, 300)) +
   facet_wrap(~Diagnosis) +
@@
 #| message: false
-gctpl_roc <- roc(Dysglycemia ~ plasma_glu_val, data = df)
-gctcap_roc <- roc(Dysglycemia ~ hour_capillary_glucose_val, data = df)
-ogtt1_roc <- roc(Dysglycemia ~ plasma_gluc_1_hr, data = df)
-rpg_roc <- roc(Dysglycemia ~ plasma_glucose, data = df)
-rcg_roc <- roc(Dysglycemia ~ capillary, data = df)
-a1c_roc <- roc(Dysglycemia ~ hemo_a1c, data = df)
+gctpl_roc <- roc(Dysglycemia ~ `GCT 60 min plasma`, data = df)
+gctcap_roc <- roc(Dysglycemia ~ `GCT 60 min capillary`, data = df)
+ogtt1_roc <- roc(Dysglycemia ~ `OGTT 60 min`, data = df)
+rpg_roc <- roc(Dysglycemia ~ `GCT 0 min plasma`, data = df)
+rcg_roc <- roc(Dysglycemia ~ `GCT 0 min capillary`, data = df)
+a1c_roc <- roc(Dysglycemia ~ HbA1c, data = df)
 roc_table <-

### 2024-08-28 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
added ethnicity to CFFPR ROC curves

--- 2024-08-07 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
@@
     fig-cap-location: top
-    fig-width: 8
-    fig-height: 8
+    fig-width: 12
+    fig-height: 9
     page-layout: full
@@
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
   fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
   fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+```
+
+### By ethnicity
+
+#### Hispanic
+
+```{r}
+hispanic_ngt_agt <- ngt_agt %>% filter(Hispanicrace == "Yes")
+hispanic_agt_cfrd <- agt_cfrd %>% filter(Hispanicrace == "No")
+```
+
+##### NGT vs. AGT
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = hispanic_ngt_agt)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+```
+
+##### AGT vs. CFRD
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = hispanic_agt_cfrd)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+```
+
+#### Non-Hispanic
+
+```{r}
+non_hispanic_ngt_agt <-
+  ngt_agt %>% filter(Race == "Black or African American")
+non_hispanic_agt_cfrd <-
+  agt_cfrd %>% filter(Race == "Black or African American")
+```
+
+##### NGT vs. AGT
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = non_hispanic_ngt_agt)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+```
+
+##### AGT vs. CFRD
+
+Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = non_hispanic_agt_cfrd)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(coords(r, "best"))) %>%
+  fmt_number(n_sigfig = 3)
+gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
+  fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 ##### $\geq$ 200
+
+Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.
 
@@
 plot(r, print.auc = T, print.thres = T)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
+  fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 ##### $\geq$ 200
-
-Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.
 
@@
   fmt_number(n_sigfig = 3)
-gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
+gt(data.frame(ci.coords(r, "best"))) %>%
   fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
+
+### By ethnicity
+
+#### Hispanic
+
+```{r}
+hispanic_roc <- roc_df %>% filter(Hispanicrace == "Yes")
+```
+
+##### $\geq$ 140
+
+```{r}
+#| message: false
+r <- roc(agt ~ hgba1c, data = hispanic_roc)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+```
+
+##### $\geq$ 200
+
+```{r}
+#| message: false
+r <- roc(cfrd ~ hgba1c, hispanic_roc)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+```
+
+#### Non-Hispanic
+
+```{r}
+non_hispanic_roc <- roc_df %>% filter(Hispanicrace == "No")
+```
+
+##### $\geq$ 140
+
+```{r}
+#| message: false
+r <- roc(agt ~ hgba1c, data = non_hispanic_roc)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+```
+
+##### $\geq$ 200
+
+```{r}
+#| message: false
+r <- roc(cfrd ~ hgba1c, data = non_hispanic_roc)
+plot(r, print.auc = T, print.thres = T)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+```

### 2024-08-30 Scott Sagel/PEx in CF/analysis_for_yaron.qmd
updated analysis for Yaron and Scott

--- 2024-01-02 Scott Sagel/PEx in CF/analysis_for_yaron.qmd
@@
       dark: darkly
-bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Dropbox/Miscellaneous/american-medical-association.csl
+bibliography: /home/tim/Documents/Miscellaneous/zotero.bib
+csl: /home/tim/GitHub/styles/american-medical-association.csl
 editor: source
@@
 #| include: false
-library(tidyverse)
+library(dplyr)
+library(tidyr)
 library(lubridate)
 library(gtsummary)
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
-)
-knitr::opts_knit$set(root.dir = home_dir)
+library(ggpubr)
+library(rspiro)
+knitr::opts_knit$set(root.dir = "/home/tim/OneDrive/Vigers/CF/Scott Sagel/PEx in CF")
+# Colors for plots
+yellow <- "#f8b80a"
+red <- "#e30022"
+purple <- "#7b2670"
+blue <- "#022451"
 ```
@@
 ```{r data cleaning}
+#| warning: false
 # Import
-df <- read.csv("./Scott Sagel/PEx in CF/Data_Raw/ExacerbationsInCF222_DATA_2024-01-02_0857.csv",
+df <- read.csv("./Data_Raw/ExacerbationsInCF222_DATA_2024-01-02_0857.csv",
   na.strings = ""
@@
 )
+# CFTR modulators
 df$cftr_mod <- factor(df$cftr_mod, levels = 0:1, labels = c("No", "Yes"))
+df$cftr_mod_spec[df$cftr_mod == "No"] <- 5
+df$cftr_mod_spec <- factor(df$cftr_mod_spec,
+  levels = 1:5,
+  labels = c("Kalydeco", "Orkambi", "Symdeko", "Trikafta", "None")
+)
+df$Trikafta <- factor(df$cftr_mod_spec == "Trikafta",
+  levels = c(F, T), labels = c("No", "Yes")
+)
+# Other medications
+df$azithro <- factor(df$azithro, levels = 0:1, labels = c("No", "Yes"))
+df$inhaled_abx <- factor(df$inhaled_abx,
+  levels = 0:1,
+  labels = c("No", "Yes")
+)
+df$steroids_system <- factor(df$steroids_system,
+  levels = 0:1,
+  labels = c("No", "Yes")
+)
+# Tests
+df$fungal_test <- factor(df$fungal_test, levels = 0:1, labels = c("No", "Yes"))
+df$ntm_test <- factor(df$ntm_test, levels = 0:1, labels = c("No", "Yes"))
+df$viral_test <- factor(df$viral_test, levels = 0:1, labels = c("No", "Yes"))
+df$viral_test_results <- factor(df$viral_test_results,
+  levels = 0:1, labels = c("Negative", "Positive")
+)
+# % predicted FEV1
+df$sex <- factor(df$sex, levels = c(2, 1), labels = c("Male", "Female"))
+df$fev1_baseline <- rowMeans(df[, paste0("fev1_baseline", 1:6)], na.rm = T)
+# Add fill in demographic variables
+df <- df %>%
+  group_by(sid) %>%
+  fill(
+    sex, race, ethnicity, mutation1, mutation2,
+    insurance___1:insurance___8
+  ) %>%
+  ungroup()
+# Predicted lung function
+df$ppfev1_admit <- pctpred_GLIgl(
+  age = df$age_at_admit, height = df$height / 100, gender = as.numeric(df$sex),
+  FEV1 = df$fev1_admit
+)
+df$ppfev1_baseline <- pctpred_GLIgl(
+  age = df$age_at_admit, height = df$height / 100, gender = as.numeric(df$sex),
+  FEV1 = df$fev1_baseline
+)
+df$ppfev1_discharge <- pctpred_GLIgl(
+  age = df$age_at_admit, height = df$height / 100, gender = as.numeric(df$sex),
+  FEV1 = df$fev1_discharge
+)
+# Demographic variables
+df$age_group <- cut(df$age_at_admit, c(-Inf, 6, 12, 18, Inf),
+  right = F,
+  labels = c("<6 yrs", "≥6-<12 yrs", "≥12-<18 yrs", "≥18 yrs")
+)
+df$race <- factor(df$race,
+  levels = 1:8,
+  labels = c(
+    "White", "Black or African-American", "Other", "Asian", "Other",
+    "Other", "Other", "Unknown/not reported"
+  )
+)
+df$ethnicity <- factor(df$ethnicity,
+  levels = 1:3,
+  labels = c(
+    "Hispanic or Latino", "Not Hispanic or Latino", "Unknown or not reported"
+  )
+)
+df$mutation1 <- factor(df$mutation1,
+  levels = 1:15,
+  labels = c("F508del", rep("Other", times = 14))
+)
+df$mutation2 <- factor(df$mutation2,
+  levels = 1:15,
+  labels = c("F508del", rep("Other", times = 14))
+)
+df$Genotype <- factor(
+  paste0(as.character(df$mutation1), "/", as.character(df$mutation2)),
+  levels = c(
+    "F508del/F508del", "F508del/Other", "Other/F508del", "Other/Other"
+  ),
+  labels = c(
+    "F508del Homozygous", "F508del Heterozygous", "F508del Heterozygous",
+    "Other"
+  )
+)
+insurances <- c(
+  "Private insurance", "Medicare", "Medicaid", "Other",
+  "Tricare/other military health plan", "Other", "Other", "None"
+)
+df$Insurance <- apply(
+  df[, grep("insurance___", colnames(df))], 1,
+  function(r) {
+    w <- which(r == 1)
+    if (length(w) == 0) {
+      return("None")
+    } else {
+      return(paste0(insurances[which(r == 1)], collapse = "/"))
+    }
+  }
+)
 # Labels
 labels <- list(
+  "Age group at admission" = "age_group",
+  "Age at admission" = "age_at_admit", "Sex" = "sex",
+  "Race" = "race", "Ethnicity" = "ethnicity",
+  "Azithromycin" = "azithro", "Inhaled antibiotics" = "inhaled_abx",
+  "Systemic steroids" = "steroids_system",
+  "Source of respiratory culture at admit" = "admit_cx_source",
   "Aspergillus fumigatus" = "fungi___1",
@@
   "NTM testing performed?" = "ntm_test",
+  "Fungal testing performed?" = "fungal_test",
+  "Viral testing performed?" = "viral_test",
+  "Viral testing results" = "viral_test_results",
+  "Influenza A or B" = "viruses___1",
+  "Respiratory Syncytial Virus" = "viruses___2",
+  "Rhinovirus/enterovirus" = "viruses___3",
+  "SARS CoV-2" = "viruses___4",
+  "Other virus" = "viruses___5",
   "M. abscessus species" = "ntm___1",
@@
   "Other bacteria" = "admit_cx_results___9",
-  "CFTR Modulator" = "cftr_mod",
-  "Cultures Tested" = "redcap_data_access_group"
+  "On CFTR Modulator?" = "cftr_mod",
+  "CFTR Modulator Type" = "cftr_mod_spec",
+  "Cultures Tested" = "redcap_data_access_group",
+  "Length of Admission" = "days_admit",
+  "Systemic Steroids" = "steroids_system",
+  "Oral Steroids" = "steroids_oral",
+  "IV Steroids" = "steroids_iv",
+  "Baseline FEV1pp (2022)" = "ppfev1_baseline",
+  "Admission FEV1pp (2022)" = "ppfev1_admit",
+  "Discharge FEV1pp (2022)" = "ppfev1_discharge",
+  "Change in mean FEV1pp from admit to discharge" = "ppfev1_change"
 )
-labels <- setNames(names(labels), labels)
+labels <- as.list(setNames(names(labels), labels))
 # Format
@@
   select(
-    sid, redcap_event_name, redcap_data_access_group, year, cftr_mod,
+    sid,
+    redcap_data_access_group, year, sex, age_group, age_at_admit, race,
+    ethnicity, Genotype, Insurance, cftr_mod, cftr_mod_spec,
+    Trikafta, azithro, inhaled_abx, steroids_system, admit_cx_source,
+    ppfev1_baseline, ppfev1_admit, ppfev1_discharge, days_admit,
     fungal_test, fungi___1:fungi___5,
     ntm_test, ntm___1:ntm___3,
-    admit_cx_results___1:admit_cx_results___9
+    viral_test, viruses___1:viruses___5,
+    admit_cx_results___1:admit_cx_results___9,
+    contains("steroids")
   ) %>%
   filter(!is.na(year))
+# Write for Scott and Yaron to check
+write.csv(df,file = "./Data_Clean/analysis_dataset.csv",row.names = F,na = "")
+```
+
+# Demographics
+
+```{r}
+#| label: tbl-table-1
+#| tbl-cap: Participant characteristics at admission
+df %>%
+  select(redcap_data_access_group:days_admit) %>%
+  tbl_summary(
+    by = year, missing_text = "(Missing)", label = labels,
+    statistic =
+      list(
+        all_continuous() ~ "{median} [{p25}, {p75}]",
+        c(age_at_admit, ppfev1_baseline, ppfev1_admit, ppfev1_discharge) ~
+          "{mean} ({sd})"
+      )
+  ) %>%
+  add_p(
+    test =
+      list(
+        c(age_at_admit, ppfev1_baseline, ppfev1_admit, ppfev1_discharge) ~
+          "t.test"
+      ),
+    test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
+  ) %>%
+  separate_p_footnotes()
+```
+
+```{r}
+#| label: tbl-table-1-eti-2022
+#| tbl-cap: Participant characteristics on ETI in 2022
+df %>%
+  filter(Trikafta == "Yes", year == 2022) %>%
+  select(redcap_data_access_group:days_admit, -year) %>%
+  tbl_summary(
+    missing_text = "(Missing)", label = labels,
+    statistic =
+      list(
+        all_continuous() ~ "{median} [{p25}, {p75}]",
+        c(age_at_admit, ppfev1_baseline, ppfev1_admit, ppfev1_discharge) ~
+          "{mean} ({sd})"
+      )
+  )
 ```
@@
 df %>%
-  filter(fungal_test == 1) %>%
+  filter(fungal_test == "Yes") %>%
   select(year, redcap_data_access_group, fungi___1:fungi___5) %>%
@@
 ```{r}
-#| label: tbl-fungi_center
-#| tbl-cap: Fungi by Center
+#| label: tbl-fungi-center
+#| tbl-cap: Fungi by center
 #| message: false
 df %>%
-  filter(fungal_test == 1) %>%
+  filter(fungal_test == "Yes") %>%
   select(year, redcap_data_access_group, fungi___1:fungi___5) %>%
@@
 ```{r}
-#| label: tbl-fungi_cftr_mod
-#| tbl-cap: Fungi by Modulator Status
+#| label: tbl-fungi-cftr-mod
+#| tbl-cap: Fungi by modulator status
 #| message: false
 df %>%
-  filter(fungal_test == 1) %>%
+  filter(fungal_test == "Yes") %>%
   select(year, cftr_mod, fungi___1:fungi___5) %>%
@@
         tbl_summary(by = cftr_mod, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-fungi-cftr-mod-spec
+#| tbl-cap: Fungi by specific modulator status
+#| message: false
+df %>%
+  filter(fungal_test == "Yes") %>%
+  select(year, cftr_mod_spec, fungi___1:fungi___5) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(
+          by = cftr_mod_spec, missing_text = "Missing",
+          label = labels
+        ) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-fungi-trikafta
+#| tbl-cap: Fungi by Trikafta status
+#| message: false
+df %>%
+  filter(fungal_test == "Yes") %>%
+  select(year, Trikafta, fungi___1:fungi___5) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
         add_overall() %>%
@@
 df %>%
-  filter(ntm_test == 1) %>%
+  filter(ntm_test == "Yes") %>%
   select(year, redcap_data_access_group, ntm___1:ntm___3) %>%
@@
 ```{r}
-#| label: tbl-ntm_center
-#| tbl-cap: NTM by Center
+#| label: tbl-ntm-center
+#| tbl-cap: NTM by center
 #| message: false
 df %>%
-  filter(ntm_test == 1) %>%
+  filter(ntm_test == "Yes") %>%
   select(year, redcap_data_access_group, ntm___1:ntm___3) %>%
@@
 ```{r}
-#| label: tbl-ntm_cftr_mod
-#| tbl-cap: NTM by Modulator Status
+#| label: tbl-ntm-cftr-mod
+#| tbl-cap: NTM by modulator status
 #| message: false
 df %>%
-  filter(ntm_test == 1) %>%
+  filter(ntm_test == "Yes") %>%
   select(year, cftr_mod, ntm___1:ntm___3) %>%
@@
       ~ .x %>%
+        tbl_summary(
+          by = cftr_mod, missing_text = "Missing",
+          label = labels
+        ) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-ntm-cftr-mod-spec
+#| tbl-cap: NTM by specific modulator status
+#| message: false
+df %>%
+  filter(ntm_test == "Yes") %>%
+  select(year, cftr_mod_spec, ntm___1:ntm___3) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(
+          by = cftr_mod_spec, missing_text = "Missing",
+          label = labels
+        ) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-ntm-trikafta
+#| tbl-cap: NTM by Trikafta status
+#| message: false
+df %>%
+  filter(ntm_test == "Yes") %>%
+  select(year, Trikafta, ntm___1:ntm___3) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+# Viruses
+
+The tables below are limited to patients who had viral testing performed.
+
+```{r}
+#| label: tbl-viruses
+#| tbl-cap: Viruses
+df %>%
+  filter(viral_test == "Yes") %>%
+  select(year, redcap_data_access_group, viruses___1:viruses___5) %>%
+  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes()
+```
+
+```{r}
+#| label: tbl-viruses-cftr-mod
+#| tbl-cap: Viruses by modulator status
+#| message: false
+df %>%
+  filter(viral_test == "Yes") %>%
+  select(year, cftr_mod, viruses___1:viruses___5) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
         tbl_summary(by = cftr_mod, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-viruses-cftr-mod-spec
+#| tbl-cap: Viruses by specific modulator status
+#| message: false
+df %>%
+  filter(viral_test == "Yes") %>%
+  select(year, cftr_mod_spec, viruses___1:viruses___5) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(
+          by = cftr_mod_spec, missing_text = "Missing",
+          label = labels
+        ) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-viruses-trikafta
+#| tbl-cap: Viruses by Trikafta status
+#| message: false
+df %>%
+  filter(viral_test == "Yes") %>%
+  select(year, Trikafta, viruses___1:viruses___5) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
         add_overall() %>%
@@
   select(
-    year, redcap_data_access_group, admit_cx_results___1:admit_cx_results___9
+    year, redcap_data_access_group,
+    admit_cx_results___1:admit_cx_results___9
   ) %>%
@@
 ```{r}
-#| label: tbl-bacteria_center
-#| tbl-cap: Bacteria by Center
+#| label: tbl-bacteria-center
+#| tbl-cap: Bacteria by center
 #| message: false
@@
 ```{r}
-#| label: tbl-bacteria_cftr_mod
-#| tbl-cap: Bacteria by Modulator Status
+#| label: tbl-bacteria-cftr-mod
+#| tbl-cap: Bacteria by modulator status
 #| message: false
@@
       ~ .x %>%
-        tbl_summary(by = cftr_mod, missing_text = "Missing", label = labels) %>%
+        tbl_summary(
+          by = cftr_mod, missing_text = "Missing",
+          label = labels
+        ) %>%
         add_overall() %>%
@@
 ```
+
+```{r}
+#| label: tbl-bacteria-cftr-mod-spec
+#| tbl-cap: Bacteria by specfic modulator status
+#| message: false
+df %>%
+  select(year, cftr_mod_spec, admit_cx_results___1:admit_cx_results___9) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(
+          by = cftr_mod_spec, missing_text = "Missing",
+          label = labels
+        ) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-bacteria-trikafta
+#| tbl-cap: Bacteria by Trikafta status
+#| message: false
+df %>%
+  select(year, Trikafta, admit_cx_results___1:admit_cx_results___9) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+# Days of admission
+
+```{r}
+#| label: tbl-days
+#| tbl-cap: Length of admission
+df %>%
+  select(year, days_admit) %>%
+  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes()
+```
+
+```{r}
+#| label: tbl-days-center
+#| tbl-cap: Length of admission by center
+#| message: false
+df %>%
+  select(year, redcap_data_access_group, days_admit) %>%
+  tbl_strata(
+    strata = redcap_data_access_group,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-days-cftr-mod
+#| tbl-cap: Length of admission by modulator status
+#| message: false
+df %>%
+  select(year, cftr_mod, days_admit) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(
+          by = cftr_mod, missing_text = "Missing", label = labels
+        ) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-days-cftr-mod-spec
+#| tbl-cap: Length of admission by specific modulator status
+#| message: false
+df %>%
+  select(year, cftr_mod_spec, days_admit) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(
+          by = cftr_mod_spec, missing_text = "Missing", label = labels
+        ) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-days-trikafta
+#| tbl-cap: Length of admission by Trikafta status
+#| message: false
+df %>%
+  select(year, Trikafta, days_admit) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+# Steroids
+
+```{r}
+#| label: tbl-steroids
+#| tbl-cap: Steroids
+df %>%
+  select(year, contains("steroids")) %>%
+  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes()
+```
+
+```{r}
+#| label: tbl-steroids-center
+#| tbl-cap: Steroids by center
+#| message: false
+df %>%
+  select(year, redcap_data_access_group, contains("steroids")) %>%
+  tbl_strata(
+    strata = redcap_data_access_group,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-steroids-cftr-mod
+#| tbl-cap: Steroids by modulator status
+#| message: false
+df %>%
+  select(year, cftr_mod, contains("steroids")) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(
+          by = cftr_mod, missing_text = "Missing", label = labels
+        ) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-steroids-cftr-mod-spec
+#| tbl-cap: Steroids by specific modulator status
+#| message: false
+df %>%
+  select(year, cftr_mod_spec, contains("steroids")) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(
+          by = cftr_mod_spec, missing_text = "Missing", label = labels
+        ) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+```{r}
+#| label: tbl-steroids-trikafta
+#| tbl-cap: Steroids by Trikafta status
+#| message: false
+df %>%
+  select(year, Trikafta, contains("steroids")) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
+        add_overall() %>%
+        add_p() %>%
+        separate_p_footnotes()
+  )
+```
+
+# Plots
+
+```{r}
+#| label: fig-num-hosps
+#| fig-cap: Number of hospitalizations
+df %>%
+  group_by(year) %>%
+  summarise(n = n()) %>%
+  ggplot(aes(x = factor(year), y = n, fill = factor(year))) +
+  geom_bar(position = "dodge", stat = "identity") +
+  geom_label(aes(label = n), nudge_y = -20, fill = "white") +
+  ylab("Num. of Hospitalizations") +
+  theme_bw() +
+  theme(legend.position = "none", axis.title.x = element_blank()) +
+  scale_fill_manual(values = c(purple, yellow))
+```
+
+```{r}
+#| label: fig-perc-hosp-2-plus
+#| fig-cap: Percent hospitalized 2 or more times in both years
+df %>%
+  group_by(sid, year) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  group_by(year) %>%
+  summarise(
+    num = sum(n >= 2), total = n(),
+    perc = mean(n >= 2) * 100, .groups = "drop"
+  ) %>%
+  ggplot(aes(x = factor(year), y = perc, fill = factor(year))) +
+  geom_bar(position = "dodge", stat = "identity") +
+  geom_label(aes(label = scales::percent(perc, scale = 1, accuracy = 0.1)),
+    nudge_y = -5, fill = "white"
+  ) +
+  geom_label(aes(label = paste0(num, "/", total)),
+    nudge_y = -8, fill = "white"
+  ) +
+  geom_signif(
+    xmin = 1, xmax = 2, y_position = 50, map_signif_level = T,
+    annotation = "<0.001"
+  ) +
+  xlab("Site") +
+  ylab("Percent With >= 2 Hospitalizations") +
+  theme_bw() +
+  theme(legend.position = "none") +
+  scale_fill_manual(values = c(purple, yellow))
+```
+
+```{r}
+#| label: fig-num-hosp-2-plus
+#| fig-cap: Number hospitalized 2 or more times in both years
+df %>%
+  group_by(sid, year) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  group_by(year) %>%
+  summarise(
+    num = sum(n >= 2), total = n(),
+    .groups = "drop"
+  ) %>%
+  ggplot(aes(x = factor(year), y = num, fill = factor(year))) +
+  geom_bar(position = "dodge", stat = "identity") +
+  geom_label(aes(label = num), nudge_y = -5, fill = "white") +
+  xlab("Site") +
+  ylab("Number With >= 2 Hospitalizations") +
+  theme_bw() +
+  theme(legend.position = "none") +
+  scale_fill_manual(values = c(purple, yellow))
+```
+
+```{r}
+#| label: fig-fev1-year
+#| fig-cap: FEV1 % predicted by year
+#| warning: false
+df %>%
+  group_by(year, sid) %>%
+  rename(
+    Admission = ppfev1_admit, Discharge = ppfev1_discharge,
+    Baseline = ppfev1_baseline
+  ) %>%
+  pivot_longer(c(Admission, Discharge, Baseline)) %>%
+  mutate(name = factor(name,
+    levels = c("Baseline", "Admission", "Discharge")
+  )) %>%
+  ggplot(aes(
+    x = name, y = value,
+    fill = factor(year)
+  )) +
+  geom_boxplot() +
+  geom_pwc(
+    aes(group = year),
+    tip.length = 0.01,
+    method = "t_test", label = "p.format"
+  ) +
+  ylab("FEV1 % predicted") +
+  theme_bw() +
+  theme(
+    legend.title = element_blank(), legend.position = "bottom",
+    axis.title.x = element_blank()
+  ) +
+  scale_fill_manual(values = c(purple, yellow))
+```
+
+FEV1 percent predicted was calculated using the GLI global (2022) equations with the `rspiro` R package.
+
+```{r}
+#| label: fig-return-90-100
+#| fig-cap: Proportions who returned to >= 90% and >= 100% at discharge
+df %>%
+  mutate(
+    returned_90 = ppfev1_discharge >= ppfev1_baseline * 0.9,
+    returned_100 = ppfev1_discharge >= ppfev1_baseline
+  ) %>%
+  group_by(year) %>%
+  summarise(
+    n_returned_90 = sum(returned_90, na.rm = T),
+    n_returned_100 = sum(returned_100, na.rm = T),
+    total = sum(!is.na(ppfev1_discharge) & !is.na(ppfev1_baseline)),
+    returned_90 = mean(returned_90, na.rm = T) * 100,
+    returned_100 = mean(returned_100, na.rm = T) * 100,
+  ) %>%
+  ggplot(aes(x = factor(year))) +
+  geom_bar(aes(y = returned_90, color = "Returned to 90%"),
+    position = "dodge", stat = "identity", fill = yellow
+  ) +
+  geom_label(
+    aes(
+      y = returned_90,
+      label = scales::percent(returned_90, accuracy = 0.1, scale = 1)
+    ),
+    nudge_y = -5, fill = "white"
+  ) +
+  geom_label(aes(y = returned_90, label = paste0(n_returned_90, "/", total)),
+    nudge_y = -11, fill = "white"
+  ) +
+  geom_bar(aes(y = returned_100, color = "Returned to 100%"),
+    position = "dodge", stat = "identity", fill = purple
+  ) +
+  geom_label(
+    aes(
+      y = returned_100,
+      label = scales::percent(returned_100, accuracy = 0.1, scale = 1)
+    ),
+    nudge_y = -10, fill = "white"
+  ) +
+  geom_label(aes(y = returned_100, label = paste0(n_returned_100, "/", total)),
+    nudge_y = -16, fill = "white"
+  ) +
+  ylab("Percent") +
+  xlab("Year") +
+  theme_classic() +
+  scale_color_manual(
+    name = element_blank(), values = c(purple, yellow),
+    labels = c("Returned to 100%", "Returned to 90%")
+  )
+```
+
+The above plot represents the proportion of admissions where the patient returned to close to baseline lung function, and was not aggregated by patient.
+In other words, multiple admissions for the same patient were considered independent in this case.
+
+# Comparisons by race
+
+Due to small numbers, Asian participants were included in the "Other" race category for the following analyses.
+Also, Hispanic ethnicity was considered its own racial category and superseded other racial groups for these analyses (i.e., if someone had both "Other" for race and "Hispanic or Latino" for ethnicity, they were considered Hispanic).
+
+```{r}
+# Make dataframe for race comparisons
+df$race <- as.character(df$race)
+df$race[df$ethnicity == "Hispanic or Latino"] <- "Hispanic or Latino"
+df$race <- factor(df$race,
+  levels = c(
+    "White", "Black or African-American", "Hispanic or Latino",
+    "Unknown/not reported", "Other", "Asian"
+  ),
+  labels = c(
+    "White", "Black or African-American", "Hispanic or Latino",
+    "Unknown/Other", "Unknown/Other", "Unknown/Other"
+  )
+)
+race_df <- df %>%
+  group_by(sid) %>%
+  filter(row_number() == 1) %>%
+  ungroup()
+```
+
+## CFTR modulator
+
+```{r}
+#| label: tbl-cftr-mod-overall
+#| tbl-cap: Any CFTR modulator prescription by race
+race_df %>%
+  select(race, cftr_mod) %>%
+  tbl_summary(by = race, label = labels, missing_text = "(Missing)") %>%
+  add_p() %>%
+  add_overall()
+```
+
+```{r}
+#| label: tbl-cftr-mod-year
+#| tbl-cap: Any CFTR modulator prescription by race and year
+race_df %>%
+  select(year, race, cftr_mod) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(
+          by = race, label = labels, missing_text = "(Missing)"
+        ) %>%
+        add_p() %>%
+        add_overall()
+  )
+```
+
+```{r}
+#| label: tbl-cftr-mod-specific-overall
+#| tbl-cap: CFTR modulator type by race
+race_df %>%
+  select(race, cftr_mod_spec) %>%
+  tbl_summary(by = race, label = labels, missing_text = "(Missing)") %>%
+  add_p(test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)) %>%
+  add_overall()
+```
+
+```{r}
+#| label: tbl-cftr-mod-specific-year
+#| tbl-cap: CFTR modulator type by race and year
+race_df %>%
+  select(race, cftr_mod_spec, year) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(
+          by = race, label = labels, missing_text = "(Missing)"
+        ) %>%
+        add_p(test.args = all_tests("fisher.test") ~
+          list(simulate.p.value = T)) %>%
+        add_overall()
+  )
+```
+
+```{r}
+#| label: tbl-eti-2022
+#| tbl-cap: ETI by race in 2022
+race_df %>%
+  filter(year == 2022) %>%
+  select(race, Trikafta) %>%
+  tbl_summary(by = race, label = labels, missing_text = "(Missing)") %>%
+  add_p(test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)) %>%
+  add_overall()
+```
+
+## Genotype
+
+```{r}
+#| label: tbl-genotype-overall
+#| tbl-cap: Genotype by race
+race_df %>%
+  select(race, Genotype) %>%
+  tbl_summary(by = race, label = labels, missing_text = "(Missing)") %>%
+  add_p() %>%
+  add_overall()
+```
+
+```{r}
+#| label: tbl-genotype-year
+#| tbl-cap: Genotype by race and year
+race_df %>%
+  select(year, race, Genotype) %>%
+  tbl_strata(
+    strata = year,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(
+          by = race, label = labels, missing_text = "(Missing)"
+        ) %>%
+        add_p() %>%
+        add_overall()
+  )
+```
+
+# FEV1 response (from admission to discharge) on ETI vs not on ETI
+
+Hospitalizations within a person were considered independent in the table below.
+
+```{r}
+#| label: tbl-fev1-change-eti
+#| tbl-cap: Change in FEV1pp by ETI status
+df %>%
+  mutate(ppfev1_change = ppfev1_discharge - ppfev1_admit) %>%
+  filter(year == 2022) %>%
+  select(ppfev1_change, Trikafta) %>%
+  tbl_summary(
+    by = Trikafta, label = labels, missing_text = "(Missing)",
+    statistic = list(all_continuous() ~ "{mean} ({sd})")
+  ) %>%
+  add_p(test = list(all_continuous() ~ "t.test")) %>%
+  add_overall()
+```

### 2024-08-30 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
added full tables to GCT analysis

@@
 df$Diagnosis[df$`OGTT 120 min` >= 200] <- "CFRD"
-df[which(rowSums(is.na(df[, grep("OGTT", colnames(df))])) == 3), "Diagnosis"] <- NA
+df[which(rowSums(is.na(df[, grep("OGTT", colnames(df))])) == 3), "Diagnosis"] <-
+  NA
 df$Diagnosis <- factor(df$Diagnosis,
@@
 )
+df$CFRD <- factor(df$`OGTT 120 min` >= 200,
+  levels = c(F, T),
+  labels = c("No", "Yes")
+)
 # Write for Christine to check
@@
   file = "./Christine Chan/CFF GCT Study/Data_Clean/gct_analysis_data.csv",
-          row.names = F,na = "")
-# Remove those without OGTT data
-no_ogtt <- df$ID[which(is.na(df$Diagnosis))]
-df <- df %>% filter(!ID %in% no_ogtt)
+  row.names = F, na = ""
+)
 ```
@@
 
-A total of `r length(no_ogtt)` participants were excluded because they had no OGTT data.
+```{r}
+#| label: table1a
+#| tbl-cap: Participant characteristics (including those missing OGTT data)
+#| message: false
+df %>%
+  select(-ID, -Dysglycemia) %>%
+  tbl_summary(missing_text = "(Missing)")
+# Now remove those without OGTT
+df <- df %>% filter(!is.na(Diagnosis))
+```
 
 ```{r}
-#| label: table1a
+#| label: table1b
 #| tbl-cap: Participant characteristics by diagnosis
@@
 ```{r}
-#| label: table1b
+#| label: table1c
 #| tbl-cap: Participant characteristics by dysglycemia
@@
   facet_wrap(~Diagnosis) +
+  theme_bw() +
   theme(legend.position = "none")
@@
 ```{r}
-#| label: fig-point
-#| fig-cap: Point plot of OGTT glucose values by diabetes status
+#| label: fig-dot
+#| fig-cap: Dot plot of OGTT glucose values by diabetes status
 #| warning: false
-point_p <- ggplot(
+dot_p <- ggplot(
   plot_df,
@@
   facet_wrap(~Diagnosis) +
+  theme_bw() +
   theme(legend.position = "none")
-point_p
+dot_p
 ```
@@
 
+## Dysglycemia
+
 Dysglycemia was defined as any diagnosis on OGTT other than normal glucose tolerance (NGT).
+
+### Youden indices
 
@@
 
+### ROC curves
+
 ```{r}
@@
 ```
+
+### Full cutoff tables
+
+#### GCTpl
+
+```{r}
+gt(data.frame(coords(gctpl_roc, "all", ret = "all"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+#### GCTcap
+
+```{r}
+gt(data.frame(coords(gctcap_roc, "all", ret = "all"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+#### OGTT 1 hour
+
+```{r}
+gt(data.frame(coords(ogtt1_roc, "all", ret = "all"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+#### RPG
+
+```{r}
+gt(data.frame(coords(rpg_roc, "all", ret = "all"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+#### RCG
+
+```{r}
+gt(data.frame(coords(rcg_roc, "all", ret = "all"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+#### HbA1c
+
+```{r}
+gt(data.frame(coords(a1c_roc, "all", ret = "all"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+### Predefined cutoffs
 
@@
 ```
+
+```{r}
+#| label: fig-dot-cutoffs
+#| fig-cap:  Dot plot of oral glucose challenge test plasma glucose (GCTpl) (mg/dl) by diabetes status
+#| warning: false
+dot_p <- ggplot(
+  df,
+  aes(x = Diagnosis, y = `GCT 60 min plasma`)
+) +
+  geom_point() +
+  geom_hline(yintercept = 140, linetype = "dashed") +
+  geom_hline(yintercept = 150, linetype = "dashed") +
+  geom_hline(yintercept = 160, linetype = "dashed") +
+  theme_bw()
+dot_p
+```
+
+## CFRD
+
+CFRD was defined as a two hour OGTT $\geq$ 200 mg/dL.
+
+### Youden indices
+
+```{r}
+#| label: tbl-roc-cfrd
+#| tbl-cap: Accuracy of different screening tests to predict CFRD
+#| message: false
+gctpl_roc <- roc(CFRD ~ `GCT 60 min plasma`, data = df)
+gctcap_roc <- roc(CFRD ~ `GCT 60 min capillary`, data = df)
+ogtt1_roc <- roc(CFRD ~ `OGTT 60 min`, data = df)
+rpg_roc <- roc(CFRD ~ `GCT 0 min plasma`, data = df)
+rcg_roc <- roc(CFRD ~ `GCT 0 min capillary`, data = df)
+a1c_roc <- roc(CFRD ~ HbA1c, data = df)
+roc_table <-
+  lapply(
+    list(gctpl_roc, gctcap_roc, ogtt1_roc, rpg_roc, rcg_roc, a1c_roc),
+    function(roc) {
+      coords(roc, "best", ret = "all")
+    }
+  )
+roc_table <- do.call(rbind, roc_table)
+rownames(roc_table) <- c("GCTpl", "GCTcap", "OGTT 1 hr", "RPG", "RCG", "A1c")
+gt(roc_table, rownames_to_stub = T) %>% fmt_number(decimals = 3)
+```
+
+### ROC curves
+
+```{r}
+invisible(lapply(
+  list(gctpl_roc, gctcap_roc, ogtt1_roc, rpg_roc, rcg_roc, a1c_roc),
+  function(roc) {
+    print(plot(roc, print.auc = T, print.thres = T))
+  }
+))
+```
+
+### Full cutoff tables
+
+#### GCTpl
+
+```{r}
+gt(data.frame(coords(gctpl_roc, "all", ret = "all"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+#### GCTcap
+
+```{r}
+gt(data.frame(coords(gctcap_roc, "all", ret = "all"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+#### OGTT 1 hour
+
+```{r}
+gt(data.frame(coords(ogtt1_roc, "all", ret = "all"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+#### RPG
+
+```{r}
+gt(data.frame(coords(rpg_roc, "all", ret = "all"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+#### RCG
+
+```{r}
+gt(data.frame(coords(rcg_roc, "all", ret = "all"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+#### HbA1c
+
+```{r}
+gt(data.frame(coords(a1c_roc, "all", ret = "all"))) %>%
+  fmt_number(n_sigfig = 3)
+```

### 2024-09-06 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.qmd
CGM pre/post Trikafta responses

--- 2024-05-06 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
@@
 author: "Tim Vigers"
-date: "`r format(Sys.time(), '%d %B %Y')`"
-output:
-  html_document:
-    number_sections: yes
-    toc: yes
-    toc_depth: 5
-    toc_float: yes
+date: "today"
+date-format: long
+format:
+  html:
+    toc: true
+    toc-depth: 5
+    toc-float: true
+    code-fold: true
+    self-contained: true
+    fig-cap-location: top
+    page-layout: full
+    theme:
+      light: flatly
+      dark: darkly
+bibliography: /home/tim/Documents/Miscellaneous/zotero.bib
+csl: /home/tim/GitHub/styles/american-medical-association.csl
+editor: source
 ---
@@
 ```{r setup, include=FALSE}
-library(tidyverse)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(ggpmisc)
+library(lubridate)
 library(gtsummary)
-library(arsenal)
 library(rspiro)
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood",
+  "Linux" = "/home/tim/OneDrive/Vigers/CF/Colleen Wood"
 )
@@
 
+```{r data cleaning}
+#| include: false
+load("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/analysis_dataset.rdata")
+# Ethnicity
+chco_demo <- read.csv("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Raw/Data From CHCO/Registry/Copy of Earliest Trikafta Dates in Registry_updated 2.2.21 cc modified_cfrdyn.csv")
+chco_demo$Hispanic <- NA
+chco_demo$Hispanic[chco_demo$Hispanic.Latinx] <- "Yes"
+chco_demo$Hispanic[!chco_demo$Hispanic.Latinx] <- "No"
+chco_demo <- chco_demo[, c("CFF.ID", "Hispanic")]
+
+montana_demo <- read.csv("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/montana_demo.csv")
+montana_demo$Hispanic <- montana_demo$Is.Patient.of.Hispanic.Origin
+montana_demo <- montana_demo[, c("CFF.ID", "Hispanic")]
+
+uw_demo <- read.csv("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/uw_demo.csv")
+uw_demo$Hispanic <- uw_demo$Is.Patient.of.Hispanic.Origin
+uw_demo <- uw_demo[, c("CFF.ID", "Hispanic")]
+
+eth <- do.call(rbind, list(chco_demo, montana_demo, uw_demo))
+colnames(eth) <- c("CFF ID", "Hispanic")
+
+df <- left_join(df, eth)
+```
+
 ```{r}
-load("./Colleen Wood/Prepost Trikafta/Data_Cleaned/analysis_dataset.rdata")
-# Baseline HbA1c, BMI group pre- and post-modulator
+# Baseline HbA1c, BMI group pre- and post-modulator, age group
 df <- df %>%
@@
     `Baseline HbA1c` = HbA1c[`Pre or Post` == "Pre"],
+    `Age group at baseline` = Age[`Pre or Post` == "Pre"],
     `Baseline BMI` = `BMI Percentile`[`Pre or Post` == "Pre"],
@@
   ungroup()
-df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf), right = F)
+df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf),
+  right = F
+)
+df$`Age group at baseline` <- cut(df$`Age group at baseline`,
+  c(-Inf, 18, Inf),
+  right = F
+)
 df$`Baseline BMI` <- cut(df$`Baseline BMI`, c(-Inf, 85, Inf), right = F)
 df$`Post BMI` <- cut(df$`Post BMI`, c(-Inf, 85, Inf), right = F)
-# Pre-modulator BMI percentile
-df <- df %>%
-  group_by(`CFF ID`) %>%
-  mutate(`Baseline HbA1c` = HbA1c[`Pre or Post` == "Pre"]) %>%
-  ungroup()
-df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf), right = F)
 # Race neutral equations
@@
 )
+# TDD per kg
+df$TDD <- df$TDD / df$Weight
 # Columns formats
@@
     ppFVCgl, `G-tube in past 12 months`, `Pancreatic Status`, Site,
-    Genotype, `Age Group`, Race
+    Genotype, `Age Group`, Race, Hispanic
   ) %>%
   tbl_summary(
-    statistic = list(c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ "{mean} ± {sd}"),
+    statistic = list(c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~
+      "{mean} ± {sd}"),
     missing_text = "Missing"
@@
           c("Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile") ~
+            c("{median} [{p25}, {p75}]")
+        ),
+        include = -`CFF ID`, digits = list(everything() ~ 2),
+        missing_text = "Missing"
+      ) %>%
+      add_p(
+        test = list(
+          c(
+            "Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile",
+            "TDD"
+          ) ~ "paired.wilcox.test",
+          c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ "paired.t.test"
+        ),
+        group = `CFF ID`
+      ) %>%
+      separate_p_footnotes()
+  )
+# Plot
+ggplot(df, aes(x = `Pre or Post`, y = HbA1c, group = `CFF ID`)) +
+  geom_point(alpha = 0.2) +
+  geom_line(alpha = 0.2) +
+  theme_classic() +
+  theme(axis.title.x = element_blank()) +
+  facet_wrap(~Site)
+```
+
+### By age group at baseline
+
+```{r}
+# Table
+df %>%
+  select(
+    `CFF ID`, `Pre or Post`, `Age group at baseline`, `Num. A1c Values`, HbA1c, Time,
+    Weight, `BMI Percentile`, ppFEV1, ppFEV1gl, ppFVC, ppFVCgl, TDD,
+  ) %>%
+  tbl_strata(
+    strata = `Age group at baseline`,
+    .tbl_fun = ~ .x %>%
+      tbl_summary(
+        by = `Pre or Post`,
+        type = list(
+          c(
+            "Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile",
+            "ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl", "TDD"
+          ) ~ "continuous2"
+        ),
+        statistic = list(
+          c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~
+            c("{mean} ± {sd}", "{min}, {max}"),
+          c("Num. A1c Values", "HbA1c", "Time", "Weight", "BMI Percentile") ~
             c("{median} [{p25}, {p75}]", "{min}, {max}")
@@
   theme(axis.title.x = element_blank()) +
-  facet_wrap(~Site)
+  facet_wrap(~`Age group at baseline`)
 ```
@@
     statistic = list(
-      everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
+      everything() ~ c("{median} [{p25}, {p75}]")
     ),
@@
   separate_p_footnotes()
+```
+
+#### By age group at baseline
+
+```{r}
+cgm %>%
+  select(
+    `CFF ID`, `Pre or Post`, `Age group at baseline`, all_of(cgm_vars)
+  ) %>%
+  tbl_strata(
+    strata = `Age group at baseline`,
+    .tbl_fun = ~ .x %>%
+      tbl_summary(
+        by = `Pre or Post`,
+        type = list(everything() ~ "continuous2"),
+        statistic = list(
+          everything() ~ c("{median} [{p25}, {p75}]")
+        ),
+        include = -`CFF ID`, digits = list(everything() ~ 2),
+        missing_text = "Missing"
+      ) %>%
+      add_p(
+        test = list(everything() ~ "paired.wilcox.test"),
+        group = `CFF ID`
+      ) %>%
+      separate_p_footnotes()
+  )
 ```
@@
         statistic = list(
-          everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
+          everything() ~ c("{median} [{p25}, {p75}]")
         ),
@@
   separate_p_footnotes()
+```
+
+#### By age group at baseline
+
+```{r}
+cgm %>%
+  select(
+    `CFF ID`, `Pre or Post`, `Age group at baseline`, all_of(day_cgm)
+  ) %>%
+  tbl_strata(
+    strata = `Age group at baseline`,
+    .tbl_fun = ~ .x %>%
+      tbl_summary(
+        by = `Pre or Post`,
+        type = list(everything() ~ "continuous2"),
+        statistic = list(
+          everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
+        ),
+        include = -`CFF ID`, digits = list(everything() ~ 2),
+        missing_text = "Missing"
+      ) %>%
+      add_p(
+        test = list(everything() ~ "paired.wilcox.test"),
+        group = `CFF ID`
+      ) %>%
+      separate_p_footnotes()
+  )
 ```
@@
 
+#### By age group at baseline
+
+```{r}
+cgm %>%
+  select(
+    `CFF ID`, `Pre or Post`, `Age group at baseline`, all_of(night_cgm)
+  ) %>%
+  tbl_strata(
+    strata = `Age group at baseline`,
+    .tbl_fun = ~ .x %>%
+      tbl_summary(
+        by = `Pre or Post`,
+        type = list(everything() ~ "continuous2"),
+        statistic = list(
+          everything() ~ c("{median} [{p25}, {p75}]", "{min}, {max}")
+        ),
+        include = -`CFF ID`, digits = list(everything() ~ 2),
+        missing_text = "Missing"
+      ) %>%
+      add_p(
+        test = list(everything() ~ "paired.wilcox.test"),
+        group = `CFF ID`
+      ) %>%
+      separate_p_footnotes()
+  )
+```
+
 #### By baseline HbA1c
@@
 ```
+
+# Additional analyses
+
+## Longitudinal HbA1c
+
+The vast majority of participants have 2 A1c values pre-ETI and 2 after.
+
+```{r}
+df %>%
+  mutate(`Num. A1c Values` = factor(`Num. A1c Values`)) %>%
+  select(`Pre or Post`, `Num. A1c Values`) %>%
+  tbl_summary(by = `Pre or Post`)
+```
+
+## Change in HbA1c and change in FEV1
+
+```{r}
+change <- df %>%
+  group_by(`CFF ID`) %>%
+  summarise(
+    `Baseline HbA1c` = HbA1c[`Pre or Post` == "Pre"],
+    `Baseline ppFEV1gl` = ppFEV1gl[`Pre or Post` == "Pre"],
+    `Post ppFEV1gl` = ppFEV1gl[`Pre or Post` == "Post"],
+    `Change in HbA1c` =
+      HbA1c[`Pre or Post` == "Post"] - HbA1c[`Pre or Post` == "Pre"],
+    `Change in FEV1` =
+      ppFEV1gl[`Pre or Post` == "Post"] - ppFEV1gl[`Pre or Post` == "Pre"],
+    `% Change in FEV1` =
+      ((ppFEV1gl[`Pre or Post` == "Post"] - ppFEV1gl[`Pre or Post` == "Pre"]) /
+        ppFEV1gl[`Pre or Post` == "Pre"]) * 100
+  )
+change$`HbA1c group` <- cut(change$`Change in HbA1c`,
+  breaks = c(-Inf, -0.5, 0.5, Inf), right = F,
+  labels = c("Decreased", "No change", "Increased")
+)
+ggplot(change, aes(x = `Change in FEV1`, y = `Change in HbA1c`)) +
+  geom_point() +
+  stat_poly_line(se = F) +
+  stat_poly_eq(use_label(c("eq", "R2", "p", "n"))) +
+  theme_bw()
+ggplot(change, aes(x = `% Change in FEV1`, y = `Change in HbA1c`)) +
+  geom_point() +
+  stat_poly_line(se = F) +
+  stat_poly_eq(use_label(c("eq", "R2", "p", "n"))) +
+  theme_bw()
+```
+
+## Change in FEV1 by HbA1c group
+
+```{r}
+df <- df %>%
+  group_by(`CFF ID`) %>%
+  mutate(
+    `Change in HbA1c` =
+      HbA1c[`Pre or Post` == "Post"] - HbA1c[`Pre or Post` == "Pre"]
+  )
+df$`HbA1c group` <- cut(df$`Change in HbA1c`,
+  breaks = c(-Inf, -0.5, 0.5, Inf), right = F,
+  labels = c("HbA1c decreased", "HbA1 no change", "HbA1c increased")
+)
+ggplot(df, aes(x = `Pre or Post`, y = ppFEV1gl, group = `CFF ID`)) +
+  geom_point(alpha = 0.2) +
+  geom_line(alpha = 0.2) +
+  theme_classic() +
+  theme(axis.title.x = element_blank()) +
+  facet_wrap(~`HbA1c group`)
+```

### 2024-09-06 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
CGM pre/post Trikafta responses

--- 2024-08-23 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
@@
 #| include: false
-library(tidyverse)
+library(dplyr)
+library(tidyr)
 library(readxl)
+library(gtsummary)
 home_dir <- switch(Sys.info()["sysname"],
@@
 ```{r data cleaning}
+# Import data
+chco <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv",
+  na.strings = ""
+)
+njh <- read_excel("./Data_Raw/SweatVariability2025_CFgenotypes 2024-08-21.xlsx")
+chec <- read.csv("./Data_Raw/CHEC sweat sodium and chloride.csv")
+# Numeric variables
+num_vars <- c(
+  "Na", "K", "Cl", "Amount_Left", "Amount_Right", "Na_Left", "Na_Right",
+  "K_Left", "K_Right", "Cl_Left", "Cl_Right"
+)
+chco[, num_vars] <- lapply(chco[, num_vars], function(c) {
+  suppressWarnings(as.numeric(c))
+})
+# Fix diagnoses/exclusion data for CHCO (based on Angela's review)
+diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
+exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) | diagnoses$Code == "Exclude"]
+exclude <- exclude[!is.na(exclude)]
+chco <- chco[!(chco$PatientID %in% exclude), ]
+chco$Diagnosis <- factor(
+  diagnoses$Diagnosis[match(
+    chco$PatientID,
+    diagnoses$PatientID
+  )],
+  levels = c("CFTR-related disorder", "CRMS", "Cystic Fibrosis", "Non-CF"),
+  labels = c("CFTR-RD or CRMS", "CFTR-RD or CRMS", "Cystic Fibrosis", "Non-CF")
+)
+```
 
+```{r}
+chco %>%
+  group_by(PatientID) %>%
+  filter(row_number() == 1) %>%
+  ungroup() %>%
+  select(-PatientID, -Genotypes1, -Genotypes2, -T_Tract, -Notes) %>%
+  tbl_summary(by = Diagnosis, missing_text = "(Missing)")%>%
+  add_overall()
 ```

### 2024-09-06 Colleen Wood/Pre-post Trikafta/clean_and_merge.R

@@ -1,17 +1,20 @@
-library(tidyverse)
+library(dplyr)
+library(tidyr)
+library(readr)
+library(purrr)
 library(lubridate)
 library(childsds)
 library(rspiro)
-source("/home/timvigers/GitHub/CF/genotype_class_severity.R")
+source("/home/tim/GitHub/CF/genotype_class_severity.R")
 t1_variables <- c(
   "CFF ID", "Site", "Sex", "Race", "Genotypes1", "Genotypes2", "Pancreatic Status",
   "G-tube in past 12 months", "Time", "Age",
   "BMI Percentile", "Height", "Weight", "FEV1", "FVC", "HbA1c"
 )
 # CHCO
-chco_a1c <- read_csv("/home/timvigers/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/chco_glycemic.csv", show_col_types = F)
+chco_a1c <- read_csv("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/chco_glycemic.csv", show_col_types = F)
 chco_a1c$Date <- parse_date_time(chco_a1c$Date, orders = "mdy")
-chco_bmi <- read_csv("/home/timvigers/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/chco_bmi_pft.csv", show_col_types = F)
+chco_bmi <- read_csv("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/chco_bmi_pft.csv", show_col_types = F)
 chco_bmi$Date <- parse_date_time(chco_bmi$Date, orders = "mdy")
 chco <- full_join(chco_a1c, chco_bmi)
 # Fill CFF ID by MRN
@@ -22,13 +25,13 @@ chco <- chco %>%
   filter(!is.na(`CFF ID`)) %>%
   select(-MRN)
 # Add demographics, filter
-chco_reg <- read_csv("/home/timvigers/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/chco_registry.csv", show_col_types = F)
[…]

### 2024-09-11 Christine Chan/EnVision CF/catecholamines_analysis.qmd
updates for Amir

@@
 #| include: false
-library(tidyverse)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
 library(gt)
@@
 )
-# For this analysis, we only want those with catecholamines
-df <- df %>%
-  group_by(study_id, Date) %>%
-  mutate(catecholamines = any(!is.na(Epinephrine) | !is.na(Norepinephrine))) %>%
-  filter(catecholamines) %>%
-  ungroup()
-# # Time as a factor
-df$Timepoint <- factor(df$Timepoint, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
+# Timepoint and variables lists
+timepoints <- c(0, 10, 30, 60, 90, 120, 150, 180)
+glucoses <- paste0("Glucose_", timepoints)
+insulins <- paste0("Insulin_", timepoints)
+epinephrines <- paste0("Epinephrine_", timepoints)
+norepinephrines <- paste0("Norepinephrine_", timepoints)
+# For this analysis, we only want those with catecholamines measured
+df$catecholamines <- apply(
+  df[, c(epinephrines, norepinephrines)], 1,
+  function(r) {
+    any(!is.na(r))
+  }
+)
+df <- df %>% filter(catecholamines)
+# To long format
+t <- df %>%
+  select(
+    study_id, Date,
+    all_of(c(glucoses, insulins, epinephrines, norepinephrines))
+  ) %>%
+  pivot_longer(all_of(c(glucoses, insulins, epinephrines, norepinephrines)),
+               names_to = c(".value", "Timepoint"),
+    names_pattern = "(.*)_(.*)")
 ```

### 2024-09-12 Christine Chan/EnVision CF/catecholamines_analysis.qmd
catecholamines analysis updates

--- 2024-08-02 Christine Chan/EnVision CF/catecholamines_analysis.qmd
@@
 
-- Epinephrine and norepinephrine were log transformed for modeling purposes.
+- Epinephrine, norepinephrine, and insulin were log transformed for modeling purposes.
 
@@
 )
-# For this analysis, we only want those with catecholamines
-df <- df %>%
-  group_by(study_id, Date) %>%
-  mutate(catecholamines = any(!is.na(Epinephrine) | !is.na(Norepinephrine))) %>%
-  filter(catecholamines) %>%
-  ungroup()
-# # Time as a factor
-df$Timepoint <- factor(df$Timepoint, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
+# Timepoint and variables lists
+timepoints <- c(0, 10, 30, 60, 90, 120, 150, 180)
+limited_timepoints <- c(0, 120, 150, 180)
+glucoses <- paste0("Glucose_", timepoints)
+insulins <- paste0("Insulin_", timepoints)
+epinephrines <- paste0("Epinephrine_", timepoints)
+norepinephrines <- paste0("Norepinephrine_", timepoints)
+# For this analysis, we only want those with catecholamines measured
+df$catecholamines <- apply(
+  df[, c(epinephrines, norepinephrines)], 1,
+  function(r) {
+    any(!is.na(r))
+  }
+)
+df <- df %>% filter(catecholamines)
+# Make a long dataset for mixed models
+lmm_df <- df %>%
+  select(
+    study_id, Hypo60, Hypo70,
+    all_of(c(glucoses, insulins, epinephrines, norepinephrines))
+  ) %>%
+  pivot_longer(all_of(c(glucoses, insulins, epinephrines, norepinephrines)),
+    names_to = c(".value", "Timepoint"),
+    names_pattern = "(.*)_(.*)"
+  )
 ```
@@
 df %>%
-  group_by(study_id) %>%
-  filter(row_number() == 1) %>%
-  ungroup() %>%
   select(
@@
 df %>%
-  filter(Timepoint %in% c(0, 120, 150, 180)) %>%
-  pivot_wider(
-    names_from = Timepoint, values_from = c(Epinephrine, Norepinephrine),
-    id_cols = c(study_id:cftr_mutation_2, Hypo70)
+  select(
+    Hypo70, age_visit:cftr_mutation_2, hypoglycemia_symptoms_yesno,
+    all_of(paste0("Glucose_", timepoints)),
+    all_of(paste0("Insulin_", timepoints)),
+    all_of(paste0("Epinephrine_", limited_timepoints)),
+    all_of(paste0("Norepinephrine_", limited_timepoints))
   ) %>%
-  select(Hypo70, age_visit:cftr_mutation_2, Epinephrine_0:Norepinephrine_180) %>%
   tbl_summary(by = Hypo70, missing_text = "Missing") %>%
@@
   separate_p_footnotes()
+```
+
+## Glucose
+
+Glucose was modeled using a linear mixed effect model with random intercept for participant and time treated as a categorical variable.
+
+```{r}
+#| warning: false
+# Fit
+mod <- lmer(Glucose ~ Timepoint * Hypo70 + (1 | study_id),
+  data = lmm_df
+)
+mod %>% tbl_regression(intercept = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Means") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  ylab("LS Mean Glucose") +
+  xlab("Timepoint") +
+  theme_bw() +
+  theme(legend.title = element_blank())
+```
+
+## Insulin
+
+Insulin was modeled using a linear mixed effect model with random intercept for participant and time treated as a categorical variable.
+
+```{r}
+#| warning: false
+# Fit
+mod <- lmer(log(Insulin) ~ Timepoint * Hypo70 + (1 | study_id),
+  data = lmm_df
+)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Means") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  ylab("LS Mean Insulin") +
+  xlab("Timepoint") +
+  theme_bw() +
+  theme(legend.title = element_blank())
 ```
@@
 # Fit
-mod <- lmer(log(Epinephrine) ~ Timepoint * Hypo70 + (1 | study_id), data = df)
+mod <- lmer(log(Epinephrine) ~ Timepoint * Hypo70 + (1 | study_id),
+  data = lmm_df
+)
 mod %>% tbl_regression(intercept = T, exponentiate = T)
 means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"), type = "response"))
-means %>% gt(caption = "LS Means")
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Means") %>%
+  fmt_number(decimals = 2)
 # Plot
-emmip(mod, Hypo70 ~ Timepoint, CIs = T, type = "response") +
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Mean Epinephrine") +
+  xlab("Timepoint") +
   theme_bw() +
@@
 # Fit
-mod <- lmer(log(Norepinephrine) ~ Timepoint * Hypo70 + (1 | study_id), data = df)
+mod <- lmer(log(Norepinephrine) ~ Timepoint * Hypo70 + (1 | study_id),
+  data = lmm_df
+)
 mod %>% tbl_regression(intercept = T, exponentiate = T)
 means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"), type = "response"))
-means %>% gt(caption = "LS Means")
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Means") %>%
+  fmt_number(decimals = 2)
 # Plot
-emmip(mod, Hypo70 ~ Timepoint, CIs = T, type = "response") +
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Mean Norepinephrine") +
+  xlab("Timepoint") +
   theme_bw() +
@@
 df %>%
-  filter(Timepoint %in% c(0, 120, 150, 180)) %>%
-  pivot_wider(
-    names_from = Timepoint, values_from = c(Epinephrine, Norepinephrine),
-    id_cols = c(study_id:cftr_mutation_2, Hypo60)
+  select(
+    Hypo60, age_visit:cftr_mutation_2, hypoglycemia_symptoms_yesno,
+    all_of(paste0("Glucose_", timepoints)),
+    all_of(paste0("Insulin_", timepoints)),
+    all_of(paste0("Epinephrine_", limited_timepoints)),
+    all_of(paste0("Norepinephrine_", limited_timepoints))
   ) %>%
-  select(Hypo60, age_visit:cftr_mutation_2, Epinephrine_0:Norepinephrine_180) %>%
   tbl_summary(by = Hypo60, missing_text = "Missing") %>%
@@
   separate_p_footnotes()
+```
+
+## Glucose
+
+Glucose was modeled using a linear mixed effect model with random intercept for participant and time treated as a categorical variable.
+
+```{r}
+#| warning: false
+# Fit
+mod <- lmer(Glucose ~ Timepoint * Hypo60 + (1 | study_id),
+  data = lmm_df
+)
+mod %>% tbl_regression(intercept = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Means") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  ylab("LS Mean Glucose") +
+  xlab("Timepoint") +
+  theme_bw() +
+  theme(legend.title = element_blank())
+```
+
+## Insulin
+
+Insulin was modeled using a linear mixed effect model with random intercept for participant and time treated as a categorical variable.
+
+```{r}
+#| warning: false
+# Fit
+mod <- lmer(log(Insulin) ~ Timepoint * Hypo60 + (1 | study_id),
+  data = lmm_df
+)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Means") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  ylab("LS Mean Insulin") +
+  xlab("Timepoint") +
+  theme_bw() +
+  theme(legend.title = element_blank())
 ```
@@
 # Fit
-mod <- lmer(log(Epinephrine) ~ Timepoint * Hypo60 + (1 | study_id), data = df)
+mod <- lmer(log(Epinephrine) ~ Timepoint * Hypo60 + (1 | study_id),
+  data = lmm_df
+)
 mod %>% tbl_regression(intercept = T, exponentiate = T)
 means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"), type = "response"))
-means %>% gt(caption = "LS Means")
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Means") %>%
+  fmt_number(decimals = 2)
 # Plot
-emmip(mod, Hypo60 ~ Timepoint, CIs = T, type = "response") +
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Mean Epinephrine") +
+  xlab("Timepoint") +
   theme_bw() +
@@
 # Fit
-mod <- lmer(log(Norepinephrine) ~ Timepoint * Hypo60 + (1 | study_id), data = df)
+mod <- lmer(log(Norepinephrine) ~ Timepoint * Hypo60 + (1 | study_id),
+  data = lmm_df
+)
 mod %>% tbl_regression(intercept = T, exponentiate = T)
 means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"), type = "response"))
-means %>% gt(caption = "LS Means")
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Means") %>%
+  fmt_number(decimals = 2)
 # Plot
-emmip(mod, Hypo60 ~ Timepoint, CIs = T, type = "response") +
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Mean Norepinephrine") +
+  xlab("Timepoint") +
   theme_bw() +

### 2024-09-12 Christine Chan/EnVision CF/clean_envision_data.R

@@ -2,8 +2,7 @@
 # Setup
 #-------------------------------------------------------------------------------
 # Libraries
-library(dplyr)
-library(tidyr)
+library(tidyverse)
 library(lubridate)
 library(readxl)
 library(stringi)
@@ -83,17 +82,18 @@ glucose$Timepoint <- sub("_min", "", glucose$Timepoint)
 glucose$Timepoint <- as.numeric(glucose$Timepoint)
 glucose$Date <- ymd(glucose$Date)
 #-------------------------------------------------------------------------------
-# Demographics
+# Demographics and visit data
 #-------------------------------------------------------------------------------
-demo <- redcap %>%
+visits <- redcap %>%
   select(
     study_id, date_visit, age_visit, height, weight, sex, origin_race,
-    ethnicity, redcap_data_access_group, cftr_mutation_1, cftr_mutation_2
+    ethnicity, redcap_data_access_group, cftr_mutation_1, cftr_mutation_2,
+    hypoglycemia_symptoms_yesno
   ) %>%
   filter(!is.na(date_visit)) %>%
   rename(Date = date_visit) %>%
   filter(!if_all(age_visit:ethnicity, is.na))
-demo$Date <- ymd(demo$Date)
+visits$Date <- ymd(visits$Date)
 #-------------------------------------------------------------------------------
 # Date fixes per Holly's 7/19 emails
 #-------------------------------------------------------------------------------
@@ -187,7 +187,7 @@ insulin$Insulin <- as.numeric(insulin$Insulin)
 cgm <- read.csv("./Christine Chan/EnVision CF/Data_Clean/CGM/cgm_variables.csv",
   na.strings = ""
 )
-cgm = cgm[-which(cgm$percent_cgm_wear==0),]
+cgm <- cgm[-which(cgm$percent_cgm_wear == 0), ]
 cgm$subject_id <- sub("_sensor_raw_data_upload", "", cgm$subject_id)
 cgm$study_id <- sapply(
   stri_split_fixed(str = cgm$subject_id, pattern = "_", n = 2),
@@ -203,27 +203,18 @@ cgm <- left_join(cgm, redcap %>% select(study_id, redcap_event_name, date_visit)
 cgm <- cgm %>% rename(Date = date_visit)
[…]

### 2024-09-13 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
swaet test data cleaned

--- 2024-08-23 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
@@
 ```{r data cleaning}
+# Import data
+chco <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv",
+  na.strings = ""
+)
+# Numeric variables
+num_vars <- c(
+  "Na", "K", "Cl", "Amount_Left", "Amount_Right", "Na_Left", "Na_Right",
+  "K_Left", "K_Right", "Cl_Left", "Cl_Right"
+)
+chco[, num_vars] <- lapply(chco[, num_vars], function(c) {
+  suppressWarnings(as.numeric(c))
+})
+# Fix diagnoses/exclusion data for CHCO (based on Angela's review)
+diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
+exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) | diagnoses$Code == "Exclude"]
+exclude <- exclude[!is.na(exclude)]
+chco <- chco[!(chco$PatientID %in% exclude), ]
+chco$Diagnosis <- factor(
+  diagnoses$Diagnosis[match(
+    chco$PatientID,
+    diagnoses$PatientID
+  )],
+  levels = c("CFTR-related disorder", "CRMS", "Cystic Fibrosis", "Non-CF"),
+  labels = c("CFTR-RD or CRMS", "CFTR-RD or CRMS", "Cystic Fibrosis", "Non-CF")
+)
+# Exclude those without demographics (not confident in diagnosis per Edith)
+chco <- chco %>% filter(Demographics.Available == "TRUE")
+# Limit to the columns that we need for analysis
+chco <- chco %>% select(
+  PatientID, Diagnosis, Sex, Genotypes1, Genotypes2,
+  Age.at.Test.in.Years, Method, Site, Na, K, Cl, Na_Left, Na_Right, K_Left,
+  K_Right, Cl_Left, Cl_Right
+)
+# Get each person's first test
+chco <- chco %>%
+  group_by(PatientID) %>%
+  slice_min(Age.at.Test.in.Years, na_rm = T)
+# Import NJH data
+njh <- read_excel("./Data_Raw/SweatVariability2025_CFgenotypes.xlsx")
+# Filter and select similar to CHCO
+njh <- njh %>%
+  filter(`Demographics Available`) %>%
+  select(
+    `Patient ID`, Diagnosis, Sex, `Genotype 1`, `Genotype 2`,
+    `Age at test (in years)`, `Test Method`, `Site of test`, `Sweat sodium`,
+    `Sweat potassium`, `Sweat chloride`, `Sweat sodium left arm`,
+    `Sweat sodium right arm`, `Sweat potassium left arm`,
+    `Sweat potassium right arm`, `Sweat chloride left arm`,
+    `Sweat chloride right arm`
+  )
+colnames(njh) <- colnames(chco)
+njh <- njh %>%
+  group_by(PatientID) %>%
+  slice_min(Age.at.Test.in.Years, na_rm = T)
+# Import detailed diagnoses from Katie
+diagnoses <- read.csv("./Data_Cleaned/SweatVariability2025_CFgenotypes.csv",
+  na.strings = c("", ".")
+)
+# Merge NJH and diagnoses
+diagnoses <- diagnoses %>% select(Patient.ID, Diagnosis...DataScout)
+colnames(diagnoses) <- c("PatientID", "Diagnosis")
+njh <- full_join(njh, diagnoses)
+# Import CHEC data
+chec <- read.csv("./Data_Raw/CHEC sweat sodium and chloride.csv")
+# Filter and select as above (need to add some empty columns and split genotype)
+chec$Genotypes1 <- trimws(sapply(strsplit(chec$gntyp_pair, "/"), "[[", 1))
+chec$Genotypes2 <- trimws(sapply(strsplit(chec$gntyp_pair, "/"), "[[", 2))
+chec$Diagnosis <- "CF"
+chec[, c(
+  "Method", "K", "Na_Left", "Na_Right", "K_Left", "K_Right", "Cl_Left",
+  "Cl_Right"
+)] <- NA
+chec <- chec %>%
+  select(
+    subject, Diagnosis, sex, Genotypes1, Genotypes2, age, Method, site,
+    swsavg, K, swcavg, Na_Left, Na_Right, K_Left, K_Right, Cl_Left, Cl_Right
+  )
+colnames(chec) = colnames(chco)
+# Put everything together
+df = full_join(chco,njh)
+df$PatientID = as.character(df$PatientID)
+df = full_join(df,chec)
+```
 
+```{r}
 ```

### 2024-09-13 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
Update sweat_test_analysis.qmd

--- 2024-08-23 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
@@
 library(readxl)
+library(gtsummary)
 home_dir <- switch(Sys.info()["sysname"],
@@
 )
-njh <- read_excel("./Data_Raw/SweatVariability2025_CFgenotypes 2024-08-21.xlsx")
-chec <- read.csv("./Data_Raw/CHEC sweat sodium and chloride.csv")
+# Numeric variables
+num_vars <- c(
+  "Na", "K", "Cl", "Amount_Left", "Amount_Right", "Na_Left", "Na_Right",
+  "K_Left", "K_Right", "Cl_Left", "Cl_Right"
+)
+chco[, num_vars] <- lapply(chco[, num_vars], function(c) {
+  suppressWarnings(as.numeric(c))
+})
 # Fix diagnoses/exclusion data for CHCO (based on Angela's review)
 diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
-exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) | diagnoses$Code == "Exclude"]
+exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) |
+  diagnoses$Code == "Exclude"]
 exclude <- exclude[!is.na(exclude)]
@@
 )
+# Exclude those without demographics (not confident in diagnosis per Edith)
+chco <- chco %>% filter(Demographics.Available == "TRUE")
+# Limit to the columns that we need for analysis
+chco <- chco %>% select(
+  PatientID, Diagnosis, Sex, Genotypes1, Genotypes2,
+  Age.at.Test.in.Years, Method, Site, Na, K, Cl, Na_Left, Na_Right, K_Left,
+  K_Right, Cl_Left, Cl_Right
+)
+# Get each person's first test
+chco <- chco %>%
+  group_by(PatientID) %>%
+  slice_min(Age.at.Test.in.Years, na_rm = T)
+# Import NJH data
+njh <- read_excel("./Data_Raw/SweatVariability2025_CFgenotypes.xlsx")
+# Filter and select similar to CHCO
+njh <- njh %>%
+  filter(`Demographics Available`) %>%
+  select(
+    `Patient ID`, Diagnosis, Sex, `Genotype 1`, `Genotype 2`,
+    `Age at test (in years)`, `Test Method`, `Site of test`, `Sweat sodium`,
+    `Sweat potassium`, `Sweat chloride`, `Sweat sodium left arm`,
+    `Sweat sodium right arm`, `Sweat potassium left arm`,
+    `Sweat potassium right arm`, `Sweat chloride left arm`,
+    `Sweat chloride right arm`
+  )
+colnames(njh) <- colnames(chco)
+njh <- njh %>%
+  group_by(PatientID) %>%
+  slice_min(Age.at.Test.in.Years, na_rm = T)
+# Import detailed diagnoses from Katie
+diagnoses <- read.csv("./Data_Cleaned/SweatVariability2025_CFgenotypes.csv",
+  na.strings = c("", ".")
+)
+# Merge NJH and diagnoses
+diagnoses <- diagnoses %>% select(Patient.ID, Diagnosis...DataScout)
+colnames(diagnoses) <- c("PatientID", "Diagnosis")
+njh <- full_join(njh, diagnoses)
+# Import CHEC data
+chec <- read.csv("./Data_Raw/CHEC sweat sodium and chloride.csv")
+# Filter and select as above (need to add some empty columns and split genotype)
+chec$Genotypes1 <- trimws(sapply(strsplit(chec$gntyp_pair, "/"), "[[", 1))
+chec$Genotypes2 <- trimws(sapply(strsplit(chec$gntyp_pair, "/"), "[[", 2))
+chec$Diagnosis <- "CF"
+chec[, c(
+  "Method", "K", "Na_Left", "Na_Right", "K_Left", "K_Right", "Cl_Left",
+  "Cl_Right"
+)] <- NA
+chec <- chec %>%
+  select(
+    subject, Diagnosis, sex, Genotypes1, Genotypes2, age, Method, site,
+    swsavg, K, swcavg, Na_Left, Na_Right, K_Left, K_Right, Cl_Left, Cl_Right
+  )
+colnames(chec) <- colnames(chco)
+# Put everything together
+df <- full_join(chco, njh)
+df$PatientID <- as.character(df$PatientID)
+df <- full_join(df, chec)
+# Average the left and right values
+df$NaAvg <- rowMeans(df[, c("Na_Left", "Na_Right")], na.rm = T)
+df$NaAvg[is.nan(df$NaAvg)] <- NA
+df$KAvg <- rowMeans(df[, c("K_Left", "K_Right")], na.rm = T)
+df$KAvg[is.nan(df$KAvg)] <- NA
+df$ClAvg <- rowMeans(df[, c("Cl_Left", "Cl_Right")], na.rm = T)
+df$ClAvg[is.nan(df$ClAvg)] <- NA
+df <- df %>%
+  unite(Na, c("Na", "NaAvg"), na.rm = T) %>%
+  unite(K, c("K", "KAvg"), na.rm = T) %>%
+  unite(Cl, c("Cl", "ClAvg"), na.rm = T)
+# Genotype function
+genotype_function <- read.csv("./Data_Cleaned/genotype_function.csv")
+# Get single genes and add them together
+genotype_function <-
+  genotype_function[grep("NULL", genotype_function$genotype), ]
+genotype_function$genotype <- sub("/NULL", "", genotype_function$genotype)
+df$Genotype1Function <-
+  genotype_function$percent_wildtype_cftr_function[match(
+    df$Genotypes1,
+    genotype_function$genotype
+  )]
+df$Genotype2Function <-
+  genotype_function$percent_wildtype_cftr_function[match(
+    df$Genotypes2,
+    genotype_function$genotype
+  )]
+df$TotalGenotypeFunction <- rowSums(
+  df[, c("Genotype1Function", "Genotype2Function")],
+  na.rm = T
+)
+df$TotalGenotypeFunction[df$TotalGenotypeFunction == 0] <- NA
+# Numeric variables
+df[, c("Na", "K", "Cl")] <- lapply(df[, c("Na", "K", "Cl")], as.numeric)
+# As data frame
+df <- as.data.frame(df)
+# Write for checking
+write.csv(df, "./Data_Cleaned/analysis_dataset.csv", row.names = F, na = "")
 ```
+
+# Excluding those with no genotype function estimate
+
+```{r}
+df %>%
+  filter(!is.na(TotalGenotypeFunction)) %>%
+  select(
+    Diagnosis, Genotypes1, Genotypes2, Genotype1Function, Genotype2Function,
+    TotalGenotypeFunction, Na, K, Cl
+  ) %>%
+  droplevels(.) %>%
+  tbl_summary(by = Diagnosis)
+```
+
+```{r}
+df %>%
+  filter(!is.na(TotalGenotypeFunction)) %>%
+  ggplot(
+    aes(x = Cl, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
+  ) +
+  geom_point() +
+  geom_smooth(method = "lm", se = F) +
+  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())
+
+df %>%
+  filter(!is.na(TotalGenotypeFunction)) %>%
+  ggplot(
+    aes(x = Cl, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
+  ) +
+  geom_point() +
+  geom_smooth(method = "lm", se = F) +
+  ylim(c(0, 20)) +
+  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())
+```

### 2024-09-13 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
Update sweat_test_analysis.qmd

@@
 ```{r data cleaning}
+#| message: false
 # Import data
@@
 ```{r}
+#| message: false
 df %>%
@@
 ```{r}
+#| message: false
+#| warning: false
 df %>%

### 2024-09-16 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.qmd
Hba1c longitudinal

@@
       dark: darkly
-bibliography: /home/tim/Documents/Miscellaneous/zotero.bib
-csl: /home/tim/GitHub/styles/american-medical-association.csl
+bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
+csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
 editor: source
@@
 library(gtsummary)
+library(knitr)
 library(rspiro)
+library(lmerTest)
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood",
-  "Linux" = "/home/tim/OneDrive/Vigers/CF/Colleen Wood"
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood"
 )
@@
 #| include: false
-load("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/analysis_dataset.rdata")
+load("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/analysis_dataset.rdata")
+load("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/hba1c_dataset.rdata")
 # Ethnicity
-chco_demo <- read.csv("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Raw/Data From CHCO/Registry/Copy of Earliest Trikafta Dates in Registry_updated 2.2.21 cc modified_cfrdyn.csv")
+chco_demo <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Raw/Data From CHCO/Registry/Copy of Earliest Trikafta Dates in Registry_updated 2.2.21 cc modified_cfrdyn.csv")
 chco_demo$Hispanic <- NA
@@
 chco_demo <- chco_demo[, c("CFF.ID", "Hispanic")]
-
-montana_demo <- read.csv("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/montana_demo.csv")
+montana_demo <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/montana_demo.csv")
 montana_demo$Hispanic <- montana_demo$Is.Patient.of.Hispanic.Origin
 montana_demo <- montana_demo[, c("CFF.ID", "Hispanic")]
-
-uw_demo <- read.csv("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/uw_demo.csv")
+uw_demo <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/uw_demo.csv")
 uw_demo$Hispanic <- uw_demo$Is.Patient.of.Hispanic.Origin
 uw_demo <- uw_demo[, c("CFF.ID", "Hispanic")]
-
 eth <- do.call(rbind, list(chco_demo, montana_demo, uw_demo))
 colnames(eth) <- c("CFF ID", "Hispanic")
-
 df <- left_join(df, eth)
@@
 df$`Num. A1c Values` <- as.numeric(df$`Num. A1c Values`)
+```
+
+# HbA1c longitudinal analysis
+
+```{r}
+# Simple model with random intercept (can't fit random slope)
+hba1cs$Change <- (hba1cs$Time > 0) * hba1cs$Time
+mod <- lmer(HbA1c ~ Time + Change + (1 | `CFF ID`), data = hba1cs)
+hba1c_df <- mod@frame
+hba1c_df$pop_pred <- predict(mod, re.form = NA)
+hba1c_df$pred <- predict(mod)
+# Plot
+ggplot(hba1c_df, aes(x = Time)) +
+  geom_point(
+    aes(y = HbA1c, colour = factor(`CFF ID`), group = factor(`CFF ID`)),
+    alpha = 0.1
+  ) +
+  geom_line(
+    aes(y = pred, colour = factor(`CFF ID`), group = factor(`CFF ID`)),
+    alpha = 0.1
+  ) +
+  geom_line(aes(y = pop_pred)) +
+  theme_bw() +
+  theme(legend.position = "none")
+# Print table
+kable(summary(mod)$coefficients)
 ```

### 2024-09-16 Colleen Wood/Pre-post Trikafta/clean_and_merge.R

@@ -1,20 +1,18 @@
-library(dplyr)
-library(tidyr)
-library(readr)
-library(purrr)
+library(tidyverse)
 library(lubridate)
 library(childsds)
 library(rspiro)
-source("/home/tim/GitHub/CF/genotype_class_severity.R")
+library(cgmanalysis)
+source("/Users/timvigers/GitHub/CF/genotype_class_severity.R")
 t1_variables <- c(
   "CFF ID", "Site", "Sex", "Race", "Genotypes1", "Genotypes2", "Pancreatic Status",
   "G-tube in past 12 months", "Time", "Age",
   "BMI Percentile", "Height", "Weight", "FEV1", "FVC", "HbA1c"
 )
 # CHCO
-chco_a1c <- read_csv("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/chco_glycemic.csv", show_col_types = F)
+chco_a1c <- read_csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/chco_glycemic.csv", show_col_types = F)
 chco_a1c$Date <- parse_date_time(chco_a1c$Date, orders = "mdy")
-chco_bmi <- read_csv("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/chco_bmi_pft.csv", show_col_types = F)
+chco_bmi <- read_csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/chco_bmi_pft.csv", show_col_types = F)
 chco_bmi$Date <- parse_date_time(chco_bmi$Date, orders = "mdy")
 chco <- full_join(chco_a1c, chco_bmi)
 # Fill CFF ID by MRN
@@ -25,13 +23,13 @@ chco <- chco %>%
   filter(!is.na(`CFF ID`)) %>%
   select(-MRN)
 # Add demographics, filter
-chco_reg <- read_csv("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/chco_registry.csv", show_col_types = F)
[…]

### 2024-09-18 Quarto Template.qmd
Update Quarto Template.qmd

@@
       dark: darkly
-bibliography: /home/tim/Documents/Miscellaneous/zotero.bib
-csl: /home/tim/GitHub/styles/american-medical-association.csl
+bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
+csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
 editor: source

### 2024-09-18 Jordana Hoppe/SSwAB-IT/process_data.qmd
Update process_data.qmd

@@
       dark: darkly
-bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
-csl: /home/timvigers/GitHub/styles/american-medical-association.csl
+bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
+csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
 editor: source

### 2024-09-18 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.qmd
Update pre_post_trikafta_analysis.qmd

@@
 library(lmerTest)
+library(car)
 home_dir <- switch(Sys.info()["sysname"],
@@
 df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf),
-  right = F
+  right = F, labels = c("HbA1c <6.5% pre-ETI", "HbA1c ≥6.5% post-ETI")
 )
@@
   theme_bw() +
-  theme(legend.position = "none")
+  theme(legend.position = "none") +
+  xlab("Days from ETI Start")
 # Print table
 kable(summary(mod)$coefficients)
+linearHypothesis(mod, "Time + Change = 0")
 ```
@@
   select(
-    `CFF ID`, `Pre or Post`, `Age group at baseline`, `Num. A1c Values`, HbA1c, Time,
-    Weight, `BMI Percentile`, ppFEV1, ppFEV1gl, ppFVC, ppFVCgl, TDD,
+    `CFF ID`, `Pre or Post`, `Age group at baseline`, `Num. A1c Values`, HbA1c,
+    Time, Weight, `BMI Percentile`, ppFEV1, ppFEV1gl, ppFVC, ppFVCgl, TDD,
   ) %>%
@@
   geom_line(alpha = 0.2) +
+  geom_hline(yintercept = 6.5, linetype = "dashed") +
   theme_classic() +

### 2024-09-18 Christine Chan/EnVision CF/catecholamines_analysis.qmd
envision catecholamine update

@@
 library(emmeans)
+library(GLMMadaptive)
 home_dir <- switch(Sys.info()["sysname"],
@@
 norepinephrines <- paste0("Norepinephrine_", timepoints)
+adren_survey <- paste0("adren_score_", limited_timepoints)
+neuro_survey <- paste0("neuro_score_", limited_timepoints)
+total_survey <- paste0("total_score_", limited_timepoints)
+num_symptoms_survey <- paste0("num_symptoms_", limited_timepoints)
+hypo_surveys <- c(adren_survey, neuro_survey, total_survey, num_symptoms_survey)
 # For this analysis, we only want those with catecholamines measured
@@
     study_id, Hypo60, Hypo70,
-    all_of(c(glucoses, insulins, epinephrines, norepinephrines))
+    all_of(c(glucoses, insulins, epinephrines, norepinephrines, hypo_surveys))
   ) %>%
-  pivot_longer(all_of(c(glucoses, insulins, epinephrines, norepinephrines)),
+  pivot_longer(
+    all_of(c(glucoses, insulins, epinephrines, norepinephrines, hypo_surveys)),
     names_to = c(".value", "Timepoint"),
-    names_pattern = "(.*)_(.*)"
+    names_pattern = "(.*)_([0-9]{1,3})"
   )
+lmm_df$Timepoint <- factor(lmm_df$Timepoint, levels = timepoints)
+# Remove BMI percentile for those >= 20
+df$bmi_perc[df$age_visit >= 20] <- NA
+df$bmi[df$age_visit < 20] <- NA
 ```
@@
   select(
-    redcap_data_access_group, age_visit:ethnicity, cftr_mutation_1,
-    cftr_mutation_2
+    redcap_data_access_group, age_visit:ethnicity, bmi, bmi_perc, CFTR
   ) %>%
@@
   select(
-    Hypo70, age_visit:cftr_mutation_2, hypoglycemia_symptoms_yesno,
+    Hypo70, age_visit:ethnicity, bmi, bmi_perc, CFTR,
     all_of(paste0("Glucose_", timepoints)),
@@
 
+## Hypoglycemia symptoms
+
+### `adren_score`
+
+`adren_score` was modeled using a generalized linear mixed effect model assuming a Poisson distribution for the outcome and a log link function.
+We also considered a negative binomial model, but the Poisson model was better by AIC.
+AICs for the Poisson and a zero-inflated Poisson were similar, so the Poisson was chosen for ease of interpretation.
+The following models were also determined based on AIC.
+
+```{r}
+#| warning: false
+# Fit
+# Poisson model better than NB by AIC and BIC
+mod <- glmer(adren_score ~ Timepoint * Hypo70 + (1 | study_id),
+  data = lmm_df,
+  family = "poisson"
+)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Rate", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Rates") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Rate, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Rates adren_score") +
+  xlab("Timepoint") +
+  theme_bw() +
+  theme(legend.title = element_blank())
+```
+
+### `neuro_score`
+
+`neuro_score` was modeled using a zero-inflated Poisson model with random intercept for participant and an intercept as the only fixed effect in the zero part of the model.
+
+```{r}
+#| warning: false
+# Fit
+# Poisson model no different from NB by AIC
+t <- lmm_df %>% filter(Timepoint %in% limited_timepoints)
+mod <- mixed_model(neuro_score ~ Timepoint * Hypo70,
+  random = ~ 1 | study_id, data = t,
+  family = zi.poisson(), zi_fixed = ~1
+)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Rate", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Rates") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Rate, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Rates neuro_score") +
+  xlab("Timepoint") +
+  theme_bw() +
+  theme(legend.title = element_blank())
+```
+
+### `num_symptoms`
+
+`num_symptoms` was modeled using a Poisson model with random intercept for participant.
+
+```{r}
+#| warning: false
+# Fit
+# Poisson better than NB by AIC
+mod <- glmer(num_symptoms ~ Timepoint * Hypo70 + (1 | study_id),
+  data = lmm_df,
+  family = "poisson"
+)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Rate", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Rates") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Rate, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Rates num_symptoms") +
+  xlab("Timepoint") +
+  theme_bw() +
+  theme(legend.title = element_blank())
+```
+
+### `total_score`
+
+`total_score` was modeled using a zero-inflated Poisson model with random intercept for participant and an intercept and hypoglycemia group as the fixed effects in the zero part of the model.
+
+```{r}
+#| warning: false
+# Fit
+mod <- mixed_model(total_score ~ Timepoint * Hypo70,
+  random = ~ 1 | study_id, data = t,
+  family = zi.poisson(), zi_fixed = ~Hypo70
+)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Rate", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Rates") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Rate, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Rates total_score") +
+  xlab("Timepoint") +
+  theme_bw() +
+  theme(legend.title = element_blank())
+```
+
 # Hypoglycemia defined as < 60 mg/dL at any point
@@
   select(
-    Hypo60, age_visit:cftr_mutation_2, hypoglycemia_symptoms_yesno,
+    Hypo60, age_visit:ethnicity, bmi, bmi_perc, CFTR,
     all_of(paste0("Glucose_", timepoints)),
@@
 ```
+
+## Hypoglycemia symptoms
+
+### `adren_score`
+
+`adren_score` was modeled using a generalized linear mixed effect model assuming a Poisson distribution for the outcome and a log link function.
+We also considered a negative binomial model, but the Poisson model was better by AIC.
+AICs for the Poisson and a zero-inflated Poisson were similar, so the Poisson was chosen for ease of interpretation.
+The following models were also determined based on AIC.
+
+```{r}
+#| warning: false
+# Fit
+# Poisson model better than NB by AIC and BIC
+mod <- glmer(adren_score ~ Timepoint * Hypo60 + (1 | study_id),
+  data = lmm_df,
+  family = "poisson"
+)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Rate", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Rates") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Rate, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Rates adren_score") +
+  xlab("Timepoint") +
+  theme_bw() +
+  theme(legend.title = element_blank())
+```
+
+### `neuro_score`
+
+`neuro_score` was modeled using a zero-inflated Poisson model with random intercept for participant and an intercept as the only fixed effect in the zero part of the model.
+
+```{r}
+#| warning: false
+# Fit
+# Poisson model no different from NB by AIC
+t <- lmm_df %>% filter(Timepoint %in% limited_timepoints)
+mod <- mixed_model(neuro_score ~ Timepoint * Hypo60,
+  random = ~ 1 | study_id, data = t,
+  family = zi.poisson(), zi_fixed = ~1
+)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Rate", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Rates") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Rate, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Rates neuro_score") +
+  xlab("Timepoint") +
+  theme_bw() +
+  theme(legend.title = element_blank())
+```
+
+### `num_symptoms`
+
+`num_symptoms` was modeled using a Poisson model with random intercept for participant.
+
+```{r}
+#| warning: false
+# Fit
+# Poisson better than NB by AIC
+mod <- glmer(num_symptoms ~ Timepoint * Hypo60 + (1 | study_id),
+  data = lmm_df,
+  family = "poisson"
+)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Rate", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Rates") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Rate, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Rates num_symptoms") +
+  xlab("Timepoint") +
+  theme_bw() +
+  theme(legend.title = element_blank())
+```
+
+### `total_score`
+
+`total_score` was modeled using a zero-inflated Poisson model with random intercept for participant and an intercept and hypoglycemia group as the fixed effects in the zero part of the model.
+
+```{r}
+#| warning: false
+# Fit
+mod <- mixed_model(total_score ~ Timepoint * Hypo60,
+  random = ~ 1 | study_id, data = t,
+  family = zi.poisson(), zi_fixed = ~Hypo60
+)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Rate", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Rates") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Rate, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Rates total_score") +
+  xlab("Timepoint") +
+  theme_bw() +
+  theme(legend.title = element_blank())
+```

### 2024-09-18 Christine Chan/EnVision CF/clean_envision_data.R

@@ -86,9 +86,9 @@ glucose$Date <- ymd(glucose$Date)
 #-------------------------------------------------------------------------------
 visits <- redcap %>%
   select(
-    study_id, date_visit, age_visit, height, weight, sex, origin_race,
-    ethnicity, redcap_data_access_group, cftr_mutation_1, cftr_mutation_2,
-    hypoglycemia_symptoms_yesno
+    study_id, date_visit, age_visit, height, weight, bmi_study_visit,
+    sex, origin_race, ethnicity, redcap_data_access_group,
+    cftr_mutation_1, cftr_mutation_2
   ) %>%
   filter(!is.na(date_visit)) %>%
   rename(Date = date_visit) %>%
@@ -169,6 +169,23 @@ insulin <- insulin %>%
 # Convert to numeric
 insulin$Insulin <- as.numeric(insulin$Insulin)
 #-------------------------------------------------------------------------------
+# Hypoglycemia symptom surveys
+#-------------------------------------------------------------------------------
+hypo_surveys <- c(
+  "adren_score_baseline", "adren_score_120", "adren_score_150", "adren_score_180",
+  "neuro_score_baseline", "neuro_score_120", "neuro_score_150", "neuro_score_180",
+  "total_score_baseline", "total_score_120", "total_score_150", "total_score_180",
+  "num_symptoms_base", "num_symptoms_120", "num_symptoms_150", "num_symptoms_180"
+)
+hypo_symptoms <- redcap %>%
+  select(study_id, date_visit, all_of(hypo_surveys)) %>%
+  rename(
+    Date = date_visit, adren_score_0 = adren_score_baseline,
+    neuro_score_0 = neuro_score_baseline, total_score_0 = total_score_baseline,
+    num_symptoms_0 = num_symptoms_base
+  )
+hypo_symptoms$Date <- ymd(hypo_symptoms$Date)
+#-------------------------------------------------------------------------------
 # CGM data
 #-------------------------------------------------------------------------------
 # Clean data (does not need to be run every time)
@@ -229,6 +246,7 @@ final_df <- final_df %>% pivot_wider(
 )
 final_df <- full_join(final_df, cgm)
 final_df <- full_join(final_df, visits)
+final_df <- full_join(final_df, hypo_symptoms)
 # Add HbA1c, etc.
[…]

### 2024-09-18 Christine Chan/EnVision CF/catecholamines_analysis.qmd
Update catecholamines_analysis.qmd

--- 2024-09-11 Christine Chan/EnVision CF/catecholamines_analysis.qmd
@@
 #| include: false
-library(dplyr)
-library(tidyr)
-library(ggplot2)
+library(tidyverse)
 library(gt)
@@
 knitr::opts_knit$set(root.dir = home_dir)
+emm_options(lmerTest.limit = 1e6, pbkrtest.limit = 1e6)
 ```
@@
 timepoints <- c(0, 10, 30, 60, 90, 120, 150, 180)
+limited_timepoints <- c(0, 120, 150, 180)
 glucoses <- paste0("Glucose_", timepoints)
@@
 norepinephrines <- paste0("Norepinephrine_", timepoints)
-# For this analysis, we only want those with catecholamines measured
+adren_survey <- paste0("adren_score_", limited_timepoints)
+neuro_survey <- paste0("neuro_score_", limited_timepoints)
+total_survey <- paste0("total_score_", limited_timepoints)
+num_symptoms_survey <- paste0("num_symptoms_", limited_timepoints)
+hypo_surveys <- c(adren_survey, neuro_survey, total_survey, num_symptoms_survey)
+# For some of this analysis, we only want those with catecholamines measured
 df$catecholamines <- apply(
@@
 )
-df <- df %>% filter(catecholamines)
-# To long format
-t <- df %>%
+# Make a long dataset for mixed models
+lmm_df <- df %>%
   select(
-    study_id, Date,
-    all_of(c(glucoses, insulins, epinephrines, norepinephrines))
+    study_id, catecholamines, Hypo60, Hypo70,
+    all_of(c(glucoses, insulins, epinephrines, norepinephrines, hypo_surveys))
   ) %>%
-  pivot_longer(all_of(c(glucoses, insulins, epinephrines, norepinephrines)),
+  pivot_longer(
+    all_of(c(glucoses, insulins, epinephrines, norepinephrines, hypo_surveys)),
     names_to = c(".value", "Timepoint"),
-    names_pattern = "(.*)_(.*)")
+    names_pattern = "(.*)_([0-9]{1,3})"
+  )
+lmm_df$Timepoint <- factor(lmm_df$Timepoint, levels = timepoints)
+# Remove BMI percentile for those >= 20
+df$bmi_perc[df$age_visit >= 20] <- NA
+df$bmi[df$age_visit < 20] <- NA
+# Remove 0 values
+lmm_df$Insulin[lmm_df$Insulin == 0] <- NA
 ```
@@
   group_by(study_id) %>%
-  filter(row_number() == 1) %>%
+  slice_min(age_visit) %>%
   ungroup() %>%
   select(
-    redcap_data_access_group, age_visit:ethnicity, cftr_mutation_1,
-    cftr_mutation_2
+    redcap_data_access_group, age_visit:ethnicity, bmi, bmi_perc, CFTR
   ) %>%
@@
 
-## Demographics
+## Full cohort
 
-IA0005, IA0006, IA0017, IA0042, and IA0079 had catecholamines measured at multiple timepoints but these are treated as independent in the table below.
+### Demographics
+
+Repeat visits are treated as independent in the table below.
 
@@
 df %>%
-  filter(Timepoint %in% c(0, 120, 150, 180)) %>%
-  pivot_wider(
-    names_from = Timepoint, values_from = c(Epinephrine, Norepinephrine),
-    id_cols = c(study_id:cftr_mutation_2, Hypo70)
+  select(
+    Hypo70, age_visit:ethnicity, bmi, bmi_perc, CFTR,
+    all_of(paste0("Glucose_", timepoints)),
+    all_of(paste0("Insulin_", timepoints))
   ) %>%
-  select(Hypo70, age_visit:cftr_mutation_2, Epinephrine_0:Norepinephrine_180) %>%
   tbl_summary(by = Hypo70, missing_text = "Missing") %>%
@@
 
-## Epinephrine
+### Glucose
+
+Glucose was modeled using a linear mixed effect model with random intercept for participant and time treated as a categorical variable.
+
+```{r}
+#| warning: false
+# Fit
+mod <- lmer(Glucose ~ Timepoint * Hypo70 + (1 | study_id),
+  data = lmm_df
+)
+mod %>% tbl_regression(intercept = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Means") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  ylab("LS Mean Glucose") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+### Insulin
+
+Insulin was modeled using a linear mixed effect model with random intercept for participant and time treated as a categorical variable.
+
+```{r}
+#| warning: false
+# Fit
+mod <- lmer(log(Insulin) ~ Timepoint * Hypo70 + (1 | study_id),
+  data = lmm_df
+)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Means") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  ylab("LS Mean Insulin") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+### Hypoglycemia symptoms
+
+#### `adren_score`
+
+`adren_score` was modeled using a generalized linear mixed effect model assuming a Poisson distribution for the outcome and a log link function.
+We also considered a negative binomial model, but the Poisson model was better by AIC.
+AICs for the Poisson and a zero-inflated Poisson were similar, so the Poisson was chosen for ease of interpretation.
+The following models were also determined based on AIC.
+
+```{r}
+#| warning: false
+# Fit
+# Poisson model better than NB by AIC and BIC
+mod <- glmer(adren_score ~ Timepoint * Hypo70 + (1 | study_id),
+  data = lmm_df,
+  family = "poisson"
+)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Rate", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Rates") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Rate, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Rates adren_score") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+#### `neuro_score`
+
+`neuro_score` was modeled using a zero-inflated Poisson model with random intercept for participant and an intercept as the only fixed effect in the zero part of the model.
+
+```{r}
+#| warning: false
+# Fit
+# Poisson model no different from NB by AIC
+t <- lmm_df %>% filter(Timepoint %in% limited_timepoints)
+mod <- mixed_model(neuro_score ~ Timepoint * Hypo70,
+  random = ~ 1 | study_id, data = t,
+  family = zi.poisson(), zi_fixed = ~1
+)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Rate", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Rates") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Rate, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Rates neuro_score") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+#### `num_symptoms`
+
+`num_symptoms` was modeled using a Poisson model with random intercept for participant.
+
+```{r}
+#| warning: false
+# Fit
+# Poisson better than NB by AIC
+mod <- glmer(num_symptoms ~ Timepoint * Hypo70 + (1 | study_id),
+  data = lmm_df,
+  family = "poisson"
+)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Rate", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Rates") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Rate, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Rates num_symptoms") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+#### `total_score`
+
+`total_score` was modeled using a zero-inflated Poisson model with random intercept for participant and an intercept and hypoglycemia group as the fixed effects in the zero part of the model.
+
+```{r}
+#| warning: false
+# Fit
+mod <- mixed_model(total_score ~ Timepoint * Hypo70,
+  random = ~ 1 | study_id, data = t,
+  family = zi.poisson(), zi_fixed = ~Hypo70
+)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Rate", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Rates") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Rate, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Rates total_score") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+## Catecholamines only
+
+```{r}
+cat_df <- df %>% filter(catecholamines)
+cat_lmm_df <- lmm_df %>% filter(catecholamines)
+```
+
+### Demographics
+
+IA0005, IA0006, IA0017, IA0042, and IA0079 had catecholamines measured at multiple timepoints but these are treated as independent in the table below.
+
+```{r}
+#| label: tbl-70-catecholamines
+#| tbl-cap: BG < 70 at Any Point (Catecholamines Only)
+#| warning: false
+cat_df %>%
+  select(
+    Hypo70, age_visit:ethnicity, bmi, bmi_perc, CFTR,
+    all_of(paste0("Glucose_", timepoints)),
+    all_of(paste0("Insulin_", timepoints)),
+    all_of(paste0("Epinephrine_", limited_timepoints)),
+    all_of(paste0("Norepinephrine_", limited_timepoints))
+  ) %>%
+  tbl_summary(by = Hypo70, missing_text = "Missing") %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes()
+```
+
+### Epinephrine
 
@@
 # Fit
-mod <- lmer(log(Epinephrine) ~ Timepoint * Hypo70 + (1 | study_id), data = df)
+mod <- lmer(log(Epinephrine) ~ Timepoint * Hypo70 + (1 | study_id),
+  data = cat_lmm_df
+)
 mod %>% tbl_regression(intercept = T, exponentiate = T)
 means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"), type = "response"))
-means %>% gt(caption = "LS Means")
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Means") %>%
+  fmt_number(decimals = 2)
 # Plot
-emmip(mod, Hypo70 ~ Timepoint, CIs = T, type = "response") +
-  theme_bw() +
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Mean Epinephrine") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
   theme(legend.title = element_blank())
@@
 
-## Norepinephrine
+### Norepinephrine
 
@@
 # Fit
-mod <- lmer(log(Norepinephrine) ~ Timepoint * Hypo70 + (1 | study_id), data = df)
+mod <- lmer(log(Norepinephrine) ~ Timepoint * Hypo70 + (1 | study_id),
+  data = cat_lmm_df
+)
 mod %>% tbl_regression(intercept = T, exponentiate = T)
 means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"), type = "response"))
-means %>% gt(caption = "LS Means")
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Means") %>%
+  fmt_number(decimals = 2)
 # Plot
-emmip(mod, Hypo70 ~ Timepoint, CIs = T, type = "response") +
-  theme_bw() +
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Mean Norepinephrine") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
   theme(legend.title = element_blank())
@@
 df %>%
-  filter(Timepoint %in% c(0, 120, 150, 180)) %>%
-  pivot_wider(
-    names_from = Timepoint, values_from = c(Epinephrine, Norepinephrine),
-    id_cols = c(study_id:cftr_mutation_2, Hypo60)
+  select(
+    Hypo60, age_visit:ethnicity, bmi, bmi_perc, CFTR,
+    all_of(paste0("Glucose_", timepoints)),
+    all_of(paste0("Insulin_", timepoints))
   ) %>%
-  select(Hypo60, age_visit:cftr_mutation_2, Epinephrine_0:Norepinephrine_180) %>%
   tbl_summary(by = Hypo60, missing_text = "Missing") %>%
@@
 
-## Epinephrine
+### Glucose
+
+Glucose was modeled using a linear mixed effect model with random intercept for participant and time treated as a categorical variable.
+
+```{r}
+#| warning: false
+# Fit
+mod <- lmer(Glucose ~ Timepoint * Hypo60 + (1 | study_id),
+  data = lmm_df
+)
+mod %>% tbl_regression(intercept = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Means") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  ylab("LS Mean Glucose") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+### Insulin
+
+Insulin was modeled using a linear mixed effect model with random intercept for participant and time treated as a categorical variable.
+
+```{r}
+#| warning: false
+# Fit
+mod <- lmer(log(Insulin) ~ Timepoint * Hypo60 + (1 | study_id),
+  data = lmm_df
+)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Means") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  ylab("LS Mean Insulin") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+### Hypoglycemia symptoms
+
+#### `adren_score`
+
+`adren_score` was modeled using a generalized linear mixed effect model assuming a Poisson distribution for the outcome and a log link function.
+We also considered a negative binomial model, but the Poisson model was better by AIC.
+AICs for the Poisson and a zero-inflated Poisson were similar, so the Poisson was chosen for ease of interpretation.
+The following models were also determined based on AIC.
+
+```{r}
+#| warning: false
+# Fit
+# Poisson model better than NB by AIC and BIC
+mod <- glmer(adren_score ~ Timepoint * Hypo60 + (1 | study_id),
+  data = lmm_df,
+  family = "poisson"
+)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Rate", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Rates") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Rate, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Rates adren_score") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+#### `neuro_score`
+
+`neuro_score` was modeled using a zero-inflated Poisson model with random intercept for participant and an intercept as the only fixed effect in the zero part of the model.
+
+```{r}
+#| warning: false
+# Fit
+# Poisson model no different from NB by AIC
+t <- lmm_df %>% filter(Timepoint %in% limited_timepoints)
+mod <- mixed_model(neuro_score ~ Timepoint * Hypo60,
+  random = ~ 1 | study_id, data = t,
+  family = zi.poisson(), zi_fixed = ~1
+)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Rate", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Rates") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Rate, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Rates neuro_score") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+#### `num_symptoms`
+
+`num_symptoms` was modeled using a Poisson model with random intercept for participant.
+
+```{r}
+#| warning: false
+# Fit
+# Poisson better than NB by AIC
+mod <- glmer(num_symptoms ~ Timepoint * Hypo60 + (1 | study_id),
+  data = lmm_df,
+  family = "poisson"
+)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Rate", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Rates") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Rate, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Rates num_symptoms") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+#### `total_score`
+
+`total_score` was modeled using a zero-inflated Poisson model with random intercept for participant and an intercept and hypoglycemia group as the fixed effects in the zero part of the model.
+
+```{r}
+#| warning: false
+# Fit
+mod <- mixed_model(total_score ~ Timepoint * Hypo60,
+  random = ~ 1 | study_id, data = t,
+  family = zi.poisson(), zi_fixed = ~Hypo60
+)
+mod %>% tbl_regression(intercept = T, exponentiate = T)
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"), type = "response"))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Rate", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Rates") %>%
+  fmt_number(decimals = 2)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Rate, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Rates total_score") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+## Catecholamines only
+
+### Demographics
+
+IA0005, IA0006, IA0017, IA0042, and IA0079 had catecholamines measured at multiple timepoints but these are treated as independent in the table below.
+
+```{r}
+#| label: tbl-60-catecholamines
+#| tbl-cap: BG < 60 at Any Point (Catecholamines Only)
+#| warning: false
+cat_df %>%
+  select(
+    Hypo60, age_visit:ethnicity, bmi, bmi_perc, CFTR,
+    all_of(paste0("Glucose_", timepoints)),
+    all_of(paste0("Insulin_", timepoints)),
+    all_of(paste0("Epinephrine_", limited_timepoints)),
+    all_of(paste0("Norepinephrine_", limited_timepoints))
+  ) %>%
+  tbl_summary(by = Hypo60, missing_text = "Missing") %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes()
+```
+
+### Epinephrine
 
@@
 # Fit
-mod <- lmer(log(Epinephrine) ~ Timepoint * Hypo60 + (1 | study_id), data = df)
+mod <- lmer(log(Epinephrine) ~ Timepoint * Hypo60 + (1 | study_id),
+  data = cat_lmm_df
+)
 mod %>% tbl_regression(intercept = T, exponentiate = T)
 means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"), type = "response"))
-means %>% gt(caption = "LS Means")
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Means") %>%
+  fmt_number(decimals = 2)
 # Plot
-emmip(mod, Hypo60 ~ Timepoint, CIs = T, type = "response") +
-  theme_bw() +
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Mean Epinephrine") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
   theme(legend.title = element_blank())
@@
 # Fit
-mod <- lmer(log(Norepinephrine) ~ Timepoint * Hypo60 + (1 | study_id), data = df)
+mod <- lmer(log(Norepinephrine) ~ Timepoint * Hypo60 + (1 | study_id),
+  data = cat_lmm_df
+)
 mod %>% tbl_regression(intercept = T, exponentiate = T)
 means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"), type = "response"))
-means %>% gt(caption = "LS Means")
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+means %>%
+  gt(caption = "LS Means") %>%
+  fmt_number(decimals = 2)
 # Plot
-emmip(mod, Hypo60 ~ Timepoint, CIs = T, type = "response") +
-  theme_bw() +
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = limited_timepoints) +
+  ylab("LS Mean Norepinephrine") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
   theme(legend.title = element_blank())

### 2024-09-19 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
Update cf_registry_ROC_analysis.qmd

--- 2024-08-08 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
@@
     fig-cap-location: top
-    fig-width: 8
-    fig-height: 8
+    fig-width: 12
+    fig-height: 9
     page-layout: full
-bibliography: /home/tim/Documents/Miscellaneous/zotero.bib
-csl: /home/tim/GitHub/styles/american-medical-association.csl
+bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
+csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
 editor: source
@@
 library(pROC)
+library(ggrepel)
 home_dir <- switch(Sys.info()["sysname"],
@@
 
+```{r}
+# Plotting function
+roc_plot <- function(roc_obj) {
+  c <- coords(roc_obj, "best")
+  plot_df <- coords(roc_obj)
+  point_df <- plot_df[plot_df$threshold == as.numeric(c[1]), ]
+  c <- round(as.numeric(c), 2)
+  point_label <- paste0(
+    "Threshold: ", c[1], "\n",
+    "Specificity: ", c[2], "\n",
+    "Sensitivity: ", c[3]
+  )
+  p <- ggplot(plot_df, aes(x = specificity, y = sensitivity)) +
+    scale_x_reverse() +
+    geom_line() +
+    geom_point(data = point_df) +
+    geom_label_repel(
+      data = point_df, aes(label = point_label), nudge_x = 0.25,
+      segment.color = "grey50", size = 5
+    ) +
+    xlab("Specificity") +
+    ylab("Sensitivity") +
+    theme_bw(base_size = 20)
+  return(p)
+}
+```
+
 # Data cleaning
@@
   ylab("HbA1c") +
-  theme_classic()
+  theme_classic(base_size = 20)
 ```
@@
 r <- roc(ogtt ~ hgba1c, data = ngt_agt)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(ogtt ~ hgba1c, data = agt_cfrd)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(ogtt ~ hgba1c, data = peds_ngt_agt)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(ogtt ~ hgba1c, data = peds_agt_cfrd)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(ogtt ~ hgba1c, data = adult_ngt_agt)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(ogtt ~ hgba1c, data = adult_agt_cfrd)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(ogtt ~ hgba1c, data = white_ngt_agt)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(ogtt ~ hgba1c, data = white_agt_cfrd)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(ogtt ~ hgba1c, data = black_ngt_agt)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(ogtt ~ hgba1c, data = black_agt_cfrd)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(coords(r, "best"))) %>%
@@
 r <- roc(ogtt ~ hgba1c, data = more_ngt_agt)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(ogtt ~ hgba1c, data = more_agt_cfrd)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(ogtt ~ hgba1c, data = other_ngt_agt)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(coords(r, "best"))) %>%
@@
 r <- roc(ogtt ~ hgba1c, data = other_agt_cfrd)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(coords(r, "best"))) %>%
@@
   fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+```
+
+### By ethnicity
+
+#### Hispanic
+
+```{r}
+hispanic_ngt_agt <- ngt_agt %>% filter(Hispanicrace == "Yes")
+hispanic_agt_cfrd <- agt_cfrd %>% filter(Hispanicrace == "No")
+```
+
+##### NGT vs. AGT
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = hispanic_ngt_agt)
+roc_plot(r)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+```
+
+##### AGT vs. CFRD
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = hispanic_agt_cfrd)
+roc_plot(r)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+```
+
+#### Non-Hispanic
+
+```{r}
+non_hispanic_ngt_agt <-
+  ngt_agt %>% filter(Race == "Black or African American")
+non_hispanic_agt_cfrd <-
+  agt_cfrd %>% filter(Race == "Black or African American")
+```
+
+##### NGT vs. AGT
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = non_hispanic_ngt_agt)
+roc_plot(r)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+```
+
+##### AGT vs. CFRD
+
+Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.
+
+```{r}
+#| message: false
+r <- roc(ogtt ~ hgba1c, data = non_hispanic_agt_cfrd)
+roc_plot(r)
+gt(data.frame(coords(r, "best"))) %>%
+  fmt_number(n_sigfig = 3)
+gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
+  fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(agt ~ hgba1c, data = roc_df)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(cfrd ~ hgba1c, data = roc_df)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(agt ~ hgba1c, data = peds_roc)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(cfrd ~ hgba1c, data = peds_roc)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(agt ~ hgba1c, data = adult_roc)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(cfrd ~ hgba1c, data = adult_roc)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(agt ~ hgba1c, data = white_roc)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(cfrd ~ hgba1c, data = white_roc)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(agt ~ hgba1c, data = black_roc)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 
+Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.
+
 ```{r}
@@
 r <- roc(cfrd ~ hgba1c, data = black_roc)
-plot(r, print.auc = T, print.thres = T)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+roc_plot(r)
+gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
+  fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(agt ~ hgba1c, data = more_roc)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(cfrd ~ hgba1c, data = more_roc)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
@@
 r <- roc(agt ~ hgba1c, data = other_roc)
-plot(r, print.auc = T, print.thres = T)
+roc_plot(r)
 gt(data.frame(coords(r, "best"))) %>%
@@
 
-Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.
+```{r}
+#| message: false
+r <- roc(cfrd ~ hgba1c, data = other_roc)
+roc_plot(r)
+gt(data.frame(coords(r, "best"))) %>%
+  fmt_number(n_sigfig = 3)
+gt(data.frame(ci.coords(r, "best"))) %>%
+  fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+```
+
+### By ethnicity
+
+#### Hispanic
+
+```{r}
+hispanic_roc <- roc_df %>% filter(Hispanicrace == "Yes")
+```
+
+##### $\geq$ 140
 
@@
 #| message: false
-r <- roc(cfrd ~ hgba1c, data = other_roc)
-plot(r, print.auc = T, print.thres = T)
-gt(data.frame(coords(r, "best"))) %>%
-  fmt_number(n_sigfig = 3)
-gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
-  fmt_number(n_sigfig = 3)
+r <- roc(agt ~ hgba1c, data = hispanic_roc)
+roc_plot(r)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
 ```
+
+##### $\geq$ 200
+
+```{r}
+#| message: false
+r <- roc(cfrd ~ hgba1c, hispanic_roc)
+roc_plot(r)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+```
+
+#### Non-Hispanic
+
+```{r}
+non_hispanic_roc <- roc_df %>% filter(Hispanicrace == "No")
+```
+
+##### $\geq$ 140
+
+```{r}
+#| message: false
+r <- roc(agt ~ hgba1c, data = non_hispanic_roc)
+roc_plot(r)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+```
+
+##### $\geq$ 200
+
+```{r}
+#| message: false
+r <- roc(cfrd ~ hgba1c, data = non_hispanic_roc)
+roc_plot(r)
+gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+```

### 2024-09-19 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
Update cf_registry_ROC_analysis.qmd

@@
 
+## Boxplot
+
 ```{r}
@@
   theme_classic(base_size = 20)
+```
+
+## Dotplot
+
+```{r}
+ggplot(roc_df, aes(x = `OGTT 2-hour`, y = hgba1c)) +
+  geom_jitter(alpha = 0.25) +
+  ylab("HbA1c") +
+  theme_classic(base_size = 20)
+```
+
+## Violin plot
+
+```{r}
+ggplot(roc_df, aes(x = `OGTT 2-hour`, y = hgba1c)) +
+  geom_violin(aes(fill = `OGTT 2-hour`)) +
+  ylab("HbA1c") +
+  theme_classic(base_size = 20) +
+  theme(legend.position = "none")
 ```

### 2024-09-19 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
Update phase_1_analysis.qmd

@@
       dark: darkly
-bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
-csl: /home/timvigers/GitHub/styles/american-medical-association.csl
 editor: source
@@
 Accuracy is the proportion of samples where the two tests agreed (positive/positive and negative/negative).
-Cohen's kappa accounts for agreement between the two tests by chance, and is interpreted analogously to correlation where -1 indicates perfect disagreement and 1 indicates perfect agreement [@mchughInterraterReliabilityKappa2012;@ranganathanCommonPitfallsStatistical2017].
+Cohen's kappa accounts for agreement between the two tests by chance, and is interpreted analogously to correlation where -1 indicates perfect disagreement and 1 indicates perfect agreement.
 A Cohen's kappa of NaN or NA indicates a lack of variability between the two tests.
@@
   d %>%
-    tbl_cross(percent = "cell") %>%
-    add_ci() %>%
+    tbl_summary(by = COS, percent = "cell") %>%
+    add_ci(method = list(all_categorical() ~ "wald")) %>%
     print()
@@
   cat("\n")
-  print(plot(euler(d == "Negative"), quantities = list(cex = 3)))
+  print(plot(euler(d == "Negative"),
+    quantities = list(cex = 3),
+    fills = list(fill = c("#FFC107", "#1E88E5"))
+  ))
   cat("\n")
@@
   cat("\n")
-  print(plot(euler(d == "Positive"), quantities = list(cex = 3)))
+  print(plot(euler(d == "Positive"),
+    quantities = list(cex = 3),
+    fills = list(fill = c("#FFC107", "#1E88E5"))
+  ))
   cat("\n")
@@
 ```{r results='asis'}
+#| message: false
 invisible(lapply(names(bug_list), function(b) {
+  print(b)
   cat("\n")
@@
 ```{r results='asis'}
+#| message: false
 invisible(lapply(names(bug_list), function(b) {
@@
 ```
-
-# References

### 2024-09-19 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
Update phase_1_analysis.qmd

@@
     quantities = list(cex = 3),
-    fills = list(fill = c("#FFC107", "#1E88E5"))
+    fills = list(fill = c("#FFC107","#1E88E5","#004D40"))
   ))
@@
     quantities = list(cex = 3),
-    fills = list(fill = c("#FFC107", "#1E88E5"))
+    fills = list(fill = c("#FFC107", "#1E88E5","#004D40"))
   ))

### 2024-09-23 Emily Muther/body_image_trikafta.qmd
body image trikafta n=52

@@
 
-1. Participants who answered "I don't know" or "no" on the variable `TakingTrikafta` were excluded from these analyses.
+1. Participants who answered "I don't know" or "no" on the variable `TakingTrikafta` were NOT excluded from these analyses.
 
@@
 # Filter
-df <- df %>%
-  filter(TakingTrikafta == "yes")
+# df <- df %>%
+#   filter(TakingTrikafta == "yes")
 # Trim white space

### 2024-09-23 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd

--- 2024-09-19 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
@@
     page-layout: full
-bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
-csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
 editor: source
@@
   "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
+  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
   "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/CF Registry"
@@
   geom_boxplot() +
+  geom_hline(yintercept = 5.3, linetype = "dashed") +
+  geom_hline(yintercept = 5.5, linetype = "dashed") +
+  geom_hline(yintercept = 5.7, linetype = "dashed") +
   ylab("HbA1c") +
   theme_classic(base_size = 20)
+```
+
+## Dotplot
+
+```{r}
+ggplot(roc_df, aes(x = `OGTT 2-hour`, y = hgba1c)) +
+  geom_jitter(alpha = 0.05) +
+  geom_hline(yintercept = 5.3, linetype = "dashed") +
+  geom_hline(yintercept = 5.5, linetype = "dashed") +
+  geom_hline(yintercept = 5.7, linetype = "dashed") +
+  ylab("HbA1c") +
+  theme_classic(base_size = 20)
+```
+
+## Violin plot
+
+```{r}
+ggplot(roc_df, aes(x = `OGTT 2-hour`, y = hgba1c)) +
+  geom_violin(aes(fill = `OGTT 2-hour`)) +
+  geom_hline(yintercept = 5.3, linetype = "dashed") +
+  geom_hline(yintercept = 5.5, linetype = "dashed") +
+  geom_hline(yintercept = 5.7, linetype = "dashed") +
+  ylab("HbA1c") +
+  theme_classic(base_size = 20) +
+  theme(legend.position = "none")
 ```

### 2024-10-04 Quarto Template.qmd
template update

@@
   "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/tim/OneDrive/Vigers/CF"
+  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/timviers/OneDrive/Vigers/CF"
 )

### 2024-10-04 Quarto Template.qmd
Update Quarto Template.qmd

--- 2024-09-18 Quarto Template.qmd
@@
   "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/tim/OneDrive/Vigers/CF"
+  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
 )
@@
   "Darwin" = "/Users/timvigers/GitHub",
-  "Windows" = "C:/Users/Tim/GitHub",
-  "Linux" = "/home/tim/GitHub"
+  "Windows" = "C:/Users/timvigers/GitHub",
+  "Linux" = "/home/timvigers/GitHub"
 )

### 2024-10-09 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
sweat analysis update

--- 2024-09-13 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
@@
   "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Test",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Test",
-  "Linux" = "/home/tim/OneDrive/Vigers/CF/Edith Zemanick/Sweat Test"
+  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Test",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Test"
 )
@@
 df$TotalGenotypeFunction <- rowSums(
-  df[, c("Genotype1Function", "Genotype2Function")],
-  na.rm = T
+  df[, c("Genotype1Function", "Genotype2Function")]
 )
@@
 df <- as.data.frame(df)
+# Limit to complete data
+df <- df %>%
+  select(
+    PatientID, Diagnosis, Sex, Age.at.Test.in.Years, Na:Cl,
+    Genotypes1, Genotype1Function, Genotypes2, Genotype2Function,
+    TotalGenotypeFunction
+  )
+# Set non-CF to 100% CFTR function (200 because 100 for each allele)
+df$TotalGenotypeFunction[df$Diagnosis == "Non-CF"] <- 200
 # Write for checking
@@
 
-# Excluding those with no genotype function estimate
+```{r}
+#| message: false
+#| warning: false
+df %>%
+  filter(Diagnosis == "CF") %>%
+  ggplot(
+    aes(x = Na, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
+  ) +
+  geom_point() +
+  theme()
+df %>%
+  filter(Diagnosis == "CF") %>%
+  ggplot(
+    aes(x = Cl, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
+  ) +
+  geom_point() +
+  theme()
+df %>%
+  filter(Diagnosis == "CF") %>%
+  ggplot(
+    aes(x = Na / Cl, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
+  ) +
+  geom_point() +
+  theme()
 
-```{r}
 df %>%
-  filter(!is.na(TotalGenotypeFunction)) %>%
-  select(
-    Diagnosis, Genotypes1, Genotypes2, Genotype1Function, Genotype2Function,
-    TotalGenotypeFunction, Na, K, Cl
-  ) %>%
-  droplevels(.) %>%
-  tbl_summary(by = Diagnosis)
+  filter(Diagnosis == "CF") %>%
+  ggplot(
+    aes(
+      x = Na, y = TotalGenotypeFunction, color = Age.at.Test.in.Years,
+      group = Diagnosis
+    )
+  ) +
+  geom_point() +
+  theme()
+df %>%
+  filter(Diagnosis == "CF") %>%
+  ggplot(
+    aes(
+      x = Cl, y = TotalGenotypeFunction, color = Age.at.Test.in.Years,
+      group = Diagnosis
+    )
+  ) +
+  geom_point() +
+  theme()
+df %>%
+  filter(Diagnosis == "CF") %>%
+  ggplot(
+    aes(
+      x = Na / Cl, y = TotalGenotypeFunction, color = Age.at.Test.in.Years,
+      group = Diagnosis
+    )
+  ) +
+  geom_point() +
+  theme()
 ```
@@
 df %>%
-  filter(!is.na(TotalGenotypeFunction)) %>%
-  ggplot(
-    aes(x = Cl, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
-  ) +
-  geom_point() +
-  geom_smooth(method = "lm", se = F) +
-  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())
+  filter(Diagnosis == "CF") %>%
+  mutate(
+    `Age group` = cut(Age.at.Test.in.Years, c(-Inf, 18, Inf), right = F),
+    `Cl < 60` = Cl < 60
+  ) %>%
+  select(Na, Cl, TotalGenotypeFunction, `Cl < 60`, `Age group`) %>%
+  tbl_summary(by = `Age group`, missing_text = "Missing")
+write.csv(df, file = "~/sweat_analysis_data.csv", na = "", row.names = F)
+```
 
-df %>%
-  filter(!is.na(TotalGenotypeFunction)) %>%
-  ggplot(
-    aes(x = Cl, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
-  ) +
-  geom_point() +
-  geom_smooth(method = "lm", se = F) +
-  ylim(c(0, 20)) +
-  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())
-```
+Add CRMS change in sweat and change in ratio
+
+Add CF carrier genotypes to this report
+
+Plot just the CF genotype function
+  - Are the chloride values matching up with the CFTR function as we'd expect?

### 2024-10-09 Lilah Melzer/cough_monitor_analysis.qmd
started cough monitor analysis

---
title: "Home Nighttime Cough Monitoring"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(lubridate)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/timvigers/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| message: false
# Import REDCap data
source(paste0(github_dir, "/CF/Lilah Melzer/cough_monitor_import.r"))
# Import counts
agg_counts <- read.csv(
  "./Data_Raw/Cough_w_TotalCounts_0824 - Cough_agg_w_count_0828.csv",
  na.strings = c("", "NMD"), header = F
)
# Format aggregated count data
agg_counts <- data.frame(t(agg_counts))
colnames(agg_counts) <- agg_counts[1, ]
agg_counts$`Date/Device` <- NULL
agg_counts <- agg_counts[-1, ]
colnames(agg_counts)[1] <- "sid"
agg_counts$sid <- sub("FN", "", agg_counts$sid)
# Wide to long, fix dates, count number of nights
agg_counts <- agg_counts %>%
  pivot_longer(cols = -one_of("sid"), names_to = "Date", values_to = "Coughs") %>%
  filter(!is.na(Coughs)) %>%
  mutate(Date = mdy(Date),
    Coughs = as.numeric(Coughs))%>%
  group_by(sid) %>%
  mutate(
    Night = row_number()
  ) %>%
  ungroup()
```

# Demographics

```{r}
#| label: tbl-mean-demo
#| tbl-cap: Participant characteristics
```

# Coughs per night

```{r}
#| label: tbl-mean-coughs
#| tbl-cap: Mean coughs per night
```

# Coughs over time (aggregated)

```{r}
#| label: fig-coughs-by-night
#| fig-cap: Coughs by date
ggplot(agg_counts, aes(x = Date, y = Coughs, group = sid, color = sid)) +
  geom_point() +
  geom_line(alpha = 0.1) +
  theme(legend.position = "None")
```

```{r}
#| label: fig-coughs-by-night
#| fig-cap: Coughs by nights from start
ggplot(agg_counts, aes(x = Night, y = Coughs, group = sid, color = sid)) +
  geom_point() +
  geom_line(alpha = 0.1) +
  theme(legend.position = "None")
```

#

### 2024-10-09 Lilah Melzer/cough_monitor_analysis.qmd
Update cough_monitor_analysis.qmd

@@
       dark: darkly
-bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
+bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
 csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
@@
 library(lubridate)
+library(arsenal)
 home_dir <- switch(Sys.info()["sysname"],
@@
   filter(!is.na(Coughs)) %>%
-  mutate(Date = mdy(Date),
-    Coughs = as.numeric(Coughs))%>%
+  mutate(
+    Date = mdy(Date),
+    Coughs = as.numeric(Coughs)
+  ) %>%
   group_by(sid) %>%
@@
   ungroup()
+# Demographics data
+data$Diagnosis <- factor(sub(".*arm_*", "", data$redcap_event_name),
+  levels = 1:2,
+  labels = c("CF", "Non-CF")
+)
+demo <- data %>%
+  group_by(sid) %>%
+  filter(row_number() == 1) %>%
+  select(
+    sid, Diagnosis, age_at_consent, height, weight,
+    nih_sex.factor:nih_ethnicity.factor, mutation1.factor, mutation2.factor,
+    cftr_mod.factor, cftr_mod_type.factor, mucolytic.factor:antih.factor
+  )
+# Set table options for the document
+mycontrols <-
+  tableby.control(numeric.stats = c("Nmiss2", "meansd", "medianq1q3", "range"))
 ```
@@
 
-```{r}
+```{r results='asis'}
 #| label: tbl-mean-demo
 #| tbl-cap: Participant characteristics
+f <- as.formula(paste0("Diagnosis ~", paste0(colnames(demo), collapse = "+")))
+f <- update(f, . ~ . - Diagnosis - sid)
+t1 <- tableby(f, data = demo, control = mycontrols)
+summary(t1, pfootnote = T)
 ```
@@
 
-```{r}
+```{r results='asis'}
 #| label: tbl-mean-coughs
@@
 ```{r}
-#| label: fig-coughs-by-night
+#| label: fig-coughs-by-date
 #| fig-cap: Coughs by date
@@
 ```
-
-#

### 2024-10-09 Lilah Melzer/cough_monitor_analysis.qmd
Update cough_monitor_analysis.qmd

@@
   select(
-    sid, Diagnosis, age_at_consent, height, weight,
+    sid, Diagnosis, age_at_consent, height, weight, pre_fev,
     nih_sex.factor:nih_ethnicity.factor, mutation1.factor, mutation2.factor,
@@
   )
+# Calculate FEV1% - rspiro not accepting NA values for gender, which is really
+# annoying
+fev1 <- demo %>%
+  select(age_at_consent, height, nih_sex.factor, pre_fev) %>%
+  filter(nih_sex.factor %in% c("Male", "Female"))
+fev1$pre_fev_pp <- pctpred_GLIgl(fev1$age_at_consent, fev1$height / 100,
+  factor(fev1$nih_sex.factor, levels = c("Male", "Female")),
+  FEV1 = fev1$pre_fev
+)
+demo = left_join(demo,fev1[,c("sid","pre_fev_pp")])
 # Set table options for the document

### 2024-10-16 Lilah Melzer/cough_monitor_analysis.qmd
cough monitor update

--- 2024-10-09 Lilah Melzer/cough_monitor_analysis.qmd
@@
       dark: darkly
-bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
-csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
 editor: source
@@
 library(lubridate)
+library(arsenal)
+library(rspiro)
+library(lmerTest)
+library(performance)
 home_dir <- switch(Sys.info()["sysname"],
@@
   "Darwin" = "/Users/timvigers/GitHub",
-  "Windows" = "C:/Users/timvigers/GitHub",
+  "Windows" = "C:/Users/timvigers/Documents/GitHub",
   "Linux" = "/home/timvigers/GitHub"
@@
 #| message: false
+#| warning: false
 # Import REDCap data
 source(paste0(github_dir, "/CF/Lilah Melzer/cough_monitor_import.r"))
+# Import use data
+use <- read.csv("./Data_Clean/Cough study vacation days consent date.csv",
+  na.strings = ""
+)
 # Import counts
@@
 agg_counts$sid <- sub("FN", "", agg_counts$sid)
-# Wide to long, fix dates, count number of nights
+# Wide to long, fix dates
 agg_counts <- agg_counts %>%
-  pivot_longer(cols = -one_of("sid"), names_to = "Date", values_to = "Coughs") %>%
-  filter(!is.na(Coughs)) %>%
-  mutate(Date = mdy(Date),
-    Coughs = as.numeric(Coughs))%>%
+  pivot_longer(
+    cols = -one_of("sid"), names_to = "Date", values_to = "Coughs"
+  ) %>%
+  mutate(
+    Date = mdy(Date),
+    Coughs = as.numeric(Coughs)
+  )
+# Add start date
+agg_counts$sid <- as.numeric(agg_counts$sid)
+agg_counts$Start <-
+  mdy(use$Start.date[match(agg_counts$sid, use$Participant.number)])
+# Count number of nights
+agg_counts <- agg_counts %>%
+  filter(Date >= Start) %>%
   group_by(sid) %>%
-  mutate(
-    Night = row_number()
-  ) %>%
+  mutate(Night = row_number()) %>%
   ungroup()
+# Demographics data
+data$Diagnosis <- factor(sub(".*arm_*", "", data$redcap_event_name),
+  levels = 1:2,
+  labels = c("CF", "Non-CF")
+)
+demo <- data %>%
+  group_by(sid) %>%
+  filter(row_number() == 1) %>%
+  select(
+    sid, Diagnosis, age_at_consent, height, weight, pre_fev,
+    nih_sex.factor:nih_ethnicity.factor, mutation1.factor, mutation2.factor,
+    cftr_mod.factor, cftr_mod_type.factor, mucolytic.factor:antih.factor
+  )
+# Calculate FEV1% - rspiro not accepting NA values for gender, which is really
+# annoying
+fev1 <- demo %>%
+  select(age_at_consent, height, nih_sex.factor, pre_fev) %>%
+  filter(nih_sex.factor %in% c("Male", "Female"))
+fev1$pre_fev_pp <- pctpred_GLIgl(fev1$age_at_consent, fev1$height / 100,
+  factor(fev1$nih_sex.factor, levels = c("Male", "Female")),
+  FEV1 = fev1$pre_fev
+)
+demo <- left_join(demo, fev1[, c("sid", "pre_fev_pp")])
+label(demo$pre_fev_pp) <- "Pre-bronchodilator FEV1 % Predicted"
+# CF genotype
+demo$Genotype <-
+  factor(rowSums(demo[, c("mutation1.factor", "mutation2.factor")] == "F508del"),
+    levels = c(2, 1, 0),
+    labels = c("F508del homozygous", "F508del heterozygous", "No F508del")
+  )
+demo[, grep("mutation", colnames(demo))] <- NULL
+# Add group to counts
+agg_counts$Diagnosis <-
+  demo$Diagnosis[match(agg_counts$sid, as.numeric(as.character(demo$sid)))]
+# Per night
+per_night <- agg_counts %>%
+  group_by(sid) %>%
+  summarise(
+    median_coughs = median(Coughs, na.rm = T),
+    nights_data = sum(!is.na(Coughs))
+  )
+demo$sid <- as.numeric(demo$sid)
+demo <- left_join(demo, per_night)
+label(demo$median_coughs) <- "Median Coughs per Night"
+label(demo$nights_data) <- "Nights With Cough Data"
+# Set table options for the document
+mycontrols <-
+  tableby.control(numeric.stats = c("Nmiss", "medianq1q3", "range"),
+                  numeric.test = "kwt")
 ```
@@
 
-```{r}
+```{r results='asis'}
 #| label: tbl-mean-demo
 #| tbl-cap: Participant characteristics
+f <- as.formula(paste0("Diagnosis ~", paste0(colnames(demo), collapse = "+")))
+f <- update(f, . ~ . - Diagnosis - sid - median_coughs + kwt(median_coughs))
+t1 <- tableby(f, data = demo, control = mycontrols)
+summary(t1, pfootnote = T)
 ```
 
-# Coughs per night
+# Distribution of coughs
 
 ```{r}
-#| label: tbl-mean-coughs
-#| tbl-cap: Mean coughs per night
+#| label: fig-coughs-dist
+#| fig-cap: Histogram of Coughs Across All Nights
+#| warning: false
+ggplot(agg_counts, aes(x = Coughs)) +
+  geom_histogram(binwidth = 1)
 ```
+
+The distribution of coughs per night is extremely skewed, so coughs per night, etc. are reported as medians rather than means.
 
@@
 
-```{r}
-#| label: fig-coughs-by-night
-#| fig-cap: Coughs by date
-ggplot(agg_counts, aes(x = Date, y = Coughs, group = sid, color = sid)) +
-  geom_point() +
-  geom_line(alpha = 0.1) +
-  theme(legend.position = "None")
-```
+## Plot
 
@@
 #| fig-cap: Coughs by nights from start
-ggplot(agg_counts, aes(x = Night, y = Coughs, group = sid, color = sid)) +
-  geom_point() +
+#| warning: false
+ggplot(agg_counts, aes(
+  x = Night, y = Coughs, group = factor(sid), color = Diagnosis
+)) +
+  geom_point(alpha = 0.1) +
   geom_line(alpha = 0.1) +
-  theme(legend.position = "None")
+  geom_smooth(aes(group = Diagnosis))
 ```
 
-#
+## Model
+
+```{r}
+mod <- glmer(Coughs ~ Night * Diagnosis + (1 | sid),
+  data = agg_counts, family = "poisson"
+)
+gtsummary::tbl_regression(mod, exponentiate = T)
+
+mod <- glmer(Coughs ~ Diagnosis + (1 | sid),
+  data = agg_counts, family = "poisson"
+)
+gtsummary::tbl_regression(mod, exponentiate = T)
+```

### 2024-10-17 Lilah Melzer/cough_monitor_analysis.qmd
mixed model with random slope for Lilah

@@
 mycontrols <-
-  tableby.control(numeric.stats = c("Nmiss", "medianq1q3", "range"),
-                  numeric.test = "kwt")
+  tableby.control(
+    numeric.stats = c("Nmiss", "medianq1q3", "range"),
+    numeric.test = "kwt"
+  )
 ```
@@
 ```{r}
-mod <- glmer(Coughs ~ Night * Diagnosis + (1 | sid),
-  data = agg_counts, family = "poisson"
-)
-gtsummary::tbl_regression(mod, exponentiate = T)
-
-mod <- glmer(Coughs ~ Diagnosis + (1 | sid),
+mod <- glmer(Coughs ~ Night + Diagnosis + (Night | sid),
   data = agg_counts, family = "poisson"

### 2024-10-22 Scott Sagel/PEx in CF/analysis_for_yaron.qmd
minor updates for Scott

@@
       dark: darkly
-bibliography: /home/tim/Documents/Miscellaneous/zotero.bib
-csl: /home/tim/GitHub/styles/american-medical-association.csl
 editor: source
@@
 library(rspiro)
-knitr::opts_knit$set(root.dir = "/home/tim/OneDrive/Vigers/CF/Scott Sagel/PEx in CF")
+library(patchwork)
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Scott Sagel/PEx in CF",
+  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Scott Sagel/PEx in CF",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Scott Sagel/PEx in CF"
+)
+github_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/GitHub",
+  "Windows" = "C:/Users/timvigers/GitHub",
+  "Linux" = "/home/timvigers/GitHub"
+)
+knitr::opts_knit$set(root.dir = home_dir)
 # Colors for plots
@@
   }
+)
+df$Insurance <- factor(df$Insurance,
+  levels = c(
+    "Medicaid", "Medicaid/None", "Medicaid/Other", "Medicare", "Medicare/Other",
+    "None", "Other", "Private insurance", "Private insurance/Medicaid",
+    "Private insurance/Other",
+    "Private insurance/Tricare/other military health plan",
+    "Tricare/other military health plan"
+  ),
+  labels = c(
+    "Medicaid", "Medicaid", "Medicaid", "Medicare", "Medicare", "None/Other",
+    "None/Other", "Private insurance", "Private insurance", "Private insurance",
+    "Tricare/other military health plan", "Tricare/other military health plan"
+  )
+)
+df$admit_cx_source <- factor(df$admit_cx_source,
+  levels = 1:4,
+  labels = c(
+    "Bronchoscopy", "Expectorated sputum", "Induced sputum", "Throat swab"
+  )
+)
+# Combine species
+df$any_fungi <- apply(df, 1, function(r) {
+  r <- as.numeric(r[paste0("fungi___", 1:5)])
+  any(r == 1)
+})
+df$any_fungi <- factor(df$any_fungi, levels = c(F, T), labels = c("No", "Yes"))
+df$any_ntm <- apply(df, 1, function(r) {
+  r <- as.numeric(r[paste0("ntm___", 1:3)])
+  any(r == 1)
+})
+df$any_ntm <- factor(df$any_ntm, levels = c(F, T), labels = c("No", "Yes"))
+df$any_virus <- apply(df, 1, function(r) {
+  r <- as.numeric(r[paste0("viruses___", 1:5)])
+  any(r == 1)
+})
+df$any_virus <- factor(df$any_virus, levels = c(F, T), labels = c("No", "Yes"))
+df$any_bact <- apply(df, 1, function(r) {
+  r <- as.numeric(r[paste0("admit_cx_results___", 1:9)])
+  any(r == 1)
+})
+df$any_bact <- factor(df$any_bact, levels = c(F, T), labels = c("No", "Yes"))
+
+df$no_bact <- apply(df, 1, function(r) {
+  r <- as.numeric(r[paste0("admit_cx_results___", 1:8)])
+  all(r == 0)
+})
+df$no_bact <- factor(df$no_bact, levels = c(F, T), labels = c("No", "Yes"))
+# Change in FEV1 from admission to discharge
+df$ppfev1_change <- df$ppfev1_discharge - df$ppfev1_admit
+# Format fungi, NTM, etc. columns
+df <- df %>%
+  mutate(
+    across(
+      c(
+        fungi___1:fungi___5, ntm___1:ntm___3, viruses___1:viruses___5,
+        admit_cx_results___1:admit_cx_results___9
+      ),
+      ~ factor(as.numeric(.x),
+        levels = 0:1,
+        labels = c("Not detected", "Detected")
+      )
+    )
   )
@@
   "Other fungus" = "fungi___5",
+  "Any fungi?" = "any_fungi",
   "NTM testing performed?" = "ntm_test",
@@
   "Other virus" = "viruses___5",
+  "Any virus?" = "any_virus",
   "M. abscessus species" = "ntm___1",
@@
   "Other NTM species" = "ntm___3",
+  "Any NTM?" = "any_ntm",
   "A. xylosoxidans" = "admit_cx_results___1",
@@
   "Other bacteria" = "admit_cx_results___9",
+  "Any bacteria?" = "any_bact",
+  "No CF bacteria" = "no_bact",
   "On CFTR Modulator?" = "cftr_mod",
   "CFTR Modulator Type" = "cftr_mod_spec",
-  "Cultures Tested" = "redcap_data_access_group",
+  "Site" = "redcap_data_access_group",
   "Length of Admission" = "days_admit",
@@
   "IV Steroids" = "steroids_iv",
-  "Baseline FEV1pp (2022)" = "ppfev1_baseline",
-  "Admission FEV1pp (2022)" = "ppfev1_admit",
-  "Discharge FEV1pp (2022)" = "ppfev1_discharge",
+  "Baseline FEV1pp" = "ppfev1_baseline",
+  "Admission FEV1pp" = "ppfev1_admit",
+  "Discharge FEV1pp" = "ppfev1_discharge",
   "Change in mean FEV1pp from admit to discharge" = "ppfev1_change"
@@
     Trikafta, azithro, inhaled_abx, steroids_system, admit_cx_source,
-    ppfev1_baseline, ppfev1_admit, ppfev1_discharge, days_admit,
-    fungal_test, fungi___1:fungi___5,
-    ntm_test, ntm___1:ntm___3,
-    viral_test, viruses___1:viruses___5,
-    admit_cx_results___1:admit_cx_results___9,
+    ppfev1_baseline, ppfev1_admit, ppfev1_discharge, ppfev1_change,
+    days_admit, fungal_test, fungi___1:fungi___5, any_fungi,
+    ntm_test, ntm___1:ntm___3, any_ntm,
+    viral_test, viruses___1:viruses___5, any_virus,
+    admit_cx_results___1:admit_cx_results___9, any_bact, no_bact,
     contains("steroids")
@@
   filter(!is.na(year))
-# Write for Scott and Yaron to check
-write.csv(df,file = "./Data_Clean/analysis_dataset.csv",row.names = F,na = "")
 ```
@@
 #| label: tbl-table-1
-#| tbl-cap: Participant characteristics at admission
+#| tbl-cap: Participant characteristics overall
+df %>%
+  group_by(sid) %>%
+  filter(row_number() == 1) %>%
+  ungroup() %>%
+  select(redcap_data_access_group, sex, race, ethnicity, Genotype) %>%
+  tbl_summary(missing_text = "(Missing)", label = labels)
+```
+
+```{r}
+#| label: tbl-table-1-year
+#| tbl-cap: Participant characteristics by year
+df %>%
+  group_by(sid, year) %>%
+  filter(row_number() == 1) %>%
+  ungroup() %>%
+  select(year, redcap_data_access_group, sex, race, ethnicity, Genotype) %>%
+  tbl_summary(by = year, missing_text = "(Missing)", label = labels)
+```
+
+```{r}
+#| label: tbl-table-1-indep
+#| tbl-cap: Participant characteristics at admission by year
 df %>%
@@
         all_continuous() ~ "{median} [{p25}, {p75}]",
-        c(age_at_admit, ppfev1_baseline, ppfev1_admit, ppfev1_discharge) ~
+        c(
+          age_at_admit, ppfev1_baseline, ppfev1_admit, ppfev1_discharge,
+          ppfev1_change
+        ) ~
           "{mean} ({sd})"
@@
   filter(fungal_test == "Yes") %>%
-  select(year, redcap_data_access_group, fungi___1:fungi___5) %>%
+  select(year, redcap_data_access_group, fungi___1:fungi___5, any_fungi) %>%
   tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
@@
   filter(fungal_test == "Yes") %>%
-  select(year, redcap_data_access_group, fungi___1:fungi___5) %>%
+  select(year, redcap_data_access_group, fungi___1:fungi___5, any_fungi) %>%
   tbl_strata(
@@
   filter(fungal_test == "Yes") %>%
-  select(year, cftr_mod, fungi___1:fungi___5) %>%
+  select(year, cftr_mod, fungi___1:fungi___5, any_fungi) %>%
   tbl_strata(
@@
   filter(fungal_test == "Yes") %>%
-  select(year, cftr_mod_spec, fungi___1:fungi___5) %>%
+  select(year, cftr_mod_spec, fungi___1:fungi___5, any_fungi) %>%
   tbl_strata(
@@
   filter(fungal_test == "Yes") %>%
-  select(year, Trikafta, fungi___1:fungi___5) %>%
+  select(year, Trikafta, fungi___1:fungi___5, any_fungi) %>%
   tbl_strata(
@@
   filter(ntm_test == "Yes") %>%
-  select(year, redcap_data_access_group, ntm___1:ntm___3) %>%
+  select(year, redcap_data_access_group, ntm___1:ntm___3, any_ntm) %>%
   tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
@@
   filter(ntm_test == "Yes") %>%
-  select(year, redcap_data_access_group, ntm___1:ntm___3) %>%
+  select(year, redcap_data_access_group, ntm___1:ntm___3, any_ntm) %>%
   tbl_strata(
@@
   filter(ntm_test == "Yes") %>%
-  select(year, cftr_mod, ntm___1:ntm___3) %>%
+  select(year, cftr_mod, ntm___1:ntm___3, any_ntm) %>%
   tbl_strata(
@@
   filter(ntm_test == "Yes") %>%
-  select(year, cftr_mod_spec, ntm___1:ntm___3) %>%
+  select(year, cftr_mod_spec, ntm___1:ntm___3, any_ntm) %>%
   tbl_strata(
@@
   filter(ntm_test == "Yes") %>%
-  select(year, Trikafta, ntm___1:ntm___3) %>%
+  select(year, Trikafta, ntm___1:ntm___3, any_ntm) %>%
   tbl_strata(
@@
   filter(viral_test == "Yes") %>%
-  select(year, redcap_data_access_group, viruses___1:viruses___5) %>%
+  select(year, redcap_data_access_group, viruses___1:viruses___5, any_virus) %>%
   tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
@@
   filter(viral_test == "Yes") %>%
-  select(year, cftr_mod, viruses___1:viruses___5) %>%
+  select(year, cftr_mod, viruses___1:viruses___5, any_virus) %>%
   tbl_strata(
@@
   filter(viral_test == "Yes") %>%
-  select(year, cftr_mod_spec, viruses___1:viruses___5) %>%
+  select(year, cftr_mod_spec, viruses___1:viruses___5, any_virus) %>%
   tbl_strata(
@@
   filter(viral_test == "Yes") %>%
-  select(year, Trikafta, viruses___1:viruses___5) %>%
+  select(year, Trikafta, viruses___1:viruses___5, any_virus) %>%
   tbl_strata(
@@
     year, redcap_data_access_group,
-    admit_cx_results___1:admit_cx_results___9
+    admit_cx_results___1:admit_cx_results___9, any_bact, no_bact
   ) %>%
@@
   select(
-    year, redcap_data_access_group, admit_cx_results___1:admit_cx_results___9
+    year, redcap_data_access_group, admit_cx_results___1:admit_cx_results___9,
+    any_bact, no_bact
   ) %>%
@@
 df %>%
-  select(year, cftr_mod, admit_cx_results___1:admit_cx_results___9) %>%
+  select(
+    year, cftr_mod, admit_cx_results___1:admit_cx_results___9, any_bact, no_bact
+  ) %>%
   tbl_strata(
@@
 df %>%
-  select(year, cftr_mod_spec, admit_cx_results___1:admit_cx_results___9) %>%
+  select(
+    year, cftr_mod_spec, admit_cx_results___1:admit_cx_results___9, any_bact,
+    no_bact
+  ) %>%
   tbl_strata(
@@
 df %>%
-  select(year, Trikafta, admit_cx_results___1:admit_cx_results___9) %>%
+  select(
+    year, Trikafta, admit_cx_results___1:admit_cx_results___9, any_bact, no_bact
+  ) %>%
   tbl_strata(
@@
 #| fig-cap: Number of hospitalizations
-df %>%
+fig1 <- df %>%
   group_by(year) %>%
@@
   scale_fill_manual(values = c(purple, yellow))
+fig1
 ```
@@
 #| fig-cap: Percent hospitalized 2 or more times in both years
-df %>%
+fig2 <- df %>%
   group_by(sid, year) %>%
@@
   ) +
-  xlab("Site") +
   ylab("Percent With >= 2 Hospitalizations") +
   theme_bw() +
-  theme(legend.position = "none") +
+  theme(legend.position = "none", axis.title.x = element_blank()) +
   scale_fill_manual(values = c(purple, yellow))
+fig2
+```
+
+```{r}
+#| label: fig-panel
+#| fig-cap: Hospitalizations
+fig1 + fig2
 ```
@@
 
-Hospitalizations within a person were considered independent in the table below.
+Hospitalizations within a person were considered independent in the tables below.
 
@@
 df %>%
-  mutate(ppfev1_change = ppfev1_discharge - ppfev1_admit) %>%
   filter(year == 2022) %>%

### 2024-10-23 Scott Sagel/PEx in CF/analysis_for_yaron.qmd
table 2 p values

@@
   select(year, redcap_data_access_group, sex, race, ethnicity, Genotype) %>%
-  tbl_summary(by = year, missing_text = "(Missing)", label = labels)
+  tbl_summary(by = year, missing_text = "(Missing)", label = labels) %>%
+  add_p() %>%
+  separate_p_footnotes()
 ```

### 2024-10-24 Lilah Melzer/cough_monitor_analysis.qmd
updates for lilah

@@
 library(rspiro)
+library(childsds)
 library(lmerTest)
@@
 # Import use data
-use <- read.csv("./Data_Clean/Cough study vacation days consent date.csv",
-  na.strings = ""
-)
+start_end <- read.csv("./Data_Clean/Start and end dates.csv", na.strings = "")
+start_end$Start <- mdy(start_end$Start)
+start_end$End <- mdy(start_end$End)
+vacations <- read.csv("./Data_Clean/Vacation days.csv", na.strings = "")
+vacations$Vacation.Start <- mdy(vacations$Vacation.Start)
+vacations$Vacation.End <- mdy(vacations$Vacation.End)
 # Import counts
@@
   )
-# Add start date
+# Add start and end dates, filter
 agg_counts$sid <- as.numeric(agg_counts$sid)
-agg_counts$Start <-
-  mdy(use$Start.date[match(agg_counts$sid, use$Participant.number)])
+agg_counts <- left_join(agg_counts, start_end)
+agg_counts <- agg_counts %>%
+  filter(Date >= Start & Date <= End)
+# Vacations
+agg_counts$Vacation <- apply(agg_counts, 1, function(r) {
+  id <- as.numeric(r["sid"])
+  d <- r["Date"]
+  any(d >= vacations$Vacation.Start[vacations$sid == id] &
+    d <= vacations$Vacation.End[vacations$sid == id])
+})
+agg_counts <- agg_counts %>%
+  filter(!Vacation) %>%
+  arrange(sid, Date)
+# Sick
+sick <- data %>%
+  select(sid, pulm_ex_v2.factor) %>%
+  mutate(
+    antib_dt = mdy(antib_dt), 
+    px_start = antib_dt - 7,
+    px_end = antib_dt + 28
+  )
 # Count number of nights
 agg_counts <- agg_counts %>%
-  filter(Date >= Start) %>%
   group_by(sid) %>%
@@
 fev1 <- demo %>%
-  select(age_at_consent, height, nih_sex.factor, pre_fev) %>%
+  select(age_at_consent, height, weight, nih_sex.factor, pre_fev) %>%
   filter(nih_sex.factor %in% c("Male", "Female"))
@@
 )
-demo <- left_join(demo, fev1[, c("sid", "pre_fev_pp")])
+fev1$height_perc <- sds(as.numeric(fev1$height),
+  age = fev1$age_at_consent, sex = fev1$nih_sex.factor,
+  male = "Male", female = "Female", item = "height2_20", ref = cdc.ref,
+  type = "perc"
+) * 100
+fev1$weight_perc <- sds(as.numeric(fev1$weight),
+  age = fev1$age_at_consent, sex = fev1$nih_sex.factor,
+  male = "Male", female = "Female", item = "weight2_20", ref = cdc.ref,
+  type = "perc"
+) * 100
+demo <-
+  left_join(demo, fev1[, c("sid", "height_perc", "weight_perc", "pre_fev_pp")])
 label(demo$pre_fev_pp) <- "Pre-bronchodilator FEV1 % Predicted"
+label(demo$height_perc) <- "Height %ile"
+label(demo$weight_perc) <- "Weight %ile"
 # CF genotype
@@
 label(demo$nights_data) <- "Nights With Cough Data"
+# Order columns
+demo <- demo %>%
+  select(sid, height_perc, weight_perc, pre_fev_pp, everything()) %>%
+  select(-height, -weight, -pre_fev)
+# Exclude vacation days
+
 # Set table options for the document

### 2024-10-24 Lilah Melzer/cough_monitor_import.r

@@ -4,7 +4,7 @@ home_dir <- switch(Sys.info()["sysname"],
   "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring"
 )
 # Read Data
-data <- read.csv(paste0(home_dir, "/Data_Raw/CoughMonitor230382_DATA_2024-10-09_1237.csv"), na.strings = "")
+data <- read.csv(paste0(home_dir, "/Data_Raw/CoughMonitor230382_DATA_2024-10-23_1317.csv"), na.strings = "")
 # Setting Labels
 label(data$sid) <- "Subject ID"
 label(data$redcap_event_name) <- "Event Name"

### 2024-10-24 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
stratified by age group

--- 2024-08-05 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
@@
 # Libraries
+library(Hmisc)
 library(tidyverse)
@@
   "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/tim/CF"
+  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
 )
@@
 
-The following models are adjusted for age group and insurance status. We fit a Gaussian mixed model for continuous outcomes with random intercept for each subject.
+```{r}
+# Create change in slope variable
+encounter$slope_change_days <- (encounter$Days > 0) * encounter$Days
+label(encounter$slope_change_days) <- "Change in Slope"
+encounter$Years <- encounter$Days / 365.25
+encounter$slope_change_years <- (encounter$Years > 0) * encounter$Years
+label(encounter$slope_change_years) <- "Change in Slope"
+```
 
-```{r results='asis'}
-#| message: false
-invisible(lapply(continuous_outcomes[1:3], function(o) {
-  # Formula
-  f <-
-    as.formula(paste0(
-      o,
-      "~ bs(Days, knots = 0, degree = 1) * Race + age_group + Insurance + (1 | eDWID)"
-    ))
-  m <- lmer(f, data = encounter)
-  # Print table
-  cat("\n")
-  cat("\n")
-  cat("##", o)
-  cat("\n")
-  cat("\n")
-  cat("### Model results")
-  cat("\n")
-  cat("\n")
-  m %>%
-    tbl_regression() %>%
-    print()
-  cat("\n")
-  cat("\n")
-  # Plot
-  cat("\n")
-  cat("\n")
-  cat("## Model predictions")
-  cat("\n")
-  cat("\n")
-  pred <- predict_response(m, terms = c("Days[-1095,0,730]", "Race"))
-  pred %>% print()
-  cat("\n")
-  cat("\n")
-  cat("### Plot")
-  cat("\n")
-  cat("\n")
-  pred %>%
-    plot() %>%
-    print()
-  cat("\n")
-  cat("\n")
-}))
+The following models are adjusted for age group and insurance status.
+We fit a Gaussian mixed model for continuous outcomes with random intercept for each subject (random slope did not improve the model AIC).
+
+## FEV1 % predicted
+
+### Full cohort
+
+#### GLI 2022 equations
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  gli_fev1_ppred_rn ~ Years * Race +
+    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  data = encounter
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = gli_fev1_ppred_rn, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
 ```
+
+#### GLI 2012 equations
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  GLI_FEV1_pct_predicted ~ Years * Race +
+    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  data = encounter
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = GLI_FEV1_pct_predicted, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+### Age < 18 at ETI initiation
+
+The following models were not adjusted for age group.
+
+#### GLI 2022 equations
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race +
+    Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = gli_fev1_ppred_rn, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+#### GLI 2012 equations
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  GLI_FEV1_pct_predicted ~ Years * Race +
+    slope_change_years * Race + Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = GLI_FEV1_pct_predicted, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+### Age >= 18 at ETI initiation
+
+The following models were not adjusted for age group.
+
+#### GLI 2022 equations
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  gli_fev1_ppred_rn ~ Years * Race +
+    slope_change_years * Race + Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = gli_fev1_ppred_rn, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+#### GLI 2012 equations
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  GLI_FEV1_pct_predicted ~ Years * Race +
+    slope_change_years * Race + Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = GLI_FEV1_pct_predicted, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+## FVC % predicted
+
+### Full cohort
+
+#### GLI 2022 equations
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  gli_fvc_ppred_rn ~ Years * Race +
+    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  data = encounter
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = gli_fvc_ppred_rn, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+#### GLI 2012 equations
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  GLI_FVC_pct_predicted ~ Years * Race +
+    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  data = encounter
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = GLI_FVC_pct_predicted, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+### Age < 18 at ETI initiation
+
+The following models were not adjusted for age group.
+
+#### GLI 2022 equations
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  gli_fvc_ppred_rn ~ Years * Race + slope_change_years * Race +
+    Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = gli_fvc_ppred_rn, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+#### GLI 2012 equations
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  GLI_FVC_pct_predicted ~ Years * Race +
+    slope_change_years * Race + Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = GLI_FVC_pct_predicted, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+### Age >= 18 at ETI initiation
+
+The following models were not adjusted for age group.
+
+#### GLI 2022 equations
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  gli_fvc_ppred_rn ~ Years * Race +
+    slope_change_years * Race + Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = gli_fvc_ppred_rn, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+#### GLI 2012 equations
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  GLI_FVC_pct_predicted ~ Years * Race +
+    slope_change_years * Race + Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = GLI_FVC_pct_predicted, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+## FEV1/FVC % predicted
+
+### Full cohort
+
+#### GLI 2022 equations
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  gli_fev1fvc_ppred_rn ~ Years * Race +
+    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  data = encounter
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = gli_fev1fvc_ppred_rn, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+#### GLI 2012 equations
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  GLI_FEV1FVC_pct_predicted ~ Years * Race +
+    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  data = encounter
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = GLI_FEV1FVC_pct_predicted, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+### Age < 18 at ETI initiation
+
+The following models were not adjusted for age group.
+
+#### GLI 2022 equations
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  gli_fev1fvc_ppred_rn ~ Years * Race + slope_change_years * Race +
+    Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = gli_fev1fvc_ppred_rn, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+#### GLI 2012 equations
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  GLI_FEV1FVC_pct_predicted ~ Years * Race +
+    slope_change_years * Race + Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = GLI_FEV1FVC_pct_predicted, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+### Age >= 18 at ETI initiation
+
+The following models were not adjusted for age group.
+
+#### GLI 2022 equations
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  gli_fev1fvc_ppred_rn ~ Years * Race +
+    slope_change_years * Race + Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = gli_fev1fvc_ppred_rn, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+#### GLI 2012 equations
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  GLI_FEV1FVC_pct_predicted ~ Years * Race +
+    slope_change_years * Race + Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = GLI_FEV1FVC_pct_predicted, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+## BMI %ile
+
+### Full cohort
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  bmipercentile ~ Years * Race +
+    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  data = encounter
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = bmipercentile, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+### Age < 18 at ETI initiation
+
+The following models were not adjusted for age group.
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  bmipercentile ~ Years * Race + slope_change_years * Race +
+    Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = bmipercentile, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+### Age >= 18 at ETI initiation
+
+The following models were not adjusted for age group.
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  bmipercentile ~ Years * Race +
+    slope_change_years * Race + Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = bmipercentile, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+## Height %ile
+
+### Full cohort
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  heightpercentile ~ Years * Race +
+    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  data = encounter
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = heightpercentile, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+### Age < 18 at ETI initiation
+
+The following models were not adjusted for age group.
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  heightpercentile ~ Years * Race + slope_change_years * Race +
+    Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = heightpercentile, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+### Age >= 18 at ETI initiation
+
+The following models were not adjusted for age group.
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  heightpercentile ~ Years * Race +
+    slope_change_years * Race + Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = heightpercentile, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+## Weight %ile
+
+### Full cohort
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  weightpercentile ~ Years * Race +
+    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  data = encounter
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = weightpercentile, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+### Age < 18 at ETI initiation
+
+The following models were not adjusted for age group.
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  weightpercentile ~ Years * Race + slope_change_years * Race +
+    Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = weightpercentile, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```
+
+### Age >= 18 at ETI initiation
+
+The following models were not adjusted for age group.
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  weightpercentile ~ Years * Race +
+    slope_change_years * Race + Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
+)
+# Model plot
+mod_df <- change_mod@frame
+mod_df$ind_pred <- predict(change_mod)
+mod_df$pop_pred <- predict(change_mod, re.form = NA)
+ggplot(mod_df, aes(x = Years)) +
+  geom_point(aes(y = weightpercentile, group = eDWID),
+    alpha = 0.01, show.legend = F, shape = "."
+  ) +
+  geom_line(aes(y = ind_pred, group = eDWID),
+    alpha = 0.01, show.legend = F
+  ) +
+  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  xlab("Years from ETI Start") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression()
+```

### 2024-10-24 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R

@@ -122,6 +122,11 @@ encounter$Modulator_trikafta_first_date <-
 encounter$encounterdate <- mdy(encounter$encounterdate)
 encounter$Days <- as.numeric(encounter$encounterdate -
   encounter$Modulator_trikafta_first_date)
+encounter$age_eti_start <-
+  round(encounter$encounterage - (encounter$Days / 365.25), 2)
+encounter$age_eti_group <- cut(encounter$age_eti_start,
+  breaks = c(-Inf, 18, Inf), right = FALSE
+)
 # Hospitalization or home IVs
 hospitalizations$CareEpi_StartDt <- mdy(hospitalizations$CareEpi_StartDt)
 hospitalizations$CareEpi_EndDt <- mdy(hospitalizations$CareEpi_EndDt)

### 2024-10-24 Christine Chan/CF Registry/cf_registry_a1c_CFRD_analysis.qmd
Create cf_registry_a1c_CFRD_analysis.qmd

--- 2024-09-23 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
@@
 ---
-title: "CF Registry Analysis - ROC"
+title: "HbA1c and CFRD"
 author: "Tim Vigers"
@@
 #| include: false
-library(dplyr)
-library(tidyr)
-library(ggplot2)
-library(gtsummary)
-library(gt)
-library(pROC)
-library(ggrepel)
+library(tidyverse)
 home_dir <- switch(Sys.info()["sysname"],
@@
 knitr::opts_knit$set(root.dir = home_dir)
-set.seed(1017)
 ```
@@
 # Import data
-encounter <- read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv",
+annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
   na.strings = c("", "N/A")
@@
 )
-# Age at year end
-annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
-  na.strings = c("", "N/A")
-)
-annual <- annual %>% rename(reviewyear = ReviewYear)
 # Merge
-encounter <- left_join(
-  encounter,
+annual <- left_join(
+  annual,
   demo %>% select(eDWID, Race, Hispanicrace, Gender, F508, MutClass),
@@
 )
-encounter <- left_join(
-  encounter,
-  annual %>% select(eDWID, reviewyear, Age_YrEnd),
-  by = join_by(eDWID, reviewyear)
+# CFRD by OGTT two hour
+annual$cfrd_ogtt <- factor(annual$A_twohour_bloodglucose >= 200,
+  levels = c(F, T), labels = c("No", "Yes")
 )
-# Limit to those with two hour OGTT and HbA1c at the same encounter
-# Also remove years before 2010 and those under age 10
-roc_df <- encounter %>%
-  arrange(eDWID, reviewyear, encounternum) %>%
+# CFRD by registry definition
+annual$cfrd_registry <- factor(annual$cfrd_status_annual,
+  levels = c(1:5), labels = c("No", "No", "Yes", "No", "No")
+)
+# Assume that missing data for CFRD means "No" (some of these will get filled in
+# later)
+annual$cfrd_registry[is.na(annual$cfrd_registry)] <- "No"
+annual$cfrd_ogtt[is.na(annual$cfrd_ogtt)] <- "No"
+# Select columns
+annual <- annual %>%
   select(
-    eDWID, reviewyear, Age_YrEnd, Race, Hispanicrace, Gender, F508, MutClass,
-    twohour_bloodglucose, hgba1c
+    eDWID, ReviewYear, Race, Hispanicrace, Gender, F508, MutClass,
+    Age_YrEnd, A_bmipercentile, A_hgba1c, cfrd_registry, cfrd_status_annual,
+    cfrd_ogtt, A_twohour_bloodglucose
   ) %>%
-  filter(reviewyear >= 2010, Age_YrEnd >= 10, hgba1c >= 3.5 & hgba1c <= 15) %>%
-  drop_na() %>%
   group_by(eDWID) %>%
-  filter(row_number() == n()) %>%
-  ungroup() %>%
   mutate(
-    `OGTT 2-hour` = cut(twohour_bloodglucose,
-      breaks = c(-Inf, 140, 200, Inf),
-      labels = c("<140", "[140,200)", ">=200")
-    ),
-    `Age Group` = cut(Age_YrEnd, breaks = c(10, 18, 30, Inf), right = F)
+    cfrd_registry_age = first(na.omit(Age_YrEnd[cfrd_registry == "Yes"])),
+    cfrd_ogtt_age = first(na.omit(Age_YrEnd[cfrd_ogtt == "Yes"]))
   )
-# Metrics to report
-metrics <- c(
-  "thresh", "specificity", "sensitivity", "accuracy", "tn", "tp", "fn",
-  "fp", "npv", "ppv"
-)
+# Fill in CFRD diagnoses
+annual$cfrd_registry_age[is.na(annual$cfrd_registry_age)] <- Inf
+annual$cfrd_ogtt_age[is.na(annual$cfrd_ogtt_age)] <- Inf
+annual$cfrd_registry[annual$Age_YrEnd < annual$cfrd_registry_age] <- "No"
+annual$cfrd_registry[annual$Age_YrEnd >= annual$cfrd_registry_age] <- "Yes"
+annual$cfrd_ogtt[annual$Age_YrEnd < annual$cfrd_ogtt_age] <- "No"
+annual$cfrd_ogtt[annual$Age_YrEnd >= annual$cfrd_ogtt_age] <- "Yes"
+# Split into one dataset for registry and one for OGTT
+registry <- annual %>%
+  filter(Age_YrEnd <= cfrd_registry_age) %>%
+  select(eDWID:A_hgba1c, cfrd_registry)
+ogtt <- annual %>%
+  filter(Age_YrEnd <= cfrd_ogtt_age) %>%
+  select(eDWID:A_hgba1c, cfrd_ogtt, cfrd_ogtt_age)
 ```

### 2024-10-24 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
Update cf_registry_survival_analysis.qmd

@@
     toc: true
-    toc-depth: 5
+    toc-depth: 10
     toc-float: true
@@
   "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
-  "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/CF Registry"
+  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
 )
@@
   tbl_regression()
-ggcoef_model(cox_res, show_p_values = F, signif_stars = F, colour = F)
+ggcoef_table(cox_res, show_p_values = F, add_reference_rows = T, signif_stars = F, colour = F)
 # Diagnostics
@@
   tbl_regression()
-ggcoef_model(vx770_cox_res,
-  show_p_values = F, signif_stars = F, colour = F
+ggcoef_table(vx770_cox_res,
+  show_p_values = F, add_reference_rows = T, signif_stars = F, colour = F
 )
@@
   tbl_regression()
-ggcoef_model(vx809comb_cox_res,
-  show_p_values = F, signif_stars = F, colour = F
+ggcoef_table(vx809comb_cox_res,
+  show_p_values = F, add_reference_rows = T, signif_stars = F, colour = F
 )
@@
   cat("\n")
+  cat("\n")
   cat("####", v)
@@
   # Plot base model (without time varying)
-  ggcoef_model(cox_res_split,
-    show_p_values = F, add_reference_rows = F, signif_stars = F, colour = F,
-    include = xvars[-which(xvars == v)]
+  ggcoef_table(cox_res_split,
+    show_p_values = F, add_reference_rows = T, signif_stars = F, colour = F
   ) %>%
@@
   cat("\n")
-  # Print EMMs
-  print(kable(emm, digits = 3))
-  cat("\n")
-  cat("\n")
-  # Plot marginal means
-  tvfp <- time_varying_forest_plot(emm,
-    x = "emmean", var = v,
-    var_label = labels[v]
-  ) +
-    theme(text = element_text(size = 20))
-  print(tvfp)
-  ggsave(
-    paste0(
-      "./Dissemination/Figures/", v,
-      "_time_varying_forest_plot.png"
-    ),
-    plot = tvfp, width = 12, height = 9, units = "in"
-  )
-  cat("\n")
-  cat("\n")
 }))

### 2024-10-24 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
Update racial_and_ethnic_differences_in_eti_outcomes.qmd

@@
 )
+# EMMs
+emm = emmeans(change_mod,spec = ~Years+Race)
+
 # Model plot
@@
 mod_df$ind_pred <- predict(change_mod)
-mod_df$pop_pred <- predict(change_mod, re.form = NA)
+pop_dat = mod_df %>% select(Years:Insurance)
+pop_dat$pred <- predict(change_mod, re.form = NA,newdata = pop_dat)
+pop_dat = pop_dat%>% distinct()
 ggplot(mod_df, aes(x = Years)) +
@@
   ) +
-  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  geom_line(data = pop_dat,aes(y = pred, group = Race, colour = Race)) +
   xlab("Years from ETI Start") +

### 2024-10-28 Christine Chan/CF Registry/cf_registry_a1c_CFRD_analysis.qmd
updated sweat test data for Stacey

--- 2024-09-19 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
@@
 #| include: false
-library(dplyr)
-library(tidyr)
-library(ggplot2)
+library(tidyverse)
+library(arsenal)
+library(survival)
+library(ggsurvfit)
+library(ggstats)
 library(gtsummary)
-library(gt)
-library(pROC)
-library(ggrepel)
 home_dir <- switch(Sys.info()["sysname"],
   "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
+  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
   "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/CF Registry"
@@
 )
-encounter <- left_join(
-  encounter,
-  annual %>% select(eDWID, reviewyear, Age_YrEnd),
-  by = join_by(eDWID, reviewyear)
+# CFRD by OGTT two hour
+annual$cfrd_ogtt <- factor(annual$A_twohour_bloodglucose >= 200,
+  levels = c(F, T), labels = c("CFRD-", "CFRD+")
 )
-# Limit to those with two hour OGTT and HbA1c at the same encounter
-# Also remove years before 2010 and those under age 10
-roc_df <- encounter %>%
-  arrange(eDWID, reviewyear, encounternum) %>%
+# CFRD by registry definition
+annual$cfrd_registry <- factor(annual$cfrd_status_annual,
+  levels = c(1:5), labels = c("CFRD-", "CFRD-", "CFRD+", "CFRD-", "CFRD-")
+)
+# Assume that missing data for CFRD means "CFRD-" (some of these will get filled in
+# later)
+annual$cfrd_registry[is.na(annual$cfrd_registry)] <- "CFRD-"
+annual$cfrd_ogtt[is.na(annual$cfrd_ogtt)] <- "CFRD-"
+# Modulators
+annual$A_Vx770 <- factor(annual$A_Vx770, levels = 0:1, labels = c("No", "Yes"))
+annual$A_VX809comb <- factor(annual$A_VX809comb,
+  levels = 0:1, labels = c("No", "Yes")
+)
+# Select columns
+annual <- annual %>%
   select(
-    eDWID, reviewyear, Age_YrEnd, Race, Hispanicrace, Gender, F508, MutClass,
-    twohour_bloodglucose, hgba1c
+    eDWID, ReviewYear, Race, Hispanicrace, Gender, F508, MutClass,
+    A_Vx770, A_VX809comb, Age_YrEnd, A_bmipercentile, A_hgba1c,
+    cfrd_registry, cfrd_status_annual, cfrd_ogtt, A_twohour_bloodglucose
   ) %>%
-  filter(reviewyear >= 2010, Age_YrEnd >= 10, hgba1c >= 3.5 & hgba1c <= 15) %>%
-  drop_na() %>%
   group_by(eDWID) %>%
-  filter(row_number() == n()) %>%
-  ungroup() %>%
   mutate(
-    `OGTT 2-hour` = cut(twohour_bloodglucose,
-      breaks = c(-Inf, 140, 200, Inf),
-      labels = c("<140", "[140,200)", ">=200")
-    ),
-    `Age Group` = cut(Age_YrEnd, breaks = c(10, 18, 30, Inf), right = F)
+    Age_YrEnd = round(Age_YrEnd, 2),
+    cfrd_registry_age = first(na.omit(Age_YrEnd[cfrd_registry == "CFRD+"])),
+    cfrd_ogtt_age = first(na.omit(Age_YrEnd[cfrd_ogtt == "CFRD+"]))
   )
-# Metrics to report
-metrics <- c(
-  "thresh", "specificity", "sensitivity", "accuracy", "tn", "tp", "fn",
-  "fp", "npv", "ppv"
+# Fill in CFRD diagnoses
+annual$cfrd_registry_age[is.na(annual$cfrd_registry_age)] <- Inf
+annual$cfrd_ogtt_age[is.na(annual$cfrd_ogtt_age)] <- Inf
+annual$cfrd_registry[annual$Age_YrEnd < annual$cfrd_registry_age] <- "CFRD-"
+annual$cfrd_registry[annual$Age_YrEnd >= annual$cfrd_registry_age] <- "CFRD+"
+annual$cfrd_ogtt[annual$Age_YrEnd < annual$cfrd_ogtt_age] <- "CFRD-"
+annual$cfrd_ogtt[annual$Age_YrEnd >= annual$cfrd_ogtt_age] <- "CFRD+"
+# Split into one dataset for registry and one for OGTT
+registry <- annual %>%
+  filter(Age_YrEnd <= cfrd_registry_age) %>%
+  select(eDWID:A_hgba1c, cfrd_registry)
+ogtt <- annual %>%
+  filter(Age_YrEnd <= cfrd_ogtt_age) %>%
+  select(eDWID:A_hgba1c, cfrd_ogtt, cfrd_ogtt_age)
+# Create start times
+registry$Age_YrStart <- registry$Age_YrEnd - 1
+ogtt$Age_YrStart <- ogtt$Age_YrEnd - 1
+```
+
+# By registry CFRD status
+
+## Participant demographics
+
+Categorical variables are from either the first year with CFRD or the most recent year for those without CFRD.
+
+```{r results='asis'}
+v1 <- registry %>%
+  group_by(eDWID) %>%
+  slice_max(ReviewYear)
+t1 <- tableby(cfrd_registry ~ Race + Hispanicrace + Gender + F508 + MutClass +
+  Age_YrEnd + A_bmipercentile + A_hgba1c, data = v1)
+summary(t1, pfootnote = T)
+```
+
+## Kaplan-Meier plots
+
+```{r}
+# Plot
+survfit2(Surv(Age_YrStart, Age_YrEnd, as.numeric(cfrd_registry)) ~ 1,
+  data = registry
+) %>%
+  ggsurvfit() +
+  labs(
+    x = "Age",
+    y = "Overall CFRD- probability"
+  ) +
+  add_confidence_interval() +
+  add_risktable()
+# Median "survival" time
+survfit2(Surv(Age_YrStart, Age_YrEnd, as.numeric(cfrd_registry)) ~ 1,
+  data = registry
+) %>%
+  tbl_survfit(
+    probs = 0.5,
+    label_header = "**Median time to CFRD+ (95% CI)**"
   )
@@
 
+## Cox PH model
+
 ```{r}
-# Plotting function
-roc_plot <- function(roc_obj) {
-  c <- coords(roc_obj, "best")
-  plot_df <- coords(roc_obj)
-  point_df <- plot_df[plot_df$threshold == as.numeric(c[1]), ]
-  c <- round(as.numeric(c), 2)
-  point_label <- paste0(
-    "Threshold: ", c[1], "\n",
-    "Specificity: ", c[2], "\n",
-    "Sensitivity: ", c[3]
+cph <- coxph(
+  Surv(Age_YrStart, Age_YrEnd, as.numeric(cfrd_registry)) ~
+    A_hgba1c + Gender + F508 + A_Vx770 + A_VX809comb,
+  data = registry, cluster = eDWID
 )
-  p <- ggplot(plot_df, aes(x = specificity, y = sensitivity)) +
-    scale_x_reverse() +
-    geom_line() +
-    geom_point(data = point_df) +
-    geom_label_repel(
-      data = point_df, aes(label = point_label), nudge_x = 0.25,
-      segment.color = "grey50", size = 5
-    ) +
-    xlab("Specificity") +
-    ylab("Sensitivity") +
-    theme_bw(base_size = 20)
-  return(p)
-}
+cph %>% tbl_regression(exponentiate = T)
+plot(cox.zph(cph), resid = F)
+cox.zph(cph)
+ggcoef_table(cph,
+  show_p_values = F, add_reference_rows = T, signif_stars = F, colour = F
+)
 ```
 
-# Data cleaning
+# By CFRD status based on 2 hour OGTT >= 200
 
-1.
-People who reported American Indian or Alaska Native, Asian, and Native Hawaiian or Other Pacific Islander for race were considered as "Other race" due to relatively small sample sizes for these groups.
+## Participant demographics
 
-2. Years prior to 2010 and those under age 10 at the end of a given year were excluded from the analysis.
+Categorical variables are from either the first year with CFRD or the most recent year for those without CFRD.
 
-3. HbA1cs < 3.5% or > 15% were excluded from the analysis, as these values were assumed to be entered incorrectly.
+```{r results='asis'}
+v1 <- ogtt %>%
+  group_by(eDWID) %>%
+  slice_max(ReviewYear)
+t1 <- tableby(cfrd_ogtt ~ Race + Hispanicrace + Gender + F508 + MutClass +
+  Age_YrEnd + A_bmipercentile + A_hgba1c, data = v1)
+summary(t1, pfootnote = T)
+```
 
-# Participant Characteristics
-
-## By two hour OGTT value
+## Kaplan-Meier plots
 
 ```{r}
-roc_df %>%
-  select(
-    reviewyear, hgba1c, Age_YrEnd, `Age Group`, `OGTT 2-hour`, Race,
-    Hispanicrace, Gender, F508, MutClass
+# Plot
+survfit2(Surv(Age_YrStart, Age_YrEnd, as.numeric(cfrd_ogtt)) ~ 1,
+  data = ogtt
 ) %>%
-  tbl_summary(
-    by = `OGTT 2-hour`, missing_text = "Missing",
-    type = list(reviewyear ~ "categorical"),
-    label = list(
-      hgba1c = "HbA1c", Hispanicrace = "Hispanic", MutClass = "Mutation Class",
-      reviewyear = "Year", Age_YrEnd = "Age at Year End"
+  ggsurvfit() +
+  labs(
+    x = "Age",
+    y = "Overall CFRD- probability"
+  ) +
+  add_confidence_interval() +
+  add_risktable()
+# Median "survival" time
+survfit2(Surv(Age_YrStart, Age_YrEnd, as.numeric(cfrd_ogtt)) ~ 1,
+  data = ogtt
+) %>%
+  tbl_survfit(
+    probs = 0.5,
+    label_header = "**Median time to CFRD+ (95% CI)**"
   )
-  ) %>%
-  add_overall() %>%
-  add_p(
-    test = list(reviewyear = "fisher.test"),
-    test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
-  ) %>%
-  separate_p_footnotes() %>%
-  as_gt()
 ```
 
-## By age
+## Cox PH model
 
 ```{r}
-roc_df %>%
-  select(
-    reviewyear, hgba1c, Age_YrEnd, `Age Group`, `OGTT 2-hour`, Race,
-    Hispanicrace, Gender, F508, MutClass
-  ) %>%
-  tbl_summary(
-    by = `Age Group`, missing_text = "Missing",
-    type = list(reviewyear ~ "categorical"),
-    label = list(
-      hgba1c = "HbA1c", Hispanicrace = "Hispanic", MutClass = "Mutation Class",
-      reviewyear = "Year", Age_YrEnd = "Age at Year End"
+cph <- coxph(
+  Surv(Age_YrStart, Age_YrEnd, as.numeric(cfrd_ogtt)) ~
+    A_hgba1c + Gender + F508 + A_Vx770 + A_VX809comb,
+  data = ogtt, cluster = eDWID
 )
-  ) %>%
-  add_overall() %>%
-  add_p(
-    test = list(reviewyear = "fisher.test"),
-    test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
-  ) %>%
-  separate_p_footnotes() %>%
-  as_gt()
+cph %>% tbl_regression(exponentiate = T)
+plot(cox.zph(cph), resid = F)
+cox.zph(cph)
+ggcoef_table(cph,
+  show_p_values = F, add_reference_rows = T, signif_stars = F, colour = F
+)
 ```
-
-# HbA1c by diagnosis
-
-## Boxplot
-
-```{r}
-ggplot(roc_df, aes(x = `OGTT 2-hour`, y = hgba1c)) +
-  geom_boxplot() +
-  ylab("HbA1c") +
-  theme_classic(base_size = 20)
-```
-
-## Dotplot
-
-```{r}
-ggplot(roc_df, aes(x = `OGTT 2-hour`, y = hgba1c)) +
-  geom_jitter(alpha = 0.25) +
-  ylab("HbA1c") +
-  theme_classic(base_size = 20)
-```
-
-## Violin plot
-
-```{r}
-ggplot(roc_df, aes(x = `OGTT 2-hour`, y = hgba1c)) +
-  geom_violin(aes(fill = `OGTT 2-hour`)) +
-  ylab("HbA1c") +
-  theme_classic(base_size = 20) +
-  theme(legend.position = "none")
-```
-
-# ROC analysis
-
-For each ROC curve, we performed 2000 bootstrap replicates.
-Here the median best threshold (based on Youden index), sensitivity, and specificity are presented along with the bootstrapped 95% confidence intervals.
-Column names are printed in terms of percentiles, so `threshold.2.5.` indicates the lower limit of the confidence interval, `threshold.50.` indicates the median best threshold, and `threshold.97.5.` indicates the upper limit of the confidence interval.
-All non-bootstrapped thresholds are also reported.
-
-## NGT vs. AGT and AGT vs. CFRD
-
-For the following analyses, we considered those with a two hour glucose value >= 140 and < 200 as a distinct group (AGT) and use receiver operating characteristic curves to find cutoffs that predict:
-
-1. NGT vs. AGT
-
-2. AGT vs. CFRD
-
-```{r}
-ngt_agt <- roc_df %>% filter(`OGTT 2-hour` != ">=200")
-ngt_agt$ogtt <- droplevels(ngt_agt$`OGTT 2-hour`)
-agt_cfrd <- roc_df %>% filter(`OGTT 2-hour` != "<140")
-agt_cfrd$ogtt <- droplevels(agt_cfrd$`OGTT 2-hour`)
-```
-
-### Full cohort
-
-#### NGT vs. AGT
-
-```{r}
-#| message: false
-r <- roc(ogtt ~ hgba1c, data = ngt_agt)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-#### AGT vs. CFRD
-
-```{r}
-#| message: false
-r <- roc(ogtt ~ hgba1c, data = agt_cfrd)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-### Peds vs. Adults
-
-#### Peds
-
-```{r}
-peds_ngt_agt <- ngt_agt %>% filter(Age_YrEnd < 18)
-peds_agt_cfrd <- agt_cfrd %>% filter(Age_YrEnd < 18)
-```
-
-##### NGT vs. AGT
-
-```{r}
-#| message: false
-r <- roc(ogtt ~ hgba1c, data = peds_ngt_agt)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-##### AGT vs. CFRD
-
-```{r}
-#| message: false
-r <- roc(ogtt ~ hgba1c, data = peds_agt_cfrd)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-#### Adults
-
-```{r}
-adult_ngt_agt <- ngt_agt %>% filter(Age_YrEnd >= 18)
-adult_agt_cfrd <- agt_cfrd %>% filter(Age_YrEnd >= 18)
-```
-
-##### NGT vs. AGT
-
-```{r}
-#| message: false
-r <- roc(ogtt ~ hgba1c, data = adult_ngt_agt)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-##### AGT vs. CFRD
-
-```{r}
-#| message: false
-r <- roc(ogtt ~ hgba1c, data = adult_agt_cfrd)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-### By race
-
-#### White
-
-```{r}
-white_ngt_agt <- ngt_agt %>% filter(Race == "White")
-white_agt_cfrd <- agt_cfrd %>% filter(Race == "White")
-```
-
-##### NGT vs. AGT
-
-```{r}
-#| message: false
-r <- roc(ogtt ~ hgba1c, data = white_ngt_agt)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-##### AGT vs. CFRD
-
-```{r}
-#| message: false
-r <- roc(ogtt ~ hgba1c, data = white_agt_cfrd)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-#### Black or African American
-
-```{r}
-black_ngt_agt <- ngt_agt %>% filter(Race == "Black or African American")
-black_agt_cfrd <- agt_cfrd %>% filter(Race == "Black or African American")
-```
-
-##### NGT vs. AGT
-
-```{r}
-#| message: false
-r <- roc(ogtt ~ hgba1c, data = black_ngt_agt)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-##### AGT vs. CFRD
-
-Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.
-
-```{r}
-#| message: false
-r <- roc(ogtt ~ hgba1c, data = black_agt_cfrd)
-roc_plot(r)
-gt(data.frame(coords(r, "best"))) %>%
-  fmt_number(n_sigfig = 3)
-gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
-  fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-#### More than one race
-
-```{r}
-more_ngt_agt <- ngt_agt %>% filter(Race == "More than one race")
-more_agt_cfrd <- agt_cfrd %>% filter(Race == "More than one race")
-```
-
-##### NGT vs. AGT
-
-```{r}
-#| message: false
-r <- roc(ogtt ~ hgba1c, data = more_ngt_agt)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-##### AGT vs. CFRD
-
-```{r}
-#| message: false
-r <- roc(ogtt ~ hgba1c, data = more_agt_cfrd)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-#### Other race
-
-```{r}
-other_ngt_agt <- ngt_agt %>% filter(Race == "Other race")
-other_agt_cfrd <- agt_cfrd %>% filter(Race == "Other race")
-```
-
-##### NGT vs. AGT
-
-Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.
-
-```{r}
-#| message: false
-r <- roc(ogtt ~ hgba1c, data = other_ngt_agt)
-roc_plot(r)
-gt(data.frame(coords(r, "best"))) %>%
-  fmt_number(n_sigfig = 3)
-gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
-  fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-##### AGT vs. CFRD
-
-Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.
-
-```{r}
-#| message: false
-r <- roc(ogtt ~ hgba1c, data = other_agt_cfrd)
-roc_plot(r)
-gt(data.frame(coords(r, "best"))) %>%
-  fmt_number(n_sigfig = 3)
-gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
-  fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-### By ethnicity
-
-#### Hispanic
-
-```{r}
-hispanic_ngt_agt <- ngt_agt %>% filter(Hispanicrace == "Yes")
-hispanic_agt_cfrd <- agt_cfrd %>% filter(Hispanicrace == "No")
-```
-
-##### NGT vs. AGT
-
-```{r}
-#| message: false
-r <- roc(ogtt ~ hgba1c, data = hispanic_ngt_agt)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-##### AGT vs. CFRD
-
-```{r}
-#| message: false
-r <- roc(ogtt ~ hgba1c, data = hispanic_agt_cfrd)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-#### Non-Hispanic
-
-```{r}
-non_hispanic_ngt_agt <-
-  ngt_agt %>% filter(Race == "Black or African American")
-non_hispanic_agt_cfrd <-
-  agt_cfrd %>% filter(Race == "Black or African American")
-```
-
-##### NGT vs. AGT
-
-```{r}
-#| message: false
-r <- roc(ogtt ~ hgba1c, data = non_hispanic_ngt_agt)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-##### AGT vs. CFRD
-
-Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.
-
-```{r}
-#| message: false
-r <- roc(ogtt ~ hgba1c, data = non_hispanic_agt_cfrd)
-roc_plot(r)
-gt(data.frame(coords(r, "best"))) %>%
-  fmt_number(n_sigfig = 3)
-gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
-  fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-## Two hour OGTT >= 140 or >= 200
-
-For the following analyses, we examined HbA1c cutoffs that best predict a two hour OGTT value either >= 140 or >= 200
-
-```{r}
-roc_df$agt <- as.numeric(roc_df$twohour_bloodglucose >= 140)
-roc_df$cfrd <- as.numeric(roc_df$twohour_bloodglucose >= 200)
-```
-
-### Full cohort
-
-#### $\geq$ 140
-
-```{r}
-#| message: false
-r <- roc(agt ~ hgba1c, data = roc_df)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-#### $\geq$ 200
-
-```{r}
-#| message: false
-r <- roc(cfrd ~ hgba1c, data = roc_df)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-### Peds vs. Adults
-
-#### Peds
-
-```{r}
-peds_roc <- roc_df %>% filter(Age_YrEnd < 18)
-```
-
-##### $\geq$ 140
-
-```{r}
-#| message: false
-r <- roc(agt ~ hgba1c, data = peds_roc)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-##### $\geq$ 200
-
-```{r}
-#| message: false
-r <- roc(cfrd ~ hgba1c, data = peds_roc)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-#### Adults
-
-```{r}
-adult_roc <- roc_df %>% filter(Age_YrEnd >= 18)
-```
-
-##### $\geq$ 140
-
-```{r}
-#| message: false
-r <- roc(agt ~ hgba1c, data = adult_roc)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-##### $\geq$ 200
-
-```{r}
-#| message: false
-r <- roc(cfrd ~ hgba1c, data = adult_roc)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-### By race
-
-#### White
-
-```{r}
-white_roc <- roc_df %>% filter(Race == "White")
-```
-
-##### $\geq$ 140
-
-```{r}
-#| message: false
-r <- roc(agt ~ hgba1c, data = white_roc)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-##### $\geq$ 200
-
-```{r}
-#| message: false
-r <- roc(cfrd ~ hgba1c, data = white_roc)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-#### Black or African American
-
-```{r}
-black_roc <- roc_df %>% filter(Race == "Black or African American")
-```
-
-##### $\geq$ 140
-
-```{r}
-#| message: false
-r <- roc(agt ~ hgba1c, data = black_roc)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-##### $\geq$ 200
-
-Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.
-
-```{r}
-#| message: false
-r <- roc(cfrd ~ hgba1c, data = black_roc)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
-  fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-#### More than one race
-
-```{r}
-more_roc <- roc_df %>% filter(Race == "More than one race")
-```
-
-##### $\geq$ 140
-
-```{r}
-#| message: false
-r <- roc(agt ~ hgba1c, data = more_roc)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-##### $\geq$ 200
-
-```{r}
-#| message: false
-r <- roc(cfrd ~ hgba1c, data = more_roc)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-#### Other race
-
-```{r}
-other_roc <- roc_df %>% filter(Race == "Other race")
-```
-
-##### $\geq$ 140
-
-Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.
-
-```{r}
-#| message: false
-r <- roc(agt ~ hgba1c, data = other_roc)
-roc_plot(r)
-gt(data.frame(coords(r, "best"))) %>%
-  fmt_number(n_sigfig = 3)
-gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
-  fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-##### $\geq$ 200
-
-```{r}
-#| message: false
-r <- roc(cfrd ~ hgba1c, data = other_roc)
-roc_plot(r)
-gt(data.frame(coords(r, "best"))) %>%
-  fmt_number(n_sigfig = 3)
-gt(data.frame(ci.coords(r, "best"))) %>%
-  fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-### By ethnicity
-
-#### Hispanic
-
-```{r}
-hispanic_roc <- roc_df %>% filter(Hispanicrace == "Yes")
-```
-
-##### $\geq$ 140
-
-```{r}
-#| message: false
-r <- roc(agt ~ hgba1c, data = hispanic_roc)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-##### $\geq$ 200
-
-```{r}
-#| message: false
-r <- roc(cfrd ~ hgba1c, hispanic_roc)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-#### Non-Hispanic
-
-```{r}
-non_hispanic_roc <- roc_df %>% filter(Hispanicrace == "No")
-```
-
-##### $\geq$ 140
-
-```{r}
-#| message: false
-r <- roc(agt ~ hgba1c, data = non_hispanic_roc)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-##### $\geq$ 200
-
-```{r}
-#| message: false
-r <- roc(cfrd ~ hgba1c, data = non_hispanic_roc)
-roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```

### 2024-10-28 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
updated sweat test data for Stacey

@@
 library(tidyverse)
-library(forestmodel)
 library(childsds)
@@
 library(ggstats)
-library(forestmodel)
 home_dir <- switch(Sys.info()["sysname"],

### 2024-10-28 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
updated sweat test data for Stacey

@@
 chco <- chco %>% select(
-  PatientID, Diagnosis, Sex, Genotypes1, Genotypes2,
-  Age.at.Test.in.Years, Method, Site, Na, K, Cl, Na_Left, Na_Right, K_Left,
+  PatientID, Site, Diagnosis, Sex, Genotypes1, Genotypes2,
+  Age.at.Test.in.Years, Method, Na, K, Cl, Na_Left, Na_Right, K_Left,
   K_Right, Cl_Left, Cl_Right
 )
-# Get each person's first test
-chco <- chco %>%
-  group_by(PatientID) %>%
-  slice_min(Age.at.Test.in.Years, na_rm = T)
 # Import NJH data
-njh <- read_excel("./Data_Raw/SweatVariability2025_CFgenotypes.xlsx")
+njh <- read_excel("./Data_Raw/SweatVariability2025_CFgenotypes 2024-08-21.xlsx")
 # Filter and select similar to CHCO
@@
   filter(`Demographics Available`) %>%
+  mutate(Site = "NJH") %>%
   select(
-    `Patient ID`, Diagnosis, Sex, `Genotype 1`, `Genotype 2`,
-    `Age at test (in years)`, `Test Method`, `Site of test`, `Sweat sodium`,
+    `Patient ID`, `Site of test`, Diagnosis, Sex, `Genotype 1`, `Genotype 2`,
+    `Age at test (in years)`, `Test Method`, `Sweat sodium`,
     `Sweat potassium`, `Sweat chloride`, `Sweat sodium left arm`,
@@
 colnames(diagnoses) <- c("PatientID", "Diagnosis")
+# Less detailed diagnoses
+diagnoses$Diagnosis <- factor(diagnoses$Diagnosis,
+  levels = c(
+    "Abnormal sweat choride test without diagnosis of cystic fibrosis",
+    "Abnormal sweat test",
+    "Alpha-1-antitrypsin deficiency",
+    "Alpha-1-antitrypsin deficiency carrier",
+    "Alpha-1-antitrypsin deficiency, Positive cystic fibrosis sweat test without diagnosis of cystic fibrosis",
+    "CF", "CF carrier, Alpha-1-antitrypsin deficiency",
+    "CF, alpha 1-antitrypsin deficiency",
+    "CF, alpha 1-antitrypsin deficiency carrier",
+    "CF, Alpha-1-antitrypsin deficiency",
+    "Cystic fibrosis carrier",
+    "Cystic fibrosis carrier, Alpha-1-antitrypsin deficiency",
+    "Cystic fibrosis carrier, Alpha-1-antitrypsin deficiency carrier",
+    "Cystic fibrosis screening",
+    "Cystic fibrosis transmembrane conductance regulator (CFTR)-related disorder",
+    "Positive cystic fibrosis sweat test without diagnosis of cystic fibrosis",
+    "Primary ciliary dyskinesia",
+    "Primary Ciliary dyskinesia, Cystic Fibrosis carrier"
+  ),
+  labels = c(
+    "Non-CF", "Non-CF", "Non-CF", "CF Carrier", "Non-CF", "Cystic Fibrosis",
+    "CF Carrier", "Cystic Fibrosis", "Cystic Fibrosis", "Cystic Fibrosis",
+    "CF Carrier", "CF Carrier", "CF Carrier", "Non-CF", "CFTR-RD or CRMS",
+    "Non-CF", "Non-CF", "CF Carrier"
+  )
+)
 njh <- full_join(njh, diagnoses)
@@
   select(
-    subject, Diagnosis, sex, Genotypes1, Genotypes2, age, Method, site,
+    subject, site, Diagnosis, sex, Genotypes1, Genotypes2, age, Method,
     swsavg, K, swcavg, Na_Left, Na_Right, K_Left, K_Right, Cl_Left, Cl_Right
@@
 df <- as.data.frame(df)
-# Limit to complete data
+# Get rid of unnecessary columns
 df <- df %>%
   select(
-    PatientID, Diagnosis, Sex, Age.at.Test.in.Years, Na:Cl,
+    PatientID, Site, Diagnosis, Sex, Age.at.Test.in.Years, Na:Cl,
     Genotypes1, Genotype1Function, Genotypes2, Genotype2Function,
@@
   )
+# For each ID, get the row with the most data (or first test)
+df$Missing <- rowSums(is.na(df))
+df <- df %>%
+  group_by(PatientID) %>%
+  slice_min(Missing) %>%
+  slice_min(row_number()) %>%
+  filter(!is.na(Diagnosis)) %>%
+  ungroup()
 # Set non-CF to 100% CFTR function (200 because 100 for each allele)
@@
 # Write for checking
-write.csv(df, "./Data_Cleaned/analysis_dataset.csv", row.names = F, na = "")
+write.csv(df, "./Data_Cleaned/final_analysis_dataset.csv", row.names = F, na = "")
 ```
@@
 df %>%
-  filter(Diagnosis == "CF") %>%
   ggplot(
@@
 df %>%
-  filter(Diagnosis == "CF") %>%
   ggplot(
     aes(x = Na / Cl, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
-  ) +
-  geom_point() +
-  theme()
-
-df %>%
-  filter(Diagnosis == "CF") %>%
-  ggplot(
-    aes(
-      x = Na, y = TotalGenotypeFunction, color = Age.at.Test.in.Years,
-      group = Diagnosis
-    )
-  ) +
-  geom_point() +
-  theme()
-df %>%
-  filter(Diagnosis == "CF") %>%
-  ggplot(
-    aes(
-      x = Cl, y = TotalGenotypeFunction, color = Age.at.Test.in.Years,
-      group = Diagnosis
-    )
-  ) +
-  geom_point() +
-  theme()
-df %>%
-  filter(Diagnosis == "CF") %>%
-  ggplot(
-    aes(
-      x = Na / Cl, y = TotalGenotypeFunction, color = Age.at.Test.in.Years,
-      group = Diagnosis
-    )
   ) +
@@
   select(Na, Cl, TotalGenotypeFunction, `Cl < 60`, `Age group`) %>%
-  tbl_summary(by = `Age group`, missing_text = "Missing")
-write.csv(df, file = "~/sweat_analysis_data.csv", na = "", row.names = F)
+  tbl_summary(by = `Age group`, missing_text = "Missing") %>%
+  add_p()
 ```
-
-Add CRMS change in sweat and change in ratio
-
-Add CF carrier genotypes to this report
-
-Plot just the CF genotype function
-  - Are the chloride values matching up with the CFTR function as we'd expect?

### 2024-11-01 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
Update sweat_test_analysis.qmd

@@
 chec$Genotypes2 <- trimws(sapply(strsplit(chec$gntyp_pair, "/"), "[[", 2))
-chec$Diagnosis <- "CF"
+chec$Diagnosis <- "Cystic Fibrosis"
 chec[, c(
@@
 # Write for checking
-write.csv(df, "./Data_Cleaned/final_analysis_dataset.csv", row.names = F, na = "")
+write.csv(df, "./Data_Cleaned/final_analysis_dataset.csv",
+  row.names = F, na = ""
+)
 ```
@@
 df %>%
-  filter(Diagnosis == "CF") %>%
+  filter(Diagnosis == "Cystic Fibrosis") %>%
   mutate(

### 2024-11-01 Lilah Melzer/cough_monitor_analysis.qmd
Update cough_monitor_analysis.qmd

@@
 library(Hmisc)
+library(redcapAPI)
 library(tidyverse)
-library(lubridate)
-library(arsenal)
 library(rspiro)
 library(childsds)
+library(arsenal)
 library(lmerTest)
 library(performance)
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring",
-  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring"
+```
+
+```{r data import}
+#| include: false
+# Import from REDCap
+unlockREDCap(c(rcon = "Cough Monitor 23-0382"),
+  keyring = "API_KEYs",
+  envir = 1,
+  url = "https://redcap.ucdenver.edu/api/"
 )
-github_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/GitHub",
-  "Windows" = "C:/Users/timvigers/Documents/GitHub",
-  "Linux" = "/home/timvigers/GitHub"
-)
-knitr::opts_knit$set(root.dir = home_dir)
+df <- exportRecordsTyped(rcon)
 ```
@@
 #| warning: false
-# Import REDCap data
-source(paste0(github_dir, "/CF/Lilah Melzer/cough_monitor_import.r"))
 # Import use data
-start_end <- read.csv("./Data_Clean/Start and end dates.csv", na.strings = "")
+start_end <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring/Data_Clean/Start and end dates.csv",
+  na.strings = ""
+)
 start_end$Start <- mdy(start_end$Start)
 start_end$End <- mdy(start_end$End)
-vacations <- read.csv("./Data_Clean/Vacation days.csv", na.strings = "")
+vacations <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring/Data_Clean/Vacation days.csv",
+  na.strings = ""
+)
 vacations$Vacation.Start <- mdy(vacations$Vacation.Start)
 vacations$Vacation.End <- mdy(vacations$Vacation.End)
+# Make a grid of everyone's possible dates
+dates <- start_end %>%
+  rowwise() %>%
+  transmute(sid,
+    Date = list(seq(Start, End - 1, by = "day"))
+  ) %>%
+  unnest(Date)
 # Import counts
 agg_counts <- read.csv(
-  "./Data_Raw/Cough_w_TotalCounts_0824 - Cough_agg_w_count_0828.csv",
+  "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring/Data_Raw/Cough_w_TotalCounts_0824 - Cough_agg_w_count_0828.csv",
   na.strings = c("", "NMD"), header = F
@@
   )
-# Add start and end dates, filter
+# Merge with all possible dates
 agg_counts$sid <- as.numeric(agg_counts$sid)
-agg_counts <- left_join(agg_counts, start_end)
-agg_counts <- agg_counts %>%
-  filter(Date >= Start & Date <= End)
+agg_counts <- left_join(dates, agg_counts)
 # Vacations
@@
 agg_counts <- agg_counts %>%
-  filter(!Vacation) %>%
+  filter(!Vacation | is.na(Vacation)) %>%
   arrange(sid, Date)
+agg_counts$Vacation <- NULL
 # Sick
-sick <- data %>%
-  select(sid, pulm_ex_v2.factor) %>%
-  mutate(
-    antib_dt = mdy(antib_dt), 
-    px_start = antib_dt - 7,
-    px_end = antib_dt + 28
+sick <- df %>%
+  select(sid, pulm_ex_v2, pulm_ex_spec_v2) %>%
+  filter(pulm_ex_v2 == "Yes")
+sick$pulm_ex_abx_start <- mdy(c("4/17/24", "5/8/24"))
+sick$pulm_ex_start <- sick$pulm_ex_abx_start - 7
+sick$pulm_ex_end <- sick$pulm_ex_abx_start + 28
+# Make a grid of everyone's sick days and merge
+sick <- sick %>%
+  select(sid, pulm_ex_start, pulm_ex_end) %>%
+  rowwise() %>%
+  transmute(sid,
+    Date = list(seq(pulm_ex_start, pulm_ex_end, by = "day"))
+  ) %>%
+  unnest(Date) %>%
+  mutate(Sick = T)
+sick$sid <- as.numeric(sick$sid)
+agg_counts <- left_join(agg_counts, sick)
+# Format sick column
+agg_counts$Sick[is.na(agg_counts$Sick)] <- F
+agg_counts$Sick <- factor(agg_counts$Sick,
+  levels = c(F, T),
+  labels = c("No", "Yes")
 )
@@
 # Demographics data
-data$Diagnosis <- factor(sub(".*arm_*", "", data$redcap_event_name),
-  levels = 1:2,
-  labels = c("CF", "Non-CF")
+df$Diagnosis <- factor(df$redcap_event_name,
+  levels = c(
+    "Visits (Arm 1: CF)", "Interim (Arm 1: CF)", "Visits (Arm 2: Control)",
+    "Interim (Arm 2: Control)"
+  ),
+  labels = c("CF", "CF", "Non-CF", "Non-CF")
 )
-demo <- data %>%
+df$Diagnosis <- relevel(df$Diagnosis, ref = "Non-CF")
+demo <- df %>%
   group_by(sid) %>%
@@
     sid, Diagnosis, age_at_consent, height, weight, pre_fev,
-    nih_sex.factor:nih_ethnicity.factor, mutation1.factor, mutation2.factor,
-    cftr_mod.factor, cftr_mod_type.factor, mucolytic.factor:antih.factor
+    nih_sex:nih_ethnicity, mutation1, mutation2,
+    cftr_mod, cftr_mod_type, mucolytic:antih
   )
@@
 fev1 <- demo %>%
-  select(age_at_consent, height, weight, nih_sex.factor, pre_fev) %>%
-  filter(nih_sex.factor %in% c("Male", "Female"))
+  select(sid, age_at_consent, height, weight, nih_sex, pre_fev) %>%
+  filter(nih_sex %in% c("Male", "Female"))
 fev1$pre_fev_pp <- pctpred_GLIgl(fev1$age_at_consent, fev1$height / 100,
-  factor(fev1$nih_sex.factor, levels = c("Male", "Female")),
+  factor(fev1$nih_sex, levels = c("Male", "Female")),
   FEV1 = fev1$pre_fev
@@
 fev1$height_perc <- sds(as.numeric(fev1$height),
-  age = fev1$age_at_consent, sex = fev1$nih_sex.factor,
+  age = fev1$age_at_consent, sex = fev1$nih_sex,
   male = "Male", female = "Female", item = "height2_20", ref = cdc.ref,
@@
 fev1$weight_perc <- sds(as.numeric(fev1$weight),
-  age = fev1$age_at_consent, sex = fev1$nih_sex.factor,
+  age = fev1$age_at_consent, sex = fev1$nih_sex,
   male = "Male", female = "Female", item = "weight2_20", ref = cdc.ref,
@@
 demo$Genotype <-
-  factor(rowSums(demo[, c("mutation1.factor", "mutation2.factor")] == "F508del"),
+  factor(rowSums(demo[, c("mutation1", "mutation2")] == "F508del"),
     levels = c(2, 1, 0),
@@
   demo$Diagnosis[match(agg_counts$sid, as.numeric(as.character(demo$sid)))]
+# Save
+write.csv(agg_counts, file = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring/Data_Clean/cleaned_aggregated_counts.csv", na = "", row.names = F)
 # Per night
@@
   select(-height, -weight, -pre_fev)
-# Exclude vacation days
-
 # Set table options for the document
@@
   tableby.control(
-    numeric.stats = c("Nmiss", "medianq1q3", "range"),
+    numeric.stats = c("N", "Nmiss", "medianq1q3", "range"),
+    cat.stats = c("N", "Nmiss", "countpct"),
     numeric.test = "kwt"
@@
 f <- as.formula(paste0("Diagnosis ~", paste0(colnames(demo), collapse = "+")))
-f <- update(f, . ~ . - Diagnosis - sid - median_coughs + kwt(median_coughs))
+f <- update(f, . ~ . - Diagnosis - sid)
 t1 <- tableby(f, data = demo, control = mycontrols)
 summary(t1, pfootnote = T)
+```
+
+# Data for ATS abstract
+
+```{r results='asis'}
+per_person <- agg_counts %>%
+  group_by(sid, Diagnosis) %>%
+  summarise(
+    `Nights` = n(),
+    `Nights With Data` = sum(!is.na(Coughs)),
+    `Prop. Nights With Data` = sum(!is.na(Coughs)) / n(),
+    `Median Coughs per Night` = median(Coughs, na.rm = T),
+    .groups = "drop"
+  )
+t2 <- tableby(
+  Diagnosis ~ `Nights` + `Nights With Data` + `Prop. Nights With Data` +
+    `Median Coughs per Night`,
+  data = per_person, control = mycontrols
+)
+summary(t2, pfootnote = T)
 ```
@@
 gtsummary::tbl_regression(mod, exponentiate = T)
+# check_model(mod, re_formula = NA, residual_type = "normal")
 ```
+
+### By sick status
+
+```{r}
+mod <- glmer(Coughs ~ Night + Diagnosis + Sick + (Night | sid),
+  data = agg_counts, family = "poisson"
+)
+gtsummary::tbl_regression(mod, exponentiate = T)
+# check_model(mod, re_formula = NA, residual_type = "normal")
+```

### 2024-11-04 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
minor updates - going to try just running three separate models per Kristen

@@
           by = CFRD, missing_text = "Missing",
-          type = list(A_corticosteroids1 ~ "categorical")
+          type = list(A_corticosteroids1 ~ "categorical"),
+          statistic = list(c(
+            "Age_YrEnd", "bmi_perc", "A_FEV1_pct_predicted",
+            "A_FVC_pct_predicted"
+          ) ~ "{mean} ({sd})")
         ) %>%
-        add_p() %>%
+        add_p(test = list(c(
+          "Age_YrEnd", "bmi_perc", "A_FEV1_pct_predicted",
+          "A_FVC_pct_predicted"
+        ) ~ "t.test")) %>%
         add_overall() %>%
@@
     .groups = "drop"
-  ) %>%
+  )
+kable(
+  screening %>%
+    select(ReviewYear, age_group, `Screened OGTT`) %>%
+    mutate(`Screened OGTT` = round(`Screened OGTT` * 100, 2)) %>%
+    pivot_wider(names_from = age_group, values_from = `Screened OGTT`),
+  caption = "% Screened by OGTT"
+)
+screening <- screening %>%
   pivot_longer(`Screened HbA1c`:`Screened Either`)
@@
 og <- ymd("2003-01-01")
+# Need to use a dataframe not a tibble
+annual <- as.data.frame(annual)
 # Formula for base model
@@
 cox_res %>%
-  tbl_regression()
-ggcoef_table(cox_res, show_p_values = F, add_reference_rows = T, signif_stars = F, colour = F)
+  tbl_regression(exponentiate = T)
+ggcoef_table(cox_res,
+  show_p_values = F, add_reference_rows = T, exponentiate = T,
+  signif_stars = F, colour = F
+)
 # Diagnostics
@@
 vx770_cox_res %>%
-  tbl_regression()
+  tbl_regression(exponentiate = T)
 ggcoef_table(vx770_cox_res,
-  show_p_values = F, add_reference_rows = T, signif_stars = F, colour = F
+  show_p_values = F, add_reference_rows = T, exponentiate = T,
+  signif_stars = F, colour = F
 )
@@
 vx809comb_cox_res %>%
-  tbl_regression()
+  tbl_regression(exponentiate = T)
 ggcoef_table(vx809comb_cox_res,
-  show_p_values = F, add_reference_rows = T, signif_stars = F, colour = F
+  show_p_values = F, add_reference_rows = T, exponentiate = T,
+  signif_stars = F, colour = F
 )
@@
   cox_res_split %>%
-    tbl_regression() %>%
+    tbl_regression(exponentiate = T) %>%
     print()
@@
   ggcoef_table(cox_res_split,
-    show_p_values = F, add_reference_rows = T, signif_stars = F, colour = F
+    show_p_values = F, add_reference_rows = T, exponentiate = T,
+    signif_stars = F, colour = F
   ) %>%

### 2024-11-07 Scott Sagel/PEx in CF/analysis_for_yaron.qmd
added white vs. non-white

--- 2024-10-22 Scott Sagel/PEx in CF/analysis_for_yaron.qmd
@@
   )
+)
+df$race_wnw <- factor(df$race,
+  levels = c(
+    "White", "Black or African-American", "Other", "Asian",
+    "Unknown/not reported"
+  ),
+  labels = c("White", "Non-White", "Non-White", "Non-White", "Non-White")
 )
@@
   "Age at admission" = "age_at_admit", "Sex" = "sex",
-  "Race" = "race", "Ethnicity" = "ethnicity",
-  "Azithromycin" = "azithro", "Inhaled antibiotics" = "inhaled_abx",
+  "Race" = "race", "Race (White vs. Non-White)" = "race_wnw",
+  "Ethnicity" = "ethnicity", "Azithromycin" = "azithro",
+  "Inhaled antibiotics" = "inhaled_abx",
   "Systemic steroids" = "steroids_system",
@@
     redcap_data_access_group, year, sex, age_group, age_at_admit, race,
-    ethnicity, Genotype, Insurance, cftr_mod, cftr_mod_spec,
+    race_wnw, ethnicity, Genotype, Insurance, cftr_mod, cftr_mod_spec,
     Trikafta, azithro, inhaled_abx, steroids_system, admit_cx_source,
@@
   ungroup() %>%
-  select(redcap_data_access_group, sex, race, ethnicity, Genotype) %>%
+  select(redcap_data_access_group, sex, race, race_wnw, ethnicity, Genotype) %>%
   tbl_summary(missing_text = "(Missing)", label = labels)
@@
   ungroup() %>%
-  select(year, redcap_data_access_group, sex, race, ethnicity, Genotype) %>%
-  tbl_summary(by = year, missing_text = "(Missing)", label = labels)
+  select(
+    year, redcap_data_access_group, sex, race, race_wnw, ethnicity, Genotype
+  ) %>%
+  tbl_summary(by = year, missing_text = "(Missing)", label = labels) %>%
+  add_p() %>%
+  separate_p_footnotes()
 ```
@@
         c(
-          age_at_admit, ppfev1_baseline, ppfev1_admit, ppfev1_discharge,
-          ppfev1_change
+          age_at_admit, ppfev1_baseline, ppfev1_admit, ppfev1_discharge
         ) ~

### 2024-11-08 Lilah Melzer/cough_monitor_analysis.qmd
fixed cough monitor end date

@@
 # Import use data
-start_end <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring/Data_Clean/Start and end dates.csv",
+start_end <- read.csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring/Data_Clean/Start and end dates.csv",
   na.strings = ""
@@
 start_end$End <- mdy(start_end$End)
-vacations <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring/Data_Clean/Vacation days.csv",
+vacations <- read.csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring/Data_Clean/Vacation days.csv",
   na.strings = ""
@@
 vacations$Vacation.End <- mdy(vacations$Vacation.End)
+# Those with end date after the last data pull will just end on the date of the
+# pull
+pull_date <- mdy("8/28/24")
+start_end$End[is.na(start_end$End)] <- pull_date
+start_end$End[start_end$End > pull_date] <- pull_date
 # Make a grid of everyone's possible dates
@@
   transmute(sid,
-    Date = list(seq(Start, End - 1, by = "day"))
+    Date = list(seq(Start, End, by = "day"))
   ) %>%
@@
 agg_counts <- read.csv(
-  "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring/Data_Raw/Cough_w_TotalCounts_0824 - Cough_agg_w_count_0828.csv",
+  "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring/Data_Raw/Cough_w_TotalCounts_0824 - Cough_agg_w_count_0828.csv",
   na.strings = c("", "NMD"), header = F
@@
 # Save
-write.csv(agg_counts, file = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring/Data_Clean/cleaned_aggregated_counts.csv", na = "", row.names = F)
+write.csv(agg_counts, file = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring/Data_Clean/cleaned_aggregated_counts.csv", na = "", row.names = F)
 # Per night

### 2024-11-12 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
three models only

@@
 library(ggstats)
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
-  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
-)
-github_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/GitHub",
-  "Windows" = "C:/Users/timvigers/Documents/GitHub",
-  "Linux" = "/home/timvigers/GitHub"
-)
-source(paste0(github_dir, "/shared-resources/Visualization/forest_plots.R"))
-knitr::opts_knit$set(root.dir = home_dir)
+knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry")
 ```
@@
 time_varying <- rownames(test.ph)[test.ph$p < 0.05]
+# Just split data by time period so we're only fitting 3 models
+# "[2003,2009)", "[2009,2014)", "[2014,2019)"
+annual$time_period <- cut(annual$ReviewYear,
+  breaks = c(-Inf, 2009, 2014, Inf), right = F
+)
+era1 <- coxph(f, data = annual[as.numeric(annual$time_period) == 1, ], cluster = eDWID)
+era2 <- coxph(f, data = annual[as.numeric(annual$time_period) == 2, ], cluster = eDWID)
+era3 <- coxph(f, data = annual[as.numeric(annual$time_period) == 3, ], cluster = eDWID)
 ```

### 2024-11-12 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
model comparison table and plot

@@
 annual <- as.data.frame(annual)
+annual$tstart <- as.numeric(annual$tstart) - as.numeric(og)
+annual$tstop <- as.numeric(annual$tstop) - as.numeric(og)
 # Formula for base model
@@
 )
-y <- "Surv(time = as.numeric(tstart), time2 = as.numeric(tstop), event = as.numeric(CFRD) == 2)"
+y <- "Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2)"
 f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
@@
 )
-era1 <- coxph(f, data = annual[as.numeric(annual$time_period) == 1, ], cluster = eDWID)
-era2 <- coxph(f, data = annual[as.numeric(annual$time_period) == 2, ], cluster = eDWID)
-era3 <- coxph(f, data = annual[as.numeric(annual$time_period) == 3, ], cluster = eDWID)
+era1 <- coxph(f,
+  data = annual[as.numeric(annual$time_period) == 1, ],
+  cluster = eDWID
+)
+era2 <- coxph(f,
+  data = annual[as.numeric(annual$time_period) == 2, ],
+  cluster = eDWID
+)
+era3 <- coxph(f,
+  data = annual[as.numeric(annual$time_period) == 3, ],
+  cluster = eDWID
+)
+# # KM plot
+# fit <- survfit(Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2) ~ age_group,
+#   data = annual[as.numeric(annual$time_period) == 1, ]
+# )
+# # Visualize with survminer
+# ggsurvplot(fit, data = annual[as.numeric(annual$time_period) == 1, ], risk.table = TRUE)
 ```
@@
 
+#### 2003 - 2008
+
 ```{r}
-annual$tstart <- as.numeric(annual$tstart)
-annual$tstop <- as.numeric(annual$tstop)
-annual_split <- survSplit(
-  Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
-    origin = as.numeric(og)
-  ) ~ .,
-  data = annual,
-  cut = c(
-    as.numeric(ymd("2008-12-31")) - as.numeric(og),
-    as.numeric(ymd("2013-12-31")) - as.numeric(og)
-  ),
-  episode = "tgroup"
+t1 <- era1 %>% tbl_regression(exponentiate = T)
+t1
+ggcoef_model(era1,
+  show_p_values = F, add_reference_rows = T, exponentiate = T,
+  signif_stars = F, colour = F
 )
-annual_split$tgroup <- factor(annual_split$tgroup,
-  levels = 1:3, labels = c("[2003,2009)", "[2009,2014)", "[2014,2019)")
-)
-y <- "Surv(tstart, tstop, event)"
 ```
 
-In the plots below, time-varying coefficients are excluded from the first forest plot, to show the average effect of the predictors.
-In the second forest plot, the estimated marginal means (EMMs) for the time-varying coefficients are shown.
-The EMMs are averaged over the levels of all the other variables in the model.
-In other words, the EMM represents the baseline hazard (not hazard ratio) at the average level of all the other variables in the model.
-For example, the log hazard for an average participant age [30, Inf) in the time period [2003,2009) is 1.887 and the log hazard for a similar participant in [2014,2019) is 1.401.
+#### 2009 - 2013
 
-```{r results='asis'}
-invisible(lapply(time_varying, function(v) {
-  # Formulas
-  new_f <- as.formula(paste0(
-    y, "~", paste0(xvars[-which(xvars == v)], collapse = "+"),
-    paste0("+", v, "*strata(tgroup)")
-  ))
-  new_f_emm <- as.formula(paste0(
-    "~", paste0(v, "|tgroup")
-  ))
-  # This could be done more efficiently but it seems to work
-  # Fit one stratified model and one with a regular interaction term
-  cox_res_split <- coxph(new_f, data = annual_split, cluster = eDWID)
-  # Use the regular interaction term to get means for each time period
-  cat("\n")
-  cat("\n")
-  cat("####", v)
-  cat("\n")
-  cat("\n")
-  cox_res_split %>%
-    tbl_regression(exponentiate = T) %>%
-    print()
-  cat("\n")
-  cat("\n")
-  emm <- emmeans(cox_res_split, new_f_emm, nuisance = xvars[-which(xvars == v)])
-  emm <- data.frame(emm)
-  # Plot base model (without time varying)
-  ggcoef_table(cox_res_split,
+```{r}
+t2 <- era2 %>% tbl_regression(exponentiate = T)
+t2
+ggcoef_model(era2,
   show_p_values = F, add_reference_rows = T, exponentiate = T,
   signif_stars = F, colour = F
-  ) %>%
-    print()
-  cat("\n")
-  cat("\n")
-}))
+)
 ```
+
+#### 2014 - 2018
+
+```{r}
+t3 <- era3 %>% tbl_regression(exponentiate = T)
+t3
+ggcoef_model(era3,
+  show_p_values = F, add_reference_rows = T, exponentiate = T,
+  signif_stars = F, colour = F
+)
+```
+
+#### Combined results
+
+```{r}
+tbl_merge(
+  tbls = list("2003 - 2008" = t1, "2009 - 2013" = t2, "2014 - 2018" = t3)
+)
+ggcoef_compare(
+  list("2003 - 2008" = era1, "2009 - 2013" = era2, "2014 - 2018" = era3),
+  add_reference_rows = T, exponentiate = T, type = "faceted", colour = F
+)
+```

### 2024-11-12 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
fixed table labels

@@
 tbl_merge(
-  tbls = list("2003 - 2008" = t1, "2009 - 2013" = t2, "2014 - 2018" = t3)
+  tbls = list(t1, t2, t3),
+  tab_spanner = c("2003 - 2008", "2009 - 2013", "2014 - 2018")
 )

### 2024-11-12 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
saved nice figures

--- 2024-11-12 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
@@
 )
-vx770_cox_res %>%
+t1 <- vx770_cox_res %>%
   tbl_regression(exponentiate = T)
+t1
 ggcoef_table(vx770_cox_res,
@@
 )
-vx809comb_cox_res %>%
+t2 <- vx809comb_cox_res %>%
   tbl_regression(exponentiate = T)
+t2
 ggcoef_table(vx809comb_cox_res,
@@
   signif_stars = F, colour = F
+)
+```
+
+### Combined results
+
+```{r}
+tbl_merge(
+  tbls = list(t1, t2),
+  tab_spanner = c("VX770", "VX809comb")
+)
+comb_plot = ggcoef_compare(
+  list("VX770" = vx770_cox_res, "VX809comb" = vx809comb_cox_res),
+  add_reference_rows = T, exponentiate = T, type = "faceted", colour = F
+)
+comb_plot
+ggsave("./Dissemination/Figures/vx_combined_forest.png",
+  plot = comb_plot, width = 12, height = 9, units = "in"
 )
@@
 tbl_merge(
-  tbls = list("2003 - 2008" = t1, "2009 - 2013" = t2, "2014 - 2018" = t3)
+  tbls = list(t1, t2, t3),
+  tab_spanner = c("2003 - 2008", "2009 - 2013", "2014 - 2018")
 )
-ggcoef_compare(
+comb_plot = ggcoef_compare(
   list("2003 - 2008" = era1, "2009 - 2013" = era2, "2014 - 2018" = era3),
@@
 )
+comb_plot
+ggsave("./Dissemination/Figures/time_combined_forest.png",
+  plot = comb_plot, width = 12, height = 9, units = "in"
+)
 ```

### 2024-11-14 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
a bunch of bad segmented model code

--- 2024-10-24 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
@@
 library(ggeffects)
+library(emmeans)
 # Paths
@@
 )
+cm <- chngptm(
+  formula.1 = gli_fev1_ppred_rn ~ age_group + Insurance + (1 | eDWID),
+  formula.2 = ~Years, family = "gaussian", data = encounter,
+  type = "segmented", est.method = "grid", var.type = "none"
+)
+
+t <- data.frame(encounter)
+levels(t$Race) <- c("White", "Black", "Hispanic", "Other")
+o <- lme(gli_fev1_ppred_rn ~ Years + Race + age_group + Insurance,
+  random = ~ 1 | eDWID, data = t, na.action = na.omit
+)
+os <- segmented(o, ~Years,
+  data = t,
+  fixed.parms = c(G0 = 0),
+  random = list(eDWID = pdDiag(~ 1 + Years + U))
+)
+
+emt <- emtrends(change_mod, ~ Years + slope_change_years | Race, "Years",
+  at = list(Years = c(-3, 2))
+)
+# EMMs
+emm1 <- emmeans(change_mod,
+  spec = ~ Years | Race,
+  at = list(Years = c(-3, -2, -1, 0)),
+  pbkrtest.limit = Inf
+)
+emm2 <- emmeans(change_mod,
+  spec = ~ Years + slope_change_years | Race,
+  at = list(Years = c(0, 1, 2), slope_change_years = c(0, 1, 2)),
+  pbkrtest.limit = Inf
+)
 # Model plot
@@
 mod_df$ind_pred <- predict(change_mod)
-mod_df$pop_pred <- predict(change_mod, re.form = NA)
+pop_dat <- mod_df %>% select(Years:Insurance)
+pop_dat$pred <- predict(change_mod, re.form = NA, newdata = pop_dat)
+pop_dat <- pop_dat %>% distinct()
 ggplot(mod_df, aes(x = Years)) +
@@
   ) +
-  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+  geom_line(data = pop_dat, aes(y = pred, group = Race, colour = Race)) +
   xlab("Years from ETI Start") +

### 2024-11-14 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
just need contrasts for Kam and Scott

@@
 load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData")
+# Change age group reference level
+encounter$age_group <- relevel(encounter$age_group, ref = "[18,30)")
 ```
@@
 
+# Outcome distributions
+
+## FEV1 % predicted
+
+### GLI 2022 equations
+
+```{r}
+ggplot(encounter, aes(x = gli_fev1_ppred_rn, fill = Race, color = Race)) +
+  geom_density() +
+  facet_wrap(~Race)
+ggplot(encounter, aes(x = gli_fev1_ppred_rn, fill = Race)) +
+  geom_density(alpha = 0.2)
+```
+
+### GLI 2012 equations
+
+```{r}
+ggplot(encounter, aes(x = GLI_FEV1_pct_predicted, fill = Race, color = Race)) +
+  geom_density() +
+  facet_wrap(~Race)
+ggplot(encounter, aes(x = GLI_FEV1_pct_predicted, fill = Race)) +
+  geom_density(alpha = 0.2)
+```
+
+## FVC % predicted
+
+### GLI 2022 equations
+
+```{r}
+ggplot(encounter, aes(x = gli_fvc_ppred_rn, fill = Race, color = Race)) +
+  geom_density() +
+  facet_wrap(~Race)
+ggplot(encounter, aes(x = gli_fvc_ppred_rn, fill = Race)) +
+  geom_density(alpha = 0.2)
+```
+
+### GLI 2012 equations
+
+```{r}
+ggplot(encounter, aes(x = GLI_FVC_pct_predicted, fill = Race, color = Race)) +
+  geom_density() +
+  facet_wrap(~Race)
+ggplot(encounter, aes(x = GLI_FVC_pct_predicted, fill = Race)) +
+  geom_density(alpha = 0.2)
+```
+
+## FEV1/FVC % predicted
+
+### GLI 2022 equations
+
+```{r}
+ggplot(encounter, aes(
+  x = gli_fev1fvc_ppred_rn, fill = Race,
+  color = Race
+)) +
+  geom_density() +
+  facet_wrap(~Race)
+ggplot(encounter, aes(x = gli_fev1fvc_ppred_rn, fill = Race)) +
+  geom_density(alpha = 0.2)
+```
+
+### GLI 2012 equations
+
+```{r}
+ggplot(encounter, aes(x = GLI_FEV1FVC_pct_predicted, fill = Race, color = Race)) +
+  geom_density() +
+  facet_wrap(~Race)
+ggplot(encounter, aes(x = GLI_FEV1FVC_pct_predicted, fill = Race)) +
+  geom_density(alpha = 0.2)
+```
+
+## BMI %ile (age < 20 at ETI start)
+
+```{r}
+encounter %>%
+  filter(age_eti_start < 20) %>%
+  ggplot(aes(x = bmipercentile, fill = Race, color = Race)) +
+  geom_density() +
+  facet_wrap(~Race)
+encounter %>%
+  filter(age_eti_start < 20) %>%
+  ggplot(aes(x = bmipercentile, fill = Race)) +
+  geom_density(alpha = 0.2)
+```
+
+## BMI raw value (age >= 20 at ETI start)
+
+```{r}
+encounter %>%
+  filter(age_eti_start >= 20) %>%
+  ggplot(aes(x = bmivalue, fill = Race, color = Race)) +
+  geom_density() +
+  facet_wrap(~Race)
+encounter %>%
+  filter(age_eti_start >= 20) %>%
+  ggplot(aes(x = bmivalue, fill = Race)) +
+  geom_density(alpha = 0.2)
+```
+
+## Height %ile (age < 20 at ETI start)
+
+```{r}
+encounter %>%
+  filter(age_eti_start < 20) %>%
+  ggplot(aes(x = heightpercentile, fill = Race, color = Race)) +
+  geom_density() +
+  facet_wrap(~Race)
+encounter %>%
+  filter(age_eti_start < 20) %>%
+  ggplot(aes(x = heightpercentile, fill = Race)) +
+  geom_density(alpha = 0.2)
+```
+
+## Height raw value (age >= 20 at ETI start)
+
+```{r}
+encounter %>%
+  filter(age_eti_start >= 20) %>%
+  ggplot(aes(x = height, fill = Race, color = Race)) +
+  geom_density() +
+  facet_wrap(~Race)
+encounter %>%
+  filter(age_eti_start >= 20) %>%
+  ggplot(aes(x = height, fill = Race)) +
+  geom_density(alpha = 0.2)
+```
+
+## Weight %ile (age < 20 at ETI start)
+
+```{r}
+encounter %>%
+  filter(age_eti_start < 20) %>%
+  ggplot(aes(x = weightpercentile, fill = Race, color = Race)) +
+  geom_density() +
+  facet_wrap(~Race)
+encounter %>%
+  filter(age_eti_start < 20) %>%
+  ggplot(aes(x = weightpercentile, fill = Race)) +
+  geom_density(alpha = 0.2)
+```
+
+## Weight raw value (age >= 20 at ETI start)
+
+```{r}
+encounter %>%
+  filter(age_eti_start >= 20) %>%
+  ggplot(aes(x = weight, fill = Race, color = Race)) +
+  geom_density() +
+  facet_wrap(~Race)
+encounter %>%
+  filter(age_eti_start >= 20) %>%
+  ggplot(aes(x = weight, fill = Race)) +
+  geom_density(alpha = 0.2)
+```
+
 # Segmented models
@@
 )
-cm <- chngptm(
-  formula.1 = gli_fev1_ppred_rn ~ age_group + Insurance + (1 | eDWID),
-  formula.2 = ~Years, family = "gaussian", data = encounter,
-  type = "segmented", est.method = "grid", var.type = "none"
-)
-
-t <- data.frame(encounter)
-levels(t$Race) <- c("White", "Black", "Hispanic", "Other")
-o <- lme(gli_fev1_ppred_rn ~ Years + Race + age_group + Insurance,
-  random = ~ 1 | eDWID, data = t, na.action = na.omit
-)
-os <- segmented(o, ~Years,
-  data = t,
-  fixed.parms = c(G0 = 0),
-  random = list(eDWID = pdDiag(~ 1 + Years + U))
-)
-
-emt <- emtrends(change_mod, ~ Years + slope_change_years | Race, "Years",
-  at = list(Years = c(-3, 2))
-)
-# EMMs
-emm1 <- emmeans(change_mod,
-  spec = ~ Years | Race,
-  at = list(Years = c(-3, -2, -1, 0)),
-  pbkrtest.limit = Inf
-)
-emm2 <- emmeans(change_mod,
-  spec = ~ Years + slope_change_years | Race,
-  at = list(Years = c(0, 1, 2), slope_change_years = c(0, 1, 2)),
-  pbkrtest.limit = Inf
-)
 # Model plot
-mod_df <- change_mod@frame
-mod_df$ind_pred <- predict(change_mod)
-pop_dat <- mod_df %>% select(Years:Insurance)
-pop_dat$pred <- predict(change_mod, re.form = NA, newdata = pop_dat)
-pop_dat <- pop_dat %>% distinct()
-ggplot(mod_df, aes(x = Years)) +
-  geom_point(aes(y = gli_fev1_ppred_rn, group = eDWID),
-    alpha = 0.01, show.legend = F, shape = "."
-  ) +
-  geom_line(aes(y = ind_pred, group = eDWID),
-    alpha = 0.01, show.legend = F
-  ) +
-  geom_line(data = pop_dat, aes(y = pred, group = Race, colour = Race)) +
+pop_dat <- data.frame(expand.grid(
+  Years = -3:2, Race = unique(encounter$Race), age_group = "[18,30)",
+  Insurance = "Private"
+))
+pop_dat$slope_change_years <- (pop_dat$Years > 0) * pop_dat$Years
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
   xlab("Years from ETI Start") +
+  ylab("gli_fev1_ppred_rn") +
   theme_classic()
 # Model summary
-change_mod %>% gtsummary::tbl_regression()
+change_mod %>% gtsummary::tbl_regression(intercept = T)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
+  gt() %>%
+  fmt_number(decimals = 2)
 ```
@@
 # Model plot
-mod_df <- change_mod@frame
-mod_df$ind_pred <- predict(change_mod)
-mod_df$pop_pred <- predict(change_mod, re.form = NA)
-ggplot(mod_df, aes(x = Years)) +
-  geom_point(aes(y = GLI_FEV1_pct_predicted, group = eDWID),
-    alpha = 0.01, show.legend = F, shape = "."
-  ) +
-  geom_line(aes(y = ind_pred, group = eDWID),
-    alpha = 0.01, show.legend = F
-  ) +
-  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
   xlab("Years from ETI Start") +
+  ylab("GLI_FEV1_pct_predicted") +
   theme_classic()
 # Model summary
-change_mod %>% gtsummary::tbl_regression()
+change_mod %>% gtsummary::tbl_regression(intercept = T)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
+  gt() %>%
+  fmt_number(decimals = 2)
 ```
@@
 # Model plot
-mod_df <- change_mod@frame
-mod_df$ind_pred <- predict(change_mod)
-mod_df$pop_pred <- predict(change_mod, re.form = NA)
-ggplot(mod_df, aes(x = Years)) +
-  geom_point(aes(y = gli_fev1_ppred_rn, group = eDWID),
-    alpha = 0.01, show.legend = F, shape = "."
-  ) +
-  geom_line(aes(y = ind_pred, group = eDWID),
-    alpha = 0.01, show.legend = F
-  ) +
-  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
   xlab("Years from ETI Start") +
+  ylab("gli_fev1_ppred_rn") +
   theme_classic()
 # Model summary
-change_mod %>% gtsummary::tbl_regression()
+change_mod %>% gtsummary::tbl_regression(intercept = T)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
+  gt() %>%
+  fmt_number(decimals = 2)
 ```
@@
 # Model plot
-mod_df <- change_mod@frame
-mod_df$ind_pred <- predict(change_mod)
-mod_df$pop_pred <- predict(change_mod, re.form = NA)
-ggplot(mod_df, aes(x = Years)) +
-  geom_point(aes(y = gli_fvc_ppred_rn, group = eDWID),
-    alpha = 0.01, show.legend = F, shape = "."
-  ) +
-  geom_line(aes(y = ind_pred, group = eDWID),
-    alpha = 0.01, show.legend = F
-  ) +
-  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
   xlab("Years from ETI Start") +
+  ylab("gli_fvc_ppred_rn") +
   theme_classic()
 # Model summary
-change_mod %>% gtsummary::tbl_regression()
+change_mod %>% gtsummary::tbl_regression(intercept = T)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(gli_fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
+  gt() %>%
+  fmt_number(decimals = 2)
 ```
@@
 # Model plot
-mod_df <- change_mod@frame
-mod_df$ind_pred <- predict(change_mod)
-mod_df$pop_pred <- predict(change_mod, re.form = NA)
-ggplot(mod_df, aes(x = Years)) +
-  geom_point(aes(y = GLI_FVC_pct_predicted, group = eDWID),
-    alpha = 0.01, show.legend = F, shape = "."
-  ) +
-  geom_line(aes(y = ind_pred, group = eDWID),
-    alpha = 0.01, show.legend = F
-  ) +
-  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
   xlab("Years from ETI Start") +
+  ylab("GLI_FVC_pct_predicted") +
   theme_classic()
 # Model summary
-change_mod %>% gtsummary::tbl_regression()
+change_mod %>% gtsummary::tbl_regression(intercept = T)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(GLI_FVC_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
+  gt() %>%
+  fmt_number(decimals = 2)
 ```
@@
 # Model plot
-mod_df <- change_mod@frame
-mod_df$ind_pred <- predict(change_mod)
-mod_df$pop_pred <- predict(change_mod, re.form = NA)
-ggplot(mod_df, aes(x = Years)) +
-  geom_point(aes(y = gli_fev1fvc_ppred_rn, group = eDWID),
-    alpha = 0.01, show.legend = F, shape = "."
-  ) +
-  geom_line(aes(y = ind_pred, group = eDWID),
-    alpha = 0.01, show.legend = F
-  ) +
-  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
   xlab("Years from ETI Start") +
+  ylab("gli_fev1fvc_ppred_rn") +
   theme_classic()
 # Model summary
-change_mod %>% gtsummary::tbl_regression()
+change_mod %>% gtsummary::tbl_regression(intercept = T)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(gli_fev1fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
+  gt() %>%
+  fmt_number(decimals = 2)
 ```
@@
 # Model plot
-mod_df <- change_mod@frame
-mod_df$ind_pred <- predict(change_mod)
-mod_df$pop_pred <- predict(change_mod, re.form = NA)
-ggplot(mod_df, aes(x = Years)) +
-  geom_point(aes(y = GLI_FEV1FVC_pct_predicted, group = eDWID),
-    alpha = 0.01, show.legend = F, shape = "."
-  ) +
-  geom_line(aes(y = ind_pred, group = eDWID),
-    alpha = 0.01, show.legend = F
-  ) +
-  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
   xlab("Years from ETI Start") +
+  ylab("GLI_FEV1FVC_pct_predicted") +
   theme_classic()
 # Model summary
-change_mod %>% gtsummary::tbl_regression()
+change_mod %>% gtsummary::tbl_regression(intercept = T)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(GLI_FEV1FVC_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
+  gt() %>%
+  fmt_number(decimals = 2)
 ```
@@
 # Model plot
-mod_df <- change_mod@frame
-mod_df$ind_pred <- predict(change_mod)
-mod_df$pop_pred <- predict(change_mod, re.form = NA)
-ggplot(mod_df, aes(x = Years)) +
-  geom_point(aes(y = GLI_FEV1FVC_pct_predicted, group = eDWID),
-    alpha = 0.01, show.legend = F, shape = "."
-  ) +
-  geom_line(aes(y = ind_pred, group = eDWID),
-    alpha = 0.01, show.legend = F
-  ) +
-  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
   xlab("Years from ETI Start") +
+  ylab("GLI_FEV1FVC_pct_predicted") +
   theme_classic()
 # Model summary
-change_mod %>% gtsummary::tbl_regression()
+change_mod %>% gtsummary::tbl_regression(intercept = T)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(GLI_FEV1FVC_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
+  gt() %>%
+  fmt_number(decimals = 2)
 ```
 
-## BMI %ile
+## BMI
 
-### Full cohort
+### BMI %ile (age < 20 at ETI start)
 
@@
     slope_change_years * Race + age_group + Insurance + (1 | eDWID),
-  data = encounter
+  data = encounter[encounter$age_eti_start < 20, ]
 )
 # Model plot
-mod_df <- change_mod@frame
-mod_df$ind_pred <- predict(change_mod)
-mod_df$pop_pred <- predict(change_mod, re.form = NA)
-ggplot(mod_df, aes(x = Years)) +
-  geom_point(aes(y = bmipercentile, group = eDWID),
-    alpha = 0.01, show.legend = F, shape = "."
-  ) +
-  geom_line(aes(y = ind_pred, group = eDWID),
-    alpha = 0.01, show.legend = F
-  ) +
-  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
   xlab("Years from ETI Start") +
+  ylab("bmipercentile") +
   theme_classic()
 # Model summary
-change_mod %>% gtsummary::tbl_regression()
+change_mod %>% gtsummary::tbl_regression(intercept = T)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(bmipercentile = pred, ci.ll = ll, ci.ul = ul) %>%
+  gt() %>%
+  fmt_number(decimals = 2)
 ```
 
-### Age < 18 at ETI initiation
-
-The following models were not adjusted for age group.
+### BMI raw value (age >= 20 at ETI start)
 
 ```{r}
-# Fit mixed model with change point
 change_mod <- lmer(
-  bmipercentile ~ Years * Race + slope_change_years * Race +
-    Insurance + (1 | eDWID),
-  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
+  bmivalue ~ Years * Race +
+    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_start >= 20, ]
 )
 # Model plot
-mod_df <- change_mod@frame
-mod_df$ind_pred <- predict(change_mod)
-mod_df$pop_pred <- predict(change_mod, re.form = NA)
-ggplot(mod_df, aes(x = Years)) +
-  geom_point(aes(y = bmipercentile, group = eDWID),
-    alpha = 0.01, show.legend = F, shape = "."
-  ) +
-  geom_line(aes(y = ind_pred, group = eDWID),
-    alpha = 0.01, show.legend = F
-  ) +
-  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
   xlab("Years from ETI Start") +
+  ylab("bmivalue") +
   theme_classic()
 # Model summary
-change_mod %>% gtsummary::tbl_regression()
+change_mod %>% gtsummary::tbl_regression(intercept = T)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(bmivalue = pred, ci.ll = ll, ci.ul = ul) %>%
+  gt() %>%
+  fmt_number(decimals = 2)
 ```
 
-### Age >= 18 at ETI initiation
+## Height
 
-The following models were not adjusted for age group.
-
-```{r}
-# Fit mixed model with change point
-change_mod <- lmer(
-  bmipercentile ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
-  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
-)
-# Model plot
-mod_df <- change_mod@frame
-mod_df$ind_pred <- predict(change_mod)
-mod_df$pop_pred <- predict(change_mod, re.form = NA)
-ggplot(mod_df, aes(x = Years)) +
-  geom_point(aes(y = bmipercentile, group = eDWID),
-    alpha = 0.01, show.legend = F, shape = "."
-  ) +
-  geom_line(aes(y = ind_pred, group = eDWID),
-    alpha = 0.01, show.legend = F
-  ) +
-  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
-  xlab("Years from ETI Start") +
-  theme_classic()
-# Model summary
-change_mod %>% gtsummary::tbl_regression()
-```
-
-## Height %ile
-
-### Full cohort
+### Height %ile (age < 20 at ETI start)
 
@@
     slope_change_years * Race + age_group + Insurance + (1 | eDWID),
-  data = encounter
+  data = encounter[encounter$age_eti_start < 20, ]
 )
 # Model plot
-mod_df <- change_mod@frame
-mod_df$ind_pred <- predict(change_mod)
-mod_df$pop_pred <- predict(change_mod, re.form = NA)
-ggplot(mod_df, aes(x = Years)) +
-  geom_point(aes(y = heightpercentile, group = eDWID),
-    alpha = 0.01, show.legend = F, shape = "."
-  ) +
-  geom_line(aes(y = ind_pred, group = eDWID),
-    alpha = 0.01, show.legend = F
-  ) +
-  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
   xlab("Years from ETI Start") +
+  ylab("heightpercentile") +
   theme_classic()
 # Model summary
-change_mod %>% gtsummary::tbl_regression()
+change_mod %>% gtsummary::tbl_regression(intercept = T)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(heightpercentile = pred, ci.ll = ll, ci.ul = ul) %>%
+  gt() %>%
+  fmt_number(decimals = 2)
 ```
 
-### Age < 18 at ETI initiation
-
-The following models were not adjusted for age group.
+### Height raw value (age >= 20 at ETI start)
 
 ```{r}
-# Fit mixed model with change point
 change_mod <- lmer(
-  heightpercentile ~ Years * Race + slope_change_years * Race +
-    Insurance + (1 | eDWID),
-  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
+  height ~ Years * Race +
+    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_start >= 20, ]
 )
 # Model plot
-mod_df <- change_mod@frame
-mod_df$ind_pred <- predict(change_mod)
-mod_df$pop_pred <- predict(change_mod, re.form = NA)
-ggplot(mod_df, aes(x = Years)) +
-  geom_point(aes(y = heightpercentile, group = eDWID),
-    alpha = 0.01, show.legend = F, shape = "."
-  ) +
-  geom_line(aes(y = ind_pred, group = eDWID),
-    alpha = 0.01, show.legend = F
-  ) +
-  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
   xlab("Years from ETI Start") +
+  ylab("height") +
   theme_classic()
 # Model summary
-change_mod %>% gtsummary::tbl_regression()
+change_mod %>% gtsummary::tbl_regression(intercept = T)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(height = pred, ci.ll = ll, ci.ul = ul) %>%
+  gt() %>%
+  fmt_number(decimals = 2)
 ```
 
-### Age >= 18 at ETI initiation
+## Weight
 
-The following models were not adjusted for age group.
-
-```{r}
-# Fit mixed model with change point
-change_mod <- lmer(
-  heightpercentile ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
-  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
-)
-# Model plot
-mod_df <- change_mod@frame
-mod_df$ind_pred <- predict(change_mod)
-mod_df$pop_pred <- predict(change_mod, re.form = NA)
-ggplot(mod_df, aes(x = Years)) +
-  geom_point(aes(y = heightpercentile, group = eDWID),
-    alpha = 0.01, show.legend = F, shape = "."
-  ) +
-  geom_line(aes(y = ind_pred, group = eDWID),
-    alpha = 0.01, show.legend = F
-  ) +
-  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
-  xlab("Years from ETI Start") +
-  theme_classic()
-# Model summary
-change_mod %>% gtsummary::tbl_regression()
-```
-
-## Weight %ile
-
-### Full cohort
+### Weight %ile (age < 20 at ETI start)
 
@@
     slope_change_years * Race + age_group + Insurance + (1 | eDWID),
-  data = encounter
+  data = encounter[encounter$age_eti_start < 20, ]
 )
 # Model plot
-mod_df <- change_mod@frame
-mod_df$ind_pred <- predict(change_mod)
-mod_df$pop_pred <- predict(change_mod, re.form = NA)
-ggplot(mod_df, aes(x = Years)) +
-  geom_point(aes(y = weightpercentile, group = eDWID),
-    alpha = 0.01, show.legend = F, shape = "."
-  ) +
-  geom_line(aes(y = ind_pred, group = eDWID),
-    alpha = 0.01, show.legend = F
-  ) +
-  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
   xlab("Years from ETI Start") +
+  ylab("weightpercentile") +
   theme_classic()
 # Model summary
-change_mod %>% gtsummary::tbl_regression()
+change_mod %>% gtsummary::tbl_regression(intercept = T)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(weightpercentile = pred, ci.ll = ll, ci.ul = ul) %>%
+  gt() %>%
+  fmt_number(decimals = 2)
 ```
 
-### Age < 18 at ETI initiation
-
-The following models were not adjusted for age group.
+### Weight raw value (age >= 20 at ETI start)
 
 ```{r}
-# Fit mixed model with change point
 change_mod <- lmer(
-  weightpercentile ~ Years * Race + slope_change_years * Race +
-    Insurance + (1 | eDWID),
-  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
+  weight ~ Years * Race +
+    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  data = encounter[encounter$age_eti_start >= 20, ]
 )
 # Model plot
-mod_df <- change_mod@frame
-mod_df$ind_pred <- predict(change_mod)
-mod_df$pop_pred <- predict(change_mod, re.form = NA)
-ggplot(mod_df, aes(x = Years)) +
-  geom_point(aes(y = weightpercentile, group = eDWID),
-    alpha = 0.01, show.legend = F, shape = "."
-  ) +
-  geom_line(aes(y = ind_pred, group = eDWID),
-    alpha = 0.01, show.legend = F
-  ) +
-  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
   xlab("Years from ETI Start") +
+  ylab("weight") +
   theme_classic()
 # Model summary
-change_mod %>% gtsummary::tbl_regression()
+change_mod %>% gtsummary::tbl_regression(intercept = T)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(weight = pred, ci.ll = ll, ci.ul = ul) %>%
+  gt() %>%
+  fmt_number(decimals = 2)
 ```
-
-### Age >= 18 at ETI initiation
-
-The following models were not adjusted for age group.
-
-```{r}
-# Fit mixed model with change point
-change_mod <- lmer(
-  weightpercentile ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
-  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
-)
-# Model plot
-mod_df <- change_mod@frame
-mod_df$ind_pred <- predict(change_mod)
-mod_df$pop_pred <- predict(change_mod, re.form = NA)
-ggplot(mod_df, aes(x = Years)) +
-  geom_point(aes(y = weightpercentile, group = eDWID),
-    alpha = 0.01, show.legend = F, shape = "."
-  ) +
-  geom_line(aes(y = ind_pred, group = eDWID),
-    alpha = 0.01, show.legend = F
-  ) +
-  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
-  xlab("Years from ETI Start") +
-  theme_classic()
-# Model summary
-change_mod %>% gtsummary::tbl_regression()
-```

### 2024-11-15 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
updates for Scott Sagel

@@
 library(lmerTest)
-library(splines)
-library(ggeffects)
 library(emmeans)

### 2024-11-15 Scott Sagel/PEx in CF/analysis_for_yaron.qmd
updates for Scott Sagel

@@
   geom_label(aes(label = n), nudge_y = -20, fill = "white") +
-  ylab("Num. of Hospitalizations") +
+  ylab("Number of Hospitalizations") +
   theme_bw() +
@@
   ) +
-  ylab("Percent With >= 2 Hospitalizations") +
+  ylab("Percent with >= 2 Hospitalizations in a Single Year") +
   theme_bw() +
@@
 #| fig-cap: Hospitalizations
-fig1 + fig2
+fig1 + fig2 + plot_annotation(tag_levels = c("a", "b"))
 ```

### 2024-11-19 Jordana Hoppe/Home Collection/manuscript.qmd
plot updates for Jordana

@@
 github_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/GitHub",
+  "Darwin" = "/Users/timvigers/Documents/GitHub",
   "Windows" = "C:/Users/Tim/GitHub",
@@
 labels["age"] <- "Age at Clinic Visit"
+labels <- as.list(labels)
 # Genotype format
@@
   d %>%
-    mutate(across(everything(), ~ as.numeric(. == "Present" | . == "No Growth"))) %>%
+    mutate(across(
+      everything(),
+      ~ as.numeric(. == "Present" | . == "No Growth")
+    )) %>%
     drop_na(.) %>%
-    euler(.) %>%
+    venn(.) %>%
     plot(., quantities = T) %>%
     print(.)
+  cat("\n")
+  cat("\n")
+  cat("#### Bar plot")
+  cat("\n")
+  cat("\n")
+  # Format data
+  bar_plot_df <- d %>%
+    rowwise() %>%
+    mutate(
+      "Immediate vs. Delayed" = case_when(
+        Immediate == "Present" & Delayed == "Present" ~ "Both Present",
+        Immediate == "Absent" & Delayed == "Absent" ~ "Both Absent",
+        .default = "Disagreed"
+      ),
+      "Immediate vs. Home" = case_when(
+        Immediate == "Present" & Home == "Present" ~ "Both Present",
+        Immediate == "Absent" & Home == "Absent" ~ "Both Absent",
+        .default = "Disagreed"
+      ),
+      "Delayed vs. Home" = case_when(
+        Delayed == "Present" & Home == "Present" ~ "Both Present",
+        Delayed == "Absent" & Home == "Absent" ~ "Both Absent",
+        .default = "Disagreed"
+      ),
+    ) %>%
+    select(-c(Immediate:Home)) %>%
+    pivot_longer(cols = `Immediate vs. Delayed`:`Delayed vs. Home`) %>%
+    group_by(name) %>%
+    count(value) %>%
+    mutate(pct = (n / sum(n)) * 100)
+  # Plot
+  p = ggplot(bar_plot_df, aes(x = name, fill = value, y = n)) +
+    geom_bar(stat = "identity") +
+    geom_text(aes(label = paste0(sprintf("%1.1f", pct), "%")),
+      position = position_stack(vjust = 0.5)
+    ) +
+    ylab("Samples (N)") +
+    theme_classic() +
+    theme(legend.title = element_blank(), axis.title.x = element_blank())
+  cat("\n")
+  cat("\n")
+  show(p)
   cat("\n")

### 2024-11-19 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
plot updates for Jordana

--- 2024-09-19 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
@@
 })
+db_demographics$Race <- factor(db_demographics$Race,
+  levels = c(
+    "White", "Other", "Multiple Races", "American Indian or Alaskan Native"
+  )
+)
 db_demographics$Ethnicity <- apply(db_demographics, 1, function(r) {
@@
 })
+db_demographics$Ethnicity <- factor(db_demographics$Ethnicity,
+  levels = c("Non-Hispanic", "Hispanic", "Unknown/Not Reported")
+)
+# Drop empty levels
+db_demographics <- droplevels(db_demographics)
 # Add groups to other dataframes
@@
     quantities = list(cex = 3),
-    fills = list(fill = c("#FFC107", "#1E88E5"))
+    fills = list(fill = c("#FFC107", "#1E88E5", "#004D40"))
   ))

### 2024-11-19 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
SSwAB-IT fixed age

@@
 db_demographics <- left_join(db_demographics,
-  df %>% select(record_id, cos_result_date),
+  db_clinic_obtained_values %>% select(record_id, cos_collection_date),
   by = join_by(record_id)
 )
-db_demographics$age_at_cos <- as.numeric(difftime(db_demographics$cos_result_date,
+db_demographics$age_at_cos <- as.numeric(difftime(db_demographics$cos_collection_date,
   db_demographics$dob,
@@
   tbl_summary(
-    by = group,
+    by = group, missing_text = "Missing",
     statistic = all_continuous() ~ c("{median} ({min}, {max})"),
@@
   ) %>%
-  tbl_summary(by = group) %>%
+  tbl_summary(by = group, missing_text = "Missing") %>%
   add_overall() %>%
@@
     select(agree, detailed_agree, Group) %>%
-    tbl_summary(by = Group, label = list(
+    tbl_summary(by = Group, missing_text = "Missing", label = list(
       agree ~ "Agree (Both Positive/Positive and Negative/Negative)",

### 2024-11-22 Christine Chan/EnVision CF/katie_dataset_check.qmd
dataset EDA for katie

---
title: "EnVision Data Summary"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
csl: /Users/timvigers/Documents/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(arsenal)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF")
```

```{r data cleaning}
# Import cleaned dataset
df <- read.csv("./Data_Clean/analysis_dataset.csv", na.strings = "")
# Set table options
mycontrols <-
  tableby.control(
    numeric.stats = c("N", "Nmiss", "meansd", "medianq1q3", "range"),
    cat.stats = c("N", "Nmiss", "countpct"),
    digits = 1
  )
```

```{r results='asis'}
#| label: tbl-auc-numbers-overall
#| tbl-cap: Glucose and Insulin iAUC
tbl <- tableby(
  Diagnosis ~ iAUC30gluc + iAUC60gluc + iAUC120gluc + iAUC180gluc + iAUC30ins +
    iAUC60ins + iAUC120ins + iAUC180ins,
  data = df, control = mycontrols
)
summary(tbl, pfootnote = T)
```

```{r results='asis'}
#| label: fig-auc-hist
#| fig-cap: Glucose and Insulin iAUC Histograms
df %>%
  select(
    iAUC30gluc, iAUC60gluc, iAUC120gluc, iAUC180gluc, iAUC30ins, iAUC60ins,
    iAUC120ins, iAUC180ins
  ) %>%
  pivot_longer(everything()) %>%
  mutate(
    measure = factor(sub("iAUC\\d{2,3}", "", name),
      levels = c("gluc", "ins"), labels = c("Glucose", "Insulin")
    ),
    name = factor(sub("gluc|ins", "", name),
      levels = c("iAUC30", "iAUC60", "iAUC120", "iAUC180")
    )
  ) %>%
  ggplot(., aes(x = name, y = value)) +
  geom_boxplot() +
  facet_wrap(~measure)
```

### 2024-11-22 Christine Chan/EnVision CF/clean_envision_data.R

@@ -290,7 +290,6 @@ final_df <- full_join(final_df, catecholamines)
 final_df <- final_df %>% pivot_wider(
   names_from = Timepoint, values_from = c(Glucose:Epinephrine)
 )
-final_df <- full_join(final_df, cgm)
 final_df <- full_join(final_df, visits)
 final_df <- full_join(final_df, hypo_symptoms)
 # Add OGTT data
@@ -302,6 +301,8 @@ final_df <- final_df %>%
   select(study_id, Date, ogtt_num, everything())
 final_df <- full_join(final_df, ogtts)
 final_df <- final_df %>% arrange(study_id, Date)
+# Add CGM
+final_df <- full_join(final_df, cgm)
 # Add HbA1c, etc.
 a1c <- redcap %>%
   select(study_id, date_visit, a1c_result) %>%
@@ -344,7 +345,7 @@ final_df$Diagnosis <- factor(final_df$Diagnosis,
 # iAUCs
 final_df$iAUC30gluc <- apply(final_df, 1, function(r) {
   y <- as.numeric(r[glucose[1:3]]) - as.numeric(r[glucose[1]])
-  if (is.na(last(y))) {
+  if (is.na(first(y)) | is.na(last(y))) {
     return(NA)
   } else {
     x <- c(0, 10, 30)[!is.na(y)]
@@ -355,7 +356,7 @@ final_df$iAUC30gluc <- apply(final_df, 1, function(r) {
 })
 final_df$iAUC60gluc <- apply(final_df, 1, function(r) {
   y <- as.numeric(r[glucose[1:4]]) - as.numeric(r[glucose[1]])
-  if (is.na(last(y))) {
+  if (is.na(first(y)) | is.na(last(y))) {
     return(NA)
   } else {
     x <- c(0, 10, 30, 60)[!is.na(y)]
@@ -366,7 +367,7 @@ final_df$iAUC60gluc <- apply(final_df, 1, function(r) {
 })
 final_df$iAUC120gluc <- apply(final_df, 1, function(r) {
   y <- as.numeric(r[glucose[1:6]]) - as.numeric(r[glucose[1]])
-  if (is.na(last(y))) {
+  if (is.na(first(y)) | is.na(last(y))) {
     return(NA)
   } else {
     x <- c(0, 10, 30, 60, 90, 120)[!is.na(y)]
[…]

### 2024-11-22 Christine Chan/EnVision CF/katie_dataset_check.qmd
basic AUC analysis for Katie

@@
 library(arsenal)
+library(skimr)
 knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF")
@@
 # Import cleaned dataset
-df <- read.csv("./Data_Clean/analysis_dataset.csv", na.strings = "")
+df <- read.csv("./Data_Clean/envision_analysis_dataset.csv", na.strings = "")
+# Find those missing both iAUC120ins and iAUC120gluc, and both iAUC180ins and
+# iAUC180gluc
+df$miss_120 <- factor(is.na(df$iAUC120gluc) & is.na(df$iAUC120ins),
+  levels = c(F, T), labels = c("No", "Yes")
+)
+df$miss_180 <- factor(is.na(df$iAUC180gluc) & is.na(df$iAUC180ins),
+  levels = c(F, T), labels = c("No", "Yes")
+)
+# Labels
+mylabels <- list(
+  miss_120 = "Missing iAUC120gluc and iAUC120ins?",
+  miss_180 = "Missing iAUC180gluc and iAUC180ins?"
+)
 # Set table options
@@
 
+# Demographics
+
 ```{r results='asis'}
-#| label: tbl-auc-numbers-overall
-#| tbl-cap: Glucose and Insulin iAUC
+#| label: tbl-demo
+#| tbl-cap: Demographics by Participant
 tbl <- tableby(
-  Diagnosis ~ iAUC30gluc + iAUC60gluc + iAUC120gluc + iAUC180gluc + iAUC30ins +
-    iAUC60ins + iAUC120ins + iAUC180ins,
+  ~ age_visit + height + weight + bmi + bmi_perc + origin_race + ethnicity +
+    sex + CFTR,
+  data = df %>% group_by(study_id) %>% slice_min(row_number()),
+  control = mycontrols
+)
+summary(tbl, pfootnote = T)
+```
+
+```{r results='asis'}
+#| label: tbl-demo-ogtt
+#| tbl-cap: Demographics Treating OGTTs as Independent
+tbl <- tableby(
+  ~ age_visit + height + weight + bmi + bmi_perc + origin_race + ethnicity +
+    sex + CFTR,
   data = df, control = mycontrols
@@
 
+# iAUC
+
+```{r results='asis'}
+#| label: tbl-auc-numbers-overall
+#| tbl-cap: Glucose and Insulin iAUC
+tbl <- tableby(
+  ~ iAUC30gluc + iAUC60gluc + iAUC120gluc + iAUC180gluc + iAUC30ins +
+    iAUC60ins + iAUC120ins + iAUC180ins + miss_120 + miss_180,
+  data = df, control = mycontrols
+)
+summary(tbl, pfootnote = T, labelTranslations = mylabels)
+```
+
+```{r results='asis'}
+#| label: tbl-auc-numbers-diagnosis
+#| tbl-cap: Glucose and Insulin iAUC by Diagnosis
+tbl <- tableby(
+  Diagnosis ~ iAUC30gluc + iAUC60gluc + iAUC120gluc + iAUC180gluc + iAUC30ins +
+    iAUC60ins + iAUC120ins + iAUC180ins + miss_120 + miss_180,
+  data = df, control = mycontrols
+)
+summary(tbl, pfootnote = T, labelTranslations = mylabels)
+```
+
 ```{r results='asis'}
 #| label: fig-auc-hist
-#| fig-cap: Glucose and Insulin iAUC Histograms
+#| fig-cap: Glucose and Insulin iAUC Boxplots
 df %>%
@@
   ) %>%
+  filter(!is.na(value)) %>%
   ggplot(., aes(x = name, y = value)) +
@@
 ```
+
+# EDA
+
+```{r}
+#| label: fig-skimr
+#| fig-cap: Data Skim
+skim(df)
+```

### 2024-11-22 Christine Chan/EnVision CF/clean_envision_data.R

@@ -469,6 +469,6 @@ final_df$redcap_data_access_group <- factor(
 )
 # Write
 write.csv(final_df,
-  file = "./Christine Chan/EnVision CF/Data_Clean/analysis_dataset.csv",
+  file = "./Christine Chan/EnVision CF/Data_Clean/envision_analysis_dataset.csv",
   row.names = F, na = ""
 )

### 2024-11-22 Jordana Hoppe/Home Collection/manuscript.qmd
Home collection figures updated for Jordana

@@
 library(caret)
-library(likert)
+library(ggstats)
+library(RColorBrewer)
 library(patchwork)
@@
   cat("\n")
-  d %>%
+  venn <- d %>%
     mutate(across(
@@
     venn(.) %>%
-    plot(., quantities = T) %>%
-    print(.)
+    plot(., quantities = T)
   cat("\n")
@@
   # Plot
-  p = ggplot(bar_plot_df, aes(x = name, fill = value, y = n)) +
+  bar_plot <- ggplot(bar_plot_df, aes(x = name, fill = value, y = n)) +
     geom_bar(stat = "identity") +
@@
     ylab("Samples (N)") +
+    scale_fill_manual(values = brewer.pal(3, "Accent")) +
     theme_classic() +
     theme(legend.title = element_blank(), axis.title.x = element_blank())
+  p <- bar_plot + venn + plot_annotation(tag_levels = c("a", "b"))
   cat("\n")
@@
   cat("\n")
+  # Save MSSA plot
+  if (result_pattern == "results___1") {
+    ggsave("Jordana Hoppe/Home Collection/Dissemination/Figures/mssa.png",
+      plot = p,
+      width = 12, height = , units = "in"
+    )
+    bar_plot <- ggplot(
+      bar_plot_df[bar_plot_df$name != "Delayed vs. Home", ],
+      aes(x = name, fill = value, y = n)
+    ) +
+      geom_bar(stat = "identity") +
+      geom_text(aes(label = paste0(sprintf("%1.1f", pct), "%")),
+        position = position_stack(vjust = 0.5)
+      ) +
+      ylab("Samples (N)") +
+      scale_fill_manual(values = brewer.pal(3, "Accent")) +
+      theme_classic() +
+      theme(legend.title = element_blank(), axis.title.x = element_blank())
+    p <- bar_plot + venn + plot_annotation(tag_levels = c("a", "b"))
+    ggsave("Jordana Hoppe/Home Collection/Dissemination/Figures/mssa_2.png",
+      plot = p,
+      width = 12, height = , units = "in"
+    )
+  }
   cat("### Pairwise agreement")
@@
 yn <- c(
-  "q1.factor", "q3.factor", "previous.factor", "goodsample.factor", "q4.factor",
-  "q5.factor", "q6.factor"
+  "q1.factor", "q3.factor", "q4.factor", "goodsample.factor",
+  "q6.factor", "q5.factor", "previous.factor"
 )
@@
   rename(any_of(setNames(names(labels), labels)))
-yn_df <- data.frame(lapply(yn_df, factor, levels = c("No", "Yes")), check.names = F)
-p <- plot(likert(yn_df))
+yn_df <- data.frame(lapply(yn_df, factor, levels = c("No", "Yes")),
+  check.names = F
+)
+p <- gglikert(yn_df, sort = "none", cutoff = 1, add_totals = F) +
+  scale_fill_manual(values = c("No" = "#D8B365", "Yes" = "#5AB4AC"))
+p
 ggsave("Jordana Hoppe/Home Collection/Dissemination/Figures/yn.png",
@@
 
-p_benefits <- plot(likert(benefits_df)) + ggtitle("Benefits") +
+p_benefits <- gglikert(benefits_df, add_totals = F) +
+  ggtitle("Benefits") +
   theme(
@@
     legend.position = "none"
-  )
-p_barriers <- plot(likert(barriers_df)) + ggtitle("Barriers")
+  ) +
+  scale_fill_manual(values = c("Unchecked" = "#D8B365", "Checked" = "#5AB4AC"))
+p_barriers <- gglikert(barriers_df, reverse_likert = T, add_totals = F) +
+  ggtitle("Barriers") +
+  scale_fill_manual(values = c("Unchecked" = "#5AB4AC", "Checked" = "#D8B365"))
 p <- p_benefits / p_barriers
+p
 ggsave("Jordana Hoppe/Home Collection/Dissemination/Figures/check.png",
@@
   rename(any_of(setNames(names(labels), labels)))
-p <- plot(likert(df3))
+p3 <- gglikert(as_tibble(df3))
+p3
 ggsave("Jordana Hoppe/Home Collection/Dissemination/Figures/q2.png",
-  plot = p,
+  plot = p3,
   width = 8, height = 6, units = "in"
@@
   rename(any_of(setNames(names(labels), labels)))
-p <- plot(likert(df5))
+p5 <- gglikert(df5)
+p5
 ggsave("Jordana Hoppe/Home Collection/Dissemination/Figures/difficulty.png",
+  plot = p5,
+  width = 9, height = 6, units = "in"
+)
+p <- p3 / p5 + plot_annotation(tag_levels = c("a", "b"))
+p
+ggsave("Jordana Hoppe/Home Collection/Dissemination/Figures/fig2.png",
   plot = p,

### 2024-12-05 Christine Chan/CF Registry/cf_registry_CFRD_prediction.qmd
updates for Monique

--- 2024-10-28 Christine Chan/CF Registry/cf_registry_a1c_CFRD_analysis.qmd
@@
 ---
-title: "HbA1c and CFRD"
+title: "Using HbA1c and Fasting BG to Predict Development of CFRD"
 author: "Tim Vigers"
@@
 library(gtsummary)
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
-  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
-  "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/CF Registry"
-)
-github_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/GitHub",
-  "Windows" = "C:/Users/Tim/GitHub",
-  "Linux" = "/home/tim/GitHub"
-)
-knitr::opts_knit$set(root.dir = home_dir)
+knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry")
 ```

### 2024-12-06 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
updated models and contrasts

@@
 library(Hmisc)
+library(multcomp)
 library(tidyverse)
@@
 library(lmerTest)
-library(emmeans)
-# Paths
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
-)
-knitr::opts_knit$set(root.dir = home_dir)
+library(broom)
+knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF")
 ```
@@
 label(encounter$slope_change_years) <- "Change in Slope"
+# Prior modulator use
+prior_mods <- encounter %>%
+  select(eDWID, Days, vx661comb, vx445comb, Vx770, Vx809comb) %>%
+  filter(Days < 0) %>%
+  group_by(eDWID) %>%
+  summarise(prior_modulator = any(vx661comb == 1) | any(vx445comb == 1) |
+    any(Vx770 == 1) | any(Vx809comb))
+prior_mods$prior_modulator <-
+  fct_na_value_to_level(factor(prior_mods$prior_modulator), level = "No")
+prior_mods$prior_modulator <- factor(prior_mods$prior_modulator,
+  levels = c(T, "No"),
+  labels = c("Yes", "No")
+)
+encounter <- left_join(encounter, prior_mods, by = join_by(eDWID))
+# On enzymes for pancreatic insufficiency
+encounter$isOnEnzymes <-
+  factor(encounter$isOnEnzymes, levels = 0:1, labels = c("No", "Yes"))
 ```
@@
 change_mod <- lmer(
-  gli_fev1_ppred_rn ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + age_group +
+    Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter
@@
   Years = -3:2, Race = unique(encounter$Race), age_group = "[18,30)",
-  Insurance = "Private"
+  Insurance = "Private", prior_modulator = "Yes",
+  isOnEnzymes = "Yes"
 ))
@@
   rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
-  gt() %>%
-  fmt_number(decimals = 2)
+  gt(caption = "Predicted Values at Various Timepoints") %>%
+  fmt_number(decimals = 3)
+# Linear hypotheses
+white_vs_black_pre <- glht(
+  change_mod,
+  "Years + `Years:RaceBlack or African American` = 1.195465"
+)
+white_vs_black_pre <- tidy(summary(white_vs_black_pre))
+white_vs_black_pre$contrast <- "Black vs. White Pre-ETI Slope"
+white_vs_black_pre %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+white_vs_black_post <- glht(
+  change_mod,
+  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.738885"
+)
+white_vs_black_post <- tidy(summary(white_vs_black_post))
+white_vs_black_post$contrast <- "Black vs. White Post-ETI Slope"
+white_vs_black_post %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+
+white_vs_hispanic_pre <- glht(
+  change_mod,
+  "Years + `Years:RaceHispanic or Latino` = 1.19547"
+)
+white_vs_hispanic_pre <- tidy(summary(white_vs_hispanic_pre))
+white_vs_hispanic_pre$contrast <- "Hispanic vs. White Pre-ETI Slope"
+white_vs_hispanic_pre %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+white_vs_hispanic_post <- glht(
+  change_mod,
+  "Years + `Years:RaceHispanic or Latino` + slope_change_years + `RaceHispanic or Latino:slope_change_years` = 2.738885"
+)
+white_vs_hispanic_post <- tidy(summary(white_vs_hispanic_post))
+white_vs_hispanic_post$contrast <- "Hispanic vs. White Post-ETI Slope"
+white_vs_hispanic_post %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+
+white_vs_other_pre <- glht(
+  change_mod,
+  "Years + `Years:RaceOther` = 1.19547"
+)
+white_vs_other_pre <- tidy(summary(white_vs_other_pre))
+white_vs_other_pre$contrast <- "Other vs. White Pre-ETI Slope"
+white_vs_other_pre %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+white_vs_other_post <- glht(
+  change_mod,
+  "Years + `Years:RaceOther` + slope_change_years + `RaceOther:slope_change_years` = 2.738885"
+)
+white_vs_other_post <- tidy(summary(white_vs_other_post))
+white_vs_other_post$contrast <- "Other vs. White Post-ETI Slope"
+white_vs_other_post %>%
+  gt() %>%
+  fmt_number(decimals = 3)
 ```
@@
 change_mod <- lmer(
-  GLI_FEV1_pct_predicted ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  GLI_FEV1_pct_predicted ~ Years * Race + slope_change_years * Race +
+    age_group + Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter
@@
   gt() %>%
-  fmt_number(decimals = 2)
+  fmt_number(decimals = 3)
+# Linear hypotheses
+white_vs_black_pre <- glht(
+  change_mod,
+  "Years + `Years:RaceBlack or African American` = 1.195465"
+)
+white_vs_black_pre <- tidy(summary(white_vs_black_pre))
+white_vs_black_pre$contrast <- "Black vs. White Pre-ETI Slope"
+white_vs_black_pre %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+white_vs_black_post <- glht(
+  change_mod,
+  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.738885"
+)
+white_vs_black_post <- tidy(summary(white_vs_black_post))
+white_vs_black_post$contrast <- "Black vs. White Post-ETI Slope"
+white_vs_black_post %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+
+white_vs_hispanic_pre <- glht(
+  change_mod,
+  "Years + `Years:RaceHispanic or Latino` = 1.19547"
+)
+white_vs_hispanic_pre <- tidy(summary(white_vs_hispanic_pre))
+white_vs_hispanic_pre$contrast <- "Hispanic vs. White Pre-ETI Slope"
+white_vs_hispanic_pre %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+white_vs_hispanic_post <- glht(
+  change_mod,
+  "Years + `Years:RaceHispanic or Latino` + slope_change_years + `RaceHispanic or Latino:slope_change_years` = 2.738885"
+)
+white_vs_hispanic_post <- tidy(summary(white_vs_hispanic_post))
+white_vs_hispanic_post$contrast <- "Hispanic vs. White Post-ETI Slope"
+white_vs_hispanic_post %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+
+white_vs_other_pre <- glht(
+  change_mod,
+  "Years + `Years:RaceOther` = 1.19547"
+)
+white_vs_other_pre <- tidy(summary(white_vs_other_pre))
+white_vs_other_pre$contrast <- "Other vs. White Pre-ETI Slope"
+white_vs_other_pre %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+white_vs_other_post <- glht(
+  change_mod,
+  "Years + `Years:RaceOther` + slope_change_years + `RaceOther:slope_change_years` = 2.738885"
+)
+white_vs_other_post <- tidy(summary(white_vs_other_post))
+white_vs_other_post$contrast <- "Other vs. White Post-ETI Slope"
+white_vs_other_post %>%
+  gt() %>%
+  fmt_number(decimals = 3)
 ```
@@
   gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race +
-    Insurance + (1 | eDWID),
+    Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
@@
 change_mod <- lmer(
-  GLI_FEV1_pct_predicted ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
+  GLI_FEV1_pct_predicted ~ Years * Race + slope_change_years * Race +
+    Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
@@
 change_mod <- lmer(
-  gli_fev1_ppred_rn ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
+  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race +
+    Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[18, Inf)", ]
@@
 change_mod <- lmer(
-  GLI_FEV1_pct_predicted ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
+  GLI_FEV1_pct_predicted ~ Years * Race + slope_change_years * Race +
+    Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[18, Inf)", ]
@@
 change_mod <- lmer(
-  gli_fvc_ppred_rn ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  gli_fvc_ppred_rn ~ Years * Race + slope_change_years * Race + age_group +
+    Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter
@@
 change_mod <- lmer(
-  GLI_FVC_pct_predicted ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  GLI_FVC_pct_predicted ~ Years * Race + slope_change_years * Race +
+    age_group + Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter
@@
   gli_fvc_ppred_rn ~ Years * Race + slope_change_years * Race +
-    Insurance + (1 | eDWID),
+    Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
@@
 change_mod <- lmer(
-  GLI_FVC_pct_predicted ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
+  GLI_FVC_pct_predicted ~ Years * Race + slope_change_years * Race +
+    Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
@@
 change_mod <- lmer(
-  gli_fvc_ppred_rn ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
+  gli_fvc_ppred_rn ~ Years * Race + slope_change_years * Race +
+    Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[18, Inf)", ]
@@
 change_mod <- lmer(
-  GLI_FVC_pct_predicted ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
+  GLI_FVC_pct_predicted ~ Years * Race + slope_change_years * Race +
+    Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[18, Inf)", ]
@@
 change_mod <- lmer(
-  gli_fev1fvc_ppred_rn ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  gli_fev1fvc_ppred_rn ~ Years * Race + slope_change_years * Race +
+    age_group + Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter
@@
 change_mod <- lmer(
-  GLI_FEV1FVC_pct_predicted ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  GLI_FEV1FVC_pct_predicted ~ Years * Race + slope_change_years * Race +
+    age_group + Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter
@@
   gli_fev1fvc_ppred_rn ~ Years * Race + slope_change_years * Race +
-    Insurance + (1 | eDWID),
+    Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
@@
 change_mod <- lmer(
-  GLI_FEV1FVC_pct_predicted ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
+  GLI_FEV1FVC_pct_predicted ~ Years * Race + slope_change_years * Race +
+    Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
@@
 change_mod <- lmer(
-  gli_fev1fvc_ppred_rn ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
+  gli_fev1fvc_ppred_rn ~ Years * Race + slope_change_years * Race +
+    Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[18, Inf)", ]
@@
 change_mod <- lmer(
-  GLI_FEV1FVC_pct_predicted ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
+  GLI_FEV1FVC_pct_predicted ~ Years * Race + slope_change_years * Race +
+    Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[18, Inf)", ]
@@
 change_mod <- lmer(
-  bmipercentile ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  bmipercentile ~ Years * Race + slope_change_years * Race + age_group +
+    Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter[encounter$age_eti_start < 20, ]
@@
 change_mod <- lmer(
-  bmivalue ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  bmivalue ~ Years * Race + slope_change_years * Race + age_group +
+    Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter[encounter$age_eti_start >= 20, ]
@@
 change_mod <- lmer(
-  heightpercentile ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  heightpercentile ~ Years * Race + slope_change_years * Race + age_group +
+    Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter[encounter$age_eti_start < 20, ]
@@
 change_mod <- lmer(
-  height ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  height ~ Years * Race + slope_change_years * Race + age_group +
+    Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter[encounter$age_eti_start >= 20, ]
@@
 change_mod <- lmer(
-  weightpercentile ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  weightpercentile ~ Years * Race + slope_change_years * Race + age_group +
+    Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter[encounter$age_eti_start < 20, ]
@@
 change_mod <- lmer(
-  weight ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  weight ~ Years * Race + slope_change_years * Race + age_group +
+    Insurance + prior_modulator + isOnEnzymes + (1 | eDWID),
   data = encounter[encounter$age_eti_start >= 20, ]
@@
   gt() %>%
-  fmt_number(decimals = 2)
-```
+  fmt_number(decimals = 3)
+# Linear hypotheses
+white_vs_black_pre <- glht(
+  change_mod,
+  "Years + `Years:RaceBlack or African American` = 1.195465"
+)
+white_vs_black_pre <- tidy(summary(white_vs_black_pre))
+white_vs_black_pre$contrast <- "Black vs. White Pre-ETI Slope"
+white_vs_black_pre %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+white_vs_black_post <- glht(
+  change_mod,
+  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.738885"
+)
+white_vs_black_post <- tidy(summary(white_vs_black_post))
+white_vs_black_post$contrast <- "Black vs. White Post-ETI Slope"
+white_vs_black_post %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+
+white_vs_hispanic_pre <- glht(
+  change_mod,
+  "Years + `Years:RaceHispanic or Latino` = 1.19547"
+)
+white_vs_hispanic_pre <- tidy(summary(white_vs_hispanic_pre))
+white_vs_hispanic_pre$contrast <- "Hispanic vs. White Pre-ETI Slope"
+white_vs_hispanic_pre %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+white_vs_hispanic_post <- glht(
+  change_mod,
+  "Years + `Years:RaceHispanic or Latino` + slope_change_years + `RaceHispanic or Latino:slope_change_years` = 2.738885"
+)
+white_vs_hispanic_post <- tidy(summary(white_vs_hispanic_post))
+white_vs_hispanic_post$contrast <- "Hispanic vs. White Post-ETI Slope"
+white_vs_hispanic_post %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+
+white_vs_other_pre <- glht(
+  change_mod,
+  "Years + `Years:RaceOther` = 1.19547"
+)
+white_vs_other_pre <- tidy(summary(white_vs_other_pre))
+white_vs_other_pre$contrast <- "Other vs. White Pre-ETI Slope"
+white_vs_other_pre %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+white_vs_other_post <- glht(
+  change_mod,
+  "Years + `Years:RaceOther` + slope_change_years + `RaceOther:slope_change_years` = 2.738885"
+)
+white_vs_other_post <- tidy(summary(white_vs_other_post))
+white_vs_other_post$contrast <- "Other vs. White Post-ETI Slope"
+white_vs_other_post %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+```

### 2024-12-07 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
adjust for prior modulator, pseudomonas, etc.

--- 2024-11-15 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
@@
 label(encounter$slope_change_years) <- "Change in Slope"
+# Prior modulator use
+prior_mods <- encounter %>%
+  select(eDWID, Days, vx661comb, vx445comb, Vx770, Vx809comb) %>%
+  filter(Days < 0) %>%
+  group_by(eDWID) %>%
+  summarise(prior_modulator = any(vx661comb == 1) | any(vx445comb == 1) |
+    any(Vx770 == 1) | any(Vx809comb))
+prior_mods$prior_modulator <-
+  fct_na_value_to_level(factor(prior_mods$prior_modulator), level = "No")
+prior_mods$prior_modulator <- factor(prior_mods$prior_modulator,
+  levels = c(T, "No"),
+  labels = c("Yes", "No")
+)
+encounter <- left_join(encounter, prior_mods, by = join_by(eDWID))
+# On enzymes for pancreatic insufficiency
+encounter$isOnEnzymes <-
+  factor(encounter$isOnEnzymes, levels = 0:1, labels = c("No", "Yes"))
+# Pseudomonas
+encounter$pseudomonasaeruginosa[is.na(encounter$pseudomonasaeruginosa)] <- 0
+encounter$pseudomonasaeruginosa <- factor(encounter$pseudomonasaeruginosa,
+  levels = 0:1, labels = c("No", "Yes")
+)
 ```
@@
 change_mod <- lmer(
-  gli_fev1_ppred_rn ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + age_group +
+    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    (1 | eDWID),
   data = encounter
@@
   Years = -3:2, Race = unique(encounter$Race), age_group = "[18,30)",
-  Insurance = "Private"
+  Insurance = "Private", prior_modulator = "Yes",
+  isOnEnzymes = "Yes", pseudomonasaeruginosa = "No"
 ))
@@
 change_mod <- lmer(
-  GLI_FEV1_pct_predicted ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  GLI_FEV1_pct_predicted ~ Years * Race + slope_change_years * Race +
+    age_group + Insurance + prior_modulator + isOnEnzymes +
+    pseudomonasaeruginosa + (1 | eDWID),
   data = encounter
@@
 change_mod <- lmer(
-  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race +
-    Insurance + (1 | eDWID),
+  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
+    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
@@
 change_mod <- lmer(
-  GLI_FEV1_pct_predicted ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
+  GLI_FEV1_pct_predicted ~ Years * Race + slope_change_years * Race +
+    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
@@
 change_mod <- lmer(
-  gli_fev1_ppred_rn ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
+  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
+    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[18, Inf)", ]
@@
 change_mod <- lmer(
-  GLI_FEV1_pct_predicted ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
+  GLI_FEV1_pct_predicted ~ Years * Race + slope_change_years * Race +
+    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[18, Inf)", ]
@@
 change_mod <- lmer(
-  gli_fvc_ppred_rn ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  gli_fvc_ppred_rn ~ Years * Race + slope_change_years * Race + age_group +
+    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    (1 | eDWID),
   data = encounter
@@
 change_mod <- lmer(
-  GLI_FVC_pct_predicted ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  GLI_FVC_pct_predicted ~ Years * Race + slope_change_years * Race +
+    age_group + Insurance + prior_modulator + isOnEnzymes +
+    pseudomonasaeruginosa + (1 | eDWID),
   data = encounter
@@
 change_mod <- lmer(
-  gli_fvc_ppred_rn ~ Years * Race + slope_change_years * Race +
-    Insurance + (1 | eDWID),
+  gli_fvc_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
+    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
@@
 change_mod <- lmer(
-  GLI_FVC_pct_predicted ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
+  GLI_FVC_pct_predicted ~ Years * Race + slope_change_years * Race +
+    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
@@
 change_mod <- lmer(
-  gli_fvc_ppred_rn ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
+  gli_fvc_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
+    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[18, Inf)", ]
@@
 change_mod <- lmer(
-  GLI_FVC_pct_predicted ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
+  GLI_FVC_pct_predicted ~ Years * Race + slope_change_years * Race +
+    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[18, Inf)", ]
@@
 change_mod <- lmer(
-  gli_fev1fvc_ppred_rn ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  gli_fev1fvc_ppred_rn ~ Years * Race + slope_change_years * Race + age_group +
+    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    (1 | eDWID),
   data = encounter
@@
 change_mod <- lmer(
-  GLI_FEV1FVC_pct_predicted ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  GLI_FEV1FVC_pct_predicted ~ Years * Race + slope_change_years * Race +
+    age_group + Insurance + prior_modulator + isOnEnzymes +
+    pseudomonasaeruginosa + (1 | eDWID),
   data = encounter
@@
   gli_fev1fvc_ppred_rn ~ Years * Race + slope_change_years * Race +
-    Insurance + (1 | eDWID),
+    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
@@
 change_mod <- lmer(
-  GLI_FEV1FVC_pct_predicted ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
+  GLI_FEV1FVC_pct_predicted ~ Years * Race + slope_change_years * Race +
+    Insurance + prior_modulator + isOnEnzymes +
+    pseudomonasaeruginosa + (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
@@
 change_mod <- lmer(
-  gli_fev1fvc_ppred_rn ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
+  gli_fev1fvc_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
+    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[18, Inf)", ]
@@
 change_mod <- lmer(
-  GLI_FEV1FVC_pct_predicted ~ Years * Race +
-    slope_change_years * Race + Insurance + (1 | eDWID),
+  GLI_FEV1FVC_pct_predicted ~ Years * Race + slope_change_years * Race +
+    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    (1 | eDWID),
   data = encounter[encounter$age_eti_group == "[18, Inf)", ]
@@
 change_mod <- lmer(
-  bmipercentile ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  bmipercentile ~ Years * Race + slope_change_years * Race + age_group +
+    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    (1 | eDWID),
   data = encounter[encounter$age_eti_start < 20, ]
@@
 change_mod <- lmer(
-  bmivalue ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  bmivalue ~ Years * Race + slope_change_years * Race + age_group +
+    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    (1 | eDWID),
   data = encounter[encounter$age_eti_start >= 20, ]
@@
 change_mod <- lmer(
-  heightpercentile ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  heightpercentile ~ Years * Race + slope_change_years * Race + age_group +
+    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    (1 | eDWID),
   data = encounter[encounter$age_eti_start < 20, ]
@@
 change_mod <- lmer(
-  height ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  height ~ Years * Race + slope_change_years * Race + age_group + Insurance +
+    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
   data = encounter[encounter$age_eti_start >= 20, ]
@@
 change_mod <- lmer(
-  weightpercentile ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  weightpercentile ~ Years * Race + slope_change_years * Race + age_group +
+    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    (1 | eDWID),
   data = encounter[encounter$age_eti_start < 20, ]
@@
 change_mod <- lmer(
-  weight ~ Years * Race +
-    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
+  weight ~ Years * Race + slope_change_years * Race + age_group + Insurance +
+    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
   data = encounter[encounter$age_eti_start >= 20, ]

### 2024-12-07 Lilah Melzer/cough_monitor_analysis.qmd
file paths

@@
 # Import use data
-start_end <- read.csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring/Data_Clean/Start and end dates.csv",
+start_end <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Clean/Start and end dates.csv",
   na.strings = ""
@@
 start_end$End <- mdy(start_end$End)
-vacations <- read.csv("C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring/Data_Clean/Vacation days.csv",
+vacations <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Clean/Vacation days.csv",
   na.strings = ""
@@
 agg_counts <- read.csv(
-  "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring/Data_Raw/Cough_w_TotalCounts_0824 - Cough_agg_w_count_0828.csv",
+  "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Raw/Cough_w_TotalCounts_0824 - Cough_agg_w_count_0828.csv",
   na.strings = c("", "NMD"), header = F
@@
 # Save
-write.csv(agg_counts, file = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Lilah Melzer/Home Nighttime Cough Monitoring/Data_Clean/cleaned_aggregated_counts.csv", na = "", row.names = F)
+write.csv(agg_counts, file = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Clean/cleaned_aggregated_counts.csv", na = "", row.names = F)
 # Per night

### 2024-12-10 Quarto Template.qmd
fixed template

--- 2024-09-18 Quarto Template.qmd
@@
 library(redcapAPI)
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/tim/OneDrive/Vigers/CF"
-)
 github_dir <- switch(Sys.info()["sysname"],
   "Darwin" = "/Users/timvigers/GitHub",
-  "Windows" = "C:/Users/Tim/GitHub",
-  "Linux" = "/home/tim/GitHub"
+  "Windows" = "C:/Users/timvigers/GitHub",
+  "Linux" = "/home/timvigers/GitHub"
 )
-knitr::opts_knit$set(root.dir = home_dir)
+knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF")
 ```

### 2024-12-12 Lilah Melzer/cough_monitor_analysis.qmd
cough analysis new data

@@
 # pull
-pull_date <- mdy("8/28/24")
+pull_date <- mdy("11/25/24")
 start_end$End[is.na(start_end$End)] <- pull_date
@@
 agg_counts <- read.csv(
-  "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Raw/Cough_w_TotalCounts_0824 - Cough_agg_w_count_0828.csv",
+  "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Raw/Cough_w_TotalCounts_1105 - Cough_agg_w_count_nov25.csv",
   na.strings = c("", "NMD"), header = F
@@
 agg_counts$`Date/Device` <- NULL
+agg_counts$`#nights-monitoring` = NULL
 agg_counts <- agg_counts[-1, ]

### 2024-12-13 Christine Chan/CF Registry/cfrd_1hr_ogtt_cgm_predictive.qmd
new power calculations for Christine Chan

---
title: "Redefining CFRD Screening and Diagnosis in the Highly Effective Modulator Therapy Era (CF- SuGAR)"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(rspiro)
library(childsds)
library(tidyverse)
library(tableone)
library(pROC)
library(survival)
library(performance)
knitr::opts_knit$set(root.dir = "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo")
```

```{r}
#| include: false
# read in separate data files
ogtt <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/ogtt.csv")
pft <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/pfts.csv")
a1c <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/a1c.csv")
demo <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/demographics.csv")
dob <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/dob.csv")
# merge dob and pfts
pft <- merge(pft, dob, by = "MRN", all.x = T, all.y = F)
pft <- unique(pft)
# checking for dups
# t <- ogtt[,c("MRN","FName","LName")]
# t <- unique(t)
# ct <- t %>% group_by(LName,FName) %>% count(unique(MRN))
# merge ogtt and demo
demo$FName <- NULL
demo$LName <- NULL
ogtt <- ogtt %>%
  group_by(MRN, Date.of.OGTT) %>%
  filter(row_number() == 1)
ogtt_demo <- merge(ogtt, demo, by = "MRN", all.x = T, all.y = T)
ogtt_demo$Date.of.OGTT <- mdy(ogtt_demo$Date.of.OGTT)
a1c$EncounterDate <- as.Date(a1c$Date.of.A1C, format = "%m/%d/%Y")
a1c$FName <- NULL
a1c$LName <- NULL
a1c$Date.of.OGTT <- NULL
a1c$Days.Between.A1C.and.OGTT <- NULL
ogtt_demo$EncounterDate <- ogtt_demo$Date.of.OGTT
ogtt_demo_a1c <- merge(ogtt_demo, a1c, by = c("MRN", "EncounterDate"), all.x = T, all.y = T)

# find most recent FVC that is within the 90 days prior to OGTT
pft$EncounterDate <- as.Date(pft$Date.of.PFT, format = "%m/%d/%Y")
pft$FName <- NULL
pft$LName <- NULL
alldata <- merge(ogtt_demo_a1c, pft, by = c("MRN", "EncounterDate"), all.x = T, all.y = T)
# get predicted spirometry values
alldata$FVC...Pred <- NULL
alldata$FEV1...Pred <- NULL
alldata$FEF.25.75...Pred <- NULL
# dob <- pft[,c("MRN","DOB")]
# dob <- unique(dob)
# alldata <- merge(alldata,dob,by="MRN",all.x = T,all.y = F)
alldata$DOB <- as.Date(as.character(alldata$DOB), format = "%m/%d/%Y")
alldata$ht_m <- alldata$Height..cm. / 100
alldata$sexnum <- ifelse(alldata$Sex == "Male", 1, 2)
alldata$FEV1.pred <- pred_GLI(age = alldata$Age.at.PFT, height = alldata$ht_m, gender = alldata$sexnum, param = "FEV1")
alldata$FVC.pred <- pred_GLI(age = alldata$Age.at.PFT, height = alldata$ht_m, gender = alldata$sexnum, param = "FVC")
alldata$FEF.25.75.pred <- pred_GLI(age = alldata$Age.at.PFT, height = alldata$ht_m, gender = alldata$sexnum, param = "FEF2575")
alldata$FEV1.per.pred <- (alldata$FEV1.Raw / alldata$FEV1.pred) * 100
alldata$FVC.per.pred <- (alldata$FVC.Raw / alldata$FVC.pred) * 100
alldata$FEF.25.75.per.pred <- (alldata$FEF.25.75.Raw / alldata$FEF.25.75.pred) * 100
# alldata$Sex <- NULL
# hard code some corrections to data per CC
alldata[alldata$MRN == 686476, ]$X1.hour.Glucose <- ifelse(alldata[alldata$MRN == 686476, ]$Age.At.Test..years. == 14.5, 185, alldata[alldata$MRN == 686476, ]$X1.hour.Glucose)
# severity of genotype
# Red = severe or minimal function mutation
# Orange= unknown mutation status (for now)
# Yellow= partial function or mild
# Green = not CF causing, classify as "other", will most likely be seen in people with CRMS diagnosis
alldata$class2 <- ifelse(alldata$Mutation.2 %in% c(
  "1717-1G->A", "1717-1G-A", "G542X", "Q493X", "R1162X", "R553X", "R553x", "W1089X", "W1282X", "1078delT", "R75X",
  "3659delC", "621+1G->T", "621+1G>T", "394delTT", "3120 + 1g->A", "1154InsTC", "1154insTC", "1213delT",
  "1259insA", "1288insTA", "3791delC", "E60X", "K710X", "2184delA", "CFTRdele2,3", "663delT", "Glu528",
  "1461ins4", "306insA", "R709X", "CFTRdele22-24", "2711delT", "2183AA- >G", "2183delAA>G"
), "I",
ifelse(alldata$Mutation.2 %in% c("F508", "F508del", "F508del ", "G85E", "I507", "I 507", "N1303K", "A559T", "R560T", "A561E"),
  "II",
  ifelse(alldata$Mutation.2 %in% c("G551D", "G551S", "S549N"), "III",
    ifelse(alldata$Mutation.2 %in% c("R334W", "R347P", "R117H", "R117C", "P67L", "L206W", "l206w", "I206w", "D614G", "R347H", "D1152H", "3849+10kbC>T"), "IV",
      ifelse(alldata$Mutation.2 %in% c(
        "A455E", "2789+5G->A", "3849+10CT", "3849+10kbC->T", "711+3A->G", "1898+5G->T",
        "P574H", "3272-26A->G5", "3272-26A->G"
      ), "V", "")
    )
  )
)
)
alldata$severity2 <- ifelse(alldata$Mutation.2 %in% c(
  "1717-1G->A", "1717-1G-A", "G542X", "Q493X", "R1162X", "R553X", "R553x", "W1089X", "W1282X", "1078delT", "R75X",
  "3659delC", "621+1G->T", "621+1G>T", "394delTT", "3120 + 1g->A", "1154InsTC", "1154insTC", "1213delT",
  "1259insA", "1288insTA", "3791delC", "E60X", "K710X", "2184delA", "CFTRdele2,3", "663delT", "Glu528",
  "1461ins4", "306insA", "R709X", "CFTRdele22-24", "2711delT", "2183AA- >G", "2183delAA>G", "F508", "F508del", "F508del ", "G85E",
  "I507", "I 507", "N1303K", "A559T", "R560T", "A561E", "G551D", "G551S", "S549N", "1898+1G->A", "2143delT", "2183AA-G",
  "2183delAA->G", "2184insA", "S434X", "2585delT", "3120+1G->A", "3905insT", "406-1G->A", "712-1G->T", "L1245X", "Q2X", "R1158X",
  "R851X", "V520F", "L1077P", "S489X", "F311del", "W846X", "M1101K", "2622+1G->A", "S945L", "Y1092X", "R1066C", "G178R", "2957delT",
  "CFTRdele1", "W1204X", "q493x", "W57G", "1585-8G>A", "1717-8G->A", "G1244E"
), "Severe",
ifelse(alldata$Mutation.2 %in% c(
  "R334W", "R347P", "R117H", "R117C", "P67L", "L206W", "l206w", "I206w", "D614G", "R347H", "D1152H", "3849+10kbC>T", "D110H", "R1070W",
  "A455E", "2789+5G->A", "3849+10CT", "3849+10kbC->T", "711+3A->G", "1898+5G->T", "P574H", "3272-26A->G5", "3272-26A->G", "3849G->A", "H199Y", "H199y",
  "S492F", "S531P", "L1335P", "R352Q", "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val",
  "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"
), "Mild",
ifelse(alldata$Mutation.2 %in% c(
  "1262insA", "I506T", "3363delGT", "Q1209P", "3349insT", "G178R", "L467P", "317insC", "I336K",
  "ex14a", "3500-2A->G", "c.3407_3422del16", "3349insT", "1248+1G->A", "M470V", "1812-1G->A",
  "del X22-24", "Q237H", "Q237H", "W57G", "1078delA", "G178R", "2957delT", "G576A",
  "other", "Other", "Unknown", "unknown", "CFTRdele1", "1811+1.6kb A->", "UNK", "I336K", "M470V", "S1235R", "M470V", "F508C",
  "R1066H", "P205S"
), "Unk", "")
)
)
alldata$class1 <- ifelse(alldata$Mutation.1 %in% c(
  "1717-1G->A", "1717-1G-A", "G542X", "Q493X", "R1162X", "R553X", "R553x", "W1089X", "W1282X", "1078delT", "R75X",
  "3659delC", "621+1G->T", "621+1G>T", "394delTT", "3120 + 1g->A", "1154InsTC", "1154insTC", "1213delT",
  "1259insA", "1288insTA", "3791delC", "E60X", "K710X", "2184delA", "CFTRdele2,3", "663delT", "Glu528",
  "1461ins4", "306insA", "R709X", "CFTRdele22-24", "2711delT", "2183AA- >G", "2183delAA>G"
), "I",
ifelse(alldata$Mutation.1 %in% c("F508", "F508del", "F508del ", "G85E", "I507", "I 507", "N1303K", "A559T", "R560T", "A561E"), "II",
  ifelse(alldata$Mutation.1 %in% c("G551D", "G551S", "S549N"), "III",
    ifelse(alldata$Mutation.1 %in% c("R334W", "R347P", "R117H", "R117C", "P67L", "L206W", "l206w", "I206w", "D614G", "R347H", "D1152H", "3849+10kbC>T"), "IV",
      ifelse(alldata$Mutation.1 %in% c(
        "A455E", "2789+5G->A", "3849+10CT", "3849+10kbC->T", "711+3A->G", "1898+5G->T",
        "P574H", "3272-26A->G5", "3272-26A->G"
      ), "V", "")
    )
  )
)
)
alldata$severity1 <- ifelse(alldata$Mutation.1 %in% c(
  "1717-1G->A", "1717-1G-A", "G542X", "Q493X", "R1162X", "R553X", "R553x", "W1089X", "W1282X", "1078delT", "R75X",
  "3659delC", "621+1G->T", "621+1G>T", "394delTT", "3120 + 1g->A", "1154InsTC", "1154insTC", "1213delT",
  "1259insA", "1288insTA", "3791delC", "E60X", "K710X", "2184delA", "CFTRdele2,3", "663delT", "Glu528",
  "1461ins4", "306insA", "R709X", "CFTRdele22-24", "2711delT", "2183AA- >G", "2183delAA>G", "F508", "F508del", "F508del ", "G85E",
  "I507", "I 507", "N1303K", "A559T", "R560T", "A561E", "G551D", "G551S", "S549N", "1898+1G->A", "2143delT", "2183AA-G",
  "2183delAA->G", "2184insA", "S434X", "2585delT", "3120+1G->A", "3905insT", "406-1G->A", "712-1G->T", "L1245X", "Q2X", "R1158X",
  "R851X", "V520F", "L1077P", "S489X", "F311del", "W846X", "M1101K", "2622+1G->A", "S945L",
  "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", "W1204X", "q493x", "W57G", "1585-8G>A", "1717-8G->A", "G1244E"
), "Severe",
ifelse(alldata$Mutation.1 %in% c(
  "R334W", "R347P", "R117H", "R117C", "P67L", "L206W", "l206w", "I206w", "D614G", "R347H", "D1152H", "3849+10kbC>T", "D110H", "R1070W",
  "A455E", "2789+5G->A", "3849+10CT", "3849+10kbC->T", "711+3A->G", "1898+5G->T",
  "P574H", "3272-26A->G5", "3272-26A->G", "3849G->A", "H199Y", "H199y", "S492F", "S531P",
  "L1335P", "R352Q", "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val",
  "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"
), "Mild",
ifelse(alldata$Mutation.1 %in% c(
  "1262insA", "I506T", "3363delGT", "Q1209P", "3349insT", "G178R", "L467P", "317insC", "I336K",
  "ex14a", "3500-2A->G", "c.3407_3422del16", "3349insT", "1248+1G->A", "M470V", "1812-1G->A",
  "del X22-24", "Q237H", "Q237H", "W57G", "1078delA", "G178R", "2957delT", "G576A",
  "other", "Other", "Unknown", "unknown", "CFTRdele1", "1811+1.6kb A->", "UNK", "I336K", "M470V", "S1235R", "M470V", "F508C",
  "R1066H", "P205S"
), "Unk", "")
)
)
# 3 patients have severity2="Severe" per Christine
alldata$severity2 <- ifelse(alldata$MRN %in% c(546678, 664400, 2015506), "Severe", alldata$severity2)
alldata$genoRisk <- ifelse(alldata$severity1 == "Mild" | alldata$severity2 == "Mild", "Low",
  ifelse(alldata$severity1 == "Severe" & alldata$severity2 == "Severe", "High", "")
)
# write file of people with missing genotype risk
# write.csv(alldata[alldata$genoRisk=="",],"./Christine Chan/1 hour OGTT in CFRD/Data_raw/missing_genotype_risk.csv")
# merge back all baseline characteristics
d <- alldata[, c("MRN", "FName", "LName", "Sex", "Race", "Latino.a.", "Mutation.1", "Mutation.2", "class1", "class2", "severity1", "severity2", "genoRisk")]
d <- d[!is.na(d$FName), ]
d <- unique(d)
alldata$FName <- NULL
alldata$Latino.a. <- NULL
alldata$LName <- NULL
alldata$Sex <- NULL
alldata$Race <- NULL
alldata$Mutation.1 <- NULL
alldata$Mutation.2 <- NULL
alldata$class1 <- NULL
alldata$class2 <- NULL
alldata$severity1 <- NULL
alldata$severity2 <- NULL
alldata$genoRisk <- NULL
alldata <- merge(alldata, d, by = "MRN", all.x = T, all.y = T)
# checking for dups
# t <- alldata[,c("MRN","FName","LName")]
# t <- unique(t)
# ct <- t %>% group_by(LName,FName) %>% count(unique(MRN))
# find first OGTT
alldata$Date <- alldata$EncounterDate
o <- alldata[!is.na(alldata$X1.hour.Glucose), c("MRN", "Date")]
o <- o[order(o$MRN, o$Date), ]
o <- o %>%
  group_by(MRN) %>%
  filter(row_number() == 1)
colnames(o) <- c("MRN", "Date of first OGTT")
alldata <- merge(alldata, o, by = "MRN", all.x = T, all.y = T)
alldata <- alldata[!is.na(alldata$`Date of first OGTT`), ]
# checking ppts that Collin said had CFRD
col_check <- alldata[alldata$MRN %in% c(
  1006479, 1053930, 1078408, 1136992, 1289450, 1339023, 1413821, 1695512, 664400,
  686476, 707831, 712676, 739231, 739460, 787372, 800007, 848440, 860181, 875128, 897478, 905284, 908966,
  918124, 932695, 949042, 952250, 952252, 825118
), ]
col_check <- col_check[order(col_check$MRN, col_check$Date), ]
# write.csv(col_check,"./Christine Chan/1 hour OGTT in CFRD/Data_raw/check_cfrd.csv",na="",row.names = F)
# last 2 discrepant cases after fixing date issue
col_check2 <- alldata[alldata$MRN %in% c(686476, 825118), ]
col_check2 <- col_check2[order(col_check2$MRN, col_check2$Date), ]
# write.csv(col_check2,"./Christine Chan/1 hour OGTT in CFRD/Data_raw/check_cfrd2.csv",na="",row.names = F)
# remove all visits before first OGTT
# alldata$`Date of first OGTT` <- as.Date(as.character(alldata$`Date of first OGTT`),format="%m/%d/%Y")
alldata <- alldata[alldata$Date >= alldata$`Date of first OGTT`, ]
alldata <- alldata %>% arrange(MRN, EncounterDate)
# calculate BMI
alldata$BMI <- alldata$Weight..kg. / ((alldata$Height..cm. / 100)^2)
# reorder columns
colorder <- c(
  "MRN", "FName", "LName", "Sex", "Race", "Latino.a.", "Mutation.1", "Mutation.2", "class1", "class2", "severity1", "severity2", "genoRisk",
  "Date of first OGTT", "Date", "Age.At.Test..years.", "Fasting.Glucose", "X1.hour.Glucose", "X2.hour.Glucose", "A1C",
  "DOB", "Age.at.PFT", "FVC.Raw", "Height..cm.", "Weight..kg.", "FEV1.Raw", "FEF.25.75.Raw", "ht_m",
  "sexnum", "FEV1.pred", "FVC.pred", "FEF.25.75.pred", "FEV1.per.pred", "FVC.per.pred", "FEF.25.75.per.pred", "BMI"
)
alldata <- alldata[, colorder]
alldata <- alldata[order(alldata$MRN, alldata$Date), ]
# create combined race/ethnicity
alldata$race_eth <- ifelse(alldata$Latino.a. == "TRUE", "Hispanic", alldata$Race)
# get BMI percentile
alldata$age <- ifelse(is.na(alldata$Age.at.PFT), alldata$Age.At.Test..years., alldata$Age.at.PFT)
alldata$age_bmi <- ifelse(alldata$age > 20, 20, alldata$age)
alldata$bmip <- sds(alldata$BMI,
  age = alldata$age_bmi,
  sex = alldata$Sex, male = "Male", female = "Female",
  ref = cdc.ref,
  item = "bmi",
  type = "perc"
)
# for demographics data, multiply BMI percentile by 100
alldata$bmipercentile <- alldata$bmip * 100
# classify CFRD at each visit
alldata$cfrd_visit <- NA
alldata$cfrd_visit <- ifelse(is.na(alldata$X2.hour.Glucose), NA, ifelse(alldata$X2.hour.Glucose >= 200, 1, 0))
# find people with CFRD at baseline
cfrdbase <- alldata[alldata$Date == alldata$`Date of first OGTT` & alldata$cfrd_visit == 1, ]
before_cfrd <- length(unique(alldata$MRN))
alldata <- alldata[!(alldata$MRN %in% cfrdbase$MRN), ]
after_cfrd <- length(unique(alldata$MRN))
# number excluded with CFRD at baseline
nexcl_cfrd <- before_cfrd - after_cfrd
# create variable for CFRD groups (i.e., those who developed CFRD vs those who did not)
# and get date of first OGTT with CFRD
cfrdcheck <- alldata[alldata$cfrd_visit == 1, c("MRN", "Date", "cfrd_visit")]
cfrdcheck <- cfrdcheck[order(cfrdcheck$MRN, desc(cfrdcheck$cfrd_visit)), ]
cfrdcheck <- cfrdcheck[!is.na(cfrdcheck$MRN), ]
cfrdcheck <- cfrdcheck %>%
  group_by(MRN) %>%
  filter(row_number() == 1)
colnames(cfrdcheck) <- c("MRN", "Date of CFRD diagnosis", "CFRD")
alldata <- merge(alldata, cfrdcheck, by = "MRN", all.x = T, all.y = T)
alldata$CFRD <- ifelse(is.na(alldata$CFRD), 0, 1)
# get baseline age
baseage <- alldata[alldata$Date == alldata$`Date of first OGTT`, c("MRN", "age")]
colnames(baseage) <- c("MRN", "Base_age")
alldata <- merge(alldata, baseage, by = "MRN", all.x = T, all.y = T)
# get baseline 1 hour and 2 hour glucose
basegluc <- alldata[alldata$Date == alldata$`Date of first OGTT`, c("MRN", "X1.hour.Glucose", "X2.hour.Glucose")]
colnames(basegluc) <- c("MRN", "Baseline_1hr_glucose", "Baseline_2hr_glucose")
alldata <- merge(alldata, basegluc, by = "MRN", all.x = T, all.y = T)
# dichotomize baseline 1 hour glucose
alldata$Baseline_1hr_glucose_155 <- ifelse(alldata$Baseline_1hr_glucose >= 155, 1, 0)
alldata$Baseline_1hr_glucose_140 <- ifelse(alldata$Baseline_1hr_glucose >= 140, 1, 0)
# dichotomize baseline 2 hour glucose
alldata$Baseline_2hr_glucose_140 <- ifelse(alldata$Baseline_2hr_glucose >= 140, 1, 0)
# find date of last OGTT
lastogtt <- alldata[
  !is.na(alldata$Fasting.Glucose) | !is.na(alldata$X1.hour.Glucose) | !is.na(alldata$X2.hour.Glucose),
  c("MRN", "Date")
]
lastogtt <- lastogtt[order(lastogtt$MRN, desc(lastogtt$Date)), ]
lastogtt <- lastogtt %>%
  group_by(MRN) %>%
  filter(row_number() == 1)
colnames(lastogtt) <- c("MRN", "Date of last OGTT")
alldata <- merge(alldata, lastogtt, by = "MRN", all.x = T, all.y = T)
# read in date of modulator use
mod <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/modulators.csv", na.strings = c("", " ", NA))
mod$Kalydeco.Start.Date <- as.Date(mod$Kalydeco.Start.Date, format = "%m/%d/%Y")
mod$Orkambi.Start.Date <- as.Date(mod$Orkambi.Start.Date, format = "%m/%d/%Y")
mod$Symdeko.Start.Date <- as.Date(mod$Symdeko.Start.Date, format = "%m/%d/%Y")
mod$Trikafta.Start.Date <- as.Date(mod$Trikafta.Start.Date, format = "%m/%d/%Y")
mod$modstart <- as.Date(NA)
mod$he_modstart <- as.Date(NA)
mod$me_modstart <- as.Date(NA)
for (i in 1:nrow(mod)) {
  mod[i, ]$modstart <-
    as.Date(pmin(mod[i, ]$Kalydeco.Start.Date, mod[i, ]$Orkambi.Start.Date, mod[i, ]$Symdeko.Start.Date, mod[i, ]$Trikafta.Start.Date, na.rm = T))
  mod[i, ]$he_modstart <- as.Date(pmin(mod[i, ]$Kalydeco.Start.Date, mod[i, ]$Trikafta.Start.Date, na.rm = T))
  mod[i, ]$me_modstart <- as.Date(pmin(mod[i, ]$Orkambi.Start.Date, mod[i, ]$Symdeko.Start.Date, na.rm = T))
}
modkeep <- mod[, c("MRN", "modstart", "he_modstart", "me_modstart")]
alldata <- merge(alldata, modkeep, by = "MRN", all.x = TRUE, all.y = FALSE)
alldata$onmod <- ifelse(!is.na(alldata$modstart) & alldata$modstart <= alldata$Date, 1,
  ifelse(!is.na(alldata$modstart) & alldata$modstart > alldata$Date, 0, NA)
)
alldata$onmod <- ifelse(is.na(alldata$modstart) & is.na(alldata$onmod), 0, alldata$onmod)
alldata$onhemod <- ifelse(!is.na(alldata$he_modstart) & alldata$he_modstart <= alldata$Date, 1,
  ifelse(!is.na(alldata$he_modstart) & alldata$he_modstart > alldata$Date, 0, NA)
)
alldata$onhemod <- ifelse(is.na(alldata$he_modstart) & is.na(alldata$onhemod), 0, alldata$onhemod)
alldata$onmemod <- ifelse(!is.na(alldata$me_modstart) & alldata$me_modstart <= alldata$Date, 1,
  ifelse(!is.na(alldata$me_modstart) & alldata$me_modstart > alldata$Date, 0, NA)
)
alldata$onmemod <- ifelse(is.na(alldata$me_modstart) & is.na(alldata$onmemod), 0, alldata$onmemod)
alldata$FEV.1.per.pred_FVC.per.pred <- alldata$FEV1.per.pred / alldata$FVC.per.pred
# for analyses about prediction of CFRD, we need to exclude those who only have one OGTT
# can use alldata for mixed models of PFTs and BMI
oneogtt <- alldata[alldata$`Date of first OGTT` == alldata$`Date of last OGTT`, ]
multogtt <- alldata[!(alldata$MRN %in% oneogtt$MRN), ]
# have CFRD indicator
# create variable for follow-up time
# in those who do not develop CFRD, this is the time between first and last OGTT, and they are censored
# for those who develop CFRD, this is time between first OGTT and diagnosis of CFRD
multogtt$Follow_up_time_yrs <- ifelse(multogtt$CFRD == 1, (multogtt$`Date of CFRD diagnosis` - multogtt$`Date of first OGTT`) / 365.25,
  (multogtt$`Date of last OGTT` - multogtt$`Date of first OGTT`) / 365.25
)
# get age at diagnosis of CFRD
multogtt$age_at_CFRD_dx <- ifelse(multogtt$CFRD == 1, (multogtt$`Date of CFRD diagnosis` - multogtt$DOB) / 365.25, NA)
# create categorical time
multogtt$cat_time <- as.factor(plyr::round_any(multogtt$Follow_up_time_yrs, 0.5))
# need a time variable for mixed models
multogtt$time <- as.numeric((multogtt$Date - multogtt$`Date of first OGTT`) / 365.25)
multogtt$cat_time2 <- as.factor(plyr::round_any(multogtt$time, 0.5))
# get number of participants in each cohort
n_alldata <- length(unique(alldata$MRN))
n_multogtt <- length(unique(multogtt$MRN))
# number of participants in each cohort who developed CFRD
n_alldata_cfrd <- length(unique(alldata[alldata$CFRD == 1, ]$MRN))
n_multogtt_cfrd <- length(unique(multogtt[multogtt$CFRD == 1, ]$MRN))
# create dataset with baseline only
baseline_mult <- multogtt[multogtt$Date == multogtt$`Date of first OGTT`, ]
baseline_mult$onmod_baseline <- as.factor(baseline_mult$onmod)
baseline_all <- alldata[alldata$Date == alldata$`Date of first OGTT`, ]
baseline_all$onmod_baseline <- as.factor(baseline_all$onmod)
# find patients who were put on modulators during follow-up
evermod <- alldata %>%
  group_by(MRN) %>%
  summarise(max(onmod))
baseline_all <- merge(baseline_all, evermod, by = "MRN", all.x = T, all.y = T)
baseline_all$modstarted_fup <- as.factor(ifelse(baseline_all$onmod_baseline == 0 & baseline_all$`max(onmod)` == 1, 1, 0))
baseline_mult <- merge(baseline_mult, evermod, by = "MRN", all.x = T, all.y = F)
baseline_mult$modstarted_fup <- as.factor(ifelse(baseline_mult$onmod_baseline == 0 & baseline_mult$`max(onmod)` == 1, 1, 0))
################## FILL IN MISSING A1c, BMI, PFT AT BASELINE
# find most recent A1c that is within the 90 days prior to OGTT or up to 14 days after OGTT
a1c$Date.of.A1C <- as.Date(a1c$Date.of.A1C, format = "%m/%d/%Y")
baseline_mult$A1C <- unlist(apply(baseline_mult, 1, function(r) {
  Date <- r["Date of first OGTT"]
  ID <- as.numeric(r["MRN"])
  d <- a1c %>% filter(MRN == ID)
  d$days <- difftime(d$Date.of.A1C, Date, units = "days")
  d <- d %>% filter(days <= 14 & days >= -90)
  a <- d$A1C[which.min(abs(difftime(d$Date.of.A1C, Date)))]
  if (is.na(Date) | nrow(d) == 0) {
    a <- NA
  }
  return(a)
}))
baseline_all$A1C <- unlist(apply(baseline_all, 1, function(r) {
  Date <- r["Date of first OGTT"]
  ID <- as.numeric(r["MRN"])
  d <- a1c %>% filter(MRN == ID)
  d$days <- difftime(d$Date.of.A1C, Date, units = "days")
  d <- d %>% filter(days <= 14 & days >= -90)
  a <- d$A1C[which.min(abs(difftime(d$Date.of.A1C, Date)))]
  if (is.na(Date) | nrow(d) == 0) {
    a <- NA
  }
  return(a)
}))
# find most recent FEV1 that is within the 90 days prior to OGTT or up to 14 days after OGTT
pfttemp <- alldata[!is.na(alldata$FEV1.per.pred), ]
pfttemp$Date.of.PFT <- as.Date(pfttemp$Date, format = "%m/%d/%Y")
baseline_mult$FEV1.per.pred <- unlist(apply(baseline_mult, 1, function(r) {
  Date <- r["Date of first OGTT"]
  ID <- as.numeric(r["MRN"])
  d <- pfttemp %>% filter(MRN == ID)
  d$days <- difftime(d$Date.of.PFT, Date, units = "days")
  d <- d %>% filter(days <= 14 & days >= -90)
  a <- d$FEV1.per.pred[which.min(abs(difftime(d$Date.of.PFT, Date)))]
  if (is.na(Date) | nrow(d) == 0) {
    a <- NA
  }
  return(a)
}))
baseline_all$FEV1.per.pred <- unlist(apply(baseline_all, 1, function(r) {
  Date <- r["Date of first OGTT"]
  ID <- as.numeric(r["MRN"])
  d <- pfttemp %>% filter(MRN == ID)
  d$days <- difftime(d$Date.of.PFT, Date, units = "days")
  d <- d %>% filter(days <= 14 & days >= -90)
  a <- d$FEV1.per.pred[which.min(abs(difftime(d$Date.of.PFT, Date)))]
  if (is.na(Date) | nrow(d) == 0) {
    a <- NA
  }
  return(a)
}))
# didn't add any PFTs so won't repeat code for FVC
# find most recent BMI that is within the 90 days prior to OGTT or up to 14 days after OGTT
bmitemp <- alldata[!is.na(alldata$BMI), ]
bmitemp$Date.of.BMI <- as.Date(bmitemp$Date, format = "%m/%d/%Y")
baseline_mult$BMI <- unlist(apply(baseline_mult, 1, function(r) {
  Date <- r["Date of first OGTT"]
  ID <- as.numeric(r["MRN"])
  d <- bmitemp %>% filter(MRN == ID)
  d$days <- difftime(d$Date.of.BMI, Date, units = "days")
  d <- d %>% filter(days <= 14 & days >= -90)
  a <- d$BMI[which.min(abs(difftime(d$Date.of.BMI, Date)))]
  if (is.na(Date) | nrow(d) == 0) {
    a <- NA
  }
  return(a)
}))
baseline_all$BMI <- unlist(apply(baseline_all, 1, function(r) {
  Date <- r["Date of first OGTT"]
  ID <- as.numeric(r["MRN"])
  d <- bmitemp %>% filter(MRN == ID)
  d$days <- difftime(d$Date.of.BMI, Date, units = "days")
  d <- d %>% filter(days <= 14 & days >= -90)
  a <- d$BMI[which.min(abs(difftime(d$Date.of.BMI, Date)))]
  if (is.na(Date) | nrow(d) == 0) {
    a <- NA
  }
  return(a)
}))
#############################################################
# create multogtt with time <4 years since there is little data beyond that
multogtt_reduced <- multogtt[multogtt$time <= 4, ]
##############
# DESC STATS #
##############
# table of demographics in those with at least one pre-CFRD OGTT
demovars <- c(
  "Sex", "race_eth", "age", "genoRisk", "Fasting.Glucose", "X1.hour.Glucose", "X2.hour.Glucose", "onmod_baseline", "modstarted_fup",
  "A1C", "FEV1.per.pred", "FVC.per.pred", "BMI", "bmipercentile", "FEV.1.per.pred_FVC.per.pred"
)
# Taking out these clinical variables - if they are to be added to table 1, need to figure out how
# because they come from PFT visits, not the same date as OGTT
#              "BMI","FEV1.per.pred","FVC.per.pred","FEF.25.75.per.pred","bmip")
t1_alldata <- CreateTableOne(data = baseline_all, vars = demovars)
t1_alldata <- print(t1_alldata, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T, test = F)
# table of demographics in those with at least one pre-CFRD OGTT by sex
t1_alldata_sex <- CreateTableOne(data = baseline_all, vars = demovars, strata = "Sex")
t1_alldata_sex <- print(t1_alldata_sex, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T)
# table of demographics in those with more than 1 OGTT
t1_multogtt <- CreateTableOne(data = baseline_mult, vars = c(demovars, "Follow_up_time_yrs"))
t1_multogtt <- print(t1_multogtt, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T, test = F)
# table of demographics in those with more than 1 OGTT stratified by progression
t1_multogtt_prog <- CreateTableOne(data = baseline_mult, vars = c(demovars, "Follow_up_time_yrs", "age_at_CFRD_dx"), strata = "CFRD")
t1_multogtt_prog <- print(t1_multogtt_prog, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T, test = T)
# table of demographics in those with more than 1 OGTT by sex
t1_multogtt_sex <- CreateTableOne(data = baseline_mult, vars = demovars, strata = "Sex")
t1_multogtt_sex <- print(t1_multogtt_sex, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T)
# compare baseline glucose values in those who developed CFRD and those who did not
# added other vars too
gluctable <- CreateTableOne(baseline_mult, vars = demovars, strata = "CFRD")
gluctable <- print(gluctable, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T, test = T)
########################
# LOGISTIC REGRESSIONS #
########################
# logistic regression of 1 hour and 2 hour glucose predicting cfrd
log1 <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose, family = "binomial")
or1 <- exp(cbind(OR = coef(log1), confint(log1)))
log1_adj <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose + Base_age + Sex, family = "binomial")
or1_adj <- exp(cbind(OR = coef(log1_adj), confint(log1_adj)))
# add interaction with sex
log1_adj_int <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose + Base_age + Sex + Sex * Baseline_1hr_glucose, family = "binomial")
or1_adj_int <- exp(cbind(OR = coef(log1_adj_int), confint(log1_adj_int)))
log1_cat <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose_155, family = "binomial")
or1_cat <- exp(cbind(OR = coef(log1_cat), confint(log1_cat)))
log1_cat_adj <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose_155 + Base_age + Sex, family = "binomial")
or1_cat_adj <- exp(cbind(OR = coef(log1_cat_adj), confint(log1_cat_adj)))
log1_cat_140 <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose_140, family = "binomial")
or1_cat_140 <- exp(cbind(OR = coef(log1_cat_140), confint(log1_cat_140)))
log1_cat_adj_140 <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose_140 + Base_age + Sex, family = "binomial")
or1_cat_adj_140 <- exp(cbind(OR = coef(log1_cat_adj_140), confint(log1_cat_adj_140)))
# add interaction with sex
log1_cat_adj_int <- glm(
  data = baseline_mult, CFRD ~ Baseline_1hr_glucose_155 + Base_age + Sex + Sex * Baseline_1hr_glucose_155,
  family = "binomial"
)
or1_cat_adj_int <- exp(cbind(OR = coef(log1_cat_adj_int), confint(log1_cat_adj_int)))
log2 <- glm(data = baseline_mult, CFRD ~ Baseline_2hr_glucose, family = "binomial")
or2 <- exp(cbind(OR = coef(log2), confint(log2)))
log2_adj <- glm(data = baseline_mult, CFRD ~ Baseline_2hr_glucose + Base_age + Sex, family = "binomial")
or2_adj <- exp(cbind(OR = coef(log2_adj), confint(log2_adj)))
# add interaction with sex
log2_adj_int <- glm(data = baseline_mult, CFRD ~ Baseline_2hr_glucose + Base_age + Sex + Sex * Baseline_2hr_glucose, family = "binomial")
or2_adj_int <- exp(cbind(OR = coef(log2_adj_int), confint(log2_adj_int)))
# ROC for 1 hour glucose
# get predicted values from logistic regression
baseline_mult$pred_1hr <- predict(log1, type = "response", data.frame = baseline_mult)
# perform the ROC analysis
roc1 <- roc(data = baseline_mult, response = CFRD, predictor = X1.hour.Glucose)
# find the cutoff that maximizes the Youden Index
coords1 <- coords(roc1, "b", ret = c("t", "specificity", "sensitivity"), best.method = "youden") # default
# ROC for 2 hour glucose
baseline_mult$pred_2hr <- predict(log2, type = "response", data.frame = baseline_mult)
roc2 <- roc(data = baseline_mult, response = CFRD, predictor = X2.hour.Glucose)
coords2 <- coords(roc2, "b", ret = c("t", "specificity", "sensitivity"), best.method = "youden") # default
##############
# COX MODELS #
##############
# CPH model for baseline 1 hour glucose
cox1hr <- coxph(Surv(Follow_up_time_yrs, CFRD) ~ Baseline_1hr_glucose, data = baseline_mult)
# Diagnostics
d1 <- cox.zph(cox1hr)
r2_coxsnell(cox1hr)
```

For the purposes of this power calculation, I took Laura Pyle's analysis code for the "Predictive Value of 1-Hour Glucose" paper @lorenzPredictiveValue1Hour2023 and re-fit the Cox proportional hazard model excluding pancreatic sufficient participants, in order to get an updated value for the apparent Cox-Snell $R^2$ ($R^2_{CSapp}$).

### 2024-12-18 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
new age groups SSwAB-IT

@@
 df <- full_join(db_cos_collection_results, db_sos_collection_results)
+df <- full_join(df,
+  db_selfobtained_sample[, c("record_id", "who_collected_the_sos")],
+  by = join_by(record_id)
+)
 # Add dates to demographics
@@
 )
-db_demographics$age_at_cos <- as.numeric(difftime(db_demographics$cos_collection_date,
+db_demographics$age_at_cos <-
+  as.numeric(difftime(db_demographics$cos_collection_date,
     db_demographics$dob,
@@
   )) / 365.25
+# 1-5, 6-11, 12-17 and 18+ years
 db_demographics$age_group <- cut(db_demographics$age_at_cos,
-  c(-Inf, 10, 18, 30, Inf),
+  c(-Inf, 6, 12, 18, Inf),
   right = F
@@
   select(
-    record_id,
+    record_id, who_collected_cos, who_collected_the_sos,
     staphylococcus_aureus_cos:other_result_cos,
@@
 db_feasibility <- db_feasibility %>% filter(!record_id %in% exclude)
+# Drop unused levels
+df$who_collected_cos <- droplevels(df$who_collected_cos)
 ```
@@
   add_overall() %>%
-  add_p(test = list(c("Race", "Ethnicity") ~ "fisher.test")) %>%
+  add_p(
+    test = list(c("Race", "Ethnicity") ~ "fisher.test"),
+    test.args = age_group ~ list(workspace = 2e9)
+  ) %>%
   separate_p_footnotes() %>%
@@
   cat("\n")
+  cat("\n")
   cat("####", "2x2 Table")
+  cat("\n")
   cat("\n")
@@
   cat("\n")
+  cat("\n")
   cat("####", "2x2 Table (With CIs)")
+  cat("\n")
   cat("\n")
@@
   cat("\n")
+  cat("\n")
   print(plot(euler(d == "Negative"),
@@
   cat("\n")
+  cat("\n")
   print(plot(euler(d == "Positive"),
@@
   ))
+  cat("\n")
   cat("\n")
@@
   cat("\n")
+  cat("\n")
   d %>%
@@
     add_overall() %>%
-    add_p(test.args = all_tests("fisher.test") ~ list(workspace = 2e6)) %>%
+    add_p(test.args = all_tests("fisher.test") ~ list(workspace = 2e9)) %>%
     separate_p_footnotes() %>%
     print()
+  cat("\n")
   cat("\n")
@@
   cat("\n")
+  cat("\n")
   cat("##", as.character(bug_list[b]))
+  cat("\n")
   cat("\n")
@@
   cat("\n")
+  cat("\n")
   agreement(b)
+  cat("\n")
   cat("\n")
@@
   cat("\n")
+  cat("\n")
   agreement_test(b)
+  cat("\n")
   cat("\n")
@@
   cat("\n")
+  cat("\n")
   cat("###", as.character(bug_list[b]))
+  cat("\n")
   cat("\n")
@@
   cat("\n")
+  cat("\n")
   agreement_test(b, group = "age_group")
+  cat("\n")
   cat("\n")
@@
   cat("\n")
+  cat("\n")
   agreement_test(b, group = "azithromycin")
+  cat("\n")
   cat("\n")
@@
   cat("\n")
+  cat("\n")
   agreement_test(b, group = "inhaled_tobramycin")
+  cat("\n")
   cat("\n")
@@
   cat("\n")
+  cat("\n")
   agreement_test(b, group = "inhaled_aztreonam")
+  cat("\n")
   cat("\n")
@@
   cat("\n")
+  cat("\n")
   agreement_test(b, group = "inhaled_colisitin")
+  cat("\n")
   cat("\n")
@@
   cat("\n")
+  cat("\n")
   agreement_test(b, group = "inhaled_vancomycin")
+  cat("\n")
   cat("\n")
@@
   cat("\n")
+  cat("\n")
   agreement_test(b, group = "other_inhaled_antibiotics")
+  cat("\n")
   cat("\n")
@@
   cat("\n")
+  cat("\n")
   agreement_test(b, group = "chronic_oral_antibiotic")
+  cat("\n")
   cat("\n")
@@
   cat("\n")
+  cat("\n")
   agreement_test(b, group = "cftr_modulator")
+  cat("\n")
   cat("\n")
@@
 ```
+
+## Who collected the COS
+
+```{r results='asis'}
+#| message: false
+invisible(lapply(names(bug_list), function(b) {
+  cat("\n")
+  cat("\n")
+  cat("###", as.character(bug_list[b]))
+  cat("\n")
+  cat("\n")
+  cat("####", "Agreement by who collected the COS")
+  cat("\n")
+  cat("\n")
+  agreement_test(b, group = "who_collected_cos")
+  cat("\n")
+  cat("\n")
+}))
+```
+
+## Who collected the SOS
+
+```{r results='asis'}
+#| message: false
+invisible(lapply(names(bug_list), function(b) {
+  cat("\n")
+  cat("\n")
+  cat("###", as.character(bug_list[b]))
+  cat("\n")
+  cat("\n")
+  cat("####", "Agreement by who collected the SOS")
+  cat("\n")
+  cat("\n")
+  agreement_test(b, group = "who_collected_the_sos")
+  cat("\n")
+  cat("\n")
+}))
+```

### 2025-01-02 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.qmd
minor formatting changes for Colleen

@@
 bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
-csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
+csl: /Users/timvigers/Dropbox/GitHub/styles/american-medical-association.csl
 editor: source
@@
 library(car)
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood"
-)
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
-knitr::opts_knit$set(root.dir = home_dir)
+knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood")
 ```
@@
 #| include: false
-load("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/analysis_dataset.rdata")
-load("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/hba1c_dataset.rdata")
+load("./Prepost Trikafta/Data_Cleaned/analysis_dataset.rdata")
+load("./Prepost Trikafta/Data_Cleaned/hba1c_dataset.rdata")
 # Ethnicity
-chco_demo <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Raw/Data From CHCO/Registry/Copy of Earliest Trikafta Dates in Registry_updated 2.2.21 cc modified_cfrdyn.csv")
+chco_demo <- read.csv("./Prepost Trikafta/Data_Raw/Data From CHCO/Registry/Copy of Earliest Trikafta Dates in Registry_updated 2.2.21 cc modified_cfrdyn.csv")
 chco_demo$Hispanic <- NA
@@
 chco_demo <- chco_demo[, c("CFF.ID", "Hispanic")]
-montana_demo <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/montana_demo.csv")
+montana_demo <- read.csv("./Prepost Trikafta/Data_Cleaned/montana_demo.csv")
 montana_demo$Hispanic <- montana_demo$Is.Patient.of.Hispanic.Origin
 montana_demo <- montana_demo[, c("CFF.ID", "Hispanic")]
-uw_demo <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/uw_demo.csv")
+uw_demo <- read.csv("./Prepost Trikafta/Data_Cleaned/uw_demo.csv")
 uw_demo$Hispanic <- uw_demo$Is.Patient.of.Hispanic.Origin

### 2025-01-02 Colleen Wood/Pre-post Trikafta/clean_and_merge.R

@@ -3,17 +3,19 @@ library(lubridate)
 library(childsds)
 library(rspiro)
 library(cgmanalysis)
-source("/Users/timvigers/GitHub/CF/genotype_class_severity.R")
+source("/Users/timvigers/Dropbox/GitHub/CF/genotype_class_severity.R")
+setwd("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood")
 t1_variables <- c(
   "CFF ID", "Site", "Sex", "Race", "Genotypes1", "Genotypes2", "Pancreatic Status",
   "G-tube in past 12 months", "Time", "Age",
   "BMI Percentile", "Height", "Weight", "FEV1", "FVC", "HbA1c"
 )
 # CHCO
-chco_a1c <- read_csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/chco_glycemic.csv", show_col_types = F)
+chco_a1c <- read_csv("./Prepost Trikafta/Data_Cleaned/chco_glycemic.csv", show_col_types = F)
 chco_a1c$Date <- parse_date_time(chco_a1c$Date, orders = "mdy")
-chco_bmi <- read_csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/chco_bmi_pft.csv", show_col_types = F)
+chco_bmi <- read_csv("./Prepost Trikafta/Data_Cleaned/chco_bmi_pft.csv", show_col_types = F)
 chco_bmi$Date <- parse_date_time(chco_bmi$Date, orders = "mdy")
+chco_insulin <- read_csv("./Prepost Trikafta/Data_Raw/Data From CHCO/insulin.csv", show_col_types = F)
 chco <- full_join(chco_a1c, chco_bmi)
 # Fill CFF ID by MRN
 chco <- chco %>%
@@ -23,13 +25,13 @@ chco <- chco %>%
   filter(!is.na(`CFF ID`)) %>%
   select(-MRN)
 # Add demographics, filter
[…]

### 2025-01-07 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
minor changes to GCT

@@
       dark: darkly
-bibliography: /home/tim/Documents/Miscellaneous/zotero.bib
-csl: /home/tim/GitHub/styles/american-medical-association.csl
+bibliography: /Users/timvigers/Library/CloudStorage/Dropbox/Miscellaneous/zotero.bib
+csl: /Users/timvigers/Library/CloudStorage/Dropbox/GitHub/styles/american-medical-association.csl
 editor: source
@@
   )
+# Per Christine, we can exclude C046's 140 value and keep them in the NGT
+# category because they had multiple normal OGTTs after this one and the 140 is
+# likely a transcription error
+df$`OGTT 0 min`[df$ID == "C046"] <- NA
 # Get diagnosis
@@
 )
+df$Diabetes <- factor(df$Diagnosis,
+  levels = c("NGT", "IFG", "INDET", "IGT", "CFRD"),
+  labels = c("NGT", "Pre-CFRD", "Pre-CFRD", "Pre-CFRD", "CFRD")
+)
 df$CFRD <- factor(df$`OGTT 120 min` >= 200,
@@
   labels = c("No", "Yes")
+)
+# Has OGTT vs. does not
+df$`Got OGTT and GCT?` <- apply(
+  df[, c("OGTT 120 min", "GCT 60 min plasma")], 1, function(r) {
+    !any(is.na(r))
+  }
+)
+df$`Got OGTT and GCT?` <- factor(df$`Got OGTT and GCT?`,
+  levels = c(T, F),
+  labels = c(
+    "Have GCT 1 Hour Plasma or OGTT 2 Hour",
+    "Missing GCT 1 Hour Plasma or OGTT 2 Hour"
+  )
 )
@@
 #| label: table1a
-#| tbl-cap: Participant characteristics (including those missing OGTT data)
+#| tbl-cap: Participant characteristics by OGTT status
 #| message: false
 df %>%
-  select(-ID, -Dysglycemia) %>%
-  tbl_summary(missing_text = "(Missing)")
-# Now remove those without OGTT
-df <- df %>% filter(!is.na(Diagnosis))
+  select(-ID) %>%
+  tbl_summary(by = `Got OGTT and GCT?`, missing_text = "(Missing)") %>%
+  add_overall() %>%
+  add_p() %>%
+  separate_p_footnotes()
+# Limit to 189 after this
+df <- df %>%
+  filter(`Got OGTT and GCT?` == "Have GCT 1 Hour Plasma or OGTT 2 Hour") %>%
+  select(-`Got OGTT and GCT?`)
 ```
@@
 #| label: table1b
-#| tbl-cap: Participant characteristics by diagnosis
+#| tbl-cap: Participant characteristics by CFRD status
 #| message: false
 df %>%
-  select(-ID, -Dysglycemia) %>%
-  tbl_summary(by = Diagnosis, missing_text = "(Missing)") %>%
+  select(-ID) %>%
+  tbl_summary(by = Diabetes, missing_text = "(Missing)") %>%
   add_overall() %>%
@@
 #| warning: false
-plot_df <- df %>% pivot_longer(contains("OGTT"),
+plot_df <- df %>%
+  pivot_longer(contains("OGTT"),
     names_to = "OGTT Timepoint", values_to = "Glucose"
@@
   scale_y_continuous(breaks = c(100, 140, 200, 300)) +
-  facet_wrap(~Diagnosis) +
-  theme_bw() +
-  theme(legend.position = "none")
+  facet_wrap(~Diabetes) +
+  theme_bw(base_size = 20) +
+  theme(
+    legend.position = "none",
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.x = element_blank()
+  )
 box_p
-```
-
-```{r}
-#| label: fig-dot
-#| fig-cap: Dot plot of OGTT glucose values by diabetes status
-#| warning: false
-dot_p <- ggplot(
-  plot_df,
-  aes(x = `OGTT Timepoint`, y = Glucose, color = `OGTT Timepoint`)
-) +
-  geom_jitter() +
-  scale_y_continuous(breaks = c(100, 140, 200, 300)) +
-  facet_wrap(~Diagnosis) +
-  theme_bw() +
-  theme(legend.position = "none")
-dot_p
-```
-
-```{r}
-#| label: fig-violin
-#| fig-cap: Violin plot of OGTT glucose values by diabetes status
-#| warning: false
-violin_p <- ggplot(
-  plot_df,
-  aes(x = `OGTT Timepoint`, y = Glucose, fill = `OGTT Timepoint`)
-) +
-  geom_violin() +
-  scale_y_continuous(breaks = c(100, 140, 200, 300)) +
-  facet_wrap(~Diagnosis) +
-  theme_bw() +
-  theme(legend.position = "none")
-violin_p
 ```

### 2025-01-09 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
GCT updates for Christine

@@
 library(tidyr)
+library(lubridate)
 library(ggplot2)
+library(ggpmisc)
 library(pROC)
@@
 library(gt)
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/tim/OneDrive/Vigers/CF"
-)
-github_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/GitHub",
-  "Windows" = "C:/Users/Tim/GitHub",
-  "Linux" = "/home/tim/GitHub"
-)
-knitr::opts_knit$set(root.dir = home_dir)
+knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF")
 ```
@@
     ID, HbA1c, Age, Gender, Race, Ethnicity, `Pancreatic insufficient`,
-    contains("GCT"), contains("OGTT ")
+    contains("GCT"), contains("OGTT "), "OGTT.Date.Time",
+    "Study.visit.PFTs.Date."
   )
+# Remove C044
+df <- df %>% filter(ID != "C044")
+# Time from OGTT to GCT
+df$OGTT.Date.Time <- mdy_hm(df$OGTT.Date.Time)
+df$Study.visit.PFTs.Date. <- mdy_hm(df$Study.visit.PFTs.Date.)
+df$`Days From OGTT to GCT` <-
+  as.numeric(as.Date(df$OGTT.Date.Time) - as.Date(df$Study.visit.PFTs.Date.))
+df$OGTT.Date.Time <- NULL
+df$Study.visit.PFTs.Date. <- NULL
 # Per Christine, we can exclude C046's 140 value and keep them in the NGT
@@
   select(-ID) %>%
-  tbl_summary(by = `Got OGTT and GCT?`, missing_text = "(Missing)") %>%
+  tbl_summary(by = `Got OGTT and GCT?`, missing_text = "Missing") %>%
   add_overall() %>%
@@
   select(-ID) %>%
-  tbl_summary(by = Diabetes, missing_text = "(Missing)") %>%
+  tbl_summary(by = Diabetes, missing_text = "Missing") %>%
   add_overall() %>%
@@
   select(-ID) %>%
-  tbl_summary(by = Dysglycemia) %>%
+  tbl_summary(by = Dysglycemia, missing_text = "Missing") %>%
   add_overall() %>%
@@
 plot_df <- df %>%
-  pivot_longer(contains("OGTT"),
+  pivot_longer(c(`OGTT 0 min`, `OGTT 60 min`, `OGTT 120 min`),
     names_to = "OGTT Timepoint", values_to = "Glucose"
@@
 box_p
+ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/boxplot.png",
+  box_p,
+  device = "png"
+)
+# OGTT 1 hour and GCT 1 hour regression
+scatter_p <- ggplot(df, aes(x = `GCT 60 min plasma`, y = `OGTT 60 min`)) +
+  geom_point() +
+  stat_poly_line() +
+  stat_poly_eq(use_label(c("eq", "R2")),
+    size = 5,
+    label.y = "bottom", label.x = "right"
+  ) +
+  theme_bw(base_size = 20)
+scatter_p
+ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/scatterplot.png",
+  scatter_p,
+  device = "png"
+)
 ```

### 2025-01-09 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
rename sweat test folder

---
title: "Sweat electrolytes and CFTR activity in the era of CFTR modulators"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 10
    toc-float: true
    number-sections: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(readxl)
library(gtsummary)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability")
```

```{r data cleaning}
#| message: false
# Import data
# source("/Users/timvigers/Library/CloudStorage/Dropbox/GitHub/CF/Edith Zemanick/Sweat Variability/data_import.r")
load("./Data_Cleaned/chco_data.RData")
# Exclude people without demographics
chco <- chco %>% filter(demog_avail == "True")
# Fix diagnoses/exclusion data for CHCO (based on Angela's review)
diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) |
  diagnoses$Code == "Exclude"]
exclude <- exclude[!is.na(exclude)]
chco <- chco[!(chco$patient_id %in% exclude), ]
chco$Diagnosis <- diagnoses$Diagnosis[match(chco$patient_id, diagnoses$PatientID)]
diagnosis_check <- chco %>% select(patient_id, Diagnosis, diagnosis)
# Import NJH data, remove those without demographics
njh <- read_excel("./Data_Raw/SweatVariability2025_CFgenotypes 2024-08-21.xlsx")
# Filter and select similar to CHCO
njh <- njh %>% filter(`Demographics Available`)
# Import detailed diagnoses from Katie
diagnoses <- read.csv("./Data_Cleaned/SweatVariability2025_CFgenotypes.csv",
  na.strings = c("", ".")
)
# Merge NJH and diagnoses
diagnoses <- diagnoses %>% select(Patient.ID, Diagnosis...DataScout)
colnames(diagnoses) <- c("PatientID", "Diagnosis")
# Less detailed diagnoses
diagnoses$Diagnosis <- factor(diagnoses$Diagnosis,
  levels = c(
    "Abnormal sweat choride test without diagnosis of cystic fibrosis",
    "Abnormal sweat test",
    "Alpha-1-antitrypsin deficiency",
    "Alpha-1-antitrypsin deficiency carrier",
    "Alpha-1-antitrypsin deficiency, Positive cystic fibrosis sweat test without diagnosis of cystic fibrosis",
    "CF", "CF carrier, Alpha-1-antitrypsin deficiency",
    "CF, alpha 1-antitrypsin deficiency",
    "CF, alpha 1-antitrypsin deficiency carrier",
    "CF, Alpha-1-antitrypsin deficiency",
    "Cystic fibrosis carrier",
    "Cystic fibrosis carrier, Alpha-1-antitrypsin deficiency",
    "Cystic fibrosis carrier, Alpha-1-antitrypsin deficiency carrier",
    "Cystic fibrosis screening",
    "Cystic fibrosis transmembrane conductance regulator (CFTR)-related disorder",
    "Positive cystic fibrosis sweat test without diagnosis of cystic fibrosis",
    "Primary ciliary dyskinesia",
    "Primary Ciliary dyskinesia, Cystic Fibrosis carrier"
  ),
  labels = c(
    "Non-CF", "Non-CF", "Non-CF", "CF Carrier", "Non-CF", "Cystic Fibrosis",
    "CF Carrier", "Cystic Fibrosis", "Cystic Fibrosis", "Cystic Fibrosis",
    "CF Carrier", "CF Carrier", "CF Carrier", "Non-CF", "CFTR-RD or CRMS",
    "Non-CF", "Non-CF", "CF Carrier"
  )
)
# Patient ids may not match so keep CHCO and NJH separate and check names after
```

```{r}
#| message: false
#| warning: false
df %>%
  ggplot(
    aes(x = Na, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
  ) +
  geom_point() +
  theme()
df %>%
  ggplot(
    aes(x = Cl, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
  ) +
  geom_point() +
  theme()
df %>%
  ggplot(
    aes(x = Na / Cl, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
  ) +
  geom_point() +
  theme()
```

```{r}
df %>%
  filter(Diagnosis == "Cystic Fibrosis") %>%
  mutate(
    `Age group` = cut(Age.at.Test.in.Years, c(-Inf, 18, Inf), right = F),
    `Cl < 60` = Cl < 60
  ) %>%
  select(Na, Cl, TotalGenotypeFunction, `Cl < 60`, `Age group`) %>%
  tbl_summary(by = `Age group`, missing_text = "Missing") %>%
  add_p()
```

### 2025-01-09 Edith Zemanick/Sweat Variability/sweat_test_analysis_chec.qmd
rename sweat test folder

---
title: "CHEC Study Sweat Chloride"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
csl: /home/timvigers/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(readxl)
library(ggpmisc)
library(gtsummary)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Test",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Test",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Test"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/Tim/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
source(paste0(github_dir, "/CF/genotype_class_severity.R"))
knitr::opts_knit$set(root.dir = home_dir)
```

# CHCO data

```{r}
# Edith's data
# Import exclusion data from Excel
diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) | diagnoses$Code == "Exclude"]
exclude <- exclude[!is.na(exclude)]
# Import data
df <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv",
  na.strings = ""
)
# Remove exclusions from Angela
df <- df[!(df$PatientID %in% exclude), ]
# Diagnosis per Angela
df$Diagnosis <- factor(diagnoses$Diagnosis[match(df$PatientID, diagnoses$PatientID)],
  levels = c("CFTR-related disorder", "CRMS", "Cystic Fibrosis", "Non-CF"),
  labels = c("CFTR-RD or CRMS", "CFTR-RD or CRMS", "Cystic Fibrosis", "Non-CF")
)
# Add modulator info
cftr <- read.csv("./Data_Raw/Copy of Report CFTR Modulators.csv")
df$Modulator <- apply(df, 1, function(r) {
  mod <- cftr[cftr$Patient.ID == as.numeric(r["PatientID"]), ]
  if (any(duplicated(mod$Age.at.Start..years.))) {
    mod <- mod[-which(duplicated(mod$Age.at.Start..years.)), ]
  }
  mod <- mod[order(mod$Age.at.Start..years.), ]
  age <- as.numeric(r["Age.at.Test.in.Years"])
  as.character(cut(age, c(-Inf, mod$Age.at.Start..years., Inf),
    labels = c("None", mod$Modulator), right = F
  ))
})
# Numeric variables
num_vars <- c("Amount", "Na", "K", "Cl")
num_vars <- c(paste0(num_vars, "_Left"), paste0(num_vars, "_Right"))
df[, num_vars] <- suppressWarnings(lapply(df[, num_vars], as.numeric))
# Filter per Maxie's code
exclude2 <- c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793)
no_dem <- unique(df$PatientID[df$Demographics.Available != "TRUE" | is.na(df$Demographics.Available) | df$Sex == ""])
df <- df %>%
  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks * 7) %>%
  filter(
    Age.at.Test.in.Years < 50 & Age.at.Test.in.Days >= 2,
    !(Diagnosis == "Non-CF" & PatientID %in% exclude2),
    Modulator == "None"
  ) %>%
  distinct(.) %>%
  filter(Sex != "", Demographics.Available == "TRUE")
# Genotype info
## By severity
df$geno1 <- cf_genotype_class_severity(df$Genotypes1)
df$geno2 <- cf_genotype_class_severity(df$Genotypes2)
df$geno1[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno2[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno_risk <- paste(df$geno1, df$geno2, sep = "/")
df$geno_risk <- factor(df$geno_risk)
## By function
df$geno1_class <- cf_genotype_class_severity(df$Genotypes1, out = "class")
df$geno1_class[grep("508", df$Genotypes1)] <- "F508del"
df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("I", "II", "III"), "MF")
df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("IV", "V"), "RF")
df$geno2_class <- cf_genotype_class_severity(df$Genotypes2, out = "class")
df$geno2_class[grep("508", df$Genotypes2)] <- "F508del"
df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("I", "II", "III"), "MF")
df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("IV", "V"), "RF")
df$geno_class <- factor(paste0(
  pmin(df$geno1_class, df$geno2_class), "/",
  pmax(df$geno1_class, df$geno2_class)
))
levels(df$geno_class) <- c(
  "MF/MF", "MF/MF", "MF/RF", "MF/Unknown", "MF/MF",
  "MF/RF", "MF/Unknown", "Both RF or Unknown", "Both RF or Unknown"
)
# Which arm?
df$Arm <- apply(df[, c("Cl", "Cl_Left", "Cl_Right", "Na", "Na_Left", "Na_Right", "K", "K_Left", "K_Right")], 1, function(r) {
  k <- sum(!is.na(r))
  if (k == 6) {
    return("Both")
  } else if (k == 3) {
    return("Arm Not Noted")
  } else {
    return(NA)
  }
})
# Merge left and right values
df$Amount <- apply(df[, c("Amount", "Amount_Left", "Amount_Right")], 1, function(r) {
  amt <- suppressWarnings(as.numeric(as.character(r)))
  mean(amt, na.rm = T)
})
df$Cl <- apply(df[, c("Cl", "Cl_Left", "Cl_Right")], 1, function(r) {
  cl <- suppressWarnings(as.numeric(as.character(r)))
  mean(cl, na.rm = T)
})
df$Na <- apply(df[, c("Na", "Na_Left", "Na_Right")], 1, function(r) {
  na <- suppressWarnings(as.numeric(as.character(r)))
  mean(na, na.rm = T)
})
df$K <- apply(df[, c("K", "K_Left", "K_Right")], 1, function(r) {
  k <- suppressWarnings(as.numeric(as.character(r)))
  mean(k, na.rm = T)
})
# Sodium to chloride ratio
df$`Na/Cl Ratio` <- df$Na / df$Cl
high_ratio <- unique(which(df$`Na/Cl Ratio` > 2))
high_cl <- unique(which(df$Cl > 160))
high_na <- unique(which(df$Na > 160))
df <- df[-unique(c(high_ratio, high_cl, high_na)), ]
# Sweat chloride group
df$`Sweat Chloride Group` <- cut(df$Cl, c(-Inf, 30, 60, Inf),
  right = F,
  labels = c("< 30", "30-59", ">60")
)
# Sort
df <- df %>% arrange(PatientID, Age.at.Test.in.Weeks)
# First measure for everyone, remove those with no sweat chloride
df <- df %>%
  filter(!is.na(Cl)) %>%
  group_by(PatientID) %>%
  filter(row_number() == 1) %>%
  ungroup()
# Order groups
df$Diagnosis <- factor(df$Diagnosis,
  levels = c("Cystic Fibrosis", "CFTR-RD or CRMS", "Non-CF")
)
```

## By sweat chloride group

```{r}
#| warning: false
ggplot(df, aes(x = Na)) +
  geom_histogram(aes(y = ..density.., fill = `Sweat Chloride Group`)) +
  geom_density(aes(group = `Sweat Chloride Group`)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(df, aes(x = Cl)) +
  geom_histogram(aes(y = ..density.., fill = `Sweat Chloride Group`)) +
  geom_density(aes(group = `Sweat Chloride Group`)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(df, aes(x = Na, y = Cl, colour = `Sweat Chloride Group`)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(df, aes(x = Na, y = Cl)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~`Sweat Chloride Group`, scales = "free", ncol = 2)
df %>%
  select(Na, Cl, `Sweat Chloride Group`) %>%
  tbl_summary(by = `Sweat Chloride Group`)
```

## By classification

```{r}
#| warning: false
ggplot(df, aes(x = Na)) +
  geom_histogram(aes(y = ..density.., fill = Diagnosis)) +
  geom_density(aes(group = Diagnosis)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(df, aes(x = Cl)) +
  geom_histogram(aes(y = ..density.., fill = Diagnosis)) +
  geom_density(aes(group = Diagnosis)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(df, aes(x = Na, y = Cl, colour = Diagnosis)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(df, aes(x = Na, y = Cl)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~Diagnosis, scales = "free", ncol = 2)
df %>%
  select(Na, Cl, Diagnosis) %>%
  tbl_summary(by = Diagnosis)
```

## By genotype severity

```{r}
#| warning: false
ggplot(df, aes(x = Na)) +
  geom_histogram(aes(y = ..density.., fill = geno_class)) +
  geom_density(aes(group = geno_class)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(df, aes(x = Cl)) +
  geom_histogram(aes(y = ..density.., fill = geno_class)) +
  geom_density(aes(group = geno_class)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(df, aes(x = Na, y = Cl, colour = geno_class)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(df, aes(x = Na, y = Cl)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~geno_class, scales = "free", ncol = 2)
df %>%
  select(Na, Cl, geno_class) %>%
  tbl_summary(by = geno_class)
```

# CHEC data

```{r data cleaning}
# Import CHEC data
chec <- read.csv("./Data_Raw/CHEC sweat sodium and chloride.csv")
# Chloride groups
chec$swc_group <- cut(chec$swcavg, c(-Inf, 30, 60, Inf),
  right = F,
  labels = c("< 30", "30-59", ">60")
)
chec$swc_group_pm <- cut(chec$swhcavgpm, c(-Inf, 30, 60, Inf),
  right = F,
  labels = c("< 30", "30-59", ">60")
)
```

# Pre-modulator

## By sweat chloride group

```{r}
#| warning: false
ggplot(chec, aes(x = swhsavgpm)) +
  geom_histogram(aes(y = ..density.., fill = swc_group_pm)) +
  geom_density(aes(group = swc_group_pm)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(chec, aes(x = swhcavgpm)) +
  geom_histogram(aes(y = ..density.., fill = swc_group_pm)) +
  geom_density(aes(group = swc_group_pm)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm, colour = swc_group_pm)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~swc_group_pm, scales = "free", ncol = 2)
chec %>%
  select(swhsavgpm, swhcavgpm, swc_group_pm) %>%
  tbl_summary(by = swc_group_pm)
```

## By CHEC genotype group

```{r}
#| warning: false
ggplot(chec, aes(x = swhsavgpm)) +
  geom_histogram(aes(y = ..density.., fill = genotype_group)) +
  geom_density(aes(group = genotype_group)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(chec, aes(x = swhcavgpm)) +
  geom_histogram(aes(y = ..density.., fill = genotype_group)) +
  geom_density(aes(group = genotype_group)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm, colour = genotype_group)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~genotype_group, scales = "free", ncol = 2)
chec %>%
  select(swhsavgpm, swhcavgpm, genotype_group) %>%
  tbl_summary(by = genotype_group)
```

# Post-modulator

## By sweat chloride group

```{r}
#| warning: false
ggplot(chec, aes(x = swhsavgpm)) +
  geom_histogram(aes(y = ..density.., fill = swc_group)) +
  geom_density(aes(group = swc_group)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(chec, aes(x = swhcavgpm)) +
  geom_histogram(aes(y = ..density.., fill = swc_group)) +
  geom_density(aes(group = swc_group)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(chec, aes(x = swsavg, y = swcavg, colour = swc_group)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(chec, aes(x = swsavg, y = swcavg)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~swc_group, scales = "free", ncol = 2)
chec %>%
  select(swsavg, swcavg, swc_group) %>%
  tbl_summary(by = swc_group)
```

## By CHEC genotype group

```{r}
#| warning: false
ggplot(chec, aes(x = swhsavgpm)) +
  geom_histogram(aes(y = ..density.., fill = genotype_group)) +
  geom_density(aes(group = genotype_group)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(chec, aes(x = swhcavgpm)) +
  geom_histogram(aes(y = ..density.., fill = genotype_group)) +
  geom_density(aes(group = genotype_group)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(chec, aes(x = swsavg, y = swcavg, colour = genotype_group)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(chec, aes(x = swsavg, y = swcavg)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~genotype_group, scales = "free", ncol = 2)
chec %>%
  select(swsavg, swcavg, genotype_group) %>%
  tbl_summary(by = genotype_group)
```

## By modulator type

```{r}
#| warning: false
ggplot(chec, aes(x = swhsavgpm)) +
  geom_histogram(aes(y = ..density.., fill = modqual)) +
  geom_density(aes(group = modqual)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(chec, aes(x = swhcavgpm)) +
  geom_histogram(aes(y = ..density.., fill = modqual)) +
  geom_density(aes(group = modqual)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplot(chec, aes(x = swsavg, y = swcavg, colour = modqual)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(chec, aes(x = swsavg, y = swcavg)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~modqual, scales = "free", ncol = 2)
chec %>%
  select(swsavg, swcavg, modqual) %>%
  tbl_summary(by = modqual)
```

### 2025-01-09 Edith Zemanick/Sweat Variability/data_import.r

@@ -1,0 +1,118 @@
+library(Hmisc)
+setwd("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability")
+# Read Data
+data <- read.csv("./Data_Raw/SweatVariability2025_DATA_2025-01-08_0928.csv", na.strings = "")
+# Setting Labels
+label(data$record_id) <- "Record ID"
+label(data$redcap_data_access_group) <- "Data Access Group"
+label(data$patient_id) <- "Patient ID"
+label(data$patient_name) <- "Patient Name"
+label(data$dob) <- "Birth Date"
+label(data$diagnosis) <- "Diagnosis"
+label(data$genotype1) <- "Genotype 1"
+label(data$genotype1_spec) <- "Specify mutation 1, if Other"
+label(data$tract_type1) <- "Tract type (for R117H)"
+label(data$genotype2) <- "Genotype 2"
+label(data$genotype2_spec) <- "Specify mutation 2, if Other"
+label(data$tract_type2) <- "Tract type (for R117H)"
+label(data$demog_avail) <- "Demographics Available"
+label(data$sex) <- "Sex"
+label(data$race___1) <- "Race (choice=Asian)"
+label(data$race___2) <- "Race (choice=Black/African-American)"
+label(data$race___3) <- "Race (choice=Native Hawaiian/Pacific Islander)"
+label(data$race___4) <- "Race (choice=White/Caucasian)"
+label(data$race___5) <- "Race (choice=Native American/Alaska Native)"
+label(data$race___6) <- "Race (choice=More than one race)"
+label(data$race___7) <- "Race (choice=Other)"
+label(data$race___8) <- "Race (choice=Not Reported)"
+label(data$ethnicity) <- "Ethnicity"
+label(data$patient_data_complete) <- "Complete?"
+label(data$sweat_date) <- "Sweat test date"
+label(data$age_at_test) <- "Age at test (in years)"
+label(data$method) <- "Test Method"
+label(data$site) <- "Site of test"
+label(data$amount) <- "Sweat volume/amount"
+label(data$result_na) <- "Sweat sodium"
+label(data$result_k) <- "Sweat potassium"
+label(data$result_cl_operand) <- "Chloride operand"
+label(data$result_cl) <- "Sweat chloride"
+label(data$amount_qns_left) <- "Left arm volume less than 15/QNS"
[…]

### 2025-01-10 Lilah Melzer/cough_monitor_analysis.qmd
cough monitor zero inflated models

--- 2024-12-07 Lilah Melzer/cough_monitor_analysis.qmd
@@
 library(performance)
+library(GLMMadaptive)
 ```
@@
 vacations$Vacation.End <- mdy(vacations$Vacation.End)
-# Those with end date after the last data pull will just end on the date of the
-# pull
-pull_date <- mdy("8/28/24")
+# Last date should be either the end date in REDCap or the last day with data
+# I
+pull_date <- mdy("11/05/24")
 start_end$End[is.na(start_end$End)] <- pull_date
@@
 agg_counts <- read.csv(
-  "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Raw/Cough_w_TotalCounts_0824 - Cough_agg_w_count_0828.csv",
+  "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Raw/Cough_w_TotalCounts_1105_NMDcorrected - Cough_agg_w_count_122324.csv",
   na.strings = c("", "NMD"), header = F
@@
 agg_counts$`Date/Device` <- NULL
+agg_counts$`#nights-monitoring` <- NULL
 agg_counts <- agg_counts[-1, ]
@@
   demo$Diagnosis[match(agg_counts$sid, as.numeric(as.character(demo$sid)))]
-# Save
-write.csv(agg_counts, file = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Clean/cleaned_aggregated_counts.csv", na = "", row.names = F)
-# Per night
-per_night <- agg_counts %>%
-  group_by(sid) %>%
-  summarise(
-    median_coughs = median(Coughs, na.rm = T),
-    nights_data = sum(!is.na(Coughs))
-  )
-demo$sid <- as.numeric(demo$sid)
-demo <- left_join(demo, per_night)
-label(demo$median_coughs) <- "Median Coughs per Night"
-label(demo$nights_data) <- "Nights With Cough Data"
 # Order columns
@@
     numeric.test = "kwt"
+  )
+# Exclude participant 3 per Lilah
+agg_counts <- agg_counts %>% filter(sid != 3)
+# Save
+write.csv(agg_counts,
+  file = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Clean/cleaned_aggregated_counts.csv",
+  na = "", row.names = F
 )
@@
 
-# Data for ATS abstract
+# Summary data
 
@@
   )
+# Exclude the people with 0 nights of data
+per_person <- per_person %>% filter(`Nights With Data` != 0)
 t2 <- tableby(
@@
 #| warning: false
-ggplot(agg_counts, aes(
+ggplot(agg_counts[!is.na(agg_counts$Coughs), ], aes(
   x = Night, y = Coughs, group = factor(sid), color = Diagnosis
@@
 
-## Model
+## Zero-inflated models
 
 ```{r}
-mod <- glmer(Coughs ~ Night + Diagnosis + (Night | sid),
-  data = agg_counts, family = "poisson"
+zip1 <- mixed_model(Coughs ~ Night + Diagnosis,
+  random = ~ 1 | sid, data = agg_counts,
+  family = zi.poisson(), zi_fixed = ~1,
+  zi_random = ~ 1 | sid
 )
-gtsummary::tbl_regression(mod, exponentiate = T)
-# check_model(mod, re_formula = NA, residual_type = "normal")
+zip2 <- update(zip1, zi_fixed = ~ Night + Diagnosis)
+
+anova(zip1, zip2)
+
+zinb1 <- update(zip2, family = zi.negative.binomial())
+
+anova(zip2, zinb1)
+
+summary(zinb1)
+# check_model(zinb1, re_formula = NA, residual_type = "normal")
 ```
@@
 ```{r}
-mod <- glmer(Coughs ~ Night + Diagnosis + Sick + (Night | sid),
-  data = agg_counts, family = "poisson"
+zinb2 <- update(zinb1,
+  fixed = Coughs ~ Night + Diagnosis + Sick,
+  zi_fixed = ~ Night + Diagnosis + Sick
 )
-gtsummary::tbl_regression(mod, exponentiate = T)
-# check_model(mod, re_formula = NA, residual_type = "normal")
+summary(zinb2)
+# check_model(zinb2, re_formula = NA, residual_type = "normal")
 ```

### 2025-01-14 Lilah Melzer/cough_monitor_analysis.qmd
ZINB models done

@@
 
+We compared several models using BIC.
+First, we compared a zero-inflated Poisson (ZIP) model with no fixed effects in the zero part of the model to one with time and diagnosis included as fixed effects in the zero part (both parts of the model included a random intercept for participant).
+The model with the same fixed effects in both parts was better (10088.29 vs.
+10071.32).
+Next, we compared this model to the same model structure but with a negative binomial distribution (ZINB) rather than Poisson to account for overdispersion in the non-zero part of the model.
+The ZINB was significantly better than the ZIP model (10071.32 vs.
+6810.49).
+Next, we compared the ZINB model with random intercept in both parts of the model to one with an additional random slope in the zero part, one with random slope in the count part, and one with random slope in both parts.
+The ZINB model with random slope in the non-zero part of the model but only a random intercept in the zero part was better than the simple ZINB model (6810.49 vs.
+6753.76).
+Interestingly, it was also better than the model with a random slope in both parts (6732.82 vs.
+6801.86).
+
 ```{r}
 zip1 <- mixed_model(Coughs ~ Night + Diagnosis,
-  random = ~ 1 | sid, data = agg_counts,
-  family = zi.poisson(), zi_fixed = ~1,
-  zi_random = ~ 1 | sid
+  data = agg_counts,
+  random = ~ 1 | sid, zi_random = ~ 1 | sid,
+  family = zi.poisson(), zi_fixed = ~1
 )
@@
 
-anova(zip1, zip2)
-
 zinb1 <- update(zip2, family = zi.negative.binomial())
 
-anova(zip2, zinb1)
+zinb2 <- update(zinb1, random = ~ Night | sid)
 
-summary(zinb1)
-# check_model(zinb1, re_formula = NA, residual_type = "normal")
+zinb3 <- update(zinb1, zi_random = ~ Night | sid)
+
+zinb <- zinb2
+
+summary(zinb)
 ```
@@
 
+To examine the effect of sickness on cough rate, we kept the same random effect structure as above, but added a fixed effect for sick status in both parts of the model.
+
 ```{r}
-zinb2 <- update(zinb1,
+zinb_sick <- update(zinb,
   fixed = Coughs ~ Night + Diagnosis + Sick,
@@
 )
-summary(zinb2)
-# check_model(zinb2, re_formula = NA, residual_type = "normal")
+summary(zinb_sick)
 ```

### 2025-01-14 Lilah Melzer/cough_monitor_analysis.qmd
ZINB model formatting

@@
 library(GLMMadaptive)
+library(gt)
 ```
@@
 
-summary(zinb)
+s <- summary(zinb)
+
+s$coef_table_zi %>%
+  data.frame() %>%
+  rownames_to_column() %>%
+  gt(caption = "Zero-part coefficients") %>%
+  fmt_number(decimals = 3)
+
+s$coef_table %>%
+  data.frame() %>%
+  rownames_to_column() %>%
+  gt(caption = "Count coefficients") %>%
+  fmt_number(decimals = 3)
 ```
+
+For now, the results above are still on the log scale, so a negative value in the estimate column indicates a lower odds ratio or incidence rate ratio.
+The zero-part coefficients indicate that the odds of having a night with 0 coughs were significantly lower in the CF group (log(OR) = `r round(s$coef_table_zi["DiagnosisCF",1],3)`, OR = `r round(exp(s$coef_table_zi["DiagnosisCF",1]),3)`, p = `r format.pval(s$coef_table_zi["DiagnosisCF",4],eps = 0.001,digits = 3)`).
+In other words, people with CF were more likely to cough than those without CF.
+However, the count coefficients table indicates that given a night with more than 0 coughs, people with CF coughed at a higher rate than those without CF, but this difference was not statistically significant (log(OR) = `r round(s$coef_table["DiagnosisCF",1],3)`, OR = `r round(exp(s$coef_table["DiagnosisCF",1]),3)`, p = `r format.pval(s$coef_table["DiagnosisCF",4],eps = 0.001,digits = 3)`).
 
@@
 )
-summary(zinb_sick)
+
+s <- summary(zinb_sick)
+
+s$coef_table_zi %>%
+  data.frame() %>%
+  rownames_to_column() %>%
+  gt(caption = "Zero-part coefficients") %>%
+  fmt_number(decimals = 3)
+
+s$coef_table %>%
+  data.frame() %>%
+  rownames_to_column() %>%
+  gt(caption = "Count coefficients") %>%
+  fmt_number(decimals = 3)
 ```
+
+The interpretation of these results is generally the same as above.
+The zero-part coefficients indicate that people with CF are still more likely to cough than people without CF (log(OR) = `r round(s$coef_table_zi["DiagnosisCF",1],3)`, OR = `r round(exp(s$coef_table_zi["DiagnosisCF",1]),3)`, p = `r format.pval(s$coef_table_zi["DiagnosisCF",4],eps = 0.001,digits = 3)`), even after adjusting for sickness.
+Sickness increased the odds of coughing, but the difference was not significant (log(OR) = `r round(s$coef_table_zi["SickYes",1],3)`, OR = `r round(exp(s$coef_table_zi["SickYes",1]),3)`, p = `r format.pval(s$coef_table_zi["SickYes",4],eps = 0.001,digits = 3)`).
+However, given a night with coughs, sickness increased the rate of cough significantly (log(OR) =  `r round(s$coef_table["SickYes",1],3)`, OR = `r round(exp(s$coef_table["SickYes",1]),3)`, p = `r format.pval(s$coef_table["SickYes",4],eps = 0.001,digits = 3)`).

### 2025-01-14 Lilah Melzer/cough_monitor_analysis.qmd
cough monitor sick plots

@@
   )
-# Exclude participant 3 per Lilah
-agg_counts <- agg_counts %>% filter(sid != 3)
 # Save
@@
 
+### All participants
+
 ```{r}
@@
   geom_line(alpha = 0.1) +
-  geom_smooth(aes(group = Diagnosis))
+  geom_smooth(aes(group = Diagnosis)) +
+  theme_bw()
+```
+
+### Participant 2
+
+```{r}
+#| label: fig-coughs-by-night-2
+#| fig-cap: Coughs by nights from start
+#| warning: false
+ggplot(agg_counts[agg_counts$sid == 2 & !is.na(agg_counts$Coughs), ], aes(
+  x = Night, y = Coughs, group = factor(sid), color = Sick
+)) +
+  geom_point() +
+  geom_line() +
+  theme_bw()
+```
+
+### Participant 3
+
+```{r}
+#| label: fig-coughs-by-night-3
+#| fig-cap: Coughs by nights from start
+#| warning: false
+ggplot(agg_counts[agg_counts$sid == 3 & agg_counts$Night %in% c(1:7, 96:102), ], aes(
+  x = Night, y = Coughs, group = factor(sid), color = Sick
+)) +
+  geom_point() +
+  geom_line() +
+  theme_bw()
 ```
@@
 10071.32).
+
 Next, we compared this model to the same model structure but with a negative binomial distribution (ZINB) rather than Poisson to account for overdispersion in the non-zero part of the model.
@@
 6810.49).
-Next, we compared the ZINB model with random intercept in both parts of the model to one with an additional random slope in the zero part, one with random slope in the count part, and one with random slope in both parts.
+
+Finally, we compared the ZINB model with random intercept in both parts of the model to one with an additional random slope in the zero part, one with random slope in the count part, and one with random slope in both parts.
 The ZINB model with random slope in the non-zero part of the model but only a random intercept in the zero part was better than the simple ZINB model (6810.49 vs.
@@
 The zero-part coefficients indicate that people with CF are still more likely to cough than people without CF (log(OR) = `r round(s$coef_table_zi["DiagnosisCF",1],3)`, OR = `r round(exp(s$coef_table_zi["DiagnosisCF",1]),3)`, p = `r format.pval(s$coef_table_zi["DiagnosisCF",4],eps = 0.001,digits = 3)`), even after adjusting for sickness.
-Sickness increased the odds of coughing, but the difference was not significant (log(OR) = `r round(s$coef_table_zi["SickYes",1],3)`, OR = `r round(exp(s$coef_table_zi["SickYes",1]),3)`, p = `r format.pval(s$coef_table_zi["SickYes",4],eps = 0.001,digits = 3)`).
+Sickness increased the odds of coughing, but the difference was not statistically significant (log(OR) = `r round(s$coef_table_zi["SickYes",1],3)`, OR = `r round(exp(s$coef_table_zi["SickYes",1]),3)`, p = `r format.pval(s$coef_table_zi["SickYes",4],eps = 0.001,digits = 3)`).
 However, given a night with coughs, sickness increased the rate of cough significantly (log(OR) =  `r round(s$coef_table["SickYes",1],3)`, OR = `r round(exp(s$coef_table["SickYes",1]),3)`, p = `r format.pval(s$coef_table["SickYes",4],eps = 0.001,digits = 3)`).

### 2025-01-16 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.qmd
a bunch of plot updates

@@
       dark: darkly
-bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Dropbox/GitHub/styles/american-medical-association.csl
+bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
+csl: /Users/timvigers/Documents/GitHub/styles/american-medical-association.csl
 editor: source
@@
 # Plot
-ggplot(hba1c_df, aes(x = Time)) +
+fig2=ggplot(hba1c_df, aes(x = Time)) +
   geom_point(
     aes(y = HbA1c, colour = factor(`CFF ID`), group = factor(`CFF ID`)),
-    alpha = 0.1
+    alpha = 0.3
   ) +
@@
   xlab("Days from ETI Start")
+fig2
+ggsave("./Prepost Trikafta/Dissemination/Figures/Figure 2.tiff", fig2,
+  width = 7, height = 5, units = "in"
+)
 # Print table
@@
 # Plot
-ggplot(df, aes(x = `Pre or Post`, y = HbA1c, group = `CFF ID`)) +
+fig1 = ggplot(df, aes(x = `Pre or Post`, y = HbA1c, group = `CFF ID`)) +
   geom_point(alpha = 0.2) +
@@
   facet_wrap(~`Baseline HbA1c`)
+fig1
+ggsave("./Prepost Trikafta/Dissemination/Figures/Figure 1.tiff", fig1,
+  width = 7, height = 5, units = "in"
+)
 ```
@@
 ggplot(df, aes(x = `Pre or Post`, y = ppFEV1gl, group = `CFF ID`)) +
-  geom_point(alpha = 0.2) +
+  geom_point(alpha = 0.5) +
   geom_line(alpha = 0.2) +

### 2025-01-16 Lilah Melzer/cough_monitor_analysis.qmd
a bunch of plot updates

--- 2025-01-14 Lilah Melzer/cough_monitor_analysis.qmd
@@
   geom_line(alpha = 0.1) +
-  geom_smooth(aes(group = Diagnosis))
+  geom_smooth(aes(group = Diagnosis)) +
+  theme_bw()
 ```
 
-## Zero-inflated models
+### By participant
 
-We compared several models using BIC.
-First, we compared a zero-inflated Poisson (ZIP) model with no fixed effects in the zero part of the model to one with time and diagnosis included as fixed effects in the zero part (both parts of the model included a random intercept for participant).
-The model with the same fixed effects in both parts was better (10088.29 vs.
-10071.32).
-Next, we compared this model to the same model structure but with a negative binomial distribution (ZINB) rather than Poisson to account for overdispersion in the non-zero part of the model.
-The ZINB was significantly better than the ZIP model (10071.32 vs.
-6810.49).
-Next, we compared the ZINB model with random intercept in both parts of the model to one with an additional random slope in the zero part, one with random slope in the count part, and one with random slope in both parts.
-The ZINB model with random slope in the non-zero part of the model but only a random intercept in the zero part was better than the simple ZINB model (6810.49 vs.
-6753.76).
-Interestingly, it was also better than the model with a random slope in both parts (6732.82 vs.
-6801.86).
+```{r results='asis'}
+#| label: fig-coughs-by-night-2
+#| fig-cap: Coughs by nights from start
+#| warning: false
+invisible(lapply(unique(agg_counts$sid), function(s) {
+  if (s == 3) {
+    d <- agg_counts[agg_counts$sid == s & agg_counts$Night %in% c(1:7, 96:102), ]
+  } else {
+    d <- agg_counts[agg_counts$sid == s & !is.na(agg_counts$Coughs), ]
+  }
 
-```{r}
-zip1 <- mixed_model(Coughs ~ Night + Diagnosis,
-  data = agg_counts,
-  random = ~ 1 | sid, zi_random = ~ 1 | sid,
-  family = zi.poisson(), zi_fixed = ~1
-)
-zip2 <- update(zip1, zi_fixed = ~ Night + Diagnosis)
-
-zinb1 <- update(zip2, family = zi.negative.binomial())
-
-zinb2 <- update(zinb1, random = ~ Night | sid)
-
-zinb3 <- update(zinb1, zi_random = ~ Night | sid)
-
-zinb <- zinb2
-
-summary(zinb)
+  diag <- d$Diagnosis[1]
+  p <- ggplot(d, aes(
+    x = Night, y = Coughs, group = factor(sid), color = Sick
+  )) +
+    geom_point() +
+    geom_line() +
+    theme_bw() +
+    ylim(0, 50) +
+    xlim(0, 147) +
+    scale_color_manual(values = c("#377EB8", "#E41A1C")) +
+    ggtitle(paste0("Participant ", s, " (", diag, ")"))
+  if (s == 38) {
+    p <- p + ylim(0, 270)
+  }
+  cat("\n")
+  cat("\n")
+  print(p)
+  cat("\n")
+  cat("\n")
+}))
 ```
-
-### By sick status
-
-To examine the effect of sickness on cough rate, we kept the same random effect structure as above, but added a fixed effect for sick status in both parts of the model.
-
-```{r}
-zinb_sick <- update(zinb,
-  fixed = Coughs ~ Night + Diagnosis + Sick,
-  zi_fixed = ~ Night + Diagnosis + Sick
-)
-summary(zinb_sick)
-```

### 2025-01-16 Scott Sagel/PEx in CF/analysis_for_yaron.qmd
a bunch of plot updates

--- 2024-11-07 Scott Sagel/PEx in CF/analysis_for_yaron.qmd
@@
 library(patchwork)
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Scott Sagel/PEx in CF",
-  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Scott Sagel/PEx in CF",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Scott Sagel/PEx in CF"
-)
-github_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/GitHub",
-  "Windows" = "C:/Users/timvigers/Documents/GitHub",
-  "Linux" = "/home/timvigers/GitHub"
-)
-knitr::opts_knit$set(root.dir = home_dir)
+knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Scott Sagel/PEx in CF")
 # Colors for plots
@@
 #| fig-cap: Number of hospitalizations
-fig1 <- df %>%
+p1 <- df %>%
   group_by(year) %>%
@@
   scale_fill_manual(values = c(purple, yellow))
-fig1
+p1
 ```
@@
 #| fig-cap: Percent hospitalized 2 or more times in both years
-fig2 <- df %>%
+p2 <- df %>%
   group_by(sid, year) %>%
@@
   scale_fill_manual(values = c(purple, yellow))
-fig2
+p2
 ```
@@
 #| fig-cap: Hospitalizations
-fig1 + fig2
+fig1 <- p1 + p2 + plot_annotation(tag_levels = c("a", "b"))
+fig1
+ggsave("./Dissemination/Figures/Figure 1.tiff", fig1,
+  width = 7, height = 5, units = "in"
+)
 ```
@@
 #| warning: false
-df %>%
+fig2 <- df %>%
   group_by(year, sid) %>%
@@
   scale_fill_manual(values = c(purple, yellow))
+fig2
+ggsave("./Dissemination/Figures/Figure 2.tiff", fig2,
+  width = 7, height = 5, units = "in"
+)
 ```
@@
 #| fig-cap: Proportions who returned to >= 90% and >= 100% at discharge
-df %>%
+figE1 <- df %>%
   mutate(
@@
   )
+figE1
+ggsave("./Dissemination/Figures/Figure E1.tiff", figE1,
+  width = 7, height = 5, units = "in"
+)
 ```

### 2025-01-17 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
GCT analysis calculated FEV1pp (lots of missing data)

@@
       dark: darkly
-bibliography: /Users/timvigers/Library/CloudStorage/Dropbox/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Library/CloudStorage/Dropbox/GitHub/styles/american-medical-association.csl
+bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
+csl: /Users/timvigers/Documents/GitHub/styles/american-medical-association.csl
 editor: source
@@
 library(gt)
+library(rspiro)
+library(childsds)
 knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF")
@@
     "Age" = Age.at.enrollment..years.,
-    "Pancreatic insufficient" = Pancreatic.exocrine.insufficient
+    "Pancreatic insufficient" = Pancreatic.exocrine.insufficient,
+    "Weight (kg)" = Body.weight..kg., "Height (cm)" = Height..cm.,
+    "Highest FEV1" = Highest.FEV1..last.12.months..FEV1..L.,
+    "Second Highest FEV1" =
+      Second.Highest.FEV1..12.months...FEV1..2nd.highest.in.L..,
+    "Type 2 diabetes in the family - Father" =
+      Type.2.diabetes.in.the.family...Father,
+    "Type 2 diabetes in the family - Mother" =
+      Type.2.diabetes.in.the.family...Mother,
+    "Brothers (number with diabetes)" = Brothers..number.with.diabetes.,
+    "Sisters (number with diabetes)" = Sisters..number.with.diabetes.,
+    "Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)" =
+      Number.of.other.relatives.with.diabetes..aunts..uncles..cousins..grandparents.
   ) %>%
   select(
-    ID, HbA1c, Age, Gender, Race, Ethnicity, `Pancreatic insufficient`,
+    ID, HbA1c, Age, `Weight (kg)`, `Height (cm)`, BMI,
+    Gender, Race, Ethnicity, `Pancreatic insufficient`,
+    `Type 2 diabetes in the family - Father`:`Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)`,
+    `Highest FEV1`, `Second Highest FEV1`,
     contains("GCT"), contains("OGTT "), "OGTT.Date.Time",
@@
 df <- df %>% filter(ID != "C044")
+# Calculate FEV1 pp and height, weight, and BMI percentiles
+lung_function <- df %>%
+  select(
+    ID, Age, Gender, Race, `Height (cm)`, `Highest FEV1`,
+    `Second Highest FEV1`
+  ) %>%
+  filter(!is.na(Gender))
+lung_function$`Average of Highest FEV1s` <-
+  rowMeans(lung_function[, c("Highest FEV1", "Second Highest FEV1")], na.rm = T)
+lung_function$Gender <- factor(lung_function$Gender,
+  levels = c("Male", "Female"),
+  labels = c("male", "female")
+)
+lung_function$Race <- factor(lung_function$Race,
+  levels = c(
+    "White", "Black/African American",
+    "Asian Non-Indian", "Asian Indian", "American Indian/Alaska Native",
+    "Native Hawaiian/Other Pacific Islander"
+  ),
+  labels = c(
+    "Caucasian", "African-American", "NE Asian", "SE Asian", "Other/mixed",
+    "Other/mixed"
+  )
+)
+lung_function$`Highest FEV1pp (2012)` <- pctpred_GLI(
+  age = lung_function$Age, height = lung_function$`Height (cm)` / 100,
+  gender = lung_function$Gender, ethnicity = lung_function$Race,
+  FEV1 = lung_function$`Highest FEV1`
+)
+lung_function$`Second Highest FEV1pp (2012)` <- pctpred_GLI(
+  age = lung_function$Age, height = lung_function$`Height (cm)` / 100,
+  gender = lung_function$Gender, ethnicity = lung_function$Race,
+  FEV1 = lung_function$`Second Highest FEV1`
+)
+lung_function$`Average of Highest FEV1pp (2012)` <- pctpred_GLI(
+  age = lung_function$Age, height = lung_function$`Height (cm)` / 100,
+  gender = lung_function$Gender, ethnicity = lung_function$Race,
+  FEV1 = lung_function$`Average of Highest FEV1s`
+)
+
+lung_function$`Highest FEV1pp (2022)` <- pctpred_GLIgl(
+  age = lung_function$Age, height = lung_function$`Height (cm)` / 100,
+  gender = lung_function$Gender, FEV1 = lung_function$`Highest FEV1`
+)
+lung_function$`Second Highest FEV1pp (2022)` <- pctpred_GLIgl(
+  age = lung_function$Age, height = lung_function$`Height (cm)` / 100,
+  gender = lung_function$Gender, FEV1 = lung_function$`Second Highest FEV1`
+)
+lung_function$`Average of Highest FEV1pp (2012)` <- pctpred_GLIgl(
+  age = lung_function$Age, height = lung_function$`Height (cm)` / 100,
+  gender = lung_function$Gender,
+  FEV1 = lung_function$`Average of Highest FEV1s`
+)
+
+df <- left_join(
+  df,
+  lung_function %>%
+    select(ID, `Highest FEV1pp (2012)`:`Second Highest FEV1pp (2022)`)
+)
+# Height, weight, BMI percentiles
+df$`Height %ile` <- sds(df$`Height (cm)`,
+  age = ifelse(df$Age > 20, 20, df$Age),
+  item = "height2_20", ref = cdc.ref, type = "perc",
+  male = "Male", female = "Female",
+  sex = as.character(df$Gender)
+)
+df$`Weight %ile` <- sds(df$`Weight (kg)`,
+  age = ifelse(df$Age > 20, 20, df$Age),
+  item = "weight2_20", ref = cdc.ref, type = "perc",
+  male = "Male", female = "Female",
+  sex = as.character(df$Gender)
+)
+df$`BMI %ile` <- sds(df$BMI,
+  age = ifelse(df$Age > 20, 20, df$Age),
+  item = "bmi", ref = cdc.ref, type = "perc",
+  male = "Male", female = "Female",
+  sex = as.character(df$Gender)
+)
+df <- df %>% select(-`Highest FEV1`, -`Second Highest FEV1`)
 # Time from OGTT to GCT
@@
   select(-ID) %>%
-  tbl_summary(by = `Got OGTT and GCT?`, missing_text = "Missing") %>%
+  tbl_summary(
+    by = `Got OGTT and GCT?`, missing_text = "Missing",
+    type =
+      "Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)" ~
+        "continuous2"
+  ) %>%
   add_overall() %>%
@@
   select(-ID) %>%
-  tbl_summary(by = Diabetes, missing_text = "Missing") %>%
+  tbl_summary(
+    by = Diabetes, missing_text = "Missing",
+    type =
+      "Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)" ~
+        "continuous2"
+  ) %>%
   add_overall() %>%
@@
   select(-ID) %>%
-  tbl_summary(by = Dysglycemia, missing_text = "Missing") %>%
+  tbl_summary(
+    by = Dysglycemia, missing_text = "Missing",
+    type =
+      "Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)" ~
+        "continuous2"
+  ) %>%
   add_overall() %>%

### 2025-01-17 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
removed FEV1pp calculations

@@
     "Weight (kg)" = Body.weight..kg., "Height (cm)" = Height..cm.,
-    "Highest FEV1" = Highest.FEV1..last.12.months..FEV1..L.,
-    "Second Highest FEV1" =
-      Second.Highest.FEV1..12.months...FEV1..2nd.highest.in.L..,
+    "Highest FEV1pp" = Highest.FEV1..last.12.months....FEV1.highest....,
+    "Second Highest FEV1pp" =
+      Second.Highest.FEV1..12.months....FEV1.2nd.highest....,
+    "Average of Highest FEV1pp" =
+      Average.of.the.two.highest.FEV1.s...Average.FEV1.......,
     "Type 2 diabetes in the family - Father" =
@@
     `Type 2 diabetes in the family - Father`:`Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)`,
-    `Highest FEV1`, `Second Highest FEV1`,
+    `Highest FEV1pp`, `Second Highest FEV1pp`, `Average of Highest FEV1pp`,
     contains("GCT"), contains("OGTT "), "OGTT.Date.Time",
@@
 df <- df %>% filter(ID != "C044")
-# Calculate FEV1 pp and height, weight, and BMI percentiles
-lung_function <- df %>%
-  select(
-    ID, Age, Gender, Race, `Height (cm)`, `Highest FEV1`,
-    `Second Highest FEV1`
-  ) %>%
-  filter(!is.na(Gender))
-lung_function$`Average of Highest FEV1s` <-
-  rowMeans(lung_function[, c("Highest FEV1", "Second Highest FEV1")], na.rm = T)
-lung_function$Gender <- factor(lung_function$Gender,
-  levels = c("Male", "Female"),
-  labels = c("male", "female")
-)
-lung_function$Race <- factor(lung_function$Race,
-  levels = c(
-    "White", "Black/African American",
-    "Asian Non-Indian", "Asian Indian", "American Indian/Alaska Native",
-    "Native Hawaiian/Other Pacific Islander"
-  ),
-  labels = c(
-    "Caucasian", "African-American", "NE Asian", "SE Asian", "Other/mixed",
-    "Other/mixed"
-  )
-)
-lung_function$`Highest FEV1pp (2012)` <- pctpred_GLI(
-  age = lung_function$Age, height = lung_function$`Height (cm)` / 100,
-  gender = lung_function$Gender, ethnicity = lung_function$Race,
-  FEV1 = lung_function$`Highest FEV1`
-)
-lung_function$`Second Highest FEV1pp (2012)` <- pctpred_GLI(
-  age = lung_function$Age, height = lung_function$`Height (cm)` / 100,
-  gender = lung_function$Gender, ethnicity = lung_function$Race,
-  FEV1 = lung_function$`Second Highest FEV1`
-)
-lung_function$`Average of Highest FEV1pp (2012)` <- pctpred_GLI(
-  age = lung_function$Age, height = lung_function$`Height (cm)` / 100,
-  gender = lung_function$Gender, ethnicity = lung_function$Race,
-  FEV1 = lung_function$`Average of Highest FEV1s`
-)
-
-lung_function$`Highest FEV1pp (2022)` <- pctpred_GLIgl(
-  age = lung_function$Age, height = lung_function$`Height (cm)` / 100,
-  gender = lung_function$Gender, FEV1 = lung_function$`Highest FEV1`
-)
-lung_function$`Second Highest FEV1pp (2022)` <- pctpred_GLIgl(
-  age = lung_function$Age, height = lung_function$`Height (cm)` / 100,
-  gender = lung_function$Gender, FEV1 = lung_function$`Second Highest FEV1`
-)
-lung_function$`Average of Highest FEV1pp (2012)` <- pctpred_GLIgl(
-  age = lung_function$Age, height = lung_function$`Height (cm)` / 100,
-  gender = lung_function$Gender,
-  FEV1 = lung_function$`Average of Highest FEV1s`
-)
-
-df <- left_join(
-  df,
-  lung_function %>%
-    select(ID, `Highest FEV1pp (2012)`:`Second Highest FEV1pp (2022)`)
-)
 # Height, weight, BMI percentiles
@@
 )
-df <- df %>% select(-`Highest FEV1`, -`Second Highest FEV1`)
 # Time from OGTT to GCT

### 2025-01-22 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
sweat test data problems

@@
 # Import data
-# source("/Users/timvigers/Library/CloudStorage/Dropbox/GitHub/CF/Edith Zemanick/Sweat Variability/data_import.r")
-load("./Data_Cleaned/chco_data.RData")
+chco <- read.csv("./Data_Raw/SweatVariability2025_DATA_LABELS_2025-01-21_1211.csv",
+  na.strings = ""
+)
+# Format dates
+chco$Birth.Date <- as.Date(chco$Birth.Date, "%Y-%m-%d")
+chco$Sweat.test.date <- as.Date(chco$Sweat.test.date, "%Y-%m-%d")
 # Exclude people without demographics
-chco <- chco %>% filter(demog_avail == "True")
+chco <- chco %>% filter(Demographics.Available == "True")
 # Fix diagnoses/exclusion data for CHCO (based on Angela's review)
@@
 exclude <- exclude[!is.na(exclude)]
-chco <- chco[!(chco$patient_id %in% exclude), ]
-chco$Diagnosis <- diagnoses$Diagnosis[match(chco$patient_id, diagnoses$PatientID)]
-diagnosis_check <- chco %>% select(patient_id, Diagnosis, diagnosis)
+chco <- chco[!(chco$Patient.ID %in% exclude), ]
+chco$New.Diagnosis <- diagnoses$Diagnosis[match(chco$Patient.ID, diagnoses$PatientID)]
+diagnosis_check <- chco %>% select(Patient.ID, Diagnosis, New.Diagnosis)
 # Import NJH data, remove those without demographics
-njh <- read_excel("./Data_Raw/SweatVariability2025_CFgenotypes 2024-08-21.xlsx")
+njh <- read_excel("./Data_Raw/SweatVariability2025_CFgenotypes 2024-08-21.xlsx", .name_repair = "universal")
 # Filter and select similar to CHCO
-njh <- njh %>% filter(`Demographics Available`)
+njh <- njh %>% filter(Demographics.Available)
+# Convert dates
+njh$Birth.Date <- as.Date(njh$Birth.Date, origin = "1899-12-30")
+njh$Sweat.test.date <- as.Date(njh$Sweat.test.date, origin = "1899-12-30")
 # Import detailed diagnoses from Katie
@@
 # Patient ids may not match so keep CHCO and NJH separate and check names after
+njh$New.Diagnosis <- diagnoses$Diagnosis[match(njh$Patient.ID, diagnoses$PatientID)]
+diagnosis_check_njh <- njh %>% select(Patient.ID, Diagnosis, New.Diagnosis)
+# Remove unnecessary columns and combine
+chco <- chco %>%
+  mutate(Dataset = "CHCO") %>%
+  select(
+    Dataset, Record.ID, Patient.ID, Patient.Name, Birth.Date, Diagnosis,
+    New.Diagnosis, Genotype.1, Genotype.2, Sweat.test.date,
+    Age.at.test..in.years., Site.of.test:Sweat.chloride.right.arm
+  )
+njh <- njh %>%
+  mutate(Dataset = "NJH") %>%
+  select(
+    Dataset, Record.ID, Patient.ID, Patient.Name, Birth.Date, Diagnosis,
+    New.Diagnosis, Genotype.1, Genotype.2, Sweat.test.date,
+    Age.at.test..in.years., Site.of.test:Sweat.chloride.right.arm
+  )
+df <- rbind(chco, njh)
+df <- df %>% arrange(Patient.ID, Sweat.test.date, Dataset)
+# Save
+write.csv(df, file = "./Data_Cleaned/combined_data.csv", row.names = F, na = "")
 ```

### 2025-01-24 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
updates for Monique (GCT)

@@
 library(pROC)
-library(gtsummary)
-library(gt)
 library(rspiro)
 library(childsds)
+library(arsenal)
+library(gt)
+library(reporttools)
 knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF")
@@
       Average.of.the.two.highest.FEV1.s...Average.FEV1.......,
+  ) %>%
+  mutate(
     "Type 2 diabetes in the family - Father" =
-      Type.2.diabetes.in.the.family...Father,
+      factor(Type.2.diabetes.in.the.family...Father),
     "Type 2 diabetes in the family - Mother" =
-      Type.2.diabetes.in.the.family...Mother,
-    "Brothers (number with diabetes)" = Brothers..number.with.diabetes.,
-    "Sisters (number with diabetes)" = Sisters..number.with.diabetes.,
+      factor(Type.2.diabetes.in.the.family...Mother),
+    "Brothers (number with diabetes)" = factor(Brothers..number.with.diabetes.),
+    "Sisters (number with diabetes)" = factor(Sisters..number.with.diabetes.),
     "Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)" =
@@
 )
+# Drop levels
+df$Race <- droplevels(df$Race)
+# Set table options
+mycontrols <-
+  tableby.control(
+    numeric.stats = c("N", "Nmiss2", "meansd", "range"),
+    cat.stats = c("N", "Nmiss2", "countpct")
+  )
 # Write for Christine to check
@@
 
-```{r}
+## By OGTT status
+
+```{r results='asis'}
 #| label: table1a
@@
 #| message: false
-df %>%
-  select(-ID) %>%
-  tbl_summary(
-    by = `Got OGTT and GCT?`, missing_text = "Missing",
-    type =
-      "Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)" ~
-        "continuous2"
-  ) %>%
-  add_overall() %>%
-  add_p() %>%
-  separate_p_footnotes()
+# Basic table formula
+f <- ~ HbA1c + kwt(Age, "N", "Nmiss2", "medianq1q3", "range") +
+  kwt(`Weight (kg)`, "N", "Nmiss2", "medianq1q3", "range") + `Height (cm)` +
+  kwt(BMI, "N", "Nmiss2", "medianq1q3", "range") + Gender + fe(Race) +
+  fe(Ethnicity) + fe(`Pancreatic insufficient`) +
+  fe(`Type 2 diabetes in the family - Father`) +
+  fe(`Type 2 diabetes in the family - Mother`) +
+  fe(`Brothers (number with diabetes)`) + fe(`Sisters (number with diabetes)`) +
+  kwt(
+    `Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)`,
+    "N", "Nmiss2", "medianq1q3", "range"
+  ) +
+  `Highest FEV1pp` + `Second Highest FEV1pp` + `Average of Highest FEV1pp` +
+  kwt(`GCT 0 min plasma`, "N", "Nmiss2", "medianq1q3", "range") +
+  kwt(`GCT 0 min capillary`, "N", "Nmiss2", "medianq1q3", "range") +
+  kwt(`GCT 60 min plasma`, "N", "Nmiss2", "medianq1q3", "range") +
+  kwt(`GCT 60 min capillary`, "N", "Nmiss2", "medianq1q3", "range") +
+  kwt(`OGTT 0 min`, "N", "Nmiss2", "medianq1q3", "range") +
+  kwt(`OGTT 60 min`, "N", "Nmiss2", "medianq1q3", "range") +
+  kwt(`OGTT 120 min`, "N", "Nmiss2", "medianq1q3", "range") +
+  kwt(`Height %ile`, "N", "Nmiss2", "medianq1q3", "range") +
+  kwt(`Weight %ile`, "N", "Nmiss2", "medianq1q3", "range") +
+  kwt(`BMI %ile`, "N", "Nmiss2", "medianq1q3", "range") +
+  kwt(`Days From OGTT to GCT`, "N", "Nmiss2", "medianq1q3", "range") +
+  fe(Diagnosis) + fe(Dysglycemia) + fe(Diabetes) + fe(CFRD)
+t1a <- tableby(update(f, `Got OGTT and GCT?` ~ . - `Got OGTT and GCT?`),
+  data = df, control = mycontrols
+)
+summary(t1a, pfootnote = T, digits = 2)
 # Limit to 189 after this
@@
 
-```{r}
+## By CFRD status
+
+```{r results='asis'}
 #| label: table1b
@@
 #| message: false
-df %>%
-  select(-ID) %>%
-  tbl_summary(
-    by = Diabetes, missing_text = "Missing",
-    type =
-      "Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)" ~
-        "continuous2"
-  ) %>%
-  add_overall() %>%
-  add_p() %>%
-  separate_p_footnotes()
+t1b <- tableby(update(f, Diabetes ~ . - Diabetes),
+  data = df, control = mycontrols
+)
+summary(t1b, pfootnote = T, digits = 2)
 ```
 
+### Pairwise comparisons
+
 ```{r}
+pairwise.t.test(df$`Average of Highest FEV1pp`, df$Diabetes)
+pairwise.fisher.test(df$`Pancreatic insufficient`, df$Diabetes,
+  p.adjust.method = "holm"
+)
+pairwise.t.test(df$HbA1c, df$Diabetes)
+pairwise.wilcox.test(df$`GCT 60 min plasma`, df$Diabetes)
+pairwise.wilcox.test(df$`GCT 60 min capillary`, df$Diabetes)
+pairwise.wilcox.test(df$`OGTT 0 min`, df$Diabetes)
+pairwise.wilcox.test(df$`OGTT 60 min`, df$Diabetes)
+pairwise.wilcox.test(df$`OGTT 120 min`, df$Diabetes)
+pairwise.wilcox.test(df$`Days From OGTT to GCT`, df$Diabetes)
+```
+
+## By dysglycemia status
+
+```{r results='asis'}
 #| label: table1c
@@
 #| message: false
-df %>%
-  select(-ID) %>%
-  tbl_summary(
-    by = Dysglycemia, missing_text = "Missing",
-    type =
-      "Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)" ~
-        "continuous2"
-  ) %>%
-  add_overall() %>%
-  add_p() %>%
-  separate_p_footnotes()
+t1c <- tableby(update(f, Dysglycemia ~ . - Dysglycemia),
+  data = df, control = mycontrols
+)
+summary(t1c, pfootnote = T, digits = 2)
 ```
@@
   function(roc) {
-    print(plot(roc, print.auc = T, print.thres = T))
+    print(plot(roc, print.auc = T, print.thres = T, legacy.axes = T))
   }

### 2025-01-24 Lilah Melzer/cough_monitor_analysis.qmd
started CFQ-R surveys

@@
 library(redcapAPI)
-library(tidyverse)
 library(rspiro)
@@
 library(GLMMadaptive)
-library(gt)
+library(tidyverse)
+library(ggstats)
 ```
@@
 # Last date should be either the end date in REDCap or the last day with data
-# I
 pull_date <- mdy("11/05/24")
@@
 invisible(lapply(unique(agg_counts$sid), function(s) {
-  if (s == 3) {
-    d <- agg_counts[agg_counts$sid == s & agg_counts$Night %in% c(1:7, 96:102), ]
-  } else {
-    d <- agg_counts[agg_counts$sid == s & !is.na(agg_counts$Coughs), ]
-  }
-
+  d <- agg_counts[agg_counts$sid == s, ]
   diag <- d$Diagnosis[1]
@@
     ylim(0, 50) +
-    xlim(0, 147) +
+    xlim(0, 150) +
     scale_color_manual(values = c("#377EB8", "#E41A1C")) +
@@
 ```
+
+# CFQ-R
+
+```{r}
+# CFQ-R Caregivers 6-13
+cfqr_6_13 <- df %>%
+  filter(redcap_repeat_instrument == "CFQ-R Caregivers 6-13") %>%
+  select(sid, Diagnosis, cfqr_caregivers_613_timestamp:cfqr_caregivers_613_complete)
+```

### 2025-01-29 Edith Zemanick/Sweat Variability/sweat_test_analysis_chec.qmd
Update sweat_test_analysis_chec.qmd

@@
     fig-cap-location: top
+    fig-width: 9
+    fig-height: 7
     page-layout: full
@@
       dark: darkly
-bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
-csl: /home/timvigers/GitHub/styles/american-medical-association.csl
+bibliography: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Personal/Miscellaneous/zotero.bib
+csl: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Personal/GitHub/styles/american-medical-association.csl
 editor: source
+execute:
+  warning: false
 ---
@@
 #| include: false
+library(Hmisc)
 library(tidyverse)
 library(readxl)
-library(ggpmisc)
-library(gtsummary)
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Test",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Test",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Test"
-)
-github_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/GitHub",
-  "Windows" = "C:/Users/Tim/GitHub",
-  "Linux" = "/home/timvigers/GitHub"
-)
-source(paste0(github_dir, "/CF/genotype_class_severity.R"))
-knitr::opts_knit$set(root.dir = home_dir)
+library(arsenal)
+library(ggpubr)
+library(performance)
+knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability")
 ```
-
-# CHCO data
-
-```{r}
-# Edith's data
-# Import exclusion data from Excel
-diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
-exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) | diagnoses$Code == "Exclude"]
-exclude <- exclude[!is.na(exclude)]
-# Import data
-df <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv",
-  na.strings = ""
-)
-# Remove exclusions from Angela
-df <- df[!(df$PatientID %in% exclude), ]
-# Diagnosis per Angela
-df$Diagnosis <- factor(diagnoses$Diagnosis[match(df$PatientID, diagnoses$PatientID)],
-  levels = c("CFTR-related disorder", "CRMS", "Cystic Fibrosis", "Non-CF"),
-  labels = c("CFTR-RD or CRMS", "CFTR-RD or CRMS", "Cystic Fibrosis", "Non-CF")
-)
-# Add modulator info
-cftr <- read.csv("./Data_Raw/Copy of Report CFTR Modulators.csv")
-df$Modulator <- apply(df, 1, function(r) {
-  mod <- cftr[cftr$Patient.ID == as.numeric(r["PatientID"]), ]
-  if (any(duplicated(mod$Age.at.Start..years.))) {
-    mod <- mod[-which(duplicated(mod$Age.at.Start..years.)), ]
-  }
-  mod <- mod[order(mod$Age.at.Start..years.), ]
-  age <- as.numeric(r["Age.at.Test.in.Years"])
-  as.character(cut(age, c(-Inf, mod$Age.at.Start..years., Inf),
-    labels = c("None", mod$Modulator), right = F
-  ))
-})
-# Numeric variables
-num_vars <- c("Amount", "Na", "K", "Cl")
-num_vars <- c(paste0(num_vars, "_Left"), paste0(num_vars, "_Right"))
-df[, num_vars] <- suppressWarnings(lapply(df[, num_vars], as.numeric))
-# Filter per Maxie's code
-exclude2 <- c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793)
-no_dem <- unique(df$PatientID[df$Demographics.Available != "TRUE" | is.na(df$Demographics.Available) | df$Sex == ""])
-df <- df %>%
-  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks * 7) %>%
-  filter(
-    Age.at.Test.in.Years < 50 & Age.at.Test.in.Days >= 2,
-    !(Diagnosis == "Non-CF" & PatientID %in% exclude2),
-    Modulator == "None"
-  ) %>%
-  distinct(.) %>%
-  filter(Sex != "", Demographics.Available == "TRUE")
-# Genotype info
-## By severity
-df$geno1 <- cf_genotype_class_severity(df$Genotypes1)
-df$geno2 <- cf_genotype_class_severity(df$Genotypes2)
-df$geno1[df$Diagnosis == "Non-CF"] <- "Not CF"
-df$geno2[df$Diagnosis == "Non-CF"] <- "Not CF"
-df$geno_risk <- paste(df$geno1, df$geno2, sep = "/")
-df$geno_risk <- factor(df$geno_risk)
-## By function
-df$geno1_class <- cf_genotype_class_severity(df$Genotypes1, out = "class")
-df$geno1_class[grep("508", df$Genotypes1)] <- "F508del"
-df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("I", "II", "III"), "MF")
-df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("IV", "V"), "RF")
-df$geno2_class <- cf_genotype_class_severity(df$Genotypes2, out = "class")
-df$geno2_class[grep("508", df$Genotypes2)] <- "F508del"
-df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("I", "II", "III"), "MF")
-df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("IV", "V"), "RF")
-df$geno_class <- factor(paste0(
-  pmin(df$geno1_class, df$geno2_class), "/",
-  pmax(df$geno1_class, df$geno2_class)
-))
-levels(df$geno_class) <- c(
-  "MF/MF", "MF/MF", "MF/RF", "MF/Unknown", "MF/MF",
-  "MF/RF", "MF/Unknown", "Both RF or Unknown", "Both RF or Unknown"
-)
-# Which arm?
-df$Arm <- apply(df[, c("Cl", "Cl_Left", "Cl_Right", "Na", "Na_Left", "Na_Right", "K", "K_Left", "K_Right")], 1, function(r) {
-  k <- sum(!is.na(r))
-  if (k == 6) {
-    return("Both")
-  } else if (k == 3) {
-    return("Arm Not Noted")
-  } else {
-    return(NA)
-  }
-})
-# Merge left and right values
-df$Amount <- apply(df[, c("Amount", "Amount_Left", "Amount_Right")], 1, function(r) {
-  amt <- suppressWarnings(as.numeric(as.character(r)))
-  mean(amt, na.rm = T)
-})
-df$Cl <- apply(df[, c("Cl", "Cl_Left", "Cl_Right")], 1, function(r) {
-  cl <- suppressWarnings(as.numeric(as.character(r)))
-  mean(cl, na.rm = T)
-})
-df$Na <- apply(df[, c("Na", "Na_Left", "Na_Right")], 1, function(r) {
-  na <- suppressWarnings(as.numeric(as.character(r)))
-  mean(na, na.rm = T)
-})
-df$K <- apply(df[, c("K", "K_Left", "K_Right")], 1, function(r) {
-  k <- suppressWarnings(as.numeric(as.character(r)))
-  mean(k, na.rm = T)
-})
-# Sodium to chloride ratio
-df$`Na/Cl Ratio` <- df$Na / df$Cl
-high_ratio <- unique(which(df$`Na/Cl Ratio` > 2))
-high_cl <- unique(which(df$Cl > 160))
-high_na <- unique(which(df$Na > 160))
-df <- df[-unique(c(high_ratio, high_cl, high_na)), ]
-# Sweat chloride group
-df$`Sweat Chloride Group` <- cut(df$Cl, c(-Inf, 30, 60, Inf),
-  right = F,
-  labels = c("< 30", "30-59", ">60")
-)
-# Sort
-df <- df %>% arrange(PatientID, Age.at.Test.in.Weeks)
-# First measure for everyone, remove those with no sweat chloride
-df <- df %>%
-  filter(!is.na(Cl)) %>%
-  group_by(PatientID) %>%
-  filter(row_number() == 1) %>%
-  ungroup()
-# Order groups
-df$Diagnosis <- factor(df$Diagnosis,
-  levels = c("Cystic Fibrosis", "CFTR-RD or CRMS", "Non-CF")
-)
-```
-
-## By sweat chloride group
-
-```{r}
-#| warning: false
-ggplot(df, aes(x = Na)) +
-  geom_histogram(aes(y = ..density.., fill = `Sweat Chloride Group`)) +
-  geom_density(aes(group = `Sweat Chloride Group`)) +
-  theme_bw() +
-  theme(
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-ggplot(df, aes(x = Cl)) +
-  geom_histogram(aes(y = ..density.., fill = `Sweat Chloride Group`)) +
-  geom_density(aes(group = `Sweat Chloride Group`)) +
-  theme_bw() +
-  theme(
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-ggplot(df, aes(x = Na, y = Cl, colour = `Sweat Chloride Group`)) +
-  geom_point(alpha = 0.2) +
-  stat_poly_line(se = F) +
-  stat_poly_eq(use_label(c("eq", "R2"))) +
-  theme_bw()
-ggplot(df, aes(x = Na, y = Cl)) +
-  geom_point(alpha = 0.2) +
-  stat_poly_line(se = F) +
-  stat_poly_eq(use_label(c("eq", "R2"))) +
-  theme_bw() +
-  facet_wrap(~`Sweat Chloride Group`, scales = "free", ncol = 2)
-df %>%
-  select(Na, Cl, `Sweat Chloride Group`) %>%
-  tbl_summary(by = `Sweat Chloride Group`)
-```
-
-## By classification
-
-```{r}
-#| warning: false
-ggplot(df, aes(x = Na)) +
-  geom_histogram(aes(y = ..density.., fill = Diagnosis)) +
-  geom_density(aes(group = Diagnosis)) +
-  theme_bw() +
-  theme(
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-ggplot(df, aes(x = Cl)) +
-  geom_histogram(aes(y = ..density.., fill = Diagnosis)) +
-  geom_density(aes(group = Diagnosis)) +
-  theme_bw() +
-  theme(
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-ggplot(df, aes(x = Na, y = Cl, colour = Diagnosis)) +
-  geom_point(alpha = 0.2) +
-  stat_poly_line(se = F) +
-  stat_poly_eq(use_label(c("eq", "R2"))) +
-  theme_bw()
-ggplot(df, aes(x = Na, y = Cl)) +
-  geom_point(alpha = 0.2) +
-  stat_poly_line(se = F) +
-  stat_poly_eq(use_label(c("eq", "R2"))) +
-  theme_bw() +
-  facet_wrap(~Diagnosis, scales = "free", ncol = 2)
-df %>%
-  select(Na, Cl, Diagnosis) %>%
-  tbl_summary(by = Diagnosis)
-```
-
-## By genotype severity
-
-```{r}
-#| warning: false
-ggplot(df, aes(x = Na)) +
-  geom_histogram(aes(y = ..density.., fill = geno_class)) +
-  geom_density(aes(group = geno_class)) +
-  theme_bw() +
-  theme(
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-ggplot(df, aes(x = Cl)) +
-  geom_histogram(aes(y = ..density.., fill = geno_class)) +
-  geom_density(aes(group = geno_class)) +
-  theme_bw() +
-  theme(
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-ggplot(df, aes(x = Na, y = Cl, colour = geno_class)) +
-  geom_point(alpha = 0.2) +
-  stat_poly_line(se = F) +
-  stat_poly_eq(use_label(c("eq", "R2"))) +
-  theme_bw()
-ggplot(df, aes(x = Na, y = Cl)) +
-  geom_point(alpha = 0.2) +
-  stat_poly_line(se = F) +
-  stat_poly_eq(use_label(c("eq", "R2"))) +
-  theme_bw() +
-  facet_wrap(~geno_class, scales = "free", ncol = 2)
-df %>%
-  select(Na, Cl, geno_class) %>%
-  tbl_summary(by = geno_class)
-```
-
-# CHEC data
 
@@
 # Import CHEC data
-chec <- read.csv("./Data_Raw/CHEC sweat sodium and chloride.csv")
-# Chloride groups
-chec$swc_group <- cut(chec$swcavg, c(-Inf, 30, 60, Inf),
-  right = F,
-  labels = c("< 30", "30-59", ">60")
+chec <- read.csv("./Data_Raw/CHEC sweat sodium and chloride.csv",
+  na.strings = c("", "NA")
 )
-chec$swc_group_pm <- cut(chec$swhcavgpm, c(-Inf, 30, 60, Inf),
-  right = F,
-  labels = c("< 30", "30-59", ">60")
+# Calculate ratios
+chec$pre_ratio <- chec$swhsavgpm / chec$swhcavgpm
+chec$post_ratio <- chec$swsavg / chec$swcavg
+# Long dataset for easier plotting and paired analysis
+chec_long <- chec %>%
+  rename(
+    pre_cl = swhcavgpm, pre_na = swhsavgpm, post_cl = swcavg, post_na = swsavg
+  ) %>%
+  pivot_longer(c(pre_cl, pre_na, post_cl, post_na, pre_ratio, post_ratio),
+    names_to = c("pre_post", ".value"),
+    names_pattern = "(.*)_(.*)",
+    values_to = "count"
   )
+chec_long$pre_post <- factor(chec_long$pre_post, levels = c("pre", "post"))
+# Set table options
+mycontrols <-
+  tableby.control(
+    numeric.stats = c("N", "Nmiss2", "meansd", "range"),
+    cat.stats = c("N", "Nmiss2", "countpct")
+  )
+# Labels
+label(chec_long$cl) <- "Sweat Cl"
+label(chec_long$na) <- "Sweat Na"
+label(chec_long$ratio) <- "Sweat Na:Cl"
 ```
 
-# Pre-modulator
+# Tables
 
-## By sweat chloride group
+## Table 1: Participant demographics by CFTR modulator
+
+```{r results='asis'}
+f <- modqual~sex + age + agecat_txt + race_txt + genotype_group
+t1 <- tableby(f, data = chec, control = mycontrols)
+summary(t1, pfootnote = T)
+```
+
+- I think that we currently don't have ethnicity data.
+
+## Table 2: Change in sweat electrolytes
+
+```{r results='asis'}
+pt1 <- paired(
+  pre_post ~
+    signed.rank(cl, "N", "Nmiss2", "medianq1q3", "range") +
+    signed.rank(na, "N", "Nmiss2", "medianq1q3", "range") +
+    signed.rank(ratio, "N", "Nmiss2", "medianq1q3", "range"),
+  data = chec_long, id = subject,
+  control = paired.control(signed.rank.exact = F)
+)
+summary(pt1, pfootnote = T)
+```
+
+## Table 3: Change in sweat electrolytes by modulator
+
+```{r results='asis'}
+pt2 <- paired(
+  pre_post ~
+    signed.rank(cl, "N", "Nmiss2", "medianq1q3", "range") +
+    signed.rank(na, "N", "Nmiss2", "medianq1q3", "range") +
+    signed.rank(ratio, "N", "Nmiss2", "medianq1q3", "range"),
+  data = chec_long, id = subject, strata = modqual,
+  control = paired.control(signed.rank.exact = F)
+)
+summary(pt2, pfootnote = T)
+```
+
+# Histograms and density plots
+
+## Sweat chloride (overall)
+
+### Histogram
 
 ```{r}
-#| warning: false
-ggplot(chec, aes(x = swhsavgpm)) +
-  geom_histogram(aes(y = ..density.., fill = swc_group_pm)) +
-  geom_density(aes(group = swc_group_pm)) +
+ggplot(chec_long, aes(x = cl, fill = pre_post)) +
+  geom_histogram(alpha = 0.5) +
   theme_bw() +
-  theme(
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-ggplot(chec, aes(x = swhcavgpm)) +
-  geom_histogram(aes(y = ..density.., fill = swc_group_pm)) +
-  geom_density(aes(group = swc_group_pm)) +
-  theme_bw() +
-  theme(
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm, colour = swc_group_pm)) +
-  geom_point(alpha = 0.2) +
-  stat_poly_line(se = F) +
-  stat_poly_eq(use_label(c("eq", "R2"))) +
-  theme_bw()
-ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm)) +
-  geom_point(alpha = 0.2) +
-  stat_poly_line(se = F) +
-  stat_poly_eq(use_label(c("eq", "R2"))) +
-  theme_bw() +
-  facet_wrap(~swc_group_pm, scales = "free", ncol = 2)
-chec %>%
-  select(swhsavgpm, swhcavgpm, swc_group_pm) %>%
-  tbl_summary(by = swc_group_pm)
+  scale_fill_manual(values = c("#EFB036", "#3B6790"))
 ```
 
-## By CHEC genotype group
+### Density
 
 ```{r}
-#| warning: false
-ggplot(chec, aes(x = swhsavgpm)) +
-  geom_histogram(aes(y = ..density.., fill = genotype_group)) +
-  geom_density(aes(group = genotype_group)) +
+ggplot(chec_long, aes(x = cl, fill = pre_post)) +
+  geom_density(alpha = 0.5) +
   theme_bw() +
-  theme(
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-ggplot(chec, aes(x = swhcavgpm)) +
-  geom_histogram(aes(y = ..density.., fill = genotype_group)) +
-  geom_density(aes(group = genotype_group)) +
-  theme_bw() +
-  theme(
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm, colour = genotype_group)) +
-  geom_point(alpha = 0.2) +
-  stat_poly_line(se = F) +
-  stat_poly_eq(use_label(c("eq", "R2"))) +
-  theme_bw()
-ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm)) +
-  geom_point(alpha = 0.2) +
-  stat_poly_line(se = F) +
-  stat_poly_eq(use_label(c("eq", "R2"))) +
-  theme_bw() +
-  facet_wrap(~genotype_group, scales = "free", ncol = 2)
-chec %>%
-  select(swhsavgpm, swhcavgpm, genotype_group) %>%
-  tbl_summary(by = genotype_group)
+  scale_fill_manual(values = c("#EFB036", "#3B6790"))
 ```
 
-# Post-modulator
+## Sweat sodium (overall)
 
-## By sweat chloride group
+### Histogram
 
 ```{r}
-#| warning: false
-ggplot(chec, aes(x = swhsavgpm)) +
-  geom_histogram(aes(y = ..density.., fill = swc_group)) +
-  geom_density(aes(group = swc_group)) +
+ggplot(chec_long, aes(x = na, fill = pre_post)) +
+  geom_histogram(alpha = 0.5) +
   theme_bw() +
-  theme(
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-ggplot(chec, aes(x = swhcavgpm)) +
-  geom_histogram(aes(y = ..density.., fill = swc_group)) +
-  geom_density(aes(group = swc_group)) +
-  theme_bw() +
-  theme(
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-ggplot(chec, aes(x = swsavg, y = swcavg, colour = swc_group)) +
-  geom_point(alpha = 0.2) +
-  stat_poly_line(se = F) +
-  stat_poly_eq(use_label(c("eq", "R2"))) +
-  theme_bw()
-ggplot(chec, aes(x = swsavg, y = swcavg)) +
-  geom_point(alpha = 0.2) +
-  stat_poly_line(se = F) +
-  stat_poly_eq(use_label(c("eq", "R2"))) +
-  theme_bw() +
-  facet_wrap(~swc_group, scales = "free", ncol = 2)
-chec %>%
-  select(swsavg, swcavg, swc_group) %>%
-  tbl_summary(by = swc_group)
+  scale_fill_manual(values = c("#EFB036", "#3B6790"))
 ```
 
-## By CHEC genotype group
+### Density
 
 ```{r}
-#| warning: false
-ggplot(chec, aes(x = swhsavgpm)) +
-  geom_histogram(aes(y = ..density.., fill = genotype_group)) +
-  geom_density(aes(group = genotype_group)) +
+ggplot(chec_long, aes(x = na, fill = pre_post)) +
+  geom_density(alpha = 0.5) +
   theme_bw() +
-  theme(
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-ggplot(chec, aes(x = swhcavgpm)) +
-  geom_histogram(aes(y = ..density.., fill = genotype_group)) +
-  geom_density(aes(group = genotype_group)) +
-  theme_bw() +
-  theme(
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-ggplot(chec, aes(x = swsavg, y = swcavg, colour = genotype_group)) +
-  geom_point(alpha = 0.2) +
-  stat_poly_line(se = F) +
-  stat_poly_eq(use_label(c("eq", "R2"))) +
-  theme_bw()
-ggplot(chec, aes(x = swsavg, y = swcavg)) +
-  geom_point(alpha = 0.2) +
-  stat_poly_line(se = F) +
-  stat_poly_eq(use_label(c("eq", "R2"))) +
-  theme_bw() +
-  facet_wrap(~genotype_group, scales = "free", ncol = 2)
-chec %>%
-  select(swsavg, swcavg, genotype_group) %>%
-  tbl_summary(by = genotype_group)
+  scale_fill_manual(values = c("#EFB036", "#3B6790"))
 ```
 
-## By modulator type
+## Sweat sodium:chloride ratio (overall)
+
+### Histogram
 
 ```{r}
-#| warning: false
-ggplot(chec, aes(x = swhsavgpm)) +
-  geom_histogram(aes(y = ..density.., fill = modqual)) +
-  geom_density(aes(group = modqual)) +
+ggplot(chec_long, aes(x = ratio, fill = pre_post)) +
+  geom_histogram(alpha = 0.5) +
   theme_bw() +
-  theme(
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-ggplot(chec, aes(x = swhcavgpm)) +
-  geom_histogram(aes(y = ..density.., fill = modqual)) +
-  geom_density(aes(group = modqual)) +
+  scale_fill_manual(values = c("#EFB036", "#3B6790"))
+```
+
+### Density
+
+```{r}
+ggplot(chec_long, aes(x = ratio, fill = pre_post)) +
+  geom_density(alpha = 0.5) +
   theme_bw() +
-  theme(
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-ggplot(chec, aes(x = swsavg, y = swcavg, colour = modqual)) +
-  geom_point(alpha = 0.2) +
-  stat_poly_line(se = F) +
-  stat_poly_eq(use_label(c("eq", "R2"))) +
-  theme_bw()
-ggplot(chec, aes(x = swsavg, y = swcavg)) +
-  geom_point(alpha = 0.2) +
-  stat_poly_line(se = F) +
-  stat_poly_eq(use_label(c("eq", "R2"))) +
+  scale_fill_manual(values = c("#EFB036", "#3B6790"))
+```
+
+## Sweat chloride by modulator
+
+### Histogram
+
+```{r}
+ggplot(chec_long, aes(x = cl, fill = pre_post)) +
+  geom_histogram(alpha = 0.5) +
   theme_bw() +
-  facet_wrap(~modqual, scales = "free", ncol = 2)
-chec %>%
-  select(swsavg, swcavg, modqual) %>%
-  tbl_summary(by = modqual)
+  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
+  facet_wrap(~modqual)
 ```
+
+### Density
+
+```{r}
+ggplot(chec_long, aes(x = cl, fill = pre_post)) +
+  geom_density(alpha = 0.5) +
+  theme_bw() +
+  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
+  facet_wrap(~modqual)
+```
+
+## Sweat sodium by modulator
+
+### Histogram
+
+```{r}
+ggplot(chec_long, aes(x = na, fill = pre_post)) +
+  geom_histogram(alpha = 0.5) +
+  theme_bw() +
+  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
+  facet_wrap(~modqual)
+```
+
+### Density
+
+```{r}
+ggplot(chec_long, aes(x = na, fill = pre_post)) +
+  geom_density(alpha = 0.5) +
+  theme_bw() +
+  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
+  facet_wrap(~modqual)
+```
+
+## Sweat sodium:chloride ratio by modulator
+
+### Histogram
+
+```{r}
+ggplot(chec_long, aes(x = ratio, fill = pre_post)) +
+  geom_histogram(alpha = 0.5) +
+  theme_bw() +
+  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
+  facet_wrap(~modqual)
+```
+
+### Density
+
+```{r}
+ggplot(chec_long, aes(x = ratio, fill = pre_post)) +
+  geom_density(alpha = 0.5) +
+  theme_bw() +
+  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
+  facet_wrap(~modqual)
+```
+
+# Boxplots
+
+```{r}
+ggplot(chec_long, aes(y = ratio, x = pre_post, fill = pre_post)) +
+  geom_boxplot() +
+  geom_signif(
+    comparison = list(c("pre", "post")), test = "wilcox.test",
+    test.args = list(paired = TRUE, exact = FALSE), y_position = 3.5,
+    map_signif_level = function(p) format.pval(p, digits = 3, eps = 0.001)
+  ) +
+  theme_bw() +
+  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
+  facet_wrap(~modqual)
+```
+
+# Scatterplots and models
+
+To evaluate whether the relationship between pre-modulator sweat chloride and pre-modulator sodium:chloride ratio is logarithmic, we compared three models:
+
+1. A simple linear model with no transformations.
+2. A linear model with log-transformed sweat chloride as the outcome.
+3. A quadratic model.
+
+The model residuals do *not* appear to be improved by log transforming the outcome.
+Also, the quadratic model is better than the linear model by BIC, but this does not remain the case when the influential point (CFF ID 1650133) is removed.
+
+```{r}
+pre_model_df <- chec %>%
+  select(cffid_first, swhcavgpm, pre_ratio) %>%
+  drop_na() %>%
+  unique()
+
+m1 <- lm(swhcavgpm ~ pre_ratio, data = pre_model_df)
+log_m <- lm(log(swhcavgpm) ~ pre_ratio, data = pre_model_df)
+m2 <- lm(swhcavgpm ~ pre_ratio + I(pre_ratio^2), data = pre_model_df)
+```
+
+## Linear model assumptions check
+
+```{r}
+check_model(m1)
+```
+
+## Log model assumptions check
+
+```{r}
+check_model(log_m)
+```
+
+## Quadratic model assumptions check
+
+```{r}
+check_model(m2)
+```
+
+## Model predicted values
+
+```{r}
+ggplot(pre_model_df, aes(y = swhcavgpm, x = pre_ratio)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = F, aes(color = "Linear")) +
+  geom_smooth(
+    method = "lm", formula = y ~ x + I(x^2), se = F,
+    aes(color = "Quadratic")
+  ) +
+  geom_smooth(
+    method = "glm", formula = y ~ x, se = F, aes(color = "Log Linear"),
+    method.args = list(family = gaussian(link = "log"))
+  ) +
+  theme_bw() +
+  theme(legend.title = element_blank())
+```

### 2025-01-29 Christine Chan/CF Registry/cf_registry_CFRD_trees.qmd
Create cf_registry_CFRD_trees.qmd

---
title: "HbA1c, random glucose, and CFRD"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-width: 12
    fig-height: 9
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(rpart)
library(rpart.plot)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry")
```

```{r data cleaning}
# Import data
encounters <- read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv",
  na.strings = c("", "N/A")
)
# Demographic data
demo <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv",
  na.strings = c("", "N/A")
)
# Race
races <- c(
  "White", "Black or African American", "American Indian or Alaska Native",
  "Asian", "Native Hawaiian or Other Pacific Islander", "Some other race"
)
demo$Race <- apply(demo, 1, function(r) {
  w <- which(r[paste0("Race", 1:6)] == 1)
  if (length(w) == 0) {
    "Unknown"
  } else if (length(w) > 1) {
    "More than one race"
  } else {
    races[w]
  }
})
demo$Race <- factor(demo$Race,
  levels = c(
    "White", "Black or African American", "American Indian or Alaska Native",
    "Asian", "More than one race", "Native Hawaiian or Other Pacific Islander",
    "Some other race"
  ),
  labels = c(
    "White", "Black or African American", "Other race", "Other race",
    "More than one race", "Other race", "Other race"
  )
)
demo$Hispanicrace <- factor(demo$Hispanicrace,
  levels = 1:2,
  labels = c("Yes", "No")
)
# F508
demo$F508 <- factor(demo$F508,
  levels = 1:3,
  labels = c("F508del/F508del", "F508del/Other", "Other/Other")
)
# Merge
encounters <- left_join(
  encounters,
  demo %>% select(eDWID, Race, Hispanicrace, Gender, F508, MutClass),
  by = join_by(eDWID)
)
# CFRD by OGTT two hour
encounters$cfrd_ogtt <- factor(encounters$twohour_bloodglucose >= 200,
  levels = c(F, T), labels = c("CFRD-", "CFRD+")
)
# CFRD by registry definition
encounters$cfrd_registry <- factor(encounters$cfrd_status,
  levels = c(1:5), labels = c("CFRD-", "CFRD-", "CFRD+", "CFRD-", "CFRD-")
)
# Dataset for decision trees
encounters <- encounters %>%
  select(
    eDWID, reviewyear, encounternum, rand_bloodglucose, hgba1c,
    ogttfast_bloodglucose, twohour_bloodglucose, cfrd_ogtt, cfrd_registry
  )
```

# Random forests

```{r}
tree1 <- rpart(cfrd_ogtt ~ ogttfast_bloodglucose + hgba1c, data = encounters,
               control = rpart.control(xval = 100))
rpart.plot(tree1, type = 1, extra = 8)
```

### 2025-01-29 Lilah Melzer/cough_monitor_analysis.qmd
CFQ-R mostly done, check with Edith and Lilah

@@
 # CFQ-R Caregivers 6-13
+# SAS code
+# if nmiss (phys1, phys2, phys3, phys4, phys5, phys6) <= 3 then
+# physical = (mean (phys1, phys2, phys3, phys4, phys5, phys6)-1)/3*100;
+#
+# if nmiss (emot7, emot8, emot9, emot10, emot11, emot12, emot13, emot14) <= 4 then
+# emotion = (mean (emot7, emot8, emot9, emot10, emot11, emot12, emot13, emot14)-1)/3*100;
+#
+# if nmiss (social20, social21, social22, social23, social24, social25, social26) <= 3 then
+# social = (mean (social20, social21, social22, social23, social24, social25, social26)-1)/3*100;
+#
+# if nmiss (eat15, eat17, eat19) <= 1 then
+# eat = (mean(eat15, eat17, eat19)-1)/3*100;
+#
+# if nmiss (body27, body28, body29) <= 1 then
+# body = (mean(body27, body28, body29)-1)/3*100;
+#
+# if nmiss (treat16, treat18, treat30) <= 1 then
+# treat = (mean (treat16, treat18, treat30)-1)/3*100;
+#
+# if nmiss (resp31, resp32, resp33, resp34) <= 2 then
+# respirat = (mean (resp31, resp32, resp33, resp34)-1)/3*100;
+#
+# if nmiss (digest35) = 0 then
+# digest = (mean (digest35)-1)/3*100;
+
 cfqr_6_13 <- df %>%
   filter(redcap_repeat_instrument == "CFQ-R Caregivers 6-13") %>%
-  select(sid, Diagnosis, cfqr_caregivers_613_timestamp:cfqr_caregivers_613_complete)
+  select(
+    sid, Diagnosis, cfqr_caregivers_613_timestamp:cfqr_caregivers_613_complete
+  )
+
+cfqr_6_13$cfqr_physical <- apply(
+  data.matrix(cfqr_6_13[, paste0("cfq_r_p_", 1:5)]), 1, function(r) {
+    (mean(r) - 1) / 3 * 100
+  }
+)
+cfqr_6_13$cfqr_emotion <- apply(
+  data.matrix(cfqr_6_13[, paste0("cfq_r_p_", 6:13)]), 1, function(r) {
+    (mean(r) - 1) / 3 * 100
+  }
+)
+
 ```

### 2025-01-29 Edith Zemanick/Sweat Variability/sweat_test_analysis_chec.qmd
CHEC excclude those without sweat sodium

@@
 bibliography: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Personal/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Personal/GitHub/styles/american-medical-association.csl
+csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
 editor: source
@@
 )
+# Only include people with either a pre-modulator or post-modulator sodium
+chec <- chec %>% filter(!is.na(swhsavgpm) | !is.na(swsavg))
 # Calculate ratios

### 2025-01-31 Jordana Hoppe/Home Collection/manuscript.qmd
fixed exclusions

@@
 github_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Documents/GitHub",
+  "Darwin" = "/Users/timvigers/GitHub",
   "Windows" = "C:/Users/Tim/GitHub",
@@
 survey_exclude <- c(15, 27, 30, 34, 47)
+data <- data %>% filter(!subject_id %in% data_exclude)
+survey_data <- survey_data %>% filter(!participant_id %in% survey_exclude)
 # Fix missing labels
@@
 survey_data %>%
-  filter(!participant_id %in% survey_exclude) %>%
   select(q1.factor:q6b___4.factor) %>%
@@
 data %>%
-  filter(!subject_id %in% data_exclude) %>%
   select(clinic_cx_source.factor:clinic_burk_sp.factor) %>%
@@
 data %>%
-  filter(!subject_id %in% data_exclude) %>%
   select(research_cx.factor:res_burk_sp.factor) %>%
@@
 data %>%
-  filter(!subject_id %in% data_exclude) %>%
   select(home_cx.factor:home_burk_sp.factor) %>%
@@
     ggsave("Jordana Hoppe/Home Collection/Dissemination/Figures/mssa_2.png",
-      plot = p,
-      width = 12, height = , units = "in"
+      plot = p, width = 12, height = , units = "in"
     )
@@
 )
-# Subset and label
-t <- survey_data %>%
-  filter(!participant_id %in% survey_exclude) %>%
-  select(participant_id, q1.factor:q6b___4.factor)
+# Label
 labels <- as.character(label(survey_data)[sub("\\.factor", "", colnames(t))])

### 2025-02-05 Edith Zemanick/Sweat Variability/sweat_test_analysis_chec.qmd
model ideas CHEC

@@
 library(ggpubr)
+library(lmerTest)
 library(performance)
@@
 ```
+
+# Model-based approaches
+
+```{r}
+m <- glmer(pre_post ~ na + cl + (1 | subject), data = chec_long, family = "binomial")
+m2 <- lmer(cl ~ pre_post * na + (1 | subject), data = chec_long)
+```

### 2025-02-05 Lilah Melzer/cough_monitor_analysis.qmd
need to add CFQ-R plots

@@
 library(ggstats)
+library(corrplot)
+knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor")
 ```
@@
 # Import use data
-start_end <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Clean/Start and end dates.csv",
+start_end <- read.csv("./Data_Clean/Start and end dates.csv",
   na.strings = ""
@@
 start_end$End <- mdy(start_end$End)
-vacations <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Clean/Vacation days.csv",
+vacations <- read.csv("./Data_Clean/Vacation days.csv",
   na.strings = ""
@@
 agg_counts <- read.csv(
-  "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Raw/Cough_w_TotalCounts_1105_NMDcorrected - Cough_agg_w_count_122324.csv",
+  "./Data_Raw/Cough_w_TotalCounts_1105_NMDcorrected - Cough_agg_w_count_122324.csv",
   na.strings = c("", "NMD"), header = F
@@
   select(
-    sid, Diagnosis, age_at_consent, height, weight, pre_fev,
+    sid, Diagnosis, age_at_consent, height, weight, pre_fev, post_fev,
     nih_sex:nih_ethnicity, mutation1, mutation2,
@@
 fev1 <- demo %>%
-  select(sid, age_at_consent, height, weight, nih_sex, pre_fev) %>%
+  select(sid, age_at_consent, height, weight, nih_sex, pre_fev, post_fev) %>%
   filter(nih_sex %in% c("Male", "Female"))
@@
   FEV1 = fev1$pre_fev
+)
+fev1$post_fev_pp <- pctpred_GLIgl(fev1$age_at_consent, fev1$height / 100,
+  factor(fev1$nih_sex, levels = c("Male", "Female")),
+  FEV1 = fev1$post_fev
 )
@@
 demo <-
-  left_join(demo, fev1[, c("sid", "height_perc", "weight_perc", "pre_fev_pp")])
+  left_join(
+    demo,
+    fev1[, c("sid", "height_perc", "weight_perc", "pre_fev_pp", "post_fev_pp")]
+  )
 label(demo$pre_fev_pp) <- "Pre-bronchodilator FEV1 % Predicted"
+label(demo$pre_fev_pp) <- "Post-bronchodilator FEV1 % Predicted"
 label(demo$height_perc) <- "Height %ile"
@@
   demo$Diagnosis[match(agg_counts$sid, as.numeric(as.character(demo$sid)))]
+# Import and merge sleep times from Lilah
+sleep <- read.csv("./Data_Clean/Total sleep times.csv", na.strings = "")
+sleep <- sleep %>%
+  select(
+    Participant.First.Name..or.any.uniq.identifier.,
+    Total.sleep.time..hours.
+  ) %>%
+  rename(
+    sid = Participant.First.Name..or.any.uniq.identifier.,
+    total_sleep = Total.sleep.time..hours.
+  )
+sleep$sid <- as.character(as.numeric(sub("FN", "", sleep$sid)))
+label(sleep$total_sleep) <- "Total sleep time (hours)"
+demo <- left_join(demo, sleep)
 # Order columns
 demo <- demo %>%
-  select(sid, height_perc, weight_perc, pre_fev_pp, everything()) %>%
+  select(sid, height_perc, weight_perc, pre_fev_pp, total_sleep, everything()) %>%
   select(-height, -weight, -pre_fev)
@@
 write.csv(agg_counts,
-  file = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Clean/cleaned_aggregated_counts.csv",
+  file = "./Data_Clean/cleaned_aggregated_counts.csv",
   na = "", row.names = F
@@
 
+# Basic statistics excluding 0s
+
+## Aggregated by person
+
+The comparisons below exclude any nights with missing data OR 0 coughs, as well as periods of sickness.
+This table could potentially be a little bit confusing because the numbers presented are the median of multiple medians, but the data are still so skewed without 0s that this seemed the best way to aggregate the data.
+
+```{r results='asis'}
+#| warning: false
+no_0 <- agg_counts %>%
+  filter(Sick == "No", !is.na(Coughs)) %>%
+  group_by(sid, Diagnosis) %>%
+  mutate(`Prop. Non-0 Coughs` = mean(Coughs > 0, na.rm = T)) %>%
+  filter(Coughs > 0) %>%
+  summarise(
+    `Prop. Non-0 Coughs` = `Prop. Non-0 Coughs`[1],
+    `Median Coughs` = median(Coughs),
+    `Cough IQR` = IQR(Coughs),
+    .groups = "drop"
+  )
+no_0 <- left_join(
+  no_0,
+  fev1 %>%
+    mutate(sid = as.numeric(sid)) %>%
+    select(sid, pre_fev_pp, post_fev_pp),
+  by = join_by(sid)
+)
+no_0_t <- tableby(
+  Diagnosis ~
+    kwt(`Prop. Non-0 Coughs`, "medianq1q3", "range") +
+    kwt(`Median Coughs`, "medianq1q3", "range"),
+  data = no_0
+)
+summary(no_0_t, pfootnote = T)
+M <- cor.test(no_0$`Median Coughs`, no_0$pre_fev_pp, method = "spearman")
+corrplot(cor(no_0 %>% select(`Median Coughs`, pre_fev_pp),
+  method = "spearman", use = "pairwise.complete.obs"
+))
+ggplot(no_0, aes(x = `Median Coughs`, y = pre_fev_pp)) +
+  geom_point()
+```
+
+Spearman correlation between median coughs and pre-bronchodilator FEV1 was `r round(as.numeric(M$estimate),3)` (p = `r format.pval(M$p.value,eps = 0.001,digits = 3)`).
+
+## Not aggregated
+
+```{r results='asis'}
+#| warning: false
+no_0 <- agg_counts %>% filter(Sick == "No", !is.na(Coughs), Coughs > 0)
+no_0 <- left_join(
+  no_0,
+  fev1 %>%
+    mutate(sid = as.numeric(sid)) %>%
+    select(sid, pre_fev_pp, post_fev_pp),
+  by = join_by(sid)
+)
+no_0_t <- tableby(Diagnosis ~ kwt(Coughs, "medianq1q3", "range"), data = no_0)
+summary(no_0_t, pfootnote = T)
+```
+
+Spearman correlation between median coughs and pre-bronchodilator FEV1 was `r round(as.numeric(M$estimate),3)` (p = `r format.pval(M$p.value,eps = 0.001,digits = 3)`).
+
 # CFQ-R
@@
 
-cfqr_6_13 <- df %>%
-  filter(redcap_repeat_instrument == "CFQ-R Caregivers 6-13") %>%
+# Caregivers of children ages 6-13
+# Rename columns so we can use the code from the CFF
+cfqr_parent_6_13_names <- list(
+  "cfq_r_p_1" = "phys1", "cfq_r_p_2" = "phys2", "cfq_r_p_3" = "phys3",
+  "cfq_r_p_4" = "phys4", "cfq_r_p_5" = "phys5", "cfq_r_p_6" = "emot6",
+  "cfq_r_p_7" = "emot7", "cfq_r_p_8" = "vital8", "cfq_r_p_9" = "vital9",
+  "cfq_r_p_10" = "vital10", "cfq_r_p_11" = "vital11", "cfq_r_p_12" = "vital12",
+  "cfq_r_p_13" = "school13", "cfq_r_p_14" = "phys14", "cfq_r_p_15" = "phys15",
+  "cfq_r_p_16" = "phys16", "cfq_r_p_17" = "eat17", "cfq_r_p_18" = "treat18",
+  "cfq_r_p_19" = "body19", "cfq_r_p_20" = "body20", "cfq_r_p_21" = "body21",
+  "cfq_r_p_22" = "health22", "cfq_r_p_23" = "emot23", "cfq_r_p_24" = "health24",
+  "cfq_r_p_25" = "emot25", "cfq_r_p_26" = "emot26", "cfq_r_p_27" = "school27",
+  "cfq_r_p_28" = "school28", "cfq_r_p_29" = "school29",
+  "cfq_r_p_30" = "treat30", "cfq_r_p_31" = "treat31", "cfq_r_p_32" = "health32",
+  "cfq_r_p_33" = "weight33", "cfq_r_p_34" = "resp34", "cfq_r_p_35" = "resp35",
+  "cfq_r_p_36" = "resp36", "cfq_r_p_37" = "resp37", "cfq_r_p_38" = "resp38",
+  "cfq_r_p_39" = "resp39", "cfq_r_p_40" = "resp40", "cfq_r_p_41" = "digest41",
+  "cfq_r_p_42" = "digest42", "cfq_r_p_43" = "digest43", "cfq_r_p_44" = "eat44"
+)
+cfqr_parent_6_13_names <-
+  setNames(names(cfqr_parent_6_13_names), cfqr_parent_6_13_names)
+# Some items need to be reverse scored
+cfqr_parent_6_13_reverse <- c(
+  "emot6", "vital10", "vital12", "phys15", "treat31", "health22", "health24",
+  "health32", "school28", "resp37"
+)
+# Do it!
+cfqr_parent_6_13 <- df %>%
   select(
-    sid, Diagnosis, cfqr_caregivers_613_timestamp:cfqr_caregivers_613_complete
+    sid, redcap_event_name:redcap_repeat_instance,
+    cfq_r_p_1:cfq_r_p_44
+  ) %>%
+  mutate(across(cfq_r_p_1:cfq_r_p_44, ~ as.numeric(.x))) %>%
+  rename(all_of(unlist(cfqr_parent_6_13_names))) %>%
+  mutate(across(all_of(cfqr_parent_6_13_reverse), ~ 5 - .x)) %>%
+  rowwise() %>%
+  mutate(
+    physical =
+      ifelse(
+        sum(is.na(c(phys1, phys2, phys3, phys4, phys5, phys14, phys15, phys16))) <= 4,
+        (mean(c(phys1, phys2, phys3, phys4, phys5, phys14, phys15, phys16)) - 1)
+        / 3 * 100, NA
+      ),
+    emotion =
+      ifelse(
+        sum(is.na(c(emot6, emot7, emot23, emot25, emot26))) <= 2,
+        (mean(c(emot6, emot7, emot23, emot25, emot26)) - 1)
+        / 3 * 100, NA
+      ),
+    vitality =
+      ifelse(
+        sum(is.na(c(vital8, vital9, vital10, vital11, vital12))) <= 2,
+        (mean(c(vital8, vital9, vital10, vital11, vital12)) - 1)
+        / 3 * 100, NA
+      ),
+    school =
+      ifelse(
+        sum(is.na(c(school13, school27, school28, school29))) <= 2,
+        (mean(c(school13, school27, school28, school29)) - 1)
+        / 3 * 100, NA
+      ),
+    eat =
+      ifelse(
+        sum(is.na(c(eat17, eat44))) == 0,
+        (mean(c(eat17, eat44)) - 1)
+        / 3 * 100, NA
+      ),
+    body =
+      ifelse(
+        sum(is.na(c(body19, body20, body21))) <= 1,
+        (mean(c(body19, body20, body21)) - 1)
+        / 3 * 100, NA
+      ),
+    treat =
+      ifelse(
+        sum(is.na(c(treat18, treat30, treat31))) <= 1,
+        (mean(c(treat18, treat30, treat31)) - 1)
+        / 3 * 100, NA
+      ),
+    health =
+      ifelse(
+        sum(is.na(c(health22, health24, health32))) <= 1,
+        (mean(c(health22, health24, health32)) - 1)
+        / 3 * 100, NA
+      ),
+    respirat =
+      ifelse(
+        sum(is.na(c(resp34, resp35, resp36, resp38, resp39, resp40))) <= 3,
+        (mean(c(resp34, resp35, resp36, resp38, resp39, resp40)) - 1)
+        / 3 * 100, NA
+      ),
+    digest =
+      ifelse(
+        sum(is.na(c(digest41, digest42, digest43))) <= 1,
+        (mean(c(digest41, digest42, digest43)) - 1)
+        / 3 * 100, NA
+      ),
+    weight =
+      ifelse(
+        sum(is.na(c(weight33))) == 0,
+        (mean(c(weight33)) - 1)
+        / 3 * 100, NA
       )
+  )
+check <- left_join(
+  df %>%
+    select(
+      sid, redcap_event_name:redcap_repeat_instance,
+      cfq_r_p_1:cfq_r_p_44
+    ),
+  cfqr_parent_6_13 %>%
+    select(sid, redcap_event_name:redcap_repeat_instance, physical:weight),
+  join_by(
+    sid, redcap_event_name, redcap_repeat_instrument,
+    redcap_repeat_instance
+  )
+)
+write.csv(check, file = "./Data_Clean/cfqr_parent_6_13_check.csv", na = "", row.names = F)
 
-cfqr_6_13$cfqr_physical <- apply(
-  data.matrix(cfqr_6_13[, paste0("cfq_r_p_", 1:5)]), 1, function(r) {
-    (mean(r) - 1) / 3 * 100
-  }
+cfqr_children_6_11 <- list(
+  "cfq_r_y_1" = "phys1", "cfq_r_y_2" = "phys2", "cfq_r_y_3" = "phys3",
+  "cfq_r_y_4" = "phys4", "cfq_r_y_5" = "phys5", "cfq_r_y_6" = "phys6",
+  "cfq_r_y_7" = "emot7", "cfq_r_y_8" = "emot8", "cfq_r_y_9" = "emot9",
+  "cfq_r_y_10" = "emot10", "cfq_r_y_11" = "emot11", "cfq_r_y_12" = "emot12",
+  "cfq_r_y_13" = "emot13", "cfq_r_y_14" = "emot14", "cfq_r_y_15" = "eat15",
+  "cfq_r_y_16" = "treat16", "cfq_r_y_17" = "eat17", "cfq_r_y_18" = "treat18",
+  "cfq_r_y_19" = "eat19", "cfq_r_y_20" = "social20", "cfq_r_y_21" = "social21",
+  "cfq_r_y_22" = "social22", "cfq_r_y_23" = "social23",
+  "cfq_r_y_24" = "social24", "cfq_r_y_25" = "social25",
+  "cfq_r_y_26" = "social26", "cfq_r_y_27" = "body27", "cfq_r_y_28" = "body28",
+  "cfq_r_y_29" = "body29", "cfq_r_y_30" = "treat30", "cfq_r_y_31" = "resp31",
+  "cfq_r_y_32" = "resp32", "cfq_r_y_33" = "resp33", "cfq_r_y_34" = "resp34",
+  "cfq_r_y_35" = "digest35 "
 )
-cfqr_6_13$cfqr_emotion <- apply(
-  data.matrix(cfqr_6_13[, paste0("cfq_r_p_", 6:13)]), 1, function(r) {
-    (mean(r) - 1) / 3 * 100
-  }
+cfqr_children_6_11 <-
+  setNames(names(cfqr_children_6_11), cfqr_children_6_11)
+cfqr_teens_adults <- list(
+  "cfq_r_a_1" = "phys1", "cfq_r_a_2" = "phys2", "cfq_r_a_3" = "phys3",
+  "cfq_r_a_4" = "phys4", "cfq_r_a_5" = "phys5", "cfq_r_a_6" = "vital6",
+  "cfq_r_a_7" = "emot7", "cfq_r_a_8" = "emot8", "cfq_r_a_9" = "vital9",
+  "cfq_r_a_10" = "vital10", "cfq_r_a_11" = "vital11", "cfq_r_a_12" = "emot12",
+  "cfq_r_a_13" = "phys13", "cfq_r_a_14" = "eat14", "cfq_r_a_15" = "treat15",
+  "cfq_r_a_16" = "treat16", "cfq_r_a_17" = "treat17", "cfq_r_a_18" = "health18",
+  "cfq_r_a_19" = "phys19", "cfq_r_a_20" = "phys20", "cfq_r_a_21" = "eat21",
+  "cfq_r_a_22" = "social22", "cfq_r_a_23" = "social23", "cfq_r_a_24" = "body24",
+  "cfq_r_a_25" = "body25", "cfq_r_a_26" = "body26", "cfq_r_a_27" = "social27",
+  "cfq_r_a_28" = "social28", "cfq_r_a_29" = "social29",
+  "cfq_r_a_30" = "social30", "cfq_r_a_31" = "emot31", "cfq_r_a_32" = "health32",
+  "cfq_r_a_33" = "emot33", "cfq_r_a_34" = "health34", "cfq_r_a_35" = "role35",
+  "cfq_r_a_36" = "role36", "cfq_r_a_37" = "role37", "cfq_r_a_38" = "role38",
+  "cfq_r_a_39" = "weight39", "cfq_r_a_40" = "resp40", "cfq_r_a_41" = "resp41",
+  "cfq_r_a_42" = "resp42", "cfq_r_a_43" = "resp43", "cfq_r_a_44" = "resp44",
+  "cfq_r_a_45" = "resp45", "cfq_r_a_46" = "resp46", "cfq_r_a_47" = "digest47",
+  "cfq_r_a_48" = "digest48", "cfq_r_a_49" = "digest49", "cfq_r_a_50" = "eat50"
 )
-
+cfqr_teens_adults <-
+  setNames(names(cfqr_teens_adults), cfqr_teens_adults)
 ```

### 2025-02-06 Lilah Melzer/cough_monitor_analysis.qmd
need to figure out missing CFQ-R dates

@@
 library(corrplot)
+library(patchwork)
 knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor")
@@
 ```{r}
-# CFQ-R Caregivers 6-13
-# SAS code
-# if nmiss (phys1, phys2, phys3, phys4, phys5, phys6) <= 3 then
-# physical = (mean (phys1, phys2, phys3, phys4, phys5, phys6)-1)/3*100;
-#
-# if nmiss (emot7, emot8, emot9, emot10, emot11, emot12, emot13, emot14) <= 4 then
-# emotion = (mean (emot7, emot8, emot9, emot10, emot11, emot12, emot13, emot14)-1)/3*100;
-#
-# if nmiss (social20, social21, social22, social23, social24, social25, social26) <= 3 then
-# social = (mean (social20, social21, social22, social23, social24, social25, social26)-1)/3*100;
-#
-# if nmiss (eat15, eat17, eat19) <= 1 then
-# eat = (mean(eat15, eat17, eat19)-1)/3*100;
-#
-# if nmiss (body27, body28, body29) <= 1 then
-# body = (mean(body27, body28, body29)-1)/3*100;
-#
-# if nmiss (treat16, treat18, treat30) <= 1 then
-# treat = (mean (treat16, treat18, treat30)-1)/3*100;
-#
-# if nmiss (resp31, resp32, resp33, resp34) <= 2 then
-# respirat = (mean (resp31, resp32, resp33, resp34)-1)/3*100;
-#
-# if nmiss (digest35) = 0 then
-# digest = (mean (digest35)-1)/3*100;
+# See SAS scoring instructions (from Quitter's CFQ-R website).docx for
+# instructions
 
@@
 )
-# Do it!
+# Do all the calculations!
 cfqr_parent_6_13 <- df %>%
   select(
-    sid, redcap_event_name:redcap_repeat_instance,
-    cfq_r_p_1:cfq_r_p_44
+    sid, cfqr_caregivers_613_timestamp, cfq_r_p_1:cfq_r_p_44
   ) %>%
-  mutate(across(cfq_r_p_1:cfq_r_p_44, ~ as.numeric(.x))) %>%
+  mutate(
+    Date = as.Date(cfqr_caregivers_613_timestamp),
+    across(cfq_r_p_1:cfq_r_p_44, ~ as.numeric(.x))
+  ) %>%
   rename(all_of(unlist(cfqr_parent_6_13_names))) %>%
@@
       )
-  )
-check <- left_join(
-  df %>%
-    select(
-      sid, redcap_event_name:redcap_repeat_instance,
-      cfq_r_p_1:cfq_r_p_44
-    ),
-  cfqr_parent_6_13 %>%
-    select(sid, redcap_event_name:redcap_repeat_instance, physical:weight),
-  join_by(
-    sid, redcap_event_name, redcap_repeat_instrument,
-    redcap_repeat_instance
-  )
-)
-write.csv(check, file = "./Data_Clean/cfqr_parent_6_13_check.csv", na = "", row.names = F)
-
-cfqr_children_6_11 <- list(
+  ) %>%
+  select(sid, Date, physical:weight)
+# Children aged 6 to 11
+cfqr_children_6_11_names <- list(
   "cfq_r_y_1" = "phys1", "cfq_r_y_2" = "phys2", "cfq_r_y_3" = "phys3",
@@
   "cfq_r_y_32" = "resp32", "cfq_r_y_33" = "resp33", "cfq_r_y_34" = "resp34",
-  "cfq_r_y_35" = "digest35 "
+  "cfq_r_y_35" = "digest35"
 )
-cfqr_children_6_11 <-
-  setNames(names(cfqr_children_6_11), cfqr_children_6_11)
-cfqr_teens_adults <- list(
+cfqr_children_6_11_names <-
+  setNames(names(cfqr_children_6_11_names), cfqr_children_6_11_names)
+# Some items need to be reverse scored
+cfqr_children_6_11_reverse <- c(
+  "phys1", "phys2", "phys3", "phys4", "phys5", "emot14", "treat18", "eat19",
+  "social20", "social22", "social24", "social26"
+)
+# Do all the calculations
+cfqr_children_6_11 <- df %>%
+  select(
+    sid, cfqr_children_ages_611_timestamp, cfq_r_y_1:cfq_r_y_35
+  ) %>%
+  mutate(
+    Date = as.Date(cfqr_children_ages_611_timestamp),
+    across(cfq_r_y_1:cfq_r_y_35, ~ as.numeric(.x))
+  ) %>%
+  rename(all_of(unlist(cfqr_children_6_11_names))) %>%
+  mutate(across(all_of(cfqr_children_6_11_reverse), ~ 5 - .x)) %>%
+  rowwise() %>%
+  mutate(
+    physical =
+      ifelse(
+        sum(is.na(c(phys1, phys2, phys3, phys4, phys5, phys6))) <= 3,
+        (mean(c(phys1, phys2, phys3, phys4, phys5, phys6)) - 1)
+        / 3 * 100, NA
+      ),
+    emotion =
+      ifelse(
+        sum(is.na(c(emot7, emot8, emot9, emot10, emot11, emot12, emot13, emot14))) <= 4,
+        (mean(c(emot7, emot8, emot9, emot10, emot11, emot12, emot13, emot14)) - 1)
+        / 3 * 100, NA
+      ),
+    social =
+      ifelse(
+        sum(is.na(c(social20, social21, social22, social23, social24, social25, social26))) <= 3,
+        (mean(c(social20, social21, social22, social23, social24, social25, social26)) - 1)
+        / 3 * 100, NA
+      ),
+    eat =
+      ifelse(
+        sum(is.na(c(eat15, eat17, eat19))) <= 1,
+        (mean(c(eat15, eat17, eat19)) - 1)
+        / 3 * 100, NA
+      ),
+    body =
+      ifelse(
+        sum(is.na(c(body27, body28, body29))) <= 1,
+        (mean(c(body27, body28, body29)) - 1)
+        / 3 * 100, NA
+      ),
+    treat =
+      ifelse(
+        sum(is.na(c(treat16, treat18, treat30))) <= 1,
+        (mean(c(treat16, treat18, treat30)) - 1)
+        / 3 * 100, NA
+      ),
+    respirat =
+      ifelse(
+        sum(is.na(c(resp31, resp32, resp33, resp34))) <= 2,
+        (mean(c(resp31, resp32, resp33, resp34)) - 1)
+        / 3 * 100, NA
+      ),
+    digest =
+      ifelse(
+        sum(is.na(c(digest35))) == 0,
+        (mean(c(digest35)) - 1)
+        / 3 * 100, NA
+      )
+  ) %>%
+  select(sid, Date, physical:digest)
+# Teen/Adult version
+cfqr_teens_adults_names <- list(
   "cfq_r_a_1" = "phys1", "cfq_r_a_2" = "phys2", "cfq_r_a_3" = "phys3",
@@
 )
-cfqr_teens_adults <-
-  setNames(names(cfqr_teens_adults), cfqr_teens_adults)
+cfqr_teens_adults_names <-
+  setNames(names(cfqr_teens_adults_names), cfqr_teens_adults_names)
+# Some items need to be reverse scored
+cfqr_teens_adults_reverse <- c(
+  "vital6", "vital10", "phys13", "treat15", "treat17", "health18", "social23",
+  "social28", "social30", "health32", "health34", "role35", "resp43"
+)
+# Do all the calculations
+cfqr_teens_adults <- df %>%
+  select(
+    sid, cfqr_adolescents_and_adults_14_years_timestamp, cfq_r_a_1:cfq_r_a_50
+  ) %>%
+  mutate(
+    Date = as.Date(cfqr_adolescents_and_adults_14_years_timestamp),
+    across(cfq_r_a_1:cfq_r_a_50, ~ as.numeric(.x))
+  ) %>%
+  rename(all_of(unlist(cfqr_teens_adults_names))) %>%
+  mutate(across(all_of(cfqr_teens_adults_reverse), ~ 5 - .x)) %>%
+  rowwise() %>%
+  mutate(
+    physical =
+      ifelse(
+        sum(is.na(c(phys1, phys2, phys3, phys4, phys5, phys13, phys19, phys20))) <= 4,
+        (mean(c(phys1, phys2, phys3, phys4, phys5, phys13, phys19, phys20)) - 1)
+        / 3 * 100, NA
+      ),
+    role =
+      ifelse(
+        sum(is.na(c(role35, role36, role37, role38))) <= 2,
+        (mean(c(role35, role36, role37, role38)) - 1)
+        / 3 * 100, NA
+      ),
+    vitality =
+      ifelse(
+        sum(is.na(c(vital6, vital9, vital10, vital11))) <= 2,
+        (mean(c(vital6, vital9, vital10, vital11)) - 1)
+        / 3 * 100, NA
+      ),
+    emotion =
+      ifelse(
+        sum(is.na(c(emot7, emot8, emot12, emot31, emot33))) <= 2,
+        (mean(c(emot7, emot8, emot12, emot31, emot33)) - 1)
+        / 3 * 100, NA
+      ),
+    social =
+      ifelse(
+        sum(is.na(c(social22, social23, social27, social28, social29, social30))) <= 3,
+        (mean(c(social22, social23, social27, social28, social29, social30)) - 1)
+        / 3 * 100, NA
+      ),
+    body =
+      ifelse(
+        sum(is.na(c(body24, body25, body26))) <= 1,
+        (mean(c(body24, body25, body26)) - 1)
+        / 3 * 100, NA
+      ),
+    eat =
+      ifelse(
+        sum(is.na(c(eat14, eat21, eat50))) <= 1,
+        (mean(c(eat14, eat21, eat50)) - 1)
+        / 3 * 100, NA
+      ),
+    treat =
+      ifelse(
+        sum(is.na(c(treat15, treat16, treat17))) <= 1,
+        (mean(c(treat15, treat16, treat17)) - 1)
+        / 3 * 100, NA
+      ),
+    health =
+      ifelse(
+        sum(is.na(c(health18, health32, health34))) <= 1,
+        (mean(c(health18, health32, health34)) - 1)
+        / 3 * 100, NA
+      ),
+    weight =
+      ifelse(
+        sum(is.na(c(weight39))) == 0,
+        (mean(c(weight39)) - 1)
+        / 3 * 100, NA
+      ),
+    respirat =
+      ifelse(
+        sum(is.na(c(resp40, resp41, resp42, resp44, resp45, resp46))) <= 3,
+        (mean(c(resp40, resp41, resp42, resp44, resp45, resp46)) - 1)
+        / 3 * 100, NA
+      ),
+    digest =
+      ifelse(
+        sum(is.na(c(digest47, digest48, digest49))) <= 1,
+        (mean(c(digest47, digest48, digest49)) - 1)
+        / 3 * 100, NA
+      )
+  ) %>%
+  select(sid, Date, physical:digest)
+# Write for checking
+write.csv(df %>%
+  select(
+    sid, cfqr_caregivers_613_timestamp, cfq_r_p_1:cfq_r_p_44
+  ), file = "~/cfqr_parent_6_13.csv", na = "", row.names = F)
+write.csv(df %>%
+  select(
+    sid, cfqr_children_ages_611_timestamp, cfq_r_y_1:cfq_r_y_35
+  ), file = "~/cfqr_children_6_11.csv", na = "", row.names = F)
+write.csv(df %>%
+  select(
+    sid, cfqr_adolescents_and_adults_14_years_timestamp, cfq_r_a_1:cfq_r_a_50
+  ), file = "~/cfqr_teens_adults.csv", na = "", row.names = F)
+# Add to counts for plotting
+cfqr_parent_6_13$sid <- as.numeric(cfqr_parent_6_13$sid)
+cfqr_children_6_11$sid <- as.numeric(cfqr_children_6_11$sid)
+cfqr_teens_adults$sid <- as.numeric(cfqr_teens_adults$sid)
+agg_counts <- left_join(agg_counts, cfqr_parent_6_13)
+agg_counts <- left_join(agg_counts, cfqr_children_6_11)
+agg_counts <- left_join(agg_counts, cfqr_teens_adults)
 ```
+
+## Respiratory scores
+
+### Plots by participant
+
+```{r}
+#| message: false
+#| warning: false
+invisible(lapply(unique(agg_counts$sid), function(s) {
+  d <- agg_counts[agg_counts$sid == s, ]
+  diag <- d$Diagnosis[1]
+  p <- ggplot(d, aes(
+    x = Night, y = Coughs, group = factor(sid), color = Sick
+  )) +
+    geom_point() +
+    geom_line() +
+    theme_bw() +
+    ylim(0, 50) +
+    xlim(0, 150) +
+    scale_color_manual(values = c("#377EB8", "#E41A1C")) +
+    ggtitle(paste0("Participant ", s, " (", diag, ")"))
+  if (s == 38) {
+    p <- p + ylim(0, 270)
+  }
+  p2 <- ggplot(d[!is.na(d$respirat), ], aes(
+    x = Night, y = respirat, group = factor(sid), color = Sick
+  )) +
+    geom_point() +
+    geom_line() +
+    theme_bw() +
+    ylim(50, 100) +
+    xlim(0, 150) +
+    scale_color_manual(values = c("#377EB8", "#E41A1C")) +
+    ggtitle(paste0("Participant ", s, " (", diag, ")"))
+  cat("\n")
+  cat("\n")
+  print(p + p2 + plot_layout(guides = "collect"))
+  cat("\n")
+  cat("\n")
+}))
+```

### 2025-02-06 Angela Metcalf/quantifying_cl_cf_carriers.qmd
Create quantifying_cl_cf_carriers.qmd

---
title: "Quantifying Sweat Chloride in CF Carriers"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/timvigers/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Angela Metcalf/Quantifying Sweat Chloride in CF Carriers")
```

```{r data cleaning}
#| message: false
df <- read.csv("./Data_Raw/Midpoint data.csv", na.strings = "")
df <- df %>%
  mutate(
    `Right Arm Cl` = as.numeric(Sweat.chloride..right.arm),
    `Left Arm Cl` = as.numeric(Sweat.chloride..left.arm),
    `Right Arm Na` = as.numeric(Sweat.sodium..right.arm),
    `Left Arm Na` = as.numeric(Sweat.sodium..left.arm)
  )
df$`Mean Cl` <- rowMeans(df[, c("Right Arm Cl", "Left Arm Cl")])
df$`Mean Na` <- rowMeans(df[, c("Right Arm Na", "Left Arm Na")])
```

# Plots

```{r}
ggplot(df,aes(y=`Mean Cl`))+boxplot()
```

### 2025-02-07 Jordana Hoppe/Home Collection/manuscript.qmd
plot updates finished for Jordana

@@
 library(patchwork)
+library(ggpubr)
 home_dir <- switch(Sys.info()["sysname"],
@@
     as.data.frame(.) %>%
+    arrange(desc(Freq)) %>%
     kable(.) %>%
@@
     theme(legend.title = element_blank(), axis.title.x = element_blank())
-  p <- bar_plot + venn + plot_annotation(tag_levels = c("a", "b"))
+  p <- ggarrange(venn, bar_plot, nrow = 1, labels = "auto") + bgcolor("White")
   cat("\n")
@@
       theme(legend.title = element_blank(), axis.title.x = element_blank())
-    p <- bar_plot + venn + plot_annotation(tag_levels = c("a", "b"))
+    p <- ggarrange(venn, bar_plot, nrow = 1, labels = "auto") + bgcolor("White")
     ggsave("Jordana Hoppe/Home Collection/Dissemination/Figures/mssa_2.png",
@@
 # Label
-labels <- as.character(label(survey_data)[sub("\\.factor", "", colnames(t))])
-names(labels) <- colnames(t)
+labels <- as.character(label(survey_data)[sub("\\.factor", "", colnames(survey_data))])
+names(labels) <- colnames(survey_data)
 ```
@@
 ```{r}
-yn_df <- t %>%
+yn_df <- survey_data %>%
   select(all_of(yn)) %>%
@@
 ```{r}
-check_df <- t %>%
+check_df <- survey_data %>%
   select(all_of(checkboxes)) %>%
@@
 )]
+benefits_df <- benefits_df %>%
+  pivot_longer(everything()) %>%
+  group_by(name) %>%
+  summarise(prop = mean(value == "Checked")) %>%
+  arrange(desc(prop))
+p_benefits <- ggplot(benefits_df, aes(y = reorder(name, prop), x = prop)) +
+  geom_bar(stat = "identity", fill = "#5AB4AC") +
+  geom_text(aes(label = scales::percent(prop, accuracy = 1)), position = position_stack(vjust = 0.5)) +
+  ggtitle("Benefits") +
+  scale_x_continuous(
+    breaks = c(0, 0.25, 0.5, 0.75, 1),
+    labels = scales::percent(c(0, 0.25, 0.5, 0.75, 1))
+  ) +
+  theme_bw() +
+  theme(axis.title.y = element_blank(), axis.title.x = element_blank())
 barriers_df <- check_df[, c(
@@
 )]
-
-p_benefits <- gglikert(benefits_df, add_totals = F) +
-  ggtitle("Benefits") +
-  theme(
-    axis.text.x = element_blank(), axis.title.x = element_blank(),
-    legend.position = "none"
+barriers_df <- barriers_df %>%
+  pivot_longer(everything()) %>%
+  group_by(name) %>%
+  summarise(prop = mean(value == "Checked")) %>%
+  arrange(desc(prop))
+p_barriers <- ggplot(barriers_df, aes(y = reorder(name, prop), x = prop)) +
+  geom_bar(stat = "identity", fill = "#D8B365") +
+  geom_text(aes(label = scales::percent(prop, accuracy = 1)), position = position_stack(vjust = 0.5)) +
+  ggtitle("Barriers") +
+  scale_x_continuous(
+    breaks = c(0, 0.05, 0.1, 0.15),
+    labels = scales::percent(c(0, 0.05, 0.1, 0.15))
   ) +
-  scale_fill_manual(values = c("Unchecked" = "#D8B365", "Checked" = "#5AB4AC"))
-p_barriers <- gglikert(barriers_df, reverse_likert = T, add_totals = F) +
-  ggtitle("Barriers") +
-  scale_fill_manual(values = c("Unchecked" = "#5AB4AC", "Checked" = "#D8B365"))
+  theme_bw() +
+  theme(axis.title.y = element_blank(), axis.title.x = element_blank())
 p <- p_benefits / p_barriers
@@
 ```{r}
-df3 <- t %>%
+df3 <- survey_data %>%
   select(all_of(levels3)) %>%
   rename(any_of(setNames(names(labels), labels)))
-p3 <- gglikert(as_tibble(df3))
+p3 <- gglikert(as_tibble(df3), add_totals = F)
 p3
@@
 ```{r}
-df5 <- t %>%
+df5 <- survey_data %>%
   select(all_of(levels5)) %>%
@@
 )
-p <- p3 / p5 + plot_annotation(tag_levels = c("a", "b"))
+p <- p3 / p5 + plot_annotation(tag_level = "a")
 p

### 2025-02-07 Christine Chan/CF Registry/cf_registry_CFRD_prediction.qmd
CFRD prediciton models

@@
 library(arsenal)
+library(childsds)
 library(survival)
@@
 )
+# Convert BMI to percentile (assume age > 20 is 20)
+annual$bmi_perc <- sds(annual$A_bmivalue,
+  age = ifelse(annual$Age_YrEnd > 20, 20, annual$Age_YrEnd),
+  sex = annual$Gender, item = "bmi", type = "perc", male = "M", female = "F",
+  ref = cdc.ref
+) * 100
 # Select columns
@@
     eDWID, ReviewYear, Race, Hispanicrace, Gender, F508, MutClass,
-    A_Vx770, A_VX809comb, Age_YrEnd, A_bmipercentile, A_hgba1c,
-    cfrd_registry, cfrd_status_annual, cfrd_ogtt, A_twohour_bloodglucose
+    A_Vx770, A_VX809comb, Age_YrEnd, A_bmivalue, A_bmipercentile, bmi_perc,
+    A_FEV1_pct_predicted, NumPulmExacerbation, A_IsOnEnzymes,
+    A_fast_bloodglucose, A_hgba1c, cfrd_registry, cfrd_status_annual, cfrd_ogtt,
+    A_twohour_bloodglucose
   ) %>%
@@
 ogtt$Age_YrStart <- ogtt$Age_YrEnd - 1
+# Format columns
+annual$MutClass <- factor(annual$MutClass,
+  levels = c("1-3", "4-5", "Oth", "N/A"),
+  labels = c("1-3", "4-5", "Oth", "Unknown")
+)
+annual$NumPulmExacerbation <- cut(annual$NumPulmExacerbation,
+  labels = c("0", "1", "2", ">=3"),
+  breaks = c(0, 1, 2, 3, Inf), right = F
+)
+# Labels
+labels <- list(
+  Age_YrStart = "Age at Year Start", Age_YrEnd = "Age at Year End",
+  Hispanicrace = "Hispanic", A_Vx770 = "Use of CFTR modulator Vx770",
+  A_VX809comb = "Use of CFTR modulator Vx809 combination",
+  A_FEV1_pct_predicted = "FEV1 percent predicted", MutClass = "Mutation class",
+  NumPulmExacerbation = "Admissions for PEx"
+)
+# Basic formula
+xvars <- c(
+  "Gender", "Race", "Hispanicrace", "MutClass", "A_IsOnEnzymes",
+  "A_FEV1_pct_predicted", "NumPulmExacerbation", "A_corticosteroids1"
+)
+f <- ~ A_hgba1c + A_fast_bloodglucose + Gender + Race + Hispanicrace + MutClass +
+  A_IsOnEnzymes + A_FEV1_pct_predicted + NumPulmExacerbation
 ```
@@
 cph <- coxph(
-  Surv(Age_YrStart, Age_YrEnd, as.numeric(cfrd_registry)) ~
-    A_hgba1c + Gender + F508 + A_Vx770 + A_VX809comb,
+  update(f, Surv(time = Age_YrStart, time2 = Age_YrEnd, event = as.numeric(cfrd_registry)==2) ~ .),
   data = registry, cluster = eDWID
@@
 ggcoef_table(cph,
-  show_p_values = F, add_reference_rows = T, signif_stars = F, colour = F
+  exponentiate = T, show_p_values = F, add_reference_rows = T, signif_stars = F,
+  colour = F
 )
+
+d = model.frame(cph)
+d$pred <- predict(cph, type = "survival")
 ```
@@
 cph <- coxph(
-  Surv(Age_YrStart, Age_YrEnd, as.numeric(cfrd_ogtt)) ~
-    A_hgba1c + Gender + F508 + A_Vx770 + A_VX809comb,
+  update(f, Surv(time = Age_YrStart, time2 = Age_YrEnd, event = as.numeric(cfrd_ogtt)) ~ .),
   data = ogtt, cluster = eDWID
@@
 ggcoef_table(cph,
-  show_p_values = F, add_reference_rows = T, signif_stars = F, colour = F
+  exponentiate = T, show_p_values = F, add_reference_rows = T, signif_stars = F,
+  colour = F
 )

### 2025-02-07 Jordana Hoppe/Home Collection/manuscript.qmd
home manuscript final figures

@@
   d[d == "Checked"] <- "Present"
+  d[d == "No Growth"] <- "Absent"
+  d[d == "Some Growth"] <- "Present"
   fk <- kappam.fleiss(d)
@@
   }
+  # Save no growth Venn diagram
+  if (result_pattern == "no_growth") {
+    ggsave("Jordana Hoppe/Home Collection/Dissemination/Figures/no_growth.png",
+      plot = venn,
+    )
+  }
+  cat("\n")
   cat("### Pairwise agreement")
@@
   rename(any_of(setNames(names(labels), labels)))
-p3 <- gglikert(as_tibble(df3), add_totals = F)
+p3 <- gglikert(as_tibble(df3), add_totals = F) +
+  scale_fill_manual(values = c("#5AB4AC", "#F5F5F5", "#D8B365"))
 p3
@@
   rename(any_of(setNames(names(labels), labels)))
-p5 <- gglikert(df5)
+p5 <- gglikert(df5) +
+  scale_fill_manual(values = c("#018571", "#80CDC1", "#F5F5F5", "#DFC27D", "#A6611A"))
 p5

### 2025-02-07 Quarto Template.qmd
Update Quarto Template.qmd

@@
 bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
-csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
+csl: /Users/timvigers/Documents/GitHub/styles/american-medical-association.csl
 editor: source
@@
 )
-exportBulkRecords(list(db = rcon), envir = 1)
+df <- exportRecordsTyped(rcon)
 ```

### 2025-02-11 Angela Metcalf/quantifying_cl_cf_carriers.qmd
CF carrier interim power

@@
 #| include: false
+library(redcapAPI)
 library(tidyverse)
+library(arsenal)
 github_dir <- switch(Sys.info()["sysname"],
@@
 #| message: false
-df <- read.csv("./Data_Raw/Midpoint data.csv", na.strings = "")
+# Import from REDCap
+unlockREDCap(c(rcon = "Sweat Electrolytes in CF Carriers 24-0691"),
+  keyring = "API_KEYs",
+  envir = 1,
+  url = "https://redcap.ucdenver.edu/api/"
+)
+df <- exportRecordsTyped(rcon)
+# Format
 df <- df %>%
   mutate(
-    `Right Arm Cl` = as.numeric(Sweat.chloride..right.arm),
-    `Left Arm Cl` = as.numeric(Sweat.chloride..left.arm),
-    `Right Arm Na` = as.numeric(Sweat.sodium..right.arm),
-    `Left Arm Na` = as.numeric(Sweat.sodium..left.arm)
+    `Cl Right` = as.numeric(cl_right), `Cl Left` = as.numeric(cl_left),
+    `Na Right` = as.numeric(na_right), `Na Left` = as.numeric(na_left)
   )
-df$`Mean Cl` <- rowMeans(df[, c("Right Arm Cl", "Left Arm Cl")])
-df$`Mean Na` <- rowMeans(df[, c("Right Arm Na", "Left Arm Na")])
+df$`Mean Cl` <- rowMeans(df[, c("Cl Right", "Cl Left")])
+df$`Mean Na` <- rowMeans(df[, c("Na Right", "Na Left")])
+```
+
+# Table 1
+
+```{r}
+t1 = tableby(carrier~`Mean Cl` +`Mean Na`,data = df )
+summary(t1)
 ```
@@
 ```{r}
-ggplot(df,aes(y=`Mean Cl`))+boxplot()
+ggplot(df, aes(x = carrier, y = `Mean Cl`)) +
+  geom_boxplot()
 ```

### 2025-02-11 Lilah Melzer/cough_monitor_analysis.qmd
cough monitor surveys at first visit

@@
 ```{r}
+#| message: false
 # See SAS scoring instructions (from Quitter's CFQ-R website).docx for
 # instructions
-
+df <- df %>%
+  group_by(sid) %>%
+  fill(c(date_v1_cf, date_v1_hc)) %>%
+  ungroup()
 # Caregivers of children ages 6-13
@@
   select(
-    sid, cfqr_caregivers_613_timestamp, cfq_r_p_1:cfq_r_p_44
+    sid, date_v1_cf, date_v1_hc,
+    cfqr_caregivers_613_timestamp, cfq_r_p_1:cfq_r_p_44
   ) %>%
@@
   ) %>%
-  select(sid, Date, physical:weight)
+  mutate(
+    Date = coalesce(Date, date_v1_cf),
+    Date = coalesce(Date, date_v1_hc),
+    Survey = "CFQ-R Caregivers 6-13"
+  ) %>%
+  select(
+    sid, Survey, Date, physical:weight
+  ) %>%
+  filter(!if_all(physical:weight, ~ is.na(.)))
 # Children aged 6 to 11
@@
   select(
-    sid, cfqr_children_ages_611_timestamp, cfq_r_y_1:cfq_r_y_35
+    sid, date_v1_cf, date_v1_hc,
+    cfqr_children_ages_611_timestamp, cfq_r_y_1:cfq_r_y_35
   ) %>%
@@
   ) %>%
-  select(sid, Date, physical:digest)
+  mutate(
+    Date = coalesce(Date, date_v1_cf),
+    Date = coalesce(Date, date_v1_hc),
+    Survey = "CFQ-R Children Ages 6-11"
+  ) %>%
+  select(
+    sid, Survey, Date, physical:digest
+  ) %>%
+  filter(!if_all(physical:digest, ~ is.na(.)))
 # Teen/Adult version
@@
   select(
-    sid, cfqr_adolescents_and_adults_14_years_timestamp, cfq_r_a_1:cfq_r_a_50
+    sid, date_v1_cf, date_v1_hc,
+    cfqr_adolescents_and_adults_14_years_timestamp, cfq_r_a_1:cfq_r_a_50
   ) %>%
@@
   ) %>%
-  select(sid, Date, physical:digest)
-# Write for checking
-write.csv(df %>%
+  mutate(
+    Date = coalesce(Date, date_v1_cf),
+    Date = coalesce(Date, date_v1_hc),
+    Survey = "CFQ-R Adolescents and Adults 14 Years"
+  ) %>%
   select(
-    sid, cfqr_caregivers_613_timestamp, cfq_r_p_1:cfq_r_p_44
-  ), file = "~/cfqr_parent_6_13.csv", na = "", row.names = F)
-write.csv(df %>%
-  select(
-    sid, cfqr_children_ages_611_timestamp, cfq_r_y_1:cfq_r_y_35
-  ), file = "~/cfqr_children_6_11.csv", na = "", row.names = F)
-write.csv(df %>%
-  select(
-    sid, cfqr_adolescents_and_adults_14_years_timestamp, cfq_r_a_1:cfq_r_a_50
-  ), file = "~/cfqr_teens_adults.csv", na = "", row.names = F)
+    sid, Survey, Date, physical:digest
+  ) %>%
+  filter(!if_all(physical:digest, ~ is.na(.)))
+# Combine
+surveys <- full_join(cfqr_parent_6_13, cfqr_children_6_11)
+surveys <- full_join(surveys, cfqr_teens_adults)
 # Add to counts for plotting
@@
 ```
+
+## Correlation between baseline surveys and coughs
+
+```{r}
+```

### 2025-02-12 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
clean up SSwAB-IT tables

@@
 df$who_collected_cos <- droplevels(df$who_collected_cos)
+# Shorter group labels for cleaner tables
+levels(df$group) <- c("COS/SOS/IP", "COS/SOS/DP", "SOS/COS/IP", "SOS/COS/DP")
+levels(db_demographics$group) <-
+  c("COS/SOS/IP", "COS/SOS/DP", "SOS/COS/IP", "SOS/COS/DP")
+levels(db_medications$group) <-
+  c("COS/SOS/IP", "COS/SOS/DP", "SOS/COS/IP", "SOS/COS/DP")
 ```
@@
   separate_p_footnotes() %>%
+  modify_footnote(everything() ~ NA) %>%
   as_gt()
@@
   d$detailed_agree <- paste0("COS ", d$COS, "/", "SOS ", d$SOS)
+  d$detailed_agree <- factor(d$detailed_agree,
+    levels = c(
+      "COS Negative/SOS Negative", "COS Negative/SOS Positive",
+      "COS Positive/SOS Negative", "COS Positive/SOS Positive"
+    ),
+    labels = c("COS-/SOS-", "COS-/SOS+", "COS+/SOS-", "COS+/SOS+")
+  )
   # Print
@@
     tbl_summary(by = Group, missing_text = "Missing", label = list(
-      agree ~ "Agree (Both Positive/Positive and Negative/Negative)",
+      agree ~ "Agree (+/+ and -/-)",
       detailed_agree ~ "Detailed Agreement"
@@
     separate_p_footnotes() %>%
+    modify_footnote(everything() ~ NA) %>%
     print()

### 2025-02-13 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
Update phase_1_analysis.qmd

@@
 library(eulerr)
+library(ggpubr)
 ```
@@
 ```
+
+# Manuscript Figures
+
+```{r}
+# All pathogens
+d <- df[, c("overall_cos", "overall_sos")]
+colnames(d) <- c("COS", "SOS")
+d <- d[complete.cases(d), ]
+a1 <- plot(euler(d == "Positive"),
+  quantities = list(cex = 3),
+  fills = list(fill = c("#FFC107", "#1E88E5", "#004D40"))
+)
+a2 <- ggplot(as.data.frame(table(d$COS, d$SOS)), aes(Var1, Var2)) +
+  geom_tile(color = "white", fill = "#1E88E5") +
+  geom_text(aes(label = paste0(Freq, "\n(", round(Freq / sum(Freq), 3) * 100, "%)")),
+    color = "white"
+  ) +
+  xlab("COS") +
+  ylab("SOS") +
+  theme_bw() +
+  theme(text = element_text(size = 15))
+# MRSA
+d <- df[, c("staphylococcus_aureus_cos", "staphylococcus_aureus_sos")]
+colnames(d) <- c("COS", "SOS")
+d <- d[complete.cases(d), ]
+b1 <- plot(euler(d == "Positive"),
+  quantities = list(cex = 3),
+  fills = list(fill = c("#FFC107", "#1E88E5", "#004D40"))
+)
+b2 <- ggplot(as.data.frame(table(d$COS, d$SOS)), aes(Var1, Var2)) +
+  geom_tile(color = "white", fill = "#1E88E5") +
+  geom_text(aes(label = paste0(Freq, "\n(", round(Freq / sum(Freq), 3) * 100, "%)")),
+    color = "white"
+  ) +
+  xlab("COS") +
+  ylab("SOS") +
+  theme_bw() +
+  theme(text = element_text(size = 15))
+p <- ggarrange(a1, a2, b1, b2, nrow = 2, ncol = 2) +
+  bgcolor("White")
+ggsave("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Jordana Hoppe/SSwAB-IT/Dissemination/Figures/fig.png",
+  plot = p, width =9, height = 6, units = "in"
+)
+```

### 2025-02-13 Lilah Melzer/cough_monitor_analysis.qmd
Update cough_monitor_analysis.qmd

@@
 cfqr_teens_adults$sid <- as.numeric(cfqr_teens_adults$sid)
-agg_counts <- left_join(agg_counts, cfqr_parent_6_13)
-agg_counts <- left_join(agg_counts, cfqr_children_6_11)
-agg_counts <- left_join(agg_counts, cfqr_teens_adults)
+agg_counts <- full_join(agg_counts, cfqr_parent_6_13)
+agg_counts <- full_join(agg_counts, cfqr_children_6_11)
+agg_counts <- full_join(agg_counts, cfqr_teens_adults)
 ```
@@
 ```
-
-## Correlation between baseline surveys and coughs
-
-```{r}
-```

### 2025-02-21 Edith Zemanick/Sweat Variability/sweat_test_analysis_chec.qmd
CHEC models

--- 2025-01-29 Edith Zemanick/Sweat Variability/sweat_test_analysis_chec.qmd
@@
       dark: darkly
-bibliography: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Personal/Miscellaneous/zotero.bib
-csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
+bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
+csl: /home/timvigers/GitHub/styles/american-medical-association.csl
 editor: source
@@
 library(ggpubr)
+library(lmerTest)
 library(performance)
-knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability")
+library(gt)
+knitr::opts_knit$set(root.dir = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Variability")
 ```
@@
 ```
+
+# Model-based approaches
+
+```{r}
+model_df <- chec_long %>%
+  select(subject, cl, na, ratio, pre_post) %>%
+  drop_na()
+m <- lmer(cl ~ pre_post * na + (1 | subject), data = model_df)
+# From Amalia
+m1 <- glm(pre_post ~ cl, data = model_df, family = "binomial")
+m2 <- glm(pre_post ~ cl + na, data = model_df, family = "binomial")
+m3 <- glm(pre_post ~ cl * na, data = model_df, family = "binomial")
+m4 <- glm(pre_post ~ ratio, data = model_df, family = "binomial")
+gt(compare_performance(m1, m2, m3, m4, metrics = "common")) %>%
+  fmt_number(decimals = 3)
+model_df$m1_pred <- factor(predict(m1, type = "response") > 0.5,
+  levels = c(F, T), labels = c("pred_pre", "pred_post")
+)
+model_df$m2_pred <- factor(predict(m2, type = "response") > 0.5,
+  levels = c(F, T), labels = c("pred_pre", "pred_post")
+)
+model_df$m3_pred <- factor(predict(m3, type = "response") > 0.5,
+  levels = c(F, T), labels = c("pred_pre", "pred_post")
+)
+model_df$m4_pred <- factor(predict(m4, type = "response") > 0.5,
+  levels = c(F, T), labels = c("pred_pre", "pred_post")
+)
+```
+
+I tried fitting Amalia's models with a random effect for participant, but the random effect variance was 0 and the model estimates are the exact same as a regular logistic model.
+We think that this is because the correlation in responses is accounted for by the design, so the mized effects are not necessary.
+This is worth thinking about a little bit more, but for now we'll go ahead with the simplified logistic models.
+
+## Model 1
+
+```{r}
+m1 %>% gtsummary::tbl_regression(exponentiate = T)
+table(model_df$pre_post, model_df$m1_pred)
+prop.table(table(model_df$pre_post, model_df$m1_pred))
+```
+
+## Model 2
+
+```{r}
+m2 %>% gtsummary::tbl_regression(exponentiate = T)
+table(model_df$pre_post, model_df$m2_pred)
+prop.table(table(model_df$pre_post, model_df$m2_pred))
+```
+
+## Model 3
+
+```{r}
+m3 %>% gtsummary::tbl_regression(exponentiate = T)
+table(model_df$pre_post, model_df$m3_pred)
+prop.table(table(model_df$pre_post, model_df$m3_pred))
+```
+
+## Model 4
+
+```{r}
+m4 %>% gtsummary::tbl_regression(exponentiate = T)
+```
+
+The model using the ratio alone does not really work due to the fitted probabilities issue.

### 2025-02-26 Angela Metcalf/quantifying_cl_cf_carriers.qmd
mostly file path changes

--- 2025-02-06 Angela Metcalf/quantifying_cl_cf_carriers.qmd
@@
       dark: darkly
-bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
-csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
+bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
+csl: /home/timvigers/Documents/GitHub/styles/american-medical-association.csl
 editor: source
@@
 )
-knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Angela Metcalf/Quantifying Sweat Chloride in CF Carriers")
+knitr::opts_knit$set(root.dir = "/home/timvigers/OneDrive/Vigers/CF/Angela Metcalf/Quantifying Sweat Chloride in CF Carriers")
 ```
@@
 #| message: false
-df <- read.csv("./Data_Raw/Midpoint data.csv", na.strings = "")
-df <- df %>%
-  mutate(
-    `Right Arm Cl` = as.numeric(Sweat.chloride..right.arm),
-    `Left Arm Cl` = as.numeric(Sweat.chloride..left.arm),
-    `Right Arm Na` = as.numeric(Sweat.sodium..right.arm),
-    `Left Arm Na` = as.numeric(Sweat.sodium..left.arm)
-  )
-df$`Mean Cl` <- rowMeans(df[, c("Right Arm Cl", "Left Arm Cl")])
-df$`Mean Na` <- rowMeans(df[, c("Right Arm Na", "Left Arm Na")])
+# Import from REDCap
+# unlockREDCap(c(rcon = "Sweat Electrolytes in CF Carriers 24-0691"),
+#   keyring = "API_KEYs",
+#   envir = 1,
+#   url = "https://redcap.ucdenver.edu/api/"
+# )
+# df <- exportRecordsTyped(rcon)
+# # Format
+# df <- df %>%
+#   mutate(
+#     `Cl Right` = as.numeric(cl_right), `Cl Left` = as.numeric(cl_left),
+#     `Na Right` = as.numeric(na_right), `Na Left` = as.numeric(na_left)
+#   )
+# df$`Mean Cl` <- rowMeans(df[, c("Cl Right", "Cl Left")])
+# df$`Mean Na` <- rowMeans(df[, c("Na Right", "Na Left")])
 ```
 
-# Plots
+# Confidence interval power calculations
 
 ```{r}
-ggplot(df,aes(y=`Mean Cl`))+boxplot()
+ns <- 35:45
+power <- lapply(ns, function(n) {
+  se <- qt(0.05 / 2, lower.tail = F, df = n - 1) * (12 / sqrt(n))
+  return(c(
+    "Confidence level" = 0.95, "Sample size (N)" = n,
+    "Actual Distance from Mean to Limits" = se,
+    "Standard deviation of sweat chloride" = 12
+  ))
+})
+power <- do.call(rbind, power)
+knitr::kable(power)
 ```

### 2025-02-26 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
mostly file path changes

@@
 library(ggstats)
-knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry")
+knitr::opts_knit$set(root.dir = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry")
 ```

### 2025-02-26 Christine Chan/EnVision CF/christine_CGM_OGTT_category.qmd
mostly file path changes

@@
 library(ggplot2)
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/tim/OneDrive/Vigers/CF"
-)
-knitr::opts_knit$set(root.dir = home_dir)
+knitr::opts_knit$set(root.dir = "/home/timvigers/OneDrive/Vigers/CF")
 ```
@@
 # Import
-df <- read.csv("./Christine Chan/EnVision CF/Data_Clean/analysis_dataset.csv",
+df <- read.csv("./Christine Chan/EnVision CF/Data_Clean/envision_analysis_dataset.csv",
   na.strings = ""
@@
   select(
-    redcap_data_access_group, age_visit:origin_race, cftr_mutation_1,
-    cftr_mutation_2, weight
+    redcap_data_access_group, age_visit:origin_race, CFTR, weight
   ) %>%
@@
   select(
-    redcap_data_access_group, age_visit:origin_race, cftr_mutation_1,
-    cftr_mutation_2, weight, scully140
+    redcap_data_access_group, age_visit:origin_race, CFTR, weight, scully140
   ) %>%
@@
   select(
-    redcap_data_access_group, age_visit:origin_race, cftr_mutation_1,
-    cftr_mutation_2, weight, scully180
+    redcap_data_access_group, age_visit:origin_race, CFTR, weight, scully180
   ) %>%

### 2025-02-26 Edith Zemanick/Sweat Variability/sweat_test_analysis_chec.qmd
mostly file path changes

@@
 bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
-csl: /home/timvigers/GitHub/styles/american-medical-association.csl
+csl: /home/timvigers/Documents/GitHub/styles/american-medical-association.csl
 editor: source
@@
 m3 <- glm(pre_post ~ cl * na, data = model_df, family = "binomial")
-m4 <- glm(pre_post ~ ratio, data = model_df, family = "binomial")
+m4 <- glm(pre_post ~ scale(ratio), data = model_df, family = "binomial")
 gt(compare_performance(m1, m2, m3, m4, metrics = "common")) %>%

### 2025-02-26 Lilah Melzer/cough_monitor_analysis.qmd
mostly file path changes

@@
 library(patchwork)
-knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor")
+knitr::opts_knit$set(root.dir = "/home/timvigers/OneDrive/Vigers/CF/Lilah Melzer/Lilah Cough Monitor")
 ```

### 2025-02-26 Lilah Melzer/cough_monitor_analysis.qmd
finally finished CFQR and HC survey scoring

@@
       ifelse(
-        sum(is.na(c(phys1, phys2, phys3, phys4, phys5, phys14, phys15, phys16))) <= 4,
+        sum(is.na(c(
+          phys1, phys2, phys3, phys4, phys5, phys14, phys15, phys16
+        ))) <= 4,
         (mean(c(phys1, phys2, phys3, phys4, phys5, phys14, phys15, phys16)) - 1)
@@
       ifelse(
-        sum(is.na(c(emot7, emot8, emot9, emot10, emot11, emot12, emot13, emot14))) <= 4,
-        (mean(c(emot7, emot8, emot9, emot10, emot11, emot12, emot13, emot14)) - 1)
+        sum(is.na(c(
+          emot7, emot8, emot9, emot10, emot11, emot12, emot13, emot14
+        ))) <= 4,
+        (mean(c(
+          emot7, emot8, emot9, emot10, emot11, emot12, emot13, emot14
+        )) - 1)
         / 3 * 100, NA
@@
       ifelse(
-        sum(is.na(c(social20, social21, social22, social23, social24, social25, social26))) <= 3,
-        (mean(c(social20, social21, social22, social23, social24, social25, social26)) - 1)
+        sum(is.na(c(
+          social20, social21, social22, social23, social24, social25, social26
+        ))) <= 3,
+        (mean(c(
+          social20, social21, social22, social23, social24, social25, social26
+        )) - 1)
         / 3 * 100, NA
@@
     Survey = "CFQ-R Children Ages 6-11"
+  ) %>%
+  select(
+    sid, Survey, Date, physical:digest
+  ) %>%
+  filter(!if_all(physical:digest, ~ is.na(.)))
+# Children aged 12 to 13
+cfqr_children_12_13_names <- list(
+  "cfq_r_o1" = "phys1", "cfq_r_o2" = "phys2", "cfq_r_o3" = "phys3",
+  "cfq_r_o4" = "phys4", "cfq_r_o5" = "phys5", "cfq_r_o6" = "phys6",
+  "cfq_r_o7" = "emot7", "cfq_r_o8" = "emot8", "cfq_r_or9" = "emot9",
+  "cfq_r_o10" = "emot10", "cfq_r_o11" = "emot11", "cfq_r_o12" = "emot12",
+  "cfq_r_o13" = "emot13", "cfq_r_o14" = "emot14", "cfq_r_o15" = "eat15",
+  "cfq_r_o16" = "treat16", "cfq_r_o17" = "eat17", "cfq_r_o18" = "treat18",
+  "cfq_r_o19" = "eat19", "cfq_r_o20" = "social20", "cfq_r_o21" = "social21",
+  "cfq_r_o22" = "social22", "cfq_r_o23" = "social23", "cfq_r_o24" = "social24",
+  "cfq_r_o25" = "social25", "cfq_r_o26" = "social26", "cfq_r_o27" = "body27",
+  "cfq_r_o28" = "body28", "cfq_r_o29" = "body29", "cfq_r_o30" = "treat30",
+  "cfq_r_o31" = "resp31", "cfq_r_o32" = "resp32", "cfq_r_o33" = "resp33",
+  "cfq_r_o34" = "resp34", "cfq_r_o35" = "digest35"
+)
+cfqr_children_12_13_names <-
+  setNames(names(cfqr_children_12_13_names), cfqr_children_12_13_names)
+# Some items need to be reverse scored
+cfqr_children_12_13_reverse <- c(
+  "phys1", "phys2", "phys3", "phys4", "phys5", "emot14", "treat18", "eat19",
+  "social20", "social22", "social24", "social26"
+)
+# Do all the calculations
+cfqr_children_12_13 <- df %>%
+  select(
+    sid, date_v1_cf, date_v1_hc,
+    cfqr_children_ages_1213_timestamp, cfq_r_o1:cfq_r_o35
+  ) %>%
+  mutate(
+    Date = as.Date(cfqr_children_ages_1213_timestamp),
+    across(cfq_r_o1:cfq_r_o35, ~ as.numeric(.x))
+  ) %>%
+  rename(all_of(unlist(cfqr_children_12_13_names))) %>%
+  mutate(across(all_of(cfqr_children_12_13_reverse), ~ 5 - .x)) %>%
+  rowwise() %>%
+  mutate(
+    physical =
+      ifelse(
+        sum(is.na(c(phys1, phys2, phys3, phys4, phys5, phys6))) <= 3,
+        (mean(c(phys1, phys2, phys3, phys4, phys5, phys6)) - 1)
+        / 3 * 100, NA
+      ),
+    emotion =
+      ifelse(
+        sum(is.na(c(
+          emot7, emot8, emot9, emot10, emot11, emot12, emot13, emot14
+        ))) <= 4,
+        (mean(c(
+          emot7, emot8, emot9, emot10, emot11, emot12, emot13, emot14
+        )) - 1)
+        / 3 * 100, NA
+      ),
+    social =
+      ifelse(
+        sum(is.na(c(
+          social20, social21, social22, social23, social24, social25, social26
+        ))) <= 3,
+        (mean(c(
+          social20, social21, social22, social23, social24, social25, social26
+        )) - 1)
+        / 3 * 100, NA
+      ),
+    eat =
+      ifelse(
+        sum(is.na(c(eat15, eat17, eat19))) <= 1,
+        (mean(c(eat15, eat17, eat19)) - 1)
+        / 3 * 100, NA
+      ),
+    body =
+      ifelse(
+        sum(is.na(c(body27, body28, body29))) <= 1,
+        (mean(c(body27, body28, body29)) - 1)
+        / 3 * 100, NA
+      ),
+    treat =
+      ifelse(
+        sum(is.na(c(treat16, treat18, treat30))) <= 1,
+        (mean(c(treat16, treat18, treat30)) - 1)
+        / 3 * 100, NA
+      ),
+    respirat =
+      ifelse(
+        sum(is.na(c(resp31, resp32, resp33, resp34))) <= 2,
+        (mean(c(resp31, resp32, resp33, resp34)) - 1)
+        / 3 * 100, NA
+      ),
+    digest =
+      ifelse(
+        sum(is.na(c(digest35))) == 0,
+        (mean(c(digest35)) - 1)
+        / 3 * 100, NA
+      )
+  ) %>%
+  mutate(
+    Date = coalesce(Date, date_v1_cf),
+    Date = coalesce(Date, date_v1_hc),
+    Survey = "CFQ-R Children Ages 12-13"
   ) %>%
@@
       ifelse(
-        sum(is.na(c(social22, social23, social27, social28, social29, social30))) <= 3,
-        (mean(c(social22, social23, social27, social28, social29, social30)) - 1)
+        sum(is.na(c(
+          social22, social23, social27, social28, social29, social30
+        ))) <= 3,
+        (mean(c(
+          social22, social23, social27, social28, social29, social30
+        )) - 1)
         / 3 * 100, NA
@@
   filter(!if_all(physical:digest, ~ is.na(.)))
+# Controls
+hc_surveys <- df %>%
+  select(
+    sid, date_v1_cf, date_v1_hc, hc_pediatric_cough_survey_timestamp, q1:q5
+  ) %>%
+  mutate(
+    Date = as.Date(hc_pediatric_cough_survey_timestamp),
+    Date = coalesce(Date, date_v1_cf),
+    Date = coalesce(Date, date_v1_hc),
+    Survey = "HC Cough Survey"
+  ) %>%
+  select(
+    sid, Survey, Date, q1:q5
+  ) %>%
+  filter(!if_all(q1:q5, ~ is.na(.)))
+hc_surveys$sid <- as.numeric(hc_surveys$sid)
 # Combine
-surveys <- full_join(cfqr_parent_6_13, cfqr_children_6_11)
-surveys <- full_join(surveys, cfqr_teens_adults)
+cf_surveys <- full_join(cfqr_parent_6_13, cfqr_children_6_11)
+cf_surveys <- full_join(cf_surveys, cfqr_children_12_13)
+cf_surveys <- full_join(cf_surveys, cfqr_teens_adults)
+cf_surveys$sid <- as.numeric(cf_surveys$sid)
 # Add to counts for plotting
-cfqr_parent_6_13$sid <- as.numeric(cfqr_parent_6_13$sid)
-cfqr_children_6_11$sid <- as.numeric(cfqr_children_6_11$sid)
-cfqr_teens_adults$sid <- as.numeric(cfqr_teens_adults$sid)
-agg_counts <- full_join(agg_counts, cfqr_parent_6_13)
-agg_counts <- full_join(agg_counts, cfqr_children_6_11)
-agg_counts <- full_join(agg_counts, cfqr_teens_adults)
+agg_counts <- full_join(agg_counts, cf_surveys, by = join_by(sid, Date))
+agg_counts <- full_join(agg_counts, hc_surveys, by = join_by(sid, Date, Survey))
 ```

### 2025-02-26 Lilah Melzer/cough_monitor_analysis.qmd
cough monitor surveys analyzed

@@
 library(patchwork)
+library(gt)
 knitr::opts_knit$set(root.dir = "/home/timvigers/OneDrive/Vigers/CF/Lilah Melzer/Lilah Cough Monitor")
@@
 
-# Coughs over time (aggregated)
-
-## Plot
-
-### All participants
-
-```{r}
-#| label: fig-coughs-by-night
-#| fig-cap: Coughs by nights from start
-#| warning: false
-ggplot(agg_counts[!is.na(agg_counts$Coughs), ], aes(
-  x = Night, y = Coughs, group = factor(sid), color = Diagnosis
-)) +
-  geom_point(alpha = 0.1) +
-  geom_line(alpha = 0.1) +
-  geom_smooth(aes(group = Diagnosis)) +
-  theme_bw()
-```
-
-### By participant
-
-```{r results='asis'}
-#| label: fig-coughs-by-night-2
-#| fig-cap: Coughs by nights from start
-#| warning: false
-invisible(lapply(unique(agg_counts$sid), function(s) {
-  d <- agg_counts[agg_counts$sid == s, ]
-  diag <- d$Diagnosis[1]
-  p <- ggplot(d, aes(
-    x = Night, y = Coughs, group = factor(sid), color = Sick
-  )) +
-    geom_point() +
-    geom_line() +
-    theme_bw() +
-    ylim(0, 50) +
-    xlim(0, 150) +
-    scale_color_manual(values = c("#377EB8", "#E41A1C")) +
-    ggtitle(paste0("Participant ", s, " (", diag, ")"))
-  if (s == 38) {
-    p <- p + ylim(0, 270)
-  }
-  cat("\n")
-  cat("\n")
-  print(p)
-  cat("\n")
-  cat("\n")
-}))
-```
-
 # Basic statistics excluding 0s
@@
 Spearman correlation between median coughs and pre-bronchodilator FEV1 was `r round(as.numeric(M$estimate),3)` (p = `r format.pval(M$p.value,eps = 0.001,digits = 3)`).
+
+# Coughs over time (aggregated)
+
+```{r}
+#| label: fig-coughs-by-night
+#| fig-cap: Coughs by nights from start
+#| warning: false
+ggplot(agg_counts[!is.na(agg_counts$Coughs), ], aes(
+  x = Night, y = Coughs, group = factor(sid), color = Diagnosis
+)) +
+  geom_point(alpha = 0.1) +
+  geom_line(alpha = 0.1) +
+  geom_smooth(aes(group = Diagnosis)) +
+  theme_bw()
+```
+
+# Zero-inflated models
+
+We compared several models using BIC.
+First, we compared a zero-inflated Poisson (ZIP) model with no fixed effects in the zero part of the model to one with time and diagnosis included as fixed effects in the zero part (both parts of the model included a random intercept for participant).
+The model with the same fixed effects in both parts was better (10088.29 vs.
+10071.32).
+
+Next, we compared this model to the same model structure but with a negative binomial distribution (ZINB) rather than Poisson to account for overdispersion in the non-zero part of the model.
+The ZINB was significantly better than the ZIP model (10071.32 vs.
+6810.49).
+
+Finally, we compared the ZINB model with random intercept in both parts of the model to one with an additional random slope in the zero part, one with random slope in the count part, and one with random slope in both parts.
+The ZINB model with random slope in the non-zero part of the model but only a random intercept in the zero part was better than the simple ZINB model (6810.49 vs.
+6753.76).
+Interestingly, it was also better than the model with a random slope in both parts (6732.82 vs.
+6801.86).
+
+```{r}
+zip1 <- mixed_model(Coughs ~ Night + Diagnosis,
+  data = agg_counts,
+  random = ~ 1 | sid, zi_random = ~ 1 | sid,
+  family = zi.poisson(), zi_fixed = ~1
+)
+zip2 <- update(zip1, zi_fixed = ~ Night + Diagnosis)
+
+zinb1 <- update(zip2, family = zi.negative.binomial())
+
+zinb2 <- update(zinb1, random = ~ Night | sid)
+
+zinb3 <- update(zinb1, zi_random = ~ Night | sid)
+
+zinb <- zinb2
+
+s <- summary(zinb)
+
+s$coef_table_zi %>%
+  data.frame() %>%
+  rownames_to_column() %>%
+  gt(caption = "Zero-part coefficients") %>%
+  fmt_number(decimals = 3)
+
+s$coef_table %>%
+  data.frame() %>%
+  rownames_to_column() %>%
+  gt(caption = "Count coefficients") %>%
+  fmt_number(decimals = 3)
+```
+
+For now, the results above are still on the log scale, so a negative value in the estimate column indicates a lower odds ratio or incidence rate ratio.
+The zero-part coefficients indicate that the odds of having a night with 0 coughs were significantly lower in the CF group (log(OR) = -1.277, OR = 0.279, p = 0.02).
+In other words, people with CF were more likely to cough than those without CF.
+However, the count coefficients table indicates that given a night with more than 0 coughs, people with CF coughed at a higher rate than those without CF, but this difference was not statistically significant (log(OR) = 0.591, OR = 1.805, p = 0.082).
+
+## By sick status
+
+To examine the effect of sickness on cough rate, we kept the same random effect structure as above, but added a fixed effect for sick status in both parts of the model.
+
+```{r}
+zinb_sick <- update(zinb,
+  fixed = Coughs ~ Night + Diagnosis + Sick,
+  zi_fixed = ~ Night + Diagnosis + Sick
+)
+
+s <- summary(zinb_sick)
+
+s$coef_table_zi %>%
+  data.frame() %>%
+  rownames_to_column() %>%
+  gt(caption = "Zero-part coefficients") %>%
+  fmt_number(decimals = 3)
+
+s$coef_table %>%
+  data.frame() %>%
+  rownames_to_column() %>%
+  gt(caption = "Count coefficients") %>%
+  fmt_number(decimals = 3)
+```
+
+The interpretation of these results is generally the same as above.
+The zero-part coefficients indicate that people with CF are still more likely to cough than people without CF (log(OR) = -1.301, OR = 0.272, p = 0.018), even after adjusting for sickness.
+Sickness increased the odds of coughing, but the difference was not statistically significant (log(OR) = 0.693, OR = 1.999, p = 0.346).
+However, given a night with coughs, sickness increased the rate of cough significantly (log(OR) = 1.244, OR = 3.468, p = 0.011).
 
@@
 hc_surveys$sid <- as.numeric(hc_surveys$sid)
-# Combine
-cf_surveys <- full_join(cfqr_parent_6_13, cfqr_children_6_11)
-cf_surveys <- full_join(cf_surveys, cfqr_children_12_13)
-cf_surveys <- full_join(cf_surveys, cfqr_teens_adults)
-cf_surveys$sid <- as.numeric(cf_surveys$sid)
-# Add to counts for plotting
-agg_counts <- full_join(agg_counts, cf_surveys, by = join_by(sid, Date))
-agg_counts <- full_join(agg_counts, hc_surveys, by = join_by(sid, Date, Survey))
 ```
@@
 
-### Plots by participant
+I tried to examine the association between respiratory scores and cough using similar ZINB to above, but there just isn't enough data to fit these models.
+For example, looking at the caregiver surveys, there are only two participants with both survey results and cough data on the same date.
+The two participants have 8 and 12 dates with data, but by using the model approach we would be excluding 5 participants who each took the survey only once.
+
+So, instead of the model approach I did a more simple correlation between median coughs per night and median respiratory score.
+
+### CFQ-R Caregivers 6-13
+
+#### Median scores
+
+```{r}
+#| warning: false
+cfqr_parent_6_13$sid <- as.numeric(cfqr_parent_6_13$sid)
+caregivers <-
+  full_join(agg_counts, cfqr_parent_6_13, join_by(sid, Date)) %>%
+  select(sid, Sick, Night, Coughs, respirat) %>%
+  filter(!if_all(Coughs:respirat, ~ is.na(.)))
+caregivers_med <- caregivers %>%
+  group_by(sid) %>%
+  summarise(
+    `Median Coughs` = median(Coughs, na.rm = T),
+    `Median Resp. Score` = median(respirat, na.rm = T)
+  ) %>%
+  drop_na()
+M <- cor.test(caregivers_med$`Median Coughs`,
+  caregivers_med$`Median Resp. Score`,
+  method = "spearman"
+)
+ggplot(caregivers_med, aes(
+  y = `Median Coughs`, x = `Median Resp. Score`,
+  colour = factor(sid)
+)) +
+  geom_point() +
+  theme_bw()
+```
+
+Spearman's correlation between median coughs and median respiratory symptom score was `r round(as.numeric(M$estimate),3)` (p = `r format.pval(M$p.value,digits = 3,eps=0.001)`).
+
+#### Plots by participant
 
@@
 #| warning: false
-invisible(lapply(unique(agg_counts$sid), function(s) {
-  d <- agg_counts[agg_counts$sid == s, ]
-  diag <- d$Diagnosis[1]
-  p <- ggplot(d, aes(
-    x = Night, y = Coughs, group = factor(sid), color = Sick
-  )) +
+invisible(lapply(unique(caregivers_med$sid), function(s) {
+  d <- caregivers[caregivers$sid == s, ]
+  d <- d %>% pivot_longer(c(Coughs, respirat))
+  p <- ggplot(d, aes(x = Night, y = value, color = Sick)) +
     geom_point() +
@@
     theme_bw() +
-    ylim(0, 50) +
-    xlim(0, 150) +
     scale_color_manual(values = c("#377EB8", "#E41A1C")) +
-    ggtitle(paste0("Participant ", s, " (", diag, ")"))
-  if (s == 38) {
-    p <- p + ylim(0, 270)
-  }
-  p2 <- ggplot(d[!is.na(d$respirat), ], aes(
-    x = Night, y = respirat, group = factor(sid), color = Sick
+    facet_wrap(~name, ncol = 1) +
+    ggtitle(paste0("Participant ", s))
+  cat("\n")
+  cat("\n")
+  print(p)
+  cat("\n")
+  cat("\n")
+}))
+```
+
+### CFQ-R Children Ages 12-13
+
+#### Median scores
+
+```{r}
+#| warning: false
+cfqr_children_12_13$sid <- as.numeric(cfqr_children_12_13$sid)
+children_12_13 <-
+  full_join(agg_counts, cfqr_children_12_13, join_by(sid, Date)) %>%
+  select(sid, Sick, Night, Coughs, respirat) %>%
+  filter(!if_all(Coughs:respirat, ~ is.na(.)))
+children_12_13_med <- children_12_13 %>%
+  group_by(sid) %>%
+  summarise(
+    `Median Coughs` = median(Coughs, na.rm = T),
+    `Median Resp. Score` = median(respirat, na.rm = T)
+  ) %>%
+  drop_na()
+M <- cor.test(children_12_13_med$`Median Coughs`,
+  children_12_13_med$`Median Resp. Score`,
+  method = "spearman"
+)
+ggplot(children_12_13_med, aes(
+  y = `Median Coughs`, x = `Median Resp. Score`,
+  colour = factor(sid)
 )) +
+  geom_point() +
+  theme_bw()
+```
+
+Spearman's correlation between median coughs and median respiratory symptom score was `r round(as.numeric(M$estimate),3)` (p = `r format.pval(M$p.value,digits = 3,eps=0.001)`).
+
+#### Plots by participant
+
+```{r}
+#| message: false
+#| warning: false
+invisible(lapply(unique(children_12_13_med$sid), function(s) {
+  d <- children_12_13[children_12_13$sid == s, ]
+  d <- d %>% pivot_longer(c(Coughs, respirat))
+  p <- ggplot(d, aes(x = Night, y = value, color = Sick)) +
     geom_point() +
@@
     theme_bw() +
-    ylim(50, 100) +
-    xlim(0, 150) +
     scale_color_manual(values = c("#377EB8", "#E41A1C")) +
-    ggtitle(paste0("Participant ", s, " (", diag, ")"))
+    facet_wrap(~name, ncol = 1) +
+    ggtitle(paste0("Participant ", s))
   cat("\n")
   cat("\n")
-  print(p + p2 + plot_layout(guides = "collect"))
+  print(p)
   cat("\n")
@@
 ```
+
+### CFQ-R Adolescents and Adults 14 Years
+
+#### Median scores
+
+```{r}
+#| warning: false
+cfqr_teens_adults$sid <- as.numeric(cfqr_teens_adults$sid)
+teens_adults <-
+  full_join(agg_counts, cfqr_teens_adults, join_by(sid, Date)) %>%
+  select(sid, Sick, Night, Coughs, respirat) %>%
+  filter(!if_all(Coughs:respirat, ~ is.na(.)))
+teens_adults_med <- teens_adults %>%
+  group_by(sid) %>%
+  summarise(
+    `Median Coughs` = median(Coughs, na.rm = T),
+    `Median Resp. Score` = median(respirat, na.rm = T)
+  ) %>%
+  drop_na()
+M <- cor.test(teens_adults_med$`Median Coughs`,
+  teens_adults_med$`Median Resp. Score`,
+  method = "spearman"
+)
+ggplot(teens_adults_med, aes(
+  y = `Median Coughs`, x = `Median Resp. Score`,
+  colour = factor(sid)
+)) +
+  geom_point() +
+  theme_bw()
+```
+
+Spearman's correlation between median coughs and median respiratory symptom score was `r round(as.numeric(M$estimate),3)` (p = `r format.pval(M$p.value,digits = 3,eps=0.001)`).
+
+#### Plots by participant
+
+```{r}
+#| message: false
+#| warning: false
+invisible(lapply(unique(teens_adults_med$sid), function(s) {
+  d <- teens_adults[teens_adults$sid == s, ]
+  d <- d %>% pivot_longer(c(Coughs, respirat))
+  p <- ggplot(d, aes(x = Night, y = value, color = Sick)) +
+    geom_point() +
+    geom_line() +
+    theme_bw() +
+    scale_color_manual(values = c("#377EB8", "#E41A1C")) +
+    facet_wrap(~name, ncol = 1) +
+    ggtitle(paste0("Participant ", s))
+  cat("\n")
+  cat("\n")
+  print(p)
+  cat("\n")
+  cat("\n")
+}))
+```
+
+### Healthy controls
+
+#### Median scores
+
+```{r}
+#| warning: false
+hc_surveys$sid <- as.numeric(hc_surveys$sid)
+hcs <-
+  full_join(agg_counts, hc_surveys, join_by(sid, Date)) %>%
+  select(sid, Sick, Night, Coughs, q1:q5) %>%
+  filter(!if_all(q1:q5, ~ is.na(.)))
+hcs_med <- hcs %>%
+  group_by(sid) %>%
+  summarise(
+    `Median Coughs` = median(Coughs, na.rm = T),
+    `Median Resp. Score` = median(as.numeric(q1), na.rm = T)
+  ) %>%
+  drop_na()
+M <- cor.test(hcs_med$`Median Coughs`,
+  hcs_med$`Median Resp. Score`,
+  method = "spearman"
+)
+ggplot(hcs_med, aes(
+  y = `Median Coughs`, x = `Median Resp. Score`,
+  colour = factor(sid)
+)) +
+  geom_point() +
+  theme_bw()
+```
+
+Spearman's correlation between median coughs and median respiratory symptom score was `r round(as.numeric(M$estimate),3)` (p = `r format.pval(M$p.value,digits = 3,eps=0.001)`).

### 2025-02-28 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
B&W Euler diagrams

--- 2025-02-12 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
@@
   cat("\n")
-  print(plot(euler(d == "Negative"),
-    quantities = list(cex = 3),
-    fills = list(fill = c("#FFC107", "#1E88E5", "#004D40"))
-  ))
+  print(plot(euler(d == "Negative"), quantities = list(cex = 3)))
   cat("\n")
@@
   cat("\n")
-  print(plot(euler(d == "Positive"),
-    quantities = list(cex = 3),
-    fills = list(fill = c("#FFC107", "#1E88E5", "#004D40"))
-  ))
+  print(plot(euler(d == "Positive"), quantities = list(cex = 3)))
   cat("\n")
@@
 ```
+
+# Manuscript Figures
+
+```{r}
+# All pathogens
+d_overall <- df[, c("overall_cos", "overall_sos")]
+colnames(d_overall) <- c("COS", "SOS")
+d_overall <- d_overall[complete.cases(d_overall), ]
+a1 <- plot(euler(d_overall == "Positive"), quantities = list(cex = 3))
+a2 <-
+  ggplot(as.data.frame(table(d_overall$COS, d_overall$SOS)), aes(Var1, Var2)) +
+  geom_tile(color = "white", fill = "#1E88E5") +
+  geom_text(
+    aes(label = paste0(Freq, "\n(", round(Freq / sum(Freq), 3) * 100, "%)")),
+    color = "white"
+  ) +
+  xlab("COS") +
+  ylab("SOS") +
+  theme_bw() +
+  theme(text = element_text(size = 15))
+# MRSA
+d_MSSA <- df[, c("staphylococcus_aureus_cos", "staphylococcus_aureus_sos")]
+colnames(d_MSSA) <- c("COS", "SOS")
+d_MSSA <- d_MSSA[complete.cases(d_MSSA), ]
+b1 <- plot(euler(d_MSSA == "Positive"), quantities = list(cex = 3))
+b2 <-
+  ggplot(as.data.frame(table(d_MSSA$COS, d_MSSA$SOS)), aes(Var1, Var2)) +
+  geom_tile(color = "white", fill = "#1E88E5") +
+  geom_text(aes(label = paste0(Freq, "\n(", round(Freq / sum(Freq), 3) * 100, "%)")),
+    color = "white"
+  ) +
+  xlab("COS") +
+  ylab("SOS") +
+  theme_bw() +
+  theme(text = element_text(size = 15))
+p1 <- ggarrange(a1, a2, b1, b2, nrow = 2, ncol = 2) +
+  bgcolor("White")
+ggsave("/home/timvigers/OneDrive/Vigers/CF/Jordana Hoppe/SSwAB-IT/Dissemination/Figures/fig.png",
+  plot = p1, width = 9, height = 6, units = "in"
+)
+a1 <- plot(euler(d_overall == "Positive"),
+  quantities = list(cex = 3),
+  main = "Any CF Pathogen"
+)
+b1 <- plot(euler(d_MSSA == "Positive"),
+  quantities = list(cex = 3),
+  main = "MSSA"
+)
+p2 <- ggarrange(a1, b1, nrow = 1, ncol = 2) +
+  bgcolor("White")
+ggsave("/home/timvigers/OneDrive/Vigers/CF/Jordana Hoppe/SSwAB-IT/Dissemination/Figures/euler_bw.png",
+  plot = p2, width = 9, height = 6, units = "in"
+)
+a1 <- plot(venn(d_overall == "Positive"),
+  quantities = list(cex = 3),
+  main = "Any CF Pathogen"
+)
+b1 <- plot(venn(d_MSSA == "Positive"),
+  quantities = list(cex = 3),
+  main = "MSSA"
+)
+p3 <- ggarrange(a1, b1, nrow = 1, ncol = 2) +
+  bgcolor("White")
+ggsave("/home/timvigers/OneDrive/Vigers/CF/Jordana Hoppe/SSwAB-IT/Dissemination/Figures/venn_bw.png",
+  plot = p3, width = 9, height = 6, units = "in"
+)
+```

### 2025-03-03 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
CFRD survival new BMI categories and fixed enzyme levels

@@
   })
-# Calculate percentiles
+# Get BMI categories
 annual$bmi <- annual$A_weight / ((annual$A_height / 100)^2)
-annual$age_perc <- ifelse(annual$Age_YrEnd < 20, annual$Age_YrEnd, 20)
-annual$height_perc <- sds(annual$A_height,
-  age = annual$age_perc,
-  sex = annual$Gender, male = "M", female = "F",
-  item = "height2_20", ref = cdc.ref, type = "perc"
-) * 100
-annual$bmi_perc <- sds(annual$bmi,
-  age = annual$age_perc,
+annual$bmi <- coalesce(annual$A_bmivalue, annual$bmi)
+annual$bmi_perc_cdc <- sds(annual$bmi,
+  age = annual$Age_YrEnd,
   sex = annual$Gender, male = "M", female = "F",
@@
 ) * 100
+annual$bmi_perc_who <- sds(annual$bmi,
+  age = annual$Age_YrEnd,
+  sex = annual$Gender, male = "M", female = "F",
+  item = "bmi", ref = who.ref, type = "perc"
+) * 100
+annual$bmi_perc <- coalesce(annual$bmi_perc_cdc, annual$bmi_perc_who)
+annual$bmi_cat <- apply(annual, 1, function(r) {
+  age <- as.numeric(r["Age_YrEnd"])
+  gender <- as.character(r["Gender"])
+  bmi <- as.numeric(r["bmi"])
+  bmi_perc <- as.numeric(r["bmi_perc"])
+  if (age <= 20) {
+    return(cut(bmi_perc, c(-Inf, 10, 50, 85, Inf),
+      right = F,
+      labels = c("Undernourished", "At Risk", "Target/Goal", "Overweight")
+    ))
+  } else if (age > 20) {
+    if (gender == "M") {
+      return(cut(bmi, c(-Inf, 18.5, 23, 25, Inf),
+        right = F,
+        labels = c("Undernourished", "At Risk", "Target/Goal", "Overweight")
+      ))
+    } else if (gender == "F") {
+      return(cut(bmi, c(-Inf, 18.5, 22, 25, Inf),
+        right = F,
+        labels = c("Undernourished", "At Risk", "Target/Goal", "Overweight")
+      ))
+    }
+  }
+})
+annual$bmi_cat <- relevel(annual$bmi_cat, ref = "Target/Goal")
 # Format columns
@@
 )
-annual$A_IsOnEnzymes <- relevel(annual$A_IsOnEnzymes, ref = "Yes")
-# BMI percentile to categorical
-annual$bmi_perc_cat <- cut(annual$bmi_perc, breaks = c(0, 10, 50, 85, Inf))
-annual$bmi_perc_cat <- relevel(annual$bmi_perc_cat, ref = "(50,85]")
+annual$A_IsOnEnzymes <- factor(annual$A_IsOnEnzymes, levels = c("Yes", "No"))
 # FEV1 to categorical
@@
 vars <- c(
-  "age_group", "Gender", "bmi_perc", "bmi_perc_cat", "MutClass", "Race",
+  "age_group", "Gender", "bmi_perc", "bmi_cat", "MutClass", "Race",
   "Hispanicrace", "A_IsOnEnzymes", "ppFEV1_cat", "tube_feeding",
@@
   NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
-  height_perc = "Height percentile", bmi_perc_cat = "BMI percentile",
+  height_perc = "Height percentile", bmi_cat = "BMI category",
   A_Vx770 = "Use of CFTR modulator Vx770",
@@
     ReviewYear, CFRD, Age_YrEnd, Gender, Race, Hispanicrace, bmi_perc,
-    bmi_perc_cat, MutClass, tube_feeding, liver_disease,
+    bmi_cat, MutClass, tube_feeding, liver_disease,
     A_FEV1_pct_predicted, A_FVC_pct_predicted, NumPulmExacerbation,
@@
 xvars <- c(
-  "age_group", "Gender", "Race", "Hispanicrace", "bmi_perc_cat", "MutClass",
+  "age_group", "Gender", "Race", "Hispanicrace", "bmi_cat", "MutClass",
   "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
@@
 )
-comb_plot = ggcoef_compare(
+comb_plot <- ggcoef_compare(
   list("VX770" = vx770_cox_res, "VX809comb" = vx809comb_cox_res),
@@
 )
-comb_plot = ggcoef_compare(
+comb_plot <- ggcoef_compare(
   list("2003 - 2008" = era1, "2009 - 2013" = era2, "2014 - 2018" = era3),

### 2025-03-03 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
updated plots for Christine

@@
 ggcoef_model(era1,
-  show_p_values = F, add_reference_rows = T, exponentiate = T,
-  signif_stars = F, colour = F
+  show_p_values = F, add_reference_rows = T, exponentiate = T, shape_guide = F,
+  signif_stars = F, colour = "significance", shape = "reference_row",
+  shape_values = c(16, 15)
 )
@@
 ggcoef_model(era2,
-  show_p_values = F, add_reference_rows = T, exponentiate = T,
-  signif_stars = F, colour = F
+  show_p_values = F, add_reference_rows = T, exponentiate = T, shape_guide = F,
+  signif_stars = F, colour = "significance", shape = "reference_row",
+  shape_values = c(16, 15)
 )
@@
 ggcoef_model(era3,
-  show_p_values = F, add_reference_rows = T, exponentiate = T,
-  signif_stars = F, colour = F
+  show_p_values = F, add_reference_rows = T, exponentiate = T, shape_guide = F,
+  signif_stars = F, colour = "significance", shape = "reference_row",
+  shape_values = c(16, 15)
 )
@@
   list("2003 - 2008" = era1, "2009 - 2013" = era2, "2014 - 2018" = era3),
-  add_reference_rows = T, exponentiate = T, type = "faceted", colour = F
+  add_reference_rows = T, exponentiate = T, shape_guide = F,
+  colour = "significance", shape = "reference_row", shape_values = c(16, 15),
+  type = "faceted"
 )

### 2025-03-04 Christine Chan/CF Registry/cf_registry_CFRD_prediction.qmd
joint model errors

--- 2024-12-05 Christine Chan/CF Registry/cf_registry_CFRD_prediction.qmd
@@
 #| include: false
-library(tidyverse)
-library(arsenal)
 library(survival)
-library(ggsurvfit)
-library(ggstats)
-library(gtsummary)
-knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry")
+library(nlme)
+library(JM)
+library(dplyr)
+library(tidyr)
+knitr::opts_knit$set(root.dir = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry")
 ```
@@
 )
-# Select columns
+# Get age at first CFRD (by OGTT and registry)
 annual <- annual %>%
-  select(
-    eDWID, ReviewYear, Race, Hispanicrace, Gender, F508, MutClass,
-    A_Vx770, A_VX809comb, Age_YrEnd, A_bmipercentile, A_hgba1c,
-    cfrd_registry, cfrd_status_annual, cfrd_ogtt, A_twohour_bloodglucose
-  ) %>%
   group_by(eDWID) %>%
   mutate(
-    Age_YrEnd = round(Age_YrEnd, 2),
     cfrd_registry_age = first(na.omit(Age_YrEnd[cfrd_registry == "CFRD+"])),
@@
 annual$cfrd_ogtt[annual$Age_YrEnd >= annual$cfrd_ogtt_age] <- "CFRD+"
-# Split into one dataset for registry and one for OGTT
-registry <- annual %>%
-  filter(Age_YrEnd <= cfrd_registry_age) %>%
-  select(eDWID:A_hgba1c, cfrd_registry)
-ogtt <- annual %>%
-  filter(Age_YrEnd <= cfrd_ogtt_age) %>%
-  select(eDWID:A_hgba1c, cfrd_ogtt, cfrd_ogtt_age)
 # Create start times
-registry$Age_YrStart <- registry$Age_YrEnd - 1
-ogtt$Age_YrStart <- ogtt$Age_YrEnd - 1
+annual$Age_YrStart <- annual$Age_YrEnd - 1
+# Format columns
+annual$MutClass <- factor(annual$MutClass,
+  levels = c("1-3", "4-5", "Oth", "N/A"),
+  labels = c("1-3", "4-5", "Oth", "Unknown")
+)
+annual$NumPulmExacerbation <- cut(annual$NumPulmExacerbation,
+  labels = c("0", "1", "2", ">=3"),
+  breaks = c(0, 1, 2, 3, Inf), right = F
+)
+# Labels
+labels <- list(
+  Age_YrStart = "Age at Year Start", Age_YrEnd = "Age at Year End",
+  Hispanicrace = "Hispanic", A_Vx770 = "Use of CFTR modulator Vx770",
+  A_VX809comb = "Use of CFTR modulator Vx809 combination",
+  A_FEV1_pct_predicted = "FEV1 percent predicted", MutClass = "Mutation class",
+  NumPulmExacerbation = "Admissions for PEx"
+)
+# Select columns
+annual <- annual %>%
+  select(
+    eDWID, Race, Hispanicrace, Gender, F508, MutClass,
+    ReviewYear, Age_YrStart, Age_YrEnd, A_Vx770, A_VX809comb, A_bmivalue,
+    A_FEV1_pct_predicted, NumPulmExacerbation, A_IsOnEnzymes,
+    A_fast_bloodglucose, A_hgba1c, cfrd_registry, cfrd_ogtt,
+    A_twohour_bloodglucose, cfrd_status_annual
+  )
+```
+
+# Joint models
+
+```{r}
+# Complete cases
+model_df <- annual %>%
+  select(eDWID, Gender, Age_YrStart, Age_YrEnd, A_hgba1c, cfrd_registry) %>%
+  drop_na()
+# Fit mixed model and Cox model separately
+lme_fit_a1c <- lme(A_hgba1c ~ Age_YrStart + Gender,
+  random = ~ Age_YrStart | eDWID,
+  data = model_df
+)
+cox_fit <- coxph(
+  Surv(
+    time = Age_YrStart, time2 = Age_YrEnd,
+    event = as.numeric(cfrd_registry == "CFRD+")
+  ) ~ Gender,
+  cluster = eDWID,
+  data = model_df, x = T, model = T
+)
+# Combine
+joint_fit <- jointModel(lme_fit_a1c, cox_fit,
+  timeVar = "Age_YrStart", method = "piecewise-PH-aGH"
+)
 ```
@@
 cph <- coxph(
-  Surv(Age_YrStart, Age_YrEnd, as.numeric(cfrd_registry)) ~
-    A_hgba1c + Gender + F508 + A_Vx770 + A_VX809comb,
+  update(f, Surv(time = Age_YrStart, time2 = Age_YrEnd, event = as.numeric(cfrd_registry) == 2) ~ .),
   data = registry, cluster = eDWID
@@
 ggcoef_table(cph,
-  show_p_values = F, add_reference_rows = T, signif_stars = F, colour = F
+  exponentiate = T, show_p_values = F, add_reference_rows = T, signif_stars = F,
+  colour = F
 )
+
+d <- model.frame(cph)
+d$pred <- predict(cph, type = "survival")
 ```

### 2025-03-05 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
CFRD epi paper updates

@@
 yn_vars <- c(
-  "A_IsOnEnzymes", "A_corticosteroids1", "A_feeding2",
+  "A_corticosteroids1", "A_feeding2",
   "A_Vx770", "A_VX809comb", bugs, "tube_feeding", "liver_disease"
@@
 )
-annual$A_IsOnEnzymes <- factor(annual$A_IsOnEnzymes, levels = c("Yes", "No"))
+annual$A_IsOnEnzymes <- factor(annual$A_IsOnEnzymes,
+  levels = c(1, 0), labels = c(" Yes", " No")
+)
 # FEV1 to categorical
@@
 )
-annual$ppFEV1_cat <- relevel(annual$ppFEV1_cat, ref = "[100,Inf)")
+annual$ppFEV1_cat <- factor(annual$ppFEV1_cat,
+  levels = c("[100,Inf)", "[80,100)", "[40,80)", "[0,40)")
+)
 # Males to reference group
@@
   labels[colnames(annual)[which(colnames(annual) %in% names(labels))]]
+# Remove those on modulators
+# annual <- annual %>%
+#   filter((is.na(A_Vx770) | A_Vx770 == "No") &
+#     (is.na(A_VX809comb) | A_VX809comb == "No"))
 ```
@@
 ```{r}
-annual %>%
+prev_year_age <- annual %>%
   group_by(ReviewYear, age_group) %>%
   summarise(Prevalence = mean(CFRD == "CFRD+") * 100, .groups = "drop") %>%
-  pivot_wider(names_from = age_group, values_from = Prevalence) %>%
-  kable(., digits = 2)
+  pivot_wider(names_from = age_group, values_from = Prevalence)
+kable(prev_year_age, digits = 2)
 ```
@@
   )
-kable(
-  screening %>%
+screening_table <- screening %>%
   select(ReviewYear, age_group, `Screened OGTT`) %>%
   mutate(`Screened OGTT` = round(`Screened OGTT` * 100, 2)) %>%
-    pivot_wider(names_from = age_group, values_from = `Screened OGTT`),
-  caption = "% Screened by OGTT"
-)
+  pivot_wider(names_from = age_group, values_from = `Screened OGTT`)
+kable(screening_table, caption = "% Screened by OGTT")
 screening <- screening %>%
@@
   print()
+```
+
+# Figure 1
+
+```{r}
+p <- prev_year_age %>%
+  pivot_longer(`[10,19)`:`[31, Inf)`, names_to = "Age Group") %>%
+  rename(Year = ReviewYear)
+p$name <- "Prevalence"
+i <- incidence %>%
+  select(
+    Year, `Total Incidence Age 10-18`, `Total Incidence Age 19-30`,
+    `Total Incidence Age >= 31`
+  ) %>%
+  pivot_longer(`Total Incidence Age 10-18`:`Total Incidence Age >= 31`,
+    names_to = "Age Group"
+  )
+i$name <- "Incidence"
+s <- screening %>%
+  filter(name == "Screened OGTT") %>%
+  select(ReviewYear, age_group, value) %>%
+  rename(Year = ReviewYear, `Age Group` = age_group)
+s$name <- " Screening Rates for CFRD"
+s$value <- s$value * 100
+t <- rbind(p, i, s)
+t$`Age Group` <- factor(t$`Age Group`,
+  levels = c(
+    "[10,19)", "[19,31)", "[31, Inf)", "Total Incidence Age >= 31",
+    "Total Incidence Age 10-18", "Total Incidence Age 19-30"
+  ),
+  labels = c(
+    "10-18", "19-30", ">= 31", ">= 31",
+    "10-18", "19-30"
+  )
+)
+t <- t[t$Year >= 2004, ]
+fig1 <- ggplot(t, aes(x = Year, y = value, color = `Age Group`)) +
+  geom_line() +
+  facet_wrap(~name, scales = "free", nrow = 3)
+fig1
+ggsave("./Dissemination/Figures/fig1.png",
+  plot = fig1, width = 9, height = 12, units = "in"
+)
 ```
@@
 t3
+t <- ggcoef_table(era3)
 ggcoef_model(era3,

### 2025-03-05 Lilah Melzer/cough_monitor_analysis.qmd
added baseline survey table summaries

@@
 ```{r}
+# Change reference group for easier interpretation
+# agg_counts$Diagnosis <- relevel(agg_counts$Diagnosis, ref = "CF")
+# Model selection
 zip1 <- mixed_model(Coughs ~ Night + Diagnosis,
@@
 
-So, instead of the model approach I did a more simple correlation between median coughs per night and median respiratory score.
-
-### CFQ-R Caregivers 6-13
-
-#### Median scores
-
-```{r}
-#| warning: false
-cfqr_parent_6_13$sid <- as.numeric(cfqr_parent_6_13$sid)
-caregivers <-
-  full_join(agg_counts, cfqr_parent_6_13, join_by(sid, Date)) %>%
-  select(sid, Sick, Night, Coughs, respirat) %>%
-  filter(!if_all(Coughs:respirat, ~ is.na(.)))
-caregivers_med <- caregivers %>%
+A total of `r length(unique(cfqr_children_6_11$sid))` people filled out the CFQ-R for children 6-11, with `r length(unique(cfqr_children_6_11$sid[duplicated(cfqr_children_6_11$sid)]))` filling it out more than once.
+
+A total of `r length(unique(cfqr_children_12_13$sid))` people filled out the CFQ-R for children 12-13, with `r length(unique(cfqr_children_12_13$sid[duplicated(cfqr_children_12_13$sid)]))` filling it out more than once.
+
+A total of `r length(unique(cfqr_teens_adults$sid))` people filled out the CFQ-R for teens and adults, with `r length(unique(cfqr_teens_adults$sid[duplicated(cfqr_teens_adults$sid)]))` filling it out more than once.
+
+A total of `r length(unique(cfqr_parent_6_13$sid))` people filled out the CFQ-R for teens and adults, with `r length(unique(cfqr_parent_6_13$sid[duplicated(cfqr_parent_6_13$sid)]))` filling it out more than once.
+
+## Summary of baseline values
+
+```{r results='asis'}
+# Combine CFQ-R into one
+cf1 <- cfqr_children_12_13 %>%
   group_by(sid) %>%
-  summarise(
-    `Median Coughs` = median(Coughs, na.rm = T),
-    `Median Resp. Score` = median(respirat, na.rm = T)
-  ) %>%
-  drop_na()
-M <- cor.test(caregivers_med$`Median Coughs`,
-  caregivers_med$`Median Resp. Score`,
-  method = "spearman"
-)
-ggplot(caregivers_med, aes(
-  y = `Median Coughs`, x = `Median Resp. Score`,
-  colour = factor(sid)
-)) +
-  geom_point() +
-  theme_bw()
-```
-
-Spearman's correlation between median coughs and median respiratory symptom score was `r round(as.numeric(M$estimate),3)` (p = `r format.pval(M$p.value,digits = 3,eps=0.001)`).
-
-#### Plots by participant
-
-```{r}
-#| message: false
-#| warning: false
-invisible(lapply(unique(caregivers_med$sid), function(s) {
-  d <- caregivers[caregivers$sid == s, ]
-  d <- d %>% pivot_longer(c(Coughs, respirat))
-  p <- ggplot(d, aes(x = Night, y = value, color = Sick)) +
-    geom_point() +
-    geom_line() +
-    theme_bw() +
-    scale_color_manual(values = c("#377EB8", "#E41A1C")) +
-    facet_wrap(~name, ncol = 1) +
-    ggtitle(paste0("Participant ", s))
-  cat("\n")
-  cat("\n")
-  print(p)
-  cat("\n")
-  cat("\n")
-}))
-```
-
-### CFQ-R Children Ages 12-13
-
-#### Median scores
-
-```{r}
-#| warning: false
-cfqr_children_12_13$sid <- as.numeric(cfqr_children_12_13$sid)
-children_12_13 <-
-  full_join(agg_counts, cfqr_children_12_13, join_by(sid, Date)) %>%
-  select(sid, Sick, Night, Coughs, respirat) %>%
-  filter(!if_all(Coughs:respirat, ~ is.na(.)))
-children_12_13_med <- children_12_13 %>%
+  filter(row_number() == 1) %>%
+  select(-Date) %>%
+  pivot_longer(physical:digest)
+cf2 <- cfqr_children_6_11 %>%
   group_by(sid) %>%
-  summarise(
-    `Median Coughs` = median(Coughs, na.rm = T),
-    `Median Resp. Score` = median(respirat, na.rm = T)
-  ) %>%
-  drop_na()
-M <- cor.test(children_12_13_med$`Median Coughs`,
-  children_12_13_med$`Median Resp. Score`,
-  method = "spearman"
-)
-ggplot(children_12_13_med, aes(
-  y = `Median Coughs`, x = `Median Resp. Score`,
-  colour = factor(sid)
-)) +
-  geom_point() +
-  theme_bw()
-```
-
-Spearman's correlation between median coughs and median respiratory symptom score was `r round(as.numeric(M$estimate),3)` (p = `r format.pval(M$p.value,digits = 3,eps=0.001)`).
-
-#### Plots by participant
-
-```{r}
-#| message: false
-#| warning: false
-invisible(lapply(unique(children_12_13_med$sid), function(s) {
-  d <- children_12_13[children_12_13$sid == s, ]
-  d <- d %>% pivot_longer(c(Coughs, respirat))
-  p <- ggplot(d, aes(x = Night, y = value, color = Sick)) +
-    geom_point() +
-    geom_line() +
-    theme_bw() +
-    scale_color_manual(values = c("#377EB8", "#E41A1C")) +
-    facet_wrap(~name, ncol = 1) +
-    ggtitle(paste0("Participant ", s))
-  cat("\n")
-  cat("\n")
-  print(p)
-  cat("\n")
-  cat("\n")
-}))
-```
-
-### CFQ-R Adolescents and Adults 14 Years
-
-#### Median scores
-
-```{r}
-#| warning: false
-cfqr_teens_adults$sid <- as.numeric(cfqr_teens_adults$sid)
-teens_adults <-
-  full_join(agg_counts, cfqr_teens_adults, join_by(sid, Date)) %>%
-  select(sid, Sick, Night, Coughs, respirat) %>%
-  filter(!if_all(Coughs:respirat, ~ is.na(.)))
-teens_adults_med <- teens_adults %>%
+  filter(row_number() == 1) %>%
+  select(-Date) %>%
+  pivot_longer(physical:digest)
+cf3 <- cfqr_teens_adults %>%
   group_by(sid) %>%
-  summarise(
-    `Median Coughs` = median(Coughs, na.rm = T),
-    `Median Resp. Score` = median(respirat, na.rm = T)
-  ) %>%
-  drop_na()
-M <- cor.test(teens_adults_med$`Median Coughs`,
-  teens_adults_med$`Median Resp. Score`,
-  method = "spearman"
-)
-ggplot(teens_adults_med, aes(
-  y = `Median Coughs`, x = `Median Resp. Score`,
-  colour = factor(sid)
-)) +
-  geom_point() +
-  theme_bw()
-```
-
-Spearman's correlation between median coughs and median respiratory symptom score was `r round(as.numeric(M$estimate),3)` (p = `r format.pval(M$p.value,digits = 3,eps=0.001)`).
-
-#### Plots by participant
-
-```{r}
-#| message: false
-#| warning: false
-invisible(lapply(unique(teens_adults_med$sid), function(s) {
-  d <- teens_adults[teens_adults$sid == s, ]
-  d <- d %>% pivot_longer(c(Coughs, respirat))
-  p <- ggplot(d, aes(x = Night, y = value, color = Sick)) +
-    geom_point() +
-    geom_line() +
-    theme_bw() +
-    scale_color_manual(values = c("#377EB8", "#E41A1C")) +
-    facet_wrap(~name, ncol = 1) +
-    ggtitle(paste0("Participant ", s))
-  cat("\n")
-  cat("\n")
-  print(p)
-  cat("\n")
-  cat("\n")
-}))
-```
-
-### Healthy controls
-
-#### Median scores
-
-```{r}
-#| warning: false
-hc_surveys$sid <- as.numeric(hc_surveys$sid)
-hcs <-
-  full_join(agg_counts, hc_surveys, join_by(sid, Date)) %>%
-  select(sid, Sick, Night, Coughs, q1:q5) %>%
-  filter(!if_all(q1:q5, ~ is.na(.)))
-hcs_med <- hcs %>%
+  filter(row_number() == 1) %>%
+  select(-Date) %>%
+  pivot_longer(physical:digest)
+cf4 <- cfqr_parent_6_13 %>%
   group_by(sid) %>%
-  summarise(
-    `Median Coughs` = median(Coughs, na.rm = T),
-    `Median Resp. Score` = median(as.numeric(q1), na.rm = T)
-  ) %>%
-  drop_na()
-M <- cor.test(hcs_med$`Median Coughs`,
-  hcs_med$`Median Resp. Score`,
-  method = "spearman"
-)
-ggplot(hcs_med, aes(
-  y = `Median Coughs`, x = `Median Resp. Score`,
-  colour = factor(sid)
-)) +
-  geom_point() +
-  theme_bw()
-```
-
-Spearman's correlation between median coughs and median respiratory symptom score was `r round(as.numeric(M$estimate),3)` (p = `r format.pval(M$p.value,digits = 3,eps=0.001)`).
+  filter(row_number() == 1) %>%
+  select(-Date) %>%
+  pivot_longer(physical:weight)
+# Put together then back to wide
+cfqr_baseline <- do.call(rbind, list(cf1, cf2, cf3, cf4))
+cfqr_baseline <- cfqr_baseline %>%
+  pivot_wider()
+t3 <- tableby(
+  Survey ~ notest(physical) + notest(emotion) + notest(social) + notest(eat) +
+    notest(body) + notest(treat) + notest(respirat) + notest(digest) +
+    notest(role) + notest(vitality) + notest(health) + notest(weight) +
+    notest(school),
+  data = cfqr_baseline, control = mycontrols
+)
+summary(t3)
+```
+
+# Healthy control surveys
+
+A total of `r length(unique(hc_surveys$sid))` controls filled out the surveys, with `r length(unique(hc_surveys$sid[duplicated(hc_surveys$sid)]))` filling it out more than once.
+
+## Summary of baseline values
+
+```{r results='asis'}
+hc <- hc_surveys %>%
+  group_by(sid) %>%
+  filter(row_number() == 1)
+t4 <- tableby(
+  ~ notest(q1) + notest(q2) + notest(q3) + notest(q4) +
+    notest(q5),
+  data = hc, control = mycontrols
+)
+summary(t4)
+```

### 2025-03-05 Lilah Melzer/cough_monitor_analysis.qmd
cough monitor plots finished

@@
 #| warning: false
-ggplot(agg_counts, aes(x = Coughs)) +
-  geom_histogram(binwidth = 1)
+hist <- ggplot(agg_counts, aes(x = Coughs)) +
+  geom_histogram() +
+  theme_bw() +
+  theme(axis.title.y = element_blank())
+hist
+ggsave("/home/timvigers/OneDrive/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Dissemination/Figures/histogram.png",
+  plot = hist, width = 9, height = 6, units = "in"
+)
 ```
@@
 
+## Association with coughs
+
+There were no statistically significant correlations between coughs and survey respiratory scores.
+
 # Healthy control surveys
@@
 ```
+
+## Association with coughs
+
+There were no statistically significant correlations between coughs and survey scores.
+
+# Participant figures
+
+```{r}
+#| warning: false
+cfqr_children_6_11$sid <- as.numeric(cfqr_children_6_11$sid)
+cfqr_parent_6_13$sid <- as.numeric(cfqr_parent_6_13$sid)
+p38 <- full_join(
+  agg_counts,
+  cfqr_children_6_11 %>% select(sid, Date, respirat) %>%
+    rename(`CFQ-R Children Ages 6-11` = respirat),
+  by = join_by(sid, Date)
+)
+p38 <- full_join(
+  p38,
+  cfqr_parent_6_13 %>% select(sid, Date, respirat) %>%
+    rename(`CFQ-R Caregivers 6-13` = respirat),
+  by = join_by(sid, Date)
+)
+p38 <- p38 %>%
+  filter(sid == 38) %>%
+  pivot_longer(c(Coughs, `CFQ-R Children Ages 6-11`, `CFQ-R Caregivers 6-13`)) %>%
+  filter(!is.na(value))
+p38$name <- factor(p38$name,
+  levels = c("Coughs", "CFQ-R Children Ages 6-11", "CFQ-R Caregivers 6-13")
+)
+p38_plot <- ggplot(p38, aes(x = Night, y = value)) +
+  geom_point() +
+  geom_line() +
+  theme_bw() +
+  theme(axis.title.y = element_blank()) +
+  facet_wrap(~name, nrow = 3, scales = "free")
+p38_plot
+ggsave("/home/timvigers/OneDrive/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Dissemination/Figures/participant_38.png",
+  plot = p38_plot, width = 9, height = 7, units = "in"
+)
+```

### 2025-03-06 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
cfrd survival check OGTTs and A1c

@@
 library(ggstats)
-knitr::opts_knit$set(root.dir = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry")
+knitr::opts_knit$set(root.dir = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry/")
 ```
@@
 #     (is.na(A_VX809comb) | A_VX809comb == "No"))
+```
+
+```{r}
+# Per reviewer, out of X individuals with a diagnosis of CFRD, X number had an
+# OGTT (and X had an A1c, but no OGTT) that year or the year prior.
+# This is super slow but idc idc
+t <- apply(first_cfrd, 1, function(r) {
+  id <- as.numeric(r["eDWID"])
+  year <- as.numeric(r["first_cfrd"])
+  d <- annual[annual$eDWID == id & annual$ReviewYear %in% c(year, year - 1), ]
+  if (any(!is.na(d$A_hgba1c))) {
+    a1c <- T
+  } else {
+    a1c <- F
+  }
+  if (any(!is.na(d["A_twohour_bloodglucose"]) |
+    !is.na(d["A_ogttfast_bloodglucose"]))) {
+    ogtt <- T
+  } else {
+    ogtt <- F
+  }
+  if (a1c & ogtt) {
+    return("both")
+  } else if (a1c & !ogtt) {
+    return("a1c")
+  } else if (ogtt & !a1c) {
+    return("ogtt")
+  } else if (!ogtt & !a1c) {
+    return("neither")
+  }
+})
 ```

### 2025-03-06 Lilah Melzer/cough_monitor_analysis.qmd
cough monitor tables

@@
   select(-height, -weight, -pre_fev)
+demo$nih_race <- droplevels(demo$nih_race)
+label(demo$nih_race) <- "Race"
 # Set table options for the document
@@
 #| tbl-cap: Participant characteristics
-f <- as.formula(paste0("Diagnosis ~", paste0(colnames(demo), collapse = "+")))
-f <- update(f, . ~ . - Diagnosis - sid)
-t1 <- tableby(f, data = demo, control = mycontrols)
+t1 <- tableby(Diagnosis ~ nih_sex + age_at_consent + height_perc + weight_perc +
+  nih_race + nih_ethnicity + cftr_mod + cftr_mod_type + Genotype + post_fev +
+  total_sleep, data = demo, control = mycontrols)
 summary(t1, pfootnote = T)

### 2025-03-06 Quarto Template.qmd
quarto template paths

@@
 library(redcapAPI)
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+)
 github_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/GitHub",
-  "Windows" = "C:/Users/timvigers/GitHub",
-  "Linux" = "/home/timvigers/GitHub"
+  "Darwin" = "/Users/timvigers/Documents/GitHub",
+  "Linux" = "/home/timvigers/Documents/GitHub"
 )
-knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF")
+knitr::opts_knit$set(root.dir = home_dir)
 ```
@@
 #| tbl-cap: Planets
+
 ```
@@
 #| fig-cap: Planets
+
 ```

### 2025-03-06 Lilah Melzer/cough_monitor_analysis.qmd
cough monitor saved figures

--- 2025-03-05 Lilah Melzer/cough_monitor_analysis.qmd
@@
 library(gt)
-knitr::opts_knit$set(root.dir = "/home/timvigers/OneDrive/Vigers/CF/Lilah Melzer/Lilah Cough Monitor")
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Lilah Melzer/Lilah Cough Monitor"
+)
+github_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Documents/GitHub",
+  "Linux" = "/home/timvigers/Documents/GitHub"
+)
+knitr::opts_knit$set(root.dir = home_dir)
 ```
@@
 label(demo$pre_fev_pp) <- "Pre-bronchodilator FEV1 % Predicted"
-label(demo$pre_fev_pp) <- "Post-bronchodilator FEV1 % Predicted"
+label(demo$post_fev_pp) <- "Post-bronchodilator FEV1 % Predicted"
 label(demo$height_perc) <- "Height %ile"
@@
 #| tbl-cap: Participant characteristics
-f <- as.formula(paste0("Diagnosis ~", paste0(colnames(demo), collapse = "+")))
-f <- update(f, . ~ . - Diagnosis - sid)
-t1 <- tableby(f, data = demo, control = mycontrols)
+t1 <- tableby(Diagnosis ~ nih_sex + age_at_consent + height_perc + weight_perc +
+  nih_race + nih_ethnicity + cftr_mod + cftr_mod_type + Genotype + pre_fev_pp +
+  total_sleep, data = demo, control = mycontrols)
 summary(t1, pfootnote = T)
@@
 #| warning: false
-ggplot(agg_counts, aes(x = Coughs)) +
-  geom_histogram(binwidth = 1)
+hist <- ggplot(agg_counts, aes(x = Coughs)) +
+  geom_histogram() +
+  theme_bw() +
+  theme(axis.title.y = element_blank())
+hist
+ggsave(paste0(home_dir, "/Dissemination/Figures/histogram.png"),
+  plot = hist, width = 9, height = 6, units = "in"
+)
 ```
@@
 #| warning: false
-ggplot(agg_counts[!is.na(agg_counts$Coughs), ], aes(
+fig2 <- ggplot(agg_counts[!is.na(agg_counts$Coughs), ], aes(
   x = Night, y = Coughs, group = factor(sid), color = Diagnosis
@@
   theme_bw()
+fig2
+ggsave(paste0(home_dir, "/Dissemination/Figures/fig2.png"),
+  plot = fig2, width = 9, height = 6, units = "in"
+)
 ```
@@
 ```
+
+## Association with coughs
+
+There were no statistically significant correlations between coughs and survey scores.
+
+# Participant figures
+
+```{r}
+#| warning: false
+cfqr_children_6_11$sid <- as.numeric(cfqr_children_6_11$sid)
+cfqr_parent_6_13$sid <- as.numeric(cfqr_parent_6_13$sid)
+p38 <- full_join(
+  agg_counts,
+  cfqr_children_6_11 %>% select(sid, Date, respirat) %>%
+    rename(`CFQ-R Children Ages 6-11` = respirat),
+  by = join_by(sid, Date)
+)
+p38 <- full_join(
+  p38,
+  cfqr_parent_6_13 %>% select(sid, Date, respirat) %>%
+    rename(`CFQ-R Caregivers 6-13` = respirat),
+  by = join_by(sid, Date)
+)
+p38 <- p38 %>%
+  filter(sid == 38) %>%
+  pivot_longer(c(Coughs, `CFQ-R Children Ages 6-11`, `CFQ-R Caregivers 6-13`)) %>%
+  filter(!is.na(value))
+p38$name <- factor(p38$name,
+  levels = c("Coughs", "CFQ-R Children Ages 6-11", "CFQ-R Caregivers 6-13")
+)
+p38_plot <- ggplot(p38, aes(x = Night, y = value)) +
+  geom_point() +
+  geom_line() +
+  theme_bw() +
+  theme(axis.title.y = element_blank()) +
+  facet_wrap(~name, nrow = 3, scales = "free")
+p38_plot
+ggsave(paste0(home_dir, "/Dissemination/Figures/participant_38.png"),
+  plot = p38_plot, width = 9, height = 7, units = "in"
+)
+```

### 2025-03-09 Christine Chan/CF Registry/cf_registry_CFRD_prediction.qmd
nonconformable errors

--- 2025-02-07 Christine Chan/CF Registry/cf_registry_CFRD_prediction.qmd
@@
 #| include: false
-library(tidyverse)
-library(arsenal)
-library(childsds)
 library(survival)
-library(ggsurvfit)
-library(ggstats)
-library(gtsummary)
-knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry")
+library(nlme)
+library(JM)
+library(dplyr)
+library(tidyr)
+knitr::opts_knit$set(root.dir = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry")
 ```
@@
 )
+# Exclude T1D and T2D
+annual <- annual %>% filter(!cfrd_status_annual %in% c(4, 5))
+annual$cfrd_registry <- as.numeric(annual$cfrd_status_annual == 3)
+annual$cfrd_ogtt <- as.numeric(annual$A_twohour_bloodglucose >= 200)
 # Demographic data
@@
 )
-# CFRD by OGTT two hour
-annual$cfrd_ogtt <- factor(annual$A_twohour_bloodglucose >= 200,
-  levels = c(F, T), labels = c("CFRD-", "CFRD+")
-)
-# CFRD by registry definition
-annual$cfrd_registry <- factor(annual$cfrd_status_annual,
-  levels = c(1:5), labels = c("CFRD-", "CFRD-", "CFRD+", "CFRD-", "CFRD-")
-)
 # Assume that missing data for CFRD means "CFRD-" (some of these will get filled in
 # later)
-annual$cfrd_registry[is.na(annual$cfrd_registry)] <- "CFRD-"
-annual$cfrd_ogtt[is.na(annual$cfrd_ogtt)] <- "CFRD-"
+annual$cfrd_registry[is.na(annual$cfrd_registry)] <- 0
+annual$cfrd_ogtt[is.na(annual$cfrd_ogtt)] <- 0
 # Modulators
@@
 )
-# Convert BMI to percentile (assume age > 20 is 20)
-annual$bmi_perc <- sds(annual$A_bmivalue,
-  age = ifelse(annual$Age_YrEnd > 20, 20, annual$Age_YrEnd),
-  sex = annual$Gender, item = "bmi", type = "perc", male = "M", female = "F",
-  ref = cdc.ref
-) * 100
-# Select columns
+# Get age at first CFRD (by OGTT and registry)
 annual <- annual %>%
-  select(
-    eDWID, ReviewYear, Race, Hispanicrace, Gender, F508, MutClass,
-    A_Vx770, A_VX809comb, Age_YrEnd, A_bmivalue, A_bmipercentile, bmi_perc,
-    A_FEV1_pct_predicted, NumPulmExacerbation, A_IsOnEnzymes,
-    A_fast_bloodglucose, A_hgba1c, cfrd_registry, cfrd_status_annual, cfrd_ogtt,
-    A_twohour_bloodglucose
-  ) %>%
   group_by(eDWID) %>%
   mutate(
-    Age_YrEnd = round(Age_YrEnd, 2),
-    cfrd_registry_age = first(na.omit(Age_YrEnd[cfrd_registry == "CFRD+"])),
-    cfrd_ogtt_age = first(na.omit(Age_YrEnd[cfrd_ogtt == "CFRD+"]))
+    cfrd_registry_age = first(na.omit(Age_YrEnd[cfrd_registry == 1])),
+    cfrd_ogtt_age = first(na.omit(Age_YrEnd[cfrd_ogtt == 1]))
   )
@@
 annual$cfrd_ogtt_age[is.na(annual$cfrd_ogtt_age)] <- Inf
-annual$cfrd_registry[annual$Age_YrEnd < annual$cfrd_registry_age] <- "CFRD-"
-annual$cfrd_registry[annual$Age_YrEnd >= annual$cfrd_registry_age] <- "CFRD+"
-annual$cfrd_ogtt[annual$Age_YrEnd < annual$cfrd_ogtt_age] <- "CFRD-"
-annual$cfrd_ogtt[annual$Age_YrEnd >= annual$cfrd_ogtt_age] <- "CFRD+"
-# Split into one dataset for registry and one for OGTT
-registry <- annual %>%
-  filter(Age_YrEnd <= cfrd_registry_age) %>%
-  select(eDWID:A_hgba1c, cfrd_registry)
-ogtt <- annual %>%
-  filter(Age_YrEnd <= cfrd_ogtt_age) %>%
-  select(eDWID:A_hgba1c, cfrd_ogtt, cfrd_ogtt_age)
-# Create start times
-registry$Age_YrStart <- registry$Age_YrEnd - 1
-ogtt$Age_YrStart <- ogtt$Age_YrEnd - 1
+annual$cfrd_registry[annual$Age_YrEnd < annual$cfrd_registry_age] <- 0
+annual$cfrd_registry[annual$Age_YrEnd >= annual$cfrd_registry_age] <- 1
+annual$cfrd_ogtt[annual$Age_YrEnd < annual$cfrd_ogtt_age] <- 0
+annual$cfrd_ogtt[annual$Age_YrEnd >= annual$cfrd_ogtt_age] <- 1
 # Format columns
@@
 )
+annual$A_IsOnEnzymes <- factor(annual$A_IsOnEnzymes,
+  levels = 1:0, labels = c("Yes", "No")
+)
+# Calculate bmi value
+annual$bmi <- annual$A_weight / ((annual$A_height / 100)^2)
+annual$bmi <- coalesce(annual$A_bmivalue, annual$bmi)
 # Labels
 labels <- list(
-  Age_YrStart = "Age at Year Start", Age_YrEnd = "Age at Year End",
+  Age_YrEnd = "Age at Year End",
   Hispanicrace = "Hispanic", A_Vx770 = "Use of CFTR modulator Vx770",
@@
   A_FEV1_pct_predicted = "FEV1 percent predicted", MutClass = "Mutation class",
-  NumPulmExacerbation = "Admissions for PEx"
+  NumPulmExacerbation = "Admissions for PEx",
+  A_IsOnEnzymes = "Pancreatic enzyme use", bmi = "BMI value"
 )
-# Basic formula
-xvars <- c(
-  "Gender", "Race", "Hispanicrace", "MutClass", "A_IsOnEnzymes",
-  "A_FEV1_pct_predicted", "NumPulmExacerbation", "A_corticosteroids1"
+# List of predictor variables
+predictors <- c(
+  "Gender", "Race", "Hispanicrace", "A_IsOnEnzymes",
+  "NumPulmExacerbation", "A_FEV1_pct_predicted", "bmi"
 )
-f <- ~ A_hgba1c + A_fast_bloodglucose + Gender + Race + Hispanicrace + MutClass +
-  A_IsOnEnzymes + A_FEV1_pct_predicted + NumPulmExacerbation
+# Make survival datasets for registry and OGTT diagnoses
+ogtt <- annual %>%
+  select(
+    eDWID, ReviewYear, Age_YrEnd, A_hgba1c, all_of(predictors), cfrd_ogtt
+  ) %>%
+  drop_na() %>%
+  distinct()
+surv_ogtt <- annual %>%
+  filter(Age_YrEnd <= cfrd_ogtt_age) %>%
+  group_by(eDWID) %>%
+  slice_max(row_number()) %>%
+  select(eDWID, ReviewYear, Age_YrEnd, all_of(predictors), cfrd_ogtt) %>%
+  drop_na() %>%
+  distinct()
+ogtt_include <- intersect(ogtt$eDWID, surv_ogtt$eDWID)
 ```
 
-# By registry CFRD status
+# Joint models
 
-## Participant demographics
-
-Categorical variables are from either the first year with CFRD or the most recent year for those without CFRD.
-
-```{r results='asis'}
-v1 <- registry %>%
-  group_by(eDWID) %>%
-  slice_max(ReviewYear)
-t1 <- tableby(cfrd_registry ~ Race + Hispanicrace + Gender + F508 + MutClass +
-  Age_YrEnd + A_bmipercentile + A_hgba1c, data = v1)
-summary(t1, pfootnote = T)
-```
-
-## Kaplan-Meier plots
+## CFRD diagnosis by OGTT
 
 ```{r}
-# Plot
-survfit2(Surv(Age_YrStart, Age_YrEnd, as.numeric(cfrd_registry)) ~ 1,
-  data = registry
-) %>%
-  ggsurvfit() +
-  labs(
-    x = "Age",
-    y = "Overall CFRD- probability"
-  ) +
-  add_confidence_interval() +
-  add_risktable()
-# Median "survival" time
-survfit2(Surv(Age_YrStart, Age_YrEnd, as.numeric(cfrd_registry)) ~ 1,
-  data = registry
-) %>%
-  tbl_survfit(
-    probs = 0.5,
-    label_header = "**Median time to CFRD+ (95% CI)**"
+# Formulas
+cox_f <- as.formula(paste0(
+  "Surv(time = Age_YrEnd, event = cfrd_ogtt)~",
+  paste0(predictors, collapse = "+")
+))
+lme_f <- as.formula(paste0(
+  "A_hgba1c~Age_YrEnd+",
+  paste0(predictors, collapse = "+")
+))
+# Fit Cox model first, and then pull the clusters
+cox_fit <- coxph(cox_f,
+  cluster = eDWID, x = T, model = T,
+  data = surv_ogtt[surv_ogtt$eDWID %in% ogtt_include, ]
 )
+lme_fit_a1c <- lme(lme_f,
+  random = ~ Age_YrEnd | eDWID, na.action = na.omit,
+  data = ogtt[ogtt$eDWID %in% ogtt_include, ]
+)
+# Combine
+joint_fit <- jointModel(lme_fit_a1c, cox_fit, timeVar = "Age_YrEnd")
 ```
-
-## Cox PH model
-
-```{r}
-cph <- coxph(
-  update(f, Surv(time = Age_YrStart, time2 = Age_YrEnd, event = as.numeric(cfrd_registry)==2) ~ .),
-  data = registry, cluster = eDWID
-)
-cph %>% tbl_regression(exponentiate = T)
-plot(cox.zph(cph), resid = F)
-cox.zph(cph)
-ggcoef_table(cph,
-  exponentiate = T, show_p_values = F, add_reference_rows = T, signif_stars = F,
-  colour = F
-)
-
-d = model.frame(cph)
-d$pred <- predict(cph, type = "survival")
-```
-
-# By CFRD status based on 2 hour OGTT >= 200
-
-## Participant demographics
-
-Categorical variables are from either the first year with CFRD or the most recent year for those without CFRD.
-
-```{r results='asis'}
-v1 <- ogtt %>%
-  group_by(eDWID) %>%
-  slice_max(ReviewYear)
-t1 <- tableby(cfrd_ogtt ~ Race + Hispanicrace + Gender + F508 + MutClass +
-  Age_YrEnd + A_bmipercentile + A_hgba1c, data = v1)
-summary(t1, pfootnote = T)
-```
-
-## Kaplan-Meier plots
-
-```{r}
-# Plot
-survfit2(Surv(Age_YrStart, Age_YrEnd, as.numeric(cfrd_ogtt)) ~ 1,
-  data = ogtt
-) %>%
-  ggsurvfit() +
-  labs(
-    x = "Age",
-    y = "Overall CFRD- probability"
-  ) +
-  add_confidence_interval() +
-  add_risktable()
-# Median "survival" time
-survfit2(Surv(Age_YrStart, Age_YrEnd, as.numeric(cfrd_ogtt)) ~ 1,
-  data = ogtt
-) %>%
-  tbl_survfit(
-    probs = 0.5,
-    label_header = "**Median time to CFRD+ (95% CI)**"
-  )
-```
-
-## Cox PH model
-
-```{r}
-cph <- coxph(
-  update(f, Surv(time = Age_YrStart, time2 = Age_YrEnd, event = as.numeric(cfrd_ogtt)) ~ .),
-  data = ogtt, cluster = eDWID
-)
-cph %>% tbl_regression(exponentiate = T)
-plot(cox.zph(cph), resid = F)
-cox.zph(cph)
-ggcoef_table(cph,
-  exponentiate = T, show_p_values = F, add_reference_rows = T, signif_stars = F,
-  colour = F
-)
-```

### 2025-03-13 Lilah Melzer/cough_monitor_analysis.qmd
new figures for Lilah

@@
 library(gt)
+library(ggbreak)
 home_dir <- switch(Sys.info()["sysname"],
@@
   theme_bw() +
-  theme(axis.title.y = element_blank())
+  theme(axis.title.y = element_blank()) +
+  facet_wrap(~Diagnosis, scales = "free_x")
 hist
@@
 )) +
-  geom_point(alpha = 0.1) +
-  geom_line(alpha = 0.1) +
+  geom_point(alpha = 0.2) +
+  geom_line(alpha = 0.2) +
   geom_smooth(aes(group = Diagnosis)) +
-  theme_bw()
+  ylim(c(1, 275)) +
+  xlim(c(0, 100)) +
+  theme_bw() +
+  scale_color_manual(values = c("#003092", "#FFAB5B")) +
+  scale_y_break(c(80, 265), scales = 0.25, ticklabels = c(265, 270, 275))
 fig2

### 2025-03-13 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
GCT fix pancreatic status for one person
EnVision demographics

@@
 library(reporttools)
-knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF")
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+)
+github_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Documents/GitHub",
+  "Linux" = "/home/timvigers/Documents/GitHub"
+)
+knitr::opts_knit$set(root.dir = home_dir)
 ```
@@
 )
+# Fix C088 pancreatic insufficiency
+df$`Pancreatic insufficient`[df$ID == "C088"] <- "Yes"
 # Drop levels

### 2025-03-13 Christine Chan/EnVision CF/christine_CGM_OGTT_category.qmd
EnVision demographics

@@
 library(ggplot2)
-knitr::opts_knit$set(root.dir = "/home/timvigers/OneDrive/Vigers/CF")
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+)
+github_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Documents/GitHub",
+  "Linux" = "/home/timvigers/Documents/GitHub"
+)
+knitr::opts_knit$set(root.dir = home_dir)
 ```
@@
   ungroup() %>%
-  select(
-    redcap_data_access_group, age_visit:origin_race, CFTR, weight
-  ) %>%
+  select(redcap_data_access_group:fvc, bmi_perc) %>%
   tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
@@
   ungroup() %>%
-  select(
-    redcap_data_access_group, age_visit:origin_race, CFTR, weight, scully140
-  ) %>%
+  select(redcap_data_access_group:fvc, bmi_perc, scully140) %>%
   tbl_summary(by = scully140, missing_text = "Missing") %>%
@@
   ungroup() %>%
-  select(
-    redcap_data_access_group, age_visit:origin_race, CFTR, weight, scully180
-  ) %>%
+  select(redcap_data_access_group:fvc, bmi_perc, scully180) %>%
   tbl_summary(by = scully180, missing_text = "Missing") %>%

### 2025-03-13 Christine Chan/EnVision CF/clean_envision_data.R

@@ -72,7 +72,7 @@ redcap <- read.csv(
 timepoints <- c(0, 10, 30, 60, 90, 120, 150, 180)
 glucose <- redcap %>%
   select(
-    study_id, date_visit, a1c_result, num_hospitalizations,
+    study_id, date_visit, a1c_result, 
     all_of(paste("timepoint", timepoints, "min", sep = "_"))
   ) %>%
   filter(!is.na(date_visit)) %>%
@@ -90,13 +90,14 @@ glucose$Date <- ymd(glucose$Date)
 #-------------------------------------------------------------------------------
 visits <- redcap %>%
   select(
-    study_id, date_visit, age_visit, height, weight, bmi_study_visit,
-    sex, origin_race, ethnicity, redcap_data_access_group,
-    cftr_mutation_1, cftr_mutation_2
+    study_id, redcap_data_access_group, sex, origin_race, ethnicity,
+    cftr_mutation_1, cftr_mutation_2, pancreatic_status, liver_disease,
+    pulm_ex_last3months, corrector_yes_no, family_history_diabetes,
+    date_visit, age_visit, height, weight, bmi_study_visit, fev1, fvc
   ) %>%
   filter(!is.na(date_visit)) %>%
   rename(Date = date_visit) %>%
-  filter(!if_all(age_visit:ethnicity, is.na))
+  filter(!if_all(age_visit:fvc, is.na))
 visits$Date <- ymd(visits$Date)
 #-------------------------------------------------------------------------------
 # Date fixes per Holly's 7/19 emails

### 2025-03-13 Christine Chan/EnVision CF/christine_CGM_OGTT_category.qmd
clean data for Katie to check

@@
 
+Participant `r df$study_id[which(df$scully140=="[17.5, Inf)"& df$Diagnosis=="NGT")]` were categorized as NGT by OGTT but had >=17.5% time over 140 on CGM.
+
 ## Time spent >180 mg/dL >=3.4% vs < 3.4%
@@
   tbl_summary(by = scully180, missing_text = "Missing") %>%
+  add_p() %>%
+  separate_p_footnotes() %>%
+  add_overall()
+```
+
+Participants `r df$study_id[which(df$scully180=="[3.4, Inf)"& df$Diagnosis=="NGT")]` were categorized as NGT by OGTT but had >=3.4% time over 180 on CGM.
+
+## CGM metrics by OGTT status
+
+```{r}
+df %>%
+  select(
+    Diagnosis, mean_sensor, standard_deviation, cv,
+    percent_time_over_140, percent_time_over_180, percent_time_70_140,
+    percent_time_70_180
+  ) %>%
+  tbl_summary(by = Diagnosis, missing_text = "Missing") %>%
   add_p() %>%
@@
 
+## CGM metrics by OGTT status
+
+```{r}
+df %>%
+  group_by(study_id) %>%
+  filter(row_number() == 1) %>%
+  ungroup() %>%
+  select(
+    Diagnosis, mean_sensor, standard_deviation, cv,
+    percent_time_over_140, percent_time_over_180, percent_time_70_140,
+    percent_time_70_180
+  ) %>%
+  tbl_summary(by = Diagnosis, missing_text = "Missing") %>%
+  add_p() %>%
+  separate_p_footnotes() %>%
+  add_overall()
+```
+
 # Figure
@@
 df %>%
+  filter(!is.na(Diagnosis)) %>%
   group_by(study_id) %>%

### 2025-03-13 Christine Chan/EnVision CF/clean_envision_data.R

@@ -7,7 +7,7 @@ library(lubridate)
 library(readxl)
 library(stringi)
 library(childsds)
-library(cgmanalysis)
+# library(cgmanalysis)
 library(pracma)
 # Home directory
 home_dir <- switch(Sys.info()["sysname"],
@@ -72,7 +72,7 @@ redcap <- read.csv(
 timepoints <- c(0, 10, 30, 60, 90, 120, 150, 180)
 glucose <- redcap %>%
   select(
-    study_id, date_visit, a1c_result, 
+    study_id, date_visit, a1c_result,
     all_of(paste("timepoint", timepoints, "min", sep = "_"))
   ) %>%
   filter(!is.na(date_visit)) %>%
@@ -90,10 +90,10 @@ glucose$Date <- ymd(glucose$Date)
 #-------------------------------------------------------------------------------
 visits <- redcap %>%
   select(
-    study_id, redcap_data_access_group, sex, origin_race, ethnicity,
+    study_id, redcap_data_access_group, date_visit, sex, origin_race, ethnicity,
     cftr_mutation_1, cftr_mutation_2, pancreatic_status, liver_disease,
     pulm_ex_last3months, corrector_yes_no, family_history_diabetes,
-    date_visit, age_visit, height, weight, bmi_study_visit, fev1, fvc
+    age_visit, height, weight, bmi_study_visit, fev1, fvc
   ) %>%
   filter(!is.na(date_visit)) %>%
   rename(Date = date_visit) %>%
@@ -301,19 +301,10 @@ final_df <- final_df %>%
   arrange(study_id, Date) %>%
   group_by(study_id) %>%
   mutate(ogtt_num = row_number()) %>%
-  ungroup() %>%
-  select(study_id, Date, ogtt_num, everything())
+  ungroup()
 final_df <- full_join(final_df, ogtts)
-final_df <- final_df %>% arrange(study_id, Date)
 # Add CGM
 final_df <- full_join(final_df, cgm)
-# Add HbA1c, etc.
-a1c <- redcap %>%
-  select(study_id, date_visit, a1c_result) %>%
-  rename(Date = date_visit) %>%
-  drop_na(a1c_result)
-a1c$Date <- ymd(a1c$Date)
-final_df <- full_join(final_df, a1c)
 #-------------------------------------------------------------------------------
 # Calculated fields
 #-------------------------------------------------------------------------------

### 2025-03-14 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
adult BMI values for GCT

@@
       dark: darkly
-bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Documents/GitHub/styles/american-medical-association.csl
+bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
+csl: /home/timvigers/Documents/GitHub/styles/american-medical-association.csl
 editor: source
@@
 )
+df$`BMI Value (Adults Only)` <- df$BMI
+df$`BMI Value (Adults Only)`[df$Age < 20] <- NA
 # Time from OGTT to GCT
@@
   kwt(`BMI %ile`, "N", "Nmiss2", "medianq1q3", "range") +
+  kwt(`BMI Value (Adults Only)`, "N", "Nmiss2", "medianq1q3", "range") +
   kwt(`Days From OGTT to GCT`, "N", "Nmiss2", "medianq1q3", "range") +

### 2025-03-18 Christine Chan/CF-SuGAR/power_calculations.qmd
CF-SuGAR power calculations

--- 2025-03-03 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
@@
 ---
-title: "CF Registry Analysis"
+title: "CF-SuGAR Power Calculations"
 author: "Tim Vigers"
@@
   html:
-    fig-width: 10
-    fig-height: 10
     toc: true
-    toc-depth: 10
+    toc-depth: 5
     toc-float: true
@@
     page-layout: full
+    theme:
+      light: flatly
+      dark: darkly
 editor: source
@@
 
-```{r}
+```{r setup}
 #| include: false
 library(Hmisc)
-library(zoo)
 library(tidyverse)
 library(childsds)
-library(gt)
+library(survival)
 library(gtsummary)
-library(forcats)
-library(labelled)
-library(survival)
-library(survminer)
-library(knitr)
-library(eulerr)
-library(broom)
-library(emmeans)
 library(ggstats)
-knitr::opts_knit$set(root.dir = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry")
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
+)
+github_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Documents/GitHub",
+  "Linux" = "/home/timvigers/Documents/GitHub"
+)
+knitr::opts_knit$set(root.dir = home_dir)
 ```
 
-# Data cleaning
-
-```{r}
+```{r data cleaning}
 # Import data and codebook
@@
 
-Registry fields included in our liver disease yes/no variable were:
+# Base model
 
-```{r}
-data.frame(data_dict$code_meaning[data_dict$var %in%
-  sub("A_", "", liver_fields)]) %>%
-  gt() %>%
-  tab_options(column_labels.hidden = TRUE)
-```
-
-Registry fields included in our tube feeding yes/no variable were:
-
-```{r}
-data.frame(data_dict$code_meaning[
-  paste0(data_dict$field_name, data_dict$code_value) %in%
-    sub("A_", "", tube_fields)
-]) %>%
-  gt() %>%
-  tab_options(column_labels.hidden = TRUE)
-```
-
-# CFRD prevalence
-
-## By year
-
-```{r}
-annual %>%
-  group_by(ReviewYear) %>%
-  summarise(Prevalence = mean(CFRD == "CFRD+") * 100) %>%
-  kable(., digits = 2)
-```
-
-## By year and age group
-
-```{r}
-annual %>%
-  group_by(ReviewYear, age_group) %>%
-  summarise(Prevalence = mean(CFRD == "CFRD+") * 100, .groups = "drop") %>%
-  pivot_wider(names_from = age_group, values_from = Prevalence) %>%
-  kable(., digits = 2)
-```
-
-# CFRD incidence
-
-```{r}
-# For each person, filter out all the years after first CFRD diagnosis
-annual <- annual %>% filter(ReviewYear <= first_cfrd | is.na(first_cfrd))
-# For each year, pull everyone with and without CFRD that year
-year_inc <- function(df = annual, year, py = 1000) {
-  # Filter and count
-  inc <- df %>% filter(ReviewYear == year)
-  inc <- inc %>%
-    group_by(CFRD) %>%
-    count(age_group, .drop = F)
-  inc <- inc[!is.na(inc$age_group), ]
-  # Calculations
-  new_cases <- sum(inc$n[inc$CFRD == "CFRD+"])
-  total_at_risk <- sum(inc$n)
-  total_inc <- new_cases / (total_at_risk / py)
-
-  new_cases_10_18 <- sum(inc$n[inc$CFRD == "CFRD+" &
-    inc$age_group == "[10,19)"])
-  at_risk_10_18 <- sum(inc$n[inc$age_group == "[10,19)"])
-  inc_10_18 <- new_cases_10_18 / (at_risk_10_18 / py)
-
-  new_cases_19_30 <- sum(inc$n[inc$CFRD == "CFRD+" &
-    inc$age_group == "[19,31)"])
-  at_risk_19_30 <- sum(inc$n[inc$age_group == "[19,31)"])
-  inc_19_30 <- new_cases_19_30 / (at_risk_19_30 / py)
-
-  new_cases_30 <- sum(inc$n[inc$CFRD == "CFRD+" &
-    inc$age_group == "[31, Inf)"])
-  at_risk_30 <- sum(inc$n[inc$age_group == "[31, Inf)"])
-  inc_30 <- new_cases_30 / (at_risk_30 / py)
-
-  # Nice results
-  list(
-    `Year` = year,
-    `Total New Cases` = new_cases,
-    `Total At Risk` = total_at_risk,
-    `Total Incidence` = total_inc,
-    `Total New Cases Age 10-18` = new_cases_10_18,
-    `Total At Risk Age 10-18` = at_risk_10_18,
-    `Total Incidence Age 10-18` = inc_10_18,
-    `Total New Cases Age 19-30` = new_cases_19_30,
-    `Total At Risk Age 19-30` = at_risk_19_30,
-    `Total Incidence Age 19-30` = inc_19_30,
-    `Total New Cases Age >= 31` = new_cases_30,
-    `Total At Risk Age >= 31` = at_risk_30,
-    `Total Incidence Age >= 31` = inc_30
-  )
-}
-incidence <- lapply(2003:2018, function(y) {
-  year_inc(year = y)
-})
-incidence <- data.frame(do.call(rbind, incidence), check.names = F)
-incidence <- data.frame(lapply(incidence, as.numeric), check.names = F)
-kable(incidence, digits = 2)
-```
-
-## Demographics of incident CFRD cases versus non-CFRD cases in 2004, 2011, and 2018
-
-```{r}
-#| message: false
-annual %>%
-  filter(ReviewYear %in% c(2004, 2011, 2018)) %>%
-  select(
-    ReviewYear, CFRD, Age_YrEnd, Gender, Race, Hispanicrace, bmi_perc,
-    bmi_cat, MutClass, tube_feeding, liver_disease,
-    A_FEV1_pct_predicted, A_FVC_pct_predicted, NumPulmExacerbation,
-    A_corticosteroids1, A_IsOnEnzymes, A_Vx770, A_VX809comb
-  ) %>%
-  tbl_strata(
-    strata = ReviewYear,
-    .tbl_fun =
-      ~ .x %>%
-        tbl_summary(
-          by = CFRD, missing_text = "Missing",
-          type = list(A_corticosteroids1 ~ "categorical"),
-          statistic = list(c(
-            "Age_YrEnd", "bmi_perc", "A_FEV1_pct_predicted",
-            "A_FVC_pct_predicted"
-          ) ~ "{mean} ({sd})")
-        ) %>%
-        add_p(test = list(c(
-          "Age_YrEnd", "bmi_perc", "A_FEV1_pct_predicted",
-          "A_FVC_pct_predicted"
-        ) ~ "t.test")) %>%
-        add_overall() %>%
-        separate_p_footnotes()
-  )
-```
-
-# Screening rates
-
-## By age group
-
-```{r}
-screening <- annual %>%
-  group_by(ReviewYear, age_group) %>%
-  summarise(
-    `Screened HbA1c` = mean(!is.na(A_hgba1c)),
-    `Screened OGTT` = mean(!is.na(A_twohour_bloodglucose) |
-      !is.na(A_ogttfast_bloodglucose)),
-    `Screened Either` = mean(!is.na(A_twohour_bloodglucose) |
-      !is.na(A_ogttfast_bloodglucose) | !is.na(A_hgba1c)),
-    .groups = "drop"
-  )
-kable(
-  screening %>%
-    select(ReviewYear, age_group, `Screened OGTT`) %>%
-    mutate(`Screened OGTT` = round(`Screened OGTT` * 100, 2)) %>%
-    pivot_wider(names_from = age_group, values_from = `Screened OGTT`),
-  caption = "% Screened by OGTT"
-)
-screening <- screening %>%
-  pivot_longer(`Screened HbA1c`:`Screened Either`)
-screening_plot <- ggplot(screening, aes(
-  x = ReviewYear, y = value,
-  color = age_group
-)) +
-  geom_line() +
-  xlab("Year") +
-  facet_wrap(~name, nrow = 3) +
-  theme_bw() +
-  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
-  scale_color_discrete(name = "Age Group") +
-  theme(text = element_text(size = 20), axis.title.y = element_blank())
-screening_plot
-ggsave("./Dissemination/Figures/screening_rates.png",
-  plot = screening_plot, width = 12, height = 9, units = "in"
-)
-```
-
-## Who is getting OGTTs?
-
-```{r results='asis'}
-# How many OGTTs are non-CFRD people getting 2016-2018
-ogtts <- annual %>%
-  filter(ReviewYear %in% 2016:2018, CFRD == "CFRD-") %>%
-  group_by(eDWID) %>%
-  summarise(
-    `Age Group` = last(na.omit(age_group)),
-    `OGTTs` = sum(!is.na(A_twohour_bloodglucose) |
-      !is.na(A_ogttfast_bloodglucose))
-  )
-# Add demographics
-ogtts$eDWID <- as.numeric(ogtts$eDWID)
-ogtts <- left_join(ogtts, demogs, by = join_by(eDWID))
-label(ogtts$eDWID) <- "ID"
-# Format OGTT counts
-ogtts$OGTTs <- factor(ogtts$OGTTs,
-  levels = 0:3, labels = c(
-    "No OGTT", "At least 1 OGTT", "At least 1 OGTT", "At least 1 OGTT"
-  )
-)
-# Make table for each variable
-age <- ogtts %>%
-  select(`Age Group`, OGTTs) %>%
-  tbl_summary(by = `Age Group`) %>%
-  add_ci() %>%
-  add_p() %>%
-  print()
-gender <- ogtts %>%
-  select(Gender, OGTTs) %>%
-  tbl_summary(by = Gender) %>%
-  add_ci() %>%
-  add_p() %>%
-  print()
-muts <- ogtts %>%
-  select(MutClass, OGTTs) %>%
-  tbl_summary(by = MutClass) %>%
-  add_ci() %>%
-  add_p() %>%
-  print()
-race <- ogtts %>%
-  select(Race, OGTTs) %>%
-  tbl_summary(by = Race) %>%
-  add_ci() %>%
-  add_p() %>%
-  print()
-eth <- ogtts %>%
-  select(Hispanicrace, OGTTs) %>%
-  tbl_summary(by = Hispanicrace) %>%
-  add_ci() %>%
-  add_p() %>%
-  print()
-```
-
-# Survival analysis - first CFRD diagnosis
-
-Based on CFFPR annualized data.
-These data are interval censored because we know the first visit at which people with CFRD were diagnosed, but we do not know the exact time that they developed CFRD.
-These data are also left truncated because not all participants enter the study at the same time.
-
-## Cox PH models
-
-### Base model
+- From the 2003 - 2018 registry data
 
@@
 ```
-
-### Vx770
-
-Years prior to 2011 were excluded from the following model
-
-```{r results='asis'}
-annual %>%
-  group_by(ReviewYear) %>%
-  count(A_Vx770) %>%
-  gt()
-```
-
-```{r results='asis'}
-f_vx770 <- update(f, . ~ . + A_Vx770)
-vx770_cox_res <- coxph(f_vx770,
-  data = annual[annual$ReviewYear >= 2011, ],
-  cluster = eDWID
-)
-t1 <- vx770_cox_res %>%
-  tbl_regression(exponentiate = T)
-t1
-ggcoef_table(vx770_cox_res,
-  show_p_values = F, add_reference_rows = T, exponentiate = T,
-  signif_stars = F, colour = F
-)
-```
-
-### Vx809comb
-
-Years prior to 2015 were excluded from the following model due to missing data.
-
-```{r results='asis'}
-annual %>%
-  group_by(ReviewYear) %>%
-  count(A_VX809comb) %>%
-  gt()
-```
-
-```{r}
-f_vx809comb <- update(f, . ~ . + A_VX809comb)
-vx809comb_cox_res <- coxph(f_vx809comb,
-  data = annual[annual$ReviewYear >= 2015, ],
-  cluster = eDWID
-)
-t2 <- vx809comb_cox_res %>%
-  tbl_regression(exponentiate = T)
-t2
-ggcoef_table(vx809comb_cox_res,
-  show_p_values = F, add_reference_rows = T, exponentiate = T,
-  signif_stars = F, colour = F
-)
-```
-
-### Combined results
-
-```{r}
-tbl_merge(
-  tbls = list(t1, t2),
-  tab_spanner = c("VX770", "VX809comb")
-)
-comb_plot <- ggcoef_compare(
-  list("VX770" = vx770_cox_res, "VX809comb" = vx809comb_cox_res),
-  add_reference_rows = T, exponentiate = T, type = "faceted", colour = F
-)
-comb_plot
-ggsave("./Dissemination/Figures/vx_combined_forest.png",
-  plot = comb_plot, width = 12, height = 9, units = "in"
-)
-```
-
-### Number of events by variable and year
-
-```{r results='asis'}
-invisible(lapply(c(xvars, "A_Vx770", "A_VX809comb"), function(v) {
-  cat("\n")
-  cat("#### By group")
-  cat("\n")
-  annual %>%
-    group_by_at(v) %>%
-    summarise(
-      n = length(unique(eDWID)),
-      `CFRD+` = sum(CFRD == "CFRD+"),
-      .groups = "drop"
-    ) %>%
-    gt() %>%
-    print()
-  cat("\n")
-  cat("#### By and year")
-  cat("\n")
-  annual %>%
-    group_by_at(c(v, "ReviewYear")) %>%
-    summarise(
-      n = length(unique(eDWID)),
-      `CFRD+` = sum(CFRD == "CFRD+"),
-      .groups = "drop"
-    ) %>%
-    gt() %>%
-    print()
-  cat("\n")
-}))
-```
-
-### Time varying coefficients
-
-#### 2003 - 2008
-
-```{r}
-t1 <- era1 %>% tbl_regression(exponentiate = T)
-t1
-ggcoef_model(era1,
-  show_p_values = F, add_reference_rows = T, exponentiate = T, shape_guide = F,
-  signif_stars = F, colour = "significance", shape = "reference_row",
-  shape_values = c(16, 15)
-)
-```
-
-#### 2009 - 2013
-
-```{r}
-t2 <- era2 %>% tbl_regression(exponentiate = T)
-t2
-ggcoef_model(era2,
-  show_p_values = F, add_reference_rows = T, exponentiate = T, shape_guide = F,
-  signif_stars = F, colour = "significance", shape = "reference_row",
-  shape_values = c(16, 15)
-)
-```
-
-#### 2014 - 2018
-
-```{r}
-t3 <- era3 %>% tbl_regression(exponentiate = T)
-t3
-ggcoef_model(era3,
-  show_p_values = F, add_reference_rows = T, exponentiate = T, shape_guide = F,
-  signif_stars = F, colour = "significance", shape = "reference_row",
-  shape_values = c(16, 15)
-)
-```
-
-#### Combined results
-
-```{r}
-tbl_merge(
-  tbls = list(t1, t2, t3),
-  tab_spanner = c("2003 - 2008", "2009 - 2013", "2014 - 2018")
-)
-comb_plot <- ggcoef_compare(
-  list("2003 - 2008" = era1, "2009 - 2013" = era2, "2014 - 2018" = era3),
-  add_reference_rows = T, exponentiate = T, shape_guide = F,
-  colour = "significance", shape = "reference_row", shape_values = c(16, 15),
-  type = "faceted"
-)
-comb_plot
-ggsave("./Dissemination/Figures/time_combined_forest.png",
-  plot = comb_plot, width = 12, height = 9, units = "in"
-)
-```

### 2025-03-18 Christine Chan/CF-SuGAR/power_calculations.qmd
added Laura's code

@@
 
-# Base model
+# Registry
 
-- From the 2003 - 2018 registry data
+## Base model
 
@@
 )
-# Diagnostics
-test.ph <- data.frame(cox.zph(cox_res)$table)
-test.ph <- test.ph[rownames(test.ph) != "GLOBAL", ]
-# Find which variables technically violate PH assumption
-time_varying <- rownames(test.ph)[test.ph$p < 0.05]
-# Just split data by time period so we're only fitting 3 models
-# "[2003,2009)", "[2009,2014)", "[2014,2019)"
-annual$time_period <- cut(annual$ReviewYear,
-  breaks = c(-Inf, 2009, 2014, Inf), right = F
+```
+
+## Base model plus HbA1c
+
+```{r}
+# Formula for base model
+cox_res <- coxph(update(f, . ~ . + A_hgba1c), data = annual, cluster = eDWID)
+# Forest plot and coefficients table
+cox_res %>%
+  tbl_regression(exponentiate = T)
+ggcoef_table(cox_res,
+  show_p_values = F, add_reference_rows = T, exponentiate = T,
+  signif_stars = F, colour = F
 )
-era1 <- coxph(f,
-  data = annual[as.numeric(annual$time_period) == 1, ],
-  cluster = eDWID
+```
+
+# 1 hour OGTT glucose in CFRD
+
+```{r}
+# read in separate data files
+
+ogtt <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/ogtt.csv")
+pft <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/pfts.csv")
+a1c <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/a1c.csv")
+demo <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/demographics.csv")
+dob <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/dob.csv")
+
+# merge dob and pfts
+pft <- merge(pft, dob, by = "MRN", all.x = T, all.y = F)
+pft <- unique(pft)
+
+# checking for dups
+# t <- ogtt[,c("MRN","FName","LName")]
+# t <- unique(t)
+# ct <- t %>% group_by(LName,FName) %>% count(unique(MRN))
+
+# merge ogtt and demo
+demo$FName <- NULL
+demo$LName <- NULL
+ogtt <- ogtt %>%
+  group_by(MRN, Date.of.OGTT) %>%
+  filter(row_number() == 1)
+ogtt_demo <- merge(ogtt, demo, by = "MRN", all.x = T, all.y = T)
+ogtt_demo$Date.of.OGTT <- mdy(ogtt_demo$Date.of.OGTT)
+
+a1c$EncounterDate <- as.Date(a1c$Date.of.A1C, format = "%m/%d/%Y")
+a1c$FName <- NULL
+a1c$LName <- NULL
+a1c$Date.of.OGTT <- NULL
+a1c$Days.Between.A1C.and.OGTT <- NULL
+ogtt_demo$EncounterDate <- ogtt_demo$Date.of.OGTT
+ogtt_demo_a1c <- merge(ogtt_demo, a1c, by = c("MRN", "EncounterDate"), all.x = T, all.y = T)
+
+# find most recent FVC that is within the 90 days prior to OGTT
+pft$EncounterDate <- as.Date(pft$Date.of.PFT, format = "%m/%d/%Y")
+pft$FName <- NULL
+pft$LName <- NULL
+alldata <- merge(ogtt_demo_a1c, pft, by = c("MRN", "EncounterDate"), all.x = T, all.y = T)
+
+# get predicted spirometry values
+alldata$FVC...Pred <- NULL
+alldata$FEV1...Pred <- NULL
+alldata$FEF.25.75...Pred <- NULL
+# dob <- pft[,c("MRN","DOB")]
+# dob <- unique(dob)
+# alldata <- merge(alldata,dob,by="MRN",all.x = T,all.y = F)
+alldata$DOB <- as.Date(as.character(alldata$DOB), format = "%m/%d/%Y")
+alldata$ht_m <- alldata$Height..cm. / 100
+alldata$sexnum <- ifelse(alldata$Sex == "Male", 1, 2)
+alldata$FEV1.pred <- pred_GLI(age = alldata$Age.at.PFT, height = alldata$ht_m, gender = alldata$sexnum, param = "FEV1")
+alldata$FVC.pred <- pred_GLI(age = alldata$Age.at.PFT, height = alldata$ht_m, gender = alldata$sexnum, param = "FVC")
+alldata$FEF.25.75.pred <- pred_GLI(age = alldata$Age.at.PFT, height = alldata$ht_m, gender = alldata$sexnum, param = "FEF2575")
+alldata$FEV1.per.pred <- (alldata$FEV1.Raw / alldata$FEV1.pred) * 100
+alldata$FVC.per.pred <- (alldata$FVC.Raw / alldata$FVC.pred) * 100
+alldata$FEF.25.75.per.pred <- (alldata$FEF.25.75.Raw / alldata$FEF.25.75.pred) * 100
+# alldata$Sex <- NULL
+
+# hard code some corrections to data per CC
+alldata[alldata$MRN == 686476, ]$X1.hour.Glucose <- ifelse(alldata[alldata$MRN == 686476, ]$Age.At.Test..years. == 14.5, 185, alldata[alldata$MRN == 686476, ]$X1.hour.Glucose)
+
+# severity of genotype
+
+# Red = severe or minimal function mutation
+# Orange= unknown mutation status (for now)
+# Yellow= partial function or mild
+# Green = not CF causing, classify as "other", will most likely be seen in people with CRMS diagnosis
+
+alldata$class2 <- ifelse(alldata$Mutation.2 %in% c(
+  "1717-1G->A", "1717-1G-A", "G542X", "Q493X", "R1162X", "R553X", "R553x", "W1089X", "W1282X", "1078delT", "R75X",
+  "3659delC", "621+1G->T", "621+1G>T", "394delTT", "3120 + 1g->A", "1154InsTC", "1154insTC", "1213delT",
+  "1259insA", "1288insTA", "3791delC", "E60X", "K710X", "2184delA", "CFTRdele2,3", "663delT", "Glu528",
+  "1461ins4", "306insA", "R709X", "CFTRdele22-24", "2711delT", "2183AA- >G", "2183delAA>G"
+), "I",
+ifelse(alldata$Mutation.2 %in% c("F508", "F508del", "F508del ", "G85E", "I507", "I 507", "N1303K", "A559T", "R560T", "A561E"),
+  "II",
+  ifelse(alldata$Mutation.2 %in% c("G551D", "G551S", "S549N"), "III",
+    ifelse(alldata$Mutation.2 %in% c("R334W", "R347P", "R117H", "R117C", "P67L", "L206W", "l206w", "I206w", "D614G", "R347H", "D1152H", "3849+10kbC>T"), "IV",
+      ifelse(alldata$Mutation.2 %in% c(
+        "A455E", "2789+5G->A", "3849+10CT", "3849+10kbC->T", "711+3A->G", "1898+5G->T",
+        "P574H", "3272-26A->G5", "3272-26A->G"
+      ), "V", "")
     )
-era2 <- coxph(f,
-  data = annual[as.numeric(annual$time_period) == 2, ],
-  cluster = eDWID
   )
-era3 <- coxph(f,
-  data = annual[as.numeric(annual$time_period) == 3, ],
-  cluster = eDWID
 )
-# # KM plot
-# fit <- survfit(Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2) ~ age_group,
-#   data = annual[as.numeric(annual$time_period) == 1, ]
-# )
-# # Visualize with survminer
-# ggsurvplot(fit, data = annual[as.numeric(annual$time_period) == 1, ], risk.table = TRUE)
+)
+
+alldata$severity2 <- ifelse(alldata$Mutation.2 %in% c(
+  "1717-1G->A", "1717-1G-A", "G542X", "Q493X", "R1162X", "R553X", "R553x", "W1089X", "W1282X", "1078delT", "R75X",
+  "3659delC", "621+1G->T", "621+1G>T", "394delTT", "3120 + 1g->A", "1154InsTC", "1154insTC", "1213delT",
+  "1259insA", "1288insTA", "3791delC", "E60X", "K710X", "2184delA", "CFTRdele2,3", "663delT", "Glu528",
+  "1461ins4", "306insA", "R709X", "CFTRdele22-24", "2711delT", "2183AA- >G", "2183delAA>G", "F508", "F508del", "F508del ", "G85E",
+  "I507", "I 507", "N1303K", "A559T", "R560T", "A561E", "G551D", "G551S", "S549N", "1898+1G->A", "2143delT", "2183AA-G",
+  "2183delAA->G", "2184insA", "S434X", "2585delT", "3120+1G->A", "3905insT", "406-1G->A", "712-1G->T", "L1245X", "Q2X", "R1158X",
+  "R851X", "V520F", "L1077P", "S489X", "F311del", "W846X", "M1101K", "2622+1G->A", "S945L", "Y1092X", "R1066C", "G178R", "2957delT",
+  "CFTRdele1", "W1204X", "q493x", "W57G", "1585-8G>A", "1717-8G->A", "G1244E"
+), "Severe",
+ifelse(alldata$Mutation.2 %in% c(
+  "R334W", "R347P", "R117H", "R117C", "P67L", "L206W", "l206w", "I206w", "D614G", "R347H", "D1152H", "3849+10kbC>T", "D110H", "R1070W",
+  "A455E", "2789+5G->A", "3849+10CT", "3849+10kbC->T", "711+3A->G", "1898+5G->T", "P574H", "3272-26A->G5", "3272-26A->G", "3849G->A", "H199Y", "H199y",
+  "S492F", "S531P", "L1335P", "R352Q", "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val",
+  "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"
+), "Mild",
+ifelse(alldata$Mutation.2 %in% c(
+  "1262insA", "I506T", "3363delGT", "Q1209P", "3349insT", "G178R", "L467P", "317insC", "I336K",
+  "ex14a", "3500-2A->G", "c.3407_3422del16", "3349insT", "1248+1G->A", "M470V", "1812-1G->A",
+  "del X22-24", "Q237H", "Q237H", "W57G", "1078delA", "G178R", "2957delT", "G576A",
+  "other", "Other", "Unknown", "unknown", "CFTRdele1", "1811+1.6kb A->", "UNK", "I336K", "M470V", "S1235R", "M470V", "F508C",
+  "R1066H", "P205S"
+), "Unk", "")
+)
+)
+
+alldata$class1 <- ifelse(alldata$Mutation.1 %in% c(
+  "1717-1G->A", "1717-1G-A", "G542X", "Q493X", "R1162X", "R553X", "R553x", "W1089X", "W1282X", "1078delT", "R75X",
+  "3659delC", "621+1G->T", "621+1G>T", "394delTT", "3120 + 1g->A", "1154InsTC", "1154insTC", "1213delT",
+  "1259insA", "1288insTA", "3791delC", "E60X", "K710X", "2184delA", "CFTRdele2,3", "663delT", "Glu528",
+  "1461ins4", "306insA", "R709X", "CFTRdele22-24", "2711delT", "2183AA- >G", "2183delAA>G"
+), "I",
+ifelse(alldata$Mutation.1 %in% c("F508", "F508del", "F508del ", "G85E", "I507", "I 507", "N1303K", "A559T", "R560T", "A561E"), "II",
+  ifelse(alldata$Mutation.1 %in% c("G551D", "G551S", "S549N"), "III",
+    ifelse(alldata$Mutation.1 %in% c("R334W", "R347P", "R117H", "R117C", "P67L", "L206W", "l206w", "I206w", "D614G", "R347H", "D1152H", "3849+10kbC>T"), "IV",
+      ifelse(alldata$Mutation.1 %in% c(
+        "A455E", "2789+5G->A", "3849+10CT", "3849+10kbC->T", "711+3A->G", "1898+5G->T",
+        "P574H", "3272-26A->G5", "3272-26A->G"
+      ), "V", "")
+    )
+  )
+)
+)
+
+alldata$severity1 <- ifelse(alldata$Mutation.1 %in% c(
+  "1717-1G->A", "1717-1G-A", "G542X", "Q493X", "R1162X", "R553X", "R553x", "W1089X", "W1282X", "1078delT", "R75X",
+  "3659delC", "621+1G->T", "621+1G>T", "394delTT", "3120 + 1g->A", "1154InsTC", "1154insTC", "1213delT",
+  "1259insA", "1288insTA", "3791delC", "E60X", "K710X", "2184delA", "CFTRdele2,3", "663delT", "Glu528",
+  "1461ins4", "306insA", "R709X", "CFTRdele22-24", "2711delT", "2183AA- >G", "2183delAA>G", "F508", "F508del", "F508del ", "G85E",
+  "I507", "I 507", "N1303K", "A559T", "R560T", "A561E", "G551D", "G551S", "S549N", "1898+1G->A", "2143delT", "2183AA-G",
+  "2183delAA->G", "2184insA", "S434X", "2585delT", "3120+1G->A", "3905insT", "406-1G->A", "712-1G->T", "L1245X", "Q2X", "R1158X",
+  "R851X", "V520F", "L1077P", "S489X", "F311del", "W846X", "M1101K", "2622+1G->A", "S945L",
+  "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", "W1204X", "q493x", "W57G", "1585-8G>A", "1717-8G->A", "G1244E"
+), "Severe",
+ifelse(alldata$Mutation.1 %in% c(
+  "R334W", "R347P", "R117H", "R117C", "P67L", "L206W", "l206w", "I206w", "D614G", "R347H", "D1152H", "3849+10kbC>T", "D110H", "R1070W",
+  "A455E", "2789+5G->A", "3849+10CT", "3849+10kbC->T", "711+3A->G", "1898+5G->T",
+  "P574H", "3272-26A->G5", "3272-26A->G", "3849G->A", "H199Y", "H199y", "S492F", "S531P",
+  "L1335P", "R352Q", "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val",
+  "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"
+), "Mild",
+ifelse(alldata$Mutation.1 %in% c(
+  "1262insA", "I506T", "3363delGT", "Q1209P", "3349insT", "G178R", "L467P", "317insC", "I336K",
+  "ex14a", "3500-2A->G", "c.3407_3422del16", "3349insT", "1248+1G->A", "M470V", "1812-1G->A",
+  "del X22-24", "Q237H", "Q237H", "W57G", "1078delA", "G178R", "2957delT", "G576A",
+  "other", "Other", "Unknown", "unknown", "CFTRdele1", "1811+1.6kb A->", "UNK", "I336K", "M470V", "S1235R", "M470V", "F508C",
+  "R1066H", "P205S"
+), "Unk", "")
+)
+)
+
+# 3 patients have severity2="Severe" per Christine
+alldata$severity2 <- ifelse(alldata$MRN %in% c(546678, 664400, 2015506), "Severe", alldata$severity2)
+
+alldata$genoRisk <- ifelse(alldata$severity1 == "Mild" | alldata$severity2 == "Mild", "Low",
+  ifelse(alldata$severity1 == "Severe" & alldata$severity2 == "Severe", "High", "")
+)
+
+# write file of people with missing genotype risk
+# write.csv(alldata[alldata$genoRisk=="",],"./Christine Chan/1 hour OGTT in CFRD/Data_raw/missing_genotype_risk.csv")
+
+# merge back all baseline characteristics
+d <- alldata[, c("MRN", "FName", "LName", "Sex", "Race", "Latino.a.", "Mutation.1", "Mutation.2", "class1", "class2", "severity1", "severity2", "genoRisk")]
+d <- d[!is.na(d$FName), ]
+d <- unique(d)
+alldata$FName <- NULL
+alldata$Latino.a. <- NULL
+alldata$LName <- NULL
+alldata$Sex <- NULL
+alldata$Race <- NULL
+alldata$Mutation.1 <- NULL
+alldata$Mutation.2 <- NULL
+alldata$class1 <- NULL
+alldata$class2 <- NULL
+alldata$severity1 <- NULL
+alldata$severity2 <- NULL
+alldata$genoRisk <- NULL
+alldata <- merge(alldata, d, by = "MRN", all.x = T, all.y = T)
+
+# checking for dups
+# t <- alldata[,c("MRN","FName","LName")]
+# t <- unique(t)
+# ct <- t %>% group_by(LName,FName) %>% count(unique(MRN))
+
+# find first OGTT
+alldata$Date <- alldata$EncounterDate
+o <- alldata[!is.na(alldata$X1.hour.Glucose), c("MRN", "Date")]
+o <- o[order(o$MRN, o$Date), ]
+o <- o %>%
+  group_by(MRN) %>%
+  filter(row_number() == 1)
+colnames(o) <- c("MRN", "Date of first OGTT")
+alldata <- merge(alldata, o, by = "MRN", all.x = T, all.y = T)
+alldata <- alldata[!is.na(alldata$`Date of first OGTT`), ]
+
+# checking ppts that Collin said had CFRD
+col_check <- alldata[alldata$MRN %in% c(
+  1006479, 1053930, 1078408, 1136992, 1289450, 1339023, 1413821, 1695512, 664400,
+  686476, 707831, 712676, 739231, 739460, 787372, 800007, 848440, 860181, 875128, 897478, 905284, 908966,
+  918124, 932695, 949042, 952250, 952252, 825118
+), ]
+col_check <- col_check[order(col_check$MRN, col_check$Date), ]
+# write.csv(col_check,"./Christine Chan/1 hour OGTT in CFRD/Data_raw/check_cfrd.csv",na="",row.names = F)
+
+# last 2 discrepant cases after fixing date issue
+col_check2 <- alldata[alldata$MRN %in% c(686476, 825118), ]
+col_check2 <- col_check2[order(col_check2$MRN, col_check2$Date), ]
+# write.csv(col_check2,"./Christine Chan/1 hour OGTT in CFRD/Data_raw/check_cfrd2.csv",na="",row.names = F)
+
+# remove all visits before first OGTT
+# alldata$`Date of first OGTT` <- as.Date(as.character(alldata$`Date of first OGTT`),format="%m/%d/%Y")
+alldata <- alldata[alldata$Date >= alldata$`Date of first OGTT`, ]
+alldata <- alldata %>% arrange(MRN, EncounterDate)
+
+# calculate BMI
+alldata$BMI <- alldata$Weight..kg. / ((alldata$Height..cm. / 100)^2)
+
+# reorder columns
+colorder <- c(
+  "MRN", "FName", "LName", "Sex", "Race", "Latino.a.", "Mutation.1", "Mutation.2", "class1", "class2", "severity1", "severity2", "genoRisk",
+  "Date of first OGTT", "Date", "Age.At.Test..years.", "Fasting.Glucose", "X1.hour.Glucose", "X2.hour.Glucose", "A1C",
+  "DOB", "Age.at.PFT", "FVC.Raw", "Height..cm.", "Weight..kg.", "FEV1.Raw", "FEF.25.75.Raw", "ht_m",
+  "sexnum", "FEV1.pred", "FVC.pred", "FEF.25.75.pred", "FEV1.per.pred", "FVC.per.pred", "FEF.25.75.per.pred", "BMI"
+)
+alldata <- alldata[, colorder]
+alldata <- alldata[order(alldata$MRN, alldata$Date), ]
+
+# create combined race/ethnicity
+alldata$race_eth <- ifelse(alldata$Latino.a. == "TRUE", "Hispanic", alldata$Race)
+
+# get BMI percentile
+alldata$age <- ifelse(is.na(alldata$Age.at.PFT), alldata$Age.At.Test..years., alldata$Age.at.PFT)
+alldata$age_bmi <- ifelse(alldata$age > 20, 20, alldata$age)
+alldata$bmip <- sds(alldata$BMI,
+  age = alldata$age_bmi,
+  sex = alldata$Sex, male = "Male", female = "Female",
+  ref = cdc.ref,
+  item = "bmi",
+  type = "perc"
+)
+# for demographics data, multiply BMI percentile by 100
+alldata$bmipercentile <- alldata$bmip * 100
+
+# classify CFRD at each visit
+alldata$cfrd_visit <- NA
+alldata$cfrd_visit <- ifelse(is.na(alldata$X2.hour.Glucose), NA, ifelse(alldata$X2.hour.Glucose >= 200, 1, 0))
+
+# find people with CFRD at baseline
+cfrdbase <- alldata[alldata$Date == alldata$`Date of first OGTT` & alldata$cfrd_visit == 1, ]
+before_cfrd <- length(unique(alldata$MRN))
+alldata <- alldata[!(alldata$MRN %in% cfrdbase$MRN), ]
+after_cfrd <- length(unique(alldata$MRN))
+
+# number excluded with CFRD at baseline
+nexcl_cfrd <- before_cfrd - after_cfrd
+
+# create variable for CFRD groups (i.e., those who developed CFRD vs those who did not)
+# and get date of first OGTT with CFRD
+cfrdcheck <- alldata[alldata$cfrd_visit == 1, c("MRN", "Date", "cfrd_visit")]
+cfrdcheck <- cfrdcheck[order(cfrdcheck$MRN, desc(cfrdcheck$cfrd_visit)), ]
+cfrdcheck <- cfrdcheck[!is.na(cfrdcheck$MRN), ]
+cfrdcheck <- cfrdcheck %>%
+  group_by(MRN) %>%
+  filter(row_number() == 1)
+colnames(cfrdcheck) <- c("MRN", "Date of CFRD diagnosis", "CFRD")
+alldata <- merge(alldata, cfrdcheck, by = "MRN", all.x = T, all.y = T)
+alldata$CFRD <- ifelse(is.na(alldata$CFRD), 0, 1)
+
+# get baseline age
+baseage <- alldata[alldata$Date == alldata$`Date of first OGTT`, c("MRN", "age")]
+colnames(baseage) <- c("MRN", "Base_age")
+alldata <- merge(alldata, baseage, by = "MRN", all.x = T, all.y = T)
+
+# get baseline 1 hour and 2 hour glucose
+basegluc <- alldata[alldata$Date == alldata$`Date of first OGTT`, c("MRN", "X1.hour.Glucose", "X2.hour.Glucose")]
+colnames(basegluc) <- c("MRN", "Baseline_1hr_glucose", "Baseline_2hr_glucose")
+alldata <- merge(alldata, basegluc, by = "MRN", all.x = T, all.y = T)
+
+# dichotomize baseline 1 hour glucose
+alldata$Baseline_1hr_glucose_155 <- ifelse(alldata$Baseline_1hr_glucose >= 155, 1, 0)
+alldata$Baseline_1hr_glucose_140 <- ifelse(alldata$Baseline_1hr_glucose >= 140, 1, 0)
+
+# dichotomize baseline 2 hour glucose
+alldata$Baseline_2hr_glucose_140 <- ifelse(alldata$Baseline_2hr_glucose >= 140, 1, 0)
+
+# find date of last OGTT
+lastogtt <- alldata[
+  !is.na(alldata$Fasting.Glucose) | !is.na(alldata$X1.hour.Glucose) | !is.na(alldata$X2.hour.Glucose),
+  c("MRN", "Date")
+]
+lastogtt <- lastogtt[order(lastogtt$MRN, desc(lastogtt$Date)), ]
+lastogtt <- lastogtt %>%
+  group_by(MRN) %>%
+  filter(row_number() == 1)
+colnames(lastogtt) <- c("MRN", "Date of last OGTT")
+alldata <- merge(alldata, lastogtt, by = "MRN", all.x = T, all.y = T)
+
+# read in date of modulator use
+mod <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/modulators.csv", na.strings = c("", " ", NA))
+mod$Kalydeco.Start.Date <- as.Date(mod$Kalydeco.Start.Date, format = "%m/%d/%Y")
+mod$Orkambi.Start.Date <- as.Date(mod$Orkambi.Start.Date, format = "%m/%d/%Y")
+mod$Symdeko.Start.Date <- as.Date(mod$Symdeko.Start.Date, format = "%m/%d/%Y")
+mod$Trikafta.Start.Date <- as.Date(mod$Trikafta.Start.Date, format = "%m/%d/%Y")
+mod$modstart <- as.Date(NA)
+mod$he_modstart <- as.Date(NA)
+mod$me_modstart <- as.Date(NA)
+for (i in 1:nrow(mod)) {
+  mod[i, ]$modstart <-
+    as.Date(pmin(mod[i, ]$Kalydeco.Start.Date, mod[i, ]$Orkambi.Start.Date, mod[i, ]$Symdeko.Start.Date, mod[i, ]$Trikafta.Start.Date, na.rm = T))
+  mod[i, ]$he_modstart <- as.Date(pmin(mod[i, ]$Kalydeco.Start.Date, mod[i, ]$Trikafta.Start.Date, na.rm = T))
+  mod[i, ]$me_modstart <- as.Date(pmin(mod[i, ]$Orkambi.Start.Date, mod[i, ]$Symdeko.Start.Date, na.rm = T))
+}
+modkeep <- mod[, c("MRN", "modstart", "he_modstart", "me_modstart")]
+alldata <- merge(alldata, modkeep, by = "MRN", all.x = TRUE, all.y = FALSE)
+alldata$onmod <- ifelse(!is.na(alldata$modstart) & alldata$modstart <= alldata$Date, 1,
+  ifelse(!is.na(alldata$modstart) & alldata$modstart > alldata$Date, 0, NA)
+)
+alldata$onmod <- ifelse(is.na(alldata$modstart) & is.na(alldata$onmod), 0, alldata$onmod)
+
+alldata$onhemod <- ifelse(!is.na(alldata$he_modstart) & alldata$he_modstart <= alldata$Date, 1,
+  ifelse(!is.na(alldata$he_modstart) & alldata$he_modstart > alldata$Date, 0, NA)
+)
+alldata$onhemod <- ifelse(is.na(alldata$he_modstart) & is.na(alldata$onhemod), 0, alldata$onhemod)
+
+alldata$onmemod <- ifelse(!is.na(alldata$me_modstart) & alldata$me_modstart <= alldata$Date, 1,
+  ifelse(!is.na(alldata$me_modstart) & alldata$me_modstart > alldata$Date, 0, NA)
+)
+alldata$onmemod <- ifelse(is.na(alldata$me_modstart) & is.na(alldata$onmemod), 0, alldata$onmemod)
+
+alldata$FEV.1.per.pred_FVC.per.pred <- alldata$FEV1.per.pred / alldata$FVC.per.pred
+
+# for analyses about prediction of CFRD, we need to exclude those who only have one OGTT
+# can use alldata for mixed models of PFTs and BMI
+oneogtt <- alldata[alldata$`Date of first OGTT` == alldata$`Date of last OGTT`, ]
+multogtt <- alldata[!(alldata$MRN %in% oneogtt$MRN), ]
+
+# have CFRD indicator
+# create variable for follow-up time
+# in those who do not develop CFRD, this is the time between first and last OGTT, and they are censored
+# for those who develop CFRD, this is time between first OGTT and diagnosis of CFRD
+multogtt$Follow_up_time_yrs <- ifelse(multogtt$CFRD == 1, (multogtt$`Date of CFRD diagnosis` - multogtt$`Date of first OGTT`) / 365.25,
+  (multogtt$`Date of last OGTT` - multogtt$`Date of first OGTT`) / 365.25
+)
+
+# get age at diagnosis of CFRD
+multogtt$age_at_CFRD_dx <- ifelse(multogtt$CFRD == 1, (multogtt$`Date of CFRD diagnosis` - multogtt$DOB) / 365.25, NA)
+
+# create categorical time
+multogtt$cat_time <- as.factor(round_any(multogtt$Follow_up_time_yrs, 0.5))
+# need a time variable for mixed models
+multogtt$time <- as.numeric((multogtt$Date - multogtt$`Date of first OGTT`) / 365.25)
+multogtt$cat_time2 <- as.factor(round_any(multogtt$time, 0.5))
+
+# get number of participants in each cohort
+n_alldata <- length(unique(alldata$MRN))
+n_multogtt <- length(unique(multogtt$MRN))
+# number of participants in each cohort who developed CFRD
+n_alldata_cfrd <- length(unique(alldata[alldata$CFRD == 1, ]$MRN))
+n_multogtt_cfrd <- length(unique(multogtt[multogtt$CFRD == 1, ]$MRN))
+
+# create dataset with baseline only
+baseline_mult <- multogtt[multogtt$Date == multogtt$`Date of first OGTT`, ]
+baseline_mult$onmod_baseline <- as.factor(baseline_mult$onmod)
+baseline_all <- alldata[alldata$Date == alldata$`Date of first OGTT`, ]
+baseline_all$onmod_baseline <- as.factor(baseline_all$onmod)
+# find patients who were put on modulators during follow-up
+evermod <- alldata %>%
+  group_by(MRN) %>%
+  summarise(max(onmod))
+baseline_all <- merge(baseline_all, evermod, by = "MRN", all.x = T, all.y = T)
+baseline_all$modstarted_fup <- as.factor(ifelse(baseline_all$onmod_baseline == 0 & baseline_all$`max(onmod)` == 1, 1, 0))
+baseline_mult <- merge(baseline_mult, evermod, by = "MRN", all.x = T, all.y = F)
+baseline_mult$modstarted_fup <- as.factor(ifelse(baseline_mult$onmod_baseline == 0 & baseline_mult$`max(onmod)` == 1, 1, 0))
+
+################## FILL IN MISSING A1c, BMI, PFT AT BASELINE
+
+# find most recent A1c that is within the 90 days prior to OGTT or up to 14 days after OGTT
+a1c$Date.of.A1C <- as.Date(a1c$Date.of.A1C, format = "%m/%d/%Y")
+baseline_mult$A1C <- unlist(apply(baseline_mult, 1, function(r) {
+  Date <- r["Date of first OGTT"]
+  ID <- as.numeric(r["MRN"])
+  d <- a1c %>% filter(MRN == ID)
+  d$days <- difftime(d$Date.of.A1C, Date, units = "days")
+  d <- d %>% filter(days <= 14 & days >= -90)
+  a <- d$A1C[which.min(abs(difftime(d$Date.of.A1C, Date)))]
+  if (is.na(Date) | nrow(d) == 0) {
+    a <- NA
+  }
+  return(a)
+}))
+baseline_all$A1C <- unlist(apply(baseline_all, 1, function(r) {
+  Date <- r["Date of first OGTT"]
+  ID <- as.numeric(r["MRN"])
+  d <- a1c %>% filter(MRN == ID)
+  d$days <- difftime(d$Date.of.A1C, Date, units = "days")
+  d <- d %>% filter(days <= 14 & days >= -90)
+  a <- d$A1C[which.min(abs(difftime(d$Date.of.A1C, Date)))]
+  if (is.na(Date) | nrow(d) == 0) {
+    a <- NA
+  }
+  return(a)
+}))
+
+# find most recent FEV1 that is within the 90 days prior to OGTT or up to 14 days after OGTT
+pfttemp <- alldata[!is.na(alldata$FEV1.per.pred), ]
+pfttemp$Date.of.PFT <- as.Date(pfttemp$Date, format = "%m/%d/%Y")
+baseline_mult$FEV1.per.pred <- unlist(apply(baseline_mult, 1, function(r) {
+  Date <- r["Date of first OGTT"]
+  ID <- as.numeric(r["MRN"])
+  d <- pfttemp %>% filter(MRN == ID)
+  d$days <- difftime(d$Date.of.PFT, Date, units = "days")
+  d <- d %>% filter(days <= 14 & days >= -90)
+  a <- d$FEV1.per.pred[which.min(abs(difftime(d$Date.of.PFT, Date)))]
+  if (is.na(Date) | nrow(d) == 0) {
+    a <- NA
+  }
+  return(a)
+}))
+baseline_all$FEV1.per.pred <- unlist(apply(baseline_all, 1, function(r) {
+  Date <- r["Date of first OGTT"]
+  ID <- as.numeric(r["MRN"])
+  d <- pfttemp %>% filter(MRN == ID)
+  d$days <- difftime(d$Date.of.PFT, Date, units = "days")
+  d <- d %>% filter(days <= 14 & days >= -90)
+  a <- d$FEV1.per.pred[which.min(abs(difftime(d$Date.of.PFT, Date)))]
+  if (is.na(Date) | nrow(d) == 0) {
+    a <- NA
+  }
+  return(a)
+}))
+# didn't add any PFTs so won't repeat code for FVC
+
+# find most recent BMI that is within the 90 days prior to OGTT or up to 14 days after OGTT
+bmitemp <- alldata[!is.na(alldata$BMI), ]
+bmitemp$Date.of.BMI <- as.Date(bmitemp$Date, format = "%m/%d/%Y")
+baseline_mult$BMI <- unlist(apply(baseline_mult, 1, function(r) {
+  Date <- r["Date of first OGTT"]
+  ID <- as.numeric(r["MRN"])
+  d <- bmitemp %>% filter(MRN == ID)
+  d$days <- difftime(d$Date.of.BMI, Date, units = "days")
+  d <- d %>% filter(days <= 14 & days >= -90)
+  a <- d$BMI[which.min(abs(difftime(d$Date.of.BMI, Date)))]
+  if (is.na(Date) | nrow(d) == 0) {
+    a <- NA
+  }
+  return(a)
+}))
+baseline_all$BMI <- unlist(apply(baseline_all, 1, function(r) {
+  Date <- r["Date of first OGTT"]
+  ID <- as.numeric(r["MRN"])
+  d <- bmitemp %>% filter(MRN == ID)
+  d$days <- difftime(d$Date.of.BMI, Date, units = "days")
+  d <- d %>% filter(days <= 14 & days >= -90)
+  a <- d$BMI[which.min(abs(difftime(d$Date.of.BMI, Date)))]
+  if (is.na(Date) | nrow(d) == 0) {
+    a <- NA
+  }
+  return(a)
+}))
+
+#############################################################
+
+# create multogtt with time <4 years since there is little data beyond that
+multogtt_reduced <- multogtt[multogtt$time <= 4, ]
+
+##############
+# DESC STATS #
+##############
+
+# table of demographics in those with at least one pre-CFRD OGTT
+demovars <- c(
+  "Sex", "race_eth", "age", "genoRisk", "Fasting.Glucose", "X1.hour.Glucose", "X2.hour.Glucose", "onmod_baseline", "modstarted_fup",
+  "A1C", "FEV1.per.pred", "FVC.per.pred", "BMI", "bmipercentile", "FEV.1.per.pred_FVC.per.pred"
+)
+# Taking out these clinical variables - if they are to be added to table 1, need to figure out how
+# because they come from PFT visits, not the same date as OGTT
+#              "BMI","FEV1.per.pred","FVC.per.pred","FEF.25.75.per.pred","bmip")
+t1_alldata <- CreateTableOne(data = baseline_all, vars = demovars)
+t1_alldata <- print(t1_alldata, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T, test = F)
+
+# table of demographics in those with at least one pre-CFRD OGTT by sex
+t1_alldata_sex <- CreateTableOne(data = baseline_all, vars = demovars, strata = "Sex")
+t1_alldata_sex <- print(t1_alldata_sex, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T)
+
+# table of demographics in those with more than 1 OGTT
+t1_multogtt <- CreateTableOne(data = baseline_mult, vars = c(demovars, "Follow_up_time_yrs"))
+t1_multogtt <- print(t1_multogtt, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T, test = F)
+
+# table of demographics in those with more than 1 OGTT stratified by progression
+t1_multogtt_prog <- CreateTableOne(data = baseline_mult, vars = c(demovars, "Follow_up_time_yrs", "age_at_CFRD_dx"), strata = "CFRD")
+t1_multogtt_prog <- print(t1_multogtt_prog, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T, test = T)
+
+# table of demographics in those with more than 1 OGTT by sex
+t1_multogtt_sex <- CreateTableOne(data = baseline_mult, vars = demovars, strata = "Sex")
+t1_multogtt_sex <- print(t1_multogtt_sex, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T)
+
+# compare baseline glucose values in those who developed CFRD and those who did not
+# added other vars too
+gluctable <- CreateTableOne(baseline_mult, vars = demovars, strata = "CFRD")
+gluctable <- print(gluctable, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T, test = T)
+
+########################
+# LOGISTIC REGRESSIONS #
+########################
+
+# logistic regression of 1 hour and 2 hour glucose predicting cfrd
+log1 <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose, family = "binomial")
+or1 <- exp(cbind(OR = coef(log1), confint(log1)))
+log1_adj <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose + Base_age + Sex, family = "binomial")
+or1_adj <- exp(cbind(OR = coef(log1_adj), confint(log1_adj)))
+
+# add interaction with sex
+log1_adj_int <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose + Base_age + Sex + Sex * Baseline_1hr_glucose, family = "binomial")
+or1_adj_int <- exp(cbind(OR = coef(log1_adj_int), confint(log1_adj_int)))
+
+log1_cat <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose_155, family = "binomial")
+or1_cat <- exp(cbind(OR = coef(log1_cat), confint(log1_cat)))
+log1_cat_adj <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose_155 + Base_age + Sex, family = "binomial")
+or1_cat_adj <- exp(cbind(OR = coef(log1_cat_adj), confint(log1_cat_adj)))
+
+log1_cat_140 <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose_140, family = "binomial")
+or1_cat_140 <- exp(cbind(OR = coef(log1_cat_140), confint(log1_cat_140)))
+log1_cat_adj_140 <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose_140 + Base_age + Sex, family = "binomial")
+or1_cat_adj_140 <- exp(cbind(OR = coef(log1_cat_adj_140), confint(log1_cat_adj_140)))
+
+# add interaction with sex
+log1_cat_adj_int <- glm(
+  data = baseline_mult, CFRD ~ Baseline_1hr_glucose_155 + Base_age + Sex + Sex * Baseline_1hr_glucose_155,
+  family = "binomial"
+)
+or1_cat_adj_int <- exp(cbind(OR = coef(log1_cat_adj_int), confint(log1_cat_adj_int)))
+
+log2 <- glm(data = baseline_mult, CFRD ~ Baseline_2hr_glucose, family = "binomial")
+or2 <- exp(cbind(OR = coef(log2), confint(log2)))
+
+log2_adj <- glm(data = baseline_mult, CFRD ~ Baseline_2hr_glucose + Base_age + Sex, family = "binomial")
+or2_adj <- exp(cbind(OR = coef(log2_adj), confint(log2_adj)))
+
+# add interaction with sex
+log2_adj_int <- glm(data = baseline_mult, CFRD ~ Baseline_2hr_glucose + Base_age + Sex + Sex * Baseline_2hr_glucose, family = "binomial")
+or2_adj_int <- exp(cbind(OR = coef(log2_adj_int), confint(log2_adj_int)))
+
+# ROC for 1 hour glucose
+# get predicted values from logistic regression
+baseline_mult$pred_1hr <- predict(log1, type = "response", data.frame = baseline_mult)
+# perform the ROC analysis
+roc1 <- roc(data = baseline_mult, response = CFRD, predictor = X1.hour.Glucose)
+# find the cutoff that maximizes the Youden Index
+coords1 <- coords(roc1, "b", ret = c("t", "specificity", "sensitivity"), best.method = "youden") # default
+
+# ROC for 2 hour glucose
+baseline_mult$pred_2hr <- predict(log2, type = "response", data.frame = baseline_mult)
+roc2 <- roc(data = baseline_mult, response = CFRD, predictor = X2.hour.Glucose)
+coords2 <- coords(roc2, "b", ret = c("t", "specificity", "sensitivity"), best.method = "youden") # default
+
+##############
+# COX MODELS #
+##############
+
+# CPH model for baseline 1 hour glucose
+cox1hr <- coxph(Surv(Follow_up_time_yrs, CFRD) ~ Baseline_1hr_glucose, data = baseline_mult)
 ```
+
+# Questions for Deb
+
+- Part of the reason that we split the data into 3 eras was because the PH assumption was violated in the full model. How do you want to handle that for the power calculations?

### 2025-03-18 Christine Chan/CF-SuGAR/power_calculations.qmd
NIH power calculations with Laura's original code

@@
 library(childsds)
+library(rspiro)
 library(survival)
@@
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan"
 )
@@
 # Import data and codebook
-annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
+annual <- read.csv("./CF Registry/Data_Raw/CFF19_Annualized_Del1.csv",
   na.strings = c("", "N/A")
 )
-data_dict <- readxl::read_excel("./Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
+data_dict <- readxl::read_excel("./CF Registry/Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
 data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
@@
 )
-# Get mutations
-demogs <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
+# Get mutations, race, etc.
+demogs <- read.csv("./CF Registry/Data_Raw/CFF19_DemogCFDiag_Del1.csv", na.strings = "")
 races <- c(
@@
 annual$age_group <- droplevels(annual$age_group)
-# List of covariates
-vars <- c(
-  "age_group", "Gender", "bmi_perc", "bmi_cat", "MutClass", "Race",
-  "Hispanicrace", "A_IsOnEnzymes", "ppFEV1_cat", "tube_feeding",
-  "NumPulmExacerbation", "A_pseudomonasaeruginosa", "A_fungalyeast1",
-  "A_mycocultureresults3", "A_corticosteroids1", "cfrd_status_annual",
-  "A_Vx770", "A_VX809comb", "liver_disease"
-)
 # Nice labels for plots and tables
@@
 xvars <- c(
-  "age_group", "Gender", "Race", "Hispanicrace", "bmi_cat", "MutClass",
-  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
-  "NumPulmExacerbation", "A_corticosteroids1"
+  "Age_YrEnd", "Gender", "bmi", "MutClass", "tube_feeding", "liver_disease", "ppFEV1_cat",
+  "NumPulmExacerbation"
 )
@@
 
+## 1 hour glucose alone
+
 ```{r}
@@
 
-ogtt <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/ogtt.csv")
-pft <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/pfts.csv")
-a1c <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/a1c.csv")
-demo <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/demographics.csv")
-dob <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/dob.csv")
+ogtt <- read.csv("./1 hour OGTT in CFRD/Data_raw/ogtt.csv", na.strings = "")
+pft <- read.csv("./1 hour OGTT in CFRD/Data_raw/pfts.csv", na.strings = "")
+a1c <- read.csv("./1 hour OGTT in CFRD/Data_raw/a1c.csv", na.strings = "")
+demo <- read.csv("./1 hour OGTT in CFRD/Data_raw/demographics.csv", na.strings = "")
+dob <- read.csv("./1 hour OGTT in CFRD/Data_raw/dob.csv", na.strings = "")
 
@@
 # read in date of modulator use
-mod <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/modulators.csv", na.strings = c("", " ", NA))
+mod <- read.csv("./1 hour OGTT in CFRD/Data_raw/modulators.csv", na.strings = c("", " ", NA))
 mod$Kalydeco.Start.Date <- as.Date(mod$Kalydeco.Start.Date, format = "%m/%d/%Y")
@@
 ##############
-# DESC STATS #
-##############
-
-# table of demographics in those with at least one pre-CFRD OGTT
-demovars <- c(
-  "Sex", "race_eth", "age", "genoRisk", "Fasting.Glucose", "X1.hour.Glucose", "X2.hour.Glucose", "onmod_baseline", "modstarted_fup",
-  "A1C", "FEV1.per.pred", "FVC.per.pred", "BMI", "bmipercentile", "FEV.1.per.pred_FVC.per.pred"
-)
-# Taking out these clinical variables - if they are to be added to table 1, need to figure out how
-# because they come from PFT visits, not the same date as OGTT
-#              "BMI","FEV1.per.pred","FVC.per.pred","FEF.25.75.per.pred","bmip")
-t1_alldata <- CreateTableOne(data = baseline_all, vars = demovars)
-t1_alldata <- print(t1_alldata, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T, test = F)
-
-# table of demographics in those with at least one pre-CFRD OGTT by sex
-t1_alldata_sex <- CreateTableOne(data = baseline_all, vars = demovars, strata = "Sex")
-t1_alldata_sex <- print(t1_alldata_sex, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T)
-
-# table of demographics in those with more than 1 OGTT
-t1_multogtt <- CreateTableOne(data = baseline_mult, vars = c(demovars, "Follow_up_time_yrs"))
-t1_multogtt <- print(t1_multogtt, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T, test = F)
-
-# table of demographics in those with more than 1 OGTT stratified by progression
-t1_multogtt_prog <- CreateTableOne(data = baseline_mult, vars = c(demovars, "Follow_up_time_yrs", "age_at_CFRD_dx"), strata = "CFRD")
-t1_multogtt_prog <- print(t1_multogtt_prog, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T, test = T)
-
-# table of demographics in those with more than 1 OGTT by sex
-t1_multogtt_sex <- CreateTableOne(data = baseline_mult, vars = demovars, strata = "Sex")
-t1_multogtt_sex <- print(t1_multogtt_sex, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T)
-
-# compare baseline glucose values in those who developed CFRD and those who did not
-# added other vars too
-gluctable <- CreateTableOne(baseline_mult, vars = demovars, strata = "CFRD")
-gluctable <- print(gluctable, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T, test = T)
-
-########################
-# LOGISTIC REGRESSIONS #
-########################
-
-# logistic regression of 1 hour and 2 hour glucose predicting cfrd
-log1 <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose, family = "binomial")
-or1 <- exp(cbind(OR = coef(log1), confint(log1)))
-log1_adj <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose + Base_age + Sex, family = "binomial")
-or1_adj <- exp(cbind(OR = coef(log1_adj), confint(log1_adj)))
-
-# add interaction with sex
-log1_adj_int <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose + Base_age + Sex + Sex * Baseline_1hr_glucose, family = "binomial")
-or1_adj_int <- exp(cbind(OR = coef(log1_adj_int), confint(log1_adj_int)))
-
-log1_cat <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose_155, family = "binomial")
-or1_cat <- exp(cbind(OR = coef(log1_cat), confint(log1_cat)))
-log1_cat_adj <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose_155 + Base_age + Sex, family = "binomial")
-or1_cat_adj <- exp(cbind(OR = coef(log1_cat_adj), confint(log1_cat_adj)))
-
-log1_cat_140 <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose_140, family = "binomial")
-or1_cat_140 <- exp(cbind(OR = coef(log1_cat_140), confint(log1_cat_140)))
-log1_cat_adj_140 <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose_140 + Base_age + Sex, family = "binomial")
-or1_cat_adj_140 <- exp(cbind(OR = coef(log1_cat_adj_140), confint(log1_cat_adj_140)))
-
-# add interaction with sex
-log1_cat_adj_int <- glm(
-  data = baseline_mult, CFRD ~ Baseline_1hr_glucose_155 + Base_age + Sex + Sex * Baseline_1hr_glucose_155,
-  family = "binomial"
-)
-or1_cat_adj_int <- exp(cbind(OR = coef(log1_cat_adj_int), confint(log1_cat_adj_int)))
-
-log2 <- glm(data = baseline_mult, CFRD ~ Baseline_2hr_glucose, family = "binomial")
-or2 <- exp(cbind(OR = coef(log2), confint(log2)))
-
-log2_adj <- glm(data = baseline_mult, CFRD ~ Baseline_2hr_glucose + Base_age + Sex, family = "binomial")
-or2_adj <- exp(cbind(OR = coef(log2_adj), confint(log2_adj)))
-
-# add interaction with sex
-log2_adj_int <- glm(data = baseline_mult, CFRD ~ Baseline_2hr_glucose + Base_age + Sex + Sex * Baseline_2hr_glucose, family = "binomial")
-or2_adj_int <- exp(cbind(OR = coef(log2_adj_int), confint(log2_adj_int)))
-
-# ROC for 1 hour glucose
-# get predicted values from logistic regression
-baseline_mult$pred_1hr <- predict(log1, type = "response", data.frame = baseline_mult)
-# perform the ROC analysis
-roc1 <- roc(data = baseline_mult, response = CFRD, predictor = X1.hour.Glucose)
-# find the cutoff that maximizes the Youden Index
-coords1 <- coords(roc1, "b", ret = c("t", "specificity", "sensitivity"), best.method = "youden") # default
-
-# ROC for 2 hour glucose
-baseline_mult$pred_2hr <- predict(log2, type = "response", data.frame = baseline_mult)
-roc2 <- roc(data = baseline_mult, response = CFRD, predictor = X2.hour.Glucose)
-coords2 <- coords(roc2, "b", ret = c("t", "specificity", "sensitivity"), best.method = "youden") # default
-
-##############
 # COX MODELS #
 ##############
+baseline_mult$Race <- factor(baseline_mult$Race, levels = c("Caucasian", "Black", "Other"))
+# CPH model for baseline 1 hour glucose
+f <- Surv(Follow_up_time_yrs, CFRD) ~ age + Sex + BMI + FEV1.per.pred + Baseline_1hr_glucose
+cox1hr <- coxph(f, data = baseline_mult)
+# Forest plot and coefficients table
+cox1hr %>%
+  tbl_regression(exponentiate = T)
+ggcoef_table(cox1hr,
+  show_p_values = F, add_reference_rows = T, exponentiate = T,
+  signif_stars = F, colour = F
+)
+```
 
-# CPH model for baseline 1 hour glucose
-cox1hr <- coxph(Surv(Follow_up_time_yrs, CFRD) ~ Baseline_1hr_glucose, data = baseline_mult)
+## 1 hour glucose and HbA1c
+
+```{r}
+cox1hra1c <- coxph(update(f, . ~ . + A1C), data = baseline_mult)
+# Forest plot and coefficients table
+cox1hra1c %>%
+  tbl_regression(exponentiate = T)
+ggcoef_table(cox1hra1c,
+  show_p_values = F, add_reference_rows = T, exponentiate = T,
+  signif_stars = F, colour = F
+)
 ```
+
+# Notes
+
+- There are no participants over 31 in Christine's 1 hour OGTT analysis data, so I used age as a continuous variable.
+- Because these models include age and sex, I also used raw BMI value as a continuous variable.
+- Race was removed from the models because the group sizes are so small in the one hour dataset (only 3 Black participants).
+- Pancreatic status was not available in the one hour glucose dataset, so was removed from the models.
 

### 2025-03-19 Christine Chan/CF-SuGAR/power_calculations.qmd
base models for Deb

@@
   tstart = "Interval start", tstop = "Interval stop",
-  Age_YrEnd = "Age", bmi_perc = "BMI %ile"
+  Age_YrEnd = "Age", bmi_perc = "BMI %ile",
+  A_hgba1c = "HbA1c"
 )
@@
   labels[colnames(annual)[which(colnames(annual) %in% names(labels))]]
+# Limit to after 2010
+annual <- annual %>% filter(ReviewYear >= 2010)
 ```
@@
 xvars <- c(
-  "Age_YrEnd", "Gender", "bmi", "MutClass", "tube_feeding", "liver_disease", "ppFEV1_cat",
-  "NumPulmExacerbation"
+  "age_group", "Gender", "Race", "Hispanicrace", "bmi", "MutClass",
+  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "A_FEV1_pct_predicted",
+  "NumPulmExacerbation", "A_corticosteroids1", "A_hgba1c"
 )
@@
 cox_res <- coxph(f, data = annual, cluster = eDWID)
+cox.zph(cox_res)
 # Forest plot and coefficients table
@@
 
-## Base model plus HbA1c
+## Base model with variables available in Lorenz et al.
 
@@
 # Formula for base model
-cox_res <- coxph(update(f, . ~ . + A_hgba1c), data = annual, cluster = eDWID)
+xvars <- c(
+  "age_group", "Gender", "bmi", "MutClass", "A_FEV1_pct_predicted", "A_hgba1c"
+)
+y <- "Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2)"
+f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
+cox_res <- coxph(f, data = annual, cluster = eDWID)
+cox.zph(cox_res)
 # Forest plot and coefficients table
@@
 
-# 1 hour OGTT glucose in CFRD
-
-## 1 hour glucose alone
+# Lorenz et al.
 
@@
 
-ogtt <- read.csv("./1 hour OGTT in CFRD/Data_raw/ogtt.csv", na.strings = "")
-pft <- read.csv("./1 hour OGTT in CFRD/Data_raw/pfts.csv", na.strings = "")
-a1c <- read.csv("./1 hour OGTT in CFRD/Data_raw/a1c.csv", na.strings = "")
-demo <- read.csv("./1 hour OGTT in CFRD/Data_raw/demographics.csv", na.strings = "")
-dob <- read.csv("./1 hour OGTT in CFRD/Data_raw/dob.csv", na.strings = "")
+ogtt <- read.csv("./1 hour OGTT in CFRD/Data_raw/OGTT.csv")
+pft <- read.csv("./1 hour OGTT in CFRD/Data_raw/PFTs.csv")
+a1c <- read.csv("./1 hour OGTT in CFRD/Data_raw/A1c.csv")
+demo <- read.csv("./1 hour OGTT in CFRD/Data_raw/demographics.csv")
+dob <- read.csv("./1 hour OGTT in CFRD/Data_raw/dob.csv")
 
@@
 # write file of people with missing genotype risk
-# write.csv(alldata[alldata$genoRisk=="",],"./Christine Chan/1 hour OGTT in CFRD/Data_raw/missing_genotype_risk.csv")
+# write.csv(alldata[alldata$genoRisk=="",],"./1 hour OGTT in CFRD/Data_raw/missing_genotype_risk.csv")
 
@@
 col_check <- col_check[order(col_check$MRN, col_check$Date), ]
-# write.csv(col_check,"./Christine Chan/1 hour OGTT in CFRD/Data_raw/check_cfrd.csv",na="",row.names = F)
+# write.csv(col_check,"./1 hour OGTT in CFRD/Data_raw/check_cfrd.csv",na="",row.names = F)
 
@@
 col_check2 <- col_check2[order(col_check2$MRN, col_check2$Date), ]
-# write.csv(col_check2,"./Christine Chan/1 hour OGTT in CFRD/Data_raw/check_cfrd2.csv",na="",row.names = F)
+# write.csv(col_check2,"./1 hour OGTT in CFRD/Data_raw/check_cfrd2.csv",na="",row.names = F)
 
@@
 ##############
-baseline_mult$Race <- factor(baseline_mult$Race, levels = c("Caucasian", "Black", "Other"))
+```
+
+## Base model
+
+```{r}
 # CPH model for baseline 1 hour glucose
-f <- Surv(Follow_up_time_yrs, CFRD) ~ age + Sex + BMI + FEV1.per.pred + Baseline_1hr_glucose
-cox1hr <- coxph(f, data = baseline_mult)
+cox1hr <- coxph(
+  Surv(Follow_up_time_yrs, CFRD) ~ age + Sex + BMI + class1 + FEV1.per.pred +
+    A1C,
+  data = baseline_mult
+)
 # Forest plot and coefficients table
@@
 
-## 1 hour glucose and HbA1c
+## Model with one hour OGTT
 
 ```{r}
-cox1hra1c <- coxph(update(f, . ~ . + A1C), data = baseline_mult)
+# CPH model for baseline 1 hour glucose
+cox1hr <- coxph(
+  Surv(Follow_up_time_yrs, CFRD) ~ age + Sex + BMI + class1 + FEV1.per.pred +
+    A1C + Baseline_1hr_glucose,
+  data = baseline_mult
+)
 # Forest plot and coefficients table
-cox1hra1c %>%
+cox1hr %>%
   tbl_regression(exponentiate = T)
-ggcoef_table(cox1hra1c,
+ggcoef_table(cox1hr,
   show_p_values = F, add_reference_rows = T, exponentiate = T,
@@
 
-- There are no participants over 31 in Christine's 1 hour OGTT analysis data, so I used age as a continuous variable.
-- Because these models include age and sex, I also used raw BMI value as a continuous variable.
-- Race was removed from the models because the group sizes are so small in the one hour dataset (only 3 Black participants).
-- Pancreatic status was not available in the one hour glucose dataset, so was removed from the models.
-
-# Questions for Deb
-
-- Part of the reason that we split the data into 3 eras was because the PH assumption was violated in the full model. How do you want to handle that for the power calculations?
+- I removed race from the models because the Lorenz et al. paper data only includes 8 non-white participants, so the coefficients exploded.

### 2025-03-19 Christine Chan/CF-SuGAR/power_calculations.qmd
power calculations to do list

@@
   "A_IsOnEnzymes", "tube_feeding", "liver_disease", "A_FEV1_pct_predicted",
-  "NumPulmExacerbation", "A_corticosteroids1", "A_hgba1c"
+  "NumPulmExacerbation", "A_corticosteroids1"
 )
@@
 ```{r}
+annual = annual%>% filter(Age_YrEnd<=22)
+annual$age_group = cut(annual$Age_YrEnd,c(7,11,22))
 # Formula for base model
 xvars <- c(
-  "age_group", "Gender", "bmi", "MutClass", "A_FEV1_pct_predicted", "A_hgba1c"
+  "age_group", "Gender"
 )
@@
 cox1hr <- coxph(
-  Surv(Follow_up_time_yrs, CFRD) ~ age + Sex + BMI + class1 + FEV1.per.pred +
-    A1C,
+  Surv(Follow_up_time_yrs, CFRD) ~ age + Sex ,
   data = baseline_mult
@@
 # CPH model for baseline 1 hour glucose
+baseline_mult$age_group = cut(baseline_mult$age,c(7,11,22))
+
 cox1hr <- coxph(
-  Surv(Follow_up_time_yrs, CFRD) ~ age + Sex + BMI + class1 + FEV1.per.pred +
-    A1C + Baseline_1hr_glucose,
+  Surv(Follow_up_time_yrs, CFRD) ~ age*Sex,
   data = baseline_mult
 )
+
 # Forest plot and coefficients table
@@
 - I removed race from the models because the Lorenz et al. paper data only includes 8 non-white participants, so the coefficients exploded.
+
+To Do
+
+- Plot HbA1c by age for those who do develop CFRD
+- Take out HbA1c for base model
+- Include generalized R2 for the models
+- Split age into equal tertiles
+  2 x 3 table with rows for diabetes status and age as columns, cells are numbers
+
+- PROC power has a cox reg option
+  - Base model is from the registry 
+  - 0.5, 1, 2 times the 1 hr glucose point estimate from Lorenz et al.
+  - 80, 85, 90% power
+
+- Mixed model power calculations for aim 3
+  - Repeated measures with FEV1 and BMI
+  - Mutation class and sex as predictors

### 2025-03-20 Christine Chan/CF-SuGAR/power_calculations.qmd
PROC POWER calculations for Deb

@@
     toc-float: true
-    code-fold: true
     self-contained: true
@@
     page-layout: full
-    theme:
-      light: flatly
-      dark: darkly
 editor: source
@@
 ```{r data cleaning}
+#| include: false
 # Import data and codebook
@@
 annual$bmi_perc <- coalesce(annual$bmi_perc_cdc, annual$bmi_perc_who)
-annual$bmi_cat <- apply(annual, 1, function(r) {
-  age <- as.numeric(r["Age_YrEnd"])
-  gender <- as.character(r["Gender"])
-  bmi <- as.numeric(r["bmi"])
-  bmi_perc <- as.numeric(r["bmi_perc"])
-  if (age <= 20) {
-    return(cut(bmi_perc, c(-Inf, 10, 50, 85, Inf),
-      right = F,
-      labels = c("Undernourished", "At Risk", "Target/Goal", "Overweight")
-    ))
-  } else if (age > 20) {
-    if (gender == "M") {
-      return(cut(bmi, c(-Inf, 18.5, 23, 25, Inf),
-        right = F,
-        labels = c("Undernourished", "At Risk", "Target/Goal", "Overweight")
-      ))
-    } else if (gender == "F") {
-      return(cut(bmi, c(-Inf, 18.5, 22, 25, Inf),
-        right = F,
-        labels = c("Undernourished", "At Risk", "Target/Goal", "Overweight")
-      ))
-    }
-  }
-})
-annual$bmi_cat <- relevel(annual$bmi_cat, ref = "Target/Goal")
 # Format columns
@@
 annual <- annual %>% filter(ReviewYear >= 2010)
-```
-
-# Registry
-
-## Base model
-
-```{r}
 # Origin for survival analyses
@@
 annual$tstop <- as.numeric(annual$tstop) - as.numeric(og)
-# Formula for base model
-xvars <- c(
-  "age_group", "Gender", "Race", "Hispanicrace", "bmi", "MutClass",
-  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "A_FEV1_pct_predicted",
-  "NumPulmExacerbation", "A_corticosteroids1"
-)
-y <- "Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2)"
-f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
-cox_res <- coxph(f, data = annual, cluster = eDWID)
-cox.zph(cox_res)
-# Forest plot and coefficients table
-cox_res %>%
-  tbl_regression(exponentiate = T)
-ggcoef_table(cox_res,
-  show_p_values = F, add_reference_rows = T, exponentiate = T,
-  signif_stars = F, colour = F
-)
-```
-
-## Base model with variables available in Lorenz et al.
-
-```{r}
-annual = annual%>% filter(Age_YrEnd<=22)
-annual$age_group = cut(annual$Age_YrEnd,c(7,11,22))
-# Formula for base model
-xvars <- c(
-  "age_group", "Gender"
-)
-y <- "Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2)"
-f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
-cox_res <- coxph(f, data = annual, cluster = eDWID)
-cox.zph(cox_res)
-# Forest plot and coefficients table
-cox_res %>%
-  tbl_regression(exponentiate = T)
-ggcoef_table(cox_res,
-  show_p_values = F, add_reference_rows = T, exponentiate = T,
-  signif_stars = F, colour = F
-)
-```
-
-# Lorenz et al.
-
-```{r}
+# Remove ages > 22 to match Lorenz et al.
+annual <- annual %>% filter(Age_YrEnd < 23)
+# Lorenz et al. data cleaning
 # read in separate data files
@@
 multogtt_reduced <- multogtt[multogtt$time <= 4, ]
+# rename data
+lorenz <- baseline_mult
+registry <- annual
+rm(annual, baseline_mult)
+```
 
-##############
-# COX MODELS #
-##############
+# Data cleaning
+
+- Participants over age 23 were excluded from the registry data, along with data collected prior to 2010.
+- Age groups were chosen based on the Lorenz et al. data. We split the age column into equal tertiles:
+
+```{r}
+lorenz$age_group <- cut(lorenz$age, c(7, 10.48, 13.3, 22.2), right = F)
+table(lorenz$age_group, lorenz$CFRD)
+registry$age_group <- cut(registry$Age_YrEnd, c(7, 10.48, 13.3, 22.2), right = F)
 ```
+
+# Registry base model
+
+```{r}
+registry_base <- coxph(
+  Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2) ~
+    age_group + Gender,
+  data = registry, cluster = eDWID
+)
+# Forest plot and coefficients table
+registry_base %>%
+  tbl_regression(exponentiate = T)
+ggcoef_table(registry_base,
+  show_p_values = F, add_reference_rows = T, exponentiate = T,
+  signif_stars = F, colour = F
+)
+```
+
+The above model had a generalized $R^2$ of `r as.numeric(summary(registry_base)$rsq[1])`.
+
+# Lorenz et al.
 
@@
 # CPH model for baseline 1 hour glucose
-cox1hr <- coxph(
-  Surv(Follow_up_time_yrs, CFRD) ~ age + Sex ,
-  data = baseline_mult
+lorenz_base <- coxph(
+  Surv(Follow_up_time_yrs, CFRD) ~ age_group + Sex,
+  data = lorenz
 )
 # Forest plot and coefficients table
-cox1hr %>%
+lorenz_base %>%
   tbl_regression(exponentiate = T)
-ggcoef_table(cox1hr,
+ggcoef_table(lorenz_base,
   show_p_values = F, add_reference_rows = T, exponentiate = T,
@@
 
+The above model had a generalized $R^2$ of `r as.numeric(summary(lorenz_base)$rsq[1])`.
+
 ## Model with one hour OGTT
@@
 # CPH model for baseline 1 hour glucose
-baseline_mult$age_group = cut(baseline_mult$age,c(7,11,22))
-
-cox1hr <- coxph(
-  Surv(Follow_up_time_yrs, CFRD) ~ age*Sex,
-  data = baseline_mult
+lorenz_1hr <- coxph(
+  Surv(Follow_up_time_yrs, CFRD) ~ age_group + Sex + Baseline_1hr_glucose,
+  data = lorenz
 )
-
+summary(lorenz_1hr)
 # Forest plot and coefficients table
-cox1hr %>%
+lorenz_1hr %>%
   tbl_regression(exponentiate = T)
-ggcoef_table(cox1hr,
+ggcoef_table(lorenz_1hr,
   show_p_values = F, add_reference_rows = T, exponentiate = T,
@@
 
-# Notes
+The above model had a generalized $R^2$ of `r as.numeric(summary(lorenz_1hr)$rsq[1])`.
 
-- I removed race from the models because the Lorenz et al. paper data only includes 8 non-white participants, so the coefficients exploded.
+# PROC POWER
 
-To Do
+## HR for 1 hour OGTT glucose
 
-- Plot HbA1c by age for those who do develop CFRD
-- Take out HbA1c for base model
-- Include generalized R2 for the models
-- Split age into equal tertiles
-  2 x 3 table with rows for diabetes status and age as columns, cells are numbers
+```{r}
+exp(lorenz_1hr$coefficients["Baseline_1hr_glucose"] * 0.5)
+exp(lorenz_1hr$coefficients["Baseline_1hr_glucose"])
+exp(lorenz_1hr$coefficients["Baseline_1hr_glucose"] * 2)
+```
+
+## Regression of age group and sex on 1 hour OGTT glucose
+
+```{r}
+lin_mod <- lm(Baseline_1hr_glucose ~ age_group + Sex, data = lorenz)
+summary(lin_mod)$r.squared
+```
+
+## SD of 1 hour OGTT glucose
+
+```{r}
+sd(lorenz$Baseline_1hr_glucose)
+```
+
+## Event probability
+
+```{r}
+registry %>%
+  group_by(eDWID) %>%
+  summarise(cfrd = unique(first_cfrd), .groups = "drop") %>%
+  summarise(cfrd = mean(!is.na(cfrd)))
+mean(lorenz$CFRD == 1)
+```
+
+# To Do
 

### 2025-03-20 Christine Chan/CF-SuGAR/power_calculations.qmd
update NIH power calculations

@@
 ```{r}
-exp(lorenz_1hr$coefficients["Baseline_1hr_glucose"] * 0.5)
+exp(lorenz_1hr$coefficients["Baseline_1hr_glucose"]) * 0.5
 exp(lorenz_1hr$coefficients["Baseline_1hr_glucose"])
-exp(lorenz_1hr$coefficients["Baseline_1hr_glucose"] * 2)
+exp(lorenz_1hr$coefficients["Baseline_1hr_glucose"]) * 2
 ```

### 2025-03-21 Christine Chan/CF-SuGAR/power_calculations.qmd
PROC GLIMMIX for CF-SuGAR power

@@
 library(ggstats)
+library(lmerTest)
 home_dir <- switch(Sys.info()["sysname"],
@@
 annual$tstop <- as.numeric(annual$tstop) - as.numeric(og)
-# Remove ages > 22 to match Lorenz et al.
-annual <- annual %>% filter(Age_YrEnd < 23)
 # Lorenz et al. data cleaning
@@
 
-- Participants over age 23 were excluded from the registry data, along with data collected prior to 2010.
 - Age groups were chosen based on the Lorenz et al. data. We split the age column into equal tertiles:
@@
 registry$age_group <- cut(registry$Age_YrEnd, c(7, 10.48, 13.3, 22.2), right = F)
-```
-
-# Registry base model
-
-```{r}
-registry_base <- coxph(
-  Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2) ~
-    age_group + Gender,
-  data = registry, cluster = eDWID
-)
-# Forest plot and coefficients table
-registry_base %>%
-  tbl_regression(exponentiate = T)
-ggcoef_table(registry_base,
-  show_p_values = F, add_reference_rows = T, exponentiate = T,
-  signif_stars = F, colour = F
+# Write data
+write.csv(lorenz,
+  file = "./CF-SuGAR/Background/Power Calculations/Data/lorenz.csv",
+  row.names = F, na = ""
 )
@@
 
-The above model had a generalized $R^2$ of `r as.numeric(summary(registry_base)$rsq[1])`.
+# Aim 1
 
-# Lorenz et al.
+- Power calculations for Aim 1 used only the Lorenz et al. data and no registry data.
 
@@
 
-# PROC POWER
+## PROC POWER
 
-## HR for 1 hour OGTT glucose
+### HR for 1 hour OGTT glucose
 
@@
 
-## Regression of age group and sex on 1 hour OGTT glucose
+### Regression of age group and sex on 1 hour OGTT glucose
 
@@
 
-## SD of 1 hour OGTT glucose
+### SD of 1 hour OGTT glucose
 
@@
 
-## Event probability
+### Event probability
 
@@
 
-# To Do
+# Aim 2
 
-- PROC power has a cox reg option
-  - Base model is from the registry 
-  - 0.5, 1, 2 times the 1 hr glucose point estimate from Lorenz et al.
-  - 80, 85, 90% power
+- All models below were fit using registry data.
+- One BMI value of > 200 was removed
 
-- Mixed model power calculations for aim 3
-  - Repeated measures with FEV1 and BMI
-  - Mutation class and sex as predictors
+```{r}
+registry$bmi[registry$bmi >= 200] <- NA
+# Write for SAS
+write.csv(registry,
+  file = "./CF-SuGAR/Background/Power Calculations/Data/registry.csv",
+  row.names = F, na = ""
+)
+```
+
+## BMI
+
+### Plot
+
+```{r}
+plot_df <- registry %>%
+  select(eDWID, ReviewYear, Age_YrEnd, bmi) %>%
+  drop_na()
+ggplot(plot_df, aes(x = Age_YrEnd, y = bmi, group = factor(eDWID))) +
+  geom_point(alpha = 0.1) +
+  geom_line(alpha = 0.1) +
+  theme(legend.position = element_blank()) +
+  theme_bw()
+```
+
+### Models
+
+```{r}
+mod1 <- lmer(bmi ~ Age_YrEnd + (1 | eDWID), data = registry)
+summary(mod1,ddf="Kenward-Roger")
+```
+
+## FEV1 % predicted
+
+### Plot
+
+```{r}
+plot_df <- registry %>%
+  select(eDWID, ReviewYear, Age_YrEnd, A_FEV1_pct_predicted) %>%
+  drop_na()
+ggplot(plot_df, aes(
+  x = Age_YrEnd, y = A_FEV1_pct_predicted, group = factor(eDWID)
+)) +
+  geom_point(alpha = 0.1) +
+  geom_line(alpha = 0.1) +
+  theme(legend.position = element_blank()) +
+  theme_bw()
+```
+
+### Models

### 2025-03-21 Christine Chan/CF-SuGAR/power_calculations.qmd
aim2 power for Deb

@@
 registry$bmi[registry$bmi >= 200] <- NA
+# A1c group
+registry$a1c_group <- cut(registry$A_hgba1c, c(-Inf, 5.9, Inf))
+# Remove people with CFRD
+registry <- registry %>% filter(CFRD != "CFRD+")
 # Write for SAS
@@
 mod1 <- lmer(bmi ~ Age_YrEnd + (1 | eDWID), data = registry)
-summary(mod1,ddf="Kenward-Roger")
+summary(mod1, ddf = "Kenward-Roger")
 ```

### 2025-03-21 Christine Chan/CF-SuGAR/power_calculations.qmd
mostly ready for GLIMMPSE

@@
 registry$a1c_group <- cut(registry$A_hgba1c, c(-Inf, 5.9, Inf))
+registry$a1c_group = as.numeric(registry$a1c_group=="(-Inf,5.9]")
 # Remove people with CFRD

### 2025-03-24 Christine Chan/CF Registry/cf_registry_CFRD_prediction.qmd
notes from meeting with Kristen

@@
 library(tidyr)
-knitr::opts_knit$set(root.dir = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry")
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
+)
+knitr::opts_knit$set(root.dir = home_dir)
 ```
@@
 ```
+
+- Check for HbA1c measures after the survival time (not sure if needs to be strictly less than)
+- Set parameterization = "slope" to see if change in HbA1c increases hazard 
+- Look into dynamic predictions in JMBayes
+- Could try re-fitting this with a left-censored model 
+  - If it says the same thing, it's probably okay to use the right-censored
+  - Could also try using the half interval model Kristen described

### 2025-03-26 Christine Chan/CF-SuGAR/power_calculations.qmd
CF-SuGAR power pretty much done (needs writeup though)

--- 2025-03-21 Christine Chan/CF-SuGAR/power_calculations.qmd
@@
   "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan"
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan",
+  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan"
 )
@@
 registry$bmi[registry$bmi >= 200] <- NA
+# A1c group
+registry$a1c_group <- cut(registry$A_hgba1c, c(-Inf, 5.9, Inf))
+registry$a1c_group <- relevel(registry$a1c_group, ref = "(5.9, Inf]")
+# Remove people with CFRD
+registry <- registry %>%
+  filter(CFRD != "CFRD+") %>%
+  arrange(eDWID, ReviewYear)
 # Write for SAS
@@
 )
+# Proportion of good and bad A1c
+# Overall
+prop.table(table(registry$a1c_group))
+# By person (take each person's worst)
+t <- registry %>%
+  group_by(eDWID) %>%
+  summarise(
+    a1c_group = ifelse(any(a1c_group == "(5.9, Inf]"), "(5.9, Inf]", "(-Inf,5.9]")
+  )
+prop.table(table(t$a1c_group))
 ```
@@
 
-### Plot
+### Plots
+
+#### Spaghetti
 
@@
 
+#### BMI correlation matrix by visit number
+
+```{r}
+t <- registry %>%
+  group_by(eDWID) %>%
+  mutate(visit_number = row_number()) %>%
+  filter(!is.na(bmi)) %>%
+  select(eDWID, visit_number, bmi) %>%
+  pivot_wider(names_from = visit_number, values_from = bmi)
+head(t)
+cor(data.matrix(t[, -c(1)]), use = "pairwise.complete.obs")
+```
+
 ### Models
@@
 ```{r}
-mod1 <- lmer(bmi ~ Age_YrEnd + (1 | eDWID), data = registry)
-summary(mod1,ddf="Kenward-Roger")
+mod1 <- lmer(bmi ~ a1c_group + Age_YrEnd + (Age_YrEnd | eDWID),
+  data = registry
+)
+summary(mod1, ddf = "Kenward-Roger")
 ```
@@
 
-### Models
+#### FEV1 correlation matrix by visit number
+
+```{r}
+t <- registry %>%
+  group_by(eDWID) %>%
+  mutate(visit_number = row_number()) %>%
+  filter(!is.na(A_FEV1_pct_predicted)) %>%
+  select(eDWID, visit_number, A_FEV1_pct_predicted) %>%
+  pivot_wider(names_from = visit_number, values_from = A_FEV1_pct_predicted)
+head(t)
+cor(data.matrix(t[, -c(1)]), use = "pairwise.complete.obs")
+```

### 2025-03-28 Christine Chan/CF-SuGAR/power_calculations.qmd
CF-SuGAR power calculations plus FEV1
more CF-SuGAR power stuff

@@
 # A1c group
-registry$a1c_group <- cut(registry$A_hgba1c, c(-Inf, 5.9, Inf))
-registry$a1c_group <- relevel(registry$a1c_group, ref = "(5.9, Inf]")
+registry$a1c_group <- as.numeric(registry$A_hgba1c <= 5.9)
 # Remove people with CFRD
@@
 ```{r}
-mod1 <- lmer(bmi ~ a1c_group + Age_YrEnd + (Age_YrEnd | eDWID),
-  data = registry
-)
-summary(mod1, ddf = "Kenward-Roger")
+bmi_mod <- lm(BMI ~ age + as.numeric(Baseline_1hr_glucose >= 140), data = lorenz)
+summary(bmi_mod)
+sd(resid(bmi_mod))
+prop.table(table(lorenz$Baseline_1hr_glucose >= 140))
 ```
@@
 ```
+
+### Models
+
+```{r}
+fev1_mod <- lm(FEV1.per.pred ~ age + as.numeric(Baseline_1hr_glucose >= 140), data = lorenz)
+summary(fev1_mod)
+sd(resid(fev1_mod))
+```

### 2025-04-01 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
SSwAB-IT FEV1 by age

@@
 library(ggpubr)
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Jordana Hoppe/SSwAB-IT",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Jordana Hoppe/SSwAB-IT"
+)
+knitr::opts_knit$set(root.dir = home_dir)
 ```
@@
 db_demographics <- left_join(db_demographics,
-  db_clinic_obtained_values %>% select(record_id, cos_collection_date),
+  db_clinic_obtained_values %>% select(record_id, cos_collection_date, ppfev1),
   by = join_by(record_id)
@@
 db_medications <- db_medications %>% filter(!record_id %in% exclude)
+# Fix chronic oral antibiotics
+db_medications$chronic_oral_antibiotic[
+  db_medications$other_chronic_oral_antibiotic == "Azithromycin"
+] <- "No"
 # Combine race and ethnicity columns
@@
 db_medications <- left_join(db_medications,
+  db_demographics %>% select(record_id, group),
+  by = join_by(record_id)
+)
+db_clinic_obtained_values <- left_join(db_clinic_obtained_values,
   db_demographics %>% select(record_id, group),
@@
 db_demographics %>%
-  select(group, age_at_cos, age_group, sex, Race, Ethnicity) %>%
+  select(group, age_at_cos, age_group, ppfev1, sex, Race, Ethnicity) %>%
   tbl_summary(
@@
   as_gt()
+```
+
+## By age group
+
+```{r}
+db_demographics %>%
+  mutate(age_group = cut(age_at_cos, c(-Inf, 18, Inf), right = F)) %>%
+  select(group, age_group, ppfev1, sex, Race, Ethnicity) %>%
+  tbl_summary(
+    by = age_group, missing_text = "Missing",
+    statistic = all_continuous() ~ c("{median} ({min}, {max})"),
+    label = list(
+      "sex" = "Sex", "age_at_cos" = "Age at COS"
+    )
+  ) %>%
+  add_overall() %>%
+  add_p(
+    test = list(c("Race", "Ethnicity") ~ "fisher.test")
+  ) %>%
+  separate_p_footnotes() %>%
+  as_gt()
+
+db_demographics %>%
+  mutate(age_group = cut(age_at_cos, c(-Inf, 18, Inf), right = F)) %>%
+  select(age_group, ppfev1) %>%
+  drop_na() %>%
+  ggplot(., aes(x = ppfev1)) +
+  geom_histogram() +
+  facet_wrap(~age_group)
 ```
@@
   bgcolor("White")
-ggsave("/home/timvigers/OneDrive/Vigers/CF/Jordana Hoppe/SSwAB-IT/Dissemination/Figures/fig.png",
+ggsave("./Dissemination/Figures/fig.png",
   plot = p1, width = 9, height = 6, units = "in"
@@
   bgcolor("White")
-ggsave("/home/timvigers/OneDrive/Vigers/CF/Jordana Hoppe/SSwAB-IT/Dissemination/Figures/euler_bw.png",
+ggsave("./Dissemination/Figures/euler_bw.png",
   plot = p2, width = 9, height = 6, units = "in"
@@
   bgcolor("White")
-ggsave("/home/timvigers/OneDrive/Vigers/CF/Jordana Hoppe/SSwAB-IT/Dissemination/Figures/venn_bw.png",
+ggsave("./Dissemination/Figures/venn_bw.png",
   plot = p3, width = 9, height = 6, units = "in"

### 2025-04-06 Angela Metcalf/quantifying_cl_cf_carriers.qmd
new CF carriers power calculations

--- 2025-02-06 Angela Metcalf/quantifying_cl_cf_carriers.qmd
@@
     page-layout: full
-    theme:
-      light: flatly
-      dark: darkly
-bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
-csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
 editor: source
@@
 )
-knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Angela Metcalf/Quantifying Sweat Chloride in CF Carriers")
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Angela Metcalf/Quantifying Sweat Chloride in CF Carriers",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Angela Metcalf/Quantifying Sweat Chloride in CF Carriers"
+)
+knitr::opts_knit$set(root.dir = home_dir)
 ```
@@
 #| message: false
-df <- read.csv("./Data_Raw/Midpoint data.csv", na.strings = "")
+#| warning: false
+# Import from REDCap
+unlockREDCap(c(rcon = "Sweat Electrolytes in CF Carriers 24-0691"),
+  keyring = "API_KEYs",
+  envir = 1,
+  url = "https://redcap.ucdenver.edu/api/"
+)
+df <- exportRecordsTyped(rcon)
+# Format
 df <- df %>%
   mutate(
-    `Right Arm Cl` = as.numeric(Sweat.chloride..right.arm),
-    `Left Arm Cl` = as.numeric(Sweat.chloride..left.arm),
-    `Right Arm Na` = as.numeric(Sweat.sodium..right.arm),
-    `Left Arm Na` = as.numeric(Sweat.sodium..left.arm)
+    `Cl Right` = as.numeric(cl_right), `Cl Left` = as.numeric(cl_left),
+    `Na Right` = as.numeric(na_right), `Na Left` = as.numeric(na_left)
   )
-df$`Mean Cl` <- rowMeans(df[, c("Right Arm Cl", "Left Arm Cl")])
-df$`Mean Na` <- rowMeans(df[, c("Right Arm Na", "Left Arm Na")])
+df$`Mean Cl` <- rowMeans(df[, c("Cl Right", "Cl Left")])
+df$`Mean Na` <- rowMeans(df[, c("Na Right", "Na Left")])
+# Per Angela, exclude number 15 due to additional mutation
+df <- df[-which(df$pid == 15), ]
 ```
 
-# Plots
+# Confidence interval power calculations
 
 ```{r}
-ggplot(df,aes(y=`Mean Cl`))+boxplot()
+ns <- 35:45
+sigma <- round(sd(df$`Mean Cl`, na.rm = T), 1)
+power <- lapply(ns, function(n) {
+  se <- qt(0.05 / 2, lower.tail = F, df = n - 1) * (sigma / sqrt(n))
+  return(c(
+    "Confidence level" = 0.95, "Sample size (N)" = n,
+    "Actual Distance from Mean to Limits" = se,
+    "Standard deviation of sweat chloride" = sigma
+  ))
+})
+power <- do.call(rbind, power)
+knitr::kable(power)
 ```

### 2025-04-06 Quarto Template.qmd
fixed template theme

@@
     page-layout: full
-    theme:
-      light: flatly
-      dark: darkly
 bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib

### 2025-04-06 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
Fixed sweat test data for Angela

--- 2025-01-09 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
@@
 library(Hmisc)
+library(lubridate)
+library(readxl)
 library(tidyverse)
-library(readxl)
-library(gtsummary)
-knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability")
+library(arsenal)
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Variability"
+)
+knitr::opts_knit$set(root.dir = home_dir)
 ```
@@
 #| message: false
-# Import data
-# source("/Users/timvigers/Library/CloudStorage/Dropbox/GitHub/CF/Edith Zemanick/Sweat Variability/data_import.r")
-load("./Data_Cleaned/chco_data.RData")
-# Exclude people without demographics
-chco <- chco %>% filter(demog_avail == "True")
-# Fix diagnoses/exclusion data for CHCO (based on Angela's review)
-diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
-exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) |
-  diagnoses$Code == "Exclude"]
-exclude <- exclude[!is.na(exclude)]
-chco <- chco[!(chco$patient_id %in% exclude), ]
-chco$Diagnosis <- diagnoses$Diagnosis[match(chco$patient_id, diagnoses$PatientID)]
-diagnosis_check <- chco %>% select(patient_id, Diagnosis, diagnosis)
-# Import NJH data, remove those without demographics
-njh <- read_excel("./Data_Raw/SweatVariability2025_CFgenotypes 2024-08-21.xlsx")
-# Filter and select similar to CHCO
-njh <- njh %>% filter(`Demographics Available`)
-# Import detailed diagnoses from Katie
-diagnoses <- read.csv("./Data_Cleaned/SweatVariability2025_CFgenotypes.csv",
-  na.strings = c("", ".")
-)
-# Merge NJH and diagnoses
-diagnoses <- diagnoses %>% select(Patient.ID, Diagnosis...DataScout)
-colnames(diagnoses) <- c("PatientID", "Diagnosis")
-# Less detailed diagnoses
-diagnoses$Diagnosis <- factor(diagnoses$Diagnosis,
-  levels = c(
-    "Abnormal sweat choride test without diagnosis of cystic fibrosis",
-    "Abnormal sweat test",
-    "Alpha-1-antitrypsin deficiency",
-    "Alpha-1-antitrypsin deficiency carrier",
-    "Alpha-1-antitrypsin deficiency, Positive cystic fibrosis sweat test without diagnosis of cystic fibrosis",
-    "CF", "CF carrier, Alpha-1-antitrypsin deficiency",
-    "CF, alpha 1-antitrypsin deficiency",
-    "CF, alpha 1-antitrypsin deficiency carrier",
-    "CF, Alpha-1-antitrypsin deficiency",
-    "Cystic fibrosis carrier",
-    "Cystic fibrosis carrier, Alpha-1-antitrypsin deficiency",
-    "Cystic fibrosis carrier, Alpha-1-antitrypsin deficiency carrier",
-    "Cystic fibrosis screening",
-    "Cystic fibrosis transmembrane conductance regulator (CFTR)-related disorder",
-    "Positive cystic fibrosis sweat test without diagnosis of cystic fibrosis",
-    "Primary ciliary dyskinesia",
-    "Primary Ciliary dyskinesia, Cystic Fibrosis carrier"
-  ),
-  labels = c(
-    "Non-CF", "Non-CF", "Non-CF", "CF Carrier", "Non-CF", "Cystic Fibrosis",
-    "CF Carrier", "Cystic Fibrosis", "Cystic Fibrosis", "Cystic Fibrosis",
-    "CF Carrier", "CF Carrier", "CF Carrier", "Non-CF", "CFTR-RD or CRMS",
-    "Non-CF", "Non-CF", "CF Carrier"
+# Import CHCO, filter out those with no demo data, rename columns, etc.
+chco <- read.csv("./Data_Cleaned/chco_reidentified.csv",
+  na.strings = c(
+    "", "Test not performed", "Insufficient quant.",
+    "repeat sweat chloride testing", "Please reorder testing and re",
+    "The received specimen cannot"
   )
 )
-# Patient ids may not match so keep CHCO and NJH separate and check names after
+chco <- chco %>%
+  group_by(PatientID) %>%
+  filter(Demographics.Available == "TRUE") %>%
+  # Get only the earliest test
+  slice_min(Age.at.Test.in.Weeks) %>%
+  select(
+    PatientID, Name, Genotypes1, Genotypes2, Sex, Age.at.Test.in.Years,
+    Na, K, Cl, Na_Left, Na_Right, K_Left, K_Right, Cl_Left, Cl_Right
+  ) %>%
+  rename(Age = Age.at.Test.in.Years) %>%
+  mutate(Name = tolower(gsub("[[:punct:] ]+", "", Name)))
+# Fix CHCO diagnoses per Angela
+chco_diagnoses <- read.csv("./Data_Cleaned/chco_diagnoses.csv")
+chco <- full_join(chco, chco_diagnoses, by = join_by(PatientID))
+chco$PatientID <- NULL
+# For "LESS THAN 10", replace with half the limit of detection
+chco[chco == "LESS THAN 10"] <- "5"
+chco[, c("Cl", "Cl_Left", "Cl_Right")] <-
+  lapply(chco[, c("Cl", "Cl_Left", "Cl_Right")], as.numeric)
+# Import NJH and add diagnoses from Katie
+njh <- read.csv("./Data_Cleaned/njh_full_list.csv",
+  na.strings = c(
+    "", "Test not performed", "Insufficient quant.",
+    "repeat sweat chloride testing"
+  )
+)
+njh_diagnoses <- read.csv("./Data_Cleaned/njh_diagnoses.csv", na.strings = "")
+njh_diagnoses <- njh_diagnoses %>% filter(!is.na(Diagnosis))
+njh <- full_join(njh, njh_diagnoses)
+# Filter out those with no demo data, rename columns, etc.
+njh <- njh %>%
+  group_by(Patient.ID) %>%
+  filter(Demographics.Available == "TRUE") %>%
+  # Get only the earliest test
+  slice_min(Age.at.test..in.years.) %>%
+  select(
+    Patient.ID, Patient.Name, Genotype.1, Genotype.2, Sex, Age.at.test..in.years.,
+    Sweat.sodium, Sweat.potassium, Sweat.chloride, Sweat.sodium.left.arm,
+    Sweat.sodium.right.arm, Sweat.potassium.left.arm, Sweat.potassium.right.arm,
+    Sweat.chloride.left.arm, Sweat.chloride.right.arm, Diagnosis
+  ) %>%
+  rename(
+    Name = Patient.Name, Genotypes1 = Genotype.1,
+    Genotypes2 = Genotype.2, Age = Age.at.test..in.years., Na = Sweat.sodium,
+    K = Sweat.potassium, Cl = Sweat.chloride, Na_Left = Sweat.sodium.left.arm,
+    Na_Right = Sweat.sodium.right.arm, K_Left = Sweat.potassium.left.arm,
+    K_Right = Sweat.potassium.right.arm, Cl_Left = Sweat.chloride.left.arm,
+    Cl_Right = Sweat.chloride.right.arm
+  ) %>%
+  mutate(Name = tolower(gsub("[[:punct:] ]+", "", Name)))
+njh$Patient.ID <- NULL
+# Row bind, fill diagnosis, then remove dupes
+df <- rbind(chco, njh)
+df <- df %>%
+  group_by(Name) %>%
+  fill(Diagnosis) %>%
+  distinct() %>%
+  select(Name, Diagnosis, everything()) %>%
+  arrange(Name, Age)
+# Use each person's first sweat test
+df <- df %>%
+  group_by(Name) %>%
+  slice_min(Age)
+# If there are still duplicates, exclude them
+dups <- df$Name[duplicated(df$Name)]
+df <- df[!df$Name %in% dups, ]
+# For those with left and right, average them
+df$Na_mean <- rowMeans(df[, c("Na_Left", "Na_Right")])
+df$Na <- coalesce(df$Na, df$Na_mean)
+df$Cl_mean <- rowMeans(df[, c("Cl_Left", "Cl_Right")])
+df$Cl <- coalesce(df$Cl, df$Cl_mean)
+df$K_mean <- rowMeans(df[, c("K_Left", "K_Right")])
+df$K <- coalesce(df$K, df$K_mean)
+# Get rid of unnecessary columns, remove those with no diagnosis
+df <- df %>%
+  select(Name, Diagnosis, Genotypes1, Genotypes2, Sex, Age, Na, K, Cl) %>%
+  filter(!is.na(Diagnosis))
+# Set table options
+mycontrols <-
+  tableby.control(numeric.stats = c("Nmiss", "meansd", "medianq1q3", "range"))
+# Save this for Angela et al. to check
+write.csv(df,
+  file = "./Data_Cleaned/sweat_variability_analysis_dataset.csv",
+  row.names = F, na = ""
+)
 ```
 
-```{r}
-#| message: false
-#| warning: false
-df %>%
-  ggplot(
-    aes(x = Na, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
-  ) +
-  geom_point() +
-  theme()
-df %>%
-  ggplot(
-    aes(x = Cl, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
-  ) +
-  geom_point() +
-  theme()
-df %>%
-  ggplot(
-    aes(x = Na / Cl, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
-  ) +
-  geom_point() +
-  theme()
+# Data cleaning steps
+
+- Imported re-identified CHCO data from Elin (5,046 rows and 4,748 unique IDs)
+- Removed rows with `Demographics Available` not equal to `TRUE` and selected each person's earliest test by age.
+  - This resulted in 3,123 rows with 3,084 unique IDs.
+- Merged in the diagnoses from Angela.
+- Values below the limit of detection were replaced with half the limit (e.g. "LESS THAN 10" was replaced with 5). Approximately 5% of the values were below the L.O.D.
+
+- Imported NJH data (1,958 rows with 1,748 unique IDs) and merged in diagnoses from Katie.
+- Removed 1,662 rows with `Demographics Available` not equal to `TRUE` (all of these were `FALSE`)
+  - This resulted in 172 rows with unique IDs.
+- Concatenated the CHCO and NJH data and removed duplicate rows, resulting in 3,178 rows with 3,162 unique IDs.
+- For the remaining people with multiple sweat tests (i.e. tests from both CHCO and NJH), kept only the earliest test (based on age at test). This resulted in 3,169 rows with 3,162 unique IDs.
+- The remaining 7 people with different values for the same test in the CHCO and NJH data were excluded. The final analysis dataset includes 3,155 unique participants, 3,111 of which have a diagnosis.
+- For people with both left and right arm values, the two were averaged.
+
+# Table 1: Participant Demographics by Diagnosis
+
+```{r results='asis'}
+summary(tableby(Diagnosis ~ Sex + Age + Na + K + Cl,
+  data = df, control = mycontrols
+))
 ```
 
+# Boxplots
+
 ```{r}
-df %>%
-  filter(Diagnosis == "Cystic Fibrosis") %>%
-  mutate(
-    `Age group` = cut(Age.at.Test.in.Years, c(-Inf, 18, Inf), right = F),
-    `Cl < 60` = Cl < 60
-  ) %>%
-  select(Na, Cl, TotalGenotypeFunction, `Cl < 60`, `Age group`) %>%
-  tbl_summary(by = `Age group`, missing_text = "Missing") %>%
-  add_p()
+ggplot(df, aes(x = Diagnosis, y = Cl)) +
+  geom_boxplot() +
+  theme_bw()
+ggplot(df, aes(x = Diagnosis, y = Na)) +
+  geom_boxplot() +
+  theme_bw()
+ggplot(df, aes(x = Diagnosis, y = Na / Cl)) +
+  geom_boxplot() +
+  theme_bw()
+ggplot(df, aes(x = Diagnosis, y = K)) +
+  geom_boxplot() +
+  theme_bw()
 ```

### 2025-04-06 Edith Zemanick/Sweat Variability/sweat_test_analysis_chec.qmd
CHEC analyses for NACFC

@@
     page-layout: full
-    theme:
-      light: flatly
-      dark: darkly
-bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
-csl: /home/timvigers/Documents/GitHub/styles/american-medical-association.csl
 editor: source
@@
 library(gt)
-knitr::opts_knit$set(root.dir = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Variability")
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CFEdith Zemanick/Sweat Variability"
+)
+knitr::opts_knit$set(root.dir = home_dir)
 ```
@@
 The model using the ratio alone does not really work due to the fitted probabilities issue.
+
+# NACFC
+
+## Normalized ratio by modulator status
+
+```{r results='asis'}
+# Check if ratio >= 1.2
+chec$post_ratio_1.2 <- chec$post_ratio >= 1.2
+label(chec$post_ratio_1.2) <- "Post-Modulator Sweat Na:Cl >= 1.2"
+summary(tableby(modqual ~ post_ratio + post_ratio_1.2,
+  data = chec,
+  control = list(
+    numeric.stats = c("N", "Nmiss2", "meanCI", "range"),
+    cat.stats = c("N", "Nmiss2", "countpct")
+  )
+), pfootnote = T)
+```
+
+## Scatterplots pre-modulator
+
+```{r}
+chec_long %>%
+  filter(pre_post == "pre") %>%
+  ggplot(., aes(x = ratio, y = cl)) +
+  geom_point() +
+  ylab("Cl") +
+  xlab("Na:Cl") +
+  theme_bw() +
+  facet_wrap(~modqual)
+```
+
+## Scatterplots post-modulator
+
+```{r}
+chec_long %>%
+  filter(pre_post == "post") %>%
+  ggplot(., aes(x = ratio, y = cl)) +
+  geom_point() +
+  ylab("Cl") +
+  xlab("Na:Cl") +
+  theme_bw() +
+  facet_wrap(~modqual)
+```
+
+## Boxplot
+
+```{r}
+plot_df <- chec_long
+plot_df$modqual[plot_df$pre_post == "pre"] <- "Pre-Modulator"
+plot_df$modqual <- factor(plot_df$modqual,
+  levels = c(
+    "Pre-Modulator", "Lumacaftor / Ivacaftor", "Tezacaftor / Ivacaftor",
+    "Ivacaftor", "Elexacaftor / Tezacaftor / Ivacaftor"
+  )
+)
+ggplot(plot_df,aes(x=modqual,y=ratio))+geom_boxplot()+
+  ylab("Na:Cl") +
+  theme_bw()+
+  theme(axis.title.x = element_blank())
+```
+
+## Change plot
+
+```{r}
+ggplot(chec_long, aes(x = pre_post, y = ratio, group = subject)) +
+  geom_point(alpha = 0.2) +
+  geom_line(alpha = 0.2) +
+  ylab("Na:Cl") +
+  theme_classic() +
+  theme(axis.title.x = element_blank())
+```

### 2025-04-08 Edith Zemanick/Sweat Variability/sweat_test_analysis_chec.qmd
limit CHEC analysis to those with post-mod sodium

@@
 # Only include people with either a pre-modulator or post-modulator sodium
-chec <- chec %>% filter(!is.na(swhsavgpm) | !is.na(swsavg))
+chec <- chec %>% filter(!is.na(swsavg))
 # Calculate ratios
@@
 ```{r results='asis'}
-f <- modqual~sex + age + agecat_txt + race_txt + genotype_group
-t1 <- tableby(f, data = chec, control = mycontrols)
+t1 <- tableby(
+  modqual ~ sex + kwt(age, "N", "Nmiss2", "medianq1q3") + agecat_txt +
+    race_txt + genotype_group + swcavg + swhcavgpm + swsavg + swhsavgpm,
+  data = chec, cat.test = "fe", simulate.p.value = TRUE
+)
 summary(t1, pfootnote = T)
@@
 )
-ggplot(plot_df,aes(x=modqual,y=ratio))+geom_boxplot()+
+ggplot(plot_df, aes(x = modqual, y = ratio)) +
+  geom_boxplot() +
   ylab("Na:Cl") +
-  theme_bw()+
+  theme_bw() +
   theme(axis.title.x = element_blank())

### 2025-04-14 Christine Chan/CF Registry/cf_registry_CFRD_prediction.qmd
NACFC joint model update

@@
 library(tidyr)
+library(arsenal)
+library(knitr)
 home_dir <- switch(Sys.info()["sysname"],
@@
 annual$bmi <- coalesce(annual$A_bmivalue, annual$bmi)
+annual$Gender <- factor(annual$Gender, levels = c("M", "F"))
 # Labels
@@
 )
-# Make survival datasets for registry and OGTT diagnoses
-ogtt <- annual %>%
+# Make datasets
+registry <- annual %>%
+  filter(Age_YrEnd <= cfrd_registry_age) %>%
   select(
-    eDWID, ReviewYear, Age_YrEnd, A_hgba1c, all_of(predictors), cfrd_ogtt
+    eDWID, ReviewYear, Age_YrEnd, A_hgba1c, all_of(predictors), cfrd_registry
   ) %>%
@@
   distinct()
-surv_ogtt <- annual %>%
-  filter(Age_YrEnd <= cfrd_ogtt_age) %>%
+surv_reg <- annual %>%
+  filter(Age_YrEnd <= cfrd_registry_age) %>%
   group_by(eDWID) %>%
   slice_max(row_number()) %>%
-  select(eDWID, ReviewYear, Age_YrEnd, all_of(predictors), cfrd_ogtt) %>%
+  select(eDWID, ReviewYear, Age_YrEnd, all_of(predictors), cfrd_registry) %>%
   drop_na() %>%
   distinct()
-ogtt_include <- intersect(ogtt$eDWID, surv_ogtt$eDWID)
+reg_include <- intersect(registry$eDWID, surv_reg$eDWID)
+```
+
+# Demographics
+
+```{r,results='asis'}
+t <- registry[registry$eDWID %in% reg_include, ] %>%
+  group_by(eDWID) %>%
+  summarise(n())
+
+t1 <- tableby(cfrd_registry ~ Age_YrEnd + Gender + Race + Hispanicrace + A_IsOnEnzymes + NumPulmExacerbation + A_FEV1_pct_predicted + bmi, data = surv_reg[surv_reg$eDWID %in% reg_include, ])
+summary(t1)
 ```
@@
 cox_f <- as.formula(paste0(
-  "Surv(time = Age_YrEnd, event = cfrd_ogtt)~",
+  "Surv(time = Age_YrEnd, event = cfrd_registry)~",
   paste0(predictors, collapse = "+")
@@
   cluster = eDWID, x = T, model = T,
-  data = surv_ogtt[surv_ogtt$eDWID %in% ogtt_include, ]
+  data = surv_reg[surv_reg$eDWID %in% reg_include, ]
 )
@@
   random = ~ Age_YrEnd | eDWID, na.action = na.omit,
-  data = ogtt[ogtt$eDWID %in% ogtt_include, ]
+  data = registry[registry$eDWID %in% reg_include, ]
 )
 # Combine
-joint_fit <- jointModel(lme_fit_a1c, cox_fit, timeVar = "Age_YrEnd")
+joint_fit_value <- jointModel(lme_fit_a1c, cox_fit, timeVar = "Age_YrEnd", parameterization = "value")
+# joint_fit_slope <- jointModel(lme_fit_a1c, cox_fit, timeVar = "Age_YrEnd",parameterization ="slope")
+# joint_fit_both <- jointModel(lme_fit_a1c, cox_fit, timeVar = "Age_YrEnd",parameterization ="both")
+s <- summary(joint_fit_value)
+long <- data.frame(s$`CoefTable-Long`)
+long <- long[, c(1, 4)]
+long$p.value <- format.pval(long$p.value, eps = 0.001, digits = 3)
+event <- data.frame(s$`CoefTable-Event`)
+event <- event[, c(1, 4)]
+event$Value <- exp(event$Value)
+rownames(event)[nrow(event)] <- "shape"
+event$p.value <- format.pval(event$p.value, eps = 0.001, digits = 3)
+# summary(joint_fit_slope)
+# summary(joint_fit_both)
+kable(long, caption = "Longitudinal Process", digits = 3)
+kable(event, caption = "Event Process", digits = 3)
 ```
+
+# To do
 
@@
   - Could also try using the half interval model Kristen described
+
+https://erandrinopoulou.github.io/EducationalCorner_JMpred/

### 2025-04-14 Quarto Template.qmd
NACFC joint model update

@@
 ```{r data cleaning}
-#| message: false
 # Import from REDCap

### 2025-04-14 Christine Chan/EnVision CF/fsOGTT_pre_post_ETI.qmd
NACFC fsOGTT analysis start

---
title: "fsOGTT Pre- and Post-ETI"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(readxl)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/EnVision CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
# Import cleaned EnVision data
df <- read.csv("./Data_Clean/envision_analysis_dataset.csv", na.strings = "")
# Import GEM data
gem <- read.csv(
  "./Data_Raw/CGMInHealthyControls-FsOGTTDataForMonique_DATA_2025-04-10_0903.csv",
  na.strings = ""
)
# ID linking from Monique
ids <- read_excel("./Data_Raw/ETI participant IDs_TV.xlsx")
ids <- ids[!is.na(ids$`GEM ID`), ]
ids$`EnVision ID` <- sub("-2", "", ids$`EnVision ID`)
# Select columns for EnVision per Monique and combine race and ethnicity
df <- df %>%
  select(
    study_id, Date, age_visit, sex, origin_race, ethnicity, CFTR,
    pancreatic_status, corrector___4, bmi, fev1,
    Diagnosis, Glucose_0:Glucose_180, Insulin_0:Insulin_180,
    C.Peptide_0:C.Peptide_180
  ) %>%
  unite(race_ethnicity, c(origin_race, ethnicity)) %>%
  rename(ETI = corrector___4)
df$race_ethnicity <- factor(df$race_ethnicity,
  levels = c(
    "American Indian or Alaska Native (N. S. & Central America)_Hispanic/Latino",
    "American Indian or Alaska Native (N. S. & Central America)_Not Hispanic/Latino",
    "Black or African American (Africa)_N/A",
    "Black or African American (Africa)_Not Hispanic/Latino",
    "Declined_Hispanic/Latino",
    "Declined_Not Hispanic/Latino",
    "NA_NA",
    "Native Hawaiian or Other Pacific Islander (Hawaii, Guam, Samoa, Pacific Islands)_Not Hispanic/Latino",
    "White (Europe, Middle East, North Africa)_Hispanic/Latino",
    "White (Europe, Middle East, North Africa)_N/A",
    "White (Europe, Middle East, North Africa)_Not Hispanic/Latino"
  ),
  labels = c(
    "Hispanic", "Native American", "Black/African American", "Black/African American",
    "Hispanic", "Other/Multiple", "Other/Multiple", "Other/Multiple", "Hispanic", "White", "White"
  )
)
df$pancreatic_status <- factor(df$pancreatic_status, levels = 1:2)
df$ETI <- factor(df$ETI, levels = 0:1, labels = c("No", "Yes"))
# Select the same columns from GEM
gem <- gem %>%
  mutate(ETI = 0) %>%
  select(
    subject_id, date_of_visit, age_at_visit_1, gender, race_ethnicity,
    cf_genotype, cf_pancreatic, ETI, average_bmi, fev1_percent,
    group_category, lab_ogtt_fasting, min_10_glucose, min_30_glucose,
    lab_ogtt_1_hour_glucose, min_90_glucose, lab_ogtt_2_hour_glucose,
    min_150_glucose, min_180_glucose, min_0_insulin, min_10_insulin,
    min_30_insulin, min_60_insulin, min_90_insulin, min_120_insulin,
    min_150_insulin, min_180_insulin, min_0_c_peptide, min_10_c_peptide,
    min_30_c_peptide, min_60_c_peptide, min_90_c_peptide, min_120_c_peptide,
    min_150_c_peptide, min_180_c_peptide
  ) %>%
  rename(
    study_id = subject_id, Date = date_of_visit, age_visit = age_at_visit_1,
    sex = gender, CFTR = cf_genotype, pancreatic_status = cf_pancreatic,
    bmi = average_bmi, fev1 = fev1_percent, Diagnosis = group_category,
    min_0_glucose = lab_ogtt_fasting, min_60_glucose = lab_ogtt_1_hour_glucose,
    min_120_glucose = lab_ogtt_2_hour_glucose
  ) %>%
  rename_with(
    ~ paste0("Glucose_", parse_number(.)), min_0_glucose:min_180_glucose
  ) %>%
  rename_with(
    ~ paste0("Insulin_", parse_number(.)), min_0_insulin:min_180_insulin
  ) %>%
  rename_with(
    ~ paste0("C.Peptide_", parse_number(.)), min_0_c_peptide:min_180_c_peptide
  )
# GEM factors
gem$sex <- factor(gem$sex, levels = 1:2, labels = c("Male", "Female"))
gem$race_ethnicity <- factor(gem$race_ethnicity,
  levels = 1:6,
  labels = c(
    "White", "Black/African American", "Asian", "Native American",
    "Hispanic", "Other/Multiple"
  )
)
gem$CFTR <- factor(gem$CFTR,
  levels = c(
    "2184delA/M1101K", "2585delT/G551D", "3849+10kb C>T/F508del",
    "508del/508del", "F508del/1078delA", "f508DEL/1154INStc",
    "F508del/1154insTC", "F508del/1811 + 1634A > G", "F508del/1811+1634>G",
    "F508del/2183delAA->G", "F508del/2183delAA>G", "F508del/2184insA",
    "F508del/2711delT", "F508del/2789+5G>A", "F508del/27895GA",
    "F508del/3659delC", "F508del/3849+10kbC>T", "F508del/394delTT",
    "F508del/5492F", "F508del/A561E", "F508del/E60X",
    "f508del/F508del", "F508del/F508del", "F508del/G542X",
    "F508del/G551D", "F508del/N1303K", "F508del/Q98R",
    "F508del/R1162X", "F508del/R117H 9T", "F508del/R117H 9T/5T",
    "F508del/R553X", "F508del/R709X", "F508del/S492F",
    "F508del/W1282X", "F5o8del/F508del", "G542X/1078delT",
    "G542X/W1089X", "G551D/W846X", "I507del/G551D",
    "N1303K/1154insTC", "N1303K/2789 + 5G > A", "N1303K/2789+5G>A",
    "R117H 5t/G576A and R668C on exon 13", "R117H5t/G576A and R668C on exon 13",
    "V520F/3791delC", "W1089X/S492F", "W1090X/S492F"
  ),
  labels = c(
    "Other/Other", "Other/Other", "Other/Other", "F508del homozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del homozygous",
    "F508del homozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del homozygous", "Other/Other", "Other/Other", "Other/Other",
    "Other/Other", "Other/Other", "Other/Other", "Other/Other", "Other/Other",
    "Other/Other", "Other/Other", "Other/Other", "Other/Other"
  )
)
gem$pancreatic_status <- factor(gem$pancreatic_status, levels = 1:2)
gem$ETI <- factor(gem$ETI, levels = 0:1, labels = c("No", "Yes"))
gem$Diagnosis <- factor(gem$Diagnosis,
  levels = 1:4,
  labels = c("Healthy Control", "NGT", "AGT", "CFRD")
)
df$Diagnosis <- factor(df$Diagnosis,
  levels = c("NGT", "IFG", "IGT", "INDET", "CFRD"),
  labels = c("NGT", "AGT", "AGT", "AGT", "CFRD")
)
# Limit GEM per Monique's list
gem <- gem %>% filter(study_id %in% ids$`GEM ID`)
# Convert to EnVision IDs
gem$study_id <- ids$`EnVision ID`[match(gem$study_id, ids$`GEM ID`)]
# Merge
df <- full_join(df, gem)
```

- Make the plot of number of people with OGTTs pre and post
- If people were on Ivacaftor prior to ETI we can still include those visits as pre-ETI
- Adjust models for age
- Paneled figure with results for glucose, c-peptide, and insulin
- Look at acute insulin response (change in 30 minute c peptide)/(change in 30 minute glucose)
  - Can also do across the whole time period and can use insulin in the numerator also
- See Teams folder document for detailed plan

# Table

```{r}
#| label: tbl-planets
#| tbl-cap: Planets
```

# Figure

```{r}
#| label: fig-planets
#| fig-cap: Planets
```

### 2025-04-14 Christine Chan/EnVision CF/clean_envision_data.R

@@ -92,7 +92,8 @@ visits <- redcap %>%
   select(
     study_id, redcap_data_access_group, date_visit, sex, origin_race, ethnicity,
     cftr_mutation_1, cftr_mutation_2, pancreatic_status, liver_disease,
-    pulm_ex_last3months, corrector_yes_no, family_history_diabetes,
+    pulm_ex_last3months, family_history_diabetes,
+    corrector_yes_no, corrector_start_date, corrector___1:corrector___5,
     age_visit, height, weight, bmi_study_visit, fev1, fvc
   ) %>%
   filter(!is.na(date_visit)) %>%
@@ -459,6 +460,22 @@ final_df$CFTR <- factor(final_df$CFTR,
 )
 final_df$cftr_mutation_1 <- NULL
 final_df$cftr_mutation_2 <- NULL
+# Race and ethnicity
+final_df$origin_race <- factor(final_df$origin_race,
+  levels = c("1", "2", "3", "5", "6"),
+  labels = c(
+    "White (Europe, Middle East, North Africa)",
+    "Black or African American (Africa)",
+    "American Indian or Alaska Native (N. S. & Central America)",
+    "Native Hawaiian or Other Pacific Islander (Hawaii, Guam, Samoa, Pacific Islands)",
+    "Declined"
+  )
+)
+final_df$ethnicity <- factor(final_df$ethnicity,
+  levels = 1:3, labels = c("Hispanic/Latino", "Not Hispanic/Latino", "N/A")
+)
+# Sex
+final_df$sex <- factor(final_df$sex, levels = 1:2, labels = c("Male", "Female"))
 # Final formatting
 final_df$redcap_data_access_group <- gsub("\\d", "", final_df$study_id)
 final_df$redcap_data_access_group <- factor(

### 2025-04-16 Angela Metcalf/quantifying_cl_cf_carriers.qmd
NACFC updates

@@
   )
-df$`Mean Cl` <- rowMeans(df[, c("Cl Right", "Cl Left")])
-df$`Mean Na` <- rowMeans(df[, c("Na Right", "Na Left")])
-# Per Angela, exclude number 15 due to additional mutation
-df <- df[-which(df$pid == 15), ]
+df$`Mean Cl` <- rowMeans(df[, c("Cl Right", "Cl Left")], na.rm = T)
+df$`Mean Na` <- rowMeans(df[, c("Na Right", "Na Left")], na.rm = T)
+df$Ratio <- df$`Mean Na` / df$`Mean Cl`
+# Per Angela, exclude some participants
+df <- df[!df$pid %in% c(15, 36), ]
+```
+
+# NACFC 2025
+
+```{r results='asis'}
+interim <- read.csv("./Data_Clean/Interim reduced data set.csv")
+interim$`Chloride >= 30`=interim$Chloride >= 30
+summary(tableby(
+  Carrier ~ kwt(Average.sodium, "medianq1q3", "iqr") +
+    kwt(Chloride, "medianq1q3", "iqr") + kwt(Ratio, "medianq1q3", "iqr") +
+    `Chloride >= 30`,
+  data = interim
+), pfootnote = T)
 ```

### 2025-04-16 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
NACFC updates

@@
 ```{r}
-ggplot(df, aes(x = Diagnosis, y = Cl)) +
+clp <- ggplot(df, aes(x = Diagnosis, y = Cl)) +
   geom_boxplot() +
   theme_bw()
-ggplot(df, aes(x = Diagnosis, y = Na)) +
+clp
+nap <- ggplot(df, aes(x = Diagnosis, y = Na)) +
   geom_boxplot() +
   theme_bw()
-ggplot(df, aes(x = Diagnosis, y = Na / Cl)) +
+nap
+ratiop <- ggplot(df, aes(x = Diagnosis, y = Na / Cl)) +
+  geom_boxplot() +
+  theme_bw() +
+  ylab("Sweat sodium to chloride ratio")+
+  xlab("Modulator")
+ratiop
+kp <- ggplot(df, aes(x = Diagnosis, y = K)) +
   geom_boxplot() +
   theme_bw()
-ggplot(df, aes(x = Diagnosis, y = K)) +
-  geom_boxplot() +
-  theme_bw()
+kp
 ```
+
+# NACFC 2025
+
+```{r}
+```

### 2025-04-16 Edith Zemanick/Sweat Variability/sweat_test_analysis_chec.qmd
NACFC updates

@@
 library(performance)
+library(gtsummary)
 library(gt)
+library(patchwork)
 home_dir <- switch(Sys.info()["sysname"],
   "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CFEdith Zemanick/Sweat Variability"
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Variability"
 )
@@
 label(chec$post_ratio_1.2) <- "Post-Modulator Sweat Na:Cl >= 1.2"
-summary(tableby(modqual ~ post_ratio + post_ratio_1.2,
+chec$post_cl <- cut(chec$swcavg, c(-Inf, 30, 60, Inf),
+  right = F,
+  labels = c("< 30", ">= 30 and < 60", ">= 60")
+)
+label(chec$post_cl) <- "Post-Modulator Sweat Cl Group"
+summary(tableby(modqual ~ post_cl + post_ratio + post_ratio_1.2,
   data = chec,
+  control = list(
+    numeric.stats = c("N", "Nmiss2", "meanCI", "range"),
+    cat.stats = c("N", "Nmiss2", "countpct")
+  )
+), pfootnote = T)
+```
+
+## Normalized ratio by modulator status and age
+
+We fit a logistic regression model with post-modulator ratio >= 1.2 as the binary outcome. The model with an interaction term was significantly better than the mode with only main effects (p=0.003).
+
+```{r results='asis'}
+m1 <- glm(post_ratio_1.2 ~ modqual + age, data = chec, family = "binomial")
+m1 %>% tbl_regression(., intercept = T, exponentiate = T)
+```
+
+The main effects suggest that people on LUM/IVA or TEZ/IVA were less likely to have a post-modulator ratio >= 1.2 compared to those on ETI, but there was no difference for those on Ivacaftor, after adjusting for age.
+Also, there was no effect of age on the odds of having a post-modulator ratio >= 1.2.
+
+## Normalized ratio by post-ETI sweat chloride
+
+```{r results='asis'}
+summary(tableby(post_cl ~ post_ratio + post_ratio_1.2,
+  data = chec[chec$modqual == "Elexacaftor / Tezacaftor / Ivacaftor", ],
   control = list(
@@
 ```{r}
-chec_long %>%
+sp_pre <- chec_long %>%
   filter(pre_post == "pre") %>%
@@
   geom_point() +
+  geom_hline(yintercept = 60, color = "red") +
+  geom_vline(xintercept = 1.2, color = "red") +
   ylab("Cl") +
@@
   theme_bw() +
-  facet_wrap(~modqual)
+  facet_wrap(~modqual) +
+  ggtitle("Pre-Modulator")
+sp_pre
 ```
@@
 ```{r}
-chec_long %>%
+sp_post <- chec_long %>%
   filter(pre_post == "post") %>%
@@
   theme_bw() +
-  facet_wrap(~modqual)
+  facet_wrap(~modqual) +
+  ggtitle("Post-Modulator")
+sp_post
 ```
@@
 )
-ggplot(plot_df, aes(x = modqual, y = ratio)) +
+boxp <- ggplot(plot_df, aes(x = modqual, y = ratio)) +
   geom_boxplot() +
@@
   theme(axis.title.x = element_blank())
+boxp
+```
+
+## Combined figure
+
+```{r}
+boxp / (sp_pre + sp_post)
 ```

### 2025-04-17 Christine Chan/CF Registry/cf_registry_CFRD_prediction.qmd
joint model SEs

@@
   summarise(n())
-
-t1 <- tableby(cfrd_registry ~ Age_YrEnd + Gender + Race + Hispanicrace + A_IsOnEnzymes + NumPulmExacerbation + A_FEV1_pct_predicted + bmi, data = surv_reg[surv_reg$eDWID %in% reg_include, ])
+t2 <- registry[registry$eDWID %in% reg_include, ] %>%
+  group_by(eDWID) %>%
+  filter(row_number() == 1)
+t1 <- tableby(cfrd_registry ~ Age_YrEnd + Gender + Race + Hispanicrace + A_IsOnEnzymes + NumPulmExacerbation + A_FEV1_pct_predicted + bmi, data = t2)
 summary(t1)
@@
 # Combine
-joint_fit_value <- jointModel(lme_fit_a1c, cox_fit, timeVar = "Age_YrEnd", parameterization = "value")
+joint_fit_value <- jointModel(lme_fit_a1c, cox_fit, timeVar = "Age_YrEnd")
 # joint_fit_slope <- jointModel(lme_fit_a1c, cox_fit, timeVar = "Age_YrEnd",parameterization ="slope")
@@
 long <- data.frame(s$`CoefTable-Long`)
-long <- long[, c(1, 4)]
 long$p.value <- format.pval(long$p.value, eps = 0.001, digits = 3)
 event <- data.frame(s$`CoefTable-Event`)
-event <- event[, c(1, 4)]
 event$Value <- exp(event$Value)

### 2025-04-17 Christine Chan/EnVision CF/fsOGTT_pre_post_ETI.qmd
EnVision fsOGTT

@@
 library(readxl)
+library(lmerTest)
+library(performance)
+library(gtsummary)
+library(emmeans)
 home_dir <- switch(Sys.info()["sysname"],
@@
 ```{r data cleaning}
+#| message: false
 # Import cleaned EnVision data
@@
 ids$`EnVision ID` <- sub("-2", "", ids$`EnVision ID`)
+# Get pre-/post-ETI
+df$eti_start <- ifelse(df$corrector___4 == 1, df$corrector_start_date, NA)
+df <- df %>%
+  group_by(study_id) %>%
+  fill(eti_start, .direction = "downup")
+df$pre_post <-
+  factor(df$Date >= df$eti_start, levels = c(F, T), labels = c("Pre", "Post"))
 # Select columns for EnVision per Monique and combine race and ethnicity
@@
     study_id, Date, age_visit, sex, origin_race, ethnicity, CFTR,
-    pancreatic_status, corrector___4, bmi, fev1,
+    pancreatic_status, corrector___4, pre_post, bmi, fev1,
     Diagnosis, Glucose_0:Glucose_180, Insulin_0:Insulin_180,
@@
 gem <- gem %>%
-  mutate(ETI = 0) %>%
+  mutate(ETI = 0, pre_post = "Pre") %>%
   select(
     subject_id, date_of_visit, age_at_visit_1, gender, race_ethnicity,
-    cf_genotype, cf_pancreatic, ETI, average_bmi, fev1_percent,
+    cf_genotype, cf_pancreatic, ETI, pre_post, average_bmi, fev1_percent,
     group_category, lab_ogtt_fasting, min_10_glucose, min_30_glucose,
@@
 gem$study_id <- ids$`EnVision ID`[match(gem$study_id, ids$`GEM ID`)]
-# Merge
-df <- full_join(df, gem)
+# Merge, keep only pre/post
+df <- full_join(df, gem) %>% filter(!is.na(pre_post))
+df$pre_post <- factor(df$pre_post, levels = c("Pre", "Post"))
 ```
 
-- Make the plot of number of people with OGTTs pre and post
-- If people were on Ivacaftor prior to ETI we can still include those visits as pre-ETI
+# Plot
+
+```{r}
+df %>%
+  group_by(study_id) %>%
+  count(pre_post) %>%
+  ggplot(., aes(y = study_id, x = n)) +
+  geom_point() +
+  theme_bw() +
+  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
+  facet_wrap(~pre_post)
+```
+
 - Adjust models for age
@@
 
-# Table
+# Glucose
 
 ```{r}
-#| label: tbl-planets
-#| tbl-cap: Planets
+# Make a long DF for models
+long_df <- df %>% pivot_longer(Glucose_0:Glucose_180,
+  names_to = c(".value", "timepoint"),
+  names_pattern = "(.*)_(.*)",
+  values_to = "Glucose"
+)
+long_df$timepoint <- factor(long_df$timepoint,
+  levels = c("0", "10", "30", "60", "90", "120", "150", "180")
+)
+mod1 <- lmer(
+  Glucose ~ timepoint * pre_post + age_visit + (1 | study_id),
+  data = long_df
+)
+mod1 %>% tbl_regression(.)
+means <- data.frame(emmeans(mod1, ~ timepoint + pre_post))
+knitr::kable(means)
+ggplot(means, aes(
+  x = as.numeric(as.character(timepoint)), y = emmean,
+  color = pre_post, group = pre_post
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 1) +
+  geom_line() +
+  theme_bw() +
+  xlab("Timepoint") +
+  ylab("Glucose (mg/dL)") +
+  scale_x_continuous(breaks = c(0, 10, 30, 60, 90, 120, 150, 180))
 ```
-
-# Figure
-
-```{r}
-#| label: fig-planets
-#| fig-cap: Planets
-```

### 2025-04-17 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
GCT CFRD + INDET

@@
 gctcap_roc <- roc(Dysglycemia ~ `GCT 60 min capillary`, data = df)
+ogtt0_roc <- roc(Dysglycemia ~ `OGTT 0 min`, data = df)
 ogtt1_roc <- roc(Dysglycemia ~ `OGTT 60 min`, data = df)
@@
   lapply(
-    list(gctpl_roc, gctcap_roc, ogtt1_roc, rpg_roc, rcg_roc, a1c_roc),
+    list(gctpl_roc, gctcap_roc, ogtt0_roc, ogtt1_roc, rpg_roc, rcg_roc, a1c_roc),
     function(roc) {
@@
 roc_table <- do.call(rbind, roc_table)
-rownames(roc_table) <- c("GCTpl", "GCTcap", "OGTT 1 hr", "RPG", "RCG", "A1c")
+rownames(roc_table) <- c("GCTpl", "GCTcap", "OGTT 0 min", "OGTT 1 hr", "RPG", "RCG", "A1c")
 gt(roc_table, rownames_to_stub = T) %>% fmt_number(decimals = 3)
@@
 gt(data.frame(coords(gctcap_roc, "all", ret = "all"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+#### OGTT 0 min
+
+```{r}
+gt(data.frame(coords(ogtt0_roc, "all", ret = "all"))) %>%
   fmt_number(n_sigfig = 3)
@@
 gctcap_roc <- roc(CFRD ~ `GCT 60 min capillary`, data = df)
+ogtt0_roc <- roc(CFRD ~ `OGTT 0 min`, data = df)
 ogtt1_roc <- roc(CFRD ~ `OGTT 60 min`, data = df)
@@
 invisible(lapply(
-  list(gctpl_roc, gctcap_roc, ogtt1_roc, rpg_roc, rcg_roc, a1c_roc),
+  list(gctpl_roc, gctcap_roc, ogtt0_roc, ogtt1_roc, rpg_roc, rcg_roc, a1c_roc),
   function(roc) {
@@
 ```
+
+## CFRD + INDET
+
+CFRD + INDET was defined as a one hour OGTT $\geq$ 200 mg/dL OR two hour OGTT $\geq$ 200 mg/dL.
+
+```{r}
+df$CFRD_INDET <- factor(df$Diagnosis,
+  levels = c("NGT", "IFG", "INDET", "IGT", "CFRD"),
+  labels = c("No", "No", "Yes", "No", "Yes")
+)
+```
+
+### Youden indices
+
+```{r}
+#| label: tbl-roc-cfrd-indet
+#| tbl-cap: Accuracy of different screening tests to predict CFRD + INDET
+#| message: false
+gctpl_roc <- roc(CFRD_INDET ~ `GCT 60 min plasma`, data = df)
+gctcap_roc <- roc(CFRD_INDET ~ `GCT 60 min capillary`, data = df)
+ogtt0_roc <- roc(CFRD_INDET ~ `OGTT 0 min`, data = df)
+ogtt1_roc <- roc(CFRD_INDET ~ `OGTT 60 min`, data = df)
+rpg_roc <- roc(CFRD_INDET ~ `GCT 0 min plasma`, data = df)
+rcg_roc <- roc(CFRD_INDET ~ `GCT 0 min capillary`, data = df)
+a1c_roc <- roc(CFRD_INDET ~ HbA1c, data = df)
+roc_table <-
+  lapply(
+    list(gctpl_roc, gctcap_roc, ogtt0_roc, ogtt1_roc, rpg_roc, rcg_roc, a1c_roc),
+    function(roc) {
+      coords(roc, "best", ret = "all")
+    }
+  )
+roc_table <- do.call(rbind, roc_table)
+rownames(roc_table) <- c("GCTpl", "GCTcap", "OGTT 0 min", "OGTT 1 hr", "RPG", "RCG", "A1c")
+gt(roc_table, rownames_to_stub = T) %>% fmt_number(decimals = 3)
+```
+
+### ROC curves
+
+```{r}
+invisible(lapply(
+  list(gctpl_roc, gctcap_roc, ogtt0_roc, ogtt1_roc, rpg_roc, rcg_roc, a1c_roc),
+  function(roc) {
+    print(plot(roc, print.auc = T, print.thres = T, legacy.axes = T))
+  }
+))
+```
+
+### Full cutoff tables
+
+#### GCTpl
+
+```{r}
+gt(data.frame(coords(gctpl_roc, "all", ret = "all"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+#### GCTcap
+
+```{r}
+gt(data.frame(coords(gctcap_roc, "all", ret = "all"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+#### OGTT 0 min
+
+```{r}
+gt(data.frame(coords(ogtt0_roc, "all", ret = "all"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+#### OGTT 1 hour
+
+```{r}
+gt(data.frame(coords(ogtt1_roc, "all", ret = "all"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+#### RPG
+
+```{r}
+gt(data.frame(coords(rpg_roc, "all", ret = "all"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+#### RCG
+
+```{r}
+gt(data.frame(coords(rcg_roc, "all", ret = "all"))) %>%
+  fmt_number(n_sigfig = 3)
+```
+
+#### HbA1c
+
+```{r}
+gt(data.frame(coords(a1c_roc, "all", ret = "all"))) %>%
+  fmt_number(n_sigfig = 3)
+```

### 2025-04-19 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
write results and data to .csv

@@
 gt(data.frame(coords(gctpl_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(gctpl_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/GCTpl_dysglycemia.csv"
+)
 ```
@@
 gt(data.frame(coords(gctcap_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(gctcap_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/GCTcap_dysglycemia.csv"
+)
 ```
@@
 gt(data.frame(coords(ogtt0_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(ogtt0_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/OGTT_0_dysglycemia.csv"
+)
 ```
@@
 gt(data.frame(coords(ogtt1_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(ogtt1_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/OGTT_1_dysglycemia.csv"
+)
 ```
@@
 gt(data.frame(coords(rpg_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(rpg_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/RPG_dysglycemia.csv"
+)
 ```
@@
 gt(data.frame(coords(rcg_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(rcg_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/RCG_dysglycemia.csv"
+)
 ```
@@
 gt(data.frame(coords(a1c_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(a1c_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/A1c_dysglycemia.csv"
+)
 ```
@@
 gt(data.frame(coords(gctpl_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(gctpl_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/GCTpl_CFRD.csv"
+)
 ```
@@
 gt(data.frame(coords(gctcap_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(gctcap_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/GCTcap_CFRD.csv"
+)
 ```
@@
 gt(data.frame(coords(ogtt0_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(ogtt0_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/OGTT_0_CFRD.csv"
+)
 ```
@@
 gt(data.frame(coords(ogtt1_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(ogtt1_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/OGTT_1_CFRD.csv"
+)
 ```
@@
 gt(data.frame(coords(rpg_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(rpg_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/RPG_CFRD.csv"
+)
 ```
@@
 gt(data.frame(coords(rcg_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(rcg_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/RCG_CFRD.csv"
+)
 ```
@@
 gt(data.frame(coords(a1c_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(a1c_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/A1c_CFRD.csv"
+)
 ```
@@
 gt(data.frame(coords(gctpl_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(gctpl_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/GCTpl_CFRD_INDET.csv"
+)
 ```
@@
 gt(data.frame(coords(gctcap_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(gctcap_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/GCTcap_CFRD_INDET.csv"
+)
 ```
@@
 gt(data.frame(coords(ogtt0_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(ogtt0_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/OGTT_0_CFRD_INDET.csv"
+)
 ```
@@
 gt(data.frame(coords(ogtt1_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(ogtt1_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/OGTT_1_CFRD_INDET.csv"
+)
 ```
@@
 gt(data.frame(coords(rpg_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(rpg_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/RPG_CFRD_INDET.csv"
+)
 ```
@@
 gt(data.frame(coords(rcg_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(rcg_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/RCG_CFRD_INDET.csv"
+)
 ```
@@
 gt(data.frame(coords(a1c_roc, "all", ret = "all"))) %>%
-  fmt_number(n_sigfig = 3)
+  fmt_number(n_sigfig = 4)
+write.csv(coords(a1c_roc, "all", ret = "all"),
+  row.names = F, na = "",
+  file = "./Christine Chan/CFF GCT Study/Reports/Tables/A1c_CFRD_INDET.csv"
+)
 ```

### 2025-04-25 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
Race and eti removed age from models

@@
 # Model summary
-change_mod %>% gtsummary::tbl_regression(intercept = T)
+change_mod %>%
+  gtsummary::tbl_regression(intercept = T) %>%
+  as_gt() %>%
+  fmt_number(decimals = 3)
 # Table of predicted values
@@
   change_mod,
-  "Years + `Years:RaceBlack or African American` = 1.195465"
+  "Years + `Years:RaceBlack or African American` = 1.13856555"
 )
@@
   change_mod,
-  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.738885"
+  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.63633"
 )
@@
   data = encounter
+)
+# Model plot
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
+  xlab("Years from ETI Start") +
+  ylab("GLI_FEV1_pct_predicted") +
+  theme_classic()
+# Model summary
+change_mod %>%
+  gtsummary::tbl_regression(intercept = T) %>%
+  as_gt() %>%
+  fmt_number(decimals = 3)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+# Linear hypotheses
+white_vs_black_pre <- glht(
+  change_mod,
+  "Years + `Years:RaceBlack or African American` = 1.195465"
+)
+white_vs_black_pre <- tidy(summary(white_vs_black_pre))
+white_vs_black_pre$contrast <- "Black vs. White Pre-ETI Slope"
+white_vs_black_pre %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+white_vs_black_post <- glht(
+  change_mod,
+  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.738885"
+)
+white_vs_black_post <- tidy(summary(white_vs_black_post))
+white_vs_black_post$contrast <- "Black vs. White Post-ETI Slope"
+white_vs_black_post %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+
+white_vs_hispanic_pre <- glht(
+  change_mod,
+  "Years + `Years:RaceHispanic or Latino` = 1.19547"
+)
+white_vs_hispanic_pre <- tidy(summary(white_vs_hispanic_pre))
+white_vs_hispanic_pre$contrast <- "Hispanic vs. White Pre-ETI Slope"
+white_vs_hispanic_pre %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+white_vs_hispanic_post <- glht(
+  change_mod,
+  "Years + `Years:RaceHispanic or Latino` + slope_change_years + `RaceHispanic or Latino:slope_change_years` = 2.738885"
+)
+white_vs_hispanic_post <- tidy(summary(white_vs_hispanic_post))
+white_vs_hispanic_post$contrast <- "Hispanic vs. White Post-ETI Slope"
+white_vs_hispanic_post %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+
+white_vs_other_pre <- glht(
+  change_mod,
+  "Years + `Years:RaceOther` = 1.19547"
+)
+white_vs_other_pre <- tidy(summary(white_vs_other_pre))
+white_vs_other_pre$contrast <- "Other vs. White Pre-ETI Slope"
+white_vs_other_pre %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+white_vs_other_post <- glht(
+  change_mod,
+  "Years + `Years:RaceOther` + slope_change_years + `RaceOther:slope_change_years` = 2.738885"
+)
+white_vs_other_post <- tidy(summary(white_vs_other_post))
+white_vs_other_post$contrast <- "Other vs. White Post-ETI Slope"
+white_vs_other_post %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+```
+
+### Age < 18 at ETI initiation
+
+The following models were not adjusted for age group.
+
+#### GLI 2022 equations
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
+    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
+)
+# Model plot
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
+  xlab("Years from ETI Start") +
+  ylab("gli_fev1_ppred_rn") +
+  theme_classic()
+# Model summary
+change_mod %>%
+  gtsummary::tbl_regression(intercept = T) %>%
+  as_gt() %>%
+  fmt_number(decimals = 3)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+# Linear hypotheses
+white_vs_black_pre <- glht(
+  change_mod,
+  "Years + `Years:RaceBlack or African American` = 1.195465"
+)
+white_vs_black_pre <- tidy(summary(white_vs_black_pre))
+white_vs_black_pre$contrast <- "Black vs. White Pre-ETI Slope"
+white_vs_black_pre %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+white_vs_black_post <- glht(
+  change_mod,
+  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.738885"
+)
+white_vs_black_post <- tidy(summary(white_vs_black_post))
+white_vs_black_post$contrast <- "Black vs. White Post-ETI Slope"
+white_vs_black_post %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+
+white_vs_hispanic_pre <- glht(
+  change_mod,
+  "Years + `Years:RaceHispanic or Latino` = 1.19547"
+)
+white_vs_hispanic_pre <- tidy(summary(white_vs_hispanic_pre))
+white_vs_hispanic_pre$contrast <- "Hispanic vs. White Pre-ETI Slope"
+white_vs_hispanic_pre %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+white_vs_hispanic_post <- glht(
+  change_mod,
+  "Years + `Years:RaceHispanic or Latino` + slope_change_years + `RaceHispanic or Latino:slope_change_years` = 2.738885"
+)
+white_vs_hispanic_post <- tidy(summary(white_vs_hispanic_post))
+white_vs_hispanic_post$contrast <- "Hispanic vs. White Post-ETI Slope"
+white_vs_hispanic_post %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+
+white_vs_other_pre <- glht(
+  change_mod,
+  "Years + `Years:RaceOther` = 1.19547"
+)
+white_vs_other_pre <- tidy(summary(white_vs_other_pre))
+white_vs_other_pre$contrast <- "Other vs. White Pre-ETI Slope"
+white_vs_other_pre %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+white_vs_other_post <- glht(
+  change_mod,
+  "Years + `Years:RaceOther` + slope_change_years + `RaceOther:slope_change_years` = 2.738885"
+)
+white_vs_other_post <- tidy(summary(white_vs_other_post))
+white_vs_other_post$contrast <- "Other vs. White Post-ETI Slope"
+white_vs_other_post %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+```
+
+#### GLI 2012 equations
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  GLI_FEV1_pct_predicted ~ Years * Race + slope_change_years * Race +
+    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
 )
@@
 
-### Age < 18 at ETI initiation
+### Age >= 18 at ETI initiation
 
@@
     prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
-  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
+  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
 )
@@
     (1 | eDWID),
-  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
+  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
 )
@@
 
-### Age >= 18 at ETI initiation
-
-The following models were not adjusted for age group.
-
-#### GLI 2022 equations
-
-```{r}
-# Fit mixed model with change point
-change_mod <- lmer(
-  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
-    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
-  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
-)
-# Model plot
-pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
-pop_dat$pred <- pred$fit
-pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
-pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
-ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
-  geom_line() +
-  xlab("Years from ETI Start") +
-  ylab("gli_fev1_ppred_rn") +
-  theme_classic()
-# Model summary
-change_mod %>% gtsummary::tbl_regression(intercept = T)
-# Table of predicted values
-pop_dat %>%
-  select(-slope_change_years) %>%
-  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-# Linear hypotheses
-white_vs_black_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceBlack or African American` = 1.195465"
-)
-white_vs_black_pre <- tidy(summary(white_vs_black_pre))
-white_vs_black_pre$contrast <- "Black vs. White Pre-ETI Slope"
-white_vs_black_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_black_post <- glht(
-  change_mod,
-  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.738885"
-)
-white_vs_black_post <- tidy(summary(white_vs_black_post))
-white_vs_black_post$contrast <- "Black vs. White Post-ETI Slope"
-white_vs_black_post %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-
-white_vs_hispanic_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceHispanic or Latino` = 1.19547"
-)
-white_vs_hispanic_pre <- tidy(summary(white_vs_hispanic_pre))
-white_vs_hispanic_pre$contrast <- "Hispanic vs. White Pre-ETI Slope"
-white_vs_hispanic_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_hispanic_post <- glht(
-  change_mod,
-  "Years + `Years:RaceHispanic or Latino` + slope_change_years + `RaceHispanic or Latino:slope_change_years` = 2.738885"
-)
-white_vs_hispanic_post <- tidy(summary(white_vs_hispanic_post))
-white_vs_hispanic_post$contrast <- "Hispanic vs. White Post-ETI Slope"
-white_vs_hispanic_post %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-
-white_vs_other_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceOther` = 1.19547"
-)
-white_vs_other_pre <- tidy(summary(white_vs_other_pre))
-white_vs_other_pre$contrast <- "Other vs. White Pre-ETI Slope"
-white_vs_other_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_other_post <- glht(
-  change_mod,
-  "Years + `Years:RaceOther` + slope_change_years + `RaceOther:slope_change_years` = 2.738885"
-)
-white_vs_other_post <- tidy(summary(white_vs_other_post))
-white_vs_other_post$contrast <- "Other vs. White Post-ETI Slope"
-white_vs_other_post %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-```
-
-#### GLI 2012 equations
-
-```{r}
-# Fit mixed model with change point
-change_mod <- lmer(
-  GLI_FEV1_pct_predicted ~ Years * Race + slope_change_years * Race +
-    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
-    (1 | eDWID),
-  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
-)
-# Model plot
-pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
-pop_dat$pred <- pred$fit
-pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
-pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
-ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
-  geom_line() +
-  xlab("Years from ETI Start") +
-  ylab("GLI_FEV1_pct_predicted") +
-  theme_classic()
-# Model summary
-change_mod %>% gtsummary::tbl_regression(intercept = T)
-# Table of predicted values
-pop_dat %>%
-  select(-slope_change_years) %>%
-  rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-# Linear hypotheses
-white_vs_black_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceBlack or African American` = 1.195465"
-)
-white_vs_black_pre <- tidy(summary(white_vs_black_pre))
-white_vs_black_pre$contrast <- "Black vs. White Pre-ETI Slope"
-white_vs_black_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_black_post <- glht(
-  change_mod,
-  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.738885"
-)
-white_vs_black_post <- tidy(summary(white_vs_black_post))
-white_vs_black_post$contrast <- "Black vs. White Post-ETI Slope"
-white_vs_black_post %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-
-white_vs_hispanic_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceHispanic or Latino` = 1.19547"
-)
-white_vs_hispanic_pre <- tidy(summary(white_vs_hispanic_pre))
-white_vs_hispanic_pre$contrast <- "Hispanic vs. White Pre-ETI Slope"
-white_vs_hispanic_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_hispanic_post <- glht(
-  change_mod,
-  "Years + `Years:RaceHispanic or Latino` + slope_change_years + `RaceHispanic or Latino:slope_change_years` = 2.738885"
-)
-white_vs_hispanic_post <- tidy(summary(white_vs_hispanic_post))
-white_vs_hispanic_post$contrast <- "Hispanic vs. White Post-ETI Slope"
-white_vs_hispanic_post %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-
-white_vs_other_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceOther` = 1.19547"
-)
-white_vs_other_pre <- tidy(summary(white_vs_other_pre))
-white_vs_other_pre$contrast <- "Other vs. White Pre-ETI Slope"
-white_vs_other_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_other_post <- glht(
-  change_mod,
-  "Years + `Years:RaceOther` + slope_change_years + `RaceOther:slope_change_years` = 2.738885"
-)
-white_vs_other_post <- tidy(summary(white_vs_other_post))
-white_vs_other_post$contrast <- "Other vs. White Post-ETI Slope"
-white_vs_other_post %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-```
-
 ## FVC % predicted
@@
 change_mod <- lmer(
-  bmipercentile ~ Years * Race + slope_change_years * Race + age_group +
+  bmipercentile ~ Years * Race + slope_change_years * Race +
     Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
@@
 change_mod <- lmer(
-  bmivalue ~ Years * Race + slope_change_years * Race + age_group +
+  bmivalue ~ Years * Race + slope_change_years * Race +
     Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
@@
 change_mod <- lmer(
-  heightpercentile ~ Years * Race + slope_change_years * Race + age_group +
+  heightpercentile ~ Years * Race + slope_change_years * Race +
     Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
@@
 change_mod <- lmer(
-  height ~ Years * Race + slope_change_years * Race + age_group + Insurance +
+  height ~ Years * Race + slope_change_years * Race + Insurance +
     prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
@@
 change_mod <- lmer(
-  weightpercentile ~ Years * Race + slope_change_years * Race + age_group +
+  weightpercentile ~ Years * Race + slope_change_years * Race +
     Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
@@
 change_mod <- lmer(
-  weight ~ Years * Race + slope_change_years * Race + age_group + Insurance +
+  weight ~ Years * Race + slope_change_years * Race + Insurance +
     prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
@@
 # Model summary
-change_mod %>% gtsummary::tbl_regression(intercept = T)
+change_mod %>% 
+  gtsummary::tbl_regression(intercept = T)%>%
+  as_gt() %>%
+  fmt_number(decimals = 3)
 # Table of predicted values

### 2025-05-02 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
file paths

@@
 library(ggstats)
-knitr::opts_knit$set(root.dir = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry/")
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
+  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
+)
+github_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Documents/GitHub",
+  "Linux" = "/home/timvigers/Documents/GitHub"
+)
+knitr::opts_knit$set(root.dir = home_dir)
 ```
@@
 ggsave("./Dissemination/Figures/time_combined_forest.png",
-  plot = comb_plot, width = 12, height = 9, units = "in"
+  plot = comb_plot, width = 12, height = 9, units = "in", dpi = 900
 )

### 2025-05-02 Emily Muther/mental_health_screening_psc.qmd
file paths

--- 2024-01-26 Emily Muther/mental_health_screening_psc.qmd
@@
 ---
-title: "Mental Health Screening in CF - PSC Screening"
+title: "PSC Screening in Cystic Fibrosis"
 author: "Tim Vigers"
@@
 library(tidyverse)
-library(gtsummary)
-library(ggcorrplot)
+library(readxl)
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF"
+  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Emily Muther",
+  "Linux" = "/home/timvigers/OneDrive"
+)
+github_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/timvigers/Documents/GitHub",
+  "Linux" = "/home/timvigers/Documents/GitHub"
 )
@@
 ```{r data cleaning}
-#| warning: false
 # Import
-df <- read.csv("./Emily Muther/General - Muther Research Group/Mental Health Screening/Data_Raw/PSC_Full_Data_data_11-11-22.csv",
-  check.names = F, na.strings = ""
-)
-# Remove duplicate rows
-df <- df %>% distinct()
-# Sort by ID and contact date
-df$CONTACT_DATE <- lubridate::mdy(df$CONTACT_DATE)
-df <- df %>% arrange(PAT_MRN_ID, CONTACT_DATE)
-# Exclude those with no PSC data
-df <- df %>% filter(!is.na(Score))
+df <- read_excel("./Muther Research Group - General/PSC manuscript 2025/PSC_9-20-19 thru 4-18-25_pulled 4-21-25.xlsx")
+# Fill down MRN, select columns, remove duplicates
+df <- df %>%
+  fill(`Pat Mrn Id`, .direction = "down") %>%
+  select(
+    `Pat Mrn Id`, `Payor Name`, `AGE AT CONTACT`, `Internalizing subscale`,
+    `Attention subscale`, `Externalizing subscale`, Score
+  ) %>%
+  distinct(`Pat Mrn Id`, `AGE AT CONTACT`, .keep_all = T)
 ```
 
-# Data cleaning
+# Table
 
-1. Duplicate rows were removed (n=21).
-2. All rows with missing data for `Score` were excluded (n=2319).
-3. In Table 1, demographics and first contact date are reported.
-
-# Demographics
-
-```{r table 1}
-cat_demo_vars <- c(
-  "DEPARTMENT_NAME", "ENC_T", "PAYOR_NAME"
-)
-cont_demo_vars <- c("AGE AT CONTACT")
-df %>%
-  group_by(PAT_MRN_ID) %>%
-  summarise(
-    across(all_of(cont_demo_vars), ~ first(na.omit(.x))),
-    across(all_of(cat_demo_vars), ~ first(na.omit(.x)))
-  ) %>%
-  select(-PAT_MRN_ID) %>%
-  tbl_summary(missing_text = "Missing", label = list("Age at First Contact" = "AGE AT CONTACT"))
+```{r}
+#| label: tbl-planets
+#| tbl-cap: Planets
 ```
 
-# Prevalence of behavioral health difficulties
-
-## All PSC screeners
-
-```{r table 2}
-df %>%
-  select(
-    `Internalizing subscale`:`Externalizing subscale`, Score,
-    `PSC Attention`:`PSC Totals`
-  ) %>%
-  tbl_summary(
-    missing_text = "Missing",
-    type = list(`Internalizing subscale` ~ "continuous")
-  )
-```
-
-## By person
-
-The maximum score on each subscale was calculated for all participants.
+# Figure
 
 ```{r}
-#| warning: false
-df %>%
-  mutate(across(`PSC Attention`:`PSC Totals`, ~ factor(.x, ordered = T))) %>%
-  group_by(PAT_MRN_ID) %>%
-  summarise(across(c(
-    `Internalizing subscale`:`Externalizing subscale`, Score,
-    `PSC Attention`:`PSC Totals`
-  ), ~ max(.x, na.rm = T))) %>%
-  mutate(across(
-    c(`Internalizing subscale`:`Externalizing subscale`, Score),
-    ~ case_when(.x == -Inf ~ NA, .default = .x)
-  )) %>%
-  select(-PAT_MRN_ID) %>%
-  tbl_summary(
-    missing_text = "Missing",
-    type = list(`Internalizing subscale` ~ "continuous")
-  )
+#| label: fig-planets
+#| fig-cap: Planets
 ```
-
-# Correlations
-
-```{r}
-X = df %>% 
-```

### 2025-05-02 Emily Muther/mental_health_screening_psc.qmd
PSC EDA for Emily

@@
 library(readxl)
+library(arsenal)
 home_dir <- switch(Sys.info()["sysname"],
@@
 ```{r data cleaning}
+#| message: false
 # Import
@@
   ) %>%
+  # pivot_wider(names_from = `Payor Name`,values_from = `Payor Name`)
   distinct(`Pat Mrn Id`, `AGE AT CONTACT`, .keep_all = T)
+# Convert to numeric
+df$`AGE AT CONTACT` <- as.numeric(df$`AGE AT CONTACT`)
+df$`Internalizing subscale` <- as.numeric(df$`Internalizing subscale`)
+df$`Attention subscale` <- as.numeric(df$`Attention subscale`)
+df$`Externalizing subscale` <- as.numeric(df$`Externalizing subscale`)
+df$Score <- as.numeric(df$Score)
+# Make age group
+df$`Age Group` <- cut(df$`AGE AT CONTACT`,
+  breaks = c(-Inf, 6, 10, Inf),
+  labels = c("Age < 6", "Age >= 6 and < 10", "Age >= 10"), right = F
+)
+# Set table options
+mycontrols <-
+  tableby.control(
+    numeric.stats = c("N", "Nmiss", "medianq1q3", "range"),
+    cat.stats = c("N", "Nmiss", "countpct")
+  )
 ```
 
-# Table
+# Insurance providers
+
+MEDICAID = Public
+MEDICAID NEW MEXICO = Public
+MEDICAID WYOMING = Public
+UNITED HEALTHCARE = Private
+CIGNA = Private
+ANTHEM BCBS = Private
+ROCKY MOUNTAIN HEALTH PLAN = ?
+AETNA = Private
+MEDICAID OTHER STATES = Public
+TRICARE = Tricare
+CICP = ?
+CHP+COLORADO ACCESS = ?
+CU GME HEALTH = ?
+MIDLANDS CHOICE = ?
+BRIGHT HEALTH = ?
+FIRST HEALTH GROUP = ?
+ROCKY MOUNTAIN MEDICAID = Public
+KAISER = Kaiser
+MEDICAID NEBRASKA = Public
+MEDICAID OKLAHOMA = Public
+EBMS = ?
+COFINITY = ?
+SIMPLIFIED BENEFITS ADMINISTRATORS = ?
+FRIDAY HEALTH PLANS = ?
+SELECTHEALTH = INC. = Private?
+
+# Score summary
+
+## At first visit
+
+```{r results='asis'}
+first_visit <- df %>%
+  group_by(`Pat Mrn Id`) %>%
+  slice_min(`AGE AT CONTACT`)
+t1 <- tableby(
+  `Age Group` ~ `Internalizing subscale` + `Attention subscale` +
+    `Externalizing subscale` + Score,
+  data = first_visit
+)
+summary(t1, pfootnote = T)
+```
+
+## All visits
+
+The table below treats multiple visits from the same participant as independent.
+
+```{r results='asis'}
+t1 <- tableby(
+  `Age Group` ~ `Internalizing subscale` + `Attention subscale` +
+    `Externalizing subscale` + Score,
+  data = df
+)
+summary(t1, pfootnote = T)
+```
+
+# Figures
+
+## Histograms
 
 ```{r}
-#| label: tbl-planets
-#| tbl-cap: Planets
+hist(df$`Internalizing subscale`)
+hist(df$`Attention subscale`)
+hist(df$`Externalizing subscale`)
+hist(df$Score)
 ```
 
-# Figure
+## Scores by age
 
 ```{r}
-#| label: fig-planets
-#| fig-cap: Planets
+#| warning: false
+ggplot(df, aes(x = `AGE AT CONTACT`, y = `Internalizing subscale`)) +
+  geom_point() +
+  geom_line(aes(group = `Pat Mrn Id`))
+
+ggplot(df, aes(x = `AGE AT CONTACT`, y = `Attention subscale`)) +
+  geom_point() +
+  geom_line(aes(group = `Pat Mrn Id`))
+
+ggplot(df, aes(x = `AGE AT CONTACT`, y = `Externalizing subscale`)) +
+  geom_point() +
+  geom_line(aes(group = `Pat Mrn Id`))
+
+ggplot(df, aes(x = `AGE AT CONTACT`, y = Score)) +
+  geom_point() +
+  geom_line(aes(group = `Pat Mrn Id`))
 ```

### 2025-05-05 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
improved figure 1 and graphical abstract

@@
 library(ggstats)
+library(patchwork)
 home_dir <- switch(Sys.info()["sysname"],
@@
   rename(Year = ReviewYear)
-p$name <- "Prevalence"
+p$name <- "Prevalence of CFRD"
 i <- incidence %>%
@@
   )
-i$name <- "Incidence"
+i$name <- "Incidence of CFRD"
 s <- screening %>%
@@
   rename(Year = ReviewYear, `Age Group` = age_group)
-s$name <- " Screening Rates for CFRD"
+s$name <- "Screening Rates for CFRD"
 s$value <- s$value * 100
@@
 t <- t[t$Year >= 2004, ]
-fig1 <- ggplot(t, aes(x = Year, y = value, color = `Age Group`)) +
+# Figure 1
+fig1a <- ggplot(
+  t[t$name == "Screening Rates for CFRD", ],
+  aes(x = Year, y = value, linetype = `Age Group`)
+) +
   geom_line() +
-  facet_wrap(~name, scales = "free", nrow = 3)
+  theme_bw() +
+  ylab("Percentage") +
+  ggtitle("Screening Rates for CFRD")
+fig1b <- ggplot(
+  t[t$name == "Incidence of CFRD", ],
+  aes(x = Year, y = value, linetype = `Age Group`)
+) +
+  geom_line() +
+  theme_bw() +
+  ylab("Per 1,000 Person-Years") +
+  ggtitle("Incidence of CFRD")
+fig1c <- ggplot(
+  t[t$name == "Prevalence of CFRD", ],
+  aes(x = Year, y = value, linetype = `Age Group`)
+) +
+  geom_line() +
+  theme_bw() +
+  ylab("Per 1,000 Person-Years") +
+  ggtitle("Prevalence of CFRD")
+fig1 <- fig1a + fig1b + fig1c + plot_layout(guides = "collect")
 fig1
 ggsave("./Dissemination/Figures/fig1.png",
-  plot = fig1, width = 9, height = 12, units = "in"
+  plot = fig1, width = 12, height = 3, units = "in", dpi = 900
+)
+# Graphical abstract is the same but with color
+fig1a <- ggplot(
+  t[t$name == "Screening Rates for CFRD", ],
+  aes(x = Year, y = value, color = `Age Group`)
+) +
+  geom_line() +
+  theme_bw() +
+  ylab("Percentage") +
+  ggtitle("Screening Rates for CFRD")
+fig1b <- ggplot(
+  t[t$name == "Incidence of CFRD", ],
+  aes(x = Year, y = value, color = `Age Group`)
+) +
+  geom_line() +
+  theme_bw() +
+  ylab("Per 1,000 Person-Years") +
+  ggtitle("Incidence of CFRD")
+fig1c <- ggplot(
+  t[t$name == "Prevalence of CFRD", ],
+  aes(x = Year, y = value, color = `Age Group`)
+) +
+  geom_line() +
+  theme_bw() +
+  ylab("Per 1,000 Person-Years") +
+  ggtitle("Prevalence of CFRD")
+fig1 <- fig1a + fig1b + fig1c + plot_layout(guides = "collect")
+fig1
+ggsave("./Dissemination/Figures/graphical_abstract.png",
+  plot = fig1, width = 12, height = 3, units = "in", dpi = 900
 )

### 2025-05-15 Christine Chan/CF Registry/cf_registry_CFRD_prediction.qmd
file paths etc.

@@
 library(JM)
+library()
 library(dplyr)

### 2025-05-15 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
file paths etc.

@@
   ggtitle("Prevalence of CFRD")
-fig1 <- fig1a + fig1b + fig1c + plot_layout(guides = "collect")
+fig1 <- fig1a + fig1b + fig1c + plot_layout(guides = "collect") +
+  plot_annotation(title = "Figure 1: Trends in Screening, Incidence, and Prevalence of CFRD from 2003-2018, by age group ")
 fig1
@@
   ggtitle("Prevalence of CFRD")
-fig1 <- fig1a + fig1b + fig1c + plot_layout(guides = "collect")
+fig1 <- fig1a + fig1b + fig1c + plot_layout(guides = "collect")+
+  plot_annotation(title = "US Cystic Fibrosis Foundation Patient Registry Trends in Screening, Incidence, and Prevalence of Cystic Fibrosis Related Diabetes from 2003-2018")
 fig1
 ggsave("./Dissemination/Figures/graphical_abstract.png",
-  plot = fig1, width = 12, height = 3, units = "in", dpi = 900
+  plot = fig1, width = 13, height = 3, units = "in", dpi = 900
 )

### 2025-05-15 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
file paths etc.

@@
 library(reporttools)
+library(rpart)
+library(rpart.plot)
+library(rms)
 home_dir <- switch(Sys.info()["sysname"],
@@
   labels = c(
-    "Have GCT 1 Hour Plasma or OGTT 2 Hour",
+    "Have GCT 1 Hour Plasma and OGTT 2 Hour",
     "Missing GCT 1 Hour Plasma or OGTT 2 Hour"
@@
 df <- df %>%
-  filter(`Got OGTT and GCT?` == "Have GCT 1 Hour Plasma or OGTT 2 Hour") %>%
+  filter(`Got OGTT and GCT?` == "Have GCT 1 Hour Plasma and OGTT 2 Hour") %>%
   select(-`Got OGTT and GCT?`)
@@
 ```
+
+# Nomogram
+
+## CFRD
+
+### `GCT 0 min capillary` and HbA1c
+
+```{r}
+f <- lrm(CFRD ~ `GCT 0 min capillary` + HbA1c,
+  data = df
+)
+ddist <- datadist(df)
+options(datadist = "ddist")
+nom <- nomogram(f, fun = plogis, funlabel = "Risk of CFRD",lp = F)
+plot(nom)
+```
+
+Here’s how to use a nomogram.
+Pick a point on the GCT 0 min capillary and HbA1c scales and get the associated points for each.
+Add the points together to get the associated probability based on the Total Points line.
+
+For example:
+
+- GCT 0 min capillary = 130 (40 points)
+- HbA1c = 6.6 (10 points)
+
+Add up the points:
+
+40 + 10 = 50 Total Points
+
+50 Total Points lines up at about 0.08, so this person has approximately an 8% chance of CFRD. The process is the same for the nomograms below.
+
+### `GCT 0 min capillary`, HbA1c, and `GCT 60 min plasma`
+
+```{r}
+f <- lrm(CFRD ~ `GCT 0 min capillary` + HbA1c + `GCT 60 min plasma`,
+  data = df
+)
+ddist <- datadist(df)
+options(datadist = "ddist")
+nom <- nomogram(f, fun = plogis, funlabel = "Risk of CFRD",lp = F)
+plot(nom)
+```
+
+### `GCT 0 min capillary`, and `GCT 60 min plasma`
+
+```{r}
+f <- lrm(CFRD ~ `GCT 0 min capillary` + `GCT 60 min plasma`,
+  data = df
+)
+ddist <- datadist(df)
+options(datadist = "ddist")
+nom <- nomogram(f, fun = plogis, funlabel = "Risk of CFRD",lp = F)
+plot(nom)
+```
+
+# Decision trees
+
+## Predicting CFRD
+
+```{r}
+tree <- rpart(CFRD ~ `GCT 0 min capillary` + HbA1c + `GCT 60 min plasma`,
+  data = df
+)
+rpart.plot(tree, type = 0)
+```
+
+We just don't have enough data to do this properly for CFRD.
+
+## Predicting dysglycemia
+
+```{r}
+tree <- rpart(Dysglycemia ~ `GCT 0 min capillary` + `GCT 60 min plasma`, data = df)
+rpart.plot(tree, type = 0)
+```

### 2025-05-15 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
file paths etc.

@@
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
+  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
   "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Variability"
@@
 # For "LESS THAN 10", replace with half the limit of detection
-chco[chco == "LESS THAN 10"] <- "5"
+chco[chco == "LESS THAN 10"] <- "9"
 chco[, c("Cl", "Cl_Left", "Cl_Right")] <-
@@
 - Merged in the diagnoses from Angela.
-- Values below the limit of detection were replaced with half the limit (e.g. "LESS THAN 10" was replaced with 5). Approximately 5% of the values were below the L.O.D.
+- Values below the limit of detection were replaced with half the limit (e.g. "LESS THAN 10" was replaced with 9). Approximately 5% of the values were below the L.O.D.
 

### 2025-05-15 Emily Muther/mental_health_screening_psc.qmd
file paths etc.

--- 2025-05-02 Emily Muther/mental_health_screening_psc.qmd
@@
 library(readxl)
+library(arsenal)
+library(GLMMadaptive)
 home_dir <- switch(Sys.info()["sysname"],
@@
 ```{r data cleaning}
+#| message: false
 # Import
 df <- read_excel("./Muther Research Group - General/PSC manuscript 2025/PSC_9-20-19 thru 4-18-25_pulled 4-21-25.xlsx")
-# Fill down MRN, select columns, remove duplicates
+# Classify insurance per Emily
+df$Insurance <- factor(df$`Payor Name`,
+  levels = c(
+    "MEDICAID", "MEDICAID NEW MEXICO", "MEDICAID WYOMING", "UNITED HEALTHCARE",
+    "CIGNA", "ANTHEM BCBS", "ROCKY MOUNTAIN HEALTH PLAN", "AETNA",
+    "MEDICAID OTHER STATES", "TRICARE", "CICP", "CHP+COLORADO ACCESS",
+    "CU GME HEALTH", "MIDLANDS CHOICE", "BRIGHT HEALTH", "FIRST HEALTH GROUP",
+    "ROCKY MOUNTAIN MEDICAID", "KAISER", "MEDICAID NEBRASKA",
+    "MEDICAID OKLAHOMA", "EBMS", "COFINITY", "SIMPLIFIED BENEFITS ADMINISTRATORS",
+    "FRIDAY HEALTH PLANS", "SELECTHEALTH, INC."
+  ),
+  labels = c(
+    "Public", "Public", "Public", "Private", "Private", "Private", "Public",
+    "Private", "Public", "Private", "Public", "Public", "Private", "Private",
+    "Private", "Private", "Public", "Private", "Public", "Public", "Private",
+    "Private", "Private", "Private", "Private"
+  )
+)
+# Fill down MRN and select columns
 df <- df %>%
@@
   select(
-    `Pat Mrn Id`, `Payor Name`, `AGE AT CONTACT`, `Internalizing subscale`,
+    `Pat Mrn Id`, Insurance, `AGE AT CONTACT`, `Internalizing subscale`,
     `Attention subscale`, `Externalizing subscale`, Score
   ) %>%
-  distinct(`Pat Mrn Id`, `AGE AT CONTACT`, .keep_all = T)
+  distinct() %>%
+  pivot_wider(names_from = Insurance, values_from = Insurance) %>%
+  unite("Insurance", Public, Private, `NA`, remove = T, na.rm = T)
+df$Insurance <- factor(df$Insurance,
+  levels = c("", "Private", "Public", "Public_Private"),
+  labels = c("None", "Private", "Public", "Both")
+)
+# Convert to numeric
+df$`AGE AT CONTACT` <- as.numeric(df$`AGE AT CONTACT`)
+df$`Internalizing subscale` <- as.numeric(df$`Internalizing subscale`)
+df$`Attention subscale` <- as.numeric(df$`Attention subscale`)
+df$`Externalizing subscale` <- as.numeric(df$`Externalizing subscale`)
+df$Score <- as.numeric(df$Score)
+# Make age group
+df$`Age Group` <- cut(df$`AGE AT CONTACT`,
+  breaks = c(-Inf, 6, 10, Inf),
+  labels = c("Age < 6", "Age >= 6 and < 10", "Age >= 10"), right = F
+)
+# Make variables for 0 scores on each scale
+df$`Internalizing subscale symptoms` <- factor(df$`Internalizing subscale` == 0,
+  levels = c(F, T),
+  labels = c(
+    "No internalizing symptoms (score = 0)",
+    "Any internalizing symptoms (score > 0)"
+  )
+)
+df$`Attention subscale symptoms` <- factor(df$`Attention subscale` == 0,
+  levels = c(F, T),
+  labels = c(
+    "No attention symptoms (score = 0)",
+    "Any attention symptoms (score > 0)"
+  )
+)
+df$`Externalizing subscale symptoms` <- factor(df$`Externalizing subscale`==0,
+  levels = c(F, T),
+  labels = c(
+    "No externalizing symptoms (score = 0)",
+    "Any externalizing symptoms (score > 0)"
+  )
+)
+df$`Any symptoms` <- factor(df$Score == 0,
+  levels = c(F, T),
+  labels = c(
+    "No symptoms (total score = 0)",
+    "Any symptoms (total score > 0)"
+  )
+)
+# Set table options
+mycontrols <-
+  tableby.control(
+    numeric.stats = c("N", "Nmiss", "medianq1q3", "range"),
+    cat.stats = c("N", "Nmiss", "countpct")
+  )
+# Basic formula
+f <- ~ Insurance + `Internalizing subscale` + `Internalizing subscale symptoms` +
+  `Attention subscale` + `Attention subscale symptoms` +
+  `Externalizing subscale` + `Externalizing subscale symptoms` +
+  Score + `Any symptoms`
 ```
 
-# Table
+# Score summary at first visit
+
+## By age group
+
+```{r results='asis'}
+first_visit <- df %>%
+  group_by(`Pat Mrn Id`) %>%
+  slice_min(`AGE AT CONTACT`)
+t1 <- tableby(update(f, `Age Group` ~ .), data = first_visit)
+summary(t1, pfootnote = T)
+```
+
+## By insurance type
+
+```{r results='asis'}
+first_visit <- df %>%
+  group_by(`Pat Mrn Id`) %>%
+  slice_min(`AGE AT CONTACT`)
+t1 <- tableby(update(f, Insurance ~ .), data = first_visit)
+summary(t1, pfootnote = T)
+```
+
+# Figures
+
+## Histograms
 
 ```{r}
-#| label: tbl-planets
-#| tbl-cap: Planets
+hist(df$`Internalizing subscale`)
+hist(df$`Attention subscale`)
+hist(df$`Externalizing subscale`)
+hist(df$Score)
 ```
 
-# Figure
+## Scores by age
 
 ```{r}
-#| label: fig-planets
-#| fig-cap: Planets
+#| warning: false
+ggplot(df, aes(x = `AGE AT CONTACT`, y = `Internalizing subscale`)) +
+  geom_point() +
+  geom_line(aes(group = `Pat Mrn Id`))
+
+ggplot(df, aes(x = `AGE AT CONTACT`, y = `Attention subscale`)) +
+  geom_point() +
+  geom_line(aes(group = `Pat Mrn Id`))
+
+ggplot(df, aes(x = `AGE AT CONTACT`, y = `Externalizing subscale`)) +
+  geom_point() +
+  geom_line(aes(group = `Pat Mrn Id`))
+
+ggplot(df, aes(x = `AGE AT CONTACT`, y = Score)) +
+  geom_point() +
+  geom_line(aes(group = `Pat Mrn Id`))
 ```
+
+# Mixed models
+
+## Total score
+
+```{r}
+# First determine whether we need Poisson or negative binomial
+pm1 <- mixed_model(Score ~ `AGE AT CONTACT`,
+  random = ~ 1 | `Pat Mrn Id`, data = df,
+  family = zi.poisson(), zi_fixed = ~`AGE AT CONTACT`
+)
+nbm1 <- mixed_model(Score ~ `AGE AT CONTACT`,
+  random = ~ 1 | `Pat Mrn Id`, data = df,
+  family = zi.negative.binomial(), zi_fixed = ~`AGE AT CONTACT`
+)
+# anova(pm1, nbm1, test = FALSE) - NB better by AIC and BIC
+summary(nbm1)
+```
+
+# Questions for Emily
+
+- If payor is blank, does that indicate no insurance? Or is it missing?
+- Should we ask Elin for gender and lung function data? Not sure how much work that would be for her.

### 2025-05-15 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
file paths etc.

@@
 library(broom)
-knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF")
+knitr::opts_knit$set(root.dir = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF")
 ```
@@
     type = list(`Years With Data` ~ "continuous"),
+    statistic = list(c(gli_fev1_ppred_rn,gli_fvc_ppred_rn)~"{mean} [{sd}]"),
     missing_text = "Missing"

### 2025-05-15 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R

@@ -3,7 +3,7 @@ library(lubridate)
 library(parallel)
 # Paths
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
   "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
   "Linux" = "/home/tim/CF"
 )
@@ -172,7 +172,7 @@ encounter <- encounter[-late, ]
 n_enc_5 <- nrow(encounter)
 n_people_5 <- length(unique(encounter$eDWID))
 # Lung function cutoffs
-lower <- 20
+lower <- 10
 upper <- 150
 # Remove impossible lung function values
 low_fev1 <- which(encounter$gli_fev1_ppred_rn < lower)

### 2025-05-26 Quarto Template.qmd
template file paths

@@
   "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/tim/OneDrive/Vigers/CF"
 )
 github_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Documents/GitHub",
-  "Linux" = "/home/timvigers/Documents/GitHub"
+  "Darwin" = "/Users/tim/Documents/GitHub",
+  "Windows" = "C:/Users/Tim/Documents/GitHub",
+  "Linux" = "/home/tim/Documents/GitHub"
 )

### 2025-05-27 Christine Chan/EnVision CF/cfrd_roc.qmd
basic EnVision ROC curves

---
title: "Using CGM to Predict CFRD in the Envision Cohort"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(arsenal)
library(pROC)
library(gt)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/EnVision CF",
  "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/EnVision CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
# Import cleaned dataset
df <- read.csv("./Data_Clean/envision_analysis_dataset.csv", na.strings = "")
# Limit to those with CGM data (based on mean glucose)
df <- df %>% filter(!is.na(mean_sensor))
```

- **All tables and analyses below treat CGM wears as independent**, but we will need to think about how to handle this in future analyses.

# Participant demographics

- In the cleaned EnVision dataset we have a total of `r nrow(df)` CGMs from `r length(unique(df$study_id))` unique individuals. Of these CGM wears, `r sum(is.na(df$Diagnosis))` are missing a diagnosis due to missing OGTT timepoints.
- BMI percentile was calculated with age = 20 for adults.

```{r results='asis'}
t1 <- tableby(
  CFRD ~ redcap_data_access_group + sex + origin_race + ethnicity +
    pancreatic_status + CFTR + fev1 + fvc + bmi_perc + a1c_result,
  data = df
)
summary(t1, pfootnote = T)
```

# ROC Analyses

## Mean sensor glucose

```{r}
r <- roc(CFRD ~ mean_sensor, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "best", ret = "all")) %>% fmt_number(decimals = 3)
```

## Percent time spent >140 mg/dL

```{r}
r <- roc(CFRD ~ percent_time_over_140, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "best", ret = "all")) %>% fmt_number(decimals = 3)
```

## Percent time spent >180 mg/dL

```{r}
r <- roc(CFRD ~ percent_time_over_180, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "best", ret = "all")) %>% fmt_number(decimals = 3)
```

## GMI

```{r}
r <- roc(CFRD ~ gmi, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "best", ret = "all")) %>% fmt_number(decimals = 3)
```

## SD

```{r}
r <- roc(CFRD ~ standard_deviation, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "best", ret = "all")) %>% fmt_number(decimals = 3)
```

## CV

```{r}
r <- roc(CFRD ~ cv, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "best", ret = "all")) %>% fmt_number(decimals = 3)
```

# Questions

- Are we including fasting glucose >= 126 in the definition of CFRD for this analysis?

### 2025-05-27 Christine Chan/EnVision CF/clean_envision_data.R

@@ -371,6 +371,10 @@ final_df$Diagnosis[final_df$Glucose_0 >= 126 | final_df$Glucose_120 >= 200] <-
 final_df$Diagnosis <- factor(final_df$Diagnosis,
   levels = c("NGT", "IFG", "INDET", "IGT", "CFRD")
 )
+# Create CFRD+ vs. CFRD- variable
+final_df$CFRD <- factor(final_df$Diagnosis,
+  labels = c("CFRD-", "CFRD-", "CFRD-", "CFRD-", "CFRD+")
+)
 # iAUCs
 final_df$iAUC30gluc <- apply(final_df, 1, function(r) {
   y <- as.numeric(r[glucose[1:3]]) - as.numeric(r[glucose[1]])
@@ -524,7 +528,7 @@ final_df <- final_df %>%
   arrange(redcap_data_access_group, study_id, Date) %>%
   select(
     study_id, Date, ogtt_num, redcap_event_name, redcap_data_access_group:fvc,
-    bmi, bmi_perc, CFTR, Diagnosis, a1c_result, contains("Glucose_"),
+    bmi, bmi_perc, CFTR, Diagnosis, CFRD, a1c_result, contains("Glucose_"),
     matches("iAUC\\d{2,3}gluc"), Hypo60, Hypo70, contains("Insulin_"),
     matches("iAUC\\d{2,3}ins"), homa_ir, contains("C.Peptide_"),
     contains("Glucagon_"), contains("GLP.1.Active_"), contains("GIP_"),

### 2025-05-27 Emily Muther/body_image_trikafta.qmd
more file path changes

@@
 library(broom)
-library(report)
 library(ggstats)
@@
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
   "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
@@
 # Read in data
-df <- read.spss("./Emily Muther/General/BodyImage_Trikafta/Data/TrikaftaBodyImage3.6.23.sav",
+df <- read.spss("/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Emily Muther/Muther Research Group - General/BodyImage_Trikafta/Data/TrikaftaBodyImage3.6.23.sav",
   to.data.frame = T

### 2025-05-27 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
more file path changes

@@
 # Linear hypotheses
+
 white_vs_black_pre <- glht(

### 2025-05-27 Christine Chan/EnVision CF/cfrd_roc.qmd
Envision CFRD all thresholds

@@
 plot(r, print.auc = T, print.thres = T, legacy.axes = T)
-gt(coords(r, "best", ret = "all")) %>% fmt_number(decimals = 3)
+gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
 ```

### 2025-05-29 Christine Chan/EnVision CF/christine_CGM_OGTT_category.qmd
EnVision CGM and OGTT categories

@@
 library(gtsummary)
-library(dplyr)
+library(tidyverse)
 library(pracma)
@@
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
   "Linux" = "/home/timvigers/OneDrive/Vigers/CF"

### 2025-05-29 Christine Chan/EnVision CF/envision_AGPs.R

@@ -1,7 +1,7 @@
 library(tidyverse)
 library(lubridate)
 library(hms)
-files <- list.files("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF/Data_Raw/FilesReport_CGM_2025-01-30_1306/FilesReport_CGM_2025-01-30_1306/documents", full.names = T)
+files <- list.files("/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF/Data_Raw/FilesReport_CGM_2025-01-30_1306/FilesReport_CGM_2025-01-30_1306/documents", full.names = T)
 lapply(files, function(f) {
   # ID
   id <- basename(f)
@@ -29,7 +29,7 @@ lapply(files, function(f) {
     theme_bw() +
     scale_color_discrete("Date")
   # Save
-  ggsave(paste0("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF/Reports/AGPs/", id, ".png"),
+  ggsave(paste0("/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF/Reports/AGPs/", id, ".png"),
     plot = p, width = 9, height = 6, units = "in", device = "png"
   )
 })

### 2025-05-29 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
more plots for Larry

@@
 library(rms)
+library(patchwork)
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/tim/OneDrive/Vigers/CF"
 )
 github_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Documents/GitHub",
-  "Linux" = "/home/timvigers/Documents/GitHub"
+  "Darwin" = "/Users/tim/Documents/GitHub",
+  "Windows" = "C:/Users/Tim/Documents/GitHub",
+  "Linux" = "/home/tim/Documents/GitHub"
 )
@@
 #| warning: false
+# The code below is a total mess because Larry keeps sending requests -
+# clean up later
 plot_df <- df %>%
@@
 ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/boxplot.png",
+  box_p,
+  device = "png"
+)
+# Dotplot
+dotplot <- ggplot(df, aes(x = Diagnosis, y = `GCT 60 min plasma`)) +
+  geom_point() +
+  theme_bw()
+dotplot
+dotplot +
+  theme(text = element_blank())
+# Additional boxplots/dotplot per Christine and Larry
+plot_df <- df %>%
+  pivot_longer(c(`OGTT 0 min`, `OGTT 60 min`, `OGTT 120 min`, `GCT 60 min plasma`),
+    names_to = "OGTT Timepoint", values_to = "Glucose"
+  ) %>%
+  filter(`OGTT Timepoint` != "OGTT 60 min")
+plot_df$`OGTT Timepoint` <- factor(plot_df$`OGTT Timepoint`,
+  levels = c("OGTT 0 min", "GCT 60 min plasma", "OGTT 120 min")
+)
+box_p <- ggplot(
+  plot_df,
+  aes(x = `OGTT Timepoint`, y = Glucose, fill = `OGTT Timepoint`)
+) +
+  geom_boxplot() +
+  scale_y_continuous(breaks = c(100, 140, 200, 300)) +
+  facet_wrap(~Diabetes) +
+  theme_bw(base_size = 20) +
+  theme(
+    legend.position = "none",
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.x = element_blank()
+  )
+box_p
+ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/boxplot.png",
+  box_p,
+  device = "png"
+)
+# GCTpl, RCG, and HbA1c
+plot_df <- df %>%
+  pivot_longer(c(`GCT 60 min plasma`, `GCT 0 min capillary`, HbA1c),
+    names_to = "Measure", values_to = "Value"
+  )
+box_p1 <- ggplot(
+  plot_df[plot_df$Measure == "GCT 60 min plasma", ],
+  aes(x = Diabetes, y = Value, fill = Diabetes)
+) +
+  geom_boxplot() +
+  geom_hline(yintercept = 146.5, linetype = "dashed") +
+  ylim(c(50, 300)) +
+  ylab("mg/dL") +
+  ggtitle("GCTpl") +
+  theme_bw(base_size = 15) +
+  theme(
+    legend.position = "none",
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.x = element_blank()
+  )
+box_p2 <- ggplot(
+  plot_df[plot_df$Measure == "GCT 0 min capillary", ],
+  aes(x = Diabetes, y = Value, fill = Diabetes)
+) +
+  geom_boxplot() +
+  geom_hline(yintercept = 98.5, linetype = "dashed") +
+  ylim(c(50, 300)) +
+  ggtitle("RCG") +
+  theme_bw(base_size = 15) +
+  theme(
+    legend.position = "none",
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank()
+  )
+box_p3 <- ggplot(
+  plot_df[plot_df$Measure == "HbA1c", ],
+  aes(x = Diabetes, y = Value, fill = Diabetes)
+) +
+  geom_boxplot() +
+  geom_hline(yintercept = 5.45, linetype = "dashed") +
+  ylab("%") +
+  ggtitle("HbA1c") +
+  theme_bw(base_size = 15) +
+  theme(
+    legend.position = "none",
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank()
+  )
+box_p <- box_p1 + box_p2 + box_p3
+ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/boxplot2.png",
+  box_p,
+  device = "png"
+)
+# Dotplot
+plot_df <- df %>%
+  pivot_longer(c(`GCT 60 min plasma`, `GCT 0 min capillary`, HbA1c),
+    names_to = "Measure", values_to = "Value"
+  )
+box_p1 <- ggplot(
+  plot_df[plot_df$Measure == "GCT 60 min plasma", ],
+  aes(x = Diabetes, y = Value, fill = Diabetes, colour = Diabetes)
+) +
+  geom_jitter() +
+  geom_hline(yintercept = 146.5, linetype = "dashed") +
+  ylim(c(50, 300)) +
+  ylab("mg/dL") +
+  ggtitle("GCTpl") +
+  theme_bw(base_size = 15) +
+  theme(
+    legend.position = "none",
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.x = element_blank()
+  )
+box_p2 <- ggplot(
+  plot_df[plot_df$Measure == "GCT 0 min capillary", ],
+  aes(x = Diabetes, y = Value, fill = Diabetes, colour = Diabetes)
+) +
+  geom_jitter() +
+  geom_hline(yintercept = 98.5, linetype = "dashed") +
+  ylim(c(50, 300)) +
+  ggtitle("RCG") +
+  theme_bw(base_size = 15) +
+  theme(
+    legend.position = "none",
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank()
+  )
+box_p3 <- ggplot(
+  plot_df[plot_df$Measure == "HbA1c", ],
+  aes(x = Diabetes, y = Value, fill = Diabetes, colour = Diabetes)
+) +
+  geom_jitter() +
+  geom_hline(yintercept = 5.45, linetype = "dashed") +
+  ylab("%") +
+  ggtitle("HbA1c") +
+  theme_bw(base_size = 15) +
+  theme(
+    legend.position = "none",
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank()
+  )
+box_p <- box_p1 + box_p2 + box_p3
+ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/dotplot2.png",
   box_p,
@@
 options(datadist = "ddist")
-nom <- nomogram(f, fun = plogis, funlabel = "Risk of CFRD",lp = F)
+nom <- nomogram(f, fun = plogis, funlabel = "Risk of CFRD", lp = F)
 plot(nom)

### 2025-06-09 Christine Chan/EnVision CF/cfrd_roc.qmd
updated EnVision ROC curves

@@
 library(gt)
+library(patchwork)
 home_dir <- switch(Sys.info()["sysname"],
@@
 # Import cleaned dataset
-df <- read.csv("./Data_Clean/envision_analysis_dataset.csv", na.strings = "")
-# Limit to those with CGM data (based on mean glucose)
+load("./Data_Clean/envision_analysis_dataset.RData")
+df <- final_df
+# Limit to those with CGM data (based on mean glucose) and OGTT diagnosis
 df <- df %>% filter(!is.na(mean_sensor))
+# Convert CV units to match Scully et al.
+df$cv <- df$cv * 100
 ```
@@
 - In the cleaned EnVision dataset we have a total of `r nrow(df)` CGMs from `r length(unique(df$study_id))` unique individuals.
-Of these CGM wears, `r sum(is.na(df$Diagnosis))` are missing a diagnosis due to missing OGTT timepoints.
+Of these CGM wears, `r sum(is.na(df$Diagnosis))` were missing a diagnosis due to missing OGTT timepoints.
+Those without a diagnosis were excluded from all analyses.
 - BMI percentile was calculated with age = 20 for adults.
+
+```{r}
+df <- df %>% filter(!is.na(Diagnosis))
+```
+
+## By OGTT categories
+
+### Detailed OGTT category
 
@@
 t1 <- tableby(
-  CFRD ~ redcap_data_access_group + sex + origin_race + ethnicity +
+  Diagnosis ~ redcap_data_access_group + sex + origin_race + ethnicity +
     pancreatic_status + CFTR + fev1 + fvc + bmi_perc + a1c_result,
@@
 
-# ROC Analyses
+### General OGTT category
+
+```{r results='asis'}
+t2 <- tableby(
+  Dysglycemia ~ redcap_data_access_group + sex + origin_race + ethnicity +
+    pancreatic_status + CFTR + fev1 + fvc + bmi_perc + a1c_result,
+  data = df
+)
+summary(t2, pfootnote = T)
+```
+
+### CFRD status
+
+```{r results='asis'}
+t3 <- tableby(
+  CFRD ~ redcap_data_access_group + sex + origin_race + ethnicity +
+    pancreatic_status + CFTR + fev1 + fvc + bmi_perc + a1c_result,
+  data = df
+)
+summary(t3, pfootnote = T)
+```
+
+# ROC analyses
 
@@
 
-# Questions
+## HbA1c
 
-- Are we including fasting glucose >= 126 in the definition of CFRD for this analysis?
+```{r}
+r <- roc(CFRD ~ a1c_result, data = df, quiet = T)
+plot(r, print.auc = T, print.thres = T, legacy.axes = T)
+gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
+```
+
+# Pulmonary function comparisons
+
+## By OGTT categories
+
+### Detailed OGTT category
+
+```{r results='asis'}
+t1 <- tableby(Diagnosis ~ fev1 + fvc, data = df)
+summary(t1, pfootnote = T)
+```
+
+### General OGTT category
+
+```{r results='asis'}
+t2 <- tableby(Dysglycemia ~ fev1 + fvc, data = df)
+summary(t2, pfootnote = T)
+```
+
+### CFRD status
+
+```{r results='asis'}
+t3 <- tableby(CFRD ~ fev1 + fvc, data = df)
+summary(t3, pfootnote = T)
+```
+
+## By CGM category
+
+```{r}
+df$scully_ag <- cut(df$mean_sensor, c(-Inf, 112, Inf))
+df$our_ag <- cut(df$mean_sensor, c(-Inf, 101, Inf))
+df$scully_140 <- cut(df$percent_time_over_140, c(-Inf, 17.5, Inf))
+df$our_140 <- cut(df$percent_time_over_140, c(-Inf, 11.3, Inf))
+df$scully_180 <- cut(df$percent_time_over_180, c(-Inf, 3.4, Inf))
+df$our_180 <- cut(df$percent_time_over_180, c(-Inf, 3.1, Inf))
+df$scully_sd <- cut(df$standard_deviation, c(-Inf, 30.8, Inf))
+df$our_sd <- cut(df$standard_deviation, c(-Inf, 28.8, Inf))
+df$scully_cv <- cut(df$cv, c(-Inf, 30.6, Inf))
+df$our_cv <- cut(df$cv, c(-Inf, 27.7, Inf))
+```
+
+### Mean glucose
+
+#### Scully et al. cutoff
+
+```{r results='asis'}
+tb <- tableby(scully_ag ~ fev1 + fvc, data = df)
+summary(tb, pfootnote = T)
+```
+
+#### Our cutoff
+
+```{r results='asis'}
+tb <- tableby(our_ag ~ fev1 + fvc, data = df)
+summary(tb, pfootnote = T)
+```
+
+### Percent time spent >140 mg/dL
+
+#### Scully et al. cutoff
+
+```{r results='asis'}
+tb <- tableby(scully_140 ~ fev1 + fvc, data = df)
+summary(tb, pfootnote = T)
+```
+
+#### Our cutoff
+
+```{r results='asis'}
+tb <- tableby(our_140 ~ fev1 + fvc, data = df)
+summary(tb, pfootnote = T)
+```
+
+### Percent time spent >180 mg/dL
+
+#### Scully et al. cutoff
+
+```{r results='asis'}
+tb <- tableby(scully_180 ~ fev1 + fvc, data = df)
+summary(tb, pfootnote = T)
+```
+
+#### Our cutoff
+
+```{r results='asis'}
+tb <- tableby(our_180 ~ fev1 + fvc, data = df)
+summary(tb, pfootnote = T)
+```
+
+### SD
+
+#### Scully et al. cutoff
+
+```{r results='asis'}
+tb <- tableby(scully_sd ~ fev1 + fvc, data = df)
+summary(tb, pfootnote = T)
+```
+
+#### Our cutoff
+
+```{r results='asis'}
+tb <- tableby(our_sd ~ fev1 + fvc, data = df)
+summary(tb, pfootnote = T)
+```
+
+### CV
+
+#### Scully et al. cutoff
+
+```{r results='asis'}
+tb <- tableby(scully_cv ~ fev1 + fvc, data = df)
+summary(tb, pfootnote = T)
+```
+
+#### Our cutoff
+
+```{r results='asis'}
+tb <- tableby(our_cv ~ fev1 + fvc, data = df)
+summary(tb, pfootnote = T)
+```
+
+# Figure
+
+In the figure below, red lines indicate the cutoffs identified in Scully et al. and purple lines indicate the cutoffs identified in EnVision.
+
+```{r}
+#| message: false
+# Dotplot
+box_p1 <- ggplot(
+  df,
+  aes(x = Dysglycemia, y = percent_time_over_140, colour = Dysglycemia)
+) +
+  geom_jitter() +
+  geom_hline(yintercept = 17.5, color = "red") +
+  geom_hline(yintercept = 11.3, color = "purple") +
+  ggtitle("% Time Over 140 (mg/dL)") +
+  theme_bw(base_size = 12) +
+  theme(
+    legend.position = "none",
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.x = element_blank(), axis.title.y = element_blank()
+  ) +
+  scale_color_manual(values = c("#4DA8DA", "#80D8C3", "#FFD66B"))
+box_p2 <- ggplot(
+  df,
+  aes(x = Dysglycemia, y = percent_time_over_180, colour = Dysglycemia)
+) +
+  geom_jitter() +
+  geom_hline(yintercept = 3.4, color = "red") +
+  geom_hline(yintercept = 3.1, color = "purple") +
+  ggtitle("% Time Over 180 (mg/dL)") +
+  theme_bw(base_size = 12) +
+  theme(
+    legend.position = "none",
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.x = element_blank(), axis.title.y = element_blank()
+  ) +
+  scale_color_manual(values = c("#4DA8DA", "#80D8C3", "#FFD66B"))
+box_p3 <- ggplot(
+  df,
+  aes(x = Dysglycemia, y = mean_sensor, colour = Dysglycemia)
+) +
+  geom_jitter() +
+  geom_hline(yintercept = 112, color = "red") +
+  geom_hline(yintercept = 101, color = "purple") +
+  ggtitle("Mean glucose (mg/dL)") +
+  theme_bw(base_size = 12) +
+  theme(
+    legend.position = "none",
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.x = element_blank(), axis.title.y = element_blank()
+  ) +
+  scale_color_manual(values = c("#4DA8DA", "#80D8C3", "#FFD66B"))
+box_p4 <- ggplot(
+  df,
+  aes(x = Dysglycemia, y = cv, colour = Dysglycemia)
+) +
+  geom_jitter() +
+  geom_hline(yintercept = 30.6, color = "red") +
+  geom_hline(yintercept = 27.7, color = "purple") +
+  ggtitle("Coefficient of Variation (%)") +
+  theme_bw(base_size = 12) +
+  theme(
+    legend.position = "none",
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.x = element_blank(), axis.title.y = element_blank()
+  ) +
+  scale_color_manual(values = c("#4DA8DA", "#80D8C3", "#FFD66B"))
+box_p5 <- ggplot(
+  df,
+  aes(x = Dysglycemia, y = a1c_result, colour = Dysglycemia)
+) +
+  geom_jitter() +
+  geom_hline(yintercept = 5.8, color = "red") +
+  geom_hline(yintercept = 5.75, color = "purple") +
+  ggtitle("HbA1c (%)") +
+  theme_bw(base_size = 12) +
+  theme(
+    legend.position = "none",
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.x = element_blank(), axis.title.y = element_blank()
+  ) +
+  scale_color_manual(values = c("#4DA8DA", "#80D8C3", "#FFD66B"))
+box_p <- box_p1 + box_p2 + box_p4 + box_p3 + box_p5 + plot_layout(ncol = 5)
+box_p
+```

### 2025-06-09 Christine Chan/EnVision CF/clean_envision_data.R

@@ -371,8 +371,14 @@ final_df$Diagnosis[final_df$Glucose_0 >= 126 | final_df$Glucose_120 >= 200] <-
 final_df$Diagnosis <- factor(final_df$Diagnosis,
   levels = c("NGT", "IFG", "INDET", "IGT", "CFRD")
 )
+# Dysglycemia variable
+final_df$Dysglycemia <- factor(final_df$Diagnosis,
+  levels = c("NGT", "IFG", "INDET", "IGT", "CFRD"),
+  labels = c("NGT", "AGT", "AGT", "AGT", "CFRD")
+)
 # Create CFRD+ vs. CFRD- variable
 final_df$CFRD <- factor(final_df$Diagnosis,
+  levels = c("NGT", "IFG", "INDET", "IGT", "CFRD"),
   labels = c("CFRD-", "CFRD-", "CFRD-", "CFRD-", "CFRD+")
 )
 # iAUCs
@@ -540,3 +546,4 @@ write.csv(final_df,
   file = "./Christine Chan/EnVision CF/Data_Clean/envision_analysis_dataset.csv",
   row.names = F, na = ""
 )
+save(final_df, file = "./Christine Chan/EnVision CF/Data_Clean/envision_analysis_dataset.RData")

### 2025-06-09 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
old analysis for new Na:Cl data

--- 2025-04-06 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
@@
 #| include: false
+library(knitr)
 library(Hmisc)
@@
 library(arsenal)
+library(pROC)
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Variability"
+  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Variability",
+  "Linux" = "/home/tim/OneDrive/Vigers/CF/Edith Zemanick/Sweat Variability"
 )
@@
   filter(Demographics.Available == "TRUE") %>%
-  # Get only the earliest test
-  slice_min(Age.at.Test.in.Weeks) %>%
   select(
@@
   filter(Demographics.Available == "TRUE") %>%
-  # Get only the earliest test
-  slice_min(Age.at.test..in.years.) %>%
   select(
@@
   filter(!is.na(Diagnosis))
+# Combine CFTR-related disorder and CRMS
+df$Diagnosis <- factor(df$Diagnosis,
+  levels = c("Cystic Fibrosis", "CFTR-related disorder", "CRMS", "Non-CF"),
+  labels = c("Cystic Fibrosis", "CFTR-RD or CRMS", "CFTR-RD or CRMS", "Non-CF")
+)
+# Calculate Na:Cl ratio
+df$`Na:Cl` <- df$Na / df$Cl
 # Set table options
-mycontrols <-
-  tableby.control(numeric.stats = c("Nmiss", "meansd", "medianq1q3", "range"))
+mycontrols <- tableby.control(
+  numeric.stats = c("N", "Nmiss", "meansd", "medianq1q3", "range")
+)
 # Save this for Angela et al. to check
@@
 ```{r results='asis'}
-summary(tableby(Diagnosis ~ Sex + Age + Na + K + Cl,
+summary(tableby(Diagnosis ~ Sex + Age + Na + Cl + `Na:Cl` + K,
   data = df, control = mycontrols
@@
 
+## Chloride
+
 ```{r}
-ggplot(df, aes(x = Diagnosis, y = Cl)) +
+#| warning: false
+clp <- ggplot(df, aes(x = Diagnosis, y = Cl)) +
   geom_boxplot() +
   theme_bw()
-ggplot(df, aes(x = Diagnosis, y = Na)) +
+clp
+```
+
+## Sodium
+
+```{r}
+#| warning: false
+nap <- ggplot(df, aes(x = Diagnosis, y = Na)) +
   geom_boxplot() +
   theme_bw()
-ggplot(df, aes(x = Diagnosis, y = Na / Cl)) +
+nap
+```
+
+## Na:Cl
+
+```{r}
+#| warning: false
+ratiop <- ggplot(df, aes(x = Diagnosis, y = `Na:Cl`)) +
   geom_boxplot() +
-  theme_bw()
-ggplot(df, aes(x = Diagnosis, y = K)) +
+  geom_hline(yintercept = 1, color = "red") +
+  geom_hline(yintercept = 1.25, color = "red") +
+  theme_bw() +
+  ylab("Na:Cl") +
+  xlab("Modulator")
+ratiop
+ratiop <- ggplot(df, aes(x = Diagnosis, y = `Na:Cl`)) +
   geom_boxplot() +
-  theme_bw()
+  geom_hline(yintercept = 1, color = "red") +
+  geom_hline(yintercept = 1.25, color = "red") +
+  ylim(0, 2) +
+  theme_bw() +
+  ylab("Na:Cl") +
+  xlab("Modulator")
+ratiop
 ```
+
+# Histograms and Density Plots
+
+## Chloride
+
+```{r}
+#| warning: false
+ggplot(df, aes(x = Cl, fill = Diagnosis)) +
+  geom_histogram(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
+  geom_histogram(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
+  geom_histogram(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
+  theme_bw() +
+  theme(axis.title.y = element_blank())
+ggplot(df, aes(x = Cl, fill = Diagnosis)) +
+  geom_density(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
+  geom_density(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
+  geom_density(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
+  theme_bw() +
+  theme(axis.title.y = element_blank())
+```
+
+## Sodium
+
+```{r}
+#| warning: false
+ggplot(df, aes(x = Na, fill = Diagnosis)) +
+  geom_histogram(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
+  geom_histogram(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
+  geom_histogram(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
+  theme_bw() +
+  theme(axis.title.y = element_blank())
+ggplot(df, aes(x = Na, fill = Diagnosis)) +
+  geom_density(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
+  geom_density(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
+  geom_density(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
+  theme_bw() +
+  theme(axis.title.y = element_blank())
+```
+
+## Na:Cl
+
+```{r}
+#| warning: false
+ggplot(df, aes(x = `Na:Cl`, fill = Diagnosis)) +
+  geom_histogram(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
+  geom_histogram(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
+  geom_histogram(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
+  theme_bw() +
+  theme(axis.title.y = element_blank())
+ggplot(df, aes(x = `Na:Cl`, fill = Diagnosis)) +
+  geom_density(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
+  geom_density(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
+  geom_density(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
+  theme_bw() +
+  theme(axis.title.y = element_blank())
+ggplot(df, aes(x = `Na:Cl`, fill = Diagnosis)) +
+  geom_density(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
+  geom_density(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
+  geom_density(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
+  xlim(0, 2) +
+  theme_bw() +
+  theme(axis.title.y = element_blank())
+```
+
+# Na:Cl ROC Curves
+
+```{r}
+# Fit a multiclass ROC
+mc_roc <- multiclass.roc(df$Diagnosis, df$`Na:Cl`, quiet = T)
+roc_list <- list(
+  "Cystic Fibrosis vs. Non-CF" = mc_roc$rocs[[2]],
+  "CFTR-RD or CRMS vs. Cystic Fibrosis" = mc_roc$rocs[[1]],
+  "CFTR-RD or CRMS vs. Non-CF" = mc_roc$rocs[[3]]
+)
+ggroc(roc_list, legacy.axes = T) +
+  theme_bw() +
+  theme(legend.title = element_blank())
+roc_table <- lapply(roc_list, function(r) {
+  coords(r, "best",
+    best.method = "youden",
+    ret = c(
+      "threshold", "sensitivity", "specificity", "accuracy", "tpr", "fpr",
+      "tnr", "fnr"
+    )
+  )
+})
+roc_table <- do.call(rbind, roc_table)
+kable(roc_table, digits = 3)
+```
+
+- The multiclass AUC for Na:Cl is `r round(as.numeric(auc(mc_roc)),3)`. 
+- The AUC for using Na:Cl to distinguish CF from non-CF was `r round(as.numeric(auc(mc_roc$rocs[[2]])),3)`
+- The best threshold for distinguishing CF from non-CF (based on the Youden index) was `r round(roc_table["Cystic Fibrosis vs.
+Non-CF","threshold"],3)`, and provided a sensitivity of `r round(roc_table["Cystic Fibrosis vs.
+Non-CF","sensitivity"],3)` and specificity of `r round(roc_table["Cystic Fibrosis vs.
+Non-CF","specificity"],3)`
+
+- The AUC for using Na:Cl to distinguish CF from CFTR-RD or CRMS was `r round(as.numeric(auc(mc_roc$rocs[[1]])),3)`
+- The best threshold for distinguishing CF from CFTR-RD or CRMS (based on the Youden index) was `r round(roc_table["CFTR-RD or CRMS vs.
+Cystic Fibrosis","threshold"],3)`, and provided a sensitivity of `r round(roc_table["CFTR-RD or CRMS vs.
+Cystic Fibrosis","sensitivity"],3)` and specificity of `r round(roc_table["CFTR-RD or CRMS vs.
+Cystic Fibrosis","specificity"],3)`
+
+- The AUC for using Na:Cl to distinguish non-CF from CFTR-RD or CRMS was `r round(as.numeric(auc(mc_roc$rocs[[3]])),3)`
+- The best threshold for distinguishing CF from CFTR-RD or CRMS (based on the Youden index) was `r round(roc_table["CFTR-RD or CRMS vs.
+Non-CF","threshold"],3)`, and provided a sensitivity of `r round(roc_table["CFTR-RD or CRMS vs.
+Non-CF","sensitivity"],3)` and specificity of `r round(roc_table["CFTR-RD or CRMS vs.
+Non-CF","specificity"],3)`

### 2025-06-09 Emily Muther/mental_health_screening_psc.qmd
PSC screening no more mixed models

--- 2025-05-02 Emily Muther/mental_health_screening_psc.qmd
@@
 library(readxl)
+library(arsenal)
+library(lubridate)
+library(lmerTest)
+library(GLMMadaptive)
+library(knitr)
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Emily Muther",
-  "Linux" = "/home/timvigers/OneDrive"
+  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/tim/OneDrive/Vigers/CF"
 )
 github_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Documents/GitHub",
-  "Linux" = "/home/timvigers/Documents/GitHub"
+  "Darwin" = "/Users/tim/Documents/GitHub",
+  "Windows" = "C:/Users/Tim/Documents/GitHub",
+  "Linux" = "/home/tim/Documents/GitHub"
 )
@@
 ```{r data cleaning}
+#| message: false
 # Import
-df <- read_excel("./Muther Research Group - General/PSC manuscript 2025/PSC_9-20-19 thru 4-18-25_pulled 4-21-25.xlsx")
-# Fill down MRN, select columns, remove duplicates
+df <- read_excel("./Emily Muther/Muther Research Group - General/PSC manuscript 2025/Data_Raw/PSC_9-20-19 thru 4-18-25_pulled 4-21-25.xlsx")
+df$`Contact Date` <- mdy(df$`Contact Date`)
+# Fill down MRN abd contact date
 df <- df %>%
-  fill(`Pat Mrn Id`, .direction = "down") %>%
+  fill(`Pat Mrn Id`, .direction = "down")
+df <- df %>%
+  group_by(`Pat Mrn Id`, `AGE AT CONTACT`) %>%
+  fill(`Contact Date`)
+# Classify insurance per Emily
+df$Insurance <- factor(df$`Payor Name`,
+  levels = c(
+    "MEDICAID", "MEDICAID NEW MEXICO", "MEDICAID WYOMING", "UNITED HEALTHCARE",
+    "CIGNA", "ANTHEM BCBS", "ROCKY MOUNTAIN HEALTH PLAN", "AETNA",
+    "MEDICAID OTHER STATES", "TRICARE", "CICP", "CHP+COLORADO ACCESS",
+    "CU GME HEALTH", "MIDLANDS CHOICE", "BRIGHT HEALTH", "FIRST HEALTH GROUP",
+    "ROCKY MOUNTAIN MEDICAID", "KAISER", "MEDICAID NEBRASKA",
+    "MEDICAID OKLAHOMA", "EBMS", "COFINITY", "SIMPLIFIED BENEFITS ADMINISTRATORS",
+    "FRIDAY HEALTH PLANS", "SELECTHEALTH, INC."
+  ),
+  labels = c(
+    "Public", "Public", "Public", "Private", "Private", "Private", "Public",
+    "Private", "Public", "Private", "Public", "Public", "Private", "Private",
+    "Private", "Private", "Public", "Private", "Public", "Public", "Private",
+    "Private", "Private", "Private", "Private"
+  )
+)
+# Convert to wide
+df <- df %>%
   select(
-    `Pat Mrn Id`, `Payor Name`, `AGE AT CONTACT`, `Internalizing subscale`,
-    `Attention subscale`, `Externalizing subscale`, Score
+    `Pat Mrn Id`, `Contact Date`, `AGE AT CONTACT`, Insurance,
+    `Internalizing subscale`, `Attention subscale`, `Externalizing subscale`,
+    Score
   ) %>%
-  distinct(`Pat Mrn Id`, `AGE AT CONTACT`, .keep_all = T)
+  distinct() %>%
+  pivot_wider(names_from = Insurance, values_from = Insurance) %>%
+  unite("Insurance", Public, Private, `NA`, remove = T, na.rm = T)
+df$Insurance <- factor(df$Insurance,
+  levels = c("", "Private", "Public", "Public_Private"),
+  labels = c("None", "Private", "Public", "Both")
+)
+# Add data from Elin
+demo <- read_excel("./Emily Muther/Muther Research Group - General/PSC manuscript 2025/Data_Raw/PSC Data 2025 05 21.xlsx", sheet = "Demographics")
+demo$`Pat Mrn Id` <- as.character(demo$`Pat Mrn Id`)
+demo$Genotype <- rowSums(demo[, c("Mutation 1", "Mutation 2")] == "F508del")
+demo$Genotype <- factor(demo$Genotype,
+  levels = c(2, 1, 0),
+  labels = c("F508del homozygous", "F508del heterozygous", "Other/Other")
+)
+demo$Sex <- factor(demo$Sex, levels = c("M", "F"), labels = c("Male", "Female"))
+races <- c(
+  "Asian", "Black/African-American", "White", "Native American/Alaska Native",
+  "Middle Eastern/Arabic", "Race Unknown"
+)
+demo$Race <- apply(demo %>% select(Asian:`Race Unknown`), 1, function(r) {
+  w <- which(r == T)
+  if (length(w) > 1 | w == 6) {
+    "Unknown/More than one"
+  } else {
+    races[w]
+  }
+})
+demo$Race <- factor(demo$Race)
+demo$`Hispanic or Latino` <-
+  factor(demo$`Hispanic or Latino`, levels = c(F, T), labels = c("No", "Yes"))
+visits <- read_excel("./Emily Muther/Muther Research Group - General/PSC manuscript 2025/Data_Raw/PSC Data 2025 05 21.xlsx", sheet = "Visits")
+visits$`Pat Mrn Id` <- as.character(visits$`Pat Mrn Id`)
+visits <- visits %>% select(-`FEV1 raw value`)
+# Combine levels of family income
+df <- left_join(df, demo, by = join_by(`Pat Mrn Id`))
+df <- left_join(df, visits, by = join_by(`Pat Mrn Id`, `Contact Date`))
+# Convert to numeric
+df$`AGE AT CONTACT` <- as.numeric(df$`AGE AT CONTACT`)
+df$`Internalizing subscale` <- as.numeric(df$`Internalizing subscale`)
+df$`Attention subscale` <- as.numeric(df$`Attention subscale`)
+df$`Externalizing subscale` <- as.numeric(df$`Externalizing subscale`)
+df$Score <- as.numeric(df$Score)
+# Make age group
+df$`Age Group` <- cut(df$`AGE AT CONTACT`,
+  breaks = c(-Inf, 6, 10, Inf),
+  labels = c("Age < 6", "Age >= 6 and < 10", "Age >= 10"), right = F
+)
+# Make variables for 0 scores and elevated scores on each scale
+df$`Internalizing subscale elevated` <- factor(df$`Internalizing subscale` >= 5,
+  levels = c(F, T), labels = c("Not elevated", "Elevated")
+)
+df$`Internalizing subscale symptoms` <- factor(df$`Internalizing subscale` > 0,
+  levels = c(F, T),
+  labels = c(
+    "No internalizing symptoms (score = 0)",
+    "Any internalizing symptoms (score > 0)"
+  )
+)
+df$`Attention subscale elevated` <- factor(df$`Attention subscale` >= 7,
+  levels = c(F, T), labels = c("Not elevated", "Elevated")
+)
+df$`Attention subscale symptoms` <- factor(df$`Attention subscale` > 0,
+  levels = c(F, T),
+  labels = c(
+    "No attention symptoms (score = 0)",
+    "Any attention symptoms (score > 0)"
+  )
+)
+df$`Externalizing subscale elevated` <- factor(df$`Internalizing subscale` >= 7,
+  levels = c(F, T), labels = c("Not elevated", "Elevated")
+)
+df$`Externalizing subscale symptoms` <- factor(df$`Externalizing subscale` > 0,
+  levels = c(F, T),
+  labels = c(
+    "No externalizing symptoms (score = 0)",
+    "Any externalizing symptoms (score > 0)"
+  )
+)
+df$`Score elevated` <- factor(
+  df$Score >= 15 |
+    df$`Internalizing subscale` >= 5 |
+    df$`Attention subscale` >= 7 |
+    df$`Internalizing subscale` >= 7,
+  levels = c(F, T), labels = c("Not elevated", "Elevated")
+)
+df$`Any symptoms` <- factor(df$Score > 0,
+  levels = c(F, T),
+  labels = c(
+    "No symptoms (total score = 0)",
+    "Any symptoms (total score > 0)"
+  )
+)
+# Determine if they were ever elevated on each subscale
+df <- df %>%
+  group_by(`Pat Mrn Id`) %>%
+  mutate(
+    `Internalizing subscale ever elevated` =
+      any(`Internalizing subscale elevated` == "Elevated"),
+    `Attention subscale ever elevated` =
+      any(`Attention subscale elevated` == "Elevated"),
+    `Externalizing subscale ever elevated` =
+      any(`Externalizing subscale elevated` == "Elevated"),
+    `Score ever elevated` = any(`Score elevated` == "Elevated"),
+  )
+# Set table options
+mycontrols <-
+  tableby.control(
+    numeric.stats = c("N", "Nmiss", "medianq1q3", "range"),
+    cat.stats = c("N", "Nmiss", "countpct")
+  )
+# Basic formula
+f <- ~ Genotype + Race + `Hispanic or Latino` + Insurance +
+  `Annual Family Income` + BMI + `FEV1 % predicted` + Alyftrek + Kalydeco +
+  Orkambi + Symdeko + Trikafta +
+  `Internalizing subscale` + `Internalizing subscale elevated` + `Internalizing subscale symptoms` +
+  `Attention subscale` + `Attention subscale elevated` + `Attention subscale symptoms` +
+  `Externalizing subscale` + `Externalizing subscale elevated` + `Externalizing subscale symptoms` +
+  Score + `Score elevated` + `Any symptoms`
 ```
 
-# Table
+# Score summary at first visit
+
+## By age group
+
+```{r results='asis'}
+first_visit <- df %>%
+  group_by(`Pat Mrn Id`) %>%
+  slice_min(`AGE AT CONTACT`)
+t1 <- tableby(update(f, `Age Group` ~ .), data = first_visit)
+summary(t1, pfootnote = T)
+```
+
+# Plots of participants who were ever elevated
+
+## Internalizing subscale
 
 ```{r}
-#| label: tbl-planets
-#| tbl-cap: Planets
+ggplot(
+  subset(df, `Internalizing subscale ever elevated` == T),
+  aes(x = `AGE AT CONTACT`, y = `Internalizing subscale`)
+) +
+  geom_hline(yintercept = 5, color = "red") +
+  geom_point() +
+  geom_line(aes(group = `Pat Mrn Id`)) +
+  theme_bw()
 ```
 
-# Figure
+## Attention subscale
 
 ```{r}
-#| label: fig-planets
-#| fig-cap: Planets
+ggplot(
+  subset(df, `Attention subscale ever elevated` == T),
+  aes(x = `AGE AT CONTACT`, y = `Attention subscale`)
+) +
+  geom_hline(yintercept = 7, color = "red") +
+  geom_point() +
+  geom_line(aes(group = `Pat Mrn Id`)) +
+  theme_bw()
 ```
+
+## Externalizing subscale
+
+```{r}
+ggplot(
+  subset(df, `Externalizing subscale ever elevated` == T),
+  aes(x = `AGE AT CONTACT`, y = `Externalizing subscale`)
+) +
+  geom_hline(yintercept = 7, color = "red") +
+  geom_point() +
+  geom_line(aes(group = `Pat Mrn Id`)) +
+  theme_bw()
+```
+
+## Total score
+
+```{r}
+ggplot(
+  subset(df, `Score ever elevated` == T),
+  aes(x = `AGE AT CONTACT`, y = Score)
+) +
+  geom_hline(yintercept = 15, color = "red") +
+  geom_point() +
+  geom_line(aes(group = `Pat Mrn Id`)) +
+  theme_bw()
+```

### 2025-06-09 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R
nicer flowchart for race and ETI

@@ -206,43 +206,67 @@ n_people_6 <- length(unique(encounter$eDWID))
 t1_participant <- t1_participant %>% filter(!eDWID %in% no_post)
 # Flowchart
 # Make labels
-tots1 <- paste0(n_people_1, " People\n", n_enc_1, " Encounters")
+tots1 <- paste0(
+  format(n_people_1, big.mark = ","), " People\n",
+  format(n_enc_1, big.mark = ","), " Encounters"
+)
 no_date <- paste0(
-  "No ETI Date:\n", n_people_no_eti_date, " People\n",
-  length(no_eti_date), " Encounters"
+  "No ETI Date:\n", format(n_people_no_eti_date, big.mark = ","), " People\n",
+  format(length(no_eti_date), big.mark = ","), " Encounters"
 )
-tots2 <- paste0(n_people_2, " People\n", n_enc_2, " Encounters")
+tots2 <- paste0(
+  format(n_people_2, big.mark = ","), " People\n",
+  format(n_enc_2, , big.mark = ","), " Encounters"
+)
 hosps <- paste0(
-  "Care Episodes:\n", n_people_hosp, " People\n",
-  length(hosp), " Encounters"
+  "Care Episodes:\n", format(n_people_hosp, big.mark = ","), " People\n",
+  format(length(hosp), big.mark = ","), " Encounters"
 )
-tots3 <- paste0(n_people_3, " People\n", n_enc_3, " Encounters")
+tots3 <- paste0(
+  format(n_people_3, big.mark = ","), " People\n",
+  format(n_enc_3, big.mark = ","), " Encounters"
+)
 pregs <- paste0(
-  "Pregnancies:\n", n_people_preg, " People\n",
-  length(preg), " Encounters"
+  "Pregnancies:\n", format(n_people_preg, big.mark = ","), " People\n",
+  format(length(preg), big.mark = ","), " Encounters"
 )
-tots4 <- paste0(n_people_4, " People\n", n_enc_4, " Encounters")
+tots4 <- paste0(
+  format(n_people_4, big.mark = ","), " People\n",
+  format(n_enc_4, big.mark = ","), " Encounters"
+)
 earlies <- paste0(
-  "Visits > ", 365.25 * 3, " Days BEFORE ETI:\n", n_people_early, " People\n",
-  length(early), " Encounters"
+  "Visits > ", format(365.25 * 3, big.mark = ","), " Days BEFORE ETI:\n",
[…]

### 2025-06-09 Stacey Martiniano/ntm_agreement.qmd
new NTM score

@@
     page-layout: full
-    theme:
-      light: flatly
-      dark: darkly
-bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
-csl: /Users/timvigers/Dropbox/Miscellaneous/american-medical-association.csl
 editor: source
@@
 
-```{r data cleaning}
-#| message: false
+```{r}
+#| include: false
 library(tidyverse)
@@
 library(performance)
-# Import data
-api <- read.table("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt", header = F)
-rcon <- redcapConnection(
-  url = "https://redcap.ucdenver.edu/api/",
-  token = api[1, 1]
+# REDCap
+unlockREDCap(c(rcon = "NTM CT Scoring"),
+  keyring = "API_KEYs",
+  envir = globalenv(),
+  url = "https://redcap.ucdenver.edu/api/"
 )
+# Home directory
+home_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/tim/OneDrive/Vigers/CF"
+)
+github_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/tim/Documents/GitHub",
+  "Windows" = "C:/Users/Tim/Documents/GitHub",
+  "Linux" = "/home/tim/Documents/GitHub"
+)
+knitr::opts_knit$set(root.dir = home_dir)
+```
+
+```{r data cleaning}
+#| message: false
+#| warning: false
+# Import data from REDCap
 data <- exportRecordsTyped(rcon, labels = F)
@@
 # Revert duplicate IDs back to original to match with gold standard
-training <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/training.key.csv")
-dups <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/duplicated.key.csv")
+training <- read.csv("./Stacey Martiniano/NTM CT Scoring/Data_Clean/training.key.csv")
+dups <- read.csv("./Stacey Martiniano/NTM CT Scoring/Data_Clean/duplicated.key.csv")
 training$Training_id <- as.numeric(sub("T00", "", training$Training_id))
@@
 ```{r}
+#| message: false
 # Find scans with disagreement on certain features, so Stacey can discuss them with radiologists
@@
 
+# Comparison of scores between groups
+
+```{r}
+# Import disease state
+outcomes <- read.csv("./Stacey Martiniano/NTM CT Scoring/Data_Clean/disease.key.csv", na.strings = "")
+outcomes$NTM.PD..Y.N. <- factor(outcomes$NTM.PD..Y.N.,
+  levels = c("No", "Yes"), ordered = T
+)
+# For the duplicated scans, just use the first reading and combine with the
+# singletons
+scores <- rbind(data, data_dups[data_dups$scorer_repeat == 1, ])
+# Add gold standards
+scores <- rbind(scores, data_gs[data_gs$scorer == "GS", ])
+# Add outcome info
+scores <- left_join(scores,
+  outcomes %>% select(cff_registry_id, NTM.PD..Y.N.),
+  by = join_by(scan_id == cff_registry_id)
+) %>%
+  rename(`NTM-PD` = NTM.PD..Y.N.)
+# Calculate the scores for each feature.
+# When calculating the scores, take the majority vote by scan, feature, and
+# location, since that's how they scored it.
+# For example: If one scorer detected atelectasis in the RUL, one in the RML,
+# and one in the RLL, but the majority at each location was "absent," then that
+# scan would have a 0 for atelectasis. In other words, score the locations by
+# majority first, then add the locations together for the feature score.
+feature_scores <- scores %>% group_by(scan_id, Feature, `NTM-PD`)
+```
+
 # Mixed Model Approach
@@
 # Import data
-api <- read.table("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt", header = F)
-rcon <- redcapConnection(
-  url = "https://redcap.ucdenver.edu/api/",
-  token = api[1, 1]
-)
 data <- exportRecordsTyped(rcon, labels = F)

### 2025-06-10 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
additional plots for Larry (again)
updated GCT figures

@@
 plot_df$`OGTT Timepoint` <- factor(plot_df$`OGTT Timepoint`,
-  levels = c("OGTT 0 min", "OGTT 60 min", "OGTT 120 min")
+  levels = c("OGTT 0 min", "OGTT 60 min", "OGTT 120 min"),
+  labels = c("Fasting", "OGTT-1hG", "OGTT-2hG")
 )
@@
   geom_boxplot() +
+  ylab("Glucose (mg/dL)") +
   scale_y_continuous(breaks = c(100, 140, 200, 300)) +
+  scale_fill_manual(values = RColorBrewer::brewer.pal(3, "Paired")) +
   facet_wrap(~Diabetes) +
@@
   geom_point() +
+  ylab("GCTpl (mg/dL)") +
   theme_bw()
@@
 )
-# GCTpl, RCG, and HbA1c
+# GCTpl, RCG, HbA1c, OGTT1hr
 plot_df <- df %>%
-  pivot_longer(c(`GCT 60 min plasma`, `GCT 0 min capillary`, HbA1c),
+  pivot_longer(
+    c(
+      `GCT 60 min plasma`, `GCT 0 min capillary`, HbA1c,
+      `OGTT 60 min`
+    ),
     names_to = "Measure", values_to = "Value"
@@
   )
-box_p <- box_p1 + box_p2 + box_p3
+box_p4 <- ggplot(
+  plot_df[plot_df$Measure == "OGTT 60 min", ],
+  aes(x = Diabetes, y = Value, fill = Diabetes)
+) +
+  geom_boxplot() +
+  geom_hline(yintercept = 228.5, linetype = "dashed") +
+  ylim(c(50, 300)) +
+  ggtitle("OGTT 60 min") +
+  theme_bw(base_size = 15) +
+  theme(
+    legend.position = "none",
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank()
+  )
+box_p <- box_p1 + box_p2 + box_p3 + box_p4
 ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/boxplot2.png",
@@
 # Dotplot
-plot_df <- df %>%
-  pivot_longer(c(`GCT 60 min plasma`, `GCT 0 min capillary`, HbA1c),
-    names_to = "Measure", values_to = "Value"
-  )
 box_p1 <- ggplot(
@@
   geom_jitter() +
-  geom_hline(yintercept = 146.5, linetype = "dashed") +
+  geom_hline(yintercept = 146.5, linetype = "dashed", size = 1) +
   ylim(c(50, 300)) +
-  ylab("mg/dL") +
+  ylab("Glucose (mg/dL)") +
   ggtitle("GCTpl") +
   theme_bw(base_size = 15) +
+  scale_colour_manual(values = RColorBrewer::brewer.pal(3, "Set2")) +
   theme(
-    legend.position = "none",
+    legend.position = "none", plot.title = element_text(hjust = 0.5),
     axis.text.x = element_text(angle = 45, hjust = 1),
@@
   geom_jitter() +
-  geom_hline(yintercept = 98.5, linetype = "dashed") +
+  geom_hline(yintercept = 98.5, linetype = "dashed", size = 1) +
   ylim(c(50, 300)) +
@@
   theme_bw(base_size = 15) +
+  scale_colour_manual(values = RColorBrewer::brewer.pal(3, "Set2")) +
   theme(
-    legend.position = "none",
+    legend.position = "none", plot.title = element_text(hjust = 0.5),
     axis.text.x = element_text(angle = 45, hjust = 1),
@@
 ) +
-  geom_jitter() +
-  geom_hline(yintercept = 5.45, linetype = "dashed") +
+  geom_jitter(show.legend = F) +
+  geom_hline(aes(yintercept = 5.45, linetype = "Youden Index"), size = 1) +
   ylab("%") +
@@
   theme_bw(base_size = 15) +
+  scale_colour_manual(values = RColorBrewer::brewer.pal(3, "Set2")) +
+  scale_linetype_manual(values = c("Youden Index" = "dashed")) +
   theme(
-    legend.position = "none",
+    legend.title = element_blank(), plot.title = element_text(hjust = 0.5),
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.x = element_blank()
+  )
+box_p4 <- ggplot(
+  plot_df[plot_df$Measure == "OGTT 60 min", ],
+  aes(x = Diabetes, y = Value, fill = Diabetes, colour = Diabetes)
+) +
+  geom_jitter() +
+  geom_hline(yintercept = 228.5, linetype = "dashed", size = 1) +
+  ylim(c(50, 300)) +
+  ggtitle("OGTT 1hPG") +
+  theme_bw(base_size = 15) +
+  scale_colour_manual(values = RColorBrewer::brewer.pal(3, "Set2")) +
+  theme(
+    legend.position = "none", plot.title = element_text(hjust = 0.5),
     axis.text.x = element_text(angle = 45, hjust = 1),
@@
   )
-box_p <- box_p1 + box_p2 + box_p3
+box_p <- box_p1 + box_p2 + box_p4 + box_p3 + plot_layout(ncol = 4)
 ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/dotplot2.png",
@@
   ) +
+  ylab("OGTT-1hG ") +
+  xlab("GCTpl") +
   theme_bw(base_size = 20)

### 2025-06-12 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
added data from Angela

@@
 chco$PatientID <- NULL
+# Manually add some Na and Cl values that Angela was able to find (using
+# numbers so that no PHI is on GitHub)
+chco$Na_Right[791] <- 61.3
+chco$Na_Right[826] <- 72.4
+chco$Na_Left[826] <- 72.4
 # For "LESS THAN 10", replace with half the limit of detection
@@
   arrange(Name, Age)
-# Use each person's first sweat test
+# For those with left and right, average them
+df$Na_mean <- rowMeans(df[, c("Na_Left", "Na_Right")], na.rm = T)
+df$Na <- coalesce(df$Na, df$Na_mean)
+df$Cl_mean <- rowMeans(df[, c("Cl_Left", "Cl_Right")], na.rm = T)
+df$Cl <- coalesce(df$Cl, df$Cl_mean)
+df$K_mean <- rowMeans(df[, c("K_Left", "K_Right")], na.rm = T)
+df$K <- coalesce(df$K, df$K_mean)
+# Calculate Na:Cl ratio
+df$`Na:Cl` <- df$Na / df$Cl
+# Use each person's first sweat test with both sodium and chloride
 df <- df %>%
+  filter(!is.na(`Na:Cl`)) %>%
   group_by(Name) %>%
@@
 df <- df[!df$Name %in% dups, ]
-# For those with left and right, average them
-df$Na_mean <- rowMeans(df[, c("Na_Left", "Na_Right")])
-df$Na <- coalesce(df$Na, df$Na_mean)
-df$Cl_mean <- rowMeans(df[, c("Cl_Left", "Cl_Right")])
-df$Cl <- coalesce(df$Cl, df$Cl_mean)
-df$K_mean <- rowMeans(df[, c("K_Left", "K_Right")])
-df$K <- coalesce(df$K, df$K_mean)
 # Get rid of unnecessary columns, remove those with no diagnosis
 df <- df %>%
-  select(Name, Diagnosis, Genotypes1, Genotypes2, Sex, Age, Na, K, Cl) %>%
+  select(Name, Diagnosis, Genotypes1, Genotypes2, Sex, Age, Na, K, Cl, `Na:Cl`) %>%
   filter(!is.na(Diagnosis))
@@
 )
-# Calculate Na:Cl ratio
-df$`Na:Cl` <- df$Na / df$Cl
 # Set table options

### 2025-06-12 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
started GCT flowchart

@@
 library(patchwork)
+library(ggflowchart)
 home_dir <- switch(Sys.info()["sysname"],
@@
     names_to = "OGTT Timepoint", values_to = "Glucose"
-  ) %>%
-  filter(`OGTT Timepoint` != "OGTT 60 min")
+  )
 plot_df$`OGTT Timepoint` <- factor(plot_df$`OGTT Timepoint`,
-  levels = c("OGTT 0 min", "GCT 60 min plasma", "OGTT 120 min")
+  levels = c("OGTT 0 min", "OGTT 60 min", "OGTT 120 min", "GCT 60 min plasma"),
+  labels = c("Fasting", "OGTT-1hG", "OGTT-2hG", "GCTpl")
 )
@@
   geom_boxplot() +
+  ylab("Glucose (mg/dL)") +
   scale_y_continuous(breaks = c(100, 140, 200, 300)) +
+  scale_fill_manual(values = RColorBrewer::brewer.pal(4, "Set2")) +
   facet_wrap(~Diabetes) +
@@
 )
+```
+
+# Flowchart
+
+```{r}
+flow_df <- tibble(
+  from = c(
+    "Initial Screen with GCT",
+    "Initial Screen with GCT", 
+    "GCTpl >= 147",
+    "GCTpl < 147"
+  ),
+  to = c(
+    "GCTpl >= 147", 
+    "GCTpl < 147", 
+    "Oral Glucose Tolerance Test", 
+    "Repeat Screen in 1 Year"
+  )
+)
+
+ggflowchart(flow_df)
 ```

### 2025-06-12 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
WNAR background finished

@@
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
+  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
   "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
@@
 #     (is.na(A_VX809comb) | A_VX809comb == "No"))
+# Write cleaned data
+save(annual,file = "./Data_Clean/survival_dataset.RData")
 ```

### 2025-06-12 Christine Chan/CF Registry/wnar_presentation.qmd
WNAR background finished

---
title: "The Changing Landscape of Cystic Fibrosis Related Diabetes Risk Factors Between 2003-2018"
author: "Tim Vigers"
date: "June 17, 2025"
date-format: long
format: 
  revealjs:
    embed-resources: true
editor: visual
bibliography: wnar_references.bib
csl: american-medical-association.csl
---

```{r setup}
#| include: false
library(tidyverse)
library(survival)
library(survminer)
library(gtsummary)
knitr::opts_knit$set(root.dir = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry")
```

# Background

## The *CFTR* Gene

-   Cystic fibrosis (CF) is an autosomal recessive disease caused by mutations in the cystic fibrosis transmembrane conductance regulator (*CFTR*) gene.[@grasemannCysticFibrosis2023]
-   The CFTR protein is an ion channel that regulates electrolyte balance across cell membranes.[@stoltzOriginsCysticFibrosis2015]
-   As of 2023, over 2000 CFTR variants have been described. Of these, we are certain that over 700 are disease-causing.[@grasemannCysticFibrosis2023]
-   The *CFTR* gene is expressed in the upper and lower airways, **pancreas**, intestine, biliary tree, cervix, sweat glands, and other organ systems.[@grasemannCysticFibrosis2023]

## Cystic Fibrosis-Related Diabetes

-   Because *CFTR* is highly expressed in the pancreas, *CFTR* dysfunction commonly leads to pancreatic insufficiency.[@moheetNewConceptsPathogenesis2022]
-   The pathophysiology of CFRD is not fully understood, but we think that $\beta$-cell loss related to pancreas destruction is one of the major drivers of diabetes in people with CF.[@moheetNewConceptsPathogenesis2022]
-   CFRD is the most common non-pulmonary complication in CF. A diagnosis before the age of 10 is unusual, but 50% of adults over 40 have CFRD.[@moheetNewConceptsPathogenesis2022]
-   CFRD is associated with a decline in nutritional status, lung function, and early mortality. [@moranClinicalCareGuidelines2010]

## Advancements in Treatment

- Median predicted survival increased from 33 to 44 years between 2003-2018.[@cysticfibrosisfoundationCysticFibrosisFoundation2018]
- We expect to see even greater gains with the approval of new highly effective *CFTR* modulator therapies. [@Bessonova731; @middletonElexacaftorTezacaftorIvacaftor2019]
- However, despite these advances CFRD remains prevalent and there have been few large-scale studies examining changes in the epidemiology of CFRD over time.[@moranClinicalCareGuidelines2010; @marshallEpidemiologyCysticFibrosisrelated2005; @adlerHyperglycemiaDeathCystic2011; @chamnanDiabetesDeterminantMortality2010]

# How Have CFRD Risk Factors Changed With the Introduction of Highly Effective *CFTR* Modulators?

## Known Risk Factors {.smaller}

- Earlier CFRD studies [@marshallEpidemiologyCysticFibrosisrelated2005; @moranCysticFibrosisRelated2009] found positive associations between CFRD risk and:
  - Age
  - Female sex
  - Race/ethnicity from marginalized groups
  - BMI
  - Genotype
  - Pancreatic insufficiency
  - Nutritional status
  - Liver disease
  - Worse pulmonary disease 
  - Use of corticosteroids

## Known Risk Factors (cont.)
- Other studies have identified additional potential risk factors, but these are the most well-established.
- These variables were also readily available in the United States Cystic Fibrosis Foundation Patient Registry (CFFPR) and have been collected consistently since 2003.
- Very few model selection issues (yay!)

# The Fun Part!

## Participant Demographics in 2018 {.smaller}

```{r}
#| message: false
load("./Data_Clean/survival_dataset.RData")
annual %>%
  filter(ReviewYear == 2018) %>%
  select(
    CFRD, Age_YrEnd, Gender, Race, Hispanicrace, bmi_perc,
    bmi_cat
  ) %>%
  tbl_summary(
    by = CFRD, missing_text = "Missing",
    statistic = list(c("Age_YrEnd", "bmi_perc") ~ "{mean} ({sd})")
  ) %>%
  add_p(test = list(c("Age_YrEnd", "bmi_perc") ~ "t.test")) %>%
  add_overall() %>%
  separate_p_footnotes()
```

## Participant Demographics (cont.) {.smaller}

```{r}
annual %>%
  filter(ReviewYear == 2018) %>%
  select(
    CFRD, MutClass, tube_feeding, liver_disease,
    A_FEV1_pct_predicted, NumPulmExacerbation,
    A_corticosteroids1, A_IsOnEnzymes, A_Vx770, A_VX809comb
  ) %>%
  tbl_summary(
    by = CFRD, missing_text = "Missing",
    type = list(A_corticosteroids1 ~ "categorical"),
    statistic = list(c("A_FEV1_pct_predicted") ~ "{mean} ({sd})")
  ) %>%
  add_p(test = list(c("A_FEV1_pct_predicted") ~ "t.test")) %>%
  add_overall() %>%
  separate_p_footnotes()
```

## Kaplan-Meier Plot

```{r}
# Origin for survival analyses
og <- ymd("2003-01-01")
# Need to use a dataframe not a tibble
annual <- as.data.frame(annual)
annual$tstart <- as.numeric(annual$tstart) - as.numeric(og)
annual$tstop <- as.numeric(annual$tstop) - as.numeric(og)
# KM plot
fit <- survfit(Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2) ~ age_group,
  data = annual
)
# Visualize with survminer
ggsurvplot(fit, data = annual, risk.table = TRUE, palette = "simpsons")
```

## Special Thanks To

- The Cystic Fibrosis Foundation for the use of CF Foundation Patient Registry data.  
- The patients, care providers, and clinic coordinators at CF centers.
- Kristen Miller, MS

## References

### 2025-06-12 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
added one more Na:Cl value

@@
 chco$Na_Left[826] <- 72.4
+chco$Na_Right[790] <- 21.1
+chco$Na_Left[790] <- 24.1
 # For "LESS THAN 10", replace with half the limit of detection
@@
 df$K <- coalesce(df$K, df$K_mean)
+# Change Cl under 10 to 9 per CHEC
+df$Cl[df$Cl < 10] <- 9
+df$Na[df$Na < 10] <- 9
 # Calculate Na:Cl ratio
@@
   file = "./Data_Cleaned/sweat_variability_analysis_dataset.csv",
+  row.names = F, na = ""
+)
+# Additional lists for Angela
+write.csv(
+  df %>%
+    filter(Diagnosis == "Non-CF" & Cl >= 60),
+  file = "./Data_Cleaned/sweat_variability_non_cf_cl_60.csv",
+  row.names = F, na = ""
+)
+write.csv(
+  df %>%
+    filter(Diagnosis == "Cystic Fibrosis" & `Na:Cl` >= 1.2),
+  file = "./Data_Cleaned/sweat_variability_cf_nacl_1_2.csv",
+  row.names = F, na = ""
+)
+write.csv(
+  df %>%
+    filter(Diagnosis == "Non-CF" & `Na:Cl` < 1.2),
+  file = "./Data_Cleaned/sweat_variability_non_cf_nacl_1_2.csv",
   row.names = F, na = ""
@@
 - The multiclass AUC for Na:Cl is `r round(as.numeric(auc(mc_roc)),3)`.
-- The AUC for using Na:Cl to distinguish CF from non-CF was `r round(as.numeric(auc(mc_roc$rocs[[2]])),3)`
-- The best threshold for distinguishing CF from non-CF (based on the Youden index) was `r round(roc_table["Cystic Fibrosis vs.
-Non-CF","threshold"],3)`, and provided a sensitivity of `r round(roc_table["Cystic Fibrosis vs.
-Non-CF","sensitivity"],3)` and specificity of `r round(roc_table["Cystic Fibrosis vs.
-Non-CF","specificity"],3)`
-
-- The AUC for using Na:Cl to distinguish CF from CFTR-RD or CRMS was `r round(as.numeric(auc(mc_roc$rocs[[1]])),3)`
-- The best threshold for distinguishing CF from CFTR-RD or CRMS (based on the Youden index) was `r round(roc_table["CFTR-RD or CRMS vs.
-Cystic Fibrosis","threshold"],3)`, and provided a sensitivity of `r round(roc_table["CFTR-RD or CRMS vs.
-Cystic Fibrosis","sensitivity"],3)` and specificity of `r round(roc_table["CFTR-RD or CRMS vs.
-Cystic Fibrosis","specificity"],3)`
-
-- The AUC for using Na:Cl to distinguish non-CF from CFTR-RD or CRMS was `r round(as.numeric(auc(mc_roc$rocs[[3]])),3)`
-- The best threshold for distinguishing CF from CFTR-RD or CRMS (based on the Youden index) was `r round(roc_table["CFTR-RD or CRMS vs.
-Non-CF","threshold"],3)`, and provided a sensitivity of `r round(roc_table["CFTR-RD or CRMS vs.
-Non-CF","sensitivity"],3)` and specificity of `r round(roc_table["CFTR-RD or CRMS vs.
-Non-CF","specificity"],3)`

### 2025-06-13 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
WNAR presentation mostly done

@@
 #     (is.na(A_VX809comb) | A_VX809comb == "No"))
-# Write cleaned data
-save(annual,file = "./Data_Clean/survival_dataset.RData")
 ```
@@
 annual$tstop <- as.numeric(annual$tstop) - as.numeric(og)
+# Write cleaned data
+save(annual,file = "./Data_Clean/survival_dataset.RData")
 # Formula for base model
@@
 # Diagnostics
-test.ph <- data.frame(cox.zph(cox_res)$table)
-test.ph <- test.ph[rownames(test.ph) != "GLOBAL", ]
-# Find which variables technically violate PH assumption
-time_varying <- rownames(test.ph)[test.ph$p < 0.05]
+cox.zph(cox_res)
 # Just split data by time period so we're only fitting 3 models
@@
 ```{r results='asis'}
-f_vx770 <- update(f, . ~ . + A_Vx770)
+f_vx770 <- update(f, . ~ . + A_Vx770-A_corticosteroids1-NumPulmExacerbation-age_group)
 vx770_cox_res <- coxph(f_vx770,
@@
 ```{r}
-f_vx809comb <- update(f, . ~ . + A_VX809comb)
+f_vx809comb <- update(f, . ~ . + A_VX809comb-A_corticosteroids1-NumPulmExacerbation-age_group-Hispanicrace)
 vx809comb_cox_res <- coxph(f_vx809comb,
@@
 )
+
 ```

### 2025-06-13 Christine Chan/CF Registry/wnar_presentation.qmd
WNAR presentation mostly done

@@
   revealjs:
+    scrollable: true
     embed-resources: true
-editor: visual
+editor: source
 bibliography: wnar_references.bib
@@
 library(gtsummary)
+library(ggstats)
 knitr::opts_knit$set(root.dir = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry")
@@
 -   The CFTR protein is an ion channel that regulates electrolyte balance across cell membranes.[@stoltzOriginsCysticFibrosis2015]
--   As of 2023, over 2000 CFTR variants have been described. Of these, we are certain that over 700 are disease-causing.[@grasemannCysticFibrosis2023]
+-   As of 2023, over 2000 *CFTR* variants have been described. Of these, we are certain that over 700 are disease-causing.[@grasemannCysticFibrosis2023]
 -   The *CFTR* gene is expressed in the upper and lower airways, **pancreas**, intestine, biliary tree, cervix, sweat glands, and other organ systems.[@grasemannCysticFibrosis2023]
@@
 ## Known Risk Factors (cont.)
+
 -   Other studies have identified additional potential risk factors, but these are the most well-established.
@@
 
-## Participant Demographics in 2018 {.smaller}
+## Participant Demographics in 2018 {.scrolling}
 
@@
     CFRD, Age_YrEnd, Gender, Race, Hispanicrace, bmi_perc,
-    bmi_cat
-  ) %>%
-  tbl_summary(
-    by = CFRD, missing_text = "Missing",
-    statistic = list(c("Age_YrEnd", "bmi_perc") ~ "{mean} ({sd})")
-  ) %>%
-  add_p(test = list(c("Age_YrEnd", "bmi_perc") ~ "t.test")) %>%
-  add_overall() %>%
-  separate_p_footnotes()
-```
-
-## Participant Demographics (cont.) {.smaller}
-
-```{r}
-annual %>%
-  filter(ReviewYear == 2018) %>%
-  select(
-    CFRD, MutClass, tube_feeding, liver_disease,
-    A_FEV1_pct_predicted, NumPulmExacerbation,
+    bmi_cat, MutClass, tube_feeding, liver_disease,
+    A_FEV1_pct_predicted, A_FVC_pct_predicted, NumPulmExacerbation,
     A_corticosteroids1, A_IsOnEnzymes, A_Vx770, A_VX809comb
@@
     type = list(A_corticosteroids1 ~ "categorical"),
-    statistic = list(c("A_FEV1_pct_predicted") ~ "{mean} ({sd})")
+    statistic = list(c(
+      "Age_YrEnd", "bmi_perc", "A_FEV1_pct_predicted",
+      "A_FVC_pct_predicted"
+    ) ~ "{mean} ({sd})")
   ) %>%
-  add_p(test = list(c("A_FEV1_pct_predicted") ~ "t.test")) %>%
+  add_p(test = list(c(
+    "Age_YrEnd", "bmi_perc", "A_FEV1_pct_predicted",
+    "A_FVC_pct_predicted"
+  ) ~ "t.test")) %>%
   add_overall() %>%
@@
 ```{r}
-# Origin for survival analyses
-og <- ymd("2003-01-01")
-# Need to use a dataframe not a tibble
-annual <- as.data.frame(annual)
-annual$tstart <- as.numeric(annual$tstart) - as.numeric(og)
-annual$tstop <- as.numeric(annual$tstop) - as.numeric(og)
+#| fig-width: 16
+#| fig-height: 9
 # KM plot
-fit <- survfit(Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2) ~ age_group,
-  data = annual
+fit <- survfit(
+  Surv(time = tstop, event = as.numeric(CFRD) == 2) ~ age_group,
+  data = annual %>%
+    group_by(eDWID) %>%
+    slice_max(ReviewYear)
 )
 # Visualize with survminer
-ggsurvplot(fit, data = annual, risk.table = TRUE, palette = "simpsons")
+ggsurvplot(fit,
+  data = annual, risk.table = T, cumevents = T, palette = "simpsons",
+  conf.int = T
+)
 ```
+
+## Data Setup
+
+```{r}
+xvars <- c(
+  "age_group", "Gender", "Race", "Hispanicrace", "bmi_cat", "MutClass",
+  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
+  "NumPulmExacerbation", "A_corticosteroids1"
+)
+annual %>%
+  select(eDWID, ReviewYear, tstart, tstop, all_of(xvars), CFRD) %>%
+  head(20) %>%
+  gt::gt()
+```
+
+## Data Setup Notes {.smaller}
+
+-   Because we don't have an exact date of CFRD diagnosis, the only information we have is that the diagnosis occurred between January 1st of one year and January 1st of the next.
+    -   In other words, data are interval censored.
+-   Intervals are on a discrete grid (i.e. start and stop times are the same for all participants in a given year).
+-   Data include time-varying covariates such as age group, BMI category, etc.
+-   We categorized variables such as BMI and FEV1 mostly for the sake of interpretability for a clinical audience.
+    -   For BMI in particular though, we were not sure that the effect would be linear and categorization allowed for more flexible models.
+
+## Our First Model {.scrolling}
+
+```{r}
+#| fig-width: 12
+#| fig-height: 9
+# Formula for base model
+y <- "Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2)"
+f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
+cox_res <- coxph(f, data = annual, cluster = eDWID)
+ggcoef_model(cox_res,
+  show_p_values = F, add_reference_rows = T, exponentiate = T,
+  signif_stars = F, colour = F
+)
+```
+
+## Proportional Hazards Assumption
+
+```{r}
+cox.zph(cox_res)
+```
+
+## Proportional Hazards Assumption
+
+```{r}
+plot(cox.zph(cox_res)[2], resid = F)
+```
+
+## Proportional Hazards Assumption
+
+```{r}
+plot(cox.zph(cox_res)[7], resid = F)
+```
+
+## Time Dependent Covariates and Coefficients
+
+-   The hazard function for the Cox proportional hazards model with time dependent coefficients is[@therneauUsingTimeDependent2024]:
+
+$$
+\lambda(t)=\lambda_0(t)e^{\beta(t)X}
+$$
+
+-   Where the proportional hazard assumption is[@therneauUsingTimeDependent2024]
+
+$$
+\beta(t)=\beta
+$$
+
+## Time Dependent Covariates and Coefficients
+
+HOW MUCH MATH DO I NEED HERE?
+
+-   Coefficients vs. covariates?
+-   Internal vs. external covariates?
+
+# An Approach That Did Not Work
+
+-
+
+## Eras of CFTR Modulators
+
+-   We split our data into three different eras: 2003-2008, 2009-2013, and 2014-2018.
+    -   Ivacaftor (VX-770) was approved January 2012 and lumacaftor/ivacaftor (combination VX-809) was approved July 2015.
+    -   Three eras of equal lengths of time, the first with no modulators, second with one modulator, and the third with combination therapy.
+
+## Eras of CFTR Modulators
+
+```{r}
+annual$time_period <- cut(annual$ReviewYear,
+  breaks = c(-Inf, 2009, 2014, Inf), right = F
+)
+era1 <- coxph(f,
+  data = annual[as.numeric(annual$time_period) == 1, ],
+  cluster = eDWID
+)
+t1 <- era1 %>% tbl_regression(exponentiate = T)
+era2 <- coxph(f,
+  data = annual[as.numeric(annual$time_period) == 2, ],
+  cluster = eDWID
+)
+t2 <- era2 %>% tbl_regression(exponentiate = T)
+era3 <- coxph(f,
+  data = annual[as.numeric(annual$time_period) == 3, ],
+  cluster = eDWID
+)
+t3 <- era3 %>% tbl_regression(exponentiate = T)
+tbl_merge(
+  tbls = list(t1, t2, t3),
+  tab_spanner = c("2003 - 2008", "2009 - 2013", "2014 - 2018")
+)
+```
+
+## Eras of CFTR Modulators
+
+```{r}
+#| fig-width: 12
+#| fig-height: 9
+comb_plot <- ggcoef_compare(
+  list("2003 - 2008" = era1, "2009 - 2013" = era2, "2014 - 2018" = era3),
+  add_reference_rows = T, exponentiate = T, shape_guide = F,colour=F,
+  type = "faceted"
+)
+comb_plot
+```
+
+# The Effect of *CFTR* Modulators
+
+-   Is this even worth including? From a statistics point of view it's not that interesting.
+
+## Ivacaftor
+
+```{r}
+#| fig-width: 12
+#| fig-height: 9
+f_vx770 <- update(f, . ~ . + A_Vx770 - A_corticosteroids1 -
+  NumPulmExacerbation - age_group - Hispanicrace)
+vx770_cox_res <- coxph(f_vx770,
+  data = annual[annual$ReviewYear >= 2011, ],
+  cluster = eDWID
+)
+ggcoef_model(vx770_cox_res,
+  show_p_values = F, add_reference_rows = T, exponentiate = T,
+  signif_stars = F, colour = F
+)
+```
+
+## Lumacaftor/ivacaftor
+
+```{r}
+#| fig-width: 12
+#| fig-height: 9
+f_vx809comb <- update(f, . ~ . + A_VX809comb - A_corticosteroids1 -
+  NumPulmExacerbation - age_group - Hispanicrace)
+vx809comb_cox_res <- coxph(f_vx809comb,
+  data = annual[annual$ReviewYear >= 2015, ],
+  cluster = eDWID
+)
+ggcoef_model(vx809comb_cox_res,
+  show_p_values = F, add_reference_rows = T, exponentiate = T,
+  signif_stars = F, colour = F
+)
+```
+
+## Conclusions
+
+-   The middle age group is no longer at increased risk compared to the youngest.
+-   Female and Black people with CF remain at increased risk of developing CFRD.
+-   People with CF who are in the "undernourished" BMI category are no longer at increased risk, but those who are overweight are.
+-   Other risk factors appear relatively stable over time.
+-   Ivcaftor may reduce the risk of CFRD, although combination therapy does not appear to have an effect.
+
+## Limitations and Future Directions
+
+-   Methods available for interval censored data in the `survival` R package[@therneauPackageSurvivalAnalysis2024] are slightly limited.
+    -   Other packages such as `icenReg`[@anderson-bergmanIcenRegRegressionModels2017] may be able to fit more flexible models.
+-   Categorizing variables is not always ideal, and there is some concern about loss of information.
+    -   Currently we are working on extending this analysis using joint modeling for more flexibility.
+-   Data end in 2018 but Trikafta (a triple combination therapy of Elexacaftor, Ivacaftor, and Tezacaftor) was approved in 2019.
+    -   Need to study later generations of CFTR modulators.
+-   Registry data has natural limitations because each center must track and report their own data.
 
@@
 
-- The Cystic Fibrosis Foundation for the use of CF Foundation Patient Registry data.  
+-   The Cystic Fibrosis Foundation for the use of CF Foundation Patient Registry data.\
 -   The patients, care providers, and clinic coordinators at CF centers.

### 2025-06-13 Christine Chan/CF-SuGAR/power_calculations.qmd
CF-SuGAR FEV1 power calculations

@@
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan",
-  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan"
+  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/tim/OneDrive/Vigers/CF"
 )
 github_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Documents/GitHub",
-  "Linux" = "/home/timvigers/Documents/GitHub"
+  "Darwin" = "/Users/tim/Documents/GitHub",
+  "Windows" = "C:/Users/Tim/Documents/GitHub",
+  "Linux" = "/home/tim/Documents/GitHub"
 )
@@
 ```
+
+# Change in FEV1
+
+```{r}
+encounters <- read.csv("C:/Users/vigerst/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry/Data_Raw/CFF19_encountersMerged_Del1.csv")
+# Exclude T1D and T2D
+encounters <- encounters %>%
+  filter(!cfrd_status %in% c(4, 5))
+# Find each person's first year on modulators
+encounters$Vx770[is.na(encounters$Vx770)] <- 0
+encounters$Vx809comb[is.na(encounters$Vx809comb)] <- 0
+encounters <- encounters %>%
+  arrange(eDWID, reviewyear, encounternum) %>%
+  group_by(eDWID) %>%
+  mutate(
+    modulator = factor(Vx770 == 1 | Vx809comb == 1,
+      levels = c(F, T),
+      labels = c("Not on Modulator", "On Modulator")
+    ),
+    first_modulator =
+      first(reviewyear[modulator == "On Modulator"], na_rm = T)
+  )
+# Limit to two years prior and one year post-modulator
+encounters <- encounters %>%
+  filter(
+    !is.na(first_modulator),
+    reviewyear >= first_modulator - 2 & reviewyear <= first_modulator + 1
+  )
+# Limit dataset to people who were seen regularly but not too often
+# (likely hospitalized), make a new encounter number
+year_picker <- encounters %>%
+  group_by(eDWID, reviewyear) %>%
+  summarise(n = n()) %>%
+  filter(n >= 3 & n <= 10) %>%
+  ungroup() %>%
+  mutate(id = paste(eDWID, reviewyear, sep = "_"))
+encounters <- encounters %>%
+  filter(paste(eDWID, reviewyear, sep = "_") %in% year_picker$id) %>%
+  group_by(eDWID) %>%
+  mutate(visit = row_number()) %>%
+  select(
+    eDWID, reviewyear, encounternum, visit, modulator,
+    GLI_FEV1_pct_predicted
+  )
+# Plot
+# ggplot(encounters, aes(x = visit, y = GLI_FEV1_pct_predicted, group = eDWID)) +
+#   geom_line(alpha = 0.5)
+# ggplot(encounters, aes(x = visit, y = GLI_FEV1_pct_predicted, group = eDWID)) +
+#   geom_point(alpha = 0.1) +
+#   geom_smooth(aes(color = modulator, group = modulator))
+# Write for SAS models
+write.csv(encounters, file = "C:/Users/vigerst/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF-SuGAR/Background/Power Calculations/Data/encounters.csv",row.names = F,na = "")
+```

### 2025-06-13 Christine Chan/EnVision CF/cfrd_roc.qmd
EnVision ROC CIs

@@
   "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/EnVision CF",
+  "Windows" = "C:/Users/vigerst/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/EnVision CF",
   "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/EnVision CF"
@@
 ```{r data cleaning}
+set.seed(1017)
 # Import cleaned dataset
@@
 plot(r, print.auc = T, print.thres = T, legacy.axes = T)
+ci.auc(r)
+gt(data.frame(ci.coords(r, "best",
+  ret = c(c("sensitivity", "specificity", "fp", "tp", "fn", "tn"))
+))) %>% fmt_number(decimals = 3)
 gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)

### 2025-06-15 Christine Chan/CF Registry/WNAR/wnar_presentation.qmd
added a few WNAR slides

--- 2025-06-13 Christine Chan/CF Registry/wnar_presentation.qmd
@@
 
+# What do we do when the proportional hazards assumption is violated?
+
+## Two good options
+
+1. Joint modeling of longitudinal and survival data.[@hendersonJointModellingLongitudinal2000]
+  - The gold standard, do this if you can.
+  - Complex, requires somewhat specialized knowledge.
+  - Ask Elizabeth for help with this.
+
+2. Time-dependent Cox models
+  - More appropriate for external covariates (measures that vary independent of the failure time).
+  - Or baseline covariates where the effect changes over time.[@zhangTimevaryingCovariatesCoefficients2018]
+
 ## Time Dependent Covariates and Coefficients
@@
 $$
-\lambda(t)=\lambda_0(t)e^{\beta(t)X}
+\lambda(t)=\lambda_0(t)exp(\beta(t)X)
 $$
@@
 
-HOW MUCH MATH DO I NEED HERE?
+- Another way to write this model:
 
--   Coefficients vs. covariates?
--   Internal vs. external covariates?
+$$
+\lambda(t)=\lambda_0(t)exp(g(\beta,t)X)
+$$
 
-# An Approach That Did Not Work
+- Where $g(\beta,t)$ is a function of time that we can define.[@zhangTimevaryingCovariatesCoefficients2018]
 
--
+## Time as a Step Function
+
+```{r}
+data(cancer)
+vet2 <- survSplit(Surv(time, status) ~ .,
+  data = cancer, cut = c(90, 180), episode = "tgroup", id = "id"
+)
+vfit2 <- coxph(Surv(tstart, time, status) ~ age + sex + ph.ecog +
+  pat.karno:strata(tgroup), data = vet2)
+vfit2
+```
+
+# Applying This to Our CFRD Data
+
+## The Code
+
+```{r}
+annual_split <- survSplit(Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2) ~ .,
+  data = annual, cut = c(2192, 4018), episode = "tgroup"
+)
+cox_res2 <- coxph(
+  Surv(time = tstart, time2 = tstop, event) ~
+    age_group:strata(tgroup) + Gender + Race + Hispanicrace + bmi_cat + 
+    MutClass:strata(tgroup) + A_IsOnEnzymes:strata(tgroup) + 
+    tube_feeding:strata(tgroup) + liver_disease:strata(tgroup) + 
+    ppFEV1_cat:strata(tgroup) + NumPulmExacerbation:strata(tgroup) + 
+    A_corticosteroids1:strata(tgroup),
+  data = annual_split, cluster = eDWID
+)
+```
+
+## The Results
+
+```{r}
+cox_res2 %>%
+  tbl_regression(exponentiate = T)
+```
+
+## The Interpretation
+
+![](homer.jpg)
 
@@
   list("2003 - 2008" = era1, "2009 - 2013" = era2, "2014 - 2018" = era3),
-  add_reference_rows = T, exponentiate = T, shape_guide = F,colour=F,
+  add_reference_rows = T, exponentiate = T, shape_guide = F, colour = F,
   type = "faceted"

### 2025-06-15 Christine Chan/CF Registry/WNAR/wnar_presentation.qmd
WNAR slides done

@@
 title: "The Changing Landscape of Cystic Fibrosis Related Diabetes Risk Factors Between 2003-2018"
-author: "Tim Vigers"
+author:
+  - name: Tim Vigers
+    orcid: 0000-0001-7087-8570
+    email: timothy.vigers@cuanschutz.edu
+    affiliations:
+      - name: University of Colorado Denver, Anschutz Medical Campus
 date: "June 17, 2025"
@@
 
-- Median predicted survival increased from 33 to 44 years between 2003-2018.[@cysticfibrosisfoundationCysticFibrosisFoundation2018]
+- *CFTR* modulator therapies have completely transformed CF care.
+- Median predicted survival increased from 34 years in 2004 to 46 years in 2019.[@cysticfibrosisfoundationCysticFibrosisFoundation2024]
 - We expect to see even greater gains with the approval of new highly effective *CFTR* modulator therapies. [@Bessonova731; @middletonElexacaftorTezacaftorIvacaftor2019]
@@
 
-## Known Risk Factors (cont.)
+## The CFF-PR
 
-- Other studies have identified additional potential risk factors, but these are the most well-established.
 - These variables were also readily available in the United States Cystic Fibrosis Foundation Patient Registry (CFFPR) and have been collected consistently since 2003.
+- The CFF-PR is national database that aims to collect information on everyone with CF who receive care in CF Foundation-accredited care centers.
+
+## The CFF-PR
+
+- In 2023 there were 33,288 people in the registry (about 95%).
+- The study P.I. wanted to use annualized data for this analysis.
+  - Easier to get permission for this because there are fewer PHI concerns than encounter-level data.
 - Very few model selection issues (yay!)
@@
 
-## Participant Demographics in 2018 {.scrolling}
+## New Diagnoses in 2018 {.scrolling}
 
@@
   ) %>%
-  add_p(test = list(c(
-    "Age_YrEnd", "bmi_perc", "A_FEV1_pct_predicted",
-    "A_FVC_pct_predicted"
-  ) ~ "t.test")) %>%
-  add_overall() %>%
-  separate_p_footnotes()
+  add_overall()
 ```
@@
 ## Kaplan-Meier Plot
+
+Every person in the CFF-PR from 2003-2018
 
@@
   conf.int = T
-)
+) + xlab("Days From 2003-01-01")
 ```
+
+## Data Setup {.smaller}
+
+- Because we don't have an exact date of CFRD diagnosis, the only information we have is that the diagnosis occurred between January 1st of one year and January 1st of the next.
+  - Participants had to have two consecutive rows with CFRD, but first instance was considered onset.
+- Data are interval censored.
+- Intervals are on a discrete grid (i.e. start and stop times are the same for all participants in a given year).
+- Data include time-varying covariates such as age group, BMI category, etc.
+- We categorized variables such as BMI and FEV1 mostly for the sake of interpretability for a clinical audience.
+  - For BMI in particular though, we were not sure that the effect would be linear and categorization allowed for more flexible models.
 
@@
 ```
-
-## Data Setup Notes {.smaller}
-
-- Because we don't have an exact date of CFRD diagnosis, the only information we have is that the diagnosis occurred between January 1st of one year and January 1st of the next.
-  - In other words, data are interval censored.
-- Intervals are on a discrete grid (i.e. start and stop times are the same for all participants in a given year).
-- Data include time-varying covariates such as age group, BMI category, etc.
-- We categorized variables such as BMI and FEV1 mostly for the sake of interpretability for a clinical audience.
-  - For BMI in particular though, we were not sure that the effect would be linear and categorization allowed for more flexible models.
 
@@
 
-## Two good options
+## Two Good Options
 
-1. Joint modeling of longitudinal and survival data.[@hendersonJointModellingLongitudinal2000]
+- Joint modeling of longitudinal and survival data.[@hendersonJointModellingLongitudinal2000]
   - The gold standard, do this if you can.
@@
 
-2. Time-dependent Cox models
+- Time-dependent Cox models
   - More appropriate for external covariates (measures that vary independent of the failure time).
@@
 
-## Time Dependent Covariates and Coefficients
+## Time Dependent Cox Model
 
@@
 
-## Time Dependent Covariates and Coefficients
+## Time Dependent Coefficients
 
@@
 ```{r}
+#| echo: true
 data(cancer)
@@
 ```{r}
-annual_split <- survSplit(Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2) ~ .,
+#| echo: true
+annual_split <- survSplit(
+  Surv(
+    time = tstart, time2 = tstop,
+    event = as.numeric(CFRD) == 2
+  ) ~ .,
   data = annual, cut = c(2192, 4018), episode = "tgroup"
@@
 
+# An Alternative Approach
+
 ## Eras of CFTR Modulators
@@
   - Three eras of equal lengths of time, the first with no modulators, second with one modulator, and the third with combination therapy.
+
+## Eras of CFTR Modulators
+
+- Per the authors of the `survival` package, likelihood equations use only one copy of any subject, and the program picks out the correct row of data at each time.[@therneauUsingTimeDependent2024]
+  - This is why we don't need to worry about correlated data even though subjects have multiple observations.
+  - Not losing much information by splitting our data up.
 
@@
 
-- Is this even worth including? From a statistics point of view it's not that interesting.
-
 ## Ivacaftor
@@
 - People with CF who are in the "undernourished" BMI category are no longer at increased risk, but those who are overweight are.
+
+## Conclusions
+
 - Other risk factors appear relatively stable over time.
 - Ivcaftor may reduce the risk of CFRD, although combination therapy does not appear to have an effect.
+- Time-varying Cox models strike a nice balance between clinical interpretability and statistical rigor.
+- Everyone should take a survival class!
 
@@
   - Currently we are working on extending this analysis using joint modeling for more flexibility.
+    - We will be able to use encounter-level data for this.
+
+## Limitations and Future Directions
+
 - Data end in 2018 but Trikafta (a triple combination therapy of Elexacaftor, Ivacaftor, and Tezacaftor) was approved in 2019.
@@
 
-## Special Thanks To
+## Special Thanks
 
-- The Cystic Fibrosis Foundation for the use of CF Foundation Patient Registry data.\
+- The Cystic Fibrosis Foundation for the use of CF Foundation Patient Registry data.
 - The patients, care providers, and clinic coordinators at CF centers.
+- 2025 WNAR/IMS
+- Elizabeth Juarez-Colunga, PhD
 - Kristen Miller, MS

### 2025-06-17 Christine Chan/CF Registry/WNAR/wnar_presentation.qmd
removed survival class part
final WNAR changes

@@
 - The CFTR protein is an ion channel that regulates electrolyte balance across cell membranes.[@stoltzOriginsCysticFibrosis2015]
-- As of 2023, over 2000 *CFTR* variants have been described. Of these, we are certain that over 700 are disease-causing.[@grasemannCysticFibrosis2023]
-- The *CFTR* gene is expressed in the upper and lower airways, **pancreas**, intestine, biliary tree, cervix, sweat glands, and other organ systems.[@grasemannCysticFibrosis2023]
+- CF affects approximately 40,000 people in the US and 105,000 people have been diagnosed worldwide.
+- The *CFTR* gene is expressed in the upper and lower airways, pancreas, intestine, biliary tree, cervix, sweat glands, and other organ systems.[@grasemannCysticFibrosis2023]
 
@@
 - Because *CFTR* is highly expressed in the pancreas, *CFTR* dysfunction commonly leads to pancreatic insufficiency.[@moheetNewConceptsPathogenesis2022]
-- The pathophysiology of CFRD is not fully understood, but we think that $\beta$-cell loss related to pancreas destruction is one of the major drivers of diabetes in people with CF.[@moheetNewConceptsPathogenesis2022]
+- The pathophysiology of CFRD is not fully understood, but we think that $\beta$-cell loss related to pancreas destruction is one of the major drivers of diabetes in people with CF (pwCF).[@moheetNewConceptsPathogenesis2022]
 - CFRD is the most common non-pulmonary complication in CF. A diagnosis before the age of 10 is unusual, but 50% of adults over 40 have CFRD.[@moheetNewConceptsPathogenesis2022]
@@
 - These variables were also readily available in the United States Cystic Fibrosis Foundation Patient Registry (CFFPR) and have been collected consistently since 2003.
-- The CFF-PR is national database that aims to collect information on everyone with CF who receive care in CF Foundation-accredited care centers.
+- The CFF-PR is national database that aims to collect information on everyone with CF who receives care in CF Foundation-accredited care centers.
 
@@
 
-- In 2023 there were 33,288 people in the registry (about 95%).
+- In 2023 there were 33,288 people in the registry (between 85-95% of pwCF in the US).
 - The study P.I. wanted to use annualized data for this analysis.
   - Easier to get permission for this because there are fewer PHI concerns than encounter-level data.
-- Very few model selection issues (yay!)
+- No model selection necessary (yay!)
 
@@
 - Data are interval censored.
-- Intervals are on a discrete grid (i.e. start and stop times are the same for all participants in a given year).
 - Data include time-varying covariates such as age group, BMI category, etc.
@@
 
-## Conclusions
+## Conclusions {.smaller}
 
-- The middle age group is no longer at increased risk compared to the youngest.
-- Female and Black people with CF remain at increased risk of developing CFRD.
-- People with CF who are in the "undernourished" BMI category are no longer at increased risk, but those who are overweight are.
-
-## Conclusions
-
-- Other risk factors appear relatively stable over time.
-- Ivcaftor may reduce the risk of CFRD, although combination therapy does not appear to have an effect.
 - Time-varying Cox models strike a nice balance between clinical interpretability and statistical rigor.
-- Everyone should take a survival class!
+- Some risk factors are indeed changing over time, and new treatments may reduce the risk of CFRD.
+  - These changes should be accounted for when building CFRD prediction models in the modulator era.
+  - May also impact our work on updating the CFRD screening process.
+- Easy to code and fit these models.
+  - No more work than a couple logistic models, but you're still accounting for interval censoring, etc.
 
@@
   - Other packages such as `icenReg`[@anderson-bergmanIcenRegRegressionModels2017] may be able to fit more flexible models.
-- Categorizing variables is not always ideal, and there is some concern about loss of information.
+- Categorizing variables is not always ideal, and there is always some concern about loss of information.
 - Currently we are working on extending this analysis using joint modeling for more flexibility.
@@
 
-- The Cystic Fibrosis Foundation for the use of CF Foundation Patient Registry data.
+- The Cystic Fibrosis Foundation for funding some of this work and the use of CF Foundation Patient Registry data.
 - The patients, care providers, and clinic coordinators at CF centers.
-- 2025 WNAR/IMS
+- The organizers of WNAR/IMS 2025
 - Elizabeth Juarez-Colunga, PhD

### 2025-06-25 Christine Chan/EnVision CF/cfrd_roc.qmd
EnVision ROCs limited to first visit

--- 2025-06-09 Christine Chan/EnVision CF/cfrd_roc.qmd
@@
 # Limit to those with CGM data (based on mean glucose) and OGTT diagnosis
-df <- df %>% filter(!is.na(mean_sensor))
+df <- df %>% filter(!is.na(mean_sensor),!is.na(Diagnosis))
 # Convert CV units to match Scully et al.
 df$cv <- df$cv * 100
+# First CGM and OGTT pair only
+df <- df %>%
+  group_by(study_id) %>%
+  slice_min(Date)
 ```
-
-- **All tables and analyses below treat CGM wears as independent**, but we will need to think about how to handle this in future analyses.
 
@@
 - BMI percentile was calculated with age = 20 for adults.
-
-```{r}
-df <- df %>% filter(!is.na(Diagnosis))
-```
 
@@
 plot(r, print.auc = T, print.thres = T, legacy.axes = T)
+ci.auc(r)
+gt(data.frame(ci.coords(r, "best",
+  best.policy = "random",
+  ret = c(c("sensitivity", "specificity", "fp", "tp", "fn", "tn"))
+))) %>% fmt_number(decimals = 3)
 gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)

### 2025-06-25 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
GCT manual flowchart instead
GCT absolute value of days between OGTT and GCT

@@
   as.numeric(as.Date(df$OGTT.Date.Time) - as.Date(df$Study.visit.PFTs.Date.))
+df$`Days From OGTT to GCT (Absolute Value)` <- abs(df$`Days From OGTT to GCT`)
 df$OGTT.Date.Time <- NULL
@@
   kwt(`Days From OGTT to GCT`, "N", "Nmiss2", "medianq1q3", "range") +
+  kwt(
+    `Days From OGTT to GCT (Absolute Value)`,
+    "N", "Nmiss2", "medianq1q3", "range"
+  ) +
   fe(Diagnosis) + fe(Dysglycemia) + fe(Diabetes) + fe(CFRD)
@@
 )
-```
-
-# Flowchart
-
-```{r}
-flow_df <- tibble(
-  from = c(
-    "Initial Screen with GCT",
-    "Initial Screen with GCT", 
-    "GCTpl >= 147",
-    "GCTpl < 147"
-  ),
-  to = c(
-    "GCTpl >= 147", 
-    "GCTpl < 147", 
-    "Oral Glucose Tolerance Test", 
-    "Repeat Screen in 1 Year"
-  )
-)
-
-ggflowchart(flow_df)
 ```
@@
 ```
-
-# Nomogram
-
-## CFRD
-
-### `GCT 0 min capillary` and HbA1c
-
-```{r}
-f <- lrm(CFRD ~ `GCT 0 min capillary` + HbA1c,
-  data = df
-)
-ddist <- datadist(df)
-options(datadist = "ddist")
-nom <- nomogram(f, fun = plogis, funlabel = "Risk of CFRD", lp = F)
-plot(nom)
-```
-
-Here’s how to use a nomogram.
-Pick a point on the GCT 0 min capillary and HbA1c scales and get the associated points for each.
-Add the points together to get the associated probability based on the Total Points line.
-
-For example:
-
-- GCT 0 min capillary = 130 (40 points)
-- HbA1c = 6.6 (10 points)
-
-Add up the points:
-
-40 + 10 = 50 Total Points
-
-50 Total Points lines up at about 0.08, so this person has approximately an 8% chance of CFRD. The process is the same for the nomograms below.
-
-### `GCT 0 min capillary`, HbA1c, and `GCT 60 min plasma`
-
-```{r}
-f <- lrm(CFRD ~ `GCT 0 min capillary` + HbA1c + `GCT 60 min plasma`,
-  data = df
-)
-ddist <- datadist(df)
-options(datadist = "ddist")
-nom <- nomogram(f, fun = plogis, funlabel = "Risk of CFRD", lp = F)
-plot(nom)
-```
-
-### `GCT 0 min capillary`, and `GCT 60 min plasma`
-
-```{r}
-f <- lrm(CFRD ~ `GCT 0 min capillary` + `GCT 60 min plasma`,
-  data = df
-)
-ddist <- datadist(df)
-options(datadist = "ddist")
-nom <- nomogram(f, fun = plogis, funlabel = "Risk of CFRD", lp = F)
-plot(nom)
-```
-
-# Decision trees
-
-## Predicting CFRD
-
-```{r}
-tree <- rpart(CFRD ~ `GCT 0 min capillary` + HbA1c + `GCT 60 min plasma`,
-  data = df
-)
-rpart.plot(tree, type = 0)
-```
-
-We just don't have enough data to do this properly for CFRD.
-
-## Predicting dysglycemia
-
-```{r}
-tree <- rpart(Dysglycemia ~ `GCT 0 min capillary` + `GCT 60 min plasma`, data = df)
-rpart.plot(tree, type = 0)
-```

### 2025-07-01 Christine Chan/EnVision CF/fsOGTT_pre_post_ETI.qmd
write results and data to .csv
cleaned up EnVision folder

@@
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF",
+  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF",
   "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/EnVision CF"
@@
 df$pre_post <- factor(df$pre_post, levels = c("Pre", "Post"))
+# Write for checking
+write.csv(df,
+  file = "./Data_Clean/fsOGTT_pre_post_ETI.csv",
+  row.names = F, na = ""
+)
 ```

### 2025-07-01 Edith Zemanick/Sweat Variability/sweat_test_analysis_chec.qmd
cleaned up EnVision folder

@@
 library(patchwork)
+library(ggeffects)
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
+  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
   "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Variability"
@@
 # Only include people with either a pre-modulator or post-modulator sodium
-chec <- chec %>% filter(!is.na(swsavg))
+chec <- chec %>% filter(!is.na(swsavg), site != "005 Buffalo, NY CC and Peds")
 # Calculate ratios
@@
   drop_na()
-m <- lmer(cl ~ pre_post * na + (1 | subject), data = model_df)
 # From Amalia
@@
 )
+# Tim's model
+tm <- lmer(cl ~ pre_post * na + (1 | subject), data = model_df)
 ```
@@
 The model using the ratio alone does not really work due to the fitted probabilities issue.
+
+## Tim's model
+
+```{r}
+tm %>% gtsummary::tbl_regression()
+plot(predict_response(tm, terms = c("na", "pre_post")))
+```
 

### 2025-07-01 Christine Chan/EnVision CF/cfrd_roc.qmd
simplified ROC for Zach

--- 2025-06-09 Christine Chan/EnVision CF/cfrd_roc.qmd
@@
 # Limit to those with CGM data (based on mean glucose) and OGTT diagnosis
-df <- df %>% filter(!is.na(mean_sensor))
+df <- df %>% filter(!is.na(mean_sensor), !is.na(Diagnosis))
 # Convert CV units to match Scully et al.
 df$cv <- df$cv * 100
+# First CGM and OGTT pair only
+df <- df %>%
+  group_by(study_id) %>%
+  slice_min(Date)
 ```
-
-- **All tables and analyses below treat CGM wears as independent**, but we will need to think about how to handle this in future analyses.
 
@@
 
-```{r}
-df <- df %>% filter(!is.na(Diagnosis))
-```
-
 ## By OGTT categories
@@
 t1 <- tableby(
-  Diagnosis ~ redcap_data_access_group + sex + origin_race + ethnicity +
-    pancreatic_status + CFTR + fev1 + fvc + bmi_perc + a1c_result,
+  Diagnosis ~ redcap_data_access_group + age_visit + height + weight +
+    bmi_perc + sex + origin_race + ethnicity + pancreatic_status + CFTR + fev1 +
+    fvc + a1c_result + corrector_yes_no,
   data = df
@@
 t2 <- tableby(
-  Dysglycemia ~ redcap_data_access_group + sex + origin_race + ethnicity +
-    pancreatic_status + CFTR + fev1 + fvc + bmi_perc + a1c_result,
+  Dysglycemia ~ redcap_data_access_group + age_visit + height + weight +
+    bmi_perc + sex + origin_race + ethnicity + pancreatic_status + CFTR + fev1 +
+    fvc + a1c_result + corrector_yes_no,
   data = df
@@
 t3 <- tableby(
-  CFRD ~ redcap_data_access_group + sex + origin_race + ethnicity +
-    pancreatic_status + CFTR + fev1 + fvc + bmi_perc + a1c_result,
+  CFRD ~ redcap_data_access_group + age_visit + height + weight +
+    bmi_perc + sex + origin_race + ethnicity + pancreatic_status + CFTR + fev1 +
+    fvc + a1c_result + corrector_yes_no,
   data = df
@@
 plot(r, print.auc = T, print.thres = T, legacy.axes = T)
-gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
+ci.auc(r)
+# gt(data.frame(ci.coords(r, "best",
+#   best.policy = "random",
+#   ret = c(c("sensitivity", "specificity", "fp", "tp", "fn", "tn"))
+# ))) %>% fmt_number(decimals = 3)
+gt(coords(r, "best", ret = "all")) %>% fmt_number(decimals = 3)
 ```

### 2025-07-01 Lilah Melzer/cough_monitor_analysis.qmd
cough monitor updates for Lilah

--- 2025-03-06 Lilah Melzer/cough_monitor_analysis.qmd
@@
 library(gt)
+library(ggbreak)
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor",
+  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor",
   "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Lilah Melzer/Lilah Cough Monitor"
@@
   ),
-  labels = c("CF", "CF", "Non-CF", "Non-CF")
-)
-df$Diagnosis <- relevel(df$Diagnosis, ref = "Non-CF")
+  labels = c("CF", "CF", "Healthy Control", "Healthy Control")
+)
+df$Diagnosis <- relevel(df$Diagnosis, ref = "Healthy Control")
 demo <- df %>%
@@
     nih_sex:nih_ethnicity, mutation1, mutation2,
-    cftr_mod, cftr_mod_type, mucolytic:antih
+    cftr_mod, cftr_mod_type, mucolytic:antih, pulm_ex, contains("micro_type___")
   )
@@
 #| tbl-cap: Participant characteristics
-t1 <- tableby(Diagnosis ~ nih_sex + age_at_consent + height_perc + weight_perc +
+t1 <- tableby(
+  Diagnosis ~ nih_sex + age_at_consent + height_perc + weight_perc +
     nih_race + nih_ethnicity + cftr_mod + cftr_mod_type + Genotype + pre_fev_pp +
-  total_sleep, data = demo, control = mycontrols)
+    total_sleep + pulm_ex + micro_type___1 + micro_type___2 + micro_type___3 +
+    micro_type___4,
+  data = demo, control = mycontrols
+)
 summary(t1, pfootnote = T)
@@
     `Median Coughs per Night` = median(Coughs, na.rm = T),
+    `Prop. Nights With Cough` = sum(Coughs > 0, na.rm = T) / n(),
     .groups = "drop"
@@
   Diagnosis ~ `Nights` + `Nights With Data` + `Prop. Nights With Data` +
-    `Median Coughs per Night`,
+    `Prop. Nights With Cough` + `Median Coughs per Night`,
   data = per_person, control = mycontrols
@@
 #| warning: false
-hist <- ggplot(agg_counts, aes(x = Coughs)) +
-  geom_histogram() +
+hist1 <- ggplot(
+  agg_counts[agg_counts$Diagnosis == "Healthy Control", ],
+  aes(x = Coughs)
+) +
+  geom_histogram(binwidth = 1) +
+  ylab("Nights") +
+  ggtitle("Healthy Control") +
   theme_bw() +
-  theme(axis.title.y = element_blank())
-hist
+  scale_x_break(c(80, 80), ticklabels = 80, space = 0, scales = 0.0000001)
+hist2 <- ggplot(agg_counts[agg_counts$Diagnosis == "CF", ], aes(x = Coughs)) +
+  geom_histogram(binwidth = 1) +
+  ylab("Nights") +
+  ggtitle("CF") +
+  theme_bw() +
+  scale_x_break(c(85, 260), ticklabels = c(260, 269), scales = 0.25)
+hist <- hist1 + hist2 + plot_layout(guides = "collect")
 ggsave(paste0(home_dir, "/Dissemination/Figures/histogram.png"),
@@
 fig2 <- ggplot(agg_counts[!is.na(agg_counts$Coughs), ], aes(
-  x = Night, y = Coughs, group = factor(sid), color = Diagnosis
+  x = Night, y = Coughs, group = factor(sid)
 )) +
   geom_point(alpha = 0.1) +
-  geom_line(alpha = 0.1) +
-  geom_smooth(aes(group = Diagnosis)) +
-  theme_bw()
+  geom_line(alpha = 0.2, aes(linetype = Diagnosis)) +
+  geom_smooth(color = "black", aes(group = Diagnosis, linetype = Diagnosis)) +
+  annotate("rect", xmin = 1, xmax = 29, ymin = 1, ymax = 275, alpha = 0.1) +
+  annotate("rect", xmin = 41, xmax = 76, ymin = 1, ymax = 275, alpha = 0.1) +
+  ylim(c(1, 275)) +
+  xlim(c(0, 100)) +
+  theme_bw() +
+  scale_y_break(c(80, 265), scales = 0.25, ticklabels = c(265, 270, 275))
 fig2

### 2025-07-07 Christine Chan/EnVision CF/envision_analyses.qmd
combining EnVision Analyses into 1 doc

--- 2025-07-01 Christine Chan/EnVision CF/cfrd_roc.qmd
@@
 ---
-title: "Using CGM to Predict CFRD in the Envision Cohort"
+title: "EnVision CF Analyses"
 author: "Tim Vigers"
@@
 #| include: false
+library(Hmisc)
 library(tidyverse)
@@
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF",
-  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/EnVision CF",
-  "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/EnVision CF"
+  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
+  "Linux" = "/home/tim/OneDrive/Vigers/CF"
 )
@@
 ```{r data cleaning}
+# Set seed for bootstrap
 set.seed(1017)
-# Import cleaned dataset
-load("./Data_Clean/envision_analysis_dataset.RData")
-df <- final_df
+# Import data
+# See GitHub/CF/Christine Chan/EnVision CF/clean_envision_data.R
+load("/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF/Data_Clean/envision_analysis_dataset.RData")
+```
+
+# ROC Analyses
+
+```{r}
 # Limit to those with CGM data (based on mean glucose) and OGTT diagnosis
-df <- df %>% filter(!is.na(mean_sensor), !is.na(Diagnosis))
+df <- final_df %>% filter(!is.na(mean_sensor), !is.na(Diagnosis))
 # Convert CV units to match Scully et al.
@@
 
-# Participant demographics
-
-- In the cleaned EnVision dataset we have a total of `r nrow(df)` CGMs from `r length(unique(df$study_id))` unique individuals.
+For the following analyses, we used ROC curves to assess various CGM metrics' ability to distinguish between CFRD- and CFRD+ people with CF.
+Because some people had multiple CGM wears and OGTTs, we decided to use each participant's first CGM and OGTT pair.
+In the cleaned EnVision dataset we have a total of `r nrow(df)` CGMs from `r length(unique(df$study_id))` unique individuals.
 Of these CGM wears, `r sum(is.na(df$Diagnosis))` were missing a diagnosis due to missing OGTT timepoints.
 Those without a diagnosis were excluded from all analyses.
-- BMI percentile was calculated with age = 20 for adults.
 
-## By OGTT categories
-
-### Detailed OGTT category
+## Participant Demographics by CFRD status
 
@@
 t1 <- tableby(
-  Diagnosis ~ redcap_data_access_group + age_visit + height + weight +
-    bmi_perc + sex + origin_race + ethnicity + pancreatic_status + CFTR + fev1 +
-    fvc + a1c_result + corrector_yes_no,
+  CFRD ~ fe(redcap_data_access_group) +
+    kwt(age_visit, "Nmiss", "medianq1q3", "range") +
+    kwt(height, "Nmiss", "medianq1q3", "range") +
+    kwt(weight, "Nmiss", "medianq1q3", "range") + bmi_perc + sex +
+    fe(origin_race) + fe(ethnicity) + fe(pancreatic_status) + fe(CFTR) +
+    corrector_yes_no + fev1 + fvc + a1c_result,
   data = df
@@
 
-### General OGTT category
-
-```{r results='asis'}
-t2 <- tableby(
-  Dysglycemia ~ redcap_data_access_group + age_visit + height + weight +
-    bmi_perc + sex + origin_race + ethnicity + pancreatic_status + CFTR + fev1 +
-    fvc + a1c_result + corrector_yes_no,
-  data = df
-)
-summary(t2, pfootnote = T)
-```
-
-### CFRD status
-
-```{r results='asis'}
-t3 <- tableby(
-  CFRD ~ redcap_data_access_group + age_visit + height + weight +
-    bmi_perc + sex + origin_race + ethnicity + pancreatic_status + CFTR + fev1 +
-    fvc + a1c_result + corrector_yes_no,
-  data = df
-)
-summary(t3, pfootnote = T)
-```
-
-# ROC analyses
-
-## Mean sensor glucose
+## Mean Sensor Glucose
 
@@
 
-## Percent time spent >140 mg/dL
+## Percent Time >140 mg/dL
 
@@
 
-## Percent time spent >180 mg/dL
+## Percent Time >180 mg/dL
 
@@
 
-# Pulmonary function comparisons
-
-## By OGTT categories
-
-### Detailed OGTT category
-
-```{r results='asis'}
-t1 <- tableby(Diagnosis ~ fev1 + fvc, data = df)
-summary(t1, pfootnote = T)
-```
-
-### General OGTT category
-
-```{r results='asis'}
-t2 <- tableby(Dysglycemia ~ fev1 + fvc, data = df)
-summary(t2, pfootnote = T)
-```
-
-### CFRD status
-
-```{r results='asis'}
-t3 <- tableby(CFRD ~ fev1 + fvc, data = df)
-summary(t3, pfootnote = T)
-```
-
-## By CGM category
+## Pulmonary Function Comparisons by CGM
 
@@
 
-# Figure
+## Figure
 
@@
 #| message: false
+#| warning: false
+#| fig-height: 12
+#| fig-width: 16
 # Dotplot
@@
   theme(
-    legend.position = "none",
+    legend.title = element_blank(),
     axis.text.x = element_text(angle = 45, hjust = 1),

### 2025-07-07 Christine Chan/EnVision CF/envision_analyses.qmd
added Amir's analysis to unified Envision analysis

@@
 library(patchwork)
+library(lmerTest)
+library(emmeans)
 home_dir <- switch(Sys.info()["sysname"],
@@
 ```
+
+# Hypoglycemia Comparisons
+
+```{r}
+# Timepoint and variables lists
+timepoints <- c(0, 10, 30, 60, 90, 120, 150, 180)
+limited_timepoints <- c(0, 120, 150, 180)
+glucoses <- paste0("Glucose_", timepoints)
+insulins <- paste0("Insulin_", timepoints)
+cpeptides <- paste0("C.Peptide_", timepoints)
+epinephrines <- paste0("Epinephrine_", timepoints)
+norepinephrines <- paste0("Norepinephrine_", timepoints)
+adren_survey <- paste0("adren_score_", limited_timepoints)
+neuro_survey <- paste0("neuro_score_", limited_timepoints)
+total_survey <- paste0("total_score_", limited_timepoints)
+num_symptoms_survey <- paste0("num_symptoms_", limited_timepoints)
+hypo_surveys <- c(adren_survey, neuro_survey, total_survey, num_symptoms_survey)
+# First visit dataframe
+first_visit_df <- final_df %>%
+  group_by(study_id) %>%
+  filter(row_number() == 1) %>%
+  ungroup()
+# Long data for mixed models
+glmm_df <- final_df %>%
+  select(
+    study_id, Hypo60, Hypo70,
+    all_of(c(
+      glucoses, insulins, cpeptides, epinephrines, norepinephrines,
+      hypo_surveys
+    ))
+  ) %>%
+  pivot_longer(
+    all_of(c(
+      glucoses, insulins, cpeptides, epinephrines, norepinephrines,
+      hypo_surveys
+    )),
+    names_to = c(".value", "Timepoint"),
+    names_pattern = "(.*)_([0-9]{1,3})"
+  )
+glmm_df$Timepoint <- factor(glmm_df$Timepoint, levels = timepoints)
+```
+
+## Demographics at First Visit
+
+### By Hypoglycemia <70 mg/dL at Any Point
+
+```{r results='asis'}
+t1 <- tableby(
+  Hypo70 ~
+    fe(Diagnosis) + redcap_data_access_group +
+    kwt(age_visit, "Nmiss", "medianq1q3", "range") +
+    kwt(height, "Nmiss", "medianq1q3", "range") +
+    kwt(weight, "Nmiss", "medianq1q3", "range") + bmi_perc + sex +
+    fe(origin_race) + fe(ethnicity) + fe(pancreatic_status) + fe(CFTR) +
+    corrector_yes_no + fev1 + fvc + a1c_result,
+  data = first_visit_df, control = tableby.control(simulate.p.value = T)
+)
+summary(t1, pfootnote = T)
+```
+
+### By Hypoglycemia <60 mg/dL at Any Point
+
+```{r results='asis'}
+t2 <- tableby(
+  Hypo60 ~
+    fe(Diagnosis) + redcap_data_access_group +
+    kwt(age_visit, "Nmiss", "medianq1q3", "range") +
+    kwt(height, "Nmiss", "medianq1q3", "range") +
+    kwt(weight, "Nmiss", "medianq1q3", "range") + bmi_perc + sex +
+    fe(origin_race) + fe(ethnicity) + fe(pancreatic_status) + fe(CFTR) +
+    corrector_yes_no + fev1 + fvc + a1c_result,
+  data = first_visit_df, control = tableby.control(simulate.p.value = T)
+)
+summary(t2, pfootnote = T)
+```
+
+## Glucose
+
+Glucose was modeled using a linear mixed effect model with random intercept for participant and time treated as a categorical variable.
+The results from this model are presented as estimated marginal means (also known as least-squares means) for those with hypoglycemia at some point and those with no hypoglycemia.
+
+### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(Glucose ~ Timepoint * Hypo70 + (1 | study_id),
+  data = glmm_df
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean Glucose") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean glucose at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour glucose was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour glucose was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(Glucose ~ Timepoint * Hypo60 + (1 | study_id),
+  data = glmm_df
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean Glucose") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean glucose at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour glucose was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour glucose was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+## Insulin
+
+Insulin was modeled using a linear mixed effect model with random intercept for participant and time treated as a categorical variable.
+The results from this model are presented as estimated marginal means (also known as least-squares means) for those with hypoglycemia at some point and those with no hypoglycemia.
+
+### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(Insulin ~ Timepoint * Hypo70 + (1 | study_id),
+  data = glmm_df
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean Insulin") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean insulin at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour insulin was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour insulin was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(Insulin ~ Timepoint * Hypo60 + (1 | study_id),
+  data = glmm_df
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean Insulin") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean insulin at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour insulin was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour insulin was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+## C-peptide
+
+C-peptide was modeled using a linear mixed effect model with random intercept for participant and time treated as a categorical variable.
+The results from this model are presented as estimated marginal means (also known as least-squares means) for those with hypoglycemia at some point and those with no hypoglycemia.
+
+### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(C.Peptide ~ Timepoint * Hypo70 + (1 | study_id),
+  data = glmm_df
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean C-peptide") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean C-peptide at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour C-peptide was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour C-peptide was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(C.Peptide ~ Timepoint * Hypo60 + (1 | study_id),
+  data = glmm_df
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean C-peptide") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean C-peptide at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour C-peptide was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour C-peptide was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).

### 2025-07-08 Christine Chan/EnVision CF/clean_envision_data.R
Matsuda index for EnVision

@@ -475,6 +475,11 @@ final_df$iAUC180ins <- apply(final_df, 1, function(r) {
 })
 # HOMA IR
 final_df$homa_ir <- (final_df$Glucose_0 * final_df$Insulin_0) / 405
+# Matsuda
+final_df$matsuda <- 10000 /
+  (((final_df$Glucose_0 / 18) * final_df$Insulin_0) *
+    ((rowMeans(final_df[, paste0("Glucose_", c(0, 30, 60, 90, 120))], na.rm = T) / 18) *
+      rowMeans(final_df[, paste0("Insulin_", c(0, 30, 60, 90, 120))], na.rm = T)))
 # Check for hypoglycemia
 final_df$Hypo70 <-
   apply(final_df[, glucose], 1, function(r) {
@@ -572,5 +577,7 @@ label(final_df$fev1) <- "FEV1 % Predicted"
 label(final_df$fvc) <- "FVC % Predicted"
 label(final_df$a1c_result) <- "HbA1c (%)"
 label(final_df$corrector_yes_no) <- "CFTR Modulator Status"
+label(final_df$homa_ir) <- "HOMA-IR"
+label(final_df$matsuda) <- "Matsuda Index"
 # Write
 save(final_df, file = "./Christine Chan/EnVision CF/Data_Clean/envision_analysis_dataset.RData")

### 2025-07-09 Christine Chan/EnVision CF/envision_analyses.qmd
Matsuda index for EnVision
nested models for EnVision

@@
   select(
-    study_id, Hypo60, Hypo70,
+    study_id, Hypo60, Hypo70, ogtt_num,
     all_of(c(
@@
 
-## Glucose
+## OGTT Measures by Timepoint
+
+### Glucose
 
@@
 
-### Hypoglycemia <70 mg/dL at Any Point
+#### Hypoglycemia <70 mg/dL at Any Point
 
@@
 # Fit
-mod <- lmer(Glucose ~ Timepoint * Hypo70 + (1 | study_id),
+mod <- lmer(Glucose ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
   data = glmm_df
@@
 
-### Hypoglycemia <60 mg/dL at Any Point
+#### Hypoglycemia <60 mg/dL at Any Point
 
@@
 # Fit
-mod <- lmer(Glucose ~ Timepoint * Hypo60 + (1 | study_id),
+mod <- lmer(Glucose ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
   data = glmm_df
@@
 
-## Insulin
+### Insulin
 
@@
 # Fit
-mod <- lmer(Insulin ~ Timepoint * Hypo70 + (1 | study_id),
+mod <- lmer(Insulin ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
   data = glmm_df
@@
 # Fit
-mod <- lmer(Insulin ~ Timepoint * Hypo60 + (1 | study_id),
+mod <- lmer(Insulin ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
   data = glmm_df
@@
 
-## C-peptide
+### C-peptide
 
@@
 # Fit
-mod <- lmer(C.Peptide ~ Timepoint * Hypo70 + (1 | study_id),
+mod <- lmer(C.Peptide ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
   data = glmm_df
@@
 # Fit
-mod <- lmer(C.Peptide ~ Timepoint * Hypo60 + (1 | study_id),
+mod <- lmer(C.Peptide ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
   data = glmm_df
@@
 For example, for those with no hypoglycemia the mean two hour C-peptide was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour C-peptide was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+## Whole OGTT Measures
+
+### Matsuda index
+
+The Matsuda index was calculated across the 0, 30, 60, 90, and 120 minute timepoints using the formula: $$\frac{10000}{\sqrt{(\frac{\text{fasting glucose (mg/dL)}}{18} * \text{fasting insulin}) * (\frac{\text{mean glucose (mg/dL)}}{18} * \text{mean insulin})}}$$
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(matsuda ~ Hypo70 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the Matsuda index was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia groups, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(matsuda ~ Hypo60 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the Matsuda index was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia groups, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

### 2025-07-09 Christine Chan/EnVision CF/envision_analyses.qmd
EnVision basic outline done

@@
 cpeptides <- paste0("C.Peptide_", timepoints)
+glucagons <- paste0("Glucagon_", timepoints)
+glp1s <- paste0("GLP.1.Active_", timepoints)
+gips <- paste0("GIP_", timepoints)
 epinephrines <- paste0("Epinephrine_", timepoints)
@@
     all_of(c(
-      glucoses, insulins, cpeptides, epinephrines, norepinephrines,
-      hypo_surveys
+      glucoses, insulins, cpeptides, glucagons, glp1s, gips, epinephrines,
+      norepinephrines, hypo_surveys
     ))
@@
     all_of(c(
-      glucoses, insulins, cpeptides, epinephrines, norepinephrines,
-      hypo_surveys
+      glucoses, insulins, cpeptides, glucagons, glp1s, gips, epinephrines,
+      norepinephrines, hypo_surveys
     )),
@@
 
+Glucose, insulin, C-peptide, glucagon, epinephrine, and norepinephrine were modeled using a linear mixed effect model with random effects for OGTT within participant and time treated as a categorical variable.
+The results from this model are presented as estimated marginal means (also known as least-squares means) for those with hypoglycemia at some point and those with no hypoglycemia.
+
 ### Glucose
-
-Glucose was modeled using a linear mixed effect model with random intercept for participant and time treated as a categorical variable.
-The results from this model are presented as estimated marginal means (also known as least-squares means) for those with hypoglycemia at some point and those with no hypoglycemia.
 
@@
 
-Insulin was modeled using a linear mixed effect model with random intercept for participant and time treated as a categorical variable.
-The results from this model are presented as estimated marginal means (also known as least-squares means) for those with hypoglycemia at some point and those with no hypoglycemia.
-
 #### Hypoglycemia <70 mg/dL at Any Point
@@
 
-C-peptide was modeled using a linear mixed effect model with random intercept for participant and time treated as a categorical variable.
-The results from this model are presented as estimated marginal means (also known as least-squares means) for those with hypoglycemia at some point and those with no hypoglycemia.
-
 #### Hypoglycemia <70 mg/dL at Any Point
@@
 
-## Whole OGTT Measures
+### Glucagon
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(Glucagon ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+  data = glmm_df
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean Glucagon") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean glucagon at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour glucagon was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour glucagon was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(Glucagon ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+  data = glmm_df
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean Glucagon") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean glucagon at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour glucagon was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour glucagon was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+### GLP-1
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(GLP.1.Active ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+  data = glmm_df
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean GLP-1") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean GLP-1 at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour GLP-1 was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour GLP-1 was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(GLP.1.Active ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+  data = glmm_df
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean GLP-1") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean GLP-1 at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour GLP-1 was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour GLP-1 was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+### GIP
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(GIP ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+  data = glmm_df
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean GIP") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean GIP at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour GIP was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour GIP was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(GIP ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+  data = glmm_df
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean GIP") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean GIP at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour GIP was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour GIP was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+### Epinephrine
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(Epinephrine ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+  data = glmm_df
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean Epinephrine") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean epinephrine at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour epinephrine was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour epinephrine was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(Epinephrine ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+  data = glmm_df
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean Epinephrine") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean epinephrine at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour epinephrine was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour epinephrine was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+### Norepinephrine
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(Norepinephrine ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+  data = glmm_df
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean Norepinephrine") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean norepinephrine at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour norepinephrine was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour norepinephrine was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(Norepinephrine ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+  data = glmm_df
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean Norepinephrine") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean norepinephrine at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour norepinephrine was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour norepinephrine was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+### `adren_score`
+
+The `adren_score` variable was modeled using a mixed effect Poisson model with random effects for OGTT within participant and time treated as a categorical variable.
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+# Poisson model better than NB by AIC and BIC
+mod <- glmer(adren_score ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+  data = glmm_df, family = "poisson",
+  control = glmerControl(optimizer = "bobyqa")
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean adren_score") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean norepinephrine at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour norepinephrine was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour norepinephrine was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+# Poisson model better than NB by AIC and BIC
+mod <- glmer(adren_score ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+  data = glmm_df, family = "poisson",
+  control = glmerControl(optimizer = "bobyqa")
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean adren_score") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean norepinephrine at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour norepinephrine was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour norepinephrine was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+### `neuro_score`
+
+The `neuro_score` variable was modeled using a mixed effect Poisson model with random effects for OGTT within participant and time treated as a categorical variable.
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+# Poisson model better than NB by AIC and BIC
+mod <- glmer(neuro_score ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+  data = glmm_df, family = "poisson",
+  control = glmerControl(optimizer = "bobyqa")
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean neuro_score") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean norepinephrine at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour norepinephrine was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour norepinephrine was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+# Poisson model better than NB by AIC and BIC
+mod <- glmer(neuro_score ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+  data = glmm_df, family = "poisson",
+  control = glmerControl(optimizer = "bobyqa")
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean neuro_score") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean norepinephrine at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour norepinephrine was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour norepinephrine was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+### `total_score`
+
+The `total_score` variable was modeled using a mixed effect Poisson model with random effects for OGTT within participant and time treated as a categorical variable.
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+# Poisson model better than NB by AIC and BIC
+mod <- glmer(total_score ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+  data = glmm_df, family = "poisson",
+  control = glmerControl(optimizer = "bobyqa")
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean total_score") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean norepinephrine at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour norepinephrine was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour norepinephrine was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+# Poisson model better than NB by AIC and BIC
+mod <- glmer(total_score ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+  data = glmm_df, family = "poisson",
+  control = glmerControl(optimizer = "bobyqa")
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean total_score") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean norepinephrine at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour norepinephrine was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour norepinephrine was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+### `num_symptoms`
+
+The `num_symptoms` variable was modeled using a mixed effect Poisson model with random effects for OGTT within participant and time treated as a categorical variable.
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+# Poisson model better than NB by AIC and BIC
+mod <- glmer(num_symptoms ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+  data = glmm_df, family = "poisson",
+  control = glmerControl(optimizer = "bobyqa")
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+  group = Hypo70
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean num_symptoms") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean norepinephrine at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour norepinephrine was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour norepinephrine was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+# Poisson model better than NB by AIC and BIC
+mod <- glmer(num_symptoms ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+  data = glmm_df, family = "poisson",
+  control = glmerControl(optimizer = "bobyqa")
+)
+# EM means
+means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"),
+  type = "response",
+  pbkrtest.limit = 1e6
+))
+colnames(means) <- c(
+  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+)
+# Plot
+ggplot(means, aes(
+  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+  group = Hypo60
+)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    width = 3
+  ) +
+  geom_line() +
+  scale_x_continuous(breaks = timepoints) +
+  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
+  ylab("EM Mean num_symptoms") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean norepinephrine at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour norepinephrine was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour norepinephrine was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+## OGTT Summary Measures
+
+OGTT summary measures (i.e., measures calculated using multiple OGTT timepoints) were modeled using a linear mixed effect model with random intercept for participant.
 
@@
 
-The table above shows that the Matsuda index was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia groups, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+The table above shows that the Matsuda index was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
 
@@
 
-The table above shows that the Matsuda index was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia groups, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+The table above shows that the Matsuda index was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+### HOMA-IR
+
+HOMA-IR was calculated using the formula: $$\frac{\text{fasting glucose (mg/dL)}*\text{fasting insulin}}{405}$$
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(homa_ir ~ Hypo70 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the HOMA-IR was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(homa_ir ~ Hypo60 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the HOMA-IR was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+### iAUC 30 Minute Glucose
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(iAUC30gluc ~ Hypo70 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the iAUC 30 minute glucose was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(iAUC30gluc ~ Hypo60 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the iAUC 30 minute glucose was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+### iAUC 60 Minute Glucose
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(iAUC60gluc ~ Hypo70 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the iAUC 60 minute glucose was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(iAUC60gluc ~ Hypo60 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the iAUC 60 minute glucose was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+### iAUC 120 Minute Glucose
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(iAUC120gluc ~ Hypo70 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the iAUC 120 minute glucose was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(iAUC120gluc ~ Hypo60 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the iAUC 120 minute glucose was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+### iAUC 180 Minute Glucose
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(iAUC180gluc ~ Hypo70 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the iAUC 180 minute glucose was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(iAUC180gluc ~ Hypo60 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the iAUC 180 minute glucose was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+### iAUC 30 Minute Insulin
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(iAUC30ins ~ Hypo70 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the iAUC 30 minute insulin was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(iAUC30ins ~ Hypo60 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the iAUC 30 minute insulin was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+### iAUC 60 Minute Insulin
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(iAUC60ins ~ Hypo70 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the iAUC 60 minute insulin was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(iAUC60ins ~ Hypo60 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the iAUC 60 minute insulin was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+### iAUC 120 Minute Insulin
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(iAUC120ins ~ Hypo70 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the iAUC 120 minute insulin was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(iAUC120ins ~ Hypo60 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the iAUC 120 minute insulin was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+### iAUC 180 Minute Insulin
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(iAUC180ins ~ Hypo70 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the iAUC 180 minute insulin was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(iAUC180ins ~ Hypo60 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the iAUC 180 minute insulin was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+### ODI
+
+ODI was calculated as $\frac{\text{iAUC 30 Minute Insulin}}{\text{iAUC 30 Minute Glucose}}$.
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod1 <- lmer(odi ~ Hypo70 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the ODI was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(odi ~ Hypo60 + (1 | study_id), data = final_df)
+mod %>% gtsummary::tbl_regression()
+```
+
+The table above shows that the ODI was different by `r round(summary(mod)$coefficients[2,1],2)` on average in the hypoglycemia group, although this was not statistically significant (p = `r round(summary(mod)$coefficients[2,5],2)`).

### 2025-07-09 Christine Chan/EnVision CF/clean_envision_data.R

@@ -278,7 +278,7 @@ files <- list.files("./Christine Chan/EnVision CF/Data_Clean/hOGTTs",
 ogtts <- lapply(files, function(f) {
   # Import
   d <- read.csv(f,
-    na.strings = c("", "not detectable", "not detecable", "no serum")
+    na.strings = c("", "not detectable", "not detecable", "no serum", "#VALUE!")
   )
   # Get run date
   date <- sub("./Christine Chan/EnVision CF/Data_Clean/hOGTTs/summary_", "", f)
@@ -339,7 +339,10 @@ final_df <- final_df %>%
   unite(C.Peptide_150, C.Peptide_150.x, C.Peptide_150.y, na.rm = T) %>%
   unite(C.Peptide_180, C.Peptide_180.x, C.Peptide_180.y, na.rm = T)
 final_df <- final_df %>%
-  mutate(across(contains("C.Peptide_"), ~ as.numeric(.x)))
+  mutate(across(c(
+    contains("C.Peptide_"), contains("GLP.1.Active_"),
+    contains("GIP_"), contains("Glucagon_")
+  ), ~ as.numeric(.x)))
 # Add CGM
 final_df <- full_join(final_df, cgm)
 #-------------------------------------------------------------------------------
@@ -396,6 +399,7 @@ final_df$iAUC30gluc <- apply(final_df, 1, function(r) {
     return(auc)
   }
 })
+label(final_df$iAUC30gluc) <- "iAUC 30 Minute Glucose"
 final_df$iAUC60gluc <- apply(final_df, 1, function(r) {
   y <- as.numeric(r[glucose[1:4]]) - as.numeric(r[glucose[1]])
   if (is.na(first(y)) | is.na(last(y))) {
@@ -407,6 +411,7 @@ final_df$iAUC60gluc <- apply(final_df, 1, function(r) {
     return(auc)
   }
 })
+label(final_df$iAUC60gluc) <- "iAUC 60 Minute Glucose"
 final_df$iAUC120gluc <- apply(final_df, 1, function(r) {
   y <- as.numeric(r[glucose[1:6]]) - as.numeric(r[glucose[1]])
   if (is.na(first(y)) | is.na(last(y))) {
@@ -418,6 +423,7 @@ final_df$iAUC120gluc <- apply(final_df, 1, function(r) {
     return(auc)
   }
 })
[…]

### 2025-07-10 Christine Chan/EnVision CF/envision_analyses.qmd
need to fix GLP-1 models for EnVision

@@
 
-#### Hypoglycemia <70 mg/dL at Any Point
-
-```{r}
-# Fit
-mod <- lmer(Insulin ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+Insulin values were log-transformed to improve model fit.
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(log(Insulin) ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
   data = glmm_df
@@
 # Fit
-mod <- lmer(Insulin ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+mod <- lmer(log(Insulin) ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
   data = glmm_df
@@
 
-#### Hypoglycemia <70 mg/dL at Any Point
-
-```{r}
-# Fit
-mod <- lmer(C.Peptide ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+C-peptide values were log-transformed to improve model fit.
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(log(C.Peptide) ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
   data = glmm_df
@@
 # Fit
-mod <- lmer(C.Peptide ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+mod <- lmer(log(C.Peptide) ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
   data = glmm_df
@@
 
-#### Hypoglycemia <70 mg/dL at Any Point
-
-```{r}
-# Fit
-mod <- lmer(GLP.1.Active ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+GLP-1 models are not currently converging, but Tim is working on this.
+
+<!-- #### Hypoglycemia <70 mg/dL at Any Point -->
+
+<!-- ```{r} -->
+<!-- # Fit -->
+<!-- mod <- lmer(sqrt(GLP.1.Active) ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num), -->
+<!--   data = glmm_df -->
+<!-- ) -->
+<!-- # EM means -->
+<!-- means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"), -->
+<!--   type = "response", -->
+<!--   pbkrtest.limit = 1e6 -->
+<!-- )) -->
+<!-- colnames(means) <- c( -->
+<!--   "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit" -->
+<!-- ) -->
+<!-- # Plot -->
+<!-- ggplot(means, aes( -->
+<!--   x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70, -->
+<!--   group = Hypo70 -->
+<!-- )) + -->
+<!--   geom_point() + -->
+<!--   geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`), -->
+<!--     width = 3 -->
+<!--   ) + -->
+<!--   geom_line() + -->
+<!--   scale_x_continuous(breaks = timepoints) + -->
+<!--   scale_color_manual(values = c("#66C2A5", "#FC8D62")) + -->
+<!--   ylab("EM Mean GLP-1") + -->
+<!--   xlab("Timepoint") + -->
+<!--   theme_bw(base_size = 14) + -->
+<!--   theme(legend.title = element_blank()) -->
+<!-- ``` -->
+
+<!-- The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval. -->
+
+<!-- ```{r} -->
+<!-- means %>% -->
+<!--   gt() %>% -->
+<!--   fmt_number(decimals = 2) -->
+<!-- ``` -->
+
+<!-- The table above shows the estimated mean GLP-1 at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour GLP-1 was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour GLP-1 was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).
+-->
+
+<!-- #### Hypoglycemia <60 mg/dL at Any Point -->
+
+<!-- ```{r} -->
+<!-- # Fit -->
+<!-- mod <- lmer(GLP.1.Active ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num), -->
+<!--   data = glmm_df -->
+<!-- ) -->
+<!-- # EM means -->
+<!-- means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"), -->
+<!--   type = "response", -->
+<!--   pbkrtest.limit = 1e6 -->
+<!-- )) -->
+<!-- colnames(means) <- c( -->
+<!--   "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit" -->
+<!-- ) -->
+<!-- # Plot -->
+<!-- ggplot(means, aes( -->
+<!--   x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60, -->
+<!--   group = Hypo60 -->
+<!-- )) + -->
+<!--   geom_point() + -->
+<!--   geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`), -->
+<!--     width = 3 -->
+<!--   ) + -->
+<!--   geom_line() + -->
+<!--   scale_x_continuous(breaks = timepoints) + -->
+<!--   scale_color_manual(values = c("#66C2A5", "#FC8D62")) + -->
+<!--   ylab("EM Mean GLP-1") + -->
+<!--   xlab("Timepoint") + -->
+<!--   theme_bw(base_size = 14) + -->
+<!--   theme(legend.title = element_blank()) -->
+<!-- ``` -->
+
+<!-- The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval. -->
+
+<!-- ```{r} -->
+<!-- means %>% -->
+<!--   gt() %>% -->
+<!--   fmt_number(decimals = 2) -->
+<!-- ``` -->
+
+<!-- The table above shows the estimated mean GLP-1 at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour GLP-1 was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour GLP-1 was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).
+-->
+
+### GIP
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(GIP ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
   data = glmm_df
@@
   scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
-  ylab("EM Mean GLP-1") +
-  xlab("Timepoint") +
-  theme_bw(base_size = 14) +
-  theme(legend.title = element_blank())
-```
-
-The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
-
-```{r}
-means %>%
-  gt() %>%
-  fmt_number(decimals = 2)
-```
-
-The table above shows the estimated mean GLP-1 at each time point for the two groups.
-For example, for those with no hypoglycemia the mean two hour GLP-1 was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour GLP-1 was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).
-
-#### Hypoglycemia <60 mg/dL at Any Point
-
-```{r}
-# Fit
-mod <- lmer(GLP.1.Active ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+  ylab("EM Mean GIP") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean GIP at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour GIP was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour GIP was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(GIP ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
   data = glmm_df
@@
   scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
-  ylab("EM Mean GLP-1") +
-  xlab("Timepoint") +
-  theme_bw(base_size = 14) +
-  theme(legend.title = element_blank())
-```
-
-The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
-
-```{r}
-means %>%
-  gt() %>%
-  fmt_number(decimals = 2)
-```
-
-The table above shows the estimated mean GLP-1 at each time point for the two groups.
-For example, for those with no hypoglycemia the mean two hour GLP-1 was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour GLP-1 was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).
-
-### GIP
-
-#### Hypoglycemia <70 mg/dL at Any Point
-
-```{r}
-# Fit
-mod <- lmer(GIP ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+  ylab("EM Mean GIP") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean GIP at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour GIP was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour GIP was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+### Epinephrine
+
+Epinephrine values were log-transformed to improve model fit.
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(log(Epinephrine) ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
   data = glmm_df
@@
   scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
-  ylab("EM Mean GIP") +
-  xlab("Timepoint") +
-  theme_bw(base_size = 14) +
-  theme(legend.title = element_blank())
-```
-
-The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
-
-```{r}
-means %>%
-  gt() %>%
-  fmt_number(decimals = 2)
-```
-
-The table above shows the estimated mean GIP at each time point for the two groups.
-For example, for those with no hypoglycemia the mean two hour GIP was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour GIP was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).
-
-#### Hypoglycemia <60 mg/dL at Any Point
-
-```{r}
-# Fit
-mod <- lmer(GIP ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+  ylab("EM Mean Epinephrine") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean epinephrine at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour epinephrine was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour epinephrine was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+#### Hypoglycemia <60 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(log(Epinephrine) ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
   data = glmm_df
@@
   scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
-  ylab("EM Mean GIP") +
-  xlab("Timepoint") +
-  theme_bw(base_size = 14) +
-  theme(legend.title = element_blank())
-```
-
-The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
-
-```{r}
-means %>%
-  gt() %>%
-  fmt_number(decimals = 2)
-```
-
-The table above shows the estimated mean GIP at each time point for the two groups.
-For example, for those with no hypoglycemia the mean two hour GIP was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour GIP was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).
-
-### Epinephrine
-
-#### Hypoglycemia <70 mg/dL at Any Point
-
-```{r}
-# Fit
-mod <- lmer(Epinephrine ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+  ylab("EM Mean Epinephrine") +
+  xlab("Timepoint") +
+  theme_bw(base_size = 14) +
+  theme(legend.title = element_blank())
+```
+
+The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
+
+```{r}
+means %>%
+  gt() %>%
+  fmt_number(decimals = 2)
+```
+
+The table above shows the estimated mean epinephrine at each time point for the two groups.
+For example, for those with no hypoglycemia the mean two hour epinephrine was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour epinephrine was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).
+
+### Norepinephrine
+
+Norepinephrine values were log-transformed to improve model fit.
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(log(Norepinephrine) ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
   data = glmm_df
@@
   scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
-  ylab("EM Mean Epinephrine") +
-  xlab("Timepoint") +
-  theme_bw(base_size = 14) +
-  theme(legend.title = element_blank())
-```
-
-The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
-
-```{r}
-means %>%
-  gt() %>%
-  fmt_number(decimals = 2)
-```
-
-The table above shows the estimated mean epinephrine at each time point for the two groups.
-For example, for those with no hypoglycemia the mean two hour epinephrine was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour epinephrine was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).
-
-#### Hypoglycemia <60 mg/dL at Any Point
-
-```{r}
-# Fit
-mod <- lmer(Epinephrine ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
-  data = glmm_df
-)
-# EM means
-means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"),
-  type = "response",
-  pbkrtest.limit = 1e6
-))
-colnames(means) <- c(
-  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
-)
-# Plot
-ggplot(means, aes(
-  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
-  group = Hypo60
-)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
-    width = 3
-  ) +
-  geom_line() +
-  scale_x_continuous(breaks = timepoints) +
-  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
-  ylab("EM Mean Epinephrine") +
-  xlab("Timepoint") +
-  theme_bw(base_size = 14) +
-  theme(legend.title = element_blank())
-```
-
-The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval.
-
-```{r}
-means %>%
-  gt() %>%
-  fmt_number(decimals = 2)
-```
-
-The table above shows the estimated mean epinephrine at each time point for the two groups.
-For example, for those with no hypoglycemia the mean two hour epinephrine was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour epinephrine was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).
-
-### Norepinephrine
-
-#### Hypoglycemia <70 mg/dL at Any Point
-
-```{r}
-# Fit
-mod <- lmer(Norepinephrine ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
-  data = glmm_df
-)
-# EM means
-means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"),
-  type = "response",
-  pbkrtest.limit = 1e6
-))
-colnames(means) <- c(
-  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
-)
-# Plot
-ggplot(means, aes(
-  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
-  group = Hypo70
-)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
-    width = 3
-  ) +
-  geom_line() +
-  scale_x_continuous(breaks = timepoints) +
-  scale_color_manual(values = c("#66C2A5", "#FC8D62")) +
   ylab("EM Mean Norepinephrine") +
@@
 # Fit
-mod <- lmer(Norepinephrine ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+mod <- lmer(log(Norepinephrine) ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
   data = glmm_df
@@
 
-#### Hypoglycemia <70 mg/dL at Any Point
-
-```{r}
-# Fit
-mod <- lmer(matsuda ~ Hypo70 + (1 | study_id), data = final_df)
+Matsuda index values were log-transformed to improve model fit.
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(log(matsuda) ~ Hypo70 + (1 | study_id), data = final_df)
 mod %>% gtsummary::tbl_regression()
@@
 # Fit
-mod <- lmer(matsuda ~ Hypo60 + (1 | study_id), data = final_df)
+mod <- lmer(log(matsuda) ~ Hypo60 + (1 | study_id), data = final_df)
 mod %>% gtsummary::tbl_regression()
@@
 
-#### Hypoglycemia <70 mg/dL at Any Point
-
-```{r}
-# Fit
-mod <- lmer(homa_ir ~ Hypo70 + (1 | study_id), data = final_df)
+HOMA-IR values were log-transformed to improve model fit.
+
+#### Hypoglycemia <70 mg/dL at Any Point
+
+```{r}
+# Fit
+mod <- lmer(log(homa_ir) ~ Hypo70 + (1 | study_id), data = final_df)
 mod %>% gtsummary::tbl_regression()
@@
 # Fit
-mod1 <- lmer(odi ~ Hypo70 + (1 | study_id), data = final_df)
+mod <- lmer(odi ~ Hypo70 + (1 | study_id), data = final_df)
 mod %>% gtsummary::tbl_regression()

### 2025-07-10 Christine Chan/EnVision CF/envision_analyses.qmd
removed GLP1 completely (assay issues)

@@
 
-### GLP-1
-
-GLP-1 models are not currently converging, but Tim is working on this.
-
-<!-- #### Hypoglycemia <70 mg/dL at Any Point -->
-
-<!-- ```{r} -->
-<!-- # Fit -->
-<!-- mod <- lmer(sqrt(GLP.1.Active) ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num), -->
-<!--   data = glmm_df -->
-<!-- ) -->
-<!-- # EM means -->
-<!-- means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"), -->
-<!--   type = "response", -->
-<!--   pbkrtest.limit = 1e6 -->
-<!-- )) -->
-<!-- colnames(means) <- c( -->
-<!--   "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit" -->
-<!-- ) -->
-<!-- # Plot -->
-<!-- ggplot(means, aes( -->
-<!--   x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70, -->
-<!--   group = Hypo70 -->
-<!-- )) + -->
-<!--   geom_point() + -->
-<!--   geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`), -->
-<!--     width = 3 -->
-<!--   ) + -->
-<!--   geom_line() + -->
-<!--   scale_x_continuous(breaks = timepoints) + -->
-<!--   scale_color_manual(values = c("#66C2A5", "#FC8D62")) + -->
-<!--   ylab("EM Mean GLP-1") + -->
-<!--   xlab("Timepoint") + -->
-<!--   theme_bw(base_size = 14) + -->
-<!--   theme(legend.title = element_blank()) -->
-<!-- ``` -->
-
-<!-- The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval. -->
-
-<!-- ```{r} -->
-<!-- means %>% -->
-<!--   gt() %>% -->
-<!--   fmt_number(decimals = 2) -->
-<!-- ``` -->
-
-<!-- The table above shows the estimated mean GLP-1 at each time point for the two groups.
-For example, for those with no hypoglycemia the mean two hour GLP-1 was `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="No Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour GLP-1 was `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo70=="Hypoglycemia < 70" & means$Timepoint=="120","CI Upper Limit"],2)`).
--->
-
-<!-- #### Hypoglycemia <60 mg/dL at Any Point -->
-
-<!-- ```{r} -->
-<!-- # Fit -->
-<!-- mod <- lmer(GLP.1.Active ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num), -->
-<!--   data = glmm_df -->
-<!-- ) -->
-<!-- # EM means -->
-<!-- means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"), -->
-<!--   type = "response", -->
-<!--   pbkrtest.limit = 1e6 -->
-<!-- )) -->
-<!-- colnames(means) <- c( -->
-<!--   "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit" -->
-<!-- ) -->
-<!-- # Plot -->
-<!-- ggplot(means, aes( -->
-<!--   x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60, -->
-<!--   group = Hypo60 -->
-<!-- )) + -->
-<!--   geom_point() + -->
-<!--   geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`), -->
-<!--     width = 3 -->
-<!--   ) + -->
-<!--   geom_line() + -->
-<!--   scale_x_continuous(breaks = timepoints) + -->
-<!--   scale_color_manual(values = c("#66C2A5", "#FC8D62")) + -->
-<!--   ylab("EM Mean GLP-1") + -->
-<!--   xlab("Timepoint") + -->
-<!--   theme_bw(base_size = 14) + -->
-<!--   theme(legend.title = element_blank()) -->
-<!-- ``` -->
-
-<!-- The plot above shows the estimated mean at each time point. Error bars indicate the 95% confidence interval. -->
-
-<!-- ```{r} -->
-<!-- means %>% -->
-<!--   gt() %>% -->
-<!--   fmt_number(decimals = 2) -->
-<!-- ``` -->
-
-<!-- The table above shows the estimated mean GLP-1 at each time point for the two groups.
-For example, for those with no hypoglycemia the mean two hour GLP-1 was `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="No Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`) and for those with hypoglycemia at some point the mean two hour GLP-1 was `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","Mean"],2)` (95% CI: `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Lower Limit"],2)` - `r round(means[means$Hypo60=="Hypoglycemia < 60" & means$Timepoint=="120","CI Upper Limit"],2)`).
--->
-
 ### GIP

### 2025-07-11 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
nicer flowchart for race and ETI
race and ETI contrasts and exacerbations

@@
 ```{r flowchart}
-ggflowchart(flow_chart)
+ggflowchart(flow_chart, text_size = 3, y_nudge = 0.4, x_nudge = 0.45)
 ```
@@
     type = list(`Years With Data` ~ "continuous"),
-    statistic = list(c(gli_fev1_ppred_rn,gli_fvc_ppred_rn)~"{mean} [{sd}]"),
+    statistic = list(c(gli_fev1_ppred_rn, gli_fvc_ppred_rn) ~ "{mean} [{sd}]"),
     missing_text = "Missing"
@@
   fmt_number(decimals = 3)
-# Linear hypotheses
-
-white_vs_black_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceBlack or African American` = 1.13856555"
-)
-white_vs_black_pre <- tidy(summary(white_vs_black_pre))
-white_vs_black_pre$contrast <- "Black vs. White Pre-ETI Slope"
-white_vs_black_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_black_post <- glht(
-  change_mod,
-  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.63633"
-)
-white_vs_black_post <- tidy(summary(white_vs_black_post))
-white_vs_black_post$contrast <- "Black vs. White Post-ETI Slope"
-white_vs_black_post %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-
-white_vs_hispanic_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceHispanic or Latino` = 1.19547"
-)
-white_vs_hispanic_pre <- tidy(summary(white_vs_hispanic_pre))
-white_vs_hispanic_pre$contrast <- "Hispanic vs. White Pre-ETI Slope"
-white_vs_hispanic_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_hispanic_post <- glht(
-  change_mod,
-  "Years + `Years:RaceHispanic or Latino` + slope_change_years + `RaceHispanic or Latino:slope_change_years` = 2.738885"
-)
-white_vs_hispanic_post <- tidy(summary(white_vs_hispanic_post))
-white_vs_hispanic_post$contrast <- "Hispanic vs. White Post-ETI Slope"
-white_vs_hispanic_post %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-
-white_vs_other_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceOther` = 1.19547"
-)
-white_vs_other_pre <- tidy(summary(white_vs_other_pre))
-white_vs_other_pre$contrast <- "Other vs. White Pre-ETI Slope"
-white_vs_other_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_other_post <- glht(
-  change_mod,
-  "Years + `Years:RaceOther` + slope_change_years + `RaceOther:slope_change_years` = 2.738885"
-)
-white_vs_other_post <- tidy(summary(white_vs_other_post))
-white_vs_other_post$contrast <- "Other vs. White Post-ETI Slope"
-white_vs_other_post %>%
-  gt() %>%
+# Contrasts
+aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
+aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
+hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
+hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
+nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
+other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
+contr <- rbind(
+  "Black/AA Pre-ETI" = aa_pre,
+  "Black/AA Post-ETI" = aa_pre + aa_change,
+  "H/L Pre-ETI" = hl_pre,
+  "H/L Post-ETI" = hl_pre + hl_change,
+  "NHW Pre-ETI" = nhw_pre,
+  "NHW Post-ETI" = nhw_pre + nhw_change,
+  "Other Pre-ETI" = other_pre,
+  "Other Post-ETI" = other_pre + other_change
+)
+change_mod_slopes <- glht(change_mod, linfct = contr)
+change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
+change_mod_slopes_ci$confint %>%
+  as.data.frame() %>%
+  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
   fmt_number(decimals = 3)
@@
   fmt_number(decimals = 3)
-# Linear hypotheses
-white_vs_black_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceBlack or African American` = 1.195465"
-)
-white_vs_black_pre <- tidy(summary(white_vs_black_pre))
-white_vs_black_pre$contrast <- "Black vs. White Pre-ETI Slope"
-white_vs_black_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_black_post <- glht(
-  change_mod,
-  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.738885"
-)
-white_vs_black_post <- tidy(summary(white_vs_black_post))
-white_vs_black_post$contrast <- "Black vs. White Post-ETI Slope"
-white_vs_black_post %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-
-white_vs_hispanic_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceHispanic or Latino` = 1.19547"
-)
-white_vs_hispanic_pre <- tidy(summary(white_vs_hispanic_pre))
-white_vs_hispanic_pre$contrast <- "Hispanic vs. White Pre-ETI Slope"
-white_vs_hispanic_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_hispanic_post <- glht(
-  change_mod,
-  "Years + `Years:RaceHispanic or Latino` + slope_change_years + `RaceHispanic or Latino:slope_change_years` = 2.738885"
-)
-white_vs_hispanic_post <- tidy(summary(white_vs_hispanic_post))
-white_vs_hispanic_post$contrast <- "Hispanic vs. White Post-ETI Slope"
-white_vs_hispanic_post %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-
-white_vs_other_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceOther` = 1.19547"
-)
-white_vs_other_pre <- tidy(summary(white_vs_other_pre))
-white_vs_other_pre$contrast <- "Other vs. White Pre-ETI Slope"
-white_vs_other_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_other_post <- glht(
-  change_mod,
-  "Years + `Years:RaceOther` + slope_change_years + `RaceOther:slope_change_years` = 2.738885"
-)
-white_vs_other_post <- tidy(summary(white_vs_other_post))
-white_vs_other_post$contrast <- "Other vs. White Post-ETI Slope"
-white_vs_other_post %>%
-  gt() %>%
+# Contrasts
+aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
+aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
+hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
+hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
+nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
+other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
+contr <- rbind(
+  "Black/AA Pre-ETI" = aa_pre,
+  "Black/AA Post-ETI" = aa_pre + aa_change,
+  "H/L Pre-ETI" = hl_pre,
+  "H/L Post-ETI" = hl_pre + hl_change,
+  "NHW Pre-ETI" = nhw_pre,
+  "NHW Post-ETI" = nhw_pre + nhw_change,
+  "Other Pre-ETI" = other_pre,
+  "Other Post-ETI" = other_pre + other_change
+)
+change_mod_slopes <- glht(change_mod, linfct = contr)
+change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
+change_mod_slopes_ci$confint %>%
+  as.data.frame() %>%
+  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
   fmt_number(decimals = 3)
@@
   fmt_number(decimals = 3)
-# Linear hypotheses
-white_vs_black_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceBlack or African American` = 1.195465"
-)
-white_vs_black_pre <- tidy(summary(white_vs_black_pre))
-white_vs_black_pre$contrast <- "Black vs. White Pre-ETI Slope"
-white_vs_black_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_black_post <- glht(
-  change_mod,
-  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.738885"
-)
-white_vs_black_post <- tidy(summary(white_vs_black_post))
-white_vs_black_post$contrast <- "Black vs. White Post-ETI Slope"
-white_vs_black_post %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-
-white_vs_hispanic_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceHispanic or Latino` = 1.19547"
-)
-white_vs_hispanic_pre <- tidy(summary(white_vs_hispanic_pre))
-white_vs_hispanic_pre$contrast <- "Hispanic vs. White Pre-ETI Slope"
-white_vs_hispanic_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_hispanic_post <- glht(
-  change_mod,
-  "Years + `Years:RaceHispanic or Latino` + slope_change_years + `RaceHispanic or Latino:slope_change_years` = 2.738885"
-)
-white_vs_hispanic_post <- tidy(summary(white_vs_hispanic_post))
-white_vs_hispanic_post$contrast <- "Hispanic vs. White Post-ETI Slope"
-white_vs_hispanic_post %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-
-white_vs_other_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceOther` = 1.19547"
-)
-white_vs_other_pre <- tidy(summary(white_vs_other_pre))
-white_vs_other_pre$contrast <- "Other vs. White Pre-ETI Slope"
-white_vs_other_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_other_post <- glht(
-  change_mod,
-  "Years + `Years:RaceOther` + slope_change_years + `RaceOther:slope_change_years` = 2.738885"
-)
-white_vs_other_post <- tidy(summary(white_vs_other_post))
-white_vs_other_post$contrast <- "Other vs. White Post-ETI Slope"
-white_vs_other_post %>%
-  gt() %>%
+# Contrasts
+aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
+aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
+hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
+hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
+nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
+other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
+contr <- rbind(
+  "Black/AA Pre-ETI" = aa_pre,
+  "Black/AA Post-ETI" = aa_pre + aa_change,
+  "H/L Pre-ETI" = hl_pre,
+  "H/L Post-ETI" = hl_pre + hl_change,
+  "NHW Pre-ETI" = nhw_pre,
+  "NHW Post-ETI" = nhw_pre + nhw_change,
+  "Other Pre-ETI" = other_pre,
+  "Other Post-ETI" = other_pre + other_change
+)
+change_mod_slopes <- glht(change_mod, linfct = contr)
+change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
+change_mod_slopes_ci$confint %>%
+  as.data.frame() %>%
+  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
   fmt_number(decimals = 3)
@@
   fmt_number(decimals = 3)
-# Linear hypotheses
-white_vs_black_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceBlack or African American` = 1.195465"
-)
-white_vs_black_pre <- tidy(summary(white_vs_black_pre))
-white_vs_black_pre$contrast <- "Black vs. White Pre-ETI Slope"
-white_vs_black_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_black_post <- glht(
-  change_mod,
-  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.738885"
-)
-white_vs_black_post <- tidy(summary(white_vs_black_post))
-white_vs_black_post$contrast <- "Black vs. White Post-ETI Slope"
-white_vs_black_post %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-
-white_vs_hispanic_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceHispanic or Latino` = 1.19547"
-)
-white_vs_hispanic_pre <- tidy(summary(white_vs_hispanic_pre))
-white_vs_hispanic_pre$contrast <- "Hispanic vs. White Pre-ETI Slope"
-white_vs_hispanic_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_hispanic_post <- glht(
-  change_mod,
-  "Years + `Years:RaceHispanic or Latino` + slope_change_years + `RaceHispanic or Latino:slope_change_years` = 2.738885"
-)
-white_vs_hispanic_post <- tidy(summary(white_vs_hispanic_post))
-white_vs_hispanic_post$contrast <- "Hispanic vs. White Post-ETI Slope"
-white_vs_hispanic_post %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-
-white_vs_other_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceOther` = 1.19547"
-)
-white_vs_other_pre <- tidy(summary(white_vs_other_pre))
-white_vs_other_pre$contrast <- "Other vs. White Pre-ETI Slope"
-white_vs_other_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_other_post <- glht(
-  change_mod,
-  "Years + `Years:RaceOther` + slope_change_years + `RaceOther:slope_change_years` = 2.738885"
-)
-white_vs_other_post <- tidy(summary(white_vs_other_post))
-white_vs_other_post$contrast <- "Other vs. White Post-ETI Slope"
-white_vs_other_post %>%
-  gt() %>%
+# Contrasts
+aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
+aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
+hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
+hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
+nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
+other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
+contr <- rbind(
+  "Black/AA Pre-ETI" = aa_pre,
+  "Black/AA Post-ETI" = aa_pre + aa_change,
+  "H/L Pre-ETI" = hl_pre,
+  "H/L Post-ETI" = hl_pre + hl_change,
+  "NHW Pre-ETI" = nhw_pre,
+  "NHW Post-ETI" = nhw_pre + nhw_change,
+  "Other Pre-ETI" = other_pre,
+  "Other Post-ETI" = other_pre + other_change
+)
+change_mod_slopes <- glht(change_mod, linfct = contr)
+change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
+change_mod_slopes_ci$confint %>%
+  as.data.frame() %>%
+  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
+  fmt_number(decimals = 3)
+```
+
+## Number of Pulmonary Exacerbations
+
+### Full cohort
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + age_group +
+    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    (1 | eDWID),
+  data = encounter
+)
+# Model plot
+pop_dat <- data.frame(expand.grid(
+  Years = -3:2, Race = unique(encounter$Race), age_group = "[18,30)",
+  Insurance = "Private", prior_modulator = "Yes",
+  isOnEnzymes = "Yes", pseudomonasaeruginosa = "No"
+))
+pop_dat$slope_change_years <- (pop_dat$Years > 0) * pop_dat$Years
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
+  xlab("Years from ETI Start") +
+  ylab("gli_fev1_ppred_rn") +
+  theme_classic()
+# Model summary
+change_mod %>%
+  gtsummary::tbl_regression(intercept = T) %>%
+  as_gt() %>%
+  fmt_number(decimals = 3)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
+  gt(caption = "Predicted Values at Various Timepoints") %>%
+  fmt_number(decimals = 3)
+# Contrasts
+aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
+aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
+hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
+hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
+nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
+other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
+contr <- rbind(
+  "Black/AA Pre-ETI" = aa_pre,
+  "Black/AA Post-ETI" = aa_pre + aa_change,
+  "H/L Pre-ETI" = hl_pre,
+  "H/L Post-ETI" = hl_pre + hl_change,
+  "NHW Pre-ETI" = nhw_pre,
+  "NHW Post-ETI" = nhw_pre + nhw_change,
+  "Other Pre-ETI" = other_pre,
+  "Other Post-ETI" = other_pre + other_change
+)
+change_mod_slopes <- glht(change_mod, linfct = contr)
+change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
+change_mod_slopes_ci$confint %>%
+  as.data.frame() %>%
+  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
+  fmt_number(decimals = 3)
+```
+
+### Age < 18 at ETI initiation
+
+The following models were not adjusted for age group.
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
+    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
+)
+# Model plot
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
+  xlab("Years from ETI Start") +
+  ylab("gli_fev1_ppred_rn") +
+  theme_classic()
+# Model summary
+change_mod %>%
+  gtsummary::tbl_regression(intercept = T) %>%
+  as_gt() %>%
+  fmt_number(decimals = 3)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+# Contrasts
+aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
+aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
+hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
+hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
+nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
+other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
+contr <- rbind(
+  "Black/AA Pre-ETI" = aa_pre,
+  "Black/AA Post-ETI" = aa_pre + aa_change,
+  "H/L Pre-ETI" = hl_pre,
+  "H/L Post-ETI" = hl_pre + hl_change,
+  "NHW Pre-ETI" = nhw_pre,
+  "NHW Post-ETI" = nhw_pre + nhw_change,
+  "Other Pre-ETI" = other_pre,
+  "Other Post-ETI" = other_pre + other_change
+)
+change_mod_slopes <- glht(change_mod, linfct = contr)
+change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
+change_mod_slopes_ci$confint %>%
+  as.data.frame() %>%
+  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
+  fmt_number(decimals = 3)
+```
+
+### Age >= 18 at ETI initiation
+
+The following models were not adjusted for age group.
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
+    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
+  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
+)
+# Model plot
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
+  xlab("Years from ETI Start") +
+  ylab("gli_fev1_ppred_rn") +
+  theme_classic()
+# Model summary
+change_mod %>%
+  gtsummary::tbl_regression(intercept = T) %>%
+  as_gt() %>%
+  fmt_number(decimals = 3)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
+  gt() %>%
+  fmt_number(decimals = 3)
+# Contrasts
+aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
+aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
+hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
+hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
+nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
+other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
+contr <- rbind(
+  "Black/AA Pre-ETI" = aa_pre,
+  "Black/AA Post-ETI" = aa_pre + aa_change,
+  "H/L Pre-ETI" = hl_pre,
+  "H/L Post-ETI" = hl_pre + hl_change,
+  "NHW Pre-ETI" = nhw_pre,
+  "NHW Post-ETI" = nhw_pre + nhw_change,
+  "Other Pre-ETI" = other_pre,
+  "Other Post-ETI" = other_pre + other_change
+)
+change_mod_slopes <- glht(change_mod, linfct = contr)
+change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
+change_mod_slopes_ci$confint %>%
+  as.data.frame() %>%
+  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
   fmt_number(decimals = 3)
@@
 change_mod %>%
-  gtsummary::tbl_regression(intercept = T)%>%
+  gtsummary::tbl_regression(intercept = T) %>%
   as_gt() %>%
@@
   fmt_number(decimals = 3)
-# Linear hypotheses
-white_vs_black_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceBlack or African American` = 1.195465"
-)
-white_vs_black_pre <- tidy(summary(white_vs_black_pre))
-white_vs_black_pre$contrast <- "Black vs. White Pre-ETI Slope"
-white_vs_black_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_black_post <- glht(
-  change_mod,
-  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.738885"
-)
-white_vs_black_post <- tidy(summary(white_vs_black_post))
-white_vs_black_post$contrast <- "Black vs. White Post-ETI Slope"
-white_vs_black_post %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-
-white_vs_hispanic_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceHispanic or Latino` = 1.19547"
-)
-white_vs_hispanic_pre <- tidy(summary(white_vs_hispanic_pre))
-white_vs_hispanic_pre$contrast <- "Hispanic vs. White Pre-ETI Slope"
-white_vs_hispanic_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_hispanic_post <- glht(
-  change_mod,
-  "Years + `Years:RaceHispanic or Latino` + slope_change_years + `RaceHispanic or Latino:slope_change_years` = 2.738885"
-)
-white_vs_hispanic_post <- tidy(summary(white_vs_hispanic_post))
-white_vs_hispanic_post$contrast <- "Hispanic vs. White Post-ETI Slope"
-white_vs_hispanic_post %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-
-white_vs_other_pre <- glht(
-  change_mod,
-  "Years + `Years:RaceOther` = 1.19547"
-)
-white_vs_other_pre <- tidy(summary(white_vs_other_pre))
-white_vs_other_pre$contrast <- "Other vs. White Pre-ETI Slope"
-white_vs_other_pre %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-white_vs_other_post <- glht(
-  change_mod,
-  "Years + `Years:RaceOther` + slope_change_years + `RaceOther:slope_change_years` = 2.738885"
-)
-white_vs_other_post <- tidy(summary(white_vs_other_post))
-white_vs_other_post$contrast <- "Other vs. White Post-ETI Slope"
-white_vs_other_post %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-```
+# Contrasts
+aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
+aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
+hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
+hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
+nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
+other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
+contr <- rbind(
+  "Black/AA Pre-ETI" = aa_pre,
+  "Black/AA Post-ETI" = aa_pre + aa_change,
+  "H/L Pre-ETI" = hl_pre,
+  "H/L Post-ETI" = hl_pre + hl_change,
+  "NHW Pre-ETI" = nhw_pre,
+  "NHW Post-ETI" = nhw_pre + nhw_change,
+  "Other Pre-ETI" = other_pre,
+  "Other Post-ETI" = other_pre + other_change
+)
+change_mod_slopes <- glht(change_mod, linfct = contr)
+change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
+change_mod_slopes_ci$confint %>%
+  as.data.frame() %>%
+  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
+  fmt_number(decimals = 3)
+```

### 2025-07-11 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R

@@ -103,6 +103,16 @@ encounter$cultureresults <- factor(encounter$cultureresults,
   levels = 1:3,
   labels = c("No growth/sterile culture", "Normal flora", "Microorganisms")
 )
+# Pulmonary exacerbations
+encounter$pe_assessment[encounter$pe_assessment == 5] <- NA
+encounter$PEx <- factor(encounter$pe_assessment,
+  levels = 1:4,
+  labels = c(
+    "Absent", "Mild, moderate, or severe exacerbation",
+    "Mild, moderate, or severe exacerbation",
+    "Mild, moderate, or severe exacerbation"
+  )
+)
 # Merge
 encounter <- left_join(
   encounter,

### 2025-07-11 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
removed distribution plots

@@
 
-# Outcome distributions
-
-## FEV1 % predicted
-
-### GLI 2022 equations
-
-```{r}
-ggplot(encounter, aes(x = gli_fev1_ppred_rn, fill = Race, color = Race)) +
-  geom_density() +
-  facet_wrap(~Race)
-ggplot(encounter, aes(x = gli_fev1_ppred_rn, fill = Race)) +
-  geom_density(alpha = 0.2)
-```
-
-### GLI 2012 equations
-
-```{r}
-ggplot(encounter, aes(x = GLI_FEV1_pct_predicted, fill = Race, color = Race)) +
-  geom_density() +
-  facet_wrap(~Race)
-ggplot(encounter, aes(x = GLI_FEV1_pct_predicted, fill = Race)) +
-  geom_density(alpha = 0.2)
-```
-
-## FVC % predicted
-
-### GLI 2022 equations
-
-```{r}
-ggplot(encounter, aes(x = gli_fvc_ppred_rn, fill = Race, color = Race)) +
-  geom_density() +
-  facet_wrap(~Race)
-ggplot(encounter, aes(x = gli_fvc_ppred_rn, fill = Race)) +
-  geom_density(alpha = 0.2)
-```
-
-### GLI 2012 equations
-
-```{r}
-ggplot(encounter, aes(x = GLI_FVC_pct_predicted, fill = Race, color = Race)) +
-  geom_density() +
-  facet_wrap(~Race)
-ggplot(encounter, aes(x = GLI_FVC_pct_predicted, fill = Race)) +
-  geom_density(alpha = 0.2)
-```
-
-## FEV1/FVC % predicted
-
-### GLI 2022 equations
-
-```{r}
-ggplot(encounter, aes(
-  x = gli_fev1fvc_ppred_rn, fill = Race,
-  color = Race
-)) +
-  geom_density() +
-  facet_wrap(~Race)
-ggplot(encounter, aes(x = gli_fev1fvc_ppred_rn, fill = Race)) +
-  geom_density(alpha = 0.2)
-```
-
-### GLI 2012 equations
-
-```{r}
-ggplot(encounter, aes(x = GLI_FEV1FVC_pct_predicted, fill = Race, color = Race)) +
-  geom_density() +
-  facet_wrap(~Race)
-ggplot(encounter, aes(x = GLI_FEV1FVC_pct_predicted, fill = Race)) +
-  geom_density(alpha = 0.2)
-```
-
-## BMI %ile (age < 20 at ETI start)
-
-```{r}
-encounter %>%
-  filter(age_eti_start < 20) %>%
-  ggplot(aes(x = bmipercentile, fill = Race, color = Race)) +
-  geom_density() +
-  facet_wrap(~Race)
-encounter %>%
-  filter(age_eti_start < 20) %>%
-  ggplot(aes(x = bmipercentile, fill = Race)) +
-  geom_density(alpha = 0.2)
-```
-
-## BMI raw value (age >= 20 at ETI start)
-
-```{r}
-encounter %>%
-  filter(age_eti_start >= 20) %>%
-  ggplot(aes(x = bmivalue, fill = Race, color = Race)) +
-  geom_density() +
-  facet_wrap(~Race)
-encounter %>%
-  filter(age_eti_start >= 20) %>%
-  ggplot(aes(x = bmivalue, fill = Race)) +
-  geom_density(alpha = 0.2)
-```
-
-## Height %ile (age < 20 at ETI start)
-
-```{r}
-encounter %>%
-  filter(age_eti_start < 20) %>%
-  ggplot(aes(x = heightpercentile, fill = Race, color = Race)) +
-  geom_density() +
-  facet_wrap(~Race)
-encounter %>%
-  filter(age_eti_start < 20) %>%
-  ggplot(aes(x = heightpercentile, fill = Race)) +
-  geom_density(alpha = 0.2)
-```
-
-## Height raw value (age >= 20 at ETI start)
-
-```{r}
-encounter %>%
-  filter(age_eti_start >= 20) %>%
-  ggplot(aes(x = height, fill = Race, color = Race)) +
-  geom_density() +
-  facet_wrap(~Race)
-encounter %>%
-  filter(age_eti_start >= 20) %>%
-  ggplot(aes(x = height, fill = Race)) +
-  geom_density(alpha = 0.2)
-```
-
-## Weight %ile (age < 20 at ETI start)
-
-```{r}
-encounter %>%
-  filter(age_eti_start < 20) %>%
-  ggplot(aes(x = weightpercentile, fill = Race, color = Race)) +
-  geom_density() +
-  facet_wrap(~Race)
-encounter %>%
-  filter(age_eti_start < 20) %>%
-  ggplot(aes(x = weightpercentile, fill = Race)) +
-  geom_density(alpha = 0.2)
-```
-
-## Weight raw value (age >= 20 at ETI start)
-
-```{r}
-encounter %>%
-  filter(age_eti_start >= 20) %>%
-  ggplot(aes(x = weight, fill = Race, color = Race)) +
-  geom_density() +
-  facet_wrap(~Race)
-encounter %>%
-  filter(age_eti_start >= 20) %>%
-  ggplot(aes(x = weight, fill = Race)) +
-  geom_density(alpha = 0.2)
-```
-
 # Segmented models

### 2025-07-11 Christine Chan/CF Registry/cf_registry_CFRD_prediction.qmd
CFRD joint model predictions

--- 2025-04-17 Christine Chan/CF Registry/cf_registry_CFRD_prediction.qmd
@@
 
+```{r}
+nd <- joint_fit_value$data
+nd$pred <- predict(joint_fit_value, registry[registry$eDWID %in% reg_include, ], type = "Subject",idVar = "eDWID")
+```
+
 # To do

### 2025-07-11 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
much faster outcomes processing (what was I doing?!)

@@
 
-### Full cohort
-
-```{r}
-# Fit mixed model with change point
-change_mod <- lmer(
-  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + age_group +
+<!-- ### Full cohort -->
+
+<!-- ```{r} -->
+<!-- # Fit mixed model with change point -->
+<!-- change_mod <- glmer( -->
+<!--   PEx ~ Years * Race + slope_change_years * Race + age_group + -->
+<!--     Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa + -->
+<!--     (1 | eDWID), -->
+<!--   family = "binomial", -->
+<!--   data = encounter -->
+<!-- ) -->
+<!-- # Model plot -->
+<!-- pop_dat <- data.frame(expand.grid( -->
+<!--   Years = -3:2, Race = unique(encounter$Race), age_group = "[18,30)", -->
+<!--   Insurance = "Private", prior_modulator = "Yes", -->
+<!--   isOnEnzymes = "Yes", pseudomonasaeruginosa = "No" -->
+<!-- )) -->
+<!-- pop_dat$slope_change_years <- (pop_dat$Years > 0) * pop_dat$Years -->
+<!-- pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T) -->
+<!-- pop_dat$pred <- pred$fit -->
+<!-- pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit) -->
+<!-- pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit) -->
+<!-- ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) + -->
+<!--   geom_point() + -->
+<!--   geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) + -->
+<!--   geom_line() + -->
+<!--   xlab("Years from ETI Start") + -->
+<!--   ylab("gli_fev1_ppred_rn") + -->
+<!--   theme_classic() -->
+<!-- # Model summary -->
+<!-- change_mod %>% -->
+<!--   gtsummary::tbl_regression(intercept = T) %>% -->
+<!--   as_gt() %>% -->
+<!--   fmt_number(decimals = 3) -->
+<!-- # Table of predicted values -->
+<!-- pop_dat %>% -->
+<!--   select(-slope_change_years) %>% -->
+<!--   rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>% -->
+<!--   gt(caption = "Predicted Values at Various Timepoints") %>% -->
+<!--   fmt_number(decimals = 3) -->
+<!-- # Contrasts -->
+<!-- aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0) -->
+<!-- aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0) -->
+<!-- hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0) -->
+<!-- hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0) -->
+<!-- nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0) -->
+<!-- nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0) -->
+<!-- other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0) -->
+<!-- other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1) -->
+<!-- contr <- rbind( -->
+<!--   "Black/AA Pre-ETI" = aa_pre, -->
+<!--   "Black/AA Post-ETI" = aa_pre + aa_change, -->
+<!--   "H/L Pre-ETI" = hl_pre, -->
+<!--   "H/L Post-ETI" = hl_pre + hl_change, -->
+<!--   "NHW Pre-ETI" = nhw_pre, -->
+<!--   "NHW Post-ETI" = nhw_pre + nhw_change, -->
+<!--   "Other Pre-ETI" = other_pre, -->
+<!--   "Other Post-ETI" = other_pre + other_change -->
+<!-- ) -->
+<!-- change_mod_slopes <- glht(change_mod, linfct = contr) -->
+<!-- change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha()) -->
+<!-- change_mod_slopes_ci$confint %>% -->
+<!--   as.data.frame() %>% -->
+<!--   gt(caption = "Slopes by Race", rownames_to_stub = T) %>% -->
+<!--   fmt_number(decimals = 3) -->
+<!-- ``` -->
+
+<!-- ### Age < 18 at ETI initiation -->
+
+<!-- The following models were not adjusted for age group. -->
+
+<!-- ```{r} -->
+<!-- # Fit mixed model with change point -->
+<!-- change_mod <- lmer( -->
+<!--   gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance + -->
+<!--     prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID), -->
+<!--   data = encounter[encounter$age_eti_group == "[-Inf,18)", ] -->
+<!-- ) -->
+<!-- # Model plot -->
+<!-- pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T) -->
+<!-- pop_dat$pred <- pred$fit -->
+<!-- pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit) -->
+<!-- pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit) -->
+<!-- ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) + -->
+<!--   geom_point() + -->
+<!--   geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) + -->
+<!--   geom_line() + -->
+<!--   xlab("Years from ETI Start") + -->
+<!--   ylab("gli_fev1_ppred_rn") + -->
+<!--   theme_classic() -->
+<!-- # Model summary -->
+<!-- change_mod %>% -->
+<!--   gtsummary::tbl_regression(intercept = T) %>% -->
+<!--   as_gt() %>% -->
+<!--   fmt_number(decimals = 3) -->
+<!-- # Table of predicted values -->
+<!-- pop_dat %>% -->
+<!--   select(-slope_change_years) %>% -->
+<!--   rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>% -->
+<!--   gt() %>% -->
+<!--   fmt_number(decimals = 3) -->
+<!-- # Contrasts -->
+<!-- aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0) -->
+<!-- aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0) -->
+<!-- hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0) -->
+<!-- hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0) -->
+<!-- nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0) -->
+<!-- nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0) -->
+<!-- other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0) -->
+<!-- other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1) -->
+<!-- contr <- rbind( -->
+<!--   "Black/AA Pre-ETI" = aa_pre, -->
+<!--   "Black/AA Post-ETI" = aa_pre + aa_change, -->
+<!--   "H/L Pre-ETI" = hl_pre, -->
+<!--   "H/L Post-ETI" = hl_pre + hl_change, -->
+<!--   "NHW Pre-ETI" = nhw_pre, -->
+<!--   "NHW Post-ETI" = nhw_pre + nhw_change, -->
+<!--   "Other Pre-ETI" = other_pre, -->
+<!--   "Other Post-ETI" = other_pre + other_change -->
+<!-- ) -->
+<!-- change_mod_slopes <- glht(change_mod, linfct = contr) -->
+<!-- change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha()) -->
+<!-- change_mod_slopes_ci$confint %>% -->
+<!--   as.data.frame() %>% -->
+<!--   gt(caption = "Slopes by Race", rownames_to_stub = T) %>% -->
+<!--   fmt_number(decimals = 3) -->
+<!-- ``` -->
+
+<!-- ### Age >= 18 at ETI initiation -->
+
+<!-- The following models were not adjusted for age group. -->
+
+<!-- ```{r} -->
+<!-- # Fit mixed model with change point -->
+<!-- change_mod <- lmer( -->
+<!--   gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance + -->
+<!--     prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID), -->
+<!--   data = encounter[encounter$age_eti_group == "[18, Inf)", ] -->
+<!-- ) -->
+<!-- # Model plot -->
+<!-- pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T) -->
+<!-- pop_dat$pred <- pred$fit -->
+<!-- pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit) -->
+<!-- pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit) -->
+<!-- ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) + -->
+<!--   geom_point() + -->
+<!--   geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) + -->
+<!--   geom_line() + -->
+<!--   xlab("Years from ETI Start") + -->
+<!--   ylab("gli_fev1_ppred_rn") + -->
+<!--   theme_classic() -->
+<!-- # Model summary -->
+<!-- change_mod %>% -->
+<!--   gtsummary::tbl_regression(intercept = T) %>% -->
+<!--   as_gt() %>% -->
+<!--   fmt_number(decimals = 3) -->
+<!-- # Table of predicted values -->
+<!-- pop_dat %>% -->
+<!--   select(-slope_change_years) %>% -->
+<!--   rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>% -->
+<!--   gt() %>% -->
+<!--   fmt_number(decimals = 3) -->
+<!-- # Contrasts -->
+<!-- aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0) -->
+<!-- aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0) -->
+<!-- hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0) -->
+<!-- hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0) -->
+<!-- nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0) -->
+<!-- nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0) -->
+<!-- other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0) -->
+<!-- other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1) -->
+<!-- contr <- rbind( -->
+<!--   "Black/AA Pre-ETI" = aa_pre, -->
+<!--   "Black/AA Post-ETI" = aa_pre + aa_change, -->
+<!--   "H/L Pre-ETI" = hl_pre, -->
+<!--   "H/L Post-ETI" = hl_pre + hl_change, -->
+<!--   "NHW Pre-ETI" = nhw_pre, -->
+<!--   "NHW Post-ETI" = nhw_pre + nhw_change, -->
+<!--   "Other Pre-ETI" = other_pre, -->
+<!--   "Other Post-ETI" = other_pre + other_change -->
+<!-- ) -->
+<!-- change_mod_slopes <- glht(change_mod, linfct = contr) -->
+<!-- change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha()) -->
+<!-- change_mod_slopes_ci$confint %>% -->
+<!--   as.data.frame() %>% -->
+<!--   gt(caption = "Slopes by Race", rownames_to_stub = T) %>% -->
+<!--   fmt_number(decimals = 3) -->
+<!-- ``` -->
+
+## BMI
+
+### BMI %ile (age < 20 at ETI start)
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  bmipercentile ~ Years * Race + slope_change_years * Race +
     Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
     (1 | eDWID),
-  data = encounter
-)
-# Model plot
-pop_dat <- data.frame(expand.grid(
-  Years = -3:2, Race = unique(encounter$Race), age_group = "[18,30)",
-  Insurance = "Private", prior_modulator = "Yes",
-  isOnEnzymes = "Yes", pseudomonasaeruginosa = "No"
-))
-pop_dat$slope_change_years <- (pop_dat$Years > 0) * pop_dat$Years
-pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
-pop_dat$pred <- pred$fit
-pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
-pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
-ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
-  geom_line() +
-  xlab("Years from ETI Start") +
-  ylab("gli_fev1_ppred_rn") +
+  data = encounter[encounter$age_eti_start < 20, ]
+)
+# Model plot
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
+  xlab("Years from ETI Start") +
+  ylab("bmipercentile") +
   theme_classic()
@@
   select(-slope_change_years) %>%
-  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
-  gt(caption = "Predicted Values at Various Timepoints") %>%
-  fmt_number(decimals = 3)
-# Contrasts
-aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
-aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
-hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
-hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
-nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
-nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
-other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
-other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
-contr <- rbind(
-  "Black/AA Pre-ETI" = aa_pre,
-  "Black/AA Post-ETI" = aa_pre + aa_change,
-  "H/L Pre-ETI" = hl_pre,
-  "H/L Post-ETI" = hl_pre + hl_change,
-  "NHW Pre-ETI" = nhw_pre,
-  "NHW Post-ETI" = nhw_pre + nhw_change,
-  "Other Pre-ETI" = other_pre,
-  "Other Post-ETI" = other_pre + other_change
-)
-change_mod_slopes <- glht(change_mod, linfct = contr)
-change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
-change_mod_slopes_ci$confint %>%
-  as.data.frame() %>%
-  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
-  fmt_number(decimals = 3)
-```
-
-### Age < 18 at ETI initiation
-
-The following models were not adjusted for age group.
-
-```{r}
-# Fit mixed model with change point
-change_mod <- lmer(
-  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
-    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
-  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
-)
-# Model plot
-pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
-pop_dat$pred <- pred$fit
-pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
-pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
-ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
-  geom_line() +
-  xlab("Years from ETI Start") +
-  ylab("gli_fev1_ppred_rn") +
-  theme_classic()
-# Model summary
-change_mod %>%
-  gtsummary::tbl_regression(intercept = T) %>%
-  as_gt() %>%
-  fmt_number(decimals = 3)
-# Table of predicted values
-pop_dat %>%
-  select(-slope_change_years) %>%
-  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
+  rename(bmipercentile = pred, ci.ll = ll, ci.ul = ul) %>%
   gt() %>%
@@
 
-### Age >= 18 at ETI initiation
-
-The following models were not adjusted for age group.
-
-```{r}
-# Fit mixed model with change point
-change_mod <- lmer(
-  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
-    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
-  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
-)
-# Model plot
-pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
-pop_dat$pred <- pred$fit
-pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
-pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
-ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
-  geom_line() +
-  xlab("Years from ETI Start") +
-  ylab("gli_fev1_ppred_rn") +
+### BMI raw value (age >= 20 at ETI start)
+
+```{r}
+change_mod <- lmer(
+  bmivalue ~ Years * Race + slope_change_years * Race +
+    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    (1 | eDWID),
+  data = encounter[encounter$age_eti_start >= 20, ]
+)
+# Model plot
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
+  xlab("Years from ETI Start") +
+  ylab("bmivalue") +
   theme_classic()
@@
   select(-slope_change_years) %>%
-  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
+  rename(bmivalue = pred, ci.ll = ll, ci.ul = ul) %>%
   gt() %>%
@@
 
-## BMI
-
-### BMI %ile (age < 20 at ETI start)
-
-```{r}
-# Fit mixed model with change point
-change_mod <- lmer(
-  bmipercentile ~ Years * Race + slope_change_years * Race +
+## Height
+
+### Height %ile (age < 20 at ETI start)
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  heightpercentile ~ Years * Race + slope_change_years * Race +
     Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
@@
   xlab("Years from ETI Start") +
-  ylab("bmipercentile") +
-  theme_classic()
-# Model summary
-change_mod %>%
-  gtsummary::tbl_regression(intercept = T) %>%
-  as_gt() %>%
-  fmt_number(decimals = 3)
-# Table of predicted values
-pop_dat %>%
-  select(-slope_change_years) %>%
-  rename(bmipercentile = pred, ci.ll = ll, ci.ul = ul) %>%
+  ylab("heightpercentile") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression(intercept = T)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(heightpercentile = pred, ci.ll = ll, ci.ul = ul) %>%
   gt() %>%
@@
 
-### BMI raw value (age >= 20 at ETI start)
-
-```{r}
-change_mod <- lmer(
-  bmivalue ~ Years * Race + slope_change_years * Race +
-    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
-    (1 | eDWID),
+### Height raw value (age >= 20 at ETI start)
+
+```{r}
+change_mod <- lmer(
+  height ~ Years * Race + slope_change_years * Race + Insurance +
+    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
   data = encounter[encounter$age_eti_start >= 20, ]
@@
   xlab("Years from ETI Start") +
-  ylab("bmivalue") +
-  theme_classic()
-# Model summary
-change_mod %>%
-  gtsummary::tbl_regression(intercept = T) %>%
-  as_gt() %>%
-  fmt_number(decimals = 3)
-# Table of predicted values
-pop_dat %>%
-  select(-slope_change_years) %>%
-  rename(bmivalue = pred, ci.ll = ll, ci.ul = ul) %>%
+  ylab("height") +
+  theme_classic()
+# Model summary
+change_mod %>% gtsummary::tbl_regression(intercept = T)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(height = pred, ci.ll = ll, ci.ul = ul) %>%
   gt() %>%
@@
 
-## Height
-
-### Height %ile (age < 20 at ETI start)
-
-```{r}
-# Fit mixed model with change point
-change_mod <- lmer(
-  heightpercentile ~ Years * Race + slope_change_years * Race +
+## Weight
+
+### Weight %ile (age < 20 at ETI start)
+
+```{r}
+# Fit mixed model with change point
+change_mod <- lmer(
+  weightpercentile ~ Years * Race + slope_change_years * Race +
     Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
@@
   xlab("Years from ETI Start") +
-  ylab("heightpercentile") +
-  theme_classic()
-# Model summary
-change_mod %>% gtsummary::tbl_regression(intercept = T)
-# Table of predicted values
-pop_dat %>%
-  select(-slope_change_years) %>%
-  rename(heightpercentile = pred, ci.ll = ll, ci.ul = ul) %>%
+  ylab("weightpercentile") +
+  theme_classic()
+# Model summary
+change_mod %>%
+  gtsummary::tbl_regression(intercept = T) %>%
+  as_gt() %>%
+  fmt_number(decimals = 3)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(weightpercentile = pred, ci.ll = ll, ci.ul = ul) %>%
   gt() %>%
@@
 
-### Height raw value (age >= 20 at ETI start)
-
-```{r}
-change_mod <- lmer(
-  height ~ Years * Race + slope_change_years * Race + Insurance +
+### Weight raw value (age >= 20 at ETI start)
+
+```{r}
+change_mod <- lmer(
+  weight ~ Years * Race + slope_change_years * Race + Insurance +
     prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
@@
   xlab("Years from ETI Start") +
-  ylab("height") +
-  theme_classic()
-# Model summary
-change_mod %>% gtsummary::tbl_regression(intercept = T)
-# Table of predicted values
-pop_dat %>%
-  select(-slope_change_years) %>%
-  rename(height = pred, ci.ll = ll, ci.ul = ul) %>%
+  ylab("weight") +
+  theme_classic()
+# Model summary
+change_mod %>%
+  gtsummary::tbl_regression(intercept = T) %>%
+  as_gt() %>%
+  fmt_number(decimals = 3)
+# Table of predicted values
+pop_dat %>%
+  select(-slope_change_years) %>%
+  rename(weight = pred, ci.ll = ll, ci.ul = ul) %>%
   gt() %>%
@@
 ```
-
-## Weight
-
-### Weight %ile (age < 20 at ETI start)
-
-```{r}
-# Fit mixed model with change point
-change_mod <- lmer(
-  weightpercentile ~ Years * Race + slope_change_years * Race +
-    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
-    (1 | eDWID),
-  data = encounter[encounter$age_eti_start < 20, ]
-)
-# Model plot
-pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
-pop_dat$pred <- pred$fit
-pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
-pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
-ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
-  geom_line() +
-  xlab("Years from ETI Start") +
-  ylab("weightpercentile") +
-  theme_classic()
-# Model summary
-change_mod %>%
-  gtsummary::tbl_regression(intercept = T) %>%
-  as_gt() %>%
-  fmt_number(decimals = 3)
-# Table of predicted values
-pop_dat %>%
-  select(-slope_change_years) %>%
-  rename(weightpercentile = pred, ci.ll = ll, ci.ul = ul) %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-# Contrasts
-aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
-aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
-hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
-hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
-nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
-nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
-other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
-other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
-contr <- rbind(
-  "Black/AA Pre-ETI" = aa_pre,
-  "Black/AA Post-ETI" = aa_pre + aa_change,
-  "H/L Pre-ETI" = hl_pre,
-  "H/L Post-ETI" = hl_pre + hl_change,
-  "NHW Pre-ETI" = nhw_pre,
-  "NHW Post-ETI" = nhw_pre + nhw_change,
-  "Other Pre-ETI" = other_pre,
-  "Other Post-ETI" = other_pre + other_change
-)
-change_mod_slopes <- glht(change_mod, linfct = contr)
-change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
-change_mod_slopes_ci$confint %>%
-  as.data.frame() %>%
-  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
-  fmt_number(decimals = 3)
-```
-
-### Weight raw value (age >= 20 at ETI start)
-
-```{r}
-change_mod <- lmer(
-  weight ~ Years * Race + slope_change_years * Race + Insurance +
-    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
-  data = encounter[encounter$age_eti_start >= 20, ]
-)
-# Model plot
-pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
-pop_dat$pred <- pred$fit
-pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
-pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
-ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
-  geom_line() +
-  xlab("Years from ETI Start") +
-  ylab("weight") +
-  theme_classic()
-# Model summary
-change_mod %>%
-  gtsummary::tbl_regression(intercept = T) %>%
-  as_gt() %>%
-  fmt_number(decimals = 3)
-# Table of predicted values
-pop_dat %>%
-  select(-slope_change_years) %>%
-  rename(weight = pred, ci.ll = ll, ci.ul = ul) %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-# Contrasts
-aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
-aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
-hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
-hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
-nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
-nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
-other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
-other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
-contr <- rbind(
-  "Black/AA Pre-ETI" = aa_pre,
-  "Black/AA Post-ETI" = aa_pre + aa_change,
-  "H/L Pre-ETI" = hl_pre,
-  "H/L Post-ETI" = hl_pre + hl_change,
-  "NHW Pre-ETI" = nhw_pre,
-  "NHW Post-ETI" = nhw_pre + nhw_change,
-  "Other Pre-ETI" = other_pre,
-  "Other Post-ETI" = other_pre + other_change
-)
-change_mod_slopes <- glht(change_mod, linfct = contr)
-change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
-change_mod_slopes_ci$confint %>%
-  as.data.frame() %>%
-  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
-  fmt_number(decimals = 3)
-```

### 2025-07-11 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R

@@ -1,6 +1,5 @@
 library(tidyverse)
 library(lubridate)
-library(parallel)
 # Paths
 home_dir <- switch(Sys.info()["sysname"],
   "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
@@ -140,16 +139,14 @@ encounter$age_eti_group <- cut(encounter$age_eti_start,
 # Hospitalization or home IVs
 hospitalizations$CareEpi_StartDt <- mdy(hospitalizations$CareEpi_StartDt)
 hospitalizations$CareEpi_EndDt <- mdy(hospitalizations$CareEpi_EndDt)
-cl <- makeForkCluster(detectCores() * (2 / 4))
-hosps <- parApply(cl, hospitalizations, 1, function(r) {
-  which(encounter$eDWID == r["eDWID"] &
-    encounter$encounterdate >= r["CareEpi_StartDt"] &
-    encounter$encounterdate <= r["CareEpi_EndDt"])
-})
-stopCluster(cl)
-hosps <- unique(unlist(hosps))
-encounter$hospitalized <- "No"
-encounter$hospitalized[hosps] <- "Yes"
+hospitalizations <- hospitalizations %>%
+  rowwise() %>%
+  reframe(eDWID,
+    encounterdate = seq(CareEpi_StartDt, CareEpi_EndDt, by = "day")
+  ) %>%
+  mutate(hospitalized = "Yes")
+encounter <- 
+  left_join(encounter, hospitalizations, by = join_by(eDWID, encounterdate))
 # Variables for flowchart
 n_enc_1 <- nrow(encounter)
 n_people_1 <- length(unique(encounter$eDWID))

### 2025-07-14 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
cleaned up EnVision folder
new GCT figures

@@
 ```{r}
-#| label: fig-boxplot
-#| fig-cap: Boxplot of OGTT glucose values by diabetes status
 #| warning: false
@@
   geom_point() +
+  geom_hline(yintercept = 140, linetype = "dashed") +
+  geom_hline(yintercept = 150, linetype = "dashed") +
+  geom_hline(yintercept = 160, linetype = "dashed") +
+  geom_hline(yintercept = 147, linetype = "dashed", size = 1) +
   ylab("GCTpl (mg/dL)") +
-  theme_bw()
+  theme_bw() +
+  scale_y_continuous(breaks = c(100, 140, 147, 150, 160))
+
 dotplot
@@
 plot_df <- df %>%
-  pivot_longer(c(`OGTT 0 min`, `OGTT 60 min`, `OGTT 120 min`, `GCT 60 min plasma`),
+  pivot_longer(
+    c(
+      `OGTT 0 min`, `OGTT 120 min`, `GCT 0 min plasma`, `GCT 60 min plasma`,
+      `GCT 60 min capillary`
+    ),
     names_to = "OGTT Timepoint", values_to = "Glucose"
@@
 plot_df$`OGTT Timepoint` <- factor(plot_df$`OGTT Timepoint`,
-  levels = c("OGTT 0 min", "OGTT 60 min", "OGTT 120 min", "GCT 60 min plasma"),
-  labels = c("Fasting", "OGTT-1hG", "OGTT-2hG", "GCTpl")
+  levels = c(
+    "OGTT 0 min", "OGTT 120 min", "GCT 0 min plasma", "GCT 60 min plasma",
+    "GCT 60 min capillary"
+  ),
+  labels = c("Fasting", "OGTT-2hG", "RPG", "GCTpl", "GCTcap")
 )
@@
   scale_y_continuous(breaks = c(100, 140, 200, 300)) +
-  scale_fill_manual(values = RColorBrewer::brewer.pal(4, "Set2")) +
+  scale_fill_manual(values = RColorBrewer::brewer.pal(5, "Dark2")) +
   facet_wrap(~Diabetes) +
@@
 )
-# GCTpl, RCG, HbA1c, OGTT1hr
+# RPG, GCTpl, RCG, HbA1c, OGTT1hr
 plot_df <- df %>%
@@
     c(
-      `GCT 60 min plasma`, `GCT 0 min capillary`, HbA1c,
-      `OGTT 60 min`
+      `GCT 0 min plasma`, `GCT 60 min plasma`, `GCT 0 min capillary`,
+      `GCT 60 min capillary`, HbA1c
     ),
@@
   )
-box_p1 <- ggplot(
-  plot_df[plot_df$Measure == "GCT 60 min plasma", ],
-  aes(x = Diabetes, y = Value, fill = Diabetes)
-) +
-  geom_boxplot() +
-  geom_hline(yintercept = 146.5, linetype = "dashed") +
-  ylim(c(50, 300)) +
-  ylab("mg/dL") +
-  ggtitle("GCTpl") +
-  theme_bw(base_size = 15) +
-  theme(
-    legend.position = "none",
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.title.x = element_blank()
-  )
-box_p2 <- ggplot(
-  plot_df[plot_df$Measure == "GCT 0 min capillary", ],
-  aes(x = Diabetes, y = Value, fill = Diabetes)
-) +
-  geom_boxplot() +
-  geom_hline(yintercept = 98.5, linetype = "dashed") +
-  ylim(c(50, 300)) +
-  ggtitle("RCG") +
-  theme_bw(base_size = 15) +
-  theme(
-    legend.position = "none",
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.title.x = element_blank(),
-    axis.title.y = element_blank()
-  )
-box_p3 <- ggplot(
-  plot_df[plot_df$Measure == "HbA1c", ],
-  aes(x = Diabetes, y = Value, fill = Diabetes)
-) +
-  geom_boxplot() +
-  geom_hline(yintercept = 5.45, linetype = "dashed") +
-  ylab("%") +
-  ggtitle("HbA1c") +
-  theme_bw(base_size = 15) +
-  theme(
-    legend.position = "none",
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.title.x = element_blank(),
-    axis.title.y = element_blank()
-  )
-box_p4 <- ggplot(
-  plot_df[plot_df$Measure == "OGTT 60 min", ],
-  aes(x = Diabetes, y = Value, fill = Diabetes)
-) +
-  geom_boxplot() +
-  geom_hline(yintercept = 228.5, linetype = "dashed") +
-  ylim(c(50, 300)) +
-  ggtitle("OGTT 60 min") +
-  theme_bw(base_size = 15) +
-  theme(
-    legend.position = "none",
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.title.x = element_blank(),
-    axis.title.y = element_blank()
-  )
-box_p <- box_p1 + box_p2 + box_p3 + box_p4
-ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/boxplot2.png",
-  box_p,
-  device = "png"
-)
 # Dotplot
-box_p1 <- ggplot(
-  plot_df[plot_df$Measure == "GCT 60 min plasma", ],
+dot_p1 <- ggplot(
+  plot_df[plot_df$Measure == "GCT 0 min plasma", ],
   aes(x = Diabetes, y = Value, fill = Diabetes, colour = Diabetes)
@@
   geom_jitter() +
-  geom_hline(yintercept = 146.5, linetype = "dashed", size = 1) +
+  geom_hline(yintercept = 91.5, linetype = "dashed", size = 1) +
   ylim(c(50, 300)) +
   ylab("Glucose (mg/dL)") +
-  ggtitle("GCTpl") +
+  ggtitle("RPG") +
   theme_bw(base_size = 15) +
-  scale_colour_manual(values = RColorBrewer::brewer.pal(3, "Set2")) +
+  scale_colour_manual(values = c("#00ADB5", "#F4CE14", "#222831")) +
   theme(
@@
   )
-box_p2 <- ggplot(
-  plot_df[plot_df$Measure == "GCT 0 min capillary", ],
+dot_p2 <- ggplot(
+  plot_df[plot_df$Measure == "GCT 60 min plasma", ],
   aes(x = Diabetes, y = Value, fill = Diabetes, colour = Diabetes)
@@
   geom_jitter() +
-  geom_hline(yintercept = 98.5, linetype = "dashed", size = 1) +
+  geom_hline(yintercept = 140.5, linetype = "dashed", size = 1) +
   ylim(c(50, 300)) +
-  ggtitle("RCG") +
+  ylab("Glucose (mg/dL)") +
+  ggtitle("GCTpl") +
   theme_bw(base_size = 15) +
-  scale_colour_manual(values = RColorBrewer::brewer.pal(3, "Set2")) +
+  scale_colour_manual(values = c("#00ADB5", "#F4CE14", "#222831")) +
   theme(
@@
   )
-box_p3 <- ggplot(
+dot_p3 <- ggplot(
+  plot_df[plot_df$Measure == "GCT 0 min capillary", ],
+  aes(x = Diabetes, y = Value, fill = Diabetes, colour = Diabetes)
+) +
+  geom_jitter() +
+  geom_hline(yintercept = 97.5, linetype = "dashed", size = 1) +
+  ylim(c(50, 300)) +
+  ggtitle("RCG") +
+  theme_bw(base_size = 15) +
+  scale_colour_manual(values = c("#00ADB5", "#F4CE14", "#222831")) +
+  theme(
+    legend.position = "none", plot.title = element_text(hjust = 0.5),
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank()
+  )
+dot_p4 <- ggplot(
+  plot_df[plot_df$Measure == "GCT 60 min capillary", ],
+  aes(x = Diabetes, y = Value, fill = Diabetes, colour = Diabetes)
+) +
+  geom_jitter() +
+  geom_hline(yintercept = 161.5, linetype = "dashed", size = 1) +
+  ylim(c(50, 300)) +
+  ggtitle("GCTcap") +
+  theme_bw(base_size = 15) +
+  scale_colour_manual(values = c("#00ADB5", "#F4CE14", "#222831")) +
+  theme(
+    legend.position = "none", plot.title = element_text(hjust = 0.5),
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank()
+  )
+dot_p5 <- ggplot(
   plot_df[plot_df$Measure == "HbA1c", ],
@@
   theme_bw(base_size = 15) +
-  scale_colour_manual(values = RColorBrewer::brewer.pal(3, "Set2")) +
+  scale_colour_manual(values = c("#00ADB5", "#F4CE14", "#222831")) +
   scale_linetype_manual(values = c("Youden Index" = "dashed")) +
@@
   )
-box_p4 <- ggplot(
-  plot_df[plot_df$Measure == "OGTT 60 min", ],
-  aes(x = Diabetes, y = Value, fill = Diabetes, colour = Diabetes)
-) +
-  geom_jitter() +
-  geom_hline(yintercept = 228.5, linetype = "dashed", size = 1) +
-  ylim(c(50, 300)) +
-  ggtitle("OGTT 1hPG") +
-  theme_bw(base_size = 15) +
-  scale_colour_manual(values = RColorBrewer::brewer.pal(3, "Set2")) +
-  theme(
-    legend.position = "none", plot.title = element_text(hjust = 0.5),
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.title.x = element_blank(),
-    axis.title.y = element_blank()
-  )
-box_p <- box_p1 + box_p2 + box_p4 + box_p3 + plot_layout(ncol = 4)
+dot_p <- dot_p1 + dot_p2 + dot_p3 + dot_p4 + dot_p5 + plot_layout(ncol = 5)
 ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/dotplot2.png",

### 2025-07-14 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
added DOB to sweat tests

@@
   select(
-    PatientID, Name, Genotypes1, Genotypes2, Sex, Age.at.Test.in.Years,
+    PatientID, Name, DOB, Genotypes1, Genotypes2, Sex, Age.at.Test.in.Years,
     Na, K, Cl, Na_Left, Na_Right, K_Left, K_Right, Cl_Left, Cl_Right
@@
 chco$Na_Left[790] <- 24.1
-# For "LESS THAN 10", replace with half the limit of detection
-chco[chco == "LESS THAN 10"] <- "9"
+# For "LESS THAN 10", replace with missing
+chco[chco == "LESS THAN 10"] <- NA
 chco[, c("Cl", "Cl_Left", "Cl_Right")] <-
@@
   select(
-    Patient.ID, Patient.Name, Genotype.1, Genotype.2, Sex, Age.at.test..in.years.,
-    Sweat.sodium, Sweat.potassium, Sweat.chloride, Sweat.sodium.left.arm,
-    Sweat.sodium.right.arm, Sweat.potassium.left.arm, Sweat.potassium.right.arm,
-    Sweat.chloride.left.arm, Sweat.chloride.right.arm, Diagnosis
+    Patient.ID, Patient.Name, Birth.Date, Genotype.1, Genotype.2,
+    Sex, Age.at.test..in.years., Sweat.sodium, Sweat.potassium, Sweat.chloride,
+    Sweat.sodium.left.arm, Sweat.sodium.right.arm, Sweat.potassium.left.arm,
+    Sweat.potassium.right.arm, Sweat.chloride.left.arm,
+    Sweat.chloride.right.arm, Diagnosis
   ) %>%
   rename(
-    Name = Patient.Name, Genotypes1 = Genotype.1,
+    Name = Patient.Name, Genotypes1 = Genotype.1, DOB = Birth.Date,
     Genotypes2 = Genotype.2, Age = Age.at.test..in.years., Na = Sweat.sodium,
@@
 df$K <- coalesce(df$K, df$K_mean)
-# Change Cl under 10 to 9 per CHEC
-df$Cl[df$Cl < 10] <- 9
-df$Na[df$Na < 10] <- 9
+# Change Cl under 10 to missing
+df$Cl[df$Cl < 10] <- NA
+df$Na[df$Na < 10] <- NA
 # Calculate Na:Cl ratio
@@
 df <- df %>%
-  select(Name, Diagnosis, Genotypes1, Genotypes2, Sex, Age, Na, K, Cl, `Na:Cl`) %>%
+  select(
+    Name, DOB, Diagnosis, Genotypes1, Genotypes2, Sex, Age, Na, K, Cl, `Na:Cl`
+  ) %>%
   filter(!is.na(Diagnosis))

### 2025-07-14 Emily Muther/body_image_trikafta.qmd
raw BMIs for body image and trikafta

@@
 levels(df$WeightBody_Adherence) <- c("Agree", "Agree", "Neutral", "Disagree", "Disagree")
+# Raw BMI values from Emma
+df$bmi_raw <- NA
+df$bmi_raw[df$ID == 1] <- 24.21
+df$BMI[df$ID == 1] <- NA
+df$bmi_raw[df$ID == 8] <- 22.33
+df$BMI[df$ID == 8] <- NA
+df$bmi_raw[df$ID == 11] <- 21.1
+df$BMI[df$ID == 11] <- NA
+df$bmi_raw[df$ID == 31] <- 23.64
+df$BMI[df$ID == 31] <- NA
+df$bmi_raw[df$ID == 39] <- 20.73
+df$BMI[df$ID == 39] <- NA
 # Labels
@@
 label(df$BE_WEIGHT_MEAN) <- "Mean Body Esteem Scale (Weight)"
+label(df$BMI) = "BMI %ile"
+label(df$bmi_raw) = "BMI (kg/m2)"
 ```
@@
     FRS_IdealBody, FRS_ThinkyouLook, FRS_Category, Age, Gender, Race,
-    Ethnicity, InsuranceType, BMI, FEV1
+    Ethnicity, InsuranceType, BMI, bmi_raw,FEV1
   ) %>%
@@
     type = list(
-      c("FRStotal", "FRS_IdealBody", "FRS_ThinkyouLook") ~ "continuous2",
+      c("FRStotal", "FRS_IdealBody", "FRS_ThinkyouLook","bmi_raw") ~ "continuous2",
       all_continuous() ~ "continuous2"
@@
 # Save
-ggsave("./Emily Muther/General/BodyImage_Trikafta/Reports/forest_plot.png",
-  plot = p, width = 9, height = 6, units = "in"
-)
+# ggsave("./Emily Muther/General/BodyImage_Trikafta/Reports/forest_plot.png",
+#   plot = p, width = 9, height = 6, units = "in"
+# )
 ```

### 2025-07-14 Lilah Melzer/cough_monitor_analysis.qmd
cough monitor updates

@@
   tableby.control(
-    numeric.stats = c("N", "Nmiss", "medianq1q3", "range"),
-    cat.stats = c("N", "Nmiss", "countpct"),
+    numeric.stats = c("N", "medianrange"),
+    cat.stats = c("N", "countpct"),
     numeric.test = "kwt"
@@
     `Prop. Nights With Cough` = sum(Coughs > 0, na.rm = T) / n(),
+    `Prop. Nights With 0 Cough` = sum(Coughs == 0, na.rm = T) / n(),
     .groups = "drop"
@@
   Diagnosis ~ `Nights` + `Nights With Data` + `Prop. Nights With Data` +
-    `Prop. Nights With Cough` + `Median Coughs per Night`,
+    `Prop. Nights With 0 Cough` + `Prop. Nights With Cough` +
+    `Median Coughs per Night`,
   data = per_person, control = mycontrols

### 2025-07-16 Jordana Hoppe/Home Collection/manuscript.qmd
added additional survey exclusion

@@
 github_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Documents/GitHub",
+  "Darwin" = "/Users/tim/Documents/GitHub",
   "Windows" = "C:/Users/Tim/GitHub",
@@
 data_exclude <- c("009-07", "009-17", "009-25", "009-37")
-survey_exclude <- c(15, 27, 30, 34, 47)
+survey_exclude <- c(6, 15, 27, 30, 34, 47)
 data <- data %>% filter(!subject_id %in% data_exclude)

### 2025-07-17 Emily Muther/body_image_trikafta.qmd
added BE and BMI to table 1

@@
 label(df$BE_WEIGHT_MEAN) <- "Mean Body Esteem Scale (Weight)"
-label(df$BMI) = "BMI %ile"
-label(df$bmi_raw) = "BMI (kg/m2)"
+label(df$BMI) <- "BMI %ile"
+label(df$bmi_raw) <- "BMI (kg/m2)"
 ```
@@
     FRS_IdealBody, FRS_ThinkyouLook, FRS_Category, Age, Gender, Race,
-    Ethnicity, InsuranceType, BMI, bmi_raw,FEV1
+    Ethnicity, InsuranceType, BMI, bmi_raw, FEV1, BE_APPEARANCE_MEAN,
+    BE_ATTRIBUTION_MEAN, BE_WEIGHT_MEAN
   ) %>%
@@
     type = list(
-      c("FRStotal", "FRS_IdealBody", "FRS_ThinkyouLook","bmi_raw") ~ "continuous2",
+      c("FRStotal", "FRS_IdealBody", "FRS_ThinkyouLook", "bmi_raw") ~ "continuous2",
       all_continuous() ~ "continuous2"

### 2025-07-17 Lilah Melzer/cough_monitor_analysis.qmd
added comorbidities to cough monitor

@@
     nih_sex:nih_ethnicity, mutation1, mutation2,
-    cftr_mod, cftr_mod_type, mucolytic:antih, pulm_ex, contains("micro_type___")
+    cftr_mod, cftr_mod_type, mucolytic:antih, pulm_ex, contains("comorbid_dx"),
+    contains("micro_type___")
   )
@@
     nih_race + nih_ethnicity + cftr_mod + cftr_mod_type + Genotype + pre_fev_pp +
-    total_sleep + pulm_ex + micro_type___1 + micro_type___2 + micro_type___3 +
-    micro_type___4,
+    total_sleep + pulm_ex + comorbid_dx___1 + comorbid_dx___2 +
+    comorbid_dx___3 + comorbid_dx___4 + comorbid_dx___5 + micro_type___1 +
+    micro_type___2 + micro_type___3 + micro_type___4,
   data = demo, control = mycontrols
@@
 fig2
+ggsave(paste0(home_dir, "/Dissemination/Figures/fig2.png"),
+  plot = fig2, width = 9, height = 6, units = "in"
+)
+fig2_color <- ggplot(agg_counts[!is.na(agg_counts$Coughs), ], aes(
+  x = Night, y = Coughs, group = factor(sid), color = Diagnosis
+)) +
+  geom_point(alpha = 0.1) +
+  geom_line(alpha = 0.2, aes(linetype = Diagnosis)) +
+  geom_smooth(aes(group = Diagnosis, linetype = Diagnosis)) +
+  # annotate("rect", xmin = 1, xmax = 29, ymin = 1, ymax = 275, alpha = 0.1) +
+  # annotate("rect", xmin = 41, xmax = 76, ymin = 1, ymax = 275, alpha = 0.1) +
+  ylim(c(1, 275)) +
+  xlim(c(0, 100)) +
+  theme_bw() +
+  scale_color_manual(values = c("#7570B3", "#D95F02")) +
+  scale_y_break(c(80, 265), scales = 0.25, ticklabels = c(265, 270, 275))
+fig2_color
 ggsave(paste0(home_dir, "/Dissemination/Figures/fig2.png"),

### 2025-07-22 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
logistic models for number of exacerbations

--- 2025-07-11 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
@@
 
+```{r}
+# Need to do some re-cleaning for the dataset that includes exacerbations
+# Pregnancies
+preg <- which(encounter_w_hosps$pregnant == 1)
+encounter_w_hosps <- encounter_w_hosps[-preg, ]
+# Days from ETI start
+early <- which(encounter_w_hosps$Days < -365.25 * 3)
+late <- which(encounter_w_hosps$Days > 365.25 * 2)
+encounter_w_hosps <- encounter_w_hosps[-early, ]
+encounter_w_hosps <- encounter_w_hosps[-late, ]
+# Lung function cutoffs
+lower <- 10
+upper <- 150
+# Remove impossible lung function values
+low_fev1 <- which(encounter_w_hosps$gli_fev1_ppred_rn < lower)
+high_fev1 <- which(encounter_w_hosps$gli_fev1_ppred_rn > upper)
+encounter_w_hosps$gli_fev1_ppred_rn[low_fev1] <- NA
+encounter_w_hosps$gli_fev1_ppred_rn[high_fev1] <- NA
+low_fvc <- which(encounter_w_hosps$gli_fvc_ppred_rn < lower)
+high_fvc <- which(encounter_w_hosps$gli_fvc_ppred_rn > upper)
+encounter_w_hosps$gli_fvc_ppred_rn[low_fvc] <- NA
+encounter_w_hosps$gli_fvc_ppred_rn[high_fvc] <- NA
+# Get some summary statistics by participant
+t1_participant <- encounter_w_hosps %>%
+  group_by(eDWID) %>%
+  summarise(
+    `Years With Data` = length(unique(reviewyear)),
+    `Number of Encounters` = n(),
+    `Number of Encounters Pre-ETI` = sum(Days <= 0),
+    `Number of Encounters Post-ETI` = sum(Days > 0),
+    Age = last(na.omit(encounterage[Days <= 0])),
+    across(all_of(continuous_outcomes), \(x) last(na.omit(x[Days <= 0]))),
+    Gender = names(sort(table(Gender), decreasing = TRUE))[1],
+    Race = names(sort(table(Race), decreasing = TRUE))[1],
+    Insurance = names(sort(table(Insurance), decreasing = TRUE))[1],
+    .groups = "drop"
+  )
+no_post <-
+  t1_participant$eDWID[t1_participant$`Number of Encounters Post-ETI` == 0]
+no_post_visits <- which(encounter_w_hosps$eDWID %in% no_post)
+encounter_w_hosps <- encounter_w_hosps[-no_post_visits, ]
+# Pre vs. post variable
+encounter_w_hosps$pre_post <- factor(encounter_w_hosps$Days > 0,
+  levels = c(F, T), labels = c("Pre-ETI", "Post-ETI")
+)
+# Prior modulator use
+prior_mods <- encounter_w_hosps %>%
+  select(eDWID, Days, vx661comb, vx445comb, Vx770, Vx809comb) %>%
+  filter(Days < 0) %>%
+  group_by(eDWID) %>%
+  summarise(prior_modulator = any(vx661comb == 1) | any(vx445comb == 1) |
+    any(Vx770 == 1) | any(Vx809comb))
+prior_mods$prior_modulator <-
+  fct_na_value_to_level(factor(prior_mods$prior_modulator), level = "No")
+prior_mods$prior_modulator <- factor(prior_mods$prior_modulator,
+  levels = c(T, "No"),
+  labels = c("Yes", "No")
+)
+encounter_w_hosps <-
+  left_join(encounter_w_hosps, prior_mods, by = join_by(eDWID))
+# On enzymes for pancreatic insufficiency
+encounter_w_hosps$isOnEnzymes <-
+  factor(encounter_w_hosps$isOnEnzymes, levels = 0:1, labels = c("No", "Yes"))
+# Pseudomonas
+encounter_w_hosps$pseudomonasaeruginosa[
+  is.na(encounter_w_hosps$pseudomonasaeruginosa)
+] <- 0
+encounter_w_hosps$pseudomonasaeruginosa <-
+  factor(encounter_w_hosps$pseudomonasaeruginosa,
+    levels = 0:1, labels = c("No", "Yes")
+  )
+```
+
 ### Full cohort
@@
 # Fit mixed model with change point
-change_mod <- lmer(
-  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + age_group +
+pex_mod <- glmer(
+  PEx ~ Days + Race * pre_post + age_group +
     Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
     (1 | eDWID),
-  data = encounter
+  family = "binomial",
+  data = encounter_w_hosps
 )

### 2025-07-22 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R

@@ -103,19 +103,15 @@ encounter$cultureresults <- factor(encounter$cultureresults,
   labels = c("No growth/sterile culture", "Normal flora", "Microorganisms")
 )
 # Pulmonary exacerbations
-encounter$pe_assessment[encounter$pe_assessment == 5] <- NA
-encounter$PEx <- factor(encounter$pe_assessment,
-  levels = 1:4,
-  labels = c(
-    "Absent", "Mild, moderate, or severe exacerbation",
-    "Mild, moderate, or severe exacerbation",
-    "Mild, moderate, or severe exacerbation"
-  )
+encounter$PEx <- factor(encounter$encounterlocation,
+  levels = c("Clinic", "Hospital", "Home IV", "Other", "5", "6"),
+  labels = c("No PEx", "PEx", "PEx", "No PEx", "No PEx", "No PEx")
 )
 # Merge
 encounter <- left_join(
   encounter,
-  demo %>% select(eDWID, Race, Gender, Modulator_trikafta_first_date)
+  demo %>% select(eDWID, Race, Gender, Modulator_trikafta_first_date),
+  by = join_by(eDWID)
 )
 encounter <- left_join(encounter,
   annual %>% select(eDWID, ReviewYear, Insurance, pregnant),
@@ -145,7 +141,7 @@ hospitalizations <- hospitalizations %>%
     encounterdate = seq(CareEpi_StartDt, CareEpi_EndDt, by = "day")
   ) %>%
   mutate(hospitalized = "Yes")
-encounter <- 
+encounter <-
   left_join(encounter, hospitalizations, by = join_by(eDWID, encounterdate))
 # Variables for flowchart
 n_enc_1 <- nrow(encounter)
@@ -160,6 +156,7 @@ n_people_2 <- length(unique(encounter$eDWID))
 hosp <- which(encounter$hospitalized == "Yes" |
   encounter$encounterlocation %in% c("Hospital", "Home IV"))
 n_people_hosp <- length(unique(encounter$eDWID[hosp]))
+encounter_w_hosps <- encounter
 encounter <- encounter[-hosp, ]
 n_enc_3 <- nrow(encounter)
 n_people_3 <- length(unique(encounter$eDWID))
@@ -287,6 +284,6 @@ flow_chart <- tibble(
   )
 )
 # Save
-save(encounter, t1_participant, flow_chart, continuous_outcomes,
[…]

### 2025-07-23 Christine Chan/EnVision CF/fsOGTT_pre_post_ETI.qmd
both study IDs for EnVision and GEM fsOGTT

@@
 df <- df %>%
+  mutate(gem_id = NA, study = "EnVision CF") %>%
   select(
-    study_id, Date, age_visit, sex, origin_race, ethnicity, CFTR,
+    study_id, gem_id, study, Date, age_visit, sex, origin_race, ethnicity, CFTR,
     pancreatic_status, corrector___4, pre_post, bmi, fev1,
@@
 )
-df$pancreatic_status <- factor(df$pancreatic_status, levels = 1:2)
+df$pancreatic_status <- factor(df$pancreatic_status,
+  levels = 1:2,
+  labels = c("Sufficient", "Insufficient")
+)
 df$ETI <- factor(df$ETI, levels = 0:1, labels = c("No", "Yes"))
@@
 gem <- gem %>%
-  mutate(ETI = 0, pre_post = "Pre") %>%
+  mutate(ETI = 0, pre_post = "Pre", study_id = NA, study = "GEM-CF") %>%
   select(
-    subject_id, date_of_visit, age_at_visit_1, gender, race_ethnicity,
-    cf_genotype, cf_pancreatic, ETI, pre_post, average_bmi, fev1_percent,
-    group_category, lab_ogtt_fasting, min_10_glucose, min_30_glucose,
-    lab_ogtt_1_hour_glucose, min_90_glucose, lab_ogtt_2_hour_glucose,
-    min_150_glucose, min_180_glucose, min_0_insulin, min_10_insulin,
-    min_30_insulin, min_60_insulin, min_90_insulin, min_120_insulin,
-    min_150_insulin, min_180_insulin, min_0_c_peptide, min_10_c_peptide,
-    min_30_c_peptide, min_60_c_peptide, min_90_c_peptide, min_120_c_peptide,
-    min_150_c_peptide, min_180_c_peptide
+    study_id, subject_id, study, date_of_visit, age_at_visit_1, gender,
+    race_ethnicity, cf_genotype, cf_pancreatic, ETI, pre_post, average_bmi,
+    fev1_percent, group_category, lab_ogtt_fasting, min_10_glucose,
+    min_30_glucose, lab_ogtt_1_hour_glucose, min_90_glucose,
+    lab_ogtt_2_hour_glucose, min_150_glucose, min_180_glucose,
+    min_0_insulin, min_10_insulin, min_30_insulin, min_60_insulin,
+    min_90_insulin, min_120_insulin, min_150_insulin, min_180_insulin,
+    min_0_c_peptide, min_10_c_peptide, min_30_c_peptide, min_60_c_peptide,
+    min_90_c_peptide, min_120_c_peptide, min_150_c_peptide, min_180_c_peptide
   ) %>%
   rename(
-    study_id = subject_id, Date = date_of_visit, age_visit = age_at_visit_1,
+    gem_id = subject_id, Date = date_of_visit, age_visit = age_at_visit_1,
     sex = gender, CFTR = cf_genotype, pancreatic_status = cf_pancreatic,
@@
 )
-gem$pancreatic_status <- factor(gem$pancreatic_status, levels = 1:2)
+gem$pancreatic_status <- factor(gem$pancreatic_status,
+  levels = 1:2,
+  labels = c("Sufficient", "Insufficient")
+)
 gem$ETI <- factor(gem$ETI, levels = 0:1, labels = c("No", "Yes"))
@@
 )
+# IDs
+gem$study_id <- ids$`EnVision ID`[match(gem$gem_id, ids$`GEM ID`)]
+df$gem_id <- ids$`GEM ID`[match(df$study_id, ids$`EnVision ID`)]
 # Limit GEM per Monique's list
-gem <- gem %>% filter(study_id %in% ids$`GEM ID`)
-# Convert to EnVision IDs
-gem$study_id <- ids$`EnVision ID`[match(gem$study_id, ids$`GEM ID`)]
-# Merge, keep only pre/post
-df <- full_join(df, gem) %>% filter(!is.na(pre_post))
+gem <- gem %>% filter(gem_id %in% ids$`GEM ID`)
+# Merge, keep only pre/post, sort by ID and date
+df <- rbind(df, gem) %>% filter(!is.na(pre_post))
 df$pre_post <- factor(df$pre_post, levels = c("Pre", "Post"))
+df <- df %>% arrange(study_id, Date)
 # Write for checking

### 2025-07-24 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
analysis with <= Cl excluded

@@
 )
+# Additional changes based on Angela's review
+df$Diagnosis[which(is.na(df$Name))] <- "Cystic Fibrosis"
+df$Diagnosis[400] <- "Cystic Fibrosis"
+df$Na[541] <- 69
+# Angela reviewed a random set of 75 patients and requested the changes below
+# See the "random_pull.csv" document for details
+# Changes done by row number to avoid PHI in GitHub
+df$Cl[1473] <- 103.5
+df$Na[505] <- 42.6
+df$Na[238] <- 73
+df$Na[541] <- 69
+df$Na[747] <- 88.9
+# Remove people per Angela's recommendation and duplicates based on DOB review
+df <- df[-c(
+  78, 79, 183, 257, 259, 270, 314, 342, 691, 719, 724, 737, 812, 903, 1012,
+  1066, 1126, 1151, 1202, 1209, 1222, 1224, 1340, 1428, 1456
+), ]
 # Set table options
@@
 )
-# Additional lists for Angela
-write.csv(
-  df %>%
-    filter(Diagnosis == "Non-CF" & Cl >= 60),
-  file = "./Data_Cleaned/sweat_variability_non_cf_cl_60.csv",
-  row.names = F, na = ""
-)
-write.csv(
-  df %>%
-    filter(Diagnosis == "Cystic Fibrosis" & `Na:Cl` >= 1.2),
-  file = "./Data_Cleaned/sweat_variability_cf_nacl_1_2.csv",
-  row.names = F, na = ""
-)
-write.csv(
-  df %>%
-    filter(Diagnosis == "Non-CF" & `Na:Cl` < 1.2),
-  file = "./Data_Cleaned/sweat_variability_non_cf_nacl_1_2.csv",
-  row.names = F, na = ""
-)
 ```
 
-# Data cleaning steps
+# Full Cohort
 
-- Imported re-identified CHCO data from Elin (5,046 rows and 4,748 unique IDs)
-- Removed rows with `Demographics Available` not equal to `TRUE` and selected each person's earliest test by age.
-  - This resulted in 3,123 rows with 3,084 unique IDs.
-- Merged in the diagnoses from Angela.
-- Values below the limit of detection were replaced with half the limit (e.g. "LESS THAN 10" was replaced with 9). Approximately 5% of the values were below the L.O.D.
-
-- Imported NJH data (1,958 rows with 1,748 unique IDs) and merged in diagnoses from Katie.
-- Removed 1,662 rows with `Demographics Available` not equal to `TRUE` (all of these were `FALSE`)
-  - This resulted in 172 rows with unique IDs.
-- Concatenated the CHCO and NJH data and removed duplicate rows, resulting in 3,178 rows with 3,162 unique IDs.
-- For the remaining people with multiple sweat tests (i.e. tests from both CHCO and NJH), kept only the earliest test (based on age at test). This resulted in 3,169 rows with 3,162 unique IDs.
-- The remaining 7 people with different values for the same test in the CHCO and NJH data were excluded. The final analysis dataset includes 3,155 unique participants, 3,111 of which have a diagnosis.
-- For people with both left and right arm values, the two were averaged.
-
-# Table 1: Participant Demographics by Diagnosis
+## Table 1: Participant Demographics by Diagnosis
 
@@
 
-# Boxplots
+## Boxplots
 
-## Chloride
+### Chloride
 
@@
 
-## Sodium
+### Sodium
 
@@
 
-## Na:Cl
+### Na:Cl
 
@@
 
-# Histograms and Density Plots
+## Histograms and Density Plots
 
-## Chloride
+### Chloride
 
@@
 
-# Na:Cl ROC Curves
+## Na:Cl ROC Curves
 
@@
 - The multiclass AUC for Na:Cl is `r round(as.numeric(auc(mc_roc)),3)`.
+
+### Individual ROC Curves
+
+```{r}
+plot(roc_list[[2]],
+  print.auc = T, print.thres = T,
+  main = "Cystic Fibrosis vs. Non-CF"
+)
+plot(roc_list[[1]],
+  print.auc = T, print.thres = T,
+  main = "CFTR-RD or CRMS vs. Cystic Fibrosis"
+)
+plot(roc_list[[3]],
+  print.auc = T, print.thres = T,
+  main = "CFTR-RD or CRMS vs. Non-CF"
+)
+```
+
+# Excluding Cl $\leq$ 30
+
+```{r}
+df <- df %>% filter(Cl > 30)
+```
+
+## Table 1: Participant Demographics by Diagnosis
+
+```{r results='asis'}
+summary(tableby(Diagnosis ~ Sex + Age + Na + Cl + `Na:Cl` + K,
+  data = df, control = mycontrols
+))
+```
+
+## Boxplots
+
+### Chloride
+
+```{r}
+#| warning: false
+clp <- ggplot(df, aes(x = Diagnosis, y = Cl)) +
+  geom_boxplot() +
+  theme_bw()
+clp
+```
+
+### Sodium
+
+```{r}
+#| warning: false
+nap <- ggplot(df, aes(x = Diagnosis, y = Na)) +
+  geom_boxplot() +
+  theme_bw()
+nap
+```
+
+### Na:Cl
+
+```{r}
+#| warning: false
+ratiop <- ggplot(df, aes(x = Diagnosis, y = `Na:Cl`)) +
+  geom_boxplot() +
+  geom_hline(yintercept = 1, color = "red") +
+  geom_hline(yintercept = 1.25, color = "red") +
+  theme_bw() +
+  ylab("Na:Cl") +
+  xlab("Modulator")
+ratiop
+ratiop <- ggplot(df, aes(x = Diagnosis, y = `Na:Cl`)) +
+  geom_boxplot() +
+  geom_hline(yintercept = 1, color = "red") +
+  geom_hline(yintercept = 1.25, color = "red") +
+  ylim(0, 2) +
+  theme_bw() +
+  ylab("Na:Cl") +
+  xlab("Modulator")
+ratiop
+```
+
+## Histograms and Density Plots
+
+### Chloride
+
+```{r}
+#| warning: false
+ggplot(df, aes(x = Cl, fill = Diagnosis)) +
+  geom_histogram(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
+  geom_histogram(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
+  geom_histogram(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
+  theme_bw() +
+  theme(axis.title.y = element_blank())
+ggplot(df, aes(x = Cl, fill = Diagnosis)) +
+  geom_density(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
+  geom_density(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
+  geom_density(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
+  theme_bw() +
+  theme(axis.title.y = element_blank())
+```
+
+### Sodium
+
+```{r}
+#| warning: false
+ggplot(df, aes(x = Na, fill = Diagnosis)) +
+  geom_histogram(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
+  geom_histogram(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
+  geom_histogram(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
+  theme_bw() +
+  theme(axis.title.y = element_blank())
+ggplot(df, aes(x = Na, fill = Diagnosis)) +
+  geom_density(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
+  geom_density(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
+  geom_density(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
+  theme_bw() +
+  theme(axis.title.y = element_blank())
+```
+
+### Na:Cl
+
+```{r}
+#| warning: false
+ggplot(df, aes(x = `Na:Cl`, fill = Diagnosis)) +
+  geom_histogram(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
+  geom_histogram(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
+  geom_histogram(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
+  theme_bw() +
+  theme(axis.title.y = element_blank())
+ggplot(df, aes(x = `Na:Cl`, fill = Diagnosis)) +
+  geom_density(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
+  geom_density(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
+  geom_density(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
+  theme_bw() +
+  theme(axis.title.y = element_blank())
+ggplot(df, aes(x = `Na:Cl`, fill = Diagnosis)) +
+  geom_density(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
+  geom_density(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
+  geom_density(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
+  xlim(0, 2) +
+  theme_bw() +
+  theme(axis.title.y = element_blank())
+```
+
+## Na:Cl ROC Curves
+
+```{r}
+# Fit a multiclass ROC
+mc_roc <- multiclass.roc(df$Diagnosis, df$`Na:Cl`, quiet = T)
+roc_list <- list(
+  "Cystic Fibrosis vs. Non-CF" = mc_roc$rocs[[2]],
+  "CFTR-RD or CRMS vs. Cystic Fibrosis" = mc_roc$rocs[[1]],
+  "CFTR-RD or CRMS vs. Non-CF" = mc_roc$rocs[[3]]
+)
+ggroc(roc_list, legacy.axes = T) +
+  theme_bw() +
+  theme(legend.title = element_blank())
+roc_table <- lapply(roc_list, function(r) {
+  coords(r, "best",
+    best.method = "youden",
+    ret = c(
+      "threshold", "sensitivity", "specificity", "accuracy", "tpr", "fpr",
+      "tnr", "fnr"
+    )
+  )
+})
+roc_table <- do.call(rbind, roc_table)
+kable(roc_table, digits = 3)
+```
+
+- The multiclass AUC for Na:Cl is `r round(as.numeric(auc(mc_roc)),3)`.
+
+### Individual ROC Curves
+
+```{r}
+plot(roc_list[[2]],
+  print.auc = T, print.thres = T,
+  main = "Cystic Fibrosis vs. Non-CF"
+)
+plot(roc_list[[1]],
+  print.auc = T, print.thres = T,
+  main = "CFTR-RD or CRMS vs. Cystic Fibrosis"
+)
+plot(roc_list[[3]],
+  print.auc = T, print.thres = T,
+  main = "CFTR-RD or CRMS vs. Non-CF"
+)
+```

### 2025-07-25 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
PEx models FINALLY working

--- 2025-07-11 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
@@
 library(broom)
+library(emmeans)
 knitr::opts_knit$set(root.dir = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF")
@@
 
+For some reason, exacerbation models with the encounter-level data are just not working.
+I think it may be because there are often several pulmonary exacerbation visits very close together, so including the "Days" variable is causing problems.
+In the end I used a very simple logistic regression model on annualized PEx data.
+See `/home/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R` for details on making the exacerbation dataset.
+
 ### Full cohort
@@
 ```{r}
-# Fit mixed model with change point
-change_mod <- lmer(
-  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + age_group +
-    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
-    (1 | eDWID),
-  data = encounter
-)
-# Model plot
-pop_dat <- data.frame(expand.grid(
-  Years = -3:2, Race = unique(encounter$Race), age_group = "[18,30)",
-  Insurance = "Private", prior_modulator = "Yes",
-  isOnEnzymes = "Yes", pseudomonasaeruginosa = "No"
-))
-pop_dat$slope_change_years <- (pop_dat$Years > 0) * pop_dat$Years
-pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
-pop_dat$pred <- pred$fit
-pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
-pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
-ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
-  geom_line() +
-  xlab("Years from ETI Start") +
-  ylab("gli_fev1_ppred_rn") +
-  theme_classic()
-# Model summary
-change_mod %>%
-  gtsummary::tbl_regression(intercept = T) %>%
-  as_gt() %>%
-  fmt_number(decimals = 3)
-# Table of predicted values
-pop_dat %>%
-  select(-slope_change_years) %>%
-  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
-  gt(caption = "Predicted Values at Various Timepoints") %>%
-  fmt_number(decimals = 3)
-# Contrasts
-aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
-aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
-hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
-hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
-nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
-nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
-other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
-other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
-contr <- rbind(
-  "Black/AA Pre-ETI" = aa_pre,
-  "Black/AA Post-ETI" = aa_pre + aa_change,
-  "H/L Pre-ETI" = hl_pre,
-  "H/L Post-ETI" = hl_pre + hl_change,
-  "NHW Pre-ETI" = nhw_pre,
-  "NHW Post-ETI" = nhw_pre + nhw_change,
-  "Other Pre-ETI" = other_pre,
-  "Other Post-ETI" = other_pre + other_change
-)
-change_mod_slopes <- glht(change_mod, linfct = contr)
-change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
-change_mod_slopes_ci$confint %>%
-  as.data.frame() %>%
-  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
-  fmt_number(decimals = 3)
-```
+# Zero-inflated models
+# mod_p <- mixed_model(num_pex ~ ETI * Race,
+#   data = pex_data,
+#   random = ~ 1 | eDWID, zi_random = ~ 1 | eDWID,
+#   family = zi.poisson(), zi_fixed = ~1
+# )
+# mod_nb <- mixed_model(num_pex ~ ETI * Race,
+#   data = pex_data,
+#   random = ~ 1 | eDWID, zi_random = ~ 1 | eDWID,
+#   family = zi.negative.binomial(), zi_fixed = ~1
+# )
+pex_logistic <- glmer(any_pex ~ ETI * Race + (1 | eDWID),
+  data = pex_data,
+  control = glmerControl(optimizer = "bobyqa"),
+  family = "binomial"
+)
+means <- data.frame(emmeans(pex_logistic, ~ Race | ETI, type = "response"))
+means %>%
+  gt(caption = "Probabilities by Race", rownames_to_stub = T) %>%
+  fmt_number(decimals = 3)
+```
+
+The table above shows the predicted probability of a pulmonary exacerbation pre- and post-ETI.
+So for example, NHW pwCF had a PEx probability of `r round(means$prob[means$Race=="Non-Hispanic White"&means$ETI=="Pre-ETI"]*100,1)` (95%CI: `r round(means$asymp.LCL[means$Race=="Non-Hispanic White"&means$ETI=="Pre-ETI"]*100,1)` - `r round(means$asymp.UCL[means$Race=="Non-Hispanic White"&means$ETI=="Pre-ETI"]*100,1)`)  pre-ETI and `r round(means$prob[means$Race=="Non-Hispanic White"&means$ETI=="Post-ETI"]*100,1)` (95%CI: `r round(means$asymp.LCL[means$Race=="Non-Hispanic White"&means$ETI=="Post-ETI"]*100,1)` - `r round(means$asymp.UCL[means$Race=="Non-Hispanic White"&means$ETI=="Post-ETI"]*100,1)` post-ETI.
+Conversely, Black pwCF had a PEx probability of `r round(means$prob[means$Race=="Black or African American"&means$ETI=="Pre-ETI"]*100,1)` (95%CI: `r round(means$asymp.LCL[means$Race=="Black or African American"&means$ETI=="Pre-ETI"]*100,1)` - `r round(means$asymp.UCL[means$Race=="Black or African American"&means$ETI=="Pre-ETI"]*100,1)`)  pre-ETI and `r round(means$prob[means$Race=="Black or African American"&means$ETI=="Post-ETI"]*100,1)` (95%CI: `r round(means$asymp.LCL[means$Race=="Black or African American"&means$ETI=="Post-ETI"]*100,1)` - `r round(means$asymp.UCL[means$Race=="Black or African American"&means$ETI=="Post-ETI"]*100,1)` post-ETI.
 
@@
 
-The following models were not adjusted for age group.
-
-```{r}
-# Fit mixed model with change point
-change_mod <- lmer(
-  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
-    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
-  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
-)
-# Model plot
-pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
-pop_dat$pred <- pred$fit
-pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
-pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
-ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
-  geom_line() +
-  xlab("Years from ETI Start") +
-  ylab("gli_fev1_ppred_rn") +
-  theme_classic()
-# Model summary
-change_mod %>%
-  gtsummary::tbl_regression(intercept = T) %>%
-  as_gt() %>%
-  fmt_number(decimals = 3)
-# Table of predicted values
-pop_dat %>%
-  select(-slope_change_years) %>%
-  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-# Contrasts
-aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
-aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
-hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
-hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
-nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
-nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
-other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
-other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
-contr <- rbind(
-  "Black/AA Pre-ETI" = aa_pre,
-  "Black/AA Post-ETI" = aa_pre + aa_change,
-  "H/L Pre-ETI" = hl_pre,
-  "H/L Post-ETI" = hl_pre + hl_change,
-  "NHW Pre-ETI" = nhw_pre,
-  "NHW Post-ETI" = nhw_pre + nhw_change,
-  "Other Pre-ETI" = other_pre,
-  "Other Post-ETI" = other_pre + other_change
-)
-change_mod_slopes <- glht(change_mod, linfct = contr)
-change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
-change_mod_slopes_ci$confint %>%
-  as.data.frame() %>%
-  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
+```{r}
+pex_logistic <- glmer(any_pex ~ ETI * Race + (1 | eDWID),
+  data = pex_data[pex_data$age_eti_group == "[-Inf,18)", ],
+  control = glmerControl(optimizer = "bobyqa"),
+  family = "binomial"
+)
+means <- data.frame(emmeans(pex_logistic, ~ Race | ETI, type = "response"))
+means %>%
+  gt(caption = "Probabilities by Race", rownames_to_stub = T) %>%
   fmt_number(decimals = 3)
@@
 
-The following models were not adjusted for age group.
-
-```{r}
-# Fit mixed model with change point
-change_mod <- lmer(
-  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
-    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
-  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
-)
-# Model plot
-pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
-pop_dat$pred <- pred$fit
-pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
-pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
-ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
-  geom_line() +
-  xlab("Years from ETI Start") +
-  ylab("gli_fev1_ppred_rn") +
-  theme_classic()
-# Model summary
-change_mod %>%
-  gtsummary::tbl_regression(intercept = T) %>%
-  as_gt() %>%
-  fmt_number(decimals = 3)
-# Table of predicted values
-pop_dat %>%
-  select(-slope_change_years) %>%
-  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
-  gt() %>%
-  fmt_number(decimals = 3)
-# Contrasts
-aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
-aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
-hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
-hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
-nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
-nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
-other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
-other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
-contr <- rbind(
-  "Black/AA Pre-ETI" = aa_pre,
-  "Black/AA Post-ETI" = aa_pre + aa_change,
-  "H/L Pre-ETI" = hl_pre,
-  "H/L Post-ETI" = hl_pre + hl_change,
-  "NHW Pre-ETI" = nhw_pre,
-  "NHW Post-ETI" = nhw_pre + nhw_change,
-  "Other Pre-ETI" = other_pre,
-  "Other Post-ETI" = other_pre + other_change
-)
-change_mod_slopes <- glht(change_mod, linfct = contr)
-change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
-change_mod_slopes_ci$confint %>%
-  as.data.frame() %>%
-  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
+```{r}
+pex_logistic <- glmer(any_pex ~ ETI * Race + (1 | eDWID),
+  data = pex_data[pex_data$age_eti_group == "[18, Inf)", ],
+  control = glmerControl(optimizer = "bobyqa"),
+  family = "binomial"
+)
+means <- data.frame(emmeans(pex_logistic, ~ Race | ETI, type = "response"))
+means %>%
+  gt(caption = "Probabilities by Race", rownames_to_stub = T) %>%
   fmt_number(decimals = 3)

### 2025-07-25 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R

@@ -102,11 +102,6 @@ encounter$cultureresults <- factor(encounter$cultureresults,
   levels = 1:3,
   labels = c("No growth/sterile culture", "Normal flora", "Microorganisms")
 )
-# Pulmonary exacerbations
-encounter$PEx <- factor(encounter$encounterlocation,
-  levels = c("Clinic", "Hospital", "Home IV", "Other", "5", "6"),
-  labels = c("No PEx", "PEx", "PEx", "No PEx", "No PEx", "No PEx")
-)
 # Merge
 encounter <- left_join(
   encounter,
@@ -143,6 +138,31 @@ hospitalizations <- hospitalizations %>%
   mutate(hospitalized = "Yes")
 encounter <-
   left_join(encounter, hospitalizations, by = join_by(eDWID, encounterdate))
+encounter$hospitalized[is.na(encounter$hospitalized)] <- "No"
+# Make pulmonary exacerbations dataset - models aren't working with
+# encounter-level data so use annualized info. For counting the number of
+# PEx in a year, look for "runs" of hospitalization and/or home IVs.
+pex_data <- encounter %>%
+  mutate(PEx = encounterlocation %in% c("Hospital", "Home IV") | hospitalized == "Yes") %>%
+  group_by(eDWID, reviewyear) %>%
+  summarise(
+    Race = last(Race),
+    age_eti_group = last(age_eti_group),
+    # Were they on ETI at their last visit of the year?
+    ETI = factor(sign(max(Days)),
+      levels = c(-1:1),
+      labels = c("Pre-ETI", "Pre-ETI", "Post-ETI")
+    ),
+    # Count PEx runs
+    num_pex = sum(rle(PEx)$values == "TRUE"),
+    .groups = "drop"
+  )
+pex_data$any_pex <- pex_data$num_pex > 0
+# Show this to Scott and Kam
+check <- encounter %>%
+  mutate(PEx = encounterlocation %in% c("Hospital", "Home IV") | hospitalized == "Yes") %>%
+  select(eDWID, encounterdate, encounterlocation, hospitalized, PEx) %>%
+  filter(eDWID == "900060748")
 # Variables for flowchart
 n_enc_1 <- nrow(encounter)
 n_people_1 <- length(unique(encounter$eDWID))
@@ -156,7 +176,6 @@ n_people_2 <- length(unique(encounter$eDWID))
[…]

### 2025-07-25 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
removing sweat Cl > 160
fixed ROC labels

@@
 df$K <- coalesce(df$K, df$K_mean)
-# Change Cl under 10 to missing
+# Change Cl under 10 or over 160 to missing
 df$Cl[df$Cl < 10] <- NA
+df$Cl[df$Cl > 160] <- NA
+# Remove Na under 10 as well
 df$Na[df$Na < 10] <- NA
@@
   print.auc = T, print.thres = T,
-  main = "Cystic Fibrosis vs. Non-CF"
+  main = "Cystic Fibrosis vs. CFTR-RD or CRMS"
 )
@@
   print.auc = T, print.thres = T,
-  main = "CFTR-RD or CRMS vs. Cystic Fibrosis"
+  main = "Cystic Fibrosis vs. Non-CF"
 )

### 2025-07-28 Lilah Melzer/cough_monitor_analysis.qmd
additional cough monitor figures

@@
 
+## 38
+
+Participant 38 did not have a pulmonary exacerbation.
+
 ```{r}
@@
 ```
+
+## 2
+
+Participant 2 had a pulmonary exacerbation starting on night 41.
+
+```{r}
+#| warning: false
+p2 <- full_join(
+  agg_counts,
+  cfqr_children_6_11 %>% select(sid, Date, respirat) %>%
+    rename(`CFQ-R Children Ages 6-11` = respirat),
+  by = join_by(sid, Date)
+)
+p2 <- full_join(
+  p2,
+  cfqr_parent_6_13 %>% select(sid, Date, respirat) %>%
+    rename(`CFQ-R Caregivers 6-13` = respirat),
+  by = join_by(sid, Date)
+)
+p2 <- p2 %>%
+  filter(sid == 2) %>%
+  pivot_longer(c(Coughs, `CFQ-R Children Ages 6-11`, `CFQ-R Caregivers 6-13`)) %>%
+  filter(!is.na(value))
+p2$name <- factor(p2$name,
+  levels = c("Coughs", "CFQ-R Children Ages 6-11", "CFQ-R Caregivers 6-13")
+)
+p2_plot <- ggplot(p2, aes(x = Night, y = value)) +
+  geom_point() +
+  geom_line() +
+  theme_bw() +
+  theme(axis.title.y = element_blank()) +
+  facet_wrap(~name, nrow = 3, scales = "free")
+p2_plot
+ggsave(paste0(home_dir, "/Dissemination/Figures/participant_2.png"),
+  plot = p2_plot, width = 9, height = 7, units = "in"
+)
+```
+
+## 3
+
+Participant 3 had a pulmonary exacerbation starting on night 1.
+
+```{r}
+#| warning: false
+cfqr_children_12_13$sid <- as.numeric(cfqr_children_12_13$sid)
+p3 <- full_join(
+  agg_counts,
+  cfqr_children_12_13 %>% select(sid, Date, respirat) %>%
+    rename(`CFQ-R Children Ages 12-13` = respirat),
+  by = join_by(sid, Date)
+)
+p3 <- full_join(
+  p3,
+  cfqr_parent_6_13 %>% select(sid, Date, respirat) %>%
+    rename(`CFQ-R Caregivers 6-13` = respirat),
+  by = join_by(sid, Date)
+)
+p3 <- p3 %>%
+  filter(sid == 3) %>%
+  pivot_longer(c(Coughs, `CFQ-R Children Ages 12-13`, `CFQ-R Caregivers 6-13`)) %>%
+  filter(!is.na(value))
+p3$name <- factor(p3$name,
+  levels = c("Coughs", "CFQ-R Children Ages 12-13", "CFQ-R Caregivers 6-13")
+)
+p3_plot <- ggplot(p3, aes(x = Night, y = value)) +
+  geom_point() +
+  geom_line() +
+  theme_bw() +
+  theme(axis.title.y = element_blank()) +
+  facet_wrap(~name, nrow = 3, scales = "free")
+p3_plot
+ggsave(paste0(home_dir, "/Dissemination/Figures/participant_3.png"),
+  plot = p3_plot, width = 9, height = 7, units = "in"
+)
+```

### 2025-07-28 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
updated GCT figures

@@
 # Dotplot
-dotplot <- ggplot(df, aes(x = Diagnosis, y = `GCT 60 min plasma`)) +
+dotplot <- ggplot(df, aes(
+  x = Diagnosis, y = `GCT 60 min plasma`,
+  colour = Diagnosis
+)) +
   geom_point() +
-  geom_hline(yintercept = 140, linetype = "dashed") +
-  geom_hline(yintercept = 150, linetype = "dashed") +
+  geom_hline(yintercept = 130, linetype = "dashed") +
   geom_hline(yintercept = 160, linetype = "dashed") +
@@
   theme_bw() +
-  scale_y_continuous(breaks = c(100, 140, 147, 150, 160))
-
+  scale_y_continuous(breaks = c(100, 130, 147, 160)) +
+  scale_color_manual(values = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51"))
 dotplot
-dotplot +
-  theme(text = element_blank())
+ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/dotplot.png",
+  dotplot,
+  height = 4, width = 6, units = "in", dpi = 600,
+  device = "png"
+)
 # Additional boxplots/dotplot per Christine and Larry
@@
 # Dotplot
-dot_p1 <- ggplot(
+dot_rpg <- ggplot(
   plot_df[plot_df$Measure == "GCT 0 min plasma", ],
@@
   theme(
-    legend.position = "none", plot.title = element_text(hjust = 0.5),
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.title.x = element_blank()
+    plot.title = element_text(hjust = 0.5),
+    axis.title.x = element_blank(), axis.text.x = element_blank(),
+    axis.ticks.x = element_blank()
   )
-dot_p2 <- ggplot(
+dot_gctpl <- ggplot(
   plot_df[plot_df$Measure == "GCT 60 min plasma", ],
@@
   theme(
-    legend.position = "none", plot.title = element_text(hjust = 0.5),
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.title.x = element_blank(),
-    axis.title.y = element_blank()
+    plot.title = element_text(hjust = 0.5),
+    axis.title.y = element_blank(), axis.text.y = element_blank(),
+    axis.title.x = element_blank(), axis.text.x = element_blank(),
+    axis.ticks.x = element_blank()
   )
-dot_p3 <- ggplot(
+dot_rcg <- ggplot(
   plot_df[plot_df$Measure == "GCT 0 min capillary", ],
@@
     legend.position = "none", plot.title = element_text(hjust = 0.5),
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.title.x = element_blank(),
-    axis.title.y = element_blank()
+    axis.title.y = element_blank(), axis.text.y = element_blank(),
+    axis.title.x = element_blank(), axis.text.x = element_blank(),
+    axis.ticks.x = element_blank()
   )
-dot_p4 <- ggplot(
+dot_gctcap <- ggplot(
   plot_df[plot_df$Measure == "GCT 60 min capillary", ],
@@
   theme(
-    legend.position = "none", plot.title = element_text(hjust = 0.5),
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.title.x = element_blank(),
-    axis.title.y = element_blank()
+    plot.title = element_text(hjust = 0.5),
+    axis.title.y = element_blank(), axis.text.y = element_blank(), 
+    axis.ticks.y = element_blank(),
+    axis.title.x = element_blank(), axis.text.x = element_blank(),
+    axis.ticks.x = element_blank()
   )
-dot_p5 <- ggplot(
+dot_hba1c <- ggplot(
   plot_df[plot_df$Measure == "HbA1c", ],
@@
     legend.title = element_blank(), plot.title = element_text(hjust = 0.5),
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.title.x = element_blank()
+    axis.text.x = element_blank(),
+    axis.title.x = element_blank(), axis.ticks.x = element_blank()
   )
-dot_p <- dot_p1 + dot_p2 + dot_p3 + dot_p4 + dot_p5 + plot_layout(ncol = 5)
+dot_p <- dot_rpg + dot_rcg + dot_gctpl + dot_gctcap + dot_hba1c + 
+  plot_layout(ncol = 5,guides = 'collect')
 ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/dotplot2.png",
-  box_p,
+  dot_p,
+  height = 9, width = 12, units = "in", dpi = 600,
   device = "png"

### 2025-07-28 Christine Chan/EnVision CF/cgm_and_a1c.qmd
Envision CGM and HbA1c

---
title: "The relationship between HbA1c, GMI, and average glucose in adults with CFRD"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(redcapAPI)
library(gtsummary)
library(labelled)
library(ggpmisc)
library(blandr)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/tim/OneDrive/Vigers/CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| include: false
# Import from REDCap
unlockREDCap(c(rcon = "EnVision CF HbA1c and CGM"),
  keyring = "API_KEYs",
  envir = 1,
  url = "https://redcap.ahc.umn.edu/redcap/api/"
)
df <- exportRecordsTyped(rcon)
# Start building analysis dataset
df <- df %>%
  # Format columns
  mutate(
    # Numeric
    across(c(age, bmi, cfrd_diagnosis_age, average_sensor_glucose), ~
      as.numeric(.)),
    across(c(hb_a1c_value, gmi), ~
      as.numeric(sub("%", "", .))),
    # Categorical
    combined_race = factor(race,
      levels = c(
        "White", "Black or African American", "Asian", "More Than One Race",
        "Unknown / Not Reported"
      ),
      labels = c(
        "White", "Black or African American", "Other/Unknown", "Other/Unknown",
        "Other/Unknown"
      )
    )
  ) %>%
  # Drop unnecessary columns
  select(
    subject_id, age, cfrd_diagnosis_age, bmi, combined_race, sex, pi,
    cftr_modulator_use, type_of_modulator, type_of_cgm, hb_a1c_value, gmi,
    average_sensor_glucose
  ) %>%
  # Labels
  set_variable_labels(
    age = "Age (years)", cfrd_diagnosis_age = "Age at CFRD Diagnosis (years)",
    bmi = "BMI (kg/m2)", combined_race = "Race", sex = "Sex at Birth",
    type_of_modulator = "Type of Modulator", hb_a1c_value = "HbA1c (%)",
    gmi = "GMI (%)", average_sensor_glucose = "Mean Sensor Glucose (mg/dL)",
    cftr_modulator_use = "CFTR Modulator", pi = "Pancreatic Insufficiency"
  )
```

# Participant Characteristics

```{r}
#| label: tbl-1
#| tbl-cap: Table 1
df %>%
  select(-subject_id) %>%
  tbl_summary(
    missing_text = "Missing",
    type = all_continuous() ~ "continuous2",
    statistic = list(
      all_continuous() ~ c("{median} [{p25}, {p75}]", "{min} - {max}"),
      average_sensor_glucose ~ c("{mean} ({sd})", "{min} - {max}")
    ),
  )
```

# GMI and HbA1c

```{r}
#| warning: false
ggplot(df, aes(x = hb_a1c_value, y = gmi)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2", "p", "n")))

blandr.draw(df$hb_a1c_value, df$gmi, ciDisplay = F, annotate = T)
```

# GMI and HbA1c

```{r}
#| warning: false
ggplot(df, aes(x = hb_a1c_value, y = average_sensor_glucose)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2", "p", "n")))

blandr.draw(df$hb_a1c_value, df$average_sensor_glucose,
  ciDisplay = F, annotate = T
)
```

### 2025-07-28 Quarto Template.qmd
Envision CGM and HbA1c

@@
 library(redcapAPI)
+library(gtsummary)
+library(labelled)
 home_dir <- switch(Sys.info()["sysname"],

### 2025-07-28 Lilah Melzer/cough_monitor_analysis.qmd
cough monitor figure 2 idea

--- 2025-07-17 Lilah Melzer/cough_monitor_analysis.qmd
@@
   geom_point(alpha = 0.1) +
-  geom_line(alpha = 0.2, aes(linetype = Diagnosis)) +
+  geom_point(data = agg_counts[agg_counts$Sick == "Yes", ], color = "red") +
   geom_smooth(color = "black", aes(group = Diagnosis, linetype = Diagnosis)) +
-  annotate("rect", xmin = 1, xmax = 29, ymin = 1, ymax = 275, alpha = 0.1) +
-  annotate("rect", xmin = 41, xmax = 76, ymin = 1, ymax = 275, alpha = 0.1) +
+  # annotate("rect", xmin = 1, xmax = 29, ymin = 1, ymax = 275, alpha = 0.1) +
+  # annotate("rect", xmin = 41, xmax = 76, ymin = 1, ymax = 275, alpha = 0.1) +
   ylim(c(1, 275)) +
@@
   theme(axis.title.y = element_blank()) +
-  facet_wrap(~name, nrow = 3, scales = "free")
+  facet_wrap(~name, nrow = 3, scales = "free_y")
 p38_plot
@@
 ```
+
+## 2
+
+Participant 2 had a pulmonary exacerbation starting on night 41.
+
+```{r}
+#| warning: false
+p2 <- full_join(
+  agg_counts,
+  cfqr_children_6_11 %>% select(sid, Date, respirat) %>%
+    rename(`CFQ-R Children Ages 6-11` = respirat),
+  by = join_by(sid, Date)
+)
+p2 <- full_join(
+  p2,
+  cfqr_parent_6_13 %>% select(sid, Date, respirat) %>%
+    rename(`CFQ-R Caregivers 6-13` = respirat),
+  by = join_by(sid, Date)
+)
+p2 <- p2 %>%
+  filter(sid == 2) %>%
+  pivot_longer(c(Coughs, `CFQ-R Children Ages 6-11`, `CFQ-R Caregivers 6-13`)) %>%
+  filter(!is.na(value))
+p2$name <- factor(p2$name,
+  levels = c("Coughs", "CFQ-R Children Ages 6-11", "CFQ-R Caregivers 6-13")
+)
+p2_plot <- ggplot(p2, aes(x = Night, y = value)) +
+  geom_point() +
+  geom_line() +
+  theme_bw() +
+  theme(axis.title.y = element_blank()) +
+  facet_wrap(~name, nrow = 3, scales = "free_y")
+p2_plot
+ggsave(paste0(home_dir, "/Dissemination/Figures/participant_2.png"),
+  plot = p2_plot, width = 9, height = 7, units = "in"
+)
+```
+
+## 3
+
+Participant 3 had a pulmonary exacerbation starting on night 1.
+
+```{r}
+#| warning: false
+cfqr_children_12_13$sid <- as.numeric(cfqr_children_12_13$sid)
+p3 <- full_join(
+  agg_counts,
+  cfqr_children_12_13 %>% select(sid, Date, respirat) %>%
+    rename(`CFQ-R Children Ages 12-13` = respirat),
+  by = join_by(sid, Date)
+)
+p3 <- full_join(
+  p3,
+  cfqr_parent_6_13 %>% select(sid, Date, respirat) %>%
+    rename(`CFQ-R Caregivers 6-13` = respirat),
+  by = join_by(sid, Date)
+)
+p3 <- p3 %>%
+  filter(sid == 3) %>%
+  pivot_longer(c(Coughs, `CFQ-R Children Ages 12-13`, `CFQ-R Caregivers 6-13`)) %>%
+  filter(!is.na(value))
+p3$name <- factor(p3$name,
+  levels = c("Coughs", "CFQ-R Children Ages 12-13", "CFQ-R Caregivers 6-13")
+)
+p3$Night <- as.numeric(difftime(ymd(p3$Date), ymd("2024-05-07")))
+p3_plot <- ggplot(p3, aes(x = Night, y = value)) +
+  geom_point() +
+  geom_line() +
+  theme_bw() +
+  theme(axis.title.y = element_blank()) +
+  facet_wrap(~name, nrow = 3, scales = "free_y")
+p3_plot
+ggsave(paste0(home_dir, "/Dissemination/Figures/participant_3.png"),
+  plot = p3_plot, width = 9, height = 7, units = "in"
+)
+```

### 2025-07-29 Edith Zemanick/Sweat Variability/sweat_test_analysis_chec.qmd
CHEC manuscript updates

@@
 chec$post_ratio <- chec$swsavg / chec$swcavg
+# Factor variable formatting
+chec$race_txt <- factor(chec$race_txt,
+  levels = c(
+    "White", "Black or African American", "American Indian or Alaska Native",
+    "Native Hawaiian or Other Pacific Islander", "More than One Race",
+    "Unknown or Not Reported"
+  )
+)
+chec$modqual <- factor(chec$modqual,
+  levels = c(
+    "Ivacaftor", "Lumacaftor / Ivacaftor",
+    "Tezacaftor / Ivacaftor", "Elexacaftor / Tezacaftor / Ivacaftor"
+  ),
+  labels = c("IVA", "LUM/IVA", "TEZ/IVA", "ELX/TEZ/IVA")
+)
+chec$agecat_txt <- factor(chec$agecat_txt,
+  levels = c(
+    "2-5 years", "6-11 years", "12-17 years", "18-25 years", ">=26 years"
+  )
+)
 # Long dataset for easier plotting and paired analysis
@@
   )
-chec_long$pre_post <- factor(chec_long$pre_post, levels = c("pre", "post"))
+chec_long$pre_post <- factor(chec_long$pre_post,
+  levels = c("pre", "post"),
+  labels = c("Pre-Modulator", "Post-Modulator")
+)
 # Set table options
@@
   tableby.control(
-    numeric.stats = c("N", "Nmiss2", "meansd", "range"),
-    cat.stats = c("N", "Nmiss2", "countpct")
+    numeric.stats = c("N", "Nmiss", "meansd", "medianrange"),
+    cat.stats = c("N", "Nmiss", "countpct"), digits = 1
   )
@@
 t1 <- tableby(
-  modqual ~ sex + kwt(age, "N", "Nmiss2", "medianq1q3") + agecat_txt +
-    race_txt + genotype_group + swcavg + swhcavgpm + swsavg + swhsavgpm,
-  data = chec, cat.test = "fe", simulate.p.value = TRUE
+  modqual ~ sex + kwt(age) + agecat_txt +
+    race_txt + genotype_group + swcavg + swhcavgpm + swsavg + swhsavgpm +
+    pre_ratio + post_ratio,
+  data = chec, cat.test = "fe", simulate.p.value = TRUE, control = mycontrols
 )
@@
 pt1 <- paired(
-  pre_post ~
-    signed.rank(cl, "N", "Nmiss2", "medianq1q3", "range") +
-    signed.rank(na, "N", "Nmiss2", "medianq1q3", "range") +
+  pre_post ~ cl + na +
     signed.rank(ratio, "N", "Nmiss2", "medianq1q3", "range"),
@@
 ```{r}
-ggplot(chec_long, aes(y = ratio, x = pre_post, fill = pre_post)) +
+ggplot(chec_long, aes(x = pre_post, y = ratio)) +
   geom_boxplot() +
   geom_signif(
-    comparison = list(c("pre", "post")), test = "wilcox.test",
-    test.args = list(paired = TRUE, exact = FALSE), y_position = 3.5,
-    map_signif_level = function(p) format.pval(p, digits = 3, eps = 0.001)
+    comparison = list(c("Pre-Modulator", "Post-Modulator")),
+    test = "wilcox.test", test.args = list(paired = TRUE, exact = FALSE),
+    y_position = 3.5, map_signif_level = function(p) 
+      format.pval(p, digits = 3, eps = 0.001)
   ) +
   theme_bw() +
-  scale_fill_manual(values = c("#EFB036", "#3B6790")) +
+  theme(legend.title = element_blank(), axis.title.x = element_blank()) +
+  ylab("Na:Cl Ratio") +
   facet_wrap(~modqual)
@@
   drop_na()
+model_df$cl = model_df$cl / 10
+model_df$na = model_df$na / 10
 # From Amalia
@@
 I tried fitting Amalia's models with a random effect for participant, but the random effect variance was 0 and the model estimates are the exact same as a regular logistic model.
-We think that this is because the correlation in responses is accounted for by the design, so the mized effects are not necessary.
+We think that this is because the correlation in responses is accounted for by the design, so the mixed effects are not necessary.
 This is worth thinking about a little bit more, but for now we'll go ahead with the simplified logistic models.
+
+Units in the following models are for a 10 unit change in sweat Na or Cl.
 
@@
 summary(tableby(post_cl ~ post_ratio + post_ratio_1.2,
-  data = chec[chec$modqual == "Elexacaftor / Tezacaftor / Ivacaftor", ],
+  data = chec[chec$modqual == "ELX/TEZ/IVA", ],
   control = list(
@@
 sp_pre <- chec_long %>%
-  filter(pre_post == "pre") %>%
+  filter(pre_post == "Pre-Modulator") %>%
   ggplot(., aes(x = ratio, y = cl)) +
@@
 sp_post <- chec_long %>%
-  filter(pre_post == "post") %>%
+  filter(pre_post == "Post-Modulator") %>%
   ggplot(., aes(x = ratio, y = cl)) +
@@
 plot_df$modqual <- factor(plot_df$modqual,
-  levels = c(
-    "Pre-Modulator", "Lumacaftor / Ivacaftor", "Tezacaftor / Ivacaftor",
-    "Ivacaftor", "Elexacaftor / Tezacaftor / Ivacaftor"
-  )
+  levels = c("Pre-Modulator", "IVA", "LUM/IVA", "TEZ/IVA", "ELX/TEZ/IVA")
 )
@@
 
-## Combined figure
+## Combined figures
 
@@
 boxp / (sp_pre + sp_post)
+sp_pre + sp_post
 ```

### 2025-07-29 Edith Zemanick/Sweat Variability/sweat_test_analysis_chec.qmd
small formatting changes

@@
     test = "wilcox.test", test.args = list(paired = TRUE, exact = FALSE),
-    y_position = 3.5, map_signif_level = function(p) 
+    y_position = 3.5, map_signif_level = function(p) {
       format.pval(p, digits = 3, eps = 0.001)
+    }
   ) +
@@
   drop_na()
-model_df$cl = model_df$cl / 10
-model_df$na = model_df$na / 10
+model_df$cl <- model_df$cl / 10
+model_df$na <- model_df$na / 10
 # From Amalia

### 2025-07-31 Stacey Martiniano/ntm_agreement.qmd
feature scores

@@
 library(knitr)
-library(fastDummies)
 library(lmerTest)
@@
 # Import data from REDCap
-data <- exportRecordsTyped(rcon, labels = F)
+data <- exportRecordsTyped(rcon)
 data <- data.frame(lapply(data, as.character))
@@
 
+## Total disease feature score (orange)
+
 ```{r}
 # Import disease state
-outcomes <- read.csv("./Stacey Martiniano/NTM CT Scoring/Data_Clean/disease.key.csv", na.strings = "")
+outcomes <-
+  read.csv("./Stacey Martiniano/NTM CT Scoring/Data_Clean/disease.key.csv",
+    na.strings = ""
+  )
 outcomes$NTM.PD..Y.N. <- factor(outcomes$NTM.PD..Y.N.,
@@
 )
+outcomes <- outcomes %>%
+  select(cff_registry_id, NTM.PD..Y.N.) %>%
+  rename(`NTM-PD` = NTM.PD..Y.N.)
 # For the duplicated scans, just use the first reading and combine with the
@@
 scores <- rbind(scores, data_gs[data_gs$scorer == "GS", ])
-# Add outcome info
-scores <- left_join(scores,
-  outcomes %>% select(cff_registry_id, NTM.PD..Y.N.),
-  by = join_by(scan_id == cff_registry_id)
-) %>%
-  rename(`NTM-PD` = NTM.PD..Y.N.)
 # Calculate the scores for each feature.
@@
 # majority first, then add the locations together for the feature score.
-feature_scores <- scores %>% group_by(scan_id, Feature, `NTM-PD`)
+feature_scores <- scores %>%
+  group_by(scan_id, Feature, Location) %>%
+  summarise(Majority = factor(sum(value == "Present"),
+    levels = 0:3, labels = c("Absent", "Absent", "Present", "Present")
+  ), .groups = "drop_last") %>%
+  summarise(Score = sum(Majority == "Present"), .groups = "drop")
+# Add outcome data
+feature_scores <- left_join(feature_scores, outcomes,
+  by = join_by(scan_id == cff_registry_id)
+)
+# Summary table
+feature_scores %>%
+  pivot_wider(names_from = Feature, values_from = Score) %>%
+  select(-scan_id) %>%
+  tbl_summary(by = `NTM-PD`, missing_text = "Missing")
+feature_scores %>%
+  pivot_wider(names_from = Feature, values_from = Score) %>%
+  select(-scan_id) %>%
+  tbl_summary(
+    by = `NTM-PD`, missing_text = "Missing",
+    type = list(
+      c(
+        "Atelectasis", "Bronchiectasis", , "Cavity", "Consolidation", "Cyst",
+        "Ground Glass", "Mucoid Impaction", "Nodule", "Tree In Bud",
+        "Tree In Bud OR Nodule"
+      ) ~ "continuous2"
+    ),
+    statistic = all_continuous() ~ c("{median} ({p25}, {p75})", "{min}, {max}")
+  )
 ```
-
-# Mixed Model Approach
-
-```{r data cleaning 2}
-# Import data
-data <- exportRecordsTyped(rcon, labels = F)
-data <- data.frame(lapply(data, as.character))
-data$scan_id <- as.numeric(data$scan_id)
-# Make scorer IDs by initial
-data$scorer <- paste0(
-  toupper(substr(data$reader_first_name, 1, 1)),
-  toupper(substr(data$reader_last_name, 1, 1))
-)
-data$scorer[data$scorer == "GG"] <- "GS"
-# Deidentify readers
-old <- c("AF", "DL", "JW")
-new <- c("A", "B", "C")
-data$scorer[data$scorer %in% old] <- new[match(data$scorer, old, nomatch = 0)]
-# Revert duplicate IDs back to original to match with gold standard
-training <- read.csv("./Stacey Martiniano/NTM CT Scoring/Data_Clean/training.key.csv")
-dups <- read.csv("./Stacey Martiniano/NTM CT Scoring/Data_Clean/duplicated.key.csv")
-training$Training_id <- as.numeric(sub("T00", "", training$Training_id))
-# Change duplicates to OG IDs
-old <- dups$duplicated_ssid..dates.shifted..17.days.
-new <- dups$original_ssid
-data$scan_id[data$scan_id %in% old] <- new[na.omit(match(data$scan_id, old))]
-# Change training IDs to OG IDs
-old <- training$Training_id
-new <- training$Cffid.Scoring.Set
-data$scan_id[data$scan_id %in% old] <- new[na.omit(match(data$scan_id, old))]
-# Convert to long
-data <- data %>%
-  select(
-    scan_id, scorer, scoresheet_timestamp,
-    bronchiectasis_rul:atelectasis___6
-  ) %>%
-  arrange(scan_id, scorer, scoresheet_timestamp) %>%
-  group_by(scan_id, scorer) %>%
-  mutate(scorer_repeat = row_number()) %>%
-  select(-scoresheet_timestamp) %>%
-  pivot_longer(bronchiectasis_rul:atelectasis___6,
-    names_to = c("feature", "location"),
-    names_pattern = "(.*)_(.*)$"
-  ) %>%
-  mutate(
-    feature = gsub("__", "", feature),
-    feature = str_to_title(gsub("_", " ", feature))
-  )
-data$location <- factor(data$location,
-  levels = c(
-    "1", "2", "3", "4", "5", "6", "ling", "lll",
-    "lul", "rll", "rml", "rul"
-  ),
-  labels = c(
-    "RUL", "RML", "RLL", "LUL", "Ling", "LLL",
-    "Ling", "LLL", "LUL", "RLL", "RML", "RUL"
-  )
-)
-# Format values
-data$value <- factor(data$value,
-  levels = c(
-    "Absent", "Checked", "Cylindric", "Cystic", "Unchecked", "Varicose"
-  ),
-  labels = c("Absent", "Present", "Present", "Present", "Absent", "Present")
-)
-data$value <- as.numeric(data$value) - 1
-# Separate out gold standards and duplicates
-data_dups <- data %>% filter(scan_id %in% dups$original_ssid)
-data_gs <- data %>% filter(scan_id %in% training$Cffid.Scoring.Set)
-data <- data %>%
-  filter(!scan_id %in% c(dups$original_ssid, training$Cffid.Scoring.Set))
-# Second readings only for comparison to GS
-data_gs <- data_gs %>%
-  group_by(scan_id, scorer, feature, location) %>%
-  slice_max(scorer_repeat)
-# Dummy variables for scorer, organize columns
-data <- data %>%
-  dummy_cols(select_columns = "scorer") %>%
-  select(scan_id, feature, location, scorer, scorer_A, scorer_B, scorer_C, value)
-```
-
-## Data setup
-
-```{r}
-DT::datatable(data)
-```
-
-## Model
-
-- Readers $r=1,2,3$
-- Scans $s=1,2,...,S$
-- Features $f=1,2,...,F$
-- Lobes $l=1,2,...,6$
-- Outcome $y_{rsfl}$ is binary
-- Model notation: $$y_{rsf=1}=\alpha+\beta X_l + b_1+b_2+b_3+\epsilon_{rs}$$
-
-### Bronchiectasis
-
-```{r}
-#| code-fold: show
-df <- data %>% filter(feature == "Bronchiectasis")
-model <- glmer(value ~ location + (1 | scorer),
-  family = "binomial", data = df
-)
-summary(model)
-icc(model)
-```
-
-### Mucoid Impaction
-
-```{r}
-#| code-fold: show
-df <- data %>% filter(feature == "Mucoid Impaction")
-model <- glmer(value ~ location + (1 | scorer),
-  family = "binomial", data = df
-)
-summary(model)
-icc(model)
-```
-
-### Cyst
-
-```{r}
-#| code-fold: show
-df <- data %>% filter(feature == "Cyst")
-model <- glmer(value ~ location + (1 | scorer),
-  family = "binomial", data = df
-)
-summary(model)
-icc(model)
-```
-
-### Cavity
-
-```{r}
-#| code-fold: show
-df <- data %>% filter(feature == "Cavity")
-model <- glmer(value ~ location + (1 | scorer),
-  family = "binomial", data = df
-)
-summary(model)
-icc(model)
-```
-
-### Consolidation
-
-```{r}
-#| code-fold: show
-df <- data %>% filter(feature == "Consolidation")
-model <- glmer(value ~ location + (1 | scorer),
-  family = "binomial", data = df
-)
-summary(model)
-icc(model)
-```
-
-### Ground Glass
-
-```{r}
-#| code-fold: show
-df <- data %>% filter(feature == "Ground Glass")
-model <- glmer(value ~ location + (1 | scorer),
-  family = "binomial", data = df
-)
-summary(model)
-icc(model)
-```
-
-### Nodule
-
-```{r}
-#| code-fold: show
-df <- data %>% filter(feature == "Nodule")
-model <- glmer(value ~ location + (1 | scorer),
-  family = "binomial", data = df
-)
-summary(model)
-icc(model)
-```
-
-### Tree In Bud
-
-```{r}
-#| code-fold: show
-df <- data %>% filter(feature == "Tree In Bud")
-model <- glmer(value ~ location + (1 | scorer),
-  family = "binomial", data = df
-)
-summary(model)
-icc(model)
-```
-
-### Atelectasis
-
-```{r}
-#| code-fold: show
-df <- data %>% filter(feature == "Atelectasis")
-model <- glmer(value ~ location + (1 | scorer),
-  family = "binomial", data = df
-)
-summary(model)
-icc(model)
-```

### 2025-07-31 Stacey Martiniano/ntm_agreement.qmd
need to ask Stacey about the total score thing

@@
 
-## Total disease feature score (orange)
-
 ```{r}
@@
 scores <- rbind(scores, data_gs[data_gs$scorer == "GS", ])
+```
+
+## Feature score (orange)
+
+```{r}
 # Calculate the scores for each feature.
@@
 ```
+
+## Lobe score (pink)
+
+```{r}
+# Lobe scores similar to above but we'll add over the lobes rather than the
+# features
+lobe_scores <- scores %>%
+  group_by(scan_id, Location, Feature) %>%
+  summarise(Majority = factor(sum(value == "Present"),
+    levels = 0:3, labels = c("Absent", "Absent", "Present", "Present")
+  ), .groups = "drop_last") %>%
+  summarise(Score = sum(Majority == "Present"), .groups = "drop")
+# Add outcome data
+lobe_scores <- left_join(lobe_scores, outcomes,
+  by = join_by(scan_id == cff_registry_id)
+)
+# Summary table
+lobe_scores %>%
+  pivot_wider(names_from = Location, values_from = Score) %>%
+  select(-scan_id) %>%
+  tbl_summary(by = `NTM-PD`, missing_text = "Missing")
+lobe_scores %>%
+  pivot_wider(names_from = Location, values_from = Score) %>%
+  select(-scan_id) %>%
+  tbl_summary(
+    by = `NTM-PD`, missing_text = "Missing",
+    type = list(
+      c("RUL", "RML", "RLL", "LUL", "Ling", "LLL") ~ "continuous2"
+    ),
+    statistic = all_continuous() ~ c("{median} ({p25}, {p75})", "{min}, {max}")
+  )
+```
+
+## Total score (pink)
+
+The way we are currently scoring CT scans, the lobe-wise and feature-wise sums are exactly the same.
+
+```{r}
+# Basically we are adding the two previous dataframes together
+# lobe_scores = lobe_scores %>% filter()
+# feature_scores = feature_scores %>% group_by(scan_id) %>%
+#   summarise(Total = sum(Score))
+```

### 2025-08-04 Emily Muther/body_image_trikafta.qmd
added adherence to table 1

@@
     Ethnicity, InsuranceType, BMI, bmi_raw, FEV1, BE_APPEARANCE_MEAN,
-    BE_ATTRIBUTION_MEAN, BE_WEIGHT_MEAN
+    BE_ATTRIBUTION_MEAN, BE_WEIGHT_MEAN, contains("Adherence")
+  ) %>%
+  mutate(
+    PercentAdherence = as.numeric(sub("%", "", as.character(PercentAdherence))),
+    over80 = PercentAdherence >= 80
   ) %>%
@@
     type = list(
-      c("FRStotal", "FRS_IdealBody", "FRS_ThinkyouLook", "bmi_raw") ~ "continuous2",
+      c("FRStotal", "FRS_IdealBody", "FRS_ThinkyouLook", "bmi_raw", "PercentAdherence") ~ "continuous2",
       all_continuous() ~ "continuous2"

### 2025-08-05 Edith Zemanick/Sweat Variability/sweat_test_analysis_chec.qmd
CHEC added ethnicity

@@
 # Import CHEC data
-chec <- read.csv("./Data_Raw/CHEC sweat sodium and chloride.csv",
+chec <- read.csv("./Data_Raw/CHEC sweat sodium and chloride 2024-04-19.csv",
   na.strings = c("", "NA")
+)
+ethnicity <- read.csv("./Data_Raw/CHEC sweat sodium and chloride 2025-07-30.csv",
+  na.strings = c("", "NA")
+)
+chec <- left_join(chec, ethnicity %>% select(subject, crfethnicity_txt),
+  by = join_by(subject)
 )
@@
   )
+)
+chec$crfethnicity_txt <- factor(chec$crfethnicity_txt,
+  levels = c("Not Hispanic or Latino", "Hispanic or Latino")
 )
@@
   modqual ~ sex + kwt(age) + agecat_txt +
-    race_txt + genotype_group + swcavg + swhcavgpm + swsavg + swhsavgpm +
-    pre_ratio + post_ratio,
+    race_txt + crfethnicity_txt + genotype_group + swcavg + swhcavgpm + swsavg +
+    swhsavgpm + pre_ratio + post_ratio,
   data = chec, cat.test = "fe", simulate.p.value = TRUE, control = mycontrols
@@
 ```
-
-- I think that we currently don't have ethnicity data.
 

### 2025-08-05 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
HbA1c table for Monique

--- 2025-05-15 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
@@
 github_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Documents/GitHub",
-  "Linux" = "/home/timvigers/Documents/GitHub"
+  "Darwin" = "/Users/tim/Documents/GitHub",
+  "Linux" = "/home/tim/Documents/GitHub"
 )
@@
   "A_pseudomonasaeruginosa", "A_fungalyeast1",
-  "A_pulmonarycomplications1", "A_mycocultureresults3"
+  "A_pulmonarycomplications1", "A_mycocultureresults3", "A_cultureresults3"
 )
@@
 # This is super slow but idc idc
+first_cfrd=head(first_cfrd,100)
 t <- apply(first_cfrd, 1, function(r) {
@@
 })
+first_cfrd$a1c_ogtt <- t
+cfrd_only <- left_join(first_cfrd, annual |> remove_labels(),
+  by = join_by(eDWID, first_cfrd == ReviewYear)
+)
+cfrd_only$a1c_ogtt <- factor(cfrd_only$a1c_ogtt,
+  levels = c("both", "ogtt", "a1c", "neither"),
+  labels = c("OGTT", "OGTT", "HbA1c Only", "Neither")
+)
+cfrd_only %>%
+  select(
+    a1c_ogtt, Age_YrEnd, A_weight, A_height, A_bmivalue, A_bmipercentile,
+    A_FEV1_pct_predicted, A_FVC_pct_predicted, Race, Hispanicrace, MutClass,
+    A_hgba1c, A_ogttfast_bloodglucose, A_twohour_bloodglucose, A_IsOnEnzymes,
+    liver_disease, A_cultureresults3
+  ) %>%
+  tbl_summary(by = a1c_ogtt, missing_text = "Missing") %>%
+  add_overall() %>%
+  add_p(test=list(all_continuous() ~ "oneway.test"))
 ```
@@
   ggtitle("Prevalence of CFRD")
-fig1 <- fig1a + fig1b + fig1c + plot_layout(guides = "collect")+
+fig1 <- fig1a + fig1b + fig1c + plot_layout(guides = "collect") +
   plot_annotation(title = "US Cystic Fibrosis Foundation Patient Registry Trends in Screening, Incidence, and Prevalence of Cystic Fibrosis Related Diabetes from 2003-2018")
@@
 annual$tstop <- as.numeric(annual$tstop) - as.numeric(og)
+# Write cleaned data
+save(annual, file = "./Data_Clean/survival_dataset.RData")
 # Formula for base model
@@
 ```{r results='asis'}
-f_vx770 <- update(f, . ~ . + A_Vx770)
+f_vx770 <- update(f, . ~ . + A_Vx770 - A_corticosteroids1 - NumPulmExacerbation - age_group)
 vx770_cox_res <- coxph(f_vx770,
@@
 ```{r}
-f_vx809comb <- update(f, . ~ . + A_VX809comb)
+f_vx809comb <- update(f, . ~ . + A_VX809comb - A_corticosteroids1 - NumPulmExacerbation - age_group - Hispanicrace)
 vx809comb_cox_res <- coxph(f_vx809comb,

### 2025-08-05 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
CFRD diagnosis comparisons for A1c registry paper

@@
 
+# Diagnosis comparisons for HbA1c registry paper
+
 ```{r}
@@
 # This is super slow but idc idc
-first_cfrd=head(first_cfrd,100)
 t <- apply(first_cfrd, 1, function(r) {
@@
 first_cfrd$a1c_ogtt <- t
-cfrd_only <- left_join(first_cfrd, annual |> remove_labels(),
+cfrd_only <- left_join(first_cfrd, annual %>% remove_labels(),
   by = join_by(eDWID, first_cfrd == ReviewYear)
@@
   ) %>%
-  tbl_summary(by = a1c_ogtt, missing_text = "Missing") %>%
+  tbl_summary(
+    by = a1c_ogtt, missing_text = "Missing",
+    statistic = list(all_continuous() ~ "{mean} ({sd})")
+  ) %>%
   add_overall() %>%
-  add_p(test=list(all_continuous() ~ "oneway.test"))
+  add_p(test = list(all_continuous() ~ "oneway.test")) %>%
+  modify_caption("**All Groups**")
+cfrd_only %>%
+  filter(a1c_ogtt %in% c("OGTT", "HbA1c Only")) %>%
+  mutate(a1c_ogtt = droplevels(a1c_ogtt)) %>%
+  select(
+    a1c_ogtt, Age_YrEnd, A_weight, A_height, A_bmivalue, A_bmipercentile,
+    A_FEV1_pct_predicted, A_FVC_pct_predicted, Race, Hispanicrace, MutClass,
+    A_hgba1c, A_ogttfast_bloodglucose, A_twohour_bloodglucose, A_IsOnEnzymes,
+    liver_disease, A_cultureresults3
+  ) %>%
+  tbl_summary(
+    by = a1c_ogtt, missing_text = "Missing",
+    statistic = list(all_continuous() ~ "{mean} ({sd})")
+  ) %>%
+  add_overall() %>%
+  add_p(test = list(all_continuous() ~ "oneway.test")) %>%
+  modify_caption("**OGTT vs. HbA1c Only**")
 ```
@@
   tbl_summary(by = `Age Group`) %>%
-  add_ci() %>%
+  # add_ci() %>%
   add_p() %>%
@@
   tbl_summary(by = Gender) %>%
-  add_ci() %>%
+  # add_ci() %>%
   add_p() %>%
@@
   tbl_summary(by = MutClass) %>%
-  add_ci() %>%
+  # add_ci() %>%
   add_p() %>%
@@
   tbl_summary(by = Race) %>%
-  add_ci() %>%
+  # add_ci() %>%
   add_p() %>%
@@
   tbl_summary(by = Hispanicrace) %>%
-  add_ci() %>%
+  # add_ci() %>%
   add_p() %>%

### 2025-08-06 Jordana Hoppe/Home Collection/manuscript.qmd
new home collection data

--- 2025-02-07 Jordana Hoppe/Home Collection/manuscript.qmd
@@
 home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
   "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
-  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
+  "Linux" = "/home/tim/OneDrive/Vigers/CF"
 )
 github_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Documents/GitHub",
+  "Darwin" = "/Users/tim/Documents/GitHub",
   "Windows" = "C:/Users/Tim/GitHub",
-  "Linux" = "/home/timvigers/GitHub"
+  "Linux" = "/home/tim/GitHub"
 )
@@
 source(paste0(github_dir, "/CF/Jordana Hoppe/Home Collection/home_collection_surveys.r"))
-# Exclude missing subjects
+# List subjects to exclude
 data_exclude <- c("009-07", "009-17", "009-25", "009-37")
-survey_exclude <- c(15, 27, 30, 34, 47)
-data <- data %>% filter(!subject_id %in% data_exclude)
-survey_data <- survey_data %>% filter(!participant_id %in% survey_exclude)
+survey_exclude <- c(6, 15, 27, 30, 34, 47)
 # Fix missing labels
@@
   levels = 0:2,
-  labels = c("No Growth", "Some Growth", "Some Growth")
+  labels = c("No Growth", "Any Growth", "Any Growth")
 )
@@
 )
+# Times from home collection
+data$days_to_home_received <-
+  as.numeric(difftime(as.Date(data$home_cx_date_received), as.Date(data$home_cx_date),
+    units = "days"
+  ))
+label(data$days_to_home_received) <-
+  "Days from Home Collection to Home Collection Received"
+data$days_to_home_collection <-
+  as.numeric(difftime(as.Date(data$home_cx_date), as.Date(data$clinic_date),
+    units = "days"
+  ))
+label(data$days_to_home_collection) <-
+  "Days from Clinic Collection to Home Collection"
 ```
@@
     age, pt_sex.factor, race, ethnicity.factor, mutation1.factor,
-    mutation2.factor, Genotype, rural_location.factor, fev1_pp, fvc_pp
+    mutation2.factor, Genotype, rural_location.factor, cftr_modulator.factor,
+    fev1_pp, fvc_pp, days_to_home_received, days_to_home_collection,
+    cx_past_results___1.factor:cx_past_results___11.factor
   ) %>%
@@
     label = labels, missing_text = "Missing",
-    statistic = list(all_continuous() ~ "{mean} ({sd})")
+    type = list(c(days_to_home_received, days_to_home_collection) ~
+      "continuous2"),
+    statistic = list(
+      all_continuous() ~ "{mean} ({sd})",
+      c(days_to_home_received, days_to_home_collection) ~
+        "{median} [{p25}, {p75}]"
+    )
   ) %>%
   as_kable_extra()
+data <- data %>% filter(!subject_id %in% data_exclude)
+survey_data <- survey_data %>% filter(!participant_id %in% survey_exclude)
 ```
-
-It appears that the two participants who were previously marked as unknown race are now considered more than one race.
 
@@
         Immediate == "Present" & Delayed == "Present" ~ "Both Present",
+        Immediate == "Any Growth" & Delayed == "Any Growth" ~ "Any Growth on Both",
         Immediate == "Absent" & Delayed == "Absent" ~ "Both Absent",
+        Immediate == "No Growth" & Delayed == "No Growth" ~ "No Growth on Either",
         .default = "Disagreed"
@@
         Immediate == "Present" & Home == "Present" ~ "Both Present",
+        Immediate == "Any Growth" & Home == "Any Growth" ~ "Any Growth on Both",
         Immediate == "Absent" & Home == "Absent" ~ "Both Absent",
+        Immediate == "No Growth" & Home == "No Growth" ~ "No Growth on Either",
         .default = "Disagreed"
@@
         Delayed == "Present" & Home == "Present" ~ "Both Present",
+        Delayed == "Any Growth" & Home == "Any Growth" ~ "Any Growth on Both",
         Delayed == "Absent" & Home == "Absent" ~ "Both Absent",
+        Delayed == "No Growth" & Home == "No Growth" ~ "No Growth on Either",
         .default = "Disagreed"
@@
 names(labels) <- colnames(survey_data)
+# Why didn't people prefer home collection? (for reviewer response)
+no_home_collection <- survey_data %>%
+  filter(q6.factor == "No") %>%
+  select(participant_id, q6, q6b___1:q6b___4, q6b_spec)
+colnames(no_home_collection) <- label(no_home_collection)
+kable(no_home_collection)
 ```
@@
   geom_text(aes(label = scales::percent(prop, accuracy = 1)), position = position_stack(vjust = 0.5)) +
-  ggtitle("Benefits") +
+  ggtitle("Perceived Benefits") +
   scale_x_continuous(
@@
   geom_text(aes(label = scales::percent(prop, accuracy = 1)), position = position_stack(vjust = 0.5)) +
-  ggtitle("Barriers") +
+  ggtitle("Perceived Barriers") +
   scale_x_continuous(
@@
   rename(any_of(setNames(names(labels), labels)))
-p3 <- gglikert(as_tibble(df3), add_totals = F)
+p3 <- gglikert(as_tibble(df3), add_totals = F) +
+  scale_fill_manual(values = c("#7895B2", "#AEBDCA", "#E8DFCA")) +
+  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
 p3
@@
   rename(any_of(setNames(names(labels), labels)))
-p5 <- gglikert(df5)
+p5 <- gglikert(df5) +
+  scale_fill_manual(values = c("#018571", "#80CDC1", "#F5F5F5", "#DFC27D", "#A6611A")) +
+  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
 p5

### 2025-08-06 Jordana Hoppe/Home Collection/home_collection_data.r

@@ -3,14 +3,14 @@ library(Hmisc)
 # Set WD
 switch(Sys.info()[["sysname"]],
   Windows = {
-    home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
+    home_dir <- "C:/Users/tim/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
   },
   Darwin = {
     home_dir <- "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
   }
 )
 # Read Data
-data <- read.csv(paste0(home_dir, "/", "Jordana Hoppe/Home Collection/Data_Raw/HomeCollection220661_DATA_2024-01-29_0945.csv"))
+data <- read.csv(paste0(home_dir, "/", "Jordana Hoppe/Home Collection/Data_Raw/HomeCollection220661_DATA_2025-08-06_1441.csv"))
 # Setting Labels
 label(data$subject_id) <- "Subject ID"
 label(data$pt_initials) <- "Patient Initials"
@@ -30,6 +30,24 @@ label(data$tract_type1) <- "Tract type"
 label(data$mutation2) <- "Mutation 2"
 label(data$mutation2_spec) <- "Specify mutation 2, if Other"
 label(data$tract_type2) <- "Tract type "
+label(data$cftr_modulator) <- "CFTR Modulator Status"
+label(data$cx_past_results___1) <- "Any positive culture results in the past 12 months? (choice=Methicillin sensitive Staph aureus)"
+label(data$cx_past_results___2) <- "Any positive culture results in the past 12 months? (choice=Methicillin resistant Staph aureus)"
+label(data$cx_past_results___16) <- "Any positive culture results in the past 12 months? (choice=A. xylosoxidans)"
+label(data$cx_past_results___17) <- "Any positive culture results in the past 12 months? (choice=Achromobacter not xylosoxidans)"
+label(data$cx_past_results___10) <- "Any positive culture results in the past 12 months? (choice=Acinetobacter)"
+label(data$cx_past_results___3) <- "Any positive culture results in the past 12 months? (choice=H. influenzae)"
+label(data$cx_past_results___4) <- "Any positive culture results in the past 12 months? (choice=P. aeruginosa)"
+label(data$cx_past_results___5) <- "Any positive culture results in the past 12 months? (choice=P. aeruginosa (mucoid))"
[…]

### 2025-08-07 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
genotype severity analysis

@@
 )
+github_dir <- switch(Sys.info()["sysname"],
+  "Darwin" = "/Users/tim/Documents/GitHub",
+  "Windows" = "C:/Users/Tim/Documents/GitHub",
+  "Linux" = "/home/tim/Documents/GitHub"
+)
+source(paste0(github_dir, "/CF/genotype_class_severity.R"))
 knitr::opts_knit$set(root.dir = home_dir)
@@
 ), ]
+# Get genotype severity
+## By severity
+df$geno_risk_1 <- cf_genotype_class_severity(df$Genotypes1)
+df$geno_risk_2 <- cf_genotype_class_severity(df$Genotypes2)
+df$geno_risk_1[df$Diagnosis == "Non-CF"] <- "Not CF"
+df$geno_risk_2[df$Diagnosis == "Non-CF"] <- "Not CF"
+df$geno_risk <- paste(df$geno_risk_1, df$geno_risk_2, sep = "/")
+df$geno_risk <- factor(df$geno_risk,
+  levels = c(
+    "Mild/Mild", "Mild/Severe", "Mild/Unknown", "Severe/Mild", "Severe/Severe",
+    "Severe/Unknown", "Unknown/Severe", "Unknown/Unknown", "Not CF/Not CF"
+  ),
+  labels = c("M/M", "M/S", "M/U", "M/S", "S/S", "S/U", "S/U", "U", "Not CF")
+)
+## By function
+df$geno1_class <- cf_genotype_class_severity(df$Genotypes1, out = "class")
+df$geno1_class[grep("508", df$Genotypes1)] <- "F508del"
+df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("I", "II", "III"), "MF")
+df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("IV", "V"), "RF")
+df$geno2_class <- cf_genotype_class_severity(df$Genotypes2, out = "class")
+df$geno2_class[grep("508", df$Genotypes2)] <- "F508del"
+df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("I", "II", "III"), "MF")
+df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("IV", "V"), "RF")
+df$geno1_class[df$Diagnosis == "Non-CF"] <- "Not CF"
+df$geno2_class[df$Diagnosis == "Non-CF"] <- "Not CF"
+df$geno_class <- paste(df$geno1_class, df$geno2_class, sep = "/")
+df$geno_class <- factor(df$geno_class,
+  levels = c(
+    "F508del/F508del", "F508del/MF", "F508del/RF", "F508del/Unknown",
+    "MF/F508del", "MF/MF", "MF/RF", "MF/Unknown", "RF/F508del", "RF/MF",
+    "RF/Unknown", "Unknown/F508del", "Unknown/MF", "Unknown/Unknown",
+    "Not CF/Not CF"
+  ),
+  labels = c(
+    "MF/MF", "MF/MF", "MF/RF", "MF/U", "MF/MF", "MF/MF", "MF/RF", "MF/U",
+    "MF/RF", "MF/RF", "RF/U", "MF/U", "MF/U", "U", "Not CF"
+  )
+)
+# Organize columns
+df <- df %>%
+  select(Name:Genotypes2, geno_risk, geno_class, everything()) %>%
+  select(-geno_risk_1, -geno_risk_2, -geno1_class, -geno2_class) %>%
+  rename(`Genotype Risk` = geno_risk, `Genotype Class` = geno_class)
 # Set table options
@@
 
+## Table 2: Participant Demographics by Genotype Class
+
+In the table below "MF" stands for minimal function, "RF" for residual function, and "U" indicates unknown.
+
+```{r results='asis'}
+summary(tableby(`Genotype Class` ~ Sex + Age + Na + Cl + `Na:Cl` + K,
+  data = df, control = mycontrols
+))
+```
+
+## Table 3: Participant Demographics by Genotype Risk
+
+In the table below "M" stands for mild, "S" for severe, and "U" indicates unknown.
+
+```{r results='asis'}
+summary(tableby(`Genotype Risk` ~ Sex + Age + Na + Cl + `Na:Cl` + K,
+  data = df, control = mycontrols
+))
+```
+
 ## Boxplots
@@
 ### Chloride
+
+#### By diagnosis
 
@@
 
+#### By genotype class
+
+```{r}
+#| warning: false
+clp <- ggplot(df, aes(x = `Genotype Class`, y = Cl)) +
+  geom_boxplot() +
+  theme_bw()
+clp
+```
+
+#### By genotype risk
+
+```{r}
+#| warning: false
+clp <- ggplot(df, aes(x = `Genotype Risk`, y = Cl)) +
+  geom_boxplot() +
+  theme_bw()
+clp
+```
+
 ### Sodium
+
+#### By diagnosis
 
@@
 
+#### By genotype class
+
+```{r}
+#| warning: false
+nap <- ggplot(df, aes(x = `Genotype Class`, y = Na)) +
+  geom_boxplot() +
+  theme_bw()
+nap
+```
+
+#### By genotype risk
+
+```{r}
+#| warning: false
+nap <- ggplot(df, aes(x = `Genotype Risk`, y = Na)) +
+  geom_boxplot() +
+  theme_bw()
+nap
+```
+
 ### Na:Cl
+
+#### By diagnosis
 
@@
 
-## Histograms and Density Plots
-
-### Chloride
+#### By genotype class
 
@@
 #| warning: false
-ggplot(df, aes(x = Cl, fill = Diagnosis)) +
-  geom_histogram(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
-  geom_histogram(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
-  geom_histogram(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
+ratiop <- ggplot(df, aes(x = `Genotype Class`, y = `Na:Cl`)) +
+  geom_boxplot() +
+  geom_hline(yintercept = 1, color = "red") +
+  geom_hline(yintercept = 1.25, color = "red") +
   theme_bw() +
-  theme(axis.title.y = element_blank())
-ggplot(df, aes(x = Cl, fill = Diagnosis)) +
-  geom_density(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
-  geom_density(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
-  geom_density(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
+  ylab("Na:Cl") +
+  xlab("Modulator")
+ratiop
+ratiop <- ggplot(df, aes(x = `Genotype Class`, y = `Na:Cl`)) +
+  geom_boxplot() +
+  geom_hline(yintercept = 1, color = "red") +
+  geom_hline(yintercept = 1.25, color = "red") +
+  ylim(0, 2) +
   theme_bw() +
-  theme(axis.title.y = element_blank())
+  ylab("Na:Cl") +
+  xlab("Modulator")
+ratiop
 ```
 
-### Sodium
+#### By genotype risk
 
@@
 #| warning: false
-ggplot(df, aes(x = Na, fill = Diagnosis)) +
-  geom_histogram(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
-  geom_histogram(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
-  geom_histogram(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
+ratiop <- ggplot(df, aes(x = `Genotype Risk`, y = `Na:Cl`)) +
+  geom_boxplot() +
+  geom_hline(yintercept = 1, color = "red") +
+  geom_hline(yintercept = 1.25, color = "red") +
   theme_bw() +
-  theme(axis.title.y = element_blank())
-ggplot(df, aes(x = Na, fill = Diagnosis)) +
-  geom_density(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
-  geom_density(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
-  geom_density(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
+  ylab("Na:Cl") +
+  xlab("Modulator")
+ratiop
+ratiop <- ggplot(df, aes(x = `Genotype Risk`, y = `Na:Cl`)) +
+  geom_boxplot() +
+  geom_hline(yintercept = 1, color = "red") +
+  geom_hline(yintercept = 1.25, color = "red") +
+  ylim(0, 2) +
   theme_bw() +
-  theme(axis.title.y = element_blank())
-```
-
-### Na:Cl
-
-```{r}
-#| warning: false
-ggplot(df, aes(x = `Na:Cl`, fill = Diagnosis)) +
-  geom_histogram(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
-  geom_histogram(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
-  geom_histogram(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
-  theme_bw() +
-  theme(axis.title.y = element_blank())
-ggplot(df, aes(x = `Na:Cl`, fill = Diagnosis)) +
-  geom_density(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
-  geom_density(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
-  geom_density(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
-  theme_bw() +
-  theme(axis.title.y = element_blank())
-ggplot(df, aes(x = `Na:Cl`, fill = Diagnosis)) +
-  geom_density(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
-  geom_density(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
-  geom_density(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
-  xlim(0, 2) +
-  theme_bw() +
-  theme(axis.title.y = element_blank())
+  ylab("Na:Cl") +
+  xlab("Modulator")
+ratiop
 ```

### 2025-08-07 genotype_class_severity.R

@@ -1,60 +1,71 @@
-cf_genotype_class_severity = function(genotypes,out = "severity",na_unknown = T){
+cf_genotype_class_severity <- function(genotypes, out = "severity", na_unknown = T) {
   # Make table
-  I = c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
-        '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
-        '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
-        '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G')
-  I = unique(tolower(gsub("[[:punct:]]|\\s","",I)))
-  I = matrix(c(I,class = rep("I",length(I))),ncol = 2)
-  II = c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E')
-  II = unique(tolower(gsub("[[:punct:]]|\\s","",II)))
-  II = matrix(c(II,class = rep("II",length(II))),ncol = 2)
-  III = c('G551D', 'G551S', 'S549N')
-  III = unique(tolower(gsub("[[:punct:]]|\\s","",III)))
-  III = matrix(c(III,class = rep("III",length(III))),ncol = 2)
-  IV = c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T')
-  IV = unique(tolower(gsub("[[:punct:]]|\\s","",IV)))
-  IV = matrix(c(IV,class = rep("IV",length(IV))),ncol = 2)
-  V = c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
-        'P574H', '3272-26A->G5')
-  V = unique(tolower(gsub("[[:punct:]]|\\s","",V)))
-  V = matrix(c(V,class = rep("V",length(V))),ncol = 2)
-  unknown = c("unknown","I506T","35002ag","26221ga")
[…]

### 2025-08-08 Jordana Hoppe/Home Collection/manuscript.qmd
home collection manuscript update days to home collection

@@
   ))
+data$days_to_home_received[data$subject_id == "009-49"] <- NA
 label(data$days_to_home_received) <-

### 2025-08-08 Stacey Martiniano/ntm_agreement.qmd
need to ask Stacey about the total score thing
NTM added score totals and scatterplots

--- 2025-07-31 Stacey Martiniano/ntm_agreement.qmd
@@
 library(performance)
+library(gtsummary)
+library(patchwork)
 # REDCap
@@
   group_by(scan_id, Feature, Location) %>%
-  summarise(Majority = factor(sum(value == "Present"),
+  summarise(Majority = factor(sum(value == "Present", na.rm = T),
     levels = 0:3, labels = c("Absent", "Absent", "Present", "Present")
@@
   pivot_wider(names_from = Feature, values_from = Score) %>%
+  mutate(`Total Score` = rowSums(.[3:12])) %>%
   select(-scan_id) %>%
@@
         "Ground Glass", "Mucoid Impaction", "Nodule", "Tree In Bud",
-        "Tree In Bud OR Nodule"
+        "Tree In Bud OR Nodule", "Total Score"
       ) ~ "continuous2"
@@
     statistic = all_continuous() ~ c("{median} ({p25}, {p75})", "{min}, {max}")
+  ) %>%
+  add_p(
+    test.args = all_tests("wilcox.test") ~ list(exact = F)
   )
 ```
+
+## Lobe score (pink)
+
+```{r}
+# Lobe scores similar to above but we'll add over the lobes rather than the
+# features
+lobe_scores <- scores %>%
+  group_by(scan_id, Location, Feature) %>%
+  summarise(Majority = factor(sum(value == "Present", na.rm = T),
+    levels = 0:3, labels = c("Absent", "Absent", "Present", "Present")
+  ), .groups = "drop_last") %>%
+  summarise(Score = sum(Majority == "Present"), .groups = "drop")
+# Add outcome data
+lobe_scores <- left_join(lobe_scores, outcomes,
+  by = join_by(scan_id == cff_registry_id)
+)
+# Summary table
+lobe_scores %>%
+  pivot_wider(names_from = Location, values_from = Score) %>%
+  select(-scan_id) %>%
+  tbl_summary(by = `NTM-PD`, missing_text = "Missing")
+lobe_scores %>%
+  pivot_wider(names_from = Location, values_from = Score) %>%
+  mutate(`Total Score` = rowSums(.[3:8])) %>%
+  select(-scan_id) %>%
+  tbl_summary(
+    by = `NTM-PD`, missing_text = "Missing",
+    type = list(
+      c("RUL", "RML", "RLL", "LUL", "Ling", "LLL", "Total Score") ~ "continuous2"
+    ),
+    statistic = all_continuous() ~ c("{median} ({p25}, {p75})", "{min}, {max}")
+  ) %>%
+  add_p(
+    test.args = all_tests("wilcox.test") ~ list(exact = F)
+  )
+```
+
+# Scatterplots
+
+```{r}
+#| warning: false
+#| fig-width: 12
+#| fig-height: 9
+# For the duplicated scans, just use the first reading and combine with the
+# singletons
+scores <- rbind(data, data_dups[data_dups$scorer_repeat == 1, ])
+# Add gold standards
+scores <- rbind(scores, data_gs[data_gs$scorer != "GS", ])
+# Get a total by scorer
+plot_df <- scores %>%
+  group_by(scan_id, scorer) %>%
+  summarise(sum = sum(value == "Present"), .groups = "drop") %>%
+  pivot_wider(names_from = "scorer", values_from = sum)
+# Plots
+ab <- plot_df %>%
+  ggplot(., aes(x = A, y = B)) +
+  geom_point() +
+  ggtitle("Total Score") +
+  geom_abline(intercept = 0, slope = 1) +
+  coord_fixed(ratio = 1, xlim = c(0, 45), ylim = c(0, 45)) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+ac <- plot_df %>%
+  ggplot(., aes(x = A, y = C)) +
+  geom_point() +
+  ggtitle("Total Score") +
+  geom_abline(intercept = 0, slope = 1) +
+  coord_fixed(ratio = 1, xlim = c(0, 45), ylim = c(0, 45)) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+bc <- plot_df %>%
+  ggplot(., aes(x = B, y = C)) +
+  geom_point() +
+  ggtitle("Total Score") +
+  geom_abline(intercept = 0, slope = 1) +
+  coord_fixed(ratio = 1, xlim = c(0, 45), ylim = c(0, 45)) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+ab + bc + ac
+```

### 2025-08-12 Christine Chan/EnVision CF/fsOGTT_pre_post_ETI.qmd
converted c peptide units
manually fixed some ETI start dates per Monique
additional ETI date fixes

@@
 # ID linking from Monique
-ids <- read_excel("./Data_Raw/ETI participant IDs_TV.xlsx")
+ids <- read.csv("./Data_Clean/TV_ETI logitudinal patient list.csv", na.strings = "")
+ids <- ids %>%
+  select(EnVision.ID, GEM.ID) %>%
+  rename(`EnVision ID` = EnVision.ID, `GEM ID` = GEM.ID)
 ids <- ids[!is.na(ids$`GEM ID`), ]
-ids$`EnVision ID` <- sub("-2", "", ids$`EnVision ID`)
 # Get pre-/post-ETI
@@
   fill(eti_start, .direction = "downup")
+# Manually correct a few per Monique
+df$eti_start[df$study_id == "CC0001"] <- "2021-11-24"
+df$eti_start[df$study_id == "CC0010"] <- "2023-01-16"
+df$eti_start[df$study_id == "CC0022"] <- "2019-12-06"
+df$eti_start[df$study_id == "CC0023"] <- "2020-01-09"
+df$eti_start[df$study_id == "CC0026"] <- "2020-12-04"
+df$eti_start[df$study_id == "CC0059"] <- "2021-06-23"
 df$pre_post <-
   factor(df$Date >= df$eti_start, levels = c(F, T), labels = c("Pre", "Post"))
+df$pre_post[df$study_id == "CC0012"] <- "Pre"
+df$pre_post[df$study_id == "CC0017"] <- "Pre"
+df$pre_post[df$study_id == "CC0028"] <- "Pre"
 # Select columns for EnVision per Monique and combine race and ethnicity
@@
 df$ETI <- factor(df$ETI, levels = 0:1, labels = c("No", "Yes"))
+# Convert C-peptide level from pmol/L to ng/mL (multiply by 0.003)
+df[, grep("C.Peptide_.*", colnames(df))] <-
+  lapply(df[, grep("C.Peptide_.*", colnames(df))], function(c) {
+    c * 0.003
+  })
 # Select the same columns from GEM
@@
 df <- df %>% arrange(study_id, Date)
+# Calculate ODI, etc.
 # Write for checking
@@
 ```
-
-# Plot
-
-```{r}
-df %>%
-  group_by(study_id) %>%
-  count(pre_post) %>%
-  ggplot(., aes(y = study_id, x = n)) +
-  geom_point() +
-  theme_bw() +
-  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
-  facet_wrap(~pre_post)
-```
-
-- Adjust models for age
-- Paneled figure with results for glucose, c-peptide, and insulin
-- Look at acute insulin response (change in 30 minute c peptide)/(change in 30 minute glucose)
-  - Can also do across the whole time period and can use insulin in the numerator also
-- See Teams folder document for detailed plan
-
-# Glucose
-
-```{r}
-# Make a long DF for models
-long_df <- df %>% pivot_longer(Glucose_0:Glucose_180,
-  names_to = c(".value", "timepoint"),
-  names_pattern = "(.*)_(.*)",
-  values_to = "Glucose"
-)
-long_df$timepoint <- factor(long_df$timepoint,
-  levels = c("0", "10", "30", "60", "90", "120", "150", "180")
-)
-mod1 <- lmer(
-  Glucose ~ timepoint * pre_post + age_visit + (1 | study_id),
-  data = long_df
-)
-mod1 %>% tbl_regression(.)
-means <- data.frame(emmeans(mod1, ~ timepoint + pre_post))
-knitr::kable(means)
-ggplot(means, aes(
-  x = as.numeric(as.character(timepoint)), y = emmean,
-  color = pre_post, group = pre_post
-)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 1) +
-  geom_line() +
-  theme_bw() +
-  xlab("Timepoint") +
-  ylab("Glucose (mg/dL)") +
-  scale_x_continuous(breaks = c(0, 10, 30, 60, 90, 120, 150, 180))
-```

### 2025-08-12 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
fixed ETI plots

@@
   xlab("Years from ETI Start") +
-  ylab("gli_fev1_ppred_rn") +
-  theme_classic()
+  ylab("FEV1 % Predicted") +
+  theme_classic(base_size = 15)
 # Model summary
@@
   xlab("Years from ETI Start") +
-  ylab("bmipercentile") +
-  theme_classic()
+  ylab("BMI Percentile") +
+  theme_classic(base_size = 15)
 # Model summary
@@
   xlab("Years from ETI Start") +
-  ylab("bmivalue") +
-  theme_classic()
+  ylab("BMI (kg/m2)") +
+  theme_classic(base_size = 15)
 # Model summary

### 2025-08-12 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
new formatting
air formatting

@@
 library(pROC)
-home_dir <- switch(Sys.info()["sysname"],
+home_dir <- switch(
+    Sys.info()["sysname"],
     "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
@@
 )
-github_dir <- switch(Sys.info()["sysname"],
+github_dir <- switch(
+    Sys.info()["sysname"],
     "Darwin" = "/Users/tim/Documents/GitHub",
@@
 # Import CHCO, filter out those with no demo data, rename columns, etc.
-chco <- read.csv("./Data_Cleaned/chco_reidentified.csv",
+chco <- read.csv(
+    "./Data_Cleaned/chco_reidentified.csv",
     na.strings = c(
-    "", "Test not performed", "Insufficient quant.",
-    "repeat sweat chloride testing", "Please reorder testing and re",
+        "",
+        "Test not performed",
+        "Insufficient quant.",
+        "repeat sweat chloride testing",
+        "Please reorder testing and re",
         "The received specimen cannot"
@@
     select(
-    PatientID, Name, DOB, Genotypes1, Genotypes2, Sex, Age.at.Test.in.Years,
-    Na, K, Cl, Na_Left, Na_Right, K_Left, K_Right, Cl_Left, Cl_Right
+        PatientID,
+        Name,
+        DOB,
+        Genotypes1,
+        Genotypes2,
+        Sex,
+        Age.at.Test.in.Years,
+        Na,
+        K,
+        Cl,
+        Na_Left,
+        Na_Right,
+        K_Left,
+        K_Right,
+        Cl_Left,
+        Cl_Right
     ) %>%
@@
 # Import NJH and add diagnoses from Katie
-njh <- read.csv("./Data_Cleaned/njh_full_list.csv",
+njh <- read.csv(
+    "./Data_Cleaned/njh_full_list.csv",
     na.strings = c(
-    "", "Test not performed", "Insufficient quant.",
+        "",
+        "Test not performed",
+        "Insufficient quant.",
         "repeat sweat chloride testing"
@@
     select(
-    Patient.ID, Patient.Name, Birth.Date, Genotype.1, Genotype.2,
-    Sex, Age.at.test..in.years., Sweat.sodium, Sweat.potassium, Sweat.chloride,
-    Sweat.sodium.left.arm, Sweat.sodium.right.arm, Sweat.potassium.left.arm,
-    Sweat.potassium.right.arm, Sweat.chloride.left.arm,
-    Sweat.chloride.right.arm, Diagnosis
+        Patient.ID,
+        Patient.Name,
+        Birth.Date,
+        Genotype.1,
+        Genotype.2,
+        Sex,
+        Age.at.test..in.years.,
+        Sweat.sodium,
+        Sweat.potassium,
+        Sweat.chloride,
+        Sweat.sodium.left.arm,
+        Sweat.sodium.right.arm,
+        Sweat.potassium.left.arm,
+        Sweat.potassium.right.arm,
+        Sweat.chloride.left.arm,
+        Sweat.chloride.right.arm,
+        Diagnosis
     ) %>%
     rename(
-    Name = Patient.Name, Genotypes1 = Genotype.1, DOB = Birth.Date,
-    Genotypes2 = Genotype.2, Age = Age.at.test..in.years., Na = Sweat.sodium,
-    K = Sweat.potassium, Cl = Sweat.chloride, Na_Left = Sweat.sodium.left.arm,
-    Na_Right = Sweat.sodium.right.arm, K_Left = Sweat.potassium.left.arm,
-    K_Right = Sweat.potassium.right.arm, Cl_Left = Sweat.chloride.left.arm,
+        Name = Patient.Name,
+        Genotypes1 = Genotype.1,
+        DOB = Birth.Date,
+        Genotypes2 = Genotype.2,
+        Age = Age.at.test..in.years.,
+        Na = Sweat.sodium,
+        K = Sweat.potassium,
+        Cl = Sweat.chloride,
+        Na_Left = Sweat.sodium.left.arm,
+        Na_Right = Sweat.sodium.right.arm,
+        K_Left = Sweat.potassium.left.arm,
+        K_Right = Sweat.potassium.right.arm,
+        Cl_Left = Sweat.chloride.left.arm,
         Cl_Right = Sweat.chloride.right.arm
@@
     select(
-    Name, DOB, Diagnosis, Genotypes1, Genotypes2, Sex, Age, Na, K, Cl, `Na:Cl`
+        Name,
+        DOB,
+        Diagnosis,
+        Genotypes1,
+        Genotypes2,
+        Sex,
+        Age,
+        Na,
+        K,
+        Cl,
+        `Na:Cl`
     ) %>%
@@
 # Combine CFTR-related disorder and CRMS
-df$Diagnosis <- factor(df$Diagnosis,
+df$Diagnosis <- factor(
+    df$Diagnosis,
     levels = c("Cystic Fibrosis", "CFTR-related disorder", "CRMS", "Non-CF"),
-  labels = c("Cystic Fibrosis", "CFTR-RD or CRMS", "CFTR-RD or CRMS", "Non-CF")
+    labels = c(
+        "Cystic Fibrosis",
+        "CFTR-RD or CRMS",
+        "CFTR-RD or CRMS",
+        "Non-CF"
+    )
 )
@@
 # Remove people per Angela's recommendation and duplicates based on DOB review
-df <- df[-c(
-  78, 79, 183, 257, 259, 270, 314, 342, 691, 719, 724, 737, 812, 903, 1012,
-  1066, 1126, 1151, 1202, 1209, 1222, 1224, 1340, 1428, 1456
-), ]
+df <- df[
+    -c(
+        78,
+        79,
+        183,
+        257,
+        259,
+        270,
+        314,
+        342,
+        691,
+        719,
+        724,
+        737,
+        812,
+        903,
+        1012,
+        1066,
+        1126,
+        1151,
+        1202,
+        1209,
+        1222,
+        1224,
+        1340,
+        1428,
+        1456
+    ),
+]
 # Get genotype severity
@@
 df$geno_risk <- paste(df$geno_risk_1, df$geno_risk_2, sep = "/")
-df$geno_risk <- factor(df$geno_risk,
+df$geno_risk <- factor(
+    df$geno_risk,
     levels = c(
-    "Mild/Mild", "Mild/Severe", "Mild/Unknown", "Severe/Mild", "Severe/Severe",
-    "Severe/Unknown", "Unknown/Severe", "Unknown/Unknown", "Not CF/Not CF"
+        "Mild/Mild",
+        "Mild/Severe",
+        "Mild/Unknown",
+        "Severe/Mild",
+        "Severe/Severe",
+        "Severe/Unknown",
+        "Unknown/Severe",
+        "Unknown/Unknown",
+        "Not CF/Not CF"
     ),
@@
 df$geno1_class[grep("508", df$Genotypes1)] <- "F508del"
-df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("I", "II", "III"), "MF")
-df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("IV", "V"), "RF")
+df$geno1_class <- replace(
+    df$geno1_class,
+    df$geno1_class %in% c("I", "II", "III"),
+    "MF"
+)
+df$geno1_class <- replace(
+    df$geno1_class,
+    df$geno1_class %in% c("IV", "V"),
+    "RF"
+)
 df$geno2_class <- cf_genotype_class_severity(df$Genotypes2, out = "class")
 df$geno2_class[grep("508", df$Genotypes2)] <- "F508del"
-df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("I", "II", "III"), "MF")
-df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("IV", "V"), "RF")
+df$geno2_class <- replace(
+    df$geno2_class,
+    df$geno2_class %in% c("I", "II", "III"),
+    "MF"
+)
+df$geno2_class <- replace(
+    df$geno2_class,
+    df$geno2_class %in% c("IV", "V"),
+    "RF"
+)
 df$geno1_class[df$Diagnosis == "Non-CF"] <- "Not CF"
@@
 df$geno_class <- paste(df$geno1_class, df$geno2_class, sep = "/")
-df$geno_class <- factor(df$geno_class,
+df$geno_class <- factor(
+    df$geno_class,
     levels = c(
-    "F508del/F508del", "F508del/MF", "F508del/RF", "F508del/Unknown",
-    "MF/F508del", "MF/MF", "MF/RF", "MF/Unknown", "RF/F508del", "RF/MF",
-    "RF/Unknown", "Unknown/F508del", "Unknown/MF", "Unknown/Unknown",
+        "F508del/F508del",
+        "F508del/MF",
+        "F508del/RF",
+        "F508del/Unknown",
+        "MF/F508del",
+        "MF/MF",
+        "MF/RF",
+        "MF/Unknown",
+        "RF/F508del",
+        "RF/MF",
+        "RF/Unknown",
+        "Unknown/F508del",
+        "Unknown/MF",
+        "Unknown/Unknown",
         "Not CF/Not CF"
@@
     labels = c(
-    "MF/MF", "MF/MF", "MF/RF", "MF/U", "MF/MF", "MF/MF", "MF/RF", "MF/U",
-    "MF/RF", "MF/RF", "RF/U", "MF/U", "MF/U", "U", "Not CF"
+        "MF/MF",
+        "MF/MF",
+        "MF/RF",
+        "MF/U",
+        "MF/MF",
+        "MF/MF",
+        "MF/RF",
+        "MF/U",
+        "MF/RF",
+        "MF/RF",
+        "RF/U",
+        "MF/U",
+        "MF/U",
+        "U",
+        "Not CF"
     )
@@
 # Save this for Angela et al. to check
-write.csv(df,
+write.csv(
+    df,
     file = "./Data_Cleaned/sweat_variability_analysis_dataset.csv",
-  row.names = F, na = ""
+    row.names = F,
+    na = ""
 )
@@
 ```{r results='asis'}
-summary(tableby(Diagnosis ~ Sex + Age + Na + Cl + `Na:Cl` + K,
-  data = df, control = mycontrols
+summary(tableby(
+    Diagnosis ~ Sex + Age + Na + Cl + `Na:Cl` + K,
+    data = df,
+    control = mycontrols
 ))
@@
 ```{r results='asis'}
-summary(tableby(`Genotype Class` ~ Sex + Age + Na + Cl + `Na:Cl` + K,
-  data = df, control = mycontrols
+summary(tableby(
+    `Genotype Class` ~ Sex + Age + Na + Cl + `Na:Cl` + K,
+    data = df,
+    control = mycontrols
 ))
@@
 ```{r results='asis'}
-summary(tableby(`Genotype Risk` ~ Sex + Age + Na + Cl + `Na:Cl` + K,
-  data = df, control = mycontrols
+summary(tableby(
+    `Genotype Risk` ~ Sex + Age + Na + Cl + `Na:Cl` + K,
+    data = df,
+    control = mycontrols
 ))
@@
 roc_table <- lapply(roc_list, function(r) {
-  coords(r, "best",
+    coords(
+        r,
+        "best",
         best.method = "youden",
         ret = c(
-      "threshold", "sensitivity", "specificity", "accuracy", "tpr", "fpr",
-      "tnr", "fnr"
+            "threshold",
+            "sensitivity",
+            "specificity",
+            "accuracy",
+            "tpr",
+            "fpr",
+            "tnr",
+            "fnr"
         )
@@
 ```{r}
-plot(roc_list[[2]],
-  print.auc = T, print.thres = T,
+plot(
+    roc_list[[2]],
+    print.auc = T,
+    print.thres = T,
     main = "Cystic Fibrosis vs. CFTR-RD or CRMS"
 )
-plot(roc_list[[1]],
-  print.auc = T, print.thres = T,
+plot(
+    roc_list[[1]],
+    print.auc = T,
+    print.thres = T,
     main = "Cystic Fibrosis vs. Non-CF"
 )
-plot(roc_list[[3]],
-  print.auc = T, print.thres = T,
+plot(
+    roc_list[[3]],
+    print.auc = T,
+    print.thres = T,
     main = "CFTR-RD or CRMS vs. Non-CF"

### 2025-08-12 Christine Chan/EnVision CF/cgm_and_a1c.qmd
formatting for HbA1c and CGM

@@
 #| include: false
+library(keyring)
 library(tidyverse)
@@
 library(blandr)
-home_dir <- switch(Sys.info()["sysname"],
+home_dir <- switch(
+    Sys.info()["sysname"],
     "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
@@
 # Import from REDCap
-unlockREDCap(c(rcon = "EnVision CF HbA1c and CGM"),
-  keyring = "API_KEYs",
-  envir = 1,
-  url = "https://redcap.ahc.umn.edu/redcap/api/"
+token <- key_get("envision_cf_hba1c_cgm")
+rcon <- redcapConnection(
+    url = "https://redcap.ahc.umn.edu/redcap/api/",
+    token = token
 )
@@
         # Numeric
-    across(c(age, bmi, cfrd_diagnosis_age, average_sensor_glucose), ~
-      as.numeric(.)),
-    across(c(hb_a1c_value, gmi), ~
-      as.numeric(sub("%", "", .))),
+        across(
+            c(age, bmi, cfrd_diagnosis_age, average_sensor_glucose),
+            ~ as.numeric(.)
+        ),
+        across(c(hb_a1c_value, gmi), ~ as.numeric(sub("%", "", .))),
         # Categorical
-    combined_race = factor(race,
+        combined_race = factor(
+            race,
             levels = c(
-        "White", "Black or African American", "Asian", "More Than One Race",
+                "White",
+                "Black or African American",
+                "Asian",
+                "More Than One Race",
                 "Unknown / Not Reported"
@@
             labels = c(
-        "White", "Black or African American", "Other/Unknown", "Other/Unknown",
+                "White",
+                "Black or African American",
+                "Other/Unknown",
+                "Other/Unknown",
                 "Other/Unknown"
@@
     select(
-    subject_id, age, cfrd_diagnosis_age, bmi, combined_race, sex, pi,
-    cftr_modulator_use, type_of_modulator, type_of_cgm, hb_a1c_value, gmi,
+        subject_id,
+        age,
+        cfrd_diagnosis_age,
+        bmi,
+        combined_race,
+        sex,
+        pi,
+        cftr_modulator_use,
+        type_of_modulator,
+        type_of_cgm,
+        hb_a1c_value,
+        gmi,
         average_sensor_glucose
@@
     set_variable_labels(
-    age = "Age (years)", cfrd_diagnosis_age = "Age at CFRD Diagnosis (years)",
-    bmi = "BMI (kg/m2)", combined_race = "Race", sex = "Sex at Birth",
-    type_of_modulator = "Type of Modulator", hb_a1c_value = "HbA1c (%)",
-    gmi = "GMI (%)", average_sensor_glucose = "Mean Sensor Glucose (mg/dL)",
-    cftr_modulator_use = "CFTR Modulator", pi = "Pancreatic Insufficiency"
+        age = "Age (years)",
+        cfrd_diagnosis_age = "Age at CFRD Diagnosis (years)",
+        bmi = "BMI (kg/m2)",
+        combined_race = "Race",
+        sex = "Sex at Birth",
+        type_of_modulator = "Type of Modulator",
+        hb_a1c_value = "HbA1c (%)",
+        gmi = "GMI (%)",
+        average_sensor_glucose = "Mean Sensor Glucose (mg/dL)",
+        cftr_modulator_use = "CFTR Modulator",
+        pi = "Pancreatic Insufficiency"
     )
@@
 #| warning: false
-ggplot(df, aes(x = hb_a1c_value, y = gmi)) +
+ggplot(df, aes(x = gmi, y = hb_a1c_value)) +
     geom_point() +
@@
 
-blandr.draw(df$hb_a1c_value, df$gmi, ciDisplay = F, annotate = T)
+blandr.draw(df$gmi, df$hb_a1c_value, ciDisplay = F, annotate = T)
 ```
 
-# GMI and HbA1c
+# Mean sensor glucose and HbA1c
 
@@
 #| warning: false
-ggplot(df, aes(x = hb_a1c_value, y = average_sensor_glucose)) +
+ggplot(df, aes(x = average_sensor_glucose, y = hb_a1c_value)) +
     geom_point() +
@@
     stat_poly_eq(use_label(c("eq", "R2", "p", "n")))
-
-blandr.draw(df$hb_a1c_value, df$average_sensor_glucose,
-  ciDisplay = F, annotate = T
-)
 ```

### 2025-08-12 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
added method and filtered Age > 50 - need to figure out the multiple methods issue

@@
         Age.at.Test.in.Years,
+        Method,
         Na,
@@
         Age.at.test..in.years.,
+        Test.Method,
         Sweat.sodium,
@@
         Age = Age.at.test..in.years.,
+        Method = Test.Method,
         Na = Sweat.sodium,
@@
 df <- rbind(chco, njh)
+method <- df %>%
+    select(Name, Age, Method, Na, Cl) %>%
+    distinct() %>%
+    drop_na()
 df <- df %>%
+    select(-Method) %>%
     group_by(Name) %>%
@@
     ),
-    labels = c("M/M", "M/S", "M/U", "M/S", "S/S", "S/U", "S/U", "U", "Not CF")
-)
-## By function
-df$geno1_class <- cf_genotype_class_severity(df$Genotypes1, out = "class")
-df$geno1_class[grep("508", df$Genotypes1)] <- "F508del"
-df$geno1_class <- replace(
-    df$geno1_class,
-    df$geno1_class %in% c("I", "II", "III"),
-    "MF"
-)
-df$geno1_class <- replace(
-    df$geno1_class,
-    df$geno1_class %in% c("IV", "V"),
-    "RF"
-)
-df$geno2_class <- cf_genotype_class_severity(df$Genotypes2, out = "class")
-df$geno2_class[grep("508", df$Genotypes2)] <- "F508del"
-df$geno2_class <- replace(
-    df$geno2_class,
-    df$geno2_class %in% c("I", "II", "III"),
-    "MF"
-)
-df$geno2_class <- replace(
-    df$geno2_class,
-    df$geno2_class %in% c("IV", "V"),
-    "RF"
-)
-df$geno1_class[df$Diagnosis == "Non-CF"] <- "Not CF"
-df$geno2_class[df$Diagnosis == "Non-CF"] <- "Not CF"
-df$geno_class <- paste(df$geno1_class, df$geno2_class, sep = "/")
-df$geno_class <- factor(
-    df$geno_class,
-    levels = c(
-        "F508del/F508del",
-        "F508del/MF",
-        "F508del/RF",
-        "F508del/Unknown",
-        "MF/F508del",
-        "MF/MF",
-        "MF/RF",
-        "MF/Unknown",
-        "RF/F508del",
-        "RF/MF",
-        "RF/Unknown",
-        "Unknown/F508del",
-        "Unknown/MF",
-        "Unknown/Unknown",
-        "Not CF/Not CF"
-    ),
     labels = c(
-        "MF/MF",
-        "MF/MF",
-        "MF/RF",
-        "MF/U",
-        "MF/MF",
-        "MF/MF",
-        "MF/RF",
-        "MF/U",
-        "MF/RF",
-        "MF/RF",
-        "RF/U",
-        "MF/U",
-        "MF/U",
-        "U",
+        "Mild",
+        "Mild",
+        "Unknown",
+        "Mild",
+        "Severe",
+        "Unknown",
+        "Unknown",
+        "Unknown",
         "Not CF"
@@
 )
-# Organize columns
+# Create a categorical column for Cl
+df$`Cl Group` <- cut(
+    df$Cl,
+    breaks = c(-Inf, 30, 60, Inf),
+    right = F,
+    labels = c("< 30", "30-59", ">= 60")
+)
+# Add method back in
+df <- left_join(df, method)
+# Organize columns, remove age over 50 (there shouldn't be any age < 2 days)
 df <- df %>%
-    select(Name:Genotypes2, geno_risk, geno_class, everything()) %>%
-    select(-geno_risk_1, -geno_risk_2, -geno1_class, -geno2_class) %>%
-    rename(`Genotype Risk` = geno_risk, `Genotype Class` = geno_class)
+    select(Name:Genotypes2, geno_risk, everything()) %>%
+    select(-geno_risk_1, -geno_risk_2) %>%
+    rename(`Genotype Risk` = geno_risk) %>%
+    filter(Age < 50)
 # Set table options
 mycontrols <- tableby.control(
-    numeric.stats = c("N", "Nmiss", "meansd", "medianq1q3", "range")
+    numeric.stats = c("N", "Nmiss", "meansd", "medianq1q3", "iqr", "range")
 )
@@
 ```{r results='asis'}
-summary(tableby(
-    Diagnosis ~ Sex + Age + Na + Cl + `Na:Cl` + K,
+summary(
+    tableby(
+        Diagnosis ~
+            Method +
+                Sex +
+                Age +
+                fe(`Genotype Risk`) +
+                fe(`Cl Group`) +
+                Cl +
+                Na +
+                `Na:Cl` +
+                K,
+        data = df,
+        simulate.p.value = T,
+        control = mycontrols
+    ),
+    pfootnote = T
+)
+```
+
+## Table 2: Participant Demographics by Genotype Risk
+
+```{r results='asis'}
+summary(
+    tableby(
+        `Genotype Risk` ~
+            Sex +
+                Age +
+                fe(`Cl Group`) +
+                Cl +
+                Na +
+                `Na:Cl` +
+                K,
+        data = df[
+            df$`Genotype Risk` != "Unknown" & df$Diagnosis == "Cystic Fibrosis",
+        ],
+        simulate.p.value = T,
+        control = mycontrols
+    ),
+    pfootnote = T
+)
+```
+
+## Table 3: Participant Demographics by Chloride Group
+
+```{r results='asis'}
+summary(
+    tableby(
+        `Cl Group` ~ Sex + Age + Cl + Na + `Na:Cl` + K,
         data = df,
         control = mycontrols
-))
-```
-
-## Table 2: Participant Demographics by Genotype Class
-
-In the table below "MF" stands for minimal function, "RF" for residual function, and "U" indicates unknown.
-
-```{r results='asis'}
-summary(tableby(
-    `Genotype Class` ~ Sex + Age + Na + Cl + `Na:Cl` + K,
-    data = df,
-    control = mycontrols
-))
-```
-
-## Table 3: Participant Demographics by Genotype Risk
-
-In the table below "M" stands for mild, "S" for severe, and "U" indicates unknown.
-
-```{r results='asis'}
-summary(tableby(
-    `Genotype Risk` ~ Sex + Age + Na + Cl + `Na:Cl` + K,
-    data = df,
-    control = mycontrols
-))
+    ),
+    pfootnote = T
+)
 ```
@@
 clp <- ggplot(df, aes(x = Diagnosis, y = Cl)) +
-    geom_boxplot() +
-    theme_bw()
-clp
-```
-
-#### By genotype class
-
-```{r}
-#| warning: false
-clp <- ggplot(df, aes(x = `Genotype Class`, y = Cl)) +
     geom_boxplot() +
@@
 nap <- ggplot(df, aes(x = Diagnosis, y = Na)) +
-    geom_boxplot() +
-    theme_bw()
-nap
-```
-
-#### By genotype class
-
-```{r}
-#| warning: false
-nap <- ggplot(df, aes(x = `Genotype Class`, y = Na)) +
     geom_boxplot() +
@@
 ratiop <- ggplot(df, aes(x = Diagnosis, y = `Na:Cl`)) +
-    geom_boxplot() +
-    geom_hline(yintercept = 1, color = "red") +
-    geom_hline(yintercept = 1.25, color = "red") +
-    ylim(0, 2) +
-    theme_bw() +
-    ylab("Na:Cl") +
-    xlab("Modulator")
-ratiop
-```
-
-#### By genotype class
-
-```{r}
-#| warning: false
-ratiop <- ggplot(df, aes(x = `Genotype Class`, y = `Na:Cl`)) +
-    geom_boxplot() +
-    geom_hline(yintercept = 1, color = "red") +
-    geom_hline(yintercept = 1.25, color = "red") +
-    theme_bw() +
-    ylab("Na:Cl") +
-    xlab("Modulator")
-ratiop
-ratiop <- ggplot(df, aes(x = `Genotype Class`, y = `Na:Cl`)) +
     geom_boxplot() +

### 2025-08-15 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
air formatting

--- 2025-07-25 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
@@
 library(emmeans)
-knitr::opts_knit$set(root.dir = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF")
+home_dir <- switch(Sys.info()["sysname"],
+    "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
+    "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
+    "Linux" = "/home/tim/OneDrive/Vigers/CF"
+)
+github_dir <- switch(Sys.info()["sysname"],
+    "Darwin" = "/Users/tim/Documents/GitHub",
+    "Windows" = "C:/Users/Tim/Documents/GitHub",
+    "Linux" = "/home/tim/Documents/GitHub"
+)
+knitr::opts_knit$set(root.dir = home_dir)
 ```
@@
 # source("/home/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R")
-load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData")
+load(
+    "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData"
+)
 # Change age group reference level
@@
         type = list(`Years With Data` ~ "continuous"),
-    statistic = list(c(gli_fev1_ppred_rn, gli_fvc_ppred_rn) ~ "{mean} [{sd}]"),
+        statistic = list(
+            c(gli_fev1_ppred_rn, gli_fvc_ppred_rn) ~ "{mean} [{sd}]"
+        ),
         missing_text = "Missing"
@@
 encounter %>%
-  select(Race, encounterlocation, bacterialculturedone:othermicroorganisms1) %>%
+    select(
+        Race,
+        encounterlocation,
+        bacterialculturedone:othermicroorganisms1
+    ) %>%
     tbl_summary(by = Race, missing_text = "Missing")
@@
 # Create change in slope variable
-encounter$slope_change_days <- (encounter$Days > 0) * encounter$Days
-label(encounter$slope_change_days) <- "Change in Slope"
+encounter$slope_c_days <- (encounter$Days > 0) * encounter$Days
+label(encounter$slope_c_days) <- "Change in Slope"
 encounter$Years <- encounter$Days / 365.25
-encounter$slope_change_years <- (encounter$Years > 0) * encounter$Years
-label(encounter$slope_change_years) <- "Change in Slope"
+encounter$slope_c_years <- (encounter$Years > 0) * encounter$Years
+label(encounter$slope_c_years) <- "Change in Slope"
 # Prior modulator use
@@
     group_by(eDWID) %>%
-  summarise(prior_modulator = any(vx661comb == 1) | any(vx445comb == 1) |
-    any(Vx770 == 1) | any(Vx809comb))
+    summarise(
+        prior_modulator = any(vx661comb == 1) |
+            any(vx445comb == 1) |
+            any(Vx770 == 1) |
+            any(Vx809comb)
+    )
 prior_mods$prior_modulator <-
     fct_na_value_to_level(factor(prior_mods$prior_modulator), level = "No")
-prior_mods$prior_modulator <- factor(prior_mods$prior_modulator,
+prior_mods$prior_modulator <- factor(
+    prior_mods$prior_modulator,
     levels = c(T, "No"),
@@
 encounter$pseudomonasaeruginosa[is.na(encounter$pseudomonasaeruginosa)] <- 0
-encounter$pseudomonasaeruginosa <- factor(encounter$pseudomonasaeruginosa,
-  levels = 0:1, labels = c("No", "Yes")
+encounter$pseudomonasaeruginosa <- factor(
+    encounter$pseudomonasaeruginosa,
+    levels = 0:1,
+    labels = c("No", "Yes")
 )
@@
 change_mod <- lmer(
-  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + age_group +
-    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    gli_fev1_ppred_rn ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        age_group +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
         (1 | eDWID),
@@
 pop_dat <- data.frame(expand.grid(
-  Years = -3:2, Race = unique(encounter$Race), age_group = "[18,30)",
-  Insurance = "Private", prior_modulator = "Yes",
-  isOnEnzymes = "Yes", pseudomonasaeruginosa = "No"
+    Years = -3:2,
+    Race = unique(encounter$Race),
+    age_group = "[18,30)",
+    Insurance = "Private",
+    prior_modulator = "Yes",
+    isOnEnzymes = "Yes",
+    pseudomonasaeruginosa = "No"
 ))
-pop_dat$slope_change_years <- (pop_dat$Years > 0) * pop_dat$Years
-pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
-pop_dat$pred <- pred$fit
-pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
-pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
-ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
-  geom_line() +
-  xlab("Years from ETI Start") +
-  ylab("gli_fev1_ppred_rn") +
-  theme_classic()
+pop_dat$slope_c_years <- (pop_dat$Years > 0) * pop_dat$Years
+pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
+    geom_point() +
+    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+    geom_line() +
+    xlab("Years from ETI Start") +
+    ylab("FEV1 % Predicted") +
+    theme_classic(base_size = 15)
 # Model summary
@@
 pop_dat %>%
-  select(-slope_change_years) %>%
+    select(-slope_c_years) %>%
     rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
@@
 aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
-aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
+aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
 hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
-hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
+hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
 nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
-nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
 other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
-other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
-contr <- rbind(
-  "Black/AA Pre-ETI" = aa_pre,
-  "Black/AA Post-ETI" = aa_pre + aa_change,
-  "H/L Pre-ETI" = hl_pre,
-  "H/L Post-ETI" = hl_pre + hl_change,
-  "NHW Pre-ETI" = nhw_pre,
-  "NHW Post-ETI" = nhw_pre + nhw_change,
-  "Other Pre-ETI" = other_pre,
-  "Other Post-ETI" = other_pre + other_change
+other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
+contr <- rbind(
+    "Black/AA Pre-ETI" = aa_pre,
+    "Black/AA Post-ETI" = aa_pre + aa_c,
+    "H/L Pre-ETI" = hl_pre,
+    "H/L Post-ETI" = hl_pre + hl_c,
+    "NHW Pre-ETI" = nhw_pre,
+    "NHW Post-ETI" = nhw_pre + nhw_c,
+    "Other Pre-ETI" = other_pre,
+    "Other Post-ETI" = other_pre + other_c
 )
@@
 change_mod <- lmer(
-  GLI_FEV1_pct_predicted ~ Years * Race + slope_change_years * Race +
-    age_group + Insurance + prior_modulator + isOnEnzymes +
-    pseudomonasaeruginosa + (1 | eDWID),
+    GLI_FEV1_pct_predicted ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        age_group +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
+        (1 | eDWID),
     data = encounter
@@
 pop_dat %>%
-  select(-slope_change_years) %>%
+    select(-slope_c_years) %>%
     rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
@@
 change_mod <- lmer(
-  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
-    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
+    gli_fev1_ppred_rn ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
+        (1 | eDWID),
     data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
@@
 aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
-aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
+aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
 hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
-hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
+hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
 nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
-nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
 other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
-other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
-contr <- rbind(
-  "Black/AA Pre-ETI" = aa_pre,
-  "Black/AA Post-ETI" = aa_pre + aa_change,
-  "H/L Pre-ETI" = hl_pre,
-  "H/L Post-ETI" = hl_pre + hl_change,
-  "NHW Pre-ETI" = nhw_pre,
-  "NHW Post-ETI" = nhw_pre + nhw_change,
-  "Other Pre-ETI" = other_pre,
-  "Other Post-ETI" = other_pre + other_change
+other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
+contr <- rbind(
+    "Black/AA Pre-ETI" = aa_pre,
+    "Black/AA Post-ETI" = aa_pre + aa_c,
+    "H/L Pre-ETI" = hl_pre,
+    "H/L Post-ETI" = hl_pre + hl_c,
+    "NHW Pre-ETI" = nhw_pre,
+    "NHW Post-ETI" = nhw_pre + nhw_c,
+    "Other Pre-ETI" = other_pre,
+    "Other Post-ETI" = other_pre + other_c
 )
@@
 change_mod <- lmer(
-  GLI_FEV1_pct_predicted ~ Years * Race + slope_change_years * Race +
-    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    GLI_FEV1_pct_predicted ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
         (1 | eDWID),
@@
 change_mod <- lmer(
-  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
-    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
+    gli_fev1_ppred_rn ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
+        (1 | eDWID),
     data = encounter[encounter$age_eti_group == "[18, Inf)", ]
@@
 change_mod <- lmer(
-  gli_fvc_ppred_rn ~ Years * Race + slope_change_years * Race + age_group +
-    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    gli_fvc_ppred_rn ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        age_group +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
         (1 | eDWID),
@@
 pop_dat %>%
-  select(-slope_change_years) %>%
+    select(-slope_c_years) %>%
     rename(gli_fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
@@
 change_mod <- lmer(
-  GLI_FVC_pct_predicted ~ Years * Race + slope_change_years * Race +
-    age_group + Insurance + prior_modulator + isOnEnzymes +
-    pseudomonasaeruginosa + (1 | eDWID),
+    GLI_FVC_pct_predicted ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        age_group +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
+        (1 | eDWID),
     data = encounter
@@
 pop_dat %>%
-  select(-slope_change_years) %>%
+    select(-slope_c_years) %>%
     rename(GLI_FVC_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
@@
 change_mod <- lmer(
-  gli_fvc_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
-    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
+    gli_fvc_ppred_rn ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
+        (1 | eDWID),
     data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
@@
 change_mod <- lmer(
-  GLI_FVC_pct_predicted ~ Years * Race + slope_change_years * Race +
-    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    GLI_FVC_pct_predicted ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
         (1 | eDWID),
@@
 change_mod <- lmer(
-  gli_fvc_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
-    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
+    gli_fvc_ppred_rn ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
+        (1 | eDWID),
     data = encounter[encounter$age_eti_group == "[18, Inf)", ]
@@
 change_mod <- lmer(
-  gli_fev1fvc_ppred_rn ~ Years * Race + slope_change_years * Race + age_group +
-    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    gli_fev1fvc_ppred_rn ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        age_group +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
         (1 | eDWID),
@@
 pop_dat %>%
-  select(-slope_change_years) %>%
+    select(-slope_c_years) %>%
     rename(gli_fev1fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
@@
 change_mod <- lmer(
-  GLI_FEV1FVC_pct_predicted ~ Years * Race + slope_change_years * Race +
-    age_group + Insurance + prior_modulator + isOnEnzymes +
-    pseudomonasaeruginosa + (1 | eDWID),
+    GLI_FEV1FVC_pct_predicted ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        age_group +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
+        (1 | eDWID),
     data = encounter
@@
 pop_dat %>%
-  select(-slope_change_years) %>%
+    select(-slope_c_years) %>%
     rename(GLI_FEV1FVC_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
@@
 change_mod <- lmer(
-  gli_fev1fvc_ppred_rn ~ Years * Race + slope_change_years * Race +
-    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    gli_fev1fvc_ppred_rn ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
         (1 | eDWID),
@@
 change_mod <- lmer(
-  GLI_FEV1FVC_pct_predicted ~ Years * Race + slope_change_years * Race +
-    Insurance + prior_modulator + isOnEnzymes +
-    pseudomonasaeruginosa + (1 | eDWID),
+    GLI_FEV1FVC_pct_predicted ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
+        (1 | eDWID),
     data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
@@
 change_mod <- lmer(
-  gli_fev1fvc_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
-    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
+    gli_fev1fvc_ppred_rn ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
+        (1 | eDWID),
     data = encounter[encounter$age_eti_group == "[18, Inf)", ]
@@
 change_mod <- lmer(
-  GLI_FEV1FVC_pct_predicted ~ Years * Race + slope_change_years * Race +
-    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    GLI_FEV1FVC_pct_predicted ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
         (1 | eDWID),
@@
 # )
-pex_logistic <- glmer(any_pex ~ ETI * Race + (1 | eDWID),
+pex_logistic <- glmer(
+    any_pex ~ ETI * Race + (1 | eDWID),
     data = pex_data,
@@
 ```{r}
-pex_logistic <- glmer(any_pex ~ ETI * Race + (1 | eDWID),
+pex_logistic <- glmer(
+    any_pex ~ ETI * Race + (1 | eDWID),
     data = pex_data[pex_data$age_eti_group == "[-Inf,18)", ],
@@
 ```{r}
-pex_logistic <- glmer(any_pex ~ ETI * Race + (1 | eDWID),
+pex_logistic <- glmer(
+    any_pex ~ ETI * Race + (1 | eDWID),
     data = pex_data[pex_data$age_eti_group == "[18, Inf)", ],
@@
 change_mod <- lmer(
-  bmipercentile ~ Years * Race + slope_change_years * Race +
-    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    bmipercentile ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
         (1 | eDWID),
@@
     xlab("Years from ETI Start") +
-  ylab("bmipercentile") +
-  theme_classic()
+    ylab("BMI Percentile") +
+    theme_classic(base_size = 15)
 # Model summary
@@
 pop_dat %>%
-  select(-slope_change_years) %>%
+    select(-slope_c_years) %>%
     rename(bmipercentile = pred, ci.ll = ll, ci.ul = ul) %>%
@@
 change_mod <- lmer(
-  bmivalue ~ Years * Race + slope_change_years * Race +
-    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    bmivalue ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
         (1 | eDWID),
@@
     xlab("Years from ETI Start") +
-  ylab("bmivalue") +
-  theme_classic()
+    ylab("BMI (kg/m2)") +
+    theme_classic(base_size = 15)
 # Model summary
@@
 pop_dat %>%
-  select(-slope_change_years) %>%
+    select(-slope_c_years) %>%
     rename(bmivalue = pred, ci.ll = ll, ci.ul = ul) %>%
@@
 change_mod <- lmer(
-  heightpercentile ~ Years * Race + slope_change_years * Race +
-    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    heightpercentile ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
         (1 | eDWID),
@@
 pop_dat %>%
-  select(-slope_change_years) %>%
+    select(-slope_c_years) %>%
     rename(heightpercentile = pred, ci.ll = ll, ci.ul = ul) %>%
@@
 change_mod <- lmer(
-  height ~ Years * Race + slope_change_years * Race + Insurance +
-    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
+    height ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
+        (1 | eDWID),
     data = encounter[encounter$age_eti_start >= 20, ]
@@
 pop_dat %>%
-  select(-slope_change_years) %>%
+    select(-slope_c_years) %>%
     rename(height = pred, ci.ll = ll, ci.ul = ul) %>%
@@
 change_mod <- lmer(
-  weightpercentile ~ Years * Race + slope_change_years * Race +
-    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
+    weightpercentile ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
         (1 | eDWID),
@@
 pop_dat %>%
-  select(-slope_change_years) %>%
+    select(-slope_c_years) %>%
     rename(weightpercentile = pred, ci.ll = ll, ci.ul = ul) %>%
@@
 change_mod <- lmer(
-  weight ~ Years * Race + slope_change_years * Race + Insurance +
-    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
+    weight ~
+        Years *
+        Race +
+        slope_c_years * Race +
+        Insurance +
+        prior_modulator +
+        isOnEnzymes +
+        pseudomonasaeruginosa +
+        (1 | eDWID),
     data = encounter[encounter$age_eti_start >= 20, ]
@@
 pop_dat %>%
-  select(-slope_change_years) %>%
+    select(-slope_c_years) %>%
     rename(weight = pred, ci.ll = ll, ci.ul = ul) %>%
@@
 aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
-aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
+aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
 hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
-hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
+hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
 nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
-nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
 other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
-other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
-contr <- rbind(
-  "Black/AA Pre-ETI" = aa_pre,
-  "Black/AA Post-ETI" = aa_pre + aa_change,
-  "H/L Pre-ETI" = hl_pre,
-  "H/L Post-ETI" = hl_pre + hl_change,
-  "NHW Pre-ETI" = nhw_pre,
-  "NHW Post-ETI" = nhw_pre + nhw_change,
-  "Other Pre-ETI" = other_pre,
-  "Other Post-ETI" = other_pre + other_change
-)
-change_mod_slopes <- glht(change_mod, linfct = contr)
-change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
-change_mod_slopes_ci$confint %>%
-  as.data.frame() %>%
-  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
-  fmt_number(decimals = 3)
-```
+other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
+contr <- rbind(
+    "Black/AA Pre-ETI" = aa_pre,
+    "Black/AA Post-ETI" = aa_pre + aa_c,
+    "H/L Pre-ETI" = hl_pre,
+    "H/L Post-ETI" = hl_pre + hl_c,
+    "NHW Pre-ETI" = nhw_pre,
+    "NHW Post-ETI" = nhw_pre + nhw_c,
+    "Other Pre-ETI" = other_pre,
+    "Other Post-ETI" = other_pre + other_c
+)
+change_mod_slopes <- glht(change_mod, linfct = contr)
+change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
+change_mod_slopes_ci$confint %>%
+    as.data.frame() %>%
+    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
+    fmt_number(decimals = 3)
+```

### 2025-08-15 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
nicer table formatting

@@
 change_mod %>%
-    gtsummary::tbl_regression(intercept = T) %>%
-    as_gt() %>%
-    fmt_number(decimals = 3)
+    gtsummary::tbl_regression(
+        intercept = T,
+        estimate_fun = ~ style_number(.x, digits = 1),
+        pvalue_fun = ~ style_pvalue(.x, digits = 3)
+    )
 # Table of predicted values
@@
     gt(caption = "Predicted Values at Various Timepoints") %>%
-    fmt_number(decimals = 3)
+    fmt_number(decimals = 1)
 # Contrasts
@@
     gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
-    fmt_number(decimals = 3)
+    fmt_number(decimals = 1)
 ```
@@
     gt() %>%
-    fmt_number(decimals = 3)
+    fmt_number(decimals = 1)
 # Contrasts
@@
 # Model summary
-change_mod %>% gtsummary::tbl_regression(intercept = T)
+change_mod %>%
+    gtsummary::tbl_regression(
+        intercept = T,
+        estimate_fun = ~ style_number(.x, digits = 1),
+        pvalue_fun = ~ style_pvalue(.x, digits = 3)
+    )
 # Table of predicted values
@@
 change_mod %>%
-    gtsummary::tbl_regression(intercept = T) %>%
-    as_gt() %>%
-    fmt_number(decimals = 3)
+    gtsummary::tbl_regression(
+        intercept = T,
+        estimate_fun = ~ style_number(.x, digits = 2),
+        pvalue_fun = ~ style_pvalue(.x, digits = 3)
+    )
 # Table of predicted values
@@
     gt(caption = "Probabilities by Race", rownames_to_stub = T) %>%
-    fmt_number(decimals = 3)
+    fmt_number(decimals = 1)
 ```
@@
     gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
-    fmt_number(decimals = 3)
-```
+    fmt_number(decimals = 1)
+```

### 2025-08-15 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
genotype risk ROC curves, etc.

@@
 library(pROC)
-home_dir <- switch(
-    Sys.info()["sysname"],
+home_dir <- switch(Sys.info()["sysname"],
     "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
@@
 )
-github_dir <- switch(
-    Sys.info()["sysname"],
+github_dir <- switch(Sys.info()["sysname"],
     "Darwin" = "/Users/tim/Documents/GitHub",
@@
 # Add method back in
-df <- left_join(df, method)
+df <- left_join(df, method) %>% arrange(Name, desc(Method))
 # Organize columns, remove age over 50 (there shouldn't be any age < 2 days)
@@
     filter(Age < 50)
+df <- df[-which(duplicated(df$Name)), ]
 # Set table options
@@
 ```
+
+# Data Cleaning
+
+- Started with 5,046 sweat tests with 4,646 unique individuals.
+- 1,551 did not have demographics available.
+- 399 with Cl < 10 and 1 with Cl > 160
+- 156 with Na < 10
+- 19 over age 50 and 3 under age 2 days
+- 43 with insufficient quantity 
+- 25 on modulators
 
@@
 
+### By Diagnosis
+
 ```{r}
@@
 
-### Individual ROC Curves
+#### Individual ROC Curves
 
@@
 )
+```
+
+### By genotype risk
+
+```{r}
+# Fit a multiclass ROC
+gr_roc <- roc(
+    `Genotype Risk` ~ `Na:Cl`,
+    data = df,
+    levels = c("Mild", "Severe"),
+    quiet = T
+)
+plot(
+    gr_roc,
+    print.auc = T,
+    print.thres = T,
+    main = "Mild vs. Severe Genotype"
+)
+kable(coords(
+    gr_roc,
+    "best",
+    best.method = "youden",
+    ret = c(
+        "threshold",
+        "sensitivity",
+        "specificity",
+        "accuracy",
+        "tpr",
+        "fpr",
+        "tnr",
+        "fnr"
+    )
+))
 ```
@@
 ```
+
+### By genotype risk
+
+```{r}
+# Fit a multiclass ROC
+gr_roc <- roc(
+    `Genotype Risk` ~ `Na:Cl`,
+    data = df,
+    levels = c("Mild", "Severe"),
+    quiet = T
+)
+plot(
+    gr_roc,
+    print.auc = T,
+    print.thres = T,
+    main = "Mild vs. Severe Genotype"
+)
+kable(coords(
+    gr_roc,
+    "best",
+    best.method = "youden",
+    ret = c(
+        "threshold",
+        "sensitivity",
+        "specificity",
+        "accuracy",
+        "tpr",
+        "fpr",
+        "tnr",
+        "fnr"
+    )
+))
+```

### 2025-09-04 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
added PABAK, etc.

--- 2025-02-13 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
@@
 #| include: false
+library(keyring)
 library(redcapAPI)
@@
 library(caret)
+library(epiR)
 library(eulerr)
 library(ggpubr)
+home_dir <- switch(Sys.info()["sysname"],
+    "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Jordana Hoppe/SSwAB-IT",
+    "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Jordana Hoppe/SSwAB-IT"
+)
+knitr::opts_knit$set(root.dir = home_dir)
 ```
@@
 # Import from REDCap
-unlockREDCap(c(rcon = "SSwAB-IT (New)"),
-  keyring = "API_KEYs",
-  envir = 1,
+token <- key_get("SSwAB-IT (New)")
+rcon <- redcapConnection(
+    token = token,
     url = "https://redcap.ucdenver.edu/api/"
@@
 df <- full_join(db_cos_collection_results, db_sos_collection_results)
-df <- full_join(df,
+df <- full_join(
+    df,
     db_selfobtained_sample[, c("record_id", "who_collected_the_sos")],
@@
 # Add dates to demographics
-db_demographics <- left_join(db_demographics,
-  db_clinic_obtained_values %>% select(record_id, cos_collection_date),
+db_demographics <- left_join(
+    db_demographics,
+    db_clinic_obtained_values %>%
+        select(record_id, cos_collection_date, ppfev1),
     by = join_by(record_id)
@@
 db_demographics$age_at_cos <-
-  as.numeric(difftime(db_demographics$cos_collection_date,
+    as.numeric(difftime(
+        db_demographics$cos_collection_date,
         db_demographics$dob,
         units = "days"
-  )) / 365.25
+    )) /
+        365.25
 # 1-5, 6-11, 12-17 and 18+ years
-db_demographics$age_group <- cut(db_demographics$age_at_cos,
+db_demographics$age_group <- cut(
+    db_demographics$age_at_cos,
     c(-Inf, 6, 12, 18, Inf),
@@
     select(
-    record_id, who_collected_cos, who_collected_the_sos,
+        record_id,
+        who_collected_cos,
+        who_collected_the_sos,
         staphylococcus_aureus_cos:other_result_cos,
@@
 db_medications <- db_medications %>% filter(!record_id %in% exclude)
+# Fix chronic oral antibiotics
+db_medications$chronic_oral_antibiotic[
+    db_medications$other_chronic_oral_antibiotic == "Azithromycin"
+] <- "No"
 # Combine race and ethnicity columns
 races <- c(
-  "American Indian or Alaskan Native", "Asian", "Hawaiian or Pacific Islander",
-  "Black or African American", "White", "Unknown", "Other"
+    "American Indian or Alaskan Native",
+    "Asian",
+    "Hawaiian or Pacific Islander",
+    "Black or African American",
+    "White",
+    "Unknown",
+    "Other"
 )
@@
 })
-db_demographics$Race <- factor(db_demographics$Race,
+db_demographics$Race <- factor(
+    db_demographics$Race,
     levels = c(
-    "White", "Other", "Multiple Races", "American Indian or Alaskan Native"
+        "White",
+        "Other",
+        "Multiple Races",
+        "American Indian or Alaskan Native"
     )
@@
 })
-db_demographics$Ethnicity <- factor(db_demographics$Ethnicity,
+db_demographics$Ethnicity <- factor(
+    db_demographics$Ethnicity,
     levels = c("Non-Hispanic", "Hispanic", "Unknown/Not Reported")
@@
 # Add groups to other dataframes
-db_medications <- left_join(db_medications,
+db_medications <- left_join(
+    db_medications,
+    db_demographics %>% select(record_id, group),
+    by = join_by(record_id)
+)
+db_clinic_obtained_values <- left_join(
+    db_clinic_obtained_values,
     db_demographics %>% select(record_id, group),
@@
 bugs <- c(
-  "staphylococcus_aureus", "haemophilus_influenza", "pseudomonas_aeruginosa",
-  "pseudomonas_non_aero", "s_maltophilia", "burkholderia", "aspergillus", "ntm",
+    "staphylococcus_aureus",
+    "haemophilus_influenza",
+    "pseudomonas_aeruginosa",
+    "pseudomonas_non_aero",
+    "s_maltophilia",
+    "burkholderia",
+    "aspergillus",
+    "ntm",
     "other"
@@
 bug_names <- c(
-  "S. aureus", "H. influenza", "P. aeruginosa", "Pseudomonas non-aeruginosa",
-  "S. maltophilia", "Burkholderia", "Aspergillus", "NTM", "Other"
+    "S. aureus",
+    "H. influenza",
+    "P. aeruginosa",
+    "Pseudomonas non-aeruginosa",
+    "S. maltophilia",
+    "Burkholderia",
+    "Aspergillus",
+    "NTM",
+    "Other"
 )
 # Convert yes/no to positive/negative
-df[, paste0(bugs, "_cos")] <- lapply(df[, paste0(bugs, "_cos")], factor,
-  levels = c("No", "Yes"), labels = c("Negative", "Positive")
+df[, paste0(bugs, "_cos")] <- lapply(
+    df[, paste0(bugs, "_cos")],
+    factor,
+    levels = c("No", "Yes"),
+    labels = c("Negative", "Positive")
 )
-df[, paste0(bugs, "_sos")] <- lapply(df[, paste0(bugs, "_sos")], factor,
-  levels = c("No", "Yes"), labels = c("Negative", "Positive")
+df[, paste0(bugs, "_sos")] <- lapply(
+    df[, paste0(bugs, "_sos")],
+    factor,
+    levels = c("No", "Yes"),
+    labels = c("Negative", "Positive")
 )
@@
 })
-df$overall_cos <- factor(df$overall_cos,
-  levels = c(F, T), labels = c("Negative", "Positive")
+df$overall_cos <- factor(
+    df$overall_cos,
+    levels = c(F, T),
+    labels = c("Negative", "Positive")
 )
@@
 })
-df$overall_sos <- factor(df$overall_sos,
-  levels = c(F, T), labels = c("Negative", "Positive")
+df$overall_sos <- factor(
+    df$overall_sos,
+    levels = c(F, T),
+    labels = c("Negative", "Positive")
 )
@@
 # Add groups to dataframe
-df <- left_join(df, db_demographics %>% select(record_id, group, age_group),
+df <- left_join(
+    df,
+    db_demographics %>% select(record_id, group, age_group),
     by = join_by(record_id)
@@
 db_demographics %>%
-  select(group, age_at_cos, age_group, sex, Race, Ethnicity) %>%
+    select(group, age_at_cos, age_group, ppfev1, sex, Race, Ethnicity) %>%
     tbl_summary(
-    by = group, missing_text = "Missing",
-    statistic = all_continuous() ~ c("{median} ({min}, {max})"),
+        by = group,
+        missing_text = "Missing",
+        statistic = all_continuous() ~ c("{median} ({p25}, {p75})"),
         label = list(
-      "sex" = "Sex", "age_at_cos" = "Age at COS", "age_group" = "Age Group"
+            "sex" = "Sex",
+            "age_at_cos" = "Age at COS",
+            "age_group" = "Age Group"
         )
@@
     separate_p_footnotes() %>%
-  modify_footnote(everything() ~ NA) %>%
     as_gt()
+```
+
+## By age group
+
+```{r}
+db_demographics %>%
+    mutate(age_group = cut(age_at_cos, c(-Inf, 18, Inf), right = F)) %>%
+    select(group, age_group, ppfev1, sex, Race, Ethnicity) %>%
+    tbl_summary(
+        by = age_group,
+        missing_text = "Missing",
+        statistic = all_continuous() ~ c("{median} ({p25}, {p75})"),
+        label = list(
+            "sex" = "Sex",
+            "age_at_cos" = "Age at COS"
+        )
+    ) %>%
+    add_overall() %>%
+    add_p(
+        test = list(c("Race", "Ethnicity") ~ "fisher.test")
+    ) %>%
+    separate_p_footnotes() %>%
+    as_gt()
+
+db_demographics %>%
+    mutate(age_group = cut(age_at_cos, c(-Inf, 18, Inf), right = F)) %>%
+    select(age_group, ppfev1) %>%
+    drop_na() %>%
+    ggplot(., aes(x = ppfev1)) +
+    geom_histogram() +
+    facet_wrap(~age_group)
 ```
@@
     select(
-    -inhaled_colisitin_status, -inhaled_vancomycin_status,
-    -other_inhaled_status, -please_specify_mucolytic
+        -inhaled_colisitin_status,
+        -inhaled_vancomycin_status,
+        -other_inhaled_status,
+        -please_specify_mucolytic
     ) %>%
@@
     select(
-    previous_home_sample:if_no_why_not___4, -barrier_other,
-    -barriers___5, -if_no_why_not___4
+        previous_home_sample:if_no_why_not___4,
+        -barrier_other,
+        -barriers___5,
+        -if_no_why_not___4
     ) %>%
@@
     # Agreement metrics
-  c <- confusionMatrix(d$SOS,
+    c <- confusionMatrix(
+        d$SOS,
         reference = d$COS,
-    positive = "Positive", mode = "everything",
+        positive = "Positive",
+        mode = "everything",
         dnn = c("SOS", "COS")
     )
+    pabak <- epi.kappa(table(d$COS, d$SOS))
     # Print
@@
     cat("\n")
-  cat("Overall accuracy (both negative/negative and positive/positive) was ",
-    scales::percent(c$overall["Accuracy"], accuracy = 0.1), " (95% CI: ",
-    scales::percent(c$overall["AccuracyLower"], accuracy = 0.1), " - ",
-    scales::percent(c$overall["AccuracyUpper"], accuracy = 0.1), ").",
+    cat(
+        "Overall accuracy (both negative/negative and positive/positive) was ",
+        scales::percent(c$overall["Accuracy"], accuracy = 0.1),
+        " (95% CI: ",
+        scales::percent(c$overall["AccuracyLower"], accuracy = 0.1),
+        " - ",
+        scales::percent(c$overall["AccuracyUpper"], accuracy = 0.1),
+        ").",
         sep = ""
@@
     cat("\n")
-  cat("Kappa was ", round(c$overall["Kappa"], 3), ".", sep = "")
+    cat(
+        "Kappa was ",
+        round(pabak$kappa$est, 3),
+        " (95% CI: ",
+        round(pabak$kappa$lower, 3),
+        " - ",
+        round(pabak$kappa$upper, 3),
+        ")",
+        ".",
+        sep = ""
+    )
+    cat("\n")
+    cat("\n")
+    cat(
+        "Prevalence and Bias Adjusted Kappa (PABAK) was ",
+        round(pabak$pabak$est, 3),
+        " (95% CI: ",
+        round(pabak$pabak$lower, 3),
+        " - ",
+        round(pabak$pabak$upper, 3),
+        ")",
+        ".",
+        sep = ""
+    )
     cat("\n")
@@
     cat("\n")
-  print(plot(euler(d == "Negative"),
-    quantities = list(cex = 3),
-    fills = list(fill = c("#FFC107", "#1E88E5", "#004D40"))
-  ))
+    print(plot(euler(d == "Negative"), quantities = list(cex = 3)))
     cat("\n")
@@
     cat("\n")
-  print(plot(euler(d == "Positive"),
-    quantities = list(cex = 3),
-    fills = list(fill = c("#FFC107", "#1E88E5", "#004D40"))
-  ))
+    print(plot(euler(d == "Positive"), quantities = list(cex = 3)))
     cat("\n")
@@
     d$detailed_agree <- paste0("COS ", d$COS, "/", "SOS ", d$SOS)
-  d$detailed_agree <- factor(d$detailed_agree,
+    d$detailed_agree <- factor(
+        d$detailed_agree,
         levels = c(
-      "COS Negative/SOS Negative", "COS Negative/SOS Positive",
-      "COS Positive/SOS Negative", "COS Positive/SOS Positive"
+            "COS Negative/SOS Negative",
+            "COS Negative/SOS Positive",
+            "COS Positive/SOS Negative",
+            "COS Positive/SOS Positive"
         ),
@@
         select(agree, detailed_agree, Group) %>%
-    tbl_summary(by = Group, missing_text = "Missing", label = list(
+        tbl_summary(
+            by = Group,
+            missing_text = "Missing",
+            label = list(
                 agree ~ "Agree (+/+ and -/-)",
                 detailed_agree ~ "Detailed Agreement"
-    )) %>%
+            )
+        ) %>%
         add_overall() %>%
@@
 # All pathogens
-d <- df[, c("overall_cos", "overall_sos")]
-colnames(d) <- c("COS", "SOS")
-d <- d[complete.cases(d), ]
-a1 <- plot(euler(d == "Positive"),
-  quantities = list(cex = 3),
-  fills = list(fill = c("#FFC107", "#1E88E5", "#004D40"))
-)
-a2 <- ggplot(as.data.frame(table(d$COS, d$SOS)), aes(Var1, Var2)) +
+d_overall <- df[, c("overall_cos", "overall_sos")]
+colnames(d_overall) <- c("COS", "SOS")
+d_overall <- d_overall[complete.cases(d_overall), ]
+a1 <- plot(euler(d_overall == "Positive"), quantities = list(cex = 3))
+a2 <-
+    ggplot(
+        as.data.frame(table(d_overall$COS, d_overall$SOS)),
+        aes(Var1, Var2)
+    ) +
     geom_tile(color = "white", fill = "#1E88E5") +
-  geom_text(aes(label = paste0(Freq, "\n(", round(Freq / sum(Freq), 3) * 100, "%)")),
+    geom_text(
+        aes(
+            label = paste0(Freq, "\n(", round(Freq / sum(Freq), 3) * 100, "%)")
+        ),
         color = "white"
@@
 # MRSA
-d <- df[, c("staphylococcus_aureus_cos", "staphylococcus_aureus_sos")]
-colnames(d) <- c("COS", "SOS")
-d <- d[complete.cases(d), ]
-b1 <- plot(euler(d == "Positive"),
-  quantities = list(cex = 3),
-  fills = list(fill = c("#FFC107", "#1E88E5", "#004D40"))
-)
-b2 <- ggplot(as.data.frame(table(d$COS, d$SOS)), aes(Var1, Var2)) +
+d_MSSA <- df[, c("staphylococcus_aureus_cos", "staphylococcus_aureus_sos")]
+colnames(d_MSSA) <- c("COS", "SOS")
+d_MSSA <- d_MSSA[complete.cases(d_MSSA), ]
+b1 <- plot(euler(d_MSSA == "Positive"), quantities = list(cex = 3))
+b2 <-
+    ggplot(as.data.frame(table(d_MSSA$COS, d_MSSA$SOS)), aes(Var1, Var2)) +
     geom_tile(color = "white", fill = "#1E88E5") +
-  geom_text(aes(label = paste0(Freq, "\n(", round(Freq / sum(Freq), 3) * 100, "%)")),
+    geom_text(
+        aes(
+            label = paste0(Freq, "\n(", round(Freq / sum(Freq), 3) * 100, "%)")
+        ),
         color = "white"
@@
     theme(text = element_text(size = 15))
-p <- ggarrange(a1, a2, b1, b2, nrow = 2, ncol = 2) +
+p1 <- ggarrange(a1, a2, b1, b2, nrow = 2, ncol = 2) +
     bgcolor("White")
-ggsave("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Jordana Hoppe/SSwAB-IT/Dissemination/Figures/fig.png",
-  plot = p, width =9, height = 6, units = "in"
+ggsave(
+    "./Dissemination/Figures/fig.png",
+    plot = p1,
+    width = 9,
+    height = 6,
+    units = "in"
+)
+a1 <- plot(
+    euler(d_overall == "Positive"),
+    quantities = list(cex = 3),
+    main = "Any CF Pathogen"
+)
+b1 <- plot(
+    euler(d_MSSA == "Positive"),
+    quantities = list(cex = 3),
+    main = "MSSA"
+)
+p2 <- ggarrange(a1, b1, nrow = 1, ncol = 2) +
+    bgcolor("White")
+ggsave(
+    "./Dissemination/Figures/euler_bw.png",
+    plot = p2,
+    width = 9,
+    height = 6,
+    units = "in"
+)
+a1 <- plot(
+    venn(d_overall == "Positive"),
+    quantities = list(cex = 3),
+    main = "Any CF Pathogen"
+)
+b1 <- plot(
+    venn(d_MSSA == "Positive"),
+    quantities = list(cex = 3),
+    main = "MSSA"
+)
+p3 <- ggarrange(a1, b1, nrow = 1, ncol = 2) +
+    bgcolor("White")
+ggsave(
+    "./Dissemination/Figures/venn_bw.png",
+    plot = p3,
+    width = 9,
+    height = 6,
+    units = "in"
 )

### 2025-09-15 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
file paths and formatting

@@
 library(ggrepel)
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
-  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
+home_dir <- switch(
+    Sys.info()["sysname"],
+    "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
+    "Windows" = "C:/Users/tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
     "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/CF Registry"
 )
-github_dir <- switch(Sys.info()["sysname"],
+github_dir <- switch(
+    Sys.info()["sysname"],
     "Darwin" = "/Users/timvigers/GitHub",
@@
 # Import data
-encounter <- read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv",
+encounter <- read.csv(
+    "./Data_Raw/CFF19_encountersMerged_Del1.csv",
     na.strings = c("", "N/A")
@@
 # Demographic data
-demo <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv",
+demo <- read.csv(
+    "./Data_Raw/CFF19_DemogCFDiag_Del1.csv",
     na.strings = c("", "N/A")
@@
 races <- c(
-  "White", "Black or African American", "American Indian or Alaska Native",
-  "Asian", "Native Hawaiian or Other Pacific Islander", "Some other race"
+    "White",
+    "Black or African American",
+    "American Indian or Alaska Native",
+    "Asian",
+    "Native Hawaiian or Other Pacific Islander",
+    "Some other race"
 )
@@
 })
-demo$Race <- factor(demo$Race,
+demo$Race <- factor(
+    demo$Race,
     levels = c(
-    "White", "Black or African American", "American Indian or Alaska Native",
-    "Asian", "More than one race", "Native Hawaiian or Other Pacific Islander",
+        "White",
+        "Black or African American",
+        "American Indian or Alaska Native",
+        "Asian",
+        "More than one race",
+        "Native Hawaiian or Other Pacific Islander",
         "Some other race"
@@
     labels = c(
-    "White", "Black or African American", "Other race", "Other race",
-    "More than one race", "Other race", "Other race"
+        "White",
+        "Black or African American",
+        "Other race",
+        "Other race",
+        "More than one race",
+        "Other race",
+        "Other race"
     )
 )
-demo$Hispanicrace <- factor(demo$Hispanicrace,
+demo$Hispanicrace <- factor(
+    demo$Hispanicrace,
     levels = 1:2,
@@
 # F508
-demo$F508 <- factor(demo$F508,
+demo$F508 <- factor(
+    demo$F508,
     levels = 1:3,
@@
 # Age at year end
-annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
+annual <- read.csv(
+    "./Data_Raw/CFF19_Annualized_Del1.csv",
     na.strings = c("", "N/A")
@@
     select(
-    eDWID, reviewyear, Age_YrEnd, Race, Hispanicrace, Gender, F508, MutClass,
-    twohour_bloodglucose, hgba1c
+        eDWID,
+        reviewyear,
+        Age_YrEnd,
+        Race,
+        Hispanicrace,
+        Gender,
+        F508,
+        MutClass,
+        twohour_bloodglucose,
+        hgba1c
     ) %>%
-  filter(reviewyear >= 2010, Age_YrEnd >= 10, hgba1c >= 3.5 & hgba1c <= 15) %>%
+    filter(
+        reviewyear >= 2010,
+        Age_YrEnd >= 10,
+        hgba1c >= 3.5 & hgba1c <= 15
+    ) %>%
     drop_na() %>%
@@
     mutate(
-    `OGTT 2-hour` = cut(twohour_bloodglucose,
+        `OGTT 2-hour` = cut(
+            twohour_bloodglucose,
             breaks = c(-Inf, 140, 200, Inf),
@@
 metrics <- c(
-  "thresh", "specificity", "sensitivity", "accuracy", "tn", "tp", "fn",
-  "fp", "npv", "ppv"
+    "thresh",
+    "specificity",
+    "sensitivity",
+    "accuracy",
+    "tn",
+    "tp",
+    "fn",
+    "fp",
+    "npv",
+    "ppv"
 )
@@
     point_label <- paste0(
-    "Threshold: ", c[1], "\n",
-    "Specificity: ", c[2], "\n",
-    "Sensitivity: ", c[3]
+        "Threshold: ",
+        c[1],
+        "\n",
+        "Specificity: ",
+        c[2],
+        "\n",
+        "Sensitivity: ",
+        c[3]
     )
@@
         geom_label_repel(
-      data = point_df, aes(label = point_label), nudge_x = 0.25,
-      segment.color = "grey50", size = 5
+            data = point_df,
+            aes(label = point_label),
+            nudge_x = 0.25,
+            segment.color = "grey50",
+            size = 5
         ) +
@@
     select(
-    reviewyear, hgba1c, Age_YrEnd, `Age Group`, `OGTT 2-hour`, Race,
-    Hispanicrace, Gender, F508, MutClass
+        reviewyear,
+        hgba1c,
+        Age_YrEnd,
+        `Age Group`,
+        `OGTT 2-hour`,
+        Race,
+        Hispanicrace,
+        Gender,
+        F508,
+        MutClass
     ) %>%
     tbl_summary(
-    by = `OGTT 2-hour`, missing_text = "Missing",
+        by = `OGTT 2-hour`,
+        missing_text = "Missing",
         type = list(reviewyear ~ "categorical"),
         label = list(
-      hgba1c = "HbA1c", Hispanicrace = "Hispanic", MutClass = "Mutation Class",
-      reviewyear = "Year", Age_YrEnd = "Age at Year End"
+            hgba1c = "HbA1c",
+            Hispanicrace = "Hispanic",
+            MutClass = "Mutation Class",
+            reviewyear = "Year",
+            Age_YrEnd = "Age at Year End"
         )
@@
     select(
-    reviewyear, hgba1c, Age_YrEnd, `Age Group`, `OGTT 2-hour`, Race,
-    Hispanicrace, Gender, F508, MutClass
+        reviewyear,
+        hgba1c,
+        Age_YrEnd,
+        `Age Group`,
+        `OGTT 2-hour`,
+        Race,
+        Hispanicrace,
+        Gender,
+        F508,
+        MutClass
     ) %>%
     tbl_summary(
-    by = `Age Group`, missing_text = "Missing",
+        by = `Age Group`,
+        missing_text = "Missing",
         type = list(reviewyear ~ "categorical"),
         label = list(
-      hgba1c = "HbA1c", Hispanicrace = "Hispanic", MutClass = "Mutation Class",
-      reviewyear = "Year", Age_YrEnd = "Age at Year End"
+            hgba1c = "HbA1c",
+            Hispanicrace = "Hispanic",
+            MutClass = "Mutation Class",
+            reviewyear = "Year",
+            Age_YrEnd = "Age at Year End"
         )

### 2025-09-15 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
EnVision ROC updates for Christine

@@
 library(ggrepel)
-home_dir <- switch(
-    Sys.info()["sysname"],
+home_dir <- switch(Sys.info()["sysname"],
     "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
@@
 )
-github_dir <- switch(
-    Sys.info()["sysname"],
+github_dir <- switch(Sys.info()["sysname"],
     "Darwin" = "/Users/timvigers/GitHub",
@@
 ```{r data cleaning}
+setwd(home_dir)
 # Import data
@@
 annual <- annual %>% rename(reviewyear = ReviewYear)
+annual <- left_join(annual, demo %>% select(eDWID, Gender), by = join_by(eDWID))
+# Get BMI group_by
+annual$bmi_cat <- apply(annual, 1, function(r) {
+    age <- as.numeric(r["Age_YrEnd"])
+    gender <- as.character(r["Gender"])
+    bmi <- as.numeric(r["A_bmivalue"])
+    bmi_perc <- as.numeric(r["A_bmipercentile"])
+    if (age <= 20) {
+        return(cut(
+            bmi_perc,
+            c(-Inf, 10, 50, 85, Inf),
+            right = F,
+            labels = c("Undernourished", "At Risk", "Target/Goal", "Overweight")
+        ))
+    } else if (age > 20) {
+        if (gender == "M") {
+            return(cut(
+                bmi,
+                c(-Inf, 18.5, 23, 25, Inf),
+                right = F,
+                labels = c(
+                    "Undernourished",
+                    "At Risk",
+                    "Target/Goal",
+                    "Overweight"
+                )
+            ))
+        } else if (gender == "F") {
+            return(cut(
+                bmi,
+                c(-Inf, 18.5, 22, 25, Inf),
+                right = F,
+                labels = c(
+                    "Undernourished",
+                    "At Risk",
+                    "Target/Goal",
+                    "Overweight"
+                )
+            ))
+        }
+    }
+})
+annual$bmi_cat <- relevel(annual$bmi_cat, ref = "Target/Goal")
 # Merge
@@
     encounter,
-    annual %>% select(eDWID, reviewyear, Age_YrEnd),
+    annual %>%
+        select(
+            eDWID,
+            reviewyear,
+            Age_YrEnd,
+            A_IsOnEnzymes,
+            bmi_cat,
+            A_FEV1_pct_predicted,
+            A_FVC_pct_predicted
+        ),
     by = join_by(eDWID, reviewyear)
@@
         MutClass,
-        twohour_bloodglucose,
-        hgba1c
+        A_IsOnEnzymes,
+        Vx770,
+        Vx809comb,
+        height,
+        weight,
+        bmivalue,
+        bmipercentile,
+        bmi_cat,
+        A_FEV1_pct_predicted,
+        A_FVC_pct_predicted,
+        hgba1c,
+        fast_bloodglucose,
+        twohour_bloodglucose
     ) %>%
@@
     ) %>%
-    drop_na() %>%
     group_by(eDWID) %>%
@@
 roc_df %>%
-    select(
-        reviewyear,
-        hgba1c,
-        Age_YrEnd,
-        `Age Group`,
-        `OGTT 2-hour`,
-        Race,
-        Hispanicrace,
-        Gender,
-        F508,
-        MutClass
-    ) %>%
+    select(-eDWID) %>%
     tbl_summary(
@@
         missing_text = "Missing",
-        type = list(reviewyear ~ "categorical"),
+        type = list(
+            reviewyear ~ "categorical",
+            all_continuous() ~ "continuous2"
+        ),
+        statistic = list(
+            all_continuous() ~
+                c("{mean} ({sd})", "{median} ({p25}, {p75})", "{min}, {max}")
+        ),
         label = list(

### 2025-09-15 Christine Chan/EnVision CF/envision_analyses.qmd
EnVision ROC updates for Christine

@@
 library(emmeans)
+library(plotly)
 home_dir <- switch(Sys.info()["sysname"],
@@
 # See GitHub/CF/Christine Chan/EnVision CF/clean_envision_data.R
-load("/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF/Data_Clean/envision_analysis_dataset.RData")
+load(
+    "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF/Data_Clean/envision_analysis_dataset.RData"
+)
 ```
@@
 # Limit to those with CGM data (based on mean glucose) and OGTT diagnosis
-df <- final_df %>% filter(!is.na(mean_sensor), !is.na(Diagnosis))
+# Remove CC0083 per Christine (weird CGM)
+df <- final_df %>%
+    filter(!is.na(mean_sensor), !is.na(Diagnosis), study_id != "CC0083")
+# Change specific diagnoses per Christine
+df$Diagnosis[df$study_id == "WU0019"] <- c("NGT", "NGT")
+df$Diagnosis[df$study_id == "IA0080"] <- "IGT"
 # Convert CV units to match Scully et al.
@@
 # First CGM and OGTT pair only
-df <- df %>%
-  group_by(study_id) %>%
-  slice_min(Date)
+# df <- df %>%
+#     group_by(study_id) %>%
+#     slice_min(Date)
 ```
@@
 
-## Participant Demographics by CFRD status
+## Participant Demographics by Diagnosis
 
@@
 t1 <- tableby(
-  CFRD ~ fe(redcap_data_access_group) +
+    Dysglycemia ~
+        fe(redcap_data_access_group) +
         kwt(age_visit, "Nmiss", "medianq1q3", "range") +
         kwt(height, "Nmiss", "medianq1q3", "range") +
-    kwt(weight, "Nmiss", "medianq1q3", "range") + bmi_perc + sex +
-    fe(origin_race) + fe(ethnicity) + fe(pancreatic_status) + fe(CFTR) +
-    corrector_yes_no + fev1 + fvc + a1c_result,
+        kwt(weight, "Nmiss", "medianq1q3", "range") +
+        bmi_perc +
+        sex +
+        fe(origin_race) +
+        fe(ethnicity) +
+        fe(pancreatic_status) +
+        fe(CFTR) +
+        corrector_yes_no +
+        fev1 +
+        fvc +
+        a1c_result,
+    data = df,
+    simulate.p.value = TRUE
+)
+summary(t1, pfootnote = T)
+```
+
+## Participant Demographics by CFRD status
+
+```{r results='asis'}
+t1 <- tableby(
+    CFRD ~
+        fe(redcap_data_access_group) +
+        kwt(age_visit, "Nmiss", "medianq1q3", "range") +
+        kwt(height, "Nmiss", "medianq1q3", "range") +
+        kwt(weight, "Nmiss", "medianq1q3", "range") +
+        bmi_perc +
+        sex +
+        fe(origin_race) +
+        fe(ethnicity) +
+        fe(pancreatic_status) +
+        fe(CFTR) +
+        corrector_yes_no +
+        fev1 +
+        fvc +
+        a1c_result,
     data = df
@@
 ci.auc(r)
-# gt(data.frame(ci.coords(r, "best",
-#   best.policy = "random",
-#   ret = c(c("sensitivity", "specificity", "fp", "tp", "fn", "tn"))
-# ))) %>% fmt_number(decimals = 3)
 gt(coords(r, "best", ret = "all")) %>% fmt_number(decimals = 3)
+gt(data.frame(coords(
+    r,
+    "all",
+    ret = c(c(
+        "thresh",
+        "sensitivity",
+        "specificity",
+        "fp",
+        "tp",
+        "fn",
+        "tn",
+        "npv",
+        "ppv"
+    ))
+))) %>%
+    fmt_number(decimals = 3)
 ```
@@
     df,
-  aes(x = Dysglycemia, y = percent_time_over_140, colour = Dysglycemia)
+    aes(
+        x = Dysglycemia,
+        y = percent_time_over_140,
+        colour = Dysglycemia,
+        text = study_id
+    )
 ) +
@@
         axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.title.x = element_blank(), axis.title.y = element_blank()
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank()
     ) +
@@
 box_p
+ggplotly(box_p1, tooltip = c("text"))
+ggplotly(box_p2, tooltip = c("text"))
+ggplotly(box_p4, tooltip = c("text"))
+ggplotly(box_p3, tooltip = c("text"))
+ggplotly(box_p5, tooltip = c("text"))
 ```
@@
     select(
-    study_id, Hypo60, Hypo70, ogtt_num,
+        study_id,
+        Hypo60,
+        Hypo70,
+        ogtt_num,
         all_of(c(
-      glucoses, insulins, cpeptides, glucagons, glp1s, gips, epinephrines,
-      norepinephrines, hypo_surveys
+            glucoses,
+            insulins,
+            cpeptides,
+            glucagons,
+            glp1s,
+            gips,
+            epinephrines,
+            norepinephrines,
+            hypo_surveys
         ))
@@
         all_of(c(
-      glucoses, insulins, cpeptides, glucagons, glp1s, gips, epinephrines,
-      norepinephrines, hypo_surveys
+            glucoses,
+            insulins,
+            cpeptides,
+            glucagons,
+            glp1s,
+            gips,
+            epinephrines,
+            norepinephrines,
+            hypo_surveys
         )),
@@
     Hypo70 ~
-    fe(Diagnosis) + redcap_data_access_group +
+        fe(Diagnosis) +
+        redcap_data_access_group +
         kwt(age_visit, "Nmiss", "medianq1q3", "range") +
         kwt(height, "Nmiss", "medianq1q3", "range") +
-    kwt(weight, "Nmiss", "medianq1q3", "range") + bmi_perc + sex +
-    fe(origin_race) + fe(ethnicity) + fe(pancreatic_status) + fe(CFTR) +
-    corrector_yes_no + fev1 + fvc + a1c_result,
-  data = first_visit_df, control = tableby.control(simulate.p.value = T)
+        kwt(weight, "Nmiss", "medianq1q3", "range") +
+        bmi_perc +
+        sex +
+        fe(origin_race) +
+        fe(ethnicity) +
+        fe(pancreatic_status) +
+        fe(CFTR) +
+        corrector_yes_no +
+        fev1 +
+        fvc +
+        a1c_result,
+    data = first_visit_df,
+    control = tableby.control(simulate.p.value = T)
 )
@@
     Hypo60 ~
-    fe(Diagnosis) + redcap_data_access_group +
+        fe(Diagnosis) +
+        redcap_data_access_group +
         kwt(age_visit, "Nmiss", "medianq1q3", "range") +
         kwt(height, "Nmiss", "medianq1q3", "range") +
-    kwt(weight, "Nmiss", "medianq1q3", "range") + bmi_perc + sex +
-    fe(origin_race) + fe(ethnicity) + fe(pancreatic_status) + fe(CFTR) +
-    corrector_yes_no + fev1 + fvc + a1c_result,
-  data = first_visit_df, control = tableby.control(simulate.p.value = T)
+        kwt(weight, "Nmiss", "medianq1q3", "range") +
+        bmi_perc +
+        sex +
+        fe(origin_race) +
+        fe(ethnicity) +
+        fe(pancreatic_status) +
+        fe(CFTR) +
+        corrector_yes_no +
+        fev1 +
+        fvc +
+        a1c_result,
+    data = first_visit_df,
+    control = tableby.control(simulate.p.value = T)
 )
@@
 # Fit
-mod <- lmer(Glucose ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+mod <- lmer(
+    Glucose ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
     data = glmm_df
@@
 # EM means
-means <- data.frame(emmeans(mod, c("Timepoint", "Hypo70"),
+means <- data.frame(emmeans(
+    mod,
+    c("Timepoint", "Hypo70"),
     type = "response",
@@
 colnames(means) <- c(
-  "Timepoint", "Hypo70", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+    "Timepoint",
+    "Hypo70",
+    "Mean",
+    "SE",
+    "df",
+    "CI Lower Limit",
+    "CI Upper Limit"
 )
 # Plot
-ggplot(means, aes(
-  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo70,
+ggplot(
+    means,
+    aes(
+        x = as.numeric(as.character(Timepoint)),
+        y = Mean,
+        colour = Hypo70,
         group = Hypo70
-)) +
+    )
+) +
     geom_point() +
-  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    geom_errorbar(
+        aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
         width = 3
@@
 # Fit
-mod <- lmer(Glucose ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+mod <- lmer(
+    Glucose ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
     data = glmm_df
@@
 # EM means
-means <- data.frame(emmeans(mod, c("Timepoint", "Hypo60"),
+means <- data.frame(emmeans(
+    mod,
+    c("Timepoint", "Hypo60"),
     type = "response",
@@
 colnames(means) <- c(
-  "Timepoint", "Hypo60", "Mean", "SE", "df", "CI Lower Limit", "CI Upper Limit"
+    "Timepoint",
+    "Hypo60",
+    "Mean",
+    "SE",
+    "df",
+    "CI Lower Limit",
+    "CI Upper Limit"
 )
 # Plot
-ggplot(means, aes(
-  x = as.numeric(as.character(Timepoint)), y = Mean, colour = Hypo60,
+ggplot(
+    means,
+    aes(
+        x = as.numeric(as.character(Timepoint)),
+        y = Mean,
+        colour = Hypo60,
         group = Hypo60
-)) +
+    )
+) +
     geom_point() +
-  geom_errorbar(aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
+    geom_errorbar(
+        aes(ymin = `CI Lower Limit`, ymax = `CI Upper Limit`),
         width = 3
@@
 # Fit
-mod <- lmer(log(Insulin) ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+mod <- lmer(
+    log(Insulin) ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
     data = glmm_df
@@
 # Fit
-mod <- lmer(log(Insulin) ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+mod <- lmer(
+    log(Insulin) ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
     data = glmm_df
@@
 # Fit
-mod <- lmer(log(C.Peptide) ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+mod <- lmer(
+    log(C.Peptide) ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
     data = glmm_df
@@
 # Fit
-mod <- lmer(log(C.Peptide) ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+mod <- lmer(
+    log(C.Peptide) ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
     data = glmm_df
@@
 # Fit
-mod <- lmer(Glucagon ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+mod <- lmer(
+    Glucagon ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
     data = glmm_df
@@
 # Fit
-mod <- lmer(Glucagon ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+mod <- lmer(
+    Glucagon ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
     data = glmm_df
@@
 # Fit
-mod <- lmer(GIP ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+mod <- lmer(
+    GIP ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
     data = glmm_df
@@
 # Fit
-mod <- lmer(GIP ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+mod <- lmer(
+    GIP ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
     data = glmm_df
@@
 # Fit
-mod <- lmer(log(Epinephrine) ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+mod <- lmer(
+    log(Epinephrine) ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
     data = glmm_df
@@
 # Fit
-mod <- lmer(log(Epinephrine) ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+mod <- lmer(
+    log(Epinephrine) ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
     data = glmm_df
@@
 # Fit
-mod <- lmer(log(Norepinephrine) ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+mod <- lmer(
+    log(Norepinephrine) ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
     data = glmm_df
@@
 # Fit
-mod <- lmer(log(Norepinephrine) ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+mod <- lmer(
+    log(Norepinephrine) ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
     data = glmm_df
@@
 # Poisson model better than NB by AIC and BIC
-mod <- glmer(adren_score ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
-  data = glmm_df, family = "poisson",
+mod <- glmer(
+    adren_score ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+    data = glmm_df,
+    family = "poisson",
     control = glmerControl(optimizer = "bobyqa")
@@
 # Poisson model better than NB by AIC and BIC
-mod <- glmer(adren_score ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
-  data = glmm_df, family = "poisson",
+mod <- glmer(
+    adren_score ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+    data = glmm_df,
+    family = "poisson",
     control = glmerControl(optimizer = "bobyqa")
@@
 # Poisson model better than NB by AIC and BIC
-mod <- glmer(neuro_score ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
-  data = glmm_df, family = "poisson",
+mod <- glmer(
+    neuro_score ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+    data = glmm_df,
+    family = "poisson",
     control = glmerControl(optimizer = "bobyqa")
@@
 # Poisson model better than NB by AIC and BIC
-mod <- glmer(neuro_score ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
-  data = glmm_df, family = "poisson",
+mod <- glmer(
+    neuro_score ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+    data = glmm_df,
+    family = "poisson",
     control = glmerControl(optimizer = "bobyqa")
@@
 # Poisson model better than NB by AIC and BIC
-mod <- glmer(total_score ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
-  data = glmm_df, family = "poisson",
+mod <- glmer(
+    total_score ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+    data = glmm_df,
+    family = "poisson",
     control = glmerControl(optimizer = "bobyqa")
@@
 # Poisson model better than NB by AIC and BIC
-mod <- glmer(total_score ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
-  data = glmm_df, family = "poisson",
+mod <- glmer(
+    total_score ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+    data = glmm_df,
+    family = "poisson",
     control = glmerControl(optimizer = "bobyqa")
@@
 # Poisson model better than NB by AIC and BIC
-mod <- glmer(num_symptoms ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
-  data = glmm_df, family = "poisson",
+mod <- glmer(
+    num_symptoms ~ Timepoint * Hypo70 + (1 | study_id / ogtt_num),
+    data = glmm_df,
+    family = "poisson",
     control = glmerControl(optimizer = "bobyqa")
@@
 # Poisson model better than NB by AIC and BIC
-mod <- glmer(num_symptoms ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
-  data = glmm_df, family = "poisson",
+mod <- glmer(
+    num_symptoms ~ Timepoint * Hypo60 + (1 | study_id / ogtt_num),
+    data = glmm_df,
+    family = "poisson",
     control = glmerControl(optimizer = "bobyqa")

### 2025-09-15 Christine Chan/EnVision CF/clean_envision_data.R

@@ -11,7 +11,8 @@ library(childsds)
 library(cgmanalysis)
 library(pracma)
 # Home directory
-home_dir <- switch(Sys.info()["sysname"],
+home_dir <- switch(
+  Sys.info()["sysname"],
   "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
   "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
   "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
@@ -31,7 +32,9 @@ insulin <- do.call(rbind, dfs)
 insulin <- insulin %>%
   select(LastName, TestName, Result.Numeric, Collection.Date) %>%
   rename(
-    study_id = LastName, Timepoint = TestName, Insulin = Result.Numeric,
+    study_id = LastName,
+    Timepoint = TestName,
+    Insulin = Result.Numeric,
     Date = Collection.Date
   )
 # Add re-runs
@@ -41,7 +44,9 @@ new_insulin <- read.csv(
 new_insulin <- new_insulin %>%
   select(Patient.Name, Collection.Date, Test, Reported.As) %>%
   rename(
-    study_id = Patient.Name, Date = Collection.Date, Timepoint = Test,
+    study_id = Patient.Name,
+    Date = Collection.Date,
+    Timepoint = Test,
     Insulin = Reported.As
   )
 insulin <- rbind(insulin, new_insulin)
@@ -54,14 +59,19 @@ insulin$Timepoint <- as.numeric(insulin$Timepoint)
 #-------------------------------------------------------------------------------
 # Catecholamines (Excel files needed some manual cleaning prior to this)
 #-------------------------------------------------------------------------------
-files <- list.files("./Christine Chan/EnVision CF/Data_Clean/Catecholamines",
+files <- list.files(
+  "./Christine Chan/EnVision CF/Data_Clean/Catecholamines",
   full.names = T
 )
 dfs <- lapply(files, read.csv)
 catecholamines <- do.call(rbind, dfs)
 catecholamines$Date <- mdy(catecholamines$Date)
 colnames(catecholamines) <- c(
[…]

### 2025-09-16 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
updates for Monique

@@
 library(ggrepel)
-home_dir <- switch(Sys.info()["sysname"],
+home_dir <- switch(
+    Sys.info()["sysname"],
     "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
@@
 )
-github_dir <- switch(Sys.info()["sysname"],
+github_dir <- switch(
+    Sys.info()["sysname"],
     "Darwin" = "/Users/timvigers/GitHub",
@@
     by = join_by(eDWID, reviewyear)
+)
+# Format CFTR modulator variables
+encounter$Vx770[is.na(encounter$Vx770)] <- 0
+encounter$Vx770 <- factor(
+    encounter$Vx770,
+    levels = 0:1,
+    labels = c("No", "Yes")
+)
+encounter$Vx809comb[is.na(encounter$Vx809comb)] <- 0
+encounter$Vx809comb <- factor(
+    encounter$Vx809comb,
+    levels = 0:1,
+    labels = c("No", "Yes")
 )
@@
     ) %>%
+    drop_na(
+        eDWID,
+        reviewyear,
+        Age_YrEnd,
+        Race,
+        Hispanicrace,
+        Gender,
+        F508,
+        MutClass,
+        twohour_bloodglucose,
+        hgba1c
+    ) %>%
     group_by(eDWID) %>%
@@
         `Age Group` = cut(Age_YrEnd, breaks = c(10, 18, 30, Inf), right = F)
-    )
-# Metrics to report
-metrics <- c(
-    "thresh",
-    "specificity",
-    "sensitivity",
-    "accuracy",
-    "tn",
-    "tp",
-    "fn",
-    "fp",
-    "npv",
-    "ppv"
     )
@@
         ),
+        statistic = list(
+            all_continuous() ~
+                c("{mean} ({sd})", "{median} ({p25}, {p75})", "{min}, {max}")
+        ),
+        label = list(
+            hgba1c = "HbA1c",
+            Hispanicrace = "Hispanic",
+            MutClass = "Mutation Class",
+            reviewyear = "Year",
+            Age_YrEnd = "Age at Year End"
+        )
+    ) %>%
+    add_overall() %>%
+    add_p(
+        test = list(reviewyear = "fisher.test"),
+        test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
+    ) %>%
+    separate_p_footnotes() %>%
+    as_gt()
+```
+
+### Pairwise comparisons
+
+#### <140 vs. [140,200)
+
+```{r}
+roc_df %>%
+    select(`OGTT 2-hour`, Gender, MutClass, hgba1c) %>%
+    filter(`OGTT 2-hour` != ">=200") %>%
+    mutate(`OGTT 2-hour` = droplevels(`OGTT 2-hour`)) %>%
+    tbl_summary(
+        by = `OGTT 2-hour`,
+        missing_text = "Missing",
+        type = list(all_continuous() ~ "continuous2"),
+        statistic = list(
+            all_continuous() ~
+                c("{mean} ({sd})", "{median} ({p25}, {p75})", "{min}, {max}")
+        ),
+        label = list(
+            hgba1c = "HbA1c",
+            Hispanicrace = "Hispanic",
+            MutClass = "Mutation Class",
+            reviewyear = "Year",
+            Age_YrEnd = "Age at Year End"
+        )
+    ) %>%
+    add_overall() %>%
+    add_p(
+        test = list(reviewyear = "fisher.test"),
+        test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
+    ) %>%
+    separate_p_footnotes() %>%
+    as_gt()
+```
+
+#### [140,200) vs. >=200
+
+```{r}
+roc_df %>%
+    select(`OGTT 2-hour`, Gender, MutClass, hgba1c) %>%
+    filter(`OGTT 2-hour` != "<140") %>%
+    mutate(`OGTT 2-hour` = droplevels(`OGTT 2-hour`)) %>%
+    tbl_summary(
+        by = `OGTT 2-hour`,
+        missing_text = "Missing",
+        type = list(all_continuous() ~ "continuous2"),
+        statistic = list(
+            all_continuous() ~
+                c("{mean} ({sd})", "{median} ({p25}, {p75})", "{min}, {max}")
+        ),
+        label = list(
+            hgba1c = "HbA1c",
+            Hispanicrace = "Hispanic",
+            MutClass = "Mutation Class",
+            reviewyear = "Year",
+            Age_YrEnd = "Age at Year End"
+        )
+    ) %>%
+    add_overall() %>%
+    add_p(
+        test = list(reviewyear = "fisher.test"),
+        test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
+    ) %>%
+    separate_p_footnotes() %>%
+    as_gt()
+```
+
+#### <140 vs. >=200
+
+```{r}
+roc_df %>%
+    select(`OGTT 2-hour`, Gender, MutClass, hgba1c) %>%
+    filter(`OGTT 2-hour` != "[140,200)") %>%
+    mutate(`OGTT 2-hour` = droplevels(`OGTT 2-hour`)) %>%
+    tbl_summary(
+        by = `OGTT 2-hour`,
+        missing_text = "Missing",
+        type = list(all_continuous() ~ "continuous2"),
         statistic = list(
@@
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(
+    r,
+    "all",
+    ret = c(
+        "threshold",
+        "sensitivity",
+        "specificity",
+        "tpr",
+        "tnr",
+        "fpr",
+        "fnr",
+        "npv",
+        "ppv"
+    )
+))) %>%
+    fmt_number(n_sigfig = 3)
 ```
@@
     fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(
+    r,
+    "all",
+    ret = c(
+        "threshold",
+        "sensitivity",
+        "specificity",
+        "tpr",
+        "tnr",
+        "fpr",
+        "fnr",
+        "npv",
+        "ppv"
+    )
+))) %>%
+    fmt_number(n_sigfig = 3)
 ```
@@
 gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
+gt(data.frame(coords(
+    r,
+    "all",
+    ret = c(
+        "threshold",
+        "sensitivity",
+        "specificity",
+        "tpr",
+        "tnr",
+        "fpr",
+        "fnr",
+        "npv",
+        "ppv"
+    )
+))) %>%
+    fmt_number(n_sigfig = 3)
+```
+
+# FEV1 over time plot
+
+```{r}
+plot_df <- annual %>%
+    filter(
+        !is.na(cfrd_status_annual),
+        Age_YrEnd >= 10,
+        !cfrd_status_annual %in% c(4, 5)
+    ) %>%
+    group_by(eDWID) %>%
+    mutate(
+        Diagnosis = max(cfrd_status_annual),
+        `Years to CFRD or Last Visit` = case_when(
+            Diagnosis == 3 ~
+                Age_YrEnd -
+                    first(Age_YrEnd[cfrd_status_annual == 3]),
+            .default = Age_YrEnd - max(Age_YrEnd, na.rm = T)
+        )
+    )
+plot_df$Diagnosis <- factor(
+    plot_df$Diagnosis,
+    levels = 1:3,
+    labels = c("NGT", "AGT", "CFRD")
+)
+ggplot(
+    plot_df,
+    aes(
+        y = A_FEV1_pct_predicted,
+        x = `Years to CFRD or Last Visit`,
+        color = Diagnosis,
+        group = Diagnosis
+    )
+) +
+    xlim(c(-10, 10)) +
+    geom_point(alpha = 0.1, shape = ".") +
+    geom_smooth(formula = y ~ s(x, bs = "cr")) +
+    theme_bw()
+```

### 2025-09-16 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
updates for Monique

--- 2025-06-13 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
@@
 library(patchwork)
-home_dir <- switch(Sys.info()["sysname"],
+home_dir <- switch(
+    Sys.info()["sysname"],
     "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
@@
 )
-github_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Documents/GitHub",
-  "Linux" = "/home/timvigers/Documents/GitHub"
+github_dir <- switch(
+    Sys.info()["sysname"],
+    "Darwin" = "/Users/tim/Documents/GitHub",
+    "Linux" = "/home/tim/Documents/GitHub"
 )
@@
 # Import data and codebook
-annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
+annual <- read.csv(
+    "./Data_Raw/CFF19_Annualized_Del1.csv",
     na.strings = c("", "N/A")
@@
 annual <- annual[!annual$cfrd_status_annual %in% c(4, 5), ]
-annual$cfrd_status_annual <- factor(annual$cfrd_status_annual,
+annual$cfrd_status_annual <- factor(
+    annual$cfrd_status_annual,
     levels = 1:3,
@@
 races <- c(
-  "White", "Black or African American", "American Indian or Alaska Native",
-  "Asian", "Native Hawaiian or Other Pacific ", "Some other race"
+    "White",
+    "Black or African American",
+    "American Indian or Alaska Native",
+    "Asian",
+    "Native Hawaiian or Other Pacific ",
+    "Some other race"
 )
@@
 })
-demogs$Race <- factor(demogs$Race,
+demogs$Race <- factor(
+    demogs$Race,
     levels = c(
-    "White", "American Indian or Alaska Native", "Asian",
-    "Black or African American", "More than one race",
-    "Native Hawaiian or Other Pacific ", "Some other race"
+        "White",
+        "American Indian or Alaska Native",
+        "Asian",
+        "Black or African American",
+        "More than one race",
+        "Native Hawaiian or Other Pacific ",
+        "Some other race"
     ),
     labels = c(
-    "White", "Other race", "Other race",
-    "Black or African American", "More than one race",
-    "Other race", "Other race"
+        "White",
+        "Other race",
+        "Other race",
+        "Black or African American",
+        "More than one race",
+        "Other race",
+        "Other race"
     )
 )
-demogs$Hispanicrace <- factor(demogs$Hispanicrace,
+demogs$Hispanicrace <- factor(
+    demogs$Hispanicrace,
     levels = 1:2,
@@
 annual$bmi <- coalesce(annual$A_bmivalue, annual$bmi)
-annual$bmi_perc_cdc <- sds(annual$bmi,
+annual$bmi_perc_cdc <- sds(
+    annual$bmi,
     age = annual$Age_YrEnd,
-  sex = annual$Gender, male = "M", female = "F",
-  item = "bmi", ref = cdc.ref, type = "perc"
-) * 100
-annual$bmi_perc_who <- sds(annual$bmi,
+    sex = annual$Gender,
+    male = "M",
+    female = "F",
+    item = "bmi",
+    ref = cdc.ref,
+    type = "perc"
+) *
+    100
+annual$bmi_perc_who <- sds(
+    annual$bmi,
     age = annual$Age_YrEnd,
-  sex = annual$Gender, male = "M", female = "F",
-  item = "bmi", ref = who.ref, type = "perc"
-) * 100
+    sex = annual$Gender,
+    male = "M",
+    female = "F",
+    item = "bmi",
+    ref = who.ref,
+    type = "perc"
+) *
+    100
 annual$bmi_perc <- coalesce(annual$bmi_perc_cdc, annual$bmi_perc_who)
@@
     if (age <= 20) {
-    return(cut(bmi_perc, c(-Inf, 10, 50, 85, Inf),
+        return(cut(
+            bmi_perc,
+            c(-Inf, 10, 50, 85, Inf),
             right = F,
@@
         if (gender == "M") {
-      return(cut(bmi, c(-Inf, 18.5, 23, 25, Inf),
+            return(cut(
+                bmi,
+                c(-Inf, 18.5, 23, 25, Inf),
                 right = F,
-        labels = c("Undernourished", "At Risk", "Target/Goal", "Overweight")
+                labels = c(
+                    "Undernourished",
+                    "At Risk",
+                    "Target/Goal",
+                    "Overweight"
+                )
             ))
         } else if (gender == "F") {
-      return(cut(bmi, c(-Inf, 18.5, 22, 25, Inf),
+            return(cut(
+                bmi,
+                c(-Inf, 18.5, 22, 25, Inf),
                 right = F,
-        labels = c("Undernourished", "At Risk", "Target/Goal", "Overweight")
+                labels = c(
+                    "Undernourished",
+                    "At Risk",
+                    "Target/Goal",
+                    "Overweight"
+                )
             ))
@@
 # Format columns
-annual$MutClass <- factor(annual$MutClass,
+annual$MutClass <- factor(
+    annual$MutClass,
     levels = c("1-3", "4-5", "Oth", "N/A"),
@@
 )
-annual$NumPulmExacerbation <- cut(annual$NumPulmExacerbation,
+annual$NumPulmExacerbation <- cut(
+    annual$NumPulmExacerbation,
     labels = c("0", "1", "2", ">=3"),
-  breaks = c(0, 1, 2, 3, Inf), right = F
+    breaks = c(0, 1, 2, 3, Inf),
+    right = F
 )
 # Age groups 10 - 18 inclusive, 19 - 30 inclusive, >= 31
-annual$age_group <- cut(annual$Age_YrEnd,
+annual$age_group <- cut(
+    annual$Age_YrEnd,
     breaks = c(-Inf, 10, 19, 31, Inf),
@@
 bugs <- c(
-  "A_pseudomonasaeruginosa", "A_fungalyeast1",
-  "A_pulmonarycomplications1", "A_mycocultureresults3"
+    "A_pseudomonasaeruginosa",
+    "A_fungalyeast1",
+    "A_pulmonarycomplications1",
+    "A_mycocultureresults3",
+    "A_cultureresults3"
 )
@@
 yn_vars <- c(
-  "A_corticosteroids1", "A_feeding2",
-  "A_Vx770", "A_VX809comb", bugs, "tube_feeding", "liver_disease"
+    "A_corticosteroids1",
+    "A_feeding2",
+    "A_Vx770",
+    "A_VX809comb",
+    bugs,
+    "tube_feeding",
+    "liver_disease"
 )
-annual[, yn_vars] <- lapply(annual[, yn_vars], factor,
+annual[, yn_vars] <- lapply(
+    annual[, yn_vars],
+    factor,
     levels = 0:1,
@@
 )
-annual$A_IsOnEnzymes <- factor(annual$A_IsOnEnzymes,
-  levels = c(1, 0), labels = c(" Yes", " No")
+annual$A_IsOnEnzymes <- factor(
+    annual$A_IsOnEnzymes,
+    levels = c(1, 0),
+    labels = c(" Yes", " No")
 )
 # FEV1 to categorical
-annual$ppFEV1_cat <- cut(annual$A_FEV1_pct_predicted,
-  breaks = c(0, 40, 80, 100, Inf), right = F
+annual$ppFEV1_cat <- cut(
+    annual$A_FEV1_pct_predicted,
+    breaks = c(0, 40, 80, 100, Inf),
+    right = F
 )
-annual$ppFEV1_cat <- factor(annual$ppFEV1_cat,
+annual$ppFEV1_cat <- factor(
+    annual$ppFEV1_cat,
     levels = c("[100,Inf)", "[80,100)", "[40,80)", "[0,40)")
@@
 vars <- c(
-  "age_group", "Gender", "bmi_perc", "bmi_cat", "MutClass", "Race",
-  "Hispanicrace", "A_IsOnEnzymes", "ppFEV1_cat", "tube_feeding",
-  "NumPulmExacerbation", "A_pseudomonasaeruginosa", "A_fungalyeast1",
-  "A_mycocultureresults3", "A_corticosteroids1", "cfrd_status_annual",
-  "A_Vx770", "A_VX809comb", "liver_disease"
+    "age_group",
+    "Gender",
+    "bmi_perc",
+    "bmi_cat",
+    "MutClass",
+    "Race",
+    "Hispanicrace",
+    "A_IsOnEnzymes",
+    "ppFEV1_cat",
+    "tube_feeding",
+    "NumPulmExacerbation",
+    "A_pseudomonasaeruginosa",
+    "A_fungalyeast1",
+    "A_mycocultureresults3",
+    "A_corticosteroids1",
+    "cfrd_status_annual",
+    "A_Vx770",
+    "A_VX809comb",
+    "liver_disease"
 )
@@
 labels <- list(
-  cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
-  Race = "Race", Hispanicrace = "Hispanic",
-  Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
+    cfrd_status_annual = "CFRD Status",
+    MutClass = "Mutation class",
+    Race = "Race",
+    Hispanicrace = "Hispanic",
+    Gender = "Sex",
+    A_IsOnEnzymes = "Pancreatic enzyme use",
     ppFEV1_cat = "FEV1 percent predicted category",
@@
     A_FVC_pct_predicted = "FVC percent predicted",
-  tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
+    tube_feeding = "Enteric feedings",
+    liver_disease = "Liver disease",
     A_corticosteroids1 = "Use of corticosteroids",
-  NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
-  height_perc = "Height percentile", bmi_cat = "BMI category",
+    NumPulmExacerbation = "Admissions for PEx",
+    age_group = "Age group",
+    height_perc = "Height percentile",
+    bmi_cat = "BMI category",
     A_Vx770 = "Use of CFTR modulator Vx770",
@@
     A_fungalyeast1 = "Aspergillus (any species)",
-  A_mycocultureresults3 = "Microorganisms in myco culture", eDWID = "ID",
-  tstart = "Interval start", tstop = "Interval stop",
-  Age_YrEnd = "Age", bmi_perc = "BMI %ile"
+    A_mycocultureresults3 = "Microorganisms in myco culture",
+    eDWID = "ID",
+    tstart = "Interval start",
+    tstop = "Interval stop",
+    Age_YrEnd = "Age",
+    bmi_perc = "BMI %ile"
 )
@@
 # This is super slow but idc idc
-t <- apply(first_cfrd, 1, function(r) {
+cfrd_only <- apply(first_cfrd, 1, function(r) {
     id <- as.numeric(r["eDWID"])
@@
     d <- annual[annual$eDWID == id & annual$ReviewYear %in% c(year, year - 1), ]
-  if (any(!is.na(d$A_hgba1c))) {
-    a1c <- T
-  } else {
-    a1c <- F
-  }
-  if (any(!is.na(d["A_twohour_bloodglucose"]) |
-    !is.na(d["A_ogttfast_bloodglucose"]))) {
-    ogtt <- T
-  } else {
-    ogtt <- F
-  }
-  if (a1c & ogtt) {
-    return("both")
-  } else if (a1c & !ogtt) {
-    return("a1c")
-  } else if (ogtt & !a1c) {
-    return("ogtt")
-  } else if (!ogtt & !a1c) {
-    return("neither")
-  }
+    a1c <- as.numeric(last(d["A_hgba1c"], na_rm = TRUE))
+    ogtt0hr <- as.numeric(last(d["A_ogttfast_bloodglucose"], na_rm = TRUE))
+    ogtt2hr <- as.numeric(last(d["A_twohour_bloodglucose"], na_rm = TRUE))
+    return(c(id, year, a1c, ogtt0hr, ogtt2hr))
 })
+cfrd_only <- data.frame(t(cfrd_only))
+colnames(cfrd_only) <- c("id", "year", "a1c", "ogtt0hr", "ogtt2hr")
+cfrd_only <- left_join(
+    cfrd_only,
+    annual %>% mutate(eDWID = as.numeric(eDWID)),
+    by = join_by(id == eDWID, year == ReviewYear)
+)
+
+cfrd_only$a1c_ogtt <- apply(cfrd_only, 1, function(r) {
+    not_missing <- paste(
+        which(!is.na(r[c("a1c", "ogtt0hr", "ogtt2hr")])),
+        collapse = "_"
+    )
+    return(not_missing)
+})
+cfrd_only$a1c_ogtt <- factor(
+    cfrd_only$a1c_ogtt,
+    levels = c("", "1", "1_2", "1_2_3", "1_3", "2", "2_3", "3"),
+    labels = c(
+        "Neither",
+        "HbA1c Only",
+        "OGTT",
+        "OGTT",
+        "OGTT",
+        "OGTT",
+        "OGTT",
+        "OGTT"
+    )
+)
+cfrd_only %>%
+    select(
+        a1c_ogtt,
+        Age_YrEnd,
+        A_weight,
+        A_height,
+        A_bmivalue,
+        A_bmipercentile,
+        A_FEV1_pct_predicted,
+        A_FVC_pct_predicted,
+        Race,
+        Hispanicrace,
+        MutClass,
+        a1c,
+        A_ogttfast_bloodglucose,
+        A_twohour_bloodglucose,
+        A_IsOnEnzymes,
+        liver_disease,
+        A_cultureresults3
+    ) %>%
+    tbl_summary(
+        by = a1c_ogtt,
+        missing_text = "Missing",
+        statistic = list(all_continuous() ~ "{mean} ({sd})")
+    ) %>%
+    add_overall() %>%
+    add_p(test = list(all_continuous() ~ "oneway.test")) %>%
+    modify_caption("**All Groups**")
+cfrd_only %>%
+    filter(a1c_ogtt %in% c("OGTT", "HbA1c Only")) %>%
+    mutate(a1c_ogtt = droplevels(a1c_ogtt)) %>%
+    select(
+        a1c_ogtt,
+        Age_YrEnd,
+        A_weight,
+        A_height,
+        A_bmivalue,
+        A_bmipercentile,
+        A_FEV1_pct_predicted,
+        A_FVC_pct_predicted,
+        Race,
+        Hispanicrace,
+        MutClass,
+        a1c,
+        A_ogttfast_bloodglucose,
+        A_twohour_bloodglucose,
+        A_IsOnEnzymes,
+        liver_disease,
+        A_cultureresults3
+    ) %>%
+    tbl_summary(
+        by = a1c_ogtt,
+        missing_text = "Missing",
+        statistic = list(all_continuous() ~ "{mean} ({sd})")
+    ) %>%
+    add_overall() %>%
+    add_p(test = list(all_continuous() ~ "oneway.test")) %>%
+    modify_caption("**OGTT vs. HbA1c Only**")
 ```
@@
 ```{r}
-data.frame(data_dict$code_meaning[data_dict$var %in%
-  sub("A_", "", liver_fields)]) %>%
+data.frame(data_dict$code_meaning[
+    data_dict$var %in%
+        sub("A_", "", liver_fields)
+]) %>%
     gt() %>%
@@
 
-  new_cases_10_18 <- sum(inc$n[inc$CFRD == "CFRD+" &
-    inc$age_group == "[10,19)"])
+    new_cases_10_18 <- sum(inc$n[
+        inc$CFRD == "CFRD+" &
+            inc$age_group == "[10,19)"
+    ])
     at_risk_10_18 <- sum(inc$n[inc$age_group == "[10,19)"])
@@
 
-  new_cases_19_30 <- sum(inc$n[inc$CFRD == "CFRD+" &
-    inc$age_group == "[19,31)"])
+    new_cases_19_30 <- sum(inc$n[
+        inc$CFRD == "CFRD+" &
+            inc$age_group == "[19,31)"
+    ])
     at_risk_19_30 <- sum(inc$n[inc$age_group == "[19,31)"])
@@
 
-  new_cases_30 <- sum(inc$n[inc$CFRD == "CFRD+" &
-    inc$age_group == "[31, Inf)"])
+    new_cases_30 <- sum(inc$n[
+        inc$CFRD == "CFRD+" &
+            inc$age_group == "[31, Inf)"
+    ])
     at_risk_30 <- sum(inc$n[inc$age_group == "[31, Inf)"])
@@
     select(
-    ReviewYear, CFRD, Age_YrEnd, Gender, Race, Hispanicrace, bmi_perc,
-    bmi_cat, MutClass, tube_feeding, liver_disease,
-    A_FEV1_pct_predicted, A_FVC_pct_predicted, NumPulmExacerbation,
-    A_corticosteroids1, A_IsOnEnzymes, A_Vx770, A_VX809comb
+        ReviewYear,
+        CFRD,
+        Age_YrEnd,
+        Gender,
+        Race,
+        Hispanicrace,
+        bmi_perc,
+        bmi_cat,
+        MutClass,
+        tube_feeding,
+        liver_disease,
+        A_FEV1_pct_predicted,
+        A_FVC_pct_predicted,
+        NumPulmExacerbation,
+        A_corticosteroids1,
+        A_IsOnEnzymes,
+        A_Vx770,
+        A_VX809comb
     ) %>%
@@
         strata = ReviewYear,
-    .tbl_fun =
-      ~ .x %>%
+        .tbl_fun = ~ .x %>%
             tbl_summary(
-          by = CFRD, missing_text = "Missing",
+                by = CFRD,
+                missing_text = "Missing",
                 type = list(A_corticosteroids1 ~ "categorical"),
-          statistic = list(c(
-            "Age_YrEnd", "bmi_perc", "A_FEV1_pct_predicted",
+                statistic = list(
+                    c(
+                        "Age_YrEnd",
+                        "bmi_perc",
+                        "A_FEV1_pct_predicted",
                         "A_FVC_pct_predicted"
-          ) ~ "{mean} ({sd})")
+                    ) ~
+                        "{mean} ({sd})"
+                )
             ) %>%
-        add_p(test = list(c(
-          "Age_YrEnd", "bmi_perc", "A_FEV1_pct_predicted",
+            add_p(
+                test = list(
+                    c(
+                        "Age_YrEnd",
+                        "bmi_perc",
+                        "A_FEV1_pct_predicted",
                         "A_FVC_pct_predicted"
-        ) ~ "t.test")) %>%
+                    ) ~
+                        "t.test"
+                )
+            ) %>%
             add_overall() %>%
@@
         `Screened HbA1c` = mean(!is.na(A_hgba1c)),
-    `Screened OGTT` = mean(!is.na(A_twohour_bloodglucose) |
-      !is.na(A_ogttfast_bloodglucose)),
-    `Screened Either` = mean(!is.na(A_twohour_bloodglucose) |
-      !is.na(A_ogttfast_bloodglucose) | !is.na(A_hgba1c)),
+        `Screened OGTT` = mean(
+            !is.na(A_twohour_bloodglucose) |
+                !is.na(A_ogttfast_bloodglucose)
+        ),
+        `Screened Either` = mean(
+            !is.na(A_twohour_bloodglucose) |
+                !is.na(A_ogttfast_bloodglucose) |
+                !is.na(A_hgba1c)
+        ),
         .groups = "drop"
@@
     pivot_longer(`Screened HbA1c`:`Screened Either`)
-screening_plot <- ggplot(screening, aes(
-  x = ReviewYear, y = value,
+screening_plot <- ggplot(
+    screening,
+    aes(
+        x = ReviewYear,
+        y = value,
         color = age_group
-)) +
+    )
+) +
     geom_line() +
@@
 screening_plot
-ggsave("./Dissemination/Figures/screening_rates.png",
-  plot = screening_plot, width = 12, height = 9, units = "in"
+ggsave(
+    "./Dissemination/Figures/screening_rates.png",
+    plot = screening_plot,
+    width = 12,
+    height = 9,
+    units = "in"
 )
@@
         `Age Group` = last(na.omit(age_group)),
-    `OGTTs` = sum(!is.na(A_twohour_bloodglucose) |
-      !is.na(A_ogttfast_bloodglucose))
+        `OGTTs` = sum(
+            !is.na(A_twohour_bloodglucose) |
+                !is.na(A_ogttfast_bloodglucose)
+        )
     )
@@
 # Format OGTT counts
-ogtts$OGTTs <- factor(ogtts$OGTTs,
-  levels = 0:3, labels = c(
-    "No OGTT", "At least 1 OGTT", "At least 1 OGTT", "At least 1 OGTT"
+ogtts$OGTTs <- factor(
+    ogtts$OGTTs,
+    levels = 0:3,
+    labels = c(
+        "No OGTT",
+        "At least 1 OGTT",
+        "At least 1 OGTT",
+        "At least 1 OGTT"
     )
@@
     tbl_summary(by = `Age Group`) %>%
-  add_ci() %>%
+    # add_ci() %>%
     add_p() %>%
@@
     tbl_summary(by = Gender) %>%
-  add_ci() %>%
+    # add_ci() %>%
     add_p() %>%
@@
     tbl_summary(by = MutClass) %>%
-  add_ci() %>%
+    # add_ci() %>%
     add_p() %>%
@@
     tbl_summary(by = Race) %>%
-  add_ci() %>%
+    # add_ci() %>%
     add_p() %>%
@@
     tbl_summary(by = Hispanicrace) %>%
-  add_ci() %>%
+    # add_ci() %>%
     add_p() %>%
@@
     select(
-    Year, `Total Incidence Age 10-18`, `Total Incidence Age 19-30`,
+        Year,
+        `Total Incidence Age 10-18`,
+        `Total Incidence Age 19-30`,
         `Total Incidence Age >= 31`
     ) %>%
-  pivot_longer(`Total Incidence Age 10-18`:`Total Incidence Age >= 31`,
+    pivot_longer(
+        `Total Incidence Age 10-18`:`Total Incidence Age >= 31`,
         names_to = "Age Group"
@@
 t <- rbind(p, i, s)
-t$`Age Group` <- factor(t$`Age Group`,
+t$`Age Group` <- factor(
+    t$`Age Group`,
     levels = c(
-    "[10,19)", "[19,31)", "[31, Inf)", "Total Incidence Age >= 31",
-    "Total Incidence Age 10-18", "Total Incidence Age 19-30"
+        "[10,19)",
+        "[19,31)",
+        "[31, Inf)",
+        "Total Incidence Age >= 31",
+        "Total Incidence Age 10-18",
+        "Total Incidence Age 19-30"
     ),
     labels = c(
-    "10-18", "19-30", ">= 31", ">= 31",
-    "10-18", "19-30"
+        "10-18",
+        "19-30",
+        ">= 31",
+        ">= 31",
+        "10-18",
+        "19-30"
     )
@@
     ggtitle("Prevalence of CFRD")
-fig1 <- fig1a + fig1b + fig1c + plot_layout(guides = "collect") +
-  plot_annotation(title = "Figure 1: Trends in Screening, Incidence, and Prevalence of CFRD from 2003-2018, by age group ")
+fig1 <- fig1a +
+    fig1b +
+    fig1c +
+    plot_layout(guides = "collect") +
+    plot_annotation(
+        title = "Figure 1: Trends in Screening, Incidence, and Prevalence of CFRD from 2003-2018, by age group "
+    )
 fig1
-ggsave("./Dissemination/Figures/fig1.png",
-  plot = fig1, width = 12, height = 3, units = "in", dpi = 900
+ggsave(
+    "./Dissemination/Figures/fig1.png",
+    plot = fig1,
+    width = 12,
+    height = 3,
+    units = "in",
+    dpi = 900
 )
@@
     ggtitle("Prevalence of CFRD")
-fig1 <- fig1a + fig1b + fig1c + plot_layout(guides = "collect")+
-  plot_annotation(title = "US Cystic Fibrosis Foundation Patient Registry Trends in Screening, Incidence, and Prevalence of Cystic Fibrosis Related Diabetes from 2003-2018")
+fig1 <- fig1a +
+    fig1b +
+    fig1c +
+    plot_layout(guides = "collect") +
+    plot_annotation(
+        title = "US Cystic Fibrosis Foundation Patient Registry Trends in Screening, Incidence, and Prevalence of Cystic Fibrosis Related Diabetes from 2003-2018"
+    )
 fig1
-ggsave("./Dissemination/Figures/graphical_abstract.png",
-  plot = fig1, width = 13, height = 3, units = "in", dpi = 900
+ggsave(
+    "./Dissemination/Figures/graphical_abstract.png",
+    plot = fig1,
+    width = 13,
+    height = 3,
+    units = "in",
+    dpi = 900
 )
@@
 # Write cleaned data
-save(annual,file = "./Data_Clean/survival_dataset.RData")
+save(annual, file = "./Data_Clean/survival_dataset.RData")
 # Formula for base model
 xvars <- c(
-  "age_group", "Gender", "Race", "Hispanicrace", "bmi_cat", "MutClass",
-  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
-  "NumPulmExacerbation", "A_corticosteroids1"
+    "age_group",
+    "Gender",
+    "Race",
+    "Hispanicrace",
+    "bmi_cat",
+    "MutClass",
+    "A_IsOnEnzymes",
+    "tube_feeding",
+    "liver_disease",
+    "ppFEV1_cat",
+    "NumPulmExacerbation",
+    "A_corticosteroids1"
 )
@@
     tbl_regression(exponentiate = T)
-ggcoef_table(cox_res,
-  show_p_values = F, add_reference_rows = T, exponentiate = T,
-  signif_stars = F, colour = F
+ggcoef_table(
+    cox_res,
+    show_p_values = F,
+    add_reference_rows = T,
+    exponentiate = T,
+    signif_stars = F,
+    colour = F
 )
@@
 # "[2003,2009)", "[2009,2014)", "[2014,2019)"
-annual$time_period <- cut(annual$ReviewYear,
-  breaks = c(-Inf, 2009, 2014, Inf), right = F
+annual$time_period <- cut(
+    annual$ReviewYear,
+    breaks = c(-Inf, 2009, 2014, Inf),
+    right = F
 )
-era1 <- coxph(f,
+era1 <- coxph(
+    f,
     data = annual[as.numeric(annual$time_period) == 1, ],
@@
 )
-era2 <- coxph(f,
+era2 <- coxph(
+    f,
     data = annual[as.numeric(annual$time_period) == 2, ],
@@
 )
-era3 <- coxph(f,
+era3 <- coxph(
+    f,
     data = annual[as.numeric(annual$time_period) == 3, ],
@@
 ```{r results='asis'}
-f_vx770 <- update(f, . ~ . + A_Vx770-A_corticosteroids1-NumPulmExacerbation-age_group)
-vx770_cox_res <- coxph(f_vx770,
+f_vx770 <- update(
+    f,
+    . ~ . + A_Vx770 - A_corticosteroids1 - NumPulmExacerbation - age_group
+)
+vx770_cox_res <- coxph(
+    f_vx770,
     data = annual[annual$ReviewYear >= 2011, ],
@@
 t1
-ggcoef_table(vx770_cox_res,
-  show_p_values = F, add_reference_rows = T, exponentiate = T,
-  signif_stars = F, colour = F
+ggcoef_table(
+    vx770_cox_res,
+    show_p_values = F,
+    add_reference_rows = T,
+    exponentiate = T,
+    signif_stars = F,
+    colour = F
 )
@@
 ```{r}
-f_vx809comb <- update(f, . ~ . + A_VX809comb-A_corticosteroids1-NumPulmExacerbation-age_group-Hispanicrace)
-vx809comb_cox_res <- coxph(f_vx809comb,
+f_vx809comb <- update(
+    f,
+    . ~
+        . +
+            A_VX809comb -
+            A_corticosteroids1 -
+            NumPulmExacerbation -
+            age_group -
+            Hispanicrace
+)
+vx809comb_cox_res <- coxph(
+    f_vx809comb,
     data = annual[annual$ReviewYear >= 2015, ],
@@
 t2
-ggcoef_table(vx809comb_cox_res,
-  show_p_values = F, add_reference_rows = T, exponentiate = T,
-  signif_stars = F, colour = F
+ggcoef_table(
+    vx809comb_cox_res,
+    show_p_values = F,
+    add_reference_rows = T,
+    exponentiate = T,
+    signif_stars = F,
+    colour = F
 )
@@
     list("VX770" = vx770_cox_res, "VX809comb" = vx809comb_cox_res),
-  add_reference_rows = T, exponentiate = T, type = "faceted", colour = F
+    add_reference_rows = T,
+    exponentiate = T,
+    type = "faceted",
+    colour = F
 )
 comb_plot
-ggsave("./Dissemination/Figures/vx_combined_forest.png",
-  plot = comb_plot, width = 12, height = 9, units = "in"
+ggsave(
+    "./Dissemination/Figures/vx_combined_forest.png",
+    plot = comb_plot,
+    width = 12,
+    height = 9,
+    units = "in"
 )
@@
 t1
-ggcoef_model(era1,
-  show_p_values = F, add_reference_rows = T, exponentiate = T, shape_guide = F,
-  signif_stars = F, colour = "significance", shape = "reference_row",
+ggcoef_model(
+    era1,
+    show_p_values = F,
+    add_reference_rows = T,
+    exponentiate = T,
+    shape_guide = F,
+    signif_stars = F,
+    colour = "significance",
+    shape = "reference_row",
     shape_values = c(16, 15)
@@
 t2
-ggcoef_model(era2,
-  show_p_values = F, add_reference_rows = T, exponentiate = T, shape_guide = F,
-  signif_stars = F, colour = "significance", shape = "reference_row",
+ggcoef_model(
+    era2,
+    show_p_values = F,
+    add_reference_rows = T,
+    exponentiate = T,
+    shape_guide = F,
+    signif_stars = F,
+    colour = "significance",
+    shape = "reference_row",
     shape_values = c(16, 15)
@@
 t <- ggcoef_table(era3)
-ggcoef_model(era3,
-  show_p_values = F, add_reference_rows = T, exponentiate = T, shape_guide = F,
-  signif_stars = F, colour = "significance", shape = "reference_row",
+ggcoef_model(
+    era3,
+    show_p_values = F,
+    add_reference_rows = T,
+    exponentiate = T,
+    shape_guide = F,
+    signif_stars = F,
+    colour = "significance",
+    shape = "reference_row",
     shape_values = c(16, 15)
@@
     list("2003 - 2008" = era1, "2009 - 2013" = era2, "2014 - 2018" = era3),
-  add_reference_rows = T, exponentiate = T, shape_guide = F,
-  colour = "significance", shape = "reference_row", shape_values = c(16, 15),
+    add_reference_rows = T,
+    exponentiate = T,
+    shape_guide = F,
+    colour = "significance",
+    shape = "reference_row",
+    shape_values = c(16, 15),
     type = "faceted"
@@
 comb_plot
-ggsave("./Dissemination/Figures/time_combined_forest.png",
-  plot = comb_plot, width = 12, height = 9, units = "in", dpi = 900
+ggsave(
+    "./Dissemination/Figures/time_combined_forest.png",
+    plot = comb_plot,
+    width = 12,
+    height = 9,
+    units = "in",
+    dpi = 900
 )
-
 ```

### 2025-09-16 Christine Chan/CF Registry/cf_registry_CFRD_prediction.qmd
updated plot for Christine

--- 2025-04-17 Christine Chan/CF Registry/cf_registry_CFRD_prediction.qmd
@@
 library(knitr)
-home_dir <- switch(Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
+home_dir <- switch(
+  Sys.info()["sysname"],
+  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
   "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
@@
 # Import data
-annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
+annual <- read.csv(
+  "./Data_Raw/CFF19_Annualized_Del1.csv",
   na.strings = c("", "N/A")
@@
 # Demographic data
-demo <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv",
+demo <- read.csv(
+  "./Data_Raw/CFF19_DemogCFDiag_Del1.csv",
   na.strings = c("", "N/A")
@@
 races <- c(
-  "White", "Black or African American", "American Indian or Alaska Native",
-  "Asian", "Native Hawaiian or Other Pacific Islander", "Some other race"
+  "White",
+  "Black or African American",
+  "American Indian or Alaska Native",
+  "Asian",
+  "Native Hawaiian or Other Pacific Islander",
+  "Some other race"
 )
@@
 })
-demo$Race <- factor(demo$Race,
+demo$Race <- factor(
+  demo$Race,
   levels = c(
-    "White", "Black or African American", "American Indian or Alaska Native",
-    "Asian", "More than one race", "Native Hawaiian or Other Pacific Islander",
+    "White",
+    "Black or African American",
+    "American Indian or Alaska Native",
+    "Asian",
+    "More than one race",
+    "Native Hawaiian or Other Pacific Islander",
     "Some other race"
@@
   labels = c(
-    "White", "Black or African American", "Other race", "Other race",
-    "More than one race", "Other race", "Other race"
+    "White",
+    "Black or African American",
+    "Other race",
+    "Other race",
+    "More than one race",
+    "Other race",
+    "Other race"
   )
 )
-demo$Hispanicrace <- factor(demo$Hispanicrace,
+demo$Hispanicrace <- factor(
+  demo$Hispanicrace,
   levels = 1:2,
@@
 # F508
-demo$F508 <- factor(demo$F508,
+demo$F508 <- factor(
+  demo$F508,
   levels = 1:3,
@@
 annual$A_Vx770 <- factor(annual$A_Vx770, levels = 0:1, labels = c("No", "Yes"))
-annual$A_VX809comb <- factor(annual$A_VX809comb,
-  levels = 0:1, labels = c("No", "Yes")
+annual$A_VX809comb <- factor(
+  annual$A_VX809comb,
+  levels = 0:1,
+  labels = c("No", "Yes")
 )
@@
 # Format columns
-annual$MutClass <- factor(annual$MutClass,
+annual$MutClass <- factor(
+  annual$MutClass,
   levels = c("1-3", "4-5", "Oth", "N/A"),
@@
 )
-annual$NumPulmExacerbation <- cut(annual$NumPulmExacerbation,
+annual$NumPulmExacerbation <- cut(
+  annual$NumPulmExacerbation,
   labels = c("0", "1", "2", ">=3"),
-  breaks = c(0, 1, 2, 3, Inf), right = F
+  breaks = c(0, 1, 2, 3, Inf),
+  right = F
 )
-annual$A_IsOnEnzymes <- factor(annual$A_IsOnEnzymes,
-  levels = 1:0, labels = c("Yes", "No")
+annual$A_IsOnEnzymes <- factor(
+  annual$A_IsOnEnzymes,
+  levels = 1:0,
+  labels = c("Yes", "No")
 )
@@
   Age_YrEnd = "Age at Year End",
-  Hispanicrace = "Hispanic", A_Vx770 = "Use of CFTR modulator Vx770",
+  Hispanicrace = "Hispanic",
+  A_Vx770 = "Use of CFTR modulator Vx770",
   A_VX809comb = "Use of CFTR modulator Vx809 combination",
-  A_FEV1_pct_predicted = "FEV1 percent predicted", MutClass = "Mutation class",
+  A_FEV1_pct_predicted = "FEV1 percent predicted",
+  MutClass = "Mutation class",
   NumPulmExacerbation = "Admissions for PEx",
-  A_IsOnEnzymes = "Pancreatic enzyme use", bmi = "BMI value"
+  A_IsOnEnzymes = "Pancreatic enzyme use",
+  bmi = "BMI value"
 )
@@
 predictors <- c(
-  "Gender", "Race", "Hispanicrace", "A_IsOnEnzymes",
-  "NumPulmExacerbation", "A_FEV1_pct_predicted", "bmi"
+  "Gender",
+  "Race",
+  "Hispanicrace",
+  "A_IsOnEnzymes",
+  "NumPulmExacerbation",
+  "A_FEV1_pct_predicted",
+  "bmi"
 )
@@
   select(
-    eDWID, ReviewYear, Age_YrEnd, A_hgba1c, all_of(predictors), cfrd_registry
+    eDWID,
+    ReviewYear,
+    Age_YrEnd,
+    A_hgba1c,
+    all_of(predictors),
+    cfrd_registry
   ) %>%
@@
   filter(row_number() == 1)
-t1 <- tableby(cfrd_registry ~ Age_YrEnd + Gender + Race + Hispanicrace + A_IsOnEnzymes + NumPulmExacerbation + A_FEV1_pct_predicted + bmi, data = t2)
+t1 <- tableby(
+  cfrd_registry ~
+    Age_YrEnd +
+      Gender +
+      Race +
+      Hispanicrace +
+      A_IsOnEnzymes +
+      NumPulmExacerbation +
+      A_FEV1_pct_predicted +
+      bmi,
+  data = t2
+)
 summary(t1)
@@
 # Fit Cox model first, and then pull the clusters
-cox_fit <- coxph(cox_f,
-  cluster = eDWID, x = T, model = T,
+cox_fit <- coxph(
+  cox_f,
+  cluster = eDWID,
+  x = T,
+  model = T,
   data = surv_reg[surv_reg$eDWID %in% reg_include, ]
 )
-lme_fit_a1c <- lme(lme_f,
-  random = ~ Age_YrEnd | eDWID, na.action = na.omit,
+lme_fit_a1c <- lme(
+  lme_f,
+  random = ~ Age_YrEnd | eDWID,
+  na.action = na.omit,
   data = registry[registry$eDWID %in% reg_include, ]
@@
 
+```{r}
+nd <- joint_fit_value$data
+nd$pred <- predict(
+  joint_fit_value,
+  registry[registry$eDWID %in% reg_include, ],
+  type = "Subject",
+  idVar = "eDWID"
+)
+```
+
 # To do

### 2025-09-16 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
updated plot for Christine

@@
 )
-
 cfrd_only$a1c_ogtt <- apply(cfrd_only, 1, function(r) {
@@
   modify_caption("**OGTT vs. HbA1c Only**")
+# Plot by diagnosis type
+plot_df = cfrd_only %>%
+  select(id, a1c_ogtt) %>%
+  rename(eDWID = id)
+plot_df = left_join(
+  annual %>% mutate(eDWID = as.numeric(eDWID)),
+  plot_df,
+  by = join_by(eDWID)
+)
+plot_df = plot_df %>%
+  group_by(eDWID) %>%
+  mutate(`Years to CFRD` = ReviewYear - first_cfrd)
+ggplot(
+  plot_df %>% filter(!is.na(a1c_ogtt)),
+  aes(
+    y = A_FEV1_pct_predicted,
+    x = `Years to CFRD`,
+    color = a1c_ogtt,
+    group = a1c_ogtt
+  )
+) +
+  xlim(c(-10, 10)) +
+  geom_point(alpha = 0.1, shape = ".") +
+  geom_smooth(formula = y ~ s(x, bs = "cr")) +
+  theme_bw() +
+  theme(legend.title = element_blank())
 ```

### 2025-09-16 Christine Chan/CF Registry/cfrd_1hr_ogtt_cgm_predictive.qmd
updated plot for Christine

@@
 library(performance)
-knitr::opts_knit$set(root.dir = "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo")
+knitr::opts_knit$set(
+  root.dir = "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo"
+)
 ```
@@
 a1c <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/a1c.csv")
-demo <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/demographics.csv")
+demo <- read.csv(
+  "./Christine Chan/1 hour OGTT in CFRD/Data_raw/demographics.csv"
+)
 dob <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/dob.csv")
@@
 ogtt_demo$EncounterDate <- ogtt_demo$Date.of.OGTT
-ogtt_demo_a1c <- merge(ogtt_demo, a1c, by = c("MRN", "EncounterDate"), all.x = T, all.y = T)
+ogtt_demo_a1c <- merge(
+  ogtt_demo,
+  a1c,
+  by = c("MRN", "EncounterDate"),
+  all.x = T,
+  all.y = T
+)
 
@@
 pft$LName <- NULL
-alldata <- merge(ogtt_demo_a1c, pft, by = c("MRN", "EncounterDate"), all.x = T, all.y = T)
+alldata <- merge(
+  ogtt_demo_a1c,
+  pft,
+  by = c("MRN", "EncounterDate"),
+  all.x = T,
+  all.y = T
+)
 # get predicted spirometry values
@@
 alldata$sexnum <- ifelse(alldata$Sex == "Male", 1, 2)
-alldata$FEV1.pred <- pred_GLI(age = alldata$Age.at.PFT, height = alldata$ht_m, gender = alldata$sexnum, param = "FEV1")
-alldata$FVC.pred <- pred_GLI(age = alldata$Age.at.PFT, height = alldata$ht_m, gender = alldata$sexnum, param = "FVC")
-alldata$FEF.25.75.pred <- pred_GLI(age = alldata$Age.at.PFT, height = alldata$ht_m, gender = alldata$sexnum, param = "FEF2575")
+alldata$FEV1.pred <- pred_GLI(
+  age = alldata$Age.at.PFT,
+  height = alldata$ht_m,
+  gender = alldata$sexnum,
+  param = "FEV1"
+)
+alldata$FVC.pred <- pred_GLI(
+  age = alldata$Age.at.PFT,
+  height = alldata$ht_m,
+  gender = alldata$sexnum,
+  param = "FVC"
+)
+alldata$FEF.25.75.pred <- pred_GLI(
+  age = alldata$Age.at.PFT,
+  height = alldata$ht_m,
+  gender = alldata$sexnum,
+  param = "FEF2575"
+)
 alldata$FEV1.per.pred <- (alldata$FEV1.Raw / alldata$FEV1.pred) * 100
 alldata$FVC.per.pred <- (alldata$FVC.Raw / alldata$FVC.pred) * 100
-alldata$FEF.25.75.per.pred <- (alldata$FEF.25.75.Raw / alldata$FEF.25.75.pred) * 100
+alldata$FEF.25.75.per.pred <- (alldata$FEF.25.75.Raw / alldata$FEF.25.75.pred) *
+  100
 # alldata$Sex <- NULL
 # hard code some corrections to data per CC
-alldata[alldata$MRN == 686476, ]$X1.hour.Glucose <- ifelse(alldata[alldata$MRN == 686476, ]$Age.At.Test..years. == 14.5, 185, alldata[alldata$MRN == 686476, ]$X1.hour.Glucose)
+alldata[alldata$MRN == 686476, ]$X1.hour.Glucose <- ifelse(
+  alldata[alldata$MRN == 686476, ]$Age.At.Test..years. == 14.5,
+  185,
+  alldata[alldata$MRN == 686476, ]$X1.hour.Glucose
+)
 # severity of genotype
@@
 # Green = not CF causing, classify as "other", will most likely be seen in people with CRMS diagnosis
-alldata$class2 <- ifelse(alldata$Mutation.2 %in% c(
-  "1717-1G->A", "1717-1G-A", "G542X", "Q493X", "R1162X", "R553X", "R553x", "W1089X", "W1282X", "1078delT", "R75X",
-  "3659delC", "621+1G->T", "621+1G>T", "394delTT", "3120 + 1g->A", "1154InsTC", "1154insTC", "1213delT",
-  "1259insA", "1288insTA", "3791delC", "E60X", "K710X", "2184delA", "CFTRdele2,3", "663delT", "Glu528",
-  "1461ins4", "306insA", "R709X", "CFTRdele22-24", "2711delT", "2183AA- >G", "2183delAA>G"
-), "I",
-ifelse(alldata$Mutation.2 %in% c("F508", "F508del", "F508del ", "G85E", "I507", "I 507", "N1303K", "A559T", "R560T", "A561E"),
+alldata$class2 <- ifelse(
+  alldata$Mutation.2 %in%
+    c(
+      "1717-1G->A",
+      "1717-1G-A",
+      "G542X",
+      "Q493X",
+      "R1162X",
+      "R553X",
+      "R553x",
+      "W1089X",
+      "W1282X",
+      "1078delT",
+      "R75X",
+      "3659delC",
+      "621+1G->T",
+      "621+1G>T",
+      "394delTT",
+      "3120 + 1g->A",
+      "1154InsTC",
+      "1154insTC",
+      "1213delT",
+      "1259insA",
+      "1288insTA",
+      "3791delC",
+      "E60X",
+      "K710X",
+      "2184delA",
+      "CFTRdele2,3",
+      "663delT",
+      "Glu528",
+      "1461ins4",
+      "306insA",
+      "R709X",
+      "CFTRdele22-24",
+      "2711delT",
+      "2183AA- >G",
+      "2183delAA>G"
+    ),
+  "I",
+  ifelse(
+    alldata$Mutation.2 %in%
+      c(
+        "F508",
+        "F508del",
+        "F508del ",
+        "G85E",
+        "I507",
+        "I 507",
+        "N1303K",
+        "A559T",
+        "R560T",
+        "A561E"
+      ),
     "II",
-  ifelse(alldata$Mutation.2 %in% c("G551D", "G551S", "S549N"), "III",
-    ifelse(alldata$Mutation.2 %in% c("R334W", "R347P", "R117H", "R117C", "P67L", "L206W", "l206w", "I206w", "D614G", "R347H", "D1152H", "3849+10kbC>T"), "IV",
-      ifelse(alldata$Mutation.2 %in% c(
-        "A455E", "2789+5G->A", "3849+10CT", "3849+10kbC->T", "711+3A->G", "1898+5G->T",
-        "P574H", "3272-26A->G5", "3272-26A->G"
-      ), "V", "")
+    ifelse(
+      alldata$Mutation.2 %in% c("G551D", "G551S", "S549N"),
+      "III",
+      ifelse(
+        alldata$Mutation.2 %in%
+          c(
+            "R334W",
+            "R347P",
+            "R117H",
+            "R117C",
+            "P67L",
+            "L206W",
+            "l206w",
+            "I206w",
+            "D614G",
+            "R347H",
+            "D1152H",
+            "3849+10kbC>T"
+          ),
+        "IV",
+        ifelse(
+          alldata$Mutation.2 %in%
+            c(
+              "A455E",
+              "2789+5G->A",
+              "3849+10CT",
+              "3849+10kbC->T",
+              "711+3A->G",
+              "1898+5G->T",
+              "P574H",
+              "3272-26A->G5",
+              "3272-26A->G"
+            ),
+          "V",
+          ""
         )
@@
   )
-alldata$severity2 <- ifelse(alldata$Mutation.2 %in% c(
-  "1717-1G->A", "1717-1G-A", "G542X", "Q493X", "R1162X", "R553X", "R553x", "W1089X", "W1282X", "1078delT", "R75X",
-  "3659delC", "621+1G->T", "621+1G>T", "394delTT", "3120 + 1g->A", "1154InsTC", "1154insTC", "1213delT",
-  "1259insA", "1288insTA", "3791delC", "E60X", "K710X", "2184delA", "CFTRdele2,3", "663delT", "Glu528",
-  "1461ins4", "306insA", "R709X", "CFTRdele22-24", "2711delT", "2183AA- >G", "2183delAA>G", "F508", "F508del", "F508del ", "G85E",
-  "I507", "I 507", "N1303K", "A559T", "R560T", "A561E", "G551D", "G551S", "S549N", "1898+1G->A", "2143delT", "2183AA-G",
-  "2183delAA->G", "2184insA", "S434X", "2585delT", "3120+1G->A", "3905insT", "406-1G->A", "712-1G->T", "L1245X", "Q2X", "R1158X",
-  "R851X", "V520F", "L1077P", "S489X", "F311del", "W846X", "M1101K", "2622+1G->A", "S945L", "Y1092X", "R1066C", "G178R", "2957delT",
-  "CFTRdele1", "W1204X", "q493x", "W57G", "1585-8G>A", "1717-8G->A", "G1244E"
-), "Severe",
-ifelse(alldata$Mutation.2 %in% c(
-  "R334W", "R347P", "R117H", "R117C", "P67L", "L206W", "l206w", "I206w", "D614G", "R347H", "D1152H", "3849+10kbC>T", "D110H", "R1070W",
-  "A455E", "2789+5G->A", "3849+10CT", "3849+10kbC->T", "711+3A->G", "1898+5G->T", "P574H", "3272-26A->G5", "3272-26A->G", "3849G->A", "H199Y", "H199y",
-  "S492F", "S531P", "L1335P", "R352Q", "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val",
-  "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"
-), "Mild",
-ifelse(alldata$Mutation.2 %in% c(
-  "1262insA", "I506T", "3363delGT", "Q1209P", "3349insT", "G178R", "L467P", "317insC", "I336K",
-  "ex14a", "3500-2A->G", "c.3407_3422del16", "3349insT", "1248+1G->A", "M470V", "1812-1G->A",
-  "del X22-24", "Q237H", "Q237H", "W57G", "1078delA", "G178R", "2957delT", "G576A",
-  "other", "Other", "Unknown", "unknown", "CFTRdele1", "1811+1.6kb A->", "UNK", "I336K", "M470V", "S1235R", "M470V", "F508C",
-  "R1066H", "P205S"
-), "Unk", "")
+)
+alldata$severity2 <- ifelse(
+  alldata$Mutation.2 %in%
+    c(
+      "1717-1G->A",
+      "1717-1G-A",
+      "G542X",
+      "Q493X",
+      "R1162X",
+      "R553X",
+      "R553x",
+      "W1089X",
+      "W1282X",
+      "1078delT",
+      "R75X",
+      "3659delC",
+      "621+1G->T",
+      "621+1G>T",
+      "394delTT",
+      "3120 + 1g->A",
+      "1154InsTC",
+      "1154insTC",
+      "1213delT",
+      "1259insA",
+      "1288insTA",
+      "3791delC",
+      "E60X",
+      "K710X",
+      "2184delA",
+      "CFTRdele2,3",
+      "663delT",
+      "Glu528",
+      "1461ins4",
+      "306insA",
+      "R709X",
+      "CFTRdele22-24",
+      "2711delT",
+      "2183AA- >G",
+      "2183delAA>G",
+      "F508",
+      "F508del",
+      "F508del ",
+      "G85E",
+      "I507",
+      "I 507",
+      "N1303K",
+      "A559T",
+      "R560T",
+      "A561E",
+      "G551D",
+      "G551S",
+      "S549N",
+      "1898+1G->A",
+      "2143delT",
+      "2183AA-G",
+      "2183delAA->G",
+      "2184insA",
+      "S434X",
+      "2585delT",
+      "3120+1G->A",
+      "3905insT",
+      "406-1G->A",
+      "712-1G->T",
+      "L1245X",
+      "Q2X",
+      "R1158X",
+      "R851X",
+      "V520F",
+      "L1077P",
+      "S489X",
+      "F311del",
+      "W846X",
+      "M1101K",
+      "2622+1G->A",
+      "S945L",
+      "Y1092X",
+      "R1066C",
+      "G178R",
+      "2957delT",
+      "CFTRdele1",
+      "W1204X",
+      "q493x",
+      "W57G",
+      "1585-8G>A",
+      "1717-8G->A",
+      "G1244E"
+    ),
+  "Severe",
+  ifelse(
+    alldata$Mutation.2 %in%
+      c(
+        "R334W",
+        "R347P",
+        "R117H",
+        "R117C",
+        "P67L",
+        "L206W",
+        "l206w",
+        "I206w",
+        "D614G",
+        "R347H",
+        "D1152H",
+        "3849+10kbC>T",
+        "D110H",
+        "R1070W",
+        "A455E",
+        "2789+5G->A",
+        "3849+10CT",
+        "3849+10kbC->T",
+        "711+3A->G",
+        "1898+5G->T",
+        "P574H",
+        "3272-26A->G5",
+        "3272-26A->G",
+        "3849G->A",
+        "H199Y",
+        "H199y",
+        "S492F",
+        "S531P",
+        "L1335P",
+        "R352Q",
+        "D1270N",
+        "2789+2insA",
+        "3849+10kbc->T",
+        "Q98R",
+        "Phe1052Val",
+        "F1052V",
+        "Tyr1014Cys",
+        "Y1014C",
+        "T338I",
+        "5T",
+        "c.2249C>T",
+        "P750L"
+      ),
+    "Mild",
+    ifelse(
+      alldata$Mutation.2 %in%
+        c(
+          "1262insA",
+          "I506T",
+          "3363delGT",
+          "Q1209P",
+          "3349insT",
+          "G178R",
+          "L467P",
+          "317insC",
+          "I336K",
+          "ex14a",
+          "3500-2A->G",
+          "c.3407_3422del16",
+          "3349insT",
+          "1248+1G->A",
+          "M470V",
+          "1812-1G->A",
+          "del X22-24",
+          "Q237H",
+          "Q237H",
+          "W57G",
+          "1078delA",
+          "G178R",
+          "2957delT",
+          "G576A",
+          "other",
+          "Other",
+          "Unknown",
+          "unknown",
+          "CFTRdele1",
+          "1811+1.6kb A->",
+          "UNK",
+          "I336K",
+          "M470V",
+          "S1235R",
+          "M470V",
+          "F508C",
+          "R1066H",
+          "P205S"
+        ),
+      "Unk",
+      ""
     )
   )
-alldata$class1 <- ifelse(alldata$Mutation.1 %in% c(
-  "1717-1G->A", "1717-1G-A", "G542X", "Q493X", "R1162X", "R553X", "R553x", "W1089X", "W1282X", "1078delT", "R75X",
-  "3659delC", "621+1G->T", "621+1G>T", "394delTT", "3120 + 1g->A", "1154InsTC", "1154insTC", "1213delT",
-  "1259insA", "1288insTA", "3791delC", "E60X", "K710X", "2184delA", "CFTRdele2,3", "663delT", "Glu528",
-  "1461ins4", "306insA", "R709X", "CFTRdele22-24", "2711delT", "2183AA- >G", "2183delAA>G"
-), "I",
-ifelse(alldata$Mutation.1 %in% c("F508", "F508del", "F508del ", "G85E", "I507", "I 507", "N1303K", "A559T", "R560T", "A561E"), "II",
-  ifelse(alldata$Mutation.1 %in% c("G551D", "G551S", "S549N"), "III",
-    ifelse(alldata$Mutation.1 %in% c("R334W", "R347P", "R117H", "R117C", "P67L", "L206W", "l206w", "I206w", "D614G", "R347H", "D1152H", "3849+10kbC>T"), "IV",
-      ifelse(alldata$Mutation.1 %in% c(
-        "A455E", "2789+5G->A", "3849+10CT", "3849+10kbC->T", "711+3A->G", "1898+5G->T",
-        "P574H", "3272-26A->G5", "3272-26A->G"
-      ), "V", "")
+)
+alldata$class1 <- ifelse(
+  alldata$Mutation.1 %in%
+    c(
+      "1717-1G->A",
+      "1717-1G-A",
+      "G542X",
+      "Q493X",
+      "R1162X",
+      "R553X",
+      "R553x",
+      "W1089X",
+      "W1282X",
+      "1078delT",
+      "R75X",
+      "3659delC",
+      "621+1G->T",
+      "621+1G>T",
+      "394delTT",
+      "3120 + 1g->A",
+      "1154InsTC",
+      "1154insTC",
+      "1213delT",
+      "1259insA",
+      "1288insTA",
+      "3791delC",
+      "E60X",
+      "K710X",
+      "2184delA",
+      "CFTRdele2,3",
+      "663delT",
+      "Glu528",
+      "1461ins4",
+      "306insA",
+      "R709X",
+      "CFTRdele22-24",
+      "2711delT",
+      "2183AA- >G",
+      "2183delAA>G"
+    ),
+  "I",
+  ifelse(
+    alldata$Mutation.1 %in%
+      c(
+        "F508",
+        "F508del",
+        "F508del ",
+        "G85E",
+        "I507",
+        "I 507",
+        "N1303K",
+        "A559T",
+        "R560T",
+        "A561E"
+      ),
+    "II",
+    ifelse(
+      alldata$Mutation.1 %in% c("G551D", "G551S", "S549N"),
+      "III",
+      ifelse(
+        alldata$Mutation.1 %in%
+          c(
+            "R334W",
+            "R347P",
+            "R117H",
+            "R117C",
+            "P67L",
+            "L206W",
+            "l206w",
+            "I206w",
+            "D614G",
+            "R347H",
+            "D1152H",
+            "3849+10kbC>T"
+          ),
+        "IV",
+        ifelse(
+          alldata$Mutation.1 %in%
+            c(
+              "A455E",
+              "2789+5G->A",
+              "3849+10CT",
+              "3849+10kbC->T",
+              "711+3A->G",
+              "1898+5G->T",
+              "P574H",
+              "3272-26A->G5",
+              "3272-26A->G"
+            ),
+          "V",
+          ""
         )
@@
   )
-alldata$severity1 <- ifelse(alldata$Mutation.1 %in% c(
-  "1717-1G->A", "1717-1G-A", "G542X", "Q493X", "R1162X", "R553X", "R553x", "W1089X", "W1282X", "1078delT", "R75X",
-  "3659delC", "621+1G->T", "621+1G>T", "394delTT", "3120 + 1g->A", "1154InsTC", "1154insTC", "1213delT",
-  "1259insA", "1288insTA", "3791delC", "E60X", "K710X", "2184delA", "CFTRdele2,3", "663delT", "Glu528",
-  "1461ins4", "306insA", "R709X", "CFTRdele22-24", "2711delT", "2183AA- >G", "2183delAA>G", "F508", "F508del", "F508del ", "G85E",
-  "I507", "I 507", "N1303K", "A559T", "R560T", "A561E", "G551D", "G551S", "S549N", "1898+1G->A", "2143delT", "2183AA-G",
-  "2183delAA->G", "2184insA", "S434X", "2585delT", "3120+1G->A", "3905insT", "406-1G->A", "712-1G->T", "L1245X", "Q2X", "R1158X",
-  "R851X", "V520F", "L1077P", "S489X", "F311del", "W846X", "M1101K", "2622+1G->A", "S945L",
-  "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", "W1204X", "q493x", "W57G", "1585-8G>A", "1717-8G->A", "G1244E"
-), "Severe",
-ifelse(alldata$Mutation.1 %in% c(
-  "R334W", "R347P", "R117H", "R117C", "P67L", "L206W", "l206w", "I206w", "D614G", "R347H", "D1152H", "3849+10kbC>T", "D110H", "R1070W",
-  "A455E", "2789+5G->A", "3849+10CT", "3849+10kbC->T", "711+3A->G", "1898+5G->T",
-  "P574H", "3272-26A->G5", "3272-26A->G", "3849G->A", "H199Y", "H199y", "S492F", "S531P",
-  "L1335P", "R352Q", "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val",
-  "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"
-), "Mild",
-ifelse(alldata$Mutation.1 %in% c(
-  "1262insA", "I506T", "3363delGT", "Q1209P", "3349insT", "G178R", "L467P", "317insC", "I336K",
-  "ex14a", "3500-2A->G", "c.3407_3422del16", "3349insT", "1248+1G->A", "M470V", "1812-1G->A",
-  "del X22-24", "Q237H", "Q237H", "W57G", "1078delA", "G178R", "2957delT", "G576A",
-  "other", "Other", "Unknown", "unknown", "CFTRdele1", "1811+1.6kb A->", "UNK", "I336K", "M470V", "S1235R", "M470V", "F508C",
-  "R1066H", "P205S"
-), "Unk", "")
+)
+alldata$severity1 <- ifelse(
+  alldata$Mutation.1 %in%
+    c(
+      "1717-1G->A",
+      "1717-1G-A",
+      "G542X",
+      "Q493X",
+      "R1162X",
+      "R553X",
+      "R553x",
+      "W1089X",
+      "W1282X",
+      "1078delT",
+      "R75X",
+      "3659delC",
+      "621+1G->T",
+      "621+1G>T",
+      "394delTT",
+      "3120 + 1g->A",
+      "1154InsTC",
+      "1154insTC",
+      "1213delT",
+      "1259insA",
+      "1288insTA",
+      "3791delC",
+      "E60X",
+      "K710X",
+      "2184delA",
+      "CFTRdele2,3",
+      "663delT",
+      "Glu528",
+      "1461ins4",
+      "306insA",
+      "R709X",
+      "CFTRdele22-24",
+      "2711delT",
+      "2183AA- >G",
+      "2183delAA>G",
+      "F508",
+      "F508del",
+      "F508del ",
+      "G85E",
+      "I507",
+      "I 507",
+      "N1303K",
+      "A559T",
+      "R560T",
+      "A561E",
+      "G551D",
+      "G551S",
+      "S549N",
+      "1898+1G->A",
+      "2143delT",
+      "2183AA-G",
+      "2183delAA->G",
+      "2184insA",
+      "S434X",
+      "2585delT",
+      "3120+1G->A",
+      "3905insT",
+      "406-1G->A",
+      "712-1G->T",
+      "L1245X",
+      "Q2X",
+      "R1158X",
+      "R851X",
+      "V520F",
+      "L1077P",
+      "S489X",
+      "F311del",
+      "W846X",
+      "M1101K",
+      "2622+1G->A",
+      "S945L",
+      "Y1092X",
+      "R1066C",
+      "G178R",
+      "2957delT",
+      "CFTRdele1",
+      "W1204X",
+      "q493x",
+      "W57G",
+      "1585-8G>A",
+      "1717-8G->A",
+      "G1244E"
+    ),
+  "Severe",
+  ifelse(
+    alldata$Mutation.1 %in%
+      c(
+        "R334W",
+        "R347P",
+        "R117H",
+        "R117C",
+        "P67L",
+        "L206W",
+        "l206w",
+        "I206w",
+        "D614G",
+        "R347H",
+        "D1152H",
+        "3849+10kbC>T",
+        "D110H",
+        "R1070W",
+        "A455E",
+        "2789+5G->A",
+        "3849+10CT",
+        "3849+10kbC->T",
+        "711+3A->G",
+        "1898+5G->T",
+        "P574H",
+        "3272-26A->G5",
+        "3272-26A->G",
+        "3849G->A",
+        "H199Y",
+        "H199y",
+        "S492F",
+        "S531P",
+        "L1335P",
+        "R352Q",
+        "D1270N",
+        "2789+2insA",
+        "3849+10kbc->T",
+        "Q98R",
+        "Phe1052Val",
+        "F1052V",
+        "Tyr1014Cys",
+        "Y1014C",
+        "T338I",
+        "5T",
+        "c.2249C>T",
+        "P750L"
+      ),
+    "Mild",
+    ifelse(
+      alldata$Mutation.1 %in%
+        c(
+          "1262insA",
+          "I506T",
+          "3363delGT",
+          "Q1209P",
+          "3349insT",
+          "G178R",
+          "L467P",
+          "317insC",
+          "I336K",
+          "ex14a",
+          "3500-2A->G",
+          "c.3407_3422del16",
+          "3349insT",
+          "1248+1G->A",
+          "M470V",
+          "1812-1G->A",
+          "del X22-24",
+          "Q237H",
+          "Q237H",
+          "W57G",
+          "1078delA",
+          "G178R",
+          "2957delT",
+          "G576A",
+          "other",
+          "Other",
+          "Unknown",
+          "unknown",
+          "CFTRdele1",
+          "1811+1.6kb A->",
+          "UNK",
+          "I336K",
+          "M470V",
+          "S1235R",
+          "M470V",
+          "F508C",
+          "R1066H",
+          "P205S"
+        ),
+      "Unk",
+      ""
+    )
   )
@@
 # 3 patients have severity2="Severe" per Christine
-alldata$severity2 <- ifelse(alldata$MRN %in% c(546678, 664400, 2015506), "Severe", alldata$severity2)
-alldata$genoRisk <- ifelse(alldata$severity1 == "Mild" | alldata$severity2 == "Mild", "Low",
-  ifelse(alldata$severity1 == "Severe" & alldata$severity2 == "Severe", "High", "")
+alldata$severity2 <- ifelse(
+  alldata$MRN %in% c(546678, 664400, 2015506),
+  "Severe",
+  alldata$severity2
+)
+alldata$genoRisk <- ifelse(
+  alldata$severity1 == "Mild" | alldata$severity2 == "Mild",
+  "Low",
+  ifelse(
+    alldata$severity1 == "Severe" & alldata$severity2 == "Severe",
+    "High",
+    ""
+  )
 )
@@
 # merge back all baseline characteristics
-d <- alldata[, c("MRN", "FName", "LName", "Sex", "Race", "Latino.a.", "Mutation.1", "Mutation.2", "class1", "class2", "severity1", "severity2", "genoRisk")]
+d <- alldata[, c(
+  "MRN",
+  "FName",
+  "LName",
+  "Sex",
+  "Race",
+  "Latino.a.",
+  "Mutation.1",
+  "Mutation.2",
+  "class1",
+  "class2",
+  "severity1",
+  "severity2",
+  "genoRisk"
+)]
 d <- d[!is.na(d$FName), ]
@@
 # checking ppts that Collin said had CFRD
-col_check <- alldata[alldata$MRN %in% c(
-  1006479, 1053930, 1078408, 1136992, 1289450, 1339023, 1413821, 1695512, 664400,
-  686476, 707831, 712676, 739231, 739460, 787372, 800007, 848440, 860181, 875128, 897478, 905284, 908966,
-  918124, 932695, 949042, 952250, 952252, 825118
-), ]
+col_check <- alldata[
+  alldata$MRN %in%
+    c(
+      1006479,
+      1053930,
+      1078408,
+      1136992,
+      1289450,
+      1339023,
+      1413821,
+      1695512,
+      664400,
+      686476,
+      707831,
+      712676,
+      739231,
+      739460,
+      787372,
+      800007,
+      848440,
+      860181,
+      875128,
+      897478,
+      905284,
+      908966,
+      918124,
+      932695,
+      949042,
+      952250,
+      952252,
+      825118
+    ),
+]
 col_check <- col_check[order(col_check$MRN, col_check$Date), ]
@@
 colorder <- c(
-  "MRN", "FName", "LName", "Sex", "Race", "Latino.a.", "Mutation.1", "Mutation.2", "class1", "class2", "severity1", "severity2", "genoRisk",
-  "Date of first OGTT", "Date", "Age.At.Test..years.", "Fasting.Glucose", "X1.hour.Glucose", "X2.hour.Glucose", "A1C",
-  "DOB", "Age.at.PFT", "FVC.Raw", "Height..cm.", "Weight..kg.", "FEV1.Raw", "FEF.25.75.Raw", "ht_m",
-  "sexnum", "FEV1.pred", "FVC.pred", "FEF.25.75.pred", "FEV1.per.pred", "FVC.per.pred", "FEF.25.75.per.pred", "BMI"
+  "MRN",
+  "FName",
+  "LName",
+  "Sex",
+  "Race",
+  "Latino.a.",
+  "Mutation.1",
+  "Mutation.2",
+  "class1",
+  "class2",
+  "severity1",
+  "severity2",
+  "genoRisk",
+  "Date of first OGTT",
+  "Date",
+  "Age.At.Test..years.",
+  "Fasting.Glucose",
+  "X1.hour.Glucose",
+  "X2.hour.Glucose",
+  "A1C",
+  "DOB",
+  "Age.at.PFT",
+  "FVC.Raw",
+  "Height..cm.",
+  "Weight..kg.",
+  "FEV1.Raw",
+  "FEF.25.75.Raw",
+  "ht_m",
+  "sexnum",
+  "FEV1.pred",
+  "FVC.pred",
+  "FEF.25.75.pred",
+  "FEV1.per.pred",
+  "FVC.per.pred",
+  "FEF.25.75.per.pred",
+  "BMI"
 )
@@
 # create combined race/ethnicity
-alldata$race_eth <- ifelse(alldata$Latino.a. == "TRUE", "Hispanic", alldata$Race)
+alldata$race_eth <- ifelse(
+  alldata$Latino.a. == "TRUE",
+  "Hispanic",
+  alldata$Race
+)
 # get BMI percentile
-alldata$age <- ifelse(is.na(alldata$Age.at.PFT), alldata$Age.At.Test..years., alldata$Age.at.PFT)
+alldata$age <- ifelse(
+  is.na(alldata$Age.at.PFT),
+  alldata$Age.At.Test..years.,
+  alldata$Age.at.PFT
+)
 alldata$age_bmi <- ifelse(alldata$age > 20, 20, alldata$age)
-alldata$bmip <- sds(alldata$BMI,
+alldata$bmip <- sds(
+  alldata$BMI,
   age = alldata$age_bmi,
-  sex = alldata$Sex, male = "Male", female = "Female",
+  sex = alldata$Sex,
+  male = "Male",
+  female = "Female",
   ref = cdc.ref,
@@
 alldata$cfrd_visit <- NA
-alldata$cfrd_visit <- ifelse(is.na(alldata$X2.hour.Glucose), NA, ifelse(alldata$X2.hour.Glucose >= 200, 1, 0))
+alldata$cfrd_visit <- ifelse(
+  is.na(alldata$X2.hour.Glucose),
+  NA,
+  ifelse(alldata$X2.hour.Glucose >= 200, 1, 0)
+)
 # find people with CFRD at baseline
-cfrdbase <- alldata[alldata$Date == alldata$`Date of first OGTT` & alldata$cfrd_visit == 1, ]
+cfrdbase <- alldata[
+  alldata$Date == alldata$`Date of first OGTT` & alldata$cfrd_visit == 1,
+]
 before_cfrd <- length(unique(alldata$MRN))
@@
 # get baseline age
-baseage <- alldata[alldata$Date == alldata$`Date of first OGTT`, c("MRN", "age")]
+baseage <- alldata[
+  alldata$Date == alldata$`Date of first OGTT`,
+  c("MRN", "age")
+]
 colnames(baseage) <- c("MRN", "Base_age")
@@
 # get baseline 1 hour and 2 hour glucose
-basegluc <- alldata[alldata$Date == alldata$`Date of first OGTT`, c("MRN", "X1.hour.Glucose", "X2.hour.Glucose")]
+basegluc <- alldata[
+  alldata$Date == alldata$`Date of first OGTT`,
+  c("MRN", "X1.hour.Glucose", "X2.hour.Glucose")
+]
 colnames(basegluc) <- c("MRN", "Baseline_1hr_glucose", "Baseline_2hr_glucose")
@@
 # dichotomize baseline 1 hour glucose
-alldata$Baseline_1hr_glucose_155 <- ifelse(alldata$Baseline_1hr_glucose >= 155, 1, 0)
-alldata$Baseline_1hr_glucose_140 <- ifelse(alldata$Baseline_1hr_glucose >= 140, 1, 0)
+alldata$Baseline_1hr_glucose_155 <- ifelse(
+  alldata$Baseline_1hr_glucose >= 155,
+  1,
+  0
+)
+alldata$Baseline_1hr_glucose_140 <- ifelse(
+  alldata$Baseline_1hr_glucose >= 140,
+  1,
+  0
+)
 # dichotomize baseline 2 hour glucose
-alldata$Baseline_2hr_glucose_140 <- ifelse(alldata$Baseline_2hr_glucose >= 140, 1, 0)
+alldata$Baseline_2hr_glucose_140 <- ifelse(
+  alldata$Baseline_2hr_glucose >= 140,
+  1,
+  0
+)
 # find date of last OGTT
 lastogtt <- alldata[
-  !is.na(alldata$Fasting.Glucose) | !is.na(alldata$X1.hour.Glucose) | !is.na(alldata$X2.hour.Glucose),
+  !is.na(alldata$Fasting.Glucose) |
+    !is.na(alldata$X1.hour.Glucose) |
+    !is.na(alldata$X2.hour.Glucose),
   c("MRN", "Date")
@@
 # read in date of modulator use
-mod <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/modulators.csv", na.strings = c("", " ", NA))
+mod <- read.csv(
+  "./Christine Chan/1 hour OGTT in CFRD/Data_raw/modulators.csv",
+  na.strings = c("", " ", NA)
+)
 mod$Kalydeco.Start.Date <- as.Date(mod$Kalydeco.Start.Date, format = "%m/%d/%Y")
@@
   mod[i, ]$modstart <-
-    as.Date(pmin(mod[i, ]$Kalydeco.Start.Date, mod[i, ]$Orkambi.Start.Date, mod[i, ]$Symdeko.Start.Date, mod[i, ]$Trikafta.Start.Date, na.rm = T))
-  mod[i, ]$he_modstart <- as.Date(pmin(mod[i, ]$Kalydeco.Start.Date, mod[i, ]$Trikafta.Start.Date, na.rm = T))
-  mod[i, ]$me_modstart <- as.Date(pmin(mod[i, ]$Orkambi.Start.Date, mod[i, ]$Symdeko.Start.Date, na.rm = T))
+    as.Date(pmin(
+      mod[i, ]$Kalydeco.Start.Date,
+      mod[i, ]$Orkambi.Start.Date,
+      mod[i, ]$Symdeko.Start.Date,
+      mod[i, ]$Trikafta.Start.Date,
+      na.rm = T
+    ))
+  mod[i, ]$he_modstart <- as.Date(pmin(
+    mod[i, ]$Kalydeco.Start.Date,
+    mod[i, ]$Trikafta.Start.Date,
+    na.rm = T
+  ))
+  mod[i, ]$me_modstart <- as.Date(pmin(
+    mod[i, ]$Orkambi.Start.Date,
+    mod[i, ]$Symdeko.Start.Date,
+    na.rm = T
+  ))
 }
@@
 alldata <- merge(alldata, modkeep, by = "MRN", all.x = TRUE, all.y = FALSE)
-alldata$onmod <- ifelse(!is.na(alldata$modstart) & alldata$modstart <= alldata$Date, 1,
+alldata$onmod <- ifelse(
+  !is.na(alldata$modstart) & alldata$modstart <= alldata$Date,
+  1,
   ifelse(!is.na(alldata$modstart) & alldata$modstart > alldata$Date, 0, NA)
 )
-alldata$onmod <- ifelse(is.na(alldata$modstart) & is.na(alldata$onmod), 0, alldata$onmod)
-alldata$onhemod <- ifelse(!is.na(alldata$he_modstart) & alldata$he_modstart <= alldata$Date, 1,
-  ifelse(!is.na(alldata$he_modstart) & alldata$he_modstart > alldata$Date, 0, NA)
+alldata$onmod <- ifelse(
+  is.na(alldata$modstart) & is.na(alldata$onmod),
+  0,
+  alldata$onmod
 )
-alldata$onhemod <- ifelse(is.na(alldata$he_modstart) & is.na(alldata$onhemod), 0, alldata$onhemod)
-alldata$onmemod <- ifelse(!is.na(alldata$me_modstart) & alldata$me_modstart <= alldata$Date, 1,
-  ifelse(!is.na(alldata$me_modstart) & alldata$me_modstart > alldata$Date, 0, NA)
+alldata$onhemod <- ifelse(
+  !is.na(alldata$he_modstart) & alldata$he_modstart <= alldata$Date,
+  1,
+  ifelse(
+    !is.na(alldata$he_modstart) & alldata$he_modstart > alldata$Date,
+    0,
+    NA
   )
-alldata$onmemod <- ifelse(is.na(alldata$me_modstart) & is.na(alldata$onmemod), 0, alldata$onmemod)
-alldata$FEV.1.per.pred_FVC.per.pred <- alldata$FEV1.per.pred / alldata$FVC.per.pred
+)
+alldata$onhemod <- ifelse(
+  is.na(alldata$he_modstart) & is.na(alldata$onhemod),
+  0,
+  alldata$onhemod
+)
+alldata$onmemod <- ifelse(
+  !is.na(alldata$me_modstart) & alldata$me_modstart <= alldata$Date,
+  1,
+  ifelse(
+    !is.na(alldata$me_modstart) & alldata$me_modstart > alldata$Date,
+    0,
+    NA
+  )
+)
+alldata$onmemod <- ifelse(
+  is.na(alldata$me_modstart) & is.na(alldata$onmemod),
+  0,
+  alldata$onmemod
+)
+alldata$FEV.1.per.pred_FVC.per.pred <- alldata$FEV1.per.pred /
+  alldata$FVC.per.pred
 # for analyses about prediction of CFRD, we need to exclude those who only have one OGTT
 # can use alldata for mixed models of PFTs and BMI
-oneogtt <- alldata[alldata$`Date of first OGTT` == alldata$`Date of last OGTT`, ]
+oneogtt <- alldata[
+  alldata$`Date of first OGTT` == alldata$`Date of last OGTT`,
+]
 multogtt <- alldata[!(alldata$MRN %in% oneogtt$MRN), ]
@@
 # for those who develop CFRD, this is time between first OGTT and diagnosis of CFRD
-multogtt$Follow_up_time_yrs <- ifelse(multogtt$CFRD == 1, (multogtt$`Date of CFRD diagnosis` - multogtt$`Date of first OGTT`) / 365.25,
+multogtt$Follow_up_time_yrs <- ifelse(
+  multogtt$CFRD == 1,
+  (multogtt$`Date of CFRD diagnosis` - multogtt$`Date of first OGTT`) / 365.25,
   (multogtt$`Date of last OGTT` - multogtt$`Date of first OGTT`) / 365.25
@@
 # get age at diagnosis of CFRD
-multogtt$age_at_CFRD_dx <- ifelse(multogtt$CFRD == 1, (multogtt$`Date of CFRD diagnosis` - multogtt$DOB) / 365.25, NA)
+multogtt$age_at_CFRD_dx <- ifelse(
+  multogtt$CFRD == 1,
+  (multogtt$`Date of CFRD diagnosis` - multogtt$DOB) / 365.25,
+  NA
+)
 # create categorical time
-multogtt$cat_time <- as.factor(plyr::round_any(multogtt$Follow_up_time_yrs, 0.5))
+multogtt$cat_time <- as.factor(plyr::round_any(
+  multogtt$Follow_up_time_yrs,
+  0.5
+))
 # need a time variable for mixed models
-multogtt$time <- as.numeric((multogtt$Date - multogtt$`Date of first OGTT`) / 365.25)
+multogtt$time <- as.numeric(
+  (multogtt$Date - multogtt$`Date of first OGTT`) / 365.25
+)
 multogtt$cat_time2 <- as.factor(plyr::round_any(multogtt$time, 0.5))
@@
 baseline_all <- merge(baseline_all, evermod, by = "MRN", all.x = T, all.y = T)
-baseline_all$modstarted_fup <- as.factor(ifelse(baseline_all$onmod_baseline == 0 & baseline_all$`max(onmod)` == 1, 1, 0))
+baseline_all$modstarted_fup <- as.factor(ifelse(
+  baseline_all$onmod_baseline == 0 & baseline_all$`max(onmod)` == 1,
+  1,
+  0
+))
 baseline_mult <- merge(baseline_mult, evermod, by = "MRN", all.x = T, all.y = F)
-baseline_mult$modstarted_fup <- as.factor(ifelse(baseline_mult$onmod_baseline == 0 & baseline_mult$`max(onmod)` == 1, 1, 0))
+baseline_mult$modstarted_fup <- as.factor(ifelse(
+  baseline_mult$onmod_baseline == 0 & baseline_mult$`max(onmod)` == 1,
+  1,
+  0
+))
 ################## FILL IN MISSING A1c, BMI, PFT AT BASELINE
@@
 demovars <- c(
-  "Sex", "race_eth", "age", "genoRisk", "Fasting.Glucose", "X1.hour.Glucose", "X2.hour.Glucose", "onmod_baseline", "modstarted_fup",
-  "A1C", "FEV1.per.pred", "FVC.per.pred", "BMI", "bmipercentile", "FEV.1.per.pred_FVC.per.pred"
+  "Sex",
+  "race_eth",
+  "age",
+  "genoRisk",
+  "Fasting.Glucose",
+  "X1.hour.Glucose",
+  "X2.hour.Glucose",
+  "onmod_baseline",
+  "modstarted_fup",
+  "A1C",
+  "FEV1.per.pred",
+  "FVC.per.pred",
+  "BMI",
+  "bmipercentile",
+  "FEV.1.per.pred_FVC.per.pred"
 )
@@
 t1_alldata <- CreateTableOne(data = baseline_all, vars = demovars)
-t1_alldata <- print(t1_alldata, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T, test = F)
+t1_alldata <- print(
+  t1_alldata,
+  printToggle = F,
+  varLabels = TRUE,
+  missing = T,
+  showAllLevels = T,
+  test = F
+)
 # table of demographics in those with at least one pre-CFRD OGTT by sex
-t1_alldata_sex <- CreateTableOne(data = baseline_all, vars = demovars, strata = "Sex")
-t1_alldata_sex <- print(t1_alldata_sex, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T)
+t1_alldata_sex <- CreateTableOne(
+  data = baseline_all,
+  vars = demovars,
+  strata = "Sex"
+)
+t1_alldata_sex <- print(
+  t1_alldata_sex,
+  printToggle = F,
+  varLabels = TRUE,
+  missing = T,
+  showAllLevels = T
+)
 # table of demographics in those with more than 1 OGTT
-t1_multogtt <- CreateTableOne(data = baseline_mult, vars = c(demovars, "Follow_up_time_yrs"))
-t1_multogtt <- print(t1_multogtt, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T, test = F)
+t1_multogtt <- CreateTableOne(
+  data = baseline_mult,
+  vars = c(demovars, "Follow_up_time_yrs")
+)
+t1_multogtt <- print(
+  t1_multogtt,
+  printToggle = F,
+  varLabels = TRUE,
+  missing = T,
+  showAllLevels = T,
+  test = F
+)
 # table of demographics in those with more than 1 OGTT stratified by progression
-t1_multogtt_prog <- CreateTableOne(data = baseline_mult, vars = c(demovars, "Follow_up_time_yrs", "age_at_CFRD_dx"), strata = "CFRD")
-t1_multogtt_prog <- print(t1_multogtt_prog, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T, test = T)
+t1_multogtt_prog <- CreateTableOne(
+  data = baseline_mult,
+  vars = c(demovars, "Follow_up_time_yrs", "age_at_CFRD_dx"),
+  strata = "CFRD"
+)
+t1_multogtt_prog <- print(
+  t1_multogtt_prog,
+  printToggle = F,
+  varLabels = TRUE,
+  missing = T,
+  showAllLevels = T,
+  test = T
+)
 # table of demographics in those with more than 1 OGTT by sex
-t1_multogtt_sex <- CreateTableOne(data = baseline_mult, vars = demovars, strata = "Sex")
-t1_multogtt_sex <- print(t1_multogtt_sex, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T)
+t1_multogtt_sex <- CreateTableOne(
+  data = baseline_mult,
+  vars = demovars,
+  strata = "Sex"
+)
+t1_multogtt_sex <- print(
+  t1_multogtt_sex,
+  printToggle = F,
+  varLabels = TRUE,
+  missing = T,
+  showAllLevels = T
+)
 # compare baseline glucose values in those who developed CFRD and those who did not
@@
 gluctable <- CreateTableOne(baseline_mult, vars = demovars, strata = "CFRD")
-gluctable <- print(gluctable, printToggle = F, varLabels = TRUE, missing = T, showAllLevels = T, test = T)
+gluctable <- print(
+  gluctable,
+  printToggle = F,
+  varLabels = TRUE,
+  missing = T,
+  showAllLevels = T,
+  test = T
+)
 ########################
@@
 # logistic regression of 1 hour and 2 hour glucose predicting cfrd
-log1 <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose, family = "binomial")
+log1 <- glm(
+  data = baseline_mult,
+  CFRD ~ Baseline_1hr_glucose,
+  family = "binomial"
+)
 or1 <- exp(cbind(OR = coef(log1), confint(log1)))
-log1_adj <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose + Base_age + Sex, family = "binomial")
+log1_adj <- glm(
+  data = baseline_mult,
+  CFRD ~ Baseline_1hr_glucose + Base_age + Sex,
+  family = "binomial"
+)
 or1_adj <- exp(cbind(OR = coef(log1_adj), confint(log1_adj)))
 # add interaction with sex
-log1_adj_int <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose + Base_age + Sex + Sex * Baseline_1hr_glucose, family = "binomial")
+log1_adj_int <- glm(
+  data = baseline_mult,
+  CFRD ~ Baseline_1hr_glucose + Base_age + Sex + Sex * Baseline_1hr_glucose,
+  family = "binomial"
+)
 or1_adj_int <- exp(cbind(OR = coef(log1_adj_int), confint(log1_adj_int)))
-log1_cat <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose_155, family = "binomial")
+log1_cat <- glm(
+  data = baseline_mult,
+  CFRD ~ Baseline_1hr_glucose_155,
+  family = "binomial"
+)
 or1_cat <- exp(cbind(OR = coef(log1_cat), confint(log1_cat)))
-log1_cat_adj <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose_155 + Base_age + Sex, family = "binomial")
+log1_cat_adj <- glm(
+  data = baseline_mult,
+  CFRD ~ Baseline_1hr_glucose_155 + Base_age + Sex,
+  family = "binomial"
+)
 or1_cat_adj <- exp(cbind(OR = coef(log1_cat_adj), confint(log1_cat_adj)))
-log1_cat_140 <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose_140, family = "binomial")
+log1_cat_140 <- glm(
+  data = baseline_mult,
+  CFRD ~ Baseline_1hr_glucose_140,
+  family = "binomial"
+)
 or1_cat_140 <- exp(cbind(OR = coef(log1_cat_140), confint(log1_cat_140)))
-log1_cat_adj_140 <- glm(data = baseline_mult, CFRD ~ Baseline_1hr_glucose_140 + Base_age + Sex, family = "binomial")
-or1_cat_adj_140 <- exp(cbind(OR = coef(log1_cat_adj_140), confint(log1_cat_adj_140)))
+log1_cat_adj_140 <- glm(
+  data = baseline_mult,
+  CFRD ~ Baseline_1hr_glucose_140 + Base_age + Sex,
+  family = "binomial"
+)
+or1_cat_adj_140 <- exp(cbind(
+  OR = coef(log1_cat_adj_140),
+  confint(log1_cat_adj_140)
+))
 # add interaction with sex
 log1_cat_adj_int <- glm(
-  data = baseline_mult, CFRD ~ Baseline_1hr_glucose_155 + Base_age + Sex + Sex * Baseline_1hr_glucose_155,
+  data = baseline_mult,
+  CFRD ~
+    Baseline_1hr_glucose_155 + Base_age + Sex + Sex * Baseline_1hr_glucose_155,
   family = "binomial"
 )
-or1_cat_adj_int <- exp(cbind(OR = coef(log1_cat_adj_int), confint(log1_cat_adj_int)))
-log2 <- glm(data = baseline_mult, CFRD ~ Baseline_2hr_glucose, family = "binomial")
+or1_cat_adj_int <- exp(cbind(
+  OR = coef(log1_cat_adj_int),
+  confint(log1_cat_adj_int)
+))
+log2 <- glm(
+  data = baseline_mult,
+  CFRD ~ Baseline_2hr_glucose,
+  family = "binomial"
+)
 or2 <- exp(cbind(OR = coef(log2), confint(log2)))
-log2_adj <- glm(data = baseline_mult, CFRD ~ Baseline_2hr_glucose + Base_age + Sex, family = "binomial")
+log2_adj <- glm(
+  data = baseline_mult,
+  CFRD ~ Baseline_2hr_glucose + Base_age + Sex,
+  family = "binomial"
+)
 or2_adj <- exp(cbind(OR = coef(log2_adj), confint(log2_adj)))
 # add interaction with sex
-log2_adj_int <- glm(data = baseline_mult, CFRD ~ Baseline_2hr_glucose + Base_age + Sex + Sex * Baseline_2hr_glucose, family = "binomial")
+log2_adj_int <- glm(
+  data = baseline_mult,
+  CFRD ~ Baseline_2hr_glucose + Base_age + Sex + Sex * Baseline_2hr_glucose,
+  family = "binomial"
+)
 or2_adj_int <- exp(cbind(OR = coef(log2_adj_int), confint(log2_adj_int)))
@@
 # get predicted values from logistic regression
-baseline_mult$pred_1hr <- predict(log1, type = "response", data.frame = baseline_mult)
+baseline_mult$pred_1hr <- predict(
+  log1,
+  type = "response",
+  data.frame = baseline_mult
+)
 # perform the ROC analysis
@@
 # find the cutoff that maximizes the Youden Index
-coords1 <- coords(roc1, "b", ret = c("t", "specificity", "sensitivity"), best.method = "youden") # default
+coords1 <- coords(
+  roc1,
+  "b",
+  ret = c("t", "specificity", "sensitivity"),
+  best.method = "youden"
+) # default
 # ROC for 2 hour glucose
-baseline_mult$pred_2hr <- predict(log2, type = "response", data.frame = baseline_mult)
+baseline_mult$pred_2hr <- predict(
+  log2,
+  type = "response",
+  data.frame = baseline_mult
+)
 roc2 <- roc(data = baseline_mult, response = CFRD, predictor = X2.hour.Glucose)
-coords2 <- coords(roc2, "b", ret = c("t", "specificity", "sensitivity"), best.method = "youden") # default
+coords2 <- coords(
+  roc2,
+  "b",
+  ret = c("t", "specificity", "sensitivity"),
+  best.method = "youden"
+) # default
 ##############
@@
 # CPH model for baseline 1 hour glucose
-cox1hr <- coxph(Surv(Follow_up_time_yrs, CFRD) ~ Baseline_1hr_glucose, data = baseline_mult)
+cox1hr <- coxph(
+  Surv(Follow_up_time_yrs, CFRD) ~ Baseline_1hr_glucose,
+  data = baseline_mult
+)
 # Diagnostics

### 2025-09-23 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
figure updates for Monique
GCT formatting

@@
 author: "Tim Vigers"
-date: "today"
-date-format: long
+date: "now"
+date-format: "MMMM D, YYYY [at] hh:mm a"
 format:
@@
 library(ggflowchart)
-home_dir <- switch(Sys.info()["sysname"],
+home_dir <- switch(
+  Sys.info()["sysname"],
   "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
@@
 )
-github_dir <- switch(Sys.info()["sysname"],
+github_dir <- switch(
+  Sys.info()["sysname"],
   "Darwin" = "/Users/tim/Documents/GitHub",
@@
 # Import from REDCap
-df <- read.csv("./Christine Chan/CFF GCT Study/Data_Raw/CysticFibrosisScreen_DATA_LABELS_2024-04-19_1740.csv",
+df <- read.csv(
+  "./Christine Chan/CFF GCT Study/Data_Raw/CysticFibrosisScreen_DATA_LABELS_2024-04-19_1740.csv",
   na.strings = ""
@@
 races <- c(
-  "White", "Black/African American", "American Indian/Alaska Native",
-  "Asian Non-Indian", "Asian Indian", "Native Hawaiian/Other Pacific Islander"
+  "White",
+  "Black/African American",
+  "American Indian/Alaska Native",
+  "Asian Non-Indian",
+  "Asian Indian",
+  "Native Hawaiian/Other Pacific Islander"
 )
@@
 df$Race <- apply(
-  df[, grep("Ethnic.Origin.Race", colnames(df))], 1,
+  df[, grep("Ethnic.Origin.Race", colnames(df))],
+  1,
   function(r) {
@@
 df$Ethnicity <- apply(
-  df[, grep("Ethnicity", colnames(df))], 1,
+  df[, grep("Ethnicity", colnames(df))],
+  1,
   function(r) {
@@
     "Pancreatic insufficient" = Pancreatic.exocrine.insufficient,
-    "Weight (kg)" = Body.weight..kg., "Height (cm)" = Height..cm.,
+    "Weight (kg)" = Body.weight..kg.,
+    "Height (cm)" = Height..cm.,
     "Highest FEV1pp" = Highest.FEV1..last.12.months....FEV1.highest....,
-    "Second Highest FEV1pp" =
-      Second.Highest.FEV1..12.months....FEV1.2nd.highest....,
-    "Average of Highest FEV1pp" =
-      Average.of.the.two.highest.FEV1.s...Average.FEV1.......,
+    "Second Highest FEV1pp" = Second.Highest.FEV1..12.months....FEV1.2nd.highest....,
+    "Average of Highest FEV1pp" = Average.of.the.two.highest.FEV1.s...Average.FEV1.......,
   ) %>%
   mutate(
-    "Type 2 diabetes in the family - Father" =
-      factor(Type.2.diabetes.in.the.family...Father),
-    "Type 2 diabetes in the family - Mother" =
-      factor(Type.2.diabetes.in.the.family...Mother),
+    "Type 2 diabetes in the family - Father" = factor(
+      Type.2.diabetes.in.the.family...Father
+    ),
+    "Type 2 diabetes in the family - Mother" = factor(
+      Type.2.diabetes.in.the.family...Mother
+    ),
     "Brothers (number with diabetes)" = factor(Brothers..number.with.diabetes.),
     "Sisters (number with diabetes)" = factor(Sisters..number.with.diabetes.),
-    "Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)" =
-      Number.of.other.relatives.with.diabetes..aunts..uncles..cousins..grandparents.
+    "Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)" = Number.of.other.relatives.with.diabetes..aunts..uncles..cousins..grandparents.
   ) %>%
   select(
-    ID, HbA1c, Age, `Weight (kg)`, `Height (cm)`, BMI,
-    Gender, Race, Ethnicity, `Pancreatic insufficient`,
+    ID,
+    HbA1c,
+    Age,
+    `Weight (kg)`,
+    `Height (cm)`,
+    BMI,
+    Gender,
+    Race,
+    Ethnicity,
+    `Pancreatic insufficient`,
     `Type 2 diabetes in the family - Father`:`Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)`,
-    `Highest FEV1pp`, `Second Highest FEV1pp`, `Average of Highest FEV1pp`,
-    contains("GCT"), contains("OGTT "), "OGTT.Date.Time",
+    `Highest FEV1pp`,
+    `Second Highest FEV1pp`,
+    `Average of Highest FEV1pp`,
+    contains("GCT"),
+    contains("OGTT "),
+    "OGTT.Date.Time",
     "Study.visit.PFTs.Date."
@@
 # Height, weight, BMI percentiles
-df$`Height %ile` <- sds(df$`Height (cm)`,
+df$`Height %ile` <- sds(
+  df$`Height (cm)`,
   age = ifelse(df$Age > 20, 20, df$Age),
-  item = "height2_20", ref = cdc.ref, type = "perc",
-  male = "Male", female = "Female",
+  item = "height2_20",
+  ref = cdc.ref,
+  type = "perc",
+  male = "Male",
+  female = "Female",
   sex = as.character(df$Gender)
 )
-df$`Weight %ile` <- sds(df$`Weight (kg)`,
+df$`Weight %ile` <- sds(
+  df$`Weight (kg)`,
   age = ifelse(df$Age > 20, 20, df$Age),
-  item = "weight2_20", ref = cdc.ref, type = "perc",
-  male = "Male", female = "Female",
+  item = "weight2_20",
+  ref = cdc.ref,
+  type = "perc",
+  male = "Male",
+  female = "Female",
   sex = as.character(df$Gender)
 )
-df$`BMI %ile` <- sds(df$BMI,
+df$`BMI %ile` <- sds(
+  df$BMI,
   age = ifelse(df$Age > 20, 20, df$Age),
-  item = "bmi", ref = cdc.ref, type = "perc",
-  male = "Male", female = "Female",
+  item = "bmi",
+  ref = cdc.ref,
+  type = "perc",
+  male = "Male",
+  female = "Female",
   sex = as.character(df$Gender)
@@
   NA
-df$Diagnosis <- factor(df$Diagnosis,
+df$Diagnosis <- factor(
+  df$Diagnosis,
   levels = c("NGT", "IFG", "INDET", "IGT", "CFRD")
 )
-df$Dysglycemia <- factor(df$Diagnosis,
+df$Dysglycemia <- factor(
+  df$Diagnosis,
   levels = c("NGT", "IFG", "INDET", "IGT", "CFRD"),
@@
 )
-df$Diabetes <- factor(df$Diagnosis,
+df$Diabetes <- factor(
+  df$Diagnosis,
   levels = c("NGT", "IFG", "INDET", "IGT", "CFRD"),
@@
 )
-df$CFRD <- factor(df$`OGTT 120 min` >= 200,
+df$CFRD <- factor(
+  df$`OGTT 120 min` >= 200,
   levels = c(F, T),
@@
 df$`Got OGTT and GCT?` <- apply(
-  df[, c("OGTT 120 min", "GCT 60 min plasma")], 1, function(r) {
+  df[, c("OGTT 120 min", "GCT 60 min plasma")],
+  1,
+  function(r) {
     !any(is.na(r))
@@
 )
-df$`Got OGTT and GCT?` <- factor(df$`Got OGTT and GCT?`,
+df$`Got OGTT and GCT?` <- factor(
+  df$`Got OGTT and GCT?`,
   levels = c(T, F),
@@
 # Write for Christine to check
-write.csv(df,
+write.csv(
+  df,
   file = "./Christine Chan/CFF GCT Study/Data_Clean/gct_analysis_data.csv",
-  row.names = F, na = ""
+  row.names = F,
+  na = ""
 )
@@
 # Basic table formula
-f <- ~ HbA1c + kwt(Age, "N", "Nmiss2", "medianq1q3", "range") +
-  kwt(`Weight (kg)`, "N", "Nmiss2", "medianq1q3", "range") + `Height (cm)` +
-  kwt(BMI, "N", "Nmiss2", "medianq1q3", "range") + Gender + fe(Race) +
-  fe(Ethnicity) + fe(`Pancreatic insufficient`) +
+f <- ~ HbA1c +
+  kwt(Age, "N", "Nmiss2", "medianq1q3", "range") +
+  kwt(`Weight (kg)`, "N", "Nmiss2", "medianq1q3", "range") +
+  `Height (cm)` +
+  kwt(BMI, "N", "Nmiss2", "medianq1q3", "range") +
+  Gender +
+  fe(Race) +
+  fe(Ethnicity) +
+  fe(`Pancreatic insufficient`) +
   fe(`Type 2 diabetes in the family - Father`) +
   fe(`Type 2 diabetes in the family - Mother`) +
-  fe(`Brothers (number with diabetes)`) + fe(`Sisters (number with diabetes)`) +
+  fe(`Brothers (number with diabetes)`) +
+  fe(`Sisters (number with diabetes)`) +
   kwt(
     `Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)`,
-    "N", "Nmiss2", "medianq1q3", "range"
+    "N",
+    "Nmiss2",
+    "medianq1q3",
+    "range"
   ) +
-  `Highest FEV1pp` + `Second Highest FEV1pp` + `Average of Highest FEV1pp` +
+  `Highest FEV1pp` +
+  `Second Highest FEV1pp` +
+  `Average of Highest FEV1pp` +
   kwt(`GCT 0 min plasma`, "N", "Nmiss2", "medianq1q3", "range") +
@@
     `Days From OGTT to GCT (Absolute Value)`,
-    "N", "Nmiss2", "medianq1q3", "range"
+    "N",
+    "Nmiss2",
+    "medianq1q3",
+    "range"
   ) +
-  fe(Diagnosis) + fe(Dysglycemia) + fe(Diabetes) + fe(CFRD)
-t1a <- tableby(update(f, `Got OGTT and GCT?` ~ . - `Got OGTT and GCT?`),
-  data = df, control = mycontrols
+  fe(Diagnosis) +
+  fe(Dysglycemia) +
+  fe(Diabetes) +
+  fe(CFRD)
+t1a <- tableby(
+  update(f, `Got OGTT and GCT?` ~ . - `Got OGTT and GCT?`),
+  data = df,
+  control = mycontrols
 )
@@
 #| message: false
-t1b <- tableby(update(f, Diabetes ~ . - Diabetes),
-  data = df, control = mycontrols
+t1b <- tableby(
+  update(f, Diabetes ~ . - Diabetes),
+  data = df,
+  control = mycontrols
 )
@@
 pairwise.t.test(df$`Average of Highest FEV1pp`, df$Diabetes)
-pairwise.fisher.test(df$`Pancreatic insufficient`, df$Diabetes,
+pairwise.fisher.test(
+  df$`Pancreatic insufficient`,
+  df$Diabetes,
   p.adjust.method = "holm"
@@
 #| message: false
-t1c <- tableby(update(f, Dysglycemia ~ . - Dysglycemia),
-  data = df, control = mycontrols
+t1c <- tableby(
+  update(f, Dysglycemia ~ . - Dysglycemia),
+  data = df,
+  control = mycontrols
 )
@@
 plot_df <- df %>%
-  pivot_longer(c(`OGTT 0 min`, `OGTT 60 min`, `OGTT 120 min`),
-    names_to = "OGTT Timepoint", values_to = "Glucose"
-  )
-plot_df$`OGTT Timepoint` <- factor(plot_df$`OGTT Timepoint`,
+  pivot_longer(
+    c(`OGTT 0 min`, `OGTT 60 min`, `OGTT 120 min`),
+    names_to = "OGTT Timepoint",
+    values_to = "Glucose"
+  )
+plot_df$`OGTT Timepoint` <- factor(
+  plot_df$`OGTT Timepoint`,
   levels = c("OGTT 0 min", "OGTT 60 min", "OGTT 120 min"),
@@
 box_p
-ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/boxplot.png",
+ggsave(
+  "./Christine Chan/CFF GCT Study/Dissemination/Figures/boxplot.png",
   box_p,
@@
 # Dotplot
-dotplot <- ggplot(df, aes(
-  x = Diagnosis, y = `GCT 60 min plasma`,
-  colour = Diagnosis
-)) +
-  geom_point() +
+dotplot <- ggplot(
+  df,
+  aes(
+    x = Diagnosis,
+    y = `GCT 60 min plasma`
+  )
+) +
+  geom_jitter(width = 0.2) +
   geom_hline(yintercept = 130, linetype = "dashed") +
@@
   scale_y_continuous(breaks = c(100, 130, 147, 160)) +
-  scale_color_manual(values = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51"))
+  scale_color_manual(
+    values = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51")
+  )
 dotplot
-ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/dotplot.png",
+ggsave(
+  "./Christine Chan/CFF GCT Study/Dissemination/Figures/dotplot.png",
   dotplot,
-  height = 4, width = 6, units = "in", dpi = 600,
+  height = 4,
+  width = 6,
+  units = "in",
+  dpi = 600,
   device = "png"
@@
     c(
-      `OGTT 0 min`, `OGTT 120 min`, `GCT 0 min plasma`, `GCT 60 min plasma`,
-      `GCT 60 min capillary`
+      `OGTT 0 min`,
+      `OGTT 120 min`,
+      `GCT 60 min plasma`
     ),
-    names_to = "OGTT Timepoint", values_to = "Glucose"
-  )
-plot_df$`OGTT Timepoint` <- factor(plot_df$`OGTT Timepoint`,
+    names_to = "OGTT Timepoint",
+    values_to = "Glucose"
+  )
+plot_df$`OGTT Timepoint` <- factor(
+  plot_df$`OGTT Timepoint`,
   levels = c(
-    "OGTT 0 min", "OGTT 120 min", "GCT 0 min plasma", "GCT 60 min plasma",
-    "GCT 60 min capillary"
+    "OGTT 0 min",
+    "OGTT 120 min",
+    "GCT 60 min plasma"
   ),
-  labels = c("Fasting", "OGTT-2hG", "RPG", "GCTpl", "GCTcap")
+  labels = c("Fasting", "OGTT-2hG", "GCTpl")
 )
@@
     c(
-      `GCT 0 min plasma`, `GCT 60 min plasma`, `GCT 0 min capillary`,
-      `GCT 60 min capillary`, HbA1c
+      `GCT 0 min plasma`,
+      `GCT 60 min plasma`,
+      `GCT 0 min capillary`,
+      `GCT 60 min capillary`,
+      HbA1c
     ),
-    names_to = "Measure", values_to = "Value"
+    names_to = "Measure",
+    values_to = "Value"
   )
@@
     plot.title = element_text(hjust = 0.5),
-    axis.title.x = element_blank(), axis.text.x = element_blank(),
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
     axis.ticks.x = element_blank()
@@
     plot.title = element_text(hjust = 0.5),
-    axis.title.y = element_blank(), axis.text.y = element_blank(),
-    axis.title.x = element_blank(), axis.text.x = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
     axis.ticks.x = element_blank()
@@
   theme(
-    legend.position = "none", plot.title = element_text(hjust = 0.5),
-    axis.title.y = element_blank(), axis.text.y = element_blank(),
-    axis.title.x = element_blank(), axis.text.x = element_blank(),
+    legend.position = "none",
+    plot.title = element_text(hjust = 0.5),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
     axis.ticks.x = element_blank()
@@
     plot.title = element_text(hjust = 0.5),
-    axis.title.y = element_blank(), axis.text.y = element_blank(), 
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
     axis.ticks.y = element_blank(),
-    axis.title.x = element_blank(), axis.text.x = element_blank(),
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
     axis.ticks.x = element_blank()
@@
   theme(
-    legend.title = element_blank(), plot.title = element_text(hjust = 0.5),
+    legend.title = element_blank(),
+    plot.title = element_text(hjust = 0.5),
     axis.text.x = element_blank(),
-    axis.title.x = element_blank(), axis.ticks.x = element_blank()
-  )
-dot_p <- dot_rpg + dot_rcg + dot_gctpl + dot_gctcap + dot_hba1c + 
-  plot_layout(ncol = 5,guides = 'collect')
-ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/dotplot2.png",
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank()
+  )
+dot_p <- dot_rpg +
+  dot_rcg +
+  dot_gctpl +
+  dot_gctcap +
+  dot_hba1c +
+  plot_layout(ncol = 5, guides = "collect")
+ggsave(
+  "./Christine Chan/CFF GCT Study/Dissemination/Figures/dotplot2.png",
   dot_p,
-  height = 9, width = 12, units = "in", dpi = 600,
+  height = 9,
+  width = 12,
+  units = "in",
+  dpi = 600,
   device = "png"
@@
   stat_poly_line() +
-  stat_poly_eq(use_label(c("eq", "R2")),
+  stat_poly_eq(
+    use_label(c("eq", "R2")),
     size = 5,
-    label.y = "bottom", label.x = "right"
+    label.y = "bottom",
+    label.x = "right"
   ) +
@@
 scatter_p
-ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/scatterplot.png",
+ggsave(
+  "./Christine Chan/CFF GCT Study/Dissemination/Figures/scatterplot.png",
   scatter_p,
@@
   lapply(
-    list(gctpl_roc, gctcap_roc, ogtt0_roc, ogtt1_roc, rpg_roc, rcg_roc, a1c_roc),
+    list(
+      gctpl_roc,
+      gctcap_roc,
+      ogtt0_roc,
+      ogtt1_roc,
+      rpg_roc,
+      rcg_roc,
+      a1c_roc
+    ),
     function(roc) {
@@
 roc_table <- do.call(rbind, roc_table)
-rownames(roc_table) <- c("GCTpl", "GCTcap", "OGTT 0 min", "OGTT 1 hr", "RPG", "RCG", "A1c")
+rownames(roc_table) <- c(
+  "GCTpl",
+  "GCTcap",
+  "OGTT 0 min",
+  "OGTT 1 hr",
+  "RPG",
+  "RCG",
+  "A1c"
+)
 gt(roc_table, rownames_to_stub = T) %>% fmt_number(decimals = 3)
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(gctpl_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(gctpl_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/GCTpl_dysglycemia.csv"
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(gctcap_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(gctcap_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/GCTcap_dysglycemia.csv"
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(ogtt0_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(ogtt0_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/OGTT_0_dysglycemia.csv"
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(ogtt1_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(ogtt1_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/OGTT_1_dysglycemia.csv"
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(rpg_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(rpg_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/RPG_dysglycemia.csv"
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(rcg_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(rcg_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/RCG_dysglycemia.csv"
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(a1c_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(a1c_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/A1c_dysglycemia.csv"
@@
 #| message: false
-gctpl_140 <- as.data.frame(ci.coords(gctpl_roc,
+gctpl_140 <- as.data.frame(ci.coords(
+  gctpl_roc,
   x = 140,
@@
 ))
-gctpl_150 <- as.data.frame(ci.coords(gctpl_roc,
+gctpl_150 <- as.data.frame(ci.coords(
+  gctpl_roc,
   x = 150,
@@
 ))
-gctpl_160 <- as.data.frame(ci.coords(gctpl_roc,
+gctpl_160 <- as.data.frame(ci.coords(
+  gctpl_roc,
   x = 160,
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(gctpl_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(gctpl_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/GCTpl_CFRD.csv"
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(gctcap_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(gctcap_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/GCTcap_CFRD.csv"
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(ogtt0_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(ogtt0_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/OGTT_0_CFRD.csv"
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(ogtt1_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(ogtt1_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/OGTT_1_CFRD.csv"
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(rpg_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(rpg_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/RPG_CFRD.csv"
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(rcg_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(rcg_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/RCG_CFRD.csv"
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(a1c_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(a1c_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/A1c_CFRD.csv"
@@
 ```{r}
-df$CFRD_INDET <- factor(df$Diagnosis,
+df$CFRD_INDET <- factor(
+  df$Diagnosis,
   levels = c("NGT", "IFG", "INDET", "IGT", "CFRD"),
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(gctpl_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(gctpl_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/GCTpl_CFRD_INDET.csv"
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(gctcap_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(gctcap_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/GCTcap_CFRD_INDET.csv"
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(ogtt0_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(ogtt0_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/OGTT_0_CFRD_INDET.csv"
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(ogtt1_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(ogtt1_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/OGTT_1_CFRD_INDET.csv"
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(rpg_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(rpg_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/RPG_CFRD_INDET.csv"
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(rcg_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(rcg_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/RCG_CFRD_INDET.csv"
@@
   fmt_number(n_sigfig = 4)
-write.csv(coords(a1c_roc, "all", ret = "all"),
-  row.names = F, na = "",
+write.csv(
+  coords(a1c_roc, "all", ret = "all"),
+  row.names = F,
+  na = "",
   file = "./Christine Chan/CFF GCT Study/Reports/Tables/A1c_CFRD_INDET.csv"

### 2025-10-06 Christine Chan/EnVision CF/fsOGTT_pre_post_ETI.qmd
fsOGTT longitudinal formatting
fsOGTT air formatting again

@@
 # ID linking from Monique
-ids <- read.csv("./Data_Clean/TV_ETI logitudinal patient list.csv", na.strings = "")
+ids <- read.csv(
+    "./Data_Clean/TV_ETI logitudinal patient list.csv",
+    na.strings = ""
+)
 ids <- ids %>%
@@
     select(
-    study_id, gem_id, study, Date, age_visit, sex, origin_race, ethnicity, CFTR,
-    pancreatic_status, corrector___4, pre_post, bmi, fev1,
-    Diagnosis, Glucose_0:Glucose_180, Insulin_0:Insulin_180,
+        study_id,
+        gem_id,
+        study,
+        Date,
+        age_visit,
+        sex,
+        origin_race,
+        ethnicity,
+        CFTR,
+        pancreatic_status,
+        corrector___4,
+        pre_post,
+        bmi,
+        fev1,
+        Diagnosis,
+        Glucose_0:Glucose_180,
+        Insulin_0:Insulin_180,
         C.Peptide_0:C.Peptide_180
@@
     rename(ETI = corrector___4)
-df$race_ethnicity <- factor(df$race_ethnicity,
+df$race_ethnicity <- factor(
+    df$race_ethnicity,
     levels = c(
@@
     labels = c(
-    "Hispanic", "Native American", "Black/African American", "Black/African American",
-    "Hispanic", "Other/Multiple", "Other/Multiple", "Other/Multiple", "Hispanic", "White", "White"
+        "Hispanic",
+        "Native American",
+        "Black/African American",
+        "Black/African American",
+        "Hispanic",
+        "Other/Multiple",
+        "Other/Multiple",
+        "Other/Multiple",
+        "Hispanic",
+        "White",
+        "White"
     )
 )
-df$pancreatic_status <- factor(df$pancreatic_status,
+df$pancreatic_status <- factor(
+    df$pancreatic_status,
     levels = 1:2,
@@
     select(
-    study_id, subject_id, study, date_of_visit, age_at_visit_1, gender,
-    race_ethnicity, cf_genotype, cf_pancreatic, ETI, pre_post, average_bmi,
-    fev1_percent, group_category, lab_ogtt_fasting, min_10_glucose,
-    min_30_glucose, lab_ogtt_1_hour_glucose, min_90_glucose,
-    lab_ogtt_2_hour_glucose, min_150_glucose, min_180_glucose,
-    min_0_insulin, min_10_insulin, min_30_insulin, min_60_insulin,
-    min_90_insulin, min_120_insulin, min_150_insulin, min_180_insulin,
-    min_0_c_peptide, min_10_c_peptide, min_30_c_peptide, min_60_c_peptide,
-    min_90_c_peptide, min_120_c_peptide, min_150_c_peptide, min_180_c_peptide
+        study_id,
+        subject_id,
+        study,
+        date_of_visit,
+        age_at_visit_1,
+        gender,
+        race_ethnicity,
+        cf_genotype,
+        cf_pancreatic,
+        ETI,
+        pre_post,
+        average_bmi,
+        fev1_percent,
+        group_category,
+        lab_ogtt_fasting,
+        min_10_glucose,
+        min_30_glucose,
+        lab_ogtt_1_hour_glucose,
+        min_90_glucose,
+        lab_ogtt_2_hour_glucose,
+        min_150_glucose,
+        min_180_glucose,
+        min_0_insulin,
+        min_10_insulin,
+        min_30_insulin,
+        min_60_insulin,
+        min_90_insulin,
+        min_120_insulin,
+        min_150_insulin,
+        min_180_insulin,
+        min_0_c_peptide,
+        min_10_c_peptide,
+        min_30_c_peptide,
+        min_60_c_peptide,
+        min_90_c_peptide,
+        min_120_c_peptide,
+        min_150_c_peptide,
+        min_180_c_peptide
     ) %>%
     rename(
-    gem_id = subject_id, Date = date_of_visit, age_visit = age_at_visit_1,
-    sex = gender, CFTR = cf_genotype, pancreatic_status = cf_pancreatic,
-    bmi = average_bmi, fev1 = fev1_percent, Diagnosis = group_category,
-    min_0_glucose = lab_ogtt_fasting, min_60_glucose = lab_ogtt_1_hour_glucose,
+        gem_id = subject_id,
+        Date = date_of_visit,
+        age_visit = age_at_visit_1,
+        sex = gender,
+        CFTR = cf_genotype,
+        pancreatic_status = cf_pancreatic,
+        bmi = average_bmi,
+        fev1 = fev1_percent,
+        Diagnosis = group_category,
+        min_0_glucose = lab_ogtt_fasting,
+        min_60_glucose = lab_ogtt_1_hour_glucose,
         min_120_glucose = lab_ogtt_2_hour_glucose
@@
     rename_with(
-    ~ paste0("Glucose_", parse_number(.)), min_0_glucose:min_180_glucose
+        ~ paste0("Glucose_", parse_number(.)),
+        min_0_glucose:min_180_glucose
     ) %>%
     rename_with(
-    ~ paste0("Insulin_", parse_number(.)), min_0_insulin:min_180_insulin
+        ~ paste0("Insulin_", parse_number(.)),
+        min_0_insulin:min_180_insulin
     ) %>%
     rename_with(
-    ~ paste0("C.Peptide_", parse_number(.)), min_0_c_peptide:min_180_c_peptide
+        ~ paste0("C.Peptide_", parse_number(.)),
+        min_0_c_peptide:min_180_c_peptide
     )
@@
 gem$sex <- factor(gem$sex, levels = 1:2, labels = c("Male", "Female"))
-gem$race_ethnicity <- factor(gem$race_ethnicity,
+gem$race_ethnicity <- factor(
+    gem$race_ethnicity,
     levels = 1:6,
     labels = c(
-    "White", "Black/African American", "Asian", "Native American",
-    "Hispanic", "Other/Multiple"
+        "White",
+        "Black/African American",
+        "Asian",
+        "Native American",
+        "Hispanic",
+        "Other/Multiple"
     )
 )
-gem$CFTR <- factor(gem$CFTR,
+gem$CFTR <- factor(
+    gem$CFTR,
     levels = c(
-    "2184delA/M1101K", "2585delT/G551D", "3849+10kb C>T/F508del",
-    "508del/508del", "F508del/1078delA", "f508DEL/1154INStc",
-    "F508del/1154insTC", "F508del/1811 + 1634A > G", "F508del/1811+1634>G",
-    "F508del/2183delAA->G", "F508del/2183delAA>G", "F508del/2184insA",
-    "F508del/2711delT", "F508del/2789+5G>A", "F508del/27895GA",
-    "F508del/3659delC", "F508del/3849+10kbC>T", "F508del/394delTT",
-    "F508del/5492F", "F508del/A561E", "F508del/E60X",
-    "f508del/F508del", "F508del/F508del", "F508del/G542X",
-    "F508del/G551D", "F508del/N1303K", "F508del/Q98R",
-    "F508del/R1162X", "F508del/R117H 9T", "F508del/R117H 9T/5T",
-    "F508del/R553X", "F508del/R709X", "F508del/S492F",
-    "F508del/W1282X", "F5o8del/F508del", "G542X/1078delT",
-    "G542X/W1089X", "G551D/W846X", "I507del/G551D",
-    "N1303K/1154insTC", "N1303K/2789 + 5G > A", "N1303K/2789+5G>A",
-    "R117H 5t/G576A and R668C on exon 13", "R117H5t/G576A and R668C on exon 13",
-    "V520F/3791delC", "W1089X/S492F", "W1090X/S492F"
+        "2184delA/M1101K",
+        "2585delT/G551D",
+        "3849+10kb C>T/F508del",
+        "508del/508del",
+        "F508del/1078delA",
+        "f508DEL/1154INStc",
+        "F508del/1154insTC",
+        "F508del/1811 + 1634A > G",
+        "F508del/1811+1634>G",
+        "F508del/2183delAA->G",
+        "F508del/2183delAA>G",
+        "F508del/2184insA",
+        "F508del/2711delT",
+        "F508del/2789+5G>A",
+        "F508del/27895GA",
+        "F508del/3659delC",
+        "F508del/3849+10kbC>T",
+        "F508del/394delTT",
+        "F508del/5492F",
+        "F508del/A561E",
+        "F508del/E60X",
+        "f508del/F508del",
+        "F508del/F508del",
+        "F508del/G542X",
+        "F508del/G551D",
+        "F508del/N1303K",
+        "F508del/Q98R",
+        "F508del/R1162X",
+        "F508del/R117H 9T",
+        "F508del/R117H 9T/5T",
+        "F508del/R553X",
+        "F508del/R709X",
+        "F508del/S492F",
+        "F508del/W1282X",
+        "F5o8del/F508del",
+        "G542X/1078delT",
+        "G542X/W1089X",
+        "G551D/W846X",
+        "I507del/G551D",
+        "N1303K/1154insTC",
+        "N1303K/2789 + 5G > A",
+        "N1303K/2789+5G>A",
+        "R117H 5t/G576A and R668C on exon 13",
+        "R117H5t/G576A and R668C on exon 13",
+        "V520F/3791delC",
+        "W1089X/S492F",
+        "W1090X/S492F"
     ),
     labels = c(
-    "Other/Other", "Other/Other", "Other/Other", "F508del homozygous",
-    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
-    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
-    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
-    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
-    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
-    "F508del heterozygous", "F508del heterozygous", "F508del homozygous",
-    "F508del homozygous", "F508del heterozygous", "F508del heterozygous",
-    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
-    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
-    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
-    "F508del homozygous", "Other/Other", "Other/Other", "Other/Other",
-    "Other/Other", "Other/Other", "Other/Other", "Other/Other", "Other/Other",
-    "Other/Other", "Other/Other", "Other/Other", "Other/Other"
+        "Other/Other",
+        "Other/Other",
+        "Other/Other",
+        "F508del homozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del homozygous",
+        "F508del homozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del heterozygous",
+        "F508del homozygous",
+        "Other/Other",
+        "Other/Other",
+        "Other/Other",
+        "Other/Other",
+        "Other/Other",
+        "Other/Other",
+        "Other/Other",
+        "Other/Other",
+        "Other/Other",
+        "Other/Other",
+        "Other/Other",
+        "Other/Other"
     )
 )
-gem$pancreatic_status <- factor(gem$pancreatic_status,
+gem$pancreatic_status <- factor(
+    gem$pancreatic_status,
     levels = 1:2,
@@
 gem$ETI <- factor(gem$ETI, levels = 0:1, labels = c("No", "Yes"))
-gem$Diagnosis <- factor(gem$Diagnosis,
+gem$Diagnosis <- factor(
+    gem$Diagnosis,
     levels = 1:4,
@@
 )
-df$Diagnosis <- factor(df$Diagnosis,
+df$Diagnosis <- factor(
+    df$Diagnosis,
     levels = c("NGT", "IFG", "IGT", "INDET", "CFRD"),
@@
 # Write for checking
-write.csv(df,
+write.csv(
+    df,
     file = "./Data_Clean/fsOGTT_pre_post_ETI.csv",
-  row.names = F, na = ""
+    row.names = F,
+    na = ""
 )

### 2025-10-07 Christine Chan/EnVision CF/envision_analyses.qmd
enVision updates on pre/post ETI, repeated measures ROC, and EDA

@@
 df$Diagnosis[df$study_id == "IA0080"] <- "IGT"
+df$Dysglycemia[df$study_id == "WU0019"] <- c("NGT", "NGT")
+df$Dysglycemia[df$study_id == "IA0080"] <- "AGT"
+df$CFRD[df$study_id == "WU0019"] <- c("CFRD-", "CFRD-")
+df$CFRD[df$study_id == "IA0080"] <- "CFRD-"
 # Convert CV units to match Scully et al.
@@
     df,
-    aes(x = Dysglycemia, y = percent_time_over_180, colour = Dysglycemia)
+    aes(
+        x = Dysglycemia,
+        y = percent_time_over_180,
+        colour = Dysglycemia,
+        text = study_id
+    )
 ) +
@@
     df,
-    aes(x = Dysglycemia, y = mean_sensor, colour = Dysglycemia)
+    aes(x = Dysglycemia, y = mean_sensor, colour = Dysglycemia, text = study_id)
 ) +
@@
     df,
-    aes(x = Dysglycemia, y = cv, colour = Dysglycemia)
+    aes(x = Dysglycemia, y = cv, colour = Dysglycemia, text = study_id)
 ) +
@@
     df,
-    aes(x = Dysglycemia, y = a1c_result, colour = Dysglycemia)
+    aes(x = Dysglycemia, y = a1c_result, colour = Dysglycemia, text = study_id)
 ) +
@@
 ggplotly(box_p2, tooltip = c("text"))
+ggplotly(box_p3, tooltip = c("text"))
 ggplotly(box_p4, tooltip = c("text"))
-ggplotly(box_p3, tooltip = c("text"))
 ggplotly(box_p5, tooltip = c("text"))

### 2025-10-07 Christine Chan/EnVision CF/clean_envision_data.R

@@ -652,21 +652,23 @@ final_df$iAUC180ins <- apply(final_df, 1, function(r) {
 label(final_df$iAUC180ins) <- "iAUC 180 Minute Insulin"
 # ODI
 final_df$odi <- final_df$iAUC30ins / final_df$iAUC30gluc
-label(final_df$odi) <- "Oral Disposition Index"
+label(final_df$odi) <- "Oral Disposition Index (Insulin)"
 # HOMA IR
 final_df$homa_ir <- (final_df$Glucose_0 * final_df$Insulin_0) / 405
 # Matsuda
 final_df$matsuda <- 10000 /
-  (((final_df$Glucose_0 / 18) * final_df$Insulin_0) *
-    ((rowMeans(
-      final_df[, paste0("Glucose_", c(0, 30, 60, 90, 120))],
-      na.rm = T
-    ) /
-      18) *
-      rowMeans(
-        final_df[, paste0("Insulin_", c(0, 30, 60, 90, 120))],
+  sqrt(
+    ((final_df$Glucose_0 / 18) * final_df$Insulin_0) *
+      ((rowMeans(
+        final_df[, paste0("Glucose_", c(0, 30, 60, 90, 120))],
         na.rm = T
-      )))
+      ) /
+        18) *
+        rowMeans(
+          final_df[, paste0("Insulin_", c(0, 30, 60, 90, 120))],
+          na.rm = T
+        ))
+  )
 # Check for hypoglycemia
 final_df$Hypo70 <-
   apply(final_df[, glucose], 1, function(r) {

### 2025-10-07 Christine Chan/EnVision CF/clean_fsOGTT_pre_post_ETI.r

@@ -1,41 +1,18 @@
----
-title: "fsOGTT Pre- and Post-ETI"
-author: "Tim Vigers"
-date: "today"
-date-format: long
-format:
-  html:
-    toc: true
-    toc-depth: 5
-    toc-float: true
-    code-fold: true
-    self-contained: true
-    fig-cap-location: top
-    page-layout: full
-editor: source
----
-
-```{r setup}
-#| include: false
 library(tidyverse)
 library(readxl)
-library(lmerTest)
-library(performance)
-library(gtsummary)
-library(emmeans)
-home_dir <- switch(Sys.info()["sysname"],
+library(Hmisc)
+library(pracma)
+home_dir <- switch(
+    Sys.info()["sysname"],
     "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF",
     "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/EnVision CF"
 )
-github_dir <- switch(Sys.info()["sysname"],
+github_dir <- switch(
+    Sys.info()["sysname"],
     "Darwin" = "/Users/timvigers/Documents/GitHub",
     "Linux" = "/home/timvigers/Documents/GitHub"
 )
-knitr::opts_knit$set(root.dir = home_dir)
-```
-
-```{r data cleaning}
-#| message: false
+setwd(home_dir)
 # Import cleaned EnVision data
 df <- read.csv("./Data_Clean/envision_analysis_dataset.csv", na.strings = "")
 # Import GEM data
@@ -84,6 +61,7 @@ df <- df %>%
         CFTR,
         pancreatic_status,
         corrector___4,
+        eti_start,
         pre_post,
         bmi,
         fev1,
@@ -135,6 +113,7 @@ df[, grep("C.Peptide_.*", colnames(df))] <-
         c * 0.003
     })
 # Select the same columns from GEM
+gem$eti_start = NA
 gem <- gem %>%
     mutate(ETI = 0, pre_post = "Pre", study_id = NA, study = "GEM-CF") %>%
     select(
@@ -148,6 +127,7 @@ gem <- gem %>%
         cf_genotype,
         cf_pancreatic,
         ETI,
+        eti_start,
         pre_post,
[…]

### 2025-10-14 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
sweat test formatting

--- 2025-08-12 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
@@
   filter(Age < 50)
+df <- df[-which(duplicated(df$Name)), ]
 # Set table options
@@
 )
+```
+
+### By genotype risk
+
+```{r}
+# Fit a multiclass ROC
+gr_roc <- roc(
+  `Genotype Risk` ~ `Na:Cl`,
+  data = df,
+  levels = c("Mild", "Severe"),
+  quiet = T
+)
+plot(
+  gr_roc,
+  print.auc = T,
+  print.thres = T,
+  main = "Mild vs. Severe Genotype"
+)
+kable(coords(
+  gr_roc,
+  "best",
+  best.method = "youden",
+  ret = c(
+    "threshold",
+    "sensitivity",
+    "specificity",
+    "accuracy",
+    "tpr",
+    "fpr",
+    "tnr",
+    "fnr"
+  )
+))
 ```
@@
 ```
+
+### By genotype risk
+
+```{r}
+# Fit a multiclass ROC
+gr_roc <- roc(
+  `Genotype Risk` ~ `Na:Cl`,
+  data = df,
+  levels = c("Mild", "Severe"),
+  quiet = T
+)
+plot(
+  gr_roc,
+  print.auc = T,
+  print.thres = T,
+  main = "Mild vs. Severe Genotype"
+)
+kable(coords(
+  gr_roc,
+  "best",
+  best.method = "youden",
+  ret = c(
+    "threshold",
+    "sensitivity",
+    "specificity",
+    "accuracy",
+    "tpr",
+    "fpr",
+    "tnr",
+    "fnr"
+  )
+))
+```

### 2025-10-15 Christine Chan/CF Registry/cf_registry_CFRD_prediction.qmd
HbA1c joint model NACFC poster

@@
 #| include: false
+library(tidyverse)
 library(survival)
@@
 library(JM)
-library(dplyr)
-library(tidyr)
 library(arsenal)
 library(knitr)
+library(ggforestplot)
+library(patchwork)
 home_dir <- switch(
@@
   annual,
-  demo %>% select(eDWID, Race, Hispanicrace, Gender, F508, MutClass),
+  demo %>% dplyr::select(eDWID, Race, Hispanicrace, Gender, F508, MutClass),
   by = join_by(eDWID)
@@
   filter(Age_YrEnd <= cfrd_registry_age) %>%
-  select(
+  dplyr::select(
     eDWID,
@@
   slice_max(row_number()) %>%
-  select(eDWID, ReviewYear, Age_YrEnd, all_of(predictors), cfrd_registry) %>%
+  dplyr::select(
+    eDWID,
+    ReviewYear,
+    Age_YrEnd,
+    all_of(predictors),
+    cfrd_registry
+  ) %>%
   drop_na() %>%
@@
 kable(event, caption = "Event Process", digits = 3)
-```
 
-```{r}
-nd <- joint_fit_value$data
-nd$pred <- predict(
-  joint_fit_value,
-  registry[registry$eDWID %in% reg_include, ],
-  type = "Subject",
-  idVar = "eDWID"
+# Plots
+plot_df = lme_fit_a1c$data
+plot_df$pred = predict(lme_fit_a1c)
+p1 = ggplot(plot_df, aes(x = Age_YrEnd, y = pred)) +
+  geom_point(alpha = 0.1, shape = ".") +
+  geom_abline(intercept = 5.61, slope = 0.005895, color = "#3BA3BD") +
+  ylab("HbA1c (%)") +
+  xlab("Age (years)") +
+  ylim(c(2.5, 15)) +
+  ggtitle("Longitudinal Process") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
+
+fp_df = data.frame(s$`CoefTable-Event`) %>%
+  rownames_to_column() %>%
+  filter(!rowname %in% c("(Intercept)", "log(shape)")) %>%
+  rename(estimate = "Value", term = "rowname") %>%
+  mutate(
+    conf.low = estimate - (1.96 * Std.Err),
+    conf.high = estimate + (1.96 * Std.Err)
   )
+fp_df$term = factor(
+  fp_df$term,
+  levels = c(
+    "GenderF",
+    "RaceBlack or African American",
+    "RaceOther race",
+    "RaceMore than one race",
+    "HispanicraceNo",
+    "A_IsOnEnzymesNo",
+    "NumPulmExacerbation1",
+    "NumPulmExacerbation2",
+    "NumPulmExacerbation>=3",
+    "A_FEV1_pct_predicted",
+    "bmi",
+    "Assoct"
+  ),
+  labels = c(
+    "Female",
+    "Black or African American",
+    "Other race",
+    "More than one race",
+    "Hispanic",
+    "Pancreatic sufficient",
+    "1 PEx in past year",
+    "2 PEx in past year",
+    ">= 3 PEx in past year",
+    "FEV1 % predicted",
+    "BMI",
+    "Association Parameter*"
+  )
+)
+fp_df$term = as.character(fp_df$term)
+fp_df$variable = fp_df$p.value <= 0.05
+fp_df$estimate = exp(fp_df$estimate)
+fp_df$conf.low = exp(fp_df$conf.low)
+fp_df$conf.high = exp(fp_df$conf.high)
+p2 = forestplot(
+  df = fp_df,
+  name = term,
+  estimate = estimate,
+  se = Std.Err,
+  xlab = "Hazard ratio for CFRD",
+  title = "Event Process",
+  colour = variable
+) +
+  scale_color_manual(values = c("black", "#3BA3BD")) +
+  theme(legend.position = "none")
+p1 + p2
 ```

### 2025-10-28 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
fixed Youden indices in dot plot

@@
   geom_jitter() +
-  geom_hline(yintercept = 91.5, linetype = "dashed", size = 1) +
+  geom_hline(yintercept = 94, linetype = "dashed", size = 1) +
   ylim(c(50, 300)) +
@@
   geom_jitter() +
-  geom_hline(yintercept = 140.5, linetype = "dashed", size = 1) +
+  geom_hline(yintercept = 147, linetype = "dashed", size = 1) +
   ylim(c(50, 300)) +
@@
   geom_jitter() +
-  geom_hline(yintercept = 97.5, linetype = "dashed", size = 1) +
+  geom_hline(yintercept = 99, linetype = "dashed", size = 1) +
   ylim(c(50, 300)) +
@@
   geom_jitter() +
-  geom_hline(yintercept = 161.5, linetype = "dashed", size = 1) +
+  geom_hline(yintercept = 201, linetype = "dashed", size = 1) +
   ylim(c(50, 300)) +
@@
   geom_jitter(show.legend = F) +
-  geom_hline(aes(yintercept = 5.45, linetype = "Youden Index"), size = 1) +
+  geom_hline(aes(yintercept = 5.5, linetype = "Youden Index"), size = 1) +
   ylab("%") +

### 2025-11-07 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
updated plot for Christine
updates for Christine

@@
 library(ggrepel)
+library(childsds)
+library(labelled)
 home_dir <- switch(
@@
 annual <- left_join(annual, demo %>% select(eDWID, Gender), by = join_by(eDWID))
-# Get BMI group_by
-annual$bmi_cat <- apply(annual, 1, function(r) {
-    age <- as.numeric(r["Age_YrEnd"])
-    gender <- as.character(r["Gender"])
-    bmi <- as.numeric(r["A_bmivalue"])
-    bmi_perc <- as.numeric(r["A_bmipercentile"])
-    if (age <= 20) {
-        return(cut(
-            bmi_perc,
-            c(-Inf, 10, 50, 85, Inf),
-            right = F,
-            labels = c("Undernourished", "At Risk", "Target/Goal", "Overweight")
-        ))
-    } else if (age > 20) {
-        if (gender == "M") {
-            return(cut(
-                bmi,
-                c(-Inf, 18.5, 23, 25, Inf),
-                right = F,
-                labels = c(
-                    "Undernourished",
-                    "At Risk",
-                    "Target/Goal",
-                    "Overweight"
-                )
-            ))
-        } else if (gender == "F") {
-            return(cut(
-                bmi,
-                c(-Inf, 18.5, 22, 25, Inf),
-                right = F,
-                labels = c(
-                    "Undernourished",
-                    "At Risk",
-                    "Target/Goal",
-                    "Overweight"
-                )
-            ))
-        }
-    }
-})
-annual$bmi_cat <- relevel(annual$bmi_cat, ref = "Target/Goal")
 # Merge
@@
       A_IsOnEnzymes,
-            bmi_cat,
       A_FEV1_pct_predicted,
@@
     weight,
-        bmivalue,
-        bmipercentile,
-        bmi_cat,
     A_FEV1_pct_predicted,
@@
   )
+# Calculate BMI percentiles
+roc_df$bmi = roc_df$weight / ((roc_df$height / 100)^2)
+roc_df$bmi_perc = sds(
+  value = roc_df$bmi,
+  age = roc_df$Age_YrEnd,
+  sex = roc_df$Gender,
+  item = "bmi",
+  type = "perc",
+  male = "M",
+  female = "F",
+  ref = cdc.ref
+) *
+  100
+# Raw values for adults
+roc_df$bmi_adult = roc_df$bmi
+roc_df$bmi_adult[roc_df$Age_YrEnd > 20] = NA
+# Get BMI group
+roc_df$bmi_cat <- apply(roc_df, 1, function(r) {
+  age <- as.numeric(r["Age_YrEnd"])
+  gender <- as.character(r["Gender"])
+  bmi <- as.numeric(r["bmi"])
+  bmi_perc <- as.numeric(r["bmi_perc"])
+  if (age <= 20) {
+    return(cut(
+      bmi_perc,
+      c(-Inf, 10, 50, 85, Inf),
+      right = F,
+      labels = c("Undernourished", "At Risk", "Target/Goal", "Overweight")
+    ))
+  } else if (age > 20) {
+    if (gender == "M") {
+      return(cut(
+        bmi,
+        c(-Inf, 18.5, 23, 25, Inf),
+        right = F,
+        labels = c(
+          "Undernourished",
+          "At Risk",
+          "Target/Goal",
+          "Overweight"
+        )
+      ))
+    } else if (gender == "F") {
+      return(cut(
+        bmi,
+        c(-Inf, 18.5, 22, 25, Inf),
+        right = F,
+        labels = c(
+          "Undernourished",
+          "At Risk",
+          "Target/Goal",
+          "Overweight"
+        )
+      ))
+    }
+  }
+})
+roc_df$bmi_cat <- relevel(roc_df$bmi_cat, ref = "Target/Goal")
+# Labels
+var_label(roc_df$bmi) = "BMI (all participants)"
+var_label(roc_df$bmi_perc) = "BMI percentile (age <= 20)"
+var_label(roc_df$bmi_adult) = "BMI value (age > 20)"
+var_label(roc_df$bmi_cat) = "BMI category"
 ```

### 2025-11-07 Christine Chan/EnVision CF/fsOGTT_pre_post_ETI.qmd
updates for Christine

@@
 library(tidyverse)
-library(readxl)
+library(gtsummary)
+library(labelled)
+library(plotly)
 library(lmerTest)
-library(performance)
-library(gtsummary)
-library(emmeans)
-home_dir <- switch(Sys.info()["sysname"],
+home_dir <- switch(
+  Sys.info()["sysname"],
   "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF",
-    "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/EnVision CF"
+  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/EnVision CF",
+  "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/EnVision CF"
 )
-github_dir <- switch(Sys.info()["sysname"],
-    "Darwin" = "/Users/timvigers/Documents/GitHub",
-    "Linux" = "/home/timvigers/Documents/GitHub"
+github_dir <- switch(
+  Sys.info()["sysname"],
+  "Darwin" = "/Users/tim/Documents/GitHub",
+  "Windows" = "C:/Users/Tim/Documents/GitHub",
+  "Linux" = "/home/tim/Documents/GitHub"
 )
@@
 ```{r data cleaning}
-#| message: false
-# Import cleaned EnVision data
-df <- read.csv("./Data_Clean/envision_analysis_dataset.csv", na.strings = "")
-# Import GEM data
-gem <- read.csv(
-    "./Data_Raw/CGMInHealthyControls-FsOGTTDataForMonique_DATA_2025-04-10_0903.csv",
-    na.strings = ""
-)
-# ID linking from Monique
-ids <- read.csv(
-    "./Data_Clean/TV_ETI logitudinal patient list.csv",
-    na.strings = ""
-)
-ids <- ids %>%
-    select(EnVision.ID, GEM.ID) %>%
-    rename(`EnVision ID` = EnVision.ID, `GEM ID` = GEM.ID)
-ids <- ids[!is.na(ids$`GEM ID`), ]
-# Get pre-/post-ETI
-df$eti_start <- ifelse(df$corrector___4 == 1, df$corrector_start_date, NA)
-df <- df %>%
-    group_by(study_id) %>%
-    fill(eti_start, .direction = "downup")
-# Manually correct a few per Monique
-df$eti_start[df$study_id == "CC0001"] <- "2021-11-24"
-df$eti_start[df$study_id == "CC0010"] <- "2023-01-16"
-df$eti_start[df$study_id == "CC0022"] <- "2019-12-06"
-df$eti_start[df$study_id == "CC0023"] <- "2020-01-09"
-df$eti_start[df$study_id == "CC0026"] <- "2020-12-04"
-df$eti_start[df$study_id == "CC0059"] <- "2021-06-23"
-df$pre_post <-
-    factor(df$Date >= df$eti_start, levels = c(F, T), labels = c("Pre", "Post"))
-df$pre_post[df$study_id == "CC0012"] <- "Pre"
-df$pre_post[df$study_id == "CC0017"] <- "Pre"
-df$pre_post[df$study_id == "CC0028"] <- "Pre"
-# Select columns for EnVision per Monique and combine race and ethnicity
-df <- df %>%
-    mutate(gem_id = NA, study = "EnVision CF") %>%
-    select(
-        study_id,
-        gem_id,
-        study,
-        Date,
-        age_visit,
-        sex,
-        origin_race,
-        ethnicity,
-        CFTR,
-        pancreatic_status,
-        corrector___4,
-        pre_post,
-        bmi,
-        fev1,
-        Diagnosis,
-        Glucose_0:Glucose_180,
-        Insulin_0:Insulin_180,
-        C.Peptide_0:C.Peptide_180
-    ) %>%
-    unite(race_ethnicity, c(origin_race, ethnicity)) %>%
-    rename(ETI = corrector___4)
-df$race_ethnicity <- factor(
-    df$race_ethnicity,
-    levels = c(
-        "American Indian or Alaska Native (N. S. & Central America)_Hispanic/Latino",
-        "American Indian or Alaska Native (N. S. & Central America)_Not Hispanic/Latino",
-        "Black or African American (Africa)_N/A",
-        "Black or African American (Africa)_Not Hispanic/Latino",
-        "Declined_Hispanic/Latino",
-        "Declined_Not Hispanic/Latino",
-        "NA_NA",
-        "Native Hawaiian or Other Pacific Islander (Hawaii, Guam, Samoa, Pacific Islands)_Not Hispanic/Latino",
-        "White (Europe, Middle East, North Africa)_Hispanic/Latino",
-        "White (Europe, Middle East, North Africa)_N/A",
-        "White (Europe, Middle East, North Africa)_Not Hispanic/Latino"
-    ),
-    labels = c(
-        "Hispanic",
-        "Native American",
-        "Black/African American",
-        "Black/African American",
-        "Hispanic",
-        "Other/Multiple",
-        "Other/Multiple",
-        "Other/Multiple",
-        "Hispanic",
-        "White",
-        "White"
-    )
-)
-df$pancreatic_status <- factor(
-    df$pancreatic_status,
-    levels = 1:2,
-    labels = c("Sufficient", "Insufficient")
-)
-df$ETI <- factor(df$ETI, levels = 0:1, labels = c("No", "Yes"))
-# Convert C-peptide level from pmol/L to ng/mL (multiply by 0.003)
-df[, grep("C.Peptide_.*", colnames(df))] <-
-    lapply(df[, grep("C.Peptide_.*", colnames(df))], function(c) {
-        c * 0.003
-    })
-# Select the same columns from GEM
-gem <- gem %>%
-    mutate(ETI = 0, pre_post = "Pre", study_id = NA, study = "GEM-CF") %>%
-    select(
-        study_id,
-        subject_id,
-        study,
-        date_of_visit,
-        age_at_visit_1,
-        gender,
-        race_ethnicity,
-        cf_genotype,
-        cf_pancreatic,
-        ETI,
-        pre_post,
-        average_bmi,
-        fev1_percent,
-        group_category,
-        lab_ogtt_fasting,
-        min_10_glucose,
-        min_30_glucose,
-        lab_ogtt_1_hour_glucose,
-        min_90_glucose,
-        lab_ogtt_2_hour_glucose,
-        min_150_glucose,
-        min_180_glucose,
-        min_0_insulin,
-        min_10_insulin,
-        min_30_insulin,
-        min_60_insulin,
-        min_90_insulin,
-        min_120_insulin,
-        min_150_insulin,
-        min_180_insulin,
-        min_0_c_peptide,
-        min_10_c_peptide,
-        min_30_c_peptide,
-        min_60_c_peptide,
-        min_90_c_peptide,
-        min_120_c_peptide,
-        min_150_c_peptide,
-        min_180_c_peptide
-    ) %>%
-    rename(
-        gem_id = subject_id,
-        Date = date_of_visit,
-        age_visit = age_at_visit_1,
-        sex = gender,
-        CFTR = cf_genotype,
-        pancreatic_status = cf_pancreatic,
-        bmi = average_bmi,
-        fev1 = fev1_percent,
-        Diagnosis = group_category,
-        min_0_glucose = lab_ogtt_fasting,
-        min_60_glucose = lab_ogtt_1_hour_glucose,
-        min_120_glucose = lab_ogtt_2_hour_glucose
-    ) %>%
-    rename_with(
-        ~ paste0("Glucose_", parse_number(.)),
-        min_0_glucose:min_180_glucose
-    ) %>%
-    rename_with(
-        ~ paste0("Insulin_", parse_number(.)),
-        min_0_insulin:min_180_insulin
-    ) %>%
-    rename_with(
-        ~ paste0("C.Peptide_", parse_number(.)),
-        min_0_c_peptide:min_180_c_peptide
-    )
-# GEM factors
-gem$sex <- factor(gem$sex, levels = 1:2, labels = c("Male", "Female"))
-gem$race_ethnicity <- factor(
-    gem$race_ethnicity,
-    levels = 1:6,
-    labels = c(
-        "White",
-        "Black/African American",
-        "Asian",
-        "Native American",
-        "Hispanic",
-        "Other/Multiple"
-    )
-)
-gem$CFTR <- factor(
-    gem$CFTR,
-    levels = c(
-        "2184delA/M1101K",
-        "2585delT/G551D",
-        "3849+10kb C>T/F508del",
-        "508del/508del",
-        "F508del/1078delA",
-        "f508DEL/1154INStc",
-        "F508del/1154insTC",
-        "F508del/1811 + 1634A > G",
-        "F508del/1811+1634>G",
-        "F508del/2183delAA->G",
-        "F508del/2183delAA>G",
-        "F508del/2184insA",
-        "F508del/2711delT",
-        "F508del/2789+5G>A",
-        "F508del/27895GA",
-        "F508del/3659delC",
-        "F508del/3849+10kbC>T",
-        "F508del/394delTT",
-        "F508del/5492F",
-        "F508del/A561E",
-        "F508del/E60X",
-        "f508del/F508del",
-        "F508del/F508del",
-        "F508del/G542X",
-        "F508del/G551D",
-        "F508del/N1303K",
-        "F508del/Q98R",
-        "F508del/R1162X",
-        "F508del/R117H 9T",
-        "F508del/R117H 9T/5T",
-        "F508del/R553X",
-        "F508del/R709X",
-        "F508del/S492F",
-        "F508del/W1282X",
-        "F5o8del/F508del",
-        "G542X/1078delT",
-        "G542X/W1089X",
-        "G551D/W846X",
-        "I507del/G551D",
-        "N1303K/1154insTC",
-        "N1303K/2789 + 5G > A",
-        "N1303K/2789+5G>A",
-        "R117H 5t/G576A and R668C on exon 13",
-        "R117H5t/G576A and R668C on exon 13",
-        "V520F/3791delC",
-        "W1089X/S492F",
-        "W1090X/S492F"
-    ),
-    labels = c(
-        "Other/Other",
-        "Other/Other",
-        "Other/Other",
-        "F508del homozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del homozygous",
-        "F508del homozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del heterozygous",
-        "F508del homozygous",
-        "Other/Other",
-        "Other/Other",
-        "Other/Other",
-        "Other/Other",
-        "Other/Other",
-        "Other/Other",
-        "Other/Other",
-        "Other/Other",
-        "Other/Other",
-        "Other/Other",
-        "Other/Other",
-        "Other/Other"
-    )
-)
-gem$pancreatic_status <- factor(
-    gem$pancreatic_status,
-    levels = 1:2,
-    labels = c("Sufficient", "Insufficient")
-)
-gem$ETI <- factor(gem$ETI, levels = 0:1, labels = c("No", "Yes"))
-gem$Diagnosis <- factor(
-    gem$Diagnosis,
-    levels = 1:4,
-    labels = c("Healthy Control", "NGT", "AGT", "CFRD")
-)
-df$Diagnosis <- factor(
-    df$Diagnosis,
-    levels = c("NGT", "IFG", "IGT", "INDET", "CFRD"),
-    labels = c("NGT", "AGT", "AGT", "AGT", "CFRD")
-)
-# IDs
-gem$study_id <- ids$`EnVision ID`[match(gem$gem_id, ids$`GEM ID`)]
-df$gem_id <- ids$`GEM ID`[match(df$study_id, ids$`EnVision ID`)]
-# Limit GEM per Monique's list
-gem <- gem %>% filter(gem_id %in% ids$`GEM ID`)
-# Merge, keep only pre/post, sort by ID and date
-df <- rbind(df, gem) %>% filter(!is.na(pre_post))
-df$pre_post <- factor(df$pre_post, levels = c("Pre", "Post"))
-df <- df %>% arrange(study_id, Date)
-# Calculate ODI, etc.
-# Write for checking
+# Import cleaned data
+load("./Data_Clean/fsOGTT_pre_post_ETI.RData")
+# Subset to people with more than one pre-ETI fsOGTT
+multi_pre_df = df |>
+  group_by(study_id) |>
+  mutate(
+    n_pre = sum(pre_post == "Pre"),
+    time_diff_first_last = as.numeric(difftime(
+      last(Date),
+      first(Date),
+      units = "days"
+    ))
+  ) |>
+  filter(n_pre > 1)
+var_label(multi_pre_df$n_pre) = "Number of fsOGTT pre-ETI"
+var_label(multi_pre_df$time_diff_first_last) = "Days From First to Last fsOGTT"
+# Save for Christine and Monique
 write.csv(
-    df,
+  multi_pre_df,
   file = "./Data_Clean/fsOGTT_pre_post_ETI.csv",
@@
 ```
+
+# Participants with > 1 pre-ETI fsOGTT
+
+## Table 1: Participant characteristics at first fsOGTT
+
+```{r}
+#| label: tbl-1
+multi_pre_df |>
+  group_by(study_id) |>
+  filter(row_number() == 1) |>
+  ungroup() |>
+  select(
+    n_pre,
+    time_diff_first_last,
+    sex,
+    age_visit,
+    bmi,
+    bmi_adult,
+    bmi_perc,
+    fev1,
+    CFTR,
+    pancreatic_status,
+    Diagnosis,
+    Insulin_0,
+    Glucose_0,
+    igi,
+    iAUC30ins_over_gluc,
+    iAUC120ins_over_gluc,
+    iAUC180ins_over_gluc,
+    odi_igi,
+    odi_iauc
+  ) |>
+  tbl_summary(type = c("Insulin_0") ~ "continuous")
+```
+
+## IGI (delta 30min insulin/delta 30min glucose)
+
+```{r}
+#| warning: false
+# Plot
+p = ggplot(
+  multi_pre_df,
+  aes(x = years_from_eti_start, y = igi, group = study_id, color = study_id)
+) +
+  geom_point() +
+  geom_line() +
+  theme_bw()
+ggplotly(p)
+# Model
+m = lmer(
+  igi ~ years_from_eti_start_model + slope_change + (1 | study_id),
+  data = multi_pre_df
+)
+m |>
+  tbl_regression(
+    exponentiate = F,
+    estimate_fun = label_style_sigfig(digits = 4),
+    pvalue_fun = label_style_pvalue(digits = 2),
+  )
+```
+
+## iAUC30ins/iAUC30gluc
+
+```{r}
+#| warning: false
+# Plot
+p = ggplot(
+  multi_pre_df,
+  aes(
+    x = years_from_eti_start,
+    y = iAUC30ins_over_gluc,
+    group = study_id,
+    color = study_id
+  )
+) +
+  geom_point() +
+  geom_line() +
+  theme_bw()
+ggplotly(p)
+# Model
+m = lmer(
+  iAUC30ins_over_gluc ~ years_from_eti_start_model +
+    slope_change +
+    (1 | study_id),
+  data = multi_pre_df
+)
+m |>
+  tbl_regression(
+    exponentiate = F,
+    estimate_fun = label_style_sigfig(digits = 4),
+    pvalue_fun = label_style_pvalue(digits = 2),
+  )
+```
+
+## iAUC120ins/iAUC120gluc
+
+```{r}
+#| warning: false
+# Plot
+p = ggplot(
+  multi_pre_df,
+  aes(
+    x = years_from_eti_start,
+    y = iAUC120ins_over_gluc,
+    group = study_id,
+    color = study_id
+  )
+) +
+  geom_point() +
+  geom_line() +
+  theme_bw()
+ggplotly(p)
+# Model
+m = lmer(
+  iAUC120ins_over_gluc ~ years_from_eti_start_model +
+    slope_change +
+    (1 | study_id),
+  data = multi_pre_df
+)
+m |>
+  tbl_regression(
+    exponentiate = F,
+    estimate_fun = label_style_sigfig(digits = 4),
+    pvalue_fun = label_style_pvalue(digits = 2),
+  )
+```
+
+## iAUC180ins/iAUC180gluc
+
+```{r}
+#| warning: false
+# Plot
+p = ggplot(
+  multi_pre_df,
+  aes(
+    x = years_from_eti_start,
+    y = iAUC180ins_over_gluc,
+    group = study_id,
+    color = study_id
+  )
+) +
+  geom_point() +
+  geom_line() +
+  theme_bw()
+ggplotly(p)
+# Model
+m = lmer(
+  iAUC180ins_over_gluc ~ years_from_eti_start_model +
+    slope_change +
+    (1 | study_id),
+  data = multi_pre_df
+)
+m |>
+  tbl_regression(
+    exponentiate = F,
+    estimate_fun = label_style_sigfig(digits = 4),
+    pvalue_fun = label_style_pvalue(digits = 2),
+  )
+```
+
+## ODI (IGI*Matsuda)
+
+```{r}
+#| warning: false
+# Plot
+p = ggplot(
+  multi_pre_df,
+  aes(x = years_from_eti_start, y = odi_igi, group = study_id, color = study_id)
+) +
+  geom_point() +
+  geom_line() +
+  theme_bw()
+ggplotly(p)
+# Model
+m = lmer(
+  odi_igi ~ years_from_eti_start_model + slope_change + (1 | study_id),
+  data = multi_pre_df
+)
+m |>
+  tbl_regression(
+    exponentiate = F,
+    estimate_fun = label_style_sigfig(digits = 4),
+    pvalue_fun = label_style_pvalue(digits = 2),
+  )
+```
+
+## ODI [(iAUC30ins/iAUC30gluc)*Matsuda]
+
+```{r}
+#| warning: false
+# Plot
+p = ggplot(
+  multi_pre_df,
+  aes(
+    x = years_from_eti_start,
+    y = odi_iauc,
+    group = study_id,
+    color = study_id
+  )
+) +
+  geom_point() +
+  geom_line() +
+  theme_bw()
+ggplotly(p)
+# Model
+m = lmer(
+  odi_iauc ~ years_from_eti_start_model + slope_change + (1 | study_id),
+  data = multi_pre_df
+)
+m |>
+  tbl_regression(
+    exponentiate = F,
+    estimate_fun = label_style_sigfig(digits = 4),
+    pvalue_fun = label_style_pvalue(digits = 2),
+  )
+```

### 2025-11-07 Christine Chan/EnVision CF/clean_fsOGTT_pre_post_ETI.r

@@ -1,7 +1,8 @@
 library(tidyverse)
 library(readxl)
-library(Hmisc)
+library(labelled)
 library(pracma)
+library(childsds)
 home_dir <- switch(
     Sys.info()["sysname"],
     "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF",
@@ -338,7 +339,7 @@ df$iAUC30gluc <- apply(df, 1, function(r) {
         return(auc)
     }
 })
-label(df$iAUC30gluc) <- "iAUC 30 Minute Glucose"
+var_label(df$iAUC30gluc) <- "iAUC 30 Minute Glucose"
 df$iAUC60gluc <- apply(df, 1, function(r) {
     y <- as.numeric(r[glucose[1:4]]) - as.numeric(r[glucose[1]])
     if (is.na(first(y)) | is.na(last(y))) {
@@ -350,7 +351,7 @@ df$iAUC60gluc <- apply(df, 1, function(r) {
         return(auc)
     }
 })
-label(df$iAUC60gluc) <- "iAUC 60 Minute Glucose"
+var_label(df$iAUC60gluc) <- "iAUC 60 Minute Glucose"
 df$iAUC120gluc <- apply(df, 1, function(r) {
     y <- as.numeric(r[glucose[1:6]]) - as.numeric(r[glucose[1]])
     if (is.na(first(y)) | is.na(last(y))) {
@@ -362,7 +363,7 @@ df$iAUC120gluc <- apply(df, 1, function(r) {
         return(auc)
     }
 })
-label(df$iAUC120gluc) <- "iAUC 120 Minute Glucose"
+var_label(df$iAUC120gluc) <- "iAUC 120 Minute Glucose"
 df$iAUC180gluc <- apply(df, 1, function(r) {
     y <- as.numeric(r[glucose]) - as.numeric(r[glucose[1]])
     if (is.na(first(y)) | is.na(last(y))) {
@@ -374,7 +375,7 @@ df$iAUC180gluc <- apply(df, 1, function(r) {
         return(auc)
     }
 })
-label(df$iAUC180gluc) <- "iAUC 180 Minute Glucose"
+var_label(df$iAUC180gluc) <- "iAUC 180 Minute Glucose"
 df$iAUC30ins <- apply(df, 1, function(r) {
     y <- as.numeric(r[insulin[1:3]]) - as.numeric(r[insulin[1]])
[…]

### 2025-11-18 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
cleaned up A1c registry report

--- 2025-09-15 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
@@
 library(ggrepel)
-home_dir <- switch(Sys.info()["sysname"],
+library(childsds)
+library(labelled)
+home_dir <- switch(
+  Sys.info()["sysname"],
   "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
@@
 )
-github_dir <- switch(Sys.info()["sysname"],
+github_dir <- switch(
+  Sys.info()["sysname"],
   "Darwin" = "/Users/timvigers/GitHub",
@@
 )
-annual <- annual %>% rename(reviewyear = ReviewYear)
-annual <- left_join(annual, demo %>% select(eDWID, Gender), by = join_by(eDWID))
-# Get BMI group_by
-annual$bmi_cat <- apply(annual, 1, function(r) {
+annual <- annual |> rename(reviewyear = ReviewYear)
+annual <- left_join(annual, demo |> select(eDWID, Gender), by = join_by(eDWID))
+# Merge
+encounter <- left_join(
+  encounter,
+  demo |> select(eDWID, Race, Hispanicrace, Gender, F508, MutClass),
+  by = join_by(eDWID)
+)
+encounter <- left_join(
+  encounter,
+  annual |>
+    select(
+      eDWID,
+      reviewyear,
+      Age_YrEnd,
+      A_IsOnEnzymes,
+      A_FEV1_pct_predicted,
+      A_FVC_pct_predicted
+    ),
+  by = join_by(eDWID, reviewyear)
+)
+# Format CFTR modulator variables
+encounter$Vx770[is.na(encounter$Vx770)] <- 0
+encounter$Vx770 <- factor(
+  encounter$Vx770,
+  levels = 0:1,
+  labels = c("No", "Yes")
+)
+encounter$Vx809comb[is.na(encounter$Vx809comb)] <- 0
+encounter$Vx809comb <- factor(
+  encounter$Vx809comb,
+  levels = 0:1,
+  labels = c("No", "Yes")
+)
+# Limit to those with two hour OGTT and HbA1c at the same encounter
+# Also remove years before 2010 and those under age 10
+roc_df <- encounter |>
+  arrange(eDWID, reviewyear, encounternum) |>
+  select(
+    eDWID,
+    reviewyear,
+    Age_YrEnd,
+    Race,
+    Hispanicrace,
+    Gender,
+    F508,
+    MutClass,
+    A_IsOnEnzymes,
+    Vx770,
+    Vx809comb,
+    height,
+    weight,
+    A_FEV1_pct_predicted,
+    A_FVC_pct_predicted,
+    hgba1c,
+    fast_bloodglucose,
+    twohour_bloodglucose,
+    Vx770,
+    Vx809comb
+  ) |>
+  filter(
+    reviewyear >= 2010,
+    Age_YrEnd >= 10,
+    hgba1c >= 3.5 & hgba1c <= 15
+  ) |>
+  drop_na(
+    eDWID,
+    reviewyear,
+    Age_YrEnd,
+    Race,
+    Hispanicrace,
+    Gender,
+    F508,
+    MutClass,
+    twohour_bloodglucose,
+    hgba1c
+  ) |>
+  group_by(eDWID) |>
+  filter(row_number() == n()) |>
+  ungroup() |>
+  mutate(
+    `OGTT 2-hour` = cut(
+      twohour_bloodglucose,
+      breaks = c(-Inf, 140, 200, Inf),
+      labels = c("<140", "[140,200)", ">=200")
+    ),
+    `Age Group` = cut(Age_YrEnd, breaks = c(10, 18, 30, Inf), right = F)
+  )
+# Calculate BMI percentiles
+roc_df$bmi = roc_df$weight / ((roc_df$height / 100)^2)
+roc_df$bmi_perc = sds(
+  value = roc_df$bmi,
+  age = roc_df$Age_YrEnd,
+  sex = roc_df$Gender,
+  item = "bmi",
+  type = "perc",
+  male = "M",
+  female = "F",
+  ref = cdc.ref
+) *
+  100
+# Raw values for adults
+roc_df$bmi_adult = roc_df$bmi
+roc_df$bmi_adult[roc_df$Age_YrEnd > 20] = NA
+# Get BMI group
+roc_df$bmi_cat <- apply(roc_df, 1, function(r) {
   age <- as.numeric(r["Age_YrEnd"])
   gender <- as.character(r["Gender"])
-    bmi <- as.numeric(r["A_bmivalue"])
-    bmi_perc <- as.numeric(r["A_bmipercentile"])
+  bmi <- as.numeric(r["bmi"])
+  bmi_perc <- as.numeric(r["bmi_perc"])
   if (age <= 20) {
@@
 })
-annual$bmi_cat <- relevel(annual$bmi_cat, ref = "Target/Goal")
-# Merge
-encounter <- left_join(
-    encounter,
-    demo %>% select(eDWID, Race, Hispanicrace, Gender, F508, MutClass),
-    by = join_by(eDWID)
-)
-encounter <- left_join(
-    encounter,
-    annual %>%
-        select(
-            eDWID,
-            reviewyear,
-            Age_YrEnd,
-            A_IsOnEnzymes,
-            bmi_cat,
-            A_FEV1_pct_predicted,
-            A_FVC_pct_predicted
-        ),
-    by = join_by(eDWID, reviewyear)
-)
-# Limit to those with two hour OGTT and HbA1c at the same encounter
-# Also remove years before 2010 and those under age 10
-roc_df <- encounter %>%
-    arrange(eDWID, reviewyear, encounternum) %>%
-    select(
-        eDWID,
-        reviewyear,
-        Age_YrEnd,
-        Race,
-        Hispanicrace,
-        Gender,
-        F508,
-        MutClass,
-        A_IsOnEnzymes,
-        Vx770,
-        Vx809comb,
-        height,
-        weight,
-        bmivalue,
-        bmipercentile,
-        bmi_cat,
-        A_FEV1_pct_predicted,
-        A_FVC_pct_predicted,
-        hgba1c,
-        fast_bloodglucose,
-        twohour_bloodglucose
-    ) %>%
-    filter(
-        reviewyear >= 2010,
-        Age_YrEnd >= 10,
-        hgba1c >= 3.5 & hgba1c <= 15
-    ) %>%
-    group_by(eDWID) %>%
-    filter(row_number() == n()) %>%
-    ungroup() %>%
-    mutate(
-        `OGTT 2-hour` = cut(
-            twohour_bloodglucose,
-            breaks = c(-Inf, 140, 200, Inf),
-            labels = c("<140", "[140,200)", ">=200")
-        ),
-        `Age Group` = cut(Age_YrEnd, breaks = c(10, 18, 30, Inf), right = F)
-    )
-# Metrics to report
-metrics <- c(
-    "thresh",
-    "specificity",
-    "sensitivity",
-    "accuracy",
-    "tn",
-    "tp",
-    "fn",
-    "fp",
-    "npv",
-    "ppv"
-)
+roc_df$bmi_cat <- relevel(roc_df$bmi_cat, ref = "Target/Goal")
+# Labels
+var_label(roc_df$bmi) = "BMI (all participants)"
+var_label(roc_df$bmi_perc) = "BMI percentile (age <= 20)"
+var_label(roc_df$bmi_adult) = "BMI value (age > 20)"
+var_label(roc_df$bmi_cat) = "BMI category"
 ```
@@
 roc_plot <- function(roc_obj) {
+  a <- ci.auc(roc_obj)
+  a <- round(as.numeric(a[1:3]), 2)
   c <- coords(roc_obj, "best")
@@
     "Sensitivity: ",
-        c[3]
-    )
-    p <- ggplot(plot_df, aes(x = specificity, y = sensitivity)) +
-        scale_x_reverse() +
+    c[3],
+    "\n",
+    "AUC: ",
+    a[2],
+    " (95%CI: ",
+    a[1],
+    ", ",
+    a[3],
+    ")"
+  )
+  p <- ggplot(plot_df, aes(x = 1 - specificity, y = sensitivity)) +
     geom_line() +
@@
     ) +
-        xlab("Specificity") +
+    xlab("1-Specificity") +
     ylab("Sensitivity") +
@@
 ```{r}
-roc_df %>%
-    select(-eDWID) %>%
+roc_df |>
+  select(-eDWID) |>
   tbl_summary(
@@
     )
-    ) %>%
-    add_overall() %>%
+  ) |>
+  add_overall() |>
   add_p(
@@
     test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
-    ) %>%
-    separate_p_footnotes() %>%
+  ) |>
+  separate_p_footnotes() |>
   as_gt()
@@
 
+### Pairwise comparisons
+
+#### <140 vs. [140,200)
+
+```{r}
+roc_df |>
+  select(`OGTT 2-hour`, Gender, MutClass, hgba1c) |>
+  filter(`OGTT 2-hour` != ">=200") |>
+  mutate(`OGTT 2-hour` = droplevels(`OGTT 2-hour`)) |>
+  tbl_summary(
+    by = `OGTT 2-hour`,
+    missing_text = "Missing",
+    type = list(all_continuous() ~ "continuous2"),
+    statistic = list(
+      all_continuous() ~
+        c("{mean} ({sd})", "{median} ({p25}, {p75})", "{min}, {max}")
+    ),
+    label = list(
+      hgba1c = "HbA1c",
+      Hispanicrace = "Hispanic",
+      MutClass = "Mutation Class",
+      reviewyear = "Year",
+      Age_YrEnd = "Age at Year End"
+    )
+  ) |>
+  add_overall() |>
+  add_p(
+    test = list(reviewyear = "fisher.test"),
+    test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
+  ) |>
+  separate_p_footnotes() |>
+  as_gt()
+```
+
+#### [140,200) vs. >=200
+
+```{r}
+roc_df |>
+  select(`OGTT 2-hour`, Gender, MutClass, hgba1c) |>
+  filter(`OGTT 2-hour` != "<140") |>
+  mutate(`OGTT 2-hour` = droplevels(`OGTT 2-hour`)) |>
+  tbl_summary(
+    by = `OGTT 2-hour`,
+    missing_text = "Missing",
+    type = list(all_continuous() ~ "continuous2"),
+    statistic = list(
+      all_continuous() ~
+        c("{mean} ({sd})", "{median} ({p25}, {p75})", "{min}, {max}")
+    ),
+    label = list(
+      hgba1c = "HbA1c",
+      Hispanicrace = "Hispanic",
+      MutClass = "Mutation Class",
+      reviewyear = "Year",
+      Age_YrEnd = "Age at Year End"
+    )
+  ) |>
+  add_overall() |>
+  add_p(
+    test = list(reviewyear = "fisher.test"),
+    test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
+  ) |>
+  separate_p_footnotes() |>
+  as_gt()
+```
+
+#### <140 vs. >=200
+
+```{r}
+roc_df |>
+  select(`OGTT 2-hour`, Gender, MutClass, hgba1c) |>
+  filter(`OGTT 2-hour` != "[140,200)") |>
+  mutate(`OGTT 2-hour` = droplevels(`OGTT 2-hour`)) |>
+  tbl_summary(
+    by = `OGTT 2-hour`,
+    missing_text = "Missing",
+    type = list(all_continuous() ~ "continuous2"),
+    statistic = list(
+      all_continuous() ~
+        c("{mean} ({sd})", "{median} ({p25}, {p75})", "{min}, {max}")
+    ),
+    label = list(
+      hgba1c = "HbA1c",
+      Hispanicrace = "Hispanic",
+      MutClass = "Mutation Class",
+      reviewyear = "Year",
+      Age_YrEnd = "Age at Year End"
+    )
+  ) |>
+  add_overall() |>
+  add_p(
+    test = list(reviewyear = "fisher.test"),
+    test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
+  ) |>
+  separate_p_footnotes() |>
+  as_gt()
+```
+
 ## By age
@@
 ```{r}
-roc_df %>%
+roc_df |>
   select(
@@
     MutClass
-    ) %>%
+  ) |>
   tbl_summary(
@@
 
-## NGT vs. AGT and AGT vs. CFRD
+## NGT vs. (AGT + CFRD) and (NGT + AGT) vs. CFRD
 
@@
 
-1. NGT vs. AGT
-
-2. AGT vs. CFRD
-
-```{r}
-ngt_agt <- roc_df %>% filter(`OGTT 2-hour` != ">=200")
-ngt_agt$ogtt <- droplevels(ngt_agt$`OGTT 2-hour`)
-agt_cfrd <- roc_df %>% filter(`OGTT 2-hour` != "<140")
-agt_cfrd$ogtt <- droplevels(agt_cfrd$`OGTT 2-hour`)
+1. NGT vs. (AGT + CFRD)
+
+2. (NGT + AGT) vs. CFRD
+
+```{r}
+ngt_agt <- roc_df |>
+  mutate(
+    ogtt = factor(
+      `OGTT 2-hour`,
+      levels = c("<140", "[140,200)", ">=200"),
+      labels = c("NGT", "AGT + CFRD", "AGT + CFRD")
+    )
+  )
+agt_cfrd <- roc_df |>
+  mutate(
+    ogtt = factor(
+      `OGTT 2-hour`,
+      levels = c("<140", "[140,200)", ">=200"),
+      labels = c("NGT + AGT", "NGT + AGT", "CFRD")
+    )
+  )
 ```
@@
 
-#### NGT vs. AGT
+#### NGT vs. (AGT + CFRD)
 
@@
 roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-#### AGT vs. CFRD
+gt(data.frame(ci.coords(r, "best"))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
+gt(data.frame(coords(
+  r,
+  "all",
+  ret = c(
+    "threshold",
+    "sensitivity",
+    "specificity",
+    "tp",
+    "tn",
+    "fp",
+    "fn",
+    "npv",
+    "ppv",
+    "tnr",
+    "fnr"
+  )
+))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
+```
+
+#### (NGT + AGT) vs. CFRD
 
@@
 roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(ci.coords(r, "best"))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
+gt(data.frame(coords(
+  r,
+  "all",
+  ret = c(
+    "threshold",
+    "sensitivity",
+    "specificity",
+    "tp",
+    "tn",
+    "fp",
+    "fn",
+    "npv",
+    "ppv",
+    "tnr",
+    "fnr"
+  )
+))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
 ```
@@
 ```{r}
-peds_ngt_agt <- ngt_agt %>% filter(Age_YrEnd < 18)
-peds_agt_cfrd <- agt_cfrd %>% filter(Age_YrEnd < 18)
-```
-
-##### NGT vs. AGT
+peds_ngt_agt <- ngt_agt |> filter(Age_YrEnd < 18)
+peds_agt_cfrd <- agt_cfrd |> filter(Age_YrEnd < 18)
+```
+
+##### NGT vs. (AGT + CFRD)
 
@@
 roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-##### AGT vs. CFRD
+gt(data.frame(ci.coords(r, "best"))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
+gt(data.frame(coords(
+  r,
+  "all",
+  ret = c(
+    "threshold",
+    "sensitivity",
+    "specificity",
+    "tp",
+    "tn",
+    "fp",
+    "fn",
+    "npv",
+    "ppv",
+    "tnr",
+    "fnr"
+  )
+))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
+```
+
+##### (NGT + AGT) vs. CFRD
 
@@
 ```{r}
-adult_ngt_agt <- ngt_agt %>% filter(Age_YrEnd >= 18)
-adult_agt_cfrd <- agt_cfrd %>% filter(Age_YrEnd >= 18)
-```
-
-##### NGT vs. AGT
+adult_ngt_agt <- ngt_agt |> filter(Age_YrEnd >= 18)
+adult_agt_cfrd <- agt_cfrd |> filter(Age_YrEnd >= 18)
+```
+
+##### NGT vs. (AGT + CFRD)
 
@@
 ```{r}
-white_ngt_agt <- ngt_agt %>% filter(Race == "White")
-white_agt_cfrd <- agt_cfrd %>% filter(Race == "White")
-```
-
-##### NGT vs. AGT
+white_ngt_agt <- ngt_agt |> filter(Race == "White")
+white_agt_cfrd <- agt_cfrd |> filter(Race == "White")
+```
+
+##### NGT vs. (AGT + CFRD)
 
@@
 ```{r}
-black_ngt_agt <- ngt_agt %>% filter(Race == "Black or African American")
-black_agt_cfrd <- agt_cfrd %>% filter(Race == "Black or African American")
-```
-
-##### NGT vs. AGT
+black_ngt_agt <- ngt_agt |> filter(Race == "Black or African American")
+black_agt_cfrd <- agt_cfrd |> filter(Race == "Black or African American")
+```
+
+##### NGT vs. (AGT + CFRD)
 
@@
 roc_plot(r)
-gt(data.frame(coords(r, "best"))) %>%
-    fmt_number(n_sigfig = 3)
-gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
-    fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "best"))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
+gt(data.frame(ci.coords(r, "best", best.policy = "random"))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
+gt(data.frame(coords(
+  r,
+  "all",
+  ret = c(
+    "threshold",
+    "sensitivity",
+    "specificity",
+    "tp",
+    "tn",
+    "fp",
+    "fn",
+    "npv",
+    "ppv",
+    "tnr",
+    "fnr"
+  )
+))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
 ```
@@
 ```{r}
-more_ngt_agt <- ngt_agt %>% filter(Race == "More than one race")
-more_agt_cfrd <- agt_cfrd %>% filter(Race == "More than one race")
-```
-
-##### NGT vs. AGT
+more_ngt_agt <- ngt_agt |> filter(Race == "More than one race")
+more_agt_cfrd <- agt_cfrd |> filter(Race == "More than one race")
+```
+
+##### NGT vs. (AGT + CFRD)
+
+Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.
 
@@
 roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-##### AGT vs. CFRD
+gt(data.frame(coords(r, "best"))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
+gt(data.frame(ci.coords(r, "best", best.policy = "random"))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
+gt(data.frame(coords(
+  r,
+  "all",
+  ret = c(
+    "threshold",
+    "sensitivity",
+    "specificity",
+    "tp",
+    "tn",
+    "fp",
+    "fn",
+    "npv",
+    "ppv",
+    "tnr",
+    "fnr"
+  )
+))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
+```
+
+##### (NGT + AGT) vs. CFRD
 
@@
 ```{r}
-other_ngt_agt <- ngt_agt %>% filter(Race == "Other race")
-other_agt_cfrd <- agt_cfrd %>% filter(Race == "Other race")
-```
-
-##### NGT vs. AGT
+other_ngt_agt <- ngt_agt |> filter(Race == "Other race")
+other_agt_cfrd <- agt_cfrd |> filter(Race == "Other race")
+```
+
+##### NGT vs. (AGT + CFRD)
 
@@
 roc_plot(r)
-gt(data.frame(coords(r, "best"))) %>%
-    fmt_number(n_sigfig = 3)
-gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
-    fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
-
-##### AGT vs. CFRD
+gt(data.frame(coords(r, "best"))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
+gt(data.frame(ci.coords(r, "best", best.policy = "random"))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
+gt(data.frame(coords(
+  r,
+  "all",
+  ret = c(
+    "threshold",
+    "sensitivity",
+    "specificity",
+    "tp",
+    "tn",
+    "fp",
+    "fn",
+    "npv",
+    "ppv",
+    "tnr",
+    "fnr"
+  )
+))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
+```
+
+##### (NGT + AGT) vs. CFRD
 
@@
 ```{r}
-hispanic_ngt_agt <- ngt_agt %>% filter(Hispanicrace == "Yes")
-hispanic_agt_cfrd <- agt_cfrd %>% filter(Hispanicrace == "No")
-```
-
-##### NGT vs. AGT
+hispanic_ngt_agt <- ngt_agt |> filter(Hispanicrace == "Yes")
+hispanic_agt_cfrd <- agt_cfrd |> filter(Hispanicrace == "No")
+```
+
+##### NGT vs. (AGT + CFRD)
 
@@
 non_hispanic_ngt_agt <-
-    ngt_agt %>% filter(Race == "Black or African American")
+  ngt_agt |> filter(Race == "Black or African American")
 non_hispanic_agt_cfrd <-
-    agt_cfrd %>% filter(Race == "Black or African American")
-```
-
-##### NGT vs. AGT
+  agt_cfrd |> filter(Race == "Black or African American")
+```
+
+##### NGT vs. (AGT + CFRD)
+
+Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.
 
@@
 ```{r}
-peds_roc <- roc_df %>% filter(Age_YrEnd < 18)
+peds_roc <- roc_df |> filter(Age_YrEnd < 18)
 ```
@@
 ```{r}
-adult_roc <- roc_df %>% filter(Age_YrEnd >= 18)
+adult_roc <- roc_df |> filter(Age_YrEnd >= 18)
 ```
@@
 ```{r}
-white_roc <- roc_df %>% filter(Race == "White")
+white_roc <- roc_df |> filter(Race == "White")
 ```
@@
 ```{r}
-black_roc <- roc_df %>% filter(Race == "Black or African American")
+black_roc <- roc_df |> filter(Race == "Black or African American")
 ```
@@
 roc_plot(r)
-gt(data.frame(ci.coords(r, "best", best.policy = "random"))) %>%
-    fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(ci.coords(r, "best", best.policy = "random"))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
+gt(data.frame(coords(
+  r,
+  "all",
+  ret = c(
+    "threshold",
+    "sensitivity",
+    "specificity",
+    "tp",
+    "tn",
+    "fp",
+    "fn",
+    "npv",
+    "ppv",
+    "tnr",
+    "fnr"
+  )
+))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
 ```
@@
 ```{r}
-more_roc <- roc_df %>% filter(Race == "More than one race")
+more_roc <- roc_df |> filter(Race == "More than one race")
 ```
@@
 ```{r}
-other_roc <- roc_df %>% filter(Race == "Other race")
+other_roc <- roc_df |> filter(Race == "Other race")
 ```
@@
 roc_plot(r)
-gt(data.frame(coords(r, "best"))) %>%
-    fmt_number(n_sigfig = 3)
-gt(data.frame(ci.coords(r, "best"))) %>%
-    fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
+gt(data.frame(coords(r, "best"))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
+gt(data.frame(ci.coords(r, "best", best.policy = "random"))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
+gt(data.frame(coords(
+  r,
+  "all",
+  ret = c(
+    "threshold",
+    "sensitivity",
+    "specificity",
+    "tp",
+    "tn",
+    "fp",
+    "fn",
+    "npv",
+    "ppv",
+    "tnr",
+    "fnr"
+  )
+))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
 ```
@@
 ```{r}
-hispanic_roc <- roc_df %>% filter(Hispanicrace == "Yes")
+hispanic_roc <- roc_df |> filter(Hispanicrace == "Yes")
 ```
@@
 ```{r}
-non_hispanic_roc <- roc_df %>% filter(Hispanicrace == "No")
+non_hispanic_roc <- roc_df |> filter(Hispanicrace == "No")
 ```
@@
 roc_plot(r)
-gt(data.frame(ci.coords(r, "best"))) %>% fmt_number(n_sigfig = 3)
-gt(data.frame(coords(r, "all", ret = "all"))) %>% fmt_number(n_sigfig = 3)
-```
+gt(data.frame(ci.coords(r, "best"))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
+gt(data.frame(coords(
+  r,
+  "all",
+  ret = c(
+    "threshold",
+    "sensitivity",
+    "specificity",
+    "tp",
+    "tn",
+    "fp",
+    "fn",
+    "npv",
+    "ppv",
+    "tnr",
+    "fnr"
+  )
+))) |>
+  fmt_number(decimals = 3, drop_trailing_zeros = T)
+```
+
+# FEV1 over time plot
+
+```{r}
+plot_df <- annual |>
+  filter(
+    !is.na(cfrd_status_annual),
+    Age_YrEnd >= 10,
+    !cfrd_status_annual %in% c(4, 5)
+  ) |>
+  group_by(eDWID) |>
+  mutate(
+    Diagnosis = max(cfrd_status_annual),
+    `Years to CFRD or Last Visit` = case_when(
+      Diagnosis == 3 ~
+        Age_YrEnd -
+        first(Age_YrEnd[cfrd_status_annual == 3]),
+      .default = Age_YrEnd - max(Age_YrEnd, na.rm = T)
+    )
+  )
+plot_df$Diagnosis <- factor(
+  plot_df$Diagnosis,
+  levels = 1:3,
+  labels = c("NGT", "AGT", "CFRD")
+)
+ggplot(
+  plot_df,
+  aes(
+    y = A_FEV1_pct_predicted,
+    x = `Years to CFRD or Last Visit`,
+    color = Diagnosis,
+    group = Diagnosis
+  )
+) +
+  xlim(c(-10, 10)) +
+  geom_point(alpha = 0.1, shape = ".") +
+  geom_smooth(formula = y ~ s(x, bs = "cr")) +
+  theme_bw()
+```

### 2025-11-19 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
started piecewise regression for HbA1c project

@@
 #| include: false
-library(dplyr)
-library(tidyr)
+library(tidyverse)
 library(ggplot2)
@@
     fast_bloodglucose,
-    twohour_bloodglucose,
-    Vx770,
-    Vx809comb
+    twohour_bloodglucose
   ) |>
@@
   100
-# Raw values for adults
-roc_df$bmi_adult = roc_df$bmi
-roc_df$bmi_adult[roc_df$Age_YrEnd > 20] = NA
 # Get BMI group
@@
 roc_df$bmi_cat <- relevel(roc_df$bmi_cat, ref = "Target/Goal")
+# Raw values for adults, percentiles for peds
+roc_df$bmi_adult = roc_df$bmi
+roc_df$bmi_adult[roc_df$Age_YrEnd <= 20] = NA
+roc_df$bmi_perc = roc_df$bmi_perc
+roc_df$bmi_perc[roc_df$Age_YrEnd > 20] = NA
 # Labels
@@
   c <- coords(roc_obj, "best")
-  plot_df <- coords(roc_obj)
-  point_df <- plot_df[plot_df$threshold == as.numeric(c[1]), ]
+  mod_df <- coords(roc_obj)
+  mod_df$one_m_spec = 1 - mod_df$specificity
+  point_df <- mod_df[mod_df$threshold == as.numeric(c[1]), ]
   c <- round(as.numeric(c), 2)
@@
   )
-  p <- ggplot(plot_df, aes(x = 1 - specificity, y = sensitivity)) +
+  p <- ggplot(mod_df, aes(x = one_m_spec, y = sensitivity)) +
     geom_line() +
+    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
     geom_point(data = point_df) +
@@
       aes(label = point_label),
-      nudge_x = 0.25,
+      nudge_x = 0.5,
+      nudge_y = -0.5,
       segment.color = "grey50",
@@
 
-## NGT vs. (AGT + CFRD) and (NGT + AGT) vs. CFRD
+## NGT vs. AGT and AGT vs. CFRD
 
@@
 
-1. NGT vs. (AGT + CFRD)
-
-2. (NGT + AGT) vs. CFRD
-
-```{r}
-ngt_agt <- roc_df |>
-  mutate(
-    ogtt = factor(
-      `OGTT 2-hour`,
-      levels = c("<140", "[140,200)", ">=200"),
-      labels = c("NGT", "AGT + CFRD", "AGT + CFRD")
-    )
-  )
-agt_cfrd <- roc_df |>
-  mutate(
-    ogtt = factor(
-      `OGTT 2-hour`,
-      levels = c("<140", "[140,200)", ">=200"),
-      labels = c("NGT + AGT", "NGT + AGT", "CFRD")
-    )
-  )
+1. NGT vs. AGT
+
+2. AGT vs. CFRD
+
+```{r}
+ngt_agt <- roc_df |> filter(`OGTT 2-hour` != ">=200")
+ngt_agt$ogtt <- droplevels(ngt_agt$`OGTT 2-hour`)
+agt_cfrd <- roc_df |> filter(`OGTT 2-hour` != "<140")
+agt_cfrd$ogtt <- droplevels(agt_cfrd$`OGTT 2-hour`)
 ```
@@
 
-#### NGT vs. (AGT + CFRD)
+#### NGT vs. AGT
 
@@
 
-#### (NGT + AGT) vs. CFRD
+#### AGT vs. CFRD
 
@@
 
-##### NGT vs. (AGT + CFRD)
+##### NGT vs. AGT
 
@@
 
-##### (NGT + AGT) vs. CFRD
+##### AGT vs. CFRD
 
@@
 ##### $\geq$ 200
+Bootstrap replicates found more than one "best" threshold, so the non-bootstrapped threshold is reported first, and a random bootstrapped "best" interval second.
 
@@
 
-# FEV1 over time plot
-
-```{r}
-plot_df <- annual |>
-  filter(
-    !is.na(cfrd_status_annual),
-    Age_YrEnd >= 10,
-    !cfrd_status_annual %in% c(4, 5)
-  ) |>
-  group_by(eDWID) |>
-  mutate(
-    Diagnosis = max(cfrd_status_annual),
-    `Years to CFRD or Last Visit` = case_when(
-      Diagnosis == 3 ~
-        Age_YrEnd -
-        first(Age_YrEnd[cfrd_status_annual == 3]),
-      .default = Age_YrEnd - max(Age_YrEnd, na.rm = T)
-    )
-  )
-plot_df$Diagnosis <- factor(
-  plot_df$Diagnosis,
+# Diagnosis comparisons for HbA1c registry paper
+
+```{r}
+#| warning: false
+#| message: false
+library(Hmisc)
+library(lmerTest)
+# Import data and codebook
+annual <- read.csv(
+  "./Data_Raw/CFF19_Annualized_Del1.csv",
+  na.strings = c("", "N/A")
+)
+data_dict <- readxl::read_excel("./Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
+data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
+# Remove those with T1D or T2D
+annual <- annual %>% arrange(eDWID, ReviewYear)
+annual <- annual[!annual$cfrd_status_annual %in% c(4, 5), ]
+annual$cfrd_status_annual <- factor(
+  annual$cfrd_status_annual,
   levels = 1:3,
-  labels = c("NGT", "AGT", "CFRD")
-)
+  labels = c("CFRD-", "CFRD-", "CFRD+")
+)
+# Get mutations
+demogs <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
+races <- c(
+  "White",
+  "Black or African American",
+  "American Indian or Alaska Native",
+  "Asian",
+  "Native Hawaiian or Other Pacific ",
+  "Some other race"
+)
+demogs$Race <- apply(demogs, 1, function(r) {
+  w <- which(r[paste0("Race", 1:6)] == 1)
+  if (length(w) > 1) {
+    return("More than one race")
+  } else {
+    return(races[w])
+  }
+})
+demogs$Race <- factor(
+  demogs$Race,
+  levels = c(
+    "White",
+    "American Indian or Alaska Native",
+    "Asian",
+    "Black or African American",
+    "More than one race",
+    "Native Hawaiian or Other Pacific ",
+    "Some other race"
+  ),
+  labels = c(
+    "White",
+    "Other race",
+    "Other race",
+    "Black or African American",
+    "More than one race",
+    "Other race",
+    "Other race"
+  )
+)
+demogs$Hispanicrace <- factor(
+  demogs$Hispanicrace,
+  levels = 1:2,
+  labels = c("Yes", "No")
+)
+demogs <- demogs %>% select(eDWID, Gender, MutClass, Race, Hispanicrace)
+annual <- left_join(annual, demogs, by = join_by(eDWID))
+# Variable formatting
+# If A_supplement_feeding is 0, then A_feeding2 is also 0 because they are not
+# on any supplemental feeding
+# Include all kinds of tube feeds
+annual$A_feeding2[annual$A_supplement_feeding == 0] <- 0
+annual$A_feeding3[annual$A_supplement_feeding == 0] <- 0
+annual$A_feeding4[annual$A_supplement_feeding == 0] <- 0
+tube_fields <- c("A_feeding2", "A_feeding3", "A_feeding4")
+annual$tube_feeding <-
+  apply(annual[, tube_fields], 1, function(r) {
+    if (all(is.na(r))) {
+      return(NA)
+    } else {
+      return(max(as.numeric(r), na.rm = T))
+    }
+  })
+# Liver disease
+liver_fields <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
+annual$liver_disease <-
+  apply(annual[, liver_fields], 1, function(r) {
+    if (all(is.na(r))) {
+      return(NA)
+    } else {
+      return(max(as.numeric(r), na.rm = T))
+    }
+  })
+# Get BMI categories
+annual$bmi <- annual$A_weight / ((annual$A_height / 100)^2)
+annual$bmi <- coalesce(annual$A_bmivalue, annual$bmi)
+annual$bmi_perc_cdc <- sds(
+  annual$bmi,
+  age = annual$Age_YrEnd,
+  sex = annual$Gender,
+  male = "M",
+  female = "F",
+  item = "bmi",
+  ref = cdc.ref,
+  type = "perc"
+) *
+  100
+annual$bmi_perc_who <- sds(
+  annual$bmi,
+  age = annual$Age_YrEnd,
+  sex = annual$Gender,
+  male = "M",
+  female = "F",
+  item = "bmi",
+  ref = who.ref,
+  type = "perc"
+) *
+  100
+annual$bmi_perc <- coalesce(annual$bmi_perc_cdc, annual$bmi_perc_who)
+annual$bmi_cat <- apply(annual, 1, function(r) {
+  age <- as.numeric(r["Age_YrEnd"])
+  gender <- as.character(r["Gender"])
+  bmi <- as.numeric(r["bmi"])
+  bmi_perc <- as.numeric(r["bmi_perc"])
+  if (age <= 20) {
+    return(cut(
+      bmi_perc,
+      c(-Inf, 10, 50, 85, Inf),
+      right = F,
+      labels = c("Undernourished", "At Risk", "Target/Goal", "Overweight")
+    ))
+  } else if (age > 20) {
+    if (gender == "M") {
+      return(cut(
+        bmi,
+        c(-Inf, 18.5, 23, 25, Inf),
+        right = F,
+        labels = c(
+          "Undernourished",
+          "At Risk",
+          "Target/Goal",
+          "Overweight"
+        )
+      ))
+    } else if (gender == "F") {
+      return(cut(
+        bmi,
+        c(-Inf, 18.5, 22, 25, Inf),
+        right = F,
+        labels = c(
+          "Undernourished",
+          "At Risk",
+          "Target/Goal",
+          "Overweight"
+        )
+      ))
+    }
+  }
+})
+annual$bmi_cat <- relevel(annual$bmi_cat, ref = "Target/Goal")
+# Raw values for adults, percentiles for peds
+annual$bmi_adult = annual$bmi
+annual$bmi_adult[annual$Age_YrEnd <= 20] = NA
+annual$bmi_perc = annual$bmi_perc
+annual$bmi_perc[annual$Age_YrEnd > 20] = NA
+# Labels
+var_label(annual$bmi) = "BMI (all participants)"
+var_label(annual$bmi_perc) = "BMI percentile (age <= 20)"
+var_label(annual$bmi_adult) = "BMI value (age > 20)"
+var_label(annual$bmi_cat) = "BMI category"
+# Format columns
+annual$MutClass <- factor(
+  annual$MutClass,
+  levels = c("1-3", "4-5", "Oth", "N/A"),
+  labels = c("1-3", "4-5", "Oth", "Unknown")
+)
+annual$NumPulmExacerbation <- cut(
+  annual$NumPulmExacerbation,
+  labels = c("0", "1", "2", ">=3"),
+  breaks = c(0, 1, 2, 3, Inf),
+  right = F
+)
+# Age groups 10 - 18 inclusive, 19 - 30 inclusive, >= 31
+annual$age_group <- cut(
+  annual$Age_YrEnd,
+  breaks = c(-Inf, 10, 19, 31, Inf),
+  right = F
+)
+# Bugs
+bugs <- c(
+  "A_pseudomonasaeruginosa",
+  "A_fungalyeast1",
+  "A_pulmonarycomplications1",
+  "A_mycocultureresults3",
+  "A_cultureresults3"
+)
+# Yes/No variables
+yn_vars <- c(
+  "A_corticosteroids1",
+  "A_feeding2",
+  "A_Vx770",
+  "A_VX809comb",
+  bugs,
+  "tube_feeding",
+  "liver_disease"
+)
+annual[, yn_vars] <- lapply(
+  annual[, yn_vars],
+  factor,
+  levels = 0:1,
+  labels = c("No", "Yes")
+)
+annual$A_IsOnEnzymes <- factor(
+  annual$A_IsOnEnzymes,
+  levels = c(1, 0),
+  labels = c(" Yes", " No")
+)
+# FEV1 to categorical
+annual$ppFEV1_cat <- cut(
+  annual$A_FEV1_pct_predicted,
+  breaks = c(0, 40, 80, 100, Inf),
+  right = F
+)
+annual$ppFEV1_cat <- factor(
+  annual$ppFEV1_cat,
+  levels = c("[100,Inf)", "[80,100)", "[40,80)", "[0,40)")
+)
+# Males to reference group
+annual$Gender <- factor(annual$Gender, levels = c("M", "F"))
+# Hispanic no as reference
+annual$Hispanicrace <- factor(annual$Hispanicrace, levels = c("No", "Yes"))
+# Get first year of CFRD
+first_cfrd <- annual %>%
+  filter(cfrd_status_annual == "CFRD+") %>%
+  group_by(eDWID) %>%
+  summarise(first_cfrd = min(ReviewYear))
+annual <- left_join(annual, first_cfrd, by = join_by(eDWID))
+# Year prior to first CFRD in registry - assume negative for CFRD. Assume all
+# years following are CFRD+
+annual$CFRD <- NA
+annual$CFRD[annual$ReviewYear < annual$first_cfrd] <- "CFRD-"
+annual$CFRD[annual$ReviewYear >= annual$first_cfrd] <- "CFRD+"
+annual$CFRD[is.na(annual$first_cfrd)] <- "CFRD-"
+annual$CFRD <- factor(annual$CFRD, levels = c("CFRD-", "CFRD+"))
+# Add formatted columns for survival analysis
+annual$tstop <- ymd(paste0(annual$ReviewYear, "-12-31"))
+annual$tstart <- ymd(paste0(annual$ReviewYear, "-01-01"))
+annual <- annual %>%
+  group_by(eDWID) %>%
+  mutate(entry_date = tstart[1]) %>%
+  ungroup()
+# Remove those under 10
+annual$age_group <- relevel(annual$age_group, ref = "[10,19)")
+annual <- annual %>% filter(Age_YrEnd != 0, Age_YrEnd >= 10)
+annual$age_group <- droplevels(annual$age_group)
+# List of covariates
+vars <- c(
+  "age_group",
+  "Gender",
+  "bmi_perc",
+  "bmi_cat",
+  "MutClass",
+  "Race",
+  "Hispanicrace",
+  "A_IsOnEnzymes",
+  "ppFEV1_cat",
+  "tube_feeding",
+  "NumPulmExacerbation",
+  "A_pseudomonasaeruginosa",
+  "A_fungalyeast1",
+  "A_mycocultureresults3",
+  "A_corticosteroids1",
+  "cfrd_status_annual",
+  "A_Vx770",
+  "A_VX809comb",
+  "liver_disease"
+)
+# Nice labels for plots and tables
+labels <- list(
+  cfrd_status_annual = "CFRD Status",
+  MutClass = "Mutation class",
+  Race = "Race",
+  Hispanicrace = "Hispanic",
+  Gender = "Sex",
+  A_IsOnEnzymes = "Pancreatic enzyme use",
+  ppFEV1_cat = "FEV1 percent predicted category",
+  A_FEV1_pct_predicted = "FEV1 percent predicted",
+  A_FVC_pct_predicted = "FVC percent predicted",
+  tube_feeding = "Enteric feedings",
+  liver_disease = "Liver disease",
+  A_corticosteroids1 = "Use of corticosteroids",
+  NumPulmExacerbation = "Admissions for PEx",
+  age_group = "Age group",
+  height_perc = "Height percentile",
+  A_Vx770 = "Use of CFTR modulator Vx770",
+  A_VX809comb = "Use of CFTR modulator Vx809 combination",
+  A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
+  A_fungalyeast1 = "Aspergillus (any species)",
+  A_mycocultureresults3 = "Microorganisms in myco culture",
+  A_cultureresults3 = "Microorganisms in culture",
+  eDWID = "ID",
+  tstart = "Interval start",
+  tstop = "Interval stop",
+  Age_YrEnd = "Age"
+)
+label(annual[, which(colnames(annual) %in% names(labels))]) <-
+  labels[colnames(annual)[which(colnames(annual) %in% names(labels))]]
+# Per reviewer, out of X individuals with a diagnosis of CFRD, X number had an
+# OGTT (and X had an A1c, but no OGTT) that year or the year prior.
+# This is super slow but idc idc
+cfrd_only <- apply(first_cfrd, 1, function(r) {
+  id <- as.numeric(r["eDWID"])
+  year <- as.numeric(r["first_cfrd"])
+  d <- annual[annual$eDWID == id & annual$ReviewYear %in% c(year, year - 1), ]
+  a1c <- as.numeric(last(d["A_hgba1c"], na_rm = TRUE))
+  ogtt0hr <- as.numeric(last(d["A_ogttfast_bloodglucose"], na_rm = TRUE))
+  ogtt2hr <- as.numeric(last(d["A_twohour_bloodglucose"], na_rm = TRUE))
+  return(c(id, year, a1c, ogtt0hr, ogtt2hr))
+})
+cfrd_only <- data.frame(t(cfrd_only))
+colnames(cfrd_only) <- c("id", "year", "a1c", "ogtt0hr", "ogtt2hr")
+cfrd_only <- left_join(
+  cfrd_only,
+  annual %>% mutate(eDWID = as.numeric(eDWID)),
+  by = join_by(id == eDWID, year == ReviewYear)
+)
+cfrd_only$a1c_ogtt <- apply(cfrd_only, 1, function(r) {
+  not_missing <- paste(
+    which(!is.na(r[c("a1c", "ogtt0hr", "ogtt2hr")])),
+    collapse = "_"
+  )
+  return(not_missing)
+})
+cfrd_only$a1c_ogtt <- factor(
+  cfrd_only$a1c_ogtt,
+  levels = c("", "1", "1_2", "1_2_3", "1_3", "2", "2_3", "3"),
+  labels = c(
+    "Neither",
+    "HbA1c Only",
+    "OGTT",
+    "OGTT",
+    "OGTT",
+    "OGTT",
+    "OGTT",
+    "OGTT"
+  )
+)
+cfrd_only %>%
+  select(
+    a1c_ogtt,
+    Age_YrEnd,
+    A_weight,
+    A_height,
+    bmi_adult,
+    bmi_perc,
+    bmi_cat,
+    A_FEV1_pct_predicted,
+    A_FVC_pct_predicted,
+    Race,
+    Hispanicrace,
+    MutClass,
+    a1c,
+    A_ogttfast_bloodglucose,
+    A_twohour_bloodglucose,
+    A_IsOnEnzymes,
+    liver_disease,
+    A_pseudomonasaeruginosa,
+    A_fungalyeast1,
+    A_mycocultureresults3,
+    A_cultureresults3
+  ) %>%
+  tbl_summary(
+    by = a1c_ogtt,
+    missing_text = "Missing",
+    statistic = list(all_continuous() ~ "{mean} ({sd})")
+  ) %>%
+  add_overall() %>%
+  add_p(test = list(all_continuous() ~ "oneway.test")) %>%
+  modify_caption("**All Groups**")
+cfrd_only %>%
+  filter(a1c_ogtt %in% c("OGTT", "HbA1c Only")) %>%
+  mutate(a1c_ogtt = droplevels(a1c_ogtt)) %>%
+  select(
+    a1c_ogtt,
+    Age_YrEnd,
+    A_weight,
+    A_height,
+    A_bmivalue,
+    A_bmipercentile,
+    A_FEV1_pct_predicted,
+    A_FVC_pct_predicted,
+    Race,
+    Hispanicrace,
+    MutClass,
+    a1c,
+    A_ogttfast_bloodglucose,
+    A_twohour_bloodglucose,
+    A_IsOnEnzymes,
+    liver_disease,
+    A_pseudomonasaeruginosa,
+    A_fungalyeast1,
+    A_mycocultureresults3,
+    A_cultureresults3
+  ) %>%
+  tbl_summary(
+    by = a1c_ogtt,
+    missing_text = "Missing",
+    statistic = list(all_continuous() ~ "{mean} ({sd})")
+  ) %>%
+  add_overall() %>%
+  add_p(test = list(all_continuous() ~ "oneway.test")) %>%
+  modify_caption("**OGTT vs. HbA1c Only**")
+```
+
+## FEV1 trends by diagnosis method
+
+### LOESS-smoothed plot
+
+```{r}
+#| warning: false
+#| message: false
+# Plot by diagnosis type
+mod_df = cfrd_only %>%
+  select(id, a1c_ogtt) %>%
+  rename(eDWID = id)
+mod_df = left_join(
+  annual %>% mutate(eDWID = as.numeric(eDWID)),
+  mod_df,
+  by = join_by(eDWID)
+)
+mod_df = mod_df %>%
+  group_by(eDWID) %>%
+  mutate(years_to_cfrd = ReviewYear - first_cfrd)
+var_label(mod_df$years_to_cfrd) = "Years to CFRD"
 ggplot(
-  plot_df,
+  mod_df %>% filter(!is.na(a1c_ogtt)),
   aes(
     y = A_FEV1_pct_predicted,
-    x = `Years to CFRD or Last Visit`,
-    color = Diagnosis,
-    group = Diagnosis
+    x = years_to_cfrd,
+    color = a1c_ogtt,
+    group = a1c_ogtt
   )
@@
   geom_smooth(formula = y ~ s(x, bs = "cr")) +
-  theme_bw()
-```
+  theme_bw() +
+  theme(legend.title = element_blank())
+```
+
+### Piecewise linear model results
+
+```{r}
+# Fit model
+mod_df$x = mod_df$years_to_cfrd * (mod_df$years_to_cfrd >= 0)
+label(mod_df$x) = "Change in Slope Post-CFRD"
+label(mod_df$years_to_cfrd) = "Slope Pre-CFRD"
+label(mod_df$a1c_ogtt) = "Diagnosis Method"
+m <- lmer(
+  A_FEV1_pct_predicted ~ years_to_cfrd *
+    a1c_ogtt +
+    x * a1c_ogtt +
+    (1 | eDWID),
+  data = mod_df
+)
+# Population-averaged plot
+
+# Model table
+m |> tbl_regression()
+```

### 2025-11-19 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
started piecewise regression for HbA1c project

--- 2025-09-16 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
@@
 
-# Diagnosis comparisons for HbA1c registry paper
-
-```{r}
-# Per reviewer, out of X individuals with a diagnosis of CFRD, X number had an
-# OGTT (and X had an A1c, but no OGTT) that year or the year prior.
-# This is super slow but idc idc
-cfrd_only <- apply(first_cfrd, 1, function(r) {
-    id <- as.numeric(r["eDWID"])
-    year <- as.numeric(r["first_cfrd"])
-    d <- annual[annual$eDWID == id & annual$ReviewYear %in% c(year, year - 1), ]
-    a1c <- as.numeric(last(d["A_hgba1c"], na_rm = TRUE))
-    ogtt0hr <- as.numeric(last(d["A_ogttfast_bloodglucose"], na_rm = TRUE))
-    ogtt2hr <- as.numeric(last(d["A_twohour_bloodglucose"], na_rm = TRUE))
-    return(c(id, year, a1c, ogtt0hr, ogtt2hr))
-})
-cfrd_only <- data.frame(t(cfrd_only))
-colnames(cfrd_only) <- c("id", "year", "a1c", "ogtt0hr", "ogtt2hr")
-cfrd_only <- left_join(
-    cfrd_only,
-    annual %>% mutate(eDWID = as.numeric(eDWID)),
-    by = join_by(id == eDWID, year == ReviewYear)
-)
-
-cfrd_only$a1c_ogtt <- apply(cfrd_only, 1, function(r) {
-    not_missing <- paste(
-        which(!is.na(r[c("a1c", "ogtt0hr", "ogtt2hr")])),
-        collapse = "_"
-    )
-    return(not_missing)
-})
-cfrd_only$a1c_ogtt <- factor(
-    cfrd_only$a1c_ogtt,
-    levels = c("", "1", "1_2", "1_2_3", "1_3", "2", "2_3", "3"),
-    labels = c(
-        "Neither",
-        "HbA1c Only",
-        "OGTT",
-        "OGTT",
-        "OGTT",
-        "OGTT",
-        "OGTT",
-        "OGTT"
-    )
-)
-cfrd_only %>%
-    select(
-        a1c_ogtt,
-        Age_YrEnd,
-        A_weight,
-        A_height,
-        A_bmivalue,
-        A_bmipercentile,
-        A_FEV1_pct_predicted,
-        A_FVC_pct_predicted,
-        Race,
-        Hispanicrace,
-        MutClass,
-        a1c,
-        A_ogttfast_bloodglucose,
-        A_twohour_bloodglucose,
-        A_IsOnEnzymes,
-        liver_disease,
-        A_cultureresults3
-    ) %>%
-    tbl_summary(
-        by = a1c_ogtt,
-        missing_text = "Missing",
-        statistic = list(all_continuous() ~ "{mean} ({sd})")
-    ) %>%
-    add_overall() %>%
-    add_p(test = list(all_continuous() ~ "oneway.test")) %>%
-    modify_caption("**All Groups**")
-cfrd_only %>%
-    filter(a1c_ogtt %in% c("OGTT", "HbA1c Only")) %>%
-    mutate(a1c_ogtt = droplevels(a1c_ogtt)) %>%
-    select(
-        a1c_ogtt,
-        Age_YrEnd,
-        A_weight,
-        A_height,
-        A_bmivalue,
-        A_bmipercentile,
-        A_FEV1_pct_predicted,
-        A_FVC_pct_predicted,
-        Race,
-        Hispanicrace,
-        MutClass,
-        a1c,
-        A_ogttfast_bloodglucose,
-        A_twohour_bloodglucose,
-        A_IsOnEnzymes,
-        liver_disease,
-        A_cultureresults3
-    ) %>%
-    tbl_summary(
-        by = a1c_ogtt,
-        missing_text = "Missing",
-        statistic = list(all_continuous() ~ "{mean} ({sd})")
-    ) %>%
-    add_overall() %>%
-    add_p(test = list(all_continuous() ~ "oneway.test")) %>%
-    modify_caption("**OGTT vs. HbA1c Only**")
-```
-
 Registry fields included in our liver disease yes/no variable were:

### 2025-11-19 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
piecewise models for HbA1c analysis

@@
 #| include: false
-library(tidyverse)
 library(ggplot2)
@@
 library(labelled)
+library(Hmisc)
+library(lmerTest)
+library(multcomp)
+library(tidyverse)
 home_dir <- switch(
@@
 #| message: false
-library(Hmisc)
-library(lmerTest)
 # Import data and codebook
@@
   group_by(eDWID) %>%
-  summarise(first_cfrd = min(ReviewYear))
+  summarise(first_cfrd = min(ReviewYear), age_dx = min(Age_YrEnd))
 annual <- left_join(annual, first_cfrd, by = join_by(eDWID))
@@
 mod_df = cfrd_only %>%
-  select(id, a1c_ogtt) %>%
+  select(id, a1c_ogtt, Age_YrEnd) %>%
   rename(eDWID = id)
 mod_df = left_join(
-  annual %>% mutate(eDWID = as.numeric(eDWID)),
+  annual %>% select(-Age_YrEnd) |> mutate(eDWID = as.numeric(eDWID)),
   mod_df,
@@
 mod_df = mod_df %>%
+  drop_na(first_cfrd) |>
   group_by(eDWID) %>%
@@
 var_label(mod_df$years_to_cfrd) = "Years to CFRD"
+var_label(mod_df$Age_YrEnd) = "Age at Diagnosis"
 ggplot(
@@
 ) +
+  xlab("Years from CFRD") +
   xlim(c(-10, 10)) +
@@
 
-```{r}
-# Fit model
+There were not enough pancreatic sufficient people in this dataset to include that varaible in the model.
+
+```{r}
+# Fit model - adjust for age at dx, sex, genotype (severe vs mild), pancreatic
+# insufficiency, age at CF diagnosis (do we know this?), number of clinic
+# visits/year, use of CFTR modulators
 mod_df$x = mod_df$years_to_cfrd * (mod_df$years_to_cfrd >= 0)
@@
 label(mod_df$a1c_ogtt) = "Diagnosis Method"
+label(mod_df$Age_YrEnd) = "Age at Diagnosis"
 m <- lmer(
@@
     x * a1c_ogtt +
+    Age_YrEnd +
+    Gender +
+    MutClass +
+    A_Vx770 +
+    A_VX809comb +
     (1 | eDWID),
@@
 # Population-averaged plot
-
+pop_dat <- data.frame(expand.grid(
+  years_to_cfrd = -10:10,
+  a1c_ogtt = unique(mod_df$a1c_ogtt),
+  Age_YrEnd = mean(mod_df$Age_YrEnd, na.rm = T),
+  Gender = "M",
+  MutClass = "1-3",
+  A_Vx770 = "No",
+  A_VX809comb = "No"
+))
+pop_dat$x <- (pop_dat$years_to_cfrd >= 0) * pop_dat$years_to_cfrd
+pred <- predict(m, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(
+  pop_dat,
+  aes(x = years_to_cfrd, y = pred, group = a1c_ogtt, colour = a1c_ogtt)
+) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
+  xlab("Years from CFRD") +
+  ylab("FEV1 % Predicted") +
+  theme_classic(base_size = 15)
 # Model table
 m |> tbl_regression()
-```
+# Contrasts
+neither_pre <- rep(0, nrow(summary(m)$coefficients))
+neither_change <- rep(0, nrow(summary(m)$coefficients))
+neither_pre[which(rownames(summary(m)$coefficients) == "years_to_cfrd")] = 1
+neither_change[which(rownames(summary(m)$coefficients) == "x")] = 1
+a1c_pre <- rep(0, nrow(summary(m)$coefficients))
+a1c_change <- rep(0, nrow(summary(m)$coefficients))
+a1c_pre[which(
+  rownames(summary(m)$coefficients) %in%
+    c("years_to_cfrd", "years_to_cfrd:a1c_ogttHbA1c Only")
+)] = 1
+a1c_change[which(
+  rownames(summary(m)$coefficients) %in%
+    c("x", "a1c_ogttHbA1c Only:x")
+)] = 1
+ogtt_pre <- rep(0, nrow(summary(m)$coefficients))
+ogtt_change <- rep(0, nrow(summary(m)$coefficients))
+ogtt_pre[which(
+  rownames(summary(m)$coefficients) %in%
+    c("years_to_cfrd", "years_to_cfrd:a1c_ogttOGTT")
+)] = 1
+ogtt_change[which(
+  rownames(summary(m)$coefficients) %in%
+    c("x", "a1c_ogttOGTT:x")
+)] = 1
+contr <- rbind(
+  "HbA1c Only Pre-CFRD" = a1c_pre,
+  "HbA1c Only Post-CFRD" = a1c_pre + a1c_change,
+  "OGTT Pre-CFRD" = ogtt_pre,
+  "OGTT Post-CFRD" = ogtt_pre + ogtt_change,
+  "Neither Pre-CFRD" = neither_pre,
+  "Neither Post-CFRD" = neither_pre + neither_change
+)
+change_mod_slopes <- glht(m, linfct = contr)
+change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
+change_mod_slopes_ci$confint %>%
+  as.data.frame() %>%
+  gt(caption = "Slopes by Diagnosis Method", rownames_to_stub = T) %>%
+  fmt_number(decimals = 2)
+```
+
+## FVC trends by diagnosis method
+
+### LOESS-smoothed plot
+
+```{r} 
+ggplot(
+  mod_df %>% filter(!is.na(a1c_ogtt)),
+  aes(
+    y = A_FVC_pct_predicted,
+    x = years_to_cfrd,
+    color = a1c_ogtt,
+    group = a1c_ogtt
+  )
+) +
+  xlab("Years from CFRD") +
+  xlim(c(-10, 10)) +
+  geom_point(alpha = 0.1, shape = ".") +
+  geom_smooth(formula = y ~ s(x, bs = "cr")) +
+  theme_bw() +
+  theme(legend.title = element_blank())
+```
+
+### Piecewise linear model results
+
+There were not enough pancreatic sufficient people in this dataset to include that varaible in the model.
+
+```{r}
+# Fit model - adjust for age at dx, sex, genotype (severe vs mild), pancreatic
+# insufficiency, age at CF diagnosis (do we know this?), number of clinic
+# visits/year, use of CFTR modulators
+mod_df$x = mod_df$years_to_cfrd * (mod_df$years_to_cfrd >= 0)
+label(mod_df$x) = "Change in Slope Post-CFRD"
+label(mod_df$years_to_cfrd) = "Slope Pre-CFRD"
+label(mod_df$a1c_ogtt) = "Diagnosis Method"
+label(mod_df$Age_YrEnd) = "Age at Diagnosis"
+m <- lmer(
+  A_FVC_pct_predicted ~ years_to_cfrd *
+    a1c_ogtt +
+    x * a1c_ogtt +
+    Age_YrEnd +
+    Gender +
+    MutClass +
+    A_Vx770 +
+    A_VX809comb +
+    (1 | eDWID),
+  data = mod_df
+)
+# Population-averaged plot
+pop_dat <- data.frame(expand.grid(
+  years_to_cfrd = -10:10,
+  a1c_ogtt = unique(mod_df$a1c_ogtt),
+  Age_YrEnd = mean(mod_df$Age_YrEnd, na.rm = T),
+  Gender = "M",
+  MutClass = "1-3",
+  A_Vx770 = "No",
+  A_VX809comb = "No"
+))
+pop_dat$x <- (pop_dat$years_to_cfrd >= 0) * pop_dat$years_to_cfrd
+pred <- predict(m, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(
+  pop_dat,
+  aes(x = years_to_cfrd, y = pred, group = a1c_ogtt, colour = a1c_ogtt)
+) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
+  xlab("Years from CFRD") +
+  ylab("FVC % Predicted") +
+  theme_classic(base_size = 15)
+# Model table
+m |> tbl_regression()
+# Contrasts
+neither_pre <- rep(0, nrow(summary(m)$coefficients))
+neither_change <- rep(0, nrow(summary(m)$coefficients))
+neither_pre[which(rownames(summary(m)$coefficients) == "years_to_cfrd")] = 1
+neither_change[which(rownames(summary(m)$coefficients) == "x")] = 1
+a1c_pre <- rep(0, nrow(summary(m)$coefficients))
+a1c_change <- rep(0, nrow(summary(m)$coefficients))
+a1c_pre[which(
+  rownames(summary(m)$coefficients) %in%
+    c("years_to_cfrd", "years_to_cfrd:a1c_ogttHbA1c Only")
+)] = 1
+a1c_change[which(
+  rownames(summary(m)$coefficients) %in%
+    c("x", "a1c_ogttHbA1c Only:x")
+)] = 1
+ogtt_pre <- rep(0, nrow(summary(m)$coefficients))
+ogtt_change <- rep(0, nrow(summary(m)$coefficients))
+ogtt_pre[which(
+  rownames(summary(m)$coefficients) %in%
+    c("years_to_cfrd", "years_to_cfrd:a1c_ogttOGTT")
+)] = 1
+ogtt_change[which(
+  rownames(summary(m)$coefficients) %in%
+    c("x", "a1c_ogttOGTT:x")
+)] = 1
+contr <- rbind(
+  "HbA1c Only Pre-CFRD" = a1c_pre,
+  "HbA1c Only Post-CFRD" = a1c_pre + a1c_change,
+  "OGTT Pre-CFRD" = ogtt_pre,
+  "OGTT Post-CFRD" = ogtt_pre + ogtt_change,
+  "Neither Pre-CFRD" = neither_pre,
+  "Neither Post-CFRD" = neither_pre + neither_change
+)
+change_mod_slopes <- glht(m, linfct = contr)
+change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
+change_mod_slopes_ci$confint %>%
+  as.data.frame() %>%
+  gt(caption = "Slopes by Diagnosis Method", rownames_to_stub = T) %>%
+  fmt_number(decimals = 2)
+```

### 2025-11-20 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
piecewise lung function models rough draft done

--- 2025-11-19 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
@@
   levels = c(1, 0),
-  labels = c(" Yes", " No")
+  labels = c("Yes", " No")
 )
@@
 var_label(mod_df$years_to_cfrd) = "Years to CFRD"
+var_label(mod_df$Age_YrEnd) = "Age at Diagnosis"
 ggplot(
-  mod_df %>% filter(!is.na(a1c_ogtt)),
+  mod_df |> filter(!is.na(a1c_ogtt)),
   aes(
@@
 ) +
-  xlim(c(-10, 10)) +
+  xlab("Years from CFRD") +
+  xlim(c(-5, 5)) +
   geom_point(alpha = 0.1, shape = ".") +
@@
 ```{r}
-# Fit model
+# Fit model - adjust for age at dx, sex, genotype (severe vs mild), pancreatic
+# insufficiency, age at CF diagnosis (do we know this?), number of clinic
+# visits/year, use of CFTR modulators
 mod_df$x = mod_df$years_to_cfrd * (mod_df$years_to_cfrd >= 0)
@@
 label(mod_df$a1c_ogtt) = "Diagnosis Method"
+label(mod_df$Age_YrEnd) = "Age at Diagnosis"
+# Write data so Christine can see missing data
+t = mod_df |>
+  select(
+    eDWID,
+    years_to_cfrd,
+    A_FEV1_pct_predicted,
+    A_FVC_pct_predicted,
+    a1c_ogtt,
+    Age_YrEnd,
+    Gender,
+    MutClass,
+    A_IsOnEnzymes,
+    A_Vx770,
+    A_VX809comb
+  ) |>
+  drop_na(A_FEV1_pct_predicted)
+write.csv(t, "./Data_Clean/fev1_model_data.csv", row.names = F, na = "")
+# Fit the model
 m <- lmer(
@@
     x * a1c_ogtt +
+    Age_YrEnd +
+    Gender +
+    MutClass +
+    A_IsOnEnzymes +
+    A_Vx770 +
+    A_VX809comb +
     (1 | eDWID),
@@
 # Population-averaged plot
-
+pop_dat <- data.frame(expand.grid(
+  years_to_cfrd = -5:5,
+  a1c_ogtt = unique(mod_df$a1c_ogtt),
+  Age_YrEnd = mean(mod_df$Age_YrEnd, na.rm = T),
+  Gender = "M",
+  MutClass = "1-3",
+  A_IsOnEnzymes = "Yes",
+  A_Vx770 = "No",
+  A_VX809comb = "No"
+))
+pop_dat$x <- (pop_dat$years_to_cfrd >= 0) * pop_dat$years_to_cfrd
+pred <- predict(m, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(
+  pop_dat,
+  aes(x = years_to_cfrd, y = pred, group = a1c_ogtt, colour = a1c_ogtt)
+) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
+  xlab("Years from CFRD") +
+  ylab("FEV1 % Predicted") +
+  theme_classic(base_size = 15)
+# Contrasts
+neither_pre <- rep(0, nrow(summary(m)$coefficients))
+neither_change <- rep(0, nrow(summary(m)$coefficients))
+neither_pre[which(rownames(summary(m)$coefficients) == "years_to_cfrd")] = 1
+neither_change[which(rownames(summary(m)$coefficients) == "x")] = 1
+a1c_pre <- rep(0, nrow(summary(m)$coefficients))
+a1c_change <- rep(0, nrow(summary(m)$coefficients))
+a1c_pre[which(
+  rownames(summary(m)$coefficients) %in%
+    c("years_to_cfrd", "years_to_cfrd:a1c_ogttHbA1c Only")
+)] = 1
+a1c_change[which(
+  rownames(summary(m)$coefficients) %in%
+    c("x", "a1c_ogttHbA1c Only:x")
+)] = 1
+ogtt_pre <- rep(0, nrow(summary(m)$coefficients))
+ogtt_change <- rep(0, nrow(summary(m)$coefficients))
+ogtt_pre[which(
+  rownames(summary(m)$coefficients) %in%
+    c("years_to_cfrd", "years_to_cfrd:a1c_ogttOGTT")
+)] = 1
+ogtt_change[which(
+  rownames(summary(m)$coefficients) %in%
+    c("x", "a1c_ogttOGTT:x")
+)] = 1
+contr <- rbind(
+  "HbA1c Only Pre-CFRD" = a1c_pre,
+  "HbA1c Only Post-CFRD" = a1c_pre + a1c_change,
+  "OGTT Pre-CFRD" = ogtt_pre,
+  "OGTT Post-CFRD" = ogtt_pre + ogtt_change,
+  "Neither Pre-CFRD" = neither_pre,
+  "Neither Post-CFRD" = neither_pre + neither_change
+)
+change_mod_slopes <- glht(m, linfct = contr)
+change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
+change_mod_slopes_ci$confint %>%
+  as.data.frame() %>%
+  gt(caption = "Slopes by Diagnosis Method", rownames_to_stub = T) %>%
+  fmt_number(decimals = 2)
 # Model table
-m |> tbl_regression()
-```
+m |> tbl_regression() |> add_n(location = "level")
+```
+
+## FVC trends by diagnosis method
+
+### LOESS-smoothed plot
+
+```{r} 
+ggplot(
+  mod_df |> filter(!is.na(a1c_ogtt)),
+  aes(
+    y = A_FVC_pct_predicted,
+    x = years_to_cfrd,
+    color = a1c_ogtt,
+    group = a1c_ogtt
+  )
+) +
+  xlab("Years from CFRD") +
+  xlim(c(-5, 5)) +
+  geom_point(alpha = 0.1, shape = ".") +
+  geom_smooth(formula = y ~ s(x, bs = "cr")) +
+  theme_bw() +
+  theme(legend.title = element_blank())
+```
+
+### Piecewise linear model results
+
+```{r}
+# Fit model - adjust for age at dx, sex, genotype (severe vs mild), pancreatic
+# insufficiency, age at CF diagnosis (do we know this?), number of clinic
+# visits/year, use of CFTR modulators
+mod_df$x = mod_df$years_to_cfrd * (mod_df$years_to_cfrd >= 0)
+label(mod_df$x) = "Change in Slope Post-CFRD"
+label(mod_df$years_to_cfrd) = "Slope Pre-CFRD"
+label(mod_df$a1c_ogtt) = "Diagnosis Method"
+label(mod_df$Age_YrEnd) = "Age at Diagnosis"
+m <- lmer(
+  A_FVC_pct_predicted ~ years_to_cfrd *
+    a1c_ogtt +
+    x * a1c_ogtt +
+    Age_YrEnd +
+    Gender +
+    MutClass +
+    A_IsOnEnzymes +
+    A_Vx770 +
+    A_VX809comb +
+    (1 | eDWID),
+  data = mod_df
+)
+# Population-averaged plot
+pop_dat <- data.frame(expand.grid(
+  years_to_cfrd = -5:5,
+  a1c_ogtt = unique(mod_df$a1c_ogtt),
+  Age_YrEnd = mean(mod_df$Age_YrEnd, na.rm = T),
+  Gender = "M",
+  MutClass = "1-3",
+  A_IsOnEnzymes = "Yes",
+  A_Vx770 = "No",
+  A_VX809comb = "No"
+))
+pop_dat$x <- (pop_dat$years_to_cfrd >= 0) * pop_dat$years_to_cfrd
+pred <- predict(m, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(
+  pop_dat,
+  aes(x = years_to_cfrd, y = pred, group = a1c_ogtt, colour = a1c_ogtt)
+) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
+  xlab("Years from CFRD") +
+  ylab("FVC % Predicted") +
+  theme_classic(base_size = 15)
+# Contrasts
+neither_pre <- rep(0, nrow(summary(m)$coefficients))
+neither_change <- rep(0, nrow(summary(m)$coefficients))
+neither_pre[which(rownames(summary(m)$coefficients) == "years_to_cfrd")] = 1
+neither_change[which(rownames(summary(m)$coefficients) == "x")] = 1
+a1c_pre <- rep(0, nrow(summary(m)$coefficients))
+a1c_change <- rep(0, nrow(summary(m)$coefficients))
+a1c_pre[which(
+  rownames(summary(m)$coefficients) %in%
+    c("years_to_cfrd", "years_to_cfrd:a1c_ogttHbA1c Only")
+)] = 1
+a1c_change[which(
+  rownames(summary(m)$coefficients) %in%
+    c("x", "a1c_ogttHbA1c Only:x")
+)] = 1
+ogtt_pre <- rep(0, nrow(summary(m)$coefficients))
+ogtt_change <- rep(0, nrow(summary(m)$coefficients))
+ogtt_pre[which(
+  rownames(summary(m)$coefficients) %in%
+    c("years_to_cfrd", "years_to_cfrd:a1c_ogttOGTT")
+)] = 1
+ogtt_change[which(
+  rownames(summary(m)$coefficients) %in%
+    c("x", "a1c_ogttOGTT:x")
+)] = 1
+contr <- rbind(
+  "HbA1c Only Pre-CFRD" = a1c_pre,
+  "HbA1c Only Post-CFRD" = a1c_pre + a1c_change,
+  "OGTT Pre-CFRD" = ogtt_pre,
+  "OGTT Post-CFRD" = ogtt_pre + ogtt_change,
+  "Neither Pre-CFRD" = neither_pre,
+  "Neither Post-CFRD" = neither_pre + neither_change
+)
+change_mod_slopes <- glht(m, linfct = contr)
+change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
+change_mod_slopes_ci$confint %>%
+  as.data.frame() %>%
+  gt(caption = "Slopes by Diagnosis Method", rownames_to_stub = T) %>%
+  fmt_number(decimals = 2)
+# Model table
+m |> tbl_regression() |> add_n(location = "level")
+```

### 2025-11-20 Christine Chan/EnVision CF/fsOGTT_pre_post_ETI.qmd
matsuda in table 1

@@
 ```{r data cleaning}
+setwd(home_dir)
 # Import cleaned data
@@
     odi_igi,
-    odi_iauc
+    odi_iauc,
+    matsuda
   ) |>

### 2025-11-20 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
new pipe

--- 2025-11-19 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
@@
 # Remove those with T1D or T2D
-annual <- annual %>% arrange(eDWID, ReviewYear)
+annual <- annual |> arrange(eDWID, ReviewYear)
 annual <- annual[!annual$cfrd_status_annual %in% c(4, 5), ]
@@
 )
-demogs <- demogs %>% select(eDWID, Gender, MutClass, Race, Hispanicrace)
+demogs <- demogs |> select(eDWID, Gender, MutClass, Race, Hispanicrace)
 annual <- left_join(annual, demogs, by = join_by(eDWID))
@@
 # Get first year of CFRD
-first_cfrd <- annual %>%
-  filter(cfrd_status_annual == "CFRD+") %>%
-  group_by(eDWID) %>%
-  summarise(first_cfrd = min(ReviewYear))
+first_cfrd <- annual |>
+  filter(cfrd_status_annual == "CFRD+") |>
+  group_by(eDWID) |>
+  summarise(first_cfrd = min(ReviewYear), age_dx = min(Age_YrEnd))
 annual <- left_join(annual, first_cfrd, by = join_by(eDWID))
@@
 annual$tstart <- ymd(paste0(annual$ReviewYear, "-01-01"))
-annual <- annual %>%
-  group_by(eDWID) %>%
-  mutate(entry_date = tstart[1]) %>%
+annual <- annual |>
+  group_by(eDWID) |>
+  mutate(entry_date = tstart[1]) |>
   ungroup()
@@
 annual$age_group <- relevel(annual$age_group, ref = "[10,19)")
-annual <- annual %>% filter(Age_YrEnd != 0, Age_YrEnd >= 10)
+annual <- annual |> filter(Age_YrEnd != 0, Age_YrEnd >= 10)
 annual$age_group <- droplevels(annual$age_group)
@@
   cfrd_only,
-  annual %>% mutate(eDWID = as.numeric(eDWID)),
+  annual |> mutate(eDWID = as.numeric(eDWID)),
   by = join_by(id == eDWID, year == ReviewYear)
@@
 )
-cfrd_only %>%
+cfrd_only |>
   select(
@@
     A_cultureresults3
-  ) %>%
+  ) |>
   tbl_summary(
@@
     statistic = list(all_continuous() ~ "{mean} ({sd})")
-  ) %>%
-  add_overall() %>%
-  add_p(test = list(all_continuous() ~ "oneway.test")) %>%
+  ) |>
+  add_overall() |>
+  add_p(test = list(all_continuous() ~ "oneway.test")) |>
   modify_caption("**All Groups**")
-cfrd_only %>%
-  filter(a1c_ogtt %in% c("OGTT", "HbA1c Only")) %>%
-  mutate(a1c_ogtt = droplevels(a1c_ogtt)) %>%
+cfrd_only |>
+  filter(a1c_ogtt %in% c("OGTT", "HbA1c Only")) |>
+  mutate(a1c_ogtt = droplevels(a1c_ogtt)) |>
   select(
@@
     statistic = list(all_continuous() ~ "{mean} ({sd})")
-  ) %>%
-  add_overall() %>%
-  add_p(test = list(all_continuous() ~ "oneway.test")) %>%
+  ) |>
+  add_overall() |>
+  add_p(test = list(all_continuous() ~ "oneway.test")) |>
   modify_caption("**OGTT vs. HbA1c Only**")
@@
 # Plot by diagnosis type
-mod_df = cfrd_only %>%
-  select(id, a1c_ogtt) %>%
+mod_df = cfrd_only |>
+  select(id, a1c_ogtt, Age_YrEnd) |>
   rename(eDWID = id)
 mod_df = left_join(
-  annual %>% mutate(eDWID = as.numeric(eDWID)),
+  annual |> select(-Age_YrEnd) |> mutate(eDWID = as.numeric(eDWID)),
   mod_df,
@@
 )
-mod_df = mod_df %>%
-  group_by(eDWID) %>%
+mod_df = mod_df |>
+  drop_na(first_cfrd) |>
+  group_by(eDWID) |>
   mutate(years_to_cfrd = ReviewYear - first_cfrd)
 var_label(mod_df$years_to_cfrd) = "Years to CFRD"
+var_label(mod_df$Age_YrEnd) = "Age at Diagnosis"
 ggplot(
-  mod_df %>% filter(!is.na(a1c_ogtt)),
+  mod_df |> filter(!is.na(a1c_ogtt)),
   aes(
@@
 # Population-averaged plot
-
+pop_dat <- data.frame(expand.grid(
+  years_to_cfrd = -5:5,
+  a1c_ogtt = unique(mod_df$a1c_ogtt),
+  Age_YrEnd = mean(mod_df$Age_YrEnd, na.rm = T),
+  Gender = "M",
+  MutClass = "1-3",
+  A_IsOnEnzymes = "Yes",
+  A_Vx770 = "No",
+  A_VX809comb = "No"
+))
+pop_dat$x <- (pop_dat$years_to_cfrd >= 0) * pop_dat$years_to_cfrd
+pred <- predict(m, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(
+  pop_dat,
+  aes(x = years_to_cfrd, y = pred, group = a1c_ogtt, colour = a1c_ogtt)
+) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
+  xlab("Years from CFRD") +
+  ylab("FEV1 % Predicted") +
+  theme_classic(base_size = 15)
+# Contrasts
+neither_pre <- rep(0, nrow(summary(m)$coefficients))
+neither_change <- rep(0, nrow(summary(m)$coefficients))
+neither_pre[which(rownames(summary(m)$coefficients) == "years_to_cfrd")] = 1
+neither_change[which(rownames(summary(m)$coefficients) == "x")] = 1
+a1c_pre <- rep(0, nrow(summary(m)$coefficients))
+a1c_change <- rep(0, nrow(summary(m)$coefficients))
+a1c_pre[which(
+  rownames(summary(m)$coefficients) %in%
+    c("years_to_cfrd", "years_to_cfrd:a1c_ogttHbA1c Only")
+)] = 1
+a1c_change[which(
+  rownames(summary(m)$coefficients) %in%
+    c("x", "a1c_ogttHbA1c Only:x")
+)] = 1
+ogtt_pre <- rep(0, nrow(summary(m)$coefficients))
+ogtt_change <- rep(0, nrow(summary(m)$coefficients))
+ogtt_pre[which(
+  rownames(summary(m)$coefficients) %in%
+    c("years_to_cfrd", "years_to_cfrd:a1c_ogttOGTT")
+)] = 1
+ogtt_change[which(
+  rownames(summary(m)$coefficients) %in%
+    c("x", "a1c_ogttOGTT:x")
+)] = 1
+contr <- rbind(
+  "HbA1c Only Pre-CFRD" = a1c_pre,
+  "HbA1c Only Post-CFRD" = a1c_pre + a1c_change,
+  "OGTT Pre-CFRD" = ogtt_pre,
+  "OGTT Post-CFRD" = ogtt_pre + ogtt_change,
+  "Neither Pre-CFRD" = neither_pre,
+  "Neither Post-CFRD" = neither_pre + neither_change
+)
+change_mod_slopes <- glht(m, linfct = contr)
+change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
+change_mod_slopes_ci$confint |>
+  as.data.frame() |>
+  gt(caption = "Slopes by Diagnosis Method", rownames_to_stub = T) |>
+  fmt_number(decimals = 2)
 # Model table
-m |> tbl_regression()
-```
+m |> tbl_regression() |> add_n(location = "level")
+```
+
+## FVC trends by diagnosis method
+
+### LOESS-smoothed plot
+
+```{r} 
+ggplot(
+  mod_df |> filter(!is.na(a1c_ogtt)),
+  aes(
+    y = A_FVC_pct_predicted,
+    x = years_to_cfrd,
+    color = a1c_ogtt,
+    group = a1c_ogtt
+  )
+) +
+  xlab("Years from CFRD") +
+  xlim(c(-5, 5)) +
+  geom_point(alpha = 0.1, shape = ".") +
+  geom_smooth(formula = y ~ s(x, bs = "cr")) +
+  theme_bw() +
+  theme(legend.title = element_blank())
+```
+
+### Piecewise linear model results
+
+```{r}
+# Fit model - adjust for age at dx, sex, genotype (severe vs mild), pancreatic
+# insufficiency, age at CF diagnosis (do we know this?), number of clinic
+# visits/year, use of CFTR modulators
+mod_df$x = mod_df$years_to_cfrd * (mod_df$years_to_cfrd >= 0)
+label(mod_df$x) = "Change in Slope Post-CFRD"
+label(mod_df$years_to_cfrd) = "Slope Pre-CFRD"
+label(mod_df$a1c_ogtt) = "Diagnosis Method"
+label(mod_df$Age_YrEnd) = "Age at Diagnosis"
+m <- lmer(
+  A_FVC_pct_predicted ~ years_to_cfrd *
+    a1c_ogtt +
+    x * a1c_ogtt +
+    Age_YrEnd +
+    Gender +
+    MutClass +
+    A_IsOnEnzymes +
+    A_Vx770 +
+    A_VX809comb +
+    (1 | eDWID),
+  data = mod_df
+)
+# Population-averaged plot
+pop_dat <- data.frame(expand.grid(
+  years_to_cfrd = -5:5,
+  a1c_ogtt = unique(mod_df$a1c_ogtt),
+  Age_YrEnd = mean(mod_df$Age_YrEnd, na.rm = T),
+  Gender = "M",
+  MutClass = "1-3",
+  A_IsOnEnzymes = "Yes",
+  A_Vx770 = "No",
+  A_VX809comb = "No"
+))
+pop_dat$x <- (pop_dat$years_to_cfrd >= 0) * pop_dat$years_to_cfrd
+pred <- predict(m, re.form = NA, newdata = pop_dat, se.fit = T)
+pop_dat$pred <- pred$fit
+pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
+pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+ggplot(
+  pop_dat,
+  aes(x = years_to_cfrd, y = pred, group = a1c_ogtt, colour = a1c_ogtt)
+) +
+  geom_point() +
+  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_line() +
+  xlab("Years from CFRD") +
+  ylab("FVC % Predicted") +
+  theme_classic(base_size = 15)
+# Contrasts
+neither_pre <- rep(0, nrow(summary(m)$coefficients))
+neither_change <- rep(0, nrow(summary(m)$coefficients))
+neither_pre[which(rownames(summary(m)$coefficients) == "years_to_cfrd")] = 1
+neither_change[which(rownames(summary(m)$coefficients) == "x")] = 1
+a1c_pre <- rep(0, nrow(summary(m)$coefficients))
+a1c_change <- rep(0, nrow(summary(m)$coefficients))
+a1c_pre[which(
+  rownames(summary(m)$coefficients) %in%
+    c("years_to_cfrd", "years_to_cfrd:a1c_ogttHbA1c Only")
+)] = 1
+a1c_change[which(
+  rownames(summary(m)$coefficients) %in%
+    c("x", "a1c_ogttHbA1c Only:x")
+)] = 1
+ogtt_pre <- rep(0, nrow(summary(m)$coefficients))
+ogtt_change <- rep(0, nrow(summary(m)$coefficients))
+ogtt_pre[which(
+  rownames(summary(m)$coefficients) %in%
+    c("years_to_cfrd", "years_to_cfrd:a1c_ogttOGTT")
+)] = 1
+ogtt_change[which(
+  rownames(summary(m)$coefficients) %in%
+    c("x", "a1c_ogttOGTT:x")
+)] = 1
+contr <- rbind(
+  "HbA1c Only Pre-CFRD" = a1c_pre,
+  "HbA1c Only Post-CFRD" = a1c_pre + a1c_change,
+  "OGTT Pre-CFRD" = ogtt_pre,
+  "OGTT Post-CFRD" = ogtt_pre + ogtt_change,
+  "Neither Pre-CFRD" = neither_pre,
+  "Neither Post-CFRD" = neither_pre + neither_change
+)
+change_mod_slopes <- glht(m, linfct = contr)
+change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
+change_mod_slopes_ci$confint |>
+  as.data.frame() |>
+  gt(caption = "Slopes by Diagnosis Method", rownames_to_stub = T) |>
+  fmt_number(decimals = 2)
+# Model table
+m |> tbl_regression() |> add_n(location = "level")
+```

### 2025-11-24 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
nicer SSwAB-IT tables

@@
 library(redcapAPI)
-library(tidyverse)
 library(gt)
@@
 library(ggpubr)
-home_dir <- switch(Sys.info()["sysname"],
+library(tidyverse)
+home_dir <- switch(
+  Sys.info()["sysname"],
   "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Jordana Hoppe/SSwAB-IT",
@@
   cat("\n")
-    d %>%
-        tbl_cross(percent = "cell") %>%
+  d1 = d %>%
+    tbl_cross(percent = "cell", digits = c(0, 1)) %>%
     print()
@@
   d %>%
-        tbl_summary(by = COS, percent = "cell") %>%
-        add_ci(method = list(all_categorical() ~ "wald")) %>%
+    tbl_summary(
+      by = COS,
+      percent = "cell",
+      digits = list("SOS" ~ c(0, 1))
+    ) %>%
+    add_ci(
+      pattern = "{stat} [{ci}]",
+      method = list(all_categorical() ~ "wald"),
+      style_fun = all_categorical() ~ label_style_number(
+        scale = 100,
+        digits = 1
+      )
+    ) %>%
     print()
@@
         detailed_agree ~ "Detailed Agreement"
-            )
+      ),
+      digits = list(c("agree", "detailed_agree") ~ c(0, 1))
     ) %>%
     add_overall() %>%
-        add_p(test.args = all_tests("fisher.test") ~ list(workspace = 2e9)) %>%
+    add_p(
+      test = list(agree ~ "chisq.test", detailed_agree ~ "fisher.test"),
+      test.args = all_tests("fisher.test") ~ list(workspace = 2e9)
+    ) %>%
     separate_p_footnotes() %>%

### 2025-11-25 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
separate report for lung function analyses

--- 2025-11-19 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
@@
 ```
-
-# Diagnosis comparisons for HbA1c registry paper
-
-```{r}
-#| warning: false
-#| message: false
-# Import data and codebook
-annual <- read.csv(
-  "./Data_Raw/CFF19_Annualized_Del1.csv",
-  na.strings = c("", "N/A")
-)
-data_dict <- readxl::read_excel("./Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
-data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
-# Remove those with T1D or T2D
-annual <- annual %>% arrange(eDWID, ReviewYear)
-annual <- annual[!annual$cfrd_status_annual %in% c(4, 5), ]
-annual$cfrd_status_annual <- factor(
-  annual$cfrd_status_annual,
-  levels = 1:3,
-  labels = c("CFRD-", "CFRD-", "CFRD+")
-)
-# Get mutations
-demogs <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
-races <- c(
-  "White",
-  "Black or African American",
-  "American Indian or Alaska Native",
-  "Asian",
-  "Native Hawaiian or Other Pacific ",
-  "Some other race"
-)
-demogs$Race <- apply(demogs, 1, function(r) {
-  w <- which(r[paste0("Race", 1:6)] == 1)
-  if (length(w) > 1) {
-    return("More than one race")
-  } else {
-    return(races[w])
-  }
-})
-demogs$Race <- factor(
-  demogs$Race,
-  levels = c(
-    "White",
-    "American Indian or Alaska Native",
-    "Asian",
-    "Black or African American",
-    "More than one race",
-    "Native Hawaiian or Other Pacific ",
-    "Some other race"
-  ),
-  labels = c(
-    "White",
-    "Other race",
-    "Other race",
-    "Black or African American",
-    "More than one race",
-    "Other race",
-    "Other race"
-  )
-)
-demogs$Hispanicrace <- factor(
-  demogs$Hispanicrace,
-  levels = 1:2,
-  labels = c("Yes", "No")
-)
-demogs <- demogs %>% select(eDWID, Gender, MutClass, Race, Hispanicrace)
-annual <- left_join(annual, demogs, by = join_by(eDWID))
-# Variable formatting
-# If A_supplement_feeding is 0, then A_feeding2 is also 0 because they are not
-# on any supplemental feeding
-# Include all kinds of tube feeds
-annual$A_feeding2[annual$A_supplement_feeding == 0] <- 0
-annual$A_feeding3[annual$A_supplement_feeding == 0] <- 0
-annual$A_feeding4[annual$A_supplement_feeding == 0] <- 0
-tube_fields <- c("A_feeding2", "A_feeding3", "A_feeding4")
-annual$tube_feeding <-
-  apply(annual[, tube_fields], 1, function(r) {
-    if (all(is.na(r))) {
-      return(NA)
-    } else {
-      return(max(as.numeric(r), na.rm = T))
-    }
-  })
-# Liver disease
-liver_fields <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
-annual$liver_disease <-
-  apply(annual[, liver_fields], 1, function(r) {
-    if (all(is.na(r))) {
-      return(NA)
-    } else {
-      return(max(as.numeric(r), na.rm = T))
-    }
-  })
-# Get BMI categories
-annual$bmi <- annual$A_weight / ((annual$A_height / 100)^2)
-annual$bmi <- coalesce(annual$A_bmivalue, annual$bmi)
-annual$bmi_perc_cdc <- sds(
-  annual$bmi,
-  age = annual$Age_YrEnd,
-  sex = annual$Gender,
-  male = "M",
-  female = "F",
-  item = "bmi",
-  ref = cdc.ref,
-  type = "perc"
-) *
-  100
-annual$bmi_perc_who <- sds(
-  annual$bmi,
-  age = annual$Age_YrEnd,
-  sex = annual$Gender,
-  male = "M",
-  female = "F",
-  item = "bmi",
-  ref = who.ref,
-  type = "perc"
-) *
-  100
-annual$bmi_perc <- coalesce(annual$bmi_perc_cdc, annual$bmi_perc_who)
-annual$bmi_cat <- apply(annual, 1, function(r) {
-  age <- as.numeric(r["Age_YrEnd"])
-  gender <- as.character(r["Gender"])
-  bmi <- as.numeric(r["bmi"])
-  bmi_perc <- as.numeric(r["bmi_perc"])
-  if (age <= 20) {
-    return(cut(
-      bmi_perc,
-      c(-Inf, 10, 50, 85, Inf),
-      right = F,
-      labels = c("Undernourished", "At Risk", "Target/Goal", "Overweight")
-    ))
-  } else if (age > 20) {
-    if (gender == "M") {
-      return(cut(
-        bmi,
-        c(-Inf, 18.5, 23, 25, Inf),
-        right = F,
-        labels = c(
-          "Undernourished",
-          "At Risk",
-          "Target/Goal",
-          "Overweight"
-        )
-      ))
-    } else if (gender == "F") {
-      return(cut(
-        bmi,
-        c(-Inf, 18.5, 22, 25, Inf),
-        right = F,
-        labels = c(
-          "Undernourished",
-          "At Risk",
-          "Target/Goal",
-          "Overweight"
-        )
-      ))
-    }
-  }
-})
-annual$bmi_cat <- relevel(annual$bmi_cat, ref = "Target/Goal")
-# Raw values for adults, percentiles for peds
-annual$bmi_adult = annual$bmi
-annual$bmi_adult[annual$Age_YrEnd <= 20] = NA
-annual$bmi_perc = annual$bmi_perc
-annual$bmi_perc[annual$Age_YrEnd > 20] = NA
-# Labels
-var_label(annual$bmi) = "BMI (all participants)"
-var_label(annual$bmi_perc) = "BMI percentile (age <= 20)"
-var_label(annual$bmi_adult) = "BMI value (age > 20)"
-var_label(annual$bmi_cat) = "BMI category"
-# Format columns
-annual$MutClass <- factor(
-  annual$MutClass,
-  levels = c("1-3", "4-5", "Oth", "N/A"),
-  labels = c("1-3", "4-5", "Oth", "Unknown")
-)
-annual$NumPulmExacerbation <- cut(
-  annual$NumPulmExacerbation,
-  labels = c("0", "1", "2", ">=3"),
-  breaks = c(0, 1, 2, 3, Inf),
-  right = F
-)
-# Age groups 10 - 18 inclusive, 19 - 30 inclusive, >= 31
-annual$age_group <- cut(
-  annual$Age_YrEnd,
-  breaks = c(-Inf, 10, 19, 31, Inf),
-  right = F
-)
-# Bugs
-bugs <- c(
-  "A_pseudomonasaeruginosa",
-  "A_fungalyeast1",
-  "A_pulmonarycomplications1",
-  "A_mycocultureresults3",
-  "A_cultureresults3"
-)
-# Yes/No variables
-yn_vars <- c(
-  "A_corticosteroids1",
-  "A_feeding2",
-  "A_Vx770",
-  "A_VX809comb",
-  bugs,
-  "tube_feeding",
-  "liver_disease"
-)
-annual[, yn_vars] <- lapply(
-  annual[, yn_vars],
-  factor,
-  levels = 0:1,
-  labels = c("No", "Yes")
-)
-annual$A_IsOnEnzymes <- factor(
-  annual$A_IsOnEnzymes,
-  levels = c(1, 0),
-  labels = c(" Yes", " No")
-)
-# FEV1 to categorical
-annual$ppFEV1_cat <- cut(
-  annual$A_FEV1_pct_predicted,
-  breaks = c(0, 40, 80, 100, Inf),
-  right = F
-)
-annual$ppFEV1_cat <- factor(
-  annual$ppFEV1_cat,
-  levels = c("[100,Inf)", "[80,100)", "[40,80)", "[0,40)")
-)
-# Males to reference group
-annual$Gender <- factor(annual$Gender, levels = c("M", "F"))
-# Hispanic no as reference
-annual$Hispanicrace <- factor(annual$Hispanicrace, levels = c("No", "Yes"))
-# Get first year of CFRD
-first_cfrd <- annual %>%
-  filter(cfrd_status_annual == "CFRD+") %>%
-  group_by(eDWID) %>%
-  summarise(first_cfrd = min(ReviewYear), age_dx = min(Age_YrEnd))
-annual <- left_join(annual, first_cfrd, by = join_by(eDWID))
-# Year prior to first CFRD in registry - assume negative for CFRD. Assume all
-# years following are CFRD+
-annual$CFRD <- NA
-annual$CFRD[annual$ReviewYear < annual$first_cfrd] <- "CFRD-"
-annual$CFRD[annual$ReviewYear >= annual$first_cfrd] <- "CFRD+"
-annual$CFRD[is.na(annual$first_cfrd)] <- "CFRD-"
-annual$CFRD <- factor(annual$CFRD, levels = c("CFRD-", "CFRD+"))
-# Add formatted columns for survival analysis
-annual$tstop <- ymd(paste0(annual$ReviewYear, "-12-31"))
-annual$tstart <- ymd(paste0(annual$ReviewYear, "-01-01"))
-annual <- annual %>%
-  group_by(eDWID) %>%
-  mutate(entry_date = tstart[1]) %>%
-  ungroup()
-# Remove those under 10
-annual$age_group <- relevel(annual$age_group, ref = "[10,19)")
-annual <- annual %>% filter(Age_YrEnd != 0, Age_YrEnd >= 10)
-annual$age_group <- droplevels(annual$age_group)
-# List of covariates
-vars <- c(
-  "age_group",
-  "Gender",
-  "bmi_perc",
-  "bmi_cat",
-  "MutClass",
-  "Race",
-  "Hispanicrace",
-  "A_IsOnEnzymes",
-  "ppFEV1_cat",
-  "tube_feeding",
-  "NumPulmExacerbation",
-  "A_pseudomonasaeruginosa",
-  "A_fungalyeast1",
-  "A_mycocultureresults3",
-  "A_corticosteroids1",
-  "cfrd_status_annual",
-  "A_Vx770",
-  "A_VX809comb",
-  "liver_disease"
-)
-# Nice labels for plots and tables
-labels <- list(
-  cfrd_status_annual = "CFRD Status",
-  MutClass = "Mutation class",
-  Race = "Race",
-  Hispanicrace = "Hispanic",
-  Gender = "Sex",
-  A_IsOnEnzymes = "Pancreatic enzyme use",
-  ppFEV1_cat = "FEV1 percent predicted category",
-  A_FEV1_pct_predicted = "FEV1 percent predicted",
-  A_FVC_pct_predicted = "FVC percent predicted",
-  tube_feeding = "Enteric feedings",
-  liver_disease = "Liver disease",
-  A_corticosteroids1 = "Use of corticosteroids",
-  NumPulmExacerbation = "Admissions for PEx",
-  age_group = "Age group",
-  height_perc = "Height percentile",
-  A_Vx770 = "Use of CFTR modulator Vx770",
-  A_VX809comb = "Use of CFTR modulator Vx809 combination",
-  A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
-  A_fungalyeast1 = "Aspergillus (any species)",
-  A_mycocultureresults3 = "Microorganisms in myco culture",
-  A_cultureresults3 = "Microorganisms in culture",
-  eDWID = "ID",
-  tstart = "Interval start",
-  tstop = "Interval stop",
-  Age_YrEnd = "Age"
-)
-label(annual[, which(colnames(annual) %in% names(labels))]) <-
-  labels[colnames(annual)[which(colnames(annual) %in% names(labels))]]
-# Per reviewer, out of X individuals with a diagnosis of CFRD, X number had an
-# OGTT (and X had an A1c, but no OGTT) that year or the year prior.
-# This is super slow but idc idc
-cfrd_only <- apply(first_cfrd, 1, function(r) {
-  id <- as.numeric(r["eDWID"])
-  year <- as.numeric(r["first_cfrd"])
-  d <- annual[annual$eDWID == id & annual$ReviewYear %in% c(year, year - 1), ]
-  a1c <- as.numeric(last(d["A_hgba1c"], na_rm = TRUE))
-  ogtt0hr <- as.numeric(last(d["A_ogttfast_bloodglucose"], na_rm = TRUE))
-  ogtt2hr <- as.numeric(last(d["A_twohour_bloodglucose"], na_rm = TRUE))
-  return(c(id, year, a1c, ogtt0hr, ogtt2hr))
-})
-cfrd_only <- data.frame(t(cfrd_only))
-colnames(cfrd_only) <- c("id", "year", "a1c", "ogtt0hr", "ogtt2hr")
-cfrd_only <- left_join(
-  cfrd_only,
-  annual %>% mutate(eDWID = as.numeric(eDWID)),
-  by = join_by(id == eDWID, year == ReviewYear)
-)
-cfrd_only$a1c_ogtt <- apply(cfrd_only, 1, function(r) {
-  not_missing <- paste(
-    which(!is.na(r[c("a1c", "ogtt0hr", "ogtt2hr")])),
-    collapse = "_"
-  )
-  return(not_missing)
-})
-cfrd_only$a1c_ogtt <- factor(
-  cfrd_only$a1c_ogtt,
-  levels = c("", "1", "1_2", "1_2_3", "1_3", "2", "2_3", "3"),
-  labels = c(
-    "Neither",
-    "HbA1c Only",
-    "OGTT",
-    "OGTT",
-    "OGTT",
-    "OGTT",
-    "OGTT",
-    "OGTT"
-  )
-)
-cfrd_only %>%
-  select(
-    a1c_ogtt,
-    Age_YrEnd,
-    A_weight,
-    A_height,
-    bmi_adult,
-    bmi_perc,
-    bmi_cat,
-    A_FEV1_pct_predicted,
-    A_FVC_pct_predicted,
-    Race,
-    Hispanicrace,
-    MutClass,
-    a1c,
-    A_ogttfast_bloodglucose,
-    A_twohour_bloodglucose,
-    A_IsOnEnzymes,
-    liver_disease,
-    A_pseudomonasaeruginosa,
-    A_fungalyeast1,
-    A_mycocultureresults3,
-    A_cultureresults3
-  ) %>%
-  tbl_summary(
-    by = a1c_ogtt,
-    missing_text = "Missing",
-    statistic = list(all_continuous() ~ "{mean} ({sd})")
-  ) %>%
-  add_overall() %>%
-  add_p(test = list(all_continuous() ~ "oneway.test")) %>%
-  modify_caption("**All Groups**")
-cfrd_only %>%
-  filter(a1c_ogtt %in% c("OGTT", "HbA1c Only")) %>%
-  mutate(a1c_ogtt = droplevels(a1c_ogtt)) %>%
-  select(
-    a1c_ogtt,
-    Age_YrEnd,
-    A_weight,
-    A_height,
-    A_bmivalue,
-    A_bmipercentile,
-    A_FEV1_pct_predicted,
-    A_FVC_pct_predicted,
-    Race,
-    Hispanicrace,
-    MutClass,
-    a1c,
-    A_ogttfast_bloodglucose,
-    A_twohour_bloodglucose,
-    A_IsOnEnzymes,
-    liver_disease,
-    A_pseudomonasaeruginosa,
-    A_fungalyeast1,
-    A_mycocultureresults3,
-    A_cultureresults3
-  ) %>%
-  tbl_summary(
-    by = a1c_ogtt,
-    missing_text = "Missing",
-    statistic = list(all_continuous() ~ "{mean} ({sd})")
-  ) %>%
-  add_overall() %>%
-  add_p(test = list(all_continuous() ~ "oneway.test")) %>%
-  modify_caption("**OGTT vs. HbA1c Only**")
-```
-
-## FEV1 trends by diagnosis method
-
-### LOESS-smoothed plot
-
-```{r}
-#| warning: false
-#| message: false
-# Plot by diagnosis type
-mod_df = cfrd_only %>%
-  select(id, a1c_ogtt, Age_YrEnd) %>%
-  rename(eDWID = id)
-mod_df = left_join(
-  annual %>% select(-Age_YrEnd) |> mutate(eDWID = as.numeric(eDWID)),
-  mod_df,
-  by = join_by(eDWID)
-)
-mod_df = mod_df %>%
-  drop_na(first_cfrd) |>
-  group_by(eDWID) %>%
-  mutate(years_to_cfrd = ReviewYear - first_cfrd)
-var_label(mod_df$years_to_cfrd) = "Years to CFRD"
-var_label(mod_df$Age_YrEnd) = "Age at Diagnosis"
-ggplot(
-  mod_df %>% filter(!is.na(a1c_ogtt)),
-  aes(
-    y = A_FEV1_pct_predicted,
-    x = years_to_cfrd,
-    color = a1c_ogtt,
-    group = a1c_ogtt
-  )
-) +
-  xlab("Years from CFRD") +
-  xlim(c(-10, 10)) +
-  geom_point(alpha = 0.1, shape = ".") +
-  geom_smooth(formula = y ~ s(x, bs = "cr")) +
-  theme_bw() +
-  theme(legend.title = element_blank())
-```
-
-### Piecewise linear model results
-
-There were not enough pancreatic sufficient people in this dataset to include that varaible in the model.
-
-```{r}
-# Fit model - adjust for age at dx, sex, genotype (severe vs mild), pancreatic
-# insufficiency, age at CF diagnosis (do we know this?), number of clinic
-# visits/year, use of CFTR modulators
-mod_df$x = mod_df$years_to_cfrd * (mod_df$years_to_cfrd >= 0)
-label(mod_df$x) = "Change in Slope Post-CFRD"
-label(mod_df$years_to_cfrd) = "Slope Pre-CFRD"
-label(mod_df$a1c_ogtt) = "Diagnosis Method"
-label(mod_df$Age_YrEnd) = "Age at Diagnosis"
-m <- lmer(
-  A_FEV1_pct_predicted ~ years_to_cfrd *
-    a1c_ogtt +
-    x * a1c_ogtt +
-    Age_YrEnd +
-    Gender +
-    MutClass +
-    A_Vx770 +
-    A_VX809comb +
-    (1 | eDWID),
-  data = mod_df
-)
-# Population-averaged plot
-pop_dat <- data.frame(expand.grid(
-  years_to_cfrd = -10:10,
-  a1c_ogtt = unique(mod_df$a1c_ogtt),
-  Age_YrEnd = mean(mod_df$Age_YrEnd, na.rm = T),
-  Gender = "M",
-  MutClass = "1-3",
-  A_Vx770 = "No",
-  A_VX809comb = "No"
-))
-pop_dat$x <- (pop_dat$years_to_cfrd >= 0) * pop_dat$years_to_cfrd
-pred <- predict(m, re.form = NA, newdata = pop_dat, se.fit = T)
-pop_dat$pred <- pred$fit
-pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
-pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
-ggplot(
-  pop_dat,
-  aes(x = years_to_cfrd, y = pred, group = a1c_ogtt, colour = a1c_ogtt)
-) +
-  geom_point() +
-  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
-  geom_line() +
-  xlab("Years from CFRD") +
-  ylab("FEV1 % Predicted") +
-  theme_classic(base_size = 15)
-# Model table
-m |> tbl_regression()
-# Contrasts
-neither_pre <- rep(0, nrow(summary(m)$coefficients))
-neither_change <- rep(0, nrow(summary(m)$coefficients))
-neither_pre[which(rownames(summary(m)$coefficients) == "years_to_cfrd")] = 1
-neither_change[which(rownames(summary(m)$coefficients) == "x")] = 1
-a1c_pre <- rep(0, nrow(summary(m)$coefficients))
-a1c_change <- rep(0, nrow(summary(m)$coefficients))
-a1c_pre[which(
-  rownames(summary(m)$coefficients) %in%
-    c("years_to_cfrd", "years_to_cfrd:a1c_ogttHbA1c Only")
-)] = 1
-a1c_change[which(
-  rownames(summary(m)$coefficients) %in%
-    c("x", "a1c_ogttHbA1c Only:x")
-)] = 1
-ogtt_pre <- rep(0, nrow(summary(m)$coefficients))
-ogtt_change <- rep(0, nrow(summary(m)$coefficients))
-ogtt_pre[which(
-  rownames(summary(m)$coefficients) %in%
-    c("years_to_cfrd", "years_to_cfrd:a1c_ogttOGTT")
-)] = 1
-ogtt_change[which(
-  rownames(summary(m)$coefficients) %in%
-    c("x", "a1c_ogttOGTT:x")
-)] = 1
-contr <- rbind(
-  "HbA1c Only Pre-CFRD" = a1c_pre,
-  "HbA1c Only Post-CFRD" = a1c_pre + a1c_change,
-  "OGTT Pre-CFRD" = ogtt_pre,
-  "OGTT Post-CFRD" = ogtt_pre + ogtt_change,
-  "Neither Pre-CFRD" = neither_pre,
-  "Neither Post-CFRD" = neither_pre + neither_change
-)
-change_mod_slopes <- glht(m, linfct = contr)
-change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
-change_mod_slopes_ci$confint %>%
-  as.data.frame() %>%
-  gt(caption = "Slopes by Diagnosis Method", rownames_to_stub = T) %>%
-  fmt_number(decimals = 2)
-```
-
-## FVC trends by diagnosis method
-
-### LOESS-smoothed plot
-
-```{r} 
-ggplot(
-  mod_df %>% filter(!is.na(a1c_ogtt)),
-  aes(
-    y = A_FVC_pct_predicted,
-    x = years_to_cfrd,
-    color = a1c_ogtt,
-    group = a1c_ogtt
-  )
-) +
-  xlab("Years from CFRD") +
-  xlim(c(-10, 10)) +
-  geom_point(alpha = 0.1, shape = ".") +
-  geom_smooth(formula = y ~ s(x, bs = "cr")) +
-  theme_bw() +
-  theme(legend.title = element_blank())
-```
-
-### Piecewise linear model results
-
-There were not enough pancreatic sufficient people in this dataset to include that varaible in the model.
-
-```{r}
-# Fit model - adjust for age at dx, sex, genotype (severe vs mild), pancreatic
-# insufficiency, age at CF diagnosis (do we know this?), number of clinic
-# visits/year, use of CFTR modulators
-mod_df$x = mod_df$years_to_cfrd * (mod_df$years_to_cfrd >= 0)
-label(mod_df$x) = "Change in Slope Post-CFRD"
-label(mod_df$years_to_cfrd) = "Slope Pre-CFRD"
-label(mod_df$a1c_ogtt) = "Diagnosis Method"
-label(mod_df$Age_YrEnd) = "Age at Diagnosis"
-m <- lmer(
-  A_FVC_pct_predicted ~ years_to_cfrd *
-    a1c_ogtt +
-    x * a1c_ogtt +
-    Age_YrEnd +
-    Gender +
-    MutClass +
-    A_Vx770 +
-    A_VX809comb +
-    (1 | eDWID),
-  data = mod_df
-)
-# Population-averaged plot
-pop_dat <- data.frame(expand.grid(
-  years_to_cfrd = -10:10,
-  a1c_ogtt = unique(mod_df$a1c_ogtt),
-  Age_YrEnd = mean(mod_df$Age_YrEnd, na.rm = T),
-  Gender = "M",
-  MutClass = "1-3",
-  A_Vx770 = "No",
-  A_VX809comb = "No"
-))
-pop_dat$x <- (pop_dat$years_to_cfrd >= 0) * pop_dat$years_to_cfrd
-pred <- predict(m, re.form = NA, newdata = pop_dat, se.fit = T)
-pop_dat$pred <- pred$fit
-pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
-pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
-ggplot(
-  pop_dat,
-  aes(x = years_to_cfrd, y = pred, group = a1c_ogtt, colour = a1c_ogtt)
-) +
-  geom_point() +
-  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
-  geom_line() +
-  xlab("Years from CFRD") +
-  ylab("FVC % Predicted") +
-  theme_classic(base_size = 15)
-# Model table
-m |> tbl_regression()
-# Contrasts
-neither_pre <- rep(0, nrow(summary(m)$coefficients))
-neither_change <- rep(0, nrow(summary(m)$coefficients))
-neither_pre[which(rownames(summary(m)$coefficients) == "years_to_cfrd")] = 1
-neither_change[which(rownames(summary(m)$coefficients) == "x")] = 1
-a1c_pre <- rep(0, nrow(summary(m)$coefficients))
-a1c_change <- rep(0, nrow(summary(m)$coefficients))
-a1c_pre[which(
-  rownames(summary(m)$coefficients) %in%
-    c("years_to_cfrd", "years_to_cfrd:a1c_ogttHbA1c Only")
-)] = 1
-a1c_change[which(
-  rownames(summary(m)$coefficients) %in%
-    c("x", "a1c_ogttHbA1c Only:x")
-)] = 1
-ogtt_pre <- rep(0, nrow(summary(m)$coefficients))
-ogtt_change <- rep(0, nrow(summary(m)$coefficients))
-ogtt_pre[which(
-  rownames(summary(m)$coefficients) %in%
-    c("years_to_cfrd", "years_to_cfrd:a1c_ogttOGTT")
-)] = 1
-ogtt_change[which(
-  rownames(summary(m)$coefficients) %in%
-    c("x", "a1c_ogttOGTT:x")
-)] = 1
-contr <- rbind(
-  "HbA1c Only Pre-CFRD" = a1c_pre,
-  "HbA1c Only Post-CFRD" = a1c_pre + a1c_change,
-  "OGTT Pre-CFRD" = ogtt_pre,
-  "OGTT Post-CFRD" = ogtt_pre + ogtt_change,
-  "Neither Pre-CFRD" = neither_pre,
-  "Neither Post-CFRD" = neither_pre + neither_change
-)
-change_mod_slopes <- glht(m, linfct = contr)
-change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
-change_mod_slopes_ci$confint %>%
-  as.data.frame() %>%
-  gt(caption = "Slopes by Diagnosis Method", rownames_to_stub = T) %>%
-  fmt_number(decimals = 2)
-```

### 2025-11-25 Christine Chan/CF Registry/cf_registry_lung_function_analysis.qmd
separate report for lung function analyses

---
title: "CF Registry Analysis - Lung Function"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-width: 12
    fig-height: 9
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(ggplot2)
library(gtsummary)
library(gt)
library(pROC)
library(ggrepel)
library(childsds)
library(labelled)
library(Hmisc)
library(lmerTest)
library(multcomp)
library(tidyverse)
home_dir <- switch(
  Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
  "Windows" = "C:/Users/tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
  "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/CF Registry"
)
github_dir <- switch(
  Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/Tim/GitHub",
  "Linux" = "/home/tim/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
set.seed(1017)
```

# Participant characteristics at CFRD diagnosis by diagnostic method

```{r}
#| warning: false
#| message: false
# Import data and codebook
setwd(home_dir)
annual <- read.csv(
  "./Data_Raw/CFF19_Annualized_Del1.csv",
  na.strings = c("", "N/A")
)
data_dict <- readxl::read_excel("./Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
# Remove those with T1D or T2D
annual <- annual |> arrange(eDWID, ReviewYear)
annual <- annual[!annual$cfrd_status_annual %in% c(4, 5), ]
annual$cfrd_status_annual <- factor(
  annual$cfrd_status_annual,
  levels = 1:3,
  labels = c("CFRD-", "CFRD-", "CFRD+")
)
# Get mutations
demogs <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
races <- c(
  "White",
  "Black or African American",
  "American Indian or Alaska Native",
  "Asian",
  "Native Hawaiian or Other Pacific ",
  "Some other race"
)
demogs$Race <- apply(demogs, 1, function(r) {
  w <- which(r[paste0("Race", 1:6)] == 1)
  if (length(w) > 1) {
    return("More than one race")
  } else {
    return(races[w])
  }
})
demogs$Race <- factor(
  demogs$Race,
  levels = c(
    "White",
    "American Indian or Alaska Native",
    "Asian",
    "Black or African American",
    "More than one race",
    "Native Hawaiian or Other Pacific ",
    "Some other race"
  ),
  labels = c(
    "White",
    "Other race",
    "Other race",
    "Black or African American",
    "More than one race",
    "Other race",
    "Other race"
  )
)
demogs$Hispanicrace <- factor(
  demogs$Hispanicrace,
  levels = 1:2,
  labels = c("Yes", "No")
)
demogs <- demogs |> select(eDWID, Gender, MutClass, Race, Hispanicrace)
annual <- left_join(annual, demogs, by = join_by(eDWID))
# Variable formatting
# If A_supplement_feeding is 0, then A_feeding2 is also 0 because they are not
# on any supplemental feeding
# Include all kinds of tube feeds
annual$A_feeding2[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding3[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding4[annual$A_supplement_feeding == 0] <- 0
tube_fields <- c("A_feeding2", "A_feeding3", "A_feeding4")
annual$tube_feeding <-
  apply(annual[, tube_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Liver disease
liver_fields <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
annual$liver_disease <-
  apply(annual[, liver_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Get BMI categories
annual$bmi <- annual$A_weight / ((annual$A_height / 100)^2)
annual$bmi <- coalesce(annual$A_bmivalue, annual$bmi)
annual$bmi_perc_cdc <- sds(
  annual$bmi,
  age = annual$Age_YrEnd,
  sex = annual$Gender,
  male = "M",
  female = "F",
  item = "bmi",
  ref = cdc.ref,
  type = "perc"
) *
  100
annual$bmi_perc_who <- sds(
  annual$bmi,
  age = annual$Age_YrEnd,
  sex = annual$Gender,
  male = "M",
  female = "F",
  item = "bmi",
  ref = who.ref,
  type = "perc"
) *
  100
annual$bmi_perc <- coalesce(annual$bmi_perc_cdc, annual$bmi_perc_who)
annual$bmi_cat <- apply(annual, 1, function(r) {
  age <- as.numeric(r["Age_YrEnd"])
  gender <- as.character(r["Gender"])
  bmi <- as.numeric(r["bmi"])
  bmi_perc <- as.numeric(r["bmi_perc"])
  if (age <= 20) {
    return(cut(
      bmi_perc,
      c(-Inf, 10, 50, 85, Inf),
      right = F,
      labels = c("Undernourished", "At Risk", "Target/Goal", "Overweight")
    ))
  } else if (age > 20) {
    if (gender == "M") {
      return(cut(
        bmi,
        c(-Inf, 18.5, 23, 25, Inf),
        right = F,
        labels = c(
          "Undernourished",
          "At Risk",
          "Target/Goal",
          "Overweight"
        )
      ))
    } else if (gender == "F") {
      return(cut(
        bmi,
        c(-Inf, 18.5, 22, 25, Inf),
        right = F,
        labels = c(
          "Undernourished",
          "At Risk",
          "Target/Goal",
          "Overweight"
        )
      ))
    }
  }
})
annual$bmi_cat <- relevel(annual$bmi_cat, ref = "Target/Goal")
# Raw values for adults, percentiles for peds
annual$bmi_adult = annual$bmi
annual$bmi_adult[annual$Age_YrEnd <= 20] = NA
annual$bmi_perc = annual$bmi_perc
annual$bmi_perc[annual$Age_YrEnd > 20] = NA
# Labels
var_label(annual$bmi) = "BMI (all participants)"
var_label(annual$bmi_perc) = "BMI percentile (age <= 20)"
var_label(annual$bmi_adult) = "BMI value (age > 20)"
var_label(annual$bmi_cat) = "BMI category"
# Format columns
annual$MutClass <- factor(
  annual$MutClass,
  levels = c("1-3", "4-5", "Oth", "N/A"),
  labels = c("1-3", "4-5", "Oth", "Unknown")
)
annual$NumPulmExacerbation <- cut(
  annual$NumPulmExacerbation,
  labels = c("0", "1", "2", ">=3"),
  breaks = c(0, 1, 2, 3, Inf),
  right = F
)
# Age groups 10 - 18 inclusive, 19 - 30 inclusive, >= 31
annual$age_group <- cut(
  annual$Age_YrEnd,
  breaks = c(-Inf, 10, 19, 31, Inf),
  right = F
)
# Bugs
bugs <- c(
  "A_pseudomonasaeruginosa",
  "A_fungalyeast1",
  "A_pulmonarycomplications1",
  "A_mycocultureresults3",
  "A_cultureresults3"
)
# Yes/No variables
annual$cfrd_treatment_options1[annual$cfrd_treatment_options1 == 2] <- NA
annual$cfrd_treatment_options4[annual$cfrd_treatment_options1 == 4] <- NA
yn_vars <- c(
  "A_corticosteroids1",
  "A_feeding2",
  "A_Vx770",
  "A_VX809comb",
  bugs,
  "tube_feeding",
  "liver_disease",
  "cfrd_treatment",
  "cfrd_treatment_options1",
  "cfrd_treatment_options2",
  "cfrd_treatment_options3",
  "cfrd_treatment_options4"
)
annual[, yn_vars] <- lapply(
  annual[, yn_vars],
  factor,
  levels = 0:1,
  labels = c("No", "Yes")
)
annual$A_IsOnEnzymes <- factor(
  annual$A_IsOnEnzymes,
  levels = c(1, 0),
  labels = c("Yes", " No")
)
# FEV1 to categorical
annual$ppFEV1_cat <- cut(
  annual$A_FEV1_pct_predicted,
  breaks = c(0, 40, 80, 100, Inf),
  right = F
)
annual$ppFEV1_cat <- factor(
  annual$ppFEV1_cat,
  levels = c("[100,Inf)", "[80,100)", "[40,80)", "[0,40)")
)
# Males to reference group
annual$Gender <- factor(annual$Gender, levels = c("M", "F"))
# Hispanic no as reference
annual$Hispanicrace <- factor(annual$Hispanicrace, levels = c("No", "Yes"))
# Get first year of CFRD
first_cfrd <- annual |>
  filter(cfrd_status_annual == "CFRD+") |>
  group_by(eDWID) |>
  summarise(first_cfrd = min(ReviewYear), age_dx = min(Age_YrEnd))
annual <- left_join(annual, first_cfrd, by = join_by(eDWID))
# Year prior to first CFRD in registry - assume negative for CFRD. Assume all
# years following are CFRD+
annual$CFRD <- NA
annual$CFRD[annual$ReviewYear < annual$first_cfrd] <- "CFRD-"
annual$CFRD[annual$ReviewYear >= annual$first_cfrd] <- "CFRD+"
annual$CFRD[is.na(annual$first_cfrd)] <- "CFRD-"
annual$CFRD <- factor(annual$CFRD, levels = c("CFRD-", "CFRD+"))
# Add formatted columns for survival analysis
annual$tstop <- ymd(paste0(annual$ReviewYear, "-12-31"))
annual$tstart <- ymd(paste0(annual$ReviewYear, "-01-01"))
annual <- annual |>
  group_by(eDWID) |>
  mutate(entry_date = tstart[1]) |>
  ungroup()
# Remove those under 10
annual$age_group <- relevel(annual$age_group, ref = "[10,19)")
annual <- annual |> filter(Age_YrEnd != 0, Age_YrEnd >= 10)
annual$age_group <- droplevels(annual$age_group)
# Get first CFRD again for people who were diagnosed before age 10
first_cfrd <- annual |>
  filter(cfrd_status_annual == "CFRD+") |>
  group_by(eDWID) |>
  summarise(first_cfrd = min(ReviewYear), age_dx = min(Age_YrEnd))
annual$first_cfrd = NULL
annual <- left_join(annual, first_cfrd, by = join_by(eDWID))
# Number of visits
annual <- annual |> group_by(eDWID) |> mutate(years_with_data = n())
# List of covariates
vars <- c(
  "age_group",
  "Gender",
  "bmi_perc",
  "bmi_cat",
  "MutClass",
  "Race",
  "Hispanicrace",
  "A_IsOnEnzymes",
  "ppFEV1_cat",
  "tube_feeding",
  "NumPulmExacerbation",
  "A_pseudomonasaeruginosa",
  "A_fungalyeast1",
  "A_mycocultureresults3",
  "A_corticosteroids1",
  "cfrd_status_annual",
  "A_Vx770",
  "A_VX809comb",
  "liver_disease"
)
# Nice labels for plots and tables
labels <- list(
  cfrd_status_annual = "CFRD Status",
  MutClass = "Mutation class",
  Race = "Race",
  Hispanicrace = "Hispanic",
  Gender = "Sex",
  A_IsOnEnzymes = "Pancreatic enzyme use",
  ppFEV1_cat = "FEV1 percent predicted category",
  A_FEV1_pct_predicted = "FEV1 percent predicted",
  A_FVC_pct_predicted = "FVC percent predicted",
  tube_feeding = "Enteric feedings",
  liver_disease = "Liver disease",
  A_corticosteroids1 = "Use of corticosteroids",
  NumPulmExacerbation = "Admissions for PEx",
  age_group = "Age group",
  height_perc = "Height percentile",
  A_Vx770 = "Use of CFTR modulator Vx770",
  A_VX809comb = "Use of CFTR modulator Vx809 combination",
  A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
  A_fungalyeast1 = "Aspergillus (any species)",
  A_mycocultureresults3 = "Microorganisms in myco culture",
  A_cultureresults3 = "Microorganisms in culture",
  eDWID = "ID",
  tstart = "Interval start",
  tstop = "Interval stop",
  Age_YrEnd = "Age",
  cfrd_treatment = "Prescribed treatment for CFRD",
  cfrd_treatment_options1 = "Dietary change used to treat CFRD during the year",
  cfrd_treatment_options2 = "Oral hypoglycemic agents used to treat CFRD during the year",
  cfrd_treatment_options3 = "Intermittent insulin used to treat CFRD during the year",
  cfrd_treatment_options4 = "Chronic insulin used to treat CFRD during the year",
  years_with_data = "Years With Data"
)
label(annual[, which(colnames(annual) %in% names(labels))]) <-
  labels[colnames(annual)[which(colnames(annual) %in% names(labels))]]
# Per reviewer, out of X individuals with a diagnosis of CFRD, X number had an
# OGTT (and X had an A1c, but no OGTT) that year or the year prior.
# This is super slow but idc idc
cfrd_only <- apply(first_cfrd, 1, function(r) {
  id <- as.numeric(r["eDWID"])
  year <- as.numeric(r["first_cfrd"])
  d <- annual[annual$eDWID == id & annual$ReviewYear %in% c(year, year - 1), ]
  a1c <- as.numeric(last(d["A_hgba1c"], na_rm = TRUE))
  ogtt0hr <- as.numeric(last(d["A_ogttfast_bloodglucose"], na_rm = TRUE))
  ogtt2hr <- as.numeric(last(d["A_twohour_bloodglucose"], na_rm = TRUE))
  return(c(id, year, a1c, ogtt0hr, ogtt2hr))
})
cfrd_only <- data.frame(t(cfrd_only))
colnames(cfrd_only) <- c("id", "year", "a1c", "ogtt0hr", "ogtt2hr")
cfrd_only <- left_join(
  cfrd_only,
  annual |> mutate(eDWID = as.numeric(eDWID)),
  by = join_by(id == eDWID, year == ReviewYear)
)
cfrd_only$a1c_ogtt <- apply(cfrd_only, 1, function(r) {
  not_missing <- paste(
    which(!is.na(r[c("a1c", "ogtt0hr", "ogtt2hr")])),
    collapse = "_"
  )
  return(not_missing)
})
cfrd_only$a1c_ogtt <- factor(
  cfrd_only$a1c_ogtt,
  levels = c("", "1", "1_2", "1_2_3", "1_3", "2", "2_3", "3"),
  labels = c(
    "Neither",
    "HbA1c Only",
    "OGTT",
    "OGTT",
    "OGTT",
    "OGTT",
    "OGTT",
    "OGTT"
  )
)
cfrd_only |>
  select(
    a1c_ogtt,
    Age_YrEnd,
    A_weight,
    A_height,
    bmi_adult,
    bmi_perc,
    bmi_cat,
    A_FEV1_pct_predicted,
    A_FVC_pct_predicted,
    Race,
    Hispanicrace,
    MutClass,
    a1c,
    A_ogttfast_bloodglucose,
    A_twohour_bloodglucose,
    A_IsOnEnzymes,
    liver_disease,
    A_pseudomonasaeruginosa,
    A_fungalyeast1,
    A_mycocultureresults3,
    A_cultureresults3,
    cfrd_treatment,
    cfrd_treatment_options1,
    cfrd_treatment_options2,
    cfrd_treatment_options3,
    cfrd_treatment_options4,
    years_with_data
  ) |>
  tbl_summary(
    by = a1c_ogtt,
    missing_text = "Missing",
    statistic = list(all_continuous() ~ "{mean} ({sd})")
  ) |>
  add_overall() |>
  add_p(
    test = list(
      all_continuous() ~ "oneway.test",
      "years_with_data" ~ "kruskal.test"
    )
  ) |>
  modify_caption("**All Groups**")
cfrd_only |>
  filter(a1c_ogtt %in% c("OGTT", "HbA1c Only")) |>
  mutate(a1c_ogtt = droplevels(a1c_ogtt)) |>
  select(
    a1c_ogtt,
    Age_YrEnd,
    A_weight,
    A_height,
    A_bmivalue,
    A_bmipercentile,
    A_FEV1_pct_predicted,
    A_FVC_pct_predicted,
    Race,
    Hispanicrace,
    MutClass,
    a1c,
    A_ogttfast_bloodglucose,
    A_twohour_bloodglucose,
    A_IsOnEnzymes,
    liver_disease,
    A_pseudomonasaeruginosa,
    A_fungalyeast1,
    A_mycocultureresults3,
    A_cultureresults3,
    cfrd_treatment,
    cfrd_treatment_options1,
    cfrd_treatment_options2,
    cfrd_treatment_options3,
    cfrd_treatment_options4,
    years_with_data
  ) |>
  tbl_summary(
    by = a1c_ogtt,
    missing_text = "Missing",
    statistic = list(all_continuous() ~ "{mean} ({sd})")
  ) |>
  add_overall() |>
  add_p(test = list(all_continuous() ~ "oneway.test")) |>
  modify_caption("**OGTT vs. HbA1c Only**")
```

## FEV1 trends by diagnosis method

### LOESS-smoothed plot

```{r}
#| warning: false
#| message: false
# Plot by diagnosis type
mod_df = cfrd_only |>
  select(id, a1c_ogtt, Age_YrEnd) |>
  rename(eDWID = id)
mod_df = left_join(
  annual |> select(-Age_YrEnd) |> mutate(eDWID = as.numeric(eDWID)),
  mod_df,
  by = join_by(eDWID)
)
mod_df = mod_df |>
  drop_na(first_cfrd) |>
  group_by(eDWID) |>
  mutate(years_to_cfrd = ReviewYear - first_cfrd)
var_label(mod_df$years_to_cfrd) = "Years to CFRD"
var_label(mod_df$Age_YrEnd) = "Age at Diagnosis"
ggplot(
  mod_df |> filter(!is.na(a1c_ogtt)),
  aes(
    y = A_FEV1_pct_predicted,
    x = years_to_cfrd,
    color = a1c_ogtt,
    group = a1c_ogtt
  )
) +
  xlab("Years from CFRD") +
  xlim(c(-5, 5)) +
  geom_point(alpha = 0.1, shape = ".") +
  geom_smooth(formula = y ~ s(x, bs = "cr")) +
  theme_bw() +
  theme(legend.title = element_blank())
```

### Piecewise linear model results

```{r}
# Fit model - adjust for age at dx, sex, genotype (severe vs mild), pancreatic
# insufficiency, age at CF diagnosis (do we know this?), number of clinic
# visits/year, use of CFTR modulators
mod_df$x = mod_df$years_to_cfrd * (mod_df$years_to_cfrd >= 0)
label(mod_df$x) = "Change in Slope Post-CFRD"
label(mod_df$years_to_cfrd) = "Slope Pre-CFRD"
label(mod_df$a1c_ogtt) = "Diagnosis Method"
label(mod_df$Age_YrEnd) = "Age at Diagnosis"
# Write data so Christine can see missing data
t = mod_df |>
  select(
    eDWID,
    years_to_cfrd,
    A_FEV1_pct_predicted,
    A_FVC_pct_predicted,
    a1c_ogtt,
    Age_YrEnd,
    Gender,
    MutClass,
    A_IsOnEnzymes,
    A_Vx770,
    A_VX809comb,
    cfrd_treatment,
    cfrd_treatment_options1,
    cfrd_treatment_options2,
    cfrd_treatment_options3,
    cfrd_treatment_options4
  ) |>
  drop_na(A_FEV1_pct_predicted)
write.csv(t, "./Data_Clean/fev1_model_data.csv", row.names = F, na = "")
# Fit the model
m <- lmer(
  A_FEV1_pct_predicted ~ years_to_cfrd *
    a1c_ogtt +
    x * a1c_ogtt +
    Age_YrEnd +
    Gender +
    MutClass +
    A_IsOnEnzymes +
    A_Vx770 +
    A_VX809comb +
    (1 | eDWID),
  data = mod_df
)
# Population-averaged plot
pop_dat <- data.frame(expand.grid(
  years_to_cfrd = -5:5,
  a1c_ogtt = unique(mod_df$a1c_ogtt),
  Age_YrEnd = mean(mod_df$Age_YrEnd, na.rm = T),
  Gender = "M",
  MutClass = "1-3",
  A_IsOnEnzymes = "Yes",
  A_Vx770 = "No",
  A_VX809comb = "No"
))
pop_dat$x <- (pop_dat$years_to_cfrd >= 0) * pop_dat$years_to_cfrd
pred <- predict(m, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(
  pop_dat,
  aes(x = years_to_cfrd, y = pred, group = a1c_ogtt, colour = a1c_ogtt)
) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from CFRD") +
  ylab("FEV1 % Predicted") +
  theme_classic(base_size = 15)
# Contrasts
neither_pre <- rep(0, nrow(summary(m)$coefficients))
neither_change <- rep(0, nrow(summary(m)$coefficients))
neither_pre[which(rownames(summary(m)$coefficients) == "years_to_cfrd")] = 1
neither_change[which(rownames(summary(m)$coefficients) == "x")] = 1
a1c_pre <- rep(0, nrow(summary(m)$coefficients))
a1c_change <- rep(0, nrow(summary(m)$coefficients))
a1c_pre[which(
  rownames(summary(m)$coefficients) %in%
    c("years_to_cfrd", "years_to_cfrd:a1c_ogttHbA1c Only")
)] = 1
a1c_change[which(
  rownames(summary(m)$coefficients) %in%
    c("x", "a1c_ogttHbA1c Only:x")
)] = 1
ogtt_pre <- rep(0, nrow(summary(m)$coefficients))
ogtt_change <- rep(0, nrow(summary(m)$coefficients))
ogtt_pre[which(
  rownames(summary(m)$coefficients) %in%
    c("years_to_cfrd", "years_to_cfrd:a1c_ogttOGTT")
)] = 1
ogtt_change[which(
  rownames(summary(m)$coefficients) %in%
    c("x", "a1c_ogttOGTT:x")
)] = 1
contr <- rbind(
  "HbA1c Only Pre-CFRD" = a1c_pre,
  "HbA1c Only Post-CFRD" = a1c_pre + a1c_change,
  "OGTT Pre-CFRD" = ogtt_pre,
  "OGTT Post-CFRD" = ogtt_pre + ogtt_change,
  "Neither Pre-CFRD" = neither_pre,
  "Neither Post-CFRD" = neither_pre + neither_change
)
change_mod_slopes <- glht(m, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint |>
  as.data.frame() |>
  gt(caption = "Slopes by Diagnosis Method", rownames_to_stub = T) |>
  fmt_number(decimals = 2)
# Model table
m |> tbl_regression() |> add_n(location = "level")
```

### Categorical model results

## FVC trends by diagnosis method

### LOESS-smoothed plot

```{r} 
ggplot(
  mod_df |> filter(!is.na(a1c_ogtt)),
  aes(
    y = A_FVC_pct_predicted,
    x = years_to_cfrd,
    color = a1c_ogtt,
    group = a1c_ogtt
  )
) +
  xlab("Years from CFRD") +
  xlim(c(-5, 5)) +
  geom_point(alpha = 0.1, shape = ".") +
  geom_smooth(formula = y ~ s(x, bs = "cr")) +
  theme_bw() +
  theme(legend.title = element_blank())
```

### Piecewise linear model results

```{r}
# Fit model - adjust for age at dx, sex, genotype (severe vs mild), pancreatic
# insufficiency, age at CF diagnosis (do we know this?), number of clinic
# visits/year, use of CFTR modulators
mod_df$x = mod_df$years_to_cfrd * (mod_df$years_to_cfrd >= 0)
label(mod_df$x) = "Change in Slope Post-CFRD"
label(mod_df$years_to_cfrd) = "Slope Pre-CFRD"
label(mod_df$a1c_ogtt) = "Diagnosis Method"
label(mod_df$Age_YrEnd) = "Age at Diagnosis"
m <- lmer(
  A_FVC_pct_predicted ~ years_to_cfrd *
    a1c_ogtt +
    x * a1c_ogtt +
    Age_YrEnd +
    Gender +
    MutClass +
    A_IsOnEnzymes +
    A_Vx770 +
    A_VX809comb +
    (1 | eDWID),
  data = mod_df
)
# Population-averaged plot
pop_dat <- data.frame(expand.grid(
  years_to_cfrd = -5:5,
  a1c_ogtt = unique(mod_df$a1c_ogtt),
  Age_YrEnd = mean(mod_df$Age_YrEnd, na.rm = T),
  Gender = "M",
  MutClass = "1-3",
  A_IsOnEnzymes = "Yes",
  A_Vx770 = "No",
  A_VX809comb = "No"
))
pop_dat$x <- (pop_dat$years_to_cfrd >= 0) * pop_dat$years_to_cfrd
pred <- predict(m, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(
  pop_dat,
  aes(x = years_to_cfrd, y = pred, group = a1c_ogtt, colour = a1c_ogtt)
) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from CFRD") +
  ylab("FVC % Predicted") +
  theme_classic(base_size = 15)
# Contrasts
neither_pre <- rep(0, nrow(summary(m)$coefficients))
neither_change <- rep(0, nrow(summary(m)$coefficients))
neither_pre[which(rownames(summary(m)$coefficients) == "years_to_cfrd")] = 1
neither_change[which(rownames(summary(m)$coefficients) == "x")] = 1
a1c_pre <- rep(0, nrow(summary(m)$coefficients))
a1c_change <- rep(0, nrow(summary(m)$coefficients))
a1c_pre[which(
  rownames(summary(m)$coefficients) %in%
    c("years_to_cfrd", "years_to_cfrd:a1c_ogttHbA1c Only")
)] = 1
a1c_change[which(
  rownames(summary(m)$coefficients) %in%
    c("x", "a1c_ogttHbA1c Only:x")
)] = 1
ogtt_pre <- rep(0, nrow(summary(m)$coefficients))
ogtt_change <- rep(0, nrow(summary(m)$coefficients))
ogtt_pre[which(
  rownames(summary(m)$coefficients) %in%
    c("years_to_cfrd", "years_to_cfrd:a1c_ogttOGTT")
)] = 1
ogtt_change[which(
  rownames(summary(m)$coefficients) %in%
    c("x", "a1c_ogttOGTT:x")
)] = 1
contr <- rbind(
  "HbA1c Only Pre-CFRD" = a1c_pre,
  "HbA1c Only Post-CFRD" = a1c_pre + a1c_change,
  "OGTT Pre-CFRD" = ogtt_pre,
  "OGTT Post-CFRD" = ogtt_pre + ogtt_change,
  "Neither Pre-CFRD" = neither_pre,
  "Neither Post-CFRD" = neither_pre + neither_change
)
change_mod_slopes <- glht(m, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint |>
  as.data.frame() |>
  gt(caption = "Slopes by Diagnosis Method", rownames_to_stub = T) |>
  fmt_number(decimals = 2)
# Model table
m |> tbl_regression() |> add_n(location = "level")
```

### 2025-12-05 Christine Chan/CF Registry/cf_registry_lung_function_analysis.qmd
fixed lung function analysis

@@
 library(lmerTest)
+library(emmeans)
 library(multcomp)
 library(tidyverse)
-home_dir <- switch(
-  Sys.info()["sysname"],
-  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
-  "Windows" = "C:/Users/tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
-  "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/CF Registry"
+library(parallel)
+library(knitr)
+base_dir <- switch(
+  Sys.info()["nodename"],
+  "togo" = "/home/tvigers/Documents/Data",
+  "Tims-MacBook-Air.local" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers",
+  "Tims-Mac-mini.local" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers",
+  "Mac" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers"
 )
-github_dir <- switch(
-  Sys.info()["sysname"],
-  "Darwin" = "/Users/timvigers/GitHub",
-  "Windows" = "C:/Users/Tim/GitHub",
-  "Linux" = "/home/tim/GitHub"
+data_dir <- switch(
+  Sys.info()["nodename"],
+  "togo" = "/CF/Christine Chan/CF Registry",
+  "Tims-MacBook-Air.local" = "/CF/Christine Chan/CF Registry",
+  "Tims-Mac-mini.local" = "/CF/Christine Chan/CF Registry",
+  "Mac" = "/CF/Christine Chan/CF Registry"
 )
+home_dir <- paste0(base_dir, data_dir)
 knitr::opts_knit$set(root.dir = home_dir)
-set.seed(1017)
 ```
-
-# Participant characteristics at CFRD diagnosis by diagnostic method
 
@@
 )
-data_dict <- readxl::read_excel("./Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
-data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
 # Remove those with T1D or T2D
@@
 annual <- left_join(annual, demogs, by = join_by(eDWID))
-# Variable formatting
-# If A_supplement_feeding is 0, then A_feeding2 is also 0 because they are not
-# on any supplemental feeding
-# Include all kinds of tube feeds
-annual$A_feeding2[annual$A_supplement_feeding == 0] <- 0
-annual$A_feeding3[annual$A_supplement_feeding == 0] <- 0
-annual$A_feeding4[annual$A_supplement_feeding == 0] <- 0
-tube_fields <- c("A_feeding2", "A_feeding3", "A_feeding4")
-annual$tube_feeding <-
-  apply(annual[, tube_fields], 1, function(r) {
-    if (all(is.na(r))) {
-      return(NA)
-    } else {
-      return(max(as.numeric(r), na.rm = T))
-    }
-  })
 # Liver disease
@@
 # Raw values for adults, percentiles for peds
-annual$bmi_adult = annual$bmi
-annual$bmi_adult[annual$Age_YrEnd <= 20] = NA
-annual$bmi_perc = annual$bmi_perc
-annual$bmi_perc[annual$Age_YrEnd > 20] = NA
+annual$bmi_adult <- annual$bmi
+annual$bmi_adult[annual$Age_YrEnd <= 20] <- NA
+annual$bmi_perc <- annual$bmi_perc
+annual$bmi_perc[annual$Age_YrEnd > 20] <- NA
 # Labels
@@
 )
-# Age groups 10 - 18 inclusive, 19 - 30 inclusive, >= 31
-annual$age_group <- cut(
-  annual$Age_YrEnd,
-  breaks = c(-Inf, 10, 19, 31, Inf),
-  right = F
-)
 # Bugs
@@
 annual$cfrd_treatment_options1[annual$cfrd_treatment_options1 == 2] <- NA
-annual$cfrd_treatment_options4[annual$cfrd_treatment_options1 == 4] <- NA
+annual$cfrd_treatment_options4[annual$cfrd_treatment_options4 == 4] <- NA
 yn_vars <- c(
@@
   "A_VX809comb",
+  "liver_disease",
   bugs,
-  "tube_feeding",
-  "liver_disease",
-  "cfrd_treatment",
-  "cfrd_treatment_options1",
   "cfrd_treatment_options2",
@@
   labels = c("Yes", " No")
-)
-# FEV1 to categorical
-annual$ppFEV1_cat <- cut(
-  annual$A_FEV1_pct_predicted,
-  breaks = c(0, 40, 80, 100, Inf),
-  right = F
-)
-annual$ppFEV1_cat <- factor(
-  annual$ppFEV1_cat,
-  levels = c("[100,Inf)", "[80,100)", "[40,80)", "[0,40)")
 )
@@
 annual$CFRD <- factor(annual$CFRD, levels = c("CFRD-", "CFRD+"))
-# Add formatted columns for survival analysis
-annual$tstop <- ymd(paste0(annual$ReviewYear, "-12-31"))
-annual$tstart <- ymd(paste0(annual$ReviewYear, "-01-01"))
-annual <- annual |>
-  group_by(eDWID) |>
-  mutate(entry_date = tstart[1]) |>
-  ungroup()
-# Remove those under 10
-annual$age_group <- relevel(annual$age_group, ref = "[10,19)")
-annual <- annual |> filter(Age_YrEnd != 0, Age_YrEnd >= 10)
-annual$age_group <- droplevels(annual$age_group)
-# Get first CFRD again for people who were diagnosed before age 10
-first_cfrd <- annual |>
-  filter(cfrd_status_annual == "CFRD+") |>
-  group_by(eDWID) |>
-  summarise(first_cfrd = min(ReviewYear), age_dx = min(Age_YrEnd))
-annual$first_cfrd = NULL
-annual <- left_join(annual, first_cfrd, by = join_by(eDWID))
-# Number of visits
-annual <- annual |> group_by(eDWID) |> mutate(years_with_data = n())
-# List of covariates
-vars <- c(
-  "age_group",
-  "Gender",
-  "bmi_perc",
-  "bmi_cat",
-  "MutClass",
-  "Race",
-  "Hispanicrace",
-  "A_IsOnEnzymes",
-  "ppFEV1_cat",
-  "tube_feeding",
-  "NumPulmExacerbation",
-  "A_pseudomonasaeruginosa",
-  "A_fungalyeast1",
-  "A_mycocultureresults3",
-  "A_corticosteroids1",
-  "cfrd_status_annual",
-  "A_Vx770",
-  "A_VX809comb",
-  "liver_disease"
-)
-# Nice labels for plots and tables
-labels <- list(
-  cfrd_status_annual = "CFRD Status",
-  MutClass = "Mutation class",
-  Race = "Race",
-  Hispanicrace = "Hispanic",
-  Gender = "Sex",
-  A_IsOnEnzymes = "Pancreatic enzyme use",
-  ppFEV1_cat = "FEV1 percent predicted category",
-  A_FEV1_pct_predicted = "FEV1 percent predicted",
-  A_FVC_pct_predicted = "FVC percent predicted",
-  tube_feeding = "Enteric feedings",
-  liver_disease = "Liver disease",
-  A_corticosteroids1 = "Use of corticosteroids",
-  NumPulmExacerbation = "Admissions for PEx",
-  age_group = "Age group",
-  height_perc = "Height percentile",
-  A_Vx770 = "Use of CFTR modulator Vx770",
-  A_VX809comb = "Use of CFTR modulator Vx809 combination",
-  A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
-  A_fungalyeast1 = "Aspergillus (any species)",
-  A_mycocultureresults3 = "Microorganisms in myco culture",
-  A_cultureresults3 = "Microorganisms in culture",
-  eDWID = "ID",
-  tstart = "Interval start",
-  tstop = "Interval stop",
-  Age_YrEnd = "Age",
-  cfrd_treatment = "Prescribed treatment for CFRD",
-  cfrd_treatment_options1 = "Dietary change used to treat CFRD during the year",
-  cfrd_treatment_options2 = "Oral hypoglycemic agents used to treat CFRD during the year",
-  cfrd_treatment_options3 = "Intermittent insulin used to treat CFRD during the year",
-  cfrd_treatment_options4 = "Chronic insulin used to treat CFRD during the year",
-  years_with_data = "Years With Data"
-)
-label(annual[, which(colnames(annual) %in% names(labels))]) <-
-  labels[colnames(annual)[which(colnames(annual) %in% names(labels))]]
-# Per reviewer, out of X individuals with a diagnosis of CFRD, X number had an
-# OGTT (and X had an A1c, but no OGTT) that year or the year prior.
-# This is super slow but idc idc
-cfrd_only <- apply(first_cfrd, 1, function(r) {
+# Check whether they had an OGTT and/or HbA1c the year of first CFRD or the year
+# prior. This is super slow but idc idc.
+cl <- makeCluster(type = "FORK", 8)
+cfrd_only <- parApply(
+  cl,
+  first_cfrd,
+  1,
+  function(r) {
     id <- as.numeric(r["eDWID"])
@@
     ogtt2hr <- as.numeric(last(d["A_twohour_bloodglucose"], na_rm = TRUE))
-  return(c(id, year, a1c, ogtt0hr, ogtt2hr))
-})
+    d <- annual[annual$eDWID == id & annual$ReviewYear %in% c(year, year + 1), ]
+    treatment = any(
+      d[, c(
+        "cfrd_treatment_options2",
+        "cfrd_treatment_options3",
+        "cfrd_treatment_options4"
+      )] ==
+        "Yes",
+      na.rm = TRUE
+    )
+    return(c(id, year, a1c, ogtt0hr, ogtt2hr, treatment))
+  }
+)
+stopCluster(cl)
 cfrd_only <- data.frame(t(cfrd_only))
-colnames(cfrd_only) <- c("id", "year", "a1c", "ogtt0hr", "ogtt2hr")
-cfrd_only <- left_join(
-  cfrd_only,
-  annual |> mutate(eDWID = as.numeric(eDWID)),
-  by = join_by(id == eDWID, year == ReviewYear)
+colnames(cfrd_only) <- c(
+  "eDWID",
+  "dx_year",
+  "a1c",
+  "ogtt0hr",
+  "ogtt2hr",
+  "treatment"
 )
+# cfrd_only_store <- cfrd_only
 cfrd_only$a1c_ogtt <- apply(cfrd_only, 1, function(r) {
@@
 )
-cfrd_only |>
+var_label(cfrd_only$a1c_ogtt) <- "Diagnosis Method"
+cfrd_only$treatment = factor(
+  cfrd_only$treatment,
+  levels = 0:1,
+  labels = c("No Treatment", "Treatment")
+)
+var_label(cfrd_only$treatment) = "CFRD Treatment Started Within 1 Year?"
+# Calculate number of visits (just for table 1)
+n_vis = annual |>
+  group_by(eDWID) |>
+  summarise(years_with_data = n())
+cfrd_only <- left_join(cfrd_only, n_vis, by = join_by(eDWID))
+```
+
+# Diagnosed with CFRD between 2004 - 2018
+
+## Participant characteristics at diagnosis
+
+```{r}
+# Create a model dataframe
+mod_df = inner_join(
+  annual,
+  cfrd_only |> filter(dx_year >= 2004),
+  by = join_by(eDWID)
+) |>
+  drop_na(a1c_ogtt) |>
+  mutate(years_to_cfrd = ReviewYear - first_cfrd) |>
+  filter(years_to_cfrd <= 5 & years_to_cfrd >= -5) |>
+  mutate(years_to_cfrd = factor(years_to_cfrd))
+var_label(mod_df$years_to_cfrd) = "Years From CFRD Dx"
+# Write data for Christine to look at
+mod_df |>
+  select(
+    eDWID,
+    Race,
+    Hispanicrace,
+    MutClass,
+    ReviewYear,
+    Age_YrEnd,
+    CFRD,
+    first_cfrd,
+    age_dx,
+    A_weight,
+    A_height,
+    bmi_adult,
+    bmi_perc,
+    bmi_cat,
+    A_FEV1_pct_predicted,
+    A_FVC_pct_predicted,
+    a1c_ogtt,
+    A_hgba1c,
+    A_ogttfast_bloodglucose,
+    A_twohour_bloodglucose,
+    A_IsOnEnzymes,
+    liver_disease,
+    A_pseudomonasaeruginosa,
+    A_fungalyeast1,
+    A_mycocultureresults3,
+    A_cultureresults3,
+    treatment,
+    cfrd_treatment_options2,
+    cfrd_treatment_options3,
+    cfrd_treatment_options4,
+    years_with_data
+  ) %>%
+  write.csv(
+    .,
+    file = "./Data_Clean/fev1_model_data_2004.csv",
+    row.names = FALSE,
+    na = ""
+  )
+# Tables
+mod_df |>
+  filter(ReviewYear == first_cfrd) |>
   select(
     a1c_ogtt,
-    Age_YrEnd,
+    age_dx,
     A_weight,
@@
     A_cultureresults3,
-    cfrd_treatment,
-    cfrd_treatment_options1,
+    treatment,
     cfrd_treatment_options2,
@@
   modify_caption("**All Groups**")
-cfrd_only |>
-  filter(a1c_ogtt %in% c("OGTT", "HbA1c Only")) |>
+mod_df |>
+  filter(ReviewYear == first_cfrd, a1c_ogtt %in% c("OGTT", "HbA1c Only")) |>
   mutate(a1c_ogtt = droplevels(a1c_ogtt)) |>
@@
     a1c_ogtt,
-    Age_YrEnd,
+    age_dx,
     A_weight,
@@
 
-## FEV1 trends by diagnosis method
+## FEV1
 
@@
 # Plot by diagnosis type
-mod_df = cfrd_only |>
-  select(id, a1c_ogtt, Age_YrEnd) |>
-  rename(eDWID = id)
-mod_df = left_join(
-  annual |> select(-Age_YrEnd) |> mutate(eDWID = as.numeric(eDWID)),
+ggplot(
   mod_df,
-  by = join_by(eDWID)
-)
-mod_df = mod_df |>
-  drop_na(first_cfrd) |>
-  group_by(eDWID) |>
-  mutate(years_to_cfrd = ReviewYear - first_cfrd)
-var_label(mod_df$years_to_cfrd) = "Years to CFRD"
-var_label(mod_df$Age_YrEnd) = "Age at Diagnosis"
-ggplot(
-  mod_df |> filter(!is.na(a1c_ogtt)),
   aes(
@@
   xlab("Years from CFRD") +
-  xlim(c(-5, 5)) +
-  geom_point(alpha = 0.1, shape = ".") +
-  geom_smooth(formula = y ~ s(x, bs = "cr")) +
-  theme_bw() +
-  theme(legend.title = element_blank())
+  geom_smooth() +
+  theme_bw()
 ```
 
-### Piecewise linear model results
+### Linear model results
+
+#### Not adjusted for treatment status
+
+Adjusted for age at diagnosis, sex, mutation class, and pancreatic enzyme use.
 
 ```{r}
-# Fit model - adjust for age at dx, sex, genotype (severe vs mild), pancreatic
-# insufficiency, age at CF diagnosis (do we know this?), number of clinic
-# visits/year, use of CFTR modulators
-mod_df$x = mod_df$years_to_cfrd * (mod_df$years_to_cfrd >= 0)
-label(mod_df$x) = "Change in Slope Post-CFRD"
-label(mod_df$years_to_cfrd) = "Slope Pre-CFRD"
-label(mod_df$a1c_ogtt) = "Diagnosis Method"
-label(mod_df$Age_YrEnd) = "Age at Diagnosis"
-# Write data so Christine can see missing data
-t = mod_df |>
-  select(
-    eDWID,
-    years_to_cfrd,
-    A_FEV1_pct_predicted,
-    A_FVC_pct_predicted,
-    a1c_ogtt,
-    Age_YrEnd,
-    Gender,
-    MutClass,
-    A_IsOnEnzymes,
-    A_Vx770,
-    A_VX809comb,
-    cfrd_treatment,
-    cfrd_treatment_options1,
-    cfrd_treatment_options2,
-    cfrd_treatment_options3,
-    cfrd_treatment_options4
-  ) |>
-  drop_na(A_FEV1_pct_predicted)
-write.csv(t, "./Data_Clean/fev1_model_data.csv", row.names = F, na = "")
 # Fit the model
@@
     a1c_ogtt +
-    x * a1c_ogtt +
-    Age_YrEnd +
+    age_dx +
     Gender +
@@
     A_IsOnEnzymes +
-    A_Vx770 +
-    A_VX809comb +
     (1 | eDWID),
@@
 )
-# Population-averaged plot
-pop_dat <- data.frame(expand.grid(
-  years_to_cfrd = -5:5,
-  a1c_ogtt = unique(mod_df$a1c_ogtt),
-  Age_YrEnd = mean(mod_df$Age_YrEnd, na.rm = T),
-  Gender = "M",
-  MutClass = "1-3",
-  A_IsOnEnzymes = "Yes",
-  A_Vx770 = "No",
-  A_VX809comb = "No"
-))
-pop_dat$x <- (pop_dat$years_to_cfrd >= 0) * pop_dat$years_to_cfrd
-pred <- predict(m, re.form = NA, newdata = pop_dat, se.fit = T)
-pop_dat$pred <- pred$fit
-pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
-pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+t3 = car::Anova(m, type = 'III', contrasts = "contr.sum")
+emm = emmeans(m, specs = ~ years_to_cfrd + a1c_ogtt, pbkrtest.limit = 1e5)
+emm = data.frame(emm)
 ggplot(
-  pop_dat,
-  aes(x = years_to_cfrd, y = pred, group = a1c_ogtt, colour = a1c_ogtt)
+  emm,
+  aes(x = years_to_cfrd, y = emmean, group = a1c_ogtt, color = a1c_ogtt)
 ) +
   geom_point() +
-  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.25) +
   geom_line() +
-  xlab("Years from CFRD") +
-  ylab("FEV1 % Predicted") +
-  theme_classic(base_size = 15)
-# Contrasts
-neither_pre <- rep(0, nrow(summary(m)$coefficients))
-neither_change <- rep(0, nrow(summary(m)$coefficients))
-neither_pre[which(rownames(summary(m)$coefficients) == "years_to_cfrd")] = 1
-neither_change[which(rownames(summary(m)$coefficients) == "x")] = 1
-a1c_pre <- rep(0, nrow(summary(m)$coefficients))
-a1c_change <- rep(0, nrow(summary(m)$coefficients))
-a1c_pre[which(
-  rownames(summary(m)$coefficients) %in%
-    c("years_to_cfrd", "years_to_cfrd:a1c_ogttHbA1c Only")
-)] = 1
-a1c_change[which(
-  rownames(summary(m)$coefficients) %in%
-    c("x", "a1c_ogttHbA1c Only:x")
-)] = 1
-ogtt_pre <- rep(0, nrow(summary(m)$coefficients))
-ogtt_change <- rep(0, nrow(summary(m)$coefficients))
-ogtt_pre[which(
-  rownames(summary(m)$coefficients) %in%
-    c("years_to_cfrd", "years_to_cfrd:a1c_ogttOGTT")
-)] = 1
-ogtt_change[which(
-  rownames(summary(m)$coefficients) %in%
-    c("x", "a1c_ogttOGTT:x")
-)] = 1
-contr <- rbind(
-  "HbA1c Only Pre-CFRD" = a1c_pre,
-  "HbA1c Only Post-CFRD" = a1c_pre + a1c_change,
-  "OGTT Pre-CFRD" = ogtt_pre,
-  "OGTT Post-CFRD" = ogtt_pre + ogtt_change,
-  "Neither Pre-CFRD" = neither_pre,
-  "Neither Post-CFRD" = neither_pre + neither_change
-)
-change_mod_slopes <- glht(m, linfct = contr)
-change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
-change_mod_slopes_ci$confint |>
-  as.data.frame() |>
-  gt(caption = "Slopes by Diagnosis Method", rownames_to_stub = T) |>
-  fmt_number(decimals = 2)
-# Model table
+  xlab("Years From CFRD Dx") +
+  ylab("Estimated Mean ppFEV1") +
+  labs(color = "Diagnosis Method") +
+  theme_bw()
+# Model tables
+kable(emm, digits = 3)
 m |> tbl_regression() |> add_n(location = "level")
+t3
 ```
 
-### Categorical model results
+The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[nrow(t3),"Pr(>Chisq)"], eps = 0.001)`).
 
-## FVC trends by diagnosis method
+#### Adjusted for treatment status
+
+Adjusted for age at diagnosis, sex, mutation class, pancreatic enzyme use, and CFRD treatment.
+
+```{r}
+# Fit the model
+m <- lmer(
+  A_FEV1_pct_predicted ~ years_to_cfrd *
+    a1c_ogtt +
+    age_dx +
+    Gender +
+    MutClass +
+    A_IsOnEnzymes +
+    treatment +
+    (1 | eDWID),
+  data = mod_df
+)
+t3 = car::Anova(m, type = 'III', contrasts = "contr.sum")
+emm = emmeans(m, specs = ~ years_to_cfrd + a1c_ogtt, pbkrtest.limit = 1e5)
+emm = data.frame(emm)
+ggplot(
+  emm,
+  aes(x = years_to_cfrd, y = emmean, group = a1c_ogtt, color = a1c_ogtt)
+) +
+  geom_point() +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.25) +
+  geom_line() +
+  xlab("Years From CFRD Dx") +
+  ylab("Estimated Mean ppFEV1") +
+  labs(color = "Diagnosis Method") +
+  theme_bw()
+# Model tables
+kable(emm, digits = 3)
+m |> tbl_regression() |> add_n(location = "level")
+t3
+```
+
+The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[nrow(t3),"Pr(>Chisq)"], eps = 0.001)`).
+
+## FVC
 
@@
 ```{r}
+#| warning: false
+#| message: false
+# Plot by diagnosis type
 ggplot(
-  mod_df |> filter(!is.na(a1c_ogtt)),
+  mod_df,
   aes(
@@
   xlab("Years from CFRD") +
-  xlim(c(-5, 5)) +
-  geom_point(alpha = 0.1, shape = ".") +
-  geom_smooth(formula = y ~ s(x, bs = "cr")) +
-  theme_bw() +
-  theme(legend.title = element_blank())
+  geom_smooth() +
+  theme_bw()
 ```
 
-### Piecewise linear model results
+### Linear model results
+
+#### Not adjusted for treatment status
+
+Adjusted for age at diagnosis, sex, mutation class, and pancreatic enzyme use.
 
 ```{r}
-# Fit model - adjust for age at dx, sex, genotype (severe vs mild), pancreatic
-# insufficiency, age at CF diagnosis (do we know this?), number of clinic
-# visits/year, use of CFTR modulators
-mod_df$x = mod_df$years_to_cfrd * (mod_df$years_to_cfrd >= 0)
-label(mod_df$x) = "Change in Slope Post-CFRD"
-label(mod_df$years_to_cfrd) = "Slope Pre-CFRD"
-label(mod_df$a1c_ogtt) = "Diagnosis Method"
-label(mod_df$Age_YrEnd) = "Age at Diagnosis"
+# Fit the model
 m <- lmer(
@@
 )
-# Population-averaged plot
-pop_dat <- data.frame(expand.grid(
-  years_to_cfrd = -5:5,
-  a1c_ogtt = unique(mod_df$a1c_ogtt),
-  Age_YrEnd = mean(mod_df$Age_YrEnd, na.rm = T),
-  Gender = "M",
-  MutClass = "1-3",
-  A_IsOnEnzymes = "Yes",
-  A_Vx770 = "No",
-  A_VX809comb = "No"
-))
-pop_dat$x <- (pop_dat$years_to_cfrd >= 0) * pop_dat$years_to_cfrd
-pred <- predict(m, re.form = NA, newdata = pop_dat, se.fit = T)
-pop_dat$pred <- pred$fit
-pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
-pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
+t3 = car::Anova(m, type = 'III', contrasts = "contr.sum")
+emm = emmeans(m, specs = ~ years_to_cfrd + a1c_ogtt, pbkrtest.limit = 1e5)
+emm = data.frame(emm)
 ggplot(
-  pop_dat,
-  aes(x = years_to_cfrd, y = pred, group = a1c_ogtt, colour = a1c_ogtt)
+  emm,
+  aes(x = years_to_cfrd, y = emmean, group = a1c_ogtt, color = a1c_ogtt)
 ) +
   geom_point() +
-  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.25) +
   geom_line() +
+  xlab("Years From CFRD Dx") +
+  ylab("Estimated Mean ppFVC") +
+  labs(color = "Diagnosis Method") +
+  theme_bw()
+# Model tables
+kable(emm, digits = 3)
+m |> tbl_regression() |> add_n(location = "level")
+t3
+```
+
+The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[nrow(t3),"Pr(>Chisq)"], eps = 0.001)`).
+
+#### Adjusted for treatment status
+
+Adjusted for age at diagnosis, sex, mutation class, pancreatic enzyme use, and CFRD treatment.
+
+```{r}
+# Fit the model
+m <- lmer(
+  A_FVC_pct_predicted ~ years_to_cfrd *
+    a1c_ogtt +
+    age_dx +
+    Gender +
+    MutClass +
+    A_IsOnEnzymes +
+    treatment +
+    (1 | eDWID),
+  data = mod_df
+)
+t3 = car::Anova(m, type = 'III', contrasts = "contr.sum")
+emm = emmeans(m, specs = ~ years_to_cfrd + a1c_ogtt, pbkrtest.limit = 1e5)
+emm = data.frame(emm)
+ggplot(
+  emm,
+  aes(x = years_to_cfrd, y = emmean, group = a1c_ogtt, color = a1c_ogtt)
+) +
+  geom_point() +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.25) +
+  geom_line() +
+  xlab("Years From CFRD Dx") +
+  ylab("Estimated Mean ppFVC") +
+  labs(color = "Diagnosis Method") +
+  theme_bw()
+# Model tables
+kable(emm, digits = 3)
+m |> tbl_regression() |> add_n(location = "level")
+t3
+```
+
+The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[nrow(t3),"Pr(>Chisq)"], eps = 0.001)`).
+
+# Diagnosed with CFRD between 2010 - 2018
+
+```{r}
+# Create a model dataframe
+mod_df = inner_join(
+  annual,
+  cfrd_only |> filter(dx_year >= 2010),
+  by = join_by(eDWID)
+) |>
+  drop_na(a1c_ogtt) |>
+  mutate(years_to_cfrd = ReviewYear - first_cfrd) |>
+  filter(years_to_cfrd <= 5 & years_to_cfrd >= -5) |>
+  mutate(years_to_cfrd = factor(years_to_cfrd))
+var_label(mod_df$years_to_cfrd) = "Years From CFRD Dx"
+```
+
+## Participant characteristics at diagnosis
+
+```{r}
+# Write data for Christine to look at
+mod_df |>
+  select(
+    eDWID,
+    Race,
+    Hispanicrace,
+    MutClass,
+    ReviewYear,
+    Age_YrEnd,
+    CFRD,
+    first_cfrd,
+    age_dx,
+    A_weight,
+    A_height,
+    bmi_adult,
+    bmi_perc,
+    bmi_cat,
+    A_FEV1_pct_predicted,
+    A_FVC_pct_predicted,
+    a1c_ogtt,
+    A_hgba1c,
+    A_ogttfast_bloodglucose,
+    A_twohour_bloodglucose,
+    A_IsOnEnzymes,
+    liver_disease,
+    A_pseudomonasaeruginosa,
+    A_fungalyeast1,
+    A_mycocultureresults3,
+    A_cultureresults3,
+    treatment,
+    cfrd_treatment_options2,
+    cfrd_treatment_options3,
+    cfrd_treatment_options4,
+    years_with_data
+  ) %>%
+  write.csv(
+    .,
+    file = "./Data_Clean/fev1_model_data_2010.csv",
+    row.names = FALSE,
+    na = ""
+  )
+# Tables
+mod_df |>
+  filter(ReviewYear == first_cfrd) |>
+  select(
+    a1c_ogtt,
+    age_dx,
+    A_weight,
+    A_height,
+    bmi_adult,
+    bmi_perc,
+    bmi_cat,
+    A_FEV1_pct_predicted,
+    A_FVC_pct_predicted,
+    Race,
+    Hispanicrace,
+    MutClass,
+    a1c,
+    A_ogttfast_bloodglucose,
+    A_twohour_bloodglucose,
+    A_IsOnEnzymes,
+    liver_disease,
+    A_pseudomonasaeruginosa,
+    A_fungalyeast1,
+    A_mycocultureresults3,
+    A_cultureresults3,
+    treatment,
+    cfrd_treatment_options2,
+    cfrd_treatment_options3,
+    cfrd_treatment_options4,
+    years_with_data
+  ) |>
+  tbl_summary(
+    by = a1c_ogtt,
+    missing_text = "Missing",
+    statistic = list(all_continuous() ~ "{mean} ({sd})")
+  ) |>
+  add_overall() |>
+  add_p(
+    test = list(
+      all_continuous() ~ "oneway.test",
+      "years_with_data" ~ "kruskal.test"
+    )
+  ) |>
+  modify_caption("**All Groups**")
+mod_df |>
+  filter(ReviewYear == first_cfrd) |>
+  filter(a1c_ogtt %in% c("OGTT", "HbA1c Only")) |>
+  mutate(a1c_ogtt = droplevels(a1c_ogtt)) |>
+  select(
+    a1c_ogtt,
+    age_dx,
+    A_weight,
+    A_height,
+    A_bmivalue,
+    A_bmipercentile,
+    A_FEV1_pct_predicted,
+    A_FVC_pct_predicted,
+    Race,
+    Hispanicrace,
+    MutClass,
+    a1c,
+    A_ogttfast_bloodglucose,
+    A_twohour_bloodglucose,
+    A_IsOnEnzymes,
+    liver_disease,
+    A_pseudomonasaeruginosa,
+    A_fungalyeast1,
+    A_mycocultureresults3,
+    A_cultureresults3,
+    treatment,
+    cfrd_treatment_options2,
+    cfrd_treatment_options3,
+    cfrd_treatment_options4,
+    years_with_data
+  ) |>
+  tbl_summary(
+    by = a1c_ogtt,
+    missing_text = "Missing",
+    statistic = list(all_continuous() ~ "{mean} ({sd})")
+  ) |>
+  add_overall() |>
+  add_p(test = list(all_continuous() ~ "oneway.test")) |>
+  modify_caption("**OGTT vs. HbA1c Only**")
+```
+
+## FEV1
+
+### LOESS-smoothed plot
+
+```{r}
+#| warning: false
+#| message: false
+# Plot by diagnosis type
+ggplot(
+  mod_df,
+  aes(
+    y = A_FEV1_pct_predicted,
+    x = years_to_cfrd,
+    color = a1c_ogtt,
+    group = a1c_ogtt
+  )
+) +
   xlab("Years from CFRD") +
-  ylab("FVC % Predicted") +
-  theme_classic(base_size = 15)
-# Contrasts
-neither_pre <- rep(0, nrow(summary(m)$coefficients))
-neither_change <- rep(0, nrow(summary(m)$coefficients))
-neither_pre[which(rownames(summary(m)$coefficients) == "years_to_cfrd")] = 1
-neither_change[which(rownames(summary(m)$coefficients) == "x")] = 1
-a1c_pre <- rep(0, nrow(summary(m)$coefficients))
-a1c_change <- rep(0, nrow(summary(m)$coefficients))
-a1c_pre[which(
-  rownames(summary(m)$coefficients) %in%
-    c("years_to_cfrd", "years_to_cfrd:a1c_ogttHbA1c Only")
-)] = 1
-a1c_change[which(
-  rownames(summary(m)$coefficients) %in%
-    c("x", "a1c_ogttHbA1c Only:x")
-)] = 1
-ogtt_pre <- rep(0, nrow(summary(m)$coefficients))
-ogtt_change <- rep(0, nrow(summary(m)$coefficients))
-ogtt_pre[which(
-  rownames(summary(m)$coefficients) %in%
-    c("years_to_cfrd", "years_to_cfrd:a1c_ogttOGTT")
-)] = 1
-ogtt_change[which(
-  rownames(summary(m)$coefficients) %in%
-    c("x", "a1c_ogttOGTT:x")
-)] = 1
-contr <- rbind(
-  "HbA1c Only Pre-CFRD" = a1c_pre,
-  "HbA1c Only Post-CFRD" = a1c_pre + a1c_change,
-  "OGTT Pre-CFRD" = ogtt_pre,
-  "OGTT Post-CFRD" = ogtt_pre + ogtt_change,
-  "Neither Pre-CFRD" = neither_pre,
-  "Neither Post-CFRD" = neither_pre + neither_change
+  geom_smooth() +
+  theme_bw()
+```
+
+### Linear model results
+
+#### Not adjusted for treatment status
+
+Adjusted for age at diagnosis, sex, mutation class, and pancreatic enzyme use.
+
+```{r}
+# Fit the model
+m <- lmer(
+  A_FEV1_pct_predicted ~ years_to_cfrd *
+    a1c_ogtt +
+    age_dx +
+    Gender +
+    MutClass +
+    A_IsOnEnzymes +
+    (1 | eDWID),
+  data = mod_df
 )
-change_mod_slopes <- glht(m, linfct = contr)
-change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
-change_mod_slopes_ci$confint |>
-  as.data.frame() |>
-  gt(caption = "Slopes by Diagnosis Method", rownames_to_stub = T) |>
-  fmt_number(decimals = 2)
-# Model table
+t3 = car::Anova(m, type = 'III', contrasts = "contr.sum")
+emm = emmeans(m, specs = ~ years_to_cfrd + a1c_ogtt, pbkrtest.limit = 1e5)
+emm = data.frame(emm)
+ggplot(
+  emm,
+  aes(x = years_to_cfrd, y = emmean, group = a1c_ogtt, color = a1c_ogtt)
+) +
+  geom_point() +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.25) +
+  geom_line() +
+  xlab("Years From CFRD Dx") +
+  ylab("Estimated Mean ppFEV1") +
+  labs(color = "Diagnosis Method") +
+  theme_bw()
+# Model tables
+kable(emm, digits = 3)
 m |> tbl_regression() |> add_n(location = "level")
+t3
 ```
+
+The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[nrow(t3),"Pr(>Chisq)"], eps = 0.001)`).
+
+#### Adjusted for treatment status
+
+Adjusted for age at diagnosis, sex, mutation class, pancreatic enzyme use, and CFRD treatment.
+
+```{r}
+# Fit the model
+m <- lmer(
+  A_FEV1_pct_predicted ~ years_to_cfrd *
+    a1c_ogtt +
+    age_dx +
+    Gender +
+    MutClass +
+    A_IsOnEnzymes +
+    treatment +
+    (1 | eDWID),
+  data = mod_df
+)
+t3 = car::Anova(m, type = 'III', contrasts = "contr.sum")
+emm = emmeans(m, specs = ~ years_to_cfrd + a1c_ogtt, pbkrtest.limit = 1e5)
+emm = data.frame(emm)
+ggplot(
+  emm,
+  aes(x = years_to_cfrd, y = emmean, group = a1c_ogtt, color = a1c_ogtt)
+) +
+  geom_point() +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.25) +
+  geom_line() +
+  xlab("Years From CFRD Dx") +
+  ylab("Estimated Mean ppFEV1") +
+  labs(color = "Diagnosis Method") +
+  theme_bw()
+# Model tables
+kable(emm, digits = 3)
+m |> tbl_regression() |> add_n(location = "level")
+t3
+```
+
+The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[nrow(t3),"Pr(>Chisq)"], eps = 0.001)`).
+
+## FVC
+
+### LOESS-smoothed plot
+
+```{r}
+#| warning: false
+#| message: false
+# Plot by diagnosis type
+ggplot(
+  mod_df,
+  aes(
+    y = A_FVC_pct_predicted,
+    x = years_to_cfrd,
+    color = a1c_ogtt,
+    group = a1c_ogtt
+  )
+) +
+  xlab("Years from CFRD") +
+  geom_smooth() +
+  theme_bw()
+```
+
+### Linear model results
+
+#### Not adjusted for treatment status
+
+Adjusted for age at diagnosis, sex, mutation class, and pancreatic enzyme use.
+
+```{r}
+# Fit the model
+m <- lmer(
+  A_FVC_pct_predicted ~ years_to_cfrd *
+    a1c_ogtt +
+    age_dx +
+    Gender +
+    MutClass +
+    A_IsOnEnzymes +
+    (1 | eDWID),
+  data = mod_df
+)
+t3 = car::Anova(m, type = 'III', contrasts = "contr.sum")
+emm = emmeans(m, specs = ~ years_to_cfrd + a1c_ogtt, pbkrtest.limit = 1e5)
+emm = data.frame(emm)
+ggplot(
+  emm,
+  aes(x = years_to_cfrd, y = emmean, group = a1c_ogtt, color = a1c_ogtt)
+) +
+  geom_point() +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.25) +
+  geom_line() +
+  xlab("Years From CFRD Dx") +
+  ylab("Estimated Mean ppFVC") +
+  labs(color = "Diagnosis Method") +
+  theme_bw()
+# Model tables
+kable(emm, digits = 3)
+m |> tbl_regression() |> add_n(location = "level")
+t3
+```
+
+The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[nrow(t3),"Pr(>Chisq)"], eps = 0.001)`).
+
+#### Adjusted for treatment status
+
+Adjusted for age at diagnosis, sex, mutation class, pancreatic enzyme use, and CFRD treatment.
+
+```{r}
+# Fit the model
+m <- lmer(
+  A_FVC_pct_predicted ~ years_to_cfrd *
+    a1c_ogtt +
+    age_dx +
+    Gender +
+    MutClass +
+    A_IsOnEnzymes +
+    treatment +
+    (1 | eDWID),
+  data = mod_df
+)
+t3 = car::Anova(m, type = 'III', contrasts = "contr.sum")
+emm = emmeans(m, specs = ~ years_to_cfrd + a1c_ogtt, pbkrtest.limit = 1e5)
+emm = data.frame(emm)
+ggplot(
+  emm,
+  aes(x = years_to_cfrd, y = emmean, group = a1c_ogtt, color = a1c_ogtt)
+) +
+  geom_point() +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.25) +
+  geom_line() +
+  xlab("Years From CFRD Dx") +
+  ylab("Estimated Mean ppFVC") +
+  labs(color = "Diagnosis Method") +
+  theme_bw()
+# Model tables
+kable(emm, digits = 3)
+m |> tbl_regression() |> add_n(location = "level")
+t3
+```
+
+The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[nrow(t3),"Pr(>Chisq)"], eps = 0.001)`).

### 2025-12-10 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
sweat test formatting
updated report per Angela's changes
updated sweat test figures

--- 2025-08-15 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
@@
     code-fold: true
-    self-contained: true
+    embed-resources: true
     fig-cap-location: top
@@
 library(pROC)
-home_dir <- switch(Sys.info()["sysname"],
+library(patchwork)
+home_dir <- switch(
+  Sys.info()["sysname"],
   "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
@@
 chco <- full_join(chco, chco_diagnoses, by = join_by(PatientID))
+# Manual fixes per Angela's 11/21/25 email
+chco$Diagnosis[chco$PatientID == 549] = "CFTR-related disorder"
+chco$Diagnosis[chco$PatientID == 650] = "Cystic Fibrosis"
+chco$Diagnosis[chco$PatientID == 653] = "CFTR-related disorder"
+chco$Diagnosis[chco$PatientID == 693] = "Cystic Fibrosis"
+chco$Diagnosis[chco$PatientID == 702] = "Cystic Fibrosis"
+chco$Diagnosis[chco$PatientID == 721] = "Cystic Fibrosis"
+chco$Diagnosis[chco$PatientID == 736] = "CFTR-related disorder"
+chco[nrow(chco) + 1, ] <- NA
+chco$PatientID[nrow(chco)] = 692
+chco$Age[nrow(chco)] = (mdy("3/24/2014") - mdy("9/26/13")) / 365.25
+chco$Diagnosis[nrow(chco)] = "CRMS"
+chco$Na_Left[nrow(chco)] = 24.1
+chco$Na_Right[nrow(chco)] = 21.1
+chco$Cl[nrow(chco)] = 30
+chco$Cl_Right[nrow(chco)] = 30
+chco[nrow(chco) + 1, ] <- NA
+chco$PatientID[nrow(chco)] = 724
+chco$Age[nrow(chco)] = (mdy("5/08/2015") - mdy("1/18/09")) / 365.25
+chco$Diagnosis[nrow(chco)] = "CFTR-related disorder"
+chco$Na_Left[nrow(chco)] = 72.4
+chco$Na_Right[nrow(chco)] = 72.4
+chco$Cl_Left[nrow(chco)] = 64
+chco$Cl_Right[nrow(chco)] = 59
 chco$PatientID <- NULL
@@
 df <- df[-which(duplicated(df$Name)), ]
+# Age groups
+df$`Age Group` = cut(
+  df$Age,
+  breaks = c(-Inf, 18, Inf),
+  labels = c("< 18", ">= 18"),
+  right = FALSE
+)
+# Specific genotypes
+df$Genotypes1[
+  !df$Genotypes1 %in%
+    c(
+      "F508del",
+      "G542X",
+      "G551D",
+      "W1282X",
+      "N1303K",
+      "R117H"
+    )
+] = "Other"
+df$Genotypes2[
+  !df$Genotypes2 %in%
+    c(
+      "F508del",
+      "G542X",
+      "G551D",
+      "W1282X",
+      "N1303K",
+      "R117H"
+    )
+] = "Other"
+df$Genotype = paste0(df$Genotypes1, "/", df$Genotypes2)
+df$Genotype = factor(
+  df$Genotype,
+  levels = c(
+    "F508del/F508del",
+    "F508del/G542X",
+    "F508del/G551D",
+    "F508del/N1303K",
+    "F508del/Other",
+    "F508del/R117H",
+    "F508del/W1282X",
+    "G542X/G551D",
+    "G542X/N1303K",
+    "G542X/Other",
+    "G542X/R117H",
+    "G551D/Other",
+    "G551D/R117H",
+    "N1303K/G542X",
+    "N1303K/N1303K",
+    "N1303K/Other",
+    "Other/F508del",
+    "Other/G542X",
+    "Other/Other",
+    "Other/W1282X",
+    "R117H/F508del",
+    "R117H/Other",
+    "W1282X/N1303K"
+  ),
+  labels = c(
+    "F508del/F508del",
+    "F508del/G542X",
+    "F508del/G551D",
+    "F508del/N1303K",
+    "F508del/Other",
+    "F508del/R117H",
+    "F508del/W1282X",
+    "Other",
+    "Other",
+    "Other",
+    "Other",
+    "Other",
+    "Other",
+    "Other",
+    "Other",
+    "Other",
+    "F508del/Other",
+    "Other",
+    "Other",
+    "Other",
+    "F508del/R117H",
+    "Other",
+    "Other"
+  )
+)
+# Order factor variables
+df$Diagnosis = factor(
+  df$Diagnosis,
+  levels = c("Non-CF", "CFTR-RD or CRMS", "Cystic Fibrosis")
+)
+df$`Genotype Risk` = factor(
+  df$`Genotype Risk`,
+  levels = c("Not CF", "Mild", "Severe", "Unknown")
+)
 # Set table options
 mycontrols <- tableby.control(
-    numeric.stats = c("N", "Nmiss", "meansd", "medianq1q3", "iqr", "range")
+  numeric.stats = c("N", "Nmiss", "meansd", "medianq1q3", "iqr", "range"),
+  simulate.p.value = TRUE
 )
@@
       Age +
+      fe(`Age Group`) +
+      fe(Genotype) +
       fe(`Genotype Risk`) +
@@
       Age +
+      fe(`Age Group`) +
+      fe(Genotype) +
       fe(`Cl Group`) +
@@
   tableby(
-        `Cl Group` ~ Sex + Age + Cl + Na + `Na:Cl` + K,
+    `Cl Group` ~ Sex +
+      Age +
+      fe(`Age Group`) +
+      fe(Genotype) +
+      Cl +
+      Na +
+      `Na:Cl` +
+      K,
     data = df,
@@
 #| warning: false
-clp <- ggplot(df, aes(x = Diagnosis, y = Cl)) +
+group_n <- df |> group_by(Diagnosis) |> summarise(n = n()) |> mutate(pos = 160)
+clp_d <- ggplot(df, aes(x = Diagnosis, y = Cl)) +
   geom_boxplot() +
+  geom_label(
+    data = group_n,
+    aes(x = Diagnosis, y = pos, label = paste("n =", n))
+  ) +
   theme_bw()
-clp
+clp_d
 ```
@@
 #| warning: false
-clp <- ggplot(df, aes(x = `Genotype Risk`, y = Cl)) +
+geno_n <- df |>
+  group_by(`Genotype Risk`) |>
+  summarise(n = n()) |>
+  mutate(pos = 160)
+clp_g <- ggplot(df, aes(x = `Genotype Risk`, y = Cl)) +
   geom_boxplot() +
+  geom_label(
+    data = geno_n,
+    aes(x = `Genotype Risk`, y = pos, label = paste("n =", n))
+  ) +
   theme_bw()
-clp
+clp_g
 ```
@@
 #| warning: false
-nap <- ggplot(df, aes(x = Diagnosis, y = Na)) +
+nap_d <- ggplot(df, aes(x = Diagnosis, y = Na)) +
   geom_boxplot() +
+  geom_label(
+    data = group_n,
+    aes(x = Diagnosis, y = pos, label = paste("n =", n))
+  ) +
   theme_bw()
-nap
+nap_d
 ```
@@
 #| warning: false
-nap <- ggplot(df, aes(x = `Genotype Risk`, y = Na)) +
+nap_g <- ggplot(df, aes(x = `Genotype Risk`, y = Na)) +
   geom_boxplot() +
+  geom_label(
+    data = geno_n,
+    aes(x = `Genotype Risk`, y = pos, label = paste("n =", n))
+  ) +
   theme_bw()
-nap
+nap_g
 ```
@@
 #| warning: false
-ratiop <- ggplot(df, aes(x = Diagnosis, y = `Na:Cl`)) +
+group_n$pos = 4
+ratiop_d <- ggplot(df, aes(x = Diagnosis, y = `Na:Cl`)) +
   geom_boxplot() +
+  geom_label(
+    data = group_n,
+    aes(x = Diagnosis, y = pos, label = paste("n =", n))
+  ) +
   geom_hline(yintercept = 1, color = "red") +
@@
   theme_bw() +
-    ylab("Na:Cl") +
-    xlab("Modulator")
-ratiop
-ratiop <- ggplot(df, aes(x = Diagnosis, y = `Na:Cl`)) +
-    geom_boxplot() +
-    geom_hline(yintercept = 1, color = "red") +
-    geom_hline(yintercept = 1.25, color = "red") +
-    ylim(0, 2) +
-    theme_bw() +
-    ylab("Na:Cl") +
-    xlab("Modulator")
-ratiop
+  ylab("Na:Cl")
+ratiop_d
 ```
@@
 #| warning: false
-ratiop <- ggplot(df, aes(x = `Genotype Risk`, y = `Na:Cl`)) +
+geno_n$pos = 4
+ratiop_g <- ggplot(df, aes(x = `Genotype Risk`, y = `Na:Cl`)) +
   geom_boxplot() +
+  geom_label(
+    data = geno_n,
+    aes(x = `Genotype Risk`, y = pos, label = paste("n =", n))
+  ) +
   geom_hline(yintercept = 1, color = "red") +
@@
   theme_bw() +
-    ylab("Na:Cl") +
-    xlab("Modulator")
-ratiop
-ratiop <- ggplot(df, aes(x = `Genotype Risk`, y = `Na:Cl`)) +
-    geom_boxplot() +
-    geom_hline(yintercept = 1, color = "red") +
-    geom_hline(yintercept = 1.25, color = "red") +
-    ylim(0, 2) +
-    theme_bw() +
-    ylab("Na:Cl") +
-    xlab("Modulator")
-ratiop
+  ylab("Na:Cl")
+ratiop_g
+```
+
+```{r}
+p1 = (clp_d + ratiop_d) / (clp_g + ratiop_g)
+p1
+ggsave(
+  filename = "./Dissemination/Figures/fig1.png",
+  plot = p1,
+  height = 9,
+  width = 12,
+  units = "in",
+  dpi = 600
+)
 ```
@@
   theme_bw() +
-    ylab("Na:Cl") +
-    xlab("Modulator")
-ratiop
-ratiop <- ggplot(df, aes(x = `Genotype Risk`, y = `Na:Cl`)) +
-    geom_boxplot() +
-    geom_hline(yintercept = 1, color = "red") +
-    geom_hline(yintercept = 1.25, color = "red") +
-    ylim(0, 2) +
-    theme_bw() +
-    ylab("Na:Cl") +
-    xlab("Modulator")
-ratiop
+  ylab("Na:Cl")
+ratiop_g
+```
+
+```{r}
+p2 = (clp_d + ratiop_d) / (clp_g + ratiop_g)
+p2
+ggsave(
+  filename = "./Dissemination/Figures/fig2.png",
+  plot = p2,
+  height = 9,
+  width = 12,
+  units = "in",
+  dpi = 600
+)
 ```

### 2025-12-16 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
updated Cl filter, etc.

@@
 #| message: false
+setwd(home_dir)
 # Import CHCO, filter out those with no demo data, rename columns, etc.
@@
   ) +
-  geom_hline(yintercept = 1, color = "red") +
-  geom_hline(yintercept = 1.25, color = "red") +
   theme_bw() +
@@
 # Fit a multiclass ROC
-mc_roc <- multiclass.roc(df$Diagnosis, df$`Na:Cl`, quiet = T)
+mc_roc <- multiclass.roc(
+  df$Diagnosis,
+  df$`Na:Cl`,
+  quiet = T,
+  levels = c("Non-CF", "CFTR-RD or CRMS", "Cystic Fibrosis")
+)
 roc_list <- list(
   "Cystic Fibrosis vs. Non-CF" = mc_roc$rocs[[2]],
-  "CFTR-RD or CRMS vs. Cystic Fibrosis" = mc_roc$rocs[[1]],
-  "CFTR-RD or CRMS vs. Non-CF" = mc_roc$rocs[[3]]
+  "CFTR-RD or CRMS vs. Cystic Fibrosis" = mc_roc$rocs[[3]],
+  "CFTR-RD or CRMS vs. Non-CF" = mc_roc$rocs[[1]]
 )
@@
 
-# Excluding Cl $\leq$ 30
-
-```{r}
-df <- df %>% filter(Cl > 30)
+# Excluding Cl < 30
+
+```{r}
+df <- df %>% filter(Cl >= 30)
 ```

### 2025-12-19 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
Air formatting for Kam

@@
 library(emmeans)
-home_dir <- switch(Sys.info()["sysname"],
+home_dir <- switch(
+  Sys.info()["sysname"],
   "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",

### 2025-12-19 Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R

@@ -1,29 +1,38 @@
 library(tidyverse)
 library(lubridate)
 # Paths
-home_dir <- switch(Sys.info()["sysname"],
+home_dir <- switch(
+  Sys.info()["sysname"],
   "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
   "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
   "Linux" = "/home/tim/CF"
 )
 setwd(home_dir)
 # Import
-encounter <- read.csv("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_encountersMerged_Del1.csv",
+encounter <- read.csv(
+  "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_encountersMerged_Del1.csv",
   na.strings = ""
 )
-annual <- read.csv("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_Annualized_Del1.csv",
+annual <- read.csv(
+  "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_Annualized_Del1.csv",
   na.strings = ""
 )
-demo <- read.csv("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_DemogCFDiag_Del1.csv",
+demo <- read.csv(
+  "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_DemogCFDiag_Del1.csv",
   na.strings = ""
 )
-hospitalizations <- read.csv("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_CareEpisodes_Del1.csv",
+hospitalizations <- read.csv(
+  "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Raw/DataDelivery_20240125/CFF22_CareEpisodes_Del1.csv",
   na.strings = ""
 )
 # List of outcomes
 continuous_outcomes <- c(
[…]

### 2025-12-22 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
update sweat variability tables

@@
   labels = c(
-    "Mild",
-    "Mild",
+    "Low",
+    "Low",
     "Unknown",
-    "Mild",
-    "Severe",
+    "Low",
+    "High",
     "Unknown",
@@
 )
+# All genotypes
+df$`Detailed Genotype` = paste0(df$Genotypes1, "/", df$Genotypes2)
 # Specific genotypes
@@
   df$Diagnosis,
-  levels = c("Non-CF", "CFTR-RD or CRMS", "Cystic Fibrosis")
+  levels = c("Cystic Fibrosis", "CFTR-RD or CRMS", "Non-CF")
 )
@@
   df$`Genotype Risk`,
-  levels = c("Not CF", "Mild", "Severe", "Unknown")
+  levels = c("Not CF", "Low", "High", "Unknown")
 )
@@
   file = "./Data_Cleaned/sweat_variability_analysis_dataset.csv",
+  row.names = F,
+  na = ""
+)
+# Make detailed genotype table for manuscript
+geno_tab = xtabs(
+  ~ `Detailed Genotype` + `Genotype Risk` + Diagnosis,
+  data = df,
+  addNA = T
+) |>
+  data.frame() |>
+  filter(Diagnosis == "Cystic Fibrosis") |>
+  select(-Diagnosis) |>
+  pivot_wider(names_from = Genotype.Risk, values_from = Freq)
+categories = c("Not CF", "Low", "High", "Unknown")
+geno_tab$Category = apply(geno_tab[, categories], 1, function(r) {
+  w = which(r != 0)
+  if (length(w) > 0) {
+    return(categories[w])
+  } else {
+    return(NA)
+  }
+})
+geno_tab$Count = rowSums(geno_tab[, c("Not CF", "Low", "High", "Unknown")])
+geno_tab$Percentage = scales::percent(geno_tab$Count / sum(geno_tab$Count))
+geno_tab$`Num. CF Participants` = paste0(
+  geno_tab$Count,
+  " (",
+  geno_tab$Percentage,
+  ")"
+)
+geno_tab = geno_tab |>
+  filter(!is.na(Category)) |>
+  select(Detailed.Genotype, `Num. CF Participants`, Category) |>
+  rename(`Detailed Genotype` = Detailed.Genotype)
+write.csv(
+  geno_tab,
+  file = "./Dissemination/Genotypes.csv",
   row.names = F,
@@
 
-## Table 2: Participant Demographics by Genotype Risk
+## Table 2: CF Demographics by Genotype Risk
 
@@
 )
-roc_list <- list(
-  "Cystic Fibrosis vs. Non-CF" = mc_roc$rocs[[2]],
-  "CFTR-RD or CRMS vs. Cystic Fibrosis" = mc_roc$rocs[[3]],
-  "CFTR-RD or CRMS vs. Non-CF" = mc_roc$rocs[[1]]
-)
+roc_list = mc_roc$rocs
+names(roc_list) <- lapply(mc_roc$rocs, function(r) {
+  paste(r$levels, collapse = " vs. ")
+})
 ggroc(roc_list, legacy.axes = T) +
@@
 plot(
+  roc_list[[1]],
+  print.auc = T,
+  print.thres = T,
+  main = names(roc_list)[1]
+)
+plot(
   roc_list[[2]],
@@
   print.thres = T,
-  main = "Cystic Fibrosis vs. CFTR-RD or CRMS"
-)
-plot(
-  roc_list[[1]],
-  print.auc = T,
-  print.thres = T,
-  main = "Cystic Fibrosis vs. Non-CF"
+  main = names(roc_list)[2]
 )
@@
   print.thres = T,
-  main = "CFTR-RD or CRMS vs. Non-CF"
+  main = names(roc_list)[3]
 )
@@
   data = df,
-  levels = c("Mild", "Severe"),
+  levels = c("Low", "High"),
   quiet = T
@@
   print.thres = T,
-  main = "Mild vs. Severe Genotype"
+  main = "Low vs. High Risk Genotype"
 )

### 2025-12-22 genotype_class_severity.R

@@ -71,7 +71,8 @@ cf_genotype_class_severity <- function(
     "R347H",
     "D1152H",
     "3849+10kbC>T",
-    "TG11-5T"
+    "TG11-5T",
+    "TG11"
   )
   IV <- unique(tolower(gsub("[[:punct:]]|\\s", "", IV)))
   IV <- matrix(c(IV, class = rep("IV", length(IV))), ncol = 2)

### 2025-12-23 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
a bunch of updates for Christine

@@
 
+P values below are FDR-adjusted for multiple comparisons.
+
 #### <140 vs. [140,200)
@@
 roc_df |>
-  select(`OGTT 2-hour`, Gender, MutClass, hgba1c) |>
+  select(-eDWID) |>
   filter(`OGTT 2-hour` != ">=200") |>
@@
     missing_text = "Missing",
-    type = list(all_continuous() ~ "continuous2"),
+    type = list(
+      reviewyear ~ "categorical",
+      all_continuous() ~ "continuous2"
+    ),
     statistic = list(
@@
   ) |>
+  add_q() |>
   separate_p_footnotes() |>
+  modify_column_hide(p.value) |>
   as_gt()
@@
 roc_df |>
-  select(`OGTT 2-hour`, Gender, MutClass, hgba1c) |>
+  select(-eDWID) |>
   filter(`OGTT 2-hour` != "<140") |>
@@
 roc_df |>
-  select(`OGTT 2-hour`, Gender, MutClass, hgba1c) |>
+  select(-eDWID) |>
   filter(`OGTT 2-hour` != "[140,200)") |>

### 2025-12-23 Christine Chan/CF Registry/cf_registry_lung_function_analysis.qmd
a bunch of updates for Christine

@@
 var_label(cfrd_only$treatment) = "CFRD Treatment Started Within 1 Year?"
-# Calculate number of visits (just for table 1)
-n_vis = annual |>
-  group_by(eDWID) |>
-  summarise(years_with_data = n())
-cfrd_only <- left_join(cfrd_only, n_vis, by = join_by(eDWID))
 ```
@@
     cfrd_treatment_options4,
-    years_with_data
+    NumFEV1,
+    NumFVC
   ) %>%
@@
     cfrd_treatment_options4,
-    years_with_data
+    NumFEV1,
+    NumFVC
   ) |>
@@
   add_overall() |>
-  add_p(
-    test = list(
-      all_continuous() ~ "oneway.test",
-      "years_with_data" ~ "kruskal.test"
-    )
-  ) |>
+  add_p() |>
   modify_caption("**All Groups**")
@@
 
-### LOESS-smoothed plot
-
-```{r}
-#| warning: false
-#| message: false
-# Plot by diagnosis type
-ggplot(
-  mod_df,
-  aes(
-    y = A_FEV1_pct_predicted,
-    x = years_to_cfrd,
-    color = a1c_ogtt,
-    group = a1c_ogtt
-  )
-) +
-  xlab("Years from CFRD") +
-  geom_smooth() +
-  theme_bw()
-```
-
-### Linear model results
-
-#### Not adjusted for treatment status
-
-Adjusted for age at diagnosis, sex, mutation class, and pancreatic enzyme use.
+### Not adjusted for treatment status
+
+Adjusted for age at diagnosis, number of FEV1 measures, sex, mutation class, and pancreatic enzyme use.
 
@@
     age_dx +
+    NumFEV1 +
     Gender +
@@
 emm = emmeans(m, specs = ~ years_to_cfrd + a1c_ogtt, pbkrtest.limit = 1e5)
-emm = data.frame(emm)
 ggplot(
-  emm,
+  data.frame(emm),
   aes(x = years_to_cfrd, y = emmean, group = a1c_ogtt, color = a1c_ogtt)
@@
 # Model tables
-kable(emm, digits = 3)
 m |> tbl_regression() |> add_n(location = "level")
-t3
 ```
@@
 
-#### Adjusted for treatment status
-
-Adjusted for age at diagnosis, sex, mutation class, pancreatic enzyme use, and CFRD treatment.
+```{r}
+# Contrasts - probably a more elegant way to do this but oh well
+neither_m_5 = rep(0, times = 33)
+neither_m_5[1] = 1
+neither_m_1 = rep(0, times = 33)
+neither_m_1[5] = 1
+neither_0 = rep(0, times = 33)
+neither_0[6] = 1
+neither_1 = rep(0, times = 33)
+neither_1[7] = 1
+neither_5 = rep(0, times = 33)
+neither_5[11] = 1
+hba1c_m_5 = rep(0, times = 33)
+hba1c_m_5[12] = 1
+hba1c_m_1 = rep(0, times = 33)
+hba1c_m_1[16] = 1
+hba1c_0 = rep(0, times = 33)
+hba1c_0[17] = 1
+hba1c_1 = rep(0, times = 33)
+hba1c_1[18] = 1
+hba1c_5 = rep(0, times = 33)
+hba1c_5[22] = 1
+ogtt_m_5 = rep(0, times = 33)
+ogtt_m_5[23] = 1
+ogtt_m_1 = rep(0, times = 33)
+ogtt_m_1[27] = 1
+ogtt_0 = rep(0, times = 33)
+ogtt_0[28] = 1
+ogtt_1 = rep(0, times = 33)
+ogtt_1[29] = 1
+ogtt_5 = rep(0, times = 33)
+ogtt_5[33] = 1
+# Test contrasts
+F <- contrast(
+  emm,
+  list(
+    "Neither: -5 to 0 years" = neither_0 - neither_m_5,
+    "Neither: 0 to 5 years" = neither_5 - neither_0,
+    "Neither: -1 to 0 years" = neither_0 - neither_m_1,
+    "Neither: 0 to 1 years" = neither_1 - neither_0,
+    "HbA1c Only: -5 to 0 years" = hba1c_0 - hba1c_m_5,
+    "HbA1c Only: 0 to 5 years" = hba1c_5 - hba1c_0,
+    "HbA1c Only: -1 to 0 years" = hba1c_0 - hba1c_m_1,
+    "HbA1c Only: 0 to 1 years" = hba1c_1 - hba1c_0,
+    "OGTT: -5 to 0 years" = ogtt_0 - ogtt_m_5,
+    "OGTT: 0 to 5 years" = ogtt_5 - ogtt_0,
+    "OGTT: -1 to 0 years" = ogtt_0 - ogtt_m_1,
+    "OGTT: 0 to 1 years" = ogtt_1 - ogtt_0
+  )
+)
+kable(F, digits = 3)
+kable(data.frame(emm), digits = 3)
+```
+
+The table above indicates the change in average lung function from a given timepoint to another within each group.
+For example, a value of -13.9 for "Neither: -5 to 0 years" indicates that among people without HbA1c or OGTT, lung function declined by 13 percentage points on average from 5 years prior to diagnosis to time of diagnosis.
+
+### Adjusted for treatment status
+
+Adjusted for age at diagnosis, number of FEV1 measures, sex, mutation class, pancreatic enzyme use, and CFRD treatment.
 
@@
   theme_bw()
-# Model tables
-kable(emm, digits = 3)
+# Model table
 m |> tbl_regression() |> add_n(location = "level")
-t3
 ```
@@
 
+```{r}
+# Test contrasts
+F <- contrast(
+  emm,
+  list(
+    "Neither: -5 to 0 years" = neither_0 - neither_m_5,
+    "Neither: 0 to 5 years" = neither_5 - neither_0,
+    "Neither: -1 to 0 years" = neither_0 - neither_m_1,
+    "Neither: 0 to 1 years" = neither_1 - neither_0,
+    "HbA1c Only: -5 to 0 years" = hba1c_0 - hba1c_m_5,
+    "HbA1c Only: 0 to 5 years" = hba1c_5 - hba1c_0,
+    "HbA1c Only: -1 to 0 years" = hba1c_0 - hba1c_m_1,
+    "HbA1c Only: 0 to 1 years" = hba1c_1 - hba1c_0,
+    "OGTT: -5 to 0 years" = ogtt_0 - ogtt_m_5,
+    "OGTT: 0 to 5 years" = ogtt_5 - ogtt_0,
+    "OGTT: -1 to 0 years" = ogtt_0 - ogtt_m_1,
+    "OGTT: 0 to 1 years" = ogtt_1 - ogtt_0
+  )
+)
+kable(F, digits = 3)
+kable(data.frame(emm), digits = 3)
+```
+
+The table above indicates the change in average lung function from a given timepoint to another within each group.
+For example, a value of -13.9 for "Neither: -5 to 0 years" indicates that among people without HbA1c or OGTT, lung function declined by 13 percentage points on average from 5 years prior to diagnosis to time of diagnosis.
+
 ## FVC
 
-### LOESS-smoothed plot
-
-```{r}
-#| warning: false
-#| message: false
-# Plot by diagnosis type
-ggplot(
-  mod_df,
-  aes(
-    y = A_FVC_pct_predicted,
-    x = years_to_cfrd,
-    color = a1c_ogtt,
-    group = a1c_ogtt
-  )
-) +
-  xlab("Years from CFRD") +
-  geom_smooth() +
-  theme_bw()
-```
-
-### Linear model results
-
-#### Not adjusted for treatment status
-
-Adjusted for age at diagnosis, sex, mutation class, and pancreatic enzyme use.
+### Not adjusted for treatment status
+
+Adjusted for age at diagnosis, number of FVC measures, sex, mutation class, and pancreatic enzyme use.
 
@@
     age_dx +
+    NumFVC +
     Gender +
@@
 
-#### Adjusted for treatment status
-
-Adjusted for age at diagnosis, sex, mutation class, pancreatic enzyme use, and CFRD treatment.
+```{r}
+# Test contrasts
+F <- contrast(
+  emm,
+  list(
+    "Neither: -5 to 0 years" = neither_0 - neither_m_5,
+    "Neither: 0 to 5 years" = neither_5 - neither_0,
+    "Neither: -1 to 0 years" = neither_0 - neither_m_1,
+    "Neither: 0 to 1 years" = neither_1 - neither_0,
+    "HbA1c Only: -5 to 0 years" = hba1c_0 - hba1c_m_5,
+    "HbA1c Only: 0 to 5 years" = hba1c_5 - hba1c_0,
+    "HbA1c Only: -1 to 0 years" = hba1c_0 - hba1c_m_1,
+    "HbA1c Only: 0 to 1 years" = hba1c_1 - hba1c_0,
+    "OGTT: -5 to 0 years" = ogtt_0 - ogtt_m_5,
+    "OGTT: 0 to 5 years" = ogtt_5 - ogtt_0,
+    "OGTT: -1 to 0 years" = ogtt_0 - ogtt_m_1,
+    "OGTT: 0 to 1 years" = ogtt_1 - ogtt_0
+  )
+)
+kable(F, digits = 3)
+kable(data.frame(emm), digits = 3)
+```
+
+The table above indicates the change in average lung function from a given timepoint to another within each group.
+For example, a value of -13.9 for "Neither: -5 to 0 years" indicates that among people without HbA1c or OGTT, lung function declined by 13 percentage points on average from 5 years prior to diagnosis to time of diagnosis.
+
+### Adjusted for treatment status
+
+Adjusted for age at diagnosis, number of FVC measures, sex, mutation class, pancreatic enzyme use, and CFRD treatment.
 
@@
 The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[nrow(t3),"Pr(>Chisq)"], eps = 0.001)`).
+
+```{r}
+# Test contrasts
+F <- contrast(
+  emm,
+  list(
+    "Neither: -5 to 0 years" = neither_0 - neither_m_5,
+    "Neither: 0 to 5 years" = neither_5 - neither_0,
+    "Neither: -1 to 0 years" = neither_0 - neither_m_1,
+    "Neither: 0 to 1 years" = neither_1 - neither_0,
+    "HbA1c Only: -5 to 0 years" = hba1c_0 - hba1c_m_5,
+    "HbA1c Only: 0 to 5 years" = hba1c_5 - hba1c_0,
+    "HbA1c Only: -1 to 0 years" = hba1c_0 - hba1c_m_1,
+    "HbA1c Only: 0 to 1 years" = hba1c_1 - hba1c_0,
+    "OGTT: -5 to 0 years" = ogtt_0 - ogtt_m_5,
+    "OGTT: 0 to 5 years" = ogtt_5 - ogtt_0,
+    "OGTT: -1 to 0 years" = ogtt_0 - ogtt_m_1,
+    "OGTT: 0 to 1 years" = ogtt_1 - ogtt_0
+  )
+)
+kable(F, digits = 3)
+kable(data.frame(emm), digits = 3)
+```
+
+The table above indicates the change in average lung function from a given timepoint to another within each group.
+For example, a value of -13.9 for "Neither: -5 to 0 years" indicates that among people without HbA1c or OGTT, lung function declined by 13 percentage points on average from 5 years prior to diagnosis to time of diagnosis.
 
@@
 
-#### Adjusted for treatment status
-
-Adjusted for age at diagnosis, sex, mutation class, pancreatic enzyme use, and CFRD treatment.
+```{r}
+# Test contrasts
+F <- contrast(
+  emm,
+  list(
+    "Neither: -5 to 0 years" = neither_0 - neither_m_5,
+    "Neither: 0 to 5 years" = neither_5 - neither_0,
+    "Neither: -1 to 0 years" = neither_0 - neither_m_1,
+    "Neither: 0 to 1 years" = neither_1 - neither_0,
+    "HbA1c Only: -5 to 0 years" = hba1c_0 - hba1c_m_5,
+    "HbA1c Only: 0 to 5 years" = hba1c_5 - hba1c_0,
+    "HbA1c Only: -1 to 0 years" = hba1c_0 - hba1c_m_1,
+    "HbA1c Only: 0 to 1 years" = hba1c_1 - hba1c_0,
+    "OGTT: -5 to 0 years" = ogtt_0 - ogtt_m_5,
+    "OGTT: 0 to 5 years" = ogtt_5 - ogtt_0,
+    "OGTT: -1 to 0 years" = ogtt_0 - ogtt_m_1,
+    "OGTT: 0 to 1 years" = ogtt_1 - ogtt_0
+  )
+)
+kable(F, digits = 3)
+kable(data.frame(emm), digits = 3)
+```
+
+The table above indicates the change in average lung function from a given timepoint to another within each group.
+For example, a value of -13.9 for "Neither: -5 to 0 years" indicates that among people without HbA1c or OGTT, lung function declined by 13 percentage points on average from 5 years prior to diagnosis to time of diagnosis.
+
+### Adjusted for treatment status
+
+Adjusted for age at diagnosis, number of FEV1 measures, sex, mutation class, pancreatic enzyme use, and CFRD treatment.
 
@@
 The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[nrow(t3),"Pr(>Chisq)"], eps = 0.001)`).
+
+```{r}
+# Test contrasts
+F <- contrast(
+  emm,
+  list(
+    "Neither: -5 to 0 years" = neither_0 - neither_m_5,
+    "Neither: 0 to 5 years" = neither_5 - neither_0,
+    "Neither: -1 to 0 years" = neither_0 - neither_m_1,
+    "Neither: 0 to 1 years" = neither_1 - neither_0,
+    "HbA1c Only: -5 to 0 years" = hba1c_0 - hba1c_m_5,
+    "HbA1c Only: 0 to 5 years" = hba1c_5 - hba1c_0,
+    "HbA1c Only: -1 to 0 years" = hba1c_0 - hba1c_m_1,
+    "HbA1c Only: 0 to 1 years" = hba1c_1 - hba1c_0,
+    "OGTT: -5 to 0 years" = ogtt_0 - ogtt_m_5,
+    "OGTT: 0 to 5 years" = ogtt_5 - ogtt_0,
+    "OGTT: -1 to 0 years" = ogtt_0 - ogtt_m_1,
+    "OGTT: 0 to 1 years" = ogtt_1 - ogtt_0
+  )
+)
+kable(F, digits = 3)
+kable(data.frame(emm), digits = 3)
+```
+
+The table above indicates the change in average lung function from a given timepoint to another within each group.
+For example, a value of -13.9 for "Neither: -5 to 0 years" indicates that among people without HbA1c or OGTT, lung function declined by 13 percentage points on average from 5 years prior to diagnosis to time of diagnosis.

### 2025-12-23 Christine Chan/EnVision CF/fsOGTT_pre_post_ETI.qmd
a bunch of updates for Christine

@@
 multi_pre_df = df |>
+  arrange(study_id, Date) |>
   group_by(study_id) |>
@@
       units = "days"
+    )),
+    time_diff_last_first = as.numeric(difftime(
+      first(Date[pre_post == "Post"]),
+      last(Date[pre_post == "Pre"]),
+      units = "days"
     ))
@@
 var_label(multi_pre_df$time_diff_first_last) = "Days From First to Last fsOGTT"
+var_label(
+  multi_pre_df$time_diff_last_first
+) = "Days From Last Pre-ETI fsOGTT to First Post-ETI fsOGTT"
+# Calculate days between pre- and post-ETI fsOGTTs
+times = multi_pre_df |>
+  group_by(study_id, pre_post) |>
+  reframe(d = difftime(Date, lag(Date), units = "days"))
 # Save for Christine and Monique
@@
 ## Table 1: Participant characteristics at first fsOGTT
+
+The median number of days between pre-ETI fsOGTTs was `r as.numeric(summary(times$d[times$pre_post=="Pre"])[3])` (IQR: `r as.numeric(summary(times$d[times$pre_post=="Pre"])[2])`, `r as.numeric(summary(times$d[times$pre_post=="Pre"])[5])`).
+The median number of days between post-ETI fsOGTTs was `r as.numeric(summary(times$d[times$pre_post=="Post"])[3])` (IQR: `r as.numeric(summary(times$d[times$pre_post=="Post"])[2])`, `r as.numeric(summary(times$d[times$pre_post=="Post"])[5])`).
 
@@
 ```
+
+## Matsuda
+
+```{r}
+#| warning: false
+# Plot
+p = ggplot(
+  multi_pre_df,
+  aes(
+    x = years_from_eti_start,
+    y = matsuda,
+    group = study_id,
+    color = study_id
+  )
+) +
+  geom_point() +
+  geom_line() +
+  theme_bw()
+ggplotly(p)
+# Model
+m = lmer(
+  matsuda ~ years_from_eti_start_model + slope_change + (1 | study_id),
+  data = multi_pre_df
+)
+m |>
+  tbl_regression(
+    exponentiate = F,
+    estimate_fun = label_style_sigfig(digits = 4),
+    pvalue_fun = label_style_pvalue(digits = 2),
+  )
+```
